### 2020-11-16 presentations/agu_abstract.tex
add abstract text file

Title: 
Fast high-rank Hessian approximation for Bayesian ice sheet inverse problems

Abstract:
Antarctic ice sheet flow is expected to be the dominant contributor to 21st century sea level rise in a warming climate. However, a critical challenge to understanding ice sheet dynamics is the presence of significant sources of uncertainty in the governing flow equations. A leading source of uncertainty is the basal sliding friction, which cannot be measured directly. Instead, it is inferred from satellite observations of the surface ice flow velocity by solving an inverse problem governed by the nonlinear Stokes forward model.

We adopt the Bayesian inference framework to solve this inverse problem and quantify the uncertainty in the solution. The Hessian of the negative log posterior is central to numerical methods for Bayesian inference. However, existing Hessian approximations are based on low-rank approximation methods, which require computing twice as many linearized forward or adjoint partial differential equation (PDE) solves as the numerical rank of the Hessian. These methods are inefficient when the numerical rank of the Hessian is large, as is the case in continental scale ice sheet inverse problems.

We present a new method for approximating the Hessian, which allows us to form a high-rank approximation using a small number of PDE solves. Our key innovation is a product-convolution approximation that takes advantage of the Hessian's local translation invariance. This leads to fast access to matrix entries, which makes the approximation well suited for hierarchical matrix compression. We apply this product-convolution/hierarchical matrix Hessian approximation to substantially reduce the computational cost of solving the Bayesian ice sheet inverse problem.

### 2020-11-16 paper/localpsf.tex
add paper directory and files

% SIAM Article Template
\documentclass[review,onefignum,onetabnum]{siamart190516}

% Information that is shared between the article and the supplement
% (title and author information, macros, packages, etc.) goes into
% ex_shared.tex. If there is no supplement, this file can be included
% directly.

\input{localpsf_shared}

% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={Fast matrix-free approximation of smoothly varying blur operators, with application to Hessians in PDE-constrained inverse problems with highly informative data},
  pdfauthor={N. Alger, N. Petra, and O. Ghattas}
}
\fi

% The next statement enables references to information in the
% supplement. See the xr-hyperref package for details.

\externaldocument{localpsf_supplement}

% FundRef data to be entered by SIAM
%<funding-group specific-use="FundRef">
%<award-group>
%<funding-source>
%<named-content content-type="funder-name"> 
%</named-content> 
%<named-content content-type="funder-identifier"> 
%</named-content>
%</funding-source>
%<award-id> </award-id>
%</award-group>
%</funding-group>

\begin{document}

\maketitle

% REQUIRED
\begin{abstract}
  We present an efficient matrix-free method for approximating smoothly varying blur operators, and apply the method to approximate Hessians in large-scale inverse problems with highly informative data. Such Hessians are high rank and therefore expensive to approximate with existing methods. We compute the impulse responses of the operator at many points by applying the operator to a small number of Dirac combs of point sources. Within each Dirac comb, the point sources must be spaced sufficiently far apart so that the impulse responses to those point sources do not overlap with each other or with the boundary. We estimate the required spacing of point sources by applying the operator to a small number of constant, linear, and quadratic functions, then postprocessing the results. Once computed, the impulse responses are used to form a low-rank approximation of the spatially varying impulse response function (SVIR), using a continuous analog of the interpolative matrix decomposition. Smoothly varying blur operators have a low rank SVIR, even if the operator itself is high rank. We convert the resulting approximation to hierarchical matrix format, where it is inverted using fast hierarchical matrix arithmetic and used to solve linear systems.  We apply the method to a large-scale PDE-constrained inverse problem with highly informative data. 
%  {\color{red}Numerical results demonstrate that our method substantially outperforms existing methods.}
\end{abstract}

% REQUIRED
\begin{keywords}
  example, \LaTeX
\end{keywords}

% REQUIRED
\begin{AMS}
  68Q25, 68R10, 68U05
\end{AMS}

\section{Introduction}

Let $\Omega \subset \mathbb{R}^d$ be a bounded domain, let $a:L^2(\Omega) \times L^2(\Omega) \rightarrow \mathbb{R}$,
\begin{equation}
	\label{eq:kernel_representation}
	a(v,u) := \int_\Omega \int_\Omega v(y) A(y,x) u(x) dx dy,
\end{equation}
be a bilinear form with integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$, and let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be the corresponding linear operator generated by currying. That is, for each $u\in L^2(\Omega)$, the linear functional $\mathcal{A}u:L^2(\Omega)\rightarrow \mathbb{R}$ is defined by
\begin{equation*}
	(\mathcal{A}u)(v) := a(v,u).
\end{equation*}
We seek to approximate operators $\mathcal{A}$ in the ``matrix free'' setting, in which access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$. Here ($\mathcal{A}^Tv)(u) := (\mathcal{A}u)(v)$. Integral kernel representation \eqref{eq:kernel_representation}, is assumed to exist,
%e.g., by the Schwarz kernel theorem,
but one cannot easily access entries $A(y,x)$ of the kernel. Such operators arise, for example, as Schur complements in reduced space methods for solving partial differential equations (PDEs), and as Hessians in distributed parameter PDE-constrained inverse problems.

The \emph{point spread function} (PSF) of $\mathcal{A}$ at $x\in \Omega$ is
\begin{equation*}
	\varphi_x(y) := \langle \mathcal{A},\delta_x\rangle^*(y+x) = A(y+x,x).
\end{equation*}
Here $\langle \mathcal{A}, \delta_x \rangle \in L^2(\Omega)'$ is the result of applying $\mathcal{A}$ to a point source $\delta_x$ (delta distribution) centered at $x$, and $\langle\mathcal{A},\delta_x\rangle^* \in L^2(\Omega)$ is the Riesz representation of $\langle\mathcal{A},\delta_x\rangle$. The PSF, $\varphi_x$, is the result of translating $\langle\mathcal{A},\delta_x\rangle^*$ to re-center it at zero instead of $x$. Recall the Riesz representation of a linear functional $\psi \in L^2(\Omega)'$ is the unique vector $\psi^* \in L^2(\Omega)$ satisfying $\left(\psi^*, v\right)_{L^2(\Omega)} = \psi(v)$ for all $v \in \L^2(\Omega)$.

In this paper we present an efficient matrix-free method for approximating operators $\mathcal{A}$ with a product-convolution approximation, $\widetilde{A}$, of the form
\begin{equation*}
	\left(\mathcal{A}u\right) \approx \left(\widetilde{\mathcal{A}}u\right) := \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right)
\end{equation*}

PSFs that satisfy the following properties:
\begin{enumerate}
	\item $\varphi_x$ varies smoothly as a function of $x$.
	\item $\varphi_x(y)$ is zero (or small) when $|y|$ is large.
	\item $\varphi_x$ is a non-negative (or mostly non-negative) function.
\end{enumerate}
If $\varphi_x$ was constant in $x$, then $\mathcal{A}$ would be the convolution operator $\left(\mathcal{A}v\right)^* = \varphi_0 \ast v$. Operators satisfying property 1 therefore act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies smoothly as $x$ changes. Operators satisfying property 2 are local, in the sense that sources at point $x$ have no impact on points far away from $x$. Operators satisfying property 3 blur inputs: their action on a point source consists of scaling the source, then spreading the energy of the scaled point source across a region nonuniformly. We call operators satisfying these properties \emph{smoothly varying blur operators}.

The idea of our method is to sample $\varphi_{x}$ at a collection of points $x_i$ by applying $\mathcal{A}$ to a sum of point sources at the points $x_i$ (a ``Dirac comb''). The points $x_i$ are chosen so that the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ and the support of $\langle\mathcal{A},\delta_{x_j}\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$. Since the point sources do not overlap, by applying $\mathcal{A}$ to one vector, we recover $\varphi_{x_i}$ for many points $x_i$. The process is repeated to get $\varphi_{x_i}$ at more points, until an error estimate is satisfied. The $\varphi_{x_i}$ are interpolated smoothly to construct $\varphi_x$ for arbitrary $x$, which yields an approximation $\widetilde{A}(y,x)$ to the kernel $A(y,x)$, and an approximation $\widetilde{\mathcal{A}}$ to the operator $\mathcal{A}$. While $A(y,x)$ is inaccessible, we can evaluate $\widetilde{A}(y,x)$ at arbitrary $(y,x)$ points. This allows us to construct a hierarchical matrix representation of $\widetilde{\mathcal{A}}$, which allows us to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations, using fast hierarchical matrix arithmetic.

Properties 1 and 2 are essential to our method. Property 3 is only required in the subprocedure for estimating the support of $\langle\mathcal{A},\delta_x\rangle^*$, and may be dropped if other methods for estimating the support are available.

The approximation $\widetilde{A}(y,x)$ may not be accurate if both $x$ and $y$ are near the boundary. Other approximation methods, if available, should be considered for the block of the operator associated with boundary-boundary interactions. Nevertheless, in our numerical results we see that approximation works well for the the inverse problem considered, without any modifications to the boundary-boundary block.

\subsection{Motivating example: PDE constrained inverse problems with highly informative data}
\label{sec:PDE_hessian_motivation}

Our motivation for this work is approximation of Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs). That is, inverse problems in which one seeks to reconstruct an unknown parameter field, $m$, from noisy observations, 
\begin{equation*}
y=f(m,u(m)) + \text{noise},
\end{equation*}
which depend on  a state variable $u$. In turn, $u$ depends on $m$ implicitly through the solution of a PDE, 
\begin{equation}
\label{eq:state_pde}
	0=g(m,u).
\end{equation}
Here $u(m)$ denotes the solution of the PDE \eqref{eq:state_pde} as a function of $m$.
%
%That is, inverse problems in which one seeks to reconstruct an unknown parameter field, $m$, from noisy measurements,
%\begin{equation*}
%	y = B(m,u) + \xi,
%\end{equation*}
%of a state variable, $u$, which depends on $m$ indirectly through the solution of a PDE,
%\begin{equation}
%\label{eq:state_equation}
%	0=F(m,u).	
%\end{equation}
%Here $\xi$ is the noise. The observations $y$ are known, while $m$, $u$, and $\xi$ are unknown. In the deterministic setting, the inverse problem of finding $m$ is commonly framed as a least squares minimization problem
%\begin{equation}
%\label{eq:minimization_problem}
%	\min_m \quad J_d(m) + R(m),
%\end{equation}
%where
%\begin{equation*}
%	J_d(m) := \frac{1}{2}\|y - B(m,u(m))\|_{\Gamma}^2
%\end{equation*}
%is a data misfit term, $\Gamma$ is the noise covariance, $R(m)$ is a regularization term, and $u(m)$ denotes the solution $u$ of the PDE \eqref{eq:state_equation} as a function of $m$, . 
%
%Approximation of the Hessian,
%\begin{equation*}
%	\mathcal{H} = \mathcal{H}_d + \mathcal{H}_R := \frac{d^2 J_d}{dm^2} + \frac{d^2 R}{dm^2},
%\end{equation*}
In the deterministic approach to inverse problems, one typically finds $m$ as the solution to the minimization problem 
\begin{equation}
\label{eq:minimization_problem}
	\min_m \quad \frac{1}{2}\|y - f(m,u(m))\|_W^2 + R(m),
\end{equation}
where $\|\cdot\|_W$ is weighted norm which depends on the noise covariance, and $R(m)$ is a regularization term. Approximation of the Hessian of the objective function, $\mathcal{H}$, allows for fast solution of \eqref{eq:minimization_problem} via Newton-type methods. Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because $\mathcal{H}^{-1}$ locally approximates the Bayesian posterior covariance for $m$. 
We only have access to $\mathcal{H}$ via its action on an arbitrary vectors $v$, i.e., evaluation of the map $v \mapsto \mathcal{H} v$. Evaluating $v \mapsto \mathcal{H} v$ is an expensive process that involves solving two auxiliary PDEs.

%The regularization term in the Hessian, $\mathcal{H}_R := \frac{d^2 R}{dm^2}$, is typically an elliptic differential operator, and is therefore easy to work with. The data misfit term, $\mathcal{H}_d = \frac{d^2 J}{dm^2}$, is difficult to work with. The data misfit term $\mathcal{H}_d$ is not directly accessible because it depends on $u$, which depends on $m$ implicitly through the solution of the PDE \eqref{eq:state_equation}. 
%
%We only have access to $\mathcal{H}_d$ via its action on an arbitrary vectors $v$, i.e., evaluation of the map $v \mapsto \mathcal{H} v$. Evaluating $v \mapsto \mathcal{H} v$ is an expensive process that involves solving two auxiliary PDEs. We show how to evaluate $v \mapsto \mathcal{H} v$ in Appendix ADSF.

The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term. Either the Lanczos method or randomized SVD are used to perform the low rank approximation using only matrix-vector products. The larger the rank of the approximation, the more matrix-vector products are required. These methods suffer from a ``data predicament''---as more informative data is used in the inversion, the numerical rank of the data misfit term in the Hessian grows. The ideal scenario from a scientific perspective (highly informative data) is therefore the worst case scenario from a computational perspective (large computational cost). In our numerical results, we show that our operator approximation method can efficiently approximate the data misfit term in the Hessian in certain PDE constrained inverse problems. The method in this paper offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because the rank of the approximation is not limited by the number of matrix-vector products used to construct the approximation. High-rank approximations of an operator can be formed using a small number of matrix-vector products.

\subsection{Operator access levels}

%In order to solve inverse or optimization problems governed by PDEs, we need to solve linear systems with the Hessian as the coefficient matrix [cite cite]. In order to do this efficiently, we have access to only Hessian-vector products. That is, we have the ability to compute the functions $x \mapsto \mathcal{H} x$ and $x \mapsto \mathcal{H}^T x$ for arbitrary vectors $x$. Therefore

A linear operator may be classified by the way in which we may access the operator.

\begin{itemize}
	\item \emph{Level 1:} A matrix representation of the operator, $\mathbf{A}$, is stored in memory in dense or sparse format.
	\item \emph{Level 2:} We can efficiently compute arbitrary matrix entries $\mathbf{A}_{ij}$ at will, but the matrix $\mathbf{A}$ is not stored in memory.
	\item \emph{Level 3:} Access to $\mathbf{A}$ is only available through an algorithm that computes matrix-vector products $\mathbf{u} \mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{u} \mapsto \mathbf{A}^T \mathbf{u}$.
\end{itemize}
In large-scale scientific computing,

As discussed in Section \ref{sec:PDE_hessian_motivation}, in inverse problems governed by PDEs, matrix entries of the Hessian are not easily computable since getting one entry would require computing an entire column of the Hessian, then throwing away all but one entry in the column. Computing a single column costs at least two PDE solves. Many such entries are needed to approximate the Hessian in inverse problems with highly informative data (i.e., for Hessians with large numerical rank) [cite cite].

 of the implicit definition of $u(m)$.

Hierarchical matrices (H, H2, HSS, HBS, HODLR), quantized tensor trains, hierarchical interpolative factorization, STRUMPACK (LBNL),

Level 1: not relevant in large-scale problems
State clearly: we do not use Level 1 at all.
This paper: Level 3 -> Level 2
H-matrix: solve on level 2

%\subsection{Existing work}
%
%
%\begin{itemize}
%	\item CITE ESCANDE, others, for a reviews of the subject.
%	\item Resolution analysis seismic papers: try to figure out size of $\varphi_x$ by matvec with random function, post-processing to get autocorrelation length. They did the constant matvec volume thing. Their goal is to get a sense of local length scales that can be resolved in the inverse problem. They do not build preconditioner
%	\item Escande Weiss scattered response paper. They interpolate scattered impulse responses using radial basis functions and study how the geometry of the points affects the approximation accuracy. Matrix-free / getting many points at once is not a concern of them. The applications they consider one has direct access to the impulse response, e.g., through formula for the kernel. 
%	\item Imaging problems: ``matvec'' requires performing an experiment
%	\item pseudodifferential operator equivalence
%	\item Laundry list of applications
%	\item scattered impulse response of escande weiss paper: radial basis functions for approximation in $y$. How to determine if the impulse responses are far enough away from each other?
%	\item my convolution paper: compute impulse responses one at a time, one impulse response per matvec: too slow! Here we compute many impulse responses using only 1 matvec.
%	\item uniformity of smoothness condition
%\end{itemize}

\subsection{Background}

\subsubsection{Distributions}

Let $\overline{\Omega}$ be the closure of $\Omega$, and let $C(\overline{\Omega})$ be the space of continuous functions mapping $\overline{\Omega}\rightarrow \mathbb{R}$. The action of $\mathcal{A}$ is extended to distributions $\mu:C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ via the formula
\begin{equation*}
	\langle\mathcal{A},\mu\rangle(w) := \int_\Omega w(y) \mu\left(A(y, \cdot)\right) dy, 
\end{equation*}
where $A(y,\cdot)$ is the function $x \mapsto A(y,x)$. This is derived formally as follows:
\begin{align*}
	\langle\mathcal{A},\mu\rangle(w) &= \int_\Omega \int_\Omega w(y) A(y,x) \textrm{``} \mu(x) \textrm{''} dx dy \\
	&= \int_\Omega w(y) \int_\Omega A(y,x) \textrm{``} \mu(x) \textrm{''} dx \\
	&= \int_\Omega w(y) \mu\left(A(y,\cdot)\right) dx.
\end{align*}
For example, the delta distribution $\delta_x$ is defined by $\delta_x(v) = v(x)$, and the action of $\mathcal{A}$ on $\delta_x$ is given by
\begin{equation*}
	\langle\mathcal{A},\delta_x\rangle(w)	= \int_\Omega w(y) A(y,x) dy.
\end{equation*}

\subsubsection{Discretization}
\label{sec:discretization}

Let $\phi_i$, $i=1,\dots,N$ be a set of finite element basis functions, and let
\begin{equation*}
	V_h = \Span\left(\phi_1, \phi_2, \dots, \phi_N\right).
\end{equation*}
be the corresponding  finite element space approximating $L^2(\Omega)$. Here $V_h$ inherits the $L^2$ inner product, and $h$ denotes the mesh size parameter. Functions $u\in L^2(\Omega)$ are approximated by functions $u_h \in V_h$. In turn, functions $u_h\in V_h$ are represented in computations by length $N$ arrays $\mathbf{u}$, such that the array entries of $\mathbf{u}$ are the coordinates of $u_h$ in the finite element basis. That is,
\begin{equation*}
	u(x) \approx u_h(x) = \sum_{i=1}^N \mathbf{u}_i \phi_i(x).
\end{equation*}
We say that the coefficient arrays $\mathbf{u}$ reside $\mathbb{R}^N_\mathbf{M}$, which is the space $\mathbb{R}^N$ with the matrix-weighted inner product $\left(\mathbf{u},\mathbf{v}\right)_\mathbf{M} := \left(\mathbf{M} \mathbf{u}\right)^T \mathbf{v}$, where
$\mathbf{M} \in \mathbb{R}^{N \times N}$ is the \emph{mass matrix} which has entries
\begin{equation*}
	\mathbf{M}_{ij} = \left(\phi_i, \phi_j\right)_{L^2(\Omega)} = \int_{\Omega} \phi_i(x) \phi_j(x) dx.
\end{equation*}
Direct calculation shows that $V_h$ is isometrically isomorphic $\mathbb{R}^N_\mathbf{M}$, in the sense that functions $u_h \in V_h$ are in bijective correspondence to coefficient vectors $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, and $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u},\mathbf{v}\right)_\mathbf{M}$. Likewise, the dual space of continuous linear functionals $\psi_h:V_h \rightarrow \mathbb{R}$, denoted $V_h'$, is isometrically isomorphic to $\mathbb{R}^N_{\mathbf{M}^{-1}}$. All operations on functions in $V_h$ and linear functionals in $V_h'$ may therefore be performed on coefficient vectors in $\mathbb{R}^N_\mathbf{M}$ and coefficient dual vectors in $\mathbb{R}^N_{\mathbf{M}^{-1}}$. We summarize coefficient vector versions of the function space operation that are used in this paper in Appendix \ref{app:discretized_operations}.

\section{Sampling impulse responses}

We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points. The challenge is to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$.

The function $\varphi_x(y)$ is undefined when $x + y \notin \Omega$, which can occur even if $x \in \Omega$ and $y \in \Omega$. Extending $\varphi_x$ by zero makes it defined for all $y \in \mathbb{R}^d$, but this extension is only reasonable if $x$ is far enough from $\partial \Omega$ that $\varphi_x(y)$ decays to zero, as a function of $y$, before $y$ reaches $\partial \Omega$. We therefore choose sample points $x_i$ from a region $\Omega_I \subset \Omega$ which is not too close to the boundary. We determine $\Omega_I$ by applying $\mathcal{A}^T$ to a boundary source, then post processing the result.

Ensuring that the points $x_i$ are well separated from each other requires estimating the supports of the functions $\langle\mathcal{A},\delta_{x}\rangle^*$. Since $\varphi_x$, and thus $\langle\mathcal{A},\delta_{x}\rangle^*$, is nonnegative, $\langle\mathcal{A},\delta_{x}\rangle^*$ is a scaled probability distribution. We compute the scaling factors for all $x$ simultaneously by applying $\mathcal{A}$ to a constant function. We then compute the mean and covariance of the normalized version of $\langle\mathcal{A},\delta_{x}\rangle^*$, for all $x$ simultaneously, by applying $\mathcal{A}^T$ to a small number of linear and quadratic functions. The support of $\langle\mathcal{A},\delta_{x}\rangle^*$ is estimated to be contained in an ellipsoid of points that are within a small number of standard deviations of the mean for a Gaussian distribution with the same mean and covariance as the normalized version of $\langle\mathcal{A},\delta_{x}\rangle^*$.

\subsection{Determination of the boundary region}
\label{sec:boundary}

Define $\Omega_B \subset \Omega$ to be the set of points that are too close to $\partial \Omega$, in the sense that $x \in \Omega_B$ if $\varphi_x(y-x) \neq 0$ for some $y \in \partial \Omega$. Further, define $\Omega_I := \Omega \setminus \Omega_B$ as the set of points that are in the interior of $\Omega$ and are not too close to $\partial B$. In Proposition \ref{prop:boundary_source} we show that $\Omega_B$ can be determined by applying $\mathcal{A}^T$ to a boundary source distribution and post-processing the result. The resulting algorithm for determining $\Omega_B$ and $\Omega_I$ in the discretized setting is shown in Algorithm \ref{alg:boundary_region}.

\begin{prop}
	\label{prop:boundary_source}
	Let $\delta_{\partial \Omega}: C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ be the boundary source distribution defined by
	\begin{equation}
	\label{eq:boundary_source}
	\delta_{\partial \Omega}(v) = \int_{\partial \Omega} v(y) dy
	\end{equation}
	for all $v \in C(\overline{\Omega})$. We have $\varphi_x(y-x) = 0$ for almost every $y \in \partial \Omega$ if and only if $\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle^*(x) = 0$.
\end{prop}

The proof of Proposition \ref{prop:boundary_source} is shown in Appendix \ref{app:proofs}.

%Straightforward analysis shows that $\varphi_x(y-x) = 0$ for $y \in \partial \Omega$ almost everywhere if and only if $\left(\mathcal{A}^T \delta_{\partial \Omega}\right)(x) = 0$, where $\delta_{\partial \Omega}$ is the uniform boundary source distribution. That is, the distribution that acts on a test function $v$ by performing a surface integral of $v$ over the boundary of the domain:
%%\begin{equation}
%%\label{eq:boundary_source}
%%	\delta_{\partial \Omega}(v) = \int_{\partial \Omega} v(y) dy.
%%\end{equation}
%We therefore compute $b  = \mathcal{A}^T \delta_{\partial \Omega}$, then identify $\Omega_B$ as the set of points $x \in \Omega$ for which $|b(x)|$ is larger than a small tolerance $\epsilon$. This is described in Algorithm \ref{alg:boundary_region}.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	{

		Construct boundary source $\boldsymbol{\delta}_{\partial \Omega} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$, $\displaystyle \left(\boldsymbol{\delta}_{\partial \Omega}\right)_i = \int_{\partial \Omega} \phi_i(s) ds$, $i=1,\dots,N$

%		\For{$i=1,\dots,N$}{
%			$\displaystyle \mathbf{b}_i \gets \int_{\partial \Omega} \phi_i(s) ds$
%		}

		$\mathbf{b} = \mathbf{M}^{-1}\mathbf{A}^T\mathbf{M}^{-1}\boldsymbol{\delta}_{\partial \Omega}$

		$\Omega_B = \{x \in \Omega : |b_h(x)| \ge \epsilon\}$

		$\Omega_I = \Omega \setminus \Omega_B$

	}
	\caption{Compute boundary region $\Omega_B$ and interior region $\Omega_I$}
	\label{alg:boundary_region}
\end{algorithm2e}

One should only expect $\widetilde{A}(y,x)$ to be accurate for $x \in \Omega_I$, $y \in \Omega$. If $\mathcal{A}$ is symmetric then one may overwrite potentially innacurate entries $\widetilde{A}(y,x)$, $x \in \Omega_b$, $y \in \Omega_I$, with the accurate entries $\widetilde{A}(x,y)$. If $\mathcal{A}$ is nonsymmetric, one may form an accurate approximation of these entries by applying our procedure to $\mathcal{A}^T$. Our method will furnish an approximation $\widetilde{A}(y,x)$ for $x \in \Omega_B$, $y \in \Omega_B$ (the block of $\widetilde{H}$ corresponding to boundary-boundary interactions), but this part of the approximation is unlikely to satisfy Condition 1 ($\varphi_x$ varies smoothly as a function of $x$).

\subsection{Mean and covariance estimation}

Let 
\begin{equation*}
	\rho_x := \langle\mathcal{A}, \delta_x\rangle^* \big/ \alpha(x)
\end{equation*}
denote the normalized version of $\langle\mathcal{A}, \delta_x\rangle^*$, where $\alpha:\Omega \rightarrow \mathbb{R}$ is the volume function
\begin{equation*}
	\alpha(x) := \int_{\Omega} \langle\mathcal{A}, \delta_x\rangle^*(z) dz.
\end{equation*}
Note that $\rho_x$ is a probability distribution on $\Omega$.

In Proposition \ref{thm:vol_mean_cov}, we see that $\alpha(x)$ may be computed for all $x$ simultaneously by applying $\mathcal{A}^T$ to a constant function. We also see that the spatially varying mean, $\mu(x)$, and covariance, $\Sigma(x)$, of $\rho_x$, defined as
\begin{align*}
	\mu(x) :=& \int_\Omega z \rho_x(z) dz \\
	\Sigma(x) :=& \int_\Omega (z - \mu(x))(z - \mu(x))^T \rho_x(z) dz
\end{align*}
may be computed for all $x$ simultaneously by applying $\mathcal{A}^T$ to $d$ linear functions, and $d(d+1)/2$ quadratic functions, respectively. In Algorithm \ref{alg:varhpi_mean_cov} we describe an algorithm for computing the volume function, mean, and covariance based on the results of Proposition \ref{thm:vol_mean_cov}.

\begin{prop}
	\label{thm:vol_mean_cov}
	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$, and let $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions, respectively:
	\begin{equation*}
		C(x) := 1, \qquad
		L^i(x) := x^i, \qquad
		Q^{ij}(x) = x^i x^j.
	\end{equation*}
	We have
	\begin{align}
		\alpha =& \left(\mathcal{A}^TC\right)^*, \label{eq:vol_mean_var_thm1}\\
		\mu^i =& \left(\mathcal{A}^T L^i\right)^* / \alpha, \label{eq:vol_mean_var_thm2}\\
		\Sigma^{ij} =& \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot\mu^j, \label{eq:vol_mean_var_thm3}
	\end{align}
	We write $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
\end{prop}

The proof of Theorem \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
%	\Input{Function $v \mapsto \mathcal{H}^Tv$}

%	\mbox{}\\

%	\Output{Spatially varying mean $\mu:\Omega \rightarrow \mathbb{R}^d$, \\ 
%		Spatially varying covariance $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$}
%	\Begin{
	{	
		Form constant function $C(x)=1$

		$\alpha = \left(\mathcal{A}^T C\right)^*$

		\For{$i=1,2,\dots,d$}{
			Form linear function $L^i(x) = x^i$

			$\mu^i = \left(\mathcal{A}^T L^i\right)^* / \alpha$
		}
		\For{$i=1,2,\dots,d$}{
			\For{$j=1,\dots,i$}{
				Form quadratic function $Q^{ij} = x^i x^j$

				$\Sigma^{ij} = \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot \mu^j$

				$\Sigma^{ij} = \Sigma^{ji}$

			}
		}

%		\Return{$\mu$, $\Sigma$}
	}
	\caption{Compute volume $\alpha_h$, mean $\mu_h$, and covariance $\Sigma_h$}
	\label{alg:varhpi_mean_cov}
\end{algorithm2e}

\paragraph{Discretization} In computations, $\alpha$, $\mu^{i}$, and $\Sigma^{ij}$ are replaced with finite element approximations $\alpha_h, \mu_h^i, \Sigma_h^{ij}\in V_h$, respectively, which have coefficient vectors $\boldsymbol{\alpha}, \boldsymbol{\mu}^{ij}, \boldsymbol{\Sigma}^{ij}\in\mathbb{R}^N_\mathbf{M}$, respectively. The functions $C$, $L^i$, and $Q^{ij}$ are replaced with their $L^2$ projections onto $V_h$, which are denoted by $C_h$, $L_h^i$, and $Q_h^{ij}$, respectively. The corresponding coefficient vectors are $\mathbf{C}$, $\mathbf{L}^i$, and $\mathbf{Q}^{ij}$, respectively. The discretized versions of \eqref{eq:vol_mean_var_thm1}, \eqref{eq:vol_mean_var_thm2}, and \eqref{eq:vol_mean_var_thm3} are
\begin{align*}
	\boldsymbol{\alpha} &= \mathbf{M}^{-1}\mathbf{A}^T\mathbf{C} \\
	\boldsymbol{\mu}^i &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{L}^i\right) / \boldsymbol{\alpha}\\
	\boldsymbol{\Sigma}^{ij} &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{Q}^{ij}\right) / \boldsymbol{\alpha} - \boldsymbol{\mu}^i\cdot \boldsymbol{\mu}^j,
\end{align*}
respectively. Here $\mathbf{f} \cdot \mathbf{g}$ and $\mathbf{f} / \mathbf{g}$ denote the element-wise multiplication and division of vectors, respectively.

\subsection{Choosing sample points}
\label{sec:choosing_sample_points}

Ideally, sample points $x_i$ and $x_j$ should be far enough apart from each other that the supports of $\left\langle\mathcal{A}, \delta_{x_i}\right\rangle^*$ and $\left\langle\mathcal{A}, \delta_{x_j}\right\rangle^*$ do not overlap. Since it is too expensive compute the support of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ directly, we instead estimate that the support of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ is contained in the ellipsoid
\begin{equation*}
	E_x := \{x' \in \Omega: (x' - \mu(x))^T \Sigma(x)^{-1} (x' - \mu(x)) \le \tau^2\},
\end{equation*}
where $\tau$ is a small number of standard deviations; we use $\tau=3$. The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations of the mean for the Gaussian distribution that has the same mean and covariance as $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$. The fraction of the mass of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though we find this bound is conservative and far less mass resides in this region in practice. 
%We call $E_x$ the \emph{unit ellipsoid} with mean $\mu(x)$ and covariance $\Sigma(x)/\tau^2$.

We select a batch of sample points, $S_b$, randomly from a finite set of candidate points $P \subset \Omega_I$. Starting from an empty set, we build $S_b$ by choosing candidate points $p$ randomly from $P$, one at a time. If $p$ is sufficiently far from all of the previously chosen points, in the sense that $E_p \cap E_q = \{\}$ for all $q \in S_b$, we add $p$ to $S_b$. Otherwise we discard $p$. This process repeats until it is no longer possible to pick a point from $P$ that is sufficiently far from the points in $S_b$. 
%This is detailed in Algorithm \ref{alg:point_choice}.

We repeat the process to construct several batches of points $S_1, S_2, \dots$, until the total number of sample points exceeds a desired threshold. We write $S = S_1 \cup \dots$ to denote the set of sample points from all batches. In our implementation, the set of candidate points $P$ is the set of all Lagrange nodes for the finite element basis functions $\phi_i$ used to discretize the problem, minus all of the Lagrange nodes that are in $\Omega_B$, minus all points in previously chosen batches.

%\begin{algorithm2e}
%	\SetAlgoNoLine
%	\SetKwInOut{Input}{Input}
%	\SetKwProg{Fn}{Function}{}{}
%
%	Start with a finite set of candidate points $P \subset \Omega_I$
%	
%	$S_b \gets \{\}$
%	
%	\While{$P$ is not empty}{
%		Pick a random point $p \in P$ and remove $p$ from $P$
%		
%		$p\text{\_is\_acceptable} \gets \text{True}$
%		
%		\For{$q \in S_b$}{
%			
%			\If{$E_p \cap E_q \neq \{\}$}{
%				
%				$p\text{\_is\_acceptable} \gets \text{False}$
%				
%				Terminate for loop
%			}
%		}
%	
%		\If{$p\text{\_is\_acceptable}$}{
%			
%			$S_b \gets S_b \cup \{p\}$
%			
%			$P \gets P \setminus \{q \in P : \mu(q) \in E_p\}$
%		}
%	}
%
%	\caption{Choosing one batch of sample points, $S_b$}
%	\label{alg:point_choice}
%\end{algorithm2e}

\subsection{Extracting impulse responses at sample points}
\label{sec:get_impulse_response}

The \emph{Dirac comb}, $\xi_b$, associated with a batch of sample points $S_b$ is the sum of Dirac distributions (point sources) centered at the points $x_i \in S_b$. That is,
\begin{equation*}
	\xi^b := \sum_{x \in S_b} \delta_{x_i}.
\end{equation*}
For each batch $S_b$, we compute as the action of $\mathcal{A}$ on the associated Dirac comb:
\begin{equation}
\label{eq:dirac_comb_H_action}
	\eta^b := \langle\mathcal{A}, \xi^b\rangle^*.
\end{equation}
By linearity, $\eta^b$ may be written as
\begin{equation}
\label{eq:phi_b}
	\eta^b = \left\langle\mathcal{A},\sum_{x_i \in S_b} \delta_{x_i}\right\rangle^* = \sum_{x_i \in S_b} \langle\mathcal{A}, \delta_{x_i}\rangle^*.
\end{equation}
Since the points $x_i$ are chosen so that the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ and the support of $\langle\mathcal{A},\delta_{x_j}\rangle^*$ do not overlap (or do not overlap much), we have
\begin{equation}
\label{eq:varphi_eval}
	\varphi_{x_i}(z) = \langle\mathcal{A}, \delta_{x_i}\rangle^*(z+x_i) = \begin{cases}
		\eta^b(z+x_i), & z+x_i \in E_{x_i} \\
		0, & \text{otherwise}.
		\end{cases}
\end{equation}
for all $x_i \in S_b$. By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover $\varphi_{x_i}$ for every point $x_i \in S_b$. 
%For $x_i \in X$ and $y \in \mathbb{R}^d$, we may evaluate $\varphi_{x_i}(y)$ by finding the batch $X_b$ that $x_i$ is contained in, then using \eqref{eq:varphi_eval}.

%\begin{algorithm2e}
%	\SetAlgoNoLine
%	\SetKwInOut{Input}{Input}
%	\SetKwInOut{Output}{Output}
%	\SetKwProg{Fn}{Function}{}{}
%	\Input{Point source locations $x_1, x_2, \dots, x_m$
%	}
%	\Output{Response $\boldsymbol{\eta}$ of $\mathbf{A}$ to Dirac comb of point sources}
%
%	\BlankLine
%
%	Construct $\boldsymbol{\xi} \in \mathbb{R}^N_{\mathbf{M}^{-1}}, \quad \boldsymbol{\xi}_i = \sum_{j=1}^m \phi_i(x_j)$, \quad $i=1,\dots,N$
%	
%	$\boldsymbol{\eta} \gets \mathbf{M}^{-1} \mathbf{A}^T \mathbf{M}^{-1}\boldsymbol{\xi}$
%
%	\caption{Computing response $\mathbf{A}$ to Dirac comb}
%	\label{alg:dirac_comb_response}
%\end{algorithm2e}

\paragraph{Discretization}
In computations, the Dirac comb $\xi^b$ is replaced by its restriction to $V_h$, denoted $\xi^b \in V_h'$. The response $\eta^b$ is replaced by the finite element approximation $\eta_h^b \in V_h$. The coefficient dual vector for $\xi_h^b$ is the vector $\boldsymbol{\xi}^b \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has entries
\begin{equation*}
	\boldsymbol{\xi}^b_i = \sum_{x_j \in S_b} \phi_i(x_j), \quad i=1,\dots,N.
\end{equation*}
The coefficient vector $\boldsymbol{\eta}^b \in \mathbb{R}^N_\mathbf{M}$ for $\eta_h^b$ is given by
\begin{equation*}
	\boldsymbol{\eta}^b = \mathbf{M}^{-1} \mathbf{A}^T \mathbf{M}^{-1}\boldsymbol{\xi}^b.
\end{equation*}

\section{Interpolation of impulse response samples}

We interpolate the functions $\varphi_{x_i}$ associated with the points $x_i \in S$, to generate functions $\varphi_x$ associated with new points $x \notin S$. This is done using a continuous generalization of the interpolative decomposition for matrices, wherein a select number of columns of a matrix are interpolated to fill the whole matrix. Interpolating $\varphi_x$ differs from interpolative low rank approximation of $\mathcal{A}$, which corresponds to interpolating $\langle\mathcal{A},\delta_x\rangle^*$. Because of the shifting that maps $\langle\mathcal{A}, \delta_x\rangle$ to $\varphi_x$, interpolation of $\varphi_{x_i}$ yields an approximation $\widetilde{\mathcal{A}}$ which can be high-rank.

The interpolation procedure requires a basis of smooth functions. We form this basis from smooth eigenmodes of the Laplacian on $\Omega$ with Neumann boundary conditions. The resulting approximation $\mathcal{A}$ may be interpreted as a weighted sum of convolution operators, with smooth spatially varying weighting functions. Kernel entries of the approximation, $\widetilde{A}(y,x)$, may be evaluated efficiently.

\subsection{Interpolative factorization of the SVIR}
\label{sec:interp_fact_svir}

The \emph{spatially varying impulse response}, denoted $T$, is the reinterpretation of $\varphi_x(y)$ as a function of both $y$ and $x$:
\begin{equation*}
	T(y,x) := \varphi_x(y) = A(y+x,x).
\end{equation*}
Since $T(y,x)$ is smooth as a function of $x$ (Condition 1), the continuous analog of the row space of $T$ should be well-approximated by the span of a small number of orthonormal smooth functions, $\{u_k\}_{k=1}^r$. Condition 1, therefore, is equivalent to the existence of a low rank approximation $\widetilde{T}$ of $T$ of the form 
\begin{equation}
	\label{eq:formula_for_A}
	T \approx \widetilde{T} = \Psi U
\end{equation}
where $\Psi:\mathbb{R}^d \times \{1,\dots,r\} \rightarrow \mathbb{R}$ is an unknown semi-discrete kernel with $r$ columns, and $U:\{1,\dots,r\} \times \Omega \rightarrow \mathbb{R}$ is the semi-discrete kernel with $r$ rows, such that the $i^\text{th}$ row of $U$ is the function $u_i$. That is,
\begin{equation*}
	U(i,x) = u_i(x).
\end{equation*}
The multiplication $K_1 K_2$ of kernels $K_1:X \times Y \rightarrow \mathbb{R}$ and $K_2:Y \times Z \rightarrow \mathbb{R}$ is defined as
\begin{equation*}
	\left(K_1 K_2\right)(x,z) := \sum_{y \in Y} K_1(x,y) K_2(y,z)
\end{equation*}
if $Y$ is a discrete set, and
\begin{equation*}
	\left(K_1 K_2\right)(x,z) := \int_Y K_1(x,y) K_2(y,z) dy	
\end{equation*}
if $Y\subset \mathbb{R}^d$ is a domain.

The number of smooth functions, $r$, required to well-approximate $T$ using \eqref{eq:formula_for_A} can be small even if the numerical rank of $\mathcal{A}$ is large. For example, let $\varphi$ be any convolution kernel, and suppose $\mathcal{A}$ is the convolution operator $\left(\mathcal{A}v\right)^* := \varphi \ast v$. The rank of $\mathcal{A}$ can be made arbitrarily large or small by changing the convolution kernel $\varphi$. At the same time, $T=\Psi U$ holds with $r=1$, where $\Psi(x,1):= \varphi(x)$ and $U(1,x):=1$.

We may determine $\Psi$ using a continuous analog of the process used to form the interpolative matrix factorization Let $S:=\{x_j\}_{j=1}^q \subset \Omega$ be the set of $q \ge r$ sample points determined by the method in Section \ref{sec:choosing_sample_points}, and define $T_S:\Omega \times \{1,\dots,q\}\rightarrow\mathbb{R}$ to be the semi-discrete kernel with $i^\text{th}$ column $\varphi_{x_i}$. That is,
\begin{equation*}
	T_S(y,i):= \varphi_{x_i}(y) = T(y,x_i).	
\end{equation*}
Further, let $U_S$ be the $r \times q$ matrix with $(i,j)$ entry given by $u_i(x_j)$, and assume that the points $x_i$ are chosen such that $U_S$ has full row rank. This is not difficult; for example, if the points $x_i$ were chosen uniformly at random then $U_S$ would have full rank almost surely. Indeed, with uniform random points, $\frac{|\Omega|}{q}\left(U_S U_S^T\right)_{ij}$ is a Monte-Carlo approximation of $\langle u_i, u_j \rangle_{L^2(\Omega)}$, where $|\Omega|$ is the total volume of the domain $\Omega$. Thus $\frac{|\Omega|}{q}\left(U_S U_S^T\right) \rightarrow I$ as $q \rightarrow \infty$, where $I$ is the identity matrix. This implies $\sigma_\text{min} \rightarrow \sqrt{q/|\Omega|}$ as $q \rightarrow \infty$, where $\sigma_\text{min}$ is the minimum singular value of $U_S$. We do not use uniform random sampling to choose the points, but the way we choose points does achieve relatively uniform coverage of the domain.

Restricting the second input of both sides of the equation \eqref{eq:formula_for_A} to points in $S$, we have
\begin{equation}
\label{eq:A_I_equals_X_U_I}
\widetilde{T}_S = \Psi U_S.
\end{equation}
Multiplying both sides of \cref{eq:A_I_equals_X_U_I} by the Moore-Penrose pseudoinverse of $U_S$, denoted $U_S^+$, and using the fact that $U_S$ has full row rank, yields
\begin{equation}
\label{eq:formula_for_X}
\widetilde{T}_S U_S^+ = \Psi.
\end{equation}
Substituting the left hand side of \eqref{eq:formula_for_X} for $\Psi$ into \eqref{eq:formula_for_A} yields the following interpolative-type low rank factorization $\widetilde{T}$ of $T$:
\begin{equation}
\label{eq:interpolative_decomposition0}
T \approx \widetilde{T} = T_S U_S^+ U.
\end{equation}
%We will quantify the accuracy of this approximation in Theorem \ref{thm:smooth_basis_err} in Section \ref{sec:error_estimate_adaptive_algorithm}.

%Briefly, the error will be small when (1) $\mathcal{U}$ well-approximates the row space of $\mathcal{T}$, and (2) the smallest singular value of $\mathcal{U}_X$ is not small.

\subsection{Smooth basis of Laplacian eigenmodes}

For the basis of smooth functions $u_i$, we use the dominant eigenfunctions of the inverse Laplacian on $\Omega$ with Neumann boundary conditions. This basis of inverse Laplacian eigenfunctions generalizes the Fourier basis to non-rectangular domains. Let $\Delta_N^{-1}$ denote the solution operator for the following PDE:
\begin{equation}
\label{eq:laplacian_pde}
\begin{cases}
-\Delta u =f, & \text{in } \Omega, \\
n \cdot \nabla u = 0, & \text{on } \partial \Omega, \\
\int_\Omega u dx = 0, &
\end{cases}	
\end{equation}
where $\Delta$ is the Laplacian and $n$ is the normal to $\partial \Omega$. That is, $\Delta_N^{-1}$ is the map from source $f$ to PDE solution $u$. The first smooth basis function used for interpolation is the constant function $u_1(x):=1$. The remaining smooth basis functions, $u_2, u_3, \dots, u_r$, are the eigenfunctions corresponding to the $r-1$ largest eigenvalues of $\Delta_N^{-1}$. These eigenfunctions may be computed efficiently using randomized SVD \cite{HMT11}, Krylov methods such as Lanczos, power iterations with deflation, or other methods. We use randomized SVD. All of these methods require the ability to apply $\Delta_N^{-1}$ to given vectors, which amounts to solving PDE \eqref{eq:laplacian_pde} with the given vector as the right hand side source. Highly efficient multigrid schemes exist for solving this PDE.
%; we use an algebraic multigrid implementation in PyAMG.

\subsection{Action of the approximate operator}

Defining
\begin{equation*}
	W := U_S^+ U,
\end{equation*}
we may write $\widetilde{T}$ as
\begin{equation*}
	\widetilde{T} = T_S W.	
\end{equation*}
Using the definition of $T_S$, we may write the entries of the resulting approximation to $A$ as
\begin{equation}
\label{eq:H_kernel_entries}
	\widetilde{A}(y,x) = \widetilde{T}(y-x,x) = \sum_{i=1}^q w_i(x) \varphi_{x_i}(y-x) 
\end{equation}
where the functions $w_i(x):=W(i,x)$ are the rows of $W$, which are linear combinations of the functions $u_i$.
%Since the Laplacian eigenfunctions $u_i$ are defined on $\Omega$, the weighting functions $w_i$ are defined on all of $\Omega$, so $T$ and $H$ are defined on $\mathbb{R}^d \times \Omega$. However, the error estimate we present in the next section will only be valid for these operators on $\mathbb{R}^d \times \Omega_I$. 
From \eqref{eq:H_kernel_entries}, we derive the action of $\widetilde{\mathcal{A}}$ as
\begin{align*}
	\left(\widetilde{\mathcal{A}}u\right)(v) &= \int_\Omega \int_\Omega v(y) \widetilde{A}(y,x)u(x) dx dy \\
	&= \int_\Omega \int_\Omega v(y) \widetilde{T}(y-x,x)u(x) dx dy \\
	&= \int_\Omega \int_\Omega v(y) \left(\sum_{i=1}^q  \varphi_{x_i}(y-x) w_i(x)\right) u(x) dx dy \\
	&= \int_\Omega v(y) \sum_{i=1}^q \int_\Omega \varphi_{x_i}(y-x) w_i(x) u(x) dx dy\\
	&= \int_\Omega v(y) \left(\sum_{i=1}^q \varphi_{x_i} \ast \left(w_i \cdot u\right)\right)(y) dy,
\end{align*}
which implies
\begin{equation*}
	\left(\widetilde{\mathcal{A}} u	\right)^* = \sum_{i=1}^q \varphi_{x_i} \ast \left(w_i \cdot u\right)
\end{equation*}
Here $f \ast g$ denotes the convolution of the functions $f$ and $g$. Thus, $\widetilde{\mathcal{A}}$ is a weighted sum of convolution operators with smooth spatially varying weighting functions.

\subsection{Efficient computation of approximation kernel entries}

We now describe how to efficiently evaluate $\widetilde{A}(x',y')$ at arbitrary points $(x',y')$. Let $b(i)$ denote the batch index such that $x_i \in S^{b(i)}$, and define
\begin{equation*}
	\zeta_i := y' - x' + x_i,
\end{equation*}
for $x_i \in S$. From \eqref{eq:H_kernel_entries}, we have
\begin{equation*}
	\widetilde{A}(y',x') = \sum_{i=1}^q w_i(x') \varphi_{x_i}(y'-x'),
\end{equation*}
and from \eqref{eq:varphi_eval}, we have
\begin{equation}
\label{eq:kernel_entry_formua}
	\varphi_{x_i}(y'-x') = \begin{cases}
		\eta^{b(i)}(\zeta_i), & \zeta_i \in E_{x_i} \\
		0, & \text{otherwise}.
	\end{cases}
\end{equation}
%
%\begin{align*}
%	\widetilde{A}(y',x') &= \sum_{i=1}^q w_i(x') \varphi_{x_i}(y'-x') \\
%	&= \sum_{i=1}^q w_i(x') \begin{cases}
%		\eta^{b(i)}(\zeta_i), & \zeta_i \in E_{x_i} \\
%		0, & \text{otherwise}
%	\end{cases} \\
%	&= \sum_{i=1}^q w_i(x')\Psi_{ik},
%\end{align*}
%where $\omega_{ik}:= w_i(x_k') = \left(\mathcal{U}_X^+ \mathcal{U}\right)_{ik}$, and 
%\begin{equation*}
%	\Psi_{ik} := \begin{cases}
%		\eta^{b(i)}(\zeta_{ik}), & \zeta_{ik} \in E_{x_i}\\
%		0, & \text{otherwise}.
%	\end{cases}
%\end{equation*}
%To compute $\omega$, first we construct $\nu \in \mathbb{R}^{r \times N}$,	$\nu_{jk} := u_j(x_k)$,
%then we set
%\begin{equation*}
%	\omega := \mathcal{U}_X^+ \nu.
%\end{equation*}
%To compute $\Psi$, we must evaluate $\eta^{b(i)}(\zeta_{ik})$ for many points $\zeta_{ik}$.
%If the mesh used to discretize the problem is non-uniform, generally $\zeta_{ik}$ will not be a degree of freedom location, even if the points $x_k'$, $y_k'$, and $x_i$ are degree of freedom locations. Thus, evaluating $\eta^{b(i)}(\zeta_{ik})$ requires interpolation. Since it is typically more efficient to perform interpolation at many points all at once, rather than interpolating points one at a time, we precompute the set of indices
%\begin{equation*}
%	\chi^b := \{(i,k): \zeta_{ik} \in E_{x_i},~x_i \in X_b,~ k=1,\dots,N\}
%\end{equation*}
%for which the interpolation $\eta^{b(i)}(\zeta_{ik})$ must be performed. Then we compute the batch of interpolations $\eta^b\left(\zeta_{\chi^b}\right)$, and set 
%\begin{equation*}
%	\Psi_{\chi^b} = \eta^b\left(\zeta_{\chi^b}\right).
%\end{equation*}
%Here $\Psi_{\chi^b}$ is the subset of entries of $\Psi$ corresponding to indices $(i,k) \in \chi^b$, and $\eta^b\left(\zeta_{\chi^b}\right)$ is the interpolation of $\eta^b$ at all of the points $\zeta_{ik}$ corresponding to indices $(i,k) \in \chi^b$. This process is repeated for each batch $b$. All other entries of $\Psi$ are set to zero. The resulting procedure for evaluating $\widetilde{H}(y_k',x_k')$ is summarized in Algorithm \ref{alg:H_entries}.

%, for each batch $X_b$ we predetermine the subset of points 
%\begin{equation*}
%	p_{ik} := y_k - x_k + x_i
%	\end{equation*}
%satisfying $p_{bik} \in E_{x_i}$ for all $x_i \in X_b$ then evaluate $\eta_b(p_{bik})$ for this subsets of points all at once. Let
%\begin{equation*}
%	\chi_b:=\{p_{bik}:p_{bik} \in E_{x_i}\}.
%\end{equation*}
%and define $\Psi \in \mathbb{R}^{q \times N}$, 
%
%so that
%\begin{equation*}
%	\widetilde{H}(y_k,x_k) = \sum_{i=1}^q w_i(x_k) \Psi_{ik}.	
%\end{equation*}
%We have 
%\begin{equation*}
%	\Psi_{ik}(\chi_b) = \sum_{b=1}^B \eta_b(p_{bik})
%\end{equation*}
%where $\eta_b(\chi)$ is the 
%
%In Algorithm asdf, we describe how to evaluate $\widetilde{H}$ at many points using this procedure.
%
%\begin{algorithm2e}
%	\SetAlgoNoLine
%	\SetKwInOut{Input}{Input}
%	\SetKwInOut{Output}{Output}
%	\SetKwProg{Fn}{Function}{}{}
%	\Input{Points $\{x_k'\}_{k=1}^n \subset \Omega$}	
%	\Output{Weighting function values $\boldsymbol{\omega} \in \mathbb{R}^{q \times n}$, $\boldsymbol{\omega}_{i,k} = w_i(x_k')$}
%	
%	\mbox{}\\
%	
%	\Fn{$\computeweightingentries{\{x'_k\}_{k=1}^n}$}{
%		$\boldsymbol{\nu} \gets \mathbf{0}_{q\times n}$
%		
%		\For{$i=1,\dots,q$}{
%			$\boldsymbol{\nu}[i,:] \gets \interpolate(\mathbf{u}_i, \{x_1',x_2',\dots,x_k'\})$
%		}
%		
%		$\boldsymbol{\omega} \gets \mathbf{U}[:,\mathtt{S}]^+ \boldsymbol{\nu}$
%		
%		\Return{$\boldsymbol{\omega}$}
%	}
%	
%	\caption{Compute many weighting function values $\omega_{ik} = w_i(x_k')$}
%	\label{alg:weighting_entries}
%\end{algorithm2e}
%
%
%\begin{algorithm2e}
%	\SetAlgoNoLine
%	\SetKwInOut{Input}{Input}
%	\SetKwInOut{Output}{Output}
%	\SetKwProg{Fn}{Function}{}{}
%	\Input{Pairs of points $\{(x_k',y_k')\}_{k=1}^n \subset \Omega \times \Omega$}	
%	\Output{Approximate kernel entries $\mathbf{z} \in \mathbb{R}^n$, $\mathbf{z}_k = \widetilde{H}(y_k',x_k')$}
%	
%	\mbox{}\\
%	
%	\Fn{$\computekernelentries{\{x'_k\}_{k=1}^n}{\{y'_k\}_{k=1}^n}$}
%	{
%		$\boldsymbol{\omega} \gets \computeweightingentries{\{x'_k\}_{k=1}^n}$
%		
%		Construct $\boldsymbol{\zeta} \in \mathbb{R}^{d \times q \times n}$, ~$\boldsymbol{\zeta}[:,i,k] = y_k' - x_k' + \mathbf{X}\left[:,\mathtt{S}[i]\right]$
%
%		\BlankLine
%		
%		\BlankLine
%		
%		$\boldsymbol{\Psi}\gets \mathbf{0}_{q \times n}$
%		
%		\For{$b=1,\dots,N_B$}{
%			$\mathtt{\chi}_b \gets \{\}$
%			
%			\For{$j \in \mathtt{S}_b$}{
%				$i \gets$ index such that $\mathtt{S}[i] = j$
%				
%				$E_j \gets$ unit ellipsoid with mean $\boldsymbol{\mu}[:,j]$ and covariance $\boldsymbol{\Sigma}[:,:,j]/\tau^2$
%				
%				$\mathtt{\chi}_b \gets \mathtt{\chi}_b \cup \{(i,k) : k \in \{1,\dots,N\}, \boldsymbol{\zeta}[:,i,k] \in E_j\}$
%				
%			}
%
%			
%			$\boldsymbol{\Psi}[\chi_b] \gets \interpolate\left(\boldsymbol{\eta}^b, \{\boldsymbol{\zeta}[:,i,k] : (i,k) \in \mathtt{\chi}^b\}\right)$
%		
%		}
%	
%		\BlankLine
%		
%		\BlankLine
%		
%		Construct $\mathbf{z} \in \mathbb{R}^n$, ~$\mathbf{z}_k = \sum_{i=1}^q \boldsymbol{\omega}_{ik} \boldsymbol{\Psi}_{ik}$
%	
%		\Return{$\mathbf{z}$}
%	}
%	\caption{Compute many approximate kernel entries $\widetilde{H}(y_k',x_k')$}
%	\label{alg:H_entries}
%\end{algorithm2e}

\section{Adaptive algorithm}
\label{sec:error_estimate_adaptive_algorithm}

In this section we bound the error in the approximation to $T$ (Theorem \ref{thm:smooth_basis_err}), and use this bound to drive an adaptive algorithm (Algorithm \ref{alg:adaptive}) for constructing the approximation.

\begin{thm}
	\label{thm:smooth_basis_err}
	Let $\sigma_\text{min}$ be the smallest singular value of $U_S$, and let
	\begin{equation*}
		\gamma(q) := \sigma_\text{min} \sqrt{\frac{q}{|\Omega|}},
	\end{equation*}
	where $|\Omega|$ is the measure of $\Omega$. We have
	\begin{equation*}
		\|T - T_S U_S^+ U\|_2 \le \left(1 + \frac{|\Omega|}{\gamma(q)} \right)\|T - TU^TU\|_2,
	\end{equation*}
	where $\sigma_\text{min}$ is the smallest singular value of $U_S$, and $\nor{\mathcal{A}}$ is the induced norm for the operator $\mathcal{A}$.
\end{thm}

\begin{proof}
	Let $\widehat{T} := TU^TU$, and let $\widehat{T}_S:\mathbb{R}^d \times \{1,\dots,q\} \rightarrow \mathbb{R}$ be the semi-discrete kernel defined by
	\begin{equation*}
		\widehat{T}_S(y,i) = \widehat{T}_S(y,x_i).
	\end{equation*}
	Using the triangle inequality yields
	\begin{equation}
		\nor{T - T_S U_S^+ U}_2 \le \nor{T - \widehat{T}_S U_S^+ U}_2 
		+ \nor{\left(T_S - \widehat{T}_S\right) U_S^+ U}_2 \label{eq:two_bound_align}
	\end{equation}
	For the first term, note that $\widehat{T}_S = T U^T U_S$, so 
	\begin{equation*}
		\widehat{T}_S U_S^+ U = T U^T U_S U_S^+ U = T U^T U = \widehat{T}.
	\end{equation*}
	Thus,
	\begin{equation}
	\label{eq:A_nogamma}
	\nor{T - \widehat{T}_S U_S^+ U }_2 = \nor{T - \widehat{T}}_2.
	\end{equation}
	We estimate the second term as:
	\begin{equation*}
		\nor{\left(T_S - \widehat{T}_S\right) U_S^+ U}_2 \le \nor{T_S - \widehat{T}_S}_2 \nor{U_S^+}_2 \nor{U}_2 = \frac{1}{\sigma_\text{min}} \nor{T_S - \widehat{T}_S}_2	
	\end{equation*}	
%	\begin{align*}
%		\nor{\left(T_S - \widehat{T}_S\right) U_S^+ U}_2 &\le \nor{T_S - \widehat{T}_S}_2 \nor{U_S^+ U}_2 \\
%		&\le \sqrt{q}\nor{T_S - \widehat{T}_S}_1 \nor{U_S^+ U}_2 \\
%		&\le \sqrt{q}\nor{T_S - \widehat{T}_S}_1 \nor{U_S^+}_2 \nor{U}_2 = \frac{\sqrt{q}}{\sigma_\text{min}} \nor{T_S - \widehat{T}_S}_1. 
%	\end{align*}
	where we used the facts that $\nor{U_S^+}_2 = 1/\sigma_\text{min}$ and $\nor{U}_2=1$. By H\"older's inequality, we have 
	\begin{equation*}
		\nor{T - \widehat{T}}_{2} \le |\Omega|^{1/2} \nor{T - \widehat{T}}_{2,1}.
	\end{equation*}

	To estimate this term further, let $\delta_{x_i}^\epsilon \in L^1(\Omega)$ be an $\epsilon$-approximation to $\delta_{x_i}$ with norm $\|\delta_{x_i}^\epsilon\|_1=1$, defined as
	\begin{equation*}
		\delta_{x_i}^\epsilon(x) := \begin{cases}
			1 / \vol(B_\epsilon(x_i)), & x \in B_\epsilon(x_i), \\
			0, & \text{otherwise},
		\end{cases}
	\end{equation*}
	where $B_\epsilon(x_i)$ is the ball of radius $\epsilon$ centered at $x_i$, and $\vol(B_\epsilon(x_i))$ is the volume of this ball.
	We have
	\begin{align*}
		\nor{T_S - \widehat{T}_S}_1 &= \max_{i \in 1,2,\dots,q} \|T(~\cdot~,x_i) - \widehat{T}(~\cdot~,x_i)\|_1\\
		&= \|T(~\cdot~,x_k) - \widehat{T}(~\cdot~,x_k)\|_1 \\
		&= \lim_{\epsilon \rightarrow 0} \nor{(T-\widehat{T}) \delta_{x_k}^\epsilon}_1 \\
		&\le \sup_{\|v\|_1=1} \nor{(T - \widehat{T}) v}_1 = \nor{T - \widehat{T}}_1,
	\end{align*}
	where $k$ is the index achieving the maximum in the first line. Thus
	\begin{equation}
	\label{eq:_gamma_A_bound}
	\nor{\left(T_S - \widehat{T}_S\right) U_S^+ U}_2 \le \frac{\sqrt{q}}{\sigma_\text{min}}\nor{T - \widehat{T}}_1.
	\end{equation}
	The proposition follows from substituting bounds \eqref{eq:A_nogamma} and \eqref{eq:_gamma_A_bound} into \eqref{eq:two_bound_align}.
\end{proof}

Theorem \ref{thm:smooth_basis_err} shows that the error in our approximation depends on (1) how well the row space of $\mathcal{U}$ captures the row space of $\mathcal{T}$ (smaller $\|\mathcal{T} - \mathcal{T}\mathcal{U}^T\mathcal{U}\|$ $\implies$ less error), and (2) the minimum singular value of $\mathcal{U}_X$ (larger $\sigma_\text{min}$ $\implies$ less error). We can decrease $\|\mathcal{T} - \mathcal{T}\mathcal{U}^T\mathcal{U}\|$ by including more functions $u_i$ in $\mathcal{U}$. We can increase $\sigma_\text{min}$ by incorporating more batches of sample points $X_b$ in the approximation. But if $\sigma_\text{min}>1$, there are diminishing returns to making $\sigma_\text{min}$ bigger. By analogy to constrained optimization, $\|\mathcal{T} - \mathcal{T}\mathcal{U}^T\mathcal{U}\|$ is like an objective function to be minimized, while ensuring $\sigma_\text{min}$ is not small is like a constraint to be satisfied. We therefore increase the accuracy of the approximation by including more functions in $\mathcal{U}$, adding more batches of sample points as needed to ensure $\sigma_\text{min} > 1$.

The error in the approximation for batches $1,\dots,b$ is estimated with the the error for the next batch
\begin{equation*}
	\mathcal{E} := \frac{\nor{\eta^{b+1} - \widetilde{\eta}^{b+1}}}{\nor{\eta^{b+1}}},
\end{equation*}
where $\eta^{b+1}$ and $\widetilde{\eta}^{b+1}$ are the responses of $\mathcal{H}$ and $\widetilde{\mathcal{H}}$, respectively, to the Dirac comb at the next batch $\mathtt{S}_{b+1}$ of sample points, that were not used in the approximation $\widetilde{\mathcal{H}}$ yet. Entries of $\widetilde{\eta}_{b+1}$ take the form
\begin{equation}
\label{eq:eta_test}
	\widetilde{\eta}^{b+1}(y_k) = \sum_{x_j \in S_{b+1}} \widetilde{H}(y_k,x_j),
\end{equation}
which we compute using Algorithm \ref{alg:H_entries} for all degree of freedom locations $y_k$ for the discretization of the problem.
If $\mathcal{E}$ is larger than a desired tolerance, we compute more eigenmodes of $\Delta_N^{-1}$ and add those modes to the basis $\mathcal{U}$. Adding more functions to $\mathcal{U}$ may change $\sigma_\text{min}$. If $\sigma_\text{min}<1$ after this change, we select new batches of sample points and incorporate them in the approximation until $\sigma_\text{min} \ge 1$ once again. This process repeats until $\mathcal{E}$ is less than the desired tolerance. Each time an additional batch is needed, we compute another batch for testing, and incorporate the batch that was previously used for testing into the approximation. Once the desired accuracy is reached, we incorporate the last batch that was used for testing into the approximation to increase the accuracy further.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
%	\Input{Function $v \mapsto \mathcal{H}v$\\
%		Function $v \mapsto \mathcal{H}^Tv$
%	}

%	\mbox{}\\

%	\Output{Responses $\eta^1, \dots, \eta^b, \eta^\text{test}$ of $\mathcal{H}$ to Dirac combs \\
%		Basis of smooth functions $u_1, \dots, u_r$ \\
%		Interpolation pseudoinverse $\mathcal{U}_X^+$}
	%	\Begin{
	{

		Determine $\Omega_I$ with Algorithm \ref{alg:boundary_region}

		Compute $\mu$ and $\Sigma$ with Algorithm \ref{alg:varhpi_mean_cov}

		Select first sample point batch $S_1$ (Section \ref{sec:choosing_sample_points})

		Compute $\eta^1$ (Section \ref{sec:get_impulse_response})

		Set $S = \{\}$ and $b=0$

%		Form $\mathbf{X} \in \mathbb{R}^{d \times N}$ as the array of all degree of freedom coordinates
%		
%		$\mathtt{C} \gets \mathtt{I}$
%		
%		$b=0$
%		
%		$\mathtt{S}_1 \gets \choosebatch{\mathtt{C}}$ \tcp*{Algorithm \ref{alg:point_choice}}
%		
%		$\mathtt{C} \gets \mathtt{C} \setminus \mathtt{S}_1$
%		
%		$\boldsymbol{\eta}_1 \gets \computediracresponse{\mathtt{S}_1}$ \tcp*{Algorithm \ref{alg:dirac_comb_response}}
%		
%		$\mathtt{S} \gets \{\}$

		\For{$k=1,2,\dots,$}{

			Compute $u_k$ as the $k^\text{th}$ eigenfunction of $\Delta_N^{-1}$

%			\tcp{$U_S$ is the matrix with entries $\left(U_S\right)_{ij} = u_i(x_j)$ for $x_j\in S$}

%			\If{$U_S$ has less columns than rows}{
%				Select more sample point batches until $U_S$ has more columns than rows
%				
%				Compute functions $\eta^b$ for the new sample point batches
%			}

			\While{$|S| < k$}{

				Set $b=b+1$

				Set $S=S\cup S_b$

				Select new sample point batch $S_{b+1}$

				Form $\eta^{b+1}$

%				Form Dirac comb $\xi^b$ associated with $S_b$ and compute $\eta^b=\mathcal{A}\xi^b$ 
%				
%				Add new columns to $U_S$ corresponding to points $x_j \in S_b$
%				
%				Add new columns to $U_S$ corresponding to points $x_j \in S_b$
%				
%				$b \gets b+1$
%			
%				$\mathtt{S} \gets \mathtt{S} \cup \mathtt{S}_b$
%				
%				$\mathtt{S}_{b+1} \gets \choosebatch{\mathtt{C}}$
%				
%				$\mathtt{C} \gets \mathtt{C} \setminus \mathtt{S}_{b+1}$
%				
%				$\boldsymbol{\eta}_{b+1} \gets \computediracresponse{\mathtt{S}_{b+1}}$

			}

			Compute $\widetilde{\eta}^{b+1}$ using formulas \eqref{eq:eta_test} and \eqref{eq:kernel_entry_formua}.

			\If{$\displaystyle\frac{\nor{\eta^{b+1} - \widetilde{\eta}^{b+1}}}{\nor{\eta^{b+1}}}$ is sufficiently small}{

				Terminate for loop

			}

		}

		Set $S = S \cup S_{b+1}$

		Add new columns to $U_S$ corresponding to points $x_j \in S_b$

%		$\mathtt{Q} \gets \mathtt{Q} \cup \mathtt{Q}_\text{test}$

%		\Return{$\mu, \Sigma$, $\eta^1, \dots, \eta^b, \eta^\text{test}$, $u_1, \dots, u_r$, $\mathcal{U}_X^+$}

	}
	\caption{Adaptive approximation of $\mathcal{H}$}
	\label{alg:adaptive}
\end{algorithm2e}

\section{Hierarchical matrix construction}

\section{Numerical results}

\section{Conclusions}
\label{sec:conclusions}

Some conclusions here.

\appendix
\section{Fast ellipsoid intersection test}
The procedure for choosing sample points relies on quickly determining whether two ellipsoids intersect. Let $E_p$ and $E_q$ be the ellipsoids defined as
\begin{align*}
	E_p :=& \{x : (x - \mu_p)^T \Sigma_p^{-1} (x - \mu_p) \le \tau^2\} \\
	E_q :=& \{x : (x - \mu_q)^T \Sigma_q^{-1} (x - \mu_q) \le \tau^2\}, \\
\end{align*}
where $\mu_p, \mu_q \in \mathbb{R}^d$, and $\Sigma_p, \Sigma_q \in \mathbb{R}^{d \times d}$ are positive definite. Let $K$ be the following one dimensional convex function:
\begin{equation*}
	K(s) := 1 - \frac{1}{\tau^2} (\mu_p - \mu_q)^T \left(\frac{1}{1-s}\Sigma_p + \frac{1}{s}\Sigma_q\right)^{-1}(\mu_p - \mu_q)	
\end{equation*}
In CITE it is shown that $E_p \cap E_q = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$. We check whether $E_p$ and $E_q$ intersect by minimizing $K(s)$ on $(0,1)$. If $K(s^*) <0$ at the minimizer $s^*$, then $E_p \cap E_q = \{\}$. Otherwise $E_p \cap E_q \neq \{\}$.

The function $K(s)$ may be evaluated quickly for many $s$ by pre-computing the solution to the generalized eigenvalue problem
\begin{equation*}
	\Sigma_p \Phi = \Sigma_q \Phi \Lambda,
\end{equation*}
where $\Phi \in \mathbb{R}^{d \times d}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_d)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$. The matrix $\Phi$ simultaneously diagonalizes $\Sigma_p$ and $\Sigma_q$, in the sense that $\Phi^T\Sigma_p\Phi = \Lambda$, and $\Phi^T\Sigma_q\Phi = I$, where $I$ is the $d \times d$ identity matrix. Using this diagonalization, and some algebraic manipulations, we may write $K(s)$ as
\begin{equation}
\label{eq:Ks_generalized}
K(s) = 1 - \frac{1}{\tau^2} \sum_{i=1}^d \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2,
\end{equation}
where $v := \Phi^T\left(\mu_p - \mu_q\right)$. We compute the generalized eigenvalue decomposition of $\Sigma_p$ and $\Sigma_q$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's algorithm (any fast 1 dimensional convex optimization routine may be used). The resulting algorithm for checking whether $E_p$ and $E_q$ intersect is summarized in Algorithm \ref{alg:ellipsoid_intersection_test}.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	\SetKwProg{Fn}{Function}{}{}
	\Input{Ellipsoid $E_p$ with mean $\mu_p$ and covariance $\Sigma_p/\tau^2$\\
		Ellipsoid $E_q$ with mean $\mu_q$ and covariance $\Sigma_q/\tau^2$ 
	}

	\Output{Boolean which is true if $E_p \cap E_q \neq \{\}$ and false otherwise}

	\mbox{}\\

	\Fn{$\ellipsoidsintersect{E_p}{E_q}$}
	{
		Solve generalized eigenvalue problem $\Sigma_p \Phi = \Sigma_q \Phi \Lambda$

		$v \gets \Phi^T\left(\mu_p - \mu_q\right)$

		$\displaystyle K^* \gets \min_{s \in (0,1)}~1 - \frac{1}{\tau^2} \sum_{i=1}^d \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2$

		\If{$K^* < 0$}{

			\Return{False}

		}
		\Else{

			\Return{True}

		}
	}
	\caption{Determining whether two ellipsoids intersect}
	\label{alg:ellipsoid_intersection_test}
\end{algorithm2e}

\section{Additional proofs}
\label{app:proofs}

\begin{proof}[Proof of Proposition \ref{prop:boundary_source}]
	Let $w \in L^2(\Omega)$ be arbitrary. We have
	\begin{align*}
		\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle(w) =& \int_\Omega w(y) \int_{\partial \Omega} A(y,x) dy dx \\
		=& \int_\Omega w(x) \int_{\partial \Omega} \varphi_x(y-x) dy dx = \left(w, z\right)_{L^2(\Omega)},
	\end{align*}
	where
	\begin{equation*}
		z(x) := \int_{\partial \Omega} \varphi_x(y-x) dy,	
	\end{equation*}
	which implies
	\begin{equation*}
		\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle^*(x) = z = \int_{\partial \Omega} \varphi_x(y-x) dy.
	\end{equation*}
	Since $\varphi_x$ is a non-negative function, the result follows from the fact that the integral of a non-negative function over a set of positive measure is zero if and only if the function is zero almost everywhere on that set.
\end{proof}

\begin{proof}[Proof of Theorem \ref{thm:vol_mean_cov}]
	Let $w \in L^2$ be arbitrary. We have
	\begin{align*}
		\left(w, \alpha\right)_{L^2(\Omega)} =& \int_\Omega w(x) \int_{\Omega} \langle\mathcal{A},\delta_x\rangle^*(z) dz dx\\
		=& \int_\Omega w(x) \int_{\Omega} A(z,x) dz dx \\
		=& \int_\Omega \int_\Omega w(x) A(z,x) C(z) dz dx \\
		=& \left(\mathcal{A}^T C\right)(w),
	\end{align*}
	which implies \eqref{eq:vol_mean_var_thm1}.

	Using similar techniques, we have
	\begin{align*}
		\left(w, \alpha \cdot \mu^i\right)_{L^2(\Omega)} =& \int_\Omega w(x) \alpha(x) \int_{\Omega} z^i \langle\mathcal{A},\delta_x\rangle^*(z)/\alpha(x) dz dx\\
		=& \int_\Omega \int_\Omega w(x) L^i(z) A(z,x) dz dx \\
		=& \left(\mathcal{A}^T L^i\right)(w),
	\end{align*}
	which implies \eqref{eq:vol_mean_var_thm2}.

	To prove \eqref{eq:vol_mean_var_thm3} we use the identity
	\begin{equation*}
		\Sigma(x) = \int_\Omega zz^T \rho_x(z) dz - \mu(x) \mu(x)^T.
	\end{equation*}
	We have
	\begin{align*}
		\left(w, \alpha \cdot \left(\Sigma^{ij}+\mu^i \cdot \mu^j\right)\right)_{L^2(\Omega)} =& \int_\Omega w(x) \alpha(x) \int_\Omega z^i z^j \langle\mathcal{A},\delta_x\rangle^*(z) / \alpha(x) dz dx \\
		=& \int_\Omega \int_\Omega w(x) Q^{ij}(z) A(z,x) dz dx \\
		=& \left(\mathcal{A}Q^{ij}\right)(w),
	\end{align*}
	which implies \eqref{eq:vol_mean_var_thm3}.

\end{proof}

\section{Discretized function space operations}
\label{app:discretized_operations}

As described in Section \ref{sec:discretization}, a function $u_h$ in a finite element space $V_h$ may be written in terms of its vector of coefficients, $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, with respect the finite element basis $\{\phi_i\}_{i=1}^N$. Recall that $\mathbf{M}$ is the mass matrix. We have
\begin{equation*}
	u(x) \approx u_h(x) = \sum_{i=1}^N \mathbf{u}_i \phi_i(x).
\end{equation*}
The finite element function space $V_h$ is isometrically isomorphic to the coefficient space $\mathbb{R}^N_\mathbf{M}$, and the dual space $V_h'$ is isometrically isomorphic to $\mathbb{R}^N_{\mathbf{M}^{-1}}$. In computations, we perform operations with coefficient vectors in $\mathbb{R}^N_\mathbf{M}$ and coefficient dual vectors in $\mathbb{R}^N_{\mathbf{M}^{-1}}$, rather than functions in $V_h$ and functionals in $V_h'$. Here we describe how to perform common function space operations in terms of coefficient vectors and coefficient dual vectors.

\begin{description}
	\item[Coefficients of a functional:] A linear functional $\psi_h \in V_h'$ has coefficient dual vector $\boldsymbol{\psi} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$, which has entries
	\begin{equation*}
		\boldsymbol{\psi}_i = \psi_h(\phi_i), \quad i=1,\dots,N.
	\end{equation*}
	The action of a linear functional $\psi_h \in V_h'$ on a function $u_h \in V_h$ is given by
	\begin{equation*}
		\psi_h(v_h) = \mathbf{v}^T \boldsymbol{\psi}.
	\end{equation*}
	\item[Riesz representation of a functional:] Let $\psi_h \in V_h'$. We have
	\begin{equation*} 
		 \psi_h^* \in V_h: \quad \left(\psi_h^*, v_h\right)_{L^2(\Omega)} = \psi_h(v_h) \quad\forall v_h \in V_h \quad \Leftrightarrow \quad \boldsymbol{\psi}^* = \mathbf{M}^{-1} \boldsymbol{\psi}.
	\end{equation*}
	The function $\psi_h^*$ is the Riesz representation of the functional $\psi_h$, and $\boldsymbol{\psi}^* \in \mathbb{R}^N_\mathbf{M}$ is the coefficient vector of $\psi_h^*$.
	\item[$L^2$ projection:] Let $f \in L^2(\Omega)$, and define $\mathbf{f}^* \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ to be the vector with entries
	\begin{equation*}
		\mathbf{f}^*_i = \int_\Omega f(x) \phi_i(x) dx, \quad i=1,\dots,N.
	\end{equation*}
	We have
	\begin{equation*}
		f_h \in V_h, \quad \left(f_h, v_h\right)_{L^2(\Omega)} = \left(f, v_h\right)_{L^2(\Omega)} \quad\forall v_h \in V_h \quad \Leftrightarrow \quad \mathbf{f} = \mathbf{M}^{-1}\mathbf{f}^*.	
	\end{equation*}
	The function $f_h$ is the $L^2$ projection of $f$ onto $V_h$, and $\mathbf{f} \in \mathbb{R}^N_\mathbf{M}$ is the coefficient vector of $f_h$.
	\item[Matrix representation of an operator:] Let $\mathcal{B}_h : V_h \rightarrow V_h'$ be a linear operator, and let $\mathbf{B} \in \mathbf{R}^{N \times N}$ be the matrix with entries
	\begin{equation*}
		\mathbf{B}_{ij} = \left(\mathcal{B}_h \phi_j\right)(\phi_i).
	\end{equation*}
	The matrix $\mathbf{B}$ maps $\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}$ by matrix multiplication, and
	\begin{equation*}
		\psi_h = \mathcal{B}_h u_h \quad \Leftrightarrow \quad \boldsymbol{\psi} = \mathbf{B} \mathbf{u}.
	\end{equation*}
	\item[Transpose:] The matrix $\mathbf{B}^T$ maps $\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}$ via matrix multiplication, and
	\begin{equation*}
		\psi_h = \mathcal{B}_h^T u_h \quad \Leftrightarrow \quad \boldsymbol{\psi} = \mathbf{B}^T \mathbf{u}.
	\end{equation*}
%	\item[Implicitly defined matrix representation:]
%	Let $\mathcal{A}_h:V_h \rightarrow V_h'$ be a finite dimensional approximation of $\mathcal{A}:L^2(\Omega) \rightarrow L^2(\Omega)'$, and let $\mathbf{A}$ be the matrix representation of $\mathcal{A}_h$. The matrix $\mathbf{A}$ is implicitly defined, in the sense that we have a algorithm to compute the matrix-vector products $\mathbf{u} \mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{A}^T \mathbf{u}$ for arbitrary $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, but we do not have direct access to matrix entries $\mathbf{A}_{ij}$. In principle, all matrix entries of $\mathbf{A}$ could be computed by applying $\mathbf{A}$ to each column of the identity matrix, but this is far too expensive to be practical.
	\item[Distributions:] Let $\mu:C\left(\overline{\Omega}\right)\rightarrow \mathbb{R}$ be a distribution. If $\mu(\phi_i)$ is well-defined for all basis vectors $\phi_i$, then the restriction of $\mu$ to domain $V_h$ is a linear functional $\mu_h\in V_h'$. The functional $\mu_h$ has a coefficient dual vector $\boldsymbol{\mu} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ with entries $\boldsymbol{\mu}_i=\mu(\phi_i)$ for $i=1,\dots,N$. 
	\item[Action of an operator on a distribution:] Since $V_h$ is finite-dimensional, the restriction, $\mu_h \in V_h$, of a distribution, $\mu$, has a a Riesz representation $\mu_h^* \in V_h$. Let $B_h$ be the integral kernel associated with an operator $\mathcal{B}_h:V_h \rightarrow V_h'$. For $w_h \in V_h$, we have
	\begin{align*}
		\langle \mathcal{B}_h, \mu_h \rangle(w) &= \int_\Omega w_h(y) \mu_h\left(B_h(y, \cdot)\right) dy \\
		&= \int_\Omega w_h(y) \left(\int_\Omega B_h(y,x) \mu_h^*(x) dx\right) dy \\
		&= \int_\Omega \int_\Omega w_h(y) B_h(y,x) \mu_h^*(x) dx dy = \left(\mathcal{B}_h \mu_h^*\right)(w).
	\end{align*}
	Going from the first to the second line we used the definition of the Riesz representation. The action of $\mathcal{B}_h$ on $\mu_h$ in the sense of distributions is therefore equal to the action of $\mathcal{B}_h$ on $\mu^*_h$ in the conventional sense. We have
	\begin{equation*}
		\psi_h = \langle \mathcal{B}_h, \mu_h \rangle \quad \Leftrightarrow \quad \psi_h = \mathcal{B}_h \mu_h^* \quad \Leftrightarrow \quad \boldsymbol{\psi} = \mathbf{B} \mathbf{M}^{-1} \boldsymbol{\mu}.
	\end{equation*}
\end{description}

\section*{Acknowledgments}
We acknowledge

\bibliographystyle{siamplain}
\bibliography{references}
\end{document}

### 2020-11-30 paper/localpsf.tex
edited abstrach

@@
 \begin{abstract}
-  We present an efficient matrix-free method for approximating smoothly varying blur operators, and apply the method to approximate Hessians in large-scale inverse problems with highly informative data.
-Such Hessians are high rank and therefore expensive to approximate with existing methods.
-We compute the impulse responses of the operator at many points by applying the operator to a small number of Dirac combs of point sources.
-Within each Dirac comb, the point sources must be spaced sufficiently far apart so that the impulse responses to those point sources do not overlap with each other or with the boundary.
+  We present an efficient matrix-free method for approximating locally translation invariant operators with non-negative integral kernels.
+The idea of the method is to compute impulse responses of the operator at a collection of well-spaced points, then interpolate these impulse responses to approximate impulse responses at other arbitrary points.
+We convert the resulting approximation to hierarchical matrix format, which allows us to perform matrix inversion or other matrix operations using fast hierarchical matrix arithmetic.
+Impulse responses are computed by applying the operator to a small number of Dirac comb ``batches'' of point sources, and special care is taken to ensure that the supports of the impulse responses within each batch do not overlap.
+We apply the method to approximate Hessians in large-scale PDE-constrained inverse problem with highly informative data.
+Numerical results demonstrate that our method substantially outperforms existing Hessian approximation methods which are based on low-rank approximation.
+Our method is able to form high quality approximations of high rank Hessians using only a small number of Hessian matrix-vector products.
+%  Within each Dirac comb, the point sources must be spaced sufficiently far apart so that the impulse responses to those point sources do not overlap.
 We estimate the required spacing of point sources by applying the operator to a small number of constant, linear, and quadratic functions, then postprocessing the results.
-Once computed, the impulse responses are used to form a low-rank approximation of the spatially varying impulse response function (SVIR), using a continuous analog of the interpolative matrix decomposition.
-Smoothly varying blur operators have a low rank SVIR, even if the operator itself is high rank.
-We convert the resulting approximation to hierarchical matrix format, where it is inverted using fast hierarchical matrix arithmetic and used to solve linear systems.
-We apply the method to a large-scale PDE-constrained inverse problem with highly informative data.
-%  {\color{red}Numerical results demonstrate that our method substantially outperforms existing methods.}
 \end{abstract}
@@
 
-Let $\Omega \subset \mathbb{R}^d$ be a bounded domain, let $a:L^2(\Omega) \times L^2(\Omega) \rightarrow \mathbb{R}$,
+Let $\Omega \subset \mathbb{R}^d$ be a bounded domain, and let $a:L^2(\Omega) \times L^2(\Omega) \rightarrow \mathbb{R}$,
 \begin{equation}
@@
 \end{equation}
-be a bilinear form with integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$, and let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be the corresponding linear operator generated by currying.
+be a bilinear form with non-negative integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$, $A(y,x) \ge 0$.
+Let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be the corresponding linear operator generated by currying.
 That is, for each $u\in L^2(\Omega)$, the linear functional $\mathcal{A}u:L^2(\Omega)\rightarrow \mathbb{R}$ is defined by

### 2020-11-30 code/src/grid_interpolate.cpp

@@ -4,7 +4,7 @@
 #include <Eigen/Dense>
 #include <Eigen/LU>
 
-//#include <unsupported/Eigen/CXX11/Tensor>
+// #include <unsupported/Eigen/CXX11/Tensor>
 
 using namespace Eigen;
 using namespace std;

### 2020-11-30 paper/localpsf.tex
editing intro

@@
 \begin{abstract}
-  We present an efficient matrix-free method for approximating locally translation invariant operators with non-negative integral kernels.
+  We present an efficient matrix-free method for approximating locally translation invariant operators with locally supported non-negative integral kernels.
 The idea of the method is to compute impulse responses of the operator at a collection of well-spaced points, then interpolate these impulse responses to approximate impulse responses at other arbitrary points.
@@
 
-Let $\Omega \subset \mathbb{R}^d$ be a bounded domain, and let $a:L^2(\Omega) \times L^2(\Omega) \rightarrow \mathbb{R}$,
-\begin{equation}
-	\label{eq:kernel_representation}
-	a(v,u) := \int_\Omega \int_\Omega v(y) A(y,x) u(x) dx dy,
-\end{equation}
-be a bilinear form with non-negative integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$, $A(y,x) \ge 0$.
-Let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be the corresponding linear operator generated by currying.
-That is, for each $u\in L^2(\Omega)$, the linear functional $\mathcal{A}u:L^2(\Omega)\rightarrow \mathbb{R}$ is defined by
-\begin{equation*}
-	(\mathcal{A}u)(v) := a(v,u).
-\end{equation*}
-We seek to approximate operators $\mathcal{A}$ in the ``matrix free'' setting, in which access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$.
+In this paper, we present a fast matrix free method for approximating locally translation-invariant operators, $\mathcal{A}$, with locally supported non-negative integral kernels.
+By ``matrix-free'', we mean that $\mathcal{A}$ is available only through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$.
+Such operators arise, for example, as Schur complements in reduced space methods for solving partial differential equations (PDEs), and as Hessians in distributed parameter PDE-constrained inverse problems.
+
+Specfically, we consider operators $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ of the form
+\begin{equation*}
+(\mathcal{A}u)(v) := \int_\Omega \int_\Omega v(y) A(y,x) u(x) dx dy,
+\end{equation*}
+where $\Omega \subset \mathbb{R}^d$ is a bounded domain, and the integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$ is non-negative in the sense that  $A(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$.
+%Let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be the corresponding linear operator generated by currying.
+That is, for each $u\in L^2(\Omega)$, the linear functional $\mathcal{A}u:L^2(\Omega)\rightarrow \mathbb{R}$ is
+%We approximate the operator $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ defined by
+%\begin{equation*}
+%	(\mathcal{A}u)(v) := a(v,u).
+%\end{equation*}
+%We seek to approximate operators $\mathcal{A}$ in the ``matrix free'' setting, in which access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$.
 Here ($\mathcal{A}^Tv)(u) := (\mathcal{A}u)(v)$.
@@
 but one cannot easily access entries $A(y,x)$ of the kernel.
-Such operators arise, for example, as Schur complements in reduced space methods for solving partial differential equations (PDEs), and as Hessians in distributed parameter PDE-constrained inverse problems.
+
+We further assume that $A$ is locally supported in the sense that the impulse response function, 
+\begin{equation*}
+	\varphi_x(y) := A(y, x)
+\end{equation*}
+is supported in a region near $x$.
 

### 2020-11-30 paper/localpsf.tex
editing intro

@@
 %e.g., by the Schwarz kernel theorem,
-but one cannot easily access entries $A(y,x)$ of the kernel.
-
-We further assume that $A$ is locally supported in the sense that the impulse response function, 
-\begin{equation*}
-	\varphi_x(y) := A(y, x)
-\end{equation*}
-is supported in a region near $x$.
+but one cannot easily access entries $A(y,x)$ of the kernel. We assume that $A$ is locally supported in the sense that the impulse response of $\mathcal{A}$ to a point source at $x$, 
+\begin{equation*}
+	\varphi_x(y) := A(y+x, x),
+\end{equation*}
+is supported in a region near $0$, and we assume that $\mathcal{A}$ is locally translation invariant in the sense that
+\begin{equation*}
+A(y+h, x+h) \approx A(y,x),
+\end{equation*}
+so long as $h$ is not too large.
 

### 2020-11-30 paper/localpsf.tex
editing intro

--- 2020-11-30 paper/localpsf.tex
@@
 %e.g., by the Schwarz kernel theorem,
-but one cannot easily access entries $A(y,x)$ of the kernel.
-
-We further assume that $A$ is locally supported in the sense that the impulse response function, 
-\begin{equation*}
-	\varphi_x(y) := A(y, x)
-\end{equation*}
-is supported in a region near $x$.
+but one cannot easily access entries $A(y,x)$ of the kernel. We assume that the impulse response of $\mathcal{A}$ is locally supported in the sense that the function 
+$
+A(~\cdot~, x)
+$
+is supported in a region near $x$, and we assume that $\mathcal{A}$ is locally translation invariant in the sense that
+\begin{equation*}
+A(y+h, x+h) \approx A(y,x),
+\end{equation*}
+so long as $h$ is not too large.
 

### 2020-11-30 paper/localpsf.tex
editing intro

@@
 
-In this paper, we present a fast matrix free method for approximating locally translation-invariant operators, $\mathcal{A}$, with locally supported non-negative integral kernels.
-By ``matrix-free'', we mean that $\mathcal{A}$ is available only through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$.
+In this paper, we present a fast matrix free method for approximating locally translation-invariant operators, $\mathcal{A}$, which have locally supported non-negative integral kernels.
 Such operators arise, for example, as Schur complements in reduced space methods for solving partial differential equations (PDEs), and as Hessians in distributed parameter PDE-constrained inverse problems.
 
-Specfically, we consider operators $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ of the form
-\begin{equation*}
+Let $\Omega \subset \mathbb{R}^d$ be a bounded domain, and let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the form
+\begin{equation}
+\label{eq:kernel_representation}
 (\mathcal{A}u)(v) := \int_\Omega \int_\Omega v(y) A(y,x) u(x) dx dy,
-\end{equation*}
-where $\Omega \subset \mathbb{R}^d$ is a bounded domain, and the integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$ is non-negative in the sense that  $A(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$.
+\end{equation}
+with integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$. We focus on operators of this form which have the following properties:
+\begin{description}
+\item[Matrix-free:] One cannot easily access kernel entries $A(y,x)$.
+Access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$.
+\item[Non-negative:] $A(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$.
+\item[Local support:] $A(y, x)$ is supported near $y=x$.
+\item[Local translation invariance:] $A(y+h, x+h) \approx A(y,x)$
+if $h$ is not too large.
+\end{description}
+Access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$.
+Integral kernel representation \eqref{eq:kernel_representation} is assumed to exist,
+%e.g., by the Schwarz kernel theorem,
+but o
+One cannot easily access entries $A(y,x)$ of the kernel.
+
+% is non-negative in the sense that  $A(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$. 
 %Let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be the corresponding linear operator generated by currying.
@@
 Here ($\mathcal{A}^Tv)(u) := (\mathcal{A}u)(v)$.
-Integral kernel representation \eqref{eq:kernel_representation}, is assumed to exist,
+Integral kernel representation \eqref{eq:kernel_representation} is assumed to exist,
 %e.g., by the Schwarz kernel theorem,
 but one cannot easily access entries $A(y,x)$ of the kernel. We assume that the impulse response of $\mathcal{A}$ is locally supported in the sense that the function 
-$
-A(~\cdot~, x)
-$
-is supported in a region near $x$, and we assume that $\mathcal{A}$ is locally translation invariant in the sense that
+$y \mapsto A(y, x)$ is supported in a region near $y=x$, and we assume that $\mathcal{A}$ is locally translation invariant in the sense that
 \begin{equation*}

### 2020-12-01 paper/localpsf.tex
editing intro

@@
 \end{description}
-Access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$.
-Integral kernel representation \eqref{eq:kernel_representation} is assumed to exist,
+If $\mathcal{A}$ were translation invariant---i.e., if the exact equality $A(y+h, x+h) = A(y,x)$ held for arbitrary $x$, $y$, $h$---then it is straightforward to show that $\mathcal{A}$ would take the form of a convolution operator $\left(\mathcal{A}u\right)(v) = \left\langle v, \varphi_p \ast u\right\rangle_{L^2(\Omega)}$.
+The convolution kernel,
+\begin{equation*}
+	\varphi_x(y) := A(y+x,x)
+\end{equation*}
+is the result of applying $\mathcal{A}$ to a point source at some fixed point $x$, then translating the resulting impulse response function re-center it at zero instead of $x$.
+Locally translation invariant operators therefore act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes.
+
+%Access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$.
+%Integral kernel representation \eqref{eq:kernel_representation} is assumed to exist,
 %e.g., by the Schwarz kernel theorem,
-but o
-One cannot easily access entries $A(y,x)$ of the kernel.
+%but o
+%One cannot easily access entries $A(y,x)$ of the kernel.
 
@@
 Here ($\mathcal{A}^Tv)(u) := (\mathcal{A}u)(v)$.
-Integral kernel representation \eqref{eq:kernel_representation} is assumed to exist,
+%Integral kernel representation \eqref{eq:kernel_representation} is assumed to exist,
 %e.g., by the Schwarz kernel theorem,
-but one cannot easily access entries $A(y,x)$ of the kernel. We assume that the impulse response of $\mathcal{A}$ is locally supported in the sense that the function 
-$y \mapsto A(y, x)$ is supported in a region near $y=x$, and we assume that $\mathcal{A}$ is locally translation invariant in the sense that
-\begin{equation*}
-A(y+h, x+h) \approx A(y,x),
-\end{equation*}
-so long as $h$ is not too large.
+%but one cannot easily access entries $A(y,x)$ of the kernel. We assume that the impulse response of $\mathcal{A}$ is locally supported in the sense that the function 
+%$y \mapsto A(y, x)$ is supported in a region near $y=x$, and we assume that $\mathcal{A}$ is locally translation invariant in the sense that
+%\begin{equation*}
+%A(y+h, x+h) \approx A(y,x),
+%\end{equation*}
+%so long as $h$ is not too large.
 

### 2020-12-01 paper/localpsf.tex
editing intro

@@
 
-In this paper, we present a fast matrix free method for approximating locally translation-invariant operators, $\mathcal{A}$, which have locally supported non-negative integral kernels.
+In this paper, we present a fast matrix free method for approximating locally translation-invariant operators, $\mathcal{A}$, that have locally supported non-negative integral kernels.
 Such operators arise, for example, as Schur complements in reduced space methods for solving partial differential equations (PDEs), and as Hessians in distributed parameter PDE-constrained inverse problems.
@@
 \end{description}
-If $\mathcal{A}$ were translation invariant---i.e., if the exact equality $A(y+h, x+h) = A(y,x)$ held for arbitrary $x$, $y$, $h$---then it is straightforward to show that $\mathcal{A}$ would take the form of a convolution operator $\left(\mathcal{A}u\right)(v) = \left\langle v, \varphi_p \ast u\right\rangle_{L^2(\Omega)}$.
+
+If $\mathcal{A}$ were translation invariant---i.e., if the exact equality $A(y+h, x+h) = A(y,x)$ held for arbitrary $x$, $y$, $h$---then it is straightforward to show that $\mathcal{A}$ would take the form of the convolution operator $\left(\mathcal{A}u\right)(v) = \left\langle v, \varphi_p \ast u\right\rangle_{L^2(\Omega)}$.
+Here $f \ast g$ denotes convolution of the functions $f$ and $g$.
 The convolution kernel,
 \begin{equation*}
-	\varphi_x(y) := A(y+x,x)
+	\varphi_x(y) := A(y+x,x),
 \end{equation*}
 is the result of applying $\mathcal{A}$ to a point source at some fixed point $x$, then translating the resulting impulse response function re-center it at zero instead of $x$.
-Locally translation invariant operators therefore act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes.
+Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes.
+We therefore sample $\varphi_{x_i}$ at a collection of points $\{x_i\}_{i=1}^r \subset \Omega$, and patch together the local convolution operators $u \mapsto \varphi_{x_i} \ast u$ using smooth weighting functions $w_i(x)$.
+This yields a product-convolution approximation, $\widetilde{\mathcal{A}} \approx \mathcal{A}$, given by
+\begin{equation*}
+	\left(\widetilde{\mathcal{A}}u\right)(v) := \left\langle v, \sum_{i=1}^r \varphi_{x_i} \ast \left(w_i \cdot u\right) \right\rangle_{L^2(\Omega)}.
+\end{equation*}
+Here $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$.
+
+While kernel entries $\mathcal{A}(y,x)$ are not easily computable, kernel entries $\widetilde{A}(y,x)$ are given by the formula
+\begin{equation*}
+	\widetilde{A}(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x).
+\end{equation*}
 

### 2020-12-01 paper/localpsf.tex
editing intro

@@
 \begin{equation*}
-	\left(\widetilde{\mathcal{A}}u\right)(v) := \left\langle v, \sum_{i=1}^r \varphi_{x_i} \ast \left(w_i \cdot u\right) \right\rangle_{L^2(\Omega)}.
-\end{equation*}
-Here $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$.
-
-While kernel entries $\mathcal{A}(y,x)$ are not easily computable, kernel entries $\widetilde{A}(y,x)$ are given by the formula
+	\left(\widetilde{\mathcal{A}}u\right)(v) := \left\langle v, \sum_{i=1}^r \varphi_{x_i} \ast \left(w_i \cdot u\right) \right\rangle_{L^2(\Omega)},
+\end{equation*}
+where $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$.
+
+While kernel entries $A(y,x)$ are not easily computable, kernel entries $\widetilde{A}(y,x)$ are given by the formula
 \begin{equation*}

### 2020-12-04 paper/localpsf.tex
editing intro

@@
 We apply the method to approximate Hessians in large-scale PDE-constrained inverse problem with highly informative data.
-Numerical results demonstrate that our method substantially outperforms existing Hessian approximation methods which are based on low-rank approximation.
+Numerical results demonstrate that our method substantially outperforms existing state-of-the-art Hessian approximation methods which are based on low-rank approximation.
 Our method is able to form high quality approximations of high rank Hessians using only a small number of Hessian matrix-vector products.
@@
 
-If $\mathcal{A}$ were translation invariant---i.e., if the exact equality $A(y+h, x+h) = A(y,x)$ held for arbitrary $x$, $y$, $h$---then it is straightforward to show that $\mathcal{A}$ would take the form of the convolution operator $\left(\mathcal{A}u\right)(v) = \left\langle v, \varphi_p \ast u\right\rangle_{L^2(\Omega)}$.
-Here $f \ast g$ denotes convolution of the functions $f$ and $g$.
-The convolution kernel,
-\begin{equation*}
-	\varphi_x(y) := A(y+x,x),
-\end{equation*}
-is the result of applying $\mathcal{A}$ to a point source at some fixed point $x$, then translating the resulting impulse response function re-center it at zero instead of $x$.
+If $\mathcal{A}$ were translation invariant (that is, if $A(y+h, x+h) = A(y,x)$ for all $x$, $y$, $h$) then it is straightforward to show that $\mathcal{A}$ would be a convolution operator, and the convolution kernel, $\varphi_x$, would be given by
+\begin{equation}
+\label{eq:convolution_kernel}
+	\varphi_x(y) := A(y+x,x).
+\end{equation}
+The function $\varphi_x$ is the result of applying $\mathcal{A}$ to a point source at some fixed point $x$, then translating the resulting impulse response function re-center it at zero instead of $x$.
 Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes.
-We therefore sample $\varphi_{x_i}$ at a collection of points $\{x_i\}_{i=1}^r \subset \Omega$, and patch together the local convolution operators $u \mapsto \varphi_{x_i} \ast u$ using smooth weighting functions $w_i(x)$.
+Locally translation invariant operators may therefore be approximated by spatially varying weighted sums of convolution operators of the form
+\begin{equation}
+\label{eq:product_convolution}
+	\left(\mathcal{A}u\right)(v) \approx \left(\widetilde{\mathcal{A}}u\right)(v) := \left\langle v, \sum_{i=1}^r \varphi_{x_i} \ast \left(w_i \cdot u\right) \right\rangle_{L^2(\Omega)},
+\end{equation}
+Where $f \cdot g$ denotes pointwise multiplication, and $f \ast g$ denotes convolution of the functions $f$ and $g$. Here
+\begin{itemize}
+	\item $\varphi_{x_i}$ is given by formula \eqref{eq:convolution_kernel}, but with $x$ replaced by $x_i$.
+	\item $\{x_i\}_{i=1}^r \subset \Omega$ is a collection of points scattered throughout the domain.
+	\item The functions $w_i(x)$ are spatially varying weighting function that interpolate the convolution kernels $\varphi_{x_i}$.
+\end{itemize}
+\eqref{eq:product_convolution}
+
+To take advantage of local translation invariance, we sample $\varphi_{x_i}$ at a collection of points $\{x_i\}_{i=1}^r \subset \Omega$, and patch together the local convolution operators $u \mapsto \varphi_{x_i} \ast u$ using smooth weighting functions $w_i(x)$.
 This yields a product-convolution approximation, $\widetilde{\mathcal{A}} \approx \mathcal{A}$, given by
@@
 \end{equation*}
-where $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$.
+where $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$. There is a long history of product-convolution operator approximation of this form; for a review of the literature,
 
@@
 \end{equation*}
+
+ $\left(\mathcal{A}u\right)(v) = \left\langle v, \varphi_p \ast u\right\rangle_{L^2(\Omega)}$. Here $f \ast g$ denotes convolution of the functions $f$ and $g$, and the convolution kernel,
 

### 2020-12-04 paper/localpsf.tex
editing intro

@@
   We present an efficient matrix-free method for approximating locally translation invariant operators with locally supported non-negative integral kernels.
-The idea of the method is to compute impulse responses of the operator at a collection of well-spaced points, then interpolate these impulse responses to approximate impulse responses at other arbitrary points.
-We convert the resulting approximation to hierarchical matrix format, which allows us to perform matrix inversion or other matrix operations using fast hierarchical matrix arithmetic.
-Impulse responses are computed by applying the operator to a small number of Dirac comb ``batches'' of point sources, and special care is taken to ensure that the supports of the impulse responses within each batch do not overlap.
+The idea of the method is to compute impulse responses of the operator at a collection of scattered points, then interpolate the computed impulse responses to approximate impulse responses at other arbitrary points.
+We convert the resulting approximation to hierarchical matrix format, which allows us to perform matrix inversion and other matrix operations using fast hierarchical matrix arithmetic.
+Impulse responses are computed by applying the operator to a small number of Dirac comb ``batches'' of point sources.
+A key innovation of our method is a matrix-free procedure for choosing as many points as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
 We apply the method to approximate Hessians in large-scale PDE-constrained inverse problem with highly informative data.
@@
 \label{eq:product_convolution}
-	\left(\mathcal{A}u\right)(v) \approx \left(\widetilde{\mathcal{A}}u\right)(v) := \left\langle v, \sum_{i=1}^r \varphi_{x_i} \ast \left(w_i \cdot u\right) \right\rangle_{L^2(\Omega)},
+	\left(\mathcal{A}u\right)(v) \approx \left(\widetilde{\mathcal{A}}u\right)(v) := \left\langle v, \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right\rangle_{L^2(\Omega)},
 \end{equation}
@@
 \begin{itemize}
-	\item $\varphi_{x_i}$ is given by formula \eqref{eq:convolution_kernel}, but with $x$ replaced by $x_i$.
+	\item $\varphi_i := \varphi_{x_i}$ is given by formula \eqref{eq:convolution_kernel}, but with $x$ replaced by $x_i$.
 	\item $\{x_i\}_{i=1}^r \subset \Omega$ is a collection of points scattered throughout the domain.
-	\item The functions $w_i(x)$ are spatially varying weighting function that interpolate the convolution kernels $\varphi_{x_i}$.
+	\item The functions $w_i(x)$ are spatially varying weighting function that interpolate the convolution kernels $\varphi_i$.
 \end{itemize}
-\eqref{eq:product_convolution}
-
-To take advantage of local translation invariance, we sample $\varphi_{x_i}$ at a collection of points $\{x_i\}_{i=1}^r \subset \Omega$, and patch together the local convolution operators $u \mapsto \varphi_{x_i} \ast u$ using smooth weighting functions $w_i(x)$.
-This yields a product-convolution approximation, $\widetilde{\mathcal{A}} \approx \mathcal{A}$, given by
-\begin{equation*}
-	\left(\widetilde{\mathcal{A}}u\right)(v) := \left\langle v, \sum_{i=1}^r \varphi_{x_i} \ast \left(w_i \cdot u\right) \right\rangle_{L^2(\Omega)},
-\end{equation*}
-where $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$. There is a long history of product-convolution operator approximation of this form; for a review of the literature,
+Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution.
+For background on product-convolution approximations, we recommend [CITES].
 

### 2020-12-04 paper/localpsf.tex
editing intro

@@
   We present an efficient matrix-free method for approximating locally translation invariant operators with locally supported non-negative integral kernels.
-The idea of the method is to compute impulse responses of the operator at a collection of scattered points, then interpolate the computed impulse responses to approximate impulse responses at other arbitrary points.
+The idea of the method is to compute impulse responses of the operator at a collection of scattered points, interpolate the computed impulse responses to approximate impulse responses at other arbitrary points.
 We convert the resulting approximation to hierarchical matrix format, which allows us to perform matrix inversion and other matrix operations using fast hierarchical matrix arithmetic.
@@
 \end{equation}
-Where $f \cdot g$ denotes pointwise multiplication, and $f \ast g$ denotes convolution of the functions $f$ and $g$. Here
+here $f \cdot g$ denotes pointwise multiplication, and $f \ast g$ denotes convolution of the functions $f$ and $g$. Here
 \begin{itemize}
@@
 Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution.
-For background on product-convolution approximations, we recommend [CITES].
-
-While kernel entries $A(y,x)$ are not easily computable, kernel entries $\widetilde{A}(y,x)$ are given by the formula
-\begin{equation*}
-	\widetilde{A}(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x).
-\end{equation*}
-
- $\left(\mathcal{A}u\right)(v) = \left\langle v, \varphi_p \ast u\right\rangle_{L^2(\Omega)}$. Here $f \ast g$ denotes convolution of the functions $f$ and $g$, and the convolution kernel,
+The more locally translation invariant an operator is, the smaller the number of terms that are required in \eqref{eq:product_convolution} to achieve an accurate approximation.
+
+Let $\widetilde{A}$ denote the integral kernel associated with $\widetilde{\mathcal{A}}$. While $A(y,x)$ is not easily computable, $\widetilde{A}(y,x)$ is given by
+\begin{equation}
+\label{eq:kernel_entries}
+	\widetilde{A}(y,x) = \sum_{i=1}^r \varphi_i(y-x) w_i(x).
+\end{equation}
+Formula \eqref{eq:kernel_entries} provides fast access to kernel entries $\widetilde{A}(y,x)$, which allows us to construct a hierarchical matrix (H-matrix) representation of $\widetilde{\mathcal{A}}$.
+Using this H-matrix representation, one can invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations with $\widetilde{\mathcal{A}}$, using fast H-matrix arithmetic.
+
+For background on product-convolution approximations, we recommend reading the following papers: [CITES].
+
+% $\left(\mathcal{A}u\right)(v) = \left\langle v, \varphi_p \ast u\right\rangle_{L^2(\Omega)}$. Here $f \ast g$ denotes convolution of the functions $f$ and $g$, and the convolution kernel,
 

### 2020-12-04 paper/localpsf.tex
editing intro
put in kriging stuff from overleaf document

@@
 \label{eq:product_convolution}
-	\left(\mathcal{A}u\right)(v) \approx \left(\widetilde{\mathcal{A}}u\right)(v) := \left\langle v, \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right\rangle_{L^2(\Omega)},
+	\left(\mathcal{A}u\right)(v) \approx \left(\widetilde{\mathcal{A}}u\right)(v) := \left\langle v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right\rangle_{L^2(\Omega)},
 \end{equation}
@@
 
-For background on product-convolution approximations, we recommend reading the following papers: [CITES].
+For background on product-convolution approximations, we recommend reading the following papers: [CITES]. For background on H-matrices, we recommend reading the following papers: [CITES].
 
@@
 
-\subsection{Motivating example: PDE constrained inverse problems with highly informative data}
+\subsection{Motivation: PDE constrained inverse problems with highly informative data}
 \label{sec:PDE_hessian_motivation}
@@
 
-\section{Interpolation of impulse response samples}
+\section{Interpolating impulse response samples}
+
+\section{Poisson PSF interpolation}
+
+We form an interpolant, $\widetilde{\varphi}_x$, of the point spread function (PSF), $\varphi_x$, as follows:
+\begin{equation*}
+	\varphi_x(y) \approx \widetilde{\varphi}_x(y) := \sum_{i=1}^q w_i(x)\varphi_{x_i}(y).
+\end{equation*}
+The functions $y \mapsto \varphi_{x_i}(y)$, are samples of the PSF at the sample points $x_i$, $i=1,\dots,q$.
+The functions $w_i(x)$, are smooth spatially varying weighting functions that solve the following optimization problem:
+\begin{equation}
+\label{eq:wi_optimization_problem}
+\begin{aligned}
+\min_{w_i} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
+\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
+\end{aligned}
+\end{equation}
+Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$.
+We refer to $\widetilde{\varphi}_x$ as the \emph{Poisson interpolant} of $\varphi_x$.
+
+Let $u_y$ denote the function $u_y(x) := \widetilde{\varphi}_x(y)$ for an arbitrary but fixed value of $y$. It is straightforward to show that $u_y$ solves the optimization problem
+\begin{equation}
+\label{eq:uy_optimization_problem}
+\begin{aligned}
+\min_{u_y} &\quad \frac{1}{2}\norm{\Delta u_y}_{L^2(\Omega)}^2 \\
+\text{such that} &\quad u_y(x_i) = \varphi_{x_i}(y), \quad i=1,\dots,q.
+\end{aligned}
+\end{equation}
+This follows from linearity of the KKT system associated with \eqref{eq:wi_optimization_problem}, which in turn follows from the fact that the the objective function in \eqref{eq:wi_optimization_problem} is quadratic, and the constraint is linear.
+Since the Laplacian measures local deviation from flatness [cite Jacobsen], from \eqref{eq:uy_optimization_problem} we see that the interpolant $x \mapsto u_y(x) = \widetilde{\varphi}_x(y)$ is the `flattest possible'' function (in a least-squares sense) that has Neumann boundary conditions and interpolates the data $\varphi_{x_i}(y)$.
+
+Optimization problems \eqref{eq:wi_optimization_problem} and \eqref{eq:uy_optimization_problem} have unique continuous solutions when the spatial dimension is $1$, $2$, or $3$.
+When inverting for parameter fields defined on $4$ or higher spatial dimensions, the solution to this optimization problem does not exist, and Poisson interpolation (as described here) should not be used\footnote{The reason for this dependence on spatial dimension is related to the fact that the fundamental solution of the biharmonic equation is continuous in $3$ or less dimensions, but singular in $4$ or more dimensions.
+In the unusual scenario where one is inverting for a parameter field in $4$ or more spatial dimensions, one could use a modified Poisson interpolation where $\Delta$ is replaced with $\Delta^\alpha$ in the objective function of \eqref{eq:wi_optimization_problem}, where $\alpha>1$ is chosen so that the fundamental solution of $\Delta^{2\alpha}$ is continuous.}.
+
+A key advantage of Poisson interpolation over other PSF interpolation methods, most notably radial basis functions [cite Escande etc], is that Poisson interpolation is \emph{geometry-aware}---Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions.
+Consider the Pine island glacier geometry shown in Figure REF.
+Regions of the glacier are separated by thin inlets of ocean.
+With radial basis function interpolation, PSF samples on one side of the inlet inappropriately influence the interpolation on the other side of the inlet.
+With Poisson interpolation, the PSF samples only influence the interpolation on their own side of the inlet.
+
+In Section SEC, we will show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$.
+Briefly, constructing all $q$ of the functions $w_i$ requires solving the Poisson equation $2q$ times with different right hand side sources, and these Poisson PDE solves can be performed efficiently with multigrid.
+Radial basis function interpolation of $\varphi_x(y)$ is cheaper per sample point because radial basis functions are given by analytic formulas.
+However, the cost of these Poisson PDE solves is typically cheap compared to the cost of computing the impulse response samples, $\varphi_{x_i}$.
+Computing the impulse response samples, $\varphi_{x_i}$, requires applying the Hessian to vectors, which in turn requires performing PDE solves of the linearized forward and ajoint equations for the inverse problem.
+These forward and adjoint PDE solves are typically much more expensive than Poisson PDE solves, because the Poisson equation is well-studied and easy to solve with multigrid.
+
+% For example,
+% \begin{itemize}
+%     \item In inverse problems where the forward model is time-dependent, computing Hessian matrix-vector products requires solving time-dependent PDEs.
+In contrast, the Poisson equation is a stationary PDE.
+%     \item In inverse problems where the parameter being inverted for resides on a boundary, computing Hessian matrix-vector products requires solving PDEs that are defined on a higher dimensional geometric domain than the Poisson PDEs that must be solved for the interpolation.
+%     \item In multi-experiment inverse problems, data from many different ``experiments'' on the same medium are used in the inversion.
+Performing a Hessian matrix-vector product requires solving two PDEs \emph{per experiment}.
+The number of Poisson solves used to construct the weighting functions does not depend on the number of experiments.
+% \end{itemize}
+% The additional cost of Poisson interpolation is, therefore, a worthwhile investment if it allows one to achieve the same level of accuracy as other interpolation methods but with less sample points.
+
+% One may also show that the weighting functions $w_i$ sum to one pointwise.
+
+\subsection{Computation of Poisson weighting functions}
+
+After discretization, optimization problem \eqref{eq:wi_optimization_problem} becomes
+\begin{equation}
+\label{eq:discrete_poisson_optimization_1}
+\begin{aligned}
+\min_{\mathbf{w}} &\quad \frac{1}{2}\mathbf{w}^T K^T M^{-1} K \mathbf{w} \\
+\text{such that} &\quad E \mathbf{w} = \mathbf{d}.
+\end{aligned}
+\end{equation}
+Here $\mathbf{w}$ is the discretization of $w_i$, $K$ is the stiffness matrix for the discretized Poisson problem with pure Neumann boundary conditions, $M$ is the mass matrix, and $\mathbf{d}=(0,\dots,0,1,0,\dots,0)$ is the length-$q$ vector with $i^\text{th}$ entry one and all other entries zero.
+Furthermore, $E$ is the pointwise observation matrix defined by $\left(E\mathbf{u}\right)_j = u(x_j)$, where $u$ denotes the function represented by the vector $\mathbf{u}$.
+Here we dropped the subscripts $i$ from variables for notational convenience.
+One must solve $q$ optimization problems of the form \eqref{eq:discrete_poisson_optimization_1}; one for each $i=1,\dots,q$.
+
+In the remainder of this section we will derive an efficient algorithm for solving \eqref{eq:discrete_poisson_optimization_1}.
+The algorithm (Algorithm REF) is based on explicit formula for the solution of \eqref{eq:discrete_poisson_optimization_1} which we present in Proposition \ref{prop:poisson_interpolation_formula}.
+The formulas in Proposition \ref{prop:poisson_interpolation_formula} rely on certain matrices that we define in Definition \ref{defn:poisson_interpolation_matrices}.
+The primary computational cost of the algorithm is the solution of Poisson problems with pure Neumann boundary conditions.
+
+\begin{defn}
+	\label{defn:poisson_interpolation_matrices}
+	Let 
+	\begin{align*}
+		\Psi :=& K^+ E^T, \\
+		\Theta :=& K^+ M \Psi \\
+		S :=& \Psi^T M \Psi,
+	\end{align*}
+	where $K^+$ is the Moore-Penrose pseudoinverse of $K$.
+Further, let $\mathbf{c}$ dente the discretization of the constant function $c(x)=1$ on $\Omega$, and let $\mathbf{1}:=(1,\dots,1)$ denote the vector of length $q$ with all entries equal to one.
+\end{defn}
+
+\begin{remark}
+	Computing $\mathbf{g} = K^+\mathbf{f}$ is equivalent to solving the discretized Poisson problem with $f$ as the right hand side source, with pure Neumann boundary conditions, and with the constraint that the solution has average value equal to zero.
+That is,
+	\begin{equation*}
+		\mathbf{g} = K^+ \mathbf{f} \quad \Leftrightarrow \quad 
+		\begin{cases}
+			\Delta g = f, & \text{in }\Omega,\\
+			\nu \cdot \nabla g = 0, & \text{on }\partial \Omega,\\
+			\int_\Omega g(x) dx = 0,
+		\end{cases}
+	\end{equation*}
+	where $\nu$ denotes the normal to the boundary.
+
+	Hence, the $i^\text{th}$ column of $\Psi$ is the result of solving the discretized pure Neumann Poisson problem, with a point source at location $x_i$.
+The $i^\text{th}$ column of $\Theta$ is the result of solving the discretized pure Neumann Poisson problem with distributed source given by the $i^\text{th}$ column of $\Psi$.
+\end{remark}
+
+\begin{prop}
+	\label{prop:poisson_interpolation_formula}
+	The solution to optimization problem \eqref{eq:discrete_poisson_optimization_1} is given by
+	\begin{equation*}
+		\mathbf{w} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c},
+	\end{equation*}
+	where
+	\begin{align*}
+		\alpha =& \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{d} \\
+		\boldsymbol{\lambda} =& -S^{-1} \left(\mathbf{d} - \alpha \mathbf{1} \right). \\
+	\end{align*}
+\end{prop}
+
+\begin{proof}
+	Let
+	\begin{equation*}
+		\mathbf{v} := K \mathbf{w}.
+	\end{equation*}
+	Since the null-space of the Laplacian with pure Neumann boundary conditions is the set of all constant functions, we have
+	\begin{equation}
+	\label{eq:definition_of_w}
+	\mathbf{w} = K^+ \mathbf{v} + \alpha \mathbf{c}
+	\end{equation}
+	for some scalar $\alpha$.
+
+	Writing optimization problem \eqref{eq:discrete_poisson_optimization_1} in terms of $\mathbf{v}$ and $\alpha$ instead of $\mathbf{w}$ yields the following equivalent optimization problem:
+	\begin{equation}
+	\label{eq:discrete_poisson_optimization_2}
+	\begin{aligned}
+	\min_{\mathbf{v}, \alpha} &\quad \frac{1}{2}\mathbf{v}^T M^{-1} \mathbf{v} \\
+	\text{such that} &\quad EK^+ \mathbf{v} + \alpha \mathbf{1} = \mathbf{d}.
+	\end{aligned}
+	\end{equation}
+	The Lagrangian for optimization problem \eqref{eq:discrete_poisson_optimization_2} is
+	\begin{equation*}
+		\mathcal{L} = \frac{1}{2}\mathbf{v}^T M^{-1} \mathbf{v} + \boldsymbol{\lambda}^T\left(EK^+ \mathbf{v} + \alpha \mathbf{1} - \mathbf{d}\right).
+	\end{equation*}
+	The solution to this optimization problem is the stationary point of the Lagrangian, i.e., the point at which the gradient of the Lagrangian is zero:
+	\begin{equation*}
+		0 = \nabla \mathcal{L} = 
+		\begin{bmatrix}
+			\nabla_\mathbf{v} \mathcal{L} \\
+			\nabla_{\boldsymbol{\lambda}} \mathcal{L} \\
+			\nabla_\alpha \mathcal{L}
+		\end{bmatrix} 
+		= \begin{bmatrix}
+			M^{-1} \mathbf{v} + K^+ E^T \boldsymbol{\lambda} \\
+			EK^+\mathbf{v} + \alpha \mathbf{1} - \mathbf{d} \\
+			\mathbf{1}^T \boldsymbol{\lambda}
+		\end{bmatrix}.
+	\end{equation*}
+	This equation may be rewritten as the following block $3 \times 3$ linear system:
+	\begin{equation*}
+		\begin{bmatrix}
+			M^{-1} & K^+ E^T & 0 \\
+			EK^+ & 0 & \mathbf{1} \\
+			0 & \mathbf{1}^T & 0
+		\end{bmatrix}
+		\begin{bmatrix}
+			\mathbf{v} \\ \boldsymbol{\lambda} \\ \alpha
+		\end{bmatrix}
+		=
+		\begin{bmatrix}
+			0 \\ \mathbf{d} \\ 0
+		\end{bmatrix}.
+	\end{equation*}
+	Performing block Gaussian elimination on this system allows us to reduce the system to the following block triangular form
+	\begin{equation*}
+		\begin{bmatrix}
+			M^{-1} & K^+ E^T & 0 \\
+			& -S & \mathbf{1} \\
+			0 & 0 & \mathbf{1}^T S^{-1} \mathbf{1}
+		\end{bmatrix}
+		\begin{bmatrix}
+			\mathbf{v} \\ \boldsymbol{\lambda} \\ \alpha
+		\end{bmatrix}
+		=
+		\begin{bmatrix}
+			0 \\ \mathbf{y} \\ \mathbf{1}^T S^{-1} y
+		\end{bmatrix}
+	\end{equation*}
+	where we recall that $S := \Psi^T M \Psi = EK^+ M K^+ E^T$. From this block triangular system, we read off the solution as:
+	\begin{align*}
+		\alpha =& \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{d} \\
+		\boldsymbol{\lambda} =& -S^{-1} \left(\mathbf{d} - \alpha \mathbf{1} \right) \\
+		\mathbf{v} =& -M K^+ E^T \boldsymbol{\lambda}
+	\end{align*}
+	Substituting these results into the definition of $\mathbf{w}$ in \eqref{eq:definition_of_w}, we have
+	\begin{equation*}
+		\mathbf{w} = -K^+ M K^+ E^T \boldsymbol{\lambda} + \alpha \mathbf{c} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c},
+	\end{equation*}
+	as required.
+\end{proof}
+
+We now design an incremental algorithm for constructing interpolants $\mathbf{u}$, in which we can add points $x_i$ more efficiently. Let
+\begin{align*}
+	\Psi :=& K^+ E^T \\
+	\Theta :=& K^+ M \Psi
+\end{align*}
+so that 
+\begin{equation*}
+	S = \Psi^T M \Psi
+\end{equation*}
+and
+\begin{equation*}
+	\mathbf{u} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c}
+\end{equation*}
+The columns of $\Psi$ are the responses of $A^+$ to point sources (delta distributions) centered at the sample points $x_i$.
+The columns of $\Theta$ are the responses of $A^+M^{-1} A$ to these point sources.
+The matrices $\Psi$ and $\Theta$ may be constructed incrementally as new points are added.
+
+The solution to this system of equations is found by the following process:
+\begin{enumerate}
+	\item Compute 
+	\begin{equation*}
+		\Psi := A^+ B^T
+	\end{equation*}
+	by solving Neumann poisson problems for point source right hand sides, with point sources located at the points $x_i$, $i=1,\dots,q$.
+	\item Compute 
+	\begin{equation*}
+		\Theta := A^+ M \Psi
+	\end{equation*}
+	by solving Neumann Poisson problems with the columns of $\Psi$ as distributed sources.
+	\item Form 
+	\begin{equation*}
+		S = \Psi^T M \Psi
+	\end{equation*}.
+	\item Compute \begin{equation*}
+		\alpha = \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{y}.
+	\end{equation*}
+	\item Compute 
+	\begin{equation*}
+		\boldsymbol{\lambda} = -S^{-1} \left(\mathbf{y} - \alpha \mathbf{1} \right).
+	\end{equation*}
+	\item Compute
+	\begin{equation*}
+		\mathbf{u} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c}
+	\end{equation*}
+\end{enumerate}
+
+\begin{thm}
+	The weighting functions form a partition of unity.
+\end{thm}
+
+\section{Interpolating impulse response samples old}
 

### 2020-12-04 paper/localpsf.tex
did many good edits to the intro

@@
 
-In this paper, we present a fast matrix free method for approximating locally translation-invariant operators, $\mathcal{A}$, that have locally supported non-negative integral kernels.
+We present a fast matrix free method for approximating locally translation-invariant operators, $\mathcal{A}$, that have locally supported non-negative integral kernels.
 Such operators arise, for example, as Schur complements in reduced space methods for solving partial differential equations (PDEs), and as Hessians in distributed parameter PDE-constrained inverse problems.
@@
 \item[Matrix-free:] One cannot easily access kernel entries $A(y,x)$.
-Access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$.
-\item[Non-negative:] $A(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$.
+Access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors.
+That is, evaluation of the functions
+\begin{equation*}
+	u \mapsto\mathcal{A}u \quad \text{and} \quad v \mapsto\mathcal{A}^Tv.
+\end{equation*} 
+\item[Non-negative kernel:] For all $(y,x) \in \Omega \times \Omega$, we have
+\begin{equation*}
+	A(y,x) \ge 0.
+\end{equation*}
 \item[Local support:] $A(y, x)$ is supported near $y=x$.
-\item[Local translation invariance:] $A(y+h, x+h) \approx A(y,x)$
-if $h$ is not too large.
+\item[Local translation invariance:] If $h$ is not too large, then
+\begin{equation}
+	\label{eq:local_translation_invariance}
+	A(y+h, x+h) \approx A(y,x).
+\end{equation}
 \end{description}
-
-If $\mathcal{A}$ were translation invariant (that is, if $A(y+h, x+h) = A(y,x)$ for all $x$, $y$, $h$) then it is straightforward to show that $\mathcal{A}$ would be a convolution operator, and the convolution kernel, $\varphi_x$, would be given by
+Local translation invariance is illustrated in Figure FIG. 
+%We qualify the degree of local translation invariance with the words ``perfectly'', ``highly'', and ``poorly''.
+%\begin{itemize}
+%\item If equality holds in \eqref{eq:local_translation_invariance} for all $x$, $y$, and $h$, then we say $\mathcal{A}$ is \emph{perfectly translation invariant}. 
+%\item If the discrepancy between left hand side and right hand side of \eqref{eq:local_translation_invariance} is small for a large set of points $x$, $y$, and $h$, then we say that $\mathcal{A}$ is \emph{highly locally translation invariant}.
+%\item If the discrepancy between left hand side and right hand side of \eqref{eq:local_translation_invariance} is large, or if the discrepancy is only small for a small set of $x$, $y$, and $h$, then we say that $\mathcal{A}$ is \emph{poorly locally translation invariant}.
+%\end{itemize} 
+If $\mathcal{A}$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then it is straightforward to show that $\mathcal{A}$ would be a convolution operator, and the convolution kernel, $\varphi_x$, would be given by
 \begin{equation}
@@
 \end{equation}
-The function $\varphi_x$ is the result of applying $\mathcal{A}$ to a point source at some fixed point $x$, then translating the resulting impulse response function re-center it at zero instead of $x$.
+The function $\varphi_x$ is the result of applying $\mathcal{A}$ to a point source at some fixed point $x \in \Omega$, then translating the resulting impulse response function re-center it at zero instead of $x$.
 Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes.
@@
 \end{equation}
-here $f \cdot g$ denotes pointwise multiplication, and $f \ast g$ denotes convolution of the functions $f$ and $g$. Here
+where $f \cdot g$ denotes pointwise multiplication, and $f \ast g$ denotes convolution of the functions $f$ and $g$. Here
 \begin{itemize}
@@
 Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution.
-The more locally translation invariant an operator is, the smaller the number of terms that are required in \eqref{eq:product_convolution} to achieve an accurate approximation.
-
-Let $\widetilde{A}$ denote the integral kernel associated with $\widetilde{\mathcal{A}}$. While $A(y,x)$ is not easily computable, $\widetilde{A}(y,x)$ is given by
+The more locally translation invariant an operator is (i.e., the smaller the discrepancy between the left hand side and right hand side in \eqref{eq:local_translation_invariance}), the smaller the number of terms that are required in \eqref{eq:product_convolution} to achieve an accurate product-convolution approximation.
+
+The core content of this paper is specifying (a) how we choose the points, $x_i$, and compute the convolution kernels, $\varphi_i$ (Section \ref{sec:choosing_sample_points}), and (b) what weighting functions $w_i$ that we use (Section \ref{sec:weighting_functions}).
+Briefly, we compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb'').
+That is, we compute
+\begin{equation*}
+	\eta = \mathcal{A} \left(\sum_{i=1}^m \delta_{x_i}\right).
+\end{equation*}
+The batch of points $x_i$ are chosen so that the support of $\mathcal{A}\delta_{x_i}$ and the support of $\mathcal{A}\delta_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$.
+Since the impulse responses do not overlap, we may post-process $\eta$ to recover $\varphi_i$ for all of the points $x_i$ in the batch---by applying $\mathcal{A}$ to one vector, we recover many $\varphi_i$.
+This is illustrated in Figure FIG.
+The process is repeated to get more batches of $\varphi_i$'s, until the approximation is sufficiently accurate or a maximum number of sample points is reached.
+The supports of the impulse responses $\mathcal{A}\delta_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions.
+This procedure for estimating the support of $\mathcal{A}\delta_x$ a-priori is the only part of the paper that requires non-negativity of the integral kernel.
+The functions $w_i$ are the defined as the smoothest functions, in a certain least-squares sense, that interpolate the points $x_i$.
+
+%A small number of terms in \eqref{eq:product_convolution} are required to accurately approximate a highly locally translation invariant operator, while a large number of terms are required to accurately approximate a poorly locally translation invariant operator.
+
+Let $\widetilde{A}$ denote the integral kernel associated with $\widetilde{\mathcal{A}}$. While $A(y,x)$ is not easily computable, $\widetilde{A}(y,x)$ is given by the formula
 \begin{equation}
@@
 \end{equation}
-Formula \eqref{eq:kernel_entries} provides fast access to kernel entries $\widetilde{A}(y,x)$, which allows us to construct a hierarchical matrix (H-matrix) representation of $\widetilde{\mathcal{A}}$.
-Using this H-matrix representation, one can invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations with $\widetilde{\mathcal{A}}$, using fast H-matrix arithmetic.
-
-For background on product-convolution approximations, we recommend reading the following papers: [CITES]. For background on H-matrices, we recommend reading the following papers: [CITES].
+Formula \eqref{eq:kernel_entries} provides fast access to kernel entries $\widetilde{A}(y,x)$, which allows us to construct a hierarchical matrix (H-matrix) representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method [CITE].
+Once the H-matrix representation is formed, one can invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations with $\widetilde{\mathcal{A}}$, using fast H-matrix arithmetic.
+
+For background on product-convolution approximations, we recommend reading the following papers: [CITES].
+In [CITE], Escande and Weiss explored the idea of approximating linear operators from scattered impulse responses using a product-convolution scheme of the same form as \eqref{eq:product_convolution}.
+In our previous work [CITE MY PAPER], we choose the point locations with an adaptive grid.
+However,
 
@@
 We call operators satisfying these properties \emph{smoothly varying blur operators}.
+
+% Computing the functions $w_i$ requires solving two Poisson PDEs for each point $x_i$, and these Poisson PDEs may be solved cheaply with multigrid.
+
+%Specifically, since the integral kernel for $\mathcal{A}$ is non-negative, $\langle\mathcal{A},\delta_x\rangle^*$ may be interpreted as a scaled probability distribution.
+We compute the mean and covariance of this probability distribution for all points $x \in \Omega$ simultaneously, by applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions.
+We then estimate the support of $\langle\mathcal{A},\delta_x\rangle^*$ to be contained within a small number of standard deviations of the mean of this probability distribution.
+
+%We interpret $\langle\mathcal{A},\delta_{x_i}\rangle^*$ as a scaled probability distribution, and estimate the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ to be contained within a small number of standard deviations of the mean of this probability distribution (this is why we require a non-negative kernel).
+The mean and covariance of $\langle\mathcal{A},\delta_{x}\rangle^*$ can be simultaneously computed for all points $x \in \Omega$ by applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions.
+
+The functions $w_i$ are the defined as the smoothest functions, in a certain least-squares sense, that interpolate the points $x_i$.
+Computing the functions $w_i$ requires solving two Poisson PDEs for each point $x_i$, and these Poisson PDEs may be solved cheaply with multigrid.
 
@@
 \section{Poisson PSF interpolation}
+\label{sec:weighting_functions}
 

### 2020-12-04 paper/localpsf.tex
did many good edits to the intro

@@
 \begin{abstract}
-  We present an efficient matrix-free method for approximating locally translation invariant operators with locally supported non-negative integral kernels.
-The idea of the method is to compute impulse responses of the operator at a collection of scattered points, interpolate the computed impulse responses to approximate impulse responses at other arbitrary points.
-We convert the resulting approximation to hierarchical matrix format, which allows us to perform matrix inversion and other matrix operations using fast hierarchical matrix arithmetic.
+  We present an efficient matrix-free method for approximating locally translation invariant operators that have locally supported non-negative integral kernels.
+The idea of the method is to compute impulse responses of the operator at a collection of scattered points, then interpolate these computed impulse responses to approximate impulse responses at other arbitrary points.
+This results in a product-convolution approximation, which we convert to hierarchical matrix format.
+Once the approximation is represented in hierarchical matrix format, we perform matrix inversion and other matrix operations using fast hierarchical matrix arithmetic.
 Impulse responses are computed by applying the operator to a small number of Dirac comb ``batches'' of point sources.
@@
 \item[Matrix-free:] One cannot easily access kernel entries $A(y,x)$.
-Access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors.
+Access to $\mathcal{A}$ is available only through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors.
 That is, evaluation of the functions
@@
 \end{equation}
-The function $\varphi_x$ is the result of applying $\mathcal{A}$ to a point source at some fixed point $x \in \Omega$, then translating the resulting impulse response function re-center it at zero instead of $x$.
+The function $\varphi_x$ is the result of applying $\mathcal{A}$ to a point source, $\delta_x$, at some fixed point $x \in \Omega$, then translating the resulting impulse response function, $\mathcal{A}\delta_x$, to re-center it at zero instead of $x$.
 Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes.
@@
 \begin{itemize}
-	\item $\varphi_i := \varphi_{x_i}$ is given by formula \eqref{eq:convolution_kernel}, but with $x$ replaced by $x_i$.
+	\item The local convolution kernels, $\varphi_i := \varphi_{x_i}$, are given by formula \eqref{eq:convolution_kernel}, but with $x$ replaced by $x_i$.
 	\item $\{x_i\}_{i=1}^r \subset \Omega$ is a collection of points scattered throughout the domain.
@@
 
-The core content of this paper is specifying (a) how we choose the points, $x_i$, and compute the convolution kernels, $\varphi_i$ (Section \ref{sec:choosing_sample_points}), and (b) what weighting functions $w_i$ that we use (Section \ref{sec:weighting_functions}).
+The core content of this paper is specifying (a) how we choose the points, $x_i$, (b) how we compute the convolution kernels, $\varphi_i$, and (c) what weighting functions $w_i$ that we use.
 Briefly, we compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb'').

### 2020-12-04 paper/localpsf.tex
did many good edits to the intro

@@
 
-Let $\widetilde{A}$ denote the integral kernel associated with $\widetilde{\mathcal{A}}$. While $A(y,x)$ is not easily computable, $\widetilde{A}(y,x)$ is given by the formula
+Once the $\varphi_i$ and $w_i$ are computed, we convert $\widetilde{\mathcal{A}}$ to hierarchical matrix (H-matrix) format.
+Let $\widetilde{A}$ denote the integral kernel associated with $\widetilde{\mathcal{A}}$.
+While $A(y,x)$ is not easily computable, $\widetilde{A}(y,x)$ is given by the formula
 \begin{equation}
@@
 \end{equation}
-Formula \eqref{eq:kernel_entries} provides fast access to kernel entries $\widetilde{A}(y,x)$, which allows us to construct a hierarchical matrix (H-matrix) representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method [CITE].
-Once the H-matrix representation is formed, one can invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations with $\widetilde{\mathcal{A}}$, using fast H-matrix arithmetic.
+Formula \eqref{eq:kernel_entries} provides fast access to kernel entries $\widetilde{A}(y,x)$, which allows us to construct a H-matrix representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method [CITE].
+Once in H-matrix format, fast H-matrix arithmetic can be used to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations.
 

### 2020-12-05 paper/localpsf.tex
edited 3 slides. Edited paper intro more

@@
 Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes.
-Locally translation invariant operators may therefore be approximated by spatially varying weighted sums of convolution operators of the form
+We therefore approximate $\mathcal{A}$ by a spatially varying weighted sum of convolution operators of the form
 \begin{equation}
@@
 	\item $\{x_i\}_{i=1}^r \subset \Omega$ is a collection of points scattered throughout the domain.
-	\item The functions $w_i(x)$ are spatially varying weighting function that interpolate the convolution kernels $\varphi_i$.
+	\item The functions $w_i(x)$ are spatially varying weighting functions that interpolate the convolution kernels $\varphi_i$.
 \end{itemize}
@@
 
-The core content of this paper is specifying (a) how we choose the points, $x_i$, (b) how we compute the convolution kernels, $\varphi_i$, and (c) what weighting functions $w_i$ that we use.
-Briefly, we compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb'').
-That is, we compute
+Our approximation is defined by:
+\begin{itemize}
+	\item how we compute the convolution kernels, $\varphi_i$, (Section \ref{sec:get_impulse_response})
+	\item how we choose the points, $x_i$, (Section \ref{sec:choosing_sample_points}) and 
+	\item what weighting functions $w_i$ that we use (Section \ref{sec:weighting_functions}). 
+\end{itemize}
+To briefly summarize, we compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb''). That is, we compute
 \begin{equation*}
@@
 The process is repeated to get more batches of $\varphi_i$'s, until the approximation is sufficiently accurate or a maximum number of sample points is reached.
-The supports of the impulse responses $\mathcal{A}\delta_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions.
-This procedure for estimating the support of $\mathcal{A}\delta_x$ a-priori is the only part of the paper that requires non-negativity of the integral kernel.
-The functions $w_i$ are the defined as the smoothest functions, in a certain least-squares sense, that interpolate the points $x_i$.
+The supports of the impulse responses $\mathcal{A}\delta_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure (described in Section \ref{eq:mean_and_covariance_estimation}) that involves applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions.
+This procedure is the only part of the paper that requires non-negativity of the integral kernel.
+The functions $w_i$ are the smoothest functions, in a least-squares sense, that interpolate the points $x_i$.
+Computing the functions $w_i$ require solving two Poisson PDEs per weighting function, and these PDE solves are done cheaply using multigrid.
 
@@
 \subsection{Mean and covariance estimation}
+\label{eq:mean_and_covariance_estimation}
 

### 2020-12-05 paper/localpsf.tex
dirac comb dual notation may be confusing, but is needed for correctness

@@
 \begin{equation*}
-	\eta = \mathcal{A} \left(\sum_{i=1}^m \delta_{x_i}\right).
-\end{equation*}
-The batch of points $x_i$ are chosen so that the support of $\mathcal{A}\delta_{x_i}$ and the support of $\mathcal{A}\delta_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$.
+	\eta = \sum_{i=1}^m A(~\cdot~, x_i) = \left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*.
+\end{equation*}
+Here $\sum_{i=1}^m \delta_{x_i}$ is the Dirac comb distribution, and $\left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*$ denotes the Riesz representation of the application of $\mathcal{A}$ to the Dirac comb.
+In computations, this amounts to computing $\boldsymbol{\eta} = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \left(\sum_{i=1}^m \boldsymbol{\delta}_{x_i}\right)$, where $\boldsymbol{\eta}$, $\mathbf{A}$, and $\boldsymbol{\delta}_{x_i}$ are discretized versions of $\eta$, $A$, and $\delta_{x_i}$, respectively, and $\mathbf{M}$ is a mass matrix.
+The batch of points $x_i$ are chosen so that the support of $\left\langle \mathcal{A}, \delta_{x_i}\right\rangle^*$ and the support of $\left\langle \mathcal{A}, \delta_{x_j}\right\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$.
 Since the impulse responses do not overlap, we may post-process $\eta$ to recover $\varphi_i$ for all of the points $x_i$ in the batch---by applying $\mathcal{A}$ to one vector, we recover many $\varphi_i$.
@@
 The process is repeated to get more batches of $\varphi_i$'s, until the approximation is sufficiently accurate or a maximum number of sample points is reached.
-The supports of the impulse responses $\mathcal{A}\delta_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure (described in Section \ref{eq:mean_and_covariance_estimation}) that involves applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions.
+The supports of the impulse responses $\left\langle \mathcal{A}, \delta_{x}\right\rangle^*$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure (described in Section \ref{eq:mean_and_covariance_estimation}) that involves applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions.
 This procedure is the only part of the paper that requires non-negativity of the integral kernel.
@@
 \end{equation}
-Formula \eqref{eq:kernel_entries} provides fast access to kernel entries $\widetilde{A}(y,x)$, which allows us to construct a H-matrix representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method [CITE].
+Formula \eqref{eq:kernel_entries} provides fast access to kernel entries $\widetilde{A}(y,x)$.
+This allows us to construct a H-matrix representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks [CITE].
 Once in H-matrix format, fast H-matrix arithmetic can be used to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations.
@@
 
-\subsection{Operator access levels}
-
-%In order to solve inverse or optimization problems governed by PDEs, we need to solve linear systems with the Hessian as the coefficient matrix [cite cite].
+%\subsection{Operator access levels}
+%
+%
+%%In order to solve inverse or optimization problems governed by PDEs, we need to solve linear systems with the Hessian as the coefficient matrix [cite cite].
 In order to do this efficiently, we have access to only Hessian-vector products.
@@
 Therefore
-
-A linear operator may be classified by the way in which we may access the operator.
-
-\begin{itemize}
-	\item \emph{Level 1:} A matrix representation of the operator, $\mathbf{A}$, is stored in memory in dense or sparse format.
-	\item \emph{Level 2:} We can efficiently compute arbitrary matrix entries $\mathbf{A}_{ij}$ at will, but the matrix $\mathbf{A}$ is not stored in memory.
-	\item \emph{Level 3:} Access to $\mathbf{A}$ is only available through an algorithm that computes matrix-vector products $\mathbf{u} \mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{u} \mapsto \mathbf{A}^T \mathbf{u}$.
-\end{itemize}
-In large-scale scientific computing,
-
-As discussed in Section \ref{sec:PDE_hessian_motivation}, in inverse problems governed by PDEs, matrix entries of the Hessian are not easily computable since getting one entry would require computing an entire column of the Hessian, then throwing away all but one entry in the column.
+%
+%A linear operator may be classified by the way in which we may access the operator. 
+%
+%
+%\begin{itemize}
+%	\item \emph{Level 1:} A matrix representation of the operator, $\mathbf{A}$, is stored in memory in dense or sparse format.
+%	\item \emph{Level 2:} We can efficiently compute arbitrary matrix entries $\mathbf{A}_{ij}$ at will, but the matrix $\mathbf{A}$ is not stored in memory.
+%	\item \emph{Level 3:} Access to $\mathbf{A}$ is only available through an algorithm that computes matrix-vector products $\mathbf{u} \mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{u} \mapsto \mathbf{A}^T \mathbf{u}$.
+%\end{itemize}
+%In large-scale scientific computing, 
+%
+%
+%As discussed in Section \ref{sec:PDE_hessian_motivation}, in inverse problems governed by PDEs, matrix entries of the Hessian are not easily computable since getting one entry would require computing an entire column of the Hessian, then throwing away all but one entry in the column.
 Computing a single column costs at least two PDE solves.
 Many such entries are needed to approximate the Hessian in inverse problems with highly informative data (i.e., for Hessians with large numerical rank) [cite cite].
-
- of the implicit definition of $u(m)$.
-
-Hierarchical matrices (H, H2, HSS, HBS, HODLR), quantized tensor trains, hierarchical interpolative factorization, STRUMPACK (LBNL),
-
-Level 1: not relevant in large-scale problems
-State clearly: we do not use Level 1 at all.
-This paper: Level 3 -> Level 2
-H-matrix: solve on level 2
-
-%\subsection{Existing work}
-%
-%
-%\begin{itemize}
-%	\item CITE ESCANDE, others, for a reviews of the subject.
-%	\item Resolution analysis seismic papers: try to figure out size of $\varphi_x$ by matvec with random function, post-processing to get autocorrelation length.
+%
+% of the implicit definition of $u(m)$. 
+%
+%Hierarchical matrices (H, H2, HSS, HBS, HODLR), quantized tensor trains, hierarchical interpolative factorization, STRUMPACK (LBNL), 
+%
+%Level 1: not relevant in large-scale problems
+%State clearly: we do not use Level 1 at all.
+%This paper: Level 3 -> Level 2
+%H-matrix: solve on level 2
+%
+%%\subsection{Existing work}
+%%
+%%
+%%\begin{itemize}
+%%	\item CITE ESCANDE, others, for a reviews of the subject.
+%%	\item Resolution analysis seismic papers: try to figure out size of $\varphi_x$ by matvec with random function, post-processing to get autocorrelation length.
 They did the constant matvec volume thing.
@@
 They do not build preconditioner
-%	\item Escande Weiss scattered response paper.
+%%	\item Escande Weiss scattered response paper.
 They interpolate scattered impulse responses using radial basis functions and study how the geometry of the points affects the approximation accuracy.
@@
 The applications they consider one has direct access to the impulse response, e.g., through formula for the kernel.
-%	\item Imaging problems: ``matvec'' requires performing an experiment
-%	\item pseudodifferential operator equivalence
-%	\item Laundry list of applications
-%	\item scattered impulse response of escande weiss paper: radial basis functions for approximation in $y$. How to determine if the impulse responses are far enough away from each other?
-%	\item my convolution paper: compute impulse responses one at a time, one impulse response per matvec: too slow! Here we compute many impulse responses using only 1 matvec.
-%	\item uniformity of smoothness condition
-%\end{itemize}
+%%	\item Imaging problems: ``matvec'' requires performing an experiment
+%%	\item pseudodifferential operator equivalence
+%%	\item Laundry list of applications
+%%	\item scattered impulse response of escande weiss paper: radial basis functions for approximation in $y$. How to determine if the impulse responses are far enough away from each other?
+%%	\item my convolution paper: compute impulse responses one at a time, one impulse response per matvec: too slow! Here we compute many impulse responses using only 1 matvec.
+%%	\item uniformity of smoothness condition
+%%\end{itemize}
 

### 2020-12-05 paper/localpsf.tex
dirac comb dual notation may be confusing, but is needed for correctness

@@
 	\item how we choose the points, $x_i$, (Section \ref{sec:choosing_sample_points}) and 
-	\item what weighting functions $w_i$ that we use (Section \ref{sec:weighting_functions}). 
+	\item what weighting functions, $w_i$, that we use (Section \ref{sec:weighting_functions}). 
 \end{itemize}
-To briefly summarize, we compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb''). That is, we compute
+Briefly, we compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb''). That is, we compute
 \begin{equation*}
@@
 The batch of points $x_i$ are chosen so that the support of $\left\langle \mathcal{A}, \delta_{x_i}\right\rangle^*$ and the support of $\left\langle \mathcal{A}, \delta_{x_j}\right\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$.
-Since the impulse responses do not overlap, we may post-process $\eta$ to recover $\varphi_i$ for all of the points $x_i$ in the batch---by applying $\mathcal{A}$ to one vector, we recover many $\varphi_i$.
+Since the impulse responses do not overlap, we may post-process $\eta$ to recover $\varphi_i$ for all of the points $x_i$ in the batch---by applying $\mathcal{A}$ to one distribution, we recover many $\varphi_i$.
 This is illustrated in Figure FIG.

### 2020-12-05 paper/localpsf.tex
more edits to intro

@@
 
-Our approximation is defined by:
-\begin{itemize}
-	\item how we compute the convolution kernels, $\varphi_i$, (Section \ref{sec:get_impulse_response})
-	\item how we choose the points, $x_i$, (Section \ref{sec:choosing_sample_points}) and 
-	\item what weighting functions, $w_i$, that we use (Section \ref{sec:weighting_functions}). 
-\end{itemize}
+Our approximation is defined by how we compute the convolution kernels, $\varphi_i$, (Section \ref{sec:get_impulse_response}) how we choose the points, $x_i$, (Section \ref{sec:choosing_sample_points}) and what weighting functions, $w_i$, that we use (Section \ref{sec:weighting_functions}).
 Briefly, we compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb''). That is, we compute
 \begin{equation*}
-	\eta = \sum_{i=1}^m A(~\cdot~, x_i) = \left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*.
+	\eta := \sum_{i=1}^m A(~\cdot~, x_i) = \left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*.
 \end{equation*}
@@
 The process is repeated to get more batches of $\varphi_i$'s, until the approximation is sufficiently accurate or a maximum number of sample points is reached.
-The supports of the impulse responses $\left\langle \mathcal{A}, \delta_{x}\right\rangle^*$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure (described in Section \ref{eq:mean_and_covariance_estimation}) that involves applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions.
+In order to choose the points $x_i$, we need to estimate the support of $\left\langle \mathcal{A}, \delta_{x_i}\right\rangle^*$ \emph{before} we compute it.
+The supports of the impulse responses $\left\langle \mathcal{A}, \delta_{x}\right\rangle^*$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions (Section \ref{eq:mean_and_covariance_estimation}).
 This procedure is the only part of the paper that requires non-negativity of the integral kernel.
-The functions $w_i$ are the smoothest functions, in a least-squares sense, that interpolate the points $x_i$.
-Computing the functions $w_i$ require solving two Poisson PDEs per weighting function, and these PDE solves are done cheaply using multigrid.
+We choose the functions $w_i$ to be the smoothest functions, in a least-squares sense, that interpolate the points $x_i$.
+Computing the functions $w_i$ require solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid.
 
@@
 \end{equation}
-Formula \eqref{eq:kernel_entries} provides fast access to kernel entries $\widetilde{A}(y,x)$.
-This allows us to construct a H-matrix representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks [CITE].
-Once in H-matrix format, fast H-matrix arithmetic can be used to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations.
+Formula \eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks [CITE].
+Once in H-matrix format, fast H-matrix arithmetic can is used to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations.
 

### 2020-12-06 paper/localpsf.tex
more edits to intro

@@
 
-Our approximation is defined by how we compute the convolution kernels, $\varphi_i$, (Section \ref{sec:get_impulse_response}) how we choose the points, $x_i$, (Section \ref{sec:choosing_sample_points}) and what weighting functions, $w_i$, that we use (Section \ref{sec:weighting_functions}).
+\subsection{Overview of the method}
+
+Our approximation method is defined by how we compute the convolution kernels, $\varphi_i$, (Section \ref{sec:get_impulse_response}) how we choose the points, $x_i$, (Section \ref{sec:choosing_sample_points}) and what weighting functions, $w_i$, that we use (Section \ref{sec:weighting_functions}).
 Briefly, we compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb''). That is, we compute
@@
 This procedure is the only part of the paper that requires non-negativity of the integral kernel.
-We choose the functions $w_i$ to be the smoothest functions, in a least-squares sense, that interpolate the points $x_i$.
-Computing the functions $w_i$ require solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid.
+
+We choose the weighting functions $w_i$ to be the smoothest functions, in a least-squares sense, that interpolate the points $x_i$.
+Computing the weighting functions $w_i$ requires solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid.
 
@@
 Once in H-matrix format, fast H-matrix arithmetic can is used to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations.
+
+\subsection{Connection to existing methods}
 
@@
 However,
+
+\subsection{old stuff}
 
@@
 
-In this paper we present an efficient matrix-free method for approximating operators $\mathcal{A}$ with a product-convolution approximation, $\widetilde{A}$, of the form
-\begin{equation*}
-	\left(\mathcal{A}u\right) \approx \left(\widetilde{\mathcal{A}}u\right) := \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right)
-\end{equation*}
-
-PSFs that satisfy the following properties:
-\begin{enumerate}
-	\item $\varphi_x$ varies smoothly as a function of $x$.
-	\item $\varphi_x(y)$ is zero (or small) when $|y|$ is large.
-	\item $\varphi_x$ is a non-negative (or mostly non-negative) function.
-\end{enumerate}
-If $\varphi_x$ was constant in $x$, then $\mathcal{A}$ would be the convolution operator $\left(\mathcal{A}v\right)^* = \varphi_0 \ast v$.
+%In this paper we present an efficient matrix-free method for approximating operators $\mathcal{A}$ with a product-convolution approximation, $\widetilde{A}$, of the form
+%\begin{equation*}
+%	\left(\mathcal{A}u\right) \approx \left(\widetilde{\mathcal{A}}u\right) := \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right)
+%\end{equation*}
+%
+%
+%PSFs that satisfy the following properties:
+%\begin{enumerate}
+%	\item $\varphi_x$ varies smoothly as a function of $x$.
+%	\item $\varphi_x(y)$ is zero (or small) when $|y|$ is large.
+%	\item $\varphi_x$ is a non-negative (or mostly non-negative) function.
+%\end{enumerate}
+%If $\varphi_x$ was constant in $x$, then $\mathcal{A}$ would be the convolution operator $\left(\mathcal{A}v\right)^* = \varphi_0 \ast v$.
 Operators satisfying property 1 therefore act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies smoothly as $x$ changes.
@@
 
-The functions $w_i$ are the defined as the smoothest functions, in a certain least-squares sense, that interpolate the points $x_i$.
+%The functions $w_i$ are the defined as the smoothest functions, in a certain least-squares sense, that interpolate the points $x_i$.
 Computing the functions $w_i$ requires solving two Poisson PDEs for each point $x_i$, and these Poisson PDEs may be solved cheaply with multigrid.
 
-The idea of our method is to sample $\varphi_{x}$ at a collection of points $x_i$ by applying $\mathcal{A}$ to a sum of point sources at the points $x_i$ (a ``Dirac comb'').
+%The idea of our method is to sample $\varphi_{x}$ at a collection of points $x_i$ by applying $\mathcal{A}$ to a sum of point sources at the points $x_i$ (a ``Dirac comb'').
 The points $x_i$ are chosen so that the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ and the support of $\langle\mathcal{A},\delta_{x_j}\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$.
@@
 
-Properties 1 and 2 are essential to our method.
+%Properties 1 and 2 are essential to our method.
 Property 3 is only required in the subprocedure for estimating the support of $\langle\mathcal{A},\delta_x\rangle^*$, and may be dropped if other methods for estimating the support are available.

### 2020-12-06 paper/localpsf.tex
more edits to intro

@@
 \label{eq:product_convolution}
-	\left(\mathcal{A}u\right)(v) \approx \left(\widetilde{\mathcal{A}}u\right)(v) := \left\langle v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right\rangle_{L^2(\Omega)},
-\end{equation}
-where $f \cdot g$ denotes pointwise multiplication, and $f \ast g$ denotes convolution of the functions $f$ and $g$. Here
+	\left(\mathcal{A}u\right)(v) \approx \left(\widetilde{\mathcal{A}}u\right)(v) := \left( v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
+\end{equation}
+where $f \cdot g$ denotes pointwise multiplication, $f \ast g$ denotes convolution, and $\left(f, g\right)_{L^2(\Omega)}$ denotes the $L^2$ inner product on $\Omega$, for functions $f,g \in L^2(\Omega)$.
+Here
 \begin{itemize}
@@
 	\item $\{x_i\}_{i=1}^r \subset \Omega$ is a collection of points scattered throughout the domain.
-	\item The functions $w_i(x)$ are spatially varying weighting functions that interpolate the convolution kernels $\varphi_i$.
+	\item The functions $w_i$ are spatially varying weighting functions that interpolate the convolution kernels $\varphi_i$.
 \end{itemize}
@@
 
-Our approximation method is defined by how we compute the convolution kernels, $\varphi_i$, (Section \ref{sec:get_impulse_response}) how we choose the points, $x_i$, (Section \ref{sec:choosing_sample_points}) and what weighting functions, $w_i$, that we use (Section \ref{sec:weighting_functions}).
-Briefly, we compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb''). That is, we compute
+Our operator approximation method is defined by how we compute the convolution kernels, $\varphi_i$, (Section \ref{sec:get_impulse_response}) how we choose the points, $x_i$, (Section \ref{sec:choosing_sample_points}) what weighting functions, $w_i$, that we use (Section \ref{sec:weighting_functions}), and how we use the resulting product-convolution approximation to perform linear algebra operations such as solving linear systems (Section \ref{sec:H_matrix}).
+In this section we summarize the answers to these questions.
+
+We compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb''). That is, we compute
 \begin{equation*}
@@
 Formula \eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks [CITE].
-Once in H-matrix format, fast H-matrix arithmetic can is used to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations.
+Once in H-matrix format, fast H-matrix arithmetic is used to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations.
 

### 2020-12-06 paper/localpsf.tex
more edits to intro

@@
 
-For background on product-convolution approximations, we recommend reading the following papers: [CITES].
+In general, product-convolution approximations are approximations of the form \eqref{eq:product_convolution}, but where $\varphi_i$ and $w_i$ may be arbitrary functions. 
+For background on product-convolution approximations, we recommend reading the following excellent papers: [CITES].
+
 In [CITE], Escande and Weiss explored the idea of approximating linear operators from scattered impulse responses using a product-convolution scheme of the same form as \eqref{eq:product_convolution}.
@@
 However,
+
+Nagy's paper: breaking domain into regular grid, using piecewise linear interpolation.
+
+Use for approximating Hessians in PDE-constrained optimization: Georg's paper, my paper.
+
+Getting many impulse responses at once (batching).
+
+Geometry-aware weighting functions vs radial basis functions
 
@@
 \end{align*}
-For example, the delta distribution $\delta_x$ is defined by $\delta_x(v) = v(x)$, and the action of $\mathcal{A}$ on $\delta_x$ is given by
-\begin{equation*}
-	\langle\mathcal{A},\delta_x\rangle(w)	= \int_\Omega w(y) A(y,x) dy.
-\end{equation*}
+For example, the delta distribution $\delta_x$ is defined by $\delta_x(v) = v(x)$, the action of $\mathcal{A}$ on $\delta_x$ is given by
+\begin{equation*}
+	\langle\mathcal{A},\delta_x\rangle(w)	= \int_\Omega w(y) A(y,x) dy,
+\end{equation*}
+and so the Riesz representation of $\langle\mathcal{A},\delta_x\rangle$ is given by $\langle\mathcal{A},\delta_x\rangle^* = A(~\cdot~,x)$.
+That is, $\langle\mathcal{A},\delta_x\rangle^*$ is the function $y \mapsto A(y,x)$.
 

### 2020-12-06 paper/localpsf.tex
edits to poiwwon weighting functions section

@@
 
-\section{Interpolating impulse response samples}
-
-\section{Poisson PSF interpolation}
+\section{Poisson interpolation weighting functions}
 \label{sec:weighting_functions}
 
-We form an interpolant, $\widetilde{\varphi}_x$, of the point spread function (PSF), $\varphi_x$, as follows:
-\begin{equation*}
-	\varphi_x(y) \approx \widetilde{\varphi}_x(y) := \sum_{i=1}^q w_i(x)\varphi_{x_i}(y).
-\end{equation*}
-The functions $y \mapsto \varphi_{x_i}(y)$, are samples of the PSF at the sample points $x_i$, $i=1,\dots,q$.
-The functions $w_i(x)$, are smooth spatially varying weighting functions that solve the following optimization problem:
+We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to the following optimization problem:
 \begin{equation}
@@
 Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$.
-We refer to $\widetilde{\varphi}_x$ as the \emph{Poisson interpolant} of $\varphi_x$.
-
-Let $u_y$ denote the function $u_y(x) := \widetilde{\varphi}_x(y)$ for an arbitrary but fixed value of $y$. It is straightforward to show that $u_y$ solves the optimization problem
+Since the Laplacian measures local deviation from flatness [cite Jacobsen], the weighting function $w_i$ is the ``flattest possible'' function (in a least-squares sense) that has Neumann boundary conditions and takes the value $1$ at point $x_i$, and zero at the points $x_j$, $j \neq i$.
+In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense.
+
+\begin{thm}
+	\label{thm:smoothest_interpolant}
+Let $u$ be the function
+\begin{equation*}
+u(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right)
+\end{equation*}
+for an arbitrary but fixed value of $y$. Then $u$ solves the optimization problem
 \begin{equation}
@@
 \begin{aligned}
-\min_{u_y} &\quad \frac{1}{2}\norm{\Delta u_y}_{L^2(\Omega)}^2 \\
-\text{such that} &\quad u_y(x_i) = \varphi_{x_i}(y), \quad i=1,\dots,q.
+\min_{u} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
+\text{such that} &\quad u(x_i) = \varphi_i(y), \quad i=1,\dots,q.
 \end{aligned}
 \end{equation}
+\end{thm}
+\begin{proof}
 This follows from linearity of the KKT system associated with \eqref{eq:wi_optimization_problem}, which in turn follows from the fact that the the objective function in \eqref{eq:wi_optimization_problem} is quadratic, and the constraint is linear.
-Since the Laplacian measures local deviation from flatness [cite Jacobsen], from \eqref{eq:uy_optimization_problem} we see that the interpolant $x \mapsto u_y(x) = \widetilde{\varphi}_x(y)$ is the `flattest possible'' function (in a least-squares sense) that has Neumann boundary conditions and interpolates the data $\varphi_{x_i}(y)$.
+\end{proof}
+
+%Since the Laplacian measures local deviation from flatness [cite Jacobsen], from \eqref{eq:uy_optimization_problem} we see that the interpolant $x \mapsto u_y(x) = \widetilde{\varphi}_x(y)$ is the ``flattest possible'' function (in a least-squares sense) that has Neumann boundary conditions and interpolates the data $\varphi_{x_i}(y)$.
 
@@
 With Poisson interpolation, the PSF samples only influence the interpolation on their own side of the inlet.
-
-In Section SEC, we will show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$.
-Briefly, constructing all $q$ of the functions $w_i$ requires solving the Poisson equation $2q$ times with different right hand side sources, and these Poisson PDE solves can be performed efficiently with multigrid.
-Radial basis function interpolation of $\varphi_x(y)$ is cheaper per sample point because radial basis functions are given by analytic formulas.
-However, the cost of these Poisson PDE solves is typically cheap compared to the cost of computing the impulse response samples, $\varphi_{x_i}$.
-Computing the impulse response samples, $\varphi_{x_i}$, requires applying the Hessian to vectors, which in turn requires performing PDE solves of the linearized forward and ajoint equations for the inverse problem.
-These forward and adjoint PDE solves are typically much more expensive than Poisson PDE solves, because the Poisson equation is well-studied and easy to solve with multigrid.
 
@@
 \subsection{Computation of Poisson weighting functions}
+
+In this section, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$.
+Briefly, constructing all $q$ of the functions $w_i$ requires solving the Poisson equation $2q$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid.
+All the existing weighting functions, $\{w_i\}_{i=1}^q$, will change if a new point, $x_{q+1}$, is added.
+We present an efficient incremental algorithm for updating these weighting functions when a new point is added, in which we only need to perform two new Poisson PDE solves (rather than $2(q+1)$ new PDE solves).
+
+%Radial basis function interpolation of $\varphi_x(y)$ is cheaper per sample point because radial basis functions are given by analytic formulas.
+However, the cost of these Poisson PDE solves is typically cheap compared to the cost of computing the impulse response samples, $\varphi_{x_i}$.
+Computing the impulse response samples, $\varphi_{x_i}$, requires applying the Hessian to vectors, which in turn requires performing PDE solves of the linearized forward and ajoint equations for the inverse problem.
+These forward and adjoint PDE solves are typically much more expensive than Poisson PDE solves, because the Poisson equation is well-studied and easy to solve with multigrid.
 

### 2020-12-06 paper/localpsf.tex
edits to poiwwon weighting functions section

@@
 In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense.
+Then in Section \ref{sec:computation_of_weighting_functions}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$.
+Constructing all $q$ of the functions $w_i$ requires solving the Poisson equation $2q$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid.
+All the existing weighting functions, $\{w_i\}_{i=1}^q$, will change if a new point, $x_{q+1}$, is added.
+However, we present an efficient incremental algorithm for updating these weighting functions when a new point is added, so that we only need to perform two new Poisson PDE solves, rather than $2(q+1)$ new PDE solves.
+
+A key advantage of Poisson interpolation over other PSF interpolation methods, most notably radial basis functions [cite Escande etc], is that Poisson interpolation is \emph{geometry-aware}---Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions.
+Consider the Pine island glacier geometry shown in Figure REF.
+Regions of the glacier are separated by thin inlets of ocean.
+With radial basis function interpolation, PSF samples on one side of the inlet inappropriately influence the interpolation on the other side of the inlet.
+With Poisson interpolation, the PSF samples only influence the interpolation on their own side of the inlet.
+
+It is important to note that optimization problems \eqref{eq:wi_optimization_problem} only has a unique continuous solution when the spatial dimension is $1$, $2$, or $3$.
+When inverting for parameter fields defined on $4$ or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used\footnote{The reason for this dependence on spatial dimension is related to the fact that the fundamental solution of the biharmonic equation is continuous in $3$ or less dimensions, but singular in $4$ or more dimensions.
+In the unusual scenario where one is inverting for a parameter field in $4$ or more spatial dimensions, one could use a modified Poisson interpolation where $\Delta$ is replaced with $\Delta^\alpha$ in the objective function of \eqref{eq:wi_optimization_problem}, where $\alpha>1$ is chosen so that the fundamental solution of $\Delta^{2\alpha}$ is continuous.}.
 
@@
 %Since the Laplacian measures local deviation from flatness [cite Jacobsen], from \eqref{eq:uy_optimization_problem} we see that the interpolant $x \mapsto u_y(x) = \widetilde{\varphi}_x(y)$ is the ``flattest possible'' function (in a least-squares sense) that has Neumann boundary conditions and interpolates the data $\varphi_{x_i}(y)$.
-
-Optimization problems \eqref{eq:wi_optimization_problem} and \eqref{eq:uy_optimization_problem} have unique continuous solutions when the spatial dimension is $1$, $2$, or $3$.
-When inverting for parameter fields defined on $4$ or higher spatial dimensions, the solution to this optimization problem does not exist, and Poisson interpolation (as described here) should not be used\footnote{The reason for this dependence on spatial dimension is related to the fact that the fundamental solution of the biharmonic equation is continuous in $3$ or less dimensions, but singular in $4$ or more dimensions.
-In the unusual scenario where one is inverting for a parameter field in $4$ or more spatial dimensions, one could use a modified Poisson interpolation where $\Delta$ is replaced with $\Delta^\alpha$ in the objective function of \eqref{eq:wi_optimization_problem}, where $\alpha>1$ is chosen so that the fundamental solution of $\Delta^{2\alpha}$ is continuous.}.
-
-A key advantage of Poisson interpolation over other PSF interpolation methods, most notably radial basis functions [cite Escande etc], is that Poisson interpolation is \emph{geometry-aware}---Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions.
-Consider the Pine island glacier geometry shown in Figure REF.
-Regions of the glacier are separated by thin inlets of ocean.
-With radial basis function interpolation, PSF samples on one side of the inlet inappropriately influence the interpolation on the other side of the inlet.
-With Poisson interpolation, the PSF samples only influence the interpolation on their own side of the inlet.
 
@@
 \subsection{Computation of Poisson weighting functions}
-
-In this section, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$.
-Briefly, constructing all $q$ of the functions $w_i$ requires solving the Poisson equation $2q$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid.
-All the existing weighting functions, $\{w_i\}_{i=1}^q$, will change if a new point, $x_{q+1}$, is added.
-We present an efficient incremental algorithm for updating these weighting functions when a new point is added, in which we only need to perform two new Poisson PDE solves (rather than $2(q+1)$ new PDE solves).
+\label{sec:computation_of_weighting_functions}
 

### 2020-12-06 paper/localpsf.tex
edits to poiwwon weighting functions section

@@
 
-A key advantage of Poisson interpolation over other PSF interpolation methods, most notably radial basis functions [cite Escande etc], is that Poisson interpolation is \emph{geometry-aware}---Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions.
+A key advantage of Poisson interpolation over other PSF interpolation methods, most notably radial basis functions [cite Escande etc], is that Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions.
 Consider the Pine island glacier geometry shown in Figure REF.

### 2020-12-08 paper/localpsf.tex
writing existing work

@@
 
-\subsection{Connection to existing methods}
-
-In general, product-convolution approximations are approximations of the form \eqref{eq:product_convolution}, but where $\varphi_i$ and $w_i$ may be arbitrary functions. 
+\subsection{Existing work}
+
+In general, product-convolution approximations take the form \eqref{eq:product_convolution}, but $\varphi_i$ and $w_i$ may be arbitrary functions.
+Product-convolution approximations have been used in a wide variety of fields going back several decades [CITE BIG LIST].
 For background on product-convolution approximations, we recommend reading the following excellent papers: [CITES].
+
+We build upon a line of work in which the convolution kernels, $\varphi_i$, are impulse responses of $\mathcal{A}$ to point sources [CITE BIG LIST].
+The primary question in methods of this type are how to choose the points $x_i$, and how to choose the weighting functions $w_i$.
+In [CITE NAGY], the impulse responses are sampled on a regular rectilinear grid, and interpolated with piecewise linear weighting functions, $w_i$, on a triangle mesh subordinate to that grid.
+Adaptive refinement of the grid was proposed in [CITE], and we demonstrated a practical algorithm for adaptive refinement in our previous work [CITE MY PAPER].
+The issue with all of these existing methods, however, is computing each $\varphi_i$ requires one matrix vector product with $\mathcal{H}$, so that $r$ matrix-vector products are required.
+This is expensive when $\mathcal{H}$ is the Hessian in a PDE-constrained inverse problem, because each matrix vector product with $\mathcal{H}$ requires solving two PDEs.
+A key innovation in this paper is that we pick points $x_i$ in such a way that many $\varphi_i$ are computed with each matrix vector product with $\mathcal{H}$, thereby greatly reducing the number of matrix-vector products with $\mathcal{H}$ that are required.
+
+We do this by taking advantage of the local support property of $\mathcal{A}$.
+
+SVIR low rank: sample columns of matrix
+
+Pros and cons of different point picking strategies depend on the application of interest.
+
+In astronomy image deblurring applications, the point source positions typically correspond to locations of stars in an image, which are fixed.
 
@@
 
+H-matrices for Hessians: Ilonas paper with KAUST people
+
+Demanet matrix probing can be interpreted as convolution-product.
+
+Resolution analysis: uses matvecs with random functions to estimate spatial covariance. Was inspiration for us.
+
 Getting many impulse responses at once (batching).
 
-Geometry-aware weighting functions vs radial basis functions
+Geometry-aware weighting functions vs radial basis functions. Kriging.
 
@@
 We only have access to $\mathcal{H}$ via its action on an arbitrary vectors $v$, i.e., evaluation of the map $v \mapsto \mathcal{H} v$.
-Evaluating $v \mapsto \mathcal{H} v$ is an expensive process that involves solving two auxiliary PDEs.
+Evaluating $v \mapsto \mathcal{H} v$ is an expensive process that involves solving two auxiliary linear PDEs.
 

### 2020-12-09 paper/localpsf.tex
I hate writing literature review

@@
 
-\subsection{Existing work}
-
-In general, product-convolution approximations take the form \eqref{eq:product_convolution}, but $\varphi_i$ and $w_i$ may be arbitrary functions.
+\subsection{Connection to existing work}
+
+In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\varphi_i$ and $w_i$ may be arbitrary functions.
 Product-convolution approximations have been used in a wide variety of fields going back several decades [CITE BIG LIST].
@@
 
-We build upon a line of work in which the convolution kernels, $\varphi_i$, are impulse responses of $\mathcal{A}$ to point sources [CITE BIG LIST].
-The primary question in methods of this type are how to choose the points $x_i$, and how to choose the weighting functions $w_i$.
-In [CITE NAGY], the impulse responses are sampled on a regular rectilinear grid, and interpolated with piecewise linear weighting functions, $w_i$, on a triangle mesh subordinate to that grid.
+We build upon a class of product convolution approximations in which the functions, $\varphi_i$, are impulse responses of $\mathcal{A}$ to point sources at a collection of points $x_i$ [CITE BIG LIST].
+%The primary question in methods of this type are how to choose the points $x_i$, and how to choose the weighting functions $w_i$.
+A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the functions $\varphi_i$ with piecewise linear interpolation [CITE NAGY], or splines [CITE SPLINES].
 Adaptive refinement of the grid was proposed in [CITE], and we demonstrated a practical algorithm for adaptive refinement in our previous work [CITE MY PAPER].
-The issue with all of these existing methods, however, is computing each $\varphi_i$ requires one matrix vector product with $\mathcal{H}$, so that $r$ matrix-vector products are required.
-This is expensive when $\mathcal{H}$ is the Hessian in a PDE-constrained inverse problem, because each matrix vector product with $\mathcal{H}$ requires solving two PDEs.
-A key innovation in this paper is that we pick points $x_i$ in such a way that many $\varphi_i$ are computed with each matrix vector product with $\mathcal{H}$, thereby greatly reducing the number of matrix-vector products with $\mathcal{H}$ that are required.
+The issue with these existing methods is that computing each $\varphi_i$ requires one matrix vector product with $\mathcal{H}$, and each of these matrix vector products is computationally costly.
+The key innovation in this paper is that we pick points $x_i$ in such a way that many $\varphi_i$ are computed with each matrix vector product.
+
+Our method of estimating the support of $\varphi_i$ was inspired by resolution analysis [CITE], in which the
+
+The goal of these existing papers is to pick points $x_i$ so that the required the number of $\varphi_i$ is minimized.
+Our goal here is to minimize the required number of matrix vector products with $\mathcal{H}$.
 

### 2020-12-09 paper/localpsf.tex
I hate writing literature review

@@
 Adaptive refinement of the grid was proposed in [CITE], and we demonstrated a practical algorithm for adaptive refinement in our previous work [CITE MY PAPER].
-The issue with these existing methods is that computing each $\varphi_i$ requires one matrix vector product with $\mathcal{H}$, and each of these matrix vector products is computationally costly.
-The key innovation in this paper is that we pick points $x_i$ in such a way that many $\varphi_i$ are computed with each matrix vector product.
-
-Our method of estimating the support of $\varphi_i$ was inspired by resolution analysis [CITE], in which the
-
-The goal of these existing papers is to pick points $x_i$ so that the required the number of $\varphi_i$ is minimized.
+Even with adaptive refinement, the number of points $x_i$ may be large, and computing the functions $\varphi_i$ requires one matrix vector product with $\mathcal{A}$ for each point $x_i$, which is computationally costly.
+A key innovation in this paper is that we pick points $x_i$ in such a way that many $\varphi_i$ are computed with each matrix vector product.
+We do this by taking advantage of the fact that the functions $\varphi_i$ are locally supported; existing methods typically do not require this local support property.
+
+Our method of estimating the support of the functions $\varphi_i$ was inspired by resolution analysis in seismic imaging [CITE].
+In resolution analysis, $\mathcal{A}$ is the Hessian for a seismic inverse problem, and the width of $\varphi_p$ is used as an estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
+For example, if the width of $\varphi_p$ is one meter, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than one meter in size.
+%in which the length scales of the functions $\varphi_x$ are estimated for all $x$ simultaneously by applying $\mathcal{A}^T$ to a small number of random noise functions, $\zeta$, and post-processing the results.
+In [CITE], the width of $\varphi_x$ is estimated to be the local autocorrelation length of the function $\mathcal{A}^T \zeta$ near $x$, where $\zeta$ is a random noise funtion.
+In this paper, rather than probing $\mathcal{A}^T$ with random noise functions, we use specific linear and quadratic functions.
+Our method achieves substantially better estimates of the supports than resolution analysis, but our method requires that $\mathcal{A}$ has a positive integral kernel, while the previous resolution analysis method does not have this requirement.
+
+%The goal of these existing papers is to pick points $x_i$ so that the required the number of $\varphi_i$ is minimized.
 Our goal here is to minimize the required number of matrix vector products with $\mathcal{H}$.
 
-We do this by taking advantage of the local support property of $\mathcal{A}$.
-
-SVIR low rank: sample columns of matrix
-
-Pros and cons of different point picking strategies depend on the application of interest.
+%We do this by taking advantage of the local support property of $\mathcal{A}$.
+
+%SVIR low rank: sample columns of matrix
+
+%Pros and cons of different point picking strategies depend on the application of interest.
 
@@
 In [CITE], Escande and Weiss explored the idea of approximating linear operators from scattered impulse responses using a product-convolution scheme of the same form as \eqref{eq:product_convolution}.
-In our previous work [CITE MY PAPER], we choose the point locations with an adaptive grid.
-However,
-
-Nagy's paper: breaking domain into regular grid, using piecewise linear interpolation.
+
+%In our previous work [CITE MY PAPER], we choose the point locations with an adaptive grid. However,
+
+%Nagy's paper: breaking domain into regular grid, using piecewise linear interpolation.
 
@@
 
-Resolution analysis: uses matvecs with random functions to estimate spatial covariance. Was inspiration for us.
-
-Getting many impulse responses at once (batching).
+%Resolution analysis: uses matvecs with random functions to estimate spatial covariance. Was inspiration for us.
+
+%Getting many impulse responses at once (batching).
 
 Geometry-aware weighting functions vs radial basis functions. Kriging.
+
+Wavelets [CITE ESCANDE] vs. H-matrices.
 

### 2020-12-09 paper/localpsf.tex
literature review gah

@@
 
-We build upon a class of product convolution approximations in which the functions, $\varphi_i$, are impulse responses of $\mathcal{A}$ to point sources at a collection of points $x_i$ [CITE BIG LIST].
+We build upon the class of product convolution approximations in which the functions, $\varphi_i$, are impulse responses of $\mathcal{A}$ to point sources at a collection of points $x_i$ [CITE BIG LIST].
 %The primary question in methods of this type are how to choose the points $x_i$, and how to choose the weighting functions $w_i$.
 A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the functions $\varphi_i$ with piecewise linear interpolation [CITE NAGY], or splines [CITE SPLINES].
-Adaptive refinement of the grid was proposed in [CITE], and we demonstrated a practical algorithm for adaptive refinement in our previous work [CITE MY PAPER].
-Even with adaptive refinement, the number of points $x_i$ may be large, and computing the functions $\varphi_i$ requires one matrix vector product with $\mathcal{A}$ for each point $x_i$, which is computationally costly.
-A key innovation in this paper is that we pick points $x_i$ in such a way that many $\varphi_i$ are computed with each matrix vector product.
+With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
+This is computationally prohibitive in our applications, because computing the functions $\varphi_i$ requires one matrix vector product with $\mathcal{A}$ for each point $x_i$, and each of these matrix vector products is extremely computationally costly.
+One can reduce the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid only in the regions where the error in the approximation is large [CITE, CITE MY PAPER].
+But even with adaptive refinement, the required number of matrix vector products with $\mathcal{A}$ can be large.
+The key innovation in this paper is that we pick points $x_i$ in such a way that many $\varphi_i$ are computed with each matrix vector product.
 We do this by taking advantage of the fact that the functions $\varphi_i$ are locally supported; existing methods typically do not require this local support property.
@@
 In resolution analysis, $\mathcal{A}$ is the Hessian for a seismic inverse problem, and the width of $\varphi_p$ is used as an estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
-For example, if the width of $\varphi_p$ is one meter, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than one meter in size.
+For example, if the width of $\varphi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
 %in which the length scales of the functions $\varphi_x$ are estimated for all $x$ simultaneously by applying $\mathcal{A}^T$ to a small number of random noise functions, $\zeta$, and post-processing the results.
@@
 In this paper, rather than probing $\mathcal{A}^T$ with random noise functions, we use specific linear and quadratic functions.
-Our method achieves substantially better estimates of the supports than resolution analysis, but our method requires that $\mathcal{A}$ has a positive integral kernel, while the previous resolution analysis method does not have this requirement.
+Our method estimates the support of $\varphi_p$ more accurately than resolution analysis, but our method requires that $\mathcal{A}$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
 
@@
 
-In astronomy image deblurring applications, the point source positions typically correspond to locations of stars in an image, which are fixed.
+%In astronomy image deblurring applications, the point source positions typically correspond to locations of stars in an image, which are fixed.
 

### 2020-12-09 paper/localpsf.tex
literature review gah

@@
 Once in H-matrix format, fast H-matrix arithmetic is used to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations.
-
-\subsection{Connection to existing work}
-
-In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\varphi_i$ and $w_i$ may be arbitrary functions.
-Product-convolution approximations have been used in a wide variety of fields going back several decades [CITE BIG LIST].
-For background on product-convolution approximations, we recommend reading the following excellent papers: [CITES].
-
-We build upon the class of product convolution approximations in which the functions, $\varphi_i$, are impulse responses of $\mathcal{A}$ to point sources at a collection of points $x_i$ [CITE BIG LIST].
-%The primary question in methods of this type are how to choose the points $x_i$, and how to choose the weighting functions $w_i$.
-A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the functions $\varphi_i$ with piecewise linear interpolation [CITE NAGY], or splines [CITE SPLINES].
-With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
-This is computationally prohibitive in our applications, because computing the functions $\varphi_i$ requires one matrix vector product with $\mathcal{A}$ for each point $x_i$, and each of these matrix vector products is extremely computationally costly.
-One can reduce the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid only in the regions where the error in the approximation is large [CITE, CITE MY PAPER].
-But even with adaptive refinement, the required number of matrix vector products with $\mathcal{A}$ can be large.
-The key innovation in this paper is that we pick points $x_i$ in such a way that many $\varphi_i$ are computed with each matrix vector product.
-We do this by taking advantage of the fact that the functions $\varphi_i$ are locally supported; existing methods typically do not require this local support property.
-
-Our method of estimating the support of the functions $\varphi_i$ was inspired by resolution analysis in seismic imaging [CITE].
-In resolution analysis, $\mathcal{A}$ is the Hessian for a seismic inverse problem, and the width of $\varphi_p$ is used as an estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
-For example, if the width of $\varphi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
-%in which the length scales of the functions $\varphi_x$ are estimated for all $x$ simultaneously by applying $\mathcal{A}^T$ to a small number of random noise functions, $\zeta$, and post-processing the results.
-In [CITE], the width of $\varphi_x$ is estimated to be the local autocorrelation length of the function $\mathcal{A}^T \zeta$ near $x$, where $\zeta$ is a random noise funtion.
-In this paper, rather than probing $\mathcal{A}^T$ with random noise functions, we use specific linear and quadratic functions.
-Our method estimates the support of $\varphi_p$ more accurately than resolution analysis, but our method requires that $\mathcal{A}$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
-
-%The goal of these existing papers is to pick points $x_i$ so that the required the number of $\varphi_i$ is minimized.
-Our goal here is to minimize the required number of matrix vector products with $\mathcal{H}$.
-
-%We do this by taking advantage of the local support property of $\mathcal{A}$.
-
-%SVIR low rank: sample columns of matrix
-
-%Pros and cons of different point picking strategies depend on the application of interest.
-
-%In astronomy image deblurring applications, the point source positions typically correspond to locations of stars in an image, which are fixed.
-
-In [CITE], Escande and Weiss explored the idea of approximating linear operators from scattered impulse responses using a product-convolution scheme of the same form as \eqref{eq:product_convolution}.
-
-%In our previous work [CITE MY PAPER], we choose the point locations with an adaptive grid. However,
-
-%Nagy's paper: breaking domain into regular grid, using piecewise linear interpolation.
-
-Use for approximating Hessians in PDE-constrained optimization: Georg's paper, my paper.
-
-H-matrices for Hessians: Ilonas paper with KAUST people
-
-Demanet matrix probing can be interpreted as convolution-product.
-
-%Resolution analysis: uses matvecs with random functions to estimate spatial covariance. Was inspiration for us.
-
-%Getting many impulse responses at once (batching).
-
-Geometry-aware weighting functions vs radial basis functions. Kriging.
-
-Wavelets [CITE ESCANDE] vs. H-matrices.
-
-\subsection{old stuff}
-
-% $\left(\mathcal{A}u\right)(v) = \left\langle v, \varphi_p \ast u\right\rangle_{L^2(\Omega)}$. Here $f \ast g$ denotes convolution of the functions $f$ and $g$, and the convolution kernel,
-
-%Access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$.
-%Integral kernel representation \eqref{eq:kernel_representation} is assumed to exist,
-%e.g., by the Schwarz kernel theorem,
-%but o
-%One cannot easily access entries $A(y,x)$ of the kernel.
-
-% is non-negative in the sense that  $A(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$. 
-%Let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be the corresponding linear operator generated by currying.
-That is, for each $u\in L^2(\Omega)$, the linear functional $\mathcal{A}u:L^2(\Omega)\rightarrow \mathbb{R}$ is
-%We approximate the operator $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ defined by
-%\begin{equation*}
-%	(\mathcal{A}u)(v) := a(v,u).
-%\end{equation*}
-%We seek to approximate operators $\mathcal{A}$ in the ``matrix free'' setting, in which access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$.
-Here ($\mathcal{A}^Tv)(u) := (\mathcal{A}u)(v)$.
-%Integral kernel representation \eqref{eq:kernel_representation} is assumed to exist,
-%e.g., by the Schwarz kernel theorem,
-%but one cannot easily access entries $A(y,x)$ of the kernel. We assume that the impulse response of $\mathcal{A}$ is locally supported in the sense that the function 
-%$y \mapsto A(y, x)$ is supported in a region near $y=x$, and we assume that $\mathcal{A}$ is locally translation invariant in the sense that
-%\begin{equation*}
-%A(y+h, x+h) \approx A(y,x),
-%\end{equation*}
-%so long as $h$ is not too large.
-
-The \emph{point spread function} (PSF) of $\mathcal{A}$ at $x\in \Omega$ is
-\begin{equation*}
-	\varphi_x(y) := \langle \mathcal{A},\delta_x\rangle^*(y+x) = A(y+x,x).
-\end{equation*}
-Here $\langle \mathcal{A}, \delta_x \rangle \in L^2(\Omega)'$ is the result of applying $\mathcal{A}$ to a point source $\delta_x$ (delta distribution) centered at $x$, and $\langle\mathcal{A},\delta_x\rangle^* \in L^2(\Omega)$ is the Riesz representation of $\langle\mathcal{A},\delta_x\rangle$.
-The PSF, $\varphi_x$, is the result of translating $\langle\mathcal{A},\delta_x\rangle^*$ to re-center it at zero instead of $x$.
-Recall the Riesz representation of a linear functional $\psi \in L^2(\Omega)'$ is the unique vector $\psi^* \in L^2(\Omega)$ satisfying $\left(\psi^*, v\right)_{L^2(\Omega)} = \psi(v)$ for all $v \in \L^2(\Omega)$.
-
-%In this paper we present an efficient matrix-free method for approximating operators $\mathcal{A}$ with a product-convolution approximation, $\widetilde{A}$, of the form
-%\begin{equation*}
-%	\left(\mathcal{A}u\right) \approx \left(\widetilde{\mathcal{A}}u\right) := \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right)
-%\end{equation*}
-%
-%
-%PSFs that satisfy the following properties:
-%\begin{enumerate}
-%	\item $\varphi_x$ varies smoothly as a function of $x$.
-%	\item $\varphi_x(y)$ is zero (or small) when $|y|$ is large.
-%	\item $\varphi_x$ is a non-negative (or mostly non-negative) function.
-%\end{enumerate}
-%If $\varphi_x$ was constant in $x$, then $\mathcal{A}$ would be the convolution operator $\left(\mathcal{A}v\right)^* = \varphi_0 \ast v$.
-Operators satisfying property 1 therefore act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies smoothly as $x$ changes.
-Operators satisfying property 2 are local, in the sense that sources at point $x$ have no impact on points far away from $x$.
-Operators satisfying property 3 blur inputs: their action on a point source consists of scaling the source, then spreading the energy of the scaled point source across a region nonuniformly.
-We call operators satisfying these properties \emph{smoothly varying blur operators}.
-
-% Computing the functions $w_i$ requires solving two Poisson PDEs for each point $x_i$, and these Poisson PDEs may be solved cheaply with multigrid.
-
-%Specifically, since the integral kernel for $\mathcal{A}$ is non-negative, $\langle\mathcal{A},\delta_x\rangle^*$ may be interpreted as a scaled probability distribution.
-We compute the mean and covariance of this probability distribution for all points $x \in \Omega$ simultaneously, by applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions.
-We then estimate the support of $\langle\mathcal{A},\delta_x\rangle^*$ to be contained within a small number of standard deviations of the mean of this probability distribution.
-
-%We interpret $\langle\mathcal{A},\delta_{x_i}\rangle^*$ as a scaled probability distribution, and estimate the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ to be contained within a small number of standard deviations of the mean of this probability distribution (this is why we require a non-negative kernel).
-The mean and covariance of $\langle\mathcal{A},\delta_{x}\rangle^*$ can be simultaneously computed for all points $x \in \Omega$ by applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions.
-
-%The functions $w_i$ are the defined as the smoothest functions, in a certain least-squares sense, that interpolate the points $x_i$.
-Computing the functions $w_i$ requires solving two Poisson PDEs for each point $x_i$, and these Poisson PDEs may be solved cheaply with multigrid.
-
-%The idea of our method is to sample $\varphi_{x}$ at a collection of points $x_i$ by applying $\mathcal{A}$ to a sum of point sources at the points $x_i$ (a ``Dirac comb'').
-The points $x_i$ are chosen so that the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ and the support of $\langle\mathcal{A},\delta_{x_j}\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$.
-Since the point sources do not overlap, by applying $\mathcal{A}$ to one vector, we recover $\varphi_{x_i}$ for many points $x_i$.
-The process is repeated to get $\varphi_{x_i}$ at more points, until an error estimate is satisfied.
-The $\varphi_{x_i}$ are interpolated smoothly to construct $\varphi_x$ for arbitrary $x$, which yields an approximation $\widetilde{A}(y,x)$ to the kernel $A(y,x)$, and an approximation $\widetilde{\mathcal{A}}$ to the operator $\mathcal{A}$.
-While $A(y,x)$ is inaccessible, we can evaluate $\widetilde{A}(y,x)$ at arbitrary $(y,x)$ points.
-This allows us to construct a hierarchical matrix representation of $\widetilde{\mathcal{A}}$, which allows us to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations, using fast hierarchical matrix arithmetic.
-
-%Properties 1 and 2 are essential to our method.
-Property 3 is only required in the subprocedure for estimating the support of $\langle\mathcal{A},\delta_x\rangle^*$, and may be dropped if other methods for estimating the support are available.
-
-The approximation $\widetilde{A}(y,x)$ may not be accurate if both $x$ and $y$ are near the boundary.
-Other approximation methods, if available, should be considered for the block of the operator associated with boundary-boundary interactions.
-Nevertheless, in our numerical results we see that approximation works well for the the inverse problem considered, without any modifications to the boundary-boundary block.
 
@@
 
-The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term.
-Either the Lanczos method or randomized SVD are used to perform the low rank approximation using only matrix-vector products.
-The larger the rank of the approximation, the more matrix-vector products are required.
-These methods suffer from a ``data predicament''---as more informative data is used in the inversion, the numerical rank of the data misfit term in the Hessian grows.
-The ideal scenario from a scientific perspective (highly informative data) is therefore the worst case scenario from a computational perspective (large computational cost).
+The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term [CITE LOTS OF STUFF].
+Either the Lanczos method or the randomized singular value decomposition are used to perform the low rank approximation using only matrix-vector products.
+These methods suffer from a ``data predicament''---as more informative data is used in the inversion, the numerical rank of the data misfit term in the Hessian grows, so more matrix-vector products are required to form the aforementioned low rank approximation.
+The ideal scenario from a scientific perspective (highly informative data) is therefore the worst case scenario from a computational perspective (large computational cost) [CITE MY DISSERTATION].
 In our numerical results, we show that our operator approximation method can efficiently approximate the data misfit term in the Hessian in certain PDE constrained inverse problems.
@@
 High-rank approximations of an operator can be formed using a small number of matrix-vector products.
+
+\subsection{Connection to existing work}
+
+In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\varphi_i$ and $w_i$ may be arbitrary functions.
+Product-convolution approximations have been used in a wide variety of fields going back several decades [CITE BIG LIST].
+For background on product-convolution approximations, we recommend reading the following excellent papers: [CITES].
+
+We build upon the class of product convolution approximations in which the functions, $\varphi_i$, are impulse responses of $\mathcal{A}$ to point sources at a collection of points $x_i$ [CITE BIG LIST].
+%The primary question in methods of this type are how to choose the points $x_i$, and how to choose the weighting functions $w_i$.
+A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the functions $\varphi_i$ with piecewise linear interpolation [CITE NAGY], or splines [CITE SPLINES].
+With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
+This is computationally prohibitive in our applications, because computing the functions $\varphi_i$ requires one matrix vector product with $\mathcal{A}$ for each point $x_i$, and each of these matrix vector products is extremely computationally costly.
+One can reduce the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid only in the regions where the error in the approximation is large [CITE, CITE MY PAPER].
+But even with adaptive refinement, the required number of matrix vector products with $\mathcal{A}$ can be large.
+The key innovation in this paper is that we pick points $x_i$ in such a way that many $\varphi_i$ are computed with each matrix vector product.
+We do this by taking advantage of the fact that the functions $\varphi_i$ are locally supported; existing methods typically do not require this local support property.
+
+Our method of estimating the support of the functions $\varphi_i$ was inspired by resolution analysis in seismic imaging [CITE].
+In resolution analysis, $\mathcal{A}$ is the Hessian for a seismic inverse problem, and the width of $\varphi_p$ is used as an estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
+For example, if the width of $\varphi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
+%in which the length scales of the functions $\varphi_x$ are estimated for all $x$ simultaneously by applying $\mathcal{A}^T$ to a small number of random noise functions, $\zeta$, and post-processing the results.
+In [CITE], the width of $\varphi_x$ is estimated to be the local autocorrelation length of the function $\mathcal{A}^T \zeta$ near $x$, where $\zeta$ is a random noise funtion.
+In this paper, rather than probing $\mathcal{A}^T$ with random noise functions, we use specific linear and quadratic functions.
+Our method estimates the support of $\varphi_p$ more accurately than resolution analysis, but our method requires that $\mathcal{A}$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
+
+%The goal of these existing papers is to pick points $x_i$ so that the required the number of $\varphi_i$ is minimized.
+Our goal here is to minimize the required number of matrix vector products with $\mathcal{H}$.
+
+%We do this by taking advantage of the local support property of $\mathcal{A}$.
+
+%SVIR low rank: sample columns of matrix
+
+%Pros and cons of different point picking strategies depend on the application of interest.
+
+%In astronomy image deblurring applications, the point source positions typically correspond to locations of stars in an image, which are fixed.
+
+In [CITE], Escande and Weiss explored the idea of approximating linear operators from scattered impulse responses using a product-convolution scheme of the same form as \eqref{eq:product_convolution}.
+
+%In our previous work [CITE MY PAPER], we choose the point locations with an adaptive grid. However,
+
+%Nagy's paper: breaking domain into regular grid, using piecewise linear interpolation.
+
+Use for approximating Hessians in PDE-constrained optimization: Georg's paper, my paper.
+
+H-matrices for Hessians: Ilonas paper with KAUST people
+
+Demanet matrix probing can be interpreted as convolution-product.
+
+%Resolution analysis: uses matvecs with random functions to estimate spatial covariance. Was inspiration for us.
+
+%Getting many impulse responses at once (batching).
+
+Geometry-aware weighting functions vs radial basis functions. Kriging.
+
+Wavelets [CITE ESCANDE] vs. H-matrices.
 
@@
 
+\section{old stuff}
+
+% $\left(\mathcal{A}u\right)(v) = \left\langle v, \varphi_p \ast u\right\rangle_{L^2(\Omega)}$. Here $f \ast g$ denotes convolution of the functions $f$ and $g$, and the convolution kernel,
+
+%Access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$.
+%Integral kernel representation \eqref{eq:kernel_representation} is assumed to exist,
+%e.g., by the Schwarz kernel theorem,
+%but o
+%One cannot easily access entries $A(y,x)$ of the kernel.
+
+% is non-negative in the sense that  $A(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$. 
+%Let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be the corresponding linear operator generated by currying.
+That is, for each $u\in L^2(\Omega)$, the linear functional $\mathcal{A}u:L^2(\Omega)\rightarrow \mathbb{R}$ is
+%We approximate the operator $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ defined by
+%\begin{equation*}
+%	(\mathcal{A}u)(v) := a(v,u).
+%\end{equation*}
+%We seek to approximate operators $\mathcal{A}$ in the ``matrix free'' setting, in which access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$.
+Here ($\mathcal{A}^Tv)(u) := (\mathcal{A}u)(v)$.
+%Integral kernel representation \eqref{eq:kernel_representation} is assumed to exist,
+%e.g., by the Schwarz kernel theorem,
+%but one cannot easily access entries $A(y,x)$ of the kernel. We assume that the impulse response of $\mathcal{A}$ is locally supported in the sense that the function 
+%$y \mapsto A(y, x)$ is supported in a region near $y=x$, and we assume that $\mathcal{A}$ is locally translation invariant in the sense that
+%\begin{equation*}
+%A(y+h, x+h) \approx A(y,x),
+%\end{equation*}
+%so long as $h$ is not too large.
+
+The \emph{point spread function} (PSF) of $\mathcal{A}$ at $x\in \Omega$ is
+\begin{equation*}
+\varphi_x(y) := \langle \mathcal{A},\delta_x\rangle^*(y+x) = A(y+x,x).
+\end{equation*}
+Here $\langle \mathcal{A}, \delta_x \rangle \in L^2(\Omega)'$ is the result of applying $\mathcal{A}$ to a point source $\delta_x$ (delta distribution) centered at $x$, and $\langle\mathcal{A},\delta_x\rangle^* \in L^2(\Omega)$ is the Riesz representation of $\langle\mathcal{A},\delta_x\rangle$.
+The PSF, $\varphi_x$, is the result of translating $\langle\mathcal{A},\delta_x\rangle^*$ to re-center it at zero instead of $x$.
+Recall the Riesz representation of a linear functional $\psi \in L^2(\Omega)'$ is the unique vector $\psi^* \in L^2(\Omega)$ satisfying $\left(\psi^*, v\right)_{L^2(\Omega)} = \psi(v)$ for all $v \in \L^2(\Omega)$.
+
+%In this paper we present an efficient matrix-free method for approximating operators $\mathcal{A}$ with a product-convolution approximation, $\widetilde{A}$, of the form
+%\begin{equation*}
+%	\left(\mathcal{A}u\right) \approx \left(\widetilde{\mathcal{A}}u\right) := \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right)
+%\end{equation*}
+%
+%
+%PSFs that satisfy the following properties:
+%\begin{enumerate}
+%	\item $\varphi_x$ varies smoothly as a function of $x$.
+%	\item $\varphi_x(y)$ is zero (or small) when $|y|$ is large.
+%	\item $\varphi_x$ is a non-negative (or mostly non-negative) function.
+%\end{enumerate}
+%If $\varphi_x$ was constant in $x$, then $\mathcal{A}$ would be the convolution operator $\left(\mathcal{A}v\right)^* = \varphi_0 \ast v$.
+Operators satisfying property 1 therefore act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies smoothly as $x$ changes.
+Operators satisfying property 2 are local, in the sense that sources at point $x$ have no impact on points far away from $x$.
+Operators satisfying property 3 blur inputs: their action on a point source consists of scaling the source, then spreading the energy of the scaled point source across a region nonuniformly.
+We call operators satisfying these properties \emph{smoothly varying blur operators}.
+
+% Computing the functions $w_i$ requires solving two Poisson PDEs for each point $x_i$, and these Poisson PDEs may be solved cheaply with multigrid.
+
+%Specifically, since the integral kernel for $\mathcal{A}$ is non-negative, $\langle\mathcal{A},\delta_x\rangle^*$ may be interpreted as a scaled probability distribution.
+We compute the mean and covariance of this probability distribution for all points $x \in \Omega$ simultaneously, by applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions.
+We then estimate the support of $\langle\mathcal{A},\delta_x\rangle^*$ to be contained within a small number of standard deviations of the mean of this probability distribution.
+
+%We interpret $\langle\mathcal{A},\delta_{x_i}\rangle^*$ as a scaled probability distribution, and estimate the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ to be contained within a small number of standard deviations of the mean of this probability distribution (this is why we require a non-negative kernel).
+The mean and covariance of $\langle\mathcal{A},\delta_{x}\rangle^*$ can be simultaneously computed for all points $x \in \Omega$ by applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions.
+
+%The functions $w_i$ are the defined as the smoothest functions, in a certain least-squares sense, that interpolate the points $x_i$.
+Computing the functions $w_i$ requires solving two Poisson PDEs for each point $x_i$, and these Poisson PDEs may be solved cheaply with multigrid.
+
+%The idea of our method is to sample $\varphi_{x}$ at a collection of points $x_i$ by applying $\mathcal{A}$ to a sum of point sources at the points $x_i$ (a ``Dirac comb'').
+The points $x_i$ are chosen so that the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ and the support of $\langle\mathcal{A},\delta_{x_j}\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$.
+Since the point sources do not overlap, by applying $\mathcal{A}$ to one vector, we recover $\varphi_{x_i}$ for many points $x_i$.
+The process is repeated to get $\varphi_{x_i}$ at more points, until an error estimate is satisfied.
+The $\varphi_{x_i}$ are interpolated smoothly to construct $\varphi_x$ for arbitrary $x$, which yields an approximation $\widetilde{A}(y,x)$ to the kernel $A(y,x)$, and an approximation $\widetilde{\mathcal{A}}$ to the operator $\mathcal{A}$.
+While $A(y,x)$ is inaccessible, we can evaluate $\widetilde{A}(y,x)$ at arbitrary $(y,x)$ points.
+This allows us to construct a hierarchical matrix representation of $\widetilde{\mathcal{A}}$, which allows us to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations, using fast hierarchical matrix arithmetic.
+
+%Properties 1 and 2 are essential to our method.
+Property 3 is only required in the subprocedure for estimating the support of $\langle\mathcal{A},\delta_x\rangle^*$, and may be dropped if other methods for estimating the support are available.
+
+The approximation $\widetilde{A}(y,x)$ may not be accurate if both $x$ and $y$ are near the boundary.
+Other approximation methods, if available, should be considered for the block of the operator associated with boundary-boundary interactions.
+Nevertheless, in our numerical results we see that approximation works well for the the inverse problem considered, without any modifications to the boundary-boundary block.
+
 \section*{Acknowledgments}

### 2020-12-10 paper/localpsf.tex
literature review gah

@@
 For example, if the width of $\varphi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
-%in which the length scales of the functions $\varphi_x$ are estimated for all $x$ simultaneously by applying $\mathcal{A}^T$ to a small number of random noise functions, $\zeta$, and post-processing the results.
 In [CITE], the width of $\varphi_x$ is estimated to be the local autocorrelation length of the function $\mathcal{A}^T \zeta$ near $x$, where $\zeta$ is a random noise funtion.
@@
 
-Use for approximating Hessians in PDE-constrained optimization: Georg's paper, my paper.
-
-H-matrices for Hessians: Ilonas paper with KAUST people
+A product-convolution type approximation of the Hessian in a seismic imaging problem was used in [CITE GEORG].
+
+Conventional H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
+There are matrix-free methods for H-matrix approximation [CITE LEXING PAPER, CITE MARTINSSON, CITE OTHERS], and these methods have been used to form H-matrix compressions of Hessians in PDE constrained inverse problems [CITE ILONA, CITE NOEMI/TUCKER HODLR].
+While these matrix-free H-matrix construction methods are asymptotically scalable in theory, the required number of matrix vector products is large in practice.
+
+%H-matrices for Hessians: Ilonas paper with KAUST people
 

### 2020-12-10 paper/localpsf.tex
literature review gah

@@
 
+Matrix-vector products with product-convolution approximations can be performed using the fast Fourier transform if the problem is discretized on a regular grid [CITE].
+The product-convolution approximation can then replace the original operator when using Krylov methods to solve linear systems with the original operator as the coefficient operator [CITE].
+However, for complex geometries, problems are rarely discretized using regular grids, so we cannot easily use the fast Fourier transform.
+Furthermore, in big-data inverse problems the Hessian is typically highly ill-conditioned, so the number of Krylov iterations required is large.
+Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are more amenable to linear algebra operations.
+Wavelet compression methods have been used for this purpose [CITE WAVELETS].
+Here we follow our previous work [CITE MY PAPER], in which we converting the product-convolution approximation to H-matrix format.
+
 Conventional H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
@@
 Geometry-aware weighting functions vs radial basis functions. Kriging.
-
-Wavelets [CITE ESCANDE] vs. H-matrices.
 

### 2020-12-10 paper/localpsf.tex
literature review gah

@@
 Product-convolution approximations have been used in a wide variety of fields going back several decades [CITE BIG LIST].
-For background on product-convolution approximations, we recommend reading the following excellent papers: [CITES].
+For background on product-convolution approximations, we recommend reading the following papers: [CITES].
 
@@
 Our method of estimating the support of the functions $\varphi_i$ was inspired by resolution analysis in seismic imaging [CITE].
-In resolution analysis, $\mathcal{A}$ is the Hessian for a seismic inverse problem, and the width of $\varphi_p$ is used as an estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
-For example, if the width of $\varphi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
+In resolution analysis, $\mathcal{A}$ is the Hessian for a seismic inverse problem, and the width of $\varphi_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
+If the width of $\varphi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
 In [CITE], the width of $\varphi_x$ is estimated to be the local autocorrelation length of the function $\mathcal{A}^T \zeta$ near $x$, where $\zeta$ is a random noise funtion.
@@
 Furthermore, in big-data inverse problems the Hessian is typically highly ill-conditioned, so the number of Krylov iterations required is large.
-Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are more amenable to linear algebra operations.
+After constructing the product-convolution approximation, it is therefore desirable to convert the product-convolution approximation to other matrix formats that are more amenable to linear algebra operations.
 Wavelet compression methods have been used for this purpose [CITE WAVELETS].
-Here we follow our previous work [CITE MY PAPER], in which we converting the product-convolution approximation to H-matrix format.
+Here we follow our previous work [CITE MY PAPER], in which we convert the product-convolution approximation to H-matrix format.
 
@@
 
-\subsection{Background}
-
-\subsubsection{Distributions}
+\subsection{Background: distributions}
 
@@
 
-\subsubsection{Discretization}
+\subsection{Background: discretization}
 \label{sec:discretization}
@@
 We summarize coefficient vector versions of the function space operation that are used in this paper in Appendix \ref{app:discretized_operations}.
+
+\subsection{Background: hierarchical matrices}
 

### 2020-12-11 paper/localpsf.tex
wrote weighting function algorithm at infinite dimensional level

@@
 
-\subsection{Determination of the boundary region}
-\label{sec:boundary}
-
-Define $\Omega_B \subset \Omega$ to be the set of points that are too close to $\partial \Omega$, in the sense that $x \in \Omega_B$ if $\varphi_x(y-x) \neq 0$ for some $y \in \partial \Omega$.
-Further, define $\Omega_I := \Omega \setminus \Omega_B$ as the set of points that are in the interior of $\Omega$ and are not too close to $\partial B$.
-In Proposition \ref{prop:boundary_source} we show that $\Omega_B$ can be determined by applying $\mathcal{A}^T$ to a boundary source distribution and post-processing the result.
-The resulting algorithm for determining $\Omega_B$ and $\Omega_I$ in the discretized setting is shown in Algorithm \ref{alg:boundary_region}.
-
-\begin{prop}
-	\label{prop:boundary_source}
-	Let $\delta_{\partial \Omega}: C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ be the boundary source distribution defined by
-	\begin{equation}
-	\label{eq:boundary_source}
-	\delta_{\partial \Omega}(v) = \int_{\partial \Omega} v(y) dy
-	\end{equation}
-	for all $v \in C(\overline{\Omega})$. We have $\varphi_x(y-x) = 0$ for almost every $y \in \partial \Omega$ if and only if $\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle^*(x) = 0$.
-\end{prop}
-
-The proof of Proposition \ref{prop:boundary_source} is shown in Appendix \ref{app:proofs}.
-
-%Straightforward analysis shows that $\varphi_x(y-x) = 0$ for $y \in \partial \Omega$ almost everywhere if and only if $\left(\mathcal{A}^T \delta_{\partial \Omega}\right)(x) = 0$, where $\delta_{\partial \Omega}$ is the uniform boundary source distribution.
-That is, the distribution that acts on a test function $v$ by performing a surface integral of $v$ over the boundary of the domain:
-%%\begin{equation}
-%%\label{eq:boundary_source}
-%%	\delta_{\partial \Omega}(v) = \int_{\partial \Omega} v(y) dy.
-%%\end{equation}
-%We therefore compute $b  = \mathcal{A}^T \delta_{\partial \Omega}$, then identify $\Omega_B$ as the set of points $x \in \Omega$ for which $|b(x)|$ is larger than a small tolerance $\epsilon$.
-This is described in Algorithm \ref{alg:boundary_region}.
-
-\begin{algorithm2e}
-	\SetAlgoNoLine
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Output}
-	{
-
-		Construct boundary source $\boldsymbol{\delta}_{\partial \Omega} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$, $\displaystyle \left(\boldsymbol{\delta}_{\partial \Omega}\right)_i = \int_{\partial \Omega} \phi_i(s) ds$, $i=1,\dots,N$
-
-%		\For{$i=1,\dots,N$}{
-%			$\displaystyle \mathbf{b}_i \gets \int_{\partial \Omega} \phi_i(s) ds$
-%		}
-
-		$\mathbf{b} = \mathbf{M}^{-1}\mathbf{A}^T\mathbf{M}^{-1}\boldsymbol{\delta}_{\partial \Omega}$
-
-		$\Omega_B = \{x \in \Omega : |b_h(x)| \ge \epsilon\}$
-
-		$\Omega_I = \Omega \setminus \Omega_B$
-
-	}
-	\caption{Compute boundary region $\Omega_B$ and interior region $\Omega_I$}
-	\label{alg:boundary_region}
-\end{algorithm2e}
-
-One should only expect $\widetilde{A}(y,x)$ to be accurate for $x \in \Omega_I$, $y \in \Omega$.
-If $\mathcal{A}$ is symmetric then one may overwrite potentially innacurate entries $\widetilde{A}(y,x)$, $x \in \Omega_b$, $y \in \Omega_I$, with the accurate entries $\widetilde{A}(x,y)$.
-If $\mathcal{A}$ is nonsymmetric, one may form an accurate approximation of these entries by applying our procedure to $\mathcal{A}^T$.
-Our method will furnish an approximation $\widetilde{A}(y,x)$ for $x \in \Omega_B$, $y \in \Omega_B$ (the block of $\widetilde{H}$ corresponding to boundary-boundary interactions), but this part of the approximation is unlikely to satisfy Condition 1 ($\varphi_x$ varies smoothly as a function of $x$).
-
 \subsection{Mean and covariance estimation}
@@
 That is,
-	\begin{equation*}
+	\begin{equation}
+	\label{eq:pure_neumann_poisson}
 		\mathbf{g} = K^+ \mathbf{f} \quad \Leftrightarrow \quad 
@@
 		\end{cases}
-	\end{equation*}
+	\end{equation}
 	where $\nu$ denotes the normal to the boundary.
@@
 
+\begin{algorithm2e}
+	\SetAlgoNoLine
+	\SetKwInOut{Input}{Input}
+	\SetKwInOut{Output}{Output}
+	%	\Input{Function $v \mapsto \mathcal{H}^Tv$}
+
+	%	\mbox{}\\
+
+	%	\Output{Spatially varying mean $\mu:\Omega \rightarrow \mathbb{R}^d$, \\ 
+	%		Spatially varying covariance $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$}
+	%	\Begin{
+	{	
+		\For{$i=1,2, \dots, r$}{
+			Compute $\psi_i$ as the solution to following Poisson problem with pure Neumann boundary conditions and a point source at $x_i$:
+			$$\begin{cases}
+			\Delta \psi_i = \delta_{x_i}, & \text{in }\Omega,\\
+			\nu \cdot \nabla \psi_i = 0, & \text{on }\partial \Omega,\\
+			\int_\Omega \psi_i(x) dx = 0.
+			\end{cases}$$
+
+			Compute $\theta_i$ as the solution to the following Poisson problem with pure Neumann boundary conditions and source $\psi_i$:
+			$$\begin{cases}
+			\Delta \theta_i = \psi_i, & \text{in }\Omega,\\
+			\nu \cdot \nabla \theta_i = 0, & \text{on }\partial \Omega,\\
+			\int_\Omega \theta_i(x) dx = 0.
+			\end{cases}$$
+		}
+
+		Form the $r \times r$ matrix $S$ with entries
+		$$S_{ij} = \left(\psi_i, \psi_j\right)_{L^2(\Omega)}.$$
+
+		\For{$i=1,2,\dots,r$}{
+			Compute
+			$$\alpha = \frac{\mathbf{1}^T S^{-1} e_i}{\mathbf{1}^T S^{-1} \mathbf{1}},$$
+			where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one, and $e_i = (0,\dots,0,1,0,\dots,0)$ is the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.
+
+			Compute 
+			\begin{equation*}
+			\lambda = -S^{-1} \left(e_i - \alpha \mathbf{1} \right).
+			\end{equation*}
+
+			Compute
+			$$w_i = -\sum_{k=1}^r \lambda_k \theta_k + \alpha C,$$
+			where we recall that $C$ is the constant function $C(x)=1$ on $\Omega$.
+		}
+	}
+	\caption{Compute weighting functions $w_i$, $i=1,\dots,r$.}
+	\label{alg:weighting_functions_incremental}
+\end{algorithm2e}
+
 The solution to this system of equations is found by the following process:
@@
 	\begin{equation*}
-		\Psi := A^+ B^T
+		\Psi := K^+ E^T
 	\end{equation*}
@@
 	\begin{equation*}
-		\Theta := A^+ M \Psi
+		\Theta := K^+ M \Psi
 	\end{equation*}
@@
 \appendix
+
+\section{Discretized algorithms}
+
 \section{Fast ellipsoid intersection test}
@@
 
+\subsection{Determination of the boundary region}
+\label{sec:boundary}
+
+Define $\Omega_B \subset \Omega$ to be the set of points that are too close to $\partial \Omega$, in the sense that $x \in \Omega_B$ if $\varphi_x(y-x) \neq 0$ for some $y \in \partial \Omega$.
+Further, define $\Omega_I := \Omega \setminus \Omega_B$ as the set of points that are in the interior of $\Omega$ and are not too close to $\partial B$.
+In Proposition \ref{prop:boundary_source} we show that $\Omega_B$ can be determined by applying $\mathcal{A}^T$ to a boundary source distribution and post-processing the result.
+The resulting algorithm for determining $\Omega_B$ and $\Omega_I$ in the discretized setting is shown in Algorithm \ref{alg:boundary_region}.
+
+\begin{prop}
+	\label{prop:boundary_source}
+	Let $\delta_{\partial \Omega}: C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ be the boundary source distribution defined by
+	\begin{equation}
+	\label{eq:boundary_source}
+	\delta_{\partial \Omega}(v) = \int_{\partial \Omega} v(y) dy
+	\end{equation}
+	for all $v \in C(\overline{\Omega})$. We have $\varphi_x(y-x) = 0$ for almost every $y \in \partial \Omega$ if and only if $\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle^*(x) = 0$.
+\end{prop}
+
+The proof of Proposition \ref{prop:boundary_source} is shown in Appendix \ref{app:proofs}.
+
+%Straightforward analysis shows that $\varphi_x(y-x) = 0$ for $y \in \partial \Omega$ almost everywhere if and only if $\left(\mathcal{A}^T \delta_{\partial \Omega}\right)(x) = 0$, where $\delta_{\partial \Omega}$ is the uniform boundary source distribution.
+That is, the distribution that acts on a test function $v$ by performing a surface integral of $v$ over the boundary of the domain:
+%%\begin{equation}
+%%\label{eq:boundary_source}
+%%	\delta_{\partial \Omega}(v) = \int_{\partial \Omega} v(y) dy.
+%%\end{equation}
+%We therefore compute $b  = \mathcal{A}^T \delta_{\partial \Omega}$, then identify $\Omega_B$ as the set of points $x \in \Omega$ for which $|b(x)|$ is larger than a small tolerance $\epsilon$.
+This is described in Algorithm \ref{alg:boundary_region}.
+
+\begin{algorithm2e}
+	\SetAlgoNoLine
+	\SetKwInOut{Input}{Input}
+	\SetKwInOut{Output}{Output}
+	{
+
+		Construct boundary source $\boldsymbol{\delta}_{\partial \Omega} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$, $\displaystyle \left(\boldsymbol{\delta}_{\partial \Omega}\right)_i = \int_{\partial \Omega} \phi_i(s) ds$, $i=1,\dots,N$
+
+		%		\For{$i=1,\dots,N$}{
+		%			$\displaystyle \mathbf{b}_i \gets \int_{\partial \Omega} \phi_i(s) ds$
+		%		}
+
+		$\mathbf{b} = \mathbf{M}^{-1}\mathbf{A}^T\mathbf{M}^{-1}\boldsymbol{\delta}_{\partial \Omega}$
+
+		$\Omega_B = \{x \in \Omega : |b_h(x)| \ge \epsilon\}$
+
+		$\Omega_I = \Omega \setminus \Omega_B$
+
+	}
+	\caption{Compute boundary region $\Omega_B$ and interior region $\Omega_I$}
+	\label{alg:boundary_region}
+\end{algorithm2e}
+
+One should only expect $\widetilde{A}(y,x)$ to be accurate for $x \in \Omega_I$, $y \in \Omega$.
+If $\mathcal{A}$ is symmetric then one may overwrite potentially inaccurate entries $\widetilde{A}(y,x)$, $x \in \Omega_b$, $y \in \Omega_I$, with the accurate entries $\widetilde{A}(x,y)$.
+If $\mathcal{A}$ is nonsymmetric, one may form an accurate approximation of these entries by applying our procedure to $\mathcal{A}^T$.
+Our method will furnish an approximation $\widetilde{A}(y,x)$ for $x \in \Omega_B$, $y \in \Omega_B$ (the block of $\widetilde{H}$ corresponding to boundary-boundary interactions), but this part of the approximation is unlikely to satisfy Condition 1 ($\varphi_x$ varies smoothly as a function of $x$).
+
 \section*{Acknowledgments}
-We acknowledge
+We thank J.J. Alger for helpful discussions
 

### 2020-12-11 paper/localpsf.tex
wrote weighting function algorithm at infinite dimensional level

@@
 		\For{$i=1,2, \dots, r$}{
-			Compute $\psi_i$ as the solution to following Poisson problem with pure Neumann boundary conditions and a point source at $x_i$:
-			$$\begin{cases}
-			\Delta \psi_i = \delta_{x_i}, & \text{in }\Omega,\\
-			\nu \cdot \nabla \psi_i = 0, & \text{on }\partial \Omega,\\
-			\int_\Omega \psi_i(x) dx = 0.
-			\end{cases}$$
-
-			Compute $\theta_i$ as the solution to the following Poisson problem with pure Neumann boundary conditions and source $\psi_i$:
-			$$\begin{cases}
-			\Delta \theta_i = \psi_i, & \text{in }\Omega,\\
-			\nu \cdot \nabla \theta_i = 0, & \text{on }\partial \Omega,\\
-			\int_\Omega \theta_i(x) dx = 0.
-			\end{cases}$$
+			Compute $\psi_i$ as the solution to the Poisson problem
+			\begin{equation*}
+				\Delta \psi_i = \delta_{x_i}, \quad \text{in }\Omega,
+			\end{equation*}
+			with Neumann boundary conditions, $\nu \cdot \nabla \psi_i = 0$ on $\partial \Omega$, and the average value zero condition, $\int_\Omega \psi_i(x) dx = 0$.
+
+			Compute $\theta_i$ as the solution to the Poisson problem
+			\begin{equation*}
+			\Delta \theta_i = \psi_i, \quad \text{in }\Omega,
+			\end{equation*}
+			with Neumann boundary conditions, $\nu \cdot \nabla \psi_i = 0$ on $\partial \Omega$, and the average value zero condition, $\int_\Omega \psi_i(x) dx = 0$.
+
 		}
@@
 
-			Compute
+			Form the weighting function
 			$$w_i = -\sum_{k=1}^r \lambda_k \theta_k + \alpha C,$$
@@
 \end{algorithm2e}
-
-The solution to this system of equations is found by the following process:
-\begin{enumerate}
-	\item Compute 
-	\begin{equation*}
-		\Psi := K^+ E^T
-	\end{equation*}
-	by solving Neumann poisson problems for point source right hand sides, with point sources located at the points $x_i$, $i=1,\dots,q$.
-	\item Compute 
-	\begin{equation*}
-		\Theta := K^+ M \Psi
-	\end{equation*}
-	by solving Neumann Poisson problems with the columns of $\Psi$ as distributed sources.
-	\item Form 
-	\begin{equation*}
-		S = \Psi^T M \Psi
-	\end{equation*}.
-	\item Compute \begin{equation*}
-		\alpha = \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{y}.
-	\end{equation*}
-	\item Compute 
-	\begin{equation*}
-		\boldsymbol{\lambda} = -S^{-1} \left(\mathbf{y} - \alpha \mathbf{1} \right).
-	\end{equation*}
-	\item Compute
-	\begin{equation*}
-		\mathbf{u} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c}
-	\end{equation*}
-\end{enumerate}
 
@@
 
+The solution to this system of equations is found by the following process:
+\begin{enumerate}
+	\item Compute 
+	\begin{equation*}
+	\Psi := K^+ E^T
+	\end{equation*}
+	by solving Neumann poisson problems for point source right hand sides, with point sources located at the points $x_i$, $i=1,\dots,q$.
+	\item Compute 
+	\begin{equation*}
+	\Theta := K^+ M \Psi
+	\end{equation*}
+	by solving Neumann Poisson problems with the columns of $\Psi$ as distributed sources.
+	\item Form 
+	\begin{equation*}
+	S = \Psi^T M \Psi
+	\end{equation*}.
+	\item Compute \begin{equation*}
+	\alpha = \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{y}.
+	\end{equation*}
+	\item Compute 
+	\begin{equation*}
+	\boldsymbol{\lambda} = -S^{-1} \left(\mathbf{y} - \alpha \mathbf{1} \right).
+	\end{equation*}
+	\item Compute
+	\begin{equation*}
+	\mathbf{u} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c}
+	\end{equation*}
+\end{enumerate}
+
 \subsection{Determination of the boundary region}

### 2020-12-11 paper/localpsf.tex
wrote weighting function algorithm at infinite dimensional level

@@
 
-We choose the weighting functions $w_i$ to be the smoothest functions, in a least-squares sense, that interpolate the points $x_i$.
+We choose the weighting functions, $w_i$, to be the solutions to the following optimization problems:
+\begin{equation}
+\label{eq:wi_optimization_problem_intro}
+\begin{aligned}
+\min_{w_i} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
+\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
+\end{aligned}
+\end{equation}
+Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$.
+These are the smoothest functions, in a least-squares sense, that interpolate the points $x_i$.
 Computing the weighting functions $w_i$ requires solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid.
+We call this interpolation scheme \emph{Poisson interpolation}.
+The advantage of Poisson interpolation over other interpolation methods, most notably radial basis function interpolation [cite Escande etc], is that Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions.
+Consider the Pine island glacier geometry shown in Figure REF.
+Regions of the glacier are separated by thin inlets of ocean.
+With radial basis function interpolation, points on one side of the inlet inappropriately influence the interpolation on the other side of the inlet.
+With Poisson interpolation, the PSF samples only influence the interpolation on their own side of the inlet.
 
@@
 Product-convolution approximations have been used in a wide variety of fields going back several decades [CITE BIG LIST].
-For background on product-convolution approximations, we recommend reading the following papers: [CITES].
+For background on product-convolution approximations, we recommend reading the following papers: [CITES, INCLUDE ESCANDE SCATTERED].
 
@@
 
-In [CITE], Escande and Weiss explored the idea of approximating linear operators from scattered impulse responses using a product-convolution scheme of the same form as \eqref{eq:product_convolution}.
+%In [CITE], Escande and Weiss explored the idea of approximating linear operators from scattered impulse responses using a product-convolution scheme of the same form as \eqref{eq:product_convolution}.
 
@@
 %Nagy's paper: breaking domain into regular grid, using piecewise linear interpolation.
-
-A product-convolution type approximation of the Hessian in a seismic imaging problem was used in [CITE GEORG].
 
@@
 While these matrix-free H-matrix construction methods are asymptotically scalable in theory, the required number of matrix vector products is large in practice.
+
+A product-convolution type approximation of the Hessian in a seismic imaging problem was used in [CITE GEORG].
 
@@
 However, we present an efficient incremental algorithm for updating these weighting functions when a new point is added, so that we only need to perform two new Poisson PDE solves, rather than $2(q+1)$ new PDE solves.
-
-A key advantage of Poisson interpolation over other PSF interpolation methods, most notably radial basis functions [cite Escande etc], is that Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions.
-Consider the Pine island glacier geometry shown in Figure REF.
-Regions of the glacier are separated by thin inlets of ocean.
-With radial basis function interpolation, PSF samples on one side of the inlet inappropriately influence the interpolation on the other side of the inlet.
-With Poisson interpolation, the PSF samples only influence the interpolation on their own side of the inlet.
 

### 2020-12-12 paper/localpsf.tex
asdf
more intro editing

@@
 \end{equation}
-The function $\varphi_x$ is the result of applying $\mathcal{A}$ to a point source, $\delta_x$, at some fixed point $x \in \Omega$, then translating the resulting impulse response function, $\mathcal{A}\delta_x$, to re-center it at zero instead of $x$.
+The function $\varphi_x$ is the result of applying $\mathcal{A}$ to a point source, $\delta_x$, at some fixed point $x \in \Omega$, then translating the resulting impulse response function to re-center it at zero instead of $x$.
 Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes.
@@
 	\item $\{x_i\}_{i=1}^r \subset \Omega$ is a collection of points scattered throughout the domain.
-	\item The functions $w_i$ are spatially varying weighting functions that interpolate the convolution kernels $\varphi_i$.
+	\item The functions $w_i$ are spatially varying weighting functions that are used to interpolate the convolution kernels $\varphi_i$.
 \end{itemize}
@@
 The process is repeated to get more batches of $\varphi_i$'s, until the approximation is sufficiently accurate or a maximum number of sample points is reached.
+
 In order to choose the points $x_i$, we need to estimate the support of $\left\langle \mathcal{A}, \delta_{x_i}\right\rangle^*$ \emph{before} we compute it.
-The supports of the impulse responses $\left\langle \mathcal{A}, \delta_{x}\right\rangle^*$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions (Section \ref{eq:mean_and_covariance_estimation}).
+The supports of the functions $\left\langle \mathcal{A}, \delta_{x}\right\rangle^*$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\mathcal{A}^T$ to a small number of specially chosen functions (three functions if $d=1$, six if $d=2$, ten if $d=3$), then post processing the results (Section \ref{eq:mean_and_covariance_estimation}).
 This procedure is the only part of the paper that requires non-negativity of the integral kernel.
@@
 Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$.
-These are the smoothest functions, in a least-squares sense, that interpolate the points $x_i$.
+Since the Laplacian measures local deviation from flatness [cite Jacobsen], the functions $w_i$ are the flattest functions, in a least-squares sense, that interpolate the points $x_i$.
 Computing the weighting functions $w_i$ requires solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid.
@@
 The advantage of Poisson interpolation over other interpolation methods, most notably radial basis function interpolation [cite Escande etc], is that Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions.
-Consider the Pine island glacier geometry shown in Figure REF.
-Regions of the glacier are separated by thin inlets of ocean.
-With radial basis function interpolation, points on one side of the inlet inappropriately influence the interpolation on the other side of the inlet.
-With Poisson interpolation, the PSF samples only influence the interpolation on their own side of the inlet.
+This is illustrated in Figure FIG.
 
@@
 The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term [CITE LOTS OF STUFF].
-Either the Lanczos method or the randomized singular value decomposition are used to perform the low rank approximation using only matrix-vector products.
+Either the Lanczos method or the randomized singular value decomposition [CITE RSVD] are used to perform the low rank approximation using only matrix-vector products.
 These methods suffer from a ``data predicament''---as more informative data is used in the inversion, the numerical rank of the data misfit term in the Hessian grows, so more matrix-vector products are required to form the aforementioned low rank approximation.
-The ideal scenario from a scientific perspective (highly informative data) is therefore the worst case scenario from a computational perspective (large computational cost) [CITE MY DISSERTATION].
-In our numerical results, we show that our operator approximation method can efficiently approximate the data misfit term in the Hessian in certain PDE constrained inverse problems.
-The method in this paper offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because the rank of the approximation is not limited by the number of matrix-vector products used to construct the approximation.
-High-rank approximations of an operator can be formed using a small number of matrix-vector products.
+The ideal scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost) [CITE MY DISSERTATION].
+%In our numerical results, we show that our operator approximation method can efficiently approximate the data misfit term in the Hessian in certain PDE constrained inverse problems. 
+The method in this paper offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because high-rank approximations of an operator can be formed using a small number of matrix-vector products.
 
@@
 With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
-This is computationally prohibitive in our applications, because computing the functions $\varphi_i$ requires one matrix vector product with $\mathcal{A}$ for each point $x_i$, and each of these matrix vector products is extremely computationally costly.
-One can reduce the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid only in the regions where the error in the approximation is large [CITE, CITE MY PAPER].
+This is computationally prohibitive in our applications, because computing the functions $\varphi_i$ requires one matrix vector product with $\mathcal{A}$ for each point $x_i$, and these matrix vector products are computationally costly.
+One can reduce the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large [CITE, CITE MY PAPER].
 But even with adaptive refinement, the required number of matrix vector products with $\mathcal{A}$ can be large.
-The key innovation in this paper is that we pick points $x_i$ in such a way that many $\varphi_i$ are computed with each matrix vector product.
+In this paper we reduce the computational cost by picking points $x_i$ such that many $\varphi_i$ are computed with each matrix vector product.
 We do this by taking advantage of the fact that the functions $\varphi_i$ are locally supported; existing methods typically do not require this local support property.
@@
 If the width of $\varphi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
-In [CITE], the width of $\varphi_x$ is estimated to be the local autocorrelation length of the function $\mathcal{A}^T \zeta$ near $x$, where $\zeta$ is a random noise funtion.
-In this paper, rather than probing $\mathcal{A}^T$ with random noise functions, we use specific linear and quadratic functions.
+In [CITE], the width of $\varphi_p$ is estimated to be the local autocorrelation length of the function $\mathcal{A}^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
+In this paper, rather than probing $\mathcal{A}^T$ with random noise functions, we use specific constant, linear, and quadratic functions.
 Our method estimates the support of $\varphi_p$ more accurately than resolution analysis, but our method requires that $\mathcal{A}$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
@@
 However, for complex geometries, problems are rarely discretized using regular grids, so we cannot easily use the fast Fourier transform.
-Furthermore, in big-data inverse problems the Hessian is typically highly ill-conditioned, so the number of Krylov iterations required is large.
-After constructing the product-convolution approximation, it is therefore desirable to convert the product-convolution approximation to other matrix formats that are more amenable to linear algebra operations.
+Furthermore, in big-data inverse problems the Hessian is typically ill-conditioned, so the required number of Krylov iterations is large.
+Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are more amenable to linear algebra operations.
 Wavelet compression methods have been used for this purpose [CITE WAVELETS].
@@
 There are matrix-free methods for H-matrix approximation [CITE LEXING PAPER, CITE MARTINSSON, CITE OTHERS], and these methods have been used to form H-matrix compressions of Hessians in PDE constrained inverse problems [CITE ILONA, CITE NOEMI/TUCKER HODLR].
-While these matrix-free H-matrix construction methods are asymptotically scalable in theory, the required number of matrix vector products is large in practice.
+While these matrix-free H-matrix construction methods are asymptotically scalable in theory, the required number of matrix vector products can be large in practice.
 
@@
 
+Consider the Pine island glacier geometry shown in Figure REF.
+Regions of the glacier are separated by thin inlets of ocean.
+With radial basis function interpolation, points on one side of the inlet inappropriately influence the interpolation on the other side of the inlet.
+With Poisson interpolation, the PSF samples only influence the interpolation on their own side of the inlet.
+
 \subsection{Determination of the boundary region}

### 2020-12-12 paper/localpsf.tex
more intro editing

@@
 This is computationally prohibitive in our applications, because computing the functions $\varphi_i$ requires one matrix vector product with $\mathcal{A}$ for each point $x_i$, and these matrix vector products are computationally costly.
-One can reduce the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large [CITE, CITE MY PAPER].
-But even with adaptive refinement, the required number of matrix vector products with $\mathcal{A}$ can be large.
-In this paper we reduce the computational cost by picking points $x_i$ such that many $\varphi_i$ are computed with each matrix vector product.
+In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large [CITE, CITE MY PAPER].
+But even with adaptive refinement, many matrix vector products with $\mathcal{A}$ may be required.
+In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\varphi_i$ are computed with each matrix vector product.
 We do this by taking advantage of the fact that the functions $\varphi_i$ are locally supported; existing methods typically do not require this local support property.
@@
 Furthermore, in big-data inverse problems the Hessian is typically ill-conditioned, so the required number of Krylov iterations is large.
-Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are more amenable to linear algebra operations.
+Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are amenable to fast linear algebra operations.
 Wavelet compression methods have been used for this purpose [CITE WAVELETS].

### 2020-12-12 paper/localpsf.tex
more intro editing

@@
 Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$.
-Since the Laplacian measures local deviation from flatness [cite Jacobsen], the functions $w_i$ are the flattest functions, in a least-squares sense, that interpolate the points $x_i$.
+Since the Laplacian measures local deviation from flatness [CITE JACOBSEN], the functions $w_i$ are the flattest functions, in a least-squares sense, that interpolate the points $x_i$.
 Computing the weighting functions $w_i$ requires solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid.

### 2020-12-17 paper/localpsf.tex
editing weighting functions section to be infinite dimensional

@@
 
-Let $\Omega \subset \mathbb{R}^d$ be a bounded domain, and let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the form
+Let $\Omega \subset \mathbb{R}^d$ be a bounded domain in dimension $d=1$, $2$, or $3$, and let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the form
 \begin{equation}
@@
 \begin{aligned}
-\min_{w_i} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
+\min_{w_i \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
 \text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
@@
 Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$.
-Since the Laplacian measures local deviation from flatness [CITE JACOBSEN], the functions $w_i$ are the flattest functions, in a least-squares sense, that interpolate the points $x_i$.
+Since the Laplacian measures local deviation from flatness [CITE JACOBSEN], the function $w_i$ is the flattest function, in a least-squares sense, that takes the value one at the sample point $x_i$, and the value zero at the other sample points $x_j$, $i \neq j$.
 Computing the weighting functions $w_i$ requires solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid.
@@
 
-\subsection{Background: discretization}
-\label{sec:discretization}
-
-Let $\phi_i$, $i=1,\dots,N$ be a set of finite element basis functions, and let
-\begin{equation*}
-	V_h = \Span\left(\phi_1, \phi_2, \dots, \phi_N\right).
-\end{equation*}
-be the corresponding  finite element space approximating $L^2(\Omega)$.
-Here $V_h$ inherits the $L^2$ inner product, and $h$ denotes the mesh size parameter.
-Functions $u\in L^2(\Omega)$ are approximated by functions $u_h \in V_h$.
-In turn, functions $u_h\in V_h$ are represented in computations by length $N$ arrays $\mathbf{u}$, such that the array entries of $\mathbf{u}$ are the coordinates of $u_h$ in the finite element basis.
-That is,
-\begin{equation*}
-	u(x) \approx u_h(x) = \sum_{i=1}^N \mathbf{u}_i \phi_i(x).
-\end{equation*}
-We say that the coefficient arrays $\mathbf{u}$ reside $\mathbb{R}^N_\mathbf{M}$, which is the space $\mathbb{R}^N$ with the matrix-weighted inner product $\left(\mathbf{u},\mathbf{v}\right)_\mathbf{M} := \left(\mathbf{M} \mathbf{u}\right)^T \mathbf{v}$, where
-$\mathbf{M} \in \mathbb{R}^{N \times N}$ is the \emph{mass matrix} which has entries
-\begin{equation*}
-	\mathbf{M}_{ij} = \left(\phi_i, \phi_j\right)_{L^2(\Omega)} = \int_{\Omega} \phi_i(x) \phi_j(x) dx.
-\end{equation*}
-Direct calculation shows that $V_h$ is isometrically isomorphic $\mathbb{R}^N_\mathbf{M}$, in the sense that functions $u_h \in V_h$ are in bijective correspondence to coefficient vectors $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, and $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u},\mathbf{v}\right)_\mathbf{M}$.
-Likewise, the dual space of continuous linear functionals $\psi_h:V_h \rightarrow \mathbb{R}$, denoted $V_h'$, is isometrically isomorphic to $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
-All operations on functions in $V_h$ and linear functionals in $V_h'$ may therefore be performed on coefficient vectors in $\mathbb{R}^N_\mathbf{M}$ and coefficient dual vectors in $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
-We summarize coefficient vector versions of the function space operation that are used in this paper in Appendix \ref{app:discretized_operations}.
-
 \subsection{Background: hierarchical matrices}
 
-\section{Sampling impulse responses}
+\section{Choosing sample points}
+\label{sec:choosing_sample_points}
 
 We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points.
+In this section we describe how we choose the points $x_i$.
 The challenge is to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$.
-
-The function $\varphi_x(y)$ is undefined when $x + y \notin \Omega$, which can occur even if $x \in \Omega$ and $y \in \Omega$.
-Extending $\varphi_x$ by zero makes it defined for all $y \in \mathbb{R}^d$, but this extension is only reasonable if $x$ is far enough from $\partial \Omega$ that $\varphi_x(y)$ decays to zero, as a function of $y$, before $y$ reaches $\partial \Omega$.
-We therefore choose sample points $x_i$ from a region $\Omega_I \subset \Omega$ which is not too close to the boundary.
-We determine $\Omega_I$ by applying $\mathcal{A}^T$ to a boundary source, then post processing the result.
 
@@
 The support of $\langle\mathcal{A},\delta_{x}\rangle^*$ is estimated to be contained in an ellipsoid of points that are within a small number of standard deviations of the mean for a Gaussian distribution with the same mean and covariance as the normalized version of $\langle\mathcal{A},\delta_{x}\rangle^*$.
+
+The function $\varphi_x(y)$ is undefined when $x + y \notin \Omega$, which can occur even if $x \in \Omega$ and $y \in \Omega$.
+Extending $\varphi_x$ by zero makes it defined for all $y \in \mathbb{R}^d$, but this extension is only reasonable if $x$ is far enough from $\partial \Omega$ that $\varphi_x(y)$ decays to zero, as a function of $y$, before $y$ reaches $\partial \Omega$.
+We therefore choose exclude sample points $x_i$ if the ellipsoid associated with that point overlaps the boundary.
 
@@
 
-\paragraph{Discretization} In computations, $\alpha$, $\mu^{i}$, and $\Sigma^{ij}$ are replaced with finite element approximations $\alpha_h, \mu_h^i, \Sigma_h^{ij}\in V_h$, respectively, which have coefficient vectors $\boldsymbol{\alpha}, \boldsymbol{\mu}^{ij}, \boldsymbol{\Sigma}^{ij}\in\mathbb{R}^N_\mathbf{M}$, respectively.
-The functions $C$, $L^i$, and $Q^{ij}$ are replaced with their $L^2$ projections onto $V_h$, which are denoted by $C_h$, $L_h^i$, and $Q_h^{ij}$, respectively.
-The corresponding coefficient vectors are $\mathbf{C}$, $\mathbf{L}^i$, and $\mathbf{Q}^{ij}$, respectively.
-The discretized versions of \eqref{eq:vol_mean_var_thm1}, \eqref{eq:vol_mean_var_thm2}, and \eqref{eq:vol_mean_var_thm3} are
-\begin{align*}
-	\boldsymbol{\alpha} &= \mathbf{M}^{-1}\mathbf{A}^T\mathbf{C} \\
-	\boldsymbol{\mu}^i &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{L}^i\right) / \boldsymbol{\alpha}\\
-	\boldsymbol{\Sigma}^{ij} &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{Q}^{ij}\right) / \boldsymbol{\alpha} - \boldsymbol{\mu}^i\cdot \boldsymbol{\mu}^j,
-\end{align*}
-respectively. Here $\mathbf{f} \cdot \mathbf{g}$ and $\mathbf{f} / \mathbf{g}$ denote the element-wise multiplication and division of vectors, respectively.
-
-\subsection{Choosing sample points}
-\label{sec:choosing_sample_points}
+\subsection{Greedy point selection algorithm}
+\label{sec:greedy_point_selection}
 
@@
 
-\paragraph{Discretization}
-In computations, the Dirac comb $\xi^b$ is replaced by its restriction to $V_h$, denoted $\xi^b \in V_h'$.
-The response $\eta^b$ is replaced by the finite element approximation $\eta_h^b \in V_h$.
-The coefficient dual vector for $\xi_h^b$ is the vector $\boldsymbol{\xi}^b \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has entries
-\begin{equation*}
-	\boldsymbol{\xi}^b_i = \sum_{x_j \in S_b} \phi_i(x_j), \quad i=1,\dots,N.
-\end{equation*}
-The coefficient vector $\boldsymbol{\eta}^b \in \mathbb{R}^N_\mathbf{M}$ for $\eta_h^b$ is given by
-\begin{equation*}
-	\boldsymbol{\eta}^b = \mathbf{M}^{-1} \mathbf{A}^T \mathbf{M}^{-1}\boldsymbol{\xi}^b.
-\end{equation*}
-
 \section{Poisson interpolation weighting functions}
@@
 Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$.
-Since the Laplacian measures local deviation from flatness [cite Jacobsen], the weighting function $w_i$ is the ``flattest possible'' function (in a least-squares sense) that has Neumann boundary conditions and takes the value $1$ at point $x_i$, and zero at the points $x_j$, $j \neq i$.
 In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense.
-Then in Section \ref{sec:computation_of_weighting_functions}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$.
+Theorem \ref{thm:smoothest_interpolant} relies on some straightforward prerequisites that we present in Definition \ref{def:stiffness_mass}, Proposition \ref{prop:equivalent_writing_of_weighting_optimization}, and Corollary \ref{cor:linearity_of_solution_operator}.
+In Theorem \ref{thm:wi_sum_to_one} we note that the weighting functions sum to one pointwise.
+
+In Section \ref{sec:computation_of_weighting_functions}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$.
 Constructing all $q$ of the functions $w_i$ requires solving the Poisson equation $2q$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid.
@@
 
-It is important to note that optimization problems \eqref{eq:wi_optimization_problem} only has a unique continuous solution when the spatial dimension is $1$, $2$, or $3$.
-When inverting for parameter fields defined on $4$ or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used\footnote{The reason for this dependence on spatial dimension is related to the fact that the fundamental solution of the biharmonic equation is continuous in $3$ or less dimensions, but singular in $4$ or more dimensions.
+It is important to note that functions in $H^2$ only have pointwise values if the spatial dimension is $d=1$, $2$, or $3$.
+Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-defined in three spatial dimensions or fewer.
+When inverting for parameter fields defined on four or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used\footnote{The reason for this dependence on spatial dimension is related to the fact that the fundamental solution of the biharmonic equation is continuous in $3$ or less dimensions, but singular in $4$ or more dimensions.
 In the unusual scenario where one is inverting for a parameter field in $4$ or more spatial dimensions, one could use a modified Poisson interpolation where $\Delta$ is replaced with $\Delta^\alpha$ in the objective function of \eqref{eq:wi_optimization_problem}, where $\alpha>1$ is chosen so that the fundamental solution of $\Delta^{2\alpha}$ is continuous.}.
+
+\begin{defn}
+	\label{def:stiffness_mass}
+	We define $\mathcal{K}:H^2(\Omega)\rightarrow H^{-2}(\Omega)$ and $\mathcal{M}:L^2(\Omega) \rightarrow L^2(\Omega)'$ to be the following stiffness and mass linear operators:
+	\begin{align*}
+	\left(\mathcal{K}u\right)(v) :=& \int_\Omega \nabla u \cdot \nabla v ~dx \\
+	\left(\mathcal{M}u\right)(v) :=& \int_\Omega u v ~dx,
+	\end{align*}
+	and $\mathcal{P}:H^2(\Omega)\rightarrow \mathbb{R}^r$ to be the following pointwise observation operator:
+	\begin{equation*}
+	(\mathcal{P}u)_i = u(x_i)
+	\end{equation*}
+	for $i=1,\dots,r$.
+\end{defn}
+
+\begin{prop}
+	\label{prop:equivalent_writing_of_weighting_optimization}
+	Let $f \in \mathbb{R}^r$ be arbitrary. The optimization problem
+	\begin{equation*}
+	\begin{aligned}
+	\min_{u \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
+	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r
+	\end{aligned}
+	\end{equation*}
+	may be rewritten as follows:
+	\begin{equation}
+		\label{eq:generic_optimization}
+		\begin{aligned}
+		\min_{u \in H^2(\Omega)} &\quad \frac{1}{2} \left(\mathcal{K}\mathcal{M}^{-1}\mathcal{K}u\right)(u) \\
+		\text{such that} &\quad 
+		\mathcal{P} u = f.
+		\end{aligned}
+	\end{equation}
+\end{prop}
+
+\begin{prop}
+	There exists a unique solution to optimization problem \eqref{eq:generic_optimization}.
+The solution to \eqref{eq:generic_optimization}, $u^*$, and the adjoint variable for enforcing the constraint, $\lambda^*$, solve the following Karun-Kush-Tucker (KKT) system:
+	\begin{equation}
+	\label{eq:weighting_kkt}
+		\begin{bmatrix}
+			\mathcal{K}\mathcal{M}^{-1}\mathcal{K} & \mathcal{P}^T \\
+			\mathcal{P} & 0
+		\end{bmatrix}
+		\begin{bmatrix}
+			u^* \\ \lambda^*
+		\end{bmatrix}
+		=
+		\begin{bmatrix}
+			0 \\ f
+		\end{bmatrix},
+	\end{equation}
+	where we use block-matrix notation to denote block-operators.
+\end{prop}
+
+\begin{cor}
+	\label{cor:linearity_of_solution_operator}
+	Let $\mathcal{S}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization}	as a function of $f$. That is, 
+	\begin{equation*}
+		\mathcal{S}(f) := u^*,
+	\end{equation*}
+	where $u^*$ is the minimizer of \eqref{eq:generic_optimization}. The operator $\mathcal{S}$ is a linear operator.
+\end{cor}
 
@@
 	\label{thm:smoothest_interpolant}
-Let $u$ be the function
+Here let $u$ be the function
 \begin{equation*}
@@
 \end{equation*}
-for an arbitrary but fixed value of $y$. Then $u$ solves the optimization problem
+for an arbitrary but fixed value of $y$, where the weighting functions $w_i$ are the solutions to optimization problems \eqref{eq:wi_optimization_problem}. Then $u$ solves the optimization problem
 \begin{equation}
@@
 \begin{proof}
-This follows from linearity of the KKT system associated with \eqref{eq:wi_optimization_problem}, which in turn follows from the fact that the the objective function in \eqref{eq:wi_optimization_problem} is quadratic, and the constraint is linear.
+	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$. The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
+	\begin{align*}
+		\mathcal{S}(f) &= \varphi_1(y) \mathcal{S}(\delta_{1j}) + \varphi_2(y) \mathcal{S}(\delta_{2j}) + \dots +\varphi_r(y) \mathcal{S}(\delta_{rj}) \\
+		&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
+		&= u
+	\end{align*}
+ 	because of linearity of the solution map, $\mathcal{S}$.
+\end{proof}
+
+\begin{thm}
+	\label{thm:wi_sum_to_one}
+	We have
+	\begin{equation*}
+		\sum_{i=1}^r w_i(x) = 1
+	\end{equation*}
+	for all $x \in \Omega$.
+\end{thm}
+
+\begin{proof}
+asdf
 \end{proof}
@@
 \end{algorithm2e}
-
-\begin{thm}
-	The weighting functions form a partition of unity.
-\end{thm}
 
@@
 
-\section{Discretized function space operations}
+\section{Discretization}
+\label{sec:discretization}
+
+Let $\phi_i$, $i=1,\dots,N$ be a set of finite element basis functions, and let
+\begin{equation*}
+V_h = \Span\left(\phi_1, \phi_2, \dots, \phi_N\right).
+\end{equation*}
+be the corresponding  finite element space approximating $L^2(\Omega)$.
+Here $V_h$ inherits the $L^2$ inner product, and $h$ denotes the mesh size parameter.
+Functions $u\in L^2(\Omega)$ are approximated by functions $u_h \in V_h$.
+In turn, functions $u_h\in V_h$ are represented in computations by length $N$ arrays $\mathbf{u}$, such that the array entries of $\mathbf{u}$ are the coordinates of $u_h$ in the finite element basis.
+That is,
+\begin{equation*}
+u(x) \approx u_h(x) = \sum_{i=1}^N \mathbf{u}_i \phi_i(x).
+\end{equation*}
+We say that the coefficient arrays $\mathbf{u}$ reside $\mathbb{R}^N_\mathbf{M}$, which is the space $\mathbb{R}^N$ with the matrix-weighted inner product $\left(\mathbf{u},\mathbf{v}\right)_\mathbf{M} := \left(\mathbf{M} \mathbf{u}\right)^T \mathbf{v}$, where
+$\mathbf{M} \in \mathbb{R}^{N \times N}$ is the \emph{mass matrix} which has entries
+\begin{equation*}
+\mathbf{M}_{ij} = \left(\phi_i, \phi_j\right)_{L^2(\Omega)} = \int_{\Omega} \phi_i(x) \phi_j(x) dx.
+\end{equation*}
+Direct calculation shows that $V_h$ is isometrically isomorphic $\mathbb{R}^N_\mathbf{M}$, in the sense that functions $u_h \in V_h$ are in bijective correspondence to coefficient vectors $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, and $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u},\mathbf{v}\right)_\mathbf{M}$.
+Likewise, the dual space of continuous linear functionals $\psi_h:V_h \rightarrow \mathbb{R}$, denoted $V_h'$, is isometrically isomorphic to $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
+All operations on functions in $V_h$ and linear functionals in $V_h'$ may therefore be performed on coefficient vectors in $\mathbb{R}^N_\mathbf{M}$ and coefficient dual vectors in $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
+We summarize coefficient vector versions of the function space operation that are used in this paper in Appendix \ref{app:discretized_operations}.
+
+\subsection{Discretized operations}
 \label{app:discretized_operations}
@@
 
+\subsection{Discretized mean and covariance estimation} In computations, $\alpha$, $\mu^{i}$, and $\Sigma^{ij}$ are replaced with finite element approximations $\alpha_h, \mu_h^i, \Sigma_h^{ij}\in V_h$, respectively, which have coefficient vectors $\boldsymbol{\alpha}, \boldsymbol{\mu}^{ij}, \boldsymbol{\Sigma}^{ij}\in\mathbb{R}^N_\mathbf{M}$, respectively.
+The functions $C$, $L^i$, and $Q^{ij}$ are replaced with their $L^2$ projections onto $V_h$, which are denoted by $C_h$, $L_h^i$, and $Q_h^{ij}$, respectively.
+The corresponding coefficient vectors are $\mathbf{C}$, $\mathbf{L}^i$, and $\mathbf{Q}^{ij}$, respectively.
+The discretized versions of \eqref{eq:vol_mean_var_thm1}, \eqref{eq:vol_mean_var_thm2}, and \eqref{eq:vol_mean_var_thm3} are
+\begin{align*}
+\boldsymbol{\alpha} &= \mathbf{M}^{-1}\mathbf{A}^T\mathbf{C} \\
+\boldsymbol{\mu}^i &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{L}^i\right) / \boldsymbol{\alpha}\\
+\boldsymbol{\Sigma}^{ij} &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{Q}^{ij}\right) / \boldsymbol{\alpha} - \boldsymbol{\mu}^i\cdot \boldsymbol{\mu}^j,
+\end{align*}
+respectively. Here $\mathbf{f} \cdot \mathbf{g}$ and $\mathbf{f} / \mathbf{g}$ denote the element-wise multiplication and division of vectors, respectively.
+
+\subsection{Discretized sampling of impulse responses}
+In computations, the Dirac comb $\xi^b$ is replaced by its restriction to $V_h$, denoted $\xi^b \in V_h'$.
+The response $\eta^b$ is replaced by the finite element approximation $\eta_h^b \in V_h$.
+The coefficient dual vector for $\xi_h^b$ is the vector $\boldsymbol{\xi}^b \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has entries
+\begin{equation*}
+\boldsymbol{\xi}^b_i = \sum_{x_j \in S_b} \phi_i(x_j), \quad i=1,\dots,N.
+\end{equation*}
+The coefficient vector $\boldsymbol{\eta}^b \in \mathbb{R}^N_\mathbf{M}$ for $\eta_h^b$ is given by
+\begin{equation*}
+\boldsymbol{\eta}^b = \mathbf{M}^{-1} \mathbf{A}^T \mathbf{M}^{-1}\boldsymbol{\xi}^b.
+\end{equation*}
+
 \section{old stuff}

### 2020-12-17 paper/localpsf.tex
editing weighting functions section to be infinite dimensional

@@
 Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-defined in three spatial dimensions or fewer.
-When inverting for parameter fields defined on four or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used\footnote{The reason for this dependence on spatial dimension is related to the fact that the fundamental solution of the biharmonic equation is continuous in $3$ or less dimensions, but singular in $4$ or more dimensions.
-In the unusual scenario where one is inverting for a parameter field in $4$ or more spatial dimensions, one could use a modified Poisson interpolation where $\Delta$ is replaced with $\Delta^\alpha$ in the objective function of \eqref{eq:wi_optimization_problem}, where $\alpha>1$ is chosen so that the fundamental solution of $\Delta^{2\alpha}$ is continuous.}.
+When inverting for parameter fields defined on four or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used.
 
@@
 	\label{cor:linearity_of_solution_operator}
-	Let $\mathcal{S}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization}	as a function of $f$. That is, 
-	\begin{equation*}
-		\mathcal{S}(f) := u^*,
-	\end{equation*}
-	where $u^*$ is the minimizer of \eqref{eq:generic_optimization}. The operator $\mathcal{S}$ is a linear operator.
+	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization}	as a function of $f$. That is, 
+	\begin{equation*}
+		\mathcal{T}(f) := u^*,
+	\end{equation*}
+	where $u^*$ is the minimizer of \eqref{eq:generic_optimization}. The operator $\mathcal{T}$ is a linear operator.
 \end{cor}
@@
 Here let $u$ be the function
-\begin{equation*}
+\begin{equation}
+\label{eq:u_defn_in_thm}
 u(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right)
-\end{equation*}
+\end{equation}
 for an arbitrary but fixed value of $y$, where the weighting functions $w_i$ are the solutions to optimization problems \eqref{eq:wi_optimization_problem}. Then $u$ solves the optimization problem
@@
 \begin{proof}
-	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$. The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
+	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.
+The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
 	\begin{align*}
-		\mathcal{S}(f) &= \varphi_1(y) \mathcal{S}(\delta_{1j}) + \varphi_2(y) \mathcal{S}(\delta_{2j}) + \dots +\varphi_r(y) \mathcal{S}(\delta_{rj}) \\
+		\mathcal{T}(f) &= \varphi_1(y) \mathcal{T}(e_1) + \varphi_2(y) \mathcal{T}(e_2) + \dots +\varphi_r(y) \mathcal{T}(e_r) \\
 		&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
-		&= u
+		&= u.
 	\end{align*}
- 	because of linearity of the solution map, $\mathcal{S}$.
+ 	In the first line we used linearity of the solution map, $\mathcal{T}$ (Corollary \ref{cor:linearity_of_solution_operator}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $u$ in \eqref{eq:u_defn_in_thm}.
 \end{proof}
@@
 \begin{proof}
-asdf
+	By linearity of $\mathcal{T}$ (Corollary \ref{cor:linearity_of_solution_operator}), we have
+	\begin{equation*}
+		\sum_{i=1}^r w_i(x) = \sum_{i=1}^r \mathcal{T}(e_i)(x) = \mathcal{T}\left(\sum_{i=1}^r e_i\right)(x) = \mathcal{T}(\mathbf{1})(x) = 1,
+	\end{equation*}
+	where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one.
+The last equality, $\mathcal{T}(\mathbf{1})(x) = 1$, holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint $C(x_i) = 1$ for all $x_i$, $i=1,\dots,r$, and yields the minimum possible value of zero for the objective function, $\frac{1}{2}\|\Delta C\|^2 = 0$.
 \end{proof}
@@
 
+\footnote{The reason for this dependence on spatial dimension is related to the fact that the fundamental solution of the biharmonic equation is continuous in $3$ or less dimensions, but singular in $4$ or more dimensions.
+In the unusual scenario where one is inverting for a parameter field in $4$ or more spatial dimensions, one could use a modified Poisson interpolation where $\Delta$ is replaced with $\Delta^\alpha$ in the objective function of \eqref{eq:wi_optimization_problem}, where $\alpha>1$ is chosen so that the fundamental solution of $\Delta^{2\alpha}$ is continuous.}
+
 % $\left(\mathcal{A}u\right)(v) = \left\langle v, \varphi_p \ast u\right\rangle_{L^2(\Omega)}$. Here $f \ast g$ denotes convolution of the functions $f$ and $g$, and the convolution kernel,

### 2020-12-17 paper/localpsf.tex
editing weighting functions section to be infinite dimensional

@@
 	Let $f \in \mathbb{R}^r$ be arbitrary. The optimization problem
-	\begin{equation*}
+	\begin{equation}
+	\label{eq:generic_optimization_initial}
 	\begin{aligned}
@@
 	\end{aligned}
-	\end{equation*}
+	\end{equation}
 	may be rewritten as follows:
@@
 	where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one.
-The last equality, $\mathcal{T}(\mathbf{1})(x) = 1$, holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint $C(x_i) = 1$ for all $x_i$, $i=1,\dots,r$, and yields the minimum possible value of zero for the objective function, $\frac{1}{2}\|\Delta C\|^2 = 0$.
+The last equality ($\mathcal{T}(\mathbf{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \mathbf{1}$, and yields the minimum possible value of zero for the objective function (i.e., $C(x_i) = 1$ for $i=1,\dots,r$,  and $\frac{1}{2}\|\Delta C\|^2 = 0$).
 \end{proof}

### 2020-12-17 paper/localpsf.tex
editing weighting functions section to be infinite dimensional

@@
 
-It is important to note that functions in $H^2$ only have pointwise values if the spatial dimension is $d=1$, $2$, or $3$.
+It is important to note that functions in $H^2$ only have well-defined pointwise values if the spatial dimension is $d=1$, $2$, or $3$.
 Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-defined in three spatial dimensions or fewer.
@@
 	\label{def:stiffness_mass}
-	We define $\mathcal{K}:H^2(\Omega)\rightarrow H^{-2}(\Omega)$ and $\mathcal{M}:L^2(\Omega) \rightarrow L^2(\Omega)'$ to be the following stiffness and mass linear operators:
+	We define $\mathcal{K}:H^2(\Omega)\rightarrow L^2(\Omega)'$ and $\mathcal{M}:L^2(\Omega) \rightarrow L^2(\Omega)'$ to be the following linear operators:
 	\begin{align*}
-	\left(\mathcal{K}u\right)(v) :=& \int_\Omega \nabla u \cdot \nabla v ~dx \\
+	\left(\mathcal{K}u\right)(v) :=& \int_\Omega v \Delta u ~dx \\
 	\left(\mathcal{M}u\right)(v) :=& \int_\Omega u v ~dx,
@@
 		\begin{aligned}
-		\min_{u \in H^2(\Omega)} &\quad \frac{1}{2} \left(\mathcal{K}\mathcal{M}^{-1}\mathcal{K}u\right)(u) \\
+		\min_{u \in H^2(\Omega)} &\quad \frac{1}{2} \left(\mathcal{K}^T\mathcal{M}^{-1}\mathcal{K}u\right)(u) \\
 		\text{such that} &\quad 
@@
 		\begin{bmatrix}
-			\mathcal{K}\mathcal{M}^{-1}\mathcal{K} & \mathcal{P}^T \\
+			\mathcal{K}^T\mathcal{M}^{-1}\mathcal{K} & \mathcal{P}^T \\
 			\mathcal{P} & 0
@@
 	where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one.
-The last equality ($\mathcal{T}(\mathbf{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \mathbf{1}$, and yields the minimum possible value of zero for the objective function (i.e., $C(x_i) = 1$ for $i=1,\dots,r$,  and $\frac{1}{2}\|\Delta C\|^2 = 0$).
+The last equality ($\mathcal{T}(\mathbf{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \mathbf{1}$, and yields the minimum possible value of zero for the objective function.
 \end{proof}
@@
 	\begin{equation*}
-		\mathbf{v} := K \mathbf{w}.
-	\end{equation*}
+		v := \mathcal{K} w_i,
+	\end{equation*}
+	and for the remainder of this proof let $e := e_i$.
 	Since the null-space of the Laplacian with pure Neumann boundary conditions is the set of all constant functions, we have
@@
 	\label{eq:definition_of_w}
-	\mathbf{w} = K^+ \mathbf{v} + \alpha \mathbf{c}
+	w_i = \mathcal{K}^+ v + \alpha C
 	\end{equation}
@@
 
-	Writing optimization problem \eqref{eq:discrete_poisson_optimization_1} in terms of $\mathbf{v}$ and $\alpha$ instead of $\mathbf{w}$ yields the following equivalent optimization problem:
+	Writing optimization problem \eqref{eq:wi_optimization_problem} in terms of $v$ and $\alpha$ instead of $w_i$ yields the following equivalent optimization problem:
 	\begin{equation}
@@
 	\begin{aligned}
-	\min_{\mathbf{v}, \alpha} &\quad \frac{1}{2}\mathbf{v}^T M^{-1} \mathbf{v} \\
-	\text{such that} &\quad EK^+ \mathbf{v} + \alpha \mathbf{1} = \mathbf{d}.
+	\min_{v \in L^2(\Omega)', \alpha \in \mathbb{R}} &\quad \frac{1}{2} v\left(\mathcal{M}^{-1} v\right) \\
+	\text{such that} &\quad \mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} = e.
 	\end{aligned}
@@
 	\begin{equation*}
-		\mathcal{L} = \frac{1}{2}\mathbf{v}^T M^{-1} \mathbf{v} + \boldsymbol{\lambda}^T\left(EK^+ \mathbf{v} + \alpha \mathbf{1} - \mathbf{d}\right).
+		\mathcal{L} = \frac{1}{2}v\left( \mathcal{M}^{-1} v\right) + \lambda^T\left(\mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} - e\right).
 	\end{equation*}
@@
 		\begin{bmatrix}
-			\nabla_\mathbf{v} \mathcal{L} \\
-			\nabla_{\boldsymbol{\lambda}} \mathcal{L} \\
+			\nabla_{v} \mathcal{L} \\
+			\nabla_{\lambda} \mathcal{L} \\
 			\nabla_\alpha \mathcal{L}
@@
 		= \begin{bmatrix}
-			M^{-1} \mathbf{v} + K^+ E^T \boldsymbol{\lambda} \\
-			EK^+\mathbf{v} + \alpha \mathbf{1} - \mathbf{d} \\
-			\mathbf{1}^T \boldsymbol{\lambda}
+			\mathcal{M}^{-1} v + \mathcal{K}^+ \mathcal{P}^T \lambda \\
+			\mathcal{P}\mathcal{K}^+v + \alpha \mathbf{1} - e \\
+			\mathbf{1}^T \lambda
 		\end{bmatrix}.
@@
 		\begin{bmatrix}
-			M^{-1} & K^+ E^T & 0 \\
-			EK^+ & 0 & \mathbf{1} \\
+			\mathcal{M}^{-1} & \mathcal{K}^+ \mathcal{P}^T & 0 \\
+			\mathcal{P}\mathcal{K}^+ & 0 & \mathbf{1} \\
 			0 & \mathbf{1}^T & 0
@@
 		\begin{bmatrix}
-			\mathbf{v} \\ \boldsymbol{\lambda} \\ \alpha
+			v \\ \lambda \\ \alpha
 		\end{bmatrix}
@@
 		\begin{bmatrix}
-			0 \\ \mathbf{d} \\ 0
+			0 \\ e \\ 0
 		\end{bmatrix}.
@@
 		\begin{bmatrix}
-			M^{-1} & K^+ E^T & 0 \\
+			\mathcal{M}^{-1} & \mathcal{K}^+ \mathcal{P}^T & 0 \\
 			& -S & \mathbf{1} \\
@@
 		\begin{bmatrix}
-			0 \\ \mathbf{y} \\ \mathbf{1}^T S^{-1} y
+			0 \\ e \\ \mathbf{1}^T S^{-1} e
 		\end{bmatrix}
 	\end{equation*}
-	where we recall that $S := \Psi^T M \Psi = EK^+ M K^+ E^T$. From this block triangular system, we read off the solution as:
+	where we recall that $S := \Psi^T \mathcal{M} \Psi = \mathcal{P}\mathcal{K}^+ \mathcal{M} \mathcal{K}^+ \mathcal{P}^T$. From this block triangular system, we read off the solution as:
 	\begin{align*}

### 2020-12-19 paper/localpsf.tex
editing weighting functions section to be infinite dimensional
figured out clear proof for infinite dimensional weighting function theorem

@@
 \begin{aligned}
-\min_{w_i \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
+\min_{w_i \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
 \text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
@@
 \end{equation}
-Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$.
+Here $\Delta$ is the Laplacian on $\Omega$, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$, and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions.
 In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense.
-Theorem \ref{thm:smoothest_interpolant} relies on some straightforward prerequisites that we present in Definition \ref{def:stiffness_mass}, Proposition \ref{prop:equivalent_writing_of_weighting_optimization}, and Corollary \ref{cor:linearity_of_solution_operator}.
 In Theorem \ref{thm:wi_sum_to_one} we note that the weighting functions sum to one pointwise.
-
-In Section \ref{sec:computation_of_weighting_functions}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$.
-Constructing all $q$ of the functions $w_i$ requires solving the Poisson equation $2q$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid.
-All the existing weighting functions, $\{w_i\}_{i=1}^q$, will change if a new point, $x_{q+1}$, is added.
-However, we present an efficient incremental algorithm for updating these weighting functions when a new point is added, so that we only need to perform two new Poisson PDE solves, rather than $2(q+1)$ new PDE solves.
-
-It is important to note that functions in $H^2$ only have well-defined pointwise values if the spatial dimension is $d=1$, $2$, or $3$.
+These theorems rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}
+
+In Algorithm \ref{alg:weighting_functions_incremental}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$.
+We prove \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}.
+Constructing all $r$ of the functions $w_i$ requires solving the Poisson equation $2r$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid.
+All the existing weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added.
+The Poisson PDE solves can be performed incrementally, so that we only need to perform two new Poisson PDE solves when a new point is added, rather than $2(r+1)$ new PDE solves.
+
+It is important to note that functions in $H^2_N(\Omega)$ only have well-defined pointwise values if the spatial dimension is $d=1$, $2$, or $3$.
 Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-defined in three spatial dimensions or fewer.
@@
 
-\begin{defn}
-	\label{def:stiffness_mass}
-	We define $\mathcal{K}:H^2(\Omega)\rightarrow L^2(\Omega)'$ and $\mathcal{M}:L^2(\Omega) \rightarrow L^2(\Omega)'$ to be the following linear operators:
-	\begin{align*}
-	\left(\mathcal{K}u\right)(v) :=& \int_\Omega v \Delta u ~dx \\
-	\left(\mathcal{M}u\right)(v) :=& \int_\Omega u v ~dx,
-	\end{align*}
-	and $\mathcal{P}:H^2(\Omega)\rightarrow \mathbb{R}^r$ to be the following pointwise observation operator:
-	\begin{equation*}
-	(\mathcal{P}u)_i = u(x_i)
-	\end{equation*}
-	for $i=1,\dots,r$.
-\end{defn}
-
-\begin{prop}
-	\label{prop:equivalent_writing_of_weighting_optimization}
-	Let $f \in \mathbb{R}^r$ be arbitrary. The optimization problem
+\begin{lem}
+	\label{lem:T_is_linear}
+	Let $f \in \mathbb{R}^r$. There exists a unique solution to the following optimization problem:
 	\begin{equation}
@@
 	\begin{aligned}
-	\min_{u \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
-	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r
+	\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
+	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r.
 	\end{aligned}
 	\end{equation}
-	may be rewritten as follows:
-	\begin{equation}
-		\label{eq:generic_optimization}
-		\begin{aligned}
-		\min_{u \in H^2(\Omega)} &\quad \frac{1}{2} \left(\mathcal{K}^T\mathcal{M}^{-1}\mathcal{K}u\right)(u) \\
-		\text{such that} &\quad 
-		\mathcal{P} u = f.
-		\end{aligned}
-	\end{equation}
-\end{prop}
-
-\begin{prop}
-	There exists a unique solution to optimization problem \eqref{eq:generic_optimization}.
-The solution to \eqref{eq:generic_optimization}, $u^*$, and the adjoint variable for enforcing the constraint, $\lambda^*$, solve the following Karun-Kush-Tucker (KKT) system:
-	\begin{equation}
-	\label{eq:weighting_kkt}
-		\begin{bmatrix}
-			\mathcal{K}^T\mathcal{M}^{-1}\mathcal{K} & \mathcal{P}^T \\
-			\mathcal{P} & 0
-		\end{bmatrix}
-		\begin{bmatrix}
-			u^* \\ \lambda^*
-		\end{bmatrix}
-		=
-		\begin{bmatrix}
-			0 \\ f
-		\end{bmatrix},
-	\end{equation}
-	where we use block-matrix notation to denote block-operators.
-\end{prop}
-
-\begin{cor}
-	\label{cor:linearity_of_solution_operator}
-	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization}	as a function of $f$. That is, 
+	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
 	\begin{equation*}
@@
 	\end{equation*}
-	where $u^*$ is the minimizer of \eqref{eq:generic_optimization}. The operator $\mathcal{T}$ is a linear operator.
-\end{cor}
+	where $u^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
+\end{lem}
 
@@
 	\label{thm:smoothest_interpolant}
-Here let $u$ be the function
+Let $y \in \mathbb{R}^d$, and let $u$ be the function
 \begin{equation}
@@
 \end{equation}
-for an arbitrary but fixed value of $y$, where the weighting functions $w_i$ are the solutions to optimization problems \eqref{eq:wi_optimization_problem}. Then $u$ solves the optimization problem
+where the weighting functions $w_i$ are the solutions to optimization problems \eqref{eq:wi_optimization_problem}. Then $u$ solves the optimization problem
 \begin{equation}
@@
 \begin{aligned}
-\min_{u} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
+\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
 \text{such that} &\quad u(x_i) = \varphi_i(y), \quad i=1,\dots,q.
@@
 	\end{align*}
- 	In the first line we used linearity of the solution map, $\mathcal{T}$ (Corollary \ref{cor:linearity_of_solution_operator}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $u$ in \eqref{eq:u_defn_in_thm}.
+ 	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $u$ in \eqref{eq:u_defn_in_thm}.
 \end{proof}
@@
 \end{proof}
+
+\begin{algorithm2e}
+	\SetAlgoNoLine
+	\SetKwInOut{Input}{Input}
+	\SetKwInOut{Output}{Output}
+	%	\Input{Function $v \mapsto \mathcal{H}^Tv$}
+
+	%	\mbox{}\\
+
+	%	\Output{Spatially varying mean $\mu:\Omega \rightarrow \mathbb{R}^d$, \\ 
+	%		Spatially varying covariance $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$}
+	%	\Begin{
+	{	
+		\For{$i=1,2, \dots, r$}{
+			Compute $\psi_i$ as the solution to the Poisson problem
+			\begin{equation*}
+			\Delta \psi_i = \delta_{x_i}, \quad \text{in }\Omega,
+			\end{equation*}
+			with Neumann boundary conditions, $\nu \cdot \nabla \psi_i = 0$ on $\partial \Omega$, and the average value zero condition, $\int_\Omega \psi_i(x) dx = 0$.
+
+			Compute $\theta_i$ as the solution to the Poisson problem
+			\begin{equation*}
+			\Delta \theta_i = \psi_i, \quad \text{in }\Omega,
+			\end{equation*}
+			with Neumann boundary conditions, $\nu \cdot \nabla \psi_i = 0$ on $\partial \Omega$, and the average value zero condition, $\int_\Omega \psi_i(x) dx = 0$.
+
+		}
+
+		Form the $r \times r$ matrix $S$ with entries
+		$$S_{ij} = \left(\psi_i, \psi_j\right)_{L^2(\Omega)}.$$
+
+		\For{$i=1,2,\dots,r$}{
+			Compute
+			$$\alpha = \frac{\mathbf{1}^T S^{-1} e_i}{\mathbf{1}^T S^{-1} \mathbf{1}},$$
+			where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one, and $e_i = (0,\dots,0,1,0,\dots,0)$ is the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.
+
+			Compute 
+			\begin{equation*}
+			\lambda = -S^{-1} \left(e_i - \alpha \mathbf{1} \right).
+			\end{equation*}
+
+			Form the weighting function $w_i$ given by
+			$$w_i(x) = \alpha - \sum_{k=1}^r \lambda_k \theta_k(x).$$
+		}
+	}
+	\caption{Compute weighting functions $w_i$, $i=1,\dots,r$.}
+	\label{alg:weighting_functions_incremental}
+\end{algorithm2e}
+
+\begin{thm}
+\label{thm:wi_alg_is_correct}
+	The weighting functions $w_i$ computed by Algorithm \ref{alg:weighting_functions_incremental} solve optimization problem \ref{eq:wi_optimization_problem}.
+\end{thm}
 
@@
 		= \begin{bmatrix}
-			\mathcal{M}^{-1} v + \mathcal{K}^+ \mathcal{P}^T \lambda \\
+			\mathcal{M}^{-1} v + \left(\mathcal{K}^+\right)^T \mathcal{P}^T \lambda \\
 			\mathcal{P}\mathcal{K}^+v + \alpha \mathbf{1} - e \\
@@
 		\begin{bmatrix}
-			\mathcal{M}^{-1} & \mathcal{K}^+ \mathcal{P}^T & 0 \\
+			\mathcal{M}^{-1} & \left(\mathcal{K}^+\right)^T \mathcal{P}^T & 0 \\
 			\mathcal{P}\mathcal{K}^+ & 0 & \mathbf{1} \\
@@
 		\begin{bmatrix}
-			\mathcal{M}^{-1} & \mathcal{K}^+ \mathcal{P}^T & 0 \\
+			\mathcal{M}^{-1} & \left(\mathcal{K}^+\right)^T \mathcal{P}^T & 0 \\
 			& -S & \mathbf{1} \\
@@
 	\end{equation*}
-	where we recall that $S := \Psi^T \mathcal{M} \Psi = \mathcal{P}\mathcal{K}^+ \mathcal{M} \mathcal{K}^+ \mathcal{P}^T$. From this block triangular system, we read off the solution as:
+	where we recall that $S := \Psi^T \mathcal{M} \Psi = \mathcal{P}\mathcal{K}^+ \mathcal{M} \left(\mathcal{K}^+\right)^T \mathcal{P}^T$. From this block triangular system, we read off the solution as:
 	\begin{align*}
@@
 The matrices $\Psi$ and $\Theta$ may be constructed incrementally as new points are added.
-
-\begin{algorithm2e}
-	\SetAlgoNoLine
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Output}
-	%	\Input{Function $v \mapsto \mathcal{H}^Tv$}
-
-	%	\mbox{}\\
-
-	%	\Output{Spatially varying mean $\mu:\Omega \rightarrow \mathbb{R}^d$, \\ 
-	%		Spatially varying covariance $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$}
-	%	\Begin{
-	{	
-		\For{$i=1,2, \dots, r$}{
-			Compute $\psi_i$ as the solution to the Poisson problem
-			\begin{equation*}
-				\Delta \psi_i = \delta_{x_i}, \quad \text{in }\Omega,
-			\end{equation*}
-			with Neumann boundary conditions, $\nu \cdot \nabla \psi_i = 0$ on $\partial \Omega$, and the average value zero condition, $\int_\Omega \psi_i(x) dx = 0$.
-
-			Compute $\theta_i$ as the solution to the Poisson problem
-			\begin{equation*}
-			\Delta \theta_i = \psi_i, \quad \text{in }\Omega,
-			\end{equation*}
-			with Neumann boundary conditions, $\nu \cdot \nabla \psi_i = 0$ on $\partial \Omega$, and the average value zero condition, $\int_\Omega \psi_i(x) dx = 0$.
-
-		}
-
-		Form the $r \times r$ matrix $S$ with entries
-		$$S_{ij} = \left(\psi_i, \psi_j\right)_{L^2(\Omega)}.$$
-
-		\For{$i=1,2,\dots,r$}{
-			Compute
-			$$\alpha = \frac{\mathbf{1}^T S^{-1} e_i}{\mathbf{1}^T S^{-1} \mathbf{1}},$$
-			where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one, and $e_i = (0,\dots,0,1,0,\dots,0)$ is the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.
-
-			Compute 
-			\begin{equation*}
-			\lambda = -S^{-1} \left(e_i - \alpha \mathbf{1} \right).
-			\end{equation*}
-
-			Form the weighting function
-			$$w_i = -\sum_{k=1}^r \lambda_k \theta_k + \alpha C,$$
-			where we recall that $C$ is the constant function $C(x)=1$ on $\Omega$.
-		}
-	}
-	\caption{Compute weighting functions $w_i$, $i=1,\dots,r$.}
-	\label{alg:weighting_functions_incremental}
-\end{algorithm2e}
 
@@
 
+\begin{defn}
+	\label{def:stiffness_mass}
+	We define $\mathcal{K}:H^2(\Omega)\rightarrow L^2(\Omega)'$ and $\mathcal{M}:L^2(\Omega) \rightarrow L^2(\Omega)'$ to be the following linear operators:
+	\begin{align*}
+	\left(\mathcal{K}u\right)(v) :=& \int_\Omega v \Delta u ~dx \\
+	\left(\mathcal{M}u\right)(v) :=& \int_\Omega u v ~dx,
+	\end{align*}
+	and $\mathcal{P}:H^2(\Omega)\rightarrow \mathbb{R}^r$ to be the following pointwise observation operator:
+	\begin{equation*}
+	(\mathcal{P}u)_i = u(x_i)
+	\end{equation*}
+	for $i=1,\dots,r$.
+\end{defn}
+
+\begin{prop}
+	\label{prop:solution_operator_is_linear}
+	Let $f \in \mathbb{R}^r$ be arbitrary. The optimization problem
+	\begin{equation}
+	\label{eq:generic_optimization_initial_old}
+	\begin{aligned}
+	\min_{u \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
+	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r
+	\end{aligned}
+	\end{equation}
+	may be rewritten as follows:
+	\begin{equation}
+	\label{eq:generic_optimization}
+	\begin{aligned}
+	\min_{u \in H^2(\Omega)} &\quad \frac{1}{2} \left(\mathcal{K}^T\mathcal{M}^{-1}\mathcal{K}u\right)(u) \\
+	\text{such that} &\quad 
+	\mathcal{P} u = f.
+	\end{aligned}
+	\end{equation}
+\end{prop}
+
+\begin{prop}
+	There exists a unique solution to optimization problem \eqref{eq:generic_optimization}.
+The solution to \eqref{eq:generic_optimization}, $u^*$, and the adjoint variable for enforcing the constraint, $\lambda^*$, solve the following Karun-Kush-Tucker (KKT) system:
+	\begin{equation}
+	\label{eq:weighting_kkt}
+	\begin{bmatrix}
+	\mathcal{K}^T\mathcal{M}^{-1}\mathcal{K} & \mathcal{P}^T \\
+	\mathcal{P} & 0
+	\end{bmatrix}
+	\begin{bmatrix}
+	u^* \\ \lambda^*
+	\end{bmatrix}
+	=
+	\begin{bmatrix}
+	0 \\ f
+	\end{bmatrix},
+	\end{equation}
+	where we use block-matrix notation to denote block-operators.
+\end{prop}
+
+\begin{cor}
+	\label{cor:linearity_of_solution_operator}
+	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization}	as a function of $f$. That is, 
+	\begin{equation*}
+	\mathcal{T}(f) := u^*,
+	\end{equation*}
+	where $u^*$ is the minimizer of \eqref{eq:generic_optimization}. The operator $\mathcal{T}$ is a linear operator.
+\end{cor}
+
 \footnote{The reason for this dependence on spatial dimension is related to the fact that the fundamental solution of the biharmonic equation is continuous in $3$ or less dimensions, but singular in $4$ or more dimensions.

### 2020-12-19 paper/localpsf.tex
figured out clear proof for infinite dimensional weighting function theorem

@@
 In Algorithm \ref{alg:weighting_functions_incremental}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$.
-We prove \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}.
+We prove correctness of Algorithm \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}.
 Constructing all $r$ of the functions $w_i$ requires solving the Poisson equation $2r$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid.
@@
 			Compute $\psi_i$ as the solution to the Poisson problem
-			\begin{equation*}
+			\begin{equation}
+			\label{eq:psi_eq}
 			\Delta \psi_i = \delta_{x_i}, \quad \text{in }\Omega,
-			\end{equation*}
-			with Neumann boundary conditions, $\nu \cdot \nabla \psi_i = 0$ on $\partial \Omega$, and the average value zero condition, $\int_\Omega \psi_i(x) dx = 0$.
+			\end{equation}
+			with Neumann zero boundary conditions, $\nu \cdot \nabla \psi_i = 0$ on $\partial \Omega$, and the average value zero condition, $\int_\Omega \psi_i(x) dx = 0$.
 
 			Compute $\theta_i$ as the solution to the Poisson problem
-			\begin{equation*}
+			\begin{equation}
+			\label{eq:theta_eq}
 			\Delta \theta_i = \psi_i, \quad \text{in }\Omega,
-			\end{equation*}
-			with Neumann boundary conditions, $\nu \cdot \nabla \psi_i = 0$ on $\partial \Omega$, and the average value zero condition, $\int_\Omega \psi_i(x) dx = 0$.
+			\end{equation}
+			with Neumann zero boundary conditions, $\nu \cdot \nabla \psi_i = 0$ on $\partial \Omega$, and the average value zero condition, $\int_\Omega \psi_i(x) dx = 0$.
 
@@
 		Form the $r \times r$ matrix $S$ with entries
-		$$S_{ij} = \left(\psi_i, \psi_j\right)_{L^2(\Omega)}.$$
+		\begin{equation}
+		\label{eq:defn_of_S_alg}
+		S_{ij} = \left(\psi_i, \psi_j\right)_{L^2(\Omega)}.
+		\end{equation}
 
@@
 			Compute 
-			\begin{equation*}
+			\begin{equation}
+			\label{eq:definition_of_mu_alg}
 			\lambda = -S^{-1} \left(e_i - \alpha \mathbf{1} \right).
-			\end{equation*}
+			\end{equation}
 
 			Form the weighting function $w_i$ given by
-			$$w_i(x) = \alpha - \sum_{k=1}^r \lambda_k \theta_k(x).$$
+			\begin{equation}
+			\label{eq:definition_of_wi_alg}
+				w_i(x) = \alpha - \sum_{k=1}^r \lambda_k \theta_k(x).
+			\end{equation}
 		}
@@
 \end{thm}
+
+\begin{proof}
+	For notational convenience, in this proof we write $w$, $\lambda$, and $e$ instead of $w_i$, $\lambda_i$, and $e_i$, respectively.
+Here subscripts for $\lambda$, and $e$ instead denote components of those vectors.
+For example, here $\lambda_j$ denotes the $j^\text{th}$ component of the vector $\lambda$, which corresponds to the $j^\text{th}$ of the vector $\lambda_i$ in Algorithm \ref{alg:weighting_functions_incremental}.
+
+	In this proof, we will repeatedly make use of the fact that
+	\begin{equation}
+	\label{eq:inline_wi_lemma}
+	\left(\psi_j, \Delta h\right)_{L^2(\Omega)} = h(x_j)
+	\end{equation}
+	for all $h\in H^2_N(\Omega)$. This follows from integration by parts and the definition of $\psi$ in \eqref{eq:psi_eq}.
+
+	From standard optimization theory, the solution to optimization problem \ref{eq:wi_optimization_problem} is the stationary point of the Lagrangian,
+	\begin{equation*}
+		\mathcal{L}(w,\lambda) = \frac{1}{2}\|\Delta w\|^2 + \sum_{j=1}^r \lambda_j\left(e_j - w(x_j)\right).
+	\end{equation*}
+	Using the definition of $w_i$ in \eqref{eq:definition_of_wi_alg}, we may write the gradient of $\mathcal{L}$ with respect to $\lambda_j$ as
+	\begin{equation*}
+		\nabla_{\lambda_j} \mathcal{L}(w, \lambda) = e_j - w(x_j) 
+		= e_j - \alpha + \sum_{k=1}^r \lambda_k \theta_k(x_j).
+	\end{equation*}
+	Using \eqref{eq:inline_wi_lemma} with $h=\theta_k$, the equation for $\theta$ in \eqref{eq:theta_eq}, and the definition of $S$ in \eqref{eq:defn_of_S_alg}, we have
+	\begin{align*}
+		\sum_{k=1}^r \lambda_k \theta_k(x_j) &= \sum_{k=1}^r \lambda_k \left(\psi_j, \Delta \theta_k\right)_{L^2(\Omega)} \\
+		&= \sum_{k=1}^r \lambda_k \left(\psi_j, \psi_k\right)_{L^2(\Omega)} = \left(S \lambda\right)_j.
+	\end{align*}
+	Combining all of these components of $\nabla_\lambda \mathcal{L}$ into a single vector and using the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} yields
+	\begin{equation*}
+		\nabla_\lambda \mathcal{L}(w, \lambda) = e - \alpha \mathbf{1} + S \lambda = e - \alpha \mathbf{1} + S \left(-S^{-1}\left(e - \alpha \mathbf{1}\right)\right) = 0.
+	\end{equation*}
+
+	Taking the directional derivative of $\mathcal{L}$ with respect to $w$ in direction $h$ and using \eqref{eq:inline_wi_lemma} yields
+	\begin{align*}
+		\nabla_w \mathcal{L}(w, \lambda)h &= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} - \sum_{j=1}^r \lambda_j h(x_j) \\
+		&= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} - \sum_{j=1}^r \lambda_j \left(\psi_j, \Delta h\right)_{L^2(\Omega)} \\
+		&= \left(\Delta w - \sum_{j=1}^r \lambda_j \psi_j,~ \Delta h\right)_{L^2(\Omega)}.
+	\end{align*}
+	Using the definition of $w$ in Algorithm \ref{alg:weighting_functions_incremental}, and the fact that the Laplacian of a constant function is zero, we have
+	\begin{align*}
+		\Delta w &= -\sum_{j=1}^r \lambda_j \Delta \theta_j \\
+		&= -\sum_{j=1}^r \lambda_j \psi_j
+	\end{align*}
+
+\end{proof}
 

### 2020-12-19 paper/localpsf.tex
finished infinite dimensional weighting function proof

@@
 In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense.
-In Theorem \ref{thm:wi_sum_to_one} we note that the weighting functions sum to one pointwise.
+In Proposition \ref{thm:wi_sum_to_one} we note that the weighting functions sum to one pointwise.
 These theorems rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}
@@
 We prove correctness of Algorithm \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}.
+Theorem \ref{thm:wi_alg_is_correct} relies on a prerequisite result which we present in Lemma \ref{lem:psi_delta_eval}.
 Constructing all $r$ of the functions $w_i$ requires solving the Poisson equation $2r$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid.
@@
 
-\begin{lem}
+\begin{lem}[Preparation for Theorem \ref{thm:smoothest_interpolant}]
 	\label{lem:T_is_linear}
@@
 
-\begin{thm}
+\begin{thm}[Optimal interpolation]
 	\label{thm:smoothest_interpolant}
@@
 
-\begin{thm}
+\begin{prop}
 	\label{thm:wi_sum_to_one}
@@
 	for all $x \in \Omega$.
-\end{thm}
+\end{prop}
 
@@
 			Compute
-			$$\alpha = \frac{\mathbf{1}^T S^{-1} e_i}{\mathbf{1}^T S^{-1} \mathbf{1}},$$
+			\begin{equation}
+			\label{eq:definition_of_alpha_alg}
+			\alpha = \frac{\mathbf{1}^T S^{-1} e_i}{\mathbf{1}^T S^{-1} \mathbf{1}},
+			\end{equation}
 			where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one, and $e_i = (0,\dots,0,1,0,\dots,0)$ is the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.
@@
 
-\begin{thm}
+\begin{lem}[Preparation for Theorem \ref{thm:wi_alg_is_correct}]
+	\label{lem:psi_delta_eval}
+	We have
+	\begin{equation}
+	\label{eq:inline_wi_lemma}
+	\left(\psi_j, \Delta h\right)_{L^2(\Omega)} = h(x_j) - \avg(h)
+	\end{equation}
+	for all $h\in H^2_N(\Omega)$, where
+	\begin{equation*}
+		\avg(h) :=	 \frac{1}{|\Omega|} \int_\Omega h ~dx.
+	\end{equation*}
+\end{lem}
+
+\begin{proof}
+	This follows from integration by parts and the definition of $\psi$ in \eqref{eq:psi_eq}.
+\end{proof}
+
+\begin{thm}[Correctness of Algorithm \ref{alg:weighting_functions_incremental}]
 \label{thm:wi_alg_is_correct}
@@
 
-	In this proof, we will repeatedly make use of the fact that
-	\begin{equation}
-	\label{eq:inline_wi_lemma}
-	\left(\psi_j, \Delta h\right)_{L^2(\Omega)} = h(x_j)
-	\end{equation}
-	for all $h\in H^2_N(\Omega)$. This follows from integration by parts and the definition of $\psi$ in \eqref{eq:psi_eq}.
-
 	From standard optimization theory, the solution to optimization problem \ref{eq:wi_optimization_problem} is the stationary point of the Lagrangian,
 	\begin{equation*}
-		\mathcal{L}(w,\lambda) = \frac{1}{2}\|\Delta w\|^2 + \sum_{j=1}^r \lambda_j\left(e_j - w(x_j)\right).
+		\mathcal{L}(w,\lambda) = \frac{1}{2}\|\Delta w\|^2 + \sum_{j=1}^r \lambda_j\left(w(x_j) - e_j\right).
 	\end{equation*}
@@
 	\begin{equation*}
-		\nabla_{\lambda_j} \mathcal{L}(w, \lambda) = e_j - w(x_j) 
-		= e_j - \alpha + \sum_{k=1}^r \lambda_k \theta_k(x_j).
-	\end{equation*}
-	Using \eqref{eq:inline_wi_lemma} with $h=\theta_k$, the equation for $\theta$ in \eqref{eq:theta_eq}, and the definition of $S$ in \eqref{eq:defn_of_S_alg}, we have
+		\nabla_{\lambda_j} \mathcal{L}(w, \lambda) = w(x_j) - e_j
+		= \alpha - \sum_{k=1}^r \lambda_k \theta_k(x_j) - e_j.
+	\end{equation*}
+	Using Lemma \ref{lem:psi_delta_eval} with $h=\theta_k$, the equation for $\theta_k$ in \eqref{eq:theta_eq}, and the definition of $S$ in \eqref{eq:defn_of_S_alg}, we have
 	\begin{align*}
-		\sum_{k=1}^r \lambda_k \theta_k(x_j) &= \sum_{k=1}^r \lambda_k \left(\psi_j, \Delta \theta_k\right)_{L^2(\Omega)} \\
+		\sum_{k=1}^r \lambda_k \theta_k(x_j) &= \sum_{k=1}^r \lambda_k \left(\psi_j, \Delta \theta_k\right)_{L^2(\Omega)}\\
 		&= \sum_{k=1}^r \lambda_k \left(\psi_j, \psi_k\right)_{L^2(\Omega)} = \left(S \lambda\right)_j.
@@
 	Combining all of these components of $\nabla_\lambda \mathcal{L}$ into a single vector and using the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} yields
-	\begin{equation*}
-		\nabla_\lambda \mathcal{L}(w, \lambda) = e - \alpha \mathbf{1} + S \lambda = e - \alpha \mathbf{1} + S \left(-S^{-1}\left(e - \alpha \mathbf{1}\right)\right) = 0.
-	\end{equation*}
-
-	Taking the directional derivative of $\mathcal{L}$ with respect to $w$ in direction $h$ and using \eqref{eq:inline_wi_lemma} yields
 	\begin{align*}
-		\nabla_w \mathcal{L}(w, \lambda)h &= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} - \sum_{j=1}^r \lambda_j h(x_j) \\
-		&= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} - \sum_{j=1}^r \lambda_j \left(\psi_j, \Delta h\right)_{L^2(\Omega)} \\
-		&= \left(\Delta w - \sum_{j=1}^r \lambda_j \psi_j,~ \Delta h\right)_{L^2(\Omega)}.
+		\nabla_\lambda \mathcal{L}(w, \lambda) &= \alpha \mathbf{1} - e - S \lambda \\
+		&= \alpha \mathbf{1} - e - S \left(-S^{-1}\left(e - \alpha \mathbf{1}\right)\right) = 0.
 	\end{align*}
+
+	Taking the directional derivative of $\mathcal{L}$ with respect to $w$ in direction $h$ and using Lemma \ref{lem:psi_delta_eval} yields
+	\begin{align}
+		\nabla_w \mathcal{L}(w, \lambda)h &= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j h(x_j) \nonumber \\
+		&= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \lambda_j \avg(h) \nonumber \\
+		&= \left(\Delta w + \sum_{j=1}^r \lambda_j \psi_j,~ \Delta h\right)_{L^2(\Omega)} + \avg(h) \mathbf{1}^T \lambda. \label{eq:L_w_38}
+	\end{align}
 	Using the definition of $w$ in Algorithm \ref{alg:weighting_functions_incremental}, and the fact that the Laplacian of a constant function is zero, we have
+	\begin{equation*}
+		\Delta w = -\sum_{j=1}^r \lambda_j \Delta \theta_j 
+		= -\sum_{j=1}^r \lambda_j \psi_j,
+	\end{equation*}
+	which implies that the first term in \eqref{eq:L_w_38} is zero. By the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} and $\alpha$ in \eqref{eq:definition_of_alpha_alg}, we have
 	\begin{align*}
-		\Delta w &= -\sum_{j=1}^r \lambda_j \Delta \theta_j \\
-		&= -\sum_{j=1}^r \lambda_j \psi_j
+		\mathbf{1}^T \lambda &= -\mathbf{1}^T S^{-1} \left(e - \alpha \mathbf{1}\right) \\
+		&= -\mathbf{1}^T S^{-1} e + \alpha \mathbf{1}^T S^{-1} \mathbf{1} \\
+		&= -\mathbf{1}^T S^{-1} e + \frac{\mathbf{1}^T S^{-1} e}{\mathbf{1}^T S^{-1} \mathbf{1}} \mathbf{1}^T S^{-1} \mathbf{1} = 0,
 	\end{align*}
-
+	which implies that the second term in \eqref{eq:L_w_38} is zero.
+Thus $\nabla_w \mathcal{L}(w, \lambda) = 0$.
+Since $\nabla \mathcal{L}(w, \lambda)=0$, Algorithm \ref{alg:weighting_functions_incremental} correctly produces the solution to optimization problem \ref{eq:wi_optimization_problem}.
 \end{proof}

### 2020-12-19 paper/localpsf.tex
finished infinite dimensional weighting function proof

@@
 In Proposition \ref{thm:wi_sum_to_one} we note that the weighting functions sum to one pointwise.
-These theorems rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}
+These results rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}
 
@@
 \end{equation}
-where the weighting functions $w_i$ are the solutions to optimization problems \eqref{eq:wi_optimization_problem}. Then $u$ solves the optimization problem
+where the weighting functions $w_i$ are the solutions to optimization problem \eqref{eq:wi_optimization_problem}. Then $u$ solves the optimization problem
 \begin{equation}
@@
 \begin{proof}
-	By linearity of $\mathcal{T}$ (Corollary \ref{cor:linearity_of_solution_operator}), we have
+	By linearity of $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), we have
 	\begin{equation*}
@@
 \begin{proof}
-	This follows from integration by parts and the definition of $\psi$ in \eqref{eq:psi_eq}.
+	This is a standard result that follows from integration by parts and the definition of $\psi$ in \eqref{eq:psi_eq}.
 \end{proof}
@@
 	\end{equation*}
+	We will prove the theorem by showing that $\nabla \mathcal{L}(w,\lambda)=0$.
+
 	Using the definition of $w_i$ in \eqref{eq:definition_of_wi_alg}, we may write the gradient of $\mathcal{L}$ with respect to $\lambda_j$ as
@@
 	which implies that the second term in \eqref{eq:L_w_38} is zero.
-Thus $\nabla_w \mathcal{L}(w, \lambda) = 0$.
-Since $\nabla \mathcal{L}(w, \lambda)=0$, Algorithm \ref{alg:weighting_functions_incremental} correctly produces the solution to optimization problem \ref{eq:wi_optimization_problem}.
+Thus $\nabla_w \mathcal{L}(w, \lambda) = 0$, so Algorithm \ref{alg:weighting_functions_incremental} produces the solution to optimization problem \ref{eq:wi_optimization_problem}.
 \end{proof}
@@
 % One may also show that the weighting functions $w_i$ sum to one pointwise.
-
-\subsection{Computation of Poisson weighting functions}
-\label{sec:computation_of_weighting_functions}
-
-%Radial basis function interpolation of $\varphi_x(y)$ is cheaper per sample point because radial basis functions are given by analytic formulas.
-However, the cost of these Poisson PDE solves is typically cheap compared to the cost of computing the impulse response samples, $\varphi_{x_i}$.
-Computing the impulse response samples, $\varphi_{x_i}$, requires applying the Hessian to vectors, which in turn requires performing PDE solves of the linearized forward and ajoint equations for the inverse problem.
-These forward and adjoint PDE solves are typically much more expensive than Poisson PDE solves, because the Poisson equation is well-studied and easy to solve with multigrid.
-
-After discretization, optimization problem \eqref{eq:wi_optimization_problem} becomes
-\begin{equation}
-\label{eq:discrete_poisson_optimization_1}
-\begin{aligned}
-\min_{\mathbf{w}} &\quad \frac{1}{2}\mathbf{w}^T K^T M^{-1} K \mathbf{w} \\
-\text{such that} &\quad E \mathbf{w} = \mathbf{d}.
-\end{aligned}
-\end{equation}
-Here $\mathbf{w}$ is the discretization of $w_i$, $K$ is the stiffness matrix for the discretized Poisson problem with pure Neumann boundary conditions, $M$ is the mass matrix, and $\mathbf{d}=(0,\dots,0,1,0,\dots,0)$ is the length-$q$ vector with $i^\text{th}$ entry one and all other entries zero.
-Furthermore, $E$ is the pointwise observation matrix defined by $\left(E\mathbf{u}\right)_j = u(x_j)$, where $u$ denotes the function represented by the vector $\mathbf{u}$.
-Here we dropped the subscripts $i$ from variables for notational convenience.
-One must solve $q$ optimization problems of the form \eqref{eq:discrete_poisson_optimization_1}; one for each $i=1,\dots,q$.
-
-In the remainder of this section we will derive an efficient algorithm for solving \eqref{eq:discrete_poisson_optimization_1}.
-The algorithm (Algorithm REF) is based on explicit formula for the solution of \eqref{eq:discrete_poisson_optimization_1} which we present in Proposition \ref{prop:poisson_interpolation_formula}.
-The formulas in Proposition \ref{prop:poisson_interpolation_formula} rely on certain matrices that we define in Definition \ref{defn:poisson_interpolation_matrices}.
-The primary computational cost of the algorithm is the solution of Poisson problems with pure Neumann boundary conditions.
-
-\begin{defn}
-	\label{defn:poisson_interpolation_matrices}
-	Let 
-	\begin{align*}
-		\Psi :=& K^+ E^T, \\
-		\Theta :=& K^+ M \Psi \\
-		S :=& \Psi^T M \Psi,
-	\end{align*}
-	where $K^+$ is the Moore-Penrose pseudoinverse of $K$.
-Further, let $\mathbf{c}$ dente the discretization of the constant function $c(x)=1$ on $\Omega$, and let $\mathbf{1}:=(1,\dots,1)$ denote the vector of length $q$ with all entries equal to one.
-\end{defn}
-
-\begin{remark}
-	Computing $\mathbf{g} = K^+\mathbf{f}$ is equivalent to solving the discretized Poisson problem with $f$ as the right hand side source, with pure Neumann boundary conditions, and with the constraint that the solution has average value equal to zero.
-That is,
-	\begin{equation}
-	\label{eq:pure_neumann_poisson}
-		\mathbf{g} = K^+ \mathbf{f} \quad \Leftrightarrow \quad 
-		\begin{cases}
-			\Delta g = f, & \text{in }\Omega,\\
-			\nu \cdot \nabla g = 0, & \text{on }\partial \Omega,\\
-			\int_\Omega g(x) dx = 0,
-		\end{cases}
-	\end{equation}
-	where $\nu$ denotes the normal to the boundary.
-
-	Hence, the $i^\text{th}$ column of $\Psi$ is the result of solving the discretized pure Neumann Poisson problem, with a point source at location $x_i$.
-The $i^\text{th}$ column of $\Theta$ is the result of solving the discretized pure Neumann Poisson problem with distributed source given by the $i^\text{th}$ column of $\Psi$.
-\end{remark}
-
-\begin{prop}
-	\label{prop:poisson_interpolation_formula}
-	The solution to optimization problem \eqref{eq:discrete_poisson_optimization_1} is given by
-	\begin{equation*}
-		\mathbf{w} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c},
-	\end{equation*}
-	where
-	\begin{align*}
-		\alpha =& \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{d} \\
-		\boldsymbol{\lambda} =& -S^{-1} \left(\mathbf{d} - \alpha \mathbf{1} \right). \\
-	\end{align*}
-\end{prop}
-
-\begin{proof}
-	Let
-	\begin{equation*}
-		v := \mathcal{K} w_i,
-	\end{equation*}
-	and for the remainder of this proof let $e := e_i$.
-	Since the null-space of the Laplacian with pure Neumann boundary conditions is the set of all constant functions, we have
-	\begin{equation}
-	\label{eq:definition_of_w}
-	w_i = \mathcal{K}^+ v + \alpha C
-	\end{equation}
-	for some scalar $\alpha$.
-
-	Writing optimization problem \eqref{eq:wi_optimization_problem} in terms of $v$ and $\alpha$ instead of $w_i$ yields the following equivalent optimization problem:
-	\begin{equation}
-	\label{eq:discrete_poisson_optimization_2}
-	\begin{aligned}
-	\min_{v \in L^2(\Omega)', \alpha \in \mathbb{R}} &\quad \frac{1}{2} v\left(\mathcal{M}^{-1} v\right) \\
-	\text{such that} &\quad \mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} = e.
-	\end{aligned}
-	\end{equation}
-	The Lagrangian for optimization problem \eqref{eq:discrete_poisson_optimization_2} is
-	\begin{equation*}
-		\mathcal{L} = \frac{1}{2}v\left( \mathcal{M}^{-1} v\right) + \lambda^T\left(\mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} - e\right).
-	\end{equation*}
-	The solution to this optimization problem is the stationary point of the Lagrangian, i.e., the point at which the gradient of the Lagrangian is zero:
-	\begin{equation*}
-		0 = \nabla \mathcal{L} = 
-		\begin{bmatrix}
-			\nabla_{v} \mathcal{L} \\
-			\nabla_{\lambda} \mathcal{L} \\
-			\nabla_\alpha \mathcal{L}
-		\end{bmatrix} 
-		= \begin{bmatrix}
-			\mathcal{M}^{-1} v + \left(\mathcal{K}^+\right)^T \mathcal{P}^T \lambda \\
-			\mathcal{P}\mathcal{K}^+v + \alpha \mathbf{1} - e \\
-			\mathbf{1}^T \lambda
-		\end{bmatrix}.
-	\end{equation*}
-	This equation may be rewritten as the following block $3 \times 3$ linear system:
-	\begin{equation*}
-		\begin{bmatrix}
-			\mathcal{M}^{-1} & \left(\mathcal{K}^+\right)^T \mathcal{P}^T & 0 \\
-			\mathcal{P}\mathcal{K}^+ & 0 & \mathbf{1} \\
-			0 & \mathbf{1}^T & 0
-		\end{bmatrix}
-		\begin{bmatrix}
-			v \\ \lambda \\ \alpha
-		\end{bmatrix}
-		=
-		\begin{bmatrix}
-			0 \\ e \\ 0
-		\end{bmatrix}.
-	\end{equation*}
-	Performing block Gaussian elimination on this system allows us to reduce the system to the following block triangular form
-	\begin{equation*}
-		\begin{bmatrix}
-			\mathcal{M}^{-1} & \left(\mathcal{K}^+\right)^T \mathcal{P}^T & 0 \\
-			& -S & \mathbf{1} \\
-			0 & 0 & \mathbf{1}^T S^{-1} \mathbf{1}
-		\end{bmatrix}
-		\begin{bmatrix}
-			v \\ \lambda \\ \alpha
-		\end{bmatrix}
-		=
-		\begin{bmatrix}
-			0 \\ e \\ \mathbf{1}^T S^{-1} e
-		\end{bmatrix}
-	\end{equation*}
-	where we recall that $S := \Psi^T \mathcal{M} \Psi = \mathcal{P}\mathcal{K}^+ \mathcal{M} \left(\mathcal{K}^+\right)^T \mathcal{P}^T$. From this block triangular system, we read off the solution as:
-	\begin{align*}
-		\alpha =& \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{d} \\
-		\boldsymbol{\lambda} =& -S^{-1} \left(\mathbf{d} - \alpha \mathbf{1} \right) \\
-		\mathbf{v} =& -M K^+ E^T \boldsymbol{\lambda}
-	\end{align*}
-	Substituting these results into the definition of $\mathbf{w}$ in \eqref{eq:definition_of_w}, we have
-	\begin{equation*}
-		\mathbf{w} = -K^+ M K^+ E^T \boldsymbol{\lambda} + \alpha \mathbf{c} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c},
-	\end{equation*}
-	as required.
-\end{proof}
-
-We now design an incremental algorithm for constructing interpolants $\mathbf{u}$, in which we can add points $x_i$ more efficiently. Let
-\begin{align*}
-	\Psi :=& K^+ E^T \\
-	\Theta :=& K^+ M \Psi
-\end{align*}
-so that 
-\begin{equation*}
-	S = \Psi^T M \Psi
-\end{equation*}
-and
-\begin{equation*}
-	\mathbf{u} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c}
-\end{equation*}
-The columns of $\Psi$ are the responses of $A^+$ to point sources (delta distributions) centered at the sample points $x_i$.
-The columns of $\Theta$ are the responses of $A^+M^{-1} A$ to these point sources.
-The matrices $\Psi$ and $\Theta$ may be constructed incrementally as new points are added.
 
@@
 
+\subsection{Computation of Poisson weighting functions}
+\label{sec:computation_of_weighting_functions}
+
+%Radial basis function interpolation of $\varphi_x(y)$ is cheaper per sample point because radial basis functions are given by analytic formulas.
+However, the cost of these Poisson PDE solves is typically cheap compared to the cost of computing the impulse response samples, $\varphi_{x_i}$.
+Computing the impulse response samples, $\varphi_{x_i}$, requires applying the Hessian to vectors, which in turn requires performing PDE solves of the linearized forward and ajoint equations for the inverse problem.
+These forward and adjoint PDE solves are typically much more expensive than Poisson PDE solves, because the Poisson equation is well-studied and easy to solve with multigrid.
+
+After discretization, optimization problem \eqref{eq:wi_optimization_problem} becomes
+\begin{equation}
+\label{eq:discrete_poisson_optimization_1}
+\begin{aligned}
+\min_{\mathbf{w}} &\quad \frac{1}{2}\mathbf{w}^T K^T M^{-1} K \mathbf{w} \\
+\text{such that} &\quad E \mathbf{w} = \mathbf{d}.
+\end{aligned}
+\end{equation}
+Here $\mathbf{w}$ is the discretization of $w_i$, $K$ is the stiffness matrix for the discretized Poisson problem with pure Neumann boundary conditions, $M$ is the mass matrix, and $\mathbf{d}=(0,\dots,0,1,0,\dots,0)$ is the length-$q$ vector with $i^\text{th}$ entry one and all other entries zero.
+Furthermore, $E$ is the pointwise observation matrix defined by $\left(E\mathbf{u}\right)_j = u(x_j)$, where $u$ denotes the function represented by the vector $\mathbf{u}$.
+Here we dropped the subscripts $i$ from variables for notational convenience.
+One must solve $q$ optimization problems of the form \eqref{eq:discrete_poisson_optimization_1}; one for each $i=1,\dots,q$.
+
+In the remainder of this section we will derive an efficient algorithm for solving \eqref{eq:discrete_poisson_optimization_1}.
+The algorithm (Algorithm REF) is based on explicit formula for the solution of \eqref{eq:discrete_poisson_optimization_1} which we present in Proposition \ref{prop:poisson_interpolation_formula}.
+The formulas in Proposition \ref{prop:poisson_interpolation_formula} rely on certain matrices that we define in Definition \ref{defn:poisson_interpolation_matrices}.
+The primary computational cost of the algorithm is the solution of Poisson problems with pure Neumann boundary conditions.
+
+\begin{defn}
+	\label{defn:poisson_interpolation_matrices}
+	Let 
+	\begin{align*}
+	\Psi :=& K^+ E^T, \\
+	\Theta :=& K^+ M \Psi \\
+	S :=& \Psi^T M \Psi,
+	\end{align*}
+	where $K^+$ is the Moore-Penrose pseudoinverse of $K$.
+Further, let $\mathbf{c}$ dente the discretization of the constant function $c(x)=1$ on $\Omega$, and let $\mathbf{1}:=(1,\dots,1)$ denote the vector of length $q$ with all entries equal to one.
+\end{defn}
+
+\begin{remark}
+	Computing $\mathbf{g} = K^+\mathbf{f}$ is equivalent to solving the discretized Poisson problem with $f$ as the right hand side source, with pure Neumann boundary conditions, and with the constraint that the solution has average value equal to zero.
+That is,
+	\begin{equation}
+	\label{eq:pure_neumann_poisson}
+	\mathbf{g} = K^+ \mathbf{f} \quad \Leftrightarrow \quad 
+	\begin{cases}
+	\Delta g = f, & \text{in }\Omega,\\
+	\nu \cdot \nabla g = 0, & \text{on }\partial \Omega,\\
+	\int_\Omega g(x) dx = 0,
+	\end{cases}
+	\end{equation}
+	where $\nu$ denotes the normal to the boundary.
+
+	Hence, the $i^\text{th}$ column of $\Psi$ is the result of solving the discretized pure Neumann Poisson problem, with a point source at location $x_i$.
+The $i^\text{th}$ column of $\Theta$ is the result of solving the discretized pure Neumann Poisson problem with distributed source given by the $i^\text{th}$ column of $\Psi$.
+\end{remark}
+
+\begin{prop}
+	\label{prop:poisson_interpolation_formula}
+	The solution to optimization problem \eqref{eq:discrete_poisson_optimization_1} is given by
+	\begin{equation*}
+	\mathbf{w} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c},
+	\end{equation*}
+	where
+	\begin{align*}
+	\alpha =& \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{d} \\
+	\boldsymbol{\lambda} =& -S^{-1} \left(\mathbf{d} - \alpha \mathbf{1} \right). \\
+	\end{align*}
+\end{prop}
+
+\begin{proof}
+	Let
+	\begin{equation*}
+	v := \mathcal{K} w_i,
+	\end{equation*}
+	and for the remainder of this proof let $e := e_i$.
+	Since the null-space of the Laplacian with pure Neumann boundary conditions is the set of all constant functions, we have
+	\begin{equation}
+	\label{eq:definition_of_w}
+	w_i = \mathcal{K}^+ v + \alpha C
+	\end{equation}
+	for some scalar $\alpha$.
+
+	Writing optimization problem \eqref{eq:wi_optimization_problem} in terms of $v$ and $\alpha$ instead of $w_i$ yields the following equivalent optimization problem:
+	\begin{equation}
+	\label{eq:discrete_poisson_optimization_2}
+	\begin{aligned}
+	\min_{v \in L^2(\Omega)', \alpha \in \mathbb{R}} &\quad \frac{1}{2} v\left(\mathcal{M}^{-1} v\right) \\
+	\text{such that} &\quad \mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} = e.
+	\end{aligned}
+	\end{equation}
+	The Lagrangian for optimization problem \eqref{eq:discrete_poisson_optimization_2} is
+	\begin{equation*}
+	\mathcal{L} = \frac{1}{2}v\left( \mathcal{M}^{-1} v\right) + \lambda^T\left(\mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} - e\right).
+	\end{equation*}
+	The solution to this optimization problem is the stationary point of the Lagrangian, i.e., the point at which the gradient of the Lagrangian is zero:
+	\begin{equation*}
+	0 = \nabla \mathcal{L} = 
+	\begin{bmatrix}
+	\nabla_{v} \mathcal{L} \\
+	\nabla_{\lambda} \mathcal{L} \\
+	\nabla_\alpha \mathcal{L}
+	\end{bmatrix} 
+	= \begin{bmatrix}
+	\mathcal{M}^{-1} v + \left(\mathcal{K}^+\right)^T \mathcal{P}^T \lambda \\
+	\mathcal{P}\mathcal{K}^+v + \alpha \mathbf{1} - e \\
+	\mathbf{1}^T \lambda
+	\end{bmatrix}.
+	\end{equation*}
+	This equation may be rewritten as the following block $3 \times 3$ linear system:
+	\begin{equation*}
+	\begin{bmatrix}
+	\mathcal{M}^{-1} & \left(\mathcal{K}^+\right)^T \mathcal{P}^T & 0 \\
+	\mathcal{P}\mathcal{K}^+ & 0 & \mathbf{1} \\
+	0 & \mathbf{1}^T & 0
+	\end{bmatrix}
+	\begin{bmatrix}
+	v \\ \lambda \\ \alpha
+	\end{bmatrix}
+	=
+	\begin{bmatrix}
+	0 \\ e \\ 0
+	\end{bmatrix}.
+	\end{equation*}
+	Performing block Gaussian elimination on this system allows us to reduce the system to the following block triangular form
+	\begin{equation*}
+	\begin{bmatrix}
+	\mathcal{M}^{-1} & \left(\mathcal{K}^+\right)^T \mathcal{P}^T & 0 \\
+	& -S & \mathbf{1} \\
+	0 & 0 & \mathbf{1}^T S^{-1} \mathbf{1}
+	\end{bmatrix}
+	\begin{bmatrix}
+	v \\ \lambda \\ \alpha
+	\end{bmatrix}
+	=
+	\begin{bmatrix}
+	0 \\ e \\ \mathbf{1}^T S^{-1} e
+	\end{bmatrix}
+	\end{equation*}
+	where we recall that $S := \Psi^T \mathcal{M} \Psi = \mathcal{P}\mathcal{K}^+ \mathcal{M} \left(\mathcal{K}^+\right)^T \mathcal{P}^T$. From this block triangular system, we read off the solution as:
+	\begin{align*}
+	\alpha =& \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{d} \\
+	\boldsymbol{\lambda} =& -S^{-1} \left(\mathbf{d} - \alpha \mathbf{1} \right) \\
+	\mathbf{v} =& -M K^+ E^T \boldsymbol{\lambda}
+	\end{align*}
+	Substituting these results into the definition of $\mathbf{w}$ in \eqref{eq:definition_of_w}, we have
+	\begin{equation*}
+	\mathbf{w} = -K^+ M K^+ E^T \boldsymbol{\lambda} + \alpha \mathbf{c} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c},
+	\end{equation*}
+	as required.
+\end{proof}
+
+We now design an incremental algorithm for constructing interpolants $\mathbf{u}$, in which we can add points $x_i$ more efficiently. Let
+\begin{align*}
+\Psi :=& K^+ E^T \\
+\Theta :=& K^+ M \Psi
+\end{align*}
+so that 
+\begin{equation*}
+S = \Psi^T M \Psi
+\end{equation*}
+and
+\begin{equation*}
+\mathbf{u} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c}
+\end{equation*}
+The columns of $\Psi$ are the responses of $A^+$ to point sources (delta distributions) centered at the sample points $x_i$.
+The columns of $\Theta$ are the responses of $A^+M^{-1} A$ to these point sources.
+The matrices $\Psi$ and $\Theta$ may be constructed incrementally as new points are added.
+
 \section*{Acknowledgments}

### 2020-12-20 paper/localpsf.tex
issues with infinite dimensional weak form regularity details

@@
 	\label{eq:inline_wi_lemma}
-	\left(\psi_j, \Delta h\right)_{L^2(\Omega)} = h(x_j) - \avg(h)
+	h(x_j) = \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \avg(h)
 	\end{equation}
@@
 	\begin{equation*}
-		\avg(h) :=	 \frac{1}{|\Omega|} \int_\Omega h ~dx.
-	\end{equation*}
+		\avg(h) :=	 \frac{1}{|\Omega|} \int_\Omega h ~dx,
+	\end{equation*}
+	and $|\Omega| = \int_\Omega 1 ~dx$ is the Lebesgue measure of $\Omega$.
 \end{lem}
@@
 \begin{proof}
-	This is a standard result that follows from integration by parts and the definition of $\psi$ in \eqref{eq:psi_eq}.
+	The standard weak form for $\psi_j$ in \eqref{eq:psi_eq} is given by the following saddle point system for $\psi_j \in H^1(\Omega)$ and $\rho \in \mathbb{R}$:
+	\begin{alignat}{5}
+		-\int_\Omega \nabla \psi_j \cdot \nabla h ~dx &~+~& \rho \int_\Omega h ~dx &~=~ h(x_j) &\quad \forall h \in H^1(\Omega) \label{eq:lem_delta_eval_feq} \\
+		\int_\Omega \psi_j ~dx & & &~=~ 0. & \nonumber
+	\end{alignat}
+	This follows from integration by parts, using the Neumann zero boundary conditions for $\psi_j$, and enforcing the average value zero constraint for $\psi_j$ with the Lagrange multiplier $\rho$.
+If $h$ is the constant function $h(x)=1$, the first equation becomes
+	\begin{equation*}
+		\rho \int_\Omega 1 ~ dx = 1,
+	\end{equation*}
+	which implies $\rho = \frac{1}{|\Omega|}$.
+Now suppose that $h\in H^2_N(\Omega)$.
+Of course, we still have $h\in H^1(\Omega)$ as well, so equation \eqref{eq:lem_delta_eval_feq} still holds.
+But now $h$ has additional regularity, which allows us to use integration by parts on the first term in \eqref{eq:lem_delta_eval_feq} to bring all of the derivatives onto $h$.
+Furthermore, $h$ has Neumann zero boundary conditions, so the boundary term that arises during integration by parts is zero.
+Performing this integration by parts, and substituting in the value for $\rho$, yields
+	\begin{equation*}
+		\int_\Omega \psi_j \Delta h ~dx + \frac{1}{|\Omega|} \int_\Omega h ~dx = h(x_j),
+	\end{equation*}
+	which is the desired result.
 \end{proof}
@@
 Here subscripts for $\lambda$, and $e$ instead denote components of those vectors.
-For example, here $\lambda_j$ denotes the $j^\text{th}$ component of the vector $\lambda$, which corresponds to the $j^\text{th}$ of the vector $\lambda_i$ in Algorithm \ref{alg:weighting_functions_incremental}.
+For example, here $\lambda_j$ denotes the $j^\text{th}$ component of the vector $\lambda$, which corresponds to the $j^\text{th}$ component of the vector $\lambda_i$ in Algorithm \ref{alg:weighting_functions_incremental}.
 
@@
 	\end{equation*}
-	We will prove the theorem by showing that $\nabla \mathcal{L}(w,\lambda)=0$.
+	The plan of this proof is to show that $\nabla \mathcal{L}(w,\lambda)=0$, which will verify that $w$ solves optimization problem \ref{eq:wi_optimization_problem}.
 

### 2020-12-20 paper/localpsf.tex
fixed infinite dimensional regularity issues

--- 2020-12-19 paper/localpsf.tex
@@
 			\label{eq:psi_eq}
-			\Delta \psi_i = \delta_{x_i}, \quad \text{in }\Omega,
+			\Delta \psi_i = \delta_{x_i} - C/|\Omega|, \quad \text{in }\Omega,
 			\end{equation}
 			with Neumann zero boundary conditions, $\nu \cdot \nabla \psi_i = 0$ on $\partial \Omega$, and the average value zero condition, $\int_\Omega \psi_i(x) dx = 0$.
+Recall $\delta_{x_i}$ is a point source (delta distribution) centered at $x_i$, $C$ is the constant function $C(x)=1$, and $|\Omega|$ is the Lebesgue measure of $\Omega$.
 
@@
 \begin{proof}
-	This is a standard result that follows from integration by parts and the definition of $\psi$ in \eqref{eq:psi_eq}.
+	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\psi_j$ onto $h$, and using the Neumann zero boundary conditions for $\psi_j$ and $h$ to eliminate boundary integral terms.
+%	
+%	The standard weak form for $\psi_j$ in \eqref{eq:psi_eq} is given by the following saddle point system for $\psi_j \in H^1(\Omega)$ and $\rho \in \mathbb{R}$:
+%	\begin{alignat}{5}
+%		-\int_\Omega \nabla \psi_j \cdot \nabla h ~dx &~+~& \rho \int_\Omega h ~dx &~=~ h(x_j) &\quad \forall h \in H^1(\Omega) \label{eq:lem_delta_eval_feq} \\
+%		\int_\Omega \psi_j ~dx & & &~=~ 0. & \nonumber
+%	\end{alignat}
+%	This follows from integration by parts, using the Neumann zero boundary conditions for $\psi_j$, and enforcing the average value zero constraint for $\psi_j$ with the Lagrange multiplier $\rho$.
+If $h$ is the constant function $h(x)=1$, the first equation becomes
+%	\begin{equation*}
+%		\rho \int_\Omega 1 ~ dx = 1,
+%	\end{equation*}
+%	which implies $\rho = \frac{1}{|\Omega|}$.
+Now suppose that $h\in H^2_N(\Omega)$.
+Of course, we still have $h\in H^1(\Omega)$ as well, so equation \eqref{eq:lem_delta_eval_feq} still holds.
+But now $h$ has additional regularity, which allows us to use integration by parts on the first term in \eqref{eq:lem_delta_eval_feq} to bring all of the derivatives onto $h$.
+Furthermore, $h$ has Neumann zero boundary conditions, so the boundary term that arises during integration by parts is zero.
+Performing this integration by parts, and substituting in the value for $\rho$, yields
+%	\begin{equation*}
+%		\int_\Omega \psi_j \Delta h ~dx + \frac{1}{|\Omega|} \int_\Omega h ~dx = h(x_j),
+%	\end{equation*}
+%	which is the desired result.
 \end{proof}

### 2020-12-20 paper/localpsf.tex
adding back algorithm for point selection

@@
 Since it is too expensive compute the support of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ directly, we instead estimate that the support of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ is contained in the ellipsoid
-\begin{equation*}
+\begin{equation}
+\label{eq:support_ellipsoid}
 	E_x := \{x' \in \Omega: (x' - \mu(x))^T \Sigma(x)^{-1} (x' - \mu(x)) \le \tau^2\},
-\end{equation*}
+\end{equation}
 where $\tau$ is a small number of standard deviations; we use $\tau=3$.
@@
 
-%\begin{algorithm2e}
-%	\SetAlgoNoLine
-%	\SetKwInOut{Input}{Input}
-%	\SetKwProg{Fn}{Function}{}{}
-%
-%	Start with a finite set of candidate points $P \subset \Omega_I$
-%	
-%	$S_b \gets \{\}$
-%	
-%	\While{$P$ is not empty}{
-%		Pick a random point $p \in P$ and remove $p$ from $P$
-%		
-%		$p\text{\_is\_acceptable} \gets \text{True}$
-%		
-%		\For{$q \in S_b$}{
-%			
-%			\If{$E_p \cap E_q \neq \{\}$}{
-%				
-%				$p\text{\_is\_acceptable} \gets \text{False}$
-%				
-%				Terminate for loop
-%			}
-%		}
-%	
-%		\If{$p\text{\_is\_acceptable}$}{
-%			
-%			$S_b \gets S_b \cup \{p\}$
-%			
-%			$P \gets P \setminus \{q \in P : \mu(q) \in E_p\}$
-%		}
-%	}
-%
-%	\caption{Choosing one batch of sample points, $S_b$}
-%	\label{alg:point_choice}
-%\end{algorithm2e}
+\begin{algorithm2e}
+	\SetAlgoNoLine
+	\SetKwInOut{Input}{Input}
+	\SetKwProg{Fn}{Function}{}{}
+
+	Start with a finite set of candidate points $P \subset \Omega_I$
+
+	Initialize new empty batch of sample points, $S_b$
+
+	\While{$P$ is not empty}{
+		Determine the point $p \in P$ that is farthest from all points in previous sample point batches
+
+		Remove $p$ from $P$
+
+		\If{$\text{point\_is\_acceptable}(p, S_b)$}{
+
+			Add $p$ to $S_b$
+
+			Remove all points $q$ with $\mu(q) \in E_p$ from $P$
+		}
+	}
+
+    \SetKwFunction{FMain}{point\_is\_acceptable}
+\SetKwProg{Fn}{Function}{:}{}
+\Fn{\FMain{$p$, $S_b$}}{
+
+	$p\text{\_is\_acceptable} \gets \text{True}$
+
+	$E_p$ is the ellipsoid associated with $p$ (see \eqref{eq:support_ellipsoid})
+
+	\For{$q \in S_b$}{
+
+		$E_q$ is the ellipsoid associated with $q$
+
+		\If{$E_p \cap E_q \neq \{\}$}{
+
+			$p\text{\_is\_acceptable} \gets \text{False}$
+
+			Terminate for loop
+		}
+	}
+	\textbf{return} $p\text{\_is\_acceptable}$ 
+}
+
+	\caption{Choosing one batch of sample points, $S_b$}
+	\label{alg:point_choice}
+\end{algorithm2e}
 

### 2020-12-21 paper/localpsf.tex
improved point picking algorithm box

@@
 	\While{$P$ is not empty}{
-		Determine the point $p \in P$ that is farthest from all points in previous sample point batches
+		Determine the point $p \in P$ that is farthest from all points in the previous sample point batches $S_1,\dots,S_{b-1}$
 
@@
 
-		\If{$\text{point\_is\_acceptable}(p, S_b)$}{
+%		\If{$\text{point\_is\_acceptable}(p, S_b)$}{
+		\If{$E_p \cap E_q \neq \{\}$ for all $q \in S_b$}{
 
@@
 
-			Remove all points $q$ with $\mu(q) \in E_p$ from $P$
+			Remove all points $x$ satisfying $\mu(x) \in E_p$ from $P$
 		}
@@
     \SetKwFunction{FMain}{point\_is\_acceptable}
-\SetKwProg{Fn}{Function}{:}{}
-\Fn{\FMain{$p$, $S_b$}}{
-
-	$p\text{\_is\_acceptable} \gets \text{True}$
-
-	$E_p$ is the ellipsoid associated with $p$ (see \eqref{eq:support_ellipsoid})
-
-	\For{$q \in S_b$}{
-
-		$E_q$ is the ellipsoid associated with $q$
-
-		\If{$E_p \cap E_q \neq \{\}$}{
-
-			$p\text{\_is\_acceptable} \gets \text{False}$
-
-			Terminate for loop
-		}
-	}
-	\textbf{return} $p\text{\_is\_acceptable}$ 
-}
+%\SetKwProg{Fn}{Function}{:}{}
+%\Fn{\FMain{$p$, $S_b$}}{
+%	
+%	$p\text{\_is\_acceptable} \gets \text{True}$
+%	
+%	$E_p$ is the ellipsoid associated with $p$ (see \eqref{eq:support_ellipsoid})
+%	
+%	\For{$q \in S_b$}{
+%		
+%		$E_q$ is the ellipsoid associated with $q$
+%		
+%		\If{$E_p \cap E_q \neq \{\}$}{
+%			
+%			$p\text{\_is\_acceptable} \gets \text{False}$
+%			
+%			Terminate for loop
+%		}
+%	}
+%	\textbf{return} $p\text{\_is\_acceptable}$ 
+%}
 

### 2020-12-21 paper/localpsf.tex
improved point picking algorithm box

@@
 
-We select a batch of sample points, $S_b$, randomly from a finite set of candidate points $P \subset \Omega_I$.
-Starting from an empty set, we build $S_b$ by choosing candidate points $p$ randomly from $P$, one at a time.
-If $p$ is sufficiently far from all of the previously chosen points, in the sense that $E_p \cap E_q = \{\}$ for all $q \in S_b$, we add $p$ to $S_b$.
+We select a batch of sample points, $S_b$, randomly from a finite set of candidate points $P$.
+Starting from an empty set, we build $S_b$ by choosing candidate points $p$ from $P$, one at a time.
+The current candidate point, $p$, is the point in $P$ that is the farthest away from all points in previous sample point baches.
+That is, the minimizer of the following optimization problem:
+\begin{equation*}
+\min_{p \in P} \max_{x \in S_1 \cup \dots \cup S_{b-1}} \|p - x\|
+\end{equation*}
+After we choose the candidate $p$, we remove $p$ from $P$.
+If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in S_b$, and if $p$ is sufficiently far from the boundary, in the sense that $E_p \cap \partial \Omega = \{\}$, then we add $p$ to $S_b$.
 Otherwise we discard $p$.
@@
 We repeat the process to construct several batches of points $S_1, S_2, \dots$, until the total number of sample points exceeds a desired threshold.
-We write $S = S_1 \cup \dots$ to denote the set of sample points from all batches.
-In our implementation, the set of candidate points $P$ is the set of all Lagrange nodes for the finite element basis functions $\phi_i$ used to discretize the problem, minus all of the Lagrange nodes that are in $\Omega_B$, minus all points in previously chosen batches.
+In our implementation, the set of candidate points $P$ is the set of all Lagrange nodes for the finite element basis functions $\phi_i$ used to discretize the problem, minus all points in previously chosen batches.
 
@@
 %		\If{$\text{point\_is\_acceptable}(p, S_b)$}{
-		\If{$E_p \cap E_q \neq \{\}$ for all $q \in S_b$}{
+		\If{$E_p \cap E_q \neq \{\}$ for all $q \in S_b$ and $E_p \cap \partial \Omega = \{\}$}{
 

### 2020-12-22 paper/localpsf.tex
organizing paper more

@@
 
-\subsection{Connection to existing work}
+\subsection{Existing work}
 
@@
 
-\subsection{Background: distributions}
-
-Let $\overline{\Omega}$ be the closure of $\Omega$, and let $C(\overline{\Omega})$ be the space of continuous functions mapping $\overline{\Omega}\rightarrow \mathbb{R}$.
-The action of $\mathcal{A}$ is extended to distributions $\mu:C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ via the formula
-\begin{equation*}
-	\langle\mathcal{A},\mu\rangle(w) := \int_\Omega w(y) \mu\left(A(y, \cdot)\right) dy, 
-\end{equation*}
-where $A(y,\cdot)$ is the function $x \mapsto A(y,x)$. This is derived formally as follows:
-\begin{align*}
-	\langle\mathcal{A},\mu\rangle(w) &= \int_\Omega \int_\Omega w(y) A(y,x) \textrm{``} \mu(x) \textrm{''} dx dy \\
-	&= \int_\Omega w(y) \int_\Omega A(y,x) \textrm{``} \mu(x) \textrm{''} dx \\
-	&= \int_\Omega w(y) \mu\left(A(y,\cdot)\right) dx.
-\end{align*}
-For example, the delta distribution $\delta_x$ is defined by $\delta_x(v) = v(x)$, the action of $\mathcal{A}$ on $\delta_x$ is given by
-\begin{equation*}
-	\langle\mathcal{A},\delta_x\rangle(w)	= \int_\Omega w(y) A(y,x) dy,
-\end{equation*}
-and so the Riesz representation of $\langle\mathcal{A},\delta_x\rangle$ is given by $\langle\mathcal{A},\delta_x\rangle^* = A(~\cdot~,x)$.
-That is, $\langle\mathcal{A},\delta_x\rangle^*$ is the function $y \mapsto A(y,x)$.
-
-\subsection{Background: hierarchical matrices}
-
-\section{Choosing sample points}
-\label{sec:choosing_sample_points}
-
-We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points.
-In this section we describe how we choose the points $x_i$.
-The challenge is to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$.
-
-Ensuring that the points $x_i$ are well separated from each other requires estimating the supports of the functions $\langle\mathcal{A},\delta_{x}\rangle^*$.
+\section{Constructing the approximation}
+\label{sec:method}
+This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\widetilde{\mathcal{A}} \approx \mathcal{A}$, given by
+\begin{equation}
+\left(\widetilde{\mathcal{A}}u\right)(v) := \left( v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)}.
+\end{equation}
+
+We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}).
+We use a greedy algorithm to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}).
+
+Ensuring that the points $x_i$ are well separated from each other requires estimating the supports of the functions $\langle\mathcal{A},\delta_{x}\rangle^*$ (Section \ref{eq:mean_and_covariance_estimation}).
 Since $\varphi_x$, and thus $\langle\mathcal{A},\delta_{x}\rangle^*$, is nonnegative, $\langle\mathcal{A},\delta_{x}\rangle^*$ is a scaled probability distribution.
@@
 
-The function $\varphi_x(y)$ is undefined when $x + y \notin \Omega$, which can occur even if $x \in \Omega$ and $y \in \Omega$.
+%The function $\varphi_x(y)$ is undefined when $x + y \notin \Omega$, which can occur even if $x \in \Omega$ and $y \in \Omega$.
 Extending $\varphi_x$ by zero makes it defined for all $y \in \mathbb{R}^d$, but this extension is only reasonable if $x$ is far enough from $\partial \Omega$ that $\varphi_x(y)$ decays to zero, as a function of $y$, before $y$ reaches $\partial \Omega$.
@@
 
-\subsection{Greedy point selection algorithm}
+\subsection{Greedy sample point selection}
 \label{sec:greedy_point_selection}
@@
 The current candidate point, $p$, is the point in $P$ that is the farthest away from all points in previous sample point baches.
-That is, the minimizer of the following optimization problem:
-\begin{equation*}
-\min_{p \in P} \max_{x \in S_1 \cup \dots \cup S_{b-1}} \|p - x\|
+That is, the maximizer of the following optimization problem:
+\begin{equation*}
+\max_{p \in P} \min_{x \in S_1 \cup \dots \cup S_{b-1}} \|p - x\|
 \end{equation*}
@@
 Otherwise we discard $p$.
-This process repeats until it is no longer possible to pick a point from $P$ that is sufficiently far from the points in $S_b$.
+This process repeats until there are no more points in $P$.
 %This is detailed in Algorithm \ref{alg:point_choice}.
@@
 We repeat the process to construct several batches of points $S_1, S_2, \dots$, until the total number of sample points exceeds a desired threshold.
-In our implementation, the set of candidate points $P$ is the set of all Lagrange nodes for the finite element basis functions $\phi_i$ used to discretize the problem, minus all points in previously chosen batches.
+In our implementation, the set of candidate points $P$ is the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, minus all points in previously chosen batches.
 
@@
 
-	Start with a finite set of candidate points $P \subset \Omega_I$
+	Start with a finite set of candidate points $P \subset \Omega$
 
@@
 
-\subsection{Extracting impulse responses at sample points}
+\subsection{Batched computation of impulse responses}
 \label{sec:get_impulse_response}
@@
 
-\section{Discretized algorithms}
-
 \section{Fast ellipsoid intersection test}
@@
 
-\section{Additional proofs}
+\section{Additional theory}
 \label{app:proofs}
@@
 
+\subsection{Background: distributions}
+
+Let $\overline{\Omega}$ be the closure of $\Omega$, and let $C(\overline{\Omega})$ be the space of continuous functions mapping $\overline{\Omega}\rightarrow \mathbb{R}$.
+The action of $\mathcal{A}$ is extended to distributions $\mu:C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ via the formula
+\begin{equation*}
+\langle\mathcal{A},\mu\rangle(w) := \int_\Omega w(y) \mu\left(A(y, \cdot)\right) dy, 
+\end{equation*}
+where $A(y,\cdot)$ is the function $x \mapsto A(y,x)$. This is derived formally as follows:
+\begin{align*}
+\langle\mathcal{A},\mu\rangle(w) &= \int_\Omega \int_\Omega w(y) A(y,x) \textrm{``} \mu(x) \textrm{''} dx dy \\
+&= \int_\Omega w(y) \int_\Omega A(y,x) \textrm{``} \mu(x) \textrm{''} dx \\
+&= \int_\Omega w(y) \mu\left(A(y,\cdot)\right) dx.
+\end{align*}
+For example, the delta distribution $\delta_x$ is defined by $\delta_x(v) = v(x)$, the action of $\mathcal{A}$ on $\delta_x$ is given by
+\begin{equation*}
+\langle\mathcal{A},\delta_x\rangle(w)	= \int_\Omega w(y) A(y,x) dy,
+\end{equation*}
+and so the Riesz representation of $\langle\mathcal{A},\delta_x\rangle$ is given by $\langle\mathcal{A},\delta_x\rangle^* = A(~\cdot~,x)$.
+That is, $\langle\mathcal{A},\delta_x\rangle^*$ is the function $y \mapsto A(y,x)$.
+
 \begin{defn}

### 2020-12-22 paper/localpsf.tex
organizing paper more

@@
 We use a greedy algorithm to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}).
-
 Ensuring that the points $x_i$ are well separated from each other requires estimating the supports of the functions $\langle\mathcal{A},\delta_{x}\rangle^*$ (Section \ref{eq:mean_and_covariance_estimation}).
-Since $\varphi_x$, and thus $\langle\mathcal{A},\delta_{x}\rangle^*$, is nonnegative, $\langle\mathcal{A},\delta_{x}\rangle^*$ is a scaled probability distribution.
-We compute the scaling factors for all $x$ simultaneously by applying $\mathcal{A}$ to a constant function.
-We then compute the mean and covariance of the normalized version of $\langle\mathcal{A},\delta_{x}\rangle^*$, for all $x$ simultaneously, by applying $\mathcal{A}^T$ to a small number of linear and quadratic functions.
-The support of $\langle\mathcal{A},\delta_{x}\rangle^*$ is estimated to be contained in an ellipsoid of points that are within a small number of standard deviations of the mean for a Gaussian distribution with the same mean and covariance as the normalized version of $\langle\mathcal{A},\delta_{x}\rangle^*$.
 
@@
 
-Let 
-\begin{equation*}
-	\rho_x := \langle\mathcal{A}, \delta_x\rangle^* \big/ \alpha(x)
-\end{equation*}
-denote the normalized version of $\langle\mathcal{A}, \delta_x\rangle^*$, where $\alpha:\Omega \rightarrow \mathbb{R}$ is the volume function
-\begin{equation*}
-	\alpha(x) := \int_{\Omega} \langle\mathcal{A}, \delta_x\rangle^*(z) dz.
-\end{equation*}
-Note that $\rho_x$ is a probability distribution on $\Omega$.
-
-In Proposition \ref{thm:vol_mean_cov}, we see that $\alpha(x)$ may be computed for all $x$ simultaneously by applying $\mathcal{A}^T$ to a constant function.
+Since $\varphi_x$, and thus $\langle\mathcal{A},\delta_{x}\rangle^*$, is nonnegative, $\langle\mathcal{A},\delta_{x}\rangle^*$ is a scaled probability distribution.
+We compute the scaling factor, denoted $\alpha(x)$, for all $x$ simultaneously by applying $\mathcal{A}$ to a constant function.
+We then compute the mean, $\mu(x)$, and covariance, $\Sigma(x)$ of the normalized version of $\langle\mathcal{A},\delta_{x}\rangle^*$, for all $x$ simultaneously, by applying $\mathcal{A}^T$ to a small number of linear and quadratic functions.
+The support of $\langle\mathcal{A},\delta_{x}\rangle^*$ is estimated to be contained in an ellipsoid of points that are within a small number of standard deviations of the mean for a Gaussian distribution with the same mean and covariance as the normalized version of $\langle\mathcal{A},\delta_{x}\rangle^*$.
+
+\begin{defn}
+	We define the spatially varying volume, $\alpha$, the scaled impulse response, $\rho_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
+	\begin{align*}
+		\alpha:\Omega \rightarrow \mathbb{R}, &\qquad \alpha(x) := \int_{\Omega} \langle\mathcal{A}, \delta_x\rangle^*(z) dz, \\
+		\rho_x:\Omega \rightarrow \mathbb{R}, &\qquad \rho_x := \langle\mathcal{A}, \delta_x\rangle^* \big/ \alpha(x), \\
+		\mu:\Omega \rightarrow \mathbb{R}^d, &\qquad \mu(x) := \int_\Omega z \rho_x(z) dz, \\
+		\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\Sigma(x) := \int_\Omega (z - \mu(x))(z - \mu(x))^T \rho_x(z) dz.
+	\end{align*}
+\end{defn}
+
+%Let 
+%\begin{equation*}
+%	\rho_x := \langle\mathcal{A}, \delta_x\rangle^* \big/ \alpha(x)
+%\end{equation*}
+%denote the normalized version of $\langle\mathcal{A}, \delta_x\rangle^*$, where $\alpha:\Omega \rightarrow \mathbb{R}$ is the volume function
+%\begin{equation*}
+%	\alpha(x) := \int_{\Omega} \langle\mathcal{A}, \delta_x\rangle^*(z) dz.
+%\end{equation*}
+%Note that $\rho_x$ is a probability distribution on $\Omega$.
+%
+%In Proposition \ref{thm:vol_mean_cov}, we see that $\alpha(x)$ may be computed for all $x$ simultaneously by applying $\mathcal{A}^T$ to a constant function.
 We also see that the spatially varying mean, $\mu(x)$, and covariance, $\Sigma(x)$, of $\rho_x$, defined as
-\begin{align*}
-	\mu(x) :=& \int_\Omega z \rho_x(z) dz \\
-	\Sigma(x) :=& \int_\Omega (z - \mu(x))(z - \mu(x))^T \rho_x(z) dz
-\end{align*}
-may be computed for all $x$ simultaneously by applying $\mathcal{A}^T$ to $d$ linear functions, and $d(d+1)/2$ quadratic functions, respectively.
+%\begin{align*}
+%	\mu(x) :=& \int_\Omega z \rho_x(z) dz \\
+%	\Sigma(x) :=& \int_\Omega (z - \mu(x))(z - \mu(x))^T \rho_x(z) dz
+%\end{align*}
+%may be computed for all $x$ simultaneously by applying $\mathcal{A}^T$ to $d$ linear functions, and $d(d+1)/2$ quadratic functions, respectively.
 In Algorithm \ref{alg:varhpi_mean_cov} we describe an algorithm for computing the volume function, mean, and covariance based on the results of Proposition \ref{thm:vol_mean_cov}.

### 2020-12-22 paper/localpsf.tex
editing construction section

@@
 
-\subsection{Mean and covariance estimation}
+\subsection{Estimating impulse response supports}
 \label{eq:mean_and_covariance_estimation}
@@
 We compute the scaling factor, denoted $\alpha(x)$, for all $x$ simultaneously by applying $\mathcal{A}$ to a constant function.
-We then compute the mean, $\mu(x)$, and covariance, $\Sigma(x)$ of the normalized version of $\langle\mathcal{A},\delta_{x}\rangle^*$, for all $x$ simultaneously, by applying $\mathcal{A}^T$ to a small number of linear and quadratic functions.
-The support of $\langle\mathcal{A},\delta_{x}\rangle^*$ is estimated to be contained in an ellipsoid of points that are within a small number of standard deviations of the mean for a Gaussian distribution with the same mean and covariance as the normalized version of $\langle\mathcal{A},\delta_{x}\rangle^*$.
+We then compute the spatially varying mean, $\mu(x)$, and covariance, $\Sigma(x)$ of the normalized version of $\langle\mathcal{A},\delta_{x}\rangle^*$, for all $x$ simultaneously, by applying $\mathcal{A}^T$ to a small number of linear and quadratic functions.
+We make the approximation that the support of $\langle\mathcal{A},\delta_{x}\rangle^*$ is contained within the ellipsoid
+\begin{equation}
+\label{eq:support_ellipsoid}
+E_x := \{x' \in \Omega: (x' - \mu(x))^T \Sigma(x)^{-1} (x' - \mu(x)) \le \tau^2\},
+\end{equation}
+where $\tau$ is a small number of standard deviations; we use $\tau=3$.
+The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations of the mean for the Gaussian distribution that has the same mean and covariance as the normalized version of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$.
+The fraction of the mass of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though we find this bound is conservative and far less mass resides in this region in practice.
 
@@
 
+%The support of $\langle\mathcal{A},\delta_{x}\rangle^*$ is estimated to be contained within a small number of standard deviations of the mean for a Gaussian distribution with the same mean and covariance as the normalized version of $\langle\mathcal{A},\delta_{x}\rangle^*$.
+
 %Let 
@@
 
-\begin{prop}
+\begin{theorem}
 	\label{thm:vol_mean_cov}
@@
 	We write $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
-\end{prop}
+\end{theorem}
 
@@
 Ideally, sample points $x_i$ and $x_j$ should be far enough apart from each other that the supports of $\left\langle\mathcal{A}, \delta_{x_i}\right\rangle^*$ and $\left\langle\mathcal{A}, \delta_{x_j}\right\rangle^*$ do not overlap.
-Since it is too expensive compute the support of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ directly, we instead estimate that the support of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ is contained in the ellipsoid
-\begin{equation}
-\label{eq:support_ellipsoid}
-	E_x := \{x' \in \Omega: (x' - \mu(x))^T \Sigma(x)^{-1} (x' - \mu(x)) \le \tau^2\},
-\end{equation}
-where $\tau$ is a small number of standard deviations; we use $\tau=3$.
+
+%Since it is too expensive compute the support of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ directly, we instead estimate that the support of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ is contained in the ellipsoid
+%\begin{equation}
+%\label{eq:support_ellipsoid}
+%	E_x := \{x' \in \Omega: (x' - \mu(x))^T \Sigma(x)^{-1} (x' - \mu(x)) \le \tau^2\},
+%\end{equation}
+%where $\tau$ is a small number of standard deviations; we use $\tau=3$.
 The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations of the mean for the Gaussian distribution that has the same mean and covariance as $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$.
 The fraction of the mass of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though we find this bound is conservative and far less mass resides in this region in practice.
+
 %We call $E_x$ the \emph{unit ellipsoid} with mean $\mu(x)$ and covariance $\Sigma(x)/\tau^2$.
@@
 \begin{equation*}
-\max_{p \in P} \min_{x \in S_1 \cup \dots \cup S_{b-1}} \|p - x\|
+\max_{p \in P} \min_{x \in S_1 \cup \dots \cup S_{b-1}} \|p - x\|.
 \end{equation*}

### 2020-12-22 paper/localpsf.tex
editing construction section

@@
 	}
-	\caption{Compute volume $\alpha_h$, mean $\mu_h$, and covariance $\Sigma_h$}
+	\caption{Compute volume $\alpha$, mean $\mu$, and covariance $\Sigma$}
 	\label{alg:varhpi_mean_cov}
@@
 
-Ideally, sample points $x_i$ and $x_j$ should be far enough apart from each other that the supports of $\left\langle\mathcal{A}, \delta_{x_i}\right\rangle^*$ and $\left\langle\mathcal{A}, \delta_{x_j}\right\rangle^*$ do not overlap.
+% sample points $x_i$ and $x_j$ should be far enough apart from each other that the supports of $\left\langle\mathcal{A}, \delta_{x_i}\right\rangle^*$ and $\left\langle\mathcal{A}, \delta_{x_j}\right\rangle^*$ do not overlap.
 
@@
 
-We select a batch of sample points, $S_b$, randomly from a finite set of candidate points $P$.
-Starting from an empty set, we build $S_b$ by choosing candidate points $p$ from $P$, one at a time.
-The current candidate point, $p$, is the point in $P$ that is the farthest away from all points in previous sample point baches.
+Starting from a finite set of candidate points $P$, we select batches of sample points, $S_b$, using a greedy algorithm.
+To build a batch $S_b$, first we initialize $S_b$ as an empty set.
+Then we select a candidate point $p \in P$ that is the farthest away from all points in previous sample point batches.
 That is, the maximizer of the following optimization problem:
@@
 \end{equation*}
-After we choose the candidate $p$, we remove $p$ from $P$.
+Once $p$ is selected, we remove $p$ from $P$.
 If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in S_b$, and if $p$ is sufficiently far from the boundary, in the sense that $E_p \cap \partial \Omega = \{\}$, then we add $p$ to $S_b$.
@@
 This process repeats until there are no more points in $P$.
-%This is detailed in Algorithm \ref{alg:point_choice}.
-
-We repeat the process to construct several batches of points $S_1, S_2, \dots$, until the total number of sample points exceeds a desired threshold.
-In our implementation, the set of candidate points $P$ is the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, minus all points in previously chosen batches.
+This is detailed in Algorithm \ref{alg:point_choice}.
+
+%We select a batches of sample points, $S_b$, from a finite set of candidate points $P$ using a greedy algorithm.
+Starting from an empty set, we build $S_b$ from candidate points $p$ from $P$, one point at a time.
+The current candidate point, $p$, is the point in $P$
+
+We repeat the process to construct several batches of points $S_1, S_2, \dots$, until the number of batches exceeds a desired threshold.
+In our implementation, the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, minus all points in previously chosen batches.
+The set $P$ is re-initialized for each batch $S_b$.
 

### 2020-12-22 paper/localpsf.tex
editing construction section

@@
 Then we select a candidate point $p \in P$ that is the farthest away from all points in previous sample point batches.
-That is, the maximizer of the following optimization problem:
+That is, $p$ is the maximizer of the following optimization problem:
 \begin{equation*}
@@
 This is detailed in Algorithm \ref{alg:point_choice}.
+We determine whether $E_p \cap E_q = \{\}$ using the fast ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.
 
@@
 We repeat the process to construct several batches of points $S_1, S_2, \dots$, until the number of batches exceeds a desired threshold.
-In our implementation, the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, minus all points in previously chosen batches.
-The set $P$ is re-initialized for each batch $S_b$.
+In our implementation, for each batch the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, minus all points in previously chosen batches.
 
@@
 \end{equation}
-for all $x_i \in S_b$. By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover $\varphi_{x_i}$ for every point $x_i \in S_b$. 
+for all $x_i \in S_b$.
+By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover $\varphi_{x_i}$ for every point $x_i \in S_b$.
+To minimize memory usage, the functions $\varphi_i$ are never isolated and stored individually.
+Instead, we store the functions $\eta^b$, and use formula \eqref{eq:varphi_eval} to perform function evaluations $z \mapsto \varphi_i(z)$ as needed.
 %For $x_i \in X$ and $y \in \mathbb{R}^d$, we may evaluate $\varphi_{x_i}(y)$ by finding the batch $X_b$ that $x_i$ is contained in, then using \eqref{eq:varphi_eval}.
@@
 \section{Fast ellipsoid intersection test}
+\label{sec:fast_ellipsoid_intersection_test}
 The procedure for choosing sample points relies on quickly determining whether two ellipsoids intersect. Let $E_p$ and $E_q$ be the ellipsoids defined as

### 2020-12-22 paper/localpsf.tex
move poisson weighting function proofs to appendix

@@
 
-\section{Poisson interpolation weighting functions}
+\subsection{Poisson interpolation weighting functions}
 \label{sec:weighting_functions}
@@
 Here $\Delta$ is the Laplacian on $\Omega$, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$, and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions.
-In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense.
-In Proposition \ref{thm:wi_sum_to_one} we note that the weighting functions sum to one pointwise.
-These results rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}
+Using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense (Theorem \ref{thm:smoothest_interpolant}).
+Also, these weighting functions sum to one pointwise (Proposition \ref{thm:wi_sum_to_one}).
 
 In Algorithm \ref{alg:weighting_functions_incremental}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$.
-We prove correctness of Algorithm \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}.
-Theorem \ref{thm:wi_alg_is_correct} relies on a prerequisite result which we present in Lemma \ref{lem:psi_delta_eval}.
 Constructing all $r$ of the functions $w_i$ requires solving the Poisson equation $2r$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid.
-All the existing weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added.
-The Poisson PDE solves can be performed incrementally, so that we only need to perform two new Poisson PDE solves when a new point is added, rather than $2(r+1)$ new PDE solves.
+All of the weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added.
+However, the dominant computational cost in constructing the weighting functions is the solution of the Poisson PDEs, and these Poisson PDE solves can be performed incrementally---we only need to perform two new Poisson PDE solves when a new point is added, rather than $2(r+1)$ new PDE solves.
 
@@
 When inverting for parameter fields defined on four or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used.
-
-\begin{lem}[Preparation for Theorem \ref{thm:smoothest_interpolant}]
-	\label{lem:T_is_linear}
-	Let $f \in \mathbb{R}^r$. There exists a unique solution to the following optimization problem:
-	\begin{equation}
-	\label{eq:generic_optimization_initial}
-	\begin{aligned}
-	\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
-	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r.
-	\end{aligned}
-	\end{equation}
-	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
-	\begin{equation*}
-	\mathcal{T}(f) := u^*,
-	\end{equation*}
-	where $u^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
-\end{lem}
+Proofs for results in this section are given in Appendix \ref{app:proofs}.
 
@@
 \end{thm}
-\begin{proof}
-	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.
-The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
-	\begin{align*}
-		\mathcal{T}(f) &= \varphi_1(y) \mathcal{T}(e_1) + \varphi_2(y) \mathcal{T}(e_2) + \dots +\varphi_r(y) \mathcal{T}(e_r) \\
-		&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
-		&= u.
-	\end{align*}
- 	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $u$ in \eqref{eq:u_defn_in_thm}.
-\end{proof}
-
-\begin{prop}
+
+\begin{prop}[Weighting functions sum to one]
 	\label{thm:wi_sum_to_one}
@@
 \end{prop}
-
-\begin{proof}
-	By linearity of $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), we have
-	\begin{equation*}
-		\sum_{i=1}^r w_i(x) = \sum_{i=1}^r \mathcal{T}(e_i)(x) = \mathcal{T}\left(\sum_{i=1}^r e_i\right)(x) = \mathcal{T}(\mathbf{1})(x) = 1,
-	\end{equation*}
-	where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one.
-The last equality ($\mathcal{T}(\mathbf{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \mathbf{1}$, and yields the minimum possible value of zero for the objective function.
-\end{proof}
 
@@
 
-\begin{lem}[Preparation for Theorem \ref{thm:wi_alg_is_correct}]
-	\label{lem:psi_delta_eval}
-	We have
-	\begin{equation}
-	\label{eq:inline_wi_lemma}
-	h(x_j) = \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \avg(h)
-	\end{equation}
-	for all $h\in H^2_N(\Omega)$, where
-	\begin{equation*}
-		\avg(h) :=	 \frac{1}{|\Omega|} \int_\Omega h ~dx,
-	\end{equation*}
-	and $|\Omega| = \int_\Omega 1 ~dx$ is the Lebesgue measure of $\Omega$.
-\end{lem}
-
-\begin{proof}
-	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\psi_j$ onto $h$, and using the Neumann zero boundary conditions for $\psi_j$ and $h$ to eliminate boundary integral terms.
 %	
@@
 %	which is the desired result.
-\end{proof}
 
@@
 \end{thm}
-
-\begin{proof}
-	For notational convenience, in this proof we write $w$, $\lambda$, and $e$ instead of $w_i$, $\lambda_i$, and $e_i$, respectively.
-Here subscripts for $\lambda$, and $e$ instead denote components of those vectors.
-For example, here $\lambda_j$ denotes the $j^\text{th}$ component of the vector $\lambda$, which corresponds to the $j^\text{th}$ component of the vector $\lambda_i$ in Algorithm \ref{alg:weighting_functions_incremental}.
-
-	From standard optimization theory, the solution to optimization problem \ref{eq:wi_optimization_problem} is the stationary point of the Lagrangian,
-	\begin{equation*}
-		\mathcal{L}(w,\lambda) = \frac{1}{2}\|\Delta w\|^2 + \sum_{j=1}^r \lambda_j\left(w(x_j) - e_j\right).
-	\end{equation*}
-	The plan of this proof is to show that $\nabla \mathcal{L}(w,\lambda)=0$, which will verify that $w$ solves optimization problem \ref{eq:wi_optimization_problem}.
-
-	Using the definition of $w_i$ in \eqref{eq:definition_of_wi_alg}, we may write the gradient of $\mathcal{L}$ with respect to $\lambda_j$ as
-	\begin{equation*}
-		\nabla_{\lambda_j} \mathcal{L}(w, \lambda) = w(x_j) - e_j
-		= \alpha - \sum_{k=1}^r \lambda_k \theta_k(x_j) - e_j.
-	\end{equation*}
-	Using Lemma \ref{lem:psi_delta_eval} with $h=\theta_k$, the equation for $\theta_k$ in \eqref{eq:theta_eq}, and the definition of $S$ in \eqref{eq:defn_of_S_alg}, we have
-	\begin{align*}
-		\sum_{k=1}^r \lambda_k \theta_k(x_j) &= \sum_{k=1}^r \lambda_k \left(\psi_j, \Delta \theta_k\right)_{L^2(\Omega)}\\
-		&= \sum_{k=1}^r \lambda_k \left(\psi_j, \psi_k\right)_{L^2(\Omega)} = \left(S \lambda\right)_j.
-	\end{align*}
-	Combining all of these components of $\nabla_\lambda \mathcal{L}$ into a single vector and using the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} yields
-	\begin{align*}
-		\nabla_\lambda \mathcal{L}(w, \lambda) &= \alpha \mathbf{1} - e - S \lambda \\
-		&= \alpha \mathbf{1} - e - S \left(-S^{-1}\left(e - \alpha \mathbf{1}\right)\right) = 0.
-	\end{align*}
-
-	Taking the directional derivative of $\mathcal{L}$ with respect to $w$ in direction $h$ and using Lemma \ref{lem:psi_delta_eval} yields
-	\begin{align}
-		\nabla_w \mathcal{L}(w, \lambda)h &= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j h(x_j) \nonumber \\
-		&= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \lambda_j \avg(h) \nonumber \\
-		&= \left(\Delta w + \sum_{j=1}^r \lambda_j \psi_j,~ \Delta h\right)_{L^2(\Omega)} + \avg(h) \mathbf{1}^T \lambda. \label{eq:L_w_38}
-	\end{align}
-	Using the definition of $w$ in Algorithm \ref{alg:weighting_functions_incremental}, and the fact that the Laplacian of a constant function is zero, we have
-	\begin{equation*}
-		\Delta w = -\sum_{j=1}^r \lambda_j \Delta \theta_j 
-		= -\sum_{j=1}^r \lambda_j \psi_j,
-	\end{equation*}
-	which implies that the first term in \eqref{eq:L_w_38} is zero. By the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} and $\alpha$ in \eqref{eq:definition_of_alpha_alg}, we have
-	\begin{align*}
-		\mathbf{1}^T \lambda &= -\mathbf{1}^T S^{-1} \left(e - \alpha \mathbf{1}\right) \\
-		&= -\mathbf{1}^T S^{-1} e + \alpha \mathbf{1}^T S^{-1} \mathbf{1} \\
-		&= -\mathbf{1}^T S^{-1} e + \frac{\mathbf{1}^T S^{-1} e}{\mathbf{1}^T S^{-1} \mathbf{1}} \mathbf{1}^T S^{-1} \mathbf{1} = 0,
-	\end{align*}
-	which implies that the second term in \eqref{eq:L_w_38} is zero.
-Thus $\nabla_w \mathcal{L}(w, \lambda) = 0$, so Algorithm \ref{alg:weighting_functions_incremental} produces the solution to optimization problem \ref{eq:wi_optimization_problem}.
-\end{proof}
 
@@
 
-\begin{proof}[Proof of Proposition \ref{prop:boundary_source}]
-	Let $w \in L^2(\Omega)$ be arbitrary. We have
+These results rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}.
+We prove correctness of Algorithm \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}.
+Theorem \ref{thm:wi_alg_is_correct} relies on a prerequisite result which we present in Lemma \ref{lem:psi_delta_eval}.
+
+\begin{lem}[Preparation for Theorem \ref{thm:smoothest_interpolant}]
+	\label{lem:T_is_linear}
+	Let $f \in \mathbb{R}^r$. There exists a unique solution to the following optimization problem:
+	\begin{equation}
+	\label{eq:generic_optimization_initial}
+	\begin{aligned}
+	\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
+	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r.
+	\end{aligned}
+	\end{equation}
+	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
+	\begin{equation*}
+	\mathcal{T}(f) := u^*,
+	\end{equation*}
+	where $u^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
+\end{lem}
+\begin{proof}
+	asdf
+\end{proof}
+
+\begin{proof}[Proof of Theorem \ref{thm:smoothest_interpolant}]
+	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.
+The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
 	\begin{align*}
-		\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle(w) =& \int_\Omega w(y) \int_{\partial \Omega} A(y,x) dy dx \\
-		=& \int_\Omega w(x) \int_{\partial \Omega} \varphi_x(y-x) dy dx = \left(w, z\right)_{L^2(\Omega)},
+	\mathcal{T}(f) &= \varphi_1(y) \mathcal{T}(e_1) + \varphi_2(y) \mathcal{T}(e_2) + \dots +\varphi_r(y) \mathcal{T}(e_r) \\
+	&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
+	&= u.
 	\end{align*}
-	where
-	\begin{equation*}
-		z(x) := \int_{\partial \Omega} \varphi_x(y-x) dy,	
-	\end{equation*}
-	which implies
-	\begin{equation*}
-		\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle^*(x) = z = \int_{\partial \Omega} \varphi_x(y-x) dy.
-	\end{equation*}
-	Since $\varphi_x$ is a non-negative function, the result follows from the fact that the integral of a non-negative function over a set of positive measure is zero if and only if the function is zero almost everywhere on that set.
+	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $u$ in \eqref{eq:u_defn_in_thm}.
 \end{proof}
+
+\begin{proof}[Proof of Proposition \ref{thm:wi_sum_to_one}]
+	By linearity of $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), we have
+	\begin{equation*}
+	\sum_{i=1}^r w_i(x) = \sum_{i=1}^r \mathcal{T}(e_i)(x) = \mathcal{T}\left(\sum_{i=1}^r e_i\right)(x) = \mathcal{T}(\mathbf{1})(x) = 1,
+	\end{equation*}
+	where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one.
+The last equality ($\mathcal{T}(\mathbf{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \mathbf{1}$, and yields the minimum possible value of zero for the objective function.
+\end{proof}
+
+\begin{lem}[Preparation for Theorem \ref{thm:wi_alg_is_correct}]
+	\label{lem:psi_delta_eval}
+	We have
+	\begin{equation}
+	\label{eq:inline_wi_lemma}
+	h(x_j) = \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \avg(h)
+	\end{equation}
+	for all $h\in H^2_N(\Omega)$, where
+	\begin{equation*}
+	\avg(h) :=	 \frac{1}{|\Omega|} \int_\Omega h ~dx,
+	\end{equation*}
+	and $|\Omega| = \int_\Omega 1 ~dx$ is the Lebesgue measure of $\Omega$.
+\end{lem}
+
+\begin{proof}
+	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\psi_j$ onto $h$, and using the Neumann zero boundary conditions for $\psi_j$ and $h$ to eliminate boundary integral terms.
+\end{proof}
+
+\begin{proof}[Proof of Theorem \ref{alg:weighting_functions_incremental}]
+	For notational convenience, in this proof we write $w$, $\lambda$, and $e$ instead of $w_i$, $\lambda_i$, and $e_i$, respectively.
+Here subscripts for $\lambda$, and $e$ instead denote components of those vectors.
+For example, here $\lambda_j$ denotes the $j^\text{th}$ component of the vector $\lambda$, which corresponds to the $j^\text{th}$ component of the vector $\lambda_i$ in Algorithm \ref{alg:weighting_functions_incremental}.
+
+	From standard optimization theory, the solution to optimization problem \ref{eq:wi_optimization_problem} is the stationary point of the Lagrangian,
+	\begin{equation*}
+	\mathcal{L}(w,\lambda) = \frac{1}{2}\|\Delta w\|^2 + \sum_{j=1}^r \lambda_j\left(w(x_j) - e_j\right).
+	\end{equation*}
+	The plan of this proof is to show that $\nabla \mathcal{L}(w,\lambda)=0$, which will verify that $w$ solves optimization problem \ref{eq:wi_optimization_problem}.
+
+	Using the definition of $w_i$ in \eqref{eq:definition_of_wi_alg}, we may write the gradient of $\mathcal{L}$ with respect to $\lambda_j$ as
+	\begin{equation*}
+	\nabla_{\lambda_j} \mathcal{L}(w, \lambda) = w(x_j) - e_j
+	= \alpha - \sum_{k=1}^r \lambda_k \theta_k(x_j) - e_j.
+	\end{equation*}
+	Using Lemma \ref{lem:psi_delta_eval} with $h=\theta_k$, the equation for $\theta_k$ in \eqref{eq:theta_eq}, and the definition of $S$ in \eqref{eq:defn_of_S_alg}, we have
+	\begin{align*}
+	\sum_{k=1}^r \lambda_k \theta_k(x_j) &= \sum_{k=1}^r \lambda_k \left(\psi_j, \Delta \theta_k\right)_{L^2(\Omega)}\\
+	&= \sum_{k=1}^r \lambda_k \left(\psi_j, \psi_k\right)_{L^2(\Omega)} = \left(S \lambda\right)_j.
+	\end{align*}
+	Combining all of these components of $\nabla_\lambda \mathcal{L}$ into a single vector and using the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} yields
+	\begin{align*}
+	\nabla_\lambda \mathcal{L}(w, \lambda) &= \alpha \mathbf{1} - e - S \lambda \\
+	&= \alpha \mathbf{1} - e - S \left(-S^{-1}\left(e - \alpha \mathbf{1}\right)\right) = 0.
+	\end{align*}
+
+	Taking the directional derivative of $\mathcal{L}$ with respect to $w$ in direction $h$ and using Lemma \ref{lem:psi_delta_eval} yields
+	\begin{align}
+	\nabla_w \mathcal{L}(w, \lambda)h &= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j h(x_j) \nonumber \\
+	&= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \lambda_j \avg(h) \nonumber \\
+	&= \left(\Delta w + \sum_{j=1}^r \lambda_j \psi_j,~ \Delta h\right)_{L^2(\Omega)} + \avg(h) \mathbf{1}^T \lambda. \label{eq:L_w_38}
+	\end{align}
+	Using the definition of $w$ in Algorithm \ref{alg:weighting_functions_incremental}, and the fact that the Laplacian of a constant function is zero, we have
+	\begin{equation*}
+	\Delta w = -\sum_{j=1}^r \lambda_j \Delta \theta_j 
+	= -\sum_{j=1}^r \lambda_j \psi_j,
+	\end{equation*}
+	which implies that the first term in \eqref{eq:L_w_38} is zero. By the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} and $\alpha$ in \eqref{eq:definition_of_alpha_alg}, we have
+	\begin{align*}
+	\mathbf{1}^T \lambda &= -\mathbf{1}^T S^{-1} \left(e - \alpha \mathbf{1}\right) \\
+	&= -\mathbf{1}^T S^{-1} e + \alpha \mathbf{1}^T S^{-1} \mathbf{1} \\
+	&= -\mathbf{1}^T S^{-1} e + \frac{\mathbf{1}^T S^{-1} e}{\mathbf{1}^T S^{-1} \mathbf{1}} \mathbf{1}^T S^{-1} \mathbf{1} = 0,
+	\end{align*}
+	which implies that the second term in \eqref{eq:L_w_38} is zero.
+Thus $\nabla_w \mathcal{L}(w, \lambda) = 0$, so Algorithm \ref{alg:weighting_functions_incremental} produces the solution to optimization problem \ref{eq:wi_optimization_problem}.
+\end{proof}
+
+%\begin{proof}[Proof of Proposition \ref{prop:boundary_source}]
+%	Let $w \in L^2(\Omega)$ be arbitrary. We have
+%	\begin{align*}
+%		\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle(w) =& \int_\Omega w(y) \int_{\partial \Omega} A(y,x) dy dx \\
+%		=& \int_\Omega w(x) \int_{\partial \Omega} \varphi_x(y-x) dy dx = \left(w, z\right)_{L^2(\Omega)},
+%	\end{align*}
+%	where
+%	\begin{equation*}
+%		z(x) := \int_{\partial \Omega} \varphi_x(y-x) dy,	
+%	\end{equation*}
+%	which implies
+%	\begin{equation*}
+%		\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle^*(x) = z = \int_{\partial \Omega} \varphi_x(y-x) dy.
+%	\end{equation*}
+%	Since $\varphi_x$ is a non-negative function, the result follows from the fact that the integral of a non-negative function over a set of positive measure is zero if and only if the function is zero almost everywhere on that set.
+%\end{proof}
 

### 2020-12-23 paper/localpsf.tex
wrote about hessian product convolutions in existing work. made more edits to Section 2

@@
 
-A product-convolution type approximation of the Hessian in a seismic imaging problem was used in [CITE GEORG].
+Product-convolution type approximations of Hessians have been used in a seismic inverse problem in [CITE GEORG], and in an advection-diffusion inverse problem in [CITE MY PAPER].
+In [DEMANET SEISMIC], matrix probing [DEMANET MATRIX] is used to approximate the Hessian in a seismic inverse problem as the sum of simple pseudodifferential operators.
+Although it is not explicitly mentioned, the approximation in [DEMANET SEISMIC] could be interpreted as a convolution-product interpolation, which is like a product-convolution approximation, except the order of pointwise multiplications and convolutions is reversed in \eqref{eq:product_convolution}.
 
@@
 
-Demanet matrix probing can be interpreted as convolution-product.
+%Demanet matrix probing can be interpreted as convolution-product.
 
@@
 
-Geometry-aware weighting functions vs radial basis functions. Kriging.
+%Geometry-aware weighting functions vs radial basis functions. Kriging.
 
@@
 \label{sec:method}
-This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\widetilde{\mathcal{A}} \approx \mathcal{A}$, given by
-\begin{equation}
-\left(\widetilde{\mathcal{A}}u\right)(v) := \left( v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)}.
-\end{equation}
-
-We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}).
-We use a greedy algorithm to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}).
-Ensuring that the points $x_i$ are well separated from each other requires estimating the supports of the functions $\langle\mathcal{A},\delta_{x}\rangle^*$ (Section \ref{eq:mean_and_covariance_estimation}).
+This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\widetilde{\mathcal{A}} \approx \mathcal{A}$, given in \eqref{eq:product_convolution}.
+\begin{itemize}
+\item We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}). 
+\item We use a greedy algorithm to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}).
+\item Ensuring that the points $x_i$ are well separated from each other requires a-priori estimates of the supports of the functions $\langle\mathcal{A},\delta_{x}\rangle^*$ (Section \ref{eq:mean_and_covariance_estimation}).
+\item The weighting functions are defined as the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
+\end{itemize}
+The complete algorithm for constructing $\widetilde{\mathcal{A}}$ is shown in Algorithm \ref{alg:construct_Atilde}.
+
+\begin{algorithm2e}
+	\SetAlgoNoLine
+	\SetKwInOut{Input}{Input}
+	\SetKwInOut{Output}{Output}
+	%	\Input{Function $v \mapsto \mathcal{H}^Tv$}
+
+	%	\mbox{}\\
+
+	%	\Output{Spatially varying mean $\mu:\Omega \rightarrow \mathbb{R}^d$, \\ 
+	%		Spatially varying covariance $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$}
+	%	\Begin{
+	{	
+		Compute spatially varying mean, $\mu$, and covariance, $\Sigma$, using Algorithm \ref{alg:varhpi_mean_cov}.
+
+		\For{$b=1,2,\dots,\text{max\_batches}$}{
+			Choose a batch of sample points, $S_b$, using Algorithm \ref{alg:point_choice}
+
+			Compute $\eta^b$ by applying $\mathcal{A}$ to the dirac comb associated with $S_b$ (Section \ref{sec:get_impulse_response})
+
+			Compute weighting functions $w_i$ using Algorithm \ref{alg:weighting_functions_incremental}.
+In lines 2 and 3 of Algorithm \ref{alg:weighting_functions_incremental}, only solve for $\psi_i$ and $\theta_i$ associated with new points $x_i$ in $S_b$.
+The functions $\psi_i$ and $\theta_i$ associated with points in previous batches have already been computed in previous iterations of this for loop.
+		}
+
+		%		\Return{$\mu$, $\Sigma$}
+	}
+	\caption{Construct product-convolution approximation $\widetilde{\mathcal{A}}$}
+	\label{alg:construct_Atilde}
+\end{algorithm2e}
 

### 2020-12-31 paper/localpsf.tex
h matrix stuff

@@
 
-Our operator approximation method is defined by how we compute the convolution kernels, $\varphi_i$, (Section \ref{sec:get_impulse_response}) how we choose the points, $x_i$, (Section \ref{sec:choosing_sample_points}) what weighting functions, $w_i$, that we use (Section \ref{sec:weighting_functions}), and how we use the resulting product-convolution approximation to perform linear algebra operations such as solving linear systems (Section \ref{sec:H_matrix}).
-In this section we summarize the answers to these questions.
+Our operator approximation method is defined by how we compute the convolution kernels, $\varphi_i$, how we choose the points, $x_i$, what weighting functions, $w_i$, that we use, and how we use the resulting product-convolution approximation to perform linear algebra operations such as solving linear systems.
+In this section we summarize the answers to these questions; we will provide detailed answers to these questions in Section \ref{sec:method} and Section \ref{sec:using_approximation}.
 
@@
 
-\section{Interpolating impulse response samples old}
-
-We interpolate the functions $\varphi_{x_i}$ associated with the points $x_i \in S$, to generate functions $\varphi_x$ associated with new points $x \notin S$.
-This is done using a continuous generalization of the interpolative decomposition for matrices, wherein a select number of columns of a matrix are interpolated to fill the whole matrix.
-Interpolating $\varphi_x$ differs from interpolative low rank approximation of $\mathcal{A}$, which corresponds to interpolating $\langle\mathcal{A},\delta_x\rangle^*$.
-Because of the shifting that maps $\langle\mathcal{A}, \delta_x\rangle$ to $\varphi_x$, interpolation of $\varphi_{x_i}$ yields an approximation $\widetilde{\mathcal{A}}$ which can be high-rank.
-
-The interpolation procedure requires a basis of smooth functions.
-We form this basis from smooth eigenmodes of the Laplacian on $\Omega$ with Neumann boundary conditions.
-The resulting approximation $\mathcal{A}$ may be interpreted as a weighted sum of convolution operators, with smooth spatially varying weighting functions.
-Kernel entries of the approximation, $\widetilde{A}(y,x)$, may be evaluated efficiently.
-
-\subsection{Interpolative factorization of the SVIR}
-\label{sec:interp_fact_svir}
-
-The \emph{spatially varying impulse response}, denoted $T$, is the reinterpretation of $\varphi_x(y)$ as a function of both $y$ and $x$:
-\begin{equation*}
-	T(y,x) := \varphi_x(y) = A(y+x,x).
-\end{equation*}
-Since $T(y,x)$ is smooth as a function of $x$ (Condition 1), the continuous analog of the row space of $T$ should be well-approximated by the span of a small number of orthonormal smooth functions, $\{u_k\}_{k=1}^r$.
-Condition 1, therefore, is equivalent to the existence of a low rank approximation $\widetilde{T}$ of $T$ of the form
-\begin{equation}
-	\label{eq:formula_for_A}
-	T \approx \widetilde{T} = \Psi U
-\end{equation}
-where $\Psi:\mathbb{R}^d \times \{1,\dots,r\} \rightarrow \mathbb{R}$ is an unknown semi-discrete kernel with $r$ columns, and $U:\{1,\dots,r\} \times \Omega \rightarrow \mathbb{R}$ is the semi-discrete kernel with $r$ rows, such that the $i^\text{th}$ row of $U$ is the function $u_i$.
-That is,
-\begin{equation*}
-	U(i,x) = u_i(x).
-\end{equation*}
-The multiplication $K_1 K_2$ of kernels $K_1:X \times Y \rightarrow \mathbb{R}$ and $K_2:Y \times Z \rightarrow \mathbb{R}$ is defined as
-\begin{equation*}
-	\left(K_1 K_2\right)(x,z) := \sum_{y \in Y} K_1(x,y) K_2(y,z)
-\end{equation*}
-if $Y$ is a discrete set, and
-\begin{equation*}
-	\left(K_1 K_2\right)(x,z) := \int_Y K_1(x,y) K_2(y,z) dy	
-\end{equation*}
-if $Y\subset \mathbb{R}^d$ is a domain.
-
-The number of smooth functions, $r$, required to well-approximate $T$ using \eqref{eq:formula_for_A} can be small even if the numerical rank of $\mathcal{A}$ is large.
-For example, let $\varphi$ be any convolution kernel, and suppose $\mathcal{A}$ is the convolution operator $\left(\mathcal{A}v\right)^* := \varphi \ast v$.
-The rank of $\mathcal{A}$ can be made arbitrarily large or small by changing the convolution kernel $\varphi$.
-At the same time, $T=\Psi U$ holds with $r=1$, where $\Psi(x,1):= \varphi(x)$ and $U(1,x):=1$.
-
-We may determine $\Psi$ using a continuous analog of the process used to form the interpolative matrix factorization Let $S:=\{x_j\}_{j=1}^q \subset \Omega$ be the set of $q \ge r$ sample points determined by the method in Section \ref{sec:choosing_sample_points}, and define $T_S:\Omega \times \{1,\dots,q\}\rightarrow\mathbb{R}$ to be the semi-discrete kernel with $i^\text{th}$ column $\varphi_{x_i}$.
-That is,
-\begin{equation*}
-	T_S(y,i):= \varphi_{x_i}(y) = T(y,x_i).	
-\end{equation*}
-Further, let $U_S$ be the $r \times q$ matrix with $(i,j)$ entry given by $u_i(x_j)$, and assume that the points $x_i$ are chosen such that $U_S$ has full row rank.
-This is not difficult; for example, if the points $x_i$ were chosen uniformly at random then $U_S$ would have full rank almost surely.
-Indeed, with uniform random points, $\frac{|\Omega|}{q}\left(U_S U_S^T\right)_{ij}$ is a Monte-Carlo approximation of $\langle u_i, u_j \rangle_{L^2(\Omega)}$, where $|\Omega|$ is the total volume of the domain $\Omega$.
-Thus $\frac{|\Omega|}{q}\left(U_S U_S^T\right) \rightarrow I$ as $q \rightarrow \infty$, where $I$ is the identity matrix.
-This implies $\sigma_\text{min} \rightarrow \sqrt{q/|\Omega|}$ as $q \rightarrow \infty$, where $\sigma_\text{min}$ is the minimum singular value of $U_S$.
-We do not use uniform random sampling to choose the points, but the way we choose points does achieve relatively uniform coverage of the domain.
-
-Restricting the second input of both sides of the equation \eqref{eq:formula_for_A} to points in $S$, we have
-\begin{equation}
-\label{eq:A_I_equals_X_U_I}
-\widetilde{T}_S = \Psi U_S.
-\end{equation}
-Multiplying both sides of \cref{eq:A_I_equals_X_U_I} by the Moore-Penrose pseudoinverse of $U_S$, denoted $U_S^+$, and using the fact that $U_S$ has full row rank, yields
-\begin{equation}
-\label{eq:formula_for_X}
-\widetilde{T}_S U_S^+ = \Psi.
-\end{equation}
-Substituting the left hand side of \eqref{eq:formula_for_X} for $\Psi$ into \eqref{eq:formula_for_A} yields the following interpolative-type low rank factorization $\widetilde{T}$ of $T$:
-\begin{equation}
-\label{eq:interpolative_decomposition0}
-T \approx \widetilde{T} = T_S U_S^+ U.
-\end{equation}
-%We will quantify the accuracy of this approximation in Theorem \ref{thm:smooth_basis_err} in Section \ref{sec:error_estimate_adaptive_algorithm}.
-
-%Briefly, the error will be small when (1) $\mathcal{U}$ well-approximates the row space of $\mathcal{T}$, and (2) the smallest singular value of $\mathcal{U}_X$ is not small.
-
-\subsection{Smooth basis of Laplacian eigenmodes}
-
-For the basis of smooth functions $u_i$, we use the dominant eigenfunctions of the inverse Laplacian on $\Omega$ with Neumann boundary conditions.
-This basis of inverse Laplacian eigenfunctions generalizes the Fourier basis to non-rectangular domains.
-Let $\Delta_N^{-1}$ denote the solution operator for the following PDE:
-\begin{equation}
-\label{eq:laplacian_pde}
-\begin{cases}
--\Delta u =f, & \text{in } \Omega, \\
-n \cdot \nabla u = 0, & \text{on } \partial \Omega, \\
-\int_\Omega u dx = 0, &
+\section{Using the approximation}
+\label{sec:using_approximation}
+
+\subsection{Evaluating kernel entries of the approximation}
+
+We now describe how to efficiently evaluate $\widetilde{A}(y,x)$ at an arbitrary point $(x,y) \in \Omega \times \Omega$. From \eqref{eq:H_kernel_entries}, we have
+\begin{equation}
+\label{eq:eval_y_x}
+\widetilde{A}(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x),
+\end{equation}
+and from \eqref{eq:varphi_eval}, we have
+\begin{equation}
+\label{eq:kernel_entry_formua}
+\varphi_{x_i}(y-x) = \begin{cases}
+\eta^{b(i)}(y - x + x_i), & y - x + x_i \in E_{x_i} \\
+0, & \text{otherwise},
 \end{cases}
 \end{equation}
-where $\Delta$ is the Laplacian and $n$ is the normal to $\partial \Omega$.
-That is, $\Delta_N^{-1}$ is the map from source $f$ to PDE solution $u$.
-The first smooth basis function used for interpolation is the constant function $u_1(x):=1$.
-The remaining smooth basis functions, $u_2, u_3, \dots, u_r$, are the eigenfunctions corresponding to the $r-1$ largest eigenvalues of $\Delta_N^{-1}$.
-These eigenfunctions may be computed efficiently using randomized SVD \cite{HMT11}, Krylov methods such as Lanczos, power iterations with deflation, or other methods.
-We use randomized SVD.
-All of these methods require the ability to apply $\Delta_N^{-1}$ to given vectors, which amounts to solving PDE \eqref{eq:laplacian_pde} with the given vector as the right hand side source.
-Highly efficient multigrid schemes exist for solving this PDE.
-%; we use an algebraic multigrid implementation in PyAMG.
-
-\subsection{Action of the approximate operator}
-
-Defining
-\begin{equation*}
-	W := U_S^+ U,
-\end{equation*}
-we may write $\widetilde{T}$ as
-\begin{equation*}
-	\widetilde{T} = T_S W.	
-\end{equation*}
-Using the definition of $T_S$, we may write the entries of the resulting approximation to $A$ as
-\begin{equation}
-\label{eq:H_kernel_entries}
-	\widetilde{A}(y,x) = \widetilde{T}(y-x,x) = \sum_{i=1}^q w_i(x) \varphi_{x_i}(y-x) 
-\end{equation}
-where the functions $w_i(x):=W(i,x)$ are the rows of $W$, which are linear combinations of the functions $u_i$.
-%Since the Laplacian eigenfunctions $u_i$ are defined on $\Omega$, the weighting functions $w_i$ are defined on all of $\Omega$, so $T$ and $H$ are defined on $\mathbb{R}^d \times \Omega$.
-However, the error estimate we present in the next section will only be valid for these operators on $\mathbb{R}^d \times \Omega_I$.
-From \eqref{eq:H_kernel_entries}, we derive the action of $\widetilde{\mathcal{A}}$ as
-\begin{align*}
-	\left(\widetilde{\mathcal{A}}u\right)(v) &= \int_\Omega \int_\Omega v(y) \widetilde{A}(y,x)u(x) dx dy \\
-	&= \int_\Omega \int_\Omega v(y) \widetilde{T}(y-x,x)u(x) dx dy \\
-	&= \int_\Omega \int_\Omega v(y) \left(\sum_{i=1}^q  \varphi_{x_i}(y-x) w_i(x)\right) u(x) dx dy \\
-	&= \int_\Omega v(y) \sum_{i=1}^q \int_\Omega \varphi_{x_i}(y-x) w_i(x) u(x) dx dy\\
-	&= \int_\Omega v(y) \left(\sum_{i=1}^q \varphi_{x_i} \ast \left(w_i \cdot u\right)\right)(y) dy,
-\end{align*}
-which implies
-\begin{equation*}
-	\left(\widetilde{\mathcal{A}} u	\right)^* = \sum_{i=1}^q \varphi_{x_i} \ast \left(w_i \cdot u\right)
-\end{equation*}
-Here $f \ast g$ denotes the convolution of the functions $f$ and $g$. Thus, $\widetilde{\mathcal{A}}$ is a weighted sum of convolution operators with smooth spatially varying weighting functions.
-
-\subsection{Efficient computation of approximation kernel entries}
-
-We now describe how to efficiently evaluate $\widetilde{A}(x',y')$ at arbitrary points $(x',y')$. Let $b(i)$ denote the batch index such that $x_i \in S^{b(i)}$, and define
-\begin{equation*}
-	\zeta_i := y' - x' + x_i,
-\end{equation*}
-for $x_i \in S$. From \eqref{eq:H_kernel_entries}, we have
-\begin{equation*}
-	\widetilde{A}(y',x') = \sum_{i=1}^q w_i(x') \varphi_{x_i}(y'-x'),
-\end{equation*}
-and from \eqref{eq:varphi_eval}, we have
-\begin{equation}
-\label{eq:kernel_entry_formua}
-	\varphi_{x_i}(y'-x') = \begin{cases}
-		\eta^{b(i)}(\zeta_i), & \zeta_i \in E_{x_i} \\
-		0, & \text{otherwise}.
-	\end{cases}
-\end{equation}
+where $b(i)$ denotes the batch index such that $x_i \in S^{b(i)}$.
+To compute $z=\widetilde{A}(y,x)$, we start with $z=0$, then loop through all batches $b$, and all sample points in each batch.
+For each sample point, we add one term in the sum in \eqref{eq:eval_y_x} to $z$, using \eqref{eq:kernel_entry_formua} as needed to evaluate $\varphi_{x_i}$.
+This is shown in Algorithm \ref{alg:eval_y_x}.
 %
@@
 %
-%\begin{algorithm2e}
-%	\SetAlgoNoLine
-%	\SetKwInOut{Input}{Input}
-%	\SetKwInOut{Output}{Output}
-%	\SetKwProg{Fn}{Function}{}{}
-%	\Input{Points $\{x_k'\}_{k=1}^n \subset \Omega$}	
-%	\Output{Weighting function values $\boldsymbol{\omega} \in \mathbb{R}^{q \times n}$, $\boldsymbol{\omega}_{i,k} = w_i(x_k')$}
-%	
-%	\mbox{}\\
-%	
-%	\Fn{$\computeweightingentries{\{x'_k\}_{k=1}^n}$}{
-%		$\boldsymbol{\nu} \gets \mathbf{0}_{q\times n}$
-%		
-%		\For{$i=1,\dots,q$}{
-%			$\boldsymbol{\nu}[i,:] \gets \interpolate(\mathbf{u}_i, \{x_1',x_2',\dots,x_k'\})$
-%		}
-%		
-%		$\boldsymbol{\omega} \gets \mathbf{U}[:,\mathtt{S}]^+ \boldsymbol{\nu}$
-%		
-%		\Return{$\boldsymbol{\omega}$}
-%	}
-%	
-%	\caption{Compute many weighting function values $\omega_{ik} = w_i(x_k')$}
-%	\label{alg:weighting_entries}
-%\end{algorithm2e}
-%
-%
+\begin{algorithm2e}
+	\SetAlgoNoLine
+
+	$z \gets 0$
+
+	\For{$b=1,2,\dots,\text{}max\_batches$}{
+		\For{$x_i \in S_b$}{
+			\If{$y - x + x_i \in E_{x_i}$}{
+				$z \gets z + w_i(x)\eta^b(y-x+x_i)$
+			}
+		}
+	}	
+	\caption{Compute approximate kernel entry $z = \widetilde{A}(y,x)$}
+	\label{alg:eval_y_x}
+\end{algorithm2e}
+
 %\begin{algorithm2e}
@@
 %	
+%	Compute $\boldsymbol{\omega} \in \mathbb{R}^{r \times n}$, $\boldsymbol{\omega}_{i,k} = w_i(x_k')$
+%	
 %	\mbox{}\\
@@
 %	}
-%	\caption{Compute many approximate kernel entries $\widetilde{H}(y_k',x_k')$}
+%	\caption{Compute many approximate kernel entries $z_k = \widetilde{A}(y_k',x_k')$}
 %	\label{alg:H_entries}
 %\end{algorithm2e}
+
+\subsection{Conversion to hierarchical matrix format}
+
+In order to use the product-convolution approximation $\widetilde{\mathcal{A}}$, we convert $\widetilde{\mathcal{A}}$ to hierarchical matrix (H-matrix) format.
+Many classes of hierarchical matrices exist---H, uniform H1, H2, HODLR, HSS, HBS, and more.
+We use H1 matrices in our numerical results, but any of the other H-matrix formats could be used instead.
+The basic ideas behind H-matrices are as follows:
+\begin{itemize} 
+	\item Often, large dense matrices of practical interest may be permuted, then partitioned into blocks recursively, in such a way that many off-diagonal blocks of the matrix are numerically low rank, even if the matrix is high rank.
+	\item Recursive algorithms can take advantage of this low rank block structure to perform matrix arithmetic fast.
+Conventional algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix addition require either $O(n^2)$ or $O(n^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(n \log^a n)$ time and memory for hierarchical matrices (here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix).
+\end{itemize}
+For more details on H-matrices, we recommend [CITES].
+
+We use the standard geometrical clustering/adaptive cross H1 matrix construction method implemented within the HLibPro software package [CITE] to convert $\widetilde{\mathcal{A}}$ into H1 matrix format.
+We now summarize this method.
+
+\paragraph{Cluster trees} First, the degrees of freedom associated with the columns of the matrix are hierarchically clustered into a \emph{cluster tree} using recursive hyperplane splitting.
+Since we use finite elements to discretize the problem, we associate the $i^\text{th}$ column of $\widetilde{\mathcal{A}}$ with the nodal point in $\mathbb{R}^d$ associated with the $i^\text{th}$ finite element basis function.
+The columns of the matrix thus correspond to a point cloud in $\mathbb{R}^d$.
+We split this point cloud into two roughly equally sized \emph{child} point clouds, using a hyperplane which is perpendicular to the coordinate axis direction in which the point cloud is widest (e.g., either the $x$, $y$, or $z$ axis in $3$D).
+The two child point clouds are split in the same way.
+This splitting process repeats recursively until the point clouds have less than a preset number of points (we use $32$ points).
+This hierarchical partitioning of the point cloud into smaller and smaller point clouds corresponds to a hierarchical partitioning of the columns of the matrix into smaller and smaller \emph{clusters} of columns, which we call a cluster tree.
+A depth-first search ordering of the cluster tree is then generated.
+When the columns of the matrix are permuted into this ordering, the columns associated with each cluster in the cluster tree are contiguous.
+
+In the same way, the degrees of freedom associated with the rows of the matrix are hierarchically clustered into another cluster tree, and an ordering for the rows is generated.
+In our examples, the degrees of freedom for the columns coincide with the degrees of freedom for the rows, so the cluster trees for the rows and columns are the same, but this is not required in general.
+
+\paragraph{Block cluster tree} Second, we partition the matrix (after reordering the rows and columns via the depth-first ordering described above) into a recursive hierarchy of mostly low rank blocks called the \emph{block cluster tree}.
+The idea is that the interaction between far-away groups of degrees of freedom may be characterized by a small number of bulk properties of those groups.
+
+\section{Interpolating impulse response samples old}
+
+We interpolate the functions $\varphi_{x_i}$ associated with the points $x_i \in S$, to generate functions $\varphi_x$ associated with new points $x \notin S$.
+This is done using a continuous generalization of the interpolative decomposition for matrices, wherein a select number of columns of a matrix are interpolated to fill the whole matrix.
+Interpolating $\varphi_x$ differs from interpolative low rank approximation of $\mathcal{A}$, which corresponds to interpolating $\langle\mathcal{A},\delta_x\rangle^*$.
+Because of the shifting that maps $\langle\mathcal{A}, \delta_x\rangle$ to $\varphi_x$, interpolation of $\varphi_{x_i}$ yields an approximation $\widetilde{\mathcal{A}}$ which can be high-rank.
+
+The interpolation procedure requires a basis of smooth functions.
+We form this basis from smooth eigenmodes of the Laplacian on $\Omega$ with Neumann boundary conditions.
+The resulting approximation $\mathcal{A}$ may be interpreted as a weighted sum of convolution operators, with smooth spatially varying weighting functions.
+Kernel entries of the approximation, $\widetilde{A}(y,x)$, may be evaluated efficiently.
+
+\subsection{Interpolative factorization of the SVIR}
+\label{sec:interp_fact_svir}
+
+The \emph{spatially varying impulse response}, denoted $T$, is the reinterpretation of $\varphi_x(y)$ as a function of both $y$ and $x$:
+\begin{equation*}
+	T(y,x) := \varphi_x(y) = A(y+x,x).
+\end{equation*}
+Since $T(y,x)$ is smooth as a function of $x$ (Condition 1), the continuous analog of the row space of $T$ should be well-approximated by the span of a small number of orthonormal smooth functions, $\{u_k\}_{k=1}^r$.
+Condition 1, therefore, is equivalent to the existence of a low rank approximation $\widetilde{T}$ of $T$ of the form
+\begin{equation}
+	\label{eq:formula_for_A}
+	T \approx \widetilde{T} = \Psi U
+\end{equation}
+where $\Psi:\mathbb{R}^d \times \{1,\dots,r\} \rightarrow \mathbb{R}$ is an unknown semi-discrete kernel with $r$ columns, and $U:\{1,\dots,r\} \times \Omega \rightarrow \mathbb{R}$ is the semi-discrete kernel with $r$ rows, such that the $i^\text{th}$ row of $U$ is the function $u_i$.
+That is,
+\begin{equation*}
+	U(i,x) = u_i(x).
+\end{equation*}
+The multiplication $K_1 K_2$ of kernels $K_1:X \times Y \rightarrow \mathbb{R}$ and $K_2:Y \times Z \rightarrow \mathbb{R}$ is defined as
+\begin{equation*}
+	\left(K_1 K_2\right)(x,z) := \sum_{y \in Y} K_1(x,y) K_2(y,z)
+\end{equation*}
+if $Y$ is a discrete set, and
+\begin{equation*}
+	\left(K_1 K_2\right)(x,z) := \int_Y K_1(x,y) K_2(y,z) dy	
+\end{equation*}
+if $Y\subset \mathbb{R}^d$ is a domain.
+
+The number of smooth functions, $r$, required to well-approximate $T$ using \eqref{eq:formula_for_A} can be small even if the numerical rank of $\mathcal{A}$ is large.
+For example, let $\varphi$ be any convolution kernel, and suppose $\mathcal{A}$ is the convolution operator $\left(\mathcal{A}v\right)^* := \varphi \ast v$.
+The rank of $\mathcal{A}$ can be made arbitrarily large or small by changing the convolution kernel $\varphi$.
+At the same time, $T=\Psi U$ holds with $r=1$, where $\Psi(x,1):= \varphi(x)$ and $U(1,x):=1$.
+
+We may determine $\Psi$ using a continuous analog of the process used to form the interpolative matrix factorization Let $S:=\{x_j\}_{j=1}^q \subset \Omega$ be the set of $q \ge r$ sample points determined by the method in Section \ref{sec:choosing_sample_points}, and define $T_S:\Omega \times \{1,\dots,q\}\rightarrow\mathbb{R}$ to be the semi-discrete kernel with $i^\text{th}$ column $\varphi_{x_i}$.
+That is,
+\begin{equation*}
+	T_S(y,i):= \varphi_{x_i}(y) = T(y,x_i).	
+\end{equation*}
+Further, let $U_S$ be the $r \times q$ matrix with $(i,j)$ entry given by $u_i(x_j)$, and assume that the points $x_i$ are chosen such that $U_S$ has full row rank.
+This is not difficult; for example, if the points $x_i$ were chosen uniformly at random then $U_S$ would have full rank almost surely.
+Indeed, with uniform random points, $\frac{|\Omega|}{q}\left(U_S U_S^T\right)_{ij}$ is a Monte-Carlo approximation of $\langle u_i, u_j \rangle_{L^2(\Omega)}$, where $|\Omega|$ is the total volume of the domain $\Omega$.
+Thus $\frac{|\Omega|}{q}\left(U_S U_S^T\right) \rightarrow I$ as $q \rightarrow \infty$, where $I$ is the identity matrix.
+This implies $\sigma_\text{min} \rightarrow \sqrt{q/|\Omega|}$ as $q \rightarrow \infty$, where $\sigma_\text{min}$ is the minimum singular value of $U_S$.
+We do not use uniform random sampling to choose the points, but the way we choose points does achieve relatively uniform coverage of the domain.
+
+Restricting the second input of both sides of the equation \eqref{eq:formula_for_A} to points in $S$, we have
+\begin{equation}
+\label{eq:A_I_equals_X_U_I}
+\widetilde{T}_S = \Psi U_S.
+\end{equation}
+Multiplying both sides of \cref{eq:A_I_equals_X_U_I} by the Moore-Penrose pseudoinverse of $U_S$, denoted $U_S^+$, and using the fact that $U_S$ has full row rank, yields
+\begin{equation}
+\label{eq:formula_for_X}
+\widetilde{T}_S U_S^+ = \Psi.
+\end{equation}
+Substituting the left hand side of \eqref{eq:formula_for_X} for $\Psi$ into \eqref{eq:formula_for_A} yields the following interpolative-type low rank factorization $\widetilde{T}$ of $T$:
+\begin{equation}
+\label{eq:interpolative_decomposition0}
+T \approx \widetilde{T} = T_S U_S^+ U.
+\end{equation}
+%We will quantify the accuracy of this approximation in Theorem \ref{thm:smooth_basis_err} in Section \ref{sec:error_estimate_adaptive_algorithm}.
+
+%Briefly, the error will be small when (1) $\mathcal{U}$ well-approximates the row space of $\mathcal{T}$, and (2) the smallest singular value of $\mathcal{U}_X$ is not small.
+
+\subsection{Smooth basis of Laplacian eigenmodes}
+
+For the basis of smooth functions $u_i$, we use the dominant eigenfunctions of the inverse Laplacian on $\Omega$ with Neumann boundary conditions.
+This basis of inverse Laplacian eigenfunctions generalizes the Fourier basis to non-rectangular domains.
+Let $\Delta_N^{-1}$ denote the solution operator for the following PDE:
+\begin{equation}
+\label{eq:laplacian_pde}
+\begin{cases}
+-\Delta u =f, & \text{in } \Omega, \\
+n \cdot \nabla u = 0, & \text{on } \partial \Omega, \\
+\int_\Omega u dx = 0, &
+\end{cases}	
+\end{equation}
+where $\Delta$ is the Laplacian and $n$ is the normal to $\partial \Omega$.
+That is, $\Delta_N^{-1}$ is the map from source $f$ to PDE solution $u$.
+The first smooth basis function used for interpolation is the constant function $u_1(x):=1$.
+The remaining smooth basis functions, $u_2, u_3, \dots, u_r$, are the eigenfunctions corresponding to the $r-1$ largest eigenvalues of $\Delta_N^{-1}$.
+These eigenfunctions may be computed efficiently using randomized SVD \cite{HMT11}, Krylov methods such as Lanczos, power iterations with deflation, or other methods.
+We use randomized SVD.
+All of these methods require the ability to apply $\Delta_N^{-1}$ to given vectors, which amounts to solving PDE \eqref{eq:laplacian_pde} with the given vector as the right hand side source.
+Highly efficient multigrid schemes exist for solving this PDE.
+%; we use an algebraic multigrid implementation in PyAMG.
+
+\subsection{Action of the approximate operator}
+
+Defining
+\begin{equation*}
+	W := U_S^+ U,
+\end{equation*}
+we may write $\widetilde{T}$ as
+\begin{equation*}
+	\widetilde{T} = T_S W.	
+\end{equation*}
+Using the definition of $T_S$, we may write the entries of the resulting approximation to $A$ as
+\begin{equation}
+\label{eq:H_kernel_entries}
+	\widetilde{A}(y,x) = \widetilde{T}(y-x,x) = \sum_{i=1}^q w_i(x) \varphi_{x_i}(y-x) 
+\end{equation}
+where the functions $w_i(x):=W(i,x)$ are the rows of $W$, which are linear combinations of the functions $u_i$.
+%Since the Laplacian eigenfunctions $u_i$ are defined on $\Omega$, the weighting functions $w_i$ are defined on all of $\Omega$, so $T$ and $H$ are defined on $\mathbb{R}^d \times \Omega$.
+However, the error estimate we present in the next section will only be valid for these operators on $\mathbb{R}^d \times \Omega_I$.
+From \eqref{eq:H_kernel_entries}, we derive the action of $\widetilde{\mathcal{A}}$ as
+\begin{align*}
+	\left(\widetilde{\mathcal{A}}u\right)(v) &= \int_\Omega \int_\Omega v(y) \widetilde{A}(y,x)u(x) dx dy \\
+	&= \int_\Omega \int_\Omega v(y) \widetilde{T}(y-x,x)u(x) dx dy \\
+	&= \int_\Omega \int_\Omega v(y) \left(\sum_{i=1}^q  \varphi_{x_i}(y-x) w_i(x)\right) u(x) dx dy \\
+	&= \int_\Omega v(y) \sum_{i=1}^q \int_\Omega \varphi_{x_i}(y-x) w_i(x) u(x) dx dy\\
+	&= \int_\Omega v(y) \left(\sum_{i=1}^q \varphi_{x_i} \ast \left(w_i \cdot u\right)\right)(y) dy,
+\end{align*}
+which implies
+\begin{equation*}
+	\left(\widetilde{\mathcal{A}} u	\right)^* = \sum_{i=1}^q \varphi_{x_i} \ast \left(w_i \cdot u\right)
+\end{equation*}
+Here $f \ast g$ denotes the convolution of the functions $f$ and $g$. Thus, $\widetilde{\mathcal{A}}$ is a weighted sum of convolution operators with smooth spatially varying weighting functions.
 

### 2021-01-01 paper/localpsf.tex
H matrix block cluster tree writing

@@
 In order to use the product-convolution approximation $\widetilde{\mathcal{A}}$, we convert $\widetilde{\mathcal{A}}$ to hierarchical matrix (H-matrix) format.
-Many classes of hierarchical matrices exist---H, uniform H1, H2, HODLR, HSS, HBS, and more.
+Many classes of hierarchical matrices exist---H1, uniform H1, H2, HODLR, HSS, HBS, and more.
 We use H1 matrices in our numerical results, but any of the other H-matrix formats could be used instead.
@@
 
-We use the standard geometrical clustering/adaptive cross H1 matrix construction method implemented within the HLibPro software package [CITE] to convert $\widetilde{\mathcal{A}}$ into H1 matrix format.
-We now summarize this method.
+We convert $\widetilde{\mathcal{A}}$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package [CITE].
+In the remainder of this section we describe this method in detail.
 
@@
 The columns of the matrix thus correspond to a point cloud in $\mathbb{R}^d$.
-We split this point cloud into two roughly equally sized \emph{child} point clouds, using a hyperplane which is perpendicular to the coordinate axis direction in which the point cloud is widest (e.g., either the $x$, $y$, or $z$ axis in $3$D).
+We split this point cloud into two equally sized \emph{child} point clouds, using a hyperplane which is perpendicular to the coordinate axis direction in which the point cloud is widest (e.g., either the $x$, $y$, or $z$ axis in $3$D).
 The two child point clouds are split in the same way.
-This splitting process repeats recursively until the point clouds have less than a preset number of points (we use $32$ points).
-This hierarchical partitioning of the point cloud into smaller and smaller point clouds corresponds to a hierarchical partitioning of the columns of the matrix into smaller and smaller \emph{clusters} of columns, which we call a cluster tree.
-A depth-first search ordering of the cluster tree is then generated.
-When the columns of the matrix are permuted into this ordering, the columns associated with each cluster in the cluster tree are contiguous.
+This splitting process repeats until the point clouds have less than a preset number of points (we use $32$ points).
+This hierarchical partitioning of the point cloud into smaller and smaller point clouds corresponds to a hierarchical partitioning of the columns of the matrix into smaller and smaller \emph{clusters} of columns.
+This hierarchical partitioning of the columns forms a binary tree, which we call a cluster tree.
+The root of the tree is the set of all columns, and the leaves of the tree are clusters of columns that are not subdivided any further.
+A depth-first search ordering of the cluster tree leaves is then generated.
+When the columns of the matrix are permuted into this depth-first ordering, the columns associated with each cluster in the cluster tree are contiguous.
 
@@
 
-\paragraph{Block cluster tree} Second, we partition the matrix (after reordering the rows and columns via the depth-first ordering described above) into a recursive hierarchy of mostly low rank blocks called the \emph{block cluster tree}.
-The idea is that the interaction between far-away groups of degrees of freedom may be characterized by a small number of bulk properties of those groups.
+\paragraph{Block cluster tree} Second, we partition the matrix into a recursive hierarchy of mostly low rank blocks called the \emph{block cluster tree}.
+The idea is that a block of the matrix is likely to be low rank if the point cloud associated with the rows of the block is far away from the point cloud associated with the columns of the block.
+This is reasonable to expect here because of the locality property of $\mathcal{A}$.
+Indeed, locality implies that many blocks of the matrix corresponding to far away point cloud clusters will be rank zero.
+
+After reordering the rows and columns of $\widetilde{\mathcal{A}}$ via the depth-first ordering described above, we partition the reordered version of $\widetilde{\mathcal{A}}$ into a tree of $2 \times 2$ block matrices recursively.
+We use a geometric admissibility condition (discussed below) to decide which blocks to subdivide, and use the cluster trees for the rows and columns to determine how to subdivide those blocks.
+For the first stage of subdivision, let $r_1$ and $r_2$ be the children row clusters for the root of the row cluster tree, let $c_1$ and $c_2$ be the children column clusters for the root of the column cluster tree.
+The matrix $\widetilde{\mathcal{A}}$ is partitioned into blocks as follows:
+\begin{equation*}
+	\begin{bmatrix}
+		\widetilde{\mathcal{A}}_{11} & \widetilde{\mathcal{A}}_{12} \\
+		\widetilde{\mathcal{A}}_{21} & \widetilde{\mathcal{A}}_{22},
+	\end{bmatrix}
+\end{equation*}
+where $\widetilde{\mathcal{A}}_{11}$ denotes the block of $\widetilde{\mathcal{A}}$ with rows $r_1$ and columns $c_1$, $\widetilde{\mathcal{A}}_{12}$ denotes the block of $\widetilde{\mathcal{A}}$ with rows $r_1$ and columns $c_2$, and similar for $\widetilde{\mathcal{A}}_{21}$ and $\widetilde{\mathcal{A}}_{22}$.
+
+We now loop through the four blocks, $\widetilde{\mathcal{A}}_{11}$, $\widetilde{\mathcal{A}}_{12}$, $\widetilde{\mathcal{A}}_{21}$, and $\widetilde{\mathcal{A}}_{22}$, and decide which blocks should be subdivided further.
+For the purpose of explanation, consider $\widetilde{\mathcal{A}}_{12}$.
+If
+\begin{equation}
+	\label{eq:weak_admissibility_cond}
+	\dist\left(r_1, c_2\right) \ge \eta \min\left(\diam\left(r_1\right), \diam\left(c_2\right)\right),
+\end{equation}
+then we mark $\widetilde{\mathcal{A}}_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone.
+Here $\dist\left(r_1, c_2\right)$ is the Euclidean distance between the axis-aligned bounding box for the point cloud associated with the row cluster $r_1$, and the axis aligned bounding box for the point cloud associated with the column cluster $c_2$.
+The quantity $\diam\left(r_1\right)$ is the diameter of the axis aligned bounding box for the point cloud associated with the row cluster $r_1$, and $\diam\left(c_2\right)$ is the analogous diameter associated with the column cluster $c_2$.
+The quantity $\eta$ is a scalar constant; we use $\eta=2.0$.
+Basically, if the point clouds associated with $r_1$ and $c_2$ are far away from each other relative to their diameters, then we expect that the corresponding block of the matrix will be low rank.
+This process is repeated for the other blocks to determine which blocks are admissible and which are not.
+For us, the diagonal blocks $\widetilde{\mathcal{A}}_{11}$ and $\widetilde{\mathcal{A}}_{22}$ are not admissible because the distance between a point cloud and itself is zero.
+
+Next, we subdivide all blocks that not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\mathcal{\widetilde{A}}$.
+But now we subdivide a block based on the two child row clusters and two child column clusters for the rows and columns of that block.
+This subdivision process continues recursively until all blocks are either admissible, or smaller than the predetermined size mentioned above.
+The resulting hierarchical partitioning of matrix blocks forms a tree (quad-tree in $2$D, or oct-tree in $3$D), which is called the \emph{block cluster tree}.
+The root of the tree is the whole matrix, internal nodes in the tree are blocks that are subdivided, and the leaves of the tree are blocks that are either expected to be low rank, or are small.
+
+Note that the cluster trees and block cluster tree are computed using only the geometric information about the degree of freedom locations, and do not depend on the values of the matrix entries of $\widetilde{\mathcal{A}}$.
+
+\paragraph{Adaptive cross low rank approximation of blocks} WORK IN PROGRESS
 

### 2021-01-02 paper/localpsf.tex
adaptive cross writing

@@
 \subsection{Evaluating kernel entries of the approximation}
+\label{sec:eval_matrix_entries}
 
@@
 In order to use the product-convolution approximation $\widetilde{\mathcal{A}}$, we convert $\widetilde{\mathcal{A}}$ to hierarchical matrix (H-matrix) format.
-Many classes of hierarchical matrices exist---H1, uniform H1, H2, HODLR, HSS, HBS, and more.
+Many classes of hierarchical matrices exist---H1, uniform H1, H2, HODLR, HSS, HBS, and more [CITES].
 We use H1 matrices in our numerical results, but any of the other H-matrix formats could be used instead.
@@
 We convert $\widetilde{\mathcal{A}}$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package [CITE].
-In the remainder of this section we describe this method in detail.
-
-\paragraph{Cluster trees} First, the degrees of freedom associated with the columns of the matrix are hierarchically clustered into a \emph{cluster tree} using recursive hyperplane splitting.
+Briefly, the H-matrix construction process that we use proceeds as follows.
+First, we construct a hierarchical partitionings of the degrees of freedom for the rows and columns of the matrix (cluster trees, Section \ref{sec:cluster_trees}).
+Second, we construct a hierarchical partitioning of the blocks of the matrix, in such a way that many of the blocks in the partitioning are expected to be low rank, and the remaining high rank blocks are small (block cluster tree, Section \ref{sec:block_cluster_tree}).
+Finally, we form low rank approximations of the blocks of the matrix that are expected to be low rank (adaptive cross approximation, Section \ref{sec:adaptive_cross}), and fill in the remaining high rank  blocks with their numerical values.
+The first two steps require geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix, but these steps do not depend on the particular values of matrix entries.
+The third step requires us to evaluate $O(n \log n)$ specially-chosen matrix entries of $\widetilde{\mathcal{A}}$, which we do using the method described in Section \ref{sec:eval_matrix_entries}.
+In the remainder of this section we describe each of these steps in more detail.
+
+\subsubsection{Cluster trees}
+\label{sec:cluster_trees}
+
+The degrees of freedom associated with the columns of the matrix are hierarchically clustered into a \emph{cluster tree} using recursive hyperplane splitting.
 Since we use finite elements to discretize the problem, we associate the $i^\text{th}$ column of $\widetilde{\mathcal{A}}$ with the nodal point in $\mathbb{R}^d$ associated with the $i^\text{th}$ finite element basis function.
@@
 The root of the tree is the set of all columns, and the leaves of the tree are clusters of columns that are not subdivided any further.
+
 A depth-first search ordering of the cluster tree leaves is then generated.
@@
 
-\paragraph{Block cluster tree} Second, we partition the matrix into a recursive hierarchy of mostly low rank blocks called the \emph{block cluster tree}.
+\subsubsection{Block cluster tree} 
+\label{sec:block_cluster_tree}
+
+We partition the matrix into a recursive hierarchy of mostly low rank blocks called the \emph{block cluster tree}.
 The idea is that a block of the matrix is likely to be low rank if the point cloud associated with the rows of the block is far away from the point cloud associated with the columns of the block.
@@
 \end{equation*}
-where $\widetilde{\mathcal{A}}_{11}$ denotes the block of $\widetilde{\mathcal{A}}$ with rows $r_1$ and columns $c_1$, $\widetilde{\mathcal{A}}_{12}$ denotes the block of $\widetilde{\mathcal{A}}$ with rows $r_1$ and columns $c_2$, and similar for $\widetilde{\mathcal{A}}_{21}$ and $\widetilde{\mathcal{A}}_{22}$.
+where $\widetilde{\mathcal{A}}_{11}$ denotes the block of $\widetilde{\mathcal{A}}$ with rows $r_1$ and columns $c_1$, $\widetilde{\mathcal{A}}_{12}$ denotes the block of $\widetilde{\mathcal{A}}$ with rows $r_1$ and columns $c_2$, and so on for $\widetilde{\mathcal{A}}_{21}$ and $\widetilde{\mathcal{A}}_{22}$.
 
@@
 
-Next, we subdivide all blocks that not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\mathcal{\widetilde{A}}$.
+Next, we subdivide all blocks that are not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\mathcal{\widetilde{A}}$.
 But now we subdivide a block based on the two child row clusters and two child column clusters for the rows and columns of that block.
@@
 
-Note that the cluster trees and block cluster tree are computed using only the geometric information about the degree of freedom locations, and do not depend on the values of the matrix entries of $\widetilde{\mathcal{A}}$.
-
-\paragraph{Adaptive cross low rank approximation of blocks} WORK IN PROGRESS
+%Note that the cluster trees and block cluster tree are computed using only the geometric information about the degree of freedom locations, and do not depend on the values of the matrix entries of $\widetilde{\mathcal{A}}$.
+
+\subsubsection{Adaptive cross low rank approximation of blocks}
+\label{sec:adaptive_cross}
+
+Once the block cluster tree has been constructed, low rank approximations of the admissible (low rank) blocks are formed using the adaptive cross method [CITE].
+Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\widetilde{\mathcal{A}}$.
+The idea of the adaptive cross method is to form a low rank approximation of $X$ by sampling a small number of rows and columns of $X$.
+\begin{itemize}
+\item Let $C \in \mathbb{R}^{m \times r}$ be a matrix consisting of a subset of $r$ columns of $X$, such that the span of the columns in $C$ approximates the column space of $X$. 
+\item Let $R\in\mathbb{R}^{r \times m}$ be a subset of the rows of $X$, such that the span of the rows in $R$ approximates the row space of $X$.
+\item Let $U \in \mathbb{R}^{r \times r}$ be the block of $X$ consisting of the intersection of the rows from $R$ with the columns from $C$.
+\end{itemize}
+Then it is well-established that
+\begin{equation}
+\label{eq:CUR}
+	X \approx C U^+ R,
+\end{equation}
+where $U^+$ is the pseudoinverse of $U$ [CITE].
+The quality of approximation \eqref{eq:CUR} depends on how well the columns of $C$ approximate the column space of $X$, and how well the rows of $R$ approximate the row space of $X$.
+
+In the adaptive cross method, ``good'' columns, $C$, and rows, $R$, are chosen via an alternating iterative process.
+An initial set of columns $C$ is chosen.
+Keeping $C$ fixed, a set of rows $R$ is chosen so that the so that determinant of the submatrix $U$ within $C$ is as large as possible.
+This may be done via either the maxvol procedure [CITE GOOD SUBMATRIX], or via QR factorization.
+Now, keeping $R$ fixed, a set of new columns $C$ is chosen so that the submatrix $U$ within $R$ is as large as possible.
+This process repeats a small number of times.
+This results in matrices $R$ and $C$ such that the error in the approximation, \eqref{eq:CUR}, is small.
+The dominant cost of this procedure is the cost of computing $kr$ columns of $X$ and $kr$ rows of $X$, where $k$ is the number of alternating iterations.
+There is also a small linear algebra overhead cost for the determinant maximization process that is performed at each step.
+For more details on adaptive cross low rank approximation, see [CITE].
+The key point is that the adaptive cross method allows us to form a rank-$r$ approximation of an $m \times m$ block of the matrix via a process that only requires accessing $O(mr)$ entries of that block.
+
+We use the adaptive cross method to form low rank approximations for each admissible block of $\widetilde{\mathcal{A}}$.
+We directly compute all entries of the small dense blocks of $\widetilde{\mathcal{A}}$ that are not admissible.
+This process requires us to compute $O(r^2 N \log N)$ entries of $\widetilde{\mathcal{A}}$, which is relatively cheap compared to the matrix-vector products with $\mathcal{A}$ that are required to form the product-convolution approximation $\widetilde{\mathcal{A}}$.
 

### 2021-01-03 paper/localpsf.tex
deleted old text from paper

@@
 Our method is able to form high quality approximations of high rank Hessians using only a small number of Hessian matrix-vector products.
-%  Within each Dirac comb, the point sources must be spaced sufficiently far apart so that the impulse responses to those point sources do not overlap.
-We estimate the required spacing of point sources by applying the operator to a small number of constant, linear, and quadratic functions, then postprocessing the results.
 \end{abstract}
@@
 Local translation invariance is illustrated in Figure FIG.
-%We qualify the degree of local translation invariance with the words ``perfectly'', ``highly'', and ``poorly''.
-%\begin{itemize}
-%\item If equality holds in \eqref{eq:local_translation_invariance} for all $x$, $y$, and $h$, then we say $\mathcal{A}$ is \emph{perfectly translation invariant}. 
-%\item If the discrepancy between left hand side and right hand side of \eqref{eq:local_translation_invariance} is small for a large set of points $x$, $y$, and $h$, then we say that $\mathcal{A}$ is \emph{highly locally translation invariant}.
-%\item If the discrepancy between left hand side and right hand side of \eqref{eq:local_translation_invariance} is large, or if the discrepancy is only small for a small set of $x$, $y$, and $h$, then we say that $\mathcal{A}$ is \emph{poorly locally translation invariant}.
-%\end{itemize} 
 If $\mathcal{A}$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then it is straightforward to show that $\mathcal{A}$ would be a convolution operator, and the convolution kernel, $\varphi_x$, would be given by
@@
 
-%A small number of terms in \eqref{eq:product_convolution} are required to accurately approximate a highly locally translation invariant operator, while a large number of terms are required to accurately approximate a poorly locally translation invariant operator.
-
 Once the $\varphi_i$ and $w_i$ are computed, we convert $\widetilde{\mathcal{A}}$ to hierarchical matrix (H-matrix) format.
@@
 Here $u(m)$ denotes the solution of the PDE \eqref{eq:state_pde} as a function of $m$.
-%
-%That is, inverse problems in which one seeks to reconstruct an unknown parameter field, $m$, from noisy measurements,
-%\begin{equation*}
-%	y = B(m,u) + \xi,
-%\end{equation*}
-%of a state variable, $u$, which depends on $m$ indirectly through the solution of a PDE,
-%\begin{equation}
-%\label{eq:state_equation}
-%	0=F(m,u).	
-%\end{equation}
-%Here $\xi$ is the noise.
-The observations $y$ are known, while $m$, $u$, and $\xi$ are unknown.
-In the deterministic setting, the inverse problem of finding $m$ is commonly framed as a least squares minimization problem
-%\begin{equation}
-%\label{eq:minimization_problem}
-%	\min_m \quad J_d(m) + R(m),
-%\end{equation}
-%where
-%\begin{equation*}
-%	J_d(m) := \frac{1}{2}\|y - B(m,u(m))\|_{\Gamma}^2
-%\end{equation*}
-%is a data misfit term, $\Gamma$ is the noise covariance, $R(m)$ is a regularization term, and $u(m)$ denotes the solution $u$ of the PDE \eqref{eq:state_equation} as a function of $m$, . 
-%
-%Approximation of the Hessian,
-%\begin{equation*}
-%	\mathcal{H} = \mathcal{H}_d + \mathcal{H}_R := \frac{d^2 J_d}{dm^2} + \frac{d^2 R}{dm^2},
-%\end{equation*}
 In the deterministic approach to inverse problems, one typically finds $m$ as the solution to the minimization problem
@@
 
-%The regularization term in the Hessian, $\mathcal{H}_R := \frac{d^2 R}{dm^2}$, is typically an elliptic differential operator, and is therefore easy to work with.
-The data misfit term, $\mathcal{H}_d = \frac{d^2 J}{dm^2}$, is difficult to work with.
-The data misfit term $\mathcal{H}_d$ is not directly accessible because it depends on $u$, which depends on $m$ implicitly through the solution of the PDE \eqref{eq:state_equation}.
-%
-%We only have access to $\mathcal{H}_d$ via its action on an arbitrary vectors $v$, i.e., evaluation of the map $v \mapsto \mathcal{H} v$.
-Evaluating $v \mapsto \mathcal{H} v$ is an expensive process that involves solving two auxiliary PDEs.
-We show how to evaluate $v \mapsto \mathcal{H} v$ in Appendix ADSF.
-
 The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term [CITE LOTS OF STUFF].
@@
 The ideal scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost) [CITE MY DISSERTATION].
-%In our numerical results, we show that our operator approximation method can efficiently approximate the data misfit term in the Hessian in certain PDE constrained inverse problems. 
 The method in this paper offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because high-rank approximations of an operator can be formed using a small number of matrix-vector products.
@@
 We build upon the class of product convolution approximations in which the functions, $\varphi_i$, are impulse responses of $\mathcal{A}$ to point sources at a collection of points $x_i$ [CITE BIG LIST].
-%The primary question in methods of this type are how to choose the points $x_i$, and how to choose the weighting functions $w_i$.
 A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the functions $\varphi_i$ with piecewise linear interpolation [CITE NAGY], or splines [CITE SPLINES].
@@
 
-%The goal of these existing papers is to pick points $x_i$ so that the required the number of $\varphi_i$ is minimized.
-Our goal here is to minimize the required number of matrix vector products with $\mathcal{H}$.
-
-%We do this by taking advantage of the local support property of $\mathcal{A}$.
-
-%SVIR low rank: sample columns of matrix
-
-%Pros and cons of different point picking strategies depend on the application of interest.
-
-%In astronomy image deblurring applications, the point source positions typically correspond to locations of stars in an image, which are fixed.
-
-%In [CITE], Escande and Weiss explored the idea of approximating linear operators from scattered impulse responses using a product-convolution scheme of the same form as \eqref{eq:product_convolution}.
-
-%In our previous work [CITE MY PAPER], we choose the point locations with an adaptive grid. However,
-
-%Nagy's paper: breaking domain into regular grid, using piecewise linear interpolation.
-
 Matrix-vector products with product-convolution approximations can be performed using the fast Fourier transform if the problem is discretized on a regular grid [CITE].
@@
 
-%H-matrices for Hessians: Ilonas paper with KAUST people
-
-%Demanet matrix probing can be interpreted as convolution-product.
-
-%Resolution analysis: uses matvecs with random functions to estimate spatial covariance. Was inspiration for us.
-
-%Getting many impulse responses at once (batching).
-
-%Geometry-aware weighting functions vs radial basis functions. Kriging.
-
-%\subsection{Operator access levels}
-%
-%
-%%In order to solve inverse or optimization problems governed by PDEs, we need to solve linear systems with the Hessian as the coefficient matrix [cite cite].
-In order to do this efficiently, we have access to only Hessian-vector products.
-That is, we have the ability to compute the functions $x \mapsto \mathcal{H} x$ and $x \mapsto \mathcal{H}^T x$ for arbitrary vectors $x$.
-Therefore
-%
-%A linear operator may be classified by the way in which we may access the operator. 
-%
-%
-%\begin{itemize}
-%	\item \emph{Level 1:} A matrix representation of the operator, $\mathbf{A}$, is stored in memory in dense or sparse format.
-%	\item \emph{Level 2:} We can efficiently compute arbitrary matrix entries $\mathbf{A}_{ij}$ at will, but the matrix $\mathbf{A}$ is not stored in memory.
-%	\item \emph{Level 3:} Access to $\mathbf{A}$ is only available through an algorithm that computes matrix-vector products $\mathbf{u} \mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{u} \mapsto \mathbf{A}^T \mathbf{u}$.
-%\end{itemize}
-%In large-scale scientific computing, 
-%
-%
-%As discussed in Section \ref{sec:PDE_hessian_motivation}, in inverse problems governed by PDEs, matrix entries of the Hessian are not easily computable since getting one entry would require computing an entire column of the Hessian, then throwing away all but one entry in the column.
-Computing a single column costs at least two PDE solves.
-Many such entries are needed to approximate the Hessian in inverse problems with highly informative data (i.e., for Hessians with large numerical rank) [cite cite].
-%
-% of the implicit definition of $u(m)$. 
-%
-%Hierarchical matrices (H, H2, HSS, HBS, HODLR), quantized tensor trains, hierarchical interpolative factorization, STRUMPACK (LBNL), 
-%
-%Level 1: not relevant in large-scale problems
-%State clearly: we do not use Level 1 at all.
-%This paper: Level 3 -> Level 2
-%H-matrix: solve on level 2
-%
-%%\subsection{Existing work}
-%%
-%%
-%%\begin{itemize}
-%%	\item CITE ESCANDE, others, for a reviews of the subject.
-%%	\item Resolution analysis seismic papers: try to figure out size of $\varphi_x$ by matvec with random function, post-processing to get autocorrelation length.
-They did the constant matvec volume thing.
-Their goal is to get a sense of local length scales that can be resolved in the inverse problem.
-They do not build preconditioner
-%%	\item Escande Weiss scattered response paper.
-They interpolate scattered impulse responses using radial basis functions and study how the geometry of the points affects the approximation accuracy.
-Matrix-free / getting many points at once is not a concern of them.
-The applications they consider one has direct access to the impulse response, e.g., through formula for the kernel.
-%%	\item Imaging problems: ``matvec'' requires performing an experiment
-%%	\item pseudodifferential operator equivalence
-%%	\item Laundry list of applications
-%%	\item scattered impulse response of escande weiss paper: radial basis functions for approximation in $y$. How to determine if the impulse responses are far enough away from each other?
-%%	\item my convolution paper: compute impulse responses one at a time, one impulse response per matvec: too slow! Here we compute many impulse responses using only 1 matvec.
-%%	\item uniformity of smoothness condition
-%%\end{itemize}
-
 \section{Constructing the approximation}
@@
 \item We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}).
+To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as individual functions.
 \item We use a greedy algorithm to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}).
-\item Ensuring that the points $x_i$ are well separated from each other requires a-priori estimates of the supports of the functions $\langle\mathcal{A},\delta_{x}\rangle^*$ (Section \ref{eq:mean_and_covariance_estimation}).
-\item The weighting functions are defined as the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
+\item We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\langle\mathcal{A},\delta_{x}\rangle^*$ (Section \ref{eq:mean_and_covariance_estimation}).
+\item For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
 \end{itemize}
@@
 	\SetKwInOut{Output}{Output}
-	%	\Input{Function $v \mapsto \mathcal{H}^Tv$}
-
-	%	\mbox{}\\
-
-	%	\Output{Spatially varying mean $\mu:\Omega \rightarrow \mathbb{R}^d$, \\ 
-	%		Spatially varying covariance $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$}
-	%	\Begin{
 	{	
@@
 
-		%		\Return{$\mu$, $\Sigma$}
 	}
@@
 \end{algorithm2e}
-
-%The function $\varphi_x(y)$ is undefined when $x + y \notin \Omega$, which can occur even if $x \in \Omega$ and $y \in \Omega$.
-Extending $\varphi_x$ by zero makes it defined for all $y \in \mathbb{R}^d$, but this extension is only reasonable if $x$ is far enough from $\partial \Omega$ that $\varphi_x(y)$ decays to zero, as a function of $y$, before $y$ reaches $\partial \Omega$.
-We therefore choose exclude sample points $x_i$ if the ellipsoid associated with that point overlaps the boundary.
 
@@
 \end{defn}
-
-%The support of $\langle\mathcal{A},\delta_{x}\rangle^*$ is estimated to be contained within a small number of standard deviations of the mean for a Gaussian distribution with the same mean and covariance as the normalized version of $\langle\mathcal{A},\delta_{x}\rangle^*$.
-
-%Let 
-%\begin{equation*}
-%	\rho_x := \langle\mathcal{A}, \delta_x\rangle^* \big/ \alpha(x)
-%\end{equation*}
-%denote the normalized version of $\langle\mathcal{A}, \delta_x\rangle^*$, where $\alpha:\Omega \rightarrow \mathbb{R}$ is the volume function
-%\begin{equation*}
-%	\alpha(x) := \int_{\Omega} \langle\mathcal{A}, \delta_x\rangle^*(z) dz.
-%\end{equation*}
-%Note that $\rho_x$ is a probability distribution on $\Omega$.
-%
-%In Proposition \ref{thm:vol_mean_cov}, we see that $\alpha(x)$ may be computed for all $x$ simultaneously by applying $\mathcal{A}^T$ to a constant function.
-We also see that the spatially varying mean, $\mu(x)$, and covariance, $\Sigma(x)$, of $\rho_x$, defined as
-%\begin{align*}
-%	\mu(x) :=& \int_\Omega z \rho_x(z) dz \\
-%	\Sigma(x) :=& \int_\Omega (z - \mu(x))(z - \mu(x))^T \rho_x(z) dz
-%\end{align*}
-%may be computed for all $x$ simultaneously by applying $\mathcal{A}^T$ to $d$ linear functions, and $d(d+1)/2$ quadratic functions, respectively.
-In Algorithm \ref{alg:varhpi_mean_cov} we describe an algorithm for computing the volume function, mean, and covariance based on the results of Proposition \ref{thm:vol_mean_cov}.
 
@@
 	\SetKwInOut{Output}{Output}
-%	\Input{Function $v \mapsto \mathcal{H}^Tv$}
-
-%	\mbox{}\\
-
-%	\Output{Spatially varying mean $\mu:\Omega \rightarrow \mathbb{R}^d$, \\ 
-%		Spatially varying covariance $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$}
-%	\Begin{
 	{	
@@
 
-%		\Return{$\mu$, $\Sigma$}
 	}
@@
 
-% sample points $x_i$ and $x_j$ should be far enough apart from each other that the supports of $\left\langle\mathcal{A}, \delta_{x_i}\right\rangle^*$ and $\left\langle\mathcal{A}, \delta_{x_j}\right\rangle^*$ do not overlap.
-
-%Since it is too expensive compute the support of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ directly, we instead estimate that the support of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ is contained in the ellipsoid
-%\begin{equation}
-%\label{eq:support_ellipsoid}
-%	E_x := \{x' \in \Omega: (x' - \mu(x))^T \Sigma(x)^{-1} (x' - \mu(x)) \le \tau^2\},
-%\end{equation}
-%where $\tau$ is a small number of standard deviations; we use $\tau=3$.
-The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations of the mean for the Gaussian distribution that has the same mean and covariance as $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$.
-The fraction of the mass of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though we find this bound is conservative and far less mass resides in this region in practice.
-
-%We call $E_x$ the \emph{unit ellipsoid} with mean $\mu(x)$ and covariance $\Sigma(x)/\tau^2$.
-
 Starting from a finite set of candidate points $P$, we select batches of sample points, $S_b$, using a greedy algorithm.
@@
 Then we select a candidate point $p \in P$ that is the farthest away from all points in previous sample point batches.
-That is, $p$ is the maximizer of the following optimization problem:
+That is, $p$ is a maximizer of the following optimization problem:
 \begin{equation*}
@@
 
-%We select a batches of sample points, $S_b$, from a finite set of candidate points $P$ using a greedy algorithm.
-Starting from an empty set, we build $S_b$ from candidate points $p$ from $P$, one point at a time.
-The current candidate point, $p$, is the point in $P$
-
 We repeat the process to construct several batches of points $S_1, S_2, \dots$, until the number of batches exceeds a desired threshold.
@@
 
-%		\If{$\text{point\_is\_acceptable}(p, S_b)$}{
 		\If{$E_p \cap E_q \neq \{\}$ for all $q \in S_b$ and $E_p \cap \partial \Omega = \{\}$}{
@@
     \SetKwFunction{FMain}{point\_is\_acceptable}
-%\SetKwProg{Fn}{Function}{:}{}
-%\Fn{\FMain{$p$, $S_b$}}{
-%	
-%	$p\text{\_is\_acceptable} \gets \text{True}$
-%	
-%	$E_p$ is the ellipsoid associated with $p$ (see \eqref{eq:support_ellipsoid})
-%	
-%	\For{$q \in S_b$}{
-%		
-%		$E_q$ is the ellipsoid associated with $q$
-%		
-%		\If{$E_p \cap E_q \neq \{\}$}{
-%			
-%			$p\text{\_is\_acceptable} \gets \text{False}$
-%			
-%			Terminate for loop
-%		}
-%	}
-%	\textbf{return} $p\text{\_is\_acceptable}$ 
-%}
-
 	\caption{Choosing one batch of sample points, $S_b$}
@@
 Instead, we store the functions $\eta^b$, and use formula \eqref{eq:varphi_eval} to perform function evaluations $z \mapsto \varphi_i(z)$ as needed.
-%For $x_i \in X$ and $y \in \mathbb{R}^d$, we may evaluate $\varphi_{x_i}(y)$ by finding the batch $X_b$ that $x_i$ is contained in, then using \eqref{eq:varphi_eval}.
-
-%\begin{algorithm2e}
-%	\SetAlgoNoLine
-%	\SetKwInOut{Input}{Input}
-%	\SetKwInOut{Output}{Output}
-%	\SetKwProg{Fn}{Function}{}{}
-%	\Input{Point source locations $x_1, x_2, \dots, x_m$
-%	}
-%	\Output{Response $\boldsymbol{\eta}$ of $\mathbf{A}$ to Dirac comb of point sources}
-%
-%	\BlankLine
-%
-%	Construct $\boldsymbol{\xi} \in \mathbb{R}^N_{\mathbf{M}^{-1}}, \quad \boldsymbol{\xi}_i = \sum_{j=1}^m \phi_i(x_j)$, \quad $i=1,\dots,N$
-%	
-%	$\boldsymbol{\eta} \gets \mathbf{M}^{-1} \mathbf{A}^T \mathbf{M}^{-1}\boldsymbol{\xi}$
-%
-%	\caption{Computing response $\mathbf{A}$ to Dirac comb}
-%	\label{alg:dirac_comb_response}
-%\end{algorithm2e}
 
@@
 \end{algorithm2e}
-
-%	
-%	The standard weak form for $\psi_j$ in \eqref{eq:psi_eq} is given by the following saddle point system for $\psi_j \in H^1(\Omega)$ and $\rho \in \mathbb{R}$:
-%	\begin{alignat}{5}
-%		-\int_\Omega \nabla \psi_j \cdot \nabla h ~dx &~+~& \rho \int_\Omega h ~dx &~=~ h(x_j) &\quad \forall h \in H^1(\Omega) \label{eq:lem_delta_eval_feq} \\
-%		\int_\Omega \psi_j ~dx & & &~=~ 0. & \nonumber
-%	\end{alignat}
-%	This follows from integration by parts, using the Neumann zero boundary conditions for $\psi_j$, and enforcing the average value zero constraint for $\psi_j$ with the Lagrange multiplier $\rho$.
-If $h$ is the constant function $h(x)=1$, the first equation becomes
-%	\begin{equation*}
-%		\rho \int_\Omega 1 ~ dx = 1,
-%	\end{equation*}
-%	which implies $\rho = \frac{1}{|\Omega|}$.
-Now suppose that $h\in H^2_N(\Omega)$.
-Of course, we still have $h\in H^1(\Omega)$ as well, so equation \eqref{eq:lem_delta_eval_feq} still holds.
-But now $h$ has additional regularity, which allows us to use integration by parts on the first term in \eqref{eq:lem_delta_eval_feq} to bring all of the derivatives onto $h$.
-Furthermore, $h$ has Neumann zero boundary conditions, so the boundary term that arises during integration by parts is zero.
-Performing this integration by parts, and substituting in the value for $\rho$, yields
-%	\begin{equation*}
-%		\int_\Omega \psi_j \Delta h ~dx + \frac{1}{|\Omega|} \int_\Omega h ~dx = h(x_j),
-%	\end{equation*}
-%	which is the desired result.
 
@@
 \end{thm}
-
-%Since the Laplacian measures local deviation from flatness [cite Jacobsen], from \eqref{eq:uy_optimization_problem} we see that the interpolant $x \mapsto u_y(x) = \widetilde{\varphi}_x(y)$ is the ``flattest possible'' function (in a least-squares sense) that has Neumann boundary conditions and interpolates the data $\varphi_{x_i}(y)$.
-
-% For example,
-% \begin{itemize}
-%     \item In inverse problems where the forward model is time-dependent, computing Hessian matrix-vector products requires solving time-dependent PDEs.
-In contrast, the Poisson equation is a stationary PDE.
-%     \item In inverse problems where the parameter being inverted for resides on a boundary, computing Hessian matrix-vector products requires solving PDEs that are defined on a higher dimensional geometric domain than the Poisson PDEs that must be solved for the interpolation.
-%     \item In multi-experiment inverse problems, data from many different ``experiments'' on the same medium are used in the inversion.
-Performing a Hessian matrix-vector product requires solving two PDEs \emph{per experiment}.
-The number of Poisson solves used to construct the weighting functions does not depend on the number of experiments.
-% \end{itemize}
-% The additional cost of Poisson interpolation is, therefore, a worthwhile investment if it allows one to achieve the same level of accuracy as other interpolation methods but with less sample points.
-
-% One may also show that the weighting functions $w_i$ sum to one pointwise.
 
@@
 This is shown in Algorithm \ref{alg:eval_y_x}.
-%
-%\begin{align*}
-%	\widetilde{A}(y',x') &= \sum_{i=1}^q w_i(x') \varphi_{x_i}(y'-x') \\
-%	&= \sum_{i=1}^q w_i(x') \begin{cases}
-%		\eta^{b(i)}(\zeta_i), & \zeta_i \in E_{x_i} \\
-%		0, & \text{otherwise}
-%	\end{cases} \\
-%	&= \sum_{i=1}^q w_i(x')\Psi_{ik},
-%\end{align*}
-%where $\omega_{ik}:= w_i(x_k') = \left(\mathcal{U}_X^+ \mathcal{U}\right)_{ik}$, and 
-%\begin{equation*}
-%	\Psi_{ik} := \begin{cases}
-%		\eta^{b(i)}(\zeta_{ik}), & \zeta_{ik} \in E_{x_i}\\
-%		0, & \text{otherwise}.
-%	\end{cases}
-%\end{equation*}
-%To compute $\omega$, first we construct $\nu \in \mathbb{R}^{r \times N}$,	$\nu_{jk} := u_j(x_k)$,
-%then we set
-%\begin{equation*}
-%	\omega := \mathcal{U}_X^+ \nu.
-%\end{equation*}
-%To compute $\Psi$, we must evaluate $\eta^{b(i)}(\zeta_{ik})$ for many points $\zeta_{ik}$.
-%If the mesh used to discretize the problem is non-uniform, generally $\zeta_{ik}$ will not be a degree of freedom location, even if the points $x_k'$, $y_k'$, and $x_i$ are degree of freedom locations.
-Thus, evaluating $\eta^{b(i)}(\zeta_{ik})$ requires interpolation.
-Since it is typically more efficient to perform interpolation at many points all at once, rather than interpolating points one at a time, we precompute the set of indices
-%\begin{equation*}
-%	\chi^b := \{(i,k): \zeta_{ik} \in E_{x_i},~x_i \in X_b,~ k=1,\dots,N\}
-%\end{equation*}
-%for which the interpolation $\eta^{b(i)}(\zeta_{ik})$ must be performed. Then we compute the batch of interpolations $\eta^b\left(\zeta_{\chi^b}\right)$, and set 
-%\begin{equation*}
-%	\Psi_{\chi^b} = \eta^b\left(\zeta_{\chi^b}\right).
-%\end{equation*}
-%Here $\Psi_{\chi^b}$ is the subset of entries of $\Psi$ corresponding to indices $(i,k) \in \chi^b$, and $\eta^b\left(\zeta_{\chi^b}\right)$ is the interpolation of $\eta^b$ at all of the points $\zeta_{ik}$ corresponding to indices $(i,k) \in \chi^b$.
-This process is repeated for each batch $b$.
-All other entries of $\Psi$ are set to zero.
-The resulting procedure for evaluating $\widetilde{H}(y_k',x_k')$ is summarized in Algorithm \ref{alg:H_entries}.
-
-%, for each batch $X_b$ we predetermine the subset of points 
-%\begin{equation*}
-%	p_{ik} := y_k - x_k + x_i
-%	\end{equation*}
-%satisfying $p_{bik} \in E_{x_i}$ for all $x_i \in X_b$ then evaluate $\eta_b(p_{bik})$ for this subsets of points all at once. Let
-%\begin{equation*}
-%	\chi_b:=\{p_{bik}:p_{bik} \in E_{x_i}\}.
-%\end{equation*}
-%and define $\Psi \in \mathbb{R}^{q \times N}$, 
-%
-%so that
-%\begin{equation*}
-%	\widetilde{H}(y_k,x_k) = \sum_{i=1}^q w_i(x_k) \Psi_{ik}.	
-%\end{equation*}
-%We have 
-%\begin{equation*}
-%	\Psi_{ik}(\chi_b) = \sum_{b=1}^B \eta_b(p_{bik})
-%\end{equation*}
-%where $\eta_b(\chi)$ is the 
-%
-%In Algorithm asdf, we describe how to evaluate $\widetilde{H}$ at many points using this procedure.
-%
+
 \begin{algorithm2e}
@@
 \end{algorithm2e}
-
-%\begin{algorithm2e}
-%	\SetAlgoNoLine
-%	\SetKwInOut{Input}{Input}
-%	\SetKwInOut{Output}{Output}
-%	\SetKwProg{Fn}{Function}{}{}
-%	\Input{Pairs of points $\{(x_k',y_k')\}_{k=1}^n \subset \Omega \times \Omega$}	
-%	\Output{Approximate kernel entries $\mathbf{z} \in \mathbb{R}^n$, $\mathbf{z}_k = \widetilde{H}(y_k',x_k')$}
-%	
-%	Compute $\boldsymbol{\omega} \in \mathbb{R}^{r \times n}$, $\boldsymbol{\omega}_{i,k} = w_i(x_k')$
-%	
-%	\mbox{}\\
-%	
-%	\Fn{$\computekernelentries{\{x'_k\}_{k=1}^n}{\{y'_k\}_{k=1}^n}$}
-%	{
-%		$\boldsymbol{\omega} \gets \computeweightingentries{\{x'_k\}_{k=1}^n}$
-%		
-%		Construct $\boldsymbol{\zeta} \in \mathbb{R}^{d \times q \times n}$, ~$\boldsymbol{\zeta}[:,i,k] = y_k' - x_k' + \mathbf{X}\left[:,\mathtt{S}[i]\right]$
-%
-%		\BlankLine
-%		
-%		\BlankLine
-%		
-%		$\boldsymbol{\Psi}\gets \mathbf{0}_{q \times n}$
-%		
-%		\For{$b=1,\dots,N_B$}{
-%			$\mathtt{\chi}_b \gets \{\}$
-%			
-%			\For{$j \in \mathtt{S}_b$}{
-%				$i \gets$ index such that $\mathtt{S}[i] = j$
-%				
-%				$E_j \gets$ unit ellipsoid with mean $\boldsymbol{\mu}[:,j]$ and covariance $\boldsymbol{\Sigma}[:,:,j]/\tau^2$
-%				
-%				$\mathtt{\chi}_b \gets \mathtt{\chi}_b \cup \{(i,k) : k \in \{1,\dots,N\}, \boldsymbol{\zeta}[:,i,k] \in E_j\}$
-%				
-%			}
-%
-%			
-%			$\boldsymbol{\Psi}[\chi_b] \gets \interpolate\left(\boldsymbol{\eta}^b, \{\boldsymbol{\zeta}[:,i,k] : (i,k) \in \mathtt{\chi}^b\}\right)$
-%		
-%		}
-%	
-%		\BlankLine
-%		
-%		\BlankLine
-%		
-%		Construct $\mathbf{z} \in \mathbb{R}^n$, ~$\mathbf{z}_k = \sum_{i=1}^q \boldsymbol{\omega}_{ik} \boldsymbol{\Psi}_{ik}$
-%	
-%		\Return{$\mathbf{z}$}
-%	}
-%	\caption{Compute many approximate kernel entries $z_k = \widetilde{A}(y_k',x_k')$}
-%	\label{alg:H_entries}
-%\end{algorithm2e}
 
@@
 The root of the tree is the whole matrix, internal nodes in the tree are blocks that are subdivided, and the leaves of the tree are blocks that are either expected to be low rank, or are small.
-
-%Note that the cluster trees and block cluster tree are computed using only the geometric information about the degree of freedom locations, and do not depend on the values of the matrix entries of $\widetilde{\mathcal{A}}$.
 
@@
 
-\section{Interpolating impulse response samples old}
-
-We interpolate the functions $\varphi_{x_i}$ associated with the points $x_i \in S$, to generate functions $\varphi_x$ associated with new points $x \notin S$.
-This is done using a continuous generalization of the interpolative decomposition for matrices, wherein a select number of columns of a matrix are interpolated to fill the whole matrix.
-Interpolating $\varphi_x$ differs from interpolative low rank approximation of $\mathcal{A}$, which corresponds to interpolating $\langle\mathcal{A},\delta_x\rangle^*$.
-Because of the shifting that maps $\langle\mathcal{A}, \delta_x\rangle$ to $\varphi_x$, interpolation of $\varphi_{x_i}$ yields an approximation $\widetilde{\mathcal{A}}$ which can be high-rank.
-
-The interpolation procedure requires a basis of smooth functions.
-We form this basis from smooth eigenmodes of the Laplacian on $\Omega$ with Neumann boundary conditions.
-The resulting approximation $\mathcal{A}$ may be interpreted as a weighted sum of convolution operators, with smooth spatially varying weighting functions.
-Kernel entries of the approximation, $\widetilde{A}(y,x)$, may be evaluated efficiently.
-
-\subsection{Interpolative factorization of the SVIR}
-\label{sec:interp_fact_svir}
-
-The \emph{spatially varying impulse response}, denoted $T$, is the reinterpretation of $\varphi_x(y)$ as a function of both $y$ and $x$:
-\begin{equation*}
-	T(y,x) := \varphi_x(y) = A(y+x,x).
-\end{equation*}
-Since $T(y,x)$ is smooth as a function of $x$ (Condition 1), the continuous analog of the row space of $T$ should be well-approximated by the span of a small number of orthonormal smooth functions, $\{u_k\}_{k=1}^r$.
-Condition 1, therefore, is equivalent to the existence of a low rank approximation $\widetilde{T}$ of $T$ of the form
-\begin{equation}
-	\label{eq:formula_for_A}
-	T \approx \widetilde{T} = \Psi U
-\end{equation}
-where $\Psi:\mathbb{R}^d \times \{1,\dots,r\} \rightarrow \mathbb{R}$ is an unknown semi-discrete kernel with $r$ columns, and $U:\{1,\dots,r\} \times \Omega \rightarrow \mathbb{R}$ is the semi-discrete kernel with $r$ rows, such that the $i^\text{th}$ row of $U$ is the function $u_i$.
-That is,
-\begin{equation*}
-	U(i,x) = u_i(x).
-\end{equation*}
-The multiplication $K_1 K_2$ of kernels $K_1:X \times Y \rightarrow \mathbb{R}$ and $K_2:Y \times Z \rightarrow \mathbb{R}$ is defined as
-\begin{equation*}
-	\left(K_1 K_2\right)(x,z) := \sum_{y \in Y} K_1(x,y) K_2(y,z)
-\end{equation*}
-if $Y$ is a discrete set, and
-\begin{equation*}
-	\left(K_1 K_2\right)(x,z) := \int_Y K_1(x,y) K_2(y,z) dy	
-\end{equation*}
-if $Y\subset \mathbb{R}^d$ is a domain.
-
-The number of smooth functions, $r$, required to well-approximate $T$ using \eqref{eq:formula_for_A} can be small even if the numerical rank of $\mathcal{A}$ is large.
-For example, let $\varphi$ be any convolution kernel, and suppose $\mathcal{A}$ is the convolution operator $\left(\mathcal{A}v\right)^* := \varphi \ast v$.
-The rank of $\mathcal{A}$ can be made arbitrarily large or small by changing the convolution kernel $\varphi$.
-At the same time, $T=\Psi U$ holds with $r=1$, where $\Psi(x,1):= \varphi(x)$ and $U(1,x):=1$.
-
-We may determine $\Psi$ using a continuous analog of the process used to form the interpolative matrix factorization Let $S:=\{x_j\}_{j=1}^q \subset \Omega$ be the set of $q \ge r$ sample points determined by the method in Section \ref{sec:choosing_sample_points}, and define $T_S:\Omega \times \{1,\dots,q\}\rightarrow\mathbb{R}$ to be the semi-discrete kernel with $i^\text{th}$ column $\varphi_{x_i}$.
-That is,
-\begin{equation*}
-	T_S(y,i):= \varphi_{x_i}(y) = T(y,x_i).	
-\end{equation*}
-Further, let $U_S$ be the $r \times q$ matrix with $(i,j)$ entry given by $u_i(x_j)$, and assume that the points $x_i$ are chosen such that $U_S$ has full row rank.
-This is not difficult; for example, if the points $x_i$ were chosen uniformly at random then $U_S$ would have full rank almost surely.
-Indeed, with uniform random points, $\frac{|\Omega|}{q}\left(U_S U_S^T\right)_{ij}$ is a Monte-Carlo approximation of $\langle u_i, u_j \rangle_{L^2(\Omega)}$, where $|\Omega|$ is the total volume of the domain $\Omega$.
-Thus $\frac{|\Omega|}{q}\left(U_S U_S^T\right) \rightarrow I$ as $q \rightarrow \infty$, where $I$ is the identity matrix.
-This implies $\sigma_\text{min} \rightarrow \sqrt{q/|\Omega|}$ as $q \rightarrow \infty$, where $\sigma_\text{min}$ is the minimum singular value of $U_S$.
-We do not use uniform random sampling to choose the points, but the way we choose points does achieve relatively uniform coverage of the domain.
-
-Restricting the second input of both sides of the equation \eqref{eq:formula_for_A} to points in $S$, we have
-\begin{equation}
-\label{eq:A_I_equals_X_U_I}
-\widetilde{T}_S = \Psi U_S.
-\end{equation}
-Multiplying both sides of \cref{eq:A_I_equals_X_U_I} by the Moore-Penrose pseudoinverse of $U_S$, denoted $U_S^+$, and using the fact that $U_S$ has full row rank, yields
-\begin{equation}
-\label{eq:formula_for_X}
-\widetilde{T}_S U_S^+ = \Psi.
-\end{equation}
-Substituting the left hand side of \eqref{eq:formula_for_X} for $\Psi$ into \eqref{eq:formula_for_A} yields the following interpolative-type low rank factorization $\widetilde{T}$ of $T$:
-\begin{equation}
-\label{eq:interpolative_decomposition0}
-T \approx \widetilde{T} = T_S U_S^+ U.
-\end{equation}
-%We will quantify the accuracy of this approximation in Theorem \ref{thm:smooth_basis_err} in Section \ref{sec:error_estimate_adaptive_algorithm}.
-
-%Briefly, the error will be small when (1) $\mathcal{U}$ well-approximates the row space of $\mathcal{T}$, and (2) the smallest singular value of $\mathcal{U}_X$ is not small.
-
-\subsection{Smooth basis of Laplacian eigenmodes}
-
-For the basis of smooth functions $u_i$, we use the dominant eigenfunctions of the inverse Laplacian on $\Omega$ with Neumann boundary conditions.
-This basis of inverse Laplacian eigenfunctions generalizes the Fourier basis to non-rectangular domains.
-Let $\Delta_N^{-1}$ denote the solution operator for the following PDE:
-\begin{equation}
-\label{eq:laplacian_pde}
-\begin{cases}
--\Delta u =f, & \text{in } \Omega, \\
-n \cdot \nabla u = 0, & \text{on } \partial \Omega, \\
-\int_\Omega u dx = 0, &
-\end{cases}	
-\end{equation}
-where $\Delta$ is the Laplacian and $n$ is the normal to $\partial \Omega$.
-That is, $\Delta_N^{-1}$ is the map from source $f$ to PDE solution $u$.
-The first smooth basis function used for interpolation is the constant function $u_1(x):=1$.
-The remaining smooth basis functions, $u_2, u_3, \dots, u_r$, are the eigenfunctions corresponding to the $r-1$ largest eigenvalues of $\Delta_N^{-1}$.
-These eigenfunctions may be computed efficiently using randomized SVD \cite{HMT11}, Krylov methods such as Lanczos, power iterations with deflation, or other methods.
-We use randomized SVD.
-All of these methods require the ability to apply $\Delta_N^{-1}$ to given vectors, which amounts to solving PDE \eqref{eq:laplacian_pde} with the given vector as the right hand side source.
-Highly efficient multigrid schemes exist for solving this PDE.
-%; we use an algebraic multigrid implementation in PyAMG.
-
-\subsection{Action of the approximate operator}
-
-Defining
-\begin{equation*}
-	W := U_S^+ U,
-\end{equation*}
-we may write $\widetilde{T}$ as
-\begin{equation*}
-	\widetilde{T} = T_S W.	
-\end{equation*}
-Using the definition of $T_S$, we may write the entries of the resulting approximation to $A$ as
-\begin{equation}
-\label{eq:H_kernel_entries}
-	\widetilde{A}(y,x) = \widetilde{T}(y-x,x) = \sum_{i=1}^q w_i(x) \varphi_{x_i}(y-x) 
-\end{equation}
-where the functions $w_i(x):=W(i,x)$ are the rows of $W$, which are linear combinations of the functions $u_i$.
-%Since the Laplacian eigenfunctions $u_i$ are defined on $\Omega$, the weighting functions $w_i$ are defined on all of $\Omega$, so $T$ and $H$ are defined on $\mathbb{R}^d \times \Omega$.
-However, the error estimate we present in the next section will only be valid for these operators on $\mathbb{R}^d \times \Omega_I$.
-From \eqref{eq:H_kernel_entries}, we derive the action of $\widetilde{\mathcal{A}}$ as
-\begin{align*}
-	\left(\widetilde{\mathcal{A}}u\right)(v) &= \int_\Omega \int_\Omega v(y) \widetilde{A}(y,x)u(x) dx dy \\
-	&= \int_\Omega \int_\Omega v(y) \widetilde{T}(y-x,x)u(x) dx dy \\
-	&= \int_\Omega \int_\Omega v(y) \left(\sum_{i=1}^q  \varphi_{x_i}(y-x) w_i(x)\right) u(x) dx dy \\
-	&= \int_\Omega v(y) \sum_{i=1}^q \int_\Omega \varphi_{x_i}(y-x) w_i(x) u(x) dx dy\\
-	&= \int_\Omega v(y) \left(\sum_{i=1}^q \varphi_{x_i} \ast \left(w_i \cdot u\right)\right)(y) dy,
-\end{align*}
-which implies
-\begin{equation*}
-	\left(\widetilde{\mathcal{A}} u	\right)^* = \sum_{i=1}^q \varphi_{x_i} \ast \left(w_i \cdot u\right)
-\end{equation*}
-Here $f \ast g$ denotes the convolution of the functions $f$ and $g$. Thus, $\widetilde{\mathcal{A}}$ is a weighted sum of convolution operators with smooth spatially varying weighting functions.
-
-\section{Adaptive algorithm}
-\label{sec:error_estimate_adaptive_algorithm}
-
-In this section we bound the error in the approximation to $T$ (Theorem \ref{thm:smooth_basis_err}), and use this bound to drive an adaptive algorithm (Algorithm \ref{alg:adaptive}) for constructing the approximation.
-
-\begin{thm}
-	\label{thm:smooth_basis_err}
-	Let $\sigma_\text{min}$ be the smallest singular value of $U_S$, and let
-	\begin{equation*}
-		\gamma(q) := \sigma_\text{min} \sqrt{\frac{q}{|\Omega|}},
-	\end{equation*}
-	where $|\Omega|$ is the measure of $\Omega$. We have
-	\begin{equation*}
-		\|T - T_S U_S^+ U\|_2 \le \left(1 + \frac{|\Omega|}{\gamma(q)} \right)\|T - TU^TU\|_2,
-	\end{equation*}
-	where $\sigma_\text{min}$ is the smallest singular value of $U_S$, and $\nor{\mathcal{A}}$ is the induced norm for the operator $\mathcal{A}$.
-\end{thm}
-
-\begin{proof}
-	Let $\widehat{T} := TU^TU$, and let $\widehat{T}_S:\mathbb{R}^d \times \{1,\dots,q\} \rightarrow \mathbb{R}$ be the semi-discrete kernel defined by
-	\begin{equation*}
-		\widehat{T}_S(y,i) = \widehat{T}_S(y,x_i).
-	\end{equation*}
-	Using the triangle inequality yields
-	\begin{equation}
-		\nor{T - T_S U_S^+ U}_2 \le \nor{T - \widehat{T}_S U_S^+ U}_2 
-		+ \nor{\left(T_S - \widehat{T}_S\right) U_S^+ U}_2 \label{eq:two_bound_align}
-	\end{equation}
-	For the first term, note that $\widehat{T}_S = T U^T U_S$, so 
-	\begin{equation*}
-		\widehat{T}_S U_S^+ U = T U^T U_S U_S^+ U = T U^T U = \widehat{T}.
-	\end{equation*}
-	Thus,
-	\begin{equation}
-	\label{eq:A_nogamma}
-	\nor{T - \widehat{T}_S U_S^+ U }_2 = \nor{T - \widehat{T}}_2.
-	\end{equation}
-	We estimate the second term as:
-	\begin{equation*}
-		\nor{\left(T_S - \widehat{T}_S\right) U_S^+ U}_2 \le \nor{T_S - \widehat{T}_S}_2 \nor{U_S^+}_2 \nor{U}_2 = \frac{1}{\sigma_\text{min}} \nor{T_S - \widehat{T}_S}_2	
-	\end{equation*}	
-%	\begin{align*}
-%		\nor{\left(T_S - \widehat{T}_S\right) U_S^+ U}_2 &\le \nor{T_S - \widehat{T}_S}_2 \nor{U_S^+ U}_2 \\
-%		&\le \sqrt{q}\nor{T_S - \widehat{T}_S}_1 \nor{U_S^+ U}_2 \\
-%		&\le \sqrt{q}\nor{T_S - \widehat{T}_S}_1 \nor{U_S^+}_2 \nor{U}_2 = \frac{\sqrt{q}}{\sigma_\text{min}} \nor{T_S - \widehat{T}_S}_1. 
-%	\end{align*}
-	where we used the facts that $\nor{U_S^+}_2 = 1/\sigma_\text{min}$ and $\nor{U}_2=1$. By H\"older's inequality, we have 
-	\begin{equation*}
-		\nor{T - \widehat{T}}_{2} \le |\Omega|^{1/2} \nor{T - \widehat{T}}_{2,1}.
-	\end{equation*}
-
-	To estimate this term further, let $\delta_{x_i}^\epsilon \in L^1(\Omega)$ be an $\epsilon$-approximation to $\delta_{x_i}$ with norm $\|\delta_{x_i}^\epsilon\|_1=1$, defined as
-	\begin{equation*}
-		\delta_{x_i}^\epsilon(x) := \begin{cases}
-			1 / \vol(B_\epsilon(x_i)), & x \in B_\epsilon(x_i), \\
-			0, & \text{otherwise},
-		\end{cases}
-	\end{equation*}
-	where $B_\epsilon(x_i)$ is the ball of radius $\epsilon$ centered at $x_i$, and $\vol(B_\epsilon(x_i))$ is the volume of this ball.
-	We have
-	\begin{align*}
-		\nor{T_S - \widehat{T}_S}_1 &= \max_{i \in 1,2,\dots,q} \|T(~\cdot~,x_i) - \widehat{T}(~\cdot~,x_i)\|_1\\
-		&= \|T(~\cdot~,x_k) - \widehat{T}(~\cdot~,x_k)\|_1 \\
-		&= \lim_{\epsilon \rightarrow 0} \nor{(T-\widehat{T}) \delta_{x_k}^\epsilon}_1 \\
-		&\le \sup_{\|v\|_1=1} \nor{(T - \widehat{T}) v}_1 = \nor{T - \widehat{T}}_1,
-	\end{align*}
-	where $k$ is the index achieving the maximum in the first line. Thus
-	\begin{equation}
-	\label{eq:_gamma_A_bound}
-	\nor{\left(T_S - \widehat{T}_S\right) U_S^+ U}_2 \le \frac{\sqrt{q}}{\sigma_\text{min}}\nor{T - \widehat{T}}_1.
-	\end{equation}
-	The proposition follows from substituting bounds \eqref{eq:A_nogamma} and \eqref{eq:_gamma_A_bound} into \eqref{eq:two_bound_align}.
-\end{proof}
-
-Theorem \ref{thm:smooth_basis_err} shows that the error in our approximation depends on (1) how well the row space of $\mathcal{U}$ captures the row space of $\mathcal{T}$ (smaller $\|\mathcal{T} - \mathcal{T}\mathcal{U}^T\mathcal{U}\|$ $\implies$ less error), and (2) the minimum singular value of $\mathcal{U}_X$ (larger $\sigma_\text{min}$ $\implies$ less error).
-We can decrease $\|\mathcal{T} - \mathcal{T}\mathcal{U}^T\mathcal{U}\|$ by including more functions $u_i$ in $\mathcal{U}$.
-We can increase $\sigma_\text{min}$ by incorporating more batches of sample points $X_b$ in the approximation.
-But if $\sigma_\text{min}>1$, there are diminishing returns to making $\sigma_\text{min}$ bigger.
-By analogy to constrained optimization, $\|\mathcal{T} - \mathcal{T}\mathcal{U}^T\mathcal{U}\|$ is like an objective function to be minimized, while ensuring $\sigma_\text{min}$ is not small is like a constraint to be satisfied.
-We therefore increase the accuracy of the approximation by including more functions in $\mathcal{U}$, adding more batches of sample points as needed to ensure $\sigma_\text{min} > 1$.
-
-The error in the approximation for batches $1,\dots,b$ is estimated with the the error for the next batch
-\begin{equation*}
-	\mathcal{E} := \frac{\nor{\eta^{b+1} - \widetilde{\eta}^{b+1}}}{\nor{\eta^{b+1}}},
-\end{equation*}
-where $\eta^{b+1}$ and $\widetilde{\eta}^{b+1}$ are the responses of $\mathcal{H}$ and $\widetilde{\mathcal{H}}$, respectively, to the Dirac comb at the next batch $\mathtt{S}_{b+1}$ of sample points, that were not used in the approximation $\widetilde{\mathcal{H}}$ yet.
-Entries of $\widetilde{\eta}_{b+1}$ take the form
-\begin{equation}
-\label{eq:eta_test}
-	\widetilde{\eta}^{b+1}(y_k) = \sum_{x_j \in S_{b+1}} \widetilde{H}(y_k,x_j),
-\end{equation}
-which we compute using Algorithm \ref{alg:H_entries} for all degree of freedom locations $y_k$ for the discretization of the problem.
-If $\mathcal{E}$ is larger than a desired tolerance, we compute more eigenmodes of $\Delta_N^{-1}$ and add those modes to the basis $\mathcal{U}$.
-Adding more functions to $\mathcal{U}$ may change $\sigma_\text{min}$.
-If $\sigma_\text{min}<1$ after this change, we select new batches of sample points and incorporate them in the approximation until $\sigma_\text{min} \ge 1$ once again.
-This process repeats until $\mathcal{E}$ is less than the desired tolerance.
-Each time an additional batch is needed, we compute another batch for testing, and incorporate the batch that was previously used for testing into the approximation.
-Once the desired accuracy is reached, we incorporate the last batch that was used for testing into the approximation to increase the accuracy further.
-
-\begin{algorithm2e}
-	\SetAlgoNoLine
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Output}
-%	\Input{Function $v \mapsto \mathcal{H}v$\\
-%		Function $v \mapsto \mathcal{H}^Tv$
-%	}
-
-%	\mbox{}\\
-
-%	\Output{Responses $\eta^1, \dots, \eta^b, \eta^\text{test}$ of $\mathcal{H}$ to Dirac combs \\
-%		Basis of smooth functions $u_1, \dots, u_r$ \\
-%		Interpolation pseudoinverse $\mathcal{U}_X^+$}
-	%	\Begin{
-	{
-
-		Determine $\Omega_I$ with Algorithm \ref{alg:boundary_region}
-
-		Compute $\mu$ and $\Sigma$ with Algorithm \ref{alg:varhpi_mean_cov}
-
-		Select first sample point batch $S_1$ (Section \ref{sec:choosing_sample_points})
-
-		Compute $\eta^1$ (Section \ref{sec:get_impulse_response})
-
-		Set $S = \{\}$ and $b=0$
-
-%		Form $\mathbf{X} \in \mathbb{R}^{d \times N}$ as the array of all degree of freedom coordinates
-%		
-%		$\mathtt{C} \gets \mathtt{I}$
-%		
-%		$b=0$
-%		
-%		$\mathtt{S}_1 \gets \choosebatch{\mathtt{C}}$ \tcp*{Algorithm \ref{alg:point_choice}}
-%		
-%		$\mathtt{C} \gets \mathtt{C} \setminus \mathtt{S}_1$
-%		
-%		$\boldsymbol{\eta}_1 \gets \computediracresponse{\mathtt{S}_1}$ \tcp*{Algorithm \ref{alg:dirac_comb_response}}
-%		
-%		$\mathtt{S} \gets \{\}$
-
-		\For{$k=1,2,\dots,$}{
-
-			Compute $u_k$ as the $k^\text{th}$ eigenfunction of $\Delta_N^{-1}$
-
-%			\tcp{$U_S$ is the matrix with entries $\left(U_S\right)_{ij} = u_i(x_j)$ for $x_j\in S$}
-
-%			\If{$U_S$ has less columns than rows}{
-%				Select more sample point batches until $U_S$ has more columns than rows
-%				
-%				Compute functions $\eta^b$ for the new sample point batches
-%			}
-
-			\While{$|S| < k$}{
-
-				Set $b=b+1$
-
-				Set $S=S\cup S_b$
-
-				Select new sample point batch $S_{b+1}$
-
-				Form $\eta^{b+1}$
-
-%				Form Dirac comb $\xi^b$ associated with $S_b$ and compute $\eta^b=\mathcal{A}\xi^b$ 
-%				
-%				Add new columns to $U_S$ corresponding to points $x_j \in S_b$
-%				
-%				Add new columns to $U_S$ corresponding to points $x_j \in S_b$
-%				
-%				$b \gets b+1$
-%			
-%				$\mathtt{S} \gets \mathtt{S} \cup \mathtt{S}_b$
-%				
-%				$\mathtt{S}_{b+1} \gets \choosebatch{\mathtt{C}}$
-%				
-%				$\mathtt{C} \gets \mathtt{C} \setminus \mathtt{S}_{b+1}$
-%				
-%				$\boldsymbol{\eta}_{b+1} \gets \computediracresponse{\mathtt{S}_{b+1}}$
-
-			}
-
-			Compute $\widetilde{\eta}^{b+1}$ using formulas \eqref{eq:eta_test} and \eqref{eq:kernel_entry_formua}.
-
-			\If{$\displaystyle\frac{\nor{\eta^{b+1} - \widetilde{\eta}^{b+1}}}{\nor{\eta^{b+1}}}$ is sufficiently small}{
-
-				Terminate for loop
-
-			}
-
-		}
-
-		Set $S = S \cup S_{b+1}$
-
-		Add new columns to $U_S$ corresponding to points $x_j \in S_b$
-
-%		$\mathtt{Q} \gets \mathtt{Q} \cup \mathtt{Q}_\text{test}$
-
-%		\Return{$\mu, \Sigma$, $\eta^1, \dots, \eta^b, \eta^\text{test}$, $u_1, \dots, u_r$, $\mathcal{U}_X^+$}
-
-	}
-	\caption{Adaptive approximation of $\mathcal{H}$}
-	\label{alg:adaptive}
-\end{algorithm2e}
-
-\section{Hierarchical matrix construction}
-
 \section{Numerical results}
+\label{sec:numerical_results}
+
+Some numerical results here.
 
@@
 
-%\begin{proof}[Proof of Proposition \ref{prop:boundary_source}]
-%	Let $w \in L^2(\Omega)$ be arbitrary. We have
-%	\begin{align*}
-%		\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle(w) =& \int_\Omega w(y) \int_{\partial \Omega} A(y,x) dy dx \\
-%		=& \int_\Omega w(x) \int_{\partial \Omega} \varphi_x(y-x) dy dx = \left(w, z\right)_{L^2(\Omega)},
-%	\end{align*}
-%	where
-%	\begin{equation*}
-%		z(x) := \int_{\partial \Omega} \varphi_x(y-x) dy,	
-%	\end{equation*}
-%	which implies
-%	\begin{equation*}
-%		\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle^*(x) = z = \int_{\partial \Omega} \varphi_x(y-x) dy.
-%	\end{equation*}
-%	Since $\varphi_x$ is a non-negative function, the result follows from the fact that the integral of a non-negative function over a set of positive measure is zero if and only if the function is zero almost everywhere on that set.
-%\end{proof}
-
 \begin{proof}[Proof of Theorem \ref{thm:vol_mean_cov}]
@@
 	\end{equation*}
-%	\item[Implicitly defined matrix representation:]
-%	Let $\mathcal{A}_h:V_h \rightarrow V_h'$ be a finite dimensional approximation of $\mathcal{A}:L^2(\Omega) \rightarrow L^2(\Omega)'$, and let $\mathbf{A}$ be the matrix representation of $\mathcal{A}_h$.
-The matrix $\mathbf{A}$ is implicitly defined, in the sense that we have a algorithm to compute the matrix-vector products $\mathbf{u} \mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{A}^T \mathbf{u}$ for arbitrary $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, but we do not have direct access to matrix entries $\mathbf{A}_{ij}$.
-In principle, all matrix entries of $\mathbf{A}$ could be computed by applying $\mathbf{A}$ to each column of the identity matrix, but this is far too expensive to be practical.
+
 	\item[Distributions:] Let $\mu:C\left(\overline{\Omega}\right)\rightarrow \mathbb{R}$ be a distribution.
@@
 
-% $\left(\mathcal{A}u\right)(v) = \left\langle v, \varphi_p \ast u\right\rangle_{L^2(\Omega)}$. Here $f \ast g$ denotes convolution of the functions $f$ and $g$, and the convolution kernel,
-
-%Access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$.
-%Integral kernel representation \eqref{eq:kernel_representation} is assumed to exist,
-%e.g., by the Schwarz kernel theorem,
-%but o
-%One cannot easily access entries $A(y,x)$ of the kernel.
-
-% is non-negative in the sense that  $A(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$. 
-%Let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be the corresponding linear operator generated by currying.
-That is, for each $u\in L^2(\Omega)$, the linear functional $\mathcal{A}u:L^2(\Omega)\rightarrow \mathbb{R}$ is
-%We approximate the operator $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ defined by
-%\begin{equation*}
-%	(\mathcal{A}u)(v) := a(v,u).
-%\end{equation*}
-%We seek to approximate operators $\mathcal{A}$ in the ``matrix free'' setting, in which access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$.
-Here ($\mathcal{A}^Tv)(u) := (\mathcal{A}u)(v)$.
-%Integral kernel representation \eqref{eq:kernel_representation} is assumed to exist,
-%e.g., by the Schwarz kernel theorem,
-%but one cannot easily access entries $A(y,x)$ of the kernel. We assume that the impulse response of $\mathcal{A}$ is locally supported in the sense that the function 
-%$y \mapsto A(y, x)$ is supported in a region near $y=x$, and we assume that $\mathcal{A}$ is locally translation invariant in the sense that
-%\begin{equation*}
-%A(y+h, x+h) \approx A(y,x),
-%\end{equation*}
-%so long as $h$ is not too large.
-
 The \emph{point spread function} (PSF) of $\mathcal{A}$ at $x\in \Omega$ is
@@
 Recall the Riesz representation of a linear functional $\psi \in L^2(\Omega)'$ is the unique vector $\psi^* \in L^2(\Omega)$ satisfying $\left(\psi^*, v\right)_{L^2(\Omega)} = \psi(v)$ for all $v \in \L^2(\Omega)$.
-
-%In this paper we present an efficient matrix-free method for approximating operators $\mathcal{A}$ with a product-convolution approximation, $\widetilde{A}$, of the form
-%\begin{equation*}
-%	\left(\mathcal{A}u\right) \approx \left(\widetilde{\mathcal{A}}u\right) := \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right)
-%\end{equation*}
-%
-%
-%PSFs that satisfy the following properties:
-%\begin{enumerate}
-%	\item $\varphi_x$ varies smoothly as a function of $x$.
-%	\item $\varphi_x(y)$ is zero (or small) when $|y|$ is large.
-%	\item $\varphi_x$ is a non-negative (or mostly non-negative) function.
-%\end{enumerate}
-%If $\varphi_x$ was constant in $x$, then $\mathcal{A}$ would be the convolution operator $\left(\mathcal{A}v\right)^* = \varphi_0 \ast v$.
-Operators satisfying property 1 therefore act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies smoothly as $x$ changes.
-Operators satisfying property 2 are local, in the sense that sources at point $x$ have no impact on points far away from $x$.
-Operators satisfying property 3 blur inputs: their action on a point source consists of scaling the source, then spreading the energy of the scaled point source across a region nonuniformly.
-We call operators satisfying these properties \emph{smoothly varying blur operators}.
-
-% Computing the functions $w_i$ requires solving two Poisson PDEs for each point $x_i$, and these Poisson PDEs may be solved cheaply with multigrid.
-
-%Specifically, since the integral kernel for $\mathcal{A}$ is non-negative, $\langle\mathcal{A},\delta_x\rangle^*$ may be interpreted as a scaled probability distribution.
-We compute the mean and covariance of this probability distribution for all points $x \in \Omega$ simultaneously, by applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions.
-We then estimate the support of $\langle\mathcal{A},\delta_x\rangle^*$ to be contained within a small number of standard deviations of the mean of this probability distribution.
-
-%We interpret $\langle\mathcal{A},\delta_{x_i}\rangle^*$ as a scaled probability distribution, and estimate the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ to be contained within a small number of standard deviations of the mean of this probability distribution (this is why we require a non-negative kernel).
-The mean and covariance of $\langle\mathcal{A},\delta_{x}\rangle^*$ can be simultaneously computed for all points $x \in \Omega$ by applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions.
-
-%The functions $w_i$ are the defined as the smoothest functions, in a certain least-squares sense, that interpolate the points $x_i$.
-Computing the functions $w_i$ requires solving two Poisson PDEs for each point $x_i$, and these Poisson PDEs may be solved cheaply with multigrid.
-
-%The idea of our method is to sample $\varphi_{x}$ at a collection of points $x_i$ by applying $\mathcal{A}$ to a sum of point sources at the points $x_i$ (a ``Dirac comb'').
-The points $x_i$ are chosen so that the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ and the support of $\langle\mathcal{A},\delta_{x_j}\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$.
-Since the point sources do not overlap, by applying $\mathcal{A}$ to one vector, we recover $\varphi_{x_i}$ for many points $x_i$.
-The process is repeated to get $\varphi_{x_i}$ at more points, until an error estimate is satisfied.
-The $\varphi_{x_i}$ are interpolated smoothly to construct $\varphi_x$ for arbitrary $x$, which yields an approximation $\widetilde{A}(y,x)$ to the kernel $A(y,x)$, and an approximation $\widetilde{\mathcal{A}}$ to the operator $\mathcal{A}$.
-While $A(y,x)$ is inaccessible, we can evaluate $\widetilde{A}(y,x)$ at arbitrary $(y,x)$ points.
-This allows us to construct a hierarchical matrix representation of $\widetilde{\mathcal{A}}$, which allows us to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations, using fast hierarchical matrix arithmetic.
-
-%Properties 1 and 2 are essential to our method.
-Property 3 is only required in the subprocedure for estimating the support of $\langle\mathcal{A},\delta_x\rangle^*$, and may be dropped if other methods for estimating the support are available.
 
@@
 
-\subsection{Determination of the boundary region}
-\label{sec:boundary}
-
-Define $\Omega_B \subset \Omega$ to be the set of points that are too close to $\partial \Omega$, in the sense that $x \in \Omega_B$ if $\varphi_x(y-x) \neq 0$ for some $y \in \partial \Omega$.
-Further, define $\Omega_I := \Omega \setminus \Omega_B$ as the set of points that are in the interior of $\Omega$ and are not too close to $\partial B$.
-In Proposition \ref{prop:boundary_source} we show that $\Omega_B$ can be determined by applying $\mathcal{A}^T$ to a boundary source distribution and post-processing the result.
-The resulting algorithm for determining $\Omega_B$ and $\Omega_I$ in the discretized setting is shown in Algorithm \ref{alg:boundary_region}.
-
-\begin{prop}
-	\label{prop:boundary_source}
-	Let $\delta_{\partial \Omega}: C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ be the boundary source distribution defined by
-	\begin{equation}
-	\label{eq:boundary_source}
-	\delta_{\partial \Omega}(v) = \int_{\partial \Omega} v(y) dy
-	\end{equation}
-	for all $v \in C(\overline{\Omega})$. We have $\varphi_x(y-x) = 0$ for almost every $y \in \partial \Omega$ if and only if $\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle^*(x) = 0$.
-\end{prop}
-
-The proof of Proposition \ref{prop:boundary_source} is shown in Appendix \ref{app:proofs}.
-
-%Straightforward analysis shows that $\varphi_x(y-x) = 0$ for $y \in \partial \Omega$ almost everywhere if and only if $\left(\mathcal{A}^T \delta_{\partial \Omega}\right)(x) = 0$, where $\delta_{\partial \Omega}$ is the uniform boundary source distribution.
-That is, the distribution that acts on a test function $v$ by performing a surface integral of $v$ over the boundary of the domain:
-%%\begin{equation}
-%%\label{eq:boundary_source}
-%%	\delta_{\partial \Omega}(v) = \int_{\partial \Omega} v(y) dy.
-%%\end{equation}
-%We therefore compute $b  = \mathcal{A}^T \delta_{\partial \Omega}$, then identify $\Omega_B$ as the set of points $x \in \Omega$ for which $|b(x)|$ is larger than a small tolerance $\epsilon$.
-This is described in Algorithm \ref{alg:boundary_region}.
-
-\begin{algorithm2e}
-	\SetAlgoNoLine
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Output}
-	{
-
-		Construct boundary source $\boldsymbol{\delta}_{\partial \Omega} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$, $\displaystyle \left(\boldsymbol{\delta}_{\partial \Omega}\right)_i = \int_{\partial \Omega} \phi_i(s) ds$, $i=1,\dots,N$
-
-		%		\For{$i=1,\dots,N$}{
-		%			$\displaystyle \mathbf{b}_i \gets \int_{\partial \Omega} \phi_i(s) ds$
-		%		}
-
-		$\mathbf{b} = \mathbf{M}^{-1}\mathbf{A}^T\mathbf{M}^{-1}\boldsymbol{\delta}_{\partial \Omega}$
-
-		$\Omega_B = \{x \in \Omega : |b_h(x)| \ge \epsilon\}$
-
-		$\Omega_I = \Omega \setminus \Omega_B$
-
-	}
-	\caption{Compute boundary region $\Omega_B$ and interior region $\Omega_I$}
-	\label{alg:boundary_region}
-\end{algorithm2e}
-
-One should only expect $\widetilde{A}(y,x)$ to be accurate for $x \in \Omega_I$, $y \in \Omega$.
-If $\mathcal{A}$ is symmetric then one may overwrite potentially inaccurate entries $\widetilde{A}(y,x)$, $x \in \Omega_b$, $y \in \Omega_I$, with the accurate entries $\widetilde{A}(x,y)$.
-If $\mathcal{A}$ is nonsymmetric, one may form an accurate approximation of these entries by applying our procedure to $\mathcal{A}^T$.
-Our method will furnish an approximation $\widetilde{A}(y,x)$ for $x \in \Omega_B$, $y \in \Omega_B$ (the block of $\widetilde{H}$ corresponding to boundary-boundary interactions), but this part of the approximation is unlikely to satisfy Condition 1 ($\varphi_x$ varies smoothly as a function of $x$).
-
 \subsection{Computation of Poisson weighting functions}
 \label{sec:computation_of_weighting_functions}
-
-%Radial basis function interpolation of $\varphi_x(y)$ is cheaper per sample point because radial basis functions are given by analytic formulas.
-However, the cost of these Poisson PDE solves is typically cheap compared to the cost of computing the impulse response samples, $\varphi_{x_i}$.
-Computing the impulse response samples, $\varphi_{x_i}$, requires applying the Hessian to vectors, which in turn requires performing PDE solves of the linearized forward and ajoint equations for the inverse problem.
-These forward and adjoint PDE solves are typically much more expensive than Poisson PDE solves, because the Poisson equation is well-studied and easy to solve with multigrid.
 

### 2021-01-03 paper/localpsf.tex
moved some h-matrix stuff to the appendix

@@
 	\item Recursive algorithms can take advantage of this low rank block structure to perform matrix arithmetic fast.
-Conventional algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix addition require either $O(n^2)$ or $O(n^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(n \log^a n)$ time and memory for hierarchical matrices (here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix).
+Conventional algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix addition require either $O(n^2)$ or $O(n^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(N \log^a N)$ time and memory for hierarchical matrices (here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix).
 \end{itemize}
@@
 We convert $\widetilde{\mathcal{A}}$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package [CITE].
-Briefly, the H-matrix construction process that we use proceeds as follows.
+Briefly, this method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix to construct a recursive block partitioning of $\widetilde{\mathcal{A}}$, such that all blocks at the finest level of the partitioning are either low rank, or are small.
+Then, the method forms low rank approximations of the blocks that are expected to be low rank by sampling specially chosen rows and columns of those blocks.
+The small high rank blocks are formed directly by evaluating their entries.
+The primary cost of constructing the approximation is the evaluation of $O(N \log N)$ matrix entries of $\widetilde{\mathcal{A}}$, which we perform using Algorithm \ref{alg:eval_y_x}.
+We describe this H-matrix construction method in detail in Appendix \ref{app:h_matrix}.
+
+\section{Numerical results}
+\label{sec:numerical_results}
+
+Some numerical results here.
+
+\section{Conclusions}
+\label{sec:conclusions}
+
+Some conclusions here.
+
+\appendix
+
+\section{Hierarchical matrix construction}
+\label{app:h_matrix}
+
+The H-matrix construction process that we use proceeds as follows.
 First, we construct a hierarchical partitionings of the degrees of freedom for the rows and columns of the matrix (cluster trees, Section \ref{sec:cluster_trees}).
@@
 The third step requires us to evaluate $O(n \log n)$ specially-chosen matrix entries of $\widetilde{\mathcal{A}}$, which we do using the method described in Section \ref{sec:eval_matrix_entries}.
-In the remainder of this section we describe each of these steps in more detail.
-
-\subsubsection{Cluster trees}
+
+\subsection{Cluster trees}
 \label{sec:cluster_trees}
@@
 
-\subsubsection{Block cluster tree} 
+\subsection{Block cluster tree} 
 \label{sec:block_cluster_tree}
@@
 
-\subsubsection{Adaptive cross low rank approximation of blocks}
+\subsection{Adaptive cross low rank approximation of blocks}
 \label{sec:adaptive_cross}
@@
 This process requires us to compute $O(r^2 N \log N)$ entries of $\widetilde{\mathcal{A}}$, which is relatively cheap compared to the matrix-vector products with $\mathcal{A}$ that are required to form the product-convolution approximation $\widetilde{\mathcal{A}}$.
-
-\section{Numerical results}
-\label{sec:numerical_results}
-
-Some numerical results here.
-
-\section{Conclusions}
-\label{sec:conclusions}
-
-Some conclusions here.
-
-\appendix
 

### 2021-01-03 paper/localpsf.tex
minor edits

@@
 We call this interpolation scheme \emph{Poisson interpolation}.
-The advantage of Poisson interpolation over other interpolation methods, most notably radial basis function interpolation [cite Escande etc], is that Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions.
+The advantage of Poisson interpolation over other interpolation methods, most notably radial basis function interpolation [CITE ESCANDE ETC], is that Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions.
 This is illustrated in Figure FIG.
@@
 The ideal scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost) [CITE MY DISSERTATION].
-The method in this paper offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because high-rank approximations of an operator can be formed using a small number of matrix-vector products.
+Although the method we present in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest.
+For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because high-rank approximations of an operator can be formed using a small number of matrix-vector products.
 

### 2021-01-04 paper/localpsf.tex
changed format of reference placeholders

@@
 Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$.
-Since the Laplacian measures local deviation from flatness [CITE JACOBSEN], the function $w_i$ is the flattest function, in a least-squares sense, that takes the value one at the sample point $x_i$, and the value zero at the other sample points $x_j$, $i \neq j$.
+Since the Laplacian measures local deviation from flatness \cite{JACOBSEN}, the function $w_i$ is the flattest function, in a least-squares sense, that takes the value one at the sample point $x_i$, and the value zero at the other sample points $x_j$, $i \neq j$.
 Computing the weighting functions $w_i$ requires solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid.
 We call this interpolation scheme \emph{Poisson interpolation}.
-The advantage of Poisson interpolation over other interpolation methods, most notably radial basis function interpolation [CITE ESCANDE ETC], is that Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions.
+The advantage of Poisson interpolation over other interpolation methods, most notably radial basis function interpolation \cite{ESCANDE}, is that Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions.
 This is illustrated in Figure FIG.
@@
 \end{equation}
-Formula \eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks [CITE].
+Formula \eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
 Once in H-matrix format, fast H-matrix arithmetic is used to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations.
@@
 
-The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term [CITE LOTS OF STUFF].
-Either the Lanczos method or the randomized singular value decomposition [CITE RSVD] are used to perform the low rank approximation using only matrix-vector products.
+The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term \cite{CCGOPAPERS}.
+Either the Lanczos method or the randomized singular value decomposition \cite{HMTRANDOM} are used to perform the low rank approximation using only matrix-vector products.
 These methods suffer from a ``data predicament''---as more informative data is used in the inversion, the numerical rank of the data misfit term in the Hessian grows, so more matrix-vector products are required to form the aforementioned low rank approximation.
-The ideal scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost) [CITE MY DISSERTATION].
+The ideal scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost) \cite{MYDISSERTATION}.
 Although the method we present in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest.
@@
 In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\varphi_i$ and $w_i$ may be arbitrary functions.
-Product-convolution approximations have been used in a wide variety of fields going back several decades [CITE BIG LIST].
-For background on product-convolution approximations, we recommend reading the following papers: [CITES, INCLUDE ESCANDE SCATTERED].
-
-We build upon the class of product convolution approximations in which the functions, $\varphi_i$, are impulse responses of $\mathcal{A}$ to point sources at a collection of points $x_i$ [CITE BIG LIST].
-A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the functions $\varphi_i$ with piecewise linear interpolation [CITE NAGY], or splines [CITE SPLINES].
+Product-convolution approximations have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}.
+For background on product-convolution approximations, we recommend reading the following papers: \cite{PRODCONVGOOD}.
+
+We build upon the class of product convolution approximations in which the functions, $\varphi_i$, are impulse responses of $\mathcal{A}$ to point sources at a collection of points $x_i$ \cite{PRODCONVLIST}.
+A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the functions $\varphi_i$ with piecewise linear interpolation \cite{NAGY}, or splines \cite{PRODCONVSPLINES}.
 With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
 This is computationally prohibitive in our applications, because computing the functions $\varphi_i$ requires one matrix vector product with $\mathcal{A}$ for each point $x_i$, and these matrix vector products are computationally costly.
-In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large [CITE, CITE MY PAPER].
+In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
 But even with adaptive refinement, many matrix vector products with $\mathcal{A}$ may be required.
@@
 
-Our method of estimating the support of the functions $\varphi_i$ was inspired by resolution analysis in seismic imaging [CITE].
+Our method of estimating the support of the functions $\varphi_i$ was inspired by resolution analysis in seismic imaging \cite{RESOLUTION}.
 In resolution analysis, $\mathcal{A}$ is the Hessian for a seismic inverse problem, and the width of $\varphi_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
 If the width of $\varphi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
-In [CITE], the width of $\varphi_p$ is estimated to be the local autocorrelation length of the function $\mathcal{A}^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
+In \cite{RESOLUTION}, the width of $\varphi_p$ is estimated to be the local autocorrelation length of the function $\mathcal{A}^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
 In this paper, rather than probing $\mathcal{A}^T$ with random noise functions, we use specific constant, linear, and quadratic functions.
@@
 
-Matrix-vector products with product-convolution approximations can be performed using the fast Fourier transform if the problem is discretized on a regular grid [CITE].
-The product-convolution approximation can then replace the original operator when using Krylov methods to solve linear systems with the original operator as the coefficient operator [CITE].
+Matrix-vector products with product-convolution approximations can be performed using the fast Fourier transform if the problem is discretized on a regular grid \cite{PRODCONVFFT}.
+The product-convolution approximation can then replace the original operator when using Krylov methods to solve linear systems with the original operator as the coefficient operator \cite{PRODCONVKRYLOV}.
 However, for complex geometries, problems are rarely discretized using regular grids, so we cannot easily use the fast Fourier transform.
@@
 Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are amenable to fast linear algebra operations.
-Wavelet compression methods have been used for this purpose [CITE WAVELETS].
-Here we follow our previous work [CITE MY PAPER], in which we convert the product-convolution approximation to H-matrix format.
+Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}.
+Here we follow our previous work \cite{PRODCONVMYPAPER}, in which we convert the product-convolution approximation to H-matrix format.
 
 Conventional H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
-There are matrix-free methods for H-matrix approximation [CITE LEXING PAPER, CITE MARTINSSON, CITE OTHERS], and these methods have been used to form H-matrix compressions of Hessians in PDE constrained inverse problems [CITE ILONA, CITE NOEMI/TUCKER HODLR].
+There are matrix-free methods for H-matrix approximation \cite{LEXINGPEELINGPROCESS}, and these methods have been used to form H-matrix compressions of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
 While these matrix-free H-matrix construction methods are asymptotically scalable in theory, the required number of matrix vector products can be large in practice.
 
-Product-convolution type approximations of Hessians have been used in a seismic inverse problem in [CITE GEORG], and in an advection-diffusion inverse problem in [CITE MY PAPER].
-In [DEMANET SEISMIC], matrix probing [DEMANET MATRIX] is used to approximate the Hessian in a seismic inverse problem as the sum of simple pseudodifferential operators.
-Although it is not explicitly mentioned, the approximation in [DEMANET SEISMIC] could be interpreted as a convolution-product interpolation, which is like a product-convolution approximation, except the order of pointwise multiplications and convolutions is reversed in \eqref{eq:product_convolution}.
+Product-convolution type approximations of Hessians have been used in a seismic inverse problem in \cite{GEORGSEISMICPRODCONV}, and in an advection-diffusion inverse problem in \cite{PRODCONVMYPAPER}.
+In \cite{DEMANETSEISMIC}, matrix probing \cite{MATRIXPROBING} is used to approximate the Hessian in a seismic inverse problem as the sum of simple pseudodifferential operators.
+Although it is not explicitly mentioned, the approximation in \cite{DEMANETSEISMIC} could be interpreted as a convolution-product interpolation, which is like a product-convolution approximation, except the order of pointwise multiplications and convolutions is reversed in \eqref{eq:product_convolution}.
 
@@
 In order to use the product-convolution approximation $\widetilde{\mathcal{A}}$, we convert $\widetilde{\mathcal{A}}$ to hierarchical matrix (H-matrix) format.
-Many classes of hierarchical matrices exist---H1, uniform H1, H2, HODLR, HSS, HBS, and more [CITES].
+Many classes of hierarchical matrices exist---H1, uniform H1, H2, HODLR, HSS, HBS, and more \cite{HMATRIX}.
 We use H1 matrices in our numerical results, but any of the other H-matrix formats could be used instead.
@@
 \end{itemize}
-For more details on H-matrices, we recommend [CITES].
-
-We convert $\widetilde{\mathcal{A}}$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package [CITE].
+For more details on H-matrices, we recommend \cite{HMATRIXGOOD}.
+
+We convert $\widetilde{\mathcal{A}}$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{HLIBPRO}.
 Briefly, this method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix to construct a recursive block partitioning of $\widetilde{\mathcal{A}}$, such that all blocks at the finest level of the partitioning are either low rank, or are small.
@@
 
-Once the block cluster tree has been constructed, low rank approximations of the admissible (low rank) blocks are formed using the adaptive cross method [CITE].
+Once the block cluster tree has been constructed, low rank approximations of the admissible (low rank) blocks are formed using the adaptive cross method \cite{ACA}.
 Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\widetilde{\mathcal{A}}$.
@@
 \end{equation}
-where $U^+$ is the pseudoinverse of $U$ [CITE].
+where $U^+$ is the pseudoinverse of $U$ \cite{CUR}.
 The quality of approximation \eqref{eq:CUR} depends on how well the columns of $C$ approximate the column space of $X$, and how well the rows of $R$ approximate the row space of $X$.
@@
 Keeping $C$ fixed, a set of rows $R$ is chosen so that the so that determinant of the submatrix $U$ within $C$ is as large as possible.
-This may be done via either the maxvol procedure [CITE GOOD SUBMATRIX], or via QR factorization.
+This may be done via either the maxvol procedure \cite{GOODSUBMATRIX}, or via QR factorization.
 Now, keeping $R$ fixed, a set of new columns $C$ is chosen so that the submatrix $U$ within $R$ is as large as possible.
@@
 There is also a small linear algebra overhead cost for the determinant maximization process that is performed at each step.
-For more details on adaptive cross low rank approximation, see [CITE].
+For more details on adaptive cross low rank approximation, see \cite{ACA}.
 The key point is that the adaptive cross method allows us to form a rank-$r$ approximation of an $m \times m$ block of the matrix via a process that only requires accessing $O(mr)$ entries of that block.
@@
 \end{equation*}
-In CITE it is shown that $E_p \cap E_q = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$.
+In \cite{ELLIPSOIDINTERSECT} it is shown that $E_p \cap E_q = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$.
 We check whether $E_p$ and $E_q$ intersect by minimizing $K(s)$ on $(0,1)$.
@@
 
-\section{Additional theory}
+\section{Proofs}
 \label{app:proofs}
 
-These results rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}.
+%These results rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}.
 We prove correctness of Algorithm \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}.
@@
 \begin{proof}
-	asdf
+	WRITE THIS PROOF
 \end{proof}
@@
 
-\section{old stuff}
+\section{DISORGANIZED OLD STUFF}
 

### 2021-01-06 paper/localpsf.tex
edits based on Tuckers comments

@@
 \item[Matrix-free:] One cannot easily access kernel entries $A(y,x)$.
-Access to $\mathcal{A}$ is available only through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors.
-That is, evaluation of the functions
+Access to $\mathcal{A}$ is available only through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary functions.
+That is, evaluation of the maps
 \begin{equation*}
@@
 \end{description}
+
+After discretization (see Appendix \ref{app:discretized_operations}), the operator $\mathcal{A}$ becomes a dense matrix, $\mathbf{A}$, that is typically too large to be built and stored.
+The first property is called ``matrix-free'' because at the discrete level it means that we can perform matrix-vector products of $\mathbf{A}$ and $\mathbf{A}^T$ with arbitrary vectors, but cannot easily access matrix entries $\mathbf{A}_{ij}$.
+In the matrix-free regime, to compute $\mathbf{A}_{ij}$ one could compute $\mathbf{v} = \mathbf{A}\mathbf{e}_j$, then extract the $i^\text{th}$ entry, $\mathbf{A}_{ij} = \mathbf{v}_i$.
+However, this process is wasteful because one computes the entire vector $\mathbf{v}$, then discards all but one component.
+Here $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero.
+In PDE constrained optimization and inverse problems, Hessians are typically available matrix free; we do not have direct access to these operators, but we can apply them to vectors via a process that involve solving two PDEs \cite{HESSIANADJOINT}.
+More generally, an operator is matrix-free if the process of applying the operator to a function involves solving a large linear system, using a timestepping procedure, or performing an iterative process.
+
 Local translation invariance is illustrated in Figure FIG.
@@
 We compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb''). That is, we compute
-\begin{equation*}
+\begin{equation}
+\label{eq:eta_intro}
 	\eta := \sum_{i=1}^m A(~\cdot~, x_i) = \left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*.
-\end{equation*}
-Here $\sum_{i=1}^m \delta_{x_i}$ is the Dirac comb distribution, and $\left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*$ denotes the Riesz representation of the application of $\mathcal{A}$ to the Dirac comb.
-In computations, this amounts to computing $\boldsymbol{\eta} = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \left(\sum_{i=1}^m \boldsymbol{\delta}_{x_i}\right)$, where $\boldsymbol{\eta}$, $\mathbf{A}$, and $\boldsymbol{\delta}_{x_i}$ are discretized versions of $\eta$, $A$, and $\delta_{x_i}$, respectively, and $\mathbf{M}$ is a mass matrix.
+\end{equation}
+Here $\sum_{i=1}^m \delta_{x_i}$ is the Dirac comb distribution, $\left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle$ denotes the application of $\mathcal{A}$ to the Dirac comb, and $\left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*$ denotes the Riesz representation of the application of $\mathcal{A}$ to the Dirac comb.
+Although the domain of $\mathcal{A}$ is defined as $L^2(\Omega)$, which does not contain the Dirac comb distribution, the action of $\mathcal{A}$ may be extended to distributions in a natural way, as described in Appendix \ref{app:distributions}.
+We write $\langle \mathcal{A}, \xi \rangle$ to denote the application of $\mathcal{A}$ to the distribution $\xi$.
+Because $\mathcal{A}$ maps $L^2(\Omega) \rightarrow L^2(\Omega)'$, applying $\mathcal{A}$ to a function or distribution results in a linear functional.
+We take the Riesz representative of the linear functional $\left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle \in L^2(\Omega)'$ to convert it into a function $\left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^* \in L^2(\Omega)$.
+In computations, the process of applying $\mathcal{A}$ to the Dirac comb, then forming the Riesz representative, amounts to computing $\boldsymbol{\eta} = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \left(\sum_{i=1}^m \boldsymbol{\delta}_{x_i}\right)$, where $\boldsymbol{\eta}$, $\mathbf{A}$, and $\boldsymbol{\delta}_{x_i}$ are discretized versions of $\eta$, $A$, and $\delta_{x_i}$, respectively, and $\mathbf{M}$ is a mass matrix.
+
 The batch of points $x_i$ are chosen so that the support of $\left\langle \mathcal{A}, \delta_{x_i}\right\rangle^*$ and the support of $\left\langle \mathcal{A}, \delta_{x_j}\right\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$.
@@
 This is illustrated in Figure FIG.
-The process is repeated to get more batches of $\varphi_i$'s, until the approximation is sufficiently accurate or a maximum number of sample points is reached.
+The process is repeated to get more batches of $\varphi_i$'s, until a desired number of batches is reached.
 
@@
 Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because $\mathcal{H}^{-1}$ locally approximates the Bayesian posterior covariance for $m$.
-We only have access to $\mathcal{H}$ via its action on an arbitrary vectors $v$, i.e., evaluation of the map $v \mapsto \mathcal{H} v$.
+%We only have access to $\mathcal{H}$ via its action on an arbitrary vectors $v$, i.e., evaluation of the map $v \mapsto \mathcal{H} v$.
 Evaluating $v \mapsto \mathcal{H} v$ is an expensive process that involves solving two auxiliary linear PDEs.
@@
 Either the Lanczos method or the randomized singular value decomposition \cite{HMTRANDOM} are used to perform the low rank approximation using only matrix-vector products.
-These methods suffer from a ``data predicament''---as more informative data is used in the inversion, the numerical rank of the data misfit term in the Hessian grows, so more matrix-vector products are required to form the aforementioned low rank approximation.
+These methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of matrix-vector products are required to form the aforementioned low rank approximation.
 The ideal scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost) \cite{MYDISSERTATION}.
@@
 where $\tau$ is a small number of standard deviations; we use $\tau=3$. 
-The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations of the mean for the Gaussian distribution that has the same mean and covariance as the normalized version of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$.
+%The ellipsoid $E_{x}$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian approximation of the normalized version of $\langle \mathcal{A}, \delta_{x}\rangle^{*}$ which has the same mean and covariance.
+The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian distribution which has the same mean and covariance as the normalized version of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$.
 The fraction of the mass of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though we find this bound is conservative and far less mass resides in this region in practice.
@@
 \begin{equation*}
-	\xi^b := \sum_{x \in S_b} \delta_{x_i}.
+	\xi^b := \sum_{x_i \in S_b} \delta_{x_i}.
 \end{equation*}
@@
 \min_{w_i \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
-\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
+\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,r.
 \end{aligned}
@@
 
-We now describe how to efficiently evaluate $\widetilde{A}(y,x)$ at an arbitrary point $(x,y) \in \Omega \times \Omega$. From \eqref{eq:H_kernel_entries}, we have
+We now describe how to efficiently evaluate $\widetilde{A}(y,x)$ at an arbitrary point $(x,y) \in \Omega \times \Omega$. From \eqref{eq:kernel_entries}, we have
 \begin{equation}
@@
 	\item Recursive algorithms can take advantage of this low rank block structure to perform matrix arithmetic fast.
-Conventional algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix addition require either $O(n^2)$ or $O(n^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(N \log^a N)$ time and memory for hierarchical matrices (here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix).
+Conventional algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix addition require either $O(N^2)$ or $O(N^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(N \log^a N)$ time and memory for hierarchical matrices.
+Here $N$ is the number of degrees of freedom for the discretized version of the problem, and $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix.
 \end{itemize}
@@
 The first two steps require geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix, but these steps do not depend on the particular values of matrix entries.
-The third step requires us to evaluate $O(n \log n)$ specially-chosen matrix entries of $\widetilde{\mathcal{A}}$, which we do using the method described in Section \ref{sec:eval_matrix_entries}.
+The third step requires us to evaluate $O(N \log N)$ specially-chosen matrix entries of $\widetilde{\mathcal{A}}$, which we do using the method described in Section \ref{sec:eval_matrix_entries}.
 
@@
 
-\section{DISORGANIZED OLD STUFF}
-
-\subsection{Background: distributions}
+\section{Background: distributions}
+\label{app:distributions}
 
@@
 That is, $\langle\mathcal{A},\delta_x\rangle^*$ is the function $y \mapsto A(y,x)$.
+
+\section{DISORGANIZED OLD STUFF}
 

### 2021-01-06 paper/localpsf.tex
trying to clarify dirac comb discussion

@@
 After discretization (see Appendix \ref{app:discretized_operations}), the operator $\mathcal{A}$ becomes a dense matrix, $\mathbf{A}$, that is typically too large to be built and stored.
-The first property is called ``matrix-free'' because at the discrete level it means that we can perform matrix-vector products of $\mathbf{A}$ and $\mathbf{A}^T$ with arbitrary vectors, but cannot easily access matrix entries $\mathbf{A}_{ij}$.
+The first property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\mathbf{A}$ and $\mathbf{A}^T$ with arbitrary vectors, but one cannot easily access matrix entries $\mathbf{A}_{ij}$.
 In the matrix-free regime, to compute $\mathbf{A}_{ij}$ one could compute $\mathbf{v} = \mathbf{A}\mathbf{e}_j$, then extract the $i^\text{th}$ entry, $\mathbf{A}_{ij} = \mathbf{v}_i$.
@@
 In PDE constrained optimization and inverse problems, Hessians are typically available matrix free; we do not have direct access to these operators, but we can apply them to vectors via a process that involve solving two PDEs \cite{HESSIANADJOINT}.
-More generally, an operator is matrix-free if the process of applying the operator to a function involves solving a large linear system, using a timestepping procedure, or performing an iterative process.
+More broadly, $\mathbf{A}$ is matrix-free if $\mathbf{A}$ is defined implicitly via its action on vectors, and the process of applying $\mathbf{A}$ to a vector requires performing a nontrivial computational procedure, such as solving a large linear system, or timestepping.
 
@@
 
-We compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb''). That is, we compute
+We compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, 
+\begin{equation*}
+	\xi := \sum_{i=1}^l \delta_{x_i},
+\end{equation*}
+where the $\delta_{x_i}$'s are point sources (delta distributions) centered at a collection of scattered points $x_i$. That is, we compute
 \begin{equation}
 \label{eq:eta_intro}
-	\eta := \sum_{i=1}^m A(~\cdot~, x_i) = \left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*.
-\end{equation}
-Here $\sum_{i=1}^m \delta_{x_i}$ is the Dirac comb distribution, $\left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle$ denotes the application of $\mathcal{A}$ to the Dirac comb, and $\left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*$ denotes the Riesz representation of the application of $\mathcal{A}$ to the Dirac comb.
-Although the domain of $\mathcal{A}$ is defined as $L^2(\Omega)$, which does not contain the Dirac comb distribution, the action of $\mathcal{A}$ may be extended to distributions in a natural way, as described in Appendix \ref{app:distributions}.
-We write $\langle \mathcal{A}, \xi \rangle$ to denote the application of $\mathcal{A}$ to the distribution $\xi$.
-Because $\mathcal{A}$ maps $L^2(\Omega) \rightarrow L^2(\Omega)'$, applying $\mathcal{A}$ to a function or distribution results in a linear functional.
-We take the Riesz representative of the linear functional $\left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle \in L^2(\Omega)'$ to convert it into a function $\left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^* \in L^2(\Omega)$.
+	\eta := \left\langle \mathcal{A}, \xi \right\rangle^* = \sum_{i=1}^l A(~\cdot~, x_i).
+\end{equation}
+Here $\left\langle \mathcal{A}, \xi \right\rangle$ denotes the result of applying the operator $\mathcal{A}$ to the distribution $\xi$, and $\left\langle \mathcal{A}, \xi \right\rangle^*$ denotes the Riesz representation of $\langle\mathcal{A}, \xi\rangle$.
+The distribution $\xi$ is called a \emph{Dirac comb}.
+Although the domain of $\mathcal{A}$ is defined as $L^2(\Omega)$, which does not contain the Dirac comb distribution, the action of $\mathcal{A}$ is extended to distributions in a natural way, as described in Appendix \ref{app:distributions}.
+We take the Riesz representative of the linear functional $\left\langle \mathcal{A}, \xi \right\rangle \in L^2(\Omega)'$ to convert it into a function $\left\langle \mathcal{A}, \xi \right\rangle^* \in L^2(\Omega)$.
 In computations, the process of applying $\mathcal{A}$ to the Dirac comb, then forming the Riesz representative, amounts to computing $\boldsymbol{\eta} = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \left(\sum_{i=1}^m \boldsymbol{\delta}_{x_i}\right)$, where $\boldsymbol{\eta}$, $\mathbf{A}$, and $\boldsymbol{\delta}_{x_i}$ are discretized versions of $\eta$, $A$, and $\delta_{x_i}$, respectively, and $\mathbf{M}$ is a mass matrix.
@@
 The batch of points $x_i$ are chosen so that the support of $\left\langle \mathcal{A}, \delta_{x_i}\right\rangle^*$ and the support of $\left\langle \mathcal{A}, \delta_{x_j}\right\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$.
-Since the impulse responses do not overlap, we may post-process $\eta$ to recover $\varphi_i$ for all of the points $x_i$ in the batch---by applying $\mathcal{A}$ to one distribution, we recover many $\varphi_i$.
+Since these supports do not overlap, we may post-process $\eta$ to recover $\varphi_i$ for all of the points $x_i$ in the batch---by applying $\mathcal{A}$ to one distribution, we recover many $\varphi_i$.
 This is illustrated in Figure FIG.

### 2021-01-06 paper/localpsf.tex
trying to clarify dirac comb discussion

@@
 The first property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\mathbf{A}$ and $\mathbf{A}^T$ with arbitrary vectors, but one cannot easily access matrix entries $\mathbf{A}_{ij}$.
-In the matrix-free regime, to compute $\mathbf{A}_{ij}$ one could compute $\mathbf{v} = \mathbf{A}\mathbf{e}_j$, then extract the $i^\text{th}$ entry, $\mathbf{A}_{ij} = \mathbf{v}_i$.
+In the matrix-free regime, to compute $\mathbf{A}_{ij}$ one could compute $\mathbf{v} = \mathbf{A}\mathbf{e}_j$, then extract $\mathbf{A}_{ij} = \mathbf{v}_i$.
 However, this process is wasteful because one computes the entire vector $\mathbf{v}$, then discards all but one component.
 Here $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero.
-In PDE constrained optimization and inverse problems, Hessians are typically available matrix free; we do not have direct access to these operators, but we can apply them to vectors via a process that involve solving two PDEs \cite{HESSIANADJOINT}.
+In PDE constrained optimization and inverse problems, Hessians are typically available matrix free; we do not have direct access to Hessian matrix entries, but the implicit function theorem allows us to apply the Hessian to a vector via a process that involves solving two PDEs \cite{HESSIANADJOINT}.
 More broadly, $\mathbf{A}$ is matrix-free if $\mathbf{A}$ is defined implicitly via its action on vectors, and the process of applying $\mathbf{A}$ to a vector requires performing a nontrivial computational procedure, such as solving a large linear system, or timestepping.
@@
 \end{equation}
-Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$.
+Here $\Delta$ is the Laplacian on $\Omega$, $H^2_N(\Omega)$ is the space of all functions in the Sobolev space $H^2(\Omega)$ with Neumann zero boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$.
 Since the Laplacian measures local deviation from flatness \cite{JACOBSEN}, the function $w_i$ is the flattest function, in a least-squares sense, that takes the value one at the sample point $x_i$, and the value zero at the other sample points $x_j$, $i \neq j$.

### 2021-01-07 paper/localpsf.tex
merged sections 2 and 3

@@
 Our operator approximation method is defined by how we compute the convolution kernels, $\varphi_i$, how we choose the points, $x_i$, what weighting functions, $w_i$, that we use, and how we use the resulting product-convolution approximation to perform linear algebra operations such as solving linear systems.
-In this section we summarize the answers to these questions; we will provide detailed answers to these questions in Section \ref{sec:method} and Section \ref{sec:using_approximation}.
+In this section we summarize the answers to these questions; we will provide detailed answers to these questions in Section \ref{sec:method}.
 
@@
 This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\widetilde{\mathcal{A}} \approx \mathcal{A}$, given in \eqref{eq:product_convolution}.
+We also describe how we evaluate kernel entries of the approximation, $\widetilde{A}(y,x)$, and how we convert $\widetilde{\mathcal{A}}$ to H-matrix format.
 \begin{itemize}
 \item We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}).
-To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as individual functions.
+To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as distinct functions.
 \item We use a greedy algorithm to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}).
@@
 The complete algorithm for constructing $\widetilde{\mathcal{A}}$ is shown in Algorithm \ref{alg:construct_Atilde}.
+The method for evaluating kernel entries of $\widetilde{A}$ is shown in Section \ref{sec:eval_matrix_entries}.
+The method for converting $\widetilde{\mathcal{A}}$ to H-matrix format is shown in Section \ref{sec:H_matrix_conversion}.
 
@@
 
-\section{Using the approximation}
-\label{sec:using_approximation}
-
 \subsection{Evaluating kernel entries of the approximation}
@@
 \subsection{Conversion to hierarchical matrix format}
+\label{sec:H_matrix_conversion}
 

### 2021-01-15 paper/localpsf.tex
lots of edits based on discussion with Noemi and Tucker today. More edits to come

@@
 We present a fast matrix free method for approximating locally translation-invariant operators, $\mathcal{A}$, that have locally supported non-negative integral kernels.
-Such operators arise, for example, as Schur complements in reduced space methods for solving partial differential equations (PDEs), and as Hessians in distributed parameter PDE-constrained inverse problems.
+Such operators arise, for example, in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, as Poincare-Steklov operators, as covariance operators, as blurring operators in imaging problems, and as Hessians in distributed parameter PDE-constrained inverse problems.
 
@@
 \end{equation}
-with integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$. We focus on operators of this form which have the following properties:
+with integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$.
+Here $\mathcal{A}u \in L^2(\Omega)'$ is the linear functional that results from applying $\mathcal{A}$ to $u\in L^2(\Omega)$, and $\left(\mathcal{A}u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
+We focus on operators of this form which have the following properties:
 \begin{description}
 \item[Matrix-free:] One cannot easily access kernel entries $A(y,x)$.
-Access to $\mathcal{A}$ is available only through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary functions.
+Access to $\mathcal{A}$ is available only through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary functions (i.e., operator actions).
 That is, evaluation of the maps
@@
 \end{equation*}
-\item[Local support:] $A(y, x)$ is supported near $y=x$.
+\item[Local support:] Let $\phi_x$ be the following \emph{impulse response}:
+\begin{equation}
+\label{eq:impulse_response_defn}
+\phi_x(y) := A(y, x).
+\end{equation} 
+We say that $\mathcal{A}$ has local support if, for all $x\in \Omega$, the support of $\phi_x$ is contained (or approximately contained) in a neighborhood of $x$.
+The smaller these supports are, the better our algorithm will perform.
 \item[Local translation invariance:] If $h$ is not too large, then
@@
 Here $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero.
-In PDE constrained optimization and inverse problems, Hessians are typically available matrix free; we do not have direct access to Hessian matrix entries, but the implicit function theorem allows us to apply the Hessian to a vector via a process that involves solving two PDEs \cite{HESSIANADJOINT}.
-More broadly, $\mathbf{A}$ is matrix-free if $\mathbf{A}$ is defined implicitly via its action on vectors, and the process of applying $\mathbf{A}$ to a vector requires performing a nontrivial computational procedure, such as solving a large linear system, or timestepping.
+In general, $\mathbf{A}$ is matrix-free if $\mathbf{A}$ is defined implicitly via its action on vectors, and the process of applying $\mathbf{A}$ to a vector requires performing a nontrivial computational procedure, such as solving a large linear system, or timestepping.
+For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessians information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
+
+Let $\delta_x$ denote the point source (delta distribution) centered at $x$.
+The function $\phi_x$ is called the impulse response because straightforward analysis shows that it may be expressed as follows:
+\begin{equation}
+\label{eq:impulse_response_delta_action}
+	\phi_x = \left\langle \mathcal{A}, \delta_x \right\rangle^*,
+\end{equation}
+where $\left\langle \mathcal{A}, \delta_x \right\rangle \in L^2(\Omega)'$ denotes the result of applying the operator $\mathcal{A}$ to the distribution $\delta_x$, and $\left\langle \mathcal{A}, \delta_x \right\rangle^* \in L^2(\Omega)$ denotes the Riesz representation of $\langle\mathcal{A}, \delta_x\rangle$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\psi \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\psi^* \in L^2(\Omega)$ such that $\psi(v) = \left(\psi^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} The local support property says, therefore, that a point source at $x$ has zero impact (or minimal impact) on points $y$ that are far away from $x$.
+
+Note that while the domain of $\mathcal{A}$ is defined as $L^2(\Omega)$, which does not contain the delta distribution, the application of $\mathcal{A}$ is extended to distributions in a natural way, as described in Appendix \ref{app:distributions}.
+%We take the Riesz representative of the linear functional $\left\langle \mathcal{A}, \delta_x \right\rangle \in L^2(\Omega)'$ to convert it into a function $\left\langle \mathcal{A}, \delta_x \right\rangle^* \in L^2(\Omega)$.
+After discretization, the process of applying $\mathcal{A}$ to a delta distribution, then forming the Riesz representative, amounts to computing $\boldsymbol{\phi}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_{x_i}$, where $\boldsymbol{\phi}_x$, $\mathbf{A}$, and $\boldsymbol{\delta}_x$ are discretized versions of $\phi_x$, $A$, and $\delta_x$, respectively, and $\mathbf{M}$ is a mass matrix (e.g., in a finite element context).
 
@@
 \label{eq:convolution_kernel}
-	\varphi_x(y) := A(y+x,x).
-\end{equation}
-The function $\varphi_x$ is the result of applying $\mathcal{A}$ to a point source, $\delta_x$, at some fixed point $x \in \Omega$, then translating the resulting impulse response function to re-center it at zero instead of $x$.
+	\varphi_x(y) := \phi_x(y+x).
+\end{equation}
+The convolution kernel $\varphi_x$ is the result of translating the impulse response $\phi_x$ to re-center it at zero instead of $x$.
 Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes.
-We therefore approximate $\mathcal{A}$ by a spatially varying weighted sum of convolution operators of the form
+We therefore approximate $\mathcal{A}$ by a spatially varying weighted sum of convolution operators, $\widetilde{\mathcal{A}} \approx \mathcal{A}$, of the form
 \begin{equation}
 \label{eq:product_convolution}
-	\left(\mathcal{A}u\right)(v) \approx \left(\widetilde{\mathcal{A}}u\right)(v) := \left( v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
+	\left(\widetilde{\mathcal{A}}u\right)(v) := \left( v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
 \end{equation}
 where $f \cdot g$ denotes pointwise multiplication, $f \ast g$ denotes convolution, and $\left(f, g\right)_{L^2(\Omega)}$ denotes the $L^2$ inner product on $\Omega$, for functions $f,g \in L^2(\Omega)$.
-Here
-\begin{itemize}
-	\item The local convolution kernels, $\varphi_i := \varphi_{x_i}$, are given by formula \eqref{eq:convolution_kernel}, but with $x$ replaced by $x_i$.
-	\item $\{x_i\}_{i=1}^r \subset \Omega$ is a collection of points scattered throughout the domain.
-	\item The functions $w_i$ are spatially varying weighting functions that are used to interpolate the convolution kernels $\varphi_i$.
-\end{itemize}
+Here we write $\varphi_i := \varphi_{x_i}$ (with $\varphi_{x_i}$ defined in \eqref{eq:convolution_kernel}) to denote local convolution kernels corresponding to a collection of points $\{x_i\}_{i=1}^r \subset \Omega$ scattered throughout the domain.
+The functions $w_i$ are spatially varying weighting functions that are used to interpolate the convolution kernels $\varphi_i$.
+Our operator approximation method is defined by how we compute the convolution kernels, $\varphi_i$, how we choose the points, $x_i$, what weighting functions, $w_i$, that we use, and how we use the resulting product-convolution approximation to perform linear algebra operations such as solving linear systems.
+In Section \ref{sec:overview_intro} we summarize the answers to these questions; we will provide detailed answers to these questions in Section \ref{sec:method}.
+
 Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution.
@@
 \subsection{Overview of the method}
-
-Our operator approximation method is defined by how we compute the convolution kernels, $\varphi_i$, how we choose the points, $x_i$, what weighting functions, $w_i$, that we use, and how we use the resulting product-convolution approximation to perform linear algebra operations such as solving linear systems.
-In this section we summarize the answers to these questions; we will provide detailed answers to these questions in Section \ref{sec:method}.
-
-We compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, 
-\begin{equation*}
-	\xi := \sum_{i=1}^l \delta_{x_i},
-\end{equation*}
-where the $\delta_{x_i}$'s are point sources (delta distributions) centered at a collection of scattered points $x_i$. That is, we compute
+\label{sec:overview_intro}
+
+We compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a Dirac comb, then post-processing the result. That is, we compute
 \begin{equation}
 \label{eq:eta_intro}
-	\eta := \left\langle \mathcal{A}, \xi \right\rangle^* = \sum_{i=1}^l A(~\cdot~, x_i).
-\end{equation}
-Here $\left\langle \mathcal{A}, \xi \right\rangle$ denotes the result of applying the operator $\mathcal{A}$ to the distribution $\xi$, and $\left\langle \mathcal{A}, \xi \right\rangle^*$ denotes the Riesz representation of $\langle\mathcal{A}, \xi\rangle$.
-The distribution $\xi$ is called a \emph{Dirac comb}.
-Although the domain of $\mathcal{A}$ is defined as $L^2(\Omega)$, which does not contain the Dirac comb distribution, the action of $\mathcal{A}$ is extended to distributions in a natural way, as described in Appendix \ref{app:distributions}.
-We take the Riesz representative of the linear functional $\left\langle \mathcal{A}, \xi \right\rangle \in L^2(\Omega)'$ to convert it into a function $\left\langle \mathcal{A}, \xi \right\rangle^* \in L^2(\Omega)$.
-In computations, the process of applying $\mathcal{A}$ to the Dirac comb, then forming the Riesz representative, amounts to computing $\boldsymbol{\eta} = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \left(\sum_{i=1}^m \boldsymbol{\delta}_{x_i}\right)$, where $\boldsymbol{\eta}$, $\mathbf{A}$, and $\boldsymbol{\delta}_{x_i}$ are discretized versions of $\eta$, $A$, and $\delta_{x_i}$, respectively, and $\mathbf{M}$ is a mass matrix.
-
-The batch of points $x_i$ are chosen so that the support of $\left\langle \mathcal{A}, \delta_{x_i}\right\rangle^*$ and the support of $\left\langle \mathcal{A}, \delta_{x_j}\right\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$.
-Since these supports do not overlap, we may post-process $\eta$ to recover $\varphi_i$ for all of the points $x_i$ in the batch---by applying $\mathcal{A}$ to one distribution, we recover many $\varphi_i$.
+\eta := \left\langle \mathcal{A}, \xi \right\rangle^*
+\end{equation}
+where 
+\begin{equation*}
+	\xi := \sum_{i=1}^l \delta_{x_i}
+\end{equation*}
+is the Dirac comb associated with point sources centered at a collection of scattered points $x_i$.
+The batch of points $x_i$ are chosen so that the support of $\phi_{x_i}$ and the support of $\phi_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$.
+
+By linearity, and the expression for $\phi_x$ in \eqref{eq:impulse_response_delta_action}, we have
+\begin{equation*}
+\eta = \sum_{i=1}^l \phi_{x_i}.
+\end{equation*}
+Since the support of $\phi_{x_i}$ and the support of $\phi_{x_j}$ do not overlap, we may post-process $\eta$ to recover $\phi_i$, and therefore $\varphi_i$, for all of the points $x_i$ in the batch---by applying $\mathcal{A}$ to one distribution, we recover many $\varphi_i$.
 This is illustrated in Figure FIG.
@@
 
-In order to choose the points $x_i$, we need to estimate the support of $\left\langle \mathcal{A}, \delta_{x_i}\right\rangle^*$ \emph{before} we compute it.
-The supports of the functions $\left\langle \mathcal{A}, \delta_{x}\right\rangle^*$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\mathcal{A}^T$ to a small number of specially chosen functions (three functions if $d=1$, six if $d=2$, ten if $d=3$), then post processing the results (Section \ref{eq:mean_and_covariance_estimation}).
+In order to choose the points $x_i$, we need to estimate the support of $\phi_{x_i}$ \emph{before} we compute it.
+The supports of the functions $\phi_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\mathcal{A}^T$ to a small number of specially chosen functions (three functions if $d=1$, six if $d=2$, ten if $d=3$), then post processing the results (Section \ref{eq:mean_and_covariance_estimation}).
 This procedure is the only part of the paper that requires non-negativity of the integral kernel.
@@
 \end{equation}
-Formula \eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
+This follows from~\eqref{eq:product_convolution} and the fact that the convolution operator $u \mapsto \varphi_i \ast u$ has $(y,x)$ kernel entry given by $\varphi(y-x)$.
+Formula~\eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
 Once in H-matrix format, fast H-matrix arithmetic is used to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations.
@@
 The method for converting $\widetilde{\mathcal{A}}$ to H-matrix format is shown in Section \ref{sec:H_matrix_conversion}.
+In the rest of the section we will need the following definitions.
+
+\begin{defn}
+	We define the spatially varying volume, $\alpha$, the scaled impulse response, $\rho_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
+	\begin{align*}
+	\alpha:\Omega \rightarrow \mathbb{R}, &\qquad \alpha(x) := \int_{\Omega} \phi_x(z) dz, \\
+	\rho_x:\Omega \rightarrow \mathbb{R}, &\qquad \rho_x := \phi_x \big/ \alpha(x), \\
+	\mu:\Omega \rightarrow \mathbb{R}^d, &\qquad \mu(x) := \int_\Omega z \rho_x(z) dz, \\
+	\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\Sigma(x) := \int_\Omega (z - \mu(x))(z - \mu(x))^T \rho_x(z) dz.
+	\end{align*}
+\end{defn}
 
@@
 
-Since $\varphi_x$, and thus $\langle\mathcal{A},\delta_{x}\rangle^*$, is nonnegative, $\langle\mathcal{A},\delta_{x}\rangle^*$ is a scaled probability distribution.
+Since $\phi_x$ is nonnegative, it is a scaled probability distribution.
 We compute the scaling factor, denoted $\alpha(x)$, for all $x$ simultaneously by applying $\mathcal{A}$ to a constant function.
-We then compute the spatially varying mean, $\mu(x)$, and covariance, $\Sigma(x)$ of the normalized version of $\langle\mathcal{A},\delta_{x}\rangle^*$, for all $x$ simultaneously, by applying $\mathcal{A}^T$ to a small number of linear and quadratic functions.
-We make the approximation that the support of $\langle\mathcal{A},\delta_{x}\rangle^*$ is contained within the ellipsoid
+We then compute the spatially varying mean, $\mu(x)$, and covariance, $\Sigma(x)$ of the normalized version of $\phi_x$, for all $x$ simultaneously, by applying $\mathcal{A}^T$ to a small number of linear and quadratic functions.
+We make the approximation that the support of $\phi_x$ is contained within the ellipsoid
 \begin{equation}
@@
 \end{equation}
-where $\tau$ is a small number of standard deviations; we use $\tau=3$. 
+where $\tau$ is a small number of standard deviations.
+The bigger $\tau$ is, the smaller the overlap between impulse responses within a batch will be; the smaller $\tau$ is, the more impulse responses we can fit within each batch.
+In practice we found that $\tau=3$ provides a good balance between these competing goals.
 %The ellipsoid $E_{x}$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian approximation of the normalized version of $\langle \mathcal{A}, \delta_{x}\rangle^{*}$ which has the same mean and covariance.
-The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian distribution which has the same mean and covariance as the normalized version of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$.
-The fraction of the mass of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though we find this bound is conservative and far less mass resides in this region in practice.
-
-\begin{defn}
-	We define the spatially varying volume, $\alpha$, the scaled impulse response, $\rho_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
-	\begin{align*}
-		\alpha:\Omega \rightarrow \mathbb{R}, &\qquad \alpha(x) := \int_{\Omega} \langle\mathcal{A}, \delta_x\rangle^*(z) dz, \\
-		\rho_x:\Omega \rightarrow \mathbb{R}, &\qquad \rho_x := \langle\mathcal{A}, \delta_x\rangle^* \big/ \alpha(x), \\
-		\mu:\Omega \rightarrow \mathbb{R}^d, &\qquad \mu(x) := \int_\Omega z \rho_x(z) dz, \\
-		\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\Sigma(x) := \int_\Omega (z - \mu(x))(z - \mu(x))^T \rho_x(z) dz.
-	\end{align*}
-\end{defn}
+The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$.
+Note that this Gaussian distribution has the same mean and covariance as  which has the same mean and covariance as the normalized version of $\phi_x$ (see Figure FIG).
+The fraction of the mass of $\phi_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though we find this bound is conservative and far less mass resides in this region in practice.
 

### 2021-01-15 paper/localpsf.tex
lots of edits based on discussion with Noemi and Tucker today. More edits to come

@@
 
-Note that while the domain of $\mathcal{A}$ is defined as $L^2(\Omega)$, which does not contain the delta distribution, the application of $\mathcal{A}$ is extended to distributions in a natural way, as described in Appendix \ref{app:distributions}.
+Note that while the domain of $\mathcal{A}$ is defined as $L^2(\Omega)$, which does not contain the delta distribution, the action of $\mathcal{A}$ may be extended to distributions in a natural way, as described in Appendix \ref{app:distributions}.
+The action of $\mathcal{A}$ on a distribution $\mu$ is written as $\langle \mathcal{A}, \mu \rangle$.
+We use this angle bracket notation to distinguish the action of $\mathcal{A}$ on distributions from the action of $\mathcal{A}$ on functions.
 %We take the Riesz representative of the linear functional $\left\langle \mathcal{A}, \delta_x \right\rangle \in L^2(\Omega)'$ to convert it into a function $\left\langle \mathcal{A}, \delta_x \right\rangle^* \in L^2(\Omega)$.
-After discretization, the process of applying $\mathcal{A}$ to a delta distribution, then forming the Riesz representative, amounts to computing $\boldsymbol{\phi}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_{x_i}$, where $\boldsymbol{\phi}_x$, $\mathbf{A}$, and $\boldsymbol{\delta}_x$ are discretized versions of $\phi_x$, $A$, and $\delta_x$, respectively, and $\mathbf{M}$ is a mass matrix (e.g., in a finite element context).
+After discretization, the process of applying $\mathcal{A}$ to a distribution, $\mu$, then forming the Riesz representative of the resulting linear functional, amounts to computing $\mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\mu}$, where $\mathbf{A}$, and $\boldsymbol{\mu}$ are discretized versions of $A$ and $\delta_x$, respectively, and $\mathbf{M}$ is a mass matrix (e.g., in a finite element context).
 
@@
 
-We compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a Dirac comb, then post-processing the result. That is, we compute
+We compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a Dirac comb, then post-processing the result. First, we compute
 \begin{equation}
@@
 The batch of points $x_i$ are chosen so that the support of $\phi_{x_i}$ and the support of $\phi_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$.
-
 By linearity, and the expression for $\phi_x$ in \eqref{eq:impulse_response_delta_action}, we have
@@
 \end{equation*}
-Since the support of $\phi_{x_i}$ and the support of $\phi_{x_j}$ do not overlap, we may post-process $\eta$ to recover $\phi_i$, and therefore $\varphi_i$, for all of the points $x_i$ in the batch---by applying $\mathcal{A}$ to one distribution, we recover many $\varphi_i$.
+We may therefore post-process $\eta$ to recover $\phi_{x_i}$ for all of the points $x_i$ in the batch.
+Since $\varphi_i$ is a translated versions of $\phi_{x_i}$, we also recover $\varphi_i$ for all points $x_i$ in the batch---by applying $\mathcal{A}$ to one distribution, we compute many $\varphi_i$.
 This is illustrated in Figure FIG.

### 2021-01-15 paper/localpsf.tex
lots of edits based on discussion with Noemi and Tucker today. More edits to come

@@
 \end{equation}
-where $\left\langle \mathcal{A}, \delta_x \right\rangle \in L^2(\Omega)'$ denotes the result of applying the operator $\mathcal{A}$ to the distribution $\delta_x$, and $\left\langle \mathcal{A}, \delta_x \right\rangle^* \in L^2(\Omega)$ denotes the Riesz representation of $\langle\mathcal{A}, \delta_x\rangle$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\psi \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\psi^* \in L^2(\Omega)$ such that $\psi(v) = \left(\psi^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} The local support property says, therefore, that a point source at $x$ has zero impact (or minimal impact) on points $y$ that are far away from $x$.
+where $\left\langle \mathcal{A}, \delta_x \right\rangle \in L^2(\Omega)'$ denotes the result of applying the operator $\mathcal{A}$ to the distribution $\delta_x$, and $\left\langle \mathcal{A}, \delta_x \right\rangle^* \in L^2(\Omega)$ denotes the Riesz representation of $\langle\mathcal{A}, \delta_x\rangle$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\psi \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\psi^* \in L^2(\Omega)$ such that $\psi(v) = \left(\psi^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} The local support property says, therefore, that a point source at $x$ has zero impact (or minimal impact) on points $y$ that are far from $x$.
 
@@
 %We take the Riesz representative of the linear functional $\left\langle \mathcal{A}, \delta_x \right\rangle \in L^2(\Omega)'$ to convert it into a function $\left\langle \mathcal{A}, \delta_x \right\rangle^* \in L^2(\Omega)$.
-After discretization, the process of applying $\mathcal{A}$ to a distribution, $\mu$, then forming the Riesz representative of the resulting linear functional, amounts to computing $\mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\mu}$, where $\mathbf{A}$, and $\boldsymbol{\mu}$ are discretized versions of $A$ and $\delta_x$, respectively, and $\mathbf{M}$ is a mass matrix (e.g., in a finite element context).
+After discretization, the process of applying $\mathcal{A}$ to the distribution $\delta_x$, then forming the Riesz representative of the resulting linear functional, amounts to computing $\boldsymbol{\phi}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x$, where $\boldsymbol{\phi}_x$, $\mathbf{A}$, and $\boldsymbol{\delta}_x$ are discretized versions of $\phi_x$, $A$ and $\delta_x$, respectively, and $\mathbf{M}$ is a mass matrix (e.g., in a finite element context).
 
@@
 We may therefore post-process $\eta$ to recover $\phi_{x_i}$ for all of the points $x_i$ in the batch.
-Since $\varphi_i$ is a translated versions of $\phi_{x_i}$, we also recover $\varphi_i$ for all points $x_i$ in the batch---by applying $\mathcal{A}$ to one distribution, we compute many $\varphi_i$.
+Since $\varphi_i$ is a translated versions of $\phi_{x_i}$, we also recover $\varphi_i$ for all points $x_i$ in the batch.
 This is illustrated in Figure FIG.
@@
 \section*{Acknowledgments}
-We thank J.J. Alger for helpful discussions
+We thank J.J. Alger and Longfei Gao for helpful discussions.
 

### 2021-01-15 paper/localpsf.tex
more edits

--- 2021-01-15 paper/localpsf.tex
@@
 \end{equation}
-where $\left\langle \mathcal{A}, \delta_x \right\rangle \in L^2(\Omega)'$ denotes the result of applying the operator $\mathcal{A}$ to the distribution $\delta_x$, and $\left\langle \mathcal{A}, \delta_x \right\rangle^* \in L^2(\Omega)$ denotes the Riesz representation of $\langle\mathcal{A}, \delta_x\rangle$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\psi \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\psi^* \in L^2(\Omega)$ such that $\psi(v) = \left(\psi^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} The local support property says, therefore, that a point source at $x$ has zero impact (or minimal impact) on points $y$ that are far away from $x$.
-
-Note that while the domain of $\mathcal{A}$ is defined as $L^2(\Omega)$, which does not contain the delta distribution, the application of $\mathcal{A}$ is extended to distributions in a natural way, as described in Appendix \ref{app:distributions}.
+where $\left\langle \mathcal{A}, \delta_x \right\rangle \in L^2(\Omega)'$ denotes the result of applying the operator $\mathcal{A}$ to the distribution $\delta_x$, and $\left\langle \mathcal{A}, \delta_x \right\rangle^* \in L^2(\Omega)$ denotes the Riesz representation of $\langle\mathcal{A}, \delta_x\rangle$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\psi \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\psi^* \in L^2(\Omega)$ such that $\psi(v) = \left(\psi^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\mathcal{A}$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, the action of $\mathcal{A}$ may be extended to distributions in a natural way, as described in Appendix \ref{app:distributions}.
+%The action of $\mathcal{A}$ on a distribution $\mu$ is written as $\langle \mathcal{A}, \mu \rangle$.
+We use this angle bracket notation to distinguish the action of $\mathcal{A}$ on distributions from the action of $\mathcal{A}$ on functions.
 %We take the Riesz representative of the linear functional $\left\langle \mathcal{A}, \delta_x \right\rangle \in L^2(\Omega)'$ to convert it into a function $\left\langle \mathcal{A}, \delta_x \right\rangle^* \in L^2(\Omega)$.
-After discretization, the process of applying $\mathcal{A}$ to a delta distribution, then forming the Riesz representative, amounts to computing $\boldsymbol{\phi}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_{x_i}$, where $\boldsymbol{\phi}_x$, $\mathbf{A}$, and $\boldsymbol{\delta}_x$ are discretized versions of $\phi_x$, $A$, and $\delta_x$, respectively, and $\mathbf{M}$ is a mass matrix (e.g., in a finite element context).
+After discretization, the process of applying $\mathcal{A}$ to the distribution $\delta_x$, then forming the Riesz representative of the resulting linear functional, amounts to computing $\boldsymbol{\phi}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x$, where $\boldsymbol{\phi}_x$, $\mathbf{A}$, and $\boldsymbol{\delta}_x$ are discretized versions of $\phi_x$, $A$ and $\delta_x$, respectively, and $\mathbf{M}$ is a discretization of the $L^2$ inner product bilinear form (e.g., a mass matrix in the finite element context).
+From \eqref{eq:impulse_response_delta_action}, we see that the local support property means that the response of $\mathcal{A}$ to a point source at $x$ is zero (or small) at points $y$ that are far from $x$.
 
@@
 where $f \cdot g$ denotes pointwise multiplication, $f \ast g$ denotes convolution, and $\left(f, g\right)_{L^2(\Omega)}$ denotes the $L^2$ inner product on $\Omega$, for functions $f,g \in L^2(\Omega)$.
-Here we write $\varphi_i := \varphi_{x_i}$ (with $\varphi_{x_i}$ defined in \eqref{eq:convolution_kernel}) to denote local convolution kernels corresponding to a collection of points $\{x_i\}_{i=1}^r \subset \Omega$ scattered throughout the domain.
+To avoid nested subscripts, we write
+\begin{equation*}
+\varphi_i := \varphi_{x_i} \quad \text{and} \quad \phi_i := \phi_{x_i},
+\end{equation*}
+with $\varphi_{x_i}$ defined in \eqref{eq:convolution_kernel} and $\phi_{x_i}$ defined in \eqref{eq:impulse_response_defn}, to denote local convolution kernels and impulse responses corresponding to a collection of points $\{x_i\}_{i=1}^r \subset \Omega$ scattered throughout the domain.
 The functions $w_i$ are spatially varying weighting functions that are used to interpolate the convolution kernels $\varphi_i$.
@@
 
-We compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a Dirac comb, then post-processing the result. That is, we compute
-\begin{equation}
-\label{eq:eta_intro}
-\eta := \left\langle \mathcal{A}, \xi \right\rangle^*
-\end{equation}
-where 
-\begin{equation*}
-	\xi := \sum_{i=1}^l \delta_{x_i}
-\end{equation*}
-is the Dirac comb associated with point sources centered at a collection of scattered points $x_i$.
-The batch of points $x_i$ are chosen so that the support of $\phi_{x_i}$ and the support of $\phi_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$.
-
-By linearity, and the expression for $\phi_x$ in \eqref{eq:impulse_response_delta_action}, we have
-\begin{equation*}
-\eta = \sum_{i=1}^l \phi_{x_i}.
-\end{equation*}
-Since the support of $\phi_{x_i}$ and the support of $\phi_{x_j}$ do not overlap, we may post-process $\eta$ to recover $\phi_i$, and therefore $\varphi_i$, for all of the points $x_i$ in the batch---by applying $\mathcal{A}$ to one distribution, we recover many $\varphi_i$.
+We compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$.
+The batch of points $x_i$ are chosen so that the support of $\phi_i$ and the support of $\phi_j$ do not overlap (or do not overlap much) if $i \neq j$.
+Because these supports do not overlap, we can post-process the response of $\mathcal{A}$ to the Dirac comb to recover the functions $\phi_i$, and therefore the functions $\varphi_i$, associated with all points $x_i$ in the batch---with one application of $\mathcal{A}$, we recover many $\varphi_i$.
 This is illustrated in Figure FIG.
@@
 
-In order to choose the points $x_i$, we need to estimate the support of $\phi_{x_i}$ \emph{before} we compute it.
+%First, we compute
+%\begin{equation}
+%\label{eq:eta_intro}
+%\eta := \left\langle \mathcal{A}, \xi \right\rangle^*
+%\end{equation}
+%where 
+%\begin{equation*}
+%	\xi := \sum_{i=1}^l \delta_{x_i}
+%\end{equation*}
+%is the Dirac comb associated with point sources centered at a collection of scattered points $x_i$.
+The batch of points $x_i$ are chosen so that the support of $\phi_i$ and the support of $\phi_j$ do not overlap (or do not overlap much) if $i \neq j$.
+By linearity, and the expression for $\phi_x$ in \eqref{eq:impulse_response_delta_action}, we have
+%\begin{equation*}
+%\eta = \sum_{i=1}^l \phi_i.
+%\end{equation*}
+%We may therefore post-process $\eta$ 
+%
+%to recover $\phi_i$ for all of the points $x_i$ in the batch.
+Since the functions $\varphi_i$ are translated versions of the functions $\phi_i$, we also recover $\varphi_i$ for all points $x_i$ in the batch.
+This is illustrated in Figure FIG.
+The process is repeated to get more batches of $\varphi_i$'s, until a desired number of batches is reached.
+
+In order to choose the points $x_i$, we need to estimate the support of $\phi_i$ \emph{before} we compute it.
 The supports of the functions $\phi_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\mathcal{A}^T$ to a small number of specially chosen functions (three functions if $d=1$, six if $d=2$, ten if $d=3$), then post processing the results (Section \ref{eq:mean_and_covariance_estimation}).
@@
 \item We use a greedy algorithm to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}).
-\item We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\langle\mathcal{A},\delta_{x}\rangle^*$ (Section \ref{eq:mean_and_covariance_estimation}).
+\item We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\psi_x$ (Section \ref{eq:mean_and_covariance_estimation}). 
 \item For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
@@
 \label{eq:phi_b}
-	\eta^b = \left\langle\mathcal{A},\sum_{x_i \in S_b} \delta_{x_i}\right\rangle^* = \sum_{x_i \in S_b} \langle\mathcal{A}, \delta_{x_i}\rangle^*.
-\end{equation}
-Since the points $x_i$ are chosen so that the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ and the support of $\langle\mathcal{A},\delta_{x_j}\rangle^*$ do not overlap (or do not overlap much), we have
+	\eta^b = \left\langle\mathcal{A},\sum_{x_i \in S_b} \delta_{x_i}\right\rangle^* = \sum_{x_i \in S_b} \langle\mathcal{A}, \delta_{x_i}\rangle^* = \sum_{x_i \in S_b} \phi_x.
+\end{equation}
+Since the points $x_i$ are chosen so that the support of $\phi_i$ and the support of $\phi_j$ do not overlap (or do not overlap much), we have
 \begin{equation}
 \label{eq:varphi_eval}
-	\varphi_{x_i}(z) = \langle\mathcal{A}, \delta_{x_i}\rangle^*(z+x_i) = \begin{cases}
+	\varphi_{x_i}(z) = \phi_i(z+x_i) = \begin{cases}
 		\eta^b(z+x_i), & z+x_i \in E_{x_i} \\
@@
 	\label{thm:smoothest_interpolant}
-Let $y \in \mathbb{R}^d$, and let $u$ be the function
+Let $y \in \mathbb{R}^d$, and let $\gamma_y$ be the function
 \begin{equation}
 \label{eq:u_defn_in_thm}
-u(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right)
-\end{equation}
-where the weighting functions $w_i$ are the solutions to optimization problem \eqref{eq:wi_optimization_problem}. Then $u$ solves the optimization problem
+\gamma_y(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right)
+\end{equation}
+where the weighting functions $w_i$ are the solutions to optimization problem \eqref{eq:wi_optimization_problem}. Then $\gamma_y$ solves the optimization problem
 \begin{equation}
@@
 \begin{aligned}
-\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
-\text{such that} &\quad u(x_i) = \varphi_i(y), \quad i=1,\dots,q.
+\min_{\gamma_y \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \gamma_y}_{L^2(\Omega)}^2 \\
+\text{such that} &\quad \gamma_y(x_i) = \varphi_i(y), \quad i=1,\dots,q.
 \end{aligned}
@@
 	\begin{aligned}
-	\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
-	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r.
+	\min_{\gamma \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \gamma}_{L^2(\Omega)}^2 \\
+	\text{such that} &\quad \gamma(x_i) = f_i, \quad i=1,\dots,r.
 	\end{aligned}
@@
 	\begin{equation*}
-	\mathcal{T}(f) := u^*,
-	\end{equation*}
-	where $u^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
+	\mathcal{T}(f) := \gamma^*,
+	\end{equation*}
+	where $\gamma^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
 \end{lem}
@@
 	&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
-	&= u.
-	\end{align*}
-	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $u$ in \eqref{eq:u_defn_in_thm}.
+	&= \gamma_y.
+	\end{align*}
+	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $\gamma_y$ in \eqref{eq:u_defn_in_thm}.
 \end{proof}

### 2021-01-16 paper/localpsf.tex
more edits

@@
 The method for converting $\widetilde{\mathcal{A}}$ to H-matrix format is shown in Section \ref{sec:H_matrix_conversion}.
-In the rest of the section we will need the following definitions.
-
-\begin{defn}
-	We define the spatially varying volume, $\alpha$, the scaled impulse response, $\rho_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
-	\begin{align*}
-	\alpha:\Omega \rightarrow \mathbb{R}, &\qquad \alpha(x) := \int_{\Omega} \phi_x(z) dz, \\
-	\rho_x:\Omega \rightarrow \mathbb{R}, &\qquad \rho_x := \phi_x \big/ \alpha(x), \\
-	\mu:\Omega \rightarrow \mathbb{R}^d, &\qquad \mu(x) := \int_\Omega z \rho_x(z) dz, \\
-	\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\Sigma(x) := \int_\Omega (z - \mu(x))(z - \mu(x))^T \rho_x(z) dz.
-	\end{align*}
-\end{defn}
 
@@
 Since $\phi_x$ is nonnegative, it is a scaled probability distribution.
-We compute the scaling factor, denoted $\alpha(x)$, for all $x$ simultaneously by applying $\mathcal{A}$ to a constant function.
-We then compute the spatially varying mean, $\mu(x)$, and covariance, $\Sigma(x)$ of the normalized version of $\phi_x$, for all $x$ simultaneously, by applying $\mathcal{A}^T$ to a small number of linear and quadratic functions.
+We compute the scaling factor, denoted $\alpha(x)$, for all $x$ simultaneously by applying $\mathcal{A}^T$ to a constant function.
+We compute the spatially varying mean, $\mu(x)$, and covariance, $\Sigma(x)$ of the normalized version of $\phi_x$, for all $x$ simultaneously, by applying $\mathcal{A}^T$ to a small number of linear and quadratic functions.
+Definitions of $\alpha$, $\mu$, and $\Sigma$ are given in Definition \ref{defn:alpha_rho_mu_sigma}.
+The procedure for computing these quantities is shown in Algorithm \ref{alg:varhpi_mean_cov}, and the theoretical basis for this procedure is shown in Theorem \ref{thm:vol_mean_cov}.
+
 We make the approximation that the support of $\phi_x$ is contained within the ellipsoid
@@
 where $\tau$ is a small number of standard deviations. 
-The bigger $\tau$ is, the smaller the overlap between impulse responses within a batch will be; the smaller $\tau$ is, the more impulse responses we can fit within each batch.
+%The ellipsoid $E_{x}$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian approximation of the normalized version of $\langle \mathcal{A}, \delta_{x}\rangle^{*}$ which has the same mean and covariance.
+The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$, i.e., the Gaussian distribution which has the same mean and covariance as the normalized version of $\phi_x$ (see Figure FIG).
+The bigger $\tau$ is, the smaller the overlap between impulse responses within a batch will be, while the smaller $\tau$ is, the more impulse responses we can fit within each batch.
 In practice we found that $\tau=3$ provides a good balance between these competing goals.
-%The ellipsoid $E_{x}$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian approximation of the normalized version of $\langle \mathcal{A}, \delta_{x}\rangle^{*}$ which has the same mean and covariance.
-The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$.
-Note that this Gaussian distribution has the same mean and covariance as  which has the same mean and covariance as the normalized version of $\phi_x$ (see Figure FIG).
 The fraction of the mass of $\phi_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though we find this bound is conservative and far less mass resides in this region in practice.
+
+\begin{defn}
+	\label{defn:alpha_rho_mu_sigma}
+	We define the spatially varying volume, $\alpha$, the scaled impulse response, $\rho_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
+	\begin{align*}
+	\alpha:\Omega \rightarrow \mathbb{R}, &\qquad \alpha(x) := \int_{\Omega} \phi_x(z) dz, \\
+	\rho_x:\Omega \rightarrow \mathbb{R}, &\qquad \rho_x := \phi_x \big/ \alpha(x), \\
+	\mu:\Omega \rightarrow \mathbb{R}^d, &\qquad \mu(x) := \int_\Omega z \rho_x(z) dz, \\
+	\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\Sigma(x) := \int_\Omega (z - \mu(x))(z - \mu(x))^T \rho_x(z) dz.
+	\end{align*}
+\end{defn}
 

### 2021-01-16 paper/localpsf.tex
more edits

@@
 In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\varphi_i$ are computed with each matrix vector product.
-We do this by taking advantage of the fact that the functions $\varphi_i$ are locally supported; existing methods typically do not require this local support property.
+
+%We do this by taking advantage of the fact that the functions $\varphi_i$ are locally supported; existing methods typically do not require this local support property.
 
@@
 In this paper, rather than probing $\mathcal{A}^T$ with random noise functions, we use specific constant, linear, and quadratic functions.
-Our method estimates the support of $\varphi_p$ more accurately than resolution analysis, but our method requires that $\mathcal{A}$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
+We observe that our method estimates the support of $\varphi_p$ more accurately and reliably than resolution analysis, but our method requires that $\mathcal{A}$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
 
@@
 However, for complex geometries, problems are rarely discretized using regular grids, so we cannot easily use the fast Fourier transform.
-Furthermore, in big-data inverse problems the Hessian is typically ill-conditioned, so the required number of Krylov iterations is large.
+Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}.
 Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are amenable to fast linear algebra operations.
@@
 
-\section{Constructing the approximation}
+\section{Constructing the approximation, $\widetilde{\mathcal{A}}$}
 \label{sec:method}
@@
 We also describe how we evaluate kernel entries of the approximation, $\widetilde{A}(y,x)$, and how we convert $\widetilde{\mathcal{A}}$ to H-matrix format.
-\begin{itemize}
-\item We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}).
+
+We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}).
 To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as distinct functions.
-\item We use a greedy algorithm to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}).
-\item We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\psi_x$ (Section \ref{eq:mean_and_covariance_estimation}). 
-\item For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
-\end{itemize}
+A greedy algorithm is used to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}).
+We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\phi_x$ (Section \ref{eq:mean_and_covariance_estimation}).
+For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
+%\begin{itemize}
+%\item We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}).
+To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as distinct functions.
+%\item We use a greedy algorithm to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}).
+%\item We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\psi_x$ (Section \ref{eq:mean_and_covariance_estimation}). 
+%\item For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
+%\end{itemize}
 The complete algorithm for constructing $\widetilde{\mathcal{A}}$ is shown in Algorithm \ref{alg:construct_Atilde}.
@@
 Since $\phi_x$ is nonnegative, it is a scaled probability distribution.
-We compute the scaling factor, denoted $\alpha(x)$, for all $x$ simultaneously by applying $\mathcal{A}^T$ to a constant function.
+Let $\alpha(x)$ denote the scaling factor, and let $\mu(x)$ and $\Sigma(x)$ denote the mean and covariance of the normalized version of $\phi_x$ (see Definition \ref{defn:alpha_rho_mu_sigma}).
+We make the approximation that the support of $\phi_x$ is contained within the ellipsoid
+\begin{equation}
+\label{eq:support_ellipsoid}
+E_x := \{x' \in \Omega: (x' - \mu(x))^T \Sigma(x)^{-1} (x' - \mu(x)) \le \tau^2\},
+\end{equation}
+where $\tau$ is a fixed constant. 
+%The ellipsoid $E_{x}$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian approximation of the normalized version of $\langle \mathcal{A}, \delta_{x}\rangle^{*}$ which has the same mean and covariance.
+The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$, i.e., the Gaussian distribution which has the same mean and covariance as the normalized version of $\phi_x$ (see Figure FIG).
+The bigger $\tau$ is, the smaller the overlap between impulse responses within a batch will be, while the smaller $\tau$ is, the more impulse responses we can fit within each batch.
+In practice we find that $\tau=3$ provides a good balance between these competing goals.
+The fraction of the ``mass'' of $\phi_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though we find this bound is conservative and far less mass resides in this region in practice.
+
+In Algorithm \ref{alg:varhpi_mean_cov} we show how to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$ for all points $x$ simultaneously, by applying $\mathcal{A}^T$ to a constant function and a small number of linear and quadratic functions.
+The theoretical basis for Algorithm \ref{alg:varhpi_mean_cov} is shown in Theorem \ref{thm:vol_mean_cov}.
+
+%We compute the scaling factor, denoted $\alpha(x)$, for all $x$ simultaneously by applying $\mathcal{A}^T$ to a constant function.
 We compute the spatially varying mean, $\mu(x)$, and covariance, $\Sigma(x)$ of the normalized version of $\phi_x$, for all $x$ simultaneously, by applying $\mathcal{A}^T$ to a small number of linear and quadratic functions.
@@
 The procedure for computing these quantities is shown in Algorithm \ref{alg:varhpi_mean_cov}, and the theoretical basis for this procedure is shown in Theorem \ref{thm:vol_mean_cov}.
-
-We make the approximation that the support of $\phi_x$ is contained within the ellipsoid
-\begin{equation}
-\label{eq:support_ellipsoid}
-E_x := \{x' \in \Omega: (x' - \mu(x))^T \Sigma(x)^{-1} (x' - \mu(x)) \le \tau^2\},
-\end{equation}
-where $\tau$ is a small number of standard deviations. 
-%The ellipsoid $E_{x}$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian approximation of the normalized version of $\langle \mathcal{A}, \delta_{x}\rangle^{*}$ which has the same mean and covariance.
-The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$, i.e., the Gaussian distribution which has the same mean and covariance as the normalized version of $\phi_x$ (see Figure FIG).
-The bigger $\tau$ is, the smaller the overlap between impulse responses within a batch will be, while the smaller $\tau$ is, the more impulse responses we can fit within each batch.
-In practice we found that $\tau=3$ provides a good balance between these competing goals.
-The fraction of the mass of $\phi_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though we find this bound is conservative and far less mass resides in this region in practice.
 

### 2021-01-16 paper/localpsf.tex
more edits

@@
 	\label{defn:alpha_rho_mu_sigma}
-	We define the spatially varying volume, $\alpha$, the scaled impulse response, $\rho_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
+	We define the spatially scaling factor, $\alpha$, the scaled impulse response, $\rho_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
 	\begin{align*}

### 2021-01-19 paper/localpsf.tex
remove redundancy between mean cov theorem and algorithm

@@
 Since $\phi_x$ is nonnegative, it is a scaled probability distribution.
-Let $\alpha(x)$ denote the scaling factor, and let $\mu(x)$ and $\Sigma(x)$ denote the mean and covariance of the normalized version of $\phi_x$ (see Definition \ref{defn:alpha_rho_mu_sigma}).
+Let $\widehat{\phi}_x$ denote the normalized version of $\phi_x$, and let $\mu(x)$ and $\Sigma(x)$ denote the mean and covariance of $\widehat{\phi}_x$, respectively (see Definition \ref{defn:alpha_rho_mu_sigma}).
 We make the approximation that the support of $\phi_x$ is contained within the ellipsoid
@@
 %The ellipsoid $E_{x}$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian approximation of the normalized version of $\langle \mathcal{A}, \delta_{x}\rangle^{*}$ which has the same mean and covariance.
-The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$, i.e., the Gaussian distribution which has the same mean and covariance as the normalized version of $\phi_x$ (see Figure FIG).
+The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\phi}_x$ (see Figure FIG).
 The bigger $\tau$ is, the smaller the overlap between impulse responses within a batch will be, while the smaller $\tau$ is, the more impulse responses we can fit within each batch.
@@
 In Algorithm \ref{alg:varhpi_mean_cov} we show how to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$ for all points $x$ simultaneously, by applying $\mathcal{A}^T$ to a constant function and a small number of linear and quadratic functions.
+Definitions required for Algorithm \ref{alg:varhpi_mean_cov} are given in Definition \ref{defn:alpha_rho_mu_sigma}, Definition \ref{defn:C_L_Q}, and Definition \ref{defn:pointwise}.
 The theoretical basis for Algorithm \ref{alg:varhpi_mean_cov} is shown in Theorem \ref{thm:vol_mean_cov}.
@@
 
-\begin{defn}
+\begin{defn}[Moments of $\phi_x$]
 	\label{defn:alpha_rho_mu_sigma}
-	We define the spatially scaling factor, $\alpha$, the scaled impulse response, $\rho_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
+	We define the spatially scaling factor, $\alpha$, the scaled impulse response, $\widehat{\phi}_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
 	\begin{align*}
 	\alpha:\Omega \rightarrow \mathbb{R}, &\qquad \alpha(x) := \int_{\Omega} \phi_x(z) dz, \\
-	\rho_x:\Omega \rightarrow \mathbb{R}, &\qquad \rho_x := \phi_x \big/ \alpha(x), \\
-	\mu:\Omega \rightarrow \mathbb{R}^d, &\qquad \mu(x) := \int_\Omega z \rho_x(z) dz, \\
-	\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\Sigma(x) := \int_\Omega (z - \mu(x))(z - \mu(x))^T \rho_x(z) dz.
+	\widehat{\phi}_x:\Omega \rightarrow \mathbb{R}, &\qquad \widehat{\phi}_x := \phi_x \big/ \alpha(x), \\
+	\mu:\Omega \rightarrow \mathbb{R}^d, &\qquad \mu(x) := \int_\Omega z \widehat{\phi}_x(z) dz, \\
+	\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\Sigma(x) := \int_\Omega (z - \mu(x))(z - \mu(x))^T \widehat{\phi}_x(z) dz.
 	\end{align*}
@@
 
-\begin{theorem}
-	\label{thm:vol_mean_cov}
-	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$, and let $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions, respectively:
+\begin{defn}[Constant, linear, and quadratic functions]	
+	\label{defn:C_L_Q}
+	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$.
+We define $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions, respectively:
 	\begin{equation*}
@@
 	\end{equation*}
-	We have
-	\begin{align}
-		\alpha =& \left(\mathcal{A}^TC\right)^*, \label{eq:vol_mean_var_thm1}\\
-		\mu^i =& \left(\mathcal{A}^T L^i\right)^* / \alpha, \label{eq:vol_mean_var_thm2}\\
-		\Sigma^{ij} =& \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot\mu^j, \label{eq:vol_mean_var_thm3}
-	\end{align}
+\end{defn}
+
+\begin{defn}[Pointwise operations]
+	\label{defn:pointwise}
 	We write $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
-\end{theorem}
-
-The proof of Theorem \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}.
+\end{defn}
+
+%\begin{theorem}
+%	\label{thm:vol_mean_cov}
+%	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$, and let $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions, respectively:
+%	\begin{equation*}
+%		C(x) := 1, \qquad
+%		L^i(x) := x^i, \qquad
+%		Q^{ij}(x) = x^i x^j.
+%	\end{equation*}
+%	We have
+%	\begin{align}
+%		\alpha =& \left(\mathcal{A}^TC\right)^*, \label{eq:vol_mean_var_thm1}\\
+%		\mu^i =& \left(\mathcal{A}^T L^i\right)^* / \alpha, \label{eq:vol_mean_var_thm2}\\
+%		\Sigma^{ij} =& \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot\mu^j, \label{eq:vol_mean_var_thm3}
+%	\end{align}
+%	We write $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
+%\end{theorem}
+%
+%The proof of Theorem \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}.
 
@@
 \end{algorithm2e}
+
+\begin{theorem}[Correctness of Algorithm \ref{alg:varhpi_mean_cov}]
+	\label{thm:vol_mean_cov}
+	Algorithm \ref{alg:varhpi_mean_cov} correctly computes $\alpha$, $\mu$, and $\Sigma$.
+\end{theorem}
+
+The proof of Theorem \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}.
 
@@
 	\end{align*}
-	which implies \eqref{eq:vol_mean_var_thm1}.
+	which implies $\alpha = \left(\mathcal{A}^TC\right)^*$.
 
@@
 	\end{align*}
-	which implies \eqref{eq:vol_mean_var_thm2}.
-
-	To prove \eqref{eq:vol_mean_var_thm3} we use the identity
-	\begin{equation*}
-		\Sigma(x) = \int_\Omega zz^T \rho_x(z) dz - \mu(x) \mu(x)^T.
-	\end{equation*}
-	We have
+	which implies $\mu^i = \left(\mathcal{A}^T L^i\right)^* / \alpha$.
+
+	Using the identity
+	\begin{equation*}
+		\Sigma(x) = \int_\Omega zz^T \widehat{\phi}_x(z) dz - \mu(x) \mu(x)^T,
+	\end{equation*}
+	we have
 	\begin{align*}
@@
 	\end{align*}
-	which implies \eqref{eq:vol_mean_var_thm3}.
+	which implies $\Sigma^{ij} = \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot\mu^j$.
 

### 2021-01-19 paper/localpsf.tex
edits

@@
 	\label{defn:alpha_rho_mu_sigma}
-	We define the spatially scaling factor, $\alpha$, the scaled impulse response, $\widehat{\phi}_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
+	We define the spatially varying scaling factor, $\alpha$, the scaled impulse response, $\widehat{\phi}_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
 	\begin{align*}
@@
 where $v := \Phi^T\left(\mu_p - \mu_q\right)$.
-We compute the generalized eigenvalue decomposition of $\Sigma_p$ and $\Sigma_q$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's algorithm (any fast 1 dimensional convex optimization routine may be used).
+We compute the generalized eigenvalue decomposition of $\Sigma_p$ and $\Sigma_q$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's algorithm \cite{BRENT} (any fast 1 dimensional convex optimization routine may be used).
 The resulting algorithm for checking whether $E_p$ and $E_q$ intersect is summarized in Algorithm \ref{alg:ellipsoid_intersection_test}.

### 2021-01-21 paper/localpsf.tex
edits
add changes based on meeting today

@@
 	{	
+		\Input{Operator $\mathcal{A}$, parameter $\text{max\_batches}$}
+		\Output{Dirac comb responses $\{\eta^b\}_{b=1}^{\text{max\_batches}}$, \\
+			weighting functions $\{w_i\}_{i=1}^r$}
+
 		Compute spatially varying mean, $\mu$, and covariance, $\Sigma$, using Algorithm \ref{alg:varhpi_mean_cov}.
@@
 	{	
+		\Input{Operator $\mathcal{A}$}
+		\Output{$\alpha$, $\mu$, and $\Sigma$}
+
+		\tcp{Compute scaling factor $\alpha$}
+
 		Form constant function $C(x)=1$
@@
 
+		\tcp{Compute mean $\mu$}
 		\For{$i=1,2,\dots,d$}{
@@
 		}
+		\tcp{Compute covariance $\Sigma$}
 		\For{$i=1,2,\dots,d$}{
@@
 	}
-	\caption{Compute volume $\alpha$, mean $\mu$, and covariance $\Sigma$}
+	\caption{Compute scaling factor $\alpha$, mean $\mu$, and covariance $\Sigma$}
 	\label{alg:varhpi_mean_cov}
@@
 	\SetKwInOut{Input}{Input}
+	\SetKwInOut{Output}{Output}
 	\SetKwProg{Fn}{Function}{}{}
 
-	Start with a finite set of candidate points $P \subset \Omega$
-
-	Initialize new empty batch of sample points, $S_b$
+	\Input{Finite set of candidate points $P\subset \Omega$, \\mean $\mu(x)$ and covariance $\Sigma(x)$, \\domain boundary information $\partial \Omega$, \\previous sample point batches $S_1, \dots, S_{b-1}$}
+	\Output{Batch of new sample points $S_b$}
+
+	Initialize empty new batch of sample points, $S_b = \{\}$
 
@@
 		\If{$E_p \cap E_q \neq \{\}$ for all $q \in S_b$ and $E_p \cap \partial \Omega = \{\}$}{
+			\tcp{$E_p$ and $E_q$ are the ellipsoids defined in \eqref{eq:support_ellipsoid}}
 
@@
 
-			Remove all points $x$ satisfying $\mu(x) \in E_p$ from $P$
+			Remove all points $p'$ satisfying $\mu(p') \in E_p$ from $P$
 		}
@@
 By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover $\varphi_{x_i}$ for every point $x_i \in S_b$.
-To minimize memory usage, the functions $\varphi_i$ are never isolated and stored individually.
+To minimize memory usage, the functions $\varphi_i$ are not stored individually.
 Instead, we store the functions $\eta^b$, and use formula \eqref{eq:varphi_eval} to perform function evaluations $z \mapsto \varphi_i(z)$ as needed.
@@
 	{	
+		\Input{Sample points $\{x_i\}_{i=1}^r$, domain geometry $\Omega$}
+		\Output{Weighting functions $\{w_i\}_{i=1}^r$}
+
+		Initialize constant function $C(x) = 1$
+
+		Compute Lebesgue measure of the domain, $|\Omega|$
+
+		\tcp{Compute harmonic functions $\{\psi_i\}_{i=1}^r$ and $\{\theta_i\}_{i=1}^r$}
+
 		\For{$i=1,2, \dots, r$}{
-			Compute $\psi_i$ as the solution to the Poisson problem
+			Initialize $\delta_{x_i}$ as the point source (delta distribution) centered at $x_i$.
+
+			Compute $\psi_i$ by solving the Neumann Poisson problem
 			\begin{equation}
 			\label{eq:psi_eq}
-			\Delta \psi_i = \delta_{x_i} - C/|\Omega|, \quad \text{in }\Omega,
-			\end{equation}
-			with Neumann zero boundary conditions, $\nu \cdot \nabla \psi_i = 0$ on $\partial \Omega$, and the average value zero condition, $\int_\Omega \psi_i(x) dx = 0$.
-Recall $\delta_{x_i}$ is a point source (delta distribution) centered at $x_i$, $C$ is the constant function $C(x)=1$, and $|\Omega|$ is the Lebesgue measure of $\Omega$.
-
-			Compute $\theta_i$ as the solution to the Poisson problem
+			\begin{cases}
+			\Delta \psi_i = \delta_{x_i} - C/|\Omega|, & \text{in }\Omega, \\
+			\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
+			\end{cases}
+			\end{equation}
+			with condition $\int_\Omega \psi_i(x) dx = 0$, where $\nu$ is the normal vector to $\partial \Omega$.
+
+			Compute $\theta_i$ by solving the Neumann Poisson problem
 			\begin{equation}
 			\label{eq:theta_eq}
-			\Delta \theta_i = \psi_i, \quad \text{in }\Omega,
-			\end{equation}
-			with Neumann zero boundary conditions, $\nu \cdot \nabla \psi_i = 0$ on $\partial \Omega$, and the average value zero condition, $\int_\Omega \psi_i(x) dx = 0$.
+						\begin{cases}
+			\Delta \theta_i = \psi_i, & \text{in }\Omega, \\
+			\nu \cdot \nabla \theta_i = 0 & \text{on } \partial \Omega,
+			\end{cases}
+			\end{equation}
+			with condition $\int_\Omega \theta_i(x) dx = 0$, where $\nu$ is the normal vector to $\partial \Omega$.
 
@@
 
+		\tcp{Compute weighting functions $\{w_i\}_{i=1}^r$}
+
 		\For{$i=1,2,\dots,r$}{
@@
 			\end{equation}
-			where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one, and $e_i = (0,\dots,0,1,0,\dots,0)$ is the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.
+			where $\mathbf{1} = (1,1,\dots,1) \in \mathbb{R}^r$, and $e_i = (0,\dots,0,1,0,\dots,0)\in \mathbb{R}^r$ is unit vector with $i^\text{th}$ component equal to one.
 
@@
 	\SetAlgoNoLine
+	\SetKwInOut{Input}{Input}
+	\SetKwInOut{Output}{Output}
+
+	\Input{points $x \in \Omega$, $y \in \Omega$, \\sample point batches $S_1, \dots, S_\text{max\_batches}$, \\Dirac comb responses $\{\eta^b\}_{b=1}^{\text{max\_batches}}$, \\weighting functions $\{w_i\}_{i=1}^r$}
+	\Output{Kernel entry $z = \widetilde{A}(y,x)$}
 
@@
 
-	\Output{Boolean which is true if $E_p \cap E_q \neq \{\}$ and false otherwise}
-
-	\mbox{}\\
-
-	\Fn{$\ellipsoidsintersect{E_p}{E_q}$}
-	{
+	\Output{Boolean $\text{ellipsoids\_intersect}$ which is true if $E_p \cap E_q \neq \{\}$ and false otherwise}
+
 	Solve generalized eigenvalue problem $\Sigma_p \Phi = \Sigma_q \Phi \Lambda$
@@
 
-			\Return{False}
+		$\text{ellipsoids\_intersect} \gets \text{False}$
 
@@
 
-			\Return{True}
-
-		}
+		$\text{ellipsoids\_intersect} \gets \text{True}$
+
 	}

### 2021-01-25 paper/localpsf.tex
edits from discussion with Noemi and Tucker just now

@@
 
+Now that we have estimates of the supports of the functions $\phi_x$, we use these estimates, combined with a greedy algorithm to choose sets of sample points, $S_b$.
+These sample points will be used in Section \ref{sec:get_impulse_response} to form Dirac combs, which are then used to compute impulse responses of $\mathcal{A}$.
 Starting from a finite set of candidate points $P$, we select batches of sample points, $S_b$, using a greedy algorithm.
@@
 Otherwise we discard $p$.
+Recall that $E_p$ and $E_q$ are ellipsoids defined in \eqref{eq:support_ellipsoid}.
 This process repeats until there are no more points in $P$.
@@
 
-The \emph{Dirac comb}, $\xi_b$, associated with a batch of sample points $S_b$ is the sum of Dirac distributions (point sources) centered at the points $x_i \in S_b$. That is,
+We compute impulse responses of $\mathcal{A}$ in batches by applying $\mathcal{A}$ to a small number Dirac combs.
+The batches of impulse responses, $\eta^b$, will ultimately be interpolated with weighting functions described in Section \ref{sec:weighting_functions} to form the desired approximation $\widetilde{\mathcal{A}}\approx \mathcal{A}$.
+The Dirac comb, $\xi_b$, associated with a batch of sample points $S_b$ is the sum of Dirac distributions (point sources) centered at the points $x_i \in S_b$.
+That is,
 \begin{equation*}
@@
 \end{equation}
-Since the points $x_i$ are chosen so that the support of $\phi_i$ and the support of $\phi_j$ do not overlap (or do not overlap much), we have
+Since the points $x_i$ are chosen so that the support of $\phi_i$ and the support of $\phi_j$ do not overlap (or do not overlap much) when $i \neq j$, from \eqref{eq:convolution_kernel} and \eqref{eq:phi_b} we have
 \begin{equation}
@@
 \label{eq:u_defn_in_thm}
-\gamma_y(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right)
+\gamma_y(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right),
 \end{equation}

### 2021-01-26 paper/localpsf.tex
edits to discretization appendix

@@
 
+\section{Background: distributions}
+\label{app:distributions}
+
+Let $\overline{\Omega}$ be the closure of $\Omega$, and let $C(\overline{\Omega})$ be the space of continuous functions mapping $\overline{\Omega}\rightarrow \mathbb{R}$.
+The action of $\mathcal{A}$ is extended to distributions $\mu:C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ via the formula
+\begin{equation*}
+\langle\mathcal{A},\mu\rangle(w) := \int_\Omega w(y) \mu\left(A(y, \cdot)\right) dy, 
+\end{equation*}
+where $A(y,\cdot)$ is the function $x \mapsto A(y,x)$. This is derived formally as follows:
+\begin{align*}
+\langle\mathcal{A},\mu\rangle(w) &= \int_\Omega \int_\Omega w(y) A(y,x) \textrm{``} \mu(x) \textrm{''} dx dy \\
+&= \int_\Omega w(y) \int_\Omega A(y,x) \textrm{``} \mu(x) \textrm{''} dx \\
+&= \int_\Omega w(y) \mu\left(A(y,\cdot)\right) dx.
+\end{align*}
+For example, the delta distribution $\delta_x$ is defined by $\delta_x(v) = v(x)$, and the action of $\mathcal{A}$ on $\delta_x$ is given by
+\begin{equation}
+\label{eq:appendix_impulse_response}
+\langle\mathcal{A},\delta_x\rangle(w)	= \int_\Omega w(y) A(y,x) dy.
+\end{equation}
+Recall that the impulse response of $\mathcal{A}$ to $\delta_x$, denoted $\phi_x$, is the Riesz representation of $\langle\mathcal{A},\delta_x\rangle$.
+From \eqref{eq:appendix_impulse_response} and the definition of the Riesz representation, we have $\phi_x = \langle\mathcal{A},\delta_x\rangle^* = A(~\cdot~,x)$.
+That is, $\phi_x$ is the function $y \mapsto A(y,x)$.
+
 \section{Discretization}
@@
 
-Let $\phi_i$, $i=1,\dots,N$ be a set of finite element basis functions, and let
-\begin{equation*}
-V_h = \Span\left(\phi_1, \phi_2, \dots, \phi_N\right).
-\end{equation*}
-be the corresponding  finite element space approximating $L^2(\Omega)$.
-Here $V_h$ inherits the $L^2$ inner product, and $h$ denotes the mesh size parameter.
+Let $f_i$, $i=1,\dots,N$ be a set of finite element basis functions, and let
+\begin{equation*}
+V_h = \Span\left(f_1, f_2, \dots, f_N\right).
+\end{equation*}
+be the corresponding  finite element space with the $L^2$ inner product.
+Here $h$ denotes the mesh size parameter.
 Functions $u\in L^2(\Omega)$ are approximated by functions $u_h \in V_h$.
@@
 That is,
-\begin{equation*}
-u(x) \approx u_h(x) = \sum_{i=1}^N \mathbf{u}_i \phi_i(x).
-\end{equation*}
-We say that the coefficient arrays $\mathbf{u}$ reside $\mathbb{R}^N_\mathbf{M}$, which is the space $\mathbb{R}^N$ with the matrix-weighted inner product $\left(\mathbf{u},\mathbf{v}\right)_\mathbf{M} := \left(\mathbf{M} \mathbf{u}\right)^T \mathbf{v}$, where
-$\mathbf{M} \in \mathbb{R}^{N \times N}$ is the \emph{mass matrix} which has entries
-\begin{equation*}
-\mathbf{M}_{ij} = \left(\phi_i, \phi_j\right)_{L^2(\Omega)} = \int_{\Omega} \phi_i(x) \phi_j(x) dx.
-\end{equation*}
-Direct calculation shows that $V_h$ is isometrically isomorphic $\mathbb{R}^N_\mathbf{M}$, in the sense that functions $u_h \in V_h$ are in bijective correspondence to coefficient vectors $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, and $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u},\mathbf{v}\right)_\mathbf{M}$.
-Likewise, the dual space of continuous linear functionals $\psi_h:V_h \rightarrow \mathbb{R}$, denoted $V_h'$, is isometrically isomorphic to $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
-All operations on functions in $V_h$ and linear functionals in $V_h'$ may therefore be performed on coefficient vectors in $\mathbb{R}^N_\mathbf{M}$ and coefficient dual vectors in $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
-We summarize coefficient vector versions of the function space operation that are used in this paper in Appendix \ref{app:discretized_operations}.
+\begin{equation}
+\label{eq:finite_bijection}
+u_h(x) = \sum_{i=1}^N \mathbf{u}_i f_i(x).
+\end{equation}
+
+Let $\mathbf{M} \in \mathbb{R}^{N \times N}$ denote the finite element \emph{mass matrix}, which has entries
+\begin{equation*}
+\mathbf{M}_{ij} = \left(f_i, f_j\right)_{L^2(\Omega)} = \int_{\Omega} f_i(x) f_j(x) dx,
+\end{equation*}
+and let $\mathbb{R}^N_\mathbf{M}$ denote the space $\mathbb{R}^N$ with the matrix-weighted inner product
+\begin{equation*}
+\left(\mathbf{u},\mathbf{v}\right)_\mathbf{M} := \left(\mathbf{M} \mathbf{u}\right)^T \mathbf{v}.
+\end{equation*}
+Direct calculation shows that $V_h$ is isometrically isomorphic $\mathbb{R}^N_\mathbf{M}$, in the sense that \eqref{eq:finite_bijection} establishes a bijective correspondence between functions $u_h \in V_h$ and coefficient vectors $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, and that this bijection preserves the inner product; i.e., $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u},\mathbf{v}\right)_\mathbf{M}$.
+Likewise, the dual space, $V_h'$, is isometrically isomorphic to $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
+These isomorphisms allow us to identify functions $u_h \in V_h$ with their corresponding coefficient arrays $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, and linear functionals $\sigma_h \in V_h'$ with their corresponding coefficient arrays $\boldsymbol{\sigma} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$.
 
@@
 
-As described in Section \ref{sec:discretization}, a function $u_h$ in a finite element space $V_h$ may be written in terms of its vector of coefficients, $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, with respect the finite element basis $\{\phi_i\}_{i=1}^N$.
+Here we list common function space operations on $V_h$ and $V_h'$, and show the corresponding matrix and vector operations on $\mathbb{R}^N_{\mathbf{M}}$ and $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
+
+%We summarize coefficient vector versions of the function space operation that are used in this paper in Appendix \ref{app:discretized_operations}.
+%
+%As described in Section \ref{sec:discretization}, a function $u_h$ in a finite element space $V_h$ may be written in terms of its vector of coefficients, $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, with respect the finite element basis $\{\phi_i\}_{i=1}^N$.
 Recall that $\mathbf{M}$ is the mass matrix.
 We have
-\begin{equation*}
-	u(x) \approx u_h(x) = \sum_{i=1}^N \mathbf{u}_i \phi_i(x).
-\end{equation*}
-The finite element function space $V_h$ is isometrically isomorphic to the coefficient space $\mathbb{R}^N_\mathbf{M}$, and the dual space $V_h'$ is isometrically isomorphic to $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
+%\begin{equation*}
+%	u(x) \approx u_h(x) = \sum_{i=1}^N \mathbf{u}_i \phi_i(x).
+%\end{equation*}
+%The finite element function space $V_h$ is isometrically isomorphic to the coefficient space $\mathbb{R}^N_\mathbf{M}$, and the dual space $V_h'$ is isometrically isomorphic to $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
 In computations, we perform operations with coefficient vectors in $\mathbb{R}^N_\mathbf{M}$ and coefficient dual vectors in $\mathbb{R}^N_{\mathbf{M}^{-1}}$, rather than functions in $V_h$ and functionals in $V_h'$.
@@
 \begin{description}
-	\item[Coefficients of a functional:] A linear functional $\psi_h \in V_h'$ has coefficient dual vector $\boldsymbol{\psi} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$, which has entries
-	\begin{equation*}
-		\boldsymbol{\psi}_i = \psi_h(\phi_i), \quad i=1,\dots,N.
-	\end{equation*}
-	The action of a linear functional $\psi_h \in V_h'$ on a function $u_h \in V_h$ is given by
-	\begin{equation*}
-		\psi_h(v_h) = \mathbf{v}^T \boldsymbol{\psi}.
-	\end{equation*}
-	\item[Riesz representation of a functional:] Let $\psi_h \in V_h'$. We have
-	\begin{equation*} 
-		 \psi_h^* \in V_h: \quad \left(\psi_h^*, v_h\right)_{L^2(\Omega)} = \psi_h(v_h) \quad\forall v_h \in V_h \quad \Leftrightarrow \quad \boldsymbol{\psi}^* = \mathbf{M}^{-1} \boldsymbol{\psi}.
-	\end{equation*}
-	The function $\psi_h^*$ is the Riesz representation of the functional $\psi_h$, and $\boldsymbol{\psi}^* \in \mathbb{R}^N_\mathbf{M}$ is the coefficient vector of $\psi_h^*$.
-	\item[$L^2$ projection:] Let $f \in L^2(\Omega)$, and define $\mathbf{f}^* \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ to be the vector with entries
-	\begin{equation*}
-		\mathbf{f}^*_i = \int_\Omega f(x) \phi_i(x) dx, \quad i=1,\dots,N.
-	\end{equation*}
-	We have
-	\begin{equation*}
-		f_h \in V_h, \quad \left(f_h, v_h\right)_{L^2(\Omega)} = \left(f, v_h\right)_{L^2(\Omega)} \quad\forall v_h \in V_h \quad \Leftrightarrow \quad \mathbf{f} = \mathbf{M}^{-1}\mathbf{f}^*.	
-	\end{equation*}
-	The function $f_h$ is the $L^2$ projection of $f$ onto $V_h$, and $\mathbf{f} \in \mathbb{R}^N_\mathbf{M}$ is the coefficient vector of $f_h$.
+	\item[Coefficients of a functional:] The coefficient dual vector, $\boldsymbol{\sigma} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$, corresponding to a linear functional $\sigma_h \in V_h'$ has components given by
+	\begin{equation*}
+		\boldsymbol{\sigma}_i = \sigma_h(f_i), \quad i=1,\dots,N.
+	\end{equation*}
+	\item[Action of a functional:] The action of a linear functional $\sigma_h \in V_h'$ on a function $u_h \in V_h$ may be computed as follows:
+	\begin{equation*}
+		\sigma_h(v_h) = \mathbf{u}^T \boldsymbol{\sigma}.
+	\end{equation*}
+	\item[Riesz representation:] Let $\sigma_h \in V_h'$. The Riesz representation of $\sigma_h$ is the unique  function $\sigma_h^* \in V_h$ satisfying 
+	\begin{equation*} 
+	\left(\sigma_h^*, v_h\right)_{L^2(\Omega)} = \sigma_h(v_h) \quad\forall v_h \in V_h.
+	\end{equation*}
+	The coefficient dual vector $\boldsymbol{\sigma}^*$ corresponding to $\sigma_h$ is given by
+	\begin{equation*} 
+		\boldsymbol{\sigma}^* = \mathbf{M}^{-1} \boldsymbol{\sigma}.
+	\end{equation*}
+	\item[Finite element $L^2$ projection:] Let $g \in L^2(\Omega)$.
+The orthogonal projection of $g$ onto $V_h$ with respect to the $L^2$ inner product is the unique function $g^\pi_h \in V_h$ satisfying
+	\begin{equation*}
+		\left(g^\pi_h, v_h\right)_{L^2(\Omega)} = \left(g, v_h\right)_{L^2(\Omega)} \quad\forall v_h \in V_h.
+	\end{equation*}
+	 Let $\left(\mathbf{g}^\pi\right)^* \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ be the vector with components
+	\begin{equation*}
+		\left(\mathbf{g}^\pi\right)^*_i = \int_\Omega g(x) f_i(x) dx, \quad i=1,\dots,N.
+	\end{equation*}
+	The coefficient vector $\mathbf{g}^\pi \in \mathbb{R}^N_{\mathbf{M}}$ corresponding to $g^\pi_h$ is given by
+	\begin{equation*}
+		\mathbf{g}^\pi = \mathbf{M}^{-1}\left(\mathbf{g}^\pi\right)^*.
+	\end{equation*}
+	\item[Finite element interpolation:] Let $p_i \in \mathbb{R}^d$ be the Lagrange node associated with the finite element basis function $f_i$.
+The finite element interpolant of a function $g \in C\left(\overline{\Omega}\right)$ is the function $g^I \in V_h$ given by
+	\begin{equation*}
+		g_h^I(x) = \sum_{i=1}^N g(p_i) f_i(x).
+	\end{equation*}
+	The coefficient vector, $\mathbf{g}$, for $g_h^I$ has components 
+	\begin{equation*}
+	\mathbf{g}_i^I = g(p_i), \quad i=1.,\dots,N.
+	\end{equation*}
+	\item[Projection vs. interpolation:]
+	Typically $g^\pi$ is a more accurate approximation of $g$ than $g^I$.
+However, $g^I$ is typically cheaper to compute than $g^\pi$ because computing $g^\pi$ requires solving a linear system, while computing $g^I$ does not.
+Interpolation is typically sufficiently accurate for practical purposes so long as the finite element basis satisfies the following Kronecker property:
+	\begin{equation*}
+		f_i(p_j) = \begin{cases}1 & i=j \\
+		0 & \text{otherwise}.\end{cases}
+	\end{equation*}
 	\item[Matrix representation of an operator:] Let $\mathcal{B}_h : V_h \rightarrow V_h'$ be a linear operator, and let $\mathbf{B} \in \mathbf{R}^{N \times N}$ be the matrix with entries
 	\begin{equation*}
-		\mathbf{B}_{ij} = \left(\mathcal{B}_h \phi_j\right)(\phi_i).
+		\mathbf{B}_{ij} = \left(\mathcal{B}_h f_j\right)(f_i).
 	\end{equation*}
@@
 	\begin{equation*}
-		\psi_h = \mathcal{B}_h u_h \quad \Leftrightarrow \quad \boldsymbol{\psi} = \mathbf{B} \mathbf{u}.
+		\sigma_h = \mathcal{B}_h u_h \quad \Leftrightarrow \quad \boldsymbol{\sigma} = \mathbf{B} \mathbf{u}.
 	\end{equation*}
@@
 	\begin{equation*}
-		\psi_h = \mathcal{B}_h^T u_h \quad \Leftrightarrow \quad \boldsymbol{\psi} = \mathbf{B}^T \mathbf{u}.
+		\sigma_h = \mathcal{B}_h^T u_h \quad \Leftrightarrow \quad \boldsymbol{\sigma} = \mathbf{B}^T \mathbf{u}.
 	\end{equation*}
@@
 	\item[Distributions:] Let $\mu:C\left(\overline{\Omega}\right)\rightarrow \mathbb{R}$ be a distribution.
-If $\mu(\phi_i)$ is well-defined for all basis vectors $\phi_i$, then the restriction of $\mu$ to domain $V_h$ is a linear functional $\mu_h\in V_h'$.
-The functional $\mu_h$ has a coefficient dual vector $\boldsymbol{\mu} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ with entries $\boldsymbol{\mu}_i=\mu(\phi_i)$ for $i=1,\dots,N$.
-	\item[Action of an operator on a distribution:] Since $V_h$ is finite-dimensional, the restriction, $\mu_h \in V_h$, of a distribution, $\mu$, has a a Riesz representation $\mu_h^* \in V_h$.
+If $\mu(f_i)$ is well-defined for all finite element basis vectors $f_i$, then the restriction of $\mu$ to domain $V_h$ is a linear functional $\mu_h\in V_h'$.
+The functional $\mu_h$ has a coefficient dual vector $\boldsymbol{\mu} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ with entries
+	\begin{equation*}
+	\boldsymbol{\mu}_i=\mu(f_i) \quad i=1,\dots,N.
+	\end{equation*} 
+	For example, let $\delta_p$ be the Dirac distribution centered at a point $p\in\Omega$.
+If the finite element basis functions, $f_i$, are continuous at $p$, then restricting the domain of $\delta_p$ to the space $V_h$ yields a linear functional in $V_h'$.
+The corresponding coefficient dual vector, $\boldsymbol{\delta}_p$, has entries $\left(\boldsymbol{\delta}_p\right)_i = f_i(p)$ for $i=1,\dots,N$.
+
+	\item[Action of an operator on a distribution:] Since $V_h$ is finite-dimensional, the restriction $\mu_h \in V_h$ of a distribution $\mu$ to the space $V_h$ has a a Riesz representation $\mu_h^* \in V_h$.
 Let $B_h$ be the integral kernel associated with an operator $\mathcal{B}_h:V_h \rightarrow V_h'$.
-For $w_h \in V_h$, we have
-	\begin{align*}
-		\langle \mathcal{B}_h, \mu_h \rangle(w) &= \int_\Omega w_h(y) \mu_h\left(B_h(y, \cdot)\right) dy \\
-		&= \int_\Omega w_h(y) \left(\int_\Omega B_h(y,x) \mu_h^*(x) dx\right) dy \\
-		&= \int_\Omega \int_\Omega w_h(y) B_h(y,x) \mu_h^*(x) dx dy = \left(\mathcal{B}_h \mu_h^*\right)(w).
+For $v_h \in V_h$, we have
+	\begin{align*}
+		\langle \mathcal{B}_h, \mu_h \rangle(v) &= \int_\Omega v_h(y) \mu_h\left(B_h(y, \cdot)\right) dy \\
+		&= \int_\Omega v_h(y) \left(\int_\Omega B_h(y,x) \mu_h^*(x) dx\right) dy \\
+		&= \int_\Omega \int_\Omega v_h(y) B_h(y,x) \mu_h^*(x) dx dy = \left(\mathcal{B}_h \mu_h^*\right)(wv).
 	\end{align*}
@@
 \end{equation*}
-
-\section{Background: distributions}
-\label{app:distributions}
-
-Let $\overline{\Omega}$ be the closure of $\Omega$, and let $C(\overline{\Omega})$ be the space of continuous functions mapping $\overline{\Omega}\rightarrow \mathbb{R}$.
-The action of $\mathcal{A}$ is extended to distributions $\mu:C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ via the formula
-\begin{equation*}
-\langle\mathcal{A},\mu\rangle(w) := \int_\Omega w(y) \mu\left(A(y, \cdot)\right) dy, 
-\end{equation*}
-where $A(y,\cdot)$ is the function $x \mapsto A(y,x)$. This is derived formally as follows:
-\begin{align*}
-\langle\mathcal{A},\mu\rangle(w) &= \int_\Omega \int_\Omega w(y) A(y,x) \textrm{``} \mu(x) \textrm{''} dx dy \\
-&= \int_\Omega w(y) \int_\Omega A(y,x) \textrm{``} \mu(x) \textrm{''} dx \\
-&= \int_\Omega w(y) \mu\left(A(y,\cdot)\right) dx.
-\end{align*}
-For example, the delta distribution $\delta_x$ is defined by $\delta_x(v) = v(x)$, the action of $\mathcal{A}$ on $\delta_x$ is given by
-\begin{equation*}
-\langle\mathcal{A},\delta_x\rangle(w)	= \int_\Omega w(y) A(y,x) dy,
-\end{equation*}
-and so the Riesz representation of $\langle\mathcal{A},\delta_x\rangle$ is given by $\langle\mathcal{A},\delta_x\rangle^* = A(~\cdot~,x)$.
-That is, $\langle\mathcal{A},\delta_x\rangle^*$ is the function $y \mapsto A(y,x)$.
 

### 2021-01-26 paper/localpsf.tex
edits to discretization appendix

@@
 	\begin{equation*}
-	\boldsymbol{\mu}_i=\mu(f_i) \quad i=1,\dots,N.
+	\boldsymbol{\mu}_i=\mu(f_i), \quad i=1,\dots,N.
 	\end{equation*} 
@@
 If the finite element basis functions, $f_i$, are continuous at $p$, then restricting the domain of $\delta_p$ to the space $V_h$ yields a linear functional in $V_h'$.
-The corresponding coefficient dual vector, $\boldsymbol{\delta}_p$, has entries $\left(\boldsymbol{\delta}_p\right)_i = f_i(p)$ for $i=1,\dots,N$.
+The corresponding coefficient dual vector, $\boldsymbol{\delta}_p$, has entries
+	\begin{equation*}
+	\left(\boldsymbol{\delta}_p\right)_i = f_i(p), \quad i=1,\dots,N.
+	\end{equation*}
 
@@
 		&= \int_\Omega v_h(y) \left(\int_\Omega B_h(y,x) \mu_h^*(x) dx\right) dy \\
-		&= \int_\Omega \int_\Omega v_h(y) B_h(y,x) \mu_h^*(x) dx dy = \left(\mathcal{B}_h \mu_h^*\right)(wv).
+		&= \int_\Omega \int_\Omega v_h(y) B_h(y,x) \mu_h^*(x) dx dy = \left(\mathcal{B}_h \mu_h^*\right)(v).
 	\end{align*}
@@
 	\begin{equation*}
-		\psi_h = \langle \mathcal{B}_h, \mu_h \rangle \quad \Leftrightarrow \quad \psi_h = \mathcal{B}_h \mu_h^* \quad \Leftrightarrow \quad \boldsymbol{\psi} = \mathbf{B} \mathbf{M}^{-1} \boldsymbol{\mu}.
+		\sigma_h = \langle \mathcal{B}_h, \mu_h \rangle \quad \Leftrightarrow \quad \sigma_h = \mathcal{B}_h \mu_h^* \quad \Leftrightarrow \quad \boldsymbol{\sigma} = \mathbf{B} \mathbf{M}^{-1} \boldsymbol{\mu}.
 	\end{equation*}
@@
 \subsection{Discretized mean and covariance estimation} In computations, $\alpha$, $\mu^{i}$, and $\Sigma^{ij}$ are replaced with finite element approximations $\alpha_h, \mu_h^i, \Sigma_h^{ij}\in V_h$, respectively, which have coefficient vectors $\boldsymbol{\alpha}, \boldsymbol{\mu}^{ij}, \boldsymbol{\Sigma}^{ij}\in\mathbb{R}^N_\mathbf{M}$, respectively.
-The functions $C$, $L^i$, and $Q^{ij}$ are replaced with their $L^2$ projections onto $V_h$, which are denoted by $C_h$, $L_h^i$, and $Q_h^{ij}$, respectively.
-The corresponding coefficient vectors are $\mathbf{C}$, $\mathbf{L}^i$, and $\mathbf{Q}^{ij}$, respectively.
+The functions $C$, $L^i$, and $Q^{ij}$ are replaced with either finite element projections onto $V_h$ (if the finite element basis does not satisfy the Kronecker property), or finite element interpolations onto $V_h$ (if the finite element basis satisfies the Kronecker property).
+In our numerical examples we use finite element basis functions that satisfy the Kronecker property, and we use finite element interpolations.
+The coefficient vectors corresponding to either the interpolations or projections are denoted $\mathbf{C}$, $\mathbf{L}^i$, and $\mathbf{Q}^{ij}$, respectively.
 The discretized versions of \eqref{eq:vol_mean_var_thm1}, \eqref{eq:vol_mean_var_thm2}, and \eqref{eq:vol_mean_var_thm3} are
@@
 \end{align*}
-respectively. Here $\mathbf{f} \cdot \mathbf{g}$ and $\mathbf{f} / \mathbf{g}$ denote the element-wise multiplication and division of vectors, respectively.
+respectively. Here $\mathbf{f} \cdot \mathbf{g}$ and $\mathbf{f} / \mathbf{g}$ denote the component-wise multiplication and division of vectors, respectively.
 

### 2021-02-18 paper/localpsf.tex
starting to work on missing proof

@@
 
-		\tcp{Compute harmonic functions $\{\psi_i\}_{i=1}^r$ and $\{\theta_i\}_{i=1}^r$}
+		\tcp{Compute Poisson impulse responses $\{\psi_i\}_{i=1}^r$ and Poisson squared impulse responses $\{\theta_i\}_{i=1}^r$}
 
@@
 			\begin{cases}
-			\Delta \psi_i = \delta_{x_i} - C/|\Omega|, & \text{in }\Omega, \\
+			\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
 			\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
@@
 			\label{eq:definition_of_alpha_alg}
-			\alpha = \frac{\mathbf{1}^T S^{-1} e_i}{\mathbf{1}^T S^{-1} \mathbf{1}},
+			\beta = \frac{\mathbf{1}^T S^{-1} e_i}{\mathbf{1}^T S^{-1} \mathbf{1}},
 			\end{equation}
@@
 			\label{eq:definition_of_mu_alg}
-			\lambda = -S^{-1} \left(e_i - \alpha \mathbf{1} \right).
+			\lambda = -S^{-1} \left(e_i - \beta \mathbf{1} \right).
 			\end{equation}
@@
 			\label{eq:definition_of_wi_alg}
-				w_i(x) = \alpha - \sum_{k=1}^r \lambda_k \theta_k(x).
+				w_i(x) = \beta - \sum_{k=1}^r \lambda_k \theta_k(x).
 			\end{equation}
@@
 
-In order to use the product-convolution approximation $\widetilde{\mathcal{A}}$, we convert $\widetilde{\mathcal{A}}$ to hierarchical matrix (H-matrix) format.
+In order to use the product-convolution approximation $\widetilde{\mathcal{A}}$, we convert $\widetilde{\mathbf{A}}$ (the discretized version of $\widetilde{\mathcal{A}}$) to hierarchical matrix (H-matrix) format.
 Many classes of hierarchical matrices exist---H1, uniform H1, H2, HODLR, HSS, HBS, and more \cite{HMATRIX}.
@@
 
-We convert $\widetilde{\mathcal{A}}$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{HLIBPRO}.
-Briefly, this method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix to construct a recursive block partitioning of $\widetilde{\mathcal{A}}$, such that all blocks at the finest level of the partitioning are either low rank, or are small.
+We convert $\widetilde{\mathbf{A}}$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{HLIBPRO}.
+Briefly, this method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix to construct a recursive block partitioning of $\widetilde{\mathbf{A}}$, such that all blocks at the finest level of the partitioning are either low rank, or are small.
 Then, the method forms low rank approximations of the blocks that are expected to be low rank by sampling specially chosen rows and columns of those blocks.
 The small high rank blocks are formed directly by evaluating their entries.
-The primary cost of constructing the approximation is the evaluation of $O(N \log N)$ matrix entries of $\widetilde{\mathcal{A}}$, which we perform using Algorithm \ref{alg:eval_y_x}.
-We describe this H-matrix construction method in detail in Appendix \ref{app:h_matrix}.
+The primary cost of constructing the approximation is the evaluation of $O(N \log N)$ entries of the kernel $\widetilde{A}$, which we perform using Algorithm \ref{alg:eval_y_x}.
+We describe the H-matrix construction in detail in Appendix \ref{app:h_matrix}.
 
@@
 The H-matrix construction process that we use proceeds as follows.
-First, we construct a hierarchical partitionings of the degrees of freedom for the rows and columns of the matrix (cluster trees, Section \ref{sec:cluster_trees}).
+First, we construct hierarchical partitionings of the degrees of freedom for the columns and rows of the matrix (cluster trees, Section \ref{sec:cluster_trees}).
 Second, we construct a hierarchical partitioning of the blocks of the matrix, in such a way that many of the blocks in the partitioning are expected to be low rank, and the remaining high rank blocks are small (block cluster tree, Section \ref{sec:block_cluster_tree}).
@@
 The first two steps require geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix, but these steps do not depend on the particular values of matrix entries.
-The third step requires us to evaluate $O(N \log N)$ specially-chosen matrix entries of $\widetilde{\mathcal{A}}$, which we do using the method described in Section \ref{sec:eval_matrix_entries}.
+The third step requires us to evaluate $O(N \log N)$ specially-chosen entries of $\widetilde{A}$, and we evaluate these entries using the method described in Section \ref{sec:eval_matrix_entries}.
 
@@
 
-The degrees of freedom associated with the columns of the matrix are hierarchically clustered into a \emph{cluster tree} using recursive hyperplane splitting.
-Since we use finite elements to discretize the problem, we associate the $i^\text{th}$ column of $\widetilde{\mathcal{A}}$ with the nodal point in $\mathbb{R}^d$ associated with the $i^\text{th}$ finite element basis function.
+We use recursive hyperplane splitting to hierarchically cluster the degrees of freedom associated with the columns and rows of the matrix into a \emph{column cluster tree} and a \emph{row cluster tree}, respectively.
+Here we describe construction of the column cluster tree; the row cluster tree is constructed similarly.
+
+Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\widetilde{\mathbf{A}}$ corresponds to the nodal point in $\mathbb{R}^d$ associated with the $i^\text{th}$ finite element basis function.
 The columns of the matrix thus correspond to a point cloud in $\mathbb{R}^d$.
@@
 This hierarchical partitioning of the point cloud into smaller and smaller point clouds corresponds to a hierarchical partitioning of the columns of the matrix into smaller and smaller \emph{clusters} of columns.
-This hierarchical partitioning of the columns forms a binary tree, which we call a cluster tree.
+This hierarchical partitioning of the columns forms a binary tree, which we call the column cluster tree.
 The root of the tree is the set of all columns, and the leaves of the tree are clusters of columns that are not subdivided any further.
 
-A depth-first search ordering of the cluster tree leaves is then generated.
+A depth-first search ordering of the column cluster tree leaves is then generated.
 When the columns of the matrix are permuted into this depth-first ordering, the columns associated with each cluster in the cluster tree are contiguous.
@@
 
-After reordering the rows and columns of $\widetilde{\mathcal{A}}$ via the depth-first ordering described above, we partition the reordered version of $\widetilde{\mathcal{A}}$ into a tree of $2 \times 2$ block matrices recursively.
+After reordering the rows and columns of $\widetilde{\mathbf{A}}$ via the depth-first ordering described above, we partition the reordered version of $\widetilde{\mathbf{A}}$ into a tree of $2 \times 2$ block matrices recursively.
 We use a geometric admissibility condition (discussed below) to decide which blocks to subdivide, and use the cluster trees for the rows and columns to determine how to subdivide those blocks.
 For the first stage of subdivision, let $r_1$ and $r_2$ be the children row clusters for the root of the row cluster tree, let $c_1$ and $c_2$ be the children column clusters for the root of the column cluster tree.
-The matrix $\widetilde{\mathcal{A}}$ is partitioned into blocks as follows:
+The matrix $\widetilde{\mathbf{A}}$ is partitioned into blocks as follows:
 \begin{equation*}
 \begin{bmatrix}
-\widetilde{\mathcal{A}}_{11} & \widetilde{\mathcal{A}}_{12} \\
-\widetilde{\mathcal{A}}_{21} & \widetilde{\mathcal{A}}_{22},
+\widetilde{\mathbf{A}}_{11} & \widetilde{\mathbf{A}}_{12} \\
+\widetilde{\mathbf{A}}_{21} & \widetilde{\mathbf{A}}_{22},
 \end{bmatrix}
 \end{equation*}
-where $\widetilde{\mathcal{A}}_{11}$ denotes the block of $\widetilde{\mathcal{A}}$ with rows $r_1$ and columns $c_1$, $\widetilde{\mathcal{A}}_{12}$ denotes the block of $\widetilde{\mathcal{A}}$ with rows $r_1$ and columns $c_2$, and so on for $\widetilde{\mathcal{A}}_{21}$ and $\widetilde{\mathcal{A}}_{22}$.
-
-We now loop through the four blocks, $\widetilde{\mathcal{A}}_{11}$, $\widetilde{\mathcal{A}}_{12}$, $\widetilde{\mathcal{A}}_{21}$, and $\widetilde{\mathcal{A}}_{22}$, and decide which blocks should be subdivided further.
-For the purpose of explanation, consider $\widetilde{\mathcal{A}}_{12}$.
+where $\widetilde{\mathbf{A}}_{11}$ denotes the block of $\widetilde{\mathbf{A}}$ with rows $r_1$ and columns $c_1$, $\widetilde{\mathbf{A}}_{12}$ denotes the block of $\widetilde{\mathbf{A}}$ with rows $r_1$ and columns $c_2$, and so on for $\widetilde{\mathbf{A}}_{21}$ and $\widetilde{\mathbf{A}}_{22}$.
+
+We now loop through the four blocks, $\widetilde{\mathbf{A}}_{11}$, $\widetilde{\mathbf{A}}_{12}$, $\widetilde{\mathbf{A}}_{21}$, and $\widetilde{\mathbf{A}}_{22}$, and decide which blocks should be subdivided further.
+For the purpose of explanation, consider $\widetilde{\mathbf{A}}_{12}$.
 If
@@
 \end{equation}
-then we mark $\widetilde{\mathcal{A}}_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone.
+then we mark $\widetilde{\mathbf{A}}_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone.
 Here $\dist\left(r_1, c_2\right)$ is the Euclidean distance between the axis-aligned bounding box for the point cloud associated with the row cluster $r_1$, and the axis aligned bounding box for the point cloud associated with the column cluster $c_2$.
@@
 This process is repeated for the other blocks to determine which blocks are admissible and which are not.
-For us, the diagonal blocks $\widetilde{\mathcal{A}}_{11}$ and $\widetilde{\mathcal{A}}_{22}$ are not admissible because the distance between a point cloud and itself is zero.
-
-Next, we subdivide all blocks that are not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\mathcal{\widetilde{A}}$.
+For us, the diagonal blocks $\widetilde{\mathbf{A}}_{11}$ and $\widetilde{\mathbf{A}}_{22}$ are not admissible because the distance between a point cloud and itself is zero.
+
+Next, we subdivide all blocks that are not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\mathbf{\widetilde{A}}$.
 But now we subdivide a block based on the two child row clusters and two child column clusters for the rows and columns of that block.
@@
 Once the block cluster tree has been constructed, low rank approximations of the admissible (low rank) blocks are formed using the adaptive cross method \cite{ACA}.
-Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\widetilde{\mathcal{A}}$.
+Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\widetilde{\mathbf{A}}$.
 The idea of the adaptive cross method is to form a low rank approximation of $X$ by sampling a small number of rows and columns of $X$.
@@
 
-We use the adaptive cross method to form low rank approximations for each admissible block of $\widetilde{\mathcal{A}}$.
-We directly compute all entries of the small dense blocks of $\widetilde{\mathcal{A}}$ that are not admissible.
-This process requires us to compute $O(r^2 N \log N)$ entries of $\widetilde{\mathcal{A}}$, which is relatively cheap compared to the matrix-vector products with $\mathcal{A}$ that are required to form the product-convolution approximation $\widetilde{\mathcal{A}}$.
+We use the adaptive cross method to form low rank approximations for each admissible block of $\widetilde{\mathbf{A}}$.
+We directly compute all entries of the small dense blocks of $\widetilde{\mathbf{A}}$ that are not admissible.
+This process requires us to compute $O(r^2 N \log N)$ entries of $\widetilde{\mathbf{A}}$, which is relatively cheap compared to the matrix-vector products with $\mathbf{A}$ that are required to form the product-convolution approximation $\widetilde{\mathbf{A}}$.
 
@@
 	\label{lem:T_is_linear}
-	Let $f \in \mathbb{R}^r$. There exists a unique solution to the following optimization problem:
+	Let $\Omega \subset \mathbb{R}^d$, $d\in\{1,2,3\}$, be an open bounded domain with a bounded extension operator, let $x_i \in \Omega$ for $i=1,\dots,r$, and let $f \in \mathbb{R}^r$.
+Consider the optimization problem
 	\begin{equation}
@@
 	\end{equation}
-	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
-	\begin{equation*}
+	A minimizer, $\gamma^*$, of \eqref{eq:generic_optimization_initial} is a function defined on $\Omega$ that interpolates the data pairs $(x_i,f_i)$, and is as flat as possible in a least squares sense.
+	The following statements hold.
+	\begin{enumerate}
+		\item[(a)] There exists a unique solution to \eqref{eq:generic_optimization_initial}.
+		\item[(b)] 	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
+		\begin{equation}
+		\label{eq:opt_soln_operator}
 		\mathcal{T}(f) := \gamma^*,
-	\end{equation*}
+		\end{equation}
 		where $\gamma^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
+	\end{enumerate}
 \end{lem}
 \begin{proof}
-	WRITE THIS PROOF
+	By the Rellich-Kondrachov theorem (see, e.g., Theorem 7.20 in \cite{ARBOGAST}), $H^2_N(\Omega)$ is compactly embedded in $C^0(\Omega)$. WORK IN PROGRESS
 \end{proof}
@@
 	\nabla_{\lambda_j} \mathcal{L}(w, \lambda) = w(x_j) - e_j
-	= \alpha - \sum_{k=1}^r \lambda_k \theta_k(x_j) - e_j.
+	= \beta - \sum_{k=1}^r \lambda_k \theta_k(x_j) - e_j.
 	\end{equation*}
@@
 	\begin{align*}
-	\nabla_\lambda \mathcal{L}(w, \lambda) &= \alpha \mathbf{1} - e - S \lambda \\
-	&= \alpha \mathbf{1} - e - S \left(-S^{-1}\left(e - \alpha \mathbf{1}\right)\right) = 0.
+	\nabla_\lambda \mathcal{L}(w, \lambda) &= \beta \mathbf{1} - e - S \lambda \\
+	&= \beta \mathbf{1} - e - S \left(-S^{-1}\left(e - \beta \mathbf{1}\right)\right) = 0.
 	\end{align*}
@@
 	\end{equation*}
-	which implies that the first term in \eqref{eq:L_w_38} is zero. By the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} and $\alpha$ in \eqref{eq:definition_of_alpha_alg}, we have
-	\begin{align*}
-	\mathbf{1}^T \lambda &= -\mathbf{1}^T S^{-1} \left(e - \alpha \mathbf{1}\right) \\
-	&= -\mathbf{1}^T S^{-1} e + \alpha \mathbf{1}^T S^{-1} \mathbf{1} \\
+	which implies that the first term in \eqref{eq:L_w_38} is zero. By the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} and $\beta$ in \eqref{eq:definition_of_alpha_alg}, we have
+	\begin{align*}
+	\mathbf{1}^T \lambda &= -\mathbf{1}^T S^{-1} \left(e - \beta \mathbf{1}\right) \\
+	&= -\mathbf{1}^T S^{-1} e + \beta \mathbf{1}^T S^{-1} \mathbf{1} \\
 	&= -\mathbf{1}^T S^{-1} e + \frac{\mathbf{1}^T S^{-1} e}{\mathbf{1}^T S^{-1} \mathbf{1}} \mathbf{1}^T S^{-1} \mathbf{1} = 0,
@@
 \begin{proof}[Proof of Theorem \ref{thm:vol_mean_cov}]
-	Let $w \in L^2$ be arbitrary. We have
-	\begin{align*}
-		\left(w, \alpha\right)_{L^2(\Omega)} =& \int_\Omega w(x) \int_{\Omega} \langle\mathcal{A},\delta_x\rangle^*(z) dz dx\\
-		=& \int_\Omega w(x) \int_{\Omega} A(z,x) dz dx \\
-		=& \int_\Omega \int_\Omega w(x) A(z,x) C(z) dz dx \\
-		=& \left(\mathcal{A}^T C\right)(w),
+	Let $v \in L^2$ be arbitrary. We have
+	\begin{align*}
+		\left(v, \alpha\right)_{L^2(\Omega)} =& \int_\Omega v(x) \int_{\Omega} \langle\mathcal{A},\delta_x\rangle^*(z) dz dx\\
+		=& \int_\Omega v(x) \int_{\Omega} A(z,x) dz dx \\
+		=& \int_\Omega \int_\Omega v(x) A(z,x) C(z) dz dx \\
+		=& \left(\mathcal{A}^T C\right)(v),
 	\end{align*}
@@
 	\begin{align*}
-		\left(w, \alpha \cdot \mu^i\right)_{L^2(\Omega)} =& \int_\Omega w(x) \alpha(x) \int_{\Omega} z^i \langle\mathcal{A},\delta_x\rangle^*(z)/\alpha(x) dz dx\\
-		=& \int_\Omega \int_\Omega w(x) L^i(z) A(z,x) dz dx \\
-		=& \left(\mathcal{A}^T L^i\right)(w),
+		\left(v, \alpha \cdot \mu^i\right)_{L^2(\Omega)} =& \int_\Omega v(x) \alpha(x) \int_{\Omega} z^i \langle\mathcal{A},\delta_x\rangle^*(z)/\alpha(x) dz dx\\
+		=& \int_\Omega \int_\Omega v(x) L^i(z) A(z,x) dz dx \\
+		=& \left(\mathcal{A}^T L^i\right)(v),
 	\end{align*}
@@
 	\begin{align*}
-		\left(w, \alpha \cdot \left(\Sigma^{ij}+\mu^i \cdot \mu^j\right)\right)_{L^2(\Omega)} =& \int_\Omega w(x) \alpha(x) \int_\Omega z^i z^j \langle\mathcal{A},\delta_x\rangle^*(z) / \alpha(x) dz dx \\
-		=& \int_\Omega \int_\Omega w(x) Q^{ij}(z) A(z,x) dz dx \\
-		=& \left(\mathcal{A}Q^{ij}\right)(w),
+		\left(v, \alpha \cdot \left(\Sigma^{ij}+\mu^i \cdot \mu^j\right)\right)_{L^2(\Omega)} =& \int_\Omega v(x) \alpha(x) \int_\Omega z^i z^j \langle\mathcal{A},\delta_x\rangle^*(z) / \alpha(x) dz dx \\
+		=& \int_\Omega \int_\Omega v(x) Q^{ij}(z) A(z,x) dz dx \\
+		=& \left(\mathcal{A}Q^{ij}\right)(v),
 	\end{align*}
@@
 \begin{equation*}
-\langle\mathcal{A},\mu\rangle(w) := \int_\Omega w(y) \mu\left(A(y, \cdot)\right) dy, 
+\langle\mathcal{A},\mu\rangle(v) := \int_\Omega v(y) \mu\left(A(y, \cdot)\right) dy, 
 \end{equation*}
@@
 \begin{align*}
-\langle\mathcal{A},\mu\rangle(w) &= \int_\Omega \int_\Omega w(y) A(y,x) \textrm{``} \mu(x) \textrm{''} dx dy \\
-&= \int_\Omega w(y) \int_\Omega A(y,x) \textrm{``} \mu(x) \textrm{''} dx \\
-&= \int_\Omega w(y) \mu\left(A(y,\cdot)\right) dx.
+\langle\mathcal{A},\mu\rangle(v) &= \int_\Omega \int_\Omega v(y) A(y,x) \textrm{``} \mu(x) \textrm{''} dx dy \\
+&= \int_\Omega v(y) \int_\Omega A(y,x) \textrm{``} \mu(x) \textrm{''} dx \\
+&= \int_\Omega v(y) \mu\left(A(y,\cdot)\right) dx.
 \end{align*}
@@
 \label{eq:appendix_impulse_response}
-\langle\mathcal{A},\delta_x\rangle(w)	= \int_\Omega w(y) A(y,x) dy.
+\langle\mathcal{A},\delta_x\rangle(v)	= \int_\Omega v(y) A(y,x) dy.
 \end{equation}
@@
 
-Let $f_i$, $i=1,\dots,N$ be a set of finite element basis functions, and let
-\begin{equation*}
-V_h = \Span\left(f_1, f_2, \dots, f_N\right).
+Let $f_i$, $i=1,\dots,N$ be a set of finite element basis functions, let $h$ denote the mesh size parameter for the finite element mesh, and let
+\begin{equation*}
+V_h = \Span\left(f_1, f_2, \dots, f_N\right)
 \end{equation*}
 be the corresponding  finite element space with the $L^2$ inner product.
-Here $h$ denotes the mesh size parameter.
 Functions $u\in L^2(\Omega)$ are approximated by functions $u_h \in V_h$.
@@
 
-Let $\mathbf{M} \in \mathbb{R}^{N \times N}$ denote the finite element \emph{mass matrix}, which has entries
+We write $\mathbf{M} \in \mathbb{R}^{N \times N}$ to denote the finite element \emph{mass matrix}, which has entries
 \begin{equation*}
@@
 \end{equation*}
-and let $\mathbb{R}^N_\mathbf{M}$ denote the space $\mathbb{R}^N$ with the matrix-weighted inner product
-\begin{equation*}
-\left(\mathbf{u},\mathbf{v}\right)_\mathbf{M} := \left(\mathbf{M} \mathbf{u}\right)^T \mathbf{v}.
-\end{equation*}
-Direct calculation shows that $V_h$ is isometrically isomorphic $\mathbb{R}^N_\mathbf{M}$, in the sense that \eqref{eq:finite_bijection} establishes a bijective correspondence between functions $u_h \in V_h$ and coefficient vectors $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, and that this bijection preserves the inner product; i.e., $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u},\mathbf{v}\right)_\mathbf{M}$.
+and we write $\mathbb{R}^N_\mathbf{M}$ to denote the space $\mathbb{R}^N$ with the matrix-weighted inner product
+\begin{equation*}
+\left(\mathbf{u},\mathbf{v}\right)_\mathbf{M} := \mathbf{v}^T\mathbf{M} \mathbf{u}.
+\end{equation*}
+Direct calculation shows that $V_h$ is isometrically isomorphic $\mathbb{R}^N_\mathbf{M}$.
+That is, equation \eqref{eq:finite_bijection} establishes a bijection between functions $u_h \in V_h$ and coefficient vectors $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, and this bijection preserves the inner product in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u},\mathbf{v}\right)_\mathbf{M}$.
 Likewise, the dual space, $V_h'$, is isometrically isomorphic to $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
-These isomorphisms allow us to identify functions $u_h \in V_h$ with their corresponding coefficient arrays $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, and linear functionals $\sigma_h \in V_h'$ with their corresponding coefficient arrays $\boldsymbol{\sigma} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$.
+
+Here vectors in $\mathbf{u} \in \mathbb{R}^N_{\mathbf{M}}$ and $\boldsymbol{\sigma} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ are viewed as column vectors, and $\mathbf{u}^T$ and $\boldsymbol{\sigma}^T$ are the row vectors that result from taking the standard matrix transposes of $\mathbf{u}$ and $\boldsymbol{\sigma}$, respectively.
+The discrete dual of a vector $\mathbf{v}$ with respect to the $\mathbf{M}$ inner product is written as $\mathbf{v}^*$, and $\mathbf{v}^*$ is viewed as a column vector.
+%These isomorphisms allow us to identify functions $u_h \in V_h$ with their corresponding coefficient arrays $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, and linear functionals $\sigma_h \in V_h'$ with their corresponding coefficient arrays $\boldsymbol{\sigma} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$.
 
@@
 
-Here we list common function space operations on $V_h$ and $V_h'$, and show the corresponding matrix and vector operations on $\mathbb{R}^N_{\mathbf{M}}$ and $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
+Here we list the function space operations on $V_h$ and $V_h'$ that are required for this paper, and show how to perform the corresponding matrix and vector operations on $\mathbb{R}^N_{\mathbf{M}}$ and $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
 
@@
 	\begin{equation*}
-		\sigma_h(v_h) = \mathbf{u}^T \boldsymbol{\sigma}.
+		\sigma_h(v_h) = \boldsymbol{\sigma}^T \mathbf{u}.
 	\end{equation*}
@@
 	\end{equation*}
-	\item[Finite element interpolation:] Let $p_i \in \mathbb{R}^d$ be the Lagrange node associated with the finite element basis function $f_i$.
+	\item[Finite element interpolation:] Let $p_i \in \mathbb{R}^d$, $i=1,\dots,N$, be the Lagrange nodes associated with the finite element basis functions $f_i$.
 The finite element interpolant of a function $g \in C\left(\overline{\Omega}\right)$ is the function $g^I \in V_h$ given by
@@
 	\item[Projection vs. interpolation:]
-	Typically $g^\pi$ is a more accurate approximation of $g$ than $g^I$.
-However, $g^I$ is typically cheaper to compute than $g^\pi$ because computing $g^\pi$ requires solving a linear system, while computing $g^I$ does not.
-Interpolation is typically sufficiently accurate for practical purposes so long as the finite element basis satisfies the following Kronecker property:
+	Typically, $g^\pi$ is a more accurate approximation of $g$ than $g^I$, but $g^I$ is cheaper to compute than $g^\pi$ because computing $g^\pi$ requires solving a linear system, while computing $g^I$ does not.
+If the finite element basis satisfies the following Kronecker property,
 	\begin{equation*}
 		f_i(p_j) = \begin{cases}1 & i=j \\
-		0 & \text{otherwise}.\end{cases}
-	\end{equation*}
+		0 & \text{otherwise},\end{cases}
+	\end{equation*}
+	then $g^I(p_i) = g(p_i)$ for $i=1,\dots,N$ (i.e., the interpolant equals the original function at the Lagrange nodes).
+Interpolation is typically sufficiently accurate for practical purposes if the Kronecker property is satisfied.
+Projection should be used if the Kronecker property is not satisfied.
 	\item[Matrix representation of an operator:] Let $\mathcal{B}_h : V_h \rightarrow V_h'$ be a linear operator, and let $\mathbf{B} \in \mathbf{R}^{N \times N}$ be the matrix with entries
@@
 	\item[Distributions:] Let $\mu:C\left(\overline{\Omega}\right)\rightarrow \mathbb{R}$ be a distribution.
-If $\mu(f_i)$ is well-defined for all finite element basis vectors $f_i$, then the restriction of $\mu$ to domain $V_h$ is a linear functional $\mu_h\in V_h'$.
+If $\mu(f_i)$ is well-defined for $i=1,\dots,N$, then the restriction of $\mu$ to domain $V_h$ is a linear functional $\mu_h\in V_h'$.
 The functional $\mu_h$ has a coefficient dual vector $\boldsymbol{\mu} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ with entries
@@
 
-	\item[Action of an operator on a distribution:] Since $V_h$ is finite-dimensional, the restriction $\mu_h \in V_h$ of a distribution $\mu$ to the space $V_h$ has a a Riesz representation $\mu_h^* \in V_h$.
+	\item[Action of an operator on a distribution:] Since $V_h$ is finite-dimensional, the restriction, $\mu_h \in V_h$, of a distribution, $\mu$, to the space $V_h$ has a a Riesz representation $\mu_h^* \in V_h$.
 Let $B_h$ be the integral kernel associated with an operator $\mathcal{B}_h:V_h \rightarrow V_h'$.
@@
 
-\subsection{Discretized mean and covariance estimation} In computations, $\alpha$, $\mu^{i}$, and $\Sigma^{ij}$ are replaced with finite element approximations $\alpha_h, \mu_h^i, \Sigma_h^{ij}\in V_h$, respectively, which have coefficient vectors $\boldsymbol{\alpha}, \boldsymbol{\mu}^{ij}, \boldsymbol{\Sigma}^{ij}\in\mathbb{R}^N_\mathbf{M}$, respectively.
+\subsection{Discretized mean and covariance estimation}
+
+\begin{algorithm2e}
+	\SetAlgoNoLine
+	\SetKwInOut{Input}{Input}
+	\SetKwInOut{Output}{Output}
+	{	
+		\Input{Operator $\mathbf{A}$}
+		\Output{$\boldsymbol{\alpha}\in \mathbb{R}^N$, $\boldsymbol{\mu}\in \mathbb{R}^{d \times N}$, and $\boldsymbol{\Sigma} \in \mathbb{R}^{d \times d \times N}$}
+
+		Form vector $\mathbf{C} \in \mathbb{R}^N_{\mathbf{M}}$ by either projecting or interpolating constant function $C(x)=1$ onto $V_h$
+
+		$\boldsymbol{\alpha} = \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{C}\right)^*$
+
+		\For{$i=1,2,\dots,d$}{
+			Form vector $\mathbf{L}^i \in \mathbb{R}^N_{\mathbf{M}}$ by either projecting or interpolating linear function $L^i(x) = x^i$ onto $V_h$
+
+			$\mathbf{\mu}^i = \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{L}^i\right)^* / \boldsymbol{\alpha}$
+		}
+		\For{$i=1,2,\dots,d$}{
+			\For{$j=1,\dots,i$}{
+				Form quadratic function $\mathbf{Q}^{ij} = x^i x^j$
+
+				$\boldsymbol{\Sigma}^{ij} = \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{Q}^{ij}\right)^* / \boldsymbol{\alpha} - \boldsymbol{\mu}^i\cdot \boldsymbol{\mu}^j$
+
+				$\boldsymbol{\Sigma}^{ij} = \boldsymbol{\Sigma}^{ji}$
+
+			}
+		}
+
+	}
+	\caption{Discretized version of Algorithm \ref{alg:varhpi_mean_cov}}
+	\label{alg:varhpi_mean_cov_discrete}
+\end{algorithm2e}
+
+In computations, $\alpha$, $\mu^{i}$, and $\Sigma^{ij}$ are replaced with finite element approximations $\alpha_h, \mu_h^i, \Sigma_h^{ij}\in V_h$, respectively, which have coefficient vectors $\boldsymbol{\alpha}, \boldsymbol{\mu}^{ij}, \boldsymbol{\Sigma}^{ij}\in\mathbb{R}^N_\mathbf{M}$, respectively.
 The functions $C$, $L^i$, and $Q^{ij}$ are replaced with either finite element projections onto $V_h$ (if the finite element basis does not satisfy the Kronecker property), or finite element interpolations onto $V_h$ (if the finite element basis satisfies the Kronecker property).
@@
 \subsection{Discretized sampling of impulse responses}
-In computations, the Dirac comb $\xi^b$ is replaced by its restriction to $V_h$, denoted $\xi^b \in V_h'$.
-The response $\eta^b$ is replaced by the finite element approximation $\eta_h^b \in V_h$.
-The coefficient dual vector for $\xi_h^b$ is the vector $\boldsymbol{\xi}^b \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has entries
-\begin{equation*}
-\boldsymbol{\xi}^b_i = \sum_{x_j \in S_b} \phi_i(x_j), \quad i=1,\dots,N.
-\end{equation*}
-The coefficient vector $\boldsymbol{\eta}^b \in \mathbb{R}^N_\mathbf{M}$ for $\eta_h^b$ is given by
+In computations, the Dirac comb coefficient dual vector for $\xi_h^b$ is the vector $\boldsymbol{\xi}^b \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has entries
+\begin{equation*}
+\boldsymbol{\xi}^b_i = \sum_{x_j \in S_b} f_i(x_j), \quad i=1,\dots,N.
+\end{equation*}
+The Dirac comb impulse response coefficient vector for $\eta_h^b$ is the vector $\boldsymbol{\eta}^b \in \mathbb{R}^N_\mathbf{M}$ given by
 \begin{equation*}

### 2021-03-08 paper/localpsf.tex
todo lists for paper from meeting

@@
 The idea of the method is to compute impulse responses of the operator at a collection of scattered points, then interpolate these computed impulse responses to approximate impulse responses at other arbitrary points.
-This results in a product-convolution approximation, which we convert to hierarchical matrix format.
-Once the approximation is represented in hierarchical matrix format, we perform matrix inversion and other matrix operations using fast hierarchical matrix arithmetic.
+This results in a product-convolution approximation, which we convert to hierarchical matrix format in order to solve linear systems and perform other matrix operations efficiently using fast hierarchical matrix arithmetic.
 Impulse responses are computed by applying the operator to a small number of Dirac comb ``batches'' of point sources.
@@
 
+\subsection{Computational cost}
+
+\begin{itemize}
+	\item computational complexity in terms of PDE solves:
+	\item cost of global low rank approximation (data scalability)
+	\item cost to build approximation
+	\item cost without compression to do a solve
+	\item cost with compression to do a solve
+	\item computational complexity for H-matrix alone vs. prod-conv + h-matrix $O(C k^2 \log N)$ PDE solves, C is big, k is big. Here $10$
+	\item HODLR: C is smaller, but k is bigger, vs. H1 peeling process
+	\item H-matrix operations cost for our method solve
+	\item irregular grid
+\end{itemize}
+
 \section{Numerical results}
 \label{sec:numerical_results}
+
+\begin{itemize}
+	\item GLR vs PCH diffusion time (H)
+	\item GLR vs PCH number of obs (Computational cost) (S)
+	\item GLR vs PCH aspect ratio (CC) (S)
+	\item mesh scalability of PCH (CC) (H)
+	\item spectrum plots $P^{-1} H - I$ (Quality of approximation) (H, not Stokes?)
+	\item number of krylov iterations: reg precond vs PCH (where?) (H)
+	\item turn on preconditioner after number of krylov iterations exceeds 15 in an iteration. Rebuild "as needed". (H)
+	\item $P - H = USV^T$ <- less good
+	\item $P^{-1} H - I = USV^T$ <- better
+	\item true parameter (H)
+	\item deterministic reconstruction (H)
+	\item noisy measurements (1\% noise). $u|_\text{top}$ (H)
+	\item recovered state at top (H)
+\end{itemize}
 

### 2021-03-17 paper/localpsf.tex
more advanced weighting function interpolation could fix boundary issues

@@
 
+\textbf{Heat equation}
 \begin{itemize}
-	\item GLR vs PCH diffusion time (H)
-	\item GLR vs PCH number of obs (Computational cost) (S)
-	\item GLR vs PCH aspect ratio (CC) (S)
-	\item mesh scalability of PCH (CC) (H)
-	\item spectrum plots $P^{-1} H - I$ (Quality of approximation) (H, not Stokes?)
-	\item number of krylov iterations: reg precond vs PCH (where?) (H)
-	\item turn on preconditioner after number of krylov iterations exceeds 15 in an iteration. Rebuild "as needed". (H)
-	\item $P - H = USV^T$ <- less good
-	\item $P^{-1} H - I = USV^T$ <- better
+	\item Hessian per-column error plots for 1, 5, and 15 batches
+	\item $H-P$ error vs number of batches for interior and whole domain
+	\item $P^{-1}H-I$: error vs rank for $P=R$ and $P=P_\text{pch}$ diffusion time, for 1, 5, and 15 batches
+	\item Krylov iterations to tols 1e-1 and 1e-6 vs diffusion parameter
+	\item Krylov method convergence plot: $R$ vs $P_\text{pch}$ vs no preconditioner
+	\item Krylov iterations to tols 1e-6 vs mesh size $h$, PCH vs no preconditioner (hold)
 	\item true parameter (H)
@@
 	\item recovered state at top (H)
+	\item mesh scalability of PCH (CC) (H)
+\end{itemize}
+
+\begin{itemize}
+	\item GLR vs PCH number of obs (Computational cost) (S)
+	\item GLR vs PCH aspect ratio (CC) (S)
+	\item turn on preconditioner after number of krylov iterations exceeds 15 in an iteration. Rebuild "as needed". (S)
 \end{itemize}

### 2021-03-17 code/src/eval_vs_small_linsolve_experiments.py

@@ -1,0 +1,29 @@
+import numpy as np
+import dolfin as dl
+from time import time
+
+mesh = dl.UnitSquareMesh(71,85)
+bbt = mesh.bounding_box_tree()
+
+V = dl.FunctionSpace(mesh, 'CG', 1)
+u = dl.Function(V)
+
+n=100000
+m=30
+u.set_allow_extrapolation(True)
+pts = [dl.Point(np.random.randn(), np.random.randn()) for _ in range(n)]
+t = time()
+uu = [u(pt) for pt in pts]
+dt_eval = time() - t
+t = time()
+ents = [bbt.compute_first_entity_collision(pt) for pt in pts]
+dt_collision = time() - t
+MM = [np.random.randn(m,m) for _ in range(n)]
+bb = [np.random.randn(m) for _ in range(n)]
+t = time()
+iMMb = [np.linalg.solve(M,b) for M, b in zip(MM, bb)]
+dt_solve = time() - t
+t = time()
+MMb = [np.dot(M,b) for M, b in zip(MM, bb)]
+dt_dot = time() - t
+print('n=', n, ', dt_eval=', dt_eval, ', dt_collision=', dt_collision, ', dt_solve=', dt_solve, ', dt_dot=', dt_dot)

### 2021-06-08 paper/localpsf.tex
refactoring

@@
 
+\section{Rational positive definite modification}
+WRITE THIS
+
 \section{Numerical results}
@@
 	\item Krylov iterations to tols 1e-1 and 1e-6 vs diffusion parameter
+	\item GLR vs PCH vs diffusion time
 	\item Krylov method convergence plot: $R$ vs $P_\text{pch}$ vs no preconditioner
-	\item Krylov iterations to tols 1e-6 vs mesh size $h$, PCH vs no preconditioner (hold)
+	\item Krylov iterations to tols 1e-6 vs mesh size $h$, PCH vs R vs no preconditioner (hold)
 	\item true parameter (H)
 	\item deterministic reconstruction (H)
-	\item noisy measurements (1\% noise). $u|_\text{top}$ (H)
+	\item noisy measurements (1\%, 5\%, 10\% noise). $u|_\text{top}$ (H)
 	\item recovered state at top (H)
 	\item mesh scalability of PCH (CC) (H)
+	\item put plot data into data directory
+	\item save function .pvd files in paraview  directory
 \end{itemize}
@@
 \begin{itemize}
-	\item GLR vs PCH number of obs (Computational cost) (S)
+	\item GLR vs PCH number of obs (Computational cost in PDE solves) (S)
 	\item GLR vs PCH aspect ratio (CC) (S)

### 2021-06-08 code/brent_minimize.py

@@ -1,114 +1,0 @@
-import numpy as np
-from numba import njit
-
-_cg = 0.3819660
-_mintol = 1e-11
-
-@njit
-def brent_minimize(func, xmin, xmax, args=(), tol=1e-7, maxiter=200):
-    # Based on version in scipy.optimize
-    xa = xmin
-    xc = xmax
-    xb = (xa + xc) / 2.
-
-    funcalls = 0
-
-    x = w = v = xb
-    fw = fv = fx = func(*((x,) + args))
-    if (xa < xc):
-        a = xa
-        b = xc
-    else:
-        a = xc
-        b = xa
-    deltax = 0.0
-    funcalls += 1
-    iter = 0
-    while (iter < maxiter):
-        tol1 = tol * np.abs(x) + _mintol
-        tol2 = 2.0 * tol1
-        xmid = 0.5 * (a + b)
-        # check for convergence
-        if np.abs(x - xmid) < (tol2 - 0.5 * (b - a)):
-            break
-        # XXX In the first iteration, rat is only bound in the true case
-        # of this conditional. This used to cause an UnboundLocalError
-        # (gh-4140). It should be set before the if (but to what?).
-        if (np.abs(deltax) <= tol1):
-            if (x >= xmid):
-                deltax = a - x  # do a golden section step
-            else:
-                deltax = b - x
-            rat = _cg * deltax
-        else:  # do a parabolic step
-            tmp1 = (x - w) * (fx - fv)
-            tmp2 = (x - v) * (fx - fw)
-            p = (x - v) * tmp2 - (x - w) * tmp1
-            tmp2 = 2.0 * (tmp2 - tmp1)
-            if (tmp2 > 0.0):
-                p = -p
-            tmp2 = np.abs(tmp2)
-            dx_temp = deltax
-            deltax = rat
-            # check parabolic fit
-            if ((p > tmp2 * (a - x)) and (p < tmp2 * (b - x)) and
-                    (np.abs(p) < np.abs(0.5 * tmp2 * dx_temp))):
-                rat = p * 1.0 / tmp2  # if parabolic step is useful.
-                u = x + rat
-                if ((u - a) < tol2 or (b - u) < tol2):
-                    if xmid - x >= 0:
-                        rat = tol1
-                    else:
-                        rat = -tol1
-            else:
-                if (x >= xmid):
-                    deltax = a - x  # if it's not do a golden section step
-                else:
-                    deltax = b - x
[…]

### 2021-06-08 code/covariance_estimation.py

@@ -1,75 +1,0 @@
-import numpy as np
-
-def unit_row_vector(d,k):
-    ek = np.zeros(d)
-    ek[k] = 1.
-    return ek.reshape((1,-1))
-
-def compute_col_moments(apply_At, apply_W, qq):
-    # qq is d-by-N array of point coordinates (d=dimension (e.g., 2 or 3), N=num_pts)
-    # apply_A(V) := A * v
-    # apply_At(v) := A^T * v
-    # apply_W(v) := W * v
-    # A is N-by-N operator, mapping X' -> X
-    # W is N-by-N mass matrix, mapping X -> X'
-    d, N = qq.shape
-    constant_function = np.ones(N)
-    col_volumes = apply_At(apply_W(constant_function))
-    apply_At_rescaled = lambda v: (1. / col_volumes) * apply_At(v)
-
-    col_means = np.zeros((d,N))
-    for ii in range(d):
-        linear_function_ii_direction = np.dot(unit_row_vector(d,ii), qq).reshape(-1)
-        col_means[ii,:] = apply_At_rescaled(apply_W(linear_function_ii_direction))
-
-    col_vars = np.zeros((d,d,N))
-    for ii in range(d):
-        for jj in range(ii+1):
-            if ii == jj:
-                v = unit_row_vector(d, ii)
-            else:
-                v = unit_row_vector(d, ii) + unit_row_vector(d, jj)
-            v = v/np.linalg.norm(v)
-            linear_function_v_direction = np.dot(v, qq).reshape(-1)
-            quadratic_function_v_direction = linear_function_v_direction**2
-            E_x = np.dot(v, col_means)
-            # E_x_b = apply_At_rescaled(apply_W(linear_function_v_direction))
-            # print('err=', np.linalg.norm(E_x_b - E_x))
-            E_x2 = apply_At_rescaled(apply_W(quadratic_function_v_direction))
-            col_vars[ii, jj, :] = E_x2 - E_x**2
-            col_vars[jj, ii, :] = col_vars[ii, jj, :]
-
-    for ii in range(d):
-        for jj in range(d):
-            if not (ii == jj):
-                col_vars[ii, jj, :] = col_vars[ii, jj, :] - col_vars[ii, ii, :]/2. - col_vars[jj, jj, :]/2.
-
-    return col_volumes, col_means, col_vars
-
-
-def distance_from_boundary_function(qq, boundary_inds):
-    dd = np.zeros(qq.shape[1])
[…]

### 2021-06-08 code/cross.py

@@ -1,133 +1,0 @@
-#https://bitbucket.org/muxas/maxvolpy/src/master/maxvolpy/cross.py
-from __future__ import absolute_import, division, print_function
-import numpy as np
-
-
-def py_cross_full(A, rank):
-    """returns cross approximation of matrix A, computed with full pivoting."""
-    rows = np.ndarray(A.shape[0], dtype=np.int32)
-    cols = np.ndarray(A.shape[1], dtype=np.int32)
-
-
-def py_lu_full_fixed(A, rank):
-    N, M = A.shape
-    U = np.ndarray((N, rank), dtype=A.dtype)
-    V = np.ndarray((rank, M), dtype=A.dtype)
-    B = A.copy()
-    for i in range(rank):
-        maxi, maxj = divmod(abs(B).argmax(), M)
-        max_val = B[maxi, maxj]
-        U[:, i] = B[:, maxj].copy()
-        V[i] = B[maxi] / max_val
-        B -= U[:, i].reshape(-1, 1) * V[i].reshape(1, -1)
-    return U, V
-
-
-def py_lu_partial_fixed(A, rank):
-    N, M = A.shape
-    U = np.ndarray((N, rank), dtype=A.dtype)
-    V = np.ndarray((rank, M), dtype=A.dtype)
-    P = [i for i in range(N)]
-    maxi = 0
-    i = 0
-    index_in_P = 0
-    while i < rank:
-        B = A[maxi] - U[maxi, :i].dot(V[:i])
-        maxj = abs(B).argmax()
-        max_val = B[maxj]
-        if max_val == 0:
-            if len(P) == 1:
-                break
-            P.pop(index_in_P)
-
-        else:
-            U[:, i] = A[:, maxj] - U[:, :i].dot(V[:i, maxj])
-            V[i] = B / max_val
-            P.pop(index_in_P)
-            index_in_P = abs(U[P, i]).argmax()
-            maxi = P[index_in_P]
-            i += 1
-    return U, V
-
-
-def aca_partial(A, tol):
-# def aca_partial(A, niter):
-    N, M = A.shape
-    rank = 10
-    U = np.ndarray((N, rank), dtype=A.dtype)
-    V = np.ndarray((rank, M), dtype=A.dtype)
-    P = [i for i in range(N)]
-    maxi = 0
-    i = 0
-    index_in_P = 0
-    eps = 1 + tol
-    while tol < eps:
-    # for _ in range(niter):
-        B = A[maxi] - U[maxi, :i].dot(V[:i])
[…]

### 2021-06-08 code/ellipsoid.py

@@ -1,50 +1,0 @@
-import numpy as np
-from numba import jit
-
-from brent_minimize import brent_minimize
-
-
-@jit(nopython=True)
-def points_which_are_not_in_ellipsoid_numba(Sigma, mu, pp, tau):
-    d = Sigma.shape[-1]
-    zz = (pp.reshape((-1,d)) - mu.reshape((1,d))).T
-    return (np.sum(zz * np.linalg.solve(Sigma, zz), axis=0) > tau**2)
-
-
-@jit(nopython=True)
-def ellipsoids_intersect(Sigma_a, Sigma_b, mu_a, mu_b, tau):
-    lambdas, Phi, v_squared = ellipsoids_intersect_helper(Sigma_a, Sigma_b, mu_a, mu_b)
-    xmin, fval, iter, funcalls = brent_minimize(K_fct, 0., 1., args=(lambdas, v_squared, tau))
-    return np.bool_(fval >= 0)
-
-
-@jit(nopython=True)
-def ellipsoids_intersect_helper(Sigma_a, Sigma_b, mu_a, mu_b):
-    lambdas, Phi = geigh_numpy(Sigma_a, Sigma_b)
-    v_squared = np.dot(Phi.T, mu_a - mu_b) ** 2
-    return lambdas, Phi, v_squared
-
-
-@jit(nopython=True)
-def K_fct(s, lambdas, v_squared, tau):
-    return 1. - (1. / tau**2) * np.sum(v_squared * ((s * (1. - s)) / (1. + s * (lambdas - 1.))))
-
-
-@jit(nopython=True)
-def K_fct_vectorized(ss, lambdas, v_squared, tau):
-    KK = np.empty(len(ss))
-    for ii in range(len(ss)):
-        KK[ii] = K_fct(ss[ii], lambdas, v_squared, tau)
-    return KK
-
-
-@jit(nopython=True)
-def geigh_numpy(A, B):
-    # Generalized eigenvalue decomposition A*Phi = B*Phi*Lambda
-    # Phi.T * B * Phi = I
-    lambdas0, Phi0 = np.linalg.eig(np.linalg.solve(B, A))
-    lambdas = lambdas0.real
-    Rayleigh_matrix = np.dot(np.dot(Phi0.real.T, B), Phi0.real)
-    scaling_factors = np.sqrt(np.diag(Rayleigh_matrix))
-    Phi = Phi0.real / scaling_factors.reshape((1,-1))
-    return lambdas, Phi

### 2021-06-08 code/ellipsoid_intersection_test.py

@@ -1,58 +1,0 @@
-import numpy as np
-from scipy.optimize import minimize_scalar
-from scipy.linalg import eigh
-from brent_minimize import brent_minimize
-from numba import njit
-from time import time
-
-@njit
-def K(x, dd, v):
-    return 1. - np.sum(v * ((dd * x * (1. - x)) / (x + dd * (1. - x))) * v)
-
-def do_ellipsoids_intersect(A, B, a, b):
-    dd, Phi = eigh(A, B, eigvals_only=False)
-    v = np.dot(Phi.T, a - b)
-    xmin, fval, iter, funcalls = brent_minimize(K, 0., 1., args=(dd, v))
-    return (fval >= 0)
-
-
-
-# d=3
-#
-# Ua, ssa, _ = np.linalg.svd(np.random.randn(d, d))
-# Ub, ssb, _ = np.linalg.svd(np.random.randn(d, d))
-#
-# A = np.dot(Ua, np.dot(np.diag(ssa), Ua.T))
-# B = np.dot(Ub, np.dot(np.diag(ssb), Ub.T))
-#
-# a = np.random.randn(d)
-# b = np.random.randn(d)
-#
-# t = time()
-# do_ellipsoids_intersect(A, B, a, b)
-# dt = time() - t
-# print('dt=', dt)
-#
-# dd, Phi = eigh(A, B, eigvals_only=False)
-# v = np.dot(Phi.T, a - b)
-#
-# N=100
-# AA = np.zeros((d,d,N))
-# BB = np.zeros((d,d,N))
-# for k in range(N):
-#     Ua, ssa, _ = np.linalg.svd(np.random.randn(d, d))
-#     Ub, ssb, _ = np.linalg.svd(np.random.randn(d, d))
-#
-#     AA[:,:,k] = np.dot(Ua, np.dot(np.diag(ssa), Ua.T))
-#     BB[:,:,k] = np.dot(Ub, np.dot(np.diag(ssb), Ub.T))
-#
-# aa = np.random.randn(d,N)
-# bb = np.random.randn(d,N)
-#
-# xx = np.random.rand(N)
-# t = time()
-# for k in range(N):
-#     do_ellipsoids_intersect(AA[:, :, k], BB[:, :, k], aa[:, k], bb[:, k])
-#     # K(xx[k],dd,v)
-# dt = time() - t
-# print('dt=', dt)

### 2021-06-08 code/gaussblur.py

@@ -1,227 +1,0 @@
-import numpy as np
-
-_run_tests = False
-# _run_tests = True
-
-def gauss_exp(A, xi):
-    eta = np.linalg.solve(A, xi)
-    return np.exp(-(1./2.) * np.sum(eta * xi, axis=0))
-
-def gauss_c0(A):
-    d = A.shape[0]
-    return np.power(2. * np.pi, -d / 2.)
-
-def gauss_const(A):
-    return gauss_c0(A) * np.power(np.linalg.det(A), -1./2.)
-
-def gauss(A, xi):
-    return gauss_const(A) * gauss_exp(A, xi)
-
-def dgauss_exp(A,xi, dA):
-    eta = np.linalg.solve(A, xi)
-    return (1./2.) * np.sum(eta * np.dot(dA, eta), axis=0) * gauss_exp(A, xi)
-
-def dgauss_const(A, dA):
-    return -(1. / 2.) * gauss_c0(A) * np.power(np.linalg.det(A), -3./2.) * np.linalg.det(A) * np.trace(np.linalg.solve(A, dA))
-
-def dgauss(A,xi, dA):
-    return dgauss_const(A, dA) * gauss_exp(A,xi) + gauss_const(A) * dgauss_exp(A,xi, dA)
-
-def grad_gauss_exp(A,xi):
-    eta = np.linalg.solve(A, xi)
-    return (1./2.) * np.einsum('n,in,jn->nij', gauss_exp(A, xi), eta, eta)
-
-def grad_gauss_const(A):
-    return -(1. / 2.) * gauss_c0(A) * np.power(np.linalg.det(A), -3./2.) * np.linalg.det(A) * np.linalg.inv(A).T
-
-def grad_gauss(A,xi):
-    return np.einsum('n,ij->nij', gauss_exp(A, xi), grad_gauss_const(A)) + gauss_const(A) * grad_gauss_exp(A, xi)
-
-def periodize_pts(qq, min=np.array([-1,-1]), max=np.array([1,1])):
-    delta = (max - min).reshape((-1, 1))
-    minc = min.reshape((-1, 1))
-    return np.mod(qq - minc, delta) + minc
-
-def periodic_displacement(p,qq, min=np.array([-1,-1]), max=np.array([1,1])):
-    p = p.reshape((-1,1))
-    qq = qq.reshape((p.shape[0], -1))
-    delta = (max-min).reshape((-1,1))
-    minc = min.reshape((-1,1))
-    return np.mod(qq - p - minc, delta) + minc
-
[…]

### 2021-06-08 code/heat_inverse_problem.py

@@ -1,221 +1,0 @@
-import numpy as np
-import fenics
-import scipy.sparse as sps
-import scipy.sparse.linalg as spla
-from fenics_to_scipy_sparse_csr_conversion import convert_fenics_csr_matrix_to_scipy_csr_matrix, vec2fct
-from fenics_interactive_impulse_response_plot import fenics_interactive_impulse_response_plot
-import mshr
-import matplotlib.pyplot as plt
-
-
-def make_amg_solver(A_petsc):
-    prec = fenics.PETScPreconditioner('hypre_amg')
-    fenics.PETScOptions.set('pc_hypre_boomeramg_relax_type_coarse', 'jacobi')
-    solver = fenics.PETScKrylovSolver('cg', prec)
-    solver.set_operator(A_petsc)
-
-    def solve_A(b_petsc, atol=0.0, rtol=1e-10, maxiter=100, verbose=False):
-        x_petsc = fenics.Vector(b_petsc)
-        solver.parameters['absolute_tolerance'] = atol
-        solver.parameters['relative_tolerance'] = rtol
-        solver.parameters['maximum_iterations'] = maxiter
-        solver.parameters['monitor_convergence'] = verbose
-        return solver.solve(x_petsc, b_petsc)
-
-    return solve_A
-
-
-def random_function(function_space_V):
-    x = fenics.Function(function_space_V)
-    x.vector()[:] = np.random.randn(function_space_V.dim())
-    return x
-
-
-class HeatInverseProblem:
-    def __init__(me, mesh_h=1e-2, finite_element_order=1, final_time_T=3e-4, num_timesteps=35, noise_level=5e-2,
-                 perform_checks=True, uniform_kappa=False, lumped_mass_matrix=False):
-        me.mesh_h = mesh_h
-        me.T = final_time_T
-        me.num_timesteps = num_timesteps
-        me.lumped_mass_matrix = lumped_mass_matrix
-
-        me.n = 1. / me.mesh_h
-        me.delta_t = me.T / me.num_timesteps
-
-        outer_circle = mshr.Circle(fenics.Point(0.5, 0.5), 0.5)
-        mesh = mshr.generate_mesh(outer_circle, me.n)
-
-        me.V = fenics.FunctionSpace(mesh, 'CG', finite_element_order)
-        me.d = mesh.geometric_dimension()
-        me.N = me.V.dim()
-
[…]

### 2021-06-08 code/localpsf.py

@@ -1,475 +1,0 @@
-import numpy as np
-import fenics
-import matplotlib.pyplot as plt
-from plot_ellipse import plot_ellipse
-from localpsf_helpers import *
-from poisson_interpolation import PoissonSquaredInterpolation
-from eval_product_convolution import BatchProductConvolution
-from make_fenics_amg_solver import make_fenics_amg_solver
-from mass_matrix import make_mass_matrix
-from fenics_function_fast_grid_evaluator import FenicsFunctionFastGridEvaluator
-import hlibpro_python_wrapper as hpro
-import scipy.sparse as sps
-
-from time import time
-
-
-def make_dirac_comb_dual_vector(points_pp, function_space_V):
-    pp = points_pp
-    V = function_space_V
-    num_pts, d = pp.shape
-    dirac_comb_dual_vector = fenics.assemble(fenics.Constant(0.0) * fenics.TestFunction(V) * fenics.dx)
-    for k in range(num_pts):
-        ps = fenics.PointSource(V, fenics.Point(pp[k,:]), 1.0)
-        ps.apply(dirac_comb_dual_vector)
-    return dirac_comb_dual_vector
-
-
-def get_hessian_dirac_comb_response(points_pp, function_space_V, apply_hessian_H, solve_mass_matrix_M):
-    apply_H = apply_hessian_H
-    solve_M = solve_mass_matrix_M
-
-    dirac_comb_dual_vector = make_dirac_comb_dual_vector(points_pp, function_space_V)
-    dirac_comb_response = fenics.Function(function_space_V)
-    dirac_comb_response.vector()[:] = solve_M(apply_H(solve_M(dirac_comb_dual_vector)))
-    return dirac_comb_response
-
-
-def compute_spatially_varying_volume(function_space_V, apply_hessian_transpose_Ht, solve_mass_matrix_M):
-    V = function_space_V
-    apply_Ht = apply_hessian_transpose_Ht
-    solve_M = solve_mass_matrix_M
-
-    vol = fenics.Function(V)
-    constant_fct = fenics.interpolate(fenics.Constant(1.0), V)
-    vol.vector()[:] = solve_M(apply_Ht(constant_fct.vector()))
-    return vol
-
-
-def compute_spatially_varying_mean(function_space_V, apply_hessain_transpose_Ht,
-                                   solve_mass_matrix_M, volume_function_vol):
-    V = function_space_V
-    apply_Ht = apply_hessain_transpose_Ht
-    solve_M = solve_mass_matrix_M
[…]

### 2021-06-08 code/localpsf_old.py

@@ -1,414 +1,0 @@
-import numpy as np
-import fenics
-import matplotlib.pyplot as plt
-from plot_ellipse import plot_ellipse
-from localpsf_helpers import *
-from poisson_interpolation import PoissonSquaredInterpolation
-from eval_product_convolution import BatchProductConvolution
-from make_fenics_amg_solver import make_fenics_amg_solver
-from mass_matrix import make_mass_matrix
-from fenics_function_fast_grid_evaluator import FenicsFunctionFastGridEvaluator
-import hlibpro_wrapper as hpro
-
-from time import time
-
-
-def make_dirac_comb_dual_vector(points_pp, function_space_V):
-    pp = points_pp
-    V = function_space_V
-    num_pts, d = pp.shape
-    dirac_comb_dual_vector = fenics.assemble(fenics.Constant(0.0) * fenics.TestFunction(V) * fenics.dx)
-    for k in range(num_pts):
-        ps = fenics.PointSource(V, fenics.Point(pp[k,:]), 1.0)
-        ps.apply(dirac_comb_dual_vector)
-    return dirac_comb_dual_vector
-
-
-def get_hessian_dirac_comb_response(points_pp, function_space_V, apply_hessian_H, solve_mass_matrix_M):
-    apply_H = apply_hessian_H
-    solve_M = solve_mass_matrix_M
-
-    dirac_comb_dual_vector = make_dirac_comb_dual_vector(points_pp, function_space_V)
-    dirac_comb_response = fenics.Function(function_space_V)
-    dirac_comb_response.vector()[:] = solve_M(apply_H(solve_M(dirac_comb_dual_vector)))
-    return dirac_comb_response
-
-
-def compute_spatially_varying_volume(function_space_V, apply_hessian_transpose_Ht, solve_mass_matrix_M):
-    V = function_space_V
-    apply_Ht = apply_hessian_transpose_Ht
-    solve_M = solve_mass_matrix_M
-
-    vol = fenics.Function(V)
-    constant_fct = fenics.interpolate(fenics.Constant(1.0), V)
-    vol.vector()[:] = solve_M(apply_Ht(constant_fct.vector()))
-    return vol
-
-
-def compute_spatially_varying_mean(function_space_V, apply_hessain_transpose_Ht,
-                                   solve_mass_matrix_M, volume_function_vol):
-    V = function_space_V
-    apply_Ht = apply_hessain_transpose_Ht
-    solve_M = solve_mass_matrix_M
-    vol = volume_function_vol
-
[…]

### 2021-06-08 code/maxvol.py

@@ -1,690 +1,0 @@
-#https://bitbucket.org/muxas/maxvolpy/src/master/maxvolpy/maxvol.py
-"""
-This submodule contains routines to find good submatrices. How good
-matrix is depends on special extreme properties of the matrix. Two of
-this properties are 1-volume and 2-volume with the following formulas:
-
-:math:`vol_1(A) = \\left|\\det(A)\\right|,\\, vol_2(A) =\
-    \\sqrt{\\max(\\det(A^HA), \\det(AA^H))}`
-"""
-
-from __future__ import absolute_import, division, print_function
-
-__all__ = ['rect_maxvol', 'maxvol', 'rect_maxvol_svd', 'maxvol_svd',
-           'rect_maxvol_qr', 'maxvol_qr']
-
-from misc import svd_cut
-
-
-def py_rect_maxvol(A, tol=1., maxK=None, min_add_K=None, minK=None,
-                   start_maxvol_iters=10, identity_submatrix=True, top_k_index=-1):
-    """
-    Python implementation of rectangular 2-volume maximization.
-
-    See Also
-    --------
-    rect_maxvol
-    """
-    # tol2 - square of parameter tol
-    tol2 = tol ** 2
-    # N - number of rows, r - number of columns of matrix A
-    N, r = A.shape
-    # some work on parameters
-    if N <= r:
-        return np.arange(N, dtype=np.int32), np.eye(N, dtype=A.dtype)
-    if maxK is None or maxK > N:
-        maxK = N
-    if maxK < r:
-        maxK = r
-    if minK is None or minK < r:
-        minK = r
-    if minK > N:
-        minK = N
-    if min_add_K is not None:
-        minK = max(minK, r + min_add_K)
-    if minK > maxK:
-        minK = maxK
-        # raise ValueError('minK value cannot be greater than maxK value')
-    if top_k_index == -1 or top_k_index > N:
-        top_k_index = N
-    if top_k_index < r:
-        top_k_index = r
-    # choose initial submatrix and coefficients according to maxvol
-    # algorithm
-    index = np.zeros(N, dtype=np.int32)
-    chosen = np.ones(top_k_index)
-    tmp_index, C = py_maxvol(A, 1.05, start_maxvol_iters, top_k_index)
-    index[:r] = tmp_index
-    chosen[tmp_index] = 0
-    C = np.asfortranarray(C)
-    # compute square 2-norms of each row in coefficients matrix C
[…]

### 2021-06-08 code/misc.py

@@ -1,87 +1,0 @@
-#https://bitbucket.org/muxas/maxvolpy/src/master/maxvolpy/misc.py
-"""
-:py:mod:`maxvolpy.misc` module contains some auxiliary routines,
-useful when working with different maxvol functions from
-:py:mod:`maxvolpy.maxvol` and when building approximations with help
-of functions from :py:mod:`maxvolpy.cross`.
-"""
-
-from __future__ import print_function, division, absolute_import
-
-__all__ = ['svd_cut', 'reduce_approximation']
-
-import numpy as np
-
-
-def svd_cut(A, tol, alpha=0., norm=2):
-    """
-    Computes SVD and cuts low singular values.
-
-    Computes singular values decomposition of matrix `A`, adds
-    regularizing parameter `alpha` to each singular value and returns
-    only largest singular values and vectors with relative tolerance
-    `tol`.
-
-    Parameters
-    ----------
-    A: numpy.ndarray
-        Real or complex matrix or matrix-like object.
-    tol: float
-        Tolerance of cutting singular values operation.
-    alpha: float, optional
-        Regularizing parameter.
-    norm: {2, 'fro'}, optional
-        Defines norm, that is chosen when cutting singular values.
-
-    Returns
-    -------
-    U: numpy.ndarray
-        Left singular vectors, corresponding to largest singular values.
-    S: numpy.ndarray
-        Largest singular values.
-    V: numpy.ndarray
-        Right singular vectors, corresponding to largest singular values.
-    """
-    U, S, V = np.linalg.svd(A, full_matrices=0)
-    S_reg = S + alpha
-    S1 = S_reg[::-1]
-    if norm == 2:
-        tmp_S = tol * S1[-1]
-        rank = S1.shape[0] - S1.searchsorted(tmp_S, 'left')
-    elif norm == 'fro':
-        S1 = S1 * S1
-        for i in range(1, S1.shape[0]):
-            S1[i] += S1[i - 1]
-        tmp_S = S1[-1] * tol * tol
-        rank = S1.shape[0] - S1.searchsorted(tmp_S, 'left')
-    else:
-        raise ValueError("Invalid parameter norm value")
-    return U[:, :rank].copy(), S_reg[:rank].copy(), V[:rank].copy()
-
-
-def reduce_approximation(U, V, tol, alpha=0.):
-    """
-    Performs `svd_cut` procedure for matrix `U.dot(V)`
-
-    Parameters
-    ----------
[…]

### 2021-06-08 code/point_rejection_sampling.py

@@ -1,69 +1,0 @@
-import numpy as np
-from gaussblur import periodic_displacement, periodize_pts, make_regular_grid_2d
-
-run_test = False
-
-def points_outside_ellipse(x, C, yy, nsigma=1.0, periodic=True):
-    if periodic:
-        z = periodic_displacement(x, yy)
-    else:
-        z = yy - x
-    return (np.sum(z * np.linalg.solve(C, z), axis=0) > nsigma)
-
-def select_well_spaced_points(qq, make_C, nsigma=1.0, periodic=True):
-    perm = np.random.permutation(qq.shape[1])
-    xx = qq[:,perm]
-
-    remaining_inds = np.arange(xx.shape[1])
-    selected_inds = []
-    while len(remaining_inds) > 1:
-        current_ind = remaining_inds[0]
-        remaining_inds = remaining_inds[1:]
-
-        x = xx[:, current_ind].reshape((-1, 1))
-        C = make_C(x)
-
-        selected_points = xx[:, selected_inds].reshape((xx.shape[0], -1))
-        remaining_points = xx[:, remaining_inds].reshape((xx.shape[0], -1))
-        select_x = np.all(points_outside_ellipse(x, C, selected_points, nsigma=nsigma, periodic=periodic))
-
-        if select_x:
-            selected_inds.append(current_ind)
-            remaining_inds = remaining_inds[points_outside_ellipse(x, C, remaining_points, nsigma=nsigma, periodic=periodic)]
-
-    return perm[selected_inds]
-
-
-if run_test:
-    import scipy.linalg as sla
-    import matplotlib.pyplot as plt
-
-    d = 2
-    n = 50
-
-    def sigma_fct(qq):
-        return (0.1 + 0.09 * np.sin(np.pi * qq[0, 0]) * np.cos(np.pi * qq[1, 0]))
-
-    def make_C(x):
-        return sigma_fct(x) * np.eye(x.shape[0])
-
-    def get_ellipse(x0, C):
-        tt = np.linspace(0,2*np.pi,200)
-        unit_circle = np.array([np.sin(tt), np.cos(tt)])
-        return x0 + np.dot(sla.sqrtm(C), unit_circle)
-
-    def plot_ellipse(x0, C):
-        ee = periodize_pts(get_ellipse(x0,C))
-        plt.plot(ee[0,:], ee[1,:],'.k', markersize=1)
-
-    qq = make_regular_grid_2d(n)
[…]

### 2021-06-08 code/point_repulsion.py

@@ -1,114 +1,0 @@
-
-import numpy as np
-import matplotlib.pyplot as plt
-from gaussblur import periodic_displacement, periodize_pts
-
-d=2
-num_pts = 200
-gamma = 0.5
-
-
-def phi(s):
-    return np.power(s, -gamma)
-
-def phi_prime(s):
-    return -gamma * np.power(s, -(gamma+1.))
-
-def force_profile(s):
-    # return 0.05*np.power(s, -1.)
-    return np.exp(-0.5*s)
-
-def get_forces(xx,CC):
-    ff = np.zeros(xx.shape)
-    for i in range(xx.shape[1]):
-        x = xx[:,i]
-        dd = periodic_displacement(x, xx)
-        ss = np.sum(dd * np.linalg.solve(CC[i,:,:], dd), axis=0)
-        ss[i] = 1.
-        ffi = force_profile(ss)
-        ffi[i] = 0.
-        ff = ff + ffi * dd
-    return ff
-
-
-xx = 2. * np.random.rand(d,num_pts) - 1
-
-def sigma_fct(qq):
-    return 2.*(0.07 + 0.04 * np.sin(np.pi * qq[0, :]) * np.cos(np.pi * qq[1, :]))
-
-CC = np.einsum('i,jk->ijk', sigma_fct(xx), np.eye(d))
-
-plt.figure()
-plt.plot(xx[0,:], xx[1,:], '.')
-CC = np.einsum('i,jk->ijk', sigma_fct(xx), np.eye(d))
-ff = get_forces(xx,CC)
-for i in range(num_pts):
-    x = xx[:,i]
-    f = ff[:,i]
-    plt.arrow(x[0], x[1], f[0], f[1])
-
-for _ in range(100):
-    CC = np.einsum('i,jk->ijk', sigma_fct(xx), np.eye(d))
-    ff = get_forces(xx,CC)
-    xx = periodize_pts(xx + 0.01*ff)
-
-###############################################################
-
-# class PointPotential:
-#     def __init__(me, xx, CC):
-#         me.xx = xx
-#         me.CC = CC
-#
-#         me.V = 0.
-#         me.GG = np.zeros(me.xx.shape)
-#         me.recompute()
-#
-#     def update_xx(me, new_xx):
-#         me.xx = new_xx.copy()
-#         me.recompute()
-#
-#     def recompute(me):
-#         me.V = 0.
[…]

### 2021-06-08 code/poisson_weighting_functions.py

@@ -1,108 +1,0 @@
-import numpy as np
-import dolfin as dl
-import scipy.linalg as sla
-import matplotlib.pyplot as plt
-from tqdm.auto import tqdm
-from neumann_poisson_solver import NeumannPoissonSolver
-
-
-def make_poisson_weighting_functions(function_space_V, points_pp):
-    print('Making Poisson weighting functions')
-    initial_points = [points_pp[k,:] for k in range(points_pp.shape[0])]
-    PSI = PoissonSquaredInterpolation(function_space_V, initial_points=initial_points)
-    ww = PSI.weighting_functions
-    return ww
-
-
-def visualize_weighting_function(weighting_functions_ww, points_pp, index_ii):
-    plt.figure()
-    w = weighting_functions_ww[index_ii]
-    cm = dl.plot(w)
-    plt.colorbar(cm)
-    plt.title('Weighting function '+ str(index_ii))
-    plt.scatter(points_pp[:, 0], points_pp[:, 1], c='k', s=2)
-    plt.scatter(points_pp[index_ii, 0], points_pp[index_ii, 1], c='r', s=2)
-
-
-class PoissonSquaredInterpolation:
-    def __init__(me, function_space_V, initial_points=None):
-        me.V = function_space_V
-        me.N = me.V.dim()
-
-        u_trial = dl.TrialFunction(me.V)
-        v_test = dl.TestFunction(me.V)
-
-        mass_form = u_trial * v_test * dl.dx
-        me.M = dl.assemble(mass_form)
-
-        me.constant_one_function = dl.interpolate(dl.Constant(1.0), me.V)
-
-        me.NPPSS = NeumannPoissonSolver(me.V)
-        me.solve_neumann_poisson = me.NPPSS.solve
-        me.solve_neumann_point_source = me.NPPSS.solve_point_source
-
-        me.points = list()
-        me.impulse_responses = list()
-        me.solve_S = lambda x: np.nan
-        me.eta = np.zeros(me.num_pts)
-        me.mu = np.nan
-        me.smooth_basis = list()
-        me.weighting_functions = list()
-
-        if initial_points is not None:
-            me.add_points(initial_points)
-
-    def add_points(me, new_points):
-        me.points = me.points + new_points
-
-        print('Poisson weighting functions: computing Poisson impulse responses')
-        new_impulse_responses = list()
-        for p in tqdm(new_points):
[…]

### 2021-06-08 code/product_convolution_hmatrix.py

@@ -1,783 +1,0 @@
-import numpy as np
-from scipy.spatial import cKDTree
-from scipy.interpolate import interpn
-from tqdm.auto import tqdm
-
-from mass_matrix import make_mass_matrix
-from make_fenics_amg_solver import make_fenics_amg_solver
-
-from nalger_helper_functions import *
-from impulse_response_moments import impulse_response_moments
-from sample_point_batches import choose_sample_point_batches
-from impulse_response_batches import compute_impulse_response_batches
-from poisson_weighting_functions import make_poisson_weighting_functions
-import hlibpro_python_wrapper as hpro
-
-
-def product_convolution_hmatrix(V_in, V_out,
-                                apply_A, apply_At,
-                                num_batches,
-                                tau=2.5,
-                                max_candidate_points=None,
-                                grid_density_multiplier=1.0,
-                                w_support_rtol=2e-2,
-                                num_extension_kernels=8,
-                                hmatrix_tol=1e-4,
-                                bct_admissibility_eta=2.0,
-                                cluster_size_cutoff=50,
-                                use_boundary_extension=True,
-                                make_positive_definite=False,
-                                return_extras=False):
-    '''Builds hierarchical matrix representation of product convolution approximation of operator
-        A: V_in -> V_out
-    using only matrix-vector products with A and A^T.
-
-    Parameters
-    ----------
-    V_in : fenics FunctionSpace
-    V_out : fenics FunctionSpace
-    apply_A : callable. Takes fenics Vector as input and returns fenics Vector as output
-        apply_A(v) computes v -> A*v.
-    apply_At : callable. Takes fenics Vector as input and returns fenics Vector as output
-        apply_At(v) computes v -> A^T*v.
-    num_batches : nonnegative int. Number of batches used in product convolution approximation
-    tau : nonnegative float
-        impulse response ellipsoid size parameter
-        k'th ellipsoid = {x: (x-mu[k,:])^T Sigma[k,:,:]^{-1} (x-mu[k,:]) <= tau}
-        support of k'th impulse response should be contained inside k'th ellipsoid
-    max_candidate_points : nonnegative int. Maximum number of candidate points to consider when choosing sample points
-    grid_density_multiplier : nonnegative float
[…]

### 2021-06-08 code/sample_point_batches.py

@@ -1,108 +1,0 @@
-import numpy as np
-import dolfin as dl
-from numba import jit
-from tqdm.auto import tqdm
-
-from ellipsoid import points_which_are_not_in_ellipsoid_numba, ellipsoids_intersect
-from nalger_helper_functions import *
-
-
-def choose_sample_point_batches(num_batches, V, mu_function, Sigma_function, tau, max_candidate_points=None):
-    '''Chooses several batches of sample points. Uses a greedy algorithm to choose as many sample points as possible per patch,
-     under the constraint that that the supports of impulse responses for a given batch do not overlap.
-
-    Parameters
-    ----------
-    num_batches: nonnegative int. Number of sample point batches
-    mu_function: fenics Function. Vector-valued. spatially varying mean function
-    Sigma_function: fenics Function. Matrix-valued. spatially varying covariance function
-    tau: positive float. Number of standard deviations for ellipsoid.
-        ellipsoid = {x : (x-mu)^T Sigma^-1 (x-mu) <= tau^2}
-    max_candidate_points: (optional) positive int. Maximum number of candidate points to pick sample points points from.
-
-    Returns
-    -------
-    point_batches: list of numpy arrays. Batches of sample points
-        point_batches[b].shape=(num_sample_points_in_batch_b, spatial_dimension).
-    mu_batches: list of numpy arrays. Impulse response means at each sample point
-        mu_batches[b].shape=(num_sample_points_in_batch_b, spatial_dimension).
-    Sigma_batches: list of numpy arrays. Impulse response covariance matrices at each sample point
-        Sigma_batches[b].shape=(num_sample_points, spatial_dimension, spatial_dimension).
-
-    '''
-    # V = mu_function.function_space()
-
-    # dof_coords = np.unique(V.tabulate_dof_coordinates(), axis=0)
-    dof_coords = V.tabulate_dof_coordinates()
-    mu_array = dlfct2array(mu_function)
-    Sigma_array = dlfct2array(Sigma_function)
-
-    if max_candidate_points is None:
-        candidate_inds = np.arange(dof_coords.shape[0])
-    else:
-        candidate_inds = np.random.permutation(V.dim())[:max_candidate_points]
-
-    point_batches = list()
-    mu_batches = list()
-    Sigma_batches = list()
-    print('Choosing sample point batches')
[…]

### 2021-06-08 code/test_localpsf.py

@@ -1,106 +1,0 @@
-from localpsf import *
-import scipy.linalg as sla
-
-#### Test geigh_numpy()
-
-def random_spd_matrix(d):
-    U, ss, _ = np.linalg.svd(np.random.randn(d, d))
-    return np.dot(U, np.dot(np.diag(ss), U.T))
-
-d=5
-A = random_spd_matrix(d)
-B = random_spd_matrix(d)
-
-lambdas_true, Phi_true = eigh(A, B)
-lambdas, Phi = geigh_numpy(A, B)
-
-err_Raya = np.linalg.norm(np.dot(Phi.T, np.dot(A, Phi)) - np.diag(lambdas))
-err_Rayb = np.linalg.norm(np.dot(Phi.T, np.dot(B, Phi)) - np.eye(5))
-
-err_Raya_true = np.linalg.norm(np.dot(Phi_true.T, np.dot(A, Phi_true)) - np.diag(lambdas_true))
-err_Rayb_true = np.linalg.norm(np.dot(Phi_true.T, np.dot(B, Phi_true)) - np.eye(5))
-
-print('err_Raya=', err_Raya, ', err_Rayb=', err_Rayb, ', err_Raya_true=', err_Raya_true, ', err_Rayb_true=', err_Rayb_true)
-
-from time import time
-t = time()
-for k in range(1000):
-    lambdas_true, Phi_true = eigh(A, B)
-dt_eigh = time() - t
-print('dt_eigh=', dt_eigh)
-
-t = time()
-for k in range(1000):
-    lambdas, Phi = geigh_numpy(A, B)
-dt_numba = time() - t
-print('dt_numba=', dt_numba)
-
-
-#### Test ellipsoids_intersect()
-
-def many_points_on_ellipsoid_boundary(Sigma, mu, N, tau):
-    d = Sigma.shape[0]
-    pp = np.random.randn(d, N)
-    pp = tau * (pp / np.linalg.norm(pp, axis=0).reshape((1,-1)))
-    return np.dot(sla.sqrtm(Sigma), pp).T + mu.reshape((1,d))
-
-
-def ellipsoids_intersect_brute_force(Sigma_a, Sigma_b, mu_a, mu_b, tau, num_pts=int(1e5)):
-    pp_a = many_points_on_ellipsoid_boundary(Sigma_a, mu_a, num_pts, tau)
[…]

### 2021-06-08 localpsf/__init__.py

@@ -1,0 +1,7 @@
+from . import ellipsoid
+from . import impulse_response_batches
+from . import impulse_response_moments
+from . import poisson_weighting_functions
+from . import product_convolution_hmatrix
+from . import sample_point_batches
+from . import heat_inverse_problem

### 2021-06-08 localpsf/brent_minimize.py

@@ -1,0 +1,133 @@
+# This is a copy of code in scipy.optimize, except modified for use with numba
+
+# https://www.scipy.org/scipylib/license.html
+# SciPy license
+# Copyright © 2001, 2002 Enthought, Inc.
+# All rights reserved.
+#
+# Copyright © 2003-2019 SciPy Developers.
+# All rights reserved.
+# Redistribution and use in source and binary forms, with or without modification, are permitted provided that the following conditions are met:
+#
+# Redistributions of source code must retain the above copyright notice, this list of conditions and the following disclaimer.
+#
+# Redistributions in binary form must reproduce the above copyright notice, this list of conditions and the following disclaimer in the documentation and/or other materials provided with the distribution.
+#
+# Neither the name of Enthought nor the names of the SciPy Developers may be used to endorse or promote products derived from this software without specific prior written permission.
+#
+# THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS “AS IS” AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE REGENTS OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
+
+
+import numpy as np
+from numba import njit
+
+_cg = 0.3819660
+_mintol = 1e-11
+
+@njit
+def brent_minimize(func, xmin, xmax, args=(), tol=1e-7, maxiter=200):
+    xa = xmin
[…]

### 2021-06-08 localpsf/ellipsoid.py

@@ -1,0 +1,50 @@
+import numpy as np
+from numba import jit
+
+from .brent_minimize import brent_minimize
+
+
+@jit(nopython=True)
+def points_which_are_not_in_ellipsoid_numba(Sigma, mu, pp, tau):
+    d = Sigma.shape[-1]
+    zz = (pp.reshape((-1,d)) - mu.reshape((1,d))).T
+    return (np.sum(zz * np.linalg.solve(Sigma, zz), axis=0) > tau**2)
+
+
+@jit(nopython=True)
+def ellipsoids_intersect(Sigma_a, Sigma_b, mu_a, mu_b, tau):
+    lambdas, Phi, v_squared = ellipsoids_intersect_helper(Sigma_a, Sigma_b, mu_a, mu_b)
+    xmin, fval, iter, funcalls = brent_minimize(K_fct, 0., 1., args=(lambdas, v_squared, tau))
+    return np.bool_(fval >= 0)
+
+
+@jit(nopython=True)
+def ellipsoids_intersect_helper(Sigma_a, Sigma_b, mu_a, mu_b):
+    lambdas, Phi = geigh_numpy(Sigma_a, Sigma_b)
+    v_squared = np.dot(Phi.T, mu_a - mu_b) ** 2
+    return lambdas, Phi, v_squared
+
+
+@jit(nopython=True)
+def K_fct(s, lambdas, v_squared, tau):
+    return 1. - (1. / tau**2) * np.sum(v_squared * ((s * (1. - s)) / (1. + s * (lambdas - 1.))))
+
+
+@jit(nopython=True)
+def K_fct_vectorized(ss, lambdas, v_squared, tau):
+    KK = np.empty(len(ss))
+    for ii in range(len(ss)):
+        KK[ii] = K_fct(ss[ii], lambdas, v_squared, tau)
+    return KK
+
+
+@jit(nopython=True)
+def geigh_numpy(A, B):
+    # Generalized eigenvalue decomposition A*Phi = B*Phi*Lambda
+    # Phi.T * B * Phi = I
+    lambdas0, Phi0 = np.linalg.eig(np.linalg.solve(B, A))
+    lambdas = lambdas0.real
+    Rayleigh_matrix = np.dot(np.dot(Phi0.real.T, B), Phi0.real)
+    scaling_factors = np.sqrt(np.diag(Rayleigh_matrix))
+    Phi = Phi0.real / scaling_factors.reshape((1,-1))
+    return lambdas, Phi

### 2021-06-08 localpsf/heat_inverse_problem.py

@@ -1,0 +1,410 @@
+import numpy as np
+import dolfin as dl
+import scipy.sparse as sps
+import scipy.sparse.linalg as spla
+import matplotlib.pyplot as plt
+from pathlib import Path
+import os
+
+from nalger_helper_functions import *
+
+
+class HeatInverseProblem:
+    def __init__(me,
+                 mesh_h=5e-2,
+                 finite_element_order=1,
+                 final_time=3e-4,
+                 num_timesteps=35,
+                 noise_level=5e-2,
+                 mesh_type='circle',
+                 conductivity_type='wiggly',
+                 initial_condition_type='angel_peak',
+                 perform_checks=True,
+                 make_plots=True,
+                 save_plots=True):
+        ########    INITIALIZE OPTIONS    ########
+
+        me.mesh_h = mesh_h
+        me.finite_element_order = finite_element_order
+        me.final_time = final_time
+        me.num_timesteps = num_timesteps
+        me.noise_level = noise_level
+        me.mesh_type = mesh_type
+        me.conductivity_type = conductivity_type
+        me.initial_condition_type = initial_condition_type
+
+        me.perform_checks = perform_checks
+        me.make_plots = make_plots
+        me.save_plots = save_plots
+
+        me.options = {'mesh_h' : me.mesh_h,
+                      'finite_element_order' : me.finite_element_order,
+                      'final_time' : me.final_time,
+                      'noise_level' : me.noise_level,
+                      'mesh_type' : me.mesh_type,
+                      'conductivity_type' : conductivity_type,
+                      'initial_condition_type' : initial_condition_type}
+
+        project_root = get_project_root()
+        me.identifier_str = str(hash(me))
+        me.save_dir = project_root / 'numerical_results' / 'heat' / me.identifier_str
+        me.data_save_dir = me.save_dir / 'data'
+        me.paraview_save_dir = me.save_dir / 'paraview'
+        me.matplotlib_save_dir = me.save_dir / 'matplotlib'
+
[…]

### 2021-06-08 localpsf/poisson_weighting_functions.py

@@ -1,0 +1,106 @@
+import numpy as np
+import dolfin as dl
+# import scipy.linalg as sla
+import matplotlib.pyplot as plt
+from tqdm.auto import tqdm
+# from neumann_poisson_solver import NeumannPoissonSolver
+
+from nalger_helper_functions import make_dense_lu_solver, NeumannPoissonSolver
+
+
+def make_poisson_weighting_functions(function_space_V, points_pp):
+    print('Making Poisson weighting functions')
+    initial_points = [points_pp[k,:] for k in range(points_pp.shape[0])]
+    PSI = PoissonSquaredInterpolation(function_space_V, initial_points=initial_points)
+    ww = PSI.weighting_functions
+    return ww
+
+
+def visualize_weighting_function(weighting_functions_ww, points_pp, index_ii):
+    plt.figure()
+    w = weighting_functions_ww[index_ii]
+    cm = dl.plot(w)
+    plt.colorbar(cm)
+    plt.title('Weighting function '+ str(index_ii))
+    plt.scatter(points_pp[:, 0], points_pp[:, 1], c='k', s=2)
+    plt.scatter(points_pp[index_ii, 0], points_pp[index_ii, 1], c='r', s=2)
+
+
+class PoissonSquaredInterpolation:
+    def __init__(me, function_space_V, initial_points=None):
+        me.V = function_space_V
+        me.N = me.V.dim()
+
+        u_trial = dl.TrialFunction(me.V)
+        v_test = dl.TestFunction(me.V)
+
+        mass_form = u_trial * v_test * dl.dx
+        me.M = dl.assemble(mass_form)
+
+        me.constant_one_function = dl.interpolate(dl.Constant(1.0), me.V)
+
+        me.NPPSS = NeumannPoissonSolver(me.V)
+        me.solve_neumann_poisson = me.NPPSS.solve
+        me.solve_neumann_point_source = me.NPPSS.solve_point_source
+
+        me.points = list()
+        me.impulse_responses = list()
+        me.solve_S = lambda x: np.nan
+        me.eta = np.zeros(me.num_pts)
+        me.mu = np.nan
+        me.smooth_basis = list()
+        me.weighting_functions = list()
+
+        if initial_points is not None:
+            me.add_points(initial_points)
+
+    def add_points(me, new_points):
+        me.points = me.points + new_points
+
+        print('Poisson weighting functions: computing Poisson impulse responses')
+        new_impulse_responses = list()
+        for p in tqdm(new_points):
[…]

### 2021-06-08 localpsf/product_convolution_hmatrix.py

@@ -1,0 +1,427 @@
+import numpy as np
+from scipy.spatial import cKDTree
+from tqdm.auto import tqdm
+
+from nalger_helper_functions import *
+import hlibpro_python_wrapper as hpro
+
+from .impulse_response_moments import impulse_response_moments
+from .sample_point_batches import choose_sample_point_batches
+from .impulse_response_batches import compute_impulse_response_batches
+from .poisson_weighting_functions import make_poisson_weighting_functions
+
+
+def product_convolution_hmatrix(V_in, V_out,
+                                apply_A, apply_At,
+                                num_batches,
+                                tau=2.5,
+                                max_candidate_points=None,
+                                grid_density_multiplier=1.0,
+                                w_support_rtol=2e-2,
+                                num_extension_kernels=8,
+                                hmatrix_tol=1e-4,
+                                bct_admissibility_eta=2.0,
+                                cluster_size_cutoff=50,
+                                use_boundary_extension=True,
+                                make_positive_definite=False,
+                                return_extras=False):
+    '''Builds hierarchical matrix representation of product convolution approximation of operator
+        A: V_in -> V_out
+    using only matrix-vector products with A and A^T.
+
+    Parameters
+    ----------
+    V_in : fenics FunctionSpace
+    V_out : fenics FunctionSpace
+    apply_A : callable. Takes fenics Vector as input and returns fenics Vector as output
+        apply_A(v) computes v -> A*v.
+    apply_At : callable. Takes fenics Vector as input and returns fenics Vector as output
+        apply_At(v) computes v -> A^T*v.
+    num_batches : nonnegative int. Number of batches used in product convolution approximation
+    tau : nonnegative float
+        impulse response ellipsoid size parameter
+        k'th ellipsoid = {x: (x-mu[k,:])^T Sigma[k,:,:]^{-1} (x-mu[k,:]) <= tau}
+        support of k'th impulse response should be contained inside k'th ellipsoid
+    max_candidate_points : nonnegative int. Maximum number of candidate points to consider when choosing sample points
+    grid_density_multiplier : nonnegative float
+        grid density scaling factor for converting functions on irregular grids to regular grid patches
[…]

### 2021-06-08 localpsf/sample_point_batches.py

@@ -1,0 +1,108 @@
+import numpy as np
+import dolfin as dl
+from numba import jit
+from tqdm.auto import tqdm
+
+from .ellipsoid import points_which_are_not_in_ellipsoid_numba, ellipsoids_intersect
+from nalger_helper_functions import *
+
+
+def choose_sample_point_batches(num_batches, V, mu_function, Sigma_function, tau, max_candidate_points=None):
+    '''Chooses several batches of sample points. Uses a greedy algorithm to choose as many sample points as possible per patch,
+     under the constraint that that the supports of impulse responses for a given batch do not overlap.
+
+    Parameters
+    ----------
+    num_batches: nonnegative int. Number of sample point batches
+    mu_function: fenics Function. Vector-valued. spatially varying mean function
+    Sigma_function: fenics Function. Matrix-valued. spatially varying covariance function
+    tau: positive float. Number of standard deviations for ellipsoid.
+        ellipsoid = {x : (x-mu)^T Sigma^-1 (x-mu) <= tau^2}
+    max_candidate_points: (optional) positive int. Maximum number of candidate points to pick sample points points from.
+
+    Returns
+    -------
+    point_batches: list of numpy arrays. Batches of sample points
+        point_batches[b].shape=(num_sample_points_in_batch_b, spatial_dimension).
+    mu_batches: list of numpy arrays. Impulse response means at each sample point
+        mu_batches[b].shape=(num_sample_points_in_batch_b, spatial_dimension).
+    Sigma_batches: list of numpy arrays. Impulse response covariance matrices at each sample point
+        Sigma_batches[b].shape=(num_sample_points, spatial_dimension, spatial_dimension).
+
+    '''
+    # V = mu_function.function_space()
+
+    # dof_coords = np.unique(V.tabulate_dof_coordinates(), axis=0)
+    dof_coords = V.tabulate_dof_coordinates()
+    mu_array = dlfct2array(mu_function)
+    Sigma_array = dlfct2array(Sigma_function)
+
+    if max_candidate_points is None:
+        candidate_inds = np.arange(dof_coords.shape[0])
+    else:
+        candidate_inds = np.random.permutation(V.dim())[:max_candidate_points]
+
+    point_batches = list()
+    mu_batches = list()
+    Sigma_batches = list()
+    print('Choosing sample point batches')
[…]

### 2021-06-08 old_stuff/covariance_estimation.py

@@ -1,0 +1,75 @@
+import numpy as np
+
+def unit_row_vector(d,k):
+    ek = np.zeros(d)
+    ek[k] = 1.
+    return ek.reshape((1,-1))
+
+def compute_col_moments(apply_At, apply_W, qq):
+    # qq is d-by-N array of point coordinates (d=dimension (e.g., 2 or 3), N=num_pts)
+    # apply_A(V) := A * v
+    # apply_At(v) := A^T * v
+    # apply_W(v) := W * v
+    # A is N-by-N operator, mapping X' -> X
+    # W is N-by-N mass matrix, mapping X -> X'
+    d, N = qq.shape
+    constant_function = np.ones(N)
+    col_volumes = apply_At(apply_W(constant_function))
+    apply_At_rescaled = lambda v: (1. / col_volumes) * apply_At(v)
+
+    col_means = np.zeros((d,N))
+    for ii in range(d):
+        linear_function_ii_direction = np.dot(unit_row_vector(d,ii), qq).reshape(-1)
+        col_means[ii,:] = apply_At_rescaled(apply_W(linear_function_ii_direction))
+
+    col_vars = np.zeros((d,d,N))
+    for ii in range(d):
+        for jj in range(ii+1):
+            if ii == jj:
+                v = unit_row_vector(d, ii)
+            else:
+                v = unit_row_vector(d, ii) + unit_row_vector(d, jj)
+            v = v/np.linalg.norm(v)
+            linear_function_v_direction = np.dot(v, qq).reshape(-1)
+            quadratic_function_v_direction = linear_function_v_direction**2
+            E_x = np.dot(v, col_means)
+            # E_x_b = apply_At_rescaled(apply_W(linear_function_v_direction))
+            # print('err=', np.linalg.norm(E_x_b - E_x))
+            E_x2 = apply_At_rescaled(apply_W(quadratic_function_v_direction))
+            col_vars[ii, jj, :] = E_x2 - E_x**2
+            col_vars[jj, ii, :] = col_vars[ii, jj, :]
+
+    for ii in range(d):
+        for jj in range(d):
+            if not (ii == jj):
+                col_vars[ii, jj, :] = col_vars[ii, jj, :] - col_vars[ii, ii, :]/2. - col_vars[jj, jj, :]/2.
+
+    return col_volumes, col_means, col_vars
+
+
+def distance_from_boundary_function(qq, boundary_inds):
+    dd = np.zeros(qq.shape[1])
[…]

### 2021-06-08 old_stuff/gaussblur.py

@@ -1,0 +1,227 @@
+import numpy as np
+
+_run_tests = False
+# _run_tests = True
+
+def gauss_exp(A, xi):
+    eta = np.linalg.solve(A, xi)
+    return np.exp(-(1./2.) * np.sum(eta * xi, axis=0))
+
+def gauss_c0(A):
+    d = A.shape[0]
+    return np.power(2. * np.pi, -d / 2.)
+
+def gauss_const(A):
+    return gauss_c0(A) * np.power(np.linalg.det(A), -1./2.)
+
+def gauss(A, xi):
+    return gauss_const(A) * gauss_exp(A, xi)
+
+def dgauss_exp(A,xi, dA):
+    eta = np.linalg.solve(A, xi)
+    return (1./2.) * np.sum(eta * np.dot(dA, eta), axis=0) * gauss_exp(A, xi)
+
+def dgauss_const(A, dA):
+    return -(1. / 2.) * gauss_c0(A) * np.power(np.linalg.det(A), -3./2.) * np.linalg.det(A) * np.trace(np.linalg.solve(A, dA))
+
+def dgauss(A,xi, dA):
+    return dgauss_const(A, dA) * gauss_exp(A,xi) + gauss_const(A) * dgauss_exp(A,xi, dA)
+
+def grad_gauss_exp(A,xi):
+    eta = np.linalg.solve(A, xi)
+    return (1./2.) * np.einsum('n,in,jn->nij', gauss_exp(A, xi), eta, eta)
+
+def grad_gauss_const(A):
+    return -(1. / 2.) * gauss_c0(A) * np.power(np.linalg.det(A), -3./2.) * np.linalg.det(A) * np.linalg.inv(A).T
+
+def grad_gauss(A,xi):
+    return np.einsum('n,ij->nij', gauss_exp(A, xi), grad_gauss_const(A)) + gauss_const(A) * grad_gauss_exp(A, xi)
+
+def periodize_pts(qq, min=np.array([-1,-1]), max=np.array([1,1])):
+    delta = (max - min).reshape((-1, 1))
+    minc = min.reshape((-1, 1))
+    return np.mod(qq - minc, delta) + minc
+
+def periodic_displacement(p,qq, min=np.array([-1,-1]), max=np.array([1,1])):
+    p = p.reshape((-1,1))
+    qq = qq.reshape((p.shape[0], -1))
+    delta = (max-min).reshape((-1,1))
+    minc = min.reshape((-1,1))
+    return np.mod(qq - p - minc, delta) + minc
+
[…]

### 2021-06-08 old_stuff/localpsf.py

@@ -1,0 +1,475 @@
+import numpy as np
+import fenics
+import matplotlib.pyplot as plt
+from plot_ellipse import plot_ellipse
+from localpsf_helpers import *
+from poisson_interpolation import PoissonSquaredInterpolation
+from eval_product_convolution import BatchProductConvolution
+from make_fenics_amg_solver import make_fenics_amg_solver
+from mass_matrix import make_mass_matrix
+from fenics_function_fast_grid_evaluator import FenicsFunctionFastGridEvaluator
+import hlibpro_python_wrapper as hpro
+import scipy.sparse as sps
+
+from time import time
+
+
+def make_dirac_comb_dual_vector(points_pp, function_space_V):
+    pp = points_pp
+    V = function_space_V
+    num_pts, d = pp.shape
+    dirac_comb_dual_vector = fenics.assemble(fenics.Constant(0.0) * fenics.TestFunction(V) * fenics.dx)
+    for k in range(num_pts):
+        ps = fenics.PointSource(V, fenics.Point(pp[k,:]), 1.0)
+        ps.apply(dirac_comb_dual_vector)
+    return dirac_comb_dual_vector
+
+
+def get_hessian_dirac_comb_response(points_pp, function_space_V, apply_hessian_H, solve_mass_matrix_M):
+    apply_H = apply_hessian_H
+    solve_M = solve_mass_matrix_M
+
+    dirac_comb_dual_vector = make_dirac_comb_dual_vector(points_pp, function_space_V)
+    dirac_comb_response = fenics.Function(function_space_V)
+    dirac_comb_response.vector()[:] = solve_M(apply_H(solve_M(dirac_comb_dual_vector)))
+    return dirac_comb_response
+
+
+def compute_spatially_varying_volume(function_space_V, apply_hessian_transpose_Ht, solve_mass_matrix_M):
+    V = function_space_V
+    apply_Ht = apply_hessian_transpose_Ht
+    solve_M = solve_mass_matrix_M
+
+    vol = fenics.Function(V)
+    constant_fct = fenics.interpolate(fenics.Constant(1.0), V)
+    vol.vector()[:] = solve_M(apply_Ht(constant_fct.vector()))
+    return vol
+
+
+def compute_spatially_varying_mean(function_space_V, apply_hessain_transpose_Ht,
+                                   solve_mass_matrix_M, volume_function_vol):
+    V = function_space_V
+    apply_Ht = apply_hessain_transpose_Ht
+    solve_M = solve_mass_matrix_M
[…]

### 2021-06-08 old_stuff/misc.py

@@ -1,0 +1,87 @@
+#https://bitbucket.org/muxas/maxvolpy/src/master/maxvolpy/misc.py
+"""
+:py:mod:`maxvolpy.misc` module contains some auxiliary routines,
+useful when working with different maxvol functions from
+:py:mod:`maxvolpy.maxvol` and when building approximations with help
+of functions from :py:mod:`maxvolpy.cross`.
+"""
+
+from __future__ import print_function, division, absolute_import
+
+__all__ = ['svd_cut', 'reduce_approximation']
+
+import numpy as np
+
+
+def svd_cut(A, tol, alpha=0., norm=2):
+    """
+    Computes SVD and cuts low singular values.
+
+    Computes singular values decomposition of matrix `A`, adds
+    regularizing parameter `alpha` to each singular value and returns
+    only largest singular values and vectors with relative tolerance
+    `tol`.
+
+    Parameters
+    ----------
+    A: numpy.ndarray
+        Real or complex matrix or matrix-like object.
+    tol: float
+        Tolerance of cutting singular values operation.
+    alpha: float, optional
+        Regularizing parameter.
+    norm: {2, 'fro'}, optional
+        Defines norm, that is chosen when cutting singular values.
+
+    Returns
+    -------
+    U: numpy.ndarray
+        Left singular vectors, corresponding to largest singular values.
+    S: numpy.ndarray
+        Largest singular values.
+    V: numpy.ndarray
+        Right singular vectors, corresponding to largest singular values.
+    """
+    U, S, V = np.linalg.svd(A, full_matrices=0)
+    S_reg = S + alpha
+    S1 = S_reg[::-1]
+    if norm == 2:
+        tmp_S = tol * S1[-1]
+        rank = S1.shape[0] - S1.searchsorted(tmp_S, 'left')
+    elif norm == 'fro':
+        S1 = S1 * S1
+        for i in range(1, S1.shape[0]):
+            S1[i] += S1[i - 1]
+        tmp_S = S1[-1] * tol * tol
+        rank = S1.shape[0] - S1.searchsorted(tmp_S, 'left')
+    else:
+        raise ValueError("Invalid parameter norm value")
+    return U[:, :rank].copy(), S_reg[:rank].copy(), V[:rank].copy()
+
+
+def reduce_approximation(U, V, tol, alpha=0.):
+    """
+    Performs `svd_cut` procedure for matrix `U.dot(V)`
+
+    Parameters
+    ----------
[…]

### 2021-06-08 old_stuff/point_rejection_sampling.py

@@ -1,0 +1,69 @@
+import numpy as np
+from gaussblur import periodic_displacement, periodize_pts, make_regular_grid_2d
+
+run_test = False
+
+def points_outside_ellipse(x, C, yy, nsigma=1.0, periodic=True):
+    if periodic:
+        z = periodic_displacement(x, yy)
+    else:
+        z = yy - x
+    return (np.sum(z * np.linalg.solve(C, z), axis=0) > nsigma)
+
+def select_well_spaced_points(qq, make_C, nsigma=1.0, periodic=True):
+    perm = np.random.permutation(qq.shape[1])
+    xx = qq[:,perm]
+
+    remaining_inds = np.arange(xx.shape[1])
+    selected_inds = []
+    while len(remaining_inds) > 1:
+        current_ind = remaining_inds[0]
+        remaining_inds = remaining_inds[1:]
+
+        x = xx[:, current_ind].reshape((-1, 1))
+        C = make_C(x)
+
+        selected_points = xx[:, selected_inds].reshape((xx.shape[0], -1))
+        remaining_points = xx[:, remaining_inds].reshape((xx.shape[0], -1))
+        select_x = np.all(points_outside_ellipse(x, C, selected_points, nsigma=nsigma, periodic=periodic))
+
+        if select_x:
+            selected_inds.append(current_ind)
+            remaining_inds = remaining_inds[points_outside_ellipse(x, C, remaining_points, nsigma=nsigma, periodic=periodic)]
+
+    return perm[selected_inds]
+
+
+if run_test:
+    import scipy.linalg as sla
+    import matplotlib.pyplot as plt
+
+    d = 2
+    n = 50
+
+    def sigma_fct(qq):
+        return (0.1 + 0.09 * np.sin(np.pi * qq[0, 0]) * np.cos(np.pi * qq[1, 0]))
+
+    def make_C(x):
+        return sigma_fct(x) * np.eye(x.shape[0])
+
+    def get_ellipse(x0, C):
+        tt = np.linspace(0,2*np.pi,200)
+        unit_circle = np.array([np.sin(tt), np.cos(tt)])
+        return x0 + np.dot(sla.sqrtm(C), unit_circle)
+
+    def plot_ellipse(x0, C):
+        ee = periodize_pts(get_ellipse(x0,C))
+        plt.plot(ee[0,:], ee[1,:],'.k', markersize=1)
+
+    qq = make_regular_grid_2d(n)
[…]

### 2021-06-08 old_stuff/point_repulsion.py

@@ -1,0 +1,114 @@
+
+import numpy as np
+import matplotlib.pyplot as plt
+from gaussblur import periodic_displacement, periodize_pts
+
+d=2
+num_pts = 200
+gamma = 0.5
+
+
+def phi(s):
+    return np.power(s, -gamma)
+
+def phi_prime(s):
+    return -gamma * np.power(s, -(gamma+1.))
+
+def force_profile(s):
+    # return 0.05*np.power(s, -1.)
+    return np.exp(-0.5*s)
+
+def get_forces(xx,CC):
+    ff = np.zeros(xx.shape)
+    for i in range(xx.shape[1]):
+        x = xx[:,i]
+        dd = periodic_displacement(x, xx)
+        ss = np.sum(dd * np.linalg.solve(CC[i,:,:], dd), axis=0)
+        ss[i] = 1.
+        ffi = force_profile(ss)
+        ffi[i] = 0.
+        ff = ff + ffi * dd
+    return ff
+
+
+xx = 2. * np.random.rand(d,num_pts) - 1
+
+def sigma_fct(qq):
+    return 2.*(0.07 + 0.04 * np.sin(np.pi * qq[0, :]) * np.cos(np.pi * qq[1, :]))
+
+CC = np.einsum('i,jk->ijk', sigma_fct(xx), np.eye(d))
+
+plt.figure()
+plt.plot(xx[0,:], xx[1,:], '.')
+CC = np.einsum('i,jk->ijk', sigma_fct(xx), np.eye(d))
+ff = get_forces(xx,CC)
+for i in range(num_pts):
+    x = xx[:,i]
+    f = ff[:,i]
+    plt.arrow(x[0], x[1], f[0], f[1])
+
+for _ in range(100):
+    CC = np.einsum('i,jk->ijk', sigma_fct(xx), np.eye(d))
+    ff = get_forces(xx,CC)
+    xx = periodize_pts(xx + 0.01*ff)
+
+###############################################################
+
+# class PointPotential:
+#     def __init__(me, xx, CC):
+#         me.xx = xx
+#         me.CC = CC
+#
+#         me.V = 0.
+#         me.GG = np.zeros(me.xx.shape)
+#         me.recompute()
+#
+#     def update_xx(me, new_xx):
+#         me.xx = new_xx.copy()
+#         me.recompute()
+#
+#     def recompute(me):
+#         me.V = 0.
[…]

### 2021-06-08 old_stuff/product_convolution_operator_old.py

@@ -1,0 +1,782 @@
+import numpy as np
+from scipy.spatial import cKDTree
+from tqdm.auto import tqdm
+
+from localpsf_helper_functions import make_fenics_amg_solver
+
+from nalger_helper_functions import *
+from impulse_response_moments import impulse_response_moments
+from sample_point_batches import choose_sample_point_batches
+from impulse_response_batches import compute_impulse_response_batches
+from poisson_weighting_functions import make_poisson_weighting_functions
+import hlibpro_python_wrapper as hpro
+
+
+def product_convolution_hmatrix(V_in, V_out,
+                                apply_A, apply_At,
+                                num_batches,
+                                tau=2.5,
+                                max_candidate_points=None,
+                                grid_density_multiplier=1.0,
+                                w_support_rtol=2e-2,
+                                num_extension_kernels=8,
+                                hmatrix_tol=1e-4,
+                                bct_admissibility_eta=2.0,
+                                cluster_size_cutoff=50,
+                                use_boundary_extension=True,
+                                make_positive_definite=False,
+                                return_extras=False):
+    '''Builds hierarchical matrix representation of product convolution approximation of operator
+        A: V_in -> V_out
+    using only matrix-vector products with A and A^T.
+
+    Parameters
+    ----------
+    V_in : fenics FunctionSpace
+    V_out : fenics FunctionSpace
+    apply_A : callable. Takes fenics Vector as input and returns fenics Vector as output
+        apply_A(v) computes v -> A*v.
+    apply_At : callable. Takes fenics Vector as input and returns fenics Vector as output
+        apply_At(v) computes v -> A^T*v.
+    num_batches : nonnegative int. Number of batches used in product convolution approximation
+    tau : nonnegative float
+        impulse response ellipsoid size parameter
+        k'th ellipsoid = {x: (x-mu[k,:])^T Sigma[k,:,:]^{-1} (x-mu[k,:]) <= tau}
+        support of k'th impulse response should be contained inside k'th ellipsoid
+    max_candidate_points : nonnegative int. Maximum number of candidate points to consider when choosing sample points
+    grid_density_multiplier : nonnegative float
[…]

### 2021-06-08 setup.py

@@ -1,0 +1,10 @@
+from distutils.core import setup
+
+setup(
+    name='localpsf',
+    version='0.1dev',
+    packages=['localpsf',],
+    license='MIT',
+    author='Nick Alger',
+    # long_description=open('README.txt').read(),
+)

### 2021-07-20 paper/localpsf.tex
wrote rational positive definite modification section

@@
 
-In order to use the product-convolution approximation $\widetilde{\mathcal{A}}$, we convert $\widetilde{\mathbf{A}}$ (the discretized version of $\widetilde{\mathcal{A}}$) to hierarchical matrix (H-matrix) format.
-Many classes of hierarchical matrices exist---H1, uniform H1, H2, HODLR, HSS, HBS, and more \cite{HMATRIX}.
-We use H1 matrices in our numerical results, but any of the other H-matrix formats could be used instead.
+We convert $\widetilde{\mathbf{A}}$ (the discretized version of $\widetilde{\mathcal{A}}$) to hierarchical matrix (H-matrix) format.
+Once in H-matrix format, useful matrix operations such as matrix factorization and inversion may be performed quickly using fast H-matrix methods.
+We use H1 matrices in our numerical results, but any of the other H-matrix formats (such as H2, HODLR, HSS, HBS, and others \cite{HMATRIX}) could be used instead.
 The basic ideas behind H-matrices are as follows:
@@
 We describe the H-matrix construction in detail in Appendix \ref{app:h_matrix}.
+
+\subsection{Rational positive semi-definite modification}
+
+In many problems of practical interest (e.g., Hessian approximation) $\mathcal{A}$ is symmetric positive semi-definite.
+However, $\widetilde{\mathbf{A}}$ is generally non-symmetric and indefinite because of errors in the product-convolution approximation.
+This is undesirable because symmetry and positive semi-definiteness are important properties of an operator which should be preserved by approximations, if possible.
+Moreover, a lack of symmetry and/or positive semi-definiteness in $\widetilde{\mathbf{A}}$ may prevent one from using many popular and highly effective algorithms to perform further useful matrix operations involving $\widetilde{\mathbf{A}}$.
+To address this issue, here we describe a method for modifying $\widetilde{\mathbf{A}}$ to force it to be symmetric positive semi-definite.
+
+We use fast H-matrix addition to symmetrize $\widetilde{\mathbf{A}}$:
+$$\widetilde{\mathbf{A}}_\text{sym} := \frac{1}{2}\left(\widetilde{\mathbf{A}} + \widetilde{\mathbf{A}}^T\right).$$
+Then we modify $\widetilde{\mathbf{A}}_\text{sym}$ to make it positive definite by forming a rational matrix function of the following form: 
+\begin{equation}
+\label{eq:rational_matrix_function}
+\widetilde{\mathbf{A}}_\text{sym}^+ := c_0 + c_1 \widetilde{\mathbf{A}}_\text{sym} + c_2 \left(\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}\right)^{-1},
+\end{equation}
+where $\mathbf{I}$ is the identity matrix which has the same shape as $\widetilde{\mathbf{A}}_\text{sym}$.
+The scalars $c_0$, $c_1$, $c_2$, and $\mu$ are chosen so that the spectrum of $\widetilde{\mathbf{A}}_\text{sym}^+$ has certain desirable properties, which we will describe later in this section.
+Forming $\widetilde{\mathbf{A}}_\text{sym}^+$ requires computing the H-matrix inverse $\left(\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}\right)^{-1}$.
+While explicit computation of matrix inverses should typically be avoided, here it is fine to compute the inverse because we are inverting an H-matrix, and there are fast, stable, scalable, and widely accepted algorithms for inverting H-matrices \cite{Hmatrixinverse}.
+
+Our choice of the constants $c_0$, $c_1$, $c_2$, and $\mu$ will depend on upper and lower bounds for the spectrum of $\widetilde{\mathbf{A}}_\text{sym}$.
+Let $m$ and $M$ be the smallest and largest eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$, respectively, so that the interval $[m, M] \subset \mathbb{R}$ contains the spectrum of $\widetilde{\mathbf{A}}_\text{sym}$.
+Note that $m < 0$ since $\widetilde{\mathbf{A}}_\text{sym}$ is indefinite.
+%Also note that $|m| \ll |M|$ if $\widetilde{A}$ is positive definite, since negative eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ are a result of approximation error. 
+We estimate $m$ and $M$ using the implicitly restarted Lanczos method [CITE], a matrix-free Krylov method.
+This is fast because computing matrix-vector products with an H-matrix is fast, and because the eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ do not cluster near $m$ or $M$ (though the eigenvalues may cluster at zero).
+Other matrix-free extremal eigenvalue methods, such as LOBPCG or shifted power iteration, could also be used.
+
+The rational matrix approximation \eqref{eq:rational_matrix_function} may be written as $\widetilde{\mathbf{A}}_\text{sym}^+ = f(\widetilde{\mathbf{A}}_\text{sym})$, where $f$ is the rational function
+\begin{equation}
+\label{eq:rational_function_scalar}
+f(\lambda) := c_0 + c_1 \lambda + \frac{c_2}{\lambda+\mu}.
+\end{equation}
+We choose the four unknown constants, $c_0, c_1, c_2$, and $\mu$, to be the solution to the following system of four equations:
+\begin{align}
+f(0) &= 0 \label{eq:desired_property_1} \\
+f'(0) &= 0 \label{eq:desired_property_2} \\
+f(m) &= |m| \label{eq:desired_property_3} \\
+f(M) &= M \label{eq:desired_property_4}
+\end{align}
+with the further condition that $\mu > |m|$ (so that the pole of $f$ is not in $[m,M]$).
+Together, these equations imply that $f(0)=0$ is the unique minimum of $f$ on $[m,M]$, which implies that $f$ is non-negative on $[m,M]$, which implies that  $f(\widetilde{\mathbf{A}}_\text{sym}) = \widetilde{\mathbf{A}}_\text{sym}^+$ is positive semi-definite.
+The equation $f(0)=0$ implies that the null space of $\widetilde{\mathbf{A}}_\text{sym}^+$ contains the null-space of $\widetilde{\mathbf{A}}_\text{sym}$.
+This is important when the operator $\mathcal{A}$ has a large (or infinite) dimensional null space, or a spectrum that clusters at zero.
+Hessians in ill-posed distributed parameter inverse problems typically have both of these properties.
+The equation $f(M)=M$ forces the moderate and large positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}^+$ to be close to the corresponding moderate and large positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$.
+I.e., $f$ modifies the ``important'' part of the spectrum of $\widetilde{\mathbf{A}}_\text{sym}$ as little as possible.
+The equation $f(m)=|m|$ ensures that the negative eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $f$.
+There is a tradeoff: if we choose a larger value for $f(m)$, then the positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}^+$ will be closer to the corresponding positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ (good), but the erroneous negative eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ will be amplified more in magnitude (bad).
+We find that $f(m)=|m|$ is an appropriate balance between these competing interests.
+
+Using the definition of $f$ in \eqref{eq:rational_function_scalar}, and basic calculus and linear algebra, we may write Equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, and \eqref{eq:desired_property_4} in the following matrix form:
+\begin{equation}
+\label{eq:rational_3_linear_system}
+\begin{bmatrix} 1 & 0 & \frac{1}{\mu} \\ 0 & 1 & -\frac{1}{\mu^2} \\ 1 & M & \frac{1}{M+\mu} \end{bmatrix} \begin{bmatrix} c_0 \\ c_1 \\ c_2 \end{bmatrix} = \begin{bmatrix} 0 \\ 0 \\ M \end{bmatrix},
+\end{equation}
+and we may write Equation \eqref{eq:desired_property_3} as follows: 
+\begin{equation}
+\label{eq:rational_extra_eq}
+c_0 + c_1 m + \frac{c_2}{m+\mu} - |m| = 0.
+\end{equation}
+Let 
+\begin{equation*}
+g(\mu) := c_0(\mu) + c_1(\mu) m + \frac{c_2(\mu)}{m+\mu} - |m|,
+\end{equation*}
+where $c_0(\mu)$, $c_1(\mu)$, and $c_2(\mu)$ denote the values of $c_0$, $c_1$, and $c_2$, respectively, that solve the linear system $\eqref{eq:rational_3_linear_system}$ for a given value of $\mu$.
+That is, $g(\mu)$ is the left hand side of \eqref{eq:rational_extra_eq}, except instead of being free variables, $c_0$, $c_1$, and $c_2$ now depend on $\mu$ through the solution of \eqref{eq:rational_3_linear_system}.
+We solve the combined system, \eqref{eq:rational_3_linear_system} and $\eqref{eq:rational_extra_eq}$, by solving the one-dimensional nonlinear root finding problem
+\begin{equation}
+\label{eq:g_mu_zero}
+g(\mu)=0.
+\end{equation}
+with bound $\mu > |m|$ and initial guess $\mu_0 = 2 |m|$.
+Each evaluation of $g$ implicitly requires solving the $3 \times 3$ linear system \eqref{eq:rational_3_linear_system}, and once the solution $\mu$ is found, we must solve \eqref{eq:rational_3_linear_system} one more time to get $c_0$, $c_1$, and $c_2$.
+%
+%Row 1 corresponds to the condition $f(M)=M$, row 2 corresponds to the condition $f(0)=0$, and row 3 corresponds to the condition $f'(0)=0$.
+To find the third row of $A$, we used the fact that the derivative of $f$ is given by $$f'(\lambda) = c_1 - \frac{c_2}{(\lambda+\mu)^2}.$$
+%
+%
+%We want $f(\lambda)$ to have the following desirable properties:
+%\begin{enumerate}
+%	\item $f(\lambda)$ is nonnegative on $[m, M]$.
+%	\item $f(\lambda) \approx \lambda$ when $\lambda$ is positive and not small.
+%	\item $f(0)=0$ is the minimum of $f(\lambda)$ on $[m, M]$.
+%	\item $f(\lambda) \le |\lambda|$ when $\lambda$ is negative.
+%\end{enumerate}
+%Property 1 ensures that $\widetilde{\mathbf{A}}_\text{sym}^+$ is positive semi-definite.
+Property 2 ensures that $f$ modifies the ``important'' part of the spectrum of $\widetilde{\mathbf{A}}_\text{sym}$ as little as possible.
+That is, the part of the spectrum of $\mathcal{A}$ that we expect to be well-approximated by the product-convolution approximation.
+Property 3 ensures that $\widetilde{\mathbf{A}}_\text{sym}^+$ preserves the null-space (and near null-space) of $\widetilde{\mathbf{A}}_\text{sym}$.
+This is important because in practice, we often want to approximate operators $\mathcal{A}$ that have a large (possibly infinite) number of zero eigenvalues, and/or a sequence of eigenvalues that clusters at zero.
+Hessians in ill-posed inverse problems typically have both of these properties.
+Property 4 ensures that $f$ does not amplify erroneous negative eigenvalues $\widetilde{\mathbf{A}}_\text{sym}$.
+If $\mathcal{A}$ is positive semi-definite, then eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ can only be negative because of errors in the product-convolution approximation.
 

### 2021-07-20 numerical_examples/heat/heat_basic_plots.py

@@ -1,0 +1,130 @@
+import numpy as np
+import dolfin as dl
+import time
+import pathlib
+from tqdm.auto import tqdm
+
+from nalger_helper_functions import *
+import hlibpro_python_wrapper as hpro
+from localpsf.heat_inverse_problem import *
+from localpsf.impulse_response_batches import visualize_impulse_response_batch
+from localpsf.product_convolution_hmatrix import product_convolution_hmatrix
+
+
+########    OPTIONS    ########
+
+HIP_options = {'mesh_h' : 1.5e-2,
+               'finite_element_order' : 1,
+               'final_time' : 3e-4,
+               'num_timesteps' : 35,
+               'noise_level' : 5e-2,
+               'mesh_type' : 'circle',
+               'conductivity_type' : 'wiggly',
+               'initial_condition_type' : 'angel_peak',
+               'perform_checks' : True,
+               'prior_correlation_length' : 0.05,
+               'regularization_parameter' : 1e-1}
+
+hmatrix_rtol = 1e-4
+krylov_tol = 1e-6
+a_reg_min = 1e-5
+a_reg_max = 1e0
+num_reg = 11
+hmatrix_num_batches = 5
+n_random_error_matvecs = 50
+
+save_dir = get_project_root() / 'numerical_examples' / 'heat' / time.ctime()
+save_dir.mkdir(parents=True, exist_ok=True)
+
+options_file = save_dir / 'options.txt'
+with open(options_file, 'w') as f:
+    print(HIP_options, file=f)
+
+
+########    SET UP HEAT INVERSE PROBLEM    ########
+
+HIP = HeatInverseProblem(**HIP_options)
+
+# Mesh
+
+plt.figure()
+dl.plot(HIP.mesh)
+plt.title('mesh')
+plt.savefig(str(save_dir / 'mesh.pdf'), bbox_inches='tight', dpi=100)
+
+file = dl.XDMFFile(str(save_dir / 'mesh.xdmf'))
+file.write(HIP.mesh)
+file.close()
+
+# Conductivity kappa
+
+plt.figure()
+cm = dl.plot(HIP.kappa, cmap='gray')
+plt.colorbar(cm)
+plt.title(r'Thermal conductivity, $\kappa$')
[…]

### 2021-07-20 paper/localpsf.tex
edits to rational positive definite section

--- 2021-06-08 paper/localpsf.tex
@@
 We describe the H-matrix construction in detail in Appendix \ref{app:h_matrix}.
+
+\subsection{Rational positive semi-definite modification}
+
+In many problems of practical interest (e.g., Hessian approximation) $\mathcal{A}$ is symmetric positive semi-definite.
+However, $\widetilde{\mathbf{A}}$ is generally non-symmetric and indefinite because of errors in the product-convolution approximation.
+This is undesirable.
+Symmetry and positive semi-definiteness are important properties which should be preserved by the approximation, if possible.
+Also, a lack of symmetry and/or positive semi-definiteness in $\widetilde{\mathbf{A}}$ may prevent one from using popular and highly effective algorithms, such as the conjugate gradient method, to perform further useful operations involving $\widetilde{\mathbf{A}}$.
+In this section we describe a method for modifying $\widetilde{\mathbf{A}}$ to make it symmetric positive semi-definite.
+
+First, we use fast H-matrix addition to symmetrize $\widetilde{\mathbf{A}}$:
+$$\widetilde{\mathbf{A}}_\text{sym} := \frac{1}{2}\left(\widetilde{\mathbf{A}} + \widetilde{\mathbf{A}}^T\right).$$
+Then we modify $\widetilde{\mathbf{A}}_\text{sym}$ to make it positive definite by forming a rational matrix function of the following form: 
+\begin{equation}
+\label{eq:rational_matrix_function}
+\widetilde{\mathbf{A}}_\text{sym}^+ := c_0 + c_1 \widetilde{\mathbf{A}}_\text{sym} + c_2 \left(\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}\right)^{-1},
+\end{equation}
+where $\mathbf{I}$ is the identity matrix which has the same shape as $\widetilde{\mathbf{A}}_\text{sym}$.
+The scalars $c_0$, $c_1$, $c_2$, and $\mu$ are chosen so that the spectrum of $\widetilde{\mathbf{A}}_\text{sym}^+$ has certain desirable properties, which we will describe later in this section.
+Forming $\widetilde{\mathbf{A}}_\text{sym}^+$ requires computing an H-matrix inverse, $\left(\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}\right)^{-1}$.
+While explicit computation of matrix inverses should typically be avoided, here it is fine to compute the inverse because $\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}$ is an H-matrix, and there are fast, stable, scalable, and widely used algorithms for inverting H-matrices \cite{Hmatrixinverse}.
+
+The rational matrix approximation \eqref{eq:rational_matrix_function} may be written as $\widetilde{\mathbf{A}}_\text{sym}^+ = f(\widetilde{\mathbf{A}}_\text{sym})$, where $f$ is the rational function
+\begin{equation}
+\label{eq:rational_function_scalar}
+f(\lambda) := c_0 + c_1 \lambda + \frac{c_2}{\lambda+\mu}.
+\end{equation}
+We choose the four unknown constants, $c_0, c_1, c_2$, and $\mu$, to be the solution to the following system of four equations:
+\begin{align}
+f(0) &= 0 \label{eq:desired_property_1} \\
+f'(0) &= 0 \label{eq:desired_property_2} \\
+f(\lambda_\text{min}) &= |\lambda_\text{min}| \label{eq:desired_property_3} \\
+f(\lambda_\text{max}) &= \lambda_\text{max}, \label{eq:desired_property_4}
+\end{align}
+where $\lambda_\text{min}$ and $\lambda_\text{max}$ are the smallest and largest eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$, respectively.
+We also require $\mu > |\lambda_\text{min}|$, so that the pole of $f$ is not in $[\lambda_\text{min},\lambda_\text{max}]$.
+Note that $\lambda_\text{min} < 0$ since $\widetilde{\mathbf{A}}_\text{sym}$ is indefinite.
+We estimate $\lambda_\text{min}$ and $\lambda_\text{max}$ using the implicitly restarted Lanczos method \cite{lanczos,scipy}, a matrix-free Krylov method.
+This is fast because computing matrix-vector products with an H-matrix is fast, and because the eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ do not cluster near $\lambda_\text{min}$ or $\lambda_\text{max}$ (though the eigenvalues may cluster at zero).
+
+Together, equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, \eqref{eq:desired_property_3},  and \eqref{eq:desired_property_4}, and the condition $\mu > |\lambda_\text{min}|$, imply that $f(0)=0$ is the unique minimum of $f$ on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that $f$ is non-negative on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that  $f(\widetilde{\mathbf{A}}_\text{sym}) = \widetilde{\mathbf{A}}_\text{sym}^+$ is positive semi-definite.
+The equation $f(0)=0$ implies that the null space of $\widetilde{\mathbf{A}}_\text{sym}^+$ contains the null-space of $\widetilde{\mathbf{A}}_\text{sym}$.
+This is important when the operator $\mathcal{A}$ has a large (or infinite) dimensional null space, or a spectrum that clusters at zero.
+Hessians in ill-posed distributed parameter inverse problems typically have both of these properties.
+The equation $f(\lambda_\text{max})=\lambda_\text{max}$ forces the moderate and large positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}^+$ to be close to the corresponding moderate and large positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$.
+I.e., $f$ modifies the ``important'' part of the spectrum of $\widetilde{\mathbf{A}}_\text{sym}$ as little as possible.
+The equation $f(\lambda_\text{min})=|\lambda_\text{min}|$ ensures that the negative eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $f$.
+There is a tradeoff: if we choose a larger value for $f(\lambda_\text{min})$, then the positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}^+$ will be closer to the corresponding positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ (good), but the erroneous negative eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ will be amplified more in magnitude (bad).
+We find that $f(\lambda_\text{min})=|\lambda_\text{min}|$ is an appropriate balance between these competing interests.
+
+Using the definition of $f$ in \eqref{eq:rational_function_scalar}, and basic calculus and linear algebra, we may write Equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, and \eqref{eq:desired_property_4} in the following matrix form:
+\begin{equation}
+\label{eq:rational_3_linear_system}
+\begin{bmatrix} 1 & 0 & 1/\mu \\ 0 & 1 & -1/\mu^2 \\ 1 & \lambda_\text{max} & 1/(\lambda_\text{max}+\mu) \end{bmatrix} \begin{bmatrix} c_0 \\ c_1 \\ c_2 \end{bmatrix} = \begin{bmatrix} 0 \\ 0 \\ \lambda_\text{max} \end{bmatrix},
+\end{equation}
+and we may write Equation \eqref{eq:desired_property_3} as follows: 
+\begin{equation}
+\label{eq:rational_extra_eq}
+c_0 + c_1 \lambda_\text{min} + \frac{c_2}{\lambda_\text{min}+\mu} - |\lambda_\text{min}| = 0.
+\end{equation}
+Let $g(\mu)$ denote the left hand side of \eqref{eq:rational_extra_eq}, but with $c_0$, $c_1$, and $c_2$ now solving \eqref{eq:rational_3_linear_system} for the given value of $\mu$.
+I.e., within $g(\mu)$ the quantities $c_0$, $c_1$, and $c_2$ are no longer free variables; they now depend on $\mu$ implicitly.
+We solve the combined system, \eqref{eq:rational_3_linear_system} and $\eqref{eq:rational_extra_eq}$, by solving the one-dimensional nonlinear root finding problem
+\begin{equation}
+\label{eq:g_mu_zero}
+g(\mu)=0
+\end{equation}
+with bound $\mu > |\lambda_\text{min}|$ and initial guess $\mu_0 = 2 |\lambda_\text{min}|$.
+Each evaluation of $g$ requires solving the $3 \times 3$ linear system \eqref{eq:rational_3_linear_system}, and once the solution $\mu$ is found, we solve \eqref{eq:rational_3_linear_system} one more time to get the final values of $c_0$, $c_1$, and $c_2$.
+%
+%Row 1 corresponds to the condition $f(M)=M$, row 2 corresponds to the condition $f(0)=0$, and row 3 corresponds to the condition $f'(0)=0$.
+To find the third row of $A$, we used the fact that the derivative of $f$ is given by $$f'(\lambda) = c_1 - \frac{c_2}{(\lambda+\mu)^2}.$$
+%
+%
+%We want $f(\lambda)$ to have the following desirable properties:
+%\begin{enumerate}
+%	\item $f(\lambda)$ is nonnegative on $[m, M]$.
+%	\item $f(\lambda) \approx \lambda$ when $\lambda$ is positive and not small.
+%	\item $f(0)=0$ is the minimum of $f(\lambda)$ on $[m, M]$.
+%	\item $f(\lambda) \le |\lambda|$ when $\lambda$ is negative.
+%\end{enumerate}
+%Property 1 ensures that $\widetilde{\mathbf{A}}_\text{sym}^+$ is positive semi-definite.
+Property 2 ensures that $f$ modifies the ``important'' part of the spectrum of $\widetilde{\mathbf{A}}_\text{sym}$ as little as possible.
+That is, the part of the spectrum of $\mathcal{A}$ that we expect to be well-approximated by the product-convolution approximation.
+Property 3 ensures that $\widetilde{\mathbf{A}}_\text{sym}^+$ preserves the null-space (and near null-space) of $\widetilde{\mathbf{A}}_\text{sym}$.
+This is important because in practice, we often want to approximate operators $\mathcal{A}$ that have a large (possibly infinite) number of zero eigenvalues, and/or a sequence of eigenvalues that clusters at zero.
+Hessians in ill-posed inverse problems typically have both of these properties.
+Property 4 ensures that $f$ does not amplify erroneous negative eigenvalues $\widetilde{\mathbf{A}}_\text{sym}$.
+If $\mathcal{A}$ is positive semi-definite, then eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ can only be negative because of errors in the product-convolution approximation.
 

### 2021-07-20 localpsf/product_convolution_hmatrix.py

@@ -164,7 +164,7 @@ def product_convolution_hmatrix(V_in, V_out,
     extras = {'vol' : vol,
               'mu' : mu,
               'Sigma' : Sigma,
-              'point_batches=' : point_batches,
+              'point_batches' : point_batches,
               'mu_batches' : mu_batches,
               'Sigma_batches' : Sigma_batches,
               'tau' : tau,

### 2021-07-20 numerical_examples/heat/heat_generate_results.py

@@ -2,6 +2,7 @@ import numpy as np
 import dolfin as dl
 import time
 import pathlib
+from tqdm.auto import tqdm
 
 from nalger_helper_functions import *
 import hlibpro_python_wrapper as hpro
@@ -12,7 +13,7 @@ from localpsf.product_convolution_hmatrix import product_convolution_hmatrix
 
 ########    OPTIONS    ########
 
-HIP_options = {'mesh_h' : 2e-2,
+HIP_options = {'mesh_h' : 1.5e-2,
                'finite_element_order' : 1,
                'final_time' : 3e-4,
                'num_timesteps' : 35,
@@ -29,7 +30,8 @@ krylov_tol = 1e-6
 a_reg_min = 1e-5
 a_reg_max = 1e0
 num_reg = 11
-all_batch_sizes = [1,5,10]
+all_batch_sizes = [1,3,6,9]
+n_random_error_matvecs = 50
 
 save_dir = get_project_root() / 'numerical_examples' / 'heat' / time.ctime()
 save_dir.mkdir(parents=True, exist_ok=True)
@@ -125,69 +127,84 @@ all_extras = list()
 for k in all_batch_sizes:
     Hd_hmatrix, extras = product_convolution_hmatrix(HIP.V, HIP.V, HIP.apply_Hd_petsc, HIP.apply_Hd_petsc, k,
                                                      hmatrix_tol=hmatrix_rtol, make_positive_definite=True,
-                                                     return_extras=True)
+                                                     return_extras=True, grid_density_multiplier=0.5)
     all_Hd_hmatrix.append(Hd_hmatrix)
     all_extras.append(extras)
 
-# Plot impulse response batches
 
-for k in range(len(all_batch_sizes)):
-    extras = all_extras[k]
-    ff_batches = extras['ff_batches']
-    mu_batches = extras['mu_batches']
-    Sigma_batches = extras['Sigma_batches']
-    point_batches = extras['point_batches']
-    tau = extras['tau']
-    for b in range(len(ff_batches)):
-        visualize_impulse_response_batch(ff_batches[k], point_batches[k], mu_batches[k], Sigma_batches[k], tau)
+# Columns error plots
+
[…]

### 2021-07-21 paper/localpsf.tex
edits to rational positive definite section

@@
 \label{eq:kernel_representation}
-(\mathcal{A}u)(v) := \int_\Omega \int_\Omega v(y) A(y,x) u(x) dx dy,
-\end{equation}
-with integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$.
+(\mathcal{A}u)(v) := \int_\Omega \int_\Omega v(y) \Phi(y,x) u(x) dx dy,
+\end{equation}
+with integral kernel $\Phi:\Omega \times \Omega \rightarrow \mathbb{R}$.
 Here $\mathcal{A}u \in L^2(\Omega)'$ is the linear functional that results from applying $\mathcal{A}$ to $u\in L^2(\Omega)$, and $\left(\mathcal{A}u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
@@
 \begin{description}
-\item[Matrix-free:] One cannot easily access kernel entries $A(y,x)$.
+\item[Matrix-free:] One cannot easily access kernel entries $\Phi(y,x)$.
 Access to $\mathcal{A}$ is available only through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary functions (i.e., operator actions).
@@
 \begin{equation*}
-	A(y,x) \ge 0.
+	\Phi(y,x) \ge 0.
 \end{equation*}
@@
 \label{eq:impulse_response_defn}
-\phi_x(y) := A(y, x).
+\phi_x(y) := \Phi(y, x).
 \end{equation} 
@@
 	\label{eq:local_translation_invariance}
-	A(y+h, x+h) \approx A(y,x).
+	\Phi(y+h, x+h) \approx \Phi(y,x).
 \end{equation}
@@
 \label{eq:u_defn_in_thm}
-\gamma_y(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right),
+\gamma_y(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{\Phi}\left(y+x,x\right),
 \end{equation}
@@
 
-We now describe how to efficiently evaluate $\widetilde{A}(y,x)$ at an arbitrary point $(x,y) \in \Omega \times \Omega$. From \eqref{eq:kernel_entries}, we have
+We now describe how to efficiently evaluate $\widetilde{\Phi}(y,x)$ at an arbitrary point $(x,y) \in \Omega \times \Omega$. From \eqref{eq:kernel_entries}, we have
 \begin{equation}
 \label{eq:eval_y_x}
-\widetilde{A}(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x),
+\widetilde{\Phi}(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x),
 \end{equation}
@@
 where $b(i)$ denotes the batch index such that $x_i \in S^{b(i)}$.
-To compute $z=\widetilde{A}(y,x)$, we start with $z=0$, then loop through all batches $b$, and all sample points in each batch.
+To compute $z=\widetilde{\Phi}(y,x)$, we start with $z=0$, then loop through all batches $b$, and all sample points in each batch.
 For each sample point, we add one term in the sum in \eqref{eq:eval_y_x} to $z$, using \eqref{eq:kernel_entry_formua} as needed to evaluate $\varphi_{x_i}$.
@@
 $$\widetilde{\mathbf{A}}_\text{sym} := \frac{1}{2}\left(\widetilde{\mathbf{A}} + \widetilde{\mathbf{A}}^T\right).$$
-Then we modify $\widetilde{\mathbf{A}}_\text{sym}$ to make it positive definite by forming a rational matrix function of the following form: 
+Then we modify $\widetilde{\mathbf{A}}_\text{sym}$ to make it positive semi-definite by forming a rational matrix function of the following form: 
 \begin{equation}
 \label{eq:rational_matrix_function}
-\widetilde{\mathbf{A}}_\text{sym}^+ := c_0 + c_1 \widetilde{\mathbf{A}}_\text{sym} + c_2 \left(\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}\right)^{-1},
+\widetilde{\mathbf{A}}_\text{sym}^+ := c_0 \mathbf{I} + c_1 \widetilde{\mathbf{A}}_\text{sym} + c_2 \left(\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}\right)^{-1},
 \end{equation}
 where $\mathbf{I}$ is the identity matrix which has the same shape as $\widetilde{\mathbf{A}}_\text{sym}$.
-The scalars $c_0$, $c_1$, $c_2$, and $\mu$ are chosen so that the spectrum of $\widetilde{\mathbf{A}}_\text{sym}^+$ has certain desirable properties, which we will describe later in this section.
+The scalars $c_0$, $c_1$, $c_2$, and $\mu$ are chosen so that the spectrum of $\widetilde{\mathbf{A}}_\text{sym}^+$ has certain desirable properties.
+Later in this section we will describe how $c_0$, $c_1$, $c_2$, and $\mu$ are chosen and how our choice of these scalars affects the spectral properties of $\widetilde{\mathbf{A}}_\text{sym}^+$.
 Forming $\widetilde{\mathbf{A}}_\text{sym}^+$ requires computing an H-matrix inverse, $\left(\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}\right)^{-1}$.
-While explicit computation of matrix inverses should typically be avoided, here it is fine to compute the inverse because $\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}$ is an H-matrix, and there are fast, stable, scalable, and widely used algorithms for inverting H-matrices \cite{Hmatrixinverse}.
+While explicit computation of matrix inverses should usually be avoided, here it is fine to compute the inverse because $\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}$ is an H-matrix.
+There exist fast, stable, scalable, and widely used algorithms for inverting H-matrices \cite{Hmatrixinverse}.
 
@@
 There is a tradeoff: if we choose a larger value for $f(\lambda_\text{min})$, then the positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}^+$ will be closer to the corresponding positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ (good), but the erroneous negative eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ will be amplified more in magnitude (bad).
+This is shown in Figure FIG.
 We find that $f(\lambda_\text{min})=|\lambda_\text{min}|$ is an appropriate balance between these competing interests.

### 2021-07-21 paper/localpsf.tex
edits to rational positive definite section based on feedback from J

@@
 This is undesirable.
-Symmetry and positive semi-definiteness are important properties which should be preserved by the approximation, if possible.
-Also, a lack of symmetry and/or positive semi-definiteness in $\widetilde{\mathbf{A}}$ may prevent one from using popular and highly effective algorithms, such as the conjugate gradient method, to perform further useful operations involving $\widetilde{\mathbf{A}}$.
+Symmetry and positive semi-definiteness are important properties which should be preserved if possible.
+Also, lacking these properties may prevent one from using popular and highly effective algorithms, such as the conjugate gradient method, to perform further useful operations involving $\widetilde{\mathbf{A}}$.
 In this section we describe a method for modifying $\widetilde{\mathbf{A}}$ to make it symmetric positive semi-definite.
@@
 where $\mathbf{I}$ is the identity matrix which has the same shape as $\widetilde{\mathbf{A}}_\text{sym}$.
-The scalars $c_0$, $c_1$, $c_2$, and $\mu$ are chosen so that the spectrum of $\widetilde{\mathbf{A}}_\text{sym}^+$ has certain desirable properties.
-Later in this section we will describe how $c_0$, $c_1$, $c_2$, and $\mu$ are chosen and how our choice of these scalars affects the spectral properties of $\widetilde{\mathbf{A}}_\text{sym}^+$.
+The scalars $c_0$, $c_1$, $c_2$, and $\mu$ are chosen so that the moderate and large positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}^+$ approximately equal the corresponding moderate and large positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$, while the negative and small positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ are modified so that $\widetilde{\mathbf{A}}_\text{sym}^+$ is positive semi-definite.
+Later in this section we will explain how $c_0$, $c_1$, $c_2$, and $\mu$ are chosen and how our choice of these scalars affects the spectral properties of $\widetilde{\mathbf{A}}_\text{sym}^+$.
 Forming $\widetilde{\mathbf{A}}_\text{sym}^+$ requires computing an H-matrix inverse, $\left(\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}\right)^{-1}$.
@@
 
-The rational matrix approximation \eqref{eq:rational_matrix_function} may be written as $\widetilde{\mathbf{A}}_\text{sym}^+ = f(\widetilde{\mathbf{A}}_\text{sym})$, where $f$ is the rational function
+Approximation \eqref{eq:rational_matrix_function} may be written as $\widetilde{\mathbf{A}}_\text{sym}^+ = f(\widetilde{\mathbf{A}}_\text{sym})$, where
 \begin{equation}
@@
 \end{equation}
-We choose the four unknown constants, $c_0, c_1, c_2$, and $\mu$, to be the solution to the following system of four equations:
+We choose $c_0, c_1, c_2$, and $\mu$, to be the solution to the following system of equations:
 \begin{align}
@@
 \label{eq:rational_extra_eq}
-c_0 + c_1 \lambda_\text{min} + \frac{c_2}{\lambda_\text{min}+\mu} - |\lambda_\text{min}| = 0.
-\end{equation}
-Let $g(\mu)$ denote the left hand side of \eqref{eq:rational_extra_eq}, but with $c_0$, $c_1$, and $c_2$ now solving \eqref{eq:rational_3_linear_system} for the given value of $\mu$.
-I.e., within $g(\mu)$ the quantities $c_0$, $c_1$, and $c_2$ are no longer free variables; they now depend on $\mu$ implicitly.
+\underbrace{c_0 + c_1 \lambda_\text{min} + \frac{c_2}{\lambda_\text{min}+\mu} - |\lambda_\text{min}|}_{g(\mu)} = 0.
+\end{equation}
+Let $g(\mu)$ denote the left hand side of \eqref{eq:rational_extra_eq}, but with $c_0$, $c_1$, and $c_2$ being the solution of \eqref{eq:rational_3_linear_system} for the given value of $\mu$.
+I.e., within $g(\mu)$ the quantities $c_0$, $c_1$, and $c_2$ are no longer free variables; they now depend on $\mu$ implicitly through the solution of a linear system.
 We solve the combined system, \eqref{eq:rational_3_linear_system} and $\eqref{eq:rational_extra_eq}$, by solving the one-dimensional nonlinear root finding problem

### 2021-07-21 paper/localpsf.tex
edits to rational positive definite section based on feedback from J

@@
 The equation $f(\lambda_\text{min})=|\lambda_\text{min}|$ ensures that the negative eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $f$.
-There is a tradeoff: if we choose a larger value for $f(\lambda_\text{min})$, then the positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}^+$ will be closer to the corresponding positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ (good), but the erroneous negative eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ will be amplified more in magnitude (bad).
+There is a tradeoff: if we choose a larger value for $f(\lambda_\text{min})$, then the positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}^+$ will be closer to the corresponding positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ (good), but the erroneous negative eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ will be amplified in magnitude (bad).
 This is shown in Figure FIG.

### 2021-07-21 paper/localpsf.tex
modify notation. Atilde -> Apc. Akernel = Phi

@@
 Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes.
-We therefore approximate $\mathcal{A}$ by a spatially varying weighted sum of convolution operators, $\widetilde{\mathcal{A}} \approx \mathcal{A}$, of the form
+We therefore approximate $\mathcal{A}$ by a spatially varying weighted sum of convolution operators, $\mathcal{A}_\text{pc} \approx \mathcal{A}$, of the form
 \begin{equation}
 \label{eq:product_convolution}
-	\left(\widetilde{\mathcal{A}}u\right)(v) := \left( v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
+	\left(\mathcal{A}_\text{pc}~u\right)(v) := \left( v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
 \end{equation}
 where $f \cdot g$ denotes pointwise multiplication, $f \ast g$ denotes convolution, and $\left(f, g\right)_{L^2(\Omega)}$ denotes the $L^2$ inner product on $\Omega$, for functions $f,g \in L^2(\Omega)$.
+Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution.
 To avoid nested subscripts, we write
@@
 
-Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution.
 The more locally translation invariant an operator is (i.e., the smaller the discrepancy between the left hand side and right hand side in \eqref{eq:local_translation_invariance}), the smaller the number of terms that are required in \eqref{eq:product_convolution} to achieve an accurate product-convolution approximation.
@@
 
-Once the $\varphi_i$ and $w_i$ are computed, we convert $\widetilde{\mathcal{A}}$ to hierarchical matrix (H-matrix) format.
-Let $\widetilde{A}$ denote the integral kernel associated with $\widetilde{\mathcal{A}}$.
-While $A(y,x)$ is not easily computable, $\widetilde{A}(y,x)$ is given by the formula
+Once the $\varphi_i$ and $w_i$ are computed, we convert $\mathcal{A}_\text{pc}$ to hierarchical matrix (H-matrix) format.
+Let $\Phi_\text{pc}$ denote the integral kernel associated with $\mathcal{A}_\text{pc}$.
+While $\Phi(y,x)$ is not easily computable, $\Phi_\text{pc}(y,x)$ is given by the formula
 \begin{equation}
 \label{eq:kernel_entries}
-	\widetilde{A}(y,x) = \sum_{i=1}^r \varphi_i(y-x) w_i(x).
+	\Phi_\text{pc}(y,x) = \sum_{i=1}^r \varphi_i(y-x) w_i(x).
 \end{equation}
 This follows from~\eqref{eq:product_convolution} and the fact that the convolution operator $u \mapsto \varphi_i \ast u$ has $(y,x)$ kernel entry given by $\varphi(y-x)$.
-Formula~\eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
-Once in H-matrix format, fast H-matrix arithmetic is used to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations.
+Formula~\eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\mathcal{A}_\text{pc}$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
+Once in H-matrix format, fast H-matrix arithmetic is used to invert $\mathcal{A}_\text{pc}$, or perform other useful matrix operations.
+
+Often we are interested in approximating operators $\mathcal{A}$ that are symmetric positive semi-definite.
+Unfortunately, $\mathcal{A}_\text{pc}$ is, in general, non-symmetric and indefinite due to errors in the product convolution approximation.
+In this case, we modify the H-matrix representation of $\mathcal{A}_\text{pc}$ to make it positive definite, by symmetrizing it, then applying a specially chosen rational matrix function to the symmetrized H-matrix.
 
@@
 
-\section{Constructing the approximation, $\widetilde{\mathcal{A}}$}
+\section{Constructing the product-convolution H-matrix approximation}
 \label{sec:method}
-This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\widetilde{\mathcal{A}} \approx \mathcal{A}$, given in \eqref{eq:product_convolution}.
-We also describe how we evaluate kernel entries of the approximation, $\widetilde{A}(y,x)$, and how we convert $\widetilde{\mathcal{A}}$ to H-matrix format.
+This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\mathcal{A}_\text{pc} \approx \mathcal{A}$, given in \eqref{eq:product_convolution}.
+We also describe how we evaluate kernel entries of the approximation, $\Phi_\text{pc}(y,x)$, how we convert $\mathcal{A}_\text{pc}$ to H-matrix format, and (optionally) how we modify $\mathcal{A}_\text{pc}$ to force it to be symmetric positive semi-definite.
 

### 2021-07-21 paper/localpsf.tex
edited rational positive definite section based on feedback from Noemi

@@
 We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}). 
-To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as distinct functions.
+%To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as distinct functions. 
 A greedy algorithm is used to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}).
@@
 %\end{itemize}
-The complete algorithm for constructing $\widetilde{\mathcal{A}}$ is shown in Algorithm \ref{alg:construct_Atilde}.
-The method for evaluating kernel entries of $\widetilde{A}$ is shown in Section \ref{sec:eval_matrix_entries}.
-The method for converting $\widetilde{\mathcal{A}}$ to H-matrix format is shown in Section \ref{sec:H_matrix_conversion}.
+The complete algorithm for constructing $\mathcal{A}_\text{pc}$ is shown in Algorithm \ref{alg:construct_Atilde}.
+The method for evaluating kernel entries of $\Phi_\text{pc}$ is shown in Section \ref{sec:eval_matrix_entries}.
+The method for converting $\mathcal{A}_\text{pc}$ to H-matrix format is shown in Section \ref{sec:H_matrix_conversion}.
+The method for modifying the resulting H-matrix to make it symmetric positive semi-definite is shown in Section \ref{sec:make_spd}.
 
@@
 \subsection{Rational positive semi-definite modification}
+\label{sec:make_spd}
 
@@
 Symmetry and positive semi-definiteness are important properties which should be preserved if possible.
-Also, lacking these properties may prevent one from using popular and highly effective algorithms, such as the conjugate gradient method, to perform further useful operations involving $\widetilde{\mathbf{A}}$.
+Also, lacking these properties may prevent one from using highly effective algorithms, such as the conjugate gradient method, to perform further useful operations involving $\widetilde{\mathbf{A}}$.
 In this section we describe a method for modifying $\widetilde{\mathbf{A}}$ to make it symmetric positive semi-definite.
@@
 Later in this section we will explain how $c_0$, $c_1$, $c_2$, and $\mu$ are chosen and how our choice of these scalars affects the spectral properties of $\widetilde{\mathbf{A}}_\text{sym}^+$.
+In Figure FIG we compare the spectrum of $\widetilde{\mathbf{A}}_\text{sym}$ to that of $\widetilde{\mathbf{A}}_\text{sym}^+$ for different choices of these scalars.
 Forming $\widetilde{\mathbf{A}}_\text{sym}^+$ requires computing an H-matrix inverse, $\left(\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}\right)^{-1}$.
-While explicit computation of matrix inverses should usually be avoided, here it is fine to compute the inverse because $\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}$ is an H-matrix.
-There exist fast, stable, scalable, and widely used algorithms for inverting H-matrices \cite{Hmatrixinverse}.
+While explicit computation of matrix inverses should usually be avoided, here it is computationally acceptable because $\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}$ is an H-matrix.
+Inversion of H-matrices is a routine procedure which only requires $O(N k^2 \log(N)^2)$ work, where $k$ is the H-matrix rank \cite{Hmatrixinverse}.
 
@@
 We choose $c_0, c_1, c_2$, and $\mu$, to be the solution to the following system of equations:
+\begin{subequations}
+	\label{eq:four_eq_system}
 \begin{align}
@@
 \end{align}
+\end{subequations}
 where $\lambda_\text{min}$ and $\lambda_\text{max}$ are the smallest and largest eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$, respectively.
@@
 Note that $\lambda_\text{min} < 0$ since $\widetilde{\mathbf{A}}_\text{sym}$ is indefinite.
-We estimate $\lambda_\text{min}$ and $\lambda_\text{max}$ using the implicitly restarted Lanczos method \cite{lanczos,scipy}, a matrix-free Krylov method.
-This is fast because computing matrix-vector products with an H-matrix is fast, and because the eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ do not cluster near $\lambda_\text{min}$ or $\lambda_\text{max}$ (though the eigenvalues may cluster at zero).
+We estimate $\lambda_\text{min}$ and $\lambda_\text{max}$ using the implicitly restarted Lanczos method \cite{lanczos,scipy}. 
+%This is fast because computing matrix-vector products with an H-matrix is fast, and because the eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ do not cluster near $\lambda_\text{min}$ or $\lambda_\text{max}$ (though the eigenvalues may cluster at zero).
 
 Together, equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, \eqref{eq:desired_property_3},  and \eqref{eq:desired_property_4}, and the condition $\mu > |\lambda_\text{min}|$, imply that $f(0)=0$ is the unique minimum of $f$ on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that $f$ is non-negative on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that  $f(\widetilde{\mathbf{A}}_\text{sym}) = \widetilde{\mathbf{A}}_\text{sym}^+$ is positive semi-definite.
-The equation $f(0)=0$ implies that the null space of $\widetilde{\mathbf{A}}_\text{sym}^+$ contains the null-space of $\widetilde{\mathbf{A}}_\text{sym}$.
+Furthermore, \eqref{eq:desired_property_1} implies that the null space of $\widetilde{\mathbf{A}}_\text{sym}^+$ contains the null-space of $\widetilde{\mathbf{A}}_\text{sym}$.
 This is important when the operator $\mathcal{A}$ has a large (or infinite) dimensional null space, or a spectrum that clusters at zero.
 Hessians in ill-posed distributed parameter inverse problems typically have both of these properties.
-The equation $f(\lambda_\text{max})=\lambda_\text{max}$ forces the moderate and large positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}^+$ to be close to the corresponding moderate and large positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$.
+Equation \eqref{eq:desired_property_4} forces the moderate and large positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}^+$ to be close to the corresponding moderate and large positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$.
 I.e., $f$ modifies the ``important'' part of the spectrum of $\widetilde{\mathbf{A}}_\text{sym}$ as little as possible.
-The equation $f(\lambda_\text{min})=|\lambda_\text{min}|$ ensures that the negative eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $f$.
+Equation \eqref{eq:desired_property_3} ensures that the negative eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $f$.
 There is a tradeoff: if we choose a larger value for $f(\lambda_\text{min})$, then the positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}^+$ will be closer to the corresponding positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ (good), but the erroneous negative eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ will be amplified in magnitude (bad).
@@
 
+Solving system \eqref{eq:four_eq_system} is computationally trivial.
 Using the definition of $f$ in \eqref{eq:rational_function_scalar}, and basic calculus and linear algebra, we may write Equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, and \eqref{eq:desired_property_4} in the following matrix form:
@@
 \label{eq:rational_extra_eq}
-\underbrace{c_0 + c_1 \lambda_\text{min} + \frac{c_2}{\lambda_\text{min}+\mu} - |\lambda_\text{min}|}_{g(\mu)} = 0.
-\end{equation}
-Let $g(\mu)$ denote the left hand side of \eqref{eq:rational_extra_eq}, but with $c_0$, $c_1$, and $c_2$ being the solution of \eqref{eq:rational_3_linear_system} for the given value of $\mu$.
+c_0 + c_1 \lambda_\text{min} + \frac{c_2}{\lambda_\text{min}+\mu} - |\lambda_\text{min}| = 0.
+\end{equation}
+Let 
+\begin{equation*}
+g(\mu) := c_0(\mu) + c_1(\mu) \lambda_\text{min} + \frac{c_2(\mu)}{\lambda_\text{min}+\mu} - |\lambda_\text{min}|
+\end{equation*}
+denote the left hand side of \eqref{eq:rational_extra_eq}, but with $c_0(\mu)$, $c_1(\mu)$, and $c_2(\mu)$ being the solution of \eqref{eq:rational_3_linear_system} for the given value of $\mu$.
 I.e., within $g(\mu)$ the quantities $c_0$, $c_1$, and $c_2$ are no longer free variables; they now depend on $\mu$ implicitly through the solution of a linear system.

### 2021-07-23 paper/localpsf.tex
H-matrix and discretization section

@@
 \section{Introduction}
+\label{sec:intro}
 
@@
 \label{eq:kernel_representation}
-(\mathcal{A}u)(v) := \int_\Omega \int_\Omega v(y) \Phi(y,x) u(x) dx dy,
-\end{equation}
-with integral kernel $\Phi:\Omega \times \Omega \rightarrow \mathbb{R}$.
+(\mathcal{A}u)(v) := \int_\Omega \int_\Omega v(y) \Aker(y,x) u(x) dx dy,
+\end{equation}
+with integral kernel $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$.
 Here $\mathcal{A}u \in L^2(\Omega)'$ is the linear functional that results from applying $\mathcal{A}$ to $u\in L^2(\Omega)$, and $\left(\mathcal{A}u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
@@
 \label{eq:u_defn_in_thm}
-\gamma_y(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{\Phi}\left(y+x,x\right),
+\gamma_y(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \AkerPc\left(y+x,x\right),
 \end{equation}
@@
 \begin{equation}
-\label{eq:eval_y_x}
-\widetilde{\Phi}(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x),
+\label{eq:eval_y_x_000}
+\AkerPc(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x),
 \end{equation}
@@
 where $b(i)$ denotes the batch index such that $x_i \in S^{b(i)}$.
-To compute $z=\widetilde{\Phi}(y,x)$, we start with $z=0$, then loop through all batches $b$, and all sample points in each batch.
+To compute $z=\AkerPc(y,x)$, we start with $z=0$, then loop through all batches $b$, and all sample points in each batch.
 For each sample point, we add one term in the sum in \eqref{eq:eval_y_x} to $z$, using \eqref{eq:kernel_entry_formua} as needed to evaluate $\varphi_{x_i}$.
@@
 
-\subsection{Conversion to hierarchical matrix format}
+\subsection{Hierarchical matrix construction}
 \label{sec:H_matrix_conversion}
 
-We convert $\widetilde{\mathbf{A}}$ (the discretized version of $\widetilde{\mathcal{A}}$) to hierarchical matrix (H-matrix) format.
+In this section we show how to construct an H-matrix representation of our product-convolution approximation, after discretizing the product-convolution approximation with finite elements.
+Let $f_1, f_2, \dots, f_N$ be a set of finite element basis functions used to discretize the problem, and let $V_h := \Span\left(f_1, f_2, \dots, f_N\right)$ be the corresponding finite element space.
+Here $h$ is the mesh size parameter for the finite element mesh used to discretize the problem.
+Further, let $p_i \in \mathbb{R}^d$, $i=1,\dots, N$ be the Lagrange nodes associated with the functions $f_i$.
+That is, the points such that $f_i(p_j)$ equals one if $i=j$, and zero otherwise.
+
+In Section \ref{sec:intro} we noted that the original integral kernel, $\Aker$, is inaccessible, while in Equation \eqref{eq:kernel_entries}  we saw that the integral kernel for our product-convolution approximation, $\AkerPc \approx \Aker$, is accessible, and may be cheaply evaluated at arbitrary pairs of points $(y,x) \in \Omega \times \Omega$ via the following formula:
+\begin{equation}
+\label{eq:eval_y_x}
+\AkerPc(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x).
+\end{equation}
+We now define a further approximation, $\AkerPcMesh \approx \AkerPc$,
+\begin{equation}
+\label{eq:defn_of_Akerpcmesh}
+\AkerPcMesh(y,x) := \sum_{i=1}^N \sum_{j=1}^N \AkerPc(p_i,p_j) f_i(y) f_j(x)
+\end{equation}
+which is the interpolation of $\AkerPc$ onto $V_h \otimes V_h$.
+Error in the combined approximation, $\AkerPcMesh \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$.
+
+Replacing $\Aker$ with $\AkerPcMesh$ within the definition of $\Aop$ in \eqref{eq:kernel_representation} yields the following operator approximation:
+\begin{equation*}
+\left(\AopPcMesh u \right)(v) := \int_\Omega \int_\Omega v(y) \AkerPcMesh(y,x) u(x) dx dy.
+\end{equation*}
+The operator $\AopPcMesh$ maps $L^2(\Omega) \rightarrow L^2(\Omega)'$, but we are interested in the restriction of this operator to functions in the finite element space (i.e., $\AopPcMesh : V_h \rightarrow V_h'$).
+%Error in the approximation $\Aop \approx \AopPcMesh$ is a combination of error in the product-convolution approximation, and error due to finite element discretization.
+Let $u_h \in V_h$ and $v_h \in V_h$, and let $\mathbf{u}$ and $\mathbf{v}$ be the coefficient vectors for $u_h$ and $v_h$, respectively, with respect to the finite element basis $f_1, \dots, f_N$.
+I.e., $u_h(x) = \sum_{i=1}^N \mathbf{u}_i f_i(x)$, and similar for $v_h$.
+By using the definition of $\AkerPcMesh$ in \eqref{eq:defn_of_Akerpcmesh} and performing algebraic manipulations, it is straightforward but tedious to show that
+\begin{equation*}
+\left(\AopPcMesh u_h \right)(v_h) = \mathbf{v}^T \mathbf{M} \AkerPcMat \mathbf{M} \mathbf{u},
+\end{equation*}
+%\begin{align*}
+%\left(\AopPcMesh u_h \right)(v_h) =& \int_\Omega \int_\Omega v_h(y) \AkerPcMesh(y,x) u_h(x) dx dy \\
+%=& \int_\Omega \int_\Omega \left(\sum_{i=1}^N \mathbf{v}_i f_i(y)\right) \left(\sum_{j=1}^N \sum_{k=1}^N \AkerPc(p_j,p_k) f_j(y) f_k(x)\right) \left(\sum_{l=1}^N \mathbf{u}_l f_l(x)\right) dx dy \\
+%=& \sum_{i,j,k,l=1}^N \mathbf{v}_i \left(\int_\Omega f_i(y) f_j(y) dy\right) \AkerPcMesh(p_j, p_k) \left(\int_\Omega f_k(x) f_l(x) dx\right) \mathbf{u}_l \\
+%=& \sum_{i,j,k,l=1}^N \mathbf{v}_i \mathbf{M}_{ij} \left(\AkerPcMat\right)_{jk} \mathbf{M}_{kl} \mathbf{u}_l \\
+%=& \mathbf{v}^T \mathbf{M} \AkerPcMat \mathbf{M} \mathbf{u},
+%\end{align*}
+where $\mathbf{M} \in \mathbb{R}^{N \times N}$ is the finite element mass matrix, which has entries
+\begin{equation*}
+\mathbf{M}_{ij} := \int_\Omega f_i(x) f_j(x) dx
+\end{equation*}
+and $\AkerPcMat \in \mathbb{R}^{N \times N}$ is the evaluation of $\AkerPcMesh$ at all pairs of finite element Lagrange nodes, which has entries 
+\begin{equation*}
+\left(\AkerPcMat\right)_{ij} := \AkerPcMesh(p_j, p_k).
+\end{equation*}
+The finite element matrix representation of $\AopPcMesh$, which we denote by $\AmatPc$, is therefore given by
+\begin{equation}
+\label{eq:Amatpc_defn}
+\AmatPc := \mathbf{M} \AkerPcMat \mathbf{M}.
+\end{equation}
+We convert $\mathbf{M}$ and $\AkerPcMat$ to H-matrix format, then multiply these matrices in \eqref{eq:Amatpc_defn} using fast H-matrix methods to form an H-matrix representation of $\AmatPc$.
 Once in H-matrix format, useful matrix operations such as matrix factorization and inversion may be performed quickly using fast H-matrix methods.
 We use H1 matrices in our numerical results, but any of the other H-matrix formats (such as H2, HODLR, HSS, HBS, and others \cite{HMATRIX}) could be used instead.
+
 The basic ideas behind H-matrices are as follows:
@@
 
-We convert $\widetilde{\mathbf{A}}$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{HLIBPRO}.
-Briefly, this method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix to construct a recursive block partitioning of $\widetilde{\mathbf{A}}$, such that all blocks at the finest level of the partitioning are either low rank, or are small.
+The mass matrix, $\mathbf{M}$, is sparse, and may therefore be converted to H-matrix format using standard H-matrix methods for sparse matrices.
+We convert $\AkerPcMat$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{HLIBPRO}.
+Briefly, this method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix (here, these locations are finite element the Lagrange nodes $p_i$, $i=1,\dots,N$) to construct a recursive block partitioning of $\AmatPc$, such that all blocks at the finest level of the partitioning are either low rank, or are small.
 Then, the method forms low rank approximations of the blocks that are expected to be low rank by sampling specially chosen rows and columns of those blocks.
 The small high rank blocks are formed directly by evaluating their entries.
-The primary cost of constructing the approximation is the evaluation of $O(N \log N)$ entries of the kernel $\widetilde{A}$, which we perform using Algorithm \ref{alg:eval_y_x}.
 We describe the H-matrix construction in detail in Appendix \ref{app:h_matrix}.
+Although $\AkerPcMat$ is dense, constructing the H-matrix approximation via this process only requires evaluation of $O(N \log N)$ entries of $\AkerPc$, and these entries are computed via the formula in Equation \eqref{eq:eval_y_x}.
 
@@
 
-In many problems of practical interest (e.g., Hessian approximation) $\mathcal{A}$ is symmetric positive semi-definite.
-However, $\widetilde{\mathbf{A}}$ is generally non-symmetric and indefinite because of errors in the product-convolution approximation.
+In many problems of practical interest (e.g., Hessian approximation) $\Amat$ is symmetric positive semi-definite.
+However, $\AmatPc$ is generally non-symmetric and indefinite because of errors in the product-convolution approximation.
 This is undesirable.
 Symmetry and positive semi-definiteness are important properties which should be preserved if possible.
-Also, lacking these properties may prevent one from using highly effective algorithms, such as the conjugate gradient method, to perform further useful operations involving $\widetilde{\mathbf{A}}$.
-In this section we describe a method for modifying $\widetilde{\mathbf{A}}$ to make it symmetric positive semi-definite.
-
-First, we use fast H-matrix addition to symmetrize $\widetilde{\mathbf{A}}$:
-$$\widetilde{\mathbf{A}}_\text{sym} := \frac{1}{2}\left(\widetilde{\mathbf{A}} + \widetilde{\mathbf{A}}^T\right).$$
-Then we modify $\widetilde{\mathbf{A}}_\text{sym}$ to make it positive semi-definite by forming a rational matrix function of the following form: 
+Also, lacking these properties may prevent one from using highly effective algorithms, such as the conjugate gradient method, to perform further useful operations involving $\AmatPc$.
+In this section we describe a method for modifying $\AmatPc$ to make it symmetric positive semi-definite.
+
+First, we use fast H-matrix addition to symmetrize $\AmatPc$:
+\begin{equation*}
+\AmatPcSym := \frac{1}{2}\left(\AmatPc + \AmatPc^T\right).
+\end{equation*}
+Then we modify $\AmatPcSym$ to make it positive semi-definite by forming a rational matrix function of the following form: 
 \begin{equation}
 \label{eq:rational_matrix_function}
-\widetilde{\mathbf{A}}_\text{sym}^+ := c_0 \mathbf{I} + c_1 \widetilde{\mathbf{A}}_\text{sym} + c_2 \left(\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}\right)^{-1},
-\end{equation}
-where $\mathbf{I}$ is the identity matrix which has the same shape as $\widetilde{\mathbf{A}}_\text{sym}$.
-The scalars $c_0$, $c_1$, $c_2$, and $\mu$ are chosen so that the moderate and large positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}^+$ approximately equal the corresponding moderate and large positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$, while the negative and small positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ are modified so that $\widetilde{\mathbf{A}}_\text{sym}^+$ is positive semi-definite.
-Later in this section we will explain how $c_0$, $c_1$, $c_2$, and $\mu$ are chosen and how our choice of these scalars affects the spectral properties of $\widetilde{\mathbf{A}}_\text{sym}^+$.
-In Figure FIG we compare the spectrum of $\widetilde{\mathbf{A}}_\text{sym}$ to that of $\widetilde{\mathbf{A}}_\text{sym}^+$ for different choices of these scalars.
-Forming $\widetilde{\mathbf{A}}_\text{sym}^+$ requires computing an H-matrix inverse, $\left(\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}\right)^{-1}$.
-While explicit computation of matrix inverses should usually be avoided, here it is computationally acceptable because $\widetilde{\mathbf{A}}_\text{sym} + \mu \mathbf{I}$ is an H-matrix.
+\AmatPcSymPlus := c_0 \mathbf{I} + c_1 \AmatPcSym + c_2 \left(\AmatPcSym + \mu \mathbf{I}\right)^{-1},
+\end{equation}
+where $\mathbf{I}$ is the identity matrix which has the same shape as $\AmatPcSym$.
+The scalars $c_0$, $c_1$, $c_2$, and $\mu$ are chosen so that the moderate and large positive eigenvalues of $\AmatPcSymPlus$ approximately equal the corresponding moderate and large positive eigenvalues of $\AmatPcSym$, while the negative and small positive eigenvalues of $\AmatPcSym$ are modified so that $\AmatPcSymPlus$ is positive semi-definite.
+Later in this section we will explain how $c_0$, $c_1$, $c_2$, and $\mu$ are chosen and how our choice of these scalars affects the spectral properties of $\AmatPcSymPlus$.
+In Figure FIG we compare the spectrum of $\AmatPcSym$ to that of $\AmatPcSymPlus$ for different choices of these scalars.
+Forming $\AmatPcSymPlus$ requires computing an H-matrix inverse, $\left(\AmatPcSym + \mu \mathbf{I}\right)^{-1}$.
+While explicit computation of matrix inverses should usually be avoided, here it is computationally acceptable because $\AmatPcSym + \mu \mathbf{I}$ is an H-matrix.
 Inversion of H-matrices is a routine procedure which only requires $O(N k^2 \log(N)^2)$ work, where $k$ is the H-matrix rank \cite{Hmatrixinverse}.
 
-Approximation \eqref{eq:rational_matrix_function} may be written as $\widetilde{\mathbf{A}}_\text{sym}^+ = f(\widetilde{\mathbf{A}}_\text{sym})$, where
+Approximation \eqref{eq:rational_matrix_function} may be written as $\AmatPcSymPlus = f(\AmatPcSym)$, where
 \begin{equation}
@@
 \end{subequations}
-where $\lambda_\text{min}$ and $\lambda_\text{max}$ are the smallest and largest eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$, respectively.
+where $\lambda_\text{min}$ and $\lambda_\text{max}$ are the smallest and largest eigenvalues of $\AmatPcSym$, respectively.
 We also require $\mu > |\lambda_\text{min}|$, so that the pole of $f$ is not in $[\lambda_\text{min},\lambda_\text{max}]$.
-Note that $\lambda_\text{min} < 0$ since $\widetilde{\mathbf{A}}_\text{sym}$ is indefinite.
+Note that $\lambda_\text{min} < 0$ since $\AmatPcSym$ is indefinite.
 We estimate $\lambda_\text{min}$ and $\lambda_\text{max}$ using the implicitly restarted Lanczos method \cite{lanczos,scipy}.
-%This is fast because computing matrix-vector products with an H-matrix is fast, and because the eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ do not cluster near $\lambda_\text{min}$ or $\lambda_\text{max}$ (though the eigenvalues may cluster at zero).
-
-Together, equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, \eqref{eq:desired_property_3},  and \eqref{eq:desired_property_4}, and the condition $\mu > |\lambda_\text{min}|$, imply that $f(0)=0$ is the unique minimum of $f$ on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that $f$ is non-negative on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that  $f(\widetilde{\mathbf{A}}_\text{sym}) = \widetilde{\mathbf{A}}_\text{sym}^+$ is positive semi-definite.
-Furthermore, \eqref{eq:desired_property_1} implies that the null space of $\widetilde{\mathbf{A}}_\text{sym}^+$ contains the null-space of $\widetilde{\mathbf{A}}_\text{sym}$.
+
+Together, equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, \eqref{eq:desired_property_3},  and \eqref{eq:desired_property_4}, and the condition $\mu > |\lambda_\text{min}|$, imply that $f(0)=0$ is the unique minimum of $f$ on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that $f$ is non-negative on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that  $\AmatPcSymPlus$ is positive semi-definite.
+Furthermore, \eqref{eq:desired_property_1} implies that the null space of $\AmatPcSymPlus$ contains the null-space of $\AmatPcSym$.
 This is important when the operator $\mathcal{A}$ has a large (or infinite) dimensional null space, or a spectrum that clusters at zero.
 Hessians in ill-posed distributed parameter inverse problems typically have both of these properties.
-Equation \eqref{eq:desired_property_4} forces the moderate and large positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}^+$ to be close to the corresponding moderate and large positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$.
-I.e., $f$ modifies the ``important'' part of the spectrum of $\widetilde{\mathbf{A}}_\text{sym}$ as little as possible.
-Equation \eqref{eq:desired_property_3} ensures that the negative eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $f$.
-There is a tradeoff: if we choose a larger value for $f(\lambda_\text{min})$, then the positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}^+$ will be closer to the corresponding positive eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ (good), but the erroneous negative eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ will be amplified in magnitude (bad).
+Equation \eqref{eq:desired_property_4} forces the moderate and large positive eigenvalues of $\AmatPcSymPlus$ to be close to the corresponding moderate and large positive eigenvalues of $\AmatPcSym$.
+I.e., $f$ modifies the ``important'' part of the spectrum of $\AmatPcSym$ as little as possible.
+Equation \eqref{eq:desired_property_3} ensures that the negative eigenvalues of $\AmatPcSym$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $f$.
+There is a tradeoff: if we choose a larger value for $f(\lambda_\text{min})$, then the positive eigenvalues of $\AmatPcSymPlus$ will be closer to the corresponding positive eigenvalues of $\AmatPcSym$ (good), but the erroneous negative eigenvalues of $\AmatPcSym$ will be amplified in magnitude (bad).
 This is shown in Figure FIG.
@@
 Each evaluation of $g$ requires solving the $3 \times 3$ linear system \eqref{eq:rational_3_linear_system}, and once the solution $\mu$ is found, we solve \eqref{eq:rational_3_linear_system} one more time to get the final values of $c_0$, $c_1$, and $c_2$.
-%
-%Row 1 corresponds to the condition $f(M)=M$, row 2 corresponds to the condition $f(0)=0$, and row 3 corresponds to the condition $f'(0)=0$.
-To find the third row of $A$, we used the fact that the derivative of $f$ is given by $$f'(\lambda) = c_1 - \frac{c_2}{(\lambda+\mu)^2}.$$
-%
-%
-%We want $f(\lambda)$ to have the following desirable properties:
-%\begin{enumerate}
-%	\item $f(\lambda)$ is nonnegative on $[m, M]$.
-%	\item $f(\lambda) \approx \lambda$ when $\lambda$ is positive and not small.
-%	\item $f(0)=0$ is the minimum of $f(\lambda)$ on $[m, M]$.
-%	\item $f(\lambda) \le |\lambda|$ when $\lambda$ is negative.
-%\end{enumerate}
-%Property 1 ensures that $\widetilde{\mathbf{A}}_\text{sym}^+$ is positive semi-definite.
-Property 2 ensures that $f$ modifies the ``important'' part of the spectrum of $\widetilde{\mathbf{A}}_\text{sym}$ as little as possible.
-That is, the part of the spectrum of $\mathcal{A}$ that we expect to be well-approximated by the product-convolution approximation.
-Property 3 ensures that $\widetilde{\mathbf{A}}_\text{sym}^+$ preserves the null-space (and near null-space) of $\widetilde{\mathbf{A}}_\text{sym}$.
-This is important because in practice, we often want to approximate operators $\mathcal{A}$ that have a large (possibly infinite) number of zero eigenvalues, and/or a sequence of eigenvalues that clusters at zero.
-Hessians in ill-posed inverse problems typically have both of these properties.
-Property 4 ensures that $f$ does not amplify erroneous negative eigenvalues $\widetilde{\mathbf{A}}_\text{sym}$.
-If $\mathcal{A}$ is positive semi-definite, then eigenvalues of $\widetilde{\mathbf{A}}_\text{sym}$ can only be negative because of errors in the product-convolution approximation.
 

### 2021-07-23 paper/localpsf.tex
H-matrix and discretization section

@@
 
-Replacing $\Aker$ with $\AkerPcMesh$ within the definition of $\Aop$ in \eqref{eq:kernel_representation} yields the following operator approximation:
-\begin{equation*}
-\left(\AopPcMesh u \right)(v) := \int_\Omega \int_\Omega v(y) \AkerPcMesh(y,x) u(x) dx dy.
-\end{equation*}
-The operator $\AopPcMesh$ maps $L^2(\Omega) \rightarrow L^2(\Omega)'$, but we are interested in the restriction of this operator to functions in the finite element space (i.e., $\AopPcMesh : V_h \rightarrow V_h'$).
+Replacing $\Aker$ with $\AkerPcMesh$ within the definition of $\Aop$ in \eqref{eq:kernel_representation} yields the following operator approximation, $\AopPcMesh$:
+\begin{equation}
+\label{eq:kernel_representation_pcmesh}
+(\AopPcMesh u)(v) := \int_\Omega \int_\Omega v(y) \AkerPcMesh(y,x) u(x) dx dy,
+\end{equation}
+Although $\AopPcMesh$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\AopPcMesh$ to functions in the finite element space (i.e., $\AopPcMesh : V_h \rightarrow V_h'$).
 %Error in the approximation $\Aop \approx \AopPcMesh$ is a combination of error in the product-convolution approximation, and error due to finite element discretization.
@@
 \begin{equation*}
-\mathbf{M}_{ij} := \int_\Omega f_i(x) f_j(x) dx
-\end{equation*}
-and $\AkerPcMat \in \mathbb{R}^{N \times N}$ is the evaluation of $\AkerPcMesh$ at all pairs of finite element Lagrange nodes, which has entries 
+\mathbf{M}_{ij} := \int_\Omega f_i(x) f_j(x) dx,
+\end{equation*}
+and $\AkerPcMat \in \mathbb{R}^{N \times N}$ is the matrix with entries
 \begin{equation*}
@@
 \end{equation}
-We convert $\mathbf{M}$ and $\AkerPcMat$ to H-matrix format, then multiply these matrices in \eqref{eq:Amatpc_defn} using fast H-matrix methods to form an H-matrix representation of $\AmatPc$.
-Once in H-matrix format, useful matrix operations such as matrix factorization and inversion may be performed quickly using fast H-matrix methods.
+
+We form an H-matrix representation of $\AmatPc$ by converting $\mathbf{M}$ and $\AkerPcMat$ to H-matrix format, then using fast H-matrix methods to multiply these matrices according to the formula in \eqref{eq:Amatpc_defn}.
+Once $\AmatPc$ is in H-matrix format, useful matrix operations such as matrix factorization and inversion may be performed quickly using fast H-matrix methods.
 We use H1 matrices in our numerical results, but any of the other H-matrix formats (such as H2, HODLR, HSS, HBS, and others \cite{HMATRIX}) could be used instead.
@@
 We convert $\AkerPcMat$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{HLIBPRO}.
-Briefly, this method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix (here, these locations are finite element the Lagrange nodes $p_i$, $i=1,\dots,N$) to construct a recursive block partitioning of $\AmatPc$, such that all blocks at the finest level of the partitioning are either low rank, or are small.
+Briefly, this method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix (here, these locations are the finite element Lagrange nodes $p_i$, $i=1,\dots,N$) to construct a recursive block partitioning of $\AkerPcMat$, such that all blocks at the finest level of the partitioning are either low rank, or are small.
 Then, the method forms low rank approximations of the blocks that are expected to be low rank by sampling specially chosen rows and columns of those blocks.

### 2021-07-26 paper/localpsf.tex
edits to h matrix construction section

@@
 \end{equation}
-We now define a further approximation, $\AkerPcMesh \approx \AkerPc$,
+We now define a further approximation, $\AkerPcMesh \approx \AkerPc$,, as follows:
 \begin{equation}
@@
 \end{equation}
-which is the interpolation of $\AkerPc$ onto $V_h \otimes V_h$.
-Error in the combined approximation, $\AkerPcMesh \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$.
+The kernel $\AkerPcMesh$ is the interpolation of $\AkerPc$ onto $V_h \otimes V_h$.
+Error in the overall kernel approximation, $\AkerPcMesh \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$.
 
@@
 Let $u_h \in V_h$ and $v_h \in V_h$, and let $\mathbf{u}$ and $\mathbf{v}$ be the coefficient vectors for $u_h$ and $v_h$, respectively, with respect to the finite element basis $f_1, \dots, f_N$.
-I.e., $u_h(x) = \sum_{i=1}^N \mathbf{u}_i f_i(x)$, and similar for $v_h$.
-By using the definition of $\AkerPcMesh$ in \eqref{eq:defn_of_Akerpcmesh} and performing algebraic manipulations, it is straightforward but tedious to show that
+I.e.,
+\begin{equation}
+\label{eq:fem_coeff_basis}
+u_h(x) = \sum_{i=1}^N \mathbf{u}_i f_i(x),
+\end{equation}
+and similar for $v_h$. Further, let $\mathbf{M} \in \mathbb{R}^{N \times N}$ be the finite element mass matrix, which has entries
+\begin{equation}
+\label{eq:mass_matrix_defn}
+\mathbf{M}_{ij} := \int_\Omega f_i(x) f_j(x) dx,
+\end{equation}
+and let $\AkerPcMat \in \mathbb{R}^{N \times N}$ be the matrix with entries
+\begin{equation}
+\label{eq:Akerpcmat_entries}
+\left(\AkerPcMat\right)_{ij} := \AkerPcMesh(p_j, p_k).
+\end{equation}
+Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, \eqref{eq:mass_matrix_defn}, and \eqref{eq:Akerpcmat_entries}, and performing algebraic manipulations, it is straightforward but tedious to show that
 \begin{equation*}
@@
 %\end{align*}
-where $\mathbf{M} \in \mathbb{R}^{N \times N}$ is the finite element mass matrix, which has entries
-\begin{equation*}
-\mathbf{M}_{ij} := \int_\Omega f_i(x) f_j(x) dx,
-\end{equation*}
-and $\AkerPcMat \in \mathbb{R}^{N \times N}$ is the matrix with entries
-\begin{equation*}
-\left(\AkerPcMat\right)_{ij} := \AkerPcMesh(p_j, p_k).
-\end{equation*}
 The finite element matrix representation of $\AopPcMesh$, which we denote by $\AmatPc$, is therefore given by
@@
 
-We form an H-matrix representation of $\AmatPc$ by converting $\mathbf{M}$ and $\AkerPcMat$ to H-matrix format, then using fast H-matrix methods to multiply these matrices according to the formula in \eqref{eq:Amatpc_defn}.
-Once $\AmatPc$ is in H-matrix format, useful matrix operations such as matrix factorization and inversion may be performed quickly using fast H-matrix methods.
+We form an H-matrix representation of $\AmatPc$ by converting $\mathbf{M}$ and $\AkerPcMat$ to H-matrix format, then using fast H-matrix methods to multiply these matrices according \eqref{eq:Amatpc_defn}.
+Once $\AmatPc$ is in H-matrix format, useful matrix operations such as matrix inversion may be performed quickly using fast H-matrix methods.
 We use H1 matrices in our numerical results, but any of the other H-matrix formats (such as H2, HODLR, HSS, HBS, and others \cite{HMATRIX}) could be used instead.
@@
 
-The mass matrix, $\mathbf{M}$, is sparse, and may therefore be converted to H-matrix format using standard H-matrix methods for sparse matrices.
-We convert $\AkerPcMat$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{HLIBPRO}.
-Briefly, this method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix (here, these locations are the finite element Lagrange nodes $p_i$, $i=1,\dots,N$) to construct a recursive block partitioning of $\AkerPcMat$, such that all blocks at the finest level of the partitioning are either low rank, or are small.
+We convert $\AkerPcMat$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within HLIBpro.
+Briefly, this method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix (here, these locations are the finite element Lagrange nodes $p_i$, $i=1,\dots,N$) to construct a recursive block partitioning of $\AkerPcMat$, such that all blocks at the finest level of the partitioning are either low rank, or small.
 Then, the method forms low rank approximations of the blocks that are expected to be low rank by sampling specially chosen rows and columns of those blocks.
@@
 Although $\AkerPcMat$ is dense, constructing the H-matrix approximation via this process only requires evaluation of $O(N \log N)$ entries of $\AkerPc$, and these entries are computed via the formula in Equation \eqref{eq:eval_y_x}.
+We convert $\mathbf{M}$ to H-matrix format using standard H-matrix methods for sparse matrices implemented within the HLIBpro software package, using the same recursive block partitioning structure as was used for $\AkerPcMat$.
 

### 2021-07-27 paper/localpsf.tex
edits

@@
 
-In this section we show how to construct an H-matrix representation of our product-convolution approximation, after discretizing the product-convolution approximation with finite elements.
-Let $f_1, f_2, \dots, f_N$ be a set of finite element basis functions used to discretize the problem, and let $V_h := \Span\left(f_1, f_2, \dots, f_N\right)$ be the corresponding finite element space.
-Here $h$ is the mesh size parameter for the finite element mesh used to discretize the problem.
+In this section we show how to construct an H-matrix representation of our product-convolution approximation, after discretizing the problem with finite elements.
+Let $f_1, f_2, \dots, f_N$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, and let $V_h := \Span\left(f_1, f_2, \dots, f_N\right)$ be the corresponding finite element space.
 Further, let $p_i \in \mathbb{R}^d$, $i=1,\dots, N$ be the Lagrange nodes associated with the functions $f_i$.
@@
 \end{equation}
-We now define a further approximation, $\AkerPcMesh \approx \AkerPc$,, as follows:
+We now define a further approximation, $\AkerPcMesh \approx \AkerPc$, as follows:
 \begin{equation}
@@
 \end{equation}
-Although $\AopPcMesh$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\AopPcMesh$ to functions in the finite element space (i.e., $\AopPcMesh : V_h \rightarrow V_h'$).
+While $\AopPcMesh$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\AopPcMesh$ to functions in the finite element space.
+That is, $\AopPcMesh : V_h \rightarrow V_h'$.
 %Error in the approximation $\Aop \approx \AopPcMesh$ is a combination of error in the product-convolution approximation, and error due to finite element discretization.
@@
 \end{equation}
-and let $\AkerPcMat \in \mathbb{R}^{N \times N}$ be the matrix with entries
+and let $\AkerPcMat \in \mathbb{R}^{N \times N}$ be following matrix of approximate kernel entries:
 \begin{equation}
 \label{eq:Akerpcmat_entries}
-\left(\AkerPcMat\right)_{ij} := \AkerPcMesh(p_j, p_k).
+\left(\AkerPcMat\right)_{ij} := \AkerPcMesh(p_i, p_j).
 \end{equation}
@@
 \begin{equation*}
-\left(\AopPcMesh u_h \right)(v_h) = \mathbf{v}^T \mathbf{M} \AkerPcMat \mathbf{M} \mathbf{u},
+\left(\AopPcMesh u_h \right)(v_h) = \mathbf{v}^T \mathbf{M} \AkerPcMat \mathbf{M} \mathbf{u}.
 \end{equation*}
@@
 
-We form an H-matrix representation of $\AmatPc$ by converting $\mathbf{M}$ and $\AkerPcMat$ to H-matrix format, then using fast H-matrix methods to multiply these matrices according \eqref{eq:Amatpc_defn}.
+We form an H-matrix representation of $\AmatPc$ by converting $\AkerPcMat$ and $\mathbf{M}$ to H-matrix format, then using fast H-matrix methods to multiply these matrices according \eqref{eq:Amatpc_defn}.
 Once $\AmatPc$ is in H-matrix format, useful matrix operations such as matrix inversion may be performed quickly using fast H-matrix methods.
@@
 	\item Recursive algorithms can take advantage of this low rank block structure to perform matrix arithmetic fast.
-Conventional algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix addition require either $O(N^2)$ or $O(N^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(N \log^a N)$ time and memory for hierarchical matrices.
-Here $N$ is the number of degrees of freedom for the discretized version of the problem, and $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix.
+Conventional algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix addition require either $O(N^2)$ or $O(N^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(N \log^a N)$ time and memory for H-matrices.
+Here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix.
 \end{itemize}
@@
 
-We convert $\AkerPcMat$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within HLIBpro.
-Briefly, this method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix (here, these locations are the finite element Lagrange nodes $p_i$, $i=1,\dots,N$) to construct a recursive block partitioning of $\AkerPcMat$, such that all blocks at the finest level of the partitioning are either low rank, or small.
+We convert $\AkerPcMat$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{HLIBPRO}.
+Briefly, this method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix (here, these locations are the finite element Lagrange nodes $p_i$, $i=1,\dots,N$) to construct a recursive block partitioning of $\AkerPcMat$, such that all blocks at the finest level of the partitioning are either expected to be low rank, or are small.
 Then, the method forms low rank approximations of the blocks that are expected to be low rank by sampling specially chosen rows and columns of those blocks.
@@
 Although $\AkerPcMat$ is dense, constructing the H-matrix approximation via this process only requires evaluation of $O(N \log N)$ entries of $\AkerPc$, and these entries are computed via the formula in Equation \eqref{eq:eval_y_x}.
-We convert $\mathbf{M}$ to H-matrix format using standard H-matrix methods for sparse matrices implemented within the HLIBpro software package, using the same recursive block partitioning structure as was used for $\AkerPcMat$.
+We convert $\mathbf{M}$ to H-matrix format using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same recursive block partitioning structure as was used for $\AkerPcMat$.
 
@@
 where $\mathbf{I}$ is the identity matrix which has the same shape as $\AmatPcSym$.
+Approximation \eqref{eq:rational_matrix_function} may be written as $\AmatPcSymPlus = \ratfct(\AmatPcSym)$, where $\ratfct$ is the following rational function:
+\begin{equation}
+\label{eq:rational_function_scalar}
+\ratfct(\lambda) := c_0 + c_1 \lambda + \frac{c_2}{\lambda+\mu}.
+\end{equation}
 The scalars $c_0$, $c_1$, $c_2$, and $\mu$ are chosen so that the moderate and large positive eigenvalues of $\AmatPcSymPlus$ approximately equal the corresponding moderate and large positive eigenvalues of $\AmatPcSym$, while the negative and small positive eigenvalues of $\AmatPcSym$ are modified so that $\AmatPcSymPlus$ is positive semi-definite.
 Later in this section we will explain how $c_0$, $c_1$, $c_2$, and $\mu$ are chosen and how our choice of these scalars affects the spectral properties of $\AmatPcSymPlus$.
-In Figure FIG we compare the spectrum of $\AmatPcSym$ to that of $\AmatPcSymPlus$ for different choices of these scalars.
+In Figure FIG we illustrate $\ratfct(\lambda)$.
 Forming $\AmatPcSymPlus$ requires computing an H-matrix inverse, $\left(\AmatPcSym + \mu \mathbf{I}\right)^{-1}$.
 While explicit computation of matrix inverses should usually be avoided, here it is computationally acceptable because $\AmatPcSym + \mu \mathbf{I}$ is an H-matrix.
-Inversion of H-matrices is a routine procedure which only requires $O(N k^2 \log(N)^2)$ work, where $k$ is the H-matrix rank \cite{Hmatrixinverse}.
-
-Approximation \eqref{eq:rational_matrix_function} may be written as $\AmatPcSymPlus = f(\AmatPcSym)$, where
-\begin{equation}
-\label{eq:rational_function_scalar}
-f(\lambda) := c_0 + c_1 \lambda + \frac{c_2}{\lambda+\mu}.
-\end{equation}
+Inversion of H-matrices is a routine procedure which only requires $O(N \log(N)^2)$ work \cite{Hmatrixinverse}.
+
 We choose $c_0, c_1, c_2$, and $\mu$, to be the solution to the following system of equations:
@@
 \begin{align}
-f(0) &= 0 \label{eq:desired_property_1} \\
-f'(0) &= 0 \label{eq:desired_property_2} \\
-f(\lambda_\text{min}) &= |\lambda_\text{min}| \label{eq:desired_property_3} \\
-f(\lambda_\text{max}) &= \lambda_\text{max}, \label{eq:desired_property_4}
+\ratfct(0) &= 0 \label{eq:desired_property_1} \\
+\ratfct'(0) &= 0 \label{eq:desired_property_2} \\
+\ratfct(\lambda_\text{min}) &= |\lambda_\text{min}| \label{eq:desired_property_3} \\
+\ratfct(\lambda_\text{max}) &= \lambda_\text{max}, \label{eq:desired_property_4}
 \end{align}
@@
 where $\lambda_\text{min}$ and $\lambda_\text{max}$ are the smallest and largest eigenvalues of $\AmatPcSym$, respectively.
-We also require $\mu > |\lambda_\text{min}|$, so that the pole of $f$ is not in $[\lambda_\text{min},\lambda_\text{max}]$.
+We also require $\mu > |\lambda_\text{min}|$, so that the pole of $\ratfct$ is not in $[\lambda_\text{min},\lambda_\text{max}]$.
 Note that $\lambda_\text{min} < 0$ since $\AmatPcSym$ is indefinite.
@@
 
-Together, equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, \eqref{eq:desired_property_3},  and \eqref{eq:desired_property_4}, and the condition $\mu > |\lambda_\text{min}|$, imply that $f(0)=0$ is the unique minimum of $f$ on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that $f$ is non-negative on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that  $\AmatPcSymPlus$ is positive semi-definite.
+Together, equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, \eqref{eq:desired_property_3},  and \eqref{eq:desired_property_4}, and the condition $\mu > |\lambda_\text{min}|$, imply that $\ratfct(0)=0$ is the unique minimum of $\ratfct$ on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that $\ratfct$ is non-negative on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that  $\AmatPcSymPlus$ is positive semi-definite.
 Furthermore, \eqref{eq:desired_property_1} implies that the null space of $\AmatPcSymPlus$ contains the null-space of $\AmatPcSym$.
@@
 Equation \eqref{eq:desired_property_4} forces the moderate and large positive eigenvalues of $\AmatPcSymPlus$ to be close to the corresponding moderate and large positive eigenvalues of $\AmatPcSym$.
-I.e., $f$ modifies the ``important'' part of the spectrum of $\AmatPcSym$ as little as possible.
-Equation \eqref{eq:desired_property_3} ensures that the negative eigenvalues of $\AmatPcSym$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $f$.
+I.e., $\ratfct$ modifies the ``important'' part of the spectrum of $\AmatPcSym$ as little as possible.
+Equation \eqref{eq:desired_property_3} ensures that the negative eigenvalues of $\AmatPcSym$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $\ratfct$.
 There is a tradeoff: if we choose a larger value for $f(\lambda_\text{min})$, then the positive eigenvalues of $\AmatPcSymPlus$ will be closer to the corresponding positive eigenvalues of $\AmatPcSym$ (good), but the erroneous negative eigenvalues of $\AmatPcSym$ will be amplified in magnitude (bad).
 This is shown in Figure FIG.
-We find that $f(\lambda_\text{min})=|\lambda_\text{min}|$ is an appropriate balance between these competing interests.
+We find that $\ratfct(\lambda_\text{min})=|\lambda_\text{min}|$ is an appropriate balance between these competing interests.
 
 Solving system \eqref{eq:four_eq_system} is computationally trivial.
-Using the definition of $f$ in \eqref{eq:rational_function_scalar}, and basic calculus and linear algebra, we may write Equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, and \eqref{eq:desired_property_4} in the following matrix form:
+Using the definition of $\ratfct$ in \eqref{eq:rational_function_scalar}, and basic calculus and linear algebra, we may write Equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, and \eqref{eq:desired_property_4} in the following matrix form:
 \begin{equation}

### 2021-07-27 paper/localpsf.tex
edits to impulse response moments section

@@
 
-We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}). 
+To briefly summarize, we extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}). 
 %To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as distinct functions. 
@@
 
-Since $\phi_x$ is nonnegative, it is a scaled probability distribution.
-Let $\widehat{\phi}_x$ denote the normalized version of $\phi_x$, and let $\mu(x)$ and $\Sigma(x)$ denote the mean and covariance of $\widehat{\phi}_x$, respectively (see Definition \ref{defn:alpha_rho_mu_sigma}).
+Because of the non-negative kernel property, for each $x \in \Omega$, the impulse response $\phi_x$ is a non-negative function $y \mapsto \phi_x(y)$.
+Hence, $\phi_x$ is a scaled probability distribution.
+Let $\alpha$ denote the scaling factor, let $\widehat{\phi}_x := \phi_x / \alpha(x)$ denote the normalized version of $\phi_x$, and let $\mu(x)$ and $\Sigma(x)$ denote the mean and covariance of $\widehat{\phi}_x$, respectively.
 We make the approximation that the support of $\phi_x$ is contained within the ellipsoid
@@
 The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\phi}_x$ (see Figure FIG).
-The bigger $\tau$ is, the smaller the overlap between impulse responses within a batch will be, while the smaller $\tau$ is, the more impulse responses we can fit within each batch.
-In practice we find that $\tau=3$ provides a good balance between these competing goals.
-The fraction of the ``mass'' of $\phi_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though we find this bound is conservative and far less mass resides in this region in practice.
-
-In Algorithm \ref{alg:varhpi_mean_cov} we show how to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$ for all points $x$ simultaneously, by applying $\mathcal{A}^T$ to a constant function and a small number of linear and quadratic functions.
-Definitions required for Algorithm \ref{alg:varhpi_mean_cov} are given in Definition \ref{defn:alpha_rho_mu_sigma}, Definition \ref{defn:C_L_Q}, and Definition \ref{defn:pointwise}.
+The quantity $\tau$ is a parameter that must be chosen by the user of the algorithm.
+The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\phi_x$.
+However, the cost of our algorithm will depend on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
+We find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results.
+The fraction of the ``mass'' of $\phi_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.
+% a good balance between these competing goals. 
+%The bigger $\tau$ is, the smaller the overlap between impulse responses within a batch will be, while the smaller $\tau$ is, the more impulse responses we can fit within each batch.
+
+The ellipsoid $E_x$ is not immediately available.
+To use $E_x$, we must first compute the scalar $\alpha(x)$, the length-$d$ vector $\mu(x)$, and the $d \times d$ matrix $\Sigma(x)$ (recall $d \in \{1,2,3\}$ is the spatial dimension of $\Omega$).
+The direct approach to compute these quantities is to apply $\Aop$ to a delta distribution centered at point $x$ to get $\phi_x$ (as per \eqref{eq:impulse_response_delta_action}), then post-process $\phi_x$ to compute these quantities.
+But this direct approach is computationally infeasible because and it requires applying $\Aop$ to one vector per point $x$ that we wish to evaluate $E_x$ at, and we will need to evaluate $E_x$ at a large number of points.
+Fortunately, it is possible to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$, \emph{for all points $x$ simultaneously}, by applying $\mathcal{A}^T$ to one constant function, $d$ linear functions, and $d(d+1)/2$ quadratic functions.
+We present this efficient method of computing $\alpha$, $\mu$, and $\Sigma$ in Theorem \ref{thm:vol_mean_cov} and summarize the method in Algorithm \ref{alg:varhpi_mean_cov}.
+
+%In Algorithm \ref{alg:varhpi_mean_cov} we show how to compute  Definitions required for Algorithm \ref{alg:varhpi_mean_cov} are given in Definition \ref{defn:alpha_rho_mu_sigma}, Definition \ref{defn:C_L_Q}, and Definition \ref{defn:pointwise}.
 The theoretical basis for Algorithm \ref{alg:varhpi_mean_cov} is shown in Theorem \ref{thm:vol_mean_cov}.
@@
 
-\begin{defn}[Moments of $\phi_x$]
-	\label{defn:alpha_rho_mu_sigma}
-	We define the spatially varying scaling factor, $\alpha$, the scaled impulse response, $\widehat{\phi}_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
-	\begin{align*}
-	\alpha:\Omega \rightarrow \mathbb{R}, &\qquad \alpha(x) := \int_{\Omega} \phi_x(z) dz, \\
-	\widehat{\phi}_x:\Omega \rightarrow \mathbb{R}, &\qquad \widehat{\phi}_x := \phi_x \big/ \alpha(x), \\
-	\mu:\Omega \rightarrow \mathbb{R}^d, &\qquad \mu(x) := \int_\Omega z \widehat{\phi}_x(z) dz, \\
-	\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\Sigma(x) := \int_\Omega (z - \mu(x))(z - \mu(x))^T \widehat{\phi}_x(z) dz.
-	\end{align*}
-\end{defn}
-
-\begin{defn}[Constant, linear, and quadratic functions]	
-	\label{defn:C_L_Q}
-	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$.
-We define $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions, respectively:
+\begin{theorem}[Impulse response moments]
+	\label{thm:vol_mean_cov}
+	Let $\alpha:\Omega \rightarrow \mathbb{R}$, $\mu:\Omega \rightarrow \mathbb{R}^d$, and $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$, be the following functions:
+	\begin{equation*}
+	\alpha(x) := \int_{\Omega} \phi_x(y) dy, \qquad
+	\mu(x) := \mathbb{E}(\widehat{\phi}_x), \qquad
+	\Sigma(x) := \mathbb{V}(\widehat{\phi}_x),
+	\end{equation*}
+	where $\mathbb{E}(\widehat{\phi}_x)$ denotes the mean of $\widehat{\phi}_x$, and $\mathbb{V}(\widehat{\phi}_x)$ denotes the variance of $\widehat{\phi}_x$.	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$.
+Let $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions:
 	\begin{equation*}
@@
 	\end{equation*}
-\end{defn}
-
-\begin{defn}[Pointwise operations]
-	\label{defn:pointwise}
-	We write $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
-\end{defn}
+	Let $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
+The following results hold:
+	\begin{align*}
+	\alpha =& \left(\mathcal{A}^T C\right)^* \\
+	\mu^i =& \left(\mathcal{A}^T L^i\right)^* / \alpha \\
+	\Sigma^{ij} =& \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot \mu^j
+	\end{align*}
+	 for $i=1,\dots, d$, $j=1,\dots,d$.
+\end{theorem}
+
+The proof of Theorem \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}. The proof is straightforward, and follows directly from the definitions.
+
+%
+%\begin{defn}[Constant, linear, and quadratic functions]	
+%	\label{defn:C_L_Q}
+%	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$.
+We define $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions, respectively:
+%	\begin{equation*}
+%	C(x) := 1, \qquad
+%	L^i(x) := x^i, \qquad
+%	Q^{ij}(x) = x^i x^j.
+%	\end{equation*}
+%\end{defn}
+%
+%\begin{defn}[Pointwise operations]
+%	\label{defn:pointwise}
+%	We write $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
+%\end{defn}
 
@@
 
-\begin{theorem}[Correctness of Algorithm \ref{alg:varhpi_mean_cov}]
-	\label{thm:vol_mean_cov}
-	Algorithm \ref{alg:varhpi_mean_cov} correctly computes $\alpha$, $\mu$, and $\Sigma$.
-\end{theorem}
-
-The proof of Theorem \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}.
-
 \subsection{Greedy sample point selection}
@@
 
-Now that we have estimates of the supports of the functions $\phi_x$, we use these estimates, combined with a greedy algorithm to choose sets of sample points, $S_b$.
+Now that we have estimates of the supports of the functions $\phi_x$, we use these estimates, combined with a greedy ellipsoid packing algorithm to choose sets of sample points, $S_b$.
 These sample points will be used in Section \ref{sec:get_impulse_response} to form Dirac combs, which are then used to compute impulse responses of $\mathcal{A}$.
-Starting from a finite set of candidate points $P$, we select batches of sample points, $S_b$, using a greedy algorithm.
-To build a batch $S_b$, first we initialize $S_b$ as an empty set.
-Then we select a candidate point $p \in P$ that is the farthest away from all points in previous sample point batches.
+We start with a finite set of candidate points $P$.
+To build a batch of sample points, $S_b$, first we initialize $S_b$ as an empty set.
+Then we select the point $p \in P$ that is the farthest away from all points in previous sample point batches.
 That is, $p$ is a maximizer of the following optimization problem:
@@
 Otherwise we discard $p$.
-Recall that $E_p$ and $E_q$ are ellipsoids defined in \eqref{eq:support_ellipsoid}.
+Recall that $E_p$ and $E_q$ are ellipsoids associated with points $p$ and $q$, respectively, as defined in \eqref{eq:support_ellipsoid}.
 This process repeats until there are no more points in $P$.
@@
 
+\begin{defn}[Moments of $\phi_x$]
+	\label{defn:alpha_rho_mu_sigma}
+	We define the spatially varying scaling factor, $\alpha$, the scaled impulse response, $\widehat{\phi}_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
+	\begin{align*}
+	\alpha:\Omega \rightarrow \mathbb{R}, &\qquad \alpha(x) := \int_{\Omega} \phi_x(z) dz, \\
+	\widehat{\phi}_x:\Omega \rightarrow \mathbb{R}, &\qquad \widehat{\phi}_x := \phi_x \big/ \alpha(x), \\
+	\mu:\Omega \rightarrow \mathbb{R}^d, &\qquad \mu(x) := \int_\Omega z \widehat{\phi}_x(z) dz, \\
+	\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\Sigma(x) := \int_\Omega (z - \mu(x))(z - \mu(x))^T \widehat{\phi}_x(z) dz.
+	\end{align*}
+\end{defn}
+
 %These results rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}.

### 2021-07-28 paper/localpsf.tex
edits to impulse response moments section

@@
 
-We present a fast matrix free method for approximating locally translation-invariant operators, $\mathcal{A}$, that have locally supported non-negative integral kernels.
+We present a fast matrix free method for approximating locally translation-invariant operators, $\Aop$, that have locally supported non-negative integral kernels.
 Such operators arise, for example, in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, as Poincare-Steklov operators, as covariance operators, as blurring operators in imaging problems, and as Hessians in distributed parameter PDE-constrained inverse problems.
 
-Let $\Omega \subset \mathbb{R}^d$ be a bounded domain in dimension $d=1$, $2$, or $3$, and let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the form
+Let $\Omega \subset \mathbb{R}^d$ be a bounded domain in dimension $d=1$, $2$, or $3$, and let $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the form
 \begin{equation}
 \label{eq:kernel_representation}
-(\mathcal{A}u)(v) := \int_\Omega \int_\Omega v(y) \Aker(y,x) u(x) dx dy,
+(\Aop u)(v) := \int_\Omega \int_\Omega v(y) \Aker(y,x) u(x) dx dy,
 \end{equation}
 with integral kernel $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$.
-Here $\mathcal{A}u \in L^2(\Omega)'$ is the linear functional that results from applying $\mathcal{A}$ to $u\in L^2(\Omega)$, and $\left(\mathcal{A}u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
+Here $\Aop u \in L^2(\Omega)'$ is the linear functional that results from applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
 We focus on operators of this form which have the following properties:
@@
 \item[Matrix-free:] One cannot easily access kernel entries $\Phi(y,x)$.
-Access to $\mathcal{A}$ is available only through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary functions (i.e., operator actions).
+Access to $\Aop$ is available only through application of $\Aop$ and $\Aop^T$ to arbitrary functions (i.e., operator actions).
 That is, evaluation of the maps
 \begin{equation*}
-	u \mapsto\mathcal{A}u \quad \text{and} \quad v \mapsto\mathcal{A}^Tv.
+	u \mapsto\Aop u \quad \text{and} \quad v \mapsto\Aop^Tv.
 \end{equation*} 
@@
 \end{equation} 
-We say that $\mathcal{A}$ has local support if, for all $x\in \Omega$, the support of $\phi_x$ is contained (or approximately contained) in a neighborhood of $x$.
+We say that $\Aop$ has local support if, for all $x\in \Omega$, the support of $\phi_x$ is contained (or approximately contained) in a neighborhood of $x$.
 The smaller these supports are, the better our algorithm will perform.
@@
 
-After discretization (see Appendix \ref{app:discretized_operations}), the operator $\mathcal{A}$ becomes a dense matrix, $\mathbf{A}$, that is typically too large to be built and stored.
+After discretization (see Appendix \ref{app:discretized_operations}), the operator $\Aop$ becomes a dense matrix, $\mathbf{A}$, that is typically too large to be built and stored.
 The first property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\mathbf{A}$ and $\mathbf{A}^T$ with arbitrary vectors, but one cannot easily access matrix entries $\mathbf{A}_{ij}$.
@@
 \label{eq:impulse_response_delta_action}
-	\phi_x = \left\langle \mathcal{A}, \delta_x \right\rangle^*,
-\end{equation}
-where $\left\langle \mathcal{A}, \delta_x \right\rangle \in L^2(\Omega)'$ denotes the result of applying the operator $\mathcal{A}$ to the distribution $\delta_x$, and $\left\langle \mathcal{A}, \delta_x \right\rangle^* \in L^2(\Omega)$ denotes the Riesz representation of $\langle\mathcal{A}, \delta_x\rangle$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\psi \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\psi^* \in L^2(\Omega)$ such that $\psi(v) = \left(\psi^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\mathcal{A}$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, the action of $\mathcal{A}$ may be extended to distributions in a natural way, as described in Appendix \ref{app:distributions}.
+	\phi_x = \left\langle \Aop, \delta_x \right\rangle^*,
+\end{equation}
+where $\left\langle \Aop, \delta_x \right\rangle \in L^2(\Omega)'$ denotes the result of applying the operator $\Aop$ to the distribution $\delta_x$, and $\left\langle \Aop, \delta_x \right\rangle^* \in L^2(\Omega)$ denotes the Riesz representation of $\langle\Aop, \delta_x\rangle$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\psi \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\psi^* \in L^2(\Omega)$ such that $\psi(v) = \left(\psi^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, the action of $\Aop$ may be extended to distributions in a natural way, as described in Appendix \ref{app:distributions}.
 %The action of $\mathcal{A}$ on a distribution $\mu$ is written as $\langle \mathcal{A}, \mu \rangle$.
@@
 %We take the Riesz representative of the linear functional $\left\langle \mathcal{A}, \delta_x \right\rangle \in L^2(\Omega)'$ to convert it into a function $\left\langle \mathcal{A}, \delta_x \right\rangle^* \in L^2(\Omega)$.
-After discretization, the process of applying $\mathcal{A}$ to the distribution $\delta_x$, then forming the Riesz representative of the resulting linear functional, amounts to computing $\boldsymbol{\phi}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x$, where $\boldsymbol{\phi}_x$, $\mathbf{A}$, and $\boldsymbol{\delta}_x$ are discretized versions of $\phi_x$, $A$ and $\delta_x$, respectively, and $\mathbf{M}$ is a discretization of the $L^2$ inner product bilinear form (e.g., a mass matrix in the finite element context).
-From \eqref{eq:impulse_response_delta_action}, we see that the local support property means that the response of $\mathcal{A}$ to a point source at $x$ is zero (or small) at points $y$ that are far from $x$.
+After discretization, the process of applying $\Aop$ to the distribution $\delta_x$, then forming the Riesz representative of the resulting linear functional, amounts to computing $\boldsymbol{\phi}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x$, where $\boldsymbol{\phi}_x$, $\mathbf{A}$, and $\boldsymbol{\delta}_x$ are discretized versions of $\phi_x$, $A$ and $\delta_x$, respectively, and $\mathbf{M}$ is a discretization of the $L^2$ inner product bilinear form (e.g., a mass matrix in the finite element context).
+From \eqref{eq:impulse_response_delta_action}, we see that the local support property means that the response of $\Aop$ to a point source at $x$ is zero (or small) at points $y$ that are far from $x$.
 
 Local translation invariance is illustrated in Figure FIG.
-If $\mathcal{A}$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then it is straightforward to show that $\mathcal{A}$ would be a convolution operator, and the convolution kernel, $\varphi_x$, would be given by
+If $\Aop$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then it is straightforward to show that $\Aop$ would be a convolution operator, and the convolution kernel, $\varphi_x$, would be given by
 \begin{equation}
@@
 Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes.
-We therefore approximate $\mathcal{A}$ by a spatially varying weighted sum of convolution operators, $\mathcal{A}_\text{pc} \approx \mathcal{A}$, of the form
+We therefore approximate $\Aop$ by a spatially varying weighted sum of convolution operators, $\AopPc \approx \Aop$, of the form
 \begin{equation}
 \label{eq:product_convolution}
-	\left(\mathcal{A}_\text{pc}~u\right)(v) := \left( v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
+	\left(\AopPc~u\right)(v) := \left( v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
 \end{equation}
@@
 
-We compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$.
+We compute one ``batch'' of $\varphi_i$'s by applying $\Aop$ to a sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$.
 The batch of points $x_i$ are chosen so that the support of $\phi_i$ and the support of $\phi_j$ do not overlap (or do not overlap much) if $i \neq j$.
-Because these supports do not overlap, we can post-process the response of $\mathcal{A}$ to the Dirac comb to recover the functions $\phi_i$, and therefore the functions $\varphi_i$, associated with all points $x_i$ in the batch---with one application of $\mathcal{A}$, we recover many $\varphi_i$.
+Because these supports do not overlap, we can post-process the response of $\Aop$ to the Dirac comb to recover the functions $\phi_i$, and therefore the functions $\varphi_i$, associated with all points $x_i$ in the batch---with one application of $\Aop$, we recover many $\varphi_i$.
 This is illustrated in Figure FIG.
@@
 In order to choose the points $x_i$, we need to estimate the support of $\phi_i$ \emph{before} we compute it.
-The supports of the functions $\phi_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\mathcal{A}^T$ to a small number of specially chosen functions (three functions if $d=1$, six if $d=2$, ten if $d=3$), then post processing the results (Section \ref{eq:mean_and_covariance_estimation}).
+The supports of the functions $\phi_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of specially chosen functions (three functions if $d=1$, six if $d=2$, ten if $d=3$), then post processing the results (Section \ref{eq:mean_and_covariance_estimation}).
 This procedure is the only part of the paper that requires non-negativity of the integral kernel.
@@
 
-Once the $\varphi_i$ and $w_i$ are computed, we convert $\mathcal{A}_\text{pc}$ to hierarchical matrix (H-matrix) format.
-Let $\Phi_\text{pc}$ denote the integral kernel associated with $\mathcal{A}_\text{pc}$.
+Once the $\varphi_i$ and $w_i$ are computed, we convert $\AopPc$ to hierarchical matrix (H-matrix) format.
+Let $\Phi_\text{pc}$ denote the integral kernel associated with $\AopPc$.
 While $\Phi(y,x)$ is not easily computable, $\Phi_\text{pc}(y,x)$ is given by the formula
@@
 This follows from~\eqref{eq:product_convolution} and the fact that the convolution operator $u \mapsto \varphi_i \ast u$ has $(y,x)$ kernel entry given by $\varphi(y-x)$.
-Formula~\eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\mathcal{A}_\text{pc}$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
-Once in H-matrix format, fast H-matrix arithmetic is used to invert $\mathcal{A}_\text{pc}$, or perform other useful matrix operations.
-
-Often we are interested in approximating operators $\mathcal{A}$ that are symmetric positive semi-definite.
-Unfortunately, $\mathcal{A}_\text{pc}$ is, in general, non-symmetric and indefinite due to errors in the product convolution approximation.
-In this case, we modify the H-matrix representation of $\mathcal{A}_\text{pc}$ to make it positive definite, by symmetrizing it, then applying a specially chosen rational matrix function to the symmetrized H-matrix.
+Formula~\eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\AopPc$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
+Once in H-matrix format, fast H-matrix arithmetic is used to invert $\AopPc$, or perform other useful matrix operations.
+
+Often we are interested in approximating operators $\Aop$ that are symmetric positive semi-definite.
+Unfortunately, $\AopPc$ is, in general, non-symmetric and indefinite due to errors in the product convolution approximation.
+In this case, we modify the H-matrix representation of $\AopPc$ to make it positive definite, by symmetrizing it, then applying a specially chosen rational matrix function to the symmetrized H-matrix.
 
@@
 
-We build upon the class of product convolution approximations in which the functions, $\varphi_i$, are impulse responses of $\mathcal{A}$ to point sources at a collection of points $x_i$ \cite{PRODCONVLIST}.
+We build upon the class of product convolution approximations in which the functions, $\varphi_i$, are impulse responses of $\Aop$ to point sources at a collection of points $x_i$ \cite{PRODCONVLIST}.
 A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the functions $\varphi_i$ with piecewise linear interpolation \cite{NAGY}, or splines \cite{PRODCONVSPLINES}.
 With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
-This is computationally prohibitive in our applications, because computing the functions $\varphi_i$ requires one matrix vector product with $\mathcal{A}$ for each point $x_i$, and these matrix vector products are computationally costly.
+This is computationally prohibitive in our applications, because computing the functions $\varphi_i$ requires one matrix vector product with $\Aop$ for each point $x_i$, and these matrix vector products are computationally costly.
 In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
-But even with adaptive refinement, many matrix vector products with $\mathcal{A}$ may be required.
+But even with adaptive refinement, many matrix vector products with $\Aop$ may be required.
 In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\varphi_i$ are computed with each matrix vector product.
@@
 Our method of estimating the support of the functions $\varphi_i$ was inspired by resolution analysis in seismic imaging \cite{RESOLUTION}.
-In resolution analysis, $\mathcal{A}$ is the Hessian for a seismic inverse problem, and the width of $\varphi_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
+In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\varphi_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
 If the width of $\varphi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
-In \cite{RESOLUTION}, the width of $\varphi_p$ is estimated to be the local autocorrelation length of the function $\mathcal{A}^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
-In this paper, rather than probing $\mathcal{A}^T$ with random noise functions, we use specific constant, linear, and quadratic functions.
-We observe that our method estimates the support of $\varphi_p$ more accurately and reliably than resolution analysis, but our method requires that $\mathcal{A}$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
+In \cite{RESOLUTION}, the width of $\varphi_p$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
+In this paper, rather than probing $\Aop^T$ with random noise functions, we use specific constant, linear, and quadratic functions.
+We observe that our method estimates the support of $\varphi_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
 
@@
 Conventional H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
-There are matrix-free methods for H-matrix approximation \cite{LEXINGPEELINGPROCESS}, and these methods have been used to form H-matrix compressions of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
+There are matrix-free methods for H-matrix approximation \cite{LEXINGPEELINGPROCESS}, and these methods have been used to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
 While these matrix-free H-matrix construction methods are asymptotically scalable in theory, the required number of matrix vector products can be large in practice.
@@
 \label{sec:method}
-This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\mathcal{A}_\text{pc} \approx \mathcal{A}$, given in \eqref{eq:product_convolution}.
-We also describe how we evaluate kernel entries of the approximation, $\Phi_\text{pc}(y,x)$, how we convert $\mathcal{A}_\text{pc}$ to H-matrix format, and (optionally) how we modify $\mathcal{A}_\text{pc}$ to force it to be symmetric positive semi-definite.
-
-To briefly summarize, we extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}). 
+This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\AopPc \approx \Aop$, given in \eqref{eq:product_convolution}.
+We also describe how we evaluate kernel entries of the approximation, $\Phi_\text{pc}(y,x)$, how we convert $\AopPc$ to H-matrix format, and (optionally) how we modify $\AopPc$ to force it to be symmetric positive semi-definite.
+
+To briefly summarize, we extract $\varphi_{x_i}$ for many points $x_i$ by applying $\Aop$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}). 
 %To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as distinct functions. 
@@
 %\end{itemize}
-The complete algorithm for constructing $\mathcal{A}_\text{pc}$ is shown in Algorithm \ref{alg:construct_Atilde}.
+The complete algorithm for constructing $\AopPc$ is shown in Algorithm \ref{alg:construct_Atilde}.
 The method for evaluating kernel entries of $\Phi_\text{pc}$ is shown in Section \ref{sec:eval_matrix_entries}.
-The method for converting $\mathcal{A}_\text{pc}$ to H-matrix format is shown in Section \ref{sec:H_matrix_conversion}.
+The method for converting $\AopPc$ to H-matrix format is shown in Section \ref{sec:H_matrix_conversion}.
 The method for modifying the resulting H-matrix to make it symmetric positive semi-definite is shown in Section \ref{sec:make_spd}.
@@
 	{	
-		\Input{Operator $\mathcal{A}$, parameter $\text{max\_batches}$}
+		\Input{Linear operator $\Aop$, parameter $\text{max\_batches}$}
 		\Output{Dirac comb responses $\{\eta^b\}_{b=1}^{\text{max\_batches}}$, \\
@@
 
-			Compute $\eta^b$ by applying $\mathcal{A}$ to the dirac comb associated with $S_b$ (Section \ref{sec:get_impulse_response})
+			Compute $\eta^b$ by applying $\Aop$ to the dirac comb associated with $S_b$ (Section \ref{sec:get_impulse_response})
 
@@
 	}
-	\caption{Construct product-convolution approximation $\widetilde{\mathcal{A}}$}
+	\caption{Construct product-convolution approximation $\AopPc$}
 	\label{alg:construct_Atilde}
@@
 To use $E_x$, we must first compute the scalar $\alpha(x)$, the length-$d$ vector $\mu(x)$, and the $d \times d$ matrix $\Sigma(x)$ (recall $d \in \{1,2,3\}$ is the spatial dimension of $\Omega$).
-The direct approach to compute these quantities is to apply $\Aop$ to a delta distribution centered at point $x$ to get $\phi_x$ (as per \eqref{eq:impulse_response_delta_action}), then post-process $\phi_x$ to compute these quantities.
+The direct approach to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$ is to apply $\Aop$ to a delta distribution centered at point $x$ to get $\phi_x$, as per \eqref{eq:impulse_response_delta_action}.
+Then one can post-process $\phi_x$ to determine $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$.
 But this direct approach is computationally infeasible because and it requires applying $\Aop$ to one vector per point $x$ that we wish to evaluate $E_x$ at, and we will need to evaluate $E_x$ at a large number of points.
-Fortunately, it is possible to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$, \emph{for all points $x$ simultaneously}, by applying $\mathcal{A}^T$ to one constant function, $d$ linear functions, and $d(d+1)/2$ quadratic functions.
+Fortunately, it is possible to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$, \emph{for all points $x$ simultaneously}, by applying $\Aop^T$ to one constant function, $d$ linear functions, and $d(d+1)/2$ quadratic functions.
 We present this efficient method of computing $\alpha$, $\mu$, and $\Sigma$ in Theorem \ref{thm:vol_mean_cov} and summarize the method in Algorithm \ref{alg:varhpi_mean_cov}.
@@
 	\mu(x) := \mathbb{E}(\widehat{\phi}_x), \qquad
-	\Sigma(x) := \mathbb{V}(\widehat{\phi}_x),
-	\end{equation*}
-	where $\mathbb{E}(\widehat{\phi}_x)$ denotes the mean of $\widehat{\phi}_x$, and $\mathbb{V}(\widehat{\phi}_x)$ denotes the variance of $\widehat{\phi}_x$.	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$.
+	\Sigma(x) := \Var(\widehat{\phi}_x),
+	\end{equation*}
+	where $\mathbb{E}(\widehat{\phi}_x)$ denotes the mean of $\widehat{\phi}_x$, and $\Var(\widehat{\phi}_x)$ denotes the variance of $\widehat{\phi}_x$.	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$.
 Let $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions:
@@
 	\begin{align*}
-	\alpha =& \left(\mathcal{A}^T C\right)^* \\
-	\mu^i =& \left(\mathcal{A}^T L^i\right)^* / \alpha \\
-	\Sigma^{ij} =& \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot \mu^j
+	\alpha =& \left(\Aop^T C\right)^* \\
+	\mu^i =& \left(\Aop^T L^i\right)^* / \alpha \\
+	\Sigma^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \alpha - \mu^i\cdot \mu^j
 	\end{align*}
@@
 
-\subsection{Batched computation of impulse responses}
+\subsection{Computation and storage of impulse responses}
 \label{sec:get_impulse_response}

### 2021-07-28 paper/localpsf.tex
edits to impulse response regular grid discussion

@@
 We compute impulse responses of $\mathcal{A}$ in batches by applying $\mathcal{A}$ to a small number Dirac combs. 
-The batches of impulse responses, $\eta^b$, will ultimately be interpolated with weighting functions described in Section \ref{sec:weighting_functions} to form the desired approximation $\widetilde{\mathcal{A}}\approx \mathcal{A}$.
-The Dirac comb, $\xi_b$, associated with a batch of sample points $S_b$ is the sum of Dirac distributions (point sources) centered at the points $x_i \in S_b$.
-That is,
+%The batches of impulse responses, $\eta^b$, will ultimately be interpolated with weighting functions described in Section \ref{sec:weighting_functions} to form the desired approximation $\widetilde{\mathcal{A}}\approx \mathcal{A}$.
+The Dirac comb, $\xi_b$, associated with a batch of sample points, $S_b$, is the sum of Dirac distributions (point sources) centered at the points $x_i \in S_b$. That is,
 \begin{equation*}
@@
 \end{equation*}
-For each batch $S_b$, we compute as the action of $\mathcal{A}$ on the associated Dirac comb:
+For each batch $S_b$, we compute the action of $\mathcal{A}$ on the associated Dirac comb:
 \begin{equation}
@@
 \end{equation}
-By linearity, $\eta^b$ may be written as
+By linearity, the resulting function $\eta^b$ may be written as
 \begin{equation}
@@
 \end{equation}
-for all $x_i \in S_b$.
-By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover $\varphi_{x_i}$ for every point $x_i \in S_b$.
-To minimize memory usage, the functions $\varphi_i$ are not stored individually.
+for all $x_i \in S_b$. By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover $\varphi_{x_i}$ for every point $x_i \in S_b$.
+
+Later we will need to perform function evaluations $z \mapsto \varphi_{x_i}(z)$ for a large number of points $z \in \Omega$.
+Using \eqref{eq:varphi_eval} to perform this task requires evaluating the function $\eta_b$ at a large number of points $z+x_i$.
+In computations the functions $\eta^b$ will be finite element functions formed on an irregular mesh, and the points $z+x_i$ will typically not be gridpoints of the mesh.
+Evaluating a finite element function at an arbitrary point requires finding which cell in the mesh that the point is in, and then evaluating several polynomial functions to evaluate the finite element function at the point in that cell.
+While this procedure is asymptotically scalable, it tends to be slow in practice \cite{FEMEVALSLOW}.
+To speed up evaluation of the function $\varphi_{x_i}$, we interpolate $\varphi_{x_i}$ onto a regular rectilinear grid on a coordinate axis aligned box that just barely contains the ellipsoid $E_{x_i}$.
+The function $\varphi_{x_i}$ is transferred from the irregular mesh to the regular rectilinear grid by evaluating $\varphi_{x_i}$ at all gridpoints in the regular rectilinear grid using \eqref{eq:varphi_eval} and the slow finite element evaluation process described previously.
+Once $\varphi_{x_i}$ has been transferred to the regular rectilinear grid, further evaluations of $\varphi_{x_i}$ are performed extremely quickly using standard algorithms for regular rectilinear grid interpolation.
+To minimize the error associated with transferring the function $\varphi_{x_i}$ from the irregular mesh to the regular rectilinear grid, the density of the regular rectilinear grid should be denser than the local density of the finite element mesh.
+In our numerical results, the distance between neighboring points in the regular rectilinear grid is chosen to be half the minimum distance between any pair of finite element Lagrange nodes that are contained in the ellipsoid $E_{x_i}$.
+This procedure is illustrated in Figure FIG.
+
+%To minimize memory usage, the functions $\varphi_i$ are not stored individually.
 Instead, we store the functions $\eta^b$, and use formula \eqref{eq:varphi_eval} to perform function evaluations $z \mapsto \varphi_i(z)$ as needed.
@@
 
-\subsection{Evaluating kernel entries of the approximation}
-\label{sec:eval_matrix_entries}
-
-We now describe how to efficiently evaluate $\widetilde{\Phi}(y,x)$ at an arbitrary point $(x,y) \in \Omega \times \Omega$. From \eqref{eq:kernel_entries}, we have
-\begin{equation}
-\label{eq:eval_y_x_000}
-\AkerPc(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x),
-\end{equation}
-and from \eqref{eq:varphi_eval}, we have
-\begin{equation}
-\label{eq:kernel_entry_formua}
-\varphi_{x_i}(y-x) = \begin{cases}
-\eta^{b(i)}(y - x + x_i), & y - x + x_i \in E_{x_i} \\
-0, & \text{otherwise},
-\end{cases}
-\end{equation}
-where $b(i)$ denotes the batch index such that $x_i \in S^{b(i)}$.
+%\subsection{Evaluating kernel entries of the approximation}
+%\label{sec:eval_matrix_entries}
+%
+%We now describe how to efficiently evaluate $\widetilde{\Phi}(y,x)$ at an arbitrary point $(x,y) \in \Omega \times \Omega$. From \eqref{eq:kernel_entries}, we have
+%\begin{equation}
+%\label{eq:eval_y_x_000}
+%\AkerPc(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x),
+%\end{equation}
+%and from \eqref{eq:varphi_eval}, we have
+%\begin{equation}
+%\label{eq:kernel_entry_formua}
+%\varphi_{x_i}(y-x) = \begin{cases}
+%\eta^{b(i)}(y - x + x_i), & y - x + x_i \in E_{x_i} \\
+%0, & \text{otherwise},
+%\end{cases}
+%\end{equation}
+%where $b(i)$ denotes the batch index such that $x_i \in S^{b(i)}$.
 To compute $z=\AkerPc(y,x)$, we start with $z=0$, then loop through all batches $b$, and all sample points in each batch.
@@
 This is shown in Algorithm \ref{alg:eval_y_x}.
-
-\begin{algorithm2e}
-	\SetAlgoNoLine
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Output}
-
-	\Input{points $x \in \Omega$, $y \in \Omega$, \\sample point batches $S_1, \dots, S_\text{max\_batches}$, \\Dirac comb responses $\{\eta^b\}_{b=1}^{\text{max\_batches}}$, \\weighting functions $\{w_i\}_{i=1}^r$}
-	\Output{Kernel entry $z = \widetilde{A}(y,x)$}
-
-	$z \gets 0$
-
-	\For{$b=1,2,\dots,\text{}max\_batches$}{
-		\For{$x_i \in S_b$}{
-			\If{$y - x + x_i \in E_{x_i}$}{
-				$z \gets z + w_i(x)\eta^b(y-x+x_i)$
-			}
-		}
-	}	
-	\caption{Compute approximate kernel entry $z = \widetilde{A}(y,x)$}
-	\label{alg:eval_y_x}
-\end{algorithm2e}
+%
+%
+%\begin{algorithm2e}
+%	\SetAlgoNoLine
+%	\SetKwInOut{Input}{Input}
+%	\SetKwInOut{Output}{Output}
+%	
+%	\Input{points $x \in \Omega$, $y \in \Omega$, \\sample point batches $S_1, \dots, S_\text{max\_batches}$, \\Dirac comb responses $\{\eta^b\}_{b=1}^{\text{max\_batches}}$, \\weighting functions $\{w_i\}_{i=1}^r$}
+%	\Output{Kernel entry $z = \widetilde{A}(y,x)$}
+%
+%	$z \gets 0$
+%	
+%	\For{$b=1,2,\dots,\text{}max\_batches$}{
+%		\For{$x_i \in S_b$}{
+%			\If{$y - x + x_i \in E_{x_i}$}{
+%				$z \gets z + w_i(x)\eta^b(y-x+x_i)$
+%			}
+%		}
+%	}	
+%	\caption{Compute approximate kernel entry $z = \widetilde{A}(y,x)$}
+%	\label{alg:eval_y_x}
+%\end{algorithm2e}
 

### 2021-07-28 paper/localpsf.tex
edits to impulse response regular grid discussion

@@
 
-Later we will need to perform function evaluations $z \mapsto \varphi_{x_i}(z)$ for a large number of points $z \in \Omega$.
+Later we will need to perform function evaluations $z \mapsto \varphi_{x_i}(z)$ for a large number of points $z$.
 Using \eqref{eq:varphi_eval} to perform this task requires evaluating the function $\eta_b$ at a large number of points $z+x_i$.
@@
 The function $\varphi_{x_i}$ is transferred from the irregular mesh to the regular rectilinear grid by evaluating $\varphi_{x_i}$ at all gridpoints in the regular rectilinear grid using \eqref{eq:varphi_eval} and the slow finite element evaluation process described previously.
-Once $\varphi_{x_i}$ has been transferred to the regular rectilinear grid, further evaluations of $\varphi_{x_i}$ are performed extremely quickly using standard algorithms for regular rectilinear grid interpolation.
-To minimize the error associated with transferring the function $\varphi_{x_i}$ from the irregular mesh to the regular rectilinear grid, the density of the regular rectilinear grid should be denser than the local density of the finite element mesh.
+Once $\varphi_{x_i}$ has been transferred to the regular rectilinear grid, further evaluations of $\varphi_{x_i}$ are performed using fast regular rectilinear grid interpolation.
+To limit the error associated with transferring the function $\varphi_{x_i}$ from the irregular mesh to the regular rectilinear grid, the density of the regular rectilinear grid should be denser than the local density of the finite element mesh.
 In our numerical results, the distance between neighboring points in the regular rectilinear grid is chosen to be half the minimum distance between any pair of finite element Lagrange nodes that are contained in the ellipsoid $E_{x_i}$.
-This procedure is illustrated in Figure FIG.
+This grid transfer procedure is illustrated in Figure FIG.
 

### 2021-07-28 paper/localpsf.tex
edits to impulse response regular grid discussion

@@
 With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
-This is computationally prohibitive in our applications, because computing the functions $\varphi_i$ requires one matrix vector product with $\Aop$ for each point $x_i$, and these matrix vector products are computationally costly.
+%This is computationally prohibitive in our applications, because computing the functions $\varphi_i$ requires one matrix vector product with $\Aop$ for each point $x_i$, and these matrix vector products are computationally costly.
 In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
@@
 We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\phi_x$ (Section \ref{eq:mean_and_covariance_estimation}).
+The support of $\phi_x$ is estimated to be contained within an ellipsoid, so the process of choosing batches of points is an ellipse packing problem.
 For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
@@
 The quantity $\tau$ is a parameter that must be chosen by the user of the algorithm.
-The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\phi_x$.
+The larger $\tau$ is, the larger the ellipsoid $E_x$ is, the more conservative the estimate is for the support of $\phi_x$.
 However, the cost of our algorithm will depend on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
@@
 To use $E_x$, we must first compute the scalar $\alpha(x)$, the length-$d$ vector $\mu(x)$, and the $d \times d$ matrix $\Sigma(x)$ (recall $d \in \{1,2,3\}$ is the spatial dimension of $\Omega$).
-The direct approach to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$ is to apply $\Aop$ to a delta distribution centered at point $x$ to get $\phi_x$, as per \eqref{eq:impulse_response_delta_action}.
+The direct approach to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\phi_x$, as per \eqref{eq:impulse_response_delta_action}.
 Then one can post-process $\phi_x$ to determine $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$.
@@
 
-\subsection{Computation and storage of impulse responses}
+\subsection{Computation, evaluation, and storage of impulse responses}
 \label{sec:get_impulse_response}
 
-We compute impulse responses of $\mathcal{A}$ in batches by applying $\mathcal{A}$ to a small number Dirac combs. 
+We compute impulse responses of $\mathcal{A}$ in batches by applying $\mathcal{A}$ to a small number Dirac combs. Each Dirac comb yields one batch of impulse responses.
 %The batches of impulse responses, $\eta^b$, will ultimately be interpolated with weighting functions described in Section \ref{sec:weighting_functions} to form the desired approximation $\widetilde{\mathcal{A}}\approx \mathcal{A}$.
@@
 \end{equation}
-Since the points $x_i$ are chosen so that the support of $\phi_i$ and the support of $\phi_j$ do not overlap (or do not overlap much) when $i \neq j$, from \eqref{eq:convolution_kernel} and \eqref{eq:phi_b} we have
+Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\phi_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\phi_j$ do not overlap when $i \neq j$, from \eqref{eq:convolution_kernel} and \eqref{eq:phi_b} we have
 \begin{equation}
@@
 \end{equation}
-for all $x_i \in S_b$. By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover $\varphi_{x_i}$ for every point $x_i \in S_b$.
-
-Later we will need to perform function evaluations $z \mapsto \varphi_{x_i}(z)$ for a large number of points $z$.
-Using \eqref{eq:varphi_eval} to perform this task requires evaluating the function $\eta_b$ at a large number of points $z+x_i$.
-In computations the functions $\eta^b$ will be finite element functions formed on an irregular mesh, and the points $z+x_i$ will typically not be gridpoints of the mesh.
-Evaluating a finite element function at an arbitrary point requires finding which cell in the mesh that the point is in, and then evaluating several polynomial functions to evaluate the finite element function at the point in that cell.
-While this procedure is asymptotically scalable, it tends to be slow in practice \cite{FEMEVALSLOW}.
-To speed up evaluation of the function $\varphi_{x_i}$, we interpolate $\varphi_{x_i}$ onto a regular rectilinear grid on a coordinate axis aligned box that just barely contains the ellipsoid $E_{x_i}$.
-The function $\varphi_{x_i}$ is transferred from the irregular mesh to the regular rectilinear grid by evaluating $\varphi_{x_i}$ at all gridpoints in the regular rectilinear grid using \eqref{eq:varphi_eval} and the slow finite element evaluation process described previously.
-Once $\varphi_{x_i}$ has been transferred to the regular rectilinear grid, further evaluations of $\varphi_{x_i}$ are performed using fast regular rectilinear grid interpolation.
-To limit the error associated with transferring the function $\varphi_{x_i}$ from the irregular mesh to the regular rectilinear grid, the density of the regular rectilinear grid should be denser than the local density of the finite element mesh.
-In our numerical results, the distance between neighboring points in the regular rectilinear grid is chosen to be half the minimum distance between any pair of finite element Lagrange nodes that are contained in the ellipsoid $E_{x_i}$.
+for all $x_i \in S_b$. By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the functions $\varphi_{x_i}$ for every point $x_i \in S_b$.
+
+To convert $\AopPc$ to H-matrix format, we will need to evaluate the function $z \mapsto \varphi_{x_i}(z)$ for a large number of points $z$.
+Using \eqref{eq:varphi_eval} to evaluate $\varphi_{x_i}(z)$ requires evaluating the function $\eta_b(z+x_i)$.
+However, in computations the function $\eta^b$ will be a finite element function formed on an irregular mesh, and the point $z+x_i$ will typically not be a gridpoint of the mesh.
+Evaluating a finite element function at an arbitrary point requires finding which cell in the mesh the point is in, then evaluating several polynomial functions.
+This procedure is asymptotically scalable, but slow in practice \cite{FEMEVALSLOW}.
+To speed up evaluation of $\varphi_{x_i}$, we interpolate $\varphi_{x_i}$ onto a regular rectilinear grid on a coordinate axis aligned box that just barely contains the ellipsoid $E_{x_i}$.
+We transfer the function $\varphi_{x_i}$ from the irregular mesh to the regular grid by evaluating $\varphi_{x_i}$ at all gridpoints in the regular grid, using \eqref{eq:varphi_eval} and the slow finite element point evaluation process described previously.
+Once $\varphi_{x_i}$ has been transferred to the regular grid, we use fast regular grid interpolation to perform further evaluations of $\varphi_{x_i}$.
+We incur interpolation error when transferring $\varphi_{x_i}$ from the irregular mesh to the regular grid.
+To limit this interpolation error, the density of the regular grid should be denser than the local density of the finite element mesh.
+In our numerical results, the distance between neighboring points in the regular grid is chosen to be half the minimum distance between any pair of finite element Lagrange nodes that are contained in $E_{x_i}$.
 This grid transfer procedure is illustrated in Figure FIG.
@@
 
-In this section we show how to construct an H-matrix representation of our product-convolution approximation, after discretizing the problem with finite elements.
+In this section we discuss how we construct an H-matrix representation of our product-convolution approximation, after discretizing the problem with finite elements.
 Let $f_1, f_2, \dots, f_N$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, and let $V_h := \Span\left(f_1, f_2, \dots, f_N\right)$ be the corresponding finite element space.
@@
 \label{eq:defn_of_Akerpcmesh}
-\AkerPcMesh(y,x) := \sum_{i=1}^N \sum_{j=1}^N \AkerPc(p_i,p_j) f_i(y) f_j(x)
+\AkerPcMesh(y,x) := \sum_{i=1}^N \sum_{j=1}^N \AkerPc(p_i,p_j) f_i(y) f_j(x).
 \end{equation}
@@
 We form an H-matrix representation of $\AmatPc$ by converting $\AkerPcMat$ and $\mathbf{M}$ to H-matrix format, then using fast H-matrix methods to multiply these matrices according \eqref{eq:Amatpc_defn}.
-Once $\AmatPc$ is in H-matrix format, useful matrix operations such as matrix inversion may be performed quickly using fast H-matrix methods.
+Once $\AmatPc$ is in H-matrix format, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods.
 We use H1 matrices in our numerical results, but any of the other H-matrix formats (such as H2, HODLR, HSS, HBS, and others \cite{HMATRIX}) could be used instead.
-
-The basic ideas behind H-matrices are as follows:
-\begin{itemize} 
-	\item Often, large dense matrices of practical interest may be permuted, then partitioned into blocks recursively, in such a way that many off-diagonal blocks of the matrix are numerically low rank, even if the matrix is high rank.
-	\item Recursive algorithms can take advantage of this low rank block structure to perform matrix arithmetic fast.
-Conventional algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix addition require either $O(N^2)$ or $O(N^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(N \log^a N)$ time and memory for H-matrices.
-Here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix.
-\end{itemize}
 For more details on H-matrices, we recommend \cite{HMATRIXGOOD}.
-
-We convert $\AkerPcMat$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{HLIBPRO}.
+We discuss H-matrices in greater detail in Appendix \ref{app:h_matrix}
+
+We convert $\AkerPcMat$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBPro software package \cite{HLIBPRO}.
 Briefly, this method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix (here, these locations are the finite element Lagrange nodes $p_i$, $i=1,\dots,N$) to construct a recursive block partitioning of $\AkerPcMat$, such that all blocks at the finest level of the partitioning are either expected to be low rank, or are small.
@@
 The small high rank blocks are formed directly by evaluating their entries.
-We describe the H-matrix construction in detail in Appendix \ref{app:h_matrix}.
-Although $\AkerPcMat$ is dense, constructing the H-matrix approximation via this process only requires evaluation of $O(N \log N)$ entries of $\AkerPc$, and these entries are computed via the formula in Equation \eqref{eq:eval_y_x}.
-We convert $\mathbf{M}$ to H-matrix format using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same recursive block partitioning structure as was used for $\AkerPcMat$.
+We describe this H-matrix construction process in detail in Appendix \ref{app:h_matrix}.
+Although $\AkerPcMat$ is a dense $N \times N$ matrix, constructing the H-matrix approximation of $\AkerPcMat$ only requires evaluation of $O(N \log N)$ entries of $\AkerPc$, and these entries are computed via the formula in Equation \eqref{eq:eval_y_x}.
+A dense representation of $\AkerPcMat$ is never formed.
+We convert $\mathbf{M}$ to H-matrix format using standard H-matrix methods for sparse matrices implemented within HLIBPro, using the same recursive block partitioning structure as was used for $\AkerPcMat$.
 
@@
 
-\section{Hierarchical matrix construction}
+\section{Hierarchical matrix details}
 \label{app:h_matrix}
 
-The H-matrix construction process that we use proceeds as follows.
+The basic ideas behind H-matrices are as follows:
+\begin{itemize} 
+	\item Often, large dense matrices of practical interest may be permuted, then partitioned into blocks recursively, in such a way that many off-diagonal blocks of the matrix are numerically low rank, even if the matrix is high rank.
+	\item Recursive algorithms can take advantage of this low rank block structure to perform matrix arithmetic fast.
+Conventional algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix addition require either $O(N^2)$ or $O(N^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(N \log^a N)$ time and memory for H-matrices.
+Here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix.
+\end{itemize}
+For more details on H-matrices, we recommend \cite{HMATRIXGOOD}.
+
+The H-matrix construction process that we use to form the matrix $\AkerPcMat$ proceeds as follows.
 First, we construct hierarchical partitionings of the degrees of freedom for the columns and rows of the matrix (cluster trees, Section \ref{sec:cluster_trees}).

### 2021-07-29 paper/localpsf.tex
edits based on comments from J

@@
 \end{equation}
-Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\phi_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\phi_j$ do not overlap when $i \neq j$, from \eqref{eq:convolution_kernel} and \eqref{eq:phi_b} we have
+Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\phi_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\phi_j$ do not overlap when $i \neq j$, from \eqref{eq:convolution_kernel} and \eqref{eq:phi_b} we have (approximately)
 \begin{equation}
@@
 		\eta^b(z+x_i), & z+x_i \in E_{x_i} \\
-		0, & \text{otherwise}.
+		0, & \text{otherwise}
 		\end{cases}
 \end{equation}
-for all $x_i \in S_b$. By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the functions $\varphi_{x_i}$ for every point $x_i \in S_b$.
-
-To convert $\AopPc$ to H-matrix format, we will need to evaluate the function $z \mapsto \varphi_{x_i}(z)$ for a large number of points $z$.
-Using \eqref{eq:varphi_eval} to evaluate $\varphi_{x_i}(z)$ requires evaluating the function $\eta_b(z+x_i)$.
-However, in computations the function $\eta^b$ will be a finite element function formed on an irregular mesh, and the point $z+x_i$ will typically not be a gridpoint of the mesh.
-Evaluating a finite element function at an arbitrary point requires finding which cell in the mesh the point is in, then evaluating several polynomial functions.
-This procedure is asymptotically scalable, but slow in practice \cite{FEMEVALSLOW}.
+for all $x_i \in S_b$. By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the convolution kernels $\varphi_{x_i}$ associated with every point $x_i \in S_b$.
+
+To convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), we will need to evaluate $\varphi_{x_i}(z)$ for a large number of points $z$.
+Per \eqref{eq:varphi_eval}, evaluating $\varphi_{x_i}(z)$ requires evaluating $\eta_b(z+x_i)$.
+This is not ideal.
+In computations, $\eta^b$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh.
+Evaluating $\eta_b(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell.
+Although this procedure is asymptotically scalable, it is slow in practice \cite{FEMEVALSLOW}.
 To speed up evaluation of $\varphi_{x_i}$, we interpolate $\varphi_{x_i}$ onto a regular rectilinear grid on a coordinate axis aligned box that just barely contains the ellipsoid $E_{x_i}$.
-We transfer the function $\varphi_{x_i}$ from the irregular mesh to the regular grid by evaluating $\varphi_{x_i}$ at all gridpoints in the regular grid, using \eqref{eq:varphi_eval} and the slow finite element point evaluation process described previously.
-Once $\varphi_{x_i}$ has been transferred to the regular grid, we use fast regular grid interpolation to perform further evaluations of $\varphi_{x_i}$.
+We transfer the function $\varphi_{x_i}$ from the irregular mesh to the regular grid by evaluating $\varphi_{x_i}$ at all gridpoints in the regular grid, using the slow process just described.
+This is an upfront cost which is done ``offline.'' After $\varphi_{x_i}$ has been transferred to the regular grid, we use fast regular grid interpolation to perform all subsequent ``online'' evaluations of $\varphi_{x_i}$.
+This yields a substantial overall speedup because the number of gridpoints in the regular grid for $\varphi_{x_i}$ is typically much smaller than the number of evaluations of $\varphi_{x_i}$ required to convert $\AopPc$ to H-matrix format.
+Having $\varphi_{x_i}$ on a regular grid also simplifies the implementation of the boundary impulse response extension procedure that we describe in Section \ref{sec:boundary_extension}.
+
 We incur interpolation error when transferring $\varphi_{x_i}$ from the irregular mesh to the regular grid.
-To limit this interpolation error, the density of the regular grid should be denser than the local density of the finite element mesh.
+To limit this interpolation error, the density of the regular grid should be as dense or denser than the local density of the finite element mesh.
 In our numerical results, the distance between neighboring points in the regular grid is chosen to be half the minimum distance between any pair of finite element Lagrange nodes that are contained in $E_{x_i}$.
-This grid transfer procedure is illustrated in Figure FIG.
+The grid transfer procedure is illustrated in Figure FIG.
 
@@
 Instead, we store the functions $\eta^b$, and use formula \eqref{eq:varphi_eval} to perform function evaluations $z \mapsto \varphi_i(z)$ as needed.
+
+\subsection{Boundary impulse response extension}
+\label{sec:boundary_extension}
 
@@
 Also, these weighting functions sum to one pointwise (Proposition \ref{thm:wi_sum_to_one}).
+Proofs for results in this section are given in Appendix \ref{app:proofs}.
 
@@
 All of the weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added.
-However, the dominant computational cost in constructing the weighting functions is the solution of the Poisson PDEs, and these Poisson PDE solves can be performed incrementally---we only need to perform two new Poisson PDE solves when a new point is added, rather than $2(r+1)$ new PDE solves.
+However, the dominant computational cost in constructing the weighting functions is the solution of the Poisson PDEs, and these Poisson PDE solves can be performed incrementally---we only need to perform two new Poisson PDE solves when a new point is added.
+%, rather than $2(r+1)$ new PDE solves.
 
@@
 When inverting for parameter fields defined on four or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used.
-Proofs for results in this section are given in Appendix \ref{app:proofs}.
-
-\begin{thm}[Optimal interpolation]
+
+\begin{thm}[Poisson weighting functions are optimal]
 	\label{thm:smoothest_interpolant}
@@
 
-\begin{prop}[Weighting functions sum to one]
+\begin{prop}[Poisson weighting functions sum to one]
 	\label{thm:wi_sum_to_one}
@@
 For more details on H-matrices, we recommend \cite{HMATRIXGOOD}.
-We discuss H-matrices in greater detail in Appendix \ref{app:h_matrix}
+We discuss H-matrices in greater detail in Appendix \ref{app:h_matrix}.
 

### 2021-07-30 paper/localpsf.tex
starting on boundary extension discussion

@@
 
-\subsection{Computation, evaluation, and storage of impulse responses}
+\subsection{Computation and storage of convolution kernels}
 \label{sec:get_impulse_response}
@@
 
-\subsection{Boundary impulse response extension}
+\subsection{Addressing convolution kernel boundary artifacts}
 \label{sec:boundary_extension}
+
+Evaluating the kernel for our product-convolution approximation requires evaluating $\varphi_{x_i}(y-x)$.
+But this may be undefined, because $y-x$ may reside outside the support of $\varphi_{x_i}$, even if $x \in \Omega$ and $y \in \Omega$.
+The default approach to deal with this problem is to extend $\varphi_{x_i}$ by zero wherever it is undefined.
+But this leads to substantial errors, known as boundary artifacts, when the sample point $x_i$ is near the boundary of the domain.
+In this case, the impulse response is artificially chopped off by the domain boundary.
+See Figure FIG for an illustration, and see \cite{ADAPTIVECONVMYPAPER} for a more detailed discussion of boundary artifacts in.
+
+To address boundary artifacts, we fill in missing/undefined portions of convolution kernels using values from convolution kernels for nearby sample points.
+This is similar to the approach used in \cite{ADAPTIVECONVMYPAPER}, except
 
@@
 
-The basic ideas behind H-matrices are as follows:
-\begin{itemize} 
-	\item Often, large dense matrices of practical interest may be permuted, then partitioned into blocks recursively, in such a way that many off-diagonal blocks of the matrix are numerically low rank, even if the matrix is high rank.
-	\item Recursive algorithms can take advantage of this low rank block structure to perform matrix arithmetic fast.
-Conventional algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix addition require either $O(N^2)$ or $O(N^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(N \log^a N)$ time and memory for H-matrices.
+Often, large dense matrices of practical interest may be permuted, then partitioned into blocks recursively, in such a way that many off-diagonal blocks of the matrix are numerically low rank, even if the matrix is high rank.
+Such matrices are known as hierarchical matrices (H-matrices).
+Many classes of H-matrices exist (H1, H2, HSS, HBS, and more), and all of these types of H-matrices could be used in conjunction with our product-convolution approximation.
+Here we use classical H1 matrices.
+For this section, when we say H-matrix, we are referring to H1 matrices.
+
+While a dense $N \times N$ matrix traditionally requires $O(N^2)$ memory to store, H-matrices may be stored using $O(N \log N)$ memory, by storing only the low rank factors for the low rank blocks.
+Recursive algorithms can take advantage of the H-matrix low rank block structure to perform matrix arithmetic fast.
+Conventional dense matrix algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix-matrix addition require either $O(N^2)$ or $O(N^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(N \log^a N)$ time and memory for H-matrices.
 Here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix.
-\end{itemize}
 For more details on H-matrices, we recommend \cite{HMATRIXGOOD}.
 
-The H-matrix construction process that we use to form the matrix $\AkerPcMat$ proceeds as follows.
+\subsection{H-matrix construction}
+
+The H-matrix construction process that we use to form the H-matrix representation of $\AkerPcMat$ proceeds as follows.
 First, we construct hierarchical partitionings of the degrees of freedom for the columns and rows of the matrix (cluster trees, Section \ref{sec:cluster_trees}).
@@
 The first two steps require geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix, but these steps do not depend on the particular values of matrix entries.
-The third step requires us to evaluate $O(N \log N)$ specially-chosen entries of $\widetilde{A}$, and we evaluate these entries using the method described in Section \ref{sec:eval_matrix_entries}.
-
-\subsection{Cluster trees}
+The third step requires us to evaluate $O(N \log N)$ specially-chosen entries of $\AkerPcMat$, and we evaluate these entries using \eqref{eq:Akerpcmat_entries}.
+
+\subsubsection{Cluster trees}
 \label{sec:cluster_trees}
@@
 
-\subsection{Block cluster tree} 
+\subsubsection{Block cluster tree} 
 \label{sec:block_cluster_tree}
@@
 
-\subsection{Adaptive cross low rank approximation of blocks}
+\subsubsection{Adaptive cross low rank approximation of blocks}
 \label{sec:adaptive_cross}

### 2021-07-31 paper/localpsf.tex
boundary extension written okish

@@
 This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\AopPc \approx \Aop$, given in \eqref{eq:product_convolution}.
-We also describe how we evaluate kernel entries of the approximation, $\Phi_\text{pc}(y,x)$, how we convert $\AopPc$ to H-matrix format, and (optionally) how we modify $\AopPc$ to force it to be symmetric positive semi-definite.
+We also describe how we convert $\AopPc$ to H-matrix format, and (optionally) how we modify $\AopPc$ to force it to be symmetric positive semi-definite.
 
@@
 
-\subsection{Estimating impulse response supports}
+\subsection{Sample point selection}
+
+\subsubsection{Estimating impulse response supports}
 \label{eq:mean_and_covariance_estimation}
@@
 
-\subsection{Greedy sample point selection}
+\subsubsection{Greedy ellipsoid packing algorithm}
 \label{sec:greedy_point_selection}
@@
 
-\subsection{Computation and storage of convolution kernels}
+\subsection{Convolution kernels}
 \label{sec:get_impulse_response}
@@
 for all $x_i \in S_b$. By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the convolution kernels $\varphi_{x_i}$ associated with every point $x_i \in S_b$.
+
+\subsubsection{Interpolation onto local regular grids}
 
@@
 Although this procedure is asymptotically scalable, it is slow in practice \cite{FEMEVALSLOW}.
+
 To speed up evaluation of $\varphi_{x_i}$, we interpolate $\varphi_{x_i}$ onto a regular rectilinear grid on a coordinate axis aligned box that just barely contains the ellipsoid $E_{x_i}$.
-We transfer the function $\varphi_{x_i}$ from the irregular mesh to the regular grid by evaluating $\varphi_{x_i}$ at all gridpoints in the regular grid, using the slow process just described.
+We transfer the function $\varphi_{x_i}$ from the irregular mesh to the regular grid by evaluating $\varphi_{x_i}$ at all gridpoints in the regular grid, using the slow process described in the previous paragraph.
 This is an upfront cost which is done ``offline.'' After $\varphi_{x_i}$ has been transferred to the regular grid, we use fast regular grid interpolation to perform all subsequent ``online'' evaluations of $\varphi_{x_i}$.
 This yields a substantial overall speedup because the number of gridpoints in the regular grid for $\varphi_{x_i}$ is typically much smaller than the number of evaluations of $\varphi_{x_i}$ required to convert $\AopPc$ to H-matrix format.
-Having $\varphi_{x_i}$ on a regular grid also simplifies the implementation of the boundary impulse response extension procedure that we describe in Section \ref{sec:boundary_extension}.
+%Having $\varphi_{x_i}$ stored on a regular grid also simplifies the implementation of the method that we use to deal with boundary artifacts, which will be described in Section \ref{sec:boundary_extension}.
 
@@
 
+When $x_i$ is near a boundary, it may occur that some gridpoints for the regular grid are in locations where $\varphi_{x_i}$ is undefined.
+In this case, we store a placeholder value for $\varphi_{x_i}$ at those gridpoints.
+We use the IEEE standard value of ``NaN'' (not a number) for this placeholder.
+These NaN values are later replaced with more appropriate values, as we will discuss in Section \ref{sec:boundary_extension}.
+
 %To minimize memory usage, the functions $\varphi_i$ are not stored individually.
@@
 
-\subsection{Addressing convolution kernel boundary artifacts}
+\subsubsection{Addressing boundary artifacts}
 \label{sec:boundary_extension}
 
-Evaluating the kernel for our product-convolution approximation requires evaluating $\varphi_{x_i}(y-x)$.
-But this may be undefined, because $y-x$ may reside outside the support of $\varphi_{x_i}$, even if $x \in \Omega$ and $y \in \Omega$.
+Evaluating the kernel for our product-convolution approximation, $\AkerPc(y,x)$, requires evaluating $\varphi_{x_i}(y-x)$.
+But $\varphi_{x_i}(y-x)$ is undefined  when $y - x + x_i$ is outside $\Omega$, which may occur even if $x \in \Omega$ and $y \in \Omega$.
 The default approach to deal with this problem is to extend $\varphi_{x_i}$ by zero wherever it is undefined.
-But this leads to substantial errors, known as boundary artifacts, when the sample point $x_i$ is near the boundary of the domain.
+But this leads to substantial errors, known as boundary artifacts, which are especially significant when the sample point $x_i$ is near the boundary of the domain.
 In this case, the impulse response is artificially chopped off by the domain boundary.
-See Figure FIG for an illustration, and see \cite{ADAPTIVECONVMYPAPER} for a more detailed discussion of boundary artifacts in.
-
-To address boundary artifacts, we fill in missing/undefined portions of convolution kernels using values from convolution kernels for nearby sample points.
-This is similar to the approach used in \cite{ADAPTIVECONVMYPAPER}, except
-
-\subsection{Poisson interpolation weighting functions}
+See Figure FIG for an illustration, and see \cite{ADAPTIVECONVMYPAPER} for more detailed discussion of boundary artifacts.
+
+While $\varphi_{x_i}$ may be undefined at a point $z$, often there is a sample point $x_j$ that is near $x_i$, but further away from the boundary than $x_i$, such that $\varphi_{x_j}$ is well defined at $z$.
+To address boundary artifacts, we fill in missing/undefined portions of convolution kernels using information from other convolution kernels associated with the $k$ nearest sample points.
+Kernels associated with nearer sample points are given precedence over kernels associated with farther away sample points.
+After the convolution kernel has been extended using information from neighboring convolution kernels, it is then further extended by zero as needed.
+
+Let $x_{n_1}, x_{n_2}, \dots, x_{n_k}$ be the $k$ nearest sample points to $x_i$, ordered by nearness, including $x_i$ itself.
+That is, $x_{n_1}=x_i$ (the point itself), $x_{n_2}$ is the nearest sample point to $x_i$ other than $x_i$ itself, $x_{n_3}$ is the second nearest sample point to $x_i$ other than $x_i$ itself, and so on.
+Also, recall that in computations that the functions $\varphi_{x_i}$ are stored on regular rectilinear grids.
+To extend $\varphi_{x_i}$, first we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\varphi_{x_i}$, but is large enough to contain the boxes associated with all of the neighbor kernels $\varphi_{x_{n_1}}, \varphi_{x_{n_2}}, \dots, \varphi_{x_{n_k}}$.
+We initialize an extended convolution kernel, $\varphi_{x_i}^E$, to zero on this expanded grid.
+Then we interpolate the functions $\varphi_{x_{n_j}}$ onto this expanded grid one at a time in reverse order.
+That is, first we interpolate the kernel, $\varphi_{x_{n_k}}$, associated with the furthest point from $x_i$ onto the grid, then the kernel, $\varphi_{x_{n_{k-1}}}$, associated with the second furthest point, and so on until we reach the original kernel $\varphi_{x_{n_1}} = \varphi_{x_i}$.
+At each step, we overwrite existing values of $\varphi_{x_i}^E$ with new values from the kernel currently being interpolated, except when the new value of $\varphi_{x_{n_j}}$ is undefined (``NaN'').
+At gridpoints where the kernel being interpolated is undefined, we leave $\varphi_{x_i}^E$ unchanged.
+In this way, we fill in undefined values of $\varphi_{x_i}$ with values from successively further and further away kernels.
+This process is illustrated in Figure FIG.
+To avoid notational clutter, and since we will henceforth only work with the extended kernel, $\varphi_{x_i}^E$, in the remainder of this paper the symbol $\varphi_{x_i}$ refers to $\varphi_{x_i}^E$.
+
+\subsection{Poisson weighting functions}
 \label{sec:weighting_functions}
@@
 
-\subsection{Hierarchical matrix construction}
+\subsection{Postprocessing the product-convolution approximation}
+
+\subsubsection{Hierarchical matrix construction}
 \label{sec:H_matrix_conversion}
@@
 
-\subsection{Rational positive semi-definite modification}
+\subsubsection{Rational positive semi-definite modification}
 \label{sec:make_spd}

### 2021-07-31 paper/localpsf.tex
boundary extension written okish

@@
 
-\subsubsection{Greedy ellipsoid packing algorithm}
+\subsubsection{Greedy ellipsoid packing}
 \label{sec:greedy_point_selection}
 
-Now that we have estimates of the supports of the functions $\phi_x$, we use these estimates, combined with a greedy ellipsoid packing algorithm to choose sets of sample points, $S_b$.
+Now that we have estimates of the supports of the functions $\phi_x$, we use these estimates, combined with a greedy ellipsoid packing algorithm, to choose sets of sample points, $S_b$.
 These sample points will be used in Section \ref{sec:get_impulse_response} to form Dirac combs, which are then used to compute impulse responses of $\mathcal{A}$.
@@
 Kernels associated with nearer sample points are given precedence over kernels associated with farther away sample points.
-After the convolution kernel has been extended using information from neighboring convolution kernels, it is then further extended by zero as needed.
-
-Let $x_{n_1}, x_{n_2}, \dots, x_{n_k}$ be the $k$ nearest sample points to $x_i$, ordered by nearness, including $x_i$ itself.
+After the convolution kernel has been extended using information from neighboring convolution kernels, it is then further extended by zero.
+
+In detail, let $x_{n_1}, x_{n_2}, \dots, x_{n_k}$ be the $k$ nearest sample points to $x_i$, ordered by nearness, including $x_i$ itself.
 That is, $x_{n_1}=x_i$ (the point itself), $x_{n_2}$ is the nearest sample point to $x_i$ other than $x_i$ itself, $x_{n_3}$ is the second nearest sample point to $x_i$ other than $x_i$ itself, and so on.
 Also, recall that in computations that the functions $\varphi_{x_i}$ are stored on regular rectilinear grids.
-To extend $\varphi_{x_i}$, first we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\varphi_{x_i}$, but is large enough to contain the boxes associated with all of the neighbor kernels $\varphi_{x_{n_1}}, \varphi_{x_{n_2}}, \dots, \varphi_{x_{n_k}}$.
+To extend $\varphi_{x_i}$, first we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\varphi_{x_i}$, and contains that grid as a subgrid, but is large enough to contain the boxes associated with all of the neighbor kernels $\varphi_{x_{n_1}}, \varphi_{x_{n_2}}, \dots, \varphi_{x_{n_k}}$.
 We initialize an extended convolution kernel, $\varphi_{x_i}^E$, to zero on this expanded grid.
 Then we interpolate the functions $\varphi_{x_{n_j}}$ onto this expanded grid one at a time in reverse order.
-That is, first we interpolate the kernel, $\varphi_{x_{n_k}}$, associated with the furthest point from $x_i$ onto the grid, then the kernel, $\varphi_{x_{n_{k-1}}}$, associated with the second furthest point, and so on until we reach the original kernel $\varphi_{x_{n_1}} = \varphi_{x_i}$.
+That is, first we interpolate the $\varphi_{x_{n_k}}$ onto the grid, then the kernel $\varphi_{x_{n_{k-1}}}$, and so on until we reach the original kernel $\varphi_{x_{n_1}} = \varphi_{x_i}$.
 At each step, we overwrite existing values of $\varphi_{x_i}^E$ with new values from the kernel currently being interpolated, except when the new value of $\varphi_{x_{n_j}}$ is undefined (``NaN'').

### 2021-07-31 paper/localpsf.tex
boundary extension written okish

@@
 Kernels associated with nearer sample points are given precedence over kernels associated with farther away sample points.
-After the convolution kernel has been extended using information from neighboring convolution kernels, it is then further extended by zero.
+%After the convolution kernel has been extended using information from neighboring convolution kernels, it is then further extended by zero.
 
@@
 Then we interpolate the functions $\varphi_{x_{n_j}}$ onto this expanded grid one at a time in reverse order.
-That is, first we interpolate the $\varphi_{x_{n_k}}$ onto the grid, then the kernel $\varphi_{x_{n_{k-1}}}$, and so on until we reach the original kernel $\varphi_{x_{n_1}} = \varphi_{x_i}$.
+That is, first we interpolate $\varphi_{x_{n_k}}$ onto the grid, then $\varphi_{x_{n_{k-1}}}$, and so on until we reach the original kernel $\varphi_{x_{n_1}} = \varphi_{x_i}$.
 At each step, we overwrite existing values of $\varphi_{x_i}^E$ with new values from the kernel currently being interpolated, except when the new value of $\varphi_{x_{n_j}}$ is undefined (``NaN'').

### 2021-07-31 paper/localpsf.tex
edits based on comments by J

@@
 Per \eqref{eq:varphi_eval}, evaluating $\varphi_{x_i}(z)$ requires evaluating $\eta_b(z+x_i)$.
-This is not ideal.
 In computations, $\eta^b$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh.
@@
 In this case, the impulse response is artificially chopped off by the domain boundary.
-See Figure FIG for an illustration, and see \cite{ADAPTIVECONVMYPAPER} for more detailed discussion of boundary artifacts.
+See Figure FIG for an illustration.
 
@@
 
-In detail, let $x_{n_1}, x_{n_2}, \dots, x_{n_k}$ be the $k$ nearest sample points to $x_i$, ordered by nearness, including $x_i$ itself.
-That is, $x_{n_1}=x_i$ (the point itself), $x_{n_2}$ is the nearest sample point to $x_i$ other than $x_i$ itself, $x_{n_3}$ is the second nearest sample point to $x_i$ other than $x_i$ itself, and so on.
+In detail, let $p_1, p_2, \dots, p_k$ be the $k$ nearest sample points to $x_i$, ordered by nearness, including $x_i$ itself.
+That is, $p_1=x_i$, $p_2$ is the nearest neighbor to $x_i$, $p_3$ is the second nearest neighbor to $x_i$, and so on.
 Also, recall that in computations that the functions $\varphi_{x_i}$ are stored on regular rectilinear grids.
-To extend $\varphi_{x_i}$, first we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\varphi_{x_i}$, and contains that grid as a subgrid, but is large enough to contain the boxes associated with all of the neighbor kernels $\varphi_{x_{n_1}}, \varphi_{x_{n_2}}, \dots, \varphi_{x_{n_k}}$.
+To extend $\varphi_{x_i}$, first we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\varphi_{x_i}$, and contains that grid as a subgrid, but is large enough to contain the boxes associated with all of the neighbor kernels, $\varphi_{p_1}, \varphi_{p_2}, \dots, \varphi_{p_k}$.
 We initialize an extended convolution kernel, $\varphi_{x_i}^E$, to zero on this expanded grid.
-Then we interpolate the functions $\varphi_{x_{n_j}}$ onto this expanded grid one at a time in reverse order.
-That is, first we interpolate $\varphi_{x_{n_k}}$ onto the grid, then $\varphi_{x_{n_{k-1}}}$, and so on until we reach the original kernel $\varphi_{x_{n_1}} = \varphi_{x_i}$.
-At each step, we overwrite existing values of $\varphi_{x_i}^E$ with new values from the kernel currently being interpolated, except when the new value of $\varphi_{x_{n_j}}$ is undefined (``NaN'').
+Then we interpolate the functions $\varphi_{p_j}$ onto this expanded grid one at a time in reverse order.
+That is, first we interpolate $\varphi_{p_k}$ onto the grid, then $\varphi_{p_{k-1}}$, and so on until we reach the original kernel $\varphi_{p_1} = \varphi_{x_i}$.
+At each step, we overwrite existing values of $\varphi_{x_i}^E$ with new values from the kernel currently being interpolated, except when the new value of $\varphi_{p_j}$ is undefined (``NaN'').
 At gridpoints where the kernel being interpolated is undefined, we leave $\varphi_{x_i}^E$ unchanged.

### 2021-08-02 paper/localpsf.tex
edits

@@
 \label{sec:method}
-This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\AopPc \approx \Aop$, given in \eqref{eq:product_convolution}.
-We also describe how we convert $\AopPc$ to H-matrix format, and (optionally) how we modify $\AopPc$ to force it to be symmetric positive semi-definite.
-
-To briefly summarize, we extract $\varphi_{x_i}$ for many points $x_i$ by applying $\Aop$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}). 
-%To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as distinct functions. 
-A greedy algorithm is used to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}).
-We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\phi_x$ (Section \ref{eq:mean_and_covariance_estimation}).
-The support of $\phi_x$ is estimated to be contained within an ellipsoid, so the process of choosing batches of points is an ellipse packing problem.
-For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
+Here we describe how we choose the sample points, $x_i$ (Section \ref{sec:sample_point_selection}), how we form the convolution kernels, $\varphi_i$ (Section \ref{sec:get_impulse_response}), and how we form the weighting functions, $w_i$ (Section \ref{sec:weighting_functions}), for the product-convolution approximation, $\AopPc \approx \Aop$, given in \eqref{eq:product_convolution}.
+We also describe how we convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), and (optionally) how we modify the resulting H-matrix, $\AmatPc$, to force it to be symmetric positive semi-definite (Section \ref{sec:make_spd}).
+The complete algorithm for constructing $\AmatPc$ is shown in Algorithm \ref{alg:construct_Atilde}.
+%The method for evaluating kernel entries of $\Phi_\text{pc}$ is shown in Section \ref{sec:eval_matrix_entries}.
+
 %\begin{itemize}
@@
 %\end{itemize}
-The complete algorithm for constructing $\AopPc$ is shown in Algorithm \ref{alg:construct_Atilde}.
+%The complete algorithm for constructing $\AopPc$ is shown in Algorithm \ref{alg:construct_Atilde}.
 The method for evaluating kernel entries of $\Phi_\text{pc}$ is shown in Section \ref{sec:eval_matrix_entries}.
@@
 \subsection{Sample point selection}
+\label{sec:sample_point_selection}
+
+We choose sample points, $x_i$, in batches.
+A greedy algorithm is used to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other.
+We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\phi_x$ (Section \ref{eq:mean_and_covariance_estimation}).
+The support of $\phi_x$ is estimated to be contained within an ellipsoid, so the process of choosing batches of points is an ellipsoid packing problem (Section \ref{sec:greedy_point_selection}).
+%In the next section (Section \ref{sec:get_impulse_response}), we will extract $\varphi_{x_i}$ for all the points $x_i$ in a given batch by applying $\Aop$ to the Dirac comb associated with that batch.
 
@@
 Then one can post-process $\phi_x$ to determine $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$.
-But this direct approach is computationally infeasible because and it requires applying $\Aop$ to one vector per point $x$ that we wish to evaluate $E_x$ at, and we will need to evaluate $E_x$ at a large number of points.
+But this direct approach is computationally infeasible because and it requires one operator action of $\Aop$ per point $x$ that we wish to evaluate $E_x$ at, and we will need to evaluate $E_x$ at a large number of points.
 Fortunately, it is possible to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$, \emph{for all points $x$ simultaneously}, by applying $\Aop^T$ to one constant function, $d$ linear functions, and $d(d+1)/2$ quadratic functions.
@@
 \end{equation*}
-Once $p$ is selected, we remove $p$ from $P$.
-If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in S_b$, and if $p$ is sufficiently far from the boundary, in the sense that $E_p \cap \partial \Omega = \{\}$, then we add $p$ to $S_b$.
+Once $p$ is selected, we remove $p$ from $P$. If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in S_b$,
+%and if $p$ is sufficiently far from the boundary, in the sense that $E_p \cap \partial \Omega = \{\}$, 
+then we add $p$ to $S_b$.
 Otherwise we discard $p$.
@@
 We use the IEEE standard value of ``NaN'' (not a number) for this placeholder.
-These NaN values are later replaced with more appropriate values, as we will discuss in Section \ref{sec:boundary_extension}.
+Later we replace these NaN values with more appropriate values (Section \ref{sec:boundary_extension}).
 
@@
 While $\varphi_{x_i}$ may be undefined at a point $z$, often there is a sample point $x_j$ that is near $x_i$, but further away from the boundary than $x_i$, such that $\varphi_{x_j}$ is well defined at $z$.
-To address boundary artifacts, we fill in missing/undefined portions of convolution kernels using information from other convolution kernels associated with the $k$ nearest sample points.
+To address boundary artifacts, we fill in missing/undefined portions of convolution kernels using information from other convolution kernels associated with the $k$ nearest sample points (we use $k=8$ in our numerical results).
 Kernels associated with nearer sample points are given precedence over kernels associated with farther away sample points.
@@
 Also, recall that in computations that the functions $\varphi_{x_i}$ are stored on regular rectilinear grids.
-To extend $\varphi_{x_i}$, first we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\varphi_{x_i}$, and contains that grid as a subgrid, but is large enough to contain the boxes associated with all of the neighbor kernels, $\varphi_{p_1}, \varphi_{p_2}, \dots, \varphi_{p_k}$.
+To extend $\varphi_{x_i}$, we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\varphi_{x_i}$, and contains that grid as a subgrid, but is large enough to contain the boxes associated with all of the neighbor kernels, $\varphi_{p_1}, \varphi_{p_2}, \dots, \varphi_{p_k}$.
 We initialize an extended convolution kernel, $\varphi_{x_i}^E$, to zero on this expanded grid.
-Then we interpolate the functions $\varphi_{p_j}$ onto this expanded grid one at a time in reverse order.
-That is, first we interpolate $\varphi_{p_k}$ onto the grid, then $\varphi_{p_{k-1}}$, and so on until we reach the original kernel $\varphi_{p_1} = \varphi_{x_i}$.
+Then we interpolate the functions $\varphi_{p_j}$ onto this expanded grid in reverse order.
+First we interpolate $\varphi_{p_k}$ onto the grid, then $\varphi_{p_{k-1}}$, and so on until the last step in which we interpolate the original kernel $\varphi_{p_1} = \varphi_{x_i}$.
 At each step, we overwrite existing values of $\varphi_{x_i}^E$ with new values from the kernel currently being interpolated, except when the new value of $\varphi_{p_j}$ is undefined (``NaN'').
@@
 \label{sec:weighting_functions}
+
+For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$.
+Constructing the weighting functions requires solving two Poisson PDEs per sample point $x_i$.
 
@@
 
-\subsection{Postprocessing the product-convolution approximation}
-
-\subsubsection{Hierarchical matrix construction}
+\subsection{Hierarchical matrix construction}
 \label{sec:H_matrix_conversion}
@@
 
-\subsubsection{Rational positive semi-definite modification}
+\subsection{Rational positive semi-definite modification}
 \label{sec:make_spd}

### 2021-08-03 paper/localpsf.tex
method section getting ok

@@
 		\Input{Linear operator $\Aop$, parameter $\text{max\_batches}$}
-		\Output{Dirac comb responses $\{\eta^b\}_{b=1}^{\text{max\_batches}}$, \\
-			weighting functions $\{w_i\}_{i=1}^r$}
-
-		Compute spatially varying mean, $\mu$, and covariance, $\Sigma$, using Algorithm \ref{alg:varhpi_mean_cov}.
+		\Output{H-matrix $\AmatPc$ (optionally, $\AmatPcSymPlus$)}
+
+		Compute $\alpha$, $\mu$, and $\Sigma$ using Algorithm \ref{alg:varhpi_mean_cov}.
 
@@
 
-			Compute $\eta^b$ by applying $\Aop$ to the dirac comb associated with $S_b$ (Section \ref{sec:get_impulse_response})
-
-			Compute weighting functions $w_i$ using Algorithm \ref{alg:weighting_functions_incremental}.
-In lines 2 and 3 of Algorithm \ref{alg:weighting_functions_incremental}, only solve for $\psi_i$ and $\theta_i$ associated with new points $x_i$ in $S_b$.
+			Compute $\eta^b$ by applying $\Aop$ to the dirac comb for $S_b$ (Section \ref{sec:get_impulse_response})
+
+		}
+
+		Form convolution kernels, $\varphi_{x_i}$, on regular grids (Section \ref{sec:convolution_kernel_regular_grid})
+
+		Modify convolution kernels to address boundary artifacts (Section \ref{sec:boundary_extension})
+
+		Compute all weighting functions $w_i$ using Algorithm \ref{alg:weighting_functions_incremental}
+%		In lines 2 and 3 of Algorithm \ref{alg:weighting_functions_incremental}, only solve for $\psi_i$ and $\theta_i$ associated with new points $x_i$ in $S_b$.
 The functions $\psi_i$ and $\theta_i$ associated with points in previous batches have already been computed in previous iterations of this for loop.
+
+		Form H-matrix $\AmatPc$ (Section \ref{sec:H_matrix_conversion})
+
+		(optional) Modify $\AmatPc$ to make $\AmatPcSymPlus$ (Section \ref{sec:make_spd})
+
 	}
-
-	}
-	\caption{Construct product-convolution approximation $\AopPc$}
+	\caption{Construct product-convolution H-matrix}
 	\label{alg:construct_Atilde}
@@
 The following results hold:
-	\begin{align*}
+	\begin{subequations}
+		\label{eq:vol_mean_var}
+	\begin{align}
 	\alpha =& \left(\Aop^T C\right)^* \\
@@
 	\Sigma^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \alpha - \mu^i\cdot \mu^j
-	\end{align*}
+	\end{align}
+	\end{subequations}
 	 for $i=1,\dots, d$, $j=1,\dots,d$.
@@
 
-		$\alpha = \left(\mathcal{A}^T C\right)^*$
+		Compute $\alpha = \left(\mathcal{A}^T C\right)^*$
 
@@
 
-			$\mu^i = \left(\mathcal{A}^T L^i\right)^* / \alpha$
+			Compute $\mu^i = \left(\mathcal{A}^T L^i\right)^* / \alpha$
 		}
@@
 			\For{$j=1,\dots,i$}{
-				Form quadratic function $Q^{ij} = x^i x^j$
-
-				$\Sigma^{ij} = \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot \mu^j$
-
-				$\Sigma^{ij} = \Sigma^{ji}$
+				Form quadratic function $Q^{ij}(x) = x^i x^j$
+
+				Compute $\Sigma^{ij} = \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot \mu^j$
+
+				Set $\Sigma^{ji} = \Sigma^{ij}$
 
@@
 
-Now that we have estimates of the supports of the functions $\phi_x$, we use these estimates, combined with a greedy ellipsoid packing algorithm, to choose sets of sample points, $S_b$.
-These sample points will be used in Section \ref{sec:get_impulse_response} to form Dirac combs, which are then used to compute impulse responses of $\mathcal{A}$.
+We use a greedy ellipsoid packing algorithm to choose sets of sample points, $S_b$, such that there is no overlap between the ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch.
+These sample point batches will be used in Section \ref{sec:get_impulse_response} to compute many impulse responses of $\Aop$ using only a small number of operator actions of $\Aop$.
+The support of $\phi_x$ is (approximately) contained in the ellipsoid $E_x$, so by applying $\mathcal{A}$ to the Dirac comb associated with all points in a batch, we will recover the impulse responses for all sample points in that batch.
+
 We start with a finite set of candidate points $P$.
@@
 \end{equation*}
+Candidate points for the first batch are chosen arbitrarily from $P$.
 Once $p$ is selected, we remove $p$ from $P$. If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in S_b$,
 %and if $p$ is sufficiently far from the boundary, in the sense that $E_p \cap \partial \Omega = \{\}$, 
-then we add $p$ to $S_b$.
-Otherwise we discard $p$.
-Recall that $E_p$ and $E_q$ are ellipsoids associated with points $p$ and $q$, respectively, as defined in \eqref{eq:support_ellipsoid}.
+then we add $p$ to $S_b$. Otherwise we discard $p$. 
+%Recall that $E_p$ and $E_q$ are ellipsoids associated with points $p$ and $q$, respectively, as defined in \eqref{eq:support_ellipsoid}. 
 This process repeats until there are no more points in $P$.
@@
 \subsubsection{Interpolation onto local regular grids}
+\label{sec:convolution_kernel_regular_grid}
 
@@
 
-For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$.
+%For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$.
 Constructing the weighting functions requires solving two Poisson PDEs per sample point $x_i$.

### 2021-08-03 paper/localpsf.tex
breaking things into further subsections
paper is ok

@@
 \begin{description}
-\item[Matrix-free:] One cannot easily access kernel entries $\Phi(y,x)$.
-Access to $\Aop$ is available only through application of $\Aop$ and $\Aop^T$ to arbitrary functions (i.e., operator actions).
-That is, evaluation of the maps
-\begin{equation*}
-	u \mapsto\Aop u \quad \text{and} \quad v \mapsto\Aop^Tv.
-\end{equation*} 
+\item[Matrix-free:] One cannot easily access kernel entries $\Phi(y,x)$. Access to $\Aop$ is available only through ``operator actions.'' That is, evaluation of the maps
+\begin{equation*}
+	u \mapsto\Aop u \quad \text{and} \quad v \mapsto\Aop^Tv,
+\end{equation*} 
+where $u$ and $v$ are arbitrary functions.
 \item[Non-negative kernel:] For all $(y,x) \in \Omega \times \Omega$, we have
@@
 The first property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\mathbf{A}$ and $\mathbf{A}^T$ with arbitrary vectors, but one cannot easily access matrix entries $\mathbf{A}_{ij}$.
-In the matrix-free regime, to compute $\mathbf{A}_{ij}$ one could compute $\mathbf{v} = \mathbf{A}\mathbf{e}_j$, then extract $\mathbf{A}_{ij} = \mathbf{v}_i$.
-However, this process is wasteful because one computes the entire vector $\mathbf{v}$, then discards all but one component.
+One could compute $\mathbf{v} = \mathbf{A}\mathbf{e}_j$, then extract $\mathbf{A}_{ij} = \mathbf{v}_i$, but this process is wasteful because one computes the entire vector $\mathbf{v}$, then discards all but one component.
 Here $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero.
 In general, $\mathbf{A}$ is matrix-free if $\mathbf{A}$ is defined implicitly via its action on vectors, and the process of applying $\mathbf{A}$ to a vector requires performing a nontrivial computational procedure, such as solving a large linear system, or timestepping.
-For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessians information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
+In reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
 
@@
 In this paper, rather than probing $\Aop^T$ with random noise functions, we use specific constant, linear, and quadratic functions.
-We observe that our method estimates the support of $\varphi_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
+Our method estimates the support of $\varphi_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
 
@@
 \label{sec:method}
-Here we describe how we choose the sample points, $x_i$ (Section \ref{sec:sample_point_selection}), how we form the convolution kernels, $\varphi_i$ (Section \ref{sec:get_impulse_response}), and how we form the weighting functions, $w_i$ (Section \ref{sec:weighting_functions}), for the product-convolution approximation, $\AopPc \approx \Aop$, given in \eqref{eq:product_convolution}.
+In this section we describe how we choose the sample points, $x_i$ (Section \ref{sec:sample_point_selection}), how we form the convolution kernels, $\varphi_i$ (Section \ref{sec:get_impulse_response}), and how we form the weighting functions, $w_i$ (Section \ref{sec:weighting_functions}), for the product-convolution approximation, $\AopPc \approx \Aop$, given in \eqref{eq:product_convolution}.
 We also describe how we convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), and (optionally) how we modify the resulting H-matrix, $\AmatPc$, to force it to be symmetric positive semi-definite (Section \ref{sec:make_spd}).
@@
 The quantity $\tau$ is a parameter that must be chosen by the user of the algorithm.
-The larger $\tau$ is, the larger the ellipsoid $E_x$ is, the more conservative the estimate is for the support of $\phi_x$.
+The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\phi_x$.
 However, the cost of our algorithm will depend on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
@@
 These sample point batches will be used in Section \ref{sec:get_impulse_response} to compute many impulse responses of $\Aop$ using only a small number of operator actions of $\Aop$.
-The support of $\phi_x$ is (approximately) contained in the ellipsoid $E_x$, so by applying $\mathcal{A}$ to the Dirac comb associated with all points in a batch, we will recover the impulse responses for all sample points in that batch.
+The support of $\phi_x$ is (approximately) contained in the ellipsoid $E_x$, so by applying $\mathcal{A}$ to the Dirac comb associated with all sample points in a batch, we will recover the impulse responses for all sample points in that batch.
 
@@
 
-We compute impulse responses of $\mathcal{A}$ in batches by applying $\mathcal{A}$ to a small number Dirac combs. Each Dirac comb yields one batch of impulse responses.
+We compute convolution kernels $\varphi_{x_i}$ in batches by applying $\Aop$ to a small number Dirac combs.
+Each application of $\Aop$ to a Dirac comb yields one batch of impulse responses, $\phi_{x_i}$, which yields one batch of convolution kernels, $\varphi_{x_i}$.
 %The batches of impulse responses, $\eta^b$, will ultimately be interpolated with weighting functions described in Section \ref{sec:weighting_functions} to form the desired approximation $\widetilde{\mathcal{A}}\approx \mathcal{A}$.
@@
 \label{eq:phi_b}
-	\eta^b = \left\langle\mathcal{A},\sum_{x_i \in S_b} \delta_{x_i}\right\rangle^* = \sum_{x_i \in S_b} \langle\mathcal{A}, \delta_{x_i}\rangle^* = \sum_{x_i \in S_b} \phi_x.
+	\eta^b = \left\langle\mathcal{A},\sum_{x_i \in S_b} \delta_{x_i}\right\rangle^* = \sum_{x_i \in S_b} \langle\mathcal{A}, \delta_{x_i}\rangle^* = \sum_{x_i \in S_b} \phi_{x_i}.
 \end{equation}
@@
 We use the IEEE standard value of ``NaN'' (not a number) for this placeholder.
-Later we replace these NaN values with more appropriate values (Section \ref{sec:boundary_extension}).
+In the next section (Section \ref{sec:boundary_extension}) we replace these NaN values with more appropriate values.
 
@@
 That is, $p_1=x_i$, $p_2$ is the nearest neighbor to $x_i$, $p_3$ is the second nearest neighbor to $x_i$, and so on.
-Also, recall that in computations that the functions $\varphi_{x_i}$ are stored on regular rectilinear grids.
+Also, recall from the previous section (Section \ref{sec:convolution_kernel_regular_grid}) that the functions $\varphi_{x_i}$ are stored on regular rectilinear grids.
 To extend $\varphi_{x_i}$, we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\varphi_{x_i}$, and contains that grid as a subgrid, but is large enough to contain the boxes associated with all of the neighbor kernels, $\varphi_{p_1}, \varphi_{p_2}, \dots, \varphi_{p_k}$.
@@
 
-\subsection{Hierarchical matrix construction}
+\subsection{Discretization and hierarchical matrix construction}
 \label{sec:H_matrix_conversion}
 
-In this section we discuss how we construct an H-matrix representation of our product-convolution approximation, after discretizing the problem with finite elements.
+Here we discuss how we construct an H-matrix representation of our product-convolution approximation, after discretizing the problem with finite elements.
+
+\subsubsection{Finite element discretization of the integral kernel}
+
 Let $f_1, f_2, \dots, f_N$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, and let $V_h := \Span\left(f_1, f_2, \dots, f_N\right)$ be the corresponding finite element space.
@@
 \end{equation}
+
+\subsubsection{Hierarchical matrix construction}
 
@@
 We estimate $\lambda_\text{min}$ and $\lambda_\text{max}$ using the implicitly restarted Lanczos method \cite{lanczos,scipy}.
+We describe how we solve system \eqref{eq:four_eq_system} in Appendix \ref{app:solve_rational_system}.
+Solving \eqref{eq:four_eq_system} reduces to solving a smooth one-dimensional optimization problem, which is computationally trivial.
 
@@
 We find that $\ratfct(\lambda_\text{min})=|\lambda_\text{min}|$ is an appropriate balance between these competing interests.
-
-Solving system \eqref{eq:four_eq_system} is computationally trivial.
-Using the definition of $\ratfct$ in \eqref{eq:rational_function_scalar}, and basic calculus and linear algebra, we may write Equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, and \eqref{eq:desired_property_4} in the following matrix form:
-\begin{equation}
-\label{eq:rational_3_linear_system}
-\begin{bmatrix} 1 & 0 & 1/\mu \\ 0 & 1 & -1/\mu^2 \\ 1 & \lambda_\text{max} & 1/(\lambda_\text{max}+\mu) \end{bmatrix} \begin{bmatrix} c_0 \\ c_1 \\ c_2 \end{bmatrix} = \begin{bmatrix} 0 \\ 0 \\ \lambda_\text{max} \end{bmatrix},
-\end{equation}
-and we may write Equation \eqref{eq:desired_property_3} as follows: 
-\begin{equation}
-\label{eq:rational_extra_eq}
-c_0 + c_1 \lambda_\text{min} + \frac{c_2}{\lambda_\text{min}+\mu} - |\lambda_\text{min}| = 0.
-\end{equation}
-Let 
-\begin{equation*}
-g(\mu) := c_0(\mu) + c_1(\mu) \lambda_\text{min} + \frac{c_2(\mu)}{\lambda_\text{min}+\mu} - |\lambda_\text{min}|
-\end{equation*}
-denote the left hand side of \eqref{eq:rational_extra_eq}, but with $c_0(\mu)$, $c_1(\mu)$, and $c_2(\mu)$ being the solution of \eqref{eq:rational_3_linear_system} for the given value of $\mu$.
-I.e., within $g(\mu)$ the quantities $c_0$, $c_1$, and $c_2$ are no longer free variables; they now depend on $\mu$ implicitly through the solution of a linear system.
-We solve the combined system, \eqref{eq:rational_3_linear_system} and $\eqref{eq:rational_extra_eq}$, by solving the one-dimensional nonlinear root finding problem
-\begin{equation}
-\label{eq:g_mu_zero}
-g(\mu)=0
-\end{equation}
-with bound $\mu > |\lambda_\text{min}|$ and initial guess $\mu_0 = 2 |\lambda_\text{min}|$.
-Each evaluation of $g$ requires solving the $3 \times 3$ linear system \eqref{eq:rational_3_linear_system}, and once the solution $\mu$ is found, we solve \eqref{eq:rational_3_linear_system} one more time to get the final values of $c_0$, $c_1$, and $c_2$.
 
@@
 Recursive algorithms can take advantage of the H-matrix low rank block structure to perform matrix arithmetic fast.
-Conventional dense matrix algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix-matrix addition require either $O(N^2)$ or $O(N^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(N \log^a N)$ time and memory for H-matrices.
+Conventional dense matrix algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix-matrix addition require either $O(N^2)$ or $O(N^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(N \log(N)^a)$ time and memory for H-matrices.
 Here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix.
@@
 This hierarchical partitioning of the point cloud into smaller and smaller point clouds corresponds to a hierarchical partitioning of the columns of the matrix into smaller and smaller \emph{clusters} of columns.
-This hierarchical partitioning of the columns forms a binary tree, which we call the column cluster tree.
+This hierarchical partitioning of the columns forms a binary tree, which is called the column cluster tree.
 The root of the tree is the set of all columns, and the leaves of the tree are clusters of columns that are not subdivided any further.
@@
 
-In the same way, the degrees of freedom associated with the rows of the matrix are hierarchically clustered into another cluster tree, and an ordering for the rows is generated.
+In the same way, the degrees of freedom associated with the rows of the matrix are hierarchically clustered into another cluster tree, and a depth-first search ordering for the rows is generated.
 In our examples, the degrees of freedom for the columns coincide with the degrees of freedom for the rows, so the cluster trees for the rows and columns are the same, but this is not required in general.
@@
 This subdivision process continues recursively until all blocks are either admissible, or smaller than the predetermined size mentioned above.
-The resulting hierarchical partitioning of matrix blocks forms a tree (quad-tree in $2$D, or oct-tree in $3$D), which is called the \emph{block cluster tree}.
+The resulting hierarchical partitioning of matrix blocks forms a tree, which is called the \emph{block cluster tree}.
 The root of the tree is the whole matrix, internal nodes in the tree are blocks that are subdivided, and the leaves of the tree are blocks that are either expected to be low rank, or are small.
@@
 \end{algorithm2e}
+
+\section{Solving for rational function scalars}
+\label{app:solve_rational_system}
+
+Using the definition of $\ratfct$ in \eqref{eq:rational_function_scalar}, and basic calculus and linear algebra, we may write Equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, and \eqref{eq:desired_property_4} in the following matrix form:
+\begin{equation}
+\label{eq:rational_3_linear_system}
+\begin{bmatrix} 1 & 0 & 1/\mu \\ 0 & 1 & -1/\mu^2 \\ 1 & \lambda_\text{max} & 1/(\lambda_\text{max}+\mu) \end{bmatrix} \begin{bmatrix} c_0 \\ c_1 \\ c_2 \end{bmatrix} = \begin{bmatrix} 0 \\ 0 \\ \lambda_\text{max} \end{bmatrix},
+\end{equation}
+and we may write Equation \eqref{eq:desired_property_3} as follows: 
+\begin{equation}
+\label{eq:rational_extra_eq}
+c_0 + c_1 \lambda_\text{min} + \frac{c_2}{\lambda_\text{min}+\mu} - |\lambda_\text{min}| = 0.
+\end{equation}
+Let 
+\begin{equation*}
+g(\mu) := c_0(\mu) + c_1(\mu) \lambda_\text{min} + \frac{c_2(\mu)}{\lambda_\text{min}+\mu} - |\lambda_\text{min}|
+\end{equation*}
+denote the left hand side of \eqref{eq:rational_extra_eq}, but with $c_0(\mu)$, $c_1(\mu)$, and $c_2(\mu)$ being the solution of \eqref{eq:rational_3_linear_system} for the given value of $\mu$.
+I.e., within $g(\mu)$ the quantities $c_0$, $c_1$, and $c_2$ are no longer free variables; they now depend on $\mu$ implicitly through the solution of a linear system.
+We solve the combined system, \eqref{eq:rational_3_linear_system} and $\eqref{eq:rational_extra_eq}$, by solving the one-dimensional nonlinear root finding problem
+\begin{equation}
+\label{eq:g_mu_zero}
+g(\mu)=0
+\end{equation}
+with bound $\mu > |\lambda_\text{min}|$ and initial guess $\mu_0 = 2 |\lambda_\text{min}|$.
+Each evaluation of $g$ requires solving the $3 \times 3$ linear system \eqref{eq:rational_3_linear_system}, and once the solution $\mu$ is found, we solve \eqref{eq:rational_3_linear_system} one more time to get the final values of $c_0$, $c_1$, and $c_2$.
 

### 2021-08-04 paper/localpsf.tex
notation in appendix fixes

@@
 
-Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\widetilde{\mathbf{A}}$ corresponds to the nodal point in $\mathbb{R}^d$ associated with the $i^\text{th}$ finite element basis function.
+Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\AkerPcMat$ corresponds to the nodal point in $\mathbb{R}^d$ associated with the $i^\text{th}$ finite element basis function.
 The columns of the matrix thus correspond to a point cloud in $\mathbb{R}^d$.
@@
 
-After reordering the rows and columns of $\widetilde{\mathbf{A}}$ via the depth-first ordering described above, we partition the reordered version of $\widetilde{\mathbf{A}}$ into a tree of $2 \times 2$ block matrices recursively.
+After reordering the rows and columns of $\AkerPcMat$ via the depth-first ordering described above, we partition the reordered version of $\AkerPcMat$ into a tree of $2 \times 2$ block matrices recursively.
 We use a geometric admissibility condition (discussed below) to decide which blocks to subdivide, and use the cluster trees for the rows and columns to determine how to subdivide those blocks.
 For the first stage of subdivision, let $r_1$ and $r_2$ be the children row clusters for the root of the row cluster tree, let $c_1$ and $c_2$ be the children column clusters for the root of the column cluster tree.
-The matrix $\widetilde{\mathbf{A}}$ is partitioned into blocks as follows:
+The matrix $\AkerPcMat$ is partitioned into blocks as follows:
 \begin{equation*}
 \begin{bmatrix}
-\widetilde{\mathbf{A}}_{11} & \widetilde{\mathbf{A}}_{12} \\
-\widetilde{\mathbf{A}}_{21} & \widetilde{\mathbf{A}}_{22},
+\left(\AkerPcMat\right)_{11} & \left(\AkerPcMat\right)_{12} \\
+\left(\AkerPcMat\right)_{21} & \left(\AkerPcMat\right)_{22},
 \end{bmatrix}
 \end{equation*}
-where $\widetilde{\mathbf{A}}_{11}$ denotes the block of $\widetilde{\mathbf{A}}$ with rows $r_1$ and columns $c_1$, $\widetilde{\mathbf{A}}_{12}$ denotes the block of $\widetilde{\mathbf{A}}$ with rows $r_1$ and columns $c_2$, and so on for $\widetilde{\mathbf{A}}_{21}$ and $\widetilde{\mathbf{A}}_{22}$.
-
-We now loop through the four blocks, $\widetilde{\mathbf{A}}_{11}$, $\widetilde{\mathbf{A}}_{12}$, $\widetilde{\mathbf{A}}_{21}$, and $\widetilde{\mathbf{A}}_{22}$, and decide which blocks should be subdivided further.
-For the purpose of explanation, consider $\widetilde{\mathbf{A}}_{12}$.
+where $\left(\AkerPcMat\right)_{11}$ denotes the block of $\AkerPcMat$ with rows $r_1$ and columns $c_1$, $\left(\AkerPcMat\right)_{12}$ denotes the block of $\AkerPcMat$ with rows $r_1$ and columns $c_2$, and so on for $\left(\AkerPcMat\right)_{21}$ and $\left(\AkerPcMat\right)_{22}$.
+
+We now loop through the four blocks, $\left(\AkerPcMat\right)_{11}$, $\left(\AkerPcMat\right)_{12}$, $\left(\AkerPcMat\right)_{21}$, and $\left(\AkerPcMat\right)_{22}$, and decide which blocks should be subdivided further.
+For the purpose of explanation, consider $\left(\AkerPcMat\right)_{12}$.
 If
@@
 \end{equation}
-then we mark $\widetilde{\mathbf{A}}_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone.
+then we mark $\left(\AkerPcMat\right)_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone.
 Here $\dist\left(r_1, c_2\right)$ is the Euclidean distance between the axis-aligned bounding box for the point cloud associated with the row cluster $r_1$, and the axis aligned bounding box for the point cloud associated with the column cluster $c_2$.
@@
 This process is repeated for the other blocks to determine which blocks are admissible and which are not.
-For us, the diagonal blocks $\widetilde{\mathbf{A}}_{11}$ and $\widetilde{\mathbf{A}}_{22}$ are not admissible because the distance between a point cloud and itself is zero.
-
-Next, we subdivide all blocks that are not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\mathbf{\widetilde{A}}$.
+For us, the diagonal blocks $\left(\AkerPcMat\right)_{11}$ and $\left(\AkerPcMat\right)_{22}$ are not admissible because the distance between a point cloud and itself is zero.
+
+Next, we subdivide all blocks that are not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\AkerPcMat$.
 But now we subdivide a block based on the two child row clusters and two child column clusters for the rows and columns of that block.
@@
 Once the block cluster tree has been constructed, low rank approximations of the admissible (low rank) blocks are formed using the adaptive cross method \cite{ACA}.
-Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\widetilde{\mathbf{A}}$.
+Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\AkerPcMat$.
 The idea of the adaptive cross method is to form a low rank approximation of $X$ by sampling a small number of rows and columns of $X$.
@@
 
-We use the adaptive cross method to form low rank approximations for each admissible block of $\widetilde{\mathbf{A}}$.
-We directly compute all entries of the small dense blocks of $\widetilde{\mathbf{A}}$ that are not admissible.
-This process requires us to compute $O(r^2 N \log N)$ entries of $\widetilde{\mathbf{A}}$, which is relatively cheap compared to the matrix-vector products with $\mathbf{A}$ that are required to form the product-convolution approximation $\widetilde{\mathbf{A}}$.
+We use the adaptive cross method to form low rank approximations for each admissible block of $\AkerPcMat$.
+We directly compute all entries of the small dense blocks of $\AkerPcMat$ that are not admissible.
+This process requires us to compute $O(r^2 N \log N)$ entries of $\AkerPcMat$, which is relatively cheap compared to the operator actions of $\Aop$ that are required to form the product-convolution approximation.
 

### 2021-08-23 paper/localpsf.tex
edits to Section 2 based on meeting with Noemi

@@
 This yields a substantial overall speedup because the number of gridpoints in the regular grid for $\varphi_{x_i}$ is typically much smaller than the number of evaluations of $\varphi_{x_i}$ required to convert $\AopPc$ to H-matrix format.
+DISCUSS MORE ABOUT COMPUTATIONAL COSTS/SAVINGS
 %Having $\varphi_{x_i}$ stored on a regular grid also simplifies the implementation of the method that we use to deal with boundary artifacts, which will be described in Section \ref{sec:boundary_extension}.
@@
 Also, recall from the previous section (Section \ref{sec:convolution_kernel_regular_grid}) that the functions $\varphi_{x_i}$ are stored on regular rectilinear grids.
-To extend $\varphi_{x_i}$, we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\varphi_{x_i}$, and contains that grid as a subgrid, but is large enough to contain the boxes associated with all of the neighbor kernels, $\varphi_{p_1}, \varphi_{p_2}, \dots, \varphi_{p_k}$.
+To extend $\varphi_{x_i}$, we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\varphi_{x_i}$, and contains that grid as a subgrid, but is large enough to contain the boxes associated with all of the neighbor kernels, $\varphi_{p_1}, \varphi_{p_2}, \dots, \varphi_{p_k}$ (recall that the functions $\varphi_{x_i}$ are all centered at zero).
 We initialize an extended convolution kernel, $\varphi_{x_i}^E$, to zero on this expanded grid.
@@
 \label{eq:kernel_representation_pcmesh}
-(\AopPcMesh u)(v) := \int_\Omega \int_\Omega v(y) \AkerPcMesh(y,x) u(x) dx dy,
+(\AopPcMesh u)(v) := \int_\Omega \int_\Omega v(y) \AkerPcMesh(y,x) u(x) dx dy.
 \end{equation}
@@
 \end{equation}
-and similar for $v_h$. Further, let $\mathbf{M} \in \mathbb{R}^{N \times N}$ be the finite element mass matrix, which has entries
-\begin{equation}
-\label{eq:mass_matrix_defn}
-\mathbf{M}_{ij} := \int_\Omega f_i(x) f_j(x) dx,
-\end{equation}
+and similar for $v_h$. Further, let $\mathbf{M} \in \mathbb{R}^{N \times N}$ be the finite element mass matrix,
+% which has entries
+%\begin{equation}
+%\label{eq:mass_matrix_defn}
+%\mathbf{M}_{ij} := \int_\Omega f_i(x) f_j(x) dx,
+%\end{equation}
 and let $\AkerPcMat \in \mathbb{R}^{N \times N}$ be following matrix of approximate kernel entries:
@@
 \end{equation*}
+% PUT INTO APPENDIX!!!!
 %\begin{align*}
@@
 
-We form an H-matrix representation of $\AmatPc$ by converting $\AkerPcMat$ and $\mathbf{M}$ to H-matrix format, then using fast H-matrix methods to multiply these matrices according \eqref{eq:Amatpc_defn}.
-Once $\AmatPc$ is in H-matrix format, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods.
+We form an H-matrix representation of $\AmatPc$ by converting $\AkerPcMat$ and $\mathbf{M}$ to H-matrix format, then using fast H-matrix methods to multiply these matrices according to \eqref{eq:Amatpc_defn}.
+Once $\AmatPc$ is in H-matrix format, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods \cite{BORMHACKBUSCHBOOK}.
 We use H1 matrices in our numerical results, but any of the other H-matrix formats (such as H2, HODLR, HSS, HBS, and others \cite{HMATRIX}) could be used instead.
 For more details on H-matrices, we recommend \cite{HMATRIXGOOD}.
-We discuss H-matrices in greater detail in Appendix \ref{app:h_matrix}.
+%We discuss H-matrices in greater detail in Appendix \ref{app:h_matrix}.
 
 We convert $\AkerPcMat$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBPro software package \cite{HLIBPRO}.
-Briefly, this method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix (here, these locations are the finite element Lagrange nodes $p_i$, $i=1,\dots,N$) to construct a recursive block partitioning of $\AkerPcMat$, such that all blocks at the finest level of the partitioning are either expected to be low rank, or are small.
-Then, the method forms low rank approximations of the blocks that are expected to be low rank by sampling specially chosen rows and columns of those blocks.
-The small high rank blocks are formed directly by evaluating their entries.
+Although $\AkerPcMat$ is a dense $N \times N$ matrix, constructing the H-matrix approximation of $\AkerPcMat$ only requires evaluation of $O(k^2 N \log N)$ entries of $\AkerPc$, and these entries are computed via the formula in Equation \eqref{eq:eval_y_x}.
+Here $k$ is the rank of the highest-rank block in the H-matrix.
+A dense representation of $\AkerPcMat$ is never formed.
 We describe this H-matrix construction process in detail in Appendix \ref{app:h_matrix}.
-Although $\AkerPcMat$ is a dense $N \times N$ matrix, constructing the H-matrix approximation of $\AkerPcMat$ only requires evaluation of $O(N \log N)$ entries of $\AkerPc$, and these entries are computed via the formula in Equation \eqref{eq:eval_y_x}.
-A dense representation of $\AkerPcMat$ is never formed.
 We convert $\mathbf{M}$ to H-matrix format using standard H-matrix methods for sparse matrices implemented within HLIBPro, using the same recursive block partitioning structure as was used for $\AkerPcMat$.
@@
 
-The H-matrix construction process that we use to form the H-matrix representation of $\AkerPcMat$ proceeds as follows.
+%The H-matrix construction method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix (here, these locations are the finite element Lagrange nodes $p_i$, $i=1,\dots,N$) to construct a recursive block partitioning of $\AkerPcMat$, such that all blocks at the finest level of the partitioning are either expected to be low rank, or are small.
+Then, the method forms low rank approximations of the blocks that are expected to be low rank by sampling specially chosen rows and columns of those blocks.
+The small high rank blocks are formed directly by evaluating their entries.
+
+In detail, the process of constructing an H-matrix representation of $\AkerPcMat$ proceeds as follows.
 First, we construct hierarchical partitionings of the degrees of freedom for the columns and rows of the matrix (cluster trees, Section \ref{sec:cluster_trees}).
@@
 
-Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\AkerPcMat$ corresponds to the nodal point in $\mathbb{R}^d$ associated with the $i^\text{th}$ finite element basis function.
+Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\AkerPcMat$ corresponds to the Lagrange node in $\mathbb{R}^d$ associated with the $i^\text{th}$ finite element basis function.
 The columns of the matrix thus correspond to a point cloud in $\mathbb{R}^d$.
@@
 \begin{equation*}
-	\Sigma_p \Phi = \Sigma_q \Phi \Lambda,
-\end{equation*}
-where $\Phi \in \mathbb{R}^{d \times d}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_d)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$.
-The matrix $\Phi$ simultaneously diagonalizes $\Sigma_p$ and $\Sigma_q$, in the sense that $\Phi^T\Sigma_p\Phi = \Lambda$, and $\Phi^T\Sigma_q\Phi = I$, where $I$ is the $d \times d$ identity matrix.
+	\Sigma_p P = \Sigma_q P \Lambda,
+\end{equation*}
+where $P \in \mathbb{R}^{d \times d}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_d)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$.
+The matrix $P$ simultaneously diagonalizes $\Sigma_p$ and $\Sigma_q$, in the sense that $P^T\Sigma_p P = \Lambda$, and $P^T\Sigma_q P = I$, where $I$ is the $d \times d$ identity matrix.
 Using this diagonalization, and some algebraic manipulations, we may write $K(s)$ as

### 2021-08-24 paper/localpsf.tex
edits to regular grid transfer section based on discussionn yesterday with Noemi. Add further dictussion about why transferring to regular grid reduces costs

@@
   pdftitle={Fast matrix-free approximation of smoothly varying blur operators, with application to Hessians in PDE-constrained inverse problems with highly informative data},
-  pdfauthor={N. Alger, N. Petra, and O. Ghattas}
+  pdfauthor={N. Alger, N. Petra, T. Hartland, and O. Ghattas}
 }
@@
 In computations, $\eta^b$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh.
-Evaluating $\eta_b(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell.
+Evaluating $\eta^b(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell.
 Although this procedure is asymptotically scalable, it is slow in practice \cite{FEMEVALSLOW}.
@@
 This is an upfront cost which is done ``offline.'' After $\varphi_{x_i}$ has been transferred to the regular grid, we use fast regular grid interpolation to perform all subsequent ``online'' evaluations of $\varphi_{x_i}$.
-This yields a substantial overall speedup because the number of gridpoints in the regular grid for $\varphi_{x_i}$ is typically much smaller than the number of evaluations of $\varphi_{x_i}$ required to convert $\AopPc$ to H-matrix format.
-DISCUSS MORE ABOUT COMPUTATIONAL COSTS/SAVINGS
-%Having $\varphi_{x_i}$ stored on a regular grid also simplifies the implementation of the method that we use to deal with boundary artifacts, which will be described in Section \ref{sec:boundary_extension}.
+This yields a substantial overall speedup because the number of finite element function evaluations required to transfer the functions $\varphi_{x_i}$ to the regular grid is $O(N)$, while the number of kernel evaluations required to convert $\AopPc$ to H-matrix format is $O(k N \log N)$.
+Here $N$ is the number of finite element degrees of freedom, and $k$ is the H-matrix rank.
+Furthermore, transferring $\varphi_{x_i}$ to a regular grid requires evaluating $\varphi_{x_i}$ at a predetermined and highly structured collection of points.
+This structure allows for faster determination of which mesh cells that points are in, and allows one to evaluate the finite element function at all regular gridpoints in parallel.
+In contrast, constructing the H-matrix requires evaluating $\varphi_{x_i}$ at an unpredictable sequence of scattered of points, which is an expensive task for finite element function evaluation, but a cheap task for regular grid interpolation.
+Having $\varphi_{x_i}$ stored on a regular grid also simplifies the implementation of the method that we use to deal with boundary artifacts, which will be described in Section \ref{sec:boundary_extension}.
 

### 2021-08-30 paper/localpsf.tex
some reorganization of intro

@@
 We present a fast matrix free method for approximating locally translation-invariant operators, $\Aop$, that have locally supported non-negative integral kernels.
-Such operators arise, for example, in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, as Poincare-Steklov operators, as covariance operators, as blurring operators in imaging problems, and as Hessians in distributed parameter PDE-constrained inverse problems.
-
-Let $\Omega \subset \mathbb{R}^d$ be a bounded domain in dimension $d=1$, $2$, or $3$, and let $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the form
+Such operators arise throughout computational science.
+Examples include Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, Poincare-Steklov operators (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems.
+
+Let $\Omega \subset \mathbb{R}^d$ be a bounded domain in dimension $d=1$, $2$, or $3$. We consider integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ of the form
 \begin{equation}
@@
 \end{equation}
-with integral kernel $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$.
+where $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel.
 Here $\Aop u \in L^2(\Omega)'$ is the linear functional that results from applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
-We focus on operators of this form which have the following properties:
+We further narrow our focus to operators which possess the following four properties:
 \begin{description}
-\item[Matrix-free:] One cannot easily access kernel entries $\Phi(y,x)$. Access to $\Aop$ is available only through ``operator actions.'' That is, evaluation of the maps
-\begin{equation*}
-	u \mapsto\Aop u \quad \text{and} \quad v \mapsto\Aop^Tv,
-\end{equation*} 
-where $u$ and $v$ are arbitrary functions.
-\item[Non-negative kernel:] For all $(y,x) \in \Omega \times \Omega$, we have
-\begin{equation*}
-	\Phi(y,x) \ge 0.
-\end{equation*}
-\item[Local support:] Let $\phi_x$ be the following \emph{impulse response}:
+\item[1) Matrix-free:] One cannot easily evaluate kernel entries $\Phi(y,x)$.
+Instead, access to $\Aop$ is available through ``operator actions.'' That is, we have a black-box computational procedure (which is typically expensive) through which we may evaluate the maps
+\begin{equation*}
+	u \mapsto\Aop u \quad \text{and} \quad v \mapsto\Aop^Tv
+\end{equation*} 
+for arbitrary functions $u$ and $v$.
+After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\mathbf{A}$, that is typically too large to be built and stored.
+This property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\mathbf{A}$ and $\mathbf{A}^T$ with arbitrary vectors, but one cannot easily access matrix entries $\mathbf{A}_{ij}$.
+One could compute $\mathbf{v} = \mathbf{A}\mathbf{e}_j$, where $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero, then extract $\mathbf{A}_{ij} = \mathbf{v}_i$.
+However, this process is wasteful because one computes the entire vector $\mathbf{v}$, then discards all but one component.
+In applications, applying $\mathbf{A}$ to a vector may involve iteratively solving a large linear system, or timestepping, or performing some other nontrivial computational procedure.
+For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
+\item[2) Local support:] Let $\phi_x$ be the following \emph{impulse response}:
 \begin{equation}
@@
 The smaller these supports are, the better our algorithm will perform.
-\item[Local translation invariance:] If $h$ is not too large, then
-\begin{equation}
-	\label{eq:local_translation_invariance}
-	\Phi(y+h, x+h) \approx \Phi(y,x).
-\end{equation}
-\end{description}
-
-After discretization (see Appendix \ref{app:discretized_operations}), the operator $\Aop$ becomes a dense matrix, $\mathbf{A}$, that is typically too large to be built and stored.
-The first property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\mathbf{A}$ and $\mathbf{A}^T$ with arbitrary vectors, but one cannot easily access matrix entries $\mathbf{A}_{ij}$.
-One could compute $\mathbf{v} = \mathbf{A}\mathbf{e}_j$, then extract $\mathbf{A}_{ij} = \mathbf{v}_i$, but this process is wasteful because one computes the entire vector $\mathbf{v}$, then discards all but one component.
-Here $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero.
-In general, $\mathbf{A}$ is matrix-free if $\mathbf{A}$ is defined implicitly via its action on vectors, and the process of applying $\mathbf{A}$ to a vector requires performing a nontrivial computational procedure, such as solving a large linear system, or timestepping.
-In reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
-
 Let $\delta_x$ denote the point source (delta distribution) centered at $x$.
@@
 After discretization, the process of applying $\Aop$ to the distribution $\delta_x$, then forming the Riesz representative of the resulting linear functional, amounts to computing $\boldsymbol{\phi}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x$, where $\boldsymbol{\phi}_x$, $\mathbf{A}$, and $\boldsymbol{\delta}_x$ are discretized versions of $\phi_x$, $A$ and $\delta_x$, respectively, and $\mathbf{M}$ is a discretization of the $L^2$ inner product bilinear form (e.g., a mass matrix in the finite element context).
-From \eqref{eq:impulse_response_delta_action}, we see that the local support property means that the response of $\Aop$ to a point source at $x$ is zero (or small) at points $y$ that are far from $x$.
-
-Local translation invariance is illustrated in Figure FIG.
-If $\Aop$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then it is straightforward to show that $\Aop$ would be a convolution operator, and the convolution kernel, $\varphi_x$, would be given by
+From \eqref{eq:impulse_response_delta_action}, we see that this local support property means that the response of $\Aop$ to a point source at $x$ is zero (or small) at points $y$ that are far from $x$.
+\item[3) Local translation invariance:] We say that $\Aop$ is locally translation invariant if
+\begin{equation}
+	\label{eq:local_translation_invariance}
+	\Phi(y+h, x+h) \approx \Phi(y,x)
+\end{equation}
+when $h$ is not too large.
+One can imagine the kernel entry $\Phi(y,x)$ as representing the strength of an interaction between a source at point $x$ and a target at point $y$.
+Local translation invariance means that the interaction strength remains roughly the same if the source and target points are both translated by the same vector.
+This is illustrated in Figure FIG.
+\item[4) Non-negative kernel:] For all $(y,x) \in \Omega \times \Omega$, we have
+\begin{equation*}
+\Phi(y,x) \ge 0.
+\end{equation*}
+\end{description}
+
+In this paper we present a computationally efficient method to form product-convolution approximations of operators possessing these four properties.
+Once the product-convolution approximation is formed, we convert it to hierarchical matrix (H-matrix) format, and use fast H-matrix methods to perform further useful linear algebra operations.
+
+\subsection{Motivation: PDE constrained inverse problems with highly informative data}
+\label{sec:PDE_hessian_motivation}
+
+Our motivation for this work is approximation of Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs).
+That is, inverse problems in which one seeks to reconstruct an unknown parameter field, $m$, from noisy observations,
+\begin{equation*}
+y=f(m,u(m)) + \text{noise},
+\end{equation*}
+which depend on  a state variable $u$. In turn, $u$ depends on $m$ implicitly through the solution of a PDE, which we write generically as follows:
+\begin{equation}
+\label{eq:state_pde}
+0=g(m,u).
+\end{equation}
+Here $u(m)$ denotes the solution of the PDE \eqref{eq:state_pde} as a function of $m$.
+In the deterministic approach to inverse problems, one typically finds $m$ as the solution to the minimization problem
+\begin{equation}
+\label{eq:minimization_problem}
+\min_m \quad \frac{1}{2}\|y - f(m,u(m))\|_W^2 + R(m),
+\end{equation}
+where $\|\cdot\|_W$ is weighted norm which depends on the noise covariance, and $R(m)$ is a regularization term.
+Approximation of the Hessian of the objective function, $\mathcal{H}$, allows for fast solution of \eqref{eq:minimization_problem} via Newton-type methods.
+Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because $\mathcal{H}^{-1}$ locally approximates the Bayesian posterior covariance for $m$.
+%We only have access to $\mathcal{H}$ via its action on an arbitrary vectors $v$, i.e., evaluation of the map $v \mapsto \mathcal{H} v$.
+Evaluating $v \mapsto \mathcal{H} v$ is an expensive process that involves solving two auxiliary linear PDEs.
+
+The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term \cite{CCGOPAPERS}.
+Either the Lanczos method or the randomized singular value decomposition \cite{HMTRANDOM} are used to perform the low rank approximation using only matrix-vector products.
+These methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of matrix-vector products are required to form the aforementioned low rank approximation.
+The ideal scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost) \cite{MYDISSERTATION}.
+Although the method we present in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest.
+For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high-rank approximations of an operator using a small number of matrix-vector products.
+
+\subsection{Product-convolution approximation}
+
+If $\Aop$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then it is straightforward to show that $\Aop$ would be the convolution operator $u \mapsto \left(\varphi_x \ast u\right)^*$, where the convolution kernel is given by
 \begin{equation}
@@
 In this case, we modify the H-matrix representation of $\AopPc$ to make it positive definite, by symmetrizing it, then applying a specially chosen rational matrix function to the symmetrized H-matrix.
-
-\subsection{Motivation: PDE constrained inverse problems with highly informative data}
-\label{sec:PDE_hessian_motivation}
-
-Our motivation for this work is approximation of Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs).
-That is, inverse problems in which one seeks to reconstruct an unknown parameter field, $m$, from noisy observations,
-\begin{equation*}
-y=f(m,u(m)) + \text{noise},
-\end{equation*}
-which depend on  a state variable $u$. In turn, $u$ depends on $m$ implicitly through the solution of a PDE, 
-\begin{equation}
-\label{eq:state_pde}
-0=g(m,u).
-\end{equation}
-Here $u(m)$ denotes the solution of the PDE \eqref{eq:state_pde} as a function of $m$.
-In the deterministic approach to inverse problems, one typically finds $m$ as the solution to the minimization problem
-\begin{equation}
-\label{eq:minimization_problem}
-\min_m \quad \frac{1}{2}\|y - f(m,u(m))\|_W^2 + R(m),
-\end{equation}
-where $\|\cdot\|_W$ is weighted norm which depends on the noise covariance, and $R(m)$ is a regularization term.
-Approximation of the Hessian of the objective function, $\mathcal{H}$, allows for fast solution of \eqref{eq:minimization_problem} via Newton-type methods.
-Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because $\mathcal{H}^{-1}$ locally approximates the Bayesian posterior covariance for $m$.
-%We only have access to $\mathcal{H}$ via its action on an arbitrary vectors $v$, i.e., evaluation of the map $v \mapsto \mathcal{H} v$.
-Evaluating $v \mapsto \mathcal{H} v$ is an expensive process that involves solving two auxiliary linear PDEs.
-
-The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term \cite{CCGOPAPERS}.
-Either the Lanczos method or the randomized singular value decomposition \cite{HMTRANDOM} are used to perform the low rank approximation using only matrix-vector products.
-These methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of matrix-vector products are required to form the aforementioned low rank approximation.
-The ideal scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost) \cite{MYDISSERTATION}.
-Although the method we present in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest.
-For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because high-rank approximations of an operator can be formed using a small number of matrix-vector products.
 
@@
 \end{equation}
-Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, \eqref{eq:mass_matrix_defn}, and \eqref{eq:Akerpcmat_entries}, and performing algebraic manipulations, it is straightforward but tedious to show that
+Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, and \eqref{eq:Akerpcmat_entries}, the definition of the mass matrix, and performing algebraic manipulations, it is straightforward but tedious to show that
 \begin{equation*}
@@
 \begin{proof}
-	By the Rellich-Kondrachov theorem (see, e.g., Theorem 7.20 in \cite{ARBOGAST}), $H^2_N(\Omega)$ is compactly embedded in $C^0(\Omega)$. WORK IN PROGRESS
+	By the Sobolev Imbedding Theorem (see, e.g., Theorem 7.20 in \cite{ARBOGAST}), $H^2_N(\Omega)$ is embedded in $C^0(\Omega)$. WORK IN PROGRESS
 \end{proof}

### 2021-08-30 paper/localpsf.tex
some reorganization of intro

@@
 However, this process is wasteful because one computes the entire vector $\mathbf{v}$, then discards all but one component.
-In applications, applying $\mathbf{A}$ to a vector may involve iteratively solving a large linear system, or timestepping, or performing some other nontrivial computational procedure.
+In applications, applying $\mathbf{A}$ to a vector may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
 For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
@@
 
-In this paper we present a computationally efficient method to form product-convolution approximations of operators possessing these four properties.
-Once the product-convolution approximation is formed, we convert it to hierarchical matrix (H-matrix) format, and use fast H-matrix methods to perform further useful linear algebra operations.
+Our overall strategy is to form a product-convolution approximation, $\AopPc$, of $\Aop$ using only operator actions, then convert $\AopPc$ into an H-matrix, which we denote by $\AmatPc$.
+Schematically,
+\begin{equation*}
+\Aop \longrightarrow \underbrace{\AopPc}_{\substack{\text{product}\\ \text{convolution}}} \longrightarrow \underbrace{\AmatPc}_{\text{H-matrix}}.
+\end{equation*}
+We will define product-convolution approximations and H-matrices, and discuss them in greater detail, throughout the paper.
+The novel contributions of our method are in the way that we form the product-convolution approximation ($\Aop \rightarrow \AopPc$).
+This is the part of the method where we use the four properties of $\Aop$ described above.
+Fast conversion of $\AopPc$ to H-matrix format ($\AopPc \rightarrow \AmatPc$) is possible because integral kernel entries for $\AopPc$ are easily accessible (unlike integral kernel entries for $\Aop$).
+Once in H-matrix format, we use fast H-matrix methods to perform further linear algebra operations.
 
@@
 These methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of matrix-vector products are required to form the aforementioned low rank approximation.
-The ideal scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost) \cite{MYDISSERTATION}.
+The ideal scenario from a scientific perspective (highly informative data) is therefore the worst case scenario from a computational perspective (large computational cost) \cite{MYDISSERTATION}.
 Although the method we present in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest.
@@
 \subsection{Product-convolution approximation}
+\label{sec:prodconv_intro}
 

### 2021-08-30 paper/localpsf.tex
some reorganization of intro

@@
 
-Our overall strategy is to form a product-convolution approximation, $\AopPc$, of $\Aop$ using only operator actions, then convert $\AopPc$ into an H-matrix, which we denote by $\AmatPc$.
+Our overall strategy is to form a product-convolution approximation, $\AopPc$, of $\Aop$ using only operator actions, then convert $\AopPc$ into a hierarchical matrix (H-matrix), which we denote by $\AmatPc$.
 Schematically,
@@
 \end{equation*}
-We will define product-convolution approximations and H-matrices, and discuss them in greater detail, throughout the paper.
+We will define product-convolution approximations in Section \ref{sec:prodconv_intro}.
+H-matrices are discussed in Appendix \ref{app:h_matrix}.
+Our method is summarized in Section \ref{sec:overview_intro}, and described in detail in Section \ref{sec:method}.
+
 The novel contributions of our method are in the way that we form the product-convolution approximation ($\Aop \rightarrow \AopPc$).
@@
 Fast conversion of $\AopPc$ to H-matrix format ($\AopPc \rightarrow \AmatPc$) is possible because integral kernel entries for $\AopPc$ are easily accessible (unlike integral kernel entries for $\Aop$).
-Once in H-matrix format, we use fast H-matrix methods to perform further linear algebra operations.
+Once in H-matrix format, one may use H-matrix methods to perform further useful linear algebra operations involving $\AmatPc$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
+These H-matrix methods are fast and scale well to large problems.
 

### 2021-08-31 paper/localpsf.tex
edits to intro

@@
   We present an efficient matrix-free method for approximating locally translation invariant operators that have locally supported non-negative integral kernels.
-The idea of the method is to compute impulse responses of the operator at a collection of scattered points, then interpolate these computed impulse responses to approximate impulse responses at other arbitrary points.
-This results in a product-convolution approximation, which we convert to hierarchical matrix format in order to solve linear systems and perform other matrix operations efficiently using fast hierarchical matrix arithmetic.
+Such operators arise, for example, as Schur complements in Schur complements methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems.
+The method computes impulse responses of the operator at a collection of scattered points, then interpolates these computed impulse responses to form a product-convolution approximation of the operator.
+The product-convolution approximation is converted to hierarchical matrix format in order to solve linear systems and perform other matrix operations efficiently.
 Impulse responses are computed by applying the operator to a small number of Dirac comb ``batches'' of point sources.
-A key innovation of our method is a matrix-free procedure for choosing as many points as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
+The key innovation of our method is a matrix-free procedure for choosing as many points as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
 We apply the method to approximate Hessians in large-scale PDE-constrained inverse problem with highly informative data.
@@
 Such operators arise throughout computational science.
-Examples include Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, Poincare-Steklov operators (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems.
+Examples include Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems.
 
@@
 After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\mathbf{A}$, that is typically too large to be built and stored.
-This property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\mathbf{A}$ and $\mathbf{A}^T$ with arbitrary vectors, but one cannot easily access matrix entries $\mathbf{A}_{ij}$.
-One could compute $\mathbf{v} = \mathbf{A}\mathbf{e}_j$, where $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero, then extract $\mathbf{A}_{ij} = \mathbf{v}_i$.
-However, this process is wasteful because one computes the entire vector $\mathbf{v}$, then discards all but one component.
+This property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\mathbf{A}$ and $\mathbf{A}^T$ with arbitrary vectors, but one cannot easily\footnote{One could compute $\mathbf{v} = \mathbf{A}\mathbf{e}_j$, where $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero, then extract $\mathbf{A}_{ij} = \mathbf{v}_i$.
+However, this process is wasteful because one computes the entire vector $\mathbf{v}$, then discards all but one component.} access matrix entries $\mathbf{A}_{ij}$.
 In applications, applying $\mathbf{A}$ to a vector may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
@@
 %We take the Riesz representative of the linear functional $\left\langle \mathcal{A}, \delta_x \right\rangle \in L^2(\Omega)'$ to convert it into a function $\left\langle \mathcal{A}, \delta_x \right\rangle^* \in L^2(\Omega)$.
-After discretization, the process of applying $\Aop$ to the distribution $\delta_x$, then forming the Riesz representative of the resulting linear functional, amounts to computing $\boldsymbol{\phi}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x$, where $\boldsymbol{\phi}_x$, $\mathbf{A}$, and $\boldsymbol{\delta}_x$ are discretized versions of $\phi_x$, $A$ and $\delta_x$, respectively, and $\mathbf{M}$ is a discretization of the $L^2$ inner product bilinear form (e.g., a mass matrix in the finite element context).
+%After discretization, the process of applying $\Aop$ to the distribution $\delta_x$, then forming the Riesz representative of the resulting linear functional, amounts to computing $\boldsymbol{\phi}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x$, where $\boldsymbol{\phi}_x$, $\mathbf{A}$, and $\boldsymbol{\delta}_x$ are discretized versions of $\phi_x$, $A$ and $\delta_x$, respectively, and $\mathbf{M}$ is a discretization of the $L^2$ inner product bilinear form (e.g., a mass matrix in the finite element context).
 From \eqref{eq:impulse_response_delta_action}, we see that this local support property means that the response of $\Aop$ to a point source at $x$ is zero (or small) at points $y$ that are far from $x$.
@@
 
-\subsection{Motivation: PDE constrained inverse problems with highly informative data}
+\subsection{Motivation: inverse problems with highly informative data}
 \label{sec:PDE_hessian_motivation}
@@
 where $\|\cdot\|_W$ is weighted norm which depends on the noise covariance, and $R(m)$ is a regularization term.
-Approximation of the Hessian of the objective function, $\mathcal{H}$, allows for fast solution of \eqref{eq:minimization_problem} via Newton-type methods.
-Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because $\mathcal{H}^{-1}$ locally approximates the Bayesian posterior covariance for $m$.
+The objective function in optimization problem \eqref{eq:minimization_problem} is defined indirectly via the implicit function theorem.
+As a result, the Hessian of the objective function, $\mathcal{H}$, is only accessible matrix-free.
+In particular, applying $\mathcal{H}$ to a vector requires solving two PDEs \cite{HESSIANACTION}.
+More accessible approximations of $\mathcal{H}$ are highly desirable, because they allow for fast solution of \eqref{eq:minimization_problem} via Newton-type methods.
+More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because $\mathcal{H}^{-1}$ locally approximates the Bayesian posterior covariance for $m$.
 %We only have access to $\mathcal{H}$ via its action on an arbitrary vectors $v$, i.e., evaluation of the map $v \mapsto \mathcal{H} v$.
@@
 The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term \cite{CCGOPAPERS}.
-Either the Lanczos method or the randomized singular value decomposition \cite{HMTRANDOM} are used to perform the low rank approximation using only matrix-vector products.
+Conventionally, either the Lanczos method or the randomized singular value decomposition \cite{HMTRANDOM} are used to perform the low rank approximation using only matrix-vector products.
 These methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of matrix-vector products are required to form the aforementioned low rank approximation.
@@
 The functions $w_i$ are spatially varying weighting functions that are used to interpolate the convolution kernels $\varphi_i$.
-Our operator approximation method is defined by how we compute the convolution kernels, $\varphi_i$, how we choose the points, $x_i$, what weighting functions, $w_i$, that we use, and how we use the resulting product-convolution approximation to perform linear algebra operations such as solving linear systems.
+Our operator approximation method is defined by how we compute the convolution kernels, $\varphi_i$, how we choose the points, $x_i$, and what weighting functions, $w_i$, that we use.
 In Section \ref{sec:overview_intro} we summarize the answers to these questions; we will provide detailed answers to these questions in Section \ref{sec:method}.

### 2021-09-07 paper/localpsf.tex
added .tex files
writing well-posedness proof

@@
 Also, these weighting functions sum to one pointwise (Proposition \ref{thm:wi_sum_to_one}).
-Proofs for results in this section are given in Appendix \ref{app:proofs}.
-
-In Algorithm \ref{alg:weighting_functions_incremental}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$.
+
+In Algorithm \ref{alg:weighting_functions_incremental}, we show how to efficiently compute the functions $w_i$. Briefly:
+\begin{enumerate}
+\item For each sample point $x_i$, we compute a function $\psi_i$ by solving the following Neumann Poisson problem:
+\begin{equation}
+\label{eq:psi_eq}
+\begin{cases}
+\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
+\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
+\end{cases}
+\end{equation}
+with condition $\int_\Omega \psi_i(x) dx = 0$.
+Here $\delta_{x_i}$ is the point source centered at $x_i$, $|\Omega|$ is the measure of the domain $\Omega$, and $\nu$ is the normal vector to $\partial \Omega$.
+\item For each function $\psi_i$, we compute a function $\theta_i$ by solving another Neumann Poisson problem:
+\begin{equation}
+\label{eq:theta_eq}
+\begin{cases}
+\Delta \theta_i = \psi_i, & \text{in }\Omega, \\
+\nu \cdot \nabla \theta_i = 0 & \text{on } \partial \Omega,
+\end{cases}
+\end{equation}
+with condition $\int_\Omega \theta_i(x) dx = 0$.
+%\item For each sample point $x_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\delta_{x_i}$ (a point source centered at $x_i$) as the right hand side source.
+We call the solution to this first Poisson problem $\psi_i$.
+%\item For each function $\psi_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\psi_i$ as the right hand side source.
+We call the solution to this second Poisson problem $\theta_i$.
+\item We form $w_i$ as a linear combination of the functions $\theta_i$, and the constant function $C(x)=1$.
+Details for this linear combination are given in Algorithm \ref{alg:weighting_functions_incremental}.
+\end{enumerate}
+In Theorem \ref{thm:wi_alg_is_correct}, we prove that the functions $w_i$ produced by this process indeed solve optimization problem \eqref{eq:wi_optimization_problem}.
+
 Constructing all $r$ of the functions $w_i$ requires solving the Poisson equation $2r$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid.
 All of the weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added.
-However, the dominant computational cost in constructing the weighting functions is the solution of the Poisson PDEs, and these Poisson PDE solves can be performed incrementally---we only need to perform two new Poisson PDE solves when a new point is added.
+However, the dominant computational cost in constructing the weighting functions is the solution of the Poisson PDEs, and these Poisson PDE solves can be performed incrementally---by storing the functions $\psi_i$ and $\theta_i$, we only need to perform two new Poisson PDE solves when a new point is added.
 %, rather than $2(r+1)$ new PDE solves.
 
-It is important to note that functions in $H^2_N(\Omega)$ only have well-defined pointwise values if the spatial dimension is $d=1$, $2$, or $3$.
-Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-defined in three spatial dimensions or fewer.
-When inverting for parameter fields defined on four or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used.
+It is important to note that functions in $H^2_N(\Omega)$ only have well-defined pointwise values if the spatial dimension of the domain is $d=1$, $2$, or $3$.
+Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-posed in three spatial dimensions or fewer.
+In four or higher spatial dimensions, one could construct appropriate weighting functions using radial basis functions, though we do not discuss this here.
 
@@
 	\label{thm:smoothest_interpolant}
-Let $y \in \mathbb{R}^d$, and let $\gamma_y$ be the function
+Let $y \in \mathbb{R}^d$, $d \in \{1,2,3\}$, and let $\gamma_y$ be the function
 \begin{equation}
@@
 
-		Initialize constant function $C(x) = 1$
-
-		Compute Lebesgue measure of the domain, $|\Omega|$
-
-		\tcp{Compute Poisson impulse responses $\{\psi_i\}_{i=1}^r$ and Poisson squared impulse responses $\{\theta_i\}_{i=1}^r$}
+%		Initialize constant function $C(x) = 1$
+
+%		Compute Lebesgue measure of the domain, $|\Omega|$
+
+%		\tcp{Compute Poisson impulse responses $\{\psi_i\}_{i=1}^r$ and Poisson squared impulse responses $\{\theta_i\}_{i=1}^r$}
 
 		\For{$i=1,2, \dots, r$}{
-			Initialize $\delta_{x_i}$ as the point source (delta distribution) centered at $x_i$.
-
-			Compute $\psi_i$ by solving the Neumann Poisson problem
-			\begin{equation}
-			\label{eq:psi_eq}
-			\begin{cases}
-			\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
-			\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
-			\end{cases}
-			\end{equation}
-			with condition $\int_\Omega \psi_i(x) dx = 0$, where $\nu$ is the normal vector to $\partial \Omega$.
-
-			Compute $\theta_i$ by solving the Neumann Poisson problem
-			\begin{equation}
-			\label{eq:theta_eq}
-						\begin{cases}
-			\Delta \theta_i = \psi_i, & \text{in }\Omega, \\
-			\nu \cdot \nabla \theta_i = 0 & \text{on } \partial \Omega,
-			\end{cases}
-			\end{equation}
-			with condition $\int_\Omega \theta_i(x) dx = 0$, where $\nu$ is the normal vector to $\partial \Omega$.
+%			Initialize $\delta_{x_i}$ as the point source (delta distribution) centered at $x_i$.
+
+			Solve the Neumann Poisson problem \eqref{eq:psi_eq} for $\psi_i$.
+
+%			Compute $\psi_i$ by solving the Neumann Poisson problem
+%			\begin{equation}
+%			\label{eq:psi_eq}
+%			\begin{cases}
+%			\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
+%			\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
+%			\end{cases}
+%			\end{equation}
+%			with condition $\int_\Omega \psi_i(x) dx = 0$, where $\nu$ is the normal vector to $\partial \Omega$.
+
+			Solve the Neumann Poisson problem \eqref{eq:theta_eq} for $\theta_i$.
+
+%			Compute $\theta_i$ by solving the Neumann Poisson problem
+%			\begin{equation}
+%			\label{eq:theta_eq}
+%						\begin{cases}
+%			\Delta \theta_i = \psi_i, & \text{in }\Omega, \\
+%			\nu \cdot \nabla \theta_i = 0 & \text{on } \partial \Omega,
+%			\end{cases}
+%			\end{equation}
+%			with condition $\int_\Omega \theta_i(x) dx = 0$, where $\nu$ is the normal vector to $\partial \Omega$.
 
@@
 
-		\tcp{Compute weighting functions $\{w_i\}_{i=1}^r$}
+%		\tcp{Compute weighting functions $\{w_i\}_{i=1}^r$}
 
@@
 Equation \eqref{eq:desired_property_3} ensures that the negative eigenvalues of $\AmatPcSym$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $\ratfct$.
-There is a tradeoff: if we choose a larger value for $f(\lambda_\text{min})$, then the positive eigenvalues of $\AmatPcSymPlus$ will be closer to the corresponding positive eigenvalues of $\AmatPcSym$ (good), but the erroneous negative eigenvalues of $\AmatPcSym$ will be amplified in magnitude (bad).
+There is a tradeoff: if we choose a larger value for $s(\lambda_\text{min})$, then the positive eigenvalues of $\AmatPcSymPlus$ will be closer to the corresponding positive eigenvalues of $\AmatPcSym$ (good), but the erroneous negative eigenvalues of $\AmatPcSym$ will be amplified in magnitude (bad).
 This is shown in Figure FIG.
@@
 
-\begin{itemize}
+c\begin{itemize}
 	\item computational complexity in terms of PDE solves:
@@
 \begin{proof}
-	By the Sobolev Imbedding Theorem (see, e.g., Theorem 7.20 in \cite{ARBOGAST}), $H^2_N(\Omega)$ is embedded in $C^0(\Omega)$. WORK IN PROGRESS
+	By the Sobolev Imbedding Theorem (see, e.g., Theorem 7.20 in \cite{ARBOGAST}), $H^2_N(\Omega)$ is continuously imbedded in $C^0(\Omega)$, so pointwise evaluation $u \mapsto u(x_i)$ of functions at a point $x_i \in \Omega$ is a continuous linear functional on $H^2_N$.
+The set
+	\begin{equation}
+	\label{eq:inverse_image_of_point}
+	\{u \in H^2_N : u(x_i)=f_i\}
+	\end{equation}
+	is therefore a closed affine set in $H^2_N$, because it is the inverse image of a point under a continuous linear functional.
+Hence the feasible set for optimization problem \eqref{eq:generic_optimization_initial},
+	\begin{equation}
+	\label{eq:wi_feasible_set}
+	X := \{u \in H^2_N : u(x_i)=f_i~\text{for}~ i=1,\dots,r\},
+	\end{equation}
+	is a closed affine set in $H^2_N$, because it is the intersection of the sets in \eqref{eq:inverse_image_of_point} for $i=1,\dots,r$.
+The feasible set $X$ is convex because it is affine, and it is non-empty because interpolating the data pairs $\{(x_i, f_i)\}_{i=1}^r$ using any smooth interpolation method yields an element of $X$.
+The objective function, $\gamma \mapsto \frac{1}{2}||\Delta \gamma||^2$, is convex because it
+
+	We now show that the objective function is strongly convex on the tangent space to $X$,
+	\begin{equation*}
+	X_0 := \{u \in H^2_N : u(x_i)=0~\text{for}~ i=1,\dots,r\}.
+	\end{equation*}
+
 \end{proof}
@@
 \section*{Acknowledgments}
-We thank J.J. Alger and Longfei Gao for helpful discussions.
+We thank J.J. Alger, Longfei Gao, and Rami Nammour for helpful discussions.
 

### 2021-09-08 paper/localpsf.tex
wrote poincare-like result proof

@@
 
+%These results rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}.
+We prove correctness of Algorithm \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}.
+Theorem \ref{thm:wi_alg_is_correct} relies on a prerequisite result which we present in Lemma \ref{lem:psi_delta_eval}.
+
+\begin{defn}[Weighting function optimization problem]
+	Let $\Omega \subset \mathbb{R}^d$, $d\in\{1,2,3\}$, be an open bounded domain with, let $x_i \in \Omega$ for $i=1,\dots,r$, and let $f \in \mathbb{R}^r$. We consider the optimization problem
+	\begin{equation}
+	\label{eq:generic_optimization_initial}
+	\begin{aligned}
+	\min_{\gamma \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \gamma}_{L^2(\Omega)}^2 \\
+	\text{such that} &\quad \gamma(x_i) = f_i, \quad i=1,\dots,r.
+	\end{aligned}
+	\end{equation}
+	A minimizer, $\gamma^*$, of \eqref{eq:generic_optimization_initial} is a function defined on $\Omega$ that interpolates the data pairs $(x_i,f_i)$, and is as flat as possible in a least squares sense.
+	We denote the objective function for optimization problem \eqref{eq:generic_optimization_initial} by 
+	\begin{equation}
+	\label{eq:wi_objective}
+	J(\gamma) := \frac{1}{2}||\Delta \gamma||^2_{L^2(\Omega)}.
+	\end{equation}
+	We denote the feasible set for optimization problem \eqref{eq:generic_optimization_initial} by 
+	\begin{equation}
+	\label{eq:wi_feasible_set}
+		X := \{u \in H^2_N : u(x_i)=f_i~\text{for}~ i=1,\dots,r\}.
+	\end{equation}
+	We denote the tangent space to the feasible set by
+	\begin{equation}
+	\label{eq:wi_tangent_space}
+	X_0 := \{u \in H^2_N : u(x_i)=0~\text{for}~ i=1,\dots,r\}.
+	\end{equation}
+\end{defn}
+
+\begin{lem}[Elliptic regularity on $X_0$]
+	\label{lem:elliptic_regularity_x0}
+	If $\partial \Omega$ is $C^{1,1}$, then there exists a constant $C > 0$ such that
+	\begin{equation*}
+		||u||_{H^2(\Omega)} \le C ||\Delta u||_{L^2(\Omega)}
+	\end{equation*}
+	for all $u \in X_0$.
+\end{lem}
+
+\begin{proof}
+	Let $\widetilde{u}$ be version of $u$ that has been shifted up or down by a constant such that $\int_\Omega \overline{u} = 0$. We have
+	\begin{align*}
+		||D^1 u||_{\left(L^2(\Omega)\right)^d}^2 + ||D^2 u||_{\left(L^2(\Omega)\right)^{d \times d}}^2 &= ||D^1 \widetilde{u}||_{\left(L^2(\Omega)\right)^d}^2 + ||D^2 \widetilde{u}||_{\left(L^2(\Omega)\right)^{d \times d}}^2\\
+		&\le ||\widetilde{u}||_{H^2(\Omega)}^2 \\
+		&\le C_1 ||\Delta \widetilde{u}||_{L^2(\Omega)}^2 \\
+		&= C_1 ||\Delta u||_{L^2(\Omega)}^2
+	\end{align*}
+	for some constant $C_1$.
+In the first and fourth lines we used the fact that the derivative of a constant is zero, in the second line we used the definition of the $H^2$ norm, and in the third line we used the standard elliptic regularity result for pure Neumann problems (see, e.g., Theorem 8.13 (b) in \cite{Arbogast}).
+
+	To complete the proof, we will now show that there exists a constant $C_2$ such that $||u||_{L^2(\Omega)}^2 \le C_2 ||\Delta u||_{L^2(\Omega)}^2$ for all $u \in X_0$.
+For this, we use a standard compactness argument, which closely follows the proof of Theorem 8.12 in \cite{ARBOGAST}.
+Suppose for the sake of contradiction that this result does not hold.
+Then there exists a sequence $\{u_k\}_{k=1}^\infty \subset X_0$ such that
+	$||u_k||_{L^2(\Omega)}=1$ and $||\Delta u_k||_{L^2(\Omega)} \le 1/k$.
+	By the result already proven, the sequence $u_k$ is bounded in $H^2(\Omega)$ as follows:
+	\begin{align*}
+	||u_k||_{H^2(\Omega)}^2 &= ||u_k||_{L^2(\Omega)}^2 + ||D^1 u||_{\left(L^2(\Omega)\right)^d}^2 + ||D^2 u||_{\left(L^2(\Omega)\right)^{d \times d}}^2 \\
+	&\le ||u_k||_{L^2(\Omega)}^2 + C_1 ||\Delta u||_{L^2(\Omega)}^2 \le 1 + C_1.
+	\end{align*}
+	Therefore, by the Banach-Alaoglu theorem, there exists a subsequence $u_{k_j}$ and function $u \in X_0$ such that
+	\begin{equation*}
+		u_{k_j} \overset{w}{\rightharpoonup} u \quad \text{weakly in }H^2(\Omega).
+	\end{equation*}
+	Let $f \in L^2(\Omega)$.
+By the Cauchy-Schwarz inequality, the linear functional $v \mapsto \int_\Omega f(x) \Delta v(x) dx$ is bounded in $H^2(\Omega)$, and is therefore an element of $\left(H^2(\Omega)\right)^*$.
+Weak convergence therefore implies
+	\begin{equation*}
+		\int_\Omega f(x) \Delta u_k(x) dx \rightarrow \int_\Omega f(x) \Delta u(x) dx.
+	\end{equation*}
+	By the assumption $||\Delta u_k||_{L^2(\Omega)} \le 1/k$ and Cauchy-Schwarz, we also have 
+	\begin{equation*}
+	\int_\Omega f(x) \Delta u_k(x) dx \rightarrow 0,
+	\end{equation*}
+	so we must have
+	\begin{equation*}
+		\int_\Omega f(x) \Delta u(x) dx = 0.
+	\end{equation*}
+	Since this holds for all $f \in L^2(\Omega)$, by the fundamental lemma of the calculus of variations, $\Delta u(x) = 0$ almost everywhere in $\Omega$.
+This, combined with the fact that $u$ has Neumann zero boundary conditions, implies that $u$ is a constant function.
+Since $u$ has value zero at least one point in each connected component of $\Omega$, we have $u=0$.
+Note that pointwise values of $u$ are well-defined since the Sobolev embedding theorem imples that $H^2(\Omega)$ is continuously imbedded in the space of continuous functions $C^0(\Omega)$ when $d < 4$, and we are considering $d \in \{1,2,3\}$ (see, e.g., Theorem 7.19 in \cite{Arbogast}).
+
+	On the other hand, by the Rellich-Kondrachov theorem, $H^2(\Omega)$ is compactly imbedded in $L^2(\Omega)$, so there must exist a sub-sub-sequence, $u_{k_{j_i}}$ such that 
+	\begin{equation*}
+		u_{k_{j_i}} \rightarrow u
+	\end{equation*}
+	strongly in $L^2(\Omega)$.
+By assumption $||u_{k_{j_i}}||_{L^2(\Omega)}=1$, so $||u||_{L^2(\Omega)}=1$.
+This contradicts the result $u=0$ that we previously derived, so we must conclude that there exists a constant $C_2 > 0$ such that
+	\begin{equation*}
+		||u||_{L^2(\Omega)} \le C_2 ||\Delta u||_{L^2(\Omega)}.
+	\end{equation*}
+	for all $u \in X_0$. Combining this with out previous result, we have
+	\begin{equation*}
+	||u||^2_{H^2(\Omega)} \le (C_1 + C_2)||\Delta u||^2
+	\end{equation*}
+	for all $u \in X_0$, which completes the proof.
+\end{proof}
+
+\begin{lem}[Well-posedness of weighting function optimization problem]
+	\label{lem:wi_existence_and_uniqueness}
+	There exists a unique solution to optimization problem \eqref{eq:generic_optimization_initial}.
+\end{lem}
+
+\begin{proof}
+	By the Sobolev Imbedding Theorem (see, e.g., Theorem 7.20 in \cite{ARBOGAST}), $H^2_N(\Omega)$ is continuously imbedded in $C^0(\Omega)$, so pointwise evaluation $u \mapsto u(x_i)$ of functions at a point $x_i \in \Omega$ is a continuous linear functional on $H^2_N$.
+The set
+	\begin{equation}
+	\label{eq:inverse_image_of_point}
+	\{u \in H^2_N : u(x_i)=f_i\}
+	\end{equation}
+	is therefore a closed affine set in $H^2_N$, because it is the inverse image of a point under a continuous linear functional.
+Hence the feasible set $X$, defined in \eqref{eq:wi_feasible_set}, is a closed affine set in $H^2_N$, because it is the intersection of the sets in \eqref{eq:inverse_image_of_point} for $i=1,\dots,r$.
+
+	The feasible set $X$ is convex because it is affine, and it is non-empty because interpolating the data pairs $\{(x_i, f_i)\}_{i=1}^r$ using any smooth interpolation method yields an element of $X$.
+Furthermore, the objective function,
+	\begin{equation}
+	J(\gamma) := \frac{1}{2}||\Delta \gamma||^2_{L^2(\Omega)},
+	\end{equation}
+	is convex because it is the composition of a convex function with a linear map. Furthermore, the elliptic regularity result in Lemma \ref{lem:elliptic_regularity_x0} implies
+	\begin{equation*}
+		J''(\gamma)v^2 = ||\Delta v||^2_{L^2(\Omega)} \ge C ||v||_{H^2(\Omega)}^2
+	\end{equation*}
+	for all $v \in X_0$, which implies that $J$ is strongly convex on $X$.
+\end{proof}
+
+\begin{lem}[Linearity of weighting function solution operator]
+	\label{lem:T_is_linear}
+	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
+		\begin{equation}
+		\label{eq:opt_soln_operator}
+		\mathcal{T}(f) := \gamma^*,
+		\end{equation}
+		where $\gamma^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
+\end{lem}
+
+\begin{proof}
+	asdf
+\end{proof}
+
+\begin{proof}[Proof of Theorem \ref{thm:smoothest_interpolant}]
+	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.
+The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
+	\begin{align*}
+	\mathcal{T}(f) &= \varphi_1(y) \mathcal{T}(e_1) + \varphi_2(y) \mathcal{T}(e_2) + \dots +\varphi_r(y) \mathcal{T}(e_r) \\
+	&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
+	&= \gamma_y.
+	\end{align*}
+	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $\gamma_y$ in \eqref{eq:u_defn_in_thm}.
+\end{proof}
+
+\begin{proof}[Proof of Proposition \ref{thm:wi_sum_to_one}]
+	By linearity of $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), we have
+	\begin{equation*}
+	\sum_{i=1}^r w_i(x) = \sum_{i=1}^r \mathcal{T}(e_i)(x) = \mathcal{T}\left(\sum_{i=1}^r e_i\right)(x) = \mathcal{T}(\mathbf{1})(x) = 1,
+	\end{equation*}
+	where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one.
+The last equality ($\mathcal{T}(\mathbf{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \mathbf{1}$, and yields the minimum possible value of zero for the objective function.
+\end{proof}
+
+\begin{lem}[Preparation for Theorem \ref{thm:wi_alg_is_correct}]
+	\label{lem:psi_delta_eval}
+	We have
+	\begin{equation}
+	\label{eq:inline_wi_lemma}
+	h(x_j) = \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \avg(h)
+	\end{equation}
+	for all $h\in H^2_N(\Omega)$, where
+	\begin{equation*}
+	\avg(h) :=	 \frac{1}{|\Omega|} \int_\Omega h ~dx,
+	\end{equation*}
+	and $|\Omega| = \int_\Omega 1 ~dx$ is the Lebesgue measure of $\Omega$.
+\end{lem}
+
+\begin{proof}
+	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\psi_j$ onto $h$, and using the Neumann zero boundary conditions for $\psi_j$ and $h$ to eliminate boundary integral terms.
+\end{proof}
+
+\begin{proof}[Proof of Theorem \ref{alg:weighting_functions_incremental}]
+	For notational convenience, in this proof we write $w$, $\lambda$, and $e$ instead of $w_i$, $\lambda_i$, and $e_i$, respectively.
+Here subscripts for $\lambda$, and $e$ instead denote components of those vectors.
+For example, here $\lambda_j$ denotes the $j^\text{th}$ component of the vector $\lambda$, which corresponds to the $j^\text{th}$ component of the vector $\lambda_i$ in Algorithm \ref{alg:weighting_functions_incremental}.
+
+	From standard optimization theory, the solution to optimization problem \ref{eq:wi_optimization_problem} is the stationary point of the Lagrangian,
+	\begin{equation*}
+	\mathcal{L}(w,\lambda) = \frac{1}{2}\|\Delta w\|^2 + \sum_{j=1}^r \lambda_j\left(w(x_j) - e_j\right).
+	\end{equation*}
+	The plan of this proof is to show that $\nabla \mathcal{L}(w,\lambda)=0$, which will verify that $w$ solves optimization problem \ref{eq:wi_optimization_problem}.
+
+	Using the definition of $w_i$ in \eqref{eq:definition_of_wi_alg}, we may write the gradient of $\mathcal{L}$ with respect to $\lambda_j$ as
+	\begin{equation*}
+	\nabla_{\lambda_j} \mathcal{L}(w, \lambda) = w(x_j) - e_j
+	= \beta - \sum_{k=1}^r \lambda_k \theta_k(x_j) - e_j.
+	\end{equation*}
+	Using Lemma \ref{lem:psi_delta_eval} with $h=\theta_k$, the equation for $\theta_k$ in \eqref{eq:theta_eq}, and the definition of $S$ in \eqref{eq:defn_of_S_alg}, we have
+	\begin{align*}
+	\sum_{k=1}^r \lambda_k \theta_k(x_j) &= \sum_{k=1}^r \lambda_k \left(\psi_j, \Delta \theta_k\right)_{L^2(\Omega)}\\
+	&= \sum_{k=1}^r \lambda_k \left(\psi_j, \psi_k\right)_{L^2(\Omega)} = \left(S \lambda\right)_j.
+	\end{align*}
+	Combining all of these components of $\nabla_\lambda \mathcal{L}$ into a single vector and using the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} yields
+	\begin{align*}
+	\nabla_\lambda \mathcal{L}(w, \lambda) &= \beta \mathbf{1} - e - S \lambda \\
+	&= \beta \mathbf{1} - e - S \left(-S^{-1}\left(e - \beta \mathbf{1}\right)\right) = 0.
+	\end{align*}
+
+	Taking the directional derivative of $\mathcal{L}$ with respect to $w$ in direction $h$ and using Lemma \ref{lem:psi_delta_eval} yields
+	\begin{align}
+	\nabla_w \mathcal{L}(w, \lambda)h &= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j h(x_j) \nonumber \\
+	&= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \lambda_j \avg(h) \nonumber \\
+	&= \left(\Delta w + \sum_{j=1}^r \lambda_j \psi_j,~ \Delta h\right)_{L^2(\Omega)} + \avg(h) \mathbf{1}^T \lambda. \label{eq:L_w_38}
+	\end{align}
+	Using the definition of $w$ in Algorithm \ref{alg:weighting_functions_incremental}, and the fact that the Laplacian of a constant function is zero, we have
+	\begin{equation*}
+	\Delta w = -\sum_{j=1}^r \lambda_j \Delta \theta_j 
+	= -\sum_{j=1}^r \lambda_j \psi_j,
+	\end{equation*}
+	which implies that the first term in \eqref{eq:L_w_38} is zero. By the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} and $\beta$ in \eqref{eq:definition_of_alpha_alg}, we have
+	\begin{align*}
+	\mathbf{1}^T \lambda &= -\mathbf{1}^T S^{-1} \left(e - \beta \mathbf{1}\right) \\
+	&= -\mathbf{1}^T S^{-1} e + \beta \mathbf{1}^T S^{-1} \mathbf{1} \\
+	&= -\mathbf{1}^T S^{-1} e + \frac{\mathbf{1}^T S^{-1} e}{\mathbf{1}^T S^{-1} \mathbf{1}} \mathbf{1}^T S^{-1} \mathbf{1} = 0,
+	\end{align*}
+	which implies that the second term in \eqref{eq:L_w_38} is zero.
+Thus $\nabla_w \mathcal{L}(w, \lambda) = 0$, so Algorithm \ref{alg:weighting_functions_incremental} produces the solution to optimization problem \ref{eq:wi_optimization_problem}.
+\end{proof}
+
 \begin{defn}[Moments of $\phi_x$]
@@
 \end{defn}
-
-%These results rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}.
-We prove correctness of Algorithm \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}.
-Theorem \ref{thm:wi_alg_is_correct} relies on a prerequisite result which we present in Lemma \ref{lem:psi_delta_eval}.
-
-\begin{lem}[Preparation for Theorem \ref{thm:smoothest_interpolant}]
-	\label{lem:T_is_linear}
-	Let $\Omega \subset \mathbb{R}^d$, $d\in\{1,2,3\}$, be an open bounded domain with a bounded extension operator, let $x_i \in \Omega$ for $i=1,\dots,r$, and let $f \in \mathbb{R}^r$.
-Consider the optimization problem
-	\begin{equation}
-	\label{eq:generic_optimization_initial}
-	\begin{aligned}
-	\min_{\gamma \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \gamma}_{L^2(\Omega)}^2 \\
-	\text{such that} &\quad \gamma(x_i) = f_i, \quad i=1,\dots,r.
-	\end{aligned}
-	\end{equation}
-	A minimizer, $\gamma^*$, of \eqref{eq:generic_optimization_initial} is a function defined on $\Omega$ that interpolates the data pairs $(x_i,f_i)$, and is as flat as possible in a least squares sense.
-	The following statements hold.
-	\begin{enumerate}
-		\item[(a)] There exists a unique solution to \eqref{eq:generic_optimization_initial}.
-		\item[(b)] 	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
-		\begin{equation}
-		\label{eq:opt_soln_operator}
-		\mathcal{T}(f) := \gamma^*,
-		\end{equation}
-		where $\gamma^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
-	\end{enumerate}
-\end{lem}
-\begin{proof}
-	By the Sobolev Imbedding Theorem (see, e.g., Theorem 7.20 in \cite{ARBOGAST}), $H^2_N(\Omega)$ is continuously imbedded in $C^0(\Omega)$, so pointwise evaluation $u \mapsto u(x_i)$ of functions at a point $x_i \in \Omega$ is a continuous linear functional on $H^2_N$.
-The set
-	\begin{equation}
-	\label{eq:inverse_image_of_point}
-	\{u \in H^2_N : u(x_i)=f_i\}
-	\end{equation}
-	is therefore a closed affine set in $H^2_N$, because it is the inverse image of a point under a continuous linear functional.
-Hence the feasible set for optimization problem \eqref{eq:generic_optimization_initial},
-	\begin{equation}
-	\label{eq:wi_feasible_set}
-	X := \{u \in H^2_N : u(x_i)=f_i~\text{for}~ i=1,\dots,r\},
-	\end{equation}
-	is a closed affine set in $H^2_N$, because it is the intersection of the sets in \eqref{eq:inverse_image_of_point} for $i=1,\dots,r$.
-The feasible set $X$ is convex because it is affine, and it is non-empty because interpolating the data pairs $\{(x_i, f_i)\}_{i=1}^r$ using any smooth interpolation method yields an element of $X$.
-The objective function, $\gamma \mapsto \frac{1}{2}||\Delta \gamma||^2$, is convex because it
-
-	We now show that the objective function is strongly convex on the tangent space to $X$,
-	\begin{equation*}
-	X_0 := \{u \in H^2_N : u(x_i)=0~\text{for}~ i=1,\dots,r\}.
-	\end{equation*}
-
-\end{proof}
-
-\begin{proof}[Proof of Theorem \ref{thm:smoothest_interpolant}]
-	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.
-The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
-	\begin{align*}
-	\mathcal{T}(f) &= \varphi_1(y) \mathcal{T}(e_1) + \varphi_2(y) \mathcal{T}(e_2) + \dots +\varphi_r(y) \mathcal{T}(e_r) \\
-	&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
-	&= \gamma_y.
-	\end{align*}
-	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $\gamma_y$ in \eqref{eq:u_defn_in_thm}.
-\end{proof}
-
-\begin{proof}[Proof of Proposition \ref{thm:wi_sum_to_one}]
-	By linearity of $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), we have
-	\begin{equation*}
-	\sum_{i=1}^r w_i(x) = \sum_{i=1}^r \mathcal{T}(e_i)(x) = \mathcal{T}\left(\sum_{i=1}^r e_i\right)(x) = \mathcal{T}(\mathbf{1})(x) = 1,
-	\end{equation*}
-	where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one.
-The last equality ($\mathcal{T}(\mathbf{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \mathbf{1}$, and yields the minimum possible value of zero for the objective function.
-\end{proof}
-
-\begin{lem}[Preparation for Theorem \ref{thm:wi_alg_is_correct}]
-	\label{lem:psi_delta_eval}
-	We have
-	\begin{equation}
-	\label{eq:inline_wi_lemma}
-	h(x_j) = \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \avg(h)
-	\end{equation}
-	for all $h\in H^2_N(\Omega)$, where
-	\begin{equation*}
-	\avg(h) :=	 \frac{1}{|\Omega|} \int_\Omega h ~dx,
-	\end{equation*}
-	and $|\Omega| = \int_\Omega 1 ~dx$ is the Lebesgue measure of $\Omega$.
-\end{lem}
-
-\begin{proof}
-	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\psi_j$ onto $h$, and using the Neumann zero boundary conditions for $\psi_j$ and $h$ to eliminate boundary integral terms.
-\end{proof}
-
-\begin{proof}[Proof of Theorem \ref{alg:weighting_functions_incremental}]
-	For notational convenience, in this proof we write $w$, $\lambda$, and $e$ instead of $w_i$, $\lambda_i$, and $e_i$, respectively.
-Here subscripts for $\lambda$, and $e$ instead denote components of those vectors.
-For example, here $\lambda_j$ denotes the $j^\text{th}$ component of the vector $\lambda$, which corresponds to the $j^\text{th}$ component of the vector $\lambda_i$ in Algorithm \ref{alg:weighting_functions_incremental}.
-
-	From standard optimization theory, the solution to optimization problem \ref{eq:wi_optimization_problem} is the stationary point of the Lagrangian,
-	\begin{equation*}
-	\mathcal{L}(w,\lambda) = \frac{1}{2}\|\Delta w\|^2 + \sum_{j=1}^r \lambda_j\left(w(x_j) - e_j\right).
-	\end{equation*}
-	The plan of this proof is to show that $\nabla \mathcal{L}(w,\lambda)=0$, which will verify that $w$ solves optimization problem \ref{eq:wi_optimization_problem}.
-
-	Using the definition of $w_i$ in \eqref{eq:definition_of_wi_alg}, we may write the gradient of $\mathcal{L}$ with respect to $\lambda_j$ as
-	\begin{equation*}
-	\nabla_{\lambda_j} \mathcal{L}(w, \lambda) = w(x_j) - e_j
-	= \beta - \sum_{k=1}^r \lambda_k \theta_k(x_j) - e_j.
-	\end{equation*}
-	Using Lemma \ref{lem:psi_delta_eval} with $h=\theta_k$, the equation for $\theta_k$ in \eqref{eq:theta_eq}, and the definition of $S$ in \eqref{eq:defn_of_S_alg}, we have
-	\begin{align*}
-	\sum_{k=1}^r \lambda_k \theta_k(x_j) &= \sum_{k=1}^r \lambda_k \left(\psi_j, \Delta \theta_k\right)_{L^2(\Omega)}\\
-	&= \sum_{k=1}^r \lambda_k \left(\psi_j, \psi_k\right)_{L^2(\Omega)} = \left(S \lambda\right)_j.
-	\end{align*}
-	Combining all of these components of $\nabla_\lambda \mathcal{L}$ into a single vector and using the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} yields
-	\begin{align*}
-	\nabla_\lambda \mathcal{L}(w, \lambda) &= \beta \mathbf{1} - e - S \lambda \\
-	&= \beta \mathbf{1} - e - S \left(-S^{-1}\left(e - \beta \mathbf{1}\right)\right) = 0.
-	\end{align*}
-
-	Taking the directional derivative of $\mathcal{L}$ with respect to $w$ in direction $h$ and using Lemma \ref{lem:psi_delta_eval} yields
-	\begin{align}
-	\nabla_w \mathcal{L}(w, \lambda)h &= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j h(x_j) \nonumber \\
-	&= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \lambda_j \avg(h) \nonumber \\
-	&= \left(\Delta w + \sum_{j=1}^r \lambda_j \psi_j,~ \Delta h\right)_{L^2(\Omega)} + \avg(h) \mathbf{1}^T \lambda. \label{eq:L_w_38}
-	\end{align}
-	Using the definition of $w$ in Algorithm \ref{alg:weighting_functions_incremental}, and the fact that the Laplacian of a constant function is zero, we have
-	\begin{equation*}
-	\Delta w = -\sum_{j=1}^r \lambda_j \Delta \theta_j 
-	= -\sum_{j=1}^r \lambda_j \psi_j,
-	\end{equation*}
-	which implies that the first term in \eqref{eq:L_w_38} is zero. By the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} and $\beta$ in \eqref{eq:definition_of_alpha_alg}, we have
-	\begin{align*}
-	\mathbf{1}^T \lambda &= -\mathbf{1}^T S^{-1} \left(e - \beta \mathbf{1}\right) \\
-	&= -\mathbf{1}^T S^{-1} e + \beta \mathbf{1}^T S^{-1} \mathbf{1} \\
-	&= -\mathbf{1}^T S^{-1} e + \frac{\mathbf{1}^T S^{-1} e}{\mathbf{1}^T S^{-1} \mathbf{1}} \mathbf{1}^T S^{-1} \mathbf{1} = 0,
-	\end{align*}
-	which implies that the second term in \eqref{eq:L_w_38} is zero.
-Thus $\nabla_w \mathcal{L}(w, \lambda) = 0$, so Algorithm \ref{alg:weighting_functions_incremental} produces the solution to optimization problem \ref{eq:wi_optimization_problem}.
-\end{proof}
 

### 2021-09-09 paper/localpsf.tex
figured out more direct proof of weighting function thing

@@
 
-In Algorithm \ref{alg:weighting_functions_incremental}, we show how to efficiently compute the functions $w_i$. Briefly:
+For each $i=1,\dots,r$, optimization problem \eqref{eq:wi_optimization_problem} is an example of the following generic omtimization problem:
+\begin{equation}
+\label{eq:wi_optimization_problem_generic}
+\begin{aligned}
+\min_{v \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta v}_{L^2(\Omega)}^2 \\
+\text{such that} &\quad v(x_j) = v_j, \quad j=1,\dots,r,
+\end{aligned}
+\end{equation}
+where $v_j \in \mathbb{R}$, $j=1,\dots,r$ are now arbitrary function values that $v$ must take at the points $x_j$.
+We solve \eqref{eq:wi_optimization_problem_generic} efficiently via the following procedure:
 \begin{enumerate}
@@
 We call the solution to this second Poisson problem $\theta_i$.
-\item We form $w_i$ as a linear combination of the functions $\theta_i$, and the constant function $C(x)=1$.
-Details for this linear combination are given in Algorithm \ref{alg:weighting_functions_incremental}.
+\item We form $f$ as follows:
+\begin{equation*}
+v(x) := \beta + \sum_{i=1}^r c_i \theta_i(x),
+\end{equation*}
+where the constants $c_i \in \mathbb{R}$, $i=1,\dots,r$, and $\beta \in \mathbb{R}$ are computed by solving the following linear system:
+\begin{equation}
+	\label{eq:interpolation_linear_system}
+	\begin{bmatrix}
+	S & \mathbf{1} \\ \mathbf{1}^T & 0
+	\end{bmatrix}
+	\begin{bmatrix}
+	\mathbf{c} \\ \beta
+	\end{bmatrix}
+	=
+	\begin{bmatrix}
+	\mathbf{v} \\ 0
+	\end{bmatrix}.
+\end{equation}
+Here $\mathbf{c} \in \mathbb{R}^r$ is the vector with $i$th component $c_i$, $\mathbf{1} \in \mathbf{R}^r$ is the column vector of all ones, $\mathbf{v} \in \mathbb{R}^r$ is the vector with $i$th component $v_i$, and $S \in \mathbb{R}^{r \times r}$ is the matrix with entries given by
+\begin{equation*}
+S_{ij} := \int_\Omega \psi_i(x) \psi_j(x) dx.
+\end{equation*}
 \end{enumerate}
-In Theorem \ref{thm:wi_alg_is_correct}, we prove that the functions $w_i$ produced by this process indeed solve optimization problem \eqref{eq:wi_optimization_problem}.
-
-Constructing all $r$ of the functions $w_i$ requires solving the Poisson equation $2r$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid.
+In Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v$ produced by this process indeed solves optimization problem \eqref{eq:wi_optimization_problem_generic}.
+
+Constructing all $r$ of the functions $w_i$ thus requires solving the Poisson equation $2r$ times with different right hand side sources to form the functions $\psi_i$ and $\theta_i$, and these Poisson PDE solves can be performed cheaply with multigrid.
 All of the weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added.
-However, the dominant computational cost in constructing the weighting functions is the solution of the Poisson PDEs, and these Poisson PDE solves can be performed incrementally---by storing the functions $\psi_i$ and $\theta_i$, we only need to perform two new Poisson PDE solves when a new point is added.
+However, the dominant computational cost in constructing the weighting functions is the solution of the Poisson PDEs, and these Poisson PDE solves can be performed incrementally---we only need to compute functions $\psi_i$ and $\theta_i$ associated with the new points.
 %, rather than $2(r+1)$ new PDE solves.

### 2021-09-09 paper/localpsf.tex
writing proof for optimization problem

@@
 where $v_j \in \mathbb{R}$, $j=1,\dots,r$ are now arbitrary function values that $v$ must take at the points $x_j$.
-We solve \eqref{eq:wi_optimization_problem_generic} efficiently via the following procedure:
+We may efficiently solve optimization problem \eqref{eq:wi_optimization_problem_generic} via the following procedure:
 \begin{enumerate}
@@
 We call the solution to this second Poisson problem $\theta_i$.
-\item We form $f$ as follows:
-\begin{equation*}
-v(x) := \beta + \sum_{i=1}^r c_i \theta_i(x),
-\end{equation*}
-where the constants $c_i \in \mathbb{R}$, $i=1,\dots,r$, and $\beta \in \mathbb{R}$ are computed by solving the following linear system:
+\item The solution to optimization problem \eqref{eq:wi_optimization_problem_generic}, which we denote by $v_*$, is formed as follows:
+\begin{equation}
+\label{eq:optimal_v}
+v_*(x) := \beta + \sum_{i=1}^r c_i \theta_i(x),
+\end{equation}
+where the constants $c_i \in \mathbb{R}$, $i=1,\dots,r$, and $\beta \in \mathbb{R}$ are the solution to the following linear system:
 \begin{equation}
@@
 Here $\mathbf{c} \in \mathbb{R}^r$ is the vector with $i$th component $c_i$, $\mathbf{1} \in \mathbf{R}^r$ is the column vector of all ones, $\mathbf{v} \in \mathbb{R}^r$ is the vector with $i$th component $v_i$, and $S \in \mathbb{R}^{r \times r}$ is the matrix with entries given by
-\begin{equation*}
+\begin{equation}
+\label{eq:S_matrix}
 S_{ij} := \int_\Omega \psi_i(x) \psi_j(x) dx.
-\end{equation*}
+\end{equation}
 \end{enumerate}
@@
 
-Constructing all $r$ of the functions $w_i$ thus requires solving the Poisson equation $2r$ times with different right hand side sources to form the functions $\psi_i$ and $\theta_i$, and these Poisson PDE solves can be performed cheaply with multigrid.
-All of the weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added.
+Constructing all $r$ of the functions $w_i$ thus requires solving the Poisson equation $2r$ times with different right hand side sources to form the functions $\psi_i$ and $\theta_i$, then post-processing these functions.
+This is shown in Algorithm \ref{alg:weighting_functions_incremental}.
+The Poisson PDE solves, while the most expensive part of this procedure, can be performed cheaply with multigrid.
+
+%All of the weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added.
 However, the dominant computational cost in constructing the weighting functions is the solution of the Poisson PDEs, and these Poisson PDE solves can be performed incrementally---we only need to compute functions $\psi_i$ and $\theta_i$ associated with the new points.
@@
 In four or higher spatial dimensions, one could construct appropriate weighting functions using radial basis functions, though we do not discuss this here.
+
+\begin{thm}[Solution of optimization problem \ref{eq:wi_optimization_problem_generic}]
+	\label{thm:wi_alg_is_correct}
+	The function $v_*$ given in \eqref{eq:optimal_v} is the unique solution to optimization problem \ref{eq:wi_optimization_problem_generic}.
+\end{thm}
+
+\begin{proof}
+	Let
+	\begin{equation*}
+		X := \{v \in H^2_N(\Omega) : v(x_i) = v_i \text{ for }i=1,\dots,r\}
+	\end{equation*}
+	denote the feasible set for optimization problem \ref{eq:wi_optimization_problem_generic}, and let 
+	\begin{equation*}
+	X_0 := \{v \in H^2_N(\Omega) : v(x_i) = 0 \text{ for }i=1,\dots,r\}
+	\end{equation*}	
+	denote the tangent space to $X$.
+Note that pointwise evaluation of functions in $H^2_N(\Omega)$ is well-defined because the Sobolev impedding theorem implies that $H^2_N(\Omega)$ is continuously imbedded in the spacde of continuous functions $C^0(\Omega)$ whenever the spatial dimension satisfies $d < 4$, and here we consider $d \in \{1,2,3\}$.
+Hence the optimization problem, the feasible set $X$ and the tangent space $X_0$ are well-defined.
+\end{proof}
 
@@
 
-		\For{$i=1,2, \dots, r$}{
+		\For{$k=1,2, \dots, r$}{
 %			Initialize $\delta_{x_i}$ as the point source (delta distribution) centered at $x_i$.
 
-			Solve the Neumann Poisson problem \eqref{eq:psi_eq} for $\psi_i$.
+			Solve the Neumann Poisson problem \eqref{eq:psi_eq} for $\psi_k$.
 
@@
 
-			Solve the Neumann Poisson problem \eqref{eq:theta_eq} for $\theta_i$.
+			Solve the Neumann Poisson problem \eqref{eq:theta_eq} for $\theta_k$.
 
@@
 
-		Form the $r \times r$ matrix $S$ with entries
-		\begin{equation}
-		\label{eq:defn_of_S_alg}
-		S_{ij} = \left(\psi_i, \psi_j\right)_{L^2(\Omega)}.
-		\end{equation}
+		Form the $r \times r$ matrix $S$ given in \eqref{eq:S_matrix}.
 
@@
 		\For{$i=1,2,\dots,r$}{
-			Compute
-			\begin{equation}
-			\label{eq:definition_of_alpha_alg}
-			\beta = \frac{\mathbf{1}^T S^{-1} e_i}{\mathbf{1}^T S^{-1} \mathbf{1}},
-			\end{equation}
-			where $\mathbf{1} = (1,1,\dots,1) \in \mathbb{R}^r$, and $e_i = (0,\dots,0,1,0,\dots,0)\in \mathbb{R}^r$ is unit vector with $i^\text{th}$ component equal to one.
-
-			Compute 
-			\begin{equation}
-			\label{eq:definition_of_mu_alg}
-			\lambda = -S^{-1} \left(e_i - \beta \mathbf{1} \right).
-			\end{equation}
+			Solve the linear system
+			\begin{equation*}
+			\begin{bmatrix}
+			S & \mathbf{1} \\ \mathbf{1}^T & 0
+			\end{bmatrix}
+			\begin{bmatrix}
+			\mathbf{c}^{(i)} \\ \beta^{(i)}
+			\end{bmatrix}
+			=
+			\begin{bmatrix}
+			\mathbf{e}^{(i)} \\ 0
+			\end{bmatrix}.
+			\end{equation*}
+			for $\mathbf{c}^{(i)}=(c_1^{(i)}, c_2^{(i)}, \dots, c_r^{(i)})$ and $\beta^{(i)}$.
+Here $\mathbf{1} = (1,1,\dots,1) \in \mathbb{R}^r$, and $\mathbf{e}^{(i)} = (0,\dots,0,1,0,\dots,0)\in \mathbb{R}^r$ is the unit vector with $i^\text{th}$ component equal to one and all other components equal to zero.
 
 			Form the weighting function $w_i$ given by
-			\begin{equation}
-			\label{eq:definition_of_wi_alg}
-				w_i(x) = \beta - \sum_{k=1}^r \lambda_k \theta_k(x).
-			\end{equation}
+			\begin{equation*}
+			w_i(x) = \beta^{(i)} + \sum_{k=1}^r c_k^{(i)} \theta_k(x).
+			\end{equation*}
+
 		}
@@
 \end{algorithm2e}
-
-\begin{thm}[Correctness of Algorithm \ref{alg:weighting_functions_incremental}]
-\label{thm:wi_alg_is_correct}
-	The weighting functions $w_i$ computed by Algorithm \ref{alg:weighting_functions_incremental} solve optimization problem \ref{eq:wi_optimization_problem}.
-\end{thm}
 

### 2021-09-10 paper/localpsf.tex
optimal weighting function proof complete, except for lemma reference

@@
 \min_{v \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta v}_{L^2(\Omega)}^2 \\
-\text{such that} &\quad v(x_j) = v_j, \quad j=1,\dots,r,
+\text{such that} &\quad v(x_j) = f_j, \quad j=1,\dots,r,
 \end{aligned}
 \end{equation}
-where $v_j \in \mathbb{R}$, $j=1,\dots,r$ are now arbitrary function values that $v$ must take at the points $x_j$.
+where $f_j \in \mathbb{R}$, $j=1,\dots,r$ are now arbitrary function values that $v$ must take at the points $x_j$.
 We may efficiently solve optimization problem \eqref{eq:wi_optimization_problem_generic} via the following procedure:
@@
 	\begin{bmatrix}
-	\mathbf{v} \\ 0
+	\mathbf{f} \\ 0
 	\end{bmatrix}.
 \end{equation}
-Here $\mathbf{c} \in \mathbb{R}^r$ is the vector with $i$th component $c_i$, $\mathbf{1} \in \mathbf{R}^r$ is the column vector of all ones, $\mathbf{v} \in \mathbb{R}^r$ is the vector with $i$th component $v_i$, and $S \in \mathbb{R}^{r \times r}$ is the matrix with entries given by
+Here $\mathbf{c} \in \mathbb{R}^r$ is the vector with $i$th component $c_i$, $\mathbf{1} \in \mathbf{R}^r$ is the column vector of all ones, $\mathbf{f} \in \mathbb{R}^r$ is the vector with $i$th component $f_i$, and $S \in \mathbb{R}^{r \times r}$ is the matrix with entries given by
 \begin{equation}
@@
 \end{enumerate}
-In Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v$ produced by this process indeed solves optimization problem \eqref{eq:wi_optimization_problem_generic}.
+In Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v_*$ produced by this process indeed solves optimization problem \eqref{eq:wi_optimization_problem_generic}.
 
@@
 This is shown in Algorithm \ref{alg:weighting_functions_incremental}.
-The Poisson PDE solves, while the most expensive part of this procedure, can be performed cheaply with multigrid.
+The Poisson PDE solves, which are the most expensive part of this procedure, can be performed cheaply with multigrid.
 
@@
 	denote the tangent space to $X$.
-Note that pointwise evaluation of functions in $H^2_N(\Omega)$ is well-defined because the Sobolev impedding theorem implies that $H^2_N(\Omega)$ is continuously imbedded in the spacde of continuous functions $C^0(\Omega)$ whenever the spatial dimension satisfies $d < 4$, and here we consider $d \in \{1,2,3\}$.
-Hence the optimization problem, the feasible set $X$ and the tangent space $X_0$ are well-defined.
+Note that pointwise values of functions in $H_N^2(\Omega)$ are well-defined since the Sobolev imbedding theorem (see, e.g., Theorem 7.19 in \cite{Arbogast}) implies that $H^2(\Omega)$ is continuously imbedded in the space of continuous functions $C^0(\Omega)$ when $d < 4$, and we are considering $d \in \{1,2,3\}$.
+The feasible set $X$ is therefore a closed affine subspace of $H^2_N(\Omega)$, the tangent space $X_0$ is a closed subspace of $H^2_N(\Omega)$, and optimization problem \ref{eq:wi_optimization_problem_generic} is well-defined.
+
+	The proof now proceeds in three steps.
+First, we show that $v_*$ is feasible ($v_* \in X$).
+Second, we show that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} if it satisfies a certain variational equation.
+Third, we verify that $v_*$ satisfies this variational equation.
+
+	\paragraph{Feasibility} To show that $v_*$ is feasible (i.e., $v_* \in X$), we must show that $v_*(x_i) = f_i$ for $i=1,\dots,r$. We have
+	\begin{align*}
+	v_*(x_i) &= \beta + \sum_{j=1}^r c_j \theta_j(x_i) \\
+	&= \beta + \sum_{j=1}^r c_j \left(\psi_i, \Delta \theta_j\right)_{L^2(\Omega)} + \avg\left(\theta_j\right) \\
+	&= \beta + \sum_{j=1}^r c_j \left(\psi_i, \psi_j\right)_{L^2(\Omega)} \\
+	&= \beta + \sum_{j=1}^r c_j S_{ij}
+	\end{align*}
+	In the first line we used the definition of $v_*$ in \eqref{eq:optimal_v}.
+In the second line we used Lemma LEM.
+In the third line we used the facts that, by construction, $\Delta \theta_j = \psi_j$ and $\int_\Omega \theta_j(x) dx = 0$.
+In the fourth line we used the definition of $S$ in \eqref{eq:S_matrix}.
+Hence the condition $v_*(x_i) = f_i$ for $i=1,\dots,r$ may be written in matrix form as
+	\begin{equation*}
+	S \mathbf{c} + \beta \mathbf{1} = \mathbf{f},
+	\end{equation*}
+	and this must hold because it is the first block row of system \eqref{eq:interpolation_linear_system} which is solved to compute $\mathbf{c}$ and $\beta$. So $v_* \in X$.
+
+	\paragraph{Optimality} Since $X$ is an affine subspace, and $X_0$ is the tangent space to $X$, any $v \in X$ may be written in the form $v = v_* + p$ for some $p \in X_0$.
+We may therefore reframe optimization problem \ref{eq:wi_optimization_problem_generic} as
+	\begin{equation}
+	\label{eq:p_optimization}
+		\min_{p \in X_0} J(p),
+	\end{equation}
+	where
+	\begin{equation*}
+		J(p) := \frac{1}{2}||\Delta (v_* + p)||^2.
+	\end{equation*}
+	Showing that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} is equivalent to showing that $0$ is the unique minimizer of \eqref{eq:p_optimization}.
+
+	By direct expansion, we have
+	\begin{equation}
+		\label{eq:Ju_expanded}
+		J(p) = J(0) + \left(\Delta u, \Delta p\right)_{L^2(\Omega)} + \frac{1}{2}||\Delta p||^2.
+	\end{equation}
+	Since $p$ has Neumann zero boundary conditions, if $\Delta p = 0$, then $p$ is constant on each connected component of $\Omega$.
+Furthermore, since $p(x_i)$ is zero for at least one point $x_i$ per connected component of $\Omega$, if $\Delta p = 0$ then $p=0$.
+Hence, by contrapositive we have
+	\begin{align*}
+	\label{eq:contrapositive_laplacian_zero}
+		p \neq 0 \quad \implies \quad& 0 < ||\Delta p||_{L^2(\Omega)}\\
+		\implies \quad& J(0) < J(p) + \left(\Delta v_*, \Delta p\right)_{L^2(\Omega)}.
+	\end{align*}
+	If $v_*$ satisfies the following variational equation,
+	\begin{equation}
+	\label{eq:variational_up}
+		(\Delta v_*, \Delta p)_{L^2(\Omega)} = 0 \quad \text{for all~} p \in X_0,
+	\end{equation} 
+	then
+	\begin{equation*}
+		p \neq 0 \quad \implies \quad J(0) < J(p).
+	\end{equation*}
+	Showing that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} therefore reduces to showing that $v_*$ satisfies variational equation \eqref{eq:variational_up}.
+
+	\paragraph{Variational equation} To show that $v_*$ satiafies variational equation \eqref{eq:variational_up}, notice that
+	\begin{align*}
+	(\Delta v_*, \Delta p)_{L^2(\Omega)} =& \sum_{i=1}^r c_i(\Delta \theta_i, \Delta p)_{L^2(\Omega)} \\
+	=& \sum_{i=1}^r c_i \left(p(x_i) + \avg\left(p\right)\right) \\
+	=& \avg\left(p\right) \sum_{i=1}^r c_i. 
+	\end{align*}
+	In the first line we used the definition of $\theta_i$ and the fact that the Laplacian of a constant is zero.
+In the second line we used Lemma LEM.
+In the third line we used the fact that $p(x_i)=0$ because $p \in X_0$.
+Hence variational equation \eqref{eq:variational_up} holds if $\sum_{i=1}^r c_i = 0$, which may be written in matrix form as $\mathbf{1}^T \mathbf{c} = 0$.
+This must hold because it is the second block row in system \eqref{eq:interpolation_linear_system}.
 \end{proof}

### 2021-09-10 paper/localpsf.tex
added lemma reference

@@
 
+\begin{lem}[Preparation for Theorem \ref{thm:wi_alg_is_correct}]
+	\label{lem:psi_delta_eval}
+	We have
+	\begin{equation}
+	\label{eq:inline_wi_lemma}
+	h(x_j) = \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \avg(h)
+	\end{equation}
+	for all $h\in H^2_N(\Omega)$, where
+	\begin{equation*}
+	\avg(h) :=	 \frac{1}{|\Omega|} \int_\Omega h ~dx,
+	\end{equation*}
+	and $|\Omega| = \int_\Omega 1 ~dx$ is the Lebesgue measure of $\Omega$.
+\end{lem}
+
+\begin{proof}
+	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\psi_j$ onto $h$, and using the Neumann zero boundary conditions for $\psi_j$ and $h$ to eliminate boundary integral terms.
+\end{proof}
+
 \begin{thm}[Solution of optimization problem \ref{eq:wi_optimization_problem_generic}]
@@
 Note that pointwise values of functions in $H_N^2(\Omega)$ are well-defined since the Sobolev imbedding theorem (see, e.g., Theorem 7.19 in \cite{Arbogast}) implies that $H^2(\Omega)$ is continuously imbedded in the space of continuous functions $C^0(\Omega)$ when $d < 4$, and we are considering $d \in \{1,2,3\}$.
-The feasible set $X$ is therefore a closed affine subspace of $H^2_N(\Omega)$, the tangent space $X_0$ is a closed subspace of $H^2_N(\Omega)$, and optimization problem \ref{eq:wi_optimization_problem_generic} is well-defined.
+The feasible set $X$ is therefore a closed affine subspace of $H^2_N(\Omega)$, the tangent space $X_0$ is a closed subspace of $H^2_N(\Omega)$, and optimization problem \eqref{eq:wi_optimization_problem_generic} is well-defined.
 
@@
 	In the first line we used the definition of $v_*$ in \eqref{eq:optimal_v}.
-In the second line we used Lemma LEM.
+In the second line we used Lemma \ref{lem:psi_delta_eval}.
 In the third line we used the facts that, by construction, $\Delta \theta_j = \psi_j$ and $\int_\Omega \theta_j(x) dx = 0$.
@@
 	(\Delta v_*, \Delta p)_{L^2(\Omega)} =& \sum_{i=1}^r c_i(\Delta \theta_i, \Delta p)_{L^2(\Omega)} \\
-	=& \sum_{i=1}^r c_i \left(p(x_i) + \avg\left(p\right)\right) \\
-	=& \avg\left(p\right) \sum_{i=1}^r c_i. 
+	=& \sum_{i=1}^r c_i \left(p(x_i) - \avg\left(p\right)\right) \\
+	=& -\avg\left(p\right) \sum_{i=1}^r c_i. 
 	\end{align*}
 	In the first line we used the definition of $\theta_i$ and the fact that the Laplacian of a constant is zero.
-In the second line we used Lemma LEM.
+In the second line we used Lemma \ref{lem:psi_delta_eval}.
 In the third line we used the fact that $p(x_i)=0$ because $p \in X_0$.
@@
 
-\begin{lem}[Preparation for Theorem \ref{thm:wi_alg_is_correct}]
-	\label{lem:psi_delta_eval}
-	We have
-	\begin{equation}
-	\label{eq:inline_wi_lemma}
-	h(x_j) = \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \avg(h)
-	\end{equation}
-	for all $h\in H^2_N(\Omega)$, where
-	\begin{equation*}
-	\avg(h) :=	 \frac{1}{|\Omega|} \int_\Omega h ~dx,
-	\end{equation*}
-	and $|\Omega| = \int_\Omega 1 ~dx$ is the Lebesgue measure of $\Omega$.
-\end{lem}
-
-\begin{proof}
-	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\psi_j$ onto $h$, and using the Neumann zero boundary conditions for $\psi_j$ and $h$ to eliminate boundary integral terms.
-\end{proof}
-
 \begin{proof}[Proof of Theorem \ref{alg:weighting_functions_incremental}]

### 2021-09-13 paper/localpsf.tex
wrote no-breakdown theorem

@@
 \end{enumerate}
-In Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v_*$ produced by this process indeed solves optimization problem \eqref{eq:wi_optimization_problem_generic}.
+In Theorem \ref{thm:wi_no_breakdown}, we prove that this procedure does not break down, so long as there is no repetition in the points $x_i$.
+Linear system \eqref{eq:interpolation_linear_system} is the KKT system that arises if one solves optimization problem \eqref{eq:wi_optimization_problem_generic} on the $r+1$-dimensional subspace spanned by the functions $\psi_i$, $i=1,\dots,r$, and the constant function $C(x)=1$, rather than solving the optimization problem on the infinite dimensional space $H^2_N(\Omega)$.
+In Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v_*$ produced by this process solves optimization problem \eqref{eq:wi_optimization_problem_generic} on $H^2_N(\Omega)$.
 
@@
 In four or higher spatial dimensions, one could construct appropriate weighting functions using radial basis functions, though we do not discuss this here.
+
+\begin{thm}[Procedure for computing $v_*$ does not break down]
+	If all the sample points $x_i$, $i=1,\dots,r$, are distinct, then PDE \eqref{eq:psi_eq} is uniquely solvable for $\psi_i$, PDE \eqref{eq:theta_eq} is uniquely solvable for $\theta_i$, and linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\mathbf{c}$ and $\beta$.
+That is, the procedure for computing $v_*$ does not break down.
+Furthermore, $v_* \in H^2_N(\Omega)$.
+\end{thm}
+
+\begin{proof}
+	Existence and uniqueness for the solutions to PDEs \eqref{eq:psi_eq} and \eqref{eq:theta_eq} follow from the standard elliptic PDE theory.
+By standard elliptic regularity theory, $\psi_i \in H^1(\Omega)$, so $\psi_i \in L^2(\Omega)$, and so all matrix entries $S_{ij}$ are finite.
+
+	Using the definition of $S$ and basic linear algebra, it is straightforward to show that if $S$ were singular, there would exist constants $a_1, a_2, \dots a_r$, which are not all zero, such that
+	\begin{equation*}
+		0 = a_1 \psi_1 + a_2 \psi_2 + \dots a_r \psi_r.
+	\end{equation*}
+	Applying the Laplacian to this equation yields
+	\begin{equation*}
+		0 = a_1 \delta_{x_1} + a_2 \delta_{x_2} + \dots a_r \delta_{x_r} - \left(\sum_{i=1}^r a_i\right)/|\Omega|,
+	\end{equation*}
+	which is impossible because a nonzero linear combination of delta distributions at different points cannot equal a constant function.
+Thus we must conclude that $S$ is non-singular.
+By block matrix factorization, non-singularity of $S$ implies that linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\mathbf{c}$ and $\beta$.
+
+	By standard elliptic regularity theory, the functions $\theta_i$ are in $H^3(\Omega)$, and therefore they are in $H^2(\Omega)$.
+By construction the functions $\theta_i$ have Neumann zero boundary conditions, so $\theta_i \in H^2_N(\Omega)$.
+Since $v_*$ is a linear combination of the functions $\theta_i$ and the constant function $C(x)=1$, $v_* \in H^2_N(\Omega)$.
+\end{proof}
 

### 2021-09-14 paper/localpsf.tex
removed a lot of commented out old text

--- 2021-09-10 paper/localpsf.tex
@@
 The product-convolution approximation is converted to hierarchical matrix format in order to solve linear systems and perform other matrix operations efficiently.
-Impulse responses are computed by applying the operator to a small number of Dirac comb ``batches'' of point sources.
+Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources.
 The key innovation of our method is a matrix-free procedure for choosing as many points as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
-We apply the method to approximate Hessians in large-scale PDE-constrained inverse problem with highly informative data.
+We apply the method to approximate Hessians in large-scale PDE-constrained inverse problems with highly informative data.
 Numerical results demonstrate that our method substantially outperforms existing state-of-the-art Hessian approximation methods which are based on low-rank approximation.
@@
 The smaller these supports are, the better our algorithm will perform.
-Let $\delta_x$ denote the point source (delta distribution) centered at $x$.
+Let $\delta_x$ denote the delta distribution (point source) centered at $x$.
 The function $\phi_x$ is called the impulse response because straightforward analysis shows that it may be expressed as follows:
@@
 \label{eq:impulse_response_delta_action}
-\phi_x = \left\langle \Aop, \delta_x \right\rangle^*,
-\end{equation}
-where $\left\langle \Aop, \delta_x \right\rangle \in L^2(\Omega)'$ denotes the result of applying the operator $\Aop$ to the distribution $\delta_x$, and $\left\langle \Aop, \delta_x \right\rangle^* \in L^2(\Omega)$ denotes the Riesz representation of $\langle\Aop, \delta_x\rangle$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\psi \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\psi^* \in L^2(\Omega)$ such that $\psi(v) = \left(\psi^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, the action of $\Aop$ may be extended to distributions in a natural way, as described in Appendix \ref{app:distributions}.
-%The action of $\mathcal{A}$ on a distribution $\mu$ is written as $\langle \mathcal{A}, \mu \rangle$.
-We use this angle bracket notation to distinguish the action of $\mathcal{A}$ on distributions from the action of $\mathcal{A}$ on functions.
-%We take the Riesz representative of the linear functional $\left\langle \mathcal{A}, \delta_x \right\rangle \in L^2(\Omega)'$ to convert it into a function $\left\langle \mathcal{A}, \delta_x \right\rangle^* \in L^2(\Omega)$.
-%After discretization, the process of applying $\Aop$ to the distribution $\delta_x$, then forming the Riesz representative of the resulting linear functional, amounts to computing $\boldsymbol{\phi}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x$, where $\boldsymbol{\phi}_x$, $\mathbf{A}$, and $\boldsymbol{\delta}_x$ are discretized versions of $\phi_x$, $A$ and $\delta_x$, respectively, and $\mathbf{M}$ is a discretization of the $L^2$ inner product bilinear form (e.g., a mass matrix in the finite element context).
+\phi_x = \left( \Aop \delta_x^* \right)^*,
+\end{equation}
+where $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying the operator $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\psi \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\psi^* \in L^2(\Omega)$ such that $\psi(v) = \left(\psi^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Phi$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
+This is described in Appendix \ref{app:distributions}.
 From \eqref{eq:impulse_response_delta_action}, we see that this local support property means that the response of $\Aop$ to a point source at $x$ is zero (or small) at points $y$ that are far from $x$.
@@
 More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because $\mathcal{H}^{-1}$ locally approximates the Bayesian posterior covariance for $m$.
-%We only have access to $\mathcal{H}$ via its action on an arbitrary vectors $v$, i.e., evaluation of the map $v \mapsto \mathcal{H} v$.
-Evaluating $v \mapsto \mathcal{H} v$ is an expensive process that involves solving two auxiliary linear PDEs.
 
@@
 
-%First, we compute
-%\begin{equation}
-%\label{eq:eta_intro}
-%\eta := \left\langle \mathcal{A}, \xi \right\rangle^*
-%\end{equation}
-%where 
-%\begin{equation*}
-%	\xi := \sum_{i=1}^l \delta_{x_i}
-%\end{equation*}
-%is the Dirac comb associated with point sources centered at a collection of scattered points $x_i$.
-The batch of points $x_i$ are chosen so that the support of $\phi_i$ and the support of $\phi_j$ do not overlap (or do not overlap much) if $i \neq j$.
-By linearity, and the expression for $\phi_x$ in \eqref{eq:impulse_response_delta_action}, we have
-%\begin{equation*}
-%\eta = \sum_{i=1}^l \phi_i.
-%\end{equation*}
-%We may therefore post-process $\eta$ 
-%
-%to recover $\phi_i$ for all of the points $x_i$ in the batch.
-Since the functions $\varphi_i$ are translated versions of the functions $\phi_i$, we also recover $\varphi_i$ for all points $x_i$ in the batch.
-This is illustrated in Figure FIG.
-The process is repeated to get more batches of $\varphi_i$'s, until a desired number of batches is reached.
-
 In order to choose the points $x_i$, we need to estimate the support of $\phi_i$ \emph{before} we compute it.
@@
 With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
-%This is computationally prohibitive in our applications, because computing the functions $\varphi_i$ requires one matrix vector product with $\Aop$ for each point $x_i$, and these matrix vector products are computationally costly.
 In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
@@
 In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\varphi_i$ are computed with each matrix vector product.
-
-%We do this by taking advantage of the fact that the functions $\varphi_i$ are locally supported; existing methods typically do not require this local support property.
 
@@
 The complete algorithm for constructing $\AmatPc$ is shown in Algorithm \ref{alg:construct_Atilde}.
-%The method for evaluating kernel entries of $\Phi_\text{pc}$ is shown in Section \ref{sec:eval_matrix_entries}.
-
-%\begin{itemize}
-%\item We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}).
-To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as distinct functions.
-%\item We use a greedy algorithm to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}).
-%\item We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\psi_x$ (Section \ref{eq:mean_and_covariance_estimation}). 
-%\item For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
-%\end{itemize}
-%The complete algorithm for constructing $\AopPc$ is shown in Algorithm \ref{alg:construct_Atilde}.
-The method for evaluating kernel entries of $\Phi_\text{pc}$ is shown in Section \ref{sec:eval_matrix_entries}.
-The method for converting $\AopPc$ to H-matrix format is shown in Section \ref{sec:H_matrix_conversion}.
-The method for modifying the resulting H-matrix to make it symmetric positive semi-definite is shown in Section \ref{sec:make_spd}.
 
@@
 
-			Compute $\eta^b$ by applying $\Aop$ to the dirac comb for $S_b$ (Section \ref{sec:get_impulse_response})
+			Compute $\combresponse_b$ by applying $\Aop$ to the dirac comb for $S_b$ (Section \ref{sec:get_impulse_response})
 
@@
 
-		Compute all weighting functions $w_i$ using Algorithm \ref{alg:weighting_functions_incremental}
-%		In lines 2 and 3 of Algorithm \ref{alg:weighting_functions_incremental}, only solve for $\psi_i$ and $\theta_i$ associated with new points $x_i$ in $S_b$.
-The functions $\psi_i$ and $\theta_i$ associated with points in previous batches have already been computed in previous iterations of this for loop.
+		Compute all weighting functions $w_i$ using Algorithm
 
@@
 The support of $\phi_x$ is estimated to be contained within an ellipsoid, so the process of choosing batches of points is an ellipsoid packing problem (Section \ref{sec:greedy_point_selection}).
-%In the next section (Section \ref{sec:get_impulse_response}), we will extract $\varphi_{x_i}$ for all the points $x_i$ in a given batch by applying $\Aop$ to the Dirac comb associated with that batch.
 
@@
 where $\tau$ is a fixed constant.
-%The ellipsoid $E_{x}$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian approximation of the normalized version of $\langle \mathcal{A}, \delta_{x}\rangle^{*}$ which has the same mean and covariance.
 The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\phi}_x$ (see Figure FIG).
@@
 The fraction of the ``mass'' of $\phi_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.
-% a good balance between these competing goals. 
-%The bigger $\tau$ is, the smaller the overlap between impulse responses within a batch will be, while the smaller $\tau$ is, the more impulse responses we can fit within each batch.
 
@@
 We present this efficient method of computing $\alpha$, $\mu$, and $\Sigma$ in Theorem \ref{thm:vol_mean_cov} and summarize the method in Algorithm \ref{alg:varhpi_mean_cov}.
-
-%In Algorithm \ref{alg:varhpi_mean_cov} we show how to compute  Definitions required for Algorithm \ref{alg:varhpi_mean_cov} are given in Definition \ref{defn:alpha_rho_mu_sigma}, Definition \ref{defn:C_L_Q}, and Definition \ref{defn:pointwise}.
-The theoretical basis for Algorithm \ref{alg:varhpi_mean_cov} is shown in Theorem \ref{thm:vol_mean_cov}.
-
-%We compute the scaling factor, denoted $\alpha(x)$, for all $x$ simultaneously by applying $\mathcal{A}^T$ to a constant function.
-We compute the spatially varying mean, $\mu(x)$, and covariance, $\Sigma(x)$ of the normalized version of $\phi_x$, for all $x$ simultaneously, by applying $\mathcal{A}^T$ to a small number of linear and quadratic functions.
-Definitions of $\alpha$, $\mu$, and $\Sigma$ are given in Definition \ref{defn:alpha_rho_mu_sigma}.
-The procedure for computing these quantities is shown in Algorithm \ref{alg:varhpi_mean_cov}, and the theoretical basis for this procedure is shown in Theorem \ref{thm:vol_mean_cov}.
 
@@
 
-%
-%\begin{defn}[Constant, linear, and quadratic functions]	
-%	\label{defn:C_L_Q}
-%	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$.
-We define $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions, respectively:
-%	\begin{equation*}
-%	C(x) := 1, \qquad
-%	L^i(x) := x^i, \qquad
-%	Q^{ij}(x) = x^i x^j.
-%	\end{equation*}
-%\end{defn}
-%
-%\begin{defn}[Pointwise operations]
-%	\label{defn:pointwise}
-%	We write $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
-%\end{defn}
-
-%\begin{theorem}
-%	\label{thm:vol_mean_cov}
-%	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$, and let $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions, respectively:
-%	\begin{equation*}
-%		C(x) := 1, \qquad
-%		L^i(x) := x^i, \qquad
-%		Q^{ij}(x) = x^i x^j.
-%	\end{equation*}
-%	We have
-%	\begin{align}
-%		\alpha =& \left(\mathcal{A}^TC\right)^*, \label{eq:vol_mean_var_thm1}\\
-%		\mu^i =& \left(\mathcal{A}^T L^i\right)^* / \alpha, \label{eq:vol_mean_var_thm2}\\
-%		\Sigma^{ij} =& \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot\mu^j, \label{eq:vol_mean_var_thm3}
-%	\end{align}
-%	We write $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
-%\end{theorem}
-%
-%The proof of Theorem \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}.
-
 \begin{algorithm2e}
@@
 Candidate points for the first batch are chosen arbitrarily from $P$.
-Once $p$ is selected, we remove $p$ from $P$. If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in S_b$,
-%and if $p$ is sufficiently far from the boundary, in the sense that $E_p \cap \partial \Omega = \{\}$, 
-then we add $p$ to $S_b$. Otherwise we discard $p$. 
-%Recall that $E_p$ and $E_q$ are ellipsoids associated with points $p$ and $q$, respectively, as defined in \eqref{eq:support_ellipsoid}. 
+Once $p$ is selected, we remove $p$ from $P$.
+If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in S_b$, then we add $p$ to $S_b$.
+Otherwise we discard $p$.
 This process repeats until there are no more points in $P$.
@@
 Each application of $\Aop$ to a Dirac comb yields one batch of impulse responses, $\phi_{x_i}$, which yields one batch of convolution kernels, $\varphi_{x_i}$.
-%The batches of impulse responses, $\eta^b$, will ultimately be interpolated with weighting functions described in Section \ref{sec:weighting_functions} to form the desired approximation $\widetilde{\mathcal{A}}\approx \mathcal{A}$.
-The Dirac comb, $\xi_b$, associated with a batch of sample points, $S_b$, is the sum of Dirac distributions (point sources) centered at the points $x_i \in S_b$. That is,
-\begin{equation*}
-	\xi^b := \sum_{x_i \in S_b} \delta_{x_i}.
+The Dirac comb, $\dcomb_b$, associated with a batch of sample points, $S_b$, is the sum of Dirac distributions (point sources) centered at the points $x_i \in S_b$.
+That is,
+\begin{equation*}
+	\dcomb_b := \sum_{x_i \in S_b} \delta_{x_i}.
 \end{equation*}
@@
 \label{eq:dirac_comb_H_action}
-	\eta^b := \langle\mathcal{A}, \xi^b\rangle^*.
-\end{equation}
-By linearity, the resulting function $\eta^b$ may be written as
+	\combresponse_b := \left(\mathcal{A} \dcomb_b^*\right)^*.
+\end{equation}
+By linearity, the resulting function $\combresponse_b$ may be written as
 \begin{equation}
 \label{eq:phi_b}
-	\eta^b = \left\langle\mathcal{A},\sum_{x_i \in S_b} \delta_{x_i}\right\rangle^* = \sum_{x_i \in S_b} \langle\mathcal{A}, \delta_{x_i}\rangle^* = \sum_{x_i \in S_b} \phi_{x_i}.
+	\combresponse_b = \left\langle\mathcal{A},\sum_{x_i \in S_b} \delta_{x_i}\right\rangle^* = \sum_{x_i \in S_b} \langle\mathcal{A}, \delta_{x_i}\rangle^* = \sum_{x_i \in S_b} \phi_{x_i}.
 \end{equation}
@@
 	\varphi_{x_i}(z) = \phi_i(z+x_i) = \begin{cases}
-		\eta^b(z+x_i), & z+x_i \in E_{x_i} \\
+		\combresponse_b(z+x_i), & z+x_i \in E_{x_i} \\
 		0, & \text{otherwise}
@@
 To convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), we will need to evaluate $\varphi_{x_i}(z)$ for a large number of points $z$.
-Per \eqref{eq:varphi_eval}, evaluating $\varphi_{x_i}(z)$ requires evaluating $\eta_b(z+x_i)$.
-In computations, $\eta^b$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh.
-Evaluating $\eta^b(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell.
+Per \eqref{eq:varphi_eval}, evaluating $\varphi_{x_i}(z)$ requires evaluating $\combresponse_b(z+x_i)$.
+In computations, $\combresponse_b$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh.
+Evaluating $\combresponse_b(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell.
 Although this procedure is asymptotically scalable, it is slow in practice \cite{FEMEVALSLOW}.
@@
 
-%To minimize memory usage, the functions $\varphi_i$ are not stored individually.
-Instead, we store the functions $\eta^b$, and use formula \eqref{eq:varphi_eval} to perform function evaluations $z \mapsto \varphi_i(z)$ as needed.
-
 \subsubsection{Addressing boundary artifacts}
@@
 Kernels associated with nearer sample points are given precedence over kernels associated with farther away sample points.
-%After the convolution kernel has been extended using information from neighboring convolution kernels, it is then further extended by zero.
 
@@
 
-%For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$.
-Constructing the weighting functions requires solving two Poisson PDEs per sample point $x_i$.
-
 We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to the following optimization problem:
@@
 \min_{v \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta v}_{L^2(\Omega)}^2 \\
-\text{such that} &\quad v(x_j) = f_j, \quad j=1,\dots,r,
+\text{such that} &\quad v(x_j) = \interpolatedvalues_j, \quad j=1,\dots,r,
 \end{aligned}
 \end{equation}
-where $f_j \in \mathbb{R}$, $j=1,\dots,r$ are now arbitrary function values that $v$ must take at the points $x_j$.
+where $\interpolatedvalues_j \in \mathbb{R}$, $j=1,\dots,r$ are now arbitrary function values that $v$ must take at the points $x_j$.
 We may efficiently solve optimization problem \eqref{eq:wi_optimization_problem_generic} via the following procedure:
@@
 Here $\delta_{x_i}$ is the point source centered at $x_i$, $|\Omega|$ is the measure of the domain $\Omega$, and $\nu$ is the normal vector to $\partial \Omega$.
-\item For each function $\psi_i$, we compute a function $\theta_i$ by solving another Neumann Poisson problem:
+\item For each function $\psi_i$, we compute a function $\secondgreens_i$ by solving another Neumann Poisson problem:
 \begin{equation}
@@
 \begin{cases}
-\Delta \theta_i = \psi_i, & \text{in }\Omega, \\
-\nu \cdot \nabla \theta_i = 0 & \text{on } \partial \Omega,
+\Delta \secondgreens_i = \psi_i, & \text{in }\Omega, \\
+\nu \cdot \nabla \secondgreens_i = 0 & \text{on } \partial \Omega,
 \end{cases}
 \end{equation}
-with condition $\int_\Omega \theta_i(x) dx = 0$.
-%\item For each sample point $x_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\delta_{x_i}$ (a point source centered at $x_i$) as the right hand side source.
-We call the solution to this first Poisson problem $\psi_i$.
-%\item For each function $\psi_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\psi_i$ as the right hand side source.
-We call the solution to this second Poisson problem $\theta_i$.
+with condition $\int_\Omega \secondgreens_i(x) dx = 0$.
 \item The solution to optimization problem \eqref{eq:wi_optimization_problem_generic}, which we denote by $v_*$, is formed as follows:
@@
 \label{eq:optimal_v}
-v_*(x) := \beta + \sum_{i=1}^r c_i \theta_i(x),
+v_*(x) := \beta + \sum_{i=1}^r c_i \secondgreens_i(x),
 \end{equation}
@@
 	\begin{bmatrix}
-	\mathbf{f} \\ 0
+	\mathbf{\interpolatedvalues} \\ 0
 	\end{bmatrix}.
 \end{equation}
-Here $\mathbf{c} \in \mathbb{R}^r$ is the vector with $i$th component $c_i$, $\mathbf{1} \in \mathbf{R}^r$ is the column vector of all ones, $\mathbf{f} \in \mathbb{R}^r$ is the vector with $i$th component $f_i$, and $S \in \mathbb{R}^{r \times r}$ is the matrix with entries given by
+Here $\mathbf{c} \in \mathbb{R}^r$ is the vector with $i$th component $c_i$, $\mathbf{1} \in \mathbf{R}^r$ is the column vector of all ones, $\mathbf{\interpolatedvalues} \in \mathbb{R}^r$ is the vector with $i$th component $\interpolatedvalues_i$, and $S \in \mathbb{R}^{r \times r}$ is the matrix with entries given by
 \begin{equation}
 \label{eq:S_matrix}
-S_{ij} := \int_\Omega \psi_i(x) \psi_j(x) dx.
+S_{ij} := \int_\Omega \firstgreens_i(x) \firstgreens_j(x) dx.
 \end{equation}
 \end{enumerate}
-In Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v_*$ produced by this process indeed solves optimization problem \eqref{eq:wi_optimization_problem_generic}.
-
-Constructing all $r$ of the functions $w_i$ thus requires solving the Poisson equation $2r$ times with different right hand side sources to form the functions $\psi_i$ and $\theta_i$, then post-processing these functions.
+In Theorem \ref{thm:wi_no_breakdown}, we prove that this procedure does not break down, so long as there is no repetition in the points $x_i$.
+Linear system \eqref{eq:interpolation_linear_system} is the KKT system that arises if one solves optimization problem \eqref{eq:wi_optimization_problem_generic} on the $r+1$-dimensional subspace spanned by the functions $\firstgreens_i$, $i=1,\dots,r$, and the constant function $C(x)=1$, rather than solving the optimization problem on the infinite dimensional space $H^2_N(\Omega)$.
+In Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v_*$ produced by this process solves optimization problem \eqref{eq:wi_optimization_problem_generic} on $H^2_N(\Omega)$.
+
+Constructing all $r$ of the functions $w_i$ thus requires solving the Poisson equation $2r$ times with different right hand side sources to form the functions $\firstgreens_i$ and $\secondgreens_i$, then post-processing these functions.
 This is shown in Algorithm \ref{alg:weighting_functions_incremental}.
 The Poisson PDE solves, which are the most expensive part of this procedure, can be performed cheaply with multigrid.
-
-%All of the weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added.
-However, the dominant computational cost in constructing the weighting functions is the solution of the Poisson PDEs, and these Poisson PDE solves can be performed incrementally---we only need to compute functions $\psi_i$ and $\theta_i$ associated with the new points.
-%, rather than $2(r+1)$ new PDE solves.
 
@@
 
+\begin{thm}[Procedure for computing $v_*$ does not break down]
+	\label{thm:wi_no_breakdown}
+	If all the sample points $x_i$, $i=1,\dots,r$, are distinct, then PDE \eqref{eq:psi_eq} is uniquely solvable for $\firstgreens_i$, PDE \eqref{eq:theta_eq} is uniquely solvable for $\secondgreens_i$, and linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\mathbf{c}$ and $\beta$.
+That is, the procedure for computing $v_*$ does not break down.
+Furthermore, $v_* \in H^2_N(\Omega)$.
+\end{thm}
+
+\begin{proof}
+	Existence and uniqueness for the solutions to PDEs \eqref{eq:psi_eq} and \eqref{eq:theta_eq} follow from the standard elliptic PDE theory.
+By standard elliptic regularity theory, $\firstgreens_i \in H^1(\Omega)$, so $\firstgreens_i \in L^2(\Omega)$, and so all matrix entries $S_{ij}$ are finite.
+
+	Using the definition of $S$ and basic linear algebra, it is straightforward to show that if $S$ were singular, there would exist constants $a_1, a_2, \dots a_r$, which are not all zero, such that
+	\begin{equation*}
+		0 = a_1 \firstgreens_1 + a_2 \firstgreens_2 + \dots a_r \firstgreens_r.
+	\end{equation*}
+	Applying the Laplacian to this equation yields
+	\begin{equation*}
+		0 = a_1 \delta_{x_1} + a_2 \delta_{x_2} + \dots a_r \delta_{x_r} - \left(\sum_{i=1}^r a_i\right)\bigg/|\Omega|,
+	\end{equation*}
+	which is impossible because a nonzero linear combination of delta distributions at different points cannot equal a constant function.
+Thus we must conclude that $S$ is non-singular.
+By block matrix factorization, non-singularity of $S$ implies that linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\mathbf{c}$ and $\beta$.
+
+	By standard elliptic regularity theory, the functions $\secondgreens_i$ are in $H^3(\Omega)$, and therefore they are in $H^2(\Omega)$.
+By construction the functions $\secondgreens_i$ have Neumann zero boundary conditions, so $\secondgreens_i \in H^2_N(\Omega)$.
+Since $v_*$ is a linear combination of the functions $\secondgreens_i$ and the constant function $C(x)=1$, $v_* \in H^2_N(\Omega)$.
+\end{proof}
+
 \begin{lem}[Preparation for Theorem \ref{thm:wi_alg_is_correct}]
@@
 	\label{eq:inline_wi_lemma}
-	h(x_j) = \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \avg(h)
+	h(x_j) = \left(\firstgreens_j, \Delta h\right)_{L^2(\Omega)} + \avg(h)
 	\end{equation}
@@
 \begin{proof}
-	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\psi_j$ onto $h$, and using the Neumann zero boundary conditions for $\psi_j$ and $h$ to eliminate boundary integral terms.
+	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\firstgreens_j$ onto $h$, and using the Neumann zero boundary conditions for $\firstgreens_j$ and $h$ to eliminate boundary integral terms.
 \end{proof}
@@
 	\begin{equation*}
-		X := \{v \in H^2_N(\Omega) : v(x_i) = v_i \text{ for }i=1,\dots,r\}
+		X := \{v \in H^2_N(\Omega) : v(x_i) = \interpolatedvalues_i \text{ for }i=1,\dots,r\}
 	\end{equation*}
@@
 
-	\paragraph{Feasibility} To show that $v_*$ is feasible (i.e., $v_* \in X$), we must show that $v_*(x_i) = f_i$ for $i=1,\dots,r$. We have
-	\begin{align*}
-	v_*(x_i) &= \beta + \sum_{j=1}^r c_j \theta_j(x_i) \\
-	&= \beta + \sum_{j=1}^r c_j \left(\psi_i, \Delta \theta_j\right)_{L^2(\Omega)} + \avg\left(\theta_j\right) \\
-	&= \beta + \sum_{j=1}^r c_j \left(\psi_i, \psi_j\right)_{L^2(\Omega)} \\
+	\paragraph{Feasibility} To show that $v_*$ is feasible (i.e., $v_* \in X$), we must show that $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,r$. We have
+	\begin{align*}
+	v_*(x_i) &= \beta + \sum_{j=1}^r c_j \secondgreens_j(x_i) \\
+	&= \beta + \sum_{j=1}^r c_j \left(\firstgreens_i, \Delta \secondgreens_j\right)_{L^2(\Omega)} + \avg\left(\secondgreens_j\right) \\
+	&= \beta + \sum_{j=1}^r c_j \left(\firstgreens_i, \firstgreens_j\right)_{L^2(\Omega)} \\
 	&= \beta + \sum_{j=1}^r c_j S_{ij}
@@
 In the second line we used Lemma \ref{lem:psi_delta_eval}.
-In the third line we used the facts that, by construction, $\Delta \theta_j = \psi_j$ and $\int_\Omega \theta_j(x) dx = 0$.
+In the third line we used the facts that, by construction, $\Delta \secondgreens_j = \firstgreens_j$ and $\int_\Omega \secondgreens_j(x) dx = 0$.
 In the fourth line we used the definition of $S$ in \eqref{eq:S_matrix}.
-Hence the condition $v_*(x_i) = f_i$ for $i=1,\dots,r$ may be written in matrix form as
-	\begin{equation*}
-	S \mathbf{c} + \beta \mathbf{1} = \mathbf{f},
+Hence the condition $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,r$ may be written in matrix form as
+	\begin{equation*}
+	S \mathbf{c} + \beta \mathbf{1} = \mathbf{\interpolatedvalues},
 	\end{equation*}
@@
 	\begin{align*}
-	(\Delta v_*, \Delta p)_{L^2(\Omega)} =& \sum_{i=1}^r c_i(\Delta \theta_i, \Delta p)_{L^2(\Omega)} \\
+	(\Delta v_*, \Delta p)_{L^2(\Omega)} =& \sum_{i=1}^r c_i(\Delta \secondgreens_i, \Delta p)_{L^2(\Omega)} \\
 	=& \sum_{i=1}^r c_i \left(p(x_i) - \avg\left(p\right)\right) \\
@@
 	\end{align*}
-	In the first line we used the definition of $\theta_i$ and the fact that the Laplacian of a constant is zero.
+	In the first line we used the definition of $\secondgreens_i$ and the fact that the Laplacian of a constant is zero.
 In the second line we used Lemma \ref{lem:psi_delta_eval}.
@@
 	\label{thm:smoothest_interpolant}
-Let $y \in \mathbb{R}^d$, $d \in \{1,2,3\}$, and let $\gamma_y$ be the function
+Let $y \in \mathbb{R}^d$, $d \in \{1,2,3\}$, and let $\horizinterpolant_y$ be the function
 \begin{equation}
 \label{eq:u_defn_in_thm}
-\gamma_y(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \AkerPc\left(y+x,x\right),
-\end{equation}
-where the weighting functions $w_i$ are the solutions to optimization problem \eqref{eq:wi_optimization_problem}. Then $\gamma_y$ solves the optimization problem
+\horizinterpolant_y(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \AkerPc\left(y+x,x\right),
+\end{equation}
+where the weighting functions $w_i$ are the solutions to optimization problem \eqref{eq:wi_optimization_problem}.
+Then for each $y \in \Omega$, the function $\horizinterpolant_y$ solves the optimization problem
 \begin{equation}
@@
 \begin{aligned}
-\min_{\gamma_y \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \gamma_y}_{L^2(\Omega)}^2 \\
-\text{such that} &\quad \gamma_y(x_i) = \varphi_i(y), \quad i=1,\dots,q.
+\min_{\horizinterpolant_y \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \horizinterpolant_y}_{L^2(\Omega)}^2 \\
+\text{such that} &\quad \horizinterpolant_y(x_i) = \varphi_i(y), \quad i=1,\dots,q.
 \end{aligned}
@@
 
-%		Initialize constant function $C(x) = 1$
-
-%		Compute Lebesgue measure of the domain, $|\Omega|$
-
-%		\tcp{Compute Poisson impulse responses $\{\psi_i\}_{i=1}^r$ and Poisson squared impulse responses $\{\theta_i\}_{i=1}^r$}
-
 		\For{$k=1,2, \dots, r$}{
-%			Initialize $\delta_{x_i}$ as the point source (delta distribution) centered at $x_i$.
-
-			Solve the Neumann Poisson problem \eqref{eq:psi_eq} for $\psi_k$.
-
-%			Compute $\psi_i$ by solving the Neumann Poisson problem
-%			\begin{equation}
-%			\label{eq:psi_eq}
-%			\begin{cases}
-%			\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
-%			\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
-%			\end{cases}
-%			\end{equation}
-%			with condition $\int_\Omega \psi_i(x) dx = 0$, where $\nu$ is the normal vector to $\partial \Omega$.
-
-			Solve the Neumann Poisson problem \eqref{eq:theta_eq} for $\theta_k$.
-
-%			Compute $\theta_i$ by solving the Neumann Poisson problem
-%			\begin{equation}
-%			\label{eq:theta_eq}
-%						\begin{cases}
-%			\Delta \theta_i = \psi_i, & \text{in }\Omega, \\
-%			\nu \cdot \nabla \theta_i = 0 & \text{on } \partial \Omega,
-%			\end{cases}
-%			\end{equation}
-%			with condition $\int_\Omega \theta_i(x) dx = 0$, where $\nu$ is the normal vector to $\partial \Omega$.
+
+			Solve the Neumann Poisson problem \eqref{eq:psi_eq} for $\firstgreens_k$.
+
+			Solve the Neumann Poisson problem \eqref{eq:theta_eq} for $\secondgreens_k$.
 
@@
 			\begin{equation*}
-			w_i(x) = \beta^{(i)} + \sum_{k=1}^r c_k^{(i)} \theta_k(x).
+			w_i(x) = \beta^{(i)} + \sum_{k=1}^r c_k^{(i)} \secondgreens_k(x).
 			\end{equation*}
@@
 
-%\subsection{Evaluating kernel entries of the approximation}
-%\label{sec:eval_matrix_entries}
-%
-%We now describe how to efficiently evaluate $\widetilde{\Phi}(y,x)$ at an arbitrary point $(x,y) \in \Omega \times \Omega$. From \eqref{eq:kernel_entries}, we have
-%\begin{equation}
-%\label{eq:eval_y_x_000}
-%\AkerPc(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x),
-%\end{equation}
-%and from \eqref{eq:varphi_eval}, we have
-%\begin{equation}
-%\label{eq:kernel_entry_formua}
-%\varphi_{x_i}(y-x) = \begin{cases}
-%\eta^{b(i)}(y - x + x_i), & y - x + x_i \in E_{x_i} \\
-%0, & \text{otherwise},
-%\end{cases}
-%\end{equation}
-%where $b(i)$ denotes the batch index such that $x_i \in S^{b(i)}$.
-To compute $z=\AkerPc(y,x)$, we start with $z=0$, then loop through all batches $b$, and all sample points in each batch.
-For each sample point, we add one term in the sum in \eqref{eq:eval_y_x} to $z$, using \eqref{eq:kernel_entry_formua} as needed to evaluate $\varphi_{x_i}$.
-This is shown in Algorithm \ref{alg:eval_y_x}.
-%
-%
-%\begin{algorithm2e}
-%	\SetAlgoNoLine
-%	\SetKwInOut{Input}{Input}
-%	\SetKwInOut{Output}{Output}
-%	
-%	\Input{points $x \in \Omega$, $y \in \Omega$, \\sample point batches $S_1, \dots, S_\text{max\_batches}$, \\Dirac comb responses $\{\eta^b\}_{b=1}^{\text{max\_batches}}$, \\weighting functions $\{w_i\}_{i=1}^r$}
-%	\Output{Kernel entry $z = \widetilde{A}(y,x)$}
-%
-%	$z \gets 0$
-%	
-%	\For{$b=1,2,\dots,\text{}max\_batches$}{
-%		\For{$x_i \in S_b$}{
-%			\If{$y - x + x_i \in E_{x_i}$}{
-%				$z \gets z + w_i(x)\eta^b(y-x+x_i)$
-%			}
-%		}
-%	}	
-%	\caption{Compute approximate kernel entry $z = \widetilde{A}(y,x)$}
-%	\label{alg:eval_y_x}
-%\end{algorithm2e}
-
 \subsection{Discretization and hierarchical matrix construction}
@@
 That is, $\AopPcMesh : V_h \rightarrow V_h'$.
-%Error in the approximation $\Aop \approx \AopPcMesh$ is a combination of error in the product-convolution approximation, and error due to finite element discretization.
 Let $u_h \in V_h$ and $v_h \in V_h$, and let $\mathbf{u}$ and $\mathbf{v}$ be the coefficient vectors for $u_h$ and $v_h$, respectively, with respect to the finite element basis $f_1, \dots, f_N$.
@@
 \end{equation}
-and similar for $v_h$. Further, let $\mathbf{M} \in \mathbb{R}^{N \times N}$ be the finite element mass matrix,
-% which has entries
-%\begin{equation}
-%\label{eq:mass_matrix_defn}
-%\mathbf{M}_{ij} := \int_\Omega f_i(x) f_j(x) dx,
-%\end{equation}
-and let $\AkerPcMat \in \mathbb{R}^{N \times N}$ be following matrix of approximate kernel entries:
+and similar for $v_h$.
+Further, let $\mathbf{M} \in \mathbb{R}^{N \times N}$ be the finite element mass matrix, and let $\AkerPcMat \in \mathbb{R}^{N \times N}$ be following matrix of approximate kernel entries:
 \begin{equation}
@@
 \end{equation*}
-% PUT INTO APPENDIX!!!!
-%\begin{align*}
-%\left(\AopPcMesh u_h \right)(v_h) =& \int_\Omega \int_\Omega v_h(y) \AkerPcMesh(y,x) u_h(x) dx dy \\
-%=& \int_\Omega \int_\Omega \left(\sum_{i=1}^N \mathbf{v}_i f_i(y)\right) \left(\sum_{j=1}^N \sum_{k=1}^N \AkerPc(p_j,p_k) f_j(y) f_k(x)\right) \left(\sum_{l=1}^N \mathbf{u}_l f_l(x)\right) dx dy \\
-%=& \sum_{i,j,k,l=1}^N \mathbf{v}_i \left(\int_\Omega f_i(y) f_j(y) dy\right) \AkerPcMesh(p_j, p_k) \left(\int_\Omega f_k(x) f_l(x) dx\right) \mathbf{u}_l \\
-%=& \sum_{i,j,k,l=1}^N \mathbf{v}_i \mathbf{M}_{ij} \left(\AkerPcMat\right)_{jk} \mathbf{M}_{kl} \mathbf{u}_l \\
-%=& \mathbf{v}^T \mathbf{M} \AkerPcMat \mathbf{M} \mathbf{u},
-%\end{align*}
 The finite element matrix representation of $\AopPcMesh$, which we denote by $\AmatPc$, is therefore given by
@@
 For more details on H-matrices, we recommend \cite{HMATRIXGOOD}.
-%We discuss H-matrices in greater detail in Appendix \ref{app:h_matrix}.
 
@@
 
-c\begin{itemize}
+\begin{itemize}
 	\item computational complexity in terms of PDE solves:
@@
 \end{itemize}
-
-\section{Rational positive definite modification}
-WRITE THIS
 
@@
 
-%The H-matrix construction method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix (here, these locations are the finite element Lagrange nodes $p_i$, $i=1,\dots,N$) to construct a recursive block partitioning of $\AkerPcMat$, such that all blocks at the finest level of the partitioning are either expected to be low rank, or are small.
-Then, the method forms low rank approximations of the blocks that are expected to be low rank by sampling specially chosen rows and columns of those blocks.
-The small high rank blocks are formed directly by evaluating their entries.
-
 In detail, the process of constructing an H-matrix representation of $\AkerPcMat$ proceeds as follows.
@@
 \label{eq:weak_admissibility_cond}
-\dist\left(r_1, c_2\right) \ge \eta \min\left(\diam\left(r_1\right), \diam\left(c_2\right)\right),
+\dist\left(r_1, c_2\right) \ge \weakadmconst \min\left(\diam\left(r_1\right), \diam\left(c_2\right)\right),
 \end{equation}
@@
 The quantity $\diam\left(r_1\right)$ is the diameter of the axis aligned bounding box for the point cloud associated with the row cluster $r_1$, and $\diam\left(c_2\right)$ is the analogous diameter associated with the column cluster $c_2$.
-The quantity $\eta$ is a scalar constant; we use $\eta=2.0$.
+The quantity $\weakadmconst$ is a scalar constant; we use $\weakadmconst=2.0$.
 Basically, if the point clouds associated with $r_1$ and $c_2$ are far away from each other relative to their diameters, then we expect that the corresponding block of the matrix will be low rank.
@@
 
-%These results rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}.
-We prove correctness of Algorithm \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}.
-Theorem \ref{thm:wi_alg_is_correct} relies on a prerequisite result which we present in Lemma \ref{lem:psi_delta_eval}.
-
-\begin{defn}[Weighting function optimization problem]
-	Let $\Omega \subset \mathbb{R}^d$, $d\in\{1,2,3\}$, be an open bounded domain with, let $x_i \in \Omega$ for $i=1,\dots,r$, and let $f \in \mathbb{R}^r$. We consider the optimization problem
-	\begin{equation}
-	\label{eq:generic_optimization_initial}
-	\begin{aligned}
-	\min_{\gamma \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \gamma}_{L^2(\Omega)}^2 \\
-	\text{such that} &\quad \gamma(x_i) = f_i, \quad i=1,\dots,r.
-	\end{aligned}
-	\end{equation}
-	A minimizer, $\gamma^*$, of \eqref{eq:generic_optimization_initial} is a function defined on $\Omega$ that interpolates the data pairs $(x_i,f_i)$, and is as flat as possible in a least squares sense.
-	We denote the objective function for optimization problem \eqref{eq:generic_optimization_initial} by 
-	\begin{equation}
-	\label{eq:wi_objective}
-	J(\gamma) := \frac{1}{2}||\Delta \gamma||^2_{L^2(\Omega)}.
-	\end{equation}
-	We denote the feasible set for optimization problem \eqref{eq:generic_optimization_initial} by 
-	\begin{equation}
-	\label{eq:wi_feasible_set}
-		X := \{u \in H^2_N : u(x_i)=f_i~\text{for}~ i=1,\dots,r\}.
-	\end{equation}
-	We denote the tangent space to the feasible set by
-	\begin{equation}
-	\label{eq:wi_tangent_space}
-	X_0 := \{u \in H^2_N : u(x_i)=0~\text{for}~ i=1,\dots,r\}.
-	\end{equation}
-\end{defn}
-
-\begin{lem}[Elliptic regularity on $X_0$]
-	\label{lem:elliptic_regularity_x0}
-	If $\partial \Omega$ is $C^{1,1}$, then there exists a constant $C > 0$ such that
-	\begin{equation*}
-		||u||_{H^2(\Omega)} \le C ||\Delta u||_{L^2(\Omega)}
-	\end{equation*}
-	for all $u \in X_0$.
-\end{lem}
-
-\begin{proof}
-	Let $\widetilde{u}$ be version of $u$ that has been shifted up or down by a constant such that $\int_\Omega \overline{u} = 0$. We have
-	\begin{align*}
-		||D^1 u||_{\left(L^2(\Omega)\right)^d}^2 + ||D^2 u||_{\left(L^2(\Omega)\right)^{d \times d}}^2 &= ||D^1 \widetilde{u}||_{\left(L^2(\Omega)\right)^d}^2 + ||D^2 \widetilde{u}||_{\left(L^2(\Omega)\right)^{d \times d}}^2\\
-		&\le ||\widetilde{u}||_{H^2(\Omega)}^2 \\
-		&\le C_1 ||\Delta \widetilde{u}||_{L^2(\Omega)}^2 \\
-		&= C_1 ||\Delta u||_{L^2(\Omega)}^2
-	\end{align*}
-	for some constant $C_1$.
-In the first and fourth lines we used the fact that the derivative of a constant is zero, in the second line we used the definition of the $H^2$ norm, and in the third line we used the standard elliptic regularity result for pure Neumann problems (see, e.g., Theorem 8.13 (b) in \cite{Arbogast}).
-
-	To complete the proof, we will now show that there exists a constant $C_2$ such that $||u||_{L^2(\Omega)}^2 \le C_2 ||\Delta u||_{L^2(\Omega)}^2$ for all $u \in X_0$.
-For this, we use a standard compactness argument, which closely follows the proof of Theorem 8.12 in \cite{ARBOGAST}.
-Suppose for the sake of contradiction that this result does not hold.
-Then there exists a sequence $\{u_k\}_{k=1}^\infty \subset X_0$ such that
-	$||u_k||_{L^2(\Omega)}=1$ and $||\Delta u_k||_{L^2(\Omega)} \le 1/k$.
-	By the result already proven, the sequence $u_k$ is bounded in $H^2(\Omega)$ as follows:
-	\begin{align*}
-	||u_k||_{H^2(\Omega)}^2 &= ||u_k||_{L^2(\Omega)}^2 + ||D^1 u||_{\left(L^2(\Omega)\right)^d}^2 + ||D^2 u||_{\left(L^2(\Omega)\right)^{d \times d}}^2 \\
-	&\le ||u_k||_{L^2(\Omega)}^2 + C_1 ||\Delta u||_{L^2(\Omega)}^2 \le 1 + C_1.
-	\end{align*}
-	Therefore, by the Banach-Alaoglu theorem, there exists a subsequence $u_{k_j}$ and function $u \in X_0$ such that
-	\begin{equation*}
-		u_{k_j} \overset{w}{\rightharpoonup} u \quad \text{weakly in }H^2(\Omega).
-	\end{equation*}
-	Let $f \in L^2(\Omega)$.
-By the Cauchy-Schwarz inequality, the linear functional $v \mapsto \int_\Omega f(x) \Delta v(x) dx$ is bounded in $H^2(\Omega)$, and is therefore an element of $\left(H^2(\Omega)\right)^*$.
-Weak convergence therefore implies
-	\begin{equation*}
-		\int_\Omega f(x) \Delta u_k(x) dx \rightarrow \int_\Omega f(x) \Delta u(x) dx.
-	\end{equation*}
-	By the assumption $||\Delta u_k||_{L^2(\Omega)} \le 1/k$ and Cauchy-Schwarz, we also have 
-	\begin{equation*}
-	\int_\Omega f(x) \Delta u_k(x) dx \rightarrow 0,
-	\end{equation*}
-	so we must have
-	\begin{equation*}
-		\int_\Omega f(x) \Delta u(x) dx = 0.
-	\end{equation*}
-	Since this holds for all $f \in L^2(\Omega)$, by the fundamental lemma of the calculus of variations, $\Delta u(x) = 0$ almost everywhere in $\Omega$.
-This, combined with the fact that $u$ has Neumann zero boundary conditions, implies that $u$ is a constant function.
-Since $u$ has value zero at least one point in each connected component of $\Omega$, we have $u=0$.
-Note that pointwise values of $u$ are well-defined since the Sobolev embedding theorem imples that $H^2(\Omega)$ is continuously imbedded in the space of continuous functions $C^0(\Omega)$ when $d < 4$, and we are considering $d \in \{1,2,3\}$ (see, e.g., Theorem 7.19 in \cite{Arbogast}).
-
-	On the other hand, by the Rellich-Kondrachov theorem, $H^2(\Omega)$ is compactly imbedded in $L^2(\Omega)$, so there must exist a sub-sub-sequence, $u_{k_{j_i}}$ such that 
-	\begin{equation*}
-		u_{k_{j_i}} \rightarrow u
-	\end{equation*}
-	strongly in $L^2(\Omega)$.
-By assumption $||u_{k_{j_i}}||_{L^2(\Omega)}=1$, so $||u||_{L^2(\Omega)}=1$.
-This contradicts the result $u=0$ that we previously derived, so we must conclude that there exists a constant $C_2 > 0$ such that
-	\begin{equation*}
-		||u||_{L^2(\Omega)} \le C_2 ||\Delta u||_{L^2(\Omega)}.
-	\end{equation*}
-	for all $u \in X_0$. Combining this with out previous result, we have
-	\begin{equation*}
-	||u||^2_{H^2(\Omega)} \le (C_1 + C_2)||\Delta u||^2
-	\end{equation*}
-	for all $u \in X_0$, which completes the proof.
-\end{proof}
-
-\begin{lem}[Well-posedness of weighting function optimization problem]
-	\label{lem:wi_existence_and_uniqueness}
-	There exists a unique solution to optimization problem \eqref{eq:generic_optimization_initial}.
-\end{lem}
-
-\begin{proof}
-	By the Sobolev Imbedding Theorem (see, e.g., Theorem 7.20 in \cite{ARBOGAST}), $H^2_N(\Omega)$ is continuously imbedded in $C^0(\Omega)$, so pointwise evaluation $u \mapsto u(x_i)$ of functions at a point $x_i \in \Omega$ is a continuous linear functional on $H^2_N$.
-The set
-	\begin{equation}
-	\label{eq:inverse_image_of_point}
-	\{u \in H^2_N : u(x_i)=f_i\}
-	\end{equation}
-	is therefore a closed affine set in $H^2_N$, because it is the inverse image of a point under a continuous linear functional.
-Hence the feasible set $X$, defined in \eqref{eq:wi_feasible_set}, is a closed affine set in $H^2_N$, because it is the intersection of the sets in \eqref{eq:inverse_image_of_point} for $i=1,\dots,r$.
-
-	The feasible set $X$ is convex because it is affine, and it is non-empty because interpolating the data pairs $\{(x_i, f_i)\}_{i=1}^r$ using any smooth interpolation method yields an element of $X$.
-Furthermore, the objective function,
-	\begin{equation}
-	J(\gamma) := \frac{1}{2}||\Delta \gamma||^2_{L^2(\Omega)},
-	\end{equation}
-	is convex because it is the composition of a convex function with a linear map. Furthermore, the elliptic regularity result in Lemma \ref{lem:elliptic_regularity_x0} implies
-	\begin{equation*}
-		J''(\gamma)v^2 = ||\Delta v||^2_{L^2(\Omega)} \ge C ||v||_{H^2(\Omega)}^2
-	\end{equation*}
-	for all $v \in X_0$, which implies that $J$ is strongly convex on $X$.
-\end{proof}
-
 \begin{lem}[Linearity of weighting function solution operator]
@@
 		\label{eq:opt_soln_operator}
-		\mathcal{T}(f) := \gamma^*,
-		\end{equation}
-		where $\gamma^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
+		\mathcal{T}(f) := \horizinterpolant_*,
+		\end{equation}
+		where $\horizinterpolant_*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
 \end{lem}
@@
 	&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
-	&= \gamma_y.
-	\end{align*}
-	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $\gamma_y$ in \eqref{eq:u_defn_in_thm}.
+	&= \horizinterpolant_y.
+	\end{align*}
+	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $\horizinterpolant_y$ in \eqref{eq:u_defn_in_thm}.
 \end{proof}
@@
 The last equality ($\mathcal{T}(\mathbf{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \mathbf{1}$, and yields the minimum possible value of zero for the objective function.
-\end{proof}
-
-\begin{proof}[Proof of Theorem \ref{alg:weighting_functions_incremental}]
-	For notational convenience, in this proof we write $w$, $\lambda$, and $e$ instead of $w_i$, $\lambda_i$, and $e_i$, respectively.
-Here subscripts for $\lambda$, and $e$ instead denote components of those vectors.
-For example, here $\lambda_j$ denotes the $j^\text{th}$ component of the vector $\lambda$, which corresponds to the $j^\text{th}$ component of the vector $\lambda_i$ in Algorithm \ref{alg:weighting_functions_incremental}.
-
-	From standard optimization theory, the solution to optimization problem \ref{eq:wi_optimization_problem} is the stationary point of the Lagrangian,
-	\begin{equation*}
-	\mathcal{L}(w,\lambda) = \frac{1}{2}\|\Delta w\|^2 + \sum_{j=1}^r \lambda_j\left(w(x_j) - e_j\right).
-	\end{equation*}
-	The plan of this proof is to show that $\nabla \mathcal{L}(w,\lambda)=0$, which will verify that $w$ solves optimization problem \ref{eq:wi_optimization_problem}.
-
-	Using the definition of $w_i$ in \eqref{eq:definition_of_wi_alg}, we may write the gradient of $\mathcal{L}$ with respect to $\lambda_j$ as
-	\begin{equation*}
-	\nabla_{\lambda_j} \mathcal{L}(w, \lambda) = w(x_j) - e_j
-	= \beta - \sum_{k=1}^r \lambda_k \theta_k(x_j) - e_j.
-	\end{equation*}
-	Using Lemma \ref{lem:psi_delta_eval} with $h=\theta_k$, the equation for $\theta_k$ in \eqref{eq:theta_eq}, and the definition of $S$ in \eqref{eq:defn_of_S_alg}, we have
-	\begin{align*}
-	\sum_{k=1}^r \lambda_k \theta_k(x_j) &= \sum_{k=1}^r \lambda_k \left(\psi_j, \Delta \theta_k\right)_{L^2(\Omega)}\\
-	&= \sum_{k=1}^r \lambda_k \left(\psi_j, \psi_k\right)_{L^2(\Omega)} = \left(S \lambda\right)_j.
-	\end{align*}
-	Combining all of these components of $\nabla_\lambda \mathcal{L}$ into a single vector and using the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} yields
-	\begin{align*}
-	\nabla_\lambda \mathcal{L}(w, \lambda) &= \beta \mathbf{1} - e - S \lambda \\
-	&= \beta \mathbf{1} - e - S \left(-S^{-1}\left(e - \beta \mathbf{1}\right)\right) = 0.
-	\end{align*}
-
-	Taking the directional derivative of $\mathcal{L}$ with respect to $w$ in direction $h$ and using Lemma \ref{lem:psi_delta_eval} yields
-	\begin{align}
-	\nabla_w \mathcal{L}(w, \lambda)h &= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j h(x_j) \nonumber \\
-	&= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \lambda_j \avg(h) \nonumber \\
-	&= \left(\Delta w + \sum_{j=1}^r \lambda_j \psi_j,~ \Delta h\right)_{L^2(\Omega)} + \avg(h) \mathbf{1}^T \lambda. \label{eq:L_w_38}
-	\end{align}
-	Using the definition of $w$ in Algorithm \ref{alg:weighting_functions_incremental}, and the fact that the Laplacian of a constant function is zero, we have
-	\begin{equation*}
-	\Delta w = -\sum_{j=1}^r \lambda_j \Delta \theta_j 
-	= -\sum_{j=1}^r \lambda_j \psi_j,
-	\end{equation*}
-	which implies that the first term in \eqref{eq:L_w_38} is zero. By the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} and $\beta$ in \eqref{eq:definition_of_alpha_alg}, we have
-	\begin{align*}
-	\mathbf{1}^T \lambda &= -\mathbf{1}^T S^{-1} \left(e - \beta \mathbf{1}\right) \\
-	&= -\mathbf{1}^T S^{-1} e + \beta \mathbf{1}^T S^{-1} \mathbf{1} \\
-	&= -\mathbf{1}^T S^{-1} e + \frac{\mathbf{1}^T S^{-1} e}{\mathbf{1}^T S^{-1} \mathbf{1}} \mathbf{1}^T S^{-1} \mathbf{1} = 0,
-	\end{align*}
-	which implies that the second term in \eqref{eq:L_w_38} is zero.
-Thus $\nabla_w \mathcal{L}(w, \lambda) = 0$, so Algorithm \ref{alg:weighting_functions_incremental} produces the solution to optimization problem \ref{eq:wi_optimization_problem}.
 \end{proof}
@@
 	which implies $\Sigma^{ij} = \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot\mu^j$.
-
 \end{proof}
@@
 The discrete dual of a vector $\mathbf{v}$ with respect to the $\mathbf{M}$ inner product is written as $\mathbf{v}^*$, and $\mathbf{v}^*$ is viewed as a column vector.
-%These isomorphisms allow us to identify functions $u_h \in V_h$ with their corresponding coefficient arrays $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, and linear functionals $\sigma_h \in V_h'$ with their corresponding coefficient arrays $\boldsymbol{\sigma} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$.
 
@@
 \subsection{Discretized sampling of impulse responses}
-In computations, the Dirac comb coefficient dual vector for $\xi_h^b$ is the vector $\boldsymbol{\xi}^b \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has entries
-\begin{equation*}
-\boldsymbol{\xi}^b_i = \sum_{x_j \in S_b} f_i(x_j), \quad i=1,\dots,N.
-\end{equation*}
-The Dirac comb impulse response coefficient vector for $\eta_h^b$ is the vector $\boldsymbol{\eta}^b \in \mathbb{R}^N_\mathbf{M}$ given by
-\begin{equation*}
-\boldsymbol{\eta}^b = \mathbf{M}^{-1} \mathbf{A}^T \mathbf{M}^{-1}\boldsymbol{\xi}^b.
+In computations, the Dirac comb coefficient dual vector for $\dcomb_h^b$ is the vector $\boldsymbol{\dcomb}^b \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has entries
+\begin{equation*}
+\boldsymbol{\dcomb}^b_i = \sum_{x_j \in S_b} f_i(x_j), \quad i=1,\dots,N.
+\end{equation*}
+The Dirac comb impulse response coefficient vector for $\combresponse^h_b$ is the vector $\boldsymbol{\combresponse}_b \in \mathbb{R}^N_\mathbf{M}$ given by
+\begin{equation*}
+\boldsymbol{\combresponse}_b = \mathbf{M}^{-1} \mathbf{A}^T \mathbf{M}^{-1}\boldsymbol{\dcomb}^b.
 \end{equation*}
@@
 \section{DISORGANIZED OLD STUFF}
+
+\begin{proof}[Proof of Theorem \ref{alg:weighting_functions_incremental}]
+	For notational convenience, in this proof we write $w$, $\lambda$, and $e$ instead of $w_i$, $\lambda_i$, and $e_i$, respectively.
+Here subscripts for $\lambda$, and $e$ instead denote components of those vectors.
+For example, here $\lambda_j$ denotes the $j^\text{th}$ component of the vector $\lambda$, which corresponds to the $j^\text{th}$ component of the vector $\lambda_i$ in Algorithm \ref{alg:weighting_functions_incremental}.
+
+	From standard optimization theory, the solution to optimization problem \ref{eq:wi_optimization_problem} is the stationary point of the Lagrangian,
+	\begin{equation*}
+	\mathcal{L}(w,\lambda) = \frac{1}{2}\|\Delta w\|^2 + \sum_{j=1}^r \lambda_j\left(w(x_j) - e_j\right).
+	\end{equation*}
+	The plan of this proof is to show that $\nabla \mathcal{L}(w,\lambda)=0$, which will verify that $w$ solves optimization problem \ref{eq:wi_optimization_problem}.
+
+	Using the definition of $w_i$ in \eqref{eq:definition_of_wi_alg}, we may write the gradient of $\mathcal{L}$ with respect to $\lambda_j$ as
+	\begin{equation*}
+	\nabla_{\lambda_j} \mathcal{L}(w, \lambda) = w(x_j) - e_j
+	= \beta - \sum_{k=1}^r \lambda_k \secondgreens_k(x_j) - e_j.
+	\end{equation*}
+	Using Lemma \ref{lem:psi_delta_eval} with $h=\secondgreens_k$, the equation for $\secondgreens_k$ in \eqref{eq:theta_eq}, and the definition of $S$ in \eqref{eq:defn_of_S_alg}, we have
+	\begin{align*}
+	\sum_{k=1}^r \lambda_k \secondgreens_k(x_j) &= \sum_{k=1}^r \lambda_k \left(\firstgreens_j, \Delta \secondgreens_k\right)_{L^2(\Omega)}\\
+	&= \sum_{k=1}^r \lambda_k \left(\firstgreens_j, \firstgreens_k\right)_{L^2(\Omega)} = \left(S \lambda\right)_j.
+	\end{align*}
+	Combining all of these components of $\nabla_\lambda \mathcal{L}$ into a single vector and using the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} yields
+	\begin{align*}
+	\nabla_\lambda \mathcal{L}(w, \lambda) &= \beta \mathbf{1} - e - S \lambda \\
+	&= \beta \mathbf{1} - e - S \left(-S^{-1}\left(e - \beta \mathbf{1}\right)\right) = 0.
+	\end{align*}
+
+	Taking the directional derivative of $\mathcal{L}$ with respect to $w$ in direction $h$ and using Lemma \ref{lem:psi_delta_eval} yields
+	\begin{align}
+	\nabla_w \mathcal{L}(w, \lambda)h &= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j h(x_j) \nonumber \\
+	&= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j \left(\firstgreens_j, \Delta h\right)_{L^2(\Omega)} + \lambda_j \avg(h) \nonumber \\
+	&= \left(\Delta w + \sum_{j=1}^r \lambda_j \firstgreens_j,~ \Delta h\right)_{L^2(\Omega)} + \avg(h) \mathbf{1}^T \lambda. \label{eq:L_w_38}
+	\end{align}
+	Using the definition of $w$ in Algorithm \ref{alg:weighting_functions_incremental}, and the fact that the Laplacian of a constant function is zero, we have
+	\begin{equation*}
+	\Delta w = -\sum_{j=1}^r \lambda_j \Delta \secondgreens_j 
+	= -\sum_{j=1}^r \lambda_j \firstgreens_j,
+	\end{equation*}
+	which implies that the first term in \eqref{eq:L_w_38} is zero. By the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} and $\beta$ in \eqref{eq:definition_of_alpha_alg}, we have
+	\begin{align*}
+	\mathbf{1}^T \lambda &= -\mathbf{1}^T S^{-1} \left(e - \beta \mathbf{1}\right) \\
+	&= -\mathbf{1}^T S^{-1} e + \beta \mathbf{1}^T S^{-1} \mathbf{1} \\
+	&= -\mathbf{1}^T S^{-1} e + \frac{\mathbf{1}^T S^{-1} e}{\mathbf{1}^T S^{-1} \mathbf{1}} \mathbf{1}^T S^{-1} \mathbf{1} = 0,
+	\end{align*}
+	which implies that the second term in \eqref{eq:L_w_38} is zero.
+Thus $\nabla_w \mathcal{L}(w, \lambda) = 0$, so Algorithm \ref{alg:weighting_functions_incremental} produces the solution to optimization problem \ref{eq:wi_optimization_problem}.
+\end{proof}
+
+\begin{defn}[Weighting function optimization problem]
+	Let $\Omega \subset \mathbb{R}^d$, $d\in\{1,2,3\}$, be an open bounded domain with, let $x_i \in \Omega$ for $i=1,\dots,r$, and let $f \in \mathbb{R}^r$. We consider the optimization problem
+	\begin{equation}
+	\label{eq:generic_optimization_initial}
+	\begin{aligned}
+	\min_{\gamma \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \gamma}_{L^2(\Omega)}^2 \\
+	\text{such that} &\quad \gamma(x_i) = f_i, \quad i=1,\dots,r.
+	\end{aligned}
+	\end{equation}
+	A minimizer, $\gamma^*$, of \eqref{eq:generic_optimization_initial} is a function defined on $\Omega$ that interpolates the data pairs $(x_i,f_i)$, and is as flat as possible in a least squares sense.
+	We denote the objective function for optimization problem \eqref{eq:generic_optimization_initial} by 
+	\begin{equation}
+	\label{eq:wi_objective}
+	J(\gamma) := \frac{1}{2}||\Delta \gamma||^2_{L^2(\Omega)}.
+	\end{equation}
+	We denote the feasible set for optimization problem \eqref{eq:generic_optimization_initial} by 
+	\begin{equation}
+	\label{eq:wi_feasible_set}
+	X := \{u \in H^2_N : u(x_i)=f_i~\text{for}~ i=1,\dots,r\}.
+	\end{equation}
+	We denote the tangent space to the feasible set by
+	\begin{equation}
+	\label{eq:wi_tangent_space}
+	X_0 := \{u \in H^2_N : u(x_i)=0~\text{for}~ i=1,\dots,r\}.
+	\end{equation}
+\end{defn}
+
+\begin{lem}[Elliptic regularity on $X_0$]
+	\label{lem:elliptic_regularity_x0}
+	If $\partial \Omega$ is $C^{1,1}$, then there exists a constant $C > 0$ such that
+	\begin{equation*}
+	||u||_{H^2(\Omega)} \le C ||\Delta u||_{L^2(\Omega)}
+	\end{equation*}
+	for all $u \in X_0$.
+\end{lem}
+
+\begin{proof}
+	Let $\widetilde{u}$ be version of $u$ that has been shifted up or down by a constant such that $\int_\Omega \overline{u} = 0$. We have
+	\begin{align*}
+	||D^1 u||_{\left(L^2(\Omega)\right)^d}^2 + ||D^2 u||_{\left(L^2(\Omega)\right)^{d \times d}}^2 &= ||D^1 \widetilde{u}||_{\left(L^2(\Omega)\right)^d}^2 + ||D^2 \widetilde{u}||_{\left(L^2(\Omega)\right)^{d \times d}}^2\\
+	&\le ||\widetilde{u}||_{H^2(\Omega)}^2 \\
+	&\le C_1 ||\Delta \widetilde{u}||_{L^2(\Omega)}^2 \\
+	&= C_1 ||\Delta u||_{L^2(\Omega)}^2
+	\end{align*}
+	for some constant $C_1$.
+In the first and fourth lines we used the fact that the derivative of a constant is zero, in the second line we used the definition of the $H^2$ norm, and in the third line we used the standard elliptic regularity result for pure Neumann problems (see, e.g., Theorem 8.13 (b) in \cite{Arbogast}).
+
+	To complete the proof, we will now show that there exists a constant $C_2$ such that $||u||_{L^2(\Omega)}^2 \le C_2 ||\Delta u||_{L^2(\Omega)}^2$ for all $u \in X_0$.
+For this, we use a standard compactness argument, which closely follows the proof of Theorem 8.12 in \cite{ARBOGAST}.
+Suppose for the sake of contradiction that this result does not hold.
+Then there exists a sequence $\{u_k\}_{k=1}^\infty \subset X_0$ such that
+	$||u_k||_{L^2(\Omega)}=1$ and $||\Delta u_k||_{L^2(\Omega)} \le 1/k$.
+	By the result already proven, the sequence $u_k$ is bounded in $H^2(\Omega)$ as follows:
+	\begin{align*}
+	||u_k||_{H^2(\Omega)}^2 &= ||u_k||_{L^2(\Omega)}^2 + ||D^1 u||_{\left(L^2(\Omega)\right)^d}^2 + ||D^2 u||_{\left(L^2(\Omega)\right)^{d \times d}}^2 \\
+	&\le ||u_k||_{L^2(\Omega)}^2 + C_1 ||\Delta u||_{L^2(\Omega)}^2 \le 1 + C_1.
+	\end{align*}
+	Therefore, by the Banach-Alaoglu theorem, there exists a subsequence $u_{k_j}$ and function $u \in X_0$ such that
+	\begin{equation*}
+	u_{k_j} \overset{w}{\rightharpoonup} u \quad \text{weakly in }H^2(\Omega).
+	\end{equation*}
+	Let $f \in L^2(\Omega)$.
+By the Cauchy-Schwarz inequality, the linear functional $v \mapsto \int_\Omega f(x) \Delta v(x) dx$ is bounded in $H^2(\Omega)$, and is therefore an element of $\left(H^2(\Omega)\right)^*$.
+Weak convergence therefore implies
+	\begin{equation*}
+	\int_\Omega f(x) \Delta u_k(x) dx \rightarrow \int_\Omega f(x) \Delta u(x) dx.
+	\end{equation*}
+	By the assumption $||\Delta u_k||_{L^2(\Omega)} \le 1/k$ and Cauchy-Schwarz, we also have 
+	\begin{equation*}
+	\int_\Omega f(x) \Delta u_k(x) dx \rightarrow 0,
+	\end{equation*}
+	so we must have
+	\begin{equation*}
+	\int_\Omega f(x) \Delta u(x) dx = 0.
+	\end{equation*}
+	Since this holds for all $f \in L^2(\Omega)$, by the fundamental lemma of the calculus of variations, $\Delta u(x) = 0$ almost everywhere in $\Omega$.
+This, combined with the fact that $u$ has Neumann zero boundary conditions, implies that $u$ is a constant function.
+Since $u$ has value zero at least one point in each connected component of $\Omega$, we have $u=0$.
+Note that pointwise values of $u$ are well-defined since the Sobolev embedding theorem imples that $H^2(\Omega)$ is continuously imbedded in the space of continuous functions $C^0(\Omega)$ when $d < 4$, and we are considering $d \in \{1,2,3\}$ (see, e.g., Theorem 7.19 in \cite{Arbogast}).
+
+	On the other hand, by the Rellich-Kondrachov theorem, $H^2(\Omega)$ is compactly imbedded in $L^2(\Omega)$, so there must exist a sub-sub-sequence, $u_{k_{j_i}}$ such that 
+	\begin{equation*}
+	u_{k_{j_i}} \rightarrow u
+	\end{equation*}
+	strongly in $L^2(\Omega)$.
+By assumption $||u_{k_{j_i}}||_{L^2(\Omega)}=1$, so $||u||_{L^2(\Omega)}=1$.
+This contradicts the result $u=0$ that we previously derived, so we must conclude that there exists a constant $C_2 > 0$ such that
+	\begin{equation*}
+	||u||_{L^2(\Omega)} \le C_2 ||\Delta u||_{L^2(\Omega)}.
+	\end{equation*}
+	for all $u \in X_0$. Combining this with out previous result, we have
+	\begin{equation*}
+	||u||^2_{H^2(\Omega)} \le (C_1 + C_2)||\Delta u||^2
+	\end{equation*}
+	for all $u \in X_0$, which completes the proof.
+\end{proof}
+
+\begin{lem}[Well-posedness of weighting function optimization problem]
+	\label{lem:wi_existence_and_uniqueness}
+	There exists a unique solution to optimization problem \eqref{eq:generic_optimization_initial}.
+\end{lem}
+
+\begin{proof}
+	By the Sobolev Imbedding Theorem (see, e.g., Theorem 7.20 in \cite{ARBOGAST}), $H^2_N(\Omega)$ is continuously imbedded in $C^0(\Omega)$, so pointwise evaluation $u \mapsto u(x_i)$ of functions at a point $x_i \in \Omega$ is a continuous linear functional on $H^2_N$.
+The set
+	\begin{equation}
+	\label{eq:inverse_image_of_point}
+	\{u \in H^2_N : u(x_i)=f_i\}
+	\end{equation}
+	is therefore a closed affine set in $H^2_N$, because it is the inverse image of a point under a continuous linear functional.
+Hence the feasible set $X$, defined in \eqref{eq:wi_feasible_set}, is a closed affine set in $H^2_N$, because it is the intersection of the sets in \eqref{eq:inverse_image_of_point} for $i=1,\dots,r$.
+
+	The feasible set $X$ is convex because it is affine, and it is non-empty because interpolating the data pairs $\{(x_i, f_i)\}_{i=1}^r$ using any smooth interpolation method yields an element of $X$.
+Furthermore, the objective function,
+	\begin{equation}
+	J(\gamma) := \frac{1}{2}||\Delta \gamma||^2_{L^2(\Omega)},
+	\end{equation}
+	is convex because it is the composition of a convex function with a linear map. Furthermore, the elliptic regularity result in Lemma \ref{lem:elliptic_regularity_x0} implies
+	\begin{equation*}
+	J''(\gamma)v^2 = ||\Delta v||^2_{L^2(\Omega)} \ge C ||v||_{H^2(\Omega)}^2
+	\end{equation*}
+	for all $v \in X_0$, which implies that $J$ is strongly convex on $X$.
+\end{proof}
 

### 2021-09-14 paper/localpsf.tex
removed old stuff section from paper

@@
 
-\section{DISORGANIZED OLD STUFF}
-
-\begin{proof}[Proof of Theorem \ref{alg:weighting_functions_incremental}]
-	For notational convenience, in this proof we write $w$, $\lambda$, and $e$ instead of $w_i$, $\lambda_i$, and $e_i$, respectively.
-Here subscripts for $\lambda$, and $e$ instead denote components of those vectors.
-For example, here $\lambda_j$ denotes the $j^\text{th}$ component of the vector $\lambda$, which corresponds to the $j^\text{th}$ component of the vector $\lambda_i$ in Algorithm \ref{alg:weighting_functions_incremental}.
-
-	From standard optimization theory, the solution to optimization problem \ref{eq:wi_optimization_problem} is the stationary point of the Lagrangian,
-	\begin{equation*}
-	\mathcal{L}(w,\lambda) = \frac{1}{2}\|\Delta w\|^2 + \sum_{j=1}^r \lambda_j\left(w(x_j) - e_j\right).
-	\end{equation*}
-	The plan of this proof is to show that $\nabla \mathcal{L}(w,\lambda)=0$, which will verify that $w$ solves optimization problem \ref{eq:wi_optimization_problem}.
-
-	Using the definition of $w_i$ in \eqref{eq:definition_of_wi_alg}, we may write the gradient of $\mathcal{L}$ with respect to $\lambda_j$ as
-	\begin{equation*}
-	\nabla_{\lambda_j} \mathcal{L}(w, \lambda) = w(x_j) - e_j
-	= \beta - \sum_{k=1}^r \lambda_k \secondgreens_k(x_j) - e_j.
-	\end{equation*}
-	Using Lemma \ref{lem:psi_delta_eval} with $h=\secondgreens_k$, the equation for $\secondgreens_k$ in \eqref{eq:theta_eq}, and the definition of $S$ in \eqref{eq:defn_of_S_alg}, we have
-	\begin{align*}
-	\sum_{k=1}^r \lambda_k \secondgreens_k(x_j) &= \sum_{k=1}^r \lambda_k \left(\firstgreens_j, \Delta \secondgreens_k\right)_{L^2(\Omega)}\\
-	&= \sum_{k=1}^r \lambda_k \left(\firstgreens_j, \firstgreens_k\right)_{L^2(\Omega)} = \left(S \lambda\right)_j.
-	\end{align*}
-	Combining all of these components of $\nabla_\lambda \mathcal{L}$ into a single vector and using the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} yields
-	\begin{align*}
-	\nabla_\lambda \mathcal{L}(w, \lambda) &= \beta \mathbf{1} - e - S \lambda \\
-	&= \beta \mathbf{1} - e - S \left(-S^{-1}\left(e - \beta \mathbf{1}\right)\right) = 0.
-	\end{align*}
-
-	Taking the directional derivative of $\mathcal{L}$ with respect to $w$ in direction $h$ and using Lemma \ref{lem:psi_delta_eval} yields
-	\begin{align}
-	\nabla_w \mathcal{L}(w, \lambda)h &= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j h(x_j) \nonumber \\
-	&= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j \left(\firstgreens_j, \Delta h\right)_{L^2(\Omega)} + \lambda_j \avg(h) \nonumber \\
-	&= \left(\Delta w + \sum_{j=1}^r \lambda_j \firstgreens_j,~ \Delta h\right)_{L^2(\Omega)} + \avg(h) \mathbf{1}^T \lambda. \label{eq:L_w_38}
-	\end{align}
-	Using the definition of $w$ in Algorithm \ref{alg:weighting_functions_incremental}, and the fact that the Laplacian of a constant function is zero, we have
-	\begin{equation*}
-	\Delta w = -\sum_{j=1}^r \lambda_j \Delta \secondgreens_j 
-	= -\sum_{j=1}^r \lambda_j \firstgreens_j,
-	\end{equation*}
-	which implies that the first term in \eqref{eq:L_w_38} is zero. By the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} and $\beta$ in \eqref{eq:definition_of_alpha_alg}, we have
-	\begin{align*}
-	\mathbf{1}^T \lambda &= -\mathbf{1}^T S^{-1} \left(e - \beta \mathbf{1}\right) \\
-	&= -\mathbf{1}^T S^{-1} e + \beta \mathbf{1}^T S^{-1} \mathbf{1} \\
-	&= -\mathbf{1}^T S^{-1} e + \frac{\mathbf{1}^T S^{-1} e}{\mathbf{1}^T S^{-1} \mathbf{1}} \mathbf{1}^T S^{-1} \mathbf{1} = 0,
-	\end{align*}
-	which implies that the second term in \eqref{eq:L_w_38} is zero.
-Thus $\nabla_w \mathcal{L}(w, \lambda) = 0$, so Algorithm \ref{alg:weighting_functions_incremental} produces the solution to optimization problem \ref{eq:wi_optimization_problem}.
-\end{proof}
-
-\begin{defn}[Weighting function optimization problem]
-	Let $\Omega \subset \mathbb{R}^d$, $d\in\{1,2,3\}$, be an open bounded domain with, let $x_i \in \Omega$ for $i=1,\dots,r$, and let $f \in \mathbb{R}^r$. We consider the optimization problem
-	\begin{equation}
-	\label{eq:generic_optimization_initial}
-	\begin{aligned}
-	\min_{\gamma \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \gamma}_{L^2(\Omega)}^2 \\
-	\text{such that} &\quad \gamma(x_i) = f_i, \quad i=1,\dots,r.
-	\end{aligned}
-	\end{equation}
-	A minimizer, $\gamma^*$, of \eqref{eq:generic_optimization_initial} is a function defined on $\Omega$ that interpolates the data pairs $(x_i,f_i)$, and is as flat as possible in a least squares sense.
-	We denote the objective function for optimization problem \eqref{eq:generic_optimization_initial} by 
-	\begin{equation}
-	\label{eq:wi_objective}
-	J(\gamma) := \frac{1}{2}||\Delta \gamma||^2_{L^2(\Omega)}.
-	\end{equation}
-	We denote the feasible set for optimization problem \eqref{eq:generic_optimization_initial} by 
-	\begin{equation}
-	\label{eq:wi_feasible_set}
-	X := \{u \in H^2_N : u(x_i)=f_i~\text{for}~ i=1,\dots,r\}.
-	\end{equation}
-	We denote the tangent space to the feasible set by
-	\begin{equation}
-	\label{eq:wi_tangent_space}
-	X_0 := \{u \in H^2_N : u(x_i)=0~\text{for}~ i=1,\dots,r\}.
-	\end{equation}
-\end{defn}
-
-\begin{lem}[Elliptic regularity on $X_0$]
-	\label{lem:elliptic_regularity_x0}
-	If $\partial \Omega$ is $C^{1,1}$, then there exists a constant $C > 0$ such that
-	\begin{equation*}
-	||u||_{H^2(\Omega)} \le C ||\Delta u||_{L^2(\Omega)}
-	\end{equation*}
-	for all $u \in X_0$.
-\end{lem}
-
-\begin{proof}
-	Let $\widetilde{u}$ be version of $u$ that has been shifted up or down by a constant such that $\int_\Omega \overline{u} = 0$. We have
-	\begin{align*}
-	||D^1 u||_{\left(L^2(\Omega)\right)^d}^2 + ||D^2 u||_{\left(L^2(\Omega)\right)^{d \times d}}^2 &= ||D^1 \widetilde{u}||_{\left(L^2(\Omega)\right)^d}^2 + ||D^2 \widetilde{u}||_{\left(L^2(\Omega)\right)^{d \times d}}^2\\
-	&\le ||\widetilde{u}||_{H^2(\Omega)}^2 \\
-	&\le C_1 ||\Delta \widetilde{u}||_{L^2(\Omega)}^2 \\
-	&= C_1 ||\Delta u||_{L^2(\Omega)}^2
-	\end{align*}
-	for some constant $C_1$.
-In the first and fourth lines we used the fact that the derivative of a constant is zero, in the second line we used the definition of the $H^2$ norm, and in the third line we used the standard elliptic regularity result for pure Neumann problems (see, e.g., Theorem 8.13 (b) in \cite{Arbogast}).
-
-	To complete the proof, we will now show that there exists a constant $C_2$ such that $||u||_{L^2(\Omega)}^2 \le C_2 ||\Delta u||_{L^2(\Omega)}^2$ for all $u \in X_0$.
-For this, we use a standard compactness argument, which closely follows the proof of Theorem 8.12 in \cite{ARBOGAST}.
-Suppose for the sake of contradiction that this result does not hold.
-Then there exists a sequence $\{u_k\}_{k=1}^\infty \subset X_0$ such that
-	$||u_k||_{L^2(\Omega)}=1$ and $||\Delta u_k||_{L^2(\Omega)} \le 1/k$.
-	By the result already proven, the sequence $u_k$ is bounded in $H^2(\Omega)$ as follows:
-	\begin{align*}
-	||u_k||_{H^2(\Omega)}^2 &= ||u_k||_{L^2(\Omega)}^2 + ||D^1 u||_{\left(L^2(\Omega)\right)^d}^2 + ||D^2 u||_{\left(L^2(\Omega)\right)^{d \times d}}^2 \\
-	&\le ||u_k||_{L^2(\Omega)}^2 + C_1 ||\Delta u||_{L^2(\Omega)}^2 \le 1 + C_1.
-	\end{align*}
-	Therefore, by the Banach-Alaoglu theorem, there exists a subsequence $u_{k_j}$ and function $u \in X_0$ such that
-	\begin{equation*}
-	u_{k_j} \overset{w}{\rightharpoonup} u \quad \text{weakly in }H^2(\Omega).
-	\end{equation*}
-	Let $f \in L^2(\Omega)$.
-By the Cauchy-Schwarz inequality, the linear functional $v \mapsto \int_\Omega f(x) \Delta v(x) dx$ is bounded in $H^2(\Omega)$, and is therefore an element of $\left(H^2(\Omega)\right)^*$.
-Weak convergence therefore implies
-	\begin{equation*}
-	\int_\Omega f(x) \Delta u_k(x) dx \rightarrow \int_\Omega f(x) \Delta u(x) dx.
-	\end{equation*}
-	By the assumption $||\Delta u_k||_{L^2(\Omega)} \le 1/k$ and Cauchy-Schwarz, we also have 
-	\begin{equation*}
-	\int_\Omega f(x) \Delta u_k(x) dx \rightarrow 0,
-	\end{equation*}
-	so we must have
-	\begin{equation*}
-	\int_\Omega f(x) \Delta u(x) dx = 0.
-	\end{equation*}
-	Since this holds for all $f \in L^2(\Omega)$, by the fundamental lemma of the calculus of variations, $\Delta u(x) = 0$ almost everywhere in $\Omega$.
-This, combined with the fact that $u$ has Neumann zero boundary conditions, implies that $u$ is a constant function.
-Since $u$ has value zero at least one point in each connected component of $\Omega$, we have $u=0$.
-Note that pointwise values of $u$ are well-defined since the Sobolev embedding theorem imples that $H^2(\Omega)$ is continuously imbedded in the space of continuous functions $C^0(\Omega)$ when $d < 4$, and we are considering $d \in \{1,2,3\}$ (see, e.g., Theorem 7.19 in \cite{Arbogast}).
-
-	On the other hand, by the Rellich-Kondrachov theorem, $H^2(\Omega)$ is compactly imbedded in $L^2(\Omega)$, so there must exist a sub-sub-sequence, $u_{k_{j_i}}$ such that 
-	\begin{equation*}
-	u_{k_{j_i}} \rightarrow u
-	\end{equation*}
-	strongly in $L^2(\Omega)$.
-By assumption $||u_{k_{j_i}}||_{L^2(\Omega)}=1$, so $||u||_{L^2(\Omega)}=1$.
-This contradicts the result $u=0$ that we previously derived, so we must conclude that there exists a constant $C_2 > 0$ such that
-	\begin{equation*}
-	||u||_{L^2(\Omega)} \le C_2 ||\Delta u||_{L^2(\Omega)}.
-	\end{equation*}
-	for all $u \in X_0$. Combining this with out previous result, we have
-	\begin{equation*}
-	||u||^2_{H^2(\Omega)} \le (C_1 + C_2)||\Delta u||^2
-	\end{equation*}
-	for all $u \in X_0$, which completes the proof.
-\end{proof}
-
-\begin{lem}[Well-posedness of weighting function optimization problem]
-	\label{lem:wi_existence_and_uniqueness}
-	There exists a unique solution to optimization problem \eqref{eq:generic_optimization_initial}.
-\end{lem}
-
-\begin{proof}
-	By the Sobolev Imbedding Theorem (see, e.g., Theorem 7.20 in \cite{ARBOGAST}), $H^2_N(\Omega)$ is continuously imbedded in $C^0(\Omega)$, so pointwise evaluation $u \mapsto u(x_i)$ of functions at a point $x_i \in \Omega$ is a continuous linear functional on $H^2_N$.
-The set
-	\begin{equation}
-	\label{eq:inverse_image_of_point}
-	\{u \in H^2_N : u(x_i)=f_i\}
-	\end{equation}
-	is therefore a closed affine set in $H^2_N$, because it is the inverse image of a point under a continuous linear functional.
-Hence the feasible set $X$, defined in \eqref{eq:wi_feasible_set}, is a closed affine set in $H^2_N$, because it is the intersection of the sets in \eqref{eq:inverse_image_of_point} for $i=1,\dots,r$.
-
-	The feasible set $X$ is convex because it is affine, and it is non-empty because interpolating the data pairs $\{(x_i, f_i)\}_{i=1}^r$ using any smooth interpolation method yields an element of $X$.
-Furthermore, the objective function,
-	\begin{equation}
-	J(\gamma) := \frac{1}{2}||\Delta \gamma||^2_{L^2(\Omega)},
-	\end{equation}
-	is convex because it is the composition of a convex function with a linear map. Furthermore, the elliptic regularity result in Lemma \ref{lem:elliptic_regularity_x0} implies
-	\begin{equation*}
-	J''(\gamma)v^2 = ||\Delta v||^2_{L^2(\Omega)} \ge C ||v||_{H^2(\Omega)}^2
-	\end{equation*}
-	for all $v \in X_0$, which implies that $J$ is strongly convex on $X$.
-\end{proof}
-
-\begin{defn}
-	\label{def:stiffness_mass}
-	We define $\mathcal{K}:H^2(\Omega)\rightarrow L^2(\Omega)'$ and $\mathcal{M}:L^2(\Omega) \rightarrow L^2(\Omega)'$ to be the following linear operators:
-	\begin{align*}
-	\left(\mathcal{K}u\right)(v) :=& \int_\Omega v \Delta u ~dx \\
-	\left(\mathcal{M}u\right)(v) :=& \int_\Omega u v ~dx,
-	\end{align*}
-	and $\mathcal{P}:H^2(\Omega)\rightarrow \mathbb{R}^r$ to be the following pointwise observation operator:
-	\begin{equation*}
-	(\mathcal{P}u)_i = u(x_i)
-	\end{equation*}
-	for $i=1,\dots,r$.
-\end{defn}
-
-\begin{prop}
-	\label{prop:solution_operator_is_linear}
-	Let $f \in \mathbb{R}^r$ be arbitrary. The optimization problem
-	\begin{equation}
-	\label{eq:generic_optimization_initial_old}
-	\begin{aligned}
-	\min_{u \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
-	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r
-	\end{aligned}
-	\end{equation}
-	may be rewritten as follows:
-	\begin{equation}
-	\label{eq:generic_optimization}
-	\begin{aligned}
-	\min_{u \in H^2(\Omega)} &\quad \frac{1}{2} \left(\mathcal{K}^T\mathcal{M}^{-1}\mathcal{K}u\right)(u) \\
-	\text{such that} &\quad 
-	\mathcal{P} u = f.
-	\end{aligned}
-	\end{equation}
-\end{prop}
-
-\begin{prop}
-	There exists a unique solution to optimization problem \eqref{eq:generic_optimization}.
-The solution to \eqref{eq:generic_optimization}, $u^*$, and the adjoint variable for enforcing the constraint, $\lambda^*$, solve the following Karun-Kush-Tucker (KKT) system:
-	\begin{equation}
-	\label{eq:weighting_kkt}
-	\begin{bmatrix}
-	\mathcal{K}^T\mathcal{M}^{-1}\mathcal{K} & \mathcal{P}^T \\
-	\mathcal{P} & 0
-	\end{bmatrix}
-	\begin{bmatrix}
-	u^* \\ \lambda^*
-	\end{bmatrix}
-	=
-	\begin{bmatrix}
-	0 \\ f
-	\end{bmatrix},
-	\end{equation}
-	where we use block-matrix notation to denote block-operators.
-\end{prop}
-
-\begin{cor}
-	\label{cor:linearity_of_solution_operator}
-	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization}	as a function of $f$. That is, 
-	\begin{equation*}
-	\mathcal{T}(f) := u^*,
-	\end{equation*}
-	where $u^*$ is the minimizer of \eqref{eq:generic_optimization}. The operator $\mathcal{T}$ is a linear operator.
-\end{cor}
-
-\footnote{The reason for this dependence on spatial dimension is related to the fact that the fundamental solution of the biharmonic equation is continuous in $3$ or less dimensions, but singular in $4$ or more dimensions.
-In the unusual scenario where one is inverting for a parameter field in $4$ or more spatial dimensions, one could use a modified Poisson interpolation where $\Delta$ is replaced with $\Delta^\alpha$ in the objective function of \eqref{eq:wi_optimization_problem}, where $\alpha>1$ is chosen so that the fundamental solution of $\Delta^{2\alpha}$ is continuous.}
-
-The \emph{point spread function} (PSF) of $\mathcal{A}$ at $x\in \Omega$ is
-\begin{equation*}
-\varphi_x(y) := \langle \mathcal{A},\delta_x\rangle^*(y+x) = A(y+x,x).
-\end{equation*}
-Here $\langle \mathcal{A}, \delta_x \rangle \in L^2(\Omega)'$ is the result of applying $\mathcal{A}$ to a point source $\delta_x$ (delta distribution) centered at $x$, and $\langle\mathcal{A},\delta_x\rangle^* \in L^2(\Omega)$ is the Riesz representation of $\langle\mathcal{A},\delta_x\rangle$.
-The PSF, $\varphi_x$, is the result of translating $\langle\mathcal{A},\delta_x\rangle^*$ to re-center it at zero instead of $x$.
-Recall the Riesz representation of a linear functional $\psi \in L^2(\Omega)'$ is the unique vector $\psi^* \in L^2(\Omega)$ satisfying $\left(\psi^*, v\right)_{L^2(\Omega)} = \psi(v)$ for all $v \in \L^2(\Omega)$.
-
-The approximation $\widetilde{A}(y,x)$ may not be accurate if both $x$ and $y$ are near the boundary.
-Other approximation methods, if available, should be considered for the block of the operator associated with boundary-boundary interactions.
-Nevertheless, in our numerical results we see that approximation works well for the the inverse problem considered, without any modifications to the boundary-boundary block.
-
-The solution to this system of equations is found by the following process:
-\begin{enumerate}
-	\item Compute 
-	\begin{equation*}
-	\Psi := K^+ E^T
-	\end{equation*}
-	by solving Neumann poisson problems for point source right hand sides, with point sources located at the points $x_i$, $i=1,\dots,q$.
-	\item Compute 
-	\begin{equation*}
-	\Theta := K^+ M \Psi
-	\end{equation*}
-	by solving Neumann Poisson problems with the columns of $\Psi$ as distributed sources.
-	\item Form 
-	\begin{equation*}
-	S = \Psi^T M \Psi
-	\end{equation*}.
-	\item Compute \begin{equation*}
-	\alpha = \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{y}.
-	\end{equation*}
-	\item Compute 
-	\begin{equation*}
-	\boldsymbol{\lambda} = -S^{-1} \left(\mathbf{y} - \alpha \mathbf{1} \right).
-	\end{equation*}
-	\item Compute
-	\begin{equation*}
-	\mathbf{u} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c}
-	\end{equation*}
-\end{enumerate}
-
-Consider the Pine island glacier geometry shown in Figure REF.
-Regions of the glacier are separated by thin inlets of ocean.
-With radial basis function interpolation, points on one side of the inlet inappropriately influence the interpolation on the other side of the inlet.
-With Poisson interpolation, the PSF samples only influence the interpolation on their own side of the inlet.
-
-\subsection{Computation of Poisson weighting functions}
-\label{sec:computation_of_weighting_functions}
-
-After discretization, optimization problem \eqref{eq:wi_optimization_problem} becomes
-\begin{equation}
-\label{eq:discrete_poisson_optimization_1}
-\begin{aligned}
-\min_{\mathbf{w}} &\quad \frac{1}{2}\mathbf{w}^T K^T M^{-1} K \mathbf{w} \\
-\text{such that} &\quad E \mathbf{w} = \mathbf{d}.
-\end{aligned}
-\end{equation}
-Here $\mathbf{w}$ is the discretization of $w_i$, $K$ is the stiffness matrix for the discretized Poisson problem with pure Neumann boundary conditions, $M$ is the mass matrix, and $\mathbf{d}=(0,\dots,0,1,0,\dots,0)$ is the length-$q$ vector with $i^\text{th}$ entry one and all other entries zero.
-Furthermore, $E$ is the pointwise observation matrix defined by $\left(E\mathbf{u}\right)_j = u(x_j)$, where $u$ denotes the function represented by the vector $\mathbf{u}$.
-Here we dropped the subscripts $i$ from variables for notational convenience.
-One must solve $q$ optimization problems of the form \eqref{eq:discrete_poisson_optimization_1}; one for each $i=1,\dots,q$.
-
-In the remainder of this section we will derive an efficient algorithm for solving \eqref{eq:discrete_poisson_optimization_1}.
-The algorithm (Algorithm REF) is based on explicit formula for the solution of \eqref{eq:discrete_poisson_optimization_1} which we present in Proposition \ref{prop:poisson_interpolation_formula}.
-The formulas in Proposition \ref{prop:poisson_interpolation_formula} rely on certain matrices that we define in Definition \ref{defn:poisson_interpolation_matrices}.
-The primary computational cost of the algorithm is the solution of Poisson problems with pure Neumann boundary conditions.
-
-\begin{defn}
-	\label{defn:poisson_interpolation_matrices}
-	Let 
-	\begin{align*}
-	\Psi :=& K^+ E^T, \\
-	\Theta :=& K^+ M \Psi \\
-	S :=& \Psi^T M \Psi,
-	\end{align*}
-	where $K^+$ is the Moore-Penrose pseudoinverse of $K$.
-Further, let $\mathbf{c}$ dente the discretization of the constant function $c(x)=1$ on $\Omega$, and let $\mathbf{1}:=(1,\dots,1)$ denote the vector of length $q$ with all entries equal to one.
-\end{defn}
-
-\begin{remark}
-	Computing $\mathbf{g} = K^+\mathbf{f}$ is equivalent to solving the discretized Poisson problem with $f$ as the right hand side source, with pure Neumann boundary conditions, and with the constraint that the solution has average value equal to zero.
-That is,
-	\begin{equation}
-	\label{eq:pure_neumann_poisson}
-	\mathbf{g} = K^+ \mathbf{f} \quad \Leftrightarrow \quad 
-	\begin{cases}
-	\Delta g = f, & \text{in }\Omega,\\
-	\nu \cdot \nabla g = 0, & \text{on }\partial \Omega,\\
-	\int_\Omega g(x) dx = 0,
-	\end{cases}
-	\end{equation}
-	where $\nu$ denotes the normal to the boundary.
-
-	Hence, the $i^\text{th}$ column of $\Psi$ is the result of solving the discretized pure Neumann Poisson problem, with a point source at location $x_i$.
-The $i^\text{th}$ column of $\Theta$ is the result of solving the discretized pure Neumann Poisson problem with distributed source given by the $i^\text{th}$ column of $\Psi$.
-\end{remark}
-
-\begin{prop}
-	\label{prop:poisson_interpolation_formula}
-	The solution to optimization problem \eqref{eq:discrete_poisson_optimization_1} is given by
-	\begin{equation*}
-	\mathbf{w} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c},
-	\end{equation*}
-	where
-	\begin{align*}
-	\alpha =& \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{d} \\
-	\boldsymbol{\lambda} =& -S^{-1} \left(\mathbf{d} - \alpha \mathbf{1} \right). \\
-	\end{align*}
-\end{prop}
-
-\begin{proof}
-	Let
-	\begin{equation*}
-	v := \mathcal{K} w_i,
-	\end{equation*}
-	and for the remainder of this proof let $e := e_i$.
-	Since the null-space of the Laplacian with pure Neumann boundary conditions is the set of all constant functions, we have
-	\begin{equation}
-	\label{eq:definition_of_w}
-	w_i = \mathcal{K}^+ v + \alpha C
-	\end{equation}
-	for some scalar $\alpha$.
-
-	Writing optimization problem \eqref{eq:wi_optimization_problem} in terms of $v$ and $\alpha$ instead of $w_i$ yields the following equivalent optimization problem:
-	\begin{equation}
-	\label{eq:discrete_poisson_optimization_2}
-	\begin{aligned}
-	\min_{v \in L^2(\Omega)', \alpha \in \mathbb{R}} &\quad \frac{1}{2} v\left(\mathcal{M}^{-1} v\right) \\
-	\text{such that} &\quad \mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} = e.
-	\end{aligned}
-	\end{equation}
-	The Lagrangian for optimization problem \eqref{eq:discrete_poisson_optimization_2} is
-	\begin{equation*}
-	\mathcal{L} = \frac{1}{2}v\left( \mathcal{M}^{-1} v\right) + \lambda^T\left(\mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} - e\right).
-	\end{equation*}
-	The solution to this optimization problem is the stationary point of the Lagrangian, i.e., the point at which the gradient of the Lagrangian is zero:
-	\begin{equation*}
-	0 = \nabla \mathcal{L} = 
-	\begin{bmatrix}
-	\nabla_{v} \mathcal{L} \\
-	\nabla_{\lambda} \mathcal{L} \\
-	\nabla_\alpha \mathcal{L}
-	\end{bmatrix} 
-	= \begin{bmatrix}
-	\mathcal{M}^{-1} v + \left(\mathcal{K}^+\right)^T \mathcal{P}^T \lambda \\
-	\mathcal{P}\mathcal{K}^+v + \alpha \mathbf{1} - e \\
-	\mathbf{1}^T \lambda
-	\end{bmatrix}.
-	\end{equation*}
-	This equation may be rewritten as the following block $3 \times 3$ linear system:
-	\begin{equation*}
-	\begin{bmatrix}
-	\mathcal{M}^{-1} & \left(\mathcal{K}^+\right)^T \mathcal{P}^T & 0 \\
-	\mathcal{P}\mathcal{K}^+ & 0 & \mathbf{1} \\
-	0 & \mathbf{1}^T & 0
-	\end{bmatrix}
-	\begin{bmatrix}
-	v \\ \lambda \\ \alpha
-	\end{bmatrix}
-	=
-	\begin{bmatrix}
-	0 \\ e \\ 0
-	\end{bmatrix}.
-	\end{equation*}
-	Performing block Gaussian elimination on this system allows us to reduce the system to the following block triangular form
-	\begin{equation*}
-	\begin{bmatrix}
-	\mathcal{M}^{-1} & \left(\mathcal{K}^+\right)^T \mathcal{P}^T & 0 \\
-	& -S & \mathbf{1} \\
-	0 & 0 & \mathbf{1}^T S^{-1} \mathbf{1}
-	\end{bmatrix}
-	\begin{bmatrix}
-	v \\ \lambda \\ \alpha
-	\end{bmatrix}
-	=
-	\begin{bmatrix}
-	0 \\ e \\ \mathbf{1}^T S^{-1} e
-	\end{bmatrix}
-	\end{equation*}
-	where we recall that $S := \Psi^T \mathcal{M} \Psi = \mathcal{P}\mathcal{K}^+ \mathcal{M} \left(\mathcal{K}^+\right)^T \mathcal{P}^T$. From this block triangular system, we read off the solution as:
-	\begin{align*}
-	\alpha =& \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{d} \\
-	\boldsymbol{\lambda} =& -S^{-1} \left(\mathbf{d} - \alpha \mathbf{1} \right) \\
-	\mathbf{v} =& -M K^+ E^T \boldsymbol{\lambda}
-	\end{align*}
-	Substituting these results into the definition of $\mathbf{w}$ in \eqref{eq:definition_of_w}, we have
-	\begin{equation*}
-	\mathbf{w} = -K^+ M K^+ E^T \boldsymbol{\lambda} + \alpha \mathbf{c} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c},
-	\end{equation*}
-	as required.
-\end{proof}
-
-We now design an incremental algorithm for constructing interpolants $\mathbf{u}$, in which we can add points $x_i$ more efficiently. Let
-\begin{align*}
-\Psi :=& K^+ E^T \\
-\Theta :=& K^+ M \Psi
-\end{align*}
-so that 
-\begin{equation*}
-S = \Psi^T M \Psi
-\end{equation*}
-and
-\begin{equation*}
-\mathbf{u} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c}
-\end{equation*}
-The columns of $\Psi$ are the responses of $A^+$ to point sources (delta distributions) centered at the sample points $x_i$.
-The columns of $\Theta$ are the responses of $A^+M^{-1} A$ to these point sources.
-The matrices $\Psi$ and $\Theta$ may be constructed incrementally as new points are added.
-
 \section*{Acknowledgments}
-We thank J.J. Alger, Longfei Gao, and Rami Nammour for helpful discussions.
+We thank J.J. Alger, Longfei Gao, Mathew Hu, and Rami Nammour for helpful discussions.
 

### 2021-09-14 paper/localpsf.tex
paper cleanup

@@
 \begin{AMS}
-  68Q25, 68R10, 68U05
+  ASDF
 \end{AMS}
@@
 \begin{description}
-\item[1) Matrix-free:] One cannot easily evaluate kernel entries $\Phi(y,x)$.
+\item[1) Matrix-free:] One cannot easily evaluate kernel entries $\Aker(y,x)$.
 Instead, access to $\Aop$ is available through ``operator actions.'' That is, we have a black-box computational procedure (which is typically expensive) through which we may evaluate the maps
 \begin{equation*}
-	u \mapsto\Aop u \quad \text{and} \quad v \mapsto\Aop^Tv
+	u \mapsto\Aop u \quad \text{and} \quad v \mapsto \Aop^Tv
 \end{equation*} 
@@
 For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
-\item[2) Local support:] Let $\phi_x$ be the following \emph{impulse response}:
+\item[2) Local support:] Let $\impulseresponse_x$ be the following \emph{impulse response}:
 \begin{equation}
 \label{eq:impulse_response_defn}
-\phi_x(y) := \Phi(y, x).
-\end{equation} 
-We say that $\Aop$ has local support if, for all $x\in \Omega$, the support of $\phi_x$ is contained (or approximately contained) in a neighborhood of $x$.
+\impulseresponse_x(y) := \Aker(y, x).
+\end{equation} 
+We say that $\Aop$ has local support if, for all $x\in \Omega$, the support of $\impulseresponse_x$ is contained (or approximately contained) in a neighborhood of $x$.
 The smaller these supports are, the better our algorithm will perform.
 Let $\delta_x$ denote the delta distribution (point source) centered at $x$.
-The function $\phi_x$ is called the impulse response because straightforward analysis shows that it may be expressed as follows:
+The function $\impulseresponse_x$ is called the impulse response because straightforward analysis shows that it may be expressed as follows:
 \begin{equation}
 	\label{eq:impulse_response_delta_action}
-\phi_x = \left( \Aop \delta_x^* \right)^*,
-\end{equation}
-where $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying the operator $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\psi \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\psi^* \in L^2(\Omega)$ such that $\psi(v) = \left(\psi^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Phi$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
+	\impulseresponse_x = \left( \Aop \delta_x^* \right)^*,
+\end{equation}
+where $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying the operator $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\psi \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\psi^* \in L^2(\Omega)$ such that $\psi(v) = \left(\psi^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
 This is described in Appendix \ref{app:distributions}.
@@
 	\label{eq:local_translation_invariance}
-	\Phi(y+h, x+h) \approx \Phi(y,x)
+	\Aker(y+h, x+h) \approx \Aker(y,x)
 \end{equation}
 when $h$ is not too large.
-One can imagine the kernel entry $\Phi(y,x)$ as representing the strength of an interaction between a source at point $x$ and a target at point $y$.
+One can imagine the kernel entry $\Aker(y,x)$ as representing the strength of an interaction between a source at point $x$ and a target at point $y$.
 Local translation invariance means that the interaction strength remains roughly the same if the source and target points are both translated by the same vector.
@@
 \begin{equation*}
-\Phi(y,x) \ge 0.
+\Aker(y,x) \ge 0.
 \end{equation*}
@@
 \begin{equation*}
-y=f(m,u(m)) + \text{noise},
+	y = f(m, u(m)) + \text{noise},
 \end{equation*}
@@
 	\label{eq:state_pde}
-0=g(m,u).
+	0 = g(m,u).
 \end{equation}
@@
 	\label{eq:convolution_kernel}
-	\varphi_x(y) := \phi_x(y+x).
-\end{equation}
-The convolution kernel $\varphi_x$ is the result of translating the impulse response $\phi_x$ to re-center it at zero instead of $x$.
+	\varphi_x(y) := \impulseresponse_x(y+x).
+\end{equation}
+The convolution kernel $\varphi_x$ is the result of translating the impulse response $\impulseresponse_x$ to re-center it at zero instead of $x$.
 Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes.
@@
 \begin{equation*}
-\varphi_i := \varphi_{x_i} \quad \text{and} \quad \phi_i := \phi_{x_i},
-\end{equation*}
-with $\varphi_{x_i}$ defined in \eqref{eq:convolution_kernel} and $\phi_{x_i}$ defined in \eqref{eq:impulse_response_defn}, to denote local convolution kernels and impulse responses corresponding to a collection of points $\{x_i\}_{i=1}^r \subset \Omega$ scattered throughout the domain.
+	\varphi_i := \varphi_{x_i} \quad \text{and} \quad \impulseresponse_i := \impulseresponse_{x_i},
+\end{equation*}
+with $\varphi_{x_i}$ defined in \eqref{eq:convolution_kernel} and $\impulseresponse_{x_i}$ defined in \eqref{eq:impulse_response_defn}, to denote local convolution kernels and impulse responses corresponding to a collection of points $\{x_i\}_{i=1}^r \subset \Omega$ scattered throughout the domain.
 The functions $w_i$ are spatially varying weighting functions that are used to interpolate the convolution kernels $\varphi_i$.
@@
 We compute one ``batch'' of $\varphi_i$'s by applying $\Aop$ to a sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$.
-The batch of points $x_i$ are chosen so that the support of $\phi_i$ and the support of $\phi_j$ do not overlap (or do not overlap much) if $i \neq j$.
-Because these supports do not overlap, we can post-process the response of $\Aop$ to the Dirac comb to recover the functions $\phi_i$, and therefore the functions $\varphi_i$, associated with all points $x_i$ in the batch---with one application of $\Aop$, we recover many $\varphi_i$.
+The batch of points $x_i$ are chosen so that the support of $\impulseresponse_i$ and the support of $\impulseresponse_j$ do not overlap (or do not overlap much) if $i \neq j$.
+Because these supports do not overlap, we can post-process the response of $\Aop$ to the Dirac comb to recover the functions $\impulseresponse_i$, and therefore the functions $\varphi_i$, associated with all points $x_i$ in the batch---with one application of $\Aop$, we recover many $\varphi_i$.
 This is illustrated in Figure FIG.
@@
 
-In order to choose the points $x_i$, we need to estimate the support of $\phi_i$ \emph{before} we compute it.
-The supports of the functions $\phi_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of specially chosen functions (three functions if $d=1$, six if $d=2$, ten if $d=3$), then post processing the results (Section \ref{eq:mean_and_covariance_estimation}).
+In order to choose the points $x_i$, we need to estimate the support of $\impulseresponse_i$ \emph{before} we compute it.
+The supports of the functions $\impulseresponse_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of specially chosen functions (three functions if $d=1$, six if $d=2$, ten if $d=3$), then post processing the results (Section \ref{eq:mean_and_covariance_estimation}).
 This procedure is the only part of the paper that requires non-negativity of the integral kernel.
@@
 Once the $\varphi_i$ and $w_i$ are computed, we convert $\AopPc$ to hierarchical matrix (H-matrix) format.
-Let $\Phi_\text{pc}$ denote the integral kernel associated with $\AopPc$.
-While $\Phi(y,x)$ is not easily computable, $\Phi_\text{pc}(y,x)$ is given by the formula
+Let $\Aker_\text{pc}$ denote the integral kernel associated with $\AopPc$.
+While $\Aker(y,x)$ is not easily computable, $\Aker_\text{pc}(y,x)$ is given by the formula
 \begin{equation}
 	\label{eq:kernel_entries}
-	\Phi_\text{pc}(y,x) = \sum_{i=1}^r \varphi_i(y-x) w_i(x).
+	\AkerPc(y,x) = \sum_{i=1}^r \varphi_i(y-x) w_i(x).
 \end{equation}
@@
 A greedy algorithm is used to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other.
-We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\phi_x$ (Section \ref{eq:mean_and_covariance_estimation}).
-The support of $\phi_x$ is estimated to be contained within an ellipsoid, so the process of choosing batches of points is an ellipsoid packing problem (Section \ref{sec:greedy_point_selection}).
+We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\impulseresponse_x$ (Section \ref{eq:mean_and_covariance_estimation}).
+The support of $\impulseresponse_x$ is estimated to be contained within an ellipsoid, so the process of choosing batches of points is an ellipsoid packing problem (Section \ref{sec:greedy_point_selection}).
 
@@
 
-Because of the non-negative kernel property, for each $x \in \Omega$, the impulse response $\phi_x$ is a non-negative function $y \mapsto \phi_x(y)$.
-Hence, $\phi_x$ is a scaled probability distribution.
-Let $\alpha$ denote the scaling factor, let $\widehat{\phi}_x := \phi_x / \alpha(x)$ denote the normalized version of $\phi_x$, and let $\mu(x)$ and $\Sigma(x)$ denote the mean and covariance of $\widehat{\phi}_x$, respectively.
-We make the approximation that the support of $\phi_x$ is contained within the ellipsoid
+Because of the non-negative kernel property, for each $x \in \Omega$, the impulse response $\impulseresponse_x$ is a non-negative function $y \mapsto \impulseresponse_x(y)$.
+Hence, $\impulseresponse_x$ is a scaled probability distribution.
+Let $\alpha$ denote the scaling factor, let $\widehat{\impulseresponse}_x := \impulseresponse_x / \alpha(x)$ denote the normalized version of $\impulseresponse_x$, and let $\mu(x)$ and $\Sigma(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
+We make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
 \begin{equation}
@@
 where $\tau$ is a fixed constant.
-The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\phi}_x$ (see Figure FIG).
+The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\impulseresponse}_x$ (see Figure FIG).
 The quantity $\tau$ is a parameter that must be chosen by the user of the algorithm.
-The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\phi_x$.
+The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$.
 However, the cost of our algorithm will depend on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
 We find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results.
-The fraction of the ``mass'' of $\phi_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.
+The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.
 
@@
 To use $E_x$, we must first compute the scalar $\alpha(x)$, the length-$d$ vector $\mu(x)$, and the $d \times d$ matrix $\Sigma(x)$ (recall $d \in \{1,2,3\}$ is the spatial dimension of $\Omega$).
-The direct approach to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\phi_x$, as per \eqref{eq:impulse_response_delta_action}.
-Then one can post-process $\phi_x$ to determine $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$.
+The direct approach to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}.
+Then one can post-process $\impulseresponse_x$ to determine $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$.
 But this direct approach is computationally infeasible because and it requires one operator action of $\Aop$ per point $x$ that we wish to evaluate $E_x$ at, and we will need to evaluate $E_x$ at a large number of points.
@@
 	\begin{equation*}
-	\alpha(x) := \int_{\Omega} \phi_x(y) dy, \qquad
-	\mu(x) := \mathbb{E}(\widehat{\phi}_x), \qquad
-	\Sigma(x) := \Var(\widehat{\phi}_x),
-	\end{equation*}
-	where $\mathbb{E}(\widehat{\phi}_x)$ denotes the mean of $\widehat{\phi}_x$, and $\Var(\widehat{\phi}_x)$ denotes the variance of $\widehat{\phi}_x$.	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$.
+		\alpha(x) := \int_{\Omega} \impulseresponse_x(y) dy, \qquad
+		\mu(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
+		\Sigma(x) := \Var(\widehat{\impulseresponse}_x),
+	\end{equation*}
+	where $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$.
 Let $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions:
@@
 These sample point batches will be used in Section \ref{sec:get_impulse_response} to compute many impulse responses of $\Aop$ using only a small number of operator actions of $\Aop$.
-The support of $\phi_x$ is (approximately) contained in the ellipsoid $E_x$, so by applying $\mathcal{A}$ to the Dirac comb associated with all sample points in a batch, we will recover the impulse responses for all sample points in that batch.
+The support of $\impulseresponse_x$ is (approximately) contained in the ellipsoid $E_x$, so by applying $\mathcal{A}$ to the Dirac comb associated with all sample points in a batch, we will recover the impulse responses for all sample points in that batch.
 
@@
 We compute convolution kernels $\varphi_{x_i}$ in batches by applying $\Aop$ to a small number Dirac combs.
-Each application of $\Aop$ to a Dirac comb yields one batch of impulse responses, $\phi_{x_i}$, which yields one batch of convolution kernels, $\varphi_{x_i}$.
+Each application of $\Aop$ to a Dirac comb yields one batch of impulse responses, $\impulseresponse_{x_i}$, which yields one batch of convolution kernels, $\varphi_{x_i}$.
 The Dirac comb, $\dcomb_b$, associated with a batch of sample points, $S_b$, is the sum of Dirac distributions (point sources) centered at the points $x_i \in S_b$.
@@
 	\label{eq:phi_b}
-	\combresponse_b = \left\langle\mathcal{A},\sum_{x_i \in S_b} \delta_{x_i}\right\rangle^* = \sum_{x_i \in S_b} \langle\mathcal{A}, \delta_{x_i}\rangle^* = \sum_{x_i \in S_b} \phi_{x_i}.
-\end{equation}
-Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\phi_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\phi_j$ do not overlap when $i \neq j$, from \eqref{eq:convolution_kernel} and \eqref{eq:phi_b} we have (approximately)
+	\combresponse_b = \left\langle\mathcal{A},\sum_{x_i \in S_b} \delta_{x_i}\right\rangle^* = \sum_{x_i \in S_b} \langle\mathcal{A}, \delta_{x_i}\rangle^* = \sum_{x_i \in S_b} \impulseresponse_{x_i}.
+\end{equation}
+Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\impulseresponse_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\impulseresponse_j$ do not overlap when $i \neq j$, from \eqref{eq:convolution_kernel} and \eqref{eq:phi_b} we have (approximately)
 \begin{equation}
 \label{eq:varphi_eval}
-	\varphi_{x_i}(z) = \phi_i(z+x_i) = \begin{cases}
+	\varphi_{x_i}(z) = \impulseresponse_i(z+x_i) = 
+	\begin{cases}
 		\combresponse_b(z+x_i), & z+x_i \in E_{x_i} \\
@@
 
-For each $i=1,\dots,r$, optimization problem \eqref{eq:wi_optimization_problem} is an example of the following generic omtimization problem:
+For each $i=1,\dots,r$, optimization problem \eqref{eq:wi_optimization_problem} is an example of the following generic optimization problem:
 \begin{equation}
@@
 \begin{enumerate}
-\item For each sample point $x_i$, we compute a function $\psi_i$ by solving the following Neumann Poisson problem:
+	\item For each sample point $x_i$, we compute a function $\firstgreens_i$ by solving the following Neumann Poisson problem:
 	\begin{equation}
@@
 		\begin{cases}
-\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
-\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
+			\Delta \firstgreens_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
+			\nu \cdot \nabla \firstgreens_i = 0 & \text{on } \partial \Omega,
 		\end{cases}
 	\end{equation}
-with condition $\int_\Omega \psi_i(x) dx = 0$.
+	with condition $\int_\Omega \firstgreens_i(x) dx = 0$.
 Here $\delta_{x_i}$ is the point source centered at $x_i$, $|\Omega|$ is the measure of the domain $\Omega$, and $\nu$ is the normal vector to $\partial \Omega$.
-\item For each function $\psi_i$, we compute a function $\secondgreens_i$ by solving another Neumann Poisson problem:
+	\item For each function $\firstgreens_i$, we compute a function $\secondgreens_i$ by solving another Neumann Poisson problem:
 	\begin{equation}
@@
 		\begin{cases}
-\Delta \secondgreens_i = \psi_i, & \text{in }\Omega, \\
+			\Delta \secondgreens_i = \firstgreens_i, & \text{in }\Omega, \\
 			\nu \cdot \nabla \secondgreens_i = 0 & \text{on } \partial \Omega,
@@
 		Form the $r \times r$ matrix $S$ given in \eqref{eq:S_matrix}.
-
-%		\tcp{Compute weighting functions $\{w_i\}_{i=1}^r$}
 
@@
 	\label{eq:rational_3_linear_system}
-\begin{bmatrix} 1 & 0 & 1/\mu \\ 0 & 1 & -1/\mu^2 \\ 1 & \lambda_\text{max} & 1/(\lambda_\text{max}+\mu) \end{bmatrix} \begin{bmatrix} c_0 \\ c_1 \\ c_2 \end{bmatrix} = \begin{bmatrix} 0 \\ 0 \\ \lambda_\text{max} \end{bmatrix},
+	\begin{bmatrix} 
+		1 & 0 & 1/\mu \\ 0 & 1 & -1/\mu^2 \\ 1 & \lambda_\text{max} & 1/(\lambda_\text{max}+\mu) 
+	\end{bmatrix}
+	\begin{bmatrix} 
+		c_0 \\ c_1 \\ c_2 
+	\end{bmatrix}
+	=
+	\begin{bmatrix} 
+		0 \\ 0 \\ \lambda_\text{max} 
+	\end{bmatrix},
 \end{equation}
@@
 
-\begin{defn}[Moments of $\phi_x$]
+\begin{defn}[Moments of $\impulseresponse_x$]
 	\label{defn:alpha_rho_mu_sigma}
-	We define the spatially varying scaling factor, $\alpha$, the scaled impulse response, $\widehat{\phi}_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
+	We define the spatially varying scaling factor, $\alpha$, the scaled impulse response, $\widehat{\impulseresponse}_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
 	\begin{align*}
-	\alpha:\Omega \rightarrow \mathbb{R}, &\qquad \alpha(x) := \int_{\Omega} \phi_x(z) dz, \\
-	\widehat{\phi}_x:\Omega \rightarrow \mathbb{R}, &\qquad \widehat{\phi}_x := \phi_x \big/ \alpha(x), \\
-	\mu:\Omega \rightarrow \mathbb{R}^d, &\qquad \mu(x) := \int_\Omega z \widehat{\phi}_x(z) dz, \\
-	\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\Sigma(x) := \int_\Omega (z - \mu(x))(z - \mu(x))^T \widehat{\phi}_x(z) dz.
+		\alpha:\Omega \rightarrow \mathbb{R}, &\qquad \alpha(x) := \int_{\Omega} \impulseresponse_x(z) dz, \\
+		\widehat{\impulseresponse}_x:\Omega \rightarrow \mathbb{R}, &\qquad \widehat{\impulseresponse}_x := \impulseresponse_x \big/ \alpha(x), \\
+		\mu:\Omega \rightarrow \mathbb{R}^d, &\qquad \mu(x) := \int_\Omega z \widehat{\impulseresponse}_x(z) dz, \\
+		\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\Sigma(x) := \int_\Omega (z - \mu(x))(z - \mu(x))^T \widehat{\impulseresponse}_x(z) dz.
 	\end{align*}
@@
 	\begin{equation*}
-		\Sigma(x) = \int_\Omega zz^T \widehat{\phi}_x(z) dz - \mu(x) \mu(x)^T,
+		\Sigma(x) = \int_\Omega zz^T \widehat{\impulseresponse}_x(z) dz - \mu(x) \mu(x)^T,
 	\end{equation*}
@@
 \end{equation}
-Recall that the impulse response of $\mathcal{A}$ to $\delta_x$, denoted $\phi_x$, is the Riesz representation of $\langle\mathcal{A},\delta_x\rangle$.
-From \eqref{eq:appendix_impulse_response} and the definition of the Riesz representation, we have $\phi_x = \langle\mathcal{A},\delta_x\rangle^* = A(~\cdot~,x)$.
-That is, $\phi_x$ is the function $y \mapsto A(y,x)$.
+Recall that the impulse response of $\mathcal{A}$ to $\delta_x$, denoted $\impulseresponse_x$, is the Riesz representation of $\langle\mathcal{A},\delta_x\rangle$.
+From \eqref{eq:appendix_impulse_response} and the definition of the Riesz representation, we have $\impulseresponse_x = \langle\mathcal{A},\delta_x\rangle^* = A(~\cdot~,x)$.
+That is, $\impulseresponse_x$ is the function $y \mapsto A(y,x)$.
 
@@
 Here we list the function space operations on $V_h$ and $V_h'$ that are required for this paper, and show how to perform the corresponding matrix and vector operations on $\mathbb{R}^N_{\mathbf{M}}$ and $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
-
-%We summarize coefficient vector versions of the function space operation that are used in this paper in Appendix \ref{app:discretized_operations}.
-%
-%As described in Section \ref{sec:discretization}, a function $u_h$ in a finite element space $V_h$ may be written in terms of its vector of coefficients, $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, with respect the finite element basis $\{\phi_i\}_{i=1}^N$.
-Recall that $\mathbf{M}$ is the mass matrix.
-We have
-%\begin{equation*}
-%	u(x) \approx u_h(x) = \sum_{i=1}^N \mathbf{u}_i \phi_i(x).
-%\end{equation*}
-%The finite element function space $V_h$ is isometrically isomorphic to the coefficient space $\mathbb{R}^N_\mathbf{M}$, and the dual space $V_h'$ is isometrically isomorphic to $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
-In computations, we perform operations with coefficient vectors in $\mathbb{R}^N_\mathbf{M}$ and coefficient dual vectors in $\mathbb{R}^N_{\mathbf{M}^{-1}}$, rather than functions in $V_h$ and functionals in $V_h'$.
-Here we describe how to perform common function space operations in terms of coefficient vectors and coefficient dual vectors.
 

### 2021-09-14 paper/localpsf.tex
paper cleanup

@@
 
-If $\Aop$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then it is straightforward to show that $\Aop$ would be the convolution operator $u \mapsto \left(\varphi_x \ast u\right)^*$, where the convolution kernel is given by
+If $\Aop$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then it is straightforward to show that $\Aop$ would be the convolution operator $u \mapsto \left(\convkernel_x \ast u\right)^*$, where the convolution kernel is given by
 \begin{equation}
 	\label{eq:convolution_kernel}
-	\varphi_x(y) := \impulseresponse_x(y+x).
-\end{equation}
-The convolution kernel $\varphi_x$ is the result of translating the impulse response $\impulseresponse_x$ to re-center it at zero instead of $x$.
-Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes.
+	\convkernel_x(y) := \impulseresponse_x(y+x).
+\end{equation}
+The convolution kernel $\convkernel_x$ is the result of translating the impulse response $\impulseresponse_x$ to re-center it at zero instead of $x$.
+Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\convkernel_x$, varies as $x$ changes.
 We therefore approximate $\Aop$ by a spatially varying weighted sum of convolution operators, $\AopPc \approx \Aop$, of the form
@@
 \label{eq:product_convolution}
-	\left(\AopPc~u\right)(v) := \left( v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
+	\left(\AopPc~u\right)(v) := \left( v,~ \sum_{i=1}^r \convkernel_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
 \end{equation}
@@
 \begin{equation*}
-	\varphi_i := \varphi_{x_i} \quad \text{and} \quad \impulseresponse_i := \impulseresponse_{x_i},
-\end{equation*}
-with $\varphi_{x_i}$ defined in \eqref{eq:convolution_kernel} and $\impulseresponse_{x_i}$ defined in \eqref{eq:impulse_response_defn}, to denote local convolution kernels and impulse responses corresponding to a collection of points $\{x_i\}_{i=1}^r \subset \Omega$ scattered throughout the domain.
-The functions $w_i$ are spatially varying weighting functions that are used to interpolate the convolution kernels $\varphi_i$.
-Our operator approximation method is defined by how we compute the convolution kernels, $\varphi_i$, how we choose the points, $x_i$, and what weighting functions, $w_i$, that we use.
+	\convkernel_i := \convkernel_{x_i} \quad \text{and} \quad \impulseresponse_i := \impulseresponse_{x_i},
+\end{equation*}
+with $\convkernel_{x_i}$ defined in \eqref{eq:convolution_kernel} and $\impulseresponse_{x_i}$ defined in \eqref{eq:impulse_response_defn}, to denote local convolution kernels and impulse responses corresponding to a collection of points $\{x_i\}_{i=1}^r \subset \Omega$ scattered throughout the domain.
+The functions $w_i$ are spatially varying weighting functions that are used to interpolate the convolution kernels $\convkernel_i$.
+Our operator approximation method is defined by how we compute the convolution kernels, $\convkernel_i$, how we choose the points, $x_i$, and what weighting functions, $w_i$, that we use.
 In Section \ref{sec:overview_intro} we summarize the answers to these questions; we will provide detailed answers to these questions in Section \ref{sec:method}.
@@
 
-We compute one ``batch'' of $\varphi_i$'s by applying $\Aop$ to a sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$.
+We compute one ``batch'' of $\convkernel_i$'s by applying $\Aop$ to a sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$.
 The batch of points $x_i$ are chosen so that the support of $\impulseresponse_i$ and the support of $\impulseresponse_j$ do not overlap (or do not overlap much) if $i \neq j$.
-Because these supports do not overlap, we can post-process the response of $\Aop$ to the Dirac comb to recover the functions $\impulseresponse_i$, and therefore the functions $\varphi_i$, associated with all points $x_i$ in the batch---with one application of $\Aop$, we recover many $\varphi_i$.
+Because these supports do not overlap, we can post-process the response of $\Aop$ to the Dirac comb to recover the functions $\impulseresponse_i$, and therefore the functions $\convkernel_i$, associated with all points $x_i$ in the batch---with one application of $\Aop$, we recover many $\convkernel_i$.
 This is illustrated in Figure FIG.
-The process is repeated to get more batches of $\varphi_i$'s, until a desired number of batches is reached.
+The process is repeated to get more batches of $\convkernel_i$'s, until a desired number of batches is reached.
 
@@
 
-Once the $\varphi_i$ and $w_i$ are computed, we convert $\AopPc$ to hierarchical matrix (H-matrix) format.
+Once the $\convkernel_i$ and $w_i$ are computed, we convert $\AopPc$ to hierarchical matrix (H-matrix) format.
 Let $\Aker_\text{pc}$ denote the integral kernel associated with $\AopPc$.
@@
 	\label{eq:kernel_entries}
-	\AkerPc(y,x) = \sum_{i=1}^r \varphi_i(y-x) w_i(x).
-\end{equation}
-This follows from~\eqref{eq:product_convolution} and the fact that the convolution operator $u \mapsto \varphi_i \ast u$ has $(y,x)$ kernel entry given by $\varphi(y-x)$.
+	\AkerPc(y,x) = \sum_{i=1}^r \convkernel_i(y-x) w_i(x).
+\end{equation}
+This follows from~\eqref{eq:product_convolution} and the fact that the convolution operator $u \mapsto \convkernel_i \ast u$ has $(y,x)$ kernel entry given by $\convkernel(y-x)$.
 Formula~\eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\AopPc$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
@@
 
-In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\varphi_i$ and $w_i$ may be arbitrary functions.
+In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
 Product-convolution approximations have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}.
@@
 
-We build upon the class of product convolution approximations in which the functions, $\varphi_i$, are impulse responses of $\Aop$ to point sources at a collection of points $x_i$ \cite{PRODCONVLIST}.
-A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the functions $\varphi_i$ with piecewise linear interpolation \cite{NAGY}, or splines \cite{PRODCONVSPLINES}.
+We build upon the class of product convolution approximations in which the functions, $\convkernel_i$, are impulse responses of $\Aop$ to point sources at a collection of points $x_i$ \cite{PRODCONVLIST}.
+A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the functions $\convkernel_i$ with piecewise linear interpolation \cite{NAGY}, or splines \cite{PRODCONVSPLINES}.
 With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
@@
 But even with adaptive refinement, many matrix vector products with $\Aop$ may be required.
-In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\varphi_i$ are computed with each matrix vector product.
-
-Our method of estimating the support of the functions $\varphi_i$ was inspired by resolution analysis in seismic imaging \cite{RESOLUTION}.
-In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\varphi_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
-If the width of $\varphi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
-In \cite{RESOLUTION}, the width of $\varphi_p$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
+In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\convkernel_i$ are computed with each matrix vector product.
+
+Our method of estimating the support of the functions $\convkernel_i$ was inspired by resolution analysis in seismic imaging \cite{RESOLUTION}.
+In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\convkernel_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
+If the width of $\convkernel_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
+In \cite{RESOLUTION}, the width of $\convkernel_p$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
 In this paper, rather than probing $\Aop^T$ with random noise functions, we use specific constant, linear, and quadratic functions.
-Our method estimates the support of $\varphi_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
+Our method estimates the support of $\convkernel_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
 
@@
 \label{sec:method}
-In this section we describe how we choose the sample points, $x_i$ (Section \ref{sec:sample_point_selection}), how we form the convolution kernels, $\varphi_i$ (Section \ref{sec:get_impulse_response}), and how we form the weighting functions, $w_i$ (Section \ref{sec:weighting_functions}), for the product-convolution approximation, $\AopPc \approx \Aop$, given in \eqref{eq:product_convolution}.
+In this section we describe how we choose the sample points, $x_i$ (Section \ref{sec:sample_point_selection}), how we form the convolution kernels, $\convkernel_i$ (Section \ref{sec:get_impulse_response}), and how we form the weighting functions, $w_i$ (Section \ref{sec:weighting_functions}), for the product-convolution approximation, $\AopPc \approx \Aop$, given in \eqref{eq:product_convolution}.
 We also describe how we convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), and (optionally) how we modify the resulting H-matrix, $\AmatPc$, to force it to be symmetric positive semi-definite (Section \ref{sec:make_spd}).
@@
 
-		Form convolution kernels, $\varphi_{x_i}$, on regular grids (Section \ref{sec:convolution_kernel_regular_grid})
+		Form convolution kernels, $\convkernel_{x_i}$, on regular grids (Section \ref{sec:convolution_kernel_regular_grid})
 
@@
 
-We compute convolution kernels $\varphi_{x_i}$ in batches by applying $\Aop$ to a small number Dirac combs.
-Each application of $\Aop$ to a Dirac comb yields one batch of impulse responses, $\impulseresponse_{x_i}$, which yields one batch of convolution kernels, $\varphi_{x_i}$.
+We compute convolution kernels $\convkernel_{x_i}$ in batches by applying $\Aop$ to a small number Dirac combs.
+Each application of $\Aop$ to a Dirac comb yields one batch of impulse responses, $\impulseresponse_{x_i}$, which yields one batch of convolution kernels, $\convkernel_{x_i}$.
 The Dirac comb, $\dcomb_b$, associated with a batch of sample points, $S_b$, is the sum of Dirac distributions (point sources) centered at the points $x_i \in S_b$.
@@
 \label{eq:varphi_eval}
-	\varphi_{x_i}(z) = \impulseresponse_i(z+x_i) = 
+	\convkernel_{x_i}(z) = \impulseresponse_i(z+x_i) = 
 	\begin{cases}
@@
 \end{equation}
-for all $x_i \in S_b$. By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the convolution kernels $\varphi_{x_i}$ associated with every point $x_i \in S_b$.
+for all $x_i \in S_b$. By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the convolution kernels $\convkernel_{x_i}$ associated with every point $x_i \in S_b$.
 
@@
 
-To convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), we will need to evaluate $\varphi_{x_i}(z)$ for a large number of points $z$.
-Per \eqref{eq:varphi_eval}, evaluating $\varphi_{x_i}(z)$ requires evaluating $\combresponse_b(z+x_i)$.
+To convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), we will need to evaluate $\convkernel_{x_i}(z)$ for a large number of points $z$.
+Per \eqref{eq:varphi_eval}, evaluating $\convkernel_{x_i}(z)$ requires evaluating $\combresponse_b(z+x_i)$.
 In computations, $\combresponse_b$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh.
@@
 
-To speed up evaluation of $\varphi_{x_i}$, we interpolate $\varphi_{x_i}$ onto a regular rectilinear grid on a coordinate axis aligned box that just barely contains the ellipsoid $E_{x_i}$.
-We transfer the function $\varphi_{x_i}$ from the irregular mesh to the regular grid by evaluating $\varphi_{x_i}$ at all gridpoints in the regular grid, using the slow process described in the previous paragraph.
-This is an upfront cost which is done ``offline.'' After $\varphi_{x_i}$ has been transferred to the regular grid, we use fast regular grid interpolation to perform all subsequent ``online'' evaluations of $\varphi_{x_i}$.
-This yields a substantial overall speedup because the number of finite element function evaluations required to transfer the functions $\varphi_{x_i}$ to the regular grid is $O(N)$, while the number of kernel evaluations required to convert $\AopPc$ to H-matrix format is $O(k N \log N)$.
+To speed up evaluation of $\convkernel_{x_i}$, we interpolate $\convkernel_{x_i}$ onto a regular rectilinear grid on a coordinate axis aligned box that just barely contains the ellipsoid $E_{x_i}$.
+We transfer the function $\convkernel_{x_i}$ from the irregular mesh to the regular grid by evaluating $\convkernel_{x_i}$ at all gridpoints in the regular grid, using the slow process described in the previous paragraph.
+This is an upfront cost which is done ``offline.'' After $\convkernel_{x_i}$ has been transferred to the regular grid, we use fast regular grid interpolation to perform all subsequent ``online'' evaluations of $\convkernel_{x_i}$.
+This yields a substantial overall speedup because the number of finite element function evaluations required to transfer the functions $\convkernel_{x_i}$ to the regular grid is $O(N)$, while the number of kernel evaluations required to convert $\AopPc$ to H-matrix format is $O(k N \log N)$.
 Here $N$ is the number of finite element degrees of freedom, and $k$ is the H-matrix rank.
-Furthermore, transferring $\varphi_{x_i}$ to a regular grid requires evaluating $\varphi_{x_i}$ at a predetermined and highly structured collection of points.
+Furthermore, transferring $\convkernel_{x_i}$ to a regular grid requires evaluating $\convkernel_{x_i}$ at a predetermined and highly structured collection of points.
 This structure allows for faster determination of which mesh cells that points are in, and allows one to evaluate the finite element function at all regular gridpoints in parallel.
-In contrast, constructing the H-matrix requires evaluating $\varphi_{x_i}$ at an unpredictable sequence of scattered of points, which is an expensive task for finite element function evaluation, but a cheap task for regular grid interpolation.
-Having $\varphi_{x_i}$ stored on a regular grid also simplifies the implementation of the method that we use to deal with boundary artifacts, which will be described in Section \ref{sec:boundary_extension}.
-
-We incur interpolation error when transferring $\varphi_{x_i}$ from the irregular mesh to the regular grid.
+In contrast, constructing the H-matrix requires evaluating $\convkernel_{x_i}$ at an unpredictable sequence of scattered of points, which is an expensive task for finite element function evaluation, but a cheap task for regular grid interpolation.
+Having $\convkernel_{x_i}$ stored on a regular grid also simplifies the implementation of the method that we use to deal with boundary artifacts, which will be described in Section \ref{sec:boundary_extension}.
+
+We incur interpolation error when transferring $\convkernel_{x_i}$ from the irregular mesh to the regular grid.
 To limit this interpolation error, the density of the regular grid should be as dense or denser than the local density of the finite element mesh.
@@
 
-When $x_i$ is near a boundary, it may occur that some gridpoints for the regular grid are in locations where $\varphi_{x_i}$ is undefined.
-In this case, we store a placeholder value for $\varphi_{x_i}$ at those gridpoints.
+When $x_i$ is near a boundary, it may occur that some gridpoints for the regular grid are in locations where $\convkernel_{x_i}$ is undefined.
+In this case, we store a placeholder value for $\convkernel_{x_i}$ at those gridpoints.
 We use the IEEE standard value of ``NaN'' (not a number) for this placeholder.
@@
 
-Evaluating the kernel for our product-convolution approximation, $\AkerPc(y,x)$, requires evaluating $\varphi_{x_i}(y-x)$.
-But $\varphi_{x_i}(y-x)$ is undefined  when $y - x + x_i$ is outside $\Omega$, which may occur even if $x \in \Omega$ and $y \in \Omega$.
-The default approach to deal with this problem is to extend $\varphi_{x_i}$ by zero wherever it is undefined.
+Evaluating the kernel for our product-convolution approximation, $\AkerPc(y,x)$, requires evaluating $\convkernel_{x_i}(y-x)$.
+But $\convkernel_{x_i}(y-x)$ is undefined  when $y - x + x_i$ is outside $\Omega$, which may occur even if $x \in \Omega$ and $y \in \Omega$.
+The default approach to deal with this problem is to extend $\convkernel_{x_i}$ by zero wherever it is undefined.
 But this leads to substantial errors, known as boundary artifacts, which are especially significant when the sample point $x_i$ is near the boundary of the domain.
@@
 
-While $\varphi_{x_i}$ may be undefined at a point $z$, often there is a sample point $x_j$ that is near $x_i$, but further away from the boundary than $x_i$, such that $\varphi_{x_j}$ is well defined at $z$.
+While $\convkernel_{x_i}$ may be undefined at a point $z$, often there is a sample point $x_j$ that is near $x_i$, but further away from the boundary than $x_i$, such that $\convkernel_{x_j}$ is well defined at $z$.
 To address boundary artifacts, we fill in missing/undefined portions of convolution kernels using information from other convolution kernels associated with the $k$ nearest sample points (we use $k=8$ in our numerical results).
@@
 That is, $p_1=x_i$, $p_2$ is the nearest neighbor to $x_i$, $p_3$ is the second nearest neighbor to $x_i$, and so on.
-Also, recall from the previous section (Section \ref{sec:convolution_kernel_regular_grid}) that the functions $\varphi_{x_i}$ are stored on regular rectilinear grids.
-To extend $\varphi_{x_i}$, we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\varphi_{x_i}$, and contains that grid as a subgrid, but is large enough to contain the boxes associated with all of the neighbor kernels, $\varphi_{p_1}, \varphi_{p_2}, \dots, \varphi_{p_k}$ (recall that the functions $\varphi_{x_i}$ are all centered at zero).
-We initialize an extended convolution kernel, $\varphi_{x_i}^E$, to zero on this expanded grid.
-Then we interpolate the functions $\varphi_{p_j}$ onto this expanded grid in reverse order.
-First we interpolate $\varphi_{p_k}$ onto the grid, then $\varphi_{p_{k-1}}$, and so on until the last step in which we interpolate the original kernel $\varphi_{p_1} = \varphi_{x_i}$.
-At each step, we overwrite existing values of $\varphi_{x_i}^E$ with new values from the kernel currently being interpolated, except when the new value of $\varphi_{p_j}$ is undefined (``NaN'').
-At gridpoints where the kernel being interpolated is undefined, we leave $\varphi_{x_i}^E$ unchanged.
-In this way, we fill in undefined values of $\varphi_{x_i}$ with values from successively further and further away kernels.
+Also, recall from the previous section (Section \ref{sec:convolution_kernel_regular_grid}) that the functions $\convkernel_{x_i}$ are stored on regular rectilinear grids.
+To extend $\convkernel_{x_i}$, we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\convkernel_{x_i}$, and contains that grid as a subgrid, but is large enough to contain the boxes associated with all of the neighbor kernels, $\convkernel_{p_1}, \convkernel_{p_2}, \dots, \convkernel_{p_k}$ (recall that the functions $\convkernel_{x_i}$ are all centered at zero).
+We initialize an extended convolution kernel, $\convkernel_{x_i}^E$, to zero on this expanded grid.
+Then we interpolate the functions $\convkernel_{p_j}$ onto this expanded grid in reverse order.
+First we interpolate $\convkernel_{p_k}$ onto the grid, then $\convkernel_{p_{k-1}}$, and so on until the last step in which we interpolate the original kernel $\convkernel_{p_1} = \convkernel_{x_i}$.
+At each step, we overwrite existing values of $\convkernel_{x_i}^E$ with new values from the kernel currently being interpolated, except when the new value of $\convkernel_{p_j}$ is undefined (``NaN'').
+At gridpoints where the kernel being interpolated is undefined, we leave $\convkernel_{x_i}^E$ unchanged.
+In this way, we fill in undefined values of $\convkernel_{x_i}$ with values from successively further and further away kernels.
 This process is illustrated in Figure FIG.
-To avoid notational clutter, and since we will henceforth only work with the extended kernel, $\varphi_{x_i}^E$, in the remainder of this paper the symbol $\varphi_{x_i}$ refers to $\varphi_{x_i}^E$.
+To avoid notational clutter, and since we will henceforth only work with the extended kernel, $\convkernel_{x_i}^E$, in the remainder of this paper the symbol $\convkernel_{x_i}$ refers to $\convkernel_{x_i}^E$.
 
@@
 Here $\Delta$ is the Laplacian on $\Omega$, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$, and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions.
-Using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense (Theorem \ref{thm:smoothest_interpolant}).
+Using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\convkernel_i$, in a least-squares sense (Theorem \ref{thm:smoothest_interpolant}).
 Also, these weighting functions sum to one pointwise (Proposition \ref{thm:wi_sum_to_one}).
@@
 		\label{eq:u_defn_in_thm}
-		\horizinterpolant_y(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \AkerPc\left(y+x,x\right),
+		\horizinterpolant_y(x) := \sum_{i=1}^r \convkernel_i(y)w_i(x) = \AkerPc\left(y+x,x\right),
 	\end{equation}
@@
 			\min_{\horizinterpolant_y \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \horizinterpolant_y}_{L^2(\Omega)}^2 \\
-			\text{such that} &\quad \horizinterpolant_y(x_i) = \varphi_i(y), \quad i=1,\dots,q.
+			\text{such that} &\quad \horizinterpolant_y(x_i) = \convkernel_i(y), \quad i=1,\dots,q.
 		\end{aligned}
@@
 	\label{eq:eval_y_x}
-	\AkerPc(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x).
+	\AkerPc(y,x) = \sum_{i=1}^r w_i(x) \convkernel_{x_i}(y-x).
 \end{equation}
@@
 \begin{proof}[Proof of Theorem \ref{thm:smoothest_interpolant}]
-	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.
+	Let $f = (\convkernel_1(y), \convkernel_2(y), \dots, \convkernel_r(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.
 The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
 	\begin{align*}
-		\mathcal{T}(f) &= \varphi_1(y) \mathcal{T}(e_1) + \varphi_2(y) \mathcal{T}(e_2) + \dots +\varphi_r(y) \mathcal{T}(e_r) \\
-		&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
+		\mathcal{T}(f) &= \convkernel_1(y) \mathcal{T}(e_1) + \convkernel_2(y) \mathcal{T}(e_2) + \dots + \convkernel_r(y) \mathcal{T}(e_r) \\
+		&= \convkernel_1(y) w_1 + \convkernel_2(y) w_2 + \dots + \convkernel_r(y) w_r \\
 		&= \horizinterpolant_y.

### 2021-09-14 paper/localpsf.tex
paper cleanup

@@
 	{	
-		\Input{Operator $\mathcal{A}$}
+		\Input{Operator $\Aop$}
 		\Output{$\alpha$, $\mu$, and $\Sigma$}
@@
 
-		Compute $\alpha = \left(\mathcal{A}^T C\right)^*$
+		Compute $\alpha = \left(\Aop^T C\right)^*$
 
@@
 
-			Compute $\mu^i = \left(\mathcal{A}^T L^i\right)^* / \alpha$
+			Compute $\mu^i = \left(\Aop^T L^i\right)^* / \alpha$
 		}
@@
 
-				Compute $\Sigma^{ij} = \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot \mu^j$
+				Compute $\Sigma^{ij} = \left(\Aop^T Q^{ij}\right)^* / \alpha - \mu^i\cdot \mu^j$
 
@@
 These sample point batches will be used in Section \ref{sec:get_impulse_response} to compute many impulse responses of $\Aop$ using only a small number of operator actions of $\Aop$.
-The support of $\impulseresponse_x$ is (approximately) contained in the ellipsoid $E_x$, so by applying $\mathcal{A}$ to the Dirac comb associated with all sample points in a batch, we will recover the impulse responses for all sample points in that batch.
+The support of $\impulseresponse_x$ is (approximately) contained in the ellipsoid $E_x$, so by applying $\Aop$ to the Dirac comb associated with all sample points in a batch, we will recover the impulse responses for all sample points in that batch.
 
@@
 \end{equation*}
-For each batch $S_b$, we compute the action of $\mathcal{A}$ on the associated Dirac comb:
+For each batch $S_b$, we compute the action of $\Aop$ on the associated Dirac comb:
 \begin{equation}
 	\label{eq:dirac_comb_H_action}
-	\combresponse_b := \left(\mathcal{A} \dcomb_b^*\right)^*.
+	\combresponse_b := \left(\Aop \dcomb_b^*\right)^*.
 \end{equation}
@@
 	\label{eq:phi_b}
-	\combresponse_b = \left\langle\mathcal{A},\sum_{x_i \in S_b} \delta_{x_i}\right\rangle^* = \sum_{x_i \in S_b} \langle\mathcal{A}, \delta_{x_i}\rangle^* = \sum_{x_i \in S_b} \impulseresponse_{x_i}.
+	\combresponse_b = \left(\Aop \left(\sum_{x_i \in S_b} \delta_{x_i}\right)^*\right)^* = \sum_{x_i \in S_b} \left(\Aop \delta_{x_i}^*\right)^* = \sum_{x_i \in S_b} \impulseresponse_{x_i}.
 \end{equation}
@@
 Furthermore, \eqref{eq:desired_property_1} implies that the null space of $\AmatPcSymPlus$ contains the null-space of $\AmatPcSym$.
-This is important when the operator $\mathcal{A}$ has a large (or infinite) dimensional null space, or a spectrum that clusters at zero.
+This is important when the operator $\Aop$ has a large (or infinite) dimensional null space, or a spectrum that clusters at zero.
 Hessians in ill-posed distributed parameter inverse problems typically have both of these properties.
@@
 The idea is that a block of the matrix is likely to be low rank if the point cloud associated with the rows of the block is far away from the point cloud associated with the columns of the block.
-This is reasonable to expect here because of the locality property of $\mathcal{A}$.
+This is reasonable to expect here because of the locality property of $\Aop$.
 Indeed, locality implies that many blocks of the matrix corresponding to far away point cloud clusters will be rank zero.
@@
 	\begin{align*}
-		\left(v, \alpha\right)_{L^2(\Omega)} =& \int_\Omega v(x) \int_{\Omega} \langle\mathcal{A},\delta_x\rangle^*(z) dz dx\\
+		\left(v, \alpha\right)_{L^2(\Omega)} =& \int_\Omega v(x) \int_{\Omega} \left(\Aop \delta_x^*\right)^*(z) dz dx\\
 		=& \int_\Omega v(x) \int_{\Omega} A(z,x) dz dx \\
 		=& \int_\Omega \int_\Omega v(x) A(z,x) C(z) dz dx \\
-		=& \left(\mathcal{A}^T C\right)(v),
+		=& \left(\Aop^T C\right)(v),
 	\end{align*}
-	which implies $\alpha = \left(\mathcal{A}^TC\right)^*$.
+	which implies $\alpha = \left(\Aop^T C\right)^*$.
 
@@
 	\begin{align*}
-		\left(v, \alpha \cdot \mu^i\right)_{L^2(\Omega)} =& \int_\Omega v(x) \alpha(x) \int_{\Omega} z^i \langle\mathcal{A},\delta_x\rangle^*(z)/\alpha(x) dz dx\\
+		\left(v, \alpha \cdot \mu^i\right)_{L^2(\Omega)} =& \int_\Omega v(x) \alpha(x) \int_{\Omega} z^i \left(\Aop \delta_x^*\right)^*(z)/\alpha(x) dz dx\\
 		=& \int_\Omega \int_\Omega v(x) L^i(z) A(z,x) dz dx \\
-		=& \left(\mathcal{A}^T L^i\right)(v),
+		=& \left(\Aop^T L^i\right)(v),
 	\end{align*}
-	which implies $\mu^i = \left(\mathcal{A}^T L^i\right)^* / \alpha$.
+	which implies $\mu^i = \left(\Aop^T L^i\right)^* / \alpha$.
 
@@
 	\begin{align*}
-		\left(v, \alpha \cdot \left(\Sigma^{ij}+\mu^i \cdot \mu^j\right)\right)_{L^2(\Omega)} =& \int_\Omega v(x) \alpha(x) \int_\Omega z^i z^j \langle\mathcal{A},\delta_x\rangle^*(z) / \alpha(x) dz dx \\
+		\left(v, \alpha \cdot \left(\Sigma^{ij}+\mu^i \cdot \mu^j\right)\right)_{L^2(\Omega)} =& \int_\Omega v(x) \alpha(x) \int_\Omega z^i z^j \left(\Aop \delta_x^*\right)^*(z) / \alpha(x) dz dx \\
 		=& \int_\Omega \int_\Omega v(x) Q^{ij}(z) A(z,x) dz dx \\
-		=& \left(\mathcal{A}Q^{ij}\right)(v),
+		=& \left(\Aop Q^{ij}\right)(v),
 	\end{align*}
-	which implies $\Sigma^{ij} = \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot\mu^j$.
+	which implies $\Sigma^{ij} = \left(\Aop^T Q^{ij}\right)^* / \alpha - \mu^i\cdot\mu^j$.
 \end{proof}
@@
 Let $\overline{\Omega}$ be the closure of $\Omega$, and let $C(\overline{\Omega})$ be the space of continuous functions mapping $\overline{\Omega}\rightarrow \mathbb{R}$.
-The action of $\mathcal{A}$ is extended to distributions $\mu:C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ via the formula
-\begin{equation*}
-	\langle\mathcal{A},\mu\rangle(v) := \int_\Omega v(y) \mu\left(A(y, \cdot)\right) dy, 
+The action of $\Aop$ is extended to distributions $\mu:C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ via the formula
+\begin{equation*}
+	\left(\Aop \mu^*\right)(v) := \int_\Omega v(y) \mu\left(A(y, \cdot)\right) dy, 
 \end{equation*}
@@
 \begin{align*}
-	\langle\mathcal{A},\mu\rangle(v) &= \int_\Omega \int_\Omega v(y) A(y,x) \textrm{``} \mu(x) \textrm{''} dx dy \\
+	\left(\Aop \mu^*\right)(v) &= \int_\Omega \int_\Omega v(y) A(y,x) \textrm{``} \mu(x) \textrm{''} dx dy \\
 	&= \int_\Omega v(y) \int_\Omega A(y,x) \textrm{``} \mu(x) \textrm{''} dx \\
@@
 \end{align*}
-For example, the delta distribution $\delta_x$ is defined by $\delta_x(v) = v(x)$, and the action of $\mathcal{A}$ on $\delta_x$ is given by
+For example, the delta distribution $\delta_x$ is defined by $\delta_x(v) = v(x)$, and the action of $\Aop$ on $\delta_x$ is given by
 \begin{equation}
 	\label{eq:appendix_impulse_response}
-	\langle\mathcal{A},\delta_x\rangle(v)	= \int_\Omega v(y) A(y,x) dy.
-\end{equation}
-Recall that the impulse response of $\mathcal{A}$ to $\delta_x$, denoted $\impulseresponse_x$, is the Riesz representation of $\langle\mathcal{A},\delta_x\rangle$.
-From \eqref{eq:appendix_impulse_response} and the definition of the Riesz representation, we have $\impulseresponse_x = \langle\mathcal{A},\delta_x\rangle^* = A(~\cdot~,x)$.
+	\left(\Aop \delta_x^*\right)(v)	= \int_\Omega v(y) A(y,x) dy.
+\end{equation}
+Recall that the impulse response of $\Aop$ to $\delta_x$, denoted $\impulseresponse_x$, is the Riesz representation of $\Aop \delta_x^*$.
+From \eqref{eq:appendix_impulse_response} and the definition of the Riesz representation, we have $\impulseresponse_x = A(~\cdot~,x)$.
 That is, $\impulseresponse_x$ is the function $y \mapsto A(y,x)$.

### 2021-09-14 paper/localpsf.tex
paper cleanup

@@
 		\label{eq:S_matrix}
-		S_{ij} := \int_\Omega \firstgreens_i(x) \firstgreens_j(x) dx.
+		S_{ij} := \left( \firstgreens_i, \firstgreens_j \right)_{L^2(\Omega)}.
 	\end{equation}
@@
 
-Let $f_1, f_2, \dots, f_N$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, and let $V_h := \Span\left(f_1, f_2, \dots, f_N\right)$ be the corresponding finite element space.
-Further, let $p_i \in \mathbb{R}^d$, $i=1,\dots, N$ be the Lagrange nodes associated with the functions $f_i$.
-That is, the points such that $f_i(p_j)$ equals one if $i=j$, and zero otherwise.
+Let $\febasis_1, \febasis_2, \dots, febasis_N$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, and let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_N\right)$ be the corresponding finite element space.
+Further, let $p_i \in \mathbb{R}^d$, $i=1,\dots, N$ be the Lagrange nodes associated with the functions $\febasis_i$.
+That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
 
@@
 	\label{eq:defn_of_Akerpcmesh}
-	\AkerPcMesh(y,x) := \sum_{i=1}^N \sum_{j=1}^N \AkerPc(p_i,p_j) f_i(y) f_j(x).
+	\AkerPcMesh(y,x) := \sum_{i=1}^N \sum_{j=1}^N \AkerPc(p_i,p_j) \febasis_i(y) \febasis_j(x).
 \end{equation}
@@
 That is, $\AopPcMesh : V_h \rightarrow V_h'$.
-Let $u_h \in V_h$ and $v_h \in V_h$, and let $\mathbf{u}$ and $\mathbf{v}$ be the coefficient vectors for $u_h$ and $v_h$, respectively, with respect to the finite element basis $f_1, \dots, f_N$.
+Let $u_h \in V_h$ and $v_h \in V_h$, and let $\mathbf{u}$ and $\mathbf{v}$ be the coefficient vectors for $u_h$ and $v_h$, respectively, with respect to the finite element basis $\febasis_1, \dots, \febasis_N$.
 I.e.,
@@
 	\label{eq:fem_coeff_basis}
-	u_h(x) = \sum_{i=1}^N \mathbf{u}_i f_i(x),
+	u_h(x) = \sum_{i=1}^N \mathbf{u}_i \febasis_i(x),
 \end{equation}
@@
 
-Let $f_i$, $i=1,\dots,N$ be a set of finite element basis functions, let $h$ denote the mesh size parameter for the finite element mesh, and let
-\begin{equation*}
-	V_h = \Span\left(f_1, f_2, \dots, f_N\right)
+Let $\febasis_i$, $i=1,\dots,N$ be a set of finite element basis functions, let $h$ denote the mesh size parameter for the finite element mesh, and let
+\begin{equation*}
+	V_h = \Span\left(\febasis_1, \febasis_2, \dots, \febasis_N\right)
 \end{equation*}
@@
 	\label{eq:finite_bijection}
-	u_h(x) = \sum_{i=1}^N \mathbf{u}_i f_i(x).
+	u_h(x) = \sum_{i=1}^N \mathbf{u}_i \febasis_i(x).
 \end{equation}
@@
 \begin{equation*}
-	\mathbf{M}_{ij} = \left(f_i, f_j\right)_{L^2(\Omega)} = \int_{\Omega} f_i(x) f_j(x) dx,
+	\mathbf{M}_{ij} = \left(\febasis_i, \febasis_j\right)_{L^2(\Omega)} = \int_{\Omega} \febasis_i(x) \febasis_j(x) dx,
 \end{equation*}
@@
 	\begin{equation*}
-		\boldsymbol{\sigma}_i = \sigma_h(f_i), \quad i=1,\dots,N.
+		\boldsymbol{\sigma}_i = \sigma_h(\febasis_i), \quad i=1,\dots,N.
 	\end{equation*}
@@
 	\begin{equation*}
-		\left(\mathbf{g}^\pi\right)^*_i = \int_\Omega g(x) f_i(x) dx, \quad i=1,\dots,N.
+		\left(\mathbf{g}^\pi\right)^*_i = \int_\Omega g(x) \febasis_i(x) dx, \quad i=1,\dots,N.
 	\end{equation*}
@@
 	\end{equation*}
-	\item[Finite element interpolation:] Let $p_i \in \mathbb{R}^d$, $i=1,\dots,N$, be the Lagrange nodes associated with the finite element basis functions $f_i$.
+	\item[Finite element interpolation:] Let $p_i \in \mathbb{R}^d$, $i=1,\dots,N$, be the Lagrange nodes associated with the finite element basis functions $\febasis_i$.
 The finite element interpolant of a function $g \in C\left(\overline{\Omega}\right)$ is the function $g^I \in V_h$ given by
 	\begin{equation*}
-		g_h^I(x) = \sum_{i=1}^N g(p_i) f_i(x).
+		g_h^I(x) = \sum_{i=1}^N g(p_i) \febasis_i(x).
 	\end{equation*}
@@
 	\begin{equation*}
-		f_i(p_j) = \begin{cases}1 & i=j \\
+		\febasis_i(p_j) = \begin{cases}1 & i=j \\
 		0 & \text{otherwise},\end{cases}
@@
 	\begin{equation*}
-		\mathbf{B}_{ij} = \left(\mathcal{B}_h f_j\right)(f_i).
+		\mathbf{B}_{ij} = \left(\mathcal{B}_h \febasis_j\right)(\febasis_i).
 	\end{equation*}
@@
 	\item[Distributions:] Let $\mu:C\left(\overline{\Omega}\right)\rightarrow \mathbb{R}$ be a distribution.
-If $\mu(f_i)$ is well-defined for $i=1,\dots,N$, then the restriction of $\mu$ to domain $V_h$ is a linear functional $\mu_h\in V_h'$.
+If $\mu(\febasis_i)$ is well-defined for $i=1,\dots,N$, then the restriction of $\mu$ to domain $V_h$ is a linear functional $\mu_h\in V_h'$.
 The functional $\mu_h$ has a coefficient dual vector $\boldsymbol{\mu} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ with entries
 	\begin{equation*}
-		\boldsymbol{\mu}_i=\mu(f_i), \quad i=1,\dots,N.
+		\boldsymbol{\mu}_i=\mu(\febasis_i), \quad i=1,\dots,N.
 	\end{equation*} 
 	For example, let $\delta_p$ be the Dirac distribution centered at a point $p\in\Omega$.
-If the finite element basis functions, $f_i$, are continuous at $p$, then restricting the domain of $\delta_p$ to the space $V_h$ yields a linear functional in $V_h'$.
+If the finite element basis functions, $\febasis_i$, are continuous at $p$, then restricting the domain of $\delta_p$ to the space $V_h$ yields a linear functional in $V_h'$.
 The corresponding coefficient dual vector, $\boldsymbol{\delta}_p$, has entries
 	\begin{equation*}
-		\left(\boldsymbol{\delta}_p\right)_i = f_i(p), \quad i=1,\dots,N.
+		\left(\boldsymbol{\delta}_p\right)_i = \febasis_i(p), \quad i=1,\dots,N.
 	\end{equation*}
@@
 \begin{equation*}
-	\boldsymbol{\dcomb}^b_i = \sum_{x_j \in S_b} f_i(x_j), \quad i=1,\dots,N.
+	\boldsymbol{\dcomb}^b_i = \sum_{x_j \in S_b} \febasis_i(x_j), \quad i=1,\dots,N.
 \end{equation*}

### 2021-09-14 paper/localpsf.tex
paper cleanup

@@
 
-		Compute $\alpha$, $\mu$, and $\Sigma$ using Algorithm \ref{alg:varhpi_mean_cov}.
+		Compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ using Algorithm \ref{alg:varhpi_mean_cov}.
 
@@
 Hence, $\impulseresponse_x$ is a scaled probability distribution.
-Let $\alpha$ denote the scaling factor, let $\widehat{\impulseresponse}_x := \impulseresponse_x / \alpha(x)$ denote the normalized version of $\impulseresponse_x$, and let $\mu(x)$ and $\Sigma(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
+Let $\spatialvol$ denote the scaling factor, let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
 We make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
@@
 	\label{eq:support_ellipsoid}
-	E_x := \{x' \in \Omega: (x' - \mu(x))^T \Sigma(x)^{-1} (x' - \mu(x)) \le \tau^2\},
+	E_x := \{x' \in \Omega: (x' - \spatialmean(x))^T \spatialcov(x)^{-1} (x' - \spatialmean(x)) \le \tau^2\},
 \end{equation}
 where $\tau$ is a fixed constant.
-The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\impulseresponse}_x$ (see Figure FIG).
+The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\impulseresponse}_x$ (see Figure FIG).
 The quantity $\tau$ is a parameter that must be chosen by the user of the algorithm.
@@
 The ellipsoid $E_x$ is not immediately available.
-To use $E_x$, we must first compute the scalar $\alpha(x)$, the length-$d$ vector $\mu(x)$, and the $d \times d$ matrix $\Sigma(x)$ (recall $d \in \{1,2,3\}$ is the spatial dimension of $\Omega$).
-The direct approach to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}.
-Then one can post-process $\impulseresponse_x$ to determine $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$.
+To use $E_x$, we must first compute the scalar $\spatialvol(x)$, the length-$d$ vector $\spatialmean(x)$, and the $d \times d$ matrix $\spatialcov(x)$ (recall $d \in \{1,2,3\}$ is the spatial dimension of $\Omega$).
+The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}.
+Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
 But this direct approach is computationally infeasible because and it requires one operator action of $\Aop$ per point $x$ that we wish to evaluate $E_x$ at, and we will need to evaluate $E_x$ at a large number of points.
-Fortunately, it is possible to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$, \emph{for all points $x$ simultaneously}, by applying $\Aop^T$ to one constant function, $d$ linear functions, and $d(d+1)/2$ quadratic functions.
-We present this efficient method of computing $\alpha$, $\mu$, and $\Sigma$ in Theorem \ref{thm:vol_mean_cov} and summarize the method in Algorithm \ref{alg:varhpi_mean_cov}.
+Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x$ simultaneously}, by applying $\Aop^T$ to one constant function, $d$ linear functions, and $d(d+1)/2$ quadratic functions.
+We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Theorem \ref{thm:vol_mean_cov} and summarize the method in Algorithm \ref{alg:varhpi_mean_cov}.
 
@@
 	\label{thm:vol_mean_cov}
-	Let $\alpha:\Omega \rightarrow \mathbb{R}$, $\mu:\Omega \rightarrow \mathbb{R}^d$, and $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$, be the following functions:
-	\begin{equation*}
-		\alpha(x) := \int_{\Omega} \impulseresponse_x(y) dy, \qquad
-		\mu(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
-		\Sigma(x) := \Var(\widehat{\impulseresponse}_x),
+	Let $\spatialvol:\Omega \rightarrow \mathbb{R}$, $\spatialmean:\Omega \rightarrow \mathbb{R}^d$, and $\spatialcov:\Omega \rightarrow \mathbb{R}^{d \times d}$, be the following functions:
+	\begin{equation*}
+		\spatialvol(x) := \int_{\Omega} \impulseresponse_x(y) dy, \qquad
+		\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
+		\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
 	\end{equation*}
@@
 		\begin{align}
-			\alpha =& \left(\Aop^T C\right)^* \\
-			\mu^i =& \left(\Aop^T L^i\right)^* / \alpha \\
-			\Sigma^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \alpha - \mu^i\cdot \mu^j
+			\spatialvol =& \left(\Aop^T C\right)^* \\
+			\spatialmean^i =& \left(\Aop^T L^i\right)^* / \spatialvol \\
+			\spatialcov^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j
 		\end{align}
@@
 		\Input{Operator $\Aop$}
-		\Output{$\alpha$, $\mu$, and $\Sigma$}
-
-		\tcp{Compute scaling factor $\alpha$}
+		\Output{$\spatialvol$, $\spatialmean$, and $\spatialcov$}
+
+		\tcp{Compute scaling factor $\spatialvol$}
 
@@
 
-		Compute $\alpha = \left(\Aop^T C\right)^*$
-
-		\tcp{Compute mean $\mu$}
+		Compute $\spatialvol = \left(\Aop^T C\right)^*$
+
+		\tcp{Compute mean $\spatialmean$}
 		\For{$i=1,2,\dots,d$}{
@@
 
-			Compute $\mu^i = \left(\Aop^T L^i\right)^* / \alpha$
-		}
-		\tcp{Compute covariance $\Sigma$}
+			Compute $\spatialmean^i = \left(\Aop^T L^i\right)^* / \spatialvol$
+		}
+		\tcp{Compute covariance $\spatialcov$}
 		\For{$i=1,2,\dots,d$}{
@@
 
-				Compute $\Sigma^{ij} = \left(\Aop^T Q^{ij}\right)^* / \alpha - \mu^i\cdot \mu^j$
-
-				Set $\Sigma^{ji} = \Sigma^{ij}$
-
-			}
-		}
-
-	}
-	\caption{Compute scaling factor $\alpha$, mean $\mu$, and covariance $\Sigma$}
+				Compute $\spatialcov^{ij} = \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j$
+
+				Set $\spatialcov^{ji} = \spatialcov^{ij}$
+
+			}
+		}
+
+	}
+	\caption{Compute scaling factor $\spatialvol$, mean $\spatialmean$, and covariance $\spatialcov$}
 	\label{alg:varhpi_mean_cov}
@@
 
-	\Input{Finite set of candidate points $P\subset \Omega$, \\mean $\mu(x)$ and covariance $\Sigma(x)$, \\domain boundary information $\partial \Omega$, \\previous sample point batches $S_1, \dots, S_{b-1}$}
+	\Input{Finite set of candidate points $P\subset \Omega$, \\spatialmean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\domain boundary information $\partial \Omega$, \\previous sample point batches $S_1, \dots, S_{b-1}$}
 	\Output{Batch of new sample points $S_b$}
@@
 
-			Remove all points $p'$ satisfying $\mu(p') \in E_p$ from $P$
+			Remove all points $p'$ satisfying $\spatialmean(p') \in E_p$ from $P$
 		}
@@
 	\label{eq:rational_matrix_function}
-	\AmatPcSymPlus := c_0 \mathbf{I} + c_1 \AmatPcSym + c_2 \left(\AmatPcSym + \mu \mathbf{I}\right)^{-1},
+	\AmatPcSymPlus := c_0 \mathbf{I} + c_1 \AmatPcSym + c_2 \left(\AmatPcSym + \ratpole \mathbf{I}\right)^{-1},
 \end{equation}
@@
 	\label{eq:rational_function_scalar}
-	\ratfct(\lambda) := c_0 + c_1 \lambda + \frac{c_2}{\lambda+\mu}.
-\end{equation}
-The scalars $c_0$, $c_1$, $c_2$, and $\mu$ are chosen so that the moderate and large positive eigenvalues of $\AmatPcSymPlus$ approximately equal the corresponding moderate and large positive eigenvalues of $\AmatPcSym$, while the negative and small positive eigenvalues of $\AmatPcSym$ are modified so that $\AmatPcSymPlus$ is positive semi-definite.
-Later in this section we will explain how $c_0$, $c_1$, $c_2$, and $\mu$ are chosen and how our choice of these scalars affects the spectral properties of $\AmatPcSymPlus$.
+	\ratfct(\lambda) := c_0 + c_1 \lambda + \frac{c_2}{\lambda+\ratpole}.
+\end{equation}
+The scalars $c_0$, $c_1$, $c_2$, and $\ratpole$ are chosen so that the moderate and large positive eigenvalues of $\AmatPcSymPlus$ approximately equal the corresponding moderate and large positive eigenvalues of $\AmatPcSym$, while the negative and small positive eigenvalues of $\AmatPcSym$ are modified so that $\AmatPcSymPlus$ is positive semi-definite.
+Later in this section we will explain how $c_0$, $c_1$, $c_2$, and $\ratpole$ are chosen and how our choice of these scalars affects the spectral properties of $\AmatPcSymPlus$.
 In Figure FIG we illustrate $\ratfct(\lambda)$.
-Forming $\AmatPcSymPlus$ requires computing an H-matrix inverse, $\left(\AmatPcSym + \mu \mathbf{I}\right)^{-1}$.
-While explicit computation of matrix inverses should usually be avoided, here it is computationally acceptable because $\AmatPcSym + \mu \mathbf{I}$ is an H-matrix.
+Forming $\AmatPcSymPlus$ requires computing an H-matrix inverse, $\left(\AmatPcSym + \ratpole \mathbf{I}\right)^{-1}$.
+While explicit computation of matrix inverses should usually be avoided, here it is computationally acceptable because $\AmatPcSym + \ratpole \mathbf{I}$ is an H-matrix.
 Inversion of H-matrices is a routine procedure which only requires $O(N \log(N)^2)$ work \cite{Hmatrixinverse}.
 
-We choose $c_0, c_1, c_2$, and $\mu$, to be the solution to the following system of equations:
+We choose $c_0, c_1, c_2$, and $\ratpole$, to be the solution to the following system of equations:
 \begin{subequations}
@@
 where $\lambda_\text{min}$ and $\lambda_\text{max}$ are the smallest and largest eigenvalues of $\AmatPcSym$, respectively.
-We also require $\mu > |\lambda_\text{min}|$, so that the pole of $\ratfct$ is not in $[\lambda_\text{min},\lambda_\text{max}]$.
-Note that $\lambda_\text{min} < 0$ since $\AmatPcSym$ is indefinite.
 We estimate $\lambda_\text{min}$ and $\lambda_\text{max}$ using the implicitly restarted Lanczos method \cite{lanczos,scipy}.
+We assume that $\lambda_\text{min} < 0$; if $\lambda_\text{min} \ge 0$, then $\AmatPcSym$ is already positive semi-definite, so we do not need to modify it with a rational function.
+We also require $\ratpole > |\lambda_\text{min}|$, so that the pole of $\ratfct$ is not in $[\lambda_\text{min},\lambda_\text{max}]$.
 We describe how we solve system \eqref{eq:four_eq_system} in Appendix \ref{app:solve_rational_system}.
@@
 
-Together, equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, \eqref{eq:desired_property_3},  and \eqref{eq:desired_property_4}, and the condition $\mu > |\lambda_\text{min}|$, imply that $\ratfct(0)=0$ is the unique minimum of $\ratfct$ on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that $\ratfct$ is non-negative on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that  $\AmatPcSymPlus$ is positive semi-definite.
+Together, equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, \eqref{eq:desired_property_3},  and \eqref{eq:desired_property_4}, and the condition $\ratpole > |\lambda_\text{min}|$, imply that $\ratfct(0)=0$ is the unique minimum of $\ratfct$ on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that $\ratfct$ is non-negative on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that  $\AmatPcSymPlus$ is positive semi-definite.
 Furthermore, \eqref{eq:desired_property_1} implies that the null space of $\AmatPcSymPlus$ contains the null-space of $\AmatPcSym$.
@@
 \begin{align*}
-	E_p :=& \{x : (x - \mu_p)^T \Sigma_p^{-1} (x - \mu_p) \le \tau^2\} \\
-	E_q :=& \{x : (x - \mu_q)^T \Sigma_q^{-1} (x - \mu_q) \le \tau^2\}, \\
+	E_p :=& \{x : (x - \spatialmean_p)^T \spatialcov_p^{-1} (x - \spatialmean_p) \le \tau^2\} \\
+	E_q :=& \{x : (x - \spatialmean_q)^T \spatialcov_q^{-1} (x - \spatialmean_q) \le \tau^2\}, \\
 \end{align*}
-where $\mu_p, \mu_q \in \mathbb{R}^d$, and $\Sigma_p, \Sigma_q \in \mathbb{R}^{d \times d}$ are positive definite. Let $K$ be the following one dimensional convex function:
-\begin{equation*}
-	K(s) := 1 - \frac{1}{\tau^2} (\mu_p - \mu_q)^T \left(\frac{1}{1-s}\Sigma_p + \frac{1}{s}\Sigma_q\right)^{-1}(\mu_p - \mu_q)	
+where $\spatialmean_p, \spatialmean_q \in \mathbb{R}^d$, and $\spatialcov_p, \spatialcov_q \in \mathbb{R}^{d \times d}$ are positive definite. Let $K$ be the following one dimensional convex function:
+\begin{equation*}
+	K(s) := 1 - \frac{1}{\tau^2} (\spatialmean_p - \spatialmean_q)^T \left(\frac{1}{1-s}\spatialcov_p + \frac{1}{s}\spatialcov_q\right)^{-1}(\spatialmean_p - \spatialmean_q)	
 \end{equation*}
@@
 \begin{equation*}
-	\Sigma_p P = \Sigma_q P \Lambda,
+	\spatialcov_p P = \spatialcov_q P \Lambda,
 \end{equation*}
 where $P \in \mathbb{R}^{d \times d}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_d)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$.
-The matrix $P$ simultaneously diagonalizes $\Sigma_p$ and $\Sigma_q$, in the sense that $P^T\Sigma_p P = \Lambda$, and $P^T\Sigma_q P = I$, where $I$ is the $d \times d$ identity matrix.
+The matrix $P$ simultaneously diagonalizes $\spatialcov_p$ and $\spatialcov_q$, in the sense that $P^T\spatialcov_p P = \Lambda$, and $P^T\spatialcov_q P = I$, where $I$ is the $d \times d$ identity matrix.
 Using this diagonalization, and some algebraic manipulations, we may write $K(s)$ as
@@
 \end{equation}
-where $v := \Phi^T\left(\mu_p - \mu_q\right)$.
-We compute the generalized eigenvalue decomposition of $\Sigma_p$ and $\Sigma_q$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's algorithm \cite{BRENT} (any fast 1 dimensional convex optimization routine may be used).
+where $v := \Phi^T\left(\spatialmean_p - \spatialmean_q\right)$.
+We compute the generalized eigenvalue decomposition of $\spatialcov_p$ and $\spatialcov_q$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's algorithm \cite{BRENT} (any fast 1 dimensional convex optimization routine may be used).
 The resulting algorithm for checking whether $E_p$ and $E_q$ intersect is summarized in Algorithm \ref{alg:ellipsoid_intersection_test}.
@@
 	\SetKwProg{Fn}{Function}{}{}
-	\Input{Ellipsoid $E_p$ with mean $\mu_p$ and covariance $\Sigma_p/\tau^2$\\
-		Ellipsoid $E_q$ with mean $\mu_q$ and covariance $\Sigma_q/\tau^2$ 
+	\Input{Ellipsoid $E_p$ with mean $\spatialmean_p$ and covariance $\spatialcov_p/\tau^2$\\
+		Ellipsoid $E_q$ with mean $\spatialmean_q$ and covariance $\spatialcov_q/\tau^2$ 
 	}
@@
 
-	Solve generalized eigenvalue problem $\Sigma_p \Phi = \Sigma_q \Phi \Lambda$
-
-	$v \gets \Phi^T\left(\mu_p - \mu_q\right)$
+	Solve generalized eigenvalue problem $\spatialcov_p \Phi = \spatialcov_q \Phi \Lambda$
+
+	$v \gets \Phi^T\left(\spatialmean_p - \spatialmean_q\right)$
 
@@
 	\begin{bmatrix} 
-		1 & 0 & 1/\mu \\ 0 & 1 & -1/\mu^2 \\ 1 & \lambda_\text{max} & 1/(\lambda_\text{max}+\mu) 
+		1 & 0 & 1/\ratpole \\ 0 & 1 & -1/\ratpole^2 \\ 1 & \lambda_\text{max} & 1/(\lambda_\text{max}+\ratpole) 
 	\end{bmatrix}
@@
 	\label{eq:rational_extra_eq}
-	c_0 + c_1 \lambda_\text{min} + \frac{c_2}{\lambda_\text{min}+\mu} - |\lambda_\text{min}| = 0.
+	c_0 + c_1 \lambda_\text{min} + \frac{c_2}{\lambda_\text{min}+\ratpole} - |\lambda_\text{min}| = 0.
 \end{equation}
@@
 \begin{equation*}
-	g(\mu) := c_0(\mu) + c_1(\mu) \lambda_\text{min} + \frac{c_2(\mu)}{\lambda_\text{min}+\mu} - |\lambda_\text{min}|
-\end{equation*}
-denote the left hand side of \eqref{eq:rational_extra_eq}, but with $c_0(\mu)$, $c_1(\mu)$, and $c_2(\mu)$ being the solution of \eqref{eq:rational_3_linear_system} for the given value of $\mu$.
-I.e., within $g(\mu)$ the quantities $c_0$, $c_1$, and $c_2$ are no longer free variables; they now depend on $\mu$ implicitly through the solution of a linear system.
+	g(\ratpole) := c_0(\ratpole) + c_1(\ratpole) \lambda_\text{min} + \frac{c_2(\ratpole)}{\lambda_\text{min}+\ratpole} - |\lambda_\text{min}|
+\end{equation*}
+denote the left hand side of \eqref{eq:rational_extra_eq}, but with $c_0(\ratpole)$, $c_1(\ratpole)$, and $c_2(\ratpole)$ being the solution of \eqref{eq:rational_3_linear_system} for the given value of $\ratpole$.
+I.e., within $g(\ratpole)$ the quantities $c_0$, $c_1$, and $c_2$ are no longer free variables; they now depend on $\ratpole$ implicitly through the solution of a linear system.
 We solve the combined system, \eqref{eq:rational_3_linear_system} and $\eqref{eq:rational_extra_eq}$, by solving the one-dimensional nonlinear root finding problem
@@
 	\label{eq:g_mu_zero}
-	g(\mu)=0
-\end{equation}
-with bound $\mu > |\lambda_\text{min}|$ and initial guess $\mu_0 = 2 |\lambda_\text{min}|$.
-Each evaluation of $g$ requires solving the $3 \times 3$ linear system \eqref{eq:rational_3_linear_system}, and once the solution $\mu$ is found, we solve \eqref{eq:rational_3_linear_system} one more time to get the final values of $c_0$, $c_1$, and $c_2$.
+	g(\ratpole)=0
+\end{equation}
+with bound $\ratpole > |\lambda_\text{min}|$ and initial guess $\ratpole_0 = 2 |\lambda_\text{min}|$.
+Each evaluation of $g$ requires solving the $3 \times 3$ linear system \eqref{eq:rational_3_linear_system}, and once the solution $\ratpole$ is found, we solve \eqref{eq:rational_3_linear_system} one more time to get the final values of $c_0$, $c_1$, and $c_2$.
 
@@
 	\label{defn:alpha_rho_mu_sigma}
-	We define the spatially varying scaling factor, $\alpha$, the scaled impulse response, $\widehat{\impulseresponse}_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
+	We define the spatially varying scaling factor, $\spatialvol$, the scaled impulse response, $\widehat{\impulseresponse}_x$, the spatially varying mean, $\spatialmean$, and the spatially varying covariance, $\spatialcov$, as follows:
 	\begin{align*}
-		\alpha:\Omega \rightarrow \mathbb{R}, &\qquad \alpha(x) := \int_{\Omega} \impulseresponse_x(z) dz, \\
-		\widehat{\impulseresponse}_x:\Omega \rightarrow \mathbb{R}, &\qquad \widehat{\impulseresponse}_x := \impulseresponse_x \big/ \alpha(x), \\
-		\mu:\Omega \rightarrow \mathbb{R}^d, &\qquad \mu(x) := \int_\Omega z \widehat{\impulseresponse}_x(z) dz, \\
-		\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\Sigma(x) := \int_\Omega (z - \mu(x))(z - \mu(x))^T \widehat{\impulseresponse}_x(z) dz.
+		\spatialvol:\Omega \rightarrow \mathbb{R}, &\qquad \spatialvol(x) := \int_{\Omega} \impulseresponse_x(z) dz, \\
+		\widehat{\impulseresponse}_x:\Omega \rightarrow \mathbb{R}, &\qquad \widehat{\impulseresponse}_x := \impulseresponse_x \big/ \spatialvol(x), \\
+		\spatialmean:\Omega \rightarrow \mathbb{R}^d, &\qquad \spatialmean(x) := \int_\Omega z \widehat{\impulseresponse}_x(z) dz, \\
+		\spatialcov:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\spatialcov(x) := \int_\Omega (z - \spatialmean(x))(z - \spatialmean(x))^T \widehat{\impulseresponse}_x(z) dz.
 	\end{align*}
@@
 	\begin{align*}
-		\left(v, \alpha\right)_{L^2(\Omega)} =& \int_\Omega v(x) \int_{\Omega} \left(\Aop \delta_x^*\right)^*(z) dz dx\\
+		\left(v, \spatialvol\right)_{L^2(\Omega)} =& \int_\Omega v(x) \int_{\Omega} \left(\Aop \delta_x^*\right)^*(z) dz dx\\
 		=& \int_\Omega v(x) \int_{\Omega} A(z,x) dz dx \\
@@
 	\end{align*}
-	which implies $\alpha = \left(\Aop^T C\right)^*$.
+	which implies $\spatialvol = \left(\Aop^T C\right)^*$.
 
@@
 	\begin{align*}
-		\left(v, \alpha \cdot \mu^i\right)_{L^2(\Omega)} =& \int_\Omega v(x) \alpha(x) \int_{\Omega} z^i \left(\Aop \delta_x^*\right)^*(z)/\alpha(x) dz dx\\
+		\left(v, \spatialvol \cdot \spatialmean^i\right)_{L^2(\Omega)} =& \int_\Omega v(x) \spatialvol(x) \int_{\Omega} z^i \left(\Aop \delta_x^*\right)^*(z)/\spatialvol(x) dz dx\\
 		=& \int_\Omega \int_\Omega v(x) L^i(z) A(z,x) dz dx \\
@@
 	\end{align*}
-	which implies $\mu^i = \left(\Aop^T L^i\right)^* / \alpha$.
+	which implies $\spatialmean^i = \left(\Aop^T L^i\right)^* / \spatialvol$.
 
@@
 	\begin{equation*}
-		\Sigma(x) = \int_\Omega zz^T \widehat{\impulseresponse}_x(z) dz - \mu(x) \mu(x)^T,
+		\spatialcov(x) = \int_\Omega zz^T \widehat{\impulseresponse}_x(z) dz - \spatialmean(x) \spatialmean(x)^T,
 	\end{equation*}
@@
 	\begin{align*}
-		\left(v, \alpha \cdot \left(\Sigma^{ij}+\mu^i \cdot \mu^j\right)\right)_{L^2(\Omega)} =& \int_\Omega v(x) \alpha(x) \int_\Omega z^i z^j \left(\Aop \delta_x^*\right)^*(z) / \alpha(x) dz dx \\
+		\left(v, \spatialvol \cdot \left(\spatialcov^{ij}+\spatialmean^i \cdot \spatialmean^j\right)\right)_{L^2(\Omega)} =& \int_\Omega v(x) \spatialvol(x) \int_\Omega z^i z^j \left(\Aop \delta_x^*\right)^*(z) / \spatialvol(x) dz dx \\
 		=& \int_\Omega \int_\Omega v(x) Q^{ij}(z) A(z,x) dz dx \\
@@
 	\end{align*}
-	which implies $\Sigma^{ij} = \left(\Aop^T Q^{ij}\right)^* / \alpha - \mu^i\cdot\mu^j$.
+	which implies $\spatialcov^{ij} = \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot\spatialmean^j$.
 \end{proof}
@@
 Let $\overline{\Omega}$ be the closure of $\Omega$, and let $C(\overline{\Omega})$ be the space of continuous functions mapping $\overline{\Omega}\rightarrow \mathbb{R}$.
-The action of $\Aop$ is extended to distributions $\mu:C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ via the formula
-\begin{equation*}
-	\left(\Aop \mu^*\right)(v) := \int_\Omega v(y) \mu\left(A(y, \cdot)\right) dy, 
+The action of $\Aop$ is extended to distributions $\genericdistribution:C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ via the formula
+\begin{equation*}
+	\left(\Aop \genericdistribution^*\right)(v) := \int_\Omega v(y) \genericdistribution\left(A(y, \cdot)\right) dy, 
 \end{equation*}
@@
 \begin{align*}
-	\left(\Aop \mu^*\right)(v) &= \int_\Omega \int_\Omega v(y) A(y,x) \textrm{``} \mu(x) \textrm{''} dx dy \\
-	&= \int_\Omega v(y) \int_\Omega A(y,x) \textrm{``} \mu(x) \textrm{''} dx \\
-	&= \int_\Omega v(y) \mu\left(A(y,\cdot)\right) dx.
+	\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) A(y,x) \textrm{``} \genericdistribution(x) \textrm{''} dx dy \\
+	&= \int_\Omega v(y) \int_\Omega A(y,x) \textrm{``} \genericdistribution(x) \textrm{''} dx \\
+	&= \int_\Omega v(y) \genericdistribution\left(A(y,\cdot)\right) dx.
 \end{align*}
@@
 
-	\item[Distributions:] Let $\mu:C\left(\overline{\Omega}\right)\rightarrow \mathbb{R}$ be a distribution.
-If $\mu(\febasis_i)$ is well-defined for $i=1,\dots,N$, then the restriction of $\mu$ to domain $V_h$ is a linear functional $\mu_h\in V_h'$.
-The functional $\mu_h$ has a coefficient dual vector $\boldsymbol{\mu} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ with entries
-	\begin{equation*}
-		\boldsymbol{\mu}_i=\mu(\febasis_i), \quad i=1,\dots,N.
+	\item[Distributions:] Let $\genericdistribution:C\left(\overline{\Omega}\right)\rightarrow \mathbb{R}$ be a distribution.
+If $\genericdistribution(\febasis_i)$ is well-defined for $i=1,\dots,N$, then the restriction of $\genericdistribution$ to domain $V_h$ is a linear functional $\genericdistribution_h\in V_h'$.
+The functional $\genericdistribution_h$ has a coefficient dual vector $\boldsymbol{\genericdistribution} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ with entries
+	\begin{equation*}
+		\boldsymbol{\genericdistribution}_i=\genericdistribution(\febasis_i), \quad i=1,\dots,N.
 	\end{equation*} 
@@
 
-	\item[Action of an operator on a distribution:] Since $V_h$ is finite-dimensional, the restriction, $\mu_h \in V_h$, of a distribution, $\mu$, to the space $V_h$ has a a Riesz representation $\mu_h^* \in V_h$.
+	\item[Action of an operator on a distribution:] Since $V_h$ is finite-dimensional, the restriction, $\genericdistribution_h \in V_h$, of a distribution, $\genericdistribution$, to the space $V_h$ has a a Riesz representation $\genericdistribution_h^* \in V_h$.
 Let $B_h$ be the integral kernel associated with an operator $\mathcal{B}_h:V_h \rightarrow V_h'$.
@@
 	\begin{align*}
-		\langle \mathcal{B}_h, \mu_h \rangle(v) &= \int_\Omega v_h(y) \mu_h\left(B_h(y, \cdot)\right) dy \\
-		&= \int_\Omega v_h(y) \left(\int_\Omega B_h(y,x) \mu_h^*(x) dx\right) dy \\
-		&= \int_\Omega \int_\Omega v_h(y) B_h(y,x) \mu_h^*(x) dx dy = \left(\mathcal{B}_h \mu_h^*\right)(v).
+		\langle \mathcal{B}_h, \genericdistribution_h \rangle(v) &= \int_\Omega v_h(y) \genericdistribution_h\left(B_h(y, \cdot)\right) dy \\
+		&= \int_\Omega v_h(y) \left(\int_\Omega B_h(y,x) \genericdistribution_h^*(x) dx\right) dy \\
+		&= \int_\Omega \int_\Omega v_h(y) B_h(y,x) \genericdistribution_h^*(x) dx dy = \left(\mathcal{B}_h \genericdistribution_h^*\right)(v).
 	\end{align*}
 	Going from the first to the second line we used the definition of the Riesz representation.
-The action of $\mathcal{B}_h$ on $\mu_h$ in the sense of distributions is therefore equal to the action of $\mathcal{B}_h$ on $\mu^*_h$ in the conventional sense.
+The action of $\mathcal{B}_h$ on $\genericdistribution_h$ in the sense of distributions is therefore equal to the action of $\mathcal{B}_h$ on $\genericdistribution^*_h$ in the conventional sense.
 We have
 	\begin{equation*}
-		\sigma_h = \langle \mathcal{B}_h, \mu_h \rangle \quad \Leftrightarrow \quad \sigma_h = \mathcal{B}_h \mu_h^* \quad \Leftrightarrow \quad \boldsymbol{\sigma} = \mathbf{B} \mathbf{M}^{-1} \boldsymbol{\mu}.
+		\sigma_h = \langle \mathcal{B}_h, \genericdistribution_h \rangle \quad \Leftrightarrow \quad \sigma_h = \mathcal{B}_h \genericdistribution_h^* \quad \Leftrightarrow \quad \boldsymbol{\sigma} = \mathbf{B} \mathbf{M}^{-1} \boldsymbol{\genericdistribution}.
 	\end{equation*}
@@
 		\Input{Operator $\mathbf{A}$}
-		\Output{$\boldsymbol{\alpha}\in \mathbb{R}^N$, $\boldsymbol{\mu}\in \mathbb{R}^{d \times N}$, and $\boldsymbol{\Sigma} \in \mathbb{R}^{d \times d \times N}$}
+		\Output{$\boldsymbol{\spatialvol}\in \mathbb{R}^N$, $\boldsymbol{\spatialmean}\in \mathbb{R}^{d \times N}$, and $\boldsymbol{\spatialcov} \in \mathbb{R}^{d \times d \times N}$}
 
@@
 
-		$\boldsymbol{\alpha} = \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{C}\right)^*$
+		$\boldsymbol{\spatialvol} = \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{C}\right)^*$
 
@@
 
-			$\mathbf{\mu}^i = \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{L}^i\right)^* / \boldsymbol{\alpha}$
+			$\mathbf{\spatialmean}^i = \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{L}^i\right)^* / \boldsymbol{\spatialvol}$
 		}
@@
 
-				$\boldsymbol{\Sigma}^{ij} = \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{Q}^{ij}\right)^* / \boldsymbol{\alpha} - \boldsymbol{\mu}^i\cdot \boldsymbol{\mu}^j$
-
-				$\boldsymbol{\Sigma}^{ij} = \boldsymbol{\Sigma}^{ji}$
+				$\boldsymbol{\spatialcov}^{ij} = \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{Q}^{ij}\right)^* / \boldsymbol{\spatialvol} - \boldsymbol{\spatialmean}^i\cdot \boldsymbol{\spatialmean}^j$
+
+				$\boldsymbol{\spatialcov}^{ij} = \boldsymbol{\spatialcov}^{ji}$
 
@@
 
-In computations, $\alpha$, $\mu^{i}$, and $\Sigma^{ij}$ are replaced with finite element approximations $\alpha_h, \mu_h^i, \Sigma_h^{ij}\in V_h$, respectively, which have coefficient vectors $\boldsymbol{\alpha}, \boldsymbol{\mu}^{ij}, \boldsymbol{\Sigma}^{ij}\in\mathbb{R}^N_\mathbf{M}$, respectively.
+In computations, $\spatialvol$, $\spatialmean^{i}$, and $\spatialcov^{ij}$ are replaced with finite element approximations $\spatialvol_h, \spatialmean_h^i, \spatialcov_h^{ij}\in V_h$, respectively, which have coefficient vectors $\boldsymbol{\spatialvol}, \boldsymbol{\spatialmean}^{ij}, \boldsymbol{\spatialcov}^{ij}\in\mathbb{R}^N_\mathbf{M}$, respectively.
 The functions $C$, $L^i$, and $Q^{ij}$ are replaced with either finite element projections onto $V_h$ (if the finite element basis does not satisfy the Kronecker property), or finite element interpolations onto $V_h$ (if the finite element basis satisfies the Kronecker property).
@@
 \begin{align*}
-	\boldsymbol{\alpha} &= \mathbf{M}^{-1}\mathbf{A}^T\mathbf{C} \\
-	\boldsymbol{\mu}^i &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{L}^i\right) / \boldsymbol{\alpha}\\
-	\boldsymbol{\Sigma}^{ij} &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{Q}^{ij}\right) / \boldsymbol{\alpha} - \boldsymbol{\mu}^i\cdot \boldsymbol{\mu}^j,
+	\boldsymbol{\spatialvol} &= \mathbf{M}^{-1}\mathbf{A}^T\mathbf{C} \\
+	\boldsymbol{\spatialmean}^i &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{L}^i\right) / \boldsymbol{\spatialvol}\\
+	\boldsymbol{\spatialcov}^{ij} &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{Q}^{ij}\right) / \boldsymbol{\spatialvol} - \boldsymbol{\spatialmean}^i\cdot \boldsymbol{\spatialmean}^j,
 \end{align*}

### 2021-09-14 paper/localpsf.tex
paper cleanup

@@
 \end{equation}
-where $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying the operator $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\psi \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\psi^* \in L^2(\Omega)$ such that $\psi(v) = \left(\psi^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
+where $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying the operator $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\sigma \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\sigma^* \in L^2(\Omega)$ such that $\sigma(v) = \left(\sigma^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
 This is described in Appendix \ref{app:distributions}.
@@
 		\For{$b=1,2,\dots,\text{max\_batches}$}{
-			Choose a batch of sample points, $S_b$, using Algorithm \ref{alg:point_choice}
-
-			Compute $\combresponse_b$ by applying $\Aop$ to the dirac comb for $S_b$ (Section \ref{sec:get_impulse_response})
+			Choose a batch of sample points, $\pointbatch_b$, using Algorithm \ref{alg:point_choice}
+
+			Compute $\combresponse_b$ by applying $\Aop$ to the dirac comb for $\pointbatch_b$ (Section \ref{sec:get_impulse_response})
 
@@
 	\end{equation*}
-	where $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$.
+	where $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denotes the normalized version of $\impulseresponse_x$, $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$.
 Let $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions:
@@
 	\end{equation*}
-	Let $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
+	Let $f/g$ denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
 The following results hold:
@@
 
-We use a greedy ellipsoid packing algorithm to choose sets of sample points, $S_b$, such that there is no overlap between the ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch.
+We use a greedy ellipsoid packing algorithm to choose sets of sample points, $\pointbatch_b$, such that there is no overlap between the ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch.
 These sample point batches will be used in Section \ref{sec:get_impulse_response} to compute many impulse responses of $\Aop$ using only a small number of operator actions of $\Aop$.
@@
 We start with a finite set of candidate points $P$.
-To build a batch of sample points, $S_b$, first we initialize $S_b$ as an empty set.
+To build a batch of sample points, $\pointbatch_b$, first we initialize $\pointbatch_b$ as an empty set.
 Then we select the point $p \in P$ that is the farthest away from all points in previous sample point batches.
@@
 \begin{equation*}
-\max_{p \in P} \min_{x \in S_1 \cup \dots \cup S_{b-1}} \|p - x\|.
+\max_{p \in P} \min_{x \in \pointbatch_1 \cup \dots \cup \pointbatch_{b-1}} \|p - x\|.
 \end{equation*}
@@
 Once $p$ is selected, we remove $p$ from $P$.
-If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in S_b$, then we add $p$ to $S_b$.
+If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_b$, then we add $p$ to $\pointbatch_b$.
 Otherwise we discard $p$.
@@
 
-We repeat the process to construct several batches of points $S_1, S_2, \dots$, until the number of batches exceeds a desired threshold.
+We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots$, until the number of batches exceeds a desired threshold.
 In our implementation, for each batch the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, minus all points in previously chosen batches.
@@
 
-	\Input{Finite set of candidate points $P\subset \Omega$, \\spatialmean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\domain boundary information $\partial \Omega$, \\previous sample point batches $S_1, \dots, S_{b-1}$}
-	\Output{Batch of new sample points $S_b$}
-
-	Initialize empty new batch of sample points, $S_b = \{\}$
+	\Input{Finite set of candidate points $P\subset \Omega$, \\spatialmean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\domain boundary information $\partial \Omega$, \\previous sample point batches $\pointbatch_1, \dots, \pointbatch_{b-1}$}
+	\Output{Batch of new sample points $\pointbatch_b$}
+
+	Initialize empty new batch of sample points, $\pointbatch_b = \{\}$
 
 	\While{$P$ is not empty}{
-		Determine the point $p \in P$ that is farthest from all points in the previous sample point batches $S_1,\dots,S_{b-1}$
+		Determine the point $p \in P$ that is farthest from all points in the previous sample point batches $\pointbatch_1,\dots,\pointbatch_{b-1}$
 
@@
 
-		\If{$E_p \cap E_q \neq \{\}$ for all $q \in S_b$ and $E_p \cap \partial \Omega = \{\}$}{
+		\If{$E_p \cap E_q \neq \{\}$ for all $q \in \pointbatch_b$ and $E_p \cap \partial \Omega = \{\}$}{
 			\tcp{$E_p$ and $E_q$ are the ellipsoids defined in \eqref{eq:support_ellipsoid}}
 
-			Add $p$ to $S_b$
+			Add $p$ to $\pointbatch_b$
 
@@
     \SetKwFunction{FMain}{point\_is\_acceptable}
-	\caption{Choosing one batch of sample points, $S_b$}
+	\caption{Choosing one batch of sample points, $\pointbatch_b$}
 	\label{alg:point_choice}
@@
 Each application of $\Aop$ to a Dirac comb yields one batch of impulse responses, $\impulseresponse_{x_i}$, which yields one batch of convolution kernels, $\convkernel_{x_i}$.
-The Dirac comb, $\dcomb_b$, associated with a batch of sample points, $S_b$, is the sum of Dirac distributions (point sources) centered at the points $x_i \in S_b$.
+The Dirac comb, $\dcomb_b$, associated with a batch of sample points, $\pointbatch_b$, is the sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_b$.
 That is,
 \begin{equation*}
-	\dcomb_b := \sum_{x_i \in S_b} \delta_{x_i}.
-\end{equation*}
-For each batch $S_b$, we compute the action of $\Aop$ on the associated Dirac comb:
+	\dcomb_b := \sum_{x_i \in \pointbatch_b} \delta_{x_i}.
+\end{equation*}
+For each batch $\pointbatch_b$, we compute the action of $\Aop$ on the associated Dirac comb:
 \begin{equation}
@@
 	\label{eq:phi_b}
-	\combresponse_b = \left(\Aop \left(\sum_{x_i \in S_b} \delta_{x_i}\right)^*\right)^* = \sum_{x_i \in S_b} \left(\Aop \delta_{x_i}^*\right)^* = \sum_{x_i \in S_b} \impulseresponse_{x_i}.
+	\combresponse_b = \left(\Aop \left(\sum_{x_i \in \pointbatch_b} \delta_{x_i}\right)^*\right)^* = \sum_{x_i \in \pointbatch_b} \left(\Aop \delta_{x_i}^*\right)^* = \sum_{x_i \in \pointbatch_b} \impulseresponse_{x_i}.
 \end{equation}
@@
 \end{equation}
-for all $x_i \in S_b$. By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the convolution kernels $\convkernel_{x_i}$ associated with every point $x_i \in S_b$.
+for all $x_i \in \pointbatch_b$.
+By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the convolution kernels $\convkernel_{x_i}$ associated with every point $x_i \in \pointbatch_b$.
 
@@
 
-Let $\febasis_1, \febasis_2, \dots, febasis_N$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, and let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_N\right)$ be the corresponding finite element space.
+Let $\febasis_1, \febasis_2, \dots, \febasis_N$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, and let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_N\right)$ be the corresponding finite element space.
 Further, let $p_i \in \mathbb{R}^d$, $i=1,\dots, N$ be the Lagrange nodes associated with the functions $\febasis_i$.
@@
 \begin{equation*}
-	\boldsymbol{\dcomb}^b_i = \sum_{x_j \in S_b} \febasis_i(x_j), \quad i=1,\dots,N.
+	\boldsymbol{\dcomb}^b_i = \sum_{x_j \in \pointbatch_b} \febasis_i(x_j), \quad i=1,\dots,N.
 \end{equation*}

### 2021-09-14 paper/localpsf.tex
fixed a lot of notational collisions

@@
 
-Let $\Omega \subset \mathbb{R}^d$ be a bounded domain in dimension $d=1$, $2$, or $3$. We consider integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ of the form
+Let $\Omega \subset \mathbb{R}^d$ be a bounded domain in dimension $\gdim=1$, $2$, or $3$. We consider integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ of the form
 \begin{equation}
@@
 \label{eq:product_convolution}
-	\left(\AopPc~u\right)(v) := \left( v,~ \sum_{i=1}^r \convkernel_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
+	\left(\AopPc~u\right)(v) := \left( v,~ \sum_{i=1}^\convrank \convkernel_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
 \end{equation}
@@
 \end{equation*}
-with $\convkernel_{x_i}$ defined in \eqref{eq:convolution_kernel} and $\impulseresponse_{x_i}$ defined in \eqref{eq:impulse_response_defn}, to denote local convolution kernels and impulse responses corresponding to a collection of points $\{x_i\}_{i=1}^r \subset \Omega$ scattered throughout the domain.
+with $\convkernel_{x_i}$ defined in \eqref{eq:convolution_kernel} and $\impulseresponse_{x_i}$ defined in \eqref{eq:impulse_response_defn}, to denote local convolution kernels and impulse responses corresponding to a collection of points $\{x_i\}_{i=1}^\convrank \subset \Omega$ scattered throughout the domain.
 The functions $w_i$ are spatially varying weighting functions that are used to interpolate the convolution kernels $\convkernel_i$.
@@
 In order to choose the points $x_i$, we need to estimate the support of $\impulseresponse_i$ \emph{before} we compute it.
-The supports of the functions $\impulseresponse_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of specially chosen functions (three functions if $d=1$, six if $d=2$, ten if $d=3$), then post processing the results (Section \ref{eq:mean_and_covariance_estimation}).
+The supports of the functions $\impulseresponse_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of specially chosen functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$), then post processing the results (Section \ref{eq:mean_and_covariance_estimation}).
 This procedure is the only part of the paper that requires non-negativity of the integral kernel.
@@
 	\label{eq:kernel_entries}
-	\AkerPc(y,x) = \sum_{i=1}^r \convkernel_i(y-x) w_i(x).
+	\AkerPc(y,x) = \sum_{i=1}^\convrank \convkernel_i(y-x) w_i(x).
 \end{equation}
@@
 
-		\For{$b=1,2,\dots,\text{max\_batches}$}{
-			Choose a batch of sample points, $\pointbatch_b$, using Algorithm \ref{alg:point_choice}
-
-			Compute $\combresponse_b$ by applying $\Aop$ to the dirac comb for $\pointbatch_b$ (Section \ref{sec:get_impulse_response})
+		\For{$k=1,2,\dots,\text{max\_batches}$}{
+			Choose a batch of sample points, $\pointbatch_k$, using Algorithm \ref{alg:point_choice}
+
+			Compute $\combresponse_k$ by applying $\Aop$ to the dirac comb for $\pointbatch_k$ (Section \ref{sec:get_impulse_response})
 
@@
 The ellipsoid $E_x$ is not immediately available.
-To use $E_x$, we must first compute the scalar $\spatialvol(x)$, the length-$d$ vector $\spatialmean(x)$, and the $d \times d$ matrix $\spatialcov(x)$ (recall $d \in \{1,2,3\}$ is the spatial dimension of $\Omega$).
+To use $E_x$, we must first compute the scalar $\spatialvol(x)$, the length-$\gdim$ vector $\spatialmean(x)$, and the $\gdim \times \gdim$ matrix $\spatialcov(x)$ (recall $\gdim \in \{1,2,3\}$ is the spatial dimension of $\Omega$).
 The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}.
@@
 But this direct approach is computationally infeasible because and it requires one operator action of $\Aop$ per point $x$ that we wish to evaluate $E_x$ at, and we will need to evaluate $E_x$ at a large number of points.
-Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x$ simultaneously}, by applying $\Aop^T$ to one constant function, $d$ linear functions, and $d(d+1)/2$ quadratic functions.
+Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions.
 We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Theorem \ref{thm:vol_mean_cov} and summarize the method in Algorithm \ref{alg:varhpi_mean_cov}.
@@
 	\label{thm:vol_mean_cov}
-	Let $\spatialvol:\Omega \rightarrow \mathbb{R}$, $\spatialmean:\Omega \rightarrow \mathbb{R}^d$, and $\spatialcov:\Omega \rightarrow \mathbb{R}^{d \times d}$, be the following functions:
+	Let $\spatialvol:\Omega \rightarrow \mathbb{R}$, $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$, and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$, be the following functions:
 	\begin{equation*}
@@
 	\end{equation*}
-	where $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denotes the normalized version of $\impulseresponse_x$, $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$.
-Let $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions:
+	where $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denotes the normalized version of $\impulseresponse_x$, $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$.
+Let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
 	\begin{equation*}
@@
 	\end{subequations}
-	 for $i=1,\dots, d$, $j=1,\dots,d$.
+	 for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$.
 \end{theorem}
@@
 		\tcp{Compute mean $\spatialmean$}
-		\For{$i=1,2,\dots,d$}{
+		\For{$i=1,2,\dots,\gdim$}{
 			Form linear function $L^i(x) = x^i$
@@
 		\tcp{Compute covariance $\spatialcov$}
-		\For{$i=1,2,\dots,d$}{
+		\For{$i=1,2,\dots,\gdim$}{
 			\For{$j=1,\dots,i$}{
@@
 
-We use a greedy ellipsoid packing algorithm to choose sets of sample points, $\pointbatch_b$, such that there is no overlap between the ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch.
+We use a greedy ellipsoid packing algorithm to choose batches of sample points, $\pointbatch_k$, such that there is no overlap between the ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch.
 These sample point batches will be used in Section \ref{sec:get_impulse_response} to compute many impulse responses of $\Aop$ using only a small number of operator actions of $\Aop$.
@@
 We start with a finite set of candidate points $P$.
-To build a batch of sample points, $\pointbatch_b$, first we initialize $\pointbatch_b$ as an empty set.
-Then we select the point $p \in P$ that is the farthest away from all points in previous sample point batches.
+To build $\pointbatch_k$, first we initialize $\pointbatch_k$ as an empty set.
+Then we select a candidate point $p \in P$ that is the farthest away from all points in previous sample point batches.
 That is, $p$ is a maximizer of the following optimization problem:
 \begin{equation*}
-\max_{p \in P} \min_{x \in \pointbatch_1 \cup \dots \cup \pointbatch_{b-1}} \|p - x\|.
+\max_{p \in P} \min_{x \in \pointbatch_1 \cup \dots \cup \pointbatch_{k-1}} \|p - x\|.
 \end{equation*}
@@
 Once $p$ is selected, we remove $p$ from $P$.
-If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_b$, then we add $p$ to $\pointbatch_b$.
+If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_k$, then we add $p$ to $\pointbatch_k$.
 Otherwise we discard $p$.
@@
 We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots$, until the number of batches exceeds a desired threshold.
-In our implementation, for each batch the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, minus all points in previously chosen batches.
+In our implementation, for each batch the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, except those points in previously chosen batches.
 
@@
 
-	\Input{Finite set of candidate points $P\subset \Omega$, \\spatialmean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\domain boundary information $\partial \Omega$, \\previous sample point batches $\pointbatch_1, \dots, \pointbatch_{b-1}$}
-	\Output{Batch of new sample points $\pointbatch_b$}
-
-	Initialize empty new batch of sample points, $\pointbatch_b = \{\}$
+	\Input{Finite set of candidate points $P\subset \Omega$, \\spatialmean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\domain boundary information $\partial \Omega$, \\previous sample point batches $\pointbatch_1, \dots, \pointbatch_{k-1}$}
+	\Output{Batch of new sample points $\pointbatch_k$}
+
+	Initialize empty new batch of sample points, $\pointbatch_k = \{\}$
 
 	\While{$P$ is not empty}{
-		Determine the point $p \in P$ that is farthest from all points in the previous sample point batches $\pointbatch_1,\dots,\pointbatch_{b-1}$
+		Determine the point $p \in P$ that is farthest from all points in the previous sample point batches $\pointbatch_1,\dots,\pointbatch_{k-1}$
 
@@
 
-		\If{$E_p \cap E_q \neq \{\}$ for all $q \in \pointbatch_b$ and $E_p \cap \partial \Omega = \{\}$}{
+		\If{$E_p \cap E_q \neq \{\}$ for all $q \in \pointbatch_k$ and $E_p \cap \partial \Omega = \{\}$}{
 			\tcp{$E_p$ and $E_q$ are the ellipsoids defined in \eqref{eq:support_ellipsoid}}
 
-			Add $p$ to $\pointbatch_b$
+			Add $p$ to $\pointbatch_k$
 
@@
     \SetKwFunction{FMain}{point\_is\_acceptable}
-	\caption{Choosing one batch of sample points, $\pointbatch_b$}
+	\caption{Choosing one batch of sample points, $\pointbatch_k$}
 	\label{alg:point_choice}
@@
 Each application of $\Aop$ to a Dirac comb yields one batch of impulse responses, $\impulseresponse_{x_i}$, which yields one batch of convolution kernels, $\convkernel_{x_i}$.
-The Dirac comb, $\dcomb_b$, associated with a batch of sample points, $\pointbatch_b$, is the sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_b$.
+The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$.
 That is,
 \begin{equation*}
-	\dcomb_b := \sum_{x_i \in \pointbatch_b} \delta_{x_i}.
-\end{equation*}
-For each batch $\pointbatch_b$, we compute the action of $\Aop$ on the associated Dirac comb:
+	\diraccomb_k := \sum_{x_i \in \pointbatch_k} \delta_{x_i}.
+\end{equation*}
+For each batch $\pointbatch_k$, we compute the action of $\Aop$ on the associated Dirac comb:
 \begin{equation}
 	\label{eq:dirac_comb_H_action}
-	\combresponse_b := \left(\Aop \dcomb_b^*\right)^*.
-\end{equation}
-By linearity, the resulting function $\combresponse_b$ may be written as
+	\combresponse_k := \left(\Aop \diraccomb_k^*\right)^*.
+\end{equation}
+By linearity, the resulting function $\combresponse_k$ may be written as
 \begin{equation}
 	\label{eq:phi_b}
-	\combresponse_b = \left(\Aop \left(\sum_{x_i \in \pointbatch_b} \delta_{x_i}\right)^*\right)^* = \sum_{x_i \in \pointbatch_b} \left(\Aop \delta_{x_i}^*\right)^* = \sum_{x_i \in \pointbatch_b} \impulseresponse_{x_i}.
+	\combresponse_k = \left(\Aop \left(\sum_{x_i \in \pointbatch_k} \delta_{x_i}\right)^*\right)^* = \sum_{x_i \in \pointbatch_k} \left(\Aop \delta_{x_i}^*\right)^* = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i}.
 \end{equation}
@@
 	\begin{cases}
-		\combresponse_b(z+x_i), & z+x_i \in E_{x_i} \\
+		\combresponse_k(z+x_i), & z+x_i \in E_{x_i} \\
 		0, & \text{otherwise}
@@
 \end{equation}
-for all $x_i \in \pointbatch_b$.
-By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the convolution kernels $\convkernel_{x_i}$ associated with every point $x_i \in \pointbatch_b$.
+for all $x_i \in \pointbatch_k$.
+By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the convolution kernels $\convkernel_{x_i}$ associated with every point $x_i \in \pointbatch_k$.
 
@@
 To convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), we will need to evaluate $\convkernel_{x_i}(z)$ for a large number of points $z$.
-Per \eqref{eq:varphi_eval}, evaluating $\convkernel_{x_i}(z)$ requires evaluating $\combresponse_b(z+x_i)$.
-In computations, $\combresponse_b$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh.
-Evaluating $\combresponse_b(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell.
+Per \eqref{eq:varphi_eval}, evaluating $\convkernel_{x_i}(z)$ requires evaluating $\combresponse_k(z+x_i)$.
+In computations, $\combresponse_k$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh.
+Evaluating $\combresponse_k(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell.
 Although this procedure is asymptotically scalable, it is slow in practice \cite{FEMEVALSLOW}.
@@
 This is an upfront cost which is done ``offline.'' After $\convkernel_{x_i}$ has been transferred to the regular grid, we use fast regular grid interpolation to perform all subsequent ``online'' evaluations of $\convkernel_{x_i}$.
-This yields a substantial overall speedup because the number of finite element function evaluations required to transfer the functions $\convkernel_{x_i}$ to the regular grid is $O(N)$, while the number of kernel evaluations required to convert $\AopPc$ to H-matrix format is $O(k N \log N)$.
-Here $N$ is the number of finite element degrees of freedom, and $k$ is the H-matrix rank.
+This yields a substantial overall speedup because the number of finite element function evaluations required to transfer the functions $\convkernel_{x_i}$ to the regular grid is $O(\fedim)$, while the number of kernel evaluations required to convert $\AopPc$ to H-matrix format is $O(k \fedim \log \fedim)$.
+Here $\fedim$ is the number of finite element degrees of freedom, and $k$ is the H-matrix rank.
 Furthermore, transferring $\convkernel_{x_i}$ to a regular grid requires evaluating $\convkernel_{x_i}$ at a predetermined and highly structured collection of points.
@@
 
-We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to the following optimization problem:
-\begin{equation}
-	\label{eq:wi_optimization_problem}
-	\begin{aligned}
-		\min_{w_i \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
-		\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,r.
-	\end{aligned}
-\end{equation}
-Here $\Delta$ is the Laplacian on $\Omega$, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$, and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions.
-Using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\convkernel_i$, in a least-squares sense (Theorem \ref{thm:smoothest_interpolant}).
-Also, these weighting functions sum to one pointwise (Proposition \ref{thm:wi_sum_to_one}).
-
-For each $i=1,\dots,r$, optimization problem \eqref{eq:wi_optimization_problem} is an example of the following generic optimization problem:
+Consider the following optimization problem,
 \begin{equation}
@@
 	\min_{v \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta v}_{L^2(\Omega)}^2 \\
-		\text{such that} &\quad v(x_j) = \interpolatedvalues_j, \quad j=1,\dots,r,
+	\text{such that} &\quad v(x_j) = \interpolatedvalues_j, \quad j=1,\dots,\convrank,
 	\end{aligned}
 \end{equation}
-where $\interpolatedvalues_j \in \mathbb{R}$, $j=1,\dots,r$ are now arbitrary function values that $v$ must take at the points $x_j$.
+in which one seeks a smooth function $v$ that takes the values $\interpolatedvalues_j \in \mathbb{R}$ at the points $x_j$.
+Here $\Delta$ is the Laplacian on $\Omega$ and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions.
+We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to optimization problem \eqref{eq:wi_optimization_problem_generic}, with
+\begin{equation*}
+	b_j = w_i(x_j) = 
+	\begin{cases}
+		1, & j=i\\
+		0, & j \neq i.
+	\end{cases}
+\end{equation*}
+Using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\convkernel_i$, in a least-squares sense (Theorem \ref{thm:smoothest_interpolant}).
+Also, these weighting functions sum to one pointwise (Proposition \ref{thm:wi_sum_to_one}).
+
+%For each $i=1,\dots,\convrank$, optimization problem \eqref{eq:wi_optimization_problem} is an example of the following generic optimization problem:
+%\begin{equation}
+%	\label{eq:wi_optimization_problem_generic}
+%	\begin{aligned}
+%		\min_{v \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta v}_{L^2(\Omega)}^2 \\
+%		\text{such that} &\quad v(x_j) = \interpolatedvalues_j, \quad j=1,\dots,\convrank,
+%	\end{aligned}
+%\end{equation}
+%where $\interpolatedvalues_j \in \mathbb{R}$, $j=1,\dots,\convrank$ are given function values that $v$ must take at the points $x_j$.
+
 We may efficiently solve optimization problem \eqref{eq:wi_optimization_problem_generic} via the following procedure:
@@
 		\label{eq:optimal_v}
-		v_*(x) := \beta + \sum_{i=1}^r c_i \secondgreens_i(x),
-	\end{equation}
-	where the constants $c_i \in \mathbb{R}$, $i=1,\dots,r$, and $\beta \in \mathbb{R}$ are the solution to the following linear system:
+		v_*(x) := \beta + \sum_{i=1}^\convrank c_i \secondgreens_i(x),
+	\end{equation}
+	where the constants $c_i \in \mathbb{R}$, $i=1,\dots,\convrank$, and $\beta \in \mathbb{R}$ are the solution to the following linear system:
 	\begin{equation}
@@
 		\begin{bmatrix}
-			S & \mathbf{1} \\ \mathbf{1}^T & 0
+			\pointinteractionmatrix & \mathbf{1} \\ \mathbf{1}^T & 0
 		\end{bmatrix}
@@
 	\end{equation}
-	Here $\mathbf{c} \in \mathbb{R}^r$ is the vector with $i$th component $c_i$, $\mathbf{1} \in \mathbf{R}^r$ is the column vector of all ones, $\mathbf{\interpolatedvalues} \in \mathbb{R}^r$ is the vector with $i$th component $\interpolatedvalues_i$, and $S \in \mathbb{R}^{r \times r}$ is the matrix with entries given by
+	Here $\mathbf{c} \in \mathbb{R}^\convrank$ is the vector with $i$th component $c_i$, $\mathbf{1} \in \mathbf{R}^\convrank$ is the column vector of all ones, $\mathbf{\interpolatedvalues} \in \mathbb{R}^\convrank$ is the vector with $i$th component $\interpolatedvalues_i$, and $\pointinteractionmatrix \in \mathbb{R}^{\convrank \times \convrank}$ is the matrix with entries given by
 	\begin{equation}
 		\label{eq:S_matrix}
-		S_{ij} := \left( \firstgreens_i, \firstgreens_j \right)_{L^2(\Omega)}.
+		\pointinteractionmatrix_{ij} := \left( \firstgreens_i, \firstgreens_j \right)_{L^2(\Omega)}.
 	\end{equation}
@@
 In Theorem \ref{thm:wi_no_breakdown}, we prove that this procedure does not break down, so long as there is no repetition in the points $x_i$.
-Linear system \eqref{eq:interpolation_linear_system} is the KKT system that arises if one solves optimization problem \eqref{eq:wi_optimization_problem_generic} on the $r+1$-dimensional subspace spanned by the functions $\firstgreens_i$, $i=1,\dots,r$, and the constant function $C(x)=1$, rather than solving the optimization problem on the infinite dimensional space $H^2_N(\Omega)$.
+Linear system \eqref{eq:interpolation_linear_system} is the KKT system that arises if one solves optimization problem \eqref{eq:wi_optimization_problem_generic} on the $\convrank+1$-dimensional subspace spanned by the functions $\firstgreens_i$, $i=1,\dots,\convrank$, and the constant function $C(x)=1$, rather than solving the optimization problem on the infinite dimensional space $H^2_N(\Omega)$.
 In Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v_*$ produced by this process solves optimization problem \eqref{eq:wi_optimization_problem_generic} on $H^2_N(\Omega)$.
 
-Constructing all $r$ of the functions $w_i$ thus requires solving the Poisson equation $2r$ times with different right hand side sources to form the functions $\firstgreens_i$ and $\secondgreens_i$, then post-processing these functions.
+Constructing all $\convrank$ of the functions $w_i$ thus requires solving the Poisson equation $2\convrank$ times with different right hand side sources to form the functions $\firstgreens_i$ and $\secondgreens_i$, then post-processing these functions.
 This is shown in Algorithm \ref{alg:weighting_functions_incremental}.
@@
 
-It is important to note that functions in $H^2_N(\Omega)$ only have well-defined pointwise values if the spatial dimension of the domain is $d=1$, $2$, or $3$.
+It is important to note that functions in $H^2_N(\Omega)$ only have well-defined pointwise values if the spatial dimension of the domain is $\gdim=1$, $2$, or $3$.
 Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-posed in three spatial dimensions or fewer.
@@
 	\label{thm:wi_no_breakdown}
-	If all the sample points $x_i$, $i=1,\dots,r$, are distinct, then PDE \eqref{eq:psi_eq} is uniquely solvable for $\firstgreens_i$, PDE \eqref{eq:theta_eq} is uniquely solvable for $\secondgreens_i$, and linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\mathbf{c}$ and $\beta$.
+	If all the sample points $x_i$, $i=1,\dots,\convrank$, are distinct, then PDE \eqref{eq:psi_eq} is uniquely solvable for $\firstgreens_i$, PDE \eqref{eq:theta_eq} is uniquely solvable for $\secondgreens_i$, and linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\mathbf{c}$ and $\beta$.
 That is, the procedure for computing $v_*$ does not break down.
@@
 	Existence and uniqueness for the solutions to PDEs \eqref{eq:psi_eq} and \eqref{eq:theta_eq} follow from the standard elliptic PDE theory.
-By standard elliptic regularity theory, $\firstgreens_i \in H^1(\Omega)$, so $\firstgreens_i \in L^2(\Omega)$, and so all matrix entries $S_{ij}$ are finite.
-
-	Using the definition of $S$ and basic linear algebra, it is straightforward to show that if $S$ were singular, there would exist constants $a_1, a_2, \dots a_r$, which are not all zero, such that
-	\begin{equation*}
-		0 = a_1 \firstgreens_1 + a_2 \firstgreens_2 + \dots a_r \firstgreens_r.
+By standard elliptic regularity theory, $\firstgreens_i \in H^1(\Omega)$, so $\firstgreens_i \in L^2(\Omega)$, and so all matrix entries $\pointinteractionmatrix_{ij}$ are finite.
+
+	Using the definition of $\pointinteractionmatrix$ and basic linear algebra, it is straightforward to show that if $\pointinteractionmatrix$ were singular, there would exist constants $a_1, a_2, \dots a_\convrank$, which are not all zero, such that
+	\begin{equation*}
+		0 = a_1 \firstgreens_1 + a_2 \firstgreens_2 + \dots a_\convrank \firstgreens_\convrank.
 	\end{equation*}
@@
 	\begin{equation*}
-		0 = a_1 \delta_{x_1} + a_2 \delta_{x_2} + \dots a_r \delta_{x_r} - \left(\sum_{i=1}^r a_i\right)\bigg/|\Omega|,
+		0 = a_1 \delta_{x_1} + a_2 \delta_{x_2} + \dots a_\convrank \delta_{x_\convrank} - \left(\sum_{i=1}^\convrank a_i\right)\bigg/|\Omega|,
 	\end{equation*}
 	which is impossible because a nonzero linear combination of delta distributions at different points cannot equal a constant function.
-Thus we must conclude that $S$ is non-singular.
-By block matrix factorization, non-singularity of $S$ implies that linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\mathbf{c}$ and $\beta$.
+Thus we must conclude that $\pointinteractionmatrix$ is non-singular.
+By block matrix factorization, non-singularity of $\pointinteractionmatrix$ implies that linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\mathbf{c}$ and $\beta$.
 
@@
 
-\begin{thm}[Solution of optimization problem \ref{eq:wi_optimization_problem_generic}]
+\begin{thm}[Solution of optimization problem \eqref{eq:wi_optimization_problem_generic}]
 	\label{thm:wi_alg_is_correct}
-	The function $v_*$ given in \eqref{eq:optimal_v} is the unique solution to optimization problem \ref{eq:wi_optimization_problem_generic}.
+	The function $v_*$ given in \eqref{eq:optimal_v} is the unique solution to optimization problem \eqref{eq:wi_optimization_problem_generic}.
 \end{thm}
@@
 	\begin{equation*}
-		X := \{v \in H^2_N(\Omega) : v(x_i) = \interpolatedvalues_i \text{ for }i=1,\dots,r\}
+		X := \{v \in H^2_N(\Omega) : v(x_i) = \interpolatedvalues_i \text{ for }i=1,\dots,\convrank\}
 	\end{equation*}
@@
 	\begin{equation*}
-		X_0 := \{v \in H^2_N(\Omega) : v(x_i) = 0 \text{ for }i=1,\dots,r\}
+		X_0 := \{v \in H^2_N(\Omega) : v(x_i) = 0 \text{ for }i=1,\dots,\convrank\}
 	\end{equation*}	
 	denote the tangent space to $X$.
-Note that pointwise values of functions in $H_N^2(\Omega)$ are well-defined since the Sobolev imbedding theorem (see, e.g., Theorem 7.19 in \cite{Arbogast}) implies that $H^2(\Omega)$ is continuously imbedded in the space of continuous functions $C^0(\Omega)$ when $d < 4$, and we are considering $d \in \{1,2,3\}$.
+Note that pointwise values of functions in $H_N^2(\Omega)$ are well-defined since the Sobolev imbedding theorem (see, e.g., Theorem 7.19 in \cite{Arbogast}) implies that $H^2(\Omega)$ is continuously imbedded in the space of continuous functions $C^0(\Omega)$ when $\gdim < 4$, and we are considering $\gdim \in \{1,2,3\}$.
 The feasible set $X$ is therefore a closed affine subspace of $H^2_N(\Omega)$, the tangent space $X_0$ is a closed subspace of $H^2_N(\Omega)$, and optimization problem \eqref{eq:wi_optimization_problem_generic} is well-defined.
@@
 
-	\paragraph{Feasibility} To show that $v_*$ is feasible (i.e., $v_* \in X$), we must show that $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,r$. We have
+	\paragraph{Feasibility} To show that $v_*$ is feasible (i.e., $v_* \in X$), we must show that $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,\convrank$. We have
 	\begin{align*}
-		v_*(x_i) &= \beta + \sum_{j=1}^r c_j \secondgreens_j(x_i) \\
-		&= \beta + \sum_{j=1}^r c_j \left(\firstgreens_i, \Delta \secondgreens_j\right)_{L^2(\Omega)} + \avg\left(\secondgreens_j\right) \\
-		&= \beta + \sum_{j=1}^r c_j \left(\firstgreens_i, \firstgreens_j\right)_{L^2(\Omega)} \\
-		&= \beta + \sum_{j=1}^r c_j S_{ij}
+		v_*(x_i) &= \beta + \sum_{j=1}^\convrank c_j \secondgreens_j(x_i) \\
+		&= \beta + \sum_{j=1}^\convrank c_j \left(\firstgreens_i, \Delta \secondgreens_j\right)_{L^2(\Omega)} + \avg\left(\secondgreens_j\right) \\
+		&= \beta + \sum_{j=1}^\convrank c_j \left(\firstgreens_i, \firstgreens_j\right)_{L^2(\Omega)} \\
+		&= \beta + \sum_{j=1}^\convrank c_j \pointinteractionmatrix_{ij}
 	\end{align*}
@@
 In the third line we used the facts that, by construction, $\Delta \secondgreens_j = \firstgreens_j$ and $\int_\Omega \secondgreens_j(x) dx = 0$.
-In the fourth line we used the definition of $S$ in \eqref{eq:S_matrix}.
-Hence the condition $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,r$ may be written in matrix form as
-	\begin{equation*}
-		S \mathbf{c} + \beta \mathbf{1} = \mathbf{\interpolatedvalues},
+In the fourth line we used the definition of $\pointinteractionmatrix$ in \eqref{eq:S_matrix}.
+Hence the condition $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,\convrank$ may be written in matrix form as
+	\begin{equation*}
+		\pointinteractionmatrix \mathbf{c} + \beta \mathbf{1} = \mathbf{\interpolatedvalues},
 	\end{equation*}
@@
 	\begin{align*}
-		(\Delta v_*, \Delta p)_{L^2(\Omega)} =& \sum_{i=1}^r c_i(\Delta \secondgreens_i, \Delta p)_{L^2(\Omega)} \\
-		=& \sum_{i=1}^r c_i \left(p(x_i) - \avg\left(p\right)\right) \\
-		=& -\avg\left(p\right) \sum_{i=1}^r c_i. 
+		(\Delta v_*, \Delta p)_{L^2(\Omega)} =& \sum_{i=1}^\convrank c_i(\Delta \secondgreens_i, \Delta p)_{L^2(\Omega)} \\
+		=& \sum_{i=1}^\convrank c_i \left(p(x_i) - \avg\left(p\right)\right) \\
+		=& -\avg\left(p\right) \sum_{i=1}^\convrank c_i. 
 	\end{align*}
@@
 In the third line we used the fact that $p(x_i)=0$ because $p \in X_0$.
-Hence variational equation \eqref{eq:variational_up} holds if $\sum_{i=1}^r c_i = 0$, which may be written in matrix form as $\mathbf{1}^T \mathbf{c} = 0$.
+Hence variational equation \eqref{eq:variational_up} holds if $\sum_{i=1}^\convrank c_i = 0$, which may be written in matrix form as $\mathbf{1}^T \mathbf{c} = 0$.
 This must hold because it is the second block row in system \eqref{eq:interpolation_linear_system}.
@@
 	\label{thm:smoothest_interpolant}
-	Let $y \in \mathbb{R}^d$, $d \in \{1,2,3\}$, and let $\horizinterpolant_y$ be the function
+	Let $y \in \mathbb{R}^\gdim$, $\gdim \in \{1,2,3\}$, and let $\horizinterpolant_y$ be the function
 	\begin{equation}
 		\label{eq:u_defn_in_thm}
-		\horizinterpolant_y(x) := \sum_{i=1}^r \convkernel_i(y)w_i(x) = \AkerPc\left(y+x,x\right),
+		\horizinterpolant_y(x) := \sum_{i=1}^\convrank \convkernel_i(y)w_i(x) = \AkerPc\left(y+x,x\right),
 	\end{equation}
@@
 	\begin{equation*}
-		\sum_{i=1}^r w_i(x) = 1
+		\sum_{i=1}^\convrank w_i(x) = 1
 	\end{equation*}
@@
 	{	
-		\Input{Sample points $\{x_i\}_{i=1}^r$, domain geometry $\Omega$}
-		\Output{Weighting functions $\{w_i\}_{i=1}^r$}
-
-		\For{$k=1,2, \dots, r$}{
+		\Input{Sample points $\{x_i\}_{i=1}^\convrank$, domain geometry $\Omega$}
+		\Output{Weighting functions $\{w_i\}_{i=1}^\convrank$}
+
+		\For{$k=1,2, \dots, \convrank$}{
 
@@
 
-		Form the $r \times r$ matrix $S$ given in \eqref{eq:S_matrix}.
-
-		\For{$i=1,2,\dots,r$}{
+		Form the $\convrank \times \convrank$ matrix $\pointinteractionmatrix$ given in \eqref{eq:S_matrix}.
+
+		\For{$i=1,2,\dots,\convrank$}{
 			Solve the linear system
@@
 			\begin{bmatrix}
-			S & \mathbf{1} \\ \mathbf{1}^T & 0
+			\pointinteractionmatrix & \mathbf{1} \\ \mathbf{1}^T & 0
 			\end{bmatrix}
@@
 			\end{equation*}
-			for $\mathbf{c}^{(i)}=(c_1^{(i)}, c_2^{(i)}, \dots, c_r^{(i)})$ and $\beta^{(i)}$.
-Here $\mathbf{1} = (1,1,\dots,1) \in \mathbb{R}^r$, and $\mathbf{e}^{(i)} = (0,\dots,0,1,0,\dots,0)\in \mathbb{R}^r$ is the unit vector with $i^\text{th}$ component equal to one and all other components equal to zero.
+			for $\mathbf{c}^{(i)}=(c_1^{(i)}, c_2^{(i)}, \dots, c_\convrank^{(i)})$ and $\beta^{(i)}$.
+Here $\mathbf{1} = (1,1,\dots,1) \in \mathbb{R}^\convrank$, and $\mathbf{e}^{(i)} = (0,\dots,0,1,0,\dots,0)\in \mathbb{R}^\convrank$ is the unit vector with $i^\text{th}$ component equal to one and all other components equal to zero.
 
@@
 			\begin{equation*}
-			w_i(x) = \beta^{(i)} + \sum_{k=1}^r c_k^{(i)} \secondgreens_k(x).
-			\end{equation*}
-
-		}
-	}
-	\caption{Compute weighting functions $w_i$, $i=1,\dots,r$.}
+			w_i(x) = \beta^{(i)} + \sum_{k=1}^\convrank c_k^{(i)} \secondgreens_k(x).
+			\end{equation*}
+
+		}
+	}
+	\caption{Compute weighting functions $w_i$, $i=1,\dots,\convrank$.}
 	\label{alg:weighting_functions_incremental}
@@
 
-Let $\febasis_1, \febasis_2, \dots, \febasis_N$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, and let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_N\right)$ be the corresponding finite element space.
-Further, let $p_i \in \mathbb{R}^d$, $i=1,\dots, N$ be the Lagrange nodes associated with the functions $\febasis_i$.
+Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, and let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space.
+Further, let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
 That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
@@
 	\label{eq:eval_y_x}
-	\AkerPc(y,x) = \sum_{i=1}^r w_i(x) \convkernel_{x_i}(y-x).
+	\AkerPc(y,x) = \sum_{i=1}^\convrank w_i(x) \convkernel_{x_i}(y-x).
 \end{equation}
@@
 	\label{eq:defn_of_Akerpcmesh}
-	\AkerPcMesh(y,x) := \sum_{i=1}^N \sum_{j=1}^N \AkerPc(p_i,p_j) \febasis_i(y) \febasis_j(x).
+	\AkerPcMesh(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \AkerPc(p_i,p_j) \febasis_i(y) \febasis_j(x).
 \end{equation}
@@
 That is, $\AopPcMesh : V_h \rightarrow V_h'$.
-Let $u_h \in V_h$ and $v_h \in V_h$, and let $\mathbf{u}$ and $\mathbf{v}$ be the coefficient vectors for $u_h$ and $v_h$, respectively, with respect to the finite element basis $\febasis_1, \dots, \febasis_N$.
+Let $u_h \in V_h$ and $v_h \in V_h$, and let $\mathbf{u}$ and $\mathbf{v}$ be the coefficient vectors for $u_h$ and $v_h$, respectively, with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$.
 I.e.,
@@
 	\label{eq:fem_coeff_basis}
-	u_h(x) = \sum_{i=1}^N \mathbf{u}_i \febasis_i(x),
+	u_h(x) = \sum_{i=1}^\fedim \mathbf{u}_i \febasis_i(x),
 \end{equation}
 and similar for $v_h$.
-Further, let $\mathbf{M} \in \mathbb{R}^{N \times N}$ be the finite element mass matrix, and let $\AkerPcMat \in \mathbb{R}^{N \times N}$ be following matrix of approximate kernel entries:
+Further, let $\mathbf{M} \in \mathbb{R}^{\fedim \times \fedim}$ be the finite element mass matrix, and let $\AkerPcMat \in \mathbb{R}^{\fedim \times \fedim}$ be following matrix of approximate kernel entries:
 \begin{equation}
@@
 We convert $\AkerPcMat$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBPro software package \cite{HLIBPRO}.
-Although $\AkerPcMat$ is a dense $N \times N$ matrix, constructing the H-matrix approximation of $\AkerPcMat$ only requires evaluation of $O(k^2 N \log N)$ entries of $\AkerPc$, and these entries are computed via the formula in Equation \eqref{eq:eval_y_x}.
-Here $k$ is the rank of the highest-rank block in the H-matrix.
+Although $\AkerPcMat$ is a dense $\fedim \times \fedim$ matrix, constructing the H-matrix approximation of $\AkerPcMat$ only requires evaluation of $O(\hmatrixrank^2 \fedim \log \fedim)$ entries of $\AkerPc$, and these entries are computed via the formula in Equation \eqref{eq:eval_y_x}.
+Here $\hmatrixrank$ is the rank of the highest-rank block in the H-matrix.
 A dense representation of $\AkerPcMat$ is never formed.
@@
 While explicit computation of matrix inverses should usually be avoided, here it is computationally acceptable because $\AmatPcSym + \ratpole \mathbf{I}$ is an H-matrix.
-Inversion of H-matrices is a routine procedure which only requires $O(N \log(N)^2)$ work \cite{Hmatrixinverse}.
+Inversion of H-matrices is a routine procedure which only requires $O(\fedim \log(\fedim)^2)$ work \cite{Hmatrixinverse}.
 
@@
 Equation \eqref{eq:desired_property_3} ensures that the negative eigenvalues of $\AmatPcSym$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $\ratfct$.
-There is a tradeoff: if we choose a larger value for $s(\lambda_\text{min})$, then the positive eigenvalues of $\AmatPcSymPlus$ will be closer to the corresponding positive eigenvalues of $\AmatPcSym$ (good), but the erroneous negative eigenvalues of $\AmatPcSym$ will be amplified in magnitude (bad).
+There is a tradeoff: if we choose a larger value for $\ratfct(\lambda_\text{min})$, then the positive eigenvalues of $\AmatPcSymPlus$ will be closer to the corresponding positive eigenvalues of $\AmatPcSym$ (good), but the erroneous negative eigenvalues of $\AmatPcSym$ will be amplified in magnitude (bad).
 This is shown in Figure FIG.
@@
 	\item cost with compression to do a solve
-	\item computational complexity for H-matrix alone vs. prod-conv + h-matrix $O(C k^2 \log N)$ PDE solves, C is big, k is big. Here $10$
+	\item computational complexity for H-matrix alone vs. prod-conv + h-matrix $O(C k^2 \log \fedim)$ PDE solves, C is big, k is big. Here $10$
 	\item HODLR: C is smaller, but k is bigger, vs. H1 peeling process
@@
 
-While a dense $N \times N$ matrix traditionally requires $O(N^2)$ memory to store, H-matrices may be stored using $O(N \log N)$ memory, by storing only the low rank factors for the low rank blocks.
+While a dense $\fedim \times \fedim$ matrix traditionally requires $O(\fedim^2)$ memory to store, H-matrices may be stored using $O(\fedim \log \fedim)$ memory, by storing only the low rank factors for the low rank blocks.
 Recursive algorithms can take advantage of the H-matrix low rank block structure to perform matrix arithmetic fast.
-Conventional dense matrix algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix-matrix addition require either $O(N^2)$ or $O(N^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(N \log(N)^a)$ time and memory for H-matrices.
+Conventional dense matrix algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix-matrix addition require either $O(\fedim^2)$ or $O(\fedim^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(\fedim \log(\fedim)^a)$ time and memory for H-matrices.
 Here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix.
@@
 The first two steps require geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix, but these steps do not depend on the particular values of matrix entries.
-The third step requires us to evaluate $O(N \log N)$ specially-chosen entries of $\AkerPcMat$, and we evaluate these entries using \eqref{eq:Akerpcmat_entries}.
+The third step requires us to evaluate $O(\fedim \log \fedim)$ specially-chosen entries of $\AkerPcMat$, and we evaluate these entries using \eqref{eq:Akerpcmat_entries}.
 
@@
 
-Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\AkerPcMat$ corresponds to the Lagrange node in $\mathbb{R}^d$ associated with the $i^\text{th}$ finite element basis function.
-The columns of the matrix thus correspond to a point cloud in $\mathbb{R}^d$.
+Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\AkerPcMat$ corresponds to the Lagrange node in $\mathbb{R}^\gdim$ associated with the $i^\text{th}$ finite element basis function.
+The columns of the matrix thus correspond to a point cloud in $\mathbb{R}^\gdim$.
 We split this point cloud into two equally sized \emph{child} point clouds, using a hyperplane which is perpendicular to the coordinate axis direction in which the point cloud is widest (e.g., either the $x$, $y$, or $z$ axis in $3$D).
@@
 The quantity $\diam\left(r_1\right)$ is the diameter of the axis aligned bounding box for the point cloud associated with the row cluster $r_1$, and $\diam\left(c_2\right)$ is the analogous diameter associated with the column cluster $c_2$.
-The quantity $\weakadmconst$ is a scalar constant; we use $\weakadmconst=2.0$.
+Here the quantity $\weakadmconst$ is a scalar constant; we use $\weakadmconst=2.0$.
 Basically, if the point clouds associated with $r_1$ and $c_2$ are far away from each other relative to their diameters, then we expect that the corresponding block of the matrix will be low rank.
@@
 \begin{itemize}
-	\item Let $C \in \mathbb{R}^{m \times r}$ be a matrix consisting of a subset of $r$ columns of $X$, such that the span of the columns in $C$ approximates the column space of $X$. 
-	\item Let $R\in\mathbb{R}^{r \times m}$ be a subset of the rows of $X$, such that the span of the rows in $R$ approximates the row space of $X$.
-	\item Let $U \in \mathbb{R}^{r \times r}$ be the block of $X$ consisting of the intersection of the rows from $R$ with the columns from $C$.
+	\item Let $C \in \mathbb{R}^{m \times \hmatrixrank}$ be a matrix consisting of a subset of $\hmatrixrank$ columns of $X$, such that the span of the columns in $C$ approximates the column space of $X$.
+	\item Let $R\in\mathbb{R}^{\hmatrixrank \times m}$ be a subset of the rows of $X$, such that the span of the rows in $R$ approximates the row space of $X$.
+	\item Let $U \in \mathbb{R}^{\hmatrixrank \times \hmatrixrank}$ be the block of $X$ consisting of the intersection of the rows from $R$ with the columns from $C$.
 \end{itemize}
@@
 This results in matrices $R$ and $C$ such that the error in the approximation, \eqref{eq:CUR}, is small.
-The dominant cost of this procedure is the cost of computing $kr$ columns of $X$ and $kr$ rows of $X$, where $k$ is the number of alternating iterations.
+The dominant cost of this procedure is the cost of computing $a \hmatrixrank$ columns of $X$ and $a\hmatrixrank$ rows of $X$, where $a$ is the number of alternating iterations.
 There is also a small linear algebra overhead cost for the determinant maximization process that is performed at each step.
 For more details on adaptive cross low rank approximation, see \cite{ACA}.
-The key point is that the adaptive cross method allows us to form a rank-$r$ approximation of an $m \times m$ block of the matrix via a process that only requires accessing $O(mr)$ entries of that block.
+The key point is that the adaptive cross method allows us to form a rank-$\hmatrixrank$ approximation of an $m \times m$ block of the matrix via a process that only requires accessing $O(m\hmatrixrank)$ entries of that block.
 
@@
 We directly compute all entries of the small dense blocks of $\AkerPcMat$ that are not admissible.
-This process requires us to compute $O(r^2 N \log N)$ entries of $\AkerPcMat$, which is relatively cheap compared to the operator actions of $\Aop$ that are required to form the product-convolution approximation.
+This process requires us to compute $O(\hmatrixrank^2 \fedim \log \fedim)$ entries of $\AkerPcMat$, which is relatively cheap compared to the operator actions of $\Aop$ that are required to form the product-convolution approximation.
 
@@
 \end{align*}
-where $\spatialmean_p, \spatialmean_q \in \mathbb{R}^d$, and $\spatialcov_p, \spatialcov_q \in \mathbb{R}^{d \times d}$ are positive definite. Let $K$ be the following one dimensional convex function:
+where $\spatialmean_p, \spatialmean_q \in \mathbb{R}^\gdim$, and $\spatialcov_p, \spatialcov_q \in \mathbb{R}^{\gdim \times \gdim}$ are positive definite.
+Let $K$ be the following one dimensional convex function:
 \begin{equation*}
@@
 \end{equation*}
-where $P \in \mathbb{R}^{d \times d}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_d)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$.
-The matrix $P$ simultaneously diagonalizes $\spatialcov_p$ and $\spatialcov_q$, in the sense that $P^T\spatialcov_p P = \Lambda$, and $P^T\spatialcov_q P = I$, where $I$ is the $d \times d$ identity matrix.
+where $P \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$.
+The matrix $P$ simultaneously diagonalizes $\spatialcov_p$ and $\spatialcov_q$, in the sense that $P^T\spatialcov_p P = \Lambda$, and $P^T\spatialcov_q P = I$, where $I$ is the $\gdim \times \gdim$ identity matrix.
 Using this diagonalization, and some algebraic manipulations, we may write $K(s)$ as
@@
 	\label{eq:Ks_generalized}
-	K(s) = 1 - \frac{1}{\tau^2} \sum_{i=1}^d \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2,
+	K(s) = 1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2,
 \end{equation}
@@
 
-	$\displaystyle K^* \gets \min_{s \in (0,1)}~1 - \frac{1}{\tau^2} \sum_{i=1}^d \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2$
+	$\displaystyle K^* \gets \min_{s \in (0,1)}~1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2$
 
@@
 \begin{equation*}
-	g(\ratpole) := c_0(\ratpole) + c_1(\ratpole) \lambda_\text{min} + \frac{c_2(\ratpole)}{\lambda_\text{min}+\ratpole} - |\lambda_\text{min}|
+	f(\ratpole) := c_0(\ratpole) + c_1(\ratpole) \lambda_\text{min} + \frac{c_2(\ratpole)}{\lambda_\text{min}+\ratpole} - |\lambda_\text{min}|
 \end{equation*}
 denote the left hand side of \eqref{eq:rational_extra_eq}, but with $c_0(\ratpole)$, $c_1(\ratpole)$, and $c_2(\ratpole)$ being the solution of \eqref{eq:rational_3_linear_system} for the given value of $\ratpole$.
-I.e., within $g(\ratpole)$ the quantities $c_0$, $c_1$, and $c_2$ are no longer free variables; they now depend on $\ratpole$ implicitly through the solution of a linear system.
+I.e., within $f(\ratpole)$ the quantities $c_0$, $c_1$, and $c_2$ are no longer free variables; they now depend on $\ratpole$ implicitly through the solution of a linear system.
 We solve the combined system, \eqref{eq:rational_3_linear_system} and $\eqref{eq:rational_extra_eq}$, by solving the one-dimensional nonlinear root finding problem
@@
 	\label{eq:g_mu_zero}
-	g(\ratpole)=0
+	f(\ratpole)=0
 \end{equation}
@@
 	\label{lem:T_is_linear}
-	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
+	Let $\mathcal{T}:\mathbb{R}^\convrank \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
 	\begin{equation}
@@
 \begin{proof}[Proof of Theorem \ref{thm:smoothest_interpolant}]
-	Let $f = (\convkernel_1(y), \convkernel_2(y), \dots, \convkernel_r(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.
+	Let $f = (\convkernel_1(y), \convkernel_2(y), \dots, \convkernel_\convrank(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $\convrank$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.
 The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
 	\begin{align*}
-		\mathcal{T}(f) &= \convkernel_1(y) \mathcal{T}(e_1) + \convkernel_2(y) \mathcal{T}(e_2) + \dots + \convkernel_r(y) \mathcal{T}(e_r) \\
-		&= \convkernel_1(y) w_1 + \convkernel_2(y) w_2 + \dots + \convkernel_r(y) w_r \\
+		\mathcal{T}(f) &= \convkernel_1(y) \mathcal{T}(e_1) + \convkernel_2(y) \mathcal{T}(e_2) + \dots + \convkernel_\convrank(y) \mathcal{T}(e_\convrank) \\
+		&= \convkernel_1(y) w_1 + \convkernel_2(y) w_2 + \dots + \convkernel_\convrank(y) w_\convrank \\
 		&= \horizinterpolant_y.
@@
 	\begin{equation*}
-		\sum_{i=1}^r w_i(x) = \sum_{i=1}^r \mathcal{T}(e_i)(x) = \mathcal{T}\left(\sum_{i=1}^r e_i\right)(x) = \mathcal{T}(\mathbf{1})(x) = 1,
-	\end{equation*}
-	where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one.
+		\sum_{i=1}^\convrank w_i(x) = \sum_{i=1}^\convrank \mathcal{T}(e_i)(x) = \mathcal{T}\left(\sum_{i=1}^\convrank e_i\right)(x) = \mathcal{T}(\mathbf{1})(x) = 1,
+	\end{equation*}
+	where $\mathbf{1} = (1,1,\dots,1)$ is the length $\convrank$ vector which has all components equal to one.
 The last equality ($\mathcal{T}(\mathbf{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \mathbf{1}$, and yields the minimum possible value of zero for the objective function.
@@
 		\widehat{\impulseresponse}_x:\Omega \rightarrow \mathbb{R}, &\qquad \widehat{\impulseresponse}_x := \impulseresponse_x \big/ \spatialvol(x), \\
-		\spatialmean:\Omega \rightarrow \mathbb{R}^d, &\qquad \spatialmean(x) := \int_\Omega z \widehat{\impulseresponse}_x(z) dz, \\
-		\spatialcov:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\spatialcov(x) := \int_\Omega (z - \spatialmean(x))(z - \spatialmean(x))^T \widehat{\impulseresponse}_x(z) dz.
+		\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim, &\qquad \spatialmean(x) := \int_\Omega z \widehat{\impulseresponse}_x(z) dz, \\
+		\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}, &\qquad\spatialcov(x) := \int_\Omega (z - \spatialmean(x))(z - \spatialmean(x))^T \widehat{\impulseresponse}_x(z) dz.
 	\end{align*}
@@
 
-Let $\febasis_i$, $i=1,\dots,N$ be a set of finite element basis functions, let $h$ denote the mesh size parameter for the finite element mesh, and let
-\begin{equation*}
-	V_h = \Span\left(\febasis_1, \febasis_2, \dots, \febasis_N\right)
+Let $\febasis_i$, $i=1,\dots,\fedim$ be a set of finite element basis functions, let $h$ denote the mesh size parameter for the finite element mesh, and let
+\begin{equation*}
+	V_h = \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)
 \end{equation*}
@@
 Functions $u\in L^2(\Omega)$ are approximated by functions $u_h \in V_h$.
-In turn, functions $u_h\in V_h$ are represented in computations by length $N$ arrays $\mathbf{u}$, such that the array entries of $\mathbf{u}$ are the coordinates of $u_h$ in the finite element basis.
+In turn, functions $u_h\in V_h$ are represented in computations by length $\fedim$ arrays $\mathbf{u}$, such that the array entries of $\mathbf{u}$ are the coordinates of $u_h$ in the finite element basis.
 That is,
@@
 	\label{eq:finite_bijection}
-	u_h(x) = \sum_{i=1}^N \mathbf{u}_i \febasis_i(x).
-\end{equation}
-
-We write $\mathbf{M} \in \mathbb{R}^{N \times N}$ to denote the finite element \emph{mass matrix}, which has entries
+	u_h(x) = \sum_{i=1}^\fedim \mathbf{u}_i \febasis_i(x).
+\end{equation}
+
+We write $\mathbf{M} \in \mathbb{R}^{\fedim \times \fedim}$ to denote the finite element \emph{mass matrix}, which has entries
 \begin{equation*}
@@
 \end{equation*}
-and we write $\mathbb{R}^N_\mathbf{M}$ to denote the space $\mathbb{R}^N$ with the matrix-weighted inner product
+and we write $\mathbb{R}^\fedim_\mathbf{M}$ to denote the space $\mathbb{R}^\fedim$ with the matrix-weighted inner product
 \begin{equation*}
@@
 \end{equation*}
-Direct calculation shows that $V_h$ is isometrically isomorphic $\mathbb{R}^N_\mathbf{M}$.
-That is, equation \eqref{eq:finite_bijection} establishes a bijection between functions $u_h \in V_h$ and coefficient vectors $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, and this bijection preserves the inner product in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u},\mathbf{v}\right)_\mathbf{M}$.
-Likewise, the dual space, $V_h'$, is isometrically isomorphic to $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
-
-Here vectors in $\mathbf{u} \in \mathbb{R}^N_{\mathbf{M}}$ and $\boldsymbol{\sigma} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ are viewed as column vectors, and $\mathbf{u}^T$ and $\boldsymbol{\sigma}^T$ are the row vectors that result from taking the standard matrix transposes of $\mathbf{u}$ and $\boldsymbol{\sigma}$, respectively.
+Direct calculation shows that $V_h$ is isometrically isomorphic $\mathbb{R}^\fedim_\mathbf{M}$.
+That is, equation \eqref{eq:finite_bijection} establishes a bijection between functions $u_h \in V_h$ and coefficient vectors $\mathbf{u} \in \mathbb{R}^\fedim_\mathbf{M}$, and this bijection preserves the inner product in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u},\mathbf{v}\right)_\mathbf{M}$.
+Likewise, the dual space, $V_h'$, is isometrically isomorphic to $\mathbb{R}^\fedim_{\mathbf{M}^{-1}}$.
+
+Here vectors in $\mathbf{u} \in \mathbb{R}^\fedim_{\mathbf{M}}$ and $\boldsymbol{\sigma} \in \mathbb{R}^\fedim_{\mathbf{M}^{-1}}$ are viewed as column vectors, and $\mathbf{u}^T$ and $\boldsymbol{\sigma}^T$ are the row vectors that result from taking the standard matrix transposes of $\mathbf{u}$ and $\boldsymbol{\sigma}$, respectively.
 The discrete dual of a vector $\mathbf{v}$ with respect to the $\mathbf{M}$ inner product is written as $\mathbf{v}^*$, and $\mathbf{v}^*$ is viewed as a column vector.
@@
 
-Here we list the function space operations on $V_h$ and $V_h'$ that are required for this paper, and show how to perform the corresponding matrix and vector operations on $\mathbb{R}^N_{\mathbf{M}}$ and $\mathbb{R}^N_{\mathbf{M}^{-1}}$.
+Here we list the function space operations on $V_h$ and $V_h'$ that are required for this paper, and show how to perform the corresponding matrix and vector operations on $\mathbb{R}^\fedim_{\mathbf{M}}$ and $\mathbb{R}^\fedim_{\mathbf{M}^{-1}}$.
 
 \begin{description}
-	\item[Coefficients of a functional:] The coefficient dual vector, $\boldsymbol{\sigma} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$, corresponding to a linear functional $\sigma_h \in V_h'$ has components given by
-	\begin{equation*}
-		\boldsymbol{\sigma}_i = \sigma_h(\febasis_i), \quad i=1,\dots,N.
+	\item[Coefficients of a functional:] The coefficient dual vector, $\boldsymbol{\sigma} \in \mathbb{R}^\fedim_{\mathbf{M}^{-1}}$, corresponding to a linear functional $\sigma_h \in V_h'$ has components given by
+	\begin{equation*}
+		\boldsymbol{\sigma}_i = \sigma_h(\febasis_i), \quad i=1,\dots,\fedim.
 	\end{equation*}
@@
 	\end{equation*}
-	 Let $\left(\mathbf{g}^\pi\right)^* \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ be the vector with components
-	\begin{equation*}
-		\left(\mathbf{g}^\pi\right)^*_i = \int_\Omega g(x) \febasis_i(x) dx, \quad i=1,\dots,N.
-	\end{equation*}
-	The coefficient vector $\mathbf{g}^\pi \in \mathbb{R}^N_{\mathbf{M}}$ corresponding to $g^\pi_h$ is given by
+	 Let $\left(\mathbf{g}^\pi\right)^* \in \mathbb{R}^\fedim_{\mathbf{M}^{-1}}$ be the vector with components
+	\begin{equation*}
+		\left(\mathbf{g}^\pi\right)^*_i = \int_\Omega g(x) \febasis_i(x) dx, \quad i=1,\dots,\fedim.
+	\end{equation*}
+	The coefficient vector $\mathbf{g}^\pi \in \mathbb{R}^\fedim_{\mathbf{M}}$ corresponding to $g^\pi_h$ is given by
 	\begin{equation*}
@@
 	\end{equation*}
-	\item[Finite element interpolation:] Let $p_i \in \mathbb{R}^d$, $i=1,\dots,N$, be the Lagrange nodes associated with the finite element basis functions $\febasis_i$.
+	\item[Finite element interpolation:] Let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots,\fedim$, be the Lagrange nodes associated with the finite element basis functions $\febasis_i$.
 The finite element interpolant of a function $g \in C\left(\overline{\Omega}\right)$ is the function $g^I \in V_h$ given by
 	\begin{equation*}
-		g_h^I(x) = \sum_{i=1}^N g(p_i) \febasis_i(x).
+		g_h^I(x) = \sum_{i=1}^\fedim g(p_i) \febasis_i(x).
 	\end{equation*}
@@
 	\begin{equation*}
-	\mathbf{g}_i^I = g(p_i), \quad i=1.,\dots,N.
+	\mathbf{g}_i^I = g(p_i), \quad i=1.,\dots,\fedim.
 	\end{equation*}
@@
 	\end{equation*}
-	then $g^I(p_i) = g(p_i)$ for $i=1,\dots,N$ (i.e., the interpolant equals the original function at the Lagrange nodes).
+	then $g^I(p_i) = g(p_i)$ for $i=1,\dots,\fedim$ (i.e., the interpolant equals the original function at the Lagrange nodes).
 Interpolation is typically sufficiently accurate for practical purposes if the Kronecker property is satisfied.
 Projection should be used if the Kronecker property is not satisfied.
-	\item[Matrix representation of an operator:] Let $\mathcal{B}_h : V_h \rightarrow V_h'$ be a linear operator, and let $\mathbf{B} \in \mathbf{R}^{N \times N}$ be the matrix with entries
+	\item[Matrix representation of an operator:] Let $\mathcal{B}_h : V_h \rightarrow V_h'$ be a linear operator, and let $\mathbf{B} \in \mathbf{R}^{\fedim \times \fedim}$ be the matrix with entries
 	\begin{equation*}
@@
 	\end{equation*}
-	The matrix $\mathbf{B}$ maps $\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}$ by matrix multiplication, and
+	The matrix $\mathbf{B}$ maps $\mathbb{R}^\fedim_\mathbf{M} \rightarrow \mathbb{R}^\fedim_{\mathbf{M}^{-1}}$ by matrix multiplication, and
 	\begin{equation*}
@@
 	\end{equation*}
-	\item[Transpose:] The matrix $\mathbf{B}^T$ maps $\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}$ via matrix multiplication, and
+	\item[Transpose:] The matrix $\mathbf{B}^T$ maps $\mathbb{R}^\fedim_\mathbf{M} \rightarrow \mathbb{R}^\fedim_{\mathbf{M}^{-1}}$ via matrix multiplication, and
 	\begin{equation*}
@@
 	\item[Distributions:] Let $\genericdistribution:C\left(\overline{\Omega}\right)\rightarrow \mathbb{R}$ be a distribution.
-If $\genericdistribution(\febasis_i)$ is well-defined for $i=1,\dots,N$, then the restriction of $\genericdistribution$ to domain $V_h$ is a linear functional $\genericdistribution_h\in V_h'$.
-The functional $\genericdistribution_h$ has a coefficient dual vector $\boldsymbol{\genericdistribution} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ with entries
-	\begin{equation*}
-		\boldsymbol{\genericdistribution}_i=\genericdistribution(\febasis_i), \quad i=1,\dots,N.
+If $\genericdistribution(\febasis_i)$ is well-defined for $i=1,\dots,\fedim$, then the restriction of $\genericdistribution$ to domain $V_h$ is a linear functional $\genericdistribution_h\in V_h'$.
+The functional $\genericdistribution_h$ has a coefficient dual vector $\boldsymbol{\genericdistribution} \in \mathbb{R}^\fedim_{\mathbf{M}^{-1}}$ with entries
+	\begin{equation*}
+		\boldsymbol{\genericdistribution}_i=\genericdistribution(\febasis_i), \quad i=1,\dots,\fedim.
 	\end{equation*} 
@@
 	\begin{equation*}
-		\left(\boldsymbol{\delta}_p\right)_i = \febasis_i(p), \quad i=1,\dots,N.
+		\left(\boldsymbol{\delta}_p\right)_i = \febasis_i(p), \quad i=1,\dots,\fedim.
 	\end{equation*}
@@
 		\Input{Operator $\mathbf{A}$}
-		\Output{$\boldsymbol{\spatialvol}\in \mathbb{R}^N$, $\boldsymbol{\spatialmean}\in \mathbb{R}^{d \times N}$, and $\boldsymbol{\spatialcov} \in \mathbb{R}^{d \times d \times N}$}
-
-		Form vector $\mathbf{C} \in \mathbb{R}^N_{\mathbf{M}}$ by either projecting or interpolating constant function $C(x)=1$ onto $V_h$
+		\Output{$\boldsymbol{\spatialvol}\in \mathbb{R}^\fedim$, $\boldsymbol{\spatialmean}\in \mathbb{R}^{\gdim \times \fedim}$, and $\boldsymbol{\spatialcov} \in \mathbb{R}^{\gdim \times \gdim \times \fedim}$}
+
+		Form vector $\mathbf{C} \in \mathbb{R}^\fedim_{\mathbf{M}}$ by either projecting or interpolating constant function $C(x)=1$ onto $V_h$
 
@@
 
-		\For{$i=1,2,\dots,d$}{
-			Form vector $\mathbf{L}^i \in \mathbb{R}^N_{\mathbf{M}}$ by either projecting or interpolating linear function $L^i(x) = x^i$ onto $V_h$
+		\For{$i=1,2,\dots,\gdim$}{
+			Form vector $\mathbf{L}^i \in \mathbb{R}^\fedim_{\mathbf{M}}$ by either projecting or interpolating linear function $L^i(x) = x^i$ onto $V_h$
 
@@
 		}
-		\For{$i=1,2,\dots,d$}{
+		\For{$i=1,2,\dots,\gdim$}{
 			\For{$j=1,\dots,i$}{
@@
 
-In computations, $\spatialvol$, $\spatialmean^{i}$, and $\spatialcov^{ij}$ are replaced with finite element approximations $\spatialvol_h, \spatialmean_h^i, \spatialcov_h^{ij}\in V_h$, respectively, which have coefficient vectors $\boldsymbol{\spatialvol}, \boldsymbol{\spatialmean}^{ij}, \boldsymbol{\spatialcov}^{ij}\in\mathbb{R}^N_\mathbf{M}$, respectively.
+In computations, $\spatialvol$, $\spatialmean^{i}$, and $\spatialcov^{ij}$ are replaced with finite element approximations $\spatialvol_h, \spatialmean_h^i, \spatialcov_h^{ij}\in V_h$, respectively, which have coefficient vectors $\boldsymbol{\spatialvol}, \boldsymbol{\spatialmean}^{ij}, \boldsymbol{\spatialcov}^{ij}\in\mathbb{R}^\fedim_\mathbf{M}$, respectively.
 The functions $C$, $L^i$, and $Q^{ij}$ are replaced with either finite element projections onto $V_h$ (if the finite element basis does not satisfy the Kronecker property), or finite element interpolations onto $V_h$ (if the finite element basis satisfies the Kronecker property).
@@
 \subsection{Discretized sampling of impulse responses}
-In computations, the Dirac comb coefficient dual vector for $\dcomb_h^b$ is the vector $\boldsymbol{\dcomb}^b \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has entries
-\begin{equation*}
-	\boldsymbol{\dcomb}^b_i = \sum_{x_j \in \pointbatch_b} \febasis_i(x_j), \quad i=1,\dots,N.
-\end{equation*}
-The Dirac comb impulse response coefficient vector for $\combresponse^h_b$ is the vector $\boldsymbol{\combresponse}_b \in \mathbb{R}^N_\mathbf{M}$ given by
-\begin{equation*}
-	\boldsymbol{\combresponse}_b = \mathbf{M}^{-1} \mathbf{A}^T \mathbf{M}^{-1}\boldsymbol{\dcomb}^b.
+In computations, the Dirac comb coefficient dual vector for $\diraccomb_h^k$ is the vector $\boldsymbol{\diraccomb}^k \in \mathbb{R}^\fedim_{\mathbf{M}^{-1}}$ which has entries
+\begin{equation*}
+	\boldsymbol{\diraccomb}^k_i = \sum_{x_j \in \pointbatch_k} \febasis_i(x_j), \quad i=1,\dots,\fedim.
+\end{equation*}
+The Dirac comb impulse response coefficient vector for $\combresponse^h_k$ is the vector $\boldsymbol{\combresponse}_k \in \mathbb{R}^\fedim_\mathbf{M}$ given by
+\begin{equation*}
+	\boldsymbol{\combresponse}_k = \mathbf{M}^{-1} \mathbf{A}^T \mathbf{M}^{-1}\boldsymbol{\diraccomb}^k.
 \end{equation*}

### 2021-09-14 paper/localpsf.tex
simplify Poisson weighting function discussion

@@
 
-We choose the weighting functions, $w_i$, to be the solutions to the following optimization problems:
-\begin{equation}
-	\label{eq:wi_optimization_problem_intro}
-	\begin{aligned}
-		\min_{w_i \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
-		\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
-	\end{aligned}
-\end{equation}
-Here $\Delta$ is the Laplacian on $\Omega$, $H^2_N(\Omega)$ is the space of all functions in the Sobolev space $H^2(\Omega)$ with Neumann zero boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$.
-Since the Laplacian measures local deviation from flatness \cite{JACOBSEN}, the function $w_i$ is the flattest function, in a least-squares sense, that takes the value one at the sample point $x_i$, and the value zero at the other sample points $x_j$, $i \neq j$.
+%We choose the weighting functions, $w_i$, to be the solutions to the following optimization problems:
+%\begin{equation}
+%	\label{eq:wi_optimization_problem_intro}
+%	\begin{aligned}
+%		\min_{w_i \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
+%		\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
+%	\end{aligned}
+%\end{equation}
+%Here $\Delta$ is the Laplacian on $\Omega$, $H^2_N(\Omega)$ is the space of all functions in the Sobolev space $H^2(\Omega)$ with Neumann zero boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$.
+Since the Laplacian measures local deviation from flatness \cite{JACOBSEN},
+
+We choose the $i$th weighting function, $w_i$, to be the flattest function, in a least-squares sense, that takes the value one at the sample point $x_i$, and the value zero at the other sample points $x_j$, $i \neq j$.
 Computing the weighting functions $w_i$ requires solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid.
@@
 \end{equation*}
-Using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\convkernel_i$, in a least-squares sense (Theorem \ref{thm:smoothest_interpolant}).
+In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' \cite{JACOBSEN} interpolation of the convolution kernels, $\convkernel_i$, in a least-squares sense.
 Also, these weighting functions sum to one pointwise (Proposition \ref{thm:wi_sum_to_one}).
-
-%For each $i=1,\dots,\convrank$, optimization problem \eqref{eq:wi_optimization_problem} is an example of the following generic optimization problem:
-%\begin{equation}
-%	\label{eq:wi_optimization_problem_generic}
-%	\begin{aligned}
-%		\min_{v \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta v}_{L^2(\Omega)}^2 \\
-%		\text{such that} &\quad v(x_j) = \interpolatedvalues_j, \quad j=1,\dots,\convrank,
-%	\end{aligned}
-%\end{equation}
-%where $\interpolatedvalues_j \in \mathbb{R}$, $j=1,\dots,\convrank$ are given function values that $v$ must take at the points $x_j$.
 
@@
 \end{enumerate}
-In Theorem \ref{thm:wi_no_breakdown}, we prove that this procedure does not break down, so long as there is no repetition in the points $x_i$.
-Linear system \eqref{eq:interpolation_linear_system} is the KKT system that arises if one solves optimization problem \eqref{eq:wi_optimization_problem_generic} on the $\convrank+1$-dimensional subspace spanned by the functions $\firstgreens_i$, $i=1,\dots,\convrank$, and the constant function $C(x)=1$, rather than solving the optimization problem on the infinite dimensional space $H^2_N(\Omega)$.
-In Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v_*$ produced by this process solves optimization problem \eqref{eq:wi_optimization_problem_generic} on $H^2_N(\Omega)$.
+Linear system \eqref{eq:interpolation_linear_system} is the KKT system that arises if one solves optimization problem \eqref{eq:wi_optimization_problem_generic} on the $\convrank+1$-dimensional subspace of $H^2_N(\Omega)$ spanned by the functions $\secondgreens_i$, $i=1,\dots,\convrank$, and the constant function $C(x)=1$.
+In Theorem \ref{thm:wi_no_breakdown}, we prove that this procedure does not break down, and in Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v_*$ produced by this procedure is the solution to optimization problem \eqref{eq:wi_optimization_problem_generic} on $H^2_N(\Omega)$.
 

### 2021-09-14 paper/localpsf.tex
Working through Tuckers comments

@@
 
-%We choose the weighting functions, $w_i$, to be the solutions to the following optimization problems:
-%\begin{equation}
-%	\label{eq:wi_optimization_problem_intro}
-%	\begin{aligned}
-%		\min_{w_i \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
-%		\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
-%	\end{aligned}
-%\end{equation}
-%Here $\Delta$ is the Laplacian on $\Omega$, $H^2_N(\Omega)$ is the space of all functions in the Sobolev space $H^2(\Omega)$ with Neumann zero boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$.
-Since the Laplacian measures local deviation from flatness \cite{JACOBSEN},
-
-We choose the $i$th weighting function, $w_i$, to be the flattest function, in a least-squares sense, that takes the value one at the sample point $x_i$, and the value zero at the other sample points $x_j$, $i \neq j$.
+We choose the $i$th weighting function, $w_i$, to be the flattest function, in a least-squares sense, that takes the value one at the sample point $x_i$, and the value zero at the other sample points $x_j$, $j \neq i$.
 Computing the weighting functions $w_i$ requires solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid.
@@
 We compute convolution kernels $\convkernel_{x_i}$ in batches by applying $\Aop$ to a small number Dirac combs.
-Each application of $\Aop$ to a Dirac comb yields one batch of impulse responses, $\impulseresponse_{x_i}$, which yields one batch of convolution kernels, $\convkernel_{x_i}$.
 The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$.
@@
 \end{equation}
-By linearity, the resulting function $\combresponse_k$ may be written as
+By linearity, the resulting function $\combresponse_k$ may be written as follows:
 \begin{equation}

### 2021-09-15 paper/localpsf.tex
Working through Tuckers comments

@@
 
-Our overall strategy is to form a product-convolution approximation, $\AopPc$, of $\Aop$ using only operator actions, then convert $\AopPc$ into a hierarchical matrix (H-matrix), which we denote by $\AmatPc$.
+Our overall strategy is to form a product-convolution approximation, $\AopPc$, of $\Aop$ using only operator actions, then convert a discretized version of $\AopPc$ into a hierarchical matrix (H-matrix), which we denote by $\AmatPc$.
 Schematically,
@@
 This is the part of the method where we use the four properties of $\Aop$ described above.
-Fast conversion of $\AopPc$ to H-matrix format ($\AopPc \rightarrow \AmatPc$) is possible because integral kernel entries for $\AopPc$ are easily accessible (unlike integral kernel entries for $\Aop$).
+Fast conversion of the discretized version of $\AopPc$ to H-matrix format ($\AopPc \rightarrow \AmatPc$) is possible because integral kernel entries for $\AopPc$ are easily accessible (unlike integral kernel entries for $\Aop$).
 Once in H-matrix format, one may use H-matrix methods to perform further useful linear algebra operations involving $\AmatPc$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
@@
 	\label{eq:minimization_problem}
-	\min_m \quad \frac{1}{2}\|y - f(m,u(m))\|_W^2 + R(m),
-\end{equation}
-where $\|\cdot\|_W$ is weighted norm which depends on the noise covariance, and $R(m)$ is a regularization term.
+	\min_m \quad \frac{1}{2}\|y - f(m,u(m))\|_W^2 + \mathcal{R}(m),
+\end{equation}
+where $\|\cdot\|_W$ is weighted norm which depends on the noise covariance, and $\mathcal{R}(m)$ is a regularization term.
 The objective function in optimization problem \eqref{eq:minimization_problem} is defined indirectly via the implicit function theorem.
-As a result, the Hessian of the objective function, $\mathcal{H}$, is only accessible matrix-free.
+As a result, information about the Hessian of the objective function, $\mathcal{H}$, is only accessible via the action of $\mathcal{H}$ action on vectors.
 In particular, applying $\mathcal{H}$ to a vector requires solving two PDEs \cite{HESSIANACTION}.
@@
 \label{sec:prodconv_intro}
-
-If $\Aop$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then it is straightforward to show that $\Aop$ would be the convolution operator $u \mapsto \left(\convkernel_x \ast u\right)^*$, where the convolution kernel is given by
+Let 
 \begin{equation}
 \label{eq:convolution_kernel}
-	\convkernel_x(y) := \impulseresponse_x(y+x).
-\end{equation}
-The convolution kernel $\convkernel_x$ is the result of translating the impulse response $\impulseresponse_x$ to re-center it at zero instead of $x$.
+\convkernel_x(y) := \impulseresponse_x(y+x)
+\end{equation}
+be the translation of $\impulseresponse_x$ to re-center it at zero instead of $x$.
+If $\Aop$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then for any point $x_0$ we would have
+\begin{align*}
+	\Aker(y,x) &= \Aker(y+(x_0-x),x+(x_0-x)) \\
+	&= \phi_{x_0}(y-x+x_0) \\
+	&= \convkernel_{x_0}(y-x)
+\end{align*}
+which would imply
+\begin{align*}
+	(\Aop u)(v) &= \int_\Omega \int_\Omega v(y) \convkernel_{x_0}(y-x) u(x) dx dy \\
+	&= \int_\Omega v(y) \left(\int_\Omega \convkernel_{x_0}(y-x) u(x) dx\right) dy \\
+	&= \left(v, \convkernel_{x_0} \ast u\right)_{L^2(\Omega)},
+\end{align*}
+where $f \ast g$ denotes convolution of the functions $f$ and $g$, and $\left(f, g\right)_{L^2(\Omega)} := \int_\Omega f(x) g(x) dx$ denotes the $L^2$ inner product of $f$ and $g$ on $\Omega$.
+Hence, if $\Aop$ were perfectly translation invariant, it would be the convolution operator $u \mapsto \left(\convkernel_{x_0} \ast u\right)^*$.
+
 Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\convkernel_x$, varies as $x$ changes.
@@
 \end{equation}
-where $f \cdot g$ denotes pointwise multiplication, $f \ast g$ denotes convolution, and $\left(f, g\right)_{L^2(\Omega)}$ denotes the $L^2$ inner product on $\Omega$, for functions $f,g \in L^2(\Omega)$.
+where $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$ (i.e., $\left(f \cdot g\right)(x) = f(x) g(x)$).
 Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution.
@@
 Hence, $\impulseresponse_x$ is a scaled probability distribution.
-Let $\spatialvol$ denote the scaling factor, let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
+Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor, let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
 We make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
@@
 The ellipsoid $E_x$ is not immediately available.
-To use $E_x$, we must first compute the scalar $\spatialvol(x)$, the length-$\gdim$ vector $\spatialmean(x)$, and the $\gdim \times \gdim$ matrix $\spatialcov(x)$ (recall $\gdim \in \{1,2,3\}$ is the spatial dimension of $\Omega$).
+To determine $E_x$, we must compute the scalar $\spatialvol(x)$, the length-$\gdim$ vector $\spatialmean(x)$, and the $\gdim \times \gdim$ matrix $\spatialcov(x)$ (recall $\gdim \in \{1,2,3\}$ is the spatial dimension of $\Omega$).
 The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}.
@@
 But this direct approach is computationally infeasible because and it requires one operator action of $\Aop$ per point $x$ that we wish to evaluate $E_x$ at, and we will need to evaluate $E_x$ at a large number of points.
-Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions.
+Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions.
 We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Theorem \ref{thm:vol_mean_cov} and summarize the method in Algorithm \ref{alg:varhpi_mean_cov}.
@@
 	\While{$P$ is not empty}{
-		Determine the point $p \in P$ that is farthest from all points in the previous sample point batches $\pointbatch_1,\dots,\pointbatch_{k-1}$
+		Determine the point $p \in P$ that is farthest from all points in previous sample point batches $\pointbatch_1,\dots,\pointbatch_{k-1}$
 

### 2021-09-15 paper/localpsf.tex
done working through Tuckers comments

@@
 This yields a substantial overall speedup because the number of finite element function evaluations required to transfer the functions $\convkernel_{x_i}$ to the regular grid is $O(\fedim)$, while the number of kernel evaluations required to convert $\AopPc$ to H-matrix format is $O(k \fedim \log \fedim)$.
-Here $\fedim$ is the number of finite element degrees of freedom, and $k$ is the H-matrix rank.
+Here $\fedim$ is the number of degrees of freedom in the finite element space used to discretize $\combresponse_k$, and $\hmatrixrank$ is the H-matrix rank.
 Furthermore, transferring $\convkernel_{x_i}$ to a regular grid requires evaluating $\convkernel_{x_i}$ at a predetermined and highly structured collection of points.
-This structure allows for faster determination of which mesh cells that points are in, and allows one to evaluate the finite element function at all regular gridpoints in parallel.
-In contrast, constructing the H-matrix requires evaluating $\convkernel_{x_i}$ at an unpredictable sequence of scattered of points, which is an expensive task for finite element function evaluation, but a cheap task for regular grid interpolation.
-Having $\convkernel_{x_i}$ stored on a regular grid also simplifies the implementation of the method that we use to deal with boundary artifacts, which will be described in Section \ref{sec:boundary_extension}.
+This structure allows for faster determination of which mesh cell that a point is in, and allows one to evaluate the finite element function at all regular gridpoints in parallel.
+In contrast, constructing the H-matrix requires evaluating $\convkernel_{x_i}$ at an unpredictable sequence of scattered points, which is an expensive task for finite element function evaluation, but a cheap task for regular grid interpolation.
+Having $\convkernel_{x_i}$ stored on a regular grid also simplifies the implementation of the method that we use to handle boundary artifacts, which will be described in Section \ref{sec:boundary_extension}.
 
@@
 
-We convert $\AkerPcMat$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBPro software package \cite{HLIBPRO}.
+We convert $\AkerPcMat$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{HLIBPRO}.
 Although $\AkerPcMat$ is a dense $\fedim \times \fedim$ matrix, constructing the H-matrix approximation of $\AkerPcMat$ only requires evaluation of $O(\hmatrixrank^2 \fedim \log \fedim)$ entries of $\AkerPc$, and these entries are computed via the formula in Equation \eqref{eq:eval_y_x}.
@@
 We describe this H-matrix construction process in detail in Appendix \ref{app:h_matrix}.
-We convert $\mathbf{M}$ to H-matrix format using standard H-matrix methods for sparse matrices implemented within HLIBPro, using the same recursive block partitioning structure as was used for $\AkerPcMat$.
+We convert $\mathbf{M}$ to H-matrix format using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same recursive block partitioning structure as was used for $\AkerPcMat$.
 

### 2021-09-15 paper/localpsf.tex
change bold package. Use G, F for greens functions

@@
 for arbitrary functions $u$ and $v$.
-After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\mathbf{A}$, that is typically too large to be built and stored.
-This property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\mathbf{A}$ and $\mathbf{A}^T$ with arbitrary vectors, but one cannot easily\footnote{One could compute $\mathbf{v} = \mathbf{A}\mathbf{e}_j$, where $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero, then extract $\mathbf{A}_{ij} = \mathbf{v}_i$.
-However, this process is wasteful because one computes the entire vector $\mathbf{v}$, then discards all but one component.} access matrix entries $\mathbf{A}_{ij}$.
-In applications, applying $\mathbf{A}$ to a vector may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
+After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\bm{A}$, that is typically too large to be built and stored.
+This property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\bm{A}$ and $\bm{A}^T$ with arbitrary vectors, but one cannot easily\footnote{One could compute $\bm{v} = \bm{A}\bm{e}_j$, where $\bm{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero, then extract $\bm{A}_{ij} = \bm{v}_i$.
+However, this process is wasteful because one computes the entire vector $\bm{v}$, then discards all but one component.} access matrix entries $\bm{A}_{ij}$.
+In applications, applying $\bm{A}$ to a vector may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
 For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
@@
 Hence, $\impulseresponse_x$ is a scaled probability distribution.
-Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor, let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
+Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor (the ``volume'' of $\impulseresponse_x$), let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
 We make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
@@
 		\label{eq:optimal_v}
-		v_*(x) := \beta + \sum_{i=1}^\convrank c_i \secondgreens_i(x),
-	\end{equation}
-	where the constants $c_i \in \mathbb{R}$, $i=1,\dots,\convrank$, and $\beta \in \mathbb{R}$ are the solution to the following linear system:
+		v_*(x) := \constcoeff + \sum_{i=1}^\convrank \secondgreenscoeff_i \secondgreens_i(x),
+	\end{equation}
+	where the constants $\secondgreenscoeff_i \in \mathbb{R}$, $i=1,\dots,\convrank$, and $\constcoeff \in \mathbb{R}$ are the solution to the following linear system:
 	\begin{equation}
@@
 		\begin{bmatrix}
-			\pointinteractionmatrix & \mathbf{1} \\ \mathbf{1}^T & 0
+			\pointinteractionmatrix & \bm{1} \\ \bm{1}^T & 0
 		\end{bmatrix}
 		\begin{bmatrix}
-			\mathbf{c} \\ \beta
+			\bm{\secondgreenscoeff} \\ \constcoeff
 		\end{bmatrix}
@@
 		\begin{bmatrix}
-			\mathbf{\interpolatedvalues} \\ 0
+			\bm{\interpolatedvalues} \\ 0
 		\end{bmatrix}.
 	\end{equation}
-	Here $\mathbf{c} \in \mathbb{R}^\convrank$ is the vector with $i$th component $c_i$, $\mathbf{1} \in \mathbf{R}^\convrank$ is the column vector of all ones, $\mathbf{\interpolatedvalues} \in \mathbb{R}^\convrank$ is the vector with $i$th component $\interpolatedvalues_i$, and $\pointinteractionmatrix \in \mathbb{R}^{\convrank \times \convrank}$ is the matrix with entries given by
+	Here $\bm{\secondgreenscoeff} \in \mathbb{R}^\convrank$ is the vector with $i$th component $\secondgreenscoeff_i$, $\bm{1} \in \bm{R}^\convrank$ is the column vector of all ones, $\bm{\interpolatedvalues} \in \mathbb{R}^\convrank$ is the vector with $i$th component $\interpolatedvalues_i$, and $\pointinteractionmatrix \in \mathbb{R}^{\convrank \times \convrank}$ is the matrix with entries given by
 	\begin{equation}
@@
 	\label{thm:wi_no_breakdown}
-	If all the sample points $x_i$, $i=1,\dots,\convrank$, are distinct, then PDE \eqref{eq:psi_eq} is uniquely solvable for $\firstgreens_i$, PDE \eqref{eq:theta_eq} is uniquely solvable for $\secondgreens_i$, and linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\mathbf{c}$ and $\beta$.
+	If all the sample points $x_i$, $i=1,\dots,\convrank$, are distinct, then PDE \eqref{eq:psi_eq} is uniquely solvable for $\firstgreens_i$, PDE \eqref{eq:theta_eq} is uniquely solvable for $\secondgreens_i$, and linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\bm{\secondgreenscoeff}$ and $\constcoeff$.
 That is, the procedure for computing $v_*$ does not break down.
@@
 Thus we must conclude that $\pointinteractionmatrix$ is non-singular.
-By block matrix factorization, non-singularity of $\pointinteractionmatrix$ implies that linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\mathbf{c}$ and $\beta$.
+By block matrix factorization, non-singularity of $\pointinteractionmatrix$ implies that linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\bm{\secondgreenscoeff}$ and $\constcoeff$.
 
@@
 	\begin{align*}
-		v_*(x_i) &= \beta + \sum_{j=1}^\convrank c_j \secondgreens_j(x_i) \\
-		&= \beta + \sum_{j=1}^\convrank c_j \left(\firstgreens_i, \Delta \secondgreens_j\right)_{L^2(\Omega)} + \avg\left(\secondgreens_j\right) \\
-		&= \beta + \sum_{j=1}^\convrank c_j \left(\firstgreens_i, \firstgreens_j\right)_{L^2(\Omega)} \\
-		&= \beta + \sum_{j=1}^\convrank c_j \pointinteractionmatrix_{ij}
+		v_*(x_i) &= \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \secondgreens_j(x_i) \\
+		&= \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \left(\firstgreens_i, \Delta \secondgreens_j\right)_{L^2(\Omega)} + \avg\left(\secondgreens_j\right) \\
+		&= \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \left(\firstgreens_i, \firstgreens_j\right)_{L^2(\Omega)} \\
+		&= \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \pointinteractionmatrix_{ij}
 	\end{align*}
@@
 	\begin{equation*}
-		\pointinteractionmatrix \mathbf{c} + \beta \mathbf{1} = \mathbf{\interpolatedvalues},
-	\end{equation*}
-	and this must hold because it is the first block row of system \eqref{eq:interpolation_linear_system} which is solved to compute $\mathbf{c}$ and $\beta$. So $v_* \in X$.
+		\pointinteractionmatrix \bm{\secondgreenscoeff} + \constcoeff \bm{1} = \bm{\interpolatedvalues},
+	\end{equation*}
+	and this must hold because it is the first block row of system \eqref{eq:interpolation_linear_system} which is solved to compute $\bm{\secondgreenscoeff}$ and $\constcoeff$. So $v_* \in X$.
 
@@
 
-	\paragraph{Variational equation} To show that $v_*$ satiafies variational equation \eqref{eq:variational_up}, notice that
+	\paragraph{Variational equation} To show that $v_*$ satisfies variational equation \eqref{eq:variational_up}, notice that
 	\begin{align*}
-		(\Delta v_*, \Delta p)_{L^2(\Omega)} =& \sum_{i=1}^\convrank c_i(\Delta \secondgreens_i, \Delta p)_{L^2(\Omega)} \\
-		=& \sum_{i=1}^\convrank c_i \left(p(x_i) - \avg\left(p\right)\right) \\
-		=& -\avg\left(p\right) \sum_{i=1}^\convrank c_i. 
+		(\Delta v_*, \Delta p)_{L^2(\Omega)} =& \sum_{i=1}^\convrank \secondgreenscoeff_i(\Delta \secondgreens_i, \Delta p)_{L^2(\Omega)} \\
+		=& \sum_{i=1}^\convrank \secondgreenscoeff_i \left(p(x_i) - \avg\left(p\right)\right) \\
+		=& -\avg\left(p\right) \sum_{i=1}^\convrank \secondgreenscoeff_i. 
 	\end{align*}
@@
 In the third line we used the fact that $p(x_i)=0$ because $p \in X_0$.
-Hence variational equation \eqref{eq:variational_up} holds if $\sum_{i=1}^\convrank c_i = 0$, which may be written in matrix form as $\mathbf{1}^T \mathbf{c} = 0$.
+Hence variational equation \eqref{eq:variational_up} holds if $\sum_{i=1}^\convrank \secondgreenscoeff_i = 0$, which may be written in matrix form as $\bm{1}^T \bm{\secondgreenscoeff} = 0$.
 This must hold because it is the second block row in system \eqref{eq:interpolation_linear_system}.
@@
 			\begin{bmatrix}
-			\pointinteractionmatrix & \mathbf{1} \\ \mathbf{1}^T & 0
+			\pointinteractionmatrix & \bm{1} \\ \bm{1}^T & 0
 			\end{bmatrix}
 			\begin{bmatrix}
-			\mathbf{c}^{(i)} \\ \beta^{(i)}
+			\bm{\secondgreenscoeff}^{(i)} \\ \constcoeff^{(i)}
 			\end{bmatrix}
@@
 			\begin{bmatrix}
-			\mathbf{e}^{(i)} \\ 0
+			\bm{e}^{(i)} \\ 0
 			\end{bmatrix}.
 			\end{equation*}
-			for $\mathbf{c}^{(i)}=(c_1^{(i)}, c_2^{(i)}, \dots, c_\convrank^{(i)})$ and $\beta^{(i)}$.
-Here $\mathbf{1} = (1,1,\dots,1) \in \mathbb{R}^\convrank$, and $\mathbf{e}^{(i)} = (0,\dots,0,1,0,\dots,0)\in \mathbb{R}^\convrank$ is the unit vector with $i^\text{th}$ component equal to one and all other components equal to zero.
+			for $\bm{\secondgreenscoeff}^{(i)}=(\secondgreenscoeff_1^{(i)}, \secondgreenscoeff_2^{(i)}, \dots, \secondgreenscoeff_\convrank^{(i)})$ and $\constcoeff^{(i)}$.
+Here $\bm{1} = (1,1,\dots,1) \in \mathbb{R}^\convrank$, and $\bm{e}^{(i)} = (0,\dots,0,1,0,\dots,0)\in \mathbb{R}^\convrank$ is the unit vector with $i^\text{th}$ component equal to one and all other components equal to zero.
 
@@
 			\begin{equation*}
-			w_i(x) = \beta^{(i)} + \sum_{k=1}^\convrank c_k^{(i)} \secondgreens_k(x).
+			w_i(x) = \constcoeff^{(i)} + \sum_{k=1}^\convrank \secondgreenscoeff_k^{(i)} \secondgreens_k(x).
 			\end{equation*}
@@
 That is, $\AopPcMesh : V_h \rightarrow V_h'$.
-Let $u_h \in V_h$ and $v_h \in V_h$, and let $\mathbf{u}$ and $\mathbf{v}$ be the coefficient vectors for $u_h$ and $v_h$, respectively, with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$.
+Let $u_h \in V_h$ and $v_h \in V_h$, and let $\bm{u}$ and $\bm{v}$ be the coefficient vectors for $u_h$ and $v_h$, respectively, with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$.
 I.e.,
@@
 	\label{eq:fem_coeff_basis}
-	u_h(x) = \sum_{i=1}^\fedim \mathbf{u}_i \febasis_i(x),
+	u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x),
 \end{equation}
 and similar for $v_h$.
-Further, let $\mathbf{M} \in \mathbb{R}^{\fedim \times \fedim}$ be the finite element mass matrix, and let $\AkerPcMat \in \mathbb{R}^{\fedim \times \fedim}$ be following matrix of approximate kernel entries:
+Further, let $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ be the finite element mass matrix, and let $\AkerPcMat \in \mathbb{R}^{\fedim \times \fedim}$ be following matrix of approximate kernel entries:
 \begin{equation}
@@
 \begin{equation*}
-	\left(\AopPcMesh u_h \right)(v_h) = \mathbf{v}^T \mathbf{M} \AkerPcMat \mathbf{M} \mathbf{u}.
+	\left(\AopPcMesh u_h \right)(v_h) = \bm{v}^T \bm{M} \AkerPcMat \bm{M} \bm{u}.
 \end{equation*}
@@
 	\label{eq:Amatpc_defn}
-	\AmatPc := \mathbf{M} \AkerPcMat \mathbf{M}.
+	\AmatPc := \bm{M} \AkerPcMat \bm{M}.
 \end{equation}
@@
 
-We form an H-matrix representation of $\AmatPc$ by converting $\AkerPcMat$ and $\mathbf{M}$ to H-matrix format, then using fast H-matrix methods to multiply these matrices according to \eqref{eq:Amatpc_defn}.
+We form an H-matrix representation of $\AmatPc$ by converting $\AkerPcMat$ and $\bm{M}$ to H-matrix format, then using fast H-matrix methods to multiply these matrices according to \eqref{eq:Amatpc_defn}.
 Once $\AmatPc$ is in H-matrix format, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods \cite{BORMHACKBUSCHBOOK}.
@@
 We describe this H-matrix construction process in detail in Appendix \ref{app:h_matrix}.
-We convert $\mathbf{M}$ to H-matrix format using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same recursive block partitioning structure as was used for $\AkerPcMat$.
+We convert $\bm{M}$ to H-matrix format using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same recursive block partitioning structure as was used for $\AkerPcMat$.
 
@@
 	\label{eq:rational_matrix_function}
-	\AmatPcSymPlus := c_0 \mathbf{I} + c_1 \AmatPcSym + c_2 \left(\AmatPcSym + \ratpole \mathbf{I}\right)^{-1},
-\end{equation}
-where $\mathbf{I}$ is the identity matrix which has the same shape as $\AmatPcSym$.
+	\AmatPcSymPlus := \ratcoeff_0 \bm{I} + \ratcoeff_1 \AmatPcSym + \ratcoeff_2 \left(\AmatPcSym + \ratpole \bm{I}\right)^{-1},
+\end{equation}
+where $\bm{I}$ is the identity matrix which has the same shape as $\AmatPcSym$.
 Approximation \eqref{eq:rational_matrix_function} may be written as $\AmatPcSymPlus = \ratfct(\AmatPcSym)$, where $\ratfct$ is the following rational function:
@@
 	\label{eq:rational_function_scalar}
-	\ratfct(\lambda) := c_0 + c_1 \lambda + \frac{c_2}{\lambda+\ratpole}.
-\end{equation}
-The scalars $c_0$, $c_1$, $c_2$, and $\ratpole$ are chosen so that the moderate and large positive eigenvalues of $\AmatPcSymPlus$ approximately equal the corresponding moderate and large positive eigenvalues of $\AmatPcSym$, while the negative and small positive eigenvalues of $\AmatPcSym$ are modified so that $\AmatPcSymPlus$ is positive semi-definite.
-Later in this section we will explain how $c_0$, $c_1$, $c_2$, and $\ratpole$ are chosen and how our choice of these scalars affects the spectral properties of $\AmatPcSymPlus$.
+	\ratfct(\lambda) := \ratcoeff_0 + \ratcoeff_1 \lambda + \frac{\ratcoeff_2}{\lambda+\ratpole}.
+\end{equation}
+The scalars $\ratcoeff_0$, $\ratcoeff_1$, $\ratcoeff_2$, and $\ratpole$ are chosen so that the moderate and large positive eigenvalues of $\AmatPcSymPlus$ approximately equal the corresponding moderate and large positive eigenvalues of $\AmatPcSym$, while the negative and small positive eigenvalues of $\AmatPcSym$ are modified so that $\AmatPcSymPlus$ is positive semi-definite.
+Later in this section we will explain how $\ratcoeff_0$, $\ratcoeff_1$, $\ratcoeff_2$, and $\ratpole$ are chosen and how our choice of these scalars affects the spectral properties of $\AmatPcSymPlus$.
 In Figure FIG we illustrate $\ratfct(\lambda)$.
-Forming $\AmatPcSymPlus$ requires computing an H-matrix inverse, $\left(\AmatPcSym + \ratpole \mathbf{I}\right)^{-1}$.
-While explicit computation of matrix inverses should usually be avoided, here it is computationally acceptable because $\AmatPcSym + \ratpole \mathbf{I}$ is an H-matrix.
+Forming $\AmatPcSymPlus$ requires computing an H-matrix inverse, $\left(\AmatPcSym + \ratpole \bm{I}\right)^{-1}$.
+While explicit computation of matrix inverses should usually be avoided, here it is computationally acceptable because $\AmatPcSym + \ratpole \bm{I}$ is an H-matrix.
 Inversion of H-matrices is a routine procedure which only requires $O(\fedim \log(\fedim)^2)$ work \cite{Hmatrixinverse}.
 
-We choose $c_0, c_1, c_2$, and $\ratpole$, to be the solution to the following system of equations:
+We choose $\ratcoeff_0, \ratcoeff_1, \ratcoeff_2$, and $\ratpole$, to be the solution to the following system of equations:
 \begin{subequations}
@@
 We use a geometric admissibility condition (discussed below) to decide which blocks to subdivide, and use the cluster trees for the rows and columns to determine how to subdivide those blocks.
-For the first stage of subdivision, let $r_1$ and $r_2$ be the children row clusters for the root of the row cluster tree, let $c_1$ and $c_2$ be the children column clusters for the root of the column cluster tree.
+For the first stage of subdivision, let $r_1$ and $r_2$ be the children row clusters for the root of the row cluster tree, let $\colcluster_1$ and $\colcluster_2$ be the children column clusters for the root of the column cluster tree.
 The matrix $\AkerPcMat$ is partitioned into blocks as follows:
@@
 \end{equation*}
-where $\left(\AkerPcMat\right)_{11}$ denotes the block of $\AkerPcMat$ with rows $r_1$ and columns $c_1$, $\left(\AkerPcMat\right)_{12}$ denotes the block of $\AkerPcMat$ with rows $r_1$ and columns $c_2$, and so on for $\left(\AkerPcMat\right)_{21}$ and $\left(\AkerPcMat\right)_{22}$.
+where $\left(\AkerPcMat\right)_{11}$ denotes the block of $\AkerPcMat$ with rows $r_1$ and columns $\colcluster_1$, $\left(\AkerPcMat\right)_{12}$ denotes the block of $\AkerPcMat$ with rows $r_1$ and columns $\colcluster_2$, and so on for $\left(\AkerPcMat\right)_{21}$ and $\left(\AkerPcMat\right)_{22}$.
 
@@
 	\label{eq:weak_admissibility_cond}
-	\dist\left(r_1, c_2\right) \ge \weakadmconst \min\left(\diam\left(r_1\right), \diam\left(c_2\right)\right),
+	\dist\left(r_1, \colcluster_2\right) \ge \weakadmconst \min\left(\diam\left(r_1\right), \diam\left(\colcluster_2\right)\right),
 \end{equation}
 then we mark $\left(\AkerPcMat\right)_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone.
-Here $\dist\left(r_1, c_2\right)$ is the Euclidean distance between the axis-aligned bounding box for the point cloud associated with the row cluster $r_1$, and the axis aligned bounding box for the point cloud associated with the column cluster $c_2$.
-The quantity $\diam\left(r_1\right)$ is the diameter of the axis aligned bounding box for the point cloud associated with the row cluster $r_1$, and $\diam\left(c_2\right)$ is the analogous diameter associated with the column cluster $c_2$.
+Here $\dist\left(r_1, \colcluster_2\right)$ is the Euclidean distance between the axis-aligned bounding box for the point cloud associated with the row cluster $r_1$, and the axis aligned bounding box for the point cloud associated with the column cluster $\colcluster_2$.
+The quantity $\diam\left(r_1\right)$ is the diameter of the axis aligned bounding box for the point cloud associated with the row cluster $r_1$, and $\diam\left(\colcluster_2\right)$ is the analogous diameter associated with the column cluster $\colcluster_2$.
 Here the quantity $\weakadmconst$ is a scalar constant; we use $\weakadmconst=2.0$.
-Basically, if the point clouds associated with $r_1$ and $c_2$ are far away from each other relative to their diameters, then we expect that the corresponding block of the matrix will be low rank.
+Basically, if the point clouds associated with $r_1$ and $\colcluster_2$ are far away from each other relative to their diameters, then we expect that the corresponding block of the matrix will be low rank.
 This process is repeated for the other blocks to determine which blocks are admissible and which are not.
@@
 	\begin{bmatrix} 
-		c_0 \\ c_1 \\ c_2 
+		\ratcoeff_0 \\ \ratcoeff_1 \\ \ratcoeff_2 
 	\end{bmatrix}
@@
 	\label{eq:rational_extra_eq}
-	c_0 + c_1 \lambda_\text{min} + \frac{c_2}{\lambda_\text{min}+\ratpole} - |\lambda_\text{min}| = 0.
+	\ratcoeff_0 + \ratcoeff_1 \lambda_\text{min} + \frac{\ratcoeff_2}{\lambda_\text{min}+\ratpole} - |\lambda_\text{min}| = 0.
 \end{equation}
@@
 \begin{equation*}
-	f(\ratpole) := c_0(\ratpole) + c_1(\ratpole) \lambda_\text{min} + \frac{c_2(\ratpole)}{\lambda_\text{min}+\ratpole} - |\lambda_\text{min}|
-\end{equation*}
-denote the left hand side of \eqref{eq:rational_extra_eq}, but with $c_0(\ratpole)$, $c_1(\ratpole)$, and $c_2(\ratpole)$ being the solution of \eqref{eq:rational_3_linear_system} for the given value of $\ratpole$.
-I.e., within $f(\ratpole)$ the quantities $c_0$, $c_1$, and $c_2$ are no longer free variables; they now depend on $\ratpole$ implicitly through the solution of a linear system.
+	f(\ratpole) := \ratcoeff_0(\ratpole) + \ratcoeff_1(\ratpole) \lambda_\text{min} + \frac{\ratcoeff_2(\ratpole)}{\lambda_\text{min}+\ratpole} - |\lambda_\text{min}|
+\end{equation*}
+denote the left hand side of \eqref{eq:rational_extra_eq}, but with $\ratcoeff_0(\ratpole)$, $\ratcoeff_1(\ratpole)$, and $\ratcoeff_2(\ratpole)$ being the solution of \eqref{eq:rational_3_linear_system} for the given value of $\ratpole$.
+I.e., within $f(\ratpole)$ the quantities $\ratcoeff_0$, $\ratcoeff_1$, and $\ratcoeff_2$ are no longer free variables; they now depend on $\ratpole$ implicitly through the solution of a linear system.
 We solve the combined system, \eqref{eq:rational_3_linear_system} and $\eqref{eq:rational_extra_eq}$, by solving the one-dimensional nonlinear root finding problem
@@
 with bound $\ratpole > |\lambda_\text{min}|$ and initial guess $\ratpole_0 = 2 |\lambda_\text{min}|$.
-Each evaluation of $g$ requires solving the $3 \times 3$ linear system \eqref{eq:rational_3_linear_system}, and once the solution $\ratpole$ is found, we solve \eqref{eq:rational_3_linear_system} one more time to get the final values of $c_0$, $c_1$, and $c_2$.
+Each evaluation of $g$ requires solving the $3 \times 3$ linear system \eqref{eq:rational_3_linear_system}, and once the solution $\ratpole$ is found, we solve \eqref{eq:rational_3_linear_system} one more time to get the final values of $\ratcoeff_0$, $\ratcoeff_1$, and $\ratcoeff_2$.
 
@@
 	\begin{equation*}
-		\sum_{i=1}^\convrank w_i(x) = \sum_{i=1}^\convrank \mathcal{T}(e_i)(x) = \mathcal{T}\left(\sum_{i=1}^\convrank e_i\right)(x) = \mathcal{T}(\mathbf{1})(x) = 1,
-	\end{equation*}
-	where $\mathbf{1} = (1,1,\dots,1)$ is the length $\convrank$ vector which has all components equal to one.
-The last equality ($\mathcal{T}(\mathbf{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \mathbf{1}$, and yields the minimum possible value of zero for the objective function.
+		\sum_{i=1}^\convrank w_i(x) = \sum_{i=1}^\convrank \mathcal{T}(e_i)(x) = \mathcal{T}\left(\sum_{i=1}^\convrank e_i\right)(x) = \mathcal{T}(\bm{1})(x) = 1,
+	\end{equation*}
+	where $\bm{1} = (1,1,\dots,1)$ is the length $\convrank$ vector which has all components equal to one.
+The last equality ($\mathcal{T}(\bm{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \bm{1}$, and yields the minimum possible value of zero for the objective function.
 \end{proof}
@@
 Functions $u\in L^2(\Omega)$ are approximated by functions $u_h \in V_h$.
-In turn, functions $u_h\in V_h$ are represented in computations by length $\fedim$ arrays $\mathbf{u}$, such that the array entries of $\mathbf{u}$ are the coordinates of $u_h$ in the finite element basis.
+In turn, functions $u_h\in V_h$ are represented in computations by length $\fedim$ arrays $\bm{u}$, such that the array entries of $\bm{u}$ are the coordinates of $u_h$ in the finite element basis.
 That is,
@@
 	\label{eq:finite_bijection}
-	u_h(x) = \sum_{i=1}^\fedim \mathbf{u}_i \febasis_i(x).
-\end{equation}
-
-We write $\mathbf{M} \in \mathbb{R}^{\fedim \times \fedim}$ to denote the finite element \emph{mass matrix}, which has entries
-\begin{equation*}
-	\mathbf{M}_{ij} = \left(\febasis_i, \febasis_j\right)_{L^2(\Omega)} = \int_{\Omega} \febasis_i(x) \febasis_j(x) dx,
-\end{equation*}
-and we write $\mathbb{R}^\fedim_\mathbf{M}$ to denote the space $\mathbb{R}^\fedim$ with the matrix-weighted inner product
-\begin{equation*}
-	\left(\mathbf{u},\mathbf{v}\right)_\mathbf{M} := \mathbf{v}^T\mathbf{M} \mathbf{u}.
-\end{equation*}
-Direct calculation shows that $V_h$ is isometrically isomorphic $\mathbb{R}^\fedim_\mathbf{M}$.
-That is, equation \eqref{eq:finite_bijection} establishes a bijection between functions $u_h \in V_h$ and coefficient vectors $\mathbf{u} \in \mathbb{R}^\fedim_\mathbf{M}$, and this bijection preserves the inner product in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u},\mathbf{v}\right)_\mathbf{M}$.
-Likewise, the dual space, $V_h'$, is isometrically isomorphic to $\mathbb{R}^\fedim_{\mathbf{M}^{-1}}$.
-
-Here vectors in $\mathbf{u} \in \mathbb{R}^\fedim_{\mathbf{M}}$ and $\boldsymbol{\sigma} \in \mathbb{R}^\fedim_{\mathbf{M}^{-1}}$ are viewed as column vectors, and $\mathbf{u}^T$ and $\boldsymbol{\sigma}^T$ are the row vectors that result from taking the standard matrix transposes of $\mathbf{u}$ and $\boldsymbol{\sigma}$, respectively.
-The discrete dual of a vector $\mathbf{v}$ with respect to the $\mathbf{M}$ inner product is written as $\mathbf{v}^*$, and $\mathbf{v}^*$ is viewed as a column vector.
+	u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
+\end{equation}
+
+We write $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ to denote the finite element \emph{mass matrix}, which has entries
+\begin{equation*}
+	\bm{M}_{ij} = \left(\febasis_i, \febasis_j\right)_{L^2(\Omega)} = \int_{\Omega} \febasis_i(x) \febasis_j(x) dx,
+\end{equation*}
+and we write $\mathbb{R}^\fedim_{\bm{M}}$ to denote the space $\mathbb{R}^\fedim$ with the matrix-weighted inner product
+\begin{equation*}
+	\left(\bm{u}, \bm{v}\right)_{\bm{M}} := \bm{v}^T \bm{M} \bm{u}.
+\end{equation*}
+Direct calculation shows that $V_h$ is isometrically isomorphic $\mathbb{R}^\fedim_{\bm{M}}$.
+That is, equation \eqref{eq:finite_bijection} establishes a bijection between functions $u_h \in V_h$ and coefficient vectors $\bm{u} \in \mathbb{R}^\fedim_{\bm{M}}$, and this bijection preserves the inner product in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\bm{u},\bm{v}\right)_{\bm{M}}$.
+Likewise, the dual space, $V_h'$, is isometrically isomorphic to $\mathbb{R}^\fedim_{\bm{M}^{-1}}$.
+
+Here vectors in $\bm{u} \in \mathbb{R}^\fedim_{\bm{M}}$ and $\bm{\sigma} \in \mathbb{R}^\fedim_{\bm{M}^{-1}}$ are viewed as column vectors, and $\bm{u}^T$ and $\bm{\sigma}^T$ are the row vectors that result from taking the standard matrix transposes of $\bm{u}$ and $\bm{\sigma}$, respectively.
+The discrete dual of a vector $\bm{v}$ with respect to the $\bm{M}$ inner product is written as $\bm{v}^*$, and $\bm{v}^*$ is viewed as a column vector.
 
@@
 
-Here we list the function space operations on $V_h$ and $V_h'$ that are required for this paper, and show how to perform the corresponding matrix and vector operations on $\mathbb{R}^\fedim_{\mathbf{M}}$ and $\mathbb{R}^\fedim_{\mathbf{M}^{-1}}$.
+Here we list the function space operations on $V_h$ and $V_h'$ that are required for this paper, and show how to perform the corresponding matrix and vector operations on $\mathbb{R}^\fedim_{\bm{M}}$ and $\mathbb{R}^\fedim_{\bm{M}^{-1}}$.
 
 \begin{description}
-	\item[Coefficients of a functional:] The coefficient dual vector, $\boldsymbol{\sigma} \in \mathbb{R}^\fedim_{\mathbf{M}^{-1}}$, corresponding to a linear functional $\sigma_h \in V_h'$ has components given by
-	\begin{equation*}
-		\boldsymbol{\sigma}_i = \sigma_h(\febasis_i), \quad i=1,\dots,\fedim.
+	\item[Coefficients of a functional:] The coefficient dual vector, $\bm{\sigma} \in \mathbb{R}^\fedim_{\bm{M}^{-1}}$, corresponding to a linear functional $\sigma_h \in V_h'$ has components given by
+	\begin{equation*}
+		\bm{\sigma}_i = \sigma_h(\febasis_i), \quad i=1,\dots,\fedim.
 	\end{equation*}
@@
 	\begin{equation*}
-		\sigma_h(v_h) = \boldsymbol{\sigma}^T \mathbf{u}.
+		\sigma_h(v_h) = \bm{\sigma}^T \bm{u}.
 	\end{equation*}
@@
 	\end{equation*}
-	The coefficient dual vector $\boldsymbol{\sigma}^*$ corresponding to $\sigma_h$ is given by
-	\begin{equation*} 
-		\boldsymbol{\sigma}^* = \mathbf{M}^{-1} \boldsymbol{\sigma}.
+	The coefficient dual vector $\bm{\sigma}^*$ corresponding to $\sigma_h$ is given by
+	\begin{equation*} 
+		\bm{\sigma}^* = \bm{M}^{-1} \bm{\sigma}.
 	\end{equation*}
@@
 	\end{equation*}
-	 Let $\left(\mathbf{g}^\pi\right)^* \in \mathbb{R}^\fedim_{\mathbf{M}^{-1}}$ be the vector with components
-	\begin{equation*}
-		\left(\mathbf{g}^\pi\right)^*_i = \int_\Omega g(x) \febasis_i(x) dx, \quad i=1,\dots,\fedim.
-	\end{equation*}
-	The coefficient vector $\mathbf{g}^\pi \in \mathbb{R}^\fedim_{\mathbf{M}}$ corresponding to $g^\pi_h$ is given by
-	\begin{equation*}
-		\mathbf{g}^\pi = \mathbf{M}^{-1}\left(\mathbf{g}^\pi\right)^*.
+	 Let $\left(\bm{g}^\pi\right)^* \in \mathbb{R}^\fedim_{\bm{M}^{-1}}$ be the vector with components
+	\begin{equation*}
+		\left(\bm{g}^\pi\right)^*_i = \int_\Omega g(x) \febasis_i(x) dx, \quad i=1,\dots,\fedim.
+	\end{equation*}
+	The coefficient vector $\bm{g}^\pi \in \mathbb{R}^\fedim_{\bm{M}}$ corresponding to $g^\pi_h$ is given by
+	\begin{equation*}
+		\bm{g}^\pi = \bm{M}^{-1}\left(\bm{g}^\pi\right)^*.
 	\end{equation*}
@@
 	\end{equation*}
-	The coefficient vector, $\mathbf{g}$, for $g_h^I$ has components 
-	\begin{equation*}
-	\mathbf{g}_i^I = g(p_i), \quad i=1.,\dots,\fedim.
+	The coefficient vector, $\bm{g}$, for $g_h^I$ has components 
+	\begin{equation*}
+	\bm{g}_i^I = g(p_i), \quad i=1.,\dots,\fedim.
 	\end{equation*}
@@
 Projection should be used if the Kronecker property is not satisfied.
-	\item[Matrix representation of an operator:] Let $\mathcal{B}_h : V_h \rightarrow V_h'$ be a linear operator, and let $\mathbf{B} \in \mathbf{R}^{\fedim \times \fedim}$ be the matrix with entries
-	\begin{equation*}
-		\mathbf{B}_{ij} = \left(\mathcal{B}_h \febasis_j\right)(\febasis_i).
-	\end{equation*}
-	The matrix $\mathbf{B}$ maps $\mathbb{R}^\fedim_\mathbf{M} \rightarrow \mathbb{R}^\fedim_{\mathbf{M}^{-1}}$ by matrix multiplication, and
-	\begin{equation*}
-		\sigma_h = \mathcal{B}_h u_h \quad \Leftrightarrow \quad \boldsymbol{\sigma} = \mathbf{B} \mathbf{u}.
-	\end{equation*}
-	\item[Transpose:] The matrix $\mathbf{B}^T$ maps $\mathbb{R}^\fedim_\mathbf{M} \rightarrow \mathbb{R}^\fedim_{\mathbf{M}^{-1}}$ via matrix multiplication, and
-	\begin{equation*}
-		\sigma_h = \mathcal{B}_h^T u_h \quad \Leftrightarrow \quad \boldsymbol{\sigma} = \mathbf{B}^T \mathbf{u}.
+	\item[Matrix representation of an operator:] Let $\mathcal{B}_h : V_h \rightarrow V_h'$ be a linear operator, and let $\bm{B} \in \bm{R}^{\fedim \times \fedim}$ be the matrix with entries
+	\begin{equation*}
+		\bm{B}_{ij} = \left(\mathcal{B}_h \febasis_j\right)(\febasis_i).
+	\end{equation*}
+	The matrix $\bm{B}$ maps $\mathbb{R}^\fedim_{\bm{M}} \rightarrow \mathbb{R}^\fedim_{\bm{M}^{-1}}$ by matrix multiplication, and
+	\begin{equation*}
+		\sigma_h = \mathcal{B}_h u_h \quad \Leftrightarrow \quad \bm{\sigma} = \bm{B} \bm{u}.
+	\end{equation*}
+	\item[Transpose:] The matrix $\bm{B}^T$ maps $\mathbb{R}^\fedim_{\bm{M}} \rightarrow \mathbb{R}^\fedim_{\bm{M}^{-1}}$ via matrix multiplication, and
+	\begin{equation*}
+		\sigma_h = \mathcal{B}_h^T u_h \quad \Leftrightarrow \quad \bm{\sigma} = \bm{B}^T \bm{u}.
 	\end{equation*}
@@
 If $\genericdistribution(\febasis_i)$ is well-defined for $i=1,\dots,\fedim$, then the restriction of $\genericdistribution$ to domain $V_h$ is a linear functional $\genericdistribution_h\in V_h'$.
-The functional $\genericdistribution_h$ has a coefficient dual vector $\boldsymbol{\genericdistribution} \in \mathbb{R}^\fedim_{\mathbf{M}^{-1}}$ with entries
-	\begin{equation*}
-		\boldsymbol{\genericdistribution}_i=\genericdistribution(\febasis_i), \quad i=1,\dots,\fedim.
+The functional $\genericdistribution_h$ has a coefficient dual vector $\bm{\genericdistribution} \in \mathbb{R}^\fedim_{\bm{M}^{-1}}$ with entries
+	\begin{equation*}
+		\bm{\genericdistribution}_i=\genericdistribution(\febasis_i), \quad i=1,\dots,\fedim.
 	\end{equation*} 
@@
 If the finite element basis functions, $\febasis_i$, are continuous at $p$, then restricting the domain of $\delta_p$ to the space $V_h$ yields a linear functional in $V_h'$.
-The corresponding coefficient dual vector, $\boldsymbol{\delta}_p$, has entries
-	\begin{equation*}
-		\left(\boldsymbol{\delta}_p\right)_i = \febasis_i(p), \quad i=1,\dots,\fedim.
+The corresponding coefficient dual vector, $\bm{\delta}_p$, has entries
+	\begin{equation*}
+		\left(\bm{\delta}_p\right)_i = \febasis_i(p), \quad i=1,\dots,\fedim.
 	\end{equation*}
@@
 	\begin{equation*}
-		\sigma_h = \langle \mathcal{B}_h, \genericdistribution_h \rangle \quad \Leftrightarrow \quad \sigma_h = \mathcal{B}_h \genericdistribution_h^* \quad \Leftrightarrow \quad \boldsymbol{\sigma} = \mathbf{B} \mathbf{M}^{-1} \boldsymbol{\genericdistribution}.
+		\sigma_h = \langle \mathcal{B}_h, \genericdistribution_h \rangle \quad \Leftrightarrow \quad \sigma_h = \mathcal{B}_h \genericdistribution_h^* \quad \Leftrightarrow \quad \bm{\sigma} = \bm{B} \bm{M}^{-1} \bm{\genericdistribution}.
 	\end{equation*}
@@
 	{	
-		\Input{Operator $\mathbf{A}$}
-		\Output{$\boldsymbol{\spatialvol}\in \mathbb{R}^\fedim$, $\boldsymbol{\spatialmean}\in \mathbb{R}^{\gdim \times \fedim}$, and $\boldsymbol{\spatialcov} \in \mathbb{R}^{\gdim \times \gdim \times \fedim}$}
-
-		Form vector $\mathbf{C} \in \mathbb{R}^\fedim_{\mathbf{M}}$ by either projecting or interpolating constant function $C(x)=1$ onto $V_h$
-
-		$\boldsymbol{\spatialvol} = \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{C}\right)^*$
+		\Input{Operator $\bm{A}$}
+		\Output{$\bm{\spatialvol}\in \mathbb{R}^\fedim$, $\bm{\spatialmean}\in \mathbb{R}^{\gdim \times \fedim}$, and $\bm{\spatialcov} \in \mathbb{R}^{\gdim \times \gdim \times \fedim}$}
+
+		Form vector $\bm{C} \in \mathbb{R}^\fedim_{\bm{M}}$ by either projecting or interpolating constant function $C(x)=1$ onto $V_h$
+
+		$\bm{\spatialvol} = \left(\bm{M}^{-1}\bm{A}^T \bm{C}\right)^*$
 
 		\For{$i=1,2,\dots,\gdim$}{
-			Form vector $\mathbf{L}^i \in \mathbb{R}^\fedim_{\mathbf{M}}$ by either projecting or interpolating linear function $L^i(x) = x^i$ onto $V_h$
-
-			$\mathbf{\spatialmean}^i = \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{L}^i\right)^* / \boldsymbol{\spatialvol}$
+			Form vector $\bm{L}^i \in \mathbb{R}^\fedim_{\bm{M}}$ by either projecting or interpolating linear function $L^i(x) = x^i$ onto $V_h$
+
+			$\bm{\spatialmean}^i = \left(\bm{M}^{-1}\bm{A}^T \bm{L}^i\right)^* / \bm{\spatialvol}$
 		}
@@
 			\For{$j=1,\dots,i$}{
-				Form quadratic function $\mathbf{Q}^{ij} = x^i x^j$
-
-				$\boldsymbol{\spatialcov}^{ij} = \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{Q}^{ij}\right)^* / \boldsymbol{\spatialvol} - \boldsymbol{\spatialmean}^i\cdot \boldsymbol{\spatialmean}^j$
-
-				$\boldsymbol{\spatialcov}^{ij} = \boldsymbol{\spatialcov}^{ji}$
+				Form quadratic function $\bm{Q}^{ij} = x^i x^j$
+
+				$\bm{\spatialcov}^{ij} = \left(\bm{M}^{-1}\bm{A}^T \bm{Q}^{ij}\right)^* / \bm{\spatialvol} - \bm{\spatialmean}^i\cdot \bm{\spatialmean}^j$
+
+				$\bm{\spatialcov}^{ij} = \bm{\spatialcov}^{ji}$
 
@@
 
-In computations, $\spatialvol$, $\spatialmean^{i}$, and $\spatialcov^{ij}$ are replaced with finite element approximations $\spatialvol_h, \spatialmean_h^i, \spatialcov_h^{ij}\in V_h$, respectively, which have coefficient vectors $\boldsymbol{\spatialvol}, \boldsymbol{\spatialmean}^{ij}, \boldsymbol{\spatialcov}^{ij}\in\mathbb{R}^\fedim_\mathbf{M}$, respectively.
+In computations, $\spatialvol$, $\spatialmean^{i}$, and $\spatialcov^{ij}$ are replaced with finite element approximations $\spatialvol_h, \spatialmean_h^i, \spatialcov_h^{ij}\in V_h$, respectively, which have coefficient vectors $\bm{\spatialvol}, \bm{\spatialmean}^{ij}, \bm{\spatialcov}^{ij}\in\mathbb{R}^\fedim_{\bm{M}}$, respectively.
 The functions $C$, $L^i$, and $Q^{ij}$ are replaced with either finite element projections onto $V_h$ (if the finite element basis does not satisfy the Kronecker property), or finite element interpolations onto $V_h$ (if the finite element basis satisfies the Kronecker property).
 In our numerical examples we use finite element basis functions that satisfy the Kronecker property, and we use finite element interpolations.
-The coefficient vectors corresponding to either the interpolations or projections are denoted $\mathbf{C}$, $\mathbf{L}^i$, and $\mathbf{Q}^{ij}$, respectively.
+The coefficient vectors corresponding to either the interpolations or projections are denoted $\bm{C}$, $\bm{L}^i$, and $\bm{Q}^{ij}$, respectively.
 The discretized versions of \eqref{eq:vol_mean_var_thm1}, \eqref{eq:vol_mean_var_thm2}, and \eqref{eq:vol_mean_var_thm3} are
 \begin{align*}
-	\boldsymbol{\spatialvol} &= \mathbf{M}^{-1}\mathbf{A}^T\mathbf{C} \\
-	\boldsymbol{\spatialmean}^i &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{L}^i\right) / \boldsymbol{\spatialvol}\\
-	\boldsymbol{\spatialcov}^{ij} &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{Q}^{ij}\right) / \boldsymbol{\spatialvol} - \boldsymbol{\spatialmean}^i\cdot \boldsymbol{\spatialmean}^j,
+	\bm{\spatialvol} &= \bm{M}^{-1}\bm{A}^T\bm{C} \\
+	\bm{\spatialmean}^i &= \left(\bm{M}^{-1}\bm{A}^T \bm{L}^i\right) / \bm{\spatialvol}\\
+	\bm{\spatialcov}^{ij} &= \left(\bm{M}^{-1}\bm{A}^T \bm{Q}^{ij}\right) / \bm{\spatialvol} - \bm{\spatialmean}^i\cdot \bm{\spatialmean}^j,
 \end{align*}
-respectively. Here $\mathbf{f} \cdot \mathbf{g}$ and $\mathbf{f} / \mathbf{g}$ denote the component-wise multiplication and division of vectors, respectively.
+respectively. Here $\bm{f} \cdot \bm{g}$ and $\bm{f} / \bm{g}$ denote the component-wise multiplication and division of vectors, respectively.
 
 \subsection{Discretized sampling of impulse responses}
-In computations, the Dirac comb coefficient dual vector for $\diraccomb_h^k$ is the vector $\boldsymbol{\diraccomb}^k \in \mathbb{R}^\fedim_{\mathbf{M}^{-1}}$ which has entries
-\begin{equation*}
-	\boldsymbol{\diraccomb}^k_i = \sum_{x_j \in \pointbatch_k} \febasis_i(x_j), \quad i=1,\dots,\fedim.
-\end{equation*}
-The Dirac comb impulse response coefficient vector for $\combresponse^h_k$ is the vector $\boldsymbol{\combresponse}_k \in \mathbb{R}^\fedim_\mathbf{M}$ given by
-\begin{equation*}
-	\boldsymbol{\combresponse}_k = \mathbf{M}^{-1} \mathbf{A}^T \mathbf{M}^{-1}\boldsymbol{\diraccomb}^k.
+In computations, the Dirac comb coefficient dual vector for $\diraccomb_h^k$ is the vector $\bm{\diraccomb}^k \in \mathbb{R}^\fedim_{\bm{M}^{-1}}$ which has entries
+\begin{equation*}
+	\bm{\diraccomb}^k_i = \sum_{x_j \in \pointbatch_k} \febasis_i(x_j), \quad i=1,\dots,\fedim.
+\end{equation*}
+The Dirac comb impulse response coefficient vector for $\combresponse^h_k$ is the vector $\bm{\combresponse}_k \in \mathbb{R}^\fedim_{\bm{M}}$ given by
+\begin{equation*}
+	\bm{\combresponse}_k = \bm{M}^{-1} \bm{A}^T \bm{M}^{-1}\bm{\diraccomb}^k.
 \end{equation*}

### 2021-09-15 paper/localpsf.tex
add proof of optimality for interpolated convolution kernel

@@
 	\label{thm:smoothest_interpolant}
-	Let $y \in \mathbb{R}^\gdim$, $\gdim \in \{1,2,3\}$, and let $\horizinterpolant_y$ be the function
-	\begin{equation}
-		\label{eq:u_defn_in_thm}
-		\horizinterpolant_y(x) := \sum_{i=1}^\convrank \convkernel_i(y)w_i(x) = \AkerPc\left(y+x,x\right),
-	\end{equation}
-	where the weighting functions $w_i$ are the solutions to optimization problem \eqref{eq:wi_optimization_problem}.
-Then for each $y \in \Omega$, the function $\horizinterpolant_y$ solves the optimization problem
+	Let 
+	\begin{equation}
+		\label{eq:conv_kernel_pc}
+		\convkernel_x^\text{pc}(y) := \sum_{i=1}^\convrank \convkernel_i(y)w_i(x)
+	\end{equation}
+	denote the local convolution kernel for the product-convolution approximation, and let $\horizinterpolant_y^\text{pc}$ be the reinterpretation of $\convkernel_x^\text{pc}(y)$ as a function of $x$ instead of $y$.
+That is,
+	\begin{equation*}
+		\horizinterpolant_y^\text{pc}(x) := \convkernel_x^\text{pc}(y).
+	\end{equation*}
+	Then for each $y \in \Omega$, the function $\horizinterpolant_y^\text{pc}$ is the unique solution to the following optimization problem:
 	\begin{equation}
@@
 		\begin{aligned}
-			\min_{\horizinterpolant_y \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \horizinterpolant_y}_{L^2(\Omega)}^2 \\
-			\text{such that} &\quad \horizinterpolant_y(x_i) = \convkernel_i(y), \quad i=1,\dots,q.
+			\min_{\horizinterpolant \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \horizinterpolant}_{L^2(\Omega)}^2 \\
+			\text{such that} &\quad \horizinterpolant(x_i) = \convkernel_i(y), \quad i=1,\dots,\convrank.
 		\end{aligned}
@@
 \end{thm}
+
+\begin{proof}
+	The optimization problem that $w_i$ solves, and optimization problem \eqref{eq:uy_optimization_problem}, both take the form of optimization problem \eqref{eq:wi_optimization_problem_generic}, but with different constraint data.
+Let $\mathcal{T}:\bm{b} \mapsto v_*$ denote the solution operator for optimization problem \eqref{eq:wi_optimization_problem_generic} as a function of the constraint data,  $\bm{b}$.
+The mapping $\mathcal{T}$ is a linear operator because the procedure we presented for solving optimization problem \eqref{eq:wi_optimization_problem_generic} is linear in the constraint data, and we proved that this procedure always yields the unique solution of optimization problem \eqref{eq:wi_optimization_problem_generic} in Theorem \ref{thm:wi_no_breakdown} and Theorem \ref{thm:wi_alg_is_correct}.
+
+	The constraint data for the optimization problem that $w_i$ solves is the unit vector $\bm{e}^{(i)}=(0,\dots,0,1,0,\dots,0) \in \mathbb{R}^\convrank$ which has $i$th component equal to one, and all other components equal to zero.
+Hence
+	\begin{equation*}
+		\label{eq:wi_via_solution_operator}
+		w_i = \mathcal{T}\left(\bm{e}^{(i)}\right).
+	\end{equation*}
+	The constraint data for optimization problem \eqref{eq:uy_optimization_problem} is the vector $(\convkernel_1(y), \convkernel_2(y), \dots, \convkernel_\convrank(y))$, so the solution, $\theta_*$, to optimization problem \eqref{eq:uy_optimization_problem} is given by
+	\begin{equation*}
+		\theta_* = \mathcal{T}\left((\convkernel_1(y), \convkernel_2(y), \dots, \convkernel_\convrank(y))\right).
+	\end{equation*}
+	Using the definition of $\horizinterpolant_y^\text{pc}(x)$ and linearity of $\mathcal{T}$, we have
+	\begin{align*}
+		\horizinterpolant_y^\text{pc}(x) &= \sum_{i=1}^\convrank \convkernel_i(y)w_i(x) \\
+		&= \sum_{i=1}^\convrank \convkernel_i(y)\mathcal{T}\left(\bm{e}^{(i)}\right)(x) \\
+		&= \mathcal{T}\left(\sum_{i=1}^\convrank \convkernel_i(y) \bm{e}^{(i)} \right)(x) \\
+		&= \mathcal{T}\left((\convkernel_1(y), \convkernel_2(y), \dots, \convkernel_\convrank(y))\right)(x) = \theta_*(x).
+	\end{align*}
+	Hence $\horizinterpolant_y^\text{pc}$ is the unique solution to optimization problem \eqref{eq:uy_optimization_problem}.
+\end{proof}
 

### 2021-09-15 paper/localpsf.tex
minor improvements

@@
 By construction the functions $\secondgreens_i$ have Neumann zero boundary conditions, so $\secondgreens_i \in H^2_N(\Omega)$.
-Since $v_*$ is a linear combination of the functions $\secondgreens_i$ and the constant function $C(x)=1$, $v_* \in H^2_N(\Omega)$.
+Since $v_*$ is a linear combination of the functions $\secondgreens_i$ and the constant function $C(x)=1$ which is in $H^2_N(\Omega)$, we have $v_* \in H^2_N(\Omega)$.
 \end{proof}
@@
 	\end{align*}
-	Hence $\horizinterpolant_y^\text{pc}$ is the unique solution to optimization problem \eqref{eq:uy_optimization_problem}.
+	Thus $\horizinterpolant_y^\text{pc}$ is the unique solution to optimization problem \eqref{eq:uy_optimization_problem}.
 \end{proof}
@@
 
-\begin{lem}[Linearity of weighting function solution operator]
-	\label{lem:T_is_linear}
-	Let $\mathcal{T}:\mathbb{R}^\convrank \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
-	\begin{equation}
-		\label{eq:opt_soln_operator}
-		\mathcal{T}(f) := \horizinterpolant_*,
-	\end{equation}
-	where $\horizinterpolant_*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
-\end{lem}
-
-\begin{proof}
-	asdf
-\end{proof}
-
-\begin{proof}[Proof of Theorem \ref{thm:smoothest_interpolant}]
-	Let $f = (\convkernel_1(y), \convkernel_2(y), \dots, \convkernel_\convrank(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $\convrank$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.
-The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
-	\begin{align*}
-		\mathcal{T}(f) &= \convkernel_1(y) \mathcal{T}(e_1) + \convkernel_2(y) \mathcal{T}(e_2) + \dots + \convkernel_\convrank(y) \mathcal{T}(e_\convrank) \\
-		&= \convkernel_1(y) w_1 + \convkernel_2(y) w_2 + \dots + \convkernel_\convrank(y) w_\convrank \\
-		&= \horizinterpolant_y.
-	\end{align*}
-	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $\horizinterpolant_y$ in \eqref{eq:u_defn_in_thm}.
-\end{proof}
-
-\begin{proof}[Proof of Proposition \ref{thm:wi_sum_to_one}]
-	By linearity of $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), we have
-	\begin{equation*}
-		\sum_{i=1}^\convrank w_i(x) = \sum_{i=1}^\convrank \mathcal{T}(e_i)(x) = \mathcal{T}\left(\sum_{i=1}^\convrank e_i\right)(x) = \mathcal{T}(\bm{1})(x) = 1,
-	\end{equation*}
-	where $\bm{1} = (1,1,\dots,1)$ is the length $\convrank$ vector which has all components equal to one.
-The last equality ($\mathcal{T}(\bm{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \bm{1}$, and yields the minimum possible value of zero for the objective function.
-\end{proof}
-
 \begin{defn}[Moments of $\impulseresponse_x$]

### 2021-09-20 paper/localpsf.tex
move Sij=Fi(xj) into lemma

@@
 
-We may efficiently solve optimization problem \eqref{eq:wi_optimization_problem_generic} via the following procedure:
+The solution to optimization problem \eqref{eq:wi_optimization_problem_generic} is given via the following procedure:
 \begin{enumerate}
@@
 
-\begin{lem}[Preparation for Theorem \ref{thm:wi_alg_is_correct}]
+\begin{lem}[Integral representation of delta distribution on $H^2_N(\Omega)$]
 	\label{lem:psi_delta_eval}
@@
 	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\firstgreens_j$ onto $h$, and using the Neumann zero boundary conditions for $\firstgreens_j$ and $h$ to eliminate boundary integral terms.
+\end{proof}
+
+\begin{lem}[Characterization of $\pointinteractionmatrix$]
+	\label{lem:S_matrix_characterizations}
+	We have
+	\begin{equation*}
+		\pointinteractionmatrix_{ij} = \secondgreens_j(x_i).
+	\end{equation*}
+\end{lem}
+
+\begin{proof}
+	We compute
+	\begin{equation*}
+		\secondgreens_j(x_i) = \left(\firstgreens_i, \Delta \secondgreens_j\right)_{L^2(\Omega)} + \avg\left(\secondgreens_j\right)
+		= \left(\firstgreens_i, \firstgreens_j\right)_{L^2(\Omega)} = \pointinteractionmatrix_{ij}.
+	\end{equation*}
+	In the first equality we used Lemma \ref{lem:psi_delta_eval}.
+In the second equality we used the facts that $\Delta \secondgreens_j = \firstgreens_j$ and $\avg(\secondgreens_j) = \left(1 / |\Omega|\right) \int_\Omega \secondgreens_j = 0$ from the definition of $\secondgreens_j$ in \eqref{eq:theta_eq}.
 \end{proof}
@@
 
-	\paragraph{Feasibility} To show that $v_*$ is feasible (i.e., $v_* \in X$), we must show that $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,\convrank$. We have
-	\begin{align*}
-		v_*(x_i) &= \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \secondgreens_j(x_i) \\
-		&= \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \left(\firstgreens_i, \Delta \secondgreens_j\right)_{L^2(\Omega)} + \avg\left(\secondgreens_j\right) \\
-		&= \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \left(\firstgreens_i, \firstgreens_j\right)_{L^2(\Omega)} \\
-		&= \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \pointinteractionmatrix_{ij}
-	\end{align*}
-	In the first line we used the definition of $v_*$ in \eqref{eq:optimal_v}.
-In the second line we used Lemma \ref{lem:psi_delta_eval}.
-In the third line we used the facts that, by construction, $\Delta \secondgreens_j = \firstgreens_j$ and $\int_\Omega \secondgreens_j(x) dx = 0$.
-In the fourth line we used the definition of $\pointinteractionmatrix$ in \eqref{eq:S_matrix}.
-Hence the condition $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,\convrank$ may be written in matrix form as
-	\begin{equation*}
-		\pointinteractionmatrix \bm{\secondgreenscoeff} + \constcoeff \bm{1} = \bm{\interpolatedvalues},
-	\end{equation*}
-	and this must hold because it is the first block row of system \eqref{eq:interpolation_linear_system} which is solved to compute $\bm{\secondgreenscoeff}$ and $\constcoeff$. So $v_* \in X$.
+	\paragraph{Feasibility} To show that $v_*$ is feasible (i.e., $v_* \in X$), we must show that $v_* \in H^2_N(\Omega)$ and $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,\convrank$.
+We already showed that $v_* \in H^2_N(\Omega)$ in Theorem \ref{thm:wi_no_breakdown}.
+Using the the definition of $v_*$ in \eqref{eq:optimal_v} and Lemma \ref{lem:S_matrix_characterizations}, we have
+	\begin{equation*}
+		v_*(x_i) = \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \secondgreens_j(x_i)
+		= \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \pointinteractionmatrix_{ij} = \left(\pointinteractionmatrix \bm{\secondgreenscoeff} + \constcoeff \bm{1}\right)_i.
+	\end{equation*}
+	The conditions $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,\convrank$ may therefore be written in matrix form as
+	$
+		\pointinteractionmatrix \bm{\secondgreenscoeff} + \constcoeff \bm{1} = \bm{\interpolatedvalues}
+	$, and this matrix equation must hold because it is the first block row of system \eqref{eq:interpolation_linear_system} which is solved to compute $\bm{\secondgreenscoeff}$ and $\constcoeff$.
+So $v_* \in X$.
 

### 2021-09-20 numerical_examples/heat/heat_generate_results.py

@@ -292,7 +292,7 @@ for ii_reg in range(len(regularization_parameters)):
 plt.figure()
 plt.loglog(regularization_parameters, num_cg_iters_none)
 plt.loglog(regularization_parameters, num_cg_iters_reg)
-for kk_batch in all_batch_sizes:
+for kk_batch in range(len(all_batch_sizes)):
     plt.loglog(regularization_parameters, all_num_cg_iters_hmatrix[:, kk_batch])
 
 plt.title(r'Conjugate gradient iterations to achieve tolerance $10^{-6}$')

### 2021-11-23 paper/localpsf.tex
added minimum ellipsoid width functionality

@@
 \hypersetup{
-  pdftitle={Fast matrix-free approximation of smoothly varying blur operators, with application to Hessians in PDE-constrained inverse problems with highly informative data},
+  pdftitle={Fast matrix-free approximation of locally translation invariant operators that have locally supported non-negative integral kernels, with application to Hessians in PDE-constrained inverse problems},
   pdfauthor={N. Alger, N. Petra, T. Hartland, and O. Ghattas}
@@
 \begin{abstract}
-	We present an efficient matrix-free method for approximating locally translation invariant operators that have locally supported non-negative integral kernels.
-Such operators arise, for example, as Schur complements in Schur complements methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems.
-The method computes impulse responses of the operator at a collection of scattered points, then interpolates these computed impulse responses to form a product-convolution approximation of the operator.
-The product-convolution approximation is converted to hierarchical matrix format in order to solve linear systems and perform other matrix operations efficiently.
+	We present a fast matrix-free method for computing hierarchical matrix (H-matrix) approximations of locally translation invariant operators that have locally supported non-negative integral kernels.
+Such operators arise, for example, as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems.
+Converting an operator to H-matrix format is highly desirable, because fast scalable H-matrix methods can be used to add, multiply, invert, and factorize H-matrices, perform matrix-vector products, solve linear systems, and perform other useful matrix operations.
+The computational cost of these operations is $O(N \log^a N)$, where the matrix is $N\times N$, and $a \in \{1,2,3\}$.
+Classical H-matrix construction methods require evaluating the integral kernel $O(N \log N)$ times, but this is not feasible in large-scale problems when the operator is only accessible via matrix-vector products (``matrix-free'').
+We overcome this problem by computing impulse responses of the operator and its transpose at a collection of scattered points, then interpolating shifted versions of these impulse responses on-the-fly to evaluate the integral kernel rapidly.
 Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources.
@@
 Numerical results demonstrate that our method substantially outperforms existing state-of-the-art Hessian approximation methods which are based on low-rank approximation.
-Our method is able to form high quality approximations of high rank Hessians using only a small number of Hessian matrix-vector products.
+We are able to form high quality approximations of high rank Hessians using only a small number of Hessian matrix-vector products.
 \end{abstract}

### 2021-11-23 localpsf/ellipsoid.py

@@ -42,7 +42,7 @@ def K_fct_vectorized(ss, lambdas, v_squared, tau):
 def geigh_numpy(A, B):
     # Generalized eigenvalue decomposition A*Phi = B*Phi*Lambda
     # Phi.T * B * Phi = I
-    lambdas0, Phi0 = np.linalg.eig(np.linalg.solve(B, A))
+    lambdas0, Phi0 = np.linalg.eig(np.linalg.solve(B, A)+0.0*1j)
     lambdas = lambdas0.real
     Rayleigh_matrix = np.dot(np.dot(Phi0.real.T, B), Phi0.real)
     scaling_factors = np.sqrt(np.diag(Rayleigh_matrix))

### 2021-11-23 localpsf/product_convolution_hmatrix.py

@@ -50,7 +50,8 @@ class ImpulseResponseBatches:
                  tau=2.5,
                  max_candidate_points=None,
                  use_lumped_mass_matrix_for_impulse_response_moments=True,
-                 num_neighbors=10):
+                 num_neighbors=10,
+                 sigma_min=1e-6):
         me.V_in = V_in
         me.V_out = V_out
         me.apply_A = apply_A
@@ -72,7 +73,7 @@ class ImpulseResponseBatches:
             me.vol, me.mu, me.Sigma = impulse_response_moments(me.V_in, me.V_out, me.apply_At, me.solve_M_in)
 
         print('Preparing c++ object')
-        mesh_out = me.V_out.mesh();
+        mesh_out = me.V_out.mesh()
         mesh_vertices = np.array(mesh_out.coordinates().T, order='F')
         mesh_cells = np.array(mesh_out.cells().T, order='F')
 
@@ -89,7 +90,12 @@ class ImpulseResponseBatches:
         me.dof2vertex_out = dl.dof_to_vertex_map(me.V_out)
 
         me.mu_array = dlfct2array(me.mu)
-        me.Sigma_array = dlfct2array(me.Sigma)
+        me.Sigma_array0 = dlfct2array(me.Sigma)
+
+        eee0, PP = np.linalg.eigh(me.Sigma_array0)
+        eee = np.max([np.ones(eee0.shape)*sigma_min**2, eee0], axis=0)
+        me.Sigma_array = np.einsum('nij,nj,nkj->nik', PP, eee, PP)
+
 
         if me.max_candidate_points is None:
             me.candidate_inds = np.arange(me.V_in.dim())
@@ -216,7 +222,7 @@ class ProductConvolutionKernelRBF:
     def __init__(me, V_in, V_out, apply_A, apply_At, num_row_batches, num_col_batches,
                  tau_rows=2.5, tau_cols=2.5,
                  num_neighbors_rows=10, num_neighbors_cols=10,
-                 symmetric=False, gamma=1e-8):
+                 symmetric=False, gamma=1e-8, sigma_min=1e-6):
         me.V_in = V_in
         me.V_out = V_out
         me.apply_A = apply_A
@@ -225,7 +231,8 @@ class ProductConvolutionKernelRBF:
         me.col_batches = ImpulseResponseBatches(V_in, V_out, apply_A, apply_At,
                                                 num_initial_batches=num_col_batches,
                                                 tau=tau_cols,
-                                                num_neighbors=num_neighbors_cols)
[…]

### 2021-11-23 numerical_examples/heat/experiments_10_8_21.py

@@ -16,10 +16,11 @@ import scipy.sparse.linalg as spla
 
 nondefault_HIP_options = {'mesh_h': 2e-2} # {'mesh_h': 3e-2}
 
-num_batches = 20
+num_batches = 10
 num_neighbors = 10
 tau = 2.5 # 4
 gamma = 1e-5
+sigma_min = 1e-6 # 1e-1 # minimum width of support ellipsoid
 
 ########    SET UP HEAT INVERSE PROBLEM    ########
 
@@ -40,7 +41,8 @@ PCK = ProductConvolutionKernelRBF(HIP.V, HIP.V, HIP.apply_Hd_petsc, HIP.apply_Hd
                                   num_neighbors_rows=num_neighbors,
                                   num_neighbors_cols=num_neighbors,
                                   symmetric=True,
-                                  gamma=gamma)
+                                  gamma=gamma,
+                                  sigma_min=sigma_min)
 
 # PCK = ProductConvolutionKernelRBF(HIP.V, HIP.V, HIP.apply_Hd_petsc, HIP.apply_Hd_petsc,
 #                                   0, num_batches,

### 2021-12-17 localpsf/impulse_response_batches.py
ccgo talk slides

@@ -61,6 +61,7 @@ class ImpulseResponseBatches:
         me.Sigma = np.einsum('nij,nj,nkj->nik', PP, eee, PP)
 
         me.point_batches = list()
+        me.vol_batches = list()
         me.mu_batches = list()
         me.Sigma_batches = list()
         me.phi_batches = list()
@@ -90,6 +91,7 @@ class ImpulseResponseBatches:
                                                  candidate_inds_ordered_by_distance, randomize=False)
 
         new_points = me.dof_coords_in[new_inds, :]
+        new_vol = me.vol[new_inds]
         new_mu = me.mu[new_inds, :]
         new_Sigma = me.Sigma[new_inds, :, :]
 
@@ -104,6 +106,7 @@ class ImpulseResponseBatches:
                                 True)
 
         me.point_batches.append(new_points)
+        me.vol_batches.append(new_vol)
         me.mu_batches.append(new_mu)
         me.Sigma_batches.append(new_Sigma)
         me.phi_batches.append(phi_vertex)

### 2022-02-09 paper/localpsf.tex
mass matrix causing negative numbers. Lumped mass no problems. Possible scaling or conditioning issue with big domain

@@
 
+\begin{itemize}
+	\item Stokes Krylov convergence (10 solves for nonlinear forward + 6 hessian matvecs for ellipsoid estimation + 10 matvecs for impulse responses + 30 matvecs for PCH6 CG solve + 100 matvecs for Reg CG solve)
+	\item Stokes error vs num batches (small number of batches; 1, 3, 6, 9 batches?) (50 matvecs for randomized error estimation +)
+	\item Velocity field
+	\item Basal friction field
+	\item Picture of impulse responses
+	\item Mesh negative numbers picture
+\end{itemize}
+
 Some numerical results here.
@@
 \section*{Acknowledgments}
-We thank J.J. Alger, Longfei Gao, Mathew Hu, and Rami Nammour for helpful discussions.
+We thank J.J. Alger, Longfei Gao, Mathew Hu, and Rami Nammour for helpful discussions. Georg Stadler (domain geometry help)
 

### 2022-02-09 localpsf/impulse_response_batches.py

@@ -18,13 +18,15 @@ class ImpulseResponseBatches:
                  max_candidate_points=None,
                  use_lumped_mass_matrix_for_impulse_response_moments=True,
                  num_neighbors=10,
-                 sigma_min=1e-6):
+                 sigma_min=1e-6,
+                 max_scale_discrepancy=1e2):
         me.V_in = V_in
         me.V_out = V_out
         me.apply_A = apply_A
         me.apply_At = apply_At
         me.max_candidate_points = max_candidate_points
         me.use_lumped_mass_matrix_for_impulse_response_moments = use_lumped_mass_matrix_for_impulse_response_moments
+        me.max_scale_discrepancy = max_scale_discrepancy
 
         print('Making mass matrices and solvers')
         me.M_in, me.solve_M_in = make_mass_matrix(me.V_in, make_solver=True)
@@ -95,11 +97,21 @@ class ImpulseResponseBatches:
         new_mu = me.mu[new_inds, :]
         new_Sigma = me.Sigma[new_inds, :, :]
 
-        phi = get_one_dirac_comb_response(new_points, me.V_in, me.V_out, me.apply_A, me.solve_M_in, me.solve_M_out)
+        min_vol = np.max(me.vol) / me.max_scale_discrepancy
+        scale_factors = 1./new_vol.copy()
+        scale_factors[new_vol < min_vol] = 0.0
+
+        # new_vol_min = min_vol*np.ones(new_vol.shape)
+        # scale_factors = 1./np.max([new_vol, new_vol_min], axis=0)
+
+        # phi = get_one_dirac_comb_response(new_points, me.V_in, me.V_out, me.apply_A, me.solve_M_in, me.solve_M_out, scale_factors=scale_factors)
+        phi = get_one_dirac_comb_response(new_points, me.V_in, me.V_out, me.apply_A, me.solve_ML_in, me.solve_ML_out,
+                                          scale_factors=scale_factors)
 
         phi_vertex = phi.vector()[me.vertex2dof_out].copy()
 
         me.cpp_object.add_batch(list(new_points),
+                                list(new_vol),
                                 list(new_mu),
                                 list(new_Sigma),
                                 phi_vertex,
@@ -196,17 +208,20 @@ class ImpulseResponseBatches:
         return me.cpp_object.num_batches()
 
 
[…]

### 2022-02-09 localpsf/product_convolution_kernel.py

@@ -9,17 +9,20 @@ class ProductConvolutionKernel:
     def __init__(me, V_in, V_out, apply_A, apply_At, num_row_batches, num_col_batches,
                  tau_rows=2.5, tau_cols=2.5,
                  num_neighbors_rows=10, num_neighbors_cols=10,
-                 symmetric=False, gamma=1e-8, sigma_min=1e-6):
+                 symmetric=False, gamma=1e-8, sigma_min=1e-6,
+                 max_scale_discrepancy=1e2):
         me.V_in = V_in
         me.V_out = V_out
         me.apply_A = apply_A
         me.apply_At = apply_At
+        me.max_scale_discrepancy = max_scale_discrepancy
 
         me.col_batches = ImpulseResponseBatches(V_in, V_out, apply_A, apply_At,
                                                 num_initial_batches=num_col_batches,
                                                 tau=tau_cols,
                                                 num_neighbors=num_neighbors_cols,
-                                                sigma_min=sigma_min)
+                                                sigma_min=sigma_min,
+                                                max_scale_discrepancy=max_scale_discrepancy)
 
         if symmetric:
             me.row_batches = me.col_batches
@@ -28,7 +31,8 @@ class ProductConvolutionKernel:
                                                     num_initial_batches=num_row_batches,
                                                     tau=tau_rows,
                                                     num_neighbors=num_neighbors_rows,
-                                                    sigma_min=sigma_min)
+                                                    sigma_min=sigma_min,
+                                                    max_scale_discrepancy=max_scale_discrepancy)
 
         me.col_coords = me.col_batches.dof_coords_in
         me.row_coords = me.col_batches.dof_coords_out

### 2022-02-22 paper/localpsf.tex
updating paper based on advances the last couple months. Work in progress...

@@
 \begin{abstract}
-	We present a fast matrix-free method for computing hierarchical matrix (H-matrix) approximations of locally translation invariant operators that have locally supported non-negative integral kernels.
-Such operators arise, for example, as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems.
-Converting an operator to H-matrix format is highly desirable, because fast scalable H-matrix methods can be used to add, multiply, invert, and factorize H-matrices, perform matrix-vector products, solve linear systems, and perform other useful matrix operations.
-The computational cost of these operations is $O(N \log^a N)$, where the matrix is $N\times N$, and $a \in \{1,2,3\}$.
-Classical H-matrix construction methods require evaluating the integral kernel $O(N \log N)$ times, but this is not feasible in large-scale problems when the operator is only accessible via matrix-vector products (``matrix-free'').
-We overcome this problem by computing impulse responses of the operator and its transpose at a collection of scattered points, then interpolating shifted versions of these impulse responses on-the-fly to evaluate the integral kernel rapidly.
+	We present a fast matrix-free method for computing hierarchical matrix (H-matrix) approximations of operators that have locally supported non-negative impulse responses.
+Such operators arise, for example, as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained inverse problems.
+	H-matrix approximations are highly desirable because fast H-matrix methods can add, multiply, invert, and factorize H-matrices, compute matrix-vector products, solve linear systems, and perform other useful matrix computations.
+%	The computational cost of these operations is $O(N \log^a N)$, where the matrix is $N\times N$, and $a \in \{1,2,3\}$. 
+	Classical H-matrix construction methods 
+%	require evaluating the integral kernel $O(N \log N)$ times, but this is not feasible 
+	are infeasible for large-scale problems in which the operator is only accessible via matrix-vector products (``matrix-free''), because these methods require evaluating the integral kernel of the operator at a large number of points.
+We overcome this problem by computing impulse responses of the operator and its transpose at a collection of scattered points, then interpolating these known impulse responses to evaluate entries of the integral kernel at other points where the impulse response is not known.
+%	We improve upon classical product-convolution approximations by translating with respect the estimated mean of the impulse response, rather than the location of the impulse response point source. 
 	Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources.
-The key innovation of our method is a matrix-free procedure for choosing as many points as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
+We construct a-priori ellipsoid estimates of the supports of the impulse responses (i.e., support ellipsoids are estimated \emph{before} any impulse responses have been computed).
+By solving an ellipsoid packing problem, we choose as many points as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
 We apply the method to approximate Hessians in large-scale PDE-constrained inverse problems with highly informative data.
 Numerical results demonstrate that our method substantially outperforms existing state-of-the-art Hessian approximation methods which are based on low-rank approximation.
-We are able to form high quality approximations of high rank Hessians using only a small number of Hessian matrix-vector products.
+We are able to form high quality approximations of high rank operators using only a small number of operator-vector products.
 \end{abstract}
@@
 \begin{keywords}
-  example, \LaTeX
+
 \end{keywords}
@@
 \begin{AMS}
-  ASDF
+
 \end{AMS}
@@
 
-Let $\Omega \subset \mathbb{R}^d$ be a bounded domain in dimension $\gdim=1$, $2$, or $3$. We consider integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ of the form
+Let $\Omega \subset \mathbb{R}^d$ be a bounded domain (in applications typically $\gdim=1$, $2$, or $3$). We consider integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ of the form
 \begin{equation}
@@
 Here $\Aop u \in L^2(\Omega)'$ is the linear functional that results from applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
-We further narrow our focus to operators which possess the following four properties:
-\begin{description}
-\item[1) Matrix-free:] One cannot easily evaluate kernel entries $\Aker(y,x)$.
+Further, let $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$,
+\begin{equation}
+\label{eq:impulse_response_defn}
+\impulseresponse_x(y) := \Aker(y, x),
+\end{equation} 
+be the \emph{impulse response} of $\Aop$ at $x$. The function $\impulseresponse_x$ is called the impulse response because straightforward analysis shows that it may be expressed as follows:
+\begin{equation}
+\label{eq:impulse_response_delta_action}
+\impulseresponse_x = \left( \Aop \delta_x^* \right)^*,
+\end{equation}
+where $\delta_x$ is the the point source (delta distribution) centered at $x$, $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying the operator $\Aop$ to $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\sigma \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\sigma^* \in L^2(\Omega)$ such that $\sigma(v) = \left(\sigma^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is formally defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
+This is described in Appendix \ref{app:distributions}.
+
+We seek to approximate operators $\Aop$ with locally supported, non-negative impulse responses that are accessible matrix-free. 
+\begin{itemize}
+	\item $\Aop$ has \textbf{locally supported impulse responses} if, for all $x\in \Omega$, the support of $\impulseresponse_x$ is contained (or approximately contained) in a small region.
+	\item  $\Aop$ has \textbf{non-negative impulse responses} if
+	\begin{equation*}
+	\impulseresponse_x(y) \ge 0
+	\end{equation*}
+	for all (or approximately all) $(y,x) \in \Omega \times \Omega$.
+	\item $\Aop$ is accessible \textbf{matrix-free} if one cannot easily evaluate kernel entries $\Aker(y,x)$.
 Instead, access to $\Aop$ is available through ``operator actions.'' That is, we have a black-box computational procedure (which is typically expensive) through which we may evaluate the maps
@@
 In applications, applying $\bm{A}$ to a vector may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
-For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
-\item[2) Local support:] Let $\impulseresponse_x$ be the following \emph{impulse response}:
-\begin{equation}
-\label{eq:impulse_response_defn}
-\impulseresponse_x(y) := \Aker(y, x).
-\end{equation} 
-We say that $\Aop$ has local support if, for all $x\in \Omega$, the support of $\impulseresponse_x$ is contained (or approximately contained) in a neighborhood of $x$.
-The smaller these supports are, the better our algorithm will perform.
+	%For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
+\end{itemize}
+
+% approximate operators which possess the following properties:
+% 
+%We further limit our focus to operators that have
+%\textbf{Local support}
+%We say that $\Aop$ has local support if, for all $x\in \Omega$, the support of $\impulseresponse_x$ is contained (or approximately contained) in a small region. 
+%The smaller these supports are, the better our algorithm will perform.
 Let $\delta_x$ denote the delta distribution (point source) centered at $x$.
-The function $\impulseresponse_x$ is called the impulse response because straightforward analysis shows that it may be expressed as follows:
-\begin{equation}
-	\label{eq:impulse_response_delta_action}
-	\impulseresponse_x = \left( \Aop \delta_x^* \right)^*,
-\end{equation}
-where $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying the operator $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\sigma \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\sigma^* \in L^2(\Omega)$ such that $\sigma(v) = \left(\sigma^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
-This is described in Appendix \ref{app:distributions}.
 From \eqref{eq:impulse_response_delta_action}, we see that this local support property means that the response of $\Aop$ to a point source at $x$ is zero (or small) at points $y$ that are far from $x$.
-\item[3) Local translation invariance:] We say that $\Aop$ is locally translation invariant if
-\begin{equation}
-	\label{eq:local_translation_invariance}
-	\Aker(y+h, x+h) \approx \Aker(y,x)
-\end{equation}
-when $h$ is not too large.
+%\item[3) Local translation invariance:] We say that $\Aop$ is locally translation invariant if
+%\begin{equation}
+%	\label{eq:local_translation_invariance}
+%	\Aker(y+h, x+h) \approx \Aker(y,x)
+%\end{equation}
+%when $h$ is not too large.
 One can imagine the kernel entry $\Aker(y,x)$ as representing the strength of an interaction between a source at point $x$ and a target at point $y$.
@@
 This is illustrated in Figure FIG.
-\item[4) Non-negative kernel:] For all $(y,x) \in \Omega \times \Omega$, we have
-\begin{equation*}
-\Aker(y,x) \ge 0.
-\end{equation*}
-\end{description}
-
-Our overall strategy is to form a product-convolution approximation, $\AopPc$, of $\Aop$ using only operator actions, then convert a discretized version of $\AopPc$ into a hierarchical matrix (H-matrix), which we denote by $\AmatPc$.
-Schematically,
-\begin{equation*}
-	\Aop \longrightarrow \underbrace{\AopPc}_{\substack{\text{product}\\ \text{convolution}}} \longrightarrow \underbrace{\AmatPc}_{\text{H-matrix}}.
+%\textbf{Non-negative impulse response:} We have
+%\begin{equation*}
+%\impulseresponse_x(y) \ge 0
+%\end{equation*}
+%for all $(y,x) \in \Omega \times \Omega$.
+
+\subsection{Idea of the method}
+
+Because of the non-negative impulse response property, $\impulseresponse_x$ may be interpreted as a scaled probability distribution.
+Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor (the ``volume'' of $\impulseresponse_x$), let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
+In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ \emph{for all $x \in \Omega$ simultaneously}, via a procedure that involves applying $\Aop^T$ to a small number of specially chosen functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).
+Knowing $V(x)$, $\mu(x)$, and $\Sigma(x)$ allows to form an ellipsoid shaped estimate for the support of $\phi_x$.
+We can predict where $\phi_x$ will be nonzero \emph{before} we compute it.
+
+Our overall strategy is to compute impulse responses $\{\impulseresponse_{x_i}\}_{i=1}^r$ of $\Aop$ to point sources at a collection of points $x_i$ using a small number of operator actions of $\Aop$.
+If $\Aop$ is non-symmetric, we also compute impulse responses $\{\impulseresponse'_{y_i}\}_{i=1}^r$ of $\Aop^T$ to point sources at a collection of points $y_i$.
+We then form a H-matrix approximation, $\boldsymbol{\Phi}_H$, of the kernel, $\Phi$, and use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation of $\Aop$, denoted $\mathbf{A}_H$.
+Forming the H-matrix approximation $\boldsymbol{\Phi}_H$ requires evaluating kernel entries, $\Phi(y,x)$, at arbitrary points $(y,x)$, and this is done by interpolating the known impulse responses $\{\impulseresponse_{x_i}\}_{i=1}^r, \{\impulseresponse'_{y_i}\}_{i=1}^r$.
+The details of the interpolation will be described in detail later, but underlying idea behind the interpolation is that
+$$\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))$$
+if $x_i$ is close to $x$.
+This is shown in Figure FIG.
+Schematically the overall framework proceeds as follows:
+\begin{equation*}
+	\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \{\impulseresponse_{x_i}\}_{i=1}^r, \{\impulseresponse'_{y_i}\}_{i=1}^r \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
 \end{equation*}
@@
 For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high-rank approximations of an operator using a small number of matrix-vector products.
+
+\subsection{Mean-displacement invariance vs. translation invariance}
+
+This is similar to the classical local translation invariance assumption, $\phi_{x_i}(y - x + x_i)$, used in product-convolution approximations.
 

### 2022-02-23 paper/localpsf.tex
updating paper based on advances the last couple months. Work in progress...

@@
 Here $\Aop u \in L^2(\Omega)'$ is the linear functional that results from applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
+
 Further, let $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$,
@@
 
-We seek to approximate operators $\Aop$ with locally supported, non-negative impulse responses that are accessible matrix-free. 
-\begin{itemize}
-	\item $\Aop$ has \textbf{locally supported impulse responses} if, for all $x\in \Omega$, the support of $\impulseresponse_x$ is contained (or approximately contained) in a small region.
-	\item  $\Aop$ has \textbf{non-negative impulse responses} if
+In this paper we form hierarchical matrix approximations of operators $\Aop$ with locally supported, non-negative impulse responses that are accessible matrix-free.
+\begin{description}
+	\item[Locally supported impulse responses:] For all $x\in \Omega$, the support of $\impulseresponse_x$ is contained (or approximately contained) in a localized region.
+	\item[Non-negative impulse responses:] We have
 	\begin{equation*}
@@
 	for all (or approximately all) $(y,x) \in \Omega \times \Omega$.
-	\item $\Aop$ is accessible \textbf{matrix-free} if one cannot easily evaluate kernel entries $\Aker(y,x)$.
+	\item[Matrix-free:] One cannot easily evaluate kernel entries $\Aker(y,x)$.
 Instead, access to $\Aop$ is available through ``operator actions.'' That is, we have a black-box computational procedure (which is typically expensive) through which we may evaluate the maps
@@
 	%For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
-\end{itemize}
+\end{description}
 
@@
 
-\subsection{Idea of the method}
-
-Because of the non-negative impulse response property, $\impulseresponse_x$ may be interpreted as a scaled probability distribution.
-Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor (the ``volume'' of $\impulseresponse_x$), let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
-In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ \emph{for all $x \in \Omega$ simultaneously}, via a procedure that involves applying $\Aop^T$ to a small number of specially chosen functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).
-Knowing $V(x)$, $\mu(x)$, and $\Sigma(x)$ allows to form an ellipsoid shaped estimate for the support of $\phi_x$.
-We can predict where $\phi_x$ will be nonzero \emph{before} we compute it.
-
-Our overall strategy is to compute impulse responses $\{\impulseresponse_{x_i}\}_{i=1}^r$ of $\Aop$ to point sources at a collection of points $x_i$ using a small number of operator actions of $\Aop$.
-If $\Aop$ is non-symmetric, we also compute impulse responses $\{\impulseresponse'_{y_i}\}_{i=1}^r$ of $\Aop^T$ to point sources at a collection of points $y_i$.
-We then form a H-matrix approximation, $\boldsymbol{\Phi}_H$, of the kernel, $\Phi$, and use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation of $\Aop$, denoted $\mathbf{A}_H$.
-Forming the H-matrix approximation $\boldsymbol{\Phi}_H$ requires evaluating kernel entries, $\Phi(y,x)$, at arbitrary points $(y,x)$, and this is done by interpolating the known impulse responses $\{\impulseresponse_{x_i}\}_{i=1}^r, \{\impulseresponse'_{y_i}\}_{i=1}^r$.
-The details of the interpolation will be described in detail later, but underlying idea behind the interpolation is that
-$$\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))$$
-if $x_i$ is close to $x$.
-This is shown in Figure FIG.
-Schematically the overall framework proceeds as follows:
-\begin{equation*}
-	\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \{\impulseresponse_{x_i}\}_{i=1}^r, \{\impulseresponse'_{y_i}\}_{i=1}^r \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
-\end{equation*}
-We will define product-convolution approximations in Section \ref{sec:prodconv_intro}.
-H-matrices are discussed in Appendix \ref{app:h_matrix}.
-Our method is summarized in Section \ref{sec:overview_intro}, and described in detail in Section \ref{sec:method}.
-
-The novel contributions of our method are in the way that we form the product-convolution approximation ($\Aop \rightarrow \AopPc$).
-This is the part of the method where we use the four properties of $\Aop$ described above.
-Fast conversion of the discretized version of $\AopPc$ to H-matrix format ($\AopPc \rightarrow \AmatPc$) is possible because integral kernel entries for $\AopPc$ are easily accessible (unlike integral kernel entries for $\Aop$).
-Once in H-matrix format, one may use H-matrix methods to perform further useful linear algebra operations involving $\AmatPc$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
-These H-matrix methods are fast and scale well to large problems.
-
 \subsection{Motivation: inverse problems with highly informative data}
@@
 For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high-rank approximations of an operator using a small number of matrix-vector products.
+
+\subsection{Idea of the method}
+Each impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative impulse response property.
+Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor (the ``volume'' of $\impulseresponse_x$), let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
+In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of polynomial functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).
+Since a probability distribution concentrates within a small number of standard deviations of its mean, knowing $\mu(x)$ and $\Sigma(x)$ permits us to form an ellipsoid shaped a-priori estimate for the support of $\phi_x$.
+In other words, we can predict an ellipsoid outside of which $\phi_x$ will be (nearly) zero, \emph{without} computing $\phi_x$.
+
+This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$ in ``batches.'' We compute one batch of $\impulseresponse_{x_i}$'s by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$.
+The batch of points $x_i$ are chosen via a greedy ellipsoid packing algorithm so that the support of $\impulseresponse_{x_i}$ and the support of $\impulseresponse_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$.
+Because these supports do not overlap, we can post-process the response of $\Aop$ to the Dirac comb to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i$ in the batch---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$.
+This is illustrated in Figure FIG.
+The process is repeated to get more batches, until a desired number of impulse responses is reached.
+
+Once the impulse responses $\{\impulseresponse_{x_i}\}_{i=1}^r$ are computed, we may approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation.
+The details of the interpolation will be described in detail later, but the key idea is that
+$$\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))$$
+if $x_i$ is close to $x$.
+This is shown in Figure FIG.
+We therefore approximate $\phi_x(y)$ by a weighted linear combination of the values $\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for the $k$-nearest points $x_i$ to $x$.
+Here $k$ is a user-defined parameter; we use $k=10$.
+The weights in the linear combination are determined via radial basis function interpolation.
+
+The ability to rapidly evaluate approximations to arbitrary kernel entries $\Phi(y,x)$ allows us to construct an H-matrix representation of $\Phi$, which we denote $\boldsymbol{\Phi}_H$, using the conventional adaptive cross H-matrix construction method.
+In this method, one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
+We then use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation of $\Aop$, denoted $\mathbf{A}_H$.
+Schematically the overall framework proceeds as follows:
+\begin{equation*}
+\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \{\impulseresponse_{x_i}\}_{i=1}^r \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
+\end{equation*}
+Once in H-matrix format, one may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
+These H-matrix methods are fast and scale well to large problems.
+
+Often we are interested in approximating operators $\Aop$ that are symmetric positive semi-definite.
+Unfortunately, the approximation is, in general, non-symmetric and indefinite due to errors in the approximation.
+In this case, we modify the H-matrix representation of $\mathbf{A}_H$ to make it symmetric positive definite, by symmetrizing it, then applying a specially chosen rational matrix function to the symmetrized H-matrix.
+
+%Forming the H-matrix approximation $\boldsymbol{\Phi}_H$ requires evaluating kernel entries, $\Phi(y,x)$, at arbitrary points $(y,x)$, and this is done by interpolating the known impulse responses $\{\impulseresponse_{x_i}\}_{i=1}^r, \{\impulseresponse'_{y_i}\}_{i=1}^r$.
+%
+%%For each batch, we solve an ellipsoid packing problem to choose as many points $x_i$ as possible, while ensuring that the supports of the impulse responses $\impulseresponse_{x_i}$ do not overlap for points $x_i$ in the batch.
+Then we apply $\Aop$ to a weighted sum of delta distributions centered at the points $x_i$ (Dirac comb).
+From the resulting
+%
+%
+%Our overall strategy is as follows.
+First, we compute the functions $V:\Omega \rightarrow \mathbb{R}$, $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$ by applying $\Aop^T$ to a small number of polynomial functions.
+Next, we solve ellipsoid packing problems to choose batches of points $x_i$ such that
+%to compute impulse responses $\{\impulseresponse_{x_i}\}_{i=1}^r$ of $\Aop$ to point sources at a collection of points $x_i$ using a small number of operator actions of $\Aop$.
+If $\Aop$ is non-symmetric, we also compute impulse responses $\{\impulseresponse'_{y_i}\}_{i=1}^r$ of $\Aop^T$ to point sources at a collection of points $y_i$.
+We then form a H-matrix approximation, $\boldsymbol{\Phi}_H$, of the kernel, $\Phi$, and use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation of $\Aop$, denoted $\mathbf{A}_H$.
+%Schematically the overall framework proceeds as follows:
+%\begin{equation*}
+%\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \{\impulseresponse_{x_i}\}_{i=1}^r \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
+%\end{equation*}
+%We will define product-convolution approximations in Section \ref{sec:prodconv_intro}.
+H-matrices are discussed in Appendix \ref{app:h_matrix}.
+Our method is summarized in Section \ref{sec:overview_intro}, and described in detail in Section \ref{sec:method}.
+%
+%The novel contributions of our method are in the way that we form the product-convolution approximation ($\Aop \rightarrow \AopPc$).
+This is the part of the method where we use the four properties of $\Aop$ described above.
+Fast conversion of the discretized version of $\AopPc$ to H-matrix format ($\AopPc \rightarrow \AmatPc$) is possible because integral kernel entries for $\AopPc$ are easily accessible (unlike integral kernel entries for $\Aop$).
+Once in H-matrix format, one may use H-matrix methods to perform further useful linear algebra operations involving $\AmatPc$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
+These H-matrix methods are fast and scale well to large problems.
 

### 2022-02-23 paper/localpsf.tex
updating paper based on advances the last couple months. Work in progress...

@@
 
-\subsection{Idea of the method}
+\subsection{Overview of the method}
 Each impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative impulse response property.
@@
 
-\subsection{Overview of the method}
+\subsection{Overview of the method (old)}
 \label{sec:overview_intro}

### 2022-02-23 paper/localpsf.tex
updating paper based on advances the last couple months. Work in progress...

@@
 \subsection{Overview of the method}
-Each impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative impulse response property.
+The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative impulse response property.
 Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor (the ``volume'' of $\impulseresponse_x$), let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
 In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of polynomial functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).
-Since a probability distribution concentrates within a small number of standard deviations of its mean, knowing $\mu(x)$ and $\Sigma(x)$ permits us to form an ellipsoid shaped a-priori estimate for the support of $\phi_x$.
-In other words, we can predict an ellipsoid outside of which $\phi_x$ will be (nearly) zero, \emph{without} computing $\phi_x$.
-
-This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$ in ``batches.'' We compute one batch of $\impulseresponse_{x_i}$'s by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$.
+Since a probability distribution concentrates within a small number of standard deviations of its mean, knowing $\mu(x)$ and $\Sigma(x)$ permits us to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
+%That is, we can predict an ellipsoid outside of which $\phi_x$ will be (nearly) zero, \emph{without} computing $\phi_x$.
+
+This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches.'' We compute one batch of $\impulseresponse_{x_i}$'s by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$.
 The batch of points $x_i$ are chosen via a greedy ellipsoid packing algorithm so that the support of $\impulseresponse_{x_i}$ and the support of $\impulseresponse_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$.
@@
 
-Once the impulse responses $\{\impulseresponse_{x_i}\}_{i=1}^r$ are computed, we may approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation.
+Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation.
 The details of the interpolation will be described in detail later, but the key idea is that
-$$\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))$$
+\begin{equation}
+\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))
+\end{equation}
 if $x_i$ is close to $x$.
@@
 
-The ability to rapidly evaluate approximations to arbitrary kernel entries $\Phi(y,x)$ allows us to construct an H-matrix representation of $\Phi$, which we denote $\boldsymbol{\Phi}_H$, using the conventional adaptive cross H-matrix construction method.
+The ability to rapidly evaluate approximate kernel entries $\Phi(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Phi}_H$, of $\Phi$, using the conventional adaptive cross H-matrix construction method.
 In this method, one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
@@
 \end{equation*}
-Once in H-matrix format, one may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
+One may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
 These H-matrix methods are fast and scale well to large problems.
 
-Often we are interested in approximating operators $\Aop$ that are symmetric positive semi-definite.
-Unfortunately, the approximation is, in general, non-symmetric and indefinite due to errors in the approximation.
-In this case, we modify the H-matrix representation of $\mathbf{A}_H$ to make it symmetric positive definite, by symmetrizing it, then applying a specially chosen rational matrix function to the symmetrized H-matrix.
+Often we are interested in approximating operators $\Aop$ that are symmetric, or symmetric positive semi-definite.
+If $\Aop$ is symmetric, impulse responses of $\Aop$ are also impulse responses of $\Aop^T$, which may be interpolated in the variable $y$ instead of $x$ to yield a different approximation of $\Phi(y,x)$.
+In this case, we combine information from impulse responses of $\Aop$ and $\Aop^T$ in the interpolation procedure for approximating $\Phi(y,x)$, to get a more accurate approximation at negligible additional cost.
+Intuitively, impulse responses of $\Aop$ can be thought of as ``columns'' of $\Phi$, impulse responses of $\Aop^T$ can be thought of as ``rows'' of $\Phi$, and $\Phi(y,x)$ can be thought of as an unknown entry of $\Phi$.
+When $\Aop$ is symmetric, computing a column of $\Aop$ also gives us a row of $\Aop$ for free, so we approximate $\Phi(y,x)$ using information from a combination of known columns and known rows, rather than just known columns.
+
+When $\Aop$ is symmetric positive semi-definite, $\mathbf{A}_H$ will generally be indefinite due to errors in the approximation.
+In this case, we modify the H-matrix representation of $\mathbf{A}_H$ to make it positive semi-definite, by applying a specially chosen rational matrix function to $\mathbf{A}_H$.
 
@@
 
-\subsection{Mean-displacement invariance vs. translation invariance}
+\subsection{Differences in interpolation}
+
+\begin{equation}
+	\phi_x(y) \approx \phi_{x_i}(y)
+\end{equation}
+
+\begin{equation}
+\phi_x(y) \approx \phi_{x_i}(y-x+x_i)
+\end{equation}
+
+\begin{equation}
+\phi_x(y) \approx \phi_{x_i}\left(y-\mu(x)+\mu(x_i)\right)
+\end{equation}
 

### 2022-02-24 paper/localpsf.tex
updating paper based on advances the last couple months. Work in progress...

@@
 
-\begin{equation}
-	\phi_x(y) \approx \phi_{x_i}(y)
-\end{equation}
+Consider the following approximations:
+\begin{align}
+	\phi_x(y) &\approx \phi_{x_i}(y) \label{eq:translate1}\\
+	\phi_x(y) &\approx \phi_{x_i}(y-x+x_i) \label{eq:translate2}\\
+	\phi_x(y) &\approx \phi_{x_i}\left(y-\mu(x)+\mu(x_i)\right). \label{eq:translate3}
+\end{align}
+
+First one: low rank. Second one: local translation invariance/product-convolution. Third one: our proposal.
+
+Approximation \eqref{eq:translate3} is an improvement upon \eqref{eq:translate1} and \eqref{eq:translate2} that generalizes them both.
+If \eqref{eq:translate1} holds, then $\mu(x) \approx \mu(x_i)$, and so \eqref{eq:translate3} holds.
+If \eqref{eq:translate2} holds, then $\mu(x_i)-\mu(x) \approx x_i - x$ because translating a distribution translates its mean, so again \eqref{eq:translate3} holds.
+
+the first one holds, then $\mu(x) = \mu(x_i)$, and so the third one holds. If the second one holds, then $\mu(x) - \mu(x_i) = x - x_i$, and so the third one holds.
+
+Notice that the shift of the
 

### 2022-02-24 paper/localpsf.tex
updating paper based on advances the last couple months. Work in progress...

@@
 Approximation \eqref{eq:translate3} is an improvement upon \eqref{eq:translate1} and \eqref{eq:translate2} that generalizes them both.
-If \eqref{eq:translate1} holds, then $\mu(x) \approx \mu(x_i)$, and so \eqref{eq:translate3} holds.
-If \eqref{eq:translate2} holds, then $\mu(x_i)-\mu(x) \approx x_i - x$ because translating a distribution translates its mean, so again \eqref{eq:translate3} holds.
+On one hand, if \eqref{eq:translate1} holds, then $\mu(x) \approx \mu(x_i)$, and so \eqref{eq:translate3} holds.
+On the other hand, translating a distribution translates its mean, so if \eqref{eq:translate2} holds, then $\mu(x_i)-\mu(x) \approx x_i - x$, so again \eqref{eq:translate3} holds.
 

### 2022-02-24 paper/localpsf.tex
updating paper based on advances the last couple months. Work in progress...

@@
 The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative impulse response property.
-Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor (the ``volume'' of $\impulseresponse_x$), let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
+Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
+\begin{equation*}
+\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy
+\end{equation*}
+denote the spatially varying scaling factor (the ``volume'' of $\impulseresponse_x$), let 
+\begin{equation*}
+\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)
+\end{equation*}
+denote the normalized version of $\impulseresponse_x$, and let $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d},$
+\begin{equation*}
+\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
+\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
+\end{equation*}
+denote the spatially varying mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
 In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of polynomial functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).
+
 Since a probability distribution concentrates within a small number of standard deviations of its mean, knowing $\mu(x)$ and $\Sigma(x)$ permits us to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
-%That is, we can predict an ellipsoid outside of which $\phi_x$ will be (nearly) zero, \emph{without} computing $\phi_x$.
-
 This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches.'' We compute one batch of $\impulseresponse_{x_i}$'s by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$.
@@
 We then use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation of $\Aop$, denoted $\mathbf{A}_H$.
-Schematically the overall framework proceeds as follows:
+Schematically the overall H-matrix approximation proceeds as follows:
 \begin{equation*}
@@
 If $\Aop$ is symmetric, impulse responses of $\Aop$ are also impulse responses of $\Aop^T$, which may be interpolated in the variable $y$ instead of $x$ to yield a different approximation of $\Phi(y,x)$.
-In this case, we combine information from impulse responses of $\Aop$ and $\Aop^T$ in the interpolation procedure for approximating $\Phi(y,x)$, to get a more accurate approximation at negligible additional cost.
+In this case, we combine information from impulse responses of both $\Aop$ and $\Aop^T$ in the interpolation procedure for approximating $\Phi(y,x)$, to get a more accurate approximation at negligible additional cost.
 Intuitively, impulse responses of $\Aop$ can be thought of as ``columns'' of $\Phi$, impulse responses of $\Aop^T$ can be thought of as ``rows'' of $\Phi$, and $\Phi(y,x)$ can be thought of as an unknown entry of $\Phi$.
@@
 
-When $\Aop$ is symmetric positive semi-definite, $\mathbf{A}_H$ will generally be indefinite due to errors in the approximation.
+When $\Aop$ is symmetric positive semi-definite, $\mathbf{A}_H$ may be indefinite due to errors in the approximation.
 In this case, we modify the H-matrix representation of $\mathbf{A}_H$ to make it positive semi-definite, by applying a specially chosen rational matrix function to $\mathbf{A}_H$.
@@
 
-\subsection{Differences in interpolation}
+\subsection{Local mean-displacement invariance}
 
@@
 \end{align}
-
-First one: low rank. Second one: local translation invariance/product-convolution. Third one: our proposal.
-
-Approximation \eqref{eq:translate3} is an improvement upon \eqref{eq:translate1} and \eqref{eq:translate2} that generalizes them both.
+Operators with smooth integral kernels satisfy \eqref{eq:translate1}. These operators can be well-approximated by low-rank decompositions of the form
+\begin{equation}
+	\label{eq:low_rank_interpolation}
+	\Phi(y,x) \approx \sum_{i=1}^r w_i(x) \phi_{x_i}(y)
+\end{equation}
+where $w_i$ are smooth weighting functions that interpolate known impulse responses $\phi_{x_i}$.
+Unfortunately, the required rank, $r$, for these interpolative-type decompositions can be extremely large in practical applications, such as Hessians in inverse problems with highly informative data.
+Moreover, if $\Aop$ is numerically low rank, then one would typically use more accurate and robust methods matrix-free methods for forming low-rank approximations, such as randomized SVD or Lanczos.
+
+Operators that satisfy \eqref{eq:translate2} are called ``locally translation-invariant.'' Kernels for these operators can be well-approximated by decompositions of the form
+\begin{equation}
+	\Phi(y,x) \approx \sum_{i=1}^r w_i(x) \phi_{x_i}(y-x)
+\end{equation}
+which correspond to operator approximations of the form
+\begin{equation}
+\label{eq:product_convolution}
+\left(\Aop~u\right)(v) \approx \left( v,~ \sum_{i=1}^\convrank \convkernel_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
+\end{equation}
+where $f \ast g$ denotes convolution, and $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$ (i.e., $\left(f \cdot g\right)(x) = f(x) g(x)$).
+Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution.
+
+We use approximation \eqref{eq:translate3}, which is an improvement upon \eqref{eq:translate1} and \eqref{eq:translate2} that generalizes both.
 On one hand, if \eqref{eq:translate1} holds, then $\mu(x) \approx \mu(x_i)$, and so \eqref{eq:translate3} holds.
@@
 \begin{equation}
-\label{eq:product_convolution}
 	\left(\AopPc~u\right)(v) := \left( v,~ \sum_{i=1}^\convrank \convkernel_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},

### 2022-02-24 paper/localpsf.tex
local mean displacement section of paper written. reads well

@@
 
-Consider the following approximations:
+Let $x$ and $x'$ be points in $\Omega$ that are close to each other, and consider the following three approximations:
 \begin{align}
-	\phi_x(y) &\approx \phi_{x_i}(y) \label{eq:translate1}\\
-	\phi_x(y) &\approx \phi_{x_i}(y-x+x_i) \label{eq:translate2}\\
-	\phi_x(y) &\approx \phi_{x_i}\left(y-\mu(x)+\mu(x_i)\right). \label{eq:translate3}
+	\phi_x(y) &\approx \phi_{x'}(y) \label{eq:translate1}\\
+	\phi_x(y) &\approx \phi_{x'}(y-x+x') \label{eq:translate2}\\
+	\phi_x(y) &\approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right). \label{eq:translate3}
 \end{align}
-Operators with smooth integral kernels satisfy \eqref{eq:translate1}. These operators can be well-approximated by low-rank decompositions of the form
-\begin{equation}
-	\label{eq:low_rank_interpolation}
-	\Phi(y,x) \approx \sum_{i=1}^r w_i(x) \phi_{x_i}(y)
-\end{equation}
-where $w_i$ are smooth weighting functions that interpolate known impulse responses $\phi_{x_i}$.
-Unfortunately, the required rank, $r$, for these interpolative-type decompositions can be extremely large in practical applications, such as Hessians in inverse problems with highly informative data.
-Moreover, if $\Aop$ is numerically low rank, then one would typically use more accurate and robust methods matrix-free methods for forming low-rank approximations, such as randomized SVD or Lanczos.
-
-Operators that satisfy \eqref{eq:translate2} are called ``locally translation-invariant.'' Kernels for these operators can be well-approximated by decompositions of the form
-\begin{equation}
-	\Phi(y,x) \approx \sum_{i=1}^r w_i(x) \phi_{x_i}(y-x)
-\end{equation}
-which correspond to operator approximations of the form
-\begin{equation}
-\label{eq:product_convolution}
-\left(\Aop~u\right)(v) \approx \left( v,~ \sum_{i=1}^\convrank \convkernel_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
-\end{equation}
-where $f \ast g$ denotes convolution, and $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$ (i.e., $\left(f \cdot g\right)(x) = f(x) g(x)$).
+These approximations are illustrated in Figure FIG.
+
+Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank interpolative factorization of the kernel $\Phi$.
+% of the form
+%\begin{equation}
+%	\label{eq:low_rank_interpolation}
+%	\Phi(y,x) \approx \sum_{i=1}^r w_i(x) \phi_{x_i}(y)
+%\end{equation}
+%where $\phi_{x_i}$ are known impulse responses at a collection of points $x_i$ scattered throughout $\Omega$, and $w_i$ are smooth weighting functions that interpolate the $\phi_{x_i}$. 
+However, such approximations are not suitable if $\Aop$ is high rank.
+%, such as Hessians in inverse problems with highly informative data. 
+%Moreover, if $\Aop$ is numerically low rank, then one would typically use more accurate and robust matrix-free methods for forming low-rank approximations, such as randomized SVD or Lanczos.
+Operators that satisfy \eqref{eq:translate2} are called ``locally translation-invariant'' because integral kernel entries $\Phi(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$.
+Locally translation-invariant operators act like convolutions locally, and can therefore be well-approximated by so-called product-convolution approximations---weighted sums of convolution operators with spatially varying weights.
+Product-convolution approximations can be high-rank even if the number of convolution operators used in the approximation is small.
+
+% Kernels for locally translation invariant operators can be well-approximated by decompositions of the form
+%\begin{equation}
+%	\Phi(y,x) \approx \sum_{i=1}^r w_i(x) \phi_{x_i}(y-x)
+%\end{equation}
+%which correspond to operator approximations of the form
+%\begin{equation}
+%\label{eq:product_convolution}
+%\left(\Aop~u\right)(v) \approx \left( v,~ \sum_{i=1}^\convrank \convkernel_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)}.
+%\end{equation}
+%Here $f \ast g$ denotes convolution, and $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$ (i.e., $\left(f \cdot g\right)(x) = f(x) g(x)$).
 Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution.
+Product-convolution approximations can be high-rank even if $r$ is small.
 
@@
 On the other hand, translating a distribution translates its mean, so if \eqref{eq:translate2} holds, then $\mu(x_i)-\mu(x) \approx x_i - x$, so again \eqref{eq:translate3} holds.
-
-the first one holds, then $\mu(x) = \mu(x_i)$, and so the third one holds. If the second one holds, then $\mu(x) - \mu(x_i) = x - x_i$, and so the third one holds.
-
-Notice that the shift of the
-
-\begin{equation}
-\phi_x(y) \approx \phi_{x_i}(y-x+x_i)
-\end{equation}
-
-\begin{equation}
-\phi_x(y) \approx \phi_{x_i}\left(y-\mu(x)+\mu(x_i)\right)
-\end{equation}
-
-This is similar to the classical local translation invariance assumption, $\phi_{x_i}(y - x + x_i)$, used in product-convolution approximations.
+But approximation \eqref{eq:translate3} may continue to hold in situation where neither \eqref{eq:translate1} nor \eqref{eq:translate2} holds.
+For example, \eqref{eq:translate3} will hold when an operator is locally translation-invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case (see Figure FIG).
+Additionally, \eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions.
+In contrast, \eqref{eq:translate2} does not naturally generalize to operators that map between function spaces on different domains, because the formula $y-x+x'$ requires $y \in \Omega_2$ and $-x+x' \in \Omega_1$ to be added together.
+We call operators that satisfy \eqref{eq:translate3} ``locally mean-displacement invariant.''
 

### 2022-02-24 paper/localpsf.tex
local mean displacement section of paper written. reads well

@@
 \end{align}
-These approximations are illustrated in Figure FIG.
-
+Approximations \eqref{eq:translate1} and \eqref{eq:translate2} are standard. Approximation \eqref{eq:translate3}, which we use,  is new. These approximations are illustrated in Figure FIG.
+
+Approximation \eqref{eq:translate1} follows from a smoothness assumption on the kernel $\Phi$.
 Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank interpolative factorization of the kernel $\Phi$.

### 2022-02-25 paper/localpsf.tex
local mean displacement section of paper written. reads well

@@
 
-Approximation \eqref{eq:translate1} follows from a smoothness assumption on the kernel $\Phi$.
-Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank interpolative factorization of the kernel $\Phi$.
+Approximation \eqref{eq:translate1} follows from the assumption that $\Phi$ is smooth.
+Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank interpolative factorization of $\Phi$.
 % of the form
@@
 %where $\phi_{x_i}$ are known impulse responses at a collection of points $x_i$ scattered throughout $\Omega$, and $w_i$ are smooth weighting functions that interpolate the $\phi_{x_i}$. 
-However, such approximations are not suitable if $\Aop$ is high rank.
-%, such as Hessians in inverse problems with highly informative data. 
+%In problems of practical interest, \eqref{eq:translate1} typically holds \emph{asymptotically} in the limit as $x'$ approaches $x$. 
+However, even if $\Phi$ is smooth (which is typically the case asymptotically) the relevant Lipschitz constant for continuity, and hence the required rank in the low rank approximation, can be large, and this makes algorithms based on assumption \eqref{eq:translate1} highly inefficient.
+This is the case, for example, for Hessians in PDE constrained inverse problems with highly informative data.
+As the data become more informative, typically the ``mass'' of $\Phi(y,x)$ concentrates nearer and nearer to the ``diagonal'' $y=x$, the Lipschitz constant increases, and the required rank blows up.
 %Moreover, if $\Aop$ is numerically low rank, then one would typically use more accurate and robust matrix-free methods for forming low-rank approximations, such as randomized SVD or Lanczos.
+
 Operators that satisfy \eqref{eq:translate2} are called ``locally translation-invariant'' because integral kernel entries $\Phi(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$.
@@
 For example, \eqref{eq:translate3} will hold when an operator is locally translation-invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case (see Figure FIG).
-Additionally, \eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions.
+Additionally, \eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions.
 In contrast, \eqref{eq:translate2} does not naturally generalize to operators that map between function spaces on different domains, because the formula $y-x+x'$ requires $y \in \Omega_2$ and $-x+x' \in \Omega_1$ to be added together.

### 2022-02-28 paper/localpsf.tex
writing

@@
 Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation.
-The details of the interpolation will be described in detail later, but the key idea is that
+The details of the interpolation will be described in Section SECTION, but the key idea is that
 \begin{equation}
@@
 One may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
+WHY DO THESE OPERATIONS: KRYLOV SOLVERS, SAMPLING STRATEGIES, CAN'T DO WITH A.
 These H-matrix methods are fast and scale well to large problems.
@@
 
+\subsection{Approximating kernel entries $\Phi(y,x)$}
+
+We approximate kernel entries $\Phi(y,x)$ by preconditioned radial basis function interpolation of the impulse response values
+\begin{equation}
+\label{eq:phi_mu_shift32}
+	\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right),
+\end{equation}
+where the points $x_i$ are sample points at which $\phi_{x_i}$ is known.
+Interpolation using all sample points $x_i$ is too expensive, so we only use the $k$ nearest sample points to $x$, excluding points $x_i$ for which \eqref{eq:phi_mu_shift32} is not defined.
+That is, we exclude points $x_i$ such that $y - \mu(x) + \mu(x_i) \notin \Omega$.
+In numerical experiments we use $k=10$.
+The $k$ nearest sample points are found by querying a precomputed KD-tree [CITE] of all sample points.
+We check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) [CITE] of the mesh cells used to discretize the problem.
+
+Once the $k$ nearest sample points, $\{x_i\}_{i=1}^k$, are found, we approximate $\Phi(y,x)$ using preconditioned radial basis function interpolation of the following form:
+\begin{equation}
+	\Phi(y,x) \approx \sum_{i=1}^k w_i b\left(\|x-x_i\|\right).
+\end{equation}
+For radial basis functions, $b_i$, we use polyharmonic splines. That is, 
+\begin{equation}
+b_i(r) = r^d
+\end{equation}
+if $d$ is odd, and
+\begin{equation}
+	b_i(r) = \begin{cases}
+		r^d \log r, & r \neq 0 \\
+		0 & r = 0
+	\end{cases}
+\end{equation}
+if $d$ is even.
+The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
+\begin{equation}
+	Bw = f,
+\end{equation}
+where the coefficient matrix $B \in \mathbb{R}^{k \times k}$ has entries 
+\begin{equation*}
+	B_{ij} = b\left(\|x_i - x_j\|\right),
+\end{equation*}
+and the right hand side vector $f \in \mathbb{R}^k$ has entries
+\begin{equation}
+	f_i = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
+\end{equation}
+%The result of this interpolation is to approximate $\Phi(y,x)$ with a linear combination of the shifted impulse response values $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$. 
+By including the factor $\frac{V(x)}{V(x_i)}$, we are implicitly performing polyharmonic spline radial basis function interpolation of the ``volume preconditioned'' function $\frac{1}{V(x)}\Phi(y,x)$, then rescaling the result by $V(x)$ to get an approximation of $\Phi(y,x)$.
+Volume preconditioning improves the approximation considerably when $V(x)$ varies widely in magnitude, as is commonly the case for many operators of practical interest.
+In the Hessian for the Stokes inverse problem, for example, $V(x)$ varies over six orders of magnitude.
+
+The point $y - \mu(x) + \mu(x_i)$ typically is not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints.
+Although it is advisable to avoid evaluating finite element functions at arbitrary points [CITE], here we are forced to do this.
+We evaluate $\frac{1}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ via the following process.
+First, we determine whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ by checking if
+\begin{equation}
+\label{eq:ellipsoid_check93}
+	(y - \mu(x))^T \Sigma^{-1}(x) (y - \mu(x))  \le \tau^2.
+\end{equation}
+If \eqref{eq:ellipsoid_check93} does not hold, then $y - \mu(x) + \mu(x_i) \notin E_{x_i}$, so we use $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)=0$ for the purpose of interpolation.
+If \eqref{eq:ellipsoid_check93} does hold, then $y - \mu(x) + \mu(x_i) \in E_{x_i}$.
+In this case, we look up the batch index $b$ for the batch of sample points containing $x_i$, and evaluate
+\begin{equation}
+	\frac{1}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right) = \eta_b\left(y - \mu(x) + \mu(x_i)\right).
+\end{equation}
+To evaluate $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$, we recall the mesh cell that $y - \mu(x) + \mu(x_i)$ resides in.
+This mesh cell was already computed as a side effect of querying the AABB tree of mesh cells when we checked whether $y - \mu(x) + \mu(x_i) \in \Omega$.
+We then evaluate the relevant finite element basis functions on the cell to evaluate $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$.
+
+If $\Aop$ is symmetric, impulse responses of $\Aop$ are also impulse responses of $\Aop^T$, and we can use this fact to improve the interpolation accuracy at negligible additional cost.
+
+Let $\phi'_{y_i}(x) := \Phi(y_i,x)$ denote the impulse response of $\Aop^T$ to a point source at $y_i$.
+
+Let $\{y_i\}_{i=1}^r$ be the $k$ nearest sample points to $y$, and let . In the case where $\Aop$ is symmetric, knowing
+
+In the case where $\Aop$ is symmetric, we have
+\begin{equation}
+	\impulseresponse_{x_i}(y) = \Phi(y,x_i) = \Phi(x_i,y) = \impulseresponse'_{x_i}(y)
+\end{equation}
+knowing the impulse response $\impulseresponse_{x_i}$ of $\Aop$ at $x_i$
+
 \subsection{Product-convolution approximation}

### 2022-03-01 paper/localpsf.tex
draft of interpolation section

@@
 
-We approximate kernel entries $\Phi(y,x)$ by preconditioned radial basis function interpolation of the impulse response values
+We approximate kernel entries $\Phi(y,x)$ by preconditioned radial basis function interpolation of known impulse response values
 \begin{equation}
 \label{eq:phi_mu_shift32}
-	\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right),
-\end{equation}
-where the points $x_i$ are sample points at which $\phi_{x_i}$ is known.
+	\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
+\end{equation}
 Interpolation using all sample points $x_i$ is too expensive, so we only use the $k$ nearest sample points to $x$, excluding points $x_i$ for which \eqref{eq:phi_mu_shift32} is not defined.
@@
 \begin{equation}
-	\Phi(y,x) \approx \sum_{i=1}^k w_i b\left(\|x-x_i\|\right).
+	\Phi(y,x) \approx \sum_{i=1}^k w_i~ b\left(\|x-x_i\|\right).
 \end{equation}
@@
 %The result of this interpolation is to approximate $\Phi(y,x)$ with a linear combination of the shifted impulse response values $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$. 
-By including the factor $\frac{V(x)}{V(x_i)}$, we are implicitly performing polyharmonic spline radial basis function interpolation of the ``volume preconditioned'' function $\frac{1}{V(x)}\Phi(y,x)$, then rescaling the result by $V(x)$ to get an approximation of $\Phi(y,x)$.
+By including the factor $\frac{V(x)}{V(x_i)}$, we are implicitly interpolating the ``volume preconditioned'' function $\frac{1}{V(x)}\Phi(y,x)$, then rescaling the result by $V(x)$ to get an approximation of $\Phi(y,x)$.
 Volume preconditioning improves the approximation considerably when $V(x)$ varies widely in magnitude, as is commonly the case for many operators of practical interest.
-In the Hessian for the Stokes inverse problem, for example, $V(x)$ varies over six orders of magnitude.
-
-The point $y - \mu(x) + \mu(x_i)$ typically is not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints.
-Although it is advisable to avoid evaluating finite element functions at arbitrary points [CITE], here we are forced to do this.
+In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.
+
+The point $y - \mu(x) + \mu(x_i)$ is typically not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints.
+Although it is advisable to avoid evaluating finite element functions at arbitrary points [CITE], here we have no choice.
 We evaluate $\frac{1}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ via the following process.
@@
 
-If $\Aop$ is symmetric, impulse responses of $\Aop$ are also impulse responses of $\Aop^T$, and we can use this fact to improve the interpolation accuracy at negligible additional cost.
-
-Let $\phi'_{y_i}(x) := \Phi(y_i,x)$ denote the impulse response of $\Aop^T$ to a point source at $y_i$.
-
-Let $\{y_i\}_{i=1}^r$ be the $k$ nearest sample points to $y$, and let . In the case where $\Aop$ is symmetric, knowing
-
-In the case where $\Aop$ is symmetric, we have
-\begin{equation}
-	\impulseresponse_{x_i}(y) = \Phi(y,x_i) = \Phi(x_i,y) = \impulseresponse'_{x_i}(y)
-\end{equation}
-knowing the impulse response $\impulseresponse_{x_i}$ of $\Aop$ at $x_i$
+If $\Aop$ is symmetric, impulse responses of $\Aop$ are also impulse responses of $\Aop^T$.
+Let $\{y_i\}_{i=1}^k$ be the $k$ nearest sample points to $y$, excluding points $y_i$ such that $x - \mu(y) + \mu(y_i) \notin \Omega$.
+We could form an alternative approximation of $\Phi(y,x)$ by interpolating the values
+\begin{equation}
+\label{eq:phi_mu_shift_y}
+\phi_{y_i}\left(x - \mu(y) + \mu(y_i)\right),
+\end{equation}
+in the manner described above, but with $x$ and $y$ now taking opposite roles.
+By combining both of these approximations, we can improve the interpolation accuracy at negligible additional cost.
+We combine both approximations by placing function values
+\begin{equation*}
+	\left(\widetilde{f}_x\right)_i := \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right), \qquad i=1,\dots,k,
+\end{equation*}
+at the points $x_i-x$, placing function values
+\begin{equation*}
+\left(\widetilde{f}_y\right)_i := \frac{V(y)}{V(y_i)}\phi_{y_i}\left(x - \mu(y) + \mu(y_i)\right), \qquad i=1,\dots,k,
+\end{equation*}
+at the points $y_i-y$, and interpolating the combination of all $2k$ of these points and values at zero.
+
+To that end, define 
+\begin{equation*}
+\widetilde{x_i} := x_i - x \qquad \text{and} \qquad \widetilde{y_i} := y_i - y.
+\end{equation*}
+and let
+\begin{equation*}
+	\widetilde{B} = \begin{bmatrix}
+		\widetilde{B}_{xx} & \widetilde{B}_{xy} \\
+		\widetilde{B}_{yx} & \widetilde{B}_{xx}
+	\end{bmatrix}, \qquad \widetilde{w} = \begin{bmatrix}\widetilde{w}_x \\ \widetilde{w}_y\end{bmatrix}, \qquad \widetilde{f} = \begin{bmatrix}
+		\widetilde{f}_x \\ \widetilde{f}_y
+	\end{bmatrix},
+\end{equation*}
+be the $2k \times 2k$ block matrix and $2k \times 1$ block vectors, respectively, with block entries of $\widetilde{B}$ given by
+\begin{align*}
+\left(\widetilde{B}_{xx}\right)_{ij} = b\left(\|\widetilde{x}_i - \widetilde{x}_j\|\right), &\qquad
+\left(\widetilde{B}_{xy}\right)_{ij} = b\left(\|\widetilde{x}_i - \widetilde{y}_j\|\right), \\
+\left(\widetilde{B}_{yx}\right)_{ij} = b\left(\|\widetilde{y}_i - \widetilde{x}_j\|\right), &\qquad \left(\widetilde{B}_{yy}\right)_{ij} = b\left(\|\widetilde{y}_i - \widetilde{y}_j\|\right),
+\end{align*}
+for $i,j=1,\dots,k$. We approximate $\Phi(y,x)$ as follows:
+\begin{equation}
+\Phi(y,x) \approx \sum_{i=1}^k \left(w_x\right)_i~ b\left(\|\widetilde{x}_i\|\right) + \sum_{i=1}^k \left(w_y\right)_i~ b\left(\|\widetilde{y}_i\|\right),
+\end{equation}
+where $\widetilde{w}$ is the solution of the following regularized least-squares problem,
+\begin{equation*}
+	\min_{\widetilde{w}} \frac{1}{2}\|\widetilde{B} \widetilde{w} - \widetilde{f}\|^2 + \frac{\alpha}{2}\|\widetilde{w}\|^2,
+\end{equation*}
+for a small regularization parameter $\alpha$. 
+We solve the regularized least squares problem by QR factorization of the block $4k \times 2k$ matrix $\begin{bmatrix}\widetilde{B} & I\end{bmatrix}^T$.
+We regularize the problem because of numerical issues that arise if there are points $\widetilde{x}_i$ and $\widetilde{y}_j$ in the interpolation such that $\widetilde{x}_i \approx \widetilde{y}_j$, a situation that arises often.
+Indeed, situations where $\widetilde{x}_i = \widetilde{y}_j$ are inevitable when a regular grid is used to discretize the problem.
+In this case, $\widetilde{B}$ is (nearly) singular, so it is numerically ill-advised to solve the standard interpolation equation $\widetilde{B} \widetilde{w} = \widetilde{f}$.
+
+%In principle, this equation is still solvable if $\left(\widetilde{f}_x\right)_i = \left(\widetilde{f}_x\right)_j$, which should hold in principle.
+However, errors incurred at any stage of the approximation process, particularly the assumption that the support of $\phi_x$ is contained in the ellipsoid $E_x$, will cause small discrepancies between $\left(\widetilde{f}_x\right)_i$ and $\left(\widetilde{f}_x\right)_j$, and these errors get amplified by the ill-conditioning.
+
+%	 or $\widetilde{x}_i \approx \widetilde{y}_j$. 
+%	
+%	while $\left(\widetilde{f}_x\right)_i \neq \left(\widetilde{f}_x\right)_j$
+%
+% $\widetilde{B}\widetilde{w}=\widetilde{f}$
+
+%and let $\widetilde{B}_{xx}$, $\widetilde{B}_{xy}$, $\widetilde{B}_{yx}$, and $\widetilde{B}_{yy}$ be $k \times k$ matrices with the following entries:
+%\begin{align*}
+%	\left(\widetilde{B}_{xx}\right)_{ij} &= b\left(\|\widetilde{x}_i - \widetilde{x}_j\|\right) \\	\left(\widetilde{B}_{xy}\right)_{ij} &= b\left(\|\widetilde{x}_i - \widetilde{y}_j\|\right) \\
+%\end{align*}
+%
+%
+%\begin{equation*}
+%	\begin{bmatrix}
+%		B_{xx} & B_{xy} \\ B_{yx} & B_{yy}
+%	\end{bmatrix}
+%\end{equation*}
+%
+%Let $\phi'_{y_i}(x) := \Phi(y_i,x)$ denote the impulse response of $\Aop^T$ to a point source at $y_i$. 
+%
+%Let $\{y_i\}_{i=1}^r$ be the $k$ nearest sample points to $y$, and let . In the case where $\Aop$ is symmetric, knowing 
+%
+%In the case where $\Aop$ is symmetric, we have
+%\begin{equation}
+%	\impulseresponse_{x_i}(y) = \Phi(y,x_i) = \Phi(x_i,y) = \impulseresponse'_{x_i}(y)
+%\end{equation}
+%knowing the impulse response $\impulseresponse_{x_i}$ of $\Aop$ at $x_i$
 

### 2022-03-01 paper/localpsf.tex
draft of interpolation section

@@
 
-We approximate kernel entries $\Phi(y,x)$ by preconditioned radial basis function interpolation of known impulse response values
-\begin{equation}
-\label{eq:phi_mu_shift32}
-	\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
-\end{equation}
-Interpolation using all sample points $x_i$ is too expensive, so we only use the $k$ nearest sample points to $x$, excluding points $x_i$ for which \eqref{eq:phi_mu_shift32} is not defined.
-That is, we exclude points $x_i$ such that $y - \mu(x) + \mu(x_i) \notin \Omega$.
+As discussed in Section SECTION, we have
+\begin{equation*}
+	\Phi(y,x) = \phi_x(y) \approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right)
+\end{equation*}
+if $x'$ is close to $x$. Hence, we approximate the kernel entry $\Phi(y,x)$ by preconditioned radial basis function interpolation of the following (point,value) pairs:
+\begin{equation}
+	\left(x_i, \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)\right).
+\end{equation}
+Here $x_i \in \Omega$ are sample points at which the impulse response $\phi_{x_i}$ is known, and $x$ is the point we seek to evaluate the interpolation at.
+%. Specifically, we interpolate $(\text{point}, \text{value})$ pairs $(x_i, f_i)$ at the point $x$, where $x_i$ are sample points, and $f_i$ are given as follows:
+%\begin{equation}
+%\label{eq:phi_mu_shift32}
+%	f_i := \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
+%\end{equation}
+Interpolation using all sample points $x_i$ is too expensive, so we only use the $k$ nearest sample points to $x$.
 In numerical experiments we use $k=10$.
 The $k$ nearest sample points are found by querying a precomputed KD-tree [CITE] of all sample points.
+We exclude points $x_i$ for which $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ is not defined.
+In other words, if $y - \mu(x) + \mu(x_i) \notin \Omega$, then the point $x_i$ is considered invalid, and not used in the interpolation, even it is close to $x$.
 We check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) [CITE] of the mesh cells used to discretize the problem.
 
-Once the $k$ nearest sample points, $\{x_i\}_{i=1}^k$, are found, we approximate $\Phi(y,x)$ using preconditioned radial basis function interpolation of the following form:
+Once the $k$ nearest (valid) sample points, $\{x_i\}_{i=1}^k$, are found, we approximate $\Phi(y,x)$ using preconditioned radial basis function interpolation of the following form:
 \begin{equation}
@@
 The point $y - \mu(x) + \mu(x_i)$ is typically not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints.
-Although it is advisable to avoid evaluating finite element functions at arbitrary points [CITE], here we have no choice.
+Although it is typically recommended that one avoid evaluating finite element functions at arbitrary points [CITE], here we have no choice.
 We evaluate $\frac{1}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ via the following process.
@@
 \end{equation}
-If \eqref{eq:ellipsoid_check93} does not hold, then $y - \mu(x) + \mu(x_i) \notin E_{x_i}$, so we use $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)=0$ for the purpose of interpolation.
+If \eqref{eq:ellipsoid_check93} does not hold, then $y - \mu(x) + \mu(x_i) \notin E_{x_i}$, so we assume $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)=0$ for the purpose of interpolation.
 If \eqref{eq:ellipsoid_check93} does hold, then $y - \mu(x) + \mu(x_i) \in E_{x_i}$.
@@
 Let $\{y_i\}_{i=1}^k$ be the $k$ nearest sample points to $y$, excluding points $y_i$ such that $x - \mu(y) + \mu(y_i) \notin \Omega$.
-We could form an alternative approximation of $\Phi(y,x)$ by interpolating the values
+We could form an alternative approximation of $\Phi(y,x)$ by interpolating the (point, value) pairs
 \begin{equation}
 \label{eq:phi_mu_shift_y}
-\phi_{y_i}\left(x - \mu(y) + \mu(y_i)\right),
-\end{equation}
-in the manner described above, but with $x$ and $y$ now taking opposite roles.
-By combining both of these approximations, we can improve the interpolation accuracy at negligible additional cost.
-We combine both approximations by placing function values
+\left(y_i,\phi_{y_i}\left(x - \mu(y) + \mu(y_i)\right)\right)
+\end{equation}
+at the point $y$, in the manner described above but with $x$ and $y$ now taking opposite roles.
+We improve the interpolation accuracy at negligible additional cost by combining both of these approximations.
+Specifically, we combine both approximations by placing function values
 \begin{equation*}
@@
 
-To that end, define 
+In detail, define 
 \begin{equation*}

### 2022-03-01 paper/localpsf.tex
draft of interpolation section

@@
 
-Let $x$ and $x'$ be points in $\Omega$ that are close to each other, and consider the following three approximations:
+Let $x$ and $x'$ be points in $\Omega$ that are close to each other, and consider the following three approximations, ordered by increasing level of sophistication:
 \begin{align}
@@
 
-As discussed in Section SECTION, we have
-\begin{equation*}
-	\Phi(y,x) = \phi_x(y) \approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right)
-\end{equation*}
-if $x'$ is close to $x$. Hence, we approximate the kernel entry $\Phi(y,x)$ by preconditioned radial basis function interpolation of the following (point,value) pairs:
-\begin{equation}
-	\left(x_i, \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)\right).
-\end{equation}
-Here $x_i \in \Omega$ are sample points at which the impulse response $\phi_{x_i}$ is known, and $x$ is the point we seek to evaluate the interpolation at.
+Given a pair of points $(y,x) \subset \Omega \times \Omega$, we define the function
+\begin{equation}
+\label{eq:fxyxp}
+	f_{y,x}(x') := \frac{V(x)}{V(x')}\phi_{x_i}\left(y - \mu(x) + \mu(x')\right).
+\end{equation}
+It follows from inspection that 
+\begin{equation*}
+	\Phi(y,x) = f_{y,x}(x),
+\end{equation*}
+and from the discussion in Section SEC, we expect that $\Phi(y,x) \approx f_{y,x}(x')$ when $x'$ is close to $x$.
+The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section SEC, is a form of preconditioning that improves the quality of this approximation in situations where $V(x)$ varies rapidly.
+%, as is commonly the case for many operators of practical interest. In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.
+%As discussed in Section SECTION, we make the approximation
+%\begin{equation*}
+%	\Phi(y,x) = \phi_x(y) \approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right)
+%\end{equation*}
+%if $x'$ is close to $x$. Hence, w
+
+We approximate $\Phi(y,x)$ by performing radial basis function interpolation of the function $f_{y,x}(x')$ in the variable $x'$.
+The data used in the interpolation are the following (point,value) pairs
+\begin{equation}
+	\left(x_i, f_{y,x}(x_i)\right), \qquad i=1,\dots,k,
+\end{equation}
+and $(x, f_{y,x}(x))$ is the (point,value) pair we approximate via interpolation.
+Here $\{x_i\}_{i=1}^k \subset \Omega$ are the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined.
+%a collection of sample points at which the impulse response $\phi_{x_i}$ is known.
 %. Specifically, we interpolate $(\text{point}, \text{value})$ pairs $(x_i, f_i)$ at the point $x$, where $x_i$ are sample points, and $f_i$ are given as follows:
@@
 %\end{equation}
-Interpolation using all sample points $x_i$ is too expensive, so we only use the $k$ nearest sample points to $x$.
-In numerical experiments we use $k=10$.
-The $k$ nearest sample points are found by querying a precomputed KD-tree [CITE] of all sample points.
-We exclude points $x_i$ for which $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ is not defined.
-In other words, if $y - \mu(x) + \mu(x_i) \notin \Omega$, then the point $x_i$ is considered invalid, and not used in the interpolation, even it is close to $x$.
+%We must perform a new interpolation every time we evaluate $\Phi(y,x)$ at a new pair of points $(y,x)$, and constructing the H-matrix approximation of $\Phi$ requires evaluating $\Phi(y,x)$ at a large number of pairs of points $(y,x)$.
+%Interpolation using all sample points $x_i$ is too expensive, so we only use the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined. 
+In numerical experiments we use $k=10$. The $k$ nearest sample points are found by querying a precomputed KD-tree [CITE] of all sample points.
+
+From the definition of $f_{y,x}$ in \eqref{eq:fxyxp}, we see that $f_{y,x}(x_i)$ is only defined if $y - \mu(x) + \mu(x_i) \in \Omega$, because $\phi_x$ is only defined on $\Omega$.
+But it may happen that $y - \mu(x) + \mu(x_i) \notin \Omega$, even if $y,x,x_i \in \Omega$.
+In this case, the point $x_i$ is considered invalid, and not used in the interpolation, even if it is close to $x$.
 We check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) [CITE] of the mesh cells used to discretize the problem.
 
-Once the $k$ nearest (valid) sample points, $\{x_i\}_{i=1}^k$, are found, we approximate $\Phi(y,x)$ using preconditioned radial basis function interpolation of the following form:
-\begin{equation}
-	\Phi(y,x) \approx \sum_{i=1}^k w_i~ b\left(\|x-x_i\|\right).
-\end{equation}
-For radial basis functions, $b_i$, we use polyharmonic splines. That is, 
-\begin{equation}
-b_i(r) = r^d
-\end{equation}
-if $d$ is odd, and
-\begin{equation}
-	b_i(r) = \begin{cases}
-		r^d \log r, & r \neq 0 \\
-		0 & r = 0
+We approximate $\Phi(y,x)$ using radial basis function interpolation of the following form:
+\begin{equation}
+	\Phi(y,x) = f_{y,x}(x) \approx \sum_{i=1}^k w_i~ b\left(\|x-x_i\|\right),
+\end{equation}
+where $b_i$ are radial basis functions, and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
+\begin{equation}
+Bw = f,
+\end{equation}
+where the coefficient matrix $B \in \mathbb{R}^{k \times k}$ has entries 
+\begin{equation*}
+B_{ij} = b\left(\|x_i - x_j\|\right),
+\end{equation*}
+and the right hand side vector $f \in \mathbb{R}^k$ has entries $f_i = f_{y,x}(x_i)$, $i=1,\dots,k$. We use polyharmonic spline radial basis functions [CITE], which are defined as follows:
+\begin{equation}
+	b_i(r) := \begin{cases}
+		r^d, & d=1,3,5,\dots \\
+		r^d \log r & d=2,4,6,\dots
 	\end{cases}
 \end{equation}
-if $d$ is even.
-The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
-\begin{equation}
-	Bw = f,
-\end{equation}
-where the coefficient matrix $B \in \mathbb{R}^{k \times k}$ has entries 
-\begin{equation*}
-	B_{ij} = b\left(\|x_i - x_j\|\right),
-\end{equation*}
-and the right hand side vector $f \in \mathbb{R}^k$ has entries
-\begin{equation}
-	f_i = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
-\end{equation}
+for $r>0$, and $b_i(0):=0$.
+%\begin{equation}
+%	f_i = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
+%\end{equation}
 %The result of this interpolation is to approximate $\Phi(y,x)$ with a linear combination of the shifted impulse response values $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$. 
-By including the factor $\frac{V(x)}{V(x_i)}$, we are implicitly interpolating the ``volume preconditioned'' function $\frac{1}{V(x)}\Phi(y,x)$, then rescaling the result by $V(x)$ to get an approximation of $\Phi(y,x)$.
+%By including the factor $\frac{V(x)}{V(x_i)}$, we are implicitly interpolating the ``volume preconditioned'' function $\frac{1}{V(x)}\Phi(y,x)$, then rescaling the result by $V(x)$ to get an approximation of $\Phi(y,x)$.
 Volume preconditioning improves the approximation considerably when $V(x)$ varies widely in magnitude, as is commonly the case for many operators of practical interest.
@@
 
-The point $y - \mu(x) + \mu(x_i)$ is typically not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints.
-Although it is typically recommended that one avoid evaluating finite element functions at arbitrary points [CITE], here we have no choice.
-We evaluate $\frac{1}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ via the following process.
-First, we determine whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ by checking if
+%In order to compute $f_{y,x}(x_i)$, we must evaluate $\impulseresponse_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$, and we do this by evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$.
+Here $b$ is the index of the batch that contains the sample point $x_i$, and $\eta_b$ is the corresponding impulse response batch function that we have computed and stored.
+%Although it is typically recommended that one avoid evaluating finite element functions at arbitrary points [CITE], here we have no choice. 
+We evaluate $f_{y,x}(x_i)$ via the following process. First, we determine whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ by checking if
 \begin{equation}
@@
 \end{equation}
-If \eqref{eq:ellipsoid_check93} does not hold, then $y - \mu(x) + \mu(x_i) \notin E_{x_i}$, so we assume $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)=0$ for the purpose of interpolation.
+If \eqref{eq:ellipsoid_check93} does not hold, then $y - \mu(x) + \mu(x_i) \notin E_{x_i}$ (i.e., $y - \mu(x) + \mu(x_i)$ is outside the estimated support of $\impulseresponse_{x_i}$), so we set $f_{y,x}(x_i)=0$ for the purpose of interpolation.
 If \eqref{eq:ellipsoid_check93} does hold, then $y - \mu(x) + \mu(x_i) \in E_{x_i}$.
-In this case, we look up the batch index $b$ for the batch of sample points containing $x_i$, and evaluate
-\begin{equation}
-	\frac{1}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right) = \eta_b\left(y - \mu(x) + \mu(x_i)\right).
-\end{equation}
-To evaluate $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$, we recall the mesh cell that $y - \mu(x) + \mu(x_i)$ resides in.
-This mesh cell was already computed as a side effect of querying the AABB tree of mesh cells when we checked whether $y - \mu(x) + \mu(x_i) \in \Omega$.
-We then evaluate the relevant finite element basis functions on the cell to evaluate $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$.
+In this case, we look up the batch index $b$ for the batch of sample points that contains $x_i$, and evaluate
+\begin{equation}
+	f_{y,x}(x_i) = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right) = V(x)\eta_b\left(y - \mu(x) + \mu(x_i)\right),
+\end{equation}
+where $\eta_b$ is the impulse response batch function that we have computed and stored.
+Evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$ requires some care, because the point $y - \mu(x) + \mu(x_i)$ is typically not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints.
+To evaluate $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$, we determine which mesh cell contains $y - \mu(x) + \mu(x_i)$, then we evaluate the relevant finite element basis functions on that mesh cell.
+Fortunately, the mesh cell containing $y - \mu(x) + \mu(x_i)$ was already determined as a side effect of querying the AABB tree of mesh cells when we checked whether $y - \mu(x) + \mu(x_i) \in \Omega$.
 
@@
 \label{eq:phi_mu_shift_y}
-\left(y_i,\phi_{y_i}\left(x - \mu(y) + \mu(y_i)\right)\right)
-\end{equation}
-at the point $y$, in the manner described above but with $x$ and $y$ now taking opposite roles.
+\left(y_i,f_{x,y}(y_i)\right), \qquad i=1,\dots,k,
+\end{equation}
+in the same manner as before, but with $x$ and $y$ now taking opposite roles.
 We improve the interpolation accuracy at negligible additional cost by combining both of these approximations.
@@
 \end{equation*}
-be the $2k \times 2k$ block matrix and $2k \times 1$ block vectors, respectively, with block entries of $\widetilde{B}$ given by
+be a $2k \times 2k$ block matrix and $2k \times 1$ block vectors, respectively, with the block entries of $\widetilde{B}$ be given by
 \begin{align*}
@@
 We solve the regularized least squares problem by QR factorization of the block $4k \times 2k$ matrix $\begin{bmatrix}\widetilde{B} & I\end{bmatrix}^T$.
-We regularize the problem because of numerical issues that arise if there are points $\widetilde{x}_i$ and $\widetilde{y}_j$ in the interpolation such that $\widetilde{x}_i \approx \widetilde{y}_j$, a situation that arises often.
+We regularize the problem because of numerical issues that arise if there are points $\widetilde{x}_i$ and $\widetilde{y}_j$ in the interpolation such that $\widetilde{x}_i \approx \widetilde{y}_j$, a situation that arises often in practice.
 Indeed, situations where $\widetilde{x}_i = \widetilde{y}_j$ are inevitable when a regular grid is used to discretize the problem.

### 2022-03-01 paper/localpsf.tex
moved old stuff to end. rearranging sections

@@
 Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation.
-The details of the interpolation will be described in Section SECTION, but the key idea is that
+The details of the interpolation will be described in Section SECTION, but the key idea is something that we call \emph{local mean-displacement invariance}, which is the widely applicable assumption that
 \begin{equation}
@@
 if $x_i$ is close to $x$.
-This is shown in Figure FIG.
-We therefore approximate $\phi_x(y)$ by a weighted linear combination of the values $\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for the $k$-nearest points $x_i$ to $x$.
+This is shown in Figure FIG, and discussed in greater detail in Section SEC.
+We approximate $\phi_x(y)$ by a weighted linear combination of the values $\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for the $k$-nearest points $x_i$ to $x$.
 Here $k$ is a user-defined parameter; we use $k=10$.
@@
 The ability to rapidly evaluate approximate kernel entries $\Phi(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Phi}_H$, of $\Phi$, using the conventional adaptive cross H-matrix construction method.
-In this method, one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
+In this method, one forms low-rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
 We then use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation of $\Aop$, denoted $\mathbf{A}_H$.
@@
 \end{align}
-Approximations \eqref{eq:translate1} and \eqref{eq:translate2} are standard. Approximation \eqref{eq:translate3}, which we use,  is new. These approximations are illustrated in Figure FIG.
+Approximations \eqref{eq:translate1} and \eqref{eq:translate2} are standard.
+Approximation \eqref{eq:translate3}, which we use,  is new.
+We call operators that satisfy \eqref{eq:translate3} ``locally mean-displacement invariant.'' These approximations are illustrated in Figure FIG.
 
@@
 %In problems of practical interest, \eqref{eq:translate1} typically holds \emph{asymptotically} in the limit as $x'$ approaches $x$. 
-However, even if $\Phi$ is smooth (which is typically the case asymptotically) the relevant Lipschitz constant for continuity, and hence the required rank in the low rank approximation, can be large, and this makes algorithms based on assumption \eqref{eq:translate1} highly inefficient.
+However, even if $\Phi$ is smooth (which is typically the case asymptotically) the relevant Lipschitz constant for continuity, and hence the required rank in the low rank approximation, can be large, which makes algorithms based on assumption \eqref{eq:translate1} computationally expensive.
 This is the case, for example, for Hessians in PDE constrained inverse problems with highly informative data.
-As the data become more informative, typically the ``mass'' of $\Phi(y,x)$ concentrates nearer and nearer to the ``diagonal'' $y=x$, the Lipschitz constant increases, and the required rank blows up.
+As the data become more informative, typically the ``mass'' of $\Phi(y,x)$ concentrates nearer and nearer to the ``diagonal'' $y=x$, the Lipschitz constant increases, and the required rank in the approximation blows up.
 %Moreover, if $\Aop$ is numerically low rank, then one would typically use more accurate and robust matrix-free methods for forming low-rank approximations, such as randomized SVD or Lanczos.
@@
 On the other hand, translating a distribution translates its mean, so if \eqref{eq:translate2} holds, then $\mu(x_i)-\mu(x) \approx x_i - x$, so again \eqref{eq:translate3} holds.
-But approximation \eqref{eq:translate3} may continue to hold in situation where neither \eqref{eq:translate1} nor \eqref{eq:translate2} holds.
+Furthermore, approximation \eqref{eq:translate3} can hold in situations where neither \eqref{eq:translate1} nor \eqref{eq:translate2} holds.
 For example, \eqref{eq:translate3} will hold when an operator is locally translation-invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case (see Figure FIG).
@@
 In contrast, \eqref{eq:translate2} does not naturally generalize to operators that map between function spaces on different domains, because the formula $y-x+x'$ requires $y \in \Omega_2$ and $-x+x' \in \Omega_1$ to be added together.
-We call operators that satisfy \eqref{eq:translate3} ``locally mean-displacement invariant.''
+
+\subsection{Existing work}
+
+In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
+Product-convolution approximations have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}.
+For background on product-convolution approximations, we recommend reading the following papers: \cite{PRODCONVGOOD}.
+
+We build upon the class of product convolution approximations in which the functions, $\convkernel_i$, are impulse responses of $\Aop$ to point sources at a collection of points $x_i$ \cite{PRODCONVLIST}.
+A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the functions $\convkernel_i$ with piecewise linear interpolation \cite{NAGY}, or splines \cite{PRODCONVSPLINES}.
+With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
+In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
+But even with adaptive refinement, many matrix vector products with $\Aop$ may be required.
+In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\convkernel_i$ are computed with each matrix vector product.
+
+Our method of estimating the support of the functions $\convkernel_i$ was inspired by resolution analysis in seismic imaging \cite{RESOLUTION}.
+In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\convkernel_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
+If the width of $\convkernel_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
+In \cite{RESOLUTION}, the width of $\convkernel_p$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
+In this paper, rather than probing $\Aop^T$ with random noise functions, we use specific constant, linear, and quadratic functions.
+Our method estimates the support of $\convkernel_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
+
+Matrix-vector products with product-convolution approximations can be performed using the fast Fourier transform if the problem is discretized on a regular grid \cite{PRODCONVFFT}.
+The product-convolution approximation can then replace the original operator when using Krylov methods to solve linear systems with the original operator as the coefficient operator \cite{PRODCONVKRYLOV}.
+However, for complex geometries, problems are rarely discretized using regular grids, so we cannot easily use the fast Fourier transform.
+Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}.
+Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are amenable to fast linear algebra operations.
+Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}.
+Here we follow our previous work \cite{PRODCONVMYPAPER}, in which we convert the product-convolution approximation to H-matrix format.
+
+Conventional H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
+There are matrix-free methods for H-matrix approximation \cite{LEXINGPEELINGPROCESS}, and these methods have been used to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
+While these matrix-free H-matrix construction methods are asymptotically scalable in theory, the required number of matrix vector products can be large in practice.
+
+Product-convolution type approximations of Hessians have been used in a seismic inverse problem in \cite{GEORGSEISMICPRODCONV}, and in an advection-diffusion inverse problem in \cite{PRODCONVMYPAPER}.
+In \cite{DEMANETSEISMIC}, matrix probing \cite{MATRIXPROBING} is used to approximate the Hessian in a seismic inverse problem as the sum of simple pseudodifferential operators.
+Although it is not explicitly mentioned, the approximation in \cite{DEMANETSEISMIC} could be interpreted as a convolution-product interpolation, which is like a product-convolution approximation, except the order of pointwise multiplications and convolutions is reversed in \eqref{eq:product_convolution}.
+
+\section{Constructing the product-convolution H-matrix approximation}
+\label{sec:method}
+In this section we describe how we choose the sample points, $x_i$ (Section \ref{sec:sample_point_selection}), how we form the convolution kernels, $\convkernel_i$ (Section \ref{sec:get_impulse_response}), and how we form the weighting functions, $w_i$ (Section \ref{sec:weighting_functions}), for the product-convolution approximation, $\AopPc \approx \Aop$, given in \eqref{eq:product_convolution}.
+We also describe how we convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), and (optionally) how we modify the resulting H-matrix, $\AmatPc$, to force it to be symmetric positive semi-definite (Section \ref{sec:make_spd}).
+The complete algorithm for constructing $\AmatPc$ is shown in Algorithm \ref{alg:construct_Atilde}.
+
+\begin{algorithm2e}
+	\SetAlgoNoLine
+	\SetKwInOut{Input}{Input}
+	\SetKwInOut{Output}{Output}
+	{	
+		\Input{Linear operator $\Aop$, parameter $\text{max\_batches}$}
+		\Output{H-matrix $\AmatPc$ (optionally, $\AmatPcSymPlus$)}
+
+		Compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ using Algorithm \ref{alg:varhpi_mean_cov}.
+
+		\For{$k=1,2,\dots,\text{max\_batches}$}{
+			Choose a batch of sample points, $\pointbatch_k$, using Algorithm \ref{alg:point_choice}
+
+			Compute $\combresponse_k$ by applying $\Aop$ to the dirac comb for $\pointbatch_k$ (Section \ref{sec:get_impulse_response})
+
+		}
+
+		Form convolution kernels, $\convkernel_{x_i}$, on regular grids (Section \ref{sec:convolution_kernel_regular_grid})
+
+		Modify convolution kernels to address boundary artifacts (Section \ref{sec:boundary_extension})
+
+		Compute all weighting functions $w_i$ using Algorithm
+
+		Form H-matrix $\AmatPc$ (Section \ref{sec:H_matrix_conversion})
+
+		(optional) Modify $\AmatPc$ to make $\AmatPcSymPlus$ (Section \ref{sec:make_spd})
+
+	}
+	\caption{Construct product-convolution H-matrix}
+	\label{alg:construct_Atilde}
+\end{algorithm2e}
+
+\subsection{Sample point selection}
+\label{sec:sample_point_selection}
+
+We choose sample points, $x_i$, in batches.
+A greedy algorithm is used to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other.
+We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\impulseresponse_x$ (Section \ref{eq:mean_and_covariance_estimation}).
+The support of $\impulseresponse_x$ is estimated to be contained within an ellipsoid, so the process of choosing batches of points is an ellipsoid packing problem (Section \ref{sec:greedy_point_selection}).
+
+\subsubsection{Estimating impulse response supports}
+\label{eq:mean_and_covariance_estimation}
+
+Because of the non-negative kernel property, for each $x \in \Omega$, the impulse response $\impulseresponse_x$ is a non-negative function $y \mapsto \impulseresponse_x(y)$.
+Hence, $\impulseresponse_x$ is a scaled probability distribution.
+Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor (the ``volume'' of $\impulseresponse_x$), let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
+We make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
+\begin{equation}
+	\label{eq:support_ellipsoid}
+	E_x := \{x' \in \Omega: (x' - \spatialmean(x))^T \spatialcov(x)^{-1} (x' - \spatialmean(x)) \le \tau^2\},
+\end{equation}
+where $\tau$ is a fixed constant.
+The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\impulseresponse}_x$ (see Figure FIG).
+The quantity $\tau$ is a parameter that must be chosen by the user of the algorithm.
+The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$.
+However, the cost of our algorithm will depend on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
+We find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results.
+The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.
+
+The ellipsoid $E_x$ is not immediately available.
+To determine $E_x$, we must compute the scalar $\spatialvol(x)$, the length-$\gdim$ vector $\spatialmean(x)$, and the $\gdim \times \gdim$ matrix $\spatialcov(x)$ (recall $\gdim \in \{1,2,3\}$ is the spatial dimension of $\Omega$).
+The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}.
+Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
+But this direct approach is computationally infeasible because and it requires one operator action of $\Aop$ per point $x$ that we wish to evaluate $E_x$ at, and we will need to evaluate $E_x$ at a large number of points.
+Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions.
+We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Theorem \ref{thm:vol_mean_cov} and summarize the method in Algorithm \ref{alg:varhpi_mean_cov}.
+
+\begin{theorem}[Impulse response moments]
+	\label{thm:vol_mean_cov}
+	Let $\spatialvol:\Omega \rightarrow \mathbb{R}$, $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$, and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$, be the following functions:
+	\begin{equation*}
+		\spatialvol(x) := \int_{\Omega} \impulseresponse_x(y) dy, \qquad
+		\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
+		\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
+	\end{equation*}
+	where $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denotes the normalized version of $\impulseresponse_x$, $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$.
+Let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
+	\begin{equation*}
+		C(x) := 1, \qquad
+		L^i(x) := x^i, \qquad
+		Q^{ij}(x) = x^i x^j.
+	\end{equation*}
+	Let $f/g$ denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
+The following results hold:
+	\begin{subequations}
+		\label{eq:vol_mean_var}
+		\begin{align}
+			\spatialvol =& \left(\Aop^T C\right)^* \\
+			\spatialmean^i =& \left(\Aop^T L^i\right)^* / \spatialvol \\
+			\spatialcov^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j
+		\end{align}
+	\end{subequations}
+	 for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$.
+\end{theorem}
+
+The proof of Theorem \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}. The proof is straightforward, and follows directly from the definitions.
+
+\begin{algorithm2e}
+	\SetAlgoNoLine
+	\SetKwInOut{Input}{Input}
+	\SetKwInOut{Output}{Output}
+	{	
+		\Input{Operator $\Aop$}
+		\Output{$\spatialvol$, $\spatialmean$, and $\spatialcov$}
+
+		\tcp{Compute scaling factor $\spatialvol$}
+
+		Form constant function $C(x)=1$
+
+		Compute $\spatialvol = \left(\Aop^T C\right)^*$
+
+		\tcp{Compute mean $\spatialmean$}
+		\For{$i=1,2,\dots,\gdim$}{
+			Form linear function $L^i(x) = x^i$
+
+			Compute $\spatialmean^i = \left(\Aop^T L^i\right)^* / \spatialvol$
+		}
+		\tcp{Compute covariance $\spatialcov$}
+		\For{$i=1,2,\dots,\gdim$}{
+			\For{$j=1,\dots,i$}{
+				Form quadratic function $Q^{ij}(x) = x^i x^j$
+
+				Compute $\spatialcov^{ij} = \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j$
+
+				Set $\spatialcov^{ji} = \spatialcov^{ij}$
+
+			}
+		}
+
+	}
+	\caption{Compute scaling factor $\spatialvol$, mean $\spatialmean$, and covariance $\spatialcov$}
+	\label{alg:varhpi_mean_cov}
+\end{algorithm2e}
+
+\subsubsection{Greedy ellipsoid packing}
+\label{sec:greedy_point_selection}
+
+We use a greedy ellipsoid packing algorithm to choose batches of sample points, $\pointbatch_k$, such that there is no overlap between the ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch.
+These sample point batches will be used in Section \ref{sec:get_impulse_response} to compute many impulse responses of $\Aop$ using only a small number of operator actions of $\Aop$.
+The support of $\impulseresponse_x$ is (approximately) contained in the ellipsoid $E_x$, so by applying $\Aop$ to the Dirac comb associated with all sample points in a batch, we will recover the impulse responses for all sample points in that batch.
+
+We start with a finite set of candidate points $P$.
+To build $\pointbatch_k$, first we initialize $\pointbatch_k$ as an empty set.
+Then we select a candidate point $p \in P$ that is the farthest away from all points in previous sample point batches.
+That is, $p$ is a maximizer of the following optimization problem:
+\begin{equation*}
+\max_{p \in P} \min_{x \in \pointbatch_1 \cup \dots \cup \pointbatch_{k-1}} \|p - x\|.
+\end{equation*}
+Candidate points for the first batch are chosen arbitrarily from $P$.
+Once $p$ is selected, we remove $p$ from $P$.
+If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_k$, then we add $p$ to $\pointbatch_k$.
+Otherwise we discard $p$.
+This process repeats until there are no more points in $P$.
+This is detailed in Algorithm \ref{alg:point_choice}.
+We determine whether $E_p \cap E_q = \{\}$ using the fast ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.
+
+We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots$, until the number of batches exceeds a desired threshold.
+In our implementation, for each batch the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, except those points in previously chosen batches.
+
+\begin{algorithm2e}
+	\SetAlgoNoLine
+	\SetKwInOut{Input}{Input}
+	\SetKwInOut{Output}{Output}
+	\SetKwProg{Fn}{Function}{}{}
+
+	\Input{Finite set of candidate points $P\subset \Omega$, \\spatialmean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\domain boundary information $\partial \Omega$, \\previous sample point batches $\pointbatch_1, \dots, \pointbatch_{k-1}$}
+	\Output{Batch of new sample points $\pointbatch_k$}
+
+	Initialize empty new batch of sample points, $\pointbatch_k = \{\}$
+
+	\While{$P$ is not empty}{
+		Determine the point $p \in P$ that is farthest from all points in previous sample point batches $\pointbatch_1,\dots,\pointbatch_{k-1}$
+
+		Remove $p$ from $P$
+
+		\If{$E_p \cap E_q \neq \{\}$ for all $q \in \pointbatch_k$ and $E_p \cap \partial \Omega = \{\}$}{
+			\tcp{$E_p$ and $E_q$ are the ellipsoids defined in \eqref{eq:support_ellipsoid}}
+
+			Add $p$ to $\pointbatch_k$
+
+			Remove all points $p'$ satisfying $\spatialmean(p') \in E_p$ from $P$
+		}
+	}
+
+    \SetKwFunction{FMain}{point\_is\_acceptable}
+	\caption{Choosing one batch of sample points, $\pointbatch_k$}
+	\label{alg:point_choice}
+\end{algorithm2e}
+
+\subsection{Convolution kernels}
+\label{sec:get_impulse_response}
+
+We compute convolution kernels $\convkernel_{x_i}$ in batches by applying $\Aop$ to a small number Dirac combs.
+The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$.
+That is,
+\begin{equation*}
+	\diraccomb_k := \sum_{x_i \in \pointbatch_k} \delta_{x_i}.
+\end{equation*}
+For each batch $\pointbatch_k$, we compute the action of $\Aop$ on the associated Dirac comb:
+\begin{equation}
+	\label{eq:dirac_comb_H_action}
+	\combresponse_k := \left(\Aop \diraccomb_k^*\right)^*.
+\end{equation}
+By linearity, the resulting function $\combresponse_k$ may be written as follows:
+\begin{equation}
+	\label{eq:phi_b}
+	\combresponse_k = \left(\Aop \left(\sum_{x_i \in \pointbatch_k} \delta_{x_i}\right)^*\right)^* = \sum_{x_i \in \pointbatch_k} \left(\Aop \delta_{x_i}^*\right)^* = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i}.
+\end{equation}
+Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\impulseresponse_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\impulseresponse_j$ do not overlap when $i \neq j$, from \eqref{eq:convolution_kernel} and \eqref{eq:phi_b} we have (approximately)
+\begin{equation}
+\label{eq:varphi_eval}
+	\convkernel_{x_i}(z) = \impulseresponse_i(z+x_i) = 
+	\begin{cases}
+		\combresponse_k(z+x_i), & z+x_i \in E_{x_i} \\
+		0, & \text{otherwise}
+	\end{cases}
+\end{equation}
+for all $x_i \in \pointbatch_k$.
+By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the convolution kernels $\convkernel_{x_i}$ associated with every point $x_i \in \pointbatch_k$.
 
@@
 
+\subsubsection{Improvements when $\Aop$ is symmetric}
 If $\Aop$ is symmetric, impulse responses of $\Aop$ are also impulse responses of $\Aop^T$.
@@
 We solve the regularized least squares problem by QR factorization of the block $4k \times 2k$ matrix $\begin{bmatrix}\widetilde{B} & I\end{bmatrix}^T$.
-We regularize the problem because of numerical issues that arise if there are points $\widetilde{x}_i$ and $\widetilde{y}_j$ in the interpolation such that $\widetilde{x}_i \approx \widetilde{y}_j$, a situation that arises often in practice.
+We regularize the problem because of numerical issues that arise if there are points $\widetilde{x}_i$ and $\widetilde{y}_j$ in the interpolation that are very close to each other, a situation that arises often in practice.
 Indeed, situations where $\widetilde{x}_i = \widetilde{y}_j$ are inevitable when a regular grid is used to discretize the problem.
@@
 
-\subsection{Product-convolution approximation}
-\label{sec:prodconv_intro}
-Let 
-\begin{equation}
-\label{eq:convolution_kernel}
-\convkernel_x(y) := \impulseresponse_x(y+x)
-\end{equation}
-be the translation of $\impulseresponse_x$ to re-center it at zero instead of $x$.
-If $\Aop$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then for any point $x_0$ we would have
-\begin{align*}
-	\Aker(y,x) &= \Aker(y+(x_0-x),x+(x_0-x)) \\
-	&= \phi_{x_0}(y-x+x_0) \\
-	&= \convkernel_{x_0}(y-x)
-\end{align*}
-which would imply
-\begin{align*}
-	(\Aop u)(v) &= \int_\Omega \int_\Omega v(y) \convkernel_{x_0}(y-x) u(x) dx dy \\
-	&= \int_\Omega v(y) \left(\int_\Omega \convkernel_{x_0}(y-x) u(x) dx\right) dy \\
-	&= \left(v, \convkernel_{x_0} \ast u\right)_{L^2(\Omega)},
-\end{align*}
-where $f \ast g$ denotes convolution of the functions $f$ and $g$, and $\left(f, g\right)_{L^2(\Omega)} := \int_\Omega f(x) g(x) dx$ denotes the $L^2$ inner product of $f$ and $g$ on $\Omega$.
-Hence, if $\Aop$ were perfectly translation invariant, it would be the convolution operator $u \mapsto \left(\convkernel_{x_0} \ast u\right)^*$.
-
-Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\convkernel_x$, varies as $x$ changes.
-We therefore approximate $\Aop$ by a spatially varying weighted sum of convolution operators, $\AopPc \approx \Aop$, of the form
-\begin{equation}
-	\left(\AopPc~u\right)(v) := \left( v,~ \sum_{i=1}^\convrank \convkernel_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
-\end{equation}
-where $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$ (i.e., $\left(f \cdot g\right)(x) = f(x) g(x)$).
-Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution.
-To avoid nested subscripts, we write
-\begin{equation*}
-	\convkernel_i := \convkernel_{x_i} \quad \text{and} \quad \impulseresponse_i := \impulseresponse_{x_i},
-\end{equation*}
-with $\convkernel_{x_i}$ defined in \eqref{eq:convolution_kernel} and $\impulseresponse_{x_i}$ defined in \eqref{eq:impulse_response_defn}, to denote local convolution kernels and impulse responses corresponding to a collection of points $\{x_i\}_{i=1}^\convrank \subset \Omega$ scattered throughout the domain.
-The functions $w_i$ are spatially varying weighting functions that are used to interpolate the convolution kernels $\convkernel_i$.
-Our operator approximation method is defined by how we compute the convolution kernels, $\convkernel_i$, how we choose the points, $x_i$, and what weighting functions, $w_i$, that we use.
-In Section \ref{sec:overview_intro} we summarize the answers to these questions; we will provide detailed answers to these questions in Section \ref{sec:method}.
-
-The more locally translation invariant an operator is (i.e., the smaller the discrepancy between the left hand side and right hand side in \eqref{eq:local_translation_invariance}), the smaller the number of terms that are required in \eqref{eq:product_convolution} to achieve an accurate product-convolution approximation.
-
-\subsection{Overview of the method (old)}
-\label{sec:overview_intro}
-
-We compute one ``batch'' of $\convkernel_i$'s by applying $\Aop$ to a sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$.
-The batch of points $x_i$ are chosen so that the support of $\impulseresponse_i$ and the support of $\impulseresponse_j$ do not overlap (or do not overlap much) if $i \neq j$.
-Because these supports do not overlap, we can post-process the response of $\Aop$ to the Dirac comb to recover the functions $\impulseresponse_i$, and therefore the functions $\convkernel_i$, associated with all points $x_i$ in the batch---with one application of $\Aop$, we recover many $\convkernel_i$.
-This is illustrated in Figure FIG.
-The process is repeated to get more batches of $\convkernel_i$'s, until a desired number of batches is reached.
-
-In order to choose the points $x_i$, we need to estimate the support of $\impulseresponse_i$ \emph{before} we compute it.
-The supports of the functions $\impulseresponse_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of specially chosen functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$), then post processing the results (Section \ref{eq:mean_and_covariance_estimation}).
-This procedure is the only part of the paper that requires non-negativity of the integral kernel.
-
-We choose the $i$th weighting function, $w_i$, to be the flattest function, in a least-squares sense, that takes the value one at the sample point $x_i$, and the value zero at the other sample points $x_j$, $j \neq i$.
-Computing the weighting functions $w_i$ requires solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid.
-We call this interpolation scheme \emph{Poisson interpolation}.
-The advantage of Poisson interpolation over other interpolation methods, most notably radial basis function interpolation \cite{ESCANDE}, is that Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions.
-This is illustrated in Figure FIG.
-
-Once the $\convkernel_i$ and $w_i$ are computed, we convert $\AopPc$ to hierarchical matrix (H-matrix) format.
-Let $\Aker_\text{pc}$ denote the integral kernel associated with $\AopPc$.
-While $\Aker(y,x)$ is not easily computable, $\Aker_\text{pc}(y,x)$ is given by the formula
-\begin{equation}
-	\label{eq:kernel_entries}
-	\AkerPc(y,x) = \sum_{i=1}^\convrank \convkernel_i(y-x) w_i(x).
-\end{equation}
-This follows from~\eqref{eq:product_convolution} and the fact that the convolution operator $u \mapsto \convkernel_i \ast u$ has $(y,x)$ kernel entry given by $\convkernel(y-x)$.
-Formula~\eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\AopPc$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
-Once in H-matrix format, fast H-matrix arithmetic is used to invert $\AopPc$, or perform other useful matrix operations.
-
-Often we are interested in approximating operators $\Aop$ that are symmetric positive semi-definite.
-Unfortunately, $\AopPc$ is, in general, non-symmetric and indefinite due to errors in the product convolution approximation.
-In this case, we modify the H-matrix representation of $\AopPc$ to make it positive definite, by symmetrizing it, then applying a specially chosen rational matrix function to the symmetrized H-matrix.
-
-\subsection{Existing work}
-
-In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
-Product-convolution approximations have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}.
-For background on product-convolution approximations, we recommend reading the following papers: \cite{PRODCONVGOOD}.
-
-We build upon the class of product convolution approximations in which the functions, $\convkernel_i$, are impulse responses of $\Aop$ to point sources at a collection of points $x_i$ \cite{PRODCONVLIST}.
-A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the functions $\convkernel_i$ with piecewise linear interpolation \cite{NAGY}, or splines \cite{PRODCONVSPLINES}.
-With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
-In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
-But even with adaptive refinement, many matrix vector products with $\Aop$ may be required.
-In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\convkernel_i$ are computed with each matrix vector product.
-
-Our method of estimating the support of the functions $\convkernel_i$ was inspired by resolution analysis in seismic imaging \cite{RESOLUTION}.
-In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\convkernel_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
-If the width of $\convkernel_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
-In \cite{RESOLUTION}, the width of $\convkernel_p$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
-In this paper, rather than probing $\Aop^T$ with random noise functions, we use specific constant, linear, and quadratic functions.
-Our method estimates the support of $\convkernel_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
-
-Matrix-vector products with product-convolution approximations can be performed using the fast Fourier transform if the problem is discretized on a regular grid \cite{PRODCONVFFT}.
-The product-convolution approximation can then replace the original operator when using Krylov methods to solve linear systems with the original operator as the coefficient operator \cite{PRODCONVKRYLOV}.
-However, for complex geometries, problems are rarely discretized using regular grids, so we cannot easily use the fast Fourier transform.
-Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}.
-Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are amenable to fast linear algebra operations.
-Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}.
-Here we follow our previous work \cite{PRODCONVMYPAPER}, in which we convert the product-convolution approximation to H-matrix format.
-
-Conventional H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
-There are matrix-free methods for H-matrix approximation \cite{LEXINGPEELINGPROCESS}, and these methods have been used to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
-While these matrix-free H-matrix construction methods are asymptotically scalable in theory, the required number of matrix vector products can be large in practice.
-
-Product-convolution type approximations of Hessians have been used in a seismic inverse problem in \cite{GEORGSEISMICPRODCONV}, and in an advection-diffusion inverse problem in \cite{PRODCONVMYPAPER}.
-In \cite{DEMANETSEISMIC}, matrix probing \cite{MATRIXPROBING} is used to approximate the Hessian in a seismic inverse problem as the sum of simple pseudodifferential operators.
-Although it is not explicitly mentioned, the approximation in \cite{DEMANETSEISMIC} could be interpreted as a convolution-product interpolation, which is like a product-convolution approximation, except the order of pointwise multiplications and convolutions is reversed in \eqref{eq:product_convolution}.
-
-\section{Constructing the product-convolution H-matrix approximation}
-\label{sec:method}
-In this section we describe how we choose the sample points, $x_i$ (Section \ref{sec:sample_point_selection}), how we form the convolution kernels, $\convkernel_i$ (Section \ref{sec:get_impulse_response}), and how we form the weighting functions, $w_i$ (Section \ref{sec:weighting_functions}), for the product-convolution approximation, $\AopPc \approx \Aop$, given in \eqref{eq:product_convolution}.
-We also describe how we convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), and (optionally) how we modify the resulting H-matrix, $\AmatPc$, to force it to be symmetric positive semi-definite (Section \ref{sec:make_spd}).
-The complete algorithm for constructing $\AmatPc$ is shown in Algorithm \ref{alg:construct_Atilde}.
-
-\begin{algorithm2e}
-	\SetAlgoNoLine
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Output}
-	{	
-		\Input{Linear operator $\Aop$, parameter $\text{max\_batches}$}
-		\Output{H-matrix $\AmatPc$ (optionally, $\AmatPcSymPlus$)}
-
-		Compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ using Algorithm \ref{alg:varhpi_mean_cov}.
-
-		\For{$k=1,2,\dots,\text{max\_batches}$}{
-			Choose a batch of sample points, $\pointbatch_k$, using Algorithm \ref{alg:point_choice}
-
-			Compute $\combresponse_k$ by applying $\Aop$ to the dirac comb for $\pointbatch_k$ (Section \ref{sec:get_impulse_response})
-
-		}
-
-		Form convolution kernels, $\convkernel_{x_i}$, on regular grids (Section \ref{sec:convolution_kernel_regular_grid})
-
-		Modify convolution kernels to address boundary artifacts (Section \ref{sec:boundary_extension})
-
-		Compute all weighting functions $w_i$ using Algorithm
-
-		Form H-matrix $\AmatPc$ (Section \ref{sec:H_matrix_conversion})
-
-		(optional) Modify $\AmatPc$ to make $\AmatPcSymPlus$ (Section \ref{sec:make_spd})
-
-	}
-	\caption{Construct product-convolution H-matrix}
-	\label{alg:construct_Atilde}
-\end{algorithm2e}
-
-\subsection{Sample point selection}
-\label{sec:sample_point_selection}
-
-We choose sample points, $x_i$, in batches.
-A greedy algorithm is used to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other.
-We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\impulseresponse_x$ (Section \ref{eq:mean_and_covariance_estimation}).
-The support of $\impulseresponse_x$ is estimated to be contained within an ellipsoid, so the process of choosing batches of points is an ellipsoid packing problem (Section \ref{sec:greedy_point_selection}).
-
-\subsubsection{Estimating impulse response supports}
-\label{eq:mean_and_covariance_estimation}
-
-Because of the non-negative kernel property, for each $x \in \Omega$, the impulse response $\impulseresponse_x$ is a non-negative function $y \mapsto \impulseresponse_x(y)$.
-Hence, $\impulseresponse_x$ is a scaled probability distribution.
-Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor (the ``volume'' of $\impulseresponse_x$), let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
-We make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
-\begin{equation}
-	\label{eq:support_ellipsoid}
-	E_x := \{x' \in \Omega: (x' - \spatialmean(x))^T \spatialcov(x)^{-1} (x' - \spatialmean(x)) \le \tau^2\},
-\end{equation}
-where $\tau$ is a fixed constant.
-The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\impulseresponse}_x$ (see Figure FIG).
-The quantity $\tau$ is a parameter that must be chosen by the user of the algorithm.
-The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$.
-However, the cost of our algorithm will depend on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
-We find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results.
-The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.
-
-The ellipsoid $E_x$ is not immediately available.
-To determine $E_x$, we must compute the scalar $\spatialvol(x)$, the length-$\gdim$ vector $\spatialmean(x)$, and the $\gdim \times \gdim$ matrix $\spatialcov(x)$ (recall $\gdim \in \{1,2,3\}$ is the spatial dimension of $\Omega$).
-The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}.
-Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
-But this direct approach is computationally infeasible because and it requires one operator action of $\Aop$ per point $x$ that we wish to evaluate $E_x$ at, and we will need to evaluate $E_x$ at a large number of points.
-Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions.
-We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Theorem \ref{thm:vol_mean_cov} and summarize the method in Algorithm \ref{alg:varhpi_mean_cov}.
-
-\begin{theorem}[Impulse response moments]
-	\label{thm:vol_mean_cov}
-	Let $\spatialvol:\Omega \rightarrow \mathbb{R}$, $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$, and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$, be the following functions:
-	\begin{equation*}
-		\spatialvol(x) := \int_{\Omega} \impulseresponse_x(y) dy, \qquad
-		\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
-		\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
-	\end{equation*}
-	where $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denotes the normalized version of $\impulseresponse_x$, $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$.
-Let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
-	\begin{equation*}
-		C(x) := 1, \qquad
-		L^i(x) := x^i, \qquad
-		Q^{ij}(x) = x^i x^j.
-	\end{equation*}
-	Let $f/g$ denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
-The following results hold:
-	\begin{subequations}
-		\label{eq:vol_mean_var}
-		\begin{align}
-			\spatialvol =& \left(\Aop^T C\right)^* \\
-			\spatialmean^i =& \left(\Aop^T L^i\right)^* / \spatialvol \\
-			\spatialcov^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j
-		\end{align}
-	\end{subequations}
-	 for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$.
-\end{theorem}
-
-The proof of Theorem \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}. The proof is straightforward, and follows directly from the definitions.
-
-\begin{algorithm2e}
-	\SetAlgoNoLine
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Output}
-	{	
-		\Input{Operator $\Aop$}
-		\Output{$\spatialvol$, $\spatialmean$, and $\spatialcov$}
-
-		\tcp{Compute scaling factor $\spatialvol$}
-
-		Form constant function $C(x)=1$
-
-		Compute $\spatialvol = \left(\Aop^T C\right)^*$
-
-		\tcp{Compute mean $\spatialmean$}
-		\For{$i=1,2,\dots,\gdim$}{
-			Form linear function $L^i(x) = x^i$
-
-			Compute $\spatialmean^i = \left(\Aop^T L^i\right)^* / \spatialvol$
-		}
-		\tcp{Compute covariance $\spatialcov$}
-		\For{$i=1,2,\dots,\gdim$}{
-			\For{$j=1,\dots,i$}{
-				Form quadratic function $Q^{ij}(x) = x^i x^j$
-
-				Compute $\spatialcov^{ij} = \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j$
-
-				Set $\spatialcov^{ji} = \spatialcov^{ij}$
-
-			}
-		}
-
-	}
-	\caption{Compute scaling factor $\spatialvol$, mean $\spatialmean$, and covariance $\spatialcov$}
-	\label{alg:varhpi_mean_cov}
-\end{algorithm2e}
-
-\subsubsection{Greedy ellipsoid packing}
-\label{sec:greedy_point_selection}
-
-We use a greedy ellipsoid packing algorithm to choose batches of sample points, $\pointbatch_k$, such that there is no overlap between the ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch.
-These sample point batches will be used in Section \ref{sec:get_impulse_response} to compute many impulse responses of $\Aop$ using only a small number of operator actions of $\Aop$.
-The support of $\impulseresponse_x$ is (approximately) contained in the ellipsoid $E_x$, so by applying $\Aop$ to the Dirac comb associated with all sample points in a batch, we will recover the impulse responses for all sample points in that batch.
-
-We start with a finite set of candidate points $P$.
-To build $\pointbatch_k$, first we initialize $\pointbatch_k$ as an empty set.
-Then we select a candidate point $p \in P$ that is the farthest away from all points in previous sample point batches.
-That is, $p$ is a maximizer of the following optimization problem:
-\begin{equation*}
-\max_{p \in P} \min_{x \in \pointbatch_1 \cup \dots \cup \pointbatch_{k-1}} \|p - x\|.
-\end{equation*}
-Candidate points for the first batch are chosen arbitrarily from $P$.
-Once $p$ is selected, we remove $p$ from $P$.
-If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_k$, then we add $p$ to $\pointbatch_k$.
-Otherwise we discard $p$.
-This process repeats until there are no more points in $P$.
-This is detailed in Algorithm \ref{alg:point_choice}.
-We determine whether $E_p \cap E_q = \{\}$ using the fast ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.
-
-We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots$, until the number of batches exceeds a desired threshold.
-In our implementation, for each batch the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, except those points in previously chosen batches.
-
-\begin{algorithm2e}
-	\SetAlgoNoLine
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Output}
-	\SetKwProg{Fn}{Function}{}{}
-
-	\Input{Finite set of candidate points $P\subset \Omega$, \\spatialmean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\domain boundary information $\partial \Omega$, \\previous sample point batches $\pointbatch_1, \dots, \pointbatch_{k-1}$}
-	\Output{Batch of new sample points $\pointbatch_k$}
-
-	Initialize empty new batch of sample points, $\pointbatch_k = \{\}$
-
-	\While{$P$ is not empty}{
-		Determine the point $p \in P$ that is farthest from all points in previous sample point batches $\pointbatch_1,\dots,\pointbatch_{k-1}$
-
-		Remove $p$ from $P$
-
-		\If{$E_p \cap E_q \neq \{\}$ for all $q \in \pointbatch_k$ and $E_p \cap \partial \Omega = \{\}$}{
-			\tcp{$E_p$ and $E_q$ are the ellipsoids defined in \eqref{eq:support_ellipsoid}}
-
-			Add $p$ to $\pointbatch_k$
-
-			Remove all points $p'$ satisfying $\spatialmean(p') \in E_p$ from $P$
-		}
-	}
-
-    \SetKwFunction{FMain}{point\_is\_acceptable}
-	\caption{Choosing one batch of sample points, $\pointbatch_k$}
-	\label{alg:point_choice}
-\end{algorithm2e}
-
-\subsection{Convolution kernels}
-\label{sec:get_impulse_response}
-
-We compute convolution kernels $\convkernel_{x_i}$ in batches by applying $\Aop$ to a small number Dirac combs.
-The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$.
-That is,
-\begin{equation*}
-	\diraccomb_k := \sum_{x_i \in \pointbatch_k} \delta_{x_i}.
-\end{equation*}
-For each batch $\pointbatch_k$, we compute the action of $\Aop$ on the associated Dirac comb:
-\begin{equation}
-	\label{eq:dirac_comb_H_action}
-	\combresponse_k := \left(\Aop \diraccomb_k^*\right)^*.
-\end{equation}
-By linearity, the resulting function $\combresponse_k$ may be written as follows:
-\begin{equation}
-	\label{eq:phi_b}
-	\combresponse_k = \left(\Aop \left(\sum_{x_i \in \pointbatch_k} \delta_{x_i}\right)^*\right)^* = \sum_{x_i \in \pointbatch_k} \left(\Aop \delta_{x_i}^*\right)^* = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i}.
-\end{equation}
-Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\impulseresponse_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\impulseresponse_j$ do not overlap when $i \neq j$, from \eqref{eq:convolution_kernel} and \eqref{eq:phi_b} we have (approximately)
-\begin{equation}
-\label{eq:varphi_eval}
-	\convkernel_{x_i}(z) = \impulseresponse_i(z+x_i) = 
-	\begin{cases}
-		\combresponse_k(z+x_i), & z+x_i \in E_{x_i} \\
-		0, & \text{otherwise}
-	\end{cases}
-\end{equation}
-for all $x_i \in \pointbatch_k$.
-By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the convolution kernels $\convkernel_{x_i}$ associated with every point $x_i \in \pointbatch_k$.
-
-\subsubsection{Interpolation onto local regular grids}
-\label{sec:convolution_kernel_regular_grid}
-
-To convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), we will need to evaluate $\convkernel_{x_i}(z)$ for a large number of points $z$.
-Per \eqref{eq:varphi_eval}, evaluating $\convkernel_{x_i}(z)$ requires evaluating $\combresponse_k(z+x_i)$.
-In computations, $\combresponse_k$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh.
-Evaluating $\combresponse_k(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell.
-Although this procedure is asymptotically scalable, it is slow in practice \cite{FEMEVALSLOW}.
-
-To speed up evaluation of $\convkernel_{x_i}$, we interpolate $\convkernel_{x_i}$ onto a regular rectilinear grid on a coordinate axis aligned box that just barely contains the ellipsoid $E_{x_i}$.
-We transfer the function $\convkernel_{x_i}$ from the irregular mesh to the regular grid by evaluating $\convkernel_{x_i}$ at all gridpoints in the regular grid, using the slow process described in the previous paragraph.
-This is an upfront cost which is done ``offline.'' After $\convkernel_{x_i}$ has been transferred to the regular grid, we use fast regular grid interpolation to perform all subsequent ``online'' evaluations of $\convkernel_{x_i}$.
-This yields a substantial overall speedup because the number of finite element function evaluations required to transfer the functions $\convkernel_{x_i}$ to the regular grid is $O(\fedim)$, while the number of kernel evaluations required to convert $\AopPc$ to H-matrix format is $O(k \fedim \log \fedim)$.
-Here $\fedim$ is the number of degrees of freedom in the finite element space used to discretize $\combresponse_k$, and $\hmatrixrank$ is the H-matrix rank.
-Furthermore, transferring $\convkernel_{x_i}$ to a regular grid requires evaluating $\convkernel_{x_i}$ at a predetermined and highly structured collection of points.
-This structure allows for faster determination of which mesh cell that a point is in, and allows one to evaluate the finite element function at all regular gridpoints in parallel.
-In contrast, constructing the H-matrix requires evaluating $\convkernel_{x_i}$ at an unpredictable sequence of scattered points, which is an expensive task for finite element function evaluation, but a cheap task for regular grid interpolation.
-Having $\convkernel_{x_i}$ stored on a regular grid also simplifies the implementation of the method that we use to handle boundary artifacts, which will be described in Section \ref{sec:boundary_extension}.
-
-We incur interpolation error when transferring $\convkernel_{x_i}$ from the irregular mesh to the regular grid.
-To limit this interpolation error, the density of the regular grid should be as dense or denser than the local density of the finite element mesh.
-In our numerical results, the distance between neighboring points in the regular grid is chosen to be half the minimum distance between any pair of finite element Lagrange nodes that are contained in $E_{x_i}$.
-The grid transfer procedure is illustrated in Figure FIG.
-
-When $x_i$ is near a boundary, it may occur that some gridpoints for the regular grid are in locations where $\convkernel_{x_i}$ is undefined.
-In this case, we store a placeholder value for $\convkernel_{x_i}$ at those gridpoints.
-We use the IEEE standard value of ``NaN'' (not a number) for this placeholder.
-In the next section (Section \ref{sec:boundary_extension}) we replace these NaN values with more appropriate values.
-
-\subsubsection{Addressing boundary artifacts}
-\label{sec:boundary_extension}
-
-Evaluating the kernel for our product-convolution approximation, $\AkerPc(y,x)$, requires evaluating $\convkernel_{x_i}(y-x)$.
-But $\convkernel_{x_i}(y-x)$ is undefined  when $y - x + x_i$ is outside $\Omega$, which may occur even if $x \in \Omega$ and $y \in \Omega$.
-The default approach to deal with this problem is to extend $\convkernel_{x_i}$ by zero wherever it is undefined.
-But this leads to substantial errors, known as boundary artifacts, which are especially significant when the sample point $x_i$ is near the boundary of the domain.
-In this case, the impulse response is artificially chopped off by the domain boundary.
-See Figure FIG for an illustration.
-
-While $\convkernel_{x_i}$ may be undefined at a point $z$, often there is a sample point $x_j$ that is near $x_i$, but further away from the boundary than $x_i$, such that $\convkernel_{x_j}$ is well defined at $z$.
-To address boundary artifacts, we fill in missing/undefined portions of convolution kernels using information from other convolution kernels associated with the $k$ nearest sample points (we use $k=8$ in our numerical results).
-Kernels associated with nearer sample points are given precedence over kernels associated with farther away sample points.
-
-In detail, let $p_1, p_2, \dots, p_k$ be the $k$ nearest sample points to $x_i$, ordered by nearness, including $x_i$ itself.
-That is, $p_1=x_i$, $p_2$ is the nearest neighbor to $x_i$, $p_3$ is the second nearest neighbor to $x_i$, and so on.
-Also, recall from the previous section (Section \ref{sec:convolution_kernel_regular_grid}) that the functions $\convkernel_{x_i}$ are stored on regular rectilinear grids.
-To extend $\convkernel_{x_i}$, we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\convkernel_{x_i}$, and contains that grid as a subgrid, but is large enough to contain the boxes associated with all of the neighbor kernels, $\convkernel_{p_1}, \convkernel_{p_2}, \dots, \convkernel_{p_k}$ (recall that the functions $\convkernel_{x_i}$ are all centered at zero).
-We initialize an extended convolution kernel, $\convkernel_{x_i}^E$, to zero on this expanded grid.
-Then we interpolate the functions $\convkernel_{p_j}$ onto this expanded grid in reverse order.
-First we interpolate $\convkernel_{p_k}$ onto the grid, then $\convkernel_{p_{k-1}}$, and so on until the last step in which we interpolate the original kernel $\convkernel_{p_1} = \convkernel_{x_i}$.
-At each step, we overwrite existing values of $\convkernel_{x_i}^E$ with new values from the kernel currently being interpolated, except when the new value of $\convkernel_{p_j}$ is undefined (``NaN'').
-At gridpoints where the kernel being interpolated is undefined, we leave $\convkernel_{x_i}^E$ unchanged.
-In this way, we fill in undefined values of $\convkernel_{x_i}$ with values from successively further and further away kernels.
-This process is illustrated in Figure FIG.
-To avoid notational clutter, and since we will henceforth only work with the extended kernel, $\convkernel_{x_i}^E$, in the remainder of this paper the symbol $\convkernel_{x_i}$ refers to $\convkernel_{x_i}^E$.
-
-\subsection{Poisson weighting functions}
-\label{sec:weighting_functions}
-
-Consider the following optimization problem,
-\begin{equation}
-	\label{eq:wi_optimization_problem_generic}
-	\begin{aligned}
-	\min_{v \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta v}_{L^2(\Omega)}^2 \\
-	\text{such that} &\quad v(x_j) = \interpolatedvalues_j, \quad j=1,\dots,\convrank,
-	\end{aligned}
-\end{equation}
-in which one seeks a smooth function $v$ that takes the values $\interpolatedvalues_j \in \mathbb{R}$ at the points $x_j$.
-Here $\Delta$ is the Laplacian on $\Omega$ and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions.
-We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to optimization problem \eqref{eq:wi_optimization_problem_generic}, with
-\begin{equation*}
-	b_j = w_i(x_j) = 
-	\begin{cases}
-		1, & j=i\\
-		0, & j \neq i.
-	\end{cases}
-\end{equation*}
-In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' \cite{JACOBSEN} interpolation of the convolution kernels, $\convkernel_i$, in a least-squares sense.
-Also, these weighting functions sum to one pointwise (Proposition \ref{thm:wi_sum_to_one}).
-
-The solution to optimization problem \eqref{eq:wi_optimization_problem_generic} is given via the following procedure:
-\begin{enumerate}
-	\item For each sample point $x_i$, we compute a function $\firstgreens_i$ by solving the following Neumann Poisson problem:
-	\begin{equation}
-		\label{eq:psi_eq}
-		\begin{cases}
-			\Delta \firstgreens_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
-			\nu \cdot \nabla \firstgreens_i = 0 & \text{on } \partial \Omega,
-		\end{cases}
-	\end{equation}
-	with condition $\int_\Omega \firstgreens_i(x) dx = 0$.
-Here $\delta_{x_i}$ is the point source centered at $x_i$, $|\Omega|$ is the measure of the domain $\Omega$, and $\nu$ is the normal vector to $\partial \Omega$.
-	\item For each function $\firstgreens_i$, we compute a function $\secondgreens_i$ by solving another Neumann Poisson problem:
-	\begin{equation}
-		\label{eq:theta_eq}
-		\begin{cases}
-			\Delta \secondgreens_i = \firstgreens_i, & \text{in }\Omega, \\
-			\nu \cdot \nabla \secondgreens_i = 0 & \text{on } \partial \Omega,
-		\end{cases}
-	\end{equation}
-	with condition $\int_\Omega \secondgreens_i(x) dx = 0$.
-	\item The solution to optimization problem \eqref{eq:wi_optimization_problem_generic}, which we denote by $v_*$, is formed as follows:
-	\begin{equation}
-		\label{eq:optimal_v}
-		v_*(x) := \constcoeff + \sum_{i=1}^\convrank \secondgreenscoeff_i \secondgreens_i(x),
-	\end{equation}
-	where the constants $\secondgreenscoeff_i \in \mathbb{R}$, $i=1,\dots,\convrank$, and $\constcoeff \in \mathbb{R}$ are the solution to the following linear system:
-	\begin{equation}
-		\label{eq:interpolation_linear_system}
-		\begin{bmatrix}
-			\pointinteractionmatrix & \bm{1} \\ \bm{1}^T & 0
-		\end{bmatrix}
-		\begin{bmatrix}
-			\bm{\secondgreenscoeff} \\ \constcoeff
-		\end{bmatrix}
-		=
-		\begin{bmatrix}
-			\bm{\interpolatedvalues} \\ 0
-		\end{bmatrix}.
-	\end{equation}
-	Here $\bm{\secondgreenscoeff} \in \mathbb{R}^\convrank$ is the vector with $i$th component $\secondgreenscoeff_i$, $\bm{1} \in \bm{R}^\convrank$ is the column vector of all ones, $\bm{\interpolatedvalues} \in \mathbb{R}^\convrank$ is the vector with $i$th component $\interpolatedvalues_i$, and $\pointinteractionmatrix \in \mathbb{R}^{\convrank \times \convrank}$ is the matrix with entries given by
-	\begin{equation}
-		\label{eq:S_matrix}
-		\pointinteractionmatrix_{ij} := \left( \firstgreens_i, \firstgreens_j \right)_{L^2(\Omega)}.
-	\end{equation}
-\end{enumerate}
-Linear system \eqref{eq:interpolation_linear_system} is the KKT system that arises if one solves optimization problem \eqref{eq:wi_optimization_problem_generic} on the $\convrank+1$-dimensional subspace of $H^2_N(\Omega)$ spanned by the functions $\secondgreens_i$, $i=1,\dots,\convrank$, and the constant function $C(x)=1$.
-In Theorem \ref{thm:wi_no_breakdown}, we prove that this procedure does not break down, and in Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v_*$ produced by this procedure is the solution to optimization problem \eqref{eq:wi_optimization_problem_generic} on $H^2_N(\Omega)$.
-
-Constructing all $\convrank$ of the functions $w_i$ thus requires solving the Poisson equation $2\convrank$ times with different right hand side sources to form the functions $\firstgreens_i$ and $\secondgreens_i$, then post-processing these functions.
-This is shown in Algorithm \ref{alg:weighting_functions_incremental}.
-The Poisson PDE solves, which are the most expensive part of this procedure, can be performed cheaply with multigrid.
-
-It is important to note that functions in $H^2_N(\Omega)$ only have well-defined pointwise values if the spatial dimension of the domain is $\gdim=1$, $2$, or $3$.
-Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-posed in three spatial dimensions or fewer.
-In four or higher spatial dimensions, one could construct appropriate weighting functions using radial basis functions, though we do not discuss this here.
-
-\begin{thm}[Procedure for computing $v_*$ does not break down]
-	\label{thm:wi_no_breakdown}
-	If all the sample points $x_i$, $i=1,\dots,\convrank$, are distinct, then PDE \eqref{eq:psi_eq} is uniquely solvable for $\firstgreens_i$, PDE \eqref{eq:theta_eq} is uniquely solvable for $\secondgreens_i$, and linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\bm{\secondgreenscoeff}$ and $\constcoeff$.
-That is, the procedure for computing $v_*$ does not break down.
-Furthermore, $v_* \in H^2_N(\Omega)$.
-\end{thm}
-
-\begin{proof}
-	Existence and uniqueness for the solutions to PDEs \eqref{eq:psi_eq} and \eqref{eq:theta_eq} follow from the standard elliptic PDE theory.
-By standard elliptic regularity theory, $\firstgreens_i \in H^1(\Omega)$, so $\firstgreens_i \in L^2(\Omega)$, and so all matrix entries $\pointinteractionmatrix_{ij}$ are finite.
-
-	Using the definition of $\pointinteractionmatrix$ and basic linear algebra, it is straightforward to show that if $\pointinteractionmatrix$ were singular, there would exist constants $a_1, a_2, \dots a_\convrank$, which are not all zero, such that
-	\begin{equation*}
-		0 = a_1 \firstgreens_1 + a_2 \firstgreens_2 + \dots a_\convrank \firstgreens_\convrank.
-	\end{equation*}
-	Applying the Laplacian to this equation yields
-	\begin{equation*}
-		0 = a_1 \delta_{x_1} + a_2 \delta_{x_2} + \dots a_\convrank \delta_{x_\convrank} - \left(\sum_{i=1}^\convrank a_i\right)\bigg/|\Omega|,
-	\end{equation*}
-	which is impossible because a nonzero linear combination of delta distributions at different points cannot equal a constant function.
-Thus we must conclude that $\pointinteractionmatrix$ is non-singular.
-By block matrix factorization, non-singularity of $\pointinteractionmatrix$ implies that linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\bm{\secondgreenscoeff}$ and $\constcoeff$.
-
-	By standard elliptic regularity theory, the functions $\secondgreens_i$ are in $H^3(\Omega)$, and therefore they are in $H^2(\Omega)$.
-By construction the functions $\secondgreens_i$ have Neumann zero boundary conditions, so $\secondgreens_i \in H^2_N(\Omega)$.
-Since $v_*$ is a linear combination of the functions $\secondgreens_i$ and the constant function $C(x)=1$ which is in $H^2_N(\Omega)$, we have $v_* \in H^2_N(\Omega)$.
-\end{proof}
-
-\begin{lem}[Integral representation of delta distribution on $H^2_N(\Omega)$]
-	\label{lem:psi_delta_eval}
-	We have
-	\begin{equation}
-		\label{eq:inline_wi_lemma}
-		h(x_j) = \left(\firstgreens_j, \Delta h\right)_{L^2(\Omega)} + \avg(h)
-	\end{equation}
-	for all $h\in H^2_N(\Omega)$, where
-	\begin{equation*}
-		\avg(h) :=	 \frac{1}{|\Omega|} \int_\Omega h ~dx,
-	\end{equation*}
-	and $|\Omega| = \int_\Omega 1 ~dx$ is the Lebesgue measure of $\Omega$.
-\end{lem}
-
-\begin{proof}
-	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\firstgreens_j$ onto $h$, and using the Neumann zero boundary conditions for $\firstgreens_j$ and $h$ to eliminate boundary integral terms.
-\end{proof}
-
-\begin{lem}[Characterization of $\pointinteractionmatrix$]
-	\label{lem:S_matrix_characterizations}
-	We have
-	\begin{equation*}
-		\pointinteractionmatrix_{ij} = \secondgreens_j(x_i).
-	\end{equation*}
-\end{lem}
-
-\begin{proof}
-	We compute
-	\begin{equation*}
-		\secondgreens_j(x_i) = \left(\firstgreens_i, \Delta \secondgreens_j\right)_{L^2(\Omega)} + \avg\left(\secondgreens_j\right)
-		= \left(\firstgreens_i, \firstgreens_j\right)_{L^2(\Omega)} = \pointinteractionmatrix_{ij}.
-	\end{equation*}
-	In the first equality we used Lemma \ref{lem:psi_delta_eval}.
-In the second equality we used the facts that $\Delta \secondgreens_j = \firstgreens_j$ and $\avg(\secondgreens_j) = \left(1 / |\Omega|\right) \int_\Omega \secondgreens_j = 0$ from the definition of $\secondgreens_j$ in \eqref{eq:theta_eq}.
-\end{proof}
-
-\begin{thm}[Solution of optimization problem \eqref{eq:wi_optimization_problem_generic}]
-	\label{thm:wi_alg_is_correct}
-	The function $v_*$ given in \eqref{eq:optimal_v} is the unique solution to optimization problem \eqref{eq:wi_optimization_problem_generic}.
-\end{thm}
-
-\begin{proof}
-	Let
-	\begin{equation*}
-		X := \{v \in H^2_N(\Omega) : v(x_i) = \interpolatedvalues_i \text{ for }i=1,\dots,\convrank\}
-	\end{equation*}
-	denote the feasible set for optimization problem \ref{eq:wi_optimization_problem_generic}, and let 
-	\begin{equation*}
-		X_0 := \{v \in H^2_N(\Omega) : v(x_i) = 0 \text{ for }i=1,\dots,\convrank\}
-	\end{equation*}	
-	denote the tangent space to $X$.
-Note that pointwise values of functions in $H_N^2(\Omega)$ are well-defined since the Sobolev imbedding theorem (see, e.g., Theorem 7.19 in \cite{Arbogast}) implies that $H^2(\Omega)$ is continuously imbedded in the space of continuous functions $C^0(\Omega)$ when $\gdim < 4$, and we are considering $\gdim \in \{1,2,3\}$.
-The feasible set $X$ is therefore a closed affine subspace of $H^2_N(\Omega)$, the tangent space $X_0$ is a closed subspace of $H^2_N(\Omega)$, and optimization problem \eqref{eq:wi_optimization_problem_generic} is well-defined.
-
-	The proof now proceeds in three steps.
-First, we show that $v_*$ is feasible ($v_* \in X$).
-Second, we show that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} if it satisfies a certain variational equation.
-Third, we verify that $v_*$ satisfies this variational equation.
-
-	\paragraph{Feasibility} To show that $v_*$ is feasible (i.e., $v_* \in X$), we must show that $v_* \in H^2_N(\Omega)$ and $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,\convrank$.
-We already showed that $v_* \in H^2_N(\Omega)$ in Theorem \ref{thm:wi_no_breakdown}.
-Using the the definition of $v_*$ in \eqref{eq:optimal_v} and Lemma \ref{lem:S_matrix_characterizations}, we have
-	\begin{equation*}
-		v_*(x_i) = \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \secondgreens_j(x_i)
-		= \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \pointinteractionmatrix_{ij} = \left(\pointinteractionmatrix \bm{\secondgreenscoeff} + \constcoeff \bm{1}\right)_i.
-	\end{equation*}
-	The conditions $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,\convrank$ may therefore be written in matrix form as
-	$
-		\pointinteractionmatrix \bm{\secondgreenscoeff} + \constcoeff \bm{1} = \bm{\interpolatedvalues}
-	$, and this matrix equation must hold because it is the first block row of system \eqref{eq:interpolation_linear_system} which is solved to compute $\bm{\secondgreenscoeff}$ and $\constcoeff$.
-So $v_* \in X$.
-
-	\paragraph{Optimality} Since $X$ is an affine subspace, and $X_0$ is the tangent space to $X$, any $v \in X$ may be written in the form $v = v_* + p$ for some $p \in X_0$.
-We may therefore reframe optimization problem \ref{eq:wi_optimization_problem_generic} as
-	\begin{equation}
-		\label{eq:p_optimization}
-		\min_{p \in X_0} J(p),
-	\end{equation}
-	where
-	\begin{equation*}
-		J(p) := \frac{1}{2}||\Delta (v_* + p)||^2.
-	\end{equation*}
-	Showing that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} is equivalent to showing that $0$ is the unique minimizer of \eqref{eq:p_optimization}.
-
-	By direct expansion, we have
-	\begin{equation}
-		\label{eq:Ju_expanded}
-		J(p) = J(0) + \left(\Delta u, \Delta p\right)_{L^2(\Omega)} + \frac{1}{2}||\Delta p||^2.
-	\end{equation}
-	Since $p$ has Neumann zero boundary conditions, if $\Delta p = 0$, then $p$ is constant on each connected component of $\Omega$.
-Furthermore, since $p(x_i)$ is zero for at least one point $x_i$ per connected component of $\Omega$, if $\Delta p = 0$ then $p=0$.
-Hence, by contrapositive we have
-	\begin{align*}
-		\label{eq:contrapositive_laplacian_zero}
-		p \neq 0 \quad \implies \quad& 0 < ||\Delta p||_{L^2(\Omega)}\\
-		\implies \quad& J(0) < J(p) + \left(\Delta v_*, \Delta p\right)_{L^2(\Omega)}.
-	\end{align*}
-	If $v_*$ satisfies the following variational equation,
-	\begin{equation}
-		\label{eq:variational_up}
-		(\Delta v_*, \Delta p)_{L^2(\Omega)} = 0 \quad \text{for all~} p \in X_0,
-	\end{equation} 
-	then
-	\begin{equation*}
-		p \neq 0 \quad \implies \quad J(0) < J(p).
-	\end{equation*}
-	Showing that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} therefore reduces to showing that $v_*$ satisfies variational equation \eqref{eq:variational_up}.
-
-	\paragraph{Variational equation} To show that $v_*$ satisfies variational equation \eqref{eq:variational_up}, notice that
-	\begin{align*}
-		(\Delta v_*, \Delta p)_{L^2(\Omega)} =& \sum_{i=1}^\convrank \secondgreenscoeff_i(\Delta \secondgreens_i, \Delta p)_{L^2(\Omega)} \\
-		=& \sum_{i=1}^\convrank \secondgreenscoeff_i \left(p(x_i) - \avg\left(p\right)\right) \\
-		=& -\avg\left(p\right) \sum_{i=1}^\convrank \secondgreenscoeff_i. 
-	\end{align*}
-	In the first line we used the definition of $\secondgreens_i$ and the fact that the Laplacian of a constant is zero.
-In the second line we used Lemma \ref{lem:psi_delta_eval}.
-In the third line we used the fact that $p(x_i)=0$ because $p \in X_0$.
-Hence variational equation \eqref{eq:variational_up} holds if $\sum_{i=1}^\convrank \secondgreenscoeff_i = 0$, which may be written in matrix form as $\bm{1}^T \bm{\secondgreenscoeff} = 0$.
-This must hold because it is the second block row in system \eqref{eq:interpolation_linear_system}.
-\end{proof}
-
-\begin{thm}[Poisson weighting functions are optimal]
-	\label{thm:smoothest_interpolant}
-	Let 
-	\begin{equation}
-		\label{eq:conv_kernel_pc}
-		\convkernel_x^\text{pc}(y) := \sum_{i=1}^\convrank \convkernel_i(y)w_i(x)
-	\end{equation}
-	denote the local convolution kernel for the product-convolution approximation, and let $\horizinterpolant_y^\text{pc}$ be the reinterpretation of $\convkernel_x^\text{pc}(y)$ as a function of $x$ instead of $y$.
-That is,
-	\begin{equation*}
-		\horizinterpolant_y^\text{pc}(x) := \convkernel_x^\text{pc}(y).
-	\end{equation*}
-	Then for each $y \in \Omega$, the function $\horizinterpolant_y^\text{pc}$ is the unique solution to the following optimization problem:
-	\begin{equation}
-		\label{eq:uy_optimization_problem}
-		\begin{aligned}
-			\min_{\horizinterpolant \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \horizinterpolant}_{L^2(\Omega)}^2 \\
-			\text{such that} &\quad \horizinterpolant(x_i) = \convkernel_i(y), \quad i=1,\dots,\convrank.
-		\end{aligned}
-	\end{equation}
-\end{thm}
-
-\begin{proof}
-	The optimization problem that $w_i$ solves, and optimization problem \eqref{eq:uy_optimization_problem}, both take the form of optimization problem \eqref{eq:wi_optimization_problem_generic}, but with different constraint data.
-Let $\mathcal{T}:\bm{b} \mapsto v_*$ denote the solution operator for optimization problem \eqref{eq:wi_optimization_problem_generic} as a function of the constraint data,  $\bm{b}$.
-The mapping $\mathcal{T}$ is a linear operator because the procedure we presented for solving optimization problem \eqref{eq:wi_optimization_problem_generic} is linear in the constraint data, and we proved that this procedure always yields the unique solution of optimization problem \eqref{eq:wi_optimization_problem_generic} in Theorem \ref{thm:wi_no_breakdown} and Theorem \ref{thm:wi_alg_is_correct}.
-
-	The constraint data for the optimization problem that $w_i$ solves is the unit vector $\bm{e}^{(i)}=(0,\dots,0,1,0,\dots,0) \in \mathbb{R}^\convrank$ which has $i$th component equal to one, and all other components equal to zero.
-Hence
-	\begin{equation*}
-		\label{eq:wi_via_solution_operator}
-		w_i = \mathcal{T}\left(\bm{e}^{(i)}\right).
-	\end{equation*}
-	The constraint data for optimization problem \eqref{eq:uy_optimization_problem} is the vector $(\convkernel_1(y), \convkernel_2(y), \dots, \convkernel_\convrank(y))$, so the solution, $\theta_*$, to optimization problem \eqref{eq:uy_optimization_problem} is given by
-	\begin{equation*}
-		\theta_* = \mathcal{T}\left((\convkernel_1(y), \convkernel_2(y), \dots, \convkernel_\convrank(y))\right).
-	\end{equation*}
-	Using the definition of $\horizinterpolant_y^\text{pc}(x)$ and linearity of $\mathcal{T}$, we have
-	\begin{align*}
-		\horizinterpolant_y^\text{pc}(x) &= \sum_{i=1}^\convrank \convkernel_i(y)w_i(x) \\
-		&= \sum_{i=1}^\convrank \convkernel_i(y)\mathcal{T}\left(\bm{e}^{(i)}\right)(x) \\
-		&= \mathcal{T}\left(\sum_{i=1}^\convrank \convkernel_i(y) \bm{e}^{(i)} \right)(x) \\
-		&= \mathcal{T}\left((\convkernel_1(y), \convkernel_2(y), \dots, \convkernel_\convrank(y))\right)(x) = \theta_*(x).
-	\end{align*}
-	Thus $\horizinterpolant_y^\text{pc}$ is the unique solution to optimization problem \eqref{eq:uy_optimization_problem}.
-\end{proof}
-
-\begin{prop}[Poisson weighting functions sum to one]
-	\label{thm:wi_sum_to_one}
-	We have
-	\begin{equation*}
-		\sum_{i=1}^\convrank w_i(x) = 1
-	\end{equation*}
-	for all $x \in \Omega$.
-\end{prop}
-
-\begin{algorithm2e}
-	\SetAlgoNoLine
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Output}
-	{	
-		\Input{Sample points $\{x_i\}_{i=1}^\convrank$, domain geometry $\Omega$}
-		\Output{Weighting functions $\{w_i\}_{i=1}^\convrank$}
-
-		\For{$k=1,2, \dots, \convrank$}{
-
-			Solve the Neumann Poisson problem \eqref{eq:psi_eq} for $\firstgreens_k$.
-
-			Solve the Neumann Poisson problem \eqref{eq:theta_eq} for $\secondgreens_k$.
-
-		}
-
-		Form the $\convrank \times \convrank$ matrix $\pointinteractionmatrix$ given in \eqref{eq:S_matrix}.
-
-		\For{$i=1,2,\dots,\convrank$}{
-			Solve the linear system
-			\begin{equation*}
-			\begin{bmatrix}
-			\pointinteractionmatrix & \bm{1} \\ \bm{1}^T & 0
-			\end{bmatrix}
-			\begin{bmatrix}
-			\bm{\secondgreenscoeff}^{(i)} \\ \constcoeff^{(i)}
-			\end{bmatrix}
-			=
-			\begin{bmatrix}
-			\bm{e}^{(i)} \\ 0
-			\end{bmatrix}.
-			\end{equation*}
-			for $\bm{\secondgreenscoeff}^{(i)}=(\secondgreenscoeff_1^{(i)}, \secondgreenscoeff_2^{(i)}, \dots, \secondgreenscoeff_\convrank^{(i)})$ and $\constcoeff^{(i)}$.
-Here $\bm{1} = (1,1,\dots,1) \in \mathbb{R}^\convrank$, and $\bm{e}^{(i)} = (0,\dots,0,1,0,\dots,0)\in \mathbb{R}^\convrank$ is the unit vector with $i^\text{th}$ component equal to one and all other components equal to zero.
-
-			Form the weighting function $w_i$ given by
-			\begin{equation*}
-			w_i(x) = \constcoeff^{(i)} + \sum_{k=1}^\convrank \secondgreenscoeff_k^{(i)} \secondgreens_k(x).
-			\end{equation*}
-
-		}
-	}
-	\caption{Compute weighting functions $w_i$, $i=1,\dots,\convrank$.}
-	\label{alg:weighting_functions_incremental}
-\end{algorithm2e}
-
 \subsection{Discretization and hierarchical matrix construction}
@@
 
+\section{OLD CONTENT}
+
+\subsection{Product-convolution approximation}
+\label{sec:prodconv_intro}
+Let 
+\begin{equation}
+\label{eq:convolution_kernel}
+\convkernel_x(y) := \impulseresponse_x(y+x)
+\end{equation}
+be the translation of $\impulseresponse_x$ to re-center it at zero instead of $x$.
+If $\Aop$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then for any point $x_0$ we would have
+\begin{align*}
+\Aker(y,x) &= \Aker(y+(x_0-x),x+(x_0-x)) \\
+&= \phi_{x_0}(y-x+x_0) \\
+&= \convkernel_{x_0}(y-x)
+\end{align*}
+which would imply
+\begin{align*}
+(\Aop u)(v) &= \int_\Omega \int_\Omega v(y) \convkernel_{x_0}(y-x) u(x) dx dy \\
+&= \int_\Omega v(y) \left(\int_\Omega \convkernel_{x_0}(y-x) u(x) dx\right) dy \\
+&= \left(v, \convkernel_{x_0} \ast u\right)_{L^2(\Omega)},
+\end{align*}
+where $f \ast g$ denotes convolution of the functions $f$ and $g$, and $\left(f, g\right)_{L^2(\Omega)} := \int_\Omega f(x) g(x) dx$ denotes the $L^2$ inner product of $f$ and $g$ on $\Omega$.
+Hence, if $\Aop$ were perfectly translation invariant, it would be the convolution operator $u \mapsto \left(\convkernel_{x_0} \ast u\right)^*$.
+
+Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\convkernel_x$, varies as $x$ changes.
+We therefore approximate $\Aop$ by a spatially varying weighted sum of convolution operators, $\AopPc \approx \Aop$, of the form
+\begin{equation}
+\left(\AopPc~u\right)(v) := \left( v,~ \sum_{i=1}^\convrank \convkernel_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
+\end{equation}
+where $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$ (i.e., $\left(f \cdot g\right)(x) = f(x) g(x)$).
+Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution.
+To avoid nested subscripts, we write
+\begin{equation*}
+\convkernel_i := \convkernel_{x_i} \quad \text{and} \quad \impulseresponse_i := \impulseresponse_{x_i},
+\end{equation*}
+with $\convkernel_{x_i}$ defined in \eqref{eq:convolution_kernel} and $\impulseresponse_{x_i}$ defined in \eqref{eq:impulse_response_defn}, to denote local convolution kernels and impulse responses corresponding to a collection of points $\{x_i\}_{i=1}^\convrank \subset \Omega$ scattered throughout the domain.
+The functions $w_i$ are spatially varying weighting functions that are used to interpolate the convolution kernels $\convkernel_i$.
+Our operator approximation method is defined by how we compute the convolution kernels, $\convkernel_i$, how we choose the points, $x_i$, and what weighting functions, $w_i$, that we use.
+In Section \ref{sec:overview_intro} we summarize the answers to these questions; we will provide detailed answers to these questions in Section \ref{sec:method}.
+
+The more locally translation invariant an operator is (i.e., the smaller the discrepancy between the left hand side and right hand side in \eqref{eq:local_translation_invariance}), the smaller the number of terms that are required in \eqref{eq:product_convolution} to achieve an accurate product-convolution approximation.
+
+\subsection{Overview of the method (old)}
+\label{sec:overview_intro}
+
+We compute one ``batch'' of $\convkernel_i$'s by applying $\Aop$ to a sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$.
+The batch of points $x_i$ are chosen so that the support of $\impulseresponse_i$ and the support of $\impulseresponse_j$ do not overlap (or do not overlap much) if $i \neq j$.
+Because these supports do not overlap, we can post-process the response of $\Aop$ to the Dirac comb to recover the functions $\impulseresponse_i$, and therefore the functions $\convkernel_i$, associated with all points $x_i$ in the batch---with one application of $\Aop$, we recover many $\convkernel_i$.
+This is illustrated in Figure FIG.
+The process is repeated to get more batches of $\convkernel_i$'s, until a desired number of batches is reached.
+
+In order to choose the points $x_i$, we need to estimate the support of $\impulseresponse_i$ \emph{before} we compute it.
+The supports of the functions $\impulseresponse_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of specially chosen functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$), then post processing the results (Section \ref{eq:mean_and_covariance_estimation}).
+This procedure is the only part of the paper that requires non-negativity of the integral kernel.
+
+We choose the $i$th weighting function, $w_i$, to be the flattest function, in a least-squares sense, that takes the value one at the sample point $x_i$, and the value zero at the other sample points $x_j$, $j \neq i$.
+Computing the weighting functions $w_i$ requires solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid.
+We call this interpolation scheme \emph{Poisson interpolation}.
+The advantage of Poisson interpolation over other interpolation methods, most notably radial basis function interpolation \cite{ESCANDE}, is that Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions.
+This is illustrated in Figure FIG.
+
+Once the $\convkernel_i$ and $w_i$ are computed, we convert $\AopPc$ to hierarchical matrix (H-matrix) format.
+Let $\Aker_\text{pc}$ denote the integral kernel associated with $\AopPc$.
+While $\Aker(y,x)$ is not easily computable, $\Aker_\text{pc}(y,x)$ is given by the formula
+\begin{equation}
+\label{eq:kernel_entries}
+\AkerPc(y,x) = \sum_{i=1}^\convrank \convkernel_i(y-x) w_i(x).
+\end{equation}
+This follows from~\eqref{eq:product_convolution} and the fact that the convolution operator $u \mapsto \convkernel_i \ast u$ has $(y,x)$ kernel entry given by $\convkernel(y-x)$.
+Formula~\eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\AopPc$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
+Once in H-matrix format, fast H-matrix arithmetic is used to invert $\AopPc$, or perform other useful matrix operations.
+
+Often we are interested in approximating operators $\Aop$ that are symmetric positive semi-definite.
+Unfortunately, $\AopPc$ is, in general, non-symmetric and indefinite due to errors in the product convolution approximation.
+In this case, we modify the H-matrix representation of $\AopPc$ to make it positive definite, by symmetrizing it, then applying a specially chosen rational matrix function to the symmetrized H-matrix.
+
+\subsection{Poisson weighting functions}
+\label{sec:weighting_functions}
+
+Consider the following optimization problem,
+\begin{equation}
+\label{eq:wi_optimization_problem_generic}
+\begin{aligned}
+\min_{v \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta v}_{L^2(\Omega)}^2 \\
+\text{such that} &\quad v(x_j) = \interpolatedvalues_j, \quad j=1,\dots,\convrank,
+\end{aligned}
+\end{equation}
+in which one seeks a smooth function $v$ that takes the values $\interpolatedvalues_j \in \mathbb{R}$ at the points $x_j$.
+Here $\Delta$ is the Laplacian on $\Omega$ and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions.
+We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to optimization problem \eqref{eq:wi_optimization_problem_generic}, with
+\begin{equation*}
+b_j = w_i(x_j) = 
+\begin{cases}
+1, & j=i\\
+0, & j \neq i.
+\end{cases}
+\end{equation*}
+In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' \cite{JACOBSEN} interpolation of the convolution kernels, $\convkernel_i$, in a least-squares sense.
+Also, these weighting functions sum to one pointwise (Proposition \ref{thm:wi_sum_to_one}).
+
+The solution to optimization problem \eqref{eq:wi_optimization_problem_generic} is given via the following procedure:
+\begin{enumerate}
+	\item For each sample point $x_i$, we compute a function $\firstgreens_i$ by solving the following Neumann Poisson problem:
+	\begin{equation}
+	\label{eq:psi_eq}
+	\begin{cases}
+	\Delta \firstgreens_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
+	\nu \cdot \nabla \firstgreens_i = 0 & \text{on } \partial \Omega,
+	\end{cases}
+	\end{equation}
+	with condition $\int_\Omega \firstgreens_i(x) dx = 0$.
+Here $\delta_{x_i}$ is the point source centered at $x_i$, $|\Omega|$ is the measure of the domain $\Omega$, and $\nu$ is the normal vector to $\partial \Omega$.
+	\item For each function $\firstgreens_i$, we compute a function $\secondgreens_i$ by solving another Neumann Poisson problem:
+	\begin{equation}
+	\label{eq:theta_eq}
+	\begin{cases}
+	\Delta \secondgreens_i = \firstgreens_i, & \text{in }\Omega, \\
+	\nu \cdot \nabla \secondgreens_i = 0 & \text{on } \partial \Omega,
+	\end{cases}
+	\end{equation}
+	with condition $\int_\Omega \secondgreens_i(x) dx = 0$.
+	\item The solution to optimization problem \eqref{eq:wi_optimization_problem_generic}, which we denote by $v_*$, is formed as follows:
+	\begin{equation}
+	\label{eq:optimal_v}
+	v_*(x) := \constcoeff + \sum_{i=1}^\convrank \secondgreenscoeff_i \secondgreens_i(x),
+	\end{equation}
+	where the constants $\secondgreenscoeff_i \in \mathbb{R}$, $i=1,\dots,\convrank$, and $\constcoeff \in \mathbb{R}$ are the solution to the following linear system:
+	\begin{equation}
+	\label{eq:interpolation_linear_system}
+	\begin{bmatrix}
+	\pointinteractionmatrix & \bm{1} \\ \bm{1}^T & 0
+	\end{bmatrix}
+	\begin{bmatrix}
+	\bm{\secondgreenscoeff} \\ \constcoeff
+	\end{bmatrix}
+	=
+	\begin{bmatrix}
+	\bm{\interpolatedvalues} \\ 0
+	\end{bmatrix}.
+	\end{equation}
+	Here $\bm{\secondgreenscoeff} \in \mathbb{R}^\convrank$ is the vector with $i$th component $\secondgreenscoeff_i$, $\bm{1} \in \bm{R}^\convrank$ is the column vector of all ones, $\bm{\interpolatedvalues} \in \mathbb{R}^\convrank$ is the vector with $i$th component $\interpolatedvalues_i$, and $\pointinteractionmatrix \in \mathbb{R}^{\convrank \times \convrank}$ is the matrix with entries given by
+	\begin{equation}
+	\label{eq:S_matrix}
+	\pointinteractionmatrix_{ij} := \left( \firstgreens_i, \firstgreens_j \right)_{L^2(\Omega)}.
+	\end{equation}
+\end{enumerate}
+Linear system \eqref{eq:interpolation_linear_system} is the KKT system that arises if one solves optimization problem \eqref{eq:wi_optimization_problem_generic} on the $\convrank+1$-dimensional subspace of $H^2_N(\Omega)$ spanned by the functions $\secondgreens_i$, $i=1,\dots,\convrank$, and the constant function $C(x)=1$.
+In Theorem \ref{thm:wi_no_breakdown}, we prove that this procedure does not break down, and in Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v_*$ produced by this procedure is the solution to optimization problem \eqref{eq:wi_optimization_problem_generic} on $H^2_N(\Omega)$.
+
+Constructing all $\convrank$ of the functions $w_i$ thus requires solving the Poisson equation $2\convrank$ times with different right hand side sources to form the functions $\firstgreens_i$ and $\secondgreens_i$, then post-processing these functions.
+This is shown in Algorithm \ref{alg:weighting_functions_incremental}.
+The Poisson PDE solves, which are the most expensive part of this procedure, can be performed cheaply with multigrid.
+
+It is important to note that functions in $H^2_N(\Omega)$ only have well-defined pointwise values if the spatial dimension of the domain is $\gdim=1$, $2$, or $3$.
+Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-posed in three spatial dimensions or fewer.
+In four or higher spatial dimensions, one could construct appropriate weighting functions using radial basis functions, though we do not discuss this here.
+
+\begin{thm}[Procedure for computing $v_*$ does not break down]
+	\label{thm:wi_no_breakdown}
+	If all the sample points $x_i$, $i=1,\dots,\convrank$, are distinct, then PDE \eqref{eq:psi_eq} is uniquely solvable for $\firstgreens_i$, PDE \eqref{eq:theta_eq} is uniquely solvable for $\secondgreens_i$, and linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\bm{\secondgreenscoeff}$ and $\constcoeff$.
+That is, the procedure for computing $v_*$ does not break down.
+Furthermore, $v_* \in H^2_N(\Omega)$.
+\end{thm}
+
+\begin{proof}
+	Existence and uniqueness for the solutions to PDEs \eqref{eq:psi_eq} and \eqref{eq:theta_eq} follow from the standard elliptic PDE theory.
+By standard elliptic regularity theory, $\firstgreens_i \in H^1(\Omega)$, so $\firstgreens_i \in L^2(\Omega)$, and so all matrix entries $\pointinteractionmatrix_{ij}$ are finite.
+
+	Using the definition of $\pointinteractionmatrix$ and basic linear algebra, it is straightforward to show that if $\pointinteractionmatrix$ were singular, there would exist constants $a_1, a_2, \dots a_\convrank$, which are not all zero, such that
+	\begin{equation*}
+	0 = a_1 \firstgreens_1 + a_2 \firstgreens_2 + \dots a_\convrank \firstgreens_\convrank.
+	\end{equation*}
+	Applying the Laplacian to this equation yields
+	\begin{equation*}
+	0 = a_1 \delta_{x_1} + a_2 \delta_{x_2} + \dots a_\convrank \delta_{x_\convrank} - \left(\sum_{i=1}^\convrank a_i\right)\bigg/|\Omega|,
+	\end{equation*}
+	which is impossible because a nonzero linear combination of delta distributions at different points cannot equal a constant function.
+Thus we must conclude that $\pointinteractionmatrix$ is non-singular.
+By block matrix factorization, non-singularity of $\pointinteractionmatrix$ implies that linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\bm{\secondgreenscoeff}$ and $\constcoeff$.
+
+	By standard elliptic regularity theory, the functions $\secondgreens_i$ are in $H^3(\Omega)$, and therefore they are in $H^2(\Omega)$.
+By construction the functions $\secondgreens_i$ have Neumann zero boundary conditions, so $\secondgreens_i \in H^2_N(\Omega)$.
+Since $v_*$ is a linear combination of the functions $\secondgreens_i$ and the constant function $C(x)=1$ which is in $H^2_N(\Omega)$, we have $v_* \in H^2_N(\Omega)$.
+\end{proof}
+
+\begin{lem}[Integral representation of delta distribution on $H^2_N(\Omega)$]
+	\label{lem:psi_delta_eval}
+	We have
+	\begin{equation}
+	\label{eq:inline_wi_lemma}
+	h(x_j) = \left(\firstgreens_j, \Delta h\right)_{L^2(\Omega)} + \avg(h)
+	\end{equation}
+	for all $h\in H^2_N(\Omega)$, where
+	\begin{equation*}
+	\avg(h) :=	 \frac{1}{|\Omega|} \int_\Omega h ~dx,
+	\end{equation*}
+	and $|\Omega| = \int_\Omega 1 ~dx$ is the Lebesgue measure of $\Omega$.
+\end{lem}
+
+\begin{proof}
+	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\firstgreens_j$ onto $h$, and using the Neumann zero boundary conditions for $\firstgreens_j$ and $h$ to eliminate boundary integral terms.
+\end{proof}
+
+\begin{lem}[Characterization of $\pointinteractionmatrix$]
+	\label{lem:S_matrix_characterizations}
+	We have
+	\begin{equation*}
+	\pointinteractionmatrix_{ij} = \secondgreens_j(x_i).
+	\end{equation*}
+\end{lem}
+
+\begin{proof}
+	We compute
+	\begin{equation*}
+	\secondgreens_j(x_i) = \left(\firstgreens_i, \Delta \secondgreens_j\right)_{L^2(\Omega)} + \avg\left(\secondgreens_j\right)
+	= \left(\firstgreens_i, \firstgreens_j\right)_{L^2(\Omega)} = \pointinteractionmatrix_{ij}.
+	\end{equation*}
+	In the first equality we used Lemma \ref{lem:psi_delta_eval}.
+In the second equality we used the facts that $\Delta \secondgreens_j = \firstgreens_j$ and $\avg(\secondgreens_j) = \left(1 / |\Omega|\right) \int_\Omega \secondgreens_j = 0$ from the definition of $\secondgreens_j$ in \eqref{eq:theta_eq}.
+\end{proof}
+
+\begin{thm}[Solution of optimization problem \eqref{eq:wi_optimization_problem_generic}]
+	\label{thm:wi_alg_is_correct}
+	The function $v_*$ given in \eqref{eq:optimal_v} is the unique solution to optimization problem \eqref{eq:wi_optimization_problem_generic}.
+\end{thm}
+
+\begin{proof}
+	Let
+	\begin{equation*}
+	X := \{v \in H^2_N(\Omega) : v(x_i) = \interpolatedvalues_i \text{ for }i=1,\dots,\convrank\}
+	\end{equation*}
+	denote the feasible set for optimization problem \ref{eq:wi_optimization_problem_generic}, and let 
+	\begin{equation*}
+	X_0 := \{v \in H^2_N(\Omega) : v(x_i) = 0 \text{ for }i=1,\dots,\convrank\}
+	\end{equation*}	
+	denote the tangent space to $X$.
+Note that pointwise values of functions in $H_N^2(\Omega)$ are well-defined since the Sobolev imbedding theorem (see, e.g., Theorem 7.19 in \cite{Arbogast}) implies that $H^2(\Omega)$ is continuously imbedded in the space of continuous functions $C^0(\Omega)$ when $\gdim < 4$, and we are considering $\gdim \in \{1,2,3\}$.
+The feasible set $X$ is therefore a closed affine subspace of $H^2_N(\Omega)$, the tangent space $X_0$ is a closed subspace of $H^2_N(\Omega)$, and optimization problem \eqref{eq:wi_optimization_problem_generic} is well-defined.
+
+	The proof now proceeds in three steps.
+First, we show that $v_*$ is feasible ($v_* \in X$).
+Second, we show that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} if it satisfies a certain variational equation.
+Third, we verify that $v_*$ satisfies this variational equation.
+
+	\paragraph{Feasibility} To show that $v_*$ is feasible (i.e., $v_* \in X$), we must show that $v_* \in H^2_N(\Omega)$ and $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,\convrank$.
+We already showed that $v_* \in H^2_N(\Omega)$ in Theorem \ref{thm:wi_no_breakdown}.
+Using the the definition of $v_*$ in \eqref{eq:optimal_v} and Lemma \ref{lem:S_matrix_characterizations}, we have
+	\begin{equation*}
+	v_*(x_i) = \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \secondgreens_j(x_i)
+	= \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \pointinteractionmatrix_{ij} = \left(\pointinteractionmatrix \bm{\secondgreenscoeff} + \constcoeff \bm{1}\right)_i.
+	\end{equation*}
+	The conditions $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,\convrank$ may therefore be written in matrix form as
+	$
+	\pointinteractionmatrix \bm{\secondgreenscoeff} + \constcoeff \bm{1} = \bm{\interpolatedvalues}
+	$, and this matrix equation must hold because it is the first block row of system \eqref{eq:interpolation_linear_system} which is solved to compute $\bm{\secondgreenscoeff}$ and $\constcoeff$.
+So $v_* \in X$.
+
+	\paragraph{Optimality} Since $X$ is an affine subspace, and $X_0$ is the tangent space to $X$, any $v \in X$ may be written in the form $v = v_* + p$ for some $p \in X_0$.
+We may therefore reframe optimization problem \ref{eq:wi_optimization_problem_generic} as
+	\begin{equation}
+	\label{eq:p_optimization}
+	\min_{p \in X_0} J(p),
+	\end{equation}
+	where
+	\begin{equation*}
+	J(p) := \frac{1}{2}||\Delta (v_* + p)||^2.
+	\end{equation*}
+	Showing that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} is equivalent to showing that $0$ is the unique minimizer of \eqref{eq:p_optimization}.
+
+	By direct expansion, we have
+	\begin{equation}
+	\label{eq:Ju_expanded}
+	J(p) = J(0) + \left(\Delta u, \Delta p\right)_{L^2(\Omega)} + \frac{1}{2}||\Delta p||^2.
+	\end{equation}
+	Since $p$ has Neumann zero boundary conditions, if $\Delta p = 0$, then $p$ is constant on each connected component of $\Omega$.
+Furthermore, since $p(x_i)$ is zero for at least one point $x_i$ per connected component of $\Omega$, if $\Delta p = 0$ then $p=0$.
+Hence, by contrapositive we have
+	\begin{align*}
+	\label{eq:contrapositive_laplacian_zero}
+	p \neq 0 \quad \implies \quad& 0 < ||\Delta p||_{L^2(\Omega)}\\
+	\implies \quad& J(0) < J(p) + \left(\Delta v_*, \Delta p\right)_{L^2(\Omega)}.
+	\end{align*}
+	If $v_*$ satisfies the following variational equation,
+	\begin{equation}
+	\label{eq:variational_up}
+	(\Delta v_*, \Delta p)_{L^2(\Omega)} = 0 \quad \text{for all~} p \in X_0,
+	\end{equation} 
+	then
+	\begin{equation*}
+	p \neq 0 \quad \implies \quad J(0) < J(p).
+	\end{equation*}
+	Showing that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} therefore reduces to showing that $v_*$ satisfies variational equation \eqref{eq:variational_up}.
+
+	\paragraph{Variational equation} To show that $v_*$ satisfies variational equation \eqref{eq:variational_up}, notice that
+	\begin{align*}
+	(\Delta v_*, \Delta p)_{L^2(\Omega)} =& \sum_{i=1}^\convrank \secondgreenscoeff_i(\Delta \secondgreens_i, \Delta p)_{L^2(\Omega)} \\
+	=& \sum_{i=1}^\convrank \secondgreenscoeff_i \left(p(x_i) - \avg\left(p\right)\right) \\
+	=& -\avg\left(p\right) \sum_{i=1}^\convrank \secondgreenscoeff_i. 
+	\end{align*}
+	In the first line we used the definition of $\secondgreens_i$ and the fact that the Laplacian of a constant is zero.
+In the second line we used Lemma \ref{lem:psi_delta_eval}.
+In the third line we used the fact that $p(x_i)=0$ because $p \in X_0$.
+Hence variational equation \eqref{eq:variational_up} holds if $\sum_{i=1}^\convrank \secondgreenscoeff_i = 0$, which may be written in matrix form as $\bm{1}^T \bm{\secondgreenscoeff} = 0$.
+This must hold because it is the second block row in system \eqref{eq:interpolation_linear_system}.
+\end{proof}
+
+\begin{thm}[Poisson weighting functions are optimal]
+	\label{thm:smoothest_interpolant}
+	Let 
+	\begin{equation}
+	\label{eq:conv_kernel_pc}
+	\convkernel_x^\text{pc}(y) := \sum_{i=1}^\convrank \convkernel_i(y)w_i(x)
+	\end{equation}
+	denote the local convolution kernel for the product-convolution approximation, and let $\horizinterpolant_y^\text{pc}$ be the reinterpretation of $\convkernel_x^\text{pc}(y)$ as a function of $x$ instead of $y$.
+That is,
+	\begin{equation*}
+	\horizinterpolant_y^\text{pc}(x) := \convkernel_x^\text{pc}(y).
+	\end{equation*}
+	Then for each $y \in \Omega$, the function $\horizinterpolant_y^\text{pc}$ is the unique solution to the following optimization problem:
+	\begin{equation}
+	\label{eq:uy_optimization_problem}
+	\begin{aligned}
+	\min_{\horizinterpolant \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \horizinterpolant}_{L^2(\Omega)}^2 \\
+	\text{such that} &\quad \horizinterpolant(x_i) = \convkernel_i(y), \quad i=1,\dots,\convrank.
+	\end{aligned}
+	\end{equation}
+\end{thm}
+
+\begin{proof}
+	The optimization problem that $w_i$ solves, and optimization problem \eqref{eq:uy_optimization_problem}, both take the form of optimization problem \eqref{eq:wi_optimization_problem_generic}, but with different constraint data.
+Let $\mathcal{T}:\bm{b} \mapsto v_*$ denote the solution operator for optimization problem \eqref{eq:wi_optimization_problem_generic} as a function of the constraint data,  $\bm{b}$.
+The mapping $\mathcal{T}$ is a linear operator because the procedure we presented for solving optimization problem \eqref{eq:wi_optimization_problem_generic} is linear in the constraint data, and we proved that this procedure always yields the unique solution of optimization problem \eqref{eq:wi_optimization_problem_generic} in Theorem \ref{thm:wi_no_breakdown} and Theorem \ref{thm:wi_alg_is_correct}.
+
+	The constraint data for the optimization problem that $w_i$ solves is the unit vector $\bm{e}^{(i)}=(0,\dots,0,1,0,\dots,0) \in \mathbb{R}^\convrank$ which has $i$th component equal to one, and all other components equal to zero.
+Hence
+	\begin{equation*}
+	\label{eq:wi_via_solution_operator}
+	w_i = \mathcal{T}\left(\bm{e}^{(i)}\right).
+	\end{equation*}
+	The constraint data for optimization problem \eqref{eq:uy_optimization_problem} is the vector $(\convkernel_1(y), \convkernel_2(y), \dots, \convkernel_\convrank(y))$, so the solution, $\theta_*$, to optimization problem \eqref{eq:uy_optimization_problem} is given by
+	\begin{equation*}
+	\theta_* = \mathcal{T}\left((\convkernel_1(y), \convkernel_2(y), \dots, \convkernel_\convrank(y))\right).
+	\end{equation*}
+	Using the definition of $\horizinterpolant_y^\text{pc}(x)$ and linearity of $\mathcal{T}$, we have
+	\begin{align*}
+	\horizinterpolant_y^\text{pc}(x) &= \sum_{i=1}^\convrank \convkernel_i(y)w_i(x) \\
+	&= \sum_{i=1}^\convrank \convkernel_i(y)\mathcal{T}\left(\bm{e}^{(i)}\right)(x) \\
+	&= \mathcal{T}\left(\sum_{i=1}^\convrank \convkernel_i(y) \bm{e}^{(i)} \right)(x) \\
+	&= \mathcal{T}\left((\convkernel_1(y), \convkernel_2(y), \dots, \convkernel_\convrank(y))\right)(x) = \theta_*(x).
+	\end{align*}
+	Thus $\horizinterpolant_y^\text{pc}$ is the unique solution to optimization problem \eqref{eq:uy_optimization_problem}.
+\end{proof}
+
+\begin{prop}[Poisson weighting functions sum to one]
+	\label{thm:wi_sum_to_one}
+	We have
+	\begin{equation*}
+	\sum_{i=1}^\convrank w_i(x) = 1
+	\end{equation*}
+	for all $x \in \Omega$.
+\end{prop}
+
+\begin{algorithm2e}
+	\SetAlgoNoLine
+	\SetKwInOut{Input}{Input}
+	\SetKwInOut{Output}{Output}
+	{	
+		\Input{Sample points $\{x_i\}_{i=1}^\convrank$, domain geometry $\Omega$}
+		\Output{Weighting functions $\{w_i\}_{i=1}^\convrank$}
+
+		\For{$k=1,2, \dots, \convrank$}{
+
+			Solve the Neumann Poisson problem \eqref{eq:psi_eq} for $\firstgreens_k$.
+
+			Solve the Neumann Poisson problem \eqref{eq:theta_eq} for $\secondgreens_k$.
+
+		}
+
+		Form the $\convrank \times \convrank$ matrix $\pointinteractionmatrix$ given in \eqref{eq:S_matrix}.
+
+		\For{$i=1,2,\dots,\convrank$}{
+			Solve the linear system
+			\begin{equation*}
+			\begin{bmatrix}
+			\pointinteractionmatrix & \bm{1} \\ \bm{1}^T & 0
+			\end{bmatrix}
+			\begin{bmatrix}
+			\bm{\secondgreenscoeff}^{(i)} \\ \constcoeff^{(i)}
+			\end{bmatrix}
+			=
+			\begin{bmatrix}
+			\bm{e}^{(i)} \\ 0
+			\end{bmatrix}.
+			\end{equation*}
+			for $\bm{\secondgreenscoeff}^{(i)}=(\secondgreenscoeff_1^{(i)}, \secondgreenscoeff_2^{(i)}, \dots, \secondgreenscoeff_\convrank^{(i)})$ and $\constcoeff^{(i)}$.
+Here $\bm{1} = (1,1,\dots,1) \in \mathbb{R}^\convrank$, and $\bm{e}^{(i)} = (0,\dots,0,1,0,\dots,0)\in \mathbb{R}^\convrank$ is the unit vector with $i^\text{th}$ component equal to one and all other components equal to zero.
+
+			Form the weighting function $w_i$ given by
+			\begin{equation*}
+			w_i(x) = \constcoeff^{(i)} + \sum_{k=1}^\convrank \secondgreenscoeff_k^{(i)} \secondgreens_k(x).
+			\end{equation*}
+
+		}
+	}
+	\caption{Compute weighting functions $w_i$, $i=1,\dots,\convrank$.}
+	\label{alg:weighting_functions_incremental}
+\end{algorithm2e}
+
+\subsubsection{Interpolation onto local regular grids}
+\label{sec:convolution_kernel_regular_grid}
+
+To convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), we will need to evaluate $\convkernel_{x_i}(z)$ for a large number of points $z$.
+Per \eqref{eq:varphi_eval}, evaluating $\convkernel_{x_i}(z)$ requires evaluating $\combresponse_k(z+x_i)$.
+In computations, $\combresponse_k$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh.
+Evaluating $\combresponse_k(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell.
+Although this procedure is asymptotically scalable, it is slow in practice \cite{FEMEVALSLOW}.
+
+To speed up evaluation of $\convkernel_{x_i}$, we interpolate $\convkernel_{x_i}$ onto a regular rectilinear grid on a coordinate axis aligned box that just barely contains the ellipsoid $E_{x_i}$.
+We transfer the function $\convkernel_{x_i}$ from the irregular mesh to the regular grid by evaluating $\convkernel_{x_i}$ at all gridpoints in the regular grid, using the slow process described in the previous paragraph.
+This is an upfront cost which is done ``offline.'' After $\convkernel_{x_i}$ has been transferred to the regular grid, we use fast regular grid interpolation to perform all subsequent ``online'' evaluations of $\convkernel_{x_i}$.
+This yields a substantial overall speedup because the number of finite element function evaluations required to transfer the functions $\convkernel_{x_i}$ to the regular grid is $O(\fedim)$, while the number of kernel evaluations required to convert $\AopPc$ to H-matrix format is $O(k \fedim \log \fedim)$.
+Here $\fedim$ is the number of degrees of freedom in the finite element space used to discretize $\combresponse_k$, and $\hmatrixrank$ is the H-matrix rank.
+Furthermore, transferring $\convkernel_{x_i}$ to a regular grid requires evaluating $\convkernel_{x_i}$ at a predetermined and highly structured collection of points.
+This structure allows for faster determination of which mesh cell that a point is in, and allows one to evaluate the finite element function at all regular gridpoints in parallel.
+In contrast, constructing the H-matrix requires evaluating $\convkernel_{x_i}$ at an unpredictable sequence of scattered points, which is an expensive task for finite element function evaluation, but a cheap task for regular grid interpolation.
+Having $\convkernel_{x_i}$ stored on a regular grid also simplifies the implementation of the method that we use to handle boundary artifacts, which will be described in Section \ref{sec:boundary_extension}.
+
+We incur interpolation error when transferring $\convkernel_{x_i}$ from the irregular mesh to the regular grid.
+To limit this interpolation error, the density of the regular grid should be as dense or denser than the local density of the finite element mesh.
+In our numerical results, the distance between neighboring points in the regular grid is chosen to be half the minimum distance between any pair of finite element Lagrange nodes that are contained in $E_{x_i}$.
+The grid transfer procedure is illustrated in Figure FIG.
+
+When $x_i$ is near a boundary, it may occur that some gridpoints for the regular grid are in locations where $\convkernel_{x_i}$ is undefined.
+In this case, we store a placeholder value for $\convkernel_{x_i}$ at those gridpoints.
+We use the IEEE standard value of ``NaN'' (not a number) for this placeholder.
+In the next section (Section \ref{sec:boundary_extension}) we replace these NaN values with more appropriate values.
+
+\subsubsection{Addressing boundary artifacts}
+\label{sec:boundary_extension}
+
+Evaluating the kernel for our product-convolution approximation, $\AkerPc(y,x)$, requires evaluating $\convkernel_{x_i}(y-x)$.
+But $\convkernel_{x_i}(y-x)$ is undefined  when $y - x + x_i$ is outside $\Omega$, which may occur even if $x \in \Omega$ and $y \in \Omega$.
+The default approach to deal with this problem is to extend $\convkernel_{x_i}$ by zero wherever it is undefined.
+But this leads to substantial errors, known as boundary artifacts, which are especially significant when the sample point $x_i$ is near the boundary of the domain.
+In this case, the impulse response is artificially chopped off by the domain boundary.
+See Figure FIG for an illustration.
+
+While $\convkernel_{x_i}$ may be undefined at a point $z$, often there is a sample point $x_j$ that is near $x_i$, but further away from the boundary than $x_i$, such that $\convkernel_{x_j}$ is well defined at $z$.
+To address boundary artifacts, we fill in missing/undefined portions of convolution kernels using information from other convolution kernels associated with the $k$ nearest sample points (we use $k=8$ in our numerical results).
+Kernels associated with nearer sample points are given precedence over kernels associated with farther away sample points.
+
+In detail, let $p_1, p_2, \dots, p_k$ be the $k$ nearest sample points to $x_i$, ordered by nearness, including $x_i$ itself.
+That is, $p_1=x_i$, $p_2$ is the nearest neighbor to $x_i$, $p_3$ is the second nearest neighbor to $x_i$, and so on.
+Also, recall from the previous section (Section \ref{sec:convolution_kernel_regular_grid}) that the functions $\convkernel_{x_i}$ are stored on regular rectilinear grids.
+To extend $\convkernel_{x_i}$, we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\convkernel_{x_i}$, and contains that grid as a subgrid, but is large enough to contain the boxes associated with all of the neighbor kernels, $\convkernel_{p_1}, \convkernel_{p_2}, \dots, \convkernel_{p_k}$ (recall that the functions $\convkernel_{x_i}$ are all centered at zero).
+We initialize an extended convolution kernel, $\convkernel_{x_i}^E$, to zero on this expanded grid.
+Then we interpolate the functions $\convkernel_{p_j}$ onto this expanded grid in reverse order.
+First we interpolate $\convkernel_{p_k}$ onto the grid, then $\convkernel_{p_{k-1}}$, and so on until the last step in which we interpolate the original kernel $\convkernel_{p_1} = \convkernel_{x_i}$.
+At each step, we overwrite existing values of $\convkernel_{x_i}^E$ with new values from the kernel currently being interpolated, except when the new value of $\convkernel_{p_j}$ is undefined (``NaN'').
+At gridpoints where the kernel being interpolated is undefined, we leave $\convkernel_{x_i}^E$ unchanged.
+In this way, we fill in undefined values of $\convkernel_{x_i}$ with values from successively further and further away kernels.
+This process is illustrated in Figure FIG.
+To avoid notational clutter, and since we will henceforth only work with the extended kernel, $\convkernel_{x_i}^E$, in the remainder of this paper the symbol $\convkernel_{x_i}$ refers to $\convkernel_{x_i}^E$.
+
 \section*{Acknowledgments}

### 2022-03-01 paper/localpsf.tex
moved old stuff to end. rearranging sections

@@
 
-\subsubsection{Estimating impulse response supports}
+\subsubsection{Impulse response moments and support estimation}
 \label{eq:mean_and_covariance_estimation}
 
-Because of the non-negative kernel property, for each $x \in \Omega$, the impulse response $\impulseresponse_x$ is a non-negative function $y \mapsto \impulseresponse_x(y)$.
+Because of the non-negative impulse response property, for each $x \in \Omega$, the impulse response $\impulseresponse_x$ is a non-negative function $y \mapsto \impulseresponse_x(y)$.
 Hence, $\impulseresponse_x$ is a scaled probability distribution.
@@
 
-\subsection{Convolution kernels}
+\subsection{Impulse response batches}
 \label{sec:get_impulse_response}
@@
 
-\subsection{Approximating kernel entries $\Phi(y,x)$}
+\subsection{Approximate integral kernel entries}
 
@@
 
-\subsubsection{Improvements when $\Aop$ is symmetric}
+\subsubsection{Improvements based on symmetry}
 If $\Aop$ is symmetric, impulse responses of $\Aop$ are also impulse responses of $\Aop^T$.

### 2022-03-02 paper/localpsf.tex
impulse response batches

@@
 
-We present a fast matrix free method for approximating locally translation-invariant operators, $\Aop$, that have locally supported non-negative integral kernels.
+We present a fast matrix free method for approximating operators, $\Aop$, that have locally supported non-negative impulse responses.
 Such operators arise throughout computational science.
@@
 
-We compute convolution kernels $\convkernel_{x_i}$ in batches by applying $\Aop$ to a small number Dirac combs.
-The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$.
-That is,
-\begin{equation*}
-	\diraccomb_k := \sum_{x_i \in \pointbatch_k} \delta_{x_i}.
-\end{equation*}
-For each batch $\pointbatch_k$, we compute the action of $\Aop$ on the associated Dirac comb:
+We compute impulse responses $\impulseresponse_{x_i}$ in batches by applying $\Aop$ to a small number Dirac combs.
+The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the following weighted sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$:
+\begin{equation*}
+	\diraccomb_k := \sum_{x_i \in \pointbatch_k} \delta_{x_i} / V(x_i).
+\end{equation*}
+The support of an impulse response is estimated to be contained within an ellipsoid, but this is just an approximation.
+A small portion of the impulse response will reside outside of the ellipsoid.
+If the impulse responses vary widely in scale, this neglected portion of an impulse response at a point $x_i$ with large $V(x_i)$ could overwhelm nearby impulse responses at points $x_j$ with small $V(x_j)$.
+Each point source is weighted by $1/V(x_i)$ so that the impulse responses within a batch are all comparable in scale.
+
+For each batch $\pointbatch_k$, we compute the \emph{impulse response batch}, $\eta_k$, by applying $\Aop$ to the associated Dirac comb:
 \begin{equation}
@@
 	\label{eq:phi_b}
-	\combresponse_k = \left(\Aop \left(\sum_{x_i \in \pointbatch_k} \delta_{x_i}\right)^*\right)^* = \sum_{x_i \in \pointbatch_k} \left(\Aop \delta_{x_i}^*\right)^* = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i}.
+	\combresponse_k = \left(\Aop \left(\sum_{x_i \in \pointbatch_k} \delta_{x_i}/V(x_i)\right)^*\right)^* = \sum_{x_i \in \pointbatch_k} \left(\Aop \delta_{x_i}^*\right)^* / V(x_i) = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
 \end{equation}
@@
 \label{eq:varphi_eval}
-	\convkernel_{x_i}(z) = \impulseresponse_i(z+x_i) = 
+	\impulseresponse_{x_i}(z) = 
 	\begin{cases}
-		\combresponse_k(z+x_i), & z+x_i \in E_{x_i} \\
+		V(x_i)\combresponse_k(z), & z \in E_{x_i} \\
 		0, & \text{otherwise}
@@
 for all $x_i \in \pointbatch_k$.
-By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the convolution kernels $\convkernel_{x_i}$ associated with every point $x_i \in \pointbatch_k$.
+By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the impulse responses $\impulseresponse_{x_i}$ associated with every point $x_i \in \pointbatch_k$.
 

### 2022-03-02 paper/localpsf.tex
impulse response batches
writing

@@
 	Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources. 
-We construct a-priori ellipsoid estimates of the supports of the impulse responses (i.e., support ellipsoids are estimated \emph{before} any impulse responses have been computed).
+%	We construct a-priori ellipsoid estimates of the supports of the impulse responses (i.e., support ellipsoids are estimated \emph{before} any impulse responses have been computed). 
 	By solving an ellipsoid packing problem, we choose as many points as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap. 
@@
 Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation.
-The details of the interpolation will be described in Section SECTION, but the key idea is something that we call \emph{local mean-displacement invariance}, which is the widely applicable assumption that
+The details of the interpolation will be described in Section SECTION, but the key principle behind the interpolation is an assumption that we call \emph{local mean-displacement invariance}.
+Specifically, we make the assumption that
 \begin{equation}
@@
 This is shown in Figure FIG, and discussed in greater detail in Section SEC.
-We approximate $\phi_x(y)$ by a weighted linear combination of the values $\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for the $k$-nearest points $x_i$ to $x$.
-Here $k$ is a user-defined parameter; we use $k=10$.
+We approximate $\phi_x(y)$ by a weighted linear combination of the values $\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for the a small number of sample points $x_i$ near $x$.
 The weights in the linear combination are determined via radial basis function interpolation.
@@
 
-\section{Constructing the product-convolution H-matrix approximation}
+\section{Method}
 \label{sec:method}
-In this section we describe how we choose the sample points, $x_i$ (Section \ref{sec:sample_point_selection}), how we form the convolution kernels, $\convkernel_i$ (Section \ref{sec:get_impulse_response}), and how we form the weighting functions, $w_i$ (Section \ref{sec:weighting_functions}), for the product-convolution approximation, $\AopPc \approx \Aop$, given in \eqref{eq:product_convolution}.
-We also describe how we convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), and (optionally) how we modify the resulting H-matrix, $\AmatPc$, to force it to be symmetric positive semi-definite (Section \ref{sec:make_spd}).
-The complete algorithm for constructing $\AmatPc$ is shown in Algorithm \ref{alg:construct_Atilde}.
+In this section we describe how we compute the impulse response moments $V,\mu,\Sigma$ and use these moments to estimate the supports of the impulse responses $\phi_x$, how we choose the sample points $x_i$ (Section \ref{sec:sample_point_selection}), how we compute the impulse response batches $\phi_{x_i}$ (Section \ref{sec:get_impulse_response}), how we interpolate impulse response to approximate integral kernel entries $\Phi(y,x)$, and how we construct the H-matrix approximation $\mathbf{A}_h$ of the operator $\mathcal{A}$ after the problem has been discretized with finite elements.
+We also describe how to (optionally) modify $\mathbf{A}_h$ to force it to be symmetric positive semi-definite.
+The complete algorithm is shown in Algorithm \ref{alg:construct_Atilde}.
 
@@
 
-\subsection{Sample point selection}
-\label{sec:sample_point_selection}
-
-We choose sample points, $x_i$, in batches.
-A greedy algorithm is used to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other.
-We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\impulseresponse_x$ (Section \ref{eq:mean_and_covariance_estimation}).
-The support of $\impulseresponse_x$ is estimated to be contained within an ellipsoid, so the process of choosing batches of points is an ellipsoid packing problem (Section \ref{sec:greedy_point_selection}).
-
-\subsubsection{Impulse response moments and support estimation}
+\subsection{Impulse response moments and support estimation}
 \label{eq:mean_and_covariance_estimation}
@@
 
-\subsubsection{Greedy ellipsoid packing}
-\label{sec:greedy_point_selection}
-
-We use a greedy ellipsoid packing algorithm to choose batches of sample points, $\pointbatch_k$, such that there is no overlap between the ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch.
-These sample point batches will be used in Section \ref{sec:get_impulse_response} to compute many impulse responses of $\Aop$ using only a small number of operator actions of $\Aop$.
+\subsection{Sample point selection via greedy ellipsoid packing}
+\label{sec:sample_point_selection}
+
+We choose sample points, $x_i$, in batches $\pointbatch_k$. 
+%A greedy algorithm is used to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other.
+We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\impulseresponse_x$ (Section \ref{eq:mean_and_covariance_estimation}).
+The support of $\impulseresponse_x$ is estimated to be contained within an ellipsoid, so the process of choosing batches of points is an ellipsoid packing problem (Section \ref{sec:greedy_point_selection}).
+We use a greedy ellipsoid packing algorithm to choose as many points as possible per batch, while ensuring that there is no overlap between the support ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch.
+%to choose batches of sample points, $\pointbatch_k$, such that there is no overlap between the ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch. 
+%These sample point batches will be used in Section \ref{sec:get_impulse_response} to compute many impulse responses of $\Aop$ using only a small number of operator actions of $\Aop$.
 The support of $\impulseresponse_x$ is (approximately) contained in the ellipsoid $E_x$, so by applying $\Aop$ to the Dirac comb associated with all sample points in a batch, we will recover the impulse responses for all sample points in that batch.
@@
 Candidate points for the first batch are chosen arbitrarily from $P$.
-Once $p$ is selected, we remove $p$ from $P$.
-If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_k$, then we add $p$ to $\pointbatch_k$.
+Once $p$ is selected, we remove $p$ from $P$. Then we perform the following two checks:
+\begin{enumerate}
+	\item We check whether $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_k$.
+	\item We make sure that $V(p)$ is not too small, in the sense $V(p) > \tau V_\text{max}$, where $V_\text{max}$ is the largest value of $V(q)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\tau=10^{-5}$).
+\end{enumerate}
+If $p$ passes both these checks (i.e., if $p$ is sufficiently far from the other points and $V(p)$ is not too small) then we add $p$ to $\pointbatch_k$.
 Otherwise we discard $p$.
@@
 This is detailed in Algorithm \ref{alg:point_choice}.
-We determine whether $E_p \cap E_q = \{\}$ using the fast ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.
+We determine whether $E_p \cap E_q = \{\}$ using the ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.
 
 We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots$, until the number of batches exceeds a desired threshold.
-In our implementation, for each batch the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, except those points in previously chosen batches.
+In our implementation, for each batch the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem.
+%, except those points in previously chosen batches.
 
@@
 
-We compute impulse responses $\impulseresponse_{x_i}$ in batches by applying $\Aop$ to a small number Dirac combs.
+We compute impulse responses $\impulseresponse_{x_i}$ in batches by applying $\Aop$ to a Dirac combs.
 The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the following weighted sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$:
@@
 \end{equation*}
+We compute the \emph{impulse response batch}, $\eta_k$, by applying $\Aop$ to the Dirac comb:
+\begin{equation}
+	\label{eq:dirac_comb_H_action}
+	\combresponse_k := \left(\Aop \diraccomb_k^*\right)^*.
+\end{equation}
+Each point source $\delta_{x_i}$ is scaled by $1/V(x_i)$ so that the resulting scaled impulse responses within a batch are all comparable in magnitude.
 The support of an impulse response is estimated to be contained within an ellipsoid, but this is just an approximation.
 A small portion of the impulse response will reside outside of the ellipsoid.
-If the impulse responses vary widely in scale, this neglected portion of an impulse response at a point $x_i$ with large $V(x_i)$ could overwhelm nearby impulse responses at points $x_j$ with small $V(x_j)$.
-Each point source is weighted by $1/V(x_i)$ so that the impulse responses within a batch are all comparable in scale.
-
-For each batch $\pointbatch_k$, we compute the \emph{impulse response batch}, $\eta_k$, by applying $\Aop$ to the associated Dirac comb:
-\begin{equation}
-	\label{eq:dirac_comb_H_action}
-	\combresponse_k := \left(\Aop \diraccomb_k^*\right)^*.
-\end{equation}
-By linearity, the resulting function $\combresponse_k$ may be written as follows:
+Without this scaling, the neglected portion of an impulse response corresponding to a point $x_i$ could overwhelm the impulse response corresponding to a nearby point $x_j$, when $V(x_i)$ is much larger than $V(x_j)$.
+We are not in danger of dividing by zero here, because our ellipsoid packing procedure from Section SEC excludes $x_i$ from consideration to be a sample point if $V(x_i)$ is smaller than a predetermined threshold.
+
+By linearity, the impulse response batch $\combresponse_k$ may be written as follows:
 \begin{equation}
@@
 
-Given a pair of points $(y,x) \subset \Omega \times \Omega$, we define the function
+Here we describe how we compute approximate kernel entries $\Phi(y,x)$ for arbitrary pairs of points $(y,x) \subset \Omega \times \Omega$. To that end, we define the function
 \begin{equation}

### 2022-03-02 paper/localpsf.tex
writing

@@
 This is the case, for example, for Hessians in PDE constrained inverse problems with highly informative data.
-As the data become more informative, typically the ``mass'' of $\Phi(y,x)$ concentrates nearer and nearer to the ``diagonal'' $y=x$, the Lipschitz constant increases, and the required rank in the approximation blows up.
+As the data become more informative, typically the ``mass'' of $\Phi(y,x)$ concentrates nearer and nearer to the ``diagonal'' $y=x$, the Lipschitz constant increases, and the required rank in the approximation increases.
 %Moreover, if $\Aop$ is numerically low rank, then one would typically use more accurate and robust matrix-free methods for forming low-rank approximations, such as randomized SVD or Lanczos.
@@
 
-We start with a finite set of candidate points $P$.
-To build $\pointbatch_k$, first we initialize $\pointbatch_k$ as an empty set.
-Then we select a candidate point $p \in P$ that is the farthest away from all points in previous sample point batches.
+We start with a finite set of candidate points $P$ and build $\pointbatch_k$ incrementally with points selected from $P$.
+For simplicity of explanation, here $B_k$ and $P$ are mutable sets that we add points to and remove points from.
+First we initialize $\pointbatch_k$ as an empty set.
+Then we select the candidate point $p \in P$ that is the farthest away from all points in previous sample point batches.
 That is, $p$ is a maximizer of the following optimization problem:
@@
 \end{equation*}
-Candidate points for the first batch are chosen arbitrarily from $P$.
+Candidate points for the first batch are chosen randomly from $P$.
 Once $p$ is selected, we remove $p$ from $P$. Then we perform the following two checks:
@@
 	\item We check whether $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_k$.
-	\item We make sure that $V(p)$ is not too small, in the sense $V(p) > \tau V_\text{max}$, where $V_\text{max}$ is the largest value of $V(q)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\tau=10^{-5}$).
+	\item We make sure that $V(p)$ is not too small, by checking whether $V(p) > \tau V_\text{max}$.
+Here $V_\text{max}$ is the largest value of $V(q)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\tau=10^{-5}$).
 \end{enumerate}
-If $p$ passes both these checks (i.e., if $p$ is sufficiently far from the other points and $V(p)$ is not too small) then we add $p$ to $\pointbatch_k$.
+If $p$ passes both these checks (i.e., if $p$ is sufficiently far from other points in the batch, and $V(p)$ is not too small) then we add $p$ to $\pointbatch_k$.
 Otherwise we discard $p$.
@@
 This is detailed in Algorithm \ref{alg:point_choice}.
-We determine whether $E_p \cap E_q = \{\}$ using the ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.
+We check whether $E_p \cap E_q = \{\}$ using the ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.
 
 We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots$, until the number of batches exceeds a desired threshold.
-In our implementation, for each batch the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem.
-%, except those points in previously chosen batches.
+In our implementation, for each batch the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, except for points in previously chosen batches.
 
@@
 \end{equation}
-Each point source $\delta_{x_i}$ is scaled by $1/V(x_i)$ so that the resulting scaled impulse responses within a batch are all comparable in magnitude.
-The support of an impulse response is estimated to be contained within an ellipsoid, but this is just an approximation.
-A small portion of the impulse response will reside outside of the ellipsoid.
-Without this scaling, the neglected portion of an impulse response corresponding to a point $x_i$ could overwhelm the impulse response corresponding to a nearby point $x_j$, when $V(x_i)$ is much larger than $V(x_j)$.
-We are not in danger of dividing by zero here, because our ellipsoid packing procedure from Section SEC excludes $x_i$ from consideration to be a sample point if $V(x_i)$ is smaller than a predetermined threshold.
-
 By linearity, the impulse response batch $\combresponse_k$ may be written as follows:
@@
 By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the impulse responses $\impulseresponse_{x_i}$ associated with every point $x_i \in \pointbatch_k$.
+
+Each point source $\delta_{x_i}$ in the Dirac comb is scaled by $1/V(x_i)$ so that the resulting scaled impulse responses within $\eta_k$ are comparable in magnitude.
+Note that we are not in danger of dividing by zero here, because our ellipsoid packing procedure from Section SEC excludes $x_i$ from consideration as a sample point if $V(x_i)$ is smaller than a predetermined threshold.
+Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, could overwhelm $\phi_{x_j}$ for a nearby point $x_j$ simply because $V(x_i)$ is much larger than $V(x_j)$.
+
+%Note that the support of $\impulseresponse_{x_i}$ is estimated to be contained within the ellipsoid $E_{x_i}$, but this is just an approximation. 
+%%The impulse response $\impulseresponse_{x_i}$ may be nonzero inside the ellipsoid $E_{x_j}$ associated with another point $x_j$ in the batch, and this will cause error in our recovery of $\phi_{x_j}$.
+%Without this scaling, the neglected portion of an impulse response corresponding to a point $x_i$ could overwhelm the impulse response corresponding to a nearby point $x_j$ if $V(x_i)$ is much larger than $V(x_j)$.
 

### 2022-03-02 paper/localpsf.tex
writing

@@
 Note that we are not in danger of dividing by zero here, because our ellipsoid packing procedure from Section SEC excludes $x_i$ from consideration as a sample point if $V(x_i)$ is smaller than a predetermined threshold.
-Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, could overwhelm $\phi_{x_j}$ for a nearby point $x_j$ simply because $V(x_i)$ is much larger than $V(x_j)$.
+Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, will overwhelm $\phi_{x_j}$ for a nearby point $x_j$ if $V(x_i)$ is much larger than $V(x_j)$.
 

### 2022-03-03 paper/localpsf.tex
add mean displacement invariance diagram

@@
 Approximation \eqref{eq:translate3}, which we use, is new.
-We call operators that satisfy \eqref{eq:translate3} ``locally mean-displacement invariant.'' These approximations are illustrated in Figure FIG.
+We call operators that satisfy \eqref{eq:translate3} ``locally mean-displacement invariant.'' We illustrate local mean-displacement invariance in Figure \ref{fig:mean_displacement_invariance}.
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale=1.0]{mean_displacement_invariance.pdf}
+	\end{center}
+	\caption{An operator is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \mu(x) + \mu(x')\right)$ when $x$ is close to $x'$.}
+	\label{fig:mean_displacement_invariance}
+\end{figure}
 
@@
 Furthermore, approximation \eqref{eq:translate3} can hold in situations where neither \eqref{eq:translate1} nor \eqref{eq:translate2} holds.
-For example, \eqref{eq:translate3} will hold when an operator is locally translation-invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case (see Figure FIG).
+For example, \eqref{eq:translate3} will hold when an operator is locally translation-invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case.
 Additionally, \eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions.
@@
 Each point source $\delta_{x_i}$ in the Dirac comb is scaled by $1/V(x_i)$ so that the resulting scaled impulse responses within $\eta_k$ are comparable in magnitude.
-Note that we are not in danger of dividing by zero here, because our ellipsoid packing procedure from Section SEC excludes $x_i$ from consideration as a sample point if $V(x_i)$ is smaller than a predetermined threshold.
-Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, will overwhelm $\phi_{x_j}$ for a nearby point $x_j$ if $V(x_i)$ is much larger than $V(x_j)$.
+Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section SEC excludes $x_i$ from consideration as a sample point if $V(x_i)$ is smaller than a predetermined threshold.
+Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, may overwhelm $\phi_{x_j}$ for a nearby point $x_j$ if $V(x_i)$ is much larger than $V(x_j)$.
 

### 2022-03-04 paper/localpsf.tex
editing writing

@@
 \end{align}
-Approximations \eqref{eq:translate1} and \eqref{eq:translate2} are standard.
-Approximation \eqref{eq:translate3}, which we use, is new.
-We call operators that satisfy \eqref{eq:translate3} ``locally mean-displacement invariant.'' We illustrate local mean-displacement invariance in Figure \ref{fig:mean_displacement_invariance}.
+We use approximation \eqref{eq:translate3}, which we call ``locally mean-displacement invariance.'' This is illustrated in Figure \ref{fig:mean_displacement_invariance}.
 
@@
 \label{sec:method}
-In this section we describe how we compute the impulse response moments $V,\mu,\Sigma$ and use these moments to estimate the supports of the impulse responses $\phi_x$, how we choose the sample points $x_i$ (Section \ref{sec:sample_point_selection}), how we compute the impulse response batches $\phi_{x_i}$ (Section \ref{sec:get_impulse_response}), how we interpolate impulse response to approximate integral kernel entries $\Phi(y,x)$, and how we construct the H-matrix approximation $\mathbf{A}_h$ of the operator $\mathcal{A}$ after the problem has been discretized with finite elements.
+In this section we describe how we compute the impulse response moments $V,\mu,\Sigma$ and use these moments to estimate the supports of the impulse responses $\phi_x$ (Section \ref{eq:mean_and_covariance_estimation}), the greedy ellipsoid packing procedure we use to choose batches of sample points $x_i$ (Section \ref{sec:sample_point_selection}), how we compute impulse responses, $\phi_{x_i}$, in batches (Section \ref{sec:get_impulse_response}), how we interpolate the $\phi_{x_i}$ to approximate integral kernel entries $\Phi(y,x)$ (Section \ref{sec:approximate_kernel_entries}), finite element discretization of $\Phi$ (Section \ref{sec:finite_element_kernel}), and how we construct the H-matrix approximation $\mathbf{A}_h$ of the operator $\mathcal{A}$.
 We also describe how to (optionally) modify $\mathbf{A}_h$ to force it to be symmetric positive semi-definite.
@@
 
-\subsection{Impulse response moments and support estimation}
+\subsection{Impulse response support ellipsoids}
 \label{eq:mean_and_covariance_estimation}
 
-Because of the non-negative impulse response property, for each $x \in \Omega$, the impulse response $\impulseresponse_x$ is a non-negative function $y \mapsto \impulseresponse_x(y)$.
+%Because of the non-negative impulse response property, for each $x \in \Omega$, the impulse response $\impulseresponse_x$ is a non-negative function $y \mapsto \impulseresponse_x(y)$.
 Hence, $\impulseresponse_x$ is a scaled probability distribution.
 Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor (the ``volume'' of $\impulseresponse_x$), let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
+
 We make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
@@
 where $\tau$ is a fixed constant.
-The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\impulseresponse}_x$ (see Figure FIG).
+The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\impulseresponse}_x$.
+
 The quantity $\tau$ is a parameter that must be chosen by the user of the algorithm.
 The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$.
-However, the cost of our algorithm will depend on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
+However, in Section \ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
 We find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results.
@@
 
-The ellipsoid $E_x$ is not immediately available.
-To determine $E_x$, we must compute the scalar $\spatialvol(x)$, the length-$\gdim$ vector $\spatialmean(x)$, and the $\gdim \times \gdim$ matrix $\spatialcov(x)$ (recall $\gdim \in \{1,2,3\}$ is the spatial dimension of $\Omega$).
+\section{Impulse response moments}
+
+%The ellipsoid $E_x$ is not immediately available. 
+To determine $E_x$, we must compute the scalar $\spatialvol(x)$, the length-$\gdim$ vector $\spatialmean(x)$, and the $\gdim \times \gdim$ matrix $\spatialcov(x)$.
+%(recall $\gdim$ is the spatial dimension of $\Omega$).
 The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}.
 Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
-But this direct approach is computationally infeasible because and it requires one operator action of $\Aop$ per point $x$ that we wish to evaluate $E_x$ at, and we will need to evaluate $E_x$ at a large number of points.
+But this direct approach is computationally infeasible because it requires one operator action of $\Aop$ per point $x$ that we wish to determine $E_x$ at, and we will need to determine $E_x$ at a large number of points.
 Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions.
-We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Theorem \ref{thm:vol_mean_cov} and summarize the method in Algorithm \ref{alg:varhpi_mean_cov}.
+We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Theorem \ref{thm:vol_mean_cov} and summarize the method for computing these moments in Algorithm \ref{alg:varhpi_mean_cov}.
 
@@
 	\label{thm:vol_mean_cov}
-	Let $\spatialvol:\Omega \rightarrow \mathbb{R}$, $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$, and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$, be the following functions:
-	\begin{equation*}
-	\spatialvol(x) := \int_{\Omega} \impulseresponse_x(y) dy, \qquad
-	\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
-	\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
-	\end{equation*}
-	where $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denotes the normalized version of $\impulseresponse_x$, $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$.
+%	Let $\spatialvol:\Omega \rightarrow \mathbb{R}$, $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$, and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$, be the following functions:
+%	\begin{equation*}
+%	\spatialvol(x) := \int_{\Omega} \impulseresponse_x(y) dy, \qquad
+%	\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
+%	\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
+%	\end{equation*}
+%	where $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denotes the normalized version of $\impulseresponse_x$, $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.
+	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$.
 Let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
@@
 The support of $\impulseresponse_x$ is estimated to be contained within an ellipsoid, so the process of choosing batches of points is an ellipsoid packing problem (Section \ref{sec:greedy_point_selection}).
-We use a greedy ellipsoid packing algorithm to choose as many points as possible per batch, while ensuring that there is no overlap between the support ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch.
+We use a greedy ellipsoid packing algorithm to choose as many points as possible per batch, while ensuring that there is no overlap between the support ellipsoids, $E_{x_i}$, associated with the sample points within a batch.
 %to choose batches of sample points, $\pointbatch_k$, such that there is no overlap between the ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch. 
@@
 We start with a finite set of candidate points $P$ and build $\pointbatch_k$ incrementally with points selected from $P$.
-For simplicity of explanation, here $B_k$ and $P$ are mutable sets that we add points to and remove points from.
+For simplicity of explanation, here $\pointbatch_k$ and $P$ are mutable sets that we add points to and remove points from.
 First we initialize $\pointbatch_k$ as an empty set.
-Then we select the candidate point $p \in P$ that is the farthest away from all points in previous sample point batches.
-That is, $p$ is a maximizer of the following optimization problem:
-\begin{equation*}
-\max_{p \in P} \min_{x \in \pointbatch_1 \cup \dots \cup \pointbatch_{k-1}} \|p - x\|.
-\end{equation*}
-Candidate points for the first batch are chosen randomly from $P$.
+Then we select the candidate point $p \in P$ that is the farthest away from all points in previous sample point batches $S_1 \cup S_2 \cup \dots \cup S_{k-1}$.
+%That is, $p$ is a maximizer of the following optimization problem:
+%\begin{equation*}
+%\max_{p \in P} \min_{x \in \pointbatch_1 \cup \dots \cup \pointbatch_{k-1}} \|p - x\|.
+%\end{equation*}
+Candidate points for the first batch $S_1$ are chosen randomly from $P$.
 Once $p$ is selected, we remove $p$ from $P$. Then we perform the following two checks:
@@
 
-We compute impulse responses $\impulseresponse_{x_i}$ in batches by applying $\Aop$ to a Dirac combs.
+We compute impulse responses $\impulseresponse_{x_i}$ in batches by applying $\Aop$ to Dirac combs.
 The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the following weighted sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$:
@@
 	\label{eq:dirac_comb_H_action}
-	\combresponse_k := \left(\Aop \diraccomb_k^*\right)^*.
-\end{equation}
-By linearity, the impulse response batch $\combresponse_k$ may be written as follows:
-\begin{equation}
-	\label{eq:phi_b}
-	\combresponse_k = \left(\Aop \left(\sum_{x_i \in \pointbatch_k} \delta_{x_i}/V(x_i)\right)^*\right)^* = \sum_{x_i \in \pointbatch_k} \left(\Aop \delta_{x_i}^*\right)^* / V(x_i) = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
-\end{equation}
-Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\impulseresponse_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\impulseresponse_j$ do not overlap when $i \neq j$, from \eqref{eq:convolution_kernel} and \eqref{eq:phi_b} we have (approximately)
+	\combresponse_k := \left(\Aop \diraccomb_k^*\right)^*
+	=\sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
+\end{equation}
+The last equality in \eqref{eq:dirac_comb_H_action} follows from linearity and the definition of $\impulseresponse_{x_i}$ in \eqref{eq:impulse_response_delta_action}.
+%\begin{equation}
+%\label{eq:phi_b}
+%\combresponse_k = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
+%\end{equation}
+%\begin{equation}
+%	\label{eq:phi_b}
+%	\combresponse_k = \left(\Aop \left(\sum_{x_i \in \pointbatch_k} \delta_{x_i}/V(x_i)\right)^*\right)^* = \sum_{x_i \in \pointbatch_k} \left(\Aop \delta_{x_i}^*\right)^* / V(x_i) = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
+%\end{equation}
+Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\impulseresponse_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\impulseresponse_j$ do not overlap when $i \neq j$, we have (approximately)
 \begin{equation}
 \label{eq:varphi_eval}
-	\impulseresponse_{x_i}(z) = 
+	\impulseresponse_{x_i}(z) / V(x_i) = 
 	\begin{cases}
-		V(x_i)\combresponse_k(z), & z \in E_{x_i} \\
+		\combresponse_k(z), & z \in E_{x_i} \\
 		0, & \text{otherwise}
@@
 for all $x_i \in \pointbatch_k$.
-By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the impulse responses $\impulseresponse_{x_i}$ associated with every point $x_i \in \pointbatch_k$.
+By performing one matrix-vector product, $\xi_k \mapsto \left(\Aop \diraccomb_k^*\right)^*$, we recover the impulse responses $\impulseresponse_{x_i}$ associated with every point $x_i \in \pointbatch_k$.
 
 Each point source $\delta_{x_i}$ in the Dirac comb is scaled by $1/V(x_i)$ so that the resulting scaled impulse responses within $\eta_k$ are comparable in magnitude.
-Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section SEC excludes $x_i$ from consideration as a sample point if $V(x_i)$ is smaller than a predetermined threshold.
+Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section \ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $V(x_i)$ is smaller than a predetermined threshold.
 Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, may overwhelm $\phi_{x_j}$ for a nearby point $x_j$ if $V(x_i)$ is much larger than $V(x_j)$.
@@
 \subsection{Approximate integral kernel entries}
-
-Here we describe how we compute approximate kernel entries $\Phi(y,x)$ for arbitrary pairs of points $(y,x) \subset \Omega \times \Omega$. To that end, we define the function
+\label{sec:approximate_kernel_entries}
+
+In Section \ref{sec:intro} we noted that the integral kernel, $\Aker$, is inaccessible. Here we describe how we compute approximate kernel entries,
+\begin{equation*}
+\widetilde{\Phi}(y,x) \approx \Phi(y,x),
+\end{equation*}
+for arbitrary pairs of points $(y,x) \subset \Omega \times \Omega$.
+% while Section \ref{sec:approximate_kernel_entries} we showed how we can cheaply evaluate approximate kernel entries at arbitrary pairs of points $(y,x) \in \Omega \times \Omega$.
+To that end, we define the function
 \begin{equation}
@@
 
-We approximate $\Phi(y,x)$ by performing radial basis function interpolation of the function $f_{y,x}(x')$ in the variable $x'$.
+We evaluate the approximate kernel entry $\widetilde{\Phi}(y,x)$ by performing radial basis function interpolation of the function $f_{y,x}(x')$ in the variable $x'$.
 The data used in the interpolation are the following (point,value) pairs
@@
 \end{equation}
-and $(x, f_{y,x}(x))$ is the (point,value) pair we approximate via interpolation.
+and $(x, f_{y,x}(x)) = (x,\Phi(y,x))$ is the (point,value) pair we approximate via interpolation.
 Here $\{x_i\}_{i=1}^k \subset \Omega$ are the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined.
@@
 %Interpolation using all sample points $x_i$ is too expensive, so we only use the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined. 
-In numerical experiments we use $k=10$. The $k$ nearest sample points are found by querying a precomputed KD-tree [CITE] of all sample points.
+In numerical experiments we use $k=10$. The $k$ nearest sample points are found by querying a precomputed KD-tree \cite{kdtree} of all sample points.
 
@@
 In this case, the point $x_i$ is considered invalid, and not used in the interpolation, even if it is close to $x$.
-We check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) [CITE] of the mesh cells used to discretize the problem.
-
-We approximate $\Phi(y,x)$ using radial basis function interpolation of the following form:
-\begin{equation}
-\Phi(y,x) = f_{y,x}(x) \approx \sum_{i=1}^k w_i~ b\left(\|x-x_i\|\right),
+We check whether $x_i$ is valid (i.e., whether $y - \mu(x) + \mu(x_i) \in \Omega$) by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
+
+Having found the $k$ nearest valid sample points $x_i$ to $x$, we approximate $\Phi(y,x)$ using radial basis function interpolation of the following form:
+\begin{equation}
+\widetilde{\Phi}(y,x) := \sum_{i=1}^k w_i~ b\left(\|x-x_i\|\right),
 \end{equation}
@@
 \begin{equation*}
-B_{ij} = b\left(\|x_i - x_j\|\right),
-\end{equation*}
-and the right hand side vector $f \in \mathbb{R}^k$ has entries $f_i = f_{y,x}(x_i)$, $i=1,\dots,k$. We use polyharmonic spline radial basis functions [CITE], which are defined as follows:
+B_{ij} := b\left(\|x_i - x_j\|\right), \qquad i,j=1,\dots,k,
+\end{equation*}
+and the right hand side vector $f \in \mathbb{R}^k$ has entries 
+\begin{equation*}
+f_i = f_{y,x}(x_i), \qquad i=1,\dots,k.
+\end{equation*}
+We use polyharmonic spline radial basis functions \cite{polyspline}, which are defined as follows:
 \begin{equation}
@@
 %Although it is typically recommended that one avoid evaluating finite element functions at arbitrary points [CITE], here we have no choice. 
-We evaluate $f_{y,x}(x_i)$ via the following process. First, we determine whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ by checking if
+To generate the right hand side vector $f$, we evaluate $f_{y,x}(x_i)$ at the points $x_i$ via the following process. First, we determine whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ by checking if
 \begin{equation}
@@
 \begin{equation}
-f_{y,x}(x_i) = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right) = V(x)\eta_b\left(y - \mu(x) + \mu(x_i)\right),
-\end{equation}
-where $\eta_b$ is the impulse response batch function that we have computed and stored.
+f_{y,x}(x_i) = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right) = V(x)\eta_b\left(y - \mu(x) + \mu(x_i)\right).
+\end{equation}
+Recall from Section \ref{sec:get_impulse_response} that $\eta_b$ is the impulse response batch that we have computed and stored.
 Evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$ requires some care, because the point $y - \mu(x) + \mu(x_i)$ is typically not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints.
@@
 \subsubsection{Improvements based on symmetry}
+\label{sec:symmetric_improvements}
 If $\Aop$ is symmetric, impulse responses of $\Aop$ are also impulse responses of $\Aop^T$.
@@
 in the same manner as before, but with $x$ and $y$ now taking opposite roles.
-We improve the interpolation accuracy at negligible additional cost by combining both of these approximations.
-Specifically, we combine both approximations by placing function values
-\begin{equation*}
-\left(\widetilde{f}_x\right)_i := \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right), \qquad i=1,\dots,k,
+We combine both of these interpolation schemes to improve the approximation accuracy at negligible additional.
+The two interpolation schemes are combined by placing function values
+\begin{equation*}
+\left(\widehat{f}_x\right)_i := \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right), \qquad i=1,\dots,k,
 \end{equation*}
@@
 \begin{equation*}
-\left(\widetilde{f}_y\right)_i := \frac{V(y)}{V(y_i)}\phi_{y_i}\left(x - \mu(y) + \mu(y_i)\right), \qquad i=1,\dots,k,
-\end{equation*}
-at the points $y_i-y$, and interpolating the combination of all $2k$ of these points and values at zero.
+\left(\widehat{f}_y\right)_i := \frac{V(y)}{V(y_i)}\phi_{y_i}\left(x - \mu(y) + \mu(y_i)\right), \qquad i=1,\dots,k,
+\end{equation*}
+at the points $y_i-y$, and interpolating the combination of these $2k$ points and values at zero.
 
@@
 \begin{equation*}
-\widetilde{x_i} := x_i - x \qquad \text{and} \qquad \widetilde{y_i} := y_i - y.
+\widehat{x_i} := x_i - x \qquad \text{and} \qquad \widehat{y_i} := y_i - y.
 \end{equation*}
@@
 \begin{equation*}
-\widetilde{B} = \begin{bmatrix}
-\widetilde{B}_{xx} & \widetilde{B}_{xy} \\
-\widetilde{B}_{yx} & \widetilde{B}_{xx}
-\end{bmatrix}, \qquad \widetilde{w} = \begin{bmatrix}\widetilde{w}_x \\ \widetilde{w}_y\end{bmatrix}, \qquad \widetilde{f} = \begin{bmatrix}
-\widetilde{f}_x \\ \widetilde{f}_y
+\widehat{B} = \begin{bmatrix}
+\widehat{B}_{xx} & \widehat{B}_{xy} \\
+\widehat{B}_{yx} & \widehat{B}_{xx}
+\end{bmatrix}, \qquad \widehat{w} = \begin{bmatrix}\widehat{w}_x \\ \widehat{w}_y\end{bmatrix}, \qquad \widehat{f} = \begin{bmatrix}
+\widehat{f}_x \\ \widehat{f}_y
 \end{bmatrix},
 \end{equation*}
-be a $2k \times 2k$ block matrix and $2k \times 1$ block vectors, respectively, with the block entries of $\widetilde{B}$ be given by
+be a $2k \times 2k$ block matrix and $2k \times 1$ block vectors, respectively, with the block entries of $\widehat{B}$ be given by
 \begin{align*}
-\left(\widetilde{B}_{xx}\right)_{ij} = b\left(\|\widetilde{x}_i - \widetilde{x}_j\|\right), &\qquad
-\left(\widetilde{B}_{xy}\right)_{ij} = b\left(\|\widetilde{x}_i - \widetilde{y}_j\|\right), \\
-\left(\widetilde{B}_{yx}\right)_{ij} = b\left(\|\widetilde{y}_i - \widetilde{x}_j\|\right), &\qquad \left(\widetilde{B}_{yy}\right)_{ij} = b\left(\|\widetilde{y}_i - \widetilde{y}_j\|\right),
+\left(\widehat{B}_{xx}\right)_{ij} = b\left(\|\widehat{x}_i - \widehat{x}_j\|\right), &\qquad
+\left(\widehat{B}_{xy}\right)_{ij} = b\left(\|\widehat{x}_i - \widehat{y}_j\|\right), \\
+\left(\widehat{B}_{yx}\right)_{ij} = b\left(\|\widehat{y}_i - \widehat{x}_j\|\right), &\qquad \left(\widehat{B}_{yy}\right)_{ij} = b\left(\|\widehat{y}_i - \widehat{y}_j\|\right),
 \end{align*}
@@
 \begin{equation}
-\Phi(y,x) \approx \sum_{i=1}^k \left(w_x\right)_i~ b\left(\|\widetilde{x}_i\|\right) + \sum_{i=1}^k \left(w_y\right)_i~ b\left(\|\widetilde{y}_i\|\right),
-\end{equation}
-where $\widetilde{w}$ is the solution of the following regularized least-squares problem,
-\begin{equation*}
-\min_{\widetilde{w}} \frac{1}{2}\|\widetilde{B} \widetilde{w} - \widetilde{f}\|^2 + \frac{\alpha}{2}\|\widetilde{w}\|^2,
+\widetilde{\Phi}(y,x) := \sum_{i=1}^k \left(w_x\right)_i~ b\left(\|\widehat{x}_i\|\right) + \sum_{i=1}^k \left(w_y\right)_i~ b\left(\|\widehat{y}_i\|\right),
+\end{equation}
+where $\widehat{w}$ is the solution of the following regularized least-squares problem,
+\begin{equation*}
+\min_{\widehat{w}} \frac{1}{2}\|\widehat{B} \widehat{w} - \widehat{f}\|^2 + \frac{\alpha}{2}\|\widehat{w}\|^2,
 \end{equation*}
 for a small regularization parameter $\alpha$. 
-We solve the regularized least squares problem by QR factorization of the block $4k \times 2k$ matrix $\begin{bmatrix}\widetilde{B} & I\end{bmatrix}^T$.
-We regularize the problem because of numerical issues that arise if there are points $\widetilde{x}_i$ and $\widetilde{y}_j$ in the interpolation that are very close to each other, a situation that arises often in practice.
-Indeed, situations where $\widetilde{x}_i = \widetilde{y}_j$ are inevitable when a regular grid is used to discretize the problem.
-In this case, $\widetilde{B}$ is (nearly) singular, so it is numerically ill-advised to solve the standard interpolation equation $\widetilde{B} \widetilde{w} = \widetilde{f}$.
+We solve the regularized least squares problem by QR factorization of the block $4k \times 2k$ matrix $\begin{bmatrix}\widehat{B} & I\end{bmatrix}^T$.
+We regularize the problem because of numerical issues that arise if there are points $\widehat{x}_i$ and $\widehat{y}_j$ in the interpolation that are very close to each other, a situation that arises often in practice.
+Indeed, situations where $\widehat{x}_i = \widehat{y}_j$ are inevitable when a regular grid is used to discretize the problem.
+In this case, $\widehat{B}$ is (nearly) singular, so it is numerically ill-advised to solve the standard interpolation equation $\widehat{B} \widehat{w} = \widehat{f}$.
 
@@
 
-\subsection{Discretization and hierarchical matrix construction}
-\label{sec:H_matrix_conversion}
-
-Here we discuss how we construct an H-matrix representation of our product-convolution approximation, after discretizing the problem with finite elements.
-
-\subsubsection{Finite element discretization of the integral kernel}
+%\subsection{Discretization and hierarchical matrix construction}
+%\label{sec:H_matrix_conversion}
+%
+%Here we discuss how we construct an H-matrix representation of our product-convolution approximation, after discretizing the problem with finite elements.
+
+\subsection{Finite element discretization of the integral kernel}
+\label{sec:finite_element_kernel}
 
@@
 
-In Section \ref{sec:intro} we noted that the original integral kernel, $\Aker$, is inaccessible, while in Equation \eqref{eq:kernel_entries}  we saw that the integral kernel for our product-convolution approximation, $\AkerPc \approx \Aker$, is accessible, and may be cheaply evaluated at arbitrary pairs of points $(y,x) \in \Omega \times \Omega$ via the following formula:
-\begin{equation}
-	\label{eq:eval_y_x}
-	\AkerPc(y,x) = \sum_{i=1}^\convrank w_i(x) \convkernel_{x_i}(y-x).
-\end{equation}
-We now define a further approximation, $\AkerPcMesh \approx \AkerPc$, as follows:
+We now define a further approximation, $\widetilde{\Phi}^h \approx \widetilde{\Phi}$, as follows:
 \begin{equation}
 	\label{eq:defn_of_Akerpcmesh}
-	\AkerPcMesh(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \AkerPc(p_i,p_j) \febasis_i(y) \febasis_j(x).
-\end{equation}
-The kernel $\AkerPcMesh$ is the interpolation of $\AkerPc$ onto $V_h \otimes V_h$.
-Error in the overall kernel approximation, $\AkerPcMesh \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$.
-
-Replacing $\Aker$ with $\AkerPcMesh$ within the definition of $\Aop$ in \eqref{eq:kernel_representation} yields the following operator approximation, $\AopPcMesh$:
-\begin{equation}
-	\label{eq:kernel_representation_pcmesh}
-	(\AopPcMesh u)(v) := \int_\Omega \int_\Omega v(y) \AkerPcMesh(y,x) u(x) dx dy.
-\end{equation}
-While $\AopPcMesh$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\AopPcMesh$ to functions in the finite element space.
-That is, $\AopPcMesh : V_h \rightarrow V_h'$.
+	\widetilde{\Phi}^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \widetilde{\Phi}(p_i,p_j) \febasis_i(y) \febasis_j(x).
+\end{equation}
+The kernel $\widetilde{\Phi}^h$ is the interpolation of $\widetilde{\Phi}$ onto $V_h \otimes V_h$.
+Error in the overall kernel approximation, $\widetilde{\Phi}^h \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$.
+
 Let $u_h \in V_h$ and $v_h \in V_h$, and let $\bm{u}$ and $\bm{v}$ be the coefficient vectors for $u_h$ and $v_h$, respectively, with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$.
@@
 and similar for $v_h$.
-Further, let $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ be the finite element mass matrix, and let $\AkerPcMat \in \mathbb{R}^{\fedim \times \fedim}$ be following matrix of approximate kernel entries:
+Replacing $\Aker$ with $\widetilde{\Phi}^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward but tedious algebraic manipulations, yields the following approximation:
+\begin{equation}
+	\label{eq:kernel_representation_pcmesh}
+	(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
+	= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u}.
+\end{equation}
+%While $\Aop$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\Aop$ to functions in the finite element space. 
+where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the following matrix of approximate kernel entries:
 \begin{equation}
 	\label{eq:Akerpcmat_entries}
-	\left(\AkerPcMat\right)_{ij} := \AkerPcMesh(p_i, p_j).
-\end{equation}
-Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, and \eqref{eq:Akerpcmat_entries}, the definition of the mass matrix, and performing algebraic manipulations, it is straightforward but tedious to show that
-\begin{equation*}
-	\left(\AopPcMesh u_h \right)(v_h) = \bm{v}^T \bm{M} \AkerPcMat \bm{M} \bm{u}.
-\end{equation*}
-The finite element matrix representation of $\AopPcMesh$, which we denote by $\AmatPc$, is therefore given by
+	\mathbf{\Phi}_{ij} := \widetilde{\Phi}(p_i, p_j),
+\end{equation}
+and $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ is the finite element mass matrix which has entries
+\begin{equation*}
+\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
+\end{equation*}
+%Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, and \eqref{eq:Akerpcmat_entries}, the definition of the mass matrix, and performing algebraic manipulations, it is straightforward but tedious to show that
+%\begin{equation*}
+%	\left(\AopPcMesh u_h \right)(v_h) = \bm{v}^T \bm{M} \AkerPcMat \bm{M} \bm{u}.
+%\end{equation*}
+The finite element matrix representation of of our approximation of $\Aop$ is therefore given by
 \begin{equation}
 	\label{eq:Amatpc_defn}
-	\AmatPc := \bm{M} \AkerPcMat \bm{M}.
-\end{equation}
-
-\subsubsection{Hierarchical matrix construction}
-
-We form an H-matrix representation of $\AmatPc$ by converting $\AkerPcMat$ and $\bm{M}$ to H-matrix format, then using fast H-matrix methods to multiply these matrices according to \eqref{eq:Amatpc_defn}.
-Once $\AmatPc$ is in H-matrix format, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods \cite{BORMHACKBUSCHBOOK}.
+	\mathbf{A} := \bm{M} \mathbf{\Phi} \bm{M}.
+\end{equation}
+
+\subsection{Hierarchical matrix construction}
+
+We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by converting $\mathbf{\Phi}$ and $\bm{M}$ to H-matrix format, then using fast H-matrix methods to multiply these matrices according to \eqref{eq:Amatpc_defn}.
+Once $\mathbf{A}_H$ is formed, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods \cite{BORMHACKBUSCHBOOK}.
 We use H1 matrices in our numerical results, but any of the other H-matrix formats (such as H2, HODLR, HSS, HBS, and others \cite{HMATRIX}) could be used instead.
@@
 
-We convert $\AkerPcMat$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{HLIBPRO}.
-Although $\AkerPcMat$ is a dense $\fedim \times \fedim$ matrix, constructing the H-matrix approximation of $\AkerPcMat$ only requires evaluation of $O(\hmatrixrank^2 \fedim \log \fedim)$ entries of $\AkerPc$, and these entries are computed via the formula in Equation \eqref{eq:eval_y_x}.
+We convert $\mathbf{\Phi}$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{HLIBPRO}.
+Although $\mathbf{\Phi}$ is a dense $\fedim \times \fedim$ matrix, constructing the H-matrix approximation only requires evaluation of $O(\hmatrixrank^2 \fedim \log \fedim)$ entries of $\mathbf{\Phi}$, and these entries are computed via interpolation procedure described in Section \ref{sec:approximate_kernel_entries}.
 Here $\hmatrixrank$ is the rank of the highest-rank block in the H-matrix.
-A dense representation of $\AkerPcMat$ is never formed.
+A dense representation of $\mathbf{\Phi}$ is never formed.
 We describe this H-matrix construction process in detail in Appendix \ref{app:h_matrix}.
-We convert $\bm{M}$ to H-matrix format using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same recursive block partitioning structure as was used for $\AkerPcMat$.
+We convert $\bm{M}$ to H-matrix format using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same recursive block partitioning structure as was used for $\mathbf{\Phi}$.
 
@@
 
-In many problems of practical interest (e.g., Hessian approximation) $\Amat$ is symmetric positive semi-definite.
-However, $\AmatPc$ is generally non-symmetric and indefinite because of errors in the product-convolution approximation.
+In many problems of practical interest (e.g., Hessian approximation) $\Aop$ is symmetric positive semi-definite.
+However, $\mathbf{A}_H$ is generally non-symmetric and indefinite because of approximation error.
 This is undesirable.
 Symmetry and positive semi-definiteness are important properties which should be preserved if possible.
-Also, lacking these properties may prevent one from using highly effective algorithms, such as the conjugate gradient method, to perform further useful operations involving $\AmatPc$.
-In this section we describe a method for modifying $\AmatPc$ to make it symmetric positive semi-definite.
-
-First, we use fast H-matrix addition to symmetrize $\AmatPc$:
-\begin{equation*}
-	\AmatPcSym := \frac{1}{2}\left(\AmatPc + \AmatPc^T\right).
-\end{equation*}
-Then we modify $\AmatPcSym$ to make it positive semi-definite by forming a rational matrix function of the following form: 
+Also, lacking these properties may prevent one from using highly effective algorithms, such as the conjugate gradient method, to perform further useful operations involving $\mathbf{A}_H$.
+In this section we describe a method for modifying $\mathbf{A}_H$ to make it symmetric positive semi-definite.
+
+Because of the modified interpolation for symmetric matrices described in Section \ref{sec:symmetric_improvements}, $\mathbf{A}_H$ is nearly (but not completely) symmetric.
+We further symmetrize $\mathbf{A}_H$ as follows:
+\begin{equation*}
+	\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
+\end{equation*}
+This step is fast because adding H-matrices is a comparatively inexpensive operation.
+
+We modify $\mathbf{A}_H^{\text{sym}}$ to make it positive semi-definite by forming a rational matrix function of the following form: 
 \begin{equation}
 	\label{eq:rational_matrix_function}
-	\AmatPcSymPlus := \ratcoeff_0 \bm{I} + \ratcoeff_1 \AmatPcSym + \ratcoeff_2 \left(\AmatPcSym + \ratpole \bm{I}\right)^{-1},
-\end{equation}
-where $\bm{I}$ is the identity matrix which has the same shape as $\AmatPcSym$.
-Approximation \eqref{eq:rational_matrix_function} may be written as $\AmatPcSymPlus = \ratfct(\AmatPcSym)$, where $\ratfct$ is the following rational function:
+	\mathbf{A}_H^{\text{sym}+} := \ratcoeff_0 \bm{I} + \ratcoeff_1 \mathbf{A}_H^{\text{sym}} + \ratcoeff_2 \left(\mathbf{A}_H^{\text{sym}} + \ratpole \bm{I}\right)^{-1},
+\end{equation}
+where $\bm{I}$ is the identity matrix which has the same shape as $\mathbf{A}_H^{\text{sym}}$.
+Approximation \eqref{eq:rational_matrix_function} may be written as $\mathbf{A}_H^{\text{sym}+} = \ratfct(\mathbf{A}_H^{\text{sym}})$, where $\ratfct$ is the following scalar rational function:
 \begin{equation}
 	\label{eq:rational_function_scalar}
-	\ratfct(\lambda) := \ratcoeff_0 + \ratcoeff_1 \lambda + \frac{\ratcoeff_2}{\lambda+\ratpole}.
-\end{equation}
-The scalars $\ratcoeff_0$, $\ratcoeff_1$, $\ratcoeff_2$, and $\ratpole$ are chosen so that the moderate and large positive eigenvalues of $\AmatPcSymPlus$ approximately equal the corresponding moderate and large positive eigenvalues of $\AmatPcSym$, while the negative and small positive eigenvalues of $\AmatPcSym$ are modified so that $\AmatPcSymPlus$ is positive semi-definite.
-Later in this section we will explain how $\ratcoeff_0$, $\ratcoeff_1$, $\ratcoeff_2$, and $\ratpole$ are chosen and how our choice of these scalars affects the spectral properties of $\AmatPcSymPlus$.
+	\ratfct(\lambda) := \ratcoeff_0 + \ratcoeff_1 \lambda + \frac{\ratcoeff_2}{\lambda+\ratpole},
+\end{equation}
+which is extended to matrices by functional calculus.
+The scalars $\ratcoeff_0$, $\ratcoeff_1$, $\ratcoeff_2$, and $\ratpole$ are chosen so that the moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym}+}$ approximately equal the corresponding moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$, while the negative and small positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$ are modified so that $\mathbf{A}_H^{\text{sym}+}$ is positive semi-definite.
+Later in this section we will explain how $\ratcoeff_0$, $\ratcoeff_1$, $\ratcoeff_2$, and $\ratpole$ are chosen and how our choice of these scalars affects the spectral properties of $\mathbf{A}_H^{\text{sym}+}$.
 In Figure FIG we illustrate $\ratfct(\lambda)$.
-Forming $\AmatPcSymPlus$ requires computing an H-matrix inverse, $\left(\AmatPcSym + \ratpole \bm{I}\right)^{-1}$.
-While explicit computation of matrix inverses should usually be avoided, here it is computationally acceptable because $\AmatPcSym + \ratpole \bm{I}$ is an H-matrix.
+Forming $\mathbf{A}_H^{\text{sym}+}$ requires computing an H-matrix inverse, $\left(\mathbf{A}_H^{\text{sym}} + \ratpole \bm{I}\right)^{-1}$.
+While explicit computation of matrix inverses should usually be avoided, here it is computationally acceptable because $\mathbf{A}_H^{\text{sym}} + \ratpole \bm{I}$ is an H-matrix.
 Inversion of H-matrices is a routine procedure which only requires $O(\fedim \log(\fedim)^2)$ work \cite{Hmatrixinverse}.
@@
 \end{subequations}
-where $\lambda_\text{min}$ and $\lambda_\text{max}$ are the smallest and largest eigenvalues of $\AmatPcSym$, respectively.
+where $\lambda_\text{min}$ and $\lambda_\text{max}$ are the smallest and largest eigenvalues of $\mathbf{A}_H^{\text{sym}}$, respectively.
 We estimate $\lambda_\text{min}$ and $\lambda_\text{max}$ using the implicitly restarted Lanczos method \cite{lanczos,scipy}.
-We assume that $\lambda_\text{min} < 0$; if $\lambda_\text{min} \ge 0$, then $\AmatPcSym$ is already positive semi-definite, so we do not need to modify it with a rational function.
+We assume that $\lambda_\text{min} < 0$; if $\lambda_\text{min} \ge 0$, then $\mathbf{A}_H^{\text{sym}}$ is already positive semi-definite, so we do not need to modify it with a rational function.
 We also require $\ratpole > |\lambda_\text{min}|$, so that the pole of $\ratfct$ is not in $[\lambda_\text{min},\lambda_\text{max}]$.
@@
 Together, equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, \eqref{eq:desired_property_3},  and \eqref{eq:desired_property_4}, and the condition $\ratpole > |\lambda_\text{min}|$, imply that $\ratfct(0)=0$ is the unique minimum of $\ratfct$ on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that $\ratfct$ is non-negative on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that  $\AmatPcSymPlus$ is positive semi-definite.
-Furthermore, \eqref{eq:desired_property_1} implies that the null space of $\AmatPcSymPlus$ contains the null-space of $\AmatPcSym$.
+Furthermore, \eqref{eq:desired_property_1} implies that the null space of $\mathbf{A}_H^{\text{sym+}}$ contains the null-space of $\mathbf{A}_H^{\text{sym}}$.
 This is important when the operator $\Aop$ has a large (or infinite) dimensional null space, or a spectrum that clusters at zero.
 Hessians in ill-posed distributed parameter inverse problems typically have both of these properties.
-Equation \eqref{eq:desired_property_4} forces the moderate and large positive eigenvalues of $\AmatPcSymPlus$ to be close to the corresponding moderate and large positive eigenvalues of $\AmatPcSym$.
-I.e., $\ratfct$ modifies the ``important'' part of the spectrum of $\AmatPcSym$ as little as possible.
-Equation \eqref{eq:desired_property_3} ensures that the negative eigenvalues of $\AmatPcSym$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $\ratfct$.
-There is a tradeoff: if we choose a larger value for $\ratfct(\lambda_\text{min})$, then the positive eigenvalues of $\AmatPcSymPlus$ will be closer to the corresponding positive eigenvalues of $\AmatPcSym$ (good), but the erroneous negative eigenvalues of $\AmatPcSym$ will be amplified in magnitude (bad).
+Equation \eqref{eq:desired_property_4} forces the moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym+}}$ to be close to the corresponding moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
+I.e., $\ratfct$ modifies the ``important'' part of the spectrum of $\mathbf{A}_H^{\text{sym}}$ as little as possible.
+Equation \eqref{eq:desired_property_3} ensures that the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $\ratfct$.
+There is a tradeoff: if we choose a larger value for $\ratfct(\lambda_\text{min})$, then the positive eigenvalues of $\mathbf{A}_H^{\text{sym+}}$ will be closer to the corresponding positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$ (good), but the erroneous negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ will be amplified in magnitude (bad).
 This is shown in Figure FIG.

### 2022-03-04 paper/localpsf.tex
editing writing

@@
 denote the spatially varying scaling factor (the ``volume'' of $\impulseresponse_x$), let 
-\begin{equation*}
+\begin{equation}
+\label{eq:scaled_impulse_response}
 \widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)
-\end{equation*}
+\end{equation}
 denote the normalized version of $\impulseresponse_x$, and let $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d},$
@@
 if $x_i$ is close to $x$.
-This is shown in Figure FIG, and discussed in greater detail in Section SEC.
+This is shown in Figure FIG.
 We approximate $\phi_x(y)$ by a weighted linear combination of the values $\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for the a small number of sample points $x_i$ near $x$.
@@
 		\Input{Linear operator $\Aop$, parameter $\text{max\_batches}$}
-		\Output{H-matrix $\AmatPc$ (optionally, $\AmatPcSymPlus$)}
+		\Output{H-matrix $\mathbf{A}_H$ (optionally, $\mathbf{A}_H^{\text{sym}+}$)}
 
@@
 
-		Form convolution kernels, $\convkernel_{x_i}$, on regular grids (Section \ref{sec:convolution_kernel_regular_grid})
-
-		Modify convolution kernels to address boundary artifacts (Section \ref{sec:boundary_extension})
-
-		Compute all weighting functions $w_i$ using Algorithm
-
-		Form H-matrix $\AmatPc$ (Section \ref{sec:H_matrix_conversion})
-
-		(optional) Modify $\AmatPc$ to make $\AmatPcSymPlus$ (Section \ref{sec:make_spd})
+		Form H-matrix approximation $\mathbf{\Phi}_H$ of integral kernel (Section \ref{sec:h_matrix_construction_short})
+
+		Form H-matrix approximation $\mathbf{A}_H$ of $\Aop$ (Section \ref{sec:h_matrix_construction_short})
+
+		(optional) Modify $\mathbf{A}_H$ to make $\mathbf{A}_H^{\text{sym}+}$ (Section \ref{sec:make_spd})
 
@@
 	\combresponse_k := \left(\Aop \diraccomb_k^*\right)^*
-	=\sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
-\end{equation}
-The last equality in \eqref{eq:dirac_comb_H_action} follows from linearity and the definition of $\impulseresponse_{x_i}$ in \eqref{eq:impulse_response_delta_action}.
+	=\sum_{x_i \in \pointbatch_k} \widehat{\impulseresponse}_{x_i}.
+\end{equation}
+The last equality in \eqref{eq:dirac_comb_H_action} follows from linearity and the definition of $\widehat{\impulseresponse}_{x_i}$ in \eqref{eq:scaled_impulse_response} and \eqref{eq:impulse_response_delta_action}.
 %\begin{equation}
@@
 \label{eq:varphi_eval}
-	\impulseresponse_{x_i}(z) / V(x_i) = 
+	\impulseresponse_{x_i}(z) / V(x_i) = \widehat{\impulseresponse}_{x_i}(z) =
 	\begin{cases}
@@
 
-We now define a further approximation, $\widetilde{\Phi}^h \approx \widetilde{\Phi}$, as follows:
+We define a further approximation, $\widetilde{\Phi}^h \approx \widetilde{\Phi}$, as follows:
 \begin{equation}
@@
 \subsection{Hierarchical matrix construction}
+\label{sec:h_matrix_construction_short}
 

### 2022-03-04 paper/localpsf.tex
editing writing

@@
 \end{equation*}
-for arbitrary pairs of points $(y,x) \subset \Omega \times \Omega$.
+for arbitrary pairs of points $(y,x) \subset \Omega \times \Omega$ via interpolation.
 % while Section \ref{sec:approximate_kernel_entries} we showed how we can cheaply evaluate approximate kernel entries at arbitrary pairs of points $(y,x) \in \Omega \times \Omega$.
@@
 
-We evaluate the approximate kernel entry $\widetilde{\Phi}(y,x)$ by performing radial basis function interpolation of the function $f_{y,x}(x')$ in the variable $x'$.
+We evaluate $\widetilde{\Phi}(y,x)$ by performing radial basis function interpolation of the function $f_{y,x}(x')$ in the variable $x'$.
 The data used in the interpolation are the following (point,value) pairs
@@
 
+\subsubsection{Radial basis function interpolation}
+
 Having found the $k$ nearest valid sample points $x_i$ to $x$, we approximate $\Phi(y,x)$ using radial basis function interpolation of the following form:
@@
 %Although it is typically recommended that one avoid evaluating finite element functions at arbitrary points [CITE], here we have no choice.
+
+%\subsubsection{Evaluating $f_{y,x}(x_i)$}
+
 To generate the right hand side vector $f$, we evaluate $f_{y,x}(x_i)$ at the points $x_i$ via the following process. First, we determine whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ by checking if
@@
 \end{equation}
-Recall from Section \ref{sec:get_impulse_response} that $\eta_b$ is the impulse response batch that we have computed and stored.
-Evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$ requires some care, because the point $y - \mu(x) + \mu(x_i)$ is typically not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints.
+%Recall from Section \ref{sec:get_impulse_response} that $\eta_b$ is the impulse response batch that we have computed and stored.
+Evaluating the impulse response batch $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$ requires some care, because the point $y - \mu(x) + \mu(x_i)$ is generally not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints.
 To evaluate $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$, we determine which mesh cell contains $y - \mu(x) + \mu(x_i)$, then we evaluate the relevant finite element basis functions on that mesh cell.
@@
 
-\subsubsection{Improvements based on symmetry}
+\subsubsection{Improved interpolation for symmetric operators}
 \label{sec:symmetric_improvements}
@@
 Indeed, situations where $\widehat{x}_i = \widehat{y}_j$ are inevitable when a regular grid is used to discretize the problem.
-In this case, $\widehat{B}$ is (nearly) singular, so it is numerically ill-advised to solve the standard interpolation equation $\widehat{B} \widehat{w} = \widehat{f}$.
+In this case, $\widehat{B}$ is (nearly) singular, so it is numerically ill-advised to solve the standard (un-regularized) interpolation equation $\widehat{B} \widehat{w} = \widehat{f}$.
 

### 2022-03-04 paper/localpsf.tex
shortening and simplifying radial basis function writing

@@
 \end{equation}
-where the coefficient matrix $B \in \mathbb{R}^{k \times k}$ has entries 
-\begin{equation*}
-B_{ij} := b\left(\|x_i - x_j\|\right), \qquad i,j=1,\dots,k,
-\end{equation*}
-and the right hand side vector $f \in \mathbb{R}^k$ has entries 
-\begin{equation*}
-f_i = f_{y,x}(x_i), \qquad i=1,\dots,k.
-\end{equation*}
+where the coefficient matrix $B \in \mathbb{R}^{k \times k}$ and right hand side vector $f \in \mathbb{R}^k$ have entries 
+\begin{align*}
+B_{ij} :=& b\left(\|x_i - x_j\|\right),\\
+f_i :=& f_{y,x}(x_i),
+\end{align*}
+for $i,j=1,\dots,k$.
+% has entries 
+%\begin{equation*}
+%f_i = f_{y,x}(x_i), \qquad i=1,\dots,k.
+%\end{equation*}
 We use polyharmonic spline radial basis functions \cite{polyspline}, which are defined as follows:
@@
 
-To generate the right hand side vector $f$, we evaluate $f_{y,x}(x_i)$ at the points $x_i$ via the following process. First, we determine whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ by checking if
+We evaluate $f_{y,x}(x_i)$ via the following process. First, we determine whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ by checking if
 \begin{equation}
@@
 \end{equation}
-If \eqref{eq:ellipsoid_check93} does not hold, then $y - \mu(x) + \mu(x_i) \notin E_{x_i}$ (i.e., $y - \mu(x) + \mu(x_i)$ is outside the estimated support of $\impulseresponse_{x_i}$), so we set $f_{y,x}(x_i)=0$ for the purpose of interpolation.
-If \eqref{eq:ellipsoid_check93} does hold, then $y - \mu(x) + \mu(x_i) \in E_{x_i}$.
-In this case, we look up the batch index $b$ for the batch of sample points that contains $x_i$, and evaluate
+If \eqref{eq:ellipsoid_check93} does not hold, then $y - \mu(x) + \mu(x_i)$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_{y,x}(x_i)=0$ for the purpose of interpolation.
+If \eqref{eq:ellipsoid_check93} does hold, then we look up the batch index $b$ for such that $x_i \in S_b$, and evaluate
 \begin{equation}
@@
 %Recall from Section \ref{sec:get_impulse_response} that $\eta_b$ is the impulse response batch that we have computed and stored.
-Evaluating the impulse response batch $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$ requires some care, because the point $y - \mu(x) + \mu(x_i)$ is generally not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints.
-To evaluate $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$, we determine which mesh cell contains $y - \mu(x) + \mu(x_i)$, then we evaluate the relevant finite element basis functions on that mesh cell.
+%Evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$ requires care, because 
+When the problem is discretized with finite elements, $y - \mu(x) + \mu(x_i)$ is generally not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints.
+To evaluate $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$, we therefore determine which mesh cell contains $y - \mu(x) + \mu(x_i)$, then we evaluate the relevant finite element basis functions on that mesh cell.
 Fortunately, the mesh cell containing $y - \mu(x) + \mu(x_i)$ was already determined as a side effect of querying the AABB tree of mesh cells when we checked whether $y - \mu(x) + \mu(x_i) \in \Omega$.

### 2022-03-04 paper/localpsf.tex
about to try eliminating f_xy(xi) notation

@@
 \subsection{Local mean-displacement invariance}
+\label{eq:local_mean_displacement_invariance}
 
@@
 
-In Section \ref{sec:intro} we noted that the integral kernel, $\Aker$, is inaccessible. Here we describe how we compute approximate kernel entries,
-\begin{equation*}
-\widetilde{\Phi}(y,x) \approx \Phi(y,x),
-\end{equation*}
-for arbitrary pairs of points $(y,x) \subset \Omega \times \Omega$ via interpolation.
+%In Section \ref{sec:intro} we noted that computing $\Aker\left(y,x\right)$ is computationally costly. 
+This section describes how we compute approximate kernel entries
+\begin{equation*}
+\widetilde{\Phi}(y,x) \approx \Phi(y,x).
+\end{equation*}
+%for arbitrary pairs of points $(y,x) \subset \Omega \times \Omega$.
 % while Section \ref{sec:approximate_kernel_entries} we showed how we can cheaply evaluate approximate kernel entries at arbitrary pairs of points $(y,x) \in \Omega \times \Omega$.
-To that end, we define the function
+Let
 \begin{equation}
@@
 \end{equation}
-It follows from inspection that 
-\begin{equation*}
-\Phi(y,x) = f_{y,x}(x),
-\end{equation*}
-and from the discussion in Section SEC, we expect that $\Phi(y,x) \approx f_{y,x}(x')$ when $x'$ is close to $x$.
-The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section SEC, is a form of preconditioning that improves the quality of this approximation in situations where $V(x)$ varies rapidly.
+It follows from inspection that $\Phi(y,x) = f_{y,x}(x)$,
+and from the discussion in Section \ref{eq:local_mean_displacement_invariance}, we expect 
+$
+\Phi(y,x) \approx f_{y,x}(x')
+$
+when $x'$ is close to $x$.
+The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section \ref{eq:local_mean_displacement_invariance}, is a form of preconditioning that improves the quality of this approximation in situations where $V(x)$ varies rapidly.
 %, as is commonly the case for many operators of practical interest. In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.

### 2022-03-04 paper/localpsf.tex
editing kernel entries section

@@
 
-\subsection{Approximate integral kernel entries}
+\subsection{Approximate integral kernel entries $\widetilde{\Phi}(y,x)$}
 \label{sec:approximate_kernel_entries}
@@
 %In Section \ref{sec:intro} we noted that computing $\Aker\left(y,x\right)$ is computationally costly. 
-This section describes how we compute approximate kernel entries
-\begin{equation*}
-\widetilde{\Phi}(y,x) \approx \Phi(y,x).
-\end{equation*}
+%This section describes how we compute approximate kernel entries
+%\begin{equation*}
+%\widetilde{\Phi}(y,x) \approx \Phi(y,x)
+%\end{equation*}
+%for arbitrary pairs of points $(y,x) \in \Omega \times \Omega$. 
+Given $(y,x)\in \Omega \times \Omega$, let
+\begin{equation}
+\label{eq:fxyxp}
+f_i := \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right), \qquad i=1,\dots,k,
+\end{equation}
 %for arbitrary pairs of points $(y,x) \subset \Omega \times \Omega$.
 % while Section \ref{sec:approximate_kernel_entries} we showed how we can cheaply evaluate approximate kernel entries at arbitrary pairs of points $(y,x) \in \Omega \times \Omega$.
-Let
-\begin{equation}
-\label{eq:fxyxp}
-f_{y,x}(x') := \frac{V(x)}{V(x')}\phi_{x_i}\left(y - \mu(x) + \mu(x')\right).
-\end{equation}
-It follows from inspection that $\Phi(y,x) = f_{y,x}(x)$,
-and from the discussion in Section \ref{eq:local_mean_displacement_invariance}, we expect 
-$
-\Phi(y,x) \approx f_{y,x}(x')
-$
-when $x'$ is close to $x$.
+where $\{x_i\}_{i=1}^k$ are the $k$ nearest sample points to $x$, excluding sample points for which $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ is undefined.
+That is, sample points such that $y-\mu(x)+\mu(x_i) \notin \Omega$.
+In numerical experiments we use $k=10$.
+We efficiently find the $k$ nearest sample points by querying a precomputed KD-tree \cite{kdtree} of all sample points, and check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
+%Per the discussion in Section \ref{eq:local_mean_displacement_invariance}, $\Phi(y,x) \approx f_i$ when $x_i$ is close to $x$.
+The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section \ref{eq:local_mean_displacement_invariance}, is a form of preconditioning that improves the quality of this approximation in situations where $V(x)$ varies widely.
+
+We approximate $\Phi(y,x)$
+%compute the approximate kernel entry,
+%\begin{equation*}
+%\widetilde{\Phi}(y,x) \approx \Phi(y,x)
+%\end{equation*}
+by interpolating the (point,value) pairs
+\begin{equation}
+\left(x_i, f_i\right), \qquad i=1,\dots,k
+\end{equation}
+at the point $x$ using radial basis function interpolation.
+%The approximation $\widetilde{\Phi}(y,x)$ is the value of this interpolation at $x$.
+%In numerical experiments we use $k=10$. The $k$ nearest sample points are found by querying a precomputed KD-tree \cite{kdtree} of all sample points. 
+%
+%Let
+%It follows from inspection that $\Phi(y,x) = f_{y,x}(x)$,
+%and from the discussion in Section \ref{eq:local_mean_displacement_invariance}, we expect 
+%$
+%\Phi(y,x) \approx f_{y,x}(x')
+%$
+%when $x'$ is close to $x$.
 The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section \ref{eq:local_mean_displacement_invariance}, is a form of preconditioning that improves the quality of this approximation in situations where $V(x)$ varies rapidly.
-%, as is commonly the case for many operators of practical interest. In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.
-%As discussed in Section SECTION, we make the approximation
-%\begin{equation*}
-%	\Phi(y,x) = \phi_x(y) \approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right)
-%\end{equation*}
-%if $x'$ is close to $x$. Hence, w
-
-We evaluate $\widetilde{\Phi}(y,x)$ by performing radial basis function interpolation of the function $f_{y,x}(x')$ in the variable $x'$.
+%%, as is commonly the case for many operators of practical interest. In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.
+%%As discussed in Section SECTION, we make the approximation
+%%\begin{equation*}
+%%	\Phi(y,x) = \phi_x(y) \approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right)
+%%\end{equation*}
+%%if $x'$ is close to $x$. Hence, w
+%
+%We evaluate $\widetilde{\Phi}(y,x)$ by performing radial basis function interpolation of the function $f_{y,x}(x')$ in the variable $x'$.
 The data used in the interpolation are the following (point,value) pairs
-\begin{equation}
-\left(x_i, f_{y,x}(x_i)\right), \qquad i=1,\dots,k,
-\end{equation}
-and $(x, f_{y,x}(x)) = (x,\Phi(y,x))$ is the (point,value) pair we approximate via interpolation.
-Here $\{x_i\}_{i=1}^k \subset \Omega$ are the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined.
-%a collection of sample points at which the impulse response $\phi_{x_i}$ is known.
-%. Specifically, we interpolate $(\text{point}, \text{value})$ pairs $(x_i, f_i)$ at the point $x$, where $x_i$ are sample points, and $f_i$ are given as follows:
 %\begin{equation}
-%\label{eq:phi_mu_shift32}
-%	f_i := \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
+%\left(x_i, f_{y,x}(x_i)\right), \qquad i=1,\dots,k,
 %\end{equation}
-%We must perform a new interpolation every time we evaluate $\Phi(y,x)$ at a new pair of points $(y,x)$, and constructing the H-matrix approximation of $\Phi$ requires evaluating $\Phi(y,x)$ at a large number of pairs of points $(y,x)$.
-%Interpolation using all sample points $x_i$ is too expensive, so we only use the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined. 
-In numerical experiments we use $k=10$. The $k$ nearest sample points are found by querying a precomputed KD-tree \cite{kdtree} of all sample points.
-
-From the definition of $f_{y,x}$ in \eqref{eq:fxyxp}, we see that $f_{y,x}(x_i)$ is only defined if $y - \mu(x) + \mu(x_i) \in \Omega$, because $\phi_x$ is only defined on $\Omega$.
+%and $(x, f_{y,x}(x)) = (x,\Phi(y,x))$ is the (point,value) pair we approximate via interpolation.
+%Here $\{x_i\}_{i=1}^k \subset \Omega$ are the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined.
+%%a collection of sample points at which the impulse response $\phi_{x_i}$ is known.
+%%. Specifically, we interpolate $(\text{point}, \text{value})$ pairs $(x_i, f_i)$ at the point $x$, where $x_i$ are sample points, and $f_i$ are given as follows:
+%%\begin{equation}
+%%\label{eq:phi_mu_shift32}
+%%	f_i := \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
+%%\end{equation}
+%%We must perform a new interpolation every time we evaluate $\Phi(y,x)$ at a new pair of points $(y,x)$, and constructing the H-matrix approximation of $\Phi$ requires evaluating $\Phi(y,x)$ at a large number of pairs of points $(y,x)$.
+%%Interpolation using all sample points $x_i$ is too expensive, so we only use the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined. 
+%In numerical experiments we use $k=10$. The $k$ nearest sample points are found by querying a precomputed KD-tree \cite{kdtree} of all sample points. 
+%
+%From the definition of $f_{y,x}$ in \eqref{eq:fxyxp}, we see that $f_{y,x}(x_i)$ is only defined if $y - \mu(x) + \mu(x_i) \in \Omega$, because $\phi_x$ is only defined on $\Omega$.
 But it may happen that $y - \mu(x) + \mu(x_i) \notin \Omega$, even if $y,x,x_i \in \Omega$.
@@
 We check whether $x_i$ is valid (i.e., whether $y - \mu(x) + \mu(x_i) \in \Omega$) by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
-
-\subsubsection{Radial basis function interpolation}
-
-Having found the $k$ nearest valid sample points $x_i$ to $x$, we approximate $\Phi(y,x)$ using radial basis function interpolation of the following form:
-\begin{equation}
-\widetilde{\Phi}(y,x) := \sum_{i=1}^k w_i~ b\left(\|x-x_i\|\right),
-\end{equation}
-where $b_i$ are radial basis functions, and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
+%\subsubsection{Radial basis function interpolation details}
+%Having found the $k$ nearest valid sample points $x_i$ to $x$, we approximate $\Phi(y,x)$ using radial basis function interpolation of the following form:
+In detail, we form the approximation $\widetilde{\Phi}(y,x) \approx \Phi(y,x)$ as follows:
+\begin{equation}
+\widetilde{\Phi}(y,x) := \sum_{i=1}^k w_i~ b\left(\|x-x_i\|\right).
+\end{equation}
+Here $b_i$ are radial basis functions and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
 \begin{equation}
@@
 \end{equation}
-where the coefficient matrix $B \in \mathbb{R}^{k \times k}$ and right hand side vector $f \in \mathbb{R}^k$ have entries 
-\begin{align*}
-B_{ij} :=& b\left(\|x_i - x_j\|\right),\\
-f_i :=& f_{y,x}(x_i),
-\end{align*}
-for $i,j=1,\dots,k$.
+where $B \in \mathbb{R}^{k \times k}$ has entries
+$
+B_{ij} := b\left(\|x_i - x_j\|\right),
+$
+and $f \in \mathbb{R}^k$ has entries $f_i$ from \eqref{eq:fxyxp}.
 % has entries 
@@
 \end{equation}
-for $r>0$, and $b_i(0):=0$.
+for $r>0$, and $b_i(0):=0$. For more details on radial basis function interpolation and polyharmonic splines, see \cite{rbf}.
 %\begin{equation}
@@
 
-We evaluate $f_{y,x}(x_i)$ via the following process. First, we determine whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ by checking if
+We evaluate $f_i$ via the following process. First, we determine whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ by checking if
 \begin{equation}
 \label{eq:ellipsoid_check93}
-(y - \mu(x))^T \Sigma^{-1}(x) (y - \mu(x))  \le \tau^2.
-\end{equation}
-If \eqref{eq:ellipsoid_check93} does not hold, then $y - \mu(x) + \mu(x_i)$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_{y,x}(x_i)=0$ for the purpose of interpolation.
-If \eqref{eq:ellipsoid_check93} does hold, then we look up the batch index $b$ for such that $x_i \in S_b$, and evaluate
-\begin{equation}
-f_{y,x}(x_i) = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right) = V(x)\eta_b\left(y - \mu(x) + \mu(x_i)\right).
-\end{equation}
+(y - \mu(x_i))^T \Sigma^{-1}(x_i) (y - \mu(x_i))  \le \tau^2.
+\end{equation}
+If \eqref{eq:ellipsoid_check93} does not hold, then $y - \mu(x) + \mu(x_i)$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_i=0$.
+If \eqref{eq:ellipsoid_check93} does hold, then we look up the batch index $b$ such that $x_i \in S_b$, and evaluate
+\begin{equation}
+f_i = V(x)\eta_b\left(y - \mu(x) + \mu(x_i)\right),
+\end{equation}
+as per \eqref{eq:fxyxp} and \eqref{eq:varphi_eval}.
 %Recall from Section \ref{sec:get_impulse_response} that $\eta_b$ is the impulse response batch that we have computed and stored.
 %Evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$ requires care, because
-When the problem is discretized with finite elements, $y - \mu(x) + \mu(x_i)$ is generally not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints.
-To evaluate $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$, we therefore determine which mesh cell contains $y - \mu(x) + \mu(x_i)$, then we evaluate the relevant finite element basis functions on that mesh cell.
+
+Note that $y - \mu(x) + \mu(x_i)$ is generally not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints. Hence,
+evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$, requires determining which mesh cell contains $y - \mu(x) + \mu(x_i)$, then evaluating the relevant finite element basis functions on that mesh cell.
 Fortunately, the mesh cell containing $y - \mu(x) + \mu(x_i)$ was already determined as a side effect of querying the AABB tree of mesh cells when we checked whether $y - \mu(x) + \mu(x_i) \in \Omega$.
@@
 Let $\{y_i\}_{i=1}^k$ be the $k$ nearest sample points to $y$, excluding points $y_i$ such that $x - \mu(y) + \mu(y_i) \notin \Omega$.
-We could form an alternative approximation of $\Phi(y,x)$ by interpolating the (point, value) pairs
-\begin{equation}
-\label{eq:phi_mu_shift_y}
-\left(y_i,f_{x,y}(y_i)\right), \qquad i=1,\dots,k,
-\end{equation}
-in the same manner as before, but with $x$ and $y$ now taking opposite roles.
-We combine both of these interpolation schemes to improve the approximation accuracy at negligible additional.
+We could form an alternative approximation of $\Phi(y,x)$ by performing the same interpolation as before, but with $x$ and $y$ taking opposite roles.
+We combine both of these interpolation schemes to improve the approximation accuracy at negligible additional cost.
 The two interpolation schemes are combined by placing function values
@@
 We solve the regularized least squares problem by QR factorization of the block $4k \times 2k$ matrix $\begin{bmatrix}\widehat{B} & I\end{bmatrix}^T$.
-We regularize the problem because of numerical issues that arise if there are points $\widehat{x}_i$ and $\widehat{y}_j$ in the interpolation that are very close to each other, a situation that arises often in practice.
-Indeed, situations where $\widehat{x}_i = \widehat{y}_j$ are inevitable when a regular grid is used to discretize the problem.
-In this case, $\widehat{B}$ is (nearly) singular, so it is numerically ill-advised to solve the standard (un-regularized) interpolation equation $\widehat{B} \widehat{w} = \widehat{f}$.
+We regularize the problem because of numerical issues that arise if there are points $\widehat{x}_i$ and $\widehat{y}_j$ in the interpolation that are equal or nearly equal, a situation that arises often in practice.
+%Indeed, situations where $\widehat{x}_i = \widehat{y}_j$ are inevitable when a regular grid is used to discretize the problem. 
+In this case, $\widehat{B}$ is (nearly) singular, so it is ill-advised to solve the un-regularized interpolation system $\widehat{B} \widehat{w} = \widehat{f}$.
+
+\subsubsection{Cost to evaluate one approximate kernel entry}
+
+In the non-symmetric case, evaluating $\widetilde{\Phi}(y,x)$ requires us to perform one $k$-nearest neighbor query of the KD-tree of sample points, $k$ point collision queries of the AABB-tree of mesh cells, and solve a $k \times k$ linear system.
+In the symmetric case, we must perform two $k$-nearest neighbor queries of the KD-tree of sample points, $2k$ point collision queries of the AABB-tree of mesh cells, and solve a $4k \times 2k$ linear least squares problem via QR factorization.
+In both cases, this requires $O(k \log r + k \log N_\text{cell} + k^3)$ work, where $r$ is the number of sample points, and $N_\text{cell}$ is the number of cells in the mesh used to discretize the problem.
+Since $k$ is small, this is cheap.
 

### 2022-03-04 paper/localpsf.tex
editing kernel entries section

@@
 %Having found the $k$ nearest valid sample points $x_i$ to $x$, we approximate $\Phi(y,x)$ using radial basis function interpolation of the following form:
-In detail, we form the approximation $\widetilde{\Phi}(y,x) \approx \Phi(y,x)$ as follows:
+In detail, we form $\widetilde{\Phi}(y,x) \approx \Phi(y,x)$ as follows:
 \begin{equation}
@@
 r^d, & d=1,3,5,\dots \\
-r^d \log r & d=2,4,6,\dots
+r^d \log r, & d=2,4,6,\dots
 \end{cases}
@@
 \end{equation*}
-be a $2k \times 2k$ block matrix and $2k \times 1$ block vectors, respectively, with the block entries of $\widehat{B}$ be given by
+be a $2k \times 2k$ block matrix and $2k \times 1$ block vectors, respectively, with the block entries of $\widehat{B}$ given by
 \begin{align*}

### 2022-03-04 paper/localpsf.tex
moved rational positive definite section to appendix
editing writing

@@
 
-Each point source $\delta_{x_i}$ in the Dirac comb is scaled by $1/V(x_i)$ so that the resulting scaled impulse responses within $\eta_k$ are comparable in magnitude.
+Each point source $\delta_{x_i}$ is scaled by $1/V(x_i)$ so that the resulting scaled impulse responses within $\eta_k$ are comparable in magnitude.
 Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section \ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $V(x_i)$ is smaller than a predetermined threshold.
@@
 Note that $y - \mu(x) + \mu(x_i)$ is generally not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints. Hence,
-evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$, requires determining which mesh cell contains $y - \mu(x) + \mu(x_i)$, then evaluating the relevant finite element basis functions on that mesh cell.
+evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$ requires determining which mesh cell contains $y - \mu(x) + \mu(x_i)$, then evaluating finite element basis functions on that mesh cell.
 Fortunately, the mesh cell containing $y - \mu(x) + \mu(x_i)$ was already determined as a side effect of querying the AABB tree of mesh cells when we checked whether $y - \mu(x) + \mu(x_i) \in \Omega$.
 
-\subsubsection{Improved interpolation for symmetric operators}
+\subsubsection{Improvement for symmetric operators}
 \label{sec:symmetric_improvements}
@@
 \end{equation*}
-at the points $y_i-y$, and interpolating the combination of these $2k$ points and values at zero.
+at the points $y_i-y$, and using radial basis functions to interpolate the combination of these $2k$ points and values at zero.
 
@@
 
-\subsection{Rational positive semi-definite modification}
-\label{sec:make_spd}
-
-In many problems of practical interest (e.g., Hessian approximation) $\Aop$ is symmetric positive semi-definite.
-However, $\mathbf{A}_H$ is generally non-symmetric and indefinite because of approximation error.
-This is undesirable.
-Symmetry and positive semi-definiteness are important properties which should be preserved if possible.
-Also, lacking these properties may prevent one from using highly effective algorithms, such as the conjugate gradient method, to perform further useful operations involving $\mathbf{A}_H$.
-In this section we describe a method for modifying $\mathbf{A}_H$ to make it symmetric positive semi-definite.
-
-Because of the modified interpolation for symmetric matrices described in Section \ref{sec:symmetric_improvements}, $\mathbf{A}_H$ is nearly (but not completely) symmetric.
-We further symmetrize $\mathbf{A}_H$ as follows:
-\begin{equation*}
-	\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
-\end{equation*}
-This step is fast because adding H-matrices is a comparatively inexpensive operation.
-
-We modify $\mathbf{A}_H^{\text{sym}}$ to make it positive semi-definite by forming a rational matrix function of the following form: 
-\begin{equation}
-	\label{eq:rational_matrix_function}
-	\mathbf{A}_H^{\text{sym}+} := \ratcoeff_0 \bm{I} + \ratcoeff_1 \mathbf{A}_H^{\text{sym}} + \ratcoeff_2 \left(\mathbf{A}_H^{\text{sym}} + \ratpole \bm{I}\right)^{-1},
-\end{equation}
-where $\bm{I}$ is the identity matrix which has the same shape as $\mathbf{A}_H^{\text{sym}}$.
-Approximation \eqref{eq:rational_matrix_function} may be written as $\mathbf{A}_H^{\text{sym}+} = \ratfct(\mathbf{A}_H^{\text{sym}})$, where $\ratfct$ is the following scalar rational function:
-\begin{equation}
-	\label{eq:rational_function_scalar}
-	\ratfct(\lambda) := \ratcoeff_0 + \ratcoeff_1 \lambda + \frac{\ratcoeff_2}{\lambda+\ratpole},
-\end{equation}
-which is extended to matrices by functional calculus.
-The scalars $\ratcoeff_0$, $\ratcoeff_1$, $\ratcoeff_2$, and $\ratpole$ are chosen so that the moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym}+}$ approximately equal the corresponding moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$, while the negative and small positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$ are modified so that $\mathbf{A}_H^{\text{sym}+}$ is positive semi-definite.
-Later in this section we will explain how $\ratcoeff_0$, $\ratcoeff_1$, $\ratcoeff_2$, and $\ratpole$ are chosen and how our choice of these scalars affects the spectral properties of $\mathbf{A}_H^{\text{sym}+}$.
-In Figure FIG we illustrate $\ratfct(\lambda)$.
-Forming $\mathbf{A}_H^{\text{sym}+}$ requires computing an H-matrix inverse, $\left(\mathbf{A}_H^{\text{sym}} + \ratpole \bm{I}\right)^{-1}$.
-While explicit computation of matrix inverses should usually be avoided, here it is computationally acceptable because $\mathbf{A}_H^{\text{sym}} + \ratpole \bm{I}$ is an H-matrix.
-Inversion of H-matrices is a routine procedure which only requires $O(\fedim \log(\fedim)^2)$ work \cite{Hmatrixinverse}.
-
-We choose $\ratcoeff_0, \ratcoeff_1, \ratcoeff_2$, and $\ratpole$, to be the solution to the following system of equations:
-\begin{subequations}
-	\label{eq:four_eq_system}
-	\begin{align}
-		\ratfct(0) &= 0 \label{eq:desired_property_1} \\
-		\ratfct'(0) &= 0 \label{eq:desired_property_2} \\
-		\ratfct(\lambda_\text{min}) &= |\lambda_\text{min}| \label{eq:desired_property_3} \\
-		\ratfct(\lambda_\text{max}) &= \lambda_\text{max}, \label{eq:desired_property_4}
-	\end{align}
-\end{subequations}
-where $\lambda_\text{min}$ and $\lambda_\text{max}$ are the smallest and largest eigenvalues of $\mathbf{A}_H^{\text{sym}}$, respectively.
-We estimate $\lambda_\text{min}$ and $\lambda_\text{max}$ using the implicitly restarted Lanczos method \cite{lanczos,scipy}.
-We assume that $\lambda_\text{min} < 0$; if $\lambda_\text{min} \ge 0$, then $\mathbf{A}_H^{\text{sym}}$ is already positive semi-definite, so we do not need to modify it with a rational function.
-We also require $\ratpole > |\lambda_\text{min}|$, so that the pole of $\ratfct$ is not in $[\lambda_\text{min},\lambda_\text{max}]$.
-We describe how we solve system \eqref{eq:four_eq_system} in Appendix \ref{app:solve_rational_system}.
-Solving \eqref{eq:four_eq_system} reduces to solving a smooth one-dimensional optimization problem, which is computationally trivial.
-
-Together, equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, \eqref{eq:desired_property_3},  and \eqref{eq:desired_property_4}, and the condition $\ratpole > |\lambda_\text{min}|$, imply that $\ratfct(0)=0$ is the unique minimum of $\ratfct$ on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that $\ratfct$ is non-negative on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that  $\AmatPcSymPlus$ is positive semi-definite.
-Furthermore, \eqref{eq:desired_property_1} implies that the null space of $\mathbf{A}_H^{\text{sym+}}$ contains the null-space of $\mathbf{A}_H^{\text{sym}}$.
-This is important when the operator $\Aop$ has a large (or infinite) dimensional null space, or a spectrum that clusters at zero.
-Hessians in ill-posed distributed parameter inverse problems typically have both of these properties.
-Equation \eqref{eq:desired_property_4} forces the moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym+}}$ to be close to the corresponding moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
-I.e., $\ratfct$ modifies the ``important'' part of the spectrum of $\mathbf{A}_H^{\text{sym}}$ as little as possible.
-Equation \eqref{eq:desired_property_3} ensures that the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $\ratfct$.
-There is a tradeoff: if we choose a larger value for $\ratfct(\lambda_\text{min})$, then the positive eigenvalues of $\mathbf{A}_H^{\text{sym+}}$ will be closer to the corresponding positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$ (good), but the erroneous negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ will be amplified in magnitude (bad).
-This is shown in Figure FIG.
-We find that $\ratfct(\lambda_\text{min})=|\lambda_\text{min}|$ is an appropriate balance between these competing interests.
-
 \subsection{Computational cost}
@@
 
-\section{Solving for rational function scalars}
+\subsection{Rational positive semi-definite modification}
+\label{sec:make_spd}
+
+In many problems of practical interest (e.g., Hessian approximation) $\Aop$ is symmetric positive semi-definite.
+However, $\mathbf{A}_H$ is generally non-symmetric and indefinite because of approximation error.
+This is undesirable.
+Symmetry and positive semi-definiteness are important properties which should be preserved if possible.
+Also, lacking these properties may prevent one from using highly effective algorithms, such as the conjugate gradient method, to perform further useful operations involving $\mathbf{A}_H$.
+In this section we describe a method for modifying $\mathbf{A}_H$ to make it symmetric positive semi-definite.
+
+Because of the modified interpolation for symmetric matrices described in Section \ref{sec:symmetric_improvements}, $\mathbf{A}_H$ is nearly (but not completely) symmetric.
+We further symmetrize $\mathbf{A}_H$ as follows:
+\begin{equation*}
+\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
+\end{equation*}
+This step is fast because adding H-matrices is a comparatively inexpensive operation.
+
+We modify $\mathbf{A}_H^{\text{sym}}$ to make it positive semi-definite by forming a rational matrix function of the following form: 
+\begin{equation}
+\label{eq:rational_matrix_function}
+\mathbf{A}_H^{\text{sym}+} := \ratcoeff_0 \bm{I} + \ratcoeff_1 \mathbf{A}_H^{\text{sym}} + \ratcoeff_2 \left(\mathbf{A}_H^{\text{sym}} + \ratpole \bm{I}\right)^{-1},
+\end{equation}
+where $\bm{I}$ is the identity matrix which has the same shape as $\mathbf{A}_H^{\text{sym}}$.
+Approximation \eqref{eq:rational_matrix_function} may be written as $\mathbf{A}_H^{\text{sym}+} = \ratfct(\mathbf{A}_H^{\text{sym}})$, where $\ratfct$ is the following scalar rational function:
+\begin{equation}
+\label{eq:rational_function_scalar}
+\ratfct(\lambda) := \ratcoeff_0 + \ratcoeff_1 \lambda + \frac{\ratcoeff_2}{\lambda+\ratpole},
+\end{equation}
+which is extended to matrices by functional calculus.
+The scalars $\ratcoeff_0$, $\ratcoeff_1$, $\ratcoeff_2$, and $\ratpole$ are chosen so that the moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym}+}$ approximately equal the corresponding moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$, while the negative and small positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$ are modified so that $\mathbf{A}_H^{\text{sym}+}$ is positive semi-definite.
+Later in this section we will explain how $\ratcoeff_0$, $\ratcoeff_1$, $\ratcoeff_2$, and $\ratpole$ are chosen and how our choice of these scalars affects the spectral properties of $\mathbf{A}_H^{\text{sym}+}$.
+In Figure FIG we illustrate $\ratfct(\lambda)$.
+Forming $\mathbf{A}_H^{\text{sym}+}$ requires computing an H-matrix inverse, $\left(\mathbf{A}_H^{\text{sym}} + \ratpole \bm{I}\right)^{-1}$.
+While explicit computation of matrix inverses should usually be avoided, here it is computationally acceptable because $\mathbf{A}_H^{\text{sym}} + \ratpole \bm{I}$ is an H-matrix.
+Inversion of H-matrices is a routine procedure which only requires $O(\fedim \log(\fedim)^2)$ work \cite{Hmatrixinverse}.
+
+We choose $\ratcoeff_0, \ratcoeff_1, \ratcoeff_2$, and $\ratpole$, to be the solution to the following system of equations:
+\begin{subequations}
+	\label{eq:four_eq_system}
+	\begin{align}
+	\ratfct(0) &= 0 \label{eq:desired_property_1} \\
+	\ratfct'(0) &= 0 \label{eq:desired_property_2} \\
+	\ratfct(\lambda_\text{min}) &= |\lambda_\text{min}| \label{eq:desired_property_3} \\
+	\ratfct(\lambda_\text{max}) &= \lambda_\text{max}, \label{eq:desired_property_4}
+	\end{align}
+\end{subequations}
+where $\lambda_\text{min}$ and $\lambda_\text{max}$ are the smallest and largest eigenvalues of $\mathbf{A}_H^{\text{sym}}$, respectively.
+We estimate $\lambda_\text{min}$ and $\lambda_\text{max}$ using the implicitly restarted Lanczos method \cite{lanczos,scipy}.
+We assume that $\lambda_\text{min} < 0$; if $\lambda_\text{min} \ge 0$, then $\mathbf{A}_H^{\text{sym}}$ is already positive semi-definite, so we do not need to modify it with a rational function.
+We also require $\ratpole > |\lambda_\text{min}|$, so that the pole of $\ratfct$ is not in $[\lambda_\text{min},\lambda_\text{max}]$.
+We describe how we solve system \eqref{eq:four_eq_system} in Appendix \ref{app:solve_rational_system}.
+Solving \eqref{eq:four_eq_system} reduces to solving a smooth one-dimensional optimization problem, which is computationally trivial.
+
+Together, equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, \eqref{eq:desired_property_3},  and \eqref{eq:desired_property_4}, and the condition $\ratpole > |\lambda_\text{min}|$, imply that $\ratfct(0)=0$ is the unique minimum of $\ratfct$ on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that $\ratfct$ is non-negative on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that  $\AmatPcSymPlus$ is positive semi-definite.
+Furthermore, \eqref{eq:desired_property_1} implies that the null space of $\mathbf{A}_H^{\text{sym+}}$ contains the null-space of $\mathbf{A}_H^{\text{sym}}$.
+This is important when the operator $\Aop$ has a large (or infinite) dimensional null space, or a spectrum that clusters at zero.
+Hessians in ill-posed distributed parameter inverse problems typically have both of these properties.
+Equation \eqref{eq:desired_property_4} forces the moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym+}}$ to be close to the corresponding moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
+I.e., $\ratfct$ modifies the ``important'' part of the spectrum of $\mathbf{A}_H^{\text{sym}}$ as little as possible.
+Equation \eqref{eq:desired_property_3} ensures that the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $\ratfct$.
+There is a tradeoff: if we choose a larger value for $\ratfct(\lambda_\text{min})$, then the positive eigenvalues of $\mathbf{A}_H^{\text{sym+}}$ will be closer to the corresponding positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$ (good), but the erroneous negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ will be amplified in magnitude (bad).
+This is shown in Figure FIG.
+We find that $\ratfct(\lambda_\text{min})=|\lambda_\text{min}|$ is an appropriate balance between these competing interests.
+
+\subsection{Solving for rational function scalars}
 \label{app:solve_rational_system}

### 2022-03-04 paper/localpsf.tex
writing is choppy after all this editing

@@
 That is, sample points such that $y-\mu(x)+\mu(x_i) \notin \Omega$.
-In numerical experiments we use $k=10$.
-We efficiently find the $k$ nearest sample points by querying a precomputed KD-tree \cite{kdtree} of all sample points, and check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
-%Per the discussion in Section \ref{eq:local_mean_displacement_invariance}, $\Phi(y,x) \approx f_i$ when $x_i$ is close to $x$.
-The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section \ref{eq:local_mean_displacement_invariance}, is a form of preconditioning that improves the quality of this approximation in situations where $V(x)$ varies widely.
+One should choose a small $k$; in numerical experiments we use $k=10$.
+We find the $k$ nearest sample points by querying a precomputed KD-tree \cite{kdtree} of all sample points.
+We check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
+Per Section \ref{eq:local_mean_displacement_invariance}, $\Phi(y,x) \approx f_i$ when $x_i$ is close to $x$.
+The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section \ref{eq:local_mean_displacement_invariance}, improves the approximation if $V(x)$ varies widely.
 

### 2022-03-04 paper/localpsf.tex
add back some details to RBF section

@@
 Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions.
-We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Theorem \ref{thm:vol_mean_cov} and summarize the method for computing these moments in Algorithm \ref{alg:varhpi_mean_cov}.
-
-\begin{theorem}[Impulse response moments]
+We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Proposition \ref{thm:vol_mean_cov} and summarize the method for computing these moments in Algorithm \ref{alg:varhpi_mean_cov}.
+
+\begin{prop}[Impulse response moments]
 	\label{thm:vol_mean_cov}
@@
 	for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$.
-\end{theorem}
-
-The proof of Theorem \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}. The proof is straightforward, and follows directly from the definitions.
+\end{prop}
+
+The proof of Proposition \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}. The proof is straightforward, and follows directly from the definitions.
 
@@
 We check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
-Per Section \ref{eq:local_mean_displacement_invariance}, $\Phi(y,x) \approx f_i$ when $x_i$ is close to $x$.
-The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section \ref{eq:local_mean_displacement_invariance}, improves the approximation if $V(x)$ varies widely.
+Per the discussion in Section \ref{eq:local_mean_displacement_invariance}, if we ignore the factor $\frac{V(x)}{V(x_i)}$, then $\Phi(y,x) \approx f_i$ when $x_i$ is close to $x$.
+The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section \ref{eq:local_mean_displacement_invariance}, is a form of preconditioning that improves this approximation if $V(x)$ varies widely.
 
@@
 \end{equation}
-at the point $x$ using radial basis function interpolation.
+at the point $x$. This is done using the following radial basis function interpolation scheme:
 %The approximation $\widetilde{\Phi}(y,x)$ is the value of this interpolation at $x$.
@@
 %Having found the $k$ nearest valid sample points $x_i$ to $x$, we approximate $\Phi(y,x)$ using radial basis function interpolation of the following form:
-In detail, we form $\widetilde{\Phi}(y,x) \approx \Phi(y,x)$ as follows:
-\begin{equation}
-\widetilde{\Phi}(y,x) := \sum_{i=1}^k w_i~ b\left(\|x-x_i\|\right).
-\end{equation}
-Here $b_i$ are radial basis functions and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
+%In detail, we form $\widetilde{\Phi}(y,x) \approx \Phi(y,x)$ as follows:
+\begin{equation}
+\Phi(y,x) \approx \widetilde{\Phi}(y,x) := \sum_{i=1}^k w_i~ b\left(\|x-x_i\|\right),
+\end{equation}
+where $b_i$ are radial basis functions and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
 \begin{equation}
@@
 \end{equation}
-for $r>0$, and $b_i(0):=0$. For more details on radial basis function interpolation and polyharmonic splines, see \cite{rbf}.
+for $r>0$, and $b_i(0):=0$.
+Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a certain least-squares sense.
+For more details on radial basis function interpolation and polyharmonic splines, see \cite{rbf}.
 %\begin{equation}
@@
 In the symmetric case, we must perform two $k$-nearest neighbor queries of the KD-tree of sample points, $2k$ point collision queries of the AABB-tree of mesh cells, and solve a $4k \times 2k$ linear least squares problem via QR factorization.
-In both cases, this requires $O(k \log r + k \log N_\text{cell} + k^3)$ work, where $r$ is the number of sample points, and $N_\text{cell}$ is the number of cells in the mesh used to discretize the problem.
+Both of these cases require $O(k \log r + k \log N_\text{cell} + k^3)$ work, where $r$ is the number of sample points, and $N_\text{cell}$ is the number of cells in the mesh used to discretize the problem.
 Since $k$ is small, this is cheap.
@@
 We convert $\mathbf{\Phi}$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{HLIBPRO}.
-Although $\mathbf{\Phi}$ is a dense $\fedim \times \fedim$ matrix, constructing the H-matrix approximation only requires evaluation of $O(\hmatrixrank^2 \fedim \log \fedim)$ entries of $\mathbf{\Phi}$, and these entries are computed via interpolation procedure described in Section \ref{sec:approximate_kernel_entries}.
+Although $\mathbf{\Phi}$ is a dense $\fedim \times \fedim$ matrix, constructing the H-matrix approximation only requires evaluation of $O(\hmatrixrank^2 \fedim \log \fedim)$ entries $\mathbf{\Phi}_{ij} = \widetilde{\Phi}(p_i, p_j)$, and these entries are computed via the procedure described in Section \ref{sec:approximate_kernel_entries}.
 Here $\hmatrixrank$ is the rank of the highest-rank block in the H-matrix.
@@
 
-In detail, the process of constructing an H-matrix representation of $\AkerPcMat$ proceeds as follows.
+The process of constructing an H-matrix representation of $\mathbf{\Phi}$ proceeds as follows.
 First, we construct hierarchical partitionings of the degrees of freedom for the columns and rows of the matrix (cluster trees, Section \ref{sec:cluster_trees}).
@@
 
-\begin{proof}[Proof of Theorem \ref{thm:vol_mean_cov}]
+\begin{proof}[Proof of Proposition \ref{thm:vol_mean_cov}]
 	Let $v \in L^2$ be arbitrary. We have

### 2022-03-04 paper/localpsf.tex
add back some details to RBF section

@@
 where $\{x_i\}_{i=1}^k$ are the $k$ nearest sample points to $x$, excluding sample points for which $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ is undefined.
-That is, sample points such that $y-\mu(x)+\mu(x_i) \notin \Omega$.
-One should choose a small $k$; in numerical experiments we use $k=10$.
+That is, excluding sample points such that $y-\mu(x)+\mu(x_i) \notin \Omega$.
+The number $k$ is a small user defined parameter; in numerical experiments we use $k=10$.
 We find the $k$ nearest sample points by querying a precomputed KD-tree \cite{kdtree} of all sample points.
 We check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
-Per the discussion in Section \ref{eq:local_mean_displacement_invariance}, if we ignore the factor $\frac{V(x)}{V(x_i)}$, then $\Phi(y,x) \approx f_i$ when $x_i$ is close to $x$.
-The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section \ref{eq:local_mean_displacement_invariance}, is a form of preconditioning that improves this approximation if $V(x)$ varies widely.
-
-We approximate $\Phi(y,x)$
-%compute the approximate kernel entry,
-%\begin{equation*}
-%\widetilde{\Phi}(y,x) \approx \Phi(y,x)
-%\end{equation*}
-by interpolating the (point,value) pairs
-\begin{equation}
-\left(x_i, f_i\right), \qquad i=1,\dots,k
-\end{equation}
-at the point $x$. This is done using the following radial basis function interpolation scheme:
+
+The idea here is that $\Phi(y,x) \approx f_i$ when $x_i$ is near $x$, so we approximate $\Phi(y,x)$ by interpolating the (point,value) pairs
+\begin{equation}
+\left(x_i, f_i\right), \qquad i=1,\dots,k,
+\end{equation}
+at the point $x$.
+To see that $\Phi(y,x) \approx f_i$, recall from Section \ref{eq:local_mean_displacement_invariance} that $\Phi(y,x) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ if $x_i$ is near $x$.
+Inclusion of the extra factor $\frac{V(x)}{V(x_i)}$ in \eqref{eq:fxyxp} is a form of preconditioning that makes this approximation more robust in the case where $V(x)$ varies widely.
+%Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section \ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $V(x_i)$ is smaller than a predetermined threshold.
+
+\subsubsection{Radial basis function interpolation}
+
+We use the following radial basis function interpolation scheme:
 %The approximation $\widetilde{\Phi}(y,x)$ is the value of this interpolation at $x$.

### 2022-03-04 paper/localpsf.tex
add back some details to RBF section

@@
 
-\subsubsection{Radial basis function interpolation}
+%\subsubsection{Radial basis function interpolation}
 
@@
 \end{equation}
-where $b_i$ are radial basis functions and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
+Here $b_i$ are radial basis functions and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
 \begin{equation}
@@
 We could form an alternative approximation of $\Phi(y,x)$ by performing the same interpolation as before, but with $x$ and $y$ taking opposite roles.
-We combine both of these interpolation schemes to improve the approximation accuracy at negligible additional cost.
-The two interpolation schemes are combined by placing function values
+Both of these interpolation schemes can be combined to improve the approximation accuracy at negligible additional cost.
+We combine these interpolation schemes by placing function values
 \begin{equation*}
@@
 \end{equation*}
-at the points $y_i-y$, and using radial basis functions to interpolate the combination of these $2k$ points and values at zero.
+at the points $y_i-y$, and interpolating the combination of these $2k$ points and values at zero.
 

### 2022-03-07 paper/localpsf.tex
commit changes before meeting

@@
 Furthermore, approximation \eqref{eq:translate3} can hold in situations where neither \eqref{eq:translate1} nor \eqref{eq:translate2} holds.
-For example, \eqref{eq:translate3} will hold when an operator is locally translation-invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case.
+For example, because the expected value is a linear operator, \eqref{eq:translate3} will hold when an $\Aop$ is locally translation-invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case.
 Additionally, \eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions.

### 2022-03-07 paper/localpsf.tex
paper edits

@@
 
+\subsection{Local mean-displacement invariance}
+\label{eq:local_mean_displacement_invariance}
+
+Let $x$ and $x'$ be points in $\Omega$ that are ``close'' to each other, and consider the following three approximations, ordered by increasing level of sophistication:
+\begin{align}
+\phi_x(y) &\approx \phi_{x'}(y) \label{eq:translate1}\\
+\phi_x(y) &\approx \phi_{x'}(y-x+x') \label{eq:translate2}\\
+\phi_x(y) &\approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right). \label{eq:translate3}
+\end{align}
+Approximation \eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close. 
+%This approximation follows from the assumption that $\Phi$ is smooth. 
+Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank interpolative factorization of $\Phi$.
+% of the form
+%\begin{equation}
+%	\label{eq:low_rank_interpolation}
+%	\Phi(y,x) \approx \sum_{i=1}^r w_i(x) \phi_{x_i}(y)
+%\end{equation}
+%where $\phi_{x_i}$ are known impulse responses at a collection of points $x_i$ scattered throughout $\Omega$, and $w_i$ are smooth weighting functions that interpolate the $\phi_{x_i}$. 
+%In problems of practical interest, \eqref{eq:translate1} typically holds \emph{asymptotically} in the limit as $x'$ approaches $x$. 
+However, 
+%even if $\Phi$ is smooth (which is typically the case asymptotically),
+%the relevant Lipschitz constant for continuity, and hence 
+the required rank in the low rank approximation can be large, which makes algorithms based on assumption \eqref{eq:translate1} computationally expensive.
+This is the case, for example, for Hessians in PDE constrained inverse problems with highly informative data, as discussed in Section \ref{sec:PDE_hessian_motivation}.
+%As the data become more informative, typically the ``mass'' of $\Phi(y,x)$ concentrates nearer and nearer to the ``diagonal'' $y=x$, the Lipschitz constant increases, and the required rank in the approximation increases.
+%Moreover, if $\Aop$ is numerically low rank, then one would typically use more accurate and robust matrix-free methods for forming low-rank approximations, such as randomized SVD or Lanczos.
+
+Operators that satisfy \eqref{eq:translate2} are called ``locally translation-invariant'' because integral kernel entries $\Phi(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$.
+Locally translation-invariant operators act like convolutions locally, and can therefore be well-approximated by so-called product-convolution approximations---weighted sums of convolution operators with spatially varying weights.
+Product-convolution approximations can be high-rank even if the number of convolution operators used in the approximation is small.
+
+% Kernels for locally translation invariant operators can be well-approximated by decompositions of the form
+%\begin{equation}
+%	\Phi(y,x) \approx \sum_{i=1}^r w_i(x) \phi_{x_i}(y-x)
+%\end{equation}
+%which correspond to operator approximations of the form
+%\begin{equation}
+%\label{eq:product_convolution}
+%\left(\Aop~u\right)(v) \approx \left( v,~ \sum_{i=1}^\convrank \convkernel_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)}.
+%\end{equation}
+%Here $f \ast g$ denotes convolution, and $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$ (i.e., $\left(f \cdot g\right)(x) = f(x) g(x)$).
+Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution.
+Product-convolution approximations can be high-rank even if $r$ is small.
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale=1.0]{mean_displacement_invariance.pdf}
+	\end{center}
+	\caption{An operator is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \mu(x) + \mu(x')\right)$ when $x$ is close to $x'$.}
+	\label{fig:mean_displacement_invariance}
+\end{figure}
+
+Our method is based on approximation \eqref{eq:translate3}, which is an improvement upon \eqref{eq:translate1} and \eqref{eq:translate2} that generalizes both.
+We call \eqref{eq:translate3} ``local mean-displacement invariance,'' and illustrate \eqref{eq:translate3} in Figure \ref{fig:mean_displacement_invariance}.
+On one hand, if \eqref{eq:translate1} holds, then $\mu(x) \approx \mu(x')$, and so \eqref{eq:translate3} holds.
+On the other hand, translating a distribution translates its mean, so if \eqref{eq:translate2} holds, then $\mu(x')-\mu(x) \approx x' - x$, so again \eqref{eq:translate3} holds.
+But approximation \eqref{eq:translate3} can hold in situations where neither \eqref{eq:translate1} nor \eqref{eq:translate2} holds.
+For example, because the expected value is a linear operator, \eqref{eq:translate3} will hold when an $\Aop$ is locally translation-invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case.
+Additionally, \eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions.
+In contrast, \eqref{eq:translate2} does not naturally generalize to operators that map between function spaces on different domains, because the formula $y-x+x'$ requires vectors in $\Omega_2$ and $\Omega_1$ to be added together.
+
 \subsection{Overview of the method}
@@
 Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation.
-The details of the interpolation will be described in Section SECTION, but the key principle behind the interpolation is an assumption that we call \emph{local mean-displacement invariance}.
-Specifically, we make the assumption that
-\begin{equation}
-\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))
-\end{equation}
-if $x_i$ is close to $x$.
-This is shown in Figure FIG.
-We approximate $\phi_x(y)$ by a weighted linear combination of the values $\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for the a small number of sample points $x_i$ near $x$.
-The weights in the linear combination are determined via radial basis function interpolation.
+The details of the interpolation will be described in Section \ref{sec:approximate_kernel_entries}, but the key principle behind the interpolation is the local mean displacement invariance assumption discussed in Section \ref{eq:local_mean_displacement_invariance}.
+%Specifically, we make the assumption that
+%\begin{equation}
+%\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))
+%\end{equation}
+%if $x_i$ is close to $x$. This is shown in Figure \ref{fig:mean_displacement_invariance}. Motivation for this assumption is discussed in Section \ref{eq:local_mean_displacement_invariance}. 
+We approximate $\Phi(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for a collection of sample points $x_i$ near $x$.
+The weights in the linear combination are determined via preconditioned radial basis function interpolation.
 
@@
 
-Often we are interested in approximating operators $\Aop$ that are symmetric, or symmetric positive semi-definite.
-If $\Aop$ is symmetric, impulse responses of $\Aop$ are also impulse responses of $\Aop^T$, which may be interpolated in the variable $y$ instead of $x$ to yield a different approximation of $\Phi(y,x)$.
-In this case, we combine information from impulse responses of both $\Aop$ and $\Aop^T$ in the interpolation procedure for approximating $\Phi(y,x)$, to get a more accurate approximation at negligible additional cost.
+%Often we are interested in approximating operators $\Aop$ that are symmetric, or symmetric positive semi-definite.
+
+If $\Aop$ is symmetric, we take advantage of symmetry to construct a more accurate approximation at negligible additional cost. In this case, 
+%impulse responses of $\Aop$ are also impulse responses of $\Aop^T$, which may be interpolated in the variable $y$ instead of $x$ to yield a different approximation of $\Phi(y,x)$. 
+we combine information from impulse responses of both $\Aop$ and $\Aop^T$ in the interpolation procedure.
+% for approximating $\Phi(y,x)$, to get a more accurate approximation at negligible additional cost.
 Intuitively, impulse responses of $\Aop$ can be thought of as ``columns'' of $\Phi$, impulse responses of $\Aop^T$ can be thought of as ``rows'' of $\Phi$, and $\Phi(y,x)$ can be thought of as an unknown entry of $\Phi$.
-When $\Aop$ is symmetric, computing a column of $\Aop$ also gives us a row of $\Aop$ for free, so we approximate $\Phi(y,x)$ using information from a combination of known columns and known rows, rather than just known columns.
-
-When $\Aop$ is symmetric positive semi-definite, $\mathbf{A}_H$ may be indefinite due to errors in the approximation.
+Therefore, when $\Aop$ is symmetric, computing a column of $\Aop$ also gives us a row of $\Aop$ for free, so we approximate $\Phi(y,x)$ using information from a combination of known columns and rows, rather than just known columns.
+
+When $\Aop$ is positive semi-definite, $\mathbf{A}_H$ may be indefinite due to errors in the approximation.
 In this case, we modify the H-matrix representation of $\mathbf{A}_H$ to make it positive semi-definite, by applying a specially chosen rational matrix function to $\mathbf{A}_H$.
+This is described in Appendix \ref{sec:make_spd}.
 
@@
 
-\subsection{Local mean-displacement invariance}
-\label{eq:local_mean_displacement_invariance}
-
-Let $x$ and $x'$ be points in $\Omega$ that are close to each other, and consider the following three approximations, ordered by increasing level of sophistication:
-\begin{align}
-	\phi_x(y) &\approx \phi_{x'}(y) \label{eq:translate1}\\
-	\phi_x(y) &\approx \phi_{x'}(y-x+x') \label{eq:translate2}\\
-	\phi_x(y) &\approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right). \label{eq:translate3}
-\end{align}
-We use approximation \eqref{eq:translate3}, which we call ``locally mean-displacement invariance.'' This is illustrated in Figure \ref{fig:mean_displacement_invariance}.
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[scale=1.0]{mean_displacement_invariance.pdf}
-	\end{center}
-	\caption{An operator is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \mu(x) + \mu(x')\right)$ when $x$ is close to $x'$.}
-	\label{fig:mean_displacement_invariance}
-\end{figure}
-
-Approximation \eqref{eq:translate1} follows from the assumption that $\Phi$ is smooth.
-Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank interpolative factorization of $\Phi$.
-% of the form
-%\begin{equation}
-%	\label{eq:low_rank_interpolation}
-%	\Phi(y,x) \approx \sum_{i=1}^r w_i(x) \phi_{x_i}(y)
-%\end{equation}
-%where $\phi_{x_i}$ are known impulse responses at a collection of points $x_i$ scattered throughout $\Omega$, and $w_i$ are smooth weighting functions that interpolate the $\phi_{x_i}$. 
-%In problems of practical interest, \eqref{eq:translate1} typically holds \emph{asymptotically} in the limit as $x'$ approaches $x$. 
-However, even if $\Phi$ is smooth (which is typically the case asymptotically) the relevant Lipschitz constant for continuity, and hence the required rank in the low rank approximation, can be large, which makes algorithms based on assumption \eqref{eq:translate1} computationally expensive.
-This is the case, for example, for Hessians in PDE constrained inverse problems with highly informative data.
-As the data become more informative, typically the ``mass'' of $\Phi(y,x)$ concentrates nearer and nearer to the ``diagonal'' $y=x$, the Lipschitz constant increases, and the required rank in the approximation increases.
-%Moreover, if $\Aop$ is numerically low rank, then one would typically use more accurate and robust matrix-free methods for forming low-rank approximations, such as randomized SVD or Lanczos.
-
-Operators that satisfy \eqref{eq:translate2} are called ``locally translation-invariant'' because integral kernel entries $\Phi(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$.
-Locally translation-invariant operators act like convolutions locally, and can therefore be well-approximated by so-called product-convolution approximations---weighted sums of convolution operators with spatially varying weights.
-Product-convolution approximations can be high-rank even if the number of convolution operators used in the approximation is small.
-
-% Kernels for locally translation invariant operators can be well-approximated by decompositions of the form
-%\begin{equation}
-%	\Phi(y,x) \approx \sum_{i=1}^r w_i(x) \phi_{x_i}(y-x)
-%\end{equation}
-%which correspond to operator approximations of the form
-%\begin{equation}
-%\label{eq:product_convolution}
-%\left(\Aop~u\right)(v) \approx \left( v,~ \sum_{i=1}^\convrank \convkernel_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)}.
-%\end{equation}
-%Here $f \ast g$ denotes convolution, and $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$ (i.e., $\left(f \cdot g\right)(x) = f(x) g(x)$).
-Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution.
-Product-convolution approximations can be high-rank even if $r$ is small.
-
-We use approximation \eqref{eq:translate3}, which is an improvement upon \eqref{eq:translate1} and \eqref{eq:translate2} that generalizes both.
-On one hand, if \eqref{eq:translate1} holds, then $\mu(x) \approx \mu(x_i)$, and so \eqref{eq:translate3} holds.
-On the other hand, translating a distribution translates its mean, so if \eqref{eq:translate2} holds, then $\mu(x_i)-\mu(x) \approx x_i - x$, so again \eqref{eq:translate3} holds.
-Furthermore, approximation \eqref{eq:translate3} can hold in situations where neither \eqref{eq:translate1} nor \eqref{eq:translate2} holds.
-For example, because the expected value is a linear operator, \eqref{eq:translate3} will hold when an $\Aop$ is locally translation-invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case.
-Additionally, \eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions.
-In contrast, \eqref{eq:translate2} does not naturally generalize to operators that map between function spaces on different domains, because the formula $y-x+x'$ requires $y \in \Omega_2$ and $-x+x' \in \Omega_1$ to be added together.
-
 \subsection{Existing work}
@@
 \label{sec:method}
-In this section we describe how we compute the impulse response moments $V,\mu,\Sigma$ and use these moments to estimate the supports of the impulse responses $\phi_x$ (Section \ref{eq:mean_and_covariance_estimation}), the greedy ellipsoid packing procedure we use to choose batches of sample points $x_i$ (Section \ref{sec:sample_point_selection}), how we compute impulse responses, $\phi_{x_i}$, in batches (Section \ref{sec:get_impulse_response}), how we interpolate the $\phi_{x_i}$ to approximate integral kernel entries $\Phi(y,x)$ (Section \ref{sec:approximate_kernel_entries}), finite element discretization of $\Phi$ (Section \ref{sec:finite_element_kernel}), and how we construct the H-matrix approximation $\mathbf{A}_h$ of the operator $\mathcal{A}$.
+In this section we describe how we compute the impulse response moments $V,\mu,\Sigma$ and use these moments to estimate the supports of the impulse responses $\phi_x$ (Section \ref{sec:moments}), the greedy ellipsoid packing procedure we use to choose batches of sample points $x_i$ (Section \ref{sec:sample_point_selection}), how we compute impulse responses, $\phi_{x_i}$, in batches (Section \ref{sec:get_impulse_response}), how we interpolate the $\phi_{x_i}$ to approximate integral kernel entries $\Phi(y,x)$ (Section \ref{sec:approximate_kernel_entries}), finite element discretization of $\Phi$ (Section \ref{sec:finite_element_kernel}), and how we construct the H-matrix approximation $\mathbf{A}_h$ of the operator $\mathcal{A}$.
 We also describe how to (optionally) modify $\mathbf{A}_h$ to force it to be symmetric positive semi-definite.
@@
 \section{Impulse response moments}
+\label{sec:moments}
 
@@
 at the point $x$.
-To see that $\Phi(y,x) \approx f_i$, recall from Section \ref{eq:local_mean_displacement_invariance} that $\Phi(y,x) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ if $x_i$ is near $x$.
+To see that $\Phi(y,x) \approx f_i$, recall from Section \ref{eq:local_mean_displacement_invariance} that $\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ if $x_i$ is near $x$.
 Inclusion of the extra factor $\frac{V(x)}{V(x_i)}$ in \eqref{eq:fxyxp} is a form of preconditioning that makes this approximation more robust in the case where $V(x)$ varies widely.

### 2022-03-07 paper/localpsf.tex
paper edits

@@
 We approximate $\Phi(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for a collection of sample points $x_i$ near $x$.
-The weights in the linear combination are determined via preconditioned radial basis function interpolation.
+The weights are determined by preconditioned radial basis function interpolation.
 
@@
 One may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
+These H-matrix methods are fast and scale well to large problems.
 WHY DO THESE OPERATIONS: KRYLOV SOLVERS, SAMPLING STRATEGIES, CAN'T DO WITH A.
-These H-matrix methods are fast and scale well to large problems.
 
@@
 In \cite{RESOLUTION}, the width of $\convkernel_p$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
-In this paper, rather than probing $\Aop^T$ with random noise functions, we use specific constant, linear, and quadratic functions.
+In this paper, rather than probing $\Aop^T$ with random noise functions, we use polynomial functions.
 Our method estimates the support of $\convkernel_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
@@
 Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}.
-Here we follow our previous work \cite{PRODCONVMYPAPER}, in which we convert the product-convolution approximation to H-matrix format.
+Here we follow our previous work \cite{PRODCONVMYPAPER}, in which we converted product-convolution approximations into H-matrix format.
 
@@
 In \cite{DEMANETSEISMIC}, matrix probing \cite{MATRIXPROBING} is used to approximate the Hessian in a seismic inverse problem as the sum of simple pseudodifferential operators.
-Although it is not explicitly mentioned, the approximation in \cite{DEMANETSEISMIC} could be interpreted as a convolution-product interpolation, which is like a product-convolution approximation, except the order of pointwise multiplications and convolutions is reversed in \eqref{eq:product_convolution}.
+Although it is not explicitly mentioned, the approximation in \cite{DEMANETSEISMIC} could be interpreted as a convolution-product approximation (a product-convolution approximation of the transpose of the operator).
+%, which is like a product-convolution approximation, except the order of pointwise multiplications and convolutions is reversed in \eqref{eq:product_convolution}.
 
@@
 \label{sec:method}
-In this section we describe how we compute the impulse response moments $V,\mu,\Sigma$ and use these moments to estimate the supports of the impulse responses $\phi_x$ (Section \ref{sec:moments}), the greedy ellipsoid packing procedure we use to choose batches of sample points $x_i$ (Section \ref{sec:sample_point_selection}), how we compute impulse responses, $\phi_{x_i}$, in batches (Section \ref{sec:get_impulse_response}), how we interpolate the $\phi_{x_i}$ to approximate integral kernel entries $\Phi(y,x)$ (Section \ref{sec:approximate_kernel_entries}), finite element discretization of $\Phi$ (Section \ref{sec:finite_element_kernel}), and how we construct the H-matrix approximation $\mathbf{A}_h$ of the operator $\mathcal{A}$.
-We also describe how to (optionally) modify $\mathbf{A}_h$ to force it to be symmetric positive semi-definite.
-The complete algorithm is shown in Algorithm \ref{alg:construct_Atilde}.
+In this section we detail our method for computing the H-matrix approximation, $\mathbf{A}_H$, of $\Aop$. This section is organized as follows:
+\begin{description}
+	\item[Section \ref{eq:mean_and_covariance_estimation}:] We describe our ellipsoid estimate for the support of $\phi_x$.
+	\item[Section \ref{sec:moments}:] We show how to compute $V,\mu,\Sigma$.
+	\item[Section \ref{sec:sample_point_selection}:] We present the greedy ellipsoid packing procedure used to choose batches, $\pointbatch_b$, of sample points $x_i$.
+	\item[Section \ref{sec:get_impulse_response}:] We describe how to compute $\phi_{x_i}$'s, in batches, $\eta_b$.
+	\item[Section \ref{sec:approximate_kernel_entries}:] We show how we interpolate $\phi_{x_i}$'s to approximate $\Phi(y,x)$.
+	\item[Section \ref{sec:finite_element_kernel}:] We discuss finite element discretization of $\Phi$.
+	\item[Section \ref{sec:h_matrix_construction_short}:] We describe how we construct the H-matrices $\mathbf{\Phi}_H$ and $\mathbf{A}_H$.
+	\item[Section \ref{sec:overall_cost}:] We review the total computational cost to construct $\mathbf{A}_H$, and list the cost to perform further linear algebra tasks involving $\mathbf{A}_H$ using H-matrix methods.
+\end{description}
+The optional method for making $\mathbf{A}_H$ positive semi-definite is described in Appendix \ref{sec:make_spd}.
+The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}.
+
+%we compute the impulse response moments $V,\mu,\Sigma$ and use these moments to estimate the supports of the impulse responses $\phi_x$ (Section \ref{sec:moments}), the greedy ellipsoid packing procedure we use to choose batches of sample points $x_i$ (Section \ref{sec:sample_point_selection}), how we compute impulse responses, $\phi_{x_i}$, in batches (Section \ref{sec:get_impulse_response}), how we interpolate the $\phi_{x_i}$ to approximate integral kernel entries $\Phi(y,x)$ (Section \ref{sec:approximate_kernel_entries}), finite element discretization of $\Phi$ (Section \ref{sec:finite_element_kernel}), and how we construct the H-matrix approximation $\mathbf{A}_h$ of the operator $\mathcal{A}$.
+We also describe how to (optionally) modify $\mathbf{A}_H$ to force it to be symmetric positive semi-definite.
+%The complete algorithm is shown in Algorithm \ref{alg:construct_Atilde}.
 
@@
 
-			Compute $\combresponse_k$ by applying $\Aop$ to the dirac comb for $\pointbatch_k$ (Section \ref{sec:get_impulse_response})
+			Compute $\combresponse_k$ by applying $\Aop$ to the Dirac comb for $\pointbatch_k$ (Section \ref{sec:get_impulse_response})
 
@@
 
-\section{Impulse response moments}
+\subsection{Impulse response moments}
 \label{sec:moments}
@@
 \subsection{Computational cost}
+\label{sec:overall_cost}
 

### 2022-03-08 paper/localpsf.tex
paper edits

@@
 But approximation \eqref{eq:translate3} can hold in situations where neither \eqref{eq:translate1} nor \eqref{eq:translate2} holds.
-For example, because the expected value is a linear operator, \eqref{eq:translate3} will hold when an $\Aop$ is locally translation-invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case.
+For example, because the expected value commutes with affine transformations, \eqref{eq:translate3} will hold when an $\Aop$ is locally translation-invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case.
 Additionally, \eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions.
@@
 
+%Recall from Section \ref{eq:local_mean_displacement_invariance} that if $x'$ is near $x$, then
+%\begin{equation*}
+%\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x')\right).
+%\end{equation*}
+
 %In Section \ref{sec:intro} we noted that computing $\Aker\left(y,x\right)$ is computationally costly. 
@@
 where $\{x_i\}_{i=1}^k$ are the $k$ nearest sample points to $x$, excluding sample points for which $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ is undefined.
-That is, excluding sample points such that $y-\mu(x)+\mu(x_i) \notin \Omega$.
+We approximate $\Phi(y,x)$ by interpolating the (point,value) pairs
+\begin{equation}
+\left(x_i, f_i\right), \qquad i=1,\dots,k,
+\end{equation}
+at the point $x$.
+The idea here is that $\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ (see Section \ref{eq:local_mean_displacement_invariance}), so we approximate $\Phi(y,x)$ by interpolating the values $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$.
+Inclusion of the factor $\frac{V(x)}{V(x_i)}$ in \eqref{eq:fxyxp} is a form of preconditioning that makes this approximation more robust in the case where $V(x)$ varies widely.
+
+To find the $k$ nearest sample points to $x$, we query a precomputed KD-tree \cite{kdtree} of all sample points.
 The number $k$ is a small user defined parameter; in numerical experiments we use $k=10$.
-We find the $k$ nearest sample points by querying a precomputed KD-tree \cite{kdtree} of all sample points.
+The function $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ is defined if and only if $y-\mu(x)+\mu(x_i) \in \Omega$.
 We check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
-
-The idea here is that $\Phi(y,x) \approx f_i$ when $x_i$ is near $x$, so we approximate $\Phi(y,x)$ by interpolating the (point,value) pairs
-\begin{equation}
-\left(x_i, f_i\right), \qquad i=1,\dots,k,
-\end{equation}
-at the point $x$.
+%Since $\Phi(y,x) \approx f_i$ when $x_i$ is near $x$, we approximate $\Phi(y,x)$ by interpolating the (point,value) pairs
+%\begin{equation}
+%\left(x_i, f_i\right), \qquad i=1,\dots,k,
+%\end{equation}
+%at the point $x$.
 To see that $\Phi(y,x) \approx f_i$, recall from Section \ref{eq:local_mean_displacement_invariance} that $\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ if $x_i$ is near $x$.
@@
 %Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section \ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $V(x_i)$ is smaller than a predetermined threshold.
-
 %\subsubsection{Radial basis function interpolation}
 
-We use the following radial basis function interpolation scheme:
+For interpolation, we use the following radial basis function scheme:
 %The approximation $\widetilde{\Phi}(y,x)$ is the value of this interpolation at $x$.
@@
 \end{equation}
-Here $b_i$ are radial basis functions and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
+where $b_i$ are radial basis functions and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
 \begin{equation}
@@
 %\end{equation*}
-We use polyharmonic spline radial basis functions \cite{polyspline}, which are defined as follows:
+We use polyharmonic spline radial basis functions \cite{polyspline}:
 \begin{equation}
@@
 for $r>0$, and $b_i(0):=0$.
-Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a certain least-squares sense.
+Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a least-squares sense.
 For more details on radial basis function interpolation and polyharmonic splines, see \cite{rbf}.

### 2022-03-08 paper/localpsf.tex
paper edits. improved interpolation section a lot

@@
 where $\{x_i\}_{i=1}^k$ are the $k$ nearest sample points to $x$, excluding sample points for which $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ is undefined.
+Here $k$ is a small user defined parameter; e.g., $k=10$.
 We approximate $\Phi(y,x)$ by interpolating the (point,value) pairs
@@
 at the point $x$.
-The idea here is that $\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ (see Section \ref{eq:local_mean_displacement_invariance}), so we approximate $\Phi(y,x)$ by interpolating the values $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$.
-Inclusion of the factor $\frac{V(x)}{V(x_i)}$ in \eqref{eq:fxyxp} is a form of preconditioning that makes this approximation more robust in the case where $V(x)$ varies widely.
+The idea is that $\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ per the discussion in Section \ref{eq:local_mean_displacement_invariance}, so we approximate $\Phi(y,x)$ by interpolating the values $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$.
+Inclusion of the factor $\frac{V(x)}{V(x_i)}$ in \eqref{eq:fxyxp} is a form of preconditioning that makes this approximation more robust if $V(x)$ varies widely.
 
 To find the $k$ nearest sample points to $x$, we query a precomputed KD-tree \cite{kdtree} of all sample points.
-The number $k$ is a small user defined parameter; in numerical experiments we use $k=10$.
-The function $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ is defined if and only if $y-\mu(x)+\mu(x_i) \in \Omega$.
+The quantity $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ is defined if and only if $y-\mu(x)+\mu(x_i) \in \Omega$.
 We check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
@@
 
-We evaluate $f_i$ via the following process. First, we determine whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ by checking if
-\begin{equation}
-\label{eq:ellipsoid_check93}
-(y - \mu(x_i))^T \Sigma^{-1}(x_i) (y - \mu(x_i))  \le \tau^2.
-\end{equation}
-If \eqref{eq:ellipsoid_check93} does not hold, then $y - \mu(x) + \mu(x_i)$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_i=0$.
-If \eqref{eq:ellipsoid_check93} does hold, then we look up the batch index $b$ such that $x_i \in S_b$, and evaluate
+To evaluate $f_i$, we check whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ using \eqref{eq:support_ellipsoid}. If $y - \mu(x) + \mu(x_i) \notin E_{x_i}$, then
+% by checking if
+%\begin{equation}
+%\label{eq:ellipsoid_check93}
+%(y - \mu(x_i))^T \Sigma^{-1}(x_i) (y - \mu(x_i))  \le \tau^2.
+%\end{equation}
+%If \eqref{eq:ellipsoid_check93} does not hold, 
+then $y - \mu(x) + \mu(x_i)$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_i=0$.
+% If \eqref{eq:ellipsoid_check93} does hold, then 
+If $y - \mu(x) + \mu(x_i) \in E_{x_i}$, we look up the batch index $b$ such that $x_i \in S_b$, and evaluate
 \begin{equation}
@@
 \end{equation}
-as per \eqref{eq:fxyxp} and \eqref{eq:varphi_eval}.
+per \eqref{eq:fxyxp} and \eqref{eq:varphi_eval}.
 %Recall from Section \ref{sec:get_impulse_response} that $\eta_b$ is the impulse response batch that we have computed and stored.

### 2022-03-08 paper/localpsf.tex
paper edits. improved interpolation section a lot

@@
 
-If $\Aop$ is symmetric, we take advantage of symmetry to construct a more accurate approximation at negligible additional cost. In this case, 
+If $\Aop$ is symmetric, we take advantage of symmetry to construct a more accurate approximation at negligible additional cost
+%. In this case, 
 %impulse responses of $\Aop$ are also impulse responses of $\Aop^T$, which may be interpolated in the variable $y$ instead of $x$ to yield a different approximation of $\Phi(y,x)$. 
-we combine information from impulse responses of both $\Aop$ and $\Aop^T$ in the interpolation procedure.
+%we combine 
+by combining information from impulse responses of both $\Aop$ and $\Aop^T$ in the interpolation procedure.
 % for approximating $\Phi(y,x)$, to get a more accurate approximation at negligible additional cost.

### 2022-03-09 paper/localpsf.tex
paper edits

@@
 Given $(y,x)\in \Omega \times \Omega$, let
+\begin{equation*}
+	z_i := y - \mu(x) + \mu(x_i), \qquad i=1,\dots,k,
+\end{equation*}
+where $\{x_i\}_{i=1}^k$ are the $k$ nearest sample points to $x$, excluding sample points $x_i$ for which $z_i \notin \Omega$. Here $k$ is a small user defined parameter; e.g., $k=10$. Further, let
 \begin{equation}
 \label{eq:fxyxp}
-f_i := \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right), \qquad i=1,\dots,k,
+f_i := \frac{V(x)}{V(x_i)}\phi_{x_i}\left(z_i\right), \qquad i=1,\dots,k,
 \end{equation}
@@
 % while Section \ref{sec:approximate_kernel_entries} we showed how we can cheaply evaluate approximate kernel entries at arbitrary pairs of points $(y,x) \in \Omega \times \Omega$.
-where $\{x_i\}_{i=1}^k$ are the $k$ nearest sample points to $x$, excluding sample points for which $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ is undefined.
-Here $k$ is a small user defined parameter; e.g., $k=10$.
+Note that $f_i$ is well-defined, because $\phi_{x_i}\left(z_i\right)$ is defined if and only if $z_i \in \Omega$, and $V(x_i)>0$ by our sample point picking procedure in Section \ref{sec:sample_point_selection}.
 We approximate $\Phi(y,x)$ by interpolating the (point,value) pairs
@@
 at the point $x$.
-The idea is that $\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ per the discussion in Section \ref{eq:local_mean_displacement_invariance}, so we approximate $\Phi(y,x)$ by interpolating the values $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$.
-Inclusion of the factor $\frac{V(x)}{V(x_i)}$ in \eqref{eq:fxyxp} is a form of preconditioning that makes this approximation more robust if $V(x)$ varies widely.
+The idea is that $\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(z_i\right)$ per the discussion in Section \ref{eq:local_mean_displacement_invariance}, so we approximate $\Phi(y,x)$ by interpolating the values $\phi_{x_i}\left(z_i\right)$.
+The factor $\frac{V(x)}{V(x_i)}$ in \eqref{eq:fxyxp} preconditions the approximation to make it more robust if $V(x)$ varies widely.
 
 To find the $k$ nearest sample points to $x$, we query a precomputed KD-tree \cite{kdtree} of all sample points.
-The quantity $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ is defined if and only if $y-\mu(x)+\mu(x_i) \in \Omega$.
-We check whether $y - \mu(x) + \mu(x_i) \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
+We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
 %Since $\Phi(y,x) \approx f_i$ when $x_i$ is near $x$, we approximate $\Phi(y,x)$ by interpolating the (point,value) pairs
@@
 
-For interpolation, we use the following radial basis function scheme:
+Interpolation is performed using the following radial basis function scheme:
 %The approximation $\widetilde{\Phi}(y,x)$ is the value of this interpolation at $x$.
@@
 for $r>0$, and $b_i(0):=0$.
-Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a least-squares sense.
+This yields the smoothest interpolant of the given points and values, in a least-squares sense.
 For more details on radial basis function interpolation and polyharmonic splines, see \cite{rbf}.
@@
 
-To evaluate $f_i$, we check whether $y - \mu(x) + \mu(x_i) \in E_{x_i}$ using \eqref{eq:support_ellipsoid}. If $y - \mu(x) + \mu(x_i) \notin E_{x_i}$, then
+To evaluate $f_i$, we check whether $z_i \in E_{x_i}$ using \eqref{eq:support_ellipsoid}. If $z_i \notin E_{x_i}$, then
 % by checking if
@@
 %If \eqref{eq:ellipsoid_check93} does not hold, 
-then $y - \mu(x) + \mu(x_i)$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_i=0$.
+then $z_i$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_i=0$.
 % If \eqref{eq:ellipsoid_check93} does hold, then 
-If $y - \mu(x) + \mu(x_i) \in E_{x_i}$, we look up the batch index $b$ such that $x_i \in S_b$, and evaluate
-\begin{equation}
-f_i = V(x)\eta_b\left(y - \mu(x) + \mu(x_i)\right),
-\end{equation}
-per \eqref{eq:fxyxp} and \eqref{eq:varphi_eval}.
+If $z_i \in E_{x_i}$, we look up the batch index $b$ such that $x_i \in S_b$, and evaluate $f_i$ via the formula
+\begin{equation}
+f_i = V(x)\eta_b\left(z_i\right),
+\end{equation}
+per \eqref{eq:varphi_eval}.
 %Recall from Section \ref{sec:get_impulse_response} that $\eta_b$ is the impulse response batch that we have computed and stored.
@@
 
-Note that $y - \mu(x) + \mu(x_i)$ is generally not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints. Hence,
-evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$ requires determining which mesh cell contains $y - \mu(x) + \mu(x_i)$, then evaluating finite element basis functions on that mesh cell.
-Fortunately, the mesh cell containing $y - \mu(x) + \mu(x_i)$ was already determined as a side effect of querying the AABB tree of mesh cells when we checked whether $y - \mu(x) + \mu(x_i) \in \Omega$.
+Note that $z_i$ is generally not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints. Hence,
+evaluating $\eta_b\left(z_i\right)$ requires determining which mesh cell contains $z_i$, then evaluating finite element basis functions on that mesh cell.
+Fortunately, the mesh cell containing $z_i$ was already determined as a side effect of querying the AABB tree of mesh cells when we checked whether $z_i \in \Omega$.
 
@@
 	\label{eq:Amatpc_defn}
-	\mathbf{A} := \bm{M} \mathbf{\Phi} \bm{M}.
+	\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
 \end{equation}
@@
 
-We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by converting $\mathbf{\Phi}$ and $\bm{M}$ to H-matrix format, then using fast H-matrix methods to multiply these matrices according to \eqref{eq:Amatpc_defn}.
+We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming H-matrix representations $\mathbf{\Phi}_H$ and $\bm{M}_H$ of $\mathbf{\Phi}$ and $\bm{M}$, respectively, then using fast H-matrix methods to multiply these matrices per \eqref{eq:Amatpc_defn} to form
+\begin{equation*}
+\mathbf{A}_H = \bm{M}_H \mathbf{\Phi}_H \bm{M}_H.
+\end{equation*}
 Once $\mathbf{A}_H$ is formed, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods \cite{BORMHACKBUSCHBOOK}.
@@
 
-We convert $\mathbf{\Phi}$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{HLIBPRO}.
-Although $\mathbf{\Phi}$ is a dense $\fedim \times \fedim$ matrix, constructing the H-matrix approximation only requires evaluation of $O(\hmatrixrank^2 \fedim \log \fedim)$ entries $\mathbf{\Phi}_{ij} = \widetilde{\Phi}(p_i, p_j)$, and these entries are computed via the procedure described in Section \ref{sec:approximate_kernel_entries}.
+We form $\mathbf{\Phi}_H$ using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{HLIBPRO}.
+Although $\mathbf{\Phi}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Phi}_H$ only requires evaluation of $O(\hmatrixrank^2 \fedim \log \fedim)$ entries $\mathbf{\Phi}_{ij} = \widetilde{\Phi}(p_i, p_j)$, and these entries are computed via the procedure described in Section \ref{sec:approximate_kernel_entries}.
 Here $\hmatrixrank$ is the rank of the highest-rank block in the H-matrix.
@@
 We describe this H-matrix construction process in detail in Appendix \ref{app:h_matrix}.
-We convert $\bm{M}$ to H-matrix format using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same recursive block partitioning structure as was used for $\mathbf{\Phi}$.
+We form $\bm{M}_H$ using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same recursive block partitioning structure as was used for $\mathbf{\Phi}$.
 
@@
 \label{sec:overall_cost}
+
+Computing $V$, $\mu$, and $\Sigma$ require $1$, $d$, and $d(d+1)/2$ operator actions of $\Aop$, respectively.
+Computing the impulse response batches $\eta_b$ requires one operator action of $\Aop$ per batch.
+Picking the sample points
 
@@
 
-Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\AkerPcMat$ corresponds to the Lagrange node in $\mathbb{R}^\gdim$ associated with the $i^\text{th}$ finite element basis function.
+Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\mathbf{\Phi}$ corresponds to the Lagrange node in $\mathbb{R}^\gdim$ associated with the $i^\text{th}$ finite element basis function.
 The columns of the matrix thus correspond to a point cloud in $\mathbb{R}^\gdim$.
@@
 
-After reordering the rows and columns of $\AkerPcMat$ via the depth-first ordering described above, we partition the reordered version of $\AkerPcMat$ into a tree of $2 \times 2$ block matrices recursively.
+After reordering the rows and columns of $\mathbf{\Phi}$ via the depth-first ordering described above, we partition the reordered version of $\mathbf{\Phi}$ into a tree of $2 \times 2$ block matrices recursively.
 We use a geometric admissibility condition (discussed below) to decide which blocks to subdivide, and use the cluster trees for the rows and columns to determine how to subdivide those blocks.
 For the first stage of subdivision, let $r_1$ and $r_2$ be the children row clusters for the root of the row cluster tree, let $\colcluster_1$ and $\colcluster_2$ be the children column clusters for the root of the column cluster tree.
-The matrix $\AkerPcMat$ is partitioned into blocks as follows:
+The matrix $\mathbf{\Phi}$ is partitioned into blocks as follows:
 \begin{equation*}
 	\begin{bmatrix}
-		\left(\AkerPcMat\right)_{11} & \left(\AkerPcMat\right)_{12} \\
-		\left(\AkerPcMat\right)_{21} & \left(\AkerPcMat\right)_{22},
-	\end{bmatrix}
-\end{equation*}
-where $\left(\AkerPcMat\right)_{11}$ denotes the block of $\AkerPcMat$ with rows $r_1$ and columns $\colcluster_1$, $\left(\AkerPcMat\right)_{12}$ denotes the block of $\AkerPcMat$ with rows $r_1$ and columns $\colcluster_2$, and so on for $\left(\AkerPcMat\right)_{21}$ and $\left(\AkerPcMat\right)_{22}$.
-
-We now loop through the four blocks, $\left(\AkerPcMat\right)_{11}$, $\left(\AkerPcMat\right)_{12}$, $\left(\AkerPcMat\right)_{21}$, and $\left(\AkerPcMat\right)_{22}$, and decide which blocks should be subdivided further.
-For the purpose of explanation, consider $\left(\AkerPcMat\right)_{12}$.
+		\mathbf{\Phi}_{11} & \mathbf{\Phi}_{12} \\
+		\mathbf{\Phi}_{21} & \mathbf{\Phi}_{22}
+	\end{bmatrix},
+\end{equation*}
+where $\mathbf{\Phi}_{11}$ denotes the block of $\mathbf{\Phi}$ with rows $r_1$ and columns $\colcluster_1$, $\mathbf{\Phi}_{12}$ denotes the block of $\mathbf{\Phi}$ with rows $r_1$ and columns $\colcluster_2$, and so on for $\mathbf{\Phi}_{21}$ and $\mathbf{\Phi}_{22}$.
+
+We now loop through the four blocks, $\mathbf{\Phi}_{11}$, $\mathbf{\Phi}_{12}$, $\mathbf{\Phi}_{21}$, and $\mathbf{\Phi}_{22}$, and decide which blocks should be subdivided further.
+For the purpose of explanation, consider $\mathbf{\Phi}_{12}$.
 If
@@
 \end{equation}
-then we mark $\left(\AkerPcMat\right)_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone.
+then we mark $\mathbf{\Phi}_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone.
 Here $\dist\left(r_1, \colcluster_2\right)$ is the Euclidean distance between the axis-aligned bounding box for the point cloud associated with the row cluster $r_1$, and the axis aligned bounding box for the point cloud associated with the column cluster $\colcluster_2$.
@@
 This process is repeated for the other blocks to determine which blocks are admissible and which are not.
-For us, the diagonal blocks $\left(\AkerPcMat\right)_{11}$ and $\left(\AkerPcMat\right)_{22}$ are not admissible because the distance between a point cloud and itself is zero.
-
-Next, we subdivide all blocks that are not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\AkerPcMat$.
+For us, the diagonal blocks $\mathbf{\Phi}_{11}$ and $\mathbf{\Phi}_{22}$ are not admissible because the distance between a point cloud and itself is zero.
+
+Next, we subdivide all blocks that are not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\mathbf{\Phi}$.
 But now we subdivide a block based on the two child row clusters and two child column clusters for the rows and columns of that block.
@@
 Once the block cluster tree has been constructed, low rank approximations of the admissible (low rank) blocks are formed using the adaptive cross method \cite{ACA}.
-Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\AkerPcMat$.
+Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\mathbf{\Phi}$.
 The idea of the adaptive cross method is to form a low rank approximation of $X$ by sampling a small number of rows and columns of $X$.
@@
 
-We use the adaptive cross method to form low rank approximations for each admissible block of $\AkerPcMat$.
-We directly compute all entries of the small dense blocks of $\AkerPcMat$ that are not admissible.
-This process requires us to compute $O(\hmatrixrank^2 \fedim \log \fedim)$ entries of $\AkerPcMat$, which is relatively cheap compared to the operator actions of $\Aop$ that are required to form the product-convolution approximation.
+We use the adaptive cross method to form low rank approximations for each admissible block of $\mathbf{\Phi}$.
+We directly compute all entries of the small dense blocks of $\mathbf{\Phi}$ that are not admissible.
+This process requires us to compute $O(\hmatrixrank^2 \fedim \log \fedim)$ entries of $\mathbf{\Phi}$, which is relatively cheap compared to the operator actions of $\Aop$ that are required to form the product-convolution approximation.
 
@@
 
-	Solve generalized eigenvalue problem $\spatialcov_p \Phi = \spatialcov_q \Phi \Lambda$
-
-	$v \gets \Phi^T\left(\spatialmean_p - \spatialmean_q\right)$
+	Solve generalized eigenvalue problem $\spatialcov_p P = \spatialcov_q P \Lambda$
+
+	$v \gets P^T\left(\spatialmean_p - \spatialmean_q\right)$
 

### 2022-03-09 paper/localpsf.tex
paper edits

@@
 
-\subsection{Overview of the method}
+\subsection{Outline of the method}
+\label{sec:intro_overview}
+
 The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative impulse response property.
@@
 \label{sec:method}
-In this section we detail our method for computing the H-matrix approximation, $\mathbf{A}_H$, of $\Aop$. This section is organized as follows:
+In this section we detail our method for computing the H-matrix approximation, $\mathbf{A}_H$, of $\Aop$, as outlined in Section \ref{sec:intro_overview}.
+%The supports of the impulse responses $\phi_x$ are estimated to be contained within ellipsoids (Section \ref{eq:mean_and_covariance_estimation}).
+The shapes of the ellipsoids depend on spatially varying volume, mean, and covariance functions, $V,\mu,\Sigma$, which we calculate by applying $\Aop$ to a small number of polynomial functions (Section \ref{sec:moments}).
+We use the ellipsoid support estimates in a greedy ellipsoid packing procedure to choose batches, $\pointbatch_b$, of sample points $x_i$ (Section \ref{sec:sample_point_selection}).
+Batches, $\eta_b$, of impulse responses, $\phi_{x_i}$, are computed by applying $\Aop$ to Dirac combs associated with the batches of sample points $S_b$.
+The impulse response batch $\eta_b$ contains the information about all impulse responses $\phi_{x_i}$ associated with points $x_i \in S_b$.
+Hence, once a small number of impulse response batches $\eta_b$ are computed, they can be used to perform pointwise evaluations $z \mapsto \phi_{x_i}(z)$ for a large number of impulse responses, $\phi_{x_i}$ (Section \ref{sec:get_impulse_response}).
+%
+%In Section \ref{eq:mean_and_covariance_estimation} we describe our ellipsoid estimate for the support of $\phi_x$.
+The ellipsoid shape depends on the functions $V,\mu,\Sigma$.
+We show how to compute these functions in Section \ref{sec:moments}.
+The ellipsoid support estimates are then used in a greedy ellipsoid packing procedure to choose batches, $\pointbatch_b$, of sample points $x_i$ (Section \ref{sec:sample_point_selection}).
+This section is organized as follows:
 \begin{description}
@@
 \label{sec:fast_ellipsoid_intersection_test}
-The procedure for choosing sample points relies on quickly determining whether two ellipsoids intersect. Let $E_p$ and $E_q$ be the ellipsoids defined as
+The procedure for choosing sample points relies on quickly determining whether two ellipsoids intersect. Let $E_1$ and $E_2$ be the ellipsoids defined as
 \begin{align*}
-	E_p :=& \{x : (x - \spatialmean_p)^T \spatialcov_p^{-1} (x - \spatialmean_p) \le \tau^2\} \\
-	E_q :=& \{x : (x - \spatialmean_q)^T \spatialcov_q^{-1} (x - \spatialmean_q) \le \tau^2\}, \\
+	E_1 :=& \{x : (x - \spatialmean_1)^T \spatialcov_1^{-1} (x - \spatialmean_1) \le \tau^2\} \\
+	E_2 :=& \{x : (x - \spatialmean_2)^T \spatialcov_2^{-1} (x - \spatialmean_2) \le \tau^2\}, \\
 \end{align*}
-where $\spatialmean_p, \spatialmean_q \in \mathbb{R}^\gdim$, and $\spatialcov_p, \spatialcov_q \in \mathbb{R}^{\gdim \times \gdim}$ are positive definite.
+where $\spatialmean_1, \spatialmean_2 \in \mathbb{R}^\gdim$, and $\spatialcov_1, \spatialcov_2 \in \mathbb{R}^{\gdim \times \gdim}$ are positive definite.
 Let $K$ be the following one dimensional convex function:
 \begin{equation*}
-	K(s) := 1 - \frac{1}{\tau^2} (\spatialmean_p - \spatialmean_q)^T \left(\frac{1}{1-s}\spatialcov_p + \frac{1}{s}\spatialcov_q\right)^{-1}(\spatialmean_p - \spatialmean_q)	
-\end{equation*}
-In \cite{ELLIPSOIDINTERSECT} it is shown that $E_p \cap E_q = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$.
-We check whether $E_p$ and $E_q$ intersect by minimizing $K(s)$ on $(0,1)$.
-If $K(s^*) <0$ at the minimizer $s^*$, then $E_p \cap E_q = \{\}$.
-Otherwise $E_p \cap E_q \neq \{\}$.
+	K(s) := 1 - \frac{1}{\tau^2} (\spatialmean_1 - \spatialmean_2)^T \left(\frac{1}{1-s}\spatialcov_1 + \frac{1}{s}\spatialcov_2\right)^{-1}(\spatialmean_1 - \spatialmean_2)	
+\end{equation*}
+In \cite{ELLIPSOIDINTERSECT} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$.
+We check whether $E_1$ and $E_2$ intersect by minimizing $K(s)$ on $(0,1)$.
+If $K(s^*) <0$ at the minimizer $s^*$, then $E_1 \cap E_2 = \{\}$.
+Otherwise $E_1 \cap E_2 \neq \{\}$.
 
@@
 \begin{equation*}
-	\spatialcov_p P = \spatialcov_q P \Lambda,
+	\spatialcov_1 P = \spatialcov_2 P \Lambda,
 \end{equation*}
 where $P \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$.
-The matrix $P$ simultaneously diagonalizes $\spatialcov_p$ and $\spatialcov_q$, in the sense that $P^T\spatialcov_p P = \Lambda$, and $P^T\spatialcov_q P = I$, where $I$ is the $\gdim \times \gdim$ identity matrix.
+The matrix $P$ simultaneously diagonalizes $\spatialcov_1$ and $\spatialcov_2$, in the sense that $P^T\spatialcov_1 P = \Lambda$, and $P^T\spatialcov_2 P = I$, where $I$ is the $\gdim \times \gdim$ identity matrix.
 Using this diagonalization, and some algebraic manipulations, we may write $K(s)$ as
@@
 \end{equation}
-where $v := \Phi^T\left(\spatialmean_p - \spatialmean_q\right)$.
-We compute the generalized eigenvalue decomposition of $\spatialcov_p$ and $\spatialcov_q$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's algorithm \cite{BRENT} (any fast 1 dimensional convex optimization routine may be used).
-The resulting algorithm for checking whether $E_p$ and $E_q$ intersect is summarized in Algorithm \ref{alg:ellipsoid_intersection_test}.
+where $v := \Phi^T\left(\spatialmean_1 - \spatialmean_2\right)$.
+We compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's algorithm \cite{BRENT} (any fast 1 dimensional convex optimization routine may be used).
+The resulting algorithm for checking whether $E_1$ and $E_2$ intersect is summarized in Algorithm \ref{alg:ellipsoid_intersection_test}.
 
@@
 	\SetKwProg{Fn}{Function}{}{}
-	\Input{Ellipsoid $E_p$ with mean $\spatialmean_p$ and covariance $\spatialcov_p/\tau^2$\\
-		Ellipsoid $E_q$ with mean $\spatialmean_q$ and covariance $\spatialcov_q/\tau^2$ 
+	\Input{Ellipsoid $E_1$ with mean $\spatialmean_1$ and covariance $\spatialcov_1/\tau^2$\\
+		Ellipsoid $E_2$ with mean $\spatialmean_2$ and covariance $\spatialcov_2/\tau^2$ 
 	}
 
-	\Output{Boolean $\text{ellipsoids\_intersect}$ which is true if $E_p \cap E_q \neq \{\}$ and false otherwise}
-
-	Solve generalized eigenvalue problem $\spatialcov_p P = \spatialcov_q P \Lambda$
-
-	$v \gets P^T\left(\spatialmean_p - \spatialmean_q\right)$
+	\Output{Boolean $\text{ellipsoids\_intersect}$ which is true if $E_1 \cap E_2 \neq \{\}$ and false otherwise}
+
+	Solve generalized eigenvalue problem $\spatialcov_1 P = \spatialcov_2 P \Lambda$
+
+	$v \gets P^T\left(\spatialmean_1 - \spatialmean_2\right)$
 

### 2022-03-11 paper/localpsf.tex
impulse response moments section in intro

@@
 
-\subsection{Local mean-displacement invariance}
+\subsection{Impulse response moments}
+\label{sec:intromoments}
+
+The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative impulse response property.
+Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
+\begin{equation*}
+\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy
+\end{equation*}
+denote the spatially varying scaling factor (the ``volume'' of $\impulseresponse_x$), let 
+\begin{equation}
+\label{eq:scaled_impulse_response}
+\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)
+\end{equation}
+denote the normalized version of $\impulseresponse_x$, and let $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d},$
+\begin{equation*}
+\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
+\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
+\end{equation*}
+denote the spatially varying mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
+
+\subsection{Volume preconditioned local mean-displacement invariance}
 \label{eq:local_mean_displacement_invariance}
 
-Let $x$ and $x'$ be points in $\Omega$ that are ``close'' to each other, and consider the following three approximations, ordered by increasing level of sophistication:
+Let $x$ and $x'$ be points in $\Omega$ that are ``close'' to each other, and consider the following four approximations, ordered by increasing level of sophistication:
 \begin{align}
@@
 \phi_x(y) &\approx \phi_{x'}(y-x+x') \label{eq:translate2}\\
-\phi_x(y) &\approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right). \label{eq:translate3}
+\phi_x(y) &\approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right) \label{eq:translate3}\\
+\frac{\phi_x(y)}{V(x)} &\approx \frac{\phi_{x'}\left(y-\mu(x)+\mu(x')\right)}{V(x')}. \label{eq:translate4}
 \end{align}
@@
 %This approximation follows from the assumption that $\Phi$ is smooth. 
-Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank interpolative factorization of $\Phi$.
+Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank interpolative factorization/Nystrom approximation of $\Phi$ \cite{CUR,NYSTROM}.
 % of the form
@@
 the required rank in the low rank approximation can be large, which makes algorithms based on assumption \eqref{eq:translate1} computationally expensive.
-This is the case, for example, for Hessians in PDE constrained inverse problems with highly informative data, as discussed in Section \ref{sec:PDE_hessian_motivation}.
+This is the case for Hessians in PDE constrained inverse problems with highly informative data, as discussed in Section \ref{sec:PDE_hessian_motivation}.
 %As the data become more informative, typically the ``mass'' of $\Phi(y,x)$ concentrates nearer and nearer to the ``diagonal'' $y=x$, the Lipschitz constant increases, and the required rank in the approximation increases.
@@
 
-The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative impulse response property.
-Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
-\begin{equation*}
-\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy
-\end{equation*}
-denote the spatially varying scaling factor (the ``volume'' of $\impulseresponse_x$), let 
-\begin{equation}
-\label{eq:scaled_impulse_response}
-\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)
-\end{equation}
-denote the normalized version of $\impulseresponse_x$, and let $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d},$
-\begin{equation*}
-\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
-\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
-\end{equation*}
-denote the spatially varying mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
 In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of polynomial functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).
-
 Since a probability distribution concentrates within a small number of standard deviations of its mean, knowing $\mu(x)$ and $\Sigma(x)$ permits us to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
@@
 \begin{equation*}
-\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \{\impulseresponse_{x_i}\}_{i=1}^r \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
+\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \{x_i\}_{i=1}^r \longrightarrow \{\impulseresponse_{x_i}\}_{i=1}^r \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
 \end{equation*}
@@
 
-In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
+%In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
 Product-convolution approximations have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}.
 For background on product-convolution approximations, we recommend reading the following papers: \cite{PRODCONVGOOD}.
-
-We build upon the class of product convolution approximations in which the functions, $\convkernel_i$, are impulse responses of $\Aop$ to point sources at a collection of points $x_i$ \cite{PRODCONVLIST}.
-A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the functions $\convkernel_i$ with piecewise linear interpolation \cite{NAGY}, or splines \cite{PRODCONVSPLINES}.
+We build upon the class of product convolution approximations in which the convolution kernels are shifted impulse responses of $\Aop$ to point sources at a collection of points $x_i$ \cite{PRODCONVLIST}.
+A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, or splines \cite{PRODCONVSPLINES}.
 With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
@@
 
-Our method of estimating the support of the functions $\convkernel_i$ was inspired by resolution analysis in seismic imaging \cite{RESOLUTION}.
-In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\convkernel_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
-If the width of $\convkernel_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
-In \cite{RESOLUTION}, the width of $\convkernel_p$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
+Our method of estimating the support of the impulse responses was inspired by resolution analysis in seismic imaging \cite{RESOLUTION}.
+In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
+If the width of $\phi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
+In \cite{RESOLUTION}, the width of $\phi_p$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
 In this paper, rather than probing $\Aop^T$ with random noise functions, we use polynomial functions.
-Our method estimates the support of $\convkernel_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
+Our method estimates the support of $\phi_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
 
@@
 
-	\Input{Finite set of candidate points $P\subset \Omega$, \\spatialmean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\domain boundary information $\partial \Omega$, \\previous sample point batches $\pointbatch_1, \dots, \pointbatch_{k-1}$}
+	\Input{Finite set of candidate points $P\subset \Omega$, \\spatially varying mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\domain boundary information $\partial \Omega$, \\previous sample point batches $\pointbatch_1, \dots, \pointbatch_{k-1}$}
 	\Output{Batch of new sample points $\pointbatch_k$}
@@
 %\end{equation*}
-The finite element matrix representation of of our approximation of $\Aop$ is therefore given by
+The finite element matrix representation of our approximation of $\Aop$ is therefore given by
 \begin{equation}
@@
 
-Let $\overline{\Omega}$ be the closure of $\Omega$, and let $C(\overline{\Omega})$ be the space of continuous functions mapping $\overline{\Omega}\rightarrow \mathbb{R}$.
-The action of $\Aop$ is extended to distributions $\genericdistribution:C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ via the formula
-\begin{equation*}
-	\left(\Aop \genericdistribution^*\right)(v) := \int_\Omega v(y) \genericdistribution\left(A(y, \cdot)\right) dy, 
-\end{equation*}
-where $A(y,\cdot)$ is the function $x \mapsto A(y,x)$. This is derived formally as follows:
+%Let $\overline{\Omega}$ be the closure of $\Omega$, and let $C(\overline{\Omega})$ be the space of continuous functions mapping $\overline{\Omega}\rightarrow \mathbb{R}$. 
+The action of $\Aop$ is extended to distributions $\rho$
+%$\genericdistribution:C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ 
+via the formula
+\begin{equation*}
+	\left(\Aop \genericdistribution^*\right)(v) := \int_\Omega v(y) \genericdistribution\left(\Phi(y, \cdot)\right) dy, 
+\end{equation*}
+where $\Phi(y,\cdot)$ is the function $x \mapsto \Phi(y,x)$. This is derived formally as follows:
 \begin{align*}
-	\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) A(y,x) \textrm{``} \genericdistribution(x) \textrm{''} dx dy \\
-	&= \int_\Omega v(y) \int_\Omega A(y,x) \textrm{``} \genericdistribution(x) \textrm{''} dx \\
-	&= \int_\Omega v(y) \genericdistribution\left(A(y,\cdot)\right) dx.
+	\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Phi(y,x) \textrm{``} \genericdistribution(x) \textrm{''} dx dy \\
+	&= \int_\Omega v(y) \int_\Omega \Phi(y,x) \textrm{``} \genericdistribution(x) \textrm{''} dx \\
+	&= \int_\Omega v(y) \genericdistribution\left(\Phi(y,\cdot)\right) dx.
 \end{align*}
@@
 	\label{eq:appendix_impulse_response}
-	\left(\Aop \delta_x^*\right)(v)	= \int_\Omega v(y) A(y,x) dy.
+	\left(\Aop \delta_x^*\right)(v)	= \int_\Omega v(y) \Phi(y,x) dy.
 \end{equation}
 Recall that the impulse response of $\Aop$ to $\delta_x$, denoted $\impulseresponse_x$, is the Riesz representation of $\Aop \delta_x^*$. 
-From \eqref{eq:appendix_impulse_response} and the definition of the Riesz representation, we have $\impulseresponse_x = A(~\cdot~,x)$.
+%From \eqref{eq:appendix_impulse_response} and the definition of the Riesz representation, we have $\impulseresponse_x = A(~\cdot~,x)$.
 That is, $\impulseresponse_x$ is the function $y \mapsto A(y,x)$.

### 2022-03-11 paper/localpsf.tex
moded impulse response moments definition into its own section in the intro. Added normalized local mean displacement invariance discussion

@@
 
-In this paper we form hierarchical matrix approximations of operators $\Aop$ with locally supported, non-negative impulse responses that are accessible matrix-free.
+In this paper we form hierarchical matrix approximations, $\mathbf{A}_H$, of operators $\Aop$ that satisfy the following three properties:
 \begin{description}
@@
 \end{description} 
+Once the hierarchical matrix approximation $\mathbf{A}_H$ is constructed, one may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
+These H-matrix methods are fast and scale well to large problems.
+WHY DO THESE OPERATIONS: KRYLOV SOLVERS, SAMPLING STRATEGIES, CAN'T DO WITH A.
 
@@
 \end{equation*}
-denote the spatially varying scaling factor (the ``volume'' of $\impulseresponse_x$), let 
-\begin{equation}
-\label{eq:scaled_impulse_response}
-\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)
-\end{equation}
-denote the normalized version of $\impulseresponse_x$, and let $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d},$
-\begin{equation*}
-\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
-\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
-\end{equation*}
-denote the spatially varying mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
-
-\subsection{Volume preconditioned local mean-displacement invariance}
+denote the spatially varying scaling factor,
+%, let 
+%\begin{equation}
+%\label{eq:scaled_impulse_response}
+%\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)
+%\end{equation}
+%denote the normalized version of $\impulseresponse_x$. 
+and let $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d},$
+\begin{equation*}
+\spatialmean(x) := \mathbb{E}(\impulseresponse_x / V(x)), \qquad
+\spatialcov(x) := \Var(\impulseresponse_x /V(x)),
+\end{equation*}
+to be the spatially varying mean and covariance of the normalized version of $\impulseresponse_x$, respectively.
+For each point $x$, imagine $\phi_x$ as a pile of sand.
+The scalar $V(x)$ is the mass of the sand pile, the vector $\mu(x)$ is the center of mass of the sand pile, and the matrix $\Sigma(x)$ contains information about how wide the sand pile is in each spatial direction.
+Throughout this paper we will use these spatially varying moment functions extensively.
+
+\subsection{Local mean-displacement invariance}
 \label{eq:local_mean_displacement_invariance}
@@
 \phi_x(y) &\approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right) \label{eq:translate3}\\
-\frac{\phi_x(y)}{V(x)} &\approx \frac{\phi_{x'}\left(y-\mu(x)+\mu(x')\right)}{V(x')}. \label{eq:translate4}
+\phi_x(y)/V(x) &\approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right) / V(x'). \label{eq:translate4}
+%\frac{\phi_x(y)}{V(x)} &\approx \frac{\phi_{x'}\left(y-\mu(x)+\mu(x')\right)}{V(x')}. \label{eq:translate4}
 \end{align}
@@
 %the relevant Lipschitz constant for continuity, and hence 
-the required rank in the low rank approximation can be large, which makes algorithms based on assumption \eqref{eq:translate1} computationally expensive.
+the required rank in the low rank approximation can be large, which makes algorithms based on \eqref{eq:translate1} computationally expensive.
 This is the case for Hessians in PDE constrained inverse problems with highly informative data, as discussed in Section \ref{sec:PDE_hessian_motivation}.
@@
 
-Our method is based on approximation \eqref{eq:translate3}, which is an improvement upon \eqref{eq:translate1} and \eqref{eq:translate2} that generalizes both.
-We call \eqref{eq:translate3} ``local mean-displacement invariance,'' and illustrate \eqref{eq:translate3} in Figure \ref{fig:mean_displacement_invariance}.
+Approximation \eqref{eq:translate3} improves upon \eqref{eq:translate1} and \eqref{eq:translate2}, and generalizes both.
 On one hand, if \eqref{eq:translate1} holds, then $\mu(x) \approx \mu(x')$, and so \eqref{eq:translate3} holds.
@@
 In contrast, \eqref{eq:translate2} does not naturally generalize to operators that map between function spaces on different domains, because the formula $y-x+x'$ requires vectors in $\Omega_2$ and $\Omega_1$ to be added together.
+We call \eqref{eq:translate3} ``local mean-displacement invariance,'' and illustrate \eqref{eq:translate3} in Figure \ref{fig:mean_displacement_invariance}.
+
+We use approximation \eqref{eq:translate4}, which is the same as \eqref{eq:translate3}, except for the extra factors of $1/V(x)$ on the left hand side and $1/V(x')$ on the right hand side.
+These factors make the approximation more robust if $V(x)$ varies widely.
+Approximation \eqref{eq:translate4} is equivalent to \eqref{eq:translate3}, but with $\phi_x$ replaced by the its normalized version, $\phi_x/V(x)$.
+We call \eqref{eq:translate4} ``normalized local mean-displacement invariance.''
 
@@
 Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation.
-The details of the interpolation will be described in Section \ref{sec:approximate_kernel_entries}, but the key principle behind the interpolation is the local mean displacement invariance assumption discussed in Section \ref{eq:local_mean_displacement_invariance}.
+The details of the interpolation will be described in Section \ref{sec:approximate_kernel_entries}, but the key principle behind the interpolation is the preconditioned local mean displacement invariance assumption discussed in Section \ref{eq:local_mean_displacement_invariance}.
 %Specifically, we make the assumption that
@@
 %if $x_i$ is close to $x$. This is shown in Figure \ref{fig:mean_displacement_invariance}. Motivation for this assumption is discussed in Section \ref{eq:local_mean_displacement_invariance}. 
-We approximate $\Phi(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for a collection of sample points $x_i$ near $x$.
+We approximate $\Phi(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{V(x)}{V(x_i)}\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for a collection of sample points $x_i$ near $x$.
 The weights are determined by preconditioned radial basis function interpolation.
@@
 \end{equation*}
-One may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
-These H-matrix methods are fast and scale well to large problems.
-WHY DO THESE OPERATIONS: KRYLOV SOLVERS, SAMPLING STRATEGIES, CAN'T DO WITH A.
 
@@
 where $\tau$ is a fixed constant.
-The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\impulseresponse}_x$.
+The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as the normalized version of $\impulseresponse_x$.
 
@@
 \end{equation}
-at the point $x$.
-The idea is that $\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(z_i\right)$ per the discussion in Section \ref{eq:local_mean_displacement_invariance}, so we approximate $\Phi(y,x)$ by interpolating the values $\phi_{x_i}\left(z_i\right)$.
-The factor $\frac{V(x)}{V(x_i)}$ in \eqref{eq:fxyxp} preconditions the approximation to make it more robust if $V(x)$ varies widely.
+at the point $x$. The idea is that $\Phi(y,x) = f_i$ per the discussion in Section \ref{eq:local_mean_displacement_invariance}, so we approximate $\Phi(y,x)$ by interpolating the values $f_i$.
 

### 2022-03-13 paper/localpsf.tex
moded impulse response moments definition into its own section in the intro. Added normalized local mean displacement invariance discussion
major edits to beginning of intro. intro beginning section is acceptable now

@@
 \begin{abstract}
-	We present a fast matrix-free method for computing hierarchical matrix (H-matrix) approximations of operators that have locally supported non-negative impulse responses.
+	We present a fast matrix-free method for computing hierarchical matrix (H-matrix) approximations of operators that have locally supported non-negative integral kernels.
 Such operators arise, for example, as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained inverse problems.
@@
 
-We present a fast matrix free method for approximating operators, $\Aop$, that have locally supported non-negative impulse responses.
-Such operators arise throughout computational science.
-Examples include Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems.
-
-Let $\Omega \subset \mathbb{R}^d$ be a bounded domain (in applications typically $\gdim=1$, $2$, or $3$). We consider integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ of the form
+We present a fast matrix free method for approximating operators, $\Aop$, that have locally supported non-negative integral kernels.
+Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems.
+
+Let $\Omega \subset \mathbb{R}^d$ be a bounded domain (in applications typically $\gdim=1$, $2$, or $3$) and let $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the following form:
 \begin{equation}
 	\label{eq:kernel_representation}
-	(\Aop u)(v) := \int_\Omega \int_\Omega v(y) \Aker(y,x) u(x) dx dy,
-\end{equation}
-where $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel.
-Here $\Aop u \in L^2(\Omega)'$ is the linear functional that results from applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
-
-Further, let $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$,
+	(\Aop u)(v) := \int_\Omega \int_\Omega v(y) \Aker(y,x) u(x) dx dy.
+\end{equation}
+Here $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, $\Aop u \in L^2(\Omega)'$ is the linear functional that results from applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
+We focus on operators $\Aop$ that are accessible \emph{matrix-free}.
+One cannot easily compute kernel entries $\Aker(y,x)$.
+Instead,
+%access to $\Aop$ is available through ``operator actions.'' That is,
+we have a black-box computational procedure through which we may evaluate the maps
+\begin{equation*}
+u \mapsto\Aop u \quad \text{and} \quad v \mapsto \Aop^Tv
+\end{equation*} 
+for arbitrary functions $u$ and $v$. 
+%After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\bm{A}$, that is typically too large to be built and stored.
+This property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\bm{A}$ and $\bm{A}^T$ with arbitrary vectors, but one cannot easily\footnote{One could compute $\bm{v} = \bm{A}\bm{e}_j$, where $\bm{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero, then extract $\bm{A}_{ij} = \bm{v}_i$.
+However, this process is wasteful because one computes the entire vector $\bm{v}$, then discards all but one component.} access matrix entries $\bm{A}_{ij}$.
+In applications, evaluating these maps may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure. 
+%For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
+
+Our method belongs to a class of methods that approximate an operator by interpolating impulse responses of the operator at a scattered collection of points.
+The \emph{impulse response} of $\Aop$ associated with the point $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$,
 \begin{equation}
@@
 \end{equation} 
-be the \emph{impulse response} of $\Aop$ at $x$. The function $\impulseresponse_x$ is called the impulse response because straightforward analysis shows that it may be expressed as follows:
+which results from keeping $x$ fixed and reinterpreting $\Phi(y,x)$ as a function of $y$.
+%for a collection of points $x_i$ scattered throughout $\Omega$.
+%Further, let $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$,
+%\begin{equation}
+%\label{eq:impulse_response_defn}
+%\impulseresponse_x(y) := \Aker(y, x),
+%\end{equation} 
+%be the \emph{impulse response} of $\Aop$ at $x$. 
+The function $\impulseresponse_x$ is called the impulse response because straightforward analysis shows that it may be expressed as follows:
 \begin{equation}
@@
 \end{equation}
-where $\delta_x$ is the the point source (delta distribution) centered at $x$, $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying the operator $\Aop$ to $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\sigma \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\sigma^* \in L^2(\Omega)$ such that $\sigma(v) = \left(\sigma^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is formally defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
+where $\delta_x$ is the the delta distribution (i.e., point source, impulse) centered at $x$, $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying the operator $\Aop$ to $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\sigma \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\sigma^* \in L^2(\Omega)$ such that $\sigma(v) = \left(\sigma^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is formally defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
 This is described in Appendix \ref{app:distributions}.
 
-In this paper we form hierarchical matrix approximations, $\mathbf{A}_H$, of operators $\Aop$ that satisfy the following three properties:
-\begin{description}
-	\item[Locally supported impulse responses:] For all $x\in \Omega$, the support of $\impulseresponse_x$ is contained (or approximately contained) in a localized region.
-	\item[Non-negative impulse responses:] We have
-	\begin{equation*}
-	\impulseresponse_x(y) \ge 0
-	\end{equation*}
-	for all (or approximately all) $(y,x) \in \Omega \times \Omega$.
-	\item[Matrix-free:] One cannot easily evaluate kernel entries $\Aker(y,x)$.
+Our method assumes that the supports of the $\impulseresponse_x$ are contained (or approximately contained) in localized regions.
+Our method also assumes that the integral kernel is non-negative, i.e., $\Phi(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$.
+Because of the non-negativity assumption, we are able to estimate the supports of the functions $\phi_x$ a-priori by applying $\Aop$ to a small number of polynomial functions.
+This, in combination with the locality assumption, allows us to compute impulse responses in batches.
+We recover the impulse responses $\phi_{x_i}$ for a large number of points $x_i$ with a small number of operator actions $v \mapsto \Aop v$.
+We also leverage the non-negativity assumption to perform a more sophisticated and accurate interpolation of the impulse responses.
+
+The output of our method is a hierarchical matrix (H-matrix) approximation, $\mathbf{A}_H$, of a discretization of $\Aop$.
+One may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$ that cannot easily be done using $\Aop$ directly, because $\Aop$ is only available matrix-free.
+These operations include matrix-matrix addition, matrix-matrix multiplication, (faster) matrix-vector products, matrix factorization, and matrix inversion.
+H-matrix methods are fast and scale well to large problems.
+The ability to perform these operations permits, for example, fast solution of Newton linear systems in optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
+
+%
+% Because we build a H-matrix, and 
+%
+%In this paper we form hierarchical matrix approximations, $\mathbf{A}_H$, of operators $\Aop$ that satisfy the following three properties:
+%\begin{description}
+%	\item[Locally supported impulse responses:] For all $x\in \Omega$, the support of $\impulseresponse_x$ is contained (or approximately contained) in a localized region.
+%	\item[Non-negative impulse responses:] We have
+%	\begin{equation*}
+%	\impulseresponse_x(y) \ge 0
+%	\end{equation*}
+%	for all (or approximately all) $(y,x) \in \Omega \times \Omega$.
+%%	\item[Matrix-free:] One cannot easily evaluate kernel entries $\Aker(y,x)$.
 Instead, access to $\Aop$ is available through ``operator actions.'' That is, we have a black-box computational procedure (which is typically expensive) through which we may evaluate the maps
-	\begin{equation*}
-	u \mapsto\Aop u \quad \text{and} \quad v \mapsto \Aop^Tv
-	\end{equation*} 
-	for arbitrary functions $u$ and $v$.
+%%	\begin{equation*}
+%%	u \mapsto\Aop u \quad \text{and} \quad v \mapsto \Aop^Tv
+%%	\end{equation*} 
+%%	for arbitrary functions $u$ and $v$.
 After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\bm{A}$, that is typically too large to be built and stored.
@@
 In applications, applying $\bm{A}$ to a vector may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
-	%For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
-\end{description} 
-Once the hierarchical matrix approximation $\mathbf{A}_H$ is constructed, one may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
+%%	%For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
+%\end{description} 
+%Once the hierarchical matrix approximation $\mathbf{A}_H$ is constructed, one may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
 These H-matrix methods are fast and scale well to large problems.
 WHY DO THESE OPERATIONS: KRYLOV SOLVERS, SAMPLING STRATEGIES, CAN'T DO WITH A.
-
-% approximate operators which possess the following properties:
 %
-%We further limit our focus to operators that have
-%\textbf{Local support}
-%We say that $\Aop$ has local support if, for all $x\in \Omega$, the support of $\impulseresponse_x$ is contained (or approximately contained) in a small region. 
-%The smaller these supports are, the better our algorithm will perform.
+%% approximate operators which possess the following properties:
+%% 
+%%We further limit our focus to operators that have
+%%\textbf{Local support}
+%%We say that $\Aop$ has local support if, for all $x\in \Omega$, the support of $\impulseresponse_x$ is contained (or approximately contained) in a small region. 
+%%The smaller these supports are, the better our algorithm will perform.
 Let $\delta_x$ denote the delta distribution (point source) centered at $x$.
 From \eqref{eq:impulse_response_delta_action}, we see that this local support property means that the response of $\Aop$ to a point source at $x$ is zero (or small) at points $y$ that are far from $x$.
-%\item[3) Local translation invariance:] We say that $\Aop$ is locally translation invariant if
-%\begin{equation}
-%	\label{eq:local_translation_invariance}
-%	\Aker(y+h, x+h) \approx \Aker(y,x)
-%\end{equation}
-%when $h$ is not too large.
+%%\item[3) Local translation invariance:] We say that $\Aop$ is locally translation invariant if
+%%\begin{equation}
+%%	\label{eq:local_translation_invariance}
+%%	\Aker(y+h, x+h) \approx \Aker(y,x)
+%%\end{equation}
+%%when $h$ is not too large.
 One can imagine the kernel entry $\Aker(y,x)$ as representing the strength of an interaction between a source at point $x$ and a target at point $y$.
@@
 This is illustrated in Figure FIG.
-%\textbf{Non-negative impulse response:} We have
-%\begin{equation*}
-%\impulseresponse_x(y) \ge 0
-%\end{equation*}
-%for all $(y,x) \in \Omega \times \Omega$.
+%%\textbf{Non-negative impulse response:} We have
+%%\begin{equation*}
+%%\impulseresponse_x(y) \ge 0
+%%\end{equation*}
+%%for all $(y,x) \in \Omega \times \Omega$.
 
@@
 
-\subsection{Impulse response moments}
+\subsection{Impulse responses and their moments}
 \label{sec:intromoments}
@@
 We approximate $\Phi(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{V(x)}{V(x_i)}\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for a collection of sample points $x_i$ near $x$.
-The weights are determined by preconditioned radial basis function interpolation.
+The weights are determined by radial basis function interpolation.
 

### 2022-03-13 paper/localpsf.tex
major edits to beginning of intro. intro beginning section is acceptable now

@@
 This, in combination with the locality assumption, allows us to compute impulse responses in batches.
-We recover the impulse responses $\phi_{x_i}$ for a large number of points $x_i$ with a small number of operator actions $v \mapsto \Aop v$.
+We recover the impulse responses $\phi_{x_i}$ for a large number of points $x_i$ with a small number of operator actions $u \mapsto \Aop u$.
 We also leverage the non-negativity assumption to perform a more sophisticated and accurate interpolation of the impulse responses.
@@
 The output of our method is a hierarchical matrix (H-matrix) approximation, $\mathbf{A}_H$, of a discretization of $\Aop$.
-One may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$ that cannot easily be done using $\Aop$ directly, because $\Aop$ is only available matrix-free.
+One may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$ that cannot easily be done using $\Aop$, because $\Aop$ is only available matrix-free.
 These operations include matrix-matrix addition, matrix-matrix multiplication, (faster) matrix-vector products, matrix factorization, and matrix inversion.

### 2022-03-14 paper/localpsf.tex
intro edits

@@
 
-Our method belongs to a class of methods that approximate an operator by interpolating impulse responses of the operator at a scattered collection of points.
+Our method belongs to a class of methods that approximate an operator by interpolating impulse responses of the operator associated with a scattered collection of points.
 The \emph{impulse response} of $\Aop$ associated with the point $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$,
@@
 
-Our method assumes that the supports of the $\impulseresponse_x$ are contained (or approximately contained) in localized regions.
-Our method also assumes that the integral kernel is non-negative, i.e., $\Phi(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$.
-Because of the non-negativity assumption, we are able to estimate the supports of the functions $\phi_x$ a-priori by applying $\Aop$ to a small number of polynomial functions.
-This, in combination with the locality assumption, allows us to compute impulse responses in batches.
-We recover the impulse responses $\phi_{x_i}$ for a large number of points $x_i$ with a small number of operator actions $u \mapsto \Aop u$.
-We also leverage the non-negativity assumption to perform a more sophisticated and accurate interpolation of the impulse responses.
+We assume:
+\begin{itemize}
+	\item The supports of the $\impulseresponse_x$ are contained (or approximately contained) in localized regions.
+	\item The integral kernel is non-negative, i.e., $\Phi(y,x) \ge 0$ for all (or nearly all) $(y,x) \in \Omega \times \Omega$. 
+\end{itemize}
+Because of these assumptions, we are able to estimate the supports of the functions $\phi_x$ a-priori by applying $\Aop$ to a small number of polynomial functions.
+We use the support estimates to compute $\phi_{x_i}$'s in batches, which allows us to recover $\phi_{x_i}$'s for a large number of points $x_i$ using a small number of operator actions $u \mapsto \Aop u$.
+The support estimates also allow us to interpolate the impulse responses with a new sophisticated and accurate interpolation scheme.
 
 The output of our method is a hierarchical matrix (H-matrix) approximation, $\mathbf{A}_H$, of a discretization of $\Aop$.
-One may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$ that cannot easily be done using $\Aop$, because $\Aop$ is only available matrix-free.
-These operations include matrix-matrix addition, matrix-matrix multiplication, (faster) matrix-vector products, matrix factorization, and matrix inversion.
-H-matrix methods are fast and scale well to large problems.
+While operator actions of $\Aop$ on a vectors is typically computationally expensive, fast H-matrix methods permit us to perform matrix-vector products with $\mathbf{A}_H$ cheaply.
+Furthermore, one may use H-matrix methods to perform useful linear algebra operations involving $\mathbf{A}_H$ that cannot easily be done using $\Aop$, because $\Aop$ is only available matrix-free.
+These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion.
+These H-matrix methods are fast and scale well to large problems.
 The ability to perform these operations permits, for example, fast solution of Newton linear systems in optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
@@
 Although the method we present in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest.
-For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high-rank approximations of an operator using a small number of matrix-vector products.
+For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high rank approximations of an operator using a small number of matrix-vector products.
 

### 2022-03-14 paper/localpsf.tex
intro edits

@@
 
-We present a fast matrix free method for approximating operators, $\Aop$, that have locally supported non-negative integral kernels.
+We present a fast matrix free method for approximating operators that have locally supported non-negative integral kernels.
 Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems.
@@
 \end{equation}
-Here $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, $\Aop u \in L^2(\Omega)'$ is the linear functional that results from applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
-We focus on operators $\Aop$ that are accessible \emph{matrix-free}.
-One cannot easily compute kernel entries $\Aker(y,x)$.
-Instead,
+The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, $\Aop u \in L^2(\Omega)'$ is the linear functional that results from applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
+We seek to approximate operators $\Aop$ that are accessible \emph{matrix-free}.
+That is,
 %access to $\Aop$ is available through ``operator actions.'' That is,
@@
 \begin{equation*}
-u \mapsto\Aop u \quad \text{and} \quad v \mapsto \Aop^Tv
-\end{equation*} 
-for arbitrary functions $u$ and $v$. 
+u \mapsto\Aop u \qquad \text{and} \qquad v \mapsto \Aop^Tv
+\end{equation*} 
+for arbitrary functions $u$ and $v$, but we cannot easily compute kernel entries $\Aker(y,x)$.
+In applications, evaluating these maps may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
+After discretization, $\Aop$ becomes a dense matrix that is typically too large to be built and stored.
+This property is called ``matrix-free'' because at the discrete level it means one can compute matrix-vector products with the matrix representation of $\Aop$, but cannot easily compute individual matrix entries.
 %After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\bm{A}$, that is typically too large to be built and stored.
@@
 However, this process is wasteful because one computes the entire vector $\bm{v}$, then discards all but one component.} access matrix entries $\bm{A}_{ij}$.
-In applications, evaluating these maps may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure. 
+
 %For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
@@
 \end{equation} 
-which results from keeping $x$ fixed and reinterpreting $\Phi(y,x)$ as a function of $y$.
+which results from keeping $x$ fixed and viewing $\Phi(y,x)$ as a function of $y$.
 %for a collection of points $x_i$ scattered throughout $\Omega$.
@@
 \end{equation}
-where $\delta_x$ is the the delta distribution (i.e., point source, impulse) centered at $x$, $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying the operator $\Aop$ to $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\sigma \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\sigma^* \in L^2(\Omega)$ such that $\sigma(v) = \left(\sigma^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is formally defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
+where $\delta_x$ is the the delta distribution (i.e., point source, impulse) centered at $x$, $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\sigma \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\sigma^* \in L^2(\Omega)$ such that $\sigma(v) = \left(\sigma^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
 This is described in Appendix \ref{app:distributions}.
@@
 \begin{itemize}
-	\item The supports of the $\impulseresponse_x$ are contained (or approximately contained) in localized regions.
-	\item The integral kernel is non-negative, i.e., $\Phi(y,x) \ge 0$ for all (or nearly all) $(y,x) \in \Omega \times \Omega$. 
+	\item The supports of the impulse responses $\impulseresponse_x$ are contained (or approximately contained) in localized regions.
+	\item The integral kernel is non-negative in the sense that $$\Phi(y,x) \ge 0$$ for all (or nearly all) $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
+The operator $\Aop$ need not be positive semi-definite to use our method, and positive semi-definite operators need not have a non-negative integral kernel.}
 \end{itemize}
-Because of these assumptions, we are able to estimate the supports of the functions $\phi_x$ a-priori by applying $\Aop$ to a small number of polynomial functions.
-We use the support estimates to compute $\phi_{x_i}$'s in batches, which allows us to recover $\phi_{x_i}$'s for a large number of points $x_i$ using a small number of operator actions $u \mapsto \Aop u$.
-The support estimates also allow us to interpolate the impulse responses with a new sophisticated and accurate interpolation scheme.
+Because of these assumptions, we are able to estimate the supports of the impulse responses a-priori by applying $\Aop$ to a small number of polynomial functions.
+We use the support estimates to compute scattered impulse responses in batches.
+This allows us to recover $\phi_{x_i}$'s for a large number of points $x_i$ using a small number of operator actions $u \mapsto \Aop u$.
+The support estimates also allow us to interpolate the impulse responses with a new more sophisticated and accurate interpolation scheme.
 
 The output of our method is a hierarchical matrix (H-matrix) approximation, $\mathbf{A}_H$, of a discretization of $\Aop$.
-While operator actions of $\Aop$ on a vectors is typically computationally expensive, fast H-matrix methods permit us to perform matrix-vector products with $\mathbf{A}_H$ cheaply.
+While operator actions of $\Aop$ on vectors are typically computationally expensive, fast H-matrix methods permit us to perform matrix-vector products with $\mathbf{A}_H$ cheaply.
 Furthermore, one may use H-matrix methods to perform useful linear algebra operations involving $\mathbf{A}_H$ that cannot easily be done using $\Aop$, because $\Aop$ is only available matrix-free.
@@
 These H-matrix methods are fast and scale well to large problems.
-The ability to perform these operations permits, for example, fast solution of Newton linear systems in optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
+The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
 
@@
 
-\subsection{Impulse responses and their moments}
+\subsection{Impulse response moments}
 \label{sec:intromoments}
 
-The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative impulse response property.
-Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
+The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative integral kernel property. Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
 \begin{equation*}
@@
 to be the spatially varying mean and covariance of the normalized version of $\impulseresponse_x$, respectively.
-For each point $x$, imagine $\phi_x$ as a pile of sand.
-The scalar $V(x)$ is the mass of the sand pile, the vector $\mu(x)$ is the center of mass of the sand pile, and the matrix $\Sigma(x)$ contains information about how wide the sand pile is in each spatial direction.
-Throughout this paper we will use these spatially varying moment functions extensively.
+Imagine $\phi_x$ as a mound of dirt piled on a localized region in $\Omega$, with $\phi_x(y)$ being the height of the dirt mound at location $y$.
+The scalar $V(x)$ is the total volume of dirt in the mound, the vector $\mu(x)$ is the center of mass of the dirt mound, and the matrix $\Sigma(x)$ contains information about how wide the dirt mound is in each spatial direction.
+In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of polynomial functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).
 
@@
 
-In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of polynomial functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).
 Since a probability distribution concentrates within a small number of standard deviations of its mean, knowing $\mu(x)$ and $\Sigma(x)$ permits us to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.

### 2022-03-15 paper/localpsf.tex
add impulse response batches figure

@@
 We use the support estimates to compute scattered impulse responses in batches.
-This allows us to recover $\phi_{x_i}$'s for a large number of points $x_i$ using a small number of operator actions $u \mapsto \Aop u$.
+This batching allows us to recover many impulse responses using a small number of operator actions $u \mapsto \Aop u$.
 The support estimates also allow us to interpolate the impulse responses with a new more sophisticated and accurate interpolation scheme.
@@
 
+\begin{figure}
+	\begin{subfigure}{0.32\textwidth}
+		\centering
+		\includegraphics[scale=0.33]{batch1.png}
+%		\caption{Non-symmetric}
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\centering
+		\includegraphics[scale=0.33]{batch12.png}
+		%		\caption{Non-symmetric}
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\centering
+		\includegraphics[scale=0.33]{batch19.png}
+%		\caption{Non-symmetric}
+	\end{subfigure}
+	\caption{Batches, $\eta_b$, of normalized impulse responses, $\phi_{x_i}/V(x_i)$, for the Stokes Hessian.
+Left: 2nd batch.
+Mid: 13th batch.
+Right: 19th batch.
+Brightness scale differs between the subfigures.
+Black stars indicate the sample point locations, $x_i$.
+Dashed gray ellipses indicate the support ellipsoids, $E_{x_i}$.
+The large circle is the boundary of the domain $\Omega$.
+Early batches tend to have lots of ellipsoids near the boundary, as compared to later batches.
+This is because our greedy ellipsoid packing algorithm tries to pick sample points that are far away from sample points in previously chosen batches.}
+	\label{fig:batches_intro}
+\end{figure}
+
 Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation.

### 2022-03-15 paper/localpsf.tex
restructuring paper

@@
 
-Our method belongs to a class of methods that approximate an operator by interpolating impulse responses of the operator associated with a scattered collection of points.
+Our method, described in Section \ref{sec:method}, belongs to a class of methods that approximate an operator by interpolating impulse responses of the operator associated with a scattered collection of points.
 The \emph{impulse response} of $\Aop$ associated with the point $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$,
@@
 Because of these assumptions, we are able to estimate the supports of the impulse responses a-priori by applying $\Aop$ to a small number of polynomial functions.
-We use the support estimates to compute scattered impulse responses in batches.
+We use the support estimates to compute scattered impulse responses in batches (see Figure \ref{fig:batches_intro}).
 This batching allows us to recover many impulse responses using a small number of operator actions $u \mapsto \Aop u$.
 The support estimates also allow us to interpolate the impulse responses with a new more sophisticated and accurate interpolation scheme.
+
+\begin{figure}
+	\begin{subfigure}{0.32\textwidth}
+		\centering
+		\includegraphics[scale=0.33]{batch1.png}
+		%		\caption{Non-symmetric}
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\centering
+		\includegraphics[scale=0.33]{batch12.png}
+		%		\caption{Non-symmetric}
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\centering
+		\includegraphics[scale=0.33]{batch19.png}
+		%		\caption{Non-symmetric}
+	\end{subfigure}
+	\caption{Batches, $\eta_b$, of normalized impulse responses, $\phi_{x_i}/V(x_i)$, for the Stokes inverse problem Hessian.
+Left: 2nd batch.
+Mid: 13th batch.
+Right: 19th batch.
+Brightness scale differs between the subfigures.
+Black stars indicate the sample point locations, $x_i$.
+Dashed gray ellipses indicate the support ellipsoids, $E_{x_i}$.
+The large circle is the boundary of the domain $\Omega$.
+Early batches tend to have lots of ellipsoids near the boundary, as compared to later batches.
+This is because our greedy ellipsoid packing algorithm tries to pick sample points that are far away from sample points in previously chosen batches.}
+	\label{fig:batches_intro}
+\end{figure}
 
@@
 
+\subsection{Existing work}
+
 The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term \cite{CCGOPAPERS}.
@@
 For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high rank approximations of an operator using a small number of matrix-vector products.
+
+%In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
+Product-convolution approximations have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}.
+For background on product-convolution approximations, we recommend reading the following papers: \cite{PRODCONVGOOD}.
+We build upon the class of product convolution approximations in which the convolution kernels are shifted impulse responses of $\Aop$ to point sources at a collection of points $x_i$ \cite{PRODCONVLIST}.
+A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, or splines \cite{PRODCONVSPLINES}.
+With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
+In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
+But even with adaptive refinement, many matrix vector products with $\Aop$ may be required.
+In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\convkernel_i$ are computed with each matrix vector product.
+
+Our method of estimating the support of the impulse responses was inspired by resolution analysis in seismic imaging \cite{RESOLUTION}.
+In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
+If the width of $\phi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
+In \cite{RESOLUTION}, the width of $\phi_p$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
+In this paper, rather than probing $\Aop^T$ with random noise functions, we use polynomial functions.
+Our method estimates the support of $\phi_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
+
+Matrix-vector products with product-convolution approximations can be performed using the fast Fourier transform if the problem is discretized on a regular grid \cite{PRODCONVFFT}.
+The product-convolution approximation can then replace the original operator when using Krylov methods to solve linear systems with the original operator as the coefficient operator \cite{PRODCONVKRYLOV}.
+However, for complex geometries, problems are rarely discretized using regular grids, so we cannot easily use the fast Fourier transform.
+Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}.
+Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are amenable to fast linear algebra operations.
+Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}.
+Here we follow our previous work \cite{PRODCONVMYPAPER}, in which we converted product-convolution approximations into H-matrix format.
+
+Conventional H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
+There are matrix-free methods for H-matrix approximation \cite{LEXINGPEELINGPROCESS}, and these methods have been used to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
+While these matrix-free H-matrix construction methods are asymptotically scalable in theory, the required number of matrix vector products can be large in practice.
+
+Product-convolution type approximations of Hessians have been used in a seismic inverse problem in \cite{GEORGSEISMICPRODCONV}, and in an advection-diffusion inverse problem in \cite{PRODCONVMYPAPER}.
+In \cite{DEMANETSEISMIC}, matrix probing \cite{MATRIXPROBING} is used to approximate the Hessian in a seismic inverse problem as the sum of simple pseudodifferential operators.
+Although it is not explicitly mentioned, the approximation in \cite{DEMANETSEISMIC} could be interpreted as a convolution-product approximation (a product-convolution approximation of the transpose of the operator).
+%, which is like a product-convolution approximation, except the order of pointwise multiplications and convolutions is reversed in \eqref{eq:product_convolution}.
+
+\section{Preliminaries}
 
@@
 The scalar $V(x)$ is the total volume of dirt in the mound, the vector $\mu(x)$ is the center of mass of the dirt mound, and the matrix $\Sigma(x)$ contains information about how wide the dirt mound is in each spatial direction.
-In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of polynomial functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).
+
+%The ellipsoid $E_x$ is not immediately available. 
+%(recall $\gdim$ is the spatial dimension of $\Omega$).
+The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}.
+Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
+But this direct approach is computationally infeasible because it requires one operator action of $\Aop$ per point $x$, and we will need to know $V(x)$, $\mu(x)$, and $\Sigma(x)$ at a large number of points $x$.
+Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions.
+We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Proposition \ref{thm:vol_mean_cov} and summarize the method for computing these moments in Algorithm \ref{alg:varhpi_mean_cov}.
+
+\begin{prop}[Impulse response moments]
+	\label{thm:vol_mean_cov}
+	%	Let $\spatialvol:\Omega \rightarrow \mathbb{R}$, $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$, and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$, be the following functions:
+	%	\begin{equation*}
+	%	\spatialvol(x) := \int_{\Omega} \impulseresponse_x(y) dy, \qquad
+	%	\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
+	%	\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
+	%	\end{equation*}
+	%	where $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denotes the normalized version of $\impulseresponse_x$, $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.
+	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$.
+Let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
+	\begin{equation*}
+	C(x) := 1, \qquad
+	L^i(x) := x^i, \qquad
+	Q^{ij}(x) = x^i x^j.
+	\end{equation*}
+	Let $f/g$ denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
+The following results hold:
+	\begin{subequations}
+		\label{eq:vol_mean_var}
+		\begin{align}
+		\spatialvol =& \left(\Aop^T C\right)^* \\
+		\spatialmean^i =& \left(\Aop^T L^i\right)^* / \spatialvol \\
+		\spatialcov^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j
+		\end{align}
+	\end{subequations}
+	for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$.
+\end{prop}
+
+The proof of Proposition \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}. The proof is straightforward, and follows directly from the definitions.
 
@@
 %This approximation follows from the assumption that $\Phi$ is smooth. 
-Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank interpolative factorization/Nystrom approximation of $\Phi$ \cite{CUR,NYSTROM}.
+Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank interpolative factorization/Nystrom approximation \cite{CUR,NYSTROM}.
 % of the form
@@
 
-\subsection{Outline of the method}
-\label{sec:intro_overview}
+\subsection{Finite element discretization of the integral kernel}
+\label{sec:finite_element_kernel}
+
+Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, and let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space.
+Further, let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
+That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
+
+We define a further approximation, $\widetilde{\Phi}^h \approx \widetilde{\Phi}$, as follows:
+\begin{equation}
+\label{eq:defn_of_Akerpcmesh}
+\widetilde{\Phi}^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \widetilde{\Phi}(p_i,p_j) \febasis_i(y) \febasis_j(x).
+\end{equation}
+The kernel $\widetilde{\Phi}^h$ is the interpolation of $\widetilde{\Phi}$ onto $V_h \otimes V_h$.
+Error in the overall kernel approximation, $\widetilde{\Phi}^h \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$.
+
+Let $u_h \in V_h$ and $v_h \in V_h$, and let $\bm{u}$ and $\bm{v}$ be the coefficient vectors for $u_h$ and $v_h$, respectively, with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$.
+I.e.,
+\begin{equation}
+\label{eq:fem_coeff_basis}
+u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x),
+\end{equation}
+and similar for $v_h$.
+Replacing $\Aker$ with $\widetilde{\Phi}^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward but tedious algebraic manipulations, yields the following approximation:
+\begin{equation}
+\label{eq:kernel_representation_pcmesh}
+(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
+= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u}.
+\end{equation}
+%While $\Aop$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\Aop$ to functions in the finite element space. 
+where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the following matrix of approximate kernel entries:
+\begin{equation}
+\label{eq:Akerpcmat_entries}
+\mathbf{\Phi}_{ij} := \widetilde{\Phi}(p_i, p_j),
+\end{equation}
+and $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ is the finite element mass matrix which has entries
+\begin{equation*}
+\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
+\end{equation*}
+%Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, and \eqref{eq:Akerpcmat_entries}, the definition of the mass matrix, and performing algebraic manipulations, it is straightforward but tedious to show that
+%\begin{equation*}
+%	\left(\AopPcMesh u_h \right)(v_h) = \bm{v}^T \bm{M} \AkerPcMat \bm{M} \bm{u}.
+%\end{equation*}
+The finite element matrix representation of our approximation of $\Aop$ is therefore given by
+\begin{equation}
+\label{eq:Amatpc_defn}
+\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
+\end{equation}
+
+\section{Method}
+\label{sec:method}
 
@@
 The process is repeated to get more batches, until a desired number of impulse responses is reached.
-
-\begin{figure}
-	\begin{subfigure}{0.32\textwidth}
-		\centering
-		\includegraphics[scale=0.33]{batch1.png}
-%		\caption{Non-symmetric}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\centering
-		\includegraphics[scale=0.33]{batch12.png}
-		%		\caption{Non-symmetric}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\centering
-		\includegraphics[scale=0.33]{batch19.png}
-%		\caption{Non-symmetric}
-	\end{subfigure}
-	\caption{Batches, $\eta_b$, of normalized impulse responses, $\phi_{x_i}/V(x_i)$, for the Stokes Hessian.
-Left: 2nd batch.
-Mid: 13th batch.
-Right: 19th batch.
-Brightness scale differs between the subfigures.
-Black stars indicate the sample point locations, $x_i$.
-Dashed gray ellipses indicate the support ellipsoids, $E_{x_i}$.
-The large circle is the boundary of the domain $\Omega$.
-Early batches tend to have lots of ellipsoids near the boundary, as compared to later batches.
-This is because our greedy ellipsoid packing algorithm tries to pick sample points that are far away from sample points in previously chosen batches.}
-	\label{fig:batches_intro}
-\end{figure}
 
@@
 
-\subsection{Existing work}
-
-%In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
-Product-convolution approximations have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}.
-For background on product-convolution approximations, we recommend reading the following papers: \cite{PRODCONVGOOD}.
-We build upon the class of product convolution approximations in which the convolution kernels are shifted impulse responses of $\Aop$ to point sources at a collection of points $x_i$ \cite{PRODCONVLIST}.
-A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, or splines \cite{PRODCONVSPLINES}.
-With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
-In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
-But even with adaptive refinement, many matrix vector products with $\Aop$ may be required.
-In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\convkernel_i$ are computed with each matrix vector product.
-
-Our method of estimating the support of the impulse responses was inspired by resolution analysis in seismic imaging \cite{RESOLUTION}.
-In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
-If the width of $\phi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
-In \cite{RESOLUTION}, the width of $\phi_p$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
-In this paper, rather than probing $\Aop^T$ with random noise functions, we use polynomial functions.
-Our method estimates the support of $\phi_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
-
-Matrix-vector products with product-convolution approximations can be performed using the fast Fourier transform if the problem is discretized on a regular grid \cite{PRODCONVFFT}.
-The product-convolution approximation can then replace the original operator when using Krylov methods to solve linear systems with the original operator as the coefficient operator \cite{PRODCONVKRYLOV}.
-However, for complex geometries, problems are rarely discretized using regular grids, so we cannot easily use the fast Fourier transform.
-Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}.
-Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are amenable to fast linear algebra operations.
-Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}.
-Here we follow our previous work \cite{PRODCONVMYPAPER}, in which we converted product-convolution approximations into H-matrix format.
-
-Conventional H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
-There are matrix-free methods for H-matrix approximation \cite{LEXINGPEELINGPROCESS}, and these methods have been used to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
-While these matrix-free H-matrix construction methods are asymptotically scalable in theory, the required number of matrix vector products can be large in practice.
-
-Product-convolution type approximations of Hessians have been used in a seismic inverse problem in \cite{GEORGSEISMICPRODCONV}, and in an advection-diffusion inverse problem in \cite{PRODCONVMYPAPER}.
-In \cite{DEMANETSEISMIC}, matrix probing \cite{MATRIXPROBING} is used to approximate the Hessian in a seismic inverse problem as the sum of simple pseudodifferential operators.
-Although it is not explicitly mentioned, the approximation in \cite{DEMANETSEISMIC} could be interpreted as a convolution-product approximation (a product-convolution approximation of the transpose of the operator).
-%, which is like a product-convolution approximation, except the order of pointwise multiplications and convolutions is reversed in \eqref{eq:product_convolution}.
-
-\section{Method}
-\label{sec:method}
 In this section we detail our method for computing the H-matrix approximation, $\mathbf{A}_H$, of $\Aop$, as outlined in Section \ref{sec:intro_overview}.
@@
 
-\subsection{Impulse response moments}
-\label{sec:moments}
-
-%The ellipsoid $E_x$ is not immediately available. 
 To determine $E_x$, we must compute the scalar $\spatialvol(x)$, the length-$\gdim$ vector $\spatialmean(x)$, and the $\gdim \times \gdim$ matrix $\spatialcov(x)$.
-%(recall $\gdim$ is the spatial dimension of $\Omega$).
-The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}.
-Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
-But this direct approach is computationally infeasible because it requires one operator action of $\Aop$ per point $x$ that we wish to determine $E_x$ at, and we will need to determine $E_x$ at a large number of points.
-Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions.
-We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Proposition \ref{thm:vol_mean_cov} and summarize the method for computing these moments in Algorithm \ref{alg:varhpi_mean_cov}.
-
-\begin{prop}[Impulse response moments]
-	\label{thm:vol_mean_cov}
-%	Let $\spatialvol:\Omega \rightarrow \mathbb{R}$, $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$, and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$, be the following functions:
-%	\begin{equation*}
-%	\spatialvol(x) := \int_{\Omega} \impulseresponse_x(y) dy, \qquad
-%	\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
-%	\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
-%	\end{equation*}
-%	where $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denotes the normalized version of $\impulseresponse_x$, $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.
-	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$.
-Let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
-	\begin{equation*}
-	C(x) := 1, \qquad
-	L^i(x) := x^i, \qquad
-	Q^{ij}(x) = x^i x^j.
-	\end{equation*}
-	Let $f/g$ denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
-The following results hold:
-	\begin{subequations}
-		\label{eq:vol_mean_var}
-		\begin{align}
-		\spatialvol =& \left(\Aop^T C\right)^* \\
-		\spatialmean^i =& \left(\Aop^T L^i\right)^* / \spatialvol \\
-		\spatialcov^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j
-		\end{align}
-	\end{subequations}
-	for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$.
-\end{prop}
-
-The proof of Proposition \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}. The proof is straightforward, and follows directly from the definitions.
+In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of polynomial functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).
 
@@
 
-\subsection{Finite element discretization of the integral kernel}
-\label{sec:finite_element_kernel}
-
-Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, and let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space.
-Further, let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
-That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
-
-We define a further approximation, $\widetilde{\Phi}^h \approx \widetilde{\Phi}$, as follows:
-\begin{equation}
-	\label{eq:defn_of_Akerpcmesh}
-	\widetilde{\Phi}^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \widetilde{\Phi}(p_i,p_j) \febasis_i(y) \febasis_j(x).
-\end{equation}
-The kernel $\widetilde{\Phi}^h$ is the interpolation of $\widetilde{\Phi}$ onto $V_h \otimes V_h$.
-Error in the overall kernel approximation, $\widetilde{\Phi}^h \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$.
-
-Let $u_h \in V_h$ and $v_h \in V_h$, and let $\bm{u}$ and $\bm{v}$ be the coefficient vectors for $u_h$ and $v_h$, respectively, with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$.
-I.e.,
-\begin{equation}
-\label{eq:fem_coeff_basis}
-u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x),
-\end{equation}
-and similar for $v_h$.
-Replacing $\Aker$ with $\widetilde{\Phi}^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward but tedious algebraic manipulations, yields the following approximation:
-\begin{equation}
-	\label{eq:kernel_representation_pcmesh}
-	(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
-	= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u}.
-\end{equation}
-%While $\Aop$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\Aop$ to functions in the finite element space. 
-where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the following matrix of approximate kernel entries:
-\begin{equation}
-	\label{eq:Akerpcmat_entries}
-	\mathbf{\Phi}_{ij} := \widetilde{\Phi}(p_i, p_j),
-\end{equation}
-and $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ is the finite element mass matrix which has entries
-\begin{equation*}
-\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
-\end{equation*}
-%Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, and \eqref{eq:Akerpcmat_entries}, the definition of the mass matrix, and performing algebraic manipulations, it is straightforward but tedious to show that
-%\begin{equation*}
-%	\left(\AopPcMesh u_h \right)(v_h) = \bm{v}^T \bm{M} \AkerPcMat \bm{M} \bm{u}.
-%\end{equation*}
-The finite element matrix representation of our approximation of $\Aop$ is therefore given by
-\begin{equation}
-	\label{eq:Amatpc_defn}
-	\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
-\end{equation}
-
 \subsection{Hierarchical matrix construction}

### 2022-03-15 paper/localpsf.tex
restructuring paper

@@
 Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
-But this direct approach is computationally infeasible because it requires one operator action of $\Aop$ per point $x$, and we will need to know $V(x)$, $\mu(x)$, and $\Sigma(x)$ at a large number of points $x$.
+But this direct approach is computationally infeasible because it requires one operator action of $\Aop$ per point $x$, and we will need to know $V(x)$, $\mu(x)$, and $\Sigma(x)$ at all points $x$.
 Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions.
-We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Proposition \ref{thm:vol_mean_cov} and summarize the method for computing these moments in Algorithm \ref{alg:varhpi_mean_cov}.
-
-\begin{prop}[Impulse response moments]
-	\label{thm:vol_mean_cov}
-	%	Let $\spatialvol:\Omega \rightarrow \mathbb{R}$, $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$, and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$, be the following functions:
-	%	\begin{equation*}
-	%	\spatialvol(x) := \int_{\Omega} \impulseresponse_x(y) dy, \qquad
-	%	\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
-	%	\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
-	%	\end{equation*}
-	%	where $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denotes the normalized version of $\impulseresponse_x$, $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.
-	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$.
-Let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
+Specifically, let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$, and let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
 \begin{equation*}
@@
 \end{equation*}
-	Let $f/g$ denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
-The following results hold:
+The following results (derived in Appendix \ref{app:proofs}) follow straightforwardly from the definitions:
 \begin{subequations}
@@
 for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$.
-\end{prop}
-
-The proof of Proposition \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}. The proof is straightforward, and follows directly from the definitions.
+Here $f/g$ denotes pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denotes pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
+
+%We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Proposition \ref{thm:vol_mean_cov} and summarize the method for computing these moments in Algorithm \ref{alg:varhpi_mean_cov}.
+
+%\begin{prop}[Impulse response moments]
+%	\label{thm:vol_mean_cov}
+%	%	Let $\spatialvol:\Omega \rightarrow \mathbb{R}$, $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$, and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$, be the following functions:
+%	%	\begin{equation*}
+%	%	\spatialvol(x) := \int_{\Omega} \impulseresponse_x(y) dy, \qquad
+%	%	\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
+%	%	\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
+%	%	\end{equation*}
+%	%	where $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denotes the normalized version of $\impulseresponse_x$, $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.
+%	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$.
+Let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
+%	\begin{equation*}
+%	C(x) := 1, \qquad
+%	L^i(x) := x^i, \qquad
+%	Q^{ij}(x) = x^i x^j.
+%	\end{equation*}
+%	Let $f/g$ denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
+The following results hold:
+%	\begin{subequations}
+%		\label{eq:vol_mean_var}
+%		\begin{align}
+%		\spatialvol =& \left(\Aop^T C\right)^* \\
+%		\spatialmean^i =& \left(\Aop^T L^i\right)^* / \spatialvol \\
+%		\spatialcov^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j
+%		\end{align}
+%	\end{subequations}
+%	for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$.
+%\end{prop}
+
+%The proof of Proposition \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}. The proof is straightforward, and follows directly from the definitions.
 
@@
 
-Since a probability distribution concentrates within a small number of standard deviations of its mean, knowing $\mu(x)$ and $\Sigma(x)$ permits us to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
-This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches.'' We compute one batch of $\impulseresponse_{x_i}$'s by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$.
-The batch of points $x_i$ are chosen via a greedy ellipsoid packing algorithm so that the support of $\impulseresponse_{x_i}$ and the support of $\impulseresponse_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$.
-Because these supports do not overlap, we can post-process the response of $\Aop$ to the Dirac comb to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i$ in the batch---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$.
-This is illustrated in Figure FIG.
+The results from Section \ref{sec:intromoments} allow us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions.
+Since a probability distribution concentrates within a small number of standard deviations of its mean, knowing $\mu(x)$ and $\Sigma(x)$ permits us to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$ (Section \ref{eq:mean_and_covariance_estimation}).
+This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}).
+We compute one batch of $\impulseresponse_{x_i}$'s by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$.
+The batch of points $x_i$ are chosen via a greedy ellipsoid packing algorithm so that the support of $\impulseresponse_{x_i}$ and the support of $\impulseresponse_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$ (Section \ref{sec:sample_point_selection}).
+Because these supports do not overlap, we can post-process the response of $\Aop$ to the Dirac comb to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i$ in the batch---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}).
 The process is repeated to get more batches, until a desired number of impulse responses is reached.
 
-Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation.
-The details of the interpolation will be described in Section \ref{sec:approximate_kernel_entries}, but the key principle behind the interpolation is the preconditioned local mean displacement invariance assumption discussed in Section \ref{eq:local_mean_displacement_invariance}.
+Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation (Section \ref{sec:approximate_kernel_entries}).
+The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section \ref{eq:local_mean_displacement_invariance}.
 %Specifically, we make the assumption that
@@
 
-The ability to rapidly evaluate approximate kernel entries $\Phi(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Phi}_H$, of $\Phi$, using the conventional adaptive cross H-matrix construction method.
-In this method, one forms low-rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
-We then use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation of $\Aop$, denoted $\mathbf{A}_H$.
-Schematically the overall H-matrix approximation proceeds as follows:
-\begin{equation*}
-\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \{x_i\}_{i=1}^r \longrightarrow \{\impulseresponse_{x_i}\}_{i=1}^r \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
-\end{equation*}
-
-%Often we are interested in approximating operators $\Aop$ that are symmetric, or symmetric positive semi-definite.
-
 If $\Aop$ is symmetric, we take advantage of symmetry to construct a more accurate approximation at negligible additional cost
@@
 %we combine 
-by combining information from impulse responses of both $\Aop$ and $\Aop^T$ in the interpolation procedure.
+by combining information from impulse responses of both $\Aop$ and $\Aop^T$ in the interpolation procedure (Section \ref{sec:symmetric_improvements}).
 % for approximating $\Phi(y,x)$, to get a more accurate approximation at negligible additional cost.
@@
 
+The ability to rapidly evaluate approximate kernel entries $\Phi(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Phi}_H$, of $\Phi$, using the conventional adaptive cross H-matrix construction method (Section \ref{sec:h_matrix_construction_short}).
+In this method, one forms low-rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
+We then use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation of $\Aop$, denoted $\mathbf{A}_H$.
+
 When $\Aop$ is positive semi-definite, $\mathbf{A}_H$ may be indefinite due to errors in the approximation.
@@
 This is described in Appendix \ref{sec:make_spd}.
+
+Schematically the overall H-matrix approximation proceeds as follows:
+\begin{equation*}
+\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \{x_i\}_{i=1}^r \longrightarrow \underbrace{\{\impulseresponse_{x_i}\}_{i=1}^r}_{\text{as }\eta_b\text{'s}} \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
+\end{equation*}
+The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}.
+
+%Often we are interested in approximating operators $\Aop$ that are symmetric, or symmetric positive semi-definite.
 
@@
 
-In this section we detail our method for computing the H-matrix approximation, $\mathbf{A}_H$, of $\Aop$, as outlined in Section \ref{sec:intro_overview}.
+%In this section we detail our method for computing the H-matrix approximation, $\mathbf{A}_H$, of $\Aop$, as outlined in Section \ref{sec:intro_overview}.
 %The supports of the impulse responses $\phi_x$ are estimated to be contained within ellipsoids (Section \ref{eq:mean_and_covariance_estimation}).
@@
 The ellipsoid support estimates are then used in a greedy ellipsoid packing procedure to choose batches, $\pointbatch_b$, of sample points $x_i$ (Section \ref{sec:sample_point_selection}).
-This section is organized as follows:
-\begin{description}
-	\item[Section \ref{eq:mean_and_covariance_estimation}:] We describe our ellipsoid estimate for the support of $\phi_x$.
-	\item[Section \ref{sec:moments}:] We show how to compute $V,\mu,\Sigma$.
-	\item[Section \ref{sec:sample_point_selection}:] We present the greedy ellipsoid packing procedure used to choose batches, $\pointbatch_b$, of sample points $x_i$.
-	\item[Section \ref{sec:get_impulse_response}:] We describe how to compute $\phi_{x_i}$'s, in batches, $\eta_b$.
-	\item[Section \ref{sec:approximate_kernel_entries}:] We show how we interpolate $\phi_{x_i}$'s to approximate $\Phi(y,x)$.
-	\item[Section \ref{sec:finite_element_kernel}:] We discuss finite element discretization of $\Phi$.
-	\item[Section \ref{sec:h_matrix_construction_short}:] We describe how we construct the H-matrices $\mathbf{\Phi}_H$ and $\mathbf{A}_H$.
-	\item[Section \ref{sec:overall_cost}:] We review the total computational cost to construct $\mathbf{A}_H$, and list the cost to perform further linear algebra tasks involving $\mathbf{A}_H$ using H-matrix methods.
-\end{description}
-The optional method for making $\mathbf{A}_H$ positive semi-definite is described in Appendix \ref{sec:make_spd}.
+%This section is organized as follows:
+%\begin{description}
+%	\item[Section \ref{eq:mean_and_covariance_estimation}:] We describe our ellipsoid estimate for the support of $\phi_x$.
+%	\item[Section \ref{sec:moments}:] We show how to compute $V,\mu,\Sigma$.
+%	\item[Section \ref{sec:sample_point_selection}:] We present the greedy ellipsoid packing procedure used to choose batches, $\pointbatch_b$, of sample points $x_i$.
+%	\item[Section \ref{sec:get_impulse_response}:] We describe how to compute $\phi_{x_i}$'s, in batches, $\eta_b$.
+%	\item[Section \ref{sec:approximate_kernel_entries}:] We show how we interpolate $\phi_{x_i}$'s to approximate $\Phi(y,x)$.
+%	\item[Section \ref{sec:finite_element_kernel}:] We discuss finite element discretization of $\Phi$.
+%	\item[Section \ref{sec:h_matrix_construction_short}:] We describe how we construct the H-matrices $\mathbf{\Phi}_H$ and $\mathbf{A}_H$.
+%	\item[Section \ref{sec:overall_cost}:] We review the total computational cost to construct $\mathbf{A}_H$, and list the cost to perform further linear algebra tasks involving $\mathbf{A}_H$ using H-matrix methods.
+%\end{description}
+%The optional method for making $\mathbf{A}_H$ positive semi-definite is described in Appendix \ref{sec:make_spd}.
 The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}.

### 2022-03-15 paper/localpsf.tex
restructuring paper

@@
 This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}).
-We compute one batch of $\impulseresponse_{x_i}$'s by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$.
-The batch of points $x_i$ are chosen via a greedy ellipsoid packing algorithm so that the support of $\impulseresponse_{x_i}$ and the support of $\impulseresponse_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$ (Section \ref{sec:sample_point_selection}).
-Because these supports do not overlap, we can post-process the response of $\Aop$ to the Dirac comb to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i$ in the batch---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}).
+We compute the impulse response batch $\eta_b$ by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section \ref{sec:get_impulse_response}).
+The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support of $\impulseresponse_{x_i}$ and the support of $\impulseresponse_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$ (Section \ref{sec:sample_point_selection}).
+Because these supports do not overlap, we can post-process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}).
 The process is repeated to get more batches, until a desired number of impulse responses is reached.
 
-Once the scattered impulse responses $\impulseresponse_{x_i}$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation (Section \ref{sec:approximate_kernel_entries}).
+Once the impulse response batches $\eta_b$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation (Section \ref{sec:approximate_kernel_entries}).
 The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section \ref{eq:local_mean_displacement_invariance}.
@@
 When $\Aop$ is positive semi-definite, $\mathbf{A}_H$ may be indefinite due to errors in the approximation.
-In this case, we modify the H-matrix representation of $\mathbf{A}_H$ to make it positive semi-definite, by applying a specially chosen rational matrix function to $\mathbf{A}_H$.
+In this case, one may (optionally) modify the H-matrix representation of $\mathbf{A}_H$ to make it positive semi-definite, by applying a specially chosen rational matrix function to $\mathbf{A}_H$.
 This is described in Appendix \ref{sec:make_spd}.
@@
 \begin{equation*}
-\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \{x_i\}_{i=1}^r \longrightarrow \underbrace{\{\impulseresponse_{x_i}\}_{i=1}^r}_{\text{as }\eta_b\text{'s}} \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
-\end{equation*}
-The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}.
+\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \underbrace{\{x_i\}_{i=1}^r}_{\text{as }S_b\text{'s}} \longrightarrow \underbrace{\{\impulseresponse_{x_i}\}_{i=1}^r}_{\text{as }\eta_b\text{'s}} \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
+\end{equation*}
+The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}. The cost to construct $\mathbf{A}_H$ is discussed in Section \ref{sec:overall_cost}.
 
@@
 
-\subsubsection{Cost to evaluate one approximate kernel entry}
-
-In the non-symmetric case, evaluating $\widetilde{\Phi}(y,x)$ requires us to perform one $k$-nearest neighbor query of the KD-tree of sample points, $k$ point collision queries of the AABB-tree of mesh cells, and solve a $k \times k$ linear system.
-In the symmetric case, we must perform two $k$-nearest neighbor queries of the KD-tree of sample points, $2k$ point collision queries of the AABB-tree of mesh cells, and solve a $4k \times 2k$ linear least squares problem via QR factorization.
-Both of these cases require $O(k \log r + k \log N_\text{cell} + k^3)$ work, where $r$ is the number of sample points, and $N_\text{cell}$ is the number of cells in the mesh used to discretize the problem.
-Since $k$ is small, this is cheap.
-
 %In principle, this equation is still solvable if $\left(\widetilde{f}_x\right)_i = \left(\widetilde{f}_x\right)_j$, which should hold in principle.
@@
 
+In the non-symmetric case, evaluating $\widetilde{\Phi}(y,x)$ requires us to perform one $k$-nearest neighbor query of the KD-tree of sample points, $k$ point collision queries of the AABB-tree of mesh cells, and solve a $k \times k$ linear system.
+In the symmetric case, we must perform two $k$-nearest neighbor queries of the KD-tree of sample points, $2k$ point collision queries of the AABB-tree of mesh cells, and solve a $4k \times 2k$ linear least squares problem via QR factorization.
+Both of these cases require $O(k \log r + k \log N_\text{cell} + k^3)$ work, where $r$ is the number of sample points, and $N_\text{cell}$ is the number of cells in the mesh used to discretize the problem.
+Since $k$ is small, this is cheap.
+
 Computing $V$, $\mu$, and $\Sigma$ require $1$, $d$, and $d(d+1)/2$ operator actions of $\Aop$, respectively.

### 2022-03-15 paper/localpsf.tex
restructuring paper

@@
 
-Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, and let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space.
-Further, let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
+Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
 That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
-
-We define a further approximation, $\widetilde{\Phi}^h \approx \widetilde{\Phi}$, as follows:
+For $u_h \in V_h$ we write $\bm{u}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$.
+I.e.,
+\begin{equation}
+\label{eq:fem_coeff_basis}
+u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
+\end{equation}
+
+%Further, let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
+That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
+
+%We define the finite element approximation of the kernel, $\Phi^h \approx \Phi$, as follows:
+The interpolation of $\Phi$ onto $V_h \otimes V_h$, which we denote by $\Phi^h$, is given as follows:
 \begin{equation}
 \label{eq:defn_of_Akerpcmesh}
-\widetilde{\Phi}^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \widetilde{\Phi}(p_i,p_j) \febasis_i(y) \febasis_j(x).
-\end{equation}
-The kernel $\widetilde{\Phi}^h$ is the interpolation of $\widetilde{\Phi}$ onto $V_h \otimes V_h$.
+\Phi(y,x) \approx \Phi^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \Phi(p_i,p_j) \febasis_i(y) \febasis_j(x).
+\end{equation}
+%The kernel $\Phi^h$ is the interpolation of $\Phi$ onto $V_h \otimes V_h$.
 Error in the overall kernel approximation, $\widetilde{\Phi}^h \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$.
-
-Let $u_h \in V_h$ and $v_h \in V_h$, and let $\bm{u}$ and $\bm{v}$ be the coefficient vectors for $u_h$ and $v_h$, respectively, with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$.
-I.e.,
-\begin{equation}
-\label{eq:fem_coeff_basis}
-u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x),
-\end{equation}
-and similar for $v_h$.
-Replacing $\Aker$ with $\widetilde{\Phi}^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward but tedious algebraic manipulations, yields the following approximation:
+Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations (see Appendix \ref{app:proofs} for details), yields the following approximation:
 \begin{equation}
@@
 %While $\Aop$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\Aop$ to functions in the finite element space. 
-where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the following matrix of approximate kernel entries:
+where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the matrix of kernel entries evaluated at pairs of Lagrange nodes:
 \begin{equation}
 \label{eq:Akerpcmat_entries}
-\mathbf{\Phi}_{ij} := \widetilde{\Phi}(p_i, p_j),
+\mathbf{\Phi}_{ij} := \Phi(p_i, p_j),
 \end{equation}
@@
 \end{equation*}
+%Matrix entries $\Phi(p_i, p_j)$ are expensive to compute because $\Aop$ is only available matrix-free.
 %Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, and \eqref{eq:Akerpcmat_entries}, the definition of the mass matrix, and performing algebraic manipulations, it is straightforward but tedious to show that
@@
 %\end{equation*}
-The finite element matrix representation of our approximation of $\Aop$ is therefore given by
+Our matrix representation of $\Aop$ is therefore given by
 \begin{equation}
@@
 \end{equation}
+Note that $\mathbf{A}$ is not a Galerkin projection of $\Aop$ onto $V_h$ because of interpolation error in the kernel approximation $\Phi^h \approx \Phi$.
+
+%We neglect this interpolation error because it is the same magnitude as the finite element approximation error. 
+%Note that the matrices $\mathbf{A}$ and $\mathbf{\Phi}$ are dense $N \times N$ matrix that only exist in theory.
+Building and storing these matrices is computationally intractible.
+Computing individual matrix entries $\mathbf{A}_{ij}$ and $\mathbf{\Phi}_{ij}$ is computationally expensive.
 
@@
 
-The ability to rapidly evaluate approximate kernel entries $\Phi(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Phi}_H$, of $\Phi$, using the conventional adaptive cross H-matrix construction method (Section \ref{sec:h_matrix_construction_short}).
+The ability to rapidly evaluate approximate kernel entries $\Phi(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Phi}_H \approx \mathbf{\Phi}$, using the conventional adaptive cross H-matrix construction method (Section \ref{sec:h_matrix_construction_short}).
 In this method, one forms low-rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
-We then use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation of $\Aop$, denoted $\mathbf{A}_H$.
+We then use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$.
 

### 2022-03-15 paper/localpsf.tex
restructuring paper

@@
 
-\section{Preliminaries}
+\section{Key concepts}
+\label{sec:prerequisites}
+
+In this section we present three key concepts that our method is based on.
+The method itself will be presented in Section \ref{sec:method}.
+First, we define moments of the impulse responses $\phi_x$, and show how these moments can be computed efficiently (Section \ref{sec:intromoments}).
+Our method will use impulse response moments to estimate the supports of the functions $\phi_x$.
+Second, we describe an impulse response approximation that we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}).
+Our method will interpolate impulse responses using this approximation.
+Third, we discuss how discretization of continuous functions by finite elements leads to matrix discretizations $\mathbf{A}$ and $\mathbf{\Phi}$ of $\Aop$ and $\Phi$, respectively (Section \ref{sec:finite_element_kernel}).
+The goal of our method is to compute H-matrix approximations of these matrices.
+%While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead with minor modifications.
 
@@
 Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
-But this direct approach is computationally infeasible because it requires one operator action of $\Aop$ per point $x$, and we will need to know $V(x)$, $\mu(x)$, and $\Sigma(x)$ at all points $x$.
+But this direct approach is computationally infeasible because it requires one operator action of $\Aop$ per point $x$, and our method will need to know $V(x)$, $\mu(x)$, and $\Sigma(x)$ at all points $x$.
 Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions.
@@
 \subsection{Local mean-displacement invariance}
-\label{eq:local_mean_displacement_invariance}
+\label{sec:local_mean_displacement_invariance}
 
@@
 
-\subsection{Finite element discretization of the integral kernel}
+\subsection{Finite element discretization}
 \label{sec:finite_element_kernel}
@@
 %While $\Aop$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\Aop$ to functions in the finite element space. 
-where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the matrix of kernel entries evaluated at pairs of Lagrange nodes:
+where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
 \begin{equation}
@@
 \end{equation}
-Note that $\mathbf{A}$ is not a Galerkin projection of $\Aop$ onto $V_h$ because of interpolation error in the kernel approximation $\Phi^h \approx \Phi$.
-
-%We neglect this interpolation error because it is the same magnitude as the finite element approximation error. 
+Note that $\mathbf{A}$ is not the Galerkin projection of $\Aop$ onto $V_h$ because of the approximation $\Phi^h \approx \Phi$.
+%because of the approximation $\Phi^h \approx \Phi$. 
+Hence, $\mathbf{A}$ may differ slightly from the implicitly defined matrix that performs the mapping $\mathbf{u} \rightarrow \mathbf{v}$ corresponding to the operator action $v=\Aop u$.
+
+%The discrete version of the operator action $v = \Aop u$ is a mapping of vectors $\mathbf{u} \rightarrow \mathbf{v}$, which is done through an implicit computational procedure.
+Typically the implicitly defined matrix that performs this mapping is the Galerkin projection of $\Aop$ onto $V_h$.
+
+%We neglect this error because it is the same magnitude as the error in finite element approximation of functions $u_h \approx u$. 
 %Note that the matrices $\mathbf{A}$ and $\mathbf{\Phi}$ are dense $N \times N$ matrix that only exist in theory.
@@
 Once the impulse response batches $\eta_b$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation (Section \ref{sec:approximate_kernel_entries}).
-The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section \ref{eq:local_mean_displacement_invariance}.
+The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section \ref{sec:local_mean_displacement_invariance}.
 %Specifically, we make the assumption that

### 2022-03-16 paper/localpsf.tex
restructuring paper

@@
 
-\subsection{Existing work}
-
-The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term \cite{CCGOPAPERS}.
-Conventionally, either the Lanczos method or the randomized singular value decomposition \cite{HMTRANDOM} are used to perform the low rank approximation using only matrix-vector products.
-These methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of matrix-vector products are required to form the aforementioned low rank approximation.
+\subsection{Background}
+
+The most popular matrix-free operator approximation methods are based on low rank approximation \cite{CCGOPAPERS}.
+Conventionally, either the Lanczos method, or the randomized singular value decomposition \cite{HMTRANDOM}, or CUR approximation/skeletonization \cite{CUR,MAHONEY} are used to perform the low rank approximation using only matrix-vector products.
+However, methods based on low rank approximation are not suitable for high rank operators, and many operators of practical interest are high rank.
+In the context of Hessians, methods based on low rank approximation suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of matrix-vector products are required to form a low rank approximation of the Hessian.
 The ideal scenario from a scientific perspective (highly informative data) is therefore the worst case scenario from a computational perspective (large computational cost) \cite{MYDISSERTATION}.
-Although the method we present in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest.
-For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high rank approximations of an operator using a small number of matrix-vector products.
 
 %In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
+Our work builds upon a history of ``product-convolution`` approximations---weighted sums of convolution operators with spatially varying weights.
+A single convolution operator may be either high rank or low rank, so product-convolution approximations can (where applicable) approximate high rank operators with a small number of convolution operators.
 Product-convolution approximations have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}.
-For background on product-convolution approximations, we recommend reading the following papers: \cite{PRODCONVGOOD}.
-We build upon the class of product convolution approximations in which the convolution kernels are shifted impulse responses of $\Aop$ to point sources at a collection of points $x_i$ \cite{PRODCONVLIST}.
-A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, or splines \cite{PRODCONVSPLINES}.
+For background, we recommend reading the following papers: \cite{PRODCONVGOOD}.
+We specifically build upon the class of product convolution approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i$ \cite{PRODCONVLIST}.
+A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, splines \cite{PRODCONVSPLINES}, or radial basis functions \cite{PRODCONVRBF}.
 With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
 In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
-But even with adaptive refinement, many matrix vector products with $\Aop$ may be required.
-In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\convkernel_i$ are computed with each matrix vector product.
+But even with adaptive refinement, many matrix vector products with $\Aop$ are required.
+In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\phi_i$ are computed with each matrix vector product.
 
@@
 First, we define moments of the impulse responses $\phi_x$, and show how these moments can be computed efficiently (Section \ref{sec:intromoments}).
-Our method will use impulse response moments to estimate the supports of the functions $\phi_x$.
+Our method will use these moments extensively.
 Second, we describe an impulse response approximation that we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}).
@@
 
+Computing $V$, $\mu$, and $\Sigma$ require $1$, $d$, and $d(d+1)/2$ operator actions of $\Aop$, respectively.
+Computing the impulse response batches $\eta_b$ requires one operator action of $\Aop$ per batch.
+Picking the sample points
+
 In the non-symmetric case, evaluating $\widetilde{\Phi}(y,x)$ requires us to perform one $k$-nearest neighbor query of the KD-tree of sample points, $k$ point collision queries of the AABB-tree of mesh cells, and solve a $k \times k$ linear system.
@@
 Since $k$ is small, this is cheap.
-
-Computing $V$, $\mu$, and $\Sigma$ require $1$, $d$, and $d(d+1)/2$ operator actions of $\Aop$, respectively.
-Computing the impulse response batches $\eta_b$ requires one operator action of $\Aop$ per batch.
-Picking the sample points
 
@@
 
+Although the method we present in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest.
+For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high rank approximations of an operator using a small number of matrix-vector products.
+
 \appendix

### 2022-03-16 paper/localpsf.tex
restructuring paper

@@
 
-In this section we present three key concepts that our method is based on.
+In this section we present the key concepts that our method is based on.
 The method itself will be presented in Section \ref{sec:method}.
-First, we define moments of the impulse responses $\phi_x$, and show how these moments can be computed efficiently (Section \ref{sec:intromoments}).
+First, we define moments of the impulse responses, $\phi_x$, and show how these moments can be computed efficiently (Section \ref{sec:intromoments}).
 Our method will use these moments extensively.
-Second, we describe an impulse response approximation that we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}).
+Second, we present ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{eq:mean_and_covariance_estimation}).
+Our method will use these ellipsoid estimates to choose sets of impulse responses that can be computed in batches.
+Third, we describe an impulse response approximation that we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}).
 Our method will interpolate impulse responses using this approximation.
-Third, we discuss how discretization of continuous functions by finite elements leads to matrix discretizations $\mathbf{A}$ and $\mathbf{\Phi}$ of $\Aop$ and $\Phi$, respectively (Section \ref{sec:finite_element_kernel}).
-The goal of our method is to compute H-matrix approximations of these matrices.
+Fourth, we discuss how discretization of continuous functions by finite elements leads to matrix discretizations $\mathbf{A}$ and $\mathbf{\Phi}$ of $\Aop$ and $\Phi$, respectively (Section \ref{sec:finite_element_kernel}).
+The primary function of our method is to compute H-matrix approximations of these matrices.
 %While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead with minor modifications.
@@
 
+\subsection{Impulse response support ellipsoids}
+\label{eq:mean_and_covariance_estimation}
+
+%Because of the non-negative impulse response property, for each $x \in \Omega$, the impulse response $\impulseresponse_x$ is a non-negative function $y \mapsto \impulseresponse_x(y)$.
+Hence, $\impulseresponse_x$ is a scaled probability distribution.
+Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor (the ``volume'' of $\impulseresponse_x$), let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
+
+We make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
+\begin{equation}
+\label{eq:support_ellipsoid}
+E_x := \{x' \in \Omega: (x' - \spatialmean(x))^T \spatialcov(x)^{-1} (x' - \spatialmean(x)) \le \tau^2\},
+\end{equation}
+where $\tau$ is a fixed constant.
+The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as the normalized version of $\impulseresponse_x$.
+
+The quantity $\tau$ is a parameter that must be chosen by the user of the algorithm.
+The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$.
+However, in Section \ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
+We find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results.
+The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.
+
+%To determine $E_x$, we must compute the scalar $\spatialvol(x)$, the length-$\gdim$ vector $\spatialmean(x)$, and the $\gdim \times \gdim$ matrix $\spatialcov(x)$.
+In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of polynomial functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).
+
 \subsection{Local mean-displacement invariance}
@@
 
-The results from Section \ref{sec:intromoments} allow us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions.
-Since a probability distribution concentrates within a small number of standard deviations of its mean, knowing $\mu(x)$ and $\Sigma(x)$ permits us to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$ (Section \ref{eq:mean_and_covariance_estimation}).
+The results from Section \ref{sec:intromoments} allow us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions, and the results of Section \ref{eq:mean_and_covariance_estimation} permit us to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
 This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}).
@@
 \end{algorithm2e}
-
-\subsection{Impulse response support ellipsoids}
-\label{eq:mean_and_covariance_estimation}
-
-%Because of the non-negative impulse response property, for each $x \in \Omega$, the impulse response $\impulseresponse_x$ is a non-negative function $y \mapsto \impulseresponse_x(y)$.
-Hence, $\impulseresponse_x$ is a scaled probability distribution.
-Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor (the ``volume'' of $\impulseresponse_x$), let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
-
-We make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
-\begin{equation}
-\label{eq:support_ellipsoid}
-E_x := \{x' \in \Omega: (x' - \spatialmean(x))^T \spatialcov(x)^{-1} (x' - \spatialmean(x)) \le \tau^2\},
-\end{equation}
-where $\tau$ is a fixed constant.
-The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as the normalized version of $\impulseresponse_x$.
-
-The quantity $\tau$ is a parameter that must be chosen by the user of the algorithm.
-The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$.
-However, in Section \ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
-We find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results.
-The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.
-
-To determine $E_x$, we must compute the scalar $\spatialvol(x)$, the length-$\gdim$ vector $\spatialmean(x)$, and the $\gdim \times \gdim$ matrix $\spatialcov(x)$.
-In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of polynomial functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).
 

### 2022-03-16 paper/localpsf.tex
restructuring paper

@@
 The output of our method is a hierarchical matrix (H-matrix) approximation, $\mathbf{A}_H$, of a discretization of $\Aop$.
+H-matrices are a matrix format in which the matrix is re-ordered then recursively subdivided into blocks, so that many off-diagonal blocks of the matrix are low rank.
+For more details about H-matrices, see Appendix \ref{app:h_matrix}.
 While operator actions of $\Aop$ on vectors are typically computationally expensive, fast H-matrix methods permit us to perform matrix-vector products with $\mathbf{A}_H$ cheaply.
@@
 
-\section{Key concepts}
+\section{Main concepts}
 \label{sec:prerequisites}
 
-In this section we present the key concepts that our method is based on.
+In this section we present the three main concepts that our method is based on.
 The method itself will be presented in Section \ref{sec:method}.
@@
 Our method will interpolate impulse responses using this approximation.
-Fourth, we discuss how discretization of continuous functions by finite elements leads to matrix discretizations $\mathbf{A}$ and $\mathbf{\Phi}$ of $\Aop$ and $\Phi$, respectively (Section \ref{sec:finite_element_kernel}).
-The primary function of our method is to compute H-matrix approximations of these matrices.
-%While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead with minor modifications.
 
@@
 Here $f/g$ denotes pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denotes pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
+The resulting method form computing $V$, $\mu$, and $\Sigma$ is shown in Algorithm \ref{alg:varhpi_mean_cov}.
+
+\begin{algorithm2e}
+	\SetAlgoNoLine
+	\SetKwInOut{Input}{Input}
+	\SetKwInOut{Output}{Output}
+	{	
+		\Input{Operator $\Aop$}
+		\Output{$\spatialvol$, $\spatialmean$, and $\spatialcov$}
+
+		\tcp{Compute scaling factor $\spatialvol$}
+
+		Form constant function $C(x)=1$
+
+		Compute $\spatialvol = \left(\Aop^T C\right)^*$
+
+		\tcp{Compute mean $\spatialmean$}
+		\For{$i=1,2,\dots,\gdim$}{
+			Form linear function $L^i(x) = x^i$
+
+			Compute $\spatialmean^i = \left(\Aop^T L^i\right)^* / \spatialvol$
+		}
+		\tcp{Compute covariance $\spatialcov$}
+		\For{$i=1,2,\dots,\gdim$}{
+			\For{$j=1,\dots,i$}{
+				Form quadratic function $Q^{ij}(x) = x^i x^j$
+
+				Compute $\spatialcov^{ij} = \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j$
+
+				Set $\spatialcov^{ji} = \spatialcov^{ij}$
+
+			}
+		}
+
+	}
+	\caption{Compute scaling factor $\spatialvol$, mean $\spatialmean$, and covariance $\spatialcov$}
+	\label{alg:varhpi_mean_cov}
+\end{algorithm2e}
 
@@
 
-\subsection{Finite element discretization}
-\label{sec:finite_element_kernel}
-
-Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
-That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
-For $u_h \in V_h$ we write $\bm{u}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$.
-I.e.,
-\begin{equation}
-\label{eq:fem_coeff_basis}
-u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
-\end{equation}
-
-%Further, let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
-That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
-
-%We define the finite element approximation of the kernel, $\Phi^h \approx \Phi$, as follows:
-The interpolation of $\Phi$ onto $V_h \otimes V_h$, which we denote by $\Phi^h$, is given as follows:
-\begin{equation}
-\label{eq:defn_of_Akerpcmesh}
-\Phi(y,x) \approx \Phi^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \Phi(p_i,p_j) \febasis_i(y) \febasis_j(x).
-\end{equation}
-%The kernel $\Phi^h$ is the interpolation of $\Phi$ onto $V_h \otimes V_h$.
-Error in the overall kernel approximation, $\widetilde{\Phi}^h \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$.
-Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations (see Appendix \ref{app:proofs} for details), yields the following approximation:
-\begin{equation}
-\label{eq:kernel_representation_pcmesh}
-(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
-= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u}.
-\end{equation}
-%While $\Aop$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\Aop$ to functions in the finite element space. 
-where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
-\begin{equation}
-\label{eq:Akerpcmat_entries}
-\mathbf{\Phi}_{ij} := \Phi(p_i, p_j),
-\end{equation}
-and $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ is the finite element mass matrix which has entries
-\begin{equation*}
-\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
-\end{equation*}
-%Matrix entries $\Phi(p_i, p_j)$ are expensive to compute because $\Aop$ is only available matrix-free.
-%Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, and \eqref{eq:Akerpcmat_entries}, the definition of the mass matrix, and performing algebraic manipulations, it is straightforward but tedious to show that
-%\begin{equation*}
-%	\left(\AopPcMesh u_h \right)(v_h) = \bm{v}^T \bm{M} \AkerPcMat \bm{M} \bm{u}.
-%\end{equation*}
-Our matrix representation of $\Aop$ is therefore given by
-\begin{equation}
-\label{eq:Amatpc_defn}
-\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
-\end{equation}
-Note that $\mathbf{A}$ is not the Galerkin projection of $\Aop$ onto $V_h$ because of the approximation $\Phi^h \approx \Phi$.
-%because of the approximation $\Phi^h \approx \Phi$. 
-Hence, $\mathbf{A}$ may differ slightly from the implicitly defined matrix that performs the mapping $\mathbf{u} \rightarrow \mathbf{v}$ corresponding to the operator action $v=\Aop u$.
-
-%The discrete version of the operator action $v = \Aop u$ is a mapping of vectors $\mathbf{u} \rightarrow \mathbf{v}$, which is done through an implicit computational procedure.
-Typically the implicitly defined matrix that performs this mapping is the Galerkin projection of $\Aop$ onto $V_h$.
-
-%We neglect this error because it is the same magnitude as the error in finite element approximation of functions $u_h \approx u$. 
-%Note that the matrices $\mathbf{A}$ and $\mathbf{\Phi}$ are dense $N \times N$ matrix that only exist in theory.
-Building and storing these matrices is computationally intractible.
-Computing individual matrix entries $\mathbf{A}_{ij}$ and $\mathbf{\Phi}_{ij}$ is computationally expensive.
-
 \section{Method}
@@
 
-The results from Section \ref{sec:intromoments} allow us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions, and the results of Section \ref{eq:mean_and_covariance_estimation} permit us to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
+After discretization with finite elements\footnote{While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead, with minor modifications to our method.} (Section \ref{sec:finite_element_kernel}), $\Aop$ and $\Phi$ become matrices $\mathbf{A}$ and $\mathbf{\Phi}$, respectively.
+Because $\Aop$ is only available matrix-free and $\Phi$ is only defined implicitly in terms of $\Aop$, the matrices $\mathbf{A}$ and $\mathbf{\Phi}$ are not easily accessible.
+Directly computing and storing these matrices is computationally intractible.
+In this section, we present our method for constructing accessible H-matrix approximations, $\mathbf{A}_H$ and $\mathbf{\Phi}_H$ of $\mathbf{A}$ and $\mathbf{\Phi}$, respectively, using a small number of operator actions of $\Aop$ and $\Aop^T$.
+
+%Fourth, we discuss how discretization of continuous functions by finite elements leads to matrix discretizations $\mathbf{A}$ and $\mathbf{\Phi}$ of $\Aop$ and $\Phi$, respectively (Section \ref{sec:finite_element_kernel}).
+The primary function of our method is to compute H-matrix approximations of these matrices.
+%While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead with minor modifications.
+
+\subsection{Overview}
+\label{sec:method_overview}
+
+Algorithm \ref{alg:varhpi_mean_cov} from Section \ref{sec:intromoments} allows us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} from Section \ref{eq:mean_and_covariance_estimation} uses $V$, $\mu$, and $\Sigma$ to construct an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
 This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}).
@@
 
-\begin{algorithm2e}
-	\SetAlgoNoLine
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Output}
-	{	
-		\Input{Operator $\Aop$}
-		\Output{$\spatialvol$, $\spatialmean$, and $\spatialcov$}
-
-		\tcp{Compute scaling factor $\spatialvol$}
-
-		Form constant function $C(x)=1$
-
-		Compute $\spatialvol = \left(\Aop^T C\right)^*$
-
-		\tcp{Compute mean $\spatialmean$}
-		\For{$i=1,2,\dots,\gdim$}{
-			Form linear function $L^i(x) = x^i$
-
-			Compute $\spatialmean^i = \left(\Aop^T L^i\right)^* / \spatialvol$
-		}
-		\tcp{Compute covariance $\spatialcov$}
-		\For{$i=1,2,\dots,\gdim$}{
-			\For{$j=1,\dots,i$}{
-				Form quadratic function $Q^{ij}(x) = x^i x^j$
-
-				Compute $\spatialcov^{ij} = \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j$
-
-				Set $\spatialcov^{ji} = \spatialcov^{ij}$
-
-			}
-		}
-
-	}
-	\caption{Compute scaling factor $\spatialvol$, mean $\spatialmean$, and covariance $\spatialcov$}
-	\label{alg:varhpi_mean_cov}
-\end{algorithm2e}
+\subsection{Finite element discretization}
+\label{sec:finite_element_kernel}
+
+Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
+That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
+For $u_h \in V_h$ we write $\bm{u}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$.
+I.e.,
+\begin{equation}
+\label{eq:fem_coeff_basis}
+u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
+\end{equation}
+
+%Further, let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
+That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
+
+%We define the finite element approximation of the kernel, $\Phi^h \approx \Phi$, as follows:
+The interpolation of $\Phi$ onto $V_h \otimes V_h$, which we denote by $\Phi^h$, is given as follows:
+\begin{equation}
+\label{eq:defn_of_Akerpcmesh}
+\Phi(y,x) \approx \Phi^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \Phi(p_i,p_j) \febasis_i(y) \febasis_j(x).
+\end{equation}
+%The kernel $\Phi^h$ is the interpolation of $\Phi$ onto $V_h \otimes V_h$.
+Error in the overall kernel approximation, $\widetilde{\Phi}^h \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$.
+Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations (see Appendix \ref{app:proofs} for details), yields the following approximation:
+\begin{equation}
+\label{eq:kernel_representation_pcmesh}
+(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
+= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u}.
+\end{equation}
+%While $\Aop$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\Aop$ to functions in the finite element space. 
+where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
+\begin{equation}
+\label{eq:Akerpcmat_entries}
+\mathbf{\Phi}_{ij} := \Phi(p_i, p_j),
+\end{equation}
+and $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ is the sparse finite element mass matrix which has entries
+\begin{equation*}
+\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
+\end{equation*}
+%Matrix entries $\Phi(p_i, p_j)$ are expensive to compute because $\Aop$ is only available matrix-free.
+%Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, and \eqref{eq:Akerpcmat_entries}, the definition of the mass matrix, and performing algebraic manipulations, it is straightforward but tedious to show that
+%\begin{equation*}
+%	\left(\AopPcMesh u_h \right)(v_h) = \bm{v}^T \bm{M} \AkerPcMat \bm{M} \bm{u}.
+%\end{equation*}
+Our matrix representation, $\mathbf{A}$, of $\Aop$ is therefore given by
+\begin{equation}
+\label{eq:Amatpc_defn}
+\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
+\end{equation}
+Note that $\mathbf{A}$ is not exactly the Galerkin projection of $\Aop$ onto $V_h$ because of the approximation $\Phi^h \approx \Phi$.
+%because of the approximation $\Phi^h \approx \Phi$. 
+Hence, $\mathbf{A}$ may differ slightly from the implicitly defined matrix that performs the mapping $\mathbf{u} \rightarrow \mathbf{v}$ corresponding to the discrete version of the operator action $v=\Aop u$.
+%It is computationally intractible to build and store the matrices $\mathbf{A}$ and $\mathbf{\Phi}$, and individual matrix entries $\mathbf{A}_{ij}$ and $\mathbf{\Phi}_{ij}$ are not easily accessible.
+In the remainder of this section, we show how to build accessible approximations of these matrices.
+
+%The discrete version of the operator action $v = \Aop u$ is a mapping of vectors $\mathbf{u} \rightarrow \mathbf{v}$, which is done through an implicit computational procedure.
+Typically the implicitly defined matrix that performs this mapping is the Galerkin projection of $\Aop$ onto $V_h$.
+
+%We neglect this error because it is the same magnitude as the error in finite element approximation of functions $u_h \approx u$. 
+%Note that the matrices $\mathbf{A}$ and $\mathbf{\Phi}$ are dense $N \times N$ matrix that only exist in theory.
+Building and storing these matrices is computationally intractible.
+Computing individual matrix entries $\mathbf{A}_{ij}$ and $\mathbf{\Phi}_{ij}$ is computationally expensive.
 

### 2022-03-16 paper/localpsf.tex
restructuring paper

@@
 The output of our method is a hierarchical matrix (H-matrix) approximation, $\mathbf{A}_H$, of a discretization of $\Aop$.
-H-matrices are a matrix format in which the matrix is re-ordered then recursively subdivided into blocks, so that many off-diagonal blocks of the matrix are low rank.
+H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank.
 For more details about H-matrices, see Appendix \ref{app:h_matrix}.
@@
 Our method will use these ellipsoid estimates to choose sets of impulse responses that can be computed in batches.
-Third, we describe an impulse response approximation that we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}).
+Third, we describe a method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}).
 Our method will interpolate impulse responses using this approximation.
@@
 
-Algorithm \ref{alg:varhpi_mean_cov} from Section \ref{sec:intromoments} allows us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} from Section \ref{eq:mean_and_covariance_estimation} uses $V$, $\mu$, and $\Sigma$ to construct an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
+Algorithm \ref{alg:varhpi_mean_cov} from Section \ref{sec:intromoments} allows us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} from Section \ref{eq:mean_and_covariance_estimation} uses $V$, $\mu$, and $\Sigma$ to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
 This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}).

### 2022-03-17 paper/localpsf.tex
restructuring paper

@@
 
-\subsection{Motivation: inverse problems with highly informative data}
+\subsection{Motivation: PDE constrained inverse problems}
 \label{sec:PDE_hessian_motivation}
@@
 Product-convolution approximations have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}.
-For background, we recommend reading the following papers: \cite{PRODCONVGOOD}.
+Often product-convolution methods are known as ``point spread function'' (PSF) methods, particularly in applied fields such as optics and image processing.
+There, the ``operator'' $\Aop$ is a physical process of distortion and blurring that occurs as rays of light pass through an optical system like a telescope or microscope.
+However, ``PSF'' is also sometimes used as an umbrella term to refer to other approximation methods that focus on interpolating impulse responses or impulse response features, including CUR-based low rank approximation \cite{NAGYPSFCUR}.
+For background on product-convolution and PSF methods, we recommend reading the following papers: \cite{PRODCONVGOOD}.
+
 We specifically build upon the class of product convolution approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i$ \cite{PRODCONVLIST}.
 A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, splines \cite{PRODCONVSPLINES}, or radial basis functions \cite{PRODCONVRBF}.
-With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
+With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate approximation.
 In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
-But even with adaptive refinement, many matrix vector products with $\Aop$ are required.
-In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\phi_i$ are computed with each matrix vector product.
-
-Our method of estimating the support of the impulse responses was inspired by resolution analysis in seismic imaging \cite{RESOLUTION}.
-In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$.
-If the width of $\phi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
-In \cite{RESOLUTION}, the width of $\phi_p$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
+But even with adaptive refinement, many operator actions of $\Aop$ are required.
+
+In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\phi_{x_i}$ are computed with each operator action of $\Aop$.
+This requires us to estimate the supports of the impulse responses, $\phi_x$, \emph{before} we compute them.
+We were inspired by resolution analysis in seismic imaging \cite{RESOLUTION}.
+In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_x$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $x$.
+If the width of $\phi_x$ is ten meters, then features of the parameter near $x$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
+In \cite{RESOLUTION}, the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
 In this paper, rather than probing $\Aop^T$ with random noise functions, we use polynomial functions.
-Our method estimates the support of $\phi_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
+Our method estimates the support of $\phi_x$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
 
@@
 
+MOMENTS PERMIT IMPROVED INTERPOLATION.
+IMPROVED INTERPOLATION REQUIRES H-MATRIX.
+H-MATRIX WITH IMPROVED INTERPOLATION ALSO SOLVES BOUNDARY ARTIFACT PROBLEM.
+ALSO DOESNT REQUIRE REGULAR GRID.
+ALSO SCALABLE TO LARGE NUMBER OF IMPULSE RESPONSES.
+FFT HAS FAST MATVEC BUT NOT OTHER OPTIONS.
+
 Conventional H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
@@
 
-In this section we present the three main concepts that our method is based on.
+In this section we present three main concepts that our method is based on.
 The method itself will be presented in Section \ref{sec:method}.
@@
 
-The quantity $\tau$ is a parameter that must be chosen by the user of the algorithm.
+The quantity $\tau$ is a parameter that must be chosen by the user.
 The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$.

### 2022-03-18 paper/localpsf.tex
restructuring paper

@@
 %In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
-Our work builds upon a history of ``product-convolution`` approximations---weighted sums of convolution operators with spatially varying weights.
+Product-convolution approximations are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
 A single convolution operator may be either high rank or low rank, so product-convolution approximations can (where applicable) approximate high rank operators with a small number of convolution operators.
-Product-convolution approximations have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}.
+Our work, while not exactly a product-convolution approximation, builds upon a history of product-convolution approximations, which have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}.
 Often product-convolution methods are known as ``point spread function'' (PSF) methods, particularly in applied fields such as optics and image processing.
@@
 
-We specifically build upon the class of product convolution approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i$ \cite{PRODCONVLIST}.
+We specifically build upon a class of product convolution approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i$ \cite{PRODCONVLIST}.
 A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, splines \cite{PRODCONVSPLINES}, or radial basis functions \cite{PRODCONVRBF}.
@@
 In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
-But even with adaptive refinement, many operator actions of $\Aop$ are required.
+But even with adaptive refinement, many points $x_i$, and hence operator actions of $\Aop$, are required.
 
@@
 This requires us to estimate the supports of the impulse responses, $\phi_x$, \emph{before} we compute them.
-We were inspired by resolution analysis in seismic imaging \cite{RESOLUTION}.
+We were inspired to do this by techniques from resolution analysis in seismic imaging \cite{RESOLUTION}.
 In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_x$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $x$.
 If the width of $\phi_x$ is ten meters, then features of the parameter near $x$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
-In \cite{RESOLUTION}, the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion.
-In this paper, rather than probing $\Aop^T$ with random noise functions, we use polynomial functions.
-Our method estimates the support of $\phi_x$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.
-
-Matrix-vector products with product-convolution approximations can be performed using the fast Fourier transform if the problem is discretized on a regular grid \cite{PRODCONVFFT}.
-The product-convolution approximation can then replace the original operator when using Krylov methods to solve linear systems with the original operator as the coefficient operator \cite{PRODCONVKRYLOV}.
-However, for complex geometries, problems are rarely discretized using regular grids, so we cannot easily use the fast Fourier transform.
-Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}.
+In \cite{RESOLUTION}, the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion.
+In this paper, rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Sections \ref{sec:intromoments} and \ref{eq:mean_and_covariance_estimation}).
+Our method estimates the support of $\phi_x$ more accurately and reliably than the techniques used in resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the techniques from resolution analysis do not have this requirement.
+%(and moreover, the seismic Hessian does not have a non-negative integral kernel).
+
+Convolutions can be performed using the fast Fourier transform (FFT) \cite{PRODCONVFFT}, allowing product-convolution approximations to be used as a fast surrogates for $\Aop$.
+But often it is necessary to perform further linear algebra operations involving $\Aop$, such as solving linear systems with $\Aop$ as the coefficient operator.
+In this case, one can perform the necessary linear algebra operations using Krylov methods that only use matrix-vector products, replacing $\Aop$ with its product-convolution approximation \cite{PRODCONVKRYLOV}.
+The Krylov method is accelerated because performing matrix vector products with the product-convolution approximation is typically faster than performing operator actions with $\Aop$.
+However, there are several problems with this approach.
+First, the number of Krylov iterations can be large, particularly when $\Aop$ is high-rank.
+Second, complex geometries problems are rarely discretized using regular grids, but the FFT can only be used for functions that are sampled on a regular grid.
+Hence, using the FFT to perform convolutions requires transferring discretized functions from an irregular mesh to a regular grid and back.
+Third, the number of FFTs that must be performed is proportional to the number of convolution kernels.
+If a large number of convolution kernels are used, this approach can be slow.
+%Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}. 
 Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are amenable to fast linear algebra operations.
-Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}.
+Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}, because operators that are amenable to product-convolution approximation are often sparse in the wavelet domain.
 Here we follow our previous work \cite{PRODCONVMYPAPER}, in which we converted product-convolution approximations into H-matrix format.
+
+Since our end goal is constructing a H-matrix, we are freed from strict adherence to the product-convolution format.
+This allows us to modify the product-convolution approach to address well-known deficiencies.
+In particular, product-convolution approximations are based on an assumption of local translation invariance, while our method is based on an assumption that we call ``normalized local mean displacement invariance,'' which is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
+Also, product-convolution approximations have a well-known issues near boundaries called boundary artifacts, which are caused by the boundary artifically ``cutting off'' impulse responses.
+The product-convolution method requires one to extend the impulse responses outside of the domain where they are undefined.
+Various extension techniques have been used, such as extension by zero, and reflection across the boundary, yet no known extension technique is fully satisfactory \cite{BOUNDARYETENSION}.
+In our method we do not have an issue with boundary artifacts because we temporarily exclude impulse responses from the approximation procedure whenever the approximation procedure attempts to evaluate the impulse response at a location where it is undefined.
+In exchange for these benefits, our approximation can no longer be written as a weighted sum of convolution operators, so we cannot apply our approximation to vectors via the FFT.
+But this is fine because our approximation is an H-matrix, which can be applied to vectors even faster.
 

### 2022-03-18 paper/localpsf.tex
restructuring paper

@@
 This allows us to modify the product-convolution approach to address well-known deficiencies.
-In particular, product-convolution approximations are based on an assumption of local translation invariance, while our method is based on an assumption that we call ``normalized local mean displacement invariance,'' which is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
+Product-convolution approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, while our method is based on an assumption that we call ``normalized local mean displacement invariance,'' which is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
 Also, product-convolution approximations have a well-known issues near boundaries called boundary artifacts, which are caused by the boundary artifically ``cutting off'' impulse responses.

### 2022-03-18 paper/localpsf.tex
restructuring paper

@@
 
-MOMENTS PERMIT IMPROVED INTERPOLATION.
+%MOMENTS PERMIT IMPROVED INTERPOLATION.
 IMPROVED INTERPOLATION REQUIRES H-MATRIX.
@@
 
-Conventional H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
-There are matrix-free methods for H-matrix approximation \cite{LEXINGPEELINGPROCESS}, and these methods have been used to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
-While these matrix-free H-matrix construction methods are asymptotically scalable in theory, the required number of matrix vector products can be large in practice.
+Since we are building an H-matrix approximation of $\Aop$, it is natural to ask why we need to go through an elaborate process to compute impulse responses and interpolate them.
+Can we build the H-matrix directly?
+In short, no, not efficiently.
+The reason is that classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
+Our impulse response interpolation lets us quickly compute approximations to matrix entries, which allows us to use these classical H-matrix construction methods.
+There are newer matrix-free methods for H-matrix approximation based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}, and these methods have been used to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
+While these matrix-free H-matrix construction methods are asymptotically scalable in theory, from a practical point of view the required number of matrix vector products is unreasonably large.
+Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make these methods more feasible in the future.
 

### 2022-03-18 paper/localpsf.tex
restructuring paper

@@
 The most popular matrix-free operator approximation methods are based on low rank approximation \cite{CCGOPAPERS}.
-Conventionally, either the Lanczos method, or the randomized singular value decomposition \cite{HMTRANDOM}, or CUR approximation/skeletonization \cite{CUR,MAHONEY} are used to perform the low rank approximation using only matrix-vector products.
-However, methods based on low rank approximation are not suitable for high rank operators, and many operators of practical interest are high rank.
-In the context of Hessians, methods based on low rank approximation suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of matrix-vector products are required to form a low rank approximation of the Hessian.
-The ideal scenario from a scientific perspective (highly informative data) is therefore the worst case scenario from a computational perspective (large computational cost) \cite{MYDISSERTATION}.
+This includes the Lanczos method, the randomized singular value decomposition \cite{HMTRANDOM}, and CUR approximation/skeletonization \cite{CUR,MAHONEY}.
+However, low rank approximation is not suitable for high rank operators, and many operators of practical interest are high rank.
+In the context of Hessians, methods based on low rank approximation suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator actions are required to form an accurate ``low rank'' approximation \cite{MYDISSERTATION}.
+%The ideal scenario from a scientific perspective (highly informative data) is therefore the worst case scenario from a computational perspective (large computational cost) \cite{MYDISSERTATION}.
 
 %In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
-Product-convolution approximations are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
-A single convolution operator may be either high rank or low rank, so product-convolution approximations can (where applicable) approximate high rank operators with a small number of convolution operators.
-Our work, while not exactly a product-convolution approximation, builds upon a history of product-convolution approximations, which have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}.
-Often product-convolution methods are known as ``point spread function'' (PSF) methods, particularly in applied fields such as optics and image processing.
-There, the ``operator'' $\Aop$ is a physical process of distortion and blurring that occurs as rays of light pass through an optical system like a telescope or microscope.
-However, ``PSF'' is also sometimes used as an umbrella term to refer to other approximation methods that focus on interpolating impulse responses or impulse response features, including CUR-based low rank approximation \cite{NAGYPSFCUR}.
-For background on product-convolution and PSF methods, we recommend reading the following papers: \cite{PRODCONVGOOD}.
-
-We specifically build upon a class of product convolution approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i$ \cite{PRODCONVLIST}.
-A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, splines \cite{PRODCONVSPLINES}, or radial basis functions \cite{PRODCONVRBF}.
-With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate approximation.
+Product-convolution (PC) approximations are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
+A single convolution operator may be high rank, so PC approximations can (where applicable) approximate high rank operators efficiently.
+PC approximations have been used in a variety of fields going back several decades \cite{PRODCONVLIST}.
+Often PC methods are known as ``point spread function'' (PSF) methods, particularly in optics and image processing.
+There, the ``operator'' $\Aop$ is a physical process of distortion and blurring of light rays passing through an optical system such as a telescope or microscope.
+However, ``PSF'' is sometimes used as an umbrella term to refer to other approximation methods that focus on interpolating impulse responses or impulse response features.
+%, including CUR-based low rank approximation \cite{NAGYPSFCUR}.
+For background on PC and PSF methods, we recommend the following papers: \cite{PRODCONVGOOD}.
+
+Our method is not a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i \in \Omega$ \cite{PRODCONVLIST}.
+A popular choice for PC methods in this class is to choose the $x_i$ to be nodes in a regular grid \cite{NAGY}.
+%, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, splines \cite{PRODCONVSPLINES}, or radial basis functions \cite{PRODCONVRBF}. 
+With a regular grid, a large number of $x_i$ are typically required to achieve an accurate approximation.
 In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
-But even with adaptive refinement, many points $x_i$, and hence operator actions of $\Aop$, are required.
-
-In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\phi_{x_i}$ are computed with each operator action of $\Aop$.
-This requires us to estimate the supports of the impulse responses, $\phi_x$, \emph{before} we compute them.
-We were inspired to do this by techniques from resolution analysis in seismic imaging \cite{RESOLUTION}.
+But even with adaptive refinement, many $x_i$, and hence operator actions of $\Aop$, are required.
+
+In this paper, 
+%rather than using adaptive refinement, 
+we reduce the computational cost by picking points $x_i$ in a way that allows many $\phi_{x_i}$ to be computed with each operator action of $\Aop$.
+This requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
+We were inspired by techniques from resolution analysis in seismic imaging \cite{RESOLUTION}.
 In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_x$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $x$.
@@
 In \cite{RESOLUTION}, the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion.
-In this paper, rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Sections \ref{sec:intromoments} and \ref{eq:mean_and_covariance_estimation}).
-Our method estimates the support of $\phi_x$ more accurately and reliably than the techniques used in resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the techniques from resolution analysis do not have this requirement.
+Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Sections \ref{sec:intromoments} and \ref{eq:mean_and_covariance_estimation}).
+Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing, but our method requires that $\Aop$ has a positive integral kernel, while random noise probing does not have this requirement.
 %(and moreover, the seismic Hessian does not have a non-negative integral kernel).
 
-Convolutions can be performed using the fast Fourier transform (FFT) \cite{PRODCONVFFT}, allowing product-convolution approximations to be used as a fast surrogates for $\Aop$.
-But often it is necessary to perform further linear algebra operations involving $\Aop$, such as solving linear systems with $\Aop$ as the coefficient operator.
-In this case, one can perform the necessary linear algebra operations using Krylov methods that only use matrix-vector products, replacing $\Aop$ with its product-convolution approximation \cite{PRODCONVKRYLOV}.
+Convolutions can be performed using the fast Fourier transform (FFT) \cite{PRODCONVFFT}, which allows PC approximations to be used as a fast surrogates for $\Aop$.
+Often it is necessary to perform further linear algebra operations involving $\Aop$, such as solving linear systems with $\Aop$ as the coefficient operator.
+Conventionally, one performs these linear algebra operations using Krylov methods, replacing $\Aop$ with its PC approximation \cite{PRODCONVKRYLOV}.
 The Krylov method is accelerated because performing matrix vector products with the product-convolution approximation is typically faster than performing operator actions with $\Aop$.
@@
 First, the number of Krylov iterations can be large, particularly when $\Aop$ is high-rank.
-Second, complex geometries problems are rarely discretized using regular grids, but the FFT can only be used for functions that are sampled on a regular grid.
-Hence, using the FFT to perform convolutions requires transferring discretized functions from an irregular mesh to a regular grid and back.
-Third, the number of FFTs that must be performed is proportional to the number of convolution kernels.
-If a large number of convolution kernels are used, this approach can be slow.
+Second, complicated geometries are rarely discretized using regular grids, so using the FFT to perform convolutions requires transferring discretized functions back and forth between the irregular mesh and a regular grid.
+Third, the computational cost grows with the number of convolution kernels used.
+%If a large number of convolution kernels are used, this approach can be slow.
 %Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}. 
-Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are amenable to fast linear algebra operations.
-Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}, because operators that are amenable to product-convolution approximation are often sparse in the wavelet domain.
-Here we follow our previous work \cite{PRODCONVMYPAPER}, in which we converted product-convolution approximations into H-matrix format.
-
-Since our end goal is constructing a H-matrix, we are freed from strict adherence to the product-convolution format.
-This allows us to modify the product-convolution approach to address well-known deficiencies.
-Product-convolution approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, while our method is based on an assumption that we call ``normalized local mean displacement invariance,'' which is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
-Also, product-convolution approximations have a well-known issues near boundaries called boundary artifacts, which are caused by the boundary artifically ``cutting off'' impulse responses.
-The product-convolution method requires one to extend the impulse responses outside of the domain where they are undefined.
-Various extension techniques have been used, such as extension by zero, and reflection across the boundary, yet no known extension technique is fully satisfactory \cite{BOUNDARYETENSION}.
+Thus, 
+%after constructing the product-convolution approximation, 
+it is desirable to convert the PC approximation to other matrix formats that are amenable to faster linear algebra operations.
+Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}, because operators that are amenable to PC approximation are often sparse in the wavelet domain.
+It is straightforward to show that convolution-product approximations (transposes of PC approximations) are equivalent to low rank approximation of the pseudo-differential symbol of the operator, so one could perform linear algebra operations involving PC approximations using fast methods for pseudo-differential operator symbol calculus \cite{YINGDEMANET}.
+Here, we follow our previous work \cite{PRODCONVMYPAPER}, in which we converted PC approximations into H-matrix format.
+
+Since our end goal is constructing a H-matrix, we are freed from strict adherence to the PC format.
+This allows us to modify the PC approach to address well-known deficiencies.
+PC approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, while our method is based on an assumption that we call ``normalized local mean displacement invariance,'' which is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
+Also, PC approximations have well-known issues near boundaries called boundary artifacts, which are caused by the boundary artifically ``cutting off'' impulse responses.
+One must extend the impulse responses outside of the domain where they are undefined, but
+%. Various extension techniques have been used, such as extension by zero, and reflection across the boundary, yet 
+no known extension technique is fully satisfactory \cite{BOUNDARYETENSION}.
 In our method we do not have an issue with boundary artifacts because we temporarily exclude impulse responses from the approximation procedure whenever the approximation procedure attempts to evaluate the impulse response at a location where it is undefined.
@@
 
-Product-convolution type approximations of Hessians have been used in a seismic inverse problem in \cite{GEORGSEISMICPRODCONV}, and in an advection-diffusion inverse problem in \cite{PRODCONVMYPAPER}.
+%Product-convolution type approximations of Hessians have been used in a seismic inverse problem in \cite{GEORGSEISMICPRODCONV}, and in an advection-diffusion inverse problem in \cite{PRODCONVMYPAPER}.
 In \cite{DEMANETSEISMIC}, matrix probing \cite{MATRIXPROBING} is used to approximate the Hessian in a seismic inverse problem as the sum of simple pseudodifferential operators.

### 2022-03-18 paper/localpsf.tex
restructuring paper

@@
 Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}, because operators that are amenable to PC approximation are often sparse in the wavelet domain.
-It is straightforward to show that convolution-product approximations (transposes of PC approximations) are equivalent to low rank approximation of the pseudo-differential symbol of the operator, so one could perform linear algebra operations involving PC approximations using fast methods for pseudo-differential operator symbol calculus \cite{YINGDEMANET}.
+Convolution-product approximations (transposes of PC approximations) are equivalent to low rank approximation of the pseudo-differential symbol of the operator, so one could perform linear algebra operations involving PC approximations using fast methods for pseudo-differential operator symbol calculus \cite{YINGDEMANET}.
 Here, we follow our previous work \cite{PRODCONVMYPAPER}, in which we converted PC approximations into H-matrix format.
 
-Since our end goal is constructing a H-matrix, we are freed from strict adherence to the PC format.
+Since our end goal is constructing an H-matrix, we are freed from strict adherence to the PC format.
 This allows us to modify the PC approach to address well-known deficiencies.
 PC approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, while our method is based on an assumption that we call ``normalized local mean displacement invariance,'' which is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
-Also, PC approximations have well-known issues near boundaries called boundary artifacts, which are caused by the boundary artifically ``cutting off'' impulse responses.
+PC approximations have well-known issues near boundaries called boundary artifacts, which are caused by the boundary artifically ``cutting off'' impulse responses.
 One must extend the impulse responses outside of the domain where they are undefined, but
@@
 In our method we do not have an issue with boundary artifacts because we temporarily exclude impulse responses from the approximation procedure whenever the approximation procedure attempts to evaluate the impulse response at a location where it is undefined.
-In exchange for these benefits, our approximation can no longer be written as a weighted sum of convolution operators, so we cannot apply our approximation to vectors via the FFT.
-But this is fine because our approximation is an H-matrix, which can be applied to vectors even faster.
+Our approximation cannot be written as a weighted sum of convolution operators, so we cannot apply our approximation to vectors via the FFT.
+This is fine because our approximation is an H-matrix, which can be applied to vectors even faster.
 
@@
 
-Since we are building an H-matrix approximation of $\Aop$, it is natural to ask why we need to go through an elaborate process to compute impulse responses and interpolate them.
-Can we build the H-matrix directly?
-In short, no, not efficiently.
+Since we are building an H-matrix approximation of $\Aop$, it is natural to ask why we need to go through an elaborate process of computing impulse responses and interpolating them. 
+%Can we build the H-matrix directly? In short, no, not efficiently. 
 The reason is that classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
@@
 While these matrix-free H-matrix construction methods are asymptotically scalable in theory, from a practical point of view the required number of matrix vector products is unreasonably large.
-Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make these methods more feasible in the future.
+Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods more feasible in the future.
 

### 2022-03-18 paper/localpsf.tex
background section of paper is acceptable

@@
 Convolution-product approximations (transposes of PC approximations) are equivalent to low rank approximation of the pseudo-differential symbol of the operator, so one could perform linear algebra operations involving PC approximations using fast methods for pseudo-differential operator symbol calculus \cite{YINGDEMANET}.
+There are also other matrix-free operator approximation methods based on wavelets \cite{HERRMANN}, and pseudo-differential operators \cite{DEMANET}, that do not use PC approximation.
 Here, we follow our previous work \cite{PRODCONVMYPAPER}, in which we converted PC approximations into H-matrix format.
@@
 
-After discretization with finite elements\footnote{While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead, with minor modifications to our method.} (Section \ref{sec:finite_element_kernel}), $\Aop$ and $\Phi$ become matrices $\mathbf{A}$ and $\mathbf{\Phi}$, respectively.
+After discretization with finite elements\footnote{While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead, with minor modifications to our method.} (Section \ref{sec:finite_element_kernel}), $\Aop$ and $\Phi$ become large dense matrices $\mathbf{A}$ and $\mathbf{\Phi}$, respectively.
 Because $\Aop$ is only available matrix-free and $\Phi$ is only defined implicitly in terms of $\Aop$, the matrices $\mathbf{A}$ and $\mathbf{\Phi}$ are not easily accessible.

### 2022-03-18 paper/localpsf.tex
background section of paper is good

@@
 Convolution-product approximations (transposes of PC approximations) are equivalent to low rank approximation of the pseudo-differential symbol of the operator, so one could perform linear algebra operations involving PC approximations using fast methods for pseudo-differential operator symbol calculus \cite{YINGDEMANET}.
-There are also other matrix-free operator approximation methods based on wavelets \cite{HERRMANN}, and pseudo-differential operators \cite{DEMANET}, that do not use PC approximation.
+Other matrix-free operator approximation methods based on wavelets \cite{HERRMANN} and pseudo-differential operators \cite{DEMANET} do not use PC approximation.
 Here, we follow our previous work \cite{PRODCONVMYPAPER}, in which we converted PC approximations into H-matrix format.
@@
 PC approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, while our method is based on an assumption that we call ``normalized local mean displacement invariance,'' which is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
-PC approximations have well-known issues near boundaries called boundary artifacts, which are caused by the boundary artifically ``cutting off'' impulse responses.
-One must extend the impulse responses outside of the domain where they are undefined, but
+PC approximations have well-known issues near boundaries called boundary artifacts, which are caused by the boundary artificially ``cutting off'' impulse responses.
+%One must extend the impulse responses outside of the domain where they are undefined, but
 %. Various extension techniques have been used, such as extension by zero, and reflection across the boundary, yet 
-no known extension technique is fully satisfactory \cite{BOUNDARYETENSION}.
+%no known extension technique is fully satisfactory \cite{BOUNDARYETENSION}. 
 In our method we do not have an issue with boundary artifacts because we temporarily exclude impulse responses from the approximation procedure whenever the approximation procedure attempts to evaluate the impulse response at a location where it is undefined.
@@
 
-Since we are building an H-matrix approximation of $\Aop$, it is natural to ask why we need to go through an elaborate process of computing impulse responses and interpolating them. 
+Since we are building an H-matrix approximation of $\Aop$, it is natural to ask why we need to compute impulse responses and interpolate them. Why not build the H-matrix directly?
 %Can we build the H-matrix directly? In short, no, not efficiently. 
@@
 There are newer matrix-free methods for H-matrix approximation based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}, and these methods have been used to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
-While these matrix-free H-matrix construction methods are asymptotically scalable in theory, from a practical point of view the required number of matrix vector products is unreasonably large.
+While these matrix-free H-matrix construction methods are asymptotically scalable in theory, in practice the required number of matrix vector products is unreasonably large.
 Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods more feasible in the future.

### 2022-03-18 paper/localpsf.tex
edit abstract

@@
 Such operators arise, for example, as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained inverse problems.
-	H-matrix approximations are highly desirable because fast H-matrix methods can add, multiply, invert, and factorize H-matrices, compute matrix-vector products, solve linear systems, and perform other useful matrix computations.
+%	H-matrix approximations are highly desirable because 
+	Fast H-matrix methods can add, multiply, invert, and factorize H-matrices, compute matrix-vector products, solve linear systems, and perform other useful matrix computations. 
 %	The computational cost of these operations is $O(N \log^a N)$, where the matrix is $N\times N$, and $a \in \{1,2,3\}$. 
-	Classical H-matrix construction methods 
+%	Classical H-matrix construction methods 
 %	require evaluating the integral kernel $O(N \log N)$ times, but this is not feasible 
-	are infeasible for large-scale problems in which the operator is only accessible via matrix-vector products (``matrix-free''), because these methods require evaluating the integral kernel of the operator at a large number of points.
-We overcome this problem by computing impulse responses of the operator and its transpose at a collection of scattered points, then interpolating these known impulse responses to evaluate entries of the integral kernel at other points where the impulse response is not known.
+%	are infeasible for large-scale problems in which the operator is only accessible via matrix-vector products (``matrix-free''), because these methods require evaluating the integral kernel of the operator at a large number of points.
+We overcome this problem by c
+	Our method computes impulse responses of the operator at a collection of scattered points, then interpolates these known impulse responses to evaluate entries of the integral kernel at other points where the impulse response is not known.
 %	We improve upon classical product-convolution approximations by translating with respect the estimated mean of the impulse response, rather than the location of the impulse response point source. 
@@
 Numerical results demonstrate that our method substantially outperforms existing state-of-the-art Hessian approximation methods which are based on low-rank approximation.
-We are able to form high quality approximations of high rank operators using only a small number of operator-vector products.
+We are able to form high quality approximations of high rank operators using only a small number of operator actions products.
 \end{abstract}

### 2022-03-18 paper/localpsf.tex
preparing document to send out to friends for comments

@@
 
-Our method is not a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i \in \Omega$ \cite{PRODCONVLIST}.
+Our method is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i \in \Omega$ \cite{PRODCONVLIST}.
 A popular choice for PC methods in this class is to choose the $x_i$ to be nodes in a regular grid \cite{NAGY}.

### 2022-03-21 paper/localpsf.tex
about to remove old stuff from paper

@@
 In our method we do not have an issue with boundary artifacts because we temporarily exclude impulse responses from the approximation procedure whenever the approximation procedure attempts to evaluate the impulse response at a location where it is undefined.
-Our approximation cannot be written as a weighted sum of convolution operators, so we cannot apply our approximation to vectors via the FFT.
+In exchange for these benefits, our approximation cannot be written as a weighted sum of convolution operators, so we cannot apply our approximation to vectors via the FFT.
 This is fine because our approximation is an H-matrix, which can be applied to vectors even faster.

### 2022-03-21 paper/localpsf.tex
remove scaled impulse response widehat variable

@@
 	\combresponse_k := \left(\Aop \diraccomb_k^*\right)^*
-	=\sum_{x_i \in \pointbatch_k} \widehat{\impulseresponse}_{x_i}.
-\end{equation}
-The last equality in \eqref{eq:dirac_comb_H_action} follows from linearity and the definition of $\widehat{\impulseresponse}_{x_i}$ in \eqref{eq:scaled_impulse_response} and \eqref{eq:impulse_response_delta_action}.
+	=\sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
+\end{equation}
+The last equality in \eqref{eq:dirac_comb_H_action} follows from linearity and the definition of $\impulseresponse_{x_i}$ in \eqref{eq:impulse_response_delta_action}.
 %\begin{equation}
@@
 \label{eq:varphi_eval}
-	\impulseresponse_{x_i}(z) / V(x_i) = \widehat{\impulseresponse}_{x_i}(z) =
+	\impulseresponse_{x_i}(z) =
 	\begin{cases}
-		\combresponse_k(z), & z \in E_{x_i} \\
+		\combresponse_k(z) V(x_i), & z \in E_{x_i} \\
 		0, & \text{otherwise}

### 2022-03-21 paper/localpsf.tex
removing old stuff

@@
 Such operators arise, for example, as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained inverse problems.
-%	H-matrix approximations are highly desirable because 
 Fast H-matrix methods can add, multiply, invert, and factorize H-matrices, compute matrix-vector products, solve linear systems, and perform other useful matrix computations.
-%	The computational cost of these operations is $O(N \log^a N)$, where the matrix is $N\times N$, and $a \in \{1,2,3\}$. 
-%	Classical H-matrix construction methods 
-%	require evaluating the integral kernel $O(N \log N)$ times, but this is not feasible 
-%	are infeasible for large-scale problems in which the operator is only accessible via matrix-vector products (``matrix-free''), because these methods require evaluating the integral kernel of the operator at a large number of points.
-We overcome this problem by c
 Our method computes impulse responses of the operator at a collection of scattered points, then interpolates these known impulse responses to evaluate entries of the integral kernel at other points where the impulse response is not known.
-%	We improve upon classical product-convolution approximations by translating with respect the estimated mean of the impulse response, rather than the location of the impulse response point source. 
 Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources.
-%	We construct a-priori ellipsoid estimates of the supports of the impulse responses (i.e., support ellipsoids are estimated \emph{before} any impulse responses have been computed). 
 By solving an ellipsoid packing problem, we choose as many points as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
@@
 We seek to approximate operators $\Aop$ that are accessible \emph{matrix-free}.
-That is,
-%access to $\Aop$ is available through ``operator actions.'' That is,
-we have a black-box computational procedure through which we may evaluate the maps
+That is, we have a black-box computational procedure through which we may evaluate the maps
 \begin{equation*}
@@
 This property is called ``matrix-free'' because at the discrete level it means one can compute matrix-vector products with the matrix representation of $\Aop$, but cannot easily compute individual matrix entries.
-%After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\bm{A}$, that is typically too large to be built and stored.
-This property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\bm{A}$ and $\bm{A}^T$ with arbitrary vectors, but one cannot easily\footnote{One could compute $\bm{v} = \bm{A}\bm{e}_j$, where $\bm{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero, then extract $\bm{A}_{ij} = \bm{v}_i$.
-However, this process is wasteful because one computes the entire vector $\bm{v}$, then discards all but one component.} access matrix entries $\bm{A}_{ij}$.
-
-%For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
 
@@
 which results from keeping $x$ fixed and viewing $\Phi(y,x)$ as a function of $y$.
-%for a collection of points $x_i$ scattered throughout $\Omega$.
-%Further, let $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$,
-%\begin{equation}
-%\label{eq:impulse_response_defn}
-%\impulseresponse_x(y) := \Aker(y, x),
-%\end{equation} 
-%be the \emph{impulse response} of $\Aop$ at $x$. 
 The function $\impulseresponse_x$ is called the impulse response because straightforward analysis shows that it may be expressed as follows:
@@
 		\includegraphics[scale=0.33]{batch1.png}
-		%		\caption{Non-symmetric}
 	\end{subfigure}
@@
 		\includegraphics[scale=0.33]{batch12.png}
-		%		\caption{Non-symmetric}
 	\end{subfigure}
@@
 		\includegraphics[scale=0.33]{batch19.png}
-		%		\caption{Non-symmetric}
 	\end{subfigure}
-	\caption{Batches, $\eta_b$, of normalized impulse responses, $\phi_{x_i}/V(x_i)$, for the Stokes inverse problem Hessian.
+	\caption{Batches of normalized impulse responses for the Stokes inverse problem Hessian.
 Left: 2nd batch.
@@
 Brightness scale differs between the subfigures.
-Black stars indicate the sample point locations, $x_i$.
-Dashed gray ellipses indicate the support ellipsoids, $E_{x_i}$.
-The large circle is the boundary of the domain $\Omega$.
+Black stars indicate the sample point locations at which the point source is applied.
+Dashed gray ellipses indicate the estimated support ellipsoids.
+The large circle is the boundary of $\Omega$.
 Early batches tend to have lots of ellipsoids near the boundary, as compared to later batches.
@@
 
-%
-% Because we build a H-matrix, and 
-%
-%In this paper we form hierarchical matrix approximations, $\mathbf{A}_H$, of operators $\Aop$ that satisfy the following three properties:
-%\begin{description}
-%	\item[Locally supported impulse responses:] For all $x\in \Omega$, the support of $\impulseresponse_x$ is contained (or approximately contained) in a localized region.
-%	\item[Non-negative impulse responses:] We have
-%	\begin{equation*}
-%	\impulseresponse_x(y) \ge 0
-%	\end{equation*}
-%	for all (or approximately all) $(y,x) \in \Omega \times \Omega$.
-%%	\item[Matrix-free:] One cannot easily evaluate kernel entries $\Aker(y,x)$.
-Instead, access to $\Aop$ is available through ``operator actions.'' That is, we have a black-box computational procedure (which is typically expensive) through which we may evaluate the maps
-%%	\begin{equation*}
-%%	u \mapsto\Aop u \quad \text{and} \quad v \mapsto \Aop^Tv
-%%	\end{equation*} 
-%%	for arbitrary functions $u$ and $v$.
-After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\bm{A}$, that is typically too large to be built and stored.
-This property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\bm{A}$ and $\bm{A}^T$ with arbitrary vectors, but one cannot easily\footnote{One could compute $\bm{v} = \bm{A}\bm{e}_j$, where $\bm{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero, then extract $\bm{A}_{ij} = \bm{v}_i$.
-However, this process is wasteful because one computes the entire vector $\bm{v}$, then discards all but one component.} access matrix entries $\bm{A}_{ij}$.
-In applications, applying $\bm{A}$ to a vector may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
-%%	%For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
-%\end{description} 
-%Once the hierarchical matrix approximation $\mathbf{A}_H$ is constructed, one may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
-These H-matrix methods are fast and scale well to large problems.
-WHY DO THESE OPERATIONS: KRYLOV SOLVERS, SAMPLING STRATEGIES, CAN'T DO WITH A.
-%
-%% approximate operators which possess the following properties:
-%% 
-%%We further limit our focus to operators that have
-%%\textbf{Local support}
-%%We say that $\Aop$ has local support if, for all $x\in \Omega$, the support of $\impulseresponse_x$ is contained (or approximately contained) in a small region. 
-%%The smaller these supports are, the better our algorithm will perform.
-Let $\delta_x$ denote the delta distribution (point source) centered at $x$.
-From \eqref{eq:impulse_response_delta_action}, we see that this local support property means that the response of $\Aop$ to a point source at $x$ is zero (or small) at points $y$ that are far from $x$.
-%%\item[3) Local translation invariance:] We say that $\Aop$ is locally translation invariant if
-%%\begin{equation}
-%%	\label{eq:local_translation_invariance}
-%%	\Aker(y+h, x+h) \approx \Aker(y,x)
-%%\end{equation}
-%%when $h$ is not too large.
-One can imagine the kernel entry $\Aker(y,x)$ as representing the strength of an interaction between a source at point $x$ and a target at point $y$.
-Local translation invariance means that the interaction strength remains roughly the same if the source and target points are both translated by the same vector.
-This is illustrated in Figure FIG.
-%%\textbf{Non-negative impulse response:} We have
-%%\begin{equation*}
-%%\impulseresponse_x(y) \ge 0
-%%\end{equation*}
-%%for all $(y,x) \in \Omega \times \Omega$.
-
 \subsection{Motivation: PDE constrained inverse problems}
@@
 In the context of Hessians, methods based on low rank approximation suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator actions are required to form an accurate ``low rank'' approximation \cite{MYDISSERTATION}.
-%The ideal scenario from a scientific perspective (highly informative data) is therefore the worst case scenario from a computational perspective (large computational cost) \cite{MYDISSERTATION}.
-
-%In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
+
 Product-convolution (PC) approximations are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
@@
 However, ``PSF'' is sometimes used as an umbrella term to refer to other approximation methods that focus on interpolating impulse responses or impulse response features.
-%, including CUR-based low rank approximation \cite{NAGYPSFCUR}.
 For background on PC and PSF methods, we recommend the following papers: \cite{PRODCONVGOOD}.
@@
 A popular choice for PC methods in this class is to choose the $x_i$ to be nodes in a regular grid \cite{NAGY}.
-%, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, splines \cite{PRODCONVSPLINES}, or radial basis functions \cite{PRODCONVRBF}. 
 With a regular grid, a large number of $x_i$ are typically required to achieve an accurate approximation.
@@
 
-In this paper, 
-%rather than using adaptive refinement, 
-we reduce the computational cost by picking points $x_i$ in a way that allows many $\phi_{x_i}$ to be computed with each operator action of $\Aop$.
+In this paper, we reduce the computational cost by picking points $x_i$ in a way that allows many $\phi_{x_i}$ to be computed with each operator action of $\Aop$.
 This requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
@@
 Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing, but our method requires that $\Aop$ has a positive integral kernel, while random noise probing does not have this requirement.
-%(and moreover, the seismic Hessian does not have a non-negative integral kernel).
 
@@
 Third, the computational cost grows with the number of convolution kernels used.
-%If a large number of convolution kernels are used, this approach can be slow.
-%Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}. 
-Thus, 
-%after constructing the product-convolution approximation, 
-it is desirable to convert the PC approximation to other matrix formats that are amenable to faster linear algebra operations.
+Thus, it is desirable to convert the PC approximation to other matrix formats that are amenable to faster linear algebra operations.
 Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}, because operators that are amenable to PC approximation are often sparse in the wavelet domain.
@@
 PC approximations have well-known issues near boundaries called boundary artifacts, which are caused by the boundary artificially ``cutting off'' impulse responses.
-%One must extend the impulse responses outside of the domain where they are undefined, but
-%. Various extension techniques have been used, such as extension by zero, and reflection across the boundary, yet 
-%no known extension technique is fully satisfactory \cite{BOUNDARYETENSION}. 
 In our method we do not have an issue with boundary artifacts because we temporarily exclude impulse responses from the approximation procedure whenever the approximation procedure attempts to evaluate the impulse response at a location where it is undefined.
@@
 
-%MOMENTS PERMIT IMPROVED INTERPOLATION.
-IMPROVED INTERPOLATION REQUIRES H-MATRIX.
-H-MATRIX WITH IMPROVED INTERPOLATION ALSO SOLVES BOUNDARY ARTIFACT PROBLEM.
-ALSO DOESNT REQUIRE REGULAR GRID.
-ALSO SCALABLE TO LARGE NUMBER OF IMPULSE RESPONSES.
-FFT HAS FAST MATVEC BUT NOT OTHER OPTIONS.
-
-Since we are building an H-matrix approximation of $\Aop$, it is natural to ask why we need to compute impulse responses and interpolate them. Why not build the H-matrix directly?
-%Can we build the H-matrix directly? In short, no, not efficiently. 
+Since we are building an H-matrix approximation of $\Aop$, it is natural to ask why we need to compute impulse responses and interpolate them.
+Why not build the H-matrix directly?
 The reason is that classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
@@
 Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods more feasible in the future.
-
-%Product-convolution type approximations of Hessians have been used in a seismic inverse problem in \cite{GEORGSEISMICPRODCONV}, and in an advection-diffusion inverse problem in \cite{PRODCONVMYPAPER}.
-In \cite{DEMANETSEISMIC}, matrix probing \cite{MATRIXPROBING} is used to approximate the Hessian in a seismic inverse problem as the sum of simple pseudodifferential operators.
-Although it is not explicitly mentioned, the approximation in \cite{DEMANETSEISMIC} could be interpreted as a convolution-product approximation (a product-convolution approximation of the transpose of the operator).
-%, which is like a product-convolution approximation, except the order of pointwise multiplications and convolutions is reversed in \eqref{eq:product_convolution}.
 
@@
 \end{equation*}
-denote the spatially varying scaling factor,
-%, let 
-%\begin{equation}
-%\label{eq:scaled_impulse_response}
-%\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)
-%\end{equation}
-%denote the normalized version of $\impulseresponse_x$. 
-and let $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d},$
+denote the spatially varying scaling factor, and let $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d},$
 \begin{equation*}
@@
 
-%The ellipsoid $E_x$ is not immediately available. 
-%(recall $\gdim$ is the spatial dimension of $\Omega$).
 The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}.
@@
 
-%We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Proposition \ref{thm:vol_mean_cov} and summarize the method for computing these moments in Algorithm \ref{alg:varhpi_mean_cov}.
-
-%\begin{prop}[Impulse response moments]
-%	\label{thm:vol_mean_cov}
-%	%	Let $\spatialvol:\Omega \rightarrow \mathbb{R}$, $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$, and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$, be the following functions:
-%	%	\begin{equation*}
-%	%	\spatialvol(x) := \int_{\Omega} \impulseresponse_x(y) dy, \qquad
-%	%	\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
-%	%	\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
-%	%	\end{equation*}
-%	%	where $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denotes the normalized version of $\impulseresponse_x$, $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.
-%	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$.
-Let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
-%	\begin{equation*}
-%	C(x) := 1, \qquad
-%	L^i(x) := x^i, \qquad
-%	Q^{ij}(x) = x^i x^j.
-%	\end{equation*}
-%	Let $f/g$ denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
-The following results hold:
-%	\begin{subequations}
-%		\label{eq:vol_mean_var}
-%		\begin{align}
-%		\spatialvol =& \left(\Aop^T C\right)^* \\
-%		\spatialmean^i =& \left(\Aop^T L^i\right)^* / \spatialvol \\
-%		\spatialcov^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j
-%		\end{align}
-%	\end{subequations}
-%	for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$.
-%\end{prop}
-
-%The proof of Proposition \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}. The proof is straightforward, and follows directly from the definitions.
-
 \subsection{Impulse response support ellipsoids}
 \label{eq:mean_and_covariance_estimation}
-
-%Because of the non-negative impulse response property, for each $x \in \Omega$, the impulse response $\impulseresponse_x$ is a non-negative function $y \mapsto \impulseresponse_x(y)$.
-Hence, $\impulseresponse_x$ is a scaled probability distribution.
-Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor (the ``volume'' of $\impulseresponse_x$), let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
 
@@
 The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.
-
-%To determine $E_x$, we must compute the scalar $\spatialvol(x)$, the length-$\gdim$ vector $\spatialmean(x)$, and the $\gdim \times \gdim$ matrix $\spatialcov(x)$.
-In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of polynomial functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).
 
@@
 \phi_x(y)/V(x) &\approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right) / V(x'). \label{eq:translate4}
-%\frac{\phi_x(y)}{V(x)} &\approx \frac{\phi_{x'}\left(y-\mu(x)+\mu(x')\right)}{V(x')}. \label{eq:translate4}
 \end{align}
 Approximation \eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close.
-%This approximation follows from the assumption that $\Phi$ is smooth. 
-Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank interpolative factorization/Nystrom approximation \cite{CUR,NYSTROM}.
-% of the form
-%\begin{equation}
-%	\label{eq:low_rank_interpolation}
-%	\Phi(y,x) \approx \sum_{i=1}^r w_i(x) \phi_{x_i}(y)
-%\end{equation}
-%where $\phi_{x_i}$ are known impulse responses at a collection of points $x_i$ scattered throughout $\Omega$, and $w_i$ are smooth weighting functions that interpolate the $\phi_{x_i}$. 
-%In problems of practical interest, \eqref{eq:translate1} typically holds \emph{asymptotically} in the limit as $x'$ approaches $x$. 
-However, 
-%even if $\Phi$ is smooth (which is typically the case asymptotically),
-%the relevant Lipschitz constant for continuity, and hence 
-the required rank in the low rank approximation can be large, which makes algorithms based on \eqref{eq:translate1} computationally expensive.
-This is the case for Hessians in PDE constrained inverse problems with highly informative data, as discussed in Section \ref{sec:PDE_hessian_motivation}.
-%As the data become more informative, typically the ``mass'' of $\Phi(y,x)$ concentrates nearer and nearer to the ``diagonal'' $y=x$, the Lipschitz constant increases, and the required rank in the approximation increases.
-%Moreover, if $\Aop$ is numerically low rank, then one would typically use more accurate and robust matrix-free methods for forming low-rank approximations, such as randomized SVD or Lanczos.
+Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank interpolative factorization/CUR approximation \cite{CUR,NYSTROM}.
+However, the required rank in the low rank approximation can be large, which makes algorithms based on \eqref{eq:translate1} expensive.
 
 Operators that satisfy \eqref{eq:translate2} are called ``locally translation-invariant'' because integral kernel entries $\Phi(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$.
-Locally translation-invariant operators act like convolutions locally, and can therefore be well-approximated by so-called product-convolution approximations---weighted sums of convolution operators with spatially varying weights.
-Product-convolution approximations can be high-rank even if the number of convolution operators used in the approximation is small.
-
-% Kernels for locally translation invariant operators can be well-approximated by decompositions of the form
-%\begin{equation}
-%	\Phi(y,x) \approx \sum_{i=1}^r w_i(x) \phi_{x_i}(y-x)
-%\end{equation}
-%which correspond to operator approximations of the form
-%\begin{equation}
-%\label{eq:product_convolution}
-%\left(\Aop~u\right)(v) \approx \left( v,~ \sum_{i=1}^\convrank \convkernel_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)}.
-%\end{equation}
-%Here $f \ast g$ denotes convolution, and $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$ (i.e., $\left(f \cdot g\right)(x) = f(x) g(x)$).
-Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution.
-Product-convolution approximations can be high-rank even if $r$ is small.
+It is straightforward to show that if equality holds in \eqref{eq:translate2}, then $\Aop$ is a convolution operator.
+Locally translation-invariant operators act like convolutions locally, and can therefore be well-approximated by PC approximations.
 
@@
 
-%Fourth, we discuss how discretization of continuous functions by finite elements leads to matrix discretizations $\mathbf{A}$ and $\mathbf{\Phi}$ of $\Aop$ and $\Phi$, respectively (Section \ref{sec:finite_element_kernel}).
-The primary function of our method is to compute H-matrix approximations of these matrices.
-%While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead with minor modifications.
-
 \subsection{Overview}
@@
 The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section \ref{sec:local_mean_displacement_invariance}.
-%Specifically, we make the assumption that
-%\begin{equation}
-%\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))
-%\end{equation}
-%if $x_i$ is close to $x$. This is shown in Figure \ref{fig:mean_displacement_invariance}. Motivation for this assumption is discussed in Section \ref{eq:local_mean_displacement_invariance}. 
 We approximate $\Phi(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{V(x)}{V(x_i)}\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for a collection of sample points $x_i$ near $x$.
@@
 
-If $\Aop$ is symmetric, we take advantage of symmetry to construct a more accurate approximation at negligible additional cost
-%. In this case, 
-%impulse responses of $\Aop$ are also impulse responses of $\Aop^T$, which may be interpolated in the variable $y$ instead of $x$ to yield a different approximation of $\Phi(y,x)$. 
-%we combine 
-by combining information from impulse responses of both $\Aop$ and $\Aop^T$ in the interpolation procedure (Section \ref{sec:symmetric_improvements}).
-% for approximating $\Phi(y,x)$, to get a more accurate approximation at negligible additional cost.
+If $\Aop$ is symmetric, we take advantage of symmetry to construct a more accurate approximation at negligible additional cost by combining information from impulse responses of both $\Aop$ and $\Aop^T$ in the interpolation procedure (Section \ref{sec:symmetric_improvements}).
 Intuitively, impulse responses of $\Aop$ can be thought of as ``columns'' of $\Phi$, impulse responses of $\Aop^T$ can be thought of as ``rows'' of $\Phi$, and $\Phi(y,x)$ can be thought of as an unknown entry of $\Phi$.
@@
 The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}. The cost to construct $\mathbf{A}_H$ is discussed in Section \ref{sec:overall_cost}.
-
-%Often we are interested in approximating operators $\Aop$ that are symmetric, or symmetric positive semi-definite.
-
-%Forming the H-matrix approximation $\boldsymbol{\Phi}_H$ requires evaluating kernel entries, $\Phi(y,x)$, at arbitrary points $(y,x)$, and this is done by interpolating the known impulse responses $\{\impulseresponse_{x_i}\}_{i=1}^r, \{\impulseresponse'_{y_i}\}_{i=1}^r$.
-%
-%%For each batch, we solve an ellipsoid packing problem to choose as many points $x_i$ as possible, while ensuring that the supports of the impulse responses $\impulseresponse_{x_i}$ do not overlap for points $x_i$ in the batch.
-Then we apply $\Aop$ to a weighted sum of delta distributions centered at the points $x_i$ (Dirac comb).
-From the resulting
-%
-%
-%Our overall strategy is as follows.
-First, we compute the functions $V:\Omega \rightarrow \mathbb{R}$, $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$ by applying $\Aop^T$ to a small number of polynomial functions.
-Next, we solve ellipsoid packing problems to choose batches of points $x_i$ such that
-%to compute impulse responses $\{\impulseresponse_{x_i}\}_{i=1}^r$ of $\Aop$ to point sources at a collection of points $x_i$ using a small number of operator actions of $\Aop$.
-If $\Aop$ is non-symmetric, we also compute impulse responses $\{\impulseresponse'_{y_i}\}_{i=1}^r$ of $\Aop^T$ to point sources at a collection of points $y_i$.
-We then form a H-matrix approximation, $\boldsymbol{\Phi}_H$, of the kernel, $\Phi$, and use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation of $\Aop$, denoted $\mathbf{A}_H$.
-%Schematically the overall framework proceeds as follows:
-%\begin{equation*}
-%\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \{\impulseresponse_{x_i}\}_{i=1}^r \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
-%\end{equation*}
-%We will define product-convolution approximations in Section \ref{sec:prodconv_intro}.
-H-matrices are discussed in Appendix \ref{app:h_matrix}.
-Our method is summarized in Section \ref{sec:overview_intro}, and described in detail in Section \ref{sec:method}.
-%
-%The novel contributions of our method are in the way that we form the product-convolution approximation ($\Aop \rightarrow \AopPc$).
-This is the part of the method where we use the four properties of $\Aop$ described above.
-Fast conversion of the discretized version of $\AopPc$ to H-matrix format ($\AopPc \rightarrow \AmatPc$) is possible because integral kernel entries for $\AopPc$ are easily accessible (unlike integral kernel entries for $\Aop$).
-Once in H-matrix format, one may use H-matrix methods to perform further useful linear algebra operations involving $\AmatPc$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
-These H-matrix methods are fast and scale well to large problems.
-
-%In this section we detail our method for computing the H-matrix approximation, $\mathbf{A}_H$, of $\Aop$, as outlined in Section \ref{sec:intro_overview}.
-%The supports of the impulse responses $\phi_x$ are estimated to be contained within ellipsoids (Section \ref{eq:mean_and_covariance_estimation}).
-The shapes of the ellipsoids depend on spatially varying volume, mean, and covariance functions, $V,\mu,\Sigma$, which we calculate by applying $\Aop$ to a small number of polynomial functions (Section \ref{sec:moments}).
-We use the ellipsoid support estimates in a greedy ellipsoid packing procedure to choose batches, $\pointbatch_b$, of sample points $x_i$ (Section \ref{sec:sample_point_selection}).
-Batches, $\eta_b$, of impulse responses, $\phi_{x_i}$, are computed by applying $\Aop$ to Dirac combs associated with the batches of sample points $S_b$.
-The impulse response batch $\eta_b$ contains the information about all impulse responses $\phi_{x_i}$ associated with points $x_i \in S_b$.
-Hence, once a small number of impulse response batches $\eta_b$ are computed, they can be used to perform pointwise evaluations $z \mapsto \phi_{x_i}(z)$ for a large number of impulse responses, $\phi_{x_i}$ (Section \ref{sec:get_impulse_response}).
-%
-%In Section \ref{eq:mean_and_covariance_estimation} we describe our ellipsoid estimate for the support of $\phi_x$.
-The ellipsoid shape depends on the functions $V,\mu,\Sigma$.
-We show how to compute these functions in Section \ref{sec:moments}.
-The ellipsoid support estimates are then used in a greedy ellipsoid packing procedure to choose batches, $\pointbatch_b$, of sample points $x_i$ (Section \ref{sec:sample_point_selection}).
-%This section is organized as follows:
-%\begin{description}
-%	\item[Section \ref{eq:mean_and_covariance_estimation}:] We describe our ellipsoid estimate for the support of $\phi_x$.
-%	\item[Section \ref{sec:moments}:] We show how to compute $V,\mu,\Sigma$.
-%	\item[Section \ref{sec:sample_point_selection}:] We present the greedy ellipsoid packing procedure used to choose batches, $\pointbatch_b$, of sample points $x_i$.
-%	\item[Section \ref{sec:get_impulse_response}:] We describe how to compute $\phi_{x_i}$'s, in batches, $\eta_b$.
-%	\item[Section \ref{sec:approximate_kernel_entries}:] We show how we interpolate $\phi_{x_i}$'s to approximate $\Phi(y,x)$.
-%	\item[Section \ref{sec:finite_element_kernel}:] We discuss finite element discretization of $\Phi$.
-%	\item[Section \ref{sec:h_matrix_construction_short}:] We describe how we construct the H-matrices $\mathbf{\Phi}_H$ and $\mathbf{A}_H$.
-%	\item[Section \ref{sec:overall_cost}:] We review the total computational cost to construct $\mathbf{A}_H$, and list the cost to perform further linear algebra tasks involving $\mathbf{A}_H$ using H-matrix methods.
-%\end{description}
-%The optional method for making $\mathbf{A}_H$ positive semi-definite is described in Appendix \ref{sec:make_spd}.
-The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}.
-
-%we compute the impulse response moments $V,\mu,\Sigma$ and use these moments to estimate the supports of the impulse responses $\phi_x$ (Section \ref{sec:moments}), the greedy ellipsoid packing procedure we use to choose batches of sample points $x_i$ (Section \ref{sec:sample_point_selection}), how we compute impulse responses, $\phi_{x_i}$, in batches (Section \ref{sec:get_impulse_response}), how we interpolate the $\phi_{x_i}$ to approximate integral kernel entries $\Phi(y,x)$ (Section \ref{sec:approximate_kernel_entries}), finite element discretization of $\Phi$ (Section \ref{sec:finite_element_kernel}), and how we construct the H-matrix approximation $\mathbf{A}_h$ of the operator $\mathcal{A}$.
-We also describe how to (optionally) modify $\mathbf{A}_H$ to force it to be symmetric positive semi-definite.
-%The complete algorithm is shown in Algorithm \ref{alg:construct_Atilde}.
 
@@
 
-%Further, let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
-That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
-
-%We define the finite element approximation of the kernel, $\Phi^h \approx \Phi$, as follows:
 The interpolation of $\Phi$ onto $V_h \otimes V_h$, which we denote by $\Phi^h$, is given as follows:
@@
 \end{equation}
-%The kernel $\Phi^h$ is the interpolation of $\Phi$ onto $V_h \otimes V_h$.
-Error in the overall kernel approximation, $\widetilde{\Phi}^h \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$.
 Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations (see Appendix \ref{app:proofs} for details), yields the following approximation:
@@
 \end{equation}
-%While $\Aop$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\Aop$ to functions in the finite element space. 
 where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
@@
 \end{equation*}
-%Matrix entries $\Phi(p_i, p_j)$ are expensive to compute because $\Aop$ is only available matrix-free.
-%Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, and \eqref{eq:Akerpcmat_entries}, the definition of the mass matrix, and performing algebraic manipulations, it is straightforward but tedious to show that
-%\begin{equation*}
-%	\left(\AopPcMesh u_h \right)(v_h) = \bm{v}^T \bm{M} \AkerPcMat \bm{M} \bm{u}.
-%\end{equation*}
 Our matrix representation, $\mathbf{A}$, of $\Aop$ is therefore given by
@@
 Note that $\mathbf{A}$ is not exactly the Galerkin projection of $\Aop$ onto $V_h$ because of the approximation $\Phi^h \approx \Phi$.
-%because of the approximation $\Phi^h \approx \Phi$. 
 Hence, $\mathbf{A}$ may differ slightly from the implicitly defined matrix that performs the mapping $\mathbf{u} \rightarrow \mathbf{v}$ corresponding to the discrete version of the operator action $v=\Aop u$.
-%It is computationally intractible to build and store the matrices $\mathbf{A}$ and $\mathbf{\Phi}$, and individual matrix entries $\mathbf{A}_{ij}$ and $\mathbf{\Phi}_{ij}$ are not easily accessible.
-In the remainder of this section, we show how to build accessible approximations of these matrices.
-
-%The discrete version of the operator action $v = \Aop u$ is a mapping of vectors $\mathbf{u} \rightarrow \mathbf{v}$, which is done through an implicit computational procedure.
-Typically the implicitly defined matrix that performs this mapping is the Galerkin projection of $\Aop$ onto $V_h$.
-
-%We neglect this error because it is the same magnitude as the error in finite element approximation of functions $u_h \approx u$. 
-%Note that the matrices $\mathbf{A}$ and $\mathbf{\Phi}$ are dense $N \times N$ matrix that only exist in theory.
-Building and storing these matrices is computationally intractible.
-Computing individual matrix entries $\mathbf{A}_{ij}$ and $\mathbf{\Phi}_{ij}$ is computationally expensive.
 
@@
 We choose sample points, $x_i$, in batches $\pointbatch_k$.
-%A greedy algorithm is used to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other.
-We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\impulseresponse_x$ (Section \ref{eq:mean_and_covariance_estimation}).
-The support of $\impulseresponse_x$ is estimated to be contained within an ellipsoid, so the process of choosing batches of points is an ellipsoid packing problem (Section \ref{sec:greedy_point_selection}).
 We use a greedy ellipsoid packing algorithm to choose as many points as possible per batch, while ensuring that there is no overlap between the support ellipsoids, $E_{x_i}$, associated with the sample points within a batch.
-%to choose batches of sample points, $\pointbatch_k$, such that there is no overlap between the ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch. 
-%These sample point batches will be used in Section \ref{sec:get_impulse_response} to compute many impulse responses of $\Aop$ using only a small number of operator actions of $\Aop$.
-The support of $\impulseresponse_x$ is (approximately) contained in the ellipsoid $E_x$, so by applying $\Aop$ to the Dirac comb associated with all sample points in a batch, we will recover the impulse responses for all sample points in that batch.
 
@@
 Then we select the candidate point $p \in P$ that is the farthest away from all points in previous sample point batches $S_1 \cup S_2 \cup \dots \cup S_{k-1}$.
-%That is, $p$ is a maximizer of the following optimization problem:
-%\begin{equation*}
-%\max_{p \in P} \min_{x \in \pointbatch_1 \cup \dots \cup \pointbatch_{k-1}} \|p - x\|.
-%\end{equation*}
 Candidate points for the first batch $S_1$ are chosen randomly from $P$.
@@
 The last equality in \eqref{eq:dirac_comb_H_action} follows from linearity and the definition of $\impulseresponse_{x_i}$ in \eqref{eq:impulse_response_delta_action}.
-%\begin{equation}
-%\label{eq:phi_b}
-%\combresponse_k = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
-%\end{equation}
-%\begin{equation}
-%	\label{eq:phi_b}
-%	\combresponse_k = \left(\Aop \left(\sum_{x_i \in \pointbatch_k} \delta_{x_i}/V(x_i)\right)^*\right)^* = \sum_{x_i \in \pointbatch_k} \left(\Aop \delta_{x_i}^*\right)^* / V(x_i) = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
-%\end{equation}
 Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\impulseresponse_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\impulseresponse_j$ do not overlap when $i \neq j$, we have (approximately)
@@
 
-%Note that the support of $\impulseresponse_{x_i}$ is estimated to be contained within the ellipsoid $E_{x_i}$, but this is just an approximation. 
-%%The impulse response $\impulseresponse_{x_i}$ may be nonzero inside the ellipsoid $E_{x_j}$ associated with another point $x_j$ in the batch, and this will cause error in our recovery of $\phi_{x_j}$.
-%Without this scaling, the neglected portion of an impulse response corresponding to a point $x_i$ could overwhelm the impulse response corresponding to a nearby point $x_j$ if $V(x_i)$ is much larger than $V(x_j)$.
-
 \subsection{Approximate integral kernel entries $\widetilde{\Phi}(y,x)$}
@@
 
-%Recall from Section \ref{eq:local_mean_displacement_invariance} that if $x'$ is near $x$, then
-%\begin{equation*}
-%\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x')\right).
-%\end{equation*}
-
-%In Section \ref{sec:intro} we noted that computing $\Aker\left(y,x\right)$ is computationally costly. 
-%This section describes how we compute approximate kernel entries
-%\begin{equation*}
-%\widetilde{\Phi}(y,x) \approx \Phi(y,x)
-%\end{equation*}
-%for arbitrary pairs of points $(y,x) \in \Omega \times \Omega$. 
 Given $(y,x)\in \Omega \times \Omega$, let
@@
 \end{equation}
-%for arbitrary pairs of points $(y,x) \subset \Omega \times \Omega$.
-% while Section \ref{sec:approximate_kernel_entries} we showed how we can cheaply evaluate approximate kernel entries at arbitrary pairs of points $(y,x) \in \Omega \times \Omega$.
 Note that $f_i$ is well-defined, because $\phi_{x_i}\left(z_i\right)$ is defined if and only if $z_i \in \Omega$, and $V(x_i)>0$ by our sample point picking procedure in Section \ref{sec:sample_point_selection}.
@@
 We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
-%Since $\Phi(y,x) \approx f_i$ when $x_i$ is near $x$, we approximate $\Phi(y,x)$ by interpolating the (point,value) pairs
-%\begin{equation}
-%\left(x_i, f_i\right), \qquad i=1,\dots,k,
-%\end{equation}
-%at the point $x$.
-To see that $\Phi(y,x) \approx f_i$, recall from Section \ref{eq:local_mean_displacement_invariance} that $\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ if $x_i$ is near $x$.
-Inclusion of the extra factor $\frac{V(x)}{V(x_i)}$ in \eqref{eq:fxyxp} is a form of preconditioning that makes this approximation more robust in the case where $V(x)$ varies widely.
-%Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section \ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $V(x_i)$ is smaller than a predetermined threshold.
-%\subsubsection{Radial basis function interpolation}
 
 Interpolation is performed using the following radial basis function scheme:
-%The approximation $\widetilde{\Phi}(y,x)$ is the value of this interpolation at $x$.
-%In numerical experiments we use $k=10$. The $k$ nearest sample points are found by querying a precomputed KD-tree \cite{kdtree} of all sample points. 
-%
-%Let
-%It follows from inspection that $\Phi(y,x) = f_{y,x}(x)$,
-%and from the discussion in Section \ref{eq:local_mean_displacement_invariance}, we expect 
-%$
-%\Phi(y,x) \approx f_{y,x}(x')
-%$
-%when $x'$ is close to $x$.
-The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section \ref{eq:local_mean_displacement_invariance}, is a form of preconditioning that improves the quality of this approximation in situations where $V(x)$ varies rapidly.
-%%, as is commonly the case for many operators of practical interest. In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.
-%%As discussed in Section SECTION, we make the approximation
-%%\begin{equation*}
-%%	\Phi(y,x) = \phi_x(y) \approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right)
-%%\end{equation*}
-%%if $x'$ is close to $x$. Hence, w
-%
-%We evaluate $\widetilde{\Phi}(y,x)$ by performing radial basis function interpolation of the function $f_{y,x}(x')$ in the variable $x'$.
-The data used in the interpolation are the following (point,value) pairs
-%\begin{equation}
-%\left(x_i, f_{y,x}(x_i)\right), \qquad i=1,\dots,k,
-%\end{equation}
-%and $(x, f_{y,x}(x)) = (x,\Phi(y,x))$ is the (point,value) pair we approximate via interpolation.
-%Here $\{x_i\}_{i=1}^k \subset \Omega$ are the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined.
-%%a collection of sample points at which the impulse response $\phi_{x_i}$ is known.
-%%. Specifically, we interpolate $(\text{point}, \text{value})$ pairs $(x_i, f_i)$ at the point $x$, where $x_i$ are sample points, and $f_i$ are given as follows:
-%%\begin{equation}
-%%\label{eq:phi_mu_shift32}
-%%	f_i := \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
-%%\end{equation}
-%%We must perform a new interpolation every time we evaluate $\Phi(y,x)$ at a new pair of points $(y,x)$, and constructing the H-matrix approximation of $\Phi$ requires evaluating $\Phi(y,x)$ at a large number of pairs of points $(y,x)$.
-%%Interpolation using all sample points $x_i$ is too expensive, so we only use the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined. 
-%In numerical experiments we use $k=10$. The $k$ nearest sample points are found by querying a precomputed KD-tree \cite{kdtree} of all sample points. 
-%
-%From the definition of $f_{y,x}$ in \eqref{eq:fxyxp}, we see that $f_{y,x}(x_i)$ is only defined if $y - \mu(x) + \mu(x_i) \in \Omega$, because $\phi_x$ is only defined on $\Omega$.
-But it may happen that $y - \mu(x) + \mu(x_i) \notin \Omega$, even if $y,x,x_i \in \Omega$.
-In this case, the point $x_i$ is considered invalid, and not used in the interpolation, even if it is close to $x$.
-We check whether $x_i$ is valid (i.e., whether $y - \mu(x) + \mu(x_i) \in \Omega$) by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
-%\subsubsection{Radial basis function interpolation details}
-%Having found the $k$ nearest valid sample points $x_i$ to $x$, we approximate $\Phi(y,x)$ using radial basis function interpolation of the following form:
-%In detail, we form $\widetilde{\Phi}(y,x) \approx \Phi(y,x)$ as follows:
 \begin{equation}
@@
 $
-and $f \in \mathbb{R}^k$ has entries $f_i$ from \eqref{eq:fxyxp}.
-% has entries 
-%\begin{equation*}
-%f_i = f_{y,x}(x_i), \qquad i=1,\dots,k.
-%\end{equation*}
-We use polyharmonic spline radial basis functions \cite{polyspline}:
+and $f \in \mathbb{R}^k$ has entries $f_i$ from \eqref{eq:fxyxp}. We use polyharmonic spline radial basis functions \cite{polyspline}:
 \begin{equation}
@@
 For more details on radial basis function interpolation and polyharmonic splines, see \cite{rbf}.
-%\begin{equation}
-%	f_i = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
-%\end{equation}
-%The result of this interpolation is to approximate $\Phi(y,x)$ with a linear combination of the shifted impulse response values $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$. 
-%By including the factor $\frac{V(x)}{V(x_i)}$, we are implicitly interpolating the ``volume preconditioned'' function $\frac{1}{V(x)}\Phi(y,x)$, then rescaling the result by $V(x)$ to get an approximation of $\Phi(y,x)$.
-Volume preconditioning improves the approximation considerably when $V(x)$ varies widely in magnitude, as is commonly the case for many operators of practical interest.
-In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.
-
-%In order to compute $f_{y,x}(x_i)$, we must evaluate $\impulseresponse_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$, and we do this by evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$.
-Here $b$ is the index of the batch that contains the sample point $x_i$, and $\eta_b$ is the corresponding impulse response batch function that we have computed and stored.
-%Although it is typically recommended that one avoid evaluating finite element functions at arbitrary points [CITE], here we have no choice.
-
-%\subsubsection{Evaluating $f_{y,x}(x_i)$}
-
-To evaluate $f_i$, we check whether $z_i \in E_{x_i}$ using \eqref{eq:support_ellipsoid}. If $z_i \notin E_{x_i}$, then
-% by checking if
-%\begin{equation}
-%\label{eq:ellipsoid_check93}
-%(y - \mu(x_i))^T \Sigma^{-1}(x_i) (y - \mu(x_i))  \le \tau^2.
-%\end{equation}
-%If \eqref{eq:ellipsoid_check93} does not hold, 
-then $z_i$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_i=0$.
-% If \eqref{eq:ellipsoid_check93} does hold, then 
+
+To evaluate $f_i$, we check whether $z_i \in E_{x_i}$ using \eqref{eq:support_ellipsoid}.
+If $z_i \notin E_{x_i}$, then $z_i$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_i=0$.
 If $z_i \in E_{x_i}$, we look up the batch index $b$ such that $x_i \in S_b$, and evaluate $f_i$ via the formula
@@
 per \eqref{eq:varphi_eval}.
-%Recall from Section \ref{sec:get_impulse_response} that $\eta_b$ is the impulse response batch that we have computed and stored.
-%Evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$ requires care, because
 
@@
 We regularize the problem because of numerical issues that arise if there are points $\widehat{x}_i$ and $\widehat{y}_j$ in the interpolation that are equal or nearly equal, a situation that arises often in practice.
-%Indeed, situations where $\widehat{x}_i = \widehat{y}_j$ are inevitable when a regular grid is used to discretize the problem. 
 In this case, $\widehat{B}$ is (nearly) singular, so it is ill-advised to solve the un-regularized interpolation system $\widehat{B} \widehat{w} = \widehat{f}$.
-
-%In principle, this equation is still solvable if $\left(\widetilde{f}_x\right)_i = \left(\widetilde{f}_x\right)_j$, which should hold in principle.
-However, errors incurred at any stage of the approximation process, particularly the assumption that the support of $\phi_x$ is contained in the ellipsoid $E_x$, will cause small discrepancies between $\left(\widetilde{f}_x\right)_i$ and $\left(\widetilde{f}_x\right)_j$, and these errors get amplified by the ill-conditioning.
-
-%	 or $\widetilde{x}_i \approx \widetilde{y}_j$. 
-%	
-%	while $\left(\widetilde{f}_x\right)_i \neq \left(\widetilde{f}_x\right)_j$
-%
-% $\widetilde{B}\widetilde{w}=\widetilde{f}$
-
-%and let $\widetilde{B}_{xx}$, $\widetilde{B}_{xy}$, $\widetilde{B}_{yx}$, and $\widetilde{B}_{yy}$ be $k \times k$ matrices with the following entries:
-%\begin{align*}
-%	\left(\widetilde{B}_{xx}\right)_{ij} &= b\left(\|\widetilde{x}_i - \widetilde{x}_j\|\right) \\	\left(\widetilde{B}_{xy}\right)_{ij} &= b\left(\|\widetilde{x}_i - \widetilde{y}_j\|\right) \\
-%\end{align*}
-%
-%
-%\begin{equation*}
-%	\begin{bmatrix}
-%		B_{xx} & B_{xy} \\ B_{yx} & B_{yy}
-%	\end{bmatrix}
-%\end{equation*}
-%
-%Let $\phi'_{y_i}(x) := \Phi(y_i,x)$ denote the impulse response of $\Aop^T$ to a point source at $y_i$. 
-%
-%Let $\{y_i\}_{i=1}^r$ be the $k$ nearest sample points to $y$, and let . In the case where $\Aop$ is symmetric, knowing 
-%
-%In the case where $\Aop$ is symmetric, we have
-%\begin{equation}
-%	\impulseresponse_{x_i}(y) = \Phi(y,x_i) = \Phi(x_i,y) = \impulseresponse'_{x_i}(y)
-%\end{equation}
-%knowing the impulse response $\impulseresponse_{x_i}$ of $\Aop$ at $x_i$
-
-%\subsection{Discretization and hierarchical matrix construction}
-%\label{sec:H_matrix_conversion}
-%
-%Here we discuss how we construct an H-matrix representation of our product-convolution approximation, after discretizing the problem with finite elements.
 
@@
 Since $k$ is small, this is cheap.
+
+\section{Content to add}
 
@@
 The first two steps require geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix, but these steps do not depend on the particular values of matrix entries.
-The third step requires us to evaluate $O(\fedim \log \fedim)$ specially-chosen entries of $\AkerPcMat$, and we evaluate these entries using \eqref{eq:Akerpcmat_entries}.
+The third step requires us to evaluate $O(\fedim \log \fedim)$ specially-chosen entries of $\mathbf{\Phi}$, and we evaluate these entries using \eqref{eq:Akerpcmat_entries}.
 
@@
 
+\subsection{disorganized march 21 22}
+
+H-matrix approximations are highly desirable because
+
+The computational cost of these operations is $O(N \log^a N)$, where the matrix is $N\times N$, and $a \in \{1,2,3\}$. 
+Classical H-matrix construction methods 
+require evaluating the integral kernel $O(N \log N)$ times, but this is not feasible 
+are infeasible for large-scale problems in which the operator is only accessible via matrix-vector products (``matrix-free''), because these methods require evaluating the integral kernel of the operator at a large number of points.
+We overcome this problem by c
+
+We improve upon classical product-convolution approximations by translating with respect the estimated mean of the impulse response, rather than the location of the impulse response point source.
+
+We construct a-priori ellipsoid estimates of the supports of the impulse responses (i.e., support ellipsoids are estimated \emph{before} any impulse responses have been computed).
+
+access to $\Aop$ is available through ``operator actions.'' That is,
+
+After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\bm{A}$, that is typically too large to be built and stored.
+This property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\bm{A}$ and $\bm{A}^T$ with arbitrary vectors, but one cannot easily\footnote{One could compute $\bm{v} = \bm{A}\bm{e}_j$, where $\bm{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero, then extract $\bm{A}_{ij} = \bm{v}_i$.
+However, this process is wasteful because one computes the entire vector $\bm{v}$, then discards all but one component.} access matrix entries $\bm{A}_{ij}$.
+
+For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
+
+for a collection of points $x_i$ scattered throughout $\Omega$.
+Further, let $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$,
+\begin{equation}
+%\label{eq:impulse_response_defn}
+\impulseresponse_x(y) := \Aker(y, x),
+\end{equation} 
+be the \emph{impulse response} of $\Aop$ at $x$.
+
+Because we build a H-matrix, and
+
+In this paper we form hierarchical matrix approximations, $\mathbf{A}_H$, of operators $\Aop$ that satisfy the following three properties:
+\begin{description}
+	\item[Locally supported impulse responses:] For all $x\in \Omega$, the support of $\impulseresponse_x$ is contained (or approximately contained) in a localized region.
+	\item[Non-negative impulse responses:] We have
+	\begin{equation*}
+	\impulseresponse_x(y) \ge 0
+	\end{equation*}
+	for all (or approximately all) $(y,x) \in \Omega \times \Omega$.
+	\item[Matrix-free:] One cannot easily evaluate kernel entries $\Aker(y,x)$.
+Instead, access to $\Aop$ is available through ``operator actions.'' That is, we have a black-box computational procedure (which is typically expensive) through which we may evaluate the maps
+	\begin{equation*}
+	u \mapsto\Aop u \quad \text{and} \quad v \mapsto \Aop^Tv
+	\end{equation*} 
+	for arbitrary functions $u$ and $v$.
+After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\bm{A}$, that is typically too large to be built and stored.
+This property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\bm{A}$ and $\bm{A}^T$ with arbitrary vectors, but one cannot easily\footnote{One could compute $\bm{v} = \bm{A}\bm{e}_j$, where $\bm{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero, then extract $\bm{A}_{ij} = \bm{v}_i$.
+However, this process is wasteful because one computes the entire vector $\bm{v}$, then discards all but one component.} access matrix entries $\bm{A}_{ij}$.
+In applications, applying $\bm{A}$ to a vector may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
+	For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
+\end{description} 
+Once the hierarchical matrix approximation $\mathbf{A}_H$ is constructed, one may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
+These H-matrix methods are fast and scale well to large problems.
+WHY DO THESE OPERATIONS: KRYLOV SOLVERS, SAMPLING STRATEGIES, CAN'T DO WITH A.
+
+ approximate operators which possess the following properties:
+
+We further limit our focus to operators that have
+\textbf{Local support}
+We say that $\Aop$ has local support if, for all $x\in \Omega$, the support of $\impulseresponse_x$ is contained (or approximately contained) in a small region. 
+The smaller these supports are, the better our algorithm will perform.
+Let $\delta_x$ denote the delta distribution (point source) centered at $x$.
+From \eqref{eq:impulse_response_delta_action}, we see that this local support property means that the response of $\Aop$ to a point source at $x$ is zero (or small) at points $y$ that are far from $x$.
+%\item[3) Local translation invariance:] We say that $\Aop$ is locally translation invariant if
+\begin{equation}
+%	\label{eq:local_translation_invariance}
+	\Aker(y+h, x+h) \approx \Aker(y,x)
+\end{equation}
+when $h$ is not too large.
+One can imagine the kernel entry $\Aker(y,x)$ as representing the strength of an interaction between a source at point $x$ and a target at point $y$.
+Local translation invariance means that the interaction strength remains roughly the same if the source and target points are both translated by the same vector.
+This is illustrated in Figure FIG.
+\textbf{Non-negative impulse response:} We have
+\begin{equation*}
+\impulseresponse_x(y) \ge 0
+\end{equation*}
+for all $(y,x) \in \Omega \times \Omega$.
+
 \subsection{Product-convolution approximation}

### 2022-03-22 paper/localpsf.tex
edits based on discussion with Noemi

@@
 This includes the Lanczos method, the randomized singular value decomposition \cite{HMTRANDOM}, and CUR approximation/skeletonization \cite{CUR,MAHONEY}.
-However, low rank approximation is not suitable for high rank operators, and many operators of practical interest are high rank.
+However, low rank approximation is not suitable for high rank operators, and often operators of practical interest are high rank.
 In the context of Hessians, methods based on low rank approximation suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator actions are required to form an accurate ``low rank'' approximation \cite{MYDISSERTATION}.
@@
 
-Our method is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i \in \Omega$ \cite{PRODCONVLIST}.
-A popular choice for PC methods in this class is to choose the $x_i$ to be nodes in a regular grid \cite{NAGY}.
+The method proposed in this paper is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i \in \Omega$ \cite{PRODCONVLIST}.
+A common choice for PC methods in this class is to choose the $x_i$ to be nodes in a regular grid \cite{NAGY}.
 With a regular grid, a large number of $x_i$ are typically required to achieve an accurate approximation.
 In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
-But even with adaptive refinement, many $x_i$, and hence operator actions of $\Aop$, are required.
-
-In this paper, we reduce the computational cost by picking points $x_i$ in a way that allows many $\phi_{x_i}$ to be computed with each operator action of $\Aop$.
+However, even with adaptive refinement, many $x_i$, and hence operator actions of $\Aop$, are required.
+
+In this paper, we reduce the computational cost further by picking points $x_i$ in a way that allows many $\phi_{x_i}$ to be computed with each operator action of $\Aop$.
 This requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
-We were inspired by techniques from resolution analysis in seismic imaging \cite{RESOLUTION}.
-In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_x$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $x$.
-If the width of $\phi_x$ is ten meters, then features of the parameter near $x$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
+This idea was inspired by techniques from resolution analysis in seismic imaging \cite{RESOLUTION}.
+In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_x$ is used to estimate the minimum length scale on which features of the parameter can be inferred near the point $x$.
+If the width of $\phi_x$ is ten meters, for instance, then features of the parameter near $x$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
 In \cite{RESOLUTION}, the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion.
 Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Sections \ref{sec:intromoments} and \ref{eq:mean_and_covariance_estimation}).
-Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing, but our method requires that $\Aop$ has a positive integral kernel, while random noise probing does not have this requirement.
-
-Convolutions can be performed using the fast Fourier transform (FFT) \cite{PRODCONVFFT}, which allows PC approximations to be used as a fast surrogates for $\Aop$.
+Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a positive integral kernel.
+
+Conventionally, PC approximations are applied to vectors using the fast Fourier transform (FFT) \cite{PRODCONVFFT}, which allows PC approximations to be used as a fast surrogates for $\Aop$.
 Often it is necessary to perform further linear algebra operations involving $\Aop$, such as solving linear systems with $\Aop$ as the coefficient operator.
-Conventionally, one performs these linear algebra operations using Krylov methods, replacing $\Aop$ with its PC approximation \cite{PRODCONVKRYLOV}.
+One may perform these linear algebra operations using Krylov methods, replacing $\Aop$ with its PC approximation \cite{PRODCONVKRYLOV}.
 The Krylov method is accelerated because performing matrix vector products with the product-convolution approximation is typically faster than performing operator actions with $\Aop$.
@@
 Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}, because operators that are amenable to PC approximation are often sparse in the wavelet domain.
-Convolution-product approximations (transposes of PC approximations) are equivalent to low rank approximation of the pseudo-differential symbol of the operator, so one could perform linear algebra operations involving PC approximations using fast methods for pseudo-differential operator symbol calculus \cite{YINGDEMANET}.
-Other matrix-free operator approximation methods based on wavelets \cite{HERRMANN} and pseudo-differential operators \cite{DEMANET} do not use PC approximation.
-Here, we follow our previous work \cite{PRODCONVMYPAPER}, in which we converted PC approximations into H-matrix format.
-
+While not commonly done, in theory linear algebra operations involving convolution-product approximations (transposes of PC approximations) could be done using fast methods for pseudo-differential operator symbol calculus \cite{YINGDEMANET} since these are equivalent to low rank approximation of the pseudo-differential symbol of the operator.
+There are other relevant matrix-free operator approximation methods based on wavelets \cite{HERRMANN} and pseudo-differential operators \cite{DEMANET}, however these do not use PC approximation.
+
+In this paper, we build on previous work \cite{PRODCONVMYPAPER}, in which we converted PC approximations into H-matrix format.
 Since our end goal is constructing an H-matrix, we are freed from strict adherence to the PC format.
-This allows us to modify the PC approach to address well-known deficiencies.
-PC approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, while our method is based on an assumption that we call ``normalized local mean displacement invariance,'' which is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
-PC approximations have well-known issues near boundaries called boundary artifacts, which are caused by the boundary artificially ``cutting off'' impulse responses.
-In our method we do not have an issue with boundary artifacts because we temporarily exclude impulse responses from the approximation procedure whenever the approximation procedure attempts to evaluate the impulse response at a location where it is undefined.
-In exchange for these benefits, our approximation cannot be written as a weighted sum of convolution operators, so we cannot apply our approximation to vectors via the FFT.
-This is fine because our approximation is an H-matrix, which can be applied to vectors even faster.
-
-Since we are building an H-matrix approximation of $\Aop$, it is natural to ask why we need to compute impulse responses and interpolate them.
+This allows us to modify the PC approach to make the following improvements:
+\begin{itemize}
+\item PC approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, while the method we propose is based on an assumption that we call ``normalized local mean displacement invariance.'' This is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
+\item PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulse responses.
+We avoid this issue by temporarily excluding impulse responses from the approximation procedure whenever the approximation procedure attempts to evaluate the impulse response at a location where it is undefined.
+\end{itemize}
+In exchange for these improvements, the proposed approximation cannot be written as a weighted sum of convolution operators, hence we cannot apply this approximation to vectors via the FFT.
+However, since our approximation is an H-matrix, it can be applied to vectors efficiently via H-matrix methods.
+
+Since we are building an H-matrix approximation of $\Aop$, it is natural to ask why we compute impulse responses and interpolate them.
 Why not build the H-matrix directly?
-The reason is that classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
-Our impulse response interpolation lets us quickly compute approximations to matrix entries, which allows us to use these classical H-matrix construction methods.
-There are newer matrix-free methods for H-matrix approximation based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}, and these methods have been used to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
-While these matrix-free H-matrix construction methods are asymptotically scalable in theory, in practice the required number of matrix vector products is unreasonably large.
-Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods more feasible in the future.
-
-\section{Main concepts}
+The reason is that \emph{classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products}.
+The impulse response interpolation allows fast evaluation of approximate matrix entries needed by the classical H-matrix construction methods.
+Latest developments and improvements in matrix-free H-matrix construction and approximation are based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}.
+These methods have, for instance, been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
+One issue with this approach is that, while these matrix-free H-matrix construction methods are asymptotically scalable in theory, in practice the required number of matrix vector products is large\footnote{DO CALCULATION}.
+Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods based on the peeling process more feasible in the future.
+
+\subsection{Finite element discretization}
+\label{sec:finite_element_kernel}
+
+Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element\footnote{While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead, with minor modifications to our method.} basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
+That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
+For $u_h \in V_h$ we write $\bm{u}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$.
+I.e.,
+\begin{equation}
+\label{eq:fem_coeff_basis}
+u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
+\end{equation}
+
+The interpolation of $\Phi$ onto $V_h \otimes V_h$, which we denote by $\Phi^h$, is given as follows:
+\begin{equation}
+\label{eq:defn_of_Akerpcmesh}
+\Phi(y,x) \approx \Phi^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \Phi(p_i,p_j) \febasis_i(y) \febasis_j(x).
+\end{equation}
+Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations (see Appendix \ref{app:proofs} for details), yields the following approximation:
+\begin{equation}
+\label{eq:kernel_representation_pcmesh}
+(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
+= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u}.
+\end{equation}
+where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
+\begin{equation}
+\label{eq:Akerpcmat_entries}
+\mathbf{\Phi}_{ij} := \Phi(p_i, p_j),
+\end{equation}
+and $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ is the sparse finite element mass matrix which has entries
+\begin{equation*}
+\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
+\end{equation*}
+Our matrix representation, $\mathbf{A}$, of $\Aop$ is therefore given by
+\begin{equation}
+\label{eq:Amatpc_defn}
+\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
+\end{equation}
+Note that $\mathbf{A}$ is not exactly the Galerkin projection of $\Aop$ onto $V_h$ because of the approximation $\Phi^h \approx \Phi$.
+Hence, $\mathbf{A}$ may differ slightly from the implicitly defined matrix that performs the mapping $\mathbf{u} \rightarrow \mathbf{v}$ corresponding to the discrete version of the operator action $v=\Aop u$.
+
+Because $\Aop$ is only available matrix-free and $\Phi$ is only defined implicitly in terms of $\Aop$, the matrices $\mathbf{A}$ and $\mathbf{\Phi}$ are not easily accessible.
+Directly computing and storing these matrices is computationally intractible.
+The primary goal of this paper is building an H-matrix approximation $\mathbf{A}_H$ of $\mathbf{A}$, using a small number of operator actions of $\Aop$ and $\Aop^T$.
+
+%After discretization with finite elements\footnote{While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead, with minor modifications to our method.} (Section \ref{sec:finite_element_kernel}), $\Aop$ and $\Phi$ become large dense matrices $\mathbf{A}$ and $\mathbf{\Phi}$, respectively.
+
+\section{Efficient matrix-free approximation of operators via impulse response interpolation: main ideas}
 \label{sec:prerequisites}
@@
 
-\section{Method}
+\section{Efficient matrix-free approximation of operators via impulse response interpolation: construction of the H-matrix}
 \label{sec:method}
 
-After discretization with finite elements\footnote{While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead, with minor modifications to our method.} (Section \ref{sec:finite_element_kernel}), $\Aop$ and $\Phi$ become large dense matrices $\mathbf{A}$ and $\mathbf{\Phi}$, respectively.
-Because $\Aop$ is only available matrix-free and $\Phi$ is only defined implicitly in terms of $\Aop$, the matrices $\mathbf{A}$ and $\mathbf{\Phi}$ are not easily accessible.
-Directly computing and storing these matrices is computationally intractible.
 In this section, we present our method for constructing accessible H-matrix approximations, $\mathbf{A}_H$ and $\mathbf{\Phi}_H$ of $\mathbf{A}$ and $\mathbf{\Phi}$, respectively, using a small number of operator actions of $\Aop$ and $\Aop^T$.
-
-\subsection{Overview}
-\label{sec:method_overview}
 
@@
 \end{algorithm2e}
-
-\subsection{Finite element discretization}
-\label{sec:finite_element_kernel}
-
-Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
-That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
-For $u_h \in V_h$ we write $\bm{u}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$.
-I.e.,
-\begin{equation}
-\label{eq:fem_coeff_basis}
-u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
-\end{equation}
-
-The interpolation of $\Phi$ onto $V_h \otimes V_h$, which we denote by $\Phi^h$, is given as follows:
-\begin{equation}
-\label{eq:defn_of_Akerpcmesh}
-\Phi(y,x) \approx \Phi^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \Phi(p_i,p_j) \febasis_i(y) \febasis_j(x).
-\end{equation}
-Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations (see Appendix \ref{app:proofs} for details), yields the following approximation:
-\begin{equation}
-\label{eq:kernel_representation_pcmesh}
-(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
-= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u}.
-\end{equation}
-where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
-\begin{equation}
-\label{eq:Akerpcmat_entries}
-\mathbf{\Phi}_{ij} := \Phi(p_i, p_j),
-\end{equation}
-and $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ is the sparse finite element mass matrix which has entries
-\begin{equation*}
-\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
-\end{equation*}
-Our matrix representation, $\mathbf{A}$, of $\Aop$ is therefore given by
-\begin{equation}
-\label{eq:Amatpc_defn}
-\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
-\end{equation}
-Note that $\mathbf{A}$ is not exactly the Galerkin projection of $\Aop$ onto $V_h$ because of the approximation $\Phi^h \approx \Phi$.
-Hence, $\mathbf{A}$ may differ slightly from the implicitly defined matrix that performs the mapping $\mathbf{u} \rightarrow \mathbf{v}$ corresponding to the discrete version of the operator action $v=\Aop u$.
 

### 2022-03-23 paper/localpsf.tex
moved discretization to intro

@@
 
-Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element\footnote{While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead, with minor modifications to our method.} basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
+In computations, functions are discretized and replaced by finite dimensional vectors, and operators mapping between infinite dimensional spaces are replaced by operators mapping between finite dimensional spaces.
+In this paper we discretize using finite elements.
+With minor modifications, our method could be used with other discretization schemes such as finite differences or finite volumes.
+
+Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space under the $L^2$ inner product, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
 That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
-For $u_h \in V_h$ we write $\bm{u}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$.
-I.e.,
+Further, let $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ denote the sparse finite element mass matrix which has entries
+\begin{equation*}
+\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
+\end{equation*}
+For $u_h \in V_h \subset L^2(\Omega)$ we write $\bm{u} \in \mathbb{R}^N_\mathbf{M}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$, i.e.,
 \begin{equation}
@@
 \end{equation}
+Linear functionals $\rho_h \in V_h' \supset L^2(\Omega)'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^N_{\mathbf{M}^{-1}}$, with entries $\left(\boldsymbol{\rho}_h\right)_i = \rho_h(\febasis_i)$ for $i=1,\dots,N$.
+Here $\mathbb{R}^N_\mathbf{M}$ is the vector space $\mathbb{R}^N$ with the inner product $(\mathbf{u},\mathbf{v})_\mathbf{M} := \mathbf{u}^T \mathbf{M} \mathbf{v}$, and $\mathbb{R}^N_{\mathbf{M}^{-1}}$ is the analogous space with $\mathbf{M}^{-1}$ replacing $\mathbf{M}$.
+Direct calculation shows that $\mathbf{R}^N_\mathbf{M}$ and $\mathbf{R}^N_{\mathbf{M}^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively, in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u}, \mathbf{v}\right)_\mathbf{M}$, $\left(\mu_h, \nu_h\right)_{L^2(\Omega)} = \left(\boldsymbol{\mu}, \boldsymbol{\nu}\right)_{\mathbf{M}^{-1}}$, and $\mu_h(u) = \boldsymbol{\mu}^T\mathbf{u}$, for $u_h,v_h \in V_h$ and $\mu_h, \nu_h \in V_h'$.
+The Riesz representation $\mu_h^* \in V_h$ of $\mu \in V_h'$ becomes $\boldsymbol{\mu}^* = \mathbf{M}^{-1} \boldsymbol{\mu}$.
+After Galerkin approximation, the operator $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$ becomes an operator $\mathbf{A}:\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has matrix entries $\mathbf{A}_{ij} = \left(\Aop \febasis_j\right)(\febasis_i)$.
+Notice that $\mathbf{A}$ is dense, and computing one matrix entry $\mathbf{A}_{ij}$ requires an action of $\Aop$, which is already computationally expensive (though we get the entire $j$th column as a side effect of computing $\mathbf{A}_{ij}$).
+Hence methods for approximating $\mathbf{A}$ are computationally intractable if they require accessing scattered matrix entries from many different rows and columns of $\mathbf{A}$.
+Computing and storing a matrix representation of $\mathbf{A}$ is out of the question.
+
+If functions in $V_h$ are continuous at $x$, then the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta} \in V_h'$ with entries $\boldsymbol{\delta}_i = \febasis_i(x)$, $i=1,\dots,N$.
+The Riesz representative of the discrete delta distribution is $\boldsymbol{\delta}^* = \mathbf{M}^{-1} \boldsymbol{\delta}_x$, and the formula for the impulse response from \eqref{eq:impulse_response_delta_action} becomes $\boldsymbol{\impulseresponse}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x$.
+
+\subsubsection{Discretized integral kernel}
 
@@
 \end{equation}
-Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations (see Appendix \ref{app:proofs} for details), yields the following approximation:
+Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations, yields the following approximation:
 \begin{equation}
@@
 (\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
-= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u}.
+= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u},
 \end{equation}
@@
 \end{equation}
-and $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ is the sparse finite element mass matrix which has entries
-\begin{equation*}
-\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
-\end{equation*}
-Our matrix representation, $\mathbf{A}$, of $\Aop$ is therefore given by
+Based on \eqref{eq:kernel_representation_pcmesh}, we define the following matrix approximation, $\widetilde{\mathbf{A}} \approx \mathbf{A}$ as follows:
 \begin{equation}
 \label{eq:Amatpc_defn}
-\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
-\end{equation}
-Note that $\mathbf{A}$ is not exactly the Galerkin projection of $\Aop$ onto $V_h$ because of the approximation $\Phi^h \approx \Phi$.
-Hence, $\mathbf{A}$ may differ slightly from the implicitly defined matrix that performs the mapping $\mathbf{u} \rightarrow \mathbf{v}$ corresponding to the discrete version of the operator action $v=\Aop u$.
-
-Because $\Aop$ is only available matrix-free and $\Phi$ is only defined implicitly in terms of $\Aop$, the matrices $\mathbf{A}$ and $\mathbf{\Phi}$ are not easily accessible.
-Directly computing and storing these matrices is computationally intractible.
-The primary goal of this paper is building an H-matrix approximation $\mathbf{A}_H$ of $\mathbf{A}$, using a small number of operator actions of $\Aop$ and $\Aop^T$.
+\widetilde{\mathbf{A}} = \bm{M} \mathbf{\Phi} \bm{M}.
+\end{equation}
+Note that $\widetilde{\mathbf{A}}$ differs slightly from $\mathbf{A}$ (the Galerkin projection of $\Aop$ onto $V_h$) because of the approximation $\Phi^h \approx \Phi$. 
+%Hence, $\mathbf{A}$ may differ slightly from the implicitly defined matrix that performs the mapping $\mathbf{u} \rightarrow \mathbf{v}$ corresponding to the discrete version of the operator action $v=\Aop u$.
+%Because $\Aop$ is only available matrix-free and $\Phi$ is only defined implicitly in terms of $\Aop$, 
+The matrices $\widetilde{\mathbf{A}}$ and $\mathbf{\Phi}$ are, of course, not easily accessible.  
+%The primary goal of this paper is building an H-matrix approximation $\mathbf{A}_H$ of $\widetilde{\mathbf{A}}$, using a small number of operator actions of $\Aop$ and $\Aop^T$.
 
@@
 
-\section{Efficient matrix-free approximation of operators via impulse response interpolation: main ideas}
+\subsubsection{Lumped mass matrix and discrete kernel non-negativity}
+
+\section{Key innovations}
 \label{sec:prerequisites}
@@
 
-\section{Efficient matrix-free approximation of operators via impulse response interpolation: construction of the H-matrix}
+\section{Method}
 \label{sec:method}
 
-In this section, we present our method for constructing accessible H-matrix approximations, $\mathbf{A}_H$ and $\mathbf{\Phi}_H$ of $\mathbf{A}$ and $\mathbf{\Phi}$, respectively, using a small number of operator actions of $\Aop$ and $\Aop^T$.
+%In this section, we present our method for constructing accessible H-matrix approximations, $\mathbf{A}_H$ and $\mathbf{\Phi}_H$ of $\mathbf{A}$ and $\mathbf{\Phi}$, respectively, using a small number of operator actions of $\Aop$ and $\Aop^T$.
 

### 2022-03-23 paper/localpsf.tex
wrote lumped mass section

@@
 In \cite{RESOLUTION}, the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion.
-Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Sections \ref{sec:intromoments} and \ref{eq:mean_and_covariance_estimation}).
+Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Section \ref{sec:intromoments}).
 Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a positive integral kernel.
@@
 Here $\mathbb{R}^N_\mathbf{M}$ is the vector space $\mathbb{R}^N$ with the inner product $(\mathbf{u},\mathbf{v})_\mathbf{M} := \mathbf{u}^T \mathbf{M} \mathbf{v}$, and $\mathbb{R}^N_{\mathbf{M}^{-1}}$ is the analogous space with $\mathbf{M}^{-1}$ replacing $\mathbf{M}$.
-Direct calculation shows that $\mathbf{R}^N_\mathbf{M}$ and $\mathbf{R}^N_{\mathbf{M}^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively, in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u}, \mathbf{v}\right)_\mathbf{M}$, $\left(\mu_h, \nu_h\right)_{L^2(\Omega)} = \left(\boldsymbol{\mu}, \boldsymbol{\nu}\right)_{\mathbf{M}^{-1}}$, and $\mu_h(u) = \boldsymbol{\mu}^T\mathbf{u}$, for $u_h,v_h \in V_h$ and $\mu_h, \nu_h \in V_h'$.
-The Riesz representation $\mu_h^* \in V_h$ of $\mu \in V_h'$ becomes $\boldsymbol{\mu}^* = \mathbf{M}^{-1} \boldsymbol{\mu}$.
+Direct calculation shows that $\mathbf{R}^N_\mathbf{M}$ and $\mathbf{R}^N_{\mathbf{M}^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively, in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u}, \mathbf{v}\right)_\mathbf{M}$, $\left(\rho_h, \nu_h\right)_{L^2(\Omega)} = \left(\boldsymbol{\rho}, \boldsymbol{\nu}\right)_{\mathbf{M}^{-1}}$, and $\rho_h(u) = \boldsymbol{\rho}^T\mathbf{u}$, for $u_h,v_h \in V_h$ and $\rho_h, \nu_h \in V_h'$.
+The Riesz representation $\rho_h^* \in V_h$ of $\rho \in V_h'$ becomes $\boldsymbol{\rho}^* = \mathbf{M}^{-1} \boldsymbol{\rho}$.
+
 After Galerkin approximation, the operator $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$ becomes an operator $\mathbf{A}:\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has matrix entries $\mathbf{A}_{ij} = \left(\Aop \febasis_j\right)(\febasis_i)$.
@@
 
-If functions in $V_h$ are continuous at $x$, then the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta} \in V_h'$ with entries $\boldsymbol{\delta}_i = \febasis_i(x)$, $i=1,\dots,N$.
-The Riesz representative of the discrete delta distribution is $\boldsymbol{\delta}^* = \mathbf{M}^{-1} \boldsymbol{\delta}_x$, and the formula for the impulse response from \eqref{eq:impulse_response_delta_action} becomes $\boldsymbol{\impulseresponse}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x$.
+If functions in $V_h$ are continuous at $x$, then the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta} \in V_h'$ with entries $\boldsymbol{\delta}_i = \febasis_i(x)$ for $i=1,\dots,N$.
+The Riesz representative of the discrete delta distribution is $\boldsymbol{\delta}^* = \mathbf{M}^{-1} \boldsymbol{\delta}_x$, and the formula for the impulse response from \eqref{eq:impulse_response_delta_action} becomes
+\begin{equation}
+\label{eq:discrete_kernel}
+\boldsymbol{\impulseresponse}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x.
+\end{equation}
 
@@
 
+At the continuum level, the integral kernel $\Phi$ is assumed to be non-negative.
+However, per \eqref{eq:discrete_kernel}, when we compute impulse responses we are effectively operating with a discrete integral kernel which has $(y,x)$ entry
+\begin{equation*}
+\boldsymbol{\delta}_y^T \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x.
+\end{equation*}
+Even if the discretization $\Aop \rightarrow \mathbf{A}$ preserves integral kernel positivity, the inverse mass matrices $\mathbf{M}^{-1}$ will typically contain negative numbers.
+If there are too many negative numbers, or if the negative numbers are too large, our algorithm will be less robust.
+We therefore recommend replacing the mass matrix $\mathbf{M}$ with a positive diagonal \emph{lumped mass} approximation, and using the resulting lumped mass matrix everywhere throughout the approximation method.
+For typical finite element meshes, lumped mass matrix approximations are typically spectrally equivalent to the mass matrix \cite{MASSLUMP2}.
+In our numerical results, we use the lumped mass matrix constructed by replacing off-diagonal entries of the mass matrix by zero.
+Other more advanced mass lumping techniques may be used \cite{MASSLUMP2}.
+
 \section{Key innovations}
@@
 
-In this section we present three main concepts that our method is based on.
+In this section we present two key innovations that our method is based on.
 The method itself will be presented in Section \ref{sec:method}.
-First, we define moments of the impulse responses, $\phi_x$, and show how these moments can be computed efficiently (Section \ref{sec:intromoments}).
-Our method will use these moments extensively.
-Second, we present ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{eq:mean_and_covariance_estimation}).
+First, we define moments of the impulse responses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{sec:intromoments}).
 Our method will use these ellipsoid estimates to choose sets of impulse responses that can be computed in batches.
-Third, we describe a method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}).
+Second, we describe an improved method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}).
 Our method will interpolate impulse responses using this approximation.
 
-\subsection{Impulse response moments}
+\subsection{Impulse response moments and support estimation}
 \label{sec:intromoments}
@@
 
-\subsection{Impulse response support ellipsoids}
-\label{eq:mean_and_covariance_estimation}
+%\subsection{Impulse response support ellipsoids}
+%\label{eq:mean_and_covariance_estimation}
 
@@
 
-Algorithm \ref{alg:varhpi_mean_cov} from Section \ref{sec:intromoments} allows us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} from Section \ref{eq:mean_and_covariance_estimation} uses $V$, $\mu$, and $\Sigma$ to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
+Algorithm \ref{alg:varhpi_mean_cov} from Section \ref{sec:intromoments} allows us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} uses $V$, $\mu$, and $\Sigma$ to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
 This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}).
@@
 \label{sec:overall_cost}
+
+WORK IN PROGRESS BELOW HERE!!!
 

### 2022-03-23 paper/localpsf.tex
add sentence to abstract about classical H-matrix construction limitations for matrix-free operators

@@
 Such operators arise, for example, as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained inverse problems.
-Fast H-matrix methods can add, multiply, invert, and factorize H-matrices, compute matrix-vector products, solve linear systems, and perform other useful matrix computations.
-Our method computes impulse responses of the operator at a collection of scattered points, then interpolates these known impulse responses to evaluate entries of the integral kernel at other points where the impulse response is not known.
+H-matrix approximations are highly desirable because fast H-matrix methods can add, multiply, invert, and factorize H-matrices, compute matrix-vector products, solve linear systems, and perform other useful matrix computations.
+Classical H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
+Our method computes impulse responses of the operator at a collection of scattered points, then interpolates these known impulse responses to evaluate entries of the integral kernel at other points where the impulse response is not known, which allows us to construct the H-matrix.
 Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources.
@@
 Numerical results demonstrate that our method substantially outperforms existing state-of-the-art Hessian approximation methods which are based on low-rank approximation.
-We are able to form high quality approximations of high rank operators using only a small number of operator actions products.
+We are able to form high quality approximations of high rank operators using only a small number of operator actions.
 \end{abstract}

### 2022-03-23 paper/localpsf.tex
minor changes

@@
 Here $f/g$ denotes pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denotes pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
-The resulting method form computing $V$, $\mu$, and $\Sigma$ is shown in Algorithm \ref{alg:varhpi_mean_cov}.
+The resulting method for computing discretized versions of $V$, $\mu$, and $\Sigma$ is shown in Algorithm \ref{alg:varhpi_mean_cov}.
 
@@
 	{	
-		\Input{Operator $\Aop$}
-		\Output{$\spatialvol$, $\spatialmean$, and $\spatialcov$}
-
-		\tcp{Compute scaling factor $\spatialvol$}
-
-		Form constant function $C(x)=1$
-
-		Compute $\spatialvol = \left(\Aop^T C\right)^*$
-
-		\tcp{Compute mean $\spatialmean$}
+		\Input{Operator $\bm{A}^T$}
+		\Output{$\bm{\spatialvol}\in \mathbb{R}^\fedim$, $\bm{\spatialmean}\in \mathbb{R}^{\gdim \times \fedim}$, and $\bm{\spatialcov} \in \mathbb{R}^{\gdim \times \gdim \times \fedim}$}
+
+		Form vector $\bm{C} \in \mathbb{R}^\fedim_{\bm{M}}$ by either projecting or interpolating constant function $C(x)=1$ onto $V_h$
+
+		$\bm{\spatialvol} \gets \bm{M}^{-1}\bm{A}^T \bm{C}$
+
 		\For{$i=1,2,\dots,\gdim$}{
-			Form linear function $L^i(x) = x^i$
-
-			Compute $\spatialmean^i = \left(\Aop^T L^i\right)^* / \spatialvol$
+			Form vector $\bm{L}^i \in \mathbb{R}^\fedim_{\bm{M}}$ by either projecting or interpolating linear function $L^i(x) = x^i$ onto $V_h$
+
+			$\bm{\spatialmean}^i \gets \left(\bm{M}^{-1}\bm{A}^T \bm{L}^i\right) / \bm{\spatialvol}$
 		}
-		\tcp{Compute covariance $\spatialcov$}
 		\For{$i=1,2,\dots,\gdim$}{
 			\For{$j=1,\dots,i$}{
-				Form quadratic function $Q^{ij}(x) = x^i x^j$
-
-				Compute $\spatialcov^{ij} = \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j$
-
-				Set $\spatialcov^{ji} = \spatialcov^{ij}$
+				Form quadratic function $\bm{Q}^{ij} = x^i x^j$
+
+				$\bm{\spatialcov}^{ij} \gets \left(\bm{M}^{-1}\bm{A}^T \bm{Q}^{ij}\right) / \bm{\spatialvol} - \bm{\spatialmean}^i\cdot \bm{\spatialmean}^j$
+
+				$\bm{\spatialcov}^{ji} \gets \bm{\spatialcov}^{ij}$
 
@@
 	}
-	\caption{Compute scaling factor $\spatialvol$, mean $\spatialmean$, and covariance $\spatialcov$}
+	\caption{Compute discretized scaling factor $\mathbf{V}$, mean $\boldsymbol{\mu}$, and covariance $\boldsymbol{\Sigma}$}
 	\label{alg:varhpi_mean_cov}
 \end{algorithm2e}
+
+%\begin{algorithm2e}
+%	\SetAlgoNoLine
+%	\SetKwInOut{Input}{Input}
+%	\SetKwInOut{Output}{Output}
+%	{	
+%		\Input{Operator $\Aop$}
+%		\Output{$\spatialvol$, $\spatialmean$, and $\spatialcov$}
+%		
+%		\tcp{Compute scaling factor $\spatialvol$}
+%		
+%		Form constant function $C(x)=1$
+%		
+%		Compute $\spatialvol = \left(\Aop^T C\right)^*$
+%		
+%		\tcp{Compute mean $\spatialmean$}
+%		\For{$i=1,2,\dots,\gdim$}{
+%			Form linear function $L^i(x) = x^i$
+%			
+%			Compute $\spatialmean^i = \left(\Aop^T L^i\right)^* / \spatialvol$
+%		}
+%		\tcp{Compute covariance $\spatialcov$}
+%		\For{$i=1,2,\dots,\gdim$}{
+%			\For{$j=1,\dots,i$}{
+%				Form quadratic function $Q^{ij}(x) = x^i x^j$
+%				
+%				Compute $\spatialcov^{ij} = \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j$
+%				
+%				Set $\spatialcov^{ji} = \spatialcov^{ij}$
+%				
+%			}
+%		}
+%		
+%	}
+%	\caption{Compute scaling factor $\spatialvol$, mean $\spatialmean$, and covariance $\spatialcov$}
+%	\label{alg:varhpi_mean_cov}
+%\end{algorithm2e}
 
@@
 
-		Compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ using Algorithm \ref{alg:varhpi_mean_cov}.
+		Compute $\bm{\spatialvol}$, $\bm{\spatialmean}$, and $\bm{\spatialcov}$ using Algorithm \ref{alg:varhpi_mean_cov}.
 
@@
 	\item We check whether $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_k$.
-	\item We make sure that $V(p)$ is not too small, by checking whether $V(p) > \tau V_\text{max}$.
-Here $V_\text{max}$ is the largest value of $V(q)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\tau=10^{-5}$).
+	\item We make sure that $V(p)$ is not too small, by checking whether $V(p) > \epsilon V_\text{max}$.
+Here $V_\text{max}$ is the largest value of $V(q)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\epsilon=10^{-5}$).
 \end{enumerate}

### 2022-03-23 paper/localpsf.tex
minor changes

@@
 
+\begin{tabular}{lll}
+	Variable name & Approximate size & Variable \\
+	\hline
+	$N$ & $10^3$--$10^9$ & Number of finite element degrees of freedom \\
+	$b$ & $5$--$25$ & number of impulse response batches \\
+	$r$ & $10$--$30$ & H-matrix rank \\
+	$k$ & $5$--$15$ & number of nearest neighbors \\
+	$\{\eta_b\}_{b=1}^\text{num\_batches}$ & $\text{num\_batches}$ actions of $\mathcal{A}$
+\end{tabular}
+
+$k_\text{hmatrix}$, $k_\text{nbrs}$, $k_\text{batches}$
+
+\begin{tabular}{ll}
+	$V$ & $1$ action of $\mathcal{A}^T$ \\
+	$\mu$ & $d$ actions of $\mathcal{A}^T$ \\
+	$\Sigma$ & $d(d+1)/2$ actions of $\mathcal{A}^T$ \\
+	$\{\eta_b\}_{b=1}^\text{num\_batches}$ & $\text{num\_batches}$ actions of $\mathcal{A}$
+\end{tabular}
+
 Computing $V$, $\mu$, and $\Sigma$ require $1$, $d$, and $d(d+1)/2$ operator actions of $\Aop$, respectively.

### 2022-03-25 paper/localpsf.tex
updated discretization section, big changes to that section

@@
 In computations, functions are discretized and replaced by finite dimensional vectors, and operators mapping between infinite dimensional spaces are replaced by operators mapping between finite dimensional spaces.
-In this paper we discretize using finite elements.
-With minor modifications, our method could be used with other discretization schemes such as finite differences or finite volumes.
-
-Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space under the $L^2$ inner product, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
-That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
-Further, let $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ denote the sparse finite element mass matrix which has entries
-\begin{equation*}
-\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
-\end{equation*}
+In this paper we discretize using continuous Galerkin finite elements satisfying the Kronecker property (defined below).
+With minor modifications, our method could be used with more general finite element methods, or other discretization schemes such as finite differences or finite volumes.
+
+Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of continuous Galerkin finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space under the $L^2$ inner product, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
+We assume that the finite element basis satisfies the Kronecker property, i.e., $\febasis_i(p_i)=1$ and $\febasis_i(p_j)=0$ if $i\neq j$.
 For $u_h \in V_h \subset L^2(\Omega)$ we write $\bm{u} \in \mathbb{R}^N_\mathbf{M}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$, i.e.,
@@
 \end{equation}
-Linear functionals $\rho_h \in V_h' \supset L^2(\Omega)'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^N_{\mathbf{M}^{-1}}$, with entries $\left(\boldsymbol{\rho}_h\right)_i = \rho_h(\febasis_i)$ for $i=1,\dots,N$.
-Here $\mathbb{R}^N_\mathbf{M}$ is the vector space $\mathbb{R}^N$ with the inner product $(\mathbf{u},\mathbf{v})_\mathbf{M} := \mathbf{u}^T \mathbf{M} \mathbf{v}$, and $\mathbb{R}^N_{\mathbf{M}^{-1}}$ is the analogous space with $\mathbf{M}^{-1}$ replacing $\mathbf{M}$.
-Direct calculation shows that $\mathbf{R}^N_\mathbf{M}$ and $\mathbf{R}^N_{\mathbf{M}^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively, in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u}, \mathbf{v}\right)_\mathbf{M}$, $\left(\rho_h, \nu_h\right)_{L^2(\Omega)} = \left(\boldsymbol{\rho}, \boldsymbol{\nu}\right)_{\mathbf{M}^{-1}}$, and $\rho_h(u) = \boldsymbol{\rho}^T\mathbf{u}$, for $u_h,v_h \in V_h$ and $\rho_h, \nu_h \in V_h'$.
-The Riesz representation $\rho_h^* \in V_h$ of $\rho \in V_h'$ becomes $\boldsymbol{\rho}^* = \mathbf{M}^{-1} \boldsymbol{\rho}$.
-
-After Galerkin approximation, the operator $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$ becomes an operator $\mathbf{A}:\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has matrix entries $\mathbf{A}_{ij} = \left(\Aop \febasis_j\right)(\febasis_i)$.
-Notice that $\mathbf{A}$ is dense, and computing one matrix entry $\mathbf{A}_{ij}$ requires an action of $\Aop$, which is already computationally expensive (though we get the entire $j$th column as a side effect of computing $\mathbf{A}_{ij}$).
-Hence methods for approximating $\mathbf{A}$ are computationally intractable if they require accessing scattered matrix entries from many different rows and columns of $\mathbf{A}$.
-Computing and storing a matrix representation of $\mathbf{A}$ is out of the question.
-
-If functions in $V_h$ are continuous at $x$, then the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta} \in V_h'$ with entries $\boldsymbol{\delta}_i = \febasis_i(x)$ for $i=1,\dots,N$.
-The Riesz representative of the discrete delta distribution is $\boldsymbol{\delta}^* = \mathbf{M}^{-1} \boldsymbol{\delta}_x$, and the formula for the impulse response from \eqref{eq:impulse_response_delta_action} becomes
+Linear functionals $\rho_h \in V_h' \supset L^2(\Omega)'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^N_{\mathbf{M}^{-1}}$, with entries 
+\begin{equation*}
+\left(\boldsymbol{\rho}_h\right)_i = \rho_h(\febasis_i), \qquad i=1,\dots,N. 
+\end{equation*}
+Here $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries
+\begin{equation*}
+\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx, \qquad i,j=1,\dots,N,
+\end{equation*}
+the space $\mathbb{R}^N_\mathbf{M}$ is $\mathbb{R}^N$ with the inner product $(\mathbf{u},\mathbf{v})_\mathbf{M} := \mathbf{u}^T \mathbf{M} \mathbf{v}$, and $\mathbb{R}^N_{\mathbf{M}^{-1}}$ is the analogous space with $\mathbf{M}^{-1}$ replacing $\mathbf{M}$.
+Direct calculation shows that $\mathbf{R}^N_\mathbf{M}$ and $\mathbf{R}^N_{\mathbf{M}^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively.
+%, in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u}, \mathbf{v}\right)_\mathbf{M}$, $\left(\rho_h, \nu_h\right)_{L^2(\Omega)} = \left(\boldsymbol{\rho}, \boldsymbol{\nu}\right)_{\mathbf{M}^{-1}}$, and $\rho_h(u) = \boldsymbol{\rho}^T\mathbf{u}$, for $u_h,v_h \in V_h$ and $\rho_h, \nu_h \in V_h'$.
+
+After discretization, the operator $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$ is replaced by an operator $A_h:V_h \rightarrow V_h'$, which becomes an operator
+\begin{equation*}
+\mathbf{A}:\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}
+\end{equation*}
+under the isomorphism discussed above.
+Our method is agnostic to the computational procedure for approximating $\Aop$ with $\mathbf{A}$.
+What is important is that we do not have direct access to matrix entries $\mathbf{A}_{ij}$.
+Rather, we have a computational procedure that allows us to compute matrix-vector products $\mathbf{u}\mapsto \mathbf{A}\mathbf{u}$, and computing these matrix-vector products is costly.
+%, and each of these matrix-vector products is costly to compute. 
+%For example, some computational procedures possess the Galerkin property, in which case $\mathbf{A}_{ij} = \left(\Aop \febasis_j\right)(\febasis_i)$. 
+%Our approximation method is agnostic to the computational procedure for performing these matrix-vector products. 
+%What is important is that $\mathbf{A}$ is dense, only available via matrix-vector products, and each of these matrix-vector products is costly to compute. 
+Of course, matrix entries can be computed via a matrix-vector products as $\mathbf{A}_{ij} = \left(\mathbf{A}\mathbf{e}_j\right)_i$,
+%\begin{equation*}
+%\mathbf{A}_{ij} = \left(\mathbf{A}\mathbf{e}_j\right)_i,
+%\end{equation*}
+where $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)^T$ is the length $N$ unit vector with one in the $j$th coordinate and zeros elsewhere.
+But computing the matrix-vector product $\mathbf{e}_j \mapsto \mathbf{A}\mathbf{e}_j$ is costly, and therefore wasteful if we do not use other matrix entries in the $j$th column of $\mathbf{A}$.
+Hence, methods for approximating $\mathbf{A}$ are computationally intractable if they require accessing scattered matrix entries from many different rows and columns of $\mathbf{A}$.
+The operator $A_h:V_h \rightarrow V_h'$ can be written in integral kernel form as
+\begin{equation}
+(A_h u_h)(v_h) := \int_\Omega \int_\Omega v_h(y) \Phi_h(y,x) u_h(x) dx dy
+\end{equation}
+for some kernel $\Phi_h$ which may differ from $\Phi$ due to discretization error.
+Since the functions $V_h$ are continuous at $x$, the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta}_x \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ with entries $\left(\boldsymbol{\delta}_x\right)_i = \febasis_i(x)$ for $i=1,\dots,N$.
+Additionally, it is straightforward to show that the Riesz representation, $\rho_h^* \in V_h$, of a functional $\rho \in V_h'$ has coefficient vector
+\begin{equation*}
+\boldsymbol{\rho}^* = \mathbf{M}^{-1} \boldsymbol{\rho}.
+\end{equation*}
+Therefore, the formula for the impulse response from \eqref{eq:impulse_response_delta_action} becomes 
 \begin{equation}
 \label{eq:discrete_kernel}
-\boldsymbol{\impulseresponse}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x.
-\end{equation}
-
-\subsubsection{Discretized integral kernel}
-
-The interpolation of $\Phi$ onto $V_h \otimes V_h$, which we denote by $\Phi^h$, is given as follows:
-\begin{equation}
-\label{eq:defn_of_Akerpcmesh}
-\Phi(y,x) \approx \Phi^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \Phi(p_i,p_j) \febasis_i(y) \febasis_j(x).
-\end{equation}
-Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations, yields the following approximation:
-\begin{equation}
-\label{eq:kernel_representation_pcmesh}
-(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
-= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u},
-\end{equation}
-where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
+\boldsymbol{\impulseresponse}_x = \left(A_h \delta_x^*\right)^* =  \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x,
+\end{equation}
+and the $(y,x)$ kernel entry of $\Phi_h$ may be written as
+\begin{equation}
+\label{eq:approx_kernel_entry_81}
+	\Phi_h(y,x) = \boldsymbol{\delta}_y^T \boldsymbol{\impulseresponse}_x = \boldsymbol{\delta}_y^T \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x.
+\end{equation}
+Let $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ be the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
 \begin{equation}
 \label{eq:Akerpcmat_entries}
-\mathbf{\Phi}_{ij} := \Phi(p_i, p_j),
-\end{equation}
-Based on \eqref{eq:kernel_representation_pcmesh}, we define the following matrix approximation, $\widetilde{\mathbf{A}} \approx \mathbf{A}$ as follows:
-\begin{equation}
-\label{eq:Amatpc_defn}
-\widetilde{\mathbf{A}} = \bm{M} \mathbf{\Phi} \bm{M}.
-\end{equation}
-Note that $\widetilde{\mathbf{A}}$ differs slightly from $\mathbf{A}$ (the Galerkin projection of $\Aop$ onto $V_h$) because of the approximation $\Phi^h \approx \Phi$. 
-%Hence, $\mathbf{A}$ may differ slightly from the implicitly defined matrix that performs the mapping $\mathbf{u} \rightarrow \mathbf{v}$ corresponding to the discrete version of the operator action $v=\Aop u$.
-%Because $\Aop$ is only available matrix-free and $\Phi$ is only defined implicitly in terms of $\Aop$, 
-The matrices $\widetilde{\mathbf{A}}$ and $\mathbf{\Phi}$ are, of course, not easily accessible.  
-%The primary goal of this paper is building an H-matrix approximation $\mathbf{A}_H$ of $\widetilde{\mathbf{A}}$, using a small number of operator actions of $\Aop$ and $\Aop^T$.
-
-%After discretization with finite elements\footnote{While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead, with minor modifications to our method.} (Section \ref{sec:finite_element_kernel}), $\Aop$ and $\Phi$ become large dense matrices $\mathbf{A}$ and $\mathbf{\Phi}$, respectively.
+\mathbf{\Phi}_{ij} := \Phi_h(p_i, p_j).
+\end{equation}
+Because of the Kronecker property of the finite element basis, we have $\boldsymbol{\delta}_{p_i} = \mathbf{e}_i$. Thus from \eqref{eq:approx_kernel_entry_81} we have
+$
+	\Phi_h(p_i,p_j) = \left(\mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1}\right)_{ij},
+$
+which implies
+\begin{equation}
+\label{eq:boldA}
+	\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
+\end{equation}
+Broadly, our method will construct an H-matrix approximation of $\mathbf{A}$ by forming H-matrix approximations of $\boldsymbol{\Phi}$ and $\mathbf{M}$ (or a lumped mass version of $\mathbf{M}$), then multiplying these matrices per \eqref{eq:boldA}.
+The central challenge is that classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Phi}_{ij}$, but these matrix entries are not easily accessible.
+The bulk of our method is therefore dedicated to forming approximations of these matrix entries that can be evaluated rapidly.
+
+%\begin{align*}
+%	\text{continuous} \quad &\longrightarrow \quad \text{discrete} \\
+%	\delta_x \quad &\longrightarrow \quad \boldsymbol{\delta}_x \\
+%	\delta_x^* \quad &\longrightarrow \quad \mathbf{M}^{-1}\boldsymbol{\delta}_x \\
+%	\Aop \delta_x^* \quad &\longrightarrow \quad \mathbf{A}\mathbf{M}^{-1}\boldsymbol{\delta}_x \\
+%	\impulseresponse_x = \left(\Aop \delta_x^*\right)^* \quad &\longrightarrow \quad \mathbf{M}^{-1}\mathbf{A}\mathbf{M}^{-1}\boldsymbol{\delta}_x = \boldsymbol{\impulseresponse}_x \\
+%	\Phi(y,x) \quad &\longrightarrow \quad \boldsymbol{\delta}_y^T\mathbf{M}^{-1}\mathbf{A}\mathbf{M}^{-1}\boldsymbol{\delta}_x
+%\end{align*}
+%Since the Lagrange finite element nodes have the Kronecker property, that is $\febasis_i(p_i) = 1$ and $\febasis_i(p_j) = 1$ for $i \neq j$, the vector $\boldsymbol{\delta}_{p_i}$ is the unit vector with one in the $i$'th component and zeros in all other components.
+Hence
+%\begin{equation*}
+%	\Phi(p_i,p_j) = \left(\mathbf{M}^{-1}\mathbf{A}\mathbf{M}^{-1}\right)_{ij}
+%\end{equation*}
+
+%\subsubsection{Discretized integral kernel}
+%
+%The interpolation of $\Phi$ onto $V_h \otimes V_h$, which we denote by $\Phi^h$, is given as follows:
+%\begin{equation}
+%\label{eq:defn_of_Akerpcmesh}
+%\Phi(y,x) \approx \Phi^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \Phi(p_i,p_j) \febasis_i(y) \febasis_j(x).
+%\end{equation}
+%Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations, yields the following approximation:
+%\begin{equation}
+%\label{eq:kernel_representation_pcmesh}
+%(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
+%= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u},
+%\end{equation}
+%where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
+%\begin{equation}
+%\label{eq:Akerpcmat_entries}
+%\mathbf{\Phi}_{ij} := \Phi(p_i, p_j),
+%\end{equation}
+%Based on \eqref{eq:kernel_representation_pcmesh}, we define the following matrix approximation, $\widetilde{\mathbf{A}} \approx \mathbf{A}$ as follows:
+%\begin{equation}
+%\label{eq:Amatpc_defn}
+%\widetilde{\mathbf{A}} = \bm{M} \mathbf{\Phi} \bm{M}.
+%\end{equation}
+%Note that $\widetilde{\mathbf{A}}$ differs slightly from $\mathbf{A}$ (the Galerkin projection of $\Aop$ onto $V_h$) because of the approximation $\Phi^h \approx \Phi$. 
+%%Hence, $\mathbf{A}$ may differ slightly from the implicitly defined matrix that performs the mapping $\mathbf{u} \rightarrow \mathbf{v}$ corresponding to the discrete version of the operator action $v=\Aop u$.
+%%Because $\Aop$ is only available matrix-free and $\Phi$ is only defined implicitly in terms of $\Aop$, 
+%The matrices $\widetilde{\mathbf{A}}$ and $\mathbf{\Phi}$ are, of course, not easily accessible.  
+%%The primary goal of this paper is building an H-matrix approximation $\mathbf{A}_H$ of $\widetilde{\mathbf{A}}$, using a small number of operator actions of $\Aop$ and $\Aop^T$.
+%
+%%After discretization with finite elements\footnote{While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead, with minor modifications to our method.} (Section \ref{sec:finite_element_kernel}), $\Aop$ and $\Phi$ become large dense matrices $\mathbf{A}$ and $\mathbf{\Phi}$, respectively.
 
@@
 At the continuum level, the integral kernel $\Phi$ is assumed to be non-negative.
-However, per \eqref{eq:discrete_kernel}, when we compute impulse responses we are effectively operating with a discrete integral kernel which has $(y,x)$ entry
-\begin{equation*}
-\boldsymbol{\delta}_y^T \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x.
-\end{equation*}
-Even if the discretization $\Aop \rightarrow \mathbf{A}$ preserves integral kernel positivity, the inverse mass matrices $\mathbf{M}^{-1}$ will typically contain negative numbers.
+However, when we compute impulse responses we are effectively operating with the discrete integral kernel shown in \eqref{eq:approx_kernel_entry_81}.
+%,  which has $(y,x)$ entry
+%\begin{equation*}
+%\boldsymbol{\delta}_y^T \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x.
+%\end{equation*}
+%Even if the discretization $\Aop \rightarrow \mathbf{A}$ preserves integral kernel non-negativity (which it typically will not), 
+The inverse mass matrices in this formula, $\mathbf{M}^{-1}$, will typically contain negative numbers.
 If there are too many negative numbers, or if the negative numbers are too large, our algorithm will be less robust.
 We therefore recommend replacing the mass matrix $\mathbf{M}$ with a positive diagonal \emph{lumped mass} approximation, and using the resulting lumped mass matrix everywhere throughout the approximation method.
-For typical finite element meshes, lumped mass matrix approximations are typically spectrally equivalent to the mass matrix \cite{MASSLUMP2}.
+For typical finite element meshes, lumped mass matrix approximations are spectrally equivalent to the mass matrix \cite{MASSLUMP2}.
 In our numerical results, we use the lumped mass matrix constructed by replacing off-diagonal entries of the mass matrix by zero.
@@
 
-We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming H-matrix representations $\mathbf{\Phi}_H$ and $\bm{M}_H$ of $\mathbf{\Phi}$ and $\bm{M}$, respectively, then using fast H-matrix methods to multiply these matrices per \eqref{eq:Amatpc_defn} to form
+We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming H-matrix representations $\mathbf{\Phi}_H$ and $\bm{M}_H$ of $\mathbf{\Phi}$ and $\bm{M}$, respectively, then using fast H-matrix methods to multiply these matrices per \eqref{eq:boldA} to form
 \begin{equation*}

### 2022-03-25 paper/localpsf.tex
fix typo

@@
 \begin{equation*}
-\left(\boldsymbol{\rho}_h\right)_i = \rho_h(\febasis_i), \qquad i=1,\dots,N. 
+\boldsymbol{\rho}_i = \rho_h(\febasis_i), \qquad i=1,\dots,N. 
 \end{equation*}
@@
 Hence, methods for approximating $\mathbf{A}$ are computationally intractable if they require accessing scattered matrix entries from many different rows and columns of $\mathbf{A}$.
+
 The operator $A_h:V_h \rightarrow V_h'$ can be written in integral kernel form as
@@
 \end{equation*}
-to be the spatially varying mean and covariance of the normalized version of $\impulseresponse_x$, respectively.
+be the spatially varying mean and covariance of the normalized version of $\impulseresponse_x$, respectively.
 Imagine $\phi_x$ as a mound of dirt piled on a localized region in $\Omega$, with $\phi_x(y)$ being the height of the dirt mound at location $y$.

### 2022-03-25 paper/localpsf.tex
minor changes

@@
 We assume that the finite element basis satisfies the Kronecker property, i.e., $\febasis_i(p_i)=1$ and $\febasis_i(p_j)=0$ if $i\neq j$.
-For $u_h \in V_h \subset L^2(\Omega)$ we write $\bm{u} \in \mathbb{R}^N_\mathbf{M}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$, i.e.,
+For $u_h \in V_h$ we write $\bm{u} \in \mathbb{R}^N_\mathbf{M}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$, i.e.,
 \begin{equation}
@@
 \end{equation}
-Linear functionals $\rho_h \in V_h' \supset L^2(\Omega)'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^N_{\mathbf{M}^{-1}}$, with entries 
+Linear functionals $\rho_h \in V_h'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^N_{\mathbf{M}^{-1}}$, with entries 
 \begin{equation*}

### 2022-03-28 paper/localpsf.tex
about to reorganize intro based on Noemi's suggestions

@@
 	}
-	\caption{Construct product-convolution H-matrix}
+	\caption{Construct H-matrix approximation}
 	\label{alg:construct_Atilde}
@@
 
-	\Input{Finite set of candidate points $P\subset \Omega$, \\spatially varying mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\domain boundary information $\partial \Omega$, \\previous sample point batches $\pointbatch_1, \dots, \pointbatch_{k-1}$}
+	\Input{Finite set of candidate points $P\subset \Omega$, \\spatially varying mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\previous sample point batches $\pointbatch_1, \dots, \pointbatch_{k-1}$}
 	\Output{Batch of new sample points $\pointbatch_k$}
@@
 
-		\If{$E_p \cap E_q \neq \{\}$ for all $q \in \pointbatch_k$ and $E_p \cap \partial \Omega = \{\}$}{
+		\If{$E_p \cap E_q \neq \{\}$ for all $q \in \pointbatch_k$}{
 			\tcp{$E_p$ and $E_q$ are the ellipsoids defined in \eqref{eq:support_ellipsoid}}

### 2022-03-28 paper/localpsf.tex
reorganizing intro based on Noemi's suggestions

@@
 
-We present a fast matrix free method for approximating operators that have locally supported non-negative integral kernels.
+We present a fast \emph{matrix free} method for approximating operators, $\Aop$, that have locally supported non-negative integral kernels.
 Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems.
-
-Let $\Omega \subset \mathbb{R}^d$ be a bounded domain (in applications typically $\gdim=1$, $2$, or $3$) and let $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the following form:
-\begin{equation}
-	\label{eq:kernel_representation}
-	(\Aop u)(v) := \int_\Omega \int_\Omega v(y) \Aker(y,x) u(x) dx dy.
-\end{equation}
-The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, $\Aop u \in L^2(\Omega)'$ is the linear functional that results from applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
-We seek to approximate operators $\Aop$ that are accessible \emph{matrix-free}.
-That is, we have a black-box computational procedure through which we may evaluate the maps
-\begin{equation*}
-u \mapsto\Aop u \qquad \text{and} \qquad v \mapsto \Aop^Tv
-\end{equation*} 
-for arbitrary functions $u$ and $v$, but we cannot easily compute kernel entries $\Aker(y,x)$.
+By ``matrix free,'' we mean that we have a black-box computational procedure through which we may apply $\Aop$ and $\Aop^T$ to arbitrary functions,
+\begin{equation}
+\label{eq:eval_maps}
+u \mapsto\Aop u \qquad \text{and} \qquad v \mapsto \Aop^Tv,
+\end{equation} 
+but we do not have any further access to $\Aop$.
+In particular, we cannot easily access entries of $\Aop$'s integral kernel.
+In applications, evaluating the maps in \eqref{eq:eval_maps} may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
+After discretization, $\Aop$ becomes a dense matrix that is typically too large to be built and stored.
+This property is called ``matrix-free'' because at the discrete level it means one can compute matrix-vector products with the matrix representation of $\Aop$, but cannot easily compute individual matrix entries.
+
+%We seek to approximate operators $\Aop$ that are accessible \emph{matrix-free}. That is, we have a black-box computational procedure through which we may evaluate the maps
+%\begin{equation*}
+%u \mapsto\Aop u \qquad \text{and} \qquad v \mapsto \Aop^Tv
+%\end{equation*} 
+%for arbitrary functions $u$ and $v$, but we cannot easily compute kernel entries $\Aker(y,x)$.
 In applications, evaluating these maps may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
@@
 
-Our method, described in Section \ref{sec:method}, belongs to a class of methods that approximate an operator by interpolating impulse responses of the operator associated with a scattered collection of points.
-The \emph{impulse response} of $\Aop$ associated with the point $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$,
-\begin{equation}
-\label{eq:impulse_response_defn}
-\impulseresponse_x(y) := \Aker(y, x),
-\end{equation} 
-which results from keeping $x$ fixed and viewing $\Phi(y,x)$ as a function of $y$.
-The function $\impulseresponse_x$ is called the impulse response because straightforward analysis shows that it may be expressed as follows:
-\begin{equation}
-\label{eq:impulse_response_delta_action}
-\impulseresponse_x = \left( \Aop \delta_x^* \right)^*,
-\end{equation}
-where $\delta_x$ is the the delta distribution (i.e., point source, impulse) centered at $x$, $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\sigma \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\sigma^* \in L^2(\Omega)$ such that $\sigma(v) = \left(\sigma^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
-This is described in Appendix \ref{app:distributions}.
-
-We assume:
-\begin{itemize}
-	\item The supports of the impulse responses $\impulseresponse_x$ are contained (or approximately contained) in localized regions.
-	\item The integral kernel is non-negative in the sense that $$\Phi(y,x) \ge 0$$ for all (or nearly all) $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
-The operator $\Aop$ need not be positive semi-definite to use our method, and positive semi-definite operators need not have a non-negative integral kernel.}
-\end{itemize}
-Because of these assumptions, we are able to estimate the supports of the impulse responses a-priori by applying $\Aop$ to a small number of polynomial functions.
-We use the support estimates to compute scattered impulse responses in batches (see Figure \ref{fig:batches_intro}).
-This batching allows us to recover many impulse responses using a small number of operator actions $u \mapsto \Aop u$.
-The support estimates also allow us to interpolate the impulse responses with a new more sophisticated and accurate interpolation scheme.
+The output of our method is a hierarchical matrix (H-matrix).
+H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank.
+For more details about H-matrices, see Appendix \ref{app:h_matrix}.
+H-matrix methods permit us to perform matrix-vector products extremely cheaply.
+Furthermore, one may use H-matrix methods to perform useful linear algebra operations involving the H-matrix that cannot easily be done using the original operator.
+These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion.
+These H-matrix methods are fast and scale well to large problems.
+The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
+
+Since we are building an H-matrix, it is natural to ask why we do not build the H-matrix directly.
+The reason is that \emph{classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products}.
+The basic idea behind our method is constructing the H-matrix via impulse response interpolation.
+The impulse response, $\phi_x$, associated with a point $x$ is the function that results from applying $\Aop$ to a delta distribution (point source, impulse) centered at $x$.
+We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with collections of points scattered throughout the domain (see Figure \ref{fig:batches_intro}).
+Then we interpolate these impulse responses to evaluate arbitrary entries of the operator's integral kernel.
+This allows fast evaluation of approximate matrix entries needed by classical H-matrix construction methods.
 
@@
 Dashed gray ellipses indicate the estimated support ellipsoids.
-The large circle is the boundary of $\Omega$.
-Early batches tend to have lots of ellipsoids near the boundary, as compared to later batches.
-This is because our greedy ellipsoid packing algorithm tries to pick sample points that are far away from sample points in previously chosen batches.}
+The large circle is the boundary of the domain.
+%		Early batches tend to have lots of ellipsoids near the boundary, as compared to later batches. 
+%		This is because our greedy ellipsoid packing algorithm tries to pick sample points that are far away from sample points in previously chosen batches.
+	}
 	\label{fig:batches_intro}
@@
 
-The output of our method is a hierarchical matrix (H-matrix) approximation, $\mathbf{A}_H$, of a discretization of $\Aop$.
-H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank.
-For more details about H-matrices, see Appendix \ref{app:h_matrix}.
-While operator actions of $\Aop$ on vectors are typically computationally expensive, fast H-matrix methods permit us to perform matrix-vector products with $\mathbf{A}_H$ cheaply.
-Furthermore, one may use H-matrix methods to perform useful linear algebra operations involving $\mathbf{A}_H$ that cannot easily be done using $\Aop$, because $\Aop$ is only available matrix-free.
-These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion.
-These H-matrix methods are fast and scale well to large problems.
-The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
-
-\subsection{Motivation: PDE constrained inverse problems}
-\label{sec:PDE_hessian_motivation}
+%\subsection{Motivation: PDE constrained inverse problems}
+%\label{sec:PDE_hessian_motivation}
 
@@
 The objective function in optimization problem \eqref{eq:minimization_problem} is defined indirectly via the implicit function theorem.
-As a result, information about the Hessian of the objective function, $\mathcal{H}$, is only accessible via the action of $\mathcal{H}$ action on vectors.
-In particular, applying $\mathcal{H}$ to a vector requires solving two PDEs \cite{HESSIANACTION}.
-More accessible approximations of $\mathcal{H}$ are highly desirable, because they allow for fast solution of \eqref{eq:minimization_problem} via Newton-type methods.
-More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because $\mathcal{H}^{-1}$ locally approximates the Bayesian posterior covariance for $m$.
-
-\subsection{Background}
+As a result, information about the Hessian of the objective function is only accessible via the action of the Hessian on vectors.
+In particular, applying the Hessian to a vector requires solving two PDEs \cite{HESSIANACTION}.
+More accessible approximations of the Hessian are highly desirable, because they allow for fast solution of \eqref{eq:minimization_problem} via Newton-type methods.
+More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because inverse of the Hessian locally approximates the Bayesian posterior covariance for $m$.
+
+%\subsection{Background}
 
@@
 
-Since we are building an H-matrix approximation of $\Aop$, it is natural to ask why we compute impulse responses and interpolate them.
-Why not build the H-matrix directly?
-The reason is that \emph{classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products}.
-The impulse response interpolation allows fast evaluation of approximate matrix entries needed by the classical H-matrix construction methods.
 Latest developments and improvements in matrix-free H-matrix construction and approximation are based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}.
@@
 Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods based on the peeling process more feasible in the future.
+
+\section{Setting}
+
+Let $\Omega \subset \mathbb{R}^d$ be a bounded domain (in applications typically $\gdim=1$, $2$, or $3$) and let $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the following form:
+\begin{equation}
+	\label{eq:kernel_representation}
+	(\Aop u)(v) := \int_\Omega \int_\Omega v(y) \Aker(y,x) u(x) dx dy.
+\end{equation}
+The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, $\Aop u \in L^2(\Omega)'$ is the linear functional that results from applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
+
+%We seek to approximate operators $\Aop$ that are accessible \emph{matrix-free}. That is, we have a black-box computational procedure through which we may evaluate the maps
+%\begin{equation*}
+%u \mapsto\Aop u \qquad \text{and} \qquad v \mapsto \Aop^Tv
+%\end{equation*} 
+%for arbitrary functions $u$ and $v$, but we cannot easily compute kernel entries $\Aker(y,x)$.
+In applications, evaluating these maps may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
+After discretization, $\Aop$ becomes a dense matrix that is typically too large to be built and stored.
+This property is called ``matrix-free'' because at the discrete level it means one can compute matrix-vector products with the matrix representation of $\Aop$, but cannot easily compute individual matrix entries.
+
+Our method, described in Section \ref{sec:method}, belongs to a class of methods that approximate an operator by interpolating impulse responses of the operator associated with a scattered collection of points.
+The \emph{impulse response} of $\Aop$ associated with the point $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$,
+\begin{equation}
+\label{eq:impulse_response_defn}
+\impulseresponse_x(y) := \Aker(y, x),
+\end{equation} 
+which results from keeping $x$ fixed and viewing $\Phi(y,x)$ as a function of $y$.
+The function $\impulseresponse_x$ is called the impulse response because straightforward analysis shows that it may be expressed as follows:
+\begin{equation}
+\label{eq:impulse_response_delta_action}
+\impulseresponse_x = \left( \Aop \delta_x^* \right)^*,
+\end{equation}
+where $\delta_x$ is the the delta distribution (i.e., point source, impulse) centered at $x$, $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\sigma \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\sigma^* \in L^2(\Omega)$ such that $\sigma(v) = \left(\sigma^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
+This is described in Appendix \ref{app:distributions}.
+
+We assume:
+\begin{itemize}
+	\item The supports of the impulse responses $\impulseresponse_x$ are contained (or approximately contained) in localized regions.
+	\item The integral kernel is non-negative in the sense that $$\Phi(y,x) \ge 0$$ for all (or nearly all) $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
+The operator $\Aop$ need not be positive semi-definite to use our method, and positive semi-definite operators need not have a non-negative integral kernel.}
+\end{itemize}
+Because of these assumptions, we are able to estimate the supports of the impulse responses a-priori by applying $\Aop$ to a small number of polynomial functions.
+We use the support estimates to compute scattered impulse responses in batches (see Figure \ref{fig:batches_intro}).
+This batching allows us to recover many impulse responses using a small number of operator actions $u \mapsto \Aop u$.
+The support estimates also allow us to interpolate the impulse responses with a new more sophisticated and accurate interpolation scheme.
 

### 2022-03-28 paper/localpsf.tex
reorganizing intro based on Noemi's suggestions

@@
 
-Let $\Omega \subset \mathbb{R}^d$ be a bounded domain (in applications typically $\gdim=1$, $2$, or $3$) and let $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the following form:
+Let $\Omega \subset \mathbb{R}^d$ be a bounded domain (in applications typically $\gdim=1$, $2$, or $3$).
+We seek to approximate integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ that may be written in the following form:
 \begin{equation}
@@
 \end{equation}
-The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, $\Aop u \in L^2(\Omega)'$ is the linear functional that results from applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
+The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, which is not easily accessible to us.
+The linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
+Here $L^2(\Omega)'$ denotes the topological dual of $L^2(\Omega)$ (space of all continuous linear functionals).
 
@@
 
-Our method, described in Section \ref{sec:method}, belongs to a class of methods that approximate an operator by interpolating impulse responses of the operator associated with a scattered collection of points.
+%Our method, described in Section \ref{sec:method}, belongs to a class of methods that approximate an operator by interpolating impulse responses of the operator associated with a scattered collection of points.
 The \emph{impulse response} of $\Aop$ associated with the point $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$, 
@@
 
-We assume:
-\begin{itemize}
-	\item The supports of the impulse responses $\impulseresponse_x$ are contained (or approximately contained) in localized regions.
-	\item The integral kernel is non-negative in the sense that $$\Phi(y,x) \ge 0$$ for all (or nearly all) $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
+We assume (1) the supports of the impulse responses $\impulseresponse_x$ are contained (or approximately contained) in localized regions, and (2) the integral kernel is non-negative in the sense that $$\Phi(y,x) \ge 0$$ for all (or nearly all) $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
 The operator $\Aop$ need not be positive semi-definite to use our method, and positive semi-definite operators need not have a non-negative integral kernel.}
-\end{itemize}
 Because of these assumptions, we are able to estimate the supports of the impulse responses a-priori by applying $\Aop$ to a small number of polynomial functions.
@@
 We assume that the finite element basis satisfies the Kronecker property, i.e., $\febasis_i(p_i)=1$ and $\febasis_i(p_j)=0$ if $i\neq j$.
-For $u_h \in V_h$ we write $\bm{u} \in \mathbb{R}^N_\mathbf{M}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$, i.e.,
+For $u_h \in V_h$ we write $\bm{u} \in \mathbb{R}^N_\mathbf{M}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis, i.e.,
 \begin{equation}

### 2022-03-29 paper/localpsf.tex
reorganizing intro based on Noemi's suggestions

@@
 For more details about H-matrices, see Appendix \ref{app:h_matrix}.
-H-matrix methods permit us to perform matrix-vector products extremely cheaply.
+H-matrix methods permit us to perform matrix-vector products cheaply.
 Furthermore, one may use H-matrix methods to perform useful linear algebra operations involving the H-matrix that cannot easily be done using the original operator.
@@
 Then we interpolate these impulse responses to evaluate arbitrary entries of the operator's integral kernel.
-This allows fast evaluation of approximate matrix entries needed by classical H-matrix construction methods.
+This allows fast evaluation of approximate matrix entries, which are needed by the classical H-matrix construction methods.
 
@@
 As a result, information about the Hessian of the objective function is only accessible via the action of the Hessian on vectors.
-In particular, applying the Hessian to a vector requires solving two PDEs \cite{HESSIANACTION}.
+In particular, applying the Hessian to a vector requires solving two linear PDEs which take the form of linearizations of \eqref{eq:state_pde} \cite{HESSIANACTION}.
 More accessible approximations of the Hessian are highly desirable, because they allow for fast solution of \eqref{eq:minimization_problem} via Newton-type methods.
-More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because inverse of the Hessian locally approximates the Bayesian posterior covariance for $m$.
+More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because inverse of the Hessian locally approximates the Bayesian posterior covariance for $m$ \cite{LAPLACEAPPROXIMATION}.
 
@@
 While not commonly done, in theory linear algebra operations involving convolution-product approximations (transposes of PC approximations) could be done using fast methods for pseudo-differential operator symbol calculus \cite{YINGDEMANET} since these are equivalent to low rank approximation of the pseudo-differential symbol of the operator.
-There are other relevant matrix-free operator approximation methods based on wavelets \cite{HERRMANN} and pseudo-differential operators \cite{DEMANET}, however these do not use PC approximation.
+We also note other relevant matrix-free operator approximation methods based on wavelets \cite{HERRMANN} and pseudo-differential operators \cite{DEMANET} which do not use PC approximation.
 
@@
 
-Latest developments and improvements in matrix-free H-matrix construction and approximation are based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}.
+Latest developments and improvements in matrix-free H-matrix construction and approximation are based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}, which we do not use.
 These methods have, for instance, been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
-One issue with this approach is that, while these matrix-free H-matrix construction methods are asymptotically scalable in theory, in practice the required number of matrix vector products is large\footnote{DO CALCULATION}.
+One issue with this approach is that, while these methods are asymptotically scalable in theory, in practice the required number of matrix vector products is large\footnote{DO CALCULATION}.
 Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods based on the peeling process more feasible in the future.
@@
 \end{equation}
-The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, which is not easily accessible to us.
+The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, which exists in principle but is not easily accessible to us.
+Here $L^2(\Omega)'$ denotes the topological dual of $L^2(\Omega)$.
 The linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
-Here $L^2(\Omega)'$ denotes the topological dual of $L^2(\Omega)$ (space of all continuous linear functionals).
+
+The action of $\Aop$ may be extended to distributions if the kernel $\Phi$ is sufficiently regular.
+We derive the extension formally as follows.
+Suppose $\rho \in L^2(\Omega)'$, and let $\rho^* \in L^2(\Omega)$ denote the Riesz representative\footnote{Recall that the Riesz representative of a functional $\sigma \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\sigma^* \in L^2(\Omega)$ such that $\sigma(v) = \left(\sigma^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} of $\rho$ with respect to the $L^2(\Omega)$ inner product.
+We have
+\begin{align*}
+\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Phi(y,x) \genericdistribution^*(x) dx dy \\
+&= \int_\Omega v(y) \int_\Omega \Phi(y,x) \genericdistribution^*(x) dx \\
+&= \int_\Omega v(y) \genericdistribution\left(\Phi(y,\cdot)\right) dy 
+%= \left(v,\genericdistribution\left(\Phi(y,\cdot)\right)\right)_{L^2(\Omega)},
+\end{align*}
+where $\Phi(y,\cdot)$ denotes the function $x \mapsto \Phi(y,x)$.
+Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions.
+If instead $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ is a distribution, then the Riesz representative of $\rho$ may not exist, so the above derivation does not make sense.
+However, if $\Phi$ is sufficiently regular such that the function $y \mapsto \rho\left(\Phi(y,~\cdot~)\right)$ is well-defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the last line in the derivation is well-defined.
+Hence, we use the last line of this derivation to \emph{define} the action of $\Aop$ on the distribution $\rho$ as follows:
+\begin{equation}
+\label{eq:action_on_distribution}
+\left(\Aop \genericdistribution^*\right)(v) := \int_\Omega v(y) \genericdistribution\left(\Phi(y, \cdot)\right) dy.
+\end{equation}
+Here ``$\Aop \genericdistribution^*$'' is an abuse of notation that means the action of $\Aop$ on the distribution $\rho$, per this definition.
+
+%We assume that $\Phi$ is sufficiently regular such that the action of $\Aop$ on delta distributions is well-defined, because
+
+%This is derived formally as follows:
+%\begin{align*}
+%\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Phi(y,x) \textrm{``} \genericdistribution^*(x) \textrm{''} dx dy \\
+%&= \int_\Omega v(y) \int_\Omega \Phi(y,x) \textrm{``} \genericdistribution^*(x) \textrm{''} dx \\
+%&= \int_\Omega v(y) \genericdistribution\left(\Phi(y,\cdot)\right) dx.
+%\end{align*}
+%Here $\textrm{``}\rho^*\textrm{''}$ formally denotes the Riesz representative of $\rho$, which, of course, does not exist if $\rho$ is a pure distribution.
+
+%an abuse of notation which views $\rho$ as a ``function'' of $x$, rather than a linear functional.
+That is, $\rho^*$ would be the Riesz representative of $\rho$, but for the obstruction that this Riesz representative does not exist on all of $L^2(\Omega)$.
+
+%For example, the delta distribution $\delta_x$ is defined by $\delta_x(v) = v(x)$, and the action of $\Aop$ on $\delta_x$ is given by
+%\begin{equation}
+%\label{eq:appendix_impulse_response}
+%\left(\Aop \delta_x^*\right)(v)	= \int_\Omega v(y) \Phi(y,x) dy.
+%\end{equation}
+%Recall that the impulse response of $\Aop$ to $\delta_x$, denoted $\impulseresponse_x$, is the Riesz representation of $\Aop \delta_x^*$. 
+%From \eqref{eq:appendix_impulse_response} and the definition of the Riesz representation, we have $\impulseresponse_x = A(~\cdot~,x)$.
+That is, $\impulseresponse_x$ is the function $y \mapsto A(y,x)$.
 
@@
 %Our method, described in Section \ref{sec:method}, belongs to a class of methods that approximate an operator by interpolating impulse responses of the operator associated with a scattered collection of points.
+Let $\delta_x$ denote the delta distribution\footnote{Recall that the delta distribution $\delta_x:\mathcal{D}(\Omega)\rightarrow \mathbb{R}$ is the linear functional defined by $\delta_x(v) = v(x)$ for all $v\in \mathcal{D}(\Omega)$.} (i.e., point source, impulse) $\delta_x$, centered at the point $x$.
 The \emph{impulse response} of $\Aop$ associated with the point $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$, 
+\begin{equation}
+\label{eq:impulse_response_delta_action}
+\impulseresponse_x = \left( \Aop \delta_x^* \right)^*,
+\end{equation}
+that is formed by applying $\Aop$ to $\delta_x$, then taking the Riesz representation of the result.
+Using \eqref{eq:action_on_distribution} and the definition of the delta distribution, we also see that $\phi_x$ is the function
 \begin{equation}
@@
 which results from keeping $x$ fixed and viewing $\Phi(y,x)$ as a function of $y$.
-The function $\impulseresponse_x$ is called the impulse response because straightforward analysis shows that it may be expressed as follows:
-\begin{equation}
-\label{eq:impulse_response_delta_action}
-\impulseresponse_x = \left( \Aop \delta_x^* \right)^*,
-\end{equation}
-where $\delta_x$ is the the delta distribution (i.e., point source, impulse) centered at $x$, $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\sigma \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\sigma^* \in L^2(\Omega)$ such that $\sigma(v) = \left(\sigma^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
+
+%The function $\impulseresponse_x$ is called the impulse response because, using \eqref{eq:action_on_distribution} and the definition of the delta function, we have:
+
+%where $\delta_x$ is the the delta distribution (i.e., point source, impulse) centered at $x$, $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.
+
+%Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
 This is described in Appendix \ref{app:distributions}.
 
-We assume (1) the supports of the impulse responses $\impulseresponse_x$ are contained (or approximately contained) in localized regions, and (2) the integral kernel is non-negative in the sense that $$\Phi(y,x) \ge 0$$ for all (or nearly all) $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
+In this paper, we focus on approximating operators that satisfy the following conditions: 
+\begin{enumerate}
+	\item The kernel $\Phi$ is sufficiently regular such that $\phi_x$ is a well-defined function for almost all $x\in \Omega$.
+	\item The supports of the impulse responses $\impulseresponse_x$ are contained (or approximately contained) in localized regions.
+	\item The integral kernel is non-negative in the sense that $$\Phi(y,x) \ge 0$$ for all (or nearly all) $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
 The operator $\Aop$ need not be positive semi-definite to use our method, and positive semi-definite operators need not have a non-negative integral kernel.}
-Because of these assumptions, we are able to estimate the supports of the impulse responses a-priori by applying $\Aop$ to a small number of polynomial functions.
-We use the support estimates to compute scattered impulse responses in batches (see Figure \ref{fig:batches_intro}).
+\end{enumerate}
+Because of these conditions, we are able to estimate the supports of the impulse responses a-priori by applying $\Aop^T$ \footnote{Recall that the transpose of $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ satisfying $\left(\Aop u\right)(v) = \left(\Aop^T v\right)(u)$ for all $u,v\in L^2(\Omega)$.} to a small number of polynomial functions (Section \ref{sec:intromoments}).
+We use the support estimates to compute scattered impulse responses in batches (Section \ref{sec:get_impulse_response}, also Figure \ref{fig:batches_intro}).
 This batching allows us to recover many impulse responses using a small number of operator actions $u \mapsto \Aop u$.
-The support estimates also allow us to interpolate the impulse responses with a new more sophisticated and accurate interpolation scheme.
+The support estimates also allow us to interpolate the impulse responses with a new more sophisticated and accurate interpolation scheme (Sections \ref{sec:local_mean_displacement_invariance} and \ref{sec:approximate_kernel_entries}).
 
@@
 the space $\mathbb{R}^N_\mathbf{M}$ is $\mathbb{R}^N$ with the inner product $(\mathbf{u},\mathbf{v})_\mathbf{M} := \mathbf{u}^T \mathbf{M} \mathbf{v}$, and $\mathbb{R}^N_{\mathbf{M}^{-1}}$ is the analogous space with $\mathbf{M}^{-1}$ replacing $\mathbf{M}$.
-Direct calculation shows that $\mathbf{R}^N_\mathbf{M}$ and $\mathbf{R}^N_{\mathbf{M}^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively.
+Direct calculation shows that $\mathbb{R}^N_\mathbf{M}$ and $\mathbb{R}^N_{\mathbf{M}^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively.
 %, in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u}, \mathbf{v}\right)_\mathbf{M}$, $\left(\rho_h, \nu_h\right)_{L^2(\Omega)} = \left(\boldsymbol{\rho}, \boldsymbol{\nu}\right)_{\mathbf{M}^{-1}}$, and $\rho_h(u) = \boldsymbol{\rho}^T\mathbf{u}$, for $u_h,v_h \in V_h$ and $\rho_h, \nu_h \in V_h'$.
@@
 What is important is that we do not have direct access to matrix entries $\mathbf{A}_{ij}$.
-Rather, we have a computational procedure that allows us to compute matrix-vector products $\mathbf{u}\mapsto \mathbf{A}\mathbf{u}$, and computing these matrix-vector products is costly.
+Rather, we have a computational procedure that allows us to compute matrix-vector products $\mathbf{u}\mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{v}\mapsto \mathbf{A}^T\mathbf{v}$, and computing these matrix-vector products is costly.
 %, and each of these matrix-vector products is costly to compute. 
@@
 
-\section{Method}
+\section{Operator approximation algorithm}
 \label{sec:method}

### 2022-03-29 paper/localpsf.tex
reorganizing intro based on Noemi's suggestions
add hmatrix figure. move figures to intro

@@
 
-We present a fast \emph{matrix free} method for approximating operators, $\Aop$, that have locally supported non-negative integral kernels.
-Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems.
+We present a fast \emph{matrix free} method for approximating operators, $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$, that have locally supported non-negative integral kernels.
+Here $\Omega \subset \mathbb{R}^d$ is a bounded domain, $d$ is the spatial dimension, and $L^2(\Omega)'$ is the topological dual space of $L^2(\Omega)$.
+Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps) \cite{DTNMAP}, covariance operators in spatial statistics \cite{COVARIANCEOPERATORS}, blurring operators in imaging \cite{BLURRING}, and Hessians in distributed parameter PDE-constrained optimization and inverse problems \cite{PDEHESSIAN}.
+
 By ``matrix free,'' we mean that we have a black-box computational procedure through which we may apply $\Aop$ and $\Aop^T$ to arbitrary functions,
@@
 The output of our method is a hierarchical matrix (H-matrix).
-H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank.
+H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank (see Figure \ref{fig:hmatrix_intro}).
 For more details about H-matrices, see Appendix \ref{app:h_matrix}.
@@
 The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
-
 Since we are building an H-matrix, it is natural to ask why we do not build the H-matrix directly.
 The reason is that \emph{classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products}.
-The basic idea behind our method is constructing the H-matrix via impulse response interpolation.
-The impulse response, $\phi_x$, associated with a point $x$ is the function that results from applying $\Aop$ to a delta distribution (point source, impulse) centered at $x$.
-We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with collections of points scattered throughout the domain (see Figure \ref{fig:batches_intro}).
-Then we interpolate these impulse responses to evaluate arbitrary entries of the operator's integral kernel.
-This allows fast evaluation of approximate matrix entries, which are needed by the classical H-matrix construction methods.
 
@@
 
+The basic idea behind our method is impulse response interpolation.
+The impulse response, $\phi_x$, associated with a point $x$ is the function that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
+We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with collections of points scattered throughout the domain (see Figure \ref{fig:batches_intro}).
+Then we interpolate these impulse responses to approximate arbitrary entries of the operator's integral kernel.
+This allows fast evaluation of approximate matrix entries, which are needed by the classical H-matrix construction methods.
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale=0.5]{bct1.eps}
+	\end{center}
+	\caption{Hierarchical matrix block structure for the Stokes Hessian. Red blocks are dense. All other blocks are low rank.
+	}
+	\label{fig:hmatrix_intro}
+\end{figure}
+
 %\subsection{Motivation: PDE constrained inverse problems}
@@
 
-The method proposed in this paper is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i \in \Omega$ \cite{PRODCONVLIST}.
+The method proposed in this paper is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of scattered points $x_i \in \Omega$ \cite{PRODCONVLIST}.
 A common choice for PC methods in this class is to choose the $x_i$ to be nodes in a regular grid \cite{NAGY}.
@@
 In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
-However, even with adaptive refinement, many $x_i$, and hence operator actions of $\Aop$, are required.
+However, computing each impulse response required one action of $\Aop$.
+Even with adaptive refinement, many $x_i$, and hence operator actions of $\Aop$, were required.
 
@@
 
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale=1.0]{mean_displacement_invariance.pdf}
+	\end{center}
+	\caption{An operator is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \mu(x) + \mu(x')\right)$ when $x$ is close to $x'$.}
+	\label{fig:mean_displacement_invariance}
+\end{figure}
+
 In this paper, we build on previous work \cite{PRODCONVMYPAPER}, in which we converted PC approximations into H-matrix format.
@@
 \begin{itemize}
-	\item PC approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, while the method we propose is based on an assumption that we call ``normalized local mean displacement invariance.'' This is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
+	\item PC approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, while the method we propose is based on an assumption that we call ``normalized local mean displacement invariance'' (see Figure \ref{fig:mean_displacement_invariance}).
+This is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
 	\item PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulse responses.
@@
 
-Latest developments and improvements in matrix-free H-matrix construction and approximation are based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}, which we do not use.
+Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}, which we do not use.
 These methods have, for instance, been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
 One issue with this approach is that, while these methods are asymptotically scalable in theory, in practice the required number of matrix vector products is large\footnote{DO CALCULATION}.
-Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods based on the peeling process more feasible in the future.
+Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods based on the peeling process more efficient in the future.
 
@@
 The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, which exists in principle but is not easily accessible to us. 
-Here $L^2(\Omega)'$ denotes the topological dual of $L^2(\Omega)$.
+%We write $L^2(\Omega)'$ to denote the topological dual of $L^2(\Omega)$. 
 The linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
@@
 The action of $\Aop$ may be extended to distributions if the kernel $\Phi$ is sufficiently regular.
-We derive the extension formally as follows.
+We derive the extension as follows.
 Suppose $\rho \in L^2(\Omega)'$, and let $\rho^* \in L^2(\Omega)$ denote the Riesz representative\footnote{Recall that the Riesz representative of a functional $\sigma \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\sigma^* \in L^2(\Omega)$ such that $\sigma(v) = \left(\sigma^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} of $\rho$ with respect to the $L^2(\Omega)$ inner product.
 We have
-\begin{align*}
-\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Phi(y,x) \genericdistribution^*(x) dx dy \\
-&= \int_\Omega v(y) \int_\Omega \Phi(y,x) \genericdistribution^*(x) dx \\
-&= \int_\Omega v(y) \genericdistribution\left(\Phi(y,\cdot)\right) dy 
-%= \left(v,\genericdistribution\left(\Phi(y,\cdot)\right)\right)_{L^2(\Omega)},
-\end{align*}
+\begin{subequations}
+	\label{eq:extension_to_distributions}
+	\begin{align}
+		\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Phi(y,x) \genericdistribution^*(x) dx ~dy \\
+		&= \int_\Omega v(y) \int_\Omega \Phi(y,x) \genericdistribution^*(x) dx ~dy \\
+		&= \int_\Omega v(y) \genericdistribution\left(\Phi(y,\cdot)\right) dy,
+	\end{align}
+\end{subequations}
+%\begin{align*}
+%\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Phi(y,x) \genericdistribution^*(x) dx dy \\
+%&= \int_\Omega v(y) \int_\Omega \Phi(y,x) \genericdistribution^*(x) dx \\
+%&= \int_\Omega v(y) \genericdistribution\left(\Phi(y,\cdot)\right) dy 
+%%= \left(v,\genericdistribution\left(\Phi(y,\cdot)\right)\right)_{L^2(\Omega)},
+%\end{align*}
 where $\Phi(y,\cdot)$ denotes the function $x \mapsto \Phi(y,x)$.
-Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions.
-If instead $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ is a distribution, then the Riesz representative of $\rho$ may not exist, so the above derivation does not make sense.
+Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions, and instead suppose $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ is a distribution.
+In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in \eqref{eq:extension_to_distributions} is not valid.
 However, if $\Phi$ is sufficiently regular such that the function $y \mapsto \rho\left(\Phi(y,~\cdot~)\right)$ is well-defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the last line in the derivation is well-defined.
@@
 \end{equation}
-Here ``$\Aop \genericdistribution^*$'' is an abuse of notation that means the action of $\Aop$ on the distribution $\rho$, per this definition.
+Since $\genericdistribution^*$ may not exist, here ``$\Aop \genericdistribution^*$'' is simply a notation that means ``the action of $\Aop$ on the distribution $\rho$,'' per \eqref{eq:action_on_distribution}.
 
@@
 %Our method, described in Section \ref{sec:method}, belongs to a class of methods that approximate an operator by interpolating impulse responses of the operator associated with a scattered collection of points.
-Let $\delta_x$ denote the delta distribution\footnote{Recall that the delta distribution $\delta_x:\mathcal{D}(\Omega)\rightarrow \mathbb{R}$ is the linear functional defined by $\delta_x(v) = v(x)$ for all $v\in \mathcal{D}(\Omega)$.} (i.e., point source, impulse) $\delta_x$, centered at the point $x$.
-The \emph{impulse response} of $\Aop$ associated with the point $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$, 
+Let $\delta_x$ denote the delta distribution\footnote{Recall that the delta distribution $\delta_x:\mathcal{D}(\Omega)\rightarrow \mathbb{R}$ is the linear functional defined by $\delta_x(v) = v(x)$ for all $v\in \mathcal{D}(\Omega)$.} (i.e., point source, impulse) centered at the point $x \in \Omega$.
+The \emph{impulse response} of $\Aop$ associated with $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$, 
 \begin{equation}
 \label{eq:impulse_response_delta_action}
-\impulseresponse_x = \left( \Aop \delta_x^* \right)^*,
-\end{equation}
-that is formed by applying $\Aop$ to $\delta_x$, then taking the Riesz representation of the result.
-Using \eqref{eq:action_on_distribution} and the definition of the delta distribution, we also see that $\phi_x$ is the function
+\impulseresponse_x := \left( \Aop \delta_x^* \right)^*,
+\end{equation}
+that is formed by applying $\Aop$ to $\delta_x$, then taking the Riesz representation of the resulting linear functional.
+Using \eqref{eq:action_on_distribution} and the definition of the delta distribution, we see that $\phi_x$ may also be written as the function
 \begin{equation}
 \label{eq:impulse_response_defn}
-\impulseresponse_x(y) := \Aker(y, x),
+\impulseresponse_x(y) = \Aker(y, x),
 \end{equation} 
@@
 \begin{enumerate}
-	\item The kernel $\Phi$ is sufficiently regular such that $\phi_x$ is a well-defined function for almost all $x\in \Omega$.
-	\item The supports of the impulse responses $\impulseresponse_x$ are contained (or approximately contained) in localized regions.
-	\item The integral kernel is non-negative in the sense that $$\Phi(y,x) \ge 0$$ for all (or nearly all) $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
+	\item The kernel $\Phi$ is sufficiently regular such that $\phi_x$ is a well-defined function for all $x\in \Omega$.
+	\item The supports of the impulse responses $\impulseresponse_x$ are contained in localized regions.
+	\item The integral kernel is non-negative in the sense that $$\Phi(y,x) \ge 0$$ for all $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
 The operator $\Aop$ need not be positive semi-definite to use our method, and positive semi-definite operators need not have a non-negative integral kernel.}
 \end{enumerate}
-Because of these conditions, we are able to estimate the supports of the impulse responses a-priori by applying $\Aop^T$ \footnote{Recall that the transpose of $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ satisfying $\left(\Aop u\right)(v) = \left(\Aop^T v\right)(u)$ for all $u,v\in L^2(\Omega)$.} to a small number of polynomial functions (Section \ref{sec:intromoments}).
+Our method may still perform well if these conditions are relaxed slightly.
+For example, it is fine if the support of the impulse response $\impulseresponse_x$ is not perfectly contained in a localized region, so long as the bulk of the ``mass'' of the impulse response is contained in a localized region.
+The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$, as long as it is non-negative for the vast majority of pairs of points $(y,x)$, and as long as the negative numbers are comparatively small.
+If these conditions are violated, our method will incur additional error.
+If these conditions are violated too much, our method may fail.
+Not all operators of interest (and in particular not all Hessians) satisfy these conditions, but many operators of practical interest do satisfy these conditions, and our method is highly effective for approximating these operators.
+%Because of these conditions, we are able to estimate the supports of the impulse responses a-priori by applying $\Aop^T$ \footnote{Recall that the transpose of $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ satisfying $\left(\Aop u\right)(v) = \left(\Aop^T v\right)(u)$ for all $u,v\in L^2(\Omega)$.} to a small number of polynomial functions (Section \ref{sec:intromoments}).
 We use the support estimates to compute scattered impulse responses in batches (Section \ref{sec:get_impulse_response}, also Figure \ref{fig:batches_intro}).
@@
 %What is important is that $\mathbf{A}$ is dense, only available via matrix-vector products, and each of these matrix-vector products is costly to compute. 
-Of course, matrix entries can be computed via a matrix-vector products as $\mathbf{A}_{ij} = \left(\mathbf{A}\mathbf{e}_j\right)_i$,
+Of course, matrix entries can be computed via matrix-vector products as $\mathbf{A}_{ij} = \left(\mathbf{A}\mathbf{e}_j\right)_i$,
 %\begin{equation*}
@@
 \end{equation}
-for some kernel $\Phi_h$ which may differ from $\Phi$ due to discretization error.
+for some kernel $\Phi_h$ which we do not know, and which may differ from $\Phi$ due to discretization error.
 Since the functions $V_h$ are continuous at $x$, the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta}_x \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ with entries $\left(\boldsymbol{\delta}_x\right)_i = \febasis_i(x)$ for $i=1,\dots,N$.
-Additionally, it is straightforward to show that the Riesz representation, $\rho_h^* \in V_h$, of a functional $\rho \in V_h'$ has coefficient vector
+Additionally, it is straightforward to verify that the Riesz representation, $\rho_h^* \in V_h$, of a functional $\rho \in V_h'$ has coefficient vector
 \begin{equation*}
@@
 \end{equation}
-Let $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ be the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
+Now define $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ to be the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
 \begin{equation}
@@
 \end{equation}
-Broadly, our method will construct an H-matrix approximation of $\mathbf{A}$ by forming H-matrix approximations of $\boldsymbol{\Phi}$ and $\mathbf{M}$ (or a lumped mass version of $\mathbf{M}$), then multiplying these matrices per \eqref{eq:boldA}.
+Broadly, our method will construct an H-matrix approximation of $\mathbf{A}$ by forming H-matrix approximations of $\boldsymbol{\Phi}$ and $\mathbf{M}$ (or a lumped mass version of $\mathbf{M}$), then multiplying these matrices with H-matrix methods per \eqref{eq:boldA}.
 The central challenge is that classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Phi}_{ij}$, but these matrix entries are not easily accessible.
@@
 In this section we present two key innovations that our method is based on.
-The method itself will be presented in Section \ref{sec:method}.
+The approximation algorithm will be presented in Section \ref{sec:method}.
 First, we define moments of the impulse responses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{sec:intromoments}).
@@
 
-\begin{figure}
-	\begin{center}
-		\includegraphics[scale=1.0]{mean_displacement_invariance.pdf}
-	\end{center}
-	\caption{An operator is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \mu(x) + \mu(x')\right)$ when $x$ is close to $x'$.}
-	\label{fig:mean_displacement_invariance}
-\end{figure}
-
 Approximation \eqref{eq:translate3} improves upon \eqref{eq:translate1} and \eqref{eq:translate2}, and generalizes both.

### 2022-03-29 paper/localpsf.tex
define transpose in intro

@@
 We present a fast \emph{matrix free} method for approximating operators, $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$, that have locally supported non-negative integral kernels.
-Here $\Omega \subset \mathbb{R}^d$ is a bounded domain, $d$ is the spatial dimension, and $L^2(\Omega)'$ is the topological dual space of $L^2(\Omega)$.
+Here $\Omega \subset \mathbb{R}^d$ is a bounded domain, $d$ is the spatial dimension, and $L^2(\Omega)'$ is the space of continuous linear functionals on $L^2(\Omega)$ (i.e., the topological dual space).
 Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps) \cite{DTNMAP}, covariance operators in spatial statistics \cite{COVARIANCEOPERATORS}, blurring operators in imaging \cite{BLURRING}, and Hessians in distributed parameter PDE-constrained optimization and inverse problems \cite{PDEHESSIAN}.
 
-By ``matrix free,'' we mean that we have a black-box computational procedure through which we may apply $\Aop$ and $\Aop^T$ to arbitrary functions,
+By ``matrix free,'' we mean that we have a black-box computational procedure through which we may apply $\Aop$ and its transpose\footnote{Recall that the transpose $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(v) = \left(\Aop v\right)(u)$ for all $u,v \in L^2(\Omega)$, where the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(v)$ is the result of applying that linear functional to $v \in L^2(\Omega)$, and similar for the analogous operations with $\Aop^T$.
+}, $\Aop^T$, to arbitrary functions,
 \begin{equation}
@@
 After discretization, $\Aop$ becomes a dense matrix that is typically too large to be built and stored.
-This property is called ``matrix-free'' because at the discrete level it means one can compute matrix-vector products with the matrix representation of $\Aop$, but cannot easily compute individual matrix entries.
+This property is called ``matrix-free'' because at the discrete level it means one can compute matrix-vector products with the matrix representations of $\Aop$ and $\Aop^T$, but cannot easily compute individual matrix entries.
 
@@
 	\end{center}
-	\caption{An operator is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \mu(x) + \mu(x')\right)$ when $x$ is close to $x'$.}
+	\caption{An operator is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \mu(x) + \mu(x')\right)$ when $x$ is close to $x'$.
+Here $\mu(x)$ is the mean (center of mass) of $\phi_x$, and similar for $\mu(x')$.}
 	\label{fig:mean_displacement_invariance}
@@
 %We write $L^2(\Omega)'$ to denote the topological dual of $L^2(\Omega)$. 
-The linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
-
+%The linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
 The action of $\Aop$ may be extended to distributions if the kernel $\Phi$ is sufficiently regular.

### 2022-03-29 paper/localpsf.tex
editing intro

@@
 
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale=0.4]{bct1.eps}
+	\end{center}
+	\caption{Hierarchical matrix block structure for the Stokes Hessian. The red blocks (tiny) are dense. All other blocks are low rank.
+	}
+	\label{fig:hmatrix_intro}
+\end{figure}
+
 The output of our method is a hierarchical matrix (H-matrix).
@@
 The reason is that \emph{classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products}.
+
+Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}, which we do not use.
+These alternative methods have, for instance, been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
+One issue with this approach is that, while these methods are asymptotically scalable in theory, in practice the required number of matrix vector products is large\footnote{DO CALCULATION}.
+Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods based on the peeling process more efficient in the future.
 
@@
 The basic idea behind our method is impulse response interpolation.
-The impulse response, $\phi_x$, associated with a point $x$ is the function that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
+The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(v) = \left(\rho^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
 We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with collections of points scattered throughout the domain (see Figure \ref{fig:batches_intro}).
@@
 This allows fast evaluation of approximate matrix entries, which are needed by the classical H-matrix construction methods.
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[scale=0.5]{bct1.eps}
-	\end{center}
-	\caption{Hierarchical matrix block structure for the Stokes Hessian. Red blocks are dense. All other blocks are low rank.
-	}
-	\label{fig:hmatrix_intro}
-\end{figure}
 
@@
 \begin{itemize}
-	\item PC approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, while the method we propose is based on an assumption that we call ``normalized local mean displacement invariance'' (see Figure \ref{fig:mean_displacement_invariance}).
+	\item PC approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, while the method we propose is based on an assumption that we call ``local mean displacement invariance'' (see Figure \ref{fig:mean_displacement_invariance}).
 This is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
@@
 
-Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}, which we do not use.
-These methods have, for instance, been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
-One issue with this approach is that, while these methods are asymptotically scalable in theory, in practice the required number of matrix vector products is large\footnote{DO CALCULATION}.
-Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods based on the peeling process more efficient in the future.
-
 \section{Setting}
@@
 %The linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
+
 The action of $\Aop$ may be extended to distributions if the kernel $\Phi$ is sufficiently regular.
 We derive the extension as follows.
-Suppose $\rho \in L^2(\Omega)'$, and let $\rho^* \in L^2(\Omega)$ denote the Riesz representative\footnote{Recall that the Riesz representative of a functional $\sigma \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\sigma^* \in L^2(\Omega)$ such that $\sigma(v) = \left(\sigma^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} of $\rho$ with respect to the $L^2(\Omega)$ inner product.
+Suppose $\rho \in L^2(\Omega)'$, and let $\rho^* \in L^2(\Omega)$ denote the Riesz representative of $\rho$ with respect to the $L^2(\Omega)$ inner product.
 We have

### 2022-03-30 paper/localpsf.tex
done editing based on Noemi's comments
editing intro

@@
 The output of our method is a hierarchical matrix (H-matrix).
-H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank (see Figure \ref{fig:hmatrix_intro}).
-For more details about H-matrices, see Appendix \ref{app:h_matrix}.
+H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high-rank (see Figure \ref{fig:hmatrix_intro} and Appendix \ref{app:h_matrix}).
 H-matrix methods permit us to perform matrix-vector products cheaply.
@@
 These H-matrix methods are fast and scale well to large problems.
+Typically, the work and memory required to perform these operations for an $N \times N$ H-matrix scales as $O(r^2 N \log(N)^a)$, where $r$ is the largest rank of the off-diagonal blocks, and $a \in \{0,1,2,3\}$.
+The exact cost depends on the type of H-matrix used and operation performed; for more details see \cite{HMATRIXCOST}.
 The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
-Since we are building an H-matrix, it is natural to ask why we do not build the H-matrix directly.
-The reason is that \emph{classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products}.
-
-Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}, which we do not use.
-These alternative methods have, for instance, been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
-One issue with this approach is that, while these methods are asymptotically scalable in theory, in practice the required number of matrix vector products is large\footnote{DO CALCULATION}.
-Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods based on the peeling process more efficient in the future.
+However, \emph{classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free}.
 
@@
 However, since our approximation is an H-matrix, it can be applied to vectors efficiently via H-matrix methods.
+
+Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}, which we do not use here.
+These alternative methods have, for instance, been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
+One issue with this approach is that, while these methods are asymptotically scalable in theory, in practice the required number of matrix vector products is large\footnote{DO CALCULATION}.
+Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods based on the peeling process more efficient in the future.
 
@@
 We already showed that $v_* \in H^2_N(\Omega)$ in Theorem \ref{thm:wi_no_breakdown}.
-Using the the definition of $v_*$ in \eqref{eq:optimal_v} and Lemma \ref{lem:S_matrix_characterizations}, we have
+Using the definition of $v_*$ in \eqref{eq:optimal_v} and Lemma \ref{lem:S_matrix_characterizations}, we have
 	\begin{equation*}

### 2022-03-30 paper/localpsf.tex
editing intro

@@
 
-\section{Setting}
+\section{Preliminaries}
 
@@
 The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, which exists in principle but is not easily accessible to us.
+In this section we describe how to extend the action of $\Aop$ to distributions , and define impulse responses (Section \ref{sec:impulse_response_2}).
+Then we state the conditions on $\Aop$ that our method requires (Section \ref{sec:conditions_2}).
+Finally, we detail finite element discretization (Section \ref{sec:finite_element_kernel}).
 %We write $L^2(\Omega)'$ to denote the topological dual of $L^2(\Omega)$. 
 %The linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
+
+\subsection{Action on distributions and impulse responses}
+\label{sec:impulse_response_2}
 
@@
 
-In this paper, we focus on approximating operators that satisfy the following conditions: 
+\subsection{Required conditions}
+\label{sec:conditions_2}
+
+We focus on approximating operators that satisfy the following conditions: 
 \begin{enumerate}
@@
 
-\subsubsection{Lumped mass matrix and discrete kernel non-negativity}
+%\subsubsection{Lumped mass matrix and discrete kernel non-negativity}
 

### 2022-03-30 paper/localpsf.tex
editing intro

@@
 In particular, we cannot easily access entries of $\Aop$'s integral kernel.
-In applications, evaluating the maps in \eqref{eq:eval_maps} may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
+In applications, evaluating the maps in \eqref{eq:eval_maps} may involve iteratively solving a large linear system, timestepping, or performing some other costly computational procedure.
 After discretization, $\Aop$ becomes a dense matrix that is typically too large to be built and stored.
@@
 
-The output of our method is a hierarchical matrix (H-matrix).
+The output of our method is a hierarchical matrix (H-matrix) approximation of a discretized version of $\Aop$.
 H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high-rank (see Figure \ref{fig:hmatrix_intro} and Appendix \ref{app:h_matrix}).
@@
 The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
-However, \emph{classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free}.
 
@@
 
-The basic idea behind our method is impulse response interpolation.
+Since our end goal is an H-matrix, it is natural to ask why we do not build the H-matrix directly using classical H-matrix construction methods.
+The reason is that \emph{classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free}.
+We overcome this difficulty using impulse response interpolation.
 The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(v) = \left(\rho^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
@@
 The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, which exists in principle but is not easily accessible to us.
-In this section we describe how to extend the action of $\Aop$ to distributions , and define impulse responses (Section \ref{sec:impulse_response_2}).
+In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define the impulse responses (Section \ref{sec:impulse_response_2}).
 Then we state the conditions on $\Aop$ that our method requires (Section \ref{sec:conditions_2}).
@@
 
-\subsection{Action on distributions and impulse responses}
+\subsection{Distributions and impulse responses}
 \label{sec:impulse_response_2}
@@
 
-At the continuum level, the integral kernel $\Phi$ is assumed to be non-negative.
+\paragraph{Lumped mass matrix} At the continuum level, the integral kernel $\Phi$ is assumed to be non-negative.
 However, when we compute impulse responses we are effectively operating with the discrete integral kernel shown in \eqref{eq:approx_kernel_entry_81}.

### 2022-03-31 paper/localpsf.tex
editing intro
update paper based on Tucker's comments. Also, working on automated sigma_min based on nearest neighbors

@@
 Such operators arise, for example, as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained inverse problems.
-H-matrix approximations are highly desirable because fast H-matrix methods can add, multiply, invert, and factorize H-matrices, compute matrix-vector products, solve linear systems, and perform other useful matrix computations.
+H-matrix approximations are highly desirable because tereh are fast H-matrix methods to add, multiply, invert, and factorize H-matrices, compute matrix-vector products, solve linear systems, and perform other useful matrix computations.
 Classical H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
@@
 Numerical results demonstrate that our method substantially outperforms existing state-of-the-art Hessian approximation methods which are based on low-rank approximation.
-We are able to form high quality approximations of high rank operators using only a small number of operator actions.
+We are able to form high quality approximations of high rank operators using only a small number of operator applies.
 \end{abstract}
@@
 More accessible approximations of the Hessian are highly desirable, because they allow for fast solution of \eqref{eq:minimization_problem} via Newton-type methods.
-More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because inverse of the Hessian locally approximates the Bayesian posterior covariance for $m$ \cite{LAPLACEAPPROXIMATION}.
+More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because the inverse of the Hessian locally approximates the Bayesian posterior covariance for $m$ \cite{LAPLACEAPPROXIMATION}.
 
@@
 
-Conventionally, PC approximations are applied to vectors using the fast Fourier transform (FFT) \cite{PRODCONVFFT}, which allows PC approximations to be used as a fast surrogates for $\Aop$.
+Conventionally, PC approximations are applied to vectors using the fast Fourier transform (FFT) \cite{PRODCONVFFT}, which allows PC approximations to be used as fast surrogates for $\Aop$.
 Often it is necessary to perform further linear algebra operations involving $\Aop$, such as solving linear systems with $\Aop$ as the coefficient operator.
@@
 for some kernel $\Phi_h$ which we do not know, and which may differ from $\Phi$ due to discretization error.
-Since the functions $V_h$ are continuous at $x$, the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta}_x \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ with entries $\left(\boldsymbol{\delta}_x\right)_i = \febasis_i(x)$ for $i=1,\dots,N$.
+Since the functions in $V_h$ are continuous at $x$, the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta}_x \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ with entries $\left(\boldsymbol{\delta}_x\right)_i = \febasis_i(x)$ for $i=1,\dots,N$.
 Additionally, it is straightforward to verify that the Riesz representation, $\rho_h^* \in V_h$, of a functional $\rho \in V_h'$ has coefficient vector

### 2022-03-31 localpsf/impulse_response_batches.py

@@ -1,6 +1,7 @@
 import numpy as np
 import dolfin as dl
 import matplotlib.pyplot as plt
+from scipy.spatial import KDTree
 from tqdm.auto import tqdm
 
 from .impulse_response_moments import impulse_response_moments
@@ -14,7 +15,7 @@ class ImpulseResponseBatches:
     def __init__(me, V_in, V_out,
                  apply_A, apply_At,
                  num_initial_batches=5,
-                 tau=2.5,
+                 tau=3.0,
                  max_candidate_points=None,
                  use_lumped_mass_matrix_for_impulse_response_moments=True,
                  num_neighbors=10,
@@ -45,11 +46,17 @@ class ImpulseResponseBatches:
         me.dof_coords_in = me.V_in.tabulate_dof_coordinates()
         me.dof_coords_out = me.V_out.tabulate_dof_coordinates()
 
+        N_out, d_out = me.dof_coords_out.shape
+
+        dof_coords_out_kdtree = KDTree(me.dof_coords_out)
+        closest_distances, _ = dof_coords_out_kdtree.query(me.dof_coords_out, 2)
+        sigma_mins = closest_distances[:,1].reshape((N_out,1)) # shape=(N,1)
+
         me.vertex2dof_out = dl.vertex_to_dof_map(me.V_out)
         me.dof2vertex_out = dl.dof_to_vertex_map(me.V_out)
 
-        eee0, PP = np.linalg.eigh(me.Sigma0)
-        eee = np.max([np.ones(eee0.shape)*sigma_min**2, eee0], axis=0)
+        eee0, PP = np.linalg.eigh(me.Sigma0) # eee0.shape=(N,d), PP.shape=(N,d,d)
+        eee = np.max([np.ones((1,d_out))*sigma_mins**2, eee0], axis=0)
         me.Sigma = np.einsum('nij,nj,nkj->nik', PP, eee, PP)
 
         mesh_vertex_vol   = [me.vol[k]       for k in me.vertex2dof_out]

### 2022-03-31 localpsf/product_convolution_kernel.py

@@ -7,7 +7,7 @@ import hlibpro_python_wrapper as hpro
 
 class ProductConvolutionKernel:
     def __init__(me, V_in, V_out, apply_A, apply_At, num_row_batches, num_col_batches,
-                 tau_rows=2.5, tau_cols=2.5,
+                 tau_rows=3.0, tau_cols=3.0,
                  num_neighbors_rows=10, num_neighbors_cols=10,
                  symmetric=False, gamma=1e-8, sigma_min=1e-6,
                  max_scale_discrepancy=1e2):
@@ -16,6 +16,7 @@ class ProductConvolutionKernel:
         me.apply_A = apply_A
         me.apply_At = apply_At
         me.max_scale_discrepancy = max_scale_discrepancy
+
 
         me.col_batches = ImpulseResponseBatches(V_in, V_out, apply_A, apply_At,
                                                 num_initial_batches=num_col_batches,

### 2022-04-04 paper/localpsf.tex
working through next round of paper comments by Noemi

@@
 In applications, evaluating the maps in \eqref{eq:eval_maps} may involve iteratively solving a large linear system, timestepping, or performing some other costly computational procedure.
-After discretization, $\Aop$ becomes a dense matrix that is typically too large to be built and stored.
+%After discretization, $\Aop$ becomes a dense matrix that is typically too large to be built and stored.
 This property is called ``matrix-free'' because at the discrete level it means one can compute matrix-vector products with the matrix representations of $\Aop$ and $\Aop^T$, but cannot easily compute individual matrix entries.
@@
 
-Our motivation for this work is approximation of Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs).
+Our motivation is Hessian approximation in distributed parameter inverse problems governed by partial differential equations (PDEs).
+That is, inverse problems in which one seeks to reconstruct an unknown parameter field from noisy observations of a state variable that depends on the parameter implicitly through the solution of a PDE.
+In the deterministic approach to inverse problems, one solves a regularized minimization problem to find the parameter that best fits the data.
+The objective function in the minimization problem is defined implicitly, so information about the Hessian of the objective function is only accessible via application of the Hessian to vectors \cite{HESSIANACTION}.
+%In particular, applying the Hessian to a vector requires solving two linear PDEs which take the form of linearizations of \eqref{eq:state_pde} \cite{HESSIANACTION}. 
+More accessible Hessian approximations are highly desirable because they permit fast solution of the minimization problem via Newton-type methods.
+More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because the inverse of the Hessian locally approximates the Bayesian posterior covariance for the parameter \cite{LAPLACEAPPROXIMATION}.
+
+%Our motivation for this work is approximation of Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs).
 That is, inverse problems in which one seeks to reconstruct an unknown parameter field, $m$, from noisy observations,
-\begin{equation*}
-y = f(m, u(m)) + \text{noise},
-\end{equation*}
-which depend on  a state variable $u$. In turn, $u$ depends on $m$ implicitly through the solution of a PDE, which we write generically as follows:
-\begin{equation}
-\label{eq:state_pde}
-0 = g(m,u).
-\end{equation}
-Here $u(m)$ denotes the solution of the PDE \eqref{eq:state_pde} as a function of $m$.
+%\begin{equation*}
+%y = f(m, u(m)) + \text{noise},
+%\end{equation*}
+%which depend on  a state variable $u$. In turn, $u$ depends on $m$ implicitly through the solution of a PDE, which we write generically as follows:
+%\begin{equation}
+%\label{eq:state_pde}
+%0 = g(m,u).
+%\end{equation}
+%Here $u(m)$ denotes the solution of the PDE \eqref{eq:state_pde} as a function of $m$.
 In the deterministic approach to inverse problems, one typically finds $m$ as the solution to the minimization problem
-\begin{equation}
-\label{eq:minimization_problem}
-\min_m \quad \frac{1}{2}\|y - f(m,u(m))\|_W^2 + \mathcal{R}(m),
-\end{equation}
-where $\|\cdot\|_W$ is weighted norm which depends on the noise covariance, and $\mathcal{R}(m)$ is a regularization term.
+%\begin{equation}
+%\label{eq:minimization_problem}
+%\min_m \quad \frac{1}{2}\|y - f(m,u(m))\|_W^2 + \mathcal{R}(m),
+%\end{equation}
+%where $\|\cdot\|_W$ is weighted norm which depends on the noise covariance, and $\mathcal{R}(m)$ is a regularization term.
 The objective function in optimization problem \eqref{eq:minimization_problem} is defined indirectly via the implicit function theorem.
@@
 In the context of Hessians, methods based on low rank approximation suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator actions are required to form an accurate ``low rank'' approximation \cite{MYDISSERTATION}.
+
+%Product-convolution (PC) approximations can (where applicable) approximate high rank operators efficiently.
 
@@
 \end{equation}
+Here the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(v)$ is the result of applying that linear functional to $v \in L^2(\Omega)$.
 The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, which exists in principle but is not easily accessible to us.
@@
 		&= \int_\Omega v(y) \int_\Omega \Phi(y,x) \genericdistribution^*(x) dx ~dy \\
-		&= \int_\Omega v(y) \genericdistribution\left(\Phi(y,\cdot)\right) dy,
+		&= \int_\Omega v(y) \genericdistribution\left(\Phi(y,\cdot)\right) dy, \label{eq:action_on_distribution}
 	\end{align}
@@
 In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in \eqref{eq:extension_to_distributions} is not valid.
-However, if $\Phi$ is sufficiently regular such that the function $y \mapsto \rho\left(\Phi(y,~\cdot~)\right)$ is well-defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the last line in the derivation is well-defined.
-Hence, we use the last line of this derivation to \emph{define} the action of $\Aop$ on the distribution $\rho$ as follows:
-\begin{equation}
-\label{eq:action_on_distribution}
-\left(\Aop \genericdistribution^*\right)(v) := \int_\Omega v(y) \genericdistribution\left(\Phi(y, \cdot)\right) dy.
-\end{equation}
-Since $\genericdistribution^*$ may not exist, here ``$\Aop \genericdistribution^*$'' is simply a notation that means ``the action of $\Aop$ on the distribution $\rho$,'' per \eqref{eq:action_on_distribution}.
+However, if $\Phi$ is sufficiently regular such that the function $y \mapsto \rho\left(\Phi(y,~\cdot~)\right)$ is well-defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the the right hand side of \eqref{eq:action_on_distribution} is well-defined.
+Hence, we \emph{define} the action of $\Aop$ on the distribution $\rho$ to be the right hand side of \eqref{eq:action_on_distribution}.
+For convenience of notation, we denote this action by ``$\Aop \genericdistribution^*$,'' even if $\rho^*$ does not exist.
+%\begin{equation}
+%\label{eq:action_on_distribution}
+%\left(\Aop \genericdistribution^*\right)(v) := \int_\Omega v(y) \genericdistribution\left(\Phi(y, \cdot)\right) dy.
+%\end{equation}
+%Since $\genericdistribution^*$ may not exist if $\rho$ is a distribution, 
+%We write ``$\Aop \genericdistribution^*$'' to denote ``the action of $\Aop$ on the distribution $\rho$,'' as defined by \eqref{eq:action_on_distribution}.
 

### 2022-04-05 paper/localpsf.tex
Massive cuts and edits to intro

@@
 
-%We seek to approximate operators $\Aop$ that are accessible \emph{matrix-free}. That is, we have a black-box computational procedure through which we may evaluate the maps
-%\begin{equation*}
-%u \mapsto\Aop u \qquad \text{and} \qquad v \mapsto \Aop^Tv
-%\end{equation*} 
-%for arbitrary functions $u$ and $v$, but we cannot easily compute kernel entries $\Aker(y,x)$.
-In applications, evaluating these maps may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
-After discretization, $\Aop$ becomes a dense matrix that is typically too large to be built and stored.
-This property is called ``matrix-free'' because at the discrete level it means one can compute matrix-vector products with the matrix representation of $\Aop$, but cannot easily compute individual matrix entries.
+Our motivation is Hessian approximation in distributed parameter inverse problems governed by partial differential equations (PDEs).
+That is, inverse problems in which one seeks to reconstruct an unknown parameter field from noisy observations of a state variable that depends on the parameter implicitly through the solution of a PDE.
+In the deterministic approach to the inverse problem, one solves a regularized minimization problem to find the parameter that best fits the data.
+The objective function in the minimization problem is defined implicitly, so information about the Hessian of the objective function is only accessible via application of the Hessian to vectors \cite{HESSIANACTION}.
+%In particular, applying the Hessian to a vector requires solving two linear PDEs which take the form of linearizations of \eqref{eq:state_pde} \cite{HESSIANACTION}. 
+%More accessible Hessian approximations are highly desirable because they permit fast solution of the minimization problem via Newton-type methods.
+More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because the inverse of the Hessian locally approximates the Bayesian posterior covariance for the parameter \cite{LAPLACEAPPROXIMATION}.
+The most popular matrix-free Hessian approximation methods are based on low rank approximation \cite{CCGOPAPERS}. 
+%This includes the Lanczos method, the randomized singular value decomposition \cite{HMTRANDOM}, and CUR approximation/skeletonization \cite{CUR,MAHONEY}. 
+However, low rank approximation methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator actions are required to form an accurate ``low rank'' approximation \cite{MYDISSERTATION}.
 
@@
 
-Since our end goal is an H-matrix, it is natural to ask why we do not build the H-matrix directly using classical H-matrix construction methods.
-The reason is that \emph{classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free}.
-We overcome this difficulty using impulse response interpolation.
+%Since our end goal is an H-matrix, it is natural to ask why we do not build the H-matrix directly using classical H-matrix construction methods. The reason is that 
+Classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.
+Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS,ILONAHMATRIX,NOEMIHMATRIX}, which we do not use here.
+%These alternative methods have, for instance, been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}. 
+While methods based on the peeling process are asymptotically scalable in theory, in practice the required number of matrix vector products is large\footnote{DO CALCULATION}.
+Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods based on the peeling process more efficient in the future.
+
+We overcome this difficulty using \emph{impulse response interpolation}.
 The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(v) = \left(\rho^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
 We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with collections of points scattered throughout the domain (see Figure \ref{fig:batches_intro}).
-Then we interpolate these impulse responses to approximate arbitrary entries of the operator's integral kernel.
+Then we interpolate shifted versions of these impulse responses to approximate arbitrary entries of the operator's integral kernel.
 This allows fast evaluation of approximate matrix entries, which are needed by the classical H-matrix construction methods.
+
+Picking the batches of points $x_i$ requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
+The idea of estimating the supports of the functions $\phi_x$ a-priori was inspired by techniques from resolution analysis in seismic imaging, in which the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion  \cite{RESOLUTION}.
+Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Section \ref{sec:intromoments}).
+Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a positive integral kernel.
+
+%We pick batches of points $x_i$ in such that many $\phi_{x_i}$ are computed with each operator action of $\Aop$.
+
+%We seek to approximate operators $\Aop$ that are accessible \emph{matrix-free}. That is, we have a black-box computational procedure through which we may evaluate the maps
+%\begin{equation*}
+%u \mapsto\Aop u \qquad \text{and} \qquad v \mapsto \Aop^Tv
+%\end{equation*} 
+%for arbitrary functions $u$ and $v$, but we cannot easily compute kernel entries $\Aker(y,x)$.
+In applications, evaluating these maps may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
+After discretization, $\Aop$ becomes a dense matrix that is typically too large to be built and stored.
+This property is called ``matrix-free'' because at the discrete level it means one can compute matrix-vector products with the matrix representation of $\Aop$, but cannot easily compute individual matrix entries.
+
+%Product-convolution (PC) approximations can (where applicable) approximate high rank operators efficiently.
+
+Our impulse response interpolation procedure may be categorized as a ``point spread function'' (PSF) method that is loosely based on so-called ``product-convolution'' (PC) approximations, which are are approximations of an operator by weighted sums of convolution operators with spatially varying weights \cite{PRODCONVLIST}.
+PC and PSF methods have a long history dating back several decades; for background we recommend the following papers: \cite{PRODCONVGOOD}.
+While PC approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, the method we propose is based on an assumption that we call ``local mean displacement invariance'' (see Figure \ref{fig:mean_displacement_invariance}).
+This is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
+Furthermore, PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulse responses.
+We avoid this issue by temporarily excluding impulse responses from the approximation procedure whenever the approximation procedure attempts to evaluate the impulse response at a location where it is undefined.
+
+%A single convolution operator may be high rank, so PC approximations can (where applicable) approximate high rank operators efficiently. 
+%The method proposed in this paper is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of scattered points $x_i \in \Omega$ \cite{PRODCONVLIST}.
+%Often PC methods are known as ``point spread function'' (PSF) methods, particularly in optics and image processing.
+There, the ``operator'' $\Aop$ is a physical process of distortion and blurring of light rays passing through an optical system such as a telescope or microscope, and the local convolution kernels are shifted\footnote{The impulse responses are shifted so that each convolution kernel is centered at zero instead of the location of the point source.} impulse responses of the system to point sources of light.
+%, which characterize the local distortion and blurring, and the weighting functions interpolate between these convolution kernels. 
+%PSF is also used as an umbrella term to refer to other approximation methods that focus on interpolating impulse responses or impulse response features.
+For background on PC and PSF methods, we recommend the following papers: \cite{PRODCONVGOOD}.
+
+%The method proposed in this paper is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses $\phi_{x_i}$ associated with a collection of scattered points $x_i \in \Omega$ \cite{PRODCONVLIST}.
+%A common choice for PC methods based on scattered impulse responses is to choose the $x_i$ to be nodes in a regular grid \cite{NAGY}.
+With a regular grid, a large number of $x_i$ are typically required to achieve an accurate approximation.
+%In our previous work, we started with a coarse grid of points $x_i$, then adaptively refined the grid in the regions where the error in the approximation was large \cite{PRODCONVMYPAPER}.
+However, computing each $\phi_{x_i}$ required one action of $\Aop$.
+Even with adaptive refinement, many $x_i$, and hence operator actions of $\Aop$, were required.
+In this paper, we reduce the computational cost further by picking points $x_i$ in a way that allows many $\phi_{x_i}$ to be computed with each operator action of $\Aop$.
+This requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
+The idea of estimating the supports of the functions $\phi_x$ a-priori was inspired by techniques from resolution analysis in seismic imaging, in which
+%In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_x$ is used to estimate the minimum length scale on which features of the parameter can be inferred near the point $x$.
+If the width of $\phi_x$ is ten meters, for instance, then features of the parameter near $x$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
+In \cite{RESOLUTION},
+%the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion  \cite{RESOLUTION}.
+Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Section \ref{sec:intromoments}).
+Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a positive integral kernel.
 
@@
 %\label{sec:PDE_hessian_motivation}
-
-Our motivation is Hessian approximation in distributed parameter inverse problems governed by partial differential equations (PDEs).
-That is, inverse problems in which one seeks to reconstruct an unknown parameter field from noisy observations of a state variable that depends on the parameter implicitly through the solution of a PDE.
-In the deterministic approach to inverse problems, one solves a regularized minimization problem to find the parameter that best fits the data.
-The objective function in the minimization problem is defined implicitly, so information about the Hessian of the objective function is only accessible via application of the Hessian to vectors \cite{HESSIANACTION}.
-%In particular, applying the Hessian to a vector requires solving two linear PDEs which take the form of linearizations of \eqref{eq:state_pde} \cite{HESSIANACTION}. 
-More accessible Hessian approximations are highly desirable because they permit fast solution of the minimization problem via Newton-type methods.
-More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because the inverse of the Hessian locally approximates the Bayesian posterior covariance for the parameter \cite{LAPLACEAPPROXIMATION}.
 
@@
 
-The most popular matrix-free operator approximation methods are based on low rank approximation \cite{CCGOPAPERS}.
-This includes the Lanczos method, the randomized singular value decomposition \cite{HMTRANDOM}, and CUR approximation/skeletonization \cite{CUR,MAHONEY}.
-However, low rank approximation is not suitable for high rank operators, and often operators of practical interest are high rank.
-In the context of Hessians, methods based on low rank approximation suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator actions are required to form an accurate ``low rank'' approximation \cite{MYDISSERTATION}.
-
-%Product-convolution (PC) approximations can (where applicable) approximate high rank operators efficiently.
-
-Product-convolution (PC) approximations are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
-A single convolution operator may be high rank, so PC approximations can (where applicable) approximate high rank operators efficiently.
-PC approximations have been used in a variety of fields going back several decades \cite{PRODCONVLIST}.
-Often PC methods are known as ``point spread function'' (PSF) methods, particularly in optics and image processing.
-There, the ``operator'' $\Aop$ is a physical process of distortion and blurring of light rays passing through an optical system such as a telescope or microscope.
-However, ``PSF'' is sometimes used as an umbrella term to refer to other approximation methods that focus on interpolating impulse responses or impulse response features.
-For background on PC and PSF methods, we recommend the following papers: \cite{PRODCONVGOOD}.
-
+%PC approximations have been used in a variety of fields going back several decades \cite{PRODCONVLIST}; for background on PC and PSF methods, we recommend the following papers: \cite{PRODCONVGOOD}.
 The method proposed in this paper is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of scattered points $x_i \in \Omega$ \cite{PRODCONVLIST}.
@@
 
-In this paper, we reduce the computational cost further by picking points $x_i$ in a way that allows many $\phi_{x_i}$ to be computed with each operator action of $\Aop$.
+% A single convolution operator may be high rank, so PC approximations can (where applicable) approximate high rank operators efficiently.
+PC approximations have been used in a variety of fields going back several decades \cite{PRODCONVLIST}.
+Often PC methods are known as ``point spread function'' (PSF) methods, particularly in optics and image processing.
+There, the ``operator'' $\Aop$ is a physical process of distortion and blurring of light rays passing through an optical system such as a telescope or microscope.
+However, ``PSF'' is sometimes used as an umbrella term to refer to other approximation methods that focus on interpolating impulse responses or impulse response features.
+For background on PC and PSF methods, we recommend the following papers: \cite{PRODCONVGOOD}.
+
+%The method proposed in this paper is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of scattered points $x_i \in \Omega$ \cite{PRODCONVLIST}.
+A common choice for PC methods in this class is to choose the $x_i$ to be nodes in a regular grid \cite{NAGY}.
+With a regular grid, a large number of $x_i$ are typically required to achieve an accurate approximation.
+In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
+However, computing each impulse response required one action of $\Aop$.
+Even with adaptive refinement, many $x_i$, and hence operator actions of $\Aop$, were required.
+
+%In this paper, we reduce the computational cost further by picking points $x_i$ in a way that allows many $\phi_{x_i}$ to be computed with each operator action of $\Aop$.
 This requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
@@
 
-Conventionally, PC approximations are applied to vectors using the fast Fourier transform (FFT) \cite{PRODCONVFFT}, which allows PC approximations to be used as fast surrogates for $\Aop$.
+%Conventionally, PC approximations are applied to vectors using the fast Fourier transform (FFT) \cite{PRODCONVFFT}, which allows PC approximations to be used as fast surrogates for $\Aop$.
 Often it is necessary to perform further linear algebra operations involving $\Aop$, such as solving linear systems with $\Aop$ as the coefficient operator.
@@
 Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}, because operators that are amenable to PC approximation are often sparse in the wavelet domain.
-While not commonly done, in theory linear algebra operations involving convolution-product approximations (transposes of PC approximations) could be done using fast methods for pseudo-differential operator symbol calculus \cite{YINGDEMANET} since these are equivalent to low rank approximation of the pseudo-differential symbol of the operator.
+%While not commonly done, in theory linear algebra operations involving convolution-product approximations (transposes of PC approximations) could be done using fast methods for pseudo-differential operator symbol calculus \cite{YINGDEMANET} since these are equivalent to low rank approximation of the pseudo-differential symbol of the operator.
 We also note other relevant matrix-free operator approximation methods based on wavelets \cite{HERRMANN} and pseudo-differential operators \cite{DEMANET} which do not use PC approximation.
@@
 
-In this paper, we build on previous work \cite{PRODCONVMYPAPER}, in which we converted PC approximations into H-matrix format.
-Since our end goal is constructing an H-matrix, we are freed from strict adherence to the PC format.
-This allows us to modify the PC approach to make the following improvements:
-\begin{itemize}
-	\item PC approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, while the method we propose is based on an assumption that we call ``local mean displacement invariance'' (see Figure \ref{fig:mean_displacement_invariance}).
-This is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
-	\item PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulse responses.
-We avoid this issue by temporarily excluding impulse responses from the approximation procedure whenever the approximation procedure attempts to evaluate the impulse response at a location where it is undefined.
-\end{itemize}
-In exchange for these improvements, the proposed approximation cannot be written as a weighted sum of convolution operators, hence we cannot apply this approximation to vectors via the FFT.
+%In this paper, we build on previous work \cite{PRODCONVMYPAPER}, in which we converted PC approximations into H-matrix format.
+
+%In exchange for these improvements, the proposed approximation cannot be written as a weighted sum of convolution operators, hence we cannot apply this approximation to vectors via the FFT.
 However, since our approximation is an H-matrix, it can be applied to vectors efficiently via H-matrix methods.
-
-Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}, which we do not use here.
-These alternative methods have, for instance, been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
-One issue with this approach is that, while these methods are asymptotically scalable in theory, in practice the required number of matrix vector products is large\footnote{DO CALCULATION}.
-Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods based on the peeling process more efficient in the future.
 

### 2022-04-06 paper/localpsf.tex
Massive cuts and edits to intro

@@
 %This includes the Lanczos method, the randomized singular value decomposition \cite{HMTRANDOM}, and CUR approximation/skeletonization \cite{CUR,MAHONEY}. 
-However, low rank approximation methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator actions are required to form an accurate ``low rank'' approximation \cite{MYDISSERTATION}.
+However, low rank approximation methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation \cite{MYDISSERTATION}.
+High-rank matrix free approximation methods are needed.
 
@@
 
-The output of our method is a hierarchical matrix (H-matrix) approximation of a discretized version of $\Aop$.
+Our method forms a hierarchical matrix (H-matrix) approximation of a discretized version of $\Aop$.
 H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high-rank (see Figure \ref{fig:hmatrix_intro} and Appendix \ref{app:h_matrix}).
@@
 
-We overcome this difficulty using \emph{impulse response interpolation}.
+We construct the H-matrix using \emph{impulse response interpolation}.
 The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(v) = \left(\rho^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.

### 2022-04-15 paper/localpsf.tex
updates to paper

@@
 Such operators arise, for example, as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained inverse problems.
-H-matrix approximations are highly desirable because tereh are fast H-matrix methods to add, multiply, invert, and factorize H-matrices, compute matrix-vector products, solve linear systems, and perform other useful matrix computations.
+H-matrix approximations are highly desirable because there are fast H-matrix methods to add, multiply, invert, and factorize H-matrices, compute matrix-vector products, solve linear systems, and perform other useful matrix computations.
 Classical H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
@@
 
-Our impulse response interpolation procedure may be categorized as a ``point spread function'' (PSF) method that is loosely based on so-called ``product-convolution'' (PC) approximations, which are are approximations of an operator by weighted sums of convolution operators with spatially varying weights \cite{PRODCONVLIST}.
+Our impulse response interpolation procedure may be categorized as a ``point spread function'' (PSF) method that is loosely based on so-called ``product-convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights \cite{PRODCONVLIST}.
 PC and PSF methods have a long history dating back several decades; for background we recommend the following papers: \cite{PRODCONVGOOD}.
@@
 Furthermore, PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulse responses.
-We avoid this issue by temporarily excluding impulse responses from the approximation procedure whenever the approximation procedure attempts to evaluate the impulse response at a location where it is undefined.
+We avoid this issue by temporarily excluding impulse responses from the approximation procedure whenever an attempt is made to evaluate the impulse response at a location where it is undefined.
 
@@
 In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in \eqref{eq:extension_to_distributions} is not valid.
-However, if $\Phi$ is sufficiently regular such that the function $y \mapsto \rho\left(\Phi(y,~\cdot~)\right)$ is well-defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the the right hand side of \eqref{eq:action_on_distribution} is well-defined.
+However, if $\Phi$ is sufficiently regular such that the function $y \mapsto \rho\left(\Phi(y,~\cdot~)\right)$ is well-defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of \eqref{eq:action_on_distribution} is well-defined.
 Hence, we \emph{define} the action of $\Aop$ on the distribution $\rho$ to be the right hand side of \eqref{eq:action_on_distribution}.
@@
 In this section we present two key innovations that our method is based on.
-The approximation algorithm will be presented in Section \ref{sec:method}.
+%The approximation algorithm will be presented in Section \ref{sec:method}. 
 First, we define moments of the impulse responses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{sec:intromoments}).
@@
 
-\subsection{Impulse response moments and support estimation}
+\subsection{Impulse response moments and ellipsoid support estimate}
 \label{sec:intromoments}
@@
 The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative integral kernel property. Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
-\begin{equation*}
+\begin{equation}
+\label{eq:define_vol}
 \spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy
-\end{equation*}
-denote the spatially varying scaling factor, and let $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d},$
-\begin{equation*}
-\spatialmean(x) := \mathbb{E}(\impulseresponse_x / V(x)), \qquad
-\spatialcov(x) := \Var(\impulseresponse_x /V(x)),
-\end{equation*}
-be the spatially varying mean and covariance of the normalized version of $\impulseresponse_x$, respectively.
+\end{equation}
+denote the spatially varying scaling factor, and for $i,j=1,\dots,d$ define $\mu:\Omega \rightarrow \mathbb{R}^d$ and $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$ as follows:
+\begin{align}
+\spatialmean^i(x) :=& \int_\Omega (\impulseresponse_p(y) / V(x)) y^i ~dy \label{eq:define_mean} \\
+\spatialcov^{ij}(x) :=& \int_\Omega (\impulseresponse_p(y) / V(x)) \left(y^i - \mu^i(x)\right) \left(y^j - \mu^j(x)\right) ~dy, \label{eq:define_cov}
+\end{align}
+where $q^i$ denotes the $i^\text{th}$ component of the vector $q$, and $M^{ij}$ denotes the $(i,j)$ entry of the matrix $M$.
+The vector $\mu(p)\in \mathbb{R}^d$ and the matrix $\Sigma(x) \in \mathbb{R}^{d \times d}$ are the mean and covariance of the normalized version of $\impulseresponse_x$, respectively.
 Imagine $\phi_x$ as a mound of dirt piled on a localized region in $\Omega$, with $\phi_x(y)$ being the height of the dirt mound at location $y$.
@@
 Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
-But this direct approach is computationally infeasible because it requires one operator action of $\Aop$ per point $x$, and our method will need to know $V(x)$, $\mu(x)$, and $\Sigma(x)$ at all points $x$.
-Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions.
-Specifically, let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$, and let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
+However, this direct approach is not feasible because we need to know $V(x)$, $\mu(x)$, and $\Sigma(x)$ before we compute $\phi_x$, in order choose the point $x$.
+Computing $\phi_x$ in order to determine $V(x)$, $\mu(x)$, and $\Sigma(x)$ would be extremely computationally expensive, and defeat the purpose of our algorithm, which is to reduce the computational cost by computing impulse responses in batches.
+Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ indirectly, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions.
+This may be motivated by analogy to matrices.
+If $\mathbf{B}\in \mathbb{R}^{N \times N}$ is a matrix with $i$th column $\mathbf{b}_i$ and $\mathbf{f} \in \mathbb{R}^N$, then
+\begin{equation*}
+\mathbf{B}^T \mathbf{f} = \begin{bmatrix}
+\horzbar & \mathbf{b}_1^T & \horzbar \\
+\horzbar & \mathbf{b}_2^T & \horzbar \\
+& \vdots & \\
+\horzbar & \mathbf{b}_N^T & \horzbar
+\end{bmatrix}
+\mathbf{f} = 
+\begin{bmatrix}
+\mathbf{b}_1^T \mathbf{f} \\
+\mathbf{b}_2^T \mathbf{f} \\
+\vdots \\
+\mathbf{b}_N^T \mathbf{f}
+\end{bmatrix}.
+\end{equation*}
+By computing one matrix-vector product of $\mathbf{B}^T$ with $\mathbf{f}$, we compute the inner product of each column of $\mathbf{B}$ with $\mathbf{f}$ simputaneously.
+The operator case is analogous, with the impulse response $\phi_x$ taking the place of matrix columns.
+We have
+\begin{equation}
+\label{eq:operator_simultaneous_ips}
+\left(\Aop^T f\right)^*(x) = \int_\Omega K(y,x) f(y) dy = \left(\phi_x, f\right)_{L^2(\Omega)}.
+\end{equation}
+By computing one operator application of $\Aop^T$ on $f$, we compute the inner product of each impulse response $\phi_x$ with $f$, for all points $x$ simultaneously.
+
+Let $C$, $L^i$, and $Q^{ij}$ be the following constant, linear, and quadratic functions:
 \begin{equation*}
@@
 L^i(x) := x^i, \qquad
-Q^{ij}(x) = x^i x^j.
-\end{equation*}
-The following results (derived in Appendix \ref{app:proofs}) follow straightforwardly from the definitions:
+Q^{ij}(x) = x^i x^j
+\end{equation*}
+for $i=1,\dots,d$. Using the definition of $V$ in \eqref{eq:define_vol} and using \eqref{eq:operator_simultaneous_ips}, we have
+\begin{equation*}
+	V(x) = \int_\Omega \phi_x(y) \cdot 1 ~ dy = \left(\phi_x, C\right)_{L^2(\Omega)} = \left(\mathcal{A}^T C\right)^*(x).
+\end{equation*}
+Hence, we may compute $V(x)$ for all $x$ simultaneously by applying $\Aop^T$ to $C$.
+Analogous manipulations show that $\mu(x)$ may be computed for all points $x$ simultaneously by applying $\Aop^T$ to the functions $L^i$, and $\Sigma(x)$ may be computed for all for all points $x$ simultaneously by applying $\Aop^T$ to the functions $Q^{ij}$.
+We have
 \begin{subequations}
@@
 \end{subequations}
-for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$.
+for $i,j=1,\dots, \gdim$.
 Here $f/g$ denotes pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denotes pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
@@
 
-We make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
+Next, we make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
 \begin{equation}
@@
 
-The quantity $\tau$ is a parameter that must be chosen by the user.
+The quantity $\tau$ is a parameter that must be chosen appropriately.
 The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$.
-However, in Section \ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
-We find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results.
+However, in Section \ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (the more ellipsoids the better), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
+In practice, we find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results.
 The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.

### 2022-04-15 paper/localpsf.tex
updates to paper

@@
 \end{equation}
-at the point $x$. The idea is that $\Phi(y,x) = f_i$ per the discussion in Section \ref{eq:local_mean_displacement_invariance}, so we approximate $\Phi(y,x)$ by interpolating the values $f_i$.
+at the point $x$. The idea is that $\Phi(y,x) = f_i$ per the discussion in Section \ref{sec:local_mean_displacement_invariance}, so we approximate $\Phi(y,x)$ by interpolating the values $f_i$.
 
@@
 For more details on radial basis function interpolation and polyharmonic splines, see \cite{rbf}.
+The interpolation procedure is illustrate in Figure \ref{fig:rbf_interpolation}.
+
+\begin{figure}
+	\begin{subfigure}{0.49\textwidth}
+		\centering
+		\includegraphics[scale=0.22]{interpolation_rbf.pdf}
+		\caption{}
+	\end{subfigure}
+	\begin{subfigure}{0.49\textwidth}
+		\centering
+		\includegraphics[scale=0.22]{interpolation_rbf_boundary.pdf}
+		\caption{}
+	\end{subfigure}
+	\caption{(a) To compute the kernel entry $\Phi(y,x)$, we use radial basis functions to interpolate the values $\frac{V(x)}{V(x_i)}\phi_{x_i}(y-\mu(x)-\mu(x_i))$ for the $k$ nearest sample points $x_i$ to $x$.
+(b) We exclude sample points for which $y-\mu(x)_\mu(x_i)$ is outside of the domain.}
+	\label{fig:rbf_interpolation}
+\end{figure}
 

### 2022-04-15 paper/localpsf.tex
minor fixes in paper

@@
 \end{equation*}
-By computing one matrix-vector product of $\mathbf{B}^T$ with $\mathbf{f}$, we compute the inner product of each column of $\mathbf{B}$ with $\mathbf{f}$ simputaneously.
+By computing one matrix-vector product of $\mathbf{B}^T$ with $\mathbf{f}$, we compute the inner product of each column of $\mathbf{B}$ with $\mathbf{f}$ simultaneously.
 The operator case is analogous, with the impulse response $\phi_x$ taking the place of matrix columns.
@@
 Analogous manipulations show that $\mu(x)$ may be computed for all points $x$ simultaneously by applying $\Aop^T$ to the functions $L^i$, and $\Sigma(x)$ may be computed for all for all points $x$ simultaneously by applying $\Aop^T$ to the functions $Q^{ij}$.
-We have
+To summarize the results of doing this, we have
 \begin{subequations}

### 2022-04-18 paper/localpsf.tex
update paper: adding references. Wrote up new positive definite modification stuff

@@
 Here $\Omega \subset \mathbb{R}^d$ is a bounded domain, $d$ is the spatial dimension, and $L^2(\Omega)'$ is the space of continuous linear functionals on $L^2(\Omega)$ (i.e., the topological dual space).
-Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps) \cite{DTNMAP}, covariance operators in spatial statistics \cite{COVARIANCEOPERATORS}, blurring operators in imaging \cite{BLURRING}, and Hessians in distributed parameter PDE-constrained optimization and inverse problems \cite{PDEHESSIAN}.
+Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{le1997non,saad1999distributed,TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps) \cite{DTNMAP}, covariance operators in spatial statistics \cite{GeogaEtAl20,ChenStein21,LindgrenRueLindstrom11}, blurring operators in imaging \cite{DenisEtAl11,NagyOleary98}, and Hessians in distributed parameter PDE-constrained optimization and inverse problems \cite{AlgerEtAl19,IsaacEtAl15}.
 
@@
 More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because the inverse of the Hessian locally approximates the Bayesian posterior covariance for the parameter \cite{LAPLACEAPPROXIMATION}.
-The most popular matrix-free Hessian approximation methods are based on low rank approximation \cite{CCGOPAPERS}. 
+The most popular matrix-free Hessian approximation methods are based on low rank approximation \cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15} because there are efficient methods for constructing low rank approximations of a matrix using only matrix-vector products \cite{Cheng05,HalkoMartinssonTropp11}.
 %This includes the Lanczos method, the randomized singular value decomposition \cite{HMTRANDOM}, and CUR approximation/skeletonization \cite{CUR,MAHONEY}. 
-However, low rank approximation methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation \cite{MYDISSERTATION}.
+However, low rank approximation methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation \cite{Alger19,AlgerEtAl17}.
 High-rank matrix free approximation methods are needed.
@@
 
-Our method forms a hierarchical matrix (H-matrix) approximation of a discretized version of $\Aop$.
+Our method forms a hierarchical matrix (H-matrix) \cite{Hackbusch99,} approximation of a discretized version of $\Aop$.
 H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high-rank (see Figure \ref{fig:hmatrix_intro} and Appendix \ref{app:h_matrix}).
@@
 Typically, the work and memory required to perform these operations for an $N \times N$ H-matrix scales as $O(r^2 N \log(N)^a)$, where $r$ is the largest rank of the off-diagonal blocks, and $a \in \{0,1,2,3\}$.
-The exact cost depends on the type of H-matrix used and operation performed; for more details see \cite{HMATRIXCOST}.
+The exact cost depends on the type of H-matrix used and operation performed; for more details see \cite{GrasedyckHackbusch03,Hackbusch15}.
 The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
@@
 %Since our end goal is an H-matrix, it is natural to ask why we do not build the H-matrix directly using classical H-matrix construction methods. The reason is that 
-Classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.
-Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS,ILONAHMATRIX,NOEMIHMATRIX}, which we do not use here.
-%These alternative methods have, for instance, been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}. 
-While methods based on the peeling process are asymptotically scalable in theory, in practice the required number of matrix vector products is large\footnote{DO CALCULATION}.
+Classical H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.
+Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}, which we do not use here.
+These alternative methods have, been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{AmbartsumyanEtAl20,HartlandEtAl22}. 
+While methods based on the peeling process are asymptotically scalable in theory, in practice the required number of matrix vector products is large.
 Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods based on the peeling process more efficient in the future.
@@
 Picking the batches of points $x_i$ requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
-The idea of estimating the supports of the functions $\phi_x$ a-priori was inspired by techniques from resolution analysis in seismic imaging, in which the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion  \cite{RESOLUTION}.
+The idea of estimating the supports of the functions $\phi_x$ a-priori was inspired by techniques from resolution analysis in seismic imaging, in which the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion  \cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
 Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Section \ref{sec:intromoments}).
@@
 
-Our impulse response interpolation procedure may be categorized as a ``point spread function'' (PSF) method that is loosely based on so-called ``product-convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights \cite{PRODCONVLIST}.
-PC and PSF methods have a long history dating back several decades; for background we recommend the following papers: \cite{PRODCONVGOOD}.
-While PC approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, the method we propose is based on an assumption that we call ``local mean displacement invariance'' (see Figure \ref{fig:mean_displacement_invariance}).
+Our impulse response interpolation procedure may be categorized as a ``point spread function'' (PSF) method that is loosely based on so-called ``product-convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
+PC and PSF methods have a long history dating back several decades.
+We note the following papers (among many others),  \cite{Adorf94,AlgerEtAl19,BigotEscandeWeiss19,NagyOleary98,FishEtAl96,EscandeWeiss12,EscandeWeiss15,EscandeWeiss22}, in which the convolution kernels are constructed from sampling impulse responses of the operator to scattered point sources.
+For background on PC and PSF methods, we recommend the following papers which provide well-written reviews of the topic: \cite{DenisEtAl15,EscandeWeiss17,GentileCourbinMeylan13}.
+While PC approximations are based on an assumption of local translation invariance, the method we propose is based on an assumption that we call ``local mean displacement invariance'' (see Figure \ref{fig:mean_displacement_invariance}).
 This is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
@@
 \end{align}
-where $q^i$ denotes the $i^\text{th}$ component of the vector $q$, and $M^{ij}$ denotes the $(i,j)$ entry of the matrix $M$.
-The vector $\mu(p)\in \mathbb{R}^d$ and the matrix $\Sigma(x) \in \mathbb{R}^{d \times d}$ are the mean and covariance of the normalized version of $\impulseresponse_x$, respectively.
-Imagine $\phi_x$ as a mound of dirt piled on a localized region in $\Omega$, with $\phi_x(y)$ being the height of the dirt mound at location $y$.
+where $\mu^i(x)$ denotes the $i^\text{th}$ component of the vector $\mu(x)$, and $\Sigma^{ij}(x)$ denotes the $(i,j)$ entry of the matrix $\Sigma(x)$.
+The vector $\mu(x)\in \mathbb{R}^d$ and the matrix $\Sigma(x) \in \mathbb{R}^{d \times d}$ are the mean and covariance of the normalized version of $\impulseresponse_x$, respectively. 
+%Imagine $\phi_x$ as a mound of dirt piled on a localized region in $\Omega$, with $\phi_x(y)$ being the height of the dirt mound at location $y$.
 The scalar $V(x)$ is the total volume of dirt in the mound, the vector $\mu(x)$ is the center of mass of the dirt mound, and the matrix $\Sigma(x)$ contains information about how wide the dirt mound is in each spatial direction.
 
-The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}.
+The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to obtain $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}.
 Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
@@
 Computing $\phi_x$ in order to determine $V(x)$, $\mu(x)$, and $\Sigma(x)$ would be extremely computationally expensive, and defeat the purpose of our algorithm, which is to reduce the computational cost by computing impulse responses in batches.
-Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ indirectly, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions.
+Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ indirectly, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions (e.g., 6 total operator applies in two spatial dimensions and 10 in three spatial dimensions).
 This may be motivated by analogy to matrices.
-If $\mathbf{B}\in \mathbb{R}^{N \times N}$ is a matrix with $i$th column $\mathbf{b}_i$ and $\mathbf{f} \in \mathbb{R}^N$, then
+If $\mathbf{B}\in \mathbb{R}^{N \times N}$ is a matrix with $i^\text{th}$ column $\mathbf{b}_i$ and $\mathbf{f} \in \mathbb{R}^N$, then
 \begin{equation*}
@@
 By computing one matrix-vector product of $\mathbf{B}^T$ with $\mathbf{f}$, we compute the inner product of each column of $\mathbf{B}$ with $\mathbf{f}$ simultaneously.
-The operator case is analogous, with the impulse response $\phi_x$ taking the place of matrix columns.
+The operator case is analogous, with the impulse response $\phi_x$ taking the place of a matrix column.
 We have
@@
 \end{equation}
-By computing one operator application of $\Aop^T$ on $f$, we compute the inner product of each impulse response $\phi_x$ with $f$, for all points $x$ simultaneously.
+By computing one operator application of $\Aop^T$ to $f$, we compute the inner product of each impulse response $\phi_x$ with $f$, for all points $x$ simultaneously.
 
@@
 L^i(x) := x^i, \qquad
-Q^{ij}(x) = x^i x^j
-\end{equation*}
-for $i=1,\dots,d$. Using the definition of $V$ in \eqref{eq:define_vol} and using \eqref{eq:operator_simultaneous_ips}, we have
-\begin{equation*}
-	V(x) = \int_\Omega \phi_x(y) \cdot 1 ~ dy = \left(\phi_x, C\right)_{L^2(\Omega)} = \left(\mathcal{A}^T C\right)^*(x).
+Q^{ij}(x) := x^i x^j
+\end{equation*}
+for $i,j=1,\dots,d$. Using the definition of $V$ in \eqref{eq:define_vol} and using \eqref{eq:operator_simultaneous_ips}, we have
+\begin{equation*}
+	V(x) = \int_\Omega \phi_x(y) C(y) ~ dy = \left(\phi_x, C\right)_{L^2(\Omega)} = \left(\mathcal{A}^T C\right)^*(x).
 \end{equation*}
 Hence, we may compute $V(x)$ for all $x$ simultaneously by applying $\Aop^T$ to $C$.
-Analogous manipulations show that $\mu(x)$ may be computed for all points $x$ simultaneously by applying $\Aop^T$ to the functions $L^i$, and $\Sigma(x)$ may be computed for all for all points $x$ simultaneously by applying $\Aop^T$ to the functions $Q^{ij}$.
+Analogous manipulations show that $\mu(x)$ and $\Sigma(x)$ may be computed for all points $x$ simultaneously by applying $\Aop^T$ to the functions $L^i$ and $Q^{ij}$, respectively.
 To summarize the results of doing this, we have
@@
 
+%WOODBURY: \cite{Hager89}.
+
 In many problems of practical interest (e.g., Hessian approximation) $\Aop$ is symmetric positive semi-definite.
@@
 Also, lacking these properties may prevent one from using highly effective algorithms, such as the conjugate gradient method, to perform further useful operations involving $\mathbf{A}_H$.
+
+If $\Aop$ is symmetric positive definite, we symmetrize $\mathbf{A}_H$ as follows
+\begin{equation*}
+\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
+\end{equation*}
+Then we compute the negative eigenvalues, of $\mathbf{A}_H$ that are less than (i.e., more negative than) a small magnitude negative cutoff value $c$, and also compute the corresponding eigenvectors.
+That is, the eigenpairs $\{(\lambda_i,u_i)\}_{i=1}^{N_\text{bad}}$ with
+\begin{equation*}
+	\lambda_i < c < 0.
+\end{equation*}
+These outlying eigenpairs may be computed using a Krylov method, or the randomized singular value decomposition.
+We then flip the sign of these ``bad'' negative eigenvalues to be positive as follows:
+\begin{equation*}
+	\mathbf{A}_H^{\text{sym}+} = \mathbf{A}_H^{\text{sym}} + \sum_{i=1}^{N_\text{bad}} 2 |\lambda_i| u_i u_i^T.
+\end{equation*}
+The matrix $\mathbf{A}_H^{\text{sym}+}$ approximates the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$.
+A H-matrix representation of $\mathbf{A}_H^{\text{sym}+}$ is formed using H-matrix addition, and used in place of $\mathbf{A}_H$ in further linear algebra operations.
+
+The cutoff $c$ is required when $\Aop$ has a large number of zero eigenvalues, or a sequence of eigenvalues that cluster at zero, a situation that occurs commonly in practice.
+We seek to correct negative eigenvalues caused by errors in our approximation procedure.
+But even without our approximation, numerical rounding and discretization errors may cause large numbers of zero or near-zero eigenvalues to become slightly negative.
+Numerically, we cannot hope to correct these small magnitude negative eigenvalues because there are so many of them, and because the spectral gap between the smallest negative eigenvalue and the first non-negative eigenvalue is small.
+
+For Hessians in ill-posed inverse problems, the Hessian is the sum of the Hessian of a data misfit term, and the Hessian of a regularization term.
+The Hessian of the regularization term is known and easy to manipulate, and one seeks to approximate the Hessian of the data misfit term with our method.
+In this case, a reasonable choice for the cutoff parameter is $c = -|\lambda_R^\text{min}|/2$, where $\lambda_R^\text{min}$ is the smallest magnitude eigenvalue of the Hessian of the regularization term.
+We use this cutoff parameter in our numerical experiments.
+
+\section{old rational}
+
 In this section we describe a method for modifying $\mathbf{A}_H$ to make it symmetric positive semi-definite.
@@
 
+\FloatBarrier
+
 \bibliographystyle{siamplain}
-\bibliography{references}
+\bibliography{localpsf}
+
 \end{document}

### 2022-04-18 paper/localpsf_old3.tex
about to remove old stuff from paper

--- 2022-04-15 paper/localpsf.tex
@@
 Such operators arise, for example, as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained inverse problems.
-H-matrix approximations are highly desirable because there are fast H-matrix methods to add, multiply, invert, and factorize H-matrices, compute matrix-vector products, solve linear systems, and perform other useful matrix computations.
-Classical H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
-Our method computes impulse responses of the operator at a collection of scattered points, then interpolates these known impulse responses to evaluate entries of the integral kernel at other points where the impulse response is not known, which allows us to construct the H-matrix.
+Fast H-matrix methods can add, multiply, invert, and factorize H-matrices, compute matrix-vector products, solve linear systems, and perform other useful matrix computations.
+Our method computes impulse responses of the operator at a collection of scattered points, then interpolates these known impulse responses to evaluate entries of the integral kernel at other points where the impulse response is not known.
 Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources.
@@
 Numerical results demonstrate that our method substantially outperforms existing state-of-the-art Hessian approximation methods which are based on low-rank approximation.
-We are able to form high quality approximations of high rank operators using only a small number of operator applies.
+We are able to form high quality approximations of high rank operators using only a small number of operator actions products.
 \end{abstract}
@@
 
-We present a fast \emph{matrix free} method for approximating operators, $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$, that have locally supported non-negative integral kernels.
-Here $\Omega \subset \mathbb{R}^d$ is a bounded domain, $d$ is the spatial dimension, and $L^2(\Omega)'$ is the space of continuous linear functionals on $L^2(\Omega)$ (i.e., the topological dual space).
-Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps) \cite{DTNMAP}, covariance operators in spatial statistics \cite{COVARIANCEOPERATORS}, blurring operators in imaging \cite{BLURRING}, and Hessians in distributed parameter PDE-constrained optimization and inverse problems \cite{PDEHESSIAN}.
-
-By ``matrix free,'' we mean that we have a black-box computational procedure through which we may apply $\Aop$ and its transpose\footnote{Recall that the transpose $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(v) = \left(\Aop v\right)(u)$ for all $u,v \in L^2(\Omega)$, where the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(v)$ is the result of applying that linear functional to $v \in L^2(\Omega)$, and similar for the analogous operations with $\Aop^T$.
-}, $\Aop^T$, to arbitrary functions,
-\begin{equation}
-\label{eq:eval_maps}
-u \mapsto\Aop u \qquad \text{and} \qquad v \mapsto \Aop^Tv,
-\end{equation} 
-but we do not have any further access to $\Aop$.
-In particular, we cannot easily access entries of $\Aop$'s integral kernel.
-In applications, evaluating the maps in \eqref{eq:eval_maps} may involve iteratively solving a large linear system, timestepping, or performing some other costly computational procedure.
-%After discretization, $\Aop$ becomes a dense matrix that is typically too large to be built and stored.
-This property is called ``matrix-free'' because at the discrete level it means one can compute matrix-vector products with the matrix representations of $\Aop$ and $\Aop^T$, but cannot easily compute individual matrix entries.
-
-Our motivation is Hessian approximation in distributed parameter inverse problems governed by partial differential equations (PDEs).
-That is, inverse problems in which one seeks to reconstruct an unknown parameter field from noisy observations of a state variable that depends on the parameter implicitly through the solution of a PDE.
-In the deterministic approach to the inverse problem, one solves a regularized minimization problem to find the parameter that best fits the data.
-The objective function in the minimization problem is defined implicitly, so information about the Hessian of the objective function is only accessible via application of the Hessian to vectors \cite{HESSIANACTION}.
-%In particular, applying the Hessian to a vector requires solving two linear PDEs which take the form of linearizations of \eqref{eq:state_pde} \cite{HESSIANACTION}. 
-%More accessible Hessian approximations are highly desirable because they permit fast solution of the minimization problem via Newton-type methods.
-More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because the inverse of the Hessian locally approximates the Bayesian posterior covariance for the parameter \cite{LAPLACEAPPROXIMATION}.
-The most popular matrix-free Hessian approximation methods are based on low rank approximation \cite{CCGOPAPERS}. 
-%This includes the Lanczos method, the randomized singular value decomposition \cite{HMTRANDOM}, and CUR approximation/skeletonization \cite{CUR,MAHONEY}. 
-However, low rank approximation methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation \cite{MYDISSERTATION}.
-High-rank matrix free approximation methods are needed.
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[scale=0.4]{bct1.eps}
-	\end{center}
-	\caption{Hierarchical matrix block structure for the Stokes Hessian. The red blocks (tiny) are dense. All other blocks are low rank.
-	}
-	\label{fig:hmatrix_intro}
-\end{figure}
-
-Our method forms a hierarchical matrix (H-matrix) approximation of a discretized version of $\Aop$.
-H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high-rank (see Figure \ref{fig:hmatrix_intro} and Appendix \ref{app:h_matrix}).
-H-matrix methods permit us to perform matrix-vector products cheaply.
-Furthermore, one may use H-matrix methods to perform useful linear algebra operations involving the H-matrix that cannot easily be done using the original operator.
-These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion.
-These H-matrix methods are fast and scale well to large problems.
-Typically, the work and memory required to perform these operations for an $N \times N$ H-matrix scales as $O(r^2 N \log(N)^a)$, where $r$ is the largest rank of the off-diagonal blocks, and $a \in \{0,1,2,3\}$.
-The exact cost depends on the type of H-matrix used and operation performed; for more details see \cite{HMATRIXCOST}.
-The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
+We present a fast matrix free method for approximating operators that have locally supported non-negative integral kernels.
+Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems.
+
+Let $\Omega \subset \mathbb{R}^d$ be a bounded domain (in applications typically $\gdim=1$, $2$, or $3$) and let $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the following form:
+\begin{equation}
+	\label{eq:kernel_representation}
+	(\Aop u)(v) := \int_\Omega \int_\Omega v(y) \Aker(y,x) u(x) dx dy.
+\end{equation}
+The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, $\Aop u \in L^2(\Omega)'$ is the linear functional that results from applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
+We seek to approximate operators $\Aop$ that are accessible \emph{matrix-free}.
+That is, we have a black-box computational procedure through which we may evaluate the maps
+\begin{equation*}
+u \mapsto\Aop u \qquad \text{and} \qquad v \mapsto \Aop^Tv
+\end{equation*} 
+for arbitrary functions $u$ and $v$, but we cannot easily compute kernel entries $\Aker(y,x)$.
+In applications, evaluating these maps may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
+After discretization, $\Aop$ becomes a dense matrix that is typically too large to be built and stored.
+This property is called ``matrix-free'' because at the discrete level it means one can compute matrix-vector products with the matrix representation of $\Aop$, but cannot easily compute individual matrix entries.
+
+Our method, described in Section \ref{sec:method}, belongs to a class of methods that approximate an operator by interpolating impulse responses of the operator associated with a scattered collection of points.
+The \emph{impulse response} of $\Aop$ associated with the point $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$,
+\begin{equation}
+\label{eq:impulse_response_defn}
+\impulseresponse_x(y) := \Aker(y, x),
+\end{equation} 
+which results from keeping $x$ fixed and viewing $\Phi(y,x)$ as a function of $y$.
+The function $\impulseresponse_x$ is called the impulse response because straightforward analysis shows that it may be expressed as follows:
+\begin{equation}
+\label{eq:impulse_response_delta_action}
+\impulseresponse_x = \left( \Aop \delta_x^* \right)^*,
+\end{equation}
+where $\delta_x$ is the the delta distribution (i.e., point source, impulse) centered at $x$, $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\sigma \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\sigma^* \in L^2(\Omega)$ such that $\sigma(v) = \left(\sigma^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
+This is described in Appendix \ref{app:distributions}.
+
+We assume:
+\begin{itemize}
+	\item The supports of the impulse responses $\impulseresponse_x$ are contained (or approximately contained) in localized regions.
+	\item The integral kernel is non-negative in the sense that $$\Phi(y,x) \ge 0$$ for all (or nearly all) $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
+The operator $\Aop$ need not be positive semi-definite to use our method, and positive semi-definite operators need not have a non-negative integral kernel.}
+\end{itemize}
+Because of these assumptions, we are able to estimate the supports of the impulse responses a-priori by applying $\Aop$ to a small number of polynomial functions.
+We use the support estimates to compute scattered impulse responses in batches (see Figure \ref{fig:batches_intro}).
+This batching allows us to recover many impulse responses using a small number of operator actions $u \mapsto \Aop u$.
+The support estimates also allow us to interpolate the impulse responses with a new more sophisticated and accurate interpolation scheme.
 
@@
 Dashed gray ellipses indicate the estimated support ellipsoids.
-The large circle is the boundary of the domain.
-		%		Early batches tend to have lots of ellipsoids near the boundary, as compared to later batches. 
-		%		This is because our greedy ellipsoid packing algorithm tries to pick sample points that are far away from sample points in previously chosen batches.
-	}
+The large circle is the boundary of $\Omega$.
+Early batches tend to have lots of ellipsoids near the boundary, as compared to later batches.
+This is because our greedy ellipsoid packing algorithm tries to pick sample points that are far away from sample points in previously chosen batches.}
 	\label{fig:batches_intro}
@@
 
-%Since our end goal is an H-matrix, it is natural to ask why we do not build the H-matrix directly using classical H-matrix construction methods. The reason is that 
-Classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.
-Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS,ILONAHMATRIX,NOEMIHMATRIX}, which we do not use here.
-%These alternative methods have, for instance, been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}. 
-While methods based on the peeling process are asymptotically scalable in theory, in practice the required number of matrix vector products is large\footnote{DO CALCULATION}.
-Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods based on the peeling process more efficient in the future.
-
-We construct the H-matrix using \emph{impulse response interpolation}.
-The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(v) = \left(\rho^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
-We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with collections of points scattered throughout the domain (see Figure \ref{fig:batches_intro}).
-Then we interpolate shifted versions of these impulse responses to approximate arbitrary entries of the operator's integral kernel.
-This allows fast evaluation of approximate matrix entries, which are needed by the classical H-matrix construction methods.
-
-Picking the batches of points $x_i$ requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
-The idea of estimating the supports of the functions $\phi_x$ a-priori was inspired by techniques from resolution analysis in seismic imaging, in which the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion  \cite{RESOLUTION}.
-Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Section \ref{sec:intromoments}).
-Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a positive integral kernel.
-
-%We pick batches of points $x_i$ in such that many $\phi_{x_i}$ are computed with each operator action of $\Aop$.
-
-%We seek to approximate operators $\Aop$ that are accessible \emph{matrix-free}. That is, we have a black-box computational procedure through which we may evaluate the maps
-%\begin{equation*}
-%u \mapsto\Aop u \qquad \text{and} \qquad v \mapsto \Aop^Tv
-%\end{equation*} 
-%for arbitrary functions $u$ and $v$, but we cannot easily compute kernel entries $\Aker(y,x)$.
-In applications, evaluating these maps may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
-After discretization, $\Aop$ becomes a dense matrix that is typically too large to be built and stored.
-This property is called ``matrix-free'' because at the discrete level it means one can compute matrix-vector products with the matrix representation of $\Aop$, but cannot easily compute individual matrix entries.
-
-%Product-convolution (PC) approximations can (where applicable) approximate high rank operators efficiently.
-
-Our impulse response interpolation procedure may be categorized as a ``point spread function'' (PSF) method that is loosely based on so-called ``product-convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights \cite{PRODCONVLIST}.
-PC and PSF methods have a long history dating back several decades; for background we recommend the following papers: \cite{PRODCONVGOOD}.
-While PC approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, the method we propose is based on an assumption that we call ``local mean displacement invariance'' (see Figure \ref{fig:mean_displacement_invariance}).
-This is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
-Furthermore, PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulse responses.
-We avoid this issue by temporarily excluding impulse responses from the approximation procedure whenever an attempt is made to evaluate the impulse response at a location where it is undefined.
-
-%A single convolution operator may be high rank, so PC approximations can (where applicable) approximate high rank operators efficiently. 
-%The method proposed in this paper is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of scattered points $x_i \in \Omega$ \cite{PRODCONVLIST}.
-%Often PC methods are known as ``point spread function'' (PSF) methods, particularly in optics and image processing.
-There, the ``operator'' $\Aop$ is a physical process of distortion and blurring of light rays passing through an optical system such as a telescope or microscope, and the local convolution kernels are shifted\footnote{The impulse responses are shifted so that each convolution kernel is centered at zero instead of the location of the point source.} impulse responses of the system to point sources of light.
-%, which characterize the local distortion and blurring, and the weighting functions interpolate between these convolution kernels. 
-%PSF is also used as an umbrella term to refer to other approximation methods that focus on interpolating impulse responses or impulse response features.
-For background on PC and PSF methods, we recommend the following papers: \cite{PRODCONVGOOD}.
-
-%The method proposed in this paper is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses $\phi_{x_i}$ associated with a collection of scattered points $x_i \in \Omega$ \cite{PRODCONVLIST}.
-%A common choice for PC methods based on scattered impulse responses is to choose the $x_i$ to be nodes in a regular grid \cite{NAGY}.
-With a regular grid, a large number of $x_i$ are typically required to achieve an accurate approximation.
-%In our previous work, we started with a coarse grid of points $x_i$, then adaptively refined the grid in the regions where the error in the approximation was large \cite{PRODCONVMYPAPER}.
-However, computing each $\phi_{x_i}$ required one action of $\Aop$.
-Even with adaptive refinement, many $x_i$, and hence operator actions of $\Aop$, were required.
-In this paper, we reduce the computational cost further by picking points $x_i$ in a way that allows many $\phi_{x_i}$ to be computed with each operator action of $\Aop$.
-This requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
-The idea of estimating the supports of the functions $\phi_x$ a-priori was inspired by techniques from resolution analysis in seismic imaging, in which
-%In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_x$ is used to estimate the minimum length scale on which features of the parameter can be inferred near the point $x$.
-If the width of $\phi_x$ is ten meters, for instance, then features of the parameter near $x$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
-In \cite{RESOLUTION},
-%the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion  \cite{RESOLUTION}.
-Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Section \ref{sec:intromoments}).
-Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a positive integral kernel.
-
-%\subsection{Motivation: PDE constrained inverse problems}
-%\label{sec:PDE_hessian_motivation}
-
-%Our motivation for this work is approximation of Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs).
+The output of our method is a hierarchical matrix (H-matrix) approximation, $\mathbf{A}_H$, of a discretization of $\Aop$.
+H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank.
+For more details about H-matrices, see Appendix \ref{app:h_matrix}.
+While operator actions of $\Aop$ on vectors are typically computationally expensive, fast H-matrix methods permit us to perform matrix-vector products with $\mathbf{A}_H$ cheaply.
+Furthermore, one may use H-matrix methods to perform useful linear algebra operations involving $\mathbf{A}_H$ that cannot easily be done using $\Aop$, because $\Aop$ is only available matrix-free.
+These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion.
+These H-matrix methods are fast and scale well to large problems.
+The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
+
+\subsection{Motivation: PDE constrained inverse problems}
+\label{sec:PDE_hessian_motivation}
+
+Our motivation for this work is approximation of Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs).
 That is, inverse problems in which one seeks to reconstruct an unknown parameter field, $m$, from noisy observations,
-%\begin{equation*}
-%y = f(m, u(m)) + \text{noise},
-%\end{equation*}
-%which depend on  a state variable $u$. In turn, $u$ depends on $m$ implicitly through the solution of a PDE, which we write generically as follows:
-%\begin{equation}
-%\label{eq:state_pde}
-%0 = g(m,u).
-%\end{equation}
-%Here $u(m)$ denotes the solution of the PDE \eqref{eq:state_pde} as a function of $m$.
+\begin{equation*}
+	y = f(m, u(m)) + \text{noise},
+\end{equation*}
+which depend on  a state variable $u$. In turn, $u$ depends on $m$ implicitly through the solution of a PDE, which we write generically as follows:
+\begin{equation}
+	\label{eq:state_pde}
+	0 = g(m,u).
+\end{equation}
+Here $u(m)$ denotes the solution of the PDE \eqref{eq:state_pde} as a function of $m$.
 In the deterministic approach to inverse problems, one typically finds $m$ as the solution to the minimization problem
-%\begin{equation}
-%\label{eq:minimization_problem}
-%\min_m \quad \frac{1}{2}\|y - f(m,u(m))\|_W^2 + \mathcal{R}(m),
-%\end{equation}
-%where $\|\cdot\|_W$ is weighted norm which depends on the noise covariance, and $\mathcal{R}(m)$ is a regularization term.
+\begin{equation}
+	\label{eq:minimization_problem}
+	\min_m \quad \frac{1}{2}\|y - f(m,u(m))\|_W^2 + \mathcal{R}(m),
+\end{equation}
+where $\|\cdot\|_W$ is weighted norm which depends on the noise covariance, and $\mathcal{R}(m)$ is a regularization term.
 The objective function in optimization problem \eqref{eq:minimization_problem} is defined indirectly via the implicit function theorem.
-As a result, information about the Hessian of the objective function is only accessible via the action of the Hessian on vectors.
-In particular, applying the Hessian to a vector requires solving two linear PDEs which take the form of linearizations of \eqref{eq:state_pde} \cite{HESSIANACTION}.
-More accessible approximations of the Hessian are highly desirable, because they allow for fast solution of \eqref{eq:minimization_problem} via Newton-type methods.
-More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because the inverse of the Hessian locally approximates the Bayesian posterior covariance for $m$ \cite{LAPLACEAPPROXIMATION}.
-
-%\subsection{Background}
-
-%PC approximations have been used in a variety of fields going back several decades \cite{PRODCONVLIST}; for background on PC and PSF methods, we recommend the following papers: \cite{PRODCONVGOOD}.
-The method proposed in this paper is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of scattered points $x_i \in \Omega$ \cite{PRODCONVLIST}.
-A common choice for PC methods in this class is to choose the $x_i$ to be nodes in a regular grid \cite{NAGY}.
-With a regular grid, a large number of $x_i$ are typically required to achieve an accurate approximation.
-In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
-However, computing each impulse response required one action of $\Aop$.
-Even with adaptive refinement, many $x_i$, and hence operator actions of $\Aop$, were required.
-
-% A single convolution operator may be high rank, so PC approximations can (where applicable) approximate high rank operators efficiently.
+As a result, information about the Hessian of the objective function, $\mathcal{H}$, is only accessible via the action of $\mathcal{H}$ action on vectors.
+In particular, applying $\mathcal{H}$ to a vector requires solving two PDEs \cite{HESSIANACTION}.
+More accessible approximations of $\mathcal{H}$ are highly desirable, because they allow for fast solution of \eqref{eq:minimization_problem} via Newton-type methods.
+More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because $\mathcal{H}^{-1}$ locally approximates the Bayesian posterior covariance for $m$.
+
+\subsection{Background}
+
+The most popular matrix-free operator approximation methods are based on low rank approximation \cite{CCGOPAPERS}.
+This includes the Lanczos method, the randomized singular value decomposition \cite{HMTRANDOM}, and CUR approximation/skeletonization \cite{CUR,MAHONEY}.
+However, low rank approximation is not suitable for high rank operators, and many operators of practical interest are high rank.
+In the context of Hessians, methods based on low rank approximation suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator actions are required to form an accurate ``low rank'' approximation \cite{MYDISSERTATION}.
+%The ideal scenario from a scientific perspective (highly informative data) is therefore the worst case scenario from a computational perspective (large computational cost) \cite{MYDISSERTATION}.
+
+%In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
+Product-convolution (PC) approximations are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
+A single convolution operator may be high rank, so PC approximations can (where applicable) approximate high rank operators efficiently.
 PC approximations have been used in a variety of fields going back several decades \cite{PRODCONVLIST}.
@@
 However, ``PSF'' is sometimes used as an umbrella term to refer to other approximation methods that focus on interpolating impulse responses or impulse response features.
+%, including CUR-based low rank approximation \cite{NAGYPSFCUR}.
 For background on PC and PSF methods, we recommend the following papers: \cite{PRODCONVGOOD}.
 
-%The method proposed in this paper is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of scattered points $x_i \in \Omega$ \cite{PRODCONVLIST}.
-A common choice for PC methods in this class is to choose the $x_i$ to be nodes in a regular grid \cite{NAGY}.
+Our method is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i \in \Omega$ \cite{PRODCONVLIST}.
+A popular choice for PC methods in this class is to choose the $x_i$ to be nodes in a regular grid \cite{NAGY}.
+%, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, splines \cite{PRODCONVSPLINES}, or radial basis functions \cite{PRODCONVRBF}. 
 With a regular grid, a large number of $x_i$ are typically required to achieve an accurate approximation.
 In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
-However, computing each impulse response required one action of $\Aop$.
-Even with adaptive refinement, many $x_i$, and hence operator actions of $\Aop$, were required.
-
-%In this paper, we reduce the computational cost further by picking points $x_i$ in a way that allows many $\phi_{x_i}$ to be computed with each operator action of $\Aop$.
+But even with adaptive refinement, many $x_i$, and hence operator actions of $\Aop$, are required.
+
+In this paper, 
+%rather than using adaptive refinement, 
+we reduce the computational cost by picking points $x_i$ in a way that allows many $\phi_{x_i}$ to be computed with each operator action of $\Aop$.
 This requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
-This idea was inspired by techniques from resolution analysis in seismic imaging \cite{RESOLUTION}.
-In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_x$ is used to estimate the minimum length scale on which features of the parameter can be inferred near the point $x$.
-If the width of $\phi_x$ is ten meters, for instance, then features of the parameter near $x$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
+We were inspired by techniques from resolution analysis in seismic imaging \cite{RESOLUTION}.
+In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_x$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $x$.
+If the width of $\phi_x$ is ten meters, then features of the parameter near $x$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
 In \cite{RESOLUTION}, the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion.
-Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Section \ref{sec:intromoments}).
-Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a positive integral kernel.
-
-%Conventionally, PC approximations are applied to vectors using the fast Fourier transform (FFT) \cite{PRODCONVFFT}, which allows PC approximations to be used as fast surrogates for $\Aop$.
+Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Sections \ref{sec:intromoments} and \ref{eq:mean_and_covariance_estimation}).
+Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing, but our method requires that $\Aop$ has a positive integral kernel, while random noise probing does not have this requirement.
+%(and moreover, the seismic Hessian does not have a non-negative integral kernel).
+
+Convolutions can be performed using the fast Fourier transform (FFT) \cite{PRODCONVFFT}, which allows PC approximations to be used as a fast surrogates for $\Aop$.
 Often it is necessary to perform further linear algebra operations involving $\Aop$, such as solving linear systems with $\Aop$ as the coefficient operator.
-One may perform these linear algebra operations using Krylov methods, replacing $\Aop$ with its PC approximation \cite{PRODCONVKRYLOV}.
+Conventionally, one performs these linear algebra operations using Krylov methods, replacing $\Aop$ with its PC approximation \cite{PRODCONVKRYLOV}.
 The Krylov method is accelerated because performing matrix vector products with the product-convolution approximation is typically faster than performing operator actions with $\Aop$.
@@
 Third, the computational cost grows with the number of convolution kernels used.
-Thus, it is desirable to convert the PC approximation to other matrix formats that are amenable to faster linear algebra operations.
+%If a large number of convolution kernels are used, this approach can be slow.
+%Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}. 
+Thus, 
+%after constructing the product-convolution approximation, 
+it is desirable to convert the PC approximation to other matrix formats that are amenable to faster linear algebra operations.
 Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}, because operators that are amenable to PC approximation are often sparse in the wavelet domain.
-%While not commonly done, in theory linear algebra operations involving convolution-product approximations (transposes of PC approximations) could be done using fast methods for pseudo-differential operator symbol calculus \cite{YINGDEMANET} since these are equivalent to low rank approximation of the pseudo-differential symbol of the operator.
-We also note other relevant matrix-free operator approximation methods based on wavelets \cite{HERRMANN} and pseudo-differential operators \cite{DEMANET} which do not use PC approximation.
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[scale=1.0]{mean_displacement_invariance.pdf}
-	\end{center}
-	\caption{An operator is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \mu(x) + \mu(x')\right)$ when $x$ is close to $x'$.
-Here $\mu(x)$ is the mean (center of mass) of $\phi_x$, and similar for $\mu(x')$.}
-	\label{fig:mean_displacement_invariance}
-\end{figure}
-
-%In this paper, we build on previous work \cite{PRODCONVMYPAPER}, in which we converted PC approximations into H-matrix format.
-
-%In exchange for these improvements, the proposed approximation cannot be written as a weighted sum of convolution operators, hence we cannot apply this approximation to vectors via the FFT.
-However, since our approximation is an H-matrix, it can be applied to vectors efficiently via H-matrix methods.
-
-\section{Preliminaries}
-
-Let $\Omega \subset \mathbb{R}^d$ be a bounded domain (in applications typically $\gdim=1$, $2$, or $3$).
-We seek to approximate integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ that may be written in the following form:
-\begin{equation}
-\label{eq:kernel_representation}
-(\Aop u)(v) := \int_\Omega \int_\Omega v(y) \Aker(y,x) u(x) dx dy.
-\end{equation}
-Here the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(v)$ is the result of applying that linear functional to $v \in L^2(\Omega)$.
-The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, which exists in principle but is not easily accessible to us.
-In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define the impulse responses (Section \ref{sec:impulse_response_2}).
-Then we state the conditions on $\Aop$ that our method requires (Section \ref{sec:conditions_2}).
-Finally, we detail finite element discretization (Section \ref{sec:finite_element_kernel}).
-%We write $L^2(\Omega)'$ to denote the topological dual of $L^2(\Omega)$. 
-%The linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
-
-\subsection{Distributions and impulse responses}
-\label{sec:impulse_response_2}
-
-The action of $\Aop$ may be extended to distributions if the kernel $\Phi$ is sufficiently regular.
-We derive the extension as follows.
-Suppose $\rho \in L^2(\Omega)'$, and let $\rho^* \in L^2(\Omega)$ denote the Riesz representative of $\rho$ with respect to the $L^2(\Omega)$ inner product.
-We have
-\begin{subequations}
-	\label{eq:extension_to_distributions}
-	\begin{align}
-		\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Phi(y,x) \genericdistribution^*(x) dx ~dy \\
-		&= \int_\Omega v(y) \int_\Omega \Phi(y,x) \genericdistribution^*(x) dx ~dy \\
-		&= \int_\Omega v(y) \genericdistribution\left(\Phi(y,\cdot)\right) dy, \label{eq:action_on_distribution}
-	\end{align}
-\end{subequations}
-%\begin{align*}
-%\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Phi(y,x) \genericdistribution^*(x) dx dy \\
-%&= \int_\Omega v(y) \int_\Omega \Phi(y,x) \genericdistribution^*(x) dx \\
-%&= \int_\Omega v(y) \genericdistribution\left(\Phi(y,\cdot)\right) dy 
-%%= \left(v,\genericdistribution\left(\Phi(y,\cdot)\right)\right)_{L^2(\Omega)},
-%\end{align*}
-where $\Phi(y,\cdot)$ denotes the function $x \mapsto \Phi(y,x)$.
-Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions, and instead suppose $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ is a distribution.
-In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in \eqref{eq:extension_to_distributions} is not valid.
-However, if $\Phi$ is sufficiently regular such that the function $y \mapsto \rho\left(\Phi(y,~\cdot~)\right)$ is well-defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of \eqref{eq:action_on_distribution} is well-defined.
-Hence, we \emph{define} the action of $\Aop$ on the distribution $\rho$ to be the right hand side of \eqref{eq:action_on_distribution}.
-For convenience of notation, we denote this action by ``$\Aop \genericdistribution^*$,'' even if $\rho^*$ does not exist.
+Convolution-product approximations (transposes of PC approximations) are equivalent to low rank approximation of the pseudo-differential symbol of the operator, so one could perform linear algebra operations involving PC approximations using fast methods for pseudo-differential operator symbol calculus \cite{YINGDEMANET}.
+Other matrix-free operator approximation methods based on wavelets \cite{HERRMANN} and pseudo-differential operators \cite{DEMANET} do not use PC approximation.
+Here, we follow our previous work \cite{PRODCONVMYPAPER}, in which we converted PC approximations into H-matrix format.
+
+Since our end goal is constructing an H-matrix, we are freed from strict adherence to the PC format.
+This allows us to modify the PC approach to address well-known deficiencies.
+PC approximations are based on an assumption of local translation invariance \cite{LOCATRANSLATIONINVARIANCE}, while our method is based on an assumption that we call ``normalized local mean displacement invariance,'' which is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
+PC approximations have well-known issues near boundaries called boundary artifacts, which are caused by the boundary artificially ``cutting off'' impulse responses.
+%One must extend the impulse responses outside of the domain where they are undefined, but
+%. Various extension techniques have been used, such as extension by zero, and reflection across the boundary, yet 
+%no known extension technique is fully satisfactory \cite{BOUNDARYETENSION}. 
+In our method we do not have an issue with boundary artifacts because we temporarily exclude impulse responses from the approximation procedure whenever the approximation procedure attempts to evaluate the impulse response at a location where it is undefined.
+In exchange for these benefits, our approximation cannot be written as a weighted sum of convolution operators, so we cannot apply our approximation to vectors via the FFT.
+This is fine because our approximation is an H-matrix, which can be applied to vectors even faster.
+
+%MOMENTS PERMIT IMPROVED INTERPOLATION.
+IMPROVED INTERPOLATION REQUIRES H-MATRIX.
+H-MATRIX WITH IMPROVED INTERPOLATION ALSO SOLVES BOUNDARY ARTIFACT PROBLEM.
+ALSO DOESNT REQUIRE REGULAR GRID.
+ALSO SCALABLE TO LARGE NUMBER OF IMPULSE RESPONSES.
+FFT HAS FAST MATVEC BUT NOT OTHER OPTIONS.
+
+Since we are building an H-matrix approximation of $\Aop$, it is natural to ask why we need to compute impulse responses and interpolate them. Why not build the H-matrix directly?
+%Can we build the H-matrix directly? In short, no, not efficiently. 
+The reason is that classical H-matrix construction methods \cite{HMATRIXCONSTRUCTION} require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
+Our impulse response interpolation lets us quickly compute approximations to matrix entries, which allows us to use these classical H-matrix construction methods.
+There are newer matrix-free methods for H-matrix approximation based on the so-called ``peeling process'' \cite{LEXINGPEELINGPROCESS}, and these methods have been used to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}.
+While these matrix-free H-matrix construction methods are asymptotically scalable in theory, in practice the required number of matrix vector products is unreasonably large.
+Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods more feasible in the future.
+
+%Product-convolution type approximations of Hessians have been used in a seismic inverse problem in \cite{GEORGSEISMICPRODCONV}, and in an advection-diffusion inverse problem in \cite{PRODCONVMYPAPER}.
+In \cite{DEMANETSEISMIC}, matrix probing \cite{MATRIXPROBING} is used to approximate the Hessian in a seismic inverse problem as the sum of simple pseudodifferential operators.
+Although it is not explicitly mentioned, the approximation in \cite{DEMANETSEISMIC} could be interpreted as a convolution-product approximation (a product-convolution approximation of the transpose of the operator).
+%, which is like a product-convolution approximation, except the order of pointwise multiplications and convolutions is reversed in \eqref{eq:product_convolution}.
+
+\section{Main concepts}
+\label{sec:prerequisites}
+
+In this section we present three main concepts that our method is based on.
+The method itself will be presented in Section \ref{sec:method}.
+First, we define moments of the impulse responses, $\phi_x$, and show how these moments can be computed efficiently (Section \ref{sec:intromoments}).
+Our method will use these moments extensively.
+Second, we present ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{eq:mean_and_covariance_estimation}).
+Our method will use these ellipsoid estimates to choose sets of impulse responses that can be computed in batches.
+Third, we describe a method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}).
+Our method will interpolate impulse responses using this approximation.
+
+\subsection{Impulse response moments}
+\label{sec:intromoments}
+
+The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative integral kernel property. Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
+\begin{equation*}
+\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy
+\end{equation*}
+denote the spatially varying scaling factor,
+%, let 
 %\begin{equation}
-%\label{eq:action_on_distribution}
-%\left(\Aop \genericdistribution^*\right)(v) := \int_\Omega v(y) \genericdistribution\left(\Phi(y, \cdot)\right) dy.
+%\label{eq:scaled_impulse_response}
+%\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)
 %\end{equation}
-%Since $\genericdistribution^*$ may not exist if $\rho$ is a distribution, 
-%We write ``$\Aop \genericdistribution^*$'' to denote ``the action of $\Aop$ on the distribution $\rho$,'' as defined by \eqref{eq:action_on_distribution}.
-
-%We assume that $\Phi$ is sufficiently regular such that the action of $\Aop$ on delta distributions is well-defined, because
-
-%This is derived formally as follows:
-%\begin{align*}
-%\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Phi(y,x) \textrm{``} \genericdistribution^*(x) \textrm{''} dx dy \\
-%&= \int_\Omega v(y) \int_\Omega \Phi(y,x) \textrm{``} \genericdistribution^*(x) \textrm{''} dx \\
-%&= \int_\Omega v(y) \genericdistribution\left(\Phi(y,\cdot)\right) dx.
-%\end{align*}
-%Here $\textrm{``}\rho^*\textrm{''}$ formally denotes the Riesz representative of $\rho$, which, of course, does not exist if $\rho$ is a pure distribution.
-
-%an abuse of notation which views $\rho$ as a ``function'' of $x$, rather than a linear functional.
-That is, $\rho^*$ would be the Riesz representative of $\rho$, but for the obstruction that this Riesz representative does not exist on all of $L^2(\Omega)$.
-
-%For example, the delta distribution $\delta_x$ is defined by $\delta_x(v) = v(x)$, and the action of $\Aop$ on $\delta_x$ is given by
-%\begin{equation}
-%\label{eq:appendix_impulse_response}
-%\left(\Aop \delta_x^*\right)(v)	= \int_\Omega v(y) \Phi(y,x) dy.
-%\end{equation}
-%Recall that the impulse response of $\Aop$ to $\delta_x$, denoted $\impulseresponse_x$, is the Riesz representation of $\Aop \delta_x^*$. 
-%From \eqref{eq:appendix_impulse_response} and the definition of the Riesz representation, we have $\impulseresponse_x = A(~\cdot~,x)$.
-That is, $\impulseresponse_x$ is the function $y \mapsto A(y,x)$.
-
-%We seek to approximate operators $\Aop$ that are accessible \emph{matrix-free}. That is, we have a black-box computational procedure through which we may evaluate the maps
-%\begin{equation*}
-%u \mapsto\Aop u \qquad \text{and} \qquad v \mapsto \Aop^Tv
-%\end{equation*} 
-%for arbitrary functions $u$ and $v$, but we cannot easily compute kernel entries $\Aker(y,x)$.
-In applications, evaluating these maps may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
-After discretization, $\Aop$ becomes a dense matrix that is typically too large to be built and stored.
-This property is called ``matrix-free'' because at the discrete level it means one can compute matrix-vector products with the matrix representation of $\Aop$, but cannot easily compute individual matrix entries.
-
-%Our method, described in Section \ref{sec:method}, belongs to a class of methods that approximate an operator by interpolating impulse responses of the operator associated with a scattered collection of points.
-Let $\delta_x$ denote the delta distribution\footnote{Recall that the delta distribution $\delta_x:\mathcal{D}(\Omega)\rightarrow \mathbb{R}$ is the linear functional defined by $\delta_x(v) = v(x)$ for all $v\in \mathcal{D}(\Omega)$.} (i.e., point source, impulse) centered at the point $x \in \Omega$.
-The \emph{impulse response} of $\Aop$ associated with $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$, 
-\begin{equation}
-\label{eq:impulse_response_delta_action}
-\impulseresponse_x := \left( \Aop \delta_x^* \right)^*,
-\end{equation}
-that is formed by applying $\Aop$ to $\delta_x$, then taking the Riesz representation of the resulting linear functional.
-Using \eqref{eq:action_on_distribution} and the definition of the delta distribution, we see that $\phi_x$ may also be written as the function
-\begin{equation}
-\label{eq:impulse_response_defn}
-\impulseresponse_x(y) = \Aker(y, x),
-\end{equation} 
-which results from keeping $x$ fixed and viewing $\Phi(y,x)$ as a function of $y$.
-
-%The function $\impulseresponse_x$ is called the impulse response because, using \eqref{eq:action_on_distribution} and the definition of the delta function, we have:
-
-%where $\delta_x$ is the the delta distribution (i.e., point source, impulse) centered at $x$, $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.
-
-%Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
-This is described in Appendix \ref{app:distributions}.
-
-\subsection{Required conditions}
-\label{sec:conditions_2}
-
-We focus on approximating operators that satisfy the following conditions: 
-\begin{enumerate}
-	\item The kernel $\Phi$ is sufficiently regular such that $\phi_x$ is a well-defined function for all $x\in \Omega$.
-	\item The supports of the impulse responses $\impulseresponse_x$ are contained in localized regions.
-	\item The integral kernel is non-negative in the sense that $$\Phi(y,x) \ge 0$$ for all $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
-The operator $\Aop$ need not be positive semi-definite to use our method, and positive semi-definite operators need not have a non-negative integral kernel.}
-\end{enumerate}
-Our method may still perform well if these conditions are relaxed slightly.
-For example, it is fine if the support of the impulse response $\impulseresponse_x$ is not perfectly contained in a localized region, so long as the bulk of the ``mass'' of the impulse response is contained in a localized region.
-The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$, as long as it is non-negative for the vast majority of pairs of points $(y,x)$, and as long as the negative numbers are comparatively small.
-If these conditions are violated, our method will incur additional error.
-If these conditions are violated too much, our method may fail.
-Not all operators of interest (and in particular not all Hessians) satisfy these conditions, but many operators of practical interest do satisfy these conditions, and our method is highly effective for approximating these operators.
-%Because of these conditions, we are able to estimate the supports of the impulse responses a-priori by applying $\Aop^T$ \footnote{Recall that the transpose of $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ satisfying $\left(\Aop u\right)(v) = \left(\Aop^T v\right)(u)$ for all $u,v\in L^2(\Omega)$.} to a small number of polynomial functions (Section \ref{sec:intromoments}).
-We use the support estimates to compute scattered impulse responses in batches (Section \ref{sec:get_impulse_response}, also Figure \ref{fig:batches_intro}).
-This batching allows us to recover many impulse responses using a small number of operator actions $u \mapsto \Aop u$.
-The support estimates also allow us to interpolate the impulse responses with a new more sophisticated and accurate interpolation scheme (Sections \ref{sec:local_mean_displacement_invariance} and \ref{sec:approximate_kernel_entries}).
-
-\subsection{Finite element discretization}
-\label{sec:finite_element_kernel}
-
-In computations, functions are discretized and replaced by finite dimensional vectors, and operators mapping between infinite dimensional spaces are replaced by operators mapping between finite dimensional spaces.
-In this paper we discretize using continuous Galerkin finite elements satisfying the Kronecker property (defined below).
-With minor modifications, our method could be used with more general finite element methods, or other discretization schemes such as finite differences or finite volumes.
-
-Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of continuous Galerkin finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space under the $L^2$ inner product, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
-We assume that the finite element basis satisfies the Kronecker property, i.e., $\febasis_i(p_i)=1$ and $\febasis_i(p_j)=0$ if $i\neq j$.
-For $u_h \in V_h$ we write $\bm{u} \in \mathbb{R}^N_\mathbf{M}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis, i.e.,
-\begin{equation}
-\label{eq:fem_coeff_basis}
-u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
-\end{equation}
-Linear functionals $\rho_h \in V_h'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^N_{\mathbf{M}^{-1}}$, with entries 
-\begin{equation*}
-\boldsymbol{\rho}_i = \rho_h(\febasis_i), \qquad i=1,\dots,N. 
-\end{equation*}
-Here $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries
-\begin{equation*}
-\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx, \qquad i,j=1,\dots,N,
-\end{equation*}
-the space $\mathbb{R}^N_\mathbf{M}$ is $\mathbb{R}^N$ with the inner product $(\mathbf{u},\mathbf{v})_\mathbf{M} := \mathbf{u}^T \mathbf{M} \mathbf{v}$, and $\mathbb{R}^N_{\mathbf{M}^{-1}}$ is the analogous space with $\mathbf{M}^{-1}$ replacing $\mathbf{M}$.
-Direct calculation shows that $\mathbb{R}^N_\mathbf{M}$ and $\mathbb{R}^N_{\mathbf{M}^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively.
-%, in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u}, \mathbf{v}\right)_\mathbf{M}$, $\left(\rho_h, \nu_h\right)_{L^2(\Omega)} = \left(\boldsymbol{\rho}, \boldsymbol{\nu}\right)_{\mathbf{M}^{-1}}$, and $\rho_h(u) = \boldsymbol{\rho}^T\mathbf{u}$, for $u_h,v_h \in V_h$ and $\rho_h, \nu_h \in V_h'$.
-
-After discretization, the operator $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$ is replaced by an operator $A_h:V_h \rightarrow V_h'$, which becomes an operator
-\begin{equation*}
-\mathbf{A}:\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}
-\end{equation*}
-under the isomorphism discussed above.
-Our method is agnostic to the computational procedure for approximating $\Aop$ with $\mathbf{A}$.
-What is important is that we do not have direct access to matrix entries $\mathbf{A}_{ij}$.
-Rather, we have a computational procedure that allows us to compute matrix-vector products $\mathbf{u}\mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{v}\mapsto \mathbf{A}^T\mathbf{v}$, and computing these matrix-vector products is costly.
-%, and each of these matrix-vector products is costly to compute. 
-%For example, some computational procedures possess the Galerkin property, in which case $\mathbf{A}_{ij} = \left(\Aop \febasis_j\right)(\febasis_i)$. 
-%Our approximation method is agnostic to the computational procedure for performing these matrix-vector products. 
-%What is important is that $\mathbf{A}$ is dense, only available via matrix-vector products, and each of these matrix-vector products is costly to compute. 
-Of course, matrix entries can be computed via matrix-vector products as $\mathbf{A}_{ij} = \left(\mathbf{A}\mathbf{e}_j\right)_i$,
-%\begin{equation*}
-%\mathbf{A}_{ij} = \left(\mathbf{A}\mathbf{e}_j\right)_i,
-%\end{equation*}
-where $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)^T$ is the length $N$ unit vector with one in the $j$th coordinate and zeros elsewhere.
-But computing the matrix-vector product $\mathbf{e}_j \mapsto \mathbf{A}\mathbf{e}_j$ is costly, and therefore wasteful if we do not use other matrix entries in the $j$th column of $\mathbf{A}$.
-Hence, methods for approximating $\mathbf{A}$ are computationally intractable if they require accessing scattered matrix entries from many different rows and columns of $\mathbf{A}$.
-
-The operator $A_h:V_h \rightarrow V_h'$ can be written in integral kernel form as
-\begin{equation}
-(A_h u_h)(v_h) := \int_\Omega \int_\Omega v_h(y) \Phi_h(y,x) u_h(x) dx dy
-\end{equation}
-for some kernel $\Phi_h$ which we do not know, and which may differ from $\Phi$ due to discretization error.
-Since the functions in $V_h$ are continuous at $x$, the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta}_x \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ with entries $\left(\boldsymbol{\delta}_x\right)_i = \febasis_i(x)$ for $i=1,\dots,N$.
-Additionally, it is straightforward to verify that the Riesz representation, $\rho_h^* \in V_h$, of a functional $\rho \in V_h'$ has coefficient vector
-\begin{equation*}
-\boldsymbol{\rho}^* = \mathbf{M}^{-1} \boldsymbol{\rho}.
-\end{equation*}
-Therefore, the formula for the impulse response from \eqref{eq:impulse_response_delta_action} becomes 
-\begin{equation}
-\label{eq:discrete_kernel}
-\boldsymbol{\impulseresponse}_x = \left(A_h \delta_x^*\right)^* =  \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x,
-\end{equation}
-and the $(y,x)$ kernel entry of $\Phi_h$ may be written as
-\begin{equation}
-\label{eq:approx_kernel_entry_81}
-	\Phi_h(y,x) = \boldsymbol{\delta}_y^T \boldsymbol{\impulseresponse}_x = \boldsymbol{\delta}_y^T \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x.
-\end{equation}
-Now define $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ to be the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
-\begin{equation}
-\label{eq:Akerpcmat_entries}
-\mathbf{\Phi}_{ij} := \Phi_h(p_i, p_j).
-\end{equation}
-Because of the Kronecker property of the finite element basis, we have $\boldsymbol{\delta}_{p_i} = \mathbf{e}_i$. Thus from \eqref{eq:approx_kernel_entry_81} we have
-$
-	\Phi_h(p_i,p_j) = \left(\mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1}\right)_{ij},
-$
-which implies
-\begin{equation}
-\label{eq:boldA}
-	\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
-\end{equation}
-Broadly, our method will construct an H-matrix approximation of $\mathbf{A}$ by forming H-matrix approximations of $\boldsymbol{\Phi}$ and $\mathbf{M}$ (or a lumped mass version of $\mathbf{M}$), then multiplying these matrices with H-matrix methods per \eqref{eq:boldA}.
-The central challenge is that classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Phi}_{ij}$, but these matrix entries are not easily accessible.
-The bulk of our method is therefore dedicated to forming approximations of these matrix entries that can be evaluated rapidly.
-
-%\begin{align*}
-%	\text{continuous} \quad &\longrightarrow \quad \text{discrete} \\
-%	\delta_x \quad &\longrightarrow \quad \boldsymbol{\delta}_x \\
-%	\delta_x^* \quad &\longrightarrow \quad \mathbf{M}^{-1}\boldsymbol{\delta}_x \\
-%	\Aop \delta_x^* \quad &\longrightarrow \quad \mathbf{A}\mathbf{M}^{-1}\boldsymbol{\delta}_x \\
-%	\impulseresponse_x = \left(\Aop \delta_x^*\right)^* \quad &\longrightarrow \quad \mathbf{M}^{-1}\mathbf{A}\mathbf{M}^{-1}\boldsymbol{\delta}_x = \boldsymbol{\impulseresponse}_x \\
-%	\Phi(y,x) \quad &\longrightarrow \quad \boldsymbol{\delta}_y^T\mathbf{M}^{-1}\mathbf{A}\mathbf{M}^{-1}\boldsymbol{\delta}_x
-%\end{align*}
-%Since the Lagrange finite element nodes have the Kronecker property, that is $\febasis_i(p_i) = 1$ and $\febasis_i(p_j) = 1$ for $i \neq j$, the vector $\boldsymbol{\delta}_{p_i}$ is the unit vector with one in the $i$'th component and zeros in all other components.
-Hence
-%\begin{equation*}
-%	\Phi(p_i,p_j) = \left(\mathbf{M}^{-1}\mathbf{A}\mathbf{M}^{-1}\right)_{ij}
-%\end{equation*}
-
-%\subsubsection{Discretized integral kernel}
-%
-%The interpolation of $\Phi$ onto $V_h \otimes V_h$, which we denote by $\Phi^h$, is given as follows:
-%\begin{equation}
-%\label{eq:defn_of_Akerpcmesh}
-%\Phi(y,x) \approx \Phi^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \Phi(p_i,p_j) \febasis_i(y) \febasis_j(x).
-%\end{equation}
-%Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations, yields the following approximation:
-%\begin{equation}
-%\label{eq:kernel_representation_pcmesh}
-%(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
-%= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u},
-%\end{equation}
-%where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
-%\begin{equation}
-%\label{eq:Akerpcmat_entries}
-%\mathbf{\Phi}_{ij} := \Phi(p_i, p_j),
-%\end{equation}
-%Based on \eqref{eq:kernel_representation_pcmesh}, we define the following matrix approximation, $\widetilde{\mathbf{A}} \approx \mathbf{A}$ as follows:
-%\begin{equation}
-%\label{eq:Amatpc_defn}
-%\widetilde{\mathbf{A}} = \bm{M} \mathbf{\Phi} \bm{M}.
-%\end{equation}
-%Note that $\widetilde{\mathbf{A}}$ differs slightly from $\mathbf{A}$ (the Galerkin projection of $\Aop$ onto $V_h$) because of the approximation $\Phi^h \approx \Phi$. 
-%%Hence, $\mathbf{A}$ may differ slightly from the implicitly defined matrix that performs the mapping $\mathbf{u} \rightarrow \mathbf{v}$ corresponding to the discrete version of the operator action $v=\Aop u$.
-%%Because $\Aop$ is only available matrix-free and $\Phi$ is only defined implicitly in terms of $\Aop$, 
-%The matrices $\widetilde{\mathbf{A}}$ and $\mathbf{\Phi}$ are, of course, not easily accessible.  
-%%The primary goal of this paper is building an H-matrix approximation $\mathbf{A}_H$ of $\widetilde{\mathbf{A}}$, using a small number of operator actions of $\Aop$ and $\Aop^T$.
-%
-%%After discretization with finite elements\footnote{While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead, with minor modifications to our method.} (Section \ref{sec:finite_element_kernel}), $\Aop$ and $\Phi$ become large dense matrices $\mathbf{A}$ and $\mathbf{\Phi}$, respectively.
-
-%\subsubsection{Lumped mass matrix and discrete kernel non-negativity}
-
-\paragraph{Lumped mass matrix} At the continuum level, the integral kernel $\Phi$ is assumed to be non-negative.
-However, when we compute impulse responses we are effectively operating with the discrete integral kernel shown in \eqref{eq:approx_kernel_entry_81}.
-%,  which has $(y,x)$ entry
-%\begin{equation*}
-%\boldsymbol{\delta}_y^T \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x.
-%\end{equation*}
-%Even if the discretization $\Aop \rightarrow \mathbf{A}$ preserves integral kernel non-negativity (which it typically will not), 
-The inverse mass matrices in this formula, $\mathbf{M}^{-1}$, will typically contain negative numbers.
-If there are too many negative numbers, or if the negative numbers are too large, our algorithm will be less robust.
-We therefore recommend replacing the mass matrix $\mathbf{M}$ with a positive diagonal \emph{lumped mass} approximation, and using the resulting lumped mass matrix everywhere throughout the approximation method.
-For typical finite element meshes, lumped mass matrix approximations are spectrally equivalent to the mass matrix \cite{MASSLUMP2}.
-In our numerical results, we use the lumped mass matrix constructed by replacing off-diagonal entries of the mass matrix by zero.
-Other more advanced mass lumping techniques may be used \cite{MASSLUMP2}.
-
-\section{Key innovations}
-\label{sec:prerequisites}
-
-In this section we present two key innovations that our method is based on.
-%The approximation algorithm will be presented in Section \ref{sec:method}. 
-First, we define moments of the impulse responses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{sec:intromoments}).
-Our method will use these ellipsoid estimates to choose sets of impulse responses that can be computed in batches.
-Second, we describe an improved method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}).
-Our method will interpolate impulse responses using this approximation.
-
-\subsection{Impulse response moments and ellipsoid support estimate}
-\label{sec:intromoments}
-
-The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative integral kernel property. Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
-\begin{equation}
-\label{eq:define_vol}
-\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy
-\end{equation}
-denote the spatially varying scaling factor, and for $i,j=1,\dots,d$ define $\mu:\Omega \rightarrow \mathbb{R}^d$ and $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$ as follows:
-\begin{align}
-\spatialmean^i(x) :=& \int_\Omega (\impulseresponse_p(y) / V(x)) y^i ~dy \label{eq:define_mean} \\
-\spatialcov^{ij}(x) :=& \int_\Omega (\impulseresponse_p(y) / V(x)) \left(y^i - \mu^i(x)\right) \left(y^j - \mu^j(x)\right) ~dy, \label{eq:define_cov}
-\end{align}
-where $q^i$ denotes the $i^\text{th}$ component of the vector $q$, and $M^{ij}$ denotes the $(i,j)$ entry of the matrix $M$.
-The vector $\mu(p)\in \mathbb{R}^d$ and the matrix $\Sigma(x) \in \mathbb{R}^{d \times d}$ are the mean and covariance of the normalized version of $\impulseresponse_x$, respectively.
+%denote the normalized version of $\impulseresponse_x$. 
+and let $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d},$
+\begin{equation*}
+\spatialmean(x) := \mathbb{E}(\impulseresponse_x / V(x)), \qquad
+\spatialcov(x) := \Var(\impulseresponse_x /V(x)),
+\end{equation*}
+to be the spatially varying mean and covariance of the normalized version of $\impulseresponse_x$, respectively.
 Imagine $\phi_x$ as a mound of dirt piled on a localized region in $\Omega$, with $\phi_x(y)$ being the height of the dirt mound at location $y$.
@@
 
+%The ellipsoid $E_x$ is not immediately available. 
+%(recall $\gdim$ is the spatial dimension of $\Omega$).
 The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}.
 Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
-However, this direct approach is not feasible because we need to know $V(x)$, $\mu(x)$, and $\Sigma(x)$ before we compute $\phi_x$, in order choose the point $x$.
-Computing $\phi_x$ in order to determine $V(x)$, $\mu(x)$, and $\Sigma(x)$ would be extremely computationally expensive, and defeat the purpose of our algorithm, which is to reduce the computational cost by computing impulse responses in batches.
-Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ indirectly, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions.
-This may be motivated by analogy to matrices.
-If $\mathbf{B}\in \mathbb{R}^{N \times N}$ is a matrix with $i$th column $\mathbf{b}_i$ and $\mathbf{f} \in \mathbb{R}^N$, then
-\begin{equation*}
-\mathbf{B}^T \mathbf{f} = \begin{bmatrix}
-\horzbar & \mathbf{b}_1^T & \horzbar \\
-\horzbar & \mathbf{b}_2^T & \horzbar \\
-& \vdots & \\
-\horzbar & \mathbf{b}_N^T & \horzbar
-\end{bmatrix}
-\mathbf{f} = 
-\begin{bmatrix}
-\mathbf{b}_1^T \mathbf{f} \\
-\mathbf{b}_2^T \mathbf{f} \\
-\vdots \\
-\mathbf{b}_N^T \mathbf{f}
-\end{bmatrix}.
-\end{equation*}
-By computing one matrix-vector product of $\mathbf{B}^T$ with $\mathbf{f}$, we compute the inner product of each column of $\mathbf{B}$ with $\mathbf{f}$ simputaneously.
-The operator case is analogous, with the impulse response $\phi_x$ taking the place of matrix columns.
-We have
-\begin{equation}
-\label{eq:operator_simultaneous_ips}
-\left(\Aop^T f\right)^*(x) = \int_\Omega K(y,x) f(y) dy = \left(\phi_x, f\right)_{L^2(\Omega)}.
-\end{equation}
-By computing one operator application of $\Aop^T$ on $f$, we compute the inner product of each impulse response $\phi_x$ with $f$, for all points $x$ simultaneously.
-
-Let $C$, $L^i$, and $Q^{ij}$ be the following constant, linear, and quadratic functions:
+But this direct approach is computationally infeasible because it requires one operator action of $\Aop$ per point $x$, and our method will need to know $V(x)$, $\mu(x)$, and $\Sigma(x)$ at all points $x$.
+Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions.
+Specifically, let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$, and let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
 \begin{equation*}
@@
 L^i(x) := x^i, \qquad
-Q^{ij}(x) = x^i x^j
-\end{equation*}
-for $i=1,\dots,d$. Using the definition of $V$ in \eqref{eq:define_vol} and using \eqref{eq:operator_simultaneous_ips}, we have
-\begin{equation*}
-	V(x) = \int_\Omega \phi_x(y) \cdot 1 ~ dy = \left(\phi_x, C\right)_{L^2(\Omega)} = \left(\mathcal{A}^T C\right)^*(x).
-\end{equation*}
-Hence, we may compute $V(x)$ for all $x$ simultaneously by applying $\Aop^T$ to $C$.
-Analogous manipulations show that $\mu(x)$ may be computed for all points $x$ simultaneously by applying $\Aop^T$ to the functions $L^i$, and $\Sigma(x)$ may be computed for all for all points $x$ simultaneously by applying $\Aop^T$ to the functions $Q^{ij}$.
-We have
+Q^{ij}(x) = x^i x^j.
+\end{equation*}
+The following results (derived in Appendix \ref{app:proofs}) follow straightforwardly from the definitions:
 \begin{subequations}
@@
 \end{subequations}
-for $i,j=1,\dots, \gdim$.
+for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$.
 Here $f/g$ denotes pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denotes pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
-The resulting method for computing discretized versions of $V$, $\mu$, and $\Sigma$ is shown in Algorithm \ref{alg:varhpi_mean_cov}.
+The resulting method form computing $V$, $\mu$, and $\Sigma$ is shown in Algorithm \ref{alg:varhpi_mean_cov}.
 
@@
 	{	
-		\Input{Operator $\bm{A}^T$}
-		\Output{$\bm{\spatialvol}\in \mathbb{R}^\fedim$, $\bm{\spatialmean}\in \mathbb{R}^{\gdim \times \fedim}$, and $\bm{\spatialcov} \in \mathbb{R}^{\gdim \times \gdim \times \fedim}$}
-
-		Form vector $\bm{C} \in \mathbb{R}^\fedim_{\bm{M}}$ by either projecting or interpolating constant function $C(x)=1$ onto $V_h$
-
-		$\bm{\spatialvol} \gets \bm{M}^{-1}\bm{A}^T \bm{C}$
-
+		\Input{Operator $\Aop$}
+		\Output{$\spatialvol$, $\spatialmean$, and $\spatialcov$}
+
+		\tcp{Compute scaling factor $\spatialvol$}
+
+		Form constant function $C(x)=1$
+
+		Compute $\spatialvol = \left(\Aop^T C\right)^*$
+
+		\tcp{Compute mean $\spatialmean$}
 		\For{$i=1,2,\dots,\gdim$}{
-			Form vector $\bm{L}^i \in \mathbb{R}^\fedim_{\bm{M}}$ by either projecting or interpolating linear function $L^i(x) = x^i$ onto $V_h$
-
-			$\bm{\spatialmean}^i \gets \left(\bm{M}^{-1}\bm{A}^T \bm{L}^i\right) / \bm{\spatialvol}$
+			Form linear function $L^i(x) = x^i$
+
+			Compute $\spatialmean^i = \left(\Aop^T L^i\right)^* / \spatialvol$
 		}
+		\tcp{Compute covariance $\spatialcov$}
 		\For{$i=1,2,\dots,\gdim$}{
 			\For{$j=1,\dots,i$}{
-				Form quadratic function $\bm{Q}^{ij} = x^i x^j$
-
-				$\bm{\spatialcov}^{ij} \gets \left(\bm{M}^{-1}\bm{A}^T \bm{Q}^{ij}\right) / \bm{\spatialvol} - \bm{\spatialmean}^i\cdot \bm{\spatialmean}^j$
-
-				$\bm{\spatialcov}^{ji} \gets \bm{\spatialcov}^{ij}$
+				Form quadratic function $Q^{ij}(x) = x^i x^j$
+
+				Compute $\spatialcov^{ij} = \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j$
+
+				Set $\spatialcov^{ji} = \spatialcov^{ij}$
 
@@
 	}
-	\caption{Compute discretized scaling factor $\mathbf{V}$, mean $\boldsymbol{\mu}$, and covariance $\boldsymbol{\Sigma}$}
+	\caption{Compute scaling factor $\spatialvol$, mean $\spatialmean$, and covariance $\spatialcov$}
 	\label{alg:varhpi_mean_cov}
@@
 
-%\begin{algorithm2e}
-%	\SetAlgoNoLine
-%	\SetKwInOut{Input}{Input}
-%	\SetKwInOut{Output}{Output}
-%	{	
-%		\Input{Operator $\Aop$}
-%		\Output{$\spatialvol$, $\spatialmean$, and $\spatialcov$}
-%		
-%		\tcp{Compute scaling factor $\spatialvol$}
-%		
-%		Form constant function $C(x)=1$
-%		
-%		Compute $\spatialvol = \left(\Aop^T C\right)^*$
-%		
-%		\tcp{Compute mean $\spatialmean$}
-%		\For{$i=1,2,\dots,\gdim$}{
-%			Form linear function $L^i(x) = x^i$
-%			
-%			Compute $\spatialmean^i = \left(\Aop^T L^i\right)^* / \spatialvol$
-%		}
-%		\tcp{Compute covariance $\spatialcov$}
-%		\For{$i=1,2,\dots,\gdim$}{
-%			\For{$j=1,\dots,i$}{
-%				Form quadratic function $Q^{ij}(x) = x^i x^j$
-%				
-%				Compute $\spatialcov^{ij} = \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j$
-%				
-%				Set $\spatialcov^{ji} = \spatialcov^{ij}$
-%				
-%			}
-%		}
-%		
-%	}
-%	\caption{Compute scaling factor $\spatialvol$, mean $\spatialmean$, and covariance $\spatialcov$}
-%	\label{alg:varhpi_mean_cov}
-%\end{algorithm2e}
-
-%\subsection{Impulse response support ellipsoids}
-%\label{eq:mean_and_covariance_estimation}
-
-Next, we make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
+%We present this efficient method of computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ in Proposition \ref{thm:vol_mean_cov} and summarize the method for computing these moments in Algorithm \ref{alg:varhpi_mean_cov}.
+
+%\begin{prop}[Impulse response moments]
+%	\label{thm:vol_mean_cov}
+%	%	Let $\spatialvol:\Omega \rightarrow \mathbb{R}$, $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$, and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$, be the following functions:
+%	%	\begin{equation*}
+%	%	\spatialvol(x) := \int_{\Omega} \impulseresponse_x(y) dy, \qquad
+%	%	\spatialmean(x) := \mathbb{E}(\widehat{\impulseresponse}_x), \qquad
+%	%	\spatialcov(x) := \Var(\widehat{\impulseresponse}_x),
+%	%	\end{equation*}
+%	%	where $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denotes the normalized version of $\impulseresponse_x$, $\mathbb{E}(\widehat{\impulseresponse}_x)$ denotes the mean of $\widehat{\impulseresponse}_x$, and $\Var(\widehat{\impulseresponse}_x)$ denotes the variance of $\widehat{\impulseresponse}_x$.
+%	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^\gdim\right)$.
+Let $C$, $\{L^i\}_{i=1}^\gdim$, and ${\{Q^{ij}\}_{i=1}^\gdim}_{j=1}^\gdim$ be the following constant, linear, and quadratic functions:
+%	\begin{equation*}
+%	C(x) := 1, \qquad
+%	L^i(x) := x^i, \qquad
+%	Q^{ij}(x) = x^i x^j.
+%	\end{equation*}
+%	Let $f/g$ denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
+The following results hold:
+%	\begin{subequations}
+%		\label{eq:vol_mean_var}
+%		\begin{align}
+%		\spatialvol =& \left(\Aop^T C\right)^* \\
+%		\spatialmean^i =& \left(\Aop^T L^i\right)^* / \spatialvol \\
+%		\spatialcov^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j
+%		\end{align}
+%	\end{subequations}
+%	for $i=1,\dots, \gdim$, $j=1,\dots,\gdim$.
+%\end{prop}
+
+%The proof of Proposition \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}. The proof is straightforward, and follows directly from the definitions.
+
+\subsection{Impulse response support ellipsoids}
+\label{eq:mean_and_covariance_estimation}
+
+%Because of the non-negative impulse response property, for each $x \in \Omega$, the impulse response $\impulseresponse_x$ is a non-negative function $y \mapsto \impulseresponse_x(y)$.
+Hence, $\impulseresponse_x$ is a scaled probability distribution.
+Let $\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy$ denote the scaling factor (the ``volume'' of $\impulseresponse_x$), let $\widehat{\impulseresponse}_x := \impulseresponse_x / \spatialvol(x)$ denote the normalized version of $\impulseresponse_x$, and let $\spatialmean(x)$ and $\spatialcov(x)$ denote the mean and covariance of $\widehat{\impulseresponse}_x$, respectively.
+
+We make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
 \begin{equation}
@@
 
-The quantity $\tau$ is a parameter that must be chosen appropriately.
+The quantity $\tau$ is a parameter that must be chosen by the user.
 The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$.
-However, in Section \ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (the more ellipsoids the better), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
-In practice, we find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results.
+However, in Section \ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
+We find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results.
 The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.
+
+%To determine $E_x$, we must compute the scalar $\spatialvol(x)$, the length-$\gdim$ vector $\spatialmean(x)$, and the $\gdim \times \gdim$ matrix $\spatialcov(x)$.
+In Section \ref{eq:mean_and_covariance_estimation} we will see that it is possible to compute $V(x)$, $\mu(x)$, and $\Sigma(x)$ for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of polynomial functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$).
 
@@
 \phi_x(y)/V(x) &\approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right) / V(x'). \label{eq:translate4}
+%\frac{\phi_x(y)}{V(x)} &\approx \frac{\phi_{x'}\left(y-\mu(x)+\mu(x')\right)}{V(x')}. \label{eq:translate4}
 \end{align}
 Approximation \eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close. 
-Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank interpolative factorization/CUR approximation \cite{CUR,NYSTROM}.
-However, the required rank in the low rank approximation can be large, which makes algorithms based on \eqref{eq:translate1} expensive.
+%This approximation follows from the assumption that $\Phi$ is smooth. 
+Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank interpolative factorization/Nystrom approximation \cite{CUR,NYSTROM}.
+% of the form
+%\begin{equation}
+%	\label{eq:low_rank_interpolation}
+%	\Phi(y,x) \approx \sum_{i=1}^r w_i(x) \phi_{x_i}(y)
+%\end{equation}
+%where $\phi_{x_i}$ are known impulse responses at a collection of points $x_i$ scattered throughout $\Omega$, and $w_i$ are smooth weighting functions that interpolate the $\phi_{x_i}$. 
+%In problems of practical interest, \eqref{eq:translate1} typically holds \emph{asymptotically} in the limit as $x'$ approaches $x$. 
+However, 
+%even if $\Phi$ is smooth (which is typically the case asymptotically),
+%the relevant Lipschitz constant for continuity, and hence 
+the required rank in the low rank approximation can be large, which makes algorithms based on \eqref{eq:translate1} computationally expensive.
+This is the case for Hessians in PDE constrained inverse problems with highly informative data, as discussed in Section \ref{sec:PDE_hessian_motivation}.
+%As the data become more informative, typically the ``mass'' of $\Phi(y,x)$ concentrates nearer and nearer to the ``diagonal'' $y=x$, the Lipschitz constant increases, and the required rank in the approximation increases.
+%Moreover, if $\Aop$ is numerically low rank, then one would typically use more accurate and robust matrix-free methods for forming low-rank approximations, such as randomized SVD or Lanczos.
 
 Operators that satisfy \eqref{eq:translate2} are called ``locally translation-invariant'' because integral kernel entries $\Phi(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$.
-It is straightforward to show that if equality holds in \eqref{eq:translate2}, then $\Aop$ is a convolution operator.
-Locally translation-invariant operators act like convolutions locally, and can therefore be well-approximated by PC approximations.
+Locally translation-invariant operators act like convolutions locally, and can therefore be well-approximated by so-called product-convolution approximations---weighted sums of convolution operators with spatially varying weights.
+Product-convolution approximations can be high-rank even if the number of convolution operators used in the approximation is small.
+
+% Kernels for locally translation invariant operators can be well-approximated by decompositions of the form
+%\begin{equation}
+%	\Phi(y,x) \approx \sum_{i=1}^r w_i(x) \phi_{x_i}(y-x)
+%\end{equation}
+%which correspond to operator approximations of the form
+%\begin{equation}
+%\label{eq:product_convolution}
+%\left(\Aop~u\right)(v) \approx \left( v,~ \sum_{i=1}^\convrank \convkernel_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)}.
+%\end{equation}
+%Here $f \ast g$ denotes convolution, and $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$ (i.e., $\left(f \cdot g\right)(x) = f(x) g(x)$).
+Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution.
+Product-convolution approximations can be high-rank even if $r$ is small.
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale=1.0]{mean_displacement_invariance.pdf}
+	\end{center}
+	\caption{An operator is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \mu(x) + \mu(x')\right)$ when $x$ is close to $x'$.}
+	\label{fig:mean_displacement_invariance}
+\end{figure}
 
@@
 
-\section{Operator approximation algorithm}
+\section{Method}
 \label{sec:method}
 
-%In this section, we present our method for constructing accessible H-matrix approximations, $\mathbf{A}_H$ and $\mathbf{\Phi}_H$ of $\mathbf{A}$ and $\mathbf{\Phi}$, respectively, using a small number of operator actions of $\Aop$ and $\Aop^T$.
-
-Algorithm \ref{alg:varhpi_mean_cov} from Section \ref{sec:intromoments} allows us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} uses $V$, $\mu$, and $\Sigma$ to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
+After discretization with finite elements\footnote{While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead, with minor modifications to our method.} (Section \ref{sec:finite_element_kernel}), $\Aop$ and $\Phi$ become large dense matrices $\mathbf{A}$ and $\mathbf{\Phi}$, respectively.
+Because $\Aop$ is only available matrix-free and $\Phi$ is only defined implicitly in terms of $\Aop$, the matrices $\mathbf{A}$ and $\mathbf{\Phi}$ are not easily accessible.
+Directly computing and storing these matrices is computationally intractible.
+In this section, we present our method for constructing accessible H-matrix approximations, $\mathbf{A}_H$ and $\mathbf{\Phi}_H$ of $\mathbf{A}$ and $\mathbf{\Phi}$, respectively, using a small number of operator actions of $\Aop$ and $\Aop^T$.
+
+%Fourth, we discuss how discretization of continuous functions by finite elements leads to matrix discretizations $\mathbf{A}$ and $\mathbf{\Phi}$ of $\Aop$ and $\Phi$, respectively (Section \ref{sec:finite_element_kernel}).
+The primary function of our method is to compute H-matrix approximations of these matrices.
+%While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead with minor modifications.
+
+\subsection{Overview}
+\label{sec:method_overview}
+
+Algorithm \ref{alg:varhpi_mean_cov} from Section \ref{sec:intromoments} allows us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} from Section \ref{eq:mean_and_covariance_estimation} uses $V$, $\mu$, and $\Sigma$ to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
 This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}).
@@
 The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section \ref{sec:local_mean_displacement_invariance}.
+%Specifically, we make the assumption that
+%\begin{equation}
+%\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}(y - \mu(x) + \mu(x_i))
+%\end{equation}
+%if $x_i$ is close to $x$. This is shown in Figure \ref{fig:mean_displacement_invariance}. Motivation for this assumption is discussed in Section \ref{eq:local_mean_displacement_invariance}. 
 We approximate $\Phi(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{V(x)}{V(x_i)}\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for a collection of sample points $x_i$ near $x$.
@@
 
-If $\Aop$ is symmetric, we take advantage of symmetry to construct a more accurate approximation at negligible additional cost by combining information from impulse responses of both $\Aop$ and $\Aop^T$ in the interpolation procedure (Section \ref{sec:symmetric_improvements}).
+If $\Aop$ is symmetric, we take advantage of symmetry to construct a more accurate approximation at negligible additional cost
+%. In this case, 
+%impulse responses of $\Aop$ are also impulse responses of $\Aop^T$, which may be interpolated in the variable $y$ instead of $x$ to yield a different approximation of $\Phi(y,x)$. 
+%we combine 
+by combining information from impulse responses of both $\Aop$ and $\Aop^T$ in the interpolation procedure (Section \ref{sec:symmetric_improvements}).
+% for approximating $\Phi(y,x)$, to get a more accurate approximation at negligible additional cost.
 Intuitively, impulse responses of $\Aop$ can be thought of as ``columns'' of $\Phi$, impulse responses of $\Aop^T$ can be thought of as ``rows'' of $\Phi$, and $\Phi(y,x)$ can be thought of as an unknown entry of $\Phi$.
@@
 The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}. The cost to construct $\mathbf{A}_H$ is discussed in Section \ref{sec:overall_cost}.
+
+%Often we are interested in approximating operators $\Aop$ that are symmetric, or symmetric positive semi-definite.
+
+%Forming the H-matrix approximation $\boldsymbol{\Phi}_H$ requires evaluating kernel entries, $\Phi(y,x)$, at arbitrary points $(y,x)$, and this is done by interpolating the known impulse responses $\{\impulseresponse_{x_i}\}_{i=1}^r, \{\impulseresponse'_{y_i}\}_{i=1}^r$.
+%
+%%For each batch, we solve an ellipsoid packing problem to choose as many points $x_i$ as possible, while ensuring that the supports of the impulse responses $\impulseresponse_{x_i}$ do not overlap for points $x_i$ in the batch.
+Then we apply $\Aop$ to a weighted sum of delta distributions centered at the points $x_i$ (Dirac comb).
+From the resulting
+%
+%
+%Our overall strategy is as follows.
+First, we compute the functions $V:\Omega \rightarrow \mathbb{R}$, $\mu:\Omega \rightarrow \mathbb{R}^d$, $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$ by applying $\Aop^T$ to a small number of polynomial functions.
+Next, we solve ellipsoid packing problems to choose batches of points $x_i$ such that
+%to compute impulse responses $\{\impulseresponse_{x_i}\}_{i=1}^r$ of $\Aop$ to point sources at a collection of points $x_i$ using a small number of operator actions of $\Aop$.
+If $\Aop$ is non-symmetric, we also compute impulse responses $\{\impulseresponse'_{y_i}\}_{i=1}^r$ of $\Aop^T$ to point sources at a collection of points $y_i$.
+We then form a H-matrix approximation, $\boldsymbol{\Phi}_H$, of the kernel, $\Phi$, and use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation of $\Aop$, denoted $\mathbf{A}_H$.
+%Schematically the overall framework proceeds as follows:
+%\begin{equation*}
+%\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \{\impulseresponse_{x_i}\}_{i=1}^r \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
+%\end{equation*}
+%We will define product-convolution approximations in Section \ref{sec:prodconv_intro}.
+H-matrices are discussed in Appendix \ref{app:h_matrix}.
+Our method is summarized in Section \ref{sec:overview_intro}, and described in detail in Section \ref{sec:method}.
+%
+%The novel contributions of our method are in the way that we form the product-convolution approximation ($\Aop \rightarrow \AopPc$).
+This is the part of the method where we use the four properties of $\Aop$ described above.
+Fast conversion of the discretized version of $\AopPc$ to H-matrix format ($\AopPc \rightarrow \AmatPc$) is possible because integral kernel entries for $\AopPc$ are easily accessible (unlike integral kernel entries for $\Aop$).
+Once in H-matrix format, one may use H-matrix methods to perform further useful linear algebra operations involving $\AmatPc$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
+These H-matrix methods are fast and scale well to large problems.
+
+%In this section we detail our method for computing the H-matrix approximation, $\mathbf{A}_H$, of $\Aop$, as outlined in Section \ref{sec:intro_overview}.
+%The supports of the impulse responses $\phi_x$ are estimated to be contained within ellipsoids (Section \ref{eq:mean_and_covariance_estimation}).
+The shapes of the ellipsoids depend on spatially varying volume, mean, and covariance functions, $V,\mu,\Sigma$, which we calculate by applying $\Aop$ to a small number of polynomial functions (Section \ref{sec:moments}).
+We use the ellipsoid support estimates in a greedy ellipsoid packing procedure to choose batches, $\pointbatch_b$, of sample points $x_i$ (Section \ref{sec:sample_point_selection}).
+Batches, $\eta_b$, of impulse responses, $\phi_{x_i}$, are computed by applying $\Aop$ to Dirac combs associated with the batches of sample points $S_b$.
+The impulse response batch $\eta_b$ contains the information about all impulse responses $\phi_{x_i}$ associated with points $x_i \in S_b$.
+Hence, once a small number of impulse response batches $\eta_b$ are computed, they can be used to perform pointwise evaluations $z \mapsto \phi_{x_i}(z)$ for a large number of impulse responses, $\phi_{x_i}$ (Section \ref{sec:get_impulse_response}).
+%
+%In Section \ref{eq:mean_and_covariance_estimation} we describe our ellipsoid estimate for the support of $\phi_x$.
+The ellipsoid shape depends on the functions $V,\mu,\Sigma$.
+We show how to compute these functions in Section \ref{sec:moments}.
+The ellipsoid support estimates are then used in a greedy ellipsoid packing procedure to choose batches, $\pointbatch_b$, of sample points $x_i$ (Section \ref{sec:sample_point_selection}).
+%This section is organized as follows:
+%\begin{description}
+%	\item[Section \ref{eq:mean_and_covariance_estimation}:] We describe our ellipsoid estimate for the support of $\phi_x$.
+%	\item[Section \ref{sec:moments}:] We show how to compute $V,\mu,\Sigma$.
+%	\item[Section \ref{sec:sample_point_selection}:] We present the greedy ellipsoid packing procedure used to choose batches, $\pointbatch_b$, of sample points $x_i$.
+%	\item[Section \ref{sec:get_impulse_response}:] We describe how to compute $\phi_{x_i}$'s, in batches, $\eta_b$.
+%	\item[Section \ref{sec:approximate_kernel_entries}:] We show how we interpolate $\phi_{x_i}$'s to approximate $\Phi(y,x)$.
+%	\item[Section \ref{sec:finite_element_kernel}:] We discuss finite element discretization of $\Phi$.
+%	\item[Section \ref{sec:h_matrix_construction_short}:] We describe how we construct the H-matrices $\mathbf{\Phi}_H$ and $\mathbf{A}_H$.
+%	\item[Section \ref{sec:overall_cost}:] We review the total computational cost to construct $\mathbf{A}_H$, and list the cost to perform further linear algebra tasks involving $\mathbf{A}_H$ using H-matrix methods.
+%\end{description}
+%The optional method for making $\mathbf{A}_H$ positive semi-definite is described in Appendix \ref{sec:make_spd}.
+The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}.
+
+%we compute the impulse response moments $V,\mu,\Sigma$ and use these moments to estimate the supports of the impulse responses $\phi_x$ (Section \ref{sec:moments}), the greedy ellipsoid packing procedure we use to choose batches of sample points $x_i$ (Section \ref{sec:sample_point_selection}), how we compute impulse responses, $\phi_{x_i}$, in batches (Section \ref{sec:get_impulse_response}), how we interpolate the $\phi_{x_i}$ to approximate integral kernel entries $\Phi(y,x)$ (Section \ref{sec:approximate_kernel_entries}), finite element discretization of $\Phi$ (Section \ref{sec:finite_element_kernel}), and how we construct the H-matrix approximation $\mathbf{A}_h$ of the operator $\mathcal{A}$.
+We also describe how to (optionally) modify $\mathbf{A}_H$ to force it to be symmetric positive semi-definite.
+%The complete algorithm is shown in Algorithm \ref{alg:construct_Atilde}.
 
@@
 
-		Compute $\bm{\spatialvol}$, $\bm{\spatialmean}$, and $\bm{\spatialcov}$ using Algorithm \ref{alg:varhpi_mean_cov}.
+		Compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ using Algorithm \ref{alg:varhpi_mean_cov}.
 
@@
 	}
-	\caption{Construct H-matrix approximation}
+	\caption{Construct product-convolution H-matrix}
 	\label{alg:construct_Atilde}
@@
 
+\subsection{Finite element discretization}
+\label{sec:finite_element_kernel}
+
+Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
+That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
+For $u_h \in V_h$ we write $\bm{u}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis $\febasis_1, \dots, \febasis_\fedim$.
+I.e.,
+\begin{equation}
+\label{eq:fem_coeff_basis}
+u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
+\end{equation}
+
+%Further, let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
+That is, the points such that $\febasis_i(p_j)$ equals one if $i=j$, and zero otherwise.
+
+%We define the finite element approximation of the kernel, $\Phi^h \approx \Phi$, as follows:
+The interpolation of $\Phi$ onto $V_h \otimes V_h$, which we denote by $\Phi^h$, is given as follows:
+\begin{equation}
+\label{eq:defn_of_Akerpcmesh}
+\Phi(y,x) \approx \Phi^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \Phi(p_i,p_j) \febasis_i(y) \febasis_j(x).
+\end{equation}
+%The kernel $\Phi^h$ is the interpolation of $\Phi$ onto $V_h \otimes V_h$.
+Error in the overall kernel approximation, $\widetilde{\Phi}^h \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$.
+Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations (see Appendix \ref{app:proofs} for details), yields the following approximation:
+\begin{equation}
+\label{eq:kernel_representation_pcmesh}
+(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
+= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u}.
+\end{equation}
+%While $\Aop$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\Aop$ to functions in the finite element space. 
+where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
+\begin{equation}
+\label{eq:Akerpcmat_entries}
+\mathbf{\Phi}_{ij} := \Phi(p_i, p_j),
+\end{equation}
+and $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ is the sparse finite element mass matrix which has entries
+\begin{equation*}
+\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx.
+\end{equation*}
+%Matrix entries $\Phi(p_i, p_j)$ are expensive to compute because $\Aop$ is only available matrix-free.
+%Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, and \eqref{eq:Akerpcmat_entries}, the definition of the mass matrix, and performing algebraic manipulations, it is straightforward but tedious to show that
+%\begin{equation*}
+%	\left(\AopPcMesh u_h \right)(v_h) = \bm{v}^T \bm{M} \AkerPcMat \bm{M} \bm{u}.
+%\end{equation*}
+Our matrix representation, $\mathbf{A}$, of $\Aop$ is therefore given by
+\begin{equation}
+\label{eq:Amatpc_defn}
+\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
+\end{equation}
+Note that $\mathbf{A}$ is not exactly the Galerkin projection of $\Aop$ onto $V_h$ because of the approximation $\Phi^h \approx \Phi$.
+%because of the approximation $\Phi^h \approx \Phi$. 
+Hence, $\mathbf{A}$ may differ slightly from the implicitly defined matrix that performs the mapping $\mathbf{u} \rightarrow \mathbf{v}$ corresponding to the discrete version of the operator action $v=\Aop u$.
+%It is computationally intractible to build and store the matrices $\mathbf{A}$ and $\mathbf{\Phi}$, and individual matrix entries $\mathbf{A}_{ij}$ and $\mathbf{\Phi}_{ij}$ are not easily accessible.
+In the remainder of this section, we show how to build accessible approximations of these matrices.
+
+%The discrete version of the operator action $v = \Aop u$ is a mapping of vectors $\mathbf{u} \rightarrow \mathbf{v}$, which is done through an implicit computational procedure.
+Typically the implicitly defined matrix that performs this mapping is the Galerkin projection of $\Aop$ onto $V_h$.
+
+%We neglect this error because it is the same magnitude as the error in finite element approximation of functions $u_h \approx u$. 
+%Note that the matrices $\mathbf{A}$ and $\mathbf{\Phi}$ are dense $N \times N$ matrix that only exist in theory.
+Building and storing these matrices is computationally intractible.
+Computing individual matrix entries $\mathbf{A}_{ij}$ and $\mathbf{\Phi}_{ij}$ is computationally expensive.
+
 \subsection{Sample point selection via greedy ellipsoid packing}
@@
 We choose sample points, $x_i$, in batches $\pointbatch_k$. 
+%A greedy algorithm is used to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other.
+We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\impulseresponse_x$ (Section \ref{eq:mean_and_covariance_estimation}).
+The support of $\impulseresponse_x$ is estimated to be contained within an ellipsoid, so the process of choosing batches of points is an ellipsoid packing problem (Section \ref{sec:greedy_point_selection}).
 We use a greedy ellipsoid packing algorithm to choose as many points as possible per batch, while ensuring that there is no overlap between the support ellipsoids, $E_{x_i}$, associated with the sample points within a batch.
+%to choose batches of sample points, $\pointbatch_k$, such that there is no overlap between the ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch. 
+%These sample point batches will be used in Section \ref{sec:get_impulse_response} to compute many impulse responses of $\Aop$ using only a small number of operator actions of $\Aop$.
+The support of $\impulseresponse_x$ is (approximately) contained in the ellipsoid $E_x$, so by applying $\Aop$ to the Dirac comb associated with all sample points in a batch, we will recover the impulse responses for all sample points in that batch.
 
@@
 Then we select the candidate point $p \in P$ that is the farthest away from all points in previous sample point batches $S_1 \cup S_2 \cup \dots \cup S_{k-1}$.
+%That is, $p$ is a maximizer of the following optimization problem:
+%\begin{equation*}
+%\max_{p \in P} \min_{x \in \pointbatch_1 \cup \dots \cup \pointbatch_{k-1}} \|p - x\|.
+%\end{equation*}
 Candidate points for the first batch $S_1$ are chosen randomly from $P$.
@@
 	\item We check whether $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_k$.
-	\item We make sure that $V(p)$ is not too small, by checking whether $V(p) > \epsilon V_\text{max}$.
-Here $V_\text{max}$ is the largest value of $V(q)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\epsilon=10^{-5}$).
+	\item We make sure that $V(p)$ is not too small, by checking whether $V(p) > \tau V_\text{max}$.
+Here $V_\text{max}$ is the largest value of $V(q)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\tau=10^{-5}$).
 \end{enumerate}
@@
 
-	\Input{Finite set of candidate points $P\subset \Omega$, \\spatially varying mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\previous sample point batches $\pointbatch_1, \dots, \pointbatch_{k-1}$}
+	\Input{Finite set of candidate points $P\subset \Omega$, \\spatially varying mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\domain boundary information $\partial \Omega$, \\previous sample point batches $\pointbatch_1, \dots, \pointbatch_{k-1}$}
 	\Output{Batch of new sample points $\pointbatch_k$}
@@
 
-		\If{$E_p \cap E_q \neq \{\}$ for all $q \in \pointbatch_k$}{
+		\If{$E_p \cap E_q \neq \{\}$ for all $q \in \pointbatch_k$ and $E_p \cap \partial \Omega = \{\}$}{
 			\tcp{$E_p$ and $E_q$ are the ellipsoids defined in \eqref{eq:support_ellipsoid}}
@@
 The last equality in \eqref{eq:dirac_comb_H_action} follows from linearity and the definition of $\impulseresponse_{x_i}$ in \eqref{eq:impulse_response_delta_action}.
+%\begin{equation}
+%\label{eq:phi_b}
+%\combresponse_k = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
+%\end{equation}
+%\begin{equation}
+%	\label{eq:phi_b}
+%	\combresponse_k = \left(\Aop \left(\sum_{x_i \in \pointbatch_k} \delta_{x_i}/V(x_i)\right)^*\right)^* = \sum_{x_i \in \pointbatch_k} \left(\Aop \delta_{x_i}^*\right)^* / V(x_i) = \sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
+%\end{equation}
 Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\impulseresponse_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\impulseresponse_j$ do not overlap when $i \neq j$, we have (approximately)
@@
 
+%Note that the support of $\impulseresponse_{x_i}$ is estimated to be contained within the ellipsoid $E_{x_i}$, but this is just an approximation. 
+%%The impulse response $\impulseresponse_{x_i}$ may be nonzero inside the ellipsoid $E_{x_j}$ associated with another point $x_j$ in the batch, and this will cause error in our recovery of $\phi_{x_j}$.
+%Without this scaling, the neglected portion of an impulse response corresponding to a point $x_i$ could overwhelm the impulse response corresponding to a nearby point $x_j$ if $V(x_i)$ is much larger than $V(x_j)$.
+
 \subsection{Approximate integral kernel entries $\widetilde{\Phi}(y,x)$}
@@
 
+%Recall from Section \ref{eq:local_mean_displacement_invariance} that if $x'$ is near $x$, then
+%\begin{equation*}
+%\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x')\right).
+%\end{equation*}
+
+%In Section \ref{sec:intro} we noted that computing $\Aker\left(y,x\right)$ is computationally costly. 
+%This section describes how we compute approximate kernel entries
+%\begin{equation*}
+%\widetilde{\Phi}(y,x) \approx \Phi(y,x)
+%\end{equation*}
+%for arbitrary pairs of points $(y,x) \in \Omega \times \Omega$. 
 Given $(y,x)\in \Omega \times \Omega$, let
@@
 \end{equation}
+%for arbitrary pairs of points $(y,x) \subset \Omega \times \Omega$.
+% while Section \ref{sec:approximate_kernel_entries} we showed how we can cheaply evaluate approximate kernel entries at arbitrary pairs of points $(y,x) \in \Omega \times \Omega$.
 Note that $f_i$ is well-defined, because $\phi_{x_i}\left(z_i\right)$ is defined if and only if $z_i \in \Omega$, and $V(x_i)>0$ by our sample point picking procedure in Section \ref{sec:sample_point_selection}.
@@
 \end{equation}
-at the point $x$. The idea is that $\Phi(y,x) = f_i$ per the discussion in Section \ref{sec:local_mean_displacement_invariance}, so we approximate $\Phi(y,x)$ by interpolating the values $f_i$.
+at the point $x$. The idea is that $\Phi(y,x) = f_i$ per the discussion in Section \ref{eq:local_mean_displacement_invariance}, so we approximate $\Phi(y,x)$ by interpolating the values $f_i$.
 
@@
 We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
+%Since $\Phi(y,x) \approx f_i$ when $x_i$ is near $x$, we approximate $\Phi(y,x)$ by interpolating the (point,value) pairs
+%\begin{equation}
+%\left(x_i, f_i\right), \qquad i=1,\dots,k,
+%\end{equation}
+%at the point $x$.
+To see that $\Phi(y,x) \approx f_i$, recall from Section \ref{eq:local_mean_displacement_invariance} that $\Phi(y,x) = \phi_x(y) \approx \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$ if $x_i$ is near $x$.
+Inclusion of the extra factor $\frac{V(x)}{V(x_i)}$ in \eqref{eq:fxyxp} is a form of preconditioning that makes this approximation more robust in the case where $V(x)$ varies widely.
+%Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section \ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $V(x_i)$ is smaller than a predetermined threshold.
+%\subsubsection{Radial basis function interpolation}
 
 Interpolation is performed using the following radial basis function scheme:
+%The approximation $\widetilde{\Phi}(y,x)$ is the value of this interpolation at $x$.
+%In numerical experiments we use $k=10$. The $k$ nearest sample points are found by querying a precomputed KD-tree \cite{kdtree} of all sample points. 
+%
+%Let
+%It follows from inspection that $\Phi(y,x) = f_{y,x}(x)$,
+%and from the discussion in Section \ref{eq:local_mean_displacement_invariance}, we expect 
+%$
+%\Phi(y,x) \approx f_{y,x}(x')
+%$
+%when $x'$ is close to $x$.
+The factor $\frac{V(x)}{V(x_i)}$, which was not discussed in Section \ref{eq:local_mean_displacement_invariance}, is a form of preconditioning that improves the quality of this approximation in situations where $V(x)$ varies rapidly.
+%%, as is commonly the case for many operators of practical interest. In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.
+%%As discussed in Section SECTION, we make the approximation
+%%\begin{equation*}
+%%	\Phi(y,x) = \phi_x(y) \approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right)
+%%\end{equation*}
+%%if $x'$ is close to $x$. Hence, w
+%
+%We evaluate $\widetilde{\Phi}(y,x)$ by performing radial basis function interpolation of the function $f_{y,x}(x')$ in the variable $x'$.
+The data used in the interpolation are the following (point,value) pairs
+%\begin{equation}
+%\left(x_i, f_{y,x}(x_i)\right), \qquad i=1,\dots,k,
+%\end{equation}
+%and $(x, f_{y,x}(x)) = (x,\Phi(y,x))$ is the (point,value) pair we approximate via interpolation.
+%Here $\{x_i\}_{i=1}^k \subset \Omega$ are the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined.
+%%a collection of sample points at which the impulse response $\phi_{x_i}$ is known.
+%%. Specifically, we interpolate $(\text{point}, \text{value})$ pairs $(x_i, f_i)$ at the point $x$, where $x_i$ are sample points, and $f_i$ are given as follows:
+%%\begin{equation}
+%%\label{eq:phi_mu_shift32}
+%%	f_i := \phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
+%%\end{equation}
+%%We must perform a new interpolation every time we evaluate $\Phi(y,x)$ at a new pair of points $(y,x)$, and constructing the H-matrix approximation of $\Phi$ requires evaluating $\Phi(y,x)$ at a large number of pairs of points $(y,x)$.
+%%Interpolation using all sample points $x_i$ is too expensive, so we only use the $k$ nearest sample points to $x$, excluding sample points for which $f_{y,x}(x_i)$ is not defined. 
+%In numerical experiments we use $k=10$. The $k$ nearest sample points are found by querying a precomputed KD-tree \cite{kdtree} of all sample points. 
+%
+%From the definition of $f_{y,x}$ in \eqref{eq:fxyxp}, we see that $f_{y,x}(x_i)$ is only defined if $y - \mu(x) + \mu(x_i) \in \Omega$, because $\phi_x$ is only defined on $\Omega$.
+But it may happen that $y - \mu(x) + \mu(x_i) \notin \Omega$, even if $y,x,x_i \in \Omega$.
+In this case, the point $x_i$ is considered invalid, and not used in the interpolation, even if it is close to $x$.
+We check whether $x_i$ is valid (i.e., whether $y - \mu(x) + \mu(x_i) \in \Omega$) by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
+%\subsubsection{Radial basis function interpolation details}
+%Having found the $k$ nearest valid sample points $x_i$ to $x$, we approximate $\Phi(y,x)$ using radial basis function interpolation of the following form:
+%In detail, we form $\widetilde{\Phi}(y,x) \approx \Phi(y,x)$ as follows:
 \begin{equation}
@@
 $
-and $f \in \mathbb{R}^k$ has entries $f_i$ from \eqref{eq:fxyxp}. We use polyharmonic spline radial basis functions \cite{polyspline}:
+and $f \in \mathbb{R}^k$ has entries $f_i$ from \eqref{eq:fxyxp}.
+% has entries 
+%\begin{equation*}
+%f_i = f_{y,x}(x_i), \qquad i=1,\dots,k.
+%\end{equation*}
+We use polyharmonic spline radial basis functions \cite{polyspline}:
 \begin{equation}
@@
 For more details on radial basis function interpolation and polyharmonic splines, see \cite{rbf}.
-The interpolation procedure is illustrate in Figure \ref{fig:rbf_interpolation}.
-
-\begin{figure}
-	\begin{subfigure}{0.49\textwidth}
-		\centering
-		\includegraphics[scale=0.22]{interpolation_rbf.pdf}
-		\caption{}
-	\end{subfigure}
-	\begin{subfigure}{0.49\textwidth}
-		\centering
-		\includegraphics[scale=0.22]{interpolation_rbf_boundary.pdf}
-		\caption{}
-	\end{subfigure}
-	\caption{(a) To compute the kernel entry $\Phi(y,x)$, we use radial basis functions to interpolate the values $\frac{V(x)}{V(x_i)}\phi_{x_i}(y-\mu(x)-\mu(x_i))$ for the $k$ nearest sample points $x_i$ to $x$.
-(b) We exclude sample points for which $y-\mu(x)_\mu(x_i)$ is outside of the domain.}
-	\label{fig:rbf_interpolation}
-\end{figure}
-
-To evaluate $f_i$, we check whether $z_i \in E_{x_i}$ using \eqref{eq:support_ellipsoid}.
-If $z_i \notin E_{x_i}$, then $z_i$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_i=0$.
+%\begin{equation}
+%	f_i = \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right).
+%\end{equation}
+%The result of this interpolation is to approximate $\Phi(y,x)$ with a linear combination of the shifted impulse response values $\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$. 
+%By including the factor $\frac{V(x)}{V(x_i)}$, we are implicitly interpolating the ``volume preconditioned'' function $\frac{1}{V(x)}\Phi(y,x)$, then rescaling the result by $V(x)$ to get an approximation of $\Phi(y,x)$.
+Volume preconditioning improves the approximation considerably when $V(x)$ varies widely in magnitude, as is commonly the case for many operators of practical interest.
+In the Hessian for the Stokes inverse problem, for example, we observe that $V(x)$ varies over six orders of magnitude.
+
+%In order to compute $f_{y,x}(x_i)$, we must evaluate $\impulseresponse_{x_i}\left(y - \mu(x) + \mu(x_i)\right)$, and we do this by evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$.
+Here $b$ is the index of the batch that contains the sample point $x_i$, and $\eta_b$ is the corresponding impulse response batch function that we have computed and stored.
+%Although it is typically recommended that one avoid evaluating finite element functions at arbitrary points [CITE], here we have no choice.
+
+%\subsubsection{Evaluating $f_{y,x}(x_i)$}
+
+To evaluate $f_i$, we check whether $z_i \in E_{x_i}$ using \eqref{eq:support_ellipsoid}. If $z_i \notin E_{x_i}$, then
+% by checking if
+%\begin{equation}
+%\label{eq:ellipsoid_check93}
+%(y - \mu(x_i))^T \Sigma^{-1}(x_i) (y - \mu(x_i))  \le \tau^2.
+%\end{equation}
+%If \eqref{eq:ellipsoid_check93} does not hold, 
+then $z_i$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_i=0$.
+% If \eqref{eq:ellipsoid_check93} does hold, then 
 If $z_i \in E_{x_i}$, we look up the batch index $b$ such that $x_i \in S_b$, and evaluate $f_i$ via the formula
@@
 per \eqref{eq:varphi_eval}.
+%Recall from Section \ref{sec:get_impulse_response} that $\eta_b$ is the impulse response batch that we have computed and stored.
+%Evaluating $\eta_b\left(y - \mu(x) + \mu(x_i)\right)$ requires care, because
 
@@
 We regularize the problem because of numerical issues that arise if there are points $\widehat{x}_i$ and $\widehat{y}_j$ in the interpolation that are equal or nearly equal, a situation that arises often in practice.
+%Indeed, situations where $\widehat{x}_i = \widehat{y}_j$ are inevitable when a regular grid is used to discretize the problem. 
 In this case, $\widehat{B}$ is (nearly) singular, so it is ill-advised to solve the un-regularized interpolation system $\widehat{B} \widehat{w} = \widehat{f}$.
+
+%In principle, this equation is still solvable if $\left(\widetilde{f}_x\right)_i = \left(\widetilde{f}_x\right)_j$, which should hold in principle.
+However, errors incurred at any stage of the approximation process, particularly the assumption that the support of $\phi_x$ is contained in the ellipsoid $E_x$, will cause small discrepancies between $\left(\widetilde{f}_x\right)_i$ and $\left(\widetilde{f}_x\right)_j$, and these errors get amplified by the ill-conditioning.
+
+%	 or $\widetilde{x}_i \approx \widetilde{y}_j$. 
+%	
+%	while $\left(\widetilde{f}_x\right)_i \neq \left(\widetilde{f}_x\right)_j$
+%
+% $\widetilde{B}\widetilde{w}=\widetilde{f}$
+
+%and let $\widetilde{B}_{xx}$, $\widetilde{B}_{xy}$, $\widetilde{B}_{yx}$, and $\widetilde{B}_{yy}$ be $k \times k$ matrices with the following entries:
+%\begin{align*}
+%	\left(\widetilde{B}_{xx}\right)_{ij} &= b\left(\|\widetilde{x}_i - \widetilde{x}_j\|\right) \\	\left(\widetilde{B}_{xy}\right)_{ij} &= b\left(\|\widetilde{x}_i - \widetilde{y}_j\|\right) \\
+%\end{align*}
+%
+%
+%\begin{equation*}
+%	\begin{bmatrix}
+%		B_{xx} & B_{xy} \\ B_{yx} & B_{yy}
+%	\end{bmatrix}
+%\end{equation*}
+%
+%Let $\phi'_{y_i}(x) := \Phi(y_i,x)$ denote the impulse response of $\Aop^T$ to a point source at $y_i$. 
+%
+%Let $\{y_i\}_{i=1}^r$ be the $k$ nearest sample points to $y$, and let . In the case where $\Aop$ is symmetric, knowing 
+%
+%In the case where $\Aop$ is symmetric, we have
+%\begin{equation}
+%	\impulseresponse_{x_i}(y) = \Phi(y,x_i) = \Phi(x_i,y) = \impulseresponse'_{x_i}(y)
+%\end{equation}
+%knowing the impulse response $\impulseresponse_{x_i}$ of $\Aop$ at $x_i$
+
+%\subsection{Discretization and hierarchical matrix construction}
+%\label{sec:H_matrix_conversion}
+%
+%Here we discuss how we construct an H-matrix representation of our product-convolution approximation, after discretizing the problem with finite elements.
 
@@
 
-We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming H-matrix representations $\mathbf{\Phi}_H$ and $\bm{M}_H$ of $\mathbf{\Phi}$ and $\bm{M}$, respectively, then using fast H-matrix methods to multiply these matrices per \eqref{eq:boldA} to form
+We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming H-matrix representations $\mathbf{\Phi}_H$ and $\bm{M}_H$ of $\mathbf{\Phi}$ and $\bm{M}$, respectively, then using fast H-matrix methods to multiply these matrices per \eqref{eq:Amatpc_defn} to form
 \begin{equation*}
@@
 
-WORK IN PROGRESS BELOW HERE!!!
-
-\begin{tabular}{lll}
-	Variable name & Approximate size & Variable \\
-	\hline
-	$N$ & $10^3$--$10^9$ & Number of finite element degrees of freedom \\
-	$b$ & $5$--$25$ & number of impulse response batches \\
-	$r$ & $10$--$30$ & H-matrix rank \\
-	$k$ & $5$--$15$ & number of nearest neighbors \\
-	$\{\eta_b\}_{b=1}^\text{num\_batches}$ & $\text{num\_batches}$ actions of $\mathcal{A}$
-\end{tabular}
-
-$k_\text{hmatrix}$, $k_\text{nbrs}$, $k_\text{batches}$
-
-\begin{tabular}{ll}
-	$V$ & $1$ action of $\mathcal{A}^T$ \\
-	$\mu$ & $d$ actions of $\mathcal{A}^T$ \\
-	$\Sigma$ & $d(d+1)/2$ actions of $\mathcal{A}^T$ \\
-	$\{\eta_b\}_{b=1}^\text{num\_batches}$ & $\text{num\_batches}$ actions of $\mathcal{A}$
-\end{tabular}
-
 Computing $V$, $\mu$, and $\Sigma$ require $1$, $d$, and $d(d+1)/2$ operator actions of $\Aop$, respectively.
@@
 Since $k$ is small, this is cheap.
-
-\section{Content to add}
 
@@
 The first two steps require geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix, but these steps do not depend on the particular values of matrix entries.
-The third step requires us to evaluate $O(\fedim \log \fedim)$ specially-chosen entries of $\mathbf{\Phi}$, and we evaluate these entries using \eqref{eq:Akerpcmat_entries}.
+The third step requires us to evaluate $O(\fedim \log \fedim)$ specially-chosen entries of $\AkerPcMat$, and we evaluate these entries using \eqref{eq:Akerpcmat_entries}.
 
@@
 We already showed that $v_* \in H^2_N(\Omega)$ in Theorem \ref{thm:wi_no_breakdown}.
-Using the definition of $v_*$ in \eqref{eq:optimal_v} and Lemma \ref{lem:S_matrix_characterizations}, we have
+Using the the definition of $v_*$ in \eqref{eq:optimal_v} and Lemma \ref{lem:S_matrix_characterizations}, we have
 	\begin{equation*}

### 2022-04-18 paper/localpsf_old95.tex
about to remove old stuff from paper

--- 2022-04-18 paper/localpsf.tex
@@
 \end{equation*}
-These outlying eigenpairs may be computed using a Krylov method, or the randomized singular value decomposition.
-We then flip the sign of these ``bad'' negative eigenvalues to be positive as follows:
+These outlying ``bad'' eigenpairs may be computed using a Krylov method (we use Lanczos), or the randomized singular value decomposition.
+We then flip the sign of these negative eigenvalues to be positive as follows:
 \begin{equation*}
@@
 We seek to correct negative eigenvalues caused by errors in our approximation procedure.
-But even without our approximation, numerical rounding and discretization errors may cause large numbers of zero or near-zero eigenvalues to become slightly negative.
-Numerically, we cannot hope to correct these small magnitude negative eigenvalues because there are so many of them, and because the spectral gap between the smallest negative eigenvalue and the first non-negative eigenvalue is small.
+But even without our approximation, rounding and discretization errors may cause large numbers of zero or near-zero eigenvalues to become slightly negative.
+Numerically, we cannot hope to correct these slightly negative eigenvalues because there are so many of them, and because the spectral gap between the smallest negative eigenvalue and the first non-negative eigenvalue is small.
 

### 2022-04-18 paper/localpsf.tex
deleted old stuff from paper

@@
 
-\subsubsection{Improvement for symmetric operators}
-\label{sec:symmetric_improvements}
-If $\Aop$ is symmetric, impulse responses of $\Aop$ are also impulse responses of $\Aop^T$.
-Let $\{y_i\}_{i=1}^k$ be the $k$ nearest sample points to $y$, excluding points $y_i$ such that $x - \mu(y) + \mu(y_i) \notin \Omega$.
-We could form an alternative approximation of $\Phi(y,x)$ by performing the same interpolation as before, but with $x$ and $y$ taking opposite roles.
-Both of these interpolation schemes can be combined to improve the approximation accuracy at negligible additional cost.
-We combine these interpolation schemes by placing function values
-\begin{equation*}
-\left(\widehat{f}_x\right)_i := \frac{V(x)}{V(x_i)}\phi_{x_i}\left(y - \mu(x) + \mu(x_i)\right), \qquad i=1,\dots,k,
-\end{equation*}
-at the points $x_i-x$, placing function values
-\begin{equation*}
-\left(\widehat{f}_y\right)_i := \frac{V(y)}{V(y_i)}\phi_{y_i}\left(x - \mu(y) + \mu(y_i)\right), \qquad i=1,\dots,k,
-\end{equation*}
-at the points $y_i-y$, and interpolating the combination of these $2k$ points and values at zero.
-
-In detail, define 
-\begin{equation*}
-\widehat{x_i} := x_i - x \qquad \text{and} \qquad \widehat{y_i} := y_i - y.
-\end{equation*}
-and let
-\begin{equation*}
-\widehat{B} = \begin{bmatrix}
-\widehat{B}_{xx} & \widehat{B}_{xy} \\
-\widehat{B}_{yx} & \widehat{B}_{xx}
-\end{bmatrix}, \qquad \widehat{w} = \begin{bmatrix}\widehat{w}_x \\ \widehat{w}_y\end{bmatrix}, \qquad \widehat{f} = \begin{bmatrix}
-\widehat{f}_x \\ \widehat{f}_y
-\end{bmatrix},
-\end{equation*}
-be a $2k \times 2k$ block matrix and $2k \times 1$ block vectors, respectively, with the block entries of $\widehat{B}$ given by
-\begin{align*}
-\left(\widehat{B}_{xx}\right)_{ij} = b\left(\|\widehat{x}_i - \widehat{x}_j\|\right), &\qquad
-\left(\widehat{B}_{xy}\right)_{ij} = b\left(\|\widehat{x}_i - \widehat{y}_j\|\right), \\
-\left(\widehat{B}_{yx}\right)_{ij} = b\left(\|\widehat{y}_i - \widehat{x}_j\|\right), &\qquad \left(\widehat{B}_{yy}\right)_{ij} = b\left(\|\widehat{y}_i - \widehat{y}_j\|\right),
-\end{align*}
-for $i,j=1,\dots,k$. We approximate $\Phi(y,x)$ as follows:
-\begin{equation}
-\widetilde{\Phi}(y,x) := \sum_{i=1}^k \left(w_x\right)_i~ b\left(\|\widehat{x}_i\|\right) + \sum_{i=1}^k \left(w_y\right)_i~ b\left(\|\widehat{y}_i\|\right),
-\end{equation}
-where $\widehat{w}$ is the solution of the following regularized least-squares problem,
-\begin{equation*}
-\min_{\widehat{w}} \frac{1}{2}\|\widehat{B} \widehat{w} - \widehat{f}\|^2 + \frac{\alpha}{2}\|\widehat{w}\|^2,
-\end{equation*}
-for a small regularization parameter $\alpha$. 
-We solve the regularized least squares problem by QR factorization of the block $4k \times 2k$ matrix $\begin{bmatrix}\widehat{B} & I\end{bmatrix}^T$.
-We regularize the problem because of numerical issues that arise if there are points $\widehat{x}_i$ and $\widehat{y}_j$ in the interpolation that are equal or nearly equal, a situation that arises often in practice.
-In this case, $\widehat{B}$ is (nearly) singular, so it is ill-advised to solve the un-regularized interpolation system $\widehat{B} \widehat{w} = \widehat{f}$.
-
 \subsection{Hierarchical matrix construction}
@@
 
-\subsection{Rational positive semi-definite modification}
+\section{Rational positive semi-definite modification}
 \label{sec:make_spd}
@@
 
-\section{old rational}
-
-In this section we describe a method for modifying $\mathbf{A}_H$ to make it symmetric positive semi-definite.
-
-Because of the modified interpolation for symmetric matrices described in Section \ref{sec:symmetric_improvements}, $\mathbf{A}_H$ is nearly (but not completely) symmetric.
-We further symmetrize $\mathbf{A}_H$ as follows:
-\begin{equation*}
-\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
-\end{equation*}
-This step is fast because adding H-matrices is a comparatively inexpensive operation.
-
-We modify $\mathbf{A}_H^{\text{sym}}$ to make it positive semi-definite by forming a rational matrix function of the following form: 
-\begin{equation}
-\label{eq:rational_matrix_function}
-\mathbf{A}_H^{\text{sym}+} := \ratcoeff_0 \bm{I} + \ratcoeff_1 \mathbf{A}_H^{\text{sym}} + \ratcoeff_2 \left(\mathbf{A}_H^{\text{sym}} + \ratpole \bm{I}\right)^{-1},
-\end{equation}
-where $\bm{I}$ is the identity matrix which has the same shape as $\mathbf{A}_H^{\text{sym}}$.
-Approximation \eqref{eq:rational_matrix_function} may be written as $\mathbf{A}_H^{\text{sym}+} = \ratfct(\mathbf{A}_H^{\text{sym}})$, where $\ratfct$ is the following scalar rational function:
-\begin{equation}
-\label{eq:rational_function_scalar}
-\ratfct(\lambda) := \ratcoeff_0 + \ratcoeff_1 \lambda + \frac{\ratcoeff_2}{\lambda+\ratpole},
-\end{equation}
-which is extended to matrices by functional calculus.
-The scalars $\ratcoeff_0$, $\ratcoeff_1$, $\ratcoeff_2$, and $\ratpole$ are chosen so that the moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym}+}$ approximately equal the corresponding moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$, while the negative and small positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$ are modified so that $\mathbf{A}_H^{\text{sym}+}$ is positive semi-definite.
-Later in this section we will explain how $\ratcoeff_0$, $\ratcoeff_1$, $\ratcoeff_2$, and $\ratpole$ are chosen and how our choice of these scalars affects the spectral properties of $\mathbf{A}_H^{\text{sym}+}$.
-In Figure FIG we illustrate $\ratfct(\lambda)$.
-Forming $\mathbf{A}_H^{\text{sym}+}$ requires computing an H-matrix inverse, $\left(\mathbf{A}_H^{\text{sym}} + \ratpole \bm{I}\right)^{-1}$.
-While explicit computation of matrix inverses should usually be avoided, here it is computationally acceptable because $\mathbf{A}_H^{\text{sym}} + \ratpole \bm{I}$ is an H-matrix.
-Inversion of H-matrices is a routine procedure which only requires $O(\fedim \log(\fedim)^2)$ work \cite{Hmatrixinverse}.
-
-We choose $\ratcoeff_0, \ratcoeff_1, \ratcoeff_2$, and $\ratpole$, to be the solution to the following system of equations:
-\begin{subequations}
-	\label{eq:four_eq_system}
-	\begin{align}
-	\ratfct(0) &= 0 \label{eq:desired_property_1} \\
-	\ratfct'(0) &= 0 \label{eq:desired_property_2} \\
-	\ratfct(\lambda_\text{min}) &= |\lambda_\text{min}| \label{eq:desired_property_3} \\
-	\ratfct(\lambda_\text{max}) &= \lambda_\text{max}, \label{eq:desired_property_4}
-	\end{align}
-\end{subequations}
-where $\lambda_\text{min}$ and $\lambda_\text{max}$ are the smallest and largest eigenvalues of $\mathbf{A}_H^{\text{sym}}$, respectively.
-We estimate $\lambda_\text{min}$ and $\lambda_\text{max}$ using the implicitly restarted Lanczos method \cite{lanczos,scipy}.
-We assume that $\lambda_\text{min} < 0$; if $\lambda_\text{min} \ge 0$, then $\mathbf{A}_H^{\text{sym}}$ is already positive semi-definite, so we do not need to modify it with a rational function.
-We also require $\ratpole > |\lambda_\text{min}|$, so that the pole of $\ratfct$ is not in $[\lambda_\text{min},\lambda_\text{max}]$.
-We describe how we solve system \eqref{eq:four_eq_system} in Appendix \ref{app:solve_rational_system}.
-Solving \eqref{eq:four_eq_system} reduces to solving a smooth one-dimensional optimization problem, which is computationally trivial.
-
-Together, equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, \eqref{eq:desired_property_3},  and \eqref{eq:desired_property_4}, and the condition $\ratpole > |\lambda_\text{min}|$, imply that $\ratfct(0)=0$ is the unique minimum of $\ratfct$ on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that $\ratfct$ is non-negative on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that  $\AmatPcSymPlus$ is positive semi-definite.
-Furthermore, \eqref{eq:desired_property_1} implies that the null space of $\mathbf{A}_H^{\text{sym+}}$ contains the null-space of $\mathbf{A}_H^{\text{sym}}$.
-This is important when the operator $\Aop$ has a large (or infinite) dimensional null space, or a spectrum that clusters at zero.
-Hessians in ill-posed distributed parameter inverse problems typically have both of these properties.
-Equation \eqref{eq:desired_property_4} forces the moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym+}}$ to be close to the corresponding moderate and large positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
-I.e., $\ratfct$ modifies the ``important'' part of the spectrum of $\mathbf{A}_H^{\text{sym}}$ as little as possible.
-Equation \eqref{eq:desired_property_3} ensures that the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $\ratfct$.
-There is a tradeoff: if we choose a larger value for $\ratfct(\lambda_\text{min})$, then the positive eigenvalues of $\mathbf{A}_H^{\text{sym+}}$ will be closer to the corresponding positive eigenvalues of $\mathbf{A}_H^{\text{sym}}$ (good), but the erroneous negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ will be amplified in magnitude (bad).
-This is shown in Figure FIG.
-We find that $\ratfct(\lambda_\text{min})=|\lambda_\text{min}|$ is an appropriate balance between these competing interests.
-
-\subsection{Solving for rational function scalars}
-\label{app:solve_rational_system}
-
-Using the definition of $\ratfct$ in \eqref{eq:rational_function_scalar}, and basic calculus and linear algebra, we may write Equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, and \eqref{eq:desired_property_4} in the following matrix form:
-\begin{equation}
-	\label{eq:rational_3_linear_system}
-	\begin{bmatrix} 
-		1 & 0 & 1/\ratpole \\ 0 & 1 & -1/\ratpole^2 \\ 1 & \lambda_\text{max} & 1/(\lambda_\text{max}+\ratpole) 
-	\end{bmatrix}
-	\begin{bmatrix} 
-		\ratcoeff_0 \\ \ratcoeff_1 \\ \ratcoeff_2 
-	\end{bmatrix}
-	=
-	\begin{bmatrix} 
-		0 \\ 0 \\ \lambda_\text{max} 
-	\end{bmatrix},
-\end{equation}
-and we may write Equation \eqref{eq:desired_property_3} as follows: 
-\begin{equation}
-	\label{eq:rational_extra_eq}
-	\ratcoeff_0 + \ratcoeff_1 \lambda_\text{min} + \frac{\ratcoeff_2}{\lambda_\text{min}+\ratpole} - |\lambda_\text{min}| = 0.
-\end{equation}
-Let 
-\begin{equation*}
-	f(\ratpole) := \ratcoeff_0(\ratpole) + \ratcoeff_1(\ratpole) \lambda_\text{min} + \frac{\ratcoeff_2(\ratpole)}{\lambda_\text{min}+\ratpole} - |\lambda_\text{min}|
-\end{equation*}
-denote the left hand side of \eqref{eq:rational_extra_eq}, but with $\ratcoeff_0(\ratpole)$, $\ratcoeff_1(\ratpole)$, and $\ratcoeff_2(\ratpole)$ being the solution of \eqref{eq:rational_3_linear_system} for the given value of $\ratpole$.
-I.e., within $f(\ratpole)$ the quantities $\ratcoeff_0$, $\ratcoeff_1$, and $\ratcoeff_2$ are no longer free variables; they now depend on $\ratpole$ implicitly through the solution of a linear system.
-We solve the combined system, \eqref{eq:rational_3_linear_system} and $\eqref{eq:rational_extra_eq}$, by solving the one-dimensional nonlinear root finding problem
-\begin{equation}
-	\label{eq:g_mu_zero}
-	f(\ratpole)=0
-\end{equation}
-with bound $\ratpole > |\lambda_\text{min}|$ and initial guess $\ratpole_0 = 2 |\lambda_\text{min}|$.
-Each evaluation of $g$ requires solving the $3 \times 3$ linear system \eqref{eq:rational_3_linear_system}, and once the solution $\ratpole$ is found, we solve \eqref{eq:rational_3_linear_system} one more time to get the final values of $\ratcoeff_0$, $\ratcoeff_1$, and $\ratcoeff_2$.
-
-\section{Proofs}
-\label{app:proofs}
-
-\begin{defn}[Moments of $\impulseresponse_x$]
-	\label{defn:alpha_rho_mu_sigma}
-	We define the spatially varying scaling factor, $\spatialvol$, the scaled impulse response, $\widehat{\impulseresponse}_x$, the spatially varying mean, $\spatialmean$, and the spatially varying covariance, $\spatialcov$, as follows:
-	\begin{align*}
-		\spatialvol:\Omega \rightarrow \mathbb{R}, &\qquad \spatialvol(x) := \int_{\Omega} \impulseresponse_x(z) dz, \\
-		\widehat{\impulseresponse}_x:\Omega \rightarrow \mathbb{R}, &\qquad \widehat{\impulseresponse}_x := \impulseresponse_x \big/ \spatialvol(x), \\
-		\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim, &\qquad \spatialmean(x) := \int_\Omega z \widehat{\impulseresponse}_x(z) dz, \\
-		\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}, &\qquad\spatialcov(x) := \int_\Omega (z - \spatialmean(x))(z - \spatialmean(x))^T \widehat{\impulseresponse}_x(z) dz.
-	\end{align*}
-\end{defn}
-
-\begin{proof}[Proof of Proposition \ref{thm:vol_mean_cov}]
-	Let $v \in L^2$ be arbitrary. We have
-	\begin{align*}
-		\left(v, \spatialvol\right)_{L^2(\Omega)} =& \int_\Omega v(x) \int_{\Omega} \left(\Aop \delta_x^*\right)^*(z) dz dx\\
-		=& \int_\Omega v(x) \int_{\Omega} A(z,x) dz dx \\
-		=& \int_\Omega \int_\Omega v(x) A(z,x) C(z) dz dx \\
-		=& \left(\Aop^T C\right)(v),
-	\end{align*}
-	which implies $\spatialvol = \left(\Aop^T C\right)^*$.
-
-	Using similar techniques, we have
-	\begin{align*}
-		\left(v, \spatialvol \cdot \spatialmean^i\right)_{L^2(\Omega)} =& \int_\Omega v(x) \spatialvol(x) \int_{\Omega} z^i \left(\Aop \delta_x^*\right)^*(z)/\spatialvol(x) dz dx\\
-		=& \int_\Omega \int_\Omega v(x) L^i(z) A(z,x) dz dx \\
-		=& \left(\Aop^T L^i\right)(v),
-	\end{align*}
-	which implies $\spatialmean^i = \left(\Aop^T L^i\right)^* / \spatialvol$.
-
-	Using the identity
-	\begin{equation*}
-		\spatialcov(x) = \int_\Omega zz^T \widehat{\impulseresponse}_x(z) dz - \spatialmean(x) \spatialmean(x)^T,
-	\end{equation*}
-	we have
-	\begin{align*}
-		\left(v, \spatialvol \cdot \left(\spatialcov^{ij}+\spatialmean^i \cdot \spatialmean^j\right)\right)_{L^2(\Omega)} =& \int_\Omega v(x) \spatialvol(x) \int_\Omega z^i z^j \left(\Aop \delta_x^*\right)^*(z) / \spatialvol(x) dz dx \\
-		=& \int_\Omega \int_\Omega v(x) Q^{ij}(z) A(z,x) dz dx \\
-		=& \left(\Aop Q^{ij}\right)(v),
-	\end{align*}
-	which implies $\spatialcov^{ij} = \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot\spatialmean^j$.
-\end{proof}
-
-\section{Background: distributions}
-\label{app:distributions}
-
-%Let $\overline{\Omega}$ be the closure of $\Omega$, and let $C(\overline{\Omega})$ be the space of continuous functions mapping $\overline{\Omega}\rightarrow \mathbb{R}$. 
-The action of $\Aop$ is extended to distributions $\rho$
-%$\genericdistribution:C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ 
-via the formula
-\begin{equation*}
-	\left(\Aop \genericdistribution^*\right)(v) := \int_\Omega v(y) \genericdistribution\left(\Phi(y, \cdot)\right) dy, 
-\end{equation*}
-where $\Phi(y,\cdot)$ is the function $x \mapsto \Phi(y,x)$. This is derived formally as follows:
-\begin{align*}
-	\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Phi(y,x) \textrm{``} \genericdistribution(x) \textrm{''} dx dy \\
-	&= \int_\Omega v(y) \int_\Omega \Phi(y,x) \textrm{``} \genericdistribution(x) \textrm{''} dx \\
-	&= \int_\Omega v(y) \genericdistribution\left(\Phi(y,\cdot)\right) dx.
-\end{align*}
-For example, the delta distribution $\delta_x$ is defined by $\delta_x(v) = v(x)$, and the action of $\Aop$ on $\delta_x$ is given by
-\begin{equation}
-	\label{eq:appendix_impulse_response}
-	\left(\Aop \delta_x^*\right)(v)	= \int_\Omega v(y) \Phi(y,x) dy.
-\end{equation}
-Recall that the impulse response of $\Aop$ to $\delta_x$, denoted $\impulseresponse_x$, is the Riesz representation of $\Aop \delta_x^*$. 
-%From \eqref{eq:appendix_impulse_response} and the definition of the Riesz representation, we have $\impulseresponse_x = A(~\cdot~,x)$.
-That is, $\impulseresponse_x$ is the function $y \mapsto A(y,x)$.
-
-\section{Discretization}
-\label{sec:discretization}
-
-Let $\febasis_i$, $i=1,\dots,\fedim$ be a set of finite element basis functions, let $h$ denote the mesh size parameter for the finite element mesh, and let
-\begin{equation*}
-	V_h = \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)
-\end{equation*}
-be the corresponding  finite element space with the $L^2$ inner product.
-Functions $u\in L^2(\Omega)$ are approximated by functions $u_h \in V_h$.
-In turn, functions $u_h\in V_h$ are represented in computations by length $\fedim$ arrays $\bm{u}$, such that the array entries of $\bm{u}$ are the coordinates of $u_h$ in the finite element basis.
-That is,
-\begin{equation}
-	\label{eq:finite_bijection}
-	u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
-\end{equation}
-
-We write $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ to denote the finite element \emph{mass matrix}, which has entries
-\begin{equation*}
-	\bm{M}_{ij} = \left(\febasis_i, \febasis_j\right)_{L^2(\Omega)} = \int_{\Omega} \febasis_i(x) \febasis_j(x) dx,
-\end{equation*}
-and we write $\mathbb{R}^\fedim_{\bm{M}}$ to denote the space $\mathbb{R}^\fedim$ with the matrix-weighted inner product
-\begin{equation*}
-	\left(\bm{u}, \bm{v}\right)_{\bm{M}} := \bm{v}^T \bm{M} \bm{u}.
-\end{equation*}
-Direct calculation shows that $V_h$ is isometrically isomorphic $\mathbb{R}^\fedim_{\bm{M}}$.
-That is, equation \eqref{eq:finite_bijection} establishes a bijection between functions $u_h \in V_h$ and coefficient vectors $\bm{u} \in \mathbb{R}^\fedim_{\bm{M}}$, and this bijection preserves the inner product in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\bm{u},\bm{v}\right)_{\bm{M}}$.
-Likewise, the dual space, $V_h'$, is isometrically isomorphic to $\mathbb{R}^\fedim_{\bm{M}^{-1}}$.
-
-Here vectors in $\bm{u} \in \mathbb{R}^\fedim_{\bm{M}}$ and $\bm{\sigma} \in \mathbb{R}^\fedim_{\bm{M}^{-1}}$ are viewed as column vectors, and $\bm{u}^T$ and $\bm{\sigma}^T$ are the row vectors that result from taking the standard matrix transposes of $\bm{u}$ and $\bm{\sigma}$, respectively.
-The discrete dual of a vector $\bm{v}$ with respect to the $\bm{M}$ inner product is written as $\bm{v}^*$, and $\bm{v}^*$ is viewed as a column vector.
-
-\subsection{Discretized operations}
-\label{app:discretized_operations}
-
-Here we list the function space operations on $V_h$ and $V_h'$ that are required for this paper, and show how to perform the corresponding matrix and vector operations on $\mathbb{R}^\fedim_{\bm{M}}$ and $\mathbb{R}^\fedim_{\bm{M}^{-1}}$.
-
-\begin{description}
-	\item[Coefficients of a functional:] The coefficient dual vector, $\bm{\sigma} \in \mathbb{R}^\fedim_{\bm{M}^{-1}}$, corresponding to a linear functional $\sigma_h \in V_h'$ has components given by
-	\begin{equation*}
-		\bm{\sigma}_i = \sigma_h(\febasis_i), \quad i=1,\dots,\fedim.
-	\end{equation*}
-	\item[Action of a functional:] The action of a linear functional $\sigma_h \in V_h'$ on a function $u_h \in V_h$ may be computed as follows:
-	\begin{equation*}
-		\sigma_h(v_h) = \bm{\sigma}^T \bm{u}.
-	\end{equation*}
-	\item[Riesz representation:] Let $\sigma_h \in V_h'$. The Riesz representation of $\sigma_h$ is the unique  function $\sigma_h^* \in V_h$ satisfying 
-	\begin{equation*} 
-		\left(\sigma_h^*, v_h\right)_{L^2(\Omega)} = \sigma_h(v_h) \quad\forall v_h \in V_h.
-	\end{equation*}
-	The coefficient dual vector $\bm{\sigma}^*$ corresponding to $\sigma_h$ is given by
-	\begin{equation*} 
-		\bm{\sigma}^* = \bm{M}^{-1} \bm{\sigma}.
-	\end{equation*}
-	\item[Finite element $L^2$ projection:] Let $g \in L^2(\Omega)$.
-The orthogonal projection of $g$ onto $V_h$ with respect to the $L^2$ inner product is the unique function $g^\pi_h \in V_h$ satisfying
-	\begin{equation*}
-		\left(g^\pi_h, v_h\right)_{L^2(\Omega)} = \left(g, v_h\right)_{L^2(\Omega)} \quad\forall v_h \in V_h.
-	\end{equation*}
-	 Let $\left(\bm{g}^\pi\right)^* \in \mathbb{R}^\fedim_{\bm{M}^{-1}}$ be the vector with components
-	\begin{equation*}
-		\left(\bm{g}^\pi\right)^*_i = \int_\Omega g(x) \febasis_i(x) dx, \quad i=1,\dots,\fedim.
-	\end{equation*}
-	The coefficient vector $\bm{g}^\pi \in \mathbb{R}^\fedim_{\bm{M}}$ corresponding to $g^\pi_h$ is given by
-	\begin{equation*}
-		\bm{g}^\pi = \bm{M}^{-1}\left(\bm{g}^\pi\right)^*.
-	\end{equation*}
-	\item[Finite element interpolation:] Let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots,\fedim$, be the Lagrange nodes associated with the finite element basis functions $\febasis_i$.
-The finite element interpolant of a function $g \in C\left(\overline{\Omega}\right)$ is the function $g^I \in V_h$ given by
-	\begin{equation*}
-		g_h^I(x) = \sum_{i=1}^\fedim g(p_i) \febasis_i(x).
-	\end{equation*}
-	The coefficient vector, $\bm{g}$, for $g_h^I$ has components 
-	\begin{equation*}
-	\bm{g}_i^I = g(p_i), \quad i=1.,\dots,\fedim.
-	\end{equation*}
-	\item[Projection vs. interpolation:]
-	Typically, $g^\pi$ is a more accurate approximation of $g$ than $g^I$, but $g^I$ is cheaper to compute than $g^\pi$ because computing $g^\pi$ requires solving a linear system, while computing $g^I$ does not.
-If the finite element basis satisfies the following Kronecker property,
-	\begin{equation*}
-		\febasis_i(p_j) = \begin{cases}1 & i=j \\
-		0 & \text{otherwise},\end{cases}
-	\end{equation*}
-	then $g^I(p_i) = g(p_i)$ for $i=1,\dots,\fedim$ (i.e., the interpolant equals the original function at the Lagrange nodes).
-Interpolation is typically sufficiently accurate for practical purposes if the Kronecker property is satisfied.
-Projection should be used if the Kronecker property is not satisfied.
-	\item[Matrix representation of an operator:] Let $\mathcal{B}_h : V_h \rightarrow V_h'$ be a linear operator, and let $\bm{B} \in \bm{R}^{\fedim \times \fedim}$ be the matrix with entries
-	\begin{equation*}
-		\bm{B}_{ij} = \left(\mathcal{B}_h \febasis_j\right)(\febasis_i).
-	\end{equation*}
-	The matrix $\bm{B}$ maps $\mathbb{R}^\fedim_{\bm{M}} \rightarrow \mathbb{R}^\fedim_{\bm{M}^{-1}}$ by matrix multiplication, and
-	\begin{equation*}
-		\sigma_h = \mathcal{B}_h u_h \quad \Leftrightarrow \quad \bm{\sigma} = \bm{B} \bm{u}.
-	\end{equation*}
-	\item[Transpose:] The matrix $\bm{B}^T$ maps $\mathbb{R}^\fedim_{\bm{M}} \rightarrow \mathbb{R}^\fedim_{\bm{M}^{-1}}$ via matrix multiplication, and
-	\begin{equation*}
-		\sigma_h = \mathcal{B}_h^T u_h \quad \Leftrightarrow \quad \bm{\sigma} = \bm{B}^T \bm{u}.
-	\end{equation*}
-
-	\item[Distributions:] Let $\genericdistribution:C\left(\overline{\Omega}\right)\rightarrow \mathbb{R}$ be a distribution.
-If $\genericdistribution(\febasis_i)$ is well-defined for $i=1,\dots,\fedim$, then the restriction of $\genericdistribution$ to domain $V_h$ is a linear functional $\genericdistribution_h\in V_h'$.
-The functional $\genericdistribution_h$ has a coefficient dual vector $\bm{\genericdistribution} \in \mathbb{R}^\fedim_{\bm{M}^{-1}}$ with entries
-	\begin{equation*}
-		\bm{\genericdistribution}_i=\genericdistribution(\febasis_i), \quad i=1,\dots,\fedim.
-	\end{equation*} 
-	For example, let $\delta_p$ be the Dirac distribution centered at a point $p\in\Omega$.
-If the finite element basis functions, $\febasis_i$, are continuous at $p$, then restricting the domain of $\delta_p$ to the space $V_h$ yields a linear functional in $V_h'$.
-The corresponding coefficient dual vector, $\bm{\delta}_p$, has entries
-	\begin{equation*}
-		\left(\bm{\delta}_p\right)_i = \febasis_i(p), \quad i=1,\dots,\fedim.
-	\end{equation*}
-
-	\item[Action of an operator on a distribution:] Since $V_h$ is finite-dimensional, the restriction, $\genericdistribution_h \in V_h$, of a distribution, $\genericdistribution$, to the space $V_h$ has a a Riesz representation $\genericdistribution_h^* \in V_h$.
-Let $B_h$ be the integral kernel associated with an operator $\mathcal{B}_h:V_h \rightarrow V_h'$.
-For $v_h \in V_h$, we have
-	\begin{align*}
-		\langle \mathcal{B}_h, \genericdistribution_h \rangle(v) &= \int_\Omega v_h(y) \genericdistribution_h\left(B_h(y, \cdot)\right) dy \\
-		&= \int_\Omega v_h(y) \left(\int_\Omega B_h(y,x) \genericdistribution_h^*(x) dx\right) dy \\
-		&= \int_\Omega \int_\Omega v_h(y) B_h(y,x) \genericdistribution_h^*(x) dx dy = \left(\mathcal{B}_h \genericdistribution_h^*\right)(v).
-	\end{align*}
-	Going from the first to the second line we used the definition of the Riesz representation.
-The action of $\mathcal{B}_h$ on $\genericdistribution_h$ in the sense of distributions is therefore equal to the action of $\mathcal{B}_h$ on $\genericdistribution^*_h$ in the conventional sense.
-We have
-	\begin{equation*}
-		\sigma_h = \langle \mathcal{B}_h, \genericdistribution_h \rangle \quad \Leftrightarrow \quad \sigma_h = \mathcal{B}_h \genericdistribution_h^* \quad \Leftrightarrow \quad \bm{\sigma} = \bm{B} \bm{M}^{-1} \bm{\genericdistribution}.
-	\end{equation*}
-\end{description}
-
-\subsection{Discretized mean and covariance estimation}
-
-\begin{algorithm2e}
-	\SetAlgoNoLine
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Output}
-	{	
-		\Input{Operator $\bm{A}$}
-		\Output{$\bm{\spatialvol}\in \mathbb{R}^\fedim$, $\bm{\spatialmean}\in \mathbb{R}^{\gdim \times \fedim}$, and $\bm{\spatialcov} \in \mathbb{R}^{\gdim \times \gdim \times \fedim}$}
-
-		Form vector $\bm{C} \in \mathbb{R}^\fedim_{\bm{M}}$ by either projecting or interpolating constant function $C(x)=1$ onto $V_h$
-
-		$\bm{\spatialvol} = \left(\bm{M}^{-1}\bm{A}^T \bm{C}\right)^*$
-
-		\For{$i=1,2,\dots,\gdim$}{
-			Form vector $\bm{L}^i \in \mathbb{R}^\fedim_{\bm{M}}$ by either projecting or interpolating linear function $L^i(x) = x^i$ onto $V_h$
-
-			$\bm{\spatialmean}^i = \left(\bm{M}^{-1}\bm{A}^T \bm{L}^i\right)^* / \bm{\spatialvol}$
-		}
-		\For{$i=1,2,\dots,\gdim$}{
-			\For{$j=1,\dots,i$}{
-				Form quadratic function $\bm{Q}^{ij} = x^i x^j$
-
-				$\bm{\spatialcov}^{ij} = \left(\bm{M}^{-1}\bm{A}^T \bm{Q}^{ij}\right)^* / \bm{\spatialvol} - \bm{\spatialmean}^i\cdot \bm{\spatialmean}^j$
-
-				$\bm{\spatialcov}^{ij} = \bm{\spatialcov}^{ji}$
-
-			}
-		}
-
-	}
-	\caption{Discretized version of Algorithm \ref{alg:varhpi_mean_cov}}
-	\label{alg:varhpi_mean_cov_discrete}
-\end{algorithm2e}
-
-In computations, $\spatialvol$, $\spatialmean^{i}$, and $\spatialcov^{ij}$ are replaced with finite element approximations $\spatialvol_h, \spatialmean_h^i, \spatialcov_h^{ij}\in V_h$, respectively, which have coefficient vectors $\bm{\spatialvol}, \bm{\spatialmean}^{ij}, \bm{\spatialcov}^{ij}\in\mathbb{R}^\fedim_{\bm{M}}$, respectively.
-The functions $C$, $L^i$, and $Q^{ij}$ are replaced with either finite element projections onto $V_h$ (if the finite element basis does not satisfy the Kronecker property), or finite element interpolations onto $V_h$ (if the finite element basis satisfies the Kronecker property).
-In our numerical examples we use finite element basis functions that satisfy the Kronecker property, and we use finite element interpolations.
-The coefficient vectors corresponding to either the interpolations or projections are denoted $\bm{C}$, $\bm{L}^i$, and $\bm{Q}^{ij}$, respectively.
-The discretized versions of \eqref{eq:vol_mean_var_thm1}, \eqref{eq:vol_mean_var_thm2}, and \eqref{eq:vol_mean_var_thm3} are
-\begin{align*}
-	\bm{\spatialvol} &= \bm{M}^{-1}\bm{A}^T\bm{C} \\
-	\bm{\spatialmean}^i &= \left(\bm{M}^{-1}\bm{A}^T \bm{L}^i\right) / \bm{\spatialvol}\\
-	\bm{\spatialcov}^{ij} &= \left(\bm{M}^{-1}\bm{A}^T \bm{Q}^{ij}\right) / \bm{\spatialvol} - \bm{\spatialmean}^i\cdot \bm{\spatialmean}^j,
-\end{align*}
-respectively. Here $\bm{f} \cdot \bm{g}$ and $\bm{f} / \bm{g}$ denote the component-wise multiplication and division of vectors, respectively.
-
-\subsection{Discretized sampling of impulse responses}
-In computations, the Dirac comb coefficient dual vector for $\diraccomb_h^k$ is the vector $\bm{\diraccomb}^k \in \mathbb{R}^\fedim_{\bm{M}^{-1}}$ which has entries
-\begin{equation*}
-	\bm{\diraccomb}^k_i = \sum_{x_j \in \pointbatch_k} \febasis_i(x_j), \quad i=1,\dots,\fedim.
-\end{equation*}
-The Dirac comb impulse response coefficient vector for $\combresponse^h_k$ is the vector $\bm{\combresponse}_k \in \mathbb{R}^\fedim_{\bm{M}}$ given by
-\begin{equation*}
-	\bm{\combresponse}_k = \bm{M}^{-1} \bm{A}^T \bm{M}^{-1}\bm{\diraccomb}^k.
-\end{equation*}
-
-\section{OLD CONTENT}
-
-\subsection{disorganized march 21 22}
-
-H-matrix approximations are highly desirable because
-
-The computational cost of these operations is $O(N \log^a N)$, where the matrix is $N\times N$, and $a \in \{1,2,3\}$. 
-Classical H-matrix construction methods 
-require evaluating the integral kernel $O(N \log N)$ times, but this is not feasible 
-are infeasible for large-scale problems in which the operator is only accessible via matrix-vector products (``matrix-free''), because these methods require evaluating the integral kernel of the operator at a large number of points.
-We overcome this problem by c
-
-We improve upon classical product-convolution approximations by translating with respect the estimated mean of the impulse response, rather than the location of the impulse response point source.
-
-We construct a-priori ellipsoid estimates of the supports of the impulse responses (i.e., support ellipsoids are estimated \emph{before} any impulse responses have been computed).
-
-access to $\Aop$ is available through ``operator actions.'' That is,
-
-After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\bm{A}$, that is typically too large to be built and stored.
-This property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\bm{A}$ and $\bm{A}^T$ with arbitrary vectors, but one cannot easily\footnote{One could compute $\bm{v} = \bm{A}\bm{e}_j$, where $\bm{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero, then extract $\bm{A}_{ij} = \bm{v}_i$.
-However, this process is wasteful because one computes the entire vector $\bm{v}$, then discards all but one component.} access matrix entries $\bm{A}_{ij}$.
-
-For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
-
-for a collection of points $x_i$ scattered throughout $\Omega$.
-Further, let $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$,
-\begin{equation}
-%\label{eq:impulse_response_defn}
-\impulseresponse_x(y) := \Aker(y, x),
-\end{equation} 
-be the \emph{impulse response} of $\Aop$ at $x$.
-
-Because we build a H-matrix, and
-
-In this paper we form hierarchical matrix approximations, $\mathbf{A}_H$, of operators $\Aop$ that satisfy the following three properties:
-\begin{description}
-	\item[Locally supported impulse responses:] For all $x\in \Omega$, the support of $\impulseresponse_x$ is contained (or approximately contained) in a localized region.
-	\item[Non-negative impulse responses:] We have
-	\begin{equation*}
-	\impulseresponse_x(y) \ge 0
-	\end{equation*}
-	for all (or approximately all) $(y,x) \in \Omega \times \Omega$.
-	\item[Matrix-free:] One cannot easily evaluate kernel entries $\Aker(y,x)$.
-Instead, access to $\Aop$ is available through ``operator actions.'' That is, we have a black-box computational procedure (which is typically expensive) through which we may evaluate the maps
-	\begin{equation*}
-	u \mapsto\Aop u \quad \text{and} \quad v \mapsto \Aop^Tv
-	\end{equation*} 
-	for arbitrary functions $u$ and $v$.
-After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\bm{A}$, that is typically too large to be built and stored.
-This property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\bm{A}$ and $\bm{A}^T$ with arbitrary vectors, but one cannot easily\footnote{One could compute $\bm{v} = \bm{A}\bm{e}_j$, where $\bm{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero, then extract $\bm{A}_{ij} = \bm{v}_i$.
-However, this process is wasteful because one computes the entire vector $\bm{v}$, then discards all but one component.} access matrix entries $\bm{A}_{ij}$.
-In applications, applying $\bm{A}$ to a vector may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
-	For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}.
-\end{description} 
-Once the hierarchical matrix approximation $\mathbf{A}_H$ is constructed, one may use H-matrix methods to perform further useful linear algebra operations involving $\mathbf{A}_H$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion.
-These H-matrix methods are fast and scale well to large problems.
-WHY DO THESE OPERATIONS: KRYLOV SOLVERS, SAMPLING STRATEGIES, CAN'T DO WITH A.
-
- approximate operators which possess the following properties:
-
-We further limit our focus to operators that have
-\textbf{Local support}
-We say that $\Aop$ has local support if, for all $x\in \Omega$, the support of $\impulseresponse_x$ is contained (or approximately contained) in a small region. 
-The smaller these supports are, the better our algorithm will perform.
-Let $\delta_x$ denote the delta distribution (point source) centered at $x$.
-From \eqref{eq:impulse_response_delta_action}, we see that this local support property means that the response of $\Aop$ to a point source at $x$ is zero (or small) at points $y$ that are far from $x$.
-%\item[3) Local translation invariance:] We say that $\Aop$ is locally translation invariant if
-\begin{equation}
-%	\label{eq:local_translation_invariance}
-	\Aker(y+h, x+h) \approx \Aker(y,x)
-\end{equation}
-when $h$ is not too large.
-One can imagine the kernel entry $\Aker(y,x)$ as representing the strength of an interaction between a source at point $x$ and a target at point $y$.
-Local translation invariance means that the interaction strength remains roughly the same if the source and target points are both translated by the same vector.
-This is illustrated in Figure FIG.
-\textbf{Non-negative impulse response:} We have
-\begin{equation*}
-\impulseresponse_x(y) \ge 0
-\end{equation*}
-for all $(y,x) \in \Omega \times \Omega$.
-
-\subsection{Product-convolution approximation}
-\label{sec:prodconv_intro}
-Let 
-\begin{equation}
-\label{eq:convolution_kernel}
-\convkernel_x(y) := \impulseresponse_x(y+x)
-\end{equation}
-be the translation of $\impulseresponse_x$ to re-center it at zero instead of $x$.
-If $\Aop$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then for any point $x_0$ we would have
-\begin{align*}
-\Aker(y,x) &= \Aker(y+(x_0-x),x+(x_0-x)) \\
-&= \phi_{x_0}(y-x+x_0) \\
-&= \convkernel_{x_0}(y-x)
-\end{align*}
-which would imply
-\begin{align*}
-(\Aop u)(v) &= \int_\Omega \int_\Omega v(y) \convkernel_{x_0}(y-x) u(x) dx dy \\
-&= \int_\Omega v(y) \left(\int_\Omega \convkernel_{x_0}(y-x) u(x) dx\right) dy \\
-&= \left(v, \convkernel_{x_0} \ast u\right)_{L^2(\Omega)},
-\end{align*}
-where $f \ast g$ denotes convolution of the functions $f$ and $g$, and $\left(f, g\right)_{L^2(\Omega)} := \int_\Omega f(x) g(x) dx$ denotes the $L^2$ inner product of $f$ and $g$ on $\Omega$.
-Hence, if $\Aop$ were perfectly translation invariant, it would be the convolution operator $u \mapsto \left(\convkernel_{x_0} \ast u\right)^*$.
-
-Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\convkernel_x$, varies as $x$ changes.
-We therefore approximate $\Aop$ by a spatially varying weighted sum of convolution operators, $\AopPc \approx \Aop$, of the form
-\begin{equation}
-\left(\AopPc~u\right)(v) := \left( v,~ \sum_{i=1}^\convrank \convkernel_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
-\end{equation}
-where $f \cdot g$ denotes pointwise multiplication of the functions $f$ and $g$ (i.e., $\left(f \cdot g\right)(x) = f(x) g(x)$).
-Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution.
-To avoid nested subscripts, we write
-\begin{equation*}
-\convkernel_i := \convkernel_{x_i} \quad \text{and} \quad \impulseresponse_i := \impulseresponse_{x_i},
-\end{equation*}
-with $\convkernel_{x_i}$ defined in \eqref{eq:convolution_kernel} and $\impulseresponse_{x_i}$ defined in \eqref{eq:impulse_response_defn}, to denote local convolution kernels and impulse responses corresponding to a collection of points $\{x_i\}_{i=1}^\convrank \subset \Omega$ scattered throughout the domain.
-The functions $w_i$ are spatially varying weighting functions that are used to interpolate the convolution kernels $\convkernel_i$.
-Our operator approximation method is defined by how we compute the convolution kernels, $\convkernel_i$, how we choose the points, $x_i$, and what weighting functions, $w_i$, that we use.
-In Section \ref{sec:overview_intro} we summarize the answers to these questions; we will provide detailed answers to these questions in Section \ref{sec:method}.
-
-The more locally translation invariant an operator is (i.e., the smaller the discrepancy between the left hand side and right hand side in \eqref{eq:local_translation_invariance}), the smaller the number of terms that are required in \eqref{eq:product_convolution} to achieve an accurate product-convolution approximation.
-
-\subsection{Overview of the method (old)}
-\label{sec:overview_intro}
-
-We compute one ``batch'' of $\convkernel_i$'s by applying $\Aop$ to a sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$.
-The batch of points $x_i$ are chosen so that the support of $\impulseresponse_i$ and the support of $\impulseresponse_j$ do not overlap (or do not overlap much) if $i \neq j$.
-Because these supports do not overlap, we can post-process the response of $\Aop$ to the Dirac comb to recover the functions $\impulseresponse_i$, and therefore the functions $\convkernel_i$, associated with all points $x_i$ in the batch---with one application of $\Aop$, we recover many $\convkernel_i$.
-This is illustrated in Figure FIG.
-The process is repeated to get more batches of $\convkernel_i$'s, until a desired number of batches is reached.
-
-In order to choose the points $x_i$, we need to estimate the support of $\impulseresponse_i$ \emph{before} we compute it.
-The supports of the functions $\impulseresponse_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of specially chosen functions (three functions if $\gdim=1$, six if $\gdim=2$, ten if $\gdim=3$), then post processing the results (Section \ref{eq:mean_and_covariance_estimation}).
-This procedure is the only part of the paper that requires non-negativity of the integral kernel.
-
-We choose the $i$th weighting function, $w_i$, to be the flattest function, in a least-squares sense, that takes the value one at the sample point $x_i$, and the value zero at the other sample points $x_j$, $j \neq i$.
-Computing the weighting functions $w_i$ requires solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid.
-We call this interpolation scheme \emph{Poisson interpolation}.
-The advantage of Poisson interpolation over other interpolation methods, most notably radial basis function interpolation \cite{ESCANDE}, is that Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions.
-This is illustrated in Figure FIG.
-
-Once the $\convkernel_i$ and $w_i$ are computed, we convert $\AopPc$ to hierarchical matrix (H-matrix) format.
-Let $\Aker_\text{pc}$ denote the integral kernel associated with $\AopPc$.
-While $\Aker(y,x)$ is not easily computable, $\Aker_\text{pc}(y,x)$ is given by the formula
-\begin{equation}
-\label{eq:kernel_entries}
-\AkerPc(y,x) = \sum_{i=1}^\convrank \convkernel_i(y-x) w_i(x).
-\end{equation}
-This follows from~\eqref{eq:product_convolution} and the fact that the convolution operator $u \mapsto \convkernel_i \ast u$ has $(y,x)$ kernel entry given by $\convkernel(y-x)$.
-Formula~\eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\AopPc$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
-Once in H-matrix format, fast H-matrix arithmetic is used to invert $\AopPc$, or perform other useful matrix operations.
-
-Often we are interested in approximating operators $\Aop$ that are symmetric positive semi-definite.
-Unfortunately, $\AopPc$ is, in general, non-symmetric and indefinite due to errors in the product convolution approximation.
-In this case, we modify the H-matrix representation of $\AopPc$ to make it positive definite, by symmetrizing it, then applying a specially chosen rational matrix function to the symmetrized H-matrix.
-
-\subsection{Poisson weighting functions}
-\label{sec:weighting_functions}
-
-Consider the following optimization problem,
-\begin{equation}
-\label{eq:wi_optimization_problem_generic}
-\begin{aligned}
-\min_{v \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta v}_{L^2(\Omega)}^2 \\
-\text{such that} &\quad v(x_j) = \interpolatedvalues_j, \quad j=1,\dots,\convrank,
-\end{aligned}
-\end{equation}
-in which one seeks a smooth function $v$ that takes the values $\interpolatedvalues_j \in \mathbb{R}$ at the points $x_j$.
-Here $\Delta$ is the Laplacian on $\Omega$ and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions.
-We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to optimization problem \eqref{eq:wi_optimization_problem_generic}, with
-\begin{equation*}
-b_j = w_i(x_j) = 
-\begin{cases}
-1, & j=i\\
-0, & j \neq i.
-\end{cases}
-\end{equation*}
-In Theorem \ref{thm:smoothest_interpolant}, we show that using these weighting functions yields the ``flattest possible'' \cite{JACOBSEN} interpolation of the convolution kernels, $\convkernel_i$, in a least-squares sense.
-Also, these weighting functions sum to one pointwise (Proposition \ref{thm:wi_sum_to_one}).
-
-The solution to optimization problem \eqref{eq:wi_optimization_problem_generic} is given via the following procedure:
-\begin{enumerate}
-	\item For each sample point $x_i$, we compute a function $\firstgreens_i$ by solving the following Neumann Poisson problem:
-	\begin{equation}
-	\label{eq:psi_eq}
-	\begin{cases}
-	\Delta \firstgreens_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
-	\nu \cdot \nabla \firstgreens_i = 0 & \text{on } \partial \Omega,
-	\end{cases}
-	\end{equation}
-	with condition $\int_\Omega \firstgreens_i(x) dx = 0$.
-Here $\delta_{x_i}$ is the point source centered at $x_i$, $|\Omega|$ is the measure of the domain $\Omega$, and $\nu$ is the normal vector to $\partial \Omega$.
-	\item For each function $\firstgreens_i$, we compute a function $\secondgreens_i$ by solving another Neumann Poisson problem:
-	\begin{equation}
-	\label{eq:theta_eq}
-	\begin{cases}
-	\Delta \secondgreens_i = \firstgreens_i, & \text{in }\Omega, \\
-	\nu \cdot \nabla \secondgreens_i = 0 & \text{on } \partial \Omega,
-	\end{cases}
-	\end{equation}
-	with condition $\int_\Omega \secondgreens_i(x) dx = 0$.
-	\item The solution to optimization problem \eqref{eq:wi_optimization_problem_generic}, which we denote by $v_*$, is formed as follows:
-	\begin{equation}
-	\label{eq:optimal_v}
-	v_*(x) := \constcoeff + \sum_{i=1}^\convrank \secondgreenscoeff_i \secondgreens_i(x),
-	\end{equation}
-	where the constants $\secondgreenscoeff_i \in \mathbb{R}$, $i=1,\dots,\convrank$, and $\constcoeff \in \mathbb{R}$ are the solution to the following linear system:
-	\begin{equation}
-	\label{eq:interpolation_linear_system}
-	\begin{bmatrix}
-	\pointinteractionmatrix & \bm{1} \\ \bm{1}^T & 0
-	\end{bmatrix}
-	\begin{bmatrix}
-	\bm{\secondgreenscoeff} \\ \constcoeff
-	\end{bmatrix}
-	=
-	\begin{bmatrix}
-	\bm{\interpolatedvalues} \\ 0
-	\end{bmatrix}.
-	\end{equation}
-	Here $\bm{\secondgreenscoeff} \in \mathbb{R}^\convrank$ is the vector with $i$th component $\secondgreenscoeff_i$, $\bm{1} \in \bm{R}^\convrank$ is the column vector of all ones, $\bm{\interpolatedvalues} \in \mathbb{R}^\convrank$ is the vector with $i$th component $\interpolatedvalues_i$, and $\pointinteractionmatrix \in \mathbb{R}^{\convrank \times \convrank}$ is the matrix with entries given by
-	\begin{equation}
-	\label{eq:S_matrix}
-	\pointinteractionmatrix_{ij} := \left( \firstgreens_i, \firstgreens_j \right)_{L^2(\Omega)}.
-	\end{equation}
-\end{enumerate}
-Linear system \eqref{eq:interpolation_linear_system} is the KKT system that arises if one solves optimization problem \eqref{eq:wi_optimization_problem_generic} on the $\convrank+1$-dimensional subspace of $H^2_N(\Omega)$ spanned by the functions $\secondgreens_i$, $i=1,\dots,\convrank$, and the constant function $C(x)=1$.
-In Theorem \ref{thm:wi_no_breakdown}, we prove that this procedure does not break down, and in Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v_*$ produced by this procedure is the solution to optimization problem \eqref{eq:wi_optimization_problem_generic} on $H^2_N(\Omega)$.
-
-Constructing all $\convrank$ of the functions $w_i$ thus requires solving the Poisson equation $2\convrank$ times with different right hand side sources to form the functions $\firstgreens_i$ and $\secondgreens_i$, then post-processing these functions.
-This is shown in Algorithm \ref{alg:weighting_functions_incremental}.
-The Poisson PDE solves, which are the most expensive part of this procedure, can be performed cheaply with multigrid.
-
-It is important to note that functions in $H^2_N(\Omega)$ only have well-defined pointwise values if the spatial dimension of the domain is $\gdim=1$, $2$, or $3$.
-Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-posed in three spatial dimensions or fewer.
-In four or higher spatial dimensions, one could construct appropriate weighting functions using radial basis functions, though we do not discuss this here.
-
-\begin{thm}[Procedure for computing $v_*$ does not break down]
-	\label{thm:wi_no_breakdown}
-	If all the sample points $x_i$, $i=1,\dots,\convrank$, are distinct, then PDE \eqref{eq:psi_eq} is uniquely solvable for $\firstgreens_i$, PDE \eqref{eq:theta_eq} is uniquely solvable for $\secondgreens_i$, and linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\bm{\secondgreenscoeff}$ and $\constcoeff$.
-That is, the procedure for computing $v_*$ does not break down.
-Furthermore, $v_* \in H^2_N(\Omega)$.
-\end{thm}
-
-\begin{proof}
-	Existence and uniqueness for the solutions to PDEs \eqref{eq:psi_eq} and \eqref{eq:theta_eq} follow from the standard elliptic PDE theory.
-By standard elliptic regularity theory, $\firstgreens_i \in H^1(\Omega)$, so $\firstgreens_i \in L^2(\Omega)$, and so all matrix entries $\pointinteractionmatrix_{ij}$ are finite.
-
-	Using the definition of $\pointinteractionmatrix$ and basic linear algebra, it is straightforward to show that if $\pointinteractionmatrix$ were singular, there would exist constants $a_1, a_2, \dots a_\convrank$, which are not all zero, such that
-	\begin{equation*}
-	0 = a_1 \firstgreens_1 + a_2 \firstgreens_2 + \dots a_\convrank \firstgreens_\convrank.
-	\end{equation*}
-	Applying the Laplacian to this equation yields
-	\begin{equation*}
-	0 = a_1 \delta_{x_1} + a_2 \delta_{x_2} + \dots a_\convrank \delta_{x_\convrank} - \left(\sum_{i=1}^\convrank a_i\right)\bigg/|\Omega|,
-	\end{equation*}
-	which is impossible because a nonzero linear combination of delta distributions at different points cannot equal a constant function.
-Thus we must conclude that $\pointinteractionmatrix$ is non-singular.
-By block matrix factorization, non-singularity of $\pointinteractionmatrix$ implies that linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\bm{\secondgreenscoeff}$ and $\constcoeff$.
-
-	By standard elliptic regularity theory, the functions $\secondgreens_i$ are in $H^3(\Omega)$, and therefore they are in $H^2(\Omega)$.
-By construction the functions $\secondgreens_i$ have Neumann zero boundary conditions, so $\secondgreens_i \in H^2_N(\Omega)$.
-Since $v_*$ is a linear combination of the functions $\secondgreens_i$ and the constant function $C(x)=1$ which is in $H^2_N(\Omega)$, we have $v_* \in H^2_N(\Omega)$.
-\end{proof}
-
-\begin{lem}[Integral representation of delta distribution on $H^2_N(\Omega)$]
-	\label{lem:psi_delta_eval}
-	We have
-	\begin{equation}
-	\label{eq:inline_wi_lemma}
-	h(x_j) = \left(\firstgreens_j, \Delta h\right)_{L^2(\Omega)} + \avg(h)
-	\end{equation}
-	for all $h\in H^2_N(\Omega)$, where
-	\begin{equation*}
-	\avg(h) :=	 \frac{1}{|\Omega|} \int_\Omega h ~dx,
-	\end{equation*}
-	and $|\Omega| = \int_\Omega 1 ~dx$ is the Lebesgue measure of $\Omega$.
-\end{lem}
-
-\begin{proof}
-	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\firstgreens_j$ onto $h$, and using the Neumann zero boundary conditions for $\firstgreens_j$ and $h$ to eliminate boundary integral terms.
-\end{proof}
-
-\begin{lem}[Characterization of $\pointinteractionmatrix$]
-	\label{lem:S_matrix_characterizations}
-	We have
-	\begin{equation*}
-	\pointinteractionmatrix_{ij} = \secondgreens_j(x_i).
-	\end{equation*}
-\end{lem}
-
-\begin{proof}
-	We compute
-	\begin{equation*}
-	\secondgreens_j(x_i) = \left(\firstgreens_i, \Delta \secondgreens_j\right)_{L^2(\Omega)} + \avg\left(\secondgreens_j\right)
-	= \left(\firstgreens_i, \firstgreens_j\right)_{L^2(\Omega)} = \pointinteractionmatrix_{ij}.
-	\end{equation*}
-	In the first equality we used Lemma \ref{lem:psi_delta_eval}.
-In the second equality we used the facts that $\Delta \secondgreens_j = \firstgreens_j$ and $\avg(\secondgreens_j) = \left(1 / |\Omega|\right) \int_\Omega \secondgreens_j = 0$ from the definition of $\secondgreens_j$ in \eqref{eq:theta_eq}.
-\end{proof}
-
-\begin{thm}[Solution of optimization problem \eqref{eq:wi_optimization_problem_generic}]
-	\label{thm:wi_alg_is_correct}
-	The function $v_*$ given in \eqref{eq:optimal_v} is the unique solution to optimization problem \eqref{eq:wi_optimization_problem_generic}.
-\end{thm}
-
-\begin{proof}
-	Let
-	\begin{equation*}
-	X := \{v \in H^2_N(\Omega) : v(x_i) = \interpolatedvalues_i \text{ for }i=1,\dots,\convrank\}
-	\end{equation*}
-	denote the feasible set for optimization problem \ref{eq:wi_optimization_problem_generic}, and let 
-	\begin{equation*}
-	X_0 := \{v \in H^2_N(\Omega) : v(x_i) = 0 \text{ for }i=1,\dots,\convrank\}
-	\end{equation*}	
-	denote the tangent space to $X$.
-Note that pointwise values of functions in $H_N^2(\Omega)$ are well-defined since the Sobolev imbedding theorem (see, e.g., Theorem 7.19 in \cite{Arbogast}) implies that $H^2(\Omega)$ is continuously imbedded in the space of continuous functions $C^0(\Omega)$ when $\gdim < 4$, and we are considering $\gdim \in \{1,2,3\}$.
-The feasible set $X$ is therefore a closed affine subspace of $H^2_N(\Omega)$, the tangent space $X_0$ is a closed subspace of $H^2_N(\Omega)$, and optimization problem \eqref{eq:wi_optimization_problem_generic} is well-defined.
-
-	The proof now proceeds in three steps.
-First, we show that $v_*$ is feasible ($v_* \in X$).
-Second, we show that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} if it satisfies a certain variational equation.
-Third, we verify that $v_*$ satisfies this variational equation.
-
-	\paragraph{Feasibility} To show that $v_*$ is feasible (i.e., $v_* \in X$), we must show that $v_* \in H^2_N(\Omega)$ and $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,\convrank$.
-We already showed that $v_* \in H^2_N(\Omega)$ in Theorem \ref{thm:wi_no_breakdown}.
-Using the definition of $v_*$ in \eqref{eq:optimal_v} and Lemma \ref{lem:S_matrix_characterizations}, we have
-	\begin{equation*}
-	v_*(x_i) = \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \secondgreens_j(x_i)
-	= \constcoeff + \sum_{j=1}^\convrank \secondgreenscoeff_j \pointinteractionmatrix_{ij} = \left(\pointinteractionmatrix \bm{\secondgreenscoeff} + \constcoeff \bm{1}\right)_i.
-	\end{equation*}
-	The conditions $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,\convrank$ may therefore be written in matrix form as
-	$
-	\pointinteractionmatrix \bm{\secondgreenscoeff} + \constcoeff \bm{1} = \bm{\interpolatedvalues}
-	$, and this matrix equation must hold because it is the first block row of system \eqref{eq:interpolation_linear_system} which is solved to compute $\bm{\secondgreenscoeff}$ and $\constcoeff$.
-So $v_* \in X$.
-
-	\paragraph{Optimality} Since $X$ is an affine subspace, and $X_0$ is the tangent space to $X$, any $v \in X$ may be written in the form $v = v_* + p$ for some $p \in X_0$.
-We may therefore reframe optimization problem \ref{eq:wi_optimization_problem_generic} as
-	\begin{equation}
-	\label{eq:p_optimization}
-	\min_{p \in X_0} J(p),
-	\end{equation}
-	where
-	\begin{equation*}
-	J(p) := \frac{1}{2}||\Delta (v_* + p)||^2.
-	\end{equation*}
-	Showing that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} is equivalent to showing that $0$ is the unique minimizer of \eqref{eq:p_optimization}.
-
-	By direct expansion, we have
-	\begin{equation}
-	\label{eq:Ju_expanded}
-	J(p) = J(0) + \left(\Delta u, \Delta p\right)_{L^2(\Omega)} + \frac{1}{2}||\Delta p||^2.
-	\end{equation}
-	Since $p$ has Neumann zero boundary conditions, if $\Delta p = 0$, then $p$ is constant on each connected component of $\Omega$.
-Furthermore, since $p(x_i)$ is zero for at least one point $x_i$ per connected component of $\Omega$, if $\Delta p = 0$ then $p=0$.
-Hence, by contrapositive we have
-	\begin{align*}
-	\label{eq:contrapositive_laplacian_zero}
-	p \neq 0 \quad \implies \quad& 0 < ||\Delta p||_{L^2(\Omega)}\\
-	\implies \quad& J(0) < J(p) + \left(\Delta v_*, \Delta p\right)_{L^2(\Omega)}.
-	\end{align*}
-	If $v_*$ satisfies the following variational equation,
-	\begin{equation}
-	\label{eq:variational_up}
-	(\Delta v_*, \Delta p)_{L^2(\Omega)} = 0 \quad \text{for all~} p \in X_0,
-	\end{equation} 
-	then
-	\begin{equation*}
-	p \neq 0 \quad \implies \quad J(0) < J(p).
-	\end{equation*}
-	Showing that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} therefore reduces to showing that $v_*$ satisfies variational equation \eqref{eq:variational_up}.
-
-	\paragraph{Variational equation} To show that $v_*$ satisfies variational equation \eqref{eq:variational_up}, notice that
-	\begin{align*}
-	(\Delta v_*, \Delta p)_{L^2(\Omega)} =& \sum_{i=1}^\convrank \secondgreenscoeff_i(\Delta \secondgreens_i, \Delta p)_{L^2(\Omega)} \\
-	=& \sum_{i=1}^\convrank \secondgreenscoeff_i \left(p(x_i) - \avg\left(p\right)\right) \\
-	=& -\avg\left(p\right) \sum_{i=1}^\convrank \secondgreenscoeff_i. 
-	\end{align*}
-	In the first line we used the definition of $\secondgreens_i$ and the fact that the Laplacian of a constant is zero.
-In the second line we used Lemma \ref{lem:psi_delta_eval}.
-In the third line we used the fact that $p(x_i)=0$ because $p \in X_0$.
-Hence variational equation \eqref{eq:variational_up} holds if $\sum_{i=1}^\convrank \secondgreenscoeff_i = 0$, which may be written in matrix form as $\bm{1}^T \bm{\secondgreenscoeff} = 0$.
-This must hold because it is the second block row in system \eqref{eq:interpolation_linear_system}.
-\end{proof}
-
-\begin{thm}[Poisson weighting functions are optimal]
-	\label{thm:smoothest_interpolant}
-	Let 
-	\begin{equation}
-	\label{eq:conv_kernel_pc}
-	\convkernel_x^\text{pc}(y) := \sum_{i=1}^\convrank \convkernel_i(y)w_i(x)
-	\end{equation}
-	denote the local convolution kernel for the product-convolution approximation, and let $\horizinterpolant_y^\text{pc}$ be the reinterpretation of $\convkernel_x^\text{pc}(y)$ as a function of $x$ instead of $y$.
-That is,
-	\begin{equation*}
-	\horizinterpolant_y^\text{pc}(x) := \convkernel_x^\text{pc}(y).
-	\end{equation*}
-	Then for each $y \in \Omega$, the function $\horizinterpolant_y^\text{pc}$ is the unique solution to the following optimization problem:
-	\begin{equation}
-	\label{eq:uy_optimization_problem}
-	\begin{aligned}
-	\min_{\horizinterpolant \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \horizinterpolant}_{L^2(\Omega)}^2 \\
-	\text{such that} &\quad \horizinterpolant(x_i) = \convkernel_i(y), \quad i=1,\dots,\convrank.
-	\end{aligned}
-	\end{equation}
-\end{thm}
-
-\begin{proof}
-	The optimization problem that $w_i$ solves, and optimization problem \eqref{eq:uy_optimization_problem}, both take the form of optimization problem \eqref{eq:wi_optimization_problem_generic}, but with different constraint data.
-Let $\mathcal{T}:\bm{b} \mapsto v_*$ denote the solution operator for optimization problem \eqref{eq:wi_optimization_problem_generic} as a function of the constraint data,  $\bm{b}$.
-The mapping $\mathcal{T}$ is a linear operator because the procedure we presented for solving optimization problem \eqref{eq:wi_optimization_problem_generic} is linear in the constraint data, and we proved that this procedure always yields the unique solution of optimization problem \eqref{eq:wi_optimization_problem_generic} in Theorem \ref{thm:wi_no_breakdown} and Theorem \ref{thm:wi_alg_is_correct}.
-
-	The constraint data for the optimization problem that $w_i$ solves is the unit vector $\bm{e}^{(i)}=(0,\dots,0,1,0,\dots,0) \in \mathbb{R}^\convrank$ which has $i$th component equal to one, and all other components equal to zero.
-Hence
-	\begin{equation*}
-	\label{eq:wi_via_solution_operator}
-	w_i = \mathcal{T}\left(\bm{e}^{(i)}\right).
-	\end{equation*}
-	The constraint data for optimization problem \eqref{eq:uy_optimization_problem} is the vector $(\convkernel_1(y), \convkernel_2(y), \dots, \convkernel_\convrank(y))$, so the solution, $\theta_*$, to optimization problem \eqref{eq:uy_optimization_problem} is given by
-	\begin{equation*}
-	\theta_* = \mathcal{T}\left((\convkernel_1(y), \convkernel_2(y), \dots, \convkernel_\convrank(y))\right).
-	\end{equation*}
-	Using the definition of $\horizinterpolant_y^\text{pc}(x)$ and linearity of $\mathcal{T}$, we have
-	\begin{align*}
-	\horizinterpolant_y^\text{pc}(x) &= \sum_{i=1}^\convrank \convkernel_i(y)w_i(x) \\
-	&= \sum_{i=1}^\convrank \convkernel_i(y)\mathcal{T}\left(\bm{e}^{(i)}\right)(x) \\
-	&= \mathcal{T}\left(\sum_{i=1}^\convrank \convkernel_i(y) \bm{e}^{(i)} \right)(x) \\
-	&= \mathcal{T}\left((\convkernel_1(y), \convkernel_2(y), \dots, \convkernel_\convrank(y))\right)(x) = \theta_*(x).
-	\end{align*}
-	Thus $\horizinterpolant_y^\text{pc}$ is the unique solution to optimization problem \eqref{eq:uy_optimization_problem}.
-\end{proof}
-
-\begin{prop}[Poisson weighting functions sum to one]
-	\label{thm:wi_sum_to_one}
-	We have
-	\begin{equation*}
-	\sum_{i=1}^\convrank w_i(x) = 1
-	\end{equation*}
-	for all $x \in \Omega$.
-\end{prop}
-
-\begin{algorithm2e}
-	\SetAlgoNoLine
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Output}
-	{	
-		\Input{Sample points $\{x_i\}_{i=1}^\convrank$, domain geometry $\Omega$}
-		\Output{Weighting functions $\{w_i\}_{i=1}^\convrank$}
-
-		\For{$k=1,2, \dots, \convrank$}{
-
-			Solve the Neumann Poisson problem \eqref{eq:psi_eq} for $\firstgreens_k$.
-
-			Solve the Neumann Poisson problem \eqref{eq:theta_eq} for $\secondgreens_k$.
-
-		}
-
-		Form the $\convrank \times \convrank$ matrix $\pointinteractionmatrix$ given in \eqref{eq:S_matrix}.
-
-		\For{$i=1,2,\dots,\convrank$}{
-			Solve the linear system
-			\begin{equation*}
-			\begin{bmatrix}
-			\pointinteractionmatrix & \bm{1} \\ \bm{1}^T & 0
-			\end{bmatrix}
-			\begin{bmatrix}
-			\bm{\secondgreenscoeff}^{(i)} \\ \constcoeff^{(i)}
-			\end{bmatrix}
-			=
-			\begin{bmatrix}
-			\bm{e}^{(i)} \\ 0
-			\end{bmatrix}.
-			\end{equation*}
-			for $\bm{\secondgreenscoeff}^{(i)}=(\secondgreenscoeff_1^{(i)}, \secondgreenscoeff_2^{(i)}, \dots, \secondgreenscoeff_\convrank^{(i)})$ and $\constcoeff^{(i)}$.
-Here $\bm{1} = (1,1,\dots,1) \in \mathbb{R}^\convrank$, and $\bm{e}^{(i)} = (0,\dots,0,1,0,\dots,0)\in \mathbb{R}^\convrank$ is the unit vector with $i^\text{th}$ component equal to one and all other components equal to zero.
-
-			Form the weighting function $w_i$ given by
-			\begin{equation*}
-			w_i(x) = \constcoeff^{(i)} + \sum_{k=1}^\convrank \secondgreenscoeff_k^{(i)} \secondgreens_k(x).
-			\end{equation*}
-
-		}
-	}
-	\caption{Compute weighting functions $w_i$, $i=1,\dots,\convrank$.}
-	\label{alg:weighting_functions_incremental}
-\end{algorithm2e}
-
-\subsubsection{Interpolation onto local regular grids}
-\label{sec:convolution_kernel_regular_grid}
-
-To convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), we will need to evaluate $\convkernel_{x_i}(z)$ for a large number of points $z$.
-Per \eqref{eq:varphi_eval}, evaluating $\convkernel_{x_i}(z)$ requires evaluating $\combresponse_k(z+x_i)$.
-In computations, $\combresponse_k$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh.
-Evaluating $\combresponse_k(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell.
-Although this procedure is asymptotically scalable, it is slow in practice \cite{FEMEVALSLOW}.
-
-To speed up evaluation of $\convkernel_{x_i}$, we interpolate $\convkernel_{x_i}$ onto a regular rectilinear grid on a coordinate axis aligned box that just barely contains the ellipsoid $E_{x_i}$.
-We transfer the function $\convkernel_{x_i}$ from the irregular mesh to the regular grid by evaluating $\convkernel_{x_i}$ at all gridpoints in the regular grid, using the slow process described in the previous paragraph.
-This is an upfront cost which is done ``offline.'' After $\convkernel_{x_i}$ has been transferred to the regular grid, we use fast regular grid interpolation to perform all subsequent ``online'' evaluations of $\convkernel_{x_i}$.
-This yields a substantial overall speedup because the number of finite element function evaluations required to transfer the functions $\convkernel_{x_i}$ to the regular grid is $O(\fedim)$, while the number of kernel evaluations required to convert $\AopPc$ to H-matrix format is $O(k \fedim \log \fedim)$.
-Here $\fedim$ is the number of degrees of freedom in the finite element space used to discretize $\combresponse_k$, and $\hmatrixrank$ is the H-matrix rank.
-Furthermore, transferring $\convkernel_{x_i}$ to a regular grid requires evaluating $\convkernel_{x_i}$ at a predetermined and highly structured collection of points.
-This structure allows for faster determination of which mesh cell that a point is in, and allows one to evaluate the finite element function at all regular gridpoints in parallel.
-In contrast, constructing the H-matrix requires evaluating $\convkernel_{x_i}$ at an unpredictable sequence of scattered points, which is an expensive task for finite element function evaluation, but a cheap task for regular grid interpolation.
-Having $\convkernel_{x_i}$ stored on a regular grid also simplifies the implementation of the method that we use to handle boundary artifacts, which will be described in Section \ref{sec:boundary_extension}.
-
-We incur interpolation error when transferring $\convkernel_{x_i}$ from the irregular mesh to the regular grid.
-To limit this interpolation error, the density of the regular grid should be as dense or denser than the local density of the finite element mesh.
-In our numerical results, the distance between neighboring points in the regular grid is chosen to be half the minimum distance between any pair of finite element Lagrange nodes that are contained in $E_{x_i}$.
-The grid transfer procedure is illustrated in Figure FIG.
-
-When $x_i$ is near a boundary, it may occur that some gridpoints for the regular grid are in locations where $\convkernel_{x_i}$ is undefined.
-In this case, we store a placeholder value for $\convkernel_{x_i}$ at those gridpoints.
-We use the IEEE standard value of ``NaN'' (not a number) for this placeholder.
-In the next section (Section \ref{sec:boundary_extension}) we replace these NaN values with more appropriate values.
-
-\subsubsection{Addressing boundary artifacts}
-\label{sec:boundary_extension}
-
-Evaluating the kernel for our product-convolution approximation, $\AkerPc(y,x)$, requires evaluating $\convkernel_{x_i}(y-x)$.
-But $\convkernel_{x_i}(y-x)$ is undefined  when $y - x + x_i$ is outside $\Omega$, which may occur even if $x \in \Omega$ and $y \in \Omega$.
-The default approach to deal with this problem is to extend $\convkernel_{x_i}$ by zero wherever it is undefined.
-But this leads to substantial errors, known as boundary artifacts, which are especially significant when the sample point $x_i$ is near the boundary of the domain.
-In this case, the impulse response is artificially chopped off by the domain boundary.
-See Figure FIG for an illustration.
-
-While $\convkernel_{x_i}$ may be undefined at a point $z$, often there is a sample point $x_j$ that is near $x_i$, but further away from the boundary than $x_i$, such that $\convkernel_{x_j}$ is well defined at $z$.
-To address boundary artifacts, we fill in missing/undefined portions of convolution kernels using information from other convolution kernels associated with the $k$ nearest sample points (we use $k=8$ in our numerical results).
-Kernels associated with nearer sample points are given precedence over kernels associated with farther away sample points.
-
-In detail, let $p_1, p_2, \dots, p_k$ be the $k$ nearest sample points to $x_i$, ordered by nearness, including $x_i$ itself.
-That is, $p_1=x_i$, $p_2$ is the nearest neighbor to $x_i$, $p_3$ is the second nearest neighbor to $x_i$, and so on.
-Also, recall from the previous section (Section \ref{sec:convolution_kernel_regular_grid}) that the functions $\convkernel_{x_i}$ are stored on regular rectilinear grids.
-To extend $\convkernel_{x_i}$, we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\convkernel_{x_i}$, and contains that grid as a subgrid, but is large enough to contain the boxes associated with all of the neighbor kernels, $\convkernel_{p_1}, \convkernel_{p_2}, \dots, \convkernel_{p_k}$ (recall that the functions $\convkernel_{x_i}$ are all centered at zero).
-We initialize an extended convolution kernel, $\convkernel_{x_i}^E$, to zero on this expanded grid.
-Then we interpolate the functions $\convkernel_{p_j}$ onto this expanded grid in reverse order.
-First we interpolate $\convkernel_{p_k}$ onto the grid, then $\convkernel_{p_{k-1}}$, and so on until the last step in which we interpolate the original kernel $\convkernel_{p_1} = \convkernel_{x_i}$.
-At each step, we overwrite existing values of $\convkernel_{x_i}^E$ with new values from the kernel currently being interpolated, except when the new value of $\convkernel_{p_j}$ is undefined (``NaN'').
-At gridpoints where the kernel being interpolated is undefined, we leave $\convkernel_{x_i}^E$ unchanged.
-In this way, we fill in undefined values of $\convkernel_{x_i}$ with values from successively further and further away kernels.
-This process is illustrated in Figure FIG.
-To avoid notational clutter, and since we will henceforth only work with the extended kernel, $\convkernel_{x_i}^E$, in the remainder of this paper the symbol $\convkernel_{x_i}$ refers to $\convkernel_{x_i}^E$.
-
 \section*{Acknowledgments}

### 2022-04-19 paper/localpsf.tex
incredibly productive day adding references

@@
 Typically, the work and memory required to perform these operations for an $N \times N$ H-matrix scales as $O(r^2 N \log(N)^a)$, where $r$ is the largest rank of the off-diagonal blocks, and $a \in \{0,1,2,3\}$.
-The exact cost depends on the type of H-matrix used and operation performed; for more details see \cite{GrasedyckHackbusch03,Hackbusch15}.
+The exact cost depends on the type of H-matrix used and operation performed; for more details see \cite{GrasedyckHackbusch03}.
 The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
@@
 %Since our end goal is an H-matrix, it is natural to ask why we do not build the H-matrix directly using classical H-matrix construction methods. The reason is that 
-Classical H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.
+Classical H-matrix construction methods (described in Appendix \ref{sec:H_matrix_construction}) require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.
 Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}, which we do not use here.
@@
 PC and PSF methods have a long history dating back several decades.
-We note the following papers (among many others),  \cite{Adorf94,AlgerEtAl19,BigotEscandeWeiss19,NagyOleary98,FishEtAl96,EscandeWeiss12,EscandeWeiss15,EscandeWeiss22}, in which the convolution kernels are constructed from sampling impulse responses of the operator to scattered point sources.
+We note the following papers (among many others),  \cite{Adorf94,AlgerEtAl19,BigotEscandeWeiss19,NagyOleary98,FishEtAl96,EscandeWeiss12,EscandeWeiss15,EscandeWeiss22,ZhuLiFomelEtAl16}, in which the convolution kernels are constructed from sampling impulse responses of the operator to scattered point sources.
 For background on PC and PSF methods, we recommend the following papers which provide well-written reviews of the topic: \cite{DenisEtAl15,EscandeWeiss17,GentileCourbinMeylan13}.
@@
 
-If $\Aop$ is symmetric, we take advantage of symmetry to construct a more accurate approximation at negligible additional cost by combining information from impulse responses of both $\Aop$ and $\Aop^T$ in the interpolation procedure (Section \ref{sec:symmetric_improvements}).
+%If $\Aop$ is symmetric, we take advantage of symmetry to construct a more accurate approximation at negligible additional cost by combining information from impulse responses of both $\Aop$ and $\Aop^T$ in the interpolation procedure (Section \ref{sec:symmetric_improvements}).
 Intuitively, impulse responses of $\Aop$ can be thought of as ``columns'' of $\Phi$, impulse responses of $\Aop^T$ can be thought of as ``rows'' of $\Phi$, and $\Phi(y,x)$ can be thought of as an unknown entry of $\Phi$.
@@
 
-When $\Aop$ is positive semi-definite, $\mathbf{A}_H$ may be indefinite due to errors in the approximation.
-In this case, one may (optionally) modify the H-matrix representation of $\mathbf{A}_H$ to make it positive semi-definite, by applying a specially chosen rational matrix function to $\mathbf{A}_H$.
+When $\Aop$ is symmetric positive semi-definite, $\mathbf{A}_H$ may be non-symmetric and indefinite due to errors in the approximation.
+In this case, one may (optionally) modify the H-matrix representation of $\mathbf{A}_H$ to make it symmetric and nearly positive semi-definite.
+The matrix is symmetrized by computing $\left(\mathbf{A}_H+\mathbf{A}_H^T\right)/2$ using H-matrix addition.
+Negative eigenvalues that are large or medium in magnitude are computed, then flipped in sign to become positive.
+Small negative eigenvalues may remain because they are difficult to compute numerically.
 This is described in Appendix \ref{sec:make_spd}.
@@
 
-To find the $k$ nearest sample points to $x$, we query a precomputed KD-tree \cite{kdtree} of all sample points.
-We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{aabbtree} of the mesh cells used to discretize the problem.
+To find the $k$ nearest sample points to $x$, we query a precomputed k-d tree \cite{Bentley75} of all sample points.
+We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{Ericson04} of the mesh cells used to discretize the problem.
 
@@
 $
-and $f \in \mathbb{R}^k$ has entries $f_i$ from \eqref{eq:fxyxp}. We use polyharmonic spline radial basis functions \cite{polyspline}:
+and $f \in \mathbb{R}^k$ has entries $f_i$ from \eqref{eq:fxyxp}. We use polyharmonic spline radial basis functions, which are defined as follows:
 \begin{equation}
@@
 for $r>0$, and $b_i(0):=0$.
-This yields the smoothest interpolant of the given points and values, in a least-squares sense.
-For more details on radial basis function interpolation and polyharmonic splines, see \cite{rbf}.
+This yields the smoothest interpolant of the given points and values, in a certain least-squares sense \cite{Duchon77}.
+For more details on radial basis function interpolation, see \cite{Wendland04}.
 The interpolation procedure is illustrate in Figure \ref{fig:rbf_interpolation}.
@@
 \begin{figure}
-	\begin{subfigure}{0.49\textwidth}
 	\centering
-		\includegraphics[scale=0.22]{interpolation_rbf.pdf}
-		\caption{}
-	\end{subfigure}
-	\begin{subfigure}{0.49\textwidth}
-		\centering
-		\includegraphics[scale=0.22]{interpolation_rbf_boundary.pdf}
-		\caption{}
-	\end{subfigure}
-	\caption{(a) To compute the kernel entry $\Phi(y,x)$, we use radial basis functions to interpolate the values $\frac{V(x)}{V(x_i)}\phi_{x_i}(y-\mu(x)-\mu(x_i))$ for the $k$ nearest sample points $x_i$ to $x$.
-(b) We exclude sample points for which $y-\mu(x)_\mu(x_i)$ is outside of the domain.}
+	\includegraphics[scale=0.2]{interpolation_rbf_boundary.pdf}
+%	\begin{subfigure}{0.49\textwidth}
+%		\centering
+%		\includegraphics[scale=0.22]{interpolation_rbf.pdf}
+%		\caption{}
+%	\end{subfigure}
+%	\begin{subfigure}{0.49\textwidth}
+%		\centering
+%		\includegraphics[scale=0.22]{interpolation_rbf_boundary.pdf}
+%		\caption{}
+%	\end{subfigure}
+	\caption{To compute the kernel entry $\Phi(y,x)$, we use radial basis functions to interpolate the values $\frac{V(x)}{V(x_i)}\phi_{x_i}(y-\mu(x)-\mu(x_i))$ for the $k$ nearest sample points $x_i$ to $x$, excluding sample points for which $y-\mu(x)-\mu(x_i)$ is outside of the domain.}
 	\label{fig:rbf_interpolation}
@@
 We use H1 matrices in our numerical results, but any of the other H-matrix formats (such as H2, HODLR, HSS, HBS, and others \cite{HMATRIX}) could be used instead.
-For more details on H-matrices, we recommend \cite{HMATRIXGOOD}.
-
-We form $\mathbf{\Phi}_H$ using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{HLIBPRO}.
+For more details on H-matrices, we recommend \cite{Hackbusch15}.
+
+We form $\mathbf{\Phi}_H$ using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{Kriemann08}.
 Although $\mathbf{\Phi}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Phi}_H$ only requires evaluation of $O(\hmatrixrank^2 \fedim \log \fedim)$ entries $\mathbf{\Phi}_{ij} = \widetilde{\Phi}(p_i, p_j)$, and these entries are computed via the procedure described in Section \ref{sec:approximate_kernel_entries}.
@@
 
-\section{Content to add}
-
-\begin{itemize}
-	\item computational complexity in terms of PDE solves:
-	\item cost of global low rank approximation (data scalability)
-	\item cost to build approximation
-	\item cost without compression to do a solve
-	\item cost with compression to do a solve
-	\item computational complexity for H-matrix alone vs. prod-conv + h-matrix $O(C k^2 \log \fedim)$ PDE solves, C is big, k is big. Here $10$
-	\item HODLR: C is smaller, but k is bigger, vs. H1 peeling process
-	\item H-matrix operations cost for our method solve
-	\item irregular grid
-\end{itemize}
+%\section{Content to add}
+%
+%\begin{itemize}
+%	\item computational complexity in terms of PDE solves:
+%	\item cost of global low rank approximation (data scalability)
+%	\item cost to build approximation
+%	\item cost without compression to do a solve
+%	\item cost with compression to do a solve
+%	\item computational complexity for H-matrix alone vs. prod-conv + h-matrix $O(C k^2 \log \fedim)$ PDE solves, C is big, k is big. Here $10$
+%	\item HODLR: C is smaller, but k is bigger, vs. H1 peeling process
+%	\item H-matrix operations cost for our method solve
+%	\item irregular grid
+%\end{itemize}
 
@@
 
-\textbf{Heat equation}
-\begin{itemize}
-	\item Hessian per-column error plots for 1, 5, and 15 batches
-	\item $H-P$ error vs number of batches for interior and whole domain
-	\item $P^{-1}H-I$: error vs rank for $P=R$ and $P=P_\text{pch}$ diffusion time, for 1, 5, and 15 batches
-	\item Krylov iterations to tols 1e-1 and 1e-6 vs diffusion parameter
-	\item GLR vs PCH vs diffusion time
-	\item Krylov method convergence plot: $R$ vs $P_\text{pch}$ vs no preconditioner
-	\item Krylov iterations to tols 1e-6 vs mesh size $h$, PCH vs R vs no preconditioner (hold)
-	\item true parameter (H)
-	\item deterministic reconstruction (H)
-	\item noisy measurements (1\%, 5\%, 10\% noise). $u|_\text{top}$ (H)
-	\item recovered state at top (H)
-	\item mesh scalability of PCH (CC) (H)
-	\item put plot data into data directory
-	\item save function .pvd files in paraview  directory
-\end{itemize}
-
-\begin{itemize}
-	\item GLR vs PCH number of obs (Computational cost in PDE solves) (S)
-	\item GLR vs PCH aspect ratio (CC) (S)
-	\item turn on preconditioner after number of krylov iterations exceeds 15 in an iteration. Rebuild "as needed". (S)
-\end{itemize}
-
-\begin{itemize}
-	\item Stokes Krylov convergence (10 solves for nonlinear forward + 6 hessian matvecs for ellipsoid estimation + 10 matvecs for impulse responses + 30 matvecs for PCH6 CG solve + 100 matvecs for Reg CG solve)
-	\item Stokes error vs num batches (small number of batches; 1, 3, 6, 9 batches?) (50 matvecs for randomized error estimation +)
-	\item Velocity field
-	\item Basal friction field
-	\item Picture of impulse responses
-	\item Mesh negative numbers picture
-\end{itemize}
+%\textbf{Heat equation}
+%\begin{itemize}
+%	\item Hessian per-column error plots for 1, 5, and 15 batches
+%	\item $H-P$ error vs number of batches for interior and whole domain
+%	\item $P^{-1}H-I$: error vs rank for $P=R$ and $P=P_\text{pch}$ diffusion time, for 1, 5, and 15 batches
+%	\item Krylov iterations to tols 1e-1 and 1e-6 vs diffusion parameter
+%	\item GLR vs PCH vs diffusion time
+%	\item Krylov method convergence plot: $R$ vs $P_\text{pch}$ vs no preconditioner
+%	\item Krylov iterations to tols 1e-6 vs mesh size $h$, PCH vs R vs no preconditioner (hold)
+%	\item true parameter (H)
+%	\item deterministic reconstruction (H)
+%	\item noisy measurements (1\%, 5\%, 10\% noise). $u|_\text{top}$ (H)
+%	\item recovered state at top (H)
+%	\item mesh scalability of PCH (CC) (H)
+%	\item put plot data into data directory
+%	\item save function .pvd files in paraview  directory
+%\end{itemize}
+%
+%\begin{itemize}
+%	\item GLR vs PCH number of obs (Computational cost in PDE solves) (S)
+%	\item GLR vs PCH aspect ratio (CC) (S)
+%	\item turn on preconditioner after number of krylov iterations exceeds 15 in an iteration. Rebuild "as needed". (S)
+%\end{itemize}
+%
+%\begin{itemize}
+%	\item Stokes Krylov convergence (10 solves for nonlinear forward + 6 hessian matvecs for ellipsoid estimation + 10 matvecs for impulse responses + 30 matvecs for PCH6 CG solve + 100 matvecs for Reg CG solve)
+%	\item Stokes error vs num batches (small number of batches; 1, 3, 6, 9 batches?) (50 matvecs for randomized error estimation +)
+%	\item Velocity field
+%	\item Basal friction field
+%	\item Picture of impulse responses
+%	\item Mesh negative numbers picture
+%\end{itemize}
 
@@
 \subsection{H-matrix construction}
+\label{sec:H_matrix_construction}
 
@@
 \end{equation}
-where $U^+$ is the pseudoinverse of $U$ \cite{CUR}.
+where $U^+$ is the pseudoinverse of $U$ \cite{goreinov1997theory,MahoneyDrineas09}.
 The quality of approximation \eqref{eq:CUR} depends on how well the columns of $C$ approximate the column space of $X$, and how well the rows of $R$ approximate the row space of $X$.
@@
 An initial set of columns $C$ is chosen.
-Keeping $C$ fixed, a set of rows $R$ is chosen so that the so that determinant of the submatrix $U$ within $C$ is as large as possible.
-This may be done via either the maxvol procedure \cite{GOODSUBMATRIX}, or via QR factorization.
+Keeping $C$ fixed, a set of rows $R$ is chosen so that the so that determinant of the submatrix $U$ within $C$ is as large as possible \cite{GoreinovEtAl10}.
 Now, keeping $R$ fixed, a set of new columns $C$ is chosen so that the submatrix $U$ within $R$ is as large as possible.
@@
 There is also a small linear algebra overhead cost for the determinant maximization process that is performed at each step.
-For more details on adaptive cross low rank approximation, see \cite{ACA}.
+%For more details on adaptive cross low rank approximation, see \cite{ACA}.
 The key point is that the adaptive cross method allows us to form a rank-$\hmatrixrank$ approximation of an $m \times m$ block of the matrix via a process that only requires accessing $O(m\hmatrixrank)$ entries of that block.
@@
 \end{equation*}
-In \cite{ELLIPSOIDINTERSECT} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$.
+In \cite{GilitschenskiHanebeck12} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$.
 We check whether $E_1$ and $E_2$ intersect by minimizing $K(s)$ on $(0,1)$.
@@
 where $v := \Phi^T\left(\spatialmean_1 - \spatialmean_2\right)$.
-We compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's algorithm \cite{BRENT} (any fast 1 dimensional convex optimization routine may be used).
+We compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's algorithm \cite{Brent71} (any fast 1 dimensional convex optimization routine may be used).
 The resulting algorithm for checking whether $E_1$ and $E_2$ intersect is summarized in Algorithm \ref{alg:ellipsoid_intersection_test}.
@@
 \end{equation*}
-These outlying ``bad'' eigenpairs may be computed using a Krylov method (we use Lanczos), or the randomized singular value decomposition.
+These outlying ``bad'' eigenpairs are computed using the Lanczos method \cite{Komzsik03,Lanczos50}.
 We then flip the sign of these negative eigenvalues to be positive as follows:
@@
 \section*{Acknowledgments}
-We thank J.J. Alger, Longfei Gao, Mathew Hu, and Rami Nammour for helpful discussions. Georg Stadler (domain geometry help)
+We thank J.J. Alger, Longfei Gao, Mathew Hu, and Rami Nammour for helpful discussions. We thank Trevor Heise for editing suggestions. We thank Georg Stadler for domain geometry help.
 

### 2022-04-19 paper/localpsf.tex
add lumped mass matrix hughes reference

@@
 If there are too many negative numbers, or if the negative numbers are too large, our algorithm will be less robust.
-We therefore recommend replacing the mass matrix $\mathbf{M}$ with a positive diagonal \emph{lumped mass} approximation, and using the resulting lumped mass matrix everywhere throughout the approximation method.
-For typical finite element meshes, lumped mass matrix approximations are spectrally equivalent to the mass matrix \cite{MASSLUMP2}.
+We therefore recommend replacing the mass matrix $\mathbf{M}$ with a positive diagonal \emph{lumped mass} approximation.
+%For typical finite element meshes, lumped mass matrix approximations are spectrally equivalent to the mass matrix \cite{MASSLUMP2}. 
 In our numerical results, we use the lumped mass matrix constructed by replacing off-diagonal entries of the mass matrix by zero.
-Other more advanced mass lumping techniques may be used \cite{MASSLUMP2}.
+Other more advanced mass lumping techniques may be used.
+See \cite{Hughes12} for a discussion of lumped mass matrices.
 

### 2022-04-26 paper/localpsf.tex
update changes during discussion with Tucker

@@
 \end{equation*}
+Then we use the high order rational function spectral projection method from \cite{HackbuschKress07} to replace the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ with zero (or approximately zero) while keeping the positive eigenvalues unchanged (approximately unchanged).
+
 Then we compute the negative eigenvalues, of $\mathbf{A}_H$ that are less than (i.e., more negative than) a small magnitude negative cutoff value $c$, and also compute the corresponding eigenvectors.

### 2022-04-26 numerical_examples/heat/error_vs_krylov_iter.py

@@ -51,11 +51,12 @@ PCK = ProductConvolutionKernel(HIP.V, HIP.V, HIP.apply_Hd_petsc, HIP.apply_Hd_pe
 
 A_pch_nonsym, extras = make_hmatrix_from_kernel(PCK, hmatrix_tol=hmatrix_tol)
 
-_, eeA_max, _ = spla.svds(A_pch_nonsym.as_linear_operator(), k=3, which='LM')
-eA_max = np.max(eeA_max)
+# _, eeA_max, _ = spla.svds(A_pch_nonsym.as_linear_operator(), k=3, which='LM')
+# eA_max = np.max(eeA_max)
 
 # cutoff = -1e-2*eA_max
 
+# A_pch = A_pch_nonsym.spd(rtol=1e-3, atol=1e-5)
 A_pch = A_pch_nonsym.spd()
 
 # min_reg_param = 1e-3

### 2022-04-27 paper/localpsf.tex
make figure and writeup for improved spd modification

--- 2022-04-19 paper/localpsf.tex
@@
 
-If $\Aop$ is symmetric positive definite, we symmetrize $\mathbf{A}_H$ as follows
-\begin{equation*}
-\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
-\end{equation*}
-Then we compute the negative eigenvalues, of $\mathbf{A}_H$ that are less than (i.e., more negative than) a small magnitude negative cutoff value $c$, and also compute the corresponding eigenvectors.
+If $\Aop$ is symmetric positive semidefinite, we symmetrize $\mathbf{A}_H$ as follows
+\begin{equation*}
+\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right),
+\end{equation*}
+then we use the high order rational function spectral projection method from \cite{HackbuschKress07} to replace (or approximately replace) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ with zero.
+In detail, let $f$ be the following scalar rational function,
+\begin{equation*}
+	f(\lambda) := \begin{cases}
+		\lambda, & \lambda \ge 0, \\
+		0, & \lambda < 0,
+	\end{cases}
+\end{equation*}
+and extend $f$ to matrices via functional calculus.
+The matrix $f(\mathbf{A}_H^{\text{sym}})$ is the ideal positive semidefinite modification of $\mathbf{A}_H^{\text{sym}}$, because $f(\mathbf{A}_H^{\text{sym}})$ is the same as $\mathbf{A}_H^{\text{sym}}$ except all the negative eigenvalues are replaced with zero.
+Computing $f(\mathbf{A}_H^{\text{sym}})$ directly is difficult, but we may approximate $f$ by the high order rational function
+\begin{equation*}
+	f(\lambda) \approx \widetilde{f}(\lambda) := \frac{\lambda}{1 + \left(\frac{2 \lambda - M}{M}\right)^{2^k} },
+\end{equation*}
+where $k$ is a positive integer, and $M$ is a positive real number that is larger than the largest eigenvalue of $\mathbf{A}_H^{\text{sym}}$.
+In practice we use $k=5$, and $M=1.5~ \lambda_\text{max}$, where $\lambda_\text{max}$ is the largest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, which is computed using the Lanczos iteration.
+If $k$ is sufficiently large, then $\widetilde{f}(\lambda) \approx \lambda$ for $\lambda \in (0,M)$ and $\widetilde{f}(\lambda) \approx 0$ for $\lambda$ outside of $[0,M]$, so $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ is a good approximation of  $f(\mathbf{A}_H^{\text{sym}})$.
+This is illustrated in Figure \ref{fig:spd_rat}.
+\begin{figure}
+	\centering
+	\includegraphics[scale=0.5]{spd_rational_function.pdf}
+	\caption{Rational function $\widetilde{f}(\lambda)$ for $M=1$ and $k=3,5,7$. As $k$ increases, $\widetilde{f}(\lambda)$ approaches the sawtooth function supported on $(0,M)$. }
+	\label{fig:spd_rat}
+\end{figure}
+
+Using $\widetilde{f}$ instead of $f$ permits us to efficiently compute $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ using H-matrix methods.
+Replacing constants with scaled identity matrices, scalar powers and multiplication with matrix powers and multiplication, and scalar division with matrix inversion, the extension of $\widetilde{f}$ to matrices becomes
+\begin{equation}
+\label{eq:rational343}
+\widetilde{f}(\mathbf{A}_H^{\text{sym}}) = \left(\mathbf{I} + \mathbf{X}^{2^k}\right)^{-1} \mathbf{A}_H^{\text{sym}},
+\end{equation}
+where
+\begin{equation*}
+	\mathbf{X} := \left(2 \mathbf{A}_H^{\text{sym}} - M \mathbf{I}\right)/M.
+\end{equation*}
+We compute $\mathbf{X}$ using fast H-matrix scalar multiplication and addition. Then we form the $\mathbf{X}^{2^k}$ via repeated squaring,
+\begin{align*}
+	\mathbf{X}^2 &= \mathbf{X} \mathbf{X} \\
+	\mathbf{X}^4 &= \mathbf{X}^2 \mathbf{X}^2 \\
+	&\vdots \\
+	\mathbf{X}^{2^k} &= \mathbf{X}^{2^{k-1}} \mathbf{X}^{2^{k-1}},
+\end{align*}
+where each of these matrix squarings is performed using fast H-matrix multiplication.
+Finally, $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ is formed from $\mathbf{X}^{2^k}$ and $\mathbf{A}_H^{\text{sym}}$ per \eqref{eq:rational343}, using fast H-matrix operations to perform the required matrix-matrix addition, matrix inversion, and matrix-matrix multiplication.
+At the end, the matrix is symmetrized once again, to correct for any small non-symmetries that are introduced due to truncation error in the H-matrix arithmetic.
+Once constructed, we use the symmetric positive semidefinite matrix $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ in place of $\mathbf{A}_H$ in further linear algebra operations.
+
+% the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ with zero (or approximate zero) while keeping the positive eigenvalues unchanged (approximately unchanged). Ideally
+%
+%
+%Then we compute the negative eigenvalues, of $\mathbf{A}_H$ that are less than (i.e., more negative than) a small magnitude negative cutoff value $c$, and also compute the corresponding eigenvectors.
 That is, the eigenpairs $\{(\lambda_i,u_i)\}_{i=1}^{N_\text{bad}}$ with
-\begin{equation*}
-	\lambda_i < c < 0.
-\end{equation*}
-These outlying ``bad'' eigenpairs are computed using the Lanczos method \cite{Komzsik03,Lanczos50}.
-We then flip the sign of these negative eigenvalues to be positive as follows:
-\begin{equation*}
-	\mathbf{A}_H^{\text{sym}+} = \mathbf{A}_H^{\text{sym}} + \sum_{i=1}^{N_\text{bad}} 2 |\lambda_i| u_i u_i^T.
-\end{equation*}
-The matrix $\mathbf{A}_H^{\text{sym}+}$ approximates the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$.
+%\begin{equation*}
+%	\lambda_i < c < 0.
+%\end{equation*}
+%These outlying ``bad'' eigenpairs are computed using the Lanczos method \cite{Komzsik03,Lanczos50}.
+%We then flip the sign of these negative eigenvalues to be positive as follows:
+%\begin{equation*}
+%	\mathbf{A}_H^{\text{sym}+} = \mathbf{A}_H^{\text{sym}} + \sum_{i=1}^{N_\text{bad}} 2 |\lambda_i| u_i u_i^T.
+%\end{equation*}
+%The matrix $\mathbf{A}_H^{\text{sym}+}$ approximates the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$.
 A H-matrix representation of $\mathbf{A}_H^{\text{sym}+}$ is formed using H-matrix addition, and used in place of $\mathbf{A}_H$ in further linear algebra operations.
-
-The cutoff $c$ is required when $\Aop$ has a large number of zero eigenvalues, or a sequence of eigenvalues that cluster at zero, a situation that occurs commonly in practice.
+%
+%The cutoff $c$ is required when $\Aop$ has a large number of zero eigenvalues, or a sequence of eigenvalues that cluster at zero, a situation that occurs commonly in practice.
 We seek to correct negative eigenvalues caused by errors in our approximation procedure.
@@
 Numerically, we cannot hope to correct these slightly negative eigenvalues because there are so many of them, and because the spectral gap between the smallest negative eigenvalue and the first non-negative eigenvalue is small.
-
-For Hessians in ill-posed inverse problems, the Hessian is the sum of the Hessian of a data misfit term, and the Hessian of a regularization term.
+%
+%For Hessians in ill-posed inverse problems, the Hessian is the sum of the Hessian of a data misfit term, and the Hessian of a regularization term.
 The Hessian of the regularization term is known and easy to manipulate, and one seeks to approximate the Hessian of the data misfit term with our method.

### 2022-04-27 numerical_examples/heat/error_vs_krylov_iter.py

@@ -201,387 +201,3 @@ plt.legend(legend)
 if save_figures:
     plt.savefig(save_dir / 'relative_residual_vs_krylov_iter.pdf', bbox_inches='tight', dpi=100)
 
-
-# ###################################################################################################
-# ###################################################################################################
-# ###################################################################################################
-# ###################################################################################################
-#
-# ##############3
-#
-# class LinopWithCounter(spla.LinearOperator):
-#     def __init__(me, linop):
-#         me.matvec_counter = 0
-#         me.rmatvec_counter = 0
-#         me.linop = linop
-#         me.shape = linop.shape
-#         me.dtype = linop.dtype
-#
-#     def matvec(me, X):
-#         me.matvec_counter += 1
-#         print('matvec_counter=', me.matvec_counter)
-#         return me.linop.matvec(X)
-#         # if len(X.shape)==1:
-#         #     X = X.reshape((-1,1))
-#         # Y = np.zeros((me.shape[0], X.shape[1]))
-#         # for k in range(X.shape[1]):
-#         #     me.matvec_counter += 1
-#         #     print('matvec_counter=', me.matvec_counter)
-#         #     Y[:,k] = me.linop.matvec(X[:,k])
-#         # return Y
-#
-#     def rmatvec(me, X):
-#         if len(X.shape) == 1:
-#             X = X.reshape((-1, 1))
-#         Y = np.zeros((me.shape[1], X.shape[1]))
-#         for k in range(X.shape[1]):
-#             me.rmatvec_counter += 1
-#             print('rmatvec_counter=', me.rmatvec_counter)
-#             Y[:, k] = me.linop.rmatvec(X[:, k])
-#         return Y
-#
-#     def reset_counters(me):
-#         me.matvec_counter = 0
-#         me.rmatvec_counter = 0
-#
[…]

### 2022-04-27 numerical_examples/heat/example_heat.py

@@ -54,11 +54,12 @@ PCK.col_batches.visualize_impulse_response_batch(0)
 
 A_pch_nonsym, extras = make_hmatrix_from_kernel(PCK, hmatrix_tol=1e-3)
 
-min_reg_param = 1e-3
-eeR_min, _ = spla.eigsh(min_reg_param * HIP.R0_scipy, k=3, which='SM')
-eR_min = np.min(eeR_min)
+# min_reg_param = 1e-3
+# eeR_min, _ = spla.eigsh(min_reg_param * HIP.R0_scipy, k=3, which='SM')
+# eR_min = np.min(eeR_min)
 
-A_pch = A_pch_nonsym.spd(cutoff=0.8*eR_min)
+# A_pch = A_pch_nonsym.spd(cutoff=0.8*eR_min)
+A_pch = A_pch_nonsym.spd()
 
 Phi_pch = extras['A_kernel_hmatrix']
 

### 2022-04-27 numerical_examples/heat/ellipsoid_intersection_figure.py
add ellipsoid intersection figure

@@ -1,0 +1,99 @@
+import numpy as np
+from scipy.linalg import eigh
+from scipy.optimize import minimize_scalar
+import matplotlib.pyplot as plt
+from matplotlib.patches import Ellipse
+
+
+def ellipsoid_intersection_test(Sigma_A, Sigma_B, mu_A, mu_B, tau):
+    lambdas, Phi, v_squared = ellipsoid_intersection_test_helper(Sigma_A, Sigma_B, mu_A, mu_B)
+    res = minimize_scalar(ellipsoid_K_function,
+                          bracket=[0.0, 0.5, 1.0],
+                          args=(lambdas, v_squared, tau))
+    return (res.fun[0] >= 0)
+
+
+def ellipsoid_intersection_test_helper(Sigma_A, Sigma_B, mu_A, mu_B):
+    lambdas, Phi = eigh(Sigma_A, b=Sigma_B)
+    v_squared = np.dot(Phi.T, mu_A - mu_B) ** 2
+    return lambdas, Phi, v_squared
+
+
+def ellipsoid_K_function(ss, lambdas, v_squared, tau):
+    ss = np.array(ss).reshape((-1,1))
+    lambdas = np.array(lambdas).reshape((1,-1))
+    v_squared = np.array(v_squared).reshape((1,-1))
+    return 1.-(1./tau**2)*np.sum(v_squared*((ss*(1.-ss))/(1.+ss*(lambdas-1.))), axis=1)
+
+
+
+tau=1.3
+
+Sigma_A = np.array([[0.7, 0.02 ], [0.02, 1.1]])
+Sigma_B = np.array([[ 0.6, -0.5], [-0.5, 0.7]])
+
+mu_A = np.array([ 0.9, -0.75])
+mu_B0 = np.array([-1.4, -0.6])
+
+
+def plot_ellipse(mu, Sigma, n_std_tau, ax=None, **kwargs):
+    # Usage:
+    #   https://github.com/NickAlger/nalger_helper_functions/tree/master/jupyter_notebooks/plot_ellipse.ipynb
+    if ax is None:
+        ax = plt.gca()
+
+    ee, V = np.linalg.eigh(Sigma)
+    e_big = ee[1]
+    e_small = ee[0]
+    v_big = V[:, 1]
+    theta = np.arctan(v_big[1] / v_big[0]) * 180. / np.pi
+
[…]

### 2022-04-27 paper/localpsf.tex
add ellipsoid intersection figure
k=7

@@
 
+\begin{figure}
+	\begin{subfigure}{0.32\textwidth}
+		\centering
+		\includegraphics[scale=0.40]{ellipsoids_intersect0.pdf}
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\centering
+		\includegraphics[scale=0.40]{ellipsoids_intersect1.pdf}
+	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\centering
+		\includegraphics[scale=0.40]{ellipsoids_intersect2.pdf}
+	\end{subfigure}
+	\caption{The ellipsoids intersect if $K(s) \ge 0$ for all $s \in (0,1)$.
+	}
+	\label{fig:ellipsoid_intersection_test}
+\end{figure}
+
 The function $K(s)$ may be evaluated quickly for many $s$ by pre-computing the solution to the generalized eigenvalue problem
@@
 \begin{equation*}
-	f(\lambda) \approx \widetilde{f}(\lambda) := \frac{\lambda}{1 + \left(\frac{2 \lambda - M}{M}\right)^{2^k} },
-\end{equation*}
-where $k$ is a positive integer, and $M$ is a positive real number that is larger than the largest eigenvalue of $\mathbf{A}_H^{\text{sym}}$.
-In practice we use $k=5$, and $M=1.5~ \lambda_\text{max}$, where $\lambda_\text{max}$ is the largest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, which is computed using the Lanczos iteration.
-If $k$ is sufficiently large, then $\widetilde{f}(\lambda) \approx \lambda$ for $\lambda \in (0,M)$ and $\widetilde{f}(\lambda) \approx 0$ for $\lambda$ outside of $[0,M]$, so $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ is a good approximation of  $f(\mathbf{A}_H^{\text{sym}})$.
+	f(\lambda) \approx \widetilde{f}(\lambda) := \frac{\lambda}{1 + \left(\frac{2 \lambda - (b+a)}{b-a}\right)^{2^k} },
+\end{equation*}
+where $k$ is a positive integer, $a$ is either zero, or a small positive real number, and $b$ is a positive real number that is larger than the largest eigenvalue of $\mathbf{A}_H^{\text{sym}}$.
+If $k$ is sufficiently large, then $\widetilde{f}(\lambda) \approx \lambda$ for $\lambda \in (a,b)$ and $\widetilde{f}(\lambda) \approx 0$ for $\lambda$ outside of $[a,b]$, so $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ is a good approximation of  $f(\mathbf{A}_H^{\text{sym}})$.
 This is illustrated in Figure \ref{fig:spd_rat}.
+In practice we use $k=7$, $a=0$ and $b=1.5~ \lambda_\text{max}$, where $\lambda_\text{max}$ is the largest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, which is computed using the Lanczos iteration.
 \begin{figure}
@@
 \begin{equation*}
-	\mathbf{X} := \left(2 \mathbf{A}_H^{\text{sym}} - M \mathbf{I}\right)/M.
+	\mathbf{X} := \left(2 \mathbf{A}_H^{\text{sym}} - (b+a) \mathbf{I}\right)/(b-a).
 \end{equation*}

### 2022-04-27 numerical_examples/heat/error_vs_krylov_iter.py

@@ -57,7 +57,36 @@ A_pch_nonsym, extras = make_hmatrix_from_kernel(PCK, hmatrix_tol=hmatrix_tol)
 # cutoff = -1e-2*eA_max
 
 # A_pch = A_pch_nonsym.spd(rtol=1e-3, atol=1e-5)
-A_pch = A_pch_nonsym.spd()
+A_pch0 = A_pch_nonsym.sym()
+A_pch3 = A_pch_nonsym.spd(k=3, a_factor=0.0)
+A_pch5 = A_pch_nonsym.spd(k=5, a_factor=0.0)
+A_pch7 = A_pch_nonsym.spd(k=7, a_factor=0.0)
+
+A_pch = A_pch7
+
+A0_dense = build_dense_matrix_from_matvecs(A_pch0.matvec, A_pch0.shape[1])
+A3_dense = build_dense_matrix_from_matvecs(A_pch3.matvec, A_pch3.shape[1])
+A5_dense = build_dense_matrix_from_matvecs(A_pch5.matvec, A_pch5.shape[1])
+A7_dense = build_dense_matrix_from_matvecs(A_pch7.matvec, A_pch7.shape[1])
+
+
+import scipy.linalg as sla
+ee0, uu0 = sla.eigh(A0_dense)
+ee3, uu3 = sla.eigh(A3_dense)
+ee5, uu5 = sla.eigh(A5_dense)
+ee7, uu7 = sla.eigh(A7_dense)
+
+print('np.min(ee0)=', np.min(ee0))
+print('np.min(ee3)=', np.min(ee3))
+print('np.min(ee5)=', np.min(ee5))
+print('np.min(ee7)=', np.min(ee7))
+
+plt.figure()
+plt.plot(ee0)
+plt.plot(ee3)
+plt.plot(ee5)
+plt.plot(ee7)
+plt.legend(['ee0', 'ee3', 'ee5', 'ee7'])
 
 # min_reg_param = 1e-3
 # eeR_min, _ = spla.eigsh(min_reg_param * HIP.R0_scipy, k=3, which='SM')

### 2022-04-29 paper/localpsf.tex
wrote up spd modification technique in appendix

@@
 
-If $\Aop$ is symmetric positive semidefinite, we symmetrize $\mathbf{A}_H$ as follows
-\begin{equation*}
-\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right),
-\end{equation*}
-then we use the high order rational function spectral projection method from \cite{HackbuschKress07} to replace (or approximately replace) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ with zero.
-In detail, let $f$ be the following scalar rational function,
-\begin{equation*}
-	f(\lambda) := \begin{cases}
-		\lambda, & \lambda \ge 0, \\
-		0, & \lambda < 0,
+If $\Aop$ is symmetric, we symmetrize $\mathbf{A}_H$ as follows
+\begin{equation*}
+\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
+\end{equation*}
+If $\Aop$ is also positive semi-definite, we use the high order rational function spectral projection method from \cite{HackbuschKress07} to flip (or approximately flip) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ to be positive.
+In detail, let $\Pi$ be the indicator function for the interval $(a,b)$,
+\begin{equation*}
+	\Pi(\lambda) = \begin{cases}
+		1, & \lambda \in (a,b), \\
+		0, & \text{otherwise},
 	\end{cases}
 \end{equation*}
-and extend $f$ to matrices via functional calculus.
-The matrix $f(\mathbf{A}_H^{\text{sym}})$ is the ideal positive semidefinite modification of $\mathbf{A}_H^{\text{sym}}$, because $f(\mathbf{A}_H^{\text{sym}})$ is the same as $\mathbf{A}_H^{\text{sym}}$ except all the negative eigenvalues are replaced with zero.
-Computing $f(\mathbf{A}_H^{\text{sym}})$ directly is difficult, but we may approximate $f$ by the high order rational function
-\begin{equation*}
-	f(\lambda) \approx \widetilde{f}(\lambda) := \frac{\lambda}{1 + \left(\frac{2 \lambda - (b+a)}{b-a}\right)^{2^k} },
-\end{equation*}
-where $k$ is a positive integer, $a$ is either zero, or a small positive real number, and $b$ is a positive real number that is larger than the largest eigenvalue of $\mathbf{A}_H^{\text{sym}}$.
-If $k$ is sufficiently large, then $\widetilde{f}(\lambda) \approx \lambda$ for $\lambda \in (a,b)$ and $\widetilde{f}(\lambda) \approx 0$ for $\lambda$ outside of $[a,b]$, so $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ is a good approximation of  $f(\mathbf{A}_H^{\text{sym}})$.
+and extend $\Pi$ to matrices via functional calculus.
+The matrix $\Pi(\mathbf{A}_H^{\text{sym}})$ is the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to eigenvalues of $\mathbf{A}_H^{\text{sym}}$ that reside in $(a,b)$.
+If we choose $a < \lambda_\text{min}$ and $b=0$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, then the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$ is given by
+\begin{equation}
+\label{eq:I_Pi_abs}
+	|\mathbf{A}_H^{\text{sym}}| = \left(\mathbf{I} - 2\Pi(\mathbf{A}_H^{\text{sym}})\right)\mathbf{A}_H^{\text{sym}}.
+\end{equation}
+That is, $|\mathbf{A}_H^{\text{sym}}|$ is the matrix that has the same eigenvectors as $\mathbf{A}_H^{\text{sym}}$, but the eigenvalues of $|\mathbf{A}_H^{\text{sym}}|$ are the absolute values of the eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
+
+Computing $\Pi(\mathbf{A}_H^{\text{sym}})$ directly is difficult, but we may approximate $\Pi$ by the following high order rational function:
+\begin{equation*}
+	\Pi(\lambda) \approx \Pi_k(\lambda) := \frac{1}{1 + T(\lambda)^{2^k} }, \qquad \text{where} \qquad T(\lambda) := \frac{2 \lambda - (b+a)}{b-a}
+\end{equation*}
+and $k$ is a positive integer; the larger $k$ is, the better the approximation $\Pi \approx \Pi_k$.
 This is illustrated in Figure \ref{fig:spd_rat}.
-In practice we use $k=7$, $a=0$ and $b=1.5~ \lambda_\text{max}$, where $\lambda_\text{max}$ is the largest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, which is computed using the Lanczos iteration.
+For $k \ge 1$, $\Pi_k(\lambda)> 1/2$ for $\lambda \in (a,b)$, so choosing $a < \lambda_\text{min}$, $b=0$, and using $\Pi_k$ instead of $\Pi$ in \eqref{eq:I_Pi_abs} yields a matrix that is positive semidefinite.
+Replacing constants with scaled identity matrices, scalar powers and multiplication with matrix powers and multiplication, and scalar division with matrix inversion, the extension of $\Pi_k$ to matrices becomes
+\begin{equation}
+\label{eq:rational343}
+\Pi_k\left(\mathbf{A}_H^{\text{sym}}\right) = \left(\mathbf{I} + \mathbf{T}^{2^k}\right)^{-1},
+\end{equation}
+where
+\begin{equation*}
+\mathbf{T} := \left(2 \mathbf{A}_H^{\text{sym}} - (b+a) \mathbf{I}\right)/(b-a).
+\end{equation*}
+
 \begin{figure}
@@
 	\includegraphics[scale=0.5]{spd_rational_function.pdf}
-	\caption{Rational function $\widetilde{f}(\lambda)$ for $M=1$ and $k=3,5,7$. As $k$ increases, $\widetilde{f}(\lambda)$ approaches the sawtooth function supported on $(0,M)$. }
+	\caption{Rational function $\Pi_k(\lambda)$ for $a=-1$, $b=0$, and $k \in \{1,2,3\}$. As $k$ increases, $\Pi_k$ approaches the indicator function for the set $(a,b)$. }
 	\label{fig:spd_rat}
@@
 
-Using $\widetilde{f}$ instead of $f$ permits us to efficiently compute $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ using H-matrix methods.
-Replacing constants with scaled identity matrices, scalar powers and multiplication with matrix powers and multiplication, and scalar division with matrix inversion, the extension of $\widetilde{f}$ to matrices becomes
-\begin{equation}
-\label{eq:rational343}
-\widetilde{f}(\mathbf{A}_H^{\text{sym}}) = \left(\mathbf{I} + \mathbf{X}^{2^k}\right)^{-1} \mathbf{A}_H^{\text{sym}},
-\end{equation}
-where
-\begin{equation*}
-	\mathbf{X} := \left(2 \mathbf{A}_H^{\text{sym}} - (b+a) \mathbf{I}\right)/(b-a).
-\end{equation*}
-We compute $\mathbf{X}$ using fast H-matrix scalar multiplication and addition. Then we form the $\mathbf{X}^{2^k}$ via repeated squaring,
-\begin{align*}
-	\mathbf{X}^2 &= \mathbf{X} \mathbf{X} \\
-	\mathbf{X}^4 &= \mathbf{X}^2 \mathbf{X}^2 \\
-	&\vdots \\
-	\mathbf{X}^{2^k} &= \mathbf{X}^{2^{k-1}} \mathbf{X}^{2^{k-1}},
-\end{align*}
-where each of these matrix squarings is performed using fast H-matrix multiplication.
-Finally, $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ is formed from $\mathbf{X}^{2^k}$ and $\mathbf{A}_H^{\text{sym}}$ per \eqref{eq:rational343}, using fast H-matrix operations to perform the required matrix-matrix addition, matrix inversion, and matrix-matrix multiplication.
-At the end, the matrix is symmetrized once again, to correct for any small non-symmetries that are introduced due to truncation error in the H-matrix arithmetic.
-Once constructed, we use the symmetric positive semidefinite matrix $\widetilde{f}(\mathbf{A}_H^{\text{sym}})$ in place of $\mathbf{A}_H$ in further linear algebra operations.
+Using $\Pi_k$ instead of $\Pi$ permits us to efficiently approximate $|\mathbf{A}_H^{\text{sym}}|$ using H-matrix methods. 
+First, we compute $\lambda_\text{min}$ using the Lanczos method, and set $a=\gamma \lambda_\text{min}$ and $b=0$, where $\gamma\ge 1$ is a parameter that must be chosen.
+We use $\gamma=1.5$, though the method is relatively insensitive to $\gamma$ and works well for any $\gamma \in [1,2]$.
+Next, we compute $\mathbf{T}$ using fast H-matrix scalar multiplication and addition.
+We form the $\mathbf{T}^{2^k}$ via repeated squaring, $\mathbf{T}^2 = \mathbf{T} \mathbf{T}$, $\mathbf{T}^4 = \mathbf{T}^2 \mathbf{T}^2$, $\mathbf{T}^8 = \mathbf{T}^4 \mathbf{T}^4$ and so on, where each of these matrix squarings is performed using fast H-matrix multiplication.
+We do not need $\Pi_k$ to be an accurate approximation of $\Pi$ since the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ are erroneous, and $\Pi_k(\lambda)$ rapidly decays to zero for positive $\lambda$, even if $k$ is small.
+Hence, we recommend using $k=1$ or $2$ (we use $k=1$ in our numerical results).
+Larger $k$ is unnecessary, and may make the method unstable.
+The approximate spectral projector $\Pi_k(\mathbf{A}_H^{\text{sym}})$ is formed from $\mathbf{T}^{2^k}$ per \eqref{eq:rational343}, using fast H-matrix operations to perform the required matrix-matrix addition and matrix inversion.
+Finally, a positive semi-definite approximation $\mathbf{A}_H^{\text{sym}+}$ of $\mathbf{A}_H^{\text{sym}}$ is formed as
+\begin{equation*}
+\mathbf{A}_H^{\text{sym}+} := \left(\mathbf{I} - 2\Pi(\mathbf{A}_H^{\text{sym}})\right)\mathbf{A}_H^{\text{sym}},
+\end{equation*}
+where fast H-matrix operations are used to perform the necessary matrix-matrix subtraction and multiplication.
+At the end, $\mathbf{A}_H^{\text{sym}+}$ is symmetrized once again, to correct for any small non-symmetries that are introduced due to truncation error in the H-matrix arithmetic.
+Once constructed, we use the symmetric positive semidefinite matrix $\mathbf{A}_H^{\text{sym}+}$ in place of $\mathbf{A}_H$ in further linear algebra operations.
 

### 2022-04-29 numerical_examples/heat/error_vs_krylov_iter.py

@@ -75,6 +75,14 @@ A_pch = A_pch_nonsym.spd(k=1)
 
 # cutoff = -1e-2*eA_max
 
+####
+
+# A_hmatrix = A_pch.sym()
+# k=2
+# N = 2**k
+# beta = np.pow(2, -N)
+# alpha = ()
+
 #########################################
 #
 # A_hmatrix = A_pch_nonsym

### 2022-04-29 numerical_examples/heat/spd_figure.py

@@ -1,27 +1,52 @@
 import numpy as np
 import matplotlib.pyplot as plt
 
-tt = np.linspace(-1, 2, 1000)
-ff1 = tt/(1. + (2.*tt - 1.)**(2**1))
-ff2 = tt/(1. + (2.*tt - 1.)**(2**2))
-ff3 = tt/(1. + (2.*tt - 1.)**(2**3))
-ff4 = tt/(1. + (2.*tt - 1.)**(2**4))
-ff5 = tt/(1. + (2.*tt - 1.)**(2**5))
-ff6 = tt/(1. + (2.*tt - 1.)**(2**6))
-ff7 = tt/(1. + (2.*tt - 1.)**(2**7))
+save=True
+all_kk=[1,2,3]
+a = -1.0
+b = 0.0
 
+linestyles = ['solid', 'dashed', 'dotted', 'dashdot', 'loosely dotted', 'loosely dashed']
+
+all_ff = []
+tt = np.linspace(a - (b-a), b + (b-a), 1000)
+for kk in all_kk:
+    ff = 1. / (1. + ((2. * tt - (b + a))/(b - a)) ** (2 ** kk))
+    all_ff.append(ff)
+
+
+legend = []
 plt.figure(figsize=(8,4))
-plt.plot(tt, ff3, '-.k')
-plt.plot(tt, ff5,'--k')
-plt.plot(tt, ff7, 'k')
-plt.legend(['k=3', 'k=5', 'k=7'])
+for ii in range(len(all_kk)):
+    kk = all_kk[ii]
+    ff = all_ff[ii]
+    plt.plot(tt, ff, c='k', linestyle=linestyles[ii])
+    legend.append('k=' + str(kk))
+
+# for kk, ff in zip(all_kk, all_ff):
+#     plt.plot(tt, ff)
+#     legend.append('k='+str(kk))
+plt.legend(legend)
 plt.xlabel(r'$\lambda$')
-plt.ylabel(r'$\widetilde{f}(\lambda)$')
-plt.title(r'Rational function $\widetilde{f}(\lambda)$')
+plt.ylabel(r'$\Pi_k(\lambda)$')
+plt.title(r'Rational function $\Pi_k(\lambda)$')
 
-plt.savefig('spd_rational_function.pdf', bbox_inches='tight', dpi=100)
+if save:
+    plt.savefig('spd_rational_function.pdf', bbox_inches='tight', dpi=100)
[…]

### 2022-04-29 paper/localpsf.tex
fix typo
edits to positive definite appendix

@@
 		1, & \lambda \in (a,b), \\
-		0, & \text{otherwise},
+		0, & \text{otherwise}.
 	\end{cases}
 \end{equation*}
-and extend $\Pi$ to matrices via functional calculus.
-The matrix $\Pi(\mathbf{A}_H^{\text{sym}})$ is the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to eigenvalues of $\mathbf{A}_H^{\text{sym}}$ that reside in $(a,b)$.
-If we choose $a < \lambda_\text{min}$ and $b=0$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, then the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$ is given by
+Using functional calculus to extend the domain of the function $\Pi$ from scalars to matrices, we define
+\begin{equation*}
+\mathbf{\Pi} := \Pi(\mathbf{A}_H^{\text{sym}}).
+\end{equation*}
+The matrix $\mathbf{\Pi}$ is the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to eigenvalues of $\mathbf{A}_H^{\text{sym}}$ that reside in $(a,b)$.
+If we choose $a < \lambda_\text{min}$ and $b=0$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, then $\mathbf{\Pi}$ is the spectral projector onto the eigenspace associated with negative eigenvalues, and $\mathbf{\Pi}\mathbf{A}_H^{\text{sym}}$ is the ``negative component'' of $\mathbf{A}_H^{\text{sym}}$ in which all the negative eigenvalues are the same but all the positive eigenvalues are replaced by zero.
+Subtracting a negative number from itself twice flips its sign, so the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$ is given by
 \begin{equation}
 \label{eq:I_Pi_abs}
-	|\mathbf{A}_H^{\text{sym}}| = \left(\mathbf{I} - 2\Pi(\mathbf{A}_H^{\text{sym}})\right)\mathbf{A}_H^{\text{sym}}.
-\end{equation}
-That is, $|\mathbf{A}_H^{\text{sym}}|$ is the matrix that has the same eigenvectors as $\mathbf{A}_H^{\text{sym}}$, but the eigenvalues of $|\mathbf{A}_H^{\text{sym}}|$ are the absolute values of the eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
-
-Computing $\Pi(\mathbf{A}_H^{\text{sym}})$ directly is difficult, but we may approximate $\Pi$ by the following high order rational function:
-\begin{equation*}
-	\Pi(\lambda) \approx \Pi_k(\lambda) := \frac{1}{1 + T(\lambda)^{2^k} }, \qquad \text{where} \qquad T(\lambda) := \frac{2 \lambda - (b+a)}{b-a}
-\end{equation*}
-and $k$ is a positive integer; the larger $k$ is, the better the approximation $\Pi \approx \Pi_k$.
-This is illustrated in Figure \ref{fig:spd_rat}.
-For $k \ge 1$, $\Pi_k(\lambda)> 1/2$ for $\lambda \in (a,b)$, so choosing $a < \lambda_\text{min}$, $b=0$, and using $\Pi_k$ instead of $\Pi$ in \eqref{eq:I_Pi_abs} yields a matrix that is positive semidefinite.
-Replacing constants with scaled identity matrices, scalar powers and multiplication with matrix powers and multiplication, and scalar division with matrix inversion, the extension of $\Pi_k$ to matrices becomes
-\begin{equation}
-\label{eq:rational343}
-\Pi_k\left(\mathbf{A}_H^{\text{sym}}\right) = \left(\mathbf{I} + \mathbf{T}^{2^k}\right)^{-1},
-\end{equation}
-where
-\begin{equation*}
-\mathbf{T} := \left(2 \mathbf{A}_H^{\text{sym}} - (b+a) \mathbf{I}\right)/(b-a).
-\end{equation*}
+	|\mathbf{A}_H^{\text{sym}}| = \left(\mathbf{I} - 2\mathbf{\Pi}\right)\mathbf{A}_H^{\text{sym}},
+\end{equation}
+where $\mathbf{I}$ is the identity matrix of the appropriate size.
+By ``matrix absolute value,'' we mean that $|\mathbf{A}_H^{\text{sym}}|$ is the matrix that has the same eigenvectors as $\mathbf{A}_H^{\text{sym}}$, but the eigenvalues of $|\mathbf{A}_H^{\text{sym}}|$ are the absolute values of the eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
 
@@
 
-Using $\Pi_k$ instead of $\Pi$ permits us to efficiently approximate $|\mathbf{A}_H^{\text{sym}}|$ using H-matrix methods. 
+Computing $\mathbf{\Pi}=\Pi(\mathbf{A}_H^{\text{sym}})$ directly is difficult, but we may approximate $\Pi$ by the following high order rational function:
+\begin{equation*}
+	\Pi(\lambda) \approx \Pi_k(\lambda) := \frac{1}{1 + T(\lambda)^{2^k} }, \qquad \text{where} \qquad T(\lambda) := \frac{2 \lambda - (b+a)}{b-a},
+\end{equation*}
+and $k$ is a positive integer; the larger $k$ is, the better the approximation $\Pi \approx \Pi_k$.
+This is illustrated in Figure \ref{fig:spd_rat}.
+Replacing scalar operations with matrix operations, we have
+\begin{equation}
+\label{eq:rational343}
+\mathbf{\Pi}_k := \Pi_k\left(\mathbf{A}_H^{\text{sym}}\right) = \left(\mathbf{I} + \mathbf{T}^{2^k}\right)^{-1},
+\end{equation}
+where
+\begin{equation*}
+\mathbf{T} := \left(2 \mathbf{A}_H^{\text{sym}} - (b+a) \mathbf{I}\right)/(b-a).
+\end{equation*}
+Using $\mathbf{\Pi}_k$ instead of $\mathbf{\Pi}$ in \eqref{eq:I_Pi_abs} allows us to efficiently approximate $|\mathbf{A}_H^{\text{sym}}|$ using H-matrix methods. 
 First, we compute $\lambda_\text{min}$ using the Lanczos method, and set $a=\gamma \lambda_\text{min}$ and $b=0$, where $\gamma\ge 1$ is a parameter that must be chosen.
@@
 We form the $\mathbf{T}^{2^k}$ via repeated squaring, $\mathbf{T}^2 = \mathbf{T} \mathbf{T}$, $\mathbf{T}^4 = \mathbf{T}^2 \mathbf{T}^2$, $\mathbf{T}^8 = \mathbf{T}^4 \mathbf{T}^4$ and so on, where each of these matrix squarings is performed using fast H-matrix multiplication.
-We do not need $\Pi_k$ to be an accurate approximation of $\Pi$ since the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ are erroneous, and $\Pi_k(\lambda)$ rapidly decays to zero for positive $\lambda$, even if $k$ is small.
-Hence, we recommend using $k=1$ or $2$ (we use $k=1$ in our numerical results).
-Larger $k$ is unnecessary, and may make the method unstable.
-The approximate spectral projector $\Pi_k(\mathbf{A}_H^{\text{sym}})$ is formed from $\mathbf{T}^{2^k}$ per \eqref{eq:rational343}, using fast H-matrix operations to perform the required matrix-matrix addition and matrix inversion.
+We recommend using $k=1$ or $2$ (we use $k=1$ in our numerical results).
+Larger $k$ is unnecessary and may make the method unstable.
+The approximate spectral projector $\mathbf{\Pi}_k$ is formed from $\mathbf{T}^{2^k}$ per \eqref{eq:rational343}, using fast H-matrix operations to perform the required matrix-matrix addition and matrix inversion.
 Finally, a positive semi-definite approximation $\mathbf{A}_H^{\text{sym}+}$ of $\mathbf{A}_H^{\text{sym}}$ is formed as
 \begin{equation*}
-\mathbf{A}_H^{\text{sym}+} := \left(\mathbf{I} - 2\Pi(\mathbf{A}_H^{\text{sym}})\right)\mathbf{A}_H^{\text{sym}},
+\mathbf{A}_H^{\text{sym}+} := \left(\mathbf{I} - 2\mathbf{\Pi}_k)\right)\mathbf{A}_H^{\text{sym}},
 \end{equation*}
 where fast H-matrix operations are used to perform the necessary matrix-matrix subtraction and multiplication.
-At the end, $\mathbf{A}_H^{\text{sym}+}$ is symmetrized once again, to correct for any small non-symmetries that are introduced due to truncation error in the H-matrix arithmetic.
+Note that for any $k \ge 1$, $\Pi_k(\lambda)> 1/2$ for $\lambda \in (a,b)$, so $\mathbf{A}_H^{\text{sym}+}$ is positive semi definite\footnote{Some indefiniteness may be introduced by truncation error in the H-matrix operations, but the negative eigenvalues caused by this truncation error are typically extremely small.}.
+At the end, $\mathbf{A}_H^{\text{sym}+}$ is symmetrized once again, to correct for any small non-symmetries that are introduced by truncation error in the H-matrix operations.
 Once constructed, we use the symmetric positive semidefinite matrix $\mathbf{A}_H^{\text{sym}+}$ in place of $\mathbf{A}_H$ in further linear algebra operations.

### 2022-05-02 paper/localpsf.tex
minor edits

@@
 These H-matrix methods are fast and scale well to large problems.
-Typically, the work and memory required to perform these operations for an $N \times N$ H-matrix scales as $O(r^2 N \log(N)^a)$, where $r$ is the largest rank of the off-diagonal blocks, and $a \in \{0,1,2,3\}$.
-The exact cost depends on the type of H-matrix used and operation performed; for more details see \cite{GrasedyckHackbusch03}.
+Typically, the work and memory required to perform these operations for an $N \times N$ H-matrix scales as $O(N \log(N)^a)$, where $a \in \{0,1,2,3\}$.
+The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks; for more details see \cite{GrasedyckHackbusch03}.
 The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
@@
 
-Computing $\mathbf{\Pi}=\Pi(\mathbf{A}_H^{\text{sym}})$ directly is difficult, but we may approximate $\Pi$ by the following high order rational function:
-\begin{equation*}
-	\Pi(\lambda) \approx \Pi_k(\lambda) := \frac{1}{1 + T(\lambda)^{2^k} }, \qquad \text{where} \qquad T(\lambda) := \frac{2 \lambda - (b+a)}{b-a},
-\end{equation*}
-and $k$ is a positive integer; the larger $k$ is, the better the approximation $\Pi \approx \Pi_k$.
+Computing $\mathbf{\Pi}=\Pi(\mathbf{A}_H^{\text{sym}})$ directly is difficult, but we may approximate $\Pi$ by a high order rational function $\Pi_k \approx \Pi(\lambda)$ which is defined as follows:
+\begin{equation*}
+	 \Pi_k(\lambda) := \frac{1}{1 + T(\lambda)^{2^k} }, \qquad \text{where} \qquad T(\lambda) := \frac{2 \lambda - (b+a)}{b-a}.
+\end{equation*}
+Here $k$ is a positive integer; the larger $k$ is, the better the approximation $\Pi \approx \Pi_k$.
 This is illustrated in Figure \ref{fig:spd_rat}.
-Replacing scalar operations with matrix operations, we have
+Replacing scalar operations with matrix operations, we have an approximation $\mathbf{\Pi}_k \approx \mathbf{\Pi}$ given by
 \begin{equation}

### 2022-05-02 paper/localpsf.tex
minor edits

@@
 \end{equation*}
-If $\Aop$ is also positive semi-definite, we use the high order rational function spectral projection method from \cite{HackbuschKress07} to flip (or approximately flip) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ to be positive.
+If $\Aop$ is also positive semi-definite, we adapt the high order rational function spectral projection method from \cite{HackbuschKress07} to flip (or approximately flip) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ to be positive.
 In detail, let $\Pi$ be the indicator function for the interval $(a,b)$,
@@
 \begin{equation*}
-\mathbf{A}_H^{\text{sym}+} := \left(\mathbf{I} - 2\mathbf{\Pi}_k)\right)\mathbf{A}_H^{\text{sym}},
+\mathbf{A}_H^{\text{sym}+} := \left(\mathbf{I} - 2\mathbf{\Pi}_k\right)\mathbf{A}_H^{\text{sym}},
 \end{equation*}

### 2022-05-03 paper/localpsf.tex
minor edits

@@
 
-Let $x$ and $x'$ be points in $\Omega$ that are ``close'' to each other, and consider the following four approximations, ordered by increasing level of sophistication:
+Let $x$ and $x'$ be points in $\Omega$ that are ``close'' to each other, and consider the following approximations:
 \begin{align}
@@
 \end{align}
+These are four different ways to approximate an impulse response by a nearby impulse response, with each successive approximation building upon the previous ones.
+Our method uses \eqref{eq:translate4}, which is the most sophisticated of these approximations.
+
 Approximation \eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close.

### 2022-05-10 paper/localpsf.tex
minor edits
wrote about recycling krylov information

@@
 Approximation \eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close.
-Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank interpolative factorization/CUR approximation \cite{CUR,NYSTROM}.
+Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank CUR approximation \cite{CUR,NYSTROM}.
 However, the required rank in the low rank approximation can be large, which makes algorithms based on \eqref{eq:translate1} expensive.
@@
 
-\section{Fast ellipsoid intersection test}
+\section{Ellipsoid intersection test}
 \label{sec:fast_ellipsoid_intersection_test}
@@
 
+\section{Recycling Krylov information via low rank updates}
+\label{sec:recycle_krylov}
+
+In applications, often one wants to use the H-matrix approximation $\mathbf{A}_H$ to build preconditioners for solving a sequence of slowly varying linear systems of the form
+\begin{equation}
+\label{eq:Ax_equals_b}
+\mathbf{B} \mathbf{u} = \mathbf{f},
+\end{equation}
+where each of these linear systems is solved using an iterative method.
+If $\mathbf{B}=\mathbf{A}$, then $\mathbf{A}_H^{-1}$ may be used to precondition this linear system.
+If $\mathbf{B}$ is built from a combination of $\mathbf{A}$ and other operators, then $\mathbf{A}_H$ may be combined with H-matrix approximations of the other operators to construct a preconditioner for $\mathbf{B}$.
+For example, in a Newton-Krylov method for solving an optimization problem, one uses a Krylov method such as conjugate gradient to solve a sequence of systems of the form \eqref{eq:Ax_equals_b}, where $\mathbf{B}$ is the Hessian of the objective function in the optimization problem, and $\mathbf{f}$ is the negative gradient of the objective function.
+Both the Hessian and the gradient are evaluated at the current Newton iterate, so $\mathbf{B}$ and $\mathbf{f}$ change as the Newton iterations proceed.
+The Hessian is typically the sum of a data-misfit term which is available matrix-free, and a regularization term which is available directly as a sparse matrix.
+The data misfit term in the Hessian is approximated with our method to form $\mathbf{A}_H$, the regularization term is approximated using standard H-matrix methods for sparse matrices, and then these two terms are added together and inverted using H-matrix methods to form a preconditioner for the overall Hessian.
+
+Let 
+\begin{equation*}
+\mathbf{P} \approx \mathbf{B}^{-1}
+\end{equation*}
+denote the H-matrix preconditioner used to accelerate the solution of \eqref{eq:Ax_equals_b} with an iterative method.
+The iterative method will typically involve, as a sub-procedure, applying $\mathbf{B}$ to a collection of vectors $\{\mathbf{x}_i\}_{i=1}^m$, thereby generating vectors $\{\mathbf{y}_i = \mathbf{B} \mathbf{x}_i\}_{i=1}^m$.
+Instead of rebuilding the H-matrix preconditioner for each successive linear system, we recommend \emph{recycling} the information generated by the iterative method while solving the current linear system, in order to update the preconditioner $P$ for the next linear system.
+In detail, let $\mathbf{X}$ and $\mathbf{Y}$ denote the $N \times m$ matrices with columns given by the vectors $\mathbf{x}_i$ and $\mathbf{y}_i$, respectively, so that the information learned about $\mathbf{B}$ while solving the current linear system may be summarized by the equation:
+\begin{equation*}
+\mathbf{B}\mathbf{X} = \mathbf{Y}.
+\end{equation*}
+For convenience, define 
+\begin{equation*}
+\mathbf{R} := \mathbf{X}-\mathbf{P}\mathbf{Y}, \qquad \text{and} \qquad \boldsymbol{\Theta} := \left(\mathbf{X}^T\mathbf{Y}\right)^{-1}.
+\end{equation*}
+One of the following three formulas may be used to perform a low rank update of $\mathbf{P}$:
+\begin{align}
+	\mathbf{P} &\gets \mathbf{P} + \mathbf{R}\left(\mathbf{Y}^T \mathbf{Y}\right)^{-1} \mathbf{Y}^T \label{eq:res_update} \\
+	\mathbf{P} &\gets \mathbf{P} + \mathbf{R} \left(\mathbf{R}^T \mathbf{Y}\right)^{-1} \mathbf{R}^T \label{eq:srk_update} \\
+	\mathbf{P} &\gets \left(\mathbf{I} - \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T\right)\mathbf{P}\left(\mathbf{I} - \mathbf{Y}\boldsymbol{\Theta}\mathbf{X}^T\right) + \mathbf{X}\boldsymbol{\Theta}\mathbf{X}^T \label{eq:dfp_update}
+\end{align}
+In all three cases, direct calculation shows that we have $\mathbf{P}\mathbf{Y}=\mathbf{X}$ after the update, which implies $\mathbf{P}^{-1} \mathbf{X} = \mathbf{B}\mathbf{X}$.
+Formulas \eqref{eq:res_update} and \eqref{eq:srk_update} perform rank-$k$ updates of $\mathbf{P}$, while formula \eqref{eq:dfp_update} performs a rank-$2k$ update.
+Formula \eqref{eq:res_update} does not preserve symmetry, and should therefore be used when $\mathbf{B}$ is not symmetric.
+Formula \eqref{eq:srk_update} preserves symmetry but not positive definiteness, and should therefore be used when $\mathbf{B}$ is symmetric and indefinite.
+Formula \eqref{eq:dfp_update} preserves both symmetry and positive definiteness, and should therefore be used when $\mathbf{B}$ is symmetric positive definite.
+Formula \eqref{eq:srk_update} is known as the ``symmetric rank-$k$'' update formula.
+Formula \eqref{eq:dfp_update} is known as the DFP update formula, named after it's inventor Davidson, and polularizers Fletcher and Powell.
+For more details on these update formulas, we recommend the discussion in Chapter 6 of \cite{NocedalWright99}.
+In \cite{NocedalWright99}, one-vector versions of formulas \eqref{eq:srk_update} and \eqref{eq:dfp_update} are discussed thoroughly; the versions here are straightforward generalizations to many-vector updates.
+
+%to update $\mathbf{B}^{(k)} \approx \left(\mathbf{A}^{(k)}\right)^{-1}$ denote the H-matrix preconditioner used for the $k$th linear system, and set $\mathbf{B}^{(1)}:= \mathbf{A}_H^{-1}$ denote the H-matrix approximation used to precondition the first linear system.
+Rather than re-build the H-matrix preconditioner for each subsequent linear system, one may instead form a preconditioner, $\mathbf{B}^{(k+1)}$ update, by \emph{recycling} the information generated by the iterative method used to solve the current system.
+%
+%In practice it is often necessary to solve a sequence of linear systems of the form 
+%\begin{equation*}
+%	\mathbf{A} \mathbf{x} = \mathbf{b}
+%\end{equation*}
+%where either $\mathbf{A}$, or $\mathbf{b}$, or both, are slowly varying. 
+%
+%\begin{equation*}
+%	\mathbf{B}\mathbf{X} = \mathbf{Y}
+%\end{equation*}
+%
+%
+%
+%\begin{equation*}
+%\mathbf{Q} := \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T
+%\end{equation*}
+%
+%\begin{equation}
+%	\mathbf{B}_{k+1} = \left(\mathbf{I} - \mathbf{Y}\boldsymbol{\Theta}\mathbf{X}^T\right)\mathbf{B}_k\left(\mathbf{I} - \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T\right) + \mathbf{Y}\boldsymbol{\Theta}\mathbf{Y}^T
+%\end{equation}
+
 \section*{Acknowledgments}

### 2022-05-10 paper/localpsf.tex
fix typo

@@
 The data misfit term in the Hessian is approximated with our method to form $\mathbf{A}_H$, the regularization term is approximated using standard H-matrix methods for sparse matrices, and then these two terms are added together and inverted using H-matrix methods to form a preconditioner for the overall Hessian.
-
 Let
@@
 The iterative method will typically involve, as a sub-procedure, applying $\mathbf{B}$ to a collection of vectors $\{\mathbf{x}_i\}_{i=1}^m$, thereby generating vectors $\{\mathbf{y}_i = \mathbf{B} \mathbf{x}_i\}_{i=1}^m$.
-Instead of rebuilding the H-matrix preconditioner for each successive linear system, we recommend \emph{recycling} the information generated by the iterative method while solving the current linear system, in order to update the preconditioner $P$ for the next linear system.
+Instead of rebuilding the H-matrix preconditioner for each successive linear system, we recommend \emph{recycling} the information generated by the iterative method while solving the current linear system, in order to update the preconditioner $\mathbf{P}$ for the next linear system.
 In detail, let $\mathbf{X}$ and $\mathbf{Y}$ denote the $N \times m$ matrices with columns given by the vectors $\mathbf{x}_i$ and $\mathbf{y}_i$, respectively, so that the information learned about $\mathbf{B}$ while solving the current linear system may be summarized by the equation:

### 2022-05-10 paper/localpsf.tex
update low rank update appendix based on Tucker's comments

--- 2022-05-03 paper/localpsf.tex
@@
 
+\section{Recycling Krylov information via low rank updates}
+\label{sec:recycle_krylov}
+
+In applications, often one wants to use the H-matrix approximation $\mathbf{A}_H$ to build preconditioners for solving a sequence of slowly varying linear systems of the form
+\begin{equation}
+\label{eq:Ax_equals_b}
+\mathbf{B}^{(k)} \mathbf{u}^{(k)} = \mathbf{f}^{(k)},
+\end{equation}
+where each of these linear systems is solved using an iterative method.
+By ``slowly varying,'' we mean that $\mathbf{B}^{(k+1)}$ does not differ much from $\mathbf{B}^{(k)}$.
+If $\mathbf{B}^{(k)}=\mathbf{A}$, then our H-matrix approximation $\mathbf{A}_H^{-1}$ may be used to precondition this linear system.
+If $\mathbf{B}^{(k)}$ is built from a combination of $\mathbf{A}$ and other operators, then $\mathbf{A}_H$ may be combined with H-matrix approximations of the other operators to construct a preconditioner for $\mathbf{B}^{(k)}$.
+For example, in a Newton-Krylov method for solving an optimization problem, one uses a Krylov method such as conjugate gradient to solve a sequence of systems of the form \eqref{eq:Ax_equals_b}, where $\mathbf{B}^{(k)}$ is the Hessian of the objective function in the optimization problem, and $\mathbf{f}^{(k)}$ is the negative gradient of the objective function.
+The Hessian and the gradient are evaluated at the $k$th Newton iterate, so both $\mathbf{B}^{(k)}$ and $\mathbf{f}^{(k)}$ change as the Newton iterations proceed.
+The Hessian is typically the sum of a data misfit term which is only available matrix-free, and a regularization term which is available directly as a sparse matrix.
+The data misfit term in the Hessian is approximated with our method to form $\mathbf{A}_H$, the regularization term is approximated using standard H-matrix methods for sparse matrices, and then these two terms are added together and inverted using H-matrix methods to form a preconditioner for the overall Hessian.
+Let
+\begin{equation*}
+\mathbf{P}^{(k)} \approx \left(\mathbf{B}^{(k)}\right)^{-1}
+\end{equation*}
+denote the H-matrix preconditioner used to accelerate the solution of \eqref{eq:Ax_equals_b} with an iterative method.
+The iterative method for solving the $k$th linear system will typically involve, as a sub-procedure, applying $\mathbf{B}^{(k)}$ to a collection of vectors $\{\mathbf{x}_i\}_{i=1}^m$, thereby generating vectors $\{\mathbf{y}_i = \mathbf{B}^{(k)} \mathbf{x}_i\}_{i=1}^m$.
+Instead of rebuilding the H-matrix preconditioner for each successive linear system, we recommend \emph{recycling} the information generated by the iterative method while solving the current linear system, performing a low-rank update to $\mathbf{P}^{(k)}$ to create $\mathbf{P}^{(k+1)}$.
+In detail, let $\mathbf{X}$ and $\mathbf{Y}$ denote the $N \times m$ matrices with columns given by the vectors $\mathbf{x}_i$ and $\mathbf{y}_i$, respectively, so that the information learned about $\mathbf{B}^{(k)}$ while solving the current linear system may be summarized by the equation:
+\begin{equation*}
+\mathbf{B}^{(k)}\mathbf{X} = \mathbf{Y}.
+\end{equation*}
+For convenience, define 
+\begin{equation*}
+\mathbf{R} := \mathbf{X}-\mathbf{P}^{(k)}\mathbf{Y}, \qquad \text{and} \qquad \boldsymbol{\Theta} := \left(\mathbf{X}^T\mathbf{Y}\right)^{-1}.
+\end{equation*}
+One of the following three formulas may be used to form $\mathbf{P}^{(k+1)}$ from $\mathbf{P}^{(k)}$:
+\begin{align}
+	\mathbf{P}^{(k+1)} &= \mathbf{P}^{(k)} + \mathbf{R}\left(\mathbf{Y}^T \mathbf{Y}\right)^{-1} \mathbf{Y}^T \label{eq:res_update} \\
+	\mathbf{P}^{(k+1)} &= \mathbf{P}^{(k)} + \mathbf{R} \left(\mathbf{R}^T \mathbf{Y}\right)^{-1} \mathbf{R}^T \label{eq:srk_update} \\
+	\mathbf{P}^{(k+1)} &= \left(\mathbf{I} - \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T\right)\mathbf{P}^{(k)}\left(\mathbf{I} - \mathbf{Y}\boldsymbol{\Theta}\mathbf{X}^T\right) + \mathbf{X}\boldsymbol{\Theta}\mathbf{X}^T \label{eq:dfp_update}
+\end{align}
+In all three cases, direct calculation shows that we have $\mathbf{P}^{(k)}\mathbf{Y}=\mathbf{X}$ after the update, which implies $\left(\mathbf{P}^{(k)}\right)^{-1} \mathbf{X} = \mathbf{B}^{(k)}\mathbf{X}$.
+Formulas \eqref{eq:res_update} and \eqref{eq:srk_update} perform rank-$k$ updates of $\mathbf{P}^{(k)}$, while formula \eqref{eq:dfp_update} performs a rank-$2k$ update.
+Formula \eqref{eq:res_update} does not preserve symmetry, and should therefore be used when $\mathbf{B}^{(k)}$ is not symmetric.
+Formula \eqref{eq:srk_update} preserves symmetry but not positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric and indefinite.
+Formula \eqref{eq:dfp_update} preserves both symmetry and positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric positive definite.
+Formula \eqref{eq:srk_update} is known as the ``symmetric rank-$k$'' update formula.
+Formula \eqref{eq:dfp_update} is known as the DFP update formula, named after it's inventor Davidson, and polularizers Fletcher and Powell.
+For more details on these update formulas, we recommend the discussion in Chapter 6 of \cite{NocedalWright99}.
+In \cite{NocedalWright99}, one-vector versions of formulas \eqref{eq:srk_update} and \eqref{eq:dfp_update} are discussed thoroughly; the versions here are straightforward generalizations to many-vector updates.
+
+%to update $\mathbf{B}^{(k)} \approx \left(\mathbf{A}^{(k)}\right)^{-1}$ denote the H-matrix preconditioner used for the $k$th linear system, and set $\mathbf{B}^{(1)}:= \mathbf{A}_H^{-1}$ denote the H-matrix approximation used to precondition the first linear system.
+Rather than re-build the H-matrix preconditioner for each subsequent linear system, one may instead form a preconditioner, $\mathbf{B}^{(k+1)}$ update, by \emph{recycling} the information generated by the iterative method used to solve the current system.
+%
+%In practice it is often necessary to solve a sequence of linear systems of the form 
+%\begin{equation*}
+%	\mathbf{A} \mathbf{x} = \mathbf{b}
+%\end{equation*}
+%where either $\mathbf{A}$, or $\mathbf{b}$, or both, are slowly varying. 
+%
+%\begin{equation*}
+%	\mathbf{B}\mathbf{X} = \mathbf{Y}
+%\end{equation*}
+%
+%
+%
+%\begin{equation*}
+%\mathbf{Q} := \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T
+%\end{equation*}
+%
+%\begin{equation}
+%	\mathbf{B}_{k+1} = \left(\mathbf{I} - \mathbf{Y}\boldsymbol{\Theta}\mathbf{X}^T\right)\mathbf{B}_k\left(\mathbf{I} - \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T\right) + \mathbf{Y}\boldsymbol{\Theta}\mathbf{Y}^T
+%\end{equation}
+
 \section*{Acknowledgments}

### 2022-05-11 paper/localpsf.tex
fix typos, minor wordsmithing, in low rank update appendix

@@
 \end{align}
-In all three cases, direct calculation shows that we have $\mathbf{P}^{(k)}\mathbf{Y}=\mathbf{X}$ after the update, which implies $\left(\mathbf{P}^{(k)}\right)^{-1} \mathbf{X} = \mathbf{B}^{(k)}\mathbf{X}$.
-Formulas \eqref{eq:res_update} and \eqref{eq:srk_update} perform rank-$k$ updates of $\mathbf{P}^{(k)}$, while formula \eqref{eq:dfp_update} performs a rank-$2k$ update.
+In all three cases, direct calculation shows that we have $\mathbf{P}^{(k+1)}\mathbf{Y}=\mathbf{X}$, which implies $\left(\mathbf{P}^{(k+1)}\right)^{-1} \mathbf{X} = \mathbf{B}^{(k)}\mathbf{X}$.
+Formulas \eqref{eq:res_update} and \eqref{eq:srk_update} perform rank-$m$ updates, while formula \eqref{eq:dfp_update} performs a rank-$2m$ update.
 Formula \eqref{eq:res_update} does not preserve symmetry, and should therefore be used when $\mathbf{B}^{(k)}$ is not symmetric.
@@
 Formula \eqref{eq:dfp_update} preserves both symmetry and positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric positive definite.
-Formula \eqref{eq:srk_update} is known as the ``symmetric rank-$k$'' update formula.
+Formula \eqref{eq:srk_update} is known as the ``symmetric rank-$m$'' update formula.
 Formula \eqref{eq:dfp_update} is known as the DFP update formula, named after it's inventor Davidson, and polularizers Fletcher and Powell.

### 2022-05-11 numerical_examples/stokes/stokes_inverse_problem_cylinder.py

@@ -13,11 +13,20 @@ from nalger_helper_functions import *
 from localpsf.product_convolution_kernel import ProductConvolutionKernel
 from localpsf.product_convolution_hmatrix import make_hmatrix_from_kernel, product_convolution_hmatrix
 
+# For Tucker's computer:
+# import sys
+# sys.path.append('/home/tucker/software/ucm-ice/')
+# sys.path.append('/home/tucker/software/hippylib/')
+# End For Tucker
+
+# For Nick's computer:
+import meshio
 import sys
-sys.path.append('/home/tucker/software/ucm-ice/')
-sys.path.append('/home/tucker/software/hippylib/')
-
-
+sys.path.append('../../../hippylib/')
+import hippylib as hp
+sys.path.append('../../../ucm-ice/')
+# from iceModel import *
+# End For Nick
 
 import hippylib as hp
 from iceModel import *

### 2022-05-12 paper/localpsf.tex
fix typos, minor wordsmithing, in low rank update appendix

@@
 \begin{align*}
-	E_1 :=& \{x : (x - \spatialmean_1)^T \spatialcov_1^{-1} (x - \spatialmean_1) \le \tau^2\} \\
+	E_1 :=& \{x : (x - \spatialmean_1)^T \spatialcov_1^{-1} (x - \spatialmean_1) \le \tau^2\}, \\
 	E_2 :=& \{x : (x - \spatialmean_2)^T \spatialcov_2^{-1} (x - \spatialmean_2) \le \tau^2\}, \\
@@
 \end{align}
-In all three cases, direct calculation shows that we have $\mathbf{P}^{(k+1)}\mathbf{Y}=\mathbf{X}$, which implies $\left(\mathbf{P}^{(k+1)}\right)^{-1} \mathbf{X} = \mathbf{B}^{(k)}\mathbf{X}$.
+In all three cases, direct calculation shows that we have $\mathbf{P}^{(k+1)}\mathbf{Y}=\mathbf{X} = \left(\mathbf{B}^{(k)}\right)^{-1} \mathbf{Y}$.
 Formulas \eqref{eq:res_update} and \eqref{eq:srk_update} perform rank-$m$ updates, while formula \eqref{eq:dfp_update} performs a rank-$2m$ update.
@@
 Formula \eqref{eq:dfp_update} preserves both symmetry and positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric positive definite.
-Formula \eqref{eq:srk_update} is known as the ``symmetric rank-$m$'' update formula.
-Formula \eqref{eq:dfp_update} is known as the DFP update formula, named after it's inventor Davidson, and polularizers Fletcher and Powell.
+When $m=1$, formula \eqref{eq:srk_update} is known as the SR1 (``symmetric rank-$1$'') update formula, and formula \eqref{eq:dfp_update} is known as the DFP update formula, which is named after it's inventor Davidson, and popularizers Fletcher and Powell.
 For more details on these update formulas, we recommend the discussion in Chapter 6 of \cite{NocedalWright99}.

### 2022-05-16 paper/localpsf.tex
update low rank update appendix based on discussion with Noemi

@@
 
-In applications, often one wants to use the H-matrix approximation $\mathbf{A}_H$ to build preconditioners for solving a sequence of slowly varying linear systems of the form
+REFER TO BIROS PAPER
+
+In applications, often one wants to use the H-matrix approximation $\mathbf{A}_H$ to build preconditioners for solving a sequence of slowly varying linear systems, where the $k$th linear system has the form
 \begin{equation}
 \label{eq:Ax_equals_b}
-\mathbf{B}^{(k)} \mathbf{u}^{(k)} = \mathbf{f}^{(k)},
-\end{equation}
-where each of these linear systems is solved using an iterative method.
-By ``slowly varying,'' we mean that $\mathbf{B}^{(k+1)}$ does not differ much from $\mathbf{B}^{(k)}$.
-If $\mathbf{B}^{(k)}=\mathbf{A}$, then our H-matrix approximation $\mathbf{A}_H^{-1}$ may be used to precondition this linear system.
-If $\mathbf{B}^{(k)}$ is built from a combination of $\mathbf{A}$ and other operators, then $\mathbf{A}_H$ may be combined with H-matrix approximations of the other operators to construct a preconditioner for $\mathbf{B}^{(k)}$.
-For example, in a Newton-Krylov method for solving an optimization problem, one uses a Krylov method such as conjugate gradient to solve a sequence of systems of the form \eqref{eq:Ax_equals_b}, where $\mathbf{B}^{(k)}$ is the Hessian of the objective function in the optimization problem, and $\mathbf{f}^{(k)}$ is the negative gradient of the objective function.
-The Hessian and the gradient are evaluated at the $k$th Newton iterate, so both $\mathbf{B}^{(k)}$ and $\mathbf{f}^{(k)}$ change as the Newton iterations proceed.
-The Hessian is typically the sum of a data misfit term which is only available matrix-free, and a regularization term which is available directly as a sparse matrix.
-The data misfit term in the Hessian is approximated with our method to form $\mathbf{A}_H$, the regularization term is approximated using standard H-matrix methods for sparse matrices, and then these two terms are added together and inverted using H-matrix methods to form a preconditioner for the overall Hessian.
-Let
+\mathbf{B}^{(k)} \mathbf{u}^{(k)} = \mathbf{f}^{(k)}.
+\end{equation}
+Each of these linear systems is solved using an iterative method. By ``slowly varying,'' we mean that $\mathbf{B}^{(k+1)}$ does not differ much from $\mathbf{B}^{(k)}$. 
+%If $\mathbf{B}^{(k)}=\mathbf{A}$, then our H-matrix approximation $\mathbf{A}_H^{-1}$ may be used to precondition this linear system. 
+%In our target applications, $\mathbf{B}^{(k)}$ is built from a combination of $\mathbf{A}$ and other operators, then $\mathbf{A}_H$ may be combined with H-matrix approximations of the other operators to construct a preconditioner for $\mathbf{B}^{(k)}$.
+For example, in a Newton-Krylov method for solving an optimization problem, one uses a Krylov method such as conjugate gradient to solve a sequence of systems of the form \eqref{eq:Ax_equals_b}, where $\mathbf{B}^{(k)}$ and $\mathbf{f}^{(k)}$ are the Hessian and the negative gradient of the objective function in the optimization problem, respectively, evaluated at the $k$th Newton iterate.
+%The Hessian and the gradient are evaluated at the $k$th Newton iterate, hence both $\mathbf{B}^{(k)}$ and $\mathbf{f}^{(k)}$ change as the Newton iterations proceed. 
+As the optimization solver converges, $\mathbf{B}^{(k)}$ converges to the Hessian at the optimal point, and therefore changes little from one iteration to the next.
+Since $\mathbf{B}^{(k)}$ varies slowly, it is reasonable to assume that Krylov information about $\mathbf{B}^{(k)}$ can be leveraged to update the preconditioner from $\mathbf{B}^{(k)}$ in order to build a preconditioner for $\mathbf{B}^{(k+1)}$.
+%In the context of an inverse problem, the Hessian is typically the sum of two terms: the Hessian of the data misfit term, which is only available matrix-free, and the Hessian of a regularization term, which is available directly as a sparse matrix.
+The Hessian of the data misfit term is approximated with our method to form $\mathbf{A}_H$, the Hessian of the regularization term is approximated using standard H-matrix methods for sparse matrices, and then these two terms are added together and inverted using H-matrix methods to form a preconditioner for the overall Hessian.
+Concretely, let 
 \begin{equation*}
@@
 \end{equation*}
-denote the H-matrix preconditioner used to accelerate the solution of \eqref{eq:Ax_equals_b} with an iterative method.
-The iterative method for solving the $k$th linear system will typically involve, as a sub-procedure, applying $\mathbf{B}^{(k)}$ to a collection of vectors $\{\mathbf{x}_i\}_{i=1}^m$, thereby generating vectors $\{\mathbf{y}_i = \mathbf{B}^{(k)} \mathbf{x}_i\}_{i=1}^m$.
-Instead of rebuilding the H-matrix preconditioner for each successive linear system, we recommend \emph{recycling} the information generated by the iterative method while solving the current linear system, performing a low-rank update to $\mathbf{P}^{(k)}$ to create $\mathbf{P}^{(k+1)}$.
-In detail, let $\mathbf{X}$ and $\mathbf{Y}$ denote the $N \times m$ matrices with columns given by the vectors $\mathbf{x}_i$ and $\mathbf{y}_i$, respectively, so that the information learned about $\mathbf{B}^{(k)}$ while solving the current linear system may be summarized by the equation:
+denote a H-matrix preconditioner used to accelerate the solution of \eqref{eq:Ax_equals_b} with an iterative method.
+The iterative method for solving the $k$th linear system will typically involve, as a sub-problem, applying $\mathbf{B}^{(k)}$ to a collection of vectors $\{\mathbf{x}_i\}_{i=1}^m$, thereby generating vectors $\{\mathbf{y}_i = \mathbf{B}^{(k)} \mathbf{x}_i\}_{i=1}^m$.
+Instead of rebuilding the H-matrix preconditioner for each successive linear system, we \emph{recycle} the information generated by the iterative method while solving the current linear system, performing a low-rank update to $\mathbf{P}^{(k)}$ to create $\mathbf{P}^{(k+1)}$.
+To do so, we build on the rank-1 update formulas discussed in \cite[Chapter 6]{NocedalWright99}, which we generalize to many vector ``rank-$m$'' updates.
+In what follows, let $\mathbf{X}$ and $\mathbf{Y}$ denote the $N \times m$ matrices with columns given by the vectors $\mathbf{x}_i$ and $\mathbf{y}_i$, respectively.
+The matrix-vector products with $\mathbf{B}^{(k)}$ performed while solving the $k$th linear system may be summarized by the equation:
 \begin{equation*}
@@
 \end{equation*}
-For convenience, define 
-\begin{equation*}
-\mathbf{R} := \mathbf{X}-\mathbf{P}^{(k)}\mathbf{Y}, \qquad \text{and} \qquad \boldsymbol{\Theta} := \left(\mathbf{X}^T\mathbf{Y}\right)^{-1}.
-\end{equation*}
-One of the following three formulas may be used to form $\mathbf{P}^{(k+1)}$ from $\mathbf{P}^{(k)}$:
+To form $\mathbf{P}^{(k+1)}$ from $\mathbf{P}^{(k)}$, one may use one of the following three update formulas:
 \begin{align}
@@
 	\mathbf{P}^{(k+1)} &= \mathbf{P}^{(k)} + \mathbf{R} \left(\mathbf{R}^T \mathbf{Y}\right)^{-1} \mathbf{R}^T \label{eq:srk_update} \\
-	\mathbf{P}^{(k+1)} &= \left(\mathbf{I} - \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T\right)\mathbf{P}^{(k)}\left(\mathbf{I} - \mathbf{Y}\boldsymbol{\Theta}\mathbf{X}^T\right) + \mathbf{X}\boldsymbol{\Theta}\mathbf{X}^T \label{eq:dfp_update}
+	\mathbf{P}^{(k+1)} &= \left(\mathbf{I} - \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T\right)\mathbf{P}^{(k)}\left(\mathbf{I} - \mathbf{Y}\boldsymbol{\Theta}\mathbf{X}^T\right) + \mathbf{X}\boldsymbol{\Theta}\mathbf{X}^T, \label{eq:dfp_update}
 \end{align}
-In all three cases, direct calculation shows that we have $\mathbf{P}^{(k+1)}\mathbf{Y}=\mathbf{X} = \left(\mathbf{B}^{(k)}\right)^{-1} \mathbf{Y}$.
-Formulas \eqref{eq:res_update} and \eqref{eq:srk_update} perform rank-$m$ updates, while formula \eqref{eq:dfp_update} performs a rank-$2m$ update.
+where $\mathbf{R} := \mathbf{X}-\mathbf{P}^{(k)}\mathbf{Y}$ and $\boldsymbol{\Theta} := \left(\mathbf{X}^T\mathbf{Y}\right)^{-1}$.
+We note that for all three cases, the update is exact for $\mathbf{B}^{k}$, in the sense that 
+\begin{equation*}
+\mathbf{P}^{(k+1)}\mathbf{Y}=\mathbf{X} = \left(\mathbf{B}^{(k)}\right)^{-1} \mathbf{Y}.
+\end{equation*}
+Because the matrices $\mathbf{B}^{k}$ vary slowly, we expect that $\mathbf{P}^{(k+1)}$ is a good approximation for $\left(\mathbf{B}^{(k+1)}\right)^{-1}$ on the space spanned by the vectors $\mathbf{y}_i$.
+One should choose which formula to use based on the properties of the matrix $\mathbf{B}^{(k)}$.
 Formula \eqref{eq:res_update} does not preserve symmetry, and should therefore be used when $\mathbf{B}^{(k)}$ is not symmetric.
 Formula \eqref{eq:srk_update} preserves symmetry but not positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric and indefinite.
-Formula \eqref{eq:dfp_update} preserves both symmetry and positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric positive definite.
-When $m=1$, formula \eqref{eq:srk_update} is known as the SR1 (``symmetric rank-$1$'') update formula, and formula \eqref{eq:dfp_update} is known as the DFP update formula, which is named after it's inventor Davidson, and popularizers Fletcher and Powell.
-For more details on these update formulas, we recommend the discussion in Chapter 6 of \cite{NocedalWright99}.
-In \cite{NocedalWright99}, one-vector versions of formulas \eqref{eq:srk_update} and \eqref{eq:dfp_update} are discussed thoroughly; the versions here are straightforward generalizations to many-vector updates.
+Formula \eqref{eq:dfp_update}, which is used in this paper, preserves both symmetry and positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric positive definite.
+Formulas \eqref{eq:res_update} and \eqref{eq:srk_update} perform rank-$m$ updates, while formula \eqref{eq:dfp_update} performs a rank-$2m$ update.
+We note that when $m=1$, formulas \eqref{eq:srk_update} are \eqref{eq:dfp_update} are known as SR1 (symmetric rank-$1$) and DFP (Davidson-Fletcher-Powell) updates, respectively \cite[Chapter 6]{NocedalWright99}.
+
+%formula, which is named after it's inventor Davidson, and popularizers Fletcher and Powell \cite{DFP}.
+
+%For more details on these update formulas, we recommend .
+
+%In \cite{NocedalWright99}, one-vector versions of formulas \eqref{eq:srk_update} and \eqref{eq:dfp_update} are discussed thoroughly; the versions here are straightforward generalizations to many-vector updates.
 

### 2022-05-16 numerical_examples/stokes/Scratch.py

@@ -16,6 +16,7 @@ from localpsf.product_convolution_hmatrix import make_hmatrix_from_kernel, produ
 
 
 from stokes_inverse_problem_cylinder import *
+# from .stokes_inverse_problem_cylinder import *
 import dolfin as dl
 
 import scipy.sparse.linalg as spla

### 2022-05-16 paper/localpsf.tex
update rational positive definte appendix based on discussion with Noemi

@@
 \end{equation*}
-If $\Aop$ is also positive semi-definite, we adapt the high order rational function spectral projection method from \cite{HackbuschKress07} to flip (or approximately flip) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ to be positive.
+If $\Aop$ is also positive semi-definite, we adapt the high order rational function spectral projection method from \cite[Section 13.8]{HackbuschKress07} to flip (or approximately flip) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ to be positive.
 In detail, let $\Pi$ be the indicator function for the interval $(a,b)$,
@@
 
-Computing $\mathbf{\Pi}=\Pi(\mathbf{A}_H^{\text{sym}})$ directly is difficult, but we may approximate $\Pi$ by a high order rational function $\Pi_k \approx \Pi(\lambda)$ which is defined as follows:
+Computing $\mathbf{\Pi}=\Pi(\mathbf{A}_H^{\text{sym}})$ directly is difficult, but we may approximate $\Pi$ by a high order rational function $\Pi_k \approx \Pi$ which is defined as follows:
 \begin{equation*}
@@
 \begin{equation*}
-\mathbf{T} := \left(2 \mathbf{A}_H^{\text{sym}} - (b+a) \mathbf{I}\right)/(b-a).
+\mathbf{T} := T(\mathbf{A}_H^{\text{sym}}) = \left(2 \mathbf{A}_H^{\text{sym}} - (b+a) \mathbf{I}\right)/(b-a).
 \end{equation*}
 Using $\mathbf{\Pi}_k$ instead of $\mathbf{\Pi}$ in \eqref{eq:I_Pi_abs} allows us to efficiently approximate $|\mathbf{A}_H^{\text{sym}}|$ using H-matrix methods. 
+This approximation procedure is as follows.
 First, we compute $\lambda_\text{min}$ using the Lanczos method, and set $a=\gamma \lambda_\text{min}$ and $b=0$, where $\gamma\ge 1$ is a parameter that must be chosen.
-We use $\gamma=1.5$, though the method is relatively insensitive to $\gamma$ and works well for any $\gamma \in [1,2]$.
+%Because of error in the rational function approximation, if $\gamma=1$, the negative eigenvalues of largest magnitude will be approximately zeroed out instead of flipped. 
+The method is relatively insensitive to $\gamma$ and numerically we find that the method works well for any $\gamma \in [1,2]$. We use $\gamma=1.5$ in our numerical results. 
 Next, we compute $\mathbf{T}$ using fast H-matrix scalar multiplication and addition.
-We form the $\mathbf{T}^{2^k}$ via repeated squaring, $\mathbf{T}^2 = \mathbf{T} \mathbf{T}$, $\mathbf{T}^4 = \mathbf{T}^2 \mathbf{T}^2$, $\mathbf{T}^8 = \mathbf{T}^4 \mathbf{T}^4$ and so on, where each of these matrix squarings is performed using fast H-matrix multiplication.
+We form the $\mathbf{T}^{2^k}$ via repeated squaring, $\mathbf{T}^2 = \mathbf{T} \mathbf{T}$, $\mathbf{T}^4 = \mathbf{T}^2 \mathbf{T}^2$, $\mathbf{T}^8 = \mathbf{T}^4 \mathbf{T}^4$ and so on, where each of these matrix squarings is performed using H-matrix multiplication, which is fast.
 We recommend using $k=1$ or $2$ (we use $k=1$ in our numerical results).

### 2022-05-26 paper/localpsf.tex
writing computational cost section of paper. Also modified NewtonCGPCH.py

@@
 
+The computational cost of our approximation method may be divided into two categories: (1) the cost of building the H-matrix approximation, and (2) the cost of using the H-matrix approximation to perform further linear algebra operations.
+The cost to build the H-matrix approximation may further be broken down into (1a) the cost of computing the impulse response moments and batches, which involves applying $\Aop$ to vectors, and (1b) the cost of building the H-matrix after the impulse response moments and batches have been computed, which does not involve applying $\Aop$ to vectors.
+In practice, (1a) is the dominant cost, because our method is targeted at applications in which applying the operator $\Aop$ to a vector involves performing an expensive computational procedure such as solving a large-scale partial differential equation.
+Conceptually, our method was designed to construct an approximation that is as accurate as possible, using as few operator applies as possible, subject to the constraint that the further costs associated with building and using the approximation are near-linear (and therefore scalable) in $N$.
+
+\paragraph{(1a): Impulse response moments and batches} Computing $V$, $\mu$, and $\Sigma$ require applying $\Aop$ to $1$, $d$, and $d(d+1)/2$ vectors, respectively.
+Concretely, in one spatial dimensions ($d=1$), computing $V$, $\mu$, and $\Sigma$ requires $1+1+1=3$ operator applies.
+In two spatial dimensions, this cost is $1+2+3=6$ operator applies.
+In three spatial dimensions, this cost is $1+3+6=10$ operator applies.
+Computing each impulse response batch requires applying $\Aop$ to $1$ vector, so computing the impulse response batches $\{\eta_i\}_{i=1}^{n_b}$ requires $n_b$ operator applies.
+In total, computing the impulse response moments and batches requires
+\begin{equation*}
+	1 + d + d(d+1)/2 + n_b \qquad \text{operator applies.}
+\end{equation*}
+In a typical application one might have $d=2$ and $n_b=5$, in which case $11$ operator applies would be required.
+Choosing one sample point batch via the greedy ellipsoid packing algorithm described in Section \ref{sec:sample_point_selection} may require performing $O(Ns)$ ellipsoid intersection tests, where $s$ is the number of sample points in the batch.
+Choosing all of the sample points therefore requires performing
+\begin{equation*}
+	O(Nm) \qquad \text{ellipsoid intersection tests,}
+\end{equation*}
+where $m$ is the total number of sample points in all batches.
+The multiplicative dependence of $N$ with $m$ is undesirable, and reducing this cost is a target for future work.
+The cost of picking sample points could be reduced, for example, by ``thinning'' the sample points.
+That is, choosing batches from a smaller subset of $N_\text{thin} << N$ points, rather than all $N$ finite element Lagrange nodes.
+The cost of picking one batch could be reduced from $O(N s)$ to $O(N \log s)$ using more advanced computational geometry techniques, such as organizing the ellipsoids associated with points that have already been picked in a batch into an AABB tree, then using this AABB tree to accelerate the required ellipsoid intersection checks.
+This would reduce the cost of choosing all sample point batches from $O(Nm)$ to $O(N n_b \log s)$.
+However, in practice the cost for picking sample points is small compared to other parts of the approximation algorithm, so we do not pursue these improvements in this paper.
+
+\paragraph{(1b): }
+
+In our view, the goal of the approximation method is to reduce the number of operator applies as much as possible
+
+we consider the operator applies to be the dominant cost to a vector is an expensive computational procedure.
+For example, in our numerical results we demonstrate the method for applications in which applying $\Aop$ to a vector requires solving two partial differential equations, which are defined on a domain with a larger spacetime dimension than that of $\Omega$.
+
 WORK IN PROGRESS BELOW HERE!!!

### 2022-05-26 numerical_examples/stokes/NewtonCGPCH.py

@@ -29,7 +29,7 @@ import numpy as np
 
 class NCGConvergenceInfo:
     def __init__(me, extra_info=None):
-        me._extra_info=extra_info
+        me.extra_info=extra_info
         me._cg_iteration = list()
         me._hessian_matvecs = list()
         me._total_cost = list()
@@ -40,13 +40,11 @@ class NCGConvergenceInfo:
         me._step_length_alpha = list()
         me._tolcg = list()
 
+        me.labels = ['cg_iteration', 'hessian_matvecs', 'total_cost', 'misfit_cost', 'reg_cost', 'mg_mhat', 'gradnorm', 'step_length_alpha', 'tolcg']
+
     @property
     def num_newton_iterations(me):
         return len(me.cg_iteration)
-
-    @property
-    def extra_info(me):
-        return me._extra_info
 
     @property
     def cg_iteration(me):
@@ -84,6 +82,19 @@ class NCGConvergenceInfo:
     def tolcg(me):
         return np.array(me._tolcg)
 
+    @property
+    def data(me):
+        return np.array([me.cg_iteration,
+                         me.hessian_matvecs,
+                         me.total_cost,
+                         me.misfit_cost,
+                         me.reg_cost,
+                         me.mg_mhat,
+                         me.gradnorm,
+                         me.step_length_alpha,
+                         me.tolcg]).T
+
+
     def add_iteration(me,
                       cg_iteration,
                       hessian_matvecs,
@@ -104,16 +115,20 @@ class NCGConvergenceInfo:
         me._step_length_alpha.append(step_length_alpha)
         me._tolcg.append(tolcg)
 
-    def print(me):
-        print('Newton CG convergence information:')
-        if me._extra_info is not None:
-            print(me._extra_info)
+    def print(me, return_string=False):
+        print_string = 'Newton CG convergence information:'
 
-        print("\n{0:3} {1:3} {2:15} {3:15} {4:15} {5:15} {6:14} {7:14} {8:14}".format(
-            "It", "cg_it", "cost", "misfit", "reg", "(g,dm)", "||g||L2", "alpha", "tolcg"))
+        print_string += '\n'
+
+        if me.extra_info is not None:
+            print_string += me._extra_info
+
+        print_string += "\n{0:3} {1:3} {2:15} {3:15} {4:15} {5:15} {6:14} {7:14} {8:14}".format(
[…]

### 2022-05-26 paper/localpsf.tex
writing computational cost section of paper.

@@
 In a typical application one might have $d=2$ and $n_b=5$, in which case $11$ operator applies would be required.
-Choosing one sample point batch via the greedy ellipsoid packing algorithm described in Section \ref{sec:sample_point_selection} may require performing $O(Ns)$ ellipsoid intersection tests, where $s$ is the number of sample points in the batch.
-Choosing all of the sample points therefore requires performing
-\begin{equation*}
-	O(Nm) \qquad \text{ellipsoid intersection tests,}
+Choosing one sample point batch via the greedy ellipsoid packing algorithm described in Section \ref{sec:sample_point_selection} may require performing up to $Ns$ ellipsoid intersection tests, where $s$ is the number of sample points in the batch.
+Choosing all of the sample points therefore requires performing at most
+\begin{equation*}
+	Nm \qquad \text{ellipsoid intersection tests,}
 \end{equation*}
@@
 That is, choosing batches from a smaller subset of $N_\text{thin} << N$ points, rather than all $N$ finite element Lagrange nodes.
-The cost of picking one batch could be reduced from $O(N s)$ to $O(N \log s)$ using more advanced computational geometry techniques, such as organizing the ellipsoids associated with points that have already been picked in a batch into an AABB tree, then using this AABB tree to accelerate the required ellipsoid intersection checks.
-This would reduce the cost of choosing all sample point batches from $O(Nm)$ to $O(N n_b \log s)$.
+The cost of picking one batch could be reduced from $N s$ to $O(N \log s)$ using more advanced computational geometry techniques, such as organizing the ellipsoids associated with points that have already been picked in a batch into an AABB tree, then using this AABB tree to accelerate the required ellipsoid intersection checks.
+This would reduce the cost of choosing all sample point batches from $Nm$ to $O(N n_b \log s)$.
 However, in practice the cost for picking sample points is small compared to other parts of the approximation algorithm, so we do not pursue these improvements in this paper.
 
-\paragraph{(1b): }
+\paragraph{(1b): Building the H-matrix} The classical H-matrix construction technique, which is described in Appendix \ref{app:h_matrix}, requires evaluating $O(k_h^2 N \log N)$ matrix entries of the approximation, where $k_h$ is the H-matrix rank (maximum rank among the blocks of the H-matrix).
+To evaluate one matrix entry, first one must perform $k_n$ nearest neighbors queries of the sample points KD tree, and each of these queries requires $O(\log m)$ work.
+Next, one must finding the mesh cells that $k_n$ distinct points reside in using our mesh AABB tree, which requires $O(k_n \log N)$ work, and evaluate finite element basis functions on each of those cells, which requires $O(k_n)$ work.
+Finally, one must solve a $k_n \times k_n$ linear system to perform the radial basis function interpolation, which requires $O(k_n^3)$ work.
+Overall, building the H-matrix therefore requires
+\begin{equation*}
+	O\left(\left(k_h^2 N \log N\right)\left(k_n \log N + k_n^3\right)\right) \qquad \text{work}.
+\end{equation*}
+This cost is nearly linear in $N$, scalable with respect to the number of sample points $m$, and quite fast in practice.
 

### 2022-05-27 paper/localpsf.tex
wrote computational cost section of paper. Ready for editing pass

@@
 Conceptually, our method was designed to construct an approximation that is as accurate as possible, using as few operator applies as possible, subject to the constraint that the further costs associated with building and using the approximation are near-linear (and therefore scalable) in $N$.
+We now describe these costs in detail.
+For convenience, variable symbols and their approximate sizes are listed in Table \ref{tab:vars}.
+
+\begin{table}
+	\begin{tabular}{lll}
+		Symbol & Approximate size & Variable name \\
+		\hline
+		$N$ & $10^2$--$10^9$ & Number of finite element degrees of freedom \\
+		$n_b$ & $1$--$25$ & Number of impulse response batches \\
+		$k_h$ & $5$--$50$ & H-matrix rank \\
+		$k_n$ & $5$--$15$ & Number of nearest neighbors for RBF interpolation \\
+		$d$ & $1$--$3$ & Spatial dimension \\
+		$r$ & $10^1$--$10^4$ & Classical matrix rank \\
+		$m$ & $10^1$--$10^4$ & Total number of sample points (all batches) \\
+		$s$ & $1$--$500$ & Number of sample points in one batch \\
+		$l$ & $1$--$2$ & Rational function parameter for SPD modification
+	\end{tabular}
+	\caption{Symbols used for variables in computational cost estimates, and their approximate sizes.}
+	\label{tab:vars}
+\end{table}
 
@@
 Computing each impulse response batch requires applying $\Aop$ to $1$ vector, so computing the impulse response batches $\{\eta_i\}_{i=1}^{n_b}$ requires $n_b$ operator applies.
-In total, computing the impulse response moments and batches requires
+In total, computing the impulse response moments and batches therefore requires
 \begin{equation*}
@@
 In a typical application one might have $d=2$ and $n_b=5$, in which case $11$ operator applies would be required.
-Choosing one sample point batch via the greedy ellipsoid packing algorithm described in Section \ref{sec:sample_point_selection} may require performing up to $Ns$ ellipsoid intersection tests, where $s$ is the number of sample points in the batch.
+In contrast, matrix-free low-rank approximation methods require $O(r)$ operator applies, where $r$ is the classical numerical rank of $\Aop$.
+Since $r$ may be in the thousands, low rank approximation may require thousands of operator applies.
+
+Computing the impulse response batches also requires choosing sample point batches via the greedy ellipsoid packing algorithm described in Section \ref{sec:sample_point_selection}.
+Choosing one batch of sample points may require performing up to $Ns$ ellipsoid intersection tests, where $s$ is the number of sample points in the batch.
 Choosing all of the sample points therefore requires performing at most
@@
 The multiplicative dependence of $N$ with $m$ is undesirable, and reducing this cost is a target for future work.
-The cost of picking sample points could be reduced, for example, by ``thinning'' the sample points.
-That is, choosing batches from a smaller subset of $N_\text{thin} << N$ points, rather than all $N$ finite element Lagrange nodes.
+The cost of picking sample points could be reduced, for example, by ``thinning'' the set of candidate points that the sample points are picked from.
+That is, choosing sample points from a smaller subset of $N_\text{thin} << N$ points, rather than all $N$ finite element Lagrange nodes.
 The cost of picking one batch could be reduced from $N s$ to $O(N \log s)$ using more advanced computational geometry techniques, such as organizing the ellipsoids associated with points that have already been picked in a batch into an AABB tree, then using this AABB tree to accelerate the required ellipsoid intersection checks.
 This would reduce the cost of choosing all sample point batches from $Nm$ to $O(N n_b \log s)$.
-However, in practice the cost for picking sample points is small compared to other parts of the approximation algorithm, so we do not pursue these improvements in this paper.
+In practice the cost for picking sample points is small compared to other parts of the approximation algorithm, so we do not pursue these improvements in this paper.
 
 \paragraph{(1b): Building the H-matrix} The classical H-matrix construction technique, which is described in Appendix \ref{app:h_matrix}, requires evaluating $O(k_h^2 N \log N)$ matrix entries of the approximation, where $k_h$ is the H-matrix rank (maximum rank among the blocks of the H-matrix).
-To evaluate one matrix entry, first one must perform $k_n$ nearest neighbors queries of the sample points KD tree, and each of these queries requires $O(\log m)$ work.
-Next, one must finding the mesh cells that $k_n$ distinct points reside in using our mesh AABB tree, which requires $O(k_n \log N)$ work, and evaluate finite element basis functions on each of those cells, which requires $O(k_n)$ work.
-Finally, one must solve a $k_n \times k_n$ linear system to perform the radial basis function interpolation, which requires $O(k_n^3)$ work.
+To evaluate one matrix entry, first one must find the $k_n$ nearest sample points to a given point, where $k_n$ is the number of impulses used in the RBF interpolation.
+This is done using a precomputed KD tree of sample points, and requires $O(k_n \log m)$ work.
+Next, one must find the mesh cells that the points $\{z_i\}_{i=1}^{k+n}$ reside.
+This is done using our mesh AABB tree, and requires $O(k_n \log N)$ work.
+Next, we must evaluate finite element basis functions on each of those cells, which requires $O(k_n)$ work.
+Finally, we must solve a $k_n \times k_n$ linear system to perform the radial basis function interpolation, which requires $O(k_n^3)$ work.
 Overall, building the H-matrix therefore requires
@@
 
-In our view, the goal of the approximation method is to reduce the number of operator applies as much as possible
-
-we consider the operator applies to be the dominant cost to a vector is an expensive computational procedure.
-For example, in our numerical results we demonstrate the method for applications in which applying $\Aop$ to a vector requires solving two partial differential equations, which are defined on a domain with a larger spacetime dimension than that of $\Omega$.
-
-WORK IN PROGRESS BELOW HERE!!!
-
-\begin{tabular}{lll}
-	Variable name & Approximate size & Variable \\
-	\hline
-	$N$ & $10^3$--$10^9$ & Number of finite element degrees of freedom \\
-	$b$ & $5$--$25$ & number of impulse response batches \\
-	$r$ & $10$--$30$ & H-matrix rank \\
-	$k$ & $5$--$15$ & number of nearest neighbors \\
-	$\{\eta_b\}_{b=1}^\text{num\_batches}$ & $\text{num\_batches}$ actions of $\mathcal{A}$
-\end{tabular}
-
-$k_\text{hmatrix}$, $k_\text{nbrs}$, $k_\text{batches}$
-
-\begin{tabular}{ll}
-	$V$ & $1$ action of $\mathcal{A}^T$ \\
-	$\mu$ & $d$ actions of $\mathcal{A}^T$ \\
-	$\Sigma$ & $d(d+1)/2$ actions of $\mathcal{A}^T$ \\
-	$\{\eta_b\}_{b=1}^\text{num\_batches}$ & $\text{num\_batches}$ actions of $\mathcal{A}$
-\end{tabular}
-
-Computing $V$, $\mu$, and $\Sigma$ require $1$, $d$, and $d(d+1)/2$ operator actions of $\Aop$, respectively.
+\paragraph{(2): H-matrix operations} H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion require
+\begin{equation*}
+	O(k_h^2 N \log(N)^a) \qquad \text{work},
+\end{equation*}
+where $a \in \{0,1,2,3\}$.
+The precise cost depends on the type of H-matrix used and the operation being performed; for more details see \cite{GrasedyckHackbusch03}.
+The rational method for modifying $\mathbf{A}_H$ to be positive semi-definite, described in Appendix \ref{sec:make_spd}, is performed using two matrix-matrix additions, $l+1$ matrix-matrix multiplications, and one matrix inversion, where $2^l$ is the order of the rational function used.
+
+%$k_\text{hmatrix}$, $k_\text{nbrs}$, $k_\text{batches}$
+%
+%\begin{tabular}{ll}
+%	$V$ & $1$ action of $\mathcal{A}^T$ \\
+%	$\mu$ & $d$ actions of $\mathcal{A}^T$ \\
+%	$\Sigma$ & $d(d+1)/2$ actions of $\mathcal{A}^T$ \\
+%	$\{\eta_b\}_{b=1}^\text{num\_batches}$ & $\text{num\_batches}$ actions of $\mathcal{A}$
+%\end{tabular}
+%
+%
+%Computing $V$, $\mu$, and $\Sigma$ require $1$, $d$, and $d(d+1)/2$ operator actions of $\Aop$, respectively.
 Computing the impulse response batches $\eta_b$ requires one operator action of $\Aop$ per batch.
 Picking the sample points
-
-In the non-symmetric case, evaluating $\widetilde{\Phi}(y,x)$ requires us to perform one $k$-nearest neighbor query of the KD-tree of sample points, $k$ point collision queries of the AABB-tree of mesh cells, and solve a $k \times k$ linear system.
+%
+%In the non-symmetric case, evaluating $\widetilde{\Phi}(y,x)$ requires us to perform one $k$-nearest neighbor query of the KD-tree of sample points, $k$ point collision queries of the AABB-tree of mesh cells, and solve a $k \times k$ linear system.
 In the symmetric case, we must perform two $k$-nearest neighbor queries of the KD-tree of sample points, $2k$ point collision queries of the AABB-tree of mesh cells, and solve a $4k \times 2k$ linear least squares problem via QR factorization.

### 2022-05-28 paper/localpsf.tex
editing pass on computational cost section

@@
 We use H1 matrices in our numerical results, but any of the other H-matrix formats (such as H2, HODLR, HSS, HBS, and others \cite{HMATRIX}) could be used instead.
-For more details on H-matrices, we recommend \cite{Hackbusch15}.
+For more details on H-matrices, see \cite{Hackbusch15}.
 
@@
 
-The computational cost of our approximation method may be divided into two categories: (1) the cost of building the H-matrix approximation, and (2) the cost of using the H-matrix approximation to perform further linear algebra operations.
-The cost to build the H-matrix approximation may further be broken down into (1a) the cost of computing the impulse response moments and batches, which involves applying $\Aop$ to vectors, and (1b) the cost of building the H-matrix after the impulse response moments and batches have been computed, which does not involve applying $\Aop$ to vectors.
-In practice, (1a) is the dominant cost, because our method is targeted at applications in which applying the operator $\Aop$ to a vector involves performing an expensive computational procedure such as solving a large-scale partial differential equation.
-Conceptually, our method was designed to construct an approximation that is as accurate as possible, using as few operator applies as possible, subject to the constraint that the further costs associated with building and using the approximation are near-linear (and therefore scalable) in $N$.
+The computational cost of our method may be divided into the costs to perform the following tasks: 
+\begin{enumerate}
+	\item Computing the impulse response moments and batches. This includes picking the sample point batches via the greed ellipsoid packing procedure.
+	\item Constructing the H-matrix $\mathbf{A}_H$.
+	\item Using $\mathbf{A}_H$ to perform further linear algebra operations. This includes, optionally, modifying $\mathbf{A}_H$ to make it symmetric positive semi-definite.
+\end{enumerate}
+In practice, 1 is the dominant cost, because 1 is the only task that requires applying $\Aop$ to vectors, and our method is targeted at applications in which applying $\Aop$ to a vector requires an expensive computational procedure such as solving a large-scale partial differential equation.
+Hence, our method was designed to construct an approximation $\mathbf{A}_H$ that is as accurate as possible, using as few operator applies as possible, subject to the constraint that the further costs associated with building and using $\mathbf{A}_H$ are near-linear (and therefore scalable) in $N$.
 We now describe these costs in detail.
@@
 		$s$ & $1$--$500$ & Number of sample points in one batch \\
-		$l$ & $1$--$2$ & Rational function parameter for SPD modification
+		$l$ & $1$--$2$ & Rational function parameter for SPSD modification
 	\end{tabular}
@@
 
-\paragraph{(1a): Impulse response moments and batches} Computing $V$, $\mu$, and $\Sigma$ require applying $\Aop$ to $1$, $d$, and $d(d+1)/2$ vectors, respectively.
+\paragraph{1: Impulse response moments and batches} Computing $V$, $\mu$, and $\Sigma$ requires applying $\Aop$ to $1$, $d$, and $d(d+1)/2$ vectors, respectively.
 Concretely, in one spatial dimensions ($d=1$), computing $V$, $\mu$, and $\Sigma$ requires $1+1+1=3$ operator applies.
@@
 where $m$ is the total number of sample points in all batches.
-The multiplicative dependence of $N$ with $m$ is undesirable, and reducing this cost is a target for future work.
-The cost of picking sample points could be reduced, for example, by ``thinning'' the set of candidate points that the sample points are picked from.
-That is, choosing sample points from a smaller subset of $N_\text{thin} << N$ points, rather than all $N$ finite element Lagrange nodes.
-The cost of picking one batch could be reduced from $N s$ to $O(N \log s)$ using more advanced computational geometry techniques, such as organizing the ellipsoids associated with points that have already been picked in a batch into an AABB tree, then using this AABB tree to accelerate the required ellipsoid intersection checks.
-This would reduce the cost of choosing all sample point batches from $Nm$ to $O(N n_b \log s)$.
-In practice the cost for picking sample points is small compared to other parts of the approximation algorithm, so we do not pursue these improvements in this paper.
-
-\paragraph{(1b): Building the H-matrix} The classical H-matrix construction technique, which is described in Appendix \ref{app:h_matrix}, requires evaluating $O(k_h^2 N \log N)$ matrix entries of the approximation, where $k_h$ is the H-matrix rank (maximum rank among the blocks of the H-matrix).
+The multiplicative dependence of $N$ with $m$ is undesirable since $m$ may be large, and reducing this cost is a target for future work.
+Several improvements are possible.
+The cost of picking sample points could be reduced by ``thinning'' the set of candidate points that the sample points are picked from.
+That is, choosing sample points from a smaller subset of $N_\text{thin} << N$ points.
+The cost to perform the ellipsoid intersection tests could be reduced by first checking whether the ellipsoid bounding boxes intersect before performing the ellipsoid intersection test of Appendix \ref{sec:fast_ellipsoid_intersection_test}.
+The cost of picking one batch could be reduced from $N s$ to $O(N \log s)$ using more advanced computational geometry techniques, such as organizing the ellipsoids associated with points that have already been picked into a dynamic AABB tree, then using this AABB tree to accelerate the required ellipsoid intersection tests.
+This would reduce the cost of choosing all sample point batches to $O(N n_b \log s)$.
+However, from a practical perspective, the cost for picking sample points is small compared to other parts of the approximation algorithm, so we do not pursue these improvements in this paper.
+
+\paragraph{2: Building the H-matrix} The classical H-matrix construction technique, which is described in Appendix \ref{app:h_matrix}, requires evaluating $O(k_h^2 N \log N)$ matrix entries of the approximation, where $k_h$ is the H-matrix rank (maximum rank among the blocks of the H-matrix).
 To evaluate one matrix entry, first one must find the $k_n$ nearest sample points to a given point, where $k_n$ is the number of impulses used in the RBF interpolation.
@@
 Finally, we must solve a $k_n \times k_n$ linear system to perform the radial basis function interpolation, which requires $O(k_n^3)$ work.
-Overall, building the H-matrix therefore requires
+Overall building the H-matrix therefore requires
 \begin{equation*}
@@
 \end{equation*}
-This cost is nearly linear in $N$, scalable with respect to the number of sample points $m$, and quite fast in practice.
+This follows from the above discussion, and the fact that $N \ge m$.
 

### 2022-05-28 paper/localpsf.tex
editing pass on computational cost section

@@
 
-\paragraph{(2): H-matrix operations} H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion require
+\paragraph{(3): H-matrix operations} H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion require
 \begin{equation*}

### 2022-05-28 paper/localpsf.tex
refactoring variable names. Also, deleting old text

@@
 We present a fast \emph{matrix free} method for approximating operators, $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$, that have locally supported non-negative integral kernels.
-Here $\Omega \subset \mathbb{R}^d$ is a bounded domain, $d$ is the spatial dimension, and $L^2(\Omega)'$ is the space of continuous linear functionals on $L^2(\Omega)$ (i.e., the topological dual space).
+Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, $\gdim$ is the spatial dimension, and $L^2(\Omega)'$ is the space of continuous linear functionals on $L^2(\Omega)$ (i.e., the topological dual space).
 Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{le1997non,saad1999distributed,TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps) \cite{DTNMAP}, covariance operators in spatial statistics \cite{GeogaEtAl20,ChenStein21,LindgrenRueLindstrom11}, blurring operators in imaging \cite{DenisEtAl11,NagyOleary98}, and Hessians in distributed parameter PDE-constrained optimization and inverse problems \cite{AlgerEtAl19,IsaacEtAl15}.
@@
 We form $\mathbf{\Phi}_H$ using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{Kriemann08}.
-Although $\mathbf{\Phi}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Phi}_H$ only requires evaluation of $O(\hmatrixrank^2 \fedim \log \fedim)$ entries $\mathbf{\Phi}_{ij} = \widetilde{\Phi}(p_i, p_j)$, and these entries are computed via the procedure described in Section \ref{sec:approximate_kernel_entries}.
-Here $\hmatrixrank$ is the rank of the highest-rank block in the H-matrix.
+Although $\mathbf{\Phi}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Phi}_H$ only requires evaluation of $O(\hrank^2 \fedim \log \fedim)$ entries $\mathbf{\Phi}_{ij} = \widetilde{\Phi}(p_i, p_j)$, and these entries are computed via the procedure described in Section \ref{sec:approximate_kernel_entries}.
+Here $\hrank$ is the rank of the highest-rank block in the H-matrix.
 A dense representation of $\mathbf{\Phi}$ is never formed.
@@
 		$n_b$ & $1$--$25$ & Number of impulse response batches \\
-		$k_h$ & $5$--$50$ & H-matrix rank \\
+		$\hrank$ & $5$--$50$ & H-matrix rank \\
 		$k_n$ & $5$--$15$ & Number of nearest neighbors for RBF interpolation \\
@@
 \begin{equation*}
-	1 + d + d(d+1)/2 + n_b \qquad \text{operator applies.}
+	1 + \gdim + \gdim(\gdim+1)/2 + n_b \qquad \text{operator applies.}
 \end{equation*}
@@
 
-\paragraph{2: Building the H-matrix} The classical H-matrix construction technique, which is described in Appendix \ref{app:h_matrix}, requires evaluating $O(k_h^2 N \log N)$ matrix entries of the approximation, where $k_h$ is the H-matrix rank (maximum rank among the blocks of the H-matrix).
+\paragraph{2: Building the H-matrix} The classical H-matrix construction technique, which is described in Appendix \ref{app:h_matrix}, requires evaluating $O(\hrank^2 N \log N)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank (maximum rank among the blocks of the H-matrix).
 To evaluate one matrix entry, first one must find the $k_n$ nearest sample points to a given point, where $k_n$ is the number of impulses used in the RBF interpolation.
@@
 \begin{equation*}
-	O\left(\left(k_h^2 N \log N\right)\left(k_n \log N + k_n^3\right)\right) \qquad \text{work}.
+	O\left(\left(\hrank^2 N \log N\right)\left(k_n \log N + k_n^3\right)\right) \qquad \text{work}.
 \end{equation*}
@@
 \begin{equation*}
-	O(k_h^2 N \log(N)^a) \qquad \text{work},
+	O(\hrank^2 N \log(N)^a) \qquad \text{work},
 \end{equation*}
@@
 \begin{itemize}
-	\item Let $C \in \mathbb{R}^{m \times \hmatrixrank}$ be a matrix consisting of a subset of $\hmatrixrank$ columns of $X$, such that the span of the columns in $C$ approximates the column space of $X$.
-	\item Let $R\in\mathbb{R}^{\hmatrixrank \times m}$ be a subset of the rows of $X$, such that the span of the rows in $R$ approximates the row space of $X$.
-	\item Let $U \in \mathbb{R}^{\hmatrixrank \times \hmatrixrank}$ be the block of $X$ consisting of the intersection of the rows from $R$ with the columns from $C$.
+	\item Let $C \in \mathbb{R}^{m \times \hrank}$ be a matrix consisting of a subset of $\hrank$ columns of $X$, such that the span of the columns in $C$ approximates the column space of $X$. 
+	\item Let $R\in\mathbb{R}^{\hrank \times m}$ be a subset of the rows of $X$, such that the span of the rows in $R$ approximates the row space of $X$.
+	\item Let $U \in \mathbb{R}^{\hrank \times \hrank}$ be the block of $X$ consisting of the intersection of the rows from $R$ with the columns from $C$.
 \end{itemize}
@@
 This results in matrices $R$ and $C$ such that the error in the approximation, \eqref{eq:CUR}, is small.
-The dominant cost of this procedure is the cost of computing $a \hmatrixrank$ columns of $X$ and $a\hmatrixrank$ rows of $X$, where $a$ is the number of alternating iterations.
+The dominant cost of this procedure is the cost of computing $a \hrank$ columns of $X$ and $a\hrank$ rows of $X$, where $a$ is the number of alternating iterations.
 There is also a small linear algebra overhead cost for the determinant maximization process that is performed at each step.
 %For more details on adaptive cross low rank approximation, see \cite{ACA}.
-The key point is that the adaptive cross method allows us to form a rank-$\hmatrixrank$ approximation of an $m \times m$ block of the matrix via a process that only requires accessing $O(m\hmatrixrank)$ entries of that block.
+The key point is that the adaptive cross method allows us to form a rank-$\hrank$ approximation of an $m \times m$ block of the matrix via a process that only requires accessing $O(m\hrank)$ entries of that block.
 
@@
 We directly compute all entries of the small dense blocks of $\mathbf{\Phi}$ that are not admissible.
-This process requires us to compute $O(\hmatrixrank^2 \fedim \log \fedim)$ entries of $\mathbf{\Phi}$, which is relatively cheap compared to the operator actions of $\Aop$ that are required to form the product-convolution approximation.
+This process requires us to compute $O(\hrank^2 \fedim \log \fedim)$ entries of $\mathbf{\Phi}$, which is relatively cheap compared to the operator actions of $\Aop$ that are required to form the product-convolution approximation.
 

### 2022-05-29 paper/localpsf.tex
good editing pass over computational cost section

@@
 In applications, evaluating the maps in \eqref{eq:eval_maps} may involve iteratively solving a large linear system, timestepping, or performing some other costly computational procedure.
-%After discretization, $\Aop$ becomes a dense matrix that is typically too large to be built and stored.
-This property is called ``matrix-free'' because at the discrete level it means one can compute matrix-vector products with the matrix representations of $\Aop$ and $\Aop^T$, but cannot easily compute individual matrix entries.
 
@@
 The objective function in the minimization problem is defined implicitly, so information about the Hessian of the objective function is only accessible via application of the Hessian to vectors \cite{HESSIANACTION}.
-%In particular, applying the Hessian to a vector requires solving two linear PDEs which take the form of linearizations of \eqref{eq:state_pde} \cite{HESSIANACTION}. 
-%More accessible Hessian approximations are highly desirable because they permit fast solution of the minimization problem via Newton-type methods.
-More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because the inverse of the Hessian locally approximates the Bayesian posterior covariance for the parameter \cite{LAPLACEAPPROXIMATION}.
 The most popular matrix-free Hessian approximation methods are based on low rank approximation \cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15} because there are efficient methods for constructing low rank approximations of a matrix using only matrix-vector products \cite{Cheng05,HalkoMartinssonTropp11}.
-%This includes the Lanczos method, the randomized singular value decomposition \cite{HMTRANDOM}, and CUR approximation/skeletonization \cite{CUR,MAHONEY}. 
 However, low rank approximation methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation \cite{Alger19,AlgerEtAl17}.
@@
 These H-matrix methods are fast and scale well to large problems.
-Typically, the work and memory required to perform these operations for an $N \times N$ H-matrix scales as $O(N \log(N)^a)$, where $a \in \{0,1,2,3\}$.
+Typically, the work and memory required to perform these operations for an $\fedim \times \fedim$ H-matrix scales as $O(\fedim \log(\fedim)^a)$, where $a \in \{0,1,2,3\}$.
 The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks; for more details see \cite{GrasedyckHackbusch03}.
@@
 The large circle is the boundary of the domain.
-		%		Early batches tend to have lots of ellipsoids near the boundary, as compared to later batches. 
-		%		This is because our greedy ellipsoid packing algorithm tries to pick sample points that are far away from sample points in previously chosen batches.
 	}
@@
 
-%Since our end goal is an H-matrix, it is natural to ask why we do not build the H-matrix directly using classical H-matrix construction methods. The reason is that 
 Classical H-matrix construction methods (described in Appendix \ref{sec:H_matrix_construction}) require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.
@@
 Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a positive integral kernel.
-
-%We pick batches of points $x_i$ in such that many $\phi_{x_i}$ are computed with each operator action of $\Aop$.
-
-%We seek to approximate operators $\Aop$ that are accessible \emph{matrix-free}. That is, we have a black-box computational procedure through which we may evaluate the maps
-%\begin{equation*}
-%u \mapsto\Aop u \qquad \text{and} \qquad v \mapsto \Aop^Tv
-%\end{equation*} 
-%for arbitrary functions $u$ and $v$, but we cannot easily compute kernel entries $\Aker(y,x)$.
-In applications, evaluating these maps may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
-After discretization, $\Aop$ becomes a dense matrix that is typically too large to be built and stored.
-This property is called ``matrix-free'' because at the discrete level it means one can compute matrix-vector products with the matrix representation of $\Aop$, but cannot easily compute individual matrix entries.
-
-%Product-convolution (PC) approximations can (where applicable) approximate high rank operators efficiently.
 
@@
 
-%A single convolution operator may be high rank, so PC approximations can (where applicable) approximate high rank operators efficiently. 
-%The method proposed in this paper is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of scattered points $x_i \in \Omega$ \cite{PRODCONVLIST}.
-%Often PC methods are known as ``point spread function'' (PSF) methods, particularly in optics and image processing.
-There, the ``operator'' $\Aop$ is a physical process of distortion and blurring of light rays passing through an optical system such as a telescope or microscope, and the local convolution kernels are shifted\footnote{The impulse responses are shifted so that each convolution kernel is centered at zero instead of the location of the point source.} impulse responses of the system to point sources of light.
-%, which characterize the local distortion and blurring, and the weighting functions interpolate between these convolution kernels. 
-%PSF is also used as an umbrella term to refer to other approximation methods that focus on interpolating impulse responses or impulse response features.
-For background on PC and PSF methods, we recommend the following papers: \cite{PRODCONVGOOD}.
-
-%The method proposed in this paper is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses $\phi_{x_i}$ associated with a collection of scattered points $x_i \in \Omega$ \cite{PRODCONVLIST}.
-%A common choice for PC methods based on scattered impulse responses is to choose the $x_i$ to be nodes in a regular grid \cite{NAGY}.
-With a regular grid, a large number of $x_i$ are typically required to achieve an accurate approximation.
-%In our previous work, we started with a coarse grid of points $x_i$, then adaptively refined the grid in the regions where the error in the approximation was large \cite{PRODCONVMYPAPER}.
-However, computing each $\phi_{x_i}$ required one action of $\Aop$.
-Even with adaptive refinement, many $x_i$, and hence operator actions of $\Aop$, were required.
-In this paper, we reduce the computational cost further by picking points $x_i$ in a way that allows many $\phi_{x_i}$ to be computed with each operator action of $\Aop$.
-This requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
-The idea of estimating the supports of the functions $\phi_x$ a-priori was inspired by techniques from resolution analysis in seismic imaging, in which
-%In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_x$ is used to estimate the minimum length scale on which features of the parameter can be inferred near the point $x$.
-If the width of $\phi_x$ is ten meters, for instance, then features of the parameter near $x$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
-In \cite{RESOLUTION},
-%the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion  \cite{RESOLUTION}.
-Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Section \ref{sec:intromoments}).
-Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a positive integral kernel.
-
-%\subsection{Motivation: PDE constrained inverse problems}
-%\label{sec:PDE_hessian_motivation}
-
-%Our motivation for this work is approximation of Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs).
-That is, inverse problems in which one seeks to reconstruct an unknown parameter field, $m$, from noisy observations,
-%\begin{equation*}
-%y = f(m, u(m)) + \text{noise},
-%\end{equation*}
-%which depend on  a state variable $u$. In turn, $u$ depends on $m$ implicitly through the solution of a PDE, which we write generically as follows:
-%\begin{equation}
-%\label{eq:state_pde}
-%0 = g(m,u).
-%\end{equation}
-%Here $u(m)$ denotes the solution of the PDE \eqref{eq:state_pde} as a function of $m$.
-In the deterministic approach to inverse problems, one typically finds $m$ as the solution to the minimization problem
-%\begin{equation}
-%\label{eq:minimization_problem}
-%\min_m \quad \frac{1}{2}\|y - f(m,u(m))\|_W^2 + \mathcal{R}(m),
-%\end{equation}
-%where $\|\cdot\|_W$ is weighted norm which depends on the noise covariance, and $\mathcal{R}(m)$ is a regularization term.
-The objective function in optimization problem \eqref{eq:minimization_problem} is defined indirectly via the implicit function theorem.
-As a result, information about the Hessian of the objective function is only accessible via the action of the Hessian on vectors.
-In particular, applying the Hessian to a vector requires solving two linear PDEs which take the form of linearizations of \eqref{eq:state_pde} \cite{HESSIANACTION}.
-More accessible approximations of the Hessian are highly desirable, because they allow for fast solution of \eqref{eq:minimization_problem} via Newton-type methods.
-More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because the inverse of the Hessian locally approximates the Bayesian posterior covariance for $m$ \cite{LAPLACEAPPROXIMATION}.
-
-%\subsection{Background}
-
-%PC approximations have been used in a variety of fields going back several decades \cite{PRODCONVLIST}; for background on PC and PSF methods, we recommend the following papers: \cite{PRODCONVGOOD}.
-The method proposed in this paper is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of scattered points $x_i \in \Omega$ \cite{PRODCONVLIST}.
-A common choice for PC methods in this class is to choose the $x_i$ to be nodes in a regular grid \cite{NAGY}.
-With a regular grid, a large number of $x_i$ are typically required to achieve an accurate approximation.
-In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
-However, computing each impulse response required one action of $\Aop$.
-Even with adaptive refinement, many $x_i$, and hence operator actions of $\Aop$, were required.
-
-% A single convolution operator may be high rank, so PC approximations can (where applicable) approximate high rank operators efficiently.
-PC approximations have been used in a variety of fields going back several decades \cite{PRODCONVLIST}.
-Often PC methods are known as ``point spread function'' (PSF) methods, particularly in optics and image processing.
-There, the ``operator'' $\Aop$ is a physical process of distortion and blurring of light rays passing through an optical system such as a telescope or microscope.
-However, ``PSF'' is sometimes used as an umbrella term to refer to other approximation methods that focus on interpolating impulse responses or impulse response features.
-For background on PC and PSF methods, we recommend the following papers: \cite{PRODCONVGOOD}.
-
-%The method proposed in this paper is not exactly a PC approximation, but it builds upon a class of PC approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of scattered points $x_i \in \Omega$ \cite{PRODCONVLIST}.
-A common choice for PC methods in this class is to choose the $x_i$ to be nodes in a regular grid \cite{NAGY}.
-With a regular grid, a large number of $x_i$ are typically required to achieve an accurate approximation.
-In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}.
-However, computing each impulse response required one action of $\Aop$.
-Even with adaptive refinement, many $x_i$, and hence operator actions of $\Aop$, were required.
-
-%In this paper, we reduce the computational cost further by picking points $x_i$ in a way that allows many $\phi_{x_i}$ to be computed with each operator action of $\Aop$.
-This requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
-This idea was inspired by techniques from resolution analysis in seismic imaging \cite{RESOLUTION}.
-In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_x$ is used to estimate the minimum length scale on which features of the parameter can be inferred near the point $x$.
-If the width of $\phi_x$ is ten meters, for instance, then features of the parameter near $x$ can be accurately inferred from data if those features are larger than roughly ten meters in size.
-In \cite{RESOLUTION}, the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion.
-Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Section \ref{sec:intromoments}).
-Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a positive integral kernel.
-
-%Conventionally, PC approximations are applied to vectors using the fast Fourier transform (FFT) \cite{PRODCONVFFT}, which allows PC approximations to be used as fast surrogates for $\Aop$.
-Often it is necessary to perform further linear algebra operations involving $\Aop$, such as solving linear systems with $\Aop$ as the coefficient operator.
-One may perform these linear algebra operations using Krylov methods, replacing $\Aop$ with its PC approximation \cite{PRODCONVKRYLOV}.
-The Krylov method is accelerated because performing matrix vector products with the product-convolution approximation is typically faster than performing operator actions with $\Aop$.
-However, there are several problems with this approach.
-First, the number of Krylov iterations can be large, particularly when $\Aop$ is high-rank.
-Second, complicated geometries are rarely discretized using regular grids, so using the FFT to perform convolutions requires transferring discretized functions back and forth between the irregular mesh and a regular grid.
-Third, the computational cost grows with the number of convolution kernels used.
-Thus, it is desirable to convert the PC approximation to other matrix formats that are amenable to faster linear algebra operations.
-Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}, because operators that are amenable to PC approximation are often sparse in the wavelet domain.
-%While not commonly done, in theory linear algebra operations involving convolution-product approximations (transposes of PC approximations) could be done using fast methods for pseudo-differential operator symbol calculus \cite{YINGDEMANET} since these are equivalent to low rank approximation of the pseudo-differential symbol of the operator.
-We also note other relevant matrix-free operator approximation methods based on wavelets \cite{HERRMANN} and pseudo-differential operators \cite{DEMANET} which do not use PC approximation.
-
 \begin{figure}
@@
 	\end{center}
-	\caption{An operator is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \mu(x) + \mu(x')\right)$ when $x$ is close to $x'$.
-Here $\mu(x)$ is the mean (center of mass) of $\phi_x$, and similar for $\mu(x')$.}
+	\caption{An operator is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$.
+Here $\spatialmean(x)$ is the mean (center of mass) of $\phi_x$, and similar for $\spatialmean(x')$.}
 	\label{fig:mean_displacement_invariance}
@@
 
-%In this paper, we build on previous work \cite{PRODCONVMYPAPER}, in which we converted PC approximations into H-matrix format.
-
-%In exchange for these improvements, the proposed approximation cannot be written as a weighted sum of convolution operators, hence we cannot apply this approximation to vectors via the FFT.
-However, since our approximation is an H-matrix, it can be applied to vectors efficiently via H-matrix methods.
-
 \section{Preliminaries}
 
-Let $\Omega \subset \mathbb{R}^d$ be a bounded domain (in applications typically $\gdim=1$, $2$, or $3$).
+Let $\Omega \subset \mathbb{R}^\gdim$ be a bounded domain (in applications typically $\gdim=1$, $2$, or $3$).
 We seek to approximate integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ that may be written in the following form:
@@
 Finally, we detail finite element discretization (Section \ref{sec:finite_element_kernel}).
-%We write $L^2(\Omega)'$ to denote the topological dual of $L^2(\Omega)$. 
-%The linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$.
 
@@
 
-The action of $\Aop$ may be extended to distributions if the kernel $\Phi$ is sufficiently regular.
+The action of $\Aop$ may be extended to distributions if the kernel $\Aker$ is sufficiently regular.
 We derive the extension as follows.
@@
 	\begin{align}
-		\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Phi(y,x) \genericdistribution^*(x) dx ~dy \\
-		&= \int_\Omega v(y) \int_\Omega \Phi(y,x) \genericdistribution^*(x) dx ~dy \\
-		&= \int_\Omega v(y) \genericdistribution\left(\Phi(y,\cdot)\right) dy, \label{eq:action_on_distribution}
+		\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Aker(y,x) \genericdistribution^*(x) dx ~dy \\
+		&= \int_\Omega v(y) \int_\Omega \Aker(y,x) \genericdistribution^*(x) dx ~dy \\
+		&= \int_\Omega v(y) \genericdistribution\left(\Aker(y,\cdot)\right) dy, \label{eq:action_on_distribution}
 	\end{align}
 \end{subequations}
-%\begin{align*}
-%\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Phi(y,x) \genericdistribution^*(x) dx dy \\
-%&= \int_\Omega v(y) \int_\Omega \Phi(y,x) \genericdistribution^*(x) dx \\
-%&= \int_\Omega v(y) \genericdistribution\left(\Phi(y,\cdot)\right) dy 
-%%= \left(v,\genericdistribution\left(\Phi(y,\cdot)\right)\right)_{L^2(\Omega)},
-%\end{align*}
-where $\Phi(y,\cdot)$ denotes the function $x \mapsto \Phi(y,x)$.
+where $\Aker(y,\cdot)$ denotes the function $x \mapsto \Aker(y,x)$.
 Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions, and instead suppose $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ is a distribution.
 In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in \eqref{eq:extension_to_distributions} is not valid.
-However, if $\Phi$ is sufficiently regular such that the function $y \mapsto \rho\left(\Phi(y,~\cdot~)\right)$ is well-defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of \eqref{eq:action_on_distribution} is well-defined.
+However, if $\Aker$ is sufficiently regular such that the function $y \mapsto \rho\left(\Aker(y,~\cdot~)\right)$ is well-defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of \eqref{eq:action_on_distribution} is well-defined.
 Hence, we \emph{define} the action of $\Aop$ on the distribution $\rho$ to be the right hand side of \eqref{eq:action_on_distribution}.
 For convenience of notation, we denote this action by ``$\Aop \genericdistribution^*$,'' even if $\rho^*$ does not exist.
-%\begin{equation}
-%\label{eq:action_on_distribution}
-%\left(\Aop \genericdistribution^*\right)(v) := \int_\Omega v(y) \genericdistribution\left(\Phi(y, \cdot)\right) dy.
-%\end{equation}
-%Since $\genericdistribution^*$ may not exist if $\rho$ is a distribution, 
-%We write ``$\Aop \genericdistribution^*$'' to denote ``the action of $\Aop$ on the distribution $\rho$,'' as defined by \eqref{eq:action_on_distribution}.
-
-%We assume that $\Phi$ is sufficiently regular such that the action of $\Aop$ on delta distributions is well-defined, because
-
-%This is derived formally as follows:
-%\begin{align*}
-%\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Phi(y,x) \textrm{``} \genericdistribution^*(x) \textrm{''} dx dy \\
-%&= \int_\Omega v(y) \int_\Omega \Phi(y,x) \textrm{``} \genericdistribution^*(x) \textrm{''} dx \\
-%&= \int_\Omega v(y) \genericdistribution\left(\Phi(y,\cdot)\right) dx.
-%\end{align*}
-%Here $\textrm{``}\rho^*\textrm{''}$ formally denotes the Riesz representative of $\rho$, which, of course, does not exist if $\rho$ is a pure distribution.
-
-%an abuse of notation which views $\rho$ as a ``function'' of $x$, rather than a linear functional.
-That is, $\rho^*$ would be the Riesz representative of $\rho$, but for the obstruction that this Riesz representative does not exist on all of $L^2(\Omega)$.
-
-%For example, the delta distribution $\delta_x$ is defined by $\delta_x(v) = v(x)$, and the action of $\Aop$ on $\delta_x$ is given by
-%\begin{equation}
-%\label{eq:appendix_impulse_response}
-%\left(\Aop \delta_x^*\right)(v)	= \int_\Omega v(y) \Phi(y,x) dy.
-%\end{equation}
-%Recall that the impulse response of $\Aop$ to $\delta_x$, denoted $\impulseresponse_x$, is the Riesz representation of $\Aop \delta_x^*$. 
-%From \eqref{eq:appendix_impulse_response} and the definition of the Riesz representation, we have $\impulseresponse_x = A(~\cdot~,x)$.
-That is, $\impulseresponse_x$ is the function $y \mapsto A(y,x)$.
-
-%We seek to approximate operators $\Aop$ that are accessible \emph{matrix-free}. That is, we have a black-box computational procedure through which we may evaluate the maps
-%\begin{equation*}
-%u \mapsto\Aop u \qquad \text{and} \qquad v \mapsto \Aop^Tv
-%\end{equation*} 
-%for arbitrary functions $u$ and $v$, but we cannot easily compute kernel entries $\Aker(y,x)$.
-In applications, evaluating these maps may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure.
-After discretization, $\Aop$ becomes a dense matrix that is typically too large to be built and stored.
-This property is called ``matrix-free'' because at the discrete level it means one can compute matrix-vector products with the matrix representation of $\Aop$, but cannot easily compute individual matrix entries.
-
-%Our method, described in Section \ref{sec:method}, belongs to a class of methods that approximate an operator by interpolating impulse responses of the operator associated with a scattered collection of points.
+
 Let $\delta_x$ denote the delta distribution\footnote{Recall that the delta distribution $\delta_x:\mathcal{D}(\Omega)\rightarrow \mathbb{R}$ is the linear functional defined by $\delta_x(v) = v(x)$ for all $v\in \mathcal{D}(\Omega)$.} (i.e., point source, impulse) centered at the point $x \in \Omega$.
@@
 \end{equation} 
-which results from keeping $x$ fixed and viewing $\Phi(y,x)$ as a function of $y$.
-
-%The function $\impulseresponse_x$ is called the impulse response because, using \eqref{eq:action_on_distribution} and the definition of the delta function, we have:
-
-%where $\delta_x$ is the the delta distribution (i.e., point source, impulse) centered at $x$, $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.
-
-%Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Aker$ is sufficiently regular then the action of $\Aop$ may be extended to distributions.
-This is described in Appendix \ref{app:distributions}.
+which results from keeping $x$ fixed and viewing $\Aker(y,x)$ as a function of $y$.
 
@@
 \begin{enumerate}
-	\item The kernel $\Phi$ is sufficiently regular such that $\phi_x$ is a well-defined function for all $x\in \Omega$.
+	\item The kernel $\Aker$ is sufficiently regular such that $\phi_x$ is a well-defined function for all $x\in \Omega$.
 	\item The supports of the impulse responses $\impulseresponse_x$ are contained in localized regions.
-	\item The integral kernel is non-negative in the sense that $$\Phi(y,x) \ge 0$$ for all $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
+	\item The integral kernel is non-negative in the sense that $$\Aker(y,x) \ge 0$$ for all $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
 The operator $\Aop$ need not be positive semi-definite to use our method, and positive semi-definite operators need not have a non-negative integral kernel.}
@@
 Not all operators of interest (and in particular not all Hessians) satisfy these conditions, but many operators of practical interest do satisfy these conditions, and our method is highly effective for approximating these operators.
-%Because of these conditions, we are able to estimate the supports of the impulse responses a-priori by applying $\Aop^T$ \footnote{Recall that the transpose of $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ satisfying $\left(\Aop u\right)(v) = \left(\Aop^T v\right)(u)$ for all $u,v\in L^2(\Omega)$.} to a small number of polynomial functions (Section \ref{sec:intromoments}).
-We use the support estimates to compute scattered impulse responses in batches (Section \ref{sec:get_impulse_response}, also Figure \ref{fig:batches_intro}).
-This batching allows us to recover many impulse responses using a small number of operator actions $u \mapsto \Aop u$.
-The support estimates also allow us to interpolate the impulse responses with a new more sophisticated and accurate interpolation scheme (Sections \ref{sec:local_mean_displacement_invariance} and \ref{sec:approximate_kernel_entries}).
 
@@
 We assume that the finite element basis satisfies the Kronecker property, i.e., $\febasis_i(p_i)=1$ and $\febasis_i(p_j)=0$ if $i\neq j$.
-For $u_h \in V_h$ we write $\bm{u} \in \mathbb{R}^N_\mathbf{M}$ to denote the coefficient vector for $u_h$ with respect to the finite element basis, i.e.,
+For $u_h \in V_h$ we write $\bm{u} \in \mathbb{R}^\nsamplepts_\massmatrix$ to denote the coefficient vector for $u_h$ with respect to the finite element basis, i.e.,
 \begin{equation}
@@
 \end{equation}
-Linear functionals $\rho_h \in V_h'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^N_{\mathbf{M}^{-1}}$, with entries 
-\begin{equation*}
-\boldsymbol{\rho}_i = \rho_h(\febasis_i), \qquad i=1,\dots,N. 
-\end{equation*}
-Here $\bm{M} \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries
-\begin{equation*}
-\bm{M}_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx, \qquad i,j=1,\dots,N,
-\end{equation*}
-the space $\mathbb{R}^N_\mathbf{M}$ is $\mathbb{R}^N$ with the inner product $(\mathbf{u},\mathbf{v})_\mathbf{M} := \mathbf{u}^T \mathbf{M} \mathbf{v}$, and $\mathbb{R}^N_{\mathbf{M}^{-1}}$ is the analogous space with $\mathbf{M}^{-1}$ replacing $\mathbf{M}$.
-Direct calculation shows that $\mathbb{R}^N_\mathbf{M}$ and $\mathbb{R}^N_{\mathbf{M}^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively.
-%, in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u}, \mathbf{v}\right)_\mathbf{M}$, $\left(\rho_h, \nu_h\right)_{L^2(\Omega)} = \left(\boldsymbol{\rho}, \boldsymbol{\nu}\right)_{\mathbf{M}^{-1}}$, and $\rho_h(u) = \boldsymbol{\rho}^T\mathbf{u}$, for $u_h,v_h \in V_h$ and $\rho_h, \nu_h \in V_h'$.
+Linear functionals $\rho_h \in V_h'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^\nsamplepts_{\massmatrix^{-1}}$, with entries 
+\begin{equation*}
+\boldsymbol{\rho}_i = \rho_h(\febasis_i), \qquad i=1,\dots,\nsamplepts. 
+\end{equation*}
+Here $\massmatrix \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries
+\begin{equation*}
+\massmatrix_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx, \qquad i,j=1,\dots,\fedim,
+\end{equation*}
+the space $\mathbb{R}^\fedim_\massmatrix$ is $\mathbb{R}^\fedim$ with the inner product $(\mathbf{u},\mathbf{v})_\massmatrix := \mathbf{u}^T \massmatrix \mathbf{v}$, and $\mathbb{R}^\fedim_{\massmatrix^{-1}}$ is the analogous space with $\massmatrix^{-1}$ replacing $\massmatrix$.
+Direct calculation shows that $\mathbb{R}^\fedim_\massmatrix$ and $\mathbb{R}^\fedim_{\massmatrix^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively.
 
@@
 \begin{equation*}
-\mathbf{A}:\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}
+\mathbf{A}:\mathbb{R}^\fedim_\massmatrix \rightarrow \mathbb{R}^\fedim_{\massmatrix^{-1}}
 \end{equation*}
@@
 Rather, we have a computational procedure that allows us to compute matrix-vector products $\mathbf{u}\mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{v}\mapsto \mathbf{A}^T\mathbf{v}$, and computing these matrix-vector products is costly.
-%, and each of these matrix-vector products is costly to compute. 
-%For example, some computational procedures possess the Galerkin property, in which case $\mathbf{A}_{ij} = \left(\Aop \febasis_j\right)(\febasis_i)$. 
-%Our approximation method is agnostic to the computational procedure for performing these matrix-vector products. 
-%What is important is that $\mathbf{A}$ is dense, only available via matrix-vector products, and each of these matrix-vector products is costly to compute. 
 Of course, matrix entries can be computed via matrix-vector products as $\mathbf{A}_{ij} = \left(\mathbf{A}\mathbf{e}_j\right)_i$,
-%\begin{equation*}
-%\mathbf{A}_{ij} = \left(\mathbf{A}\mathbf{e}_j\right)_i,
-%\end{equation*}
-where $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)^T$ is the length $N$ unit vector with one in the $j$th coordinate and zeros elsewhere.
+where $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)^T$ is the length $\fedim$ unit vector with one in the $j$th coordinate and zeros elsewhere.
 But computing the matrix-vector product $\mathbf{e}_j \mapsto \mathbf{A}\mathbf{e}_j$ is costly, and therefore wasteful if we do not use other matrix entries in the $j$th column of $\mathbf{A}$.
@@
 \begin{equation}
-(A_h u_h)(v_h) := \int_\Omega \int_\Omega v_h(y) \Phi_h(y,x) u_h(x) dx dy
-\end{equation}
-for some kernel $\Phi_h$ which we do not know, and which may differ from $\Phi$ due to discretization error.
-Since the functions in $V_h$ are continuous at $x$, the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta}_x \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ with entries $\left(\boldsymbol{\delta}_x\right)_i = \febasis_i(x)$ for $i=1,\dots,N$.
+(A_h u_h)(v_h) := \int_\Omega \int_\Omega v_h(y) \Aker_h(y,x) u_h(x) dx dy
+\end{equation}
+for some kernel $\Aker_h$ which we do not know, and which may differ from $\Aker$ due to discretization error.
+Since the functions in $V_h$ are continuous at $x$, the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta}_x \in \mathbb{R}^\fedim_{\massmatrix^{-1}}$ with entries $\left(\boldsymbol{\delta}_x\right)_i = \febasis_i(x)$ for $i=1,\dots,\fedim$.
 Additionally, it is straightforward to verify that the Riesz representation, $\rho_h^* \in V_h$, of a functional $\rho \in V_h'$ has coefficient vector
 \begin{equation*}
-\boldsymbol{\rho}^* = \mathbf{M}^{-1} \boldsymbol{\rho}.
+\boldsymbol{\rho}^* = \massmatrix^{-1} \boldsymbol{\rho}.
 \end{equation*}
@@
 \label{eq:discrete_kernel}
-\boldsymbol{\impulseresponse}_x = \left(A_h \delta_x^*\right)^* =  \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x,
-\end{equation}
-and the $(y,x)$ kernel entry of $\Phi_h$ may be written as
+\boldsymbol{\impulseresponse}_x = \left(A_h \delta_x^*\right)^* =  \massmatrix^{-1}\mathbf{A} \massmatrix^{-1} \boldsymbol{\delta}_x,
+\end{equation}
+and the $(y,x)$ kernel entry of $\Aker_h$ may be written as
 \begin{equation}
 \label{eq:approx_kernel_entry_81}
-	\Phi_h(y,x) = \boldsymbol{\delta}_y^T \boldsymbol{\impulseresponse}_x = \boldsymbol{\delta}_y^T \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x.
-\end{equation}
-Now define $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ to be the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
+	\Aker_h(y,x) = \boldsymbol{\delta}_y^T \boldsymbol{\impulseresponse}_x = \boldsymbol{\delta}_y^T \massmatrix^{-1}\mathbf{A} \massmatrix^{-1} \boldsymbol{\delta}_x.
+\end{equation}
+Now define $\mathbf{\Aker} \in \mathbb{R}^{\fedim \times \fedim}$ to be the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
 \begin{equation}
 \label{eq:Akerpcmat_entries}
-\mathbf{\Phi}_{ij} := \Phi_h(p_i, p_j).
+\mathbf{\Aker}_{ij} := \Aker_h(p_i, p_j).
 \end{equation}
@@
 $
-	\Phi_h(p_i,p_j) = \left(\mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1}\right)_{ij},
+	\Aker_h(p_i,p_j) = \left(\massmatrix^{-1}\mathbf{A} \massmatrix^{-1}\right)_{ij},
 $
@@
 \label{eq:boldA}
-	\mathbf{A} = \bm{M} \mathbf{\Phi} \bm{M}.
-\end{equation}
-Broadly, our method will construct an H-matrix approximation of $\mathbf{A}$ by forming H-matrix approximations of $\boldsymbol{\Phi}$ and $\mathbf{M}$ (or a lumped mass version of $\mathbf{M}$), then multiplying these matrices with H-matrix methods per \eqref{eq:boldA}.
-The central challenge is that classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Phi}_{ij}$, but these matrix entries are not easily accessible.
+	\mathbf{A} = \massmatrix \mathbf{\Aker} \massmatrix.
+\end{equation}
+Broadly, our method will construct an H-matrix approximation of $\mathbf{A}$ by forming H-matrix approximations of $\boldsymbol{\Aker}$ and $\massmatrix$ (or a lumped mass version of $\massmatrix$), then multiplying these matrices with H-matrix methods per \eqref{eq:boldA}.
+The central challenge is that classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Aker}_{ij}$, but these matrix entries are not easily accessible.
 The bulk of our method is therefore dedicated to forming approximations of these matrix entries that can be evaluated rapidly.
 
-%\begin{align*}
-%	\text{continuous} \quad &\longrightarrow \quad \text{discrete} \\
-%	\delta_x \quad &\longrightarrow \quad \boldsymbol{\delta}_x \\
-%	\delta_x^* \quad &\longrightarrow \quad \mathbf{M}^{-1}\boldsymbol{\delta}_x \\
-%	\Aop \delta_x^* \quad &\longrightarrow \quad \mathbf{A}\mathbf{M}^{-1}\boldsymbol{\delta}_x \\
-%	\impulseresponse_x = \left(\Aop \delta_x^*\right)^* \quad &\longrightarrow \quad \mathbf{M}^{-1}\mathbf{A}\mathbf{M}^{-1}\boldsymbol{\delta}_x = \boldsymbol{\impulseresponse}_x \\
-%	\Phi(y,x) \quad &\longrightarrow \quad \boldsymbol{\delta}_y^T\mathbf{M}^{-1}\mathbf{A}\mathbf{M}^{-1}\boldsymbol{\delta}_x
-%\end{align*}
-%Since the Lagrange finite element nodes have the Kronecker property, that is $\febasis_i(p_i) = 1$ and $\febasis_i(p_j) = 1$ for $i \neq j$, the vector $\boldsymbol{\delta}_{p_i}$ is the unit vector with one in the $i$'th component and zeros in all other components.
-Hence
-%\begin{equation*}
-%	\Phi(p_i,p_j) = \left(\mathbf{M}^{-1}\mathbf{A}\mathbf{M}^{-1}\right)_{ij}
-%\end{equation*}
-
-%\subsubsection{Discretized integral kernel}
-%
-%The interpolation of $\Phi$ onto $V_h \otimes V_h$, which we denote by $\Phi^h$, is given as follows:
-%\begin{equation}
-%\label{eq:defn_of_Akerpcmesh}
-%\Phi(y,x) \approx \Phi^h(y,x) := \sum_{i=1}^\fedim \sum_{j=1}^\fedim \Phi(p_i,p_j) \febasis_i(y) \febasis_j(x).
-%\end{equation}
-%Replacing $\Aker$ with $\Phi^h$ within the definition of $\Aop$ in \eqref{eq:kernel_representation}, then performing straightforward algebraic manipulations, yields the following approximation:
-%\begin{equation}
-%\label{eq:kernel_representation_pcmesh}
-%(\Aop u_h)(v_h) \approx \int_\Omega \int_\Omega v_h(y) \Phi^h(y,x) u_h(x) dx dy
-%= \bm{v}^T \bm{M} \mathbf{\Phi} \bm{M} \bm{u},
-%\end{equation}
-%where $\mathbf{\Phi} \in \mathbb{R}^{\fedim \times \fedim}$ is the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
-%\begin{equation}
-%\label{eq:Akerpcmat_entries}
-%\mathbf{\Phi}_{ij} := \Phi(p_i, p_j),
-%\end{equation}
-%Based on \eqref{eq:kernel_representation_pcmesh}, we define the following matrix approximation, $\widetilde{\mathbf{A}} \approx \mathbf{A}$ as follows:
-%\begin{equation}
-%\label{eq:Amatpc_defn}
-%\widetilde{\mathbf{A}} = \bm{M} \mathbf{\Phi} \bm{M}.
-%\end{equation}
-%Note that $\widetilde{\mathbf{A}}$ differs slightly from $\mathbf{A}$ (the Galerkin projection of $\Aop$ onto $V_h$) because of the approximation $\Phi^h \approx \Phi$. 
-%%Hence, $\mathbf{A}$ may differ slightly from the implicitly defined matrix that performs the mapping $\mathbf{u} \rightarrow \mathbf{v}$ corresponding to the discrete version of the operator action $v=\Aop u$.
-%%Because $\Aop$ is only available matrix-free and $\Phi$ is only defined implicitly in terms of $\Aop$, 
-%The matrices $\widetilde{\mathbf{A}}$ and $\mathbf{\Phi}$ are, of course, not easily accessible.  
-%%The primary goal of this paper is building an H-matrix approximation $\mathbf{A}_H$ of $\widetilde{\mathbf{A}}$, using a small number of operator actions of $\Aop$ and $\Aop^T$.
-%
-%%After discretization with finite elements\footnote{While we use finite elements, other discretization schemes such as finite differences or finite volumes could be used instead, with minor modifications to our method.} (Section \ref{sec:finite_element_kernel}), $\Aop$ and $\Phi$ become large dense matrices $\mathbf{A}$ and $\mathbf{\Phi}$, respectively.
-
-%\subsubsection{Lumped mass matrix and discrete kernel non-negativity}
-
-\paragraph{Lumped mass matrix} At the continuum level, the integral kernel $\Phi$ is assumed to be non-negative.
+\paragraph{Lumped mass matrix} At the continuum level, the integral kernel $\Aker$ is assumed to be non-negative.
 However, when we compute impulse responses we are effectively operating with the discrete integral kernel shown in \eqref{eq:approx_kernel_entry_81}.
-%,  which has $(y,x)$ entry
-%\begin{equation*}
-%\boldsymbol{\delta}_y^T \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x.
-%\end{equation*}
-%Even if the discretization $\Aop \rightarrow \mathbf{A}$ preserves integral kernel non-negativity (which it typically will not), 
-The inverse mass matrices in this formula, $\mathbf{M}^{-1}$, will typically contain negative numbers.
+The inverse mass matrices in this formula, $\massmatrix^{-1}$, will typically contain negative numbers.
 If there are too many negative numbers, or if the negative numbers are too large, our algorithm will be less robust.
-We therefore recommend replacing the mass matrix $\mathbf{M}$ with a positive diagonal \emph{lumped mass} approximation.
-%For typical finite element meshes, lumped mass matrix approximations are spectrally equivalent to the mass matrix \cite{MASSLUMP2}. 
+We therefore recommend replacing the mass matrix $\massmatrix$ with a positive diagonal \emph{lumped mass} approximation.
 In our numerical results, we use the lumped mass matrix constructed by replacing off-diagonal entries of the mass matrix by zero.
@@
 In this section we present two key innovations that our method is based on.
-%The approximation algorithm will be presented in Section \ref{sec:method}. 
 First, we define moments of the impulse responses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{sec:intromoments}).
@@
 \end{equation}
-denote the spatially varying scaling factor, and for $i,j=1,\dots,d$ define $\mu:\Omega \rightarrow \mathbb{R}^d$ and $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$ as follows:
+denote the spatially varying scaling factor, and for $i,j=1,\dots,\gdim$ define $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$ and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$ as follows:
 \begin{align}
 \spatialmean^i(x) :=& \int_\Omega (\impulseresponse_p(y) / V(x)) y^i ~dy \label{eq:define_mean} \\
-\spatialcov^{ij}(x) :=& \int_\Omega (\impulseresponse_p(y) / V(x)) \left(y^i - \mu^i(x)\right) \left(y^j - \mu^j(x)\right) ~dy, \label{eq:define_cov}
+\spatialcov^{ij}(x) :=& \int_\Omega (\impulseresponse_p(y) / V(x)) \left(y^i - \spatialmean^i(x)\right) \left(y^j - \spatialmean^j(x)\right) ~dy, \label{eq:define_cov}
 \end{align}
-where $\mu^i(x)$ denotes the $i^\text{th}$ component of the vector $\mu(x)$, and $\Sigma^{ij}(x)$ denotes the $(i,j)$ entry of the matrix $\Sigma(x)$.
-The vector $\mu(x)\in \mathbb{R}^d$ and the matrix $\Sigma(x) \in \mathbb{R}^{d \times d}$ are the mean and covariance of the normalized version of $\impulseresponse_x$, respectively. 
-%Imagine $\phi_x$ as a mound of dirt piled on a localized region in $\Omega$, with $\phi_x(y)$ being the height of the dirt mound at location $y$.
-The scalar $V(x)$ is the total volume of dirt in the mound, the vector $\mu(x)$ is the center of mass of the dirt mound, and the matrix $\Sigma(x)$ contains information about how wide the dirt mound is in each spatial direction.
+where $\spatialmean^i(x)$ denotes the $i^\text{th}$ component of the vector $\spatialmean(x)$, and $\spatialcov^{ij}(x)$ denotes the $(i,j)$ entry of the matrix $\spatialcov(x)$.
+The vector $\spatialmean(x)\in \mathbb{R}^\gdim$ and the matrix $\spatialcov(x) \in \mathbb{R}^{\gdim \times \gdim}$ are the mean and covariance of the normalized version of $\impulseresponse_x$, respectively.
 
@@
 Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
-However, this direct approach is not feasible because we need to know $V(x)$, $\mu(x)$, and $\Sigma(x)$ before we compute $\phi_x$, in order choose the point $x$.
-Computing $\phi_x$ in order to determine $V(x)$, $\mu(x)$, and $\Sigma(x)$ would be extremely computationally expensive, and defeat the purpose of our algorithm, which is to reduce the computational cost by computing impulse responses in batches.
+However, this direct approach is not feasible because we need to know $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ before we compute $\phi_x$, in order choose the point $x$.
+Computing $\phi_x$ in order to determine $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ would be extremely computationally expensive, and defeat the purpose of our algorithm, which is to reduce the computational cost by computing impulse responses in batches.
 Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ indirectly, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions (e.g., 6 total operator applies in two spatial dimensions and 10 in three spatial dimensions).
 This may be motivated by analogy to matrices.
-If $\mathbf{B}\in \mathbb{R}^{N \times N}$ is a matrix with $i^\text{th}$ column $\mathbf{b}_i$ and $\mathbf{f} \in \mathbb{R}^N$, then
+If $\mathbf{B}\in \mathbb{R}^{\fedim \times \fedim}$ is a matrix with $i^\text{th}$ column $\mathbf{b}_i$ and $\mathbf{f} \in \mathbb{R}^\fedim$, then
 \begin{equation*}
@@
 & \vdots & \\
-\horzbar & \mathbf{b}_N^T & \horzbar
+\horzbar & \mathbf{b}_\fedim^T & \horzbar
 \end{bmatrix}
@@
 \vdots \\
-\mathbf{b}_N^T \mathbf{f}
+\mathbf{b}_\fedim^T \mathbf{f}
 \end{bmatrix}.
@@
 \end{equation*}
-for $i,j=1,\dots,d$. Using the definition of $V$ in \eqref{eq:define_vol} and using \eqref{eq:operator_simultaneous_ips}, we have
-\begin{equation*}
-	V(x) = \int_\Omega \phi_x(y) C(y) ~ dy = \left(\phi_x, C\right)_{L^2(\Omega)} = \left(\mathcal{A}^T C\right)^*(x).
-\end{equation*}
-Hence, we may compute $V(x)$ for all $x$ simultaneously by applying $\Aop^T$ to $C$.
-Analogous manipulations show that $\mu(x)$ and $\Sigma(x)$ may be computed for all points $x$ simultaneously by applying $\Aop^T$ to the functions $L^i$ and $Q^{ij}$, respectively.
+for $i,j=1,\dots,\gdim$. Using the definition of $\spatialvol$ in \eqref{eq:define_vol} and using \eqref{eq:operator_simultaneous_ips}, we have
+\begin{equation*}
+	\spatialvol(x) = \int_\Omega \phi_x(y) C(y) ~ dy = \left(\phi_x, C\right)_{L^2(\Omega)} = \left(\mathcal{A}^T C\right)^*(x).
+\end{equation*}
+Hence, we may compute $\spatialvol(x)$ for all $x$ simultaneously by applying $\Aop^T$ to $C$.
+Analogous manipulations show that $\spatialmean(x)$ and $\spatialcov(x)$ may be computed for all points $x$ simultaneously by applying $\Aop^T$ to the functions $L^i$ and $Q^{ij}$, respectively.
 To summarize the results of doing this, we have
@@
 Here $f/g$ denotes pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denotes pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
-The resulting method for computing discretized versions of $V$, $\mu$, and $\Sigma$ is shown in Algorithm \ref{alg:varhpi_mean_cov}.
+The resulting method for computing discretized versions of $\spatialvol$, $\spatialmean$, and $\spatialcov$ is shown in Algorithm \ref{alg:varhpi_mean_cov}.
 
@@
 
-		Form vector $\bm{C} \in \mathbb{R}^\fedim_{\bm{M}}$ by either projecting or interpolating constant function $C(x)=1$ onto $V_h$
-
-		$\bm{\spatialvol} \gets \bm{M}^{-1}\bm{A}^T \bm{C}$
+		Form vector $\bm{C} \in \mathbb{R}^\fedim_{\massmatrix}$ by either projecting or interpolating constant function $C(x)=1$ onto $V_h$
+
+		$\bm{\spatialvol} \gets \massmatrix^{-1}\bm{A}^T \bm{C}$
 
 		\For{$i=1,2,\dots,\gdim$}{
-			Form vector $\bm{L}^i \in \mathbb{R}^\fedim_{\bm{M}}$ by either projecting or interpolating linear function $L^i(x) = x^i$ onto $V_h$
-
-			$\bm{\spatialmean}^i \gets \left(\bm{M}^{-1}\bm{A}^T \bm{L}^i\right) / \bm{\spatialvol}$
+			Form vector $\bm{L}^i \in \mathbb{R}^\fedim_{\massmatrix}$ by either projecting or interpolating linear function $L^i(x) = x^i$ onto $V_h$
+
+			$\bm{\spatialmean}^i \gets \left(\massmatrix^{-1}\bm{A}^T \bm{L}^i\right) / \bm{\spatialvol}$
 		}
@@
 
-				$\bm{\spatialcov}^{ij} \gets \left(\bm{M}^{-1}\bm{A}^T \bm{Q}^{ij}\right) / \bm{\spatialvol} - \bm{\spatialmean}^i\cdot \bm{\spatialmean}^j$
+				$\bm{\spatialcov}^{ij} \gets \left(\massmatrix^{-1}\bm{A}^T \bm{Q}^{ij}\right) / \bm{\spatialvol} - \bm{\spatialmean}^i\cdot \bm{\spatialmean}^j$
 
@@
 	}
-	\caption{Compute discretized scaling factor $\mathbf{V}$, mean $\boldsymbol{\mu}$, and covariance $\boldsymbol{\Sigma}$}
+	\caption{Compute discretized scaling factor $\mathbf{\spatialvol}$, mean $\boldsymbol{\spatialmean}$, and covariance $\boldsymbol{\spatialcov}$}
 	\label{alg:varhpi_mean_cov}
 \end{algorithm2e}
-
-%\begin{algorithm2e}
-%	\SetAlgoNoLine
-%	\SetKwInOut{Input}{Input}
-%	\SetKwInOut{Output}{Output}
-%	{	
-%		\Input{Operator $\Aop$}
-%		\Output{$\spatialvol$, $\spatialmean$, and $\spatialcov$}
-%		
-%		\tcp{Compute scaling factor $\spatialvol$}
-%		
-%		Form constant function $C(x)=1$
-%		
-%		Compute $\spatialvol = \left(\Aop^T C\right)^*$
-%		
-%		\tcp{Compute mean $\spatialmean$}
-%		\For{$i=1,2,\dots,\gdim$}{
-%			Form linear function $L^i(x) = x^i$
-%			
-%			Compute $\spatialmean^i = \left(\Aop^T L^i\right)^* / \spatialvol$
-%		}
-%		\tcp{Compute covariance $\spatialcov$}
-%		\For{$i=1,2,\dots,\gdim$}{
-%			\For{$j=1,\dots,i$}{
-%				Form quadratic function $Q^{ij}(x) = x^i x^j$
-%				
-%				Compute $\spatialcov^{ij} = \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j$
-%				
-%				Set $\spatialcov^{ji} = \spatialcov^{ij}$
-%				
-%			}
-%		}
-%		
-%	}
-%	\caption{Compute scaling factor $\spatialvol$, mean $\spatialmean$, and covariance $\spatialcov$}
-%	\label{alg:varhpi_mean_cov}
-%\end{algorithm2e}
-
-%\subsection{Impulse response support ellipsoids}
-%\label{eq:mean_and_covariance_estimation}
 
@@
 \phi_x(y) &\approx \phi_{x'}(y-x+x') \label{eq:translate2}\\
-\phi_x(y) &\approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right) \label{eq:translate3}\\
-\phi_x(y)/V(x) &\approx \phi_{x'}\left(y-\mu(x)+\mu(x')\right) / V(x'). \label{eq:translate4}
+\phi_x(y) &\approx \phi_{x'}\left(y-\spatialmean(x)+\spatialmean(x')\right) \label{eq:translate3}\\
+\phi_x(y)/\spatialvol(x) &\approx \phi_{x'}\left(y-\spatialmean(x)+\spatialmean(x')\right) / \spatialvol(x'). \label{eq:translate4}
 \end{align}
@@
 
-Operators that satisfy \eqref{eq:translate2} are called ``locally translation-invariant'' because integral kernel entries $\Phi(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$.
+Operators that satisfy \eqref{eq:translate2} are called ``locally translation-invariant'' because integral kernel entries $\Aker(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$.
 It is straightforward to show that if equality holds in \eqref{eq:translate2}, then $\Aop$ is a convolution operator.
@@
 Approximation \eqref{eq:translate3} improves upon \eqref{eq:translate1} and \eqref{eq:translate2}, and generalizes both.
-On one hand, if \eqref{eq:translate1} holds, then $\mu(x) \approx \mu(x')$, and so \eqref{eq:translate3} holds.
-On the other hand, translating a distribution translates its mean, so if \eqref{eq:translate2} holds, then $\mu(x')-\mu(x) \approx x' - x$, so again \eqref{eq:translate3} holds.
+On one hand, if \eqref{eq:translate1} holds, then $\spatialmean(x) \approx \spatialmean(x')$, and so \eqref{eq:translate3} holds.
+On the other hand, translating a distribution translates its mean, so if \eqref{eq:translate2} holds, then $\spatialmean(x')-\spatialmean(x) \approx x' - x$, so again \eqref{eq:translate3} holds.
 But approximation \eqref{eq:translate3} can hold in situations where neither \eqref{eq:translate1} nor \eqref{eq:translate2} holds.
@@
 
-We use approximation \eqref{eq:translate4}, which is the same as \eqref{eq:translate3}, except for the extra factors of $1/V(x)$ on the left hand side and $1/V(x')$ on the right hand side.
-These factors make the approximation more robust if $V(x)$ varies widely.
-Approximation \eqref{eq:translate4} is equivalent to \eqref{eq:translate3}, but with $\phi_x$ replaced by the its normalized version, $\phi_x/V(x)$.
+We use approximation \eqref{eq:translate4}, which is the same as \eqref{eq:translate3}, except for the extra factors of $1/\spatialvol(x)$ on the left hand side and $1/\spatialvol(x')$ on the right hand side.
+These factors make the approximation more robust if $\spatialvol(x)$ varies widely.
+Approximation \eqref{eq:translate4} is equivalent to \eqref{eq:translate3}, but with $\phi_x$ replaced by the its normalized version, $\phi_x/\spatialvol(x)$.
 We call \eqref{eq:translate4} ``normalized local mean-displacement invariance.''
@@
 
-%In this section, we present our method for constructing accessible H-matrix approximations, $\mathbf{A}_H$ and $\mathbf{\Phi}_H$ of $\mathbf{A}$ and $\mathbf{\Phi}$, respectively, using a small number of operator actions of $\Aop$ and $\Aop^T$.
-
-Algorithm \ref{alg:varhpi_mean_cov} from Section \ref{sec:intromoments} allows us to compute $V$, $\mu$, and $\Sigma$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} uses $V$, $\mu$, and $\Sigma$ to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
+Algorithm \ref{alg:varhpi_mean_cov} from Section \ref{sec:intromoments} allows us to compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} uses $\spatialvol$, $\spatialmean$, and $\spatialcov$ to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
 This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}).
@@
 
-Once the impulse response batches $\eta_b$ are computed, we approximate the integral kernel $\Phi(y,x)$ at arbitrary points $(y,x)$ by interpolation (Section \ref{sec:approximate_kernel_entries}).
+Once the impulse response batches $\eta_b$ are computed, we approximate the integral kernel $\Aker(y,x)$ at arbitrary points $(y,x)$ by interpolation (Section \ref{sec:approximate_kernel_entries}).
 The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section \ref{sec:local_mean_displacement_invariance}.
-We approximate $\Phi(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{V(x)}{V(x_i)}\phi_{x_i}(y - \mu(x) + \mu(x_i))$ for a collection of sample points $x_i$ near $x$.
+We approximate $\Aker(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}(y - \spatialmean(x) + \spatialmean(x_i))$ for a collection of sample points $x_i$ near $x$.
 The weights are determined by radial basis function interpolation.
 
-%If $\Aop$ is symmetric, we take advantage of symmetry to construct a more accurate approximation at negligible additional cost by combining information from impulse responses of both $\Aop$ and $\Aop^T$ in the interpolation procedure (Section \ref{sec:symmetric_improvements}).
-Intuitively, impulse responses of $\Aop$ can be thought of as ``columns'' of $\Phi$, impulse responses of $\Aop^T$ can be thought of as ``rows'' of $\Phi$, and $\Phi(y,x)$ can be thought of as an unknown entry of $\Phi$.
-Therefore, when $\Aop$ is symmetric, computing a column of $\Aop$ also gives us a row of $\Aop$ for free, so we approximate $\Phi(y,x)$ using information from a combination of known columns and rows, rather than just known columns.
-
-The ability to rapidly evaluate approximate kernel entries $\Phi(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Phi}_H \approx \mathbf{\Phi}$, using the conventional adaptive cross H-matrix construction method (Section \ref{sec:h_matrix_construction_short}).
+The ability to rapidly evaluate approximate kernel entries $\Aker(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Aker}_H \approx \mathbf{\Aker}$, using the conventional adaptive cross H-matrix construction method (Section \ref{sec:h_matrix_construction_short}).
 In this method, one forms low-rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
-We then use H-matrix methods to convert $\boldsymbol{\Phi}_H$ into an H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$.
+We then use H-matrix methods to convert $\boldsymbol{\Aker}_H$ into an H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$.
 
@@
 \begin{equation*}
-\Aop \longrightarrow V, \mu, \Sigma \longrightarrow \underbrace{\{x_i\}_{i=1}^r}_{\text{as }S_b\text{'s}} \longrightarrow \underbrace{\{\impulseresponse_{x_i}\}_{i=1}^r}_{\text{as }\eta_b\text{'s}} \longrightarrow \boldsymbol{\Phi}_H \longrightarrow \mathbf{A}_H.
+\Aop \longrightarrow \spatialvol, \spatialmean, \spatialcov \longrightarrow \underbrace{\{x_i\}_{i=1}^r}_{\text{as }S_b\text{'s}} \longrightarrow \underbrace{\{\impulseresponse_{x_i}\}_{i=1}^r}_{\text{as }\eta_b\text{'s}} \longrightarrow \boldsymbol{\Aker}_H \longrightarrow \mathbf{A}_H.
 \end{equation*}
@@
 	{	
-		\Input{Linear operator $\Aop$, parameter $\text{max\_batches}$}
+		\Input{Linear operator $\Aop$, parameter $\nbatch$}
 		\Output{H-matrix $\mathbf{A}_H$ (optionally, $\mathbf{A}_H^{\text{sym}+}$)}
@@
 
-		\For{$k=1,2,\dots,\text{max\_batches}$}{
+		\For{$k=1,2,\dots,\nbatch$}{
 			Choose a batch of sample points, $\pointbatch_k$, using Algorithm \ref{alg:point_choice}
@@
 
-		Form H-matrix approximation $\mathbf{\Phi}_H$ of integral kernel (Section \ref{sec:h_matrix_construction_short})
+		Form H-matrix approximation $\boldsymbol{\Aker}_H$ of integral kernel (Section \ref{sec:h_matrix_construction_short})
 
@@
 	\item We check whether $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_k$.
-	\item We make sure that $V(p)$ is not too small, by checking whether $V(p) > \epsilon V_\text{max}$.
-Here $V_\text{max}$ is the largest value of $V(q)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\epsilon=10^{-5}$).
+	\item We make sure that $\spatialvol(p)$ is not too small, by checking whether $\spatialvol(p) > \epsilon \spatialvol_\text{max}$.
+Here $\spatialvol_\text{max}$ is the largest value of $\spatialvol(q)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\epsilon=10^{-5}$).
 \end{enumerate}
-If $p$ passes both these checks (i.e., if $p$ is sufficiently far from other points in the batch, and $V(p)$ is not too small) then we add $p$ to $\pointbatch_k$.
+If $p$ passes both these checks (i.e., if $p$ is sufficiently far from other points in the batch, and $\spatialvol(p)$ is not too small) then we add $p$ to $\pointbatch_k$.
 Otherwise we discard $p$.
@@
 
-We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots$, until the number of batches exceeds a desired threshold.
+We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots$, until the number of batches equals t desired threshold $\nbatch$.
 In our implementation, for each batch the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, except for points in previously chosen batches.
@@
 \begin{equation*}
-	\diraccomb_k := \sum_{x_i \in \pointbatch_k} \delta_{x_i} / V(x_i).
+	\diraccomb_k := \sum_{x_i \in \pointbatch_k} \delta_{x_i} / \spatialvol(x_i).
 \end{equation*}
@@
 	\combresponse_k := \left(\Aop \diraccomb_k^*\right)^*
-	=\sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / V(x_i).
+	=\sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / \spatialvol(x_i).
 \end{equation}
@@
 	\begin{cases}
-		\combresponse_k(z) V(x_i), & z \in E_{x_i} \\
+		\combresponse_k(z) \spatialvol(x_i), & z \in E_{x_i} \\
 		0, & \text{otherwise}
@@
 
-Each point source $\delta_{x_i}$ is scaled by $1/V(x_i)$ so that the resulting scaled impulse responses within $\eta_k$ are comparable in magnitude.
-Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section \ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $V(x_i)$ is smaller than a predetermined threshold.
-Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, may overwhelm $\phi_{x_j}$ for a nearby point $x_j$ if $V(x_i)$ is much larger than $V(x_j)$.
-
-\subsection{Approximate integral kernel entries $\widetilde{\Phi}(y,x)$}
+Each point source $\delta_{x_i}$ is scaled by $1/\spatialvol(x_i)$ so that the resulting scaled impulse responses within $\eta_k$ are comparable in magnitude.
+Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section \ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $\spatialvol(x_i)$ is smaller than a predetermined threshold.
+Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, may overwhelm $\phi_{x_j}$ for a nearby point $x_j$ if $\spatialvol(x_i)$ is much larger than $\spatialvol(x_j)$.
+
+\subsection{Approximate integral kernel entries}
 \label{sec:approximate_kernel_entries}
@@
 \begin{equation*}
-	z_i := y - \mu(x) + \mu(x_i), \qquad i=1,\dots,k,
-\end{equation*}
-where $\{x_i\}_{i=1}^k$ are the $k$ nearest sample points to $x$, excluding sample points $x_i$ for which $z_i \notin \Omega$. Here $k$ is a small user defined parameter; e.g., $k=10$. Further, let
+	z_i := y - \spatialmean(x) + \spatialmean(x_i), \qquad i=1,\dots,\numnbr,
+\end{equation*}
+where $\{x_i\}_{i=1}^{\numnbr}$ are the $\numnbr$ nearest sample points to $x$, excluding sample points $x_i$ for which $z_i \notin \Omega$.
+Here $\numnbr$ is a small user defined parameter; e.g., $\numnbr=10$.
+Further, let
 \begin{equation}
 \label{eq:fxyxp}
-f_i := \frac{V(x)}{V(x_i)}\phi_{x_i}\left(z_i\right), \qquad i=1,\dots,k,
-\end{equation}
-Note that $f_i$ is well-defined, because $\phi_{x_i}\left(z_i\right)$ is defined if and only if $z_i \in \Omega$, and $V(x_i)>0$ by our sample point picking procedure in Section \ref{sec:sample_point_selection}.
-We approximate $\Phi(y,x)$ by interpolating the (point,value) pairs
-\begin{equation}
-\left(x_i, f_i\right), \qquad i=1,\dots,k,
-\end{equation}
-at the point $x$. The idea is that $\Phi(y,x) = f_i$ per the discussion in Section \ref{sec:local_mean_displacement_invariance}, so we approximate $\Phi(y,x)$ by interpolating the values $f_i$.
-
-To find the $k$ nearest sample points to $x$, we query a precomputed k-d tree \cite{Bentley75} of all sample points.
+f_i := \frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}\left(z_i\right), \qquad i=1,\dots,\numnbr,
+\end{equation}
+Note that $f_i$ is well-defined, because $\phi_{x_i}\left(z_i\right)$ is defined if and only if $z_i \in \Omega$, and $\spatialvol(x_i)>0$ by our sample point picking procedure in Section \ref{sec:sample_point_selection}.
+We approximate $\Aker(y,x)$ by interpolating the (point,value) pairs
+\begin{equation}
+\left(x_i, f_i\right), \qquad i=1,\dots,\numnbr,
+\end{equation}
+at the point $x$. The idea is that $\Aker(y,x) = f_i$ per the discussion in Section \ref{sec:local_mean_displacement_invariance}, so we approximate $\Aker(y,x)$ by interpolating the values $f_i$.
+
+To find the $\numnbr$ nearest sample points to $x$, we query a precomputed k-d tree \cite{Bentley75} of all sample points.
 We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{Ericson04} of the mesh cells used to discretize the problem.
@@
 \begin{equation}
-\Phi(y,x) \approx \widetilde{\Phi}(y,x) := \sum_{i=1}^k w_i~ b\left(\|x-x_i\|\right),
-\end{equation}
-where $b_i$ are radial basis functions and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_k)^T$, is found as the solution to the $k \times k$ linear system
+\Aker(y,x) \approx \widetilde{\Aker}(y,x) := \sum_{i=1}^{\numnbr} w_i~ b\left(\|x-x_i\|\right),
+\end{equation}
+where $b_i$ are radial basis functions and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_{\numnbr})^T$, is found as the solution to the $\numnbr \times \numnbr$ linear system
 \begin{equation}
@@
 \end{equation}
-where $B \in \mathbb{R}^{k \times k}$ has entries
+where $B \in \mathbb{R}^{\numnbr \times \numnbr}$ has entries
 $
@@
 $
-and $f \in \mathbb{R}^k$ has entries $f_i$ from \eqref{eq:fxyxp}. We use polyharmonic spline radial basis functions, which are defined as follows:
+and $f \in \mathbb{R}^\numnbr$ has entries $f_i$ from \eqref{eq:fxyxp}. We use polyharmonic spline radial basis functions, which are defined as follows:
 \begin{equation}
 b_i(r) := \begin{cases}
-r^d, & d=1,3,5,\dots \\
-r^d \log r, & d=2,4,6,\dots
+r^\gdim, & \gdim=1,3,5,\dots \\
+r^\gdim \log r, & \gdim=2,4,6,\dots
 \end{cases}
@@
 	\includegraphics[scale=0.2]{interpolation_rbf_boundary.pdf}
-%	\begin{subfigure}{0.49\textwidth}
-%		\centering
-%		\includegraphics[scale=0.22]{interpolation_rbf.pdf}
-%		\caption{}
-%	\end{subfigure}
-%	\begin{subfigure}{0.49\textwidth}
-%		\centering
-%		\includegraphics[scale=0.22]{interpolation_rbf_boundary.pdf}
-%		\caption{}
-%	\end{subfigure}
-	\caption{To compute the kernel entry $\Phi(y,x)$, we use radial basis functions to interpolate the values $\frac{V(x)}{V(x_i)}\phi_{x_i}(y-\mu(x)-\mu(x_i))$ for the $k$ nearest sample points $x_i$ to $x$, excluding sample points for which $y-\mu(x)-\mu(x_i)$ is outside of the domain.}
+	\caption{To compute the kernel entry $\Aker(y,x)$, we use radial basis functions to interpolate the values $\frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}(y-\spatialmean(x)-\spatialmean(x_i))$ for the $\numnbr$ nearest sample points $x_i$ to $x$, excluding sample points for which $y-\spatialmean(x)-\spatialmean(x_i)$ is outside of the domain.}
 	\label{fig:rbf_interpolation}
@@
 \begin{equation}
-f_i = V(x)\eta_b\left(z_i\right),
+f_i = \spatialvol(x)\eta_b\left(z_i\right),
 \end{equation}
@@
 
-We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming H-matrix representations $\mathbf{\Phi}_H$ and $\bm{M}_H$ of $\mathbf{\Phi}$ and $\bm{M}$, respectively, then using fast H-matrix methods to multiply these matrices per \eqref{eq:boldA} to form
-\begin{equation*}
-\mathbf{A}_H = \bm{M}_H \mathbf{\Phi}_H \bm{M}_H.
+We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming H-matrix representations $\mathbf{\Aker}_H$ and $\massmatrix_H$ of $\mathbf{\Aker}$ and $\massmatrix$, respectively, then using fast H-matrix methods to multiply these matrices per \eqref{eq:boldA} to form
+\begin{equation*}
+\mathbf{A}_H = \massmatrix_H \mathbf{\Aker}_H \massmatrix_H.
 \end{equation*}
@@
 
-We form $\mathbf{\Phi}_H$ using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{Kriemann08}.
-Although $\mathbf{\Phi}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Phi}_H$ only requires evaluation of $O(\hrank^2 \fedim \log \fedim)$ entries $\mathbf{\Phi}_{ij} = \widetilde{\Phi}(p_i, p_j)$, and these entries are computed via the procedure described in Section \ref{sec:approximate_kernel_entries}.
+We form $\mathbf{\Aker}_H$ using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{Kriemann08}.
+Although $\mathbf{\Aker}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Aker}_H$ only requires evaluation of $O(\hrank^2 \fedim \log \fedim)$ entries $\mathbf{\Aker}_{ij} = \widetilde{\Aker}(p_i, p_j)$, and these entries are computed via the procedure described in Section \ref{sec:approximate_kernel_entries}.
 Here $\hrank$ is the rank of the highest-rank block in the H-matrix.
-A dense representation of $\mathbf{\Phi}$ is never formed.
+A dense representation of $\mathbf{\Aker}$ is never formed.
 We describe this H-matrix construction process in detail in Appendix \ref{app:h_matrix}.
-We form $\bm{M}_H$ using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same recursive block partitioning structure as was used for $\mathbf{\Phi}$.
-
-\subsection{Computational cost}
+We form $\massmatrix_H$ using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same recursive block partitioning structure as was used for $\mathbf{\Aker}$.
+
+\section{Computational cost}
 \label{sec:overall_cost}
@@
 \begin{enumerate}
-	\item Computing the impulse response moments and batches. This includes picking the sample point batches via the greed ellipsoid packing procedure.
+	\item Computing the impulse response moments and batches.
 	\item Constructing the H-matrix $\mathbf{A}_H$.
-	\item Using $\mathbf{A}_H$ to perform further linear algebra operations. This includes, optionally, modifying $\mathbf{A}_H$ to make it symmetric positive semi-definite.
+	\item Using $\mathbf{A}_H$ to perform further linear algebra operations. This may optionally include modifying $\mathbf{A}_H$ to make it symmetric positive semi-definite.
 \end{enumerate}
 In practice, 1 is the dominant cost, because 1 is the only task that requires applying $\Aop$ to vectors, and our method is targeted at applications in which applying $\Aop$ to a vector requires an expensive computational procedure such as solving a large-scale partial differential equation.
-Hence, our method was designed to construct an approximation $\mathbf{A}_H$ that is as accurate as possible, using as few operator applies as possible, subject to the constraint that the further costs associated with building and using $\mathbf{A}_H$ are near-linear (and therefore scalable) in $N$.
-We now describe these costs in detail.
-For convenience, variable symbols and their approximate sizes are listed in Table \ref{tab:vars}.
+All operations that do not require applications of $\Aop$ to vectors are nearly-linear, and therefore scalable, in $N$.
+We now describe these costs in detail. For convenience, Table \ref{tab:vars} lists variable symbols and their approximate sizes.
 
@@
 		\hline
-		$N$ & $10^2$--$10^9$ & Number of finite element degrees of freedom \\
-		$n_b$ & $1$--$25$ & Number of impulse response batches \\
+		$\fedim$ & $10^2$--$10^9$ & Number of finite element degrees of freedom \\
+		$\nbatch$ & $1$--$25$ & Number of impulse response batches \\
 		$\hrank$ & $5$--$50$ & H-matrix rank \\
-		$k_n$ & $5$--$15$ & Number of nearest neighbors for RBF interpolation \\
-		$d$ & $1$--$3$ & Spatial dimension \\
-		$r$ & $10^1$--$10^4$ & Classical matrix rank \\
-		$m$ & $10^1$--$10^4$ & Total number of sample points (all batches) \\
-		$s$ & $1$--$500$ & Number of sample points in one batch \\
-		$l$ & $1$--$2$ & Rational function parameter for SPSD modification
+		$\numnbr$ & $5$--$15$ & Number of nearest neighbors for RBF interpolation \\
+		$\gdim$ & $1$--$3$ & Spatial dimension \\
+%		$\classicalrank$ & $10^1$--$10^4$ & Classical matrix rank \\
+		$\nsamplepts$ & $10^1$--$10^4$ & Total number of sample points (all batches) \\
+		$|\pointbatch_i|$ & $1$--$500$ & Number of sample points in the $i$th batch \\
+		$\ratord$ & $1$--$2$ & Rational function parameter for SPSD modification
 	\end{tabular}
-	\caption{Symbols used for variables in computational cost estimates, and their approximate sizes.}
+	\caption{Symbols used for variables in computational cost estimates, and approximate ranges for their sizes in practice.}
 	\label{tab:vars}
@@
 
-\paragraph{1: Impulse response moments and batches} Computing $V$, $\mu$, and $\Sigma$ requires applying $\Aop$ to $1$, $d$, and $d(d+1)/2$ vectors, respectively.
-Concretely, in one spatial dimensions ($d=1$), computing $V$, $\mu$, and $\Sigma$ requires $1+1+1=3$ operator applies.
-In two spatial dimensions, this cost is $1+2+3=6$ operator applies.
-In three spatial dimensions, this cost is $1+3+6=10$ operator applies.
-Computing each impulse response batch requires applying $\Aop$ to $1$ vector, so computing the impulse response batches $\{\eta_i\}_{i=1}^{n_b}$ requires $n_b$ operator applies.
+\paragraph{1: Impulse response moments and batches} Computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ requires applying $\Aop$ to $1$, $\gdim$, and $\gdim(\gdim+1)/2$ vectors, respectively.
+Computing each impulse response batch requires applying $\Aop$ to $1$ vector, so computing the impulse response batches $\{\eta_i\}_{i=1}^{\nbatch}$ requires $\nbatch$ operator applies.
 In total, computing the impulse response moments and batches therefore requires
 \begin{equation*}
-	1 + \gdim + \gdim(\gdim+1)/2 + n_b \qquad \text{operator applies.}
-\end{equation*}
-In a typical application one might have $d=2$ and $n_b=5$, in which case $11$ operator applies would be required.
-In contrast, matrix-free low-rank approximation methods require $O(r)$ operator applies, where $r$ is the classical numerical rank of $\Aop$.
-Since $r$ may be in the thousands, low rank approximation may require thousands of operator applies.
+	1 + \gdim + \gdim(\gdim+1)/2 + \nbatch \qquad \text{operator applies.}
+\end{equation*}
+In a typical application one might have $\gdim=2$ and $\nbatch=5$, in which case $11$ operator applies are required.
 
 Computing the impulse response batches also requires choosing sample point batches via the greedy ellipsoid packing algorithm described in Section \ref{sec:sample_point_selection}.
-Choosing one batch of sample points may require performing up to $Ns$ ellipsoid intersection tests, where $s$ is the number of sample points in the batch.
+Choosing the $i$th batch of sample points may require performing up to $\fedim |\pointbatch_i|$ ellipsoid intersection tests, where $|\pointbatch_i|$ is the number of sample points in the $i$th batch.
 Choosing all of the sample points therefore requires performing at most
 \begin{equation*}
-	Nm \qquad \text{ellipsoid intersection tests,}
-\end{equation*}
-where $m$ is the total number of sample points in all batches.
-The multiplicative dependence of $N$ with $m$ is undesirable since $m$ may be large, and reducing this cost is a target for future work.
+	\fedim \nsamplepts \qquad \text{ellipsoid intersection tests,}
+\end{equation*}
+where $\nsamplepts$ is the total number of sample points in all batches.
+The multiplicative dependence of $\fedim$ with $\nsamplepts$ is undesirable since $\nsamplepts$ may be large, and reducing this cost is a target for future work.
 Several improvements are possible.
 The cost of picking sample points could be reduced by ``thinning'' the set of candidate points that the sample points are picked from.
-That is, choosing sample points from a smaller subset of $N_\text{thin} << N$ points.
+That is, choosing sample points from a smaller subset of $\fedim_\text{thin} << \fedim$ points.
 The cost to perform the ellipsoid intersection tests could be reduced by first checking whether the ellipsoid bounding boxes intersect before performing the ellipsoid intersection test of Appendix \ref{sec:fast_ellipsoid_intersection_test}.
-The cost of picking one batch could be reduced from $N s$ to $O(N \log s)$ using more advanced computational geometry techniques, such as organizing the ellipsoids associated with points that have already been picked into a dynamic AABB tree, then using this AABB tree to accelerate the required ellipsoid intersection tests.
-This would reduce the cost of choosing all sample point batches to $O(N n_b \log s)$.
+The cost of picking the $i$th batch could be reduced to $O(\fedim \log |\pointbatch_i|)$ using more advanced computational geometry techniques, such as organizing the ellipsoids associated with points that have already been picked into a dynamic AABB tree, then using this AABB tree to accelerate the required ellipsoid intersection tests.
 However, from a practical perspective, the cost for picking sample points is small compared to other parts of the approximation algorithm, so we do not pursue these improvements in this paper.
 
-\paragraph{2: Building the H-matrix} The classical H-matrix construction technique, which is described in Appendix \ref{app:h_matrix}, requires evaluating $O(\hrank^2 N \log N)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank (maximum rank among the blocks of the H-matrix).
-To evaluate one matrix entry, first one must find the $k_n$ nearest sample points to a given point, where $k_n$ is the number of impulses used in the RBF interpolation.
-This is done using a precomputed KD tree of sample points, and requires $O(k_n \log m)$ work.
-Next, one must find the mesh cells that the points $\{z_i\}_{i=1}^{k+n}$ reside.
-This is done using our mesh AABB tree, and requires $O(k_n \log N)$ work.
-Next, we must evaluate finite element basis functions on each of those cells, which requires $O(k_n)$ work.
-Finally, we must solve a $k_n \times k_n$ linear system to perform the radial basis function interpolation, which requires $O(k_n^3)$ work.
-Overall building the H-matrix therefore requires
-\begin{equation*}
-	O\left(\left(\hrank^2 N \log N\right)\left(k_n \log N + k_n^3\right)\right) \qquad \text{work}.
-\end{equation*}
-This follows from the above discussion, and the fact that $N \ge m$.
-
-\paragraph{(3): H-matrix operations} H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion require
-\begin{equation*}
-	O(\hrank^2 N \log(N)^a) \qquad \text{work},
+\paragraph{2: Building the H-matrix} The classical H-matrix construction technique, which is described in Appendix \ref{app:h_matrix}, requires evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank (maximum rank among the blocks of the H-matrix).
+To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulses used in the RBF interpolation.
+This is done using a precomputed KD tree of sample points, and requires $O(\numnbr \log \nsamplepts)$ work.
+Second, one must find the mesh cells that the points $\{z_i\}_{i=1}^{\numnbr}$ reside.
+This is done using our mesh AABB tree, and requires $O(\numnbr \log \fedim)$ work.
+Third, we must evaluate finite element basis functions on each of those cells, which requires $O(\numnbr)$ work.
+Finally, we must solve a $\numnbr \times \numnbr$ linear system to perform the radial basis function interpolation, which requires $O(\numnbr^3)$ work.
+Building the H-matrix therefore requires
+\begin{equation*}
+	O\left(\left(\hrank^2 \fedim \log \fedim\right)\left(\numnbr \log \fedim + \numnbr^3\right)\right) \qquad \text{work}.
+\end{equation*}
+This follows from the above discussion, and the fact that $\fedim \ge \nsamplepts$.
+
+\paragraph{(3): H-matrix operations} It is well-known that H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion require
+\begin{equation*}
+	O(\hrank^2 \fedim \log(\fedim)^a) \qquad \text{work},
 \end{equation*}
@@
 The precise cost depends on the type of H-matrix used and the operation being performed; for more details see \cite{GrasedyckHackbusch03}.
-The rational method for modifying $\mathbf{A}_H$ to be positive semi-definite, described in Appendix \ref{sec:make_spd}, is performed using two matrix-matrix additions, $l+1$ matrix-matrix multiplications, and one matrix inversion, where $2^l$ is the order of the rational function used.
-
-%$k_\text{hmatrix}$, $k_\text{nbrs}$, $k_\text{batches}$
-%
-%\begin{tabular}{ll}
-%	$V$ & $1$ action of $\mathcal{A}^T$ \\
-%	$\mu$ & $d$ actions of $\mathcal{A}^T$ \\
-%	$\Sigma$ & $d(d+1)/2$ actions of $\mathcal{A}^T$ \\
-%	$\{\eta_b\}_{b=1}^\text{num\_batches}$ & $\text{num\_batches}$ actions of $\mathcal{A}$
-%\end{tabular}
-%
-%
-%Computing $V$, $\mu$, and $\Sigma$ require $1$, $d$, and $d(d+1)/2$ operator actions of $\Aop$, respectively.
-Computing the impulse response batches $\eta_b$ requires one operator action of $\Aop$ per batch.
-Picking the sample points
-%
-%In the non-symmetric case, evaluating $\widetilde{\Phi}(y,x)$ requires us to perform one $k$-nearest neighbor query of the KD-tree of sample points, $k$ point collision queries of the AABB-tree of mesh cells, and solve a $k \times k$ linear system.
-In the symmetric case, we must perform two $k$-nearest neighbor queries of the KD-tree of sample points, $2k$ point collision queries of the AABB-tree of mesh cells, and solve a $4k \times 2k$ linear least squares problem via QR factorization.
-Both of these cases require $O(k \log r + k \log N_\text{cell} + k^3)$ work, where $r$ is the number of sample points, and $N_\text{cell}$ is the number of cells in the mesh used to discretize the problem.
-Since $k$ is small, this is cheap.
-
-%\section{Content to add}
-%
-%\begin{itemize}
-%	\item computational complexity in terms of PDE solves:
-%	\item cost of global low rank approximation (data scalability)
-%	\item cost to build approximation
-%	\item cost without compression to do a solve
-%	\item cost with compression to do a solve
-%	\item computational complexity for H-matrix alone vs. prod-conv + h-matrix $O(C k^2 \log \fedim)$ PDE solves, C is big, k is big. Here $10$
-%	\item HODLR: C is smaller, but k is bigger, vs. H1 peeling process
-%	\item H-matrix operations cost for our method solve
-%	\item irregular grid
-%\end{itemize}
+The rational method for modifying $\mathbf{A}_H$ to be positive semi-definite, described in Appendix \ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the order of the rational function used.
 
@@
 
-The process of constructing an H-matrix representation of $\mathbf{\Phi}$ proceeds as follows.
+The process of constructing an H-matrix representation of $\mathbf{\Aker}$ proceeds as follows.
 First, we construct hierarchical partitionings of the degrees of freedom for the columns and rows of the matrix (cluster trees, Section \ref{sec:cluster_trees}).
@@
 The first two steps require geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix, but these steps do not depend on the particular values of matrix entries.
-The third step requires us to evaluate $O(\fedim \log \fedim)$ specially-chosen entries of $\mathbf{\Phi}$, and we evaluate these entries using \eqref{eq:Akerpcmat_entries}.
+The third step requires us to evaluate $O(\fedim \log \fedim)$ specially-chosen entries of $\mathbf{\Aker}$, and we evaluate these entries using \eqref{eq:Akerpcmat_entries}.
 
@@
 
-Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\mathbf{\Phi}$ corresponds to the Lagrange node in $\mathbb{R}^\gdim$ associated with the $i^\text{th}$ finite element basis function.
+Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\mathbf{\Aker}$ corresponds to the Lagrange node in $\mathbb{R}^\gdim$ associated with the $i\text{th}$ finite element basis function.
 The columns of the matrix thus correspond to a point cloud in $\mathbb{R}^\gdim$.
@@
 
-After reordering the rows and columns of $\mathbf{\Phi}$ via the depth-first ordering described above, we partition the reordered version of $\mathbf{\Phi}$ into a tree of $2 \times 2$ block matrices recursively.
+After reordering the rows and columns of $\mathbf{\Aker}$ via the depth-first ordering described above, we partition the reordered version of $\mathbf{\Aker}$ into a tree of $2 \times 2$ block matrices recursively.
 We use a geometric admissibility condition (discussed below) to decide which blocks to subdivide, and use the cluster trees for the rows and columns to determine how to subdivide those blocks.
 For the first stage of subdivision, let $r_1$ and $r_2$ be the children row clusters for the root of the row cluster tree, let $\colcluster_1$ and $\colcluster_2$ be the children column clusters for the root of the column cluster tree.
-The matrix $\mathbf{\Phi}$ is partitioned into blocks as follows:
+The matrix $\mathbf{\Aker}$ is partitioned into blocks as follows:
 \begin{equation*}
 	\begin{bmatrix}
-		\mathbf{\Phi}_{11} & \mathbf{\Phi}_{12} \\
-		\mathbf{\Phi}_{21} & \mathbf{\Phi}_{22}
+		\mathbf{\Aker}_{11} & \mathbf{\Aker}_{12} \\
+		\mathbf{\Aker}_{21} & \mathbf{\Aker}_{22}
 	\end{bmatrix},
 \end{equation*}
-where $\mathbf{\Phi}_{11}$ denotes the block of $\mathbf{\Phi}$ with rows $r_1$ and columns $\colcluster_1$, $\mathbf{\Phi}_{12}$ denotes the block of $\mathbf{\Phi}$ with rows $r_1$ and columns $\colcluster_2$, and so on for $\mathbf{\Phi}_{21}$ and $\mathbf{\Phi}_{22}$.
-
-We now loop through the four blocks, $\mathbf{\Phi}_{11}$, $\mathbf{\Phi}_{12}$, $\mathbf{\Phi}_{21}$, and $\mathbf{\Phi}_{22}$, and decide which blocks should be subdivided further.
-For the purpose of explanation, consider $\mathbf{\Phi}_{12}$.
+where $\mathbf{\Aker}_{11}$ denotes the block of $\mathbf{\Aker}$ with rows $r_1$ and columns $\colcluster_1$, $\mathbf{\Aker}_{12}$ denotes the block of $\mathbf{\Aker}$ with rows $r_1$ and columns $\colcluster_2$, and so on for $\mathbf{\Aker}_{21}$ and $\mathbf{\Aker}_{22}$.
+
+We now loop through the four blocks, $\mathbf{\Aker}_{11}$, $\mathbf{\Aker}_{12}$, $\mathbf{\Aker}_{21}$, and $\mathbf{\Aker}_{22}$, and decide which blocks should be subdivided further.
+For the purpose of explanation, consider $\mathbf{\Aker}_{12}$.
 If
@@
 \end{equation}
-then we mark $\mathbf{\Phi}_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone.
+then we mark $\mathbf{\Aker}_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone.
 Here $\dist\left(r_1, \colcluster_2\right)$ is the Euclidean distance between the axis-aligned bounding box for the point cloud associated with the row cluster $r_1$, and the axis aligned bounding box for the point cloud associated with the column cluster $\colcluster_2$.
@@
 This process is repeated for the other blocks to determine which blocks are admissible and which are not.
-For us, the diagonal blocks $\mathbf{\Phi}_{11}$ and $\mathbf{\Phi}_{22}$ are not admissible because the distance between a point cloud and itself is zero.
-
-Next, we subdivide all blocks that are not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\mathbf{\Phi}$.
+For us, the diagonal blocks $\mathbf{\Aker}_{11}$ and $\mathbf{\Aker}_{22}$ are not admissible because the distance between a point cloud and itself is zero.
+
+Next, we subdivide all blocks that are not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\mathbf{\Aker}$.
 But now we subdivide a block based on the two child row clusters and two child column clusters for the rows and columns of that block.
@@
 Once the block cluster tree has been constructed, low rank approximations of the admissible (low rank) blocks are formed using the adaptive cross method \cite{ACA}.
-Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\mathbf{\Phi}$.
+Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\mathbf{\Aker}$.
 The idea of the adaptive cross method is to form a low rank approximation of $X$ by sampling a small number of rows and columns of $X$.
@@
 There is also a small linear algebra overhead cost for the determinant maximization process that is performed at each step.
-%For more details on adaptive cross low rank approximation, see \cite{ACA}.
 The key point is that the adaptive cross method allows us to form a rank-$\hrank$ approximation of an $m \times m$ block of the matrix via a process that only requires accessing $O(m\hrank)$ entries of that block.
 
-We use the adaptive cross method to form low rank approximations for each admissible block of $\mathbf{\Phi}$.
-We directly compute all entries of the small dense blocks of $\mathbf{\Phi}$ that are not admissible.
-This process requires us to compute $O(\hrank^2 \fedim \log \fedim)$ entries of $\mathbf{\Phi}$, which is relatively cheap compared to the operator actions of $\Aop$ that are required to form the product-convolution approximation.
+We use the adaptive cross method to form low rank approximations for each admissible block of $\mathbf{\Aker}$.
+We directly compute all entries of the small dense blocks of $\mathbf{\Aker}$ that are not admissible.
+This process requires us to compute $O(\hrank^2 \fedim \log \fedim)$ entries of $\mathbf{\Aker}$, which is relatively cheap compared to the operator actions of $\Aop$ that are required to form the product-convolution approximation.
 
@@
 \end{equation}
-where $v := \Phi^T\left(\spatialmean_1 - \spatialmean_2\right)$.
+where $v := P^T\left(\spatialmean_1 - \spatialmean_2\right)$.
 We compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's algorithm \cite{Brent71} (any fast 1 dimensional convex optimization routine may be used).
@@
 
-%WOODBURY: \cite{Hager89}.
-
 In many problems of practical interest (e.g., Hessian approximation) $\Aop$ is symmetric positive semi-definite.
@@
 	\includegraphics[scale=0.5]{spd_rational_function.pdf}
-	\caption{Rational function $\Pi_k(\lambda)$ for $a=-1$, $b=0$, and $k \in \{1,2,3\}$. As $k$ increases, $\Pi_k$ approaches the indicator function for the set $(a,b)$. }
+	\caption{Rational function $\Pi_\ratord(\lambda)$ for $a=-1$, $b=0$, and $\ratord \in \{1,2,3\}$. As $\ratord$ increases, $\Pi_\ratord$ approaches the indicator function for the set $(a,b)$. }
 	\label{fig:spd_rat}
@@
 
-Computing $\mathbf{\Pi}=\Pi(\mathbf{A}_H^{\text{sym}})$ directly is difficult, but we may approximate $\Pi$ by a high order rational function $\Pi_k \approx \Pi$ which is defined as follows:
-\begin{equation*}
-	 \Pi_k(\lambda) := \frac{1}{1 + T(\lambda)^{2^k} }, \qquad \text{where} \qquad T(\lambda) := \frac{2 \lambda - (b+a)}{b-a}.
-\end{equation*}
-Here $k$ is a positive integer; the larger $k$ is, the better the approximation $\Pi \approx \Pi_k$.
+Computing $\mathbf{\Pi}=\Pi(\mathbf{A}_H^{\text{sym}})$ directly is difficult, but we may approximate $\Pi$ by a high order rational function $\Pi_\ratord \approx \Pi$ which is defined as follows:
+\begin{equation*}
+	 \Pi_\ratord(\lambda) := \frac{1}{1 + T(\lambda)^{2^\ratord} }, \qquad \text{where} \qquad T(\lambda) := \frac{2 \lambda - (b+a)}{b-a}.
+\end{equation*}
+Here $\ratord$ is a positive integer; the larger $\ratord$ is, the better the approximation $\Pi \approx \Pi_\ratord$.
 This is illustrated in Figure \ref{fig:spd_rat}.
-Replacing scalar operations with matrix operations, we have an approximation $\mathbf{\Pi}_k \approx \mathbf{\Pi}$ given by
+Replacing scalar operations with matrix operations, we have an approximation $\mathbf{\Pi}_\ratord \approx \mathbf{\Pi}$ given by
 \begin{equation}
 \label{eq:rational343}
-\mathbf{\Pi}_k := \Pi_k\left(\mathbf{A}_H^{\text{sym}}\right) = \left(\mathbf{I} + \mathbf{T}^{2^k}\right)^{-1},
+\mathbf{\Pi}_\ratord := \Pi_\ratord\left(\mathbf{A}_H^{\text{sym}}\right) = \left(\mathbf{I} + \mathbf{T}^{2^\ratord}\right)^{-1},
 \end{equation}
@@
 \end{equation*}
-Using $\mathbf{\Pi}_k$ instead of $\mathbf{\Pi}$ in \eqref{eq:I_Pi_abs} allows us to efficiently approximate $|\mathbf{A}_H^{\text{sym}}|$ using H-matrix methods. 
+Using $\mathbf{\Pi}_\ratord$ instead of $\mathbf{\Pi}$ in \eqref{eq:I_Pi_abs} allows us to efficiently approximate $|\mathbf{A}_H^{\text{sym}}|$ using H-matrix methods. 
 This approximation procedure is as follows.
 First, we compute $\lambda_\text{min}$ using the Lanczos method, and set $a=\gamma \lambda_\text{min}$ and $b=0$, where $\gamma\ge 1$ is a parameter that must be chosen.
-%Because of error in the rational function approximation, if $\gamma=1$, the negative eigenvalues of largest magnitude will be approximately zeroed out instead of flipped. 
 The method is relatively insensitive to $\gamma$ and numerically we find that the method works well for any $\gamma \in [1,2]$. We use $\gamma=1.5$ in our numerical results. 
 Next, we compute $\mathbf{T}$ using fast H-matrix scalar multiplication and addition.
-We form the $\mathbf{T}^{2^k}$ via repeated squaring, $\mathbf{T}^2 = \mathbf{T} \mathbf{T}$, $\mathbf{T}^4 = \mathbf{T}^2 \mathbf{T}^2$, $\mathbf{T}^8 = \mathbf{T}^4 \mathbf{T}^4$ and so on, where each of these matrix squarings is performed using H-matrix multiplication, which is fast.
-We recommend using $k=1$ or $2$ (we use $k=1$ in our numerical results).
-Larger $k$ is unnecessary and may make the method unstable.
-The approximate spectral projector $\mathbf{\Pi}_k$ is formed from $\mathbf{T}^{2^k}$ per \eqref{eq:rational343}, using fast H-matrix operations to perform the required matrix-matrix addition and matrix inversion.
+We form the $\mathbf{T}^{2^\ratord}$ via repeated squaring, $\mathbf{T}^2 = \mathbf{T} \mathbf{T}$, $\mathbf{T}^4 = \mathbf{T}^2 \mathbf{T}^2$, $\mathbf{T}^8 = \mathbf{T}^4 \mathbf{T}^4$ and so on, where each of these matrix squarings is performed using H-matrix multiplication, which is fast.
+We recommend using $\ratord=1$ or $2$ (we use $\ratord=1$ in our numerical results).
+Larger $\ratord$ is unnecessary and may make the method unstable.
+The approximate spectral projector $\mathbf{\Pi}_\ratord$ is formed from $\mathbf{T}^{2^\ratord}$ per \eqref{eq:rational343}, using fast H-matrix operations to perform the required matrix-matrix addition and matrix inversion.
 Finally, a positive semi-definite approximation $\mathbf{A}_H^{\text{sym}+}$ of $\mathbf{A}_H^{\text{sym}}$ is formed as
 \begin{equation*}
-\mathbf{A}_H^{\text{sym}+} := \left(\mathbf{I} - 2\mathbf{\Pi}_k\right)\mathbf{A}_H^{\text{sym}},
+\mathbf{A}_H^{\text{sym}+} := \left(\mathbf{I} - 2\mathbf{\Pi}_\ratord\right)\mathbf{A}_H^{\text{sym}},
 \end{equation*}
 where fast H-matrix operations are used to perform the necessary matrix-matrix subtraction and multiplication.
-Note that for any $k \ge 1$, $\Pi_k(\lambda)> 1/2$ for $\lambda \in (a,b)$, so $\mathbf{A}_H^{\text{sym}+}$ is positive semi definite\footnote{Some indefiniteness may be introduced by truncation error in the H-matrix operations, but the negative eigenvalues caused by this truncation error are typically extremely small.}.
+Note that for any $\ratord \ge 1$, $\Pi_\ratord(\lambda)> 1/2$ for $\lambda \in (a,b)$, so $\mathbf{A}_H^{\text{sym}+}$ is positive semi definite\footnote{Some indefiniteness may be introduced by truncation error in the H-matrix operations, but the negative eigenvalues caused by this truncation error are typically extremely small.}.
 At the end, $\mathbf{A}_H^{\text{sym}+}$ is symmetrized once again, to correct for any small non-symmetries that are introduced by truncation error in the H-matrix operations.
@@
 
-% the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ with zero (or approximate zero) while keeping the positive eigenvalues unchanged (approximately unchanged). Ideally
-%
-%
-%Then we compute the negative eigenvalues, of $\mathbf{A}_H$ that are less than (i.e., more negative than) a small magnitude negative cutoff value $c$, and also compute the corresponding eigenvectors.
-That is, the eigenpairs $\{(\lambda_i,u_i)\}_{i=1}^{N_\text{bad}}$ with
-%\begin{equation*}
-%	\lambda_i < c < 0.
-%\end{equation*}
-%These outlying ``bad'' eigenpairs are computed using the Lanczos method \cite{Komzsik03,Lanczos50}.
-%We then flip the sign of these negative eigenvalues to be positive as follows:
-%\begin{equation*}
-%	\mathbf{A}_H^{\text{sym}+} = \mathbf{A}_H^{\text{sym}} + \sum_{i=1}^{N_\text{bad}} 2 |\lambda_i| u_i u_i^T.
-%\end{equation*}
-%The matrix $\mathbf{A}_H^{\text{sym}+}$ approximates the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$.
-A H-matrix representation of $\mathbf{A}_H^{\text{sym}+}$ is formed using H-matrix addition, and used in place of $\mathbf{A}_H$ in further linear algebra operations.
-%
-%The cutoff $c$ is required when $\Aop$ has a large number of zero eigenvalues, or a sequence of eigenvalues that cluster at zero, a situation that occurs commonly in practice.
-We seek to correct negative eigenvalues caused by errors in our approximation procedure.
-But even without our approximation, rounding and discretization errors may cause large numbers of zero or near-zero eigenvalues to become slightly negative.
-Numerically, we cannot hope to correct these slightly negative eigenvalues because there are so many of them, and because the spectral gap between the smallest negative eigenvalue and the first non-negative eigenvalue is small.
-%
-%For Hessians in ill-posed inverse problems, the Hessian is the sum of the Hessian of a data misfit term, and the Hessian of a regularization term.
-The Hessian of the regularization term is known and easy to manipulate, and one seeks to approximate the Hessian of the data misfit term with our method.
-In this case, a reasonable choice for the cutoff parameter is $c = -|\lambda_R^\text{min}|/2$, where $\lambda_R^\text{min}$ is the smallest magnitude eigenvalue of the Hessian of the regularization term.
-We use this cutoff parameter in our numerical experiments.
-
 \section{Recycling Krylov information via low rank updates}
@@
 
-REFER TO BIROS PAPER
-
 In applications, often one wants to use the H-matrix approximation $\mathbf{A}_H$ to build preconditioners for solving a sequence of slowly varying linear systems, where the $k$th linear system has the form
@@
 Each of these linear systems is solved using an iterative method. By ``slowly varying,'' we mean that $\mathbf{B}^{(k+1)}$ does not differ much from $\mathbf{B}^{(k)}$. 
-%If $\mathbf{B}^{(k)}=\mathbf{A}$, then our H-matrix approximation $\mathbf{A}_H^{-1}$ may be used to precondition this linear system. 
-%In our target applications, $\mathbf{B}^{(k)}$ is built from a combination of $\mathbf{A}$ and other operators, then $\mathbf{A}_H$ may be combined with H-matrix approximations of the other operators to construct a preconditioner for $\mathbf{B}^{(k)}$.
 For example, in a Newton-Krylov method for solving an optimization problem, one uses a Krylov method such as conjugate gradient to solve a sequence of systems of the form \eqref{eq:Ax_equals_b}, where $\mathbf{B}^{(k)}$ and $\mathbf{f}^{(k)}$ are the Hessian and the negative gradient of the objective function in the optimization problem, respectively, evaluated at the $k$th Newton iterate.
-%The Hessian and the gradient are evaluated at the $k$th Newton iterate, hence both $\mathbf{B}^{(k)}$ and $\mathbf{f}^{(k)}$ change as the Newton iterations proceed. 
 As the optimization solver converges, $\mathbf{B}^{(k)}$ converges to the Hessian at the optimal point, and therefore changes little from one iteration to the next.
 Since $\mathbf{B}^{(k)}$ varies slowly, it is reasonable to assume that Krylov information about $\mathbf{B}^{(k)}$ can be leveraged to update the preconditioner from $\mathbf{B}^{(k)}$ in order to build a preconditioner for $\mathbf{B}^{(k+1)}$.
-%In the context of an inverse problem, the Hessian is typically the sum of two terms: the Hessian of the data misfit term, which is only available matrix-free, and the Hessian of a regularization term, which is available directly as a sparse matrix.
-The Hessian of the data misfit term is approximated with our method to form $\mathbf{A}_H$, the Hessian of the regularization term is approximated using standard H-matrix methods for sparse matrices, and then these two terms are added together and inverted using H-matrix methods to form a preconditioner for the overall Hessian.
 Concretely, let 
@@
 To do so, we build on the rank-1 update formulas discussed in \cite[Chapter 6]{NocedalWright99}, which we generalize to many vector ``rank-$m$'' updates.
-In what follows, let $\mathbf{X}$ and $\mathbf{Y}$ denote the $N \times m$ matrices with columns given by the vectors $\mathbf{x}_i$ and $\mathbf{y}_i$, respectively.
+In what follows, let $\mathbf{X}$ and $\mathbf{Y}$ denote the $\fedim \times m$ matrices with columns given by the vectors $\mathbf{x}_i$ and $\mathbf{y}_i$, respectively.
 The matrix-vector products with $\mathbf{B}^{(k)}$ performed while solving the $k$th linear system may be summarized by the equation:
@@
 
-%formula, which is named after it's inventor Davidson, and popularizers Fletcher and Powell \cite{DFP}.
-
-%For more details on these update formulas, we recommend .
-
-%In \cite{NocedalWright99}, one-vector versions of formulas \eqref{eq:srk_update} and \eqref{eq:dfp_update} are discussed thoroughly; the versions here are straightforward generalizations to many-vector updates.
-
-%to update $\mathbf{B}^{(k)} \approx \left(\mathbf{A}^{(k)}\right)^{-1}$ denote the H-matrix preconditioner used for the $k$th linear system, and set $\mathbf{B}^{(1)}:= \mathbf{A}_H^{-1}$ denote the H-matrix approximation used to precondition the first linear system.
-Rather than re-build the H-matrix preconditioner for each subsequent linear system, one may instead form a preconditioner, $\mathbf{B}^{(k+1)}$ update, by \emph{recycling} the information generated by the iterative method used to solve the current system.
-%
-%In practice it is often necessary to solve a sequence of linear systems of the form 
-%\begin{equation*}
-%	\mathbf{A} \mathbf{x} = \mathbf{b}
-%\end{equation*}
-%where either $\mathbf{A}$, or $\mathbf{b}$, or both, are slowly varying. 
-%
-%\begin{equation*}
-%	\mathbf{B}\mathbf{X} = \mathbf{Y}
-%\end{equation*}
-%
-%
-%
-%\begin{equation*}
-%\mathbf{Q} := \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T
-%\end{equation*}
-%
-%\begin{equation}
-%	\mathbf{B}_{k+1} = \left(\mathbf{I} - \mathbf{Y}\boldsymbol{\Theta}\mathbf{X}^T\right)\mathbf{B}_k\left(\mathbf{I} - \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T\right) + \mathbf{Y}\boldsymbol{\Theta}\mathbf{Y}^T
-%\end{equation}
-
 \section*{Acknowledgments}

### 2022-06-01 paper/localpsf.tex
edit computational cost section based on discussion with Noemi and Tucker

@@
 
-The computational cost of our method may be divided into the costs to perform the following tasks: 
-\begin{enumerate}
-	\item Computing the impulse response moments and batches.
-	\item Constructing the H-matrix $\mathbf{A}_H$.
-	\item Using $\mathbf{A}_H$ to perform further linear algebra operations. This may optionally include modifying $\mathbf{A}_H$ to make it symmetric positive semi-definite.
-\end{enumerate}
-In practice, 1 is the dominant cost, because 1 is the only task that requires applying $\Aop$ to vectors, and our method is targeted at applications in which applying $\Aop$ to a vector requires an expensive computational procedure such as solving a large-scale partial differential equation.
-All operations that do not require applications of $\Aop$ to vectors are nearly-linear, and therefore scalable, in $N$.
+The computational cost of our method may be divided into the costs to perform the following tasks: (1) Computing impulse response moments and batches; (2) Building the H-matrix; (3) Performing linear algebra operations with the H-matrix.
+This may optionally include modifying the H-matrix to make it symmetric positive semi-definite.
+In practice, (1) is the dominant cost, because (1) is the only task that requires applying $\Aop$ to vectors, and our method is targeted at applications in which applying $\Aop$ to a vector requires an expensive computational procedure such as solving a large-scale partial differential equation.
+All operations that do not require applications of $\Aop$ to vectors are nearly-linear, and therefore scalable, in the size of the problem, $N$.
 We now describe these costs in detail. For convenience, Table \ref{tab:vars} lists variable symbols and their approximate sizes.
@@
 	\begin{tabular}{lll}
-		Symbol & Approximate size & Variable name \\
+		Symbol & Typical size & Variable name \\
 		\hline
@@
 
-\paragraph{1: Impulse response moments and batches} Computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ requires applying $\Aop$ to $1$, $\gdim$, and $\gdim(\gdim+1)/2$ vectors, respectively.
-Computing each impulse response batch requires applying $\Aop$ to $1$ vector, so computing the impulse response batches $\{\eta_i\}_{i=1}^{\nbatch}$ requires $\nbatch$ operator applies.
+\paragraph{(1) Computing impulse response moments and batches} Computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ requires applying $\Aop$ to $1$, $\gdim$, and $\gdim(\gdim+1)/2$ vectors, respectively.
+Computing each impulse response batch requires applying $\Aop$ to one vector, so computing the impulse response batches $\{\eta_i\}_{i=1}^{\nbatch}$ requires $\nbatch$ operator applies.
 In total, computing the impulse response moments and batches therefore requires
@@
 \end{equation*}
-In a typical application one might have $\gdim=2$ and $\nbatch=5$, in which case $11$ operator applies are required.
+In a typical application one might have $\gdim=2$ and $\nbatch=5$, in which case a modest $11$ operator applies are required.
 
@@
 The multiplicative dependence of $\fedim$ with $\nsamplepts$ is undesirable since $\nsamplepts$ may be large, and reducing this cost is a target for future work.
-Several improvements are possible.
+%Several improvements are possible.
 The cost of picking sample points could be reduced by ``thinning'' the set of candidate points that the sample points are picked from.
@@
 
-\paragraph{2: Building the H-matrix} The classical H-matrix construction technique, which is described in Appendix \ref{app:h_matrix}, requires evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank (maximum rank among the blocks of the H-matrix).
+\paragraph{(2) Building the H-matrix} The classical H-matrix construction technique, which is described in Appendix \ref{app:h_matrix}, requires evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank (maximum rank among the blocks of the H-matrix).
 To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulses used in the RBF interpolation.
-This is done using a precomputed KD tree of sample points, and requires $O(\numnbr \log \nsamplepts)$ work.
+This is done using a precomputed KD tree of sample points, and requires $O(\numnbr \log \nsamplepts)$ floating point and logical elementary operations.
 Second, one must find the mesh cells that the points $\{z_i\}_{i=1}^{\numnbr}$ reside.
-This is done using our mesh AABB tree, and requires $O(\numnbr \log \fedim)$ work.
-Third, we must evaluate finite element basis functions on each of those cells, which requires $O(\numnbr)$ work.
-Finally, we must solve a $\numnbr \times \numnbr$ linear system to perform the radial basis function interpolation, which requires $O(\numnbr^3)$ work.
-Building the H-matrix therefore requires
-\begin{equation*}
-	O\left(\left(\hrank^2 \fedim \log \fedim\right)\left(\numnbr \log \fedim + \numnbr^3\right)\right) \qquad \text{work}.
+This is done using an AABB tree of mesh cells, and requires $O(\numnbr \log \fedim)$ elementary operations.
+Third, we must evaluate finite element basis functions on each of those cells, which requires $O(\numnbr)$ elementary operations.
+Finally, we must solve a $\numnbr \times \numnbr$ linear system to perform the radial basis function interpolation, which requires $O(\numnbr^3)$ elementary operations, which typically is negligible.
+Therefore, building the H-matrix requires
+\begin{equation*}
+	O\left(\left(\hrank^2 \fedim \log \fedim\right)\left(\numnbr \log \fedim + \numnbr^3\right)\right) \qquad \text{elementary operations}.
 \end{equation*}
@@
 
-\paragraph{(3): H-matrix operations} It is well-known that H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion require
-\begin{equation*}
-	O(\hrank^2 \fedim \log(\fedim)^a) \qquad \text{work},
+\paragraph{(3) Performing linear algebra operations with the H-matrix} It is well-known that H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion require
+\begin{equation*}
+	O(\hrank^2 \fedim \log(\fedim)^a) \qquad \text{elementary operations},
 \end{equation*}

### 2022-06-06 paper/localpsf.tex
edits to paper based on discussion with Noemi and Tucker

@@
 
-\paragraph{(2) Building the H-matrix} The classical H-matrix construction technique, which is described in Appendix \ref{app:h_matrix}, requires evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank (maximum rank among the blocks of the H-matrix).
+\paragraph{(2) Building the H-matrix} The classical H-matrix construction technique, which is described in Appendix \ref{app:h_matrix}, requires evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix.
 To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulses used in the RBF interpolation.
@@
 \end{equation*}
-This follows from the above discussion, and the fact that $\fedim \ge \nsamplepts$.
-
-\paragraph{(3) Performing linear algebra operations with the H-matrix} It is well-known that H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion require
-\begin{equation*}
+%This follows from the above discussion, and the fact that $\fedim \ge \nsamplepts$.
+
+\paragraph{(3) Performing linear algebra operations with the H-matrix} It is well-known that H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion, and low rank updates require
+\begin{equation}
+\label{eq:H_matrix_cost_operations}
 	O(\hrank^2 \fedim \log(\fedim)^a) \qquad \text{elementary operations},
-\end{equation*}
-where $a \in \{0,1,2,3\}$.
-The precise cost depends on the type of H-matrix used and the operation being performed; for more details see \cite{GrasedyckHackbusch03}.
+\end{equation}
+where $a \in \{0,1,2,3\}$ \cite{GrasedyckHackbusch03}.
+The precise cost depends on the type of H-matrix used and the operation being performed.
+In this paper, we use one matrix-matrix addition to add the H-matrix approximation of the data misfit term in the Hessian to the regularization term in the Hessian.
+Then, we use one matrix inversion to form the inverse of the Hessian, which is used as a preconditioner for the Newton linear system.
+We perform low-rank updates to the H-matrix as the Newton iterations proceed, as per the discussion in Appendix \ref{sec:recycle_krylov}.
 The rational method for modifying $\mathbf{A}_H$ to be positive semi-definite, described in Appendix \ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the order of the rational function used.

### 2022-07-13 paper/localpsf.tex
improved ncg info

@@
 
-%\textbf{Heat equation}
-%\begin{itemize}
-%	\item Hessian per-column error plots for 1, 5, and 15 batches
-%	\item $H-P$ error vs number of batches for interior and whole domain
-%	\item $P^{-1}H-I$: error vs rank for $P=R$ and $P=P_\text{pch}$ diffusion time, for 1, 5, and 15 batches
-%	\item Krylov iterations to tols 1e-1 and 1e-6 vs diffusion parameter
-%	\item GLR vs PCH vs diffusion time
-%	\item Krylov method convergence plot: $R$ vs $P_\text{pch}$ vs no preconditioner
-%	\item Krylov iterations to tols 1e-6 vs mesh size $h$, PCH vs R vs no preconditioner (hold)
-%	\item true parameter (H)
-%	\item deterministic reconstruction (H)
-%	\item noisy measurements (1\%, 5\%, 10\% noise). $u|_\text{top}$ (H)
-%	\item recovered state at top (H)
-%	\item mesh scalability of PCH (CC) (H)
-%	\item put plot data into data directory
-%	\item save function .pvd files in paraview  directory
-%\end{itemize}
-%
-%\begin{itemize}
-%	\item GLR vs PCH number of obs (Computational cost in PDE solves) (S)
-%	\item GLR vs PCH aspect ratio (CC) (S)
-%	\item turn on preconditioner after number of krylov iterations exceeds 15 in an iteration. Rebuild "as needed". (S)
-%\end{itemize}
-%
-%\begin{itemize}
-%	\item Stokes Krylov convergence (10 solves for nonlinear forward + 6 hessian matvecs for ellipsoid estimation + 10 matvecs for impulse responses + 30 matvecs for PCH6 CG solve + 100 matvecs for Reg CG solve)
-%	\item Stokes error vs num batches (small number of batches; 1, 3, 6, 9 batches?) (50 matvecs for randomized error estimation +)
-%	\item Velocity field
-%	\item Basal friction field
-%	\item Picture of impulse responses
-%	\item Mesh negative numbers picture
-%\end{itemize}
+\textbf{Heat equation}
+\begin{itemize}
+	\item Hessian per-column error plots for 1, 5, and 15 batches
+	\item $H-P$ error vs number of batches for interior and whole domain
+	\item $P^{-1}H-I$: error vs rank for $P=R$ and $P=P_\text{pch}$ diffusion time, for 1, 5, and 15 batches
+	\item Krylov iterations to tols 1e-1 and 1e-6 vs diffusion parameter
+	\item GLR vs PCH vs diffusion time
+	\item Krylov method convergence plot: $R$ vs $P_\text{pch}$ vs no preconditioner
+	\item Krylov iterations to tols 1e-6 vs mesh size $h$, PCH vs R vs no preconditioner (hold)
+	\item true parameter (H)
+	\item deterministic reconstruction (H)
+	\item noisy measurements (1\%, 5\%, 10\% noise). $u|_\text{top}$ (H)
+	\item recovered state at top (H)
+	\item mesh scalability of PCH (CC) (H)
+	\item put plot data into data directory
+	\item save function .pvd files in paraview  directory
+\end{itemize}
+
+\begin{itemize}
+	\item GLR vs PCH number of obs (Computational cost in PDE solves) (S)
+	\item GLR vs PCH aspect ratio (CC) (S)
+	\item turn on preconditioner after number of krylov iterations exceeds 15 in an iteration. Rebuild "as needed". (S)
+\end{itemize}
+
+\begin{itemize}
+	\item Stokes Krylov convergence (10 solves for nonlinear forward + 6 hessian matvecs for ellipsoid estimation + 10 matvecs for impulse responses + 30 matvecs for PCH6 CG solve + 100 matvecs for Reg CG solve)
+	\item Stokes error vs num batches (small number of batches; 1, 3, 6, 9 batches?) (50 matvecs for randomized error estimation +)
+	\item Velocity field
+	\item Basal friction field
+	\item Picture of impulse responses
+	\item Mesh negative numbers picture
+\end{itemize}
+
+\begin{itemize}
+	\item mtrue
+	\item utrue (no arrows)
+	\item utrue (yes arrows)
+	\item reconstruction iterations 1,3,4,8: PCH vs REG
+	\item urecovered (arrows, 3D)
+	\item morozov plot
+	\item Newton convergence (iter, numcg, numPDEsolves (num preconditioner build), gradnorm)
+	\item impulse response batches
+	\item 
+\end{itemize}
 

### 2022-07-18 paper/localpsf.tex
tightening up paper. commenting out stuff and reducing focus on H-matrices

--- 2022-06-06 paper/localpsf.tex
@@
 \begin{abstract}
-	We present a fast matrix-free method for computing hierarchical matrix (H-matrix) approximations of operators that have locally supported non-negative integral kernels.
+	We present a fast matrix-free method for approximating operators that have locally supported non-negative integral kernels.
 Such operators arise, for example, as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained inverse problems.
-H-matrix approximations are highly desirable because there are fast H-matrix methods to add, multiply, invert, and factorize H-matrices, compute matrix-vector products, solve linear systems, and perform other useful matrix computations.
-Classical H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
-Our method computes impulse responses of the operator at a collection of scattered points, then interpolates these known impulse responses to evaluate entries of the integral kernel at other points where the impulse response is not known, which allows us to construct the H-matrix.
+Our method computes impulse responses of the operator at a collection of scattered points, then interpolates these known impulse responses to evaluate entries of the integral kernel at other points where the impulse response is not known.
+%	Classical H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
 	Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources.
 By solving an ellipsoid packing problem, we choose as many points as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
-	We apply the method to approximate Hessians in large-scale PDE-constrained inverse problems with highly informative data.
-Numerical results demonstrate that our method substantially outperforms existing state-of-the-art Hessian approximation methods which are based on low-rank approximation.
-We are able to form high quality approximations of high rank operators using only a small number of operator applies.
+	The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
+Fast H-matrix methods are used to perform further matrix computations including matrix-matrix addition and multiplication, matrix inversion and factorization, and matrix-vector products in nearly-linear complexity.
+	We apply the method to approximate the Hessian in a large-scale PDE-constrained inverse problem with highly informative data.
+Numerical results demonstrate that our method substantially outperforms existing methods.
+We are able to approximate high rank operators using only a small number of operator applies.
 \end{abstract}
@@
 However, low rank approximation methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation \cite{Alger19,AlgerEtAl17}.
-High-rank matrix free approximation methods are needed.
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[scale=0.4]{bct1.eps}
-	\end{center}
-	\caption{Hierarchical matrix block structure for the Stokes Hessian. The red blocks (tiny) are dense. All other blocks are low rank.
-	}
-	\label{fig:hmatrix_intro}
-\end{figure}
-
-Our method forms a hierarchical matrix (H-matrix) \cite{Hackbusch99,} approximation of a discretized version of $\Aop$.
-H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high-rank (see Figure \ref{fig:hmatrix_intro} and Appendix \ref{app:h_matrix}).
-H-matrix methods permit us to perform matrix-vector products cheaply.
-Furthermore, one may use H-matrix methods to perform useful linear algebra operations involving the H-matrix that cannot easily be done using the original operator.
-These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion.
-These H-matrix methods are fast and scale well to large problems.
-Typically, the work and memory required to perform these operations for an $\fedim \times \fedim$ H-matrix scales as $O(\fedim \log(\fedim)^a)$, where $a \in \{0,1,2,3\}$.
-The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks; for more details see \cite{GrasedyckHackbusch03}.
-The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
+High-rank approximation methods are needed.
 
@@
 		\centering
-		\includegraphics[scale=0.33]{batch1.png}
+		\includegraphics[scale=0.33]{impulse_batch1.png}
 	\end{subfigure}
@@
 		\centering
-		\includegraphics[scale=0.33]{batch12.png}
+		\includegraphics[scale=0.33]{impulse_batch2.png}
 	\end{subfigure}
@@
 		\centering
-		\includegraphics[scale=0.33]{batch19.png}
+		\includegraphics[scale=0.33]{impulse_batch3.png}
 	\end{subfigure}
 	\caption{Batches of normalized impulse responses for the Stokes inverse problem Hessian. 
-Left: 2nd batch.
-Mid: 13th batch.
-Right: 19th batch.
-Brightness scale differs between the subfigures.
-Black stars indicate the sample point locations at which the point source is applied.
-Dashed gray ellipses indicate the estimated support ellipsoids.
-The large circle is the boundary of the domain.
+%	Brightness scale differs between the subfigures. 
+	Black stars indicate point source locations. Dashed gray ellipses indicate the estimated support ellipsoids. The large circle is the boundary of the domain. 
 	}
@@
 
-Classical H-matrix construction methods (described in Appendix \ref{sec:H_matrix_construction}) require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.
-Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}, which we do not use here.
-These alternative methods have, been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{AmbartsumyanEtAl20,HartlandEtAl22}. 
-While methods based on the peeling process are asymptotically scalable in theory, in practice the required number of matrix vector products is large.
-Matrix-free H-matrix construction methods are currently a topic of considerable research, so future algorithmic improvements may make matrix-free H-matrix methods based on the peeling process more efficient in the future.
-
-We construct the H-matrix using \emph{impulse response interpolation}.
+The central challenge we face is that existing scalable high-rank approximation methods typically require the ability to rapidly evaluate arbitrary entries of $\Aop$'s integral kernel, and we cannot do this because $\Aop$ is only accessible matrix-free.
+We address this challenge using \emph{impulse response interpolation}.
 The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(v) = \left(\rho^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
-We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with collections of points scattered throughout the domain (see Figure \ref{fig:batches_intro}).
-Then we interpolate shifted versions of these impulse responses to approximate arbitrary entries of the operator's integral kernel.
-This allows fast evaluation of approximate matrix entries, which are needed by the classical H-matrix construction methods.
-
-Picking the batches of points $x_i$ requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
+We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with batches of points scattered throughout the domain (see Figure \ref{fig:batches_intro}).
+Then we interpolate translated versions of these impulse responses to approximate arbitrary entries of the operator's integral kernel.
+%This allows fast evaluation of approximate matrix entries, which are needed by high-rank approximation methods, in particular, classical H-matrix construction methods.
+
+Picking the batches of points requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
 The idea of estimating the supports of the functions $\phi_x$ a-priori was inspired by techniques from resolution analysis in seismic imaging, in which the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion  \cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
@@
 We avoid this issue by temporarily excluding impulse responses from the approximation procedure whenever an attempt is made to evaluate the impulse response at a location where it is undefined.
+
+%\begin{figure}
+%	\begin{center}
+%		\includegraphics[scale=0.4]{bct1.eps}
+%	\end{center}
+%	\caption{Hierarchical matrix block structure for the Stokes Hessian. The red blocks (tiny) are dense. All other blocks are low rank.
+%	}
+%	\label{fig:hmatrix_intro}
+%\end{figure}
+
+The ability to rapidly evaluate approximations of arbitrary entries of $\Aop$'s integral kernel allows us to form a hierarchical matrix (H-matrix) \cite{Hackbusch99,} approximation of a discretized version of $\Aop$.
+H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high-rank.
+%(see Figure \ref{fig:hmatrix_intro} and Appendix \ref{app:h_matrix}).
+H-matrix methods permit us to perform matrix-vector products cheaply.
+Furthermore, one may use H-matrix methods to perform useful linear algebra operations involving the H-matrix that cannot easily be done using the original operator.
+These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion.
+%These H-matrix methods are fast and scale well to large problems. 
+The work and memory required to perform these operations for an $\fedim \times \fedim$ H-matrix is scales as $O(\fedim \log(\fedim)^a)$, where $a \in \{0,1,2,3\}$.
+That is, the cost is \emph{nearly-linear} in $N$.
+The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks; for more details see \cite{GrasedyckHackbusch03}.
+The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
+
+%Classical H-matrix construction methods 
+%(described in Appendix \ref{sec:H_matrix_construction}) 
+%require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.
+Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}, which we do not use here.
+These alternative methods have been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{AmbartsumyanEtAl20,HartlandEtAl22}. 
+Methods based on the peeling process are asymptotically scalable (typically the number of matrix vector products required scales as $O(\log N)$), but in practice the required number of matrix vector products is large.
+Improving the peeling process to reduce its cost is currently an area of active research \cite{PEELINGRESEARCH}.
+%This is an area of active research, so more efficient and practical matrix-free H-matrix construction methods based on the peeling process may be discovered in the future.
+%Matrix-free H-matrix construction methods are currently a topic of considerable research, so more efficient matrix-free H-matrix methods based on the peeling process may be discovered in the future.
 
@@
 Here $f/g$ denotes pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denotes pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
-The resulting method for computing discretized versions of $\spatialvol$, $\spatialmean$, and $\spatialcov$ is shown in Algorithm \ref{alg:varhpi_mean_cov}.
-
-\begin{algorithm2e}
-	\SetAlgoNoLine
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Output}
-	{	
-		\Input{Operator $\bm{A}^T$}
-		\Output{$\bm{\spatialvol}\in \mathbb{R}^\fedim$, $\bm{\spatialmean}\in \mathbb{R}^{\gdim \times \fedim}$, and $\bm{\spatialcov} \in \mathbb{R}^{\gdim \times \gdim \times \fedim}$}
-
-		Form vector $\bm{C} \in \mathbb{R}^\fedim_{\massmatrix}$ by either projecting or interpolating constant function $C(x)=1$ onto $V_h$
-
-		$\bm{\spatialvol} \gets \massmatrix^{-1}\bm{A}^T \bm{C}$
-
-		\For{$i=1,2,\dots,\gdim$}{
-			Form vector $\bm{L}^i \in \mathbb{R}^\fedim_{\massmatrix}$ by either projecting or interpolating linear function $L^i(x) = x^i$ onto $V_h$
-
-			$\bm{\spatialmean}^i \gets \left(\massmatrix^{-1}\bm{A}^T \bm{L}^i\right) / \bm{\spatialvol}$
-		}
-		\For{$i=1,2,\dots,\gdim$}{
-			\For{$j=1,\dots,i$}{
-				Form quadratic function $\bm{Q}^{ij} = x^i x^j$
-
-				$\bm{\spatialcov}^{ij} \gets \left(\massmatrix^{-1}\bm{A}^T \bm{Q}^{ij}\right) / \bm{\spatialvol} - \bm{\spatialmean}^i\cdot \bm{\spatialmean}^j$
-
-				$\bm{\spatialcov}^{ji} \gets \bm{\spatialcov}^{ij}$
-
-			}
-		}
-
-	}
-	\caption{Compute discretized scaling factor $\mathbf{\spatialvol}$, mean $\boldsymbol{\spatialmean}$, and covariance $\boldsymbol{\spatialcov}$}
-	\label{alg:varhpi_mean_cov}
-\end{algorithm2e}
+%The resulting method for computing discretized versions of $\spatialvol$, $\spatialmean$, and $\spatialcov$ is shown in Algorithm \ref{alg:varhpi_mean_cov}.
+
+%\begin{algorithm2e}
+%	\SetAlgoNoLine
+%	\SetKwInOut{Input}{Input}
+%	\SetKwInOut{Output}{Output}
+%	{	
+%		\Input{Operator $\bm{A}^T$}
+%		\Output{$\bm{\spatialvol}\in \mathbb{R}^\fedim$, $\bm{\spatialmean}\in \mathbb{R}^{\gdim \times \fedim}$, and $\bm{\spatialcov} \in \mathbb{R}^{\gdim \times \gdim \times \fedim}$}
+%		
+%		Form vector $\bm{C} \in \mathbb{R}^\fedim_{\massmatrix}$ by either projecting or interpolating constant function $C(x)=1$ onto $V_h$
+%		
+%		$\bm{\spatialvol} \gets \massmatrix^{-1}\bm{A}^T \bm{C}$
+%		
+%		\For{$i=1,2,\dots,\gdim$}{
+%			Form vector $\bm{L}^i \in \mathbb{R}^\fedim_{\massmatrix}$ by either projecting or interpolating linear function $L^i(x) = x^i$ onto $V_h$
+%			
+%			$\bm{\spatialmean}^i \gets \left(\massmatrix^{-1}\bm{A}^T \bm{L}^i\right) / \bm{\spatialvol}$
+%		}
+%		\For{$i=1,2,\dots,\gdim$}{
+%			\For{$j=1,\dots,i$}{
+%				Form quadratic function $\bm{Q}^{ij} = x^i x^j$
+%				
+%				$\bm{\spatialcov}^{ij} \gets \left(\massmatrix^{-1}\bm{A}^T \bm{Q}^{ij}\right) / \bm{\spatialvol} - \bm{\spatialmean}^i\cdot \bm{\spatialmean}^j$
+%				
+%				$\bm{\spatialcov}^{ji} \gets \bm{\spatialcov}^{ij}$
+%				
+%			}
+%		}
+%		
+%	}
+%	\caption{Compute discretized scaling factor $\mathbf{\spatialvol}$, mean $\boldsymbol{\spatialmean}$, and covariance $\boldsymbol{\spatialcov}$}
+%	\label{alg:varhpi_mean_cov}
+%\end{algorithm2e}
 
@@
 When $\Aop$ is symmetric positive semi-definite, $\mathbf{A}_H$ may be non-symmetric and indefinite due to errors in the approximation.
-In this case, one may (optionally) modify the H-matrix representation of $\mathbf{A}_H$ to make it symmetric and nearly positive semi-definite.
+In this case, one may (optionally) modify the H-matrix representation of $\mathbf{A}_H$ to make it symmetric positive semi-definite using the rational method described in Appendix \ref{sec:make_spd}.
+
+%.
 The matrix is symmetrized by computing $\left(\mathbf{A}_H+\mathbf{A}_H^T\right)/2$ using H-matrix addition.
-Negative eigenvalues that are large or medium in magnitude are computed, then flipped in sign to become positive.
-Small negative eigenvalues may remain because they are difficult to compute numerically.
-This is described in Appendix \ref{sec:make_spd}.
+Negative eigenvalues modified to be positive using a rational method that is described in Appendix \ref{sec:make_spd}.
 
@@
 
-		Compute $\bm{\spatialvol}$, $\bm{\spatialmean}$, and $\bm{\spatialcov}$ using Algorithm \ref{alg:varhpi_mean_cov}.
+		Compute $V, \mu$, and $\Sigma$ (Section \ref{sec:intromoments})
+
+%		Compute $\bm{\spatialvol}$, $\bm{\spatialmean}$, and $\bm{\spatialcov}$ using Algorithm \ref{alg:varhpi_mean_cov}.
 
@@
 \end{equation*}
-Once $\mathbf{A}_H$ is formed, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods \cite{BORMHACKBUSCHBOOK}.
+%Once $\mathbf{A}_H$ is formed, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods \cite{BORMHACKBUSCHBOOK}.
 We use H1 matrices in our numerical results, but any of the other H-matrix formats (such as H2, HODLR, HSS, HBS, and others \cite{HMATRIX}) could be used instead.
@@
 We form $\mathbf{\Aker}_H$ using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{Kriemann08}.
-Although $\mathbf{\Aker}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Aker}_H$ only requires evaluation of $O(\hrank^2 \fedim \log \fedim)$ entries $\mathbf{\Aker}_{ij} = \widetilde{\Aker}(p_i, p_j)$, and these entries are computed via the procedure described in Section \ref{sec:approximate_kernel_entries}.
+Although $\mathbf{\Aker}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Aker}_H$ only requires evaluation of $O(\hrank^2 \fedim \log \fedim)$ entries $\mathbf{\Aker}_{ij} = \widetilde{\Aker}(p_i, p_j)$, and these entries are computed via the radial basis function interpolation method described in Section \ref{sec:approximate_kernel_entries}.
 Here $\hrank$ is the rank of the highest-rank block in the H-matrix.
 A dense representation of $\mathbf{\Aker}$ is never formed. 
-We describe this H-matrix construction process in detail in Appendix \ref{app:h_matrix}.
+%We describe this H-matrix construction process in detail in Appendix \ref{app:h_matrix}. 
 We form $\massmatrix_H$ using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same recursive block partitioning structure as was used for $\mathbf{\Aker}$.
@@
 
-\paragraph{(2) Building the H-matrix} The classical H-matrix construction technique, which is described in Appendix \ref{app:h_matrix}, requires evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix.
+\paragraph{(2) Building the H-matrix} The classical H-matrix construction technique
+%, which is described in Appendix \ref{app:h_matrix}, 
+requires evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix.
 To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulses used in the RBF interpolation.
@@
 Then, we use one matrix inversion to form the inverse of the Hessian, which is used as a preconditioner for the Newton linear system.
-We perform low-rank updates to the H-matrix as the Newton iterations proceed, as per the discussion in Appendix \ref{sec:recycle_krylov}.
+%We perform low-rank updates to the H-matrix as the Newton iterations proceed, as per the discussion in Appendix \ref{sec:recycle_krylov}.
 The rational method for modifying $\mathbf{A}_H$ to be positive semi-definite, described in Appendix \ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the order of the rational function used.
@@
 %\end{itemize}
+%
+%\begin{itemize}
+%	\item mtrue
+%	\item utrue (no arrows)
+%	\item utrue (yes arrows)
+%	\item reconstruction iterations 1,3,4,8: PCH vs REG
+%	\item urecovered (arrows, 3D)
+%	\item morozov plot
+%	\item Newton convergence (iter, numcg, numPDEsolves (num preconditioner build), gradnorm)
+%	\item impulse response batches
+%	\item 
+%\end{itemize}
 
@@
 
-\section{Hierarchical matrix details}
-\label{app:h_matrix}
-
-Often, large dense matrices of practical interest may be permuted, then partitioned into blocks recursively, in such a way that many off-diagonal blocks of the matrix are numerically low rank, even if the matrix is high rank.
+%\section{Hierarchical matrix details}
+%\label{app:h_matrix}
+%
+%Often, large dense matrices of practical interest may be permuted, then partitioned into blocks recursively, in such a way that many off-diagonal blocks of the matrix are numerically low rank, even if the matrix is high rank.
 Such matrices are known as hierarchical matrices (H-matrices).
@@
 For this section, when we say H-matrix, we are referring to H1 matrices.
-
-While a dense $\fedim \times \fedim$ matrix traditionally requires $O(\fedim^2)$ memory to store, H-matrices may be stored using $O(\fedim \log \fedim)$ memory, by storing only the low rank factors for the low rank blocks.
+%
+%While a dense $\fedim \times \fedim$ matrix traditionally requires $O(\fedim^2)$ memory to store, H-matrices may be stored using $O(\fedim \log \fedim)$ memory, by storing only the low rank factors for the low rank blocks.
 Recursive algorithms can take advantage of the H-matrix low rank block structure to perform matrix arithmetic fast.
@@
 For more details on H-matrices, we recommend \cite{HMATRIXGOOD}.
-
-\subsection{H-matrix construction}
-\label{sec:H_matrix_construction}
-
-The process of constructing an H-matrix representation of $\mathbf{\Aker}$ proceeds as follows.
+%
+%
+%\subsection{H-matrix construction}
+%\label{sec:H_matrix_construction}
+%
+%The process of constructing an H-matrix representation of $\mathbf{\Aker}$ proceeds as follows.
 First, we construct hierarchical partitionings of the degrees of freedom for the columns and rows of the matrix (cluster trees, Section \ref{sec:cluster_trees}).
@@
 The third step requires us to evaluate $O(\fedim \log \fedim)$ specially-chosen entries of $\mathbf{\Aker}$, and we evaluate these entries using \eqref{eq:Akerpcmat_entries}.
-
-\subsubsection{Cluster trees}
-\label{sec:cluster_trees}
-
-We use recursive hyperplane splitting to hierarchically cluster the degrees of freedom associated with the columns and rows of the matrix into a \emph{column cluster tree} and a \emph{row cluster tree}, respectively.
+%
+%
+%\subsubsection{Cluster trees}
+%\label{sec:cluster_trees}
+%
+%We use recursive hyperplane splitting to hierarchically cluster the degrees of freedom associated with the columns and rows of the matrix into a \emph{column cluster tree} and a \emph{row cluster tree}, respectively.
 Here we describe construction of the column cluster tree; the row cluster tree is constructed similarly.
-
-Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\mathbf{\Aker}$ corresponds to the Lagrange node in $\mathbb{R}^\gdim$ associated with the $i\text{th}$ finite element basis function.
+%
+%Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\mathbf{\Aker}$ corresponds to the Lagrange node in $\mathbb{R}^\gdim$ associated with the $i\text{th}$ finite element basis function.
 The columns of the matrix thus correspond to a point cloud in $\mathbb{R}^\gdim$.
@@
 The root of the tree is the set of all columns, and the leaves of the tree are clusters of columns that are not subdivided any further.
-
-A depth-first search ordering of the column cluster tree leaves is then generated.
+%
+%A depth-first search ordering of the column cluster tree leaves is then generated.
 When the columns of the matrix are permuted into this depth-first ordering, the columns associated with each cluster in the cluster tree are contiguous.
-
-In the same way, the degrees of freedom associated with the rows of the matrix are hierarchically clustered into another cluster tree, and a depth-first search ordering for the rows is generated.
+%
+%In the same way, the degrees of freedom associated with the rows of the matrix are hierarchically clustered into another cluster tree, and a depth-first search ordering for the rows is generated.
 In our examples, the degrees of freedom for the columns coincide with the degrees of freedom for the rows, so the cluster trees for the rows and columns are the same, but this is not required in general.
-
-\subsubsection{Block cluster tree} 
-\label{sec:block_cluster_tree}
-
-We partition the matrix into a recursive hierarchy of mostly low rank blocks called the \emph{block cluster tree}.
+%
+%
+%\subsubsection{Block cluster tree} 
+%\label{sec:block_cluster_tree}
+%
+%We partition the matrix into a recursive hierarchy of mostly low rank blocks called the \emph{block cluster tree}.
 The idea is that a block of the matrix is likely to be low rank if the point cloud associated with the rows of the block is far away from the point cloud associated with the columns of the block.
@@
 Indeed, locality implies that many blocks of the matrix corresponding to far away point cloud clusters will be rank zero.
-
-After reordering the rows and columns of $\mathbf{\Aker}$ via the depth-first ordering described above, we partition the reordered version of $\mathbf{\Aker}$ into a tree of $2 \times 2$ block matrices recursively.
+%
+%After reordering the rows and columns of $\mathbf{\Aker}$ via the depth-first ordering described above, we partition the reordered version of $\mathbf{\Aker}$ into a tree of $2 \times 2$ block matrices recursively.
 We use a geometric admissibility condition (discussed below) to decide which blocks to subdivide, and use the cluster trees for the rows and columns to determine how to subdivide those blocks.
@@
 The matrix $\mathbf{\Aker}$ is partitioned into blocks as follows:
-\begin{equation*}
-	\begin{bmatrix}
-		\mathbf{\Aker}_{11} & \mathbf{\Aker}_{12} \\
-		\mathbf{\Aker}_{21} & \mathbf{\Aker}_{22}
-	\end{bmatrix},
-\end{equation*}
-where $\mathbf{\Aker}_{11}$ denotes the block of $\mathbf{\Aker}$ with rows $r_1$ and columns $\colcluster_1$, $\mathbf{\Aker}_{12}$ denotes the block of $\mathbf{\Aker}$ with rows $r_1$ and columns $\colcluster_2$, and so on for $\mathbf{\Aker}_{21}$ and $\mathbf{\Aker}_{22}$.
-
-We now loop through the four blocks, $\mathbf{\Aker}_{11}$, $\mathbf{\Aker}_{12}$, $\mathbf{\Aker}_{21}$, and $\mathbf{\Aker}_{22}$, and decide which blocks should be subdivided further.
+%\begin{equation*}
+%	\begin{bmatrix}
+%		\mathbf{\Aker}_{11} & \mathbf{\Aker}_{12} \\
+%		\mathbf{\Aker}_{21} & \mathbf{\Aker}_{22}
+%	\end{bmatrix},
+%\end{equation*}
+%where $\mathbf{\Aker}_{11}$ denotes the block of $\mathbf{\Aker}$ with rows $r_1$ and columns $\colcluster_1$, $\mathbf{\Aker}_{12}$ denotes the block of $\mathbf{\Aker}$ with rows $r_1$ and columns $\colcluster_2$, and so on for $\mathbf{\Aker}_{21}$ and $\mathbf{\Aker}_{22}$.
+%
+%We now loop through the four blocks, $\mathbf{\Aker}_{11}$, $\mathbf{\Aker}_{12}$, $\mathbf{\Aker}_{21}$, and $\mathbf{\Aker}_{22}$, and decide which blocks should be subdivided further.
 For the purpose of explanation, consider $\mathbf{\Aker}_{12}$.
 If
-\begin{equation}
-	\label{eq:weak_admissibility_cond}
-	\dist\left(r_1, \colcluster_2\right) \ge \weakadmconst \min\left(\diam\left(r_1\right), \diam\left(\colcluster_2\right)\right),
-\end{equation}
-then we mark $\mathbf{\Aker}_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone.
+%\begin{equation}
+%	\label{eq:weak_admissibility_cond}
+%	\dist\left(r_1, \colcluster_2\right) \ge \weakadmconst \min\left(\diam\left(r_1\right), \diam\left(\colcluster_2\right)\right),
+%\end{equation}
+%then we mark $\mathbf{\Aker}_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone.
 Here $\dist\left(r_1, \colcluster_2\right)$ is the Euclidean distance between the axis-aligned bounding box for the point cloud associated with the row cluster $r_1$, and the axis aligned bounding box for the point cloud associated with the column cluster $\colcluster_2$.
@@
 For us, the diagonal blocks $\mathbf{\Aker}_{11}$ and $\mathbf{\Aker}_{22}$ are not admissible because the distance between a point cloud and itself is zero.
-
-Next, we subdivide all blocks that are not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\mathbf{\Aker}$.
+%
+%Next, we subdivide all blocks that are not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\mathbf{\Aker}$.
 But now we subdivide a block based on the two child row clusters and two child column clusters for the rows and columns of that block.
@@
 The root of the tree is the whole matrix, internal nodes in the tree are blocks that are subdivided, and the leaves of the tree are blocks that are either expected to be low rank, or are small.
-
-\subsubsection{Adaptive cross low rank approximation of blocks}
-\label{sec:adaptive_cross}
-
-Once the block cluster tree has been constructed, low rank approximations of the admissible (low rank) blocks are formed using the adaptive cross method \cite{ACA}.
+%
+%
+%\subsubsection{Adaptive cross low rank approximation of blocks}
+%\label{sec:adaptive_cross}
+%
+%Once the block cluster tree has been constructed, low rank approximations of the admissible (low rank) blocks are formed using the adaptive cross method \cite{ACA}.
 Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\mathbf{\Aker}$.
 The idea of the adaptive cross method is to form a low rank approximation of $X$ by sampling a small number of rows and columns of $X$.
-\begin{itemize}
-	\item Let $C \in \mathbb{R}^{m \times \hrank}$ be a matrix consisting of a subset of $\hrank$ columns of $X$, such that the span of the columns in $C$ approximates the column space of $X$. 
-	\item Let $R\in\mathbb{R}^{\hrank \times m}$ be a subset of the rows of $X$, such that the span of the rows in $R$ approximates the row space of $X$.
-	\item Let $U \in \mathbb{R}^{\hrank \times \hrank}$ be the block of $X$ consisting of the intersection of the rows from $R$ with the columns from $C$.
-\end{itemize}
-Then it is well-established that
-\begin{equation}
-	\label{eq:CUR}
-	X \approx C U^+ R,
-\end{equation}
-where $U^+$ is the pseudoinverse of $U$ \cite{goreinov1997theory,MahoneyDrineas09}.
+%\begin{itemize}
+%	\item Let $C \in \mathbb{R}^{m \times \hrank}$ be a matrix consisting of a subset of $\hrank$ columns of $X$, such that the span of the columns in $C$ approximates the column space of $X$. 
+%	\item Let $R\in\mathbb{R}^{\hrank \times m}$ be a subset of the rows of $X$, such that the span of the rows in $R$ approximates the row space of $X$.
+%	\item Let $U \in \mathbb{R}^{\hrank \times \hrank}$ be the block of $X$ consisting of the intersection of the rows from $R$ with the columns from $C$.
+%\end{itemize}
+%Then it is well-established that
+%\begin{equation}
+%	\label{eq:CUR}
+%	X \approx C U^+ R,
+%\end{equation}
+%where $U^+$ is the pseudoinverse of $U$ \cite{goreinov1997theory,MahoneyDrineas09}.
 The quality of approximation \eqref{eq:CUR} depends on how well the columns of $C$ approximate the column space of $X$, and how well the rows of $R$ approximate the row space of $X$.
-
-In the adaptive cross method, ``good'' columns, $C$, and rows, $R$, are chosen via an alternating iterative process.
+%
+%In the adaptive cross method, ``good'' columns, $C$, and rows, $R$, are chosen via an alternating iterative process.
 An initial set of columns $C$ is chosen.
@@
 There is also a small linear algebra overhead cost for the determinant maximization process that is performed at each step.
-The key point is that the adaptive cross method allows us to form a rank-$\hrank$ approximation of an $m \times m$ block of the matrix via a process that only requires accessing $O(m\hrank)$ entries of that block.
-
-We use the adaptive cross method to form low rank approximations for each admissible block of $\mathbf{\Aker}$.
+%The key point is that the adaptive cross method allows us to form a rank-$\hrank$ approximation of an $m \times m$ block of the matrix via a process that only requires accessing $O(m\hrank)$ entries of that block.
+%
+%We use the adaptive cross method to form low rank approximations for each admissible block of $\mathbf{\Aker}$.
 We directly compute all entries of the small dense blocks of $\mathbf{\Aker}$ that are not admissible.
@@
 
-\section{Recycling Krylov information via low rank updates}
-\label{sec:recycle_krylov}
-
-In applications, often one wants to use the H-matrix approximation $\mathbf{A}_H$ to build preconditioners for solving a sequence of slowly varying linear systems, where the $k$th linear system has the form
-\begin{equation}
-\label{eq:Ax_equals_b}
-\mathbf{B}^{(k)} \mathbf{u}^{(k)} = \mathbf{f}^{(k)}.
-\end{equation}
-Each of these linear systems is solved using an iterative method. By ``slowly varying,'' we mean that $\mathbf{B}^{(k+1)}$ does not differ much from $\mathbf{B}^{(k)}$. 
-For example, in a Newton-Krylov method for solving an optimization problem, one uses a Krylov method such as conjugate gradient to solve a sequence of systems of the form \eqref{eq:Ax_equals_b}, where $\mathbf{B}^{(k)}$ and $\mathbf{f}^{(k)}$ are the Hessian and the negative gradient of the objective function in the optimization problem, respectively, evaluated at the $k$th Newton iterate.
-As the optimization solver converges, $\mathbf{B}^{(k)}$ converges to the Hessian at the optimal point, and therefore changes little from one iteration to the next.
+%\section{Recycling Krylov information via low rank updates}
+%\label{sec:recycle_krylov}
+%
+%In applications, often one wants to use the H-matrix approximation $\mathbf{A}_H$ to build preconditioners for solving a sequence of slowly varying linear systems, where the $k$th linear system has the form
+%\begin{equation}
+%\label{eq:Ax_equals_b}
+%\mathbf{B}^{(k)} \mathbf{u}^{(k)} = \mathbf{f}^{(k)}.
+%\end{equation}
+%Each of these linear systems is solved using an iterative method. By ``slowly varying,'' we mean that $\mathbf{B}^{(k+1)}$ does not differ much from $\mathbf{B}^{(k)}$. 
+%For example, in a Newton-Krylov method for solving an optimization problem, one uses a Krylov method such as conjugate gradient to solve a sequence of systems of the form \eqref{eq:Ax_equals_b}, where $\mathbf{B}^{(k)}$ and $\mathbf{f}^{(k)}$ are the Hessian and the negative gradient of the objective function in the optimization problem, respectively, evaluated at the $k$th Newton iterate.
+%As the optimization solver converges, $\mathbf{B}^{(k)}$ converges to the Hessian at the optimal point, and therefore changes little from one iteration to the next.
 Since $\mathbf{B}^{(k)}$ varies slowly, it is reasonable to assume that Krylov information about $\mathbf{B}^{(k)}$ can be leveraged to update the preconditioner from $\mathbf{B}^{(k)}$ in order to build a preconditioner for $\mathbf{B}^{(k+1)}$.
-Concretely, let 
-\begin{equation*}
-\mathbf{P}^{(k)} \approx \left(\mathbf{B}^{(k)}\right)^{-1}
-\end{equation*}
-denote a H-matrix preconditioner used to accelerate the solution of \eqref{eq:Ax_equals_b} with an iterative method.
+%Concretely, let 
+%\begin{equation*}
+%\mathbf{P}^{(k)} \approx \left(\mathbf{B}^{(k)}\right)^{-1}
+%\end{equation*}
+%denote a H-matrix preconditioner used to accelerate the solution of \eqref{eq:Ax_equals_b} with an iterative method.
 The iterative method for solving the $k$th linear system will typically involve, as a sub-problem, applying $\mathbf{B}^{(k)}$ to a collection of vectors $\{\mathbf{x}_i\}_{i=1}^m$, thereby generating vectors $\{\mathbf{y}_i = \mathbf{B}^{(k)} \mathbf{x}_i\}_{i=1}^m$.
@@
 The matrix-vector products with $\mathbf{B}^{(k)}$ performed while solving the $k$th linear system may be summarized by the equation:
-\begin{equation*}
-\mathbf{B}^{(k)}\mathbf{X} = \mathbf{Y}.
-\end{equation*}
-To form $\mathbf{P}^{(k+1)}$ from $\mathbf{P}^{(k)}$, one may use one of the following three update formulas:
-\begin{align}
-	\mathbf{P}^{(k+1)} &= \mathbf{P}^{(k)} + \mathbf{R}\left(\mathbf{Y}^T \mathbf{Y}\right)^{-1} \mathbf{Y}^T \label{eq:res_update} \\
-	\mathbf{P}^{(k+1)} &= \mathbf{P}^{(k)} + \mathbf{R} \left(\mathbf{R}^T \mathbf{Y}\right)^{-1} \mathbf{R}^T \label{eq:srk_update} \\
-	\mathbf{P}^{(k+1)} &= \left(\mathbf{I} - \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T\right)\mathbf{P}^{(k)}\left(\mathbf{I} - \mathbf{Y}\boldsymbol{\Theta}\mathbf{X}^T\right) + \mathbf{X}\boldsymbol{\Theta}\mathbf{X}^T, \label{eq:dfp_update}
-\end{align}
-where $\mathbf{R} := \mathbf{X}-\mathbf{P}^{(k)}\mathbf{Y}$ and $\boldsymbol{\Theta} := \left(\mathbf{X}^T\mathbf{Y}\right)^{-1}$.
-We note that for all three cases, the update is exact for $\mathbf{B}^{k}$, in the sense that 
-\begin{equation*}
-\mathbf{P}^{(k+1)}\mathbf{Y}=\mathbf{X} = \left(\mathbf{B}^{(k)}\right)^{-1} \mathbf{Y}.
-\end{equation*}
-Because the matrices $\mathbf{B}^{k}$ vary slowly, we expect that $\mathbf{P}^{(k+1)}$ is a good approximation for $\left(\mathbf{B}^{(k+1)}\right)^{-1}$ on the space spanned by the vectors $\mathbf{y}_i$.
+%\begin{equation*}
+%\mathbf{B}^{(k)}\mathbf{X} = \mathbf{Y}.
+%\end{equation*}
+%To form $\mathbf{P}^{(k+1)}$ from $\mathbf{P}^{(k)}$, one may use one of the following three update formulas:
+%\begin{align}
+%	\mathbf{P}^{(k+1)} &= \mathbf{P}^{(k)} + \mathbf{R}\left(\mathbf{Y}^T \mathbf{Y}\right)^{-1} \mathbf{Y}^T \label{eq:res_update} \\
+%	\mathbf{P}^{(k+1)} &= \mathbf{P}^{(k)} + \mathbf{R} \left(\mathbf{R}^T \mathbf{Y}\right)^{-1} \mathbf{R}^T \label{eq:srk_update} \\
+%	\mathbf{P}^{(k+1)} &= \left(\mathbf{I} - \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T\right)\mathbf{P}^{(k)}\left(\mathbf{I} - \mathbf{Y}\boldsymbol{\Theta}\mathbf{X}^T\right) + \mathbf{X}\boldsymbol{\Theta}\mathbf{X}^T, \label{eq:dfp_update}
+%\end{align}
+%where $\mathbf{R} := \mathbf{X}-\mathbf{P}^{(k)}\mathbf{Y}$ and $\boldsymbol{\Theta} := \left(\mathbf{X}^T\mathbf{Y}\right)^{-1}$.
+%We note that for all three cases, the update is exact for $\mathbf{B}^{k}$, in the sense that 
+%\begin{equation*}
+%\mathbf{P}^{(k+1)}\mathbf{Y}=\mathbf{X} = \left(\mathbf{B}^{(k)}\right)^{-1} \mathbf{Y}.
+%\end{equation*}
+%Because the matrices $\mathbf{B}^{k}$ vary slowly, we expect that $\mathbf{P}^{(k+1)}$ is a good approximation for $\left(\mathbf{B}^{(k+1)}\right)^{-1}$ on the space spanned by the vectors $\mathbf{y}_i$.
 One should choose which formula to use based on the properties of the matrix $\mathbf{B}^{(k)}$.
@@
 Formulas \eqref{eq:res_update} and \eqref{eq:srk_update} perform rank-$m$ updates, while formula \eqref{eq:dfp_update} performs a rank-$2m$ update.
-We note that when $m=1$, formulas \eqref{eq:srk_update} are \eqref{eq:dfp_update} are known as SR1 (symmetric rank-$1$) and DFP (Davidson-Fletcher-Powell) updates, respectively \cite[Chapter 6]{NocedalWright99}.
+%We note that when $m=1$, formulas \eqref{eq:srk_update} are \eqref{eq:dfp_update} are known as SR1 (symmetric rank-$1$) and DFP (Davidson-Fletcher-Powell) updates, respectively \cite[Chapter 6]{NocedalWright99}.
 

### 2022-07-18 paper/localpsf.tex
ellipsoid intersection appendix improvements: bounding box precheck

@@
 \label{eq:operator_simultaneous_ips}
-\left(\Aop^T f\right)^*(x) = \int_\Omega K(y,x) f(y) dy = \left(\phi_x, f\right)_{L^2(\Omega)}.
+\left(\Aop^T f\right)^*(x) = \int_\Omega \Aker(y,x) f(y) dy = \left(\phi_x, f\right)_{L^2(\Omega)}.
 \end{equation}
@@
 \end{algorithm2e}
+
+We may further reduce the cost of performing many ellipsoid intersection tests by first checking whether the coordinate axis aligned bounding boxes for the ellipsoids intersect.
+Checking whether two axis-aligned bounding boxes intersect is extremely cheap.
+If the bounding boxes do not intersect, then the ellipsoids do not intersect, so no further action is necessary.
+If the bounding boxes do intersect, the ellipsoids may or may not intersect, so one must perform the more expensive ellipsoid intersection test described above
 

### 2022-07-19 paper/localpsf.tex
ellipsoid intersection appendix improvements: bounding box precheck

--- 2022-07-18 paper/localpsf.tex
@@
 \end{algorithm2e}
+
+\subsection{Ellipsoid bounding box test}
+
+We may reduce the average cost of performing ellipsoid intersection tests by first checking whether the coordinate axis aligned bounding boxes for the ellipsoids intersect.
+If the bounding boxes do not intersect, then the ellipsoids do not intersect, so no further action is necessary.
+If the bounding boxes intersect, the ellipsoids might intersect, or they might not intersect, so one must perform the complete ellipsoid intersection test described in Section \ref{sec:fast_ellipsoid_intersection_test}.
+
+The ellipsoid 
+$
+E = \{x : (x - \spatialmean)^T \spatialcov^{-1} (x - \spatialmean) \le \tau^2\}
+$
+is bounded by the box
+\begin{equation*}
+	\prod_{i=1}^\gdim \left[ \spatialmean_i -\sqrt{\Sigma_{ii}}/\tau, \spatialmean_i +\sqrt{\Sigma_{ii}}/\tau \right],
+\end{equation*}
+where $\Sigma_{ii}$ is the $i$th diagonal entry of $\Sigma$\footnote{See \url{https://math.stackexchange.com/q/3928964}}. The boxes
+\begin{equation*}
+	\prod_{i=1}^\gdim \left[a_i^-, a_i^+\right], \quad \text{and} \quad \prod_{i=1}^\gdim \left[b_i^-, b_i^+\right]
+\end{equation*}
+intersect if and only if $a_i^- \le b_i^+$ and $b_i^- \le a_i^+$ for $i=1,\dots,d$\footnote{See \url{https://stackoverflow.com/a/65107290/484944}}.
 

### 2022-07-19 paper/gradVsStokesSolvesPlot.tex
standalone tikz files to be used with \include to generate various tables and tikz figures

\begin{figure}[H]
\begin{center} 
\begin{tikzpicture}[baseline, trim axis right]
\begin{axis}[width=0.99\textwidth, height=0.75\textwidth,
            ymode=log,
            ymin=5e-9,
            ymax=1,
            xmax=900,
            xtick={0, 100, 200, 300, 400, 500, 600, 700, 800, 900},
            xlabel={Cumulative Stokes solves},
            ylabel={$||g_{k}||_{2}/||g_{0}||_{2}$, relative norm of the gradient},
            title={Convergence of the preconditioned inexact Newton-CG},
			legend pos=north east,
]
\addlegendentry{None}
\addplot[dash dot, line width=0.5mm] table {gradVsCumulativeStokes_none.dat};
\addlegendentry{Reg}
\addplot[dotted, line width=0.5mm] table {gradVsCumulativeStokes_reg.dat};
\addlegendentry{PSF (5)}
\addplot[red, line width=0.5mm] table {gradVsCumulativeStokes_psf.dat};
\end{axis}
\end{tikzpicture}
\end{center}
\end{figure}

### 2022-07-20 paper/localpsf.tex
add missing impulse response files

@@
 
-We may reduce the average cost of performing ellipsoid intersection tests by first checking whether the coordinate axis aligned bounding boxes for the ellipsoids intersect.
+One may reduce the average cost of performing ellipsoid intersection tests by first checking whether the coordinate axis aligned bounding boxes for the ellipsoids intersect.
 If the bounding boxes do not intersect, then the ellipsoids do not intersect, so no further action is necessary.
@@
 
-The ellipsoid 
+The smallest coordinate axis aligned box that contains the ellipsoid 
 $
@@
 $
-is bounded by the box
-\begin{equation*}
-	\prod_{i=1}^\gdim \left[ \spatialmean_i -\sqrt{\Sigma_{ii}}/\tau, \spatialmean_i +\sqrt{\Sigma_{ii}}/\tau \right],
+is given as follows:
+\begin{equation*}
+	\prod_{i=1}^\gdim \left[ \spatialmean_i -\tau\sqrt{\Sigma_{ii}}, \spatialmean_i +\tau\sqrt{\Sigma_{ii}} \right],
 \end{equation*}
@@
 \end{equation*}
-intersect if and only if $a_i^- \le b_i^+$ and $b_i^- \le a_i^+$ for $i=1,\dots,d$\footnote{See \url{https://stackoverflow.com/a/65107290/484944}}.
+intersect if and only if $a_i^- \le b_i^+$ and $b_i^- \le a_i^+$ for $i=1,\dots,d$.
 

### 2022-07-20 paper/NewtonIteratePlot.tex

\begin{figure}[H]
	\begin{center}
	\begin{tabular}[c]{c c}
		\begin{tabular}[c]{c c c c}
			\includegraphics[scale=.1125]{m0_reg.png} &
			\includegraphics[scale=.1125]{m3_reg.png} &
			\includegraphics[scale=.1125]{m4_reg.png} &
			\includegraphics[scale=.1125]{m5_reg.png} \\
			\includegraphics[scale=.1125]{m0_none.png} &
			\includegraphics[scale=.1125]{m3_none.png} &
			\includegraphics[scale=.1125]{m4_none.png} &
			\includegraphics[scale=.1125]{m5_none.png} \\
			\includegraphics[scale=.1125]{m0_psf.png} &
			\includegraphics[scale=.1125]{m3_psf.png} &
			\includegraphics[scale=.1125]{m4_psf.png} &
			\includegraphics[scale=.1125]{m5_psf.png} 
		\end{tabular} 
		&
		\begin{tabular}[c]{c}
		 \\
		\includegraphics[scale=.25]{mColorBar.png} \\
		\end{tabular}
	\end{tabular}
	\caption{The initial parameter (left column), the parameter after three (second column)- four (third column) and five (right column) Newton iteration, the parameter after four Newton iterations for regularization preconditioning (upper row), no preconditioning (middle row) and psf preconditioning (lower row).}
	\end{center}
\end{figure}

### 2022-07-20 paper/PreconditionedSpectrumPlot.tex

\begin{figure}[H]
\begin{center} 
\begin{tikzpicture}[baseline, trim axis right]
\begin{axis}[width=0.99\textwidth, height=0.75\textwidth,
            ymode=log,
            xmin=-50,
            xmax=1450,
            xlabel={$k$, generalized eigenvalue \#},
            ylabel={$\lambda_{k}$},
            xtick={0, 200, 400, 600, 800, 1000, 1200, 1400},
			legend pos=north east,
			title={Generalized eigenvalues, $Hu_{k}=\lambda_{k}Bu_{k}$}
]
\addlegendentry{REG}
\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues.dat};
\addlegendentry{PSF (1)}
\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues.dat};
\addlegendentry{PSF (5)}
\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues.dat};
\addlegendentry{PSF (25)}
\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues.dat};
\end{axis}
\end{tikzpicture}
\end{center}
\end{figure}

### 2022-07-20 paper/RegularizationRobustnessPlot.tex

\begin{figure}[H]
\begin{center} 
\begin{tikzpicture}[baseline, trim axis right]
\begin{axis}[width=0.99\textwidth, height=0.75\textwidth,
            xmode=log,
            ymode=log,
            xlabel={$\gamma$, regularization parameter},
            ylabel={\# of CG iterations},
            %legend pos=north east,
			title={CG iterations to achieve $||H x-b||/||b||<10^{-6}$}
]
\addlegendentry{REG}
\addplot[dotted, line width=0.5mm] table[x=gamma, y=reg] {CGitsvsGamma.dat};
\addlegendentry{None}
\addplot[dash dot, line width=0.5mm] table[x=gamma, y=none] {CGitsvsGamma.dat};
\addlegendentry{PSF (1)}
\addplot[blue, line width=0.5mm] table[x=gamma, y=psf1] {CGitsvsGamma.dat};
\addlegendentry{PSF (5)}
\addplot[red, line width=0.5mm] table[x=gamma, y=psf5] {CGitsvsGamma.dat};
\addlegendentry{PSF (25)}
\addplot[purple, line width=0.5mm] table[x=gamma, y=psf25] {CGitsvsGamma.dat};
\end{axis}
\end{tikzpicture}
\end{center}
\end{figure}

### 2022-07-20 paper/localpsf.tex
pgfplots to shared and plots incorporated into main tex file

@@
 Some numerical results here.
+
+\begin{figure}[H]
+	\begin{center}
+		\begin{tabular}[c]{c c c}
+			%\includegraphics[scale=.15]{recoveredPressure.png} &
+			%\includegraphics[scale=.15]{truePressure.png} &
+			%\includegraphics[scale=.18]{pressureColorBar.png} \\
+			\includegraphics[scale=.15]{recoveredVelocity_glyphs.png} &
+			\includegraphics[scale=.15]{trueVelocity_glyphs.png} &
+			\includegraphics[scale=.18]{velocityColorBar.png} \\
+			\includegraphics[scale=.15]{recoveredVelocity_noglyphs.png} &
+			\includegraphics[scale=.15]{trueVelocity_noglyphs.png} &
+			\includegraphics[scale=.18]{velocityColorBar.png}
+		\end{tabular}
+		\caption{Reconstructed (left column) and true (right column) pressure (top row) and velocity fields (middle and bottom rows).}
+	\end{center}
+\end{figure}
+
+\begin{table}[H]
+	\begin{center}
+		\begingroup
+		\setlength{\tabcolsep}{8pt}
+		\renewcommand{\arraystretch}{1.25}
+		\begin{tabular}{c| c c c | c c c | c c c}
+			&  \multicolumn{3}{|c|}{PSF} & \multicolumn{3}{|c|}{REG} & \multicolumn{3}{|c}{NONE} \\
+			\hline
+			Iter & 
+			\#CG & \#Stokes & $||g||$ & 
+			\#CG & \#Stokes & $||g||$ & 
+			\#CG & \#Stokes & $||g||$ \\
+			\verb~0~ &
+			\verb~1~ & \verb~4~ & \verb~1.95e+7~ &
+			\verb~1~ & \verb~4~ & \verb~1.95e+7~ &
+			\verb~1~ & \verb~4~ & \verb~1.95e+7~ \\
+			\verb~1~ &
+			\verb~2~ & \verb~6~ & \verb~5.76e+6~ &
+			\verb~2~ & \verb~6~ & \verb~7.90e+6~ &
+			\verb~2~ & \verb~6~ & \verb~5.76e+6~ \\
+			\verb~2~ &
+			\verb~4~ & \verb~10~ & \verb~2.36e+6~ &
+			\verb~5~ & \verb~12~ & \verb~3.87e+6~ &
+			\verb~4~ & \verb~10~ & \verb~2.36e+6~ \\
+			\verb~3~ &
+			\verb~2~ & \verb~6+22~ & \verb~5.79e+5~ &
+			\verb~13~ & \verb~28~ & \verb~1.62e+6~ &
+			\verb~13~ & \verb~28~ & \verb~5.79e+5~ \\
+			\verb~4~ &
+			\verb~5~ & \verb~12~ & \verb~5.61e+4~ &
+			\verb~42~ & \verb~86~ & \verb~4.78e+5~ &
+			\verb~33~ & \verb~68~ & \verb~9.05e+4~ \\
+			\verb~5~ &
+			\verb~10~ & \verb~22~ & \verb~3.45e+3~ &
+			\verb~84~ & \verb~170~ & \verb~7.72e+4~ &
+			\verb~57~ & \verb~116~ & \verb~6.21e+3~ \\
+			\verb~6~ &
+			\verb~14~ & \verb~30~ & \verb~2.66e+1~ &
+			\verb~125~ & \verb~252~ & \verb~5.04e+3~ &
+			\verb~76~ & \verb~154~ & \verb~1.05e+2~ \\
+			\verb~7~ &
+			\verb~0~ & \verb~2~ & \verb~4.09e-2~ &
+			\verb~194~ & \verb~390~ & \verb~7.91e+1~ &
+			\verb~117~ & \verb~236~ & \verb~2.29e-1~ \\
+			\verb~8~ & 
+			\verb~---~ & \verb~---~ & \verb~---~ &
+			\verb~0~ & \verb~2~ & \verb~1.60e-1~ &
+			\verb~0~ & \verb~2~ & \verb~4.08e-2~ \\
+			\hline
+			Total & 
+			\verb~38~ & \verb~114~ & \verb~---~ &
+			\verb~466~ & \verb~950~ & \verb~---~ &
+			\verb~303~ & \verb~624~ & \verb~---~ \\
+		\end{tabular}
+		\endgroup
+	\end{center}
+\end{table}
+
+\begin{figure}[H]
+\begin{subfigure}{0.49\textwidth}
+	\centering
+	\begin{tikzpicture}[baseline, trim axis right]
+	\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
+	ymode=log,
+	ymin=1e-11,
+	ymax=2,
+	xmax=750,
+	axis x line*=bottom,
+	axis y line*=left,
+	xtick={0, 200, 400, 600},
+	xlabel={CG iteration},
+	ylabel={Relative residual norm},
+	title={CG convergence},
+	legend style={at={(1.1,1)},anchor=north east},
+	%legend pos=north east,
+	legend style={font=\small}
+	]
+	\addlegendentry{None}
+	\addplot[dash dot, line width=0.5mm] table {cg_relres_none.dat};
+	\addlegendentry{Reg}
+	\addplot[dotted, line width=0.5mm] table {cg_relres_reg.dat};
+	\addlegendentry{PSF (1)}
+	\addplot[blue, line width=0.5mm] table {cg_relres_p1.dat};
+	\addlegendentry{PSF (5)}
+	\addplot[red, line width=0.5mm] table {cg_relres_p5.dat};
+	\addlegendentry{PSF (25)}
+	\addplot[purple, line width=0.5mm] table {cg_relres_p25.dat};
+	\end{axis}
+	\end{tikzpicture}
+\end{subfigure}
+	\begin{subfigure}{0.49\textwidth}
+		\centering
+		\begin{tikzpicture}[baseline, trim axis right]
+		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
+		xmode=log,
+		ymode=log,
+		xlabel={$\gamma$, regularization parameter},
+		ylabel={\# of CG iterations},
+		axis y line*=right,
+		axis x line*=bottom,
+		ylabel near ticks, yticklabel pos=right,
+		%ylabel near ticks, ylabel shift={-6pt},
+		%legend pos=north east,
+		title={Regularization robustness}
+		%title={CG iterations to achieve $||H x-b||/||b||<10^{-6}$}
+		]
+		%\addlegendentry{None}
+		\addplot[dash dot, line width=0.5mm] table[x=gamma, y=none] {CGitsvsGamma.dat};
+		%\addlegendentry{REG}
+		\addplot[dotted, line width=0.5mm] table[x=gamma, y=reg] {CGitsvsGamma.dat};
+		%\addlegendentry{PSF (1)}
+		\addplot[blue, line width=0.5mm] table[x=gamma, y=psf1] {CGitsvsGamma.dat};
+		%\addlegendentry{PSF (5)}
+		\addplot[red, line width=0.5mm] table[x=gamma, y=psf5] {CGitsvsGamma.dat};
+		%\addlegendentry{PSF (25)}
+		\addplot[purple, line width=0.5mm] table[x=gamma, y=psf25] {CGitsvsGamma.dat};
+		\end{axis}
+		\end{tikzpicture}
+	\end{subfigure}
+\end{figure}
+
+\begin{figure}[H]
+	\begin{subfigure}{0.59\textwidth}
+	\begin{center}
+		\begin{tikzpicture}[baseline, trim axis right]
+		\begin{axis}[width=0.99\linewidth, height=0.8\linewidth,
+		ymode=log,
+		xmin=-50,
+		xmax=1450,
+		ymax=1.e4,
+		xlabel={$k$, generalized eigenvalue \#},
+		ylabel={$\lambda_{k}$},
+		xtick={0, 300, 600, 900, 1200},
+		%legend pos=north east,
+		legend style={font=\small},
+		ylabel near ticks, ylabel shift={-6pt},
+		axis x line*=bottom,
+		axis y line*=left,
+		legend style={at={(0.95,1)},anchor=north east},
+		title={Generalized eigenvalues, $Hu_{k}=\lambda_{k}Pu_{k}$}
+		]
+		\addlegendentry{REG}
+		\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues.dat};
+		\addlegendentry{PSF (1)}
+		\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues.dat};
+		\addlegendentry{PSF (5)}
+		\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues.dat};
+		\addlegendentry{PSF (25)}
+		\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues.dat};
+		\end{axis}
+		\end{tikzpicture}
+	\end{center}
+\end{subfigure}
+\begin{subfigure}{0.39\textwidth}
+		\begin{center}
+		\begingroup
+		\setlength{\tabcolsep}{4pt}
+		\renewcommand{\arraystretch}{1.25}
+		\begin{tabular}{c| c c}
+			& $\kappa(P^{-1}H)$ &	$\kappa(P^{-1}H_{\text{gn}})$ \\
+			\hline 
+			REG        & \verb~7.0e+3~ & \verb~7.0e+3~ \\
+			NONE       & \verb~8.7e+2~ & \verb~9.2e+2~ \\
+			PSF $(1)$  & \verb~1.8e+1~ & \verb~1.8e+1~ \\
+			PSF $(5)$  & \verb~1.4e+1~ & \verb~1.4e+1~ \\
+			PSF $(25)$ & \verb~1.0e+1~ & \verb~1.1e+1~
+		\end{tabular}
+		\endgroup
+	\end{center}
+\end{subfigure}
+\end{figure}
 

### 2022-07-21 paper/localpsf.tex
consistent ordering of plots and tables (psf, reg, none)
added m1_psf, m3_psf, m4_psf, mfinal_psf Newton iterates, incorporated them into the tex file, also added images with ColorBars so that we do not need to have 6 includegraphics calls for the ice mountain, true velocity and true parameter plots. I need to adjust these so that the colorbar is the same size on all three but this will require redoing the paraview plots and I would like to defer that effort for a later date

@@
 \begin{figure}[H]
-	\begin{center}
-		\begin{tabular}[c]{c c c}
-			%\includegraphics[scale=.15]{recoveredPressure.png} &
-			%\includegraphics[scale=.15]{truePressure.png} &
-			%\includegraphics[scale=.18]{pressureColorBar.png} \\
-			\includegraphics[scale=.15]{recoveredVelocity_glyphs.png} &
-			\includegraphics[scale=.15]{trueVelocity_glyphs.png} &
-			\includegraphics[scale=.18]{velocityColorBar.png} \\
-			\includegraphics[scale=.15]{recoveredVelocity_noglyphs.png} &
-			\includegraphics[scale=.15]{trueVelocity_noglyphs.png} &
-			\includegraphics[scale=.18]{velocityColorBar.png}
-		\end{tabular}
-		\caption{Reconstructed (left column) and true (right column) pressure (top row) and velocity fields (middle and bottom rows).}
-	\end{center}
+	\begin{subfigure}{0.275\textwidth}
+    \includegraphics[scale=0.10]{meshHeight_view0_edges_withColorBar.png}	
+    \end{subfigure}
+    \begin{subfigure}{0.275\textwidth}
+    \includegraphics[scale=0.10]{trueVelocity_glyphs_withColorBar.png}
+    \end{subfigure}
+    \begin{subfigure}{0.275\textwidth}
+    \includegraphics[scale=0.135]{mtrue_withColorBar.png}
+    \end{subfigure}
 \end{figure}
@@
 		\begingroup
-		\setlength{\tabcolsep}{8pt}
+		\setlength{\tabcolsep}{4pt}
 		\renewcommand{\arraystretch}{1.25}
@@
 \begin{figure}[H]
+	\begin{subfigure}{0.225\textwidth}
+		\includegraphics[scale=0.08]{m1_psf.png}	
+	\end{subfigure}
+	\begin{subfigure}{0.225\textwidth}
+		\includegraphics[scale=0.08]{m3_psf.png}
+	\end{subfigure}
+	\begin{subfigure}{0.225\textwidth}
+		\includegraphics[scale=0.08]{m4_psf.png}
+	\end{subfigure}
+	\begin{subfigure}{0.225\textwidth}
+	\includegraphics[scale=0.08]{m6_psf.png}
+\end{subfigure}
+\begin{subfigure}{0.05\textwidth}
+	\includegraphics[scale=0.1]{mColorBar.png}
+\end{subfigure}
+\caption{The parameter after $1$ (left), $3$ (second from left), $4$ (third from left) Newton iterates and the final Newton iterate (right).} 
+\end{figure}
+
+\begin{figure}[H]
 \begin{subfigure}{0.49\textwidth}
@@
 	]
-	\addlegendentry{None}
-	\addplot[dash dot, line width=0.5mm] table {cg_relres_none.dat};
-	\addlegendentry{Reg}
-	\addplot[dotted, line width=0.5mm] table {cg_relres_reg.dat};
 	\addlegendentry{PSF (1)}
@@
     \addplot[purple, line width=0.5mm] table {cg_relres_p25.dat};
+	\addlegendentry{Reg}
+    \addplot[dotted, line width=0.5mm] table {cg_relres_reg.dat};
+	\addlegendentry{None}
+	\addplot[dash dot, line width=0.5mm] table {cg_relres_none.dat};
 	\end{axis}
@@
 		]
-		\addlegendentry{REG}
-		\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues.dat};
 		\addlegendentry{PSF (1)}
@@
         \addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues.dat};
+		\addlegendentry{REG}
+		\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues.dat};
 		\end{axis}
@@
 			\hline 
-			REG        & \verb~7.0e+3~ & \verb~7.0e+3~ \\
-			NONE       & \verb~8.7e+2~ & \verb~9.2e+2~ \\
 			PSF $(1)$  & \verb~1.8e+1~ & \verb~1.8e+1~ \\
             PSF $(5)$  & \verb~1.4e+1~ & \verb~1.4e+1~ \\
-			PSF $(25)$ & \verb~1.0e+1~ & \verb~1.1e+1~
+            PSF $(25)$ & \verb~1.0e+1~ & \verb~1.1e+1~ \\
+			REG        & \verb~7.0e+3~ & \verb~7.0e+3~ \\
+			NONE       & \verb~8.7e+2~ & \verb~9.2e+2~ 
 		\end{tabular}

### 2022-07-21 paper/localpsf.tex
some paper updates

@@
 	\begin{center}
-		\includegraphics[scale=1.0]{mean_displacement_invariance.pdf}
+		\includegraphics[scale=0.75]{mean_displacement_invariance.pdf}
 	\end{center}
@@
 \begin{figure}[H]
-	\begin{subfigure}{0.275\textwidth}
-    \includegraphics[scale=0.10]{meshHeight_view0_edges_withColorBar.png}	
-    \end{subfigure}
-    \begin{subfigure}{0.275\textwidth}
-    \includegraphics[scale=0.10]{trueVelocity_glyphs_withColorBar.png}
-    \end{subfigure}
-    \begin{subfigure}{0.275\textwidth}
-    \includegraphics[scale=0.135]{mtrue_withColorBar.png}
+	\begin{subfigure}{0.32\textwidth}
+    \includegraphics[scale=0.12]{meshHeight_view0_edges_withColorBar.png}	
+    \end{subfigure}
+    \begin{subfigure}{0.32\textwidth}
+    \includegraphics[scale=0.12]{trueVelocity_glyphs_withColorBar.png}
+    \end{subfigure}
+    \begin{subfigure}{0.32\textwidth}
+    \includegraphics[scale=0.162]{mtrue_withColorBar.png}
     \end{subfigure}
@@
 \end{equation*}
-In \cite{GilitschenskiHanebeck12} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$.
-We check whether $E_1$ and $E_2$ intersect by minimizing $K(s)$ on $(0,1)$.
-If $K(s^*) <0$ at the minimizer $s^*$, then $E_1 \cap E_2 = \{\}$.
-Otherwise $E_1 \cap E_2 \neq \{\}$.
+In \cite{GilitschenskiHanebeck12} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$ (see Figure \ref{fig:ellipsoid_intersection_test}).
 
 \begin{figure}
-	\begin{subfigure}{0.32\textwidth}
+	\begin{subfigure}{0.37\textwidth}
 		\centering
-		\includegraphics[scale=0.40]{ellipsoids_intersect0.pdf}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[scale=1.0]{ellipsoids_intersect0.pdf}
+	\end{subfigure}
+	\begin{subfigure}{0.30\textwidth}
 		\centering
-		\includegraphics[scale=0.40]{ellipsoids_intersect1.pdf}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[scale=1.0]{ellipsoids_intersect1.pdf}
+	\end{subfigure}
+	\begin{subfigure}{0.30\textwidth}
 		\centering
-		\includegraphics[scale=0.40]{ellipsoids_intersect2.pdf}
+		\includegraphics[scale=1.0]{ellipsoids_intersect2.pdf}
 	\end{subfigure}
@@
 where $v := P^T\left(\spatialmean_1 - \spatialmean_2\right)$.
-We compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's algorithm \cite{Brent71} (any fast 1 dimensional convex optimization routine may be used).
-The resulting algorithm for checking whether $E_1$ and $E_2$ intersect is summarized in Algorithm \ref{alg:ellipsoid_intersection_test}.
-
-\begin{algorithm2e}
-	\SetAlgoNoLine
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Output}
-	\SetKwProg{Fn}{Function}{}{}
-	\Input{Ellipsoid $E_1$ with mean $\spatialmean_1$ and covariance $\spatialcov_1/\tau^2$\\
-		Ellipsoid $E_2$ with mean $\spatialmean_2$ and covariance $\spatialcov_2/\tau^2$ 
-	}
-
-	\Output{Boolean $\text{ellipsoids\_intersect}$ which is true if $E_1 \cap E_2 \neq \{\}$ and false otherwise}
-
-	Solve generalized eigenvalue problem $\spatialcov_1 P = \spatialcov_2 P \Lambda$
-
-	$v \gets P^T\left(\spatialmean_1 - \spatialmean_2\right)$
-
-	$\displaystyle K^* \gets \min_{s \in (0,1)}~1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2$
-
-	\If{$K^* < 0$}{
-
-		$\text{ellipsoids\_intersect} \gets \text{False}$
-
-	}
-	\Else{
-
-		$\text{ellipsoids\_intersect} \gets \text{True}$
-
-	}
-	\caption{Determining whether two ellipsoids intersect}
-	\label{alg:ellipsoid_intersection_test}
-\end{algorithm2e}
+
+To check whether $E_1$ and $E_2$ intersect, we compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, form $v$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's minimization algorithm \cite{Brent71} (any fast 1 dimensional convex optimization routine may be used).
+If $K(s^*) <0$ at the minimizer $s^*$, then $E_1 \cap E_2 = \{\}$.
+Otherwise $E_1 \cap E_2 \neq \{\}$.
+%The resulting algorithm for checking whether $E_1$ and $E_2$ intersect is summarized in Algorithm \ref{alg:ellipsoid_intersection_test}.
+
+%\begin{algorithm2e}
+%	\SetAlgoNoLine
+%	\SetKwInOut{Input}{Input}
+%	\SetKwInOut{Output}{Output}
+%	\SetKwProg{Fn}{Function}{}{}
+%	\Input{Ellipsoid $E_1$ with mean $\spatialmean_1$ and covariance $\spatialcov_1/\tau^2$\\
+%		Ellipsoid $E_2$ with mean $\spatialmean_2$ and covariance $\spatialcov_2/\tau^2$ 
+%	}
+%	
+%	\Output{Boolean $\text{ellipsoids\_intersect}$ which is true if $E_1 \cap E_2 \neq \{\}$ and false otherwise}
+%	
+%	
+%	Solve generalized eigenvalue problem $\spatialcov_1 P = \spatialcov_2 P \Lambda$
+%		
+%	$v \gets P^T\left(\spatialmean_1 - \spatialmean_2\right)$
+%		
+%	$\displaystyle K^* \gets \min_{s \in (0,1)}~1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2$
+%		
+%	\If{$K^* < 0$}{
+%			
+%		$\text{ellipsoids\_intersect} \gets \text{False}$
+%			
+%	}
+%	\Else{
+%			
+%		$\text{ellipsoids\_intersect} \gets \text{True}$
+%			
+%	}
+%	\caption{Determining whether two ellipsoids intersect}
+%	\label{alg:ellipsoid_intersection_test}
+%\end{algorithm2e}
 
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[scale=0.5]{spd_rational_function.pdf}
-	\caption{Rational function $\Pi_\ratord(\lambda)$ for $a=-1$, $b=0$, and $\ratord \in \{1,2,3\}$. As $\ratord$ increases, $\Pi_\ratord$ approaches the indicator function for the set $(a,b)$. }
-	\label{fig:spd_rat}
-\end{figure}
+%\begin{figure}
+%	\centering
+%	\includegraphics[scale=0.5]{spd_rational_function.pdf}
+%	\caption{Rational function $\Pi_\ratord(\lambda)$ for $a=-1$, $b=0$, and $\ratord \in \{1,2,3\}$. As $\ratord$ increases, $\Pi_\ratord$ approaches the indicator function for the set $(a,b)$. }
+%	\label{fig:spd_rat}
+%\end{figure}
 
@@
 Here $\ratord$ is a positive integer; the larger $\ratord$ is, the better the approximation $\Pi \approx \Pi_\ratord$. 
-This is illustrated in Figure \ref{fig:spd_rat}.
+%This is illustrated in Figure \ref{fig:spd_rat}. 
 Replacing scalar operations with matrix operations, we have an approximation $\mathbf{\Pi}_\ratord \approx \mathbf{\Pi}$ given by

### 2022-07-21 numerical_examples/misc/ellipsoid_intersection_figure.py

@@ -1,0 +1,130 @@
+import numpy as np
+from scipy.linalg import eigh
+from scipy.optimize import minimize_scalar
+import matplotlib.pyplot as plt
+from matplotlib.patches import Ellipse
+from matplotlib.ticker import FormatStrFormatter
+
+from localpsf import localpsf_root
+
+save_dir = localpsf_root / 'numerical_examples' / 'misc' / 'data'
+save_dir.mkdir(parents=True, exist_ok=True)
+save_dir_str = str(save_dir)
+
+
+def ellipsoid_intersection_test(Sigma_A, Sigma_B, mu_A, mu_B, tau):
+    lambdas, Phi, v_squared = ellipsoid_intersection_test_helper(Sigma_A, Sigma_B, mu_A, mu_B)
+    res = minimize_scalar(ellipsoid_K_function,
+                          bracket=[0.0, 0.5, 1.0],
+                          args=(lambdas, v_squared, tau))
+    return (res.fun[0] >= 0)
+
+
+def ellipsoid_intersection_test_helper(Sigma_A, Sigma_B, mu_A, mu_B):
+    lambdas, Phi = eigh(Sigma_A, b=Sigma_B)
+    v_squared = np.dot(Phi.T, mu_A - mu_B) ** 2
+    return lambdas, Phi, v_squared
+
+
+def ellipsoid_K_function(ss, lambdas, v_squared, tau):
+    ss = np.array(ss).reshape((-1,1))
+    lambdas = np.array(lambdas).reshape((1,-1))
+    v_squared = np.array(v_squared).reshape((1,-1))
+    return 1.-(1./tau**2)*np.sum(v_squared*((ss*(1.-ss))/(1.+ss*(lambdas-1.))), axis=1)
+
+
+
+tau=1.3
+
+Sigma_A = np.array([[0.7, 0.02 ], [0.02, 1.1]])
+Sigma_B = np.array([[ 0.6, -0.5], [-0.5, 0.7]])
+
+mu_A = np.array([ 0.9, -0.75])
+mu_B0 = np.array([-1.4, 0.6])
+
+
+def plot_ellipse(mu, Sigma, n_std_tau, ax=None, **kwargs):
+    # Usage:
+    #   https://github.com/NickAlger/nalger_helper_functions/tree/master/jupyter_notebooks/plot_ellipse.ipynb
+    if ax is None:
+        ax = plt.gca()
+
[…]

### 2022-07-22 paper/localpsf.tex
modify figures in paper
edit paper a small amount
have included the Stokes model description draft, included one more digit in the condition number table

@@
 
-Some numerical results here.
+We next test the local psf framework on an inverse problem governed by a PDE which models steady-state ice sheet flow.
+The domain that is covered by ice is denoted by
+$\Omega\subset\mathbb{R}^{d}$, $d=3$,
+$\Gamma_{b}$, $\Gamma_{l}$, and $\Gamma_{t}$ denote
+the basal, lateral and top parts of the domain boundary
+$\partial\Omega$.
+The governing equations are the linear incompressible Stokes
+equations,
+
+\begin{subequations}\label{Stokeseqn}
+	\begin{align}
+	-\nabla\cdot\bm{\sigma}_{\bm{u}}
+	&=\rho\bm{g}\,\,\,\text{ in }\Omega, \label{Stokeseqn:1}\\
+	\nabla\cdot\bm{u}
+	&=0\,\,\,\,\,\,\text{ in }\Omega, \label{Stokeseqn:2}\\
+	\bm{\sigma}_{\bm{u}}\bm{n}
+	&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{t}, \label{Stokeseqn:3}\\
+	\bm{u}\cdot\bm{n} =0 \text{ and } \bm{T}\left(\bm{\sigma}_{\bm{u}}\bm{n}+\exp\left(m\right)\bm{u}\right)
+	&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{b},\label{Stokeseqn:4} \\
+	\left(\bm{\sigma}_{\bm{u}}\bm{n}+\mu\bm{u}\right)&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{l},\label{Stokeseqn:5}	
+	\end{align}
+\end{subequations}
+whose solution determines the ice flow velocity $\bm{u}$ and pressure $p$. Here, $m$ is the logarithmic
+basal sliding parameter field, $\rho\bm{g}$ the body force density due to gravity, $\bm{n}$ is the outward normal and $\bm{T}$ is a tangential projection operator. 
+We employ Glen's flow law
+(CITE GLEN) %\cite{glen1955},
+which is a constitutive law for ice that relates the stress
+tensor $\bm{\sigma}_{\bm{u}}$ and the strain rate tensor
+$\bm{\dot{\varepsilon}}_{\bm{u}}= \frac{1}{2}\left(
+\bm{\nabla}\bm{u}+\bm{\nabla}\bm{u}^{\top} \right)$, $\bm{\sigma}_{\bm{u}}= 2\eta
+\bm{\dot{\varepsilon}}_{\bm{u}} -\bm{I}p$, where $\eta=\frac{1}{2}A^{-1}$ is the effective viscosity, $\bm{I}$ is the unit matrix,
+and $A$ is a flow rate factor. Glen's exponent has been chosen $n=1$, which provides for a linear Stokes model.
 
@@
 			\hline 
-			PSF $(1)$  & \verb~1.8e+1~ & \verb~1.8e+1~ \\
-            PSF $(5)$  & \verb~1.4e+1~ & \verb~1.4e+1~ \\
-            PSF $(25)$ & \verb~1.0e+1~ & \verb~1.1e+1~ \\
-			REG        & \verb~7.0e+3~ & \verb~7.0e+3~ \\
-			NONE       & \verb~8.7e+2~ & \verb~9.2e+2~ 
+			PSF $(1)$  & \verb~1.75e+1~ & \verb~1.76e+1~ \\
+            PSF $(5)$  & \verb~1.39e+1~ & \verb~1.39e+1~ \\
+            PSF $(25)$ & \verb~1.03e+1~ & \verb~1.06e+1~ \\
+			REG        & \verb~7.02e+3~ & \verb~7.02e+3~ \\
+			NONE       & \verb~8.71e+2~ & \verb~9.22e+2~ 
 		\end{tabular}

### 2022-07-23 paper/localpsf.tex

@@
 	\begin{center}
-		\includegraphics[scale=0.75]{mean_displacement_invariance.pdf}
+		\includegraphics[scale=0.85]{mean_displacement_invariance.pdf}
 	\end{center}
@@
 	&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{t}, \label{Stokeseqn:3}\\
-	\bm{u}\cdot\bm{n} =0 \text{ and } \bm{T}\left(\bm{\sigma}_{\bm{u}}\bm{n}+\exp\left(m\right)\bm{u}\right)
+	\bm{u}\cdot\bm{n} =0 \text{ and } \bm{T}\left(\bm{\sigma}_{\bm{u}}\bm{n}+\exp\left(\beta\right)\bm{u}\right)
 	&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{b},\label{Stokeseqn:4} \\
-	\left(\bm{\sigma}_{\bm{u}}\bm{n}+\mu\bm{u}\right)&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{l},\label{Stokeseqn:5}	
+	\left(\bm{\sigma}_{\bm{u}}\bm{n}+\omega\bm{u}\right)&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{l},\label{Stokeseqn:5}	
 	\end{align}
 \end{subequations}
-whose solution determines the ice flow velocity $\bm{u}$ and pressure $p$. Here, $m$ is the logarithmic
+whose solution determines the ice flow velocity $\bm{u}$ and pressure $p$. Here, $\beta$ is the logarithmic
 basal sliding parameter field, $\rho\bm{g}$ the body force density due to gravity, $\bm{n}$ is the outward normal and $\bm{T}$ is a tangential projection operator. 
@@
 \bm{\dot{\varepsilon}}_{\bm{u}} -\bm{I}p$, where $\eta=\frac{1}{2}A^{-1}$ is the effective viscosity, $\bm{I}$ is the unit matrix,
-and $A$ is a flow rate factor. Glen's exponent has been chosen $n=1$, which provides for a linear Stokes model.
+and $A$ is a flow rate factor. Glen's exponent has been chosen $n=1$, which provides for a linear Stokes model. The domain geometry $\Omega$ is shown in Figure \ref{fig:iceMountainMesh}.
+
+The inverse problem that we seek to solve is
+
+\begin{align*}
+\min_{\beta}J(\beta)&:=\frac{1}{2}\int_{\Gamma_{t}}||\bm{T}(\bm{u}-\bm{u}_{d})||_{2}^{2}\,dS+\frac{1}{2}\int_{\Omega}\beta\mathcal{L}^{2}\beta\,dx,\\
+\text{where }\bm{u}& \text{ solves \ref{Stokeseqn}}.
+\end{align*}
+Here $\mathcal{L}$ is a regularizing invertible elliptic differential operator, where $s=\mathcal{L}^{-1}u$ if and only if
+
+\begin{align*}
+-\gamma\Delta u +\delta\,u&=s\,\,\,\,\text{ in }\Gamma_{b},\\
+\gamma \bm{\nabla}u\cdot\bm{n}+\mu\,u&=0\,\,\,\text{ on }\partial\Gamma_{b}.
+\end{align*}
+The observation data $\bm{u}_{d}(\bm{x})$ is generated synthetically, as follows, we propose a true log basal friction field $\beta_{\text{true}}(\bm{x})$, solve equation \ref{Stokeseqn} to obtain $\bm{u}_{\text{true}}(\bm{x})$, we then add random relative noise to obtain the noisy data $\bm{u}_{d}$
+
+\begin{align*}
+\left(\bm{u}_{d}(\bm{x})\right)_{i}&=\left(\bm{u}_{\text{true}}(\bm{x})\right)_{i}+\varepsilon\, \xi_{i}(\bm{x})|\left(\bm{u}_{\text{true}}(\bm{x})\right)_{i}|\,\,\,i=1,\dots,d,
+\end{align*}
+where $\xi_{i}(\bm{x})$ are standard normal random numbers and $\varepsilon>0$ sets the noise level.
+The inverse problem is solved with an inexact-Newton CG scheme, wherein each Hessian $\mathcal{A}$ action, requires the linearized PDE solutions, and so we apply the psf framework to generate Hessian preconditioners and compare the performance with other techniques, such as preconditioning by the regularization operator.
 
 \begin{figure}[H]
-	\begin{subfigure}{0.32\textwidth}
-    \includegraphics[scale=0.12]{meshHeight_view0_edges_withColorBar.png}	
-    \end{subfigure}
-    \begin{subfigure}{0.32\textwidth}
-    \includegraphics[scale=0.12]{trueVelocity_glyphs_withColorBar.png}
-    \end{subfigure}
-    \begin{subfigure}{0.32\textwidth}
-    \includegraphics[scale=0.162]{mtrue_withColorBar.png}
+	\begin{subfigure}{0.6\textwidth}
+		\includegraphics[scale=0.25]{meshHeight_view2_edges.png}
+	\end{subfigure}
+	\begin{subfigure}{0.38\textwidth}
+		\includegraphics[scale=0.15]{meshHeight_view0_edges_withColorBar.png}
+	\end{subfigure}
+\caption{TO DO: description of ice mountain}
+\label{fig:iceMountainMesh}
+\end{figure}
+
+\begin{figure}[H]
+    \begin{subfigure}{0.49\textwidth}
+		\includegraphics[scale=0.243]{mtrue_withColorBar.png}
+	\end{subfigure}
+    \begin{subfigure}{0.49\textwidth}
+    	\includegraphics[scale=0.18]{trueVelocity_glyphs_withColorBar.png}
     \end{subfigure}
@@
 		\begingroup
-		\setlength{\tabcolsep}{4pt}
+		\setlength{\tabcolsep}{3pt}
 		\renewcommand{\arraystretch}{1.25}
@@
 			\verb~0~ &
-			\verb~1~ & \verb~4~ & \verb~1.95e+7~ &
-			\verb~1~ & \verb~4~ & \verb~1.95e+7~ &
-			\verb~1~ & \verb~4~ & \verb~1.95e+7~ \\
+			\verb~1~ & \verb~4~ & \verb~2.0e+7~ &
+			\verb~1~ & \verb~4~ & \verb~2.0e+7~ &
+			\verb~1~ & \verb~4~ & \verb~2.0e+7~ \\
 			\verb~1~ &
-			\verb~2~ & \verb~6~ & \verb~5.76e+6~ &
-			\verb~2~ & \verb~6~ & \verb~7.90e+6~ &
-			\verb~2~ & \verb~6~ & \verb~5.76e+6~ \\
+			\verb~2~ & \verb~6~ & \verb~5.8e+6~ &
+			\verb~2~ & \verb~6~ & \verb~7.9e+6~ &
+			\verb~2~ & \verb~6~ & \verb~5.8e+6~ \\
 			\verb~2~ &
-			\verb~4~ & \verb~10~ & \verb~2.36e+6~ &
-			\verb~5~ & \verb~12~ & \verb~3.87e+6~ &
-			\verb~4~ & \verb~10~ & \verb~2.36e+6~ \\
+			\verb~4~ & \verb~10~ & \verb~2.4e+6~ &
+			\verb~5~ & \verb~12~ & \verb~3.9e+6~ &
+			\verb~4~ & \verb~10~ & \verb~2.4e+6~ \\
 			\verb~3~ &
-			\verb~2~ & \verb~6+22~ & \verb~5.79e+5~ &
-			\verb~13~ & \verb~28~ & \verb~1.62e+6~ &
-			\verb~13~ & \verb~28~ & \verb~5.79e+5~ \\
+			\verb~2~ & \verb~6+22~ & \verb~5.8e+5~ &
+			\verb~13~ & \verb~28~ & \verb~1.6e+6~ &
+			\verb~13~ & \verb~28~ & \verb~5.8e+5~ \\
 			\verb~4~ &
-			\verb~5~ & \verb~12~ & \verb~5.61e+4~ &
-			\verb~42~ & \verb~86~ & \verb~4.78e+5~ &
-			\verb~33~ & \verb~68~ & \verb~9.05e+4~ \\
+			\verb~5~ & \verb~12~ & \verb~5.6e+4~ &
+			\verb~42~ & \verb~86~ & \verb~4.8e+5~ &
+			\verb~33~ & \verb~68~ & \verb~9.1e+4~ \\
 			\verb~5~ &
-			\verb~10~ & \verb~22~ & \verb~3.45e+3~ &
-			\verb~84~ & \verb~170~ & \verb~7.72e+4~ &
-			\verb~57~ & \verb~116~ & \verb~6.21e+3~ \\
+			\verb~10~ & \verb~22~ & \verb~3.5e+3~ &
+			\verb~84~ & \verb~170~ & \verb~7.7e+4~ &
+			\verb~57~ & \verb~116~ & \verb~6.2e+3~ \\
 			\verb~6~ &
-			\verb~14~ & \verb~30~ & \verb~2.66e+1~ &
-			\verb~125~ & \verb~252~ & \verb~5.04e+3~ &
-			\verb~76~ & \verb~154~ & \verb~1.05e+2~ \\
+			\verb~14~ & \verb~30~ & \verb~2.7e+1~ &
+			\verb~125~ & \verb~252~ & \verb~5.0e+3~ &
+			\verb~76~ & \verb~154~ & \verb~1.1e+2~ \\
 			\verb~7~ &
-			\verb~0~ & \verb~2~ & \verb~4.09e-2~ &
-			\verb~194~ & \verb~390~ & \verb~7.91e+1~ &
-			\verb~117~ & \verb~236~ & \verb~2.29e-1~ \\
+			\verb~0~ & \verb~2~ & \verb~4.1e-2~ &
+			\verb~194~ & \verb~390~ & \verb~7.9e+1~ &
+			\verb~117~ & \verb~236~ & \verb~2.3e-1~ \\
 			\verb~8~ & 
 			\verb~---~ & \verb~---~ & \verb~---~ &
-			\verb~0~ & \verb~2~ & \verb~1.60e-1~ &
-			\verb~0~ & \verb~2~ & \verb~4.08e-2~ \\
+			\verb~0~ & \verb~2~ & \verb~1.6e-1~ &
+			\verb~0~ & \verb~2~ & \verb~4.1e-2~ \\
 			\hline
@@
 \begin{figure}[H]
-	\begin{subfigure}{0.225\textwidth}
-		\includegraphics[scale=0.08]{m1_psf.png}	
-	\end{subfigure}
-	\begin{subfigure}{0.225\textwidth}
-		\includegraphics[scale=0.08]{m3_psf.png}
-	\end{subfigure}
-	\begin{subfigure}{0.225\textwidth}
-		\includegraphics[scale=0.08]{m4_psf.png}
-	\end{subfigure}
-	\begin{subfigure}{0.225\textwidth}
-	\includegraphics[scale=0.08]{m6_psf.png}
-\end{subfigure}
-\begin{subfigure}{0.05\textwidth}
-	\includegraphics[scale=0.1]{mColorBar.png}
-\end{subfigure}
+	\begin{subfigure}{0.24\textwidth}
+		\includegraphics[scale=0.1]{m1_psf.png}	
+	\end{subfigure}
+	\begin{subfigure}{0.24\textwidth}
+		\includegraphics[scale=0.1]{m3_psf.png}
+	\end{subfigure}
+	\begin{subfigure}{0.24\textwidth}
+		\includegraphics[scale=0.1]{m4_psf.png}
+	\end{subfigure}
+	\begin{subfigure}{0.24\textwidth}
+	\includegraphics[scale=0.1]{m6_psf.png}
+\end{subfigure}
+%\begin{subfigure}{0.05\textwidth}
+%	\includegraphics[scale=0.1]{mColorBar.png}
+%\end{subfigure}
 \caption{The parameter after $1$ (left), $3$ (second from left), $4$ (third from left) Newton iterates and the final Newton iterate (right).} 

### 2022-07-23 paper/localpsf.tex
paper edits in numerical results section. resolved merge conflict

@@
 
-We next test the local psf framework on an inverse problem governed by a PDE which models steady-state ice sheet flow.
-The domain that is covered by ice is denoted by
-$\Omega\subset\mathbb{R}^{d}$, $d=3$,
-$\Gamma_{b}$, $\Gamma_{l}$, and $\Gamma_{t}$ denote
-the basal, lateral and top parts of the domain boundary
-$\partial\Omega$.
+We approximate the Gauss-Newton Hessian in an inverse problem governed by a PDE which models steady-state ice sheet flow.
+Briefly, a sheet of ice is sliding down a mountain (see Figure \ref{fig:ice_mountain_mesh}).
+Given observations of the tangential component of the ice velocity on the top surface of the ice, we invert for the logarithm of the unknown spatially varying basal friction parameter which governs how easily the ice slides along the ground underneath it.
+
+\begin{figure}[H]
+	\begin{subfigure}{0.6\textwidth}
+		\includegraphics[scale=0.25]{meshHeight_view2_edges.png}
+		\caption{Bird's eye view}
+	\end{subfigure}
+	\begin{subfigure}{0.38\textwidth}
+		\includegraphics[scale=0.15]{meshHeight_view0_edges_withColorBar.png}
+		\caption{Top-down view}
+	\end{subfigure}
+	\caption{The ice sheet, discretized by a mesh of tetrahedra. Color indicates the height of the base of the ice sheet (i.e., the mountain topography).}
+	\label{fig:ice_mountain_mesh}
+\end{figure}
+
+The domain that is filled with ice is denoted by
+$\icedomain \subset\mathbb{R}^{3}$. The basal, lateral and top parts of $\partial \icedomain$ are denoted by $\Gamma_{b}$, $\Gamma_{l}$, and $\Gamma_{t}$, respectively.
 The governing equations are the linear incompressible Stokes
@@
 	-\nabla\cdot\bm{\sigma}_{\bm{u}}
-	&=\rho\bm{g}\,\,\,\text{ in }\Omega, \label{Stokeseqn:1}\\
+	&=\rho\bm{g}\,\,\,\text{ in }\icedomain, \label{Stokeseqn:1}\\
 	\nabla\cdot\bm{u}
-	&=0\,\,\,\,\,\,\text{ in }\Omega, \label{Stokeseqn:2}\\
+	&=0\,\,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:2}\\
 	\bm{\sigma}_{\bm{u}}\bm{n}
 	&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{t}, \label{Stokeseqn:3}\\
-	\bm{u}\cdot\bm{n} =0 \text{ and } \bm{T}\left(\bm{\sigma}_{\bm{u}}\bm{n}+\exp\left(\beta\right)\bm{u}\right)
+	\bm{u}\cdot\bm{n} =0 \quad\text{and}\quad \bm{T}\left(\bm{\sigma}_{\bm{u}}\bm{n}+\exp\left(\basalfriction\right)\bm{u}\right)
 	&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{b},\label{Stokeseqn:4} \\
-	\left(\bm{\sigma}_{\bm{u}}\bm{n}+\omega\bm{u}\right)&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{l},\label{Stokeseqn:5}	
+	\left(\bm{\sigma}_{\bm{u}}\bm{n}+\mu\bm{u}\right)&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{l}.\label{Stokeseqn:5}	
 	\end{align}
 \end{subequations}
-whose solution determines the ice flow velocity $\bm{u}$ and pressure $p$. Here, $\beta$ is the logarithmic
-basal sliding parameter field, $\rho\bm{g}$ the body force density due to gravity, $\bm{n}$ is the outward normal and $\bm{T}$ is a tangential projection operator. 
+The solution to these equations is the tuple $(\bm{u}, p)$, where $\bm{u}$ is the ice flow velocity, and $p$ is the pressure. Here, $\basalfriction$ is the unknown logarithmic
+basal sliding parameter field which is defined on the $2D$ surface $\Gamma_b$.
+The quantity $\rho\bm{g}$ is the body force density due to gravity, $\bm{n}$ is the outward normal and $\bm{T}$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary.
 We employ Glen's flow law
@@
 \bm{\dot{\varepsilon}}_{\bm{u}} -\bm{I}p$, where $\eta=\frac{1}{2}A^{-1}$ is the effective viscosity, $\bm{I}$ is the unit matrix,
-and $A$ is a flow rate factor. Glen's exponent has been chosen $n=1$, which provides for a linear Stokes model. The domain geometry $\Omega$ is shown in Figure \ref{fig:iceMountainMesh}.
+and $A$ is a flow rate factor.
+Glen's exponent has been chosen $n=1$, which provides for a linear Stokes model.
+Note that while the PDE is linear, the parameter-to-solution map, $\basalfriction \mapsto (\bm{u}, p)$, is nonlinear.
 
@@
 \begin{align*}
-\min_{\beta}J(\beta)&:=\frac{1}{2}\int_{\Gamma_{t}}||\bm{T}(\bm{u}-\bm{u}_{d})||_{2}^{2}\,dS+\frac{1}{2}\int_{\Omega}\beta\mathcal{L}^{2}\beta\,dx,\\
+\min_{\basalfriction}J(\basalfriction)&:=\frac{1}{2}\int_{\Gamma_{t}}||\bm{T}(\bm{u}-\bm{u}_{d})||_{2}^{2}\,dS+\frac{1}{2}\int_{\Omega}\basalfriction\mathcal{L}^{2}\basalfriction\,dx,\\
 \text{where }\bm{u}& \text{ solves \ref{Stokeseqn}}.
@@
 \end{align*}
-The observation data $\bm{u}_{d}(\bm{x})$ is generated synthetically, as follows, we propose a true log basal friction field $\beta_{\text{true}}(\bm{x})$, solve equation \ref{Stokeseqn} to obtain $\bm{u}_{\text{true}}(\bm{x})$, we then add random relative noise to obtain the noisy data $\bm{u}_{d}$
+The observation data $\bm{u}_{d}(\bm{x})$ is generated synthetically, as follows: we propose a true log basal friction field $\basalfriction_{\text{true}}(\bm{x})$, solve equation \ref{Stokeseqn} to obtain $\bm{u}_{\text{true}}(\bm{x})$, we then add random relative noise to obtain the noisy data $\bm{u}_{d}$
 
@@
 \begin{figure}[H]
-	\begin{subfigure}{0.6\textwidth}
-		\includegraphics[scale=0.25]{meshHeight_view2_edges.png}
-	\end{subfigure}
-	\begin{subfigure}{0.38\textwidth}
-		\includegraphics[scale=0.15]{meshHeight_view0_edges_withColorBar.png}
-	\end{subfigure}
-\caption{TO DO: description of ice mountain}
-\label{fig:iceMountainMesh}
-\end{figure}
-
-\begin{figure}[H]
 	\begin{subfigure}{0.49\textwidth}
 		\includegraphics[scale=0.243]{mtrue_withColorBar.png}
+		\caption{True $m$.}
 	\end{subfigure}
@@
 		\includegraphics[scale=0.18]{trueVelocity_glyphs_withColorBar.png}
-    \end{subfigure}
+		\caption{True $\bm{u}$.}
+	\end{subfigure}
+	\caption{True logarithmic basal sliding parameter $\basalfriction$ (left) and true velocity $\bm{u}$ (right).
+In the right plot, arrows indicate the direction of $\bm{u}$ and color indicates the magnitude of $\bm{u}$.}
+\end{figure}
+
+\begin{figure}[H]
+	\begin{subfigure}{0.24\textwidth}
+		\includegraphics[scale=0.1]{m1_psf.png}
+		%		\caption{}
+		%		\label{fig:newtoniter1}
+	\end{subfigure}
+	\begin{subfigure}{0.24\textwidth}
+		\includegraphics[scale=0.1]{m3_psf.png}
+		%		\caption{}
+		%		\label{fig:newtoniter2}
+	\end{subfigure}
+	\begin{subfigure}{0.24\textwidth}
+		\includegraphics[scale=0.1]{m4_psf.png}
+		%		\caption{}
+		%		\label{fig:newtoniter3}
+	\end{subfigure}
+	\begin{subfigure}{0.24\textwidth}
+		\includegraphics[scale=0.1]{m6_psf.png}
+		%		\caption{}
+		%		\label{fig:newtoniter4}
+	\end{subfigure}
+	%\begin{subfigure}{0.05\textwidth}
+	%	\includegraphics[scale=0.1]{mColorBar.png}
+	%\end{subfigure}
+	\caption{The parameter after $1$ (left), $3$ (second from left), $4$ (third from left) Newton iterates and the final Newton iterate (right).
+The preconditioner is constructed between Newton iterations 3 and 4.}
 \end{figure}
@@
 		\begin{tabular}{c| c c c | c c c | c c c}
-			&  \multicolumn{3}{|c|}{PSF} & \multicolumn{3}{|c|}{REG} & \multicolumn{3}{|c}{NONE} \\
+			&  \multicolumn{3}{|c|}{PSF (5)} & \multicolumn{3}{|c|}{REG} & \multicolumn{3}{|c}{NONE} \\
 			\hline
@@
 	\end{center}
+	\caption{Convergence history for solving the Stokes inverse problem using the preconditioned inexact Newton-Krylov method.
+The Newton iteration terminates when the norm of the gradient has decreased from its initial value by a factor of $10^{-8}$.
+The Gauss-Newton Hessian is used in iterations $0$--$5$, and the Hessian is used in all subsequent iterations.
+Preconditioners shown are our method with five impulse response batches (PSF (5)), regularization preconditioning (REG), and no preconditioning (NONE).
+Our preconditioner (PSF (5)) is built at iteration $3$ using the Gauss-Newton Hessian, and re-used for all subsequent iterations.
+The iterates are $\basalfriction_0$ (initial guess), $\basalfriction_1$ (parameter after one Newton step), $\basalfriction_2$ (parameter after two Newton steps), and so on.
+The columns titled \#CG show the number of conjugate gradient iterations used to solve the Newton system at $\basalfriction_k$, for the purpose of computing a search direction towards $\basalfriction_{k+1}$.
+The columns titled $||g||$ show the $l^2$ norm of the gradient at $\basalfriction_k$.
+The columns titled \#Stokes show the number of Stokes PDE solves performed in each Newton iteration.
+This consists of Stokes solves for $u$ performed during the linesearch going from $\basalfriction_{k-1}$ to $\basalfriction_k$ (including one Stokes solve at $\basalfriction_k$), plus one adjoint Stokes solve to compute the gradient at $\basalfriction_k$, plus one incremental forward and one incremental adjoint Stokes solve per CG iteration for solving the Newton system at iteration $k$.
+In the PSF (5) portion of row 3, we write $6+22$ to indicate that 6 Stokes solves were used during the standard course of the iteration, and 22 additional Stokes solves were used to build our preconditioner.
+}
 \end{table}
-
-\begin{figure}[H]
-	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.1]{m1_psf.png}	
-	\end{subfigure}
-	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.1]{m3_psf.png}
-	\end{subfigure}
-	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.1]{m4_psf.png}
-	\end{subfigure}
-	\begin{subfigure}{0.24\textwidth}
-	\includegraphics[scale=0.1]{m6_psf.png}
-\end{subfigure}
-%\begin{subfigure}{0.05\textwidth}
-%	\includegraphics[scale=0.1]{mColorBar.png}
-%\end{subfigure}
-\caption{The parameter after $1$ (left), $3$ (second from left), $4$ (third from left) Newton iterates and the final Newton iterate (right).} 
-\end{figure}
 
@@
 		\end{tikzpicture}
+		\caption{}
+		\label{fig:cg_convergence}
 	\end{subfigure}
@@
 		\end{tikzpicture}
-	\end{subfigure}
+		\caption{}
+		\label{fig:gamma_sweep}
+	\end{subfigure}
+	\caption{Left (\ref{fig:cg_convergence}) shows the convergence history for solving a linear system of the form $Hx=b$ using the preconditioned conjugate gradient method, where $b$ has random independent and identically distributed (i.i.d) entries drawn from the standard normal distribution.
+The y-axis is the relative $l^2$ norm of the residual, i.e., $||b-Hx_k||/||b||$, where $x_k$ is the $k$th iterate.
+Here the regularization parameter $\gamma$ is determined by the Morozov discrepancy principle.
+Right (\ref{fig:gamma_sweep}) shows the number of preconditioned conjugate gradient iterations required to reduce the relative $l^2$ norm of the residual below a tolerance of $10^{-6}$, for a range of different regularization parameters $\gamma$.
+Results in these figures are shown for our preconditioner with 1, 5, and 25 batches of impulse responses (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
+The preconditioner is constructed using the Gauss-Newton Hessian.
+For a fair comparison, the Hessian and Gauss-Newton Hessian are always evaluated at the same parameter $\basalfriction$, which is the optimal point for the inverse problem with $\gamma$ given by the Morozov discrepancy principle.}
 \end{figure}
@@
 		\end{center}
+		\caption{}
+		\label{fig:generalized_eigenvalues}
 	\end{subfigure}
@@
 		\end{center}
-\end{subfigure}
+		\caption{}
+		\label{fig:condition_number}
+	\end{subfigure}
+	\caption{Left (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $H u_k = \lambda_k P_k$.
+Here $H$ is the Hessian and $P$ is the PSF approximation constructed using the Gauss-Newton Hessian, $H_\text{gn}$, for 1, 5, and 25 batches, or the regularization operator.
+Right (\ref{fig:condition_number}): condition number, $\kappa$, for the preconditioned Hessian and Gauss-Newton Hessian for different preconditioners.
+All operators are evaluated at the parameter $\basalfriction$ that solves the inverse problem for $\gamma$ given by the Morozov discrepancy principle.}
 \end{figure}

### 2022-07-23 paper/localpsf.tex
including the meshHeight plot which shows the axes and ultimately the scale of the domain U

@@
 	\end{subfigure}
-	\begin{subfigure}{0.38\textwidth}
-		\includegraphics[scale=0.15]{meshHeight_view0_edges_withColorBar.png}
+	\begin{subfigure}{0.48\textwidth}
+		\includegraphics[scale=0.15]{meshHeight_view0_edges_withAxes.png}
 		\caption{Top-down view}
@@
 		\includegraphics[scale=0.243]{mtrue_withColorBar.png}
-		\caption{True $m$.}
+		\caption{True $\basalfriction$.}
 	\end{subfigure}
@@
 	%\end{subfigure}
-	\caption{The parameter after $1$ (left), $3$ (second from left), $4$ (third from left) Newton iterates and the final Newton iterate (right).
+	\caption{The log basal friction parameter $\basalfriction$ after $1$ (left), $3$ (second from left), $4$ (third from left) Newton iterates and the final Newton iterate (right).
 The preconditioner is constructed between Newton iterations 3 and 4.}

### 2022-07-23 paper/localpsf.tex
small changes to paper, adding a true velocity plot which has a colorbar the same size as the true parameter and also including an image which is notCropped in case we would like to edit it in the future
adding Glen reference to paper and bib

@@
 The quantity $\rho\bm{g}$ is the body force density due to gravity, $\bm{n}$ is the outward normal and $\bm{T}$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary.
-We employ Glen's flow law
-(CITE GLEN) %\cite{glen1955},
+We employ Glen's flow law \cite{Glen1955},
 which is a constitutive law for ice that relates the stress
@@
 	\begin{subfigure}{0.49\textwidth}
-		\includegraphics[scale=0.18]{trueVelocity_glyphs_withColorBar.png}
+		\includegraphics[scale=0.243]{trueVelocity2_glyphs.png}
 		\caption{True $\bm{u}$.}

### 2022-07-24 paper/localpsf.tex
lost of paper edits in numerical results section.

@@
 	\end{subfigure}
-	\caption{Batches of normalized impulse responses for the Stokes inverse problem Hessian. 
+	\caption{Batches of normalized impulse responses for the Stokes inverse problem data misfit Gauss-Newton Hessian, described in Section \ref{sec:numerical_results}. 
 %	Brightness scale differs between the subfigures. 
@@
 
-We approximate the Gauss-Newton Hessian in an inverse problem governed by a PDE which models steady-state ice sheet flow.
+We use our method to approximate the data misfit Gauss-Newton Hessian in an inverse problem governed by a PDE which models steady-state ice sheet flow.
 Briefly, a sheet of ice is sliding down a mountain (see Figure \ref{fig:ice_mountain_mesh}).
@@
 \begin{figure}[H]
-	\begin{subfigure}{0.6\textwidth}
-		\includegraphics[scale=0.25]{meshHeight_view2_edges.png}
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.23]{meshHeight_view2_edges.png}
 		\caption{Bird's eye view}
@@
 	\begin{subfigure}{0.48\textwidth}
-		\includegraphics[scale=0.15]{meshHeight_view0_edges_withAxes.png}
+		\includegraphics[scale=0.22]{meshHeight_view0_edges_withAxes.png}
 		\caption{Top-down view}
@@
 \end{subequations}
-The solution to these equations is the tuple $(\bm{u}, p)$, where $\bm{u}$ is the ice flow velocity, and $p$ is the pressure. Here, $\basalfriction$ is the unknown logarithmic
+The solution to these equations is the tuple $(\bm{u}, p)$, where $\bm{u}$ is the ice flow velocity field, and $p$ is the pressure field. Here, $\basalfriction$ is the unknown logarithmic
 basal sliding parameter field which is defined on the $2D$ surface $\Gamma_b$.
@@
 
-The inverse problem that we seek to solve is
-
-\begin{align*}
-\min_{\basalfriction}J(\basalfriction)&:=\frac{1}{2}\int_{\Gamma_{t}}||\bm{T}(\bm{u}-\bm{u}_{d})||_{2}^{2}\,dS+\frac{1}{2}\int_{\Omega}\basalfriction\mathcal{L}^{2}\basalfriction\,dx,\\
-\text{where }\bm{u}& \text{ solves \ref{Stokeseqn}}.
-\end{align*}
-Here $\mathcal{L}$ is a regularizing invertible elliptic differential operator, where $s=\mathcal{L}^{-1}u$ if and only if
-
+The pressure, $p$, is discretized with first order continuous Galerkin finite elements defined on a mesh of tetrahedra discretizing $U$.
+The velocity, $\bm{u}$, is discretized with second order continuous Galerkin finite elements on the same mesh, as required for discrete stability.
+The parameter $\basalfriction$ is discretized with first order scalar continuous Galerkin finite elements on the mesh of triangles that results from restricting the tetrahedral mesh to the basal boundary, $\Gamma_b$.
+
+\begin{figure}[H]
+	\begin{subfigure}{0.49\textwidth}
+		\includegraphics[scale=0.243]{mtrue_withColorBar.png}
+		\caption{$\basalfriction_\text{true}$.}
+	\end{subfigure}
+	\begin{subfigure}{0.49\textwidth}
+		\includegraphics[scale=0.243]{trueVelocity2_glyphs.png}
+		\caption{$\bm{u}_\text{true}$.}
+	\end{subfigure}
+	\caption{True logarithmic basal sliding parameter, $\basalfriction_\text{true}$, (left) and true velocity $\bm{u}_\text{true}$ (right).
+In the right plot, arrows indicate the direction of $\bm{u}_\text{true}$ and color indicates the magnitude.}
+	\label{fig:true_beta_u}
+\end{figure}
+
+The true parameter field, $\basalfriction_\text{true}$, and corresponding velocity field, $\bm{u}_\text{true}$, are shown in Figure \ref{fig:true_beta_u}.
+We generate synthetic observations by restricting the tangential component of $\bm{u}_\text{true}$ to the top surface to get
+\begin{equation*}
+\bm{y}_\text{true} = \bm{T} \bm{u}_\text{true} |_{\Gamma_t}.
+\end{equation*}
+We add $1\%$ multiplicative gaussian noise, $\boldsymbol{\xi}$, to all components of the finite element vector representing $\bm{y}_\text{true}$ to get noisy synthetic observations, $\bm{y}_\text{obs}$, which are used for the inversion.
+That is,
+\begin{equation*}
+\left[\bm{y}_\text{obs}\right]_j := \left[\bm{y}_\text{true}\right]_j + \epsilon \left[\boldsymbol{\xi}\right]_j \left|\left[\bm{y}_\text{true}\right]_j\right|
+\end{equation*}
+where $\epsilon=0.01$, and the components of the noise function, $\boldsymbol{\xi}_j \sim N(0,1)$, are independent and identically distributed (i.i.d) random numbers drawn from the standard normal distribution.
+Here $\left[\bm{f}\right]_j$ denotes the $j$th entries of the finite element vector for the discretized versions of the function $\bm{f}:\Gamma_t \rightarrow \mathbb{R}^2$, and $j$ ranges over all entries of the vector.
+Here we do not use bold to distinguish finite element vectors from their associated functions (as we did previously) because the functions $\bm{y}_\text{true}$, $\bm{y}_\text{obs}$, and $\bm{\xi}$ are already written in bold to denote the fact that they are vector valued.
+
+The inverse problem that we seek to solve takes the form
+\begin{equation}
+	\min_{\basalfriction} \quad J(\basalfriction) = J_d(\basalfriction) + J_r(\basalfriction).
+\end{equation}
+The first term,
+\begin{equation*}
+	J_d(\basalfriction) := \frac{1}{2}\int_{\Gamma_{t}}||\bm{d} - \bm{T}\bm{u}||_{2}^{2}\,dS,
+\end{equation*}
+is the data misfit, which measures the mismatch between the observed data and the predicted data based on a candidate parameter $\basalfriction$.
+Here $\bm{u} = \bm{u}(\basalfriction)$ denotes the velocity field solving $\ref{Stokeseqn}$ for the given parameter $\basalfriction$.
+The second term,
+\begin{equation*}
+	J_r(\basalfriction) := \frac{1}{2}\int_{\Omega}\basalfriction\mathcal{K}^{2}\basalfriction\,dx
+\end{equation*}
+is a Bilaplacian regularization term. Specifically, $\mathcal{K}$ is the inverse of the solution operator for the following elliptic PDE:
 \begin{align*}
@@
 \end{align*}
-The observation data $\bm{u}_{d}(\bm{x})$ is generated synthetically, as follows: we propose a true log basal friction field $\basalfriction_{\text{true}}(\bm{x})$, solve equation \ref{Stokeseqn} to obtain $\bm{u}_{\text{true}}(\bm{x})$, we then add random relative noise to obtain the noisy data $\bm{u}_{d}$
-
-\begin{align*}
-\left(\bm{u}_{d}(\bm{x})\right)_{i}&=\left(\bm{u}_{\text{true}}(\bm{x})\right)_{i}+\varepsilon\, \xi_{i}(\bm{x})|\left(\bm{u}_{\text{true}}(\bm{x})\right)_{i}|\,\,\,i=1,\dots,d,
-\end{align*}
-where $\xi_{i}(\bm{x})$ are standard normal random numbers and $\varepsilon>0$ sets the noise level.
-The inverse problem is solved with an inexact-Newton CG scheme, wherein each Hessian $\mathcal{A}$ action, requires the linearized PDE solutions, and so we apply the psf framework to generate Hessian preconditioners and compare the performance with other techniques, such as preconditioning by the regularization operator.
-
-\begin{figure}[H]
-	\begin{subfigure}{0.49\textwidth}
-		\includegraphics[scale=0.243]{mtrue_withColorBar.png}
-		\caption{True $\basalfriction$.}
-	\end{subfigure}
-	\begin{subfigure}{0.49\textwidth}
-		\includegraphics[scale=0.243]{trueVelocity2_glyphs.png}
-		\caption{True $\bm{u}$.}
-	\end{subfigure}
-	\caption{True logarithmic basal sliding parameter $\basalfriction$ (left) and true velocity $\bm{u}$ (right).
-In the right plot, arrows indicate the direction of $\bm{u}$ and color indicates the magnitude of $\bm{u}$.}
-\end{figure}
+The constant $\delta$ is chosen so that the correlation length of functions drawn from $N(0,\mathcal{L}^{-2})$ is $1/10$th of the radius of the domain.
+In all results shown except one, $\gamma$ is chosen so that the solution satisfies the Morozov discrepancy principle, that is,
+\begin{equation*}
+	\frac{1}{2}\int_{\Gamma_{t}}||\bm{d} - \bm{T}\bm{u}||_{2}^{2}\,dS = \frac{1}{2}\int_{\Gamma_{t}}||\boldsymbol{\xi}||_{2}^{2}\,dS.
+\end{equation*}
+The exception is Figure \ref{fig:gamma_sweep}, where we vary $\gamma$ to study the effect of changing $\gamma$ on the quality of our preconditioner.
+
+The inverse problem is solved with an inexact-Newton conjugate gradient (CG) scheme.
+The initial guess for the parameter, $\basalfriction_0$, is the constant function with value $\basalfriction_0(x)=10.5$, which is halfway between the maximum and minimum values for $\basalfriction_\text{true}$.
+The parameter value is updated iteratively according to the formula
+\begin{equation*}
+	\basalfriction_{k+1} = \basalfriction_k + \alpha_k p_k,
+\end{equation*}
+where the search direction $p_k$ is an approximate solution to the Newton linear system
+\begin{equation}
+	\label{eq:newton_system}
+	H p_k = - g_k,
+\end{equation}
+and $\alpha_k \in \mathbb{R}$ is a step length parameter chosen to satisfy the Armijo conditions \cite{ARMIJO}.
+Here $g_k$ is the gradient of $J$ with respect to $\basalfriction$ at $\basalfriction_k$, and $H$ is either the Hessian or Gauss-Newton Hessian of $J$ with respect to $\basalfriction$ at $\basalfriction_k$.
+The Gauss-Newton Hessian is used for the first five iterations, and the Hessian is used for all subsequent iterations.
+The method terminates when the norm of the gradient drops by a factor of $10^8$.
+The Newton system, \eqref{eq:newton_system}, is solved inexactly to the tolerance $\min \left(0.5, \sqrt{||g_k||}\right)||g_k||$, using the preconditioned conjugate gradient method, which only requires the application of the (Gauss-Newton) Hessian to vectors.
+Each application of the (Gauss-Newton) Hessian to a vector requires solving one forward and one adjoint Stokes PDE of the form \eqref{Stokeseqn} , but with different right hand sides.
+
+We use the framework described in this paper to generate Hessian preconditioners.
+We use our methodology to build H-matrix approximations of the data-misfit component of the Gauss-Newton Hessian.
+That is, the term in the Gauss-Newton Hessian that arises from the data misfit term $J_d$.
+We approximate the Gauss-Newton Hessian rather than the Hessian because the Hessian often has negative values in its integral kernel when one is away from the optimal point.
+At the discrete level, straightforward calculation shows that the regularization term in both the Hessian and the Gauss-Newton Hessian is given by
+\begin{equation*}
+	\bm{K} \bm{M}^{-1} \bm{K},
+\end{equation*}
+where $\bm{K}$ is the finite element discretization of the elliptic operator $\mathcal{K}$, and $\bm{M}$ is the finite element mass matrix.
+These matrices are sparse, so we form H-matrix approximations of these matrices using standard techniques for approximating sparse matrices with H-matrices, then invert $\bm{M}^{-1}$, and multiply these matrices together using H-matrix methods.
+We then add this H-matrix to the H-matrix for the data misfit term constructed using our method, and factor the resulting Hessian approximation using fast H-matrix factorization methods \cite{HLIBPRO}.
+We construct preconditioners using our method with 1, 5, and 25 impulse responses batches.
+The resulting preconditioners are denoted by PSF (1), PSF (5), and PSF (25), respectively.
+We compare the performance of these preconditioners with the most commonly used existing preconditioners, which are no preconditioning (denoted by NONE), and preconditioning by the regularization operator (denoted REG).
 
@@
 	%\end{subfigure}
-	\caption{The log basal friction parameter $\basalfriction$ after $1$ (left), $3$ (second from left), $4$ (third from left) Newton iterates and the final Newton iterate (right).
+	\caption{The log basal friction parameter $\basalfriction$ after $1$ (left), $3$ (second from left), and $4$ (third from left) Newton iterates and the final Newton iterate (right).
 The preconditioner is constructed between Newton iterations 3 and 4.}

### 2022-07-24 paper/localpsf.tex
more writing in numerical results section

@@
 	\end{subfigure}
-	\caption{The ice sheet, discretized by a mesh of tetrahedra. Color indicates the height of the base of the ice sheet (i.e., the mountain topography).}
+	\caption{The ice sheet, discretized by a mesh of tetrahedra.
+Color indicates the height of the base of the ice sheet (i.e., the mountain topography).
+The radius of the domain is $10$ kilometers, and the maximum height of the mountain is $2.1$ kilometers.}
 	\label{fig:ice_mountain_mesh}
@@
 where $\epsilon=0.01$, and the components of the noise function, $\boldsymbol{\xi}_j \sim N(0,1)$, are independent and identically distributed (i.i.d) random numbers drawn from the standard normal distribution.
-Here $\left[\bm{f}\right]_j$ denotes the $j$th entries of the finite element vector for the discretized versions of the function $\bm{f}:\Gamma_t \rightarrow \mathbb{R}^2$, and $j$ ranges over all entries of the vector.
+Here $\left[\bm{f}\right]_j$ denotes the $j$th entries of the finite element vector for the discretized versions of the function $\bm{f}$, and $j$ ranges over all entries of the finite element vector.
 Here we do not use bold to distinguish finite element vectors from their associated functions (as we did previously) because the functions $\bm{y}_\text{true}$, $\bm{y}_\text{obs}$, and $\bm{\xi}$ are already written in bold to denote the fact that they are vector valued.
 
-The inverse problem that we seek to solve takes the form
-\begin{equation}
+The reconstructed parameter field $\basalfriction$ is found as the solution to an optimization problem of the following form:
+\begin{equation}
+	\label{eq:deterministic_optimization_problem}
 	\min_{\basalfriction} \quad J(\basalfriction) = J_d(\basalfriction) + J_r(\basalfriction).
 \end{equation}
-The first term,
-\begin{equation*}
-	J_d(\basalfriction) := \frac{1}{2}\int_{\Gamma_{t}}||\bm{d} - \bm{T}\bm{u}||_{2}^{2}\,dS,
-\end{equation*}
-is the data misfit, which measures the mismatch between the observed data and the predicted data based on a candidate parameter $\basalfriction$.
+The first term in the objective function is
+\begin{equation*}
+	J_d(\basalfriction) := \frac{1}{2}\int_{\Gamma_{t}}||\bm{d} - \bm{T}\bm{u}||_{2}^{2}\,dS.
+\end{equation*}
+We call this term the data misfit, because it measures the mismatch between the observed data and the predicted data based on a candidate parameter $\basalfriction$.
 Here $\bm{u} = \bm{u}(\basalfriction)$ denotes the velocity field solving $\ref{Stokeseqn}$ for the given parameter $\basalfriction$.
@@
 \begin{equation*}
-	J_r(\basalfriction) := \frac{1}{2}\int_{\Omega}\basalfriction\mathcal{K}^{2}\basalfriction\,dx
-\end{equation*}
-is a Bilaplacian regularization term. Specifically, $\mathcal{K}$ is the inverse of the solution operator for the following elliptic PDE:
+	J_r(\basalfriction) := \frac{1}{2}\int_{\Omega}\left| \mathcal{L}(\basalfriction - \basalfriction_0)\right|^2\,dx,
+\end{equation*}
+is a Bilaplacian regularization term. Specifically, $\mathcal{L}$ is the inverse of the solution operator for the following elliptic PDE:
 \begin{align*}
@@
 \end{align*}
-The constant $\delta$ is chosen so that the correlation length of functions drawn from $N(0,\mathcal{L}^{-2})$ is $1/10$th of the radius of the domain.
 In all results shown except one, $\gamma$ is chosen so that the solution satisfies the Morozov discrepancy principle, that is,
@@
 \end{equation*}
+The value of $\gamma$ satisfying the Morozov discrepancy principle is $\gamma=7.3 \times 10^3$.
 The exception is Figure \ref{fig:gamma_sweep}, where we vary $\gamma$ to study the effect of changing $\gamma$ on the quality of our preconditioner.
-
-The inverse problem is solved with an inexact-Newton conjugate gradient (CG) scheme.
-The initial guess for the parameter, $\basalfriction_0$, is the constant function with value $\basalfriction_0(x)=10.5$, which is halfway between the maximum and minimum values for $\basalfriction_\text{true}$.
+The constant $\delta$ is chosen so that the correlation length of functions drawn from $N(0,\mathcal{L}^{-2})$, which is given by the formula $L=\sqrt{\gamma/\delta}$, is $1/10$th of the radius of the domain.
+The function $\basalfriction_0$ is the constant function with value $\basalfriction_0(x)=10.5$, which is halfway between the maximum and minimum values of $\basalfriction_\text{true}$.
+
+The inverse problem is solved with a preconditioned inexact-Newton conjugate gradient scheme \cite{NOCEDAWRIGHT}.
+The constant function $\basalfriction_0$ is used as the initial guess.
 The parameter value is updated iteratively according to the formula
@@
 and $\alpha_k \in \mathbb{R}$ is a step length parameter chosen to satisfy the Armijo conditions \cite{ARMIJO}.
-Here $g_k$ is the gradient of $J$ with respect to $\basalfriction$ at $\basalfriction_k$, and $H$ is either the Hessian or Gauss-Newton Hessian of $J$ with respect to $\basalfriction$ at $\basalfriction_k$.
+Here $g_k$ is the gradient of $J$ with respect to $\basalfriction$ at $\basalfriction_k$, and $H$ is either the Hessian or Gauss-Newton Hessian of $J$ with respect to $\basalfriction$.
+The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \bm{T} \bm{u}$ is replaced by its local linearization at $\basalfriction_k$.
 The Gauss-Newton Hessian is used for the first five iterations, and the Hessian is used for all subsequent iterations.
-The method terminates when the norm of the gradient drops by a factor of $10^8$.
-The Newton system, \eqref{eq:newton_system}, is solved inexactly to the tolerance $\min \left(0.5, \sqrt{||g_k||}\right)||g_k||$, using the preconditioned conjugate gradient method, which only requires the application of the (Gauss-Newton) Hessian to vectors.
-Each application of the (Gauss-Newton) Hessian to a vector requires solving one forward and one adjoint Stokes PDE of the form \eqref{Stokeseqn} , but with different right hand sides.
+The method terminates when $||g_k||/||g_0|| < 10^{-8}$.
+The Newton systems, \eqref{eq:newton_system}, are solved inexactly using an inner conjugate gradient (CG) iteration, which we terminate when $||g_k + H p_k|| < \min \left(0.5, \sqrt{||g_k||}\right)||g_k||$.
+The preconditioned conjugate gradient method only requires the application of the (Gauss-Newton) Hessian to vectors.
+Each application of the (Gauss-Newton) Hessian to a vector requires solving one forward and one adjoint Stokes PDE of the form \eqref{Stokeseqn}, but with different right hand sides.
 
 We use the framework described in this paper to generate Hessian preconditioners.
-We use our methodology to build H-matrix approximations of the data-misfit component of the Gauss-Newton Hessian.
-That is, the term in the Gauss-Newton Hessian that arises from the data misfit term $J_d$.
+We build H-matrix approximations of the data-misfit Gauss-Newton Hessian (the term in the Gauss-Newton Hessian that arises from the data misfit) using 1, 5, and 25 impulse responses batches.
 We approximate the Gauss-Newton Hessian rather than the Hessian because the Hessian often has negative values in its integral kernel when one is away from the optimal point.
-At the discrete level, straightforward calculation shows that the regularization term in both the Hessian and the Gauss-Newton Hessian is given by
-\begin{equation*}
-	\bm{K} \bm{M}^{-1} \bm{K},
-\end{equation*}
-where $\bm{K}$ is the finite element discretization of the elliptic operator $\mathcal{K}$, and $\bm{M}$ is the finite element mass matrix.
-These matrices are sparse, so we form H-matrix approximations of these matrices using standard techniques for approximating sparse matrices with H-matrices, then invert $\bm{M}^{-1}$, and multiply these matrices together using H-matrix methods.
-We then add this H-matrix to the H-matrix for the data misfit term constructed using our method, and factor the resulting Hessian approximation using fast H-matrix factorization methods \cite{HLIBPRO}.
-We construct preconditioners using our method with 1, 5, and 25 impulse responses batches.
-The resulting preconditioners are denoted by PSF (1), PSF (5), and PSF (25), respectively.
-We compare the performance of these preconditioners with the most commonly used existing preconditioners, which are no preconditioning (denoted by NONE), and preconditioning by the regularization operator (denoted REG).
+Our results show that our preconditioners, which are built using the Gauss-Newton Hessian, are also good preconditioners for the Hessian.
+For the purpose of building the Hessian approximation, the basal surface, which is a 2D surface embedded in 3D due to the mountain topography, is flattened into the plane by ignoring the height coordinate.
+This flattening is necessary because our method requires adding points on the basal surface together to get other points on the basal surface, and this is only well-defined in flat space.
+
+The H-matrix approximation of data misfit Gauss-Newton Hessian is added to a H-matrix approximation of the Hessian of the regularization term, $J_r$, to form a H-matrix approximation of the Gauss-Newton Hessian.
+For the Hessian of the regularization term, straightforward calculation shows that the discrete regularization term in both the Hessian and the Gauss-Newton Hessian is given by
+$
+(\gamma \bm{K} + \delta \bm{M}) \bm{M}^{-1} (\gamma \bm{K} + \delta \bm{M})
+$
+where $\bm{K}$ is a finite element stiffness matrix with appropriate boundary conditions, and $\bm{M}$ is the finite element mass matrix.
+These matrices are sparse, so we form H-matrix approximations of these matrices using standard techniques for approximating sparse matrices with H-matrices.
+Then we invert $\bm{M}^{-1}$, and multiply these matrices together using H-matrix methods.
+We factor the overall Gauss-Newton Hessian approximation using fast H-matrix methods, then and use the factorization as a preconditioner \cite{HLIBPRO}.
+The resulting preconditioners for 1, 5, and 25 impulse response batches are denoted by PSF (1), PSF (5), and PSF (25), respectively.
+
+In Table \ref{tab:newton_convergence_table} we study the performance of our preconditioner for accelerating the solution of optimization problem \eqref{eq:deterministic_optimization_problem}.
+We build the PSF (5) preconditioner in the third Gauss-Newton iteration, and re-use it for all subsequent Gauss-Newton and Newton iterations.
+No preconditioner is used in the three Gauss-Newton iterations before the PSF (5) preconditioner is built.
+We compare our method with the most commonly used existing preconditioners, which are no preconditioning (denoted by NONE), and preconditioning by the regularization term in the Hessian (denoted REG).
+In Figure \ref{fig:newton_iterates} we show several Newton iterates using our PSF (5) preconditioner.
+Our results show that our PSF (5) preconditioner substantially outperforms the no preconditioning and regularization preconditioning.
+Using PSF (5) reduces the total number of Stokes PDE solves by a factor of six compared to no preconditioning, and by a factor of nine compared to regularization preconditioning.
+Interestingly, we observe that no preconditioning actually outperforms regularization preconditioning here because the noise level, and hence the regularization parameter, $\gamma$, are small.
+
+Next, we build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\basalfriction$.
+We use the preconditioned conjugate gradient method to solve a linear system with the Hessian as the coefficient operator and a right hand random vector with i.i.d.
+entries drawn from the standard normal distribution.
+In Figure \ref{fig:cg_convergence} we compare the convergence of conjugate gradient for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners.
+The preconditioned conjugate gradient method converges fastest with the PSF preconditioners, with PSF (25) converging fastest, followed by PSF (5), followed by PSF (1), as expected.
+Preconditioned conjugate gradient converges much slower with no preconditioning and regularization preconditioning than it does with PSF preconditioning, with no preconditioning outperforming regularization preconditioning.
+In Figure \ref{fig:gamma_sweep}, we perform preconditioned conjugate gradient solves on linear systems of the same form, except now we vary $\gamma$.
+The performance of our PSF preconditioners is good and relatively stable over a wide range of $\gamma$'s.
+All PSF preconditioners perform the same for medium and large values of $\gamma$.
+For small values of $\gamma$ PSF (25) performs slightly better than PSF (5), which performs slightly better than PSF (1).
+As expected, regularization preconditioning performs well for large $\gamma$ and poorly for small $\gamma$.
+Our PSF preconditioners outperform no preconditioning and regularization preconditioning for medium and small $\gamma$, and perform similarly to regularization preconditioning for large $\gamma$.
 
@@
 	%\end{subfigure}
-	\caption{The log basal friction parameter $\basalfriction$ after $1$ (left), $3$ (second from left), and $4$ (third from left) Newton iterates and the final Newton iterate (right).
-The preconditioner is constructed between Newton iterations 3 and 4.}
+	\caption{The log basal friction parameter $\basalfriction$ after $1$ (left), $3$ (second from left), and $4$ (third from left) Newton iterations, and the final Newton iterate (right).
+The PSF (5) preconditioner is constructed between Newton iterations 3 and 4 and re-used for all subsequent iterations.}
+	\label{fig:newton_iterates} 
 \end{figure}
@@
 This consists of Stokes solves for $u$ performed during the linesearch going from $\basalfriction_{k-1}$ to $\basalfriction_k$ (including one Stokes solve at $\basalfriction_k$), plus one adjoint Stokes solve to compute the gradient at $\basalfriction_k$, plus one incremental forward and one incremental adjoint Stokes solve per CG iteration for solving the Newton system at iteration $k$.
-In the PSF (5) portion of row 3, we write $6+22$ to indicate that 6 Stokes solves were used during the standard course of the iteration, and 22 additional Stokes solves were used to build our preconditioner.
+In the PSF (5) portion of row 3, we write $6+22$ to indicate that 6 Stokes solves were used during the standard course of the iteration, and 22 additional Stokes solves were used to build the PSF (5) preconditioner.
 }
+	\label{tab:newton_convergence_table}
 \end{table}
@@
 The preconditioner is constructed using the Gauss-Newton Hessian.
-For a fair comparison, the Hessian and Gauss-Newton Hessian are always evaluated at the same parameter $\basalfriction$, which is the optimal point for the inverse problem with $\gamma$ given by the Morozov discrepancy principle.}
+For a fair comparison, even as $\gamma$ changes the Hessian and Gauss-Newton Hessian are always evaluated at the same parameter $\basalfriction$, which is the optimal point for the inverse problem with $\gamma=7.3 \times 10^3$, as determined by the Morozov discrepancy principle.}
+	\label{fig:krylov_convergence}
 \end{figure}

### 2022-07-24 paper/localpsf.tex
more writing in numerical results section

@@
 
-		(optional) Modify $\mathbf{A}_H$ to make $\mathbf{A}_H^{\text{sym}+}$ (Section \ref{sec:make_spd})
+		(optional) Modify $\mathbf{A}_H$ to make $\mathbf{A}_H^{\text{sym}+}$ (Appendix \ref{sec:make_spd})
 
@@
 Our PSF preconditioners outperform no preconditioning and regularization preconditioning for medium and small $\gamma$, and perform similarly to regularization preconditioning for large $\gamma$.
+
+In Figure \ref{fig:generalized_eigenvalues} we show the generalized eigenvalues $\lambda$ for the generalized eigenvalue problem
+\begin{equation*}
+	H u = \lambda P u.
+\end{equation*}
+Here $H$ is the Hessian evaluated at the reconstructed parameter $\basalfriction$ for $\gamma$ chosen to satisfy the Morozov discrepancy principle.
+The matrix $P$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations constructed at that $\basalfriction$, or the regularization Hessian.
+The PSF preconditioners cluster the eigenvalues of the Hessian near $1$, with more batches yielding better clustering.
+The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
+In Figure \ref{fig:condition_number} we show the condition number of the preconditioned Hessian $P^{-1}H$, and preconditioned Gauss-Newton Hessian, $P^{-1}H_\text{gn}$.
+The condition numbers are similar for both the Hessian and the Gauss-Newton Hessian.
+The PSF preconditioners yield the the smallest condition numbers, with more batches yielding smaller condition numbers.
+Regularization preconditioning and no preconditioning yield larger condition numbers, with regularization preconditioning actually increasing the condition number as compared to no preconditioning.
 

### 2022-07-25 paper/localpsf.tex
edits from discussion with Tucker

@@
 
-By ``matrix free,'' we mean that we have a black-box computational procedure through which we may apply $\Aop$ and its transpose\footnote{Recall that the transpose $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(v) = \left(\Aop v\right)(u)$ for all $u,v \in L^2(\Omega)$, where the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(v)$ is the result of applying that linear functional to $v \in L^2(\Omega)$, and similar for the analogous operations with $\Aop^T$.
+By ``matrix free,'' we mean that we have a black-box computational procedure through which we may apply $\Aop$ and its transpose\footnote{Recall that the transpose $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop w\right)(u)$ for all $u,w \in L^2(\Omega)$, where the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for the analogous operations with $\Aop^T$.
 }, $\Aop^T$, to arbitrary functions,
@@
 \label{eq:eval_maps}
-u \mapsto\Aop u \qquad \text{and} \qquad v \mapsto \Aop^Tv,
+u \mapsto\Aop u \qquad \text{and} \qquad w \mapsto \Aop^T w,
 \end{equation} 
@@
 We address this challenge using \emph{impulse response interpolation}.
-The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(v) = \left(\rho^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
+The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
 We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with batches of points scattered throughout the domain (see Figure \ref{fig:batches_intro}).
@@
 \label{eq:kernel_representation}
-(\Aop u)(v) := \int_\Omega \int_\Omega v(y) \Aker(y,x) u(x) dx dy.
-\end{equation}
-Here the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(v)$ is the result of applying that linear functional to $v \in L^2(\Omega)$.
+(\Aop u)(w) := \int_\Omega \int_\Omega w(y) \Aker(y,x) u(x) dx dy.
+\end{equation}
+Here the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$.
 The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, which exists in principle but is not easily accessible to us.
@@
 	\begin{align}
-		\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Aker(y,x) \genericdistribution^*(x) dx ~dy \\
-		&= \int_\Omega v(y) \int_\Omega \Aker(y,x) \genericdistribution^*(x) dx ~dy \\
-		&= \int_\Omega v(y) \genericdistribution\left(\Aker(y,\cdot)\right) dy, \label{eq:action_on_distribution}
+		\left(\Aop \genericdistribution^*\right)(w) &= \int_\Omega \int_\Omega w(y) \Aker(y,x) \genericdistribution^*(x) dx ~dy \\
+		&= \int_\Omega w(y) \int_\Omega \Aker(y,x) \genericdistribution^*(x) dx ~dy \\
+		&= \int_\Omega w(y) \genericdistribution\left(\Aker(y,\cdot)\right) dy, \label{eq:action_on_distribution}
 	\end{align}
@@
 
-Let $\delta_x$ denote the delta distribution\footnote{Recall that the delta distribution $\delta_x:\mathcal{D}(\Omega)\rightarrow \mathbb{R}$ is the linear functional defined by $\delta_x(v) = v(x)$ for all $v\in \mathcal{D}(\Omega)$.} (i.e., point source, impulse) centered at the point $x \in \Omega$.
+Let $\delta_x$ denote the delta distribution\footnote{Recall that the delta distribution $\delta_x:\mathcal{D}(\Omega)\rightarrow \mathbb{R}$ is the linear functional defined by $\delta_x(w) = w(x)$ for all $w\in \mathcal{D}(\Omega)$.} (i.e., point source, impulse) centered at the point $x \in \Omega$.
 The \emph{impulse response} of $\Aop$ associated with $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$, 
@@
 \end{equation*}
-the space $\mathbb{R}^\fedim_\massmatrix$ is $\mathbb{R}^\fedim$ with the inner product $(\mathbf{u},\mathbf{v})_\massmatrix := \mathbf{u}^T \massmatrix \mathbf{v}$, and $\mathbb{R}^\fedim_{\massmatrix^{-1}}$ is the analogous space with $\massmatrix^{-1}$ replacing $\massmatrix$.
+the space $\mathbb{R}^\fedim_\massmatrix$ is $\mathbb{R}^\fedim$ with the inner product $(\mathbf{u},\mathbf{w})_\massmatrix := \mathbf{u}^T \massmatrix \mathbf{w}$, and $\mathbb{R}^\fedim_{\massmatrix^{-1}}$ is the analogous space with $\massmatrix^{-1}$ replacing $\massmatrix$.
 Direct calculation shows that $\mathbb{R}^\fedim_\massmatrix$ and $\mathbb{R}^\fedim_{\massmatrix^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively.
@@
 What is important is that we do not have direct access to matrix entries $\mathbf{A}_{ij}$.
-Rather, we have a computational procedure that allows us to compute matrix-vector products $\mathbf{u}\mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{v}\mapsto \mathbf{A}^T\mathbf{v}$, and computing these matrix-vector products is costly.
+Rather, we have a computational procedure that allows us to compute matrix-vector products $\mathbf{u}\mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{w}\mapsto \mathbf{A}^T\mathbf{w}$, and computing these matrix-vector products is costly.
 Of course, matrix entries can be computed via matrix-vector products as $\mathbf{A}_{ij} = \left(\mathbf{A}\mathbf{e}_j\right)_i$,
@@
 \begin{equation}
-(A_h u_h)(v_h) := \int_\Omega \int_\Omega v_h(y) \Aker_h(y,x) u_h(x) dx dy
+(A_h u_h)(w_h) := \int_\Omega \int_\Omega w_h(y) \Aker_h(y,x) u_h(x) dx dy
 \end{equation}
@@
 \begin{equation*}
-\mathbf{B}^T \mathbf{f} = \begin{bmatrix}
+\mathbf{B}^T \mathbf{w} = \begin{bmatrix}
 \horzbar & \mathbf{b}_1^T & \horzbar \\
@@
 \end{bmatrix}
-\mathbf{f} = 
+\mathbf{w} = 
 \begin{bmatrix}
-\mathbf{b}_1^T \mathbf{f} \\
-\mathbf{b}_2^T \mathbf{f} \\
+\mathbf{b}_1^T \mathbf{w} \\
+\mathbf{b}_2^T \mathbf{w} \\
 \vdots \\
-\mathbf{b}_\fedim^T \mathbf{f}
+\mathbf{b}_\fedim^T \mathbf{w}
 \end{bmatrix}.
 \end{equation*}
-By computing one matrix-vector product of $\mathbf{B}^T$ with $\mathbf{f}$, we compute the inner product of each column of $\mathbf{B}$ with $\mathbf{f}$ simultaneously.
+By computing one matrix-vector product of $\mathbf{B}^T$ with $\mathbf{w}$, we compute the inner product of each column of $\mathbf{B}$ with $\mathbf{w}$ simultaneously.
 The operator case is analogous, with the impulse response $\phi_x$ taking the place of a matrix column.
@@
 \label{eq:operator_simultaneous_ips}
-\left(\Aop^T f\right)^*(x) = \int_\Omega \Aker(y,x) f(y) dy = \left(\phi_x, f\right)_{L^2(\Omega)}.
-\end{equation}
-By computing one operator application of $\Aop^T$ to $f$, we compute the inner product of each impulse response $\phi_x$ with $f$, for all points $x$ simultaneously.
+\left(\Aop^T w\right)^*(x) = \int_\Omega \Aker(y,x) w(y) dy = \left(\phi_x, w\right)_{L^2(\Omega)}.
+\end{equation}
+By computing one operator application of $\Aop^T$ to $w$, we compute the inner product of each impulse response $\phi_x$ with $w$, for all points $x$ simultaneously.
 
@@
 for $i,j=1,\dots, \gdim$.
-Here $f/g$ denotes pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denotes pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
+Here $u/w$ denotes pointwise division of functions, $\left(u/w\right)(x) = u(x)/w(x)$, and $u\cdot w$ denotes pointwise multiplication of functions, $(u \cdot w)(x) = u(x)w(x)$.
 %The resulting method for computing discretized versions of $\spatialvol$, $\spatialmean$, and $\spatialcov$ is shown in Algorithm \ref{alg:varhpi_mean_cov}.
@@
 
-We start with a finite set of candidate points $P$ and build $\pointbatch_k$ incrementally with points selected from $P$.
-For simplicity of explanation, here $\pointbatch_k$ and $P$ are mutable sets that we add points to and remove points from.
+We start with a finite set of candidate points $\candidatepts$ and build $\pointbatch_k$ incrementally with points selected from $\candidatepts$.
+For simplicity of explanation, here $\pointbatch_k$ and $\candidatepts$ are mutable sets that we add points to and remove points from.
 First we initialize $\pointbatch_k$ as an empty set.
-Then we select the candidate point $p \in P$ that is the farthest away from all points in previous sample point batches $S_1 \cup S_2 \cup \dots \cup S_{k-1}$.
-Candidate points for the first batch $S_1$ are chosen randomly from $P$.
-Once $p$ is selected, we remove $p$ from $P$. Then we perform the following two checks:
+Then we select the candidate point $\candidatepoint_i \in \candidatepts$ that is the farthest away from all points in previous sample point batches $S_1 \cup S_2 \cup \dots \cup S_{k-1}$.
+Candidate points for the first batch $S_1$ are chosen randomly from $\candidatepts$.
+Once $\candidatepoint_i$ is selected, we remove $\candidatepoint_i$ from $\candidatepts$. Then we perform the following two checks:
 \begin{enumerate}
-	\item We check whether $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_k$.
-	\item We make sure that $\spatialvol(p)$ is not too small, by checking whether $\spatialvol(p) > \epsilon \spatialvol_\text{max}$.
-Here $\spatialvol_\text{max}$ is the largest value of $\spatialvol(q)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\epsilon=10^{-5}$).
+	\item We check whether $\candidatepoint_i$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ for all $\candidatepoint_j \in \pointbatch_k$.
+	\item We make sure that $\spatialvol(\candidatepoint_i)$ is not too small, by checking whether $\spatialvol(\candidatepoint_i) > \epsilon \spatialvol_\text{max}$.
+Here $\spatialvol_\text{max}$ is the largest value of $\spatialvol(\candidatepoint_j)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\epsilon=10^{-5}$).
 \end{enumerate}
-If $p$ passes both these checks (i.e., if $p$ is sufficiently far from other points in the batch, and $\spatialvol(p)$ is not too small) then we add $p$ to $\pointbatch_k$.
-Otherwise we discard $p$.
-This process repeats until there are no more points in $P$.
-This is detailed in Algorithm \ref{alg:point_choice}.
-We check whether $E_p \cap E_q = \{\}$ using the ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.
-
-We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots$, until the number of batches equals t desired threshold $\nbatch$.
-In our implementation, for each batch the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, except for points in previously chosen batches.
-
-\begin{algorithm2e}
-	\SetAlgoNoLine
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Output}
-	\SetKwProg{Fn}{Function}{}{}
-
-	\Input{Finite set of candidate points $P\subset \Omega$, \\spatially varying mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\previous sample point batches $\pointbatch_1, \dots, \pointbatch_{k-1}$}
-	\Output{Batch of new sample points $\pointbatch_k$}
-
-	Initialize empty new batch of sample points, $\pointbatch_k = \{\}$
-
-	\While{$P$ is not empty}{
-		Determine the point $p \in P$ that is farthest from all points in previous sample point batches $\pointbatch_1,\dots,\pointbatch_{k-1}$
-
-		Remove $p$ from $P$
-
-		\If{$E_p \cap E_q \neq \{\}$ for all $q \in \pointbatch_k$}{
-			\tcp{$E_p$ and $E_q$ are the ellipsoids defined in \eqref{eq:support_ellipsoid}}
-
-			Add $p$ to $\pointbatch_k$
-
-			Remove all points $p'$ satisfying $\spatialmean(p') \in E_p$ from $P$
-		}
-	}
-
-    \SetKwFunction{FMain}{point\_is\_acceptable}
-	\caption{Choosing one batch of sample points, $\pointbatch_k$}
-	\label{alg:point_choice}
-\end{algorithm2e}
+If $\candidatepoint_i$ passes both these checks (i.e., if $\candidatepoint_i$ is sufficiently far from other points in the batch, and $\spatialvol(\candidatepoint_i)$ is not too small) then we add $\candidatepoint_i$ to $\pointbatch_k$.
+Otherwise we discard $\candidatepoint_i$.
+This process repeats until there are no more points in $\candidatepts$.
+%This is detailed in Algorithm \ref{alg:point_choice}. 
+We check whether $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ using the ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.
+We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots, \pointbatch_{\nbatch}$.
+In our implementation, for each batch, $\candidatepts$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, except for points in previously chosen batches.
+
+%\begin{algorithm2e}
+%	\SetAlgoNoLine
+%	\SetKwInOut{Input}{Input}
+%	\SetKwInOut{Output}{Output}
+%	\SetKwProg{Fn}{Function}{}{}
+%	
+%	\Input{Finite set of candidate points $\candidatepts\subset \Omega$, \\spatially varying mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\previous sample point batches $\pointbatch_1, \dots, \pointbatch_{k-1}$}
+%	\Output{Batch of new sample points $\pointbatch_k$}
+%	
+%	Initialize empty new batch of sample points, $\pointbatch_k = \{\}$
+%	
+%	\While{$\candidatepts$ is not empty}{
+%		Determine the point $\candidatepoint_i \in \candidatepts$ that is farthest from all points in previous sample point batches $\pointbatch_1,\dots,\pointbatch_{k-1}$
+%		
+%		Remove $\candidatepoint_i$ from $\candidatepts$
+%
+%	
+%		\If{$E_{\candidatepoint_i} \cap E_q \neq \{\}$ for all $\candidatepoint_j \in \pointbatch_k$}{
+%			\tcp{$E_{\candidatepoint_i}$ and $E_{\candidatepoint_j}$ are the ellipsoids defined in \eqref{eq:support_ellipsoid}}
+%			
+%			Add $\candidatepoint_i$ to $\pointbatch_k$
+%			
+%			Remove all points $\candidatepoint$ satisfying $\spatialmean(\candidatepoint) \in E_{\candidatepoint_i}$ from $\candidatepts$
+%		}
+%	}
+%
+%    \SetKwFunction{FMain}{point\_is\_acceptable}
+%	\caption{Choosing one batch of sample points, $\pointbatch_k$}
+%	\label{alg:point_choice}
+%\end{algorithm2e}
 
@@
 \begin{equation}
-\Aker(y,x) \approx \widetilde{\Aker}(y,x) := \sum_{i=1}^{\numnbr} w_i~ b\left(\|x-x_i\|\right),
-\end{equation}
-where $b_i$ are radial basis functions and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_{\numnbr})^T$, is found as the solution to the $\numnbr \times \numnbr$ linear system
-\begin{equation}
-Bw = f,
+\Aker(y,x) \approx \widetilde{\Aker}(y,x) := \sum_{i=1}^{\numnbr} \rbfweight_i~ b\left(\|x-x_i\|\right),
+\end{equation}
+where $b_i$ are radial basis functions and $\rbfweight_i$ are weights.
+The vector of weights, $\rbfweight = (\rbfweight_1, \rbfweight_2, \dots, \rbfweight_{\numnbr})^T$, is found as the solution to the $\numnbr \times \numnbr$ linear system
+\begin{equation}
+B \rbfweight = f,
 \end{equation}
@@
 \begin{equation}
-b_i(r) := \begin{cases}
+b(r) := \begin{cases}
 r^\gdim, & \gdim=1,3,5,\dots \\
@@
 \end{equation}
-for $r>0$, and $b_i(0):=0$.
+for $r>0$, and $b(0):=0$.
 This yields the smoothest interpolant of the given points and values, in a certain least-squares sense \cite{Duchon77}.
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[scale=0.2]{interpolation_rbf_boundary.pdf}
-	\caption{To compute the kernel entry $\Aker(y,x)$, we use radial basis functions to interpolate the values $\frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}(y-\spatialmean(x)-\spatialmean(x_i))$ for the $\numnbr$ nearest sample points $x_i$ to $x$, excluding sample points for which $y-\spatialmean(x)-\spatialmean(x_i)$ is outside of the domain.}
-	\label{fig:rbf_interpolation}
-\end{figure}
+%\begin{figure}
+%	\centering
+%	\includegraphics[scale=0.2]{interpolation_rbf_boundary.pdf}
+%	\caption{To compute the kernel entry $\Aker(y,x)$, we use radial basis functions to interpolate the values $\frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}(y-\spatialmean(x)-\spatialmean(x_i))$ for the $\numnbr$ nearest sample points $x_i$ to $x$, excluding sample points for which $y-\spatialmean(x)-\spatialmean(x_i)$ is outside of the domain.}
+%	\label{fig:rbf_interpolation}
+%\end{figure}
 
@@
 	\begin{align}
-	-\nabla\cdot\bm{\sigma}_{\bm{u}}
-	&=\rho\bm{g}\,\,\,\text{ in }\icedomain, \label{Stokeseqn:1}\\
-	\nabla\cdot\bm{u}
+	-\nabla \cdot \stress
+	&=\bodyforce\,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:1}\\
+	\nabla\cdot \velocity
 	&=0\,\,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:2}\\
-	\bm{\sigma}_{\bm{u}}\bm{n}
-	&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{t}, \label{Stokeseqn:3}\\
-	\bm{u}\cdot\bm{n} =0 \quad\text{and}\quad \bm{T}\left(\bm{\sigma}_{\bm{u}}\bm{n}+\exp\left(\basalfriction\right)\bm{u}\right)
-	&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{b},\label{Stokeseqn:4} \\
-	\left(\bm{\sigma}_{\bm{u}}\bm{n}+\mu\bm{u}\right)&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{l}.\label{Stokeseqn:5}	
+	\stress\normalvec
+	&=0\,\,\,\,\,\,\text{ on }\Gamma_{t}, \label{Stokeseqn:3}\\
+	\velocity \cdot\normalvec =0 \quad\text{and}\quad \tangentop\left(\stress\normalvec+\exp\left(\basalfriction\right)\velocity\right)
+	&=0\,\,\,\,\,\,\text{ on }\Gamma_{b},\label{Stokeseqn:4} \\
+	\stress\normalvec+\mu \velocity &=0\,\,\,\,\,\,\text{ on }\Gamma_{l}.\label{Stokeseqn:5}	
 	\end{align}
 \end{subequations}
-The solution to these equations is the tuple $(\bm{u}, p)$, where $\bm{u}$ is the ice flow velocity field, and $p$ is the pressure field. Here, $\basalfriction$ is the unknown logarithmic
+The solution to these equations is the pair $(\velocity, \pressure)$, where $\velocity$ is the ice flow velocity field, and $\pressure$ is the pressure field.
+Here, $\basalfriction$ is the unknown logarithmic
 basal sliding parameter field which is defined on the $2D$ surface $\Gamma_b$.
-The quantity $\rho\bm{g}$ is the body force density due to gravity, $\bm{n}$ is the outward normal and $\bm{T}$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary.
-We employ Glen's flow law \cite{Glen1955},
+The quantity $\bodyforce$ is the body force density due to gravity, $\normalvec$ is the outward normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary.
+We employ Glen's flow law \cite{Glen1955}, $\stress= 2\eta
+\dot{\strain} -I\pressure$,
 which is a constitutive law for ice that relates the stress
-tensor $\bm{\sigma}_{\bm{u}}$ and the strain rate tensor
-$\bm{\dot{\varepsilon}}_{\bm{u}}= \frac{1}{2}\left(
-\bm{\nabla}\bm{u}+\bm{\nabla}\bm{u}^{\top} \right)$, $\bm{\sigma}_{\bm{u}}= 2\eta
-\bm{\dot{\varepsilon}}_{\bm{u}} -\bm{I}p$, where $\eta=\frac{1}{2}A^{-1}$ is the effective viscosity, $\bm{I}$ is the unit matrix,
-and $A$ is a flow rate factor.
+tensor, $\stress$, to the strain rate tensor,
+$\dot{\strain}= \frac{1}{2}\left(
+\nabla\velocity+\nabla\velocity^{\top} \right)$.
+Here $\eta$ is the effective viscosity and $I$ is the unit matrix.
 Glen's exponent has been chosen $n=1$, which provides for a linear Stokes model.
-Note that while the PDE is linear, the parameter-to-solution map, $\basalfriction \mapsto (\bm{u}, p)$, is nonlinear.
-
-The pressure, $p$, is discretized with first order continuous Galerkin finite elements defined on a mesh of tetrahedra discretizing $U$.
-The velocity, $\bm{u}$, is discretized with second order continuous Galerkin finite elements on the same mesh, as required for discrete stability.
+Note that while the PDE is linear, the parameter-to-solution map, $\basalfriction \mapsto (\velocity, \pressure)$, is nonlinear.
+
+The pressure, $\pressure$, is discretized with first order continuous Galerkin finite elements defined on a mesh of tetrahedra discretizing $U$.
+The velocity, $\velocity$, is discretized with second order continuous Galerkin finite elements on the same mesh.
 The parameter $\basalfriction$ is discretized with first order scalar continuous Galerkin finite elements on the mesh of triangles that results from restricting the tetrahedral mesh to the basal boundary, $\Gamma_b$.
+The basal boundary, $\Gamma_b$ is a 2D surface embedded in 3D due to the mountain topography.
+We also generate a flattened version of $\Gamma_b$, denoted by $\Omega \subset \mathbb{R}^2$, by ignoring the height coordinate.
+The parameter $\basalfriction$ is viewed as a function on $\Gamma_b$ for the purpose of solving the Stokes equation, but is viewed as a function on $\Omega$ for the purpose of building the Hessian preconditioner and defining the regularization.
+%This flattening is necessary because our operator approximation method requires adding points on this surface together to get other points on the surface, and this is only well-defined in flat space.
 
@@
 		\includegraphics[scale=0.243]{trueVelocity2_glyphs.png}
-		\caption{$\bm{u}_\text{true}$.}
-	\end{subfigure}
-	\caption{True logarithmic basal sliding parameter, $\basalfriction_\text{true}$, (left) and true velocity $\bm{u}_\text{true}$ (right).
-In the right plot, arrows indicate the direction of $\bm{u}_\text{true}$ and color indicates the magnitude.}
+		\caption{$\velocity_\text{true}$.}
+	\end{subfigure}
+	\caption{True logarithmic basal sliding parameter, $\basalfriction_\text{true}$, (left) and true velocity $\velocity_\text{true}$ (right).
+In the right plot, arrows indicate the direction of $\velocity_\text{true}$ and color indicates the magnitude.}
 	\label{fig:true_beta_u}
@@
 
-The true parameter field, $\basalfriction_\text{true}$, and corresponding velocity field, $\bm{u}_\text{true}$, are shown in Figure \ref{fig:true_beta_u}.
-We generate synthetic observations by restricting the tangential component of $\bm{u}_\text{true}$ to the top surface to get
-\begin{equation*}
-\bm{y}_\text{true} = \bm{T} \bm{u}_\text{true} |_{\Gamma_t}.
-\end{equation*}
-We add $1\%$ multiplicative gaussian noise, $\boldsymbol{\xi}$, to all components of the finite element vector representing $\bm{y}_\text{true}$ to get noisy synthetic observations, $\bm{y}_\text{obs}$, which are used for the inversion.
+The true parameter field, $\basalfriction_\text{true}$, and corresponding velocity field, $\velocity_\text{true}$, are shown in Figure \ref{fig:true_beta_u}.
+We generate synthetic observations by restricting the tangential component of $\velocity_\text{true}$ to the top surface to get
+\begin{equation*}
+\observations_\text{true} = \tangentop \velocity_\text{true} |_{\Gamma_t}.
+\end{equation*}
+We add $1\%$ multiplicative Gaussian noise to all components of the finite element vector, $\bm{\observations}_\text{true}$, corresponding to the function $\observations_\text{true}$, which yields noisy synthetic observations, $\observations_\text{obs}$ which are used for the inversion.
 That is,
 \begin{equation*}
-\left[\bm{y}_\text{obs}\right]_j := \left[\bm{y}_\text{true}\right]_j + \epsilon \left[\boldsymbol{\xi}\right]_j \left|\left[\bm{y}_\text{true}\right]_j\right|
-\end{equation*}
-where $\epsilon=0.01$, and the components of the noise function, $\boldsymbol{\xi}_j \sim N(0,1)$, are independent and identically distributed (i.i.d) random numbers drawn from the standard normal distribution.
-Here $\left[\bm{f}\right]_j$ denotes the $j$th entries of the finite element vector for the discretized versions of the function $\bm{f}$, and $j$ ranges over all entries of the finite element vector.
-Here we do not use bold to distinguish finite element vectors from their associated functions (as we did previously) because the functions $\bm{y}_\text{true}$, $\bm{y}_\text{obs}$, and $\bm{\xi}$ are already written in bold to denote the fact that they are vector valued.
+\left(\bm{\observations}_\text{obs}\right)_j := \left(\bm{\observations}_\text{true}\right)_j + 0.01 \cdot \boldsymbol{\xi}_j \cdot \left|\left(\bm{\observations}_\text{true}\right)_j\right|,
+\end{equation*}
+where $j$ ranges over all entries of the vector of coefficients representing the finite element functions $\observations_\text{true}$ and $\observations_\text{obs}$, and $\boldsymbol{\xi}_j$ are independent and identically distributed (i.i.d) random numbers drawn from the standard normal distribution.
+We also define the noise function, $\xi$, to be the finite element function with $j$th vector entry $\boldsymbol{\xi}_j$.
 
@@
 	\label{eq:deterministic_optimization_problem}
-	\min_{\basalfriction} \quad J(\basalfriction) = J_d(\basalfriction) + J_r(\basalfriction).
+	\min_{\basalfriction} \quad J(\basalfriction) := J_d(\basalfriction) + J_r(\basalfriction).
 \end{equation}
@@
 \begin{equation*}
-	J_d(\basalfriction) := \frac{1}{2}\int_{\Gamma_{t}}||\bm{d} - \bm{T}\bm{u}||_{2}^{2}\,dS.
+	J_d(\basalfriction) := \frac{1}{2}\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS.
 \end{equation*}
 We call this term the data misfit, because it measures the mismatch between the observed data and the predicted data based on a candidate parameter $\basalfriction$.
-Here $\bm{u} = \bm{u}(\basalfriction)$ denotes the velocity field solving $\ref{Stokeseqn}$ for the given parameter $\basalfriction$.
+Here $\velocity = \velocity(\basalfriction)$ denotes the velocity field solving $\ref{Stokeseqn}$ for the given parameter $\basalfriction$.
 The second term,
@@
 \begin{align*}
--\gamma\Delta u +\delta\,u&=s\,\,\,\,\text{ in }\Gamma_{b},\\
-\gamma \bm{\nabla}u\cdot\bm{n}+\mu\,u&=0\,\,\,\text{ on }\partial\Gamma_{b}.
+-\gamma\Delta u +\delta\,u&=s\,\,\,\,\text{ in }\Omega,\\
+\gamma \nabla u\cdot\normalvec+\mu\,u&=0\,\,\,\text{ on }\partial\Omega.
 \end{align*}
+Recall that $\Omega$ is the 2D flattened version of the basal surface $\Gamma_b$.
 In all results shown except one, $\gamma$ is chosen so that the solution satisfies the Morozov discrepancy principle, that is,
 \begin{equation*}
-	\frac{1}{2}\int_{\Gamma_{t}}||\bm{d} - \bm{T}\bm{u}||_{2}^{2}\,dS = \frac{1}{2}\int_{\Gamma_{t}}||\boldsymbol{\xi}||_{2}^{2}\,dS.
+	\frac{1}{2}\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS = \frac{1}{2}\int_{\Gamma_{t}}||\xi||_{2}^{2}\,dS.
 \end{equation*}
@@
 The exception is Figure \ref{fig:gamma_sweep}, where we vary $\gamma$ to study the effect of changing $\gamma$ on the quality of our preconditioner.
-The constant $\delta$ is chosen so that the correlation length of functions drawn from $N(0,\mathcal{L}^{-2})$, which is given by the formula $L=\sqrt{\gamma/\delta}$, is $1/10$th of the radius of the domain.
+The constant $\delta$ is chosen so that the correlation length, $L$, of functions drawn from the normal distribution with covariance $\mathcal{L}^{-2}$, which is given by the formula $L=\sqrt{\gamma/\delta}$, is $1/10$th of the radius of the domain.
 The function $\basalfriction_0$ is the constant function with value $\basalfriction_0(x)=10.5$, which is halfway between the maximum and minimum values of $\basalfriction_\text{true}$.
 
-The inverse problem is solved with a preconditioned inexact-Newton conjugate gradient scheme \cite{NOCEDAWRIGHT}.
+The inverse problem is solved with a inexact Newton preconditioned conjugate gradient scheme (PCG) \cite{NOCEDAWRIGHT}.
 The constant function $\basalfriction_0$ is used as the initial guess.
 The parameter value is updated iteratively according to the formula
-\begin{equation*}
-	\basalfriction_{k+1} = \basalfriction_k + \alpha_k p_k,
-\end{equation*}
-where the search direction $p_k$ is an approximate solution to the Newton linear system
+$
+	\basalfriction_{k+1} = \basalfriction_k + \alpha_k \searchdir_k,
+$
+where $\searchdir_k$ approximately solves one of the following Newton linear systems
 \begin{equation}
 	\label{eq:newton_system}
-	H p_k = - g_k,
-\end{equation}
-and $\alpha_k \in \mathbb{R}$ is a step length parameter chosen to satisfy the Armijo conditions \cite{ARMIJO}.
-Here $g_k$ is the gradient of $J$ with respect to $\basalfriction$ at $\basalfriction_k$, and $H$ is either the Hessian or Gauss-Newton Hessian of $J$ with respect to $\basalfriction$.
-The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \bm{T} \bm{u}$ is replaced by its local linearization at $\basalfriction_k$.
-The Gauss-Newton Hessian is used for the first five iterations, and the Hessian is used for all subsequent iterations.
+	H \searchdir_k = - g_k \qquad \text{or} \qquad H_\text{gn}, \searchdir_k = - g_k
+\end{equation}
+and the step length $\alpha_k \in \mathbb{R}$ is chosen to satisfy the Armijo conditions \cite{ARMIJO}.
+Here $g_k$ is the gradient of $J$, $H$ is the Hessian of $J$, and $H_\text{gn}$ is the Gauss-Newton Hessian, and these derivatives are evaluated at $\basalfriction_k$.
+%The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
+We use $H_\text{gn}$ for the first five iterations, and $H$ for all subsequent iterations.
 The method terminates when $||g_k||/||g_0|| < 10^{-8}$.
-The Newton systems, \eqref{eq:newton_system}, are solved inexactly using an inner conjugate gradient (CG) iteration, which we terminate when $||g_k + H p_k|| < \min \left(0.5, \sqrt{||g_k||}\right)||g_k||$.
-The preconditioned conjugate gradient method only requires the application of the (Gauss-Newton) Hessian to vectors.
-Each application of the (Gauss-Newton) Hessian to a vector requires solving one forward and one adjoint Stokes PDE of the form \eqref{Stokeseqn}, but with different right hand sides.
+The systems \eqref{eq:newton_system} are solved inexactly using an inner PCG iteration, which terminates when the norm of the residual for the linear system is less than $\min \left(0.5, \sqrt{||g_k||}\right)||g_k||$.
+The PCG method only requires applying $H$ (or $H_\text{gn}$) to vectors.
+Each application of $H$ (or $H_\text{gn}$) to a vector requires solving one incremental forward and one incremental adjoint Stokes PDE of the form \eqref{Stokeseqn}, but with different right hand sides.
+
+\begin{figure}
+	\begin{subfigure}{0.24\textwidth}
+		\includegraphics[scale=0.1]{m1_psf.png}
+		%		\caption{}
+		%		\label{fig:newtoniter1}
+	\end{subfigure}
+	\begin{subfigure}{0.24\textwidth}
+		\includegraphics[scale=0.1]{m3_psf.png}
+		%		\caption{}
+		%		\label{fig:newtoniter2}
+	\end{subfigure}
+	\begin{subfigure}{0.24\textwidth}
+		\includegraphics[scale=0.1]{m4_psf.png}
+		%		\caption{}
+		%		\label{fig:newtoniter3}
+	\end{subfigure}
+	\begin{subfigure}{0.24\textwidth}
+		\includegraphics[scale=0.1]{m6_psf.png}
+		%		\caption{}
+		%		\label{fig:newtoniter4}
+	\end{subfigure}
+	%\begin{subfigure}{0.05\textwidth}
+	%	\includegraphics[scale=0.1]{mColorBar.png}
+	%\end{subfigure}
+	\caption{The log basal friction parameter $\basalfriction$ after $1$ (left), $3$ (second from left), and $4$ (third from left) Newton iterations, and the final Newton iterate (right).
+The PSF (5) preconditioner is constructed between Newton iterations 3 and 4.}
+	\label{fig:newton_iterates} 
+\end{figure}
 
@@
 Our results show that our preconditioners, which are built using the Gauss-Newton Hessian, are also good preconditioners for the Hessian.
-For the purpose of building the Hessian approximation, the basal surface, which is a 2D surface embedded in 3D due to the mountain topography, is flattened into the plane by ignoring the height coordinate.
-This flattening is necessary because our method requires adding points on the basal surface together to get other points on the basal surface, and this is only well-defined in flat space.
 
@@
 
-In Table \ref{tab:newton_convergence_table} we study the performance of our preconditioner for accelerating the solution of optimization problem \eqref{eq:deterministic_optimization_problem}.
-We build the PSF (5) preconditioner in the third Gauss-Newton iteration, and re-use it for all subsequent Gauss-Newton and Newton iterations.
-No preconditioner is used in the three Gauss-Newton iterations before the PSF (5) preconditioner is built.
-We compare our method with the most commonly used existing preconditioners, which are no preconditioning (denoted by NONE), and preconditioning by the regularization term in the Hessian (denoted REG).
-In Figure \ref{fig:newton_iterates} we show several Newton iterates using our PSF (5) preconditioner.
-Our results show that our PSF (5) preconditioner substantially outperforms the no preconditioning and regularization preconditioning.
-Using PSF (5) reduces the total number of Stokes PDE solves by a factor of six compared to no preconditioning, and by a factor of nine compared to regularization preconditioning.
-Interestingly, we observe that no preconditioning actually outperforms regularization preconditioning here because the noise level, and hence the regularization parameter, $\gamma$, are small.
-
-Next, we build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\basalfriction$.
-We use the preconditioned conjugate gradient method to solve a linear system with the Hessian as the coefficient operator and a right hand random vector with i.i.d.
-entries drawn from the standard normal distribution.
-In Figure \ref{fig:cg_convergence} we compare the convergence of conjugate gradient for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners.
-The preconditioned conjugate gradient method converges fastest with the PSF preconditioners, with PSF (25) converging fastest, followed by PSF (5), followed by PSF (1), as expected.
-Preconditioned conjugate gradient converges much slower with no preconditioning and regularization preconditioning than it does with PSF preconditioning, with no preconditioning outperforming regularization preconditioning.
-In Figure \ref{fig:gamma_sweep}, we perform preconditioned conjugate gradient solves on linear systems of the same form, except now we vary $\gamma$.
-The performance of our PSF preconditioners is good and relatively stable over a wide range of $\gamma$'s.
-All PSF preconditioners perform the same for medium and large values of $\gamma$.
-For small values of $\gamma$ PSF (25) performs slightly better than PSF (5), which performs slightly better than PSF (1).
-As expected, regularization preconditioning performs well for large $\gamma$ and poorly for small $\gamma$.
-Our PSF preconditioners outperform no preconditioning and regularization preconditioning for medium and small $\gamma$, and perform similarly to regularization preconditioning for large $\gamma$.
-
-In Figure \ref{fig:generalized_eigenvalues} we show the generalized eigenvalues $\lambda$ for the generalized eigenvalue problem
-\begin{equation*}
-	H u = \lambda P u.
-\end{equation*}
-Here $H$ is the Hessian evaluated at the reconstructed parameter $\basalfriction$ for $\gamma$ chosen to satisfy the Morozov discrepancy principle.
-The matrix $P$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations constructed at that $\basalfriction$, or the regularization Hessian.
-The PSF preconditioners cluster the eigenvalues of the Hessian near $1$, with more batches yielding better clustering.
-The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
-In Figure \ref{fig:condition_number} we show the condition number of the preconditioned Hessian $P^{-1}H$, and preconditioned Gauss-Newton Hessian, $P^{-1}H_\text{gn}$.
-The condition numbers are similar for both the Hessian and the Gauss-Newton Hessian.
-The PSF preconditioners yield the the smallest condition numbers, with more batches yielding smaller condition numbers.
-Regularization preconditioning and no preconditioning yield larger condition numbers, with regularization preconditioning actually increasing the condition number as compared to no preconditioning.
-
-\begin{figure}[H]
-	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.1]{m1_psf.png}
-		%		\caption{}
-		%		\label{fig:newtoniter1}
-	\end{subfigure}
-	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.1]{m3_psf.png}
-		%		\caption{}
-		%		\label{fig:newtoniter2}
-	\end{subfigure}
-	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.1]{m4_psf.png}
-		%		\caption{}
-		%		\label{fig:newtoniter3}
-	\end{subfigure}
-	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.1]{m6_psf.png}
-		%		\caption{}
-		%		\label{fig:newtoniter4}
-	\end{subfigure}
-	%\begin{subfigure}{0.05\textwidth}
-	%	\includegraphics[scale=0.1]{mColorBar.png}
-	%\end{subfigure}
-	\caption{The log basal friction parameter $\basalfriction$ after $1$ (left), $3$ (second from left), and $4$ (third from left) Newton iterations, and the final Newton iterate (right).
-The PSF (5) preconditioner is constructed between Newton iterations 3 and 4 and re-used for all subsequent iterations.}
-	\label{fig:newton_iterates} 
-\end{figure}
-
-\begin{table}[H]
+\begin{table}
+	{\small
 		\begin{center}
@@
 			\setlength{\tabcolsep}{3pt}
-		\renewcommand{\arraystretch}{1.25}
+			\renewcommand{\arraystretch}{1.1}
 			\begin{tabular}{c| c c c | c c c | c c c}
@@
 		\end{center}
-	\caption{Convergence history for solving the Stokes inverse problem using the preconditioned inexact Newton-Krylov method.
-The Newton iteration terminates when the norm of the gradient has decreased from its initial value by a factor of $10^{-8}$.
-The Gauss-Newton Hessian is used in iterations $0$--$5$, and the Hessian is used in all subsequent iterations.
-Preconditioners shown are our method with five impulse response batches (PSF (5)), regularization preconditioning (REG), and no preconditioning (NONE).
-Our preconditioner (PSF (5)) is built at iteration $3$ using the Gauss-Newton Hessian, and re-used for all subsequent iterations.
-The iterates are $\basalfriction_0$ (initial guess), $\basalfriction_1$ (parameter after one Newton step), $\basalfriction_2$ (parameter after two Newton steps), and so on.
+	}
+	\caption{Convergence history for solving the Stokes inverse problem using the preconditioned inexact Newton-PCG method.
+Preconditioners shown are our method with five impulse response batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
 The columns titled \#CG show the number of conjugate gradient iterations used to solve the Newton system at $\basalfriction_k$, for the purpose of computing a search direction towards $\basalfriction_{k+1}$.
 The columns titled $||g||$ show the $l^2$ norm of the gradient at $\basalfriction_k$.
-The columns titled \#Stokes show the number of Stokes PDE solves performed in each Newton iteration.
-This consists of Stokes solves for $u$ performed during the linesearch going from $\basalfriction_{k-1}$ to $\basalfriction_k$ (including one Stokes solve at $\basalfriction_k$), plus one adjoint Stokes solve to compute the gradient at $\basalfriction_k$, plus one incremental forward and one incremental adjoint Stokes solve per CG iteration for solving the Newton system at iteration $k$.
-In the PSF (5) portion of row 3, we write $6+22$ to indicate that 6 Stokes solves were used during the standard course of the iteration, and 22 additional Stokes solves were used to build the PSF (5) preconditioner.
+The columns titled \#Stokes show the total number of Stokes PDE solves performed in each Newton iteration.
+This consists of Stokes solves for $u$ performed during the linesearch going from $\basalfriction_{k-1}$ to $\basalfriction_k$, plus the adjoint Stokes solve to compute the gradient at $\basalfriction_k$, plus one incremental forward and one incremental adjoint Stokes solve per PCG iteration for solving the Newton system at iteration $k$.
+In the PSF (5) portion of row 3, we write $6+22$ to indicate that 6 Stokes solves were used during the standard course of the iteration, and 22 Stokes solves were used to build the PSF (5) preconditioner.
 }
@@
 
-\begin{figure}[H]
+In Table \ref{tab:newton_convergence_table} we study the performance of our preconditioner for accelerating the solution of optimization problem \eqref{eq:deterministic_optimization_problem}.
+We build the PSF (5) preconditioner in the third Gauss-Newton iteration, and re-use it for all subsequent Gauss-Newton and Newton iterations.
+No preconditioner is used in the three Gauss-Newton iterations before the PSF (5) preconditioner is built.
+We compare our method with the most commonly used existing preconditioners, which are no preconditioning (denoted by NONE), and preconditioning by the regularization term in the Hessian (denoted REG).
+In Figure \ref{fig:newton_iterates} we show several Newton iterates using our PSF (5) preconditioner.
+Our results show that our PSF (5) preconditioner substantially outperforms the no preconditioning and regularization preconditioning.
+Using PSF (5) reduces the total number of Stokes PDE solves by a factor of six compared to no preconditioning, and by a factor of nine compared to regularization preconditioning.
+Interestingly, we observe that no preconditioning actually outperforms regularization preconditioning here because the noise level, and hence the regularization parameter, $\gamma$, are small.
+
+\begin{figure}
 	\begin{subfigure}{0.49\textwidth}
@@
 
-\begin{figure}[H]
+Next, we build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\basalfriction$.
+We use the preconditioned conjugate gradient method to solve a linear system with the Hessian as the coefficient operator and a right hand random vector with i.i.d.
+entries drawn from the standard normal distribution.
+In Figure \ref{fig:cg_convergence} we compare the convergence of conjugate gradient for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners.
+The preconditioned conjugate gradient method converges fastest with the PSF preconditioners, with PSF (25) converging fastest, followed by PSF (5), followed by PSF (1), as expected.
+Preconditioned conjugate gradient converges much slower with no preconditioning and regularization preconditioning than it does with PSF preconditioning, with no preconditioning outperforming regularization preconditioning.
+In Figure \ref{fig:gamma_sweep}, we perform preconditioned conjugate gradient solves on linear systems of the same form, except now we vary $\gamma$.
+The performance of our PSF preconditioners is good and relatively stable over a wide range of $\gamma$'s.
+All PSF preconditioners perform the same for medium and large values of $\gamma$.
+For small values of $\gamma$ PSF (25) performs slightly better than PSF (5), which performs slightly better than PSF (1).
+As expected, regularization preconditioning performs well for large $\gamma$ and poorly for small $\gamma$.
+Our PSF preconditioners outperform no preconditioning and regularization preconditioning for medium and small $\gamma$, and perform similarly to regularization preconditioning for large $\gamma$.
+
+\begin{figure}
 	\begin{subfigure}{0.59\textwidth}
@@
 			legend style={at={(0.95,1)},anchor=north east},
-			title={Generalized eigenvalues, $Hu_{k}=\lambda_{k}Pu_{k}$}
+			title={Generalized eigenvalues, $Hu_{k}=\lambda_{k}\preconditioner u_{k}$}
 			]
@@
 			\begin{tabular}{c| c c}
-				& $\kappa(P^{-1}H)$ &	$\kappa(P^{-1}H_{\text{gn}})$ \\
+				& $\kappa(\preconditioner^{-1}H)$ &	$\kappa(\preconditioner^{-1}H_{\text{gn}})$ \\
 				\hline 
@@
 	\end{subfigure}
-	\caption{Left (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $H u_k = \lambda_k P_k$.
-Here $H$ is the Hessian and $P$ is the PSF approximation constructed using the Gauss-Newton Hessian, $H_\text{gn}$, for 1, 5, and 25 batches, or the regularization operator.
+	\caption{Left (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $H u_k = \lambda_k \preconditioner_k$.
+Here $H$ is the Hessian and $\preconditioner$ is the PSF approximation constructed using the Gauss-Newton Hessian, $H_\text{gn}$, for 1, 5, and 25 batches, or the regularization operator.
 Right (\ref{fig:condition_number}): condition number, $\kappa$, for the preconditioned Hessian and Gauss-Newton Hessian for different preconditioners.
@@
 \end{figure}
+
+In Figure \ref{fig:generalized_eigenvalues} we show the generalized eigenvalues $\lambda$ for the generalized eigenvalue problem
+\begin{equation*}
+	H u = \lambda \preconditioner u.
+\end{equation*}
+Here $H$ is the Hessian evaluated at the reconstructed parameter $\basalfriction$ for $\gamma$ chosen to satisfy the Morozov discrepancy principle.
+The matrix $\preconditioner$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations constructed at that $\basalfriction$, or the regularization Hessian.
+The PSF preconditioners cluster the eigenvalues of the Hessian near $1$, with more batches yielding better clustering.
+The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
+In Figure \ref{fig:condition_number} we show the condition number of the preconditioned Hessian $\preconditioner^{-1}H$, and preconditioned Gauss-Newton Hessian, $\preconditioner^{-1}H_\text{gn}$.
+The condition numbers are similar for both the Hessian and the Gauss-Newton Hessian.
+The PSF preconditioners yield the the smallest condition numbers, with more batches yielding smaller condition numbers.
+Regularization preconditioning and no preconditioning yield larger condition numbers, with regularization preconditioning actually increasing the condition number as compared to no preconditioning.
 
@@
 \begin{equation*}
-	\spatialcov_1 P = \spatialcov_2 P \Lambda,
-\end{equation*}
-where $P \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$.
-The matrix $P$ simultaneously diagonalizes $\spatialcov_1$ and $\spatialcov_2$, in the sense that $P^T\spatialcov_1 P = \Lambda$, and $P^T\spatialcov_2 P = I$, where $I$ is the $\gdim \times \gdim$ identity matrix.
+	\spatialcov_1 \eigenvectormatrix = \spatialcov_2 \eigenvectormatrix \Lambda,
+\end{equation*}
+where $\eigenvectormatrix \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$.
+The matrix $\eigenvectormatrix$ simultaneously diagonalizes $\spatialcov_1$ and $\spatialcov_2$, in the sense that $\eigenvectormatrix^T\spatialcov_1 \eigenvectormatrix = \Lambda$, and $\eigenvectormatrix^T\spatialcov_2 \eigenvectormatrix = I$, where $I$ is the $\gdim \times \gdim$ identity matrix.
 Using this diagonalization, and some algebraic manipulations, we may write $K(s)$ as
@@
 \end{equation}
-where $v := P^T\left(\spatialmean_1 - \spatialmean_2\right)$.
+where $v := \eigenvectormatrix^T\left(\spatialmean_1 - \spatialmean_2\right)$.
 

### 2022-07-25 paper/localpsf.tex
more paper edits. Barely under 25 pages

@@
 
-The pressure, $\pressure$, is discretized with first order continuous Galerkin finite elements defined on a mesh of tetrahedra discretizing $U$.
+The pressure, $\pressure$, is discretized with first order continuous Galerkin finite elements defined on a mesh of tetrahedra.
 The velocity, $\velocity$, is discretized with second order continuous Galerkin finite elements on the same mesh.
-The parameter $\basalfriction$ is discretized with first order scalar continuous Galerkin finite elements on the mesh of triangles that results from restricting the tetrahedral mesh to the basal boundary, $\Gamma_b$.
-The basal boundary, $\Gamma_b$ is a 2D surface embedded in 3D due to the mountain topography.
+The parameter $\basalfriction$ is discretized with first order scalar continuous Galerkin finite elements on the mesh of triangles that results from restricting the tetrahedral mesh to $\Gamma_b$.
+The basal boundary, $\Gamma_b$, is a 2D surface embedded in 3D due to the mountain topography.
 We also generate a flattened version of $\Gamma_b$, denoted by $\Omega \subset \mathbb{R}^2$, by ignoring the height coordinate.
-The parameter $\basalfriction$ is viewed as a function on $\Gamma_b$ for the purpose of solving the Stokes equation, but is viewed as a function on $\Omega$ for the purpose of building the Hessian preconditioner and defining the regularization.
+The parameter $\basalfriction$ is dually viewed as a function on $\Gamma_b$ for the purpose of solving the Stokes equations, and as a function on $\Omega$ for the purpose of building the Hessian preconditioner and defining the regularization.
 %This flattening is necessary because our operator approximation method requires adding points on this surface together to get other points on the surface, and this is only well-defined in flat space.
@@
 	\end{subfigure}
-	\caption{True logarithmic basal sliding parameter, $\basalfriction_\text{true}$, (left) and true velocity $\velocity_\text{true}$ (right).
-In the right plot, arrows indicate the direction of $\velocity_\text{true}$ and color indicates the magnitude.}
+	\caption{True parameter, $\basalfriction_\text{true}$, (left) and true velocity $\velocity_\text{true}$ (right).
+In the right plot, arrows indicate the direction of $\velocity_\text{true}$ and color indicates the magnitude of $\velocity_\text{true}$.}
 	\label{fig:true_beta_u}
@@
 We use the framework described in this paper to generate Hessian preconditioners.
-We build H-matrix approximations of the data-misfit Gauss-Newton Hessian (the term in the Gauss-Newton Hessian that arises from the data misfit) using 1, 5, and 25 impulse responses batches.
-We approximate the Gauss-Newton Hessian rather than the Hessian because the Hessian often has negative values in its integral kernel when one is away from the optimal point.
-Our results show that our preconditioners, which are built using the Gauss-Newton Hessian, are also good preconditioners for the Hessian.
-
+We build H-matrix approximations of the data-misfit Gauss-Newton Hessian (the term in $H_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 impulse responses batches.
+The resulting H-matrices, which may be indefinite due to approximation error, are modified to be symmetric positive definite via the procedure described in Appendix \ref{sec:make_spd}, with $l=2$.
+We approximate $H_\text{gn}$ rather than $H$ because $H$ often has negative values in its integral kernel when one is away from the optimal point.
+Our results show that our preconditioners are also good preconditioners for $H$.
 The H-matrix approximation of data misfit Gauss-Newton Hessian is added to a H-matrix approximation of the Hessian of the regularization term, $J_r$, to form a H-matrix approximation of the Gauss-Newton Hessian.
@@
 where $\bm{K}$ is a finite element stiffness matrix with appropriate boundary conditions, and $\bm{M}$ is the finite element mass matrix.
-These matrices are sparse, so we form H-matrix approximations of these matrices using standard techniques for approximating sparse matrices with H-matrices.
+These matrices are sparse, so we form H-matrix representations of these matrices using standard sparse H-matrix techniques.
 Then we invert $\bm{M}^{-1}$, and multiply these matrices together using H-matrix methods.
-We factor the overall Gauss-Newton Hessian approximation using fast H-matrix methods, then and use the factorization as a preconditioner \cite{HLIBPRO}.
+We factor the overall Gauss-Newton Hessian approximation, denoted $\widetilde{H}$, using fast H-matrix methods, then and use the factorization as a preconditioner \cite{HLIBPRO}.
 The resulting preconditioners for 1, 5, and 25 impulse response batches are denoted by PSF (1), PSF (5), and PSF (25), respectively.
@@
 We build the PSF (5) preconditioner in the third Gauss-Newton iteration, and re-use it for all subsequent Gauss-Newton and Newton iterations.
-No preconditioner is used in the three Gauss-Newton iterations before the PSF (5) preconditioner is built.
+No preconditioning is used in the three Gauss-Newton iterations before the PSF (5) preconditioner is built.
 We compare our method with the most commonly used existing preconditioners, which are no preconditioning (denoted by NONE), and preconditioning by the regularization term in the Hessian (denoted REG).
-In Figure \ref{fig:newton_iterates} we show several Newton iterates using our PSF (5) preconditioner.
-Our results show that our PSF (5) preconditioner substantially outperforms the no preconditioning and regularization preconditioning.
+In Figure \ref{fig:newton_iterates} we show select Newton iterates using our PSF (5) preconditioner.
 Using PSF (5) reduces the total number of Stokes PDE solves by a factor of six compared to no preconditioning, and by a factor of nine compared to regularization preconditioning.
@@
 		xlabel={CG iteration},
-		ylabel={Relative residual norm},
+		ylabel={$||b-Hx_k||/||b||$},
 		title={CG convergence},
@@
 	\end{subfigure}
-	\caption{Left (\ref{fig:cg_convergence}) shows the convergence history for solving a linear system of the form $Hx=b$ using the preconditioned conjugate gradient method, where $b$ has random independent and identically distributed (i.i.d) entries drawn from the standard normal distribution.
-The y-axis is the relative $l^2$ norm of the residual, i.e., $||b-Hx_k||/||b||$, where $x_k$ is the $k$th iterate.
-Here the regularization parameter $\gamma$ is determined by the Morozov discrepancy principle.
-Right (\ref{fig:gamma_sweep}) shows the number of preconditioned conjugate gradient iterations required to reduce the relative $l^2$ norm of the residual below a tolerance of $10^{-6}$, for a range of different regularization parameters $\gamma$.
+	\caption{Left (\ref{fig:cg_convergence}) shows the convergence history for solving $Hx=b$ using PCG, where $b$ has i.i.d. random entries drawn from the standard normal distribution. 
+%		The y-axis is $||b-Hx_k||/||b||$, where $x_k$ is the $k$th PCG iterate. 
+	Here we use $\gamma=7.3 \times 10^3$, which is determined by the Morozov discrepancy principle.
+Right (\ref{fig:gamma_sweep}) shows the number of PCG iterations required to achieve $||b-Hx_k||/||b|| < 10^{-6}$, where $x_k$ is the $k$th PCG iterate, for a range of different $\gamma$.
 Results in these figures are shown for our preconditioner with 1, 5, and 25 batches of impulse responses (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
-The preconditioner is constructed using the Gauss-Newton Hessian.
-For a fair comparison, even as $\gamma$ changes the Hessian and Gauss-Newton Hessian are always evaluated at the same parameter $\basalfriction$, which is the optimal point for the inverse problem with $\gamma=7.3 \times 10^3$, as determined by the Morozov discrepancy principle.}
+The preconditioner is constructed using $H_\text{gn}$.
+For a fair comparison, even as $\gamma$ changes, $H$ and $H_\text{gn}$ are always evaluated at the same $\basalfriction$, which is the optimal point for the inverse problem with $\gamma=7.3 \times 10^3$.}
 	\label{fig:krylov_convergence}
@@
 entries drawn from the standard normal distribution.
-In Figure \ref{fig:cg_convergence} we compare the convergence of conjugate gradient for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners.
-The preconditioned conjugate gradient method converges fastest with the PSF preconditioners, with PSF (25) converging fastest, followed by PSF (5), followed by PSF (1), as expected.
-Preconditioned conjugate gradient converges much slower with no preconditioning and regularization preconditioning than it does with PSF preconditioning, with no preconditioning outperforming regularization preconditioning.
-In Figure \ref{fig:gamma_sweep}, we perform preconditioned conjugate gradient solves on linear systems of the same form, except now we vary $\gamma$.
+In Figure \ref{fig:cg_convergence} we compare the convergence of PCG for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners.
+PCG converges fastest with the PSF preconditioners, with PSF (25) converging fastest, followed by PSF (5), followed by PSF (1), as expected.
+PCG converges much slower with no preconditioning and regularization preconditioning than it does with PSF preconditioning, with no preconditioning outperforming regularization preconditioning.
+In Figure \ref{fig:gamma_sweep}, we perform PCG solves on linear systems of the same form, except now we vary $\gamma$.
 The performance of our PSF preconditioners is good and relatively stable over a wide range of $\gamma$'s.
@@
 	\caption{Left (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $H u_k = \lambda_k \preconditioner_k$.
-Here $H$ is the Hessian and $\preconditioner$ is the PSF approximation constructed using the Gauss-Newton Hessian, $H_\text{gn}$, for 1, 5, and 25 batches, or the regularization operator.
-Right (\ref{fig:condition_number}): condition number, $\kappa$, for the preconditioned Hessian and Gauss-Newton Hessian for different preconditioners.
-All operators are evaluated at the parameter $\basalfriction$ that solves the inverse problem for $\gamma$ given by the Morozov discrepancy principle.}
+Here $H$ is the Hessian and $\preconditioner$ is the PSF approximation constructed using the Gauss-Newton Hessian, $H_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), or the regularization Hessian (REG).
+Right (\ref{fig:condition_number}): condition number, $\kappa$, for $\widetilde{H}^{-1} H$ and $\widetilde{H}^{-1} H_\text{gn}$ for these different preconditioners, and no preconditioner (NONE).
+All operators are evaluated at the $\basalfriction$ that solves the inverse problem for $\gamma=7.3 \times 10^3$, as determined by the Morozov discrepancy principle.}
 \end{figure}
@@
 In Figure \ref{fig:generalized_eigenvalues} we show the generalized eigenvalues $\lambda$ for the generalized eigenvalue problem
-\begin{equation*}
+$
 	H u = \lambda \preconditioner u.
-\end{equation*}
+$
 Here $H$ is the Hessian evaluated at the reconstructed parameter $\basalfriction$ for $\gamma$ chosen to satisfy the Morozov discrepancy principle.
@@
 \end{equation*}
-In \cite{GilitschenskiHanebeck12} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$ (see Figure \ref{fig:ellipsoid_intersection_test}).
-
-\begin{figure}
-	\begin{subfigure}{0.37\textwidth}
-		\centering
-		\includegraphics[scale=1.0]{ellipsoids_intersect0.pdf}
-	\end{subfigure}
-	\begin{subfigure}{0.30\textwidth}
-		\centering
-		\includegraphics[scale=1.0]{ellipsoids_intersect1.pdf}
-	\end{subfigure}
-	\begin{subfigure}{0.30\textwidth}
-		\centering
-		\includegraphics[scale=1.0]{ellipsoids_intersect2.pdf}
-	\end{subfigure}
-	\caption{The ellipsoids intersect if $K(s) \ge 0$ for all $s \in (0,1)$.
-	}
-	\label{fig:ellipsoid_intersection_test}
-\end{figure}
+In \cite{GilitschenskiHanebeck12} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$.
+% (see Figure \ref{fig:ellipsoid_intersection_test}).
+
+%\begin{figure}
+%	\begin{subfigure}{0.37\textwidth}
+%		\centering
+%		\includegraphics[scale=1.0]{ellipsoids_intersect0.pdf}
+%	\end{subfigure}
+%	\begin{subfigure}{0.30\textwidth}
+%		\centering
+%		\includegraphics[scale=1.0]{ellipsoids_intersect1.pdf}
+%	\end{subfigure}
+%	\begin{subfigure}{0.30\textwidth}
+%		\centering
+%		\includegraphics[scale=1.0]{ellipsoids_intersect2.pdf}
+%	\end{subfigure}
+%	\caption{The ellipsoids intersect if $K(s) \ge 0$ for all $s \in (0,1)$.
+%	}
+%	\label{fig:ellipsoid_intersection_test}
+%\end{figure}
 
@@
 where $\Sigma_{ii}$ is the $i$th diagonal entry of $\Sigma$\footnote{See \url{https://math.stackexchange.com/q/3928964}}. The boxes
-\begin{equation*}
-	\prod_{i=1}^\gdim \left[a_i^-, a_i^+\right], \quad \text{and} \quad \prod_{i=1}^\gdim \left[b_i^-, b_i^+\right]
-\end{equation*}
+$\prod_{i=1}^\gdim \left[a_i^-, a_i^+\right]$ and $\prod_{i=1}^\gdim \left[b_i^-, b_i^+\right]$
 intersect if and only if $a_i^- \le b_i^+$ and $b_i^- \le a_i^+$ for $i=1,\dots,d$.
@@
 If $\Aop$ is symmetric, we symmetrize $\mathbf{A}_H$ as follows
-\begin{equation*}
+$
 \mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
-\end{equation*}
+$
 If $\Aop$ is also positive semi-definite, we adapt the high order rational function spectral projection method from \cite[Section 13.8]{HackbuschKress07} to flip (or approximately flip) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ to be positive.
@@
 Using functional calculus to extend the domain of the function $\Pi$ from scalars to matrices, we define
-\begin{equation*}
+$
 \mathbf{\Pi} := \Pi(\mathbf{A}_H^{\text{sym}}).
-\end{equation*}
+$
 The matrix $\mathbf{\Pi}$ is the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to eigenvalues of $\mathbf{A}_H^{\text{sym}}$ that reside in $(a,b)$.
@@
 where
-\begin{equation*}
+$
 \mathbf{T} := T(\mathbf{A}_H^{\text{sym}}) = \left(2 \mathbf{A}_H^{\text{sym}} - (b+a) \mathbf{I}\right)/(b-a).
-\end{equation*}
+$
 Using $\mathbf{\Pi}_\ratord$ instead of $\mathbf{\Pi}$ in \eqref{eq:I_Pi_abs} allows us to efficiently approximate $|\mathbf{A}_H^{\text{sym}}|$ using H-matrix methods. 

### 2022-07-25 paper/localpsf.tex
more paper edits. Barely under 25 pages

@@
 
-Algorithm \ref{alg:varhpi_mean_cov} from Section \ref{sec:intromoments} allows us to compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} uses $\spatialvol$, $\spatialmean$, and $\spatialcov$ to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
+Section \ref{sec:intromoments} allows us to compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} uses $\spatialvol$, $\spatialmean$, and $\spatialcov$ to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
 This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}).
@@
 We factor the overall Gauss-Newton Hessian approximation, denoted $\widetilde{H}$, using fast H-matrix methods, then and use the factorization as a preconditioner \cite{HLIBPRO}.
-The resulting preconditioners for 1, 5, and 25 impulse response batches are denoted by PSF (1), PSF (5), and PSF (25), respectively.
+The approximations for 1, 5, and 25 impulse response batches are denoted by PSF (1), PSF (5), and PSF (25), respectively.
 

### 2022-07-25 paper/localpsf.tex
more paper edits. Barely under 25 pages

@@
 Color indicates the height of the base of the ice sheet (i.e., the mountain topography).
-The radius of the domain is $10$ kilometers, and the maximum height of the mountain is $2.1$ kilometers.}
+The radius of the domain is $10$ kilometers, the maximum height of the mountain is $2.1$ kilometers, and the average thickness of the ice sheet is 250 meters.}
 	\label{fig:ice_mountain_mesh}
@@
 The exception is Figure \ref{fig:gamma_sweep}, where we vary $\gamma$ to study the effect of changing $\gamma$ on the quality of our preconditioner.
-The constant $\delta$ is chosen so that the correlation length, $L$, of functions drawn from the normal distribution with covariance $\mathcal{L}^{-2}$, which is given by the formula $L=\sqrt{\gamma/\delta}$, is $1/10$th of the radius of the domain.
-The function $\basalfriction_0$ is the constant function with value $\basalfriction_0(x)=10.5$, which is halfway between the maximum and minimum values of $\basalfriction_\text{true}$.
-
-The inverse problem is solved with a inexact Newton preconditioned conjugate gradient scheme (PCG) \cite{NOCEDAWRIGHT}.
+The constant $\delta$ is chosen so that the correlation length, $L=\sqrt{\gamma/\delta}$, of functions drawn from the normal distribution with covariance $\mathcal{L}^{-2}$, is $1/10$th of the radius of the domain.
+The function $\basalfriction_0$ is the constant function with value $\basalfriction_0(x)=10.5$.
+%, which is halfway between the maximum and minimum values of $\basalfriction_\text{true}$.
+
+The inverse problem is solved with a inexact Newton preconditioned conjugate gradient (PCG) scheme \cite{NOCEDAWRIGHT}.
 The constant function $\basalfriction_0$ is used as the initial guess.
@@
 Our results show that our preconditioners are also good preconditioners for $H$.
-The H-matrix approximation of data misfit Gauss-Newton Hessian is added to a H-matrix approximation of the Hessian of the regularization term, $J_r$, to form a H-matrix approximation of the Gauss-Newton Hessian.
-For the Hessian of the regularization term, straightforward calculation shows that the discrete regularization term in both the Hessian and the Gauss-Newton Hessian is given by
+The H-matrix approximation of data misfit Gauss-Newton Hessian is added to a H-matrix approximation of the Hessian of the regularization term to form a H-matrix approximation of $H_\text{gn}$.
+Straightforward calculation shows that the discretized Hessian of the regularization term is given by
 $
@@
 Using PSF (5) reduces the total number of Stokes PDE solves by a factor of six compared to no preconditioning, and by a factor of nine compared to regularization preconditioning.
-Interestingly, we observe that no preconditioning actually outperforms regularization preconditioning here because the noise level, and hence the regularization parameter, $\gamma$, are small.
+Interestingly, we observe that no preconditioning actually outperforms regularization preconditioning here because the noise level (and hence $\gamma$) is small.
 

### 2022-07-26 paper/localpsf.tex
more paper edits. Barely under 25 pages

@@
 \end{equation*}
-is a Bilaplacian regularization term. Specifically, $\mathcal{L}$ is the inverse of the solution operator for the following elliptic PDE:
+is a Bilaplacian regularization term.
+Specifically, $\basalfriction_0$ is the constant function $\basalfriction_0(x)=10.5$, and $\mathcal{L}$ is the inverse of the solution operator for the following elliptic PDE:
 \begin{align*}
@@
 Recall that $\Omega$ is the 2D flattened version of the basal surface $\Gamma_b$.
-In all results shown except one, $\gamma$ is chosen so that the solution satisfies the Morozov discrepancy principle, that is,
+In all results except one, $\gamma$ is chosen so that the Morozov discrepancy principle is satisfied, i.e.,
 \begin{equation*}
@@
 The value of $\gamma$ satisfying the Morozov discrepancy principle is $\gamma=7.3 \times 10^3$.
-The exception is Figure \ref{fig:gamma_sweep}, where we vary $\gamma$ to study the effect of changing $\gamma$ on the quality of our preconditioner.
-The constant $\delta$ is chosen so that the correlation length, $L=\sqrt{\gamma/\delta}$, of functions drawn from the normal distribution with covariance $\mathcal{L}^{-2}$, is $1/10$th of the radius of the domain.
-The function $\basalfriction_0$ is the constant function with value $\basalfriction_0(x)=10.5$.
+The exception is Figure \ref{fig:gamma_sweep}, where we vary $\gamma$ to study how changing $\gamma$ impacts the effectiveness of our preconditioner.
+The constant $\delta$ is chosen so that the correlation length of functions drawn from the normal distribution with covariance $\mathcal{L}^{-2}$, given by $L=\sqrt{\gamma/\delta}$, is $1/10$th the radius of the domain.
 %, which is halfway between the maximum and minimum values of $\basalfriction_\text{true}$.
 
-The inverse problem is solved with a inexact Newton preconditioned conjugate gradient (PCG) scheme \cite{NOCEDAWRIGHT}.
-The constant function $\basalfriction_0$ is used as the initial guess.
-The parameter value is updated iteratively according to the formula
+We solve the inverse problem with an inexact Newton preconditioned conjugate gradient (PCG) scheme \cite{NOCEDAWRIGHT}, using $\basalfriction_0$ as the initial guess.
+The parameter is updated iteratively according to the formula
 $
@@
 and the step length $\alpha_k \in \mathbb{R}$ is chosen to satisfy the Armijo conditions \cite{ARMIJO}.
-Here $g_k$ is the gradient of $J$, $H$ is the Hessian of $J$, and $H_\text{gn}$ is the Gauss-Newton Hessian, and these derivatives are evaluated at $\basalfriction_k$.
+Here $g_k$ is the gradient of $J$, $H$ is the Hessian, and $H_\text{gn}$ is the Gauss-Newton Hessian, and these derivatives are evaluated at $\basalfriction_k$.
 %The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
 We use $H_\text{gn}$ for the first five iterations, and $H$ for all subsequent iterations.
-The method terminates when $||g_k||/||g_0|| < 10^{-8}$.
-The systems \eqref{eq:newton_system} are solved inexactly using an inner PCG iteration, which terminates when the norm of the residual for the linear system is less than $\min \left(0.5, \sqrt{||g_k||}\right)||g_k||$.
-The PCG method only requires applying $H$ (or $H_\text{gn}$) to vectors.
+The Newton iteration terminates when $||g_k||/||g_0|| < 10^{-8}$.
+Systems \eqref{eq:newton_system} are solved inexactly using an inner PCG iteration, which terminates when the norm of the residual for the linear system is less than $\min \left(0.5, \sqrt{||g_k||}\right)||g_k||$.
+PCG only requires applying $H$ (or $H_\text{gn}$) to vectors.
 Each application of $H$ (or $H_\text{gn}$) to a vector requires solving one incremental forward and one incremental adjoint Stokes PDE of the form \eqref{Stokeseqn}, but with different right hand sides.
@@
 We build H-matrix approximations of the data-misfit Gauss-Newton Hessian (the term in $H_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 impulse responses batches.
+These approximations are labeled by are denoted by PSF (1), PSF (5), and PSF (25), respectively.
 The resulting H-matrices, which may be indefinite due to approximation error, are modified to be symmetric positive definite via the procedure described in Appendix \ref{sec:make_spd}, with $l=2$.
@@
 where $\bm{K}$ is a finite element stiffness matrix with appropriate boundary conditions, and $\bm{M}$ is the finite element mass matrix.
-These matrices are sparse, so we form H-matrix representations of these matrices using standard sparse H-matrix techniques.
-Then we invert $\bm{M}^{-1}$, and multiply these matrices together using H-matrix methods.
+These matrices are sparse, so we form H-matrix representations of these matrices and combine them using standard sparse H-matrix techniques.
 We factor the overall Gauss-Newton Hessian approximation, denoted $\widetilde{H}$, using fast H-matrix methods, then and use the factorization as a preconditioner \cite{HLIBPRO}.
-The approximations for 1, 5, and 25 impulse response batches are denoted by PSF (1), PSF (5), and PSF (25), respectively.
 
@@
 	}
-	\caption{Convergence history for solving the Stokes inverse problem using the preconditioned inexact Newton-PCG method.
+	\caption{Convergence history for solving the Stokes inverse problem using preconditioned inexact Newton PCG.
 Preconditioners shown are our method with five impulse response batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
-The columns titled \#CG show the number of conjugate gradient iterations used to solve the Newton system at $\basalfriction_k$, for the purpose of computing a search direction towards $\basalfriction_{k+1}$.
-The columns titled $||g||$ show the $l^2$ norm of the gradient at $\basalfriction_k$.
-The columns titled \#Stokes show the total number of Stokes PDE solves performed in each Newton iteration.
-This consists of Stokes solves for $u$ performed during the linesearch going from $\basalfriction_{k-1}$ to $\basalfriction_k$, plus the adjoint Stokes solve to compute the gradient at $\basalfriction_k$, plus one incremental forward and one incremental adjoint Stokes solve per PCG iteration for solving the Newton system at iteration $k$.
+Columns titled \#CG show the number of PCG iterations used to solve the Newton system for $\widehat{\basalfriction}_k$.
+Columns titled $||g||$ show the $l^2$ norm of the gradient at $\basalfriction_k$.
+Columns titled \#Stokes show the total number of Stokes PDE solves performed in each Newton iteration.
+This consists of Stokes solves for $u$ performed during the linesearch going from $\basalfriction_{k-1}$ to $\basalfriction_k$, plus one adjoint Stokes solve to compute the gradient at $\basalfriction_k$, plus one incremental forward and one incremental adjoint Stokes solve per PCG iteration for solving the Newton system.
 In the PSF (5) portion of row 3, we write $6+22$ to indicate that 6 Stokes solves were used during the standard course of the iteration, and 22 Stokes solves were used to build the PSF (5) preconditioner.
@@
 
-In Table \ref{tab:newton_convergence_table} we study the performance of our preconditioner for accelerating the solution of optimization problem \eqref{eq:deterministic_optimization_problem}.
+Table \ref{tab:newton_convergence_table} shows the performance of our preconditioner for accelerating the solution of optimization problem \eqref{eq:deterministic_optimization_problem}.
 We build the PSF (5) preconditioner in the third Gauss-Newton iteration, and re-use it for all subsequent Gauss-Newton and Newton iterations.
-No preconditioning is used in the three Gauss-Newton iterations before the PSF (5) preconditioner is built.
-We compare our method with the most commonly used existing preconditioners, which are no preconditioning (denoted by NONE), and preconditioning by the regularization term in the Hessian (denoted REG).
+No preconditioning is used in the Gauss-Newton iterations before the PSF (5) preconditioner is built.
+We compare our method with the most commonly used existing preconditioners: no preconditioning (NONE), and preconditioning by the regularization term in the Hessian (REG).
 In Figure \ref{fig:newton_iterates} we show select Newton iterates using our PSF (5) preconditioner.
-Using PSF (5) reduces the total number of Stokes PDE solves by a factor of six compared to no preconditioning, and by a factor of nine compared to regularization preconditioning.
-Interestingly, we observe that no preconditioning actually outperforms regularization preconditioning here because the noise level (and hence $\gamma$) is small.
+Using PSF (5) reduces the total number of Stokes PDE solves by a factor of six compared to no preconditioning, and by a factor of nine compared to regularization preconditioning\footnote{Interestingly, we observe that no preconditioning outperforms regularization preconditioning here because the noise level (and hence $\gamma$) is small.}.
 
@@
 Next, we build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\basalfriction$.
-We use the preconditioned conjugate gradient method to solve a linear system with the Hessian as the coefficient operator and a right hand random vector with i.i.d.
+We use PCG to solve a linear system with the Hessian as the coefficient operator and a right hand vector with random i.i.d.
 entries drawn from the standard normal distribution.
@@
 
-In Figure \ref{fig:generalized_eigenvalues} we show the generalized eigenvalues $\lambda$ for the generalized eigenvalue problem
+In Figure \ref{fig:generalized_eigenvalues} we show the generalized eigenvalues for the generalized eigenvalue problem
 $
@@
 The matrix $\preconditioner$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations constructed at that $\basalfriction$, or the regularization Hessian.
-The PSF preconditioners cluster the eigenvalues of the Hessian near $1$, with more batches yielding better clustering.
+The PSF preconditioners cluster the eigenvalues of the Hessian near one, with more batches yielding better clustering.
 The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
-In Figure \ref{fig:condition_number} we show the condition number of the preconditioned Hessian $\preconditioner^{-1}H$, and preconditioned Gauss-Newton Hessian, $\preconditioner^{-1}H_\text{gn}$.
-The condition numbers are similar for both the Hessian and the Gauss-Newton Hessian.
+In Figure \ref{fig:condition_number} we show the condition numbers of $\preconditioner^{-1}H$ and $\preconditioner^{-1}H_\text{gn}$.
 The PSF preconditioners yield the the smallest condition numbers, with more batches yielding smaller condition numbers.
-Regularization preconditioning and no preconditioning yield larger condition numbers, with regularization preconditioning actually increasing the condition number as compared to no preconditioning.
+Regularization preconditioning and no preconditioning yield larger condition numbers.
+The condition numbers are similar for both $\preconditioner^{-1}H$ and $\preconditioner^{-1}H_\text{gn}$, demonstrating that the preconditioner built based on $H_\text{gn}$ is a good preconditioner for $H$.
 

### 2022-07-26 paper/localpsf.tex
created some space for the conslusion

@@
 Here $\numnbr$ is a small user defined parameter; e.g., $\numnbr=10$.
+To find the $\numnbr$ nearest sample points to $x$, we query a precomputed k-d tree \cite{Bentley75} of all sample points.
+We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{Ericson04} of the mesh cells used to discretize the problem.
 Further, let
@@
 \end{equation}
-Note that $f_i$ is well-defined, because $\phi_{x_i}\left(z_i\right)$ is defined if and only if $z_i \in \Omega$, and $\spatialvol(x_i)>0$ by our sample point picking procedure in Section \ref{sec:sample_point_selection}.
+Note that $\phi_{x_i}\left(z_i\right)$ is well-defined because $z_i \in \Omega$, and $\frac{\spatialvol(x)}{\spatialvol(x_i)}$ is well-defined because our sample point picking procedure in Section \ref{sec:sample_point_selection} ensures that $\spatialvol(x_i)>0$.
 We approximate $\Aker(y,x)$ by interpolating the (point,value) pairs
@@
 
-To find the $\numnbr$ nearest sample points to $x$, we query a precomputed k-d tree \cite{Bentley75} of all sample points.
-We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{Ericson04} of the mesh cells used to discretize the problem.
-
 Interpolation is performed using the following radial basis function scheme:
 \begin{equation}
-\Aker(y,x) \approx \widetilde{\Aker}(y,x) := \sum_{i=1}^{\numnbr} \rbfweight_i~ b\left(\|x-x_i\|\right),
-\end{equation}
-where $b_i$ are radial basis functions and $\rbfweight_i$ are weights.
-The vector of weights, $\rbfweight = (\rbfweight_1, \rbfweight_2, \dots, \rbfweight_{\numnbr})^T$, is found as the solution to the $\numnbr \times \numnbr$ linear system
-\begin{equation}
-B \rbfweight = f,
-\end{equation}
-where $B \in \mathbb{R}^{\numnbr \times \numnbr}$ has entries
-$
-B_{ij} := b\left(\|x_i - x_j\|\right),
-$
-and $f \in \mathbb{R}^\numnbr$ has entries $f_i$ from \eqref{eq:fxyxp}. We use polyharmonic spline radial basis functions, which are defined as follows:
-\begin{equation}
-b(r) := \begin{cases}
+\Aker(y,x) \approx \widetilde{\Aker}(y,x) := \sum_{i=1}^{\numnbr} \rbfweight_i~ \rbf\left(\|x-x_i\|\right),
+\end{equation}
+where $\rbfweight_i$ are weights, and $\rbf$ is a polyharmonic spline radial basis function, which is defined as follows:
+\begin{equation}
+\rbf(r) := \begin{cases}
 r^\gdim, & \gdim=1,3,5,\dots \\
@@
 \end{equation}
-for $r>0$, and $b(0):=0$.
-This yields the smoothest interpolant of the given points and values, in a certain least-squares sense \cite{Duchon77}.
+for $r>0$, and $\rbf(0):=0$. The vector of weights, $\rbfweight = (\rbfweight_1, \rbfweight_2, \dots, \rbfweight_{\numnbr})^T$, is found as the solution to the $\numnbr \times \numnbr$ linear system
+\begin{equation}
+B \rbfweight = f,
+\end{equation}
+where $B \in \mathbb{R}^{\numnbr \times \numnbr}$ has entries
+$
+B_{ij} := \rbf\left(\|x_i - x_j\|\right),
+$
+and $f \in \mathbb{R}^\numnbr$ has entries $f_i$ from \eqref{eq:fxyxp}.
+Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a certain least-squares sense \cite{Duchon77}.
 For more details on radial basis function interpolation, see \cite{Wendland04}.
-The interpolation procedure is illustrate in Figure \ref{fig:rbf_interpolation}.
+%The interpolation procedure is illustrate in Figure \ref{fig:rbf_interpolation}.
 
@@
 \end{equation}
-per \eqref{eq:varphi_eval}.
-
-Note that $z_i$ is generally not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints. Hence,
+per \eqref{eq:varphi_eval}. Note that $z_i$ is typically not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints. Hence,
 evaluating $\eta_b\left(z_i\right)$ requires determining which mesh cell contains $z_i$, then evaluating finite element basis functions on that mesh cell.
-Fortunately, the mesh cell containing $z_i$ was already determined as a side effect of querying the AABB tree of mesh cells when we checked whether $z_i \in \Omega$.
+Fortunately, the mesh cell containing $z_i$ was determined as a side effect of querying the AABB tree of mesh cells when we checked whether $z_i \in \Omega$.
 
@@
 %We perform low-rank updates to the H-matrix as the Newton iterations proceed, as per the discussion in Appendix \ref{sec:recycle_krylov}.
-The rational method for modifying $\mathbf{A}_H$ to be positive semi-definite, described in Appendix \ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the order of the rational function used.
+The rational method for modifying $\mathbf{A}_H$ to be positive semi-definite, described in Appendix \ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the rational function order.
 

### 2022-07-26 paper/localpsf.tex
wrote conclusion. Paper is now too long by about half a page

@@
 		\For{$k=1,2,\dots,\nbatch$}{
-			Choose a batch of sample points, $\pointbatch_k$, using Algorithm \ref{alg:point_choice}
+			Choose a batch of sample points, $\pointbatch_k$ (Section \ref{sec:sample_point_selection})
 
@@
 
-		(optional) Modify $\mathbf{A}_H$ to make $\mathbf{A}_H^{\text{sym}+}$ (Appendix \ref{sec:make_spd})
+		(optional) Modify $\mathbf{A}_H$ to make it symmetric positive definite (Appendix \ref{sec:make_spd})
 
@@
 We also generate a flattened version of $\Gamma_b$, denoted by $\Omega \subset \mathbb{R}^2$, by ignoring the height coordinate.
-The parameter $\basalfriction$ is dually viewed as a function on $\Gamma_b$ for the purpose of solving the Stokes equations, and as a function on $\Omega$ for the purpose of building the Hessian preconditioner and defining the regularization.
+The parameter $\basalfriction$ is dually viewed as a function on $\Gamma_b$ for the purpose of solving the Stokes equations, and as a function on $\Omega$ for the purpose of building our Hessian approximations and defining the regularization.
 %This flattening is necessary because our operator approximation method requires adding points on this surface together to get other points on the surface, and this is only well-defined in flat space.
@@
 \begin{equation*}
-	J_r(\basalfriction) := \frac{1}{2}\int_{\Omega}\left| \mathcal{L}(\basalfriction - \basalfriction_0)\right|^2\,dx,
+	J_r(\basalfriction) := \frac{1}{2}\int_{\Omega}\left| \mathcal{K}(\basalfriction - \basalfriction_0)\right|^2\,dx,
 \end{equation*}
 is a Bilaplacian regularization term.
-Specifically, $\basalfriction_0$ is the constant function $\basalfriction_0(x)=10.5$, and $\mathcal{L}$ is the inverse of the solution operator for the following elliptic PDE:
+Specifically, $\basalfriction_0$ is the constant function $\basalfriction_0(x)=10.5$, and $\mathcal{K}$ is the inverse of the solution operator for the following elliptic PDE:
 \begin{align*}
@@
 The exception is Figure \ref{fig:gamma_sweep}, where we vary $\gamma$ to study how changing $\gamma$ impacts the effectiveness of our preconditioner.
-The constant $\delta$ is chosen so that the correlation length of functions drawn from the normal distribution with covariance $\mathcal{L}^{-2}$, given by $L=\sqrt{\gamma/\delta}$, is $1/10$th the radius of the domain.
+The constant $\delta$ is chosen so that the correlation length of functions drawn from the normal distribution with covariance $\mathcal{K}^{-2}$, given by $L=\sqrt{\gamma/\delta}$, is $1/10$th the radius of the domain.
 %, which is halfway between the maximum and minimum values of $\basalfriction_\text{true}$.
@@
 $
-where $\searchdir_k$ approximately solves one of the following Newton linear systems
+where $\searchdir_k$ approximately solves one of the following Newton linear systems,
 \begin{equation}
 	\label{eq:newton_system}
-	H \searchdir_k = - g_k \qquad \text{or} \qquad H_\text{gn}, \searchdir_k = - g_k
-\end{equation}
-and the step length $\alpha_k \in \mathbb{R}$ is chosen to satisfy the Armijo conditions \cite{ARMIJO}.
+	H \searchdir_k = - g_k \qquad \text{or} \qquad H_\text{gn}, \searchdir_k = - g_k,
+\end{equation}
+and the step length $\alpha_k \in \mathbb{R}$ is chosen by line search to satisfy the Armijo conditions \cite{ARMIJO}.
 Here $g_k$ is the gradient of $J$, $H$ is the Hessian, and $H_\text{gn}$ is the Gauss-Newton Hessian, and these derivatives are evaluated at $\basalfriction_k$.
@@
 
-Some conclusions here.
-
-Although the method we present in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest.
-For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high rank approximations of an operator using a small number of matrix-vector products.
+We presented an efficient matrix-free method for approximating operators with locally supported non-negative integral kernels.
+The method only requires access to the operator via application of the operator to a small number of vectors; fast access to entries of the operator's matrix representation is not required.
+The idea of the method is to compute batches of impulse responses by applying the operator to Dirac combs of scattered point sources, then interpolate these known impulse responses to approximate unknown entries of the operator's integral kernel.
+The interpolation is based on a new principle we call ``local mean displacement invariance,'' which generalizes and improves upon classical local translation invariance.
+The ability to quickly approximate arbitrary integral kernel entries permits us to form a H-matrix approximation of the operator.
+Fast H-matrix arithmetic is then used to perform further linear algebra operations that cannot be performed easily with the original operator, such as matrix factorization or inversion.
+The supports of the impulse responses are estimated to be contained in ellipsoids, which are determined a-priori via a new procedure we presented that involves applying the operator to a small number of polynomial functions.
+Point source locations for the impulse response batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulse responses per batch as possible, while ensuring that the supports of the corresponding ellipsoids do not overlap.
+We applied the method to approximate the Gauss-Newton Hessian in a Stokes ice sheet inverse problem, and saw that the approximation substantially outperforms existing Hessian approximation methods.
+Although the method we presented in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest.
+For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high rank approximations of an operator using a small number of operator applications.
 

### 2022-07-27 paper/localpsf.tex
minor change to make sure it reads f^{n_k} in the radial basis function interpolation section

@@
 $
-and $f \in \mathbb{R}^\numnbr$ has entries $f_i$ from \eqref{eq:fxyxp}.
+and $f \in \mathbb{R}^{\numnbr}$ has entries $f_i$ from \eqref{eq:fxyxp}.
 Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a certain least-squares sense \cite{Duchon77}.

### 2022-07-27 paper/localpsf.tex
fixing another small typo
two more miniscule changes

@@
 We build H-matrix approximations of the data-misfit Gauss-Newton Hessian (the term in $H_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 impulse responses batches.
-These approximations are labeled by are denoted by PSF (1), PSF (5), and PSF (25), respectively.
+These approximations are denoted by PSF (1), PSF (5), and PSF (25), respectively.
 The resulting H-matrices, which may be indefinite due to approximation error, are modified to be symmetric positive definite via the procedure described in Appendix \ref{sec:make_spd}, with $l=2$.
@@
 Our results show that our preconditioners are also good preconditioners for $H$.
-The H-matrix approximation of data misfit Gauss-Newton Hessian is added to a H-matrix approximation of the Hessian of the regularization term to form a H-matrix approximation of $H_\text{gn}$.
+The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to a H-matrix approximation of the Hessian of the regularization term to form a H-matrix approximation of $H_\text{gn}$.
 Straightforward calculation shows that the discretized Hessian of the regularization term is given by
@@
 The ability to quickly approximate arbitrary integral kernel entries permits us to form a H-matrix approximation of the operator.
-Fast H-matrix arithmetic is then used to perform further linear algebra operations that cannot be performed easily with the original operator, such as matrix factorization or inversion.
+Fast H-matrix arithmetic is then used to perform further linear algebra operations that cannot be performed easily with the original operator, such as matrix factorization and inversion.
 The supports of the impulse responses are estimated to be contained in ellipsoids, which are determined a-priori via a new procedure we presented that involves applying the operator to a small number of polynomial functions.

### 2022-07-27 paper/localpsf.tex
adding a scalar descriptor word for the description of the discretization of the pressure field, in order that we do not unnecessarily distinguish this discretization from that of the parameter field
removing an unnecessary comma in between Hgn a qk in equation 6.3 for determining the search direction
removing a symbol for the Glen exponent, as it is only used once, no need to have a symbol for later reference

@@
 Here $\eta$ is the effective viscosity and $I$ is the unit matrix.
-Glen's exponent has been chosen $n=1$, which provides for a linear Stokes model.
+Glen's exponent has been chosen to be one, which provides for a linear Stokes model.
 Note that while the PDE is linear, the parameter-to-solution map, $\basalfriction \mapsto (\velocity, \pressure)$, is nonlinear.
 
-The pressure, $\pressure$, is discretized with first order continuous Galerkin finite elements defined on a mesh of tetrahedra.
+The pressure, $\pressure$, is discretized with first order scalar continuous Galerkin finite elements defined on a mesh of tetrahedra.
 The velocity, $\velocity$, is discretized with second order continuous Galerkin finite elements on the same mesh.
@@
 	\label{eq:newton_system}
-	H \searchdir_k = - g_k \qquad \text{or} \qquad H_\text{gn}, \searchdir_k = - g_k,
+	H \searchdir_k = - g_k \qquad \text{or} \qquad H_\text{gn} \searchdir_k = - g_k,
 \end{equation}

### 2022-07-27 paper/localpsf.tex
edited paper during discussion with Tucker

@@
 
-We present a fast \emph{matrix free} method for approximating operators, $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$, that have locally supported non-negative integral kernels.
+We present a fast \emph{matrix-free} method for approximating operators, $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$, that have locally supported non-negative integral kernels.
 Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, $\gdim$ is the spatial dimension, and $L^2(\Omega)'$ is the space of continuous linear functionals on $L^2(\Omega)$ (i.e., the topological dual space).
@@
 
-By ``matrix free,'' we mean that we have a black-box computational procedure through which we may apply $\Aop$ and its transpose\footnote{Recall that the transpose $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop w\right)(u)$ for all $u,w \in L^2(\Omega)$, where the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for the analogous operations with $\Aop^T$.
+By ``matrix-free,'' we mean that we have a black-box computational procedure through which we may apply $\Aop$ and its transpose\footnote{Recall that the transpose $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop w\right)(u)$ for all $u,w \in L^2(\Omega)$, where the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for the analogous operations with $\Aop^T$.
 }, $\Aop^T$, to arbitrary functions,
@@
 However, low rank approximation methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation \cite{Alger19,AlgerEtAl17}.
-High-rank approximation methods are needed.
+High rank approximation methods are needed.
 
@@
 
-The central challenge we face is that existing scalable high-rank approximation methods typically require the ability to rapidly evaluate arbitrary entries of $\Aop$'s integral kernel, and we cannot do this because $\Aop$ is only accessible matrix-free.
+The central challenge we face is that existing scalable high rank approximation methods typically require the ability to rapidly evaluate arbitrary entries of $\Aop$'s integral kernel, and we cannot do this because $\Aop$ is only accessible matrix-free.
 We address this challenge using \emph{impulse response interpolation}.
@@
 Picking the batches of points requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
-The idea of estimating the supports of the functions $\phi_x$ a-priori was inspired by techniques from resolution analysis in seismic imaging, in which the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion  \cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
+The idea of estimating the supports of the functions $\phi_x$ a-priori was inspired by techniques from resolution analysis in seismic imaging.
+There, $\Aop^T$ is applied to a random noise function, and the width of $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$  \cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
 Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Section \ref{sec:intromoments}).
@@
 
-The ability to rapidly evaluate approximations of arbitrary entries of $\Aop$'s integral kernel allows us to form a hierarchical matrix (H-matrix) \cite{Hackbusch99,} approximation of a discretized version of $\Aop$.
-H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high-rank.
+The ability to rapidly approximate entries of $\Aop$'s integral kernel allows us to form a hierarchical matrix (H-matrix) \cite{Hackbusch99,} approximation of a discretized version of $\Aop$.
+H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank.
 %(see Figure \ref{fig:hmatrix_intro} and Appendix \ref{app:h_matrix}).
-H-matrix methods permit us to perform matrix-vector products cheaply.
-Furthermore, one may use H-matrix methods to perform useful linear algebra operations involving the H-matrix that cannot easily be done using the original operator.
+H-matrix methods permit us to perform matrix-vector products cheaply, and perform other useful linear algebra operations that cannot be done easily using the original operator.
 These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion.
 %These H-matrix methods are fast and scale well to large problems. 
-The work and memory required to perform these operations for an $\fedim \times \fedim$ H-matrix is scales as $O(\fedim \log(\fedim)^a)$, where $a \in \{0,1,2,3\}$.
-That is, the cost is \emph{nearly-linear} in $N$.
+The work and memory required to perform these operations for an $\fedim \times \fedim$ H-matrix scales 
+%as $O(\fedim \log(\fedim)^a)$, where $a \in \{0,1,2,3\}$. That is, the cost is \emph{nearly-linear}
+\emph{nearly-linearly} in $N$ (i.e., $o(N^{1+\epsilon})$ for any $\epsilon > 0$).
 The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks; for more details see \cite{GrasedyckHackbusch03}.
@@
 
-Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of continuous Galerkin finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space under the $L^2$ inner product, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
-We assume that the finite element basis satisfies the Kronecker property, i.e., $\febasis_i(p_i)=1$ and $\febasis_i(p_j)=0$ if $i\neq j$.
+Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of continuous Galerkin finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space under the $L^2$ inner product, and let $\fepoint_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
+We assume that the finite element basis satisfies the Kronecker property, i.e., $\febasis_i(\fepoint_i)=1$ and $\febasis_i(\fepoint_j)=0$ if $i\neq j$.
 For $u_h \in V_h$ we write $\bm{u} \in \mathbb{R}^\nsamplepts_\massmatrix$ to denote the coefficient vector for $u_h$ with respect to the finite element basis, i.e.,
@@
 \label{eq:Akerpcmat_entries}
-\mathbf{\Aker}_{ij} := \Aker_h(p_i, p_j).
-\end{equation}
-Because of the Kronecker property of the finite element basis, we have $\boldsymbol{\delta}_{p_i} = \mathbf{e}_i$. Thus from \eqref{eq:approx_kernel_entry_81} we have
-$
-	\Aker_h(p_i,p_j) = \left(\massmatrix^{-1}\mathbf{A} \massmatrix^{-1}\right)_{ij},
+\mathbf{\Aker}_{ij} := \Aker_h(\fepoint_i, \fepoint_j).
+\end{equation}
+Because of the Kronecker property of the finite element basis, we have $\boldsymbol{\delta}_{\fepoint_i} = \mathbf{e}_i$. Thus from \eqref{eq:approx_kernel_entry_81} we have
+$
+	\Aker_h(\fepoint_i,\fepoint_j) = \left(\massmatrix^{-1}\mathbf{A} \massmatrix^{-1}\right)_{ij},
 $
@@
 Approximation \eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close.
-Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank CUR approximation \cite{CUR,NYSTROM}.
+Operators satisfying \eqref{eq:translate1} can be well-approximated via low rank CUR approximation \cite{CUR,NYSTROM}.
 However, the required rank in the low rank approximation can be large, which makes algorithms based on \eqref{eq:translate1} expensive.
@@
 The ability to rapidly evaluate approximate kernel entries $\Aker(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Aker}_H \approx \mathbf{\Aker}$, using the conventional adaptive cross H-matrix construction method (Section \ref{sec:h_matrix_construction_short}).
-In this method, one forms low-rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
+In this method, one forms low rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
 We then use H-matrix methods to convert $\boldsymbol{\Aker}_H$ into an H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$.
@@
 	\item We make sure that $\spatialvol(\candidatepoint_i)$ is not too small, by checking whether $\spatialvol(\candidatepoint_i) > \epsilon \spatialvol_\text{max}$.
-Here $\spatialvol_\text{max}$ is the largest value of $\spatialvol(\candidatepoint_j)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\epsilon=10^{-5}$).
+Here $\spatialvol_\text{max}$ is the largest value of $\spatialvol(\candidatepoint_j)$ over all points $q$ in the initial set of candidate points, and $\epsilon$ is a small threshold parameter (we use $\epsilon=10^{-5}$).
 \end{enumerate}
@@
 Here $\numnbr$ is a small user defined parameter; e.g., $\numnbr=10$.
-To find the $\numnbr$ nearest sample points to $x$, we query a precomputed k-d tree \cite{Bentley75} of all sample points.
+To find the $\numnbr$ nearest sample points to $x$, we query a precomputed kd-tree \cite{Bentley75} of all sample points.
 We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{Ericson04} of the mesh cells used to discretize the problem.
@@
 Although $\mathbf{\Aker}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Aker}_H$ only requires evaluation of $O(\hrank^2 \fedim \log \fedim)$ entries $\mathbf{\Aker}_{ij} = \widetilde{\Aker}(p_i, p_j)$, and these entries are computed via the radial basis function interpolation method described in Section \ref{sec:approximate_kernel_entries}.
-Here $\hrank$ is the rank of the highest-rank block in the H-matrix.
+Here $\hrank$ is the rank of the highest rank block in the H-matrix.
 A dense representation of $\mathbf{\Aker}$ is never formed. 
@@
 To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulses used in the RBF interpolation.
-This is done using a precomputed KD tree of sample points, and requires $O(\numnbr \log \nsamplepts)$ floating point and logical elementary operations.
+This is done using a precomputed kd-tree of sample points, and requires $O(\numnbr \log \nsamplepts)$ floating point and logical elementary operations.
 Second, one must find the mesh cells that the points $\{z_i\}_{i=1}^{\numnbr}$ reside.
@@
 Color indicates the height of the base of the ice sheet (i.e., the mountain topography).
-The radius of the domain is $10$ kilometers, the maximum height of the mountain is $2.1$ kilometers, and the average thickness of the ice sheet is 250 meters.}
+The radius of the domain is $10^4$ meters, the maximum height of the mountain is $2.1 \times 10^3$ meters, and the average thickness of the ice sheet is 250 meters.}
 	\label{fig:ice_mountain_mesh}
@@
 	&=0\,\,\,\,\,\,\text{ on }\Gamma_{b},\label{Stokeseqn:4} \\
-	\stress\normalvec+\mu \velocity &=0\,\,\,\,\,\,\text{ on }\Gamma_{l}.\label{Stokeseqn:5}	
+	\stress\normalvec+\stokesrobincoeff \velocity &=0\,\,\,\,\,\,\text{ on }\Gamma_{l}.\label{Stokeseqn:5}	
 	\end{align}
@@
 basal sliding parameter field which is defined on the $2D$ surface $\Gamma_b$.
-The quantity $\bodyforce$ is the body force density due to gravity, $\normalvec$ is the outward normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary.
+The quantity $\bodyforce$ is the body force density due to gravity, $\stokesrobincoeff=10^6$ is a Robin boundary condition constant,  $\normalvec$ is the outward normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary.
 We employ Glen's flow law \cite{Glen1955}, $\stress= 2\eta
@@
 \begin{align*}
--\gamma\Delta u +\delta\,u&=s\,\,\,\,\text{ in }\Omega,\\
-\gamma \nabla u\cdot\normalvec+\mu\,u&=0\,\,\,\text{ on }\partial\Omega.
+-\gamma\Delta u +\delta\,u&=\regrhs\,\,\,\,\text{ in }\Omega,\\
+\gamma \nabla u\cdot\normalvec+\regrobincoeff\,u&=0\,\,\,\text{ on }\partial\Omega.
 \end{align*}
 Recall that $\Omega$ is the 2D flattened version of the basal surface $\Gamma_b$.
+Here $\regrhs$ is a generic forcing term, and $\regrobincoeff=\sqrt{\delta \gamma}/1.42$ is a Robin boundary condition coefficient \cite{WhittleMatern}, and these quantities are different from $\bodyforce$ and $\stokesrobincoeff$ in the Stokes equation, \eqref{Stokeseqn}.
 In all results except one, $\gamma$ is chosen so that the Morozov discrepancy principle is satisfied, i.e.,
@@
 We solve the inverse problem with an inexact Newton preconditioned conjugate gradient (PCG) scheme \cite{NOCEDAWRIGHT}, using $\basalfriction_0$ as the initial guess.
-The parameter is updated iteratively according to the formula
-$
-	\basalfriction_{k+1} = \basalfriction_k + \alpha_k \searchdir_k,
-$
-where $\searchdir_k$ approximately solves one of the following Newton linear systems,
+The discretized parameter, $\bm{\basalfriction}$ is updated iteratively according to the formula
+$
+	\bm{\basalfriction}_{k+1} = \bm{\basalfriction}_k + \alpha_k \bm{\searchdir}_k,
+$
+where $\bm{\searchdir}_k$ approximately solves one of the following Newton linear systems,
 \begin{equation}
 	\label{eq:newton_system}
-	H \searchdir_k = - g_k \qquad \text{or} \qquad H_\text{gn} \searchdir_k = - g_k,
+	\bm{H} \bm{\searchdir}_k = - \bm{g}_k \qquad \text{or} \qquad \bm{H}_\text{gn} \bm{\searchdir}_k = - \bm{g}_k,
 \end{equation}
 and the step length $\alpha_k \in \mathbb{R}$ is chosen by line search to satisfy the Armijo conditions \cite{ARMIJO}.
-Here $g_k$ is the gradient of $J$, $H$ is the Hessian, and $H_\text{gn}$ is the Gauss-Newton Hessian, and these derivatives are evaluated at $\basalfriction_k$.
+Here $\bm{g}_k$ is the discretized gradient of $J$, $\bm{H}$ is the discretized Hessian, and $\bm{H}_\text{gn}$ is the discretized Gauss-Newton Hessian, and these derivatives are evaluated at $\bm{\basalfriction}_k$.
 %The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
-We use $H_\text{gn}$ for the first five iterations, and $H$ for all subsequent iterations.
-The Newton iteration terminates when $||g_k||/||g_0|| < 10^{-8}$.
-Systems \eqref{eq:newton_system} are solved inexactly using an inner PCG iteration, which terminates when the norm of the residual for the linear system is less than $\min \left(0.5, \sqrt{||g_k||}\right)||g_k||$.
-PCG only requires applying $H$ (or $H_\text{gn}$) to vectors.
-Each application of $H$ (or $H_\text{gn}$) to a vector requires solving one incremental forward and one incremental adjoint Stokes PDE of the form \eqref{Stokeseqn}, but with different right hand sides.
+We use $\bm{H}_\text{gn}$ for the first five iterations, and $\bm{H}$ for all subsequent iterations.
+The Newton iteration terminates when $||\bm{g}_k||/||\bm{g}_0|| < 10^{-8}$.
+Systems \eqref{eq:newton_system} are solved inexactly using an inner PCG iteration, which terminates when the norm of the residual for the linear system is less than $\min \left(0.5, \sqrt{||\bm{g}_k||}\right)||\bm{g}_k||$.
+PCG only requires applying $\bm{H}$ (or $\bm{H}_\text{gn}$) to vectors.
+Each application of $\bm{H}$ (or $\bm{H}_\text{gn}$) to a vector requires solving one incremental forward and one incremental adjoint Stokes PDE of the form \eqref{Stokeseqn}, but with different right hand sides.
 
@@
 	%\end{subfigure}
-	\caption{The log basal friction parameter $\basalfriction$ after $1$ (left), $3$ (second from left), and $4$ (third from left) Newton iterations, and the final Newton iterate (right).
+	\caption{The log basal friction parameter after $1$ (left), $3$ (second from left), and $4$ (third from left) Newton iterations, and the final Newton iterate (right).
 The PSF (5) preconditioner is constructed between Newton iterations 3 and 4.}
@@
 We use the framework described in this paper to generate Hessian preconditioners.
-We build H-matrix approximations of the data-misfit Gauss-Newton Hessian (the term in $H_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 impulse responses batches.
+We build H-matrix approximations of the data misfit Gauss-Newton Hessian (the term in $\bm{H}_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 impulse responses batches.
 These approximations are denoted by PSF (1), PSF (5), and PSF (25), respectively.
 The resulting H-matrices, which may be indefinite due to approximation error, are modified to be symmetric positive definite via the procedure described in Appendix \ref{sec:make_spd}, with $l=2$.
-We approximate $H_\text{gn}$ rather than $H$ because $H$ often has negative values in its integral kernel when one is away from the optimal point.
-Our results show that our preconditioners are also good preconditioners for $H$.
-The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to a H-matrix approximation of the Hessian of the regularization term to form a H-matrix approximation of $H_\text{gn}$.
+We approximate $\bm{H}_\text{gn}$ rather than $\bm{H}$ because $\bm{H}$ often has negative values in its integral kernel when one is away from the optimal point.
+Our results show that our preconditioners are also good preconditioners for $\bm{H}$.
+The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to a H-matrix approximation of the Hessian of the regularization term to form a H-matrix approximation of $\bm{H}_\text{gn}$.
 Straightforward calculation shows that the discretized Hessian of the regularization term is given by
@@
 These matrices are sparse, so we form H-matrix representations of these matrices and combine them using standard sparse H-matrix techniques.
-We factor the overall Gauss-Newton Hessian approximation, denoted $\widetilde{H}$, using fast H-matrix methods, then and use the factorization as a preconditioner \cite{HLIBPRO}.
+We factor the overall Gauss-Newton Hessian approximation, denoted $\bm{\preconditioner}$, using fast H-matrix methods, then and use the factorization as a preconditioner \cite{HLIBPRO}.
 
@@
 				Iter & 
-				\#CG & \#Stokes & $||g||$ & 
-				\#CG & \#Stokes & $||g||$ & 
-				\#CG & \#Stokes & $||g||$ \\
+				\#CG & \#Stokes & $||\bm{g}||$ & 
+				\#CG & \#Stokes & $||\bm{g}||$ & 
+				\#CG & \#Stokes & $||\bm{g}||$ \\
 				\verb~0~ &
@@
 Preconditioners shown are our method with five impulse response batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
-Columns titled \#CG show the number of PCG iterations used to solve the Newton system for $\widehat{\basalfriction}_k$.
-Columns titled $||g||$ show the $l^2$ norm of the gradient at $\basalfriction_k$.
+Columns titled \#CG show the number of PCG iterations used to solve the Newton system for $\bm{\widehat{\basalfriction}}_k$.
+Columns titled $||\bm{g}||$ show the $l^2$ norm of the gradient at $\bm{\basalfriction}_k$.
 Columns titled \#Stokes show the total number of Stokes PDE solves performed in each Newton iteration.
-This consists of Stokes solves for $u$ performed during the linesearch going from $\basalfriction_{k-1}$ to $\basalfriction_k$, plus one adjoint Stokes solve to compute the gradient at $\basalfriction_k$, plus one incremental forward and one incremental adjoint Stokes solve per PCG iteration for solving the Newton system.
+This consists of Stokes solves for $u$ performed during the linesearch going from $\bm{\basalfriction}_{k-1}$ to $\bm{\basalfriction}_k$, plus one adjoint Stokes solve to compute the gradient at $\bm{\basalfriction}_k$, plus one incremental forward and one incremental adjoint Stokes solve per PCG iteration for solving the Newton system.
 In the PSF (5) portion of row 3, we write $6+22$ to indicate that 6 Stokes solves were used during the standard course of the iteration, and 22 Stokes solves were used to build the PSF (5) preconditioner.
@@
 		xlabel={CG iteration},
-		ylabel={$||b-Hx_k||/||b||$},
+		ylabel={$||\bm{b}-\bm{H} \bm{x}_k||/||\bm{b}||$},
 		title={CG convergence},
@@
 	\end{subfigure}
-	\caption{Left (\ref{fig:cg_convergence}) shows the convergence history for solving $Hx=b$ using PCG, where $b$ has i.i.d. random entries drawn from the standard normal distribution. 
+	\caption{Left (\ref{fig:cg_convergence}) shows the convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
+random entries drawn from the standard normal distribution.
 %		The y-axis is $||b-Hx_k||/||b||$, where $x_k$ is the $k$th PCG iterate. 
 	Here we use $\gamma=7.3 \times 10^3$, which is determined by the Morozov discrepancy principle.
-Right (\ref{fig:gamma_sweep}) shows the number of PCG iterations required to achieve $||b-Hx_k||/||b|| < 10^{-6}$, where $x_k$ is the $k$th PCG iterate, for a range of different $\gamma$.
+Right (\ref{fig:gamma_sweep}) shows the number of PCG iterations required to achieve $||\bm{b}-\bm{H} \bm{x}_k||/||\bm{b}|| < 10^{-6}$, where $\bm{x}_k$ is the $k$th PCG iterate, for a range of different $\gamma$.
 Results in these figures are shown for our preconditioner with 1, 5, and 25 batches of impulse responses (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
-The preconditioner is constructed using $H_\text{gn}$.
-For a fair comparison, even as $\gamma$ changes, $H$ and $H_\text{gn}$ are always evaluated at the same $\basalfriction$, which is the optimal point for the inverse problem with $\gamma=7.3 \times 10^3$.}
+The preconditioner is constructed using $\bm{H}_\text{gn}$.
+For a fair comparison, even as $\gamma$ changes, $\bm{H}$ and $\bm{H}_\text{gn}$ are always evaluated at the same $\bm{\basalfriction}$, which is the optimal point for the inverse problem with $\gamma=7.3 \times 10^3$.}
 	\label{fig:krylov_convergence}
@@
 
-Next, we build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\basalfriction$.
+Next, we build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\bm{\basalfriction}$.
 We use PCG to solve a linear system with the Hessian as the coefficient operator and a right hand vector with random i.i.d.
@@
 			legend style={at={(0.95,1)},anchor=north east},
-			title={Generalized eigenvalues, $Hu_{k}=\lambda_{k}\preconditioner u_{k}$}
+			title={Generalized eigenvalues, $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$}
 			]
@@
 			\begin{tabular}{c| c c}
-				& $\kappa(\preconditioner^{-1}H)$ &	$\kappa(\preconditioner^{-1}H_{\text{gn}})$ \\
+				& $\kappa(\bm{\preconditioner}^{-1} \bm{H})$ &	$\kappa(\bm{\preconditioner}^{-1} \bm{H}_{\text{gn}})$ \\
 				\hline 
@@
 	\end{subfigure}
-	\caption{Left (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $H u_k = \lambda_k \preconditioner_k$.
-Here $H$ is the Hessian and $\preconditioner$ is the PSF approximation constructed using the Gauss-Newton Hessian, $H_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), or the regularization Hessian (REG).
-Right (\ref{fig:condition_number}): condition number, $\kappa$, for $\widetilde{H}^{-1} H$ and $\widetilde{H}^{-1} H_\text{gn}$ for these different preconditioners, and no preconditioner (NONE).
-All operators are evaluated at the $\basalfriction$ that solves the inverse problem for $\gamma=7.3 \times 10^3$, as determined by the Morozov discrepancy principle.}
+	\caption{Left (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_k = \lambda_k \bm{\preconditioner}_k$.
+Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), or the regularization Hessian (REG).
+Right (\ref{fig:condition_number}): condition number, $\kappa$, for $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$ for these different preconditioners, and no preconditioner (NONE).
+All operators are evaluated at the $\bm{\basalfriction}$ that solves the inverse problem for $\gamma=7.3 \times 10^3$, as determined by the Morozov discrepancy principle.}
 \end{figure}
@@
 $
-	H u = \lambda \preconditioner u.
-$
-Here $H$ is the Hessian evaluated at the reconstructed parameter $\basalfriction$ for $\gamma$ chosen to satisfy the Morozov discrepancy principle.
-The matrix $\preconditioner$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations constructed at that $\basalfriction$, or the regularization Hessian.
+	\bm{H} \bm{u} = \lambda \bm{\preconditioner} \bm{u}.
+$
+Here $\bm{H}$ is the Hessian evaluated at the reconstructed parameter $\bm{\basalfriction}$ for $\gamma$ chosen to satisfy the Morozov discrepancy principle.
+The matrix $\bm{\preconditioner}$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations constructed at that $\bm{\basalfriction}$, or the regularization Hessian.
 The PSF preconditioners cluster the eigenvalues of the Hessian near one, with more batches yielding better clustering.
 The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
-In Figure \ref{fig:condition_number} we show the condition numbers of $\preconditioner^{-1}H$ and $\preconditioner^{-1}H_\text{gn}$.
+In Figure \ref{fig:condition_number} we show the condition numbers of $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$.
 The PSF preconditioners yield the the smallest condition numbers, with more batches yielding smaller condition numbers.
 Regularization preconditioning and no preconditioning yield larger condition numbers.
-The condition numbers are similar for both $\preconditioner^{-1}H$ and $\preconditioner^{-1}H_\text{gn}$, demonstrating that the preconditioner built based on $H_\text{gn}$ is a good preconditioner for $H$.
+The condition numbers are similar for both $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$, demonstrating that the preconditioner built based on $\bm{H}_\text{gn}$ is a good preconditioner for $\bm{H}$.
 
@@
 Although the method we presented in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest.
-For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high rank approximations of an operator using a small number of operator applications.
+For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low rank Hessian approximation methods, because our method can form high rank approximations of an operator using a small number of operator applications.
 

### 2022-07-27 paper/localpsf.tex
changed ordering of plots/tables to reg, none, psf(1), psf(5) and psf(25)
added the reference in which (to the best of my knowledge) the 1.42 factor is obtained from, in regards to mitigating the influence/impact of imposing boundary conditions (natural for computations) on what would otherwise be a 1 to 1 correspondence with solutions of PDEs (defined on an unbounded domain) and Gaussian random fields

@@
 Recall that $\Omega$ is the 2D flattened version of the basal surface $\Gamma_b$.
-Here $\regrhs$ is a generic forcing term, and $\regrobincoeff=\sqrt{\delta \gamma}/1.42$ is a Robin boundary condition coefficient \cite{WhittleMatern}, and these quantities are different from $\bodyforce$ and $\stokesrobincoeff$ in the Stokes equation, \eqref{Stokeseqn}.
+Here $\regrhs$ is a generic forcing term, and $\regrobincoeff=\sqrt{\delta \gamma}/1.42$ is a Robin boundary condition coefficient \cite{Roininen14}, and these quantities are different from $\bodyforce$ and $\stokesrobincoeff$ in the Stokes equation, \eqref{Stokeseqn}.
 In all results except one, $\gamma$ is chosen so that the Morozov discrepancy principle is satisfied, i.e.,
@@
 		]
+		\addlegendentry{Reg}
+		\addplot[dotted, line width=0.5mm] table {cg_relres_reg.dat};
+		\addlegendentry{None}
+		\addplot[dash dot, line width=0.5mm] table {cg_relres_none.dat};
 		\addlegendentry{PSF (1)}
@@
 		\addplot[purple, line width=0.5mm] table {cg_relres_p25.dat};
-		\addlegendentry{Reg}
-		\addplot[dotted, line width=0.5mm] table {cg_relres_reg.dat};
-		\addlegendentry{None}
-		\addplot[dash dot, line width=0.5mm] table {cg_relres_none.dat};
 		\end{axis}
@@
 			]
+			\addlegendentry{REG}
+			\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues.dat};
 			\addlegendentry{PSF (1)}
@@
 			\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues.dat};
-			\addlegendentry{REG}
-			\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues.dat};
 			\end{axis}
@@
 				\hline 
+				REG        & \verb~7.02e+3~ & \verb~7.02e+3~ \\
+				NONE       & \verb~8.71e+2~ & \verb~9.22e+2~ \\
 				PSF $(1)$  & \verb~1.75e+1~ & \verb~1.76e+1~ \\
 				PSF $(5)$  & \verb~1.39e+1~ & \verb~1.39e+1~ \\
-				PSF $(25)$ & \verb~1.03e+1~ & \verb~1.06e+1~ \\
-				REG        & \verb~7.02e+3~ & \verb~7.02e+3~ \\
-				NONE       & \verb~8.71e+2~ & \verb~9.22e+2~ 
+				PSF $(25)$ & \verb~1.03e+1~ & \verb~1.06e+1~
 			\end{tabular}

### 2022-07-27 paper/localpsf.tex
small typo when stating the generalized eigenvalue problem in the caption of the plot

@@
 	\end{subfigure}
-	\caption{Left (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_k = \lambda_k \bm{\preconditioner}_k$.
+	\caption{Left (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_k = \lambda_k \bm{\preconditioner} \bm{u}_k$.
 Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), or the regularization Hessian (REG).

### 2022-07-28 paper/localpsf.tex
paper edits to abstract and intro

@@
 \begin{abstract}
-	We present a fast matrix-free method for approximating operators that have locally supported non-negative integral kernels.
-Such operators arise, for example, as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained inverse problems.
-Our method computes impulse responses of the operator at a collection of scattered points, then interpolates these known impulse responses to evaluate entries of the integral kernel at other points where the impulse response is not known.
+	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels.
+Such operators arise as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in PDE-constrained inverse problems.
+We compute impulse responses of the operator at a collection of scattered points, then interpolate translated versions of these impulse responses to evaluate entries of the integral kernel at other points where the impulse response is not known.
+The interpolation uses a new approximation we call ``local mean displacement invariance,'' which improves upon classical local translation invariance.
 %	Classical H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
 	Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources.
-By solving an ellipsoid packing problem, we choose as many points as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
+By solving an ellipsoid packing problem, we choose as many point sources as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
+Support ellipsoids for the impulse responses are estimated a-priori by applying the operator to a small number of low order polynomial functions, then post-processing the resulting functions to determine impulse response moments for all points simultaneously.
 	The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
@@
 
-We present a fast \emph{matrix-free} method for approximating operators, $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$, that have locally supported non-negative integral kernels.
+We present a fast \emph{matrix-free} point spread function (PSF) method for approximating operators, $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$, that have locally supported non-negative integral kernels.
 Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, $\gdim$ is the spatial dimension, and $L^2(\Omega)'$ is the space of continuous linear functionals on $L^2(\Omega)$ (i.e., the topological dual space).
 Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{le1997non,saad1999distributed,TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps) \cite{DTNMAP}, covariance operators in spatial statistics \cite{GeogaEtAl20,ChenStein21,LindgrenRueLindstrom11}, blurring operators in imaging \cite{DenisEtAl11,NagyOleary98}, and Hessians in distributed parameter PDE-constrained optimization and inverse problems \cite{AlgerEtAl19,IsaacEtAl15}.
-
 By ``matrix-free,'' we mean that we have a black-box computational procedure through which we may apply $\Aop$ and its transpose\footnote{Recall that the transpose $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop w\right)(u)$ for all $u,w \in L^2(\Omega)$, where the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for the analogous operations with $\Aop^T$.
@@
 \end{equation} 
-but we do not have any further access to $\Aop$.
+but we do not have further access to $\Aop$.
 In particular, we cannot easily access entries of $\Aop$'s integral kernel.
-In applications, evaluating the maps in \eqref{eq:eval_maps} may involve iteratively solving a large linear system, timestepping, or performing some other costly computational procedure.
+In applications, evaluating the maps in \eqref{eq:eval_maps} may require performing a costly computational procedure, such as solving a large linear system.
 
@@
 However, low rank approximation methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation \cite{Alger19,AlgerEtAl17}.
-High rank approximation methods are needed.
+Matrix-free high rank approximation methods are needed.
 
@@
 
-The central challenge we face is that existing scalable high rank approximation methods typically require the ability to rapidly evaluate arbitrary entries of $\Aop$'s integral kernel, and we cannot do this because $\Aop$ is only accessible matrix-free.
-We address this challenge using \emph{impulse response interpolation}.
+%Existing scalable high rank approximation methods typically require the ability to rapidly evaluate arbitrary entries of $\Aop$'s integral kernel, which we cannot do. 
+We form a high rank approximation using \emph{impulse response interpolation}.
 The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
 We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with batches of points scattered throughout the domain (see Figure \ref{fig:batches_intro}).
-Then we interpolate translated versions of these impulse responses to approximate arbitrary entries of the operator's integral kernel.
+Then we interpolate translated versions of these impulse responses to approximate entries of the operator's integral kernel.
 %This allows fast evaluation of approximate matrix entries, which are needed by high-rank approximation methods, in particular, classical H-matrix construction methods. 
-
 Picking the batches of points requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
@@
 There, $\Aop^T$ is applied to a random noise function, and the width of $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$  \cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
-Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Section \ref{sec:intromoments}).
-Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a positive integral kernel.
-
-Our impulse response interpolation procedure may be categorized as a ``point spread function'' (PSF) method that is loosely based on so-called ``product-convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
+We use polynomial functions instead of random noise functions (see Section \ref{sec:intromoments}).
+Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
+Our impulse response interpolation may be categorized as a PSF method that is loosely based on so-called ``product-convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
 PC and PSF methods have a long history dating back several decades.
 We note the following papers (among many others),  \cite{Adorf94,AlgerEtAl19,BigotEscandeWeiss19,NagyOleary98,FishEtAl96,EscandeWeiss12,EscandeWeiss15,EscandeWeiss22,ZhuLiFomelEtAl16}, in which the convolution kernels are constructed from sampling impulse responses of the operator to scattered point sources.
-For background on PC and PSF methods, we recommend the following papers which provide well-written reviews of the topic: \cite{DenisEtAl15,EscandeWeiss17,GentileCourbinMeylan13}.
-While PC approximations are based on an assumption of local translation invariance, the method we propose is based on an assumption that we call ``local mean displacement invariance'' (see Figure \ref{fig:mean_displacement_invariance}).
-This is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}).
-Furthermore, PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulse responses.
+For background on PC and PSF methods, we recommend the following papers: \cite{DenisEtAl15,EscandeWeiss17,GentileCourbinMeylan13}.
+While PC approximations are based on an assumption of local translation invariance, the method we propose is based on an assumption that we call ``local mean displacement invariance,'' which is more general than local translation invariance, and includes both CUR low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance} and Figure \ref{fig:mean_displacement_invariance}).
+%Furthermore, PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulse responses.
 We avoid this issue by temporarily excluding impulse responses from the approximation procedure whenever an attempt is made to evaluate the impulse response at a location where it is undefined.
@@
 %\end{figure}
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale=0.85]{mean_displacement_invariance.pdf}
+	\end{center}
+	\caption{$\Aop$ is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$.
+Here $\spatialmean(z)$ denotes the mean (center of mass) of $\phi_z$.}
+	\label{fig:mean_displacement_invariance}
+\end{figure}
 
@@
 \emph{nearly-linearly} in $N$ (i.e., $o(N^{1+\epsilon})$ for any $\epsilon > 0$).
-The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks; for more details see \cite{GrasedyckHackbusch03}.
+The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks (see \cite{GrasedyckHackbusch03}).
 The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
-
 %Classical H-matrix construction methods 
@@
 %require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.
-Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}, which we do not use here.
+Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}, which we do not use.
 These alternative methods have been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{AmbartsumyanEtAl20,HartlandEtAl22}. 
-Methods based on the peeling process are asymptotically scalable (typically the number of matrix vector products required scales as $O(\log N)$), but in practice the required number of matrix vector products is large.
-Improving the peeling process to reduce its cost is currently an area of active research \cite{PEELINGRESEARCH}.
+Methods based on the peeling process are asymptotically scalable,
+%(typically the number of matrix vector products required scales as $O(\log N)$), 
+but in practice the required number of matrix vector products is large. 
+%Improving the peeling process to reduce its cost is currently an area of active research \cite{PEELINGRESEARCH}.
 %This is an area of active research, so more efficient and practical matrix-free H-matrix construction methods based on the peeling process may be discovered in the future.
 %Matrix-free H-matrix construction methods are currently a topic of considerable research, so more efficient matrix-free H-matrix methods based on the peeling process may be discovered in the future.
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[scale=0.85]{mean_displacement_invariance.pdf}
-	\end{center}
-	\caption{An operator is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$.
-Here $\spatialmean(x)$ is the mean (center of mass) of $\phi_x$, and similar for $\spatialmean(x')$.}
-	\label{fig:mean_displacement_invariance}
-\end{figure}
 
@@
 
-The action of $\Aop$ may be extended to distributions if the kernel $\Aker$ is sufficiently regular.
+The action of $\Aop$ may be extended to distributions if $\Aker$ is sufficiently regular.
 We derive the extension as follows.
@@
 		\left(\Aop \genericdistribution^*\right)(w) &= \int_\Omega \int_\Omega w(y) \Aker(y,x) \genericdistribution^*(x) dx ~dy \\
-		&= \int_\Omega w(y) \int_\Omega \Aker(y,x) \genericdistribution^*(x) dx ~dy \\
-		&= \int_\Omega w(y) \genericdistribution\left(\Aker(y,\cdot)\right) dy, \label{eq:action_on_distribution}
+		&= \int_\Omega w(y) \int_\Omega \Aker(y,x) \genericdistribution^*(x) dx ~dy 
+		= \int_\Omega w(y) \genericdistribution\left(\Aker(y,\cdot)\right) dy, \label{eq:action_on_distribution}
 	\end{align}
@@
 This may be motivated by analogy to matrices.
-If $\mathbf{B}\in \mathbb{R}^{\fedim \times \fedim}$ is a matrix with $i^\text{th}$ column $\mathbf{b}_i$ and $\mathbf{f} \in \mathbb{R}^\fedim$, then
+If $\mathbf{B}\in \mathbb{R}^{\fedim \times \fedim}$ is a matrix with $i^\text{th}$ column $\mathbf{b}_i$ and $\mathbf{w} \in \mathbb{R}^\fedim$, then
 \begin{equation*}

### 2022-07-28 localpsf/tensor_experiments1.py

@@ -1,41 +1,43 @@
 import numpy as np
 
-N1 = 4
-N2 = 5
-N3 = 7
-N4 = 2
-N5 = 3
 
-r1 = 3
-r2 = 4
-r3 = 5
-r4 = 2
-r5 = 2
+####    Make random Tucker tensor A    ####
 
-U1 = np.random.randn(N1, r1)
-U2 = np.random.randn(N2, r2)
-U3 = np.random.randn(N3, r3)
-U4 = np.random.randn(N4, r4)
-U5 = np.random.randn(N5, r5)
+NN = [7,8,6,5]
+rr_tucker = [3,4,5,2]
+UU = [np.random.randn(N, r) for N, r in zip(NN, rr_tucker)]
+C = np.random.randn(*tuple(rr_tucker))
 
-C = np.random.randn(r1,r2,r3,r4,r5)
-A = np.einsum('ia,jb,kc,ld,he,abcde->ijklh', U1, U2, U3, U4, U5, C)
-
-# A = np.random.randn(N1, N2, N3, N4, N5)
+A = np.einsum('ia,jb,kc,ld,abcd->ijkl', UU[0], UU[1], UU[2], UU[3], C)
 print('A.shape=', A.shape)
 
-Q_A = np.linalg.svd(A.swapaxes(0,1).reshape((N2, N1*N3*N4*N5)))[0][:,:r2]
 
-Z = np.random.randn(N1, N1)
-ZA = np.dot(Z, A.reshape((N1, N2*N3*N4*N5))).reshape((N1, N2, N3, N4, N5))
-Q_ZA = np.linalg.svd(ZA.swapaxes(0,1).reshape((N2, N1*N3*N4*N5)))[0][:,:r2]
-# Q_ZA = np.linalg.svd(ZA.reshape((N1*N2, N3*N4*N5)))[0][:,:r1]
+####    Construct low rank basis, Q, for combined first two modes of A    ####
+
+U, ss, _ = np.linalg.svd(A.reshape((NN[0]*NN[1], NN[2]*NN[3])), full_matrices=False)
+good_inds = (ss >  np.max(ss) * 1e-10)
+Q = U[:, good_inds]
+print('Q.shape=', Q.shape)
 
 
+####    Form tensor B by contracting random square matrix into first mode of A    ####
 
-# np.einsum('')
+Z = np.random.randn(NN[0], NN[0])
+B = np.einsum('ai,ijkl->ajkl', Z, A)
 
[…]

### 2022-07-28 paper/localpsf.tex
some paper edits to shorten appendix. Almost sub 25 pages again

@@
 	\begin{center}
-		\includegraphics[scale=0.85]{mean_displacement_invariance.pdf}
+		\includegraphics[scale=0.8]{mean_displacement_invariance.pdf}
 	\end{center}
@@
 The action of $\Aop$ may be extended to distributions if $\Aker$ is sufficiently regular. 
-We derive the extension as follows.
-Suppose $\rho \in L^2(\Omega)'$, and let $\rho^* \in L^2(\Omega)$ denote the Riesz representative of $\rho$ with respect to the $L^2(\Omega)$ inner product.
-We have
+%We derive the extension as follows. 
+Let $\rho \in L^2(\Omega)'$, and let $\rho^* \in L^2(\Omega)$ denote the Riesz representative of $\rho$ with respect to the $L^2(\Omega)$ inner product. We have
 \begin{subequations}
@@
 where $\Aker(y,\cdot)$ denotes the function $x \mapsto \Aker(y,x)$.
-Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions, and instead suppose $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ is a distribution.
+Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions, and let $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ be a distribution.
 In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in \eqref{eq:extension_to_distributions} is not valid.
@@
 The function $K(s)$ may be evaluated quickly for many $s$ by pre-computing the solution to the generalized eigenvalue problem
-\begin{equation*}
+$
 	\spatialcov_1 \eigenvectormatrix = \spatialcov_2 \eigenvectormatrix \Lambda,
-\end{equation*}
+$
 where $\eigenvectormatrix \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$.
@@
 where $v := \eigenvectormatrix^T\left(\spatialmean_1 - \spatialmean_2\right)$.
-
-To check whether $E_1$ and $E_2$ intersect, we compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, form $v$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's minimization algorithm \cite{Brent71} (any fast 1 dimensional convex optimization routine may be used).
-If $K(s^*) <0$ at the minimizer $s^*$, then $E_1 \cap E_2 = \{\}$.
+To check whether $E_1$ and $E_2$ intersect, we compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, form $v$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's minimization algorithm \cite{Brent71}.
+If $K(s^*) <0$ at the minimizer, $s^*$, then $E_1 \cap E_2 = \{\}$.
 Otherwise $E_1 \cap E_2 \neq \{\}$.
@@
 If the bounding boxes intersect, the ellipsoids might intersect, or they might not intersect, so one must perform the complete ellipsoid intersection test described in Section \ref{sec:fast_ellipsoid_intersection_test}.
-
 The smallest coordinate axis aligned box that contains the ellipsoid
@@
 $
-is given as follows:
-\begin{equation*}
+is the box
+$
 	\prod_{i=1}^\gdim \left[ \spatialmean_i -\tau\sqrt{\Sigma_{ii}}, \spatialmean_i +\tau\sqrt{\Sigma_{ii}} \right],
-\end{equation*}
+$
 where $\Sigma_{ii}$ is the $i$th diagonal entry of $\Sigma$\footnote{See \url{https://math.stackexchange.com/q/3928964}}. The boxes
@@
 If $\Aop$ is also positive semi-definite, we adapt the high order rational function spectral projection method from \cite[Section 13.8]{HackbuschKress07} to flip (or approximately flip) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ to be positive.
-In detail, let $\Pi$ be the indicator function for the interval $(a,b)$,
-\begin{equation*}
-	\Pi(\lambda) = \begin{cases}
-		1, & \lambda \in (a,b), \\
-		0, & \text{otherwise}.
-	\end{cases}
-\end{equation*}
+In detail, let $\Pi$ be the indicator function for the interval $(a,b)$.
+That is, $\Pi(\lambda)=0$ for $\lambda \in (0,1)$, and $\Pi(\lambda)=0$ otherwise.
+%\begin{equation*}
+%	\Pi(\lambda) = \begin{cases}
+%		1, & \lambda \in (a,b), \\
+%		0, & \text{otherwise}.
+%	\end{cases}
+%\end{equation*}
 Using functional calculus to extend the domain of the function $\Pi$ from scalars to matrices, we define

### 2022-07-29 paper/localpsf.tex
edited conclusion. Now sub-25 pages

@@
 Our method may still perform well if these conditions are relaxed slightly.
-For example, it is fine if the support of the impulse response $\impulseresponse_x$ is not perfectly contained in a localized region, so long as the bulk of the ``mass'' of the impulse response is contained in a localized region.
+For example, it is fine if the support of $\impulseresponse_x$ is not perfectly contained in a localized region, so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
 The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$, as long as it is non-negative for the vast majority of pairs of points $(y,x)$, and as long as the negative numbers are comparatively small.
@@
 
-We presented an efficient matrix-free method for approximating operators with locally supported non-negative integral kernels.
-The method only requires access to the operator via application of the operator to a small number of vectors; fast access to entries of the operator's matrix representation is not required.
-The idea of the method is to compute batches of impulse responses by applying the operator to Dirac combs of scattered point sources, then interpolate these known impulse responses to approximate unknown entries of the operator's integral kernel.
+We presented an efficient matrix-free PSF method for approximating operators with locally supported non-negative integral kernels.
+The method only requires access to the operator via application of the operator to a small number of vectors.
+The idea of the method is to compute batches of impulse responses by applying the operator to Dirac combs of scattered point sources, then interpolate these impulse responses to approximate entries of the operator's integral kernel.
 The interpolation is based on a new principle we call ``local mean displacement invariance,'' which generalizes and improves upon classical local translation invariance.
@@
 Fast H-matrix arithmetic is then used to perform further linear algebra operations that cannot be performed easily with the original operator, such as matrix factorization and inversion.
-The supports of the impulse responses are estimated to be contained in ellipsoids, which are determined a-priori via a new procedure we presented that involves applying the operator to a small number of polynomial functions.
+The supports of the impulse responses are estimated to be contained in ellipsoids, which are determined a-priori via a new procedure that involves applying the operator to a small number of polynomial functions.
 Point source locations for the impulse response batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulse responses per batch as possible, while ensuring that the supports of the corresponding ellipsoids do not overlap.
 We applied the method to approximate the Gauss-Newton Hessian in a Stokes ice sheet inverse problem, and saw that the approximation substantially outperforms existing Hessian approximation methods.
-Although the method we presented in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest.
-For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low rank Hessian approximation methods, because our method can form high rank approximations of an operator using a small number of operator applications.
+Although the method we presented is not applicable to all Hessians, it is applicable to many Hessians of practical interest.
+For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low rank approximation, because our method can form high rank approximations of an operator using a small number of operator applies.
 

### 2022-07-29 paper/localpsf.tex
more paper edits to shorten things

@@
 \end{equation}
-Linear functionals $\rho_h \in V_h'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^\nsamplepts_{\massmatrix^{-1}}$, with entries 
-\begin{equation*}
-\boldsymbol{\rho}_i = \rho_h(\febasis_i), \qquad i=1,\dots,\nsamplepts. 
-\end{equation*}
+Linear functionals $\rho_h \in V_h'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^\nsamplepts_{\massmatrix^{-1}}$, with entries $\boldsymbol{\rho}_i = \rho_h(\febasis_i)$ for $i=1,\dots,\nsamplepts$.
+%\begin{equation*}
+%\boldsymbol{\rho}_i = \rho_h(\febasis_i), \qquad i=1,\dots,\nsamplepts. 
+%\end{equation*}
 Here $\massmatrix \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries
@@
 Additionally, it is straightforward to verify that the Riesz representation, $\rho_h^* \in V_h$, of a functional $\rho \in V_h'$ has coefficient vector
-\begin{equation*}
+$
 \boldsymbol{\rho}^* = \massmatrix^{-1} \boldsymbol{\rho}.
-\end{equation*}
+$
 Therefore, the formula for the impulse response from \eqref{eq:impulse_response_delta_action} becomes 
@@
 Broadly, our method will construct an H-matrix approximation of $\mathbf{A}$ by forming H-matrix approximations of $\boldsymbol{\Aker}$ and $\massmatrix$ (or a lumped mass version of $\massmatrix$), then multiplying these matrices with H-matrix methods per \eqref{eq:boldA}.
-The central challenge is that classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Aker}_{ij}$, but these matrix entries are not easily accessible.
+Classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Aker}_{ij}$, but these matrix entries are not easily accessible.
 The bulk of our method is therefore dedicated to forming approximations of these matrix entries that can be evaluated rapidly.
@@
 We then use H-matrix methods to convert $\boldsymbol{\Aker}_H$ into an H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$.
+Schematically the algorithm proceeds as follows:
+\begin{equation*}
+\Aop \longrightarrow \spatialvol, \spatialmean, \spatialcov \longrightarrow \underbrace{\{x_i\}_{i=1}^r}_{\text{as }S_b\text{'s}} \longrightarrow \underbrace{\{\impulseresponse_{x_i}\}_{i=1}^r}_{\text{as }\eta_b\text{'s}} \longrightarrow \boldsymbol{\Aker}_H \longrightarrow \mathbf{A}_H.
+\end{equation*}
+The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}. The cost is discussed in Section \ref{sec:overall_cost}.
 
@@
 Negative eigenvalues modified to be positive using a rational method that is described in Appendix \ref{sec:make_spd}.
-
-Schematically the overall H-matrix approximation proceeds as follows:
-\begin{equation*}
-\Aop \longrightarrow \spatialvol, \spatialmean, \spatialcov \longrightarrow \underbrace{\{x_i\}_{i=1}^r}_{\text{as }S_b\text{'s}} \longrightarrow \underbrace{\{\impulseresponse_{x_i}\}_{i=1}^r}_{\text{as }\eta_b\text{'s}} \longrightarrow \boldsymbol{\Aker}_H \longrightarrow \mathbf{A}_H.
-\end{equation*}
-The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}. The cost to construct $\mathbf{A}_H$ is discussed in Section \ref{sec:overall_cost}.
 
@@
 		\Input{Linear operator $\Aop$, parameter $\nbatch$}
-		\Output{H-matrix $\mathbf{A}_H$ (optionally, $\mathbf{A}_H^{\text{sym}+}$)}
+		\Output{H-matrix $\mathbf{A}_H$}
 

### 2022-07-29 paper/localpsf.tex
add keywords and AMS classifications to paper

@@
 \hypersetup{
-  pdftitle={Fast matrix-free approximation of locally translation invariant operators that have locally supported non-negative integral kernels, with application to Hessians in PDE-constrained inverse problems},
+  pdftitle={Efficient matrix-free point spread function approximation of operators with locally supported non-negative integral kernels, with application to Hessians in PDE constrained inverse problems},
   pdfauthor={N. Alger, N. Petra, T. Hartland, and O. Ghattas}
@@
 	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels.
-Such operators arise as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in PDE-constrained inverse problems.
-We compute impulse responses of the operator at a collection of scattered points, then interpolate translated versions of these impulse responses to evaluate entries of the integral kernel at other points where the impulse response is not known.
+Such operators arise as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in PDE constrained inverse problems.
+We compute impulse responses of the operator at a collection of scattered points, then interpolate translated versions of these impulse responses to approximate entries of the integral kernel.
 The interpolation uses a new approximation we call ``local mean displacement invariance,'' which improves upon classical local translation invariance.
@@
 By solving an ellipsoid packing problem, we choose as many point sources as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
-Support ellipsoids for the impulse responses are estimated a-priori by applying the operator to a small number of low order polynomial functions, then post-processing the resulting functions to determine impulse response moments for all points simultaneously.
+Support ellipsoids are estimated a-priori based on impulse response moments, and these moments are computed for impulse responses at all points simultaneously via a procedure that involves applying the operator to a small number of polynomial functions.
 	The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
-Fast H-matrix methods are used to perform further matrix computations including matrix-matrix addition and multiplication, matrix inversion and factorization, and matrix-vector products in nearly-linear complexity.
-	We apply the method to approximate the Hessian in a large-scale PDE-constrained inverse problem with highly informative data.
+Fast H-matrix methods are used to perform further matrix computations in nearly-linear complexity, including matrix-matrix addition and multiplication, matrix inversion and factorization, and matrix-vector products.
+	We use the method to approximate the Hessian in an ice sheet flow PDE constrained inverse problem.
 Numerical results demonstrate that our method substantially outperforms existing methods.
-We are able to approximate high rank operators using only a small number of operator applies.
+We are able to approximate high rank operators using a small number of operator applies.
 \end{abstract}
@@
 \begin{keywords}
-
+ 	data scalability, Hessian, hierarchical matrix, matrix-free, operator approximation, PDE constrained inverse problems, point spread function, product-convolution, local translation invariance, impulse response
 \end{keywords}
@@
 \begin{AMS}
-
+ 	35R30, 41A05, 41A35, 41A36, 42A85, 47A52, 47A58, 49K20, 65D12, 65F08, 65F10, 65N20, 65N21, 86A22, 86A40
 \end{AMS}
@@
 Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, $\gdim$ is the spatial dimension, and $L^2(\Omega)'$ is the space of continuous linear functionals on $L^2(\Omega)$ (i.e., the topological dual space).
-Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{le1997non,saad1999distributed,TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps) \cite{DTNMAP}, covariance operators in spatial statistics \cite{GeogaEtAl20,ChenStein21,LindgrenRueLindstrom11}, blurring operators in imaging \cite{DenisEtAl11,NagyOleary98}, and Hessians in distributed parameter PDE-constrained optimization and inverse problems \cite{AlgerEtAl19,IsaacEtAl15}.
+Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{le1997non,saad1999distributed,TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps) \cite{DTNMAP}, covariance operators in spatial statistics \cite{GeogaEtAl20,ChenStein21,LindgrenRueLindstrom11}, blurring operators in imaging \cite{DenisEtAl11,NagyOleary98}, and Hessians in distributed parameter PDE constrained optimization and inverse problems \cite{AlgerEtAl19,IsaacEtAl15}.
 By ``matrix-free,'' we mean that we have a black-box computational procedure through which we may apply $\Aop$ and its transpose\footnote{Recall that the transpose $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop w\right)(u)$ for all $u,w \in L^2(\Omega)$, where the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for the analogous operations with $\Aop^T$.
@@
 The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks (see \cite{GrasedyckHackbusch03}).
-The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
+The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
 %Classical H-matrix construction methods 

### 2022-07-29 paper/localpsf.tex
more edits. Got to just before computational cost section

@@
 For $u_h \in V_h$ we write $\bm{u} \in \mathbb{R}^\nsamplepts_\massmatrix$ to denote the coefficient vector for $u_h$ with respect to the finite element basis, i.e.,
-\begin{equation}
-\label{eq:fem_coeff_basis}
+$
+%\label{eq:fem_coeff_basis}
 u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
-\end{equation}
+$
 Linear functionals $\rho_h \in V_h'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^\nsamplepts_{\massmatrix^{-1}}$, with entries $\boldsymbol{\rho}_i = \rho_h(\febasis_i)$ for $i=1,\dots,\nsamplepts$.
@@
 Therefore, the formula for the impulse response from \eqref{eq:impulse_response_delta_action} becomes 
-\begin{equation}
-\label{eq:discrete_kernel}
+$
+%\label{eq:discrete_kernel}
 \boldsymbol{\impulseresponse}_x = \left(A_h \delta_x^*\right)^* =  \massmatrix^{-1}\mathbf{A} \massmatrix^{-1} \boldsymbol{\delta}_x,
-\end{equation}
+$
 and the $(y,x)$ kernel entry of $\Aker_h$ may be written as
-\begin{equation}
-\label{eq:approx_kernel_entry_81}
+$
+%\label{eq:approx_kernel_entry_81}
 	\Aker_h(y,x) = \boldsymbol{\delta}_y^T \boldsymbol{\impulseresponse}_x = \boldsymbol{\delta}_y^T \massmatrix^{-1}\mathbf{A} \massmatrix^{-1} \boldsymbol{\delta}_x.
-\end{equation}
+$
 Now define $\mathbf{\Aker} \in \mathbb{R}^{\fedim \times \fedim}$ to be the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
@@
 \end{equation}
-Because of the Kronecker property of the finite element basis, we have $\boldsymbol{\delta}_{\fepoint_i} = \mathbf{e}_i$. Thus from \eqref{eq:approx_kernel_entry_81} we have
+Because of the Kronecker property of the finite element basis, we have $\boldsymbol{\delta}_{\fepoint_i} = \mathbf{e}_i$. Thus, we have
 $
@@
 \end{equation}
-Broadly, our method will construct an H-matrix approximation of $\mathbf{A}$ by forming H-matrix approximations of $\boldsymbol{\Aker}$ and $\massmatrix$ (or a lumped mass version of $\massmatrix$), then multiplying these matrices with H-matrix methods per \eqref{eq:boldA}.
+Broadly, our method will construct an H-matrix approximation of $\mathbf{A}$ by forming H-matrix approximations of $\boldsymbol{\Aker}$ and $\massmatrix$ (or a lumped mass version of $\massmatrix$), then multiplying these matrices per \eqref{eq:boldA} using H-matrix methods.
 Classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Aker}_{ij}$, but these matrix entries are not easily accessible.
@@
 
-\paragraph{Lumped mass matrix} At the continuum level, the integral kernel $\Aker$ is assumed to be non-negative.
-However, when we compute impulse responses we are effectively operating with the discrete integral kernel shown in \eqref{eq:approx_kernel_entry_81}.
-The inverse mass matrices in this formula, $\massmatrix^{-1}$, will typically contain negative numbers.
+\paragraph{Lumped mass matrix} At the continuum level, $\Aker$ is assumed to be non-negative.
+However, entries of the discrete integral kernel, $\Aker_h$, involve inverse mass matrices, $\massmatrix^{-1}$, which typically contain negative numbers.
 If there are too many negative numbers, or if the negative numbers are too large, our algorithm will be less robust.
-We therefore recommend replacing the mass matrix $\massmatrix$ with a positive diagonal \emph{lumped mass} approximation.
+We therefore recommend replacing the mass matrix, $\massmatrix$, with a positive diagonal \emph{lumped mass} approximation.
 In our numerical results, we use the lumped mass matrix constructed by replacing off-diagonal entries of the mass matrix by zero.
@@
 \end{equation*}
-Hence, we may compute $\spatialvol(x)$ for all $x$ simultaneously by applying $\Aop^T$ to $C$.
+Hence, we compute $\spatialvol(x)$ for all $x$ simultaneously by applying $\Aop^T$ to $C$.
 Analogous manipulations show that $\spatialmean(x)$ and $\spatialcov(x)$ may be computed for all points $x$ simultaneously by applying $\Aop^T$ to the functions $L^i$ and $Q^{ij}$, respectively.
@@
 \end{subequations}
-for $i,j=1,\dots, \gdim$.
-Here $u/w$ denotes pointwise division of functions, $\left(u/w\right)(x) = u(x)/w(x)$, and $u\cdot w$ denotes pointwise multiplication of functions, $(u \cdot w)(x) = u(x)w(x)$.
+for $i,j=1,\dots, \gdim$. Here $u/w$ denotes pointwise division, $\left(u/w\right)(x) = u(x)/w(x)$, and $u\cdot w$ denotes pointwise multiplication, $(u \cdot w)(x) = u(x)w(x)$.
 %The resulting method for computing discretized versions of $\spatialvol$, $\spatialmean$, and $\spatialcov$ is shown in Algorithm \ref{alg:varhpi_mean_cov}.
@@
 The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as the normalized version of $\impulseresponse_x$.
-
 The quantity $\tau$ is a parameter that must be chosen appropriately.
 The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$.
-However, in Section \ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (the more ellipsoids the better), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
+However, in Section \ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids is better), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
 In practice, we find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results.
@@
 
-Section \ref{sec:intromoments} allows us to compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} uses $\spatialvol$, $\spatialmean$, and $\spatialcov$ to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$.
+We use \eqref{eq:vol_mean_var} in Section \ref{sec:intromoments} to compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ by applying $\Aop^T$ to a small number of polynomial functions.
+Then we use \eqref{eq:support_ellipsoid} to form ellipsoid shaped estimates for the support of the impulse responses $\phi_x$, \emph{without} computing them.
 This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}).
-We compute the impulse response batch $\eta_b$ by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section \ref{sec:get_impulse_response}).
-The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support of $\impulseresponse_{x_i}$ and the support of $\impulseresponse_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$ (Section \ref{sec:sample_point_selection}).
-Because these supports do not overlap, we can post-process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}).
-The process is repeated to get more batches, until a desired number of impulse responses is reached.
-
-Once the impulse response batches $\eta_b$ are computed, we approximate the integral kernel $\Aker(y,x)$ at arbitrary points $(y,x)$ by interpolation (Section \ref{sec:approximate_kernel_entries}).
+We compute one batch of impulse responses, denoted $\eta_b$, by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section \ref{sec:get_impulse_response}).
+The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support ellipsoid for $\impulseresponse_{x_i}$ and the support ellipsoid for $\impulseresponse_{x_j}$ do not overlap if $i \neq j$ (Section \ref{sec:sample_point_selection}).
+Because the supports of these impulse responses do not overlap (or do not overlap much), we can post-process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}).
+The process is repeated until a desired number of batches is reached.
+
+Once the impulse response batches $\eta_b$ are computed, we approximate the integral kernel $\Aker(y,x)$ at arbitrary points $(y,x)$ by interpolation of translated and scaled versions of the computed impulse responses (Section \ref{sec:approximate_kernel_entries}).
 The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section \ref{sec:local_mean_displacement_invariance}.
-We approximate $\Aker(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}(y - \spatialmean(x) + \spatialmean(x_i))$ for a collection of sample points $x_i$ near $x$.
+Specifically, we approximate $\Aker(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}(y - \spatialmean(x) + \spatialmean(x_i))$ for a small number of sample points $x_i$ near $x$.
 The weights are determined by radial basis function interpolation.
@@
 When $\Aop$ is symmetric positive semi-definite, $\mathbf{A}_H$ may be non-symmetric and indefinite due to errors in the approximation.
-In this case, one may (optionally) modify the H-matrix representation of $\mathbf{A}_H$ to make it symmetric positive semi-definite using the rational method described in Appendix \ref{sec:make_spd}.
+In this case, one may (optionally) modify the H-matrix representation of $\mathbf{A}_H$ to make it symmetric positive semi-definite using a rational function method that we describe in Appendix \ref{sec:make_spd}.
 
@@
 
-		Compute $V, \mu$, and $\Sigma$ (Section \ref{sec:intromoments})
+		Compute $V, \mu$, and $\Sigma$ (Equations \eqref{eq:vol_mean_var} in Section \ref{sec:intromoments})
 
@@
 Here $\numnbr$ is a small user defined parameter; e.g., $\numnbr=10$.
-To find the $\numnbr$ nearest sample points to $x$, we query a precomputed kd-tree \cite{Bentley75} of all sample points.
+We find the $\numnbr$ nearest sample points to $x$ by querying a precomputed kd-tree \cite{Bentley75} of all sample points.
 We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{Ericson04} of the mesh cells used to discretize the problem.
@@
 Note that $\phi_{x_i}\left(z_i\right)$ is well-defined because $z_i \in \Omega$, and $\frac{\spatialvol(x)}{\spatialvol(x_i)}$ is well-defined because our sample point picking procedure in Section \ref{sec:sample_point_selection} ensures that $\spatialvol(x_i)>0$.
-We approximate $\Aker(y,x)$ by interpolating the (point,value) pairs
-\begin{equation}
-\left(x_i, f_i\right), \qquad i=1,\dots,\numnbr,
-\end{equation}
-at the point $x$. The idea is that $\Aker(y,x) = f_i$ per the discussion in Section \ref{sec:local_mean_displacement_invariance}, so we approximate $\Aker(y,x)$ by interpolating the values $f_i$.
-
+Per the discussion in Section \ref{sec:local_mean_displacement_invariance}, we expect $\Aker(y,x) \approx f_i$ for $i=1,\dots,\numnbr$; the closer $x_i$ is to $x$, the better we expect the approximation to be.
+We therefore approximate $\Aker(y,x)$ by interpolating the (point,value) pairs
+$
+\left\{\left(x_i, f_i\right)\right\}_{i=1}^{\numnbr}
+$
+at the point $x$. 
+%The idea is that $\Aker(y,x) = f_i$ per the discussion in Section \ref{sec:local_mean_displacement_invariance}, so we approximate $\Aker(y,x)$ by interpolating the values $f_i$. 
 Interpolation is performed using the following radial basis function scheme:
@@
 If $z_i \in E_{x_i}$, we look up the batch index $b$ such that $x_i \in S_b$, and evaluate $f_i$ via the formula
-\begin{equation}
+$
 f_i = \spatialvol(x)\eta_b\left(z_i\right),
-\end{equation}
+$
 per \eqref{eq:varphi_eval}. Note that $z_i$ is typically not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints. Hence,
@@
 We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming H-matrix representations $\mathbf{\Aker}_H$ and $\massmatrix_H$ of $\mathbf{\Aker}$ and $\massmatrix$, respectively, then using fast H-matrix methods to multiply these matrices per \eqref{eq:boldA} to form
-\begin{equation*}
+$
 \mathbf{A}_H = \massmatrix_H \mathbf{\Aker}_H \massmatrix_H.
-\end{equation*}
+$
 %Once $\mathbf{A}_H$ is formed, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods \cite{BORMHACKBUSCHBOOK}.
@@
 We form $\mathbf{\Aker}_H$ using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{Kriemann08}.
-Although $\mathbf{\Aker}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Aker}_H$ only requires evaluation of $O(\hrank^2 \fedim \log \fedim)$ entries $\mathbf{\Aker}_{ij} = \widetilde{\Aker}(p_i, p_j)$, and these entries are computed via the radial basis function interpolation method described in Section \ref{sec:approximate_kernel_entries}.
+Although $\mathbf{\Aker}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Aker}_H$ only requires evaluation of $O(\hrank^2 \fedim \log \fedim)$ kernel entries $\mathbf{\Aker}_{ij} = \widetilde{\Aker}(\fepoint_i, \fepoint_j)$, and these entries are computed via the radial basis function interpolation method described in Section \ref{sec:approximate_kernel_entries}.
 Here $\hrank$ is the rank of the highest rank block in the H-matrix.
@@
 
-\begin{figure}[H]
+\begin{figure}
 	\begin{subfigure}{0.5\textwidth}
@@
 
-\begin{figure}[H]
+\begin{figure}
 	\begin{subfigure}{0.49\textwidth}

### 2022-07-29 paper/localpsf.tex
more edits. Got to just before computational cost section
Paper draft is good other than some minor needed reference fixes. Ready for editing pass from Noemi

@@
 \begin{equation*}
-\Aop \longrightarrow \spatialvol, \spatialmean, \spatialcov \longrightarrow \underbrace{\{x_i\}_{i=1}^r}_{\text{as }S_b\text{'s}} \longrightarrow \underbrace{\{\impulseresponse_{x_i}\}_{i=1}^r}_{\text{as }\eta_b\text{'s}} \longrightarrow \boldsymbol{\Aker}_H \longrightarrow \mathbf{A}_H.
+\Aop \longrightarrow \spatialvol, \spatialmean, \spatialcov \longrightarrow \underbrace{\{x_i\}_{i=1}^\nsamplepts}_{\text{as }S_b\text{'s}} \longrightarrow \underbrace{\{\impulseresponse_{x_i}\}_{i=1}^\nsamplepts}_{\text{as }\eta_b\text{'s}} \longrightarrow \boldsymbol{\Aker}_H \longrightarrow \mathbf{A}_H.
 \end{equation*}
@@
 This may optionally include modifying the H-matrix to make it symmetric positive semi-definite.
-In practice, (1) is the dominant cost, because (1) is the only task that requires applying $\Aop$ to vectors, and our method is targeted at applications in which applying $\Aop$ to a vector requires an expensive computational procedure such as solving a large-scale partial differential equation.
+In practice, (1) is the dominant cost, because (1) is the only task that requires applying $\Aop$ to vectors, and our method is targeted at applications in which applying $\Aop$ to a vector requires an expensive computational procedure such as solving a large-scale PDE.
 All operations that do not require applications of $\Aop$ to vectors are nearly-linear, and therefore scalable, in the size of the problem, $N$.
@@
 \paragraph{(1) Computing impulse response moments and batches} Computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ requires applying $\Aop$ to $1$, $\gdim$, and $\gdim(\gdim+1)/2$ vectors, respectively.
-Computing each impulse response batch requires applying $\Aop$ to one vector, so computing the impulse response batches $\{\eta_i\}_{i=1}^{\nbatch}$ requires $\nbatch$ operator applies.
-In total, computing the impulse response moments and batches therefore requires
+Computing each impulse response batch requires applying $\Aop$ to one vector, so computing all impulse response batches $\{\eta_i\}_{i=1}^{\nbatch}$ requires $\nbatch$ operator applies.
+In total, computing all impulse response moments and batches therefore requires
 \begin{equation*}
@@
 requires evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix.
-To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulses used in the RBF interpolation.
+To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulse responses used in the RBF interpolation.
 This is done using a precomputed kd-tree of sample points, and requires $O(\numnbr \log \nsamplepts)$ floating point and logical elementary operations.
@@
 This is done using an AABB tree of mesh cells, and requires $O(\numnbr \log \fedim)$ elementary operations.
-Third, we must evaluate finite element basis functions on each of those cells, which requires $O(\numnbr)$ elementary operations.
-Finally, we must solve a $\numnbr \times \numnbr$ linear system to perform the radial basis function interpolation, which requires $O(\numnbr^3)$ elementary operations, which typically is negligible.
+Third, one must evaluate finite element basis functions on each of those cells, which requires $O(\numnbr)$ elementary operations.
+Finally, one must solve a $\numnbr \times \numnbr$ linear system to perform the radial basis function interpolation, which requires $O(\numnbr^3)$ elementary operations, which typically is negligible.
 Therefore, building the H-matrix requires
@@
 The precise cost depends on the type of H-matrix used and the operation being performed.
-In this paper, we use one matrix-matrix addition to add the H-matrix approximation of the data misfit term in the Hessian to the regularization term in the Hessian.
-Then, we use one matrix inversion to form the inverse of the Hessian, which is used as a preconditioner for the Newton linear system.
+In the numerical results (Section \ref{sec:numerical_results}), we use one matrix-matrix addition to add the H-matrix approximation of the data misfit term in the Hessian to the regularization term in the Hessian.
+Constructing a H-matrix representation of the regularization term in the Hessian requires one matrix inversion and two matrix-matrix multiplications.
+Then, we use one matrix factorization to form a preconditioner for the Newton linear system.
 %We perform low-rank updates to the H-matrix as the Newton iterations proceed, as per the discussion in Appendix \ref{sec:recycle_krylov}.
@@
 The velocity, $\velocity$, is discretized with second order continuous Galerkin finite elements on the same mesh.
-The parameter $\basalfriction$ is discretized with first order scalar continuous Galerkin finite elements on the mesh of triangles that results from restricting the tetrahedral mesh to $\Gamma_b$.
-The basal boundary, $\Gamma_b$, is a 2D surface embedded in 3D due to the mountain topography.
+The parameter $\basalfriction$ is discretized with first order scalar continuous Galerkin finite elements on the mesh of triangles that results from restricting the tetrahedral mesh to the basal boundary, $\Gamma_b$.
+Note that $\Gamma_b$ is a 2D surface embedded in 3D due to the mountain topography.
 We also generate a flattened version of $\Gamma_b$, denoted by $\Omega \subset \mathbb{R}^2$, by ignoring the height coordinate.
@@
 \end{equation*}
-We call this term the data misfit, because it measures the mismatch between the observed data and the predicted data based on a candidate parameter $\basalfriction$.
+We call this term the data misfit, because it measures the difference between the observed data and the predicted data based on a candidate parameter $\basalfriction$.
 Here $\velocity = \velocity(\basalfriction)$ denotes the velocity field solving $\ref{Stokeseqn}$ for the given parameter $\basalfriction$.
@@
 \begin{equation*}
-	\frac{1}{2}\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS = \frac{1}{2}\int_{\Gamma_{t}}||\xi||_{2}^{2}\,dS.
+	\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS = \int_{\Gamma_{t}}||\xi||_{2}^{2}\,dS.
 \end{equation*}
@@
 The resulting H-matrices, which may be indefinite due to approximation error, are modified to be symmetric positive definite via the procedure described in Appendix \ref{sec:make_spd}, with $l=2$.
-We approximate $\bm{H}_\text{gn}$ rather than $\bm{H}$ because $\bm{H}$ often has negative values in its integral kernel when one is away from the optimal point.
+We approximate $\bm{H}_\text{gn}$ rather than $\bm{H}$ because $\bm{H}$ more often has negative values in its integral kernel, especially when one is away from the optimal point.
 Our results show that our preconditioners are also good preconditioners for $\bm{H}$.
@@
 These matrices are sparse, so we form H-matrix representations of these matrices and combine them using standard sparse H-matrix techniques.
-We factor the overall Gauss-Newton Hessian approximation, denoted $\bm{\preconditioner}$, using fast H-matrix methods, then and use the factorization as a preconditioner \cite{HLIBPRO}.
+We factor the overall Gauss-Newton Hessian approximation, denoted $\bm{\preconditioner}$, using fast H-matrix methods \cite{RONKRIEMANNEMAIL}, then and use the factorization as a preconditioner \cite{HLIBPRO}.
 
@@
 We compare our method with the most commonly used existing preconditioners: no preconditioning (NONE), and preconditioning by the regularization term in the Hessian (REG).
+Using PSF (5) reduces the total number of Stokes PDE solves by a factor of six compared to no preconditioning, and by a factor of nine compared to regularization preconditioning\footnote{Interestingly, we observe that no preconditioning outperforms regularization preconditioning here because the noise level (and hence $\gamma$) is small.}.
 In Figure \ref{fig:newton_iterates} we show select Newton iterates using our PSF (5) preconditioner.
-Using PSF (5) reduces the total number of Stokes PDE solves by a factor of six compared to no preconditioning, and by a factor of nine compared to regularization preconditioning\footnote{Interestingly, we observe that no preconditioning outperforms regularization preconditioning here because the noise level (and hence $\gamma$) is small.}.
 
@@
 The supports of the impulse responses are estimated to be contained in ellipsoids, which are determined a-priori via a new procedure that involves applying the operator to a small number of polynomial functions.
-Point source locations for the impulse response batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulse responses per batch as possible, while ensuring that the supports of the corresponding ellipsoids do not overlap.
+Point source locations for the impulse response batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulse responses per batch as possible, while ensuring that the corresponding ellipsoids do not overlap.
 We applied the method to approximate the Gauss-Newton Hessian in a Stokes ice sheet inverse problem, and saw that the approximation substantially outperforms existing Hessian approximation methods.
@@
 	E_1 :=& \{x : (x - \spatialmean_1)^T \spatialcov_1^{-1} (x - \spatialmean_1) \le \tau^2\}, \\
-	E_2 :=& \{x : (x - \spatialmean_2)^T \spatialcov_2^{-1} (x - \spatialmean_2) \le \tau^2\}, \\
+	E_2 :=& \{x : (x - \spatialmean_2)^T \spatialcov_2^{-1} (x - \spatialmean_2) \le \tau^2\},
 \end{align*}
@@
 \begin{equation*}
-	K(s) := 1 - \frac{1}{\tau^2} (\spatialmean_1 - \spatialmean_2)^T \left(\frac{1}{1-s}\spatialcov_1 + \frac{1}{s}\spatialcov_2\right)^{-1}(\spatialmean_1 - \spatialmean_2)	
+	K(s) := 1 - \frac{1}{\tau^2} (\spatialmean_1 - \spatialmean_2)^T \left(\frac{1}{1-s}\spatialcov_1 + \frac{1}{s}\spatialcov_2\right)^{-1}(\spatialmean_1 - \spatialmean_2).
 \end{equation*}

### 2022-08-01 paper/localpsf.tex
draft before taking into account Noemi's comments
remove verb font from tables
edited abstract based on Noemi's comments

@@
 
+% CITATIONS TO DO
+% - Steffen Börm, Lars Grasedyck and Wolfgang Hackbusch:
+%"Introduction to Hierarchical Matrices with Applications",
+%Engineering Analysis with Boundary Elements 27(5), pp. 405--422, 2003.
+%- Lars Grasedyck, Ronald Kriemann and Sabine Le Borne:
+%"Parallel blackbox H-LU preconditioning for elliptic boundary value problems",
+%Computing and Visualization in Sciences, Vol. 11, pp 273--291, 2007.
+%- Ronald Kriemann:
+%"H-LU Factorization on Many-Core Systems",
+%Computing and Visualization in Science, 16, pp. 105-117, 2015.
+
 \begin{document}
@@
 \begin{abstract}
-	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels.
-Such operators arise as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in PDE constrained inverse problems.
-We compute impulse responses of the operator at a collection of scattered points, then interpolate translated versions of these impulse responses to approximate entries of the integral kernel.
-The interpolation uses a new approximation we call ``local mean displacement invariance,'' which improves upon classical local translation invariance.
-%	Classical H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products.
+	We present an efficient matrix-free point spread function (PSF)-based method for approximating operators that have locally supported non-negative integral kernels.
+Such operators arise as Schur complements when solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in PDE-constrained inverse problems.
+The proposed method uses impulse responses of the underlying operator (computed at a collection of scattered points) and interpolates translated versions of these impulse responses to approximate entries of the integral kernel.
+The interpolation uses a ``local mean displacement invariance'' approximation, which improves upon classical local translation invariance.
 Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources.
@@
 Fast H-matrix methods are used to perform further matrix computations in nearly-linear complexity, including matrix-matrix addition and multiplication, matrix inversion and factorization, and matrix-vector products.
-	We use the method to approximate the Hessian in an ice sheet flow PDE constrained inverse problem.
-Numerical results demonstrate that our method substantially outperforms existing methods.
-We are able to approximate high rank operators using a small number of operator applies.
+We illustrate our approach by applying the method to approximate the Hessian in an ice sheet flow PDE constrained inverse problem.
+Numerical results demonstrate that the proposed method substantially outperforms existing methods.
+We show that high rank operators can be efficiently approximated using only a small number of operator applies.
 \end{abstract}
@@
 				\#CG & \#Stokes & $||\bm{g}||$ \\
-				\verb~0~ &
-				\verb~1~ & \verb~4~ & \verb~2.0e+7~ &
-				\verb~1~ & \verb~4~ & \verb~2.0e+7~ &
-				\verb~1~ & \verb~4~ & \verb~2.0e+7~ \\
-				\verb~1~ &
-				\verb~2~ & \verb~6~ & \verb~5.8e+6~ &
-				\verb~2~ & \verb~6~ & \verb~7.9e+6~ &
-				\verb~2~ & \verb~6~ & \verb~5.8e+6~ \\
-				\verb~2~ &
-				\verb~4~ & \verb~10~ & \verb~2.4e+6~ &
-				\verb~5~ & \verb~12~ & \verb~3.9e+6~ &
-				\verb~4~ & \verb~10~ & \verb~2.4e+6~ \\
-				\verb~3~ &
-				\verb~2~ & \verb~6+22~ & \verb~5.8e+5~ &
-				\verb~13~ & \verb~28~ & \verb~1.6e+6~ &
-				\verb~13~ & \verb~28~ & \verb~5.8e+5~ \\
-				\verb~4~ &
-				\verb~5~ & \verb~12~ & \verb~5.6e+4~ &
-				\verb~42~ & \verb~86~ & \verb~4.8e+5~ &
-				\verb~33~ & \verb~68~ & \verb~9.1e+4~ \\
-				\verb~5~ &
-				\verb~10~ & \verb~22~ & \verb~3.5e+3~ &
-				\verb~84~ & \verb~170~ & \verb~7.7e+4~ &
-				\verb~57~ & \verb~116~ & \verb~6.2e+3~ \\
-				\verb~6~ &
-				\verb~14~ & \verb~30~ & \verb~2.7e+1~ &
-				\verb~125~ & \verb~252~ & \verb~5.0e+3~ &
-				\verb~76~ & \verb~154~ & \verb~1.1e+2~ \\
-				\verb~7~ &
-				\verb~0~ & \verb~2~ & \verb~4.1e-2~ &
-				\verb~194~ & \verb~390~ & \verb~7.9e+1~ &
-				\verb~117~ & \verb~236~ & \verb~2.3e-1~ \\
-				\verb~8~ & 
-				\verb~---~ & \verb~---~ & \verb~---~ &
-				\verb~0~ & \verb~2~ & \verb~1.6e-1~ &
-				\verb~0~ & \verb~2~ & \verb~4.1e-2~ \\
+				0 &
+				1 & 4 & 2.0e+7 &
+				1 & 4 & 2.0e+7 &
+				1 & 4 & 2.0e+7 \\
+				1 &
+				2 & 6 & 5.8e+6 &
+				2 & 6 & 7.9e+6 &
+				2 & 6 & 5.8e+6 \\
+				2 &
+				4 & 10 & 2.4e+6 &
+				5 & 12 & 3.9e+6 &
+				4 & 10 & 2.4e+6 \\
+				3 &
+				2 & 6+22 & 5.8e+5 &
+				13 & 28 & 1.6e+6 &
+				13 & 28 & 5.8e+5 \\
+				4 &
+				5 & 12 & 5.6e+4 &
+				42 & 86 & 4.8e+5 &
+				33 & 68 & 9.1e+4 \\
+				5 &
+				10 & 22 & 3.5e+3 &
+				84 & 170 & 7.7e+4 &
+				57 & 116 & 6.2e+3 \\
+				6 &
+				14 & 30 & 2.7e+1 &
+				125 & 252 & 5.0e+3 &
+				76 & 154 & 1.1e+2 \\
+				7 &
+				0 & 2 & 4.1e-2 &
+				194 & 390 & 7.9e+1 &
+				117 & 236 & 2.3e-1 \\
+				8 & 
+				--- & --- & --- &
+				0 & 2 & 1.6e-1 &
+				0 & 2 & 4.1e-2 \\
 				\hline
 				Total & 
-				\verb~38~ & \verb~114~ & \verb~---~ &
-				\verb~466~ & \verb~950~ & \verb~---~ &
-				\verb~303~ & \verb~624~ & \verb~---~ \\
+				38 & 114 & --- &
+				466 & 950 & --- &
+				303 & 624 & --- \\
 			\end{tabular}
@@
 				\hline 
-				REG        & \verb~7.02e+3~ & \verb~7.02e+3~ \\
-				NONE       & \verb~8.71e+2~ & \verb~9.22e+2~ \\
-				PSF $(1)$  & \verb~1.75e+1~ & \verb~1.76e+1~ \\
-				PSF $(5)$  & \verb~1.39e+1~ & \verb~1.39e+1~ \\
-				PSF $(25)$ & \verb~1.03e+1~ & \verb~1.06e+1~
+				REG        & 7.02e+3 & 7.02e+3 \\
+				NONE       & 8.71e+2 & 9.22e+2 \\
+				PSF $(1)$  & 1.75e+1 & 1.76e+1 \\
+				PSF $(5)$  & 1.39e+1 & 1.39e+1 \\
+				PSF $(25)$ & 1.03e+1 & 1.06e+1
 			\end{tabular}

### 2022-08-01 paper/localpsf.tex
make hyphenation consistent

@@
 \begin{abstract}
-	We present an efficient matrix-free point spread function (PSF)-based method for approximating operators that have locally supported non-negative integral kernels.
-Such operators arise as Schur complements when solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in PDE-constrained inverse problems.
+	We present an efficient matrix-free point spread function (PSF) based method for approximating operators that have locally supported non-negative integral kernels.
+Such operators arise as Schur complements when solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in PDE constrained inverse problems.
 The proposed method uses impulse responses of the underlying operator (computed at a collection of scattered points) and interpolates translated versions of these impulse responses to approximate entries of the integral kernel.
@@
 The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
-Fast H-matrix methods are used to perform further matrix computations in nearly-linear complexity, including matrix-matrix addition and multiplication, matrix inversion and factorization, and matrix-vector products.
+Fast H-matrix methods are used to perform further matrix computations in nearly linear complexity, including matrix-matrix addition and multiplication, matrix inversion and factorization, and matrix-vector products.
 We illustrate our approach by applying the method to approximate the Hessian in an ice sheet flow PDE constrained inverse problem.
@@
 \begin{keywords}
- 	data scalability, Hessian, hierarchical matrix, matrix-free, operator approximation, PDE constrained inverse problems, point spread function, product-convolution, local translation invariance, impulse response
+ 	data scalability, Hessian, hierarchical matrix, matrix-free, operator approximation, PDE constrained inverse problems, point spread function, product convolution, local translation invariance, impulse response
 \end{keywords}
@@
 Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{le1997non,saad1999distributed,TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps) \cite{DTNMAP}, covariance operators in spatial statistics \cite{GeogaEtAl20,ChenStein21,LindgrenRueLindstrom11}, blurring operators in imaging \cite{DenisEtAl11,NagyOleary98}, and Hessians in distributed parameter PDE constrained optimization and inverse problems \cite{AlgerEtAl19,IsaacEtAl15}.
-By ``matrix-free,'' we mean that we have a black-box computational procedure through which we may apply $\Aop$ and its transpose\footnote{Recall that the transpose $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop w\right)(u)$ for all $u,w \in L^2(\Omega)$, where the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for the analogous operations with $\Aop^T$.
+By ``matrix-free,'' we mean that we have a black box computational procedure through which we may apply $\Aop$ and its transpose\footnote{Recall that the transpose $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop w\right)(u)$ for all $u,w \in L^2(\Omega)$, where the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for the analogous operations with $\Aop^T$.
 }, $\Aop^T$, to arbitrary functions,
@@
 Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
-Our impulse response interpolation may be categorized as a PSF method that is loosely based on so-called ``product-convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
+Our impulse response interpolation may be categorized as a PSF method that is loosely based on ``product convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
 PC and PSF methods have a long history dating back several decades.
@@
 	\end{center}
-	\caption{$\Aop$ is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$.
+	\caption{$\Aop$ is locally mean displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$.
 Here $\spatialmean(z)$ denotes the mean (center of mass) of $\phi_z$.}
@@
 %as $O(\fedim \log(\fedim)^a)$, where $a \in \{0,1,2,3\}$. That is, the cost is \emph{nearly-linear}
-\emph{nearly-linearly} in $N$ (i.e., $o(N^{1+\epsilon})$ for any $\epsilon > 0$).
+\emph{nearly linearly} in $N$ (i.e., $o(N^{1+\epsilon})$ for any $\epsilon > 0$).
 The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks (see \cite{GrasedyckHackbusch03}).
@@
 %require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.
-Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}, which we do not use.
+Recently there have been improvements in matrix-free H-matrix construction methods based on the ``peeling process'' \cite{LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}, which we do not use. 
 These alternative methods have been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{AmbartsumyanEtAl20,HartlandEtAl22}. 
@@
 In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in \eqref{eq:extension_to_distributions} is not valid.
-However, if $\Aker$ is sufficiently regular such that the function $y \mapsto \rho\left(\Aker(y,~\cdot~)\right)$ is well-defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of \eqref{eq:action_on_distribution} is well-defined.
+However, if $\Aker$ is sufficiently regular such that the function $y \mapsto \rho\left(\Aker(y,~\cdot~)\right)$ is well defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of \eqref{eq:action_on_distribution} is well defined.
 Hence, we \emph{define} the action of $\Aop$ on the distribution $\rho$ to be the right hand side of \eqref{eq:action_on_distribution}.
@@
 \begin{enumerate}
-	\item The kernel $\Aker$ is sufficiently regular such that $\phi_x$ is a well-defined function for all $x\in \Omega$.
+	\item The kernel $\Aker$ is sufficiently regular such that $\phi_x$ is a well defined function for all $x\in \Omega$.
 	\item The supports of the impulse responses $\impulseresponse_x$ are contained in localized regions.
@@
 The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to obtain $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}.
-Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
+Then one can post process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
 However, this direct approach is not feasible because we need to know $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ before we compute $\phi_x$, in order choose the point $x$.
@@
 
-\subsection{Local mean-displacement invariance}
+\subsection{Local mean displacement invariance}
 \label{sec:local_mean_displacement_invariance}
@@
 Approximation \eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close.
-Operators satisfying \eqref{eq:translate1} can be well-approximated via low rank CUR approximation \cite{CUR,NYSTROM}.
+Operators satisfying \eqref{eq:translate1} can be well approximated via low rank CUR approximation \cite{CUR,NYSTROM}.
 However, the required rank in the low rank approximation can be large, which makes algorithms based on \eqref{eq:translate1} expensive.
 
-Operators that satisfy \eqref{eq:translate2} are called ``locally translation-invariant'' because integral kernel entries $\Aker(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$.
+Operators that satisfy \eqref{eq:translate2} are called ``locally translation invariant'' because integral kernel entries $\Aker(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$.
 It is straightforward to show that if equality holds in \eqref{eq:translate2}, then $\Aop$ is a convolution operator.
-Locally translation-invariant operators act like convolutions locally, and can therefore be well-approximated by PC approximations.
+Locally translation invariant operators act like convolutions locally, and can therefore be well approximated by PC approximations.
 
@@
 But approximation \eqref{eq:translate3} can hold in situations where neither \eqref{eq:translate1} nor \eqref{eq:translate2} holds.
-For example, because the expected value commutes with affine transformations, \eqref{eq:translate3} will hold when an $\Aop$ is locally translation-invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case.
+For example, because the expected value commutes with affine transformations, \eqref{eq:translate3} will hold when an $\Aop$ is locally translation invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case.
 Additionally, \eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions.
 In contrast, \eqref{eq:translate2} does not naturally generalize to operators that map between function spaces on different domains, because the formula $y-x+x'$ requires vectors in $\Omega_2$ and $\Omega_1$ to be added together.
-We call \eqref{eq:translate3} ``local mean-displacement invariance,'' and illustrate \eqref{eq:translate3} in Figure \ref{fig:mean_displacement_invariance}.
+We call \eqref{eq:translate3} ``local mean displacement invariance,'' and illustrate \eqref{eq:translate3} in Figure \ref{fig:mean_displacement_invariance}.
 
@@
 Approximation \eqref{eq:translate4} is equivalent to \eqref{eq:translate3}, but with $\phi_x$ replaced by the its normalized version, $\phi_x/\spatialvol(x)$.
-We call \eqref{eq:translate4} ``normalized local mean-displacement invariance.''
+We call \eqref{eq:translate4} ``normalized local mean displacement invariance.''
 
@@
 The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support ellipsoid for $\impulseresponse_{x_i}$ and the support ellipsoid for $\impulseresponse_{x_j}$ do not overlap if $i \neq j$ (Section \ref{sec:sample_point_selection}).
-Because the supports of these impulse responses do not overlap (or do not overlap much), we can post-process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}).
+Because the supports of these impulse responses do not overlap (or do not overlap much), we can post process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}).
 The process is repeated until a desired number of batches is reached.
@@
 \end{equation}
-Note that $\phi_{x_i}\left(z_i\right)$ is well-defined because $z_i \in \Omega$, and $\frac{\spatialvol(x)}{\spatialvol(x_i)}$ is well-defined because our sample point picking procedure in Section \ref{sec:sample_point_selection} ensures that $\spatialvol(x_i)>0$.
+Note that $\phi_{x_i}\left(z_i\right)$ is well defined because $z_i \in \Omega$, and $\frac{\spatialvol(x)}{\spatialvol(x_i)}$ is well defined because our sample point picking procedure in Section \ref{sec:sample_point_selection} ensures that $\spatialvol(x_i)>0$.
 Per the discussion in Section \ref{sec:local_mean_displacement_invariance}, we expect $\Aker(y,x) \approx f_i$ for $i=1,\dots,\numnbr$; the closer $x_i$ is to $x$, the better we expect the approximation to be.
@@
 and $f \in \mathbb{R}^{\numnbr}$ has entries $f_i$ from \eqref{eq:fxyxp}.
-Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a certain least-squares sense \cite{Duchon77}.
+Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a certain least squares sense \cite{Duchon77}.
 For more details on radial basis function interpolation, see \cite{Wendland04}.
@@
 This may optionally include modifying the H-matrix to make it symmetric positive semi-definite.
-In practice, (1) is the dominant cost, because (1) is the only task that requires applying $\Aop$ to vectors, and our method is targeted at applications in which applying $\Aop$ to a vector requires an expensive computational procedure such as solving a large-scale PDE.
-All operations that do not require applications of $\Aop$ to vectors are nearly-linear, and therefore scalable, in the size of the problem, $N$.
+In practice, (1) is the dominant cost, because (1) is the only task that requires applying $\Aop$ to vectors, and our method is targeted at applications in which applying $\Aop$ to a vector requires an expensive computational procedure such as solving a large scale PDE.
+All operations that do not require applications of $\Aop$ to vectors are nearly linear, and therefore scalable, in the size of the problem, $N$.
 We now describe these costs in detail. For convenience, Table \ref{tab:vars} lists variable symbols and their approximate sizes.
@@
 
-\paragraph{(3) Performing linear algebra operations with the H-matrix} It is well-known that H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion, and low rank updates require
+\paragraph{(3) Performing linear algebra operations with the H-matrix} It is well known that H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion, and low rank updates require
 \begin{equation}
@@
 
-We use our method to approximate the data misfit Gauss-Newton Hessian in an inverse problem governed by a PDE which models steady-state ice sheet flow.
+We use our method to approximate the data misfit Gauss-Newton Hessian in an inverse problem governed by a PDE which models steady state ice sheet flow.
 Briefly, a sheet of ice is sliding down a mountain (see Figure \ref{fig:ice_mountain_mesh}).
@@
 		\includegraphics[scale=0.22]{meshHeight_view0_edges_withAxes.png}
-		\caption{Top-down view}
+		\caption{Top down view}
 	\end{subfigure}
@@
 Table \ref{tab:newton_convergence_table} shows the performance of our preconditioner for accelerating the solution of optimization problem \eqref{eq:deterministic_optimization_problem}.
-We build the PSF (5) preconditioner in the third Gauss-Newton iteration, and re-use it for all subsequent Gauss-Newton and Newton iterations.
+We build the PSF (5) preconditioner in the third Gauss-Newton iteration, and reuse it for all subsequent Gauss-Newton and Newton iterations.
 No preconditioning is used in the Gauss-Newton iterations before the PSF (5) preconditioner is built.
@@
 Although the method we presented is not applicable to all Hessians, it is applicable to many Hessians of practical interest.
-For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low rank approximation, because our method can form high rank approximations of an operator using a small number of operator applies.
+For these Hessians, our method offers a \emph{data scalable} alternative to conventional low rank approximation, because our method can form high rank approximations of an operator using a small number of operator applies.
 
@@
 where fast H-matrix operations are used to perform the necessary matrix-matrix subtraction and multiplication.
-Note that for any $\ratord \ge 1$, $\Pi_\ratord(\lambda)> 1/2$ for $\lambda \in (a,b)$, so $\mathbf{A}_H^{\text{sym}+}$ is positive semi definite\footnote{Some indefiniteness may be introduced by truncation error in the H-matrix operations, but the negative eigenvalues caused by this truncation error are typically extremely small.}.
+Note that for any $\ratord \ge 1$, $\Pi_\ratord(\lambda)> 1/2$ for $\lambda \in (a,b)$, so $\mathbf{A}_H^{\text{sym}+}$ is positive semi-definite\footnote{Some indefiniteness may be introduced by truncation error in the H-matrix operations, but the negative eigenvalues caused by this truncation error are typically extremely small.}.
 At the end, $\mathbf{A}_H^{\text{sym}+}$ is symmetrized once again, to correct for any small non-symmetries that are introduced by truncation error in the H-matrix operations.

### 2022-08-01 paper/localpsf.tex
wrote comments about things to do for Nick and Tucker
add and fix some references

@@
 % CITATIONS TO DO
+%
+% TASKS:
+% 1) Incorporate Ronald Kriemann's suggested citations: (Nick)
 %    - Steffen Börm, Lars Grasedyck and Wolfgang Hackbusch:
-%"Introduction to Hierarchical Matrices with Applications",
-%Engineering Analysis with Boundary Elements 27(5), pp. 405--422, 2003.
-%- Lars Grasedyck, Ronald Kriemann and Sabine Le Borne:
-%"Parallel blackbox H-LU preconditioning for elliptic boundary value problems",
-%Computing and Visualization in Sciences, Vol. 11, pp 273--291, 2007.
-%- Ronald Kriemann:
-%"H-LU Factorization on Many-Core Systems",
-%Computing and Visualization in Science, 16, pp. 105-117, 2015.
+%      "Introduction to Hierarchical Matrices with Applications",
+%       Engineering Analysis with Boundary Elements 27(5), pp. 405--422, 2003. (DONE)
+%    - Lars Grasedyck, Ronald Kriemann and Sabine Le Borne:
+%      "Parallel blackbox H-LU preconditioning for elliptic boundary value problems",
+%       Computing and Visualization in Sciences, Vol. 11, pp 273--291, 2007. (DONE)
+%    - Ronald Kriemann:
+%      "H-LU Factorization on Many-Core Systems",
+%      Computing and Visualization in Science, 16, pp. 105-117, 2015. (DONE)
+%
+% 2) Grants
+%    - Nick and Omar's grants (Omar)
+%    - Tucker and Noemi's grants (DONE)
+%
+% 3) Order citations in ascending order within brackets (Nick)
+%    e.g., [1,2,3] = good
+%          [2,3,1] = bad
+%
+% 4) Find and cite relevant papers from Noemis's group
+%    - Tucker's technical report (Tucker)
+%
+% 5) Find and cite relevant papers from Omar's group
+%    - Toby's papers (Tucker)
+%    - Noemi's 2012 stokes paper (Tucker)
+%    - Omar and Karen's recent review paper (DONE)
+%
+% 6) Look through Nick's previous PC paper for missed citations (DONE)
+
+% For citations we need:
+%    a) bibtex info
+%    b) exactly where in the paper to add the ref
+%    c) any words or sentences to add with the ref
 
@@
 In the deterministic approach to the inverse problem, one solves a regularized minimization problem to find the parameter that best fits the data.
-The objective function in the minimization problem is defined implicitly, so information about the Hessian of the objective function is only accessible via application of the Hessian to vectors \cite{HESSIANACTION}.
+The objective function in the minimization problem is defined implicitly, so information about the Hessian of the objective function is only accessible via application of the Hessian to vectors \cite{Alger19,GhattasWillcox21}.
 The most popular matrix-free Hessian approximation methods are based on low rank approximation \cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15} because there are efficient methods for constructing low rank approximations of a matrix using only matrix-vector products \cite{Cheng05,HalkoMartinssonTropp11}.
@@
 
-The ability to rapidly approximate entries of $\Aop$'s integral kernel allows us to form a hierarchical matrix (H-matrix) \cite{Hackbusch99,} approximation of a discretized version of $\Aop$.
+The ability to rapidly approximate entries of $\Aop$'s integral kernel allows us to form a hierarchical matrix (H-matrix) \cite{Hackbusch99,BormGrasedyckHackbusch03} approximation of a discretized version of $\Aop$.
 H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank.
@@
 \emph{nearly linearly} in $N$ (i.e., $o(N^{1+\epsilon})$ for any $\epsilon > 0$).
-The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks (see \cite{GrasedyckHackbusch03}).
+The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks \cite{GrasedyckHackbusch03}.
 The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
@@
 Our method uses \eqref{eq:translate4}, which is the most sophisticated of these approximations.
-
 Approximation \eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close.
-Operators satisfying \eqref{eq:translate1} can be well approximated via low rank CUR approximation \cite{CUR,NYSTROM}.
+Operators satisfying \eqref{eq:translate1} can be well approximated via low rank CUR approximation \cite{MahoneyDrineas09}.
 However, the required rank in the low rank approximation can be large, which makes algorithms based on \eqref{eq:translate1} expensive.
@@
 %Once $\mathbf{A}_H$ is formed, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods \cite{BORMHACKBUSCHBOOK}.
-We use H1 matrices in our numerical results, but any of the other H-matrix formats (such as H2, HODLR, HSS, HBS, and others \cite{HMATRIX}) could be used instead.
-For more details on H-matrices, see \cite{Hackbusch15}.
+We use H1 matrices in our numerical results, but any other H-matrix format could be used instead. For more details on H-matrices, see \cite{Hackbusch15}.
 
 We form $\mathbf{\Aker}_H$ using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{Kriemann08}.
+For details about the algorithms used for geometrical clustering, H-matrix construction, and H-matrix operations in HLIBpro, we recommend the the following papers \cite{BormGrasedyckHackbusch03,GrasedyckKriemannLeBorne08,Kriemann13}.
 Although $\mathbf{\Aker}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Aker}_H$ only requires evaluation of $O(\hrank^2 \fedim \log \fedim)$ kernel entries $\mathbf{\Aker}_{ij} = \widetilde{\Aker}(\fepoint_i, \fepoint_j)$, and these entries are computed via the radial basis function interpolation method described in Section \ref{sec:approximate_kernel_entries}.
@@
 %We describe this H-matrix construction process in detail in Appendix \ref{app:h_matrix}. 
-We form $\massmatrix_H$ using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same recursive block partitioning structure as was used for $\mathbf{\Aker}$.
+We form $\massmatrix_H$ using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same geometrical clustering block cluster tree as was used for $\mathbf{\Aker}$.
 

### 2022-08-01 paper/localpsf.tex
Paper ready for Omar's edits. Fixed references with Tucker

@@
 Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, $\gdim$ is the spatial dimension, and $L^2(\Omega)'$ is the space of continuous linear functionals on $L^2(\Omega)$ (i.e., the topological dual space).
-Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{le1997non,saad1999distributed,TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps) \cite{DTNMAP}, covariance operators in spatial statistics \cite{GeogaEtAl20,ChenStein21,LindgrenRueLindstrom11}, blurring operators in imaging \cite{DenisEtAl11,NagyOleary98}, and Hessians in distributed parameter PDE constrained optimization and inverse problems \cite{AlgerEtAl19,IsaacEtAl15}.
+Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{le1997non,saad1999distributed}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics \cite{ChenStein21,GeogaEtAl20,LindgrenRueLindstrom11}, blurring operators in imaging \cite{DenisEtAl11,NagyOleary98}, and Hessians in distributed parameter PDE constrained optimization and inverse problems \cite{AlgerEtAl19,IsaacEtAl15}.
 By ``matrix-free,'' we mean that we have a black box computational procedure through which we may apply $\Aop$ and its transpose\footnote{Recall that the transpose $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop w\right)(u)$ for all $u,w \in L^2(\Omega)$, where the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for the analogous operations with $\Aop^T$.
@@
 The most popular matrix-free Hessian approximation methods are based on low rank approximation \cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15} because there are efficient methods for constructing low rank approximations of a matrix using only matrix-vector products \cite{Cheng05,HalkoMartinssonTropp11}.
-However, low rank approximation methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation \cite{Alger19,AlgerEtAl17}.
+However, in many inverse problems of practical interest, the Hessian or other important operators are high rank \cite{AmbartsumyanEtAl20,HartlandEtAl21}.
+Low rank approximation methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation \cite{AlexanderianGloorGhattas16,Alger19,AlgerEtAl17}.
 Matrix-free high rank approximation methods are needed.
@@
 PC and PSF methods have a long history dating back several decades.
-We note the following papers (among many others),  \cite{Adorf94,AlgerEtAl19,BigotEscandeWeiss19,NagyOleary98,FishEtAl96,EscandeWeiss12,EscandeWeiss15,EscandeWeiss22,ZhuLiFomelEtAl16}, in which the convolution kernels are constructed from sampling impulse responses of the operator to scattered point sources.
+We note the following papers (among many others),  \cite{Adorf94,AlgerEtAl19,BigotEscandeWeiss19,EscandeWeiss15,EscandeWeiss22,EscandeWeiss12,FishEtAl96,NagyOleary98,ZhuLiFomelEtAl16}, in which the convolution kernels are constructed from sampling impulse responses of the operator to scattered point sources.
 For background on PC and PSF methods, we recommend the following papers: \cite{DenisEtAl15,EscandeWeiss17,GentileCourbinMeylan13}.
@@
 
-The ability to rapidly approximate entries of $\Aop$'s integral kernel allows us to form a hierarchical matrix (H-matrix) \cite{Hackbusch99,BormGrasedyckHackbusch03} approximation of a discretized version of $\Aop$.
+The ability to rapidly approximate entries of $\Aop$'s integral kernel allows us to form a hierarchical matrix (H-matrix) \cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of a discretized version of $\Aop$.
 H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank.
@@
 However, the required rank in the low rank approximation can be large, which makes algorithms based on \eqref{eq:translate1} expensive.
-
 Operators that satisfy \eqref{eq:translate2} are called ``locally translation invariant'' because integral kernel entries $\Aker(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$.
@@
 The ability to rapidly evaluate approximate kernel entries $\Aker(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Aker}_H \approx \mathbf{\Aker}$, using the conventional adaptive cross H-matrix construction method (Section \ref{sec:h_matrix_construction_short}).
-In this method, one forms low rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}.
+In this method, one forms low rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks.
 We then use H-matrix methods to convert $\boldsymbol{\Aker}_H$ into an H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$.
@@
 
-We use our method to approximate the data misfit Gauss-Newton Hessian in an inverse problem governed by a PDE which models steady state ice sheet flow.
+We use our method to approximate the data misfit Gauss-Newton Hessian in an inverse problem governed by a PDE which models steady state ice sheet flow \cite{PetraEtAl12}.
 Briefly, a sheet of ice is sliding down a mountain (see Figure \ref{fig:ice_mountain_mesh}).

### 2022-08-03 paper/localpsf.tex
minor edit to paper: appendix listed rational function order as l=1, when it is now l=2. Fixed.
fix missing references

@@
 
-We solve the inverse problem with an inexact Newton preconditioned conjugate gradient (PCG) scheme \cite{NOCEDAWRIGHT}, using $\basalfriction_0$ as the initial guess.
+We solve the inverse problem with an inexact Newton preconditioned conjugate gradient (PCG) scheme \cite[Section 7.1]{NocedalWright99}, using $\basalfriction_0$ as the initial guess.
 The discretized parameter, $\bm{\basalfriction}$ is updated iteratively according to the formula
@@
 \end{equation}
-and the step length $\alpha_k \in \mathbb{R}$ is chosen by line search to satisfy the Armijo conditions \cite{ARMIJO}.
+and the step length $\alpha_k \in \mathbb{R}$ is chosen by line search to satisfy the Armijo conditions \cite[Section 3.1]{NocedalWright99}.
 Here $\bm{g}_k$ is the discretized gradient of $J$, $\bm{H}$ is the discretized Hessian, and $\bm{H}_\text{gn}$ is the discretized Gauss-Newton Hessian, and these derivatives are evaluated at $\bm{\basalfriction}_k$.
@@
 These matrices are sparse, so we form H-matrix representations of these matrices and combine them using standard sparse H-matrix techniques.
-We factor the overall Gauss-Newton Hessian approximation, denoted $\bm{\preconditioner}$, using fast H-matrix methods \cite{RONKRIEMANNEMAIL}, then and use the factorization as a preconditioner \cite{HLIBPRO}.
+We factor the overall Gauss-Newton Hessian approximation, denoted $\bm{\preconditioner}$, using fast H-matrix methods, then and use the factorization as a preconditioner.
 
@@
 We form the $\mathbf{T}^{2^\ratord}$ via repeated squaring, $\mathbf{T}^2 = \mathbf{T} \mathbf{T}$, $\mathbf{T}^4 = \mathbf{T}^2 \mathbf{T}^2$, $\mathbf{T}^8 = \mathbf{T}^4 \mathbf{T}^4$ and so on, where each of these matrix squarings is performed using H-matrix multiplication, which is fast.
-We recommend using $\ratord=1$ or $2$ (we use $\ratord=1$ in our numerical results).
+We recommend using $\ratord=1$ or $2$.
 Larger $\ratord$ is unnecessary and may make the method unstable.

### 2022-08-21 paper/localpsf.tex
add build_stokes_matrices stuff
made some small edits after reading over the document prior to a talk at FEniCS 2022

@@
 Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{le1997non,saad1999distributed}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics \cite{ChenStein21,GeogaEtAl20,LindgrenRueLindstrom11}, blurring operators in imaging \cite{DenisEtAl11,NagyOleary98}, and Hessians in distributed parameter PDE constrained optimization and inverse problems \cite{AlgerEtAl19,IsaacEtAl15}.
-By ``matrix-free,'' we mean that we have a black box computational procedure through which we may apply $\Aop$ and its transpose\footnote{Recall that the transpose $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop w\right)(u)$ for all $u,w \in L^2(\Omega)$, where the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for the analogous operations with $\Aop^T$.
+By ``matrix-free,'' we mean that we have a black box computational procedure through which we may apply $\Aop$ and its transpose\footnote{Recall that the transpose $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for the analogous operations with $\Aop^T$.
 }, $\Aop^T$, to arbitrary functions,
@@
 \begin{align}
-\spatialmean^i(x) :=& \int_\Omega (\impulseresponse_p(y) / V(x)) y^i ~dy \label{eq:define_mean} \\
-\spatialcov^{ij}(x) :=& \int_\Omega (\impulseresponse_p(y) / V(x)) \left(y^i - \spatialmean^i(x)\right) \left(y^j - \spatialmean^j(x)\right) ~dy, \label{eq:define_cov}
+\spatialmean^i(x) :=& \int_\Omega (\impulseresponse_x(y) / V(x)) y^i ~dy \label{eq:define_mean} \\
+\spatialcov^{ij}(x) :=& \int_\Omega (\impulseresponse_x(y) / V(x)) \left(y^i - \spatialmean^i(x)\right) \left(y^j - \spatialmean^j(x)\right) ~dy, \label{eq:define_cov}
 \end{align}
@@
 But approximation \eqref{eq:translate3} can hold in situations where neither \eqref{eq:translate1} nor \eqref{eq:translate2} holds.
-For example, because the expected value commutes with affine transformations, \eqref{eq:translate3} will hold when an $\Aop$ is locally translation invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case.
+For example, because the expected value commutes with affine transformations, \eqref{eq:translate3} will hold when $\Aop$ is locally translation invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case.
 Additionally, \eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions.
@@
 These factors make the approximation more robust if $\spatialvol(x)$ varies widely.
-Approximation \eqref{eq:translate4} is equivalent to \eqref{eq:translate3}, but with $\phi_x$ replaced by the its normalized version, $\phi_x/\spatialvol(x)$.
+Approximation \eqref{eq:translate4} is equivalent to \eqref{eq:translate3}, but with $\phi_x$ replaced by its normalized version, $\phi_x/\spatialvol(x)$.
 We call \eqref{eq:translate4} ``normalized local mean displacement invariance.''
@@
 \label{eq:fxyxp}
-f_i := \frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}\left(z_i\right), \qquad i=1,\dots,\numnbr,
+f_i := \frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}\left(z_i\right), \qquad i=1,\dots,\numnbr.
 \end{equation}
@@
 At the end, $\mathbf{A}_H^{\text{sym}+}$ is symmetrized once again, to correct for any small non-symmetries that are introduced by truncation error in the H-matrix operations.
-Once constructed, we use the symmetric positive semidefinite matrix $\mathbf{A}_H^{\text{sym}+}$ in place of $\mathbf{A}_H$ in further linear algebra operations.
+Once constructed, we use the symmetric positive semi-definite matrix $\mathbf{A}_H^{\text{sym}+}$ in place of $\mathbf{A}_H$ in further linear algebra operations.
 

### 2022-08-26 paper/localpsf.tex
shaved off about a page to make space for Omar's edits

@@
 \hypersetup{
-  pdftitle={Efficient matrix-free point spread function approximation of operators with locally supported non-negative integral kernels, with application to Hessians in PDE constrained inverse problems},
+  pdftitle={Matrix-free point spread function approximation of operators with locally supported non-negative integral kernels, with application to Hessians in PDE constrained inverse problems},
   pdfauthor={N. Alger, N. Petra, T. Hartland, and O. Ghattas}
@@
 \begin{abstract}
-	We present an efficient matrix-free point spread function (PSF) based method for approximating operators that have locally supported non-negative integral kernels.
+	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels.
 Such operators arise as Schur complements when solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in PDE constrained inverse problems.
-The proposed method uses impulse responses of the underlying operator (computed at a collection of scattered points) and interpolates translated versions of these impulse responses to approximate entries of the integral kernel.
+The proposed method uses impulses of the underlying operator (computed at a collection of scattered points) and interpolates translated and scaled versions of these impulses to approximate entries of the integral kernel.
 The interpolation uses a ``local mean displacement invariance'' approximation, which improves upon classical local translation invariance.
 Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources.
-By solving an ellipsoid packing problem, we choose as many point sources as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
-Support ellipsoids are estimated a-priori based on impulse response moments, and these moments are computed for impulse responses at all points simultaneously via a procedure that involves applying the operator to a small number of polynomial functions.
+By solving an ellipsoid packing problem, we choose as many point sources as possible per batch, while ensuring that the supports of the impulses within each batch do not overlap.
+Support ellipsoids are estimated a-priori based on impulse moments, and these moments are computed for impulses at all points simultaneously via a procedure that involves applying the operator to a small number of polynomial functions.
 The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
@@
 \begin{keywords}
- 	data scalability, Hessian, hierarchical matrix, matrix-free, operator approximation, PDE constrained inverse problems, point spread function, product convolution, local translation invariance, impulse response
+ 	data scalability, Hessian, hierarchical matrix, matrix-free, operator approximation, PDE constrained inverse problems, point spread function, product convolution, local translation invariance, impulse
 \end{keywords}
@@
 
-We present a fast \emph{matrix-free} point spread function (PSF) method for approximating operators, $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$, that have locally supported non-negative integral kernels.
-Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, $\gdim$ is the spatial dimension, and $L^2(\Omega)'$ is the space of continuous linear functionals on $L^2(\Omega)$ (i.e., the topological dual space).
-Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{le1997non,saad1999distributed}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics \cite{ChenStein21,GeogaEtAl20,LindgrenRueLindstrom11}, blurring operators in imaging \cite{DenisEtAl11,NagyOleary98}, and Hessians in distributed parameter PDE constrained optimization and inverse problems \cite{AlgerEtAl19,IsaacEtAl15}.
-By ``matrix-free,'' we mean that we have a black box computational procedure through which we may apply $\Aop$ and its transpose\footnote{Recall that the transpose $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for the analogous operations with $\Aop^T$.
+We present an efficient \emph{matrix-free} point spread function (PSF) method for approximating operators, $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$, that have locally supported non-negative integral kernels.
+Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, and $L^2(\Omega)'$ is the space of real-valued continuous linear functionals on $L^2(\Omega)$.
+Such operators arise as Schur complements in Schur complement methods for solving partial differential equations (PDEs) \cite{le1997non,saad1999distributed}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics \cite{ChenStein21,GeogaEtAl20,LindgrenRueLindstrom11}, blurring operators in imaging \cite{DenisEtAl11,NagyOleary98}, and Hessians in distributed parameter PDE constrained optimization and inverse problems \cite{AlgerEtAl19,IsaacEtAl15}.
+Here ``matrix-free'' means that we may apply $\Aop$ and its transpose\footnote{Recall that $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for operations with $\Aop^T$.
 }, $\Aop^T$, to arbitrary functions,
@@
 \end{equation} 
-but we do not have further access to $\Aop$.
-In particular, we cannot easily access entries of $\Aop$'s integral kernel.
+via a black box computational procedure, 
+but we cannot easily access entries of $\Aop$'s integral kernel.
 In applications, evaluating the maps in \eqref{eq:eval_maps} may require performing a costly computational procedure, such as solving a large linear system.
 
-Our motivation is Hessian approximation in distributed parameter inverse problems governed by partial differential equations (PDEs).
-That is, inverse problems in which one seeks to reconstruct an unknown parameter field from noisy observations of a state variable that depends on the parameter implicitly through the solution of a PDE.
-In the deterministic approach to the inverse problem, one solves a regularized minimization problem to find the parameter that best fits the data.
-The objective function in the minimization problem is defined implicitly, so information about the Hessian of the objective function is only accessible via application of the Hessian to vectors \cite{Alger19,GhattasWillcox21}.
-The most popular matrix-free Hessian approximation methods are based on low rank approximation \cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15} because there are efficient methods for constructing low rank approximations of a matrix using only matrix-vector products \cite{Cheng05,HalkoMartinssonTropp11}.
-However, in many inverse problems of practical interest, the Hessian or other important operators are high rank \cite{AmbartsumyanEtAl20,HartlandEtAl21}.
-Low rank approximation methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation \cite{AlexanderianGloorGhattas16,Alger19,AlgerEtAl17}.
-Matrix-free high rank approximation methods are needed.
+Our motivation is Hessian approximation in distributed parameter inverse problems governed by PDEs. 
+%That is, inverse problems in which one seeks to reconstruct an unknown parameter field from noisy observations of a state variable that depends on the parameter implicitly through the solution of a PDE.
+The Hessian is an operator that is central to efficient methods for solving such inverse problems in both the deterministic and Bayesian settings,
+%In the deterministic approach to the inverse problem, one solves a regularized minimization problem to find the parameter that best fits the data.
+The objective function in the minimization problem is defined implicitly, so
+but information about the Hessian is only accessible via application of the Hessian to vectors \cite{Alger19,GhattasWillcox21}. 
+The most popular matrix-free Hessian approximation methods are based on low rank approximation \cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15} because there are efficient matrix-free methods for constructing low rank matrix approximations \cite{Cheng05,HalkoMartinssonTropp11}.
+However, in many inverse problems of practical interest the Hessian is high rank \cite{AmbartsumyanEtAl20,HartlandEtAl21}.
+Moreover, the numerical rank of the Hessian grows as the data in the inverse problem become more informative about the unknown parameter.
+Low rank approximation methods therefore suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation \cite{AlexanderianGloorGhattas16,Alger19,AlgerEtAl17}.
+%Matrix-free high rank approximation methods are needed.
 
@@
 	\end{subfigure}
-	\caption{Batches of normalized impulse responses for the Stokes inverse problem data misfit Gauss-Newton Hessian, described in Section \ref{sec:numerical_results}. 
+	\caption{Batches, $\eta_b$, of normalized impulses, $\impulseresponse_{x}$, for the Stokes inverse problem data misfit Gauss-Newton Hessian (Section \ref{sec:numerical_results}). 
 %	Brightness scale differs between the subfigures. 
-	Black stars indicate point source locations. Dashed gray ellipses indicate the estimated support ellipsoids. The large circle is the boundary of the domain. 
+	Black stars are point source locations. Dashed gray ellipses are estimated impulse support ellipsoids. The large circle is $\partial \Omega$. 
 	}
@@
 %Existing scalable high rank approximation methods typically require the ability to rapidly evaluate arbitrary entries of $\Aop$'s integral kernel, which we cannot do. 
-We form a high rank approximation using \emph{impulse response interpolation}.
-The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
-We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with batches of points scattered throughout the domain (see Figure \ref{fig:batches_intro}).
-Then we interpolate translated versions of these impulse responses to approximate entries of the operator's integral kernel.
+We use \emph{impulse interpolation} to form a high rank operator approximation using a small number of operator applies.
+The impulse, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
+We compute batches of impulses by applying $\Aop$ to weighted sums of delta distributions associated with batches of points scattered throughout the domain (see Figure \ref{fig:batches_intro}).
+Then we interpolate translated and scaled versions of these impulses to approximate entries of the operator's integral kernel.
 %This allows fast evaluation of approximate matrix entries, which are needed by high-rank approximation methods, in particular, classical H-matrix construction methods. 
-Picking the batches of points requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them.
-The idea of estimating the supports of the functions $\phi_x$ a-priori was inspired by techniques from resolution analysis in seismic imaging.
-There, $\Aop^T$ is applied to a random noise function, and the width of $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$  \cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
+Picking the batches of points requires us to estimate the supports of the impulses $\phi_x$ \emph{before} we compute them. 
+The idea of estimating the supports of the functions $\phi_x$ a-priori was inspired by techniques from resolution analysis in seismic imaging. There, 
+$\Aop^T$ is applied to a random noise function, and the width of $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$  \cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
 We use polynomial functions instead of random noise functions (see Section \ref{sec:intromoments}).
 Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
-Our impulse response interpolation may be categorized as a PSF method that is loosely based on ``product convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
+Our impulse interpolation may be categorized as a PSF method that is loosely based on ``product convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
 PC and PSF methods have a long history dating back several decades.
-We note the following papers (among many others),  \cite{Adorf94,AlgerEtAl19,BigotEscandeWeiss19,EscandeWeiss15,EscandeWeiss22,EscandeWeiss12,FishEtAl96,NagyOleary98,ZhuLiFomelEtAl16}, in which the convolution kernels are constructed from sampling impulse responses of the operator to scattered point sources.
+We note the following papers (among many others),  \cite{Adorf94,AlgerEtAl19,BigotEscandeWeiss19,EscandeWeiss15,EscandeWeiss22,EscandeWeiss12,FishEtAl96,NagyOleary98,ZhuLiFomelEtAl16}, in which the convolution kernels are constructed from sampling impulses of the operator to scattered point sources.
 For background on PC and PSF methods, we recommend the following papers: \cite{DenisEtAl15,EscandeWeiss17,GentileCourbinMeylan13}.
-While PC approximations are based on an assumption of local translation invariance, the method we propose is based on an assumption that we call ``local mean displacement invariance,'' which is more general than local translation invariance, and includes both CUR low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance} and Figure \ref{fig:mean_displacement_invariance}).
-%Furthermore, PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulse responses.
-We avoid this issue by temporarily excluding impulse responses from the approximation procedure whenever an attempt is made to evaluate the impulse response at a location where it is undefined.
+While PC approximations are based on an assumption of local translation invariance, the method we propose is based on an assumption that we call ``local mean displacement invariance,'' which is more general than local translation invariance, and includes both CUR low rank approximation and local translation invariance as special cases (Section \ref{sec:local_mean_displacement_invariance} and Figure \ref{fig:mean_displacement_invariance}).
+%Furthermore, PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulses.
+We avoid this issue by temporarily excluding impulses from the approximation procedure whenever an attempt is made to evaluate the impulse at a location where it is undefined.
 
@@
 	\begin{center}
-		\includegraphics[scale=0.8]{mean_displacement_invariance.pdf}
+		\includegraphics[scale=0.75]{mean_displacement_invariance.pdf}
 	\end{center}
@@
 
-Let $\Omega \subset \mathbb{R}^\gdim$ be a bounded domain (in applications typically $\gdim=1$, $2$, or $3$).
-We seek to approximate integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ that may be written in the following form:
+Let $\Omega \subset \mathbb{R}^\gdim$ be a bounded domain (typically $\gdim=1$, $2$, or $3$). We seek to approximate integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ of the form
 \begin{equation}
@@
 Here the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$.
-The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, which exists in principle but is not easily accessible to us.
-In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define the impulse responses (Section \ref{sec:impulse_response_2}).
-Then we state the conditions on $\Aop$ that our method requires (Section \ref{sec:conditions_2}).
-Finally, we detail finite element discretization (Section \ref{sec:finite_element_kernel}).
-
-\subsection{Distributions and impulse responses}
+The integral kernel, $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$, exists in principle but is not easily accessible.
+In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define impulses (Section \ref{sec:impulse_response_2}), we state the conditions on $\Aop$ that our method requires (Section \ref{sec:conditions_2}), and we detail finite element discretization (Section \ref{sec:finite_element_kernel}).
+
+\subsection{Distributions and impulses}
 \label{sec:impulse_response_2}
@@
 where $\Aker(y,\cdot)$ denotes the function $x \mapsto \Aker(y,x)$.
-Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions, and let $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ be a distribution.
+Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions and let $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ be a distribution.
 In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in \eqref{eq:extension_to_distributions} is not valid.
@@
 Hence, we \emph{define} the action of $\Aop$ on the distribution $\rho$ to be the right hand side of \eqref{eq:action_on_distribution}.
-For convenience of notation, we denote this action by ``$\Aop \genericdistribution^*$,'' even if $\rho^*$ does not exist.
-
-Let $\delta_x$ denote the delta distribution\footnote{Recall that the delta distribution $\delta_x:\mathcal{D}(\Omega)\rightarrow \mathbb{R}$ is the linear functional defined by $\delta_x(w) = w(x)$ for all $w\in \mathcal{D}(\Omega)$.} (i.e., point source, impulse) centered at the point $x \in \Omega$.
-The \emph{impulse response} of $\Aop$ associated with $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$, 
+We denote this action by ``$\Aop \genericdistribution^*$,'' even if $\rho^*$ does not exist.
+
+Let $\delta_x$ denote the delta distribution\footnote{Recall that the delta distribution $\delta_x:\mathcal{D}(\Omega)\rightarrow \mathbb{R}$ is defined by $\delta_x(w) = w(x)$ for all $w\in \mathcal{D}(\Omega)$.} (i.e., point source, impulse) centered at the point $x \in \Omega$.
+The \emph{impulse} of $\Aop$ associated with $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$, 
 \begin{equation}
@@
 that is formed by applying $\Aop$ to $\delta_x$, then taking the Riesz representation of the resulting linear functional.
-Using \eqref{eq:action_on_distribution} and the definition of the delta distribution, we see that $\phi_x$ may also be written as the function
-\begin{equation}
-\label{eq:impulse_response_defn}
-\impulseresponse_x(y) = \Aker(y, x),
-\end{equation} 
-which results from keeping $x$ fixed and viewing $\Aker(y,x)$ as a function of $y$.
+Using \eqref{eq:action_on_distribution} and the definition of the delta distribution, we see that $\phi_x$ may also be written as the function $\impulseresponse_x(y) = \Aker(y, x)$.
+%\begin{equation}
+%\label{eq:impulse_response_defn}
+%\impulseresponse_x(y) = \Aker(y, x),
+%\end{equation} 
+%which results from keeping $x$ fixed and viewing $\Aker(y,x)$ as a function of $y$.
 
@@
 \begin{enumerate}
-	\item The kernel $\Aker$ is sufficiently regular such that $\phi_x$ is a well defined function for all $x\in \Omega$.
-	\item The supports of the impulse responses $\impulseresponse_x$ are contained in localized regions.
+	\item The kernel $\Aker$ is sufficiently regular so that $\phi_x$ is well defined for all $x\in \Omega$.
+	\item The supports of the impulses $\impulseresponse_x$ are contained in localized regions.
 	\item The integral kernel is non-negative in the sense that $$\Aker(y,x) \ge 0$$ for all $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
@@
 Our method may still perform well if these conditions are relaxed slightly.
-For example, it is fine if the support of $\impulseresponse_x$ is not perfectly contained in a localized region, so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
-The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$, as long as it is non-negative for the vast majority of pairs of points $(y,x)$, and as long as the negative numbers are comparatively small.
+It is fine if the support of $\impulseresponse_x$ is not perfectly contained in a localized region, so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
+The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$, so long as it is non-negative for the vast majority of pairs of points $(y,x)$, and as long as the negative numbers are comparatively small.
 If these conditions are violated, our method will incur additional error.
 If these conditions are violated too much, our method may fail.
-Not all operators of interest (and in particular not all Hessians) satisfy these conditions, but many operators of practical interest do satisfy these conditions, and our method is highly effective for approximating these operators.
+%Not all operators of interest (and in particular not all Hessians) satisfy these conditions, but many operators of practical interest do satisfy these conditions, and our method is highly effective for approximating these operators.
 
@@
 %\end{equation*}
-Here $\massmatrix \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries
-\begin{equation*}
-\massmatrix_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx, \qquad i,j=1,\dots,\fedim,
-\end{equation*}
-the space $\mathbb{R}^\fedim_\massmatrix$ is $\mathbb{R}^\fedim$ with the inner product $(\mathbf{u},\mathbf{w})_\massmatrix := \mathbf{u}^T \massmatrix \mathbf{w}$, and $\mathbb{R}^\fedim_{\massmatrix^{-1}}$ is the analogous space with $\massmatrix^{-1}$ replacing $\massmatrix$.
+Here $\massmatrix \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries $\massmatrix_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx$ for $i,j=1,\dots,\fedim$.
+%\begin{equation*}
+%\massmatrix_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx, \qquad i,j=1,\dots,\fedim,
+%\end{equation*}
+The space $\mathbb{R}^\fedim_\massmatrix$ is $\mathbb{R}^\fedim$ with the inner product $(\mathbf{u},\mathbf{w})_\massmatrix := \mathbf{u}^T \massmatrix \mathbf{w}$, and $\mathbb{R}^\fedim_{\massmatrix^{-1}}$ is the analogous space with $\massmatrix^{-1}$ replacing $\massmatrix$.
 Direct calculation shows that $\mathbb{R}^\fedim_\massmatrix$ and $\mathbb{R}^\fedim_{\massmatrix^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively.
@@
 
-The operator $A_h:V_h \rightarrow V_h'$ can be written in integral kernel form as
-\begin{equation}
-(A_h u_h)(w_h) := \int_\Omega \int_\Omega w_h(y) \Aker_h(y,x) u_h(x) dx dy
-\end{equation}
-for some kernel $\Aker_h$ which we do not know, and which may differ from $\Aker$ due to discretization error.
+The operator $A_h:V_h \rightarrow V_h'$ can be written in integral kernel form, \eqref{eq:kernel_representation}, but with $\Aker$ replaced by a slightly different integral kernel,
+%as
+%\begin{equation}
+%(A_h u_h)(w_h) := \int_\Omega \int_\Omega w_h(y) \Aker_h(y,x) u_h(x) dx dy
+%\end{equation}
+$\Aker_h$, which we do not know, and which differs from $\Aker$ due to discretization error.
 Since the functions in $V_h$ are continuous at $x$, the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta}_x \in \mathbb{R}^\fedim_{\massmatrix^{-1}}$ with entries $\left(\boldsymbol{\delta}_x\right)_i = \febasis_i(x)$ for $i=1,\dots,\fedim$.
@@
 $
-Therefore, the formula for the impulse response from \eqref{eq:impulse_response_delta_action} becomes 
+Therefore, the formula for the impulse from \eqref{eq:impulse_response_delta_action} becomes 
 $
@@
 \paragraph{Lumped mass matrix} At the continuum level, $\Aker$ is assumed to be non-negative.
-However, entries of the discrete integral kernel, $\Aker_h$, involve inverse mass matrices, $\massmatrix^{-1}$, which typically contain negative numbers.
+However, entries of $\Aker_h$, involve inverse mass matrices, which typically contain negative numbers.
 If there are too many negative numbers, or if the negative numbers are too large, our algorithm will be less robust.
 We therefore recommend replacing the mass matrix, $\massmatrix$, with a positive diagonal \emph{lumped mass} approximation.
-In our numerical results, we use the lumped mass matrix constructed by replacing off-diagonal entries of the mass matrix by zero.
-Other more advanced mass lumping techniques may be used.
-See \cite{Hughes12} for a discussion of lumped mass matrices.
+In our numerical results, we use the lumped mass matrix constructed by replacing off-diagonal entries of the mass matrix by zero. Other mass lumping techniques may be used \cite{Hughes12}.
 
@@
 In this section we present two key innovations that our method is based on.
-First, we define moments of the impulse responses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{sec:intromoments}).
-Our method will use these ellipsoid estimates to choose sets of impulse responses that can be computed in batches.
-Second, we describe an improved method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}).
-Our method will interpolate impulse responses using this approximation.
+First, we define moments of the impulses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulses (Section \ref{sec:intromoments}).
+%Our method will use these ellipsoid estimates to choose sets of impulses that can be computed in batches. 
+Second, we describe an improved method of approximating impulses by other nearby impulses, which we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}).
+%Our method will interpolate impulses using this approximation.
 
@@
 
-The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative integral kernel property. Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
+The impulse $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative integral kernel property. Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
 \begin{equation}
@@
 However, this direct approach is not feasible because we need to know $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ before we compute $\phi_x$, in order choose the point $x$.
-Computing $\phi_x$ in order to determine $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ would be extremely computationally expensive, and defeat the purpose of our algorithm, which is to reduce the computational cost by computing impulse responses in batches.
+Computing $\phi_x$ in order to determine $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ would be extremely computationally expensive, and defeat the purpose of our algorithm, which is to reduce the computational cost by computing impulses in batches.
 Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ indirectly, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions (e.g., 6 total operator applies in two spatial dimensions and 10 in three spatial dimensions).
@@
 \horzbar & \mathbf{b}_1^T & \horzbar \\
-\horzbar & \mathbf{b}_2^T & \horzbar \\
+%\horzbar & \mathbf{b}_2^T & \horzbar \\
 & \vdots & \\
@@
 \mathbf{b}_1^T \mathbf{w} \\
-\mathbf{b}_2^T \mathbf{w} \\
+%\mathbf{b}_2^T \mathbf{w} \\
 \vdots \\
@@
 By computing one matrix-vector product of $\mathbf{B}^T$ with $\mathbf{w}$, we compute the inner product of each column of $\mathbf{B}$ with $\mathbf{w}$ simultaneously.
-The operator case is analogous, with the impulse response $\phi_x$ taking the place of a matrix column.
+The operator case is analogous, with the impulse $\phi_x$ taking the place of a matrix column.
 We have
@@
 \end{equation}
-By computing one operator application of $\Aop^T$ to $w$, we compute the inner product of each impulse response $\phi_x$ with $w$, for all points $x$ simultaneously.
+By computing one operator application of $\Aop^T$ to $w$, we compute the inner product of each impulse $\phi_x$ with $w$, for all points $x$ simultaneously.
 
@@
 Analogous manipulations show that $\spatialmean(x)$ and $\spatialcov(x)$ may be computed for all points $x$ simultaneously by applying $\Aop^T$ to the functions $L^i$ and $Q^{ij}$, respectively.
-To summarize the results of doing this, we have
+We have
 \begin{subequations}
@@
 
-Next, we make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
+We approximate the support of $\impulseresponse_x$ with the ellipsoid
 \begin{equation}
@@
 In practice, we find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results.
-The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.
+The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is typically conservative.
+% and typically far less mass resides in this region.
 
@@
 \end{align}
-These are four different ways to approximate an impulse response by a nearby impulse response, with each successive approximation building upon the previous ones.
+These are four different ways to approximate an impulse by a nearby impulse, with each successive approximation building upon the previous ones.
 Our method uses \eqref{eq:translate4}, which is the most sophisticated of these approximations.
@@
 We use \eqref{eq:vol_mean_var} in Section \ref{sec:intromoments} to compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ by applying $\Aop^T$ to a small number of polynomial functions.
-Then we use \eqref{eq:support_ellipsoid} to form ellipsoid shaped estimates for the support of the impulse responses $\phi_x$, \emph{without} computing them.
-This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}).
-We compute one batch of impulse responses, denoted $\eta_b$, by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section \ref{sec:get_impulse_response}).
+Then we use \eqref{eq:support_ellipsoid} to form ellipsoid shaped estimates for the support of the impulses $\phi_x$, \emph{without} computing them.
+This allows us to compute large numbers of impulses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}).
+We compute one batch of impulses, denoted $\eta_b$, by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section \ref{sec:get_impulse_response}).
 The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support ellipsoid for $\impulseresponse_{x_i}$ and the support ellipsoid for $\impulseresponse_{x_j}$ do not overlap if $i \neq j$ (Section \ref{sec:sample_point_selection}).
-Because the supports of these impulse responses do not overlap (or do not overlap much), we can post process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}).
+Because the supports of these impulses do not overlap (or do not overlap much), we can post process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}).
 The process is repeated until a desired number of batches is reached.
 
-Once the impulse response batches $\eta_b$ are computed, we approximate the integral kernel $\Aker(y,x)$ at arbitrary points $(y,x)$ by interpolation of translated and scaled versions of the computed impulse responses (Section \ref{sec:approximate_kernel_entries}).
+Once the impulse batches $\eta_b$ are computed, we approximate the integral kernel $\Aker(y,x)$ at arbitrary points $(y,x)$ by interpolation of translated and scaled versions of the computed impulses (Section \ref{sec:approximate_kernel_entries}).
 The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section \ref{sec:local_mean_displacement_invariance}.
@@
 We then use H-matrix methods to convert $\boldsymbol{\Aker}_H$ into an H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$.
-Schematically the algorithm proceeds as follows:
-\begin{equation*}
-\Aop \longrightarrow \spatialvol, \spatialmean, \spatialcov \longrightarrow \underbrace{\{x_i\}_{i=1}^\nsamplepts}_{\text{as }S_b\text{'s}} \longrightarrow \underbrace{\{\impulseresponse_{x_i}\}_{i=1}^\nsamplepts}_{\text{as }\eta_b\text{'s}} \longrightarrow \boldsymbol{\Aker}_H \longrightarrow \mathbf{A}_H.
-\end{equation*}
+%Schematically the algorithm proceeds as follows:
+%\begin{equation*}
+%\Aop \longrightarrow \spatialvol, \spatialmean, \spatialcov \longrightarrow \underbrace{\{x_i\}_{i=1}^\nsamplepts}_{\text{as }S_b\text{'s}} \longrightarrow \underbrace{\{\impulseresponse_{x_i}\}_{i=1}^\nsamplepts}_{\text{as }\eta_b\text{'s}} \longrightarrow \boldsymbol{\Aker}_H \longrightarrow \mathbf{A}_H.
+%\end{equation*}
 The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}. The cost is discussed in Section \ref{sec:overall_cost}.
@@
 
-We compute impulse responses $\impulseresponse_{x_i}$ in batches by applying $\Aop$ to Dirac combs.
+We compute impulses $\impulseresponse_{x_i}$ in batches by applying $\Aop$ to Dirac combs.
 The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the following weighted sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$:
@@
 \end{equation*}
-We compute the \emph{impulse response batch}, $\eta_k$, by applying $\Aop$ to the Dirac comb:
+We compute the \emph{impulse batch}, $\eta_k$, by applying $\Aop$ to the Dirac comb:
 \begin{equation}
@@
 for all $x_i \in \pointbatch_k$.
-By performing one matrix-vector product, $\xi_k \mapsto \left(\Aop \diraccomb_k^*\right)^*$, we recover the impulse responses $\impulseresponse_{x_i}$ associated with every point $x_i \in \pointbatch_k$.
-
-Each point source $\delta_{x_i}$ is scaled by $1/\spatialvol(x_i)$ so that the resulting scaled impulse responses within $\eta_k$ are comparable in magnitude.
+By performing one matrix-vector product, $\xi_k \mapsto \left(\Aop \diraccomb_k^*\right)^*$, we recover the impulses $\impulseresponse_{x_i}$ associated with every point $x_i \in \pointbatch_k$.
+
+Each point source $\delta_{x_i}$ is scaled by $1/\spatialvol(x_i)$ so that the resulting scaled impulses within $\eta_k$ are comparable in magnitude.
+Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, may overwhelm $\phi_{x_j}$ for a nearby point $x_j$ if $\spatialvol(x_i)$ is much larger than $\spatialvol(x_j)$.
 Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section \ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $\spatialvol(x_i)$ is smaller than a predetermined threshold.
-Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, may overwhelm $\phi_{x_j}$ for a nearby point $x_j$ if $\spatialvol(x_i)$ is much larger than $\spatialvol(x_j)$.
 
@@
 \end{equation*}
-where $\{x_i\}_{i=1}^{\numnbr}$ are the $\numnbr$ nearest sample points to $x$, excluding sample points $x_i$ for which $z_i \notin \Omega$.
+where $\{x_i\}_{i=1}^{\numnbr}$ are the $\numnbr$ nearest sample points to $x$, excluding points $x_i$ for which $z_i \notin \Omega$.
 Here $\numnbr$ is a small user defined parameter; e.g., $\numnbr=10$.
@@
 %The idea is that $\Aker(y,x) = f_i$ per the discussion in Section \ref{sec:local_mean_displacement_invariance}, so we approximate $\Aker(y,x)$ by interpolating the values $f_i$. 
-Interpolation is performed using the following radial basis function scheme:
+Interpolation is performed using the following radial basis function \cite{Wendland04} scheme:
 \begin{equation}
@@
 \end{equation}
-where $\rbfweight_i$ are weights, and $\rbf$ is a polyharmonic spline radial basis function, which is defined as follows:
-\begin{equation}
-\rbf(r) := \begin{cases}
-r^\gdim, & \gdim=1,3,5,\dots \\
-r^\gdim \log r, & \gdim=2,4,6,\dots
-\end{cases}
-\end{equation}
-for $r>0$, and $\rbf(0):=0$. The vector of weights, $\rbfweight = (\rbfweight_1, \rbfweight_2, \dots, \rbfweight_{\numnbr})^T$, is found as the solution to the $\numnbr \times \numnbr$ linear system
+where $\rbfweight_i$ are weights, and $\rbf$ is a polyharmonic spline radial basis function, which is defined as $\rbf(r):=r^\gdim$ if $\gdim$ is odd, and $\rbf(r):=r^\gdim \log r$ for $r>0$ and $\rbf(0):=0$ if $\gdim$ is even.
+%\begin{equation}
+%\rbf(r) := \begin{cases}
+%r^\gdim, & \gdim=1,3,5,\dots \\
+%r^\gdim \log r, & \gdim=2,4,6,\dots
+%\end{cases}
+%\end{equation}
+%for $r>0$, and $\rbf(0):=0$. 
+The vector of weights, $\rbfweight = (\rbfweight_1, \rbfweight_2, \dots, \rbfweight_{\numnbr})^T$, is found as the solution to the $\numnbr \times \numnbr$ linear system
 \begin{equation}
@@
 Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a certain least squares sense \cite{Duchon77}.
-For more details on radial basis function interpolation, see \cite{Wendland04}.
+%For more details on radial basis function interpolation, see \cite{Wendland04}. 
 %The interpolation procedure is illustrate in Figure \ref{fig:rbf_interpolation}.
@@
 
-The computational cost of our method may be divided into the costs to perform the following tasks: (1) Computing impulse response moments and batches; (2) Building the H-matrix; (3) Performing linear algebra operations with the H-matrix.
+The computational cost of our method may be divided into the costs to perform the following tasks: (1) Computing impulse moments and batches; (2) Building the H-matrix; (3) Performing linear algebra operations with the H-matrix.
 This may optionally include modifying the H-matrix to make it symmetric positive semi-definite.
@@
 		$\fedim$ & $10^2$--$10^9$ & Number of finite element degrees of freedom \\
-		$\nbatch$ & $1$--$25$ & Number of impulse response batches \\
+		$\nbatch$ & $1$--$25$ & Number of impulse batches \\
 		$\hrank$ & $5$--$50$ & H-matrix rank \\
@@
 
-\paragraph{(1) Computing impulse response moments and batches} Computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ requires applying $\Aop$ to $1$, $\gdim$, and $\gdim(\gdim+1)/2$ vectors, respectively.
-Computing each impulse response batch requires applying $\Aop$ to one vector, so computing all impulse response batches $\{\eta_i\}_{i=1}^{\nbatch}$ requires $\nbatch$ operator applies.
-In total, computing all impulse response moments and batches therefore requires
+\paragraph{(1) Computing impulse moments and batches} Computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ requires applying $\Aop$ to $1$, $\gdim$, and $\gdim(\gdim+1)/2$ vectors, respectively.
+Computing each impulse batch requires applying $\Aop$ to one vector, so computing all impulse batches $\{\eta_i\}_{i=1}^{\nbatch}$ requires $\nbatch$ operator applies.
+In total, computing all impulse moments and batches therefore requires
 \begin{equation*}
@@
 
-Computing the impulse response batches also requires choosing sample point batches via the greedy ellipsoid packing algorithm described in Section \ref{sec:sample_point_selection}.
+Computing the impulse batches also requires choosing sample point batches via the greedy ellipsoid packing algorithm described in Section \ref{sec:sample_point_selection}.
 Choosing the $i$th batch of sample points may require performing up to $\fedim |\pointbatch_i|$ ellipsoid intersection tests, where $|\pointbatch_i|$ is the number of sample points in the $i$th batch.
@@
 requires evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix.
-To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulse responses used in the RBF interpolation.
+To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulses used in the RBF interpolation.
 This is done using a precomputed kd-tree of sample points, and requires $O(\numnbr \log \nsamplepts)$ floating point and logical elementary operations.
@@
 We use the framework described in this paper to generate Hessian preconditioners.
-We build H-matrix approximations of the data misfit Gauss-Newton Hessian (the term in $\bm{H}_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 impulse responses batches.
+We build H-matrix approximations of the data misfit Gauss-Newton Hessian (the term in $\bm{H}_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 impulse batches.
 These approximations are denoted by PSF (1), PSF (5), and PSF (25), respectively.
@@
 	\caption{Convergence history for solving the Stokes inverse problem using preconditioned inexact Newton PCG.
-Preconditioners shown are our method with five impulse response batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
+Preconditioners shown are our method with five impulse batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
 Columns titled \#CG show the number of PCG iterations used to solve the Newton system for $\bm{\widehat{\basalfriction}}_k$.
@@
 Right (\ref{fig:gamma_sweep}) shows the number of PCG iterations required to achieve $||\bm{b}-\bm{H} \bm{x}_k||/||\bm{b}|| < 10^{-6}$, where $\bm{x}_k$ is the $k$th PCG iterate, for a range of different $\gamma$.
-Results in these figures are shown for our preconditioner with 1, 5, and 25 batches of impulse responses (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
+Results in these figures are shown for our preconditioner with 1, 5, and 25 batches of impulses (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
 The preconditioner is constructed using $\bm{H}_\text{gn}$.
@@
 The method only requires access to the operator via application of the operator to a small number of vectors.
-The idea of the method is to compute batches of impulse responses by applying the operator to Dirac combs of scattered point sources, then interpolate these impulse responses to approximate entries of the operator's integral kernel.
+The idea of the method is to compute batches of impulses by applying the operator to Dirac combs of scattered point sources, then interpolate these impulses to approximate entries of the operator's integral kernel.
 The interpolation is based on a new principle we call ``local mean displacement invariance,'' which generalizes and improves upon classical local translation invariance.
@@
 Fast H-matrix arithmetic is then used to perform further linear algebra operations that cannot be performed easily with the original operator, such as matrix factorization and inversion.
-The supports of the impulse responses are estimated to be contained in ellipsoids, which are determined a-priori via a new procedure that involves applying the operator to a small number of polynomial functions.
-Point source locations for the impulse response batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulse responses per batch as possible, while ensuring that the corresponding ellipsoids do not overlap.
+The supports of the impulses are estimated to be contained in ellipsoids, which are determined a-priori via a new procedure that involves applying the operator to a small number of polynomial functions.
+Point source locations for the impulse batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulses per batch as possible, while ensuring that the corresponding ellipsoids do not overlap.
 We applied the method to approximate the Gauss-Newton Hessian in a Stokes ice sheet inverse problem, and saw that the approximation substantially outperforms existing Hessian approximation methods.
@@
 
-\subsection{Ellipsoid bounding box test}
-
-One may reduce the average cost of performing ellipsoid intersection tests by first checking whether the coordinate axis aligned bounding boxes for the ellipsoids intersect.
+%\subsection{Ellipsoid bounding box test}
+
+\paragraph{Ellipsoid bounding box test} Before performing the ellipsoid intersection test described above, we check whether the coordinate axis aligned bounding boxes for the ellipsoids intersect.
 If the bounding boxes do not intersect, then the ellipsoids do not intersect, so no further action is necessary.
-If the bounding boxes intersect, the ellipsoids might intersect, or they might not intersect, so one must perform the complete ellipsoid intersection test described in Section \ref{sec:fast_ellipsoid_intersection_test}.
+If the bounding boxes intersect, the ellipsoids might or might not intersect, so we perform the complete ellipsoid intersection test described above.
 The smallest coordinate axis aligned box that contains the ellipsoid

### 2022-11-08 paper/localpsf.tex
shaved off about a page to make space for Omar's edits

@@
 \hypersetup{
-  pdftitle={Matrix-free point spread function approximation of operators with locally supported non-negative integral kernels, with application to Hessians in PDE constrained inverse problems},
+  pdftitle={Point spread function approximation of high rank Hessians with locally supported non-negative integral kernels},
   pdfauthor={N. Alger, N. Petra, T. Hartland, and O. Ghattas}
@@
 \begin{abstract}
-	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels.
-Such operators arise as Schur complements when solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in PDE constrained inverse problems.
-The proposed method uses impulses of the underlying operator (computed at a collection of scattered points) and interpolates translated and scaled versions of these impulses to approximate entries of the integral kernel.
+	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels, with particular focus on approximating high rank Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs).
+%	The method may be used to approximate other operators, including certain Schur complements, Poincare-Steklov operators, covariance operators, and blurring operators. 
+	The proposed method uses impulse responses of the underlying operator (computed at a collection of scattered points) and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel.
 The interpolation uses a ``local mean displacement invariance'' approximation, which improves upon classical local translation invariance.
 Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources.
-By solving an ellipsoid packing problem, we choose as many point sources as possible per batch, while ensuring that the supports of the impulses within each batch do not overlap.
-Support ellipsoids are estimated a-priori based on impulse moments, and these moments are computed for impulses at all points simultaneously via a procedure that involves applying the operator to a small number of polynomial functions.
+By solving an ellipsoid packing problem, we choose as many point sources as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
+Support ellipsoids are estimated a-priori based on impulse response moments, and these moments are computed for impulse responses at all points simultaneously via a procedure that involves applying the operator to a small number of polynomial functions.
 The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
 Fast H-matrix methods are used to perform further matrix computations in nearly linear complexity, including matrix-matrix addition and multiplication, matrix inversion and factorization, and matrix-vector products.
-We illustrate our approach by applying the method to approximate the Hessian in an ice sheet flow PDE constrained inverse problem.
+We illustrate our approach by applying the method to approximate the Hessians in an ice sheet flow PDE constrained inverse problem with highly informative data, and an advection dominated advection-diffusion PDE constrained inverse problem.
 Numerical results demonstrate that the proposed method substantially outperforms existing methods.
-We show that high rank operators can be efficiently approximated using only a small number of operator applies.
+We show that our method can approximate high rank operators using only a small number of operator applies.
 \end{abstract}
@@
 \begin{keywords}
- 	data scalability, Hessian, hierarchical matrix, matrix-free, operator approximation, PDE constrained inverse problems, point spread function, product convolution, local translation invariance, impulse
+ 	data scalability, Hessian, hierarchical matrix, matrix-free, operator approximation, PDE constrained inverse problems, point spread function, product convolution, local translation invariance, impulse response, moment methods
 \end{keywords}
@@
 Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, and $L^2(\Omega)'$ is the space of real-valued continuous linear functionals on $L^2(\Omega)$.
-Such operators arise as Schur complements in Schur complement methods for solving partial differential equations (PDEs) \cite{le1997non,saad1999distributed}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics \cite{ChenStein21,GeogaEtAl20,LindgrenRueLindstrom11}, blurring operators in imaging \cite{DenisEtAl11,NagyOleary98}, and Hessians in distributed parameter PDE constrained optimization and inverse problems \cite{AlgerEtAl19,IsaacEtAl15}.
+We focus on approximating Hessians in optimization and inverse problems governed by partial differential equations (PDEs) \cite{AlgerEtAl19,IsaacEtAl15}, though such operators also arise as certain Schur complements in Schur complement methods for solving partial differential equations (PDEs) \cite{le1997non,saad1999distributed}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics \cite{ChenStein21,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging \cite{DenisEtAl11,NagyOleary98}.
 Here ``matrix-free'' means that we may apply $\Aop$ and its transpose\footnote{Recall that $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for operations with $\Aop^T$.
@@
 but we cannot easily access entries of $\Aop$'s integral kernel.
-In applications, evaluating the maps in \eqref{eq:eval_maps} may require performing a costly computational procedure, such as solving a large linear system.
-
-Our motivation is Hessian approximation in distributed parameter inverse problems governed by PDEs. 
-%That is, inverse problems in which one seeks to reconstruct an unknown parameter field from noisy observations of a state variable that depends on the parameter implicitly through the solution of a PDE.
-The Hessian is an operator that is central to efficient methods for solving such inverse problems in both the deterministic and Bayesian settings,
-%In the deterministic approach to the inverse problem, one solves a regularized minimization problem to find the parameter that best fits the data.
-The objective function in the minimization problem is defined implicitly, so
-but information about the Hessian is only accessible via application of the Hessian to vectors \cite{Alger19,GhattasWillcox21}. 
-The most popular matrix-free Hessian approximation methods are based on low rank approximation \cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15} because there are efficient matrix-free methods for constructing low rank matrix approximations \cite{Cheng05,HalkoMartinssonTropp11}.
-However, in many inverse problems of practical interest the Hessian is high rank \cite{AmbartsumyanEtAl20,HartlandEtAl21}.
-Moreover, the numerical rank of the Hessian grows as the data in the inverse problem become more informative about the unknown parameter.
-Low rank approximation methods therefore suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation \cite{AlexanderianGloorGhattas16,Alger19,AlgerEtAl17}.
-%Matrix-free high rank approximation methods are needed.
+Evaluating the maps in \eqref{eq:eval_maps} may require, as a sub procedure, performing a costly computation, such as solving a large linear system or time stepping.
 
@@
 %	Brightness scale differs between the subfigures. 
-	Black stars are point source locations. Dashed gray ellipses are estimated impulse support ellipsoids. The large circle is $\partial \Omega$. 
+	Black stars are point source locations. Dashed gray ellipses are estimated impulse response support ellipsoids. The large circle is $\partial \Omega$. 
 	}
@@
 %Existing scalable high rank approximation methods typically require the ability to rapidly evaluate arbitrary entries of $\Aop$'s integral kernel, which we cannot do. 
-We use \emph{impulse interpolation} to form a high rank operator approximation using a small number of operator applies.
-The impulse, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
-We compute batches of impulses by applying $\Aop$ to weighted sums of delta distributions associated with batches of points scattered throughout the domain (see Figure \ref{fig:batches_intro}).
-Then we interpolate translated and scaled versions of these impulses to approximate entries of the operator's integral kernel.
+We use \emph{impulse response interpolation} to form a high rank operator approximation using a small number of operator applies.
+The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
+We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with batches of points scattered throughout the domain (see Figure \ref{fig:batches_intro}).
+Then we interpolate translated and scaled versions of these impulse responses to approximate entries of the operator's integral kernel.
 %This allows fast evaluation of approximate matrix entries, which are needed by high-rank approximation methods, in particular, classical H-matrix construction methods. 
-Picking the batches of points requires us to estimate the supports of the impulses $\phi_x$ \emph{before} we compute them. 
+Picking the batches of points requires us to estimate the supports of the impulse responses $\phi_x$ \emph{before} we compute them. 
 The idea of estimating the supports of the functions $\phi_x$ a-priori was inspired by techniques from resolution analysis in seismic imaging. There, 
@@
 Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
-Our impulse interpolation may be categorized as a PSF method that is loosely based on ``product convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
+Our method may be categorized as a PSF method that is loosely based on ``product convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
 PC and PSF methods have a long history dating back several decades.
-We note the following papers (among many others),  \cite{Adorf94,AlgerEtAl19,BigotEscandeWeiss19,EscandeWeiss15,EscandeWeiss22,EscandeWeiss12,FishEtAl96,NagyOleary98,ZhuLiFomelEtAl16}, in which the convolution kernels are constructed from sampling impulses of the operator to scattered point sources.
+We note the following papers (among many others),  \cite{Adorf94,AlgerEtAl19,BigotEscandeWeiss19,EscandeWeiss15,EscandeWeiss22,EscandeWeiss12,FishEtAl96,NagyOleary98,ZhuLiFomelEtAl16}, in which the convolution kernels are constructed from sampling impulse responses of the operator to scattered point sources.
 For background on PC and PSF methods, we recommend the following papers: \cite{DenisEtAl15,EscandeWeiss17,GentileCourbinMeylan13}.
-While PC approximations are based on an assumption of local translation invariance, the method we propose is based on an assumption that we call ``local mean displacement invariance,'' which is more general than local translation invariance, and includes both CUR low rank approximation and local translation invariance as special cases (Section \ref{sec:local_mean_displacement_invariance} and Figure \ref{fig:mean_displacement_invariance}).
+While PC approximations are based on an assumption of local translation invariance, the method we propose is based on an assumption we call ``local mean displacement invariance,'' which is more general than local translation invariance, and includes both CUR low rank approximation and local translation invariance as special cases (Section \ref{sec:local_mean_displacement_invariance} and Figure \ref{fig:mean_displacement_invariance}).
 %Furthermore, PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulses.
@@
 The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks \cite{GrasedyckHackbusch03}.
-The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.
+The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank Gaussian surrogate models that can be used for uncertainty quantification.
 %Classical H-matrix construction methods 
@@
 %require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.
-Recently there have been improvements in matrix-free H-matrix construction methods based on the ``peeling process'' \cite{LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}, which we do not use. 
-These alternative methods have been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{AmbartsumyanEtAl20,HartlandEtAl22}. 
-Methods based on the peeling process are asymptotically scalable,
-%(typically the number of matrix vector products required scales as $O(\log N)$), 
-but in practice the required number of matrix vector products is large. 
+
+\section{The need for high rank Hessian approximations in inverse problems governed by PDEs}
+\label{sec:hessian}
+
+In distributed parameter inverse problems governed by PDEs, one seeks to infer an unknown spatially varying parameter field from limited observations of a state variable that depends on the parameter implicitly through the solution of a PDE.
+%Inverse problems are typically ill-posed, in the sense that many different parameter fields may yield predicted observations that agree with the actual observations to within the noise level. 
+Two different frameworks are conventionally used to deal with this ill-posedness: the deterministic framework \cite{DETERMINISTIC}, and the Bayesian probabilistic framework \cite{STUART,TARANTOLA}.
+In the deterministic framework, one solves an optimization problem to find the parameter that best fits the observations, subject to appropriate regularization.
+In the probabilistic setting, Bayes' theorem combines the observations with prior information to form a posterior distribution over the space of all possible parameter fields, and computations are performed to extract statistical information about the parameter from this posterior.
+The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal to each other under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian''.
+The Hessian consists of a data misfit term (the ``data misfit Hessian''), which depends on mismatch between the actual observations and predicted observations based on a candidate parameter, and a regularization or prior term (the ``regularization Hessian'') which does not depend on the observations.
+
+Hessian approximations and preconditioners are highly desirable because the Hessian is central to efficient solution of inverse problems in both the deterministic and Bayesian settings.
+When solving the deterministic optimization problem with Newton-type methods, the Hessian is the coefficient operator for the linear system that must be solved at every Newton iteration.
+Good Hessian preconditioners reduce the number of iterations required to solve these Newton linear systems with Krylov methods such as conjugate gradient.
+In the Bayesian setting, the inverse of the Hessian is the covariance of the best local Gaussian approximation of the posterior.
+Hessian-based local Gaussian approximations of the posterior may be used as surrogate models of the posterior, or may be used as proposal distributions for Markov chain Monte-Carlo (MCMC) methods for drawing samples from the posterior.
+
+\begin{figure}
+	\begin{subfigure}{0.5\textwidth}
+		\includegraphics[scale=0.23]{informed_uninformed_modes_quadratic.pdf}
+		\caption{}
+	\end{subfigure}
+	\begin{subfigure}{0.48\textwidth}
+		\includegraphics[scale=0.1]{misfit_and_reg_hessian_eigenvalues.pdf}
+		\caption{}
+	\end{subfigure}
+	\caption{(a) The Hessian eigenstructure characterizes the directional scalings of the best local quadratic approximation to the objective function in an inverse problem.
+The larger an eigenvalue of the Hessian is, the more closely spaced the contours of the local quadratic approximation are in the corresponding eigenvector direction.
+(b) As the data in an inverse problem becomes more informative about the parameter, the eigenvalues of the data misfit Hessian increase.
+As the regularization is weakened, the eigenvalues of the regularization Hessian decrease.
+These illustrations are taken from \cite[Chapter 1]{Alger19}, with minor modifications.}
+	\label{fig:hessian_eigenstructure}
+\end{figure}
+
+Owing to the implicit dependence of predicted observations on the parameter, entries of the Hessian are not easily accessible.
+Rather, the Hessian may be applied to a vector via a computational process that involves solving a pair of forward and adjoint PDEs which are linearizations of the original PDE \cite{Alger19,GhattasWillcox21}.
+The most popular matrix-free Hessian approximation methods are based on low rank approximation of either the data misfit Hessian, or the data misfit Hessian preconditioned by the regularization Hessian  \cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15}.
+Krylov methods or randomized methods are typically used to construct these low rank matrix approximations by applying the Hessian to vectors \cite{Cheng05,HalkoMartinssonTropp11}.
+Using these methods, the required number of Hessian applications (and hence the required number of PDE solves) is proportional to the rank of the low rank approximation.
+However, in many inverse problems of practical interest the data misfit Hessian is high rank, which makes makes it extremely costly to approximate the Hessian using these low rank approximation methods \cite{AmbartsumyanEtAl20,HartlandEtAl21}.
+Examples of inverse problems with high-rank data misfit Hessians include advection dominated advection-diffusion inverse problems, high frequency wave inverse problems, and more generally, all inverse problem in which the observations are highly informative about the parameter.
+The eigenvalues of the data misfit Hessian characterize the sensitivity of the predicted observations to perturbations of the parameter in the corresponding eigenvector directions (see Figure \ref{fig:hessian_eigenstructure}), so more informative data leads to larger eigenvalues and therefore a larger numerical rank.
+Roughly speaking, the numerical rank of the data misfit Hessian is the dimension of the subspace of parameter space that is informed by the data.
+The numerical rank of the regularization preconditioned data misfit Hessian may be reduced by increasing the strength of the regularization, but this is throwing away useful information---components of the parameter that could be learned from the observations are instead reconstructed based on the regularization.
+Hence, low rank approximation methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation.
+The best case scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost) \cite{AlexanderianGloorGhattas16,Alger19,AlgerEtAl17}.
+Matrix-free high rank Hessian approximation methods are needed.
+
+Recently there have been improvements in matrix-free H-matrix construction methods in which the operator is applied to structured random vectors, and the response of the operator to those random vectors is then processed to construct the H-matrix  \cite{LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}.
+These methods have been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{AmbartsumyanEtAl20}. 
+These methods are asymptotically scalable, but in practice a large number of matrix vector products are required. 
 %Improving the peeling process to reduce its cost is currently an area of active research \cite{PEELINGRESEARCH}.
 %This is an area of active research, so more efficient and practical matrix-free H-matrix construction methods based on the peeling process may be discovered in the future.
-%Matrix-free H-matrix construction methods are currently a topic of considerable research, so more efficient matrix-free H-matrix methods based on the peeling process may be discovered in the future.
+Matrix-free H-matrix construction methods are currently an area of active research (see, e.g., \cite{LevittMartinsson}), so more efficient matrix-free H-matrix methods may be discovered in the future.
+In this paper, we form a H-matrix Hessian approximation.
+But rather than use these modern randomized methods to form the H-matrix directly, instead we exploit locality and non-negative integral kernel properties to form a PSF approximation of the data misfit Hessian, then convert the PSF approximation to H-matrix format and combine it with the regularization Hessian using classical H-matrix construction techniques.
+Not all data misfit Hessians satisfy the local non-negative integral kernel properties (for example, the wave inverse problem data misfit Hessian has a substantial amount of negative entries in its integral kernel).
+However, many data misfit Hessians of practical interest do satisfy these properties (either exactly or approximately), and our method is effective for approximating these Hessians.
+
+%Classical methods for approximating the Hessian are based on low rank approximation of either the data misfit Hessian, or the data misfit Hessian preconditioned by the inverse of the prior term in the Hessian.
+However,  More generally, the eigenvalues of the Hessian characterize how informative the data are about the component of the parameter in the corresponding parameter direction.
+The more informed the component of the parameter, the larger
+
+%In the deterministic setting, the Hessian of the objective function is central to efficient solution of the optimization problem, and in the
+
+% given the observed data is  one explores the posterior probability distribution for the parameter
+
+%Our motivation is high rank Hessian approximation in distributed parameter inverse problems governed by PDEs. 
+%The Hessian is an operator that is central to efficient methods for solving such inverse problems and is only available matrix-free; we will discuss Hessians in inverse problems governed by PDEs in greater detail in Section \ref{sec:hessian}.
+
+%Introduce inverse problem
+%
+%Role of Hessian
+%
+%Structure of Hessian that we want to exploit
+%
+%Low rank
+%
+%Need for High rank
+%
+%Low rank approximation of the Hessian:
+%\begin{itemize} 
+%	\item Hessian eigenstructure is determined by data informativeness. Bigger eigenvalue => more informed component in eigenvector direction. 
+%	\item informative data => high rank hessian
+%	\item get high rank approximation by exploiting different structure. Locality, non-negativity of kernel
+%	\item Large rank in absolute terms, even if low rank in relative terms
+%\end{itemize}
 
 \section{Preliminaries}
+
+\subsection{Integral operators, distributions and impulse responses}
+\label{sec:impulse_response_2}
 
@@
 The integral kernel, $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$, exists in principle but is not easily accessible.
-In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define impulses (Section \ref{sec:impulse_response_2}), we state the conditions on $\Aop$ that our method requires (Section \ref{sec:conditions_2}), and we detail finite element discretization (Section \ref{sec:finite_element_kernel}).
-
-\subsection{Distributions and impulses}
-\label{sec:impulse_response_2}
+In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define impulse responses (Section \ref{sec:impulse_response_2}), we state the conditions on $\Aop$ that our method requires (Section \ref{sec:conditions_2}), and we detail finite element discretization (Section \ref{sec:finite_element_kernel}).
 
@@
 Let $\delta_x$ denote the delta distribution\footnote{Recall that the delta distribution $\delta_x:\mathcal{D}(\Omega)\rightarrow \mathbb{R}$ is defined by $\delta_x(w) = w(x)$ for all $w\in \mathcal{D}(\Omega)$.} (i.e., point source, impulse) centered at the point $x \in \Omega$.
-The \emph{impulse} of $\Aop$ associated with $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$, 
+The \emph{impulse response} of $\Aop$ associated with $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$, 
 \begin{equation}
@@
 	\item The kernel $\Aker$ is sufficiently regular so that $\phi_x$ is well defined for all $x\in \Omega$.
-	\item The supports of the impulses $\impulseresponse_x$ are contained in localized regions.
+	\item The supports of the impulse responses $\impulseresponse_x$ are contained in localized regions.
 	\item The integral kernel is non-negative in the sense that $$\Aker(y,x) \ge 0$$ for all $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
@@
 Our method may still perform well if these conditions are relaxed slightly.
-It is fine if the support of $\impulseresponse_x$ is not perfectly contained in a localized region, so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
-The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$, so long as it is non-negative for the vast majority of pairs of points $(y,x)$, and as long as the negative numbers are comparatively small.
+It is fine if the support of $\impulseresponse_x$ is not perfectly contained in a localized region, as long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
+The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$, as long as it is non-negative for the vast majority of pairs of points $(y,x)$, and as long as the negative numbers are comparatively small.
 If these conditions are violated, our method will incur additional error.
@@
 $
-Therefore, the formula for the impulse from \eqref{eq:impulse_response_delta_action} becomes 
+Therefore, the formula for the impulse response from \eqref{eq:impulse_response_delta_action} becomes 
 $
@@
 In this section we present two key innovations that our method is based on.
-First, we define moments of the impulses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulses (Section \ref{sec:intromoments}).
+First, we define moments of the impulse responses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{sec:intromoments}).
 %Our method will use these ellipsoid estimates to choose sets of impulses that can be computed in batches. 
-Second, we describe an improved method of approximating impulses by other nearby impulses, which we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}).
+Second, we describe an improved method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}).
 %Our method will interpolate impulses using this approximation.
@@
 
-The impulse $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative integral kernel property. Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
+The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative integral kernel property. Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
 \begin{equation}
@@
 However, this direct approach is not feasible because we need to know $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ before we compute $\phi_x$, in order choose the point $x$.
-Computing $\phi_x$ in order to determine $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ would be extremely computationally expensive, and defeat the purpose of our algorithm, which is to reduce the computational cost by computing impulses in batches.
+Computing $\phi_x$ in order to determine $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ would be extremely computationally expensive, and defeat the purpose of our algorithm, which is to reduce the computational cost by computing impulse responses in batches.
 Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ indirectly, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions (e.g., 6 total operator applies in two spatial dimensions and 10 in three spatial dimensions).
@@
 By computing one matrix-vector product of $\mathbf{B}^T$ with $\mathbf{w}$, we compute the inner product of each column of $\mathbf{B}$ with $\mathbf{w}$ simultaneously.
-The operator case is analogous, with the impulse $\phi_x$ taking the place of a matrix column.
+The operator case is analogous, with $\phi_x$ taking the place of a matrix column.
 We have
@@
 \end{equation}
-By computing one operator application of $\Aop^T$ to $w$, we compute the inner product of each impulse $\phi_x$ with $w$, for all points $x$ simultaneously.
+By computing one operator application of $\Aop^T$ to $w$, we compute the inner product of each $\phi_x$ with $w$, for all points $x$ simultaneously.
 
@@
 \end{align}
-These are four different ways to approximate an impulse by a nearby impulse, with each successive approximation building upon the previous ones.
+These are four different ways to approximate an impulse response by a nearby impulse response, with each successive approximation building upon the previous ones.
 Our method uses \eqref{eq:translate4}, which is the most sophisticated of these approximations.
@@
 We use \eqref{eq:vol_mean_var} in Section \ref{sec:intromoments} to compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ by applying $\Aop^T$ to a small number of polynomial functions.
-Then we use \eqref{eq:support_ellipsoid} to form ellipsoid shaped estimates for the support of the impulses $\phi_x$, \emph{without} computing them.
-This allows us to compute large numbers of impulses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}).
-We compute one batch of impulses, denoted $\eta_b$, by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section \ref{sec:get_impulse_response}).
+Then we use \eqref{eq:support_ellipsoid} to form ellipsoid shaped estimates for the support of the $\phi_x$, \emph{without} computing them.
+This allows us to compute large numbers of $\impulseresponse_{x_i}$ in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}).
+We compute one batch, denoted $\eta_b$, by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section \ref{sec:get_impulse_response}).
 The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support ellipsoid for $\impulseresponse_{x_i}$ and the support ellipsoid for $\impulseresponse_{x_j}$ do not overlap if $i \neq j$ (Section \ref{sec:sample_point_selection}).
-Because the supports of these impulses do not overlap (or do not overlap much), we can post process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}).
+Because these supports do not overlap (or do not overlap much), we can post process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}).
 The process is repeated until a desired number of batches is reached.
 
-Once the impulse batches $\eta_b$ are computed, we approximate the integral kernel $\Aker(y,x)$ at arbitrary points $(y,x)$ by interpolation of translated and scaled versions of the computed impulses (Section \ref{sec:approximate_kernel_entries}).
+Once the batches $\eta_b$ are computed, we approximate the integral kernel $\Aker(y,x)$ at arbitrary points $(y,x)$ by interpolation of translated and scaled versions of the computed $\impulseresponse_{x_i}$ (Section \ref{sec:approximate_kernel_entries}).
 The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section \ref{sec:local_mean_displacement_invariance}.
@@
 
-We compute impulses $\impulseresponse_{x_i}$ in batches by applying $\Aop$ to Dirac combs.
+We compute impulse responses $\impulseresponse_{x_i}$ in batches by applying $\Aop$ to Dirac combs.
 The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the following weighted sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$:
@@
 for all $x_i \in \pointbatch_k$.
-By performing one matrix-vector product, $\xi_k \mapsto \left(\Aop \diraccomb_k^*\right)^*$, we recover the impulses $\impulseresponse_{x_i}$ associated with every point $x_i \in \pointbatch_k$.
-
-Each point source $\delta_{x_i}$ is scaled by $1/\spatialvol(x_i)$ so that the resulting scaled impulses within $\eta_k$ are comparable in magnitude.
+By performing one matrix-vector product, $\xi_k \mapsto \left(\Aop \diraccomb_k^*\right)^*$, we recover $\impulseresponse_{x_i}$ for every point $x_i \in \pointbatch_k$.
+
+Each point source $\delta_{x_i}$ is scaled by $1/\spatialvol(x_i)$ so that the resulting scaled impulse responses within $\eta_k$ are comparable in magnitude.
 Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, may overwhelm $\phi_{x_j}$ for a nearby point $x_j$ if $\spatialvol(x_i)$ is much larger than $\spatialvol(x_j)$.
@@
 		$\fedim$ & $10^2$--$10^9$ & Number of finite element degrees of freedom \\
-		$\nbatch$ & $1$--$25$ & Number of impulse batches \\
+		$\nbatch$ & $1$--$25$ & Number of batches \\
 		$\hrank$ & $5$--$50$ & H-matrix rank \\
@@
 
-\paragraph{(1) Computing impulse moments and batches} Computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ requires applying $\Aop$ to $1$, $\gdim$, and $\gdim(\gdim+1)/2$ vectors, respectively.
-Computing each impulse batch requires applying $\Aop$ to one vector, so computing all impulse batches $\{\eta_i\}_{i=1}^{\nbatch}$ requires $\nbatch$ operator applies.
-In total, computing all impulse moments and batches therefore requires
+\paragraph{(1) Computing impulse response moments and batches} Computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ requires applying $\Aop$ to $1$, $\gdim$, and $\gdim(\gdim+1)/2$ vectors, respectively.
+Computing each $\eta_i$ requires applying $\Aop$ to one vector, so computing $\{\eta_i\}_{i=1}^{\nbatch}$ requires $\nbatch$ operator applies.
+In total, computing all impulse response moments and batches therefore requires
 \begin{equation*}
@@
 
-Computing the impulse batches also requires choosing sample point batches via the greedy ellipsoid packing algorithm described in Section \ref{sec:sample_point_selection}.
+Computing the impulse response batches also requires choosing sample point batches via the greedy ellipsoid packing algorithm described in Section \ref{sec:sample_point_selection}.
 Choosing the $i$th batch of sample points may require performing up to $\fedim |\pointbatch_i|$ ellipsoid intersection tests, where $|\pointbatch_i|$ is the number of sample points in the $i$th batch.
@@
 In the numerical results (Section \ref{sec:numerical_results}), we use one matrix-matrix addition to add the H-matrix approximation of the data misfit term in the Hessian to the regularization term in the Hessian.
-Constructing a H-matrix representation of the regularization term in the Hessian requires one matrix inversion and two matrix-matrix multiplications.
+Constructing an H-matrix representation of the regularization term in the Hessian requires one matrix inversion and two matrix-matrix multiplications.
 Then, we use one matrix factorization to form a preconditioner for the Newton linear system.
@@
 We use our method to approximate the data misfit Gauss-Newton Hessian in an inverse problem governed by a PDE which models steady state ice sheet flow \cite{PetraEtAl12}.
-Briefly, a sheet of ice is sliding down a mountain (see Figure \ref{fig:ice_mountain_mesh}).
-Given observations of the tangential component of the ice velocity on the top surface of the ice, we invert for the logarithm of the unknown spatially varying basal friction parameter which governs how easily the ice slides along the ground underneath it.
+Briefly, a sheet of ice is flowing down a mountain (see Figure \ref{fig:ice_mountain_mesh}).
+Given observations of the tangential component of the ice velocity on the top surface of the ice, we invert for the logarithm of the unknown spatially varying basal friction parameter field, which governs the resistance to sliding along the base of the ice sheet.
+The inverse problem considered here is synthetic, but in practice surface velocity observations are available from satellites.
 
@@
 Here, $\basalfriction$ is the unknown logarithmic
-basal sliding parameter field which is defined on the $2D$ surface $\Gamma_b$.
+basal friction field (note that large $\basalfriction$ corresponds to high resistance to sliding) which is defined on the $2D$ surface $\Gamma_b$.
 The quantity $\bodyforce$ is the body force density due to gravity, $\stokesrobincoeff=10^6$ is a Robin boundary condition constant,  $\normalvec$ is the outward normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary.
@@
 We use the framework described in this paper to generate Hessian preconditioners.
-We build H-matrix approximations of the data misfit Gauss-Newton Hessian (the term in $\bm{H}_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 impulse batches.
+We build H-matrix approximations of the data misfit Gauss-Newton Hessian (the term in $\bm{H}_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 batches.
 These approximations are denoted by PSF (1), PSF (5), and PSF (25), respectively.
@@
 Our results show that our preconditioners are also good preconditioners for $\bm{H}$.
-The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to a H-matrix approximation of the Hessian of the regularization term to form a H-matrix approximation of $\bm{H}_\text{gn}$.
+The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to an H-matrix approximation of the Hessian of the regularization term to form an H-matrix approximation of $\bm{H}_\text{gn}$.
 Straightforward calculation shows that the discretized Hessian of the regularization term is given by
@@
 These matrices are sparse, so we form H-matrix representations of these matrices and combine them using standard sparse H-matrix techniques.
-We factor the overall Gauss-Newton Hessian approximation, denoted $\bm{\preconditioner}$, using fast H-matrix methods, then and use the factorization as a preconditioner.
+We factor the overall Gauss-Newton Hessian approximation, denoted $\bm{\preconditioner}$, using fast H-matrix methods, then use the factorization as a preconditioner.
 
@@
 	\caption{Convergence history for solving the Stokes inverse problem using preconditioned inexact Newton PCG.
-Preconditioners shown are our method with five impulse batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
+Preconditioners shown are our method with five batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
 Columns titled \#CG show the number of PCG iterations used to solve the Newton system for $\bm{\widehat{\basalfriction}}_k$.
@@
 Right (\ref{fig:gamma_sweep}) shows the number of PCG iterations required to achieve $||\bm{b}-\bm{H} \bm{x}_k||/||\bm{b}|| < 10^{-6}$, where $\bm{x}_k$ is the $k$th PCG iterate, for a range of different $\gamma$.
-Results in these figures are shown for our preconditioner with 1, 5, and 25 batches of impulses (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
+Results in these figures are shown for our preconditioner with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
 The preconditioner is constructed using $\bm{H}_\text{gn}$.
@@
 The method only requires access to the operator via application of the operator to a small number of vectors.
-The idea of the method is to compute batches of impulses by applying the operator to Dirac combs of scattered point sources, then interpolate these impulses to approximate entries of the operator's integral kernel.
+The idea of the method is to compute batches of impulse responses by applying the operator to Dirac combs of scattered point sources, then interpolate these impulse responses to approximate entries of the operator's integral kernel.
 The interpolation is based on a new principle we call ``local mean displacement invariance,'' which generalizes and improves upon classical local translation invariance.
-The ability to quickly approximate arbitrary integral kernel entries permits us to form a H-matrix approximation of the operator.
+The ability to quickly approximate arbitrary integral kernel entries permits us to form an H-matrix approximation of the operator.
 Fast H-matrix arithmetic is then used to perform further linear algebra operations that cannot be performed easily with the original operator, such as matrix factorization and inversion.
-The supports of the impulses are estimated to be contained in ellipsoids, which are determined a-priori via a new procedure that involves applying the operator to a small number of polynomial functions.
-Point source locations for the impulse batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulses per batch as possible, while ensuring that the corresponding ellipsoids do not overlap.
+The supports of the impulse responses are estimated to be contained in ellipsoids, which are determined a-priori via a new procedure that involves applying the operator to a small number of polynomial functions.
+Point source locations for the impulse response batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulse responses per batch as possible, while ensuring that the corresponding ellipsoids do not overlap.
 We applied the method to approximate the Gauss-Newton Hessian in a Stokes ice sheet inverse problem, and saw that the approximation substantially outperforms existing Hessian approximation methods.
 Although the method we presented is not applicable to all Hessians, it is applicable to many Hessians of practical interest.
-For these Hessians, our method offers a \emph{data scalable} alternative to conventional low rank approximation, because our method can form high rank approximations of an operator using a small number of operator applies.
+For these Hessians, our method offers a \emph{data scalable} alternative to conventional low rank approximation, because our method can form high rank approximations of an operator using a small number of operator applies, and thus forward and adjoint PDE solves.
 
@@
 \begin{align*}
-	E_1 :=& \{x : (x - \spatialmean_1)^T \spatialcov_1^{-1} (x - \spatialmean_1) \le \tau^2\}, \\
-	E_2 :=& \{x : (x - \spatialmean_2)^T \spatialcov_2^{-1} (x - \spatialmean_2) \le \tau^2\},
+	E_i :=& \{x : (x - \spatialmean_i)^T \spatialcov_i^{-1} (x - \spatialmean_i) \le \tau^2\}
 \end{align*}
-where $\spatialmean_1, \spatialmean_2 \in \mathbb{R}^\gdim$, and $\spatialcov_1, \spatialcov_2 \in \mathbb{R}^{\gdim \times \gdim}$ are positive definite.
+for $i\in\{1,2\}$, where $\spatialmean_1, \spatialmean_2 \in \mathbb{R}^\gdim$, and $\spatialcov_1, \spatialcov_2 \in \mathbb{R}^{\gdim \times \gdim}$ are positive definite.
 Let $K$ be the following one dimensional convex function:
@@
 
-The function $K(s)$ may be evaluated quickly for many $s$ by pre-computing the solution to the generalized eigenvalue problem
+The function $K(s)$ may be evaluated quickly for many $s$ by computing the solution to the generalized eigenvalue problem
 $
@@
 $
-where $\eigenvectormatrix \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$.
-The matrix $\eigenvectormatrix$ simultaneously diagonalizes $\spatialcov_1$ and $\spatialcov_2$, in the sense that $\eigenvectormatrix^T\spatialcov_1 \eigenvectormatrix = \Lambda$, and $\eigenvectormatrix^T\spatialcov_2 \eigenvectormatrix = I$, where $I$ is the $\gdim \times \gdim$ identity matrix.
-Using this diagonalization, and some algebraic manipulations, we may write $K(s)$ as
+where $\eigenvectormatrix \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors, and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$.
+%The matrix $\eigenvectormatrix$ simultaneously diagonalizes $\spatialcov_1$ and $\spatialcov_2$, in the sense that $\eigenvectormatrix^T\spatialcov_1 \eigenvectormatrix = \Lambda$, and $\eigenvectormatrix^T\spatialcov_2 \eigenvectormatrix = I$, where $I$ is the $\gdim \times \gdim$ identity matrix.
+Using this diagonalization, and
+After some algebraic manipulations, we have
 \begin{equation}
@@
 where $v := \eigenvectormatrix^T\left(\spatialmean_1 - \spatialmean_2\right)$.
-To check whether $E_1$ and $E_2$ intersect, we compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, form $v$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's minimization algorithm \cite{Brent71}.
+To check whether $E_1$ and $E_2$ intersect, we compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, form $v$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on $(0,1)$ using Brent's method \cite{Brent71}.
 If $K(s^*) <0$ at the minimizer, $s^*$, then $E_1 \cap E_2 = \{\}$.
@@
 
-\paragraph{Ellipsoid bounding box test} Before performing the ellipsoid intersection test described above, we check whether the coordinate axis aligned bounding boxes for the ellipsoids intersect.
-If the bounding boxes do not intersect, then the ellipsoids do not intersect, so no further action is necessary.
-If the bounding boxes intersect, the ellipsoids might or might not intersect, so we perform the complete ellipsoid intersection test described above.
-The smallest coordinate axis aligned box that contains the ellipsoid
-$
-E = \{x : (x - \spatialmean)^T \spatialcov^{-1} (x - \spatialmean) \le \tau^2\}
-$
-is the box
-$
-	\prod_{i=1}^\gdim \left[ \spatialmean_i -\tau\sqrt{\Sigma_{ii}}, \spatialmean_i +\tau\sqrt{\Sigma_{ii}} \right],
-$
-where $\Sigma_{ii}$ is the $i$th diagonal entry of $\Sigma$\footnote{See \url{https://math.stackexchange.com/q/3928964}}. The boxes
-$\prod_{i=1}^\gdim \left[a_i^-, a_i^+\right]$ and $\prod_{i=1}^\gdim \left[b_i^-, b_i^+\right]$
-intersect if and only if $a_i^- \le b_i^+$ and $b_i^- \le a_i^+$ for $i=1,\dots,d$.
+This ellipsoid intersection test is only performed if the coordinate axis aligned bounding boxes for the ellipsoids intersect. The smallest coordinate axis aligned box that contains $E_i$ is the box
+$
+	\prod_{k=1}^\gdim \left[ \spatialmean_i^{k} -\tau\sqrt{\Sigma_i^{kk}}, \spatialmean_i^{k} +\tau\sqrt{\Sigma_i^{kk}} \right],
+$
+where $\spatialmean_i^{k}$ is the $k$th entry of $\spatialmean_i$, and $\Sigma_i^{kk}$ is the $k$th diagonal entry of $\Sigma_i$\footnote{See \url{https://math.stackexchange.com/q/3928964}}. The boxes
+$\prod_{k=1}^\gdim \left[a_{-}^k, a_{+}^k\right]$ and $\prod_{k=1}^\gdim \left[b_{-}^k, b_{+}^k\right]$
+intersect if and only if $a_{-}^k \le b_{+}^k$ and $b_{-}^k \le a_{+}^k$ for $k=1,\dots,d$.
 
@@
 \label{sec:make_spd}
-
-In many problems of practical interest (e.g., Hessian approximation) $\Aop$ is symmetric positive semi-definite.
-However, $\mathbf{A}_H$ is generally non-symmetric and indefinite because of approximation error.
+In many problems $\Aop$ is symmetric positive semi-definite. However, $\mathbf{A}_H$ is non-symmetric and indefinite because of approximation error. 
 This is undesirable.
 Symmetry and positive semi-definiteness are important properties which should be preserved if possible.
-Also, lacking these properties may prevent one from using highly effective algorithms, such as the conjugate gradient method, to perform further useful operations involving $\mathbf{A}_H$.
-
-If $\Aop$ is symmetric, we symmetrize $\mathbf{A}_H$ as follows
+Also, lacking these properties may prevent one from using highly effective algorithms, such as the conjugate gradient method, to perform further operations involving $\mathbf{A}_H$.
+
+If $\Aop$ is symmetric, we symmetrize $\mathbf{A}_H$ as follows:
 $
@@
 $
-If $\Aop$ is also positive semi-definite, we adapt the high order rational function spectral projection method from \cite[Section 13.8]{HackbuschKress07} to flip (or approximately flip) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ to be positive.
-In detail, let $\Pi$ be the indicator function for the interval $(a,b)$.
-That is, $\Pi(\lambda)=0$ for $\lambda \in (0,1)$, and $\Pi(\lambda)=0$ otherwise.
-%\begin{equation*}
-%	\Pi(\lambda) = \begin{cases}
-%		1, & \lambda \in (a,b), \\
-%		0, & \text{otherwise}.
-%	\end{cases}
-%\end{equation*}
-Using functional calculus to extend the domain of the function $\Pi$ from scalars to matrices, we define
-$
-\mathbf{\Pi} := \Pi(\mathbf{A}_H^{\text{sym}}).
-$
-The matrix $\mathbf{\Pi}$ is the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to eigenvalues of $\mathbf{A}_H^{\text{sym}}$ that reside in $(a,b)$.
-If we choose $a < \lambda_\text{min}$ and $b=0$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, then $\mathbf{\Pi}$ is the spectral projector onto the eigenspace associated with negative eigenvalues, and $\mathbf{\Pi}\mathbf{A}_H^{\text{sym}}$ is the ``negative component'' of $\mathbf{A}_H^{\text{sym}}$ in which all the negative eigenvalues are the same but all the positive eigenvalues are replaced by zero.
-Subtracting a negative number from itself twice flips its sign, so the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$ is given by
-\begin{equation}
-\label{eq:I_Pi_abs}
-	|\mathbf{A}_H^{\text{sym}}| = \left(\mathbf{I} - 2\mathbf{\Pi}\right)\mathbf{A}_H^{\text{sym}},
-\end{equation}
-where $\mathbf{I}$ is the identity matrix of the appropriate size.
-By ``matrix absolute value,'' we mean that $|\mathbf{A}_H^{\text{sym}}|$ is the matrix that has the same eigenvectors as $\mathbf{A}_H^{\text{sym}}$, but the eigenvalues of $|\mathbf{A}_H^{\text{sym}}|$ are the absolute values of the eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[scale=0.5]{spd_rational_function.pdf}
-%	\caption{Rational function $\Pi_\ratord(\lambda)$ for $a=-1$, $b=0$, and $\ratord \in \{1,2,3\}$. As $\ratord$ increases, $\Pi_\ratord$ approaches the indicator function for the set $(a,b)$. }
-%	\label{fig:spd_rat}
-%\end{figure}
-
-Computing $\mathbf{\Pi}=\Pi(\mathbf{A}_H^{\text{sym}})$ directly is difficult, but we may approximate $\Pi$ by a high order rational function $\Pi_\ratord \approx \Pi$ which is defined as follows:
-\begin{equation*}
-	 \Pi_\ratord(\lambda) := \frac{1}{1 + T(\lambda)^{2^\ratord} }, \qquad \text{where} \qquad T(\lambda) := \frac{2 \lambda - (b+a)}{b-a}.
-\end{equation*}
-Here $\ratord$ is a positive integer; the larger $\ratord$ is, the better the approximation $\Pi \approx \Pi_\ratord$. 
-%This is illustrated in Figure \ref{fig:spd_rat}. 
-Replacing scalar operations with matrix operations, we have an approximation $\mathbf{\Pi}_\ratord \approx \mathbf{\Pi}$ given by
+If $\Aop$ is positive semi-definite, we adapt the rational method from \cite[Section 13.8]{HackbuschKress07} to approximate the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$, which we denote by $|\mathbf{A}_H^{\text{sym}}|$.
+That is, $|\mathbf{A}_H^{\text{sym}}|$ has the same eigenvectors as $\mathbf{A}_H^{\text{sym}}$, but the eigenvalues of $|\mathbf{A}_H^{\text{sym}}|$ are the absolute values of the eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
+%flip (or approximately flip) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ to be positive. 
+Once constructed, we use the approximation of $|\mathbf{A}_H^{\text{sym}}|$ in place of $\mathbf{A}_H$ in further linear algebra operations.
+
+In detail, let $a$ be any scalar satisfying $-a < \lambda_\text{min}$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, and set 
+\begin{equation}
+\label{eq:ratT999}
+\mathbf{T} :=  \left(2 \mathbf{A}_H^{\text{sym}} + a \mathbf{I}\right)/a,
+\end{equation}
+From the results in \cite[Section 13.8]{HackbuschKress07} it is straightforward to show that
 \begin{equation}
 \label{eq:rational343}
-\mathbf{\Pi}_\ratord := \Pi_\ratord\left(\mathbf{A}_H^{\text{sym}}\right) = \left(\mathbf{I} + \mathbf{T}^{2^\ratord}\right)^{-1},
-\end{equation}
-where
-$
-\mathbf{T} := T(\mathbf{A}_H^{\text{sym}}) = \left(2 \mathbf{A}_H^{\text{sym}} - (b+a) \mathbf{I}\right)/(b-a).
-$
-Using $\mathbf{\Pi}_\ratord$ instead of $\mathbf{\Pi}$ in \eqref{eq:I_Pi_abs} allows us to efficiently approximate $|\mathbf{A}_H^{\text{sym}}|$ using H-matrix methods. 
-This approximation procedure is as follows.
+\mathbf{\Pi}_\ratord := \left(\mathbf{I} + \mathbf{T}^{2^\ratord}\right)^{-1}
+\end{equation}
+approximates the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to negative eigenvalues. Thus, we have
+\begin{equation}
+\label{eq:rational76567}
+|\mathbf{A}_H^{\text{sym}}| \approx \left(\mathbf{I} - 2\mathbf{\Pi}_\ratord\right)\mathbf{A}_H^{\text{sym}}.
+\end{equation}
+To form an approximation of $|\mathbf{A}_H^{\text{sym}}|$, first, we compute $\lambda_\text{min}$ using the Lanczos method, and set $a=\gamma \lambda_\text{min}$, where $\gamma\ge 1$ is a parameter that must be chosen.
+The method is relatively insensitive to $\gamma$ and numerically we find that the method works well for any $\gamma \in [1,2]$. We use $\gamma=1.5$ in our numerical results. 
+Next, we compute $\mathbf{T}$ per \eqref{eq:ratT999}, compute $\mathbf{T}^{2^\ratord}$ per \eqref{eq:rational343} via repeated squaring, compute $\mathbf{\Pi}_\ratord$ per \eqref{eq:rational343}, and form an approximation of $|\mathbf{A}_H^{\text{sym}}|$ per \eqref{eq:rational76567}.
+Fast H-matrix methods are used to perform the required matrix-matrix additions and multiplications, and matrix inversion.
+We recommend using $\ratord=1$ or $2$.
+Larger $\ratord$ is unnecessary and may make the method unstable.
+
+%Finally, we approximate $|\mathbf{A}_H^{\text{sym}}|$ per \eqref{eq:rational76567}, using fast H-matrix operations to perform the required matrix-matrix addition and matrix inversion.
+
+%
+%define $
+%\mathbf{T} := T(\mathbf{A}_H^{\text{sym}}) = \left(2 \mathbf{A}_H^{\text{sym}} + a \mathbf{I}\right)/a.
+%$, where $-a < \lambda_\text{min}$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$ and define
+%
+%let $\mathbf{\Pi}$ be the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to eigenvalues of $\mathbf{A}_H^{\text{sym}}$ that reside in $(-a,0)$
+%% let $\Pi$ be the indicator function for the interval $(a,b)$. 
+%%That is, $\Pi(\lambda)=0$ for $\lambda \in (0,1)$, and $\Pi(\lambda)=0$ otherwise.
+%%\begin{equation*}
+%%	\Pi(\lambda) = \begin{cases}
+%%		1, & \lambda \in (a,b), \\
+%%		0, & \text{otherwise}.
+%%	\end{cases}
+%%\end{equation*}
+%%Using functional calculus to extend the domain of the function $\Pi$ from scalars to matrices, we define
+%%$
+%%\mathbf{\Pi} := \Pi(\mathbf{A}_H^{\text{sym}}).
+%%$
+%%The matrix $\mathbf{\Pi}$ is the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to eigenvalues of $\mathbf{A}_H^{\text{sym}}$ that reside in $(a,b)$. 
+%If we choose $-a < \lambda_\text{min}$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, then $\mathbf{\Pi}$ is the spectral projector onto the eigenspace associated with negative eigenvalues,
+%%and $\mathbf{\Pi}\mathbf{A}_H^{\text{sym}}$ is the ``negative component'' of $\mathbf{A}_H^{\text{sym}}$ in which all the negative eigenvalues are the same but all the positive eigenvalues are replaced by zero.
+Subtracting a negative number from itself twice flips its sign,
+%so the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$ is given by
+%\begin{equation}
+%\label{eq:I_Pi_abs}
+%	|\mathbf{A}_H^{\text{sym}}| = \left(\mathbf{I} - 2\mathbf{\Pi}\right)\mathbf{A}_H^{\text{sym}},
+%\end{equation}
+%where $\mathbf{I}$ is the identity matrix of the appropriate size.
+%By ``matrix absolute value,'' we mean that $|\mathbf{A}_H^{\text{sym}}|$ is the matrix that has the same eigenvectors as $\mathbf{A}_H^{\text{sym}}$, but the eigenvalues of $|\mathbf{A}_H^{\text{sym}}|$ are the absolute values of the eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
+%%\begin{figure}
+%%	\centering
+%%	\includegraphics[scale=0.5]{spd_rational_function.pdf}
+%%	\caption{Rational function $\Pi_\ratord(\lambda)$ for $a=-1$, $b=0$, and $\ratord \in \{1,2,3\}$. As $\ratord$ increases, $\Pi_\ratord$ approaches the indicator function for the set $(a,b)$. }
+%%	\label{fig:spd_rat}
+%%\end{figure}
+%Computing $\mathbf{\Pi}$ directly is difficult, but we may approximate $\mathbf{\Pi}$ by the following high order rational matrix function,
+%%$\Pi_\ratord \approx \Pi$ which is defined as follows:
+%%\begin{equation*}
+%%	 \Pi_\ratord(\lambda) := \frac{1}{1 + T(\lambda)^{2^\ratord} }, \qquad \text{where} \qquad T(\lambda) := \frac{2 \lambda - (b+a)}{b-a}.
+%%\end{equation*}
+%%Here $\ratord$ is a positive integer; the larger $\ratord$ is, the better the approximation $\Pi \approx \Pi_\ratord$. 
+%%%This is illustrated in Figure \ref{fig:spd_rat}. 
+%%Replacing scalar operations with matrix operations, we have an approximation $\mathbf{\Pi}_\ratord \approx \mathbf{\Pi}$ given by
+%\begin{equation}
+%\label{eq:rational343}
+%\mathbf{\Pi} \approx \mathbf{\Pi}_\ratord := \left(\mathbf{I} + \mathbf{T}^{2^\ratord}\right)^{-1},
+%\end{equation}
+%where
+%$
+%\mathbf{T} := T(\mathbf{A}_H^{\text{sym}}) = \left(2 \mathbf{A}_H^{\text{sym}} + a \mathbf{I}\right)/a.
+%$
+%Using $\mathbf{\Pi}_\ratord$ instead of $\mathbf{\Pi}$ in \eqref{eq:I_Pi_abs} allows us to approximate $|\mathbf{A}_H^{\text{sym}}|$.
 First, we compute $\lambda_\text{min}$ using the Lanczos method, and set $a=\gamma \lambda_\text{min}$ and $b=0$, where $\gamma\ge 1$ is a parameter that must be chosen.
-The method is relatively insensitive to $\gamma$ and numerically we find that the method works well for any $\gamma \in [1,2]$. We use $\gamma=1.5$ in our numerical results. 
-Next, we compute $\mathbf{T}$ using fast H-matrix scalar multiplication and addition.
-We form the $\mathbf{T}^{2^\ratord}$ via repeated squaring, $\mathbf{T}^2 = \mathbf{T} \mathbf{T}$, $\mathbf{T}^4 = \mathbf{T}^2 \mathbf{T}^2$, $\mathbf{T}^8 = \mathbf{T}^4 \mathbf{T}^4$ and so on, where each of these matrix squarings is performed using H-matrix multiplication, which is fast.
+%The method is relatively insensitive to $\gamma$ and numerically we find that the method works well for any $\gamma \in [1,2]$. We use $\gamma=1.5$ in our numerical results. 
+%Next, we compute $\mathbf{T}$ using fast H-matrix scalar multiplication and addition.
+We form the $\mathbf{T}^{2^\ratord}$ via repeated squaring using H-matrix multiplication.
 We recommend using $\ratord=1$ or $2$.
@@
 Finally, a positive semi-definite approximation $\mathbf{A}_H^{\text{sym}+}$ of $\mathbf{A}_H^{\text{sym}}$ is formed as
-\begin{equation*}
-\mathbf{A}_H^{\text{sym}+} := \left(\mathbf{I} - 2\mathbf{\Pi}_\ratord\right)\mathbf{A}_H^{\text{sym}},
-\end{equation*}
-where fast H-matrix operations are used to perform the necessary matrix-matrix subtraction and multiplication.
+%\begin{equation*}
+%\mathbf{A}_H^{\text{sym}+} := \left(\mathbf{I} - 2\mathbf{\Pi}_\ratord\right)\mathbf{A}_H^{\text{sym}},
+%\end{equation*}
+%where fast H-matrix operations are used to perform the necessary matrix-matrix subtraction and multiplication.
 Note that for any $\ratord \ge 1$, $\Pi_\ratord(\lambda)> 1/2$ for $\lambda \in (a,b)$, so $\mathbf{A}_H^{\text{sym}+}$ is positive semi-definite\footnote{Some indefiniteness may be introduced by truncation error in the H-matrix operations, but the negative eigenvalues caused by this truncation error are typically extremely small.}.

### 2022-11-08 localpsf/sample_point_batches.py

@@ -3,8 +3,8 @@ import dolfin as dl
 from numba import jit
 from tqdm.auto import tqdm
 
-from .ellipsoid import points_which_are_not_in_ellipsoid_numba, ellipsoids_intersect
 from nalger_helper_functions import *
+from .ellipsoid import points_which_are_not_in_ellipsoid_numba, ellipsoids_intersect
 
 
 def choose_sample_point_batches(num_batches, V, mu_function, Sigma_function, tau, max_candidate_points=None):
@@ -91,8 +91,9 @@ def choose_one_sample_point_batch(mu, Sigma, num_standard_deviations_tau, candid
         p_ind = P_inds_perm.pop()
         p_is_acceptable = True
         for x_ind in X_inds_perm:
-            if ellipsoids_intersect(Sigma_perm[x_ind, :, :], Sigma_perm[p_ind, :, :], mu_perm[x_ind, :],
-                                    mu_perm[p_ind, :], num_standard_deviations_tau):
+            if ellipsoids_intersect(Sigma_perm[x_ind, :, :], Sigma_perm[p_ind, :, :],
+                                    mu_perm[x_ind, :], mu_perm[p_ind, :],
+                                    num_standard_deviations_tau):
                 p_is_acceptable = False
                 break
         if p_is_acceptable:

### 2022-11-08 paper/localpsf.tex
hessian section paper edits

@@
 
-\section{The need for high rank Hessian approximations in inverse problems governed by PDEs}
+\section{Why we need high rank Hessian approximations}
 \label{sec:hessian}
@@
 %Inverse problems are typically ill-posed, in the sense that many different parameter fields may yield predicted observations that agree with the actual observations to within the noise level. 
-Two different frameworks are conventionally used to deal with this ill-posedness: the deterministic framework \cite{DETERMINISTIC}, and the Bayesian probabilistic framework \cite{STUART,TARANTOLA}.
+Conventionally, the inverse problem is formulated using either a deterministic framework \cite{DETERMINISTIC}, or a Bayesian probabilistic framework \cite{STUART,TARANTOLA}.
 In the deterministic framework, one solves an optimization problem to find the parameter that best fits the observations, subject to appropriate regularization.
-In the probabilistic setting, Bayes' theorem combines the observations with prior information to form a posterior distribution over the space of all possible parameter fields, and computations are performed to extract statistical information about the parameter from this posterior.
+In the probabilistic framework, Bayes' theorem is used to combine the observations with prior information to form a posterior distribution over the space of all possible parameter fields, and computations are performed to extract statistical information about the parameter from this posterior.
 The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal to each other under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian''.
@@
 
-\begin{figure}
-	\begin{subfigure}{0.5\textwidth}
-		\includegraphics[scale=0.23]{informed_uninformed_modes_quadratic.pdf}
-		\caption{}
-	\end{subfigure}
-	\begin{subfigure}{0.48\textwidth}
-		\includegraphics[scale=0.1]{misfit_and_reg_hessian_eigenvalues.pdf}
-		\caption{}
-	\end{subfigure}
-	\caption{(a) The Hessian eigenstructure characterizes the directional scalings of the best local quadratic approximation to the objective function in an inverse problem.
+%\begin{figure}
+%	\begin{subfigure}{0.5\textwidth}
+%		\includegraphics[scale=0.23]{informed_uninformed_modes_quadratic.pdf}
+%		\caption{}
+%	\end{subfigure}
+%	\begin{subfigure}{0.48\textwidth}
+%		\includegraphics[scale=0.1]{misfit_and_reg_hessian_eigenvalues.pdf}
+%		\caption{}
+%	\end{subfigure}
+%	\caption{(a) The Hessian eigenstructure characterizes the directional scalings of the best local quadratic approximation to the objective function in an inverse problem.
 The larger an eigenvalue of the Hessian is, the more closely spaced the contours of the local quadratic approximation are in the corresponding eigenvector direction.
@@
 These illustrations are taken from \cite[Chapter 1]{Alger19}, with minor modifications.}
-	\label{fig:hessian_eigenstructure}
-\end{figure}
+%	\label{fig:hessian_eigenstructure}
+%\end{figure}
 
@@
 Using these methods, the required number of Hessian applications (and hence the required number of PDE solves) is proportional to the rank of the low rank approximation.
-However, in many inverse problems of practical interest the data misfit Hessian is high rank, which makes makes it extremely costly to approximate the Hessian using these low rank approximation methods \cite{AmbartsumyanEtAl20,HartlandEtAl21}.
-Examples of inverse problems with high-rank data misfit Hessians include advection dominated advection-diffusion inverse problems, high frequency wave inverse problems, and more generally, all inverse problem in which the observations are highly informative about the parameter.
-The eigenvalues of the data misfit Hessian characterize the sensitivity of the predicted observations to perturbations of the parameter in the corresponding eigenvector directions (see Figure \ref{fig:hessian_eigenstructure}), so more informative data leads to larger eigenvalues and therefore a larger numerical rank.
+However, in many inverse problems of practical interest the rank of the data misfit Hessian is large, which makes makes it extremely costly to approximate the Hessian using these low rank approximation methods \cite{AmbartsumyanEtAl20,HartlandEtAl21}.
+
+Examples of inverse problems with high rank data misfit Hessians include advection dominated advection-diffusion inverse problems, high frequency wave inverse problems, and more generally, all inverse problem in which the observations are highly informative about the parameter.
+The eigenvalues of the data misfit Hessian characterize the sensitivity of the predicted observations to perturbations of the parameter in the corresponding eigenvector directions,
+%(see Figure \ref{fig:hessian_eigenstructure})
+so more informative data leads to larger eigenvalues and therefore a larger numerical rank.
 Roughly speaking, the numerical rank of the data misfit Hessian is the dimension of the subspace of parameter space that is informed by the data.
@@
 In this paper, we form a H-matrix Hessian approximation.
-But rather than use these modern randomized methods to form the H-matrix directly, instead we exploit locality and non-negative integral kernel properties to form a PSF approximation of the data misfit Hessian, then convert the PSF approximation to H-matrix format and combine it with the regularization Hessian using classical H-matrix construction techniques.
+But rather than use these modern methods to form the H-matrix directly, instead we exploit locality and non-negative integral kernel properties to form a PSF approximation of the data misfit Hessian, then convert the PSF approximation to H-matrix format and combine it with the regularization Hessian using classical H-matrix construction techniques.
 Not all data misfit Hessians satisfy the local non-negative integral kernel properties (for example, the wave inverse problem data misfit Hessian has a substantial amount of negative entries in its integral kernel).
@@
 
-\subsection{Integral operators, distributions and impulse responses}
-\label{sec:impulse_response_2}
-
 Let $\Omega \subset \mathbb{R}^\gdim$ be a bounded domain (typically $\gdim=1$, $2$, or $3$). We seek to approximate integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ of the form
@@
 In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define impulse responses (Section \ref{sec:impulse_response_2}), we state the conditions on $\Aop$ that our method requires (Section \ref{sec:conditions_2}), and we detail finite element discretization (Section \ref{sec:finite_element_kernel}).
+
+\subsection{Distributions and impulse responses}
+\label{sec:impulse_response_2}
 

### 2022-11-09 paper/localpsf.tex
edited hessian section based on feedback from tucker and omar

@@
 The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal to each other under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian''.
-The Hessian consists of a data misfit term (the ``data misfit Hessian''), which depends on mismatch between the actual observations and predicted observations based on a candidate parameter, and a regularization or prior term (the ``regularization Hessian'') which does not depend on the observations.
+The Hessian consists of a data misfit term (the ``data misfit Hessian''), which depends on discrepancy between the observations and the associated model predictions, and a regularization or prior term (the ``regularization Hessian'') which does not depend on the observations.
 
@@
 Good Hessian preconditioners reduce the number of iterations required to solve these Newton linear systems with Krylov methods such as conjugate gradient.
-In the Bayesian setting, the inverse of the Hessian is the covariance of the best local Gaussian approximation of the posterior.
+In the Bayesian setting, the inverse of the Hessian is the covariance of the local Gaussian approximation of the posterior generated by truncating the Taylor series of the negative log posterior after the quadratic term.
 Hessian-based local Gaussian approximations of the posterior may be used as surrogate models of the posterior, or may be used as proposal distributions for Markov chain Monte-Carlo (MCMC) methods for drawing samples from the posterior.
@@
 Using these methods, the required number of Hessian applications (and hence the required number of PDE solves) is proportional to the rank of the low rank approximation.
-However, in many inverse problems of practical interest the rank of the data misfit Hessian is large, which makes makes it extremely costly to approximate the Hessian using these low rank approximation methods \cite{AmbartsumyanEtAl20,HartlandEtAl21}.
-
-Examples of inverse problems with high rank data misfit Hessians include advection dominated advection-diffusion inverse problems, high frequency wave inverse problems, and more generally, all inverse problem in which the observations are highly informative about the parameter.
+Typically, low rank approximation methods are justified by arguing that the numerical rank of the data misfit Hessian is insensitive to the dimension of the discretized parameter.
+As the mesh used to discretize the parameter is refined, the required number of PDE solves remains the same.
+However, in many inverse problems of practical interest the numerical rank of the data misfit Hessian, while mesh independent, is still large, which makes it costly to approximate the Hessian using low rank approximation methods \cite{AmbartsumyanEtAl20,HartlandEtAl21}.
+
+Examples of inverse problems with high rank data misfit Hessians include advection dominated advection-diffusion inverse problems, high frequency wave inverse problems, and more generally, all inverse problems in which the observations highly inform the parameter.
 The eigenvalues of the data misfit Hessian characterize the sensitivity of the predicted observations to perturbations of the parameter in the corresponding eigenvector directions,
@@
 Roughly speaking, the numerical rank of the data misfit Hessian is the dimension of the subspace of parameter space that is informed by the data.
-The numerical rank of the regularization preconditioned data misfit Hessian may be reduced by increasing the strength of the regularization, but this is throwing away useful information---components of the parameter that could be learned from the observations are instead reconstructed based on the regularization.
-Hence, low rank approximation methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation.
-The best case scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost) \cite{AlexanderianGloorGhattas16,Alger19,AlgerEtAl17}.
+The numerical rank of the regularization preconditioned data misfit Hessian may be reduced by increasing the strength of the regularization, but this throws away useful information---components of the parameter that could be learned from components of the observations that have a large signal to noise ratio are instead reconstructed based on the regularization \cite{AlgerEtAl17}.
+Hence, low rank approximation methods suffer from a ``data predicament''---if the data highly inform the parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation \cite{AlexanderianGloorGhattas16,Alger19}.
+%The best case scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost) \cite{AlexanderianGloorGhattas16,Alger19,AlgerEtAl17}.
 Matrix-free high rank Hessian approximation methods are needed.
 
-Recently there have been improvements in matrix-free H-matrix construction methods in which the operator is applied to structured random vectors, and the response of the operator to those random vectors is then processed to construct the H-matrix  \cite{LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}.
+Recently there have been improvements in matrix-free H-matrix construction methods in which the operator is applied to structured random vectors, and the response of the operator to those random vectors is then processed to construct an H-matrix approximation  \cite{LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}.
 These methods have been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{AmbartsumyanEtAl20}. 
-These methods are asymptotically scalable, but in practice a large number of matrix vector products are required. 
+These methods are asymptotically scalable in the sense that the required number of matrix-vector products grows nearly linearly in the dimension of the discretized parameter and linearly in the rank of the off-diagonal blocks.
+But despite asymptotic scalability, from a practical perspective the required number of matrix-vector products is large (e.g., hundreds to thousands).
 %Improving the peeling process to reduce its cost is currently an area of active research \cite{PEELINGRESEARCH}.
@@
 Matrix-free H-matrix construction methods are currently an area of active research (see, e.g., \cite{LevittMartinsson}), so more efficient matrix-free H-matrix methods may be discovered in the future.
-In this paper, we form a H-matrix Hessian approximation.
-But rather than use these modern methods to form the H-matrix directly, instead we exploit locality and non-negative integral kernel properties to form a PSF approximation of the data misfit Hessian, then convert the PSF approximation to H-matrix format and combine it with the regularization Hessian using classical H-matrix construction techniques.
+In this paper, we form an H-matrix Hessian approximation.
+But rather than use these modern methods to form the H-matrix directly, instead we exploit locality and non-negative integral kernel properties to form a PSF approximation of the data misfit Hessian, then convert the PSF approximation to H-matrix format and combine it with the regularization Hessian using classical H-matrix techniques.
 Not all data misfit Hessians satisfy the local non-negative integral kernel properties (for example, the wave inverse problem data misfit Hessian has a substantial amount of negative entries in its integral kernel).
-However, many data misfit Hessians of practical interest do satisfy these properties (either exactly or approximately), and our method is effective for approximating these Hessians.
+However, many data misfit Hessians of practical interest do satisfy these properties (either exactly or approximately), and our method is targeted at approximating these Hessians.
 

### 2022-11-14 paper/localpsf.tex
add references to hessian section based on discussion today with Noemi and Tucker

@@
 %Inverse problems are typically ill-posed, in the sense that many different parameter fields may yield predicted observations that agree with the actual observations to within the noise level. 
-Conventionally, the inverse problem is formulated using either a deterministic framework \cite{DETERMINISTIC}, or a Bayesian probabilistic framework \cite{STUART,TARANTOLA}.
-In the deterministic framework, one solves an optimization problem to find the parameter that best fits the observations, subject to appropriate regularization.
-In the probabilistic framework, Bayes' theorem is used to combine the observations with prior information to form a posterior distribution over the space of all possible parameter fields, and computations are performed to extract statistical information about the parameter from this posterior.
-The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal to each other under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian''.
-The Hessian consists of a data misfit term (the ``data misfit Hessian''), which depends on discrepancy between the observations and the associated model predictions, and a regularization or prior term (the ``regularization Hessian'') which does not depend on the observations.
+Conventionally, the inverse problem is formulated using either a deterministic framework \cite{AsterBorchersThurber18,Vogel02},
+%Guy Chavent, Nonlinear Least Squares for Inverse Problems, Springer, 2009.
+or a Bayesian probabilistic framework \cite{KaipioSomersalo05,Stuart10a,Tarantola05}.
+In the deterministic framework, one solves an optimization problem to find the parameter that best fits the observations, subject to appropriate regularization \cite{EnglHankeNeubauer96}.
+In the probabilistic framework, Bayes' theorem combines the observations with prior information to form a posterior distribution over the space of all possible parameter fields, and computations are performed to extract statistical information about the parameter from this posterior.
+The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal to each other under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian.'' The Hessian consists of a data misfit term (the ``data misfit Hessian''), which depends on discrepancy between the observations and the associated model predictions, and a regularization or prior term (the ``regularization Hessian'') which does not depend on the observations.
+Form more details on the Hessian, see \cite{Alger19,GhattasWillcox21,VillaPetraGhattas21}.
 
 Hessian approximations and preconditioners are highly desirable because the Hessian is central to efficient solution of inverse problems in both the deterministic and Bayesian settings.
-When solving the deterministic optimization problem with Newton-type methods, the Hessian is the coefficient operator for the linear system that must be solved at every Newton iteration.
-Good Hessian preconditioners reduce the number of iterations required to solve these Newton linear systems with Krylov methods such as conjugate gradient.
-In the Bayesian setting, the inverse of the Hessian is the covariance of the local Gaussian approximation of the posterior generated by truncating the Taylor series of the negative log posterior after the quadratic term.
-Hessian-based local Gaussian approximations of the posterior may be used as surrogate models of the posterior, or may be used as proposal distributions for Markov chain Monte-Carlo (MCMC) methods for drawing samples from the posterior.
+When solving the deterministic optimization problem with Newton-type methods, the Hessian is the coefficient operator for the linear system that must be solved or approximately solved at every Newton iteration \cite{NocedalWright99}.
+Good Hessian preconditioners reduce the number of iterations required to solve these Newton linear systems with Krylov methods such as conjugate gradient \cite{Saad03}.
+In the Bayesian setting, the inverse of the Hessian is the covariance of a local Gaussian approximation of the posterior.
+Such a Gaussian distribution can be used as a direct approximations of the posterior, or as a proposal for Markov chain Monte-Carlo (MCMC) methods for drawing samples from the posterior.
+For instance, see \cite{KimEtAl21,PetraEtAl14} and the references therein.
 
@@
 Owing to the implicit dependence of predicted observations on the parameter, entries of the Hessian are not easily accessible.
-Rather, the Hessian may be applied to a vector via a computational process that involves solving a pair of forward and adjoint PDEs which are linearizations of the original PDE \cite{Alger19,GhattasWillcox21}.
-The most popular matrix-free Hessian approximation methods are based on low rank approximation of either the data misfit Hessian, or the data misfit Hessian preconditioned by the regularization Hessian  \cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15}.
-Krylov methods or randomized methods are typically used to construct these low rank matrix approximations by applying the Hessian to vectors \cite{Cheng05,HalkoMartinssonTropp11}.
-Using these methods, the required number of Hessian applications (and hence the required number of PDE solves) is proportional to the rank of the low rank approximation.
-Typically, low rank approximation methods are justified by arguing that the numerical rank of the data misfit Hessian is insensitive to the dimension of the discretized parameter.
+Rather, the Hessian may be applied to a vector via a computational process that involves solving a pair of forward and adjoint PDEs which are linearizations of the original PDE \cite{GhattasWillcox21,Gunzburger03}.
+The most popular matrix-free Hessian approximation methods are based on low rank approximation of either the data misfit Hessian, or the data misfit Hessian preconditioned by the regularization Hessian, e.g., \cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15}.
+Krylov methods such as Lanczos or randomized methods \cite{Cheng05,HalkoMartinssonTropp11} are typically used to construct these low rank approximations by applying the Hessian to vectors.
+Using these methods, the required number of Hessian applications (and hence the required number of PDE solves) is typically proportional to the rank of the low rank approximation.
+Low rank approximation methods are justified by arguing that the numerical rank of the data misfit Hessian is insensitive to the dimension of the discretized parameter.
 As the mesh used to discretize the parameter is refined, the required number of PDE solves remains the same.
-However, in many inverse problems of practical interest the numerical rank of the data misfit Hessian, while mesh independent, is still large, which makes it costly to approximate the Hessian using low rank approximation methods \cite{AmbartsumyanEtAl20,HartlandEtAl21}.
+However, in many inverse problems of practical interest the numerical rank of the data misfit Hessian, while mesh independent, is still large, which makes it costly to approximate the Hessian using low rank approximation methods \cite{AmbartsumyanEtAl20,HartlandEtAl21,IsaacEtAl15}.
 
@@
 Recently there have been improvements in matrix-free H-matrix construction methods in which the operator is applied to structured random vectors, and the response of the operator to those random vectors is then processed to construct an H-matrix approximation  \cite{LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}.
-These methods have been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{AmbartsumyanEtAl20}. 
+These methods have been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{AmbartsumyanEtAl20,HartlandEtAl22}. 
 These methods are asymptotically scalable in the sense that the required number of matrix-vector products grows nearly linearly in the dimension of the discretized parameter and linearly in the rank of the off-diagonal blocks.
@@
 %This is an area of active research, so more efficient and practical matrix-free H-matrix construction methods based on the peeling process may be discovered in the future.
-Matrix-free H-matrix construction methods are currently an area of active research (see, e.g., \cite{LevittMartinsson}), so more efficient matrix-free H-matrix methods may be discovered in the future.
+Matrix-free H-matrix construction methods are currently an area of active research (see, e.g., \cite{LevittMartinsson22}), so more efficient matrix-free H-matrix methods may be discovered in the future.
 In this paper, we form an H-matrix Hessian approximation.

### 2022-11-17 paper/localpsf.tex
paper edits based on discussion with Noemi

@@
 The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal to each other under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian.'' The Hessian consists of a data misfit term (the ``data misfit Hessian''), which depends on discrepancy between the observations and the associated model predictions, and a regularization or prior term (the ``regularization Hessian'') which does not depend on the observations.
-Form more details on the Hessian, see \cite{Alger19,GhattasWillcox21,VillaPetraGhattas21}.
+For more details on the Hessian, see \cite{Alger19,GhattasWillcox21,VillaPetraGhattas21}.
 
 Hessian approximations and preconditioners are highly desirable because the Hessian is central to efficient solution of inverse problems in both the deterministic and Bayesian settings.
-When solving the deterministic optimization problem with Newton-type methods, the Hessian is the coefficient operator for the linear system that must be solved or approximately solved at every Newton iteration \cite{NocedalWright99}.
+When solving the deterministic optimization problem with Newton-type methods, the Hessian is the coefficient operator for the linear system that must be solved or approximately solved at every Newton iteration.
 Good Hessian preconditioners reduce the number of iterations required to solve these Newton linear systems with Krylov methods such as conjugate gradient \cite{Saad03}.
 In the Bayesian setting, the inverse of the Hessian is the covariance of a local Gaussian approximation of the posterior.
-Such a Gaussian distribution can be used as a direct approximations of the posterior, or as a proposal for Markov chain Monte-Carlo (MCMC) methods for drawing samples from the posterior.
+Such a Gaussian distribution can be used as a direct approximations of the posterior, or as a proposal for Markov chain Monte-Carlo methods for drawing samples from the posterior.
 For instance, see \cite{KimEtAl21,PetraEtAl14} and the references therein.
@@
 Low rank approximation methods are justified by arguing that the numerical rank of the data misfit Hessian is insensitive to the dimension of the discretized parameter.
-As the mesh used to discretize the parameter is refined, the required number of PDE solves remains the same.
-However, in many inverse problems of practical interest the numerical rank of the data misfit Hessian, while mesh independent, is still large, which makes it costly to approximate the Hessian using low rank approximation methods \cite{AmbartsumyanEtAl20,HartlandEtAl21,IsaacEtAl15}.
-
-Examples of inverse problems with high rank data misfit Hessians include advection dominated advection-diffusion inverse problems, high frequency wave inverse problems, and more generally, all inverse problems in which the observations highly inform the parameter.
+This means that the required number of PDE solves remains the same as the mesh used to discretize the parameter is refined.
+However, in many inverse problems of practical interest the numerical rank of the data misfit Hessian, while mesh independent, is still large, which makes it costly to approximate the Hessian using low rank approximation methods \cite{AmbartsumyanEtAl20,BuiEtAl13,IsaacEtAl15}.
+
+Examples of inverse problems with high rank data misfit Hessians include advection dominated advection-diffusion inverse problems \cite{AkcelikBirosDraganescuEtAl05}\cite[Chapter 5]{Flath13}, high frequency wave inverse problems \cite{BuiEtAl13}, and more generally, all inverse problems in which the observations highly inform the parameter.
 The eigenvalues of the data misfit Hessian characterize the sensitivity of the predicted observations to perturbations of the parameter in the corresponding eigenvector directions,
 %(see Figure \ref{fig:hessian_eigenstructure})
-so more informative data leads to larger eigenvalues and therefore a larger numerical rank.
+so more informative data leads to larger eigenvalues and therefore a larger numerical rank \cite{AlexanderianGloorGhattas16}\cite[Section 1.4 and Chapter 4]{Alger19}.
 Roughly speaking, the numerical rank of the data misfit Hessian is the dimension of the subspace of parameter space that is informed by the data.
-The numerical rank of the regularization preconditioned data misfit Hessian may be reduced by increasing the strength of the regularization, but this throws away useful information---components of the parameter that could be learned from components of the observations that have a large signal to noise ratio are instead reconstructed based on the regularization \cite{AlgerEtAl17}.
-Hence, low rank approximation methods suffer from a ``data predicament''---if the data highly inform the parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation \cite{AlexanderianGloorGhattas16,Alger19}.
+The numerical rank of the regularization preconditioned data misfit Hessian may be reduced by increasing the strength of the regularization, but this throws away useful information---components of the parameter that could be learned from components of the observations that have a large signal to noise ratio are instead reconstructed based on the regularization \cite[Section 4]{AlgerEtAl17}\cite[Chapters 1 and 7]{Vogel02}.
+Hence, low rank approximation methods suffer from a predicament---if the data highly inform the parameter and the regularization is chosen appropriately, then a large number of operator applies are required to form an accurate approximation of the Hessian using low rank approximation methods.
 %The best case scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost) \cite{AlexanderianGloorGhattas16,Alger19,AlgerEtAl17}.
-Matrix-free high rank Hessian approximation methods are needed.
-
-Recently there have been improvements in matrix-free H-matrix construction methods in which the operator is applied to structured random vectors, and the response of the operator to those random vectors is then processed to construct an H-matrix approximation  \cite{LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}.
-These methods have been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{AmbartsumyanEtAl20,HartlandEtAl22}. 
-These methods are asymptotically scalable in the sense that the required number of matrix-vector products grows nearly linearly in the dimension of the discretized parameter and linearly in the rank of the off-diagonal blocks.
+High rank Hessian approximation methods are needed.
+
+Recently there have been improvements in matrix-free H-matrix construction methods in which an operator is applied to structured random vectors, and the response of the operator to those random vectors is processed to construct an H-matrix approximation  \cite{LevittMartinsson22,LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}.
+These methods (which we do not use here) have been used to approximate Hessians in PDE constrained inverse problems \cite{AmbartsumyanEtAl20,HartlandEtAl22}. 
+Although the required number of matrix-vector products grows nearly linearly in the discretized parameter dimension and linearly in the rank of the off-diagonal blocks, the required number of matrix-vector products is still large (e.g., hundreds to thousands).
+%These methods are asymptotically scalable in the sense that the required number of matrix-vector products grows nearly linearly in the dimension of the discretized parameter and linearly in the rank of the off-diagonal blocks.
 But despite asymptotic scalability, from a practical perspective the required number of matrix-vector products is large (e.g., hundreds to thousands).
@@
 %This is an area of active research, so more efficient and practical matrix-free H-matrix construction methods based on the peeling process may be discovered in the future.
-Matrix-free H-matrix construction methods are currently an area of active research (see, e.g., \cite{LevittMartinsson22}), so more efficient matrix-free H-matrix methods may be discovered in the future.
-In this paper, we form an H-matrix Hessian approximation.
-But rather than use these modern methods to form the H-matrix directly, instead we exploit locality and non-negative integral kernel properties to form a PSF approximation of the data misfit Hessian, then convert the PSF approximation to H-matrix format and combine it with the regularization Hessian using classical H-matrix techniques.
+%Matrix-free H-matrix construction methods are currently an area of active research (see, e.g., \cite{LevittMartinsson22}), so more efficient matrix-free H-matrix methods may be discovered in the future.
+In this paper, we also form an H-matrix Hessian approximation.
+However, to reduce the required number of operator actions, we first form a PSF approximation of the data misfit Hessian by exploiting locality and non-negative integral kernel properties, then form the H-matrix using classical H-matrix techniques.
 Not all data misfit Hessians satisfy the local non-negative integral kernel properties (for example, the wave inverse problem data misfit Hessian has a substantial amount of negative entries in its integral kernel).

### 2023-01-17 paper/localpsf.tex
commit after discussing with Noemi on zoom
made skeleton figures for advection diffusion with Noemi and Tucker

@@
 \label{sec:numerical_results}
+
+\subsection{Example 1: Inversion for the basal sliding coefficient in an ice sheet flow problem}
 
@@
 Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), or the regularization Hessian (REG).
-Right (\ref{fig:condition_number}): condition number, $\kappa$, for $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$ for these different preconditioners, and no preconditioner (NONE).
-All operators are evaluated at the $\bm{\basalfriction}$ that solves the inverse problem for $\gamma=7.3 \times 10^3$, as determined by the Morozov discrepancy principle.}
+Right (\ref{fig:condition_number}): condition number for $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$ for these different preconditioners, and no preconditioner (NONE).
+All operators are evaluated at the $\bm{\basalfriction}$ that solves the inverse problem for $\gamma=7.3 \times 10^3$, as determined by the Morozov discrepancy principle.
+COND(H) LABELS!}
+	\label{fig:stokes_spectrum}
 \end{figure}
@@
 The condition numbers are similar for both $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$, demonstrating that the preconditioner built based on $\bm{H}_\text{gn}$ is a good preconditioner for $\bm{H}$.
+
+\subsection{Example 2: Inversion for the initial condition in an advection-diffusion problem}
+
+%INTRODUCE EQUATIONS: REWRITE HERE vvvvvvv
+
+%The parameter-to-observable map $\iFF \,\ipar := \obsop\, u(\ipar)$ maps an initial condition $\ipar \in L^2(\D)$ to pointwise spatiotemporal observations of the concentration field $u({\vec x},t)$ through solution of the advection-diffusion equation given by
+%
+%$$
+%\begin{split}
+%u_t - \kappa\Delta u + {\vec v} \cdot \nabla u &= 0     & \quad \text{in } \D\times(0,T),\\
+%u(\cdot, 0) &= \ipar & \quad \text{in } \D,\\
+%\kappa \nabla u\cdot {\vec{n}} &= 0     & \quad \text{on } \partial\D \times (0,T).
+%\end{split}
+%$$
+%
+%Here, $\D \subset \R^d$ ($d \in \{2, 3\}$) is a bounded domain, $\kappa > 0$ is the diffusion coefficient and $T > 0$ is the final
+%time. The velocity field
+%$\vec{v}$ is computed by solving the following steady-state
+%Navier-Stokes equation with the side walls driving the flow:
+%
+%$$
+%\begin{aligned}
+%- \frac{1}{\operatorname{Re}} \Delta {\vec v} + \nabla q + {\vec v} \cdot \nabla {\vec v} &= 0 &\quad&\text{ in }\D,\\
+%\nabla \cdot {\vec v} &= 0 &&\text{ in }\D,\\
+%{\vec v} &= {\vec g} &&\text{ on } \partial\D.
+%\end{aligned}
+%$$
+%
+%Here, $q$ is pressure, $\text{Re}$ is the Reynolds number. The Dirichlet boundary data
+%${\vec g} \in \R^d$ is given by 
+%${\vec g} = {\vec e}_2$ on the left wall of the domain, 
+%${\vec g}=-{\vec e}_2$ on the right wall,  and ${\vec g} = {\vec 0}$ everywhere else.
+
+% INTRODUCE EQUATIONS: REWRITE HERE ^^^^^
+
+\begin{figure}
+	\begin{subfigure}{0.24\textwidth}
+		\includegraphics[scale=0.25]{adv_ic_true_tmp.png}
+		%		\caption{}
+		%		\label{fig:newtoniter2}
+	\end{subfigure}
+	\begin{subfigure}{0.24\textwidth}
+		\includegraphics[scale=0.25]{velocity_tmp.png}
+		%		\caption{}
+		%		\label{fig:newtoniter1}
+	\end{subfigure}
+	\begin{subfigure}{0.24\textwidth}
+		\includegraphics[scale=0.25]{adv_final_condition_true_tmp.png}
+		%		\caption{}
+		%		\label{fig:newtoniter3}
+	\end{subfigure}
+	\begin{subfigure}{0.24\textwidth}
+		\includegraphics[scale=0.25]{adv_reconstruction_tmp.png}
+		%		\caption{}
+		%		\label{fig:newtoniter4}
+	\end{subfigure}
+	%\begin{subfigure}{0.05\textwidth}
+	%	\includegraphics[scale=0.1]{mColorBar.png}
+	%\end{subfigure}
+	\caption{(a) true initial condition. (b) velocity field. (c) observed concentration at final time. (d) reconstruction of initial condition. black=0, white=1}
+	\label{fig:adv_inversion1} 
+\end{figure}
+
+\begin{figure}
+	\begin{center}
+		\includegraphics[scale=0.75]{adv_vary_kappa_tf_tmp.png}
+	\end{center}
+	\caption{Impulse responses for various diffusivities kappa and fixed simulation time Tf=0.5 (top) and simulation times with fixed kappa=1e-3 (bot).
+Colorbar is normalized independently for each subfigure}
+	\label{fig:adv_vary_kappa_tf}
+\end{figure}
+
+\begin{figure}
+	\begin{subfigure}{0.49\textwidth}
+		\centering
+		\begin{tikzpicture}[baseline, trim axis right]
+		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
+		ymode=log,
+		ymin=1e-11,
+		ymax=2,
+		xmax=750,
+		axis x line*=bottom,
+		axis y line*=left,
+		xtick={0, 200, 400, 600},
+		xlabel={CG iteration},
+		ylabel={$||\bm{b}-\bm{H} \bm{x}_k||/||\bm{b}||$},
+		title={CG convergence},
+		legend style={at={(1.1,1)},anchor=north east},
+		%legend pos=north east,
+		legend style={font=\small}
+		]
+		\addlegendentry{Reg}
+		\addplot[dotted, line width=0.5mm] table {cg_relres_reg.dat};
+		\addlegendentry{None}
+		\addplot[dash dot, line width=0.5mm] table {cg_relres_none.dat};
+		\addlegendentry{PSF (1)}
+		\addplot[blue, line width=0.5mm] table {cg_relres_p1.dat};
+		\addlegendentry{PSF (5)}
+		\addplot[red, line width=0.5mm] table {cg_relres_p5.dat};
+		\addlegendentry{PSF (25)}
+		\addplot[purple, line width=0.5mm] table {cg_relres_p25.dat};
+		\end{axis}
+		\end{tikzpicture}
+		\caption{}
+		\label{fig:adv_cg_convergence}
+	\end{subfigure}
+	\begin{subfigure}{0.49\textwidth}
+		\centering
+		\begin{tikzpicture}[baseline, trim axis right]
+		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
+		xmode=log,
+		ymode=log,
+		xlabel={$\gamma$, regularization parameter},
+		ylabel={\# of CG iterations},
+		axis y line*=right,
+		axis x line*=bottom,
+		ylabel near ticks, yticklabel pos=right,
+		%ylabel near ticks, ylabel shift={-6pt},
+		%legend pos=north east,
+		title={Regularization robustness}
+		%title={CG iterations to achieve $||H x-b||/||b||<10^{-6}$}
+		]
+		%\addlegendentry{None}
+		\addplot[dash dot, line width=0.5mm] table[x=gamma, y=none] {CGitsvsGamma.dat};
+		%\addlegendentry{REG}
+		\addplot[dotted, line width=0.5mm] table[x=gamma, y=reg] {CGitsvsGamma.dat};
+		%\addlegendentry{PSF (1)}
+		\addplot[blue, line width=0.5mm] table[x=gamma, y=psf1] {CGitsvsGamma.dat};
+		%\addlegendentry{PSF (5)}
+		\addplot[red, line width=0.5mm] table[x=gamma, y=psf5] {CGitsvsGamma.dat};
+		%\addlegendentry{PSF (25)}
+		\addplot[purple, line width=0.5mm] table[x=gamma, y=psf25] {CGitsvsGamma.dat};
+		\end{axis}
+		\end{tikzpicture}
+		\caption{}
+		\label{fig:adv_kappa_sweep}
+	\end{subfigure}
+	\caption{Left (\ref{fig:adv_cg_convergence}) shows the convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
+random entries drawn from the standard normal distribution.
+BLAH BLAH BLAH
+		%		The y-axis is $||b-Hx_k||/||b||$, where $x_k$ is the $k$th PCG iterate. 
+		Here we use $\gamma=9999 \times 10^9$, which is determined by the Morozov discrepancy principle.
+Right (\ref{fig:adv_kappa_sweep}) shows the number of PCG iterations required to achieve $||\bm{b}-\bm{H} \bm{x}_k||/||\bm{b}|| < 10^{-6}$, where $\bm{x}_k$ is the $k$th PCG iterate, for a range of different $\kappa$.
+Results in these figures are shown for our preconditioner with 5, 15, and 30 batches (PSF (5), PSF (15), and PSF(30), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
+The preconditioner is constructed using $\bm{H}_\text{gn}$.}
+	\label{fig:adv_krylov_convergence}
+\end{figure}
+
+\begin{figure}
+	\begin{subfigure}{0.59\textwidth}
+		\begin{center}
+			\begin{tikzpicture}[baseline, trim axis right]
+			\begin{axis}[width=0.99\linewidth, height=0.8\linewidth,
+			ymode=log,
+			xmin=-50,
+			xmax=1450,
+			ymax=1.e4,
+			xlabel={$k$, generalized eigenvalue \#},
+			ylabel={$\lambda_{k}$},
+			xtick={0, 300, 600, 900, 1200},
+			%legend pos=north east,
+			legend style={font=\small},
+			ylabel near ticks, ylabel shift={-6pt},
+			axis x line*=bottom,
+			axis y line*=left,
+			legend style={at={(0.95,1)},anchor=north east},
+			title={Generalized eigenvalues, $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$}
+			]
+			\addlegendentry{REG}
+			\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues.dat};
+			\addlegendentry{PSF (1)}
+			\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues.dat};
+			\addlegendentry{PSF (5)}
+			\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues.dat};
+			\addlegendentry{PSF (25)}
+			\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues.dat};
+			\end{axis}
+			\end{tikzpicture}
+		\end{center}
+		\caption{}
+		\label{fig:adv_generalized_eigenvalues}
+	\end{subfigure}
+	\begin{subfigure}{0.39\textwidth}
+		\begin{center}
+			\begingroup
+			\setlength{\tabcolsep}{4pt}
+			\renewcommand{\arraystretch}{1.25}
+			\begin{tabular}{c| c c}
+				& $\kappa(\bm{\preconditioner}^{-1} \bm{H})$\\
+				\hline 
+				REG        & 7.02e+3 \\
+				NONE       & 8.71e+2 \\
+				PSF $(5)$  & 1.75e+1 \\
+				PSF $(15)$  & 1.39e+1 \\
+				PSF $(30)$ & 1.03e+1
+			\end{tabular}
+			\endgroup
+		\end{center}
+		\caption{}
+		\label{fig:adv_condition_number}
+	\end{subfigure}
+	\caption{Left (\ref{fig:adv_generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_k = \lambda_k \bm{\preconditioner} \bm{u}_k$.
+Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 5, 15, and 30 batches (PSF (5), PSF (15), and PSF (30), respectively), or the regularization Hessian (REG).
+Right (\ref{fig:adv_condition_number}): condition number for $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$ for these different preconditioners, and no preconditioner (NONE).
+COND(H) LABELS!}
+	\label{fig:adv_spectrum}
+\end{figure}
+
+\begin{itemize}
+	\item Picture of velocity field, initial condition, final condition, reconstruction (1) (OKOK)
+	\item Impulse responses: vary time and Peclet number (3) (OKOK)
+	\item CG convergence (1) CG iters vs Peclet number (Morozov) (2) (OKOK)
+	\item Preconditioned spectrum, condition numbers? (1)
+\end{itemize}
 

### 2023-01-20 paper/localpsf.tex
work on the ice section

@@
 
-The true parameter field, $\basalfriction_\text{true}$, and corresponding velocity field, $\velocity_\text{true}$, are shown in Figure \ref{fig:true_beta_u}.
+The observations are generated by adding $1\%$ multiplicative
+Gaussian noise to the tangential component of the velocity field
+restricted to the top surface of the geometry. The true basal sliding
+coefficient field and the true velocity field , which is obtained by
+solving~\eqref{Stokeseqn}, are shown in Figure~\ref{fig:true_beta_u}.
+
+To reconstruct the basal sliding parameter field $\basalfriction$, we
+formulate an inverse problem governed by the Stokes
+problem~\eqref{Stokeseqn}. In particular this problem is formulated as
+a nonlinear least squares optimization problem, whose cost function
+consists of a misfit term (between the observations and model output)
+and a bi-Laplacian regularization term
+following~\cite{IsaacPetraStadlerEtAl15}. The regularization is
+centered at the constant function $q_0(x) = 10.5$. To mitigate
+boundary effects we use a constant coefficient Robin boundary
+condition as in~\cite{Roininen14}. The parameters are chosen so that
+the Green's function of the Hessian of the regularization has a
+characteristic length of $0.25$ of the radius of the domain. For the
+specific setup, we refer the reader
+to~[sec.22]\cite{VillaPetraGhattas21}. We choose the regularization
+parameter, $\gamma$, that controls the overall strength of the
+regularization using Morozov's discrepancy principle. The optimal
+value of $\gamma$ is $7.3 \times 10^3$, which is used for all the
+numerical results except for Figure~\ref{fig:gamma_sweep}, where we
+vary $\gamma$ to study how changing $\gamma$ impacts the effectiveness
+of the proposed preconditioner.
+
+%% from here we summarize --------
+%The true parameter field, $\basalfriction_\text{true}$, and corresponding velocity field, $\velocity_\text{true}$, are shown in Figure \ref{fig:true_beta_u}.
 We generate synthetic observations by restricting the tangential component of $\velocity_\text{true}$ to the top surface to get
-\begin{equation*}
-\observations_\text{true} = \tangentop \velocity_\text{true} |_{\Gamma_t}.
-\end{equation*}
-We add $1\%$ multiplicative Gaussian noise to all components of the finite element vector, $\bm{\observations}_\text{true}$, corresponding to the function $\observations_\text{true}$, which yields noisy synthetic observations, $\observations_\text{obs}$ which are used for the inversion.
+%\begin{equation*}
+%\observations_\text{true} = \tangentop \velocity_\text{true} |_{\Gamma_t}.
+%\end{equation*}
+%We add $1\%$ multiplicative Gaussian noise to all components of the finite element vector, $\bm{\observations}_\text{true}$, corresponding to the function $\observations_\text{true}$, which yields noisy synthetic observations, $\observations_\text{obs}$ which are used for the inversion.
 That is,
-\begin{equation*}
-\left(\bm{\observations}_\text{obs}\right)_j := \left(\bm{\observations}_\text{true}\right)_j + 0.01 \cdot \boldsymbol{\xi}_j \cdot \left|\left(\bm{\observations}_\text{true}\right)_j\right|,
-\end{equation*}
-where $j$ ranges over all entries of the vector of coefficients representing the finite element functions $\observations_\text{true}$ and $\observations_\text{obs}$, and $\boldsymbol{\xi}_j$ are independent and identically distributed (i.i.d) random numbers drawn from the standard normal distribution.
+%\begin{equation*}
+%\left(\bm{\observations}_\text{obs}\right)_j := \left(\bm{\observations}_\text{true}\right)_j + 0.01 \cdot \boldsymbol{\xi}_j \cdot \left|\left(\bm{\observations}_\text{true}\right)_j\right|,
+%\end{equation*}
+%where $j$ ranges over all entries of the vector of coefficients representing the finite element functions $\observations_\text{true}$ and $\observations_\text{obs}$, and $\boldsymbol{\xi}_j$ are independent and identically distributed (i.i.d) random numbers drawn from the standard normal distribution.
 We also define the noise function, $\xi$, to be the finite element function with $j$th vector entry $\boldsymbol{\xi}_j$.
-
-The reconstructed parameter field $\basalfriction$ is found as the solution to an optimization problem of the following form:
-\begin{equation}
-	\label{eq:deterministic_optimization_problem}
-	\min_{\basalfriction} \quad J(\basalfriction) := J_d(\basalfriction) + J_r(\basalfriction).
-\end{equation}
-The first term in the objective function is
-\begin{equation*}
-	J_d(\basalfriction) := \frac{1}{2}\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS.
-\end{equation*}
-We call this term the data misfit, because it measures the difference between the observed data and the predicted data based on a candidate parameter $\basalfriction$.
+%
+%The reconstructed parameter field $\basalfriction$ is found as the solution to an optimization problem of the following form:
+%\begin{equation}
+%	\label{eq:deterministic_optimization_problem}
+%	\min_{\basalfriction} \quad J(\basalfriction) := J_d(\basalfriction) + J_r(\basalfriction).
+%\end{equation}
+%The first term in the objective function is
+%\begin{equation*}
+%	J_d(\basalfriction) := \frac{1}{2}\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS.
+%\end{equation*}
+%We call this term the data misfit, because it measures the difference between the observed data and the predicted data based on a candidate parameter $\basalfriction$.
 Here $\velocity = \velocity(\basalfriction)$ denotes the velocity field solving $\ref{Stokeseqn}$ for the given parameter $\basalfriction$.
 The second term,
-\begin{equation*}
-	J_r(\basalfriction) := \frac{1}{2}\int_{\Omega}\left| \mathcal{K}(\basalfriction - \basalfriction_0)\right|^2\,dx,
-\end{equation*}
-is a Bilaplacian regularization term.
+%\begin{equation*}
+%	J_r(\basalfriction) := \frac{1}{2}\int_{\Omega}\left| \mathcal{K}(\basalfriction - \basalfriction_0)\right|^2\,dx,
+%\end{equation*}
+%is a Bilaplacian regularization term.
 Specifically, $\basalfriction_0$ is the constant function $\basalfriction_0(x)=10.5$, and $\mathcal{K}$ is the inverse of the solution operator for the following elliptic PDE:
-\begin{align*}
--\gamma\Delta u +\delta\,u&=\regrhs\,\,\,\,\text{ in }\Omega,\\
-\gamma \nabla u\cdot\normalvec+\regrobincoeff\,u&=0\,\,\,\text{ on }\partial\Omega.
-\end{align*}
-Recall that $\Omega$ is the 2D flattened version of the basal surface $\Gamma_b$.
+%\begin{align*}
+%-\gamma\Delta u +\delta\,u&=\regrhs\,\,\,\,\text{ in }\Omega,\\
+%\gamma \nabla u\cdot\normalvec+\regrobincoeff\,u&=0\,\,\,\text{ on }\partial\Omega.
+%\end{align*}
+%Recall that $\Omega$ is the 2D flattened version of the basal surface $\Gamma_b$.
 Here $\regrhs$ is a generic forcing term, and $\regrobincoeff=\sqrt{\delta \gamma}/1.42$ is a Robin boundary condition coefficient \cite{Roininen14}, and these quantities are different from $\bodyforce$ and $\stokesrobincoeff$ in the Stokes equation, \eqref{Stokeseqn}.
-In all results except one, $\gamma$ is chosen so that the Morozov discrepancy principle is satisfied, i.e.,
-\begin{equation*}
-	\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS = \int_{\Gamma_{t}}||\xi||_{2}^{2}\,dS.
-\end{equation*}
-The value of $\gamma$ satisfying the Morozov discrepancy principle is $\gamma=7.3 \times 10^3$.
+%In all results except one, $\gamma$ is chosen so that the Morozov discrepancy principle is satisfied, i.e.,
+%\begin{equation*}
+%	\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS = \int_{\Gamma_{t}}||\xi||_{2}^{2}\,dS.
+%\end{equation*}
+%The value of $\gamma$ satisfying the Morozov discrepancy principle is $\gamma=7.3 \times 10^3$.
 The exception is Figure \ref{fig:gamma_sweep}, where we vary $\gamma$ to study how changing $\gamma$ impacts the effectiveness of our preconditioner.
 The constant $\delta$ is chosen so that the correlation length of functions drawn from the normal distribution with covariance $\mathcal{K}^{-2}$, given by $L=\sqrt{\gamma/\delta}$, is $1/10$th the radius of the domain.
-%, which is halfway between the maximum and minimum values of $\basalfriction_\text{true}$.
-
-We solve the inverse problem with an inexact Newton preconditioned conjugate gradient (PCG) scheme \cite[Section 7.1]{NocedalWright99}, using $\basalfriction_0$ as the initial guess.
-The discretized parameter, $\bm{\basalfriction}$ is updated iteratively according to the formula
-$
-	\bm{\basalfriction}_{k+1} = \bm{\basalfriction}_k + \alpha_k \bm{\searchdir}_k,
-$
-where $\bm{\searchdir}_k$ approximately solves one of the following Newton linear systems,
+%%, which is halfway between the maximum and minimum values of $\basalfriction_\text{true}$.
+%
+
+We solve the inverse problem with an inexact Newton preconditioned
+conjugate gradient (PCG) scheme and Armijo line search for
+globalization~\cite{NocedalWright99}. The (discrete) Newton updates,
+$\bm{\searchdir}_k$, are obtained by solving 
 \begin{equation}
@@
 \end{equation}
-and the step length $\alpha_k \in \mathbb{R}$ is chosen by line search to satisfy the Armijo conditions \cite[Section 3.1]{NocedalWright99}.
-Here $\bm{g}_k$ is the discretized gradient of $J$, $\bm{H}$ is the discretized Hessian, and $\bm{H}_\text{gn}$ is the discretized Gauss-Newton Hessian, and these derivatives are evaluated at $\bm{\basalfriction}_k$.
+The initial guess is chosen as the discretization of the constant
+function $q_0$. Here $\bm{g}_k$, $\bm{H}$ and $\bm{H}_\text{gn}$ are
+the discretized gradient Hessian and Gauss-Newton Hessian of the
+inverse problem cost function, respectively.
 %The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
-We use $\bm{H}_\text{gn}$ for the first five iterations, and $\bm{H}$ for all subsequent iterations.
-The Newton iteration terminates when $||\bm{g}_k||/||\bm{g}_0|| < 10^{-8}$.
-Systems \eqref{eq:newton_system} are solved inexactly using an inner PCG iteration, which terminates when the norm of the residual for the linear system is less than $\min \left(0.5, \sqrt{||\bm{g}_k||}\right)||\bm{g}_k||$.
-PCG only requires applying $\bm{H}$ (or $\bm{H}_\text{gn}$) to vectors.
-Each application of $\bm{H}$ (or $\bm{H}_\text{gn}$) to a vector requires solving one incremental forward and one incremental adjoint Stokes PDE of the form \eqref{Stokeseqn}, but with different right hand sides.
-
+To ensure positive definiteness of the Hessian we use
+$\bm{H}_\text{gn}$ for the first five iterations, and $\bm{H}$ for all
+subsequent iterations. The Newton iterations are terminated when
+$||\bm{g}_k||/||\bm{g}_0|| < 10^{-8}$. Systems
+\eqref{eq:newton_system} are solved inexactly using an inner PCG
+iteration, which terminates when the norm of the residual for the
+linear system is less than $\min \left(0.5,
+\sqrt{||\bm{g}_k||}\right)||\bm{g}_k||$. %PCG only requires applying
+%$\bm{H}$ (or $\bm{H}_\text{gn}$) to vectors. Each application of
+%$\bm{H}$ (or $\bm{H}_\text{gn}$) to a vector requires solving one
+%incremental forward and one incremental adjoint Stokes PDE of the form
+%\eqref{Stokeseqn}, but with different right hand sides.
 \begin{figure}
@@
 The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to an H-matrix approximation of the Hessian of the regularization term to form an H-matrix approximation of $\bm{H}_\text{gn}$.
-Straightforward calculation shows that the discretized Hessian of the regularization term is given by
-$
-(\gamma \bm{K} + \delta \bm{M}) \bm{M}^{-1} (\gamma \bm{K} + \delta \bm{M})
-$
-where $\bm{K}$ is a finite element stiffness matrix with appropriate boundary conditions, and $\bm{M}$ is the finite element mass matrix.
+%
+%Straightforward calculation shows that the discretized Hessian of the regularization term is given by
+%$
+%(\gamma \bm{K} + \delta \bm{M}) \bm{M}^{-1} (\gamma \bm{K} + \delta \bm{M})
+%$
+%where $\bm{K}$ is a finite element stiffness matrix with appropriate boundary conditions, and $\bm{M}$ is the finite element mass matrix.
 These matrices are sparse, so we form H-matrix representations of these matrices and combine them using standard sparse H-matrix techniques.
-We factor the overall Gauss-Newton Hessian approximation, denoted $\bm{\preconditioner}$, using fast H-matrix methods, then use the factorization as a preconditioner.
+%
+We note that the Hessian of the regularization is a combination of
+stiffness and mass matrices, which are sparse. Therefore, we form
+H-matrix representations of these matrices and combine them using
+standard sparse H-matrix techniques.  We factor the overall
+Gauss-Newton Hessian approximation, denoted $\bm{\preconditioner}$,
+using fast H-matrix methods, then use the factorization as a
+preconditioner.
 

### 2023-01-21 paper/localpsf.tex
small edits to the tex file, cleanup, consistent use of basal sliding and no longer using basal friction and also adding the arXiv version of the HODLR paper as a reference instead of referencing the submitted HODLR paper

@@
 Recently there have been improvements in matrix-free H-matrix construction methods in which an operator is applied to structured random vectors, and the response of the operator to those random vectors is processed to construct an H-matrix approximation  \cite{LevittMartinsson22,LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}.
-These methods (which we do not use here) have been used to approximate Hessians in PDE constrained inverse problems \cite{AmbartsumyanEtAl20,HartlandEtAl22}. 
+These methods (which we do not use here) have been used to approximate Hessians in PDE constrained inverse problems \cite{AmbartsumyanEtAl20,HartlandEtAl23}. 
 Although the required number of matrix-vector products grows nearly linearly in the discretized parameter dimension and linearly in the rank of the off-diagonal blocks, the required number of matrix-vector products is still large (e.g., hundreds to thousands).
@@
 %	\item Velocity field
-%	\item Basal friction field
+%	\item basal sliding field
 %	\item Picture of impulse responses
@@
 Briefly, a sheet of ice is flowing down a mountain (see Figure \ref{fig:ice_mountain_mesh}).
-Given observations of the tangential component of the ice velocity on the top surface of the ice, we invert for the logarithm of the unknown spatially varying basal friction parameter field, which governs the resistance to sliding along the base of the ice sheet.
+Given observations of the tangential component of the ice velocity on the top surface of the ice, we invert for the logarithm of the unknown spatially varying basal sliding parameter field, which governs the resistance to sliding along the base of the ice sheet.
 The inverse problem considered here is synthetic, but in practice surface velocity observations are available from satellites.
@@
 Here, $\basalfriction$ is the unknown logarithmic
-basal friction field (note that large $\basalfriction$ corresponds to high resistance to sliding) which is defined on the $2D$ surface $\Gamma_b$.
+basal sliding field (note that large $\basalfriction$ corresponds to high resistance to sliding) which is defined on the $2$D surface $\Gamma_b$.
 The quantity $\bodyforce$ is the body force density due to gravity, $\stokesrobincoeff=10^6$ is a Robin boundary condition constant,  $\normalvec$ is the outward normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary.
@@
 %This flattening is necessary because our operator approximation method requires adding points on this surface together to get other points on the surface, and this is only well-defined in flat space.
+The observations are generated by adding $1\%$ multiplicative
+Gaussian noise to the tangential component of the velocity field
+restricted to the top surface of the geometry. The true basal sliding
+coefficient field and the true velocity field, which is obtained by
+solving~\eqref{Stokeseqn}, are shown in Figure~\ref{fig:true_beta_u}.
 
@@
 \end{figure}
-
-The observations are generated by adding $1\%$ multiplicative
-Gaussian noise to the tangential component of the velocity field
-restricted to the top surface of the geometry. The true basal sliding
-coefficient field and the true velocity field , which is obtained by
-solving~\eqref{Stokeseqn}, are shown in Figure~\ref{fig:true_beta_u}.
 
@@
 the Green's function of the Hessian of the regularization has a
-characteristic length of $0.25$ of the radius of the domain. For the
+characteristic length of $0.25$ of the domain radius. For the
 specific setup, we refer the reader
-to~[sec.22]\cite{VillaPetraGhattas21}. We choose the regularization
+to~\cite[Section 2.2]{VillaPetraGhattas21}. We choose the regularization
 parameter, $\gamma$, that controls the overall strength of the
@@
 We solve the inverse problem with an inexact Newton preconditioned
-conjugate gradient (PCG) scheme and Armijo line search for
-globalization~\cite{NocedalWright99}. The (discrete) Newton updates,
+conjugate gradient (PCG) scheme and a globalizing Armijo line search~\cite{NocedalWright99}. The (discrete) Newton updates,
 $\bm{\searchdir}_k$, are obtained by solving 
+%
 \begin{equation}
@@
 \end{equation}
-The initial guess is chosen as the discretization of the constant
+%
+wherein we choose the initial guess as the discretization of the constant
 function $q_0$. Here $\bm{g}_k$, $\bm{H}$ and $\bm{H}_\text{gn}$ are
-the discretized gradient Hessian and Gauss-Newton Hessian of the
+the discretized gradient, Hessian, and Gauss-Newton Hessian of the
 inverse problem cost function, respectively.
@@
 subsequent iterations. The Newton iterations are terminated when
-$||\bm{g}_k||/||\bm{g}_0|| < 10^{-8}$. Systems
-\eqref{eq:newton_system} are solved inexactly using an inner PCG
-iteration, which terminates when the norm of the residual for the
-linear system is less than $\min \left(0.5,
+$||\bm{g}_k||/||\bm{g}_0|| < 10^{-8}$. Systems~\eqref{eq:newton_system} are solved inexactly using an inner PCG
+iteration, which terminates when the norm of the linear system residual is less than $\min \left(0.5,
 \sqrt{||\bm{g}_k||}\right)||\bm{g}_k||$. %PCG only requires applying
@@
 	%\end{subfigure}
-	\caption{The log basal friction parameter after $1$ (left), $3$ (second from left), and $4$ (third from left) Newton iterations, and the final Newton iterate (right).
+	\caption{The log basal sliding parameter after $1$ (left), $3$ (second from left), and $4$ (third from left) Newton iterations, and the final Newton iterate (right).
 The PSF (5) preconditioner is constructed between Newton iterations 3 and 4.}
@@
 These approximations are denoted by PSF (1), PSF (5), and PSF (25), respectively.
-The resulting H-matrices, which may be indefinite due to approximation error, are modified to be symmetric positive definite via the procedure described in Appendix \ref{sec:make_spd}, with $l=2$.
-We approximate $\bm{H}_\text{gn}$ rather than $\bm{H}$ because $\bm{H}$ more often has negative values in its integral kernel, especially when one is away from the optimal point.
+The resulting H-matrices, which may be indefinite due to approximation error, are modified to be symmetric positive definite via the procedure described in Appendix~\ref{sec:make_spd}, with $l=2$.
+We approximate $\bm{H}_\text{gn}$ rather than $\bm{H}$ because $\bm{H}$ more often has negative values in its integral kernel, especially when one is away from the optimal basal sliding parameter.
 Our results show that our preconditioners are also good preconditioners for $\bm{H}$.

### 2023-01-24 paper/localpsf.tex
moved text up in the numerics

@@
 
+We use our method to approximate the Newton (or Gauss-Newton) Hessian
+in an inverse problem governed by a PDE which models steady state ice
+sheet flow~\cite{PetraEtAl12} (Section~\ref{sec:ice}) and transport
+and flow of a contaminant~\cite{PetraStadler11}
+(Section~\ref{sec:adv}). To reconstruct the unknown parameters, we
+formulate an inverse problem governed by the corresponding PDEs. In
+particular this problem is formulated as a nonlinear least squares
+optimization problem, whose cost function consists of a misfit term
+(between the observations and model output) and a bi-Laplacian
+regularization term following~\cite{VillaPetraGhattas21}. The
+regularization is centered at a constant function $q_0(x)$. To
+mitigate boundary effects
+%For stokes q_0(x) = 10.5$%
+we use a constant coefficient Robin boundary condition as
+in~\cite{Roininen14}. The parameters for the regularization are chosen
+so that the Green's function of the Hessian of the regularization has
+a characteristic length of $0.25$ of the domain radius. For the
+specific setup, we refer the reader to~\cite[Section
+  2.2]{VillaPetraGhattas21}. For both examples, we choose the
+regularization parameter, $\gamma$, that controls the overall strength
+of the regularization using Morozov's discrepancy
+principle~\cite{Vogel02}. The optimal value of $\gamma$ is used for
+%  $\gamma =  7.3 \times 10^3$ for Stokes  Figure~\ref{fig:gamma_sweep}
+all the numerical results except for the studies where study how
+changing $\gamma$ impacts the effectiveness of the proposed
+preconditioner.
+
+We solve the inverse problem with an inexact Newton preconditioned
+conjugate gradient (PCG) scheme and a globalizing Armijo line search~\cite{NocedalWright99}. The (discrete) Newton updates,
+$\bm{\searchdir}_k$, are obtained by solving 
+%
+\begin{equation}
+	\label{eq:newton_system}
+	\bm{H} \bm{\searchdir}_k = - \bm{g}_k \qquad \text{or} \qquad \bm{H}_\text{gn} \bm{\searchdir}_k = - \bm{g}_k,
+\end{equation}
+%
+wherein we choose the initial guess as the discretization of the constant
+function $q_0$. Here $\bm{g}_k$, $\bm{H}$ and $\bm{H}_\text{gn}$ are
+the discretized gradient, Hessian, and Gauss-Newton Hessian of the
+inverse problem cost function, respectively.
+%The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
+To ensure positive definiteness of the Hessian we use
+$\bm{H}_\text{gn}$ for the first five iterations, and $\bm{H}$ for all
+subsequent iterations. The Newton iterations are terminated when
+$||\bm{g}_k||/||\bm{g}_0|| < 10^{-8}$. Systems~\eqref{eq:newton_system} are solved inexactly using an inner PCG
+iteration, which terminates when the norm of the linear system residual is less than $\min \left(0.5,
+\sqrt{||\bm{g}_k||}\right)||\bm{g}_k||$.
+
+We use the framework described in this paper to generate Hessian preconditioners.
+We build H-matrix approximations of the data misfit Gauss-Newton Hessian (the term in $\bm{H}_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 batches.
+These approximations are denoted by PSF (1), PSF (5), and PSF (25), respectively.
+The resulting H-matrices, which may be indefinite due to approximation error, are modified to be symmetric positive definite via the procedure described in Appendix~\ref{sec:make_spd}, with $l=2$.
+We approximate $\bm{H}_\text{gn}$ rather than $\bm{H}$ because $\bm{H}$ more often has negative values in its integral kernel, especially when one is away from the optimal basal sliding parameter.
+Our results show that our preconditioners are also good preconditioners for $\bm{H}$.
+The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to an H-matrix approximation of the Hessian of the regularization term to form an H-matrix approximation of $\bm{H}_\text{gn}$.
+%
+%Straightforward calculation shows that the discretized Hessian of the regularization term is given by
+%$
+%(\gamma \bm{K} + \delta \bm{M}) \bm{M}^{-1} (\gamma \bm{K} + \delta \bm{M})
+%$
+%where $\bm{K}$ is a finite element stiffness matrix with appropriate boundary conditions, and $\bm{M}$ is the finite element mass matrix.
+These matrices are sparse, so we form H-matrix representations of these matrices and combine them using standard sparse H-matrix techniques.
+%
+We note that the Hessian of the regularization is a combination of
+stiffness and mass matrices, which are sparse. Therefore, we form
+H-matrix representations of these matrices and combine them using
+standard sparse H-matrix techniques.  We factor the overall
+Gauss-Newton Hessian approximation, denoted $\bm{\preconditioner}$,
+using fast H-matrix methods, then use the factorization as a
+preconditioner.
+
 \subsection{Example 1: Inversion for the basal sliding coefficient in an ice sheet flow problem}
+\label{sec:ice}
 
@@
 
-We use our method to approximate the data misfit Gauss-Newton Hessian in an inverse problem governed by a PDE which models steady state ice sheet flow \cite{PetraEtAl12}.
-Briefly, a sheet of ice is flowing down a mountain (see Figure \ref{fig:ice_mountain_mesh}).
-Given observations of the tangential component of the ice velocity on the top surface of the ice, we invert for the logarithm of the unknown spatially varying basal sliding parameter field, which governs the resistance to sliding along the base of the ice sheet.
-The inverse problem considered here is synthetic, but in practice surface velocity observations are available from satellites.
-
+For this example, we consider a sheet of ice flowing down a mountain (see Figure \ref{fig:ice_mountain_mesh}).
+Given observations of the tangential component of the ice velocity on the top surface of the ice, we invert for the logarithm of the unknown spatially varying basal sliding Robin coefficient field, which governs the resistance to sliding along the base of the ice sheet.
+%The inverse problem considered here is synthetic, but in practice surface velocity observations are available from satellites.
 \begin{figure}
@@
 \end{figure} 
-
 The domain that is filled with ice is denoted by
@@
 	\begin{align}
-	-\nabla \cdot \stress
-	&=\bodyforce\,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:1}\\
-	\nabla\cdot \velocity
-	&=0\,\,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:2}\\
-	\stress\normalvec
+	%-\nabla \cdot \stress(v, p)
+	%&=\bodyforce \,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:1}\\
+	%\nabla\cdot \velocity
+	  %&=0\,\,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:2}\\
+          -\nabla \cdot \stress(v, p)=\bodyforce \text{ and }
+          \nabla\cdot \velocity &=0 \quad \text{ in }\icedomain, \label{Stokeseqn:1}\\
+	\stress(v, p)\normalvec
 	&=0\,\,\,\,\,\,\text{ on }\Gamma_{t}, \label{Stokeseqn:3}\\
-	\velocity \cdot\normalvec =0 \quad\text{and}\quad \tangentop\left(\stress\normalvec+\exp\left(\basalfriction\right)\velocity\right)
+	\velocity \cdot\normalvec =0 \text{ and } \tangentop\left(\stress(v, p)\normalvec+\exp\left(\basalfriction\right)\velocity\right)
 	&=0\,\,\,\,\,\,\text{ on }\Gamma_{b},\label{Stokeseqn:4} \\
-	\stress\normalvec+\stokesrobincoeff \velocity &=0\,\,\,\,\,\,\text{ on }\Gamma_{l}.\label{Stokeseqn:5}	
+	\stress(v, p)\normalvec+\stokesrobincoeff \velocity &=0\,\,\,\,\,\,\text{ on }\Gamma_{l}.\label{Stokeseqn:5}	
 	\end{align}
@@
 The quantity $\bodyforce$ is the body force density due to gravity, $\stokesrobincoeff=10^6$ is a Robin boundary condition constant,  $\normalvec$ is the outward normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary.
-We employ Glen's flow law \cite{Glen1955}, $\stress= 2\eta
-\dot{\strain} -I\pressure$,
+We employ Glen's flow law \cite{Glen1955}, $\stress(v, p)= 2\eta
+\dot{\strain}(v) -I\pressure$,
 which is a constitutive law for ice that relates the stress
 tensor, $\stress$, to the strain rate tensor,
-$\dot{\strain}= \frac{1}{2}\left(
+$\dot{\strain}(v)= \frac{1}{2}\left(
 \nabla\velocity+\nabla\velocity^{\top} \right)$.
@@
 
-To reconstruct the basal sliding parameter field $\basalfriction$, we
-formulate an inverse problem governed by the Stokes
-problem~\eqref{Stokeseqn}. In particular this problem is formulated as
-a nonlinear least squares optimization problem, whose cost function
-consists of a misfit term (between the observations and model output)
-and a bi-Laplacian regularization term
-following~\cite{IsaacPetraStadlerEtAl15}. The regularization is
-centered at the constant function $q_0(x) = 10.5$. To mitigate
-boundary effects we use a constant coefficient Robin boundary
-condition as in~\cite{Roininen14}. The parameters are chosen so that
-the Green's function of the Hessian of the regularization has a
-characteristic length of $0.25$ of the domain radius. For the
-specific setup, we refer the reader
-to~\cite[Section 2.2]{VillaPetraGhattas21}. We choose the regularization
-parameter, $\gamma$, that controls the overall strength of the
-regularization using Morozov's discrepancy principle. The optimal
-value of $\gamma$ is $7.3 \times 10^3$, which is used for all the
-numerical results except for Figure~\ref{fig:gamma_sweep}, where we
-vary $\gamma$ to study how changing $\gamma$ impacts the effectiveness
-of the proposed preconditioner.
+%aaa
 
@@
 %
-
-We solve the inverse problem with an inexact Newton preconditioned
-conjugate gradient (PCG) scheme and a globalizing Armijo line search~\cite{NocedalWright99}. The (discrete) Newton updates,
-$\bm{\searchdir}_k$, are obtained by solving 
-%
-\begin{equation}
-	\label{eq:newton_system}
-	\bm{H} \bm{\searchdir}_k = - \bm{g}_k \qquad \text{or} \qquad \bm{H}_\text{gn} \bm{\searchdir}_k = - \bm{g}_k,
-\end{equation}
-%
-wherein we choose the initial guess as the discretization of the constant
-function $q_0$. Here $\bm{g}_k$, $\bm{H}$ and $\bm{H}_\text{gn}$ are
-the discretized gradient, Hessian, and Gauss-Newton Hessian of the
-inverse problem cost function, respectively.
-%The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
-To ensure positive definiteness of the Hessian we use
-$\bm{H}_\text{gn}$ for the first five iterations, and $\bm{H}$ for all
-subsequent iterations. The Newton iterations are terminated when
-$||\bm{g}_k||/||\bm{g}_0|| < 10^{-8}$. Systems~\eqref{eq:newton_system} are solved inexactly using an inner PCG
-iteration, which terminates when the norm of the linear system residual is less than $\min \left(0.5,
-\sqrt{||\bm{g}_k||}\right)||\bm{g}_k||$. %PCG only requires applying
+%PCG only requires applying
 %$\bm{H}$ (or $\bm{H}_\text{gn}$) to vectors. Each application of
@@
 \end{figure}
-
-We use the framework described in this paper to generate Hessian preconditioners.
-We build H-matrix approximations of the data misfit Gauss-Newton Hessian (the term in $\bm{H}_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 batches.
-These approximations are denoted by PSF (1), PSF (5), and PSF (25), respectively.
-The resulting H-matrices, which may be indefinite due to approximation error, are modified to be symmetric positive definite via the procedure described in Appendix~\ref{sec:make_spd}, with $l=2$.
-We approximate $\bm{H}_\text{gn}$ rather than $\bm{H}$ because $\bm{H}$ more often has negative values in its integral kernel, especially when one is away from the optimal basal sliding parameter.
-Our results show that our preconditioners are also good preconditioners for $\bm{H}$.
-The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to an H-matrix approximation of the Hessian of the regularization term to form an H-matrix approximation of $\bm{H}_\text{gn}$.
-%
-%Straightforward calculation shows that the discretized Hessian of the regularization term is given by
-%$
-%(\gamma \bm{K} + \delta \bm{M}) \bm{M}^{-1} (\gamma \bm{K} + \delta \bm{M})
-%$
-%where $\bm{K}$ is a finite element stiffness matrix with appropriate boundary conditions, and $\bm{M}$ is the finite element mass matrix.
-These matrices are sparse, so we form H-matrix representations of these matrices and combine them using standard sparse H-matrix techniques.
-%
-We note that the Hessian of the regularization is a combination of
-stiffness and mass matrices, which are sparse. Therefore, we form
-H-matrix representations of these matrices and combine them using
-standard sparse H-matrix techniques.  We factor the overall
-Gauss-Newton Hessian approximation, denoted $\bm{\preconditioner}$,
-using fast H-matrix methods, then use the factorization as a
-preconditioner.
 
@@
 This consists of Stokes solves for $u$ performed during the linesearch going from $\bm{\basalfriction}_{k-1}$ to $\bm{\basalfriction}_k$, plus one adjoint Stokes solve to compute the gradient at $\bm{\basalfriction}_k$, plus one incremental forward and one incremental adjoint Stokes solve per PCG iteration for solving the Newton system.
-In the PSF (5) portion of row 3, we write $6+22$ to indicate that 6 Stokes solves were used during the standard course of the iteration, and 22 Stokes solves were used to build the PSF (5) preconditioner.
-}
+In the PSF (5) portion of row 3, we write $6+22$ to indicate that 6 Stokes solves were used during the standard course of the iteration, and 22 Stokes solves were used to build the PSF (5) preconditioner.}
 	\label{tab:newton_convergence_table}
@@
 \subsection{Example 2: Inversion for the initial condition in an advection-diffusion problem}
+\label{sec:adv}
 

### 2023-03-07 localpsf/custom_krylov.py
wrote custom gmres function that can terminate when negative direction is encountered

@@ -1,0 +1,139 @@
+import numpy as np
+import scipy.linalg as sla
+import typing as typ
+import numpy.typing as npt
+
+
+def positive_gmres(apply_A: typ.Callable[[np.ndarray], np.ndarray],  # R^N -> R^N
+                   b: np.ndarray,  # shape=(N,)
+                   x0: np.ndarray=None,  # shape=(N,)
+                   apply_M: typ.Callable[[np.ndarray], np.ndarray]=None,  # R^N -> R^N
+                   solve_M: typ.Callable[[np.ndarray], np.ndarray]=None,  # R^N -> R^N
+                   rtol: float=1e-10,
+                   max_iter=50,
+                   terminate_negative_direction=True,
+                   display=True,
+                   callback: typ.Callable[[np.ndarray, float, float], None]=None, # callback(xk, norm_r, xt_A_x)
+                   ) -> np.ndarray:
+    '''Modified from Algorithm 6.9 in Saad iterative methods book page 172. No restarts.'''
+    assert(len(b.shape) == 1)
+    N = len(b)
+
+    if solve_M is None or apply_M is None:
+        assert(solve_M is None)
+        assert(apply_M is None)
+        apply_M = lambda x: x
+        solve_M = lambda x: x
+
+    def printmaybe(stuff):
+        if display:
+            print(stuff)
+
+    if callback is None:
+        callback = lambda xk, nr, xAx: None
+
+    norm_b = np.linalg.norm(b)
+
+    if x0 is None:
+        x0 = np.zeros(N)
+        Ax0 = np.zeros(N)
+    else:
+        assert(x0.shape == (N,))
+        Ax0 = apply_A(x0)
+    r0 = solve_M(b - Ax0)
+
+    beta = np.linalg.norm(r0)
+
+    V_big = np.zeros((N, max_iter + 1))
+    V_big[:,0] = r0 / beta
+
+    norm_r = np.linalg.norm(b - Ax0)
+    xt_A_x = np.dot(x0, Ax0)
+    callback(x0, norm_r, xt_A_x)
+    printmaybe('initial guess: norm_r/norm_b=' + str(norm_r / norm_b) + ', xt_A_x=' + str(xt_A_x))
+
+    x = x0
+    H_big = np.zeros((max_iter + 1, max_iter))
+    for jj in range(max_iter):
[…]

### 2023-03-07 localpsf/positive_definite_modifications.py

@@ -1,4 +1,6 @@
 import numpy as np
+import typing as typ
+import scipy.linalg as sla
 import scipy.sparse.linalg as spla
 
 
@@ -81,3 +83,215 @@ class WoodburyObject: # (A + U*C*V) * x = b
         y = me.solve_A(b)
         return y - me.solve_A(np.dot(me.U, me.solve_capacitance_matrix(np.dot(me.V, y))))
 
+
+def eig_coeff(eig: float,  # bad eigenvalue A*u = eig*B*u
+              tau: float,  # tolerance in (0,1)
+              method='zero',
+              ) -> float:  # correction coefficient
+    if method == 'thresh':  # eig + 1 + c = tau
+        c = tau - eig - 1.0
+    elif method == 'zero':  # eig + 1 + c = 1
+        c = -eig
+    elif method == 'flip':  # eig + 1 + c = 1 + |eig|
+        c = np.abs(eig) - eig
+    return c
+
+
+def bad_generalized_eigs(apply_A: typ.Callable[[np.ndarray], np.ndarray],  # shape=(N,N)
+                         apply_B: typ.Callable[[np.ndarray], np.ndarray],  # shape=(N,N)
+                         solve_B: typ.Callable[[np.ndarray], np.ndarray],  # shape=(N,N)
+                         N: int,
+                         tau: float = 0.5,  # tolerance in (0,1)
+                         chunk_size: int = 5,
+                         max_rank: int = 50,
+                         display: bool = False,
+                         ) -> typ.Tuple[np.ndarray,  # U=[u0, u1, ...]: shape=(N,k)
+                                        np.ndarray,  # V=[B@u0, B@u1, ...]: shape=(N,k)
+                                        np.ndarray]:  # eigs=[e0, e1, ...]: shape=(N,))
+    '''Finds geigs A@ui=ei*B@ui such that ui.T@(A + B)@ui < tau.
+
+    A is symmetric indefinite
+    B is symmetric positive definite
+    A.shape=B.shape=(N,N)
+    apply_A(x) = A @ x
+    apply_B(x) = B @ x
+    solve_B(apply_B(x)) = apply_B(solve_B(x)) = x
+    '''
+    Bop = spla.LinearOperator((N, N), matvec=apply_B)
+    iBop = spla.LinearOperator((N, N), matvec=solve_B)
+    uu: typ.List[np.ndarray] = []
[…]

### 2023-03-21 paper/localpsf.tex
wrote custom gmres function that can terminate when negative direction is encountered
added example description

@@
 
+Here we consider a time-dependent advection-diffusion equation for
+which we seek to infer an unknown initial contaminant field from
+pointwise measurements of its concentration. The problem description
+below closely follows the one in~\cite{PetraStadler11}. This PDE
+models diffusive transport in a domain $\D \subset \R^d$, which is
+depicted in Figure~\ref{fig:adv_inversion1}.  The domain
+boundaries $\partial \D$ include the outer boundaries as well as the
+internal boundaries of the rectangles, which represent buildings.  The
+parameter-to-observable map $\iFF$ in this case maps an initial
+condition $\ipar \in L^2(\D)$ to observations of the concentration
+field $u(\vec{x}, t)$ through solution of the advection-diffusion
+equation given by
+\begin{equation}\label{eq:ad}
+  \begin{aligned}
+    u_t - \kappa\Delta u + \vec{v}\cdot\nabla u &= 0 & \quad&\text{in
+    }\D\times (0,T), \\
+    u(\cdot, 0) &= \ipar  &&\text{in } \D , \\
+    \kappa\nabla u\cdot \vec{n} &= 0 &&\text{on } \partial\D \times (0,T).
+  \end{aligned}
+\end{equation}
+Here, $\kappa > 0$ is a diffusivity coefficient, and $T > 0$ is the
+final time of observations.  The velocity field $\vec{v}$, shown in
+Figure~\ref{fig:adv_inversion1}, is computed by solving the
+steady-state Navier-Stokes equations for a two dimensional flow with
+Reynolds number 50 and boundary conditions as in
+Figure~\ref{fig:adv_inversion} (left); see \cite{PetraStadler11} for
+details.  The time evolution of the state variable $u$ from a given
+initial condition $\ipar$ is illustrated in
+Figure~\ref{fig:adv_inversion1}.
+
 %INTRODUCE EQUATIONS: REWRITE HERE vvvvvvv

### 2023-04-11 paper/localpsf.tex
edits during zoom meeting with Noemi and Tucker

--- 2023-01-24 paper/localpsf.tex
@@
 We use our method to approximate the Newton (or Gauss-Newton) Hessian
-in an inverse problem governed by a PDE which models steady state ice
+in inverse problems governed by PDEs which model steady state ice
 sheet flow~\cite{PetraEtAl12} (Section~\ref{sec:ice}) and transport
 and flow of a contaminant~\cite{PetraStadler11}
-(Section~\ref{sec:adv}). To reconstruct the unknown parameters, we
-formulate an inverse problem governed by the corresponding PDEs. In
-particular this problem is formulated as a nonlinear least squares
-optimization problem, whose cost function consists of a misfit term
+(Section~\ref{sec:adv}). 
+To reconstruct the unknown parameter fields, denoted $q$, the inverse problems are formulated as nonlinear least squares
+optimization problems, whose cost functions consist of a misfit term
 (between the observations and model output) and a bi-Laplacian
@@
 so that the Green's function of the Hessian of the regularization has
-a characteristic length of $0.25$ of the domain radius. For the
+a characteristic length of $0.25$ (NOTE: CHECK THIS FOR BOTH PROBLEMS) of the domain radius. For the
 specific setup, we refer the reader to~\cite[Section
@@
 %  $\gamma =  7.3 \times 10^3$ for Stokes  Figure~\ref{fig:gamma_sweep}
-all the numerical results except for the studies where study how
+all the numerical results except for the studies where we study how
 changing $\gamma$ impacts the effectiveness of the proposed
@@
 
-We solve the inverse problem with an inexact Newton preconditioned
-conjugate gradient (PCG) scheme and a globalizing Armijo line search~\cite{NocedalWright99}. The (discrete) Newton updates,
+We solve the ice sheet inverse problem with an inexact Newton preconditioned
+conjugate gradient (PCG) scheme and a globalizing Armijo line search~\cite{NocedalWright99}. The Newton updates,
 $\bm{\searchdir}_k$, are obtained by solving 
@@
 \sqrt{||\bm{g}_k||}\right)||\bm{g}_k||$.
+The advection-diffusion inverse problem is linear, so Newton's method converges in one iteration.
+In this case the Newton linear system, \eqref{eq:newton_system}, is solved using PCG with a fixed tolerance which is specified in Section \ref{sec:adv}.
 
 We use the framework described in this paper to generate Hessian preconditioners.
-We build H-matrix approximations of the data misfit Gauss-Newton Hessian (the term in $\bm{H}_\text{gn}$ that arises from the data misfit) using 1, 5, and 25 batches.
-These approximations are denoted by PSF (1), PSF (5), and PSF (25), respectively.
-The resulting H-matrices, which may be indefinite due to approximation error, are modified to be symmetric positive definite via the procedure described in Appendix~\ref{sec:make_spd}, with $l=2$.
-We approximate $\bm{H}_\text{gn}$ rather than $\bm{H}$ because $\bm{H}$ more often has negative values in its integral kernel, especially when one is away from the optimal basal sliding parameter.
-Our results show that our preconditioners are also good preconditioners for $\bm{H}$.
-The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to an H-matrix approximation of the Hessian of the regularization term to form an H-matrix approximation of $\bm{H}_\text{gn}$.
+We build H-matrix approximations, $\mathbf{A}_H$, of the data misfit Gauss-Newton Hessian (the term in $\bm{H}_\text{gn}$ that arises from the data misfit).
+The
+%number of impulse response batches used to build these 
+approximations are indicated by ``PSF ($\nbatch$)'', where $\nbatch$ is the number of impulse response batches used to build the approximation.
+The Hessian of the regularization term is a combination of
+stiffness and mass matrices, which are sparse. Therefore, we form
+H-matrix representations of these matrices and combine them into a H-matrix approximation of the regularization term in the Hessian, $\mathbf{R}_H$, using
+standard sparse H-matrix techniques. H-matrix approximations of the Gauss-Newton Hessian, 
+$$\bm{H}_\text{gn}\approx \widetilde{\mathbf{H}}:=\mathbf{A}_H + \mathbf{R}_H,$$
+are formed by adding $\bm{A}_H$ to $\mathbf{R}_H$ using fast H-matrix arithmetic.
+The H-matrix $\widetilde{\mathbf{H}}$ may be indefinite due to approximation error, so we modify it to be symmetric positive definite via a shift and invert deflation procedure described in Appendix~\ref{sec:make_spd}.
+We factor $\bm{\preconditioner}$,
+using fast H-matrix methods, then use the factorization as a
+preconditioner.
+We approximate $\bm{H}_\text{gn}$ rather than $\bm{H}$ because $\bm{H}$ more often has negative values in its integral kernel.
+Our results show that $\widetilde{\mathbf{H}}$ is a good preconditioner for both $\bm{H}_\text{gn}$ and $\bm{H}$.
+%The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to an H-matrix approximation of the Hessian of the regularization term to form an H-matrix approximation of $\bm{H}_\text{gn}$.
 %
@@
 %
-We note that the Hessian of the regularization is a combination of
-stiffness and mass matrices, which are sparse. Therefore, we form
-H-matrix representations of these matrices and combine them using
-standard sparse H-matrix techniques.  We factor the overall
-Gauss-Newton Hessian approximation, denoted $\bm{\preconditioner}$,
-using fast H-matrix methods, then use the factorization as a
-preconditioner.
 
@@
 	\end{subfigure}
-	\caption{Left (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_k = \lambda_k \bm{\preconditioner} \bm{u}_k$.
+	\caption{Left (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \genericvec_k = \lambda_k \bm{\preconditioner} \genericvec_k$.
 Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), or the regularization Hessian (REG).
@@
 $
-	\bm{H} \bm{u} = \lambda \bm{\preconditioner} \bm{u}.
+	\bm{H} \genericvec = \lambda \bm{\preconditioner} \genericvec.
 $
@@
 
+Here we consider a time-dependent advection-diffusion equation for
+which we seek to infer the unknown spatially varying initial condition, $\ipar$, from observations of the state  
+at a final time. This PDE
+models diffusive transport in a domain $\advdomain \subset \R^d$, which is
+depicted in Figure~\ref{fig:adv_inversion1}. In this case, the state, $\concentration(x,t)$, could be interpreted as the concentration of a contaminant. The problem description
+below closely follows the one in~\cite{PetraStadler11}. The domain
+boundaries $\partial \advdomain$ include the outer boundaries as well as the
+internal boundaries of the rectangles, which represent buildings.  The
+parameter-to-observable map $\iFF$ in this case maps an initial
+condition $\ipar \in L^2(\advdomain)$ to observations of the concentration at a final time, $\concentration(x, T)$,  through solution of the advection-diffusion
+equation given by
+\begin{equation}\label{eq:ad}
+  \begin{aligned}
+    \concentration_t - \kappa\Delta \concentration + v\cdot\nabla \concentration &= 0 & \quad&\text{in
+    }\advdomain\times (0,T), \\
+    \concentration(\cdot, 0) &= \ipar  &&\text{in } \advdomain , \\
+    \kappa\nabla \concentration\cdot \nu &= 0 &&\text{on } \partial \advdomain \times (0,T).
+  \end{aligned}
+\end{equation}
+Here, $\kappa > 0$ is a diffusivity coefficient, $\nu$ is the boundary unit normal vector, and $T > 0$ is the
+final time.  The velocity field $v:\advdomain \rightarrow \mathbb{R}^d$, shown in
+Figure~\ref{fig:adv_inversion1}, is computed by solving the
+steady-state Navier-Stokes equations for a two dimensional flow with
+Reynolds number 50 and boundary conditions as in
+Figure~\ref{fig:adv_inversion} (left); see \cite{PetraStadler11} for
+details.  The time evolution of the state variable $\concentration$ from a given
+initial condition $\ipar$ is illustrated in
+Figure~\ref{fig:adv_inversion1}.
+
 %INTRODUCE EQUATIONS: REWRITE HERE vvvvvvv
 
-%The parameter-to-observable map $\iFF \,\ipar := \obsop\, u(\ipar)$ maps an initial condition $\ipar \in L^2(\D)$ to pointwise spatiotemporal observations of the concentration field $u({\vec x},t)$ through solution of the advection-diffusion equation given by
+%The parameter-to-observable map $\iFF \,\ipar := \obsop\, u(\ipar)$ maps an initial condition $\ipar \in L^2(\advdomain)$ to pointwise spatiotemporal observations of the concentration field $u({\vec x},t)$ through solution of the advection-diffusion equation given by
 %
@@
 %\begin{split}
-%u_t - \kappa\Delta u + {\vec v} \cdot \nabla u &= 0     & \quad \text{in } \D\times(0,T),\\
-%u(\cdot, 0) &= \ipar & \quad \text{in } \D,\\
-%\kappa \nabla u\cdot {\vec{n}} &= 0     & \quad \text{on } \partial\D \times (0,T).
+%u_t - \kappa\Delta u + {\vec v} \cdot \nabla u &= 0     & \quad \text{in } \advdomain\times(0,T),\\
+%u(\cdot, 0) &= \ipar & \quad \text{in } \advdomain,\\
+%\kappa \nabla u\cdot {\vec{n}} &= 0     & \quad \text{on } \partial \advdomain \times (0,T).
 %\end{split}
@@
 %
-%Here, $\D \subset \R^d$ ($d \in \{2, 3\}$) is a bounded domain, $\kappa > 0$ is the diffusion coefficient and $T > 0$ is the final
+%Here, $\advdomain \subset \R^d$ ($d \in \{2, 3\}$) is a bounded domain, $\kappa > 0$ is the diffusion coefficient and $T > 0$ is the final
 %time. The velocity field
@@
 %\begin{aligned}
-%- \frac{1}{\operatorname{Re}} \Delta {\vec v} + \nabla q + {\vec v} \cdot \nabla {\vec v} &= 0 &\quad&\text{ in }\D,\\
-%\nabla \cdot {\vec v} &= 0 &&\text{ in }\D,\\
-%{\vec v} &= {\vec g} &&\text{ on } \partial\D.
+%- \frac{1}{\operatorname{Re}} \Delta {\vec v} + \nabla q + {\vec v} \cdot \nabla {\vec v} &= 0 &\quad&\text{ in }\advdomain,\\
+%\nabla \cdot {\vec v} &= 0 &&\text{ in }\advdomain,\\
+%{\vec v} &= {\vec g} &&\text{ on } \partial \advdomain.
 %\end{aligned}
@@
 			legend style={at={(0.95,1)},anchor=north east},
-			title={Generalized eigenvalues, $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$}
+			title={Generalized eigenvalues, $\bm{H} \genericvec_{k}=\lambda_{k} \bm{\preconditioner} \genericvec_{k}$}
 			]

### 2023-05-04 paper/localpsf.tex
processed the true relative errors for the CG convergence history, including data set with initial 0 CG iteration 1 relative error (assuming initial guess for CG is 0.), removed regularization robustness plot, moved (old data) generalized eigenvalue plot, which shows eigenvalue clustering of the preconditioned Hessian

@@
 %    c) any words or sentences to add with the ref
+
+\newcommand{\todo}[1]{\noindent\emph{\textcolor{red}{To do: #1\:}}}
 
@@
 Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, and $L^2(\Omega)'$ is the space of real-valued continuous linear functionals on $L^2(\Omega)$.
-We focus on approximating Hessians in optimization and inverse problems governed by partial differential equations (PDEs) \cite{AlgerEtAl19,IsaacEtAl15}, though such operators also arise as certain Schur complements in Schur complement methods for solving partial differential equations (PDEs) \cite{le1997non,saad1999distributed}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics \cite{ChenStein21,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging \cite{DenisEtAl11,NagyOleary98}.
+We focus on approximating Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{AlgerEtAl19,IsaacEtAl15}, though such operators also arise as certain Schur complements in Schur complement methods for solving partial differential equations (PDEs)~\cite{le1997non,saad1999distributed}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics~\cite{ChenStein21,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}.
 Here ``matrix-free'' means that we may apply $\Aop$ and its transpose\footnote{Recall that $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for operations with $\Aop^T$.
@@
 but we cannot easily access entries of $\Aop$'s integral kernel.
-Evaluating the maps in \eqref{eq:eval_maps} may require, as a sub procedure, performing a costly computation, such as solving a large linear system or time stepping.
+Evaluating the maps in~\eqref{eq:eval_maps} may require, as a sub procedure, performing a costly computation, such as solving a large linear system or time stepping.
 
@@
 	\end{subfigure}
-	\caption{Batches, $\eta_b$, of normalized impulses, $\impulseresponse_{x}$, for the Stokes inverse problem data misfit Gauss-Newton Hessian (Section \ref{sec:numerical_results}). 
+	\caption{Batches, $\eta_b$, of normalized impulses, $\impulseresponse_{x}$, for the Stokes inverse problem data misfit Gauss-Newton Hessian (Section~\ref{sec:numerical_results}). 
 %	Brightness scale differs between the subfigures. 
@@
 The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
-We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with batches of points scattered throughout the domain (see Figure \ref{fig:batches_intro}).
+We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with batches of points scattered throughout the domain (see Figure~\ref{fig:batches_intro}).
 Then we interpolate translated and scaled versions of these impulse responses to approximate entries of the operator's integral kernel.
@@
 The idea of estimating the supports of the functions $\phi_x$ a-priori was inspired by techniques from resolution analysis in seismic imaging. There, 
-$\Aop^T$ is applied to a random noise function, and the width of $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$  \cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
-We use polynomial functions instead of random noise functions (see Section \ref{sec:intromoments}).
+$\Aop^T$ is applied to a random noise function, and the width of $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$~\cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
+We use polynomial functions instead of random noise functions (see Section~\ref{sec:intromoments}).
 Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
@@
 For background on PC and PSF methods, we recommend the following papers: \cite{DenisEtAl15,EscandeWeiss17,GentileCourbinMeylan13}.
-While PC approximations are based on an assumption of local translation invariance, the method we propose is based on an assumption we call ``local mean displacement invariance,'' which is more general than local translation invariance, and includes both CUR low rank approximation and local translation invariance as special cases (Section \ref{sec:local_mean_displacement_invariance} and Figure \ref{fig:mean_displacement_invariance}).
+While PC approximations are based on an assumption of local translation invariance, the method we propose is based on an assumption we call ``local mean displacement invariance,'' which is more general than local translation invariance, and includes both CUR low rank approximation and local translation invariance as special cases (Section~\ref{sec:local_mean_displacement_invariance} and Figure~\ref{fig:mean_displacement_invariance}).
 %Furthermore, PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulses.
@@
 
-The ability to rapidly approximate entries of $\Aop$'s integral kernel allows us to form a hierarchical matrix (H-matrix) \cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of a discretized version of $\Aop$.
+The ability to rapidly approximate entries of $\Aop$'s integral kernel allows us to form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of a discretized version of $\Aop$.
 H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank.
-%(see Figure \ref{fig:hmatrix_intro} and Appendix \ref{app:h_matrix}).
+%(see Figure~\ref{fig:hmatrix_intro} and Appendix~\ref{app:h_matrix}).
 H-matrix methods permit us to perform matrix-vector products cheaply, and perform other useful linear algebra operations that cannot be done easily using the original operator.
@@
 \emph{nearly linearly} in $N$ (i.e., $o(N^{1+\epsilon})$ for any $\epsilon > 0$).
-The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks \cite{GrasedyckHackbusch03}.
+The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks~\cite{GrasedyckHackbusch03}.
 The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank Gaussian surrogate models that can be used for uncertainty quantification.
 %Classical H-matrix construction methods 
-%(described in Appendix \ref{sec:H_matrix_construction}) 
+%(described in Appendix~\ref{sec:H_matrix_construction}) 
 %require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.
@@
 %Inverse problems are typically ill-posed, in the sense that many different parameter fields may yield predicted observations that agree with the actual observations to within the noise level. 
-Conventionally, the inverse problem is formulated using either a deterministic framework \cite{AsterBorchersThurber18,Vogel02},
+Conventionally, the inverse problem is formulated using either a deterministic framework~\cite{AsterBorchersThurber18,Vogel02},
 %Guy Chavent, Nonlinear Least Squares for Inverse Problems, Springer, 2009.
-or a Bayesian probabilistic framework \cite{KaipioSomersalo05,Stuart10a,Tarantola05}.
-In the deterministic framework, one solves an optimization problem to find the parameter that best fits the observations, subject to appropriate regularization \cite{EnglHankeNeubauer96}.
+or a Bayesian probabilistic framework~\cite{KaipioSomersalo05,Stuart10a,Tarantola05}.
+In the deterministic framework, one solves an optimization problem to find the parameter that best fits the observations, subject to appropriate regularization~\cite{EnglHankeNeubauer96}.
 In the probabilistic framework, Bayes' theorem combines the observations with prior information to form a posterior distribution over the space of all possible parameter fields, and computations are performed to extract statistical information about the parameter from this posterior.
 The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal to each other under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian.'' The Hessian consists of a data misfit term (the ``data misfit Hessian''), which depends on discrepancy between the observations and the associated model predictions, and a regularization or prior term (the ``regularization Hessian'') which does not depend on the observations.
-For more details on the Hessian, see \cite{Alger19,GhattasWillcox21,VillaPetraGhattas21}.
+For more details on the Hessian, see~\cite{Alger19,GhattasWillcox21,VillaPetraGhattas21}.
 
@@
 When solving the deterministic optimization problem with Newton-type methods, the Hessian is the coefficient operator for the linear system that must be solved or approximately solved at every Newton iteration.
-Good Hessian preconditioners reduce the number of iterations required to solve these Newton linear systems with Krylov methods such as conjugate gradient \cite{Saad03}.
+Good Hessian preconditioners reduce the number of iterations required to solve these Newton linear systems with Krylov methods such as conjugate gradient~\cite{Saad03}.
 In the Bayesian setting, the inverse of the Hessian is the covariance of a local Gaussian approximation of the posterior.
 Such a Gaussian distribution can be used as a direct approximations of the posterior, or as a proposal for Markov chain Monte-Carlo methods for drawing samples from the posterior.
-For instance, see \cite{KimEtAl21,PetraEtAl14} and the references therein.
+For instance, see~\cite{KimEtAl21,PetraEtAl14} and the references therein.
 
@@
 As the regularization is weakened, the eigenvalues of the regularization Hessian decrease.
-These illustrations are taken from \cite[Chapter 1]{Alger19}, with minor modifications.}
+These illustrations are taken from~\cite[Chapter 1]{Alger19}, with minor modifications.}
 %	\label{fig:hessian_eigenstructure}
@@
 Owing to the implicit dependence of predicted observations on the parameter, entries of the Hessian are not easily accessible.
-Rather, the Hessian may be applied to a vector via a computational process that involves solving a pair of forward and adjoint PDEs which are linearizations of the original PDE \cite{GhattasWillcox21,Gunzburger03}.
-The most popular matrix-free Hessian approximation methods are based on low rank approximation of either the data misfit Hessian, or the data misfit Hessian preconditioned by the regularization Hessian, e.g., \cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15}.
-Krylov methods such as Lanczos or randomized methods \cite{Cheng05,HalkoMartinssonTropp11} are typically used to construct these low rank approximations by applying the Hessian to vectors.
+Rather, the Hessian may be applied to a vector via a computational process that involves solving a pair of forward and adjoint PDEs which are linearizations of the original PDE~\cite{GhattasWillcox21,Gunzburger03}.
+The most popular matrix-free Hessian approximation methods are based on low rank approximation of either the data misfit Hessian, or the data misfit Hessian preconditioned by the regularization Hessian, e.g.,~\cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15}.
+Krylov methods such as Lanczos or randomized methods~\cite{Cheng05,HalkoMartinssonTropp11} are typically used to construct these low rank approximations by applying the Hessian to vectors.
 Using these methods, the required number of Hessian applications (and hence the required number of PDE solves) is typically proportional to the rank of the low rank approximation.
@@
 This means that the required number of PDE solves remains the same as the mesh used to discretize the parameter is refined.
-However, in many inverse problems of practical interest the numerical rank of the data misfit Hessian, while mesh independent, is still large, which makes it costly to approximate the Hessian using low rank approximation methods \cite{AmbartsumyanEtAl20,BuiEtAl13,IsaacEtAl15}.
-
-Examples of inverse problems with high rank data misfit Hessians include advection dominated advection-diffusion inverse problems \cite{AkcelikBirosDraganescuEtAl05}\cite[Chapter 5]{Flath13}, high frequency wave inverse problems \cite{BuiEtAl13}, and more generally, all inverse problems in which the observations highly inform the parameter.
+However, in many inverse problems of practical interest the numerical rank of the data misfit Hessian, while mesh independent, is still large, which makes it costly to approximate the Hessian using low rank approximation methods~\cite{AmbartsumyanEtAl20,BuiEtAl13,IsaacEtAl15}.
+
+Examples of inverse problems with high rank data misfit Hessians include advection dominated advection-diffusion inverse problems~\cite{AkcelikBirosDraganescuEtAl05}\cite[Chapter 5]{Flath13}, high frequency wave inverse problems~\cite{BuiEtAl13}, and more generally, all inverse problems in which the observations highly inform the parameter.
 The eigenvalues of the data misfit Hessian characterize the sensitivity of the predicted observations to perturbations of the parameter in the corresponding eigenvector directions,
-%(see Figure \ref{fig:hessian_eigenstructure})
-so more informative data leads to larger eigenvalues and therefore a larger numerical rank \cite{AlexanderianGloorGhattas16}\cite[Section 1.4 and Chapter 4]{Alger19}.
+%(see Figure~\ref{fig:hessian_eigenstructure})
+so more informative data leads to larger eigenvalues and therefore a larger numerical rank~\cite{AlexanderianGloorGhattas16}\cite[Section 1.4 and Chapter 4]{Alger19}.
 Roughly speaking, the numerical rank of the data misfit Hessian is the dimension of the subspace of parameter space that is informed by the data.
-The numerical rank of the regularization preconditioned data misfit Hessian may be reduced by increasing the strength of the regularization, but this throws away useful information---components of the parameter that could be learned from components of the observations that have a large signal to noise ratio are instead reconstructed based on the regularization \cite[Section 4]{AlgerEtAl17}\cite[Chapters 1 and 7]{Vogel02}.
+The numerical rank of the regularization preconditioned data misfit Hessian may be reduced by increasing the strength of the regularization, but this throws away useful information---components of the parameter that could be learned from components of the observations that have a large signal to noise ratio are instead reconstructed based on the regularization~\cite[Section 4]{AlgerEtAl17}\cite[Chapters 1 and 7]{Vogel02}.
 Hence, low rank approximation methods suffer from a predicament---if the data highly inform the parameter and the regularization is chosen appropriately, then a large number of operator applies are required to form an accurate approximation of the Hessian using low rank approximation methods.
-%The best case scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost) \cite{AlexanderianGloorGhattas16,Alger19,AlgerEtAl17}.
+%The best case scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost)~\cite{AlexanderianGloorGhattas16,Alger19,AlgerEtAl17}.
 High rank Hessian approximation methods are needed.
 
-Recently there have been improvements in matrix-free H-matrix construction methods in which an operator is applied to structured random vectors, and the response of the operator to those random vectors is processed to construct an H-matrix approximation  \cite{LevittMartinsson22,LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}.
-These methods (which we do not use here) have been used to approximate Hessians in PDE constrained inverse problems \cite{AmbartsumyanEtAl20,HartlandEtAl23}. 
+Recently there have been improvements in matrix-free H-matrix construction methods in which an operator is applied to structured random vectors, and the response of the operator to those random vectors is processed to construct an H-matrix approximation~\cite{LevittMartinsson22,LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}.
+These methods (which we do not use here) have been used to approximate Hessians in PDE constrained inverse problems~\cite{AmbartsumyanEtAl20,HartlandEtAl23}. 
 Although the required number of matrix-vector products grows nearly linearly in the discretized parameter dimension and linearly in the rank of the off-diagonal blocks, the required number of matrix-vector products is still large (e.g., hundreds to thousands).
@@
 But despite asymptotic scalability, from a practical perspective the required number of matrix-vector products is large (e.g., hundreds to thousands).
-%Improving the peeling process to reduce its cost is currently an area of active research \cite{PEELINGRESEARCH}.
+%Improving the peeling process to reduce its cost is currently an area of active research~\cite{PEELINGRESEARCH}.
 %This is an area of active research, so more efficient and practical matrix-free H-matrix construction methods based on the peeling process may be discovered in the future.
-%Matrix-free H-matrix construction methods are currently an area of active research (see, e.g., \cite{LevittMartinsson22}), so more efficient matrix-free H-matrix methods may be discovered in the future.
+%Matrix-free H-matrix construction methods are currently an area of active research (see, e.g.,~\cite{LevittMartinsson22}), so more efficient matrix-free H-matrix methods may be discovered in the future.
 In this paper, we also form an H-matrix Hessian approximation.
@@
 %Our motivation is high rank Hessian approximation in distributed parameter inverse problems governed by PDEs. 
-%The Hessian is an operator that is central to efficient methods for solving such inverse problems and is only available matrix-free; we will discuss Hessians in inverse problems governed by PDEs in greater detail in Section \ref{sec:hessian}.
+%The Hessian is an operator that is central to efficient methods for solving such inverse problems and is only available matrix-free; we will discuss Hessians in inverse problems governed by PDEs in greater detail in Section~\ref{sec:hessian}.
 
@@
 The integral kernel, $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$, exists in principle but is not easily accessible.
-In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define impulse responses (Section \ref{sec:impulse_response_2}), we state the conditions on $\Aop$ that our method requires (Section \ref{sec:conditions_2}), and we detail finite element discretization (Section \ref{sec:finite_element_kernel}).
+In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define impulse responses (Section~\ref{sec:impulse_response_2}), we state the conditions on $\Aop$ that our method requires (Section~\ref{sec:conditions_2}), and we detail finite element discretization (Section~\ref{sec:finite_element_kernel}).
 
@@
 Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions and let $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ be a distribution.
-In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in \eqref{eq:extension_to_distributions} is not valid.
-However, if $\Aker$ is sufficiently regular such that the function $y \mapsto \rho\left(\Aker(y,~\cdot~)\right)$ is well defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of \eqref{eq:action_on_distribution} is well defined.
-Hence, we \emph{define} the action of $\Aop$ on the distribution $\rho$ to be the right hand side of \eqref{eq:action_on_distribution}.
+In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in~\eqref{eq:extension_to_distributions} is not valid.
+However, if $\Aker$ is sufficiently regular such that the function $y \mapsto \rho\left(\Aker(y,~\cdot~)\right)$ is well defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of~\eqref{eq:action_on_distribution} is well defined.
+Hence, we \emph{define} the action of $\Aop$ on the distribution $\rho$ to be the right hand side of~\eqref{eq:action_on_distribution}.
 We denote this action by ``$\Aop \genericdistribution^*$,'' even if $\rho^*$ does not exist.
@@
 that is formed by applying $\Aop$ to $\delta_x$, then taking the Riesz representation of the resulting linear functional.
-Using \eqref{eq:action_on_distribution} and the definition of the delta distribution, we see that $\phi_x$ may also be written as the function $\impulseresponse_x(y) = \Aker(y, x)$.
+Using~\eqref{eq:action_on_distribution} and the definition of the delta distribution, we see that $\phi_x$ may also be written as the function $\impulseresponse_x(y) = \Aker(y, x)$.
 %\begin{equation}
@@
 
-The operator $A_h:V_h \rightarrow V_h'$ can be written in integral kernel form, \eqref{eq:kernel_representation}, but with $\Aker$ replaced by a slightly different integral kernel,
+The operator $A_h:V_h \rightarrow V_h'$ can be written in integral kernel form,~\eqref{eq:kernel_representation}, but with $\Aker$ replaced by a slightly different integral kernel,
 %as
@@
 $
-Therefore, the formula for the impulse response from \eqref{eq:impulse_response_delta_action} becomes 
+Therefore, the formula for the impulse response from~\eqref{eq:impulse_response_delta_action} becomes 
 $
@@
 \end{equation}
-Broadly, our method will construct an H-matrix approximation of $\mathbf{A}$ by forming H-matrix approximations of $\boldsymbol{\Aker}$ and $\massmatrix$ (or a lumped mass version of $\massmatrix$), then multiplying these matrices per \eqref{eq:boldA} using H-matrix methods.
+Broadly, our method will construct an H-matrix approximation of $\mathbf{A}$ by forming H-matrix approximations of $\boldsymbol{\Aker}$ and $\massmatrix$ (or a lumped mass version of $\massmatrix$), then multiplying these matrices per~\eqref{eq:boldA} using H-matrix methods.
 Classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Aker}_{ij}$, but these matrix entries are not easily accessible.
@@
 We therefore recommend replacing the mass matrix, $\massmatrix$, with a positive diagonal \emph{lumped mass} approximation.
-In our numerical results, we use the lumped mass matrix constructed by replacing off-diagonal entries of the mass matrix by zero. Other mass lumping techniques may be used \cite{Hughes12}.
+In our numerical results, we use the lumped mass matrix constructed by replacing off-diagonal entries of the mass matrix by zero. Other mass lumping techniques may be used~\cite{Hughes12}.
 
@@
 In this section we present two key innovations that our method is based on.
-First, we define moments of the impulse responses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{sec:intromoments}).
+First, we define moments of the impulse responses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section~\ref{sec:intromoments}).
 %Our method will use these ellipsoid estimates to choose sets of impulses that can be computed in batches. 
-Second, we describe an improved method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}).
+Second, we describe an improved method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section~\ref{sec:local_mean_displacement_invariance}).
 %Our method will interpolate impulses using this approximation.
@@
 
-The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to obtain $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}.
+The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to obtain $\impulseresponse_x$, as per~\eqref{eq:impulse_response_delta_action}.
 Then one can post process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
@@
 \end{equation*}
-for $i,j=1,\dots,\gdim$. Using the definition of $\spatialvol$ in \eqref{eq:define_vol} and using \eqref{eq:operator_simultaneous_ips}, we have
+for $i,j=1,\dots,\gdim$. Using the definition of $\spatialvol$ in~\eqref{eq:define_vol} and using~\eqref{eq:operator_simultaneous_ips}, we have
 \begin{equation*}
@@
 for $i,j=1,\dots, \gdim$. Here $u/w$ denotes pointwise division, $\left(u/w\right)(x) = u(x)/w(x)$, and $u\cdot w$ denotes pointwise multiplication, $(u \cdot w)(x) = u(x)w(x)$.
-%The resulting method for computing discretized versions of $\spatialvol$, $\spatialmean$, and $\spatialcov$ is shown in Algorithm \ref{alg:varhpi_mean_cov}.
+%The resulting method for computing discretized versions of $\spatialvol$, $\spatialmean$, and $\spatialcov$ is shown in Algorithm~\ref{alg:varhpi_mean_cov}.
 
@@
 The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$.
-However, in Section \ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids is better), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
+However, in Section~\ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids is better), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
 In practice, we find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results.
@@
 These are four different ways to approximate an impulse response by a nearby impulse response, with each successive approximation building upon the previous ones.
-Our method uses \eqref{eq:translate4}, which is the most sophisticated of these approximations.
-Approximation \eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close.
-Operators satisfying \eqref{eq:translate1} can be well approximated via low rank CUR approximation \cite{MahoneyDrineas09}.
-However, the required rank in the low rank approximation can be large, which makes algorithms based on \eqref{eq:translate1} expensive.
-Operators that satisfy \eqref{eq:translate2} are called ``locally translation invariant'' because integral kernel entries $\Aker(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$.
-It is straightforward to show that if equality holds in \eqref{eq:translate2}, then $\Aop$ is a convolution operator.
+Our method uses~\eqref{eq:translate4}, which is the most sophisticated of these approximations.
+Approximation~\eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close.
+Operators satisfying~\eqref{eq:translate1} can be well approximated via low rank CUR approximation~\cite{MahoneyDrineas09}.
+However, the required rank in the low rank approximation can be large, which makes algorithms based on~\eqref{eq:translate1} expensive.
+Operators that satisfy~\eqref{eq:translate2} are called ``locally translation invariant'' because integral kernel entries $\Aker(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$.
+It is straightforward to show that if equality holds in~\eqref{eq:translate2}, then $\Aop$ is a convolution operator.
 Locally translation invariant operators act like convolutions locally, and can therefore be well approximated by PC approximations.
 
-Approximation \eqref{eq:translate3} improves upon \eqref{eq:translate1} and \eqref{eq:translate2}, and generalizes both.
-On one hand, if \eqref{eq:translate1} holds, then $\spatialmean(x) \approx \spatialmean(x')$, and so \eqref{eq:translate3} holds.
-On the other hand, translating a distribution translates its mean, so if \eqref{eq:translate2} holds, then $\spatialmean(x')-\spatialmean(x) \approx x' - x$, so again \eqref{eq:translate3} holds.
-But approximation \eqref{eq:translate3} can hold in situations where neither \eqref{eq:translate1} nor \eqref{eq:translate2} holds.
-For example, because the expected value commutes with affine transformations, \eqref{eq:translate3} will hold when $\Aop$ is locally translation invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case.
-Additionally, \eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions.
-In contrast, \eqref{eq:translate2} does not naturally generalize to operators that map between function spaces on different domains, because the formula $y-x+x'$ requires vectors in $\Omega_2$ and $\Omega_1$ to be added together.
-We call \eqref{eq:translate3} ``local mean displacement invariance,'' and illustrate \eqref{eq:translate3} in Figure \ref{fig:mean_displacement_invariance}.
-
-We use approximation \eqref{eq:translate4}, which is the same as \eqref{eq:translate3}, except for the extra factors of $1/\spatialvol(x)$ on the left hand side and $1/\spatialvol(x')$ on the right hand side.
+Approximation~\eqref{eq:translate3} improves upon~\eqref{eq:translate1} and~\eqref{eq:translate2}, and generalizes both.
+On one hand, if~\eqref{eq:translate1} holds, then $\spatialmean(x) \approx \spatialmean(x')$, and so~\eqref{eq:translate3} holds.
+On the other hand, translating a distribution translates its mean, so if~\eqref{eq:translate2} holds, then $\spatialmean(x')-\spatialmean(x) \approx x' - x$, so again~\eqref{eq:translate3} holds.
+But approximation~\eqref{eq:translate3} can hold in situations where neither~\eqref{eq:translate1} nor~\eqref{eq:translate2} holds.
+For example, because the expected value commutes with affine transformations,~\eqref{eq:translate3} will hold when $\Aop$ is locally translation invariant with respect to a rotated frame of reference, while~\eqref{eq:translate2} will not hold in this case.
+Additionally,~\eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions.
+In contrast,~\eqref{eq:translate2} does not naturally generalize to operators that map between function spaces on different domains, because the formula $y-x+x'$ requires vectors in $\Omega_2$ and $\Omega_1$ to be added together.
+We call~\eqref{eq:translate3} ``local mean displacement invariance,'' and illustrate~\eqref{eq:translate3} in Figure~\ref{fig:mean_displacement_invariance}.
+
+We use approximation~\eqref{eq:translate4}, which is the same as~\eqref{eq:translate3}, except for the extra factors of $1/\spatialvol(x)$ on the left hand side and $1/\spatialvol(x')$ on the right hand side.
 These factors make the approximation more robust if $\spatialvol(x)$ varies widely.
-Approximation \eqref{eq:translate4} is equivalent to \eqref{eq:translate3}, but with $\phi_x$ replaced by its normalized version, $\phi_x/\spatialvol(x)$.
-We call \eqref{eq:translate4} ``normalized local mean displacement invariance.''
+Approximation~\eqref{eq:translate4} is equivalent to~\eqref{eq:translate3}, but with $\phi_x$ replaced by its normalized version, $\phi_x/\spatialvol(x)$.
+We call~\eqref{eq:translate4} ``normalized local mean displacement invariance.''
 
@@
 
-We use \eqref{eq:vol_mean_var} in Section \ref{sec:intromoments} to compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ by applying $\Aop^T$ to a small number of polynomial functions.
-Then we use \eqref{eq:support_ellipsoid} to form ellipsoid shaped estimates for the support of the $\phi_x$, \emph{without} computing them.
-This allows us to compute large numbers of $\impulseresponse_{x_i}$ in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}).
-We compute one batch, denoted $\eta_b$, by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section \ref{sec:get_impulse_response}).
-The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support ellipsoid for $\impulseresponse_{x_i}$ and the support ellipsoid for $\impulseresponse_{x_j}$ do not overlap if $i \neq j$ (Section \ref{sec:sample_point_selection}).
-Because these supports do not overlap (or do not overlap much), we can post process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}).
+We use~\eqref{eq:vol_mean_var} in Section~\ref{sec:intromoments} to compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ by applying $\Aop^T$ to a small number of polynomial functions.
+Then we use~\eqref{eq:support_ellipsoid} to form ellipsoid shaped estimates for the support of the $\phi_x$, \emph{without} computing them.
+This allows us to compute large numbers of $\impulseresponse_{x_i}$ in ``batches,'' $\eta_b$ (see Figure~\ref{fig:batches_intro}).
+We compute one batch, denoted $\eta_b$, by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section~\ref{sec:get_impulse_response}).
+The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support ellipsoid for $\impulseresponse_{x_i}$ and the support ellipsoid for $\impulseresponse_{x_j}$ do not overlap if $i \neq j$ (Section~\ref{sec:sample_point_selection}).
+Because these supports do not overlap (or do not overlap much), we can post process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section~\ref{sec:get_impulse_response}).
 The process is repeated until a desired number of batches is reached.
 
-Once the batches $\eta_b$ are computed, we approximate the integral kernel $\Aker(y,x)$ at arbitrary points $(y,x)$ by interpolation of translated and scaled versions of the computed $\impulseresponse_{x_i}$ (Section \ref{sec:approximate_kernel_entries}).
-The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section \ref{sec:local_mean_displacement_invariance}.
+Once the batches $\eta_b$ are computed, we approximate the integral kernel $\Aker(y,x)$ at arbitrary points $(y,x)$ by interpolation of translated and scaled versions of the computed $\impulseresponse_{x_i}$ (Section~\ref{sec:approximate_kernel_entries}).
+The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section~\ref{sec:local_mean_displacement_invariance}.
 Specifically, we approximate $\Aker(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}(y - \spatialmean(x) + \spatialmean(x_i))$ for a small number of sample points $x_i$ near $x$.
@@
 
-The ability to rapidly evaluate approximate kernel entries $\Aker(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Aker}_H \approx \mathbf{\Aker}$, using the conventional adaptive cross H-matrix construction method (Section \ref{sec:h_matrix_construction_short}).
+The ability to rapidly evaluate approximate kernel entries $\Aker(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Aker}_H \approx \mathbf{\Aker}$, using the conventional adaptive cross H-matrix construction method (Section~\ref{sec:h_matrix_construction_short}).
 In this method, one forms low rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks.
@@
 %\end{equation*}
-The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}. The cost is discussed in Section \ref{sec:overall_cost}.
+The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm~\ref{alg:construct_Atilde}. The cost is discussed in Section~\ref{sec:overall_cost}.
 
 When $\Aop$ is symmetric positive semi-definite, $\mathbf{A}_H$ may be non-symmetric and indefinite due to errors in the approximation.
-In this case, one may (optionally) modify the H-matrix representation of $\mathbf{A}_H$ to make it symmetric positive semi-definite using a rational function method that we describe in Appendix \ref{sec:make_spd}.
+In this case, one may (optionally) modify the H-matrix representation of $\mathbf{A}_H$ to make it symmetric positive semi-definite using a rational function method that we describe in Appendix~\ref{sec:make_spd}.
 
@@
 The matrix is symmetrized by computing $\left(\mathbf{A}_H+\mathbf{A}_H^T\right)/2$ using H-matrix addition.
-Negative eigenvalues modified to be positive using a rational method that is described in Appendix \ref{sec:make_spd}.
+Negative eigenvalues modified to be positive using a rational method that is described in Appendix~\ref{sec:make_spd}.
 
@@
 
-		Compute $V, \mu$, and $\Sigma$ (Equations \eqref{eq:vol_mean_var} in Section \ref{sec:intromoments})
-
-%		Compute $\bm{\spatialvol}$, $\bm{\spatialmean}$, and $\bm{\spatialcov}$ using Algorithm \ref{alg:varhpi_mean_cov}.
+		Compute $V, \mu$, and $\Sigma$ (Equations~\eqref{eq:vol_mean_var} in Section~\ref{sec:intromoments})
+
+%		Compute $\bm{\spatialvol}$, $\bm{\spatialmean}$, and $\bm{\spatialcov}$ using Algorithm~\ref{alg:varhpi_mean_cov}.
 
 		\For{$k=1,2,\dots,\nbatch$}{
-			Choose a batch of sample points, $\pointbatch_k$ (Section \ref{sec:sample_point_selection})
-
-			Compute $\combresponse_k$ by applying $\Aop$ to the Dirac comb for $\pointbatch_k$ (Section \ref{sec:get_impulse_response})
+			Choose a batch of sample points, $\pointbatch_k$ (Section~\ref{sec:sample_point_selection})
+
+			Compute $\combresponse_k$ by applying $\Aop$ to the Dirac comb for $\pointbatch_k$ (Section~\ref{sec:get_impulse_response})
 
@@
 
-		Form H-matrix approximation $\boldsymbol{\Aker}_H$ of integral kernel (Section \ref{sec:h_matrix_construction_short})
-
-		Form H-matrix approximation $\mathbf{A}_H$ of $\Aop$ (Section \ref{sec:h_matrix_construction_short})
-
-		(optional) Modify $\mathbf{A}_H$ to make it symmetric positive definite (Appendix \ref{sec:make_spd})
+		Form H-matrix approximation $\boldsymbol{\Aker}_H$ of integral kernel (Section~\ref{sec:h_matrix_construction_short})
+
+		Form H-matrix approximation $\mathbf{A}_H$ of $\Aop$ (Section~\ref{sec:h_matrix_construction_short})
+
+		(optional) Modify $\mathbf{A}_H$ to make it symmetric positive definite (Appendix~\ref{sec:make_spd})
 
@@
 This process repeats until there are no more points in $\candidatepts$.
-%This is detailed in Algorithm \ref{alg:point_choice}. 
-We check whether $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ using the ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.
+%This is detailed in Algorithm~\ref{alg:point_choice}. 
+We check whether $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ using the ellipsoid intersection test described in Appendix~\ref{sec:fast_ellipsoid_intersection_test}.
 We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots, \pointbatch_{\nbatch}$.
@@
 %		\If{$E_{\candidatepoint_i} \cap E_q \neq \{\}$ for all $\candidatepoint_j \in \pointbatch_k$}{
-%			\tcp{$E_{\candidatepoint_i}$ and $E_{\candidatepoint_j}$ are the ellipsoids defined in \eqref{eq:support_ellipsoid}}
+%			\tcp{$E_{\candidatepoint_i}$ and $E_{\candidatepoint_j}$ are the ellipsoids defined in~\eqref{eq:support_ellipsoid}}
 %			
@@
 \end{equation}
-The last equality in \eqref{eq:dirac_comb_H_action} follows from linearity and the definition of $\impulseresponse_{x_i}$ in \eqref{eq:impulse_response_delta_action}.
+The last equality in~\eqref{eq:dirac_comb_H_action} follows from linearity and the definition of $\impulseresponse_{x_i}$ in~\eqref{eq:impulse_response_delta_action}.
 Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\impulseresponse_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\impulseresponse_j$ do not overlap when $i \neq j$, we have (approximately)
@@
 Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, may overwhelm $\phi_{x_j}$ for a nearby point $x_j$ if $\spatialvol(x_i)$ is much larger than $\spatialvol(x_j)$.
-Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section \ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $\spatialvol(x_i)$ is smaller than a predetermined threshold.
+Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section~\ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $\spatialvol(x_i)$ is smaller than a predetermined threshold.
 
@@
 Here $\numnbr$ is a small user defined parameter; e.g., $\numnbr=10$.
-We find the $\numnbr$ nearest sample points to $x$ by querying a precomputed kd-tree \cite{Bentley75} of all sample points.
-We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{Ericson04} of the mesh cells used to discretize the problem.
+We find the $\numnbr$ nearest sample points to $x$ by querying a precomputed kd-tree~\cite{Bentley75} of all sample points.
+We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree)~\cite{Ericson04} of the mesh cells used to discretize the problem.
 Further, let
@@
 \end{equation}
-Note that $\phi_{x_i}\left(z_i\right)$ is well defined because $z_i \in \Omega$, and $\frac{\spatialvol(x)}{\spatialvol(x_i)}$ is well defined because our sample point picking procedure in Section \ref{sec:sample_point_selection} ensures that $\spatialvol(x_i)>0$.
-Per the discussion in Section \ref{sec:local_mean_displacement_invariance}, we expect $\Aker(y,x) \approx f_i$ for $i=1,\dots,\numnbr$; the closer $x_i$ is to $x$, the better we expect the approximation to be.
+Note that $\phi_{x_i}\left(z_i\right)$ is well defined because $z_i \in \Omega$, and $\frac{\spatialvol(x)}{\spatialvol(x_i)}$ is well defined because our sample point picking procedure in Section~\ref{sec:sample_point_selection} ensures that $\spatialvol(x_i)>0$.
+Per the discussion in Section~\ref{sec:local_mean_displacement_invariance}, we expect $\Aker(y,x) \approx f_i$ for $i=1,\dots,\numnbr$; the closer $x_i$ is to $x$, the better we expect the approximation to be.
 We therefore approximate $\Aker(y,x)$ by interpolating the (point,value) pairs
@@
 at the point $x$. 
-%The idea is that $\Aker(y,x) = f_i$ per the discussion in Section \ref{sec:local_mean_displacement_invariance}, so we approximate $\Aker(y,x)$ by interpolating the values $f_i$. 
-Interpolation is performed using the following radial basis function \cite{Wendland04} scheme:
+%The idea is that $\Aker(y,x) = f_i$ per the discussion in Section~\ref{sec:local_mean_displacement_invariance}, so we approximate $\Aker(y,x)$ by interpolating the values $f_i$. 
+Interpolation is performed using the following radial basis function~\cite{Wendland04} scheme:
 \begin{equation}
@@
 $
-and $f \in \mathbb{R}^{\numnbr}$ has entries $f_i$ from \eqref{eq:fxyxp}.
-Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a certain least squares sense \cite{Duchon77}.
-%For more details on radial basis function interpolation, see \cite{Wendland04}. 
-%The interpolation procedure is illustrate in Figure \ref{fig:rbf_interpolation}.
+and $f \in \mathbb{R}^{\numnbr}$ has entries $f_i$ from~\eqref{eq:fxyxp}.
+Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a certain least squares sense~\cite{Duchon77}.
+%For more details on radial basis function interpolation, see~\cite{Wendland04}. 
+%The interpolation procedure is illustrate in Figure~\ref{fig:rbf_interpolation}.
 
@@
 
-To evaluate $f_i$, we check whether $z_i \in E_{x_i}$ using \eqref{eq:support_ellipsoid}.
+To evaluate $f_i$, we check whether $z_i \in E_{x_i}$ using~\eqref{eq:support_ellipsoid}.
 If $z_i \notin E_{x_i}$, then $z_i$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_i=0$.
@@
 $
-per \eqref{eq:varphi_eval}. Note that $z_i$ is typically not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints. Hence,
+per~\eqref{eq:varphi_eval}. Note that $z_i$ is typically not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints. Hence,
 evaluating $\eta_b\left(z_i\right)$ requires determining which mesh cell contains $z_i$, then evaluating finite element basis functions on that mesh cell.
@@
 
-We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming H-matrix representations $\mathbf{\Aker}_H$ and $\massmatrix_H$ of $\mathbf{\Aker}$ and $\massmatrix$, respectively, then using fast H-matrix methods to multiply these matrices per \eqref{eq:boldA} to form
+We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming H-matrix representations $\mathbf{\Aker}_H$ and $\massmatrix_H$ of $\mathbf{\Aker}$ and $\massmatrix$, respectively, then using fast H-matrix methods to multiply these matrices per~\eqref{eq:boldA} to form
 $
@@
 $
-%Once $\mathbf{A}_H$ is formed, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods \cite{BORMHACKBUSCHBOOK}.
-We use H1 matrices in our numerical results, but any other H-matrix format could be used instead. For more details on H-matrices, see \cite{Hackbusch15}.
-
-We form $\mathbf{\Aker}_H$ using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{Kriemann08}.
-For details about the algorithms used for geometrical clustering, H-matrix construction, and H-matrix operations in HLIBpro, we recommend the the following papers \cite{BormGrasedyckHackbusch03,GrasedyckKriemannLeBorne08,Kriemann13}.
-Although $\mathbf{\Aker}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Aker}_H$ only requires evaluation of $O(\hrank^2 \fedim \log \fedim)$ kernel entries $\mathbf{\Aker}_{ij} = \widetilde{\Aker}(\fepoint_i, \fepoint_j)$, and these entries are computed via the radial basis function interpolation method described in Section \ref{sec:approximate_kernel_entries}.
+%Once $\mathbf{A}_H$ is formed, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods~\cite{BORMHACKBUSCHBOOK}.
+We use H1 matrices in our numerical results, but any other H-matrix format could be used instead. For more details on H-matrices, see~\cite{Hackbusch15}.
+
+We form $\mathbf{\Aker}_H$ using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package~\cite{Kriemann08}.
+For details about the algorithms used for geometrical clustering, H-matrix construction, and H-matrix operations in HLIBpro, we recommend the the following papers~\cite{BormGrasedyckHackbusch03,GrasedyckKriemannLeBorne08,Kriemann13}.
+Although $\mathbf{\Aker}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Aker}_H$ only requires evaluation of $O(\hrank^2 \fedim \log \fedim)$ kernel entries $\mathbf{\Aker}_{ij} = \widetilde{\Aker}(\fepoint_i, \fepoint_j)$, and these entries are computed via the radial basis function interpolation method described in Section~\ref{sec:approximate_kernel_entries}.
 Here $\hrank$ is the rank of the highest rank block in the H-matrix.
 A dense representation of $\mathbf{\Aker}$ is never formed. 
-%We describe this H-matrix construction process in detail in Appendix \ref{app:h_matrix}. 
+%We describe this H-matrix construction process in detail in Appendix~\ref{app:h_matrix}. 
 We form $\massmatrix_H$ using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same geometrical clustering block cluster tree as was used for $\mathbf{\Aker}$.
@@
 All operations that do not require applications of $\Aop$ to vectors are nearly linear, and therefore scalable, in the size of the problem, $N$.
-We now describe these costs in detail. For convenience, Table \ref{tab:vars} lists variable symbols and their approximate sizes.
+We now describe these costs in detail. For convenience, Table~\ref{tab:vars} lists variable symbols and their approximate sizes.
 
@@
 
-Computing the impulse response batches also requires choosing sample point batches via the greedy ellipsoid packing algorithm described in Section \ref{sec:sample_point_selection}.
+Computing the impulse response batches also requires choosing sample point batches via the greedy ellipsoid packing algorithm described in Section~\ref{sec:sample_point_selection}.
 Choosing the $i$th batch of sample points may require performing up to $\fedim |\pointbatch_i|$ ellipsoid intersection tests, where $|\pointbatch_i|$ is the number of sample points in the $i$th batch.
@@
 That is, choosing sample points from a smaller subset of $\fedim_\text{thin} << \fedim$ points.
-The cost to perform the ellipsoid intersection tests could be reduced by first checking whether the ellipsoid bounding boxes intersect before performing the ellipsoid intersection test of Appendix \ref{sec:fast_ellipsoid_intersection_test}.
+The cost to perform the ellipsoid intersection tests could be reduced by first checking whether the ellipsoid bounding boxes intersect before performing the ellipsoid intersection test of Appendix~\ref{sec:fast_ellipsoid_intersection_test}.
 The cost of picking the $i$th batch could be reduced to $O(\fedim \log |\pointbatch_i|)$ using more advanced computational geometry techniques, such as organizing the ellipsoids associated with points that have already been picked into a dynamic AABB tree, then using this AABB tree to accelerate the required ellipsoid intersection tests.
@@
 \paragraph{(2) Building the H-matrix} The classical H-matrix construction technique
-%, which is described in Appendix \ref{app:h_matrix}, 
+%, which is described in Appendix~\ref{app:h_matrix}, 
 requires evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix.
@@
 \end{equation}
-where $a \in \{0,1,2,3\}$ \cite{GrasedyckHackbusch03}.
+where $a \in \{0,1,2,3\}$~\cite{GrasedyckHackbusch03}.
 The precise cost depends on the type of H-matrix used and the operation being performed.
-In the numerical results (Section \ref{sec:numerical_results}), we use one matrix-matrix addition to add the H-matrix approximation of the data misfit term in the Hessian to the regularization term in the Hessian.
+In the numerical results (Section~\ref{sec:numerical_results}), we use one matrix-matrix addition to add the H-matrix approximation of the data misfit term in the Hessian to the regularization term in the Hessian.
 Constructing an H-matrix representation of the regularization term in the Hessian requires one matrix inversion and two matrix-matrix multiplications.
 Then, we use one matrix factorization to form a preconditioner for the Newton linear system.
-%We perform low-rank updates to the H-matrix as the Newton iterations proceed, as per the discussion in Appendix \ref{sec:recycle_krylov}.
-The rational method for modifying $\mathbf{A}_H$ to be positive semi-definite, described in Appendix \ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the rational function order.
+%We perform low-rank updates to the H-matrix as the Newton iterations proceed, as per the discussion in Appendix~\ref{sec:recycle_krylov}.
+The rational method for modifying $\mathbf{A}_H$ to be positive semi-definite, described in Appendix~\ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the rational function order.
 
@@
 The advection-diffusion inverse problem is linear, so Newton's method converges in one iteration.
-In this case the Newton linear system, \eqref{eq:newton_system}, is solved using PCG with a fixed tolerance which is specified in Section \ref{sec:adv}.
+In this case the Newton linear system,~\eqref{eq:newton_system}, is solved using PCG with a fixed tolerance which is specified in Section~\ref{sec:adv}.
 
@@
 
-For this example, we consider a sheet of ice flowing down a mountain (see Figure \ref{fig:ice_mountain_mesh}).
+For this example, we consider a sheet of ice flowing down a mountain (see Figure~\ref{fig:ice_mountain_mesh}).
 Given observations of the tangential component of the ice velocity on the top surface of the ice, we invert for the logarithm of the unknown spatially varying basal sliding Robin coefficient field, which governs the resistance to sliding along the base of the ice sheet.
 %The inverse problem considered here is synthetic, but in practice surface velocity observations are available from satellites.
+%
+%
 \begin{figure}
-	\begin{subfigure}{0.5\textwidth}
+	\begin{center} 
 	\includegraphics[scale=0.23]{meshHeight_view2_edges.png}
-		\caption{Bird's eye view}
-	\end{subfigure}
-	\begin{subfigure}{0.48\textwidth}
-		\includegraphics[scale=0.22]{meshHeight_view0_edges_withAxes.png}
-		\caption{Top down view}
-	\end{subfigure}
-	\caption{The ice sheet, discretized by a mesh of tetrahedra.
+	\end{center}
+	\caption{Bird's eye view of the ice sheet, discretized by a mesh of tetrahedra.
 Color indicates the height of the base of the ice sheet (i.e., the mountain topography).
@@
 \end{figure} 
+%
+%
 The domain that is filled with ice is denoted by
@@
 equations,
-
+%
+%
 \begin{subequations}\label{Stokeseqn}
@@
 The quantity $\bodyforce$ is the body force density due to gravity, $\stokesrobincoeff=10^6$ is a Robin boundary condition constant,  $\normalvec$ is the outward normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary.
-We employ Glen's flow law \cite{Glen1955}, $\stress(v, p)= 2\eta
+We employ Glen's flow law~\cite{Glen1955}, $\stress(v, p)= 2\eta
 \dot{\strain}(v) -I\pressure$,
@@
 %% from here we summarize --------
-%The true parameter field, $\basalfriction_\text{true}$, and corresponding velocity field, $\velocity_\text{true}$, are shown in Figure \ref{fig:true_beta_u}.
+%The true parameter field, $\basalfriction_\text{true}$, and corresponding velocity field, $\velocity_\text{true}$, are shown in Figure~\ref{fig:true_beta_u}.
 We generate synthetic observations by restricting the tangential component of $\velocity_\text{true}$ to the top surface to get
@@
 %Recall that $\Omega$ is the 2D flattened version of the basal surface $\Gamma_b$.
-Here $\regrhs$ is a generic forcing term, and $\regrobincoeff=\sqrt{\delta \gamma}/1.42$ is a Robin boundary condition coefficient \cite{Roininen14}, and these quantities are different from $\bodyforce$ and $\stokesrobincoeff$ in the Stokes equation, \eqref{Stokeseqn}.
+Here $\regrhs$ is a generic forcing term, and $\regrobincoeff=\sqrt{\delta \gamma}/1.42$ is a Robin boundary condition coefficient~\cite{Roininen14}, and these quantities are different from $\bodyforce$ and $\stokesrobincoeff$ in the Stokes equation,~\eqref{Stokeseqn}.
 %In all results except one, $\gamma$ is chosen so that the Morozov discrepancy principle is satisfied, i.e.,
@@
 %The value of $\gamma$ satisfying the Morozov discrepancy principle is $\gamma=7.3 \times 10^3$.
-The exception is Figure \ref{fig:gamma_sweep}, where we vary $\gamma$ to study how changing $\gamma$ impacts the effectiveness of our preconditioner.
+The exception is Figure~\ref{fig:gamma_sweep}, where we vary $\gamma$ to study how changing $\gamma$ impacts the effectiveness of our preconditioner.
 The constant $\delta$ is chosen so that the correlation length of functions drawn from the normal distribution with covariance $\mathcal{K}^{-2}$, given by $L=\sqrt{\gamma/\delta}$, is $1/10$th the radius of the domain.
@@
 
-Table \ref{tab:newton_convergence_table} shows the performance of our preconditioner for accelerating the solution of optimization problem \eqref{eq:deterministic_optimization_problem}.
+Table~\ref{tab:newton_convergence_table} shows the performance of our preconditioner for accelerating the solution of optimization problem~\eqref{eq:deterministic_optimization_problem}.
 We build the PSF (5) preconditioner in the third Gauss-Newton iteration, and reuse it for all subsequent Gauss-Newton and Newton iterations.
@@
 Using PSF (5) reduces the total number of Stokes PDE solves by a factor of six compared to no preconditioning, and by a factor of nine compared to regularization preconditioning\footnote{Interestingly, we observe that no preconditioning outperforms regularization preconditioning here because the noise level (and hence $\gamma$) is small.}.
-In Figure \ref{fig:newton_iterates} we show select Newton iterates using our PSF (5) preconditioner.
-
+In Figure~\ref{fig:newton_iterates} we show select Newton iterates using our PSF (5) preconditioner.
+%
+%
 \begin{figure}
@@
 		ymode=log,
-		ymin=1e-11,
+		ymin=1e-13,
 		ymax=2,
@@
 		xtick={0, 200, 400, 600},
-		xlabel={CG iteration},
-		ylabel={$||\bm{b}-\bm{H} \bm{x}_k||/||\bm{b}||$},
 		title={CG convergence},
+		xlabel={$k$, CG iteration},
+		ylabel={$||\bm{x}_k-\bm{x}_{\text{true}}||/||\bm{x}_{\text{true}}||$},
 		legend style={at={(1.1,1)},anchor=north east},
@@
 		\addlegendentry{Reg}		
-		\addplot[dotted, line width=0.5mm] table {cg_relres_reg.dat};
+		\addplot[dotted, line width=0.5mm] table {cg_relres_reg_0.05noise.dat};
 		\addlegendentry{None}
-		\addplot[dash dot, line width=0.5mm] table {cg_relres_none.dat};
+		\addplot[dash dot, line width=0.5mm] table {cg_relres_none_0.05noise.dat};
 		\addlegendentry{PSF (1)}
-		\addplot[blue, line width=0.5mm] table {cg_relres_p1.dat};
+		\addplot[blue, line width=0.5mm] table {cg_relres_psf1_0.05noise.dat};
 		\addlegendentry{PSF (5)}
-		\addplot[red, line width=0.5mm] table {cg_relres_p5.dat};
+		\addplot[red, line width=0.5mm] table {cg_relres_psf5_0.05noise.dat};
 		\addlegendentry{PSF (25)}
-		\addplot[purple, line width=0.5mm] table {cg_relres_p25.dat};
+		\addplot[purple, line width=0.5mm] table {cg_relres_psf25_0.05noise.dat};
 		\end{axis}
@@
 	\begin{subfigure}{0.49\textwidth}
-		\centering
+	\begin{center}
 		\begin{tikzpicture}[baseline, trim axis right]
 		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
-		xmode=log,
-		ymode=log,
-		xlabel={$\gamma$, regularization parameter},
-		ylabel={\# of CG iterations},
-		axis y line*=right,
-		axis x line*=bottom,
-		ylabel near ticks, yticklabel pos=right,
-		%ylabel near ticks, ylabel shift={-6pt},
-		%legend pos=north east,
-		title={Regularization robustness}
-		%title={CG iterations to achieve $||H x-b||/||b||<10^{-6}$}
-		]
-		%\addlegendentry{None}
-		\addplot[dash dot, line width=0.5mm] table[x=gamma, y=none] {CGitsvsGamma.dat};
-		%\addlegendentry{REG}
-		\addplot[dotted, line width=0.5mm] table[x=gamma, y=reg] {CGitsvsGamma.dat};
-		%\addlegendentry{PSF (1)}
-		\addplot[blue, line width=0.5mm] table[x=gamma, y=psf1] {CGitsvsGamma.dat};
-		%\addlegendentry{PSF (5)}
-		\addplot[red, line width=0.5mm] table[x=gamma, y=psf5] {CGitsvsGamma.dat};
-		%\addlegendentry{PSF (25)}
-		\addplot[purple, line width=0.5mm] table[x=gamma, y=psf25] {CGitsvsGamma.dat};
-		\end{axis}
-		\end{tikzpicture}
-		\caption{}
-		\label{fig:gamma_sweep}
-	\end{subfigure}
-	\caption{Left (\ref{fig:cg_convergence}) shows the convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
-random entries drawn from the standard normal distribution.
-%		The y-axis is $||b-Hx_k||/||b||$, where $x_k$ is the $k$th PCG iterate. 
-	Here we use $\gamma=7.3 \times 10^3$, which is determined by the Morozov discrepancy principle.
-Right (\ref{fig:gamma_sweep}) shows the number of PCG iterations required to achieve $||\bm{b}-\bm{H} \bm{x}_k||/||\bm{b}|| < 10^{-6}$, where $\bm{x}_k$ is the $k$th PCG iterate, for a range of different $\gamma$.
-Results in these figures are shown for our preconditioner with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
-The preconditioner is constructed using $\bm{H}_\text{gn}$.
-For a fair comparison, even as $\gamma$ changes, $\bm{H}$ and $\bm{H}_\text{gn}$ are always evaluated at the same $\bm{\basalfriction}$, which is the optimal point for the inverse problem with $\gamma=7.3 \times 10^3$.}
-	\label{fig:krylov_convergence}
-\end{figure}
-
-Next, we build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\bm{\basalfriction}$.
-We use PCG to solve a linear system with the Hessian as the coefficient operator and a right hand vector with random i.i.d.
-entries drawn from the standard normal distribution.
-In Figure \ref{fig:cg_convergence} we compare the convergence of PCG for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners.
-PCG converges fastest with the PSF preconditioners, with PSF (25) converging fastest, followed by PSF (5), followed by PSF (1), as expected.
-PCG converges much slower with no preconditioning and regularization preconditioning than it does with PSF preconditioning, with no preconditioning outperforming regularization preconditioning.
-In Figure \ref{fig:gamma_sweep}, we perform PCG solves on linear systems of the same form, except now we vary $\gamma$.
-The performance of our PSF preconditioners is good and relatively stable over a wide range of $\gamma$'s.
-All PSF preconditioners perform the same for medium and large values of $\gamma$.
-For small values of $\gamma$ PSF (25) performs slightly better than PSF (5), which performs slightly better than PSF (1).
-As expected, regularization preconditioning performs well for large $\gamma$ and poorly for small $\gamma$.
-Our PSF preconditioners outperform no preconditioning and regularization preconditioning for medium and small $\gamma$, and perform similarly to regularization preconditioning for large $\gamma$.
-
-\begin{figure}
-	\begin{subfigure}{0.59\textwidth}
-		\begin{center}
-			\begin{tikzpicture}[baseline, trim axis right]
-			\begin{axis}[width=0.99\linewidth, height=0.8\linewidth,
 		ymode=log,
@@
 		ymax=1.e4,
+		title={Generalized eigenvalues, $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$}
 		xlabel={$k$, generalized eigenvalue \#},
@@
 		legend style={font=\small},
-			ylabel near ticks, ylabel shift={-6pt},
+		ylabel near ticks, %ylabel shift={-6pt},
 		axis x line*=bottom,
-			axis y line*=left,
-			legend style={at={(0.95,1)},anchor=north east},
-			title={Generalized eigenvalues, $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$}
+		axis y line*=right,
+		%legend style={at={(0.95,1)},anchor=north east},
 		]
-			\addlegendentry{REG}
+		%\addlegendentry{REG}
 		\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues.dat};
-			\addlegendentry{PSF (1)}
+		%\addlegendentry{PSF (1)}
 		\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues.dat};
-			\addlegendentry{PSF (5)}
+		%\addlegendentry{PSF (5)}
 		\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues.dat};
-			\addlegendentry{PSF (25)}
+		%\addlegendentry{PSF (25)}
 		\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues.dat};
@@
     \end{subfigure}
+	\caption{Left (\ref{fig:cg_convergence}) shows the convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
+random entries drawn from the standard normal distribution.
+The reference solution $\bm{x}_{\text{true}}$ is determined by solving $\bm{H} \bm{x} = \bm{b}$ to an accuracy of ?
+using method ?
+(\todo{include details about $\bm{x}_{\text{true}}$ here}).
+	Here we use $\gamma=7.3 \times 10^3$, which is determined by the Morozov discrepancy principle.
+Results in these figures are shown for our preconditioner with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
+The preconditioner is constructed using $\bm{H}_\text{gn}$.}
+	\label{fig:krylov_convergence}
+\end{figure} 
+%
+%
+Next, we build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\bm{\basalfriction}$.
+We use PCG to solve a linear system with the Hessian as the coefficient operator and a right hand vector with random i.i.d.
+entries drawn from the standard normal distribution.
+In Figure~\ref{fig:cg_convergence} we compare the convergence of PCG for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners.
+PCG converges fastest with the PSF preconditioners, with PSF (25) converging fastest, followed by PSF (5), followed by PSF (1), as expected.
+PCG converges much slower with no preconditioning and regularization preconditioning than it does with PSF preconditioning, with no preconditioning outperforming regularization preconditioning.
+In Figure~\ref{fig:gamma_sweep}, we perform PCG solves on linear systems of the same form, except now we vary $\gamma$.
+The performance of our PSF preconditioners is good and relatively stable over a wide range of $\gamma$'s.
+All PSF preconditioners perform the same for medium and large values of $\gamma$.
+For small values of $\gamma$ PSF (25) performs slightly better than PSF (5), which performs slightly better than PSF (1).
+As expected, regularization preconditioning performs well for large $\gamma$ and poorly for small $\gamma$.
+Our PSF preconditioners outperform no preconditioning and regularization preconditioning for medium and small $\gamma$, and perform similarly to regularization preconditioning for large $\gamma$.
+
+\begin{figure}
 	\begin{subfigure}{0.39\textwidth}
@@
 
-In Figure \ref{fig:generalized_eigenvalues} we show the generalized eigenvalues for the generalized eigenvalue problem
+In Figure~\ref{fig:generalized_eigenvalues} we show the generalized eigenvalues for the generalized eigenvalue problem
 $
@@
 The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
-In Figure \ref{fig:condition_number} we show the condition numbers of $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$.
+In Figure~\ref{fig:condition_number} we show the condition numbers of $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$.
 The PSF preconditioners yield the the smallest condition numbers, with more batches yielding smaller condition numbers.
@@
 Reynolds number 50 and boundary conditions as in
-Figure~\ref{fig:adv_inversion} (left); see \cite{PetraStadler11} for
+Figure~\ref{fig:adv_inversion} (left); see~\cite{PetraStadler11} for
 details.  The time evolution of the state variable $\concentration$ from a given
@@
 Here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix.
-For more details on H-matrices, we recommend \cite{HMATRIXGOOD}.
+For more details on H-matrices, we recommend~\cite{HMATRIXGOOD}.
 %
@@
 %The process of constructing an H-matrix representation of $\mathbf{\Aker}$ proceeds as follows.
-First, we construct hierarchical partitionings of the degrees of freedom for the columns and rows of the matrix (cluster trees, Section \ref{sec:cluster_trees}).
-Second, we construct a hierarchical partitioning of the blocks of the matrix, in such a way that many of the blocks in the partitioning are expected to be low rank, and the remaining high rank blocks are small (block cluster tree, Section \ref{sec:block_cluster_tree}).
-Finally, we form low rank approximations of the blocks of the matrix that are expected to be low rank (adaptive cross approximation, Section \ref{sec:adaptive_cross}), and fill in the remaining high rank  blocks with their numerical values.
+First, we construct hierarchical partitionings of the degrees of freedom for the columns and rows of the matrix (cluster trees, Section~\ref{sec:cluster_trees}).
+Second, we construct a hierarchical partitioning of the blocks of the matrix, in such a way that many of the blocks in the partitioning are expected to be low rank, and the remaining high rank blocks are small (block cluster tree, Section~\ref{sec:block_cluster_tree}).
+Finally, we form low rank approximations of the blocks of the matrix that are expected to be low rank (adaptive cross approximation, Section~\ref{sec:adaptive_cross}), and fill in the remaining high rank  blocks with their numerical values.
 The first two steps require geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix, but these steps do not depend on the particular values of matrix entries.
-The third step requires us to evaluate $O(\fedim \log \fedim)$ specially-chosen entries of $\mathbf{\Aker}$, and we evaluate these entries using \eqref{eq:Akerpcmat_entries}.
+The third step requires us to evaluate $O(\fedim \log \fedim)$ specially-chosen entries of $\mathbf{\Aker}$, and we evaluate these entries using~\eqref{eq:Akerpcmat_entries}.
 %
@@
 %
-%Once the block cluster tree has been constructed, low rank approximations of the admissible (low rank) blocks are formed using the adaptive cross method \cite{ACA}.
+%Once the block cluster tree has been constructed, low rank approximations of the admissible (low rank) blocks are formed using the adaptive cross method~\cite{ACA}.
 Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\mathbf{\Aker}$.
@@
 %\end{equation}
-%where $U^+$ is the pseudoinverse of $U$ \cite{goreinov1997theory,MahoneyDrineas09}.
-The quality of approximation \eqref{eq:CUR} depends on how well the columns of $C$ approximate the column space of $X$, and how well the rows of $R$ approximate the row space of $X$.
+%where $U^+$ is the pseudoinverse of $U$~\cite{goreinov1997theory,MahoneyDrineas09}.
+The quality of approximation~\eqref{eq:CUR} depends on how well the columns of $C$ approximate the column space of $X$, and how well the rows of $R$ approximate the row space of $X$.
 %
@@
 An initial set of columns $C$ is chosen.
-Keeping $C$ fixed, a set of rows $R$ is chosen so that the so that determinant of the submatrix $U$ within $C$ is as large as possible \cite{GoreinovEtAl10}.
+Keeping $C$ fixed, a set of rows $R$ is chosen so that the so that determinant of the submatrix $U$ within $C$ is as large as possible~\cite{GoreinovEtAl10}.
 Now, keeping $R$ fixed, a set of new columns $C$ is chosen so that the submatrix $U$ within $R$ is as large as possible.
 This process repeats a small number of times.
-This results in matrices $R$ and $C$ such that the error in the approximation, \eqref{eq:CUR}, is small.
+This results in matrices $R$ and $C$ such that the error in the approximation,~\eqref{eq:CUR}, is small.
 The dominant cost of this procedure is the cost of computing $a \hrank$ columns of $X$ and $a\hrank$ rows of $X$, where $a$ is the number of alternating iterations.
@@
 \end{equation*}
-In \cite{GilitschenskiHanebeck12} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$.
-% (see Figure \ref{fig:ellipsoid_intersection_test}).
+In~\cite{GilitschenskiHanebeck12} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$.
+% (see Figure~\ref{fig:ellipsoid_intersection_test}).
 
@@
 where $v := \eigenvectormatrix^T\left(\spatialmean_1 - \spatialmean_2\right)$.
-To check whether $E_1$ and $E_2$ intersect, we compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, form $v$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on $(0,1)$ using Brent's method \cite{Brent71}.
+To check whether $E_1$ and $E_2$ intersect, we compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, form $v$, then minimize $K(s)$ in the form~\eqref{eq:Ks_generalized} on $(0,1)$ using Brent's method~\cite{Brent71}.
 If $K(s^*) <0$ at the minimizer, $s^*$, then $E_1 \cap E_2 = \{\}$.
 Otherwise $E_1 \cap E_2 \neq \{\}$.
-%The resulting algorithm for checking whether $E_1$ and $E_2$ intersect is summarized in Algorithm \ref{alg:ellipsoid_intersection_test}.
+%The resulting algorithm for checking whether $E_1$ and $E_2$ intersect is summarized in Algorithm~\ref{alg:ellipsoid_intersection_test}.
 
@@
 $
-If $\Aop$ is positive semi-definite, we adapt the rational method from \cite[Section 13.8]{HackbuschKress07} to approximate the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$, which we denote by $|\mathbf{A}_H^{\text{sym}}|$.
+If $\Aop$ is positive semi-definite, we adapt the rational method from~\cite[Section 13.8]{HackbuschKress07} to approximate the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$, which we denote by $|\mathbf{A}_H^{\text{sym}}|$.
 That is, $|\mathbf{A}_H^{\text{sym}}|$ has the same eigenvectors as $\mathbf{A}_H^{\text{sym}}$, but the eigenvalues of $|\mathbf{A}_H^{\text{sym}}|$ are the absolute values of the eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
@@
 \end{equation}
-From the results in \cite[Section 13.8]{HackbuschKress07} it is straightforward to show that
+From the results in~\cite[Section 13.8]{HackbuschKress07} it is straightforward to show that
 \begin{equation}
@@
 The method is relatively insensitive to $\gamma$ and numerically we find that the method works well for any $\gamma \in [1,2]$. We use $\gamma=1.5$ in our numerical results. 
-Next, we compute $\mathbf{T}$ per \eqref{eq:ratT999}, compute $\mathbf{T}^{2^\ratord}$ per \eqref{eq:rational343} via repeated squaring, compute $\mathbf{\Pi}_\ratord$ per \eqref{eq:rational343}, and form an approximation of $|\mathbf{A}_H^{\text{sym}}|$ per \eqref{eq:rational76567}.
+Next, we compute $\mathbf{T}$ per~\eqref{eq:ratT999}, compute $\mathbf{T}^{2^\ratord}$ per~\eqref{eq:rational343} via repeated squaring, compute $\mathbf{\Pi}_\ratord$ per~\eqref{eq:rational343}, and form an approximation of $|\mathbf{A}_H^{\text{sym}}|$ per~\eqref{eq:rational76567}.
 Fast H-matrix methods are used to perform the required matrix-matrix additions and multiplications, and matrix inversion.
@@
 
-%Finally, we approximate $|\mathbf{A}_H^{\text{sym}}|$ per \eqref{eq:rational76567}, using fast H-matrix operations to perform the required matrix-matrix addition and matrix inversion.
+%Finally, we approximate $|\mathbf{A}_H^{\text{sym}}|$ per~\eqref{eq:rational76567}, using fast H-matrix operations to perform the required matrix-matrix addition and matrix inversion.
 
@@
 %%Here $\ratord$ is a positive integer; the larger $\ratord$ is, the better the approximation $\Pi \approx \Pi_\ratord$. 
-%%%This is illustrated in Figure \ref{fig:spd_rat}. 
+%%%This is illustrated in Figure~\ref{fig:spd_rat}. 
 %%Replacing scalar operations with matrix operations, we have an approximation $\mathbf{\Pi}_\ratord \approx \mathbf{\Pi}$ given by
@@
 %$
-%Using $\mathbf{\Pi}_\ratord$ instead of $\mathbf{\Pi}$ in \eqref{eq:I_Pi_abs} allows us to approximate $|\mathbf{A}_H^{\text{sym}}|$.
+%Using $\mathbf{\Pi}_\ratord$ instead of $\mathbf{\Pi}$ in~\eqref{eq:I_Pi_abs} allows us to approximate $|\mathbf{A}_H^{\text{sym}}|$.
 First, we compute $\lambda_\text{min}$ using the Lanczos method, and set $a=\gamma \lambda_\text{min}$ and $b=0$, where $\gamma\ge 1$ is a parameter that must be chosen.
@@
 Larger $\ratord$ is unnecessary and may make the method unstable.
-The approximate spectral projector $\mathbf{\Pi}_\ratord$ is formed from $\mathbf{T}^{2^\ratord}$ per \eqref{eq:rational343}, using fast H-matrix operations to perform the required matrix-matrix addition and matrix inversion.
+The approximate spectral projector $\mathbf{\Pi}_\ratord$ is formed from $\mathbf{T}^{2^\ratord}$ per~\eqref{eq:rational343}, using fast H-matrix operations to perform the required matrix-matrix addition and matrix inversion.
 Finally, a positive semi-definite approximation $\mathbf{A}_H^{\text{sym}+}$ of $\mathbf{A}_H^{\text{sym}}$ is formed as
@@
 %Each of these linear systems is solved using an iterative method. By ``slowly varying,'' we mean that $\mathbf{B}^{(k+1)}$ does not differ much from $\mathbf{B}^{(k)}$. 
-%For example, in a Newton-Krylov method for solving an optimization problem, one uses a Krylov method such as conjugate gradient to solve a sequence of systems of the form \eqref{eq:Ax_equals_b}, where $\mathbf{B}^{(k)}$ and $\mathbf{f}^{(k)}$ are the Hessian and the negative gradient of the objective function in the optimization problem, respectively, evaluated at the $k$th Newton iterate.
+%For example, in a Newton-Krylov method for solving an optimization problem, one uses a Krylov method such as conjugate gradient to solve a sequence of systems of the form~\eqref{eq:Ax_equals_b}, where $\mathbf{B}^{(k)}$ and $\mathbf{f}^{(k)}$ are the Hessian and the negative gradient of the objective function in the optimization problem, respectively, evaluated at the $k$th Newton iterate.
 %As the optimization solver converges, $\mathbf{B}^{(k)}$ converges to the Hessian at the optimal point, and therefore changes little from one iteration to the next.
@@
 %\end{equation*}
-%denote a H-matrix preconditioner used to accelerate the solution of \eqref{eq:Ax_equals_b} with an iterative method.
+%denote a H-matrix preconditioner used to accelerate the solution of~\eqref{eq:Ax_equals_b} with an iterative method.
 The iterative method for solving the $k$th linear system will typically involve, as a sub-problem, applying $\mathbf{B}^{(k)}$ to a collection of vectors $\{\mathbf{x}_i\}_{i=1}^m$, thereby generating vectors $\{\mathbf{y}_i = \mathbf{B}^{(k)} \mathbf{x}_i\}_{i=1}^m$.
 Instead of rebuilding the H-matrix preconditioner for each successive linear system, we \emph{recycle} the information generated by the iterative method while solving the current linear system, performing a low-rank update to $\mathbf{P}^{(k)}$ to create $\mathbf{P}^{(k+1)}$.
-To do so, we build on the rank-1 update formulas discussed in \cite[Chapter 6]{NocedalWright99}, which we generalize to many vector ``rank-$m$'' updates.
+To do so, we build on the rank-1 update formulas discussed in~\cite[Chapter 6]{NocedalWright99}, which we generalize to many vector ``rank-$m$'' updates.
 In what follows, let $\mathbf{X}$ and $\mathbf{Y}$ denote the $\fedim \times m$ matrices with columns given by the vectors $\mathbf{x}_i$ and $\mathbf{y}_i$, respectively.
@@
 One should choose which formula to use based on the properties of the matrix $\mathbf{B}^{(k)}$.
-Formula \eqref{eq:res_update} does not preserve symmetry, and should therefore be used when $\mathbf{B}^{(k)}$ is not symmetric.
-Formula \eqref{eq:srk_update} preserves symmetry but not positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric and indefinite.
-Formula \eqref{eq:dfp_update}, which is used in this paper, preserves both symmetry and positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric positive definite.
-Formulas \eqref{eq:res_update} and \eqref{eq:srk_update} perform rank-$m$ updates, while formula \eqref{eq:dfp_update} performs a rank-$2m$ update.
-%We note that when $m=1$, formulas \eqref{eq:srk_update} are \eqref{eq:dfp_update} are known as SR1 (symmetric rank-$1$) and DFP (Davidson-Fletcher-Powell) updates, respectively \cite[Chapter 6]{NocedalWright99}.
+Formula~\eqref{eq:res_update} does not preserve symmetry, and should therefore be used when $\mathbf{B}^{(k)}$ is not symmetric.
+Formula~\eqref{eq:srk_update} preserves symmetry but not positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric and indefinite.
+Formula~\eqref{eq:dfp_update}, which is used in this paper, preserves both symmetry and positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric positive definite.
+Formulas~\eqref{eq:res_update} and~\eqref{eq:srk_update} perform rank-$m$ updates, while formula~\eqref{eq:dfp_update} performs a rank-$2m$ update.
+%We note that when $m=1$, formulas~\eqref{eq:srk_update} are~\eqref{eq:dfp_update} are known as SR1 (symmetric rank-$1$) and DFP (Davidson-Fletcher-Powell) updates, respectively~\cite[Chapter 6]{NocedalWright99}.
 

### 2023-05-04 paper/localpsf.tex
generalized eigenvalue processing scripts, including the relevant data file in the paper home directory, removing no longer needed figure labels and highlighting to do items

@@
 		xmax=1450,
-		ymax=1.e4,
-		title={Generalized eigenvalues, $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$}
+		ymax=1.e7,
+		title={Generalized eigenvalues, $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$},
 		xlabel={$k$, generalized eigenvalue \#},
@@
 		%\addlegendentry{REG}
-		\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues.dat};
+		\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues_0.05noise.dat};
+		\addplot[dash dot, line width=0.5mm] table[x=k, y=none] {generalizedEigenvalues_0.05noise.dat};
 		%\addlegendentry{PSF (1)}
-		\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues.dat};
+		\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues_0.05noise.dat};
 		%\addlegendentry{PSF (5)}
-		\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues.dat};
+		\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues_0.05noise.dat};
 		%\addlegendentry{PSF (25)}
-		\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues.dat};
+		\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues_0.05noise.dat};
 		\end{axis}
@@
 Results in these figures are shown for our preconditioner with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
-The preconditioner is constructed using $\bm{H}_\text{gn}$.}
+The preconditioner is constructed using $\bm{H}_\text{gn}$.
+	Right (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \genericvec_k = \lambda_k \bm{\preconditioner} \genericvec_k$.
+Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), or the regularization Hessian (REG).}
 	\label{fig:krylov_convergence}
@@
 \begin{figure}
-	\begin{subfigure}{0.39\textwidth}
 	\begin{center}
@@
 		\end{center}
-		\caption{}
+	\caption{Condition number for $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$ for for our preconditioner with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), no preconditioner (NONE) and regularization (REG).
+All operators are evaluated at the $\bm{\basalfriction}$ that solves the inverse problem for $\gamma=7.3 \times 10^3$, as determined by the Morozov discrepancy principle.
+\todo{COND(H) LABELS!}}
 	\label{fig:condition_number}
-	\end{subfigure}
-	\caption{Left (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \genericvec_k = \lambda_k \bm{\preconditioner} \genericvec_k$.
-Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), or the regularization Hessian (REG).
-Right (\ref{fig:condition_number}): condition number for $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$ for these different preconditioners, and no preconditioner (NONE).
-All operators are evaluated at the $\bm{\basalfriction}$ that solves the inverse problem for $\gamma=7.3 \times 10^3$, as determined by the Morozov discrepancy principle.
-COND(H) LABELS!}
-	\label{fig:stokes_spectrum}
 \end{figure}

### 2023-05-04 numerical_examples/stokes/data/stokes_deflation/Preconditionedspectrum/process.py

@@ -1,0 +1,21 @@
+import numpy as np
+
+exts = ["none", "reg", "psf1", "psf5", "psf25"]
+ndata = len(exts)
+datas = [np.loadtxt("ee_"+exts[i]+"_0.049999999999999996") for i in range(ndata)]
+N = len(datas[0])
+
+
+f = open("generalizedEigenvalues_0.05noise.dat", "w")
+f.write("k ")
+for i in range(ndata):
+    f.write(exts[i] + " ")
+f.write("\n")
+
+for i in range(N):
+    f.write("{0:d} {1:f} {2:f} {3:f} {4:f} {5:f}\n".format(i, datas[0][i], datas[1][i], datas[2][i], datas[3][i], datas[4][i]))
+
+
+
+
+

### 2023-05-04 paper/localpsf.tex
condition number table with varying noise level

@@
 			\renewcommand{\arraystretch}{1.25}
-			\begin{tabular}{c| c c}
-				& $\kappa(\bm{\preconditioner}^{-1} \bm{H})$ &	$\kappa(\bm{\preconditioner}^{-1} \bm{H}_{\text{gn}})$ \\
-				\hline 
-				REG        & 7.02e+3 & 7.02e+3 \\
-				NONE       & 8.71e+2 & 9.22e+2 \\
-				PSF $(1)$  & 1.75e+1 & 1.76e+1 \\
-				PSF $(5)$  & 1.39e+1 & 1.39e+1 \\
-				PSF $(25)$ & 1.03e+1 & 1.06e+1
+			\begin{tabular}{c| c c c c c}
+				noise    & \multicolumn{5}{c}{COND$(\bm{\preconditioner}^{-1} \bm{H})$ } \\ \cline{2-6}
+				level    & REG     &	NONE  & PSF $(1)$ & PSF $(5)$ & PSF $(25)$ \\ \hline 
+				$25\%$   & 1.68e+3 & 2.16e+3  & 1.37e+0   & 1.30e+0   & 1.21e+0    \\ 
+				$11\%$   & 9.19e+3 & 9.08e+2  & 1.85e+0   & 1.49e+0   & 1.27e+0    \\   
+				$5.0\%$  & 3.66e+4 & 4.67e+2  & 3.46e+0   & 2.62e+0   & 1.91e+0    \\ 
+				$2.2\%$  & 1.41e+5 & 8.25e+2  & 1.16e+1   & 4.99e+0   & 3.20e+0    \\  
+				$1.0\%$  & 6.06e+5 & 1.70e+3  & 3.44e+1   & 5.38e+1   & 7.06e+0    \\   
 			\end{tabular}
@@
 		\end{center}
-	\caption{Condition number for $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$ for for our preconditioner with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), no preconditioner (NONE) and regularization (REG).
-All operators are evaluated at the $\bm{\basalfriction}$ that solves the inverse problem for $\gamma=7.3 \times 10^3$, as determined by the Morozov discrepancy principle.
-\todo{COND(H) LABELS!}}
+	\caption{Condition number for $\bm{\preconditioner}^{-1} \bm{H}$ for our preconditioner with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), no preconditioner (NONE) and regularization (REG).
+All operators are evaluated at the $\bm{\basalfriction}$ that solves the inverse problem for
+	the regularization value $\gamma$ as determined by noise level by means of the Morozov discrepancy principle.}
 	\label{fig:condition_number}

### 2023-05-05 paper/localpsf.tex
noise figures :)

@@
 \begin{figure}
-	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.1]{m1_psf.png}
+	\begin{subfigure}{0.18\textwidth}
+		\includegraphics[scale=0.075]{mstar_0.25noise_cropped.png}
 		%		\caption{}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.1]{m3_psf.png}
+	\begin{subfigure}{0.18\textwidth}
+		\includegraphics[scale=0.075]{mstar_0.12noise_cropped.png}
 		%		\caption{}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.1]{m4_psf.png}
+	\begin{subfigure}{0.18\textwidth}
+		\includegraphics[scale=0.075]{mstar_0.05noise_cropped.png}
 		%		\caption{}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.1]{m6_psf.png}
+	\begin{subfigure}{0.18\textwidth}
+		\includegraphics[scale=0.075]{mstar_0.02noise_cropped.png}
 		%		\caption{}
@@
 	\end{subfigure}
-	%\begin{subfigure}{0.05\textwidth}
-	%	\includegraphics[scale=0.1]{mColorBar.png}
-	%\end{subfigure}
-	\caption{The log basal sliding parameter after $1$ (left), $3$ (second from left), and $4$ (third from left) Newton iterations, and the final Newton iterate (right).
-The PSF (5) preconditioner is constructed between Newton iterations 3 and 4.}
+	\begin{subfigure}{0.18\textwidth}
+	\includegraphics[scale=0.075]{mstar_0.01noise_cropped.png}
+	%		\caption{}
+	%		\label{fig:newtoniter4}
+    \end{subfigure}
+	\caption{The log basal sliding parameter computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $11\%$ (second from left), $5.0\%$ (third from left), $2.2\%$ (fourth from left), and $1.0\%$ (right).
+The PSF (5) preconditioner is constructed between Newton iterations 3 and 4.
+\todo{remove reference to newon iterates figure as this figure has a new significance}}
 	\label{fig:newton_iterates} 

### 2023-05-05 paper/localpsf.tex
a little bit of cleanup and consistently using \| instead of || for indicating norms

@@
 subsequent iterations. The Newton iterations are terminated when
-$||\bm{g}_k||/||\bm{g}_0|| < 10^{-8}$. Systems~\eqref{eq:newton_system} are solved inexactly using an inner PCG
+$\|\bm{g}_k\|/\|\bm{g}_0\| < 10^{-8}$. Systems~\eqref{eq:newton_system} are solved inexactly using an inner PCG
 iteration, which terminates when the norm of the linear system residual is less than $\min \left(0.5,
-\sqrt{||\bm{g}_k||}\right)||\bm{g}_k||$.
+\sqrt{\|\bm{g}_k\|}\right)\|\bm{g}_k\|$.
 The advection-diffusion inverse problem is linear, so Newton's method converges in one iteration.
@@
 		\caption{$\basalfriction_\text{true}$.}
+		\label{fig:true_beta}
 	\end{subfigure}
@@
 %\begin{equation*}
-%	J_d(\basalfriction) := \frac{1}{2}\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS.
+%	J_d(\basalfriction) := \frac{1}{2}\int_{\Gamma_{t}}\|\observations_\text{obs} - \tangentop\velocity\|_{2}^{2}\,dS.
 %\end{equation*}
@@
 %\begin{equation*}
-%	\int_{\Gamma_{t}}||\observations_\text{obs} - \tangentop\velocity||_{2}^{2}\,dS = \int_{\Gamma_{t}}||\xi||_{2}^{2}\,dS.
+%	\int_{\Gamma_{t}}\|\observations_\text{obs} - \tangentop\velocity\|_{2}^{2}\,dS = \int_{\Gamma_{t}}\|\xi\|_{2}^{2}\,dS.
 %\end{equation*}
@@
     \end{subfigure}
-	\caption{The log basal sliding parameter computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $11\%$ (second from left), $5.0\%$ (third from left), $2.2\%$ (fourth from left), and $1.0\%$ (right).
-The PSF (5) preconditioner is constructed between Newton iterations 3 and 4.
+	\caption{The log basal sliding parameter, with colorscale as in Figure~\ref{fig:true_beta}, computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $11\%$ (second from left), $5.0\%$ (middle), $2.2\%$ (second from right), and $1.0\%$ (right).
 \todo{remove reference to newon iterates figure as this figure has a new significance}}
@@
 				Iter & 
-				\#CG & \#Stokes & $||\bm{g}||$ & 
-				\#CG & \#Stokes & $||\bm{g}||$ & 
-				\#CG & \#Stokes & $||\bm{g}||$ \\
+				\#CG & \#Stokes & $\|\bm{g}\|$ & 
+				\#CG & \#Stokes & $\|\bm{g}\|$ & 
+				\#CG & \#Stokes & $\|\bm{g}\|$ \\
 				0 &
@@
 Columns titled \#CG show the number of PCG iterations used to solve the Newton system for $\bm{\widehat{\basalfriction}}_k$.
-Columns titled $||\bm{g}||$ show the $l^2$ norm of the gradient at $\bm{\basalfriction}_k$.
+Columns titled $\|\bm{g}\|$ show the $l^2$ norm of the gradient at $\bm{\basalfriction}_k$.
 Columns titled \#Stokes show the total number of Stokes PDE solves performed in each Newton iteration.
@@
 		xlabel={$k$, CG iteration},
-		ylabel={$||\bm{x}_k-\bm{x}_{\text{true}}||/||\bm{x}_{\text{true}}||$},
+		ylabel={$\|\bm{x}_k-\bm{x}_{\text{true}}\|/\|\bm{x}_{\text{true}}\|$},
 		legend style={at={(1.1,1)},anchor=north east},
@@
 		legend style={font=\small},
-		ylabel near ticks, %ylabel shift={-6pt},
+		ylabel near ticks, ylabel shift={-6pt},
 		axis x line*=bottom,
@@
 (\todo{include details about $\bm{x}_{\text{true}}$ here}).
-	Here we use $\gamma=7.3 \times 10^3$, which is determined by the Morozov discrepancy principle.
+	Here we use $\gamma=7.3 \times 10^3$, which is determined by the Morozov discrepancy principle when using $5\%$ noise.
 Results in these figures are shown for our preconditioner with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
@@
 All operators are evaluated at the $\bm{\basalfriction}$ that solves the inverse problem for
-	the regularization value $\gamma$ as determined by noise level by means of the Morozov discrepancy principle.}
+	the regularization value $\gamma$ as determined by the respective noise level and the Morozov discrepancy principle.}
 	\label{fig:condition_number}
@@
 		xlabel={CG iteration},
-		ylabel={$||\bm{b}-\bm{H} \bm{x}_k||/||\bm{b}||$},
+		ylabel={$\|\bm{b}-\bm{H} \bm{x}_k\|/\|\bm{b}\|$},
 		title={CG convergence},
@@
 		title={Regularization robustness}
-		%title={CG iterations to achieve $||H x-b||/||b||<10^{-6}$}
+		%title={CG iterations to achieve $\|H x-b\|/\|b\|<10^{-6}$}
 		]
@@
 BLAH BLAH BLAH
-		%		The y-axis is $||b-Hx_k||/||b||$, where $x_k$ is the $k$th PCG iterate. 
+		%		The y-axis is $\|b-Hx_k\|/\|b\|$, where $x_k$ is the $k$th PCG iterate. 
 		Here we use $\gamma=9999 \times 10^9$, which is determined by the Morozov discrepancy principle.
-Right (\ref{fig:adv_kappa_sweep}) shows the number of PCG iterations required to achieve $||\bm{b}-\bm{H} \bm{x}_k||/||\bm{b}|| < 10^{-6}$, where $\bm{x}_k$ is the $k$th PCG iterate, for a range of different $\kappa$.
+Right (\ref{fig:adv_kappa_sweep}) shows the number of PCG iterations required to achieve $\|\bm{b}-\bm{H} \bm{x}_k\|/\|\bm{b}\| < 10^{-6}$, where $\bm{x}_k$ is the $k$th PCG iterate, for a range of different $\kappa$.
 Results in these figures are shown for our preconditioner with 5, 15, and 30 batches (PSF (5), PSF (15), and PSF(30), respectively), regularization preconditioning (REG), and no preconditioning (NONE).

### 2023-05-06 paper/localpsf.tex
merged the ice mountain geometry figure with the true parameter and true forward solution plots as these are all data used to setup and or validate the inverse problem... also have updated the NewtonCG plot

@@
 %
-\begin{figure}
-	\begin{center} 
-	\includegraphics[scale=0.23]{meshHeight_view2_edges.png}
-	\end{center}
-	\caption{Bird's eye view of the ice sheet, discretized by a mesh of tetrahedra.
+%\begin{figure}
+%	\begin{center} 
+%	\includegraphics[scale=0.23]{meshHeight_view2_edges.png}
+%	\end{center}
+%	\caption{Bird's eye view of the ice sheet, discretized by a mesh of tetrahedra.
 Color indicates the height of the base of the ice sheet (i.e., the mountain topography).
 The radius of the domain is $10^4$ meters, the maximum height of the mountain is $2.1 \times 10^3$ meters, and the average thickness of the ice sheet is 250 meters.}
-	\label{fig:ice_mountain_mesh}
-\end{figure} 
+%	\label{fig:ice_mountain_mesh}
+%\end{figure} 
 %
@@
 \end{subequations}
+%
+%
 The solution to these equations is the pair $(\velocity, \pressure)$, where $\velocity$ is the ice flow velocity field, and $\pressure$ is the pressure field.
@@
 \begin{figure}
+	\begin{center} 
+    \begin{subfigure}{0.99\textwidth}
+    \begin{center} 
+	\includegraphics[scale=0.23]{meshHeight_view2_edges.png}
+	\end{center}
+	\caption{Ice sheet model geometry}
+	\label{fig:ice_mountain_mesh}
+    \end{subfigure} \\
 	\begin{subfigure}{0.49\textwidth}
@@
 	\end{subfigure}
-	\caption{True parameter, $\basalfriction_\text{true}$, (left) and true velocity $\velocity_\text{true}$ (right).
-In the right plot, arrows indicate the direction of $\velocity_\text{true}$ and color indicates the magnitude of $\velocity_\text{true}$.}
+    \end{center}
+	\caption{
+		Top: bird's eye view of the ice sheet, discretized by a mesh of tetrahedra.
+Color indicates the height of the base of the ice sheet (i.e., the mountain topography).
+The radius of the domain is $10^4$ meters, the maximum height of the mountain is $2.1 \times 10^3$ meters, and the average thickness of the ice sheet is 250 meters.
+		True parameter, $\basalfriction_\text{true}$, (bottom left) and true velocity $\velocity_\text{true}$ (bottom right).
+In the bottom right plot, arrows indicate the direction of $\velocity_\text{true}$ and color indicates the magnitude of $\velocity_\text{true}$.}
 	\label{fig:true_beta_u}
@@
 				0 &
-				1 & 4 & 2.0e+7 &
-				1 & 4 & 2.0e+7 &
-				1 & 4 & 2.0e+7 \\
+				1 & 4 & 1.9e+7 &
+				3 & 8 & 1.9e+7 &
+				1 & 4 & 1.9e+7 \\
 				1 &
-				2 & 6 & 5.8e+6 &
-				2 & 6 & 7.9e+6 &
-				2 & 6 & 5.8e+6 \\
+				2 & 6  & 6.1e+6 &
+				8 & 18 & 8.4e+6 &
+				2 & 6  & 6.1e+6 \\
 				2 &
-				4 & 10 & 2.4e+6 &
-				5 & 12 & 3.9e+6 &
-				4 & 10 & 2.4e+6 \\
+				3 & 8 & 2.6e+6 &
+				16 & 34 & 4.1e+6 &
+				3 & 8 & 2.6e+6 \\
 				3 &
-				2 & 6+22 & 5.8e+5 &
-				13 & 28 & 1.6e+6 &
-				13 & 28 & 5.8e+5 \\
+				1 & 4+22 & 9.2e+5 &
+				34 & 70 & 1.7e+6 &
+				11 & 24 & 9.2e+5 \\
 				4 &
-				5 & 12 & 5.6e+4 &
-				42 & 86 & 4.8e+5 &
-				33 & 68 & 9.1e+4 \\
+				2 & 6 & 1.5e+5 &
+				54 & 110 & 5.0e+5 &
+				30 & 62 & 1.4e+5 \\
 				5 &
-				10 & 22 & 3.5e+3 &
-				84 & 170 & 7.7e+4 &
-				57 & 116 & 6.2e+3 \\
+				4 & 10 & 7.1e+3 &
+				79 & 160 & 8.0e+4 &
+				34 & 70 & 1.4e+4 \\
 				6 &
-				14 & 30 & 2.7e+1 &
-				125 & 252 & 5.0e+3 &
-				76 & 154 & 1.1e+2 \\
+				5 & 12 & 6.0e+1 &
+				112 & 226 & 5.2e+3 &
+				61 & 124 & 2.3e+2 \\
 				7 &
-				0 & 2 & 4.1e-2 &
-				194 & 390 & 7.9e+1 &
-				117 & 236 & 2.3e-1 \\
+				0 & 2 & 8.9e-2 &
+				162 & 326 & 8.0e+1 &
+				84 & 170 & 7.3e-1 \\
 				8 & 
@@
 				0 & 2 & 1.6e-1 &
-				0 & 2 & 4.1e-2 \\
+				0 & 2 & 3.0e-2 \\
 				\hline
 				Total & 
-				38 & 114 & --- &
-				466 & 950 & --- &
-				303 & 624 & --- \\
+				18 & 74 & --- &
+				468 & 954 & --- &
+				226 & 470 & --- \\
 			\end{tabular}

### 2023-05-06 paper/localpsf.tex

@@
 This consists of Stokes solves for $u$ performed during the linesearch going from $\bm{\basalfriction}_{k-1}$ to $\bm{\basalfriction}_k$, plus one adjoint Stokes solve to compute the gradient at $\bm{\basalfriction}_k$, plus one incremental forward and one incremental adjoint Stokes solve per PCG iteration for solving the Newton system.
-In the PSF (5) portion of row 3, we write $6+22$ to indicate that 6 Stokes solves were used during the standard course of the iteration, and 22 Stokes solves were used to build the PSF (5) preconditioner.}
+Under PSF (5) and in row Iter 3, we write $4+22$ to indicate that $4$ Stokes solves were used during the standard course of the iteration, and $22$ Stokes solves were used to build the PSF (5) preconditioner.}
 	\label{tab:newton_convergence_table}

### 2023-05-09 paper/localpsf.tex
advection diffusion

@@
 	\begin{subfigure}{0.49\textwidth}
-		\includegraphics[scale=0.243]{mtrue_withColorBar.png}
+		\includegraphics[scale=0.22]{mtrue_withColorBar.png}
 		\caption{$\basalfriction_\text{true}$.}
@@
 	\begin{subfigure}{0.49\textwidth}
-		\includegraphics[scale=0.243]{trueVelocity2_glyphs.png}
+		\includegraphics[scale=0.22]{trueVelocity2_glyphs.png}
 		\caption{$\velocity_\text{true}$.}
@@
 \begin{figure}
-	\begin{subfigure}{0.18\textwidth}
-		\includegraphics[scale=0.075]{mstar_0.25noise_cropped.png}
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[scale=0.13]{mstar_0.25noise_cropped.png}
 		%		\caption{}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.18\textwidth}
-		\includegraphics[scale=0.075]{mstar_0.12noise_cropped.png}
-		%		\caption{}
-		%		\label{fig:newtoniter2}
-	\end{subfigure}
-	\begin{subfigure}{0.18\textwidth}
-		\includegraphics[scale=0.075]{mstar_0.05noise_cropped.png}
+%	\begin{subfigure}{0.18\textwidth}
+%		\includegraphics[scale=0.075]{mstar_0.12noise_cropped.png}
+%		%		\caption{}
+%		%		\label{fig:newtoniter2}
+%	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[scale=0.13]{mstar_0.05noise_cropped.png}
 		%		\caption{}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.18\textwidth}
-		\includegraphics[scale=0.075]{mstar_0.02noise_cropped.png}
+%	\begin{subfigure}{0.18\textwidth}
+%		\includegraphics[scale=0.075]{mstar_0.02noise_cropped.png}
+%		%		\caption{}
+%		%		\label{fig:newtoniter4}
+%	\end{subfigure}
+	\begin{subfigure}{0.32\textwidth}
+	\includegraphics[scale=0.13]{mstar_0.01noise_cropped.png}
 	%		\caption{}
@@
     \end{subfigure}
-	\begin{subfigure}{0.18\textwidth}
-	\includegraphics[scale=0.075]{mstar_0.01noise_cropped.png}
-	%		\caption{}
-	%		\label{fig:newtoniter4}
-    \end{subfigure}
-	\caption{The log basal sliding parameter, with colorscale as in Figure~\ref{fig:true_beta}, computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $11\%$ (second from left), $5.0\%$ (middle), $2.2\%$ (second from right), and $1.0\%$ (right).
+	\caption{The log basal sliding parameter, with colorscale as in Figure~\ref{fig:true_beta}, computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $5.0\%$ (middle),and $1.0\%$ (right).
 \todo{remove reference to newon iterates figure as this figure has a new significance}}
@@
 		xlabel={$k$, CG iteration},
-		ylabel={$\|\bm{x}_k-\bm{x}_{\text{true}}\|/\|\bm{x}_{\text{true}}\|$},
+		ylabel={$\|\bm{x}_k-\bm{x}\|/\|\bm{x}\|$},
 		legend style={at={(1.1,1)},anchor=north east},
@@
 		ymax=1.e7,
-		title={Generalized eigenvalues, $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$},
+		title={Generalized eigenvalues},
 		xlabel={$k$, generalized eigenvalue \#},
@@
 random entries drawn from the standard normal distribution.
-The reference solution $\bm{x}_{\text{true}}$ is determined by solving $\bm{H} \bm{x} = \bm{b}$ to an accuracy of ?
-using method ?
-(\todo{include details about $\bm{x}_{\text{true}}$ here}).
+The reference solution $\bm{x}$ is determined by solving $\bm{H} \bm{x} = \bm{b}$ to an accuracy of $10^{-12}$ using method PCG.
 	Here we use $\gamma=7.3 \times 10^3$, which is determined by the Morozov discrepancy principle when using $5\%$ noise.
@@
 The preconditioner is constructed using $\bm{H}_\text{gn}$.
-	Right (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \genericvec_k = \lambda_k \bm{\preconditioner} \genericvec_k$.
+	Right (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$.
 Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), or the regularization Hessian (REG).}
@@
 %
-Next, we build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\bm{\basalfriction}$.
+We build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\bm{\basalfriction}$.
 We use PCG to solve a linear system with the Hessian as the coefficient operator and a right hand vector with random i.i.d.
@@
 As expected, regularization preconditioning performs well for large $\gamma$ and poorly for small $\gamma$.
-Our PSF preconditioners outperform no preconditioning and regularization preconditioning for medium and small $\gamma$, and perform similarly to regularization preconditioning for large $\gamma$.
+Our PSF preconditioners outperform no preconditioning and regularization preconditioning for medium and small $\gamma$, and perform similarly to regularization preconditioning for large~$\gamma$.
 
@@
 $
-	\bm{H} \genericvec = \lambda \bm{\preconditioner} \genericvec.
+	\bm{H} \mathbf{u} = \lambda \bm{\preconditioner} \mathbf{u}.
 $

### 2023-05-09 numerical_examples/advection_diffusion/adv_deflation.py

@@ -355,6 +355,28 @@ class AdvUniverse:
         discrepancy = new_obs - me.obs
         return np.linalg.norm(discrepancy)
 
+    def vec2func(me, u_vec: np.ndarray) -> dl.Function:
+        assert(u_vec.shape == (me.N,))
+        u_func = dl.Function(me.Vh)
+        u_func.vector()[:] = u_vec.copy()
+        return u_func
+
+    @cached_property
+    def true_initial_condition_func(me) -> dl.Function:
+        return me.vec2func(me.true_initial_condition)
+
+    @cached_property
+    def obs_func(me) -> dl.Function:
+        return me.vec2func(me.obs[-1,:])
+
+    @cached_property
+    def true_obs_func(me) -> dl.Function:
+        return me.vec2func(me.true_obs[-1,:])
+
+    @cached_property
+    def noise_func(me) -> dl.Function:
+        return me.vec2func(me.noise)
+
 
 def make_adv_universe(
         noise_level: float, # 0.05, i.e., 5% noise, is typical
@@ -365,8 +387,8 @@ def make_adv_universe(
         t_init = 0.0,
         dt: float=0.1,
         prior_correlation_length: float=0.25,
-        num_checkers_x: int=6,
-        num_checkers_y: int=6,
+        num_checkers_x: int=8,
+        num_checkers_y: int=8,
         smoothing_time: float=1e-4,
 ) -> AdvUniverse:
     mesh_and_function_space = make_adv_mesh_and_function_space(num_refinements=num_mesh_refinements)
@@ -433,33 +455,32 @@ def make_adv_universe(
 import scipy.sparse.linalg as spla
 
 noise_level=0.05
-kappa=1e-4
-t_final = 1.0 # 0.5
+kappa=1e-4 # 1e-3
+t_final = 1.0 #2.0 # 1.0 # 0.5
+num_checkers = 8
 
-ADV = make_adv_universe(noise_level, kappa, t_final)
+ADV = make_adv_universe(noise_level, kappa, t_final,
+                        num_checkers_x=num_checkers,
+                        num_checkers_y=num_checkers)
 
 #
 
 plt.figure()
-obs_func = dl.Function(ADV.Vh)
-obs_func.vector()[:] = ADV.obs.reshape(-1)
[…]

### 2023-05-19 paper/localpsf.tex
some writing updates

@@
 \phi_x(y) &\approx \phi_{x'}\left(y-\spatialmean(x)+\spatialmean(x')\right) \label{eq:translate3}\\
-\phi_x(y)/\spatialvol(x) &\approx \phi_{x'}\left(y-\spatialmean(x)+\spatialmean(x')\right) / \spatialvol(x'). \label{eq:translate4}
+\phi_x(y) &\approx \phi_{x'}\left(y-\spatialmean(x)+\spatialmean(x')\right) \spatialvol(x) / \spatialvol(x'). \label{eq:translate4}
 \end{align}
@@
 Approximation~\eqref{eq:translate4} is equivalent to~\eqref{eq:translate3}, but with $\phi_x$ replaced by its normalized version, $\phi_x/\spatialvol(x)$.
-We call~\eqref{eq:translate4} ``normalized local mean displacement invariance.''
+We call~\eqref{eq:translate4} \emph{normalized local mean displacement invariance}.
+
+We have experimented with more advanced interpolation schemes involving $\Sigma$, in which one maps known impulse responses from their support ellipsoids to the unit sphere, performs interpolation on the unit sphere, then maps the resulting interpolated impulse response to the support ellipsoid associated with the target point.
+We do not use these more advanced schemes because we find that they are much less robust to negative entries in the integral kernel.
+Although we assume the integral kernel has non-negative entries (Condition 3 in Section \ref{sec:conditions_2}), in practice the integral kernel typically some negative entries due to discretization error.
+
+%but we find they are not robust, and therefore we do not use them:
+%\begin{align*}
+%	\phi_{x}(y) &\approx \phi_{x'}\left(\spatialmean(x') + \Sigma_{x'}^{1/2} \Sigma_{x}^{-1/2} \left(y - \spatialmean(x)\right)\right) \spatialvol(x) / \spatialvol(x') \\
+%	\phi_{x}(y) &\approx \phi_{x'}\left(\spatialmean(x') + \Sigma_{x'}^{1/2} \Sigma_{x}^{-1/2} \left(y - \spatialmean(x)\right)\right) \left( \spatialvol(x) \det{\Sigma^{1/2}(x')} \right) / \left( \spatialvol(x') \det{\Sigma^{1/2}(x)} \right)
+%\end{align*}
+%In these schemes, one maps known impulse responses from their support ellipsoids to the unit sphere, performs interpolation on the unit sphere, then maps the resulting interpolated impulse response to the support ellipsoid associated with the target point.
+We find that computation of $\Sigma$ is more susceptible to numerical error than computation of $V$ and $\mu$, in the sense that spurious negative numbers in $\phi_x$ cause $\Sigma(x)$ to become highly ill-conditioned or even indefinite.
+When $\Sigma$ is used for picking sample points (Section \ref{sec:sample_point_selection}), we simply avoid picking sample points $p$ with ``bad'' $\Sigma(p)$.
+These schemes require using $\Sigma_x$ at arbitrary $x$, rather than just the sample points.
+When picking sample points, one may
 

### 2023-05-23 paper/localpsf.tex
krylov iters vs kappa and tf plots

@@
 By solving an ellipsoid packing problem, we choose as many point sources as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
-Support ellipsoids are estimated a-priori based on impulse response moments, and these moments are computed for impulse responses at all points simultaneously via a procedure that involves applying the operator to a small number of polynomial functions.
+Support ellipsoids are estimated a-priori based on impulse response moments, and these moments are computed via a procedure that involves applying the operator to a small number of polynomial functions.
 The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
@@
 But approximation~\eqref{eq:translate3} can hold in situations where neither~\eqref{eq:translate1} nor~\eqref{eq:translate2} holds.
-For example, because the expected value commutes with affine transformations,~\eqref{eq:translate3} will hold when $\Aop$ is locally translation invariant with respect to a rotated frame of reference, while~\eqref{eq:translate2} will not hold in this case.
+For example, because the expected value commutes with affine transformations,~\eqref{eq:translate3} will hold when $\Aop$ is locally translation invariant with respect to a translated frame of reference, while~\eqref{eq:translate2} will not hold in this case.
 Additionally,~\eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions.
@@
 
-We use approximation~\eqref{eq:translate4}, which is the same as~\eqref{eq:translate3}, except for the extra factors of $1/\spatialvol(x)$ on the left hand side and $1/\spatialvol(x')$ on the right hand side.
-These factors make the approximation more robust if $\spatialvol(x)$ varies widely.
+We use approximation~\eqref{eq:translate4}, which is the same as~\eqref{eq:translate3}, except for the factor $\spatialvol(x)/\spatialvol(x')$.
+This factor makes the approximation more accurate if $\spatialvol(x)$ varies widely.
 Approximation~\eqref{eq:translate4} is equivalent to~\eqref{eq:translate3}, but with $\phi_x$ replaced by its normalized version, $\phi_x/\spatialvol(x)$.
@@
 
-We have experimented with more advanced interpolation schemes involving $\Sigma$, in which one maps known impulse responses from their support ellipsoids to the unit sphere, performs interpolation on the unit sphere, then maps the resulting interpolated impulse response to the support ellipsoid associated with the target point.
-We do not use these more advanced schemes because we find that they are much less robust to negative entries in the integral kernel.
-Although we assume the integral kernel has non-negative entries (Condition 3 in Section \ref{sec:conditions_2}), in practice the integral kernel typically some negative entries due to discretization error.
+\paragraph{Future extension: approximation using $\Sigma$} One may map $E_x$ to the unit sphere via the change of variables $z = \Sigma^{-1/2}\left(x\right)\left(y - \spatialmean(x)\right) / \tau$, and approximate
+\begin{equation}
+\label{eq:sigma_interpolation}
+\phi_x(z) \det{\Sigma^{1/2}\left(x\right)}/\spatialvol(x) \approx \phi_{x'}(z') \det{\Sigma^{1/2}\left(x'\right)}/\spatialvol(x').
+\end{equation}
+Notice that equality holds in \eqref{eq:sigma_interpolation} if the impulse responses are Gaussian functions.
+Although interpolation schemes based on \eqref{eq:sigma_interpolation} have the potential to be more accurate than those based on \eqref{eq:translate4}, we were unable to construct an interpolation scheme based on \eqref{eq:sigma_interpolation} that is robust to negative entries in the integral kernel.
+The integral kernel should have non-negative entries per Condition 3 in Section \ref{sec:conditions_2}, but in practice some entries are negative due to discretization error.
+%Therefore, in this paper we do not use \eqref{eq:sigma_interpolation}. 
+%Constructing interpolation schemes based on \eqref{eq:sigma_interpolation} that are robust to discretization error is a promising direction for future research.
 
@@
 
-We use~\eqref{eq:vol_mean_var} in Section~\ref{sec:intromoments} to compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ by applying $\Aop^T$ to a small number of polynomial functions.
-Then we use~\eqref{eq:support_ellipsoid} to form ellipsoid shaped estimates for the support of the $\phi_x$, \emph{without} computing them.
+We use~\eqref{eq:vol_mean_var} to compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ by applying $\Aop^T$ to polynomial functions.
+Then we use~\eqref{eq:support_ellipsoid} to form ellipsoid shaped estimates for the support of the $\phi_x$'s, \emph{without} computing them.
 This allows us to compute large numbers of $\impulseresponse_{x_i}$ in ``batches,'' $\eta_b$ (see Figure~\ref{fig:batches_intro}).
@@
 When $\Aop$ is symmetric positive semi-definite, $\mathbf{A}_H$ may be non-symmetric and indefinite due to errors in the approximation.
-In this case, one may (optionally) modify the H-matrix representation of $\mathbf{A}_H$ to make it symmetric positive semi-definite using a rational function method that we describe in Appendix~\ref{sec:make_spd}.
+In this case, we symmetrize $\mathbf{A}_H$, find directions with large negative Rayleigh quotients using a Cayley-type shift-and-invert scheme, and flip these directions so that they have positive Rayleigh quotients via a low rank correction (see Appendix~\ref{sec:make_spd}).
 
@@
 Candidate points for the first batch $S_1$ are chosen randomly from $\candidatepts$.
-Once $\candidatepoint_i$ is selected, we remove $\candidatepoint_i$ from $\candidatepts$. Then we perform the following two checks:
+Once $\candidatepoint_i$ is selected, we remove $\candidatepoint_i$ from $\candidatepts$. Then we perform the following checks:
 \begin{enumerate}
@@
 	\item We make sure that $\spatialvol(\candidatepoint_i)$ is not too small, by checking whether $\spatialvol(\candidatepoint_i) > \epsilon \spatialvol_\text{max}$.
-Here $\spatialvol_\text{max}$ is the largest value of $\spatialvol(\candidatepoint_j)$ over all points $q$ in the initial set of candidate points, and $\epsilon$ is a small threshold parameter (we use $\epsilon=10^{-5}$).
+Here $\spatialvol_\text{max}$ is the largest value of $\spatialvol(\candidatepoint_j)$ over all points $q$ in the initial set of candidate points, and $\epsilon_V$ is a small threshold parameter (we use $\epsilon_V=10^{-5}$).
+	\item We make sure that all eigenvalues of $\spatialcov(\candidatepoint_i)$ are positive, and the aspect ratio of $E_{\candidatepoint_i}$ (ratio of the largest eigenvalue of $\spatialcov(\candidatepoint_i)$ to the smallest) is bounded by a constant $1/\epsilon_\Sigma$ (we use $1/\epsilon_\Sigma=20$).
+Negative integral kernel entries due to discretization error can cause $\spatialcov(\candidatepoint_i)$ to be indefinite or highly ill-conditioned, in which case the set $E_{\candidatepoint_i}$ poorly reflects the support of $\phi_{\candidatepoint_i}$.
 \end{enumerate}
-If $\candidatepoint_i$ passes both these checks (i.e., if $\candidatepoint_i$ is sufficiently far from other points in the batch, and $\spatialvol(\candidatepoint_i)$ is not too small) then we add $\candidatepoint_i$ to $\pointbatch_k$.
-Otherwise we discard $\candidatepoint_i$.
-This process repeats until there are no more points in $\candidatepts$.
+If $\candidatepoint_i$ passes these checks 
+%(i.e., if $\candidatepoint_i$ is sufficiently far from other points in the batch, and $\spatialvol(\candidatepoint_i)$ is not too small, and the )
+then we add $\candidatepoint_i$ to $\pointbatch_k$. Otherwise we discard $\candidatepoint_i$. This process repeats until there are no more points in $\candidatepts$.  
 %This is detailed in Algorithm~\ref{alg:point_choice}. 
@@
 We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots, \pointbatch_{\nbatch}$.
-In our implementation, for each batch, $\candidatepts$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, except for points in previously chosen batches.
+For each batch, $\candidatepts$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, except for points in previous batches.
 
@@
 and $f \in \mathbb{R}^{\numnbr}$ has entries $f_i$ from~\eqref{eq:fxyxp}.
-Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a certain least squares sense~\cite{Duchon77}.
+Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a least squares sense~\cite{Duchon77}.
 %For more details on radial basis function interpolation, see~\cite{Wendland04}. 
@@
 
+{\Huge Key parameter choices: $\kappa=3.2e-3$, $T_f = 1.0$, noise level = 0.05}
+
 Here we consider a time-dependent advection-diffusion equation for

### 2023-05-26 paper/localpsf.tex
data from complete gaussian rbf runs

@@
 Krylov methods such as Lanczos or randomized methods~\cite{Cheng05,HalkoMartinssonTropp11} are typically used to construct these low rank approximations by applying the Hessian to vectors.
-Using these methods, the required number of Hessian applications (and hence the required number of PDE solves) is typically proportional to the rank of the low rank approximation.
+Using these methods, the required number of Hessian applications (and hence the required number of PDE solves) is proportional to the rank of the low rank approximation.
 Low rank approximation methods are justified by arguing that the numerical rank of the data misfit Hessian is insensitive to the dimension of the discretized parameter.
@@
 \begin{equation*}
-\mathbf{B}^T \mathbf{w} = \begin{bmatrix}
-\horzbar & \mathbf{b}_1^T & \horzbar \\
-%\horzbar & \mathbf{b}_2^T & \horzbar \\
+\mathbf{A}^T \mathbf{w} = \begin{bmatrix}
+\horzbar & \mathbf{a}_1^T & \horzbar \\
+%\horzbar & \mathbf{a}_2^T & \horzbar \\
 & \vdots & \\
-\horzbar & \mathbf{b}_\fedim^T & \horzbar
+\horzbar & \mathbf{a}_\fedim^T & \horzbar
 \end{bmatrix}
@@
 \begin{bmatrix}
-\mathbf{b}_1^T \mathbf{w} \\
-%\mathbf{b}_2^T \mathbf{w} \\
+\mathbf{a}_1^T \mathbf{w} \\
+%\mathbf{a}_2^T \mathbf{w} \\
 \vdots \\
-\mathbf{b}_\fedim^T \mathbf{w}
+\mathbf{a}_\fedim^T \mathbf{w}
 \end{bmatrix}.
 \end{equation*}
-By computing one matrix-vector product of $\mathbf{B}^T$ with $\mathbf{w}$, we compute the inner product of each column of $\mathbf{B}$ with $\mathbf{w}$ simultaneously.
+By computing one matrix-vector product of $\mathbf{A}^T$ with $\mathbf{w}$, we compute the inner product of each column of $\mathbf{A}$ with $\mathbf{w}$ simultaneously.
 The operator case is analogous, with $\phi_x$ taking the place of a matrix column.
@@
 
-\paragraph{Future extension: approximation using $\Sigma$} One may map $E_x$ to the unit sphere via the change of variables $z = \Sigma^{-1/2}\left(x\right)\left(y - \spatialmean(x)\right) / \tau$, and approximate
-\begin{equation}
-\label{eq:sigma_interpolation}
-\phi_x(z) \det{\Sigma^{1/2}\left(x\right)}/\spatialvol(x) \approx \phi_{x'}(z') \det{\Sigma^{1/2}\left(x'\right)}/\spatialvol(x').
-\end{equation}
-Notice that equality holds in \eqref{eq:sigma_interpolation} if the impulse responses are Gaussian functions.
-Although interpolation schemes based on \eqref{eq:sigma_interpolation} have the potential to be more accurate than those based on \eqref{eq:translate4}, we were unable to construct an interpolation scheme based on \eqref{eq:sigma_interpolation} that is robust to negative entries in the integral kernel.
+%\paragraph{Future extension: approximation using $\Sigma$} One may map $E_x$ to the unit sphere via the change of variables $z = \Sigma^{-1/2}\left(x\right)\left(y - \spatialmean(x)\right) / \tau$, and approximate
+%\begin{equation}
+%\label{eq:sigma_interpolation}
+%\phi_x(z) \det{\Sigma^{1/2}\left(x\right)}/\spatialvol(x) \approx \phi_{x'}(z') \det{\Sigma^{1/2}\left(x'\right)}/\spatialvol(x').
+%\end{equation}
+%%Notice that equality holds in \eqref{eq:sigma_interpolation} if the impulse responses are Gaussian functions. 
+%While interpolation schemes based on \eqref{eq:sigma_interpolation} have the potential to be more accurate than those based on \eqref{eq:translate4}, we were unable to construct an interpolation scheme based on \eqref{eq:sigma_interpolation} that is robust to negative entries in the integral kernel.
 The integral kernel should have non-negative entries per Condition 3 in Section \ref{sec:conditions_2}, but in practice some entries are negative due to discretization error.
-%Therefore, in this paper we do not use \eqref{eq:sigma_interpolation}. 
-%Constructing interpolation schemes based on \eqref{eq:sigma_interpolation} that are robust to discretization error is a promising direction for future research.
+%%Therefore, in this paper we do not use \eqref{eq:sigma_interpolation}. 
+%%Constructing interpolation schemes based on \eqref{eq:sigma_interpolation} that are robust to discretization error is a promising direction for future research.
 
@@
 When $\Aop$ is symmetric positive semi-definite, $\mathbf{A}_H$ may be non-symmetric and indefinite due to errors in the approximation.
-In this case, we symmetrize $\mathbf{A}_H$, find directions with large negative Rayleigh quotients using a Cayley-type shift-and-invert scheme, and flip these directions so that they have positive Rayleigh quotients via a low rank correction (see Appendix~\ref{sec:make_spd}).
+In this case, one may optionally symmetrize $\mathbf{A}_H$, then modify it via low rank updates to remove erroneous negative eigenvalues (Section \ref{sec:flip_eigs}).
 
@@
 
-		(optional) Modify $\mathbf{A}_H$ to make it symmetric positive definite (Appendix~\ref{sec:make_spd})
+		(optional) Modify $\mathbf{A}_H$ to make it symmetric and remove negative eigenvalues (Section \ref{sec:flip_eigs})
 
@@
 	\item We check whether $\candidatepoint_i$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ for all $\candidatepoint_j \in \pointbatch_k$.
-	\item We make sure that $\spatialvol(\candidatepoint_i)$ is not too small, by checking whether $\spatialvol(\candidatepoint_i) > \epsilon \spatialvol_\text{max}$.
+	\item We make sure that $\spatialvol(\candidatepoint_i)$ is not too small, by checking whether $\spatialvol(\candidatepoint_i) > \epsilon_V \spatialvol_\text{max}$.
 Here $\spatialvol_\text{max}$ is the largest value of $\spatialvol(\candidatepoint_j)$ over all points $q$ in the initial set of candidate points, and $\epsilon_V$ is a small threshold parameter (we use $\epsilon_V=10^{-5}$).
 	\item We make sure that all eigenvalues of $\spatialcov(\candidatepoint_i)$ are positive, and the aspect ratio of $E_{\candidatepoint_i}$ (ratio of the largest eigenvalue of $\spatialcov(\candidatepoint_i)$ to the smallest) is bounded by a constant $1/\epsilon_\Sigma$ (we use $1/\epsilon_\Sigma=20$).
-Negative integral kernel entries due to discretization error can cause $\spatialcov(\candidatepoint_i)$ to be indefinite or highly ill-conditioned, in which case the set $E_{\candidatepoint_i}$ poorly reflects the support of $\phi_{\candidatepoint_i}$.
+Negative integral kernel entries due to discretization error can cause $\spatialcov(\candidatepoint_i)$ to be indefinite or highly ill-conditioned.
+We reject these points because it is likely that the set $E_{\candidatepoint_i}$ poorly reflects the support of $\phi_{\candidatepoint_i}$.
 \end{enumerate}
@@
 %This is detailed in Algorithm~\ref{alg:point_choice}. 
-We check whether $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ using the ellipsoid intersection test described in Appendix~\ref{sec:fast_ellipsoid_intersection_test}.
+%We check whether $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ using the fast ellipsoid intersection test in \cite{GilitschenskiHanebeck12}.
+described in Appendix~\ref{sec:fast_ellipsoid_intersection_test}. 
 We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots, \pointbatch_{\nbatch}$.
@@
 \end{equation}
-where $\rbfweight_i$ are weights, and $\rbf$ is a polyharmonic spline radial basis function, which is defined as $\rbf(r):=r^\gdim$ if $\gdim$ is odd, and $\rbf(r):=r^\gdim \log r$ for $r>0$ and $\rbf(0):=0$ if $\gdim$ is even.
+where $\rbfweight_i$ are weights, and 
+$$\rbf(r):= \exp\left(-C_\text{RBF}^2\frac{r^2}{2 L^2}\right)$$
+is a Gaussian kernel radial basis function.
+Here $L:=\diam\left(\{x_i\}_{i=1}^{\numnbr}\right)$ is the diameter of the set of sample points used in the interpolation, and $C_\text{RBF}$ is a user-defined shape parameter that controls the width of the kernel function; we use $C_\text{RBF}=3.0$.
+%is a polyharmonic spline radial basis function, which is defined as $\rbf(r):=r^\gdim$ if $\gdim$ is odd, and $\rbf(r):=r^\gdim \log r$ for $r>0$ and $\rbf(0):=0$ if $\gdim$ is even.
 %\begin{equation}
@@
 and $f \in \mathbb{R}^{\numnbr}$ has entries $f_i$ from~\eqref{eq:fxyxp}. 
-Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a least squares sense~\cite{Duchon77}.
+%Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a least squares sense~\cite{Duchon77}. 
 %For more details on radial basis function interpolation, see~\cite{Wendland04}. 
@@
 We form $\massmatrix_H$ using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same geometrical clustering block cluster tree as was used for $\mathbf{\Aker}$.
+
+\subsection{Symmetrizing and flipping negative eigenvalues (optional)}
+\label{sec:flip_eigs}
+
+In many applications, one seeks to approximate an operator
+$\mathcal{H} = \Aop + \mathcal{R},$
+where $\Aop$ is a symmetric positive semi-definite operator that we approximate with our PSF method to form an H-matrix $\mathbf{A}_H$, and $\mathcal{R}$ is a symmetric positive definite operator that may be easily converted to an H-matrix $\mathbf{R}_H$ without using our PSF method.
+For example, in inverse problems $\mathcal{H}$ is the Hessian, $\Aop$ is the data misfit term in the Hessian which is dense and only available matrix-free, and $\mathcal{R}$ is the regularization term, which is typically an elliptic differential operator that becomes a sparse matrix after discretization.
+
+%In many problems $\Aop$ is symmetric positive semi-definite. 
+The PSF approximation $\mathbf{A}_H$, and therefore $\mathbf{A}_H + \mathbf{R}_H$, may be non-symmetric and indefinite because of approximation error. 
+This is undesirable because symmetry and positive semi-definiteness are important properties which should be preserved if possible.
+Also, lacking these properties may prevent one from using highly effective algorithms to perform further operations involving $\mathbf{A}_H + \mathbf{R}_H$, such as using $\mathbf{A}_H + \mathbf{R}_H$ as a preconditioner in the conjugate gradient method.
+
+We therefore modify $\mathbf{A}_H$ to make it symmetric and remove negative eigenvalues. First, we symmetrize $\mathbf{A}_H$ via
+$
+\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
+$
+Next, we find negative eigenvalues and their corresponding eigenvectors for the generalized eigenvalue problem
+$\mathbf{A}_H^{\text{sym}} \mathbf{u} = \lambda \mathbf{R}_H$
+using a Cayley shift-and-invert Krylov scheme \cite{LehoucqEtAl98}.
+We flip the signs of these eigenvalues to be positive instead of negative (i.e., $\lambda \rightarrow |\lambda|$) by performing a low rank update to $\mathbf{A}_H^{\text{sym}}$.
+The primary computational task in the Cayley shift-and-invert scheme is the solution of shifted linear systems of the form
+$\left(\mathbf{A}_H^{\text{sym}} + \mu_i \mathbf{R}_H\right) x = b,$
+for a small number of positive shifts $\mu_i$.
+We solve these linear systems by factorizing the matrices $\mathbf{A}_H^{\text{sym}} + \mu_i \mathbf{R}_H$ using fast H-matrix methods.
+We compute and flip all eigenvalues $\lambda < \tau$ which are less than some threshold $\tau \in (-1,0]$.
+By choosing $\tau > -1$, we ensure that the modified version of $\mathbf{A}_H^{\text{sym}} + \mathbf{R}_H$ is positive definite.
+Choosing $\tau=0$ would remove all erroneous negative eigenvalues.
+However, this is computationally infeasible if $\Aop$ has a large or infinite cluster of eigenvalues near zero, a common situation in practice.
+We therefore recommend choosing $\tau < 0$, and we use $\tau=-0.1$ in our numerical experiments.
 
@@
 %We perform low-rank updates to the H-matrix as the Newton iterations proceed, as per the discussion in Appendix~\ref{sec:recycle_krylov}.
-The rational method for modifying $\mathbf{A}_H$ to be positive semi-definite, described in Appendix~\ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the rational function order.
+Symmetrizing $\mathbf{A}_H$ requires one matrix-matrix addition, and flipping negative eigenvalues to be positive requires a handful (typically around 5) matrix-matrix additions and matrix factorizations.
+%semi-definite, described in Appendix~\ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the rational function order.
 
@@
 so that the Green's function of the Hessian of the regularization has
-a characteristic length of $0.25$ (NOTE: CHECK THIS FOR BOTH PROBLEMS) of the domain radius. For the
+a characteristic length of $0.25$ of the domain radius. For the
 specific setup, we refer the reader to~\cite[Section
@@
 
+%        |   0.5   |   1.0   |   2.0   
+% -------------------------------------
+% 1.0e-4 |   OK    |   OK    |   OK
+% 1.8e-4 |   OK    |   OK    |   OK
+% 3.2e-4 |   OK    |   OK    |   OK
+% 5.6e-4 |   OK    |   OK    |   5
+% 1.0e-3 |   OK    |   OK    |   5
+%
+
 Here we consider a time-dependent advection-diffusion equation for
@@
 
-\section{Rational positive semi-definite modification}
-\label{sec:make_spd}
-In many problems $\Aop$ is symmetric positive semi-definite. However, $\mathbf{A}_H$ is non-symmetric and indefinite because of approximation error. 
-This is undesirable.
+%\section{Rational positive semi-definite modification}
+%\label{sec:make_spd}
+%In many problems $\Aop$ is symmetric positive semi-definite. However, $\mathbf{A}_H$ is non-symmetric and indefinite because of approximation error. 
+%This is undesirable.
 Symmetry and positive semi-definiteness are important properties which should be preserved if possible.
 Also, lacking these properties may prevent one from using highly effective algorithms, such as the conjugate gradient method, to perform further operations involving $\mathbf{A}_H$.
-
-If $\Aop$ is symmetric, we symmetrize $\mathbf{A}_H$ as follows:
-$
-\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
-$
-If $\Aop$ is positive semi-definite, we adapt the rational method from~\cite[Section 13.8]{HackbuschKress07} to approximate the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$, which we denote by $|\mathbf{A}_H^{\text{sym}}|$.
+%
+%If $\Aop$ is symmetric, we symmetrize $\mathbf{A}_H$ as follows:
+%$
+%\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
+%$
+%If $\Aop$ is positive semi-definite, we adapt the rational method from~\cite[Section 13.8]{HackbuschKress07} to approximate the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$, which we denote by $|\mathbf{A}_H^{\text{sym}}|$.
 That is, $|\mathbf{A}_H^{\text{sym}}|$ has the same eigenvectors as $\mathbf{A}_H^{\text{sym}}$, but the eigenvalues of $|\mathbf{A}_H^{\text{sym}}|$ are the absolute values of the eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
-%flip (or approximately flip) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ to be positive. 
-Once constructed, we use the approximation of $|\mathbf{A}_H^{\text{sym}}|$ in place of $\mathbf{A}_H$ in further linear algebra operations.
-
-In detail, let $a$ be any scalar satisfying $-a < \lambda_\text{min}$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, and set 
-\begin{equation}
-\label{eq:ratT999}
-\mathbf{T} :=  \left(2 \mathbf{A}_H^{\text{sym}} + a \mathbf{I}\right)/a,
-\end{equation}
-From the results in~\cite[Section 13.8]{HackbuschKress07} it is straightforward to show that
-\begin{equation}
-\label{eq:rational343}
-\mathbf{\Pi}_\ratord := \left(\mathbf{I} + \mathbf{T}^{2^\ratord}\right)^{-1}
-\end{equation}
-approximates the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to negative eigenvalues. Thus, we have
-\begin{equation}
-\label{eq:rational76567}
-|\mathbf{A}_H^{\text{sym}}| \approx \left(\mathbf{I} - 2\mathbf{\Pi}_\ratord\right)\mathbf{A}_H^{\text{sym}}.
-\end{equation}
-To form an approximation of $|\mathbf{A}_H^{\text{sym}}|$, first, we compute $\lambda_\text{min}$ using the Lanczos method, and set $a=\gamma \lambda_\text{min}$, where $\gamma\ge 1$ is a parameter that must be chosen.
-The method is relatively insensitive to $\gamma$ and numerically we find that the method works well for any $\gamma \in [1,2]$. We use $\gamma=1.5$ in our numerical results. 
-Next, we compute $\mathbf{T}$ per~\eqref{eq:ratT999}, compute $\mathbf{T}^{2^\ratord}$ per~\eqref{eq:rational343} via repeated squaring, compute $\mathbf{\Pi}_\ratord$ per~\eqref{eq:rational343}, and form an approximation of $|\mathbf{A}_H^{\text{sym}}|$ per~\eqref{eq:rational76567}.
+%%flip (or approximately flip) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ to be positive. 
+%Once constructed, we use the approximation of $|\mathbf{A}_H^{\text{sym}}|$ in place of $\mathbf{A}_H$ in further linear algebra operations.
+%
+%In detail, let $a$ be any scalar satisfying $-a < \lambda_\text{min}$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, and set 
+%\begin{equation}
+%\label{eq:ratT999}
+%\mathbf{T} :=  \left(2 \mathbf{A}_H^{\text{sym}} + a \mathbf{I}\right)/a,
+%\end{equation}
+%From the results in~\cite[Section 13.8]{HackbuschKress07} it is straightforward to show that
+%\begin{equation}
+%\label{eq:rational343}
+%\mathbf{\Pi}_\ratord := \left(\mathbf{I} + \mathbf{T}^{2^\ratord}\right)^{-1}
+%\end{equation}
+%approximates the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to negative eigenvalues. Thus, we have
+%\begin{equation}
+%\label{eq:rational76567}
+%|\mathbf{A}_H^{\text{sym}}| \approx \left(\mathbf{I} - 2\mathbf{\Pi}_\ratord\right)\mathbf{A}_H^{\text{sym}}.
+%\end{equation}
+%To form an approximation of $|\mathbf{A}_H^{\text{sym}}|$, first, we compute $\lambda_\text{min}$ using the Lanczos method, and set $a=\gamma \lambda_\text{min}$, where $\gamma\ge 1$ is a parameter that must be chosen.
+%The method is relatively insensitive to $\gamma$ and numerically we find that the method works well for any $\gamma \in [1,2]$. We use $\gamma=1.5$ in our numerical results. 
+%Next, we compute $\mathbf{T}$ per~\eqref{eq:ratT999}, compute $\mathbf{T}^{2^\ratord}$ per~\eqref{eq:rational343} via repeated squaring, compute $\mathbf{\Pi}_\ratord$ per~\eqref{eq:rational343}, and form an approximation of $|\mathbf{A}_H^{\text{sym}}|$ per~\eqref{eq:rational76567}.
 Fast H-matrix methods are used to perform the required matrix-matrix additions and multiplications, and matrix inversion.

### 2023-05-26 localpsf/localpsf_cg1_lumped.py

@@ -463,6 +463,10 @@ class PSFObject:
         print('Building kernel hmatrix')
         kernel_hmatrix = me.psf_kernel.build_hmatrix(bct, tol=hmatrix_rtol)
 
+        z = kernel_hmatrix.matvec(np.random.randn(kernel_hmatrix.shape[1]))
+        print('np.any(np.isnan(z))=', np.any(np.isnan(z)))
+        print('np.any(np.isinf(z))=', np.any(np.isinf(z)))
+
         print('Computing hmatrix = diag(mass_lumps_out) @ kernel_hmatrix * diag(mass_lumps_in)')
         hmatrix = kernel_hmatrix.copy()
         hmatrix.mul_diag_left(me.V_out.mass_lumps)
@@ -536,6 +540,10 @@ def make_psf(
     modified_Sigma[bad_Sigmas,:,:] = np.eye(V_out.gdim).reshape((1, V_out.gdim, V_out.gdim))
 
     bad_inds = np.logical_or(bad_vols, bad_Sigmas)
+
+    print('np.any(np.isnan(modified_vol))=', np.any(np.isnan(modified_vol)))
+    print('np.any(np.isnan(modified_mu))=', np.any(np.isnan(modified_mu)))
+    print('np.any(np.isnan(modified_Sigma))=', np.any(np.isnan(modified_Sigma)))
 
     modified_IM = ImpulseMoments(modified_vol, modified_mu, modified_Sigma)
 

### 2023-05-26 numerical_examples/advection_diffusion/rerun_bad_adv.sh

@@ -3,25 +3,28 @@
 #ll, ii, jj, kk
 #nb, ns, kp, tf
 #failures:
-#[[0 0 3 2]
-# [0 0 4 0]
-# [1 0 3 2]
-# [1 0 4 2]
-# [2 0 0 1]]
+#[[0 0 0 0]
+# [0 0 1 1]
+# [1 0 0 2]
+# [1 0 3 0]
+# [2 0 4 1]]
 
-#python adv_deflation.py 0 0 1 0 'first'
-#python adv_deflation.py 0 0 1 2 'psf'
-#python adv_deflation.py 0 0 1 0 'last'
+#python adv_deflation.py 0 0 0 0 'first'
+#python adv_deflation.py 0 0 0 0 'psf'
+#python adv_deflation.py 0 0 0 0 'last'
 
-#python adv_deflation.py 0 4 0 0 'first'
-#python adv_deflation.py 0 4 0 0 'psf'
-#python adv_deflation.py 0 4 0 0 'last'
+#python adv_deflation.py 0 1 1 0 'first'
+#python adv_deflation.py 0 1 1 0 'psf'
+#python adv_deflation.py 0 1 1 0 'last'
 
-#python adv_deflation.py 0 3 2 0 'first'
-#python adv_deflation.py 0 3 2 0 'psf'
-#python adv_deflation.py 0 3 2 1 'psf'
-#python adv_deflation.py 0 3 2 0 'last'
+#python adv_deflation.py 0 0 2 0 'first'
+#python adv_deflation.py 0 0 2 1 'psf'
+#python adv_deflation.py 0 0 2 0 'last'
 
-python adv_deflation.py 0 4 2 0 'first'
-python adv_deflation.py 0 4 2 1 'psf'
-python adv_deflation.py 0 4 2 0 'last'
+#python adv_deflation.py 0 3 0 0 'first'
+#python adv_deflation.py 0 3 0 1 'psf'
+#python adv_deflation.py 0 3 0 0 'last'
+
+python adv_deflation.py 0 4 1 0 'first'
+python adv_deflation.py 0 4 1 2 'psf'
+python adv_deflation.py 0 4 1 0 'last'

### 2023-05-29 paper/localpsf.tex
edit intro and abstract of paper

@@
 \begin{abstract}
-	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels, with particular focus on approximating high rank Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs).
+	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels, with a focus on approximating high rank Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs).
 %	The method may be used to approximate other operators, including certain Schur complements, Poincare-Steklov operators, covariance operators, and blurring operators. 
-	The proposed method uses impulse responses of the underlying operator (computed at a collection of scattered points) and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel.
+	The method computes impulse responses of the operator computed at a collection of scattered points, and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel.
 The interpolation uses a ``local mean displacement invariance'' approximation, which improves upon classical local translation invariance.
-Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources.
-By solving an ellipsoid packing problem, we choose as many point sources as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
-Support ellipsoids are estimated a-priori based on impulse response moments, and these moments are computed via a procedure that involves applying the operator to a small number of polynomial functions.
+Impulse responses are computed by applying the operator to Dirac combs associated with ``batches'' of point sources.
+By solving an ellipsoid packing problem, we compute as many impulse responses as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
+Support ellipsoids are estimated a-priori based on impulse response moments, which are computed by applying the operator to a small number of polynomial functions.
 The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
-Fast H-matrix methods are used to perform further matrix computations in nearly linear complexity, including matrix-matrix addition and multiplication, matrix inversion and factorization, and matrix-vector products.
-We illustrate our approach by applying the method to approximate the Hessians in an ice sheet flow PDE constrained inverse problem with highly informative data, and an advection dominated advection-diffusion PDE constrained inverse problem.
-Numerical results demonstrate that the proposed method substantially outperforms existing methods.
-We show that our method can approximate high rank operators using only a small number of operator applies.
+Further matrix computations are performed with fast H-matrix methods.
+We apply the method to approximate the Hessians in an ice sheet flow PDE constrained inverse problem, and an advection-diffusion PDE constrained inverse problem.
+Numerical results demonstrate that the proposed method substantially outperforms existing methods, and can approximate high rank operators using only a small number of operator applies.
 \end{abstract}
@@
 Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, and $L^2(\Omega)'$ is the space of real-valued continuous linear functionals on $L^2(\Omega)$.
-We focus on approximating Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{AlgerEtAl19,IsaacEtAl15}, though such operators also arise as certain Schur complements in Schur complement methods for solving partial differential equations (PDEs)~\cite{le1997non,saad1999distributed}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics~\cite{ChenStein21,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}.
+We focus on approximating Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{AlgerEtAl19,IsaacEtAl15}.
+Such operators also arise as certain Schur complements in Schur complement methods for solving partial differential equations (PDEs)~\cite{le1997non,saad1999distributed}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics~\cite{ChenStein21,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}.
 Here ``matrix-free'' means that we may apply $\Aop$ and its transpose\footnote{Recall that $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for operations with $\Aop^T$.
@@
 %Existing scalable high rank approximation methods typically require the ability to rapidly evaluate arbitrary entries of $\Aop$'s integral kernel, which we cannot do. 
-We use \emph{impulse response interpolation} to form a high rank operator approximation using a small number of operator applies.
+We use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applies.
 The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
@@
 Picking the batches of points requires us to estimate the supports of the impulse responses $\phi_x$ \emph{before} we compute them. 
-The idea of estimating the supports of the functions $\phi_x$ a-priori was inspired by techniques from resolution analysis in seismic imaging. There, 
+In resolution analysis in seismic imaging, 
 $\Aop^T$ is applied to a random noise function, and the width of $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$~\cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
-We use polynomial functions instead of random noise functions (see Section~\ref{sec:intromoments}).
+Here, we estimate the support of the $\phi_x$ by applying $\Aop^T$ to polynomial functions instead of random noise functions (see Section~\ref{sec:intromoments}).
 Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
+
 Our method may be categorized as a PSF method that is loosely based on ``product convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
@@
 For background on PC and PSF methods, we recommend the following papers: \cite{DenisEtAl15,EscandeWeiss17,GentileCourbinMeylan13}.
-While PC approximations are based on an assumption of local translation invariance, the method we propose is based on an assumption we call ``local mean displacement invariance,'' which is more general than local translation invariance, and includes both CUR low rank approximation and local translation invariance as special cases (Section~\ref{sec:local_mean_displacement_invariance} and Figure~\ref{fig:mean_displacement_invariance}).
+While PC approximations are based on an assumption of local translation invariance, the method we propose is based on a more general assumption we call ``local mean displacement invariance'' (Section~\ref{sec:local_mean_displacement_invariance} and Figure~\ref{fig:mean_displacement_invariance}).
 %Furthermore, PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulses.
@@
 
-The ability to rapidly approximate entries of $\Aop$'s integral kernel allows us to form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of a discretized version of $\Aop$.
+The ability to rapidly approximate entries of $\Aop$'s integral kernel allows one to construct compressed representations of the discretized version of $\Aop$ using methods that require cheap access to arbitrary matrix entries.
+We form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of $\Aop$.
 H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank.
@@
 The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks~\cite{GrasedyckHackbusch03}.
-The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank Gaussian surrogate models that can be used for uncertainty quantification.
+The ability to perform these operations permits, for example, solution of Newton linear systems in PDE constrained optimization, sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank Gaussian surrogate models that can be used for uncertainty quantification.
 %Classical H-matrix construction methods 

### 2023-05-30 paper/localpsf.tex
edit intro and abstract of paper
edit abstract

@@
 \begin{abstract}
-	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels, with a focus on approximating high rank Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs).
+	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels.
+%	, with a focus on approximating high rank Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs). 
 %	The method may be used to approximate other operators, including certain Schur complements, Poincare-Steklov operators, covariance operators, and blurring operators. 
-	The method computes impulse responses of the operator computed at a collection of scattered points, and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel.
-The interpolation uses a ``local mean displacement invariance'' approximation, which improves upon classical local translation invariance.
-Impulse responses are computed by applying the operator to Dirac combs associated with ``batches'' of point sources.
+	The method computes impulse responses of the operator at a collection of scattered points, and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel.
+We compute a large number of impulse responses by applying the operator to a small number Dirac combs associated with ``batches'' of point sources.
 By solving an ellipsoid packing problem, we compute as many impulse responses as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
 Support ellipsoids are estimated a-priori based on impulse response moments, which are computed by applying the operator to a small number of polynomial functions.
+The interpolation uses a ``local mean displacement invariance'' approximation, which improves upon classical local translation invariance.
 The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
@@
 via a black box computational procedure, 
-but we cannot easily access entries of $\Aop$'s integral kernel.
-Evaluating the maps in~\eqref{eq:eval_maps} may require, as a sub procedure, performing a costly computation, such as solving a large linear system or time stepping.
+but we cannot easily access entries of $\Aop$'s integral kernel. Evaluating the maps in~\eqref{eq:eval_maps} may require, as a sub procedure, solving PDEs or performing other costly computations.
 
@@
 Here, we estimate the support of the $\phi_x$ by applying $\Aop^T$ to polynomial functions instead of random noise functions (see Section~\ref{sec:intromoments}).
-Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
+Our method estimates the support of $\phi_x$ more accurately than random noise probing, at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
 
@@
 
-The ability to rapidly approximate entries of $\Aop$'s integral kernel allows one to construct compressed representations of the discretized version of $\Aop$ using methods that require cheap access to arbitrary matrix entries.
-We form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of $\Aop$.
+The ability to rapidly approximate entries of $\Aop$'s integral kernel allows 
+%one to approximate discretized matrix representations of $\Aop$ using methods that require fast access to arbitrary matrix entries. We 
+us to form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of $\Aop$.
 H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank.

### 2023-05-30 paper/localpsf.tex
edit intro

@@
 The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
-Further matrix computations are performed with fast H-matrix methods.
+Further matrix computations are performed with H-matrix methods.
 We apply the method to approximate the Hessians in an ice sheet flow PDE constrained inverse problem, and an advection-diffusion PDE constrained inverse problem.
-Numerical results demonstrate that the proposed method substantially outperforms existing methods, and can approximate high rank operators using only a small number of operator applies.
+Numerical results demonstrate that the proposed method outperforms existing methods, and can approximate high rank operators using only a small number of operator applies.
 \end{abstract}
@@
 
-The ability to rapidly approximate entries of $\Aop$'s integral kernel allows 
+The ability to rapidly approximate entries of $\Aop$'s integral kernel allows one to approximate discretized versions of $\Aop$ using the full arsenal of tools for matrix approximation that rely on fast access to matrix entries.
 %one to approximate discretized matrix representations of $\Aop$ using methods that require fast access to arbitrary matrix entries. We 
-us to form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of $\Aop$.
+We form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of $\Aop$.
 H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank.

### 2023-05-30 paper/localpsf.tex
edit intro

@@
 The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks~\cite{GrasedyckHackbusch03}.
-The ability to perform these operations permits, for example, solution of Newton linear systems in PDE constrained optimization, sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank Gaussian surrogate models that can be used for uncertainty quantification.
+These operations can be used, for example, to solve Newton linear systems in PDE constrained optimization, sample from ill-conditioned posterior distributions in Bayesian inverse problems, and construct high rank Gaussian surrogate models that can be used for uncertainty quantification.
 %Classical H-matrix construction methods 

### 2023-05-30 paper/localpsf.tex
add advection diffusion convergence iterations table to paper

@@
 	\begin{center}
+		\begingroup
+		\setlength{\tabcolsep}{8pt}
+		\renewcommand{\arraystretch}{1.25}
+		\begin{tabular}{c| c | c c c c c}
+			& $\kappa$    & REG     &	NONE   & PSF $(1)$ & PSF $(5)$ & PSF $(25)$ \\ 
+			\hline 
+			& $1.0e-4$    & 692 & 371  & 350   & 171   & 65    \\ 
+			& $1.8e-4$    & 744 & 367  & 220   & 150   & 55    \\   
+			$T=0.5$ & $3.2e-4$    & 805 & 364  & 271   & 164   & 51    \\ 
+			& $5.6e-4$    & 867 & 365  & 400   & 171   & 45    \\  
+			& $1.0e-3$    & 820 & 383  & 139   & 82   & 37    \\ 
+			\hline
+			& $1.0e-4$    & 744 & 526  & 620   & 338   & 108    \\ 
+			& $1.8e-4$    & 769 & 528  & 554   & 248   & 98    \\   
+			$T=1.0$ & $3.2e-4$    & 778 & 542  & 401   & 233   & 101    \\ 
+			& $5.6e-4$    & 725 & 570  & 296   & 227   & 80    \\  
+			& $1.0e-3$    & 666 & 624  & 299   & 291   & 251    \\ 
+			\hline
+			& $1.0e-4$    & 718 & 707  & 616   & 596   & 194    \\ 
+			& $1.8e-4$    & 670 & 731  & 505   & 478   & 207    \\   
+			$T=2.0$ & $3.2e-4$    & 606 & 778  & 382   & 445   & 189    \\ 
+			& $5.6e-4$    & 528 & 856  & 558   & 326   & 173    \\  
+			& $1.0e-3$    & 398 & 945  & 425   & 294   & 182    \\ 
+		\end{tabular}
+		\endgroup
+	\end{center}
+	\caption{Number of preconditioned conjugate gradient iterations to achieve tolerance $10^{-7}$ for solving the Newton linear system in the advection-diffusion inverse problem.
+Iteration counts are shown for a variety of different diffusion parameters $\kappa$, simulation times $T$, and preconditioners.}
+	\label{tab:adv_convergence}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
 		\includegraphics[scale=0.75]{adv_vary_kappa_tf_tmp.png}

### 2023-05-31 paper/localpsf.tex
update adv table to use 1e-6 tolerance

--- 2023-05-30 paper/localpsf.tex
@@
 	\begin{center}
+		\begingroup
+		\setlength{\tabcolsep}{8pt}
+		\renewcommand{\arraystretch}{1.25}
+		\begin{tabular}{c| c | c c c c c}
+			& $\kappa$    & REG     &	NONE   & PSF $(1)$ & PSF $(5)$ & PSF $(25)$ \\ 
+			\hline 
+			& $1.0e-4$    & 584 & 317  & 311   & 151   & 56    \\ 
+			& $1.8e-4$    & 628 & 313  & 192   & 131   & 49    \\   
+	$T=0.5$ & $3.2e-4$    & 685 & 311  & 233   & 140   & 44    \\ 
+			& $5.6e-4$    & 734 & 310  & 334   & 144   & 40    \\  
+			& $1.0e-3$    & 702 & 324  & 122   & 71   & 33    \\ 
+			\hline
+			& $1.0e-4$    & 634 & 449  & 539   & 288   & 100    \\ 
+			& $1.8e-4$    & 665 & 449  & 493   & 212   & 84    \\   
+	$T=1.0$ & $3.2e-4$    & 681 & 459  & 350   & 202   & 90    \\ 
+			& $5.6e-4$    & 619 & 482  & 262   & 197   & 70    \\  
+			& $1.0e-3$    & 574 & 520  & 266   & 260   & 208    \\ 
+			\hline
+			& $1.0e-4$    & 609 & 591  & 548   & 520   & 165    \\ 
+			& $1.8e-4$    & 576 & 611  & 434   & 426   & 185    \\   
+	$T=2.0$ & $3.2e-4$    & 524 & 645  & 318   & 379   & 170    \\ 
+			& $5.6e-4$    & 464 & 711  & 480   & 289   & 153    \\  
+			& $1.0e-3$    & 349 & 786  & 381   & 262   & 158    \\ 
+		\end{tabular}
+		\endgroup
+	\end{center}
+	\caption{Number of preconditioned conjugate gradient iterations to achieve tolerance $10^{-6}$ for solving the Newton linear system in the advection-diffusion inverse problem.
+Iteration counts are shown for a variety of different diffusion parameters $\kappa$, simulation times $T$, and preconditioners.}
+	\label{tab:adv_convergence}
+\end{figure}
+
+\begin{figure}
+	\begin{center}
 		\includegraphics[scale=0.75]{adv_vary_kappa_tf_tmp.png}

### 2023-06-01 paper/localpsf.tex
add figures created by Tucker to paper

--- 2023-05-30 paper/localpsf.tex
@@
 
-\begin{figure}
+\begin{table}
 	\begin{center}
@@
 	\label{fig:condition_number}
-\end{figure}
+\end{table}
 
@@
 	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.25]{adv_ic_true_tmp.png}
+		\includegraphics[scale=0.14]{true_initial_condition.png}
 		%		\caption{}
@@
 	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.25]{velocity_tmp.png}
+		\includegraphics[scale=0.14]{wind_velocity_colorscheme3.png}
 		%		\caption{}
@@
 	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.25]{adv_final_condition_true_tmp.png}
+		\includegraphics[scale=0.14]{noisy_observations_T1.0_K3.2e-4.png}
 		%		\caption{}
@@
 	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.25]{adv_reconstruction_tmp.png}
+		\includegraphics[scale=0.14]{reconstructed_initial_condition_T1.0_K3.2e-4.png}
 		%		\caption{}
@@
 	%\end{subfigure}
-	\caption{(a) true initial condition. (b) velocity field. (c) observed concentration at final time. (d) reconstruction of initial condition. black=0, white=1}
+	\caption{(Left) True initial condition.
+(Middle left) Velocity field.
+Color indicates magnitude of velocity vector.
+(Middle right) Noisy observations of concentration at final time.
+(Right) Reconstructed initial condition.}
 	\label{fig:adv_inversion1} 
@@
 
+\begin{table}
+	\begin{center}
+		\begingroup
+		\setlength{\tabcolsep}{8pt}
+%		\renewcommand{\arraystretch}{1.25}
+		\renewcommand{\arraystretch}{1.0}
+		\begin{tabular}{c| c | c c c c c}
+			& $\kappa$    & REG     &	NONE   & PSF $(1)$ & PSF $(5)$ & PSF $(25)$ \\ 
+			\hline 
+			& $1.0e{-4}$    & 584 & 317  & 311   & 151   & 56    \\ 
+%			& $1.8e{-4}$    & 628 & 313  & 192   & 131   & 49    \\   
+	$T=0.5$ & $3.2e{-4}$    & 685 & 311  & 233   & 140   & 44    \\ 
+%			& $5.6e{-4}$    & 734 & 310  & 334   & 144   & 40    \\  
+			& $1.0e{-3}$    & 702 & 324  & 122   & 71   & 33    \\ 
+			\hline
+			& $1.0e{-4}$    & 634 & 449  & 539   & 288   & 100    \\ 
+%			& $1.8e{-4}$    & 665 & 449  & 493   & 212   & 84    \\   
+	$T=1.0$ & $3.2e{-4}$    & 681 & 459  & 350   & 202   & 90    \\ 
+%			& $5.6e{-4}$    & 619 & 482  & 262   & 197   & 70    \\  
+			& $1.0e{-3}$    & 574 & 520  & 266   & 260   & 208    \\ 
+			\hline
+			& $1.0e{-4}$    & 609 & 591  & 548   & 520   & 165    \\ 
+%			& $1.8e{-4}$    & 576 & 611  & 434   & 426   & 185    \\   
+	$T=2.0$ & $3.2e{-4}$    & 524 & 645  & 318   & 379   & 170    \\ 
+%			& $5.6e{-4}$    & 464 & 711  & 480   & 289   & 153    \\  
+			& $1.0e{-3}$    & 349 & 786  & 381   & 262   & 158    \\ 
+		\end{tabular}
+		\endgroup
+	\end{center}
+	\caption{Number of preconditioned conjugate gradient iterations required to achieve relative error $10^{-6}$ for solving the Newton linear system in the advection-diffusion inverse problem.
+Iteration counts are shown for a variety of different diffusion parameters $\kappa$, simulation times $T$, and preconditioners.}
+	\label{tab:adv_convergence}
+\end{table}
+
+\begin{table}
+	\centering
+	\begin{tabular}{cccc}
+		  & $T=0.5$ & $T=1.0$ & $T=2.0$ \\
+		$\kappa=1.0e{-4}$  & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-4.png} \\
+		$\kappa=3.2e{-4}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K3.2e-4.png} \\
+		$\kappa=1.0e{-3}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-3.png}
+	\end{tabular}
+	\caption{Impulse responses for a variety of different diffusion parameters $\kappa$ and simulation times $T$.}
+	\label{tbl:table_of_figures}
+\end{table}
+
 \begin{figure}
-	\begin{center}
-		\includegraphics[scale=0.75]{adv_vary_kappa_tf_tmp.png}
-	\end{center}
-	\caption{Impulse responses for various diffusivities kappa and fixed simulation time Tf=0.5 (top) and simulation times with fixed kappa=1e-3 (bot).
-Colorbar is normalized independently for each subfigure}
+	\caption{Impulse responses for various diffusivities kappa and fixed simulation time Tf=0.5 (top) and simulation times with fixed kappa=1e-3 (bot).}
 	\label{fig:adv_vary_kappa_tf}

### 2023-06-01 paper/localpsf.tex
adding plot of generalized eigenvalues and cg convergence for advection diffusion problem

@@
 
-{\Huge Key parameter choices: $\kappa=3.2e-3$, $T_f = 1.0$, noise level = 0.05}
-
 %        |   0.5   |   1.0   |   2.0   
@@
 			\hline 
-			& $1.0e{-4}$    & 584 & 317  & 311   & 151   & 56    \\ 
+			& 1.0e-4    & 584 & 317  & 311   & 151   & 56    \\ 
 %			& $1.8e{-4}$    & 628 & 313  & 192   & 131   & 49    \\   
-	$T=0.5$ & $3.2e{-4}$    & 685 & 311  & 233   & 140   & 44    \\ 
+	$T=0.5$ & 3.2e-4    & 685 & 311  & 233   & 140   & 44    \\ 
 %			& $5.6e{-4}$    & 734 & 310  & 334   & 144   & 40    \\  
-			& $1.0e{-3}$    & 702 & 324  & 122   & 71   & 33    \\ 
+			& 1.0e-3    & 702 & 324  & 122   & 71   & 33    \\ 
 			\hline
-			& $1.0e{-4}$    & 634 & 449  & 539   & 288   & 100    \\ 
+			& 1.0e-4    & 634 & 449  & 539   & 288   & 100    \\ 
 %			& $1.8e{-4}$    & 665 & 449  & 493   & 212   & 84    \\   
-	$T=1.0$ & $3.2e{-4}$    & 681 & 459  & 350   & 202   & 90    \\ 
+	$T=1.0$ & 3.2e-4    & 681 & 459  & 350   & 202   & 90    \\ 
 %			& $5.6e{-4}$    & 619 & 482  & 262   & 197   & 70    \\  
-			& $1.0e{-3}$    & 574 & 520  & 266   & 260   & 208    \\ 
+			& 1.0e-3    & 574 & 520  & 266   & 260   & 208    \\ 
 			\hline
-			& $1.0e{-4}$    & 609 & 591  & 548   & 520   & 165    \\ 
-%			& $1.8e{-4}$    & 576 & 611  & 434   & 426   & 185    \\   
-	$T=2.0$ & $3.2e{-4}$    & 524 & 645  & 318   & 379   & 170    \\ 
+			& 1.0e-4    & 609 & 591  & 548   & 520   & 165    \\ 
+%			& 1.8e{-4}$    & 576 & 611  & 434   & 426   & 185    \\   
+	$T=2.0$ & 3.2e-4    & 524 & 645  & 318   & 379   & 170    \\ 
 %			& $5.6e{-4}$    & 464 & 711  & 480   & 289   & 153    \\  
-			& $1.0e{-3}$    & 349 & 786  & 381   & 262   & 158    \\ 
+			& 1.0e-3    & 349 & 786  & 381   & 262   & 158    \\ 
 		\end{tabular}
@@
 		ymax=2,
-		xmax=750,
+		xmax=1000,
 		axis x line*=bottom,
 		axis y line*=left,
-		xtick={0, 200, 400, 600},
+		xtick={0, 200, 400, 600,800,1000},
 		xlabel={CG iteration},
-		ylabel={$\|\bm{b}-\bm{H} \bm{x}_k\|/\|\bm{b}\|$},
+		ylabel={$\|\bm{x}_k - \bm{x}\|/\|\bm{x}\|$},
 		title={CG convergence},
-		legend style={at={(1.1,1)},anchor=north east},
+		legend style={at={(1.2,1)},anchor=north east},
 		%legend pos=north east,
@@
 		\addlegendentry{Reg}
-		\addplot[dotted, line width=0.5mm] table {cg_relres_reg.dat};
+		\addplot[dotted, line width=0.5mm] table {advDiff_cg_rand_errs_reg.dat};
 		\addlegendentry{None}
-		\addplot[dash dot, line width=0.5mm] table {cg_relres_none.dat};
+		\addplot[dash dot, line width=0.5mm] table {advDiff_cg_rand_errs_none.dat};
 		\addlegendentry{PSF (1)}
-		\addplot[blue, line width=0.5mm] table {cg_relres_p1.dat};
+		\addplot[blue, line width=0.5mm] table {advDiff_cg_rand_errs_psf1.dat};
 		\addlegendentry{PSF (5)}
-		\addplot[red, line width=0.5mm] table {cg_relres_p5.dat};
+		\addplot[red, line width=0.5mm] table {advDiff_cg_rand_errs_psf5.dat};
 		\addlegendentry{PSF (25)}
-		\addplot[purple, line width=0.5mm] table {cg_relres_p25.dat};
+		\addplot[purple, line width=0.5mm] table {advDiff_cg_rand_errs_psf25.dat};
 		\end{axis}
@@
 \begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
-		xmode=log,
-		ymode=log,
-		xlabel={$\gamma$, regularization parameter},
-		ylabel={\# of CG iterations},
-		axis y line*=right,
-		axis x line*=bottom,
-		ylabel near ticks, yticklabel pos=right,
-		%ylabel near ticks, ylabel shift={-6pt},
-		%legend pos=north east,
-		title={Regularization robustness}
-		%title={CG iterations to achieve $\|H x-b\|/\|b\|<10^{-6}$}
-		]
-		%\addlegendentry{None}
-		\addplot[dash dot, line width=0.5mm] table[x=gamma, y=none] {CGitsvsGamma.dat};
-		%\addlegendentry{REG}
-		\addplot[dotted, line width=0.5mm] table[x=gamma, y=reg] {CGitsvsGamma.dat};
-		%\addlegendentry{PSF (1)}
-		\addplot[blue, line width=0.5mm] table[x=gamma, y=psf1] {CGitsvsGamma.dat};
-		%\addlegendentry{PSF (5)}
-		\addplot[red, line width=0.5mm] table[x=gamma, y=psf5] {CGitsvsGamma.dat};
-		%\addlegendentry{PSF (25)}
-		\addplot[purple, line width=0.5mm] table[x=gamma, y=psf25] {CGitsvsGamma.dat};
-		\end{axis}
-		\end{tikzpicture}
-		\caption{}
-		\label{fig:adv_kappa_sweep}
-	\end{subfigure}
-	\caption{Left (\ref{fig:adv_cg_convergence}) shows the convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
-random entries drawn from the standard normal distribution.
-BLAH BLAH BLAH
-		%		The y-axis is $\|b-Hx_k\|/\|b\|$, where $x_k$ is the $k$th PCG iterate. 
-		Here we use $\gamma=9999 \times 10^9$, which is determined by the Morozov discrepancy principle.
-Right (\ref{fig:adv_kappa_sweep}) shows the number of PCG iterations required to achieve $\|\bm{b}-\bm{H} \bm{x}_k\|/\|\bm{b}\| < 10^{-6}$, where $\bm{x}_k$ is the $k$th PCG iterate, for a range of different $\kappa$.
-Results in these figures are shown for our preconditioner with 5, 15, and 30 batches (PSF (5), PSF (15), and PSF(30), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
-The preconditioner is constructed using $\bm{H}_\text{gn}$.}
-	\label{fig:adv_krylov_convergence}
-\end{figure}
-
-\begin{figure}
-	\begin{subfigure}{0.59\textwidth}
-		\begin{center}
-			\begin{tikzpicture}[baseline, trim axis right]
-			\begin{axis}[width=0.99\linewidth, height=0.8\linewidth,
 ymode=log,
 xmin=-50,
-			xmax=1450,
-			ymax=1.e4,
+xmax=600,
+ymax=1.e6,
+title={Generalized eigenvalues},
 xlabel={$k$, generalized eigenvalue \#},
@@
 axis x line*=bottom,
-			axis y line*=left,
-			legend style={at={(0.95,1)},anchor=north east},
-			title={Generalized eigenvalues, $\bm{H} \genericvec_{k}=\lambda_{k} \bm{\preconditioner} \genericvec_{k}$}
+axis y line*=right,
+%legend style={at={(0.95,1)},anchor=north east},
 ]
-			\addlegendentry{REG}
-			\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues.dat};
-			\addlegendentry{PSF (1)}
-			\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues.dat};
-			\addlegendentry{PSF (5)}
-			\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues.dat};
-			\addlegendentry{PSF (25)}
-			\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues.dat};
+%\addlegendentry{REG}
+\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {advDiff_generalizedEigenvalues.dat};
+\addplot[dash dot, line width=0.5mm] table[x=k, y=none] {advDiff_generalizedEigenvalues.dat};
+%\addlegendentry{PSF (1)}
+\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {advDiff_generalizedEigenvalues.dat};
+%\addlegendentry{PSF (5)}
+\addplot[red, line width=0.5mm] table[x=k, y=psf5] {advDiff_generalizedEigenvalues.dat};
+%\addlegendentry{PSF (25)}
+\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {advDiff_generalizedEigenvalues.dat};
 \end{axis}
 \end{tikzpicture}
-		\end{center}
 		\caption{}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.39\textwidth}
-		\begin{center}
-			\begingroup
-			\setlength{\tabcolsep}{4pt}
-			\renewcommand{\arraystretch}{1.25}
-			\begin{tabular}{c| c c}
-				& $\kappa(\bm{\preconditioner}^{-1} \bm{H})$\\
-				\hline 
-				REG        & 7.02e+3 \\
-				NONE       & 8.71e+2 \\
-				PSF $(5)$  & 1.75e+1 \\
-				PSF $(15)$  & 1.39e+1 \\
-				PSF $(30)$ & 1.03e+1
-			\end{tabular}
-			\endgroup
-		\end{center}
-		\caption{}
-		\label{fig:adv_condition_number}
-	\end{subfigure}
-	\caption{Left (\ref{fig:adv_generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_k = \lambda_k \bm{\preconditioner} \bm{u}_k$.
-Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 5, 15, and 30 batches (PSF (5), PSF (15), and PSF (30), respectively), or the regularization Hessian (REG).
-Right (\ref{fig:adv_condition_number}): condition number for $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$ for these different preconditioners, and no preconditioner (NONE).
-COND(H) LABELS!}
-	\label{fig:adv_spectrum}
+	\caption{Left (\ref{fig:adv_cg_convergence}) shows the convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
+random entries drawn from the standard normal distribution.
+BLAH BLAH BLAH
+		%		The y-axis is $\|b-Hx_k\|/\|b\|$, where $x_k$ is the $k$th PCG iterate. 
+		Here we use $\gamma=7.88 \times 10^{-6}$, which is determined by the Morozov discrepancy principle.
+Results in these figures are shown for our preconditioner with 5, 15, and 30 batches (PSF (5), PSF (15), and PSF(30), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
+The preconditioner is constructed using $\bm{H}_\text{gn}$.}
+	\label{fig:adv_krylov_convergence}
 \end{figure}

### 2023-06-01 paper/adv_pcg_convergence.tex
move adv pcg and geigs plots to separate files

--- 2023-06-01 paper/localpsf.tex
@@
-% SIAM Article Template
-\documentclass[review,onefignum,onetabnum]{siamart190516}
+\documentclass{standalone}
+\usepackage{tikz}
+\usepackage{pgfplots}
+\usepackage{amsmath}
 
-% Information that is shared between the article and the supplement
-% (title and author information, macros, packages, etc.) goes into
-% ex_shared.tex. If there is no supplement, this file can be included
-% directly.
+% SIAM Shared Information Template
+% This is information that is shared between the main document and any
+% supplement. If no supplement is required, then this information can
+% be included directly in the main document.
 
-\input{localpsf_shared}
-
-% Optional PDF information
-\ifpdf
-\hypersetup{
-  pdftitle={Point spread function approximation of high rank Hessians with locally supported non-negative integral kernels},
-  pdfauthor={N. Alger, N. Petra, T. Hartland, and O. Ghattas}
-}
-\fi
-
-% The next statement enables references to information in the
-% supplement. See the xr-hyperref package for details.
-
-\externaldocument{localpsf_supplement}
-
-% FundRef data to be entered by SIAM
-%<funding-group specific-use="FundRef">
-%<award-group>
-%<funding-source>
-%<named-content content-type="funder-name"> 
-%</named-content> 
-%<named-content content-type="funder-identifier"> 
-%</named-content>
-%</funding-source>
-%<award-id> </award-id>
-%</award-group>
-%</funding-group>
-
-% CITATIONS TO DO
-%
-% TASKS:
-% 1) Incorporate Ronald Kriemann's suggested citations: (Nick)
-%    - Steffen Börm, Lars Grasedyck and Wolfgang Hackbusch:
-%      "Introduction to Hierarchical Matrices with Applications",
-%       Engineering Analysis with Boundary Elements 27(5), pp. 405--422, 2003. (DONE)
-%    - Lars Grasedyck, Ronald Kriemann and Sabine Le Borne:
-%      "Parallel blackbox H-LU preconditioning for elliptic boundary value problems",
-%       Computing and Visualization in Sciences, Vol. 11, pp 273--291, 2007. (DONE)
-%    - Ronald Kriemann:
-%      "H-LU Factorization on Many-Core Systems",
-%      Computing and Visualization in Science, 16, pp. 105-117, 2015. (DONE)
-%
-% 2) Grants
-%    - Nick and Omar's grants (Omar)
-%    - Tucker and Noemi's grants (DONE)
-%
-% 3) Order citations in ascending order within brackets (Nick)
-%    e.g., [1,2,3] = good
-%          [2,3,1] = bad
-%
-% 4) Find and cite relevant papers from Noemis's group
-%    - Tucker's technical report (Tucker)
-%
-% 5) Find and cite relevant papers from Omar's group
-%    - Toby's papers (Tucker)
-%    - Noemi's 2012 stokes paper (Tucker)
-%    - Omar and Karen's recent review paper (DONE)
-%
-% 6) Look through Nick's previous PC paper for missed citations (DONE)
-
-% For citations we need:
-%    a) bibtex info
-%    b) exactly where in the paper to add the ref
-%    c) any words or sentences to add with the ref
-
-\newcommand{\todo}[1]{\noindent\emph{\textcolor{red}{To do: #1\:}}}
+% Packages and macros go here
+\usepackage{lipsum}
+\usepackage{amsfonts}
+\usepackage{graphicx}
+\usepackage{epstopdf}
+\usepackage{amsmath}
+\usepackage{subcaption}
+\usepackage{pgfplots}
+\usepackage{bm}
 
 \begin{document}
-
-\maketitle
-
-% REQUIRED
-\begin{abstract}
-	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels.
-%	, with a focus on approximating high rank Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs). 
-%	The method may be used to approximate other operators, including certain Schur complements, Poincare-Steklov operators, covariance operators, and blurring operators. 
-	The method computes impulse responses of the operator at a collection of scattered points, and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel.
-We compute a large number of impulse responses by applying the operator to a small number Dirac combs associated with ``batches'' of point sources.
-By solving an ellipsoid packing problem, we compute as many impulse responses as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
-Support ellipsoids are estimated a-priori based on impulse response moments, which are computed by applying the operator to a small number of polynomial functions.
-The interpolation uses a ``local mean displacement invariance'' approximation, which improves upon classical local translation invariance.
-The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
-Further matrix computations are performed with H-matrix methods.
-We apply the method to approximate the Hessians in an ice sheet flow PDE constrained inverse problem, and an advection-diffusion PDE constrained inverse problem.
-Numerical results demonstrate that the proposed method outperforms existing methods, and can approximate high rank operators using only a small number of operator applies.
-\end{abstract}
-
-% REQUIRED
-\begin{keywords}
- 	data scalability, Hessian, hierarchical matrix, matrix-free, operator approximation, PDE constrained inverse problems, point spread function, product convolution, local translation invariance, impulse response, moment methods
-\end{keywords}
-
-% REQUIRED
-\begin{AMS}
- 	35R30, 41A05, 41A35, 41A36, 42A85, 47A52, 47A58, 49K20, 65D12, 65F08, 65F10, 65N20, 65N21, 86A22, 86A40
-\end{AMS}
-
-\section{Introduction}
-\label{sec:intro}
-
-We present an efficient \emph{matrix-free} point spread function (PSF) method for approximating operators, $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$, that have locally supported non-negative integral kernels.
-Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, and $L^2(\Omega)'$ is the space of real-valued continuous linear functionals on $L^2(\Omega)$.
-We focus on approximating Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{AlgerEtAl19,IsaacEtAl15}.
-Such operators also arise as certain Schur complements in Schur complement methods for solving partial differential equations (PDEs)~\cite{le1997non,saad1999distributed}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics~\cite{ChenStein21,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}.
-Here ``matrix-free'' means that we may apply $\Aop$ and its transpose\footnote{Recall that $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for operations with $\Aop^T$.
-}, $\Aop^T$, to arbitrary functions,
-\begin{equation}
-\label{eq:eval_maps}
-u \mapsto\Aop u \qquad \text{and} \qquad w \mapsto \Aop^T w,
-\end{equation} 
-via a black box computational procedure, 
-but we cannot easily access entries of $\Aop$'s integral kernel. Evaluating the maps in~\eqref{eq:eval_maps} may require, as a sub procedure, solving PDEs or performing other costly computations.
-
-\begin{figure}
-	\begin{subfigure}{0.32\textwidth}
-		\centering
-		\includegraphics[scale=0.33]{impulse_batch1.png}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\centering
-		\includegraphics[scale=0.33]{impulse_batch2.png}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\centering
-		\includegraphics[scale=0.33]{impulse_batch3.png}
-	\end{subfigure}
-	\caption{Batches, $\eta_b$, of normalized impulses, $\impulseresponse_{x}$, for the Stokes inverse problem data misfit Gauss-Newton Hessian (Section~\ref{sec:numerical_results}). 
-%	Brightness scale differs between the subfigures. 
-	Black stars are point source locations. Dashed gray ellipses are estimated impulse response support ellipsoids. The large circle is $\partial \Omega$. 
-	}
-	\label{fig:batches_intro}
-\end{figure}
-
-%Existing scalable high rank approximation methods typically require the ability to rapidly evaluate arbitrary entries of $\Aop$'s integral kernel, which we cannot do. 
-We use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applies.
-The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
-We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with batches of points scattered throughout the domain (see Figure~\ref{fig:batches_intro}).
-Then we interpolate translated and scaled versions of these impulse responses to approximate entries of the operator's integral kernel.
-%This allows fast evaluation of approximate matrix entries, which are needed by high-rank approximation methods, in particular, classical H-matrix construction methods. 
-Picking the batches of points requires us to estimate the supports of the impulse responses $\phi_x$ \emph{before} we compute them. 
-In resolution analysis in seismic imaging, 
-$\Aop^T$ is applied to a random noise function, and the width of $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$~\cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
-Here, we estimate the support of the $\phi_x$ by applying $\Aop^T$ to polynomial functions instead of random noise functions (see Section~\ref{sec:intromoments}).
-Our method estimates the support of $\phi_x$ more accurately than random noise probing, at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
-
-Our method may be categorized as a PSF method that is loosely based on ``product convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
-PC and PSF methods have a long history dating back several decades.
-We note the following papers (among many others), \cite{Adorf94,AlgerEtAl19,BigotEscandeWeiss19,EscandeWeiss15,EscandeWeiss22,EscandeWeiss12,FishEtAl96,NagyOleary98,ZhuLiFomelEtAl16}, in which the convolution kernels are constructed from sampling impulse responses of the operator to scattered point sources.
-For background on PC and PSF methods, we recommend the following papers: \cite{DenisEtAl15,EscandeWeiss17,GentileCourbinMeylan13}.
-While PC approximations are based on an assumption of local translation invariance, the method we propose is based on a more general assumption we call ``local mean displacement invariance'' (Section~\ref{sec:local_mean_displacement_invariance} and Figure~\ref{fig:mean_displacement_invariance}).
-%Furthermore, PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulses.
-We avoid this issue by temporarily excluding impulses from the approximation procedure whenever an attempt is made to evaluate the impulse at a location where it is undefined.
-
-%\begin{figure}
-%	\begin{center}
-%		\includegraphics[scale=0.4]{bct1.eps}
-%	\end{center}
-%	\caption{Hierarchical matrix block structure for the Stokes Hessian. The red blocks (tiny) are dense. All other blocks are low rank.
-%	}
-%	\label{fig:hmatrix_intro}
-%\end{figure}
-
-\begin{figure}
-	\begin{center}
-		\includegraphics[scale=0.75]{mean_displacement_invariance.pdf}
-	\end{center}
-	\caption{$\Aop$ is locally mean displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$.
-Here $\spatialmean(z)$ denotes the mean (center of mass) of $\phi_z$.}
-	\label{fig:mean_displacement_invariance}
-\end{figure}
-
-The ability to rapidly approximate entries of $\Aop$'s integral kernel allows one to approximate discretized versions of $\Aop$ using the full arsenal of tools for matrix approximation that rely on fast access to matrix entries.
-%one to approximate discretized matrix representations of $\Aop$ using methods that require fast access to arbitrary matrix entries. We 
-We form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of $\Aop$.
-H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank.
-%(see Figure~\ref{fig:hmatrix_intro} and Appendix~\ref{app:h_matrix}).
-H-matrix methods permit us to perform matrix-vector products cheaply, and perform other useful linear algebra operations that cannot be done easily using the original operator.
-These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion.
-%These H-matrix methods are fast and scale well to large problems. 
-The work and memory required to perform these operations for an $\fedim \times \fedim$ H-matrix scales 
-%as $O(\fedim \log(\fedim)^a)$, where $a \in \{0,1,2,3\}$. That is, the cost is \emph{nearly-linear}
-\emph{nearly linearly} in $N$ (i.e., $o(N^{1+\epsilon})$ for any $\epsilon > 0$).
-The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks~\cite{GrasedyckHackbusch03}.
-These operations can be used, for example, to solve Newton linear systems in PDE constrained optimization, sample from ill-conditioned posterior distributions in Bayesian inverse problems, and construct high rank Gaussian surrogate models that can be used for uncertainty quantification.
-%Classical H-matrix construction methods 
-%(described in Appendix~\ref{sec:H_matrix_construction}) 
-%require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.
-
-\section{Why we need high rank Hessian approximations}
-\label{sec:hessian}
-
-In distributed parameter inverse problems governed by PDEs, one seeks to infer an unknown spatially varying parameter field from limited observations of a state variable that depends on the parameter implicitly through the solution of a PDE.
-%Inverse problems are typically ill-posed, in the sense that many different parameter fields may yield predicted observations that agree with the actual observations to within the noise level. 
-Conventionally, the inverse problem is formulated using either a deterministic framework~\cite{AsterBorchersThurber18,Vogel02},
-%Guy Chavent, Nonlinear Least Squares for Inverse Problems, Springer, 2009.
-or a Bayesian probabilistic framework~\cite{KaipioSomersalo05,Stuart10a,Tarantola05}.
-In the deterministic framework, one solves an optimization problem to find the parameter that best fits the observations, subject to appropriate regularization~\cite{EnglHankeNeubauer96}.
-In the probabilistic framework, Bayes' theorem combines the observations with prior information to form a posterior distribution over the space of all possible parameter fields, and computations are performed to extract statistical information about the parameter from this posterior.
-The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal to each other under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian.'' The Hessian consists of a data misfit term (the ``data misfit Hessian''), which depends on discrepancy between the observations and the associated model predictions, and a regularization or prior term (the ``regularization Hessian'') which does not depend on the observations.
-For more details on the Hessian, see~\cite{Alger19,GhattasWillcox21,VillaPetraGhattas21}.
-
-Hessian approximations and preconditioners are highly desirable because the Hessian is central to efficient solution of inverse problems in both the deterministic and Bayesian settings.
-When solving the deterministic optimization problem with Newton-type methods, the Hessian is the coefficient operator for the linear system that must be solved or approximately solved at every Newton iteration.
-Good Hessian preconditioners reduce the number of iterations required to solve these Newton linear systems with Krylov methods such as conjugate gradient~\cite{Saad03}.
-In the Bayesian setting, the inverse of the Hessian is the covariance of a local Gaussian approximation of the posterior.
-Such a Gaussian distribution can be used as a direct approximations of the posterior, or as a proposal for Markov chain Monte-Carlo methods for drawing samples from the posterior.
-For instance, see~\cite{KimEtAl21,PetraEtAl14} and the references therein.
-
-%\begin{figure}
-%	\begin{subfigure}{0.5\textwidth}
-%		\includegraphics[scale=0.23]{informed_uninformed_modes_quadratic.pdf}
-%		\caption{}
-%	\end{subfigure}
-%	\begin{subfigure}{0.48\textwidth}
-%		\includegraphics[scale=0.1]{misfit_and_reg_hessian_eigenvalues.pdf}
-%		\caption{}
-%	\end{subfigure}
-%	\caption{(a) The Hessian eigenstructure characterizes the directional scalings of the best local quadratic approximation to the objective function in an inverse problem.
-The larger an eigenvalue of the Hessian is, the more closely spaced the contours of the local quadratic approximation are in the corresponding eigenvector direction.
-(b) As the data in an inverse problem becomes more informative about the parameter, the eigenvalues of the data misfit Hessian increase.
-As the regularization is weakened, the eigenvalues of the regularization Hessian decrease.
-These illustrations are taken from~\cite[Chapter 1]{Alger19}, with minor modifications.}
-%	\label{fig:hessian_eigenstructure}
-%\end{figure}
-
-Owing to the implicit dependence of predicted observations on the parameter, entries of the Hessian are not easily accessible.
-Rather, the Hessian may be applied to a vector via a computational process that involves solving a pair of forward and adjoint PDEs which are linearizations of the original PDE~\cite{GhattasWillcox21,Gunzburger03}.
-The most popular matrix-free Hessian approximation methods are based on low rank approximation of either the data misfit Hessian, or the data misfit Hessian preconditioned by the regularization Hessian, e.g.,~\cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15}.
-Krylov methods such as Lanczos or randomized methods~\cite{Cheng05,HalkoMartinssonTropp11} are typically used to construct these low rank approximations by applying the Hessian to vectors.
-Using these methods, the required number of Hessian applications (and hence the required number of PDE solves) is proportional to the rank of the low rank approximation.
-Low rank approximation methods are justified by arguing that the numerical rank of the data misfit Hessian is insensitive to the dimension of the discretized parameter.
-This means that the required number of PDE solves remains the same as the mesh used to discretize the parameter is refined.
-However, in many inverse problems of practical interest the numerical rank of the data misfit Hessian, while mesh independent, is still large, which makes it costly to approximate the Hessian using low rank approximation methods~\cite{AmbartsumyanEtAl20,BuiEtAl13,IsaacEtAl15}.
-
-Examples of inverse problems with high rank data misfit Hessians include advection dominated advection-diffusion inverse problems~\cite{AkcelikBirosDraganescuEtAl05}\cite[Chapter 5]{Flath13}, high frequency wave inverse problems~\cite{BuiEtAl13}, and more generally, all inverse problems in which the observations highly inform the parameter.
-The eigenvalues of the data misfit Hessian characterize the sensitivity of the predicted observations to perturbations of the parameter in the corresponding eigenvector directions,
-%(see Figure~\ref{fig:hessian_eigenstructure})
-so more informative data leads to larger eigenvalues and therefore a larger numerical rank~\cite{AlexanderianGloorGhattas16}\cite[Section 1.4 and Chapter 4]{Alger19}.
-Roughly speaking, the numerical rank of the data misfit Hessian is the dimension of the subspace of parameter space that is informed by the data.
-The numerical rank of the regularization preconditioned data misfit Hessian may be reduced by increasing the strength of the regularization, but this throws away useful information---components of the parameter that could be learned from components of the observations that have a large signal to noise ratio are instead reconstructed based on the regularization~\cite[Section 4]{AlgerEtAl17}\cite[Chapters 1 and 7]{Vogel02}.
-Hence, low rank approximation methods suffer from a predicament---if the data highly inform the parameter and the regularization is chosen appropriately, then a large number of operator applies are required to form an accurate approximation of the Hessian using low rank approximation methods.
-%The best case scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost)~\cite{AlexanderianGloorGhattas16,Alger19,AlgerEtAl17}.
-High rank Hessian approximation methods are needed.
-
-Recently there have been improvements in matrix-free H-matrix construction methods in which an operator is applied to structured random vectors, and the response of the operator to those random vectors is processed to construct an H-matrix approximation~\cite{LevittMartinsson22,LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}.
-These methods (which we do not use here) have been used to approximate Hessians in PDE constrained inverse problems~\cite{AmbartsumyanEtAl20,HartlandEtAl23}. 
-Although the required number of matrix-vector products grows nearly linearly in the discretized parameter dimension and linearly in the rank of the off-diagonal blocks, the required number of matrix-vector products is still large (e.g., hundreds to thousands).
-%These methods are asymptotically scalable in the sense that the required number of matrix-vector products grows nearly linearly in the dimension of the discretized parameter and linearly in the rank of the off-diagonal blocks.
-But despite asymptotic scalability, from a practical perspective the required number of matrix-vector products is large (e.g., hundreds to thousands).
-%Improving the peeling process to reduce its cost is currently an area of active research~\cite{PEELINGRESEARCH}.
-%This is an area of active research, so more efficient and practical matrix-free H-matrix construction methods based on the peeling process may be discovered in the future.
-%Matrix-free H-matrix construction methods are currently an area of active research (see, e.g.,~\cite{LevittMartinsson22}), so more efficient matrix-free H-matrix methods may be discovered in the future.
-In this paper, we also form an H-matrix Hessian approximation.
-However, to reduce the required number of operator actions, we first form a PSF approximation of the data misfit Hessian by exploiting locality and non-negative integral kernel properties, then form the H-matrix using classical H-matrix techniques.
-Not all data misfit Hessians satisfy the local non-negative integral kernel properties (for example, the wave inverse problem data misfit Hessian has a substantial amount of negative entries in its integral kernel).
-However, many data misfit Hessians of practical interest do satisfy these properties (either exactly or approximately), and our method is targeted at approximating these Hessians.
-
-%Classical methods for approximating the Hessian are based on low rank approximation of either the data misfit Hessian, or the data misfit Hessian preconditioned by the inverse of the prior term in the Hessian.
-However,  More generally, the eigenvalues of the Hessian characterize how informative the data are about the component of the parameter in the corresponding parameter direction.
-The more informed the component of the parameter, the larger
-
-%In the deterministic setting, the Hessian of the objective function is central to efficient solution of the optimization problem, and in the
-
-% given the observed data is  one explores the posterior probability distribution for the parameter
-
-%Our motivation is high rank Hessian approximation in distributed parameter inverse problems governed by PDEs. 
-%The Hessian is an operator that is central to efficient methods for solving such inverse problems and is only available matrix-free; we will discuss Hessians in inverse problems governed by PDEs in greater detail in Section~\ref{sec:hessian}.
-
-%Introduce inverse problem
-%
-%Role of Hessian
-%
-%Structure of Hessian that we want to exploit
-%
-%Low rank
-%
-%Need for High rank
-%
-%Low rank approximation of the Hessian:
-%\begin{itemize} 
-%	\item Hessian eigenstructure is determined by data informativeness. Bigger eigenvalue => more informed component in eigenvector direction. 
-%	\item informative data => high rank hessian
-%	\item get high rank approximation by exploiting different structure. Locality, non-negativity of kernel
-%	\item Large rank in absolute terms, even if low rank in relative terms
-%\end{itemize}
-
-\section{Preliminaries}
-
-Let $\Omega \subset \mathbb{R}^\gdim$ be a bounded domain (typically $\gdim=1$, $2$, or $3$). We seek to approximate integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ of the form
-\begin{equation}
-\label{eq:kernel_representation}
-(\Aop u)(w) := \int_\Omega \int_\Omega w(y) \Aker(y,x) u(x) dx dy.
-\end{equation}
-Here the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$.
-The integral kernel, $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$, exists in principle but is not easily accessible.
-In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define impulse responses (Section~\ref{sec:impulse_response_2}), we state the conditions on $\Aop$ that our method requires (Section~\ref{sec:conditions_2}), and we detail finite element discretization (Section~\ref{sec:finite_element_kernel}).
-
-\subsection{Distributions and impulse responses}
-\label{sec:impulse_response_2}
-
-The action of $\Aop$ may be extended to distributions if $\Aker$ is sufficiently regular. 
-%We derive the extension as follows. 
-Let $\rho \in L^2(\Omega)'$, and let $\rho^* \in L^2(\Omega)$ denote the Riesz representative of $\rho$ with respect to the $L^2(\Omega)$ inner product. We have
-\begin{subequations}
-	\label{eq:extension_to_distributions}
-	\begin{align}
-		\left(\Aop \genericdistribution^*\right)(w) &= \int_\Omega \int_\Omega w(y) \Aker(y,x) \genericdistribution^*(x) dx ~dy \\
-		&= \int_\Omega w(y) \int_\Omega \Aker(y,x) \genericdistribution^*(x) dx ~dy 
-		= \int_\Omega w(y) \genericdistribution\left(\Aker(y,\cdot)\right) dy, \label{eq:action_on_distribution}
-	\end{align}
-\end{subequations}
-where $\Aker(y,\cdot)$ denotes the function $x \mapsto \Aker(y,x)$.
-Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions and let $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ be a distribution.
-In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in~\eqref{eq:extension_to_distributions} is not valid.
-However, if $\Aker$ is sufficiently regular such that the function $y \mapsto \rho\left(\Aker(y,~\cdot~)\right)$ is well defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of~\eqref{eq:action_on_distribution} is well defined.
-Hence, we \emph{define} the action of $\Aop$ on the distribution $\rho$ to be the right hand side of~\eqref{eq:action_on_distribution}.
-We denote this action by ``$\Aop \genericdistribution^*$,'' even if $\rho^*$ does not exist.
-
-Let $\delta_x$ denote the delta distribution\footnote{Recall that the delta distribution $\delta_x:\mathcal{D}(\Omega)\rightarrow \mathbb{R}$ is defined by $\delta_x(w) = w(x)$ for all $w\in \mathcal{D}(\Omega)$.} (i.e., point source, impulse) centered at the point $x \in \Omega$.
-The \emph{impulse response} of $\Aop$ associated with $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$, 
-\begin{equation}
-\label{eq:impulse_response_delta_action}
-\impulseresponse_x := \left( \Aop \delta_x^* \right)^*,
-\end{equation}
-that is formed by applying $\Aop$ to $\delta_x$, then taking the Riesz representation of the resulting linear functional.
-Using~\eqref{eq:action_on_distribution} and the definition of the delta distribution, we see that $\phi_x$ may also be written as the function $\impulseresponse_x(y) = \Aker(y, x)$.
-%\begin{equation}
-%\label{eq:impulse_response_defn}
-%\impulseresponse_x(y) = \Aker(y, x),
-%\end{equation} 
-%which results from keeping $x$ fixed and viewing $\Aker(y,x)$ as a function of $y$.
-
-\subsection{Required conditions}
-\label{sec:conditions_2}
-
-We focus on approximating operators that satisfy the following conditions: 
-\begin{enumerate}
-	\item The kernel $\Aker$ is sufficiently regular so that $\phi_x$ is well defined for all $x\in \Omega$.
-	\item The supports of the impulse responses $\impulseresponse_x$ are contained in localized regions.
-	\item The integral kernel is non-negative in the sense that $$\Aker(y,x) \ge 0$$ for all $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
-The operator $\Aop$ need not be positive semi-definite to use our method, and positive semi-definite operators need not have a non-negative integral kernel.}
-\end{enumerate}
-Our method may still perform well if these conditions are relaxed slightly.
-It is fine if the support of $\impulseresponse_x$ is not perfectly contained in a localized region, as long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
-The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$, as long as it is non-negative for the vast majority of pairs of points $(y,x)$, and as long as the negative numbers are comparatively small.
-If these conditions are violated, our method will incur additional error.
-If these conditions are violated too much, our method may fail.
-%Not all operators of interest (and in particular not all Hessians) satisfy these conditions, but many operators of practical interest do satisfy these conditions, and our method is highly effective for approximating these operators.
-
-\subsection{Finite element discretization}
-\label{sec:finite_element_kernel}
-
-In computations, functions are discretized and replaced by finite dimensional vectors, and operators mapping between infinite dimensional spaces are replaced by operators mapping between finite dimensional spaces.
-In this paper we discretize using continuous Galerkin finite elements satisfying the Kronecker property (defined below).
-With minor modifications, our method could be used with more general finite element methods, or other discretization schemes such as finite differences or finite volumes.
-
-Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of continuous Galerkin finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space under the $L^2$ inner product, and let $\fepoint_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
-We assume that the finite element basis satisfies the Kronecker property, i.e., $\febasis_i(\fepoint_i)=1$ and $\febasis_i(\fepoint_j)=0$ if $i\neq j$.
-For $u_h \in V_h$ we write $\bm{u} \in \mathbb{R}^\nsamplepts_\massmatrix$ to denote the coefficient vector for $u_h$ with respect to the finite element basis, i.e.,
-$
-%\label{eq:fem_coeff_basis}
-u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
-$
-Linear functionals $\rho_h \in V_h'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^\nsamplepts_{\massmatrix^{-1}}$, with entries $\boldsymbol{\rho}_i = \rho_h(\febasis_i)$ for $i=1,\dots,\nsamplepts$.
-%\begin{equation*}
-%\boldsymbol{\rho}_i = \rho_h(\febasis_i), \qquad i=1,\dots,\nsamplepts. 
-%\end{equation*}
-Here $\massmatrix \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries $\massmatrix_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx$ for $i,j=1,\dots,\fedim$.
-%\begin{equation*}
-%\massmatrix_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx, \qquad i,j=1,\dots,\fedim,
-%\end{equation*}
-The space $\mathbb{R}^\fedim_\massmatrix$ is $\mathbb{R}^\fedim$ with the inner product $(\mathbf{u},\mathbf{w})_\massmatrix := \mathbf{u}^T \massmatrix \mathbf{w}$, and $\mathbb{R}^\fedim_{\massmatrix^{-1}}$ is the analogous space with $\massmatrix^{-1}$ replacing $\massmatrix$.
-Direct calculation shows that $\mathbb{R}^\fedim_\massmatrix$ and $\mathbb{R}^\fedim_{\massmatrix^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively.
-
-After discretization, the operator $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$ is replaced by an operator $A_h:V_h \rightarrow V_h'$, which becomes an operator
-\begin{equation*}
-\mathbf{A}:\mathbb{R}^\fedim_\massmatrix \rightarrow \mathbb{R}^\fedim_{\massmatrix^{-1}}
-\end{equation*}
-under the isomorphism discussed above.
-Our method is agnostic to the computational procedure for approximating $\Aop$ with $\mathbf{A}$.
-What is important is that we do not have direct access to matrix entries $\mathbf{A}_{ij}$.
-Rather, we have a computational procedure that allows us to compute matrix-vector products $\mathbf{u}\mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{w}\mapsto \mathbf{A}^T\mathbf{w}$, and computing these matrix-vector products is costly.
-Of course, matrix entries can be computed via matrix-vector products as $\mathbf{A}_{ij} = \left(\mathbf{A}\mathbf{e}_j\right)_i$,
-where $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)^T$ is the length $\fedim$ unit vector with one in the $j$th coordinate and zeros elsewhere.
-But computing the matrix-vector product $\mathbf{e}_j \mapsto \mathbf{A}\mathbf{e}_j$ is costly, and therefore wasteful if we do not use other matrix entries in the $j$th column of $\mathbf{A}$.
-Hence, methods for approximating $\mathbf{A}$ are computationally intractable if they require accessing scattered matrix entries from many different rows and columns of $\mathbf{A}$.
-
-The operator $A_h:V_h \rightarrow V_h'$ can be written in integral kernel form,~\eqref{eq:kernel_representation}, but with $\Aker$ replaced by a slightly different integral kernel,
-%as
-%\begin{equation}
-%(A_h u_h)(w_h) := \int_\Omega \int_\Omega w_h(y) \Aker_h(y,x) u_h(x) dx dy
-%\end{equation}
-$\Aker_h$, which we do not know, and which differs from $\Aker$ due to discretization error.
-Since the functions in $V_h$ are continuous at $x$, the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta}_x \in \mathbb{R}^\fedim_{\massmatrix^{-1}}$ with entries $\left(\boldsymbol{\delta}_x\right)_i = \febasis_i(x)$ for $i=1,\dots,\fedim$.
-Additionally, it is straightforward to verify that the Riesz representation, $\rho_h^* \in V_h$, of a functional $\rho \in V_h'$ has coefficient vector
-$
-\boldsymbol{\rho}^* = \massmatrix^{-1} \boldsymbol{\rho}.
-$
-Therefore, the formula for the impulse response from~\eqref{eq:impulse_response_delta_action} becomes 
-$
-%\label{eq:discrete_kernel}
-\boldsymbol{\impulseresponse}_x = \left(A_h \delta_x^*\right)^* =  \massmatrix^{-1}\mathbf{A} \massmatrix^{-1} \boldsymbol{\delta}_x,
-$
-and the $(y,x)$ kernel entry of $\Aker_h$ may be written as
-$
-%\label{eq:approx_kernel_entry_81}
-	\Aker_h(y,x) = \boldsymbol{\delta}_y^T \boldsymbol{\impulseresponse}_x = \boldsymbol{\delta}_y^T \massmatrix^{-1}\mathbf{A} \massmatrix^{-1} \boldsymbol{\delta}_x.
-$
-Now define $\mathbf{\Aker} \in \mathbb{R}^{\fedim \times \fedim}$ to be the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
-\begin{equation}
-\label{eq:Akerpcmat_entries}
-\mathbf{\Aker}_{ij} := \Aker_h(\fepoint_i, \fepoint_j).
-\end{equation}
-Because of the Kronecker property of the finite element basis, we have $\boldsymbol{\delta}_{\fepoint_i} = \mathbf{e}_i$. Thus, we have
-$
-	\Aker_h(\fepoint_i,\fepoint_j) = \left(\massmatrix^{-1}\mathbf{A} \massmatrix^{-1}\right)_{ij},
-$
-which implies
-\begin{equation}
-\label{eq:boldA}
-	\mathbf{A} = \massmatrix \mathbf{\Aker} \massmatrix.
-\end{equation}
-Broadly, our method will construct an H-matrix approximation of $\mathbf{A}$ by forming H-matrix approximations of $\boldsymbol{\Aker}$ and $\massmatrix$ (or a lumped mass version of $\massmatrix$), then multiplying these matrices per~\eqref{eq:boldA} using H-matrix methods.
-Classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Aker}_{ij}$, but these matrix entries are not easily accessible.
-The bulk of our method is therefore dedicated to forming approximations of these matrix entries that can be evaluated rapidly.
-
-\paragraph{Lumped mass matrix} At the continuum level, $\Aker$ is assumed to be non-negative.
-However, entries of $\Aker_h$, involve inverse mass matrices, which typically contain negative numbers.
-If there are too many negative numbers, or if the negative numbers are too large, our algorithm will be less robust.
-We therefore recommend replacing the mass matrix, $\massmatrix$, with a positive diagonal \emph{lumped mass} approximation.
-In our numerical results, we use the lumped mass matrix constructed by replacing off-diagonal entries of the mass matrix by zero. Other mass lumping techniques may be used~\cite{Hughes12}.
-
-\section{Key innovations}
-\label{sec:prerequisites}
-
-In this section we present two key innovations that our method is based on.
-First, we define moments of the impulse responses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section~\ref{sec:intromoments}).
-%Our method will use these ellipsoid estimates to choose sets of impulses that can be computed in batches. 
-Second, we describe an improved method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section~\ref{sec:local_mean_displacement_invariance}).
-%Our method will interpolate impulses using this approximation.
-
-\subsection{Impulse response moments and ellipsoid support estimate}
-\label{sec:intromoments}
-
-The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative integral kernel property. Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
-\begin{equation}
-\label{eq:define_vol}
-\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy
-\end{equation}
-denote the spatially varying scaling factor, and for $i,j=1,\dots,\gdim$ define $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$ and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$ as follows:
-\begin{align}
-\spatialmean^i(x) :=& \int_\Omega (\impulseresponse_x(y) / V(x)) y^i ~dy \label{eq:define_mean} \\
-\spatialcov^{ij}(x) :=& \int_\Omega (\impulseresponse_x(y) / V(x)) \left(y^i - \spatialmean^i(x)\right) \left(y^j - \spatialmean^j(x)\right) ~dy, \label{eq:define_cov}
-\end{align}
-where $\spatialmean^i(x)$ denotes the $i^\text{th}$ component of the vector $\spatialmean(x)$, and $\spatialcov^{ij}(x)$ denotes the $(i,j)$ entry of the matrix $\spatialcov(x)$.
-The vector $\spatialmean(x)\in \mathbb{R}^\gdim$ and the matrix $\spatialcov(x) \in \mathbb{R}^{\gdim \times \gdim}$ are the mean and covariance of the normalized version of $\impulseresponse_x$, respectively.
-
-The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to obtain $\impulseresponse_x$, as per~\eqref{eq:impulse_response_delta_action}.
-Then one can post process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
-However, this direct approach is not feasible because we need to know $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ before we compute $\phi_x$, in order choose the point $x$.
-Computing $\phi_x$ in order to determine $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ would be extremely computationally expensive, and defeat the purpose of our algorithm, which is to reduce the computational cost by computing impulse responses in batches.
-Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ indirectly, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions (e.g., 6 total operator applies in two spatial dimensions and 10 in three spatial dimensions).
-This may be motivated by analogy to matrices.
-If $\mathbf{B}\in \mathbb{R}^{\fedim \times \fedim}$ is a matrix with $i^\text{th}$ column $\mathbf{b}_i$ and $\mathbf{w} \in \mathbb{R}^\fedim$, then
-\begin{equation*}
-\mathbf{A}^T \mathbf{w} = \begin{bmatrix}
-\horzbar & \mathbf{a}_1^T & \horzbar \\
-%\horzbar & \mathbf{a}_2^T & \horzbar \\
-& \vdots & \\
-\horzbar & \mathbf{a}_\fedim^T & \horzbar
-\end{bmatrix}
-\mathbf{w} = 
-\begin{bmatrix}
-\mathbf{a}_1^T \mathbf{w} \\
-%\mathbf{a}_2^T \mathbf{w} \\
-\vdots \\
-\mathbf{a}_\fedim^T \mathbf{w}
-\end{bmatrix}.
-\end{equation*}
-By computing one matrix-vector product of $\mathbf{A}^T$ with $\mathbf{w}$, we compute the inner product of each column of $\mathbf{A}$ with $\mathbf{w}$ simultaneously.
-The operator case is analogous, with $\phi_x$ taking the place of a matrix column.
-We have
-\begin{equation}
-\label{eq:operator_simultaneous_ips}
-\left(\Aop^T w\right)^*(x) = \int_\Omega \Aker(y,x) w(y) dy = \left(\phi_x, w\right)_{L^2(\Omega)}.
-\end{equation}
-By computing one operator application of $\Aop^T$ to $w$, we compute the inner product of each $\phi_x$ with $w$, for all points $x$ simultaneously.
-
-Let $C$, $L^i$, and $Q^{ij}$ be the following constant, linear, and quadratic functions:
-\begin{equation*}
-C(x) := 1, \qquad
-L^i(x) := x^i, \qquad
-Q^{ij}(x) := x^i x^j
-\end{equation*}
-for $i,j=1,\dots,\gdim$. Using the definition of $\spatialvol$ in~\eqref{eq:define_vol} and using~\eqref{eq:operator_simultaneous_ips}, we have
-\begin{equation*}
-	\spatialvol(x) = \int_\Omega \phi_x(y) C(y) ~ dy = \left(\phi_x, C\right)_{L^2(\Omega)} = \left(\mathcal{A}^T C\right)^*(x).
-\end{equation*}
-Hence, we compute $\spatialvol(x)$ for all $x$ simultaneously by applying $\Aop^T$ to $C$.
-Analogous manipulations show that $\spatialmean(x)$ and $\spatialcov(x)$ may be computed for all points $x$ simultaneously by applying $\Aop^T$ to the functions $L^i$ and $Q^{ij}$, respectively.
-We have
-\begin{subequations}
-	\label{eq:vol_mean_var}
-	\begin{align}
-	\spatialvol =& \left(\Aop^T C\right)^* \\
-	\spatialmean^i =& \left(\Aop^T L^i\right)^* / \spatialvol \\
-	\spatialcov^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j
-	\end{align}
-\end{subequations}
-for $i,j=1,\dots, \gdim$. Here $u/w$ denotes pointwise division, $\left(u/w\right)(x) = u(x)/w(x)$, and $u\cdot w$ denotes pointwise multiplication, $(u \cdot w)(x) = u(x)w(x)$.
-%The resulting method for computing discretized versions of $\spatialvol$, $\spatialmean$, and $\spatialcov$ is shown in Algorithm~\ref{alg:varhpi_mean_cov}.
-
-%\begin{algorithm2e}
-%	\SetAlgoNoLine
-%	\SetKwInOut{Input}{Input}
-%	\SetKwInOut{Output}{Output}
-%	{	
-%		\Input{Operator $\bm{A}^T$}
-%		\Output{$\bm{\spatialvol}\in \mathbb{R}^\fedim$, $\bm{\spatialmean}\in \mathbb{R}^{\gdim \times \fedim}$, and $\bm{\spatialcov} \in \mathbb{R}^{\gdim \times \gdim \times \fedim}$}
-%		
-%		Form vector $\bm{C} \in \mathbb{R}^\fedim_{\massmatrix}$ by either projecting or interpolating constant function $C(x)=1$ onto $V_h$
-%		
-%		$\bm{\spatialvol} \gets \massmatrix^{-1}\bm{A}^T \bm{C}$
-%		
-%		\For{$i=1,2,\dots,\gdim$}{
-%			Form vector $\bm{L}^i \in \mathbb{R}^\fedim_{\massmatrix}$ by either projecting or interpolating linear function $L^i(x) = x^i$ onto $V_h$
-%			
-%			$\bm{\spatialmean}^i \gets \left(\massmatrix^{-1}\bm{A}^T \bm{L}^i\right) / \bm{\spatialvol}$
-%		}
-%		\For{$i=1,2,\dots,\gdim$}{
-%			\For{$j=1,\dots,i$}{
-%				Form quadratic function $\bm{Q}^{ij} = x^i x^j$
-%				
-%				$\bm{\spatialcov}^{ij} \gets \left(\massmatrix^{-1}\bm{A}^T \bm{Q}^{ij}\right) / \bm{\spatialvol} - \bm{\spatialmean}^i\cdot \bm{\spatialmean}^j$
-%				
-%				$\bm{\spatialcov}^{ji} \gets \bm{\spatialcov}^{ij}$
-%				
-%			}
-%		}
-%		
-%	}
-%	\caption{Compute discretized scaling factor $\mathbf{\spatialvol}$, mean $\boldsymbol{\spatialmean}$, and covariance $\boldsymbol{\spatialcov}$}
-%	\label{alg:varhpi_mean_cov}
-%\end{algorithm2e}
-
-We approximate the support of $\impulseresponse_x$ with the ellipsoid
-\begin{equation}
-\label{eq:support_ellipsoid}
-E_x := \{x' \in \Omega: (x' - \spatialmean(x))^T \spatialcov(x)^{-1} (x' - \spatialmean(x)) \le \tau^2\},
-\end{equation}
-where $\tau$ is a fixed constant.
-The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as the normalized version of $\impulseresponse_x$.
-The quantity $\tau$ is a parameter that must be chosen appropriately.
-The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$.
-However, in Section~\ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids is better), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
-In practice, we find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results.
-The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is typically conservative.
-% and typically far less mass resides in this region.
-
-\subsection{Local mean displacement invariance}
-\label{sec:local_mean_displacement_invariance}
-
-Let $x$ and $x'$ be points in $\Omega$ that are ``close'' to each other, and consider the following approximations:
-\begin{align}
-\phi_x(y) &\approx \phi_{x'}(y) \label{eq:translate1}\\
-\phi_x(y) &\approx \phi_{x'}(y-x+x') \label{eq:translate2}\\
-\phi_x(y) &\approx \phi_{x'}\left(y-\spatialmean(x)+\spatialmean(x')\right) \label{eq:translate3}\\
-\phi_x(y) &\approx \phi_{x'}\left(y-\spatialmean(x)+\spatialmean(x')\right) \spatialvol(x) / \spatialvol(x'). \label{eq:translate4}
-\end{align}
-These are four different ways to approximate an impulse response by a nearby impulse response, with each successive approximation building upon the previous ones.
-Our method uses~\eqref{eq:translate4}, which is the most sophisticated of these approximations.
-Approximation~\eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close.
-Operators satisfying~\eqref{eq:translate1} can be well approximated via low rank CUR approximation~\cite{MahoneyDrineas09}.
-However, the required rank in the low rank approximation can be large, which makes algorithms based on~\eqref{eq:translate1} expensive.
-Operators that satisfy~\eqref{eq:translate2} are called ``locally translation invariant'' because integral kernel entries $\Aker(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$.
-It is straightforward to show that if equality holds in~\eqref{eq:translate2}, then $\Aop$ is a convolution operator.
-Locally translation invariant operators act like convolutions locally, and can therefore be well approximated by PC approximations.
-
-Approximation~\eqref{eq:translate3} improves upon~\eqref{eq:translate1} and~\eqref{eq:translate2}, and generalizes both.
-On one hand, if~\eqref{eq:translate1} holds, then $\spatialmean(x) \approx \spatialmean(x')$, and so~\eqref{eq:translate3} holds.
-On the other hand, translating a distribution translates its mean, so if~\eqref{eq:translate2} holds, then $\spatialmean(x')-\spatialmean(x) \approx x' - x$, so again~\eqref{eq:translate3} holds.
-But approximation~\eqref{eq:translate3} can hold in situations where neither~\eqref{eq:translate1} nor~\eqref{eq:translate2} holds.
-For example, because the expected value commutes with affine transformations,~\eqref{eq:translate3} will hold when $\Aop$ is locally translation invariant with respect to a translated frame of reference, while~\eqref{eq:translate2} will not hold in this case.
-Additionally,~\eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions.
-In contrast,~\eqref{eq:translate2} does not naturally generalize to operators that map between function spaces on different domains, because the formula $y-x+x'$ requires vectors in $\Omega_2$ and $\Omega_1$ to be added together.
-We call~\eqref{eq:translate3} ``local mean displacement invariance,'' and illustrate~\eqref{eq:translate3} in Figure~\ref{fig:mean_displacement_invariance}.
-
-We use approximation~\eqref{eq:translate4}, which is the same as~\eqref{eq:translate3}, except for the factor $\spatialvol(x)/\spatialvol(x')$.
-This factor makes the approximation more accurate if $\spatialvol(x)$ varies widely.
-Approximation~\eqref{eq:translate4} is equivalent to~\eqref{eq:translate3}, but with $\phi_x$ replaced by its normalized version, $\phi_x/\spatialvol(x)$.
-We call~\eqref{eq:translate4} \emph{normalized local mean displacement invariance}.
-
-%\paragraph{Future extension: approximation using $\Sigma$} One may map $E_x$ to the unit sphere via the change of variables $z = \Sigma^{-1/2}\left(x\right)\left(y - \spatialmean(x)\right) / \tau$, and approximate
-%\begin{equation}
-%\label{eq:sigma_interpolation}
-%\phi_x(z) \det{\Sigma^{1/2}\left(x\right)}/\spatialvol(x) \approx \phi_{x'}(z') \det{\Sigma^{1/2}\left(x'\right)}/\spatialvol(x').
-%\end{equation}
-%%Notice that equality holds in \eqref{eq:sigma_interpolation} if the impulse responses are Gaussian functions. 
-%While interpolation schemes based on \eqref{eq:sigma_interpolation} have the potential to be more accurate than those based on \eqref{eq:translate4}, we were unable to construct an interpolation scheme based on \eqref{eq:sigma_interpolation} that is robust to negative entries in the integral kernel.
-The integral kernel should have non-negative entries per Condition 3 in Section \ref{sec:conditions_2}, but in practice some entries are negative due to discretization error.
-%%Therefore, in this paper we do not use \eqref{eq:sigma_interpolation}. 
-%%Constructing interpolation schemes based on \eqref{eq:sigma_interpolation} that are robust to discretization error is a promising direction for future research.
-
-%but we find they are not robust, and therefore we do not use them:
-%\begin{align*}
-%	\phi_{x}(y) &\approx \phi_{x'}\left(\spatialmean(x') + \Sigma_{x'}^{1/2} \Sigma_{x}^{-1/2} \left(y - \spatialmean(x)\right)\right) \spatialvol(x) / \spatialvol(x') \\
-%	\phi_{x}(y) &\approx \phi_{x'}\left(\spatialmean(x') + \Sigma_{x'}^{1/2} \Sigma_{x}^{-1/2} \left(y - \spatialmean(x)\right)\right) \left( \spatialvol(x) \det{\Sigma^{1/2}(x')} \right) / \left( \spatialvol(x') \det{\Sigma^{1/2}(x)} \right)
-%\end{align*}
-%In these schemes, one maps known impulse responses from their support ellipsoids to the unit sphere, performs interpolation on the unit sphere, then maps the resulting interpolated impulse response to the support ellipsoid associated with the target point.
-We find that computation of $\Sigma$ is more susceptible to numerical error than computation of $V$ and $\mu$, in the sense that spurious negative numbers in $\phi_x$ cause $\Sigma(x)$ to become highly ill-conditioned or even indefinite.
-When $\Sigma$ is used for picking sample points (Section \ref{sec:sample_point_selection}), we simply avoid picking sample points $p$ with ``bad'' $\Sigma(p)$.
-These schemes require using $\Sigma_x$ at arbitrary $x$, rather than just the sample points.
-When picking sample points, one may
-
-\section{Operator approximation algorithm}
-\label{sec:method}
-
-We use~\eqref{eq:vol_mean_var} to compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ by applying $\Aop^T$ to polynomial functions.
-Then we use~\eqref{eq:support_ellipsoid} to form ellipsoid shaped estimates for the support of the $\phi_x$'s, \emph{without} computing them.
-This allows us to compute large numbers of $\impulseresponse_{x_i}$ in ``batches,'' $\eta_b$ (see Figure~\ref{fig:batches_intro}).
-We compute one batch, denoted $\eta_b$, by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section~\ref{sec:get_impulse_response}).
-The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support ellipsoid for $\impulseresponse_{x_i}$ and the support ellipsoid for $\impulseresponse_{x_j}$ do not overlap if $i \neq j$ (Section~\ref{sec:sample_point_selection}).
-Because these supports do not overlap (or do not overlap much), we can post process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section~\ref{sec:get_impulse_response}).
-The process is repeated until a desired number of batches is reached.
-
-Once the batches $\eta_b$ are computed, we approximate the integral kernel $\Aker(y,x)$ at arbitrary points $(y,x)$ by interpolation of translated and scaled versions of the computed $\impulseresponse_{x_i}$ (Section~\ref{sec:approximate_kernel_entries}).
-The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section~\ref{sec:local_mean_displacement_invariance}.
-Specifically, we approximate $\Aker(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}(y - \spatialmean(x) + \spatialmean(x_i))$ for a small number of sample points $x_i$ near $x$.
-The weights are determined by radial basis function interpolation.
-
-The ability to rapidly evaluate approximate kernel entries $\Aker(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Aker}_H \approx \mathbf{\Aker}$, using the conventional adaptive cross H-matrix construction method (Section~\ref{sec:h_matrix_construction_short}).
-In this method, one forms low rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks.
-We then use H-matrix methods to convert $\boldsymbol{\Aker}_H$ into an H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$.
-%Schematically the algorithm proceeds as follows:
-%\begin{equation*}
-%\Aop \longrightarrow \spatialvol, \spatialmean, \spatialcov \longrightarrow \underbrace{\{x_i\}_{i=1}^\nsamplepts}_{\text{as }S_b\text{'s}} \longrightarrow \underbrace{\{\impulseresponse_{x_i}\}_{i=1}^\nsamplepts}_{\text{as }\eta_b\text{'s}} \longrightarrow \boldsymbol{\Aker}_H \longrightarrow \mathbf{A}_H.
-%\end{equation*}
-The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm~\ref{alg:construct_Atilde}. The cost is discussed in Section~\ref{sec:overall_cost}.
-
-When $\Aop$ is symmetric positive semi-definite, $\mathbf{A}_H$ may be non-symmetric and indefinite due to errors in the approximation.
-In this case, one may optionally symmetrize $\mathbf{A}_H$, then modify it via low rank updates to remove erroneous negative eigenvalues (Section \ref{sec:flip_eigs}).
-
-%.
-The matrix is symmetrized by computing $\left(\mathbf{A}_H+\mathbf{A}_H^T\right)/2$ using H-matrix addition.
-Negative eigenvalues modified to be positive using a rational method that is described in Appendix~\ref{sec:make_spd}.
-
-\begin{algorithm2e}
-	\SetAlgoNoLine
-	\SetKwInOut{Input}{Input}
-	\SetKwInOut{Output}{Output}
-	{	
-		\Input{Linear operator $\Aop$, parameter $\nbatch$}
-		\Output{H-matrix $\mathbf{A}_H$}
-
-		Compute $V, \mu$, and $\Sigma$ (Equations~\eqref{eq:vol_mean_var} in Section~\ref{sec:intromoments})
-
-%		Compute $\bm{\spatialvol}$, $\bm{\spatialmean}$, and $\bm{\spatialcov}$ using Algorithm~\ref{alg:varhpi_mean_cov}.
-
-		\For{$k=1,2,\dots,\nbatch$}{
-			Choose a batch of sample points, $\pointbatch_k$ (Section~\ref{sec:sample_point_selection})
-
-			Compute $\combresponse_k$ by applying $\Aop$ to the Dirac comb for $\pointbatch_k$ (Section~\ref{sec:get_impulse_response})
-
-		}
-
-		Form H-matrix approximation $\boldsymbol{\Aker}_H$ of integral kernel (Section~\ref{sec:h_matrix_construction_short})
-
-		Form H-matrix approximation $\mathbf{A}_H$ of $\Aop$ (Section~\ref{sec:h_matrix_construction_short})
-
-		(optional) Modify $\mathbf{A}_H$ to make it symmetric and remove negative eigenvalues (Section \ref{sec:flip_eigs})
-
-	}
-	\caption{Construct H-matrix approximation}
-	\label{alg:construct_Atilde}
-\end{algorithm2e}
-
-\subsection{Sample point selection via greedy ellipsoid packing}
-\label{sec:sample_point_selection}
-
-We choose sample points, $x_i$, in batches $\pointbatch_k$.
-We use a greedy ellipsoid packing algorithm to choose as many points as possible per batch, while ensuring that there is no overlap between the support ellipsoids, $E_{x_i}$, associated with the sample points within a batch.
-
-We start with a finite set of candidate points $\candidatepts$ and build $\pointbatch_k$ incrementally with points selected from $\candidatepts$.
-For simplicity of explanation, here $\pointbatch_k$ and $\candidatepts$ are mutable sets that we add points to and remove points from.
-First we initialize $\pointbatch_k$ as an empty set.
-Then we select the candidate point $\candidatepoint_i \in \candidatepts$ that is the farthest away from all points in previous sample point batches $S_1 \cup S_2 \cup \dots \cup S_{k-1}$.
-Candidate points for the first batch $S_1$ are chosen randomly from $\candidatepts$.
-Once $\candidatepoint_i$ is selected, we remove $\candidatepoint_i$ from $\candidatepts$. Then we perform the following checks:
-\begin{enumerate}
-	\item We check whether $\candidatepoint_i$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ for all $\candidatepoint_j \in \pointbatch_k$.
-	\item We make sure that $\spatialvol(\candidatepoint_i)$ is not too small, by checking whether $\spatialvol(\candidatepoint_i) > \epsilon_V \spatialvol_\text{max}$.
-Here $\spatialvol_\text{max}$ is the largest value of $\spatialvol(\candidatepoint_j)$ over all points $q$ in the initial set of candidate points, and $\epsilon_V$ is a small threshold parameter (we use $\epsilon_V=10^{-5}$).
-	\item We make sure that all eigenvalues of $\spatialcov(\candidatepoint_i)$ are positive, and the aspect ratio of $E_{\candidatepoint_i}$ (ratio of the largest eigenvalue of $\spatialcov(\candidatepoint_i)$ to the smallest) is bounded by a constant $1/\epsilon_\Sigma$ (we use $1/\epsilon_\Sigma=20$).
-Negative integral kernel entries due to discretization error can cause $\spatialcov(\candidatepoint_i)$ to be indefinite or highly ill-conditioned.
-We reject these points because it is likely that the set $E_{\candidatepoint_i}$ poorly reflects the support of $\phi_{\candidatepoint_i}$.
-\end{enumerate}
-If $\candidatepoint_i$ passes these checks 
-%(i.e., if $\candidatepoint_i$ is sufficiently far from other points in the batch, and $\spatialvol(\candidatepoint_i)$ is not too small, and the )
-then we add $\candidatepoint_i$ to $\pointbatch_k$. Otherwise we discard $\candidatepoint_i$. This process repeats until there are no more points in $\candidatepts$.  
-%This is detailed in Algorithm~\ref{alg:point_choice}. 
-%We check whether $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ using the fast ellipsoid intersection test in \cite{GilitschenskiHanebeck12}.
-described in Appendix~\ref{sec:fast_ellipsoid_intersection_test}. 
-We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots, \pointbatch_{\nbatch}$.
-For each batch, $\candidatepts$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, except for points in previous batches.
-
-%\begin{algorithm2e}
-%	\SetAlgoNoLine
-%	\SetKwInOut{Input}{Input}
-%	\SetKwInOut{Output}{Output}
-%	\SetKwProg{Fn}{Function}{}{}
-%	
-%	\Input{Finite set of candidate points $\candidatepts\subset \Omega$, \\spatially varying mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\previous sample point batches $\pointbatch_1, \dots, \pointbatch_{k-1}$}
-%	\Output{Batch of new sample points $\pointbatch_k$}
-%	
-%	Initialize empty new batch of sample points, $\pointbatch_k = \{\}$
-%	
-%	\While{$\candidatepts$ is not empty}{
-%		Determine the point $\candidatepoint_i \in \candidatepts$ that is farthest from all points in previous sample point batches $\pointbatch_1,\dots,\pointbatch_{k-1}$
-%		
-%		Remove $\candidatepoint_i$ from $\candidatepts$
-%
-%	
-%		\If{$E_{\candidatepoint_i} \cap E_q \neq \{\}$ for all $\candidatepoint_j \in \pointbatch_k$}{
-%			\tcp{$E_{\candidatepoint_i}$ and $E_{\candidatepoint_j}$ are the ellipsoids defined in~\eqref{eq:support_ellipsoid}}
-%			
-%			Add $\candidatepoint_i$ to $\pointbatch_k$
-%			
-%			Remove all points $\candidatepoint$ satisfying $\spatialmean(\candidatepoint) \in E_{\candidatepoint_i}$ from $\candidatepts$
-%		}
-%	}
-%
-%    \SetKwFunction{FMain}{point\_is\_acceptable}
-%	\caption{Choosing one batch of sample points, $\pointbatch_k$}
-%	\label{alg:point_choice}
-%\end{algorithm2e}
-
-\subsection{Impulse response batches}
-\label{sec:get_impulse_response}
-
-We compute impulse responses $\impulseresponse_{x_i}$ in batches by applying $\Aop$ to Dirac combs.
-The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the following weighted sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$:
-\begin{equation*}
-	\diraccomb_k := \sum_{x_i \in \pointbatch_k} \delta_{x_i} / \spatialvol(x_i).
-\end{equation*}
-We compute the \emph{impulse batch}, $\eta_k$, by applying $\Aop$ to the Dirac comb:
-\begin{equation}
-	\label{eq:dirac_comb_H_action}
-	\combresponse_k := \left(\Aop \diraccomb_k^*\right)^*
-	=\sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / \spatialvol(x_i).
-\end{equation}
-The last equality in~\eqref{eq:dirac_comb_H_action} follows from linearity and the definition of $\impulseresponse_{x_i}$ in~\eqref{eq:impulse_response_delta_action}.
-Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\impulseresponse_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\impulseresponse_j$ do not overlap when $i \neq j$, we have (approximately)
-\begin{equation}
-\label{eq:varphi_eval}
-	\impulseresponse_{x_i}(z) =
-	\begin{cases}
-		\combresponse_k(z) \spatialvol(x_i), & z \in E_{x_i} \\
-		0, & \text{otherwise}
-	\end{cases}
-\end{equation}
-for all $x_i \in \pointbatch_k$.
-By performing one matrix-vector product, $\xi_k \mapsto \left(\Aop \diraccomb_k^*\right)^*$, we recover $\impulseresponse_{x_i}$ for every point $x_i \in \pointbatch_k$.
-
-Each point source $\delta_{x_i}$ is scaled by $1/\spatialvol(x_i)$ so that the resulting scaled impulse responses within $\eta_k$ are comparable in magnitude.
-Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, may overwhelm $\phi_{x_j}$ for a nearby point $x_j$ if $\spatialvol(x_i)$ is much larger than $\spatialvol(x_j)$.
-Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section~\ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $\spatialvol(x_i)$ is smaller than a predetermined threshold.
-
-\subsection{Approximate integral kernel entries}
-\label{sec:approximate_kernel_entries}
-
-Given $(y,x)\in \Omega \times \Omega$, let
-\begin{equation*}
-	z_i := y - \spatialmean(x) + \spatialmean(x_i), \qquad i=1,\dots,\numnbr,
-\end{equation*}
-where $\{x_i\}_{i=1}^{\numnbr}$ are the $\numnbr$ nearest sample points to $x$, excluding points $x_i$ for which $z_i \notin \Omega$.
-Here $\numnbr$ is a small user defined parameter; e.g., $\numnbr=10$.
-We find the $\numnbr$ nearest sample points to $x$ by querying a precomputed kd-tree~\cite{Bentley75} of all sample points.
-We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree)~\cite{Ericson04} of the mesh cells used to discretize the problem.
-Further, let
-\begin{equation}
-\label{eq:fxyxp}
-f_i := \frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}\left(z_i\right), \qquad i=1,\dots,\numnbr.
-\end{equation}
-Note that $\phi_{x_i}\left(z_i\right)$ is well defined because $z_i \in \Omega$, and $\frac{\spatialvol(x)}{\spatialvol(x_i)}$ is well defined because our sample point picking procedure in Section~\ref{sec:sample_point_selection} ensures that $\spatialvol(x_i)>0$.
-Per the discussion in Section~\ref{sec:local_mean_displacement_invariance}, we expect $\Aker(y,x) \approx f_i$ for $i=1,\dots,\numnbr$; the closer $x_i$ is to $x$, the better we expect the approximation to be.
-We therefore approximate $\Aker(y,x)$ by interpolating the (point,value) pairs
-$
-\left\{\left(x_i, f_i\right)\right\}_{i=1}^{\numnbr}
-$
-at the point $x$. 
-%The idea is that $\Aker(y,x) = f_i$ per the discussion in Section~\ref{sec:local_mean_displacement_invariance}, so we approximate $\Aker(y,x)$ by interpolating the values $f_i$. 
-Interpolation is performed using the following radial basis function~\cite{Wendland04} scheme:
-\begin{equation}
-\Aker(y,x) \approx \widetilde{\Aker}(y,x) := \sum_{i=1}^{\numnbr} \rbfweight_i~ \rbf\left(\|x-x_i\|\right),
-\end{equation}
-where $\rbfweight_i$ are weights, and 
-$$\rbf(r):= \exp\left(-C_\text{RBF}^2\frac{r^2}{2 L^2}\right)$$
-is a Gaussian kernel radial basis function.
-Here $L:=\diam\left(\{x_i\}_{i=1}^{\numnbr}\right)$ is the diameter of the set of sample points used in the interpolation, and $C_\text{RBF}$ is a user-defined shape parameter that controls the width of the kernel function; we use $C_\text{RBF}=3.0$.
-%is a polyharmonic spline radial basis function, which is defined as $\rbf(r):=r^\gdim$ if $\gdim$ is odd, and $\rbf(r):=r^\gdim \log r$ for $r>0$ and $\rbf(0):=0$ if $\gdim$ is even.
-%\begin{equation}
-%\rbf(r) := \begin{cases}
-%r^\gdim, & \gdim=1,3,5,\dots \\
-%r^\gdim \log r, & \gdim=2,4,6,\dots
-%\end{cases}
-%\end{equation}
-%for $r>0$, and $\rbf(0):=0$. 
-The vector of weights, $\rbfweight = (\rbfweight_1, \rbfweight_2, \dots, \rbfweight_{\numnbr})^T$, is found as the solution to the $\numnbr \times \numnbr$ linear system
-\begin{equation}
-B \rbfweight = f,
-\end{equation}
-where $B \in \mathbb{R}^{\numnbr \times \numnbr}$ has entries
-$
-B_{ij} := \rbf\left(\|x_i - x_j\|\right),
-$
-and $f \in \mathbb{R}^{\numnbr}$ has entries $f_i$ from~\eqref{eq:fxyxp}. 
-%Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a least squares sense~\cite{Duchon77}. 
-%For more details on radial basis function interpolation, see~\cite{Wendland04}. 
-%The interpolation procedure is illustrate in Figure~\ref{fig:rbf_interpolation}.
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[scale=0.2]{interpolation_rbf_boundary.pdf}
-%	\caption{To compute the kernel entry $\Aker(y,x)$, we use radial basis functions to interpolate the values $\frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}(y-\spatialmean(x)-\spatialmean(x_i))$ for the $\numnbr$ nearest sample points $x_i$ to $x$, excluding sample points for which $y-\spatialmean(x)-\spatialmean(x_i)$ is outside of the domain.}
-%	\label{fig:rbf_interpolation}
-%\end{figure}
-
-To evaluate $f_i$, we check whether $z_i \in E_{x_i}$ using~\eqref{eq:support_ellipsoid}.
-If $z_i \notin E_{x_i}$, then $z_i$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_i=0$.
-If $z_i \in E_{x_i}$, we look up the batch index $b$ such that $x_i \in S_b$, and evaluate $f_i$ via the formula
-$
-f_i = \spatialvol(x)\eta_b\left(z_i\right),
-$
-per~\eqref{eq:varphi_eval}. Note that $z_i$ is typically not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints. Hence,
-evaluating $\eta_b\left(z_i\right)$ requires determining which mesh cell contains $z_i$, then evaluating finite element basis functions on that mesh cell.
-Fortunately, the mesh cell containing $z_i$ was determined as a side effect of querying the AABB tree of mesh cells when we checked whether $z_i \in \Omega$.
-
-\subsection{Hierarchical matrix construction}
-\label{sec:h_matrix_construction_short}
-
-We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming H-matrix representations $\mathbf{\Aker}_H$ and $\massmatrix_H$ of $\mathbf{\Aker}$ and $\massmatrix$, respectively, then using fast H-matrix methods to multiply these matrices per~\eqref{eq:boldA} to form
-$
-\mathbf{A}_H = \massmatrix_H \mathbf{\Aker}_H \massmatrix_H.
-$
-%Once $\mathbf{A}_H$ is formed, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods~\cite{BORMHACKBUSCHBOOK}.
-We use H1 matrices in our numerical results, but any other H-matrix format could be used instead. For more details on H-matrices, see~\cite{Hackbusch15}.
-
-We form $\mathbf{\Aker}_H$ using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package~\cite{Kriemann08}.
-For details about the algorithms used for geometrical clustering, H-matrix construction, and H-matrix operations in HLIBpro, we recommend the the following papers~\cite{BormGrasedyckHackbusch03,GrasedyckKriemannLeBorne08,Kriemann13}.
-Although $\mathbf{\Aker}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Aker}_H$ only requires evaluation of $O(\hrank^2 \fedim \log \fedim)$ kernel entries $\mathbf{\Aker}_{ij} = \widetilde{\Aker}(\fepoint_i, \fepoint_j)$, and these entries are computed via the radial basis function interpolation method described in Section~\ref{sec:approximate_kernel_entries}.
-Here $\hrank$ is the rank of the highest rank block in the H-matrix.
-A dense representation of $\mathbf{\Aker}$ is never formed. 
-%We describe this H-matrix construction process in detail in Appendix~\ref{app:h_matrix}. 
-We form $\massmatrix_H$ using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same geometrical clustering block cluster tree as was used for $\mathbf{\Aker}$.
-
-\subsection{Symmetrizing and flipping negative eigenvalues (optional)}
-\label{sec:flip_eigs}
-
-In many applications, one seeks to approximate an operator
-$\mathcal{H} = \Aop + \mathcal{R},$
-where $\Aop$ is a symmetric positive semi-definite operator that we approximate with our PSF method to form an H-matrix $\mathbf{A}_H$, and $\mathcal{R}$ is a symmetric positive definite operator that may be easily converted to an H-matrix $\mathbf{R}_H$ without using our PSF method.
-For example, in inverse problems $\mathcal{H}$ is the Hessian, $\Aop$ is the data misfit term in the Hessian which is dense and only available matrix-free, and $\mathcal{R}$ is the regularization term, which is typically an elliptic differential operator that becomes a sparse matrix after discretization.
-
-%In many problems $\Aop$ is symmetric positive semi-definite. 
-The PSF approximation $\mathbf{A}_H$, and therefore $\mathbf{A}_H + \mathbf{R}_H$, may be non-symmetric and indefinite because of approximation error. 
-This is undesirable because symmetry and positive semi-definiteness are important properties which should be preserved if possible.
-Also, lacking these properties may prevent one from using highly effective algorithms to perform further operations involving $\mathbf{A}_H + \mathbf{R}_H$, such as using $\mathbf{A}_H + \mathbf{R}_H$ as a preconditioner in the conjugate gradient method.
-
-We therefore modify $\mathbf{A}_H$ to make it symmetric and remove negative eigenvalues. First, we symmetrize $\mathbf{A}_H$ via
-$
-\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
-$
-Next, we find negative eigenvalues and their corresponding eigenvectors for the generalized eigenvalue problem
-$\mathbf{A}_H^{\text{sym}} \mathbf{u} = \lambda \mathbf{R}_H$
-using a Cayley shift-and-invert Krylov scheme \cite{LehoucqEtAl98}.
-We flip the signs of these eigenvalues to be positive instead of negative (i.e., $\lambda \rightarrow |\lambda|$) by performing a low rank update to $\mathbf{A}_H^{\text{sym}}$.
-The primary computational task in the Cayley shift-and-invert scheme is the solution of shifted linear systems of the form
-$\left(\mathbf{A}_H^{\text{sym}} + \mu_i \mathbf{R}_H\right) x = b,$
-for a small number of positive shifts $\mu_i$.
-We solve these linear systems by factorizing the matrices $\mathbf{A}_H^{\text{sym}} + \mu_i \mathbf{R}_H$ using fast H-matrix methods.
-We compute and flip all eigenvalues $\lambda < \tau$ which are less than some threshold $\tau \in (-1,0]$.
-By choosing $\tau > -1$, we ensure that the modified version of $\mathbf{A}_H^{\text{sym}} + \mathbf{R}_H$ is positive definite.
-Choosing $\tau=0$ would remove all erroneous negative eigenvalues.
-However, this is computationally infeasible if $\Aop$ has a large or infinite cluster of eigenvalues near zero, a common situation in practice.
-We therefore recommend choosing $\tau < 0$, and we use $\tau=-0.1$ in our numerical experiments.
-
-\section{Computational cost}
-\label{sec:overall_cost}
-
-The computational cost of our method may be divided into the costs to perform the following tasks: (1) Computing impulse moments and batches; (2) Building the H-matrix; (3) Performing linear algebra operations with the H-matrix.
-This may optionally include modifying the H-matrix to make it symmetric positive semi-definite.
-In practice, (1) is the dominant cost, because (1) is the only task that requires applying $\Aop$ to vectors, and our method is targeted at applications in which applying $\Aop$ to a vector requires an expensive computational procedure such as solving a large scale PDE.
-All operations that do not require applications of $\Aop$ to vectors are nearly linear, and therefore scalable, in the size of the problem, $N$.
-We now describe these costs in detail. For convenience, Table~\ref{tab:vars} lists variable symbols and their approximate sizes.
-
-\begin{table}
-	\begin{tabular}{lll}
-		Symbol & Typical size & Variable name \\
-		\hline
-		$\fedim$ & $10^2$--$10^9$ & Number of finite element degrees of freedom \\
-		$\nbatch$ & $1$--$25$ & Number of batches \\
-		$\hrank$ & $5$--$50$ & H-matrix rank \\
-		$\numnbr$ & $5$--$15$ & Number of nearest neighbors for RBF interpolation \\
-		$\gdim$ & $1$--$3$ & Spatial dimension \\
-%		$\classicalrank$ & $10^1$--$10^4$ & Classical matrix rank \\
-		$\nsamplepts$ & $10^1$--$10^4$ & Total number of sample points (all batches) \\
-		$|\pointbatch_i|$ & $1$--$500$ & Number of sample points in the $i$th batch \\
-		$\ratord$ & $1$--$2$ & Rational function parameter for SPSD modification
-	\end{tabular}
-	\caption{Symbols used for variables in computational cost estimates, and approximate ranges for their sizes in practice.}
-	\label{tab:vars}
-\end{table}
-
-\paragraph{(1) Computing impulse response moments and batches} Computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ requires applying $\Aop$ to $1$, $\gdim$, and $\gdim(\gdim+1)/2$ vectors, respectively.
-Computing each $\eta_i$ requires applying $\Aop$ to one vector, so computing $\{\eta_i\}_{i=1}^{\nbatch}$ requires $\nbatch$ operator applies.
-In total, computing all impulse response moments and batches therefore requires
-\begin{equation*}
-	1 + \gdim + \gdim(\gdim+1)/2 + \nbatch \qquad \text{operator applies.}
-\end{equation*}
-In a typical application one might have $\gdim=2$ and $\nbatch=5$, in which case a modest $11$ operator applies are required.
-
-Computing the impulse response batches also requires choosing sample point batches via the greedy ellipsoid packing algorithm described in Section~\ref{sec:sample_point_selection}.
-Choosing the $i$th batch of sample points may require performing up to $\fedim |\pointbatch_i|$ ellipsoid intersection tests, where $|\pointbatch_i|$ is the number of sample points in the $i$th batch.
-Choosing all of the sample points therefore requires performing at most
-\begin{equation*}
-	\fedim \nsamplepts \qquad \text{ellipsoid intersection tests,}
-\end{equation*}
-where $\nsamplepts$ is the total number of sample points in all batches.
-The multiplicative dependence of $\fedim$ with $\nsamplepts$ is undesirable since $\nsamplepts$ may be large, and reducing this cost is a target for future work.
-%Several improvements are possible.
-The cost of picking sample points could be reduced by ``thinning'' the set of candidate points that the sample points are picked from.
-That is, choosing sample points from a smaller subset of $\fedim_\text{thin} << \fedim$ points.
-The cost to perform the ellipsoid intersection tests could be reduced by first checking whether the ellipsoid bounding boxes intersect before performing the ellipsoid intersection test of Appendix~\ref{sec:fast_ellipsoid_intersection_test}.
-The cost of picking the $i$th batch could be reduced to $O(\fedim \log |\pointbatch_i|)$ using more advanced computational geometry techniques, such as organizing the ellipsoids associated with points that have already been picked into a dynamic AABB tree, then using this AABB tree to accelerate the required ellipsoid intersection tests.
-However, from a practical perspective, the cost for picking sample points is small compared to other parts of the approximation algorithm, so we do not pursue these improvements in this paper.
-
-\paragraph{(2) Building the H-matrix} The classical H-matrix construction technique
-%, which is described in Appendix~\ref{app:h_matrix}, 
-requires evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix.
-To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulse responses used in the RBF interpolation.
-This is done using a precomputed kd-tree of sample points, and requires $O(\numnbr \log \nsamplepts)$ floating point and logical elementary operations.
-Second, one must find the mesh cells that the points $\{z_i\}_{i=1}^{\numnbr}$ reside.
-This is done using an AABB tree of mesh cells, and requires $O(\numnbr \log \fedim)$ elementary operations.
-Third, one must evaluate finite element basis functions on each of those cells, which requires $O(\numnbr)$ elementary operations.
-Finally, one must solve a $\numnbr \times \numnbr$ linear system to perform the radial basis function interpolation, which requires $O(\numnbr^3)$ elementary operations, which typically is negligible.
-Therefore, building the H-matrix requires
-\begin{equation*}
-	O\left(\left(\hrank^2 \fedim \log \fedim\right)\left(\numnbr \log \fedim + \numnbr^3\right)\right) \qquad \text{elementary operations}.
-\end{equation*}
-%This follows from the above discussion, and the fact that $\fedim \ge \nsamplepts$.
-
-\paragraph{(3) Performing linear algebra operations with the H-matrix} It is well known that H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion, and low rank updates require
-\begin{equation}
-\label{eq:H_matrix_cost_operations}
-	O(\hrank^2 \fedim \log(\fedim)^a) \qquad \text{elementary operations},
-\end{equation}
-where $a \in \{0,1,2,3\}$~\cite{GrasedyckHackbusch03}.
-The precise cost depends on the type of H-matrix used and the operation being performed.
-In the numerical results (Section~\ref{sec:numerical_results}), we use one matrix-matrix addition to add the H-matrix approximation of the data misfit term in the Hessian to the regularization term in the Hessian.
-Constructing an H-matrix representation of the regularization term in the Hessian requires one matrix inversion and two matrix-matrix multiplications.
-Then, we use one matrix factorization to form a preconditioner for the Newton linear system.
-%We perform low-rank updates to the H-matrix as the Newton iterations proceed, as per the discussion in Appendix~\ref{sec:recycle_krylov}.
-Symmetrizing $\mathbf{A}_H$ requires one matrix-matrix addition, and flipping negative eigenvalues to be positive requires a handful (typically around 5) matrix-matrix additions and matrix factorizations.
-%semi-definite, described in Appendix~\ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the rational function order.
-
-\section{Numerical results}
-\label{sec:numerical_results}
-
-We use our method to approximate the Newton (or Gauss-Newton) Hessian
-in inverse problems governed by PDEs which model steady state ice
-sheet flow~\cite{PetraEtAl12} (Section~\ref{sec:ice}) and transport
-and flow of a contaminant~\cite{PetraStadler11}
-(Section~\ref{sec:adv}). 
-To reconstruct the unknown parameter fields, denoted $q$, the inverse problems are formulated as nonlinear least squares
-optimization problems, whose cost functions consist of a misfit term
-(between the observations and model output) and a bi-Laplacian
-regularization term following~\cite{VillaPetraGhattas21}. The
-regularization is centered at a constant function $q_0(x)$. To
-mitigate boundary effects
-%For stokes q_0(x) = 10.5$%
-we use a constant coefficient Robin boundary condition as
-in~\cite{Roininen14}. The parameters for the regularization are chosen
-so that the Green's function of the Hessian of the regularization has
-a characteristic length of $0.25$ of the domain radius. For the
-specific setup, we refer the reader to~\cite[Section
-  2.2]{VillaPetraGhattas21}. For both examples, we choose the
-regularization parameter, $\gamma$, that controls the overall strength
-of the regularization using Morozov's discrepancy
-principle~\cite{Vogel02}. The optimal value of $\gamma$ is used for
-%  $\gamma =  7.3 \times 10^3$ for Stokes  Figure~\ref{fig:gamma_sweep}
-all the numerical results except for the studies where we study how
-changing $\gamma$ impacts the effectiveness of the proposed
-preconditioner.
-
-We solve the ice sheet inverse problem with an inexact Newton preconditioned
-conjugate gradient (PCG) scheme and a globalizing Armijo line search~\cite{NocedalWright99}. The Newton updates,
-$\bm{\searchdir}_k$, are obtained by solving 
-%
-\begin{equation}
-	\label{eq:newton_system}
-	\bm{H} \bm{\searchdir}_k = - \bm{g}_k \qquad \text{or} \qquad \bm{H}_\text{gn} \bm{\searchdir}_k = - \bm{g}_k,
-\end{equation}
-%
-wherein we choose the initial guess as the discretization of the constant
-function $q_0$. Here $\bm{g}_k$, $\bm{H}$ and $\bm{H}_\text{gn}$ are
-the discretized gradient, Hessian, and Gauss-Newton Hessian of the
-inverse problem cost function, respectively.
-%The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
-To ensure positive definiteness of the Hessian we use
-$\bm{H}_\text{gn}$ for the first five iterations, and $\bm{H}$ for all
-subsequent iterations. The Newton iterations are terminated when
-$\|\bm{g}_k\|/\|\bm{g}_0\| < 10^{-8}$. Systems~\eqref{eq:newton_system} are solved inexactly using an inner PCG
-iteration, which terminates when the norm of the linear system residual is less than $\min \left(0.5,
-\sqrt{\|\bm{g}_k\|}\right)\|\bm{g}_k\|$.
-The advection-diffusion inverse problem is linear, so Newton's method converges in one iteration.
-In this case the Newton linear system,~\eqref{eq:newton_system}, is solved using PCG with a fixed tolerance which is specified in Section~\ref{sec:adv}.
-
-We use the framework described in this paper to generate Hessian preconditioners.
-We build H-matrix approximations, $\mathbf{A}_H$, of the data misfit Gauss-Newton Hessian (the term in $\bm{H}_\text{gn}$ that arises from the data misfit).
-The
-%number of impulse response batches used to build these 
-approximations are indicated by ``PSF ($\nbatch$)'', where $\nbatch$ is the number of impulse response batches used to build the approximation.
-The Hessian of the regularization term is a combination of
-stiffness and mass matrices, which are sparse. Therefore, we form
-H-matrix representations of these matrices and combine them into a H-matrix approximation of the regularization term in the Hessian, $\mathbf{R}_H$, using
-standard sparse H-matrix techniques. H-matrix approximations of the Gauss-Newton Hessian, 
-$$\bm{H}_\text{gn}\approx \widetilde{\mathbf{H}}:=\mathbf{A}_H + \mathbf{R}_H,$$
-are formed by adding $\bm{A}_H$ to $\mathbf{R}_H$ using fast H-matrix arithmetic.
-The H-matrix $\widetilde{\mathbf{H}}$ may be indefinite due to approximation error, so we modify it to be symmetric positive definite via a shift and invert deflation procedure described in Appendix~\ref{sec:make_spd}.
-We factor $\bm{\preconditioner}$,
-using fast H-matrix methods, then use the factorization as a
-preconditioner.
-We approximate $\bm{H}_\text{gn}$ rather than $\bm{H}$ because $\bm{H}$ more often has negative values in its integral kernel.
-Our results show that $\widetilde{\mathbf{H}}$ is a good preconditioner for both $\bm{H}_\text{gn}$ and $\bm{H}$.
-%The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to an H-matrix approximation of the Hessian of the regularization term to form an H-matrix approximation of $\bm{H}_\text{gn}$.
-%
-%Straightforward calculation shows that the discretized Hessian of the regularization term is given by
-%$
-%(\gamma \bm{K} + \delta \bm{M}) \bm{M}^{-1} (\gamma \bm{K} + \delta \bm{M})
-%$
-%where $\bm{K}$ is a finite element stiffness matrix with appropriate boundary conditions, and $\bm{M}$ is the finite element mass matrix.
-These matrices are sparse, so we form H-matrix representations of these matrices and combine them using standard sparse H-matrix techniques.
-%
-
-\subsection{Example 1: Inversion for the basal sliding coefficient in an ice sheet flow problem}
-\label{sec:ice}
-
-%\textbf{Heat equation}
-%\begin{itemize}
-%	\item Hessian per-column error plots for 1, 5, and 15 batches
-%	\item $H-P$ error vs number of batches for interior and whole domain
-%	\item $P^{-1}H-I$: error vs rank for $P=R$ and $P=P_\text{pch}$ diffusion time, for 1, 5, and 15 batches
-%	\item Krylov iterations to tols 1e-1 and 1e-6 vs diffusion parameter
-%	\item GLR vs PCH vs diffusion time
-%	\item Krylov method convergence plot: $R$ vs $P_\text{pch}$ vs no preconditioner
-%	\item Krylov iterations to tols 1e-6 vs mesh size $h$, PCH vs R vs no preconditioner (hold)
-%	\item true parameter (H)
-%	\item deterministic reconstruction (H)
-%	\item noisy measurements (1\%, 5\%, 10\% noise). $u|_\text{top}$ (H)
-%	\item recovered state at top (H)
-%	\item mesh scalability of PCH (CC) (H)
-%	\item put plot data into data directory
-%	\item save function .pvd files in paraview  directory
-%\end{itemize}
-%
-%\begin{itemize}
-%	\item GLR vs PCH number of obs (Computational cost in PDE solves) (S)
-%	\item GLR vs PCH aspect ratio (CC) (S)
-%	\item turn on preconditioner after number of krylov iterations exceeds 15 in an iteration. Rebuild "as needed". (S)
-%\end{itemize}
-%
-%\begin{itemize}
-%	\item Stokes Krylov convergence (10 solves for nonlinear forward + 6 hessian matvecs for ellipsoid estimation + 10 matvecs for impulse responses + 30 matvecs for PCH6 CG solve + 100 matvecs for Reg CG solve)
-%	\item Stokes error vs num batches (small number of batches; 1, 3, 6, 9 batches?) (50 matvecs for randomized error estimation +)
-%	\item Velocity field
-%	\item basal sliding field
-%	\item Picture of impulse responses
-%	\item Mesh negative numbers picture
-%\end{itemize}
-%
-%\begin{itemize}
-%	\item mtrue
-%	\item utrue (no arrows)
-%	\item utrue (yes arrows)
-%	\item reconstruction iterations 1,3,4,8: PCH vs REG
-%	\item urecovered (arrows, 3D)
-%	\item morozov plot
-%	\item Newton convergence (iter, numcg, numPDEsolves (num preconditioner build), gradnorm)
-%	\item impulse response batches
-%	\item 
-%\end{itemize}
-
-For this example, we consider a sheet of ice flowing down a mountain (see Figure~\ref{fig:ice_mountain_mesh}).
-Given observations of the tangential component of the ice velocity on the top surface of the ice, we invert for the logarithm of the unknown spatially varying basal sliding Robin coefficient field, which governs the resistance to sliding along the base of the ice sheet.
-%The inverse problem considered here is synthetic, but in practice surface velocity observations are available from satellites.
-%
-%
-%\begin{figure}
-%	\begin{center} 
-%	\includegraphics[scale=0.23]{meshHeight_view2_edges.png}
-%	\end{center}
-%	\caption{Bird's eye view of the ice sheet, discretized by a mesh of tetrahedra.
-Color indicates the height of the base of the ice sheet (i.e., the mountain topography).
-The radius of the domain is $10^4$ meters, the maximum height of the mountain is $2.1 \times 10^3$ meters, and the average thickness of the ice sheet is 250 meters.}
-%	\label{fig:ice_mountain_mesh}
-%\end{figure} 
-%
-%
-The domain that is filled with ice is denoted by
-$\icedomain \subset\mathbb{R}^{3}$. The basal, lateral and top parts of $\partial \icedomain$ are denoted by $\Gamma_{b}$, $\Gamma_{l}$, and $\Gamma_{t}$, respectively.
-The governing equations are the linear incompressible Stokes
-equations,
-%
-%
-\begin{subequations}\label{Stokeseqn}
-	\begin{align}
-	%-\nabla \cdot \stress(v, p)
-	%&=\bodyforce \,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:1}\\
-	%\nabla\cdot \velocity
-	  %&=0\,\,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:2}\\
-          -\nabla \cdot \stress(v, p)=\bodyforce \text{ and }
-          \nabla\cdot \velocity &=0 \quad \text{ in }\icedomain, \label{Stokeseqn:1}\\
-	\stress(v, p)\normalvec
-	&=0\,\,\,\,\,\,\text{ on }\Gamma_{t}, \label{Stokeseqn:3}\\
-	\velocity \cdot\normalvec =0 \text{ and } \tangentop\left(\stress(v, p)\normalvec+\exp\left(\basalfriction\right)\velocity\right)
-	&=0\,\,\,\,\,\,\text{ on }\Gamma_{b},\label{Stokeseqn:4} \\
-	\stress(v, p)\normalvec+\stokesrobincoeff \velocity &=0\,\,\,\,\,\,\text{ on }\Gamma_{l}.\label{Stokeseqn:5}	
-	\end{align}
-\end{subequations}
-%
-%
-The solution to these equations is the pair $(\velocity, \pressure)$, where $\velocity$ is the ice flow velocity field, and $\pressure$ is the pressure field.
-Here, $\basalfriction$ is the unknown logarithmic
-basal sliding field (note that large $\basalfriction$ corresponds to high resistance to sliding) which is defined on the $2$D surface $\Gamma_b$.
-The quantity $\bodyforce$ is the body force density due to gravity, $\stokesrobincoeff=10^6$ is a Robin boundary condition constant,  $\normalvec$ is the outward normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary.
-We employ Glen's flow law~\cite{Glen1955}, $\stress(v, p)= 2\eta
-\dot{\strain}(v) -I\pressure$,
-which is a constitutive law for ice that relates the stress
-tensor, $\stress$, to the strain rate tensor,
-$\dot{\strain}(v)= \frac{1}{2}\left(
-\nabla\velocity+\nabla\velocity^{\top} \right)$.
-Here $\eta$ is the effective viscosity and $I$ is the unit matrix.
-Glen's exponent has been chosen to be one, which provides for a linear Stokes model.
-Note that while the PDE is linear, the parameter-to-solution map, $\basalfriction \mapsto (\velocity, \pressure)$, is nonlinear.
-
-The pressure, $\pressure$, is discretized with first order scalar continuous Galerkin finite elements defined on a mesh of tetrahedra.
-The velocity, $\velocity$, is discretized with second order continuous Galerkin finite elements on the same mesh.
-The parameter $\basalfriction$ is discretized with first order scalar continuous Galerkin finite elements on the mesh of triangles that results from restricting the tetrahedral mesh to the basal boundary, $\Gamma_b$.
-Note that $\Gamma_b$ is a 2D surface embedded in 3D due to the mountain topography.
-We also generate a flattened version of $\Gamma_b$, denoted by $\Omega \subset \mathbb{R}^2$, by ignoring the height coordinate.
-The parameter $\basalfriction$ is dually viewed as a function on $\Gamma_b$ for the purpose of solving the Stokes equations, and as a function on $\Omega$ for the purpose of building our Hessian approximations and defining the regularization.
-%This flattening is necessary because our operator approximation method requires adding points on this surface together to get other points on the surface, and this is only well-defined in flat space.
-The observations are generated by adding $1\%$ multiplicative
-Gaussian noise to the tangential component of the velocity field
-restricted to the top surface of the geometry. The true basal sliding
-coefficient field and the true velocity field, which is obtained by
-solving~\eqref{Stokeseqn}, are shown in Figure~\ref{fig:true_beta_u}.
-
-\begin{figure}
-	\begin{center} 
-    \begin{subfigure}{0.99\textwidth}
-    \begin{center} 
-	\includegraphics[scale=0.23]{meshHeight_view2_edges.png}
-	\end{center}
-	\caption{Ice sheet model geometry}
-	\label{fig:ice_mountain_mesh}
-    \end{subfigure} \\
-	\begin{subfigure}{0.49\textwidth}
-		\includegraphics[scale=0.22]{mtrue_withColorBar.png}
-		\caption{$\basalfriction_\text{true}$.}
-		\label{fig:true_beta}
-	\end{subfigure}
-	\begin{subfigure}{0.49\textwidth}
-		\includegraphics[scale=0.22]{trueVelocity2_glyphs.png}
-		\caption{$\velocity_\text{true}$.}
-	\end{subfigure}
-    \end{center}
-	\caption{
-		Top: bird's eye view of the ice sheet, discretized by a mesh of tetrahedra.
-Color indicates the height of the base of the ice sheet (i.e., the mountain topography).
-The radius of the domain is $10^4$ meters, the maximum height of the mountain is $2.1 \times 10^3$ meters, and the average thickness of the ice sheet is 250 meters.
-		True parameter, $\basalfriction_\text{true}$, (bottom left) and true velocity $\velocity_\text{true}$ (bottom right).
-In the bottom right plot, arrows indicate the direction of $\velocity_\text{true}$ and color indicates the magnitude of $\velocity_\text{true}$.}
-	\label{fig:true_beta_u}
-\end{figure}
-
-%aaa
-
-%% from here we summarize --------
-%The true parameter field, $\basalfriction_\text{true}$, and corresponding velocity field, $\velocity_\text{true}$, are shown in Figure~\ref{fig:true_beta_u}.
-We generate synthetic observations by restricting the tangential component of $\velocity_\text{true}$ to the top surface to get
-%\begin{equation*}
-%\observations_\text{true} = \tangentop \velocity_\text{true} |_{\Gamma_t}.
-%\end{equation*}
-%We add $1\%$ multiplicative Gaussian noise to all components of the finite element vector, $\bm{\observations}_\text{true}$, corresponding to the function $\observations_\text{true}$, which yields noisy synthetic observations, $\observations_\text{obs}$ which are used for the inversion.
-That is,
-%\begin{equation*}
-%\left(\bm{\observations}_\text{obs}\right)_j := \left(\bm{\observations}_\text{true}\right)_j + 0.01 \cdot \boldsymbol{\xi}_j \cdot \left|\left(\bm{\observations}_\text{true}\right)_j\right|,
-%\end{equation*}
-%where $j$ ranges over all entries of the vector of coefficients representing the finite element functions $\observations_\text{true}$ and $\observations_\text{obs}$, and $\boldsymbol{\xi}_j$ are independent and identically distributed (i.i.d) random numbers drawn from the standard normal distribution.
-We also define the noise function, $\xi$, to be the finite element function with $j$th vector entry $\boldsymbol{\xi}_j$.
-%
-%The reconstructed parameter field $\basalfriction$ is found as the solution to an optimization problem of the following form:
-%\begin{equation}
-%	\label{eq:deterministic_optimization_problem}
-%	\min_{\basalfriction} \quad J(\basalfriction) := J_d(\basalfriction) + J_r(\basalfriction).
-%\end{equation}
-%The first term in the objective function is
-%\begin{equation*}
-%	J_d(\basalfriction) := \frac{1}{2}\int_{\Gamma_{t}}\|\observations_\text{obs} - \tangentop\velocity\|_{2}^{2}\,dS.
-%\end{equation*}
-%We call this term the data misfit, because it measures the difference between the observed data and the predicted data based on a candidate parameter $\basalfriction$.
-Here $\velocity = \velocity(\basalfriction)$ denotes the velocity field solving $\ref{Stokeseqn}$ for the given parameter $\basalfriction$.
-The second term,
-%\begin{equation*}
-%	J_r(\basalfriction) := \frac{1}{2}\int_{\Omega}\left| \mathcal{K}(\basalfriction - \basalfriction_0)\right|^2\,dx,
-%\end{equation*}
-%is a Bilaplacian regularization term.
-Specifically, $\basalfriction_0$ is the constant function $\basalfriction_0(x)=10.5$, and $\mathcal{K}$ is the inverse of the solution operator for the following elliptic PDE:
-%\begin{align*}
-%-\gamma\Delta u +\delta\,u&=\regrhs\,\,\,\,\text{ in }\Omega,\\
-%\gamma \nabla u\cdot\normalvec+\regrobincoeff\,u&=0\,\,\,\text{ on }\partial\Omega.
-%\end{align*}
-%Recall that $\Omega$ is the 2D flattened version of the basal surface $\Gamma_b$.
-Here $\regrhs$ is a generic forcing term, and $\regrobincoeff=\sqrt{\delta \gamma}/1.42$ is a Robin boundary condition coefficient~\cite{Roininen14}, and these quantities are different from $\bodyforce$ and $\stokesrobincoeff$ in the Stokes equation,~\eqref{Stokeseqn}.
-%In all results except one, $\gamma$ is chosen so that the Morozov discrepancy principle is satisfied, i.e.,
-%\begin{equation*}
-%	\int_{\Gamma_{t}}\|\observations_\text{obs} - \tangentop\velocity\|_{2}^{2}\,dS = \int_{\Gamma_{t}}\|\xi\|_{2}^{2}\,dS.
-%\end{equation*}
-%The value of $\gamma$ satisfying the Morozov discrepancy principle is $\gamma=7.3 \times 10^3$.
-The exception is Figure~\ref{fig:gamma_sweep}, where we vary $\gamma$ to study how changing $\gamma$ impacts the effectiveness of our preconditioner.
-The constant $\delta$ is chosen so that the correlation length of functions drawn from the normal distribution with covariance $\mathcal{K}^{-2}$, given by $L=\sqrt{\gamma/\delta}$, is $1/10$th the radius of the domain.
-%%, which is halfway between the maximum and minimum values of $\basalfriction_\text{true}$.
-%
-%PCG only requires applying
-%$\bm{H}$ (or $\bm{H}_\text{gn}$) to vectors. Each application of
-%$\bm{H}$ (or $\bm{H}_\text{gn}$) to a vector requires solving one
-%incremental forward and one incremental adjoint Stokes PDE of the form
-%\eqref{Stokeseqn}, but with different right hand sides.
-\begin{figure}
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[scale=0.13]{mstar_0.25noise_cropped.png}
-		%		\caption{}
-		%		\label{fig:newtoniter1}
-	\end{subfigure}
-%	\begin{subfigure}{0.18\textwidth}
-%		\includegraphics[scale=0.075]{mstar_0.12noise_cropped.png}
-%		%		\caption{}
-%		%		\label{fig:newtoniter2}
-%	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[scale=0.13]{mstar_0.05noise_cropped.png}
-		%		\caption{}
-		%		\label{fig:newtoniter3}
-	\end{subfigure}
-%	\begin{subfigure}{0.18\textwidth}
-%		\includegraphics[scale=0.075]{mstar_0.02noise_cropped.png}
-%		%		\caption{}
-%		%		\label{fig:newtoniter4}
-%	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.13]{mstar_0.01noise_cropped.png}
-	%		\caption{}
-	%		\label{fig:newtoniter4}
-    \end{subfigure}
-	\caption{The log basal sliding parameter, with colorscale as in Figure~\ref{fig:true_beta}, computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $5.0\%$ (middle),and $1.0\%$ (right).
-\todo{remove reference to newon iterates figure as this figure has a new significance}}
-	\label{fig:newton_iterates} 
-\end{figure}
-
-\begin{table}
-	{\small
-		\begin{center}
-			\begingroup
-			\setlength{\tabcolsep}{3pt}
-			\renewcommand{\arraystretch}{1.1}
-			\begin{tabular}{c| c c c | c c c | c c c}
-				&  \multicolumn{3}{|c|}{PSF (5)} & \multicolumn{3}{|c|}{REG} & \multicolumn{3}{|c}{NONE} \\
-				\hline
-				Iter & 
-				\#CG & \#Stokes & $\|\bm{g}\|$ & 
-				\#CG & \#Stokes & $\|\bm{g}\|$ & 
-				\#CG & \#Stokes & $\|\bm{g}\|$ \\
-				0 &
-				1 & 4 & 1.9e+7 &
-				3 & 8 & 1.9e+7 &
-				1 & 4 & 1.9e+7 \\
-				1 &
-				2 & 6  & 6.1e+6 &
-				8 & 18 & 8.4e+6 &
-				2 & 6  & 6.1e+6 \\
-				2 &
-				3 & 8 & 2.6e+6 &
-				16 & 34 & 4.1e+6 &
-				3 & 8 & 2.6e+6 \\
-				3 &
-				1 & 4+22 & 9.2e+5 &
-				34 & 70 & 1.7e+6 &
-				11 & 24 & 9.2e+5 \\
-				4 &
-				2 & 6 & 1.5e+5 &
-				54 & 110 & 5.0e+5 &
-				30 & 62 & 1.4e+5 \\
-				5 &
-				4 & 10 & 7.1e+3 &
-				79 & 160 & 8.0e+4 &
-				34 & 70 & 1.4e+4 \\
-				6 &
-				5 & 12 & 6.0e+1 &
-				112 & 226 & 5.2e+3 &
-				61 & 124 & 2.3e+2 \\
-				7 &
-				0 & 2 & 8.9e-2 &
-				162 & 326 & 8.0e+1 &
-				84 & 170 & 7.3e-1 \\
-				8 & 
-				--- & --- & --- &
-				0 & 2 & 1.6e-1 &
-				0 & 2 & 3.0e-2 \\
-				\hline
-				Total & 
-				18 & 74 & --- &
-				468 & 954 & --- &
-				226 & 470 & --- \\
-			\end{tabular}
-			\endgroup
-		\end{center}
-	}
-	\caption{Convergence history for solving the Stokes inverse problem using preconditioned inexact Newton PCG.
-Preconditioners shown are our method with five batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
-Columns titled \#CG show the number of PCG iterations used to solve the Newton system for $\bm{\widehat{\basalfriction}}_k$.
-Columns titled $\|\bm{g}\|$ show the $l^2$ norm of the gradient at $\bm{\basalfriction}_k$.
-Columns titled \#Stokes show the total number of Stokes PDE solves performed in each Newton iteration.
-This consists of Stokes solves for $u$ performed during the linesearch going from $\bm{\basalfriction}_{k-1}$ to $\bm{\basalfriction}_k$, plus one adjoint Stokes solve to compute the gradient at $\bm{\basalfriction}_k$, plus one incremental forward and one incremental adjoint Stokes solve per PCG iteration for solving the Newton system.
-Under PSF (5) and in row Iter 3, we write $4+22$ to indicate that $4$ Stokes solves were used during the standard course of the iteration, and $22$ Stokes solves were used to build the PSF (5) preconditioner.}
-	\label{tab:newton_convergence_table}
-\end{table}
-
-Table~\ref{tab:newton_convergence_table} shows the performance of our preconditioner for accelerating the solution of optimization problem~\eqref{eq:deterministic_optimization_problem}.
-We build the PSF (5) preconditioner in the third Gauss-Newton iteration, and reuse it for all subsequent Gauss-Newton and Newton iterations.
-No preconditioning is used in the Gauss-Newton iterations before the PSF (5) preconditioner is built.
-We compare our method with the most commonly used existing preconditioners: no preconditioning (NONE), and preconditioning by the regularization term in the Hessian (REG).
-Using PSF (5) reduces the total number of Stokes PDE solves by a factor of six compared to no preconditioning, and by a factor of nine compared to regularization preconditioning\footnote{Interestingly, we observe that no preconditioning outperforms regularization preconditioning here because the noise level (and hence $\gamma$) is small.}.
-In Figure~\ref{fig:newton_iterates} we show select Newton iterates using our PSF (5) preconditioner.
-%
-%
-\begin{figure}
-	\begin{subfigure}{0.49\textwidth}
-		\centering
-		\begin{tikzpicture}[baseline, trim axis right]
-		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
-		ymode=log,
-		ymin=1e-13,
-		ymax=2,
-		xmax=750,
-		axis x line*=bottom,
-		axis y line*=left,
-		xtick={0, 200, 400, 600},
-		title={CG convergence},
-		xlabel={$k$, CG iteration},
-		ylabel={$\|\bm{x}_k-\bm{x}\|/\|\bm{x}\|$},
-		legend style={at={(1.1,1)},anchor=north east},
-		%legend pos=north east,
-		legend style={font=\small}
-		]
-		\addlegendentry{Reg}		
-		\addplot[dotted, line width=0.5mm] table {cg_relres_reg_0.05noise.dat};
-		\addlegendentry{None}
-		\addplot[dash dot, line width=0.5mm] table {cg_relres_none_0.05noise.dat};
-		\addlegendentry{PSF (1)}
-		\addplot[blue, line width=0.5mm] table {cg_relres_psf1_0.05noise.dat};
-		\addlegendentry{PSF (5)}
-		\addplot[red, line width=0.5mm] table {cg_relres_psf5_0.05noise.dat};
-		\addlegendentry{PSF (25)}
-		\addplot[purple, line width=0.5mm] table {cg_relres_psf25_0.05noise.dat};
-		\end{axis}
-		\end{tikzpicture}
-		\caption{}
-		\label{fig:cg_convergence}
-	\end{subfigure}
-	\begin{subfigure}{0.49\textwidth}
-	\begin{center}
-		\begin{tikzpicture}[baseline, trim axis right]
-		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
-		ymode=log,
-		xmin=-50,
-		xmax=1450,
-		ymax=1.e7,
-		title={Generalized eigenvalues},
-		xlabel={$k$, generalized eigenvalue \#},
-		ylabel={$\lambda_{k}$},
-		xtick={0, 300, 600, 900, 1200},
-		%legend pos=north east,
-		legend style={font=\small},
-		ylabel near ticks, ylabel shift={-6pt},
-		axis x line*=bottom,
-		axis y line*=right,
-		%legend style={at={(0.95,1)},anchor=north east},
-		]
-		%\addlegendentry{REG}
-		\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues_0.05noise.dat};
-		\addplot[dash dot, line width=0.5mm] table[x=k, y=none] {generalizedEigenvalues_0.05noise.dat};
-		%\addlegendentry{PSF (1)}
-		\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues_0.05noise.dat};
-		%\addlegendentry{PSF (5)}
-		\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues_0.05noise.dat};
-		%\addlegendentry{PSF (25)}
-		\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues_0.05noise.dat};
-		\end{axis}
-		\end{tikzpicture}
-	\end{center}
-	\caption{}
-	\label{fig:generalized_eigenvalues}
-    \end{subfigure}
-	\caption{Left (\ref{fig:cg_convergence}) shows the convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
-random entries drawn from the standard normal distribution.
-The reference solution $\bm{x}$ is determined by solving $\bm{H} \bm{x} = \bm{b}$ to an accuracy of $10^{-12}$ using method PCG.
-	Here we use $\gamma=7.3 \times 10^3$, which is determined by the Morozov discrepancy principle when using $5\%$ noise.
-Results in these figures are shown for our preconditioner with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
-The preconditioner is constructed using $\bm{H}_\text{gn}$.
-	Right (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$.
-Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), or the regularization Hessian (REG).}
-	\label{fig:krylov_convergence}
-\end{figure} 
-%
-%
-We build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\bm{\basalfriction}$.
-We use PCG to solve a linear system with the Hessian as the coefficient operator and a right hand vector with random i.i.d.
-entries drawn from the standard normal distribution.
-In Figure~\ref{fig:cg_convergence} we compare the convergence of PCG for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners.
-PCG converges fastest with the PSF preconditioners, with PSF (25) converging fastest, followed by PSF (5), followed by PSF (1), as expected.
-PCG converges much slower with no preconditioning and regularization preconditioning than it does with PSF preconditioning, with no preconditioning outperforming regularization preconditioning.
-In Figure~\ref{fig:gamma_sweep}, we perform PCG solves on linear systems of the same form, except now we vary $\gamma$.
-The performance of our PSF preconditioners is good and relatively stable over a wide range of $\gamma$'s.
-All PSF preconditioners perform the same for medium and large values of $\gamma$.
-For small values of $\gamma$ PSF (25) performs slightly better than PSF (5), which performs slightly better than PSF (1).
-As expected, regularization preconditioning performs well for large $\gamma$ and poorly for small $\gamma$.
-Our PSF preconditioners outperform no preconditioning and regularization preconditioning for medium and small $\gamma$, and perform similarly to regularization preconditioning for large~$\gamma$.
-
-\begin{table}
-	\begin{center}
-			\begingroup
-			\setlength{\tabcolsep}{4pt}
-			\renewcommand{\arraystretch}{1.25}
-			\begin{tabular}{c| c c c c c}
-				noise    & \multicolumn{5}{c}{COND$(\bm{\preconditioner}^{-1} \bm{H})$ } \\ \cline{2-6}
-				level    & REG     &	NONE  & PSF $(1)$ & PSF $(5)$ & PSF $(25)$ \\ \hline 
-				$25\%$   & 1.68e+3 & 2.16e+3  & 1.37e+0   & 1.30e+0   & 1.21e+0    \\ 
-				$11\%$   & 9.19e+3 & 9.08e+2  & 1.85e+0   & 1.49e+0   & 1.27e+0    \\   
-				$5.0\%$  & 3.66e+4 & 4.67e+2  & 3.46e+0   & 2.62e+0   & 1.91e+0    \\ 
-				$2.2\%$  & 1.41e+5 & 8.25e+2  & 1.16e+1   & 4.99e+0   & 3.20e+0    \\  
-				$1.0\%$  & 6.06e+5 & 1.70e+3  & 3.44e+1   & 5.38e+1   & 7.06e+0    \\   
-			\end{tabular}
-			\endgroup
-		\end{center}
-	\caption{Condition number for $\bm{\preconditioner}^{-1} \bm{H}$ for our preconditioner with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), no preconditioner (NONE) and regularization (REG).
-All operators are evaluated at the $\bm{\basalfriction}$ that solves the inverse problem for
-	the regularization value $\gamma$ as determined by the respective noise level and the Morozov discrepancy principle.}
-	\label{fig:condition_number}
-\end{table}
-
-In Figure~\ref{fig:generalized_eigenvalues} we show the generalized eigenvalues for the generalized eigenvalue problem
-$
-	\bm{H} \mathbf{u} = \lambda \bm{\preconditioner} \mathbf{u}.
-$
-Here $\bm{H}$ is the Hessian evaluated at the reconstructed parameter $\bm{\basalfriction}$ for $\gamma$ chosen to satisfy the Morozov discrepancy principle.
-The matrix $\bm{\preconditioner}$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations constructed at that $\bm{\basalfriction}$, or the regularization Hessian.
-The PSF preconditioners cluster the eigenvalues of the Hessian near one, with more batches yielding better clustering.
-The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
-In Figure~\ref{fig:condition_number} we show the condition numbers of $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$.
-The PSF preconditioners yield the the smallest condition numbers, with more batches yielding smaller condition numbers.
-Regularization preconditioning and no preconditioning yield larger condition numbers.
-The condition numbers are similar for both $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$, demonstrating that the preconditioner built based on $\bm{H}_\text{gn}$ is a good preconditioner for $\bm{H}$.
-
-\subsection{Example 2: Inversion for the initial condition in an advection-diffusion problem}
-\label{sec:adv}
-
-%        |   0.5   |   1.0   |   2.0   
-% -------------------------------------
-% 1.0e-4 |   OK    |   OK    |   OK
-% 1.8e-4 |   OK    |   OK    |   OK
-% 3.2e-4 |   OK    |   OK    |   OK
-% 5.6e-4 |   OK    |   OK    |   5
-% 1.0e-3 |   OK    |   OK    |   5
-%
-
-Here we consider a time-dependent advection-diffusion equation for
-which we seek to infer the unknown spatially varying initial condition, $\ipar$, from observations of the state  
-at a final time. This PDE
-models diffusive transport in a domain $\advdomain \subset \R^d$, which is
-depicted in Figure~\ref{fig:adv_inversion1}. In this case, the state, $\concentration(x,t)$, could be interpreted as the concentration of a contaminant. The problem description
-below closely follows the one in~\cite{PetraStadler11}. The domain
-boundaries $\partial \advdomain$ include the outer boundaries as well as the
-internal boundaries of the rectangles, which represent buildings.  The
-parameter-to-observable map $\iFF$ in this case maps an initial
-condition $\ipar \in L^2(\advdomain)$ to observations of the concentration at a final time, $\concentration(x, T)$,  through solution of the advection-diffusion
-equation given by
-\begin{equation}\label{eq:ad}
-  \begin{aligned}
-    \concentration_t - \kappa\Delta \concentration + v\cdot\nabla \concentration &= 0 & \quad&\text{in
-    }\advdomain\times (0,T), \\
-    \concentration(\cdot, 0) &= \ipar  &&\text{in } \advdomain , \\
-    \kappa\nabla \concentration\cdot \nu &= 0 &&\text{on } \partial \advdomain \times (0,T).
-  \end{aligned}
-\end{equation}
-Here, $\kappa > 0$ is a diffusivity coefficient, $\nu$ is the boundary unit normal vector, and $T > 0$ is the
-final time.  The velocity field $v:\advdomain \rightarrow \mathbb{R}^d$, shown in
-Figure~\ref{fig:adv_inversion1}, is computed by solving the
-steady-state Navier-Stokes equations for a two dimensional flow with
-Reynolds number 50 and boundary conditions as in
-Figure~\ref{fig:adv_inversion} (left); see~\cite{PetraStadler11} for
-details.  The time evolution of the state variable $\concentration$ from a given
-initial condition $\ipar$ is illustrated in
-Figure~\ref{fig:adv_inversion1}.
-
-%INTRODUCE EQUATIONS: REWRITE HERE vvvvvvv
-
-%The parameter-to-observable map $\iFF \,\ipar := \obsop\, u(\ipar)$ maps an initial condition $\ipar \in L^2(\advdomain)$ to pointwise spatiotemporal observations of the concentration field $u({\vec x},t)$ through solution of the advection-diffusion equation given by
-%
-%$$
-%\begin{split}
-%u_t - \kappa\Delta u + {\vec v} \cdot \nabla u &= 0     & \quad \text{in } \advdomain\times(0,T),\\
-%u(\cdot, 0) &= \ipar & \quad \text{in } \advdomain,\\
-%\kappa \nabla u\cdot {\vec{n}} &= 0     & \quad \text{on } \partial \advdomain \times (0,T).
-%\end{split}
-%$$
-%
-%Here, $\advdomain \subset \R^d$ ($d \in \{2, 3\}$) is a bounded domain, $\kappa > 0$ is the diffusion coefficient and $T > 0$ is the final
-%time. The velocity field
-%$\vec{v}$ is computed by solving the following steady-state
-%Navier-Stokes equation with the side walls driving the flow:
-%
-%$$
-%\begin{aligned}
-%- \frac{1}{\operatorname{Re}} \Delta {\vec v} + \nabla q + {\vec v} \cdot \nabla {\vec v} &= 0 &\quad&\text{ in }\advdomain,\\
-%\nabla \cdot {\vec v} &= 0 &&\text{ in }\advdomain,\\
-%{\vec v} &= {\vec g} &&\text{ on } \partial \advdomain.
-%\end{aligned}
-%$$
-%
-%Here, $q$ is pressure, $\text{Re}$ is the Reynolds number. The Dirichlet boundary data
-%${\vec g} \in \R^d$ is given by 
-%${\vec g} = {\vec e}_2$ on the left wall of the domain, 
-%${\vec g}=-{\vec e}_2$ on the right wall,  and ${\vec g} = {\vec 0}$ everywhere else.
-
-% INTRODUCE EQUATIONS: REWRITE HERE ^^^^^
-
-\begin{figure}
-	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.14]{true_initial_condition.png}
-		%		\caption{}
-		%		\label{fig:newtoniter2}
-	\end{subfigure}
-	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.14]{wind_velocity_colorscheme3.png}
-		%		\caption{}
-		%		\label{fig:newtoniter1}
-	\end{subfigure}
-	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.14]{noisy_observations_T1.0_K3.2e-4.png}
-		%		\caption{}
-		%		\label{fig:newtoniter3}
-	\end{subfigure}
-	\begin{subfigure}{0.24\textwidth}
-		\includegraphics[scale=0.14]{reconstructed_initial_condition_T1.0_K3.2e-4.png}
-		%		\caption{}
-		%		\label{fig:newtoniter4}
-	\end{subfigure}
-	%\begin{subfigure}{0.05\textwidth}
-	%	\includegraphics[scale=0.1]{mColorBar.png}
-	%\end{subfigure}
-	\caption{(Left) True initial condition.
-(Middle left) Velocity field.
-Color indicates magnitude of velocity vector.
-(Middle right) Noisy observations of concentration at final time.
-(Right) Reconstructed initial condition.}
-	\label{fig:adv_inversion1} 
-\end{figure}
-
-\begin{table}
-	\begin{center}
-		\begingroup
-		\setlength{\tabcolsep}{8pt}
-%		\renewcommand{\arraystretch}{1.25}
-		\renewcommand{\arraystretch}{1.0}
-		\begin{tabular}{c| c | c c c c c}
-			& $\kappa$    & REG     &	NONE   & PSF $(1)$ & PSF $(5)$ & PSF $(25)$ \\ 
-			\hline 
-			& 1.0e-4    & 584 & 317  & 311   & 151   & 56    \\ 
-%			& $1.8e{-4}$    & 628 & 313  & 192   & 131   & 49    \\   
-	$T=0.5$ & 3.2e-4    & 685 & 311  & 233   & 140   & 44    \\ 
-%			& $5.6e{-4}$    & 734 & 310  & 334   & 144   & 40    \\  
-			& 1.0e-3    & 702 & 324  & 122   & 71   & 33    \\ 
-			\hline
-			& 1.0e-4    & 634 & 449  & 539   & 288   & 100    \\ 
-%			& $1.8e{-4}$    & 665 & 449  & 493   & 212   & 84    \\   
-	$T=1.0$ & 3.2e-4    & 681 & 459  & 350   & 202   & 90    \\ 
-%			& $5.6e{-4}$    & 619 & 482  & 262   & 197   & 70    \\  
-			& 1.0e-3    & 574 & 520  & 266   & 260   & 208    \\ 
-			\hline
-			& 1.0e-4    & 609 & 591  & 548   & 520   & 165    \\ 
-%			& 1.8e{-4}$    & 576 & 611  & 434   & 426   & 185    \\   
-	$T=2.0$ & 3.2e-4    & 524 & 645  & 318   & 379   & 170    \\ 
-%			& $5.6e{-4}$    & 464 & 711  & 480   & 289   & 153    \\  
-			& 1.0e-3    & 349 & 786  & 381   & 262   & 158    \\ 
-		\end{tabular}
-		\endgroup
-	\end{center}
-	\caption{Number of preconditioned conjugate gradient iterations required to achieve relative error $10^{-6}$ for solving the Newton linear system in the advection-diffusion inverse problem.
-Iteration counts are shown for a variety of different diffusion parameters $\kappa$, simulation times $T$, and preconditioners.}
-	\label{tab:adv_convergence}
-\end{table}
-
-\begin{table}
-	\centering
-	\begin{tabular}{cccc}
-		  & $T=0.5$ & $T=1.0$ & $T=2.0$ \\
-		$\kappa=1.0e{-4}$  & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-4.png} \\
-		$\kappa=3.2e{-4}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K3.2e-4.png} \\
-		$\kappa=1.0e{-3}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-3.png}
-	\end{tabular}
-	\caption{Impulse responses for a variety of different diffusion parameters $\kappa$ and simulation times $T$.}
-	\label{tbl:table_of_figures}
-\end{table}
-
-\begin{figure}
-	\caption{Impulse responses for various diffusivities kappa and fixed simulation time Tf=0.5 (top) and simulation times with fixed kappa=1e-3 (bot).}
-	\label{fig:adv_vary_kappa_tf}
-\end{figure}
-
-\begin{figure}
-	\begin{subfigure}{0.49\textwidth}
-		\centering
-		\begin{tikzpicture}[baseline, trim axis right]
-		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
+	\begin{tikzpicture}[baseline]%, trim axis right]
+		\begin{axis}[width=0.49\linewidth, height=0.5\linewidth,
 		ymode=log,
@@
 		title={CG convergence},
-		legend style={at={(1.2,1)},anchor=north east},
+		legend style={at={(1.2,1)},anchor=north east,draw=none},
 		%legend pos=north east,
@@
 	\end{tikzpicture}
-		\caption{}
-		\label{fig:adv_cg_convergence}
-	\end{subfigure}
-	\begin{subfigure}{0.49\textwidth}
-		\centering
-		\begin{tikzpicture}[baseline, trim axis right]
-\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
-ymode=log,
-xmin=-50,
-xmax=600,
-ymax=1.e6,
-title={Generalized eigenvalues},
-xlabel={$k$, generalized eigenvalue \#},
-ylabel={$\lambda_{k}$},
-xtick={0, 300, 600, 900, 1200},
-%legend pos=north east,
-legend style={font=\small},
-ylabel near ticks, ylabel shift={-6pt},
-axis x line*=bottom,
-axis y line*=right,
-%legend style={at={(0.95,1)},anchor=north east},
-]
-%\addlegendentry{REG}
-\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {advDiff_generalizedEigenvalues.dat};
-\addplot[dash dot, line width=0.5mm] table[x=k, y=none] {advDiff_generalizedEigenvalues.dat};
-%\addlegendentry{PSF (1)}
-\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {advDiff_generalizedEigenvalues.dat};
-%\addlegendentry{PSF (5)}
-\addplot[red, line width=0.5mm] table[x=k, y=psf5] {advDiff_generalizedEigenvalues.dat};
-%\addlegendentry{PSF (25)}
-\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {advDiff_generalizedEigenvalues.dat};
-\end{axis}
-\end{tikzpicture}
-		\caption{}
-		\label{fig:adv_generalized_eigenvalues}
-	\end{subfigure}
-	\caption{Left (\ref{fig:adv_cg_convergence}) shows the convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
-random entries drawn from the standard normal distribution.
-BLAH BLAH BLAH
-		%		The y-axis is $\|b-Hx_k\|/\|b\|$, where $x_k$ is the $k$th PCG iterate. 
-		Here we use $\gamma=7.88 \times 10^{-6}$, which is determined by the Morozov discrepancy principle.
-Results in these figures are shown for our preconditioner with 5, 15, and 30 batches (PSF (5), PSF (15), and PSF(30), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
-The preconditioner is constructed using $\bm{H}_\text{gn}$.}
-	\label{fig:adv_krylov_convergence}
-\end{figure}
-
-\begin{itemize}
-	\item Picture of velocity field, initial condition, final condition, reconstruction (1) (OKOK)
-	\item Impulse responses: vary time and Peclet number (3) (OKOK)
-	\item CG convergence (1) CG iters vs Peclet number (Morozov) (2) (OKOK)
-	\item Preconditioned spectrum, condition numbers? (1)
-\end{itemize}
-
-\section{Conclusions}
-\label{sec:conclusions}
-
-We presented an efficient matrix-free PSF method for approximating operators with locally supported non-negative integral kernels.
-The method only requires access to the operator via application of the operator to a small number of vectors.
-The idea of the method is to compute batches of impulse responses by applying the operator to Dirac combs of scattered point sources, then interpolate these impulse responses to approximate entries of the operator's integral kernel.
-The interpolation is based on a new principle we call ``local mean displacement invariance,'' which generalizes and improves upon classical local translation invariance.
-The ability to quickly approximate arbitrary integral kernel entries permits us to form an H-matrix approximation of the operator.
-Fast H-matrix arithmetic is then used to perform further linear algebra operations that cannot be performed easily with the original operator, such as matrix factorization and inversion.
-The supports of the impulse responses are estimated to be contained in ellipsoids, which are determined a-priori via a new procedure that involves applying the operator to a small number of polynomial functions.
-Point source locations for the impulse response batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulse responses per batch as possible, while ensuring that the corresponding ellipsoids do not overlap.
-We applied the method to approximate the Gauss-Newton Hessian in a Stokes ice sheet inverse problem, and saw that the approximation substantially outperforms existing Hessian approximation methods.
-Although the method we presented is not applicable to all Hessians, it is applicable to many Hessians of practical interest.
-For these Hessians, our method offers a \emph{data scalable} alternative to conventional low rank approximation, because our method can form high rank approximations of an operator using a small number of operator applies, and thus forward and adjoint PDE solves.
-
-\appendix
-
-%\section{Hierarchical matrix details}
-%\label{app:h_matrix}
-%
-%Often, large dense matrices of practical interest may be permuted, then partitioned into blocks recursively, in such a way that many off-diagonal blocks of the matrix are numerically low rank, even if the matrix is high rank.
-Such matrices are known as hierarchical matrices (H-matrices).
-Many classes of H-matrices exist (H1, H2, HSS, HBS, and more), and all of these types of H-matrices could be used in conjunction with our product-convolution approximation.
-Here we use classical H1 matrices.
-For this section, when we say H-matrix, we are referring to H1 matrices.
-%
-%While a dense $\fedim \times \fedim$ matrix traditionally requires $O(\fedim^2)$ memory to store, H-matrices may be stored using $O(\fedim \log \fedim)$ memory, by storing only the low rank factors for the low rank blocks.
-Recursive algorithms can take advantage of the H-matrix low rank block structure to perform matrix arithmetic fast.
-Conventional dense matrix algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix-matrix addition require either $O(\fedim^2)$ or $O(\fedim^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(\fedim \log(\fedim)^a)$ time and memory for H-matrices.
-Here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix.
-For more details on H-matrices, we recommend~\cite{HMATRIXGOOD}.
-%
-%
-%\subsection{H-matrix construction}
-%\label{sec:H_matrix_construction}
-%
-%The process of constructing an H-matrix representation of $\mathbf{\Aker}$ proceeds as follows.
-First, we construct hierarchical partitionings of the degrees of freedom for the columns and rows of the matrix (cluster trees, Section~\ref{sec:cluster_trees}).
-Second, we construct a hierarchical partitioning of the blocks of the matrix, in such a way that many of the blocks in the partitioning are expected to be low rank, and the remaining high rank blocks are small (block cluster tree, Section~\ref{sec:block_cluster_tree}).
-Finally, we form low rank approximations of the blocks of the matrix that are expected to be low rank (adaptive cross approximation, Section~\ref{sec:adaptive_cross}), and fill in the remaining high rank  blocks with their numerical values.
-The first two steps require geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix, but these steps do not depend on the particular values of matrix entries.
-The third step requires us to evaluate $O(\fedim \log \fedim)$ specially-chosen entries of $\mathbf{\Aker}$, and we evaluate these entries using~\eqref{eq:Akerpcmat_entries}.
-%
-%
-%\subsubsection{Cluster trees}
-%\label{sec:cluster_trees}
-%
-%We use recursive hyperplane splitting to hierarchically cluster the degrees of freedom associated with the columns and rows of the matrix into a \emph{column cluster tree} and a \emph{row cluster tree}, respectively.
-Here we describe construction of the column cluster tree; the row cluster tree is constructed similarly.
-%
-%Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\mathbf{\Aker}$ corresponds to the Lagrange node in $\mathbb{R}^\gdim$ associated with the $i\text{th}$ finite element basis function.
-The columns of the matrix thus correspond to a point cloud in $\mathbb{R}^\gdim$.
-We split this point cloud into two equally sized \emph{child} point clouds, using a hyperplane which is perpendicular to the coordinate axis direction in which the point cloud is widest (e.g., either the $x$, $y$, or $z$ axis in $3$D).
-The two child point clouds are split in the same way.
-This splitting process repeats until the point clouds have less than a preset number of points (we use $32$ points).
-This hierarchical partitioning of the point cloud into smaller and smaller point clouds corresponds to a hierarchical partitioning of the columns of the matrix into smaller and smaller \emph{clusters} of columns.
-This hierarchical partitioning of the columns forms a binary tree, which is called the column cluster tree.
-The root of the tree is the set of all columns, and the leaves of the tree are clusters of columns that are not subdivided any further.
-%
-%A depth-first search ordering of the column cluster tree leaves is then generated.
-When the columns of the matrix are permuted into this depth-first ordering, the columns associated with each cluster in the cluster tree are contiguous.
-%
-%In the same way, the degrees of freedom associated with the rows of the matrix are hierarchically clustered into another cluster tree, and a depth-first search ordering for the rows is generated.
-In our examples, the degrees of freedom for the columns coincide with the degrees of freedom for the rows, so the cluster trees for the rows and columns are the same, but this is not required in general.
-%
-%
-%\subsubsection{Block cluster tree} 
-%\label{sec:block_cluster_tree}
-%
-%We partition the matrix into a recursive hierarchy of mostly low rank blocks called the \emph{block cluster tree}.
-The idea is that a block of the matrix is likely to be low rank if the point cloud associated with the rows of the block is far away from the point cloud associated with the columns of the block.
-This is reasonable to expect here because of the locality property of $\Aop$.
-Indeed, locality implies that many blocks of the matrix corresponding to far away point cloud clusters will be rank zero.
-%
-%After reordering the rows and columns of $\mathbf{\Aker}$ via the depth-first ordering described above, we partition the reordered version of $\mathbf{\Aker}$ into a tree of $2 \times 2$ block matrices recursively.
-We use a geometric admissibility condition (discussed below) to decide which blocks to subdivide, and use the cluster trees for the rows and columns to determine how to subdivide those blocks.
-For the first stage of subdivision, let $r_1$ and $r_2$ be the children row clusters for the root of the row cluster tree, let $\colcluster_1$ and $\colcluster_2$ be the children column clusters for the root of the column cluster tree.
-The matrix $\mathbf{\Aker}$ is partitioned into blocks as follows:
-%\begin{equation*}
-%	\begin{bmatrix}
-%		\mathbf{\Aker}_{11} & \mathbf{\Aker}_{12} \\
-%		\mathbf{\Aker}_{21} & \mathbf{\Aker}_{22}
-%	\end{bmatrix},
-%\end{equation*}
-%where $\mathbf{\Aker}_{11}$ denotes the block of $\mathbf{\Aker}$ with rows $r_1$ and columns $\colcluster_1$, $\mathbf{\Aker}_{12}$ denotes the block of $\mathbf{\Aker}$ with rows $r_1$ and columns $\colcluster_2$, and so on for $\mathbf{\Aker}_{21}$ and $\mathbf{\Aker}_{22}$.
-%
-%We now loop through the four blocks, $\mathbf{\Aker}_{11}$, $\mathbf{\Aker}_{12}$, $\mathbf{\Aker}_{21}$, and $\mathbf{\Aker}_{22}$, and decide which blocks should be subdivided further.
-For the purpose of explanation, consider $\mathbf{\Aker}_{12}$.
-If
-%\begin{equation}
-%	\label{eq:weak_admissibility_cond}
-%	\dist\left(r_1, \colcluster_2\right) \ge \weakadmconst \min\left(\diam\left(r_1\right), \diam\left(\colcluster_2\right)\right),
-%\end{equation}
-%then we mark $\mathbf{\Aker}_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone.
-Here $\dist\left(r_1, \colcluster_2\right)$ is the Euclidean distance between the axis-aligned bounding box for the point cloud associated with the row cluster $r_1$, and the axis aligned bounding box for the point cloud associated with the column cluster $\colcluster_2$.
-The quantity $\diam\left(r_1\right)$ is the diameter of the axis aligned bounding box for the point cloud associated with the row cluster $r_1$, and $\diam\left(\colcluster_2\right)$ is the analogous diameter associated with the column cluster $\colcluster_2$.
-Here the quantity $\weakadmconst$ is a scalar constant; we use $\weakadmconst=2.0$.
-Basically, if the point clouds associated with $r_1$ and $\colcluster_2$ are far away from each other relative to their diameters, then we expect that the corresponding block of the matrix will be low rank.
-This process is repeated for the other blocks to determine which blocks are admissible and which are not.
-For us, the diagonal blocks $\mathbf{\Aker}_{11}$ and $\mathbf{\Aker}_{22}$ are not admissible because the distance between a point cloud and itself is zero.
-%
-%Next, we subdivide all blocks that are not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\mathbf{\Aker}$.
-But now we subdivide a block based on the two child row clusters and two child column clusters for the rows and columns of that block.
-This subdivision process continues recursively until all blocks are either admissible, or smaller than the predetermined size mentioned above.
-The resulting hierarchical partitioning of matrix blocks forms a tree, which is called the \emph{block cluster tree}.
-The root of the tree is the whole matrix, internal nodes in the tree are blocks that are subdivided, and the leaves of the tree are blocks that are either expected to be low rank, or are small.
-%
-%
-%\subsubsection{Adaptive cross low rank approximation of blocks}
-%\label{sec:adaptive_cross}
-%
-%Once the block cluster tree has been constructed, low rank approximations of the admissible (low rank) blocks are formed using the adaptive cross method~\cite{ACA}.
-Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\mathbf{\Aker}$.
-The idea of the adaptive cross method is to form a low rank approximation of $X$ by sampling a small number of rows and columns of $X$.
-%\begin{itemize}
-%	\item Let $C \in \mathbb{R}^{m \times \hrank}$ be a matrix consisting of a subset of $\hrank$ columns of $X$, such that the span of the columns in $C$ approximates the column space of $X$. 
-%	\item Let $R\in\mathbb{R}^{\hrank \times m}$ be a subset of the rows of $X$, such that the span of the rows in $R$ approximates the row space of $X$.
-%	\item Let $U \in \mathbb{R}^{\hrank \times \hrank}$ be the block of $X$ consisting of the intersection of the rows from $R$ with the columns from $C$.
-%\end{itemize}
-%Then it is well-established that
-%\begin{equation}
-%	\label{eq:CUR}
-%	X \approx C U^+ R,
-%\end{equation}
-%where $U^+$ is the pseudoinverse of $U$~\cite{goreinov1997theory,MahoneyDrineas09}.
-The quality of approximation~\eqref{eq:CUR} depends on how well the columns of $C$ approximate the column space of $X$, and how well the rows of $R$ approximate the row space of $X$.
-%
-%In the adaptive cross method, ``good'' columns, $C$, and rows, $R$, are chosen via an alternating iterative process.
-An initial set of columns $C$ is chosen.
-Keeping $C$ fixed, a set of rows $R$ is chosen so that the so that determinant of the submatrix $U$ within $C$ is as large as possible~\cite{GoreinovEtAl10}.
-Now, keeping $R$ fixed, a set of new columns $C$ is chosen so that the submatrix $U$ within $R$ is as large as possible.
-This process repeats a small number of times.
-This results in matrices $R$ and $C$ such that the error in the approximation,~\eqref{eq:CUR}, is small.
-The dominant cost of this procedure is the cost of computing $a \hrank$ columns of $X$ and $a\hrank$ rows of $X$, where $a$ is the number of alternating iterations.
-There is also a small linear algebra overhead cost for the determinant maximization process that is performed at each step.
-%The key point is that the adaptive cross method allows us to form a rank-$\hrank$ approximation of an $m \times m$ block of the matrix via a process that only requires accessing $O(m\hrank)$ entries of that block.
-%
-%We use the adaptive cross method to form low rank approximations for each admissible block of $\mathbf{\Aker}$.
-We directly compute all entries of the small dense blocks of $\mathbf{\Aker}$ that are not admissible.
-This process requires us to compute $O(\hrank^2 \fedim \log \fedim)$ entries of $\mathbf{\Aker}$, which is relatively cheap compared to the operator actions of $\Aop$ that are required to form the product-convolution approximation.
-
-\section{Ellipsoid intersection test}
-\label{sec:fast_ellipsoid_intersection_test}
-The procedure for choosing sample points relies on quickly determining whether two ellipsoids intersect. Let $E_1$ and $E_2$ be the ellipsoids defined as
-\begin{align*}
-	E_i :=& \{x : (x - \spatialmean_i)^T \spatialcov_i^{-1} (x - \spatialmean_i) \le \tau^2\}
-\end{align*}
-for $i\in\{1,2\}$, where $\spatialmean_1, \spatialmean_2 \in \mathbb{R}^\gdim$, and $\spatialcov_1, \spatialcov_2 \in \mathbb{R}^{\gdim \times \gdim}$ are positive definite.
-Let $K$ be the following one dimensional convex function:
-\begin{equation*}
-	K(s) := 1 - \frac{1}{\tau^2} (\spatialmean_1 - \spatialmean_2)^T \left(\frac{1}{1-s}\spatialcov_1 + \frac{1}{s}\spatialcov_2\right)^{-1}(\spatialmean_1 - \spatialmean_2).
-\end{equation*}
-In~\cite{GilitschenskiHanebeck12} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$.
-% (see Figure~\ref{fig:ellipsoid_intersection_test}).
-
-%\begin{figure}
-%	\begin{subfigure}{0.37\textwidth}
-%		\centering
-%		\includegraphics[scale=1.0]{ellipsoids_intersect0.pdf}
-%	\end{subfigure}
-%	\begin{subfigure}{0.30\textwidth}
-%		\centering
-%		\includegraphics[scale=1.0]{ellipsoids_intersect1.pdf}
-%	\end{subfigure}
-%	\begin{subfigure}{0.30\textwidth}
-%		\centering
-%		\includegraphics[scale=1.0]{ellipsoids_intersect2.pdf}
-%	\end{subfigure}
-%	\caption{The ellipsoids intersect if $K(s) \ge 0$ for all $s \in (0,1)$.
-%	}
-%	\label{fig:ellipsoid_intersection_test}
-%\end{figure}
-
-The function $K(s)$ may be evaluated quickly for many $s$ by computing the solution to the generalized eigenvalue problem
-$
-	\spatialcov_1 \eigenvectormatrix = \spatialcov_2 \eigenvectormatrix \Lambda,
-$
-where $\eigenvectormatrix \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors, and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$.
-%The matrix $\eigenvectormatrix$ simultaneously diagonalizes $\spatialcov_1$ and $\spatialcov_2$, in the sense that $\eigenvectormatrix^T\spatialcov_1 \eigenvectormatrix = \Lambda$, and $\eigenvectormatrix^T\spatialcov_2 \eigenvectormatrix = I$, where $I$ is the $\gdim \times \gdim$ identity matrix.
-Using this diagonalization, and
-After some algebraic manipulations, we have
-\begin{equation}
-	\label{eq:Ks_generalized}
-	K(s) = 1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2,
-\end{equation}
-where $v := \eigenvectormatrix^T\left(\spatialmean_1 - \spatialmean_2\right)$.
-To check whether $E_1$ and $E_2$ intersect, we compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, form $v$, then minimize $K(s)$ in the form~\eqref{eq:Ks_generalized} on $(0,1)$ using Brent's method~\cite{Brent71}.
-If $K(s^*) <0$ at the minimizer, $s^*$, then $E_1 \cap E_2 = \{\}$.
-Otherwise $E_1 \cap E_2 \neq \{\}$.
-%The resulting algorithm for checking whether $E_1$ and $E_2$ intersect is summarized in Algorithm~\ref{alg:ellipsoid_intersection_test}.
-
-%\begin{algorithm2e}
-%	\SetAlgoNoLine
-%	\SetKwInOut{Input}{Input}
-%	\SetKwInOut{Output}{Output}
-%	\SetKwProg{Fn}{Function}{}{}
-%	\Input{Ellipsoid $E_1$ with mean $\spatialmean_1$ and covariance $\spatialcov_1/\tau^2$\\
-%		Ellipsoid $E_2$ with mean $\spatialmean_2$ and covariance $\spatialcov_2/\tau^2$ 
-%	}
-%	
-%	\Output{Boolean $\text{ellipsoids\_intersect}$ which is true if $E_1 \cap E_2 \neq \{\}$ and false otherwise}
-%	
-%	
-%	Solve generalized eigenvalue problem $\spatialcov_1 P = \spatialcov_2 P \Lambda$
-%		
-%	$v \gets P^T\left(\spatialmean_1 - \spatialmean_2\right)$
-%		
-%	$\displaystyle K^* \gets \min_{s \in (0,1)}~1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2$
-%		
-%	\If{$K^* < 0$}{
-%			
-%		$\text{ellipsoids\_intersect} \gets \text{False}$
-%			
-%	}
-%	\Else{
-%			
-%		$\text{ellipsoids\_intersect} \gets \text{True}$
-%			
-%	}
-%	\caption{Determining whether two ellipsoids intersect}
-%	\label{alg:ellipsoid_intersection_test}
-%\end{algorithm2e}
-
-%\subsection{Ellipsoid bounding box test}
-
-This ellipsoid intersection test is only performed if the coordinate axis aligned bounding boxes for the ellipsoids intersect. The smallest coordinate axis aligned box that contains $E_i$ is the box
-$
-	\prod_{k=1}^\gdim \left[ \spatialmean_i^{k} -\tau\sqrt{\Sigma_i^{kk}}, \spatialmean_i^{k} +\tau\sqrt{\Sigma_i^{kk}} \right],
-$
-where $\spatialmean_i^{k}$ is the $k$th entry of $\spatialmean_i$, and $\Sigma_i^{kk}$ is the $k$th diagonal entry of $\Sigma_i$\footnote{See \url{https://math.stackexchange.com/q/3928964}}. The boxes
-$\prod_{k=1}^\gdim \left[a_{-}^k, a_{+}^k\right]$ and $\prod_{k=1}^\gdim \left[b_{-}^k, b_{+}^k\right]$
-intersect if and only if $a_{-}^k \le b_{+}^k$ and $b_{-}^k \le a_{+}^k$ for $k=1,\dots,d$.
-
-%\section{Rational positive semi-definite modification}
-%\label{sec:make_spd}
-%In many problems $\Aop$ is symmetric positive semi-definite. However, $\mathbf{A}_H$ is non-symmetric and indefinite because of approximation error. 
-%This is undesirable.
-Symmetry and positive semi-definiteness are important properties which should be preserved if possible.
-Also, lacking these properties may prevent one from using highly effective algorithms, such as the conjugate gradient method, to perform further operations involving $\mathbf{A}_H$.
-%
-%If $\Aop$ is symmetric, we symmetrize $\mathbf{A}_H$ as follows:
-%$
-%\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
-%$
-%If $\Aop$ is positive semi-definite, we adapt the rational method from~\cite[Section 13.8]{HackbuschKress07} to approximate the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$, which we denote by $|\mathbf{A}_H^{\text{sym}}|$.
-That is, $|\mathbf{A}_H^{\text{sym}}|$ has the same eigenvectors as $\mathbf{A}_H^{\text{sym}}$, but the eigenvalues of $|\mathbf{A}_H^{\text{sym}}|$ are the absolute values of the eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
-%%flip (or approximately flip) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ to be positive. 
-%Once constructed, we use the approximation of $|\mathbf{A}_H^{\text{sym}}|$ in place of $\mathbf{A}_H$ in further linear algebra operations.
-%
-%In detail, let $a$ be any scalar satisfying $-a < \lambda_\text{min}$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, and set 
-%\begin{equation}
-%\label{eq:ratT999}
-%\mathbf{T} :=  \left(2 \mathbf{A}_H^{\text{sym}} + a \mathbf{I}\right)/a,
-%\end{equation}
-%From the results in~\cite[Section 13.8]{HackbuschKress07} it is straightforward to show that
-%\begin{equation}
-%\label{eq:rational343}
-%\mathbf{\Pi}_\ratord := \left(\mathbf{I} + \mathbf{T}^{2^\ratord}\right)^{-1}
-%\end{equation}
-%approximates the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to negative eigenvalues. Thus, we have
-%\begin{equation}
-%\label{eq:rational76567}
-%|\mathbf{A}_H^{\text{sym}}| \approx \left(\mathbf{I} - 2\mathbf{\Pi}_\ratord\right)\mathbf{A}_H^{\text{sym}}.
-%\end{equation}
-%To form an approximation of $|\mathbf{A}_H^{\text{sym}}|$, first, we compute $\lambda_\text{min}$ using the Lanczos method, and set $a=\gamma \lambda_\text{min}$, where $\gamma\ge 1$ is a parameter that must be chosen.
-%The method is relatively insensitive to $\gamma$ and numerically we find that the method works well for any $\gamma \in [1,2]$. We use $\gamma=1.5$ in our numerical results. 
-%Next, we compute $\mathbf{T}$ per~\eqref{eq:ratT999}, compute $\mathbf{T}^{2^\ratord}$ per~\eqref{eq:rational343} via repeated squaring, compute $\mathbf{\Pi}_\ratord$ per~\eqref{eq:rational343}, and form an approximation of $|\mathbf{A}_H^{\text{sym}}|$ per~\eqref{eq:rational76567}.
-Fast H-matrix methods are used to perform the required matrix-matrix additions and multiplications, and matrix inversion.
-We recommend using $\ratord=1$ or $2$.
-Larger $\ratord$ is unnecessary and may make the method unstable.
-
-%Finally, we approximate $|\mathbf{A}_H^{\text{sym}}|$ per~\eqref{eq:rational76567}, using fast H-matrix operations to perform the required matrix-matrix addition and matrix inversion.
-
-%
-%define $
-%\mathbf{T} := T(\mathbf{A}_H^{\text{sym}}) = \left(2 \mathbf{A}_H^{\text{sym}} + a \mathbf{I}\right)/a.
-%$, where $-a < \lambda_\text{min}$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$ and define
-%
-%let $\mathbf{\Pi}$ be the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to eigenvalues of $\mathbf{A}_H^{\text{sym}}$ that reside in $(-a,0)$
-%% let $\Pi$ be the indicator function for the interval $(a,b)$. 
-%%That is, $\Pi(\lambda)=0$ for $\lambda \in (0,1)$, and $\Pi(\lambda)=0$ otherwise.
-%%\begin{equation*}
-%%	\Pi(\lambda) = \begin{cases}
-%%		1, & \lambda \in (a,b), \\
-%%		0, & \text{otherwise}.
-%%	\end{cases}
-%%\end{equation*}
-%%Using functional calculus to extend the domain of the function $\Pi$ from scalars to matrices, we define
-%%$
-%%\mathbf{\Pi} := \Pi(\mathbf{A}_H^{\text{sym}}).
-%%$
-%%The matrix $\mathbf{\Pi}$ is the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to eigenvalues of $\mathbf{A}_H^{\text{sym}}$ that reside in $(a,b)$. 
-%If we choose $-a < \lambda_\text{min}$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, then $\mathbf{\Pi}$ is the spectral projector onto the eigenspace associated with negative eigenvalues,
-%%and $\mathbf{\Pi}\mathbf{A}_H^{\text{sym}}$ is the ``negative component'' of $\mathbf{A}_H^{\text{sym}}$ in which all the negative eigenvalues are the same but all the positive eigenvalues are replaced by zero.
-Subtracting a negative number from itself twice flips its sign,
-%so the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$ is given by
-%\begin{equation}
-%\label{eq:I_Pi_abs}
-%	|\mathbf{A}_H^{\text{sym}}| = \left(\mathbf{I} - 2\mathbf{\Pi}\right)\mathbf{A}_H^{\text{sym}},
-%\end{equation}
-%where $\mathbf{I}$ is the identity matrix of the appropriate size.
-%By ``matrix absolute value,'' we mean that $|\mathbf{A}_H^{\text{sym}}|$ is the matrix that has the same eigenvectors as $\mathbf{A}_H^{\text{sym}}$, but the eigenvalues of $|\mathbf{A}_H^{\text{sym}}|$ are the absolute values of the eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
-%%\begin{figure}
-%%	\centering
-%%	\includegraphics[scale=0.5]{spd_rational_function.pdf}
-%%	\caption{Rational function $\Pi_\ratord(\lambda)$ for $a=-1$, $b=0$, and $\ratord \in \{1,2,3\}$. As $\ratord$ increases, $\Pi_\ratord$ approaches the indicator function for the set $(a,b)$. }
-%%	\label{fig:spd_rat}
-%%\end{figure}
-%Computing $\mathbf{\Pi}$ directly is difficult, but we may approximate $\mathbf{\Pi}$ by the following high order rational matrix function,
-%%$\Pi_\ratord \approx \Pi$ which is defined as follows:
-%%\begin{equation*}
-%%	 \Pi_\ratord(\lambda) := \frac{1}{1 + T(\lambda)^{2^\ratord} }, \qquad \text{where} \qquad T(\lambda) := \frac{2 \lambda - (b+a)}{b-a}.
-%%\end{equation*}
-%%Here $\ratord$ is a positive integer; the larger $\ratord$ is, the better the approximation $\Pi \approx \Pi_\ratord$. 
-%%%This is illustrated in Figure~\ref{fig:spd_rat}. 
-%%Replacing scalar operations with matrix operations, we have an approximation $\mathbf{\Pi}_\ratord \approx \mathbf{\Pi}$ given by
-%\begin{equation}
-%\label{eq:rational343}
-%\mathbf{\Pi} \approx \mathbf{\Pi}_\ratord := \left(\mathbf{I} + \mathbf{T}^{2^\ratord}\right)^{-1},
-%\end{equation}
-%where
-%$
-%\mathbf{T} := T(\mathbf{A}_H^{\text{sym}}) = \left(2 \mathbf{A}_H^{\text{sym}} + a \mathbf{I}\right)/a.
-%$
-%Using $\mathbf{\Pi}_\ratord$ instead of $\mathbf{\Pi}$ in~\eqref{eq:I_Pi_abs} allows us to approximate $|\mathbf{A}_H^{\text{sym}}|$.
-First, we compute $\lambda_\text{min}$ using the Lanczos method, and set $a=\gamma \lambda_\text{min}$ and $b=0$, where $\gamma\ge 1$ is a parameter that must be chosen.
-%The method is relatively insensitive to $\gamma$ and numerically we find that the method works well for any $\gamma \in [1,2]$. We use $\gamma=1.5$ in our numerical results. 
-%Next, we compute $\mathbf{T}$ using fast H-matrix scalar multiplication and addition.
-We form the $\mathbf{T}^{2^\ratord}$ via repeated squaring using H-matrix multiplication.
-We recommend using $\ratord=1$ or $2$.
-Larger $\ratord$ is unnecessary and may make the method unstable.
-The approximate spectral projector $\mathbf{\Pi}_\ratord$ is formed from $\mathbf{T}^{2^\ratord}$ per~\eqref{eq:rational343}, using fast H-matrix operations to perform the required matrix-matrix addition and matrix inversion.
-Finally, a positive semi-definite approximation $\mathbf{A}_H^{\text{sym}+}$ of $\mathbf{A}_H^{\text{sym}}$ is formed as
-%\begin{equation*}
-%\mathbf{A}_H^{\text{sym}+} := \left(\mathbf{I} - 2\mathbf{\Pi}_\ratord\right)\mathbf{A}_H^{\text{sym}},
-%\end{equation*}
-%where fast H-matrix operations are used to perform the necessary matrix-matrix subtraction and multiplication.
-Note that for any $\ratord \ge 1$, $\Pi_\ratord(\lambda)> 1/2$ for $\lambda \in (a,b)$, so $\mathbf{A}_H^{\text{sym}+}$ is positive semi-definite\footnote{Some indefiniteness may be introduced by truncation error in the H-matrix operations, but the negative eigenvalues caused by this truncation error are typically extremely small.}.
-At the end, $\mathbf{A}_H^{\text{sym}+}$ is symmetrized once again, to correct for any small non-symmetries that are introduced by truncation error in the H-matrix operations.
-Once constructed, we use the symmetric positive semi-definite matrix $\mathbf{A}_H^{\text{sym}+}$ in place of $\mathbf{A}_H$ in further linear algebra operations.
-
-%\section{Recycling Krylov information via low rank updates}
-%\label{sec:recycle_krylov}
-%
-%In applications, often one wants to use the H-matrix approximation $\mathbf{A}_H$ to build preconditioners for solving a sequence of slowly varying linear systems, where the $k$th linear system has the form
-%\begin{equation}
-%\label{eq:Ax_equals_b}
-%\mathbf{B}^{(k)} \mathbf{u}^{(k)} = \mathbf{f}^{(k)}.
-%\end{equation}
-%Each of these linear systems is solved using an iterative method. By ``slowly varying,'' we mean that $\mathbf{B}^{(k+1)}$ does not differ much from $\mathbf{B}^{(k)}$. 
-%For example, in a Newton-Krylov method for solving an optimization problem, one uses a Krylov method such as conjugate gradient to solve a sequence of systems of the form~\eqref{eq:Ax_equals_b}, where $\mathbf{B}^{(k)}$ and $\mathbf{f}^{(k)}$ are the Hessian and the negative gradient of the objective function in the optimization problem, respectively, evaluated at the $k$th Newton iterate.
-%As the optimization solver converges, $\mathbf{B}^{(k)}$ converges to the Hessian at the optimal point, and therefore changes little from one iteration to the next.
-Since $\mathbf{B}^{(k)}$ varies slowly, it is reasonable to assume that Krylov information about $\mathbf{B}^{(k)}$ can be leveraged to update the preconditioner from $\mathbf{B}^{(k)}$ in order to build a preconditioner for $\mathbf{B}^{(k+1)}$.
-%Concretely, let 
-%\begin{equation*}
-%\mathbf{P}^{(k)} \approx \left(\mathbf{B}^{(k)}\right)^{-1}
-%\end{equation*}
-%denote a H-matrix preconditioner used to accelerate the solution of~\eqref{eq:Ax_equals_b} with an iterative method.
-The iterative method for solving the $k$th linear system will typically involve, as a sub-problem, applying $\mathbf{B}^{(k)}$ to a collection of vectors $\{\mathbf{x}_i\}_{i=1}^m$, thereby generating vectors $\{\mathbf{y}_i = \mathbf{B}^{(k)} \mathbf{x}_i\}_{i=1}^m$.
-Instead of rebuilding the H-matrix preconditioner for each successive linear system, we \emph{recycle} the information generated by the iterative method while solving the current linear system, performing a low-rank update to $\mathbf{P}^{(k)}$ to create $\mathbf{P}^{(k+1)}$.
-To do so, we build on the rank-1 update formulas discussed in~\cite[Chapter 6]{NocedalWright99}, which we generalize to many vector ``rank-$m$'' updates.
-In what follows, let $\mathbf{X}$ and $\mathbf{Y}$ denote the $\fedim \times m$ matrices with columns given by the vectors $\mathbf{x}_i$ and $\mathbf{y}_i$, respectively.
-The matrix-vector products with $\mathbf{B}^{(k)}$ performed while solving the $k$th linear system may be summarized by the equation:
-%\begin{equation*}
-%\mathbf{B}^{(k)}\mathbf{X} = \mathbf{Y}.
-%\end{equation*}
-%To form $\mathbf{P}^{(k+1)}$ from $\mathbf{P}^{(k)}$, one may use one of the following three update formulas:
-%\begin{align}
-%	\mathbf{P}^{(k+1)} &= \mathbf{P}^{(k)} + \mathbf{R}\left(\mathbf{Y}^T \mathbf{Y}\right)^{-1} \mathbf{Y}^T \label{eq:res_update} \\
-%	\mathbf{P}^{(k+1)} &= \mathbf{P}^{(k)} + \mathbf{R} \left(\mathbf{R}^T \mathbf{Y}\right)^{-1} \mathbf{R}^T \label{eq:srk_update} \\
-%	\mathbf{P}^{(k+1)} &= \left(\mathbf{I} - \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T\right)\mathbf{P}^{(k)}\left(\mathbf{I} - \mathbf{Y}\boldsymbol{\Theta}\mathbf{X}^T\right) + \mathbf{X}\boldsymbol{\Theta}\mathbf{X}^T, \label{eq:dfp_update}
-%\end{align}
-%where $\mathbf{R} := \mathbf{X}-\mathbf{P}^{(k)}\mathbf{Y}$ and $\boldsymbol{\Theta} := \left(\mathbf{X}^T\mathbf{Y}\right)^{-1}$.
-%We note that for all three cases, the update is exact for $\mathbf{B}^{k}$, in the sense that 
-%\begin{equation*}
-%\mathbf{P}^{(k+1)}\mathbf{Y}=\mathbf{X} = \left(\mathbf{B}^{(k)}\right)^{-1} \mathbf{Y}.
-%\end{equation*}
-%Because the matrices $\mathbf{B}^{k}$ vary slowly, we expect that $\mathbf{P}^{(k+1)}$ is a good approximation for $\left(\mathbf{B}^{(k+1)}\right)^{-1}$ on the space spanned by the vectors $\mathbf{y}_i$.
-One should choose which formula to use based on the properties of the matrix $\mathbf{B}^{(k)}$.
-Formula~\eqref{eq:res_update} does not preserve symmetry, and should therefore be used when $\mathbf{B}^{(k)}$ is not symmetric.
-Formula~\eqref{eq:srk_update} preserves symmetry but not positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric and indefinite.
-Formula~\eqref{eq:dfp_update}, which is used in this paper, preserves both symmetry and positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric positive definite.
-Formulas~\eqref{eq:res_update} and~\eqref{eq:srk_update} perform rank-$m$ updates, while formula~\eqref{eq:dfp_update} performs a rank-$2m$ update.
-%We note that when $m=1$, formulas~\eqref{eq:srk_update} are~\eqref{eq:dfp_update} are known as SR1 (symmetric rank-$1$) and DFP (Davidson-Fletcher-Powell) updates, respectively~\cite[Chapter 6]{NocedalWright99}.
-
-\section*{Acknowledgments}
-We thank J.J. Alger, Longfei Gao, Mathew Hu, and Rami Nammour for helpful discussions. We thank Trevor Heise for editing suggestions. We thank Georg Stadler for domain geometry help.
-
-\FloatBarrier
-
-\bibliographystyle{siamplain}
-\bibliography{localpsf}
-
 \end{document}

### 2023-06-01 paper/localpsf.tex
move adv pcg and geigs plots to separate files

@@
 \begin{figure}
-	\caption{Impulse responses for various diffusivities kappa and fixed simulation time Tf=0.5 (top) and simulation times with fixed kappa=1e-3 (bot).}
+	\caption{Impulse responses for various diffusivities kappa and fixed simulation time Tf=0.5 (top) and simulation times with fixed $\kappa=1e{-3}$ (bot).}
 	\label{fig:adv_vary_kappa_tf}
@@
 \begin{figure}
-	\begin{subfigure}{0.49\textwidth}
+	\begin{subfigure}{0.52\textwidth}
 		\centering
-		\begin{tikzpicture}[baseline, trim axis right]
-		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
-		ymode=log,
-		ymin=1e-11,
-		ymax=2,
-		xmax=1000,
-		axis x line*=bottom,
-		axis y line*=left,
-		xtick={0, 200, 400, 600,800,1000},
-		xlabel={CG iteration},
-		ylabel={$\|\bm{x}_k - \bm{x}\|/\|\bm{x}\|$},
-		title={CG convergence},
-		legend style={at={(1.2,1)},anchor=north east},
-		%legend pos=north east,
-		legend style={font=\small}
-		]
-		\addlegendentry{Reg}
-		\addplot[dotted, line width=0.5mm] table {advDiff_cg_rand_errs_reg.dat};
-		\addlegendentry{None}
-		\addplot[dash dot, line width=0.5mm] table {advDiff_cg_rand_errs_none.dat};
-		\addlegendentry{PSF (1)}
-		\addplot[blue, line width=0.5mm] table {advDiff_cg_rand_errs_psf1.dat};
-		\addlegendentry{PSF (5)}
-		\addplot[red, line width=0.5mm] table {advDiff_cg_rand_errs_psf5.dat};
-		\addlegendentry{PSF (25)}
-		\addplot[purple, line width=0.5mm] table {advDiff_cg_rand_errs_psf25.dat};
-		\end{axis}
-		\end{tikzpicture}
+		\includegraphics[scale=1.0]{adv_pcg_convergence.pdf}
+%		\begin{tikzpicture}[baseline, trim axis right]
+%		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
+%		ymode=log,
+%		ymin=1e-11,
+%		ymax=2,
+%		xmax=1000,
+%		axis x line*=bottom,
+%		axis y line*=left,
+%		xtick={0, 200, 400, 600,800,1000},
+%		xlabel={CG iteration},
+%		ylabel={$\|\bm{x}_k - \bm{x}\|/\|\bm{x}\|$},
+%		title={CG convergence},
+%		legend style={at={(1.2,1)},anchor=north east},
+%		%legend pos=north east,
+%		legend style={font=\small}
+%		]
+%		\addlegendentry{Reg}
+%		\addplot[dotted, line width=0.5mm] table {advDiff_cg_rand_errs_reg.dat};
+%		\addlegendentry{None}
+%		\addplot[dash dot, line width=0.5mm] table {advDiff_cg_rand_errs_none.dat};
+%		\addlegendentry{PSF (1)}
+%		\addplot[blue, line width=0.5mm] table {advDiff_cg_rand_errs_psf1.dat};
+%		\addlegendentry{PSF (5)}
+%		\addplot[red, line width=0.5mm] table {advDiff_cg_rand_errs_psf5.dat};
+%		\addlegendentry{PSF (25)}
+%		\addplot[purple, line width=0.5mm] table {advDiff_cg_rand_errs_psf25.dat};
+%		\end{axis}
+%		\end{tikzpicture}
 		\caption{}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.49\textwidth}
+	\begin{subfigure}{0.47\textwidth}
 		\centering
-		\begin{tikzpicture}[baseline, trim axis right]
-\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
-ymode=log,
-xmin=-50,
-xmax=600,
-ymax=1.e6,
-title={Generalized eigenvalues},
-xlabel={$k$, generalized eigenvalue \#},
-ylabel={$\lambda_{k}$},
-xtick={0, 300, 600, 900, 1200},
-%legend pos=north east,
-legend style={font=\small},
-ylabel near ticks, ylabel shift={-6pt},
-axis x line*=bottom,
-axis y line*=right,
-%legend style={at={(0.95,1)},anchor=north east},
-]
-%\addlegendentry{REG}
-\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {advDiff_generalizedEigenvalues.dat};
-\addplot[dash dot, line width=0.5mm] table[x=k, y=none] {advDiff_generalizedEigenvalues.dat};
-%\addlegendentry{PSF (1)}
-\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {advDiff_generalizedEigenvalues.dat};
-%\addlegendentry{PSF (5)}
-\addplot[red, line width=0.5mm] table[x=k, y=psf5] {advDiff_generalizedEigenvalues.dat};
-%\addlegendentry{PSF (25)}
-\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {advDiff_generalizedEigenvalues.dat};
-\end{axis}
-\end{tikzpicture}
+		\includegraphics[scale=1.0]{adv_geigs_figure.pdf}
+%		\begin{tikzpicture}[baseline, trim axis right]
+%\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
+%ymode=log,
+%xmin=-50,
+%xmax=600,
+%ymax=1.e6,
+%title={Generalized eigenvalues},
+%xlabel={$k$, generalized eigenvalue \#},
+%ylabel={$\lambda_{k}$},
+%xtick={0, 300, 600, 900, 1200},
+%%legend pos=north east,
+%legend style={font=\small},
+%ylabel near ticks, ylabel shift={-6pt},
+%axis x line*=bottom,
+%axis y line*=right,
+%%legend style={at={(0.95,1)},anchor=north east},
+%]
+%%\addlegendentry{REG}
+%\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {advDiff_generalizedEigenvalues.dat};
+%\addplot[dash dot, line width=0.5mm] table[x=k, y=none] {advDiff_generalizedEigenvalues.dat};
+%%\addlegendentry{PSF (1)}
+%\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {advDiff_generalizedEigenvalues.dat};
+%%\addlegendentry{PSF (5)}
+%\addplot[red, line width=0.5mm] table[x=k, y=psf5] {advDiff_generalizedEigenvalues.dat};
+%%\addlegendentry{PSF (25)}
+%\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {advDiff_generalizedEigenvalues.dat};
+%\end{axis}
+%\end{tikzpicture}
 		\caption{}

### 2023-06-01 paper/localpsf.tex
move stokes pcg and geigs plots to separate files

@@
 		\centering
-		\includegraphics[scale=0.33]{impulse_batch1.png}
+		\includegraphics[scale=0.34]{impulse_batch1.png}
 	\end{subfigure}
@@
 		\centering
-		\includegraphics[scale=0.33]{impulse_batch2.png}
+		\includegraphics[scale=0.34]{impulse_batch2.png}
 	\end{subfigure}
@@
 		\centering
-		\includegraphics[scale=0.33]{impulse_batch3.png}
+		\includegraphics[scale=0.34]{impulse_batch3.png}
 	\end{subfigure}
@@
 \begin{figure}
-	\begin{subfigure}{0.49\textwidth}
+	\begin{subfigure}{0.52\textwidth}
 		\centering
-		\begin{tikzpicture}[baseline, trim axis right]
-		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
-		ymode=log,
-		ymin=1e-13,
-		ymax=2,
-		xmax=750,
-		axis x line*=bottom,
-		axis y line*=left,
-		xtick={0, 200, 400, 600},
-		title={CG convergence},
-		xlabel={$k$, CG iteration},
-		ylabel={$\|\bm{x}_k-\bm{x}\|/\|\bm{x}\|$},
-		legend style={at={(1.1,1)},anchor=north east},
-		%legend pos=north east,
-		legend style={font=\small}
-		]
-		\addlegendentry{Reg}		
-		\addplot[dotted, line width=0.5mm] table {cg_relres_reg_0.05noise.dat};
-		\addlegendentry{None}
-		\addplot[dash dot, line width=0.5mm] table {cg_relres_none_0.05noise.dat};
-		\addlegendentry{PSF (1)}
-		\addplot[blue, line width=0.5mm] table {cg_relres_psf1_0.05noise.dat};
-		\addlegendentry{PSF (5)}
-		\addplot[red, line width=0.5mm] table {cg_relres_psf5_0.05noise.dat};
-		\addlegendentry{PSF (25)}
-		\addplot[purple, line width=0.5mm] table {cg_relres_psf25_0.05noise.dat};
-		\end{axis}
-		\end{tikzpicture}
+		\includegraphics[scale=1.0]{stokes_pcg_convergence.pdf}
+%		\begin{tikzpicture}[baseline, trim axis right]
+%		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
+%		ymode=log,
+%		ymin=1e-13,
+%		ymax=2,
+%		xmax=750,
+%		axis x line*=bottom,
+%		axis y line*=left,
+%		xtick={0, 200, 400, 600},
+%		title={CG convergence},
+%		xlabel={$k$, CG iteration},
+%		ylabel={$\|\bm{x}_k-\bm{x}\|/\|\bm{x}\|$},
+%		legend style={at={(1.1,1)},anchor=north east},
+%		%legend pos=north east,
+%		legend style={font=\small}
+%		]
+%		\addlegendentry{Reg}		
+%		\addplot[dotted, line width=0.5mm] table {cg_relres_reg_0.05noise.dat};
+%		\addlegendentry{None}
+%		\addplot[dash dot, line width=0.5mm] table {cg_relres_none_0.05noise.dat};
+%		\addlegendentry{PSF (1)}
+%		\addplot[blue, line width=0.5mm] table {cg_relres_psf1_0.05noise.dat};
+%		\addlegendentry{PSF (5)}
+%		\addplot[red, line width=0.5mm] table {cg_relres_psf5_0.05noise.dat};
+%		\addlegendentry{PSF (25)}
+%		\addplot[purple, line width=0.5mm] table {cg_relres_psf25_0.05noise.dat};
+%		\end{axis}
+%		\end{tikzpicture}
 		\caption{}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.49\textwidth}
+	\begin{subfigure}{0.47\textwidth}
 	\begin{center}
-		\begin{tikzpicture}[baseline, trim axis right]
-		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
-		ymode=log,
-		xmin=-50,
-		xmax=1450,
-		ymax=1.e7,
-		title={Generalized eigenvalues},
-		xlabel={$k$, generalized eigenvalue \#},
-		ylabel={$\lambda_{k}$},
-		xtick={0, 300, 600, 900, 1200},
-		%legend pos=north east,
-		legend style={font=\small},
-		ylabel near ticks, ylabel shift={-6pt},
-		axis x line*=bottom,
-		axis y line*=right,
-		%legend style={at={(0.95,1)},anchor=north east},
-		]
-		%\addlegendentry{REG}
-		\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues_0.05noise.dat};
-		\addplot[dash dot, line width=0.5mm] table[x=k, y=none] {generalizedEigenvalues_0.05noise.dat};
-		%\addlegendentry{PSF (1)}
-		\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues_0.05noise.dat};
-		%\addlegendentry{PSF (5)}
-		\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues_0.05noise.dat};
-		%\addlegendentry{PSF (25)}
-		\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues_0.05noise.dat};
-		\end{axis}
-		\end{tikzpicture}
+		\includegraphics[scale=1.0]{stokes_geigs_figure.pdf}
+%		\begin{tikzpicture}[baseline, trim axis right]
+%		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
+%		ymode=log,
+%		xmin=-50,
+%		xmax=1450,
+%		ymax=1.e7,
+%		title={Generalized eigenvalues},
+%		xlabel={$k$, generalized eigenvalue \#},
+%		ylabel={$\lambda_{k}$},
+%		xtick={0, 300, 600, 900, 1200},
+%		%legend pos=north east,
+%		legend style={font=\small},
+%		ylabel near ticks, ylabel shift={-6pt},
+%		axis x line*=bottom,
+%		axis y line*=right,
+%		%legend style={at={(0.95,1)},anchor=north east},
+%		]
+%		%\addlegendentry{REG}
+%		\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues_0.05noise.dat};
+%		\addplot[dash dot, line width=0.5mm] table[x=k, y=none] {generalizedEigenvalues_0.05noise.dat};
+%		%\addlegendentry{PSF (1)}
+%		\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues_0.05noise.dat};
+%		%\addlegendentry{PSF (5)}
+%		\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues_0.05noise.dat};
+%		%\addlegendentry{PSF (25)}
+%		\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues_0.05noise.dat};
+%		\end{axis}
+%		\end{tikzpicture}
 	\end{center}

### 2023-06-01 paper/stokes_geigs_figure.tex
move stokes pcg and geigs plots to separate files

\documentclass{standalone}
\usepackage{tikz}
\usepackage{pgfplots}

\begin{document}
	\begin{tikzpicture}[baseline]%, trim axis right]
		\begin{axis}[width=0.49\linewidth, height=0.5\linewidth,
		ymode=log,
		xmin=-50,
		xmax=1450,
		ymax=1.e7,
		title={Generalized eigenvalues},
		xlabel={$k$, generalized eigenvalue \#},
		ylabel={$\lambda_{k}$},
		xtick={0, 300, 600, 900, 1200},
		%legend pos=north east,
		legend style={font=\small},
		ylabel near ticks, ylabel shift={-6pt},
		axis x line*=bottom,
		axis y line*=right,
		%legend style={at={(0.95,1)},anchor=north east},
		]
		%\addlegendentry{REG}
		\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues_0.05noise.dat};
		\addplot[dash dot, line width=0.5mm] table[x=k, y=none] {generalizedEigenvalues_0.05noise.dat};
		%\addlegendentry{PSF (1)}
		\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues_0.05noise.dat};
		%\addlegendentry{PSF (5)}
		\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues_0.05noise.dat};
		%\addlegendentry{PSF (25)}
		\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues_0.05noise.dat};
		\end{axis}
	\end{tikzpicture}
\end{document}

### 2023-06-01 paper/stokes_pcg_convergence.tex
move stokes pcg and geigs plots to separate files

--- 2023-06-01 paper/adv_pcg_convergence.tex
@@
 	ymode=log,
-		ymin=1e-11,
+	ymin=1e-13,
 	ymax=2,
-		xmax=1000,
+	xmax=750,
 	axis x line*=bottom,
 	axis y line*=left,
-		xtick={0, 200, 400, 600,800,1000},
-		xlabel={CG iteration},
-		ylabel={$\|\bm{x}_k - \bm{x}\|/\|\bm{x}\|$},
+	xtick={0, 200, 400, 600},
 	title={CG convergence},
+	xlabel={$k$, CG iteration},
+	ylabel={$\|\bm{x}_k-\bm{x}\|/\|\bm{x}\|$},
 	legend style={at={(1.2,1)},anchor=north east,draw=none},
@@
 	\addlegendentry{Reg}		
-			\addplot[dotted, line width=0.5mm] table {advDiff_cg_rand_errs_reg.dat};
+	\addplot[dotted, line width=0.5mm] table {cg_relres_reg_0.05noise.dat};
 	\addlegendentry{None}
-			\addplot[dash dot, line width=0.5mm] table {advDiff_cg_rand_errs_none.dat};
+	\addplot[dash dot, line width=0.5mm] table {cg_relres_none_0.05noise.dat};
 	\addlegendentry{PSF (1)}
-			\addplot[blue, line width=0.5mm] table {advDiff_cg_rand_errs_psf1.dat};
+	\addplot[blue, line width=0.5mm] table {cg_relres_psf1_0.05noise.dat};
 	\addlegendentry{PSF (5)}
-			\addplot[red, line width=0.5mm] table {advDiff_cg_rand_errs_psf5.dat};
+	\addplot[red, line width=0.5mm] table {cg_relres_psf5_0.05noise.dat};
 	\addlegendentry{PSF (25)}
-			\addplot[purple, line width=0.5mm] table {advDiff_cg_rand_errs_psf25.dat};
+	\addplot[purple, line width=0.5mm] table {cg_relres_psf25_0.05noise.dat};
 	\end{axis}

### 2023-06-01 paper/localpsf.tex
update stokes condition number table

@@
 				level    & REG     &	NONE  & PSF $(1)$ & PSF $(5)$ & PSF $(25)$ \\ \hline 
-				$25\%$   & 1.68e+3 & 2.16e+3  & 1.37e+0   & 1.30e+0   & 1.21e+0    \\ 
-				$11\%$   & 9.19e+3 & 9.08e+2  & 1.85e+0   & 1.49e+0   & 1.27e+0    \\   
-				$5.0\%$  & 3.66e+4 & 4.67e+2  & 3.46e+0   & 2.62e+0   & 1.91e+0    \\ 
-				$2.2\%$  & 1.41e+5 & 8.25e+2  & 1.16e+1   & 4.99e+0   & 3.20e+0    \\  
-				$1.0\%$  & 6.06e+5 & 1.70e+3  & 3.44e+1   & 5.38e+1   & 7.06e+0    \\   
+				$25\%$   & 1.01e+3 & 2.96e+3  & 1.34e+0   & 1.30e+0   & 1.18e+0    \\ 
+				$11\%$   & 7.40e+3 & 1.05e+3  & 2.27e+0   & 1.55e+0   & 1.31e+0    \\   
+				$5.0\%$  & 3.29e+4 & 4.96e+2  & 5.61e+0   & 3.06e+0   & 1.92e+0    \\ 
+				$2.2\%$  & 1.66e+5 & 8.89e+2  & 1.58e+1   & 8.07e+0   & 4.03e+0    \\  
+				$1.0\%$  & 5.36e+5 & 1.61e+3  & 7.17e+1   & 1.93e+1   & 9.19e+0    \\   
 			\end{tabular}

### 2023-06-01 paper/localpsf.tex
update stokes newton-cg convergence table

@@
 				2 &
-				3 & 8 & 2.6e+6 &
+				4 & 10 & 2.6e+6 &
 				16 & 34 & 4.1e+6 &
-				3 & 8 & 2.6e+6 \\
+				4 & 10 & 2.6e+6 \\
 				3 &
-				1 & 4+22 & 9.2e+5 &
-				34 & 70 & 1.7e+6 &
-				11 & 24 & 9.2e+5 \\
+				2 & 6+22 & 6.9e+5 &
+				34 & 70 & 1.8e+6 &
+				14 & 30 & 6.9e+5 \\
 				4 &
-				2 & 6 & 1.5e+5 &
-				54 & 110 & 5.0e+5 &
-				30 & 62 & 1.4e+5 \\
+				3 & 8 & 4.4e+4 &
+				52 & 106 & 5.6e+5 &
+				29 & 60 & 1.3e+5 \\
 				5 &
-				4 & 10 & 7.1e+3 &
-				79 & 160 & 8.0e+4 &
-				34 & 70 & 1.4e+4 \\
+				5 & 12 & 2.2e+3 &
+				79 & 160 & 9.4e+4 &
+				38 & 78 & 1.0e+4 \\
 				6 &
-				5 & 12 & 6.0e+1 &
-				112 & 226 & 5.2e+3 &
-				61 & 124 & 2.3e+2 \\
+				0 & 2 & 1.1e+1 &
+				102 & 206 & 6.5e+3 &
+				58 & 118 & 1.8e+2 \\
 				7 &
-				0 & 2 & 8.9e-2 &
-				162 & 326 & 8.0e+1 &
-				84 & 170 & 7.3e-1 \\
+				--- & --- & --- &
+				151 & 304 & 1.2e+2 &
+				0 & 2 & 5.5e-1 \\
 				8 & 
 				--- & --- & --- &
-				0 & 2 & 1.6e-1 &
-				0 & 2 & 3.0e-2 \\
+				0 & 2 & 2.9e-1 &
+				--- & --- & --- \\
 				\hline
 				Total & 
-				18 & 74 & --- &
-				468 & 954 & --- &
-				226 & 470 & --- \\
+				17 & 70 & --- &
+				445 & 908 & --- &
+				146 & 308 & --- \\
 			\end{tabular}
@@
 This consists of Stokes solves for $u$ performed during the linesearch going from $\bm{\basalfriction}_{k-1}$ to $\bm{\basalfriction}_k$, plus one adjoint Stokes solve to compute the gradient at $\bm{\basalfriction}_k$, plus one incremental forward and one incremental adjoint Stokes solve per PCG iteration for solving the Newton system.
-Under PSF (5) and in row Iter 3, we write $4+22$ to indicate that $4$ Stokes solves were used during the standard course of the iteration, and $22$ Stokes solves were used to build the PSF (5) preconditioner.}
+Under PSF (5) and in row Iter 3, we write $6+22$ to indicate that $6$ Stokes solves were used during the standard course of the iteration, and $22$ Stokes solves were used to build the PSF (5) preconditioner.}
 	\label{tab:newton_convergence_table}
@@
 		\includegraphics[scale=1.0]{stokes_pcg_convergence.pdf}
-%		\begin{tikzpicture}[baseline, trim axis right]
-%		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
-%		ymode=log,
-%		ymin=1e-13,
-%		ymax=2,
-%		xmax=750,
-%		axis x line*=bottom,
-%		axis y line*=left,
-%		xtick={0, 200, 400, 600},
-%		title={CG convergence},
-%		xlabel={$k$, CG iteration},
-%		ylabel={$\|\bm{x}_k-\bm{x}\|/\|\bm{x}\|$},
-%		legend style={at={(1.1,1)},anchor=north east},
-%		%legend pos=north east,
-%		legend style={font=\small}
-%		]
-%		\addlegendentry{Reg}		
-%		\addplot[dotted, line width=0.5mm] table {cg_relres_reg_0.05noise.dat};
-%		\addlegendentry{None}
-%		\addplot[dash dot, line width=0.5mm] table {cg_relres_none_0.05noise.dat};
-%		\addlegendentry{PSF (1)}
-%		\addplot[blue, line width=0.5mm] table {cg_relres_psf1_0.05noise.dat};
-%		\addlegendentry{PSF (5)}
-%		\addplot[red, line width=0.5mm] table {cg_relres_psf5_0.05noise.dat};
-%		\addlegendentry{PSF (25)}
-%		\addplot[purple, line width=0.5mm] table {cg_relres_psf25_0.05noise.dat};
-%		\end{axis}
-%		\end{tikzpicture}
 		\caption{}
@@
 		\includegraphics[scale=1.0]{stokes_geigs_figure.pdf}
-%		\begin{tikzpicture}[baseline, trim axis right]
-%		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
-%		ymode=log,
-%		xmin=-50,
-%		xmax=1450,
-%		ymax=1.e7,
-%		title={Generalized eigenvalues},
-%		xlabel={$k$, generalized eigenvalue \#},
-%		ylabel={$\lambda_{k}$},
-%		xtick={0, 300, 600, 900, 1200},
-%		%legend pos=north east,
-%		legend style={font=\small},
-%		ylabel near ticks, ylabel shift={-6pt},
-%		axis x line*=bottom,
-%		axis y line*=right,
-%		%legend style={at={(0.95,1)},anchor=north east},
-%		]
-%		%\addlegendentry{REG}
-%		\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues_0.05noise.dat};
-%		\addplot[dash dot, line width=0.5mm] table[x=k, y=none] {generalizedEigenvalues_0.05noise.dat};
-%		%\addlegendentry{PSF (1)}
-%		\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues_0.05noise.dat};
-%		%\addlegendentry{PSF (5)}
-%		\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues_0.05noise.dat};
-%		%\addlegendentry{PSF (25)}
-%		\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues_0.05noise.dat};
-%		\end{axis}
-%		\end{tikzpicture}
 	\end{center}

### 2023-06-01 paper/localpsf.tex
edited writing in stokes section

@@
 Picking the batches of points requires us to estimate the supports of the impulse responses $\phi_x$ \emph{before} we compute them.
-In resolution analysis in seismic imaging, 
+We form a-priori ellipsoid estimates for the supports of all the $\phi_x$ via a moment method that involves applying $\Aop^T$ to a small number of polynomial functions (see Section~\ref{sec:intromoments}).
+We were inspired by resolution analysis in seismic imaging, in which
 $\Aop^T$ is applied to a random noise function, and the width of $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$~\cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
-Here, we estimate the support of the $\phi_x$ by applying $\Aop^T$ to polynomial functions instead of random noise functions (see Section~\ref{sec:intromoments}).
-Our method estimates the support of $\phi_x$ more accurately than random noise probing, at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
-
-Our method may be categorized as a PSF method that is loosely based on ``product convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
+Our moment method estimates the support of $\phi_x$ more accurately than random noise probing, at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
+
+Our operator approximation method may be categorized as a PSF method that is loosely based on ``product convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
 PC and PSF methods have a long history dating back several decades.
@@
 %\end{equation*}
-The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm~\ref{alg:construct_Atilde}. The cost is discussed in Section~\ref{sec:overall_cost}.
 
@@
 In this case, one may optionally symmetrize $\mathbf{A}_H$, then modify it via low rank updates to remove erroneous negative eigenvalues (Section \ref{sec:flip_eigs}).
+The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm~\ref{alg:construct_Atilde}.
+The cost is discussed in Section~\ref{sec:overall_cost}.
 
@@
 \end{equation*}
+and
+\begin{equation}
+\label{eq:fxyxp}
+f_i := \frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}\left(z_i\right), \qquad i=1,\dots,\numnbr,
+\end{equation}
 where $\{x_i\}_{i=1}^{\numnbr}$ are the $\numnbr$ nearest sample points to $x$, excluding points $x_i$ for which $z_i \notin \Omega$.
@@
 We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree)~\cite{Ericson04} of the mesh cells used to discretize the problem.
-Further, let
-\begin{equation}
-\label{eq:fxyxp}
-f_i := \frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}\left(z_i\right), \qquad i=1,\dots,\numnbr.
-\end{equation}
 Note that $\phi_{x_i}\left(z_i\right)$ is well defined because $z_i \in \Omega$, and $\frac{\spatialvol(x)}{\spatialvol(x_i)}$ is well defined because our sample point picking procedure in Section~\ref{sec:sample_point_selection} ensures that $\spatialvol(x_i)>0$.
@@
 
-We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming H-matrix representations $\mathbf{\Aker}_H$ and $\massmatrix_H$ of $\mathbf{\Aker}$ and $\massmatrix$, respectively, then using fast H-matrix methods to multiply these matrices per~\eqref{eq:boldA} to form
-$
-\mathbf{A}_H = \massmatrix_H \mathbf{\Aker}_H \massmatrix_H.
-$
+We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming an H-matrix representation $\mathbf{\Aker}_H$ of $\mathbf{\Aker}$ then multiplying $\mathbf{\Aker}$ with mass matrices $\massmatrix$ per~\eqref{eq:boldA} to form
+$
+\mathbf{A}_H = \massmatrix \mathbf{\Aker}_H \massmatrix.
+$
+In our case $\massmatrix$ is a lumped mass matrix and is therefore diagonal, so these matrix-matrix multiplications are trivial.
+If a non-diagonal mass matrix is used, one may form a H-matrix representation of the mass matrix, then perform the matrix-matrix multiplications in \eqref{eq:boldA} using H-matrix methods.
 %Once $\mathbf{A}_H$ is formed, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods~\cite{BORMHACKBUSCHBOOK}.
@@
 %We describe this H-matrix construction process in detail in Appendix~\ref{app:h_matrix}. 
-We form $\massmatrix_H$ using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same geometrical clustering block cluster tree as was used for $\mathbf{\Aker}$.
+%We form $\massmatrix_H$ using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same geometrical clustering block cluster tree as was used for $\mathbf{\Aker}$.
 
@@
 This may optionally include modifying the H-matrix to make it symmetric positive semi-definite.
-In practice, (1) is the dominant cost, because (1) is the only task that requires applying $\Aop$ to vectors, and our method is targeted at applications in which applying $\Aop$ to a vector requires an expensive computational procedure such as solving a large scale PDE.
+In practice, (1) is the dominant cost, because (1) is the only task that requires applying $\Aop$ to vectors, and our method is targeted at applications in which applying $\Aop$ to a vector requires an expensive computational procedure such as solving a PDE.
 All operations that do not require applications of $\Aop$ to vectors are nearly linear, and therefore scalable, in the size of the problem, $N$.
@@
 		$\nsamplepts$ & $10^1$--$10^4$ & Total number of sample points (all batches) \\
-		$|\pointbatch_i|$ & $1$--$500$ & Number of sample points in the $i$th batch \\
-		$\ratord$ & $1$--$2$ & Rational function parameter for SPSD modification
+		$|\pointbatch_i|$ & $1$--$500$ & Number of sample points in the $i$th batch
 	\end{tabular}
@@
 where $\nsamplepts$ is the total number of sample points in all batches.
-The multiplicative dependence of $\fedim$ with $\nsamplepts$ is undesirable since $\nsamplepts$ may be large, and reducing this cost is a target for future work.
+The multiplicative dependence of $\fedim$ with $\nsamplepts$ is undesirable since $\nsamplepts$ may be large, and reducing this cost is possible with more involved computational geometry methods.
 %Several improvements are possible.
@@
 This is done using a precomputed kd-tree of sample points, and requires $O(\numnbr \log \nsamplepts)$ floating point and logical elementary operations.
-Second, one must find the mesh cells that the points $\{z_i\}_{i=1}^{\numnbr}$ reside.
+Second, one must find the mesh cells that the points $\{z_i\}_{i=1}^{\numnbr}$ reside in.
 This is done using an AABB tree of mesh cells, and requires $O(\numnbr \log \fedim)$ elementary operations.
@@
 Constructing an H-matrix representation of the regularization term in the Hessian requires one matrix inversion and two matrix-matrix multiplications.
-Then, we use one matrix factorization to form a preconditioner for the Newton linear system.
+%Then, we use one matrix factorization to form a preconditioner for the Newton linear system. 
 %We perform low-rank updates to the H-matrix as the Newton iterations proceed, as per the discussion in Appendix~\ref{sec:recycle_krylov}.
-Symmetrizing $\mathbf{A}_H$ requires one matrix-matrix addition, and flipping negative eigenvalues to be positive requires a handful (typically around 5) matrix-matrix additions and matrix factorizations.
+Symmetrizing $\mathbf{A}_H$ requires one matrix-matrix addition.
+Flipping negative eigenvalues to be positive requires a handful (typically around 5) matrix-matrix additions and matrix factorizations to factor the required shifted linear systems, and a number of factorized solves with these shifted linear systems that is proportional to the number of erroneous negative eigenvalues.
 %semi-definite, described in Appendix~\ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the rational function order.
@@
 
-We use our method to approximate the Newton (or Gauss-Newton) Hessian
+We use our method to approximate the Newton (or Gauss-Newton) Hessians
 in inverse problems governed by PDEs which model steady state ice
@@
 specific setup, we refer the reader to~\cite[Section
-  2.2]{VillaPetraGhattas21}. For both examples, we choose the
-regularization parameter, $\gamma$, that controls the overall strength
-of the regularization using Morozov's discrepancy
-principle~\cite{Vogel02}. The optimal value of $\gamma$ is used for
+  2.2]{VillaPetraGhattas21}. In all numerical results we choose the
+regularization parameter (which controls the overall strength of the regularization) using Morozov's discrepancy
+principle~\cite{Vogel02}. 
+%The optimal value of $\gamma$ is used for
 %  $\gamma =  7.3 \times 10^3$ for Stokes  Figure~\ref{fig:gamma_sweep}
-all the numerical results except for the studies where we study how
-changing $\gamma$ impacts the effectiveness of the proposed
-preconditioner.
+%all the numerical results except for the studies where we study how
+%changing $\gamma$ impacts the effectiveness of the proposed
+%preconditioner.
 
@@
 subsequent iterations. The Newton iterations are terminated when
-$\|\bm{g}_k\|/\|\bm{g}_0\| < 10^{-8}$. Systems~\eqref{eq:newton_system} are solved inexactly using an inner PCG
+$\|\bm{g}_k\|/\|\bm{g}_0\| < 10^{-6}$. Systems~\eqref{eq:newton_system} are solved inexactly using an inner PCG
 iteration, which terminates when the norm of the linear system residual is less than $\min \left(0.5,
@@
 are formed by adding $\bm{A}_H$ to $\mathbf{R}_H$ using fast H-matrix arithmetic.
-The H-matrix $\widetilde{\mathbf{H}}$ may be indefinite due to approximation error, so we modify it to be symmetric positive definite via a shift and invert deflation procedure described in Appendix~\ref{sec:make_spd}.
+The H-matrix $\widetilde{\mathbf{H}}$ may be indefinite due to approximation error, so we modify it to be symmetric positive definite via a shift and invert deflation procedure described in Section~\ref{sec:flip_eigs}.
 We factor $\bm{\preconditioner}$,
@@
 %This flattening is necessary because our operator approximation method requires adding points on this surface together to get other points on the surface, and this is only well-defined in flat space.
-The observations are generated by adding $1\%$ multiplicative
+The observations are generated by adding multiplicative
 Gaussian noise to the tangential component of the velocity field
-restricted to the top surface of the geometry. The true basal sliding
+restricted to the top surface of the geometry.
+We use $5\%$ noise in all cases, except for Figure \ref{fig:stokes_reconstructions} and Table \ref{fig:condition_number} where the noise is varied from $1\%$ to $25\%$.
+The true basal sliding
 coefficient field and the true velocity field, which is obtained by
@@
     \end{subfigure}
-	\caption{The log basal sliding parameter, with colorscale as in Figure~\ref{fig:true_beta}, computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $5.0\%$ (middle),and $1.0\%$ (right).
-\todo{remove reference to newon iterates figure as this figure has a new significance}}
-	\label{fig:newton_iterates} 
+	\caption{The log basal sliding parameter, with colorscale as in Figure~\ref{fig:true_beta}, computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $5.0\%$ (middle),and $1.0\%$ (right).}
+	\label{fig:stokes_reconstructions} 
 \end{figure}
@@
 
-Table~\ref{tab:newton_convergence_table} shows the performance of our preconditioner for accelerating the solution of optimization problem~\eqref{eq:deterministic_optimization_problem}.
+Table~\ref{tab:newton_convergence_table} shows the performance of our preconditioner for accelerating the solution of the optimization problem to reconstruct $\mathbf{q}$ from observations with $5\%$ noise.
 We build the PSF (5) preconditioner in the third Gauss-Newton iteration, and reuse it for all subsequent Gauss-Newton and Newton iterations.
@@
 We compare our method with the most commonly used existing preconditioners: no preconditioning (NONE), and preconditioning by the regularization term in the Hessian (REG).
-Using PSF (5) reduces the total number of Stokes PDE solves by a factor of six compared to no preconditioning, and by a factor of nine compared to regularization preconditioning\footnote{Interestingly, we observe that no preconditioning outperforms regularization preconditioning here because the noise level (and hence $\gamma$) is small.}.
-In Figure~\ref{fig:newton_iterates} we show select Newton iterates using our PSF (5) preconditioner.
+Using PSF (5) reduces the total number of Stokes PDE solves substantially compared to regularization preconditioning and no preconditioning\footnote{Interestingly, we observe that no preconditioning outperforms regularization preconditioning.}.
+In Figure~\ref{fig:stokes_reconstructions} we show reconstructions for $1\%$, $5\%$, and $25\%$ noise.
 %
@@
 %
-%
+
 We build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\bm{\basalfriction}$.
@@
 PCG converges much slower with no preconditioning and regularization preconditioning than it does with PSF preconditioning, with no preconditioning outperforming regularization preconditioning.
-In Figure~\ref{fig:gamma_sweep}, we perform PCG solves on linear systems of the same form, except now we vary $\gamma$.
-The performance of our PSF preconditioners is good and relatively stable over a wide range of $\gamma$'s.
-All PSF preconditioners perform the same for medium and large values of $\gamma$.
-For small values of $\gamma$ PSF (25) performs slightly better than PSF (5), which performs slightly better than PSF (1).
-As expected, regularization preconditioning performs well for large $\gamma$ and poorly for small $\gamma$.
+In Figure~\ref{fig:generalized_eigenvalues} we show the generalized eigenvalues for the generalized eigenvalue problem
+$
+\bm{H} \mathbf{u} = \lambda \bm{\preconditioner} \mathbf{u}.
+$ 
+The matrix $\bm{\preconditioner}$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations constructed at the converged solution $\bm{\basalfriction}$, or the regularization Hessian.
+The PSF preconditioners cluster the eigenvalues of the Hessian near one, with more batches yielding better clustering.
+The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
+
+%In Figure~\ref{fig:condition_number} we show the condition numbers of $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$.
+The PSF preconditioners yield the the smallest condition numbers, with more batches yielding smaller condition numbers.
+Regularization preconditioning and no preconditioning yield larger condition numbers.
+The condition numbers are similar for both $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$, demonstrating that the preconditioner built based on $\bm{H}_\text{gn}$ is a good preconditioner for $\bm{H}$.
+
+In Table~\ref{tab:condition_number}, we show the condition number of the preconditioned Hessian using our preconditioners, for a variety of noise levels ranging from $1\%$ to $25\%$.
+The condition number using our PSF preconditioners is extremely small (ranging between 1 and 10) and relatively stable over this range of noise levels.
+PSF (25) performs slightly better than PSF(5), which performs better than PSF (1), as expected.
+All PSF preconditioners outperform no preconditioning and regularization preconditioning for all noise levels.
+No preconditioning outperforms regularization preconditioning for all noise levels except $25\%$.
+We report condition number results for the preconditioned Hessian, $\mathbf{H}$, demonstrating that the PSF preconditioners built based on the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, are also good preconditioner for $\bm{H}$.
+%As expected, regularization preconditioning performs well for large $\gamma$ and poorly for small $\gamma$.
 Our PSF preconditioners outperform no preconditioning and regularization preconditioning for medium and small $\gamma$, and perform similarly to regularization preconditioning for large~$\gamma$.
@@
 All operators are evaluated at the $\bm{\basalfriction}$ that solves the inverse problem for
-	the regularization value $\gamma$ as determined by the respective noise level and the Morozov discrepancy principle.}
-	\label{fig:condition_number}
+	the respective noise level.}
+	\label{tab:condition_number}
 \end{table}
-
-In Figure~\ref{fig:generalized_eigenvalues} we show the generalized eigenvalues for the generalized eigenvalue problem
-$
-	\bm{H} \mathbf{u} = \lambda \bm{\preconditioner} \mathbf{u}.
-$
-Here $\bm{H}$ is the Hessian evaluated at the reconstructed parameter $\bm{\basalfriction}$ for $\gamma$ chosen to satisfy the Morozov discrepancy principle.
-The matrix $\bm{\preconditioner}$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations constructed at that $\bm{\basalfriction}$, or the regularization Hessian.
-The PSF preconditioners cluster the eigenvalues of the Hessian near one, with more batches yielding better clustering.
-The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
-In Figure~\ref{fig:condition_number} we show the condition numbers of $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$.
-The PSF preconditioners yield the the smallest condition numbers, with more batches yielding smaller condition numbers.
-Regularization preconditioning and no preconditioning yield larger condition numbers.
-The condition numbers are similar for both $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$, demonstrating that the preconditioner built based on $\bm{H}_\text{gn}$ is a good preconditioner for $\bm{H}$.
 
@@
 Point source locations for the impulse response batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulse responses per batch as possible, while ensuring that the corresponding ellipsoids do not overlap.
-We applied the method to approximate the Gauss-Newton Hessian in a Stokes ice sheet inverse problem, and saw that the approximation substantially outperforms existing Hessian approximation methods.
+We applied the method to approximate the Gauss-Newton Hessians in a Stokes ice sheet and advection-diffusion inverse problems, and saw that the approximation substantially outperforms existing Hessian approximation methods.
 Although the method we presented is not applicable to all Hessians, it is applicable to many Hessians of practical interest.

### 2023-06-01 paper/localpsf.tex
edited writing in stokes section

@@
 restricted to the top surface of the geometry.
-We use $5\%$ noise in all cases, except for Figure \ref{fig:stokes_reconstructions} and Table \ref{fig:condition_number} where the noise is varied from $1\%$ to $25\%$.
+We use $5\%$ noise in all cases, except for Figure \ref{fig:stokes_reconstructions} and Table \ref{tab:condition_number} where the noise is varied from $1\%$ to $25\%$.
 The true basal sliding
@@
 The condition number using our PSF preconditioners is extremely small (ranging between 1 and 10) and relatively stable over this range of noise levels.
-PSF (25) performs slightly better than PSF(5), which performs better than PSF (1), as expected.
+PSF (25) performs better than PSF(5), which performs better than PSF (1), as expected.
 All PSF preconditioners outperform no preconditioning and regularization preconditioning for all noise levels.

### 2023-06-01 paper/localpsf.tex
editing numerical results section

@@
 	\caption{
-		Top: bird's eye view of the ice sheet, discretized by a mesh of tetrahedra.
+		(Top) Bird's eye view of the ice sheet, discretized by a mesh of tetrahedra.
 Color indicates the height of the base of the ice sheet (i.e., the mountain topography).
 The radius of the domain is $10^4$ meters, the maximum height of the mountain is $2.1 \times 10^3$ meters, and the average thickness of the ice sheet is 250 meters.
-		True parameter, $\basalfriction_\text{true}$, (bottom left) and true velocity $\velocity_\text{true}$ (bottom right).
-In the bottom right plot, arrows indicate the direction of $\velocity_\text{true}$ and color indicates the magnitude of $\velocity_\text{true}$.}
+		(Bottom left) True parameter, $\basalfriction_\text{true}$.
+(Bottom right) True velocity $\velocity_\text{true}$.
+Arrows indicate the direction of $\velocity_\text{true}$ and color indicates the magnitude of $\velocity_\text{true}$.}
 	\label{fig:true_beta_u}
@@
     \end{subfigure}
-	\caption{The log basal sliding parameter, with colorscale as in Figure~\ref{fig:true_beta}, computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $5.0\%$ (middle),and $1.0\%$ (right).}
+	\caption{The log basal sliding parameter, with colorscale as in Figure~\ref{fig:true_beta}, computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $5.0\%$ (middle), and $1.0\%$ (right).}
 	\label{fig:stokes_reconstructions} 
@@
 	}
-	\caption{Convergence history for solving the Stokes inverse problem using preconditioned inexact Newton PCG.
+	\caption{Convergence history for solving the Stokes inverse problem using inexact Newton PCG.
+The Newton iteration is terminated when $||\mathbf{g}_k|| < 10^{-6} ||\mathbf{g}_0||$, where $\mathbf{g}_i$ is the gradient at the $i$th iteration.
 Preconditioners shown are our method with five batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
@@
 Columns titled \#Stokes show the total number of Stokes PDE solves performed in each Newton iteration.
-This consists of Stokes solves for $u$ performed during the linesearch going from $\bm{\basalfriction}_{k-1}$ to $\bm{\basalfriction}_k$, plus one adjoint Stokes solve to compute the gradient at $\bm{\basalfriction}_k$, plus one incremental forward and one incremental adjoint Stokes solve per PCG iteration for solving the Newton system.
+This consists of Stokes solves for $u$ performed during the line search going from $\bm{\basalfriction}_{k-1}$ to $\bm{\basalfriction}_k$, plus one adjoint Stokes solve to compute the gradient at $\bm{\basalfriction}_k$, plus one incremental forward and one incremental adjoint Stokes solve per PCG iteration for solving the Newton system.
 Under PSF (5) and in row Iter 3, we write $6+22$ to indicate that $6$ Stokes solves were used during the standard course of the iteration, and $22$ Stokes solves were used to build the PSF (5) preconditioner.}
@@
 	\caption{Left (\ref{fig:cg_convergence}) shows the convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
-random entries drawn from the standard normal distribution.
-The reference solution $\bm{x}$ is determined by solving $\bm{H} \bm{x} = \bm{b}$ to an accuracy of $10^{-12}$ using method PCG.
-	Here we use $\gamma=7.3 \times 10^3$, which is determined by the Morozov discrepancy principle when using $5\%$ noise.
+random entries drawn from the standard normal distribution and $\mathbf{H}$ is evaluated at the solution of the inverse problem.
+%	The reference solution $\bm{x}$ is determined by solving $\bm{H} \bm{x} = \bm{b}$. 
+%	Here we use $\gamma=7.3 \times 10^3$, which is determined by the Morozov discrepancy principle when using $5\%$ noise.
 	Results in these figures are shown for our preconditioner with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
@@
 The condition number using our PSF preconditioners is extremely small (ranging between 1 and 10) and relatively stable over this range of noise levels.
-PSF (25) performs better than PSF(5), which performs better than PSF (1), as expected.
+PSF (25) performs better than PSF (5), which performs better than PSF (1), as expected.
 All PSF preconditioners outperform no preconditioning and regularization preconditioning for all noise levels.
 No preconditioning outperforms regularization preconditioning for all noise levels except $25\%$.
-We report condition number results for the preconditioned Hessian, $\mathbf{H}$, demonstrating that the PSF preconditioners built based on the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, are also good preconditioner for $\bm{H}$.
+%We report condition number results for the preconditioned Hessian, $\mathbf{H}$, demonstrating that the PSF preconditioners built based on the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, are also good preconditioner for $\bm{H}$.
 %As expected, regularization preconditioning performs well for large $\gamma$ and poorly for small $\gamma$.
@@
 	\centering
+	{
+%	\setlength{\tabcolsep}{0.1em}
+%	\renewcommand{\arraystretch}{2.0}
 	\begin{tabular}{cccc}
@@
 	\end{tabular}
+	}
 	\caption{Impulse responses for a variety of different diffusion parameters $\kappa$ and simulation times $T$.}

### 2023-06-01 paper/localpsf.tex
editing numerical results section

@@
 \begin{table}
+	\centering
+	{
+		%	\setlength{\tabcolsep}{0.1em}
+		%	\renewcommand{\arraystretch}{2.0}
+		\begin{tabular}{cccc}
+			& $T=0.5$ & $T=1.0$ & $T=2.0$ \\
+			$\kappa=1.0e{-4}$  & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-4.png} \\
+			$\kappa=3.2e{-4}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K3.2e-4.png} \\
+			$\kappa=1.0e{-3}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-3.png}
+		\end{tabular}
+	}
+	\caption{Impulse responses for a variety of different diffusion parameters $\kappa$ and simulation times $T$.}
+	\label{tbl:adv_vary_kappa_tf}
+\end{table}
+
+\begin{table}
 	\begin{center}
@@
 \end{table}
-
-\begin{table}
-	\centering
-	{
-%	\setlength{\tabcolsep}{0.1em}
-%	\renewcommand{\arraystretch}{2.0}
-	\begin{tabular}{cccc}
-		  & $T=0.5$ & $T=1.0$ & $T=2.0$ \\
-		$\kappa=1.0e{-4}$  & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-4.png} \\
-		$\kappa=3.2e{-4}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K3.2e-4.png} \\
-		$\kappa=1.0e{-3}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-3.png}
-	\end{tabular}
-	}
-	\caption{Impulse responses for a variety of different diffusion parameters $\kappa$ and simulation times $T$.}
-	\label{tbl:table_of_figures}
-\end{table}
-
-\begin{figure}
-	\caption{Impulse responses for various diffusivities kappa and fixed simulation time Tf=0.5 (top) and simulation times with fixed $\kappa=1e{-3}$ (bot).}
-	\label{fig:adv_vary_kappa_tf}
-\end{figure}
 
@@
 
-\begin{itemize}
-	\item Picture of velocity field, initial condition, final condition, reconstruction (1) (OKOK)
-	\item Impulse responses: vary time and Peclet number (3) (OKOK)
-	\item CG convergence (1) CG iters vs Peclet number (Morozov) (2) (OKOK)
-	\item Preconditioned spectrum, condition numbers? (1)
-\end{itemize}
+%\begin{itemize}
+%	\item Picture of velocity field, initial condition, final condition, reconstruction (1) (OKOK)
+%	\item Impulse responses: vary time and Peclet number (3) (OKOK)
+%	\item CG convergence (1) CG iters vs Peclet number (Morozov) (2) (OKOK)
+%	\item Preconditioned spectrum, condition numbers? (1)
+%\end{itemize}
 

### 2023-06-02 paper/stokes_pcg_convergence.tex
added new data for Stokes pcg convergence and generalized eigenvalues... updated the mini tex files which generate the associated tikz figures... this is for data in e6 Stokes data directory which indicates that the inverse problem was solved to a looser tolerance of requiring a 6 orders of magnitude in the reduction of the norm of the gradient

@@
 	\addlegendentry{Reg}		
-	\addplot[dotted, line width=0.5mm] table {cg_relres_reg_0.05noise.dat};
+	\addplot[dotted, line width=0.5mm] table {Stokes_cg_rand_errs_reg_0.05noise.dat};
 	\addlegendentry{None}
-	\addplot[dash dot, line width=0.5mm] table {cg_relres_none_0.05noise.dat};
+	\addplot[dash dot, line width=0.5mm] table {Stokes_cg_rand_errs_none_0.05noise.dat};
 	\addlegendentry{PSF (1)}
-	\addplot[blue, line width=0.5mm] table {cg_relres_psf1_0.05noise.dat};
+	\addplot[blue, line width=0.5mm] table {Stokes_cg_rand_errs_psf1_0.05noise.dat};
 	\addlegendentry{PSF (5)}
-	\addplot[red, line width=0.5mm] table {cg_relres_psf5_0.05noise.dat};
+	\addplot[red, line width=0.5mm] table {Stokes_cg_rand_errs_psf5_0.05noise.dat};
 	\addlegendentry{PSF (25)}
-	\addplot[purple, line width=0.5mm] table {cg_relres_psf25_0.05noise.dat};
+	\addplot[purple, line width=0.5mm] table {Stokes_cg_rand_errs_psf25_0.05noise.dat};
 	\end{axis}

### 2023-06-02 paper/localpsf.tex
passable complete draft

@@
 		\caption{$\velocity_\text{true}$.}
+		\label{fig:stokes_velocity}
 	\end{subfigure}
@@
 	\caption{
-		(Top) Bird's eye view of the ice sheet, discretized by a mesh of tetrahedra.
+		(\ref{fig:ice_mountain_mesh}) Bird's eye view of the ice sheet, discretized by a mesh of tetrahedra.
 Color indicates the height of the base of the ice sheet (i.e., the mountain topography).
 The radius of the domain is $10^4$ meters, the maximum height of the mountain is $2.1 \times 10^3$ meters, and the average thickness of the ice sheet is 250 meters.
-		(Bottom left) True parameter, $\basalfriction_\text{true}$.
-(Bottom right) True velocity $\velocity_\text{true}$.
+		(\ref{fig:true_beta}) True parameter, $\basalfriction_\text{true}$.
+(\ref{fig:stokes_velocity}) True velocity $\velocity_\text{true}$.
 Arrows indicate the direction of $\velocity_\text{true}$ and color indicates the magnitude of $\velocity_\text{true}$.}
@@
 	Right (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$.
-Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), or the regularization Hessian (REG).}
+Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), the regularization Hessian (REG), or the identity matrix (NONE).}
 	\label{fig:krylov_convergence}
@@
 We build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\bm{\basalfriction}$.
-We use PCG to solve a linear system with the Hessian as the coefficient operator and a right hand vector with random i.i.d.
-entries drawn from the standard normal distribution.
+We use PCG to solve a linear system with the Hessian as the coefficient operator and a right hand vector with random independent and identically distributed (i.i.d.) entries drawn from the standard normal distribution.
 In Figure~\ref{fig:cg_convergence} we compare the convergence of PCG for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners.
@@
 Here, $\kappa > 0$ is a diffusivity coefficient, $\nu$ is the boundary unit normal vector, and $T > 0$ is the
-final time.  The velocity field $v:\advdomain \rightarrow \mathbb{R}^d$, shown in
-Figure~\ref{fig:adv_inversion1}, is computed by solving the
+final time.  The velocity field, $v:\advdomain \rightarrow \mathbb{R}^d$, is computed by solving the
 steady-state Navier-Stokes equations for a two dimensional flow with
-Reynolds number 50 and boundary conditions as in
-Figure~\ref{fig:adv_inversion} (left); see~\cite{PetraStadler11} for
-details.  The time evolution of the state variable $\concentration$ from a given
-initial condition $\ipar$ is illustrated in
-Figure~\ref{fig:adv_inversion1}.
+Reynolds number 50, with boundary conditions $v(x)=(0,1)$ on the left boundary, $v(x)=(0,-1)$ on the right boundary, and $v(x)=(0,0)$ on the top and bottom boundaries, as in~\cite[Section 3]{PetraStadler11}.
+In the numerical results we use a checkerboard image for the initial condition and add $5\%$ multiplicative noise to the observations at the final time, $T$.
+We use $\kappa=3.2e{-1}$ and $T=1.0$ for all results, except for Table \ref{tab:adv_convergence} and Figure \ref{tbl:adv_vary_kappa_tf} where we vary $\kappa$ and $T$ to study the impact of these parameters on the impulse responses and our preconditioner's effectiveness.
+The initial condition, velocity field, noisy observations, and reconstructed initial condition are shown in Figure~\ref{fig:adv_inversion1}.
+
+In Table \ref{tab:adv_convergence} we show the number of PCG iterations required to solve the Newton linear system to a relative error tolerance of $10^{-6}$.
+We show results for $T$ ranging from $0.5$ to $2.0$ and $\kappa$ ranging from $10^{-4}$ to $10^{-3}$, using our PSF preconditioner with $1$, $5$, and $25$ batches, regularization preconditioning, and no preconditioning.
+PSF preconditioning performs better than regularization preconditioning and no preconditioning, with more batches yielding better results.
+The PSF preconditioners perform best in the high-rank regime where $T$ and $\kappa$ are small, and the observations are highly informative about the initial condition.
+This makes sense given the effect of $T$ and $\kappa$ on the shape of the impulse responses, as shown in Figure \ref{tbl:adv_vary_kappa_tf}.
+As $\kappa$ increases, the impulse responses grow wider and as $T$ increases the impulse responses grow longer.
+This means that the impulse responses become more local as $\kappa$ and $T$ decrease, so more impulse responses can fit within each batch.
+
+In Figure \ref{fig:adv_cg_convergence} we show the convergence of PCG for solving $\mathbf{H}\mathbf{x}=\mathbf{b}$, where $\mathbf{b}$ has i.i.d.
+random entries drawn from a standard normal distribution.
+The preconditioners used, $\widetilde{H}$, are our PSF preconditioner with $1$, $5$, or $25$ batches, regularization preconditioning, and no preconditioning.
+The PSF preconditioners converge fastest, with more batches yielding faster convergence.
+In Figure \ref{fig:adv_generalized_eigenvalues}, we show the eigenvalues for the generalized eigenvalue problem $\mathbf{H}\mathbf{u}=\lambda \widetilde{\mathbf{H}}\mathbf{u}$, where the $\widetilde{H}$ are the preconditioners stated above.
+The PSF preconditioners cluster the eigenvalues of the Hessian near one, and the regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
+For the PSF preconditioner, more batches yields better clustering.
 
@@
 \begin{figure}
+	\begin{center}
 	\begin{subfigure}{0.24\textwidth}
+		\begin{center}
 		\includegraphics[scale=0.14]{true_initial_condition.png}
-		%		\caption{}
-		%		\label{fig:newtoniter2}
+	\end{center}
+				\caption{True $q$}
+				\label{fig:adv_initial_condition}
 	\end{subfigure}
 	\begin{subfigure}{0.24\textwidth}
+		\begin{center}
 		\includegraphics[scale=0.14]{wind_velocity_colorscheme3.png}
-		%		\caption{}
-		%		\label{fig:newtoniter1}
+		\end{center}
+				\caption{$v$}
+				\label{fig:adv_velocity}
 	\end{subfigure}
 	\begin{subfigure}{0.24\textwidth}
+		\begin{center}
 		\includegraphics[scale=0.14]{noisy_observations_T1.0_K3.2e-4.png}
-		%		\caption{}
-		%		\label{fig:newtoniter3}
+		\end{center}
+				\caption{Noisy observations}
+				\label{fig:adv_observations}
 	\end{subfigure}
 	\begin{subfigure}{0.24\textwidth}
+		\begin{center}
 		\includegraphics[scale=0.14]{reconstructed_initial_condition_T1.0_K3.2e-4.png}
-		%		\caption{}
-		%		\label{fig:newtoniter4}
+		\end{center}
+				\caption{Reconstructed $q$}
+				\label{fig:adv_reconstruction}
 	\end{subfigure}
+	\end{center}
 	%\begin{subfigure}{0.05\textwidth}
@@
 	%\end{subfigure}
-	\caption{(Left) True initial condition.
-(Middle left) Velocity field.
+	\caption{(\ref{fig:adv_initial_condition}) True initial condition.
+(\ref{fig:adv_velocity}) Velocity field.
 Color indicates magnitude of velocity vector.
-(Middle right) Noisy observations of concentration at final time.
-(Right) Reconstructed initial condition.}
+(\ref{fig:adv_observations}) Noisy observations of concentration at final time.
+(\ref{fig:adv_reconstruction}) Reconstructed initial condition.}
 	\label{fig:adv_inversion1} 
 \end{figure}
-
-\begin{table}
-	\centering
-	{
-		%	\setlength{\tabcolsep}{0.1em}
-		%	\renewcommand{\arraystretch}{2.0}
-		\begin{tabular}{cccc}
-			& $T=0.5$ & $T=1.0$ & $T=2.0$ \\
-			$\kappa=1.0e{-4}$  & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-4.png} \\
-			$\kappa=3.2e{-4}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K3.2e-4.png} \\
-			$\kappa=1.0e{-3}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-3.png}
-		\end{tabular}
-	}
-	\caption{Impulse responses for a variety of different diffusion parameters $\kappa$ and simulation times $T$.}
-	\label{tbl:adv_vary_kappa_tf}
-\end{table}
 
@@
 \end{table}
+
+\begin{figure}
+	\centering
+	{
+		%	\setlength{\tabcolsep}{0.1em}
+		%	\renewcommand{\arraystretch}{2.0}
+		\begin{tabular}{cccc}
+			& $T=0.5$ & $T=1.0$ & $T=2.0$ \\
+			$\kappa=1.0e{-4}$  & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-4.png} \\
+			$\kappa=3.2e{-4}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K3.2e-4.png} \\
+			$\kappa=1.0e{-3}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-3.png}
+		\end{tabular}
+	}
+	\caption{Impulse responses for a variety of different diffusion parameters $\kappa$ and simulation times $T$.}
+	\label{tbl:adv_vary_kappa_tf}
+\end{figure}
 
@@
 random entries drawn from the standard normal distribution.
-BLAH BLAH BLAH
-		%		The y-axis is $\|b-Hx_k\|/\|b\|$, where $x_k$ is the $k$th PCG iterate. 
-		Here we use $\gamma=7.88 \times 10^{-6}$, which is determined by the Morozov discrepancy principle.
-Results in these figures are shown for our preconditioner with 5, 15, and 30 batches (PSF (5), PSF (15), and PSF(30), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
-The preconditioner is constructed using $\bm{H}_\text{gn}$.}
+Right (\ref{fig:adv_generalized_eigenvalues}) shows eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$.
+Here $\bm{H}$ is the Hessian and the preconditioner $\bm{\preconditioner}$ is the PSF approximation for 1, 5, or 25 batches (PSF (1), PSF (5), and PSF (25), respectively), the regularization Hessian (REG), or the identity matrix (NONE).}
 	\label{fig:adv_krylov_convergence}
@@
 The idea of the method is to compute batches of impulse responses by applying the operator to Dirac combs of scattered point sources, then interpolate these impulse responses to approximate entries of the operator's integral kernel.
-The interpolation is based on a new principle we call ``local mean displacement invariance,'' which generalizes and improves upon classical local translation invariance.
+The interpolation is based on a new principle we call ``local mean displacement invariance,'' which generalizes classical local translation invariance.
 The ability to quickly approximate arbitrary integral kernel entries permits us to form an H-matrix approximation of the operator.
@@
 Point source locations for the impulse response batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulse responses per batch as possible, while ensuring that the corresponding ellipsoids do not overlap.
-We applied the method to approximate the Gauss-Newton Hessians in a Stokes ice sheet and advection-diffusion inverse problems, and saw that the approximation substantially outperforms existing Hessian approximation methods.
+We applied the method to approximate the Gauss-Newton Hessians in Stokes ice sheet and advection-diffusion inverse problems, and saw that the approximation substantially outperforms existing Hessian approximation methods.
 Although the method we presented is not applicable to all Hessians, it is applicable to many Hessians of practical interest.

### 2023-06-02 paper/localpsf.tex
in the middle of editing paper

@@
 Specifically, we approximate $\Aker(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}(y - \spatialmean(x) + \spatialmean(x_i))$ for a small number of sample points $x_i$ near $x$.
-The weights are determined by radial basis function interpolation.
+The weights are determined by radial basis function (RBF) interpolation.
 
@@
 If $\candidatepoint_i$ passes these checks 
-%(i.e., if $\candidatepoint_i$ is sufficiently far from other points in the batch, and $\spatialvol(\candidatepoint_i)$ is not too small, and the )
+(i.e., if $\candidatepoint_i$ is sufficiently far from other points in the batch, and $\spatialvol(\candidatepoint_i)$ and $\spatialcov(\candidatepoint_i)$ are acceptable)
 then we add $\candidatepoint_i$ to $\pointbatch_k$. Otherwise we discard $\candidatepoint_i$. This process repeats until there are no more points in $\candidatepts$.  
 %This is detailed in Algorithm~\ref{alg:point_choice}. 
-%We check whether $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ using the fast ellipsoid intersection test in \cite{GilitschenskiHanebeck12}.
-described in Appendix~\ref{sec:fast_ellipsoid_intersection_test}. 
+We check whether $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ using the fast ellipsoid intersection test in \cite{GilitschenskiHanebeck12}.
+%described in Appendix~\ref{sec:fast_ellipsoid_intersection_test}. 
 We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots, \pointbatch_{\nbatch}$.
@@
 	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[scale=0.13]{mstar_0.25noise_cropped.png}
+		\includegraphics[scale=0.20]{mstar_0.25noise_cropped.png}
 		%		\caption{}
@@
 	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[scale=0.13]{mstar_0.05noise_cropped.png}
+		\includegraphics[scale=0.20]{mstar_0.05noise_cropped.png}
 		%		\caption{}
@@
 	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.13]{mstar_0.01noise_cropped.png}
+	\includegraphics[scale=0.20]{mstar_0.01noise_cropped.png}
 	%		\caption{}
@@
 
-\appendix
+%\appendix
 
@@
 
-\section{Ellipsoid intersection test}
-\label{sec:fast_ellipsoid_intersection_test}
-The procedure for choosing sample points relies on quickly determining whether two ellipsoids intersect. Let $E_1$ and $E_2$ be the ellipsoids defined as
-\begin{align*}
-	E_i :=& \{x : (x - \spatialmean_i)^T \spatialcov_i^{-1} (x - \spatialmean_i) \le \tau^2\}
-\end{align*}
-for $i\in\{1,2\}$, where $\spatialmean_1, \spatialmean_2 \in \mathbb{R}^\gdim$, and $\spatialcov_1, \spatialcov_2 \in \mathbb{R}^{\gdim \times \gdim}$ are positive definite.
+%\section{Ellipsoid intersection test}
+%\label{sec:fast_ellipsoid_intersection_test}
+%The procedure for choosing sample points relies on quickly determining whether two ellipsoids intersect. Let $E_1$ and $E_2$ be the ellipsoids defined as
+%\begin{align*}
+%	E_i :=& \{x : (x - \spatialmean_i)^T \spatialcov_i^{-1} (x - \spatialmean_i) \le \tau^2\}
+%\end{align*}
+%for $i\in\{1,2\}$, where $\spatialmean_1, \spatialmean_2 \in \mathbb{R}^\gdim$, and $\spatialcov_1, \spatialcov_2 \in \mathbb{R}^{\gdim \times \gdim}$ are positive definite.
 Let $K$ be the following one dimensional convex function:
-\begin{equation*}
-	K(s) := 1 - \frac{1}{\tau^2} (\spatialmean_1 - \spatialmean_2)^T \left(\frac{1}{1-s}\spatialcov_1 + \frac{1}{s}\spatialcov_2\right)^{-1}(\spatialmean_1 - \spatialmean_2).
-\end{equation*}
-In~\cite{GilitschenskiHanebeck12} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$.
-% (see Figure~\ref{fig:ellipsoid_intersection_test}).
-
-%\begin{figure}
-%	\begin{subfigure}{0.37\textwidth}
-%		\centering
-%		\includegraphics[scale=1.0]{ellipsoids_intersect0.pdf}
-%	\end{subfigure}
-%	\begin{subfigure}{0.30\textwidth}
-%		\centering
-%		\includegraphics[scale=1.0]{ellipsoids_intersect1.pdf}
-%	\end{subfigure}
-%	\begin{subfigure}{0.30\textwidth}
-%		\centering
-%		\includegraphics[scale=1.0]{ellipsoids_intersect2.pdf}
-%	\end{subfigure}
-%	\caption{The ellipsoids intersect if $K(s) \ge 0$ for all $s \in (0,1)$.
-%	}
-%	\label{fig:ellipsoid_intersection_test}
-%\end{figure}
-
-The function $K(s)$ may be evaluated quickly for many $s$ by computing the solution to the generalized eigenvalue problem
-$
-	\spatialcov_1 \eigenvectormatrix = \spatialcov_2 \eigenvectormatrix \Lambda,
-$
-where $\eigenvectormatrix \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors, and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$.
-%The matrix $\eigenvectormatrix$ simultaneously diagonalizes $\spatialcov_1$ and $\spatialcov_2$, in the sense that $\eigenvectormatrix^T\spatialcov_1 \eigenvectormatrix = \Lambda$, and $\eigenvectormatrix^T\spatialcov_2 \eigenvectormatrix = I$, where $I$ is the $\gdim \times \gdim$ identity matrix.
+%\begin{equation*}
+%	K(s) := 1 - \frac{1}{\tau^2} (\spatialmean_1 - \spatialmean_2)^T \left(\frac{1}{1-s}\spatialcov_1 + \frac{1}{s}\spatialcov_2\right)^{-1}(\spatialmean_1 - \spatialmean_2).
+%\end{equation*}
+%In~\cite{GilitschenskiHanebeck12} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$.
+%% (see Figure~\ref{fig:ellipsoid_intersection_test}). 
+%
+%%\begin{figure}
+%%	\begin{subfigure}{0.37\textwidth}
+%%		\centering
+%%		\includegraphics[scale=1.0]{ellipsoids_intersect0.pdf}
+%%	\end{subfigure}
+%%	\begin{subfigure}{0.30\textwidth}
+%%		\centering
+%%		\includegraphics[scale=1.0]{ellipsoids_intersect1.pdf}
+%%	\end{subfigure}
+%%	\begin{subfigure}{0.30\textwidth}
+%%		\centering
+%%		\includegraphics[scale=1.0]{ellipsoids_intersect2.pdf}
+%%	\end{subfigure}
+%%	\caption{The ellipsoids intersect if $K(s) \ge 0$ for all $s \in (0,1)$.
+%%	}
+%%	\label{fig:ellipsoid_intersection_test}
+%%\end{figure}
+%
+%The function $K(s)$ may be evaluated quickly for many $s$ by computing the solution to the generalized eigenvalue problem
+%$
+%	\spatialcov_1 \eigenvectormatrix = \spatialcov_2 \eigenvectormatrix \Lambda,
+%$
+%where $\eigenvectormatrix \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors, and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$.
+%%The matrix $\eigenvectormatrix$ simultaneously diagonalizes $\spatialcov_1$ and $\spatialcov_2$, in the sense that $\eigenvectormatrix^T\spatialcov_1 \eigenvectormatrix = \Lambda$, and $\eigenvectormatrix^T\spatialcov_2 \eigenvectormatrix = I$, where $I$ is the $\gdim \times \gdim$ identity matrix.
 Using this diagonalization, and
-After some algebraic manipulations, we have
-\begin{equation}
-	\label{eq:Ks_generalized}
-	K(s) = 1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2,
-\end{equation}
-where $v := \eigenvectormatrix^T\left(\spatialmean_1 - \spatialmean_2\right)$.
+%After some algebraic manipulations, we have
+%\begin{equation}
+%	\label{eq:Ks_generalized}
+%	K(s) = 1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2,
+%\end{equation}
+%where $v := \eigenvectormatrix^T\left(\spatialmean_1 - \spatialmean_2\right)$.
 To check whether $E_1$ and $E_2$ intersect, we compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, form $v$, then minimize $K(s)$ in the form~\eqref{eq:Ks_generalized} on $(0,1)$ using Brent's method~\cite{Brent71}.
@@
 Otherwise $E_1 \cap E_2 \neq \{\}$.
-%The resulting algorithm for checking whether $E_1$ and $E_2$ intersect is summarized in Algorithm~\ref{alg:ellipsoid_intersection_test}.
-
-%\begin{algorithm2e}
-%	\SetAlgoNoLine
-%	\SetKwInOut{Input}{Input}
-%	\SetKwInOut{Output}{Output}
-%	\SetKwProg{Fn}{Function}{}{}
-%	\Input{Ellipsoid $E_1$ with mean $\spatialmean_1$ and covariance $\spatialcov_1/\tau^2$\\
-%		Ellipsoid $E_2$ with mean $\spatialmean_2$ and covariance $\spatialcov_2/\tau^2$ 
-%	}
-%	
-%	\Output{Boolean $\text{ellipsoids\_intersect}$ which is true if $E_1 \cap E_2 \neq \{\}$ and false otherwise}
-%	
-%	
-%	Solve generalized eigenvalue problem $\spatialcov_1 P = \spatialcov_2 P \Lambda$
-%		
-%	$v \gets P^T\left(\spatialmean_1 - \spatialmean_2\right)$
-%		
-%	$\displaystyle K^* \gets \min_{s \in (0,1)}~1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2$
-%		
-%	\If{$K^* < 0$}{
-%			
-%		$\text{ellipsoids\_intersect} \gets \text{False}$
-%			
-%	}
-%	\Else{
-%			
-%		$\text{ellipsoids\_intersect} \gets \text{True}$
-%			
-%	}
-%	\caption{Determining whether two ellipsoids intersect}
-%	\label{alg:ellipsoid_intersection_test}
-%\end{algorithm2e}
-
-%\subsection{Ellipsoid bounding box test}
-
-This ellipsoid intersection test is only performed if the coordinate axis aligned bounding boxes for the ellipsoids intersect. The smallest coordinate axis aligned box that contains $E_i$ is the box
-$
-	\prod_{k=1}^\gdim \left[ \spatialmean_i^{k} -\tau\sqrt{\Sigma_i^{kk}}, \spatialmean_i^{k} +\tau\sqrt{\Sigma_i^{kk}} \right],
-$
-where $\spatialmean_i^{k}$ is the $k$th entry of $\spatialmean_i$, and $\Sigma_i^{kk}$ is the $k$th diagonal entry of $\Sigma_i$\footnote{See \url{https://math.stackexchange.com/q/3928964}}. The boxes
-$\prod_{k=1}^\gdim \left[a_{-}^k, a_{+}^k\right]$ and $\prod_{k=1}^\gdim \left[b_{-}^k, b_{+}^k\right]$
-intersect if and only if $a_{-}^k \le b_{+}^k$ and $b_{-}^k \le a_{+}^k$ for $k=1,\dots,d$.
+%%The resulting algorithm for checking whether $E_1$ and $E_2$ intersect is summarized in Algorithm~\ref{alg:ellipsoid_intersection_test}.
+%
+%%\begin{algorithm2e}
+%%	\SetAlgoNoLine
+%%	\SetKwInOut{Input}{Input}
+%%	\SetKwInOut{Output}{Output}
+%%	\SetKwProg{Fn}{Function}{}{}
+%%	\Input{Ellipsoid $E_1$ with mean $\spatialmean_1$ and covariance $\spatialcov_1/\tau^2$\\
+%%		Ellipsoid $E_2$ with mean $\spatialmean_2$ and covariance $\spatialcov_2/\tau^2$ 
+%%	}
+%%	
+%%	\Output{Boolean $\text{ellipsoids\_intersect}$ which is true if $E_1 \cap E_2 \neq \{\}$ and false otherwise}
+%%	
+%%	
+%%	Solve generalized eigenvalue problem $\spatialcov_1 P = \spatialcov_2 P \Lambda$
+%%		
+%%	$v \gets P^T\left(\spatialmean_1 - \spatialmean_2\right)$
+%%		
+%%	$\displaystyle K^* \gets \min_{s \in (0,1)}~1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2$
+%%		
+%%	\If{$K^* < 0$}{
+%%			
+%%		$\text{ellipsoids\_intersect} \gets \text{False}$
+%%			
+%%	}
+%%	\Else{
+%%			
+%%		$\text{ellipsoids\_intersect} \gets \text{True}$
+%%			
+%%	}
+%%	\caption{Determining whether two ellipsoids intersect}
+%%	\label{alg:ellipsoid_intersection_test}
+%%\end{algorithm2e}
+%
+%%\subsection{Ellipsoid bounding box test}
+%
+%This ellipsoid intersection test is only performed if the coordinate axis aligned bounding boxes for the ellipsoids intersect. The smallest coordinate axis aligned box that contains $E_i$ is the box
+%$
+%	\prod_{k=1}^\gdim \left[ \spatialmean_i^{k} -\tau\sqrt{\Sigma_i^{kk}}, \spatialmean_i^{k} +\tau\sqrt{\Sigma_i^{kk}} \right],
+%$
+%where $\spatialmean_i^{k}$ is the $k$th entry of $\spatialmean_i$, and $\Sigma_i^{kk}$ is the $k$th diagonal entry of $\Sigma_i$\footnote{See \url{https://math.stackexchange.com/q/3928964}}.
+The boxes
+%$\prod_{k=1}^\gdim \left[a_{-}^k, a_{+}^k\right]$ and $\prod_{k=1}^\gdim \left[b_{-}^k, b_{+}^k\right]$
+%intersect if and only if $a_{-}^k \le b_{+}^k$ and $b_{-}^k \le a_{+}^k$ for $k=1,\dots,d$.
 

### 2023-06-03 paper/localpsf.tex
one editing pass complete

@@
 Support ellipsoids are estimated a-priori based on impulse response moments, which are computed by applying the operator to a small number of polynomial functions.
-The interpolation uses a ``local mean displacement invariance'' approximation, which improves upon classical local translation invariance.
+The interpolation uses a ``local mean displacement invariance'' approximation, which generalizes classical local translation invariance.
 The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
@@
 \begin{AMS}
- 	35R30, 41A05, 41A35, 41A36, 42A85, 47A52, 47A58, 49K20, 65D12, 65F08, 65F10, 65N20, 65N21, 86A22, 86A40
+ 	35R30, 41A35, 47A52, 47J06, 65D12, 65F08, 65F10, 65K10, 65N21, 86A22, 86A40
 \end{AMS}
@@
 
-We present an efficient \emph{matrix-free} point spread function (PSF) method for approximating operators, $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$, that have locally supported non-negative integral kernels.
+We present an efficient \emph{matrix-free} point spread function (PSF) method for approximating operators $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$ that have locally supported non-negative integral kernels.
 Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, and $L^2(\Omega)'$ is the space of real-valued continuous linear functionals on $L^2(\Omega)$.
@@
 via a black box computational procedure, 
-but we cannot easily access entries of $\Aop$'s integral kernel. Evaluating the maps in~\eqref{eq:eval_maps} may require, as a sub procedure, solving PDEs or performing other costly computations.
+but we cannot easily access entries of $\Aop$'s integral kernel.
+Evaluating the maps in~\eqref{eq:eval_maps} may require, as a sub procedure, solving PDEs or performing other costly computations.
+By ``non-negative integral kernel,'' we mean that entries of the integral kernel are non-negative numbers; this is not the same as positive semi-definiteness of $\Aop$.
 
@@
 	\end{subfigure}
-	\caption{Batches, $\eta_b$, of normalized impulses, $\impulseresponse_{x}$, for the Stokes inverse problem data misfit Gauss-Newton Hessian (Section~\ref{sec:numerical_results}). 
+	\caption{Batches, $\eta_b$, of normalized impulses, $\impulseresponse_{x}$, for the ice sheet inverse problem data misfit Gauss-Newton Hessian described in Section~\ref{sec:numerical_results}. 
 %	Brightness scale differs between the subfigures. 
@@
 \end{equation}
-Broadly, our method will construct an H-matrix approximation of $\mathbf{A}$ by forming H-matrix approximations of $\boldsymbol{\Aker}$ and $\massmatrix$ (or a lumped mass version of $\massmatrix$), then multiplying these matrices per~\eqref{eq:boldA} using H-matrix methods.
+Broadly, our method will construct an H-matrix approximation of $\mathbf{A}$ by forming a H-matrix approximation of $\boldsymbol{\Aker}$, then multiplying $\boldsymbol{\Aker}$ by $\massmatrix$ (or a lumped mass version of $\massmatrix$) on the left and right using H-matrix methods.
 Classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Aker}_{ij}$, but these matrix entries are not easily accessible.
@@
 However, entries of $\Aker_h$, involve inverse mass matrices, which typically contain negative numbers.
-If there are too many negative numbers, or if the negative numbers are too large, our algorithm will be less robust.
+If there are too many negative numbers, or if the negative numbers are too large, our algorithm will yield less accurate approximations and may potentially fail.
 We therefore recommend replacing the mass matrix, $\massmatrix$, with a positive diagonal \emph{lumped mass} approximation.
-In our numerical results, we use the lumped mass matrix constructed by replacing off-diagonal entries of the mass matrix by zero. Other mass lumping techniques may be used~\cite{Hughes12}.
+In our numerical results, we use the lumped mass matrix in which the $i$th diagonal entry of the lumped mass matrix is the sum of all entries in the $i$th row of the mass matrix.
+Other mass lumping techniques may be used~\cite{Hughes12}.
 
@@
 This may be motivated by analogy to matrices.
-If $\mathbf{B}\in \mathbb{R}^{\fedim \times \fedim}$ is a matrix with $i^\text{th}$ column $\mathbf{b}_i$ and $\mathbf{w} \in \mathbb{R}^\fedim$, then
+If $\mathbf{A}\in \mathbb{R}^{\fedim \times \fedim}$ is a matrix with $i^\text{th}$ column $\mathbf{a}_i$ and $\mathbf{w} \in \mathbb{R}^\fedim$, then
 \begin{equation*}
@@
 But approximation~\eqref{eq:translate3} can hold in situations where neither~\eqref{eq:translate1} nor~\eqref{eq:translate2} holds.
-For example, because the expected value commutes with affine transformations,~\eqref{eq:translate3} will hold when $\Aop$ is locally translation invariant with respect to a translated frame of reference, while~\eqref{eq:translate2} will not hold in this case.
+For example, because the expected value commutes with affine transformations,~\eqref{eq:translate3} will hold when $\Aop$ is locally translation invariant with respect to a translated and rotated frame of reference, while~\eqref{eq:translate2} will not hold in this case.
 Additionally,~\eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions.
@@
 The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support ellipsoid for $\impulseresponse_{x_i}$ and the support ellipsoid for $\impulseresponse_{x_j}$ do not overlap if $i \neq j$ (Section~\ref{sec:sample_point_selection}).
-Because these supports do not overlap (or do not overlap much), we can post process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section~\ref{sec:get_impulse_response}).
+Because these supports do not overlap (or do not overlap much), we can post process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$.
+With one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section~\ref{sec:get_impulse_response}).
 The process is repeated until a desired number of batches is reached.
@@
 Negative integral kernel entries due to discretization error can cause $\spatialcov(\candidatepoint_i)$ to be indefinite or highly ill-conditioned.
-We reject these points because it is likely that the set $E_{\candidatepoint_i}$ poorly reflects the support of $\phi_{\candidatepoint_i}$.
 \end{enumerate}
@@
 %This is detailed in Algorithm~\ref{alg:point_choice}. 
-We check whether $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ using the fast ellipsoid intersection test in \cite{GilitschenskiHanebeck12}.
+We repeat the point selection process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots, \pointbatch_{\nbatch}$.
+For each batch, $\candidatepts$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, except for points in previous batches.
+
+We check whether $E_{\candidatepoint_i} \cap E_{\candidatepoint_j} = \{\}$ in a two stage process.
+First, we check whether the axis aligned bounding boxes for the ellipsoids intersect.
+This quickly rules out intersections of ellipsoids that are far apart.
+Second, if the bounding boxes intersect, we check if the ellipsoids intersect using the ellipsoid intersection test in \cite{GilitschenskiHanebeck12}.
 %described in Appendix~\ref{sec:fast_ellipsoid_intersection_test}.
-We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots, \pointbatch_{\nbatch}$.
-For each batch, $\candidatepts$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, except for points in previous batches.
 
@@
 Here $\hrank$ is the rank of the highest rank block in the H-matrix.
-A dense representation of $\mathbf{\Aker}$ is never formed. 
+The dense matrix $\mathbf{\Aker}$ is never formed. 
 %We describe this H-matrix construction process in detail in Appendix~\ref{app:h_matrix}. 
@@
 Choosing $\tau=0$ would remove all erroneous negative eigenvalues.
-However, this is computationally infeasible if $\Aop$ has a large or infinite cluster of eigenvalues near zero, a common situation in practice.
+However, this is computationally infeasible if $\Aop$ has a large or infinite cluster of eigenvalues near zero, a common situation for Hessians in ill-posed inverse problems.
 We therefore recommend choosing $\tau < 0$, and we use $\tau=-0.1$ in our numerical experiments.
@@
 
-The computational cost of our method may be divided into the costs to perform the following tasks: (1) Computing impulse moments and batches; (2) Building the H-matrix; (3) Performing linear algebra operations with the H-matrix.
+The computational cost of our method may be divided into the costs to perform the following tasks: (1) Computing impulse response moments and batches; (2) Building the H-matrix; (3) Performing linear algebra operations with the H-matrix.
 This may optionally include modifying the H-matrix to make it symmetric positive semi-definite.
-In practice, (1) is the dominant cost, because (1) is the only task that requires applying $\Aop$ to vectors, and our method is targeted at applications in which applying $\Aop$ to a vector requires an expensive computational procedure such as solving a PDE.
+In target applications, (1) is the dominant cost because applying $\Aop$ to a vector requires an expensive computational procedure such as solving a PDE, and (1) is the only step that requires applying $\Aop$ to vectors.
+%In practice, (1) is the dominant cost, because (1) is the only task that requires applying $\Aop$ to vectors, and  
 All operations that do not require applications of $\Aop$ to vectors are nearly linear, and therefore scalable, in the size of the problem, $N$.
@@
 
-\paragraph{(2) Building the H-matrix} The classical H-matrix construction technique
+\paragraph{(2) Building the H-matrix} Classical H-matrix construction techniques
 %, which is described in Appendix~\ref{app:h_matrix}, 
-requires evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix.
+require evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix.
 To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulse responses used in the RBF interpolation.
@@
 In the numerical results (Section~\ref{sec:numerical_results}), we use one matrix-matrix addition to add the H-matrix approximation of the data misfit term in the Hessian to the regularization term in the Hessian.
-Constructing an H-matrix representation of the regularization term in the Hessian requires one matrix inversion and two matrix-matrix multiplications.
+%Constructing an H-matrix representation of the regularization term in the Hessian requires one matrix inversion and two matrix-matrix multiplications. 
 %Then, we use one matrix factorization to form a preconditioner for the Newton linear system. 
@@
 we use a constant coefficient Robin boundary condition as
-in~\cite{Roininen14}. The parameters for the regularization are chosen
+in~\cite{Roininen14}. The parameters for the bi-Laplacian operator are chosen
 so that the Green's function of the Hessian of the regularization has
@@
 The advection-diffusion inverse problem is linear, so Newton's method converges in one iteration.
-In this case the Newton linear system,~\eqref{eq:newton_system}, is solved using PCG with a fixed tolerance which is specified in Section~\ref{sec:adv}.
+In this case the Newton linear system,~\eqref{eq:newton_system}, is solved using PCG, using termination tolerances described in Section \ref{sec:adv}.
 
@@
 are formed by adding $\bm{A}_H$ to $\mathbf{R}_H$ using fast H-matrix arithmetic.
-The H-matrix $\widetilde{\mathbf{H}}$ may be indefinite due to approximation error, so we modify it to be symmetric positive definite via a shift and invert deflation procedure described in Section~\ref{sec:flip_eigs}.
+We modify $\widetilde{\mathbf{H}}$ to be symmetric and remove negative eigenvalues via the procedure described in Section~\ref{sec:flip_eigs}.
 We factor $\bm{\preconditioner}$,
@@
 	\caption{Left (\ref{fig:cg_convergence}) shows the convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
-random entries drawn from the standard normal distribution and $\mathbf{H}$ is evaluated at the solution of the inverse problem.
+random entries drawn from the standard Gaussian distribution and $\mathbf{H}$ is evaluated at the solution of the inverse problem.
 %	The reference solution $\bm{x}$ is determined by solving $\bm{H} \bm{x} = \bm{b}$. 
@@
 
-We build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\bm{\basalfriction}$.
-We use PCG to solve a linear system with the Hessian as the coefficient operator and a right hand vector with random independent and identically distributed (i.i.d.) entries drawn from the standard normal distribution.
+Next, we build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\bm{\basalfriction}$.
+We use PCG to solve a linear system with the Hessian as the coefficient operator and a right hand vector with random independent and identically distributed (i.i.d.) entries drawn from the standard Gaussian distribution.
 In Figure~\ref{fig:cg_convergence} we compare the convergence of PCG for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners.
@@
 $ 
-The matrix $\bm{\preconditioner}$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations constructed at the converged solution $\bm{\basalfriction}$, or the regularization Hessian.
+The matrix $\bm{\preconditioner}$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations, the regularization Hessian (REG), or the identity matrix (NONE).
 The PSF preconditioners cluster the eigenvalues of the Hessian near one, with more batches yielding better clustering.
-The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
+%The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
 
@@
 
-Here we consider a time-dependent advection-diffusion equation for
-which we seek to infer the unknown spatially varying initial condition, $\ipar$, from observations of the state  
-at a final time. This PDE
+Here we consider a time-dependent advection-diffusion equation in
+which we seek to infer the unknown spatially varying initial condition, $\ipar$, from noisy observations of the state  
+at a final time, $T$. This PDE
 models diffusive transport in a domain $\advdomain \subset \R^d$, which is
@@
 Reynolds number 50, with boundary conditions $v(x)=(0,1)$ on the left boundary, $v(x)=(0,-1)$ on the right boundary, and $v(x)=(0,0)$ on the top and bottom boundaries, as in~\cite[Section 3]{PetraStadler11}.
-In the numerical results we use a checkerboard image for the initial condition and add $5\%$ multiplicative noise to the observations at the final time, $T$.
+We use a checkerboard image for the initial condition and add $5\%$ multiplicative noise to the observations at the final time, $T$.
 We use $\kappa=3.2e{-1}$ and $T=1.0$ for all results, except for Table \ref{tab:adv_convergence} and Figure \ref{tbl:adv_vary_kappa_tf} where we vary $\kappa$ and $T$ to study the impact of these parameters on the impulse responses and our preconditioner's effectiveness.
@@
 
-In Table \ref{tab:adv_convergence} we show the number of PCG iterations required to solve the Newton linear system to a relative error tolerance of $10^{-6}$.
+In Table \ref{tab:adv_convergence} we show the number of PCG iterations, $k$, required to solve the Newton linear system to a relative error tolerance of $||\widehat{\mathbf{q}} - \widehat{\mathbf{q}}_k|| < 10^{-6} ||\widehat{\mathbf{q}}||$.
+The solution of the Newton system that we compare to, $\widehat{\mathbf{q}}$, is found via another PCG iteration with a relative residual tolerance of $10^{-11}$.
 We show results for $T$ ranging from $0.5$ to $2.0$ and $\kappa$ ranging from $10^{-4}$ to $10^{-3}$, using our PSF preconditioner with $1$, $5$, and $25$ batches, regularization preconditioning, and no preconditioning.
@@
 In Figure \ref{fig:adv_cg_convergence} we show the convergence of PCG for solving $\mathbf{H}\mathbf{x}=\mathbf{b}$, where $\mathbf{b}$ has i.i.d.
-random entries drawn from a standard normal distribution.
+random entries drawn from a standard Gaussian distribution.
 The preconditioners used, $\widetilde{H}$, are our PSF preconditioner with $1$, $5$, or $25$ batches, regularization preconditioning, and no preconditioning.
-The PSF preconditioners converge fastest, with more batches yielding faster convergence.
+PCG converges fastest with the PSF preconditioners, with more batches yielding faster convergence.
 In Figure \ref{fig:adv_generalized_eigenvalues}, we show the eigenvalues for the generalized eigenvalue problem $\mathbf{H}\mathbf{u}=\lambda \widetilde{\mathbf{H}}\mathbf{u}$, where the $\widetilde{H}$ are the preconditioners stated above.
@@
 	\end{center}
-	\caption{Number of preconditioned conjugate gradient iterations required to achieve relative error $10^{-6}$ for solving the Newton linear system in the advection-diffusion inverse problem.
+	\caption{Number of preconditioned conjugate gradient iterations required to achieve a relative $l^2$ error of $10^{-6}$ for solving the Newton linear system in the advection-diffusion inverse problem.
 Iteration counts are shown for a variety of different diffusion parameters $\kappa$, simulation times $T$, and preconditioners.}
@@
 	\caption{Left (\ref{fig:adv_cg_convergence}) shows the convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
-random entries drawn from the standard normal distribution.
+random entries drawn from the standard Gaussian distribution.
 Right (\ref{fig:adv_generalized_eigenvalues}) shows eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$.
@@
 Fast H-matrix arithmetic is then used to perform further linear algebra operations that cannot be performed easily with the original operator, such as matrix factorization and inversion.
-The supports of the impulse responses are estimated to be contained in ellipsoids, which are determined a-priori via a new procedure that involves applying the operator to a small number of polynomial functions.
+The supports of the impulse responses are estimated to be contained in ellipsoids, which are determined a-priori via a moment method that involves applying the operator to a small number of polynomial functions.
 Point source locations for the impulse response batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulse responses per batch as possible, while ensuring that the corresponding ellipsoids do not overlap.

### 2023-06-04 paper/localpsf.tex
edits up until Section 7.2

@@
 	The method computes impulse responses of the operator at a collection of scattered points, and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel.
-We compute a large number of impulse responses by applying the operator to a small number Dirac combs associated with ``batches'' of point sources.
+We compute a large number of impulse responses by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources.
 By solving an ellipsoid packing problem, we compute as many impulse responses as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
@@
 via a black box computational procedure, 
-but we cannot easily access entries of $\Aop$'s integral kernel.
-Evaluating the maps in~\eqref{eq:eval_maps} may require, as a sub procedure, solving PDEs or performing other costly computations.
+but cannot easily access entries of $\Aop$'s integral kernel.
+Evaluating the maps in~\eqref{eq:eval_maps} may require, as a sub procedure, solving PDEs or other costly computations.
 By ``non-negative integral kernel,'' we mean that entries of the integral kernel are non-negative numbers; this is not the same as positive semi-definiteness of $\Aop$.
@@
 %This allows fast evaluation of approximate matrix entries, which are needed by high-rank approximation methods, in particular, classical H-matrix construction methods. 
-Picking the batches of points requires us to estimate the supports of the impulse responses $\phi_x$ \emph{before} we compute them.
-We form a-priori ellipsoid estimates for the supports of all the $\phi_x$ via a moment method that involves applying $\Aop^T$ to a small number of polynomial functions (see Section~\ref{sec:intromoments}).
-We were inspired by resolution analysis in seismic imaging, in which
-$\Aop^T$ is applied to a random noise function, and the width of $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$~\cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
+Picking the batches of points requires us to estimate the supports of the impulse responses $\phi_x$ \emph{before} they are computed.
+Ellipsoid estimates for the supports of all $\phi_x$ are determined a-priori via a moment method that involves applying $\Aop^T$ to a small number of polynomial functions (see Section~\ref{sec:intromoments}).
+We are inspired by resolution analysis in seismic imaging, in which
+$\Aop^T$ is applied to a random noise function, and the width of $\phi_x$\todo{the width of the support?} is estimated to be the autocorrelation length of the resultant function near $x$~\cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
 Our moment method estimates the support of $\phi_x$ more accurately than random noise probing, at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
@@
 %one to approximate discretized matrix representations of $\Aop$ using methods that require fast access to arbitrary matrix entries. We 
-We form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of $\Aop$.
+We form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of $\Aop$\todo{here a reviewer might complain that $\Aop$ is an infinite dimensional operator and H-matrix methods are only defined for finite dimensional operators}.
 H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank.
@@
 In the probabilistic framework, Bayes' theorem combines the observations with prior information to form a posterior distribution over the space of all possible parameter fields, and computations are performed to extract statistical information about the parameter from this posterior.
-The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal to each other under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian.'' The Hessian consists of a data misfit term (the ``data misfit Hessian''), which depends on discrepancy between the observations and the associated model predictions, and a regularization or prior term (the ``regularization Hessian'') which does not depend on the observations.
+The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal to each other under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian.'' The Hessian consists of a data misfit term (the ``data misfit Hessian''), which depends on a discrepancy between the observations and the associated model predictions, and a regularization or prior term (the ``regularization Hessian'') which does not depend on the observations.
 For more details on the Hessian, see~\cite{Alger19,GhattasWillcox21,VillaPetraGhattas21}.
@@
 When solving the deterministic optimization problem with Newton-type methods, the Hessian is the coefficient operator for the linear system that must be solved or approximately solved at every Newton iteration.
-Good Hessian preconditioners reduce the number of iterations required to solve these Newton linear systems with Krylov methods such as conjugate gradient~\cite{Saad03}.
+Good Hessian preconditioners reduce the number of iterations required to solve these Newton linear systems with Krylov methods such as the conjugate gradient method~\cite{Saad03}.
 In the Bayesian setting, the inverse of the Hessian is the covariance of a local Gaussian approximation of the posterior.
-Such a Gaussian distribution can be used as a direct approximations of the posterior, or as a proposal for Markov chain Monte-Carlo methods for drawing samples from the posterior.
+This Gaussian distribution can be used directly as an approximation of the posterior, or it can be utilized as a proposal for Markov chain Monte-Carlo methods for drawing samples from the posterior.
 For instance, see~\cite{KimEtAl21,PetraEtAl14} and the references therein.
@@
 \label{eq:kernel_representation}
-(\Aop u)(w) := \int_\Omega \int_\Omega w(y) \Aker(y,x) u(x) dx dy.
+(\Aop u)(w) := \int_\Omega \int_\Omega w(y) \Aker(y,x) u(x) \mathrm{d}x \mathrm{d}y.
 \end{equation}
@@
 The integral kernel, $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$, exists in principle but is not easily accessible.
-In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define impulse responses (Section~\ref{sec:impulse_response_2}), we state the conditions on $\Aop$ that our method requires (Section~\ref{sec:conditions_2}), and we detail finite element discretization (Section~\ref{sec:finite_element_kernel}).
+In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define impulse responses (Section~\ref{sec:impulse_response_2}), we then state the conditions on $\Aop$ that our method requires (Section~\ref{sec:conditions_2}), and detail finite element discretization (Section~\ref{sec:finite_element_kernel}).
 
@@
 	\begin{align}
-		\left(\Aop \genericdistribution^*\right)(w) &= \int_\Omega \int_\Omega w(y) \Aker(y,x) \genericdistribution^*(x) dx ~dy \\
-		&= \int_\Omega w(y) \int_\Omega \Aker(y,x) \genericdistribution^*(x) dx ~dy 
-		= \int_\Omega w(y) \genericdistribution\left(\Aker(y,\cdot)\right) dy, \label{eq:action_on_distribution}
+		\left(\Aop \genericdistribution^*\right)(w) &= \int_\Omega \int_\Omega w(y) \Aker(y,x) \genericdistribution^*(x) \mathrm{d}x ~\mathrm{d}y \\
+		&= \int_\Omega w(y) \int_\Omega \Aker(y,x) \genericdistribution^*(x) \mathrm{d}x ~\mathrm{d}y 
+		= \int_\Omega w(y) \genericdistribution\left(\Aker(y,\cdot)\right) \mathrm{d}y, \label{eq:action_on_distribution}
 	\end{align}
@@
 %\end{equation*}
-Here $\massmatrix \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries $\massmatrix_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx$ for $i,j=1,\dots,\fedim$.
+Here $\massmatrix \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries $\massmatrix_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) \mathrm{d}x$ for $i,j=1,\dots,\fedim$.
 %\begin{equation*}
@@
 %Our method will use these ellipsoid estimates to choose sets of impulses that can be computed in batches. 
-Second, we describe an improved method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section~\ref{sec:local_mean_displacement_invariance}).
+Second, we describe an improved method to approximate impulse responses from other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section~\ref{sec:local_mean_displacement_invariance}).
 %Our method will interpolate impulses using this approximation.
@@
 The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative integral kernel property. Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
+%
+%
 \begin{equation}
 \label{eq:define_vol}
-\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy
+\spatialvol(x) := \int_\Omega \impulseresponse_x(y) \mathrm{d}y
 \end{equation}
+%
+%
 denote the spatially varying scaling factor, and for $i,j=1,\dots,\gdim$ define $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$ and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$ as follows:
+%
+%
 \begin{align}
-\spatialmean^i(x) :=& \int_\Omega (\impulseresponse_x(y) / V(x)) y^i ~dy \label{eq:define_mean} \\
-\spatialcov^{ij}(x) :=& \int_\Omega (\impulseresponse_x(y) / V(x)) \left(y^i - \spatialmean^i(x)\right) \left(y^j - \spatialmean^j(x)\right) ~dy, \label{eq:define_cov}
+\spatialmean^i(x) :=& \int_\Omega (\impulseresponse_x(y) / V(x)) y^i ~\mathrm{d}y \label{eq:define_mean} \\
+\spatialcov^{ij}(x) :=& \int_\Omega (\impulseresponse_x(y) / V(x)) \left(y^i - \spatialmean^i(x)\right) \left(y^j - \spatialmean^j(x)\right) ~\mathrm{d}y, \label{eq:define_cov}
 \end{align}
+%
+%
 where $\spatialmean^i(x)$ denotes the $i^\text{th}$ component of the vector $\spatialmean(x)$, and $\spatialcov^{ij}(x)$ denotes the $(i,j)$ entry of the matrix $\spatialcov(x)$.
@@
 We approximate the support of $\impulseresponse_x$ with the ellipsoid
+%
+%
 \begin{equation}
@@
 \end{equation}
+%
+%
 where $\tau$ is a fixed constant.
-The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as the normalized version of $\impulseresponse_x$.
+The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as the normalized version of $\impulseresponse_x$.
 The quantity $\tau$ is a parameter that must be chosen appropriately.
@@
 
-We therefore modify $\mathbf{A}_H$ to make it symmetric and remove negative eigenvalues. First, we symmetrize $\mathbf{A}_H$ via
+We modify $\mathbf{A}_H$ to make it symmetric and remove negative eigenvalues. First, we symmetrize $\mathbf{A}_H$ via
 $
@@
 We solve these linear systems by factorizing the matrices $\mathbf{A}_H^{\text{sym}} + \mu_i \mathbf{R}_H$ using fast H-matrix methods.
-We compute and flip all eigenvalues $\lambda < \tau$ which are less than some threshold $\tau \in (-1,0]$.
+We compute and flip all eigenvalues $\lambda < \tau$ which are less than some threshold $\tau \in (-1,0]$\todo{notation conflict, $\tau$ used as threshold here and previously as a number of standard deviations in choosing batches of points from Gaussian info of impulse responses}.
 By choosing $\tau > -1$, we ensure that the modified version of $\mathbf{A}_H^{\text{sym}} + \mathbf{R}_H$ is positive definite.
@@
 The computational cost of our method may be divided into the costs to perform the following tasks: (1) Computing impulse response moments and batches; (2) Building the H-matrix; (3) Performing linear algebra operations with the H-matrix.
-This may optionally include modifying the H-matrix to make it symmetric positive semi-definite.
+This may optionally include a symmetric positive semi-definite modification of the H-matrix.
 In target applications, (1) is the dominant cost because applying $\Aop$ to a vector requires an expensive computational procedure such as solving a PDE, and (1) is the only step that requires applying $\Aop$ to vectors.
@@
 Choosing all of the sample points therefore requires performing at most
+%
+%
 \begin{equation*}
@@
 \end{equation*}
+%
+%
 where $\nsamplepts$ is the total number of sample points in all batches.
@@
 The cost of picking the $i$th batch could be reduced to $O(\fedim \log |\pointbatch_i|)$ using more advanced computational geometry techniques, such as organizing the ellipsoids associated with points that have already been picked into a dynamic AABB tree, then using this AABB tree to accelerate the required ellipsoid intersection tests.
-However, from a practical perspective, the cost for picking sample points is small compared to other parts of the approximation algorithm, so we do not pursue these improvements in this paper.
+However, from a practical perspective, the cost for picking sample points is small compared to other parts of the approximation algorithm, so such improvements are not pursued here.
 
@@
 Therefore, building the H-matrix requires
+%
+%
 \begin{equation*}
@@
 \end{equation*}
+%
+%
 %This follows from the above discussion, and the fact that $\fedim \ge \nsamplepts$.
 
-\paragraph{(3) Performing linear algebra operations with the H-matrix} It is well known that H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion, and low rank updates require
+\paragraph{(3) Performing linear algebra operations with the H-matrix} It is well known that H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, matrix inversion, and low rank updates require
 \begin{equation}
@@
 To reconstruct the unknown parameter fields, denoted $q$, the inverse problems are formulated as nonlinear least squares
-optimization problems, whose cost functions consist of a misfit term
+optimization problems, whose cost functions consist of a data misfit term
 (between the observations and model output) and a bi-Laplacian
@@
 standard sparse H-matrix techniques. H-matrix approximations of the Gauss-Newton Hessian, 
-$$\bm{H}_\text{gn}\approx \widetilde{\mathbf{H}}:=\mathbf{A}_H + \mathbf{R}_H,$$
-are formed by adding $\bm{A}_H$ to $\mathbf{R}_H$ using fast H-matrix arithmetic.
-We modify $\widetilde{\mathbf{H}}$ to be symmetric and remove negative eigenvalues via the procedure described in Section~\ref{sec:flip_eigs}.
+$$\bm{H}_\text{gn}\approx \widetilde{\bm{H}}:=\mathbf{A}_H + \mathbf{R}_H,$$
+are formed by adding $\mathbf{A}_H$ to $\mathbf{R}_H$ using fast H-matrix arithmetic.
+We modify $\widetilde{\bm{H}}$ to be symmetric positive definite via the procedure described in Section~\ref{sec:flip_eigs}.
 We factor $\bm{\preconditioner}$,
@@
 We approximate $\bm{H}_\text{gn}$ rather than $\bm{H}$ because $\bm{H}$ more often has negative values in its integral kernel.
-Our results show that $\widetilde{\mathbf{H}}$ is a good preconditioner for both $\bm{H}_\text{gn}$ and $\bm{H}$.
+Our results show that~$\widetilde{\bm{H}}$ is a good preconditioner for both~$\bm{H}_\text{gn}$ and~$\bm{H}$.
 %The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to an H-matrix approximation of the Hessian of the regularization term to form an H-matrix approximation of $\bm{H}_\text{gn}$.
@@
 \label{sec:ice}
-
-%\textbf{Heat equation}
-%\begin{itemize}
-%	\item Hessian per-column error plots for 1, 5, and 15 batches
-%	\item $H-P$ error vs number of batches for interior and whole domain
-%	\item $P^{-1}H-I$: error vs rank for $P=R$ and $P=P_\text{pch}$ diffusion time, for 1, 5, and 15 batches
-%	\item Krylov iterations to tols 1e-1 and 1e-6 vs diffusion parameter
-%	\item GLR vs PCH vs diffusion time
-%	\item Krylov method convergence plot: $R$ vs $P_\text{pch}$ vs no preconditioner
-%	\item Krylov iterations to tols 1e-6 vs mesh size $h$, PCH vs R vs no preconditioner (hold)
-%	\item true parameter (H)
-%	\item deterministic reconstruction (H)
-%	\item noisy measurements (1\%, 5\%, 10\% noise). $u|_\text{top}$ (H)
-%	\item recovered state at top (H)
-%	\item mesh scalability of PCH (CC) (H)
-%	\item put plot data into data directory
-%	\item save function .pvd files in paraview  directory
-%\end{itemize}
-%
-%\begin{itemize}
-%	\item GLR vs PCH number of obs (Computational cost in PDE solves) (S)
-%	\item GLR vs PCH aspect ratio (CC) (S)
-%	\item turn on preconditioner after number of krylov iterations exceeds 15 in an iteration. Rebuild "as needed". (S)
-%\end{itemize}
-%
-%\begin{itemize}
-%	\item Stokes Krylov convergence (10 solves for nonlinear forward + 6 hessian matvecs for ellipsoid estimation + 10 matvecs for impulse responses + 30 matvecs for PCH6 CG solve + 100 matvecs for Reg CG solve)
-%	\item Stokes error vs num batches (small number of batches; 1, 3, 6, 9 batches?) (50 matvecs for randomized error estimation +)
-%	\item Velocity field
-%	\item basal sliding field
-%	\item Picture of impulse responses
-%	\item Mesh negative numbers picture
-%\end{itemize}
-%
-%\begin{itemize}
-%	\item mtrue
-%	\item utrue (no arrows)
-%	\item utrue (yes arrows)
-%	\item reconstruction iterations 1,3,4,8: PCH vs REG
-%	\item urecovered (arrows, 3D)
-%	\item morozov plot
-%	\item Newton convergence (iter, numcg, numPDEsolves (num preconditioner build), gradnorm)
-%	\item impulse response batches
-%	\item 
-%\end{itemize}
 
@@
 %
-The domain that is filled with ice is denoted by
-$\icedomain \subset\mathbb{R}^{3}$. The basal, lateral and top parts of $\partial \icedomain$ are denoted by $\Gamma_{b}$, $\Gamma_{l}$, and $\Gamma_{t}$, respectively.
+The region of ice is denoted by
+$\icedomain \subset\mathbb{R}^{3}$. The basal, lateral and top parts of the boundary $\partial \icedomain$ are denoted by $\Gamma_{b}$, $\Gamma_{l}$, and $\Gamma_{t}$, respectively.
 The governing equations are the linear incompressible Stokes
@@
 restricted to the top surface of the geometry.
-We use $5\%$ noise in all cases, except for Figure \ref{fig:stokes_reconstructions} and Table \ref{tab:condition_number} where the noise is varied from $1\%$ to $25\%$.
+We use $5\%$ noise in all cases, except for Figure \ref{fig:stokes_reconstructions} and Table \ref{tab:condition_number} where the noise is varied from $1\%$ to $25\%$ and the regularization determined by the Morozov discrepancy principle for each noise level.
 The true basal sliding
@@
 	\caption{Left (\ref{fig:cg_convergence}) shows the convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
-random entries drawn from the standard Gaussian distribution and $\mathbf{H}$ is evaluated at the solution of the inverse problem.
+random entries drawn from the standard Gaussian distribution and $\bm{H}$ is evaluated at the solution of the inverse problem.
 %	The reference solution $\bm{x}$ is determined by solving $\bm{H} \bm{x} = \bm{b}$. 
@@
 Reynolds number 50, with boundary conditions $v(x)=(0,1)$ on the left boundary, $v(x)=(0,-1)$ on the right boundary, and $v(x)=(0,0)$ on the top and bottom boundaries, as in~\cite[Section 3]{PetraStadler11}.
-We use a checkerboard image for the initial condition and add $5\%$ multiplicative noise to the observations at the final time, $T$.
-We use $\kappa=3.2e{-1}$ and $T=1.0$ for all results, except for Table \ref{tab:adv_convergence} and Figure \ref{tbl:adv_vary_kappa_tf} where we vary $\kappa$ and $T$ to study the impact of these parameters on the impulse responses and our preconditioner's effectiveness.
+We use a checkerboard image for the initial condition~\ref{fig:adv_initial_condition} and add $5\%$ multiplicative noise to the synthetic observations at the final time, $T$.
+We use $\kappa=3.2$e${-1}$ and $T=1.0$ for all results, except for Table \ref{tab:adv_convergence} and Figure \ref{tbl:adv_vary_kappa_tf} where we vary $\kappa$ and $T$ to study the impact of these parameters on the impulse responses and our preconditioner's effectiveness.
 The initial condition, velocity field, noisy observations, and reconstructed initial condition are shown in Figure~\ref{fig:adv_inversion1}.
 
-In Table \ref{tab:adv_convergence} we show the number of PCG iterations, $k$, required to solve the Newton linear system to a relative error tolerance of $||\widehat{\mathbf{q}} - \widehat{\mathbf{q}}_k|| < 10^{-6} ||\widehat{\mathbf{q}}||$.
+In Table~\ref{tab:adv_convergence} we show the number of PCG iterations, $k$, required to solve the Newton linear system to a relative error tolerance of $\|\widehat{\mathbf{q}} - \widehat{\mathbf{q}}_k\| < 10^{-6} \|\widehat{\mathbf{q}}\|$.
 The solution of the Newton system that we compare to, $\widehat{\mathbf{q}}$, is found via another PCG iteration with a relative residual tolerance of $10^{-11}$.
@@
 As $\kappa$ increases, the impulse responses grow wider and as $T$ increases the impulse responses grow longer.
-This means that the impulse responses become more local as $\kappa$ and $T$ decrease, so more impulse responses can fit within each batch.
-
-In Figure \ref{fig:adv_cg_convergence} we show the convergence of PCG for solving $\mathbf{H}\mathbf{x}=\mathbf{b}$, where $\mathbf{b}$ has i.i.d.
+This means that the impulse responses become more local as $\kappa$ and $T$ decrease, so more impulse responses can be computed per batch.
+
+In Figure~\ref{fig:adv_cg_convergence} we show the convergence of PCG for solving $\bm{H}\bm{x}=\bm{b}$, where $\bm{b}$ has i.i.d.
 random entries drawn from a standard Gaussian distribution.
-The preconditioners used, $\widetilde{H}$, are our PSF preconditioner with $1$, $5$, or $25$ batches, regularization preconditioning, and no preconditioning.
+The preconditioners used, $\widetilde{\bm{H}}$, are our PSF preconditioner with $1$, $5$, or $25$ batches, regularization preconditioning, and no preconditioning.
 PCG converges fastest with the PSF preconditioners, with more batches yielding faster convergence.
-In Figure \ref{fig:adv_generalized_eigenvalues}, we show the eigenvalues for the generalized eigenvalue problem $\mathbf{H}\mathbf{u}=\lambda \widetilde{\mathbf{H}}\mathbf{u}$, where the $\widetilde{H}$ are the preconditioners stated above.
+In Figure~\ref{fig:adv_generalized_eigenvalues}, we show the eigenvalues for the generalized eigenvalue problem $\bm{H}\bm{u}=\lambda \widetilde{\bm{H}}\bm{u}$, where the $\widetilde{\bm{H}}$ are the preconditioners stated above.
 The PSF preconditioners cluster the eigenvalues of the Hessian near one, and the regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
@@
 			& $T=0.5$ & $T=1.0$ & $T=2.0$ \\
-			$\kappa=1.0e{-4}$  & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-4.png} \\
-			$\kappa=3.2e{-4}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K3.2e-4.png} \\
-			$\kappa=1.0e{-3}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-3.png}
+			$\kappa=1.0$e${-4}$  & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-4.png} \\
+			$\kappa=3.2$e${-4}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K3.2e-4.png} \\
+			$\kappa=1.0$e${-3}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-3.png}
 		\end{tabular}

### 2023-06-05 paper/adv_pcg_convergence.tex
remaining edits on paper, I want to go back and review the terms that are highlighted in red. I updated a plot in order that the xlabel is consistent in Stokes and the advection diffusion example problems

@@
 		xtick={0, 200, 400, 600,800,1000},
-		xlabel={CG iteration},
+		xlabel={$k$, CG iteration},
 		ylabel={$\|\bm{x}_k - \bm{x}\|/\|\bm{x}\|$},

### 2023-06-05 paper/localpsf.tex
remaining edits on paper, I want to go back and review the terms that are highlighted in red. I updated a plot in order that the xlabel is consistent in Stokes and the advection diffusion example problems

@@
 	\end{center}
-	\caption{Number of preconditioned conjugate gradient iterations required to achieve a relative $l^2$ error of $10^{-6}$ for solving the Newton linear system in the advection-diffusion inverse problem.
+	\caption{Number of preconditioned conjugate gradient iterations required to achieve a relative $l^2$ error of $10^{-6}$ \todo{we should be more specific about what the stopping criteria is, I expect that it is based on the linear system residual and if so we should state that, in order to distinguish it from the Stokes results earlier} for solving the Newton linear system in the advection-diffusion inverse problem.
 Iteration counts are shown for a variety of different diffusion parameters $\kappa$, simulation times $T$, and preconditioners.}

### 2023-06-05 paper/adv_pcg_convergence.tex
editing pass on paper with Tucker. Paper is ready for Noemi

--- 2023-06-01 paper/adv_pcg_convergence.tex
@@
 		xtick={0, 200, 400, 600,800,1000},
-		xlabel={CG iteration},
-		ylabel={$\|\bm{x}_k - \bm{x}\|/\|\bm{x}\|$},
+		xlabel={$j$, CG iteration},
+		ylabel={$\|\bm{x}_j - \bm{x}\|/\|\bm{x}\|$},
 		title={CG convergence},

### 2023-06-05 paper/localpsf.tex
editing pass on paper with Tucker. Paper is ready for Noemi

--- 2023-06-04 paper/localpsf.tex
@@
 We are inspired by resolution analysis in seismic imaging, in which
-$\Aop^T$ is applied to a random noise function, and the width of $\phi_x$\todo{the width of the support?} is estimated to be the autocorrelation length of the resultant function near $x$~\cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
+$\Aop^T$ is applied to a random noise function, and the width of the support $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$~\cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
 Our moment method estimates the support of $\phi_x$ more accurately than random noise probing, at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
@@
 %one to approximate discretized matrix representations of $\Aop$ using methods that require fast access to arbitrary matrix entries. We 
-We form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of $\Aop$\todo{here a reviewer might complain that $\Aop$ is an infinite dimensional operator and H-matrix methods are only defined for finite dimensional operators}.
+We form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of a discretized version of $\Aop$.
 H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank.
@@
 
-We modify $\mathbf{A}_H$ to make it symmetric and remove negative eigenvalues. First, we symmetrize $\mathbf{A}_H$ via
+We modify $\mathbf{A}_H$ to make it symmetric and remove negative eigenvalues via the following procedure. First, we symmetrize $\mathbf{A}_H$ via
 $
@@
 We solve these linear systems by factorizing the matrices $\mathbf{A}_H^{\text{sym}} + \mu_i \mathbf{R}_H$ using fast H-matrix methods.
-We compute and flip all eigenvalues $\lambda < \tau$ which are less than some threshold $\tau \in (-1,0]$\todo{notation conflict, $\tau$ used as threshold here and previously as a number of standard deviations in choosing batches of points from Gaussian info of impulse responses}.
-By choosing $\tau > -1$, we ensure that the modified version of $\mathbf{A}_H^{\text{sym}} + \mathbf{R}_H$ is positive definite.
-Choosing $\tau=0$ would remove all erroneous negative eigenvalues.
+We compute and flip all eigenvalues $\lambda < \epsilon_\text{flip}$ which are less than some threshold $\epsilon_\text{flip} \in (-1,0]$.
+By choosing $\epsilon_\text{flip} > -1$, we ensure that the modified version of $\mathbf{A}_H^{\text{sym}} + \mathbf{R}_H$ is positive definite.
+Choosing $\epsilon_\text{flip}=0$ would remove all erroneous negative eigenvalues.
 However, this is computationally infeasible if $\Aop$ has a large or infinite cluster of eigenvalues near zero, a common situation for Hessians in ill-posed inverse problems.
-We therefore recommend choosing $\tau < 0$, and we use $\tau=-0.1$ in our numerical experiments.
+We therefore recommend choosing $\epsilon_\text{flip} < 0$, and we use $\epsilon_\text{flip}=-0.1$ in our numerical experiments.
 
@@
 The computational cost of our method may be divided into the costs to perform the following tasks: (1) Computing impulse response moments and batches; (2) Building the H-matrix; (3) Performing linear algebra operations with the H-matrix.
-This may optionally include a symmetric positive semi-definite modification of the H-matrix.
+This may optionally include the symmetric positive semi-definite modifications described in Section \ref{sec:flip_eigs}.
 In target applications, (1) is the dominant cost because applying $\Aop$ to a vector requires an expensive computational procedure such as solving a PDE, and (1) is the only step that requires applying $\Aop$ to vectors.
@@
   2.2]{VillaPetraGhattas21}. In all numerical results we choose the
-regularization parameter (which controls the overall strength of the regularization) using Morozov's discrepancy
+regularization parameter (which controls the overall strength of the regularization) using the Morozov discrepancy
 principle~\cite{Vogel02}. 
@@
 are formed by adding $\mathbf{A}_H$ to $\mathbf{R}_H$ using fast H-matrix arithmetic.
-We modify $\widetilde{\bm{H}}$ to be symmetric positive definite via the procedure described in Section~\ref{sec:flip_eigs}.
+We modify $\widetilde{\bm{H}}$ to be (approximately) symmetric positive semi-definite via the procedure described in Section~\ref{sec:flip_eigs}.
 We factor $\bm{\preconditioner}$,
@@
 restricted to the top surface of the geometry.
-We use $5\%$ noise in all cases, except for Figure \ref{fig:stokes_reconstructions} and Table \ref{tab:condition_number} where the noise is varied from $1\%$ to $25\%$ and the regularization determined by the Morozov discrepancy principle for each noise level.
+We use $5\%$ noise in all cases, except for Figure~\ref{fig:stokes_reconstructions} and Table \ref{tab:condition_number} where the noise is varied from $1\%$ to $25\%$ and the regularization is determined by the Morozov discrepancy principle for each noise level.
 The true basal sliding
@@
     \end{center}
-	\caption{
+	\caption{\textbf{(Ice sheet)} 
 		(\ref{fig:ice_mountain_mesh}) Bird's eye view of the ice sheet, discretized by a mesh of tetrahedra.
@@
     \end{subfigure}
-	\caption{The log basal sliding parameter, with colorscale as in Figure~\ref{fig:true_beta}, computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $5.0\%$ (middle), and $1.0\%$ (right).}
+	\caption{\textbf{(Ice sheet)} The log basal sliding parameter, with colorscale as in Figure~\ref{fig:true_beta}, computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $5.0\%$ (middle), and $1.0\%$ (right).}
 	\label{fig:stokes_reconstructions} 
@@
 	}
-	\caption{Convergence history for solving the Stokes inverse problem using inexact Newton PCG.
+	\caption{\textbf{(Ice sheet)} Convergence history for solving the Stokes inverse problem using inexact Newton PCG.
 The Newton iteration is terminated when $||\mathbf{g}_k|| < 10^{-6} ||\mathbf{g}_0||$, where $\mathbf{g}_i$ is the gradient at the $i$th iteration.
@@
 		\includegraphics[scale=1.0]{stokes_pcg_convergence.pdf}
-		\caption{}
-		\label{fig:cg_convergence}
+%		\caption{}
+%		\label{fig:cg_convergence}
 	\end{subfigure}
@@
 	\end{center}
-	\caption{}
-	\label{fig:generalized_eigenvalues}
+%	\caption{}
+%	\label{fig:generalized_eigenvalues}
     \end{subfigure}
-	\caption{Left (\ref{fig:cg_convergence}) shows the convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
+	\caption{\textbf{(Ice sheet)} Left: convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
 random entries drawn from the standard Gaussian distribution and $\bm{H}$ is evaluated at the solution of the inverse problem.
@@
 The preconditioner is constructed using $\bm{H}_\text{gn}$.
-	Right (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$.
+	Right: generalized eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$.
 Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), the regularization Hessian (REG), or the identity matrix (NONE).}
@@
 We use PCG to solve a linear system with the Hessian as the coefficient operator and a right hand vector with random independent and identically distributed (i.i.d.) entries drawn from the standard Gaussian distribution.
-In Figure~\ref{fig:cg_convergence} we compare the convergence of PCG for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners.
+In the left subfigure of Figure~\ref{fig:krylov_convergence} we compare the convergence of PCG for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners.
 PCG converges fastest with the PSF preconditioners, with PSF (25) converging fastest, followed by PSF (5), followed by PSF (1), as expected.
 PCG converges much slower with no preconditioning and regularization preconditioning than it does with PSF preconditioning, with no preconditioning outperforming regularization preconditioning.
-In Figure~\ref{fig:generalized_eigenvalues} we show the generalized eigenvalues for the generalized eigenvalue problem
+In the right subfigure of Figure~\ref{fig:krylov_convergence} we show the generalized eigenvalues for the generalized eigenvalue problem
 $
@@
 The matrix $\bm{\preconditioner}$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations, the regularization Hessian (REG), or the identity matrix (NONE).
-The PSF preconditioners cluster the eigenvalues of the Hessian near one, with more batches yielding better clustering.
+With the PSF preconditioners, the generalized eigenvalues cluster near one, with more batches yielding better clustering.
 %The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
@@
 All PSF preconditioners outperform no preconditioning and regularization preconditioning for all noise levels.
-No preconditioning outperforms regularization preconditioning for all noise levels except $25\%$.
+%No preconditioning outperforms regularization preconditioning for all noise levels except $25\%$. 
 %We report condition number results for the preconditioned Hessian, $\mathbf{H}$, demonstrating that the PSF preconditioners built based on the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, are also good preconditioner for $\bm{H}$.
@@
 		\end{center}
-	\caption{Condition number for $\bm{\preconditioner}^{-1} \bm{H}$ for our preconditioner with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), no preconditioner (NONE) and regularization (REG).
+	\caption{\textbf{(Ice sheet)} Condition number for $\bm{\preconditioner}^{-1} \bm{H}$ for our preconditioner with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), no preconditioner (NONE) and regularization (REG).
 All operators are evaluated at the $\bm{\basalfriction}$ that solves the inverse problem for
@@
 Reynolds number 50, with boundary conditions $v(x)=(0,1)$ on the left boundary, $v(x)=(0,-1)$ on the right boundary, and $v(x)=(0,0)$ on the top and bottom boundaries, as in~\cite[Section 3]{PetraStadler11}.
-We use a checkerboard image for the initial condition~\ref{fig:adv_initial_condition} and add $5\%$ multiplicative noise to the synthetic observations at the final time, $T$.
-We use $\kappa=3.2$e${-1}$ and $T=1.0$ for all results, except for Table \ref{tab:adv_convergence} and Figure \ref{tbl:adv_vary_kappa_tf} where we vary $\kappa$ and $T$ to study the impact of these parameters on the impulse responses and our preconditioner's effectiveness.
+We use a checkerboard image for the initial condition (Figure~\ref{fig:adv_initial_condition}) and add $5\%$ multiplicative noise to generate synthetic observations at the final time, $T$.
+We use $\kappa=3.2$e${-1}$ and $T=1.0$ for all results, except for Table \ref{tab:adv_convergence} and Figure~\ref{tbl:adv_vary_kappa_tf} where we vary $\kappa$ and $T$ to study the impact of these parameters on the impulse responses and our preconditioner's effectiveness.
 The initial condition, velocity field, noisy observations, and reconstructed initial condition are shown in Figure~\ref{fig:adv_inversion1}.
@@
 The PSF preconditioners perform best in the high-rank regime where $T$ and $\kappa$ are small, and the observations are highly informative about the initial condition.
-This makes sense given the effect of $T$ and $\kappa$ on the shape of the impulse responses, as shown in Figure \ref{tbl:adv_vary_kappa_tf}.
+This makes sense given the effect of $T$ and $\kappa$ on the shape of the impulse responses, as shown in Figure~\ref{tbl:adv_vary_kappa_tf}.
 As $\kappa$ increases, the impulse responses grow wider and as $T$ increases the impulse responses grow longer.
@@
 
-In Figure~\ref{fig:adv_cg_convergence} we show the convergence of PCG for solving $\bm{H}\bm{x}=\bm{b}$, where $\bm{b}$ has i.i.d.
+In the right subfigure of Figure~\ref{fig:adv_krylov_convergence} we show the convergence of PCG for solving $\bm{H}\bm{x}=\bm{b}$, where $\bm{b}$ has i.i.d.
 random entries drawn from a standard Gaussian distribution.
@@
 PCG converges fastest with the PSF preconditioners, with more batches yielding faster convergence.
-In Figure~\ref{fig:adv_generalized_eigenvalues}, we show the eigenvalues for the generalized eigenvalue problem $\bm{H}\bm{u}=\lambda \widetilde{\bm{H}}\bm{u}$, where the $\widetilde{\bm{H}}$ are the preconditioners stated above.
-The PSF preconditioners cluster the eigenvalues of the Hessian near one, and the regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
-For the PSF preconditioner, more batches yields better clustering.
+In the left subfigure of Figure~\ref{fig:adv_krylov_convergence}, we show the generalized eigenvalues for the generalized eigenvalue problem $\bm{H}\bm{u}=\lambda \widetilde{\bm{H}}\bm{u}$, where the $\widetilde{\bm{H}}$ are the preconditioners stated above.
+With the PSF preconditioners, the eigenvalues cluster near one; more batches yields better clustering.
+With the regularization preconditioner the trailing eigenvalues cluster near one, while the leading eigenvalues are amplified.
 
@@
 	%\end{subfigure}
-	\caption{(\ref{fig:adv_initial_condition}) True initial condition.
+	\caption{\textbf{(Advection-diffusion)} (\ref{fig:adv_initial_condition}) True initial condition.
 (\ref{fig:adv_velocity}) Velocity field.
@@
 	\end{center}
-	\caption{Number of preconditioned conjugate gradient iterations required to achieve a relative $l^2$ error of $10^{-6}$ for solving the Newton linear system in the advection-diffusion inverse problem.
+	\caption{\textbf{(Advection-diffusion)} Number of preconditioned conjugate gradient iterations required to achieve error tolerance $||\widehat{\bm{q}}_j - \widehat{\bm{q}}|| < 10^{-6}||\widehat{\bm{q}}||$, where $\widehat{\bm{q}}_j$ is the $j$th conjugate gradient iterate, and $\widehat{\bm{q}}$ is the solution of the Newton linear system in the advection-diffusion inverse problem.
 Iteration counts are shown for a variety of different diffusion parameters $\kappa$, simulation times $T$, and preconditioners.}
@@
 		%	\renewcommand{\arraystretch}{2.0}
-		\begin{tabular}{cccc}
-			& $T=0.5$ & $T=1.0$ & $T=2.0$ \\
-			$\kappa=1.0$e${-4}$  & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-4.png} \\
-			$\kappa=3.2$e${-4}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K3.2e-4.png} \\
-			$\kappa=1.0$e${-3}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-3.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-3.png}
+		\begin{tabular}{ccc}
+			& $T=0.5$ 
+%			& $T=1.0$ 
+			& $T=2.0$ \\
+			$\kappa=1.0$e${-4}$  & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-4.png} 
+%			& \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-4.png}
+			& \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-4.png} \\
+%			$\kappa=3.2$e${-4}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K3.2e-4.png} \\
+			$\kappa=1.0$e${-3}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-3.png} 
+%			& \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-3.png} 
+			& \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-3.png}
 		\end{tabular}
 	}
-	\caption{Impulse responses for a variety of different diffusion parameters $\kappa$ and simulation times $T$.}
+	\caption{\textbf{(Advection-diffusion)} Impulse responses for small and large diffusion parameters $\kappa$ and simulation times $T$.}
 	\label{tbl:adv_vary_kappa_tf}
@@
 %		\end{tikzpicture}
-		\caption{}
-		\label{fig:adv_cg_convergence}
+%		\caption{}
+%		\label{fig:adv_cg_convergence}
 	\end{subfigure}
@@
 %\end{tikzpicture}
-		\caption{}
-		\label{fig:adv_generalized_eigenvalues}
+%		\caption{}
+%		\label{fig:adv_generalized_eigenvalues}
 	\end{subfigure}
-	\caption{Left (\ref{fig:adv_cg_convergence}) shows the convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
+	\caption{\textbf{(Advection-diffusion)} Left: convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
 random entries drawn from the standard Gaussian distribution.
-Right (\ref{fig:adv_generalized_eigenvalues}) shows eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$.
+Right: generalized eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$.
 Here $\bm{H}$ is the Hessian and the preconditioner $\bm{\preconditioner}$ is the PSF approximation for 1, 5, or 25 batches (PSF (1), PSF (5), and PSF (25), respectively), the regularization Hessian (REG), or the identity matrix (NONE).}

### 2023-06-05 paper/stokes_pcg_convergence.tex
editing pass on paper with Tucker. Paper is ready for Noemi

@@
 	title={CG convergence},
-	xlabel={$k$, CG iteration},
-	ylabel={$\|\bm{x}_k-\bm{x}\|/\|\bm{x}\|$},
+	xlabel={$j$, CG iteration},
+	ylabel={$\|\bm{x}_j-\bm{x}\|/\|\bm{x}\|$},
 	legend style={at={(1.2,1)},anchor=north east,draw=none},

### 2023-06-06 paper/localpsf.tex
Finished edits to Numerical Results section based on Noemi's comments

@@
 Further matrix computations are performed with H-matrix methods.
-We apply the method to approximate the Hessians in an ice sheet flow PDE constrained inverse problem, and an advection-diffusion PDE constrained inverse problem.
+We apply the method to approximate the Hessians in an ice sheet flow inverse problem governed by a Stokes PDE, and a diffusive transport inverse problem governed by an advection-diffusion PDE.
 Numerical results demonstrate that the proposed method outperforms existing methods, and can approximate high rank operators using only a small number of operator applies.
@@
 $\Aop^T$ is applied to a random noise function, and the width of the support $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$~\cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
-Our moment method estimates the support of $\phi_x$ more accurately than random noise probing, at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
-
-Our operator approximation method may be categorized as a PSF method that is loosely based on ``product convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
+The moment method that we use estimates the support of $\phi_x$ more accurately than random noise probing in resolution analysis, at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
+
+The proposed operator approximation method may be categorized as a PSF method that is loosely based on ``product convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
 PC and PSF methods have a long history dating back several decades.
@@
 Not all data misfit Hessians satisfy the local non-negative integral kernel properties (for example, the wave inverse problem data misfit Hessian has a substantial amount of negative entries in its integral kernel).
-However, many data misfit Hessians of practical interest do satisfy these properties (either exactly or approximately), and our method is targeted at approximating these Hessians.
+However, many data misfit Hessians of practical interest do satisfy these properties (either exactly or approximately), and the proposed method is targeted at approximating these Hessians.
 
@@
 The integral kernel, $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$, exists in principle but is not easily accessible.
-In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define impulse responses (Section~\ref{sec:impulse_response_2}), we then state the conditions on $\Aop$ that our method requires (Section~\ref{sec:conditions_2}), and detail finite element discretization (Section~\ref{sec:finite_element_kernel}).
+In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define impulse responses (Section~\ref{sec:impulse_response_2}), we then state the conditions on $\Aop$ that the proposed method requires (Section~\ref{sec:conditions_2}), and detail finite element discretization (Section~\ref{sec:finite_element_kernel}).
 
@@
 	\item The integral kernel is non-negative in the sense that $$\Aker(y,x) \ge 0$$ for all $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
-The operator $\Aop$ need not be positive semi-definite to use our method, and positive semi-definite operators need not have a non-negative integral kernel.}
+The operator $\Aop$ need not be positive semi-definite to use the proposed method, and positive semi-definite operators need not have a non-negative integral kernel.}
 \end{enumerate}
-Our method may still perform well if these conditions are relaxed slightly.
+The proposed method may still perform well if these conditions are relaxed slightly.
 It is fine if the support of $\impulseresponse_x$ is not perfectly contained in a localized region, as long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
 The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$, as long as it is non-negative for the vast majority of pairs of points $(y,x)$, and as long as the negative numbers are comparatively small.
-If these conditions are violated, our method will incur additional error.
-If these conditions are violated too much, our method may fail.
+If these conditions are violated, the method will incur additional error.
+If these conditions are violated too much, the method may fail.
 %Not all operators of interest (and in particular not all Hessians) satisfy these conditions, but many operators of practical interest do satisfy these conditions, and our method is highly effective for approximating these operators.
@@
 In this paper we discretize using continuous Galerkin finite elements satisfying the Kronecker property (defined below).
-With minor modifications, our method could be used with more general finite element methods, or other discretization schemes such as finite differences or finite volumes.
+With minor modifications, the proposed method could be used with more general finite element methods, or other discretization schemes such as finite differences or finite volumes.
 
@@
 under the isomorphism discussed above.
-Our method is agnostic to the computational procedure for approximating $\Aop$ with $\mathbf{A}$.
+The proposed method is agnostic to the computational procedure for approximating $\Aop$ with $\mathbf{A}$.
 What is important is that we do not have direct access to matrix entries $\mathbf{A}_{ij}$.
@@
 \end{equation}
-Broadly, our method will construct an H-matrix approximation of $\mathbf{A}$ by forming a H-matrix approximation of $\boldsymbol{\Aker}$, then multiplying $\boldsymbol{\Aker}$ by $\massmatrix$ (or a lumped mass version of $\massmatrix$) on the left and right using H-matrix methods.
+Broadly, we will construct an H-matrix approximation of $\mathbf{A}$ by forming a H-matrix approximation of $\boldsymbol{\Aker}$, then multiplying $\boldsymbol{\Aker}$ by $\massmatrix$ (or a lumped mass version of $\massmatrix$) on the left and right using H-matrix methods.
 Classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Aker}_{ij}$, but these matrix entries are not easily accessible.
-The bulk of our method is therefore dedicated to forming approximations of these matrix entries that can be evaluated rapidly.
+The bulk of the proposed method is therefore dedicated to forming approximations of these matrix entries that can be evaluated rapidly.
 
@@
 However, entries of $\Aker_h$, involve inverse mass matrices, which typically contain negative numbers.
-If there are too many negative numbers, or if the negative numbers are too large, our algorithm will yield less accurate approximations and may potentially fail.
+If there are too many negative numbers, or if the negative numbers are too large, the proposed method will yield less accurate approximations and may potentially fail.
 We therefore recommend replacing the mass matrix, $\massmatrix$, with a positive diagonal \emph{lumped mass} approximation.
-In our numerical results, we use the lumped mass matrix in which the $i$th diagonal entry of the lumped mass matrix is the sum of all entries in the $i$th row of the mass matrix.
+In the numerical results, we use the lumped mass matrix in which the $i$th diagonal entry of the lumped mass matrix is the sum of all entries in the $i$th row of the mass matrix.
 Other mass lumping techniques may be used~\cite{Hughes12}.
@@
 
-In this section we present two key innovations that our method is based on.
+In this section we present two key innovations that the proposed method is based on.
 First, we define moments of the impulse responses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section~\ref{sec:intromoments}).
@@
 However, this direct approach is not feasible because we need to know $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ before we compute $\phi_x$, in order choose the point $x$.
-Computing $\phi_x$ in order to determine $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ would be extremely computationally expensive, and defeat the purpose of our algorithm, which is to reduce the computational cost by computing impulse responses in batches.
+Computing $\phi_x$ in order to determine $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ would be extremely computationally expensive, and defeat the purpose of the proposed method, which is to reduce the computational cost by computing impulse responses in batches.
 Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ indirectly, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions (e.g., 6 total operator applies in two spatial dimensions and 10 in three spatial dimensions).
@@
 The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$.
-However, in Section~\ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids is better), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
-In practice, we find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results.
+However, in Section~\ref{sec:sample_point_selection} we will see that the cost of the proposed method depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids is better), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
+In practice, we find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in the numerical results.
 The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is typically conservative.
@@
 These are four different ways to approximate an impulse response by a nearby impulse response, with each successive approximation building upon the previous ones.
-Our method uses~\eqref{eq:translate4}, which is the most sophisticated of these approximations.
+The proposed method uses~\eqref{eq:translate4}, which is the most sophisticated of these approximations.
 Approximation~\eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close.
@@
 Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, may overwhelm $\phi_{x_j}$ for a nearby point $x_j$ if $\spatialvol(x_i)$ is much larger than $\spatialvol(x_j)$.
-Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section~\ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $\spatialvol(x_i)$ is smaller than a predetermined threshold.
+Note that we are not in danger of dividing by zero, because the ellipsoid packing procedure from Section~\ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $\spatialvol(x_i)$ is smaller than a predetermined threshold.
 
@@
 We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree)~\cite{Ericson04} of the mesh cells used to discretize the problem.
-Note that $\phi_{x_i}\left(z_i\right)$ is well defined because $z_i \in \Omega$, and $\frac{\spatialvol(x)}{\spatialvol(x_i)}$ is well defined because our sample point picking procedure in Section~\ref{sec:sample_point_selection} ensures that $\spatialvol(x_i)>0$.
+Note that $\phi_{x_i}\left(z_i\right)$ is well defined because $z_i \in \Omega$, and $\frac{\spatialvol(x)}{\spatialvol(x_i)}$ is well defined because the sample point picking procedure in Section~\ref{sec:sample_point_selection} ensures that $\spatialvol(x_i)>0$.
 Per the discussion in Section~\ref{sec:local_mean_displacement_invariance}, we expect $\Aker(y,x) \approx f_i$ for $i=1,\dots,\numnbr$; the closer $x_i$ is to $x$, the better we expect the approximation to be.
@@
 
-We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming an H-matrix representation $\mathbf{\Aker}_H$ of $\mathbf{\Aker}$ then multiplying $\mathbf{\Aker}$ with mass matrices $\massmatrix$ per~\eqref{eq:boldA} to form
+We form an H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming an H-matrix representation $\mathbf{\Aker}_H$ of $\mathbf{\Aker}$ then multiplying $\mathbf{\Aker}$ with mass matrices $\massmatrix$ per~\eqref{eq:boldA} to form
 $
@@
 $
-In our case $\massmatrix$ is a lumped mass matrix and is therefore diagonal, so these matrix-matrix multiplications are trivial.
+In the numerical results, $\massmatrix$ is a lumped mass matrix and is therefore diagonal, so these matrix-matrix multiplications are trivial.
 If a non-diagonal mass matrix is used, one may form a H-matrix representation of the mass matrix, then perform the matrix-matrix multiplications in \eqref{eq:boldA} using H-matrix methods.
 %Once $\mathbf{A}_H$ is formed, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods~\cite{BORMHACKBUSCHBOOK}.
-We use H1 matrices in our numerical results, but any other H-matrix format could be used instead. For more details on H-matrices, see~\cite{Hackbusch15}.
+We use H1 matrices in the numerical results, but any other H-matrix format could be used instead. For more details on H-matrices, see~\cite{Hackbusch15}.
 
@@
 $\mathcal{H} = \Aop + \mathcal{R},$
-where $\Aop$ is a symmetric positive semi-definite operator that we approximate with our PSF method to form an H-matrix $\mathbf{A}_H$, and $\mathcal{R}$ is a symmetric positive definite operator that may be easily converted to an H-matrix $\mathbf{R}_H$ without using our PSF method.
+where $\Aop$ is a symmetric positive semi-definite operator that we approximate with the PSF method to form an H-matrix $\mathbf{A}_H$, and $\mathcal{R}$ is a symmetric positive definite operator that may be easily converted to an H-matrix $\mathbf{R}_H$ without using the PSF method.
 For example, in inverse problems $\mathcal{H}$ is the Hessian, $\Aop$ is the data misfit term in the Hessian which is dense and only available matrix-free, and $\mathcal{R}$ is the regularization term, which is typically an elliptic differential operator that becomes a sparse matrix after discretization.
@@
 However, this is computationally infeasible if $\Aop$ has a large or infinite cluster of eigenvalues near zero, a common situation for Hessians in ill-posed inverse problems.
-We therefore recommend choosing $\epsilon_\text{flip} < 0$, and we use $\epsilon_\text{flip}=-0.1$ in our numerical experiments.
+We therefore recommend choosing $\epsilon_\text{flip} < 0$, and we use $\epsilon_\text{flip}=-0.1$ in the numerical experiments.
 
@@
 
-The computational cost of our method may be divided into the costs to perform the following tasks: (1) Computing impulse response moments and batches; (2) Building the H-matrix; (3) Performing linear algebra operations with the H-matrix.
+The computational cost of the proposed method may be divided into the costs to perform the following tasks: (1) Computing impulse response moments and batches; (2) Building the H-matrix; (3) Performing linear algebra operations with the H-matrix.
 This may optionally include the symmetric positive semi-definite modifications described in Section \ref{sec:flip_eigs}.
@@
 
-We use our method to approximate the Newton (or Gauss-Newton) Hessians
+We use the proposed method to approximate the Newton (or Gauss-Newton) Hessians
 in inverse problems governed by PDEs which model steady state ice
@@
 conjugate gradient (PCG) scheme and a globalizing Armijo line search~\cite{NocedalWright99}. The Newton updates,
-$\bm{\searchdir}_k$, are obtained by solving 
+$\bm{\searchdir}$, are obtained by solving 
 %
@@
 	\label{eq:newton_system}
-	\bm{H} \bm{\searchdir}_k = - \bm{g}_k \qquad \text{or} \qquad \bm{H}_\text{gn} \bm{\searchdir}_k = - \bm{g}_k,
+	\bm{H} \bm{\searchdir} = - \bm{g} \qquad \text{or} \qquad \bm{H}_\text{gn} \bm{\searchdir} = - \bm{g},
 \end{equation}
@@
 wherein we choose the initial guess as the discretization of the constant
-function $q_0$. Here $\bm{g}_k$, $\bm{H}$ and $\bm{H}_\text{gn}$ are
+function $q_0$. Here $\bm{g}$, $\bm{H}$ and $\bm{H}_\text{gn}$ are
 the discretized gradient, Hessian, and Gauss-Newton Hessian of the
-inverse problem cost function, respectively.
+inverse problem cost function, respectively, evaluated at the current Newton iterate.
 %The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
@@
 subsequent iterations. The Newton iterations are terminated when
-$\|\bm{g}_k\|/\|\bm{g}_0\| < 10^{-6}$. Systems~\eqref{eq:newton_system} are solved inexactly using an inner PCG
-iteration, which terminates when the norm of the linear system residual is less than $\min \left(0.5,
-\sqrt{\|\bm{g}_k\|}\right)\|\bm{g}_k\|$.
-The advection-diffusion inverse problem is linear, so Newton's method converges in one iteration.
+$\|\bm{g}\| < 10^{-6} \|\bm{g}_0\|$, where $\bm{g}_0$ is the gradient evaluated at the initial guess. Systems~\eqref{eq:newton_system} are solved inexactly using an inner PCG
+iteration, which is terminated early based on the Eisenstat-Walker \cite{EisenstatWalker96} and Steihaug \cite{Steihaug83} conditions.
+The inverse problem governed by the advection-diffusion PDE is linear, hence Newton's method converges in one iteration.
 In this case the Newton linear system,~\eqref{eq:newton_system}, is solved using PCG, using termination tolerances described in Section \ref{sec:adv}.
@@
 H-matrix representations of these matrices and combine them into a H-matrix approximation of the regularization term in the Hessian, $\mathbf{R}_H$, using
-standard sparse H-matrix techniques. H-matrix approximations of the Gauss-Newton Hessian, 
+standard sparse H-matrix techniques. Then, H-matrix approximations of the Gauss-Newton Hessian, 
 $$\bm{H}_\text{gn}\approx \widetilde{\bm{H}}:=\mathbf{A}_H + \mathbf{R}_H,$$
@@
 We approximate $\bm{H}_\text{gn}$ rather than $\bm{H}$ because $\bm{H}$ more often has negative values in its integral kernel.
-Our results show that~$\widetilde{\bm{H}}$ is a good preconditioner for both~$\bm{H}_\text{gn}$ and~$\bm{H}$.
+The numerical results show that~$\widetilde{\bm{H}}$ is a good preconditioner for both~$\bm{H}_\text{gn}$ and~$\bm{H}$.
 %The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to an H-matrix approximation of the Hessian of the regularization term to form an H-matrix approximation of $\bm{H}_\text{gn}$.
@@
 Given observations of the tangential component of the ice velocity on the top surface of the ice, we invert for the logarithm of the unknown spatially varying basal sliding Robin coefficient field, which governs the resistance to sliding along the base of the ice sheet.
+The setup, which we briefly summarize, follows~\cite{IsaacEtAl15,PetraEtAl12}.
 %The inverse problem considered here is synthetic, but in practice surface velocity observations are available from satellites.
@@
 %
-The solution to these equations is the pair $(\velocity, \pressure)$, where $\velocity$ is the ice flow velocity field, and $\pressure$ is the pressure field.
+The solution to these equations is the pair $(\velocity, \pressure)$, where $\velocity$ is the ice flow velocity field\footnote{We do not use bold to denote vector or tensor fields to avoid confusion with vectors that arise from finite element discretizations, which are already denoted with bold.}, and $\pressure$ is the pressure field.
 Here, $\basalfriction$ is the unknown logarithmic
@@
 The parameter $\basalfriction$ is discretized with first order scalar continuous Galerkin finite elements on the mesh of triangles that results from restricting the tetrahedral mesh to the basal boundary, $\Gamma_b$.
-Note that $\Gamma_b$ is a 2D surface embedded in 3D due to the mountain topography.
+Note that $\Gamma_b$ is a two-dimensional surface embedded in three dimensions due to the mountain topography.
 We also generate a flattened version of $\Gamma_b$, denoted by $\Omega \subset \mathbb{R}^2$, by ignoring the height coordinate.
-The parameter $\basalfriction$ is dually viewed as a function on $\Gamma_b$ for the purpose of solving the Stokes equations, and as a function on $\Omega$ for the purpose of building our Hessian approximations and defining the regularization.
+The parameter $\basalfriction$ is viewed as a function on $\Gamma_b$ for the purpose of solving the Stokes equations, and as a function on $\Omega$ for the purpose of building Hessian approximations and defining the regularization.
 %This flattening is necessary because our operator approximation method requires adding points on this surface together to get other points on the surface, and this is only well-defined in flat space.
@@
 The true basal sliding
-coefficient field and the true velocity field, which is obtained by
+coefficient and velocity fields, which are obtained by
 solving~\eqref{Stokeseqn}, are shown in Figure~\ref{fig:true_beta_u}.
@@
     \end{center}
-	\caption{\textbf{(Ice sheet)} 
+	\caption{(Ice sheet) 
 		(\ref{fig:ice_mountain_mesh}) Bird's eye view of the ice sheet, discretized by a mesh of tetrahedra.
@@
     \end{subfigure}
-	\caption{\textbf{(Ice sheet)} The log basal sliding parameter, with colorscale as in Figure~\ref{fig:true_beta}, computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $5.0\%$ (middle), and $1.0\%$ (right).}
+	\caption{(Ice sheet) The log basal sliding parameter, with colorscale as in Figure~\ref{fig:true_beta}, computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $5.0\%$ (middle), and $1.0\%$ (right).}
 	\label{fig:stokes_reconstructions} 
@@
 	}
-	\caption{\textbf{(Ice sheet)} Convergence history for solving the Stokes inverse problem using inexact Newton PCG.
-The Newton iteration is terminated when $||\mathbf{g}_k|| < 10^{-6} ||\mathbf{g}_0||$, where $\mathbf{g}_i$ is the gradient at the $i$th iteration.
-Preconditioners shown are our method with five batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
-Columns titled \#CG show the number of PCG iterations used to solve the Newton system for $\bm{\widehat{\basalfriction}}_k$.
-Columns titled $\|\bm{g}\|$ show the $l^2$ norm of the gradient at $\bm{\basalfriction}_k$.
+	\caption{(Ice sheet) Convergence history for solving the Stokes inverse problem using inexact Newton PCG.
+The Newton iteration is terminated when $||\mathbf{g}|| < 10^{-6} ||\mathbf{g}_0||$, where $\mathbf{g}$ is the gradient at the current iteration, and $\mathbf{g}_0$ is the gradient at the initial guess.
+Preconditioners shown are the proposed PSF method with five batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
+Columns titled \#CG show the number of PCG iterations used to solve the Newton system for $\bm{\widehat{\basalfriction}}$.
+Columns titled $\|\bm{g}\|$ show the $l^2$ norm of the gradient at $\bm{\basalfriction}$.
 Columns titled \#Stokes show the total number of Stokes PDE solves performed in each Newton iteration.
-This consists of Stokes solves for $u$ performed during the line search going from $\bm{\basalfriction}_{k-1}$ to $\bm{\basalfriction}_k$, plus one adjoint Stokes solve to compute the gradient at $\bm{\basalfriction}_k$, plus one incremental forward and one incremental adjoint Stokes solve per PCG iteration for solving the Newton system.
+This consists of Stokes solves performed during the line search to arrive at the current iterate from the previous iterate, plus one adjoint Stokes solve to compute the gradient at the current iterate, plus two Stokes solves per PCG iteration when solving the Newton system to compute the next search direction.
 Under PSF (5) and in row Iter 3, we write $6+22$ to indicate that $6$ Stokes solves were used during the standard course of the iteration, and $22$ Stokes solves were used to build the PSF (5) preconditioner.}
@@
 
-Table~\ref{tab:newton_convergence_table} shows the performance of our preconditioner for accelerating the solution of the optimization problem to reconstruct $\mathbf{q}$ from observations with $5\%$ noise.
+Table~\ref{tab:newton_convergence_table} shows the performance of the preconditioner for accelerating the solution of the optimization problem to reconstruct $\mathbf{q}$ from observations with $5\%$ noise.
 We build the PSF (5) preconditioner in the third Gauss-Newton iteration, and reuse it for all subsequent Gauss-Newton and Newton iterations.
 No preconditioning is used in the Gauss-Newton iterations before the PSF (5) preconditioner is built.
-We compare our method with the most commonly used existing preconditioners: no preconditioning (NONE), and preconditioning by the regularization term in the Hessian (REG).
-Using PSF (5) reduces the total number of Stokes PDE solves substantially compared to regularization preconditioning and no preconditioning\footnote{Interestingly, we observe that no preconditioning outperforms regularization preconditioning.}.
+We compare the proposed method with the most commonly used existing preconditioners: no preconditioning (NONE), and preconditioning by the regularization term in the Hessian (REG).
+The results show that using PSF (5) reduces the total number of Stokes PDE solves substantially compared to regularization preconditioning and no preconditioning.
 In Figure~\ref{fig:stokes_reconstructions} we show reconstructions for $1\%$, $5\%$, and $25\%$ noise.
@@
     \end{subfigure}
-	\caption{\textbf{(Ice sheet)} Left: convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
+	\caption{(Ice sheet) Left: convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
 random entries drawn from the standard Gaussian distribution and $\bm{H}$ is evaluated at the solution of the inverse problem.
@@
 %	Here we use $\gamma=7.3 \times 10^3$, which is determined by the Morozov discrepancy principle when using $5\%$ noise.
-	Results in these figures are shown for our preconditioner with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
+	Results in these figures are shown for the PSF-based preconditioners with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
 The preconditioner is constructed using $\bm{H}_\text{gn}$.
@@
 In the left subfigure of Figure~\ref{fig:krylov_convergence} we compare the convergence of PCG for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners.
-PCG converges fastest with the PSF preconditioners, with PSF (25) converging fastest, followed by PSF (5), followed by PSF (1), as expected.
-PCG converges much slower with no preconditioning and regularization preconditioning than it does with PSF preconditioning, with no preconditioning outperforming regularization preconditioning.
+PCG converges fastest with the PSF-based preconditioners, with PSF (25) converging fastest, followed by PSF (5), followed by PSF (1), as expected.
+PCG converges much slower with no preconditioning and regularization preconditioning than it does with PSF-based preconditioning, with no preconditioning outperforming regularization preconditioning.
 In the right subfigure of Figure~\ref{fig:krylov_convergence} we show the generalized eigenvalues for the generalized eigenvalue problem
@@
 The matrix $\bm{\preconditioner}$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations, the regularization Hessian (REG), or the identity matrix (NONE).
-With the PSF preconditioners, the generalized eigenvalues cluster near one, with more batches yielding better clustering.
+With the PSF-based preconditioners, the generalized eigenvalues cluster near one, with more batches yielding better clustering.
 %The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
@@
 
-In Table~\ref{tab:condition_number}, we show the condition number of the preconditioned Hessian using our preconditioners, for a variety of noise levels ranging from $1\%$ to $25\%$.
-The condition number using our PSF preconditioners is extremely small (ranging between 1 and 10) and relatively stable over this range of noise levels.
-PSF (25) performs better than PSF (5), which performs better than PSF (1), as expected.
-All PSF preconditioners outperform no preconditioning and regularization preconditioning for all noise levels.
+In Table~\ref{tab:condition_number}, we show the condition number of the preconditioned Hessian for a variety of noise levels ranging from $1\%$ to $25\%$.
+Note that the condition number using PSF-based preconditioners is extremely small (ranging between 1 and 10) and relatively stable over this range of noise levels.
+As expected, PSF (25) outperforms than PSF (5), which outperforms than PSF (1).
+All PSF-based preconditioners outperform no preconditioning and regularization preconditioning for all noise levels.
 %No preconditioning outperforms regularization preconditioning for all noise levels except $25\%$. 
@@
 		\end{center}
-	\caption{\textbf{(Ice sheet)} Condition number for $\bm{\preconditioner}^{-1} \bm{H}$ for our preconditioner with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), no preconditioner (NONE) and regularization (REG).
-All operators are evaluated at the $\bm{\basalfriction}$ that solves the inverse problem for
-	the respective noise level.}
+	\caption{(Ice sheet) Condition number for $\bm{\preconditioner}^{-1} \bm{H}$ for the PSF-based preconditioners with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), no preconditioner (NONE) and regularization (REG).
+All operators are evaluated at the soutions of the inverse problems for
+	their respective noise levels.}
 	\label{tab:condition_number}
@@
 
-\subsection{Example 2: Inversion for the initial condition in an advection-diffusion problem}
+\subsection{Example 2: Inversion for the initial condition in an diffusive transport problem}
 \label{sec:adv}
@@
 We use a checkerboard image for the initial condition (Figure~\ref{fig:adv_initial_condition}) and add $5\%$ multiplicative noise to generate synthetic observations at the final time, $T$.
-We use $\kappa=3.2$e${-1}$ and $T=1.0$ for all results, except for Table \ref{tab:adv_convergence} and Figure~\ref{tbl:adv_vary_kappa_tf} where we vary $\kappa$ and $T$ to study the impact of these parameters on the impulse responses and our preconditioner's effectiveness.
 The initial condition, velocity field, noisy observations, and reconstructed initial condition are shown in Figure~\ref{fig:adv_inversion1}.
-
-In Table~\ref{tab:adv_convergence} we show the number of PCG iterations, $k$, required to solve the Newton linear system to a relative error tolerance of $\|\widehat{\mathbf{q}} - \widehat{\mathbf{q}}_k\| < 10^{-6} \|\widehat{\mathbf{q}}\|$.
+We use $\kappa=3.2$e${-1}$ and $T=1.0$ for all results, except for Table \ref{tab:adv_convergence} and Figure~\ref{tbl:adv_vary_kappa_tf} where we vary $\kappa$ and $T$ to study the impact of changing these parameters.
+
+In Table~\ref{tab:adv_convergence} we show the number of PCG iterations, $j$, required to solve the Newton linear system to a relative error tolerance of $\|\widehat{\mathbf{q}} - \widehat{\mathbf{q}}_j\| < 10^{-6} \|\widehat{\mathbf{q}}\|$.
 The solution of the Newton system that we compare to, $\widehat{\mathbf{q}}$, is found via another PCG iteration with a relative residual tolerance of $10^{-11}$.
-We show results for $T$ ranging from $0.5$ to $2.0$ and $\kappa$ ranging from $10^{-4}$ to $10^{-3}$, using our PSF preconditioner with $1$, $5$, and $25$ batches, regularization preconditioning, and no preconditioning.
-PSF preconditioning performs better than regularization preconditioning and no preconditioning, with more batches yielding better results.
-The PSF preconditioners perform best in the high-rank regime where $T$ and $\kappa$ are small, and the observations are highly informative about the initial condition.
-This makes sense given the effect of $T$ and $\kappa$ on the shape of the impulse responses, as shown in Figure~\ref{tbl:adv_vary_kappa_tf}.
-As $\kappa$ increases, the impulse responses grow wider and as $T$ increases the impulse responses grow longer.
-This means that the impulse responses become more local as $\kappa$ and $T$ decrease, so more impulse responses can be computed per batch.
-
-In the right subfigure of Figure~\ref{fig:adv_krylov_convergence} we show the convergence of PCG for solving $\bm{H}\bm{x}=\bm{b}$, where $\bm{b}$ has i.i.d.
+We show results for $T$ ranging from $0.5$ to $2.0$ and $\kappa$ ranging from $10^{-4}$ to $10^{-3}$, using the PSF-based preconditioners with $1$, $5$, and $25$ batches, regularization preconditioning, and no preconditioning.
+The results show that PSF-based preconditioning outperforms regularization preconditioning and no preconditioning, with more batches yielding better results.
+%The PSF-based preconditioners perform best in the high-rank regime where $T$ and $\kappa$ are small, and the observations are highly informative about the initial condition.
+This makes sense given the effect of $T$ and $\kappa$ on the shape of the
+
+In the left subfigure of Figure~\ref{fig:adv_krylov_convergence} we show the convergence of PCG for solving $\bm{H}\bm{x}=\bm{b}$, where $\bm{b}$ has i.i.d.
 random entries drawn from a standard Gaussian distribution.
-The preconditioners used, $\widetilde{\bm{H}}$, are our PSF preconditioner with $1$, $5$, or $25$ batches, regularization preconditioning, and no preconditioning.
-PCG converges fastest with the PSF preconditioners, with more batches yielding faster convergence.
-In the left subfigure of Figure~\ref{fig:adv_krylov_convergence}, we show the generalized eigenvalues for the generalized eigenvalue problem $\bm{H}\bm{u}=\lambda \widetilde{\bm{H}}\bm{u}$, where the $\widetilde{\bm{H}}$ are the preconditioners stated above.
-With the PSF preconditioners, the eigenvalues cluster near one; more batches yields better clustering.
+The preconditioners used, $\widetilde{\bm{H}}$, are the PSF-based preconditioners with $1$, $5$, or $25$ batches, regularization preconditioning, and no preconditioning.
+The results show that PCG converges fastest with the PSF-based preconditioners, with more batches yielding faster convergence.
+In the right subfigure of Figure~\ref{fig:adv_krylov_convergence}, we show the generalized eigenvalues for the generalized eigenvalue problem $\bm{H}\bm{u}=\lambda \widetilde{\bm{H}}\bm{u}$, where the $\widetilde{\bm{H}}$ are the preconditioners stated above.
+As can be seen, with the PSF-based preconditioners, the eigenvalues cluster near one; more batches yields better clustering.
 With the regularization preconditioner the trailing eigenvalues cluster near one, while the leading eigenvalues are amplified.
@@
 	%\end{subfigure}
-	\caption{\textbf{(Advection-diffusion)} (\ref{fig:adv_initial_condition}) True initial condition.
+	\caption{(Diffusive transport) (\ref{fig:adv_initial_condition}) True initial condition.
 (\ref{fig:adv_velocity}) Velocity field.
@@
 	\end{center}
-	\caption{\textbf{(Advection-diffusion)} Number of preconditioned conjugate gradient iterations required to achieve error tolerance $||\widehat{\bm{q}}_j - \widehat{\bm{q}}|| < 10^{-6}||\widehat{\bm{q}}||$, where $\widehat{\bm{q}}_j$ is the $j$th conjugate gradient iterate, and $\widehat{\bm{q}}$ is the solution of the Newton linear system in the advection-diffusion inverse problem.
+	\caption{(Diffusive transport) Number of PCG iterations required to solve the Newton linear system to tolerance $||\widehat{\bm{q}}_j - \widehat{\bm{q}}|| < 10^{-6}||\widehat{\bm{q}}||$, where $\widehat{\bm{q}}_j$ is the $j$th conjugate gradient iterate, and $\widehat{\bm{q}}$ is the solution of the Newton linear system.
 Iteration counts are shown for a variety of different diffusion parameters $\kappa$, simulation times $T$, and preconditioners.}
@@
 	}
-	\caption{\textbf{(Advection-diffusion)} Impulse responses for small and large diffusion parameters $\kappa$ and simulation times $T$.}
+	\caption{(Diffusive transport) Impulse responses for small and large diffusion parameters $\kappa$ and simulation times $T$.}
 	\label{tbl:adv_vary_kappa_tf}
@@
 	\end{subfigure}
-	\caption{\textbf{(Advection-diffusion)} Left: convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
+	\caption{(Diffusive transport) Left: convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
 random entries drawn from the standard Gaussian distribution.
@@
 Point source locations for the impulse response batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulse responses per batch as possible, while ensuring that the corresponding ellipsoids do not overlap.
-We applied the method to approximate the Gauss-Newton Hessians in Stokes ice sheet and advection-diffusion inverse problems, and saw that the approximation substantially outperforms existing Hessian approximation methods.
+We applied the method to approximate the Gauss-Newton Hessians in an ice sheet flow inverse problem governed by the Stokes PDE, and a diffusive transport inverse problem governed by an advection-diffusion PDE.
+We saw that the proposed approximation method substantially outperforms existing Hessian approximation methods.
 Although the method we presented is not applicable to all Hessians, it is applicable to many Hessians of practical interest.
-For these Hessians, our method offers a \emph{data scalable} alternative to conventional low rank approximation, because our method can form high rank approximations of an operator using a small number of operator applies, and thus forward and adjoint PDE solves.
+For these Hessians, the proposed method offers a \emph{data scalable} alternative to conventional low rank approximation, due to the ability to form high rank approximations of an operator using a small number of operator applies, and thus forward and adjoint PDE solves.
 
@@
 \section*{Acknowledgments}
-We thank J.J. Alger, Longfei Gao, Mathew Hu, and Rami Nammour for helpful discussions. We thank Trevor Heise for editing suggestions. We thank Georg Stadler for domain geometry help.
+We thank J.J.
+Alger, Longfei Gao, Mathew Hu, and Rami Nammour for helpful discussions.
+We thank Trevor Heise for editing suggestions.
+We thank Georg Stadler for his help with the domain setup for the ice sheet inverse problem.
 

### 2023-06-08 paper/localpsf.tex
Editing based on discussion with Noemi (everything before numerical results)

@@
 	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels.
+	The method computes impulse responses of the operator at a collection of scattered points, and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel.
+By applying the operator to Dirac combs associated with batches of point sources, we compute a large number of impulse responses using a small number of operator applies.
+We solve an ellipsoid packing problem to select as many point sources as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
+Support ellipsoids are estimated a-priori based on impulse response moments, which are computed by applying the operator to a small number of polynomial functions.
+The interpolation uses a local mean displacement invariance approximation, which generalizes classical local translation invariance.
+The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
+Further matrix computations are performed with H-matrix methods.
+We illustrate the method by approximating the Hessians in two example problems: an ice sheet flow inverse problem governed by a Stokes partial differential equation (PDE), and a diffusive transport inverse problem governed by an advection-diffusion PDE.
+Because the data in these inverse problems are highly informative about the unknown parameters, the data misfit terms in the Hessians are high rank.
+This makes conventional matrix-free Hessian approximation methods, which are based on low rank approximation, extremely expensive in terms of the required number of Hessian operator applies.
+In contrast, the numerical results show that the proposed method can form good approximations of these high rank Hessians using a small number of Hessian operator applies.
+We use the Hessian approximations as preconditioners to accelerate the solution of the inverse problems.
+Numerical results demonstrate that this approach substantially reduces the number of PDE solves required to solve the inverse problems.
+
+%	, as compared to existing methods.
+
+%	, as measured by the total number of PDE solves required to solve the inverse problem.
+
 %	, with a focus on approximating high rank Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs). 
 %	The method may be used to approximate other operators, including certain Schur complements, Poincare-Steklov operators, covariance operators, and blurring operators.
-	The method computes impulse responses of the operator at a collection of scattered points, and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel.
-We compute a large number of impulse responses by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources.
-By solving an ellipsoid packing problem, we compute as many impulse responses as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
-Support ellipsoids are estimated a-priori based on impulse response moments, which are computed by applying the operator to a small number of polynomial functions.
-The interpolation uses a ``local mean displacement invariance'' approximation, which generalizes classical local translation invariance.
-The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
-Further matrix computations are performed with H-matrix methods.
-We apply the method to approximate the Hessians in an ice sheet flow inverse problem governed by a Stokes PDE, and a diffusive transport inverse problem governed by an advection-diffusion PDE.
-Numerical results demonstrate that the proposed method outperforms existing methods, and can approximate high rank operators using only a small number of operator applies.
+
+%	that arise in inverse problems with highly informative data.
+
+%	we see that the numerical results show that the proposed method can approximate high Existing methods typically perform poorly  when the data is highly informative about the unknown parameter, because the Hessians in these problems are high rank.
+In contrast, the proposed method can cheaply approximate high rank Hessians.
+
+%	, in the sense that it can approximate Hessians in inverse problem with highly informative data, using only a small number of operator applies.
+outperforms (as measured by number of PDE solves) existing methods for Hessian approximation.
+In particular, the proposed method can approximate Hessians in inverse problem with highly informative data, using only a small number of operator applies.
 \end{abstract}

### 2023-06-08 paper/localpsf.tex
editing paper based on Noemi's comments. Found bad typo in H-matrix complexity estimate (said kh^2 N log N entries needed, should be kh N log N)

@@
 Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, and $L^2(\Omega)'$ is the space of real-valued continuous linear functionals on $L^2(\Omega)$.
-We focus on approximating Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{AlgerEtAl19,IsaacEtAl15}.
-Such operators also arise as certain Schur complements in Schur complement methods for solving partial differential equations (PDEs)~\cite{le1997non,saad1999distributed}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics~\cite{ChenStein21,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}.
+Such operators appear, for instance, as Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{AntilEtAl18,BorziSchulz11}, Schur complements in Schur complement methods for solving partial differential equations (PDEs)~\cite{le1997non,saad1999distributed}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics~\cite{ChenStein21,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}.
 Here ``matrix-free'' means that we may apply $\Aop$ and its transpose\footnote{Recall that $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for operations with $\Aop^T$.
@@
 but cannot easily access entries of $\Aop$'s integral kernel.
-Evaluating the maps in~\eqref{eq:eval_maps} may require, as a sub procedure, solving PDEs or other costly computations.
+Evaluating the maps in~\eqref{eq:eval_maps} may require solving a sub problem that involves PDEs, or performing other costly computations.
 By ``non-negative integral kernel,'' we mean that entries of the integral kernel are non-negative numbers; this is not the same as positive semi-definiteness of $\Aop$.
@@
 	\end{subfigure}
-	\caption{Batches, $\eta_b$, of normalized impulses, $\impulseresponse_{x}$, for the ice sheet inverse problem data misfit Gauss-Newton Hessian described in Section~\ref{sec:numerical_results}. 
+	\caption{Three batches, $\eta_b$, of normalized impulse responses, $\impulseresponse_{x}$, that arise from applying $\Aop$ to weighted sums of scattered point sources (see Section \ref{sec:get_impulse_response}).
+Here $\Aop$ is the ice sheet inverse problem data misfit Gauss-Newton Hessian described in Section~\ref{sec:numerical_results}.
 %	Brightness scale differs between the subfigures. 
-	Black stars are point source locations. Dashed gray ellipses are estimated impulse response support ellipsoids. The large circle is $\partial \Omega$. 
+	Black stars are point source locations.
+Shading shows the magnitude of the normalized impulse responses (darker means larger function values).
+Dashed gray ellipses are estimated impulse response support ellipsoids based on the moment method in Section \ref{sec:intromoments}.
+The large circle is $\partial \Omega$.
 	}
@@
 	\end{center}
-	\caption{$\Aop$ is locally mean displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$.
+	\caption{This figure illustrates impulse responses, $\impulseresponse_x$ and $\impulseresponse_{x'}$, that arise from applying $\Aop$ to point sources at two points, $x$ and $x'$.
+The operator $\Aop$ is locally mean displacement invariant (Section \ref{sec:local_mean_displacement_invariance}) if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$.
 Here $\spatialmean(z)$ denotes the mean (center of mass) of $\phi_z$.}
@@
 %one to approximate discretized matrix representations of $\Aop$ using methods that require fast access to arbitrary matrix entries. We 
-We form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of a discretized version of $\Aop$.
+In this work, we form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of a discretized version of $\Aop$.
 H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank.
@@
 The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks~\cite{GrasedyckHackbusch03}.
-These operations can be used, for example, to solve Newton linear systems in PDE constrained optimization, sample from ill-conditioned posterior distributions in Bayesian inverse problems, and construct high rank Gaussian surrogate models that can be used for uncertainty quantification.
+%These operations can be used, for example, to solve Newton linear systems in PDE constrained optimization, sample from ill-conditioned posterior distributions in Bayesian inverse problems, and construct high rank Gaussian surrogate models that can be used for uncertainty quantification.
+
 %Classical H-matrix construction methods 
@@
 %Inverse problems are typically ill-posed, in the sense that many different parameter fields may yield predicted observations that agree with the actual observations to within the noise level. 
-Conventionally, the inverse problem is formulated using either a deterministic framework~\cite{AsterBorchersThurber18,Vogel02},
+Conventionally, the inverse problem is formulated using either a deterministic framework~\cite{AsterBorchersThurber18,BorziSchulz11,Vogel02},
 %Guy Chavent, Nonlinear Least Squares for Inverse Problems, Springer, 2009.
 or a Bayesian probabilistic framework~\cite{KaipioSomersalo05,Stuart10a,Tarantola05}.
-In the deterministic framework, one solves an optimization problem to find the parameter that best fits the observations, subject to appropriate regularization~\cite{EnglHankeNeubauer96}.
+In the deterministic framework, one solves an optimization problem to find the parameter that best fits the observations, subject to appropriate regularization~\cite{EnglHankeNeubauer96,Vogel02}.
 In the probabilistic framework, Bayes' theorem combines the observations with prior information to form a posterior distribution over the space of all possible parameter fields, and computations are performed to extract statistical information about the parameter from this posterior.
-The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal to each other under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian.'' The Hessian consists of a data misfit term (the ``data misfit Hessian''), which depends on a discrepancy between the observations and the associated model predictions, and a regularization or prior term (the ``regularization Hessian'') which does not depend on the observations.
+The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal to each other under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian.'' The Hessian consists of a data misfit term (the \emph{data misfit Hessian}), which depends on a discrepancy between the observations and the associated model predictions, and a regularization or prior term (the ``regularization Hessian'') which does not depend on the observations.
 For more details on the Hessian, see~\cite{Alger19,GhattasWillcox21,VillaPetraGhattas21}.
@@
 
-Examples of inverse problems with high rank data misfit Hessians include advection dominated advection-diffusion inverse problems~\cite{AkcelikBirosDraganescuEtAl05}\cite[Chapter 5]{Flath13}, high frequency wave inverse problems~\cite{BuiEtAl13}, and more generally, all inverse problems in which the observations highly inform the parameter.
+Examples of inverse problems with high rank data misfit Hessians include large-scale ice sheet inverse problems \cite{HartlandEtAl23,IsaacEtAl15}, advection dominated advection-diffusion inverse problems~\cite{AkcelikBirosDraganescuEtAl05}\cite[Chapter 5]{Flath13}, high frequency wave inverse problems~\cite{BuiEtAl13}, and more generally, all inverse problems in which the observations highly inform the parameter.
 The eigenvalues of the data misfit Hessian characterize the sensitivity of the predicted observations to perturbations of the parameter in the corresponding eigenvector directions,
 %(see Figure~\ref{fig:hessian_eigenstructure})
-so more informative data leads to larger eigenvalues and therefore a larger numerical rank~\cite{AlexanderianGloorGhattas16}\cite[Section 1.4 and Chapter 4]{Alger19}.
+hence more informative data leads to larger eigenvalues and therefore a larger numerical rank~\cite{AlexanderianGloorGhattas16}\cite[Section 1.4 and Chapter 4]{Alger19}.
 Roughly speaking, the numerical rank of the data misfit Hessian is the dimension of the subspace of parameter space that is informed by the data.
@@
 In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in~\eqref{eq:extension_to_distributions} is not valid.
-However, if $\Aker$ is sufficiently regular such that the function $y \mapsto \rho\left(\Aker(y,~\cdot~)\right)$ is well defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of~\eqref{eq:action_on_distribution} is well defined.
+However, if $\Aker$ is sufficiently regular such that the function $y \mapsto \rho\left(\Aker(y,~\cdot~)\right)$ is well-defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of~\eqref{eq:action_on_distribution} is well-defined.
 Hence, we \emph{define} the action of $\Aop$ on the distribution $\rho$ to be the right hand side of~\eqref{eq:action_on_distribution}.
@@
 \begin{enumerate}
-	\item The kernel $\Aker$ is sufficiently regular so that $\phi_x$ is well defined for all $x\in \Omega$.
+	\item The kernel $\Aker$ is sufficiently regular so that $\phi_x$ is well-defined for all $x\in \Omega$.
 	\item The supports of the impulse responses $\impulseresponse_x$ are contained in localized regions.
@@
 The proposed method may still perform well if these conditions are relaxed slightly.
-It is fine if the support of $\impulseresponse_x$ is not perfectly contained in a localized region, as long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
+It is acceptable if the support of $\impulseresponse_x$ is not perfectly contained in a localized region, so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
 The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$, as long as it is non-negative for the vast majority of pairs of points $(y,x)$, and as long as the negative numbers are comparatively small.
-If these conditions are violated, the method will incur additional error.
-If these conditions are violated too much, the method may fail.
+If these conditions are violated, the method will incur additional error, and depending on the severity of the violation, the proposed method may fail.
 %Not all operators of interest (and in particular not all Hessians) satisfy these conditions, but many operators of practical interest do satisfy these conditions, and our method is highly effective for approximating these operators.
@@
 
-In computations, functions are discretized and replaced by finite dimensional vectors, and operators mapping between infinite dimensional spaces are replaced by operators mapping between finite dimensional spaces.
+In computations, functions are discretized and replaced by finite-dimensional vectors, and operators mapping between infinite-dimensional spaces are replaced by operators mapping between finite-dimensional spaces.
 In this paper we discretize using continuous Galerkin finite elements satisfying the Kronecker property (defined below).
@@
 Rather, we have a computational procedure that allows us to compute matrix-vector products $\mathbf{u}\mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{w}\mapsto \mathbf{A}^T\mathbf{w}$, and computing these matrix-vector products is costly.
+The proposed method is designed to mitigate this cost by performing as few of these matrix-vector products as possible.
 Of course, matrix entries can be computed via matrix-vector products as $\mathbf{A}_{ij} = \left(\mathbf{A}\mathbf{e}_j\right)_i$,
@@
 The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm~\ref{alg:construct_Atilde}.
-The cost is discussed in Section~\ref{sec:overall_cost}.
+The computational cost is discussed in Section~\ref{sec:overall_cost}.
 
@@
 
-			Compute $\combresponse_k$ by applying $\Aop$ to the Dirac comb for $\pointbatch_k$ (Section~\ref{sec:get_impulse_response})
+			Compute impulse response batch $\combresponse_k$ by applying $\Aop$ to the Dirac comb for $\pointbatch_k$ (Section~\ref{sec:get_impulse_response})
 
@@
 
-		Form H-matrix approximation $\boldsymbol{\Aker}_H$ of integral kernel (Section~\ref{sec:h_matrix_construction_short})
+		Form H-matrix approximation $\boldsymbol{\Aker}_H$ of integral kernel (Sections~\ref{sec:approximate_kernel_entries} and \ref{sec:h_matrix_construction_short})
 
@@
 \end{equation*}
-We compute the \emph{impulse batch}, $\eta_k$, by applying $\Aop$ to the Dirac comb:
+We compute the \emph{impulse response batch}, $\eta_k$, by applying $\Aop$ to the Dirac comb:
 \begin{equation}
@@
 
+Here we describe how to rapidly evaluate arbitrary entries of an approximation to the integral kernel by performing radial basis function interpolation of translated and scaled versions of nearby known impulse responses.
+In Section~\ref{sec:h_matrix_construction_short} we use this procedure for rapidly evaluating kernel entries to construct the H-matrix approximation of $\mathbf{A}$.
+
 Given $(y,x)\in \Omega \times \Omega$, let
@@
 We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree)~\cite{Ericson04} of the mesh cells used to discretize the problem.
-Note that $\phi_{x_i}\left(z_i\right)$ is well defined because $z_i \in \Omega$, and $\frac{\spatialvol(x)}{\spatialvol(x_i)}$ is well defined because the sample point picking procedure in Section~\ref{sec:sample_point_selection} ensures that $\spatialvol(x_i)>0$.
+Note that $\phi_{x_i}\left(z_i\right)$ is well-defined because $z_i \in \Omega$, and $\frac{\spatialvol(x)}{\spatialvol(x_i)}$ is well-defined because the sample point choosing procedure in Section~\ref{sec:sample_point_selection} ensures that $\spatialvol(x_i)>0$.
 Per the discussion in Section~\ref{sec:local_mean_displacement_invariance}, we expect $\Aker(y,x) \approx f_i$ for $i=1,\dots,\numnbr$; the closer $x_i$ is to $x$, the better we expect the approximation to be.
@@
 $
-In the numerical results, $\massmatrix$ is a lumped mass matrix and is therefore diagonal, so these matrix-matrix multiplications are trivial.
+Here we use a diagonal lumped mass matrix, so these matrix-matrix multiplications are trivial.
 If a non-diagonal mass matrix is used, one may form a H-matrix representation of the mass matrix, then perform the matrix-matrix multiplications in \eqref{eq:boldA} using H-matrix methods.
@@
 We form $\mathbf{\Aker}_H$ using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package~\cite{Kriemann08}.
-For details about the algorithms used for geometrical clustering, H-matrix construction, and H-matrix operations in HLIBpro, we recommend the the following papers~\cite{BormGrasedyckHackbusch03,GrasedyckKriemannLeBorne08,Kriemann13}.
-Although $\mathbf{\Aker}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Aker}_H$ only requires evaluation of $O(\hrank^2 \fedim \log \fedim)$ kernel entries $\mathbf{\Aker}_{ij} = \widetilde{\Aker}(\fepoint_i, \fepoint_j)$, and these entries are computed via the radial basis function interpolation method described in Section~\ref{sec:approximate_kernel_entries}.
+For details about the algorithms used for geometrical clustering, H-matrix construction, and H-matrix operations in HLIBpro, we refer the reader to~\cite{BormGrasedyckHackbusch03,GrasedyckKriemannLeBorne08,Kriemann13}.
+Although $\mathbf{\Aker}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Aker}_H$ only requires evaluation of $O(\hrank \fedim \log \fedim)$ kernel entries $\mathbf{\Aker}_{ij} = \widetilde{\Aker}(\fepoint_i, \fepoint_j)$, and these entries are computed via the radial basis function interpolation method described in Section~\ref{sec:approximate_kernel_entries}.
 Here $\hrank$ is the rank of the highest rank block in the H-matrix.
-The dense matrix $\mathbf{\Aker}$ is never formed. 
+We emphasize that the dense matrix $\mathbf{\Aker}$ is never formed. 
 %We describe this H-matrix construction process in detail in Appendix~\ref{app:h_matrix}. 
@@
 However, this is computationally infeasible if $\Aop$ has a large or infinite cluster of eigenvalues near zero, a common situation for Hessians in ill-posed inverse problems.
-We therefore recommend choosing $\epsilon_\text{flip} < 0$, and we use $\epsilon_\text{flip}=-0.1$ in the numerical experiments.
+We therefore recommend choosing $\epsilon_\text{flip} < 0$.
+In our numerical results, we use $\epsilon_\text{flip}=-0.1$.
 
@@
 
-The computational cost of the proposed method may be divided into the costs to perform the following tasks: (1) Computing impulse response moments and batches; (2) Building the H-matrix; (3) Performing linear algebra operations with the H-matrix.
+The computational cost of the proposed method may be divided into the costs to perform the following tasks: (1) Computing impulse response moments and batches (Lines 1 and 4 in Algorithm \ref{alg:construct_Atilde}); (2) Building the H-matrix (Lines 5 and 6 in Algorithm \ref{alg:construct_Atilde}); (3) Performing linear algebra operations with the H-matrix.
 This may optionally include the symmetric positive semi-definite modifications described in Section \ref{sec:flip_eigs}.
@@
 The cost of picking the $i$th batch could be reduced to $O(\fedim \log |\pointbatch_i|)$ using more advanced computational geometry techniques, such as organizing the ellipsoids associated with points that have already been picked into a dynamic AABB tree, then using this AABB tree to accelerate the required ellipsoid intersection tests.
-However, from a practical perspective, the cost for picking sample points is small compared to other parts of the approximation algorithm, so such improvements are not pursued here.
+However, from a practical perspective, the cost for choosing sample points is small compared to other parts of the approximation algorithm, hence such improvements are not pursued here.
 
@@
 %, which is described in Appendix~\ref{app:h_matrix}, 
-require evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix.
+require evaluating $O(\hrank \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix.
 To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulse responses used in the RBF interpolation.
@@
 Third, one must evaluate finite element basis functions on each of those cells, which requires $O(\numnbr)$ elementary operations.
-Finally, one must solve a $\numnbr \times \numnbr$ linear system to perform the radial basis function interpolation, which requires $O(\numnbr^3)$ elementary operations, which typically is negligible.
+Finally, one must solve a $\numnbr \times \numnbr$ linear system to perform the radial basis function interpolation, which requires $O(\numnbr^3)$ elementary operations.
 Therefore, building the H-matrix requires
@@
 \begin{equation*}
-	O\left(\left(\hrank^2 \fedim \log \fedim\right)\left(\numnbr \log \fedim + \numnbr^3\right)\right) \qquad \text{elementary operations}.
+	O\left(\left(\hrank \fedim \log \fedim\right)\left(\numnbr \log \fedim + \numnbr^3\right)\right) \qquad \text{elementary operations}.
 \end{equation*}
@@
 
-\paragraph{(3) Performing linear algebra operations with the H-matrix} It is well known that H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, matrix inversion, and low rank updates require
-\begin{equation}
-\label{eq:H_matrix_cost_operations}
-	O(\hrank^2 \fedim \log(\fedim)^a) \qquad \text{elementary operations},
-\end{equation}
-where $a \in \{0,1,2,3\}$~\cite{GrasedyckHackbusch03}.
-The precise cost depends on the type of H-matrix used and the operation being performed.
+\paragraph{(3) Performing linear algebra operations with the H-matrix} It is well known that H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, matrix inversion, and low rank updates require performing $O\left(\hrank^a \fedim \log(\fedim)^b\right)$
+%\begin{equation}
+%\label{eq:H_matrix_cost_operations}
+%ASDASDAS	O(\hrank^2 \fedim \log(\fedim)^a) \qquad \text{elementary operations},
+%\end{equation}
+elementary operations, where $a,b \in \{0,1,2,3\}$ are constants which depend on the type of H-matrix used and the operation being performed~\cite{GrasedyckHackbusch03}\cite[Section 2.1]{Kriemann13}.
+We advise the reader to be aware that asymptotic complexity estimates for H-matrix operations can be a poor indicator of practical performance because of large prefactor constants, the complicated way in which recursive H-matrix algorithms map onto hardware, and difficulties taking advantage of parallelism.
+%In our experience with several software packages for H1 matrices, we have typically observed that matrix-vector products are extremely fast, matrix-matrix addition is fast, matrix LU factorization requires a comparatively moderate amount of time, factorized solves are extremely fast, and matrix-matrix multiplication and matrix inversion are comparatively slow.
 In the numerical results (Section~\ref{sec:numerical_results}), we use one matrix-matrix addition to add the H-matrix approximation of the data misfit term in the Hessian to the regularization term in the Hessian.
@@
 %In Figure~\ref{fig:condition_number} we show the condition numbers of $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$.
-The PSF preconditioners yield the the smallest condition numbers, with more batches yielding smaller condition numbers.
+The PSF preconditioners yield the smallest condition numbers, with more batches yielding smaller condition numbers.
 Regularization preconditioning and no preconditioning yield larger condition numbers.

### 2023-06-08 paper/localpsf.tex
dug up reference for O(k^2 N log N) operation count for H-matrix construction

@@
 For details about the algorithms used for geometrical clustering, H-matrix construction, and H-matrix operations in HLIBpro, we refer the reader to~\cite{BormGrasedyckHackbusch03,GrasedyckKriemannLeBorne08,Kriemann13}.
-Although $\mathbf{\Aker}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Aker}_H$ only requires evaluation of $O(\hrank \fedim \log \fedim)$ kernel entries $\mathbf{\Aker}_{ij} = \widetilde{\Aker}(\fepoint_i, \fepoint_j)$, and these entries are computed via the radial basis function interpolation method described in Section~\ref{sec:approximate_kernel_entries}.
+Although $\mathbf{\Aker}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Aker}_H$ only requires evaluation of $O(\hrank \fedim \log \fedim)$ kernel entries $\mathbf{\Aker}_{ij} = \widetilde{\Aker}(\fepoint_i, \fepoint_j)$ (see \cite{BebendorfRjasanow03}), and these entries are computed via the radial basis function interpolation method described in Section~\ref{sec:approximate_kernel_entries}.
 Here $\hrank$ is the rank of the highest rank block in the H-matrix.
@@
 
-\paragraph{(2) Building the H-matrix} Classical H-matrix construction techniques
+\paragraph{(2) Building the H-matrix} 
+Classical H-matrix construction techniques
 %, which is described in Appendix~\ref{app:h_matrix}, 
-require evaluating $O(\hrank \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix.
+require evaluating $O(\hrank \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix \cite{BebendorfRjasanow03}.
 To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulse responses used in the RBF interpolation.
@@
 %\end{equation}
-elementary operations, where $a,b \in \{0,1,2,3\}$ are constants which depend on the type of H-matrix used and the operation being performed~\cite{GrasedyckHackbusch03}\cite[Section 2.1]{Kriemann13}.
-We advise the reader to be aware that asymptotic complexity estimates for H-matrix operations can be a poor indicator of practical performance because of large prefactor constants, the complicated way in which recursive H-matrix algorithms map onto hardware, and difficulties taking advantage of parallelism.
+elementary operations\footnote{We advise the reader to be aware that asymptotic complexity estimates for H-matrix operations can be a poor indicator of practical performance because of large prefactor constants, the complicated way in which recursive H-matrix algorithms map onto hardware, and difficulties taking advantage of parallelism.}, where $a,b \in \{0,1,2,3\}$ are constants which depend on the type of H-matrix used and the operation being performed~\cite{GrasedyckHackbusch03}\cite[Section 2.1]{Kriemann13}.
 %In our experience with several software packages for H1 matrices, we have typically observed that matrix-vector products are extremely fast, matrix-matrix addition is fast, matrix LU factorization requires a comparatively moderate amount of time, factorized solves are extremely fast, and matrix-matrix multiplication and matrix inversion are comparatively slow.

### 2023-06-08 paper/localpsf.tex
add summary to computational cost section

@@
 %semi-definite, described in Appendix~\ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the rational function order.
+
+In summary, computing all the necessary ingredients to evaluate kernel entries of the PSF approximation requires a handful of operator applies (e.g., $6+\nbatch$ operator applies in two dimensions, or $10+\nbatch$ operator applies in three dimensions, with $\nbatch$ typically in the range 1--25), plus comparatively cheap additional overhead costs, most notably performing ellipsoid intersection tests while choosing sample point batches.
+Once these ingredients are computed, no more operator applies are required, and approximate kernel entries can be evaluated fast.
+Constructing the H-matrix from kernel entries requires a number of elementary operations that scales nearly linear in $N$.
+Using the resulting H-matrix to perform further linear algebra operations also scales nearly linearly in $N$, though the details of these costs depend heavily on the type of H-matrix and operation being performed.
 

### 2023-06-08 paper/localpsf.tex
grammar, typos, and other such edits

@@
 We use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applies.
-The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
+The impulse response, $\phi_x$, associated with a point, $x$, is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
 We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with batches of points scattered throughout the domain (see Figure~\ref{fig:batches_intro}).
@@
 In this work, we form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of a discretized version of $\Aop$.
-H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank.
+H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank.
 %(see Figure~\ref{fig:hmatrix_intro} and Appendix~\ref{app:h_matrix}).
@@
 In the probabilistic framework, Bayes' theorem combines the observations with prior information to form a posterior distribution over the space of all possible parameter fields, and computations are performed to extract statistical information about the parameter from this posterior.
-The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal to each other under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian.'' The Hessian consists of a data misfit term (the \emph{data misfit Hessian}), which depends on a discrepancy between the observations and the associated model predictions, and a regularization or prior term (the ``regularization Hessian'') which does not depend on the observations.
+The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal to each other under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian.'' The Hessian consists of a data misfit term (the \emph{data misfit Hessian}), which depends on a discrepancy between the observations and the associated model predictions, and a regularization or prior term (the \emph{regularization Hessian}) which does not depend on the observations.
 For more details on the Hessian, see~\cite{Alger19,GhattasWillcox21,VillaPetraGhattas21}.
@@
 Roughly speaking, the numerical rank of the data misfit Hessian is the dimension of the subspace of parameter space that is informed by the data.
-The numerical rank of the regularization preconditioned data misfit Hessian may be reduced by increasing the strength of the regularization, but this throws away useful information---components of the parameter that could be learned from components of the observations that have a large signal to noise ratio are instead reconstructed based on the regularization~\cite[Section 4]{AlgerEtAl17}\cite[Chapters 1 and 7]{Vogel02}.
-Hence, low rank approximation methods suffer from a predicament---if the data highly inform the parameter and the regularization is chosen appropriately, then a large number of operator applies are required to form an accurate approximation of the Hessian using low rank approximation methods.
+The numerical rank of the regularization preconditioned data misfit Hessian may be reduced by increasing the strength of the regularization, but this throws away useful information.
+If the regularization is too strong, then components of the parameter that could be learned from the observations
+%components of the observations that have a large signal to noise ratio 
+would instead be reconstructed based on the regularization~\cite[Section 4]{AlgerEtAl17}\cite[Chapters 1 and 7]{Vogel02}.
+Hence, low rank approximation methods suffer from a predicament: if the data highly inform the parameter and the regularization is chosen appropriately, then a large number of operator applies are required to form an accurate approximation of the Hessian using low rank approximation methods.
 %The best case scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost)~\cite{AlexanderianGloorGhattas16,Alger19,AlgerEtAl17}.
@@
 These methods (which we do not use here) have been used to approximate Hessians in PDE constrained inverse problems~\cite{AmbartsumyanEtAl20,HartlandEtAl23}. 
-Although the required number of matrix-vector products grows nearly linearly in the discretized parameter dimension and linearly in the rank of the off-diagonal blocks, the required number of matrix-vector products is still large (e.g., hundreds to thousands).
+Although these methods are promising,
+%required number of matrix-vector products grows logarithmically in the discretized parameter dimension and linearly in the rank of the off-diagonal blocks, 
+the required number of operator applies is still large (e.g., hundreds to thousands).
 %These methods are asymptotically scalable in the sense that the required number of matrix-vector products grows nearly linearly in the dimension of the discretized parameter and linearly in the rank of the off-diagonal blocks.
@@
 In this paper, we also form an H-matrix Hessian approximation.
-However, to reduce the required number of operator actions, we first form a PSF approximation of the data misfit Hessian by exploiting locality and non-negative integral kernel properties, then form the H-matrix using classical H-matrix techniques.
+However, to reduce the required number of operator applies, we first form a PSF approximation of the data misfit Hessian by exploiting locality and non-negative integral kernel properties, then form the H-matrix using classical H-matrix techniques.
 Not all data misfit Hessians satisfy the local non-negative integral kernel properties (for example, the wave inverse problem data misfit Hessian has a substantial amount of negative entries in its integral kernel).
@@
 The integral kernel, $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$, exists in principle but is not easily accessible.
-In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define impulse responses (Section~\ref{sec:impulse_response_2}), we then state the conditions on $\Aop$ that the proposed method requires (Section~\ref{sec:conditions_2}), and detail finite element discretization (Section~\ref{sec:finite_element_kernel}).
+In this section we describe how to extend the application of $\Aop$ to distributions, which allows us to define impulse responses (Section~\ref{sec:impulse_response_2}), we then state the conditions on $\Aop$ that the proposed method requires (Section~\ref{sec:conditions_2}), and detail finite element discretization (Section~\ref{sec:finite_element_kernel}).
 
@@
 
-The action of $\Aop$ may be extended to distributions if $\Aker$ is sufficiently regular. 
+The application of $\Aop$ may be extended to distributions if $\Aker$ is sufficiently regular. 
 %We derive the extension as follows. 
@@
 However, if $\Aker$ is sufficiently regular such that the function $y \mapsto \rho\left(\Aker(y,~\cdot~)\right)$ is well-defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of~\eqref{eq:action_on_distribution} is well-defined.
-Hence, we \emph{define} the action of $\Aop$ on the distribution $\rho$ to be the right hand side of~\eqref{eq:action_on_distribution}.
-We denote this action by ``$\Aop \genericdistribution^*$,'' even if $\rho^*$ does not exist.
+Hence, we \emph{define} the application of $\Aop$ on the distribution $\rho$ to be the right hand side of~\eqref{eq:action_on_distribution}.
+We denote this operator application by ``$\Aop \genericdistribution^*$,'' even if $\rho^*$ does not exist.
 
@@
 It is acceptable if the support of $\impulseresponse_x$ is not perfectly contained in a localized region, so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
-The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$, as long as it is non-negative for the vast majority of pairs of points $(y,x)$, and as long as the negative numbers are comparatively small.
+The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$ so long as it is non-negative for the vast majority of pairs of points $(y,x)$, and so long as the negative numbers are comparatively small.
 If these conditions are violated, the method will incur additional error, and depending on the severity of the violation, the proposed method may fail.
@@
 \end{equation}
-for all $x_i \in \pointbatch_k$.
-By performing one matrix-vector product, $\xi_k \mapsto \left(\Aop \diraccomb_k^*\right)^*$, we recover $\impulseresponse_{x_i}$ for every point $x_i \in \pointbatch_k$.
+for all $x_i \in \pointbatch_k$. By applying the operator once, $\xi_k \mapsto \left(\Aop \diraccomb_k^*\right)^*$, we recover $\impulseresponse_{x_i}$ for every point $x_i \in \pointbatch_k$.
 
@@
 where $\{x_i\}_{i=1}^{\numnbr}$ are the $\numnbr$ nearest sample points to $x$, excluding points $x_i$ for which $z_i \notin \Omega$.
-Here $\numnbr$ is a small user defined parameter; e.g., $\numnbr=10$.
+Here $\numnbr$ is a small user-defined parameter, e.g., $\numnbr=10$.
 We find the $\numnbr$ nearest sample points to $x$ by querying a precomputed kd-tree~\cite{Bentley75} of all sample points.
@@
 Note that $\phi_{x_i}\left(z_i\right)$ is well-defined because $z_i \in \Omega$, and $\frac{\spatialvol(x)}{\spatialvol(x_i)}$ is well-defined because the sample point choosing procedure in Section~\ref{sec:sample_point_selection} ensures that $\spatialvol(x_i)>0$.
-Per the discussion in Section~\ref{sec:local_mean_displacement_invariance}, we expect $\Aker(y,x) \approx f_i$ for $i=1,\dots,\numnbr$; the closer $x_i$ is to $x$, the better we expect the approximation to be.
+Per the discussion in Section~\ref{sec:local_mean_displacement_invariance}, we expect $\Aker(y,x) \approx f_i$ for $i=1,\dots,\numnbr$.
+The closer $x_i$ is to $x$, the better we expect the approximation to be.
 We therefore approximate $\Aker(y,x)$ by interpolating the (point,value) pairs
@@
 is a Gaussian kernel radial basis function.
-Here $L:=\diam\left(\{x_i\}_{i=1}^{\numnbr}\right)$ is the diameter of the set of sample points used in the interpolation, and $C_\text{RBF}$ is a user-defined shape parameter that controls the width of the kernel function; we use $C_\text{RBF}=3.0$.
+Here $L:=\diam\left(\{x_i\}_{i=1}^{\numnbr}\right)$ is the diameter of the set of sample points used in the interpolation, and $C_\text{RBF}$ is a user-defined shape parameter that controls the width of the kernel function (we use $C_\text{RBF}=3.0$).
 %is a polyharmonic spline radial basis function, which is defined as $\rbf(r):=r^\gdim$ if $\gdim$ is odd, and $\rbf(r):=r^\gdim \log r$ for $r>0$ and $\rbf(0):=0$ if $\gdim$ is even.
@@
 We modify $\widetilde{\bm{H}}$ to be (approximately) symmetric positive semi-definite via the procedure described in Section~\ref{sec:flip_eigs}.
-We factor $\bm{\preconditioner}$,
+We factor $\bm{\preconditioner}$
 using fast H-matrix methods, then use the factorization as a
@@
 	\caption{(Ice sheet) 
-		(\ref{fig:ice_mountain_mesh}) Bird's eye view of the ice sheet, discretized by a mesh of tetrahedra.
+		(\ref{fig:ice_mountain_mesh}) Bird's eye view of the ice sheet discretized by a mesh of tetrahedra.
 Color indicates the height of the base of the ice sheet (i.e., the mountain topography).
@@
 		(\ref{fig:true_beta}) True parameter, $\basalfriction_\text{true}$.
-(\ref{fig:stokes_velocity}) True velocity $\velocity_\text{true}$.
+(\ref{fig:stokes_velocity}) True velocity, $\velocity_\text{true}$.
 Arrows indicate the direction of $\velocity_\text{true}$ and color indicates the magnitude of $\velocity_\text{true}$.}

### 2023-06-08 paper/localpsf.tex
paper is back under 26 pages

@@
 Here ``matrix-free'' means that we may apply $\Aop$ and its transpose\footnote{Recall that $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for operations with $\Aop^T$.
-}, $\Aop^T$, to arbitrary functions,
+}, $\Aop^T$, to functions,
 \begin{equation}
@@
 In the probabilistic framework, Bayes' theorem combines the observations with prior information to form a posterior distribution over the space of all possible parameter fields, and computations are performed to extract statistical information about the parameter from this posterior.
-The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal to each other under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian.'' The Hessian consists of a data misfit term (the \emph{data misfit Hessian}), which depends on a discrepancy between the observations and the associated model predictions, and a regularization or prior term (the \emph{regularization Hessian}) which does not depend on the observations.
+The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian.'' The Hessian consists of a data misfit term (the \emph{data misfit Hessian}), which depends on a discrepancy between the observations and the associated model predictions, and a regularization or prior term (the \emph{regularization Hessian}) which does not depend on the observations.
 For more details on the Hessian, see~\cite{Alger19,GhattasWillcox21,VillaPetraGhattas21}.
@@
 When solving the deterministic optimization problem with Newton-type methods, the Hessian is the coefficient operator for the linear system that must be solved or approximately solved at every Newton iteration.
-Good Hessian preconditioners reduce the number of iterations required to solve these Newton linear systems with Krylov methods such as the conjugate gradient method~\cite{Saad03}.
+Good Hessian preconditioners reduce the number of iterations required to solve these Newton linear systems with the conjugate gradient method~\cite{Saad03}.
 In the Bayesian setting, the inverse of the Hessian is the covariance of a local Gaussian approximation of the posterior.
-This Gaussian distribution can be used directly as an approximation of the posterior, or it can be utilized as a proposal for Markov chain Monte-Carlo methods for drawing samples from the posterior.
+This Gaussian distribution can be used directly as an approximation of the posterior, or it can be used as a proposal for Markov chain Monte-Carlo methods for drawing samples from the posterior.
 For instance, see~\cite{KimEtAl21,PetraEtAl14} and the references therein.
@@
 %(see Figure~\ref{fig:hessian_eigenstructure})
-hence more informative data leads to larger eigenvalues and therefore a larger numerical rank~\cite{AlexanderianGloorGhattas16}\cite[Section 1.4 and Chapter 4]{Alger19}.
+hence more informative data leads to larger eigenvalues and a larger numerical rank~\cite{AlexanderianGloorGhattas16}\cite[Section 1.4 and Chapter 4]{Alger19}.
 Roughly speaking, the numerical rank of the data misfit Hessian is the dimension of the subspace of parameter space that is informed by the data.
-The numerical rank of the regularization preconditioned data misfit Hessian may be reduced by increasing the strength of the regularization, but this throws away useful information.
-If the regularization is too strong, then components of the parameter that could be learned from the observations
+The numerical rank of the regularization preconditioned data misfit Hessian may be reduced by increasing the strength of the regularization, but this throws away useful information: components of the parameter that could be learned from the observations
 %components of the observations that have a large signal to noise ratio 
@@
 \end{equation}
-Here the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$.
-The integral kernel, $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$, exists in principle but is not easily accessible.
-In this section we describe how to extend the application of $\Aop$ to distributions, which allows us to define impulse responses (Section~\ref{sec:impulse_response_2}), we then state the conditions on $\Aop$ that the proposed method requires (Section~\ref{sec:conditions_2}), and detail finite element discretization (Section~\ref{sec:finite_element_kernel}).
+The linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and the scalar $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$.
+The integral kernel, $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$, exists but is not easily accessible.
+In this section we describe how to extend the domain of $\Aop$ to distributions, which allows us to define impulse responses (Section~\ref{sec:impulse_response_2}), we then state the conditions on $\Aop$ that the proposed method requires (Section~\ref{sec:conditions_2}), and detail finite element discretization (Section~\ref{sec:finite_element_kernel}).
 
@@
 
-The application of $\Aop$ may be extended to distributions if $\Aker$ is sufficiently regular. 
+The operator $\Aop$ may be applied to distributions if $\Aker$ is sufficiently regular. 
 %We derive the extension as follows. 
-Let $\rho \in L^2(\Omega)'$, and let $\rho^* \in L^2(\Omega)$ denote the Riesz representative of $\rho$ with respect to the $L^2(\Omega)$ inner product. We have
+Given $\rho \in L^2(\Omega)'$, let $\rho^* \in L^2(\Omega)$ denote the Riesz representative of $\rho$ with respect to the $L^2(\Omega)$ inner product. We have
 \begin{subequations}
@@
 where $\Aker(y,\cdot)$ denotes the function $x \mapsto \Aker(y,x)$.
-Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions and let $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ be a distribution.
-In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in~\eqref{eq:extension_to_distributions} is not valid.
+Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions and let $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ be a distribution. In this case, 
+%the Riesz representative 
+$\rho^*$ may not exist, so the derivation in~\eqref{eq:extension_to_distributions} is not valid.
 However, if $\Aker$ is sufficiently regular such that the function $y \mapsto \rho\left(\Aker(y,~\cdot~)\right)$ is well-defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of~\eqref{eq:action_on_distribution} is well-defined.
-Hence, we \emph{define} the application of $\Aop$ on the distribution $\rho$ to be the right hand side of~\eqref{eq:action_on_distribution}.
+Hence, we \emph{define} the application of $\Aop$ to the distribution $\rho$ to be the right hand side of~\eqref{eq:action_on_distribution}.
 We denote this operator application by ``$\Aop \genericdistribution^*$,'' even if $\rho^*$ does not exist.
@@
 After discretization, the operator $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$ is replaced by an operator $A_h:V_h \rightarrow V_h'$, which becomes an operator
-\begin{equation*}
+$
 \mathbf{A}:\mathbb{R}^\fedim_\massmatrix \rightarrow \mathbb{R}^\fedim_{\massmatrix^{-1}}
-\end{equation*}
+$
 under the isomorphism discussed above.
@@
 Rather, we have a computational procedure that allows us to compute matrix-vector products $\mathbf{u}\mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{w}\mapsto \mathbf{A}^T\mathbf{w}$, and computing these matrix-vector products is costly.
-The proposed method is designed to mitigate this cost by performing as few of these matrix-vector products as possible.
+The proposed method mitigates this cost by performing as few matrix-vector products as possible.
 Of course, matrix entries can be computed via matrix-vector products as $\mathbf{A}_{ij} = \left(\mathbf{A}\mathbf{e}_j\right)_i$,
@@
 We therefore recommend replacing the mass matrix, $\massmatrix$, with a positive diagonal \emph{lumped mass} approximation.
-In the numerical results, we use the lumped mass matrix in which the $i$th diagonal entry of the lumped mass matrix is the sum of all entries in the $i$th row of the mass matrix.
+Here, we use the lumped mass matrix in which the $i$th diagonal entry of the lumped mass matrix is the sum of all entries in the $i$th row of the mass matrix.
 Other mass lumping techniques may be used~\cite{Hughes12}.
@@
 These are four different ways to approximate an impulse response by a nearby impulse response, with each successive approximation building upon the previous ones.
-The proposed method uses~\eqref{eq:translate4}, which is the most sophisticated of these approximations.
+The proposed method uses~\eqref{eq:translate4}, which is the most sophisticated.
 Approximation~\eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close.
@@
 But approximation~\eqref{eq:translate3} can hold in situations where neither~\eqref{eq:translate1} nor~\eqref{eq:translate2} holds.
-For example, because the expected value commutes with affine transformations,~\eqref{eq:translate3} will hold when $\Aop$ is locally translation invariant with respect to a translated and rotated frame of reference, while~\eqref{eq:translate2} will not hold in this case.
+For example, because the expected value commutes with affine transformations,~\eqref{eq:translate3} will hold when $\Aop$ is locally translation invariant with respect to a translated and rotated frame of reference, while~\eqref{eq:translate2} will not.
 Additionally,~\eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions.
@@
 In this method, one forms low rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks.
-We then use H-matrix methods to convert $\boldsymbol{\Aker}_H$ into an H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$.
+We then convert $\boldsymbol{\Aker}_H$ into an H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$.
 %Schematically the algorithm proceeds as follows:
@@
 Given $(y,x)\in \Omega \times \Omega$, let
-\begin{equation*}
-	z_i := y - \spatialmean(x) + \spatialmean(x_i), \qquad i=1,\dots,\numnbr,
-\end{equation*}
-and
+$
+	z_i := y - \spatialmean(x) + \spatialmean(x_i)
+$
+and define
 \begin{equation}
 \label{eq:fxyxp}
-f_i := \frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}\left(z_i\right), \qquad i=1,\dots,\numnbr,
+f_i := \frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}\left(z_i\right)
 \end{equation}
-where $\{x_i\}_{i=1}^{\numnbr}$ are the $\numnbr$ nearest sample points to $x$, excluding points $x_i$ for which $z_i \notin \Omega$.
+for $i=1,\dots,\numnbr$, where $\{x_i\}_{i=1}^{\numnbr}$ are the $\numnbr$ nearest sample points to $x$, excluding points $x_i$ for which $z_i \notin \Omega$.
 Here $\numnbr$ is a small user-defined parameter, e.g., $\numnbr=10$.
@@
 where $\rbfweight_i$ are weights, and 
-$$\rbf(r):= \exp\left(-C_\text{RBF}^2\frac{r^2}{2 L^2}\right)$$
+$\rbf(r):= \exp\left(-C_\text{RBF}^2\frac{r^2}{2 L^2}\right)$
 is a Gaussian kernel radial basis function.
@@
 \end{equation}
-where $B \in \mathbb{R}^{\numnbr \times \numnbr}$ has entries
+where $B \in \mathbb{R}^{\numnbr \times \numnbr}$,
 $
@@
 %, which is described in Appendix~\ref{app:h_matrix}, 
-require evaluating $O(\hrank \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix \cite{BebendorfRjasanow03}.
+require evaluating $O(\hrank \fedim \log \fedim)$ matrix entries of the approximation \cite{BebendorfRjasanow03}, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix.
 To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulse responses used in the RBF interpolation.
@@
 This is done using an AABB tree of mesh cells, and requires $O(\numnbr \log \fedim)$ elementary operations.
-Third, one must evaluate finite element basis functions on each of those cells, which requires $O(\numnbr)$ elementary operations.
-Finally, one must solve a $\numnbr \times \numnbr$ linear system to perform the radial basis function interpolation, which requires $O(\numnbr^3)$ elementary operations.
+Third, one must evaluate finite element basis functions on those cells, which requires $O(\numnbr)$ elementary operations.
+Finally, the radial basis function interpolation requires solving a $\numnbr \times \numnbr$ linear system, which requires $O(\numnbr^3)$ elementary operations.
 Therefore, building the H-matrix requires
@@
 %\end{equation}
-elementary operations\footnote{We advise the reader to be aware that asymptotic complexity estimates for H-matrix operations can be a poor indicator of practical performance because of large prefactor constants, the complicated way in which recursive H-matrix algorithms map onto hardware, and difficulties taking advantage of parallelism.}, where $a,b \in \{0,1,2,3\}$ are constants which depend on the type of H-matrix used and the operation being performed~\cite{GrasedyckHackbusch03}\cite[Section 2.1]{Kriemann13}.
+elementary operations,
+%\footnote{We advise the reader to be aware that asymptotic complexity estimates for H-matrix operations can be a poor indicator of practical performance because of large prefactor constants, the complicated way in which recursive H-matrix algorithms map onto hardware, and difficulties taking advantage of parallelism.},
+where $a,b \in \{0,1,2,3\}$ are constants which depend on the type of H-matrix used and the operation being performed~\cite{GrasedyckHackbusch03}\cite[Section 2.1]{Kriemann13}.
 %In our experience with several software packages for H1 matrices, we have typically observed that matrix-vector products are extremely fast, matrix-matrix addition is fast, matrix LU factorization requires a comparatively moderate amount of time, factorized solves are extremely fast, and matrix-matrix multiplication and matrix inversion are comparatively slow.
@@
 Symmetrizing $\mathbf{A}_H$ requires one matrix-matrix addition.
-Flipping negative eigenvalues to be positive requires a handful (typically around 5) matrix-matrix additions and matrix factorizations to factor the required shifted linear systems, and a number of factorized solves with these shifted linear systems that is proportional to the number of erroneous negative eigenvalues.
+Flipping negative eigenvalues to be positive requires a handful (typically around 5) of matrix-matrix additions and matrix factorizations to factor the required shifted linear systems, and a number of factorized solves that is proportional to the number of erroneous negative eigenvalues.
 %semi-definite, described in Appendix~\ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the rational function order.
@@
 Here, $\basalfriction$ is the unknown logarithmic
-basal sliding field (note that large $\basalfriction$ corresponds to high resistance to sliding) which is defined on the $2$D surface $\Gamma_b$.
+basal sliding field (large $\basalfriction$ corresponds to high resistance to sliding) defined on the $2$D surface $\Gamma_b$.
 The quantity $\bodyforce$ is the body force density due to gravity, $\stokesrobincoeff=10^6$ is a Robin boundary condition constant,  $\normalvec$ is the outward normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary.
@@
 We use PCG to solve a linear system with the Hessian as the coefficient operator and a right hand vector with random independent and identically distributed (i.i.d.) entries drawn from the standard Gaussian distribution.
-In the left subfigure of Figure~\ref{fig:krylov_convergence} we compare the convergence of PCG for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners.
+In Figure~\ref{fig:krylov_convergence} (left) we compare the convergence of PCG for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners.
 PCG converges fastest with the PSF-based preconditioners, with PSF (25) converging fastest, followed by PSF (5), followed by PSF (1), as expected.
 PCG converges much slower with no preconditioning and regularization preconditioning than it does with PSF-based preconditioning, with no preconditioning outperforming regularization preconditioning.
-In the right subfigure of Figure~\ref{fig:krylov_convergence} we show the generalized eigenvalues for the generalized eigenvalue problem
+In Figure~\ref{fig:krylov_convergence} (right) we show the generalized eigenvalues for the generalized eigenvalue problem
 $
@@
 The initial condition, velocity field, noisy observations, and reconstructed initial condition are shown in Figure~\ref{fig:adv_inversion1}.
-We use $\kappa=3.2$e${-1}$ and $T=1.0$ for all results, except for Table \ref{tab:adv_convergence} and Figure~\ref{tbl:adv_vary_kappa_tf} where we vary $\kappa$ and $T$ to study the impact of changing these parameters.
+We use $\kappa=3.2$e${-1}$ and $T=1.0$ for all results, except for Table \ref{tab:adv_convergence} and Figure~\ref{tbl:adv_vary_kappa_tf} where we vary $\kappa$ and $T$.
+% to study the impact of changing these parameters.
 
@@
 
-In the left subfigure of Figure~\ref{fig:adv_krylov_convergence} we show the convergence of PCG for solving $\bm{H}\bm{x}=\bm{b}$, where $\bm{b}$ has i.i.d.
+In Figure~\ref{fig:adv_krylov_convergence} (left) we show the convergence of PCG for solving $\bm{H}\bm{x}=\bm{b}$, where $\bm{b}$ has i.i.d.
 random entries drawn from a standard Gaussian distribution.
-The preconditioners used, $\widetilde{\bm{H}}$, are the PSF-based preconditioners with $1$, $5$, or $25$ batches, regularization preconditioning, and no preconditioning.
+The preconditioners used, $\widetilde{\bm{H}}$, are the PSF-based preconditioners with $1$, $5$, or $25$ batches, the regularization Hessian, and the identity matrix (i.e., no preconditioning).
 The results show that PCG converges fastest with the PSF-based preconditioners, with more batches yielding faster convergence.
-In the right subfigure of Figure~\ref{fig:adv_krylov_convergence}, we show the generalized eigenvalues for the generalized eigenvalue problem $\bm{H}\bm{u}=\lambda \widetilde{\bm{H}}\bm{u}$, where the $\widetilde{\bm{H}}$ are the preconditioners stated above.
-As can be seen, with the PSF-based preconditioners, the eigenvalues cluster near one; more batches yields better clustering.
+In Figure~\ref{fig:adv_krylov_convergence} (right), we show the eigenvalues for the generalized eigenvalue problem $\bm{H}\bm{u}=\lambda \widetilde{\bm{H}}\bm{u}$, where the $\widetilde{\bm{H}}$ are the preconditioners stated above.
+With the PSF-based preconditioners the eigenvalues cluster near one, and more batches yields better clustering.
 With the regularization preconditioner the trailing eigenvalues cluster near one, while the leading eigenvalues are amplified.
@@
 	\end{center}
-	\caption{(Diffusive transport) Number of PCG iterations required to solve the Newton linear system to tolerance $||\widehat{\bm{q}}_j - \widehat{\bm{q}}|| < 10^{-6}||\widehat{\bm{q}}||$, where $\widehat{\bm{q}}_j$ is the $j$th conjugate gradient iterate, and $\widehat{\bm{q}}$ is the solution of the Newton linear system.
+	\caption{(Diffusive transport) Number of PCG iterations required to solve the Newton linear system to tolerance $||\widehat{\bm{q}}_j - \widehat{\bm{q}}|| < 10^{-6}||\widehat{\bm{q}}||$, where $\widehat{\bm{q}}_j$ is the $j$th iterate, and $\widehat{\bm{q}}$ is the solution of the Newton linear system.
 Iteration counts are shown for a variety of different diffusion parameters $\kappa$, simulation times $T$, and preconditioners.}
@@
 We thank Trevor Heise for editing suggestions.
-We thank Georg Stadler for his help with the domain setup for the ice sheet inverse problem.
+We thank Georg Stadler for help with the domain setup for the ice sheet problem.
 

### 2023-06-08 paper/localpsf.tex
Update abstract to include Noemi's edits, plus some of my additional edits

@@
 The method computes impulse responses of the operator at a collection of scattered points, and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel.
+By applying the operator to Dirac combs associated with batches of point sources, we compute a large number of impulse responses using a small number of operator applies.
+We solve an ellipsoid packing problem to select as many point sources as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
+Support ellipsoids are estimated a-priori based on impulse response moments, which are computed by applying the operator to a small number of polynomial functions.
+The interpolation uses a local mean displacement invariance approximation, which generalizes classical local translation invariance.
+The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
+Further matrix computations are performed with H-matrix methods.
+We apply the proposed method to approximate the Hessian operator arising in inverse problems governed by partial differential equations (PDEs).
+While for many ill-posed inverse problems the Hessian of the data misfit term exhibits a low rank structure, and hence a low rank approximation is suitable, for many problems of practical interest the numerical rank of this operator is still large.
+This makes conventional matrix-free Hessian approximation methods expensive in terms of the required number of Hessian operator applies (hence PDE solves).
+In contrast, the performance of the proposed method depends on the locality and shape of the impulse response rather than the numerical rank.
+Hessian impulse responses typically become more local as the numerical rank increases, which makes this operator a perfect candidate for the proposed approach.
+We illustrate the proposed approach with two specific problems: inversion for the basal sliding coefficient in an ice sheet flow problem and inversion for the initial condition in a diffusive transport problem.
+To demonstrate the effectiveness of the approximation, we use the proposed approach to build preconditioners to accelerate the solution of the inverse problems.
+We compare the results with classical regularization-based preconditioning and no preconditioning.
+The numerical results demonstrate the superior performance (measured in number of PDE solves) of the inverse solver when using the proposed PSF-based preconditioner.
+Finally, we present a comprehensive numerical study for the influence of the diffusion parameters and simulation time (for the diffusive transport problem) on the effectiveness of the approximation.
+These parameters have an effect on the shape of the impulse responses and on the rank of the Hessian.
+The results reveal that PSF-based preconditioners are able to form good approximations of high-rank Hessians using a small number of Hessian operator applies.
+
+%	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels.
+%	The method computes impulse responses of the operator at a collection of scattered points, and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel.
 By applying the operator to Dirac combs associated with batches of point sources, we compute a large number of impulse responses using a small number of operator applies.

### 2023-06-08 paper/localpsf.tex
Final draft being sent to Omar

@@
 This makes conventional matrix-free Hessian approximation methods expensive in terms of the required number of Hessian operator applies (hence PDE solves).
-In contrast, the performance of the proposed method depends on the locality and shape of the impulse response rather than the numerical rank.
+In contrast, the performance of the proposed method depends on the locality and shapes of the impulse responses rather than the numerical rank.
 Hessian impulse responses typically become more local as the numerical rank increases, which makes this operator a perfect candidate for the proposed approach.
@@
 These parameters have an effect on the shape of the impulse responses and on the rank of the Hessian.
-The results reveal that PSF-based preconditioners are able to form good approximations of high-rank Hessians using a small number of Hessian operator applies.
+The results reveal that PSF-based preconditioners are able to form good approximations of high-rank Hessians using a small number of operator applies.
 
@@
 \paragraph{Lumped mass matrix} At the continuum level, $\Aker$ is assumed to be non-negative.
-However, entries of $\Aker_h$, involve inverse mass matrices, which typically contain negative numbers.
-If there are too many negative numbers, or if the negative numbers are too large, the proposed method will yield less accurate approximations and may potentially fail.
+However, entries of $\boldsymbol{\Aker}$ involve inverse mass matrices, which typically contain negative numbers.
+%If there are too many negative numbers, or if the negative numbers are too large, the proposed method will yield less accurate approximations and may potentially fail. 
 We therefore recommend replacing the mass matrix, $\massmatrix$, with a positive diagonal \emph{lumped mass} approximation. 
@@
 \label{eq:define_vol}
-\spatialvol(x) := \int_\Omega \impulseresponse_x(y) \mathrm{d}y
+\spatialvol(x) := \int_\Omega \impulseresponse_x(y) \mathrm{d}y,
 \end{equation}
@@
 
-The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to obtain $\impulseresponse_x$, as per~\eqref{eq:impulse_response_delta_action}.
+The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to obtain $\impulseresponse_x$, per~\eqref{eq:impulse_response_delta_action}.
 Then one can post process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
-However, this direct approach is not feasible because we need to know $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ before we compute $\phi_x$, in order choose the point $x$.
+However, this direct approach is not feasible because we need to know $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ before we compute $\phi_x$ in order choose the point $x$.
 Computing $\phi_x$ in order to determine $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ would be extremely computationally expensive, and defeat the purpose of the proposed method, which is to reduce the computational cost by computing impulse responses in batches.
@@
 However, in Section~\ref{sec:sample_point_selection} we will see that the cost of the proposed method depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids is better), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
-In practice, we find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in the numerical results.
+In practice, we find that $\tau=3.0$ yields a reasonable balance between these competing interests, and use $\tau=3.0$ in the numerical results.
 The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is typically conservative.
@@
 
-Let $x$ and $x'$ be points in $\Omega$ that are ``close'' to each other, and consider the following approximations:
+Let $x$ and $x'$ be points in $\Omega$ that are close to each other, and consider the following approximations:
 \begin{align}
@@
 We compute one batch, denoted $\eta_b$, by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section~\ref{sec:get_impulse_response}).
-The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support ellipsoid for $\impulseresponse_{x_i}$ and the support ellipsoid for $\impulseresponse_{x_j}$ do not overlap if $i \neq j$ (Section~\ref{sec:sample_point_selection}).
+The batch of points, $S_b$, is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support ellipsoid for $\impulseresponse_{x_i}$ and the support ellipsoid for $\impulseresponse_{x_j}$ do not overlap if $i \neq j$ (Section~\ref{sec:sample_point_selection}).
 Because these supports do not overlap (or do not overlap much), we can post process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$.
@@
 	}
-	\caption{Construct H-matrix approximation}
+	\caption{Construct PSF H-matrix approximation}
 	\label{alg:construct_Atilde}
@@
 
-We compute impulse responses $\impulseresponse_{x_i}$ in batches by applying $\Aop$ to Dirac combs.
+We compute impulse responses, $\impulseresponse_{x_i}$, in batches by applying $\Aop$ to Dirac combs.
 The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the following weighted sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$:
@@
 
-Each point source $\delta_{x_i}$ is scaled by $1/\spatialvol(x_i)$ so that the resulting scaled impulse responses within $\eta_k$ are comparable in magnitude.
+Each point source, $\delta_{x_i}$, is scaled by $1/\spatialvol(x_i)$ so that the resulting scaled impulse responses within $\eta_k$ are comparable in magnitude.
 Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, may overwhelm $\phi_{x_j}$ for a nearby point $x_j$ if $\spatialvol(x_i)$ is much larger than $\spatialvol(x_j)$.
@@
 Here, $\basalfriction$ is the unknown logarithmic
-basal sliding field (large $\basalfriction$ corresponds to high resistance to sliding) defined on the $2$D surface $\Gamma_b$.
+basal sliding field (large $\basalfriction$ corresponds to high resistance to sliding) defined on the surface $\Gamma_b$.
 The quantity $\bodyforce$ is the body force density due to gravity, $\stokesrobincoeff=10^6$ is a Robin boundary condition constant,  $\normalvec$ is the outward normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary.
@@
 PCG converges fastest with the PSF-based preconditioners, with PSF (25) converging fastest, followed by PSF (5), followed by PSF (1), as expected.
-PCG converges much slower with no preconditioning and regularization preconditioning than it does with PSF-based preconditioning, with no preconditioning outperforming regularization preconditioning.
+%PCG converges much slower with no preconditioning and regularization preconditioning than it does with PSF-based preconditioning, with no preconditioning outperforming regularization preconditioning.
 In Figure~\ref{fig:krylov_convergence} (right) we show the generalized eigenvalues for the generalized eigenvalue problem
@@
 The solution of the Newton system that we compare to, $\widehat{\mathbf{q}}$, is found via another PCG iteration with a relative residual tolerance of $10^{-11}$.
-We show results for $T$ ranging from $0.5$ to $2.0$ and $\kappa$ ranging from $10^{-4}$ to $10^{-3}$, using the PSF-based preconditioners with $1$, $5$, and $25$ batches, regularization preconditioning, and no preconditioning.
+We show results for $T$ ranging from $0.5$ to $2.0$ and $\kappa$ ranging from $10^{-4}$ to $10^{-3}$ using the PSF-based preconditioners with $1$, $5$, and $25$ batches, regularization preconditioning, and no preconditioning.
 The results show that PSF-based preconditioning outperforms regularization preconditioning and no preconditioning, with more batches yielding better results.

### 2023-06-22 paper/localpsf.tex
remove commented out parts of paper

@@
 The results reveal that PSF-based preconditioners are able to form good approximations of high-rank Hessians using a small number of operator applies.
-
-%	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels.
-%	The method computes impulse responses of the operator at a collection of scattered points, and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel.
-By applying the operator to Dirac combs associated with batches of point sources, we compute a large number of impulse responses using a small number of operator applies.
-We solve an ellipsoid packing problem to select as many point sources as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
-Support ellipsoids are estimated a-priori based on impulse response moments, which are computed by applying the operator to a small number of polynomial functions.
-The interpolation uses a local mean displacement invariance approximation, which generalizes classical local translation invariance.
-The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
-Further matrix computations are performed with H-matrix methods.
-We illustrate the method by approximating the Hessians in two example problems: an ice sheet flow inverse problem governed by a Stokes partial differential equation (PDE), and a diffusive transport inverse problem governed by an advection-diffusion PDE.
-Because the data in these inverse problems are highly informative about the unknown parameters, the data misfit terms in the Hessians are high rank.
-This makes conventional matrix-free Hessian approximation methods, which are based on low rank approximation, extremely expensive in terms of the required number of Hessian operator applies.
-In contrast, the numerical results show that the proposed method can form good approximations of these high rank Hessians using a small number of Hessian operator applies.
-We use the Hessian approximations as preconditioners to accelerate the solution of the inverse problems.
-Numerical results demonstrate that this approach substantially reduces the number of PDE solves required to solve the inverse problems.
-
-%	, as compared to existing methods.
-
-%	, as measured by the total number of PDE solves required to solve the inverse problem.
-
-%	, with a focus on approximating high rank Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs). 
-%	The method may be used to approximate other operators, including certain Schur complements, Poincare-Steklov operators, covariance operators, and blurring operators.
-
-%	that arise in inverse problems with highly informative data.
-
-%	we see that the numerical results show that the proposed method can approximate high Existing methods typically perform poorly  when the data is highly informative about the unknown parameter, because the Hessians in these problems are high rank.
-In contrast, the proposed method can cheaply approximate high rank Hessians.
-
-%	, in the sense that it can approximate Hessians in inverse problem with highly informative data, using only a small number of operator applies.
-outperforms (as measured by number of PDE solves) existing methods for Hessian approximation.
-In particular, the proposed method can approximate Hessians in inverse problem with highly informative data, using only a small number of operator applies.
 \end{abstract}
@@
 Here $\Aop$ is the ice sheet inverse problem data misfit Gauss-Newton Hessian described in Section~\ref{sec:numerical_results}.
-%	Brightness scale differs between the subfigures. 
 	Black stars are point source locations.
@@
 
-%Existing scalable high rank approximation methods typically require the ability to rapidly evaluate arbitrary entries of $\Aop$'s integral kernel, which we cannot do. 
 We use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applies.
@@
 Then we interpolate translated and scaled versions of these impulse responses to approximate entries of the operator's integral kernel.
-%This allows fast evaluation of approximate matrix entries, which are needed by high-rank approximation methods, in particular, classical H-matrix construction methods. 
 Picking the batches of points requires us to estimate the supports of the impulse responses $\phi_x$ \emph{before} they are computed.
@@
 While PC approximations are based on an assumption of local translation invariance, the method we propose is based on a more general assumption we call ``local mean displacement invariance'' (Section~\ref{sec:local_mean_displacement_invariance} and Figure~\ref{fig:mean_displacement_invariance}).
-%Furthermore, PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulses.
-We avoid this issue by temporarily excluding impulses from the approximation procedure whenever an attempt is made to evaluate the impulse at a location where it is undefined.
-
-%\begin{figure}
-%	\begin{center}
-%		\includegraphics[scale=0.4]{bct1.eps}
-%	\end{center}
-%	\caption{Hierarchical matrix block structure for the Stokes Hessian. The red blocks (tiny) are dense. All other blocks are low rank.
-%	}
-%	\label{fig:hmatrix_intro}
-%\end{figure}
 
@@
 The ability to rapidly approximate entries of $\Aop$'s integral kernel allows one to approximate discretized versions of $\Aop$ using the full arsenal of tools for matrix approximation that rely on fast access to matrix entries.
-%one to approximate discretized matrix representations of $\Aop$ using methods that require fast access to arbitrary matrix entries. We 
 In this work, we form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of a discretized version of $\Aop$.
 H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank.
-%(see Figure~\ref{fig:hmatrix_intro} and Appendix~\ref{app:h_matrix}).
 H-matrix methods permit us to perform matrix-vector products cheaply, and perform other useful linear algebra operations that cannot be done easily using the original operator.
 These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion.
-%These H-matrix methods are fast and scale well to large problems. 
 The work and memory required to perform these operations for an $\fedim \times \fedim$ H-matrix scales 
-%as $O(\fedim \log(\fedim)^a)$, where $a \in \{0,1,2,3\}$. That is, the cost is \emph{nearly-linear}
 \emph{nearly linearly} in $N$ (i.e., $o(N^{1+\epsilon})$ for any $\epsilon > 0$).
 The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks~\cite{GrasedyckHackbusch03}.
-%These operations can be used, for example, to solve Newton linear systems in PDE constrained optimization, sample from ill-conditioned posterior distributions in Bayesian inverse problems, and construct high rank Gaussian surrogate models that can be used for uncertainty quantification.
-
-%Classical H-matrix construction methods 
-%(described in Appendix~\ref{sec:H_matrix_construction}) 
-%require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.
 
@@
 In distributed parameter inverse problems governed by PDEs, one seeks to infer an unknown spatially varying parameter field from limited observations of a state variable that depends on the parameter implicitly through the solution of a PDE.
-%Inverse problems are typically ill-posed, in the sense that many different parameter fields may yield predicted observations that agree with the actual observations to within the noise level. 
 Conventionally, the inverse problem is formulated using either a deterministic framework~\cite{AsterBorchersThurber18,BorziSchulz11,Vogel02},
-%Guy Chavent, Nonlinear Least Squares for Inverse Problems, Springer, 2009.
 or a Bayesian probabilistic framework~\cite{KaipioSomersalo05,Stuart10a,Tarantola05}.
@@
 For instance, see~\cite{KimEtAl21,PetraEtAl14} and the references therein.
-
-%\begin{figure}
-%	\begin{subfigure}{0.5\textwidth}
-%		\includegraphics[scale=0.23]{informed_uninformed_modes_quadratic.pdf}
-%		\caption{}
-%	\end{subfigure}
-%	\begin{subfigure}{0.48\textwidth}
-%		\includegraphics[scale=0.1]{misfit_and_reg_hessian_eigenvalues.pdf}
-%		\caption{}
-%	\end{subfigure}
-%	\caption{(a) The Hessian eigenstructure characterizes the directional scalings of the best local quadratic approximation to the objective function in an inverse problem.
-The larger an eigenvalue of the Hessian is, the more closely spaced the contours of the local quadratic approximation are in the corresponding eigenvector direction.
-(b) As the data in an inverse problem becomes more informative about the parameter, the eigenvalues of the data misfit Hessian increase.
-As the regularization is weakened, the eigenvalues of the regularization Hessian decrease.
-These illustrations are taken from~\cite[Chapter 1]{Alger19}, with minor modifications.}
-%	\label{fig:hessian_eigenstructure}
-%\end{figure}
 
@@
 The eigenvalues of the data misfit Hessian characterize the sensitivity of the predicted observations to perturbations of the parameter in the corresponding eigenvector directions,
-%(see Figure~\ref{fig:hessian_eigenstructure})
 hence more informative data leads to larger eigenvalues and a larger numerical rank~\cite{AlexanderianGloorGhattas16}\cite[Section 1.4 and Chapter 4]{Alger19}.
@@
 The numerical rank of the regularization preconditioned data misfit Hessian may be reduced by increasing the strength of the regularization, but this throws away useful information: components of the parameter that could be learned from the observations
-%components of the observations that have a large signal to noise ratio 
 would instead be reconstructed based on the regularization~\cite[Section 4]{AlgerEtAl17}\cite[Chapters 1 and 7]{Vogel02}.
 Hence, low rank approximation methods suffer from a predicament: if the data highly inform the parameter and the regularization is chosen appropriately, then a large number of operator applies are required to form an accurate approximation of the Hessian using low rank approximation methods.
-%The best case scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost)~\cite{AlexanderianGloorGhattas16,Alger19,AlgerEtAl17}.
 High rank Hessian approximation methods are needed.
@@
 Although these methods are promising,
-%required number of matrix-vector products grows logarithmically in the discretized parameter dimension and linearly in the rank of the off-diagonal blocks, 
 the required number of operator applies is still large (e.g., hundreds to thousands).
-%These methods are asymptotically scalable in the sense that the required number of matrix-vector products grows nearly linearly in the dimension of the discretized parameter and linearly in the rank of the off-diagonal blocks.
-But despite asymptotic scalability, from a practical perspective the required number of matrix-vector products is large (e.g., hundreds to thousands).
-%Improving the peeling process to reduce its cost is currently an area of active research~\cite{PEELINGRESEARCH}.
-%This is an area of active research, so more efficient and practical matrix-free H-matrix construction methods based on the peeling process may be discovered in the future.
-%Matrix-free H-matrix construction methods are currently an area of active research (see, e.g.,~\cite{LevittMartinsson22}), so more efficient matrix-free H-matrix methods may be discovered in the future.
 In this paper, we also form an H-matrix Hessian approximation.
@@
 
-%Classical methods for approximating the Hessian are based on low rank approximation of either the data misfit Hessian, or the data misfit Hessian preconditioned by the inverse of the prior term in the Hessian.
-However,  More generally, the eigenvalues of the Hessian characterize how informative the data are about the component of the parameter in the corresponding parameter direction.
-The more informed the component of the parameter, the larger
-
-%In the deterministic setting, the Hessian of the objective function is central to efficient solution of the optimization problem, and in the
-
-% given the observed data is  one explores the posterior probability distribution for the parameter
-
-%Our motivation is high rank Hessian approximation in distributed parameter inverse problems governed by PDEs. 
-%The Hessian is an operator that is central to efficient methods for solving such inverse problems and is only available matrix-free; we will discuss Hessians in inverse problems governed by PDEs in greater detail in Section~\ref{sec:hessian}.
-
-%Introduce inverse problem
-%
-%Role of Hessian
-%
-%Structure of Hessian that we want to exploit
-%
-%Low rank
-%
-%Need for High rank
-%
-%Low rank approximation of the Hessian:
-%\begin{itemize} 
-%	\item Hessian eigenstructure is determined by data informativeness. Bigger eigenvalue => more informed component in eigenvector direction. 
-%	\item informative data => high rank hessian
-%	\item get high rank approximation by exploiting different structure. Locality, non-negativity of kernel
-%	\item Large rank in absolute terms, even if low rank in relative terms
-%\end{itemize}
-
 \section{Preliminaries}
@@
 The operator $\Aop$ may be applied to distributions if $\Aker$ is sufficiently regular. 
-%We derive the extension as follows. 
 Given $\rho \in L^2(\Omega)'$, let $\rho^* \in L^2(\Omega)$ denote the Riesz representative of $\rho$ with respect to the $L^2(\Omega)$ inner product. We have
@@
 Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions and let $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ be a distribution. In this case, 
-%the Riesz representative 
 $\rho^*$ may not exist, so the derivation in~\eqref{eq:extension_to_distributions} is not valid.
@@
 Using~\eqref{eq:action_on_distribution} and the definition of the delta distribution, we see that $\phi_x$ may also be written as the function $\impulseresponse_x(y) = \Aker(y, x)$.
-%\begin{equation}
-%\label{eq:impulse_response_defn}
-%\impulseresponse_x(y) = \Aker(y, x),
-%\end{equation} 
-%which results from keeping $x$ fixed and viewing $\Aker(y,x)$ as a function of $y$.
 
@@
 If these conditions are violated, the method will incur additional error, and depending on the severity of the violation, the proposed method may fail.
-%Not all operators of interest (and in particular not all Hessians) satisfy these conditions, but many operators of practical interest do satisfy these conditions, and our method is highly effective for approximating these operators.
 
@@
 $
-%\label{eq:fem_coeff_basis}
 u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
@@
 Linear functionals $\rho_h \in V_h'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^\nsamplepts_{\massmatrix^{-1}}$, with entries $\boldsymbol{\rho}_i = \rho_h(\febasis_i)$ for $i=1,\dots,\nsamplepts$.
-%\begin{equation*}
-%\boldsymbol{\rho}_i = \rho_h(\febasis_i), \qquad i=1,\dots,\nsamplepts. 
-%\end{equation*}
 Here $\massmatrix \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries $\massmatrix_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) \mathrm{d}x$ for $i,j=1,\dots,\fedim$.
-%\begin{equation*}
-%\massmatrix_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx, \qquad i,j=1,\dots,\fedim,
-%\end{equation*}
 The space $\mathbb{R}^\fedim_\massmatrix$ is $\mathbb{R}^\fedim$ with the inner product $(\mathbf{u},\mathbf{w})_\massmatrix := \mathbf{u}^T \massmatrix \mathbf{w}$, and $\mathbb{R}^\fedim_{\massmatrix^{-1}}$ is the analogous space with $\massmatrix^{-1}$ replacing $\massmatrix$.
@@
 The operator $A_h:V_h \rightarrow V_h'$ can be written in integral kernel form,~\eqref{eq:kernel_representation}, but with $\Aker$ replaced by a slightly different integral kernel,
-%as
-%\begin{equation}
-%(A_h u_h)(w_h) := \int_\Omega \int_\Omega w_h(y) \Aker_h(y,x) u_h(x) dx dy
-%\end{equation}
 $\Aker_h$, which we do not know, and which differs from $\Aker$ due to discretization error.
@@
 $
-%\label{eq:discrete_kernel}
 \boldsymbol{\impulseresponse}_x = \left(A_h \delta_x^*\right)^* =  \massmatrix^{-1}\mathbf{A} \massmatrix^{-1} \boldsymbol{\delta}_x,
@@
 $
-%\label{eq:approx_kernel_entry_81}
 	\Aker_h(y,x) = \boldsymbol{\delta}_y^T \boldsymbol{\impulseresponse}_x = \boldsymbol{\delta}_y^T \massmatrix^{-1}\mathbf{A} \massmatrix^{-1} \boldsymbol{\delta}_x.
@@
 However, entries of $\boldsymbol{\Aker}$ involve inverse mass matrices, which typically contain negative numbers.
-%If there are too many negative numbers, or if the negative numbers are too large, the proposed method will yield less accurate approximations and may potentially fail. 
 We therefore recommend replacing the mass matrix, $\massmatrix$, with a positive diagonal \emph{lumped mass} approximation. 
@@
 First, we define moments of the impulse responses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section~\ref{sec:intromoments}).
-%Our method will use these ellipsoid estimates to choose sets of impulses that can be computed in batches. 
 Second, we describe an improved method to approximate impulse responses from other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section~\ref{sec:local_mean_displacement_invariance}).
-%Our method will interpolate impulses using this approximation.
 
@@
 for $i,j=1,\dots, \gdim$. Here $u/w$ denotes pointwise division, $\left(u/w\right)(x) = u(x)/w(x)$, and $u\cdot w$ denotes pointwise multiplication, $(u \cdot w)(x) = u(x)w(x)$.
-%The resulting method for computing discretized versions of $\spatialvol$, $\spatialmean$, and $\spatialcov$ is shown in Algorithm~\ref{alg:varhpi_mean_cov}.
-
-%\begin{algorithm2e}
-%	\SetAlgoNoLine
-%	\SetKwInOut{Input}{Input}
-%	\SetKwInOut{Output}{Output}
-%	{	
-%		\Input{Operator $\bm{A}^T$}
-%		\Output{$\bm{\spatialvol}\in \mathbb{R}^\fedim$, $\bm{\spatialmean}\in \mathbb{R}^{\gdim \times \fedim}$, and $\bm{\spatialcov} \in \mathbb{R}^{\gdim \times \gdim \times \fedim}$}
-%		
-%		Form vector $\bm{C} \in \mathbb{R}^\fedim_{\massmatrix}$ by either projecting or interpolating constant function $C(x)=1$ onto $V_h$
-%		
-%		$\bm{\spatialvol} \gets \massmatrix^{-1}\bm{A}^T \bm{C}$
-%		
-%		\For{$i=1,2,\dots,\gdim$}{
-%			Form vector $\bm{L}^i \in \mathbb{R}^\fedim_{\massmatrix}$ by either projecting or interpolating linear function $L^i(x) = x^i$ onto $V_h$
-%			
-%			$\bm{\spatialmean}^i \gets \left(\massmatrix^{-1}\bm{A}^T \bm{L}^i\right) / \bm{\spatialvol}$
-%		}
-%		\For{$i=1,2,\dots,\gdim$}{
-%			\For{$j=1,\dots,i$}{
-%				Form quadratic function $\bm{Q}^{ij} = x^i x^j$
-%				
-%				$\bm{\spatialcov}^{ij} \gets \left(\massmatrix^{-1}\bm{A}^T \bm{Q}^{ij}\right) / \bm{\spatialvol} - \bm{\spatialmean}^i\cdot \bm{\spatialmean}^j$
-%				
-%				$\bm{\spatialcov}^{ji} \gets \bm{\spatialcov}^{ij}$
-%				
-%			}
-%		}
-%		
-%	}
-%	\caption{Compute discretized scaling factor $\mathbf{\spatialvol}$, mean $\boldsymbol{\spatialmean}$, and covariance $\boldsymbol{\spatialcov}$}
-%	\label{alg:varhpi_mean_cov}
-%\end{algorithm2e}
 
@@
 The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is typically conservative.
-% and typically far less mass resides in this region.
 
@@
 
-%\paragraph{Future extension: approximation using $\Sigma$} One may map $E_x$ to the unit sphere via the change of variables $z = \Sigma^{-1/2}\left(x\right)\left(y - \spatialmean(x)\right) / \tau$, and approximate
-%\begin{equation}
-%\label{eq:sigma_interpolation}
-%\phi_x(z) \det{\Sigma^{1/2}\left(x\right)}/\spatialvol(x) \approx \phi_{x'}(z') \det{\Sigma^{1/2}\left(x'\right)}/\spatialvol(x').
-%\end{equation}
-%%Notice that equality holds in \eqref{eq:sigma_interpolation} if the impulse responses are Gaussian functions. 
-%While interpolation schemes based on \eqref{eq:sigma_interpolation} have the potential to be more accurate than those based on \eqref{eq:translate4}, we were unable to construct an interpolation scheme based on \eqref{eq:sigma_interpolation} that is robust to negative entries in the integral kernel.
-The integral kernel should have non-negative entries per Condition 3 in Section \ref{sec:conditions_2}, but in practice some entries are negative due to discretization error.
-%%Therefore, in this paper we do not use \eqref{eq:sigma_interpolation}. 
-%%Constructing interpolation schemes based on \eqref{eq:sigma_interpolation} that are robust to discretization error is a promising direction for future research.
-
-%but we find they are not robust, and therefore we do not use them:
-%\begin{align*}
-%	\phi_{x}(y) &\approx \phi_{x'}\left(\spatialmean(x') + \Sigma_{x'}^{1/2} \Sigma_{x}^{-1/2} \left(y - \spatialmean(x)\right)\right) \spatialvol(x) / \spatialvol(x') \\
-%	\phi_{x}(y) &\approx \phi_{x'}\left(\spatialmean(x') + \Sigma_{x'}^{1/2} \Sigma_{x}^{-1/2} \left(y - \spatialmean(x)\right)\right) \left( \spatialvol(x) \det{\Sigma^{1/2}(x')} \right) / \left( \spatialvol(x') \det{\Sigma^{1/2}(x)} \right)
-%\end{align*}
-%In these schemes, one maps known impulse responses from their support ellipsoids to the unit sphere, performs interpolation on the unit sphere, then maps the resulting interpolated impulse response to the support ellipsoid associated with the target point.
-We find that computation of $\Sigma$ is more susceptible to numerical error than computation of $V$ and $\mu$, in the sense that spurious negative numbers in $\phi_x$ cause $\Sigma(x)$ to become highly ill-conditioned or even indefinite.
-When $\Sigma$ is used for picking sample points (Section \ref{sec:sample_point_selection}), we simply avoid picking sample points $p$ with ``bad'' $\Sigma(p)$.
-These schemes require using $\Sigma_x$ at arbitrary $x$, rather than just the sample points.
-When picking sample points, one may
-
 \section{Operator approximation algorithm}
@@
 We then convert $\boldsymbol{\Aker}_H$ into an H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$.
-%Schematically the algorithm proceeds as follows:
-%\begin{equation*}
-%\Aop \longrightarrow \spatialvol, \spatialmean, \spatialcov \longrightarrow \underbrace{\{x_i\}_{i=1}^\nsamplepts}_{\text{as }S_b\text{'s}} \longrightarrow \underbrace{\{\impulseresponse_{x_i}\}_{i=1}^\nsamplepts}_{\text{as }\eta_b\text{'s}} \longrightarrow \boldsymbol{\Aker}_H \longrightarrow \mathbf{A}_H.
-%\end{equation*}
 
@@
 The computational cost is discussed in Section~\ref{sec:overall_cost}.
-
-%.
-The matrix is symmetrized by computing $\left(\mathbf{A}_H+\mathbf{A}_H^T\right)/2$ using H-matrix addition.
-Negative eigenvalues modified to be positive using a rational method that is described in Appendix~\ref{sec:make_spd}.
 
@@
 		Compute $V, \mu$, and $\Sigma$ (Equations~\eqref{eq:vol_mean_var} in Section~\ref{sec:intromoments})
-
-%		Compute $\bm{\spatialvol}$, $\bm{\spatialmean}$, and $\bm{\spatialcov}$ using Algorithm~\ref{alg:varhpi_mean_cov}.
 
@@
 then we add $\candidatepoint_i$ to $\pointbatch_k$. Otherwise we discard $\candidatepoint_i$. This process repeats until there are no more points in $\candidatepts$.  
-%This is detailed in Algorithm~\ref{alg:point_choice}. 
 We repeat the point selection process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots, \pointbatch_{\nbatch}$.
@@
 Second, if the bounding boxes intersect, we check if the ellipsoids intersect using the ellipsoid intersection test in \cite{GilitschenskiHanebeck12}.
-%described in Appendix~\ref{sec:fast_ellipsoid_intersection_test}.
-
-%\begin{algorithm2e}
-%	\SetAlgoNoLine
-%	\SetKwInOut{Input}{Input}
-%	\SetKwInOut{Output}{Output}
-%	\SetKwProg{Fn}{Function}{}{}
-%	
-%	\Input{Finite set of candidate points $\candidatepts\subset \Omega$, \\spatially varying mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\previous sample point batches $\pointbatch_1, \dots, \pointbatch_{k-1}$}
-%	\Output{Batch of new sample points $\pointbatch_k$}
-%	
-%	Initialize empty new batch of sample points, $\pointbatch_k = \{\}$
-%	
-%	\While{$\candidatepts$ is not empty}{
-%		Determine the point $\candidatepoint_i \in \candidatepts$ that is farthest from all points in previous sample point batches $\pointbatch_1,\dots,\pointbatch_{k-1}$
-%		
-%		Remove $\candidatepoint_i$ from $\candidatepts$
-%
-%	
-%		\If{$E_{\candidatepoint_i} \cap E_q \neq \{\}$ for all $\candidatepoint_j \in \pointbatch_k$}{
-%			\tcp{$E_{\candidatepoint_i}$ and $E_{\candidatepoint_j}$ are the ellipsoids defined in~\eqref{eq:support_ellipsoid}}
-%			
-%			Add $\candidatepoint_i$ to $\pointbatch_k$
-%			
-%			Remove all points $\candidatepoint$ satisfying $\spatialmean(\candidatepoint) \in E_{\candidatepoint_i}$ from $\candidatepts$
-%		}
-%	}
-%
-%    \SetKwFunction{FMain}{point\_is\_acceptable}
-%	\caption{Choosing one batch of sample points, $\pointbatch_k$}
-%	\label{alg:point_choice}
-%\end{algorithm2e}
 
@@
 at the point $x$. 
-%The idea is that $\Aker(y,x) = f_i$ per the discussion in Section~\ref{sec:local_mean_displacement_invariance}, so we approximate $\Aker(y,x)$ by interpolating the values $f_i$. 
 Interpolation is performed using the following radial basis function~\cite{Wendland04} scheme:
@@
 Here $L:=\diam\left(\{x_i\}_{i=1}^{\numnbr}\right)$ is the diameter of the set of sample points used in the interpolation, and $C_\text{RBF}$ is a user-defined shape parameter that controls the width of the kernel function (we use $C_\text{RBF}=3.0$).
-%is a polyharmonic spline radial basis function, which is defined as $\rbf(r):=r^\gdim$ if $\gdim$ is odd, and $\rbf(r):=r^\gdim \log r$ for $r>0$ and $\rbf(0):=0$ if $\gdim$ is even.
-%\begin{equation}
-%\rbf(r) := \begin{cases}
-%r^\gdim, & \gdim=1,3,5,\dots \\
-%r^\gdim \log r, & \gdim=2,4,6,\dots
-%\end{cases}
-%\end{equation}
-%for $r>0$, and $\rbf(0):=0$. 
 The vector of weights, $\rbfweight = (\rbfweight_1, \rbfweight_2, \dots, \rbfweight_{\numnbr})^T$, is found as the solution to the $\numnbr \times \numnbr$ linear system
@@
 and $f \in \mathbb{R}^{\numnbr}$ has entries $f_i$ from~\eqref{eq:fxyxp}.
-%Polyharmonic spline radial basis function interpolation yields the smoothest interpolant of the given points and values, in a least squares sense~\cite{Duchon77}. 
-%For more details on radial basis function interpolation, see~\cite{Wendland04}. 
-%The interpolation procedure is illustrate in Figure~\ref{fig:rbf_interpolation}.
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[scale=0.2]{interpolation_rbf_boundary.pdf}
-%	\caption{To compute the kernel entry $\Aker(y,x)$, we use radial basis functions to interpolate the values $\frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}(y-\spatialmean(x)-\spatialmean(x_i))$ for the $\numnbr$ nearest sample points $x_i$ to $x$, excluding sample points for which $y-\spatialmean(x)-\spatialmean(x_i)$ is outside of the domain.}
-%	\label{fig:rbf_interpolation}
-%\end{figure}
 
@@
 If a non-diagonal mass matrix is used, one may form a H-matrix representation of the mass matrix, then perform the matrix-matrix multiplications in \eqref{eq:boldA} using H-matrix methods.
-%Once $\mathbf{A}_H$ is formed, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods~\cite{BORMHACKBUSCHBOOK}.
 We use H1 matrices in the numerical results, but any other H-matrix format could be used instead. For more details on H-matrices, see~\cite{Hackbusch15}.
@@
 We emphasize that the dense matrix $\mathbf{\Aker}$ is never formed.
-%We describe this H-matrix construction process in detail in Appendix~\ref{app:h_matrix}. 
-%We form $\massmatrix_H$ using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same geometrical clustering block cluster tree as was used for $\mathbf{\Aker}$.
 
@@
 
-%In many problems $\Aop$ is symmetric positive semi-definite. 
 The PSF approximation $\mathbf{A}_H$, and therefore $\mathbf{A}_H + \mathbf{R}_H$, may be non-symmetric and indefinite because of approximation error. 
@@
 In target applications, (1) is the dominant cost because applying $\Aop$ to a vector requires an expensive computational procedure such as solving a PDE, and (1) is the only step that requires applying $\Aop$ to vectors.
-%In practice, (1) is the dominant cost, because (1) is the only task that requires applying $\Aop$ to vectors, and  
 All operations that do not require applications of $\Aop$ to vectors are nearly linear, and therefore scalable, in the size of the problem, $N$.
@@
 		$\gdim$ & $1$--$3$ & Spatial dimension \\
-%		$\classicalrank$ & $10^1$--$10^4$ & Classical matrix rank \\
 		$\nsamplepts$ & $10^1$--$10^4$ & Total number of sample points (all batches) \\
@@
 The multiplicative dependence of $\fedim$ with $\nsamplepts$ is undesirable since $\nsamplepts$ may be large, and reducing this cost is possible with more involved computational geometry methods.
-%Several improvements are possible.
-The cost of picking sample points could be reduced by ``thinning'' the set of candidate points that the sample points are picked from.
-That is, choosing sample points from a smaller subset of $\fedim_\text{thin} << \fedim$ points.
-The cost to perform the ellipsoid intersection tests could be reduced by first checking whether the ellipsoid bounding boxes intersect before performing the ellipsoid intersection test of Appendix~\ref{sec:fast_ellipsoid_intersection_test}.
-The cost of picking the $i$th batch could be reduced to $O(\fedim \log |\pointbatch_i|)$ using more advanced computational geometry techniques, such as organizing the ellipsoids associated with points that have already been picked into a dynamic AABB tree, then using this AABB tree to accelerate the required ellipsoid intersection tests.
 However, from a practical perspective, the cost for choosing sample points is small compared to other parts of the approximation algorithm, hence such improvements are not pursued here.
@@
 Classical H-matrix construction techniques
-%, which is described in Appendix~\ref{app:h_matrix}, 
 require evaluating $O(\hrank \fedim \log \fedim)$ matrix entries of the approximation \cite{BebendorfRjasanow03}, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix.
@@
 \end{equation*}
-%
-%
-%This follows from the above discussion, and the fact that $\fedim \ge \nsamplepts$.
 
 \paragraph{(3) Performing linear algebra operations with the H-matrix} It is well known that H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, matrix inversion, and low rank updates require performing $O\left(\hrank^a \fedim \log(\fedim)^b\right)$
-%\begin{equation}
-%\label{eq:H_matrix_cost_operations}
-%ASDASDAS	O(\hrank^2 \fedim \log(\fedim)^a) \qquad \text{elementary operations},
-%\end{equation}
 elementary operations,
-%\footnote{We advise the reader to be aware that asymptotic complexity estimates for H-matrix operations can be a poor indicator of practical performance because of large prefactor constants, the complicated way in which recursive H-matrix algorithms map onto hardware, and difficulties taking advantage of parallelism.},
 where $a,b \in \{0,1,2,3\}$ are constants which depend on the type of H-matrix used and the operation being performed~\cite{GrasedyckHackbusch03}\cite[Section 2.1]{Kriemann13}.
-%In our experience with several software packages for H1 matrices, we have typically observed that matrix-vector products are extremely fast, matrix-matrix addition is fast, matrix LU factorization requires a comparatively moderate amount of time, factorized solves are extremely fast, and matrix-matrix multiplication and matrix inversion are comparatively slow.
 In the numerical results (Section~\ref{sec:numerical_results}), we use one matrix-matrix addition to add the H-matrix approximation of the data misfit term in the Hessian to the regularization term in the Hessian.
-%Constructing an H-matrix representation of the regularization term in the Hessian requires one matrix inversion and two matrix-matrix multiplications. 
-%Then, we use one matrix factorization to form a preconditioner for the Newton linear system. 
-%We perform low-rank updates to the H-matrix as the Newton iterations proceed, as per the discussion in Appendix~\ref{sec:recycle_krylov}.
 Symmetrizing $\mathbf{A}_H$ requires one matrix-matrix addition.
 Flipping negative eigenvalues to be positive requires a handful (typically around 5) of matrix-matrix additions and matrix factorizations to factor the required shifted linear systems, and a number of factorized solves that is proportional to the number of erroneous negative eigenvalues.
-%semi-definite, described in Appendix~\ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the rational function order.
 
@@
 principle~\cite{Vogel02}.
-%The optimal value of $\gamma$ is used for
-%  $\gamma =  7.3 \times 10^3$ for Stokes  Figure~\ref{fig:gamma_sweep}
-%all the numerical results except for the studies where we study how
-%changing $\gamma$ impacts the effectiveness of the proposed
-%preconditioner.
 
@@
 inverse problem cost function, respectively, evaluated at the current Newton iterate.
-%The Gauss-Newton Hessian is Hessian of $J$, except in the data misfit, $J_d$, the nonlinear mapping $\basalfriction \mapsto \tangentop \velocity$ is replaced by its local linearization at $\basalfriction_k$.
 To ensure positive definiteness of the Hessian we use
@@
 The
-%number of impulse response batches used to build these 
 approximations are indicated by ``PSF ($\nbatch$)'', where $\nbatch$ is the number of impulse response batches used to build the approximation.
@@
 The numerical results show that~$\widetilde{\bm{H}}$ is a good preconditioner for both~$\bm{H}_\text{gn}$ and~$\bm{H}$.
-%The H-matrix approximation of the data misfit Gauss-Newton Hessian is added to an H-matrix approximation of the Hessian of the regularization term to form an H-matrix approximation of $\bm{H}_\text{gn}$.
-%
-%Straightforward calculation shows that the discretized Hessian of the regularization term is given by
-%$
-%(\gamma \bm{K} + \delta \bm{M}) \bm{M}^{-1} (\gamma \bm{K} + \delta \bm{M})
-%$
-%where $\bm{K}$ is a finite element stiffness matrix with appropriate boundary conditions, and $\bm{M}$ is the finite element mass matrix.
-These matrices are sparse, so we form H-matrix representations of these matrices and combine them using standard sparse H-matrix techniques.
-%
 
@@
 The setup, which we briefly summarize, follows~\cite{IsaacEtAl15,PetraEtAl12}.
-%The inverse problem considered here is synthetic, but in practice surface velocity observations are available from satellites.
-%
-%
-%\begin{figure}
-%	\begin{center} 
-%	\includegraphics[scale=0.23]{meshHeight_view2_edges.png}
-%	\end{center}
-%	\caption{Bird's eye view of the ice sheet, discretized by a mesh of tetrahedra.
-Color indicates the height of the base of the ice sheet (i.e., the mountain topography).
-The radius of the domain is $10^4$ meters, the maximum height of the mountain is $2.1 \times 10^3$ meters, and the average thickness of the ice sheet is 250 meters.}
-%	\label{fig:ice_mountain_mesh}
-%\end{figure} 
-%
-%
 The region of ice is denoted by
@@
 
-%aaa
-
-%% from here we summarize --------
-%The true parameter field, $\basalfriction_\text{true}$, and corresponding velocity field, $\velocity_\text{true}$, are shown in Figure~\ref{fig:true_beta_u}.
-We generate synthetic observations by restricting the tangential component of $\velocity_\text{true}$ to the top surface to get
-%\begin{equation*}
-%\observations_\text{true} = \tangentop \velocity_\text{true} |_{\Gamma_t}.
-%\end{equation*}
-%We add $1\%$ multiplicative Gaussian noise to all components of the finite element vector, $\bm{\observations}_\text{true}$, corresponding to the function $\observations_\text{true}$, which yields noisy synthetic observations, $\observations_\text{obs}$ which are used for the inversion.
-That is,
-%\begin{equation*}
-%\left(\bm{\observations}_\text{obs}\right)_j := \left(\bm{\observations}_\text{true}\right)_j + 0.01 \cdot \boldsymbol{\xi}_j \cdot \left|\left(\bm{\observations}_\text{true}\right)_j\right|,
-%\end{equation*}
-%where $j$ ranges over all entries of the vector of coefficients representing the finite element functions $\observations_\text{true}$ and $\observations_\text{obs}$, and $\boldsymbol{\xi}_j$ are independent and identically distributed (i.i.d) random numbers drawn from the standard normal distribution.
-We also define the noise function, $\xi$, to be the finite element function with $j$th vector entry $\boldsymbol{\xi}_j$.
-%
-%The reconstructed parameter field $\basalfriction$ is found as the solution to an optimization problem of the following form:
-%\begin{equation}
-%	\label{eq:deterministic_optimization_problem}
-%	\min_{\basalfriction} \quad J(\basalfriction) := J_d(\basalfriction) + J_r(\basalfriction).
-%\end{equation}
-%The first term in the objective function is
-%\begin{equation*}
-%	J_d(\basalfriction) := \frac{1}{2}\int_{\Gamma_{t}}\|\observations_\text{obs} - \tangentop\velocity\|_{2}^{2}\,dS.
-%\end{equation*}
-%We call this term the data misfit, because it measures the difference between the observed data and the predicted data based on a candidate parameter $\basalfriction$.
-Here $\velocity = \velocity(\basalfriction)$ denotes the velocity field solving $\ref{Stokeseqn}$ for the given parameter $\basalfriction$.
-The second term,
-%\begin{equation*}
-%	J_r(\basalfriction) := \frac{1}{2}\int_{\Omega}\left| \mathcal{K}(\basalfriction - \basalfriction_0)\right|^2\,dx,
-%\end{equation*}
-%is a Bilaplacian regularization term.
-Specifically, $\basalfriction_0$ is the constant function $\basalfriction_0(x)=10.5$, and $\mathcal{K}$ is the inverse of the solution operator for the following elliptic PDE:
-%\begin{align*}
-%-\gamma\Delta u +\delta\,u&=\regrhs\,\,\,\,\text{ in }\Omega,\\
-%\gamma \nabla u\cdot\normalvec+\regrobincoeff\,u&=0\,\,\,\text{ on }\partial\Omega.
-%\end{align*}
-%Recall that $\Omega$ is the 2D flattened version of the basal surface $\Gamma_b$.
-Here $\regrhs$ is a generic forcing term, and $\regrobincoeff=\sqrt{\delta \gamma}/1.42$ is a Robin boundary condition coefficient~\cite{Roininen14}, and these quantities are different from $\bodyforce$ and $\stokesrobincoeff$ in the Stokes equation,~\eqref{Stokeseqn}.
-%In all results except one, $\gamma$ is chosen so that the Morozov discrepancy principle is satisfied, i.e.,
-%\begin{equation*}
-%	\int_{\Gamma_{t}}\|\observations_\text{obs} - \tangentop\velocity\|_{2}^{2}\,dS = \int_{\Gamma_{t}}\|\xi\|_{2}^{2}\,dS.
-%\end{equation*}
-%The value of $\gamma$ satisfying the Morozov discrepancy principle is $\gamma=7.3 \times 10^3$.
-The exception is Figure~\ref{fig:gamma_sweep}, where we vary $\gamma$ to study how changing $\gamma$ impacts the effectiveness of our preconditioner.
-The constant $\delta$ is chosen so that the correlation length of functions drawn from the normal distribution with covariance $\mathcal{K}^{-2}$, given by $L=\sqrt{\gamma/\delta}$, is $1/10$th the radius of the domain.
-%%, which is halfway between the maximum and minimum values of $\basalfriction_\text{true}$.
-%
-%PCG only requires applying
-%$\bm{H}$ (or $\bm{H}_\text{gn}$) to vectors. Each application of
-%$\bm{H}$ (or $\bm{H}_\text{gn}$) to a vector requires solving one
-%incremental forward and one incremental adjoint Stokes PDE of the form
-%\eqref{Stokeseqn}, but with different right hand sides.
 \begin{figure}
@@
 		\includegraphics[scale=0.20]{mstar_0.25noise_cropped.png}
-		%		\caption{}
-		%		\label{fig:newtoniter1}
-	\end{subfigure}
-%	\begin{subfigure}{0.18\textwidth}
-%		\includegraphics[scale=0.075]{mstar_0.12noise_cropped.png}
-%		%		\caption{}
-%		%		\label{fig:newtoniter2}
-%	\end{subfigure}
+	\end{subfigure}
 	\begin{subfigure}{0.32\textwidth}
 		\includegraphics[scale=0.20]{mstar_0.05noise_cropped.png}
-		%		\caption{}
-		%		\label{fig:newtoniter3}
-	\end{subfigure}
-%	\begin{subfigure}{0.18\textwidth}
-%		\includegraphics[scale=0.075]{mstar_0.02noise_cropped.png}
-%		%		\caption{}
-%		%		\label{fig:newtoniter4}
-%	\end{subfigure}
+	\end{subfigure}
 	\begin{subfigure}{0.32\textwidth}
 	\includegraphics[scale=0.20]{mstar_0.01noise_cropped.png}
-	%		\caption{}
-	%		\label{fig:newtoniter4}
     \end{subfigure}
@@
 		\includegraphics[scale=1.0]{stokes_pcg_convergence.pdf}
-%		\caption{}
-%		\label{fig:cg_convergence}
 	\end{subfigure}
@@
 	\end{center}
-%	\caption{}
-%	\label{fig:generalized_eigenvalues}
     \end{subfigure}
@@
 random entries drawn from the standard Gaussian distribution and $\bm{H}$ is evaluated at the solution of the inverse problem.
-%	The reference solution $\bm{x}$ is determined by solving $\bm{H} \bm{x} = \bm{b}$. 
-%	Here we use $\gamma=7.3 \times 10^3$, which is determined by the Morozov discrepancy principle when using $5\%$ noise.
 	Results in these figures are shown for the PSF-based preconditioners with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
@@
 PCG converges fastest with the PSF-based preconditioners, with PSF (25) converging fastest, followed by PSF (5), followed by PSF (1), as expected.
-%PCG converges much slower with no preconditioning and regularization preconditioning than it does with PSF-based preconditioning, with no preconditioning outperforming regularization preconditioning.
 In Figure~\ref{fig:krylov_convergence} (right) we show the generalized eigenvalues for the generalized eigenvalue problem
@@
 With the PSF-based preconditioners, the generalized eigenvalues cluster near one, with more batches yielding better clustering.
-%The regularization preconditioner clusters the trailing eigenvalues but amplifies leading eigenvalues.
-
-%In Figure~\ref{fig:condition_number} we show the condition numbers of $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$.
-The PSF preconditioners yield the smallest condition numbers, with more batches yielding smaller condition numbers.
-Regularization preconditioning and no preconditioning yield larger condition numbers.
-The condition numbers are similar for both $\bm{\preconditioner}^{-1} \bm{H}$ and $\bm{\preconditioner}^{-1} \bm{H}_\text{gn}$, demonstrating that the preconditioner built based on $\bm{H}_\text{gn}$ is a good preconditioner for $\bm{H}$.
 
@@
 All PSF-based preconditioners outperform no preconditioning and regularization preconditioning for all noise levels.
-%No preconditioning outperforms regularization preconditioning for all noise levels except $25\%$. 
-%We report condition number results for the preconditioned Hessian, $\mathbf{H}$, demonstrating that the PSF preconditioners built based on the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, are also good preconditioner for $\bm{H}$.
-%As expected, regularization preconditioning performs well for large $\gamma$ and poorly for small $\gamma$.
-Our PSF preconditioners outperform no preconditioning and regularization preconditioning for medium and small $\gamma$, and perform similarly to regularization preconditioning for large~$\gamma$.
 
@@
 
-%        |   0.5   |   1.0   |   2.0   
-% -------------------------------------
-% 1.0e-4 |   OK    |   OK    |   OK
-% 1.8e-4 |   OK    |   OK    |   OK
-% 3.2e-4 |   OK    |   OK    |   OK
-% 5.6e-4 |   OK    |   OK    |   5
-% 1.0e-3 |   OK    |   OK    |   5
-%
-
 Here we consider a time-dependent advection-diffusion equation in
@@
 We use $\kappa=3.2$e${-1}$ and $T=1.0$ for all results, except for Table \ref{tab:adv_convergence} and Figure~\ref{tbl:adv_vary_kappa_tf} where we vary $\kappa$ and $T$.
-% to study the impact of changing these parameters.
 
@@
 The results show that PSF-based preconditioning outperforms regularization preconditioning and no preconditioning, with more batches yielding better results.
-%The PSF-based preconditioners perform best in the high-rank regime where $T$ and $\kappa$ are small, and the observations are highly informative about the initial condition.
-This makes sense given the effect of $T$ and $\kappa$ on the shape of the
 
@@
 With the regularization preconditioner the trailing eigenvalues cluster near one, while the leading eigenvalues are amplified.
-
-%INTRODUCE EQUATIONS: REWRITE HERE vvvvvvv
-
-%The parameter-to-observable map $\iFF \,\ipar := \obsop\, u(\ipar)$ maps an initial condition $\ipar \in L^2(\advdomain)$ to pointwise spatiotemporal observations of the concentration field $u({\vec x},t)$ through solution of the advection-diffusion equation given by
-%
-%$$
-%\begin{split}
-%u_t - \kappa\Delta u + {\vec v} \cdot \nabla u &= 0     & \quad \text{in } \advdomain\times(0,T),\\
-%u(\cdot, 0) &= \ipar & \quad \text{in } \advdomain,\\
-%\kappa \nabla u\cdot {\vec{n}} &= 0     & \quad \text{on } \partial \advdomain \times (0,T).
-%\end{split}
-%$$
-%
-%Here, $\advdomain \subset \R^d$ ($d \in \{2, 3\}$) is a bounded domain, $\kappa > 0$ is the diffusion coefficient and $T > 0$ is the final
-%time. The velocity field
-%$\vec{v}$ is computed by solving the following steady-state
-%Navier-Stokes equation with the side walls driving the flow:
-%
-%$$
-%\begin{aligned}
-%- \frac{1}{\operatorname{Re}} \Delta {\vec v} + \nabla q + {\vec v} \cdot \nabla {\vec v} &= 0 &\quad&\text{ in }\advdomain,\\
-%\nabla \cdot {\vec v} &= 0 &&\text{ in }\advdomain,\\
-%{\vec v} &= {\vec g} &&\text{ on } \partial \advdomain.
-%\end{aligned}
-%$$
-%
-%Here, $q$ is pressure, $\text{Re}$ is the Reynolds number. The Dirichlet boundary data
-%${\vec g} \in \R^d$ is given by 
-%${\vec g} = {\vec e}_2$ on the left wall of the domain, 
-%${\vec g}=-{\vec e}_2$ on the right wall,  and ${\vec g} = {\vec 0}$ everywhere else.
-
-% INTRODUCE EQUATIONS: REWRITE HERE ^^^^^
 
@@
 	\end{center}
-	%\begin{subfigure}{0.05\textwidth}
-	%	\includegraphics[scale=0.1]{mColorBar.png}
-	%\end{subfigure}
 	\caption{(Diffusive transport) (\ref{fig:adv_initial_condition}) True initial condition.
@@
 		\setlength{\tabcolsep}{8pt}
-		%		\renewcommand{\arraystretch}{1.25}
 		\renewcommand{\arraystretch}{1.0}
@@
 		\includegraphics[scale=1.0]{adv_pcg_convergence.pdf}
-%		\begin{tikzpicture}[baseline, trim axis right]
-%		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
-%		ymode=log,
-%		ymin=1e-11,
-%		ymax=2,
-%		xmax=1000,
-%		axis x line*=bottom,
-%		axis y line*=left,
-%		xtick={0, 200, 400, 600,800,1000},
-%		xlabel={CG iteration},
-%		ylabel={$\|\bm{x}_k - \bm{x}\|/\|\bm{x}\|$},
-%		title={CG convergence},
-%		legend style={at={(1.2,1)},anchor=north east},
-%		%legend pos=north east,
-%		legend style={font=\small}
-%		]
-%		\addlegendentry{Reg}
-%		\addplot[dotted, line width=0.5mm] table {advDiff_cg_rand_errs_reg.dat};
-%		\addlegendentry{None}
-%		\addplot[dash dot, line width=0.5mm] table {advDiff_cg_rand_errs_none.dat};
-%		\addlegendentry{PSF (1)}
-%		\addplot[blue, line width=0.5mm] table {advDiff_cg_rand_errs_psf1.dat};
-%		\addlegendentry{PSF (5)}
-%		\addplot[red, line width=0.5mm] table {advDiff_cg_rand_errs_psf5.dat};
-%		\addlegendentry{PSF (25)}
-%		\addplot[purple, line width=0.5mm] table {advDiff_cg_rand_errs_psf25.dat};
-%		\end{axis}
-%		\end{tikzpicture}
-%		\caption{}
-%		\label{fig:adv_cg_convergence}
 	\end{subfigure}
@@
 		\includegraphics[scale=1.0]{adv_geigs_figure.pdf}
-%		\begin{tikzpicture}[baseline, trim axis right]
-%\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
-%ymode=log,
-%xmin=-50,
-%xmax=600,
-%ymax=1.e6,
-%title={Generalized eigenvalues},
-%xlabel={$k$, generalized eigenvalue \#},
-%ylabel={$\lambda_{k}$},
-%xtick={0, 300, 600, 900, 1200},
-%%legend pos=north east,
-%legend style={font=\small},
-%ylabel near ticks, ylabel shift={-6pt},
-%axis x line*=bottom,
-%axis y line*=right,
-%%legend style={at={(0.95,1)},anchor=north east},
-%]
-%%\addlegendentry{REG}
-%\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {advDiff_generalizedEigenvalues.dat};
-%\addplot[dash dot, line width=0.5mm] table[x=k, y=none] {advDiff_generalizedEigenvalues.dat};
-%%\addlegendentry{PSF (1)}
-%\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {advDiff_generalizedEigenvalues.dat};
-%%\addlegendentry{PSF (5)}
-%\addplot[red, line width=0.5mm] table[x=k, y=psf5] {advDiff_generalizedEigenvalues.dat};
-%%\addlegendentry{PSF (25)}
-%\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {advDiff_generalizedEigenvalues.dat};
-%\end{axis}
-%\end{tikzpicture}
-%		\caption{}
-%		\label{fig:adv_generalized_eigenvalues}
 	\end{subfigure}
@@
 \end{figure}
-
-%\begin{itemize}
-%	\item Picture of velocity field, initial condition, final condition, reconstruction (1) (OKOK)
-%	\item Impulse responses: vary time and Peclet number (3) (OKOK)
-%	\item CG convergence (1) CG iters vs Peclet number (Morozov) (2) (OKOK)
-%	\item Preconditioned spectrum, condition numbers? (1)
-%\end{itemize}
 
@@
 
-%\appendix
-
-%\section{Hierarchical matrix details}
-%\label{app:h_matrix}
-%
-%Often, large dense matrices of practical interest may be permuted, then partitioned into blocks recursively, in such a way that many off-diagonal blocks of the matrix are numerically low rank, even if the matrix is high rank.
-Such matrices are known as hierarchical matrices (H-matrices).
-Many classes of H-matrices exist (H1, H2, HSS, HBS, and more), and all of these types of H-matrices could be used in conjunction with our product-convolution approximation.
-Here we use classical H1 matrices.
-For this section, when we say H-matrix, we are referring to H1 matrices.
-%
-%While a dense $\fedim \times \fedim$ matrix traditionally requires $O(\fedim^2)$ memory to store, H-matrices may be stored using $O(\fedim \log \fedim)$ memory, by storing only the low rank factors for the low rank blocks.
-Recursive algorithms can take advantage of the H-matrix low rank block structure to perform matrix arithmetic fast.
-Conventional dense matrix algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix-matrix addition require either $O(\fedim^2)$ or $O(\fedim^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(\fedim \log(\fedim)^a)$ time and memory for H-matrices.
-Here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix.
-For more details on H-matrices, we recommend~\cite{HMATRIXGOOD}.
-%
-%
-%\subsection{H-matrix construction}
-%\label{sec:H_matrix_construction}
-%
-%The process of constructing an H-matrix representation of $\mathbf{\Aker}$ proceeds as follows.
-First, we construct hierarchical partitionings of the degrees of freedom for the columns and rows of the matrix (cluster trees, Section~\ref{sec:cluster_trees}).
-Second, we construct a hierarchical partitioning of the blocks of the matrix, in such a way that many of the blocks in the partitioning are expected to be low rank, and the remaining high rank blocks are small (block cluster tree, Section~\ref{sec:block_cluster_tree}).
-Finally, we form low rank approximations of the blocks of the matrix that are expected to be low rank (adaptive cross approximation, Section~\ref{sec:adaptive_cross}), and fill in the remaining high rank  blocks with their numerical values.
-The first two steps require geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix, but these steps do not depend on the particular values of matrix entries.
-The third step requires us to evaluate $O(\fedim \log \fedim)$ specially-chosen entries of $\mathbf{\Aker}$, and we evaluate these entries using~\eqref{eq:Akerpcmat_entries}.
-%
-%
-%\subsubsection{Cluster trees}
-%\label{sec:cluster_trees}
-%
-%We use recursive hyperplane splitting to hierarchically cluster the degrees of freedom associated with the columns and rows of the matrix into a \emph{column cluster tree} and a \emph{row cluster tree}, respectively.
-Here we describe construction of the column cluster tree; the row cluster tree is constructed similarly.
-%
-%Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\mathbf{\Aker}$ corresponds to the Lagrange node in $\mathbb{R}^\gdim$ associated with the $i\text{th}$ finite element basis function.
-The columns of the matrix thus correspond to a point cloud in $\mathbb{R}^\gdim$.
-We split this point cloud into two equally sized \emph{child} point clouds, using a hyperplane which is perpendicular to the coordinate axis direction in which the point cloud is widest (e.g., either the $x$, $y$, or $z$ axis in $3$D).
-The two child point clouds are split in the same way.
-This splitting process repeats until the point clouds have less than a preset number of points (we use $32$ points).
-This hierarchical partitioning of the point cloud into smaller and smaller point clouds corresponds to a hierarchical partitioning of the columns of the matrix into smaller and smaller \emph{clusters} of columns.
-This hierarchical partitioning of the columns forms a binary tree, which is called the column cluster tree.
-The root of the tree is the set of all columns, and the leaves of the tree are clusters of columns that are not subdivided any further.
-%
-%A depth-first search ordering of the column cluster tree leaves is then generated.
-When the columns of the matrix are permuted into this depth-first ordering, the columns associated with each cluster in the cluster tree are contiguous.
-%
-%In the same way, the degrees of freedom associated with the rows of the matrix are hierarchically clustered into another cluster tree, and a depth-first search ordering for the rows is generated.
-In our examples, the degrees of freedom for the columns coincide with the degrees of freedom for the rows, so the cluster trees for the rows and columns are the same, but this is not required in general.
-%
-%
-%\subsubsection{Block cluster tree} 
-%\label{sec:block_cluster_tree}
-%
-%We partition the matrix into a recursive hierarchy of mostly low rank blocks called the \emph{block cluster tree}.
-The idea is that a block of the matrix is likely to be low rank if the point cloud associated with the rows of the block is far away from the point cloud associated with the columns of the block.
-This is reasonable to expect here because of the locality property of $\Aop$.
-Indeed, locality implies that many blocks of the matrix corresponding to far away point cloud clusters will be rank zero.
-%
-%After reordering the rows and columns of $\mathbf{\Aker}$ via the depth-first ordering described above, we partition the reordered version of $\mathbf{\Aker}$ into a tree of $2 \times 2$ block matrices recursively.
-We use a geometric admissibility condition (discussed below) to decide which blocks to subdivide, and use the cluster trees for the rows and columns to determine how to subdivide those blocks.
-For the first stage of subdivision, let $r_1$ and $r_2$ be the children row clusters for the root of the row cluster tree, let $\colcluster_1$ and $\colcluster_2$ be the children column clusters for the root of the column cluster tree.
-The matrix $\mathbf{\Aker}$ is partitioned into blocks as follows:
-%\begin{equation*}
-%	\begin{bmatrix}
-%		\mathbf{\Aker}_{11} & \mathbf{\Aker}_{12} \\
-%		\mathbf{\Aker}_{21} & \mathbf{\Aker}_{22}
-%	\end{bmatrix},
-%\end{equation*}
-%where $\mathbf{\Aker}_{11}$ denotes the block of $\mathbf{\Aker}$ with rows $r_1$ and columns $\colcluster_1$, $\mathbf{\Aker}_{12}$ denotes the block of $\mathbf{\Aker}$ with rows $r_1$ and columns $\colcluster_2$, and so on for $\mathbf{\Aker}_{21}$ and $\mathbf{\Aker}_{22}$.
-%
-%We now loop through the four blocks, $\mathbf{\Aker}_{11}$, $\mathbf{\Aker}_{12}$, $\mathbf{\Aker}_{21}$, and $\mathbf{\Aker}_{22}$, and decide which blocks should be subdivided further.
-For the purpose of explanation, consider $\mathbf{\Aker}_{12}$.
-If
-%\begin{equation}
-%	\label{eq:weak_admissibility_cond}
-%	\dist\left(r_1, \colcluster_2\right) \ge \weakadmconst \min\left(\diam\left(r_1\right), \diam\left(\colcluster_2\right)\right),
-%\end{equation}
-%then we mark $\mathbf{\Aker}_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone.
-Here $\dist\left(r_1, \colcluster_2\right)$ is the Euclidean distance between the axis-aligned bounding box for the point cloud associated with the row cluster $r_1$, and the axis aligned bounding box for the point cloud associated with the column cluster $\colcluster_2$.
-The quantity $\diam\left(r_1\right)$ is the diameter of the axis aligned bounding box for the point cloud associated with the row cluster $r_1$, and $\diam\left(\colcluster_2\right)$ is the analogous diameter associated with the column cluster $\colcluster_2$.
-Here the quantity $\weakadmconst$ is a scalar constant; we use $\weakadmconst=2.0$.
-Basically, if the point clouds associated with $r_1$ and $\colcluster_2$ are far away from each other relative to their diameters, then we expect that the corresponding block of the matrix will be low rank.
-This process is repeated for the other blocks to determine which blocks are admissible and which are not.
-For us, the diagonal blocks $\mathbf{\Aker}_{11}$ and $\mathbf{\Aker}_{22}$ are not admissible because the distance between a point cloud and itself is zero.
-%
-%Next, we subdivide all blocks that are not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\mathbf{\Aker}$.
-But now we subdivide a block based on the two child row clusters and two child column clusters for the rows and columns of that block.
-This subdivision process continues recursively until all blocks are either admissible, or smaller than the predetermined size mentioned above.
-The resulting hierarchical partitioning of matrix blocks forms a tree, which is called the \emph{block cluster tree}.
-The root of the tree is the whole matrix, internal nodes in the tree are blocks that are subdivided, and the leaves of the tree are blocks that are either expected to be low rank, or are small.
-%
-%
-%\subsubsection{Adaptive cross low rank approximation of blocks}
-%\label{sec:adaptive_cross}
-%
-%Once the block cluster tree has been constructed, low rank approximations of the admissible (low rank) blocks are formed using the adaptive cross method~\cite{ACA}.
-Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\mathbf{\Aker}$.
-The idea of the adaptive cross method is to form a low rank approximation of $X$ by sampling a small number of rows and columns of $X$.
-%\begin{itemize}
-%	\item Let $C \in \mathbb{R}^{m \times \hrank}$ be a matrix consisting of a subset of $\hrank$ columns of $X$, such that the span of the columns in $C$ approximates the column space of $X$. 
-%	\item Let $R\in\mathbb{R}^{\hrank \times m}$ be a subset of the rows of $X$, such that the span of the rows in $R$ approximates the row space of $X$.
-%	\item Let $U \in \mathbb{R}^{\hrank \times \hrank}$ be the block of $X$ consisting of the intersection of the rows from $R$ with the columns from $C$.
-%\end{itemize}
-%Then it is well-established that
-%\begin{equation}
-%	\label{eq:CUR}
-%	X \approx C U^+ R,
-%\end{equation}
-%where $U^+$ is the pseudoinverse of $U$~\cite{goreinov1997theory,MahoneyDrineas09}.
-The quality of approximation~\eqref{eq:CUR} depends on how well the columns of $C$ approximate the column space of $X$, and how well the rows of $R$ approximate the row space of $X$.
-%
-%In the adaptive cross method, ``good'' columns, $C$, and rows, $R$, are chosen via an alternating iterative process.
-An initial set of columns $C$ is chosen.
-Keeping $C$ fixed, a set of rows $R$ is chosen so that the so that determinant of the submatrix $U$ within $C$ is as large as possible~\cite{GoreinovEtAl10}.
-Now, keeping $R$ fixed, a set of new columns $C$ is chosen so that the submatrix $U$ within $R$ is as large as possible.
-This process repeats a small number of times.
-This results in matrices $R$ and $C$ such that the error in the approximation,~\eqref{eq:CUR}, is small.
-The dominant cost of this procedure is the cost of computing $a \hrank$ columns of $X$ and $a\hrank$ rows of $X$, where $a$ is the number of alternating iterations.
-There is also a small linear algebra overhead cost for the determinant maximization process that is performed at each step.
-%The key point is that the adaptive cross method allows us to form a rank-$\hrank$ approximation of an $m \times m$ block of the matrix via a process that only requires accessing $O(m\hrank)$ entries of that block.
-%
-%We use the adaptive cross method to form low rank approximations for each admissible block of $\mathbf{\Aker}$.
-We directly compute all entries of the small dense blocks of $\mathbf{\Aker}$ that are not admissible.
-This process requires us to compute $O(\hrank^2 \fedim \log \fedim)$ entries of $\mathbf{\Aker}$, which is relatively cheap compared to the operator actions of $\Aop$ that are required to form the product-convolution approximation.
-
-%\section{Ellipsoid intersection test}
-%\label{sec:fast_ellipsoid_intersection_test}
-%The procedure for choosing sample points relies on quickly determining whether two ellipsoids intersect. Let $E_1$ and $E_2$ be the ellipsoids defined as
-%\begin{align*}
-%	E_i :=& \{x : (x - \spatialmean_i)^T \spatialcov_i^{-1} (x - \spatialmean_i) \le \tau^2\}
-%\end{align*}
-%for $i\in\{1,2\}$, where $\spatialmean_1, \spatialmean_2 \in \mathbb{R}^\gdim$, and $\spatialcov_1, \spatialcov_2 \in \mathbb{R}^{\gdim \times \gdim}$ are positive definite.
-Let $K$ be the following one dimensional convex function:
-%\begin{equation*}
-%	K(s) := 1 - \frac{1}{\tau^2} (\spatialmean_1 - \spatialmean_2)^T \left(\frac{1}{1-s}\spatialcov_1 + \frac{1}{s}\spatialcov_2\right)^{-1}(\spatialmean_1 - \spatialmean_2).
-%\end{equation*}
-%In~\cite{GilitschenskiHanebeck12} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$.
-%% (see Figure~\ref{fig:ellipsoid_intersection_test}). 
-%
-%%\begin{figure}
-%%	\begin{subfigure}{0.37\textwidth}
-%%		\centering
-%%		\includegraphics[scale=1.0]{ellipsoids_intersect0.pdf}
-%%	\end{subfigure}
-%%	\begin{subfigure}{0.30\textwidth}
-%%		\centering
-%%		\includegraphics[scale=1.0]{ellipsoids_intersect1.pdf}
-%%	\end{subfigure}
-%%	\begin{subfigure}{0.30\textwidth}
-%%		\centering
-%%		\includegraphics[scale=1.0]{ellipsoids_intersect2.pdf}
-%%	\end{subfigure}
-%%	\caption{The ellipsoids intersect if $K(s) \ge 0$ for all $s \in (0,1)$.
-%%	}
-%%	\label{fig:ellipsoid_intersection_test}
-%%\end{figure}
-%
-%The function $K(s)$ may be evaluated quickly for many $s$ by computing the solution to the generalized eigenvalue problem
-%$
-%	\spatialcov_1 \eigenvectormatrix = \spatialcov_2 \eigenvectormatrix \Lambda,
-%$
-%where $\eigenvectormatrix \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors, and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$.
-%%The matrix $\eigenvectormatrix$ simultaneously diagonalizes $\spatialcov_1$ and $\spatialcov_2$, in the sense that $\eigenvectormatrix^T\spatialcov_1 \eigenvectormatrix = \Lambda$, and $\eigenvectormatrix^T\spatialcov_2 \eigenvectormatrix = I$, where $I$ is the $\gdim \times \gdim$ identity matrix.
-Using this diagonalization, and
-%After some algebraic manipulations, we have
-%\begin{equation}
-%	\label{eq:Ks_generalized}
-%	K(s) = 1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2,
-%\end{equation}
-%where $v := \eigenvectormatrix^T\left(\spatialmean_1 - \spatialmean_2\right)$.
-To check whether $E_1$ and $E_2$ intersect, we compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, form $v$, then minimize $K(s)$ in the form~\eqref{eq:Ks_generalized} on $(0,1)$ using Brent's method~\cite{Brent71}.
-If $K(s^*) <0$ at the minimizer, $s^*$, then $E_1 \cap E_2 = \{\}$.
-Otherwise $E_1 \cap E_2 \neq \{\}$.
-%%The resulting algorithm for checking whether $E_1$ and $E_2$ intersect is summarized in Algorithm~\ref{alg:ellipsoid_intersection_test}.
-%
-%%\begin{algorithm2e}
-%%	\SetAlgoNoLine
-%%	\SetKwInOut{Input}{Input}
-%%	\SetKwInOut{Output}{Output}
-%%	\SetKwProg{Fn}{Function}{}{}
-%%	\Input{Ellipsoid $E_1$ with mean $\spatialmean_1$ and covariance $\spatialcov_1/\tau^2$\\
-%%		Ellipsoid $E_2$ with mean $\spatialmean_2$ and covariance $\spatialcov_2/\tau^2$ 
-%%	}
-%%	
-%%	\Output{Boolean $\text{ellipsoids\_intersect}$ which is true if $E_1 \cap E_2 \neq \{\}$ and false otherwise}
-%%	
-%%	
-%%	Solve generalized eigenvalue problem $\spatialcov_1 P = \spatialcov_2 P \Lambda$
-%%		
-%%	$v \gets P^T\left(\spatialmean_1 - \spatialmean_2\right)$
-%%		
-%%	$\displaystyle K^* \gets \min_{s \in (0,1)}~1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2$
-%%		
-%%	\If{$K^* < 0$}{
-%%			
-%%		$\text{ellipsoids\_intersect} \gets \text{False}$
-%%			
-%%	}
-%%	\Else{
-%%			
-%%		$\text{ellipsoids\_intersect} \gets \text{True}$
-%%			
-%%	}
-%%	\caption{Determining whether two ellipsoids intersect}
-%%	\label{alg:ellipsoid_intersection_test}
-%%\end{algorithm2e}
-%
-%%\subsection{Ellipsoid bounding box test}
-%
-%This ellipsoid intersection test is only performed if the coordinate axis aligned bounding boxes for the ellipsoids intersect. The smallest coordinate axis aligned box that contains $E_i$ is the box
-%$
-%	\prod_{k=1}^\gdim \left[ \spatialmean_i^{k} -\tau\sqrt{\Sigma_i^{kk}}, \spatialmean_i^{k} +\tau\sqrt{\Sigma_i^{kk}} \right],
-%$
-%where $\spatialmean_i^{k}$ is the $k$th entry of $\spatialmean_i$, and $\Sigma_i^{kk}$ is the $k$th diagonal entry of $\Sigma_i$\footnote{See \url{https://math.stackexchange.com/q/3928964}}.
-The boxes
-%$\prod_{k=1}^\gdim \left[a_{-}^k, a_{+}^k\right]$ and $\prod_{k=1}^\gdim \left[b_{-}^k, b_{+}^k\right]$
-%intersect if and only if $a_{-}^k \le b_{+}^k$ and $b_{-}^k \le a_{+}^k$ for $k=1,\dots,d$.
-
-%\section{Rational positive semi-definite modification}
-%\label{sec:make_spd}
-%In many problems $\Aop$ is symmetric positive semi-definite. However, $\mathbf{A}_H$ is non-symmetric and indefinite because of approximation error. 
-%This is undesirable.
-Symmetry and positive semi-definiteness are important properties which should be preserved if possible.
-Also, lacking these properties may prevent one from using highly effective algorithms, such as the conjugate gradient method, to perform further operations involving $\mathbf{A}_H$.
-%
-%If $\Aop$ is symmetric, we symmetrize $\mathbf{A}_H$ as follows:
-%$
-%\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
-%$
-%If $\Aop$ is positive semi-definite, we adapt the rational method from~\cite[Section 13.8]{HackbuschKress07} to approximate the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$, which we denote by $|\mathbf{A}_H^{\text{sym}}|$.
-That is, $|\mathbf{A}_H^{\text{sym}}|$ has the same eigenvectors as $\mathbf{A}_H^{\text{sym}}$, but the eigenvalues of $|\mathbf{A}_H^{\text{sym}}|$ are the absolute values of the eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
-%%flip (or approximately flip) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ to be positive. 
-%Once constructed, we use the approximation of $|\mathbf{A}_H^{\text{sym}}|$ in place of $\mathbf{A}_H$ in further linear algebra operations.
-%
-%In detail, let $a$ be any scalar satisfying $-a < \lambda_\text{min}$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, and set 
-%\begin{equation}
-%\label{eq:ratT999}
-%\mathbf{T} :=  \left(2 \mathbf{A}_H^{\text{sym}} + a \mathbf{I}\right)/a,
-%\end{equation}
-%From the results in~\cite[Section 13.8]{HackbuschKress07} it is straightforward to show that
-%\begin{equation}
-%\label{eq:rational343}
-%\mathbf{\Pi}_\ratord := \left(\mathbf{I} + \mathbf{T}^{2^\ratord}\right)^{-1}
-%\end{equation}
-%approximates the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to negative eigenvalues. Thus, we have
-%\begin{equation}
-%\label{eq:rational76567}
-%|\mathbf{A}_H^{\text{sym}}| \approx \left(\mathbf{I} - 2\mathbf{\Pi}_\ratord\right)\mathbf{A}_H^{\text{sym}}.
-%\end{equation}
-%To form an approximation of $|\mathbf{A}_H^{\text{sym}}|$, first, we compute $\lambda_\text{min}$ using the Lanczos method, and set $a=\gamma \lambda_\text{min}$, where $\gamma\ge 1$ is a parameter that must be chosen.
-%The method is relatively insensitive to $\gamma$ and numerically we find that the method works well for any $\gamma \in [1,2]$. We use $\gamma=1.5$ in our numerical results. 
-%Next, we compute $\mathbf{T}$ per~\eqref{eq:ratT999}, compute $\mathbf{T}^{2^\ratord}$ per~\eqref{eq:rational343} via repeated squaring, compute $\mathbf{\Pi}_\ratord$ per~\eqref{eq:rational343}, and form an approximation of $|\mathbf{A}_H^{\text{sym}}|$ per~\eqref{eq:rational76567}.
-Fast H-matrix methods are used to perform the required matrix-matrix additions and multiplications, and matrix inversion.
-We recommend using $\ratord=1$ or $2$.
-Larger $\ratord$ is unnecessary and may make the method unstable.
-
-%Finally, we approximate $|\mathbf{A}_H^{\text{sym}}|$ per~\eqref{eq:rational76567}, using fast H-matrix operations to perform the required matrix-matrix addition and matrix inversion.
-
-%
-%define $
-%\mathbf{T} := T(\mathbf{A}_H^{\text{sym}}) = \left(2 \mathbf{A}_H^{\text{sym}} + a \mathbf{I}\right)/a.
-%$, where $-a < \lambda_\text{min}$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$ and define
-%
-%let $\mathbf{\Pi}$ be the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to eigenvalues of $\mathbf{A}_H^{\text{sym}}$ that reside in $(-a,0)$
-%% let $\Pi$ be the indicator function for the interval $(a,b)$. 
-%%That is, $\Pi(\lambda)=0$ for $\lambda \in (0,1)$, and $\Pi(\lambda)=0$ otherwise.
-%%\begin{equation*}
-%%	\Pi(\lambda) = \begin{cases}
-%%		1, & \lambda \in (a,b), \\
-%%		0, & \text{otherwise}.
-%%	\end{cases}
-%%\end{equation*}
-%%Using functional calculus to extend the domain of the function $\Pi$ from scalars to matrices, we define
-%%$
-%%\mathbf{\Pi} := \Pi(\mathbf{A}_H^{\text{sym}}).
-%%$
-%%The matrix $\mathbf{\Pi}$ is the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to eigenvalues of $\mathbf{A}_H^{\text{sym}}$ that reside in $(a,b)$. 
-%If we choose $-a < \lambda_\text{min}$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, then $\mathbf{\Pi}$ is the spectral projector onto the eigenspace associated with negative eigenvalues,
-%%and $\mathbf{\Pi}\mathbf{A}_H^{\text{sym}}$ is the ``negative component'' of $\mathbf{A}_H^{\text{sym}}$ in which all the negative eigenvalues are the same but all the positive eigenvalues are replaced by zero.
-Subtracting a negative number from itself twice flips its sign,
-%so the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$ is given by
-%\begin{equation}
-%\label{eq:I_Pi_abs}
-%	|\mathbf{A}_H^{\text{sym}}| = \left(\mathbf{I} - 2\mathbf{\Pi}\right)\mathbf{A}_H^{\text{sym}},
-%\end{equation}
-%where $\mathbf{I}$ is the identity matrix of the appropriate size.
-%By ``matrix absolute value,'' we mean that $|\mathbf{A}_H^{\text{sym}}|$ is the matrix that has the same eigenvectors as $\mathbf{A}_H^{\text{sym}}$, but the eigenvalues of $|\mathbf{A}_H^{\text{sym}}|$ are the absolute values of the eigenvalues of $\mathbf{A}_H^{\text{sym}}$.
-%%\begin{figure}
-%%	\centering
-%%	\includegraphics[scale=0.5]{spd_rational_function.pdf}
-%%	\caption{Rational function $\Pi_\ratord(\lambda)$ for $a=-1$, $b=0$, and $\ratord \in \{1,2,3\}$. As $\ratord$ increases, $\Pi_\ratord$ approaches the indicator function for the set $(a,b)$. }
-%%	\label{fig:spd_rat}
-%%\end{figure}
-%Computing $\mathbf{\Pi}$ directly is difficult, but we may approximate $\mathbf{\Pi}$ by the following high order rational matrix function,
-%%$\Pi_\ratord \approx \Pi$ which is defined as follows:
-%%\begin{equation*}
-%%	 \Pi_\ratord(\lambda) := \frac{1}{1 + T(\lambda)^{2^\ratord} }, \qquad \text{where} \qquad T(\lambda) := \frac{2 \lambda - (b+a)}{b-a}.
-%%\end{equation*}
-%%Here $\ratord$ is a positive integer; the larger $\ratord$ is, the better the approximation $\Pi \approx \Pi_\ratord$. 
-%%%This is illustrated in Figure~\ref{fig:spd_rat}. 
-%%Replacing scalar operations with matrix operations, we have an approximation $\mathbf{\Pi}_\ratord \approx \mathbf{\Pi}$ given by
-%\begin{equation}
-%\label{eq:rational343}
-%\mathbf{\Pi} \approx \mathbf{\Pi}_\ratord := \left(\mathbf{I} + \mathbf{T}^{2^\ratord}\right)^{-1},
-%\end{equation}
-%where
-%$
-%\mathbf{T} := T(\mathbf{A}_H^{\text{sym}}) = \left(2 \mathbf{A}_H^{\text{sym}} + a \mathbf{I}\right)/a.
-%$
-%Using $\mathbf{\Pi}_\ratord$ instead of $\mathbf{\Pi}$ in~\eqref{eq:I_Pi_abs} allows us to approximate $|\mathbf{A}_H^{\text{sym}}|$.
-First, we compute $\lambda_\text{min}$ using the Lanczos method, and set $a=\gamma \lambda_\text{min}$ and $b=0$, where $\gamma\ge 1$ is a parameter that must be chosen.
-%The method is relatively insensitive to $\gamma$ and numerically we find that the method works well for any $\gamma \in [1,2]$. We use $\gamma=1.5$ in our numerical results. 
-%Next, we compute $\mathbf{T}$ using fast H-matrix scalar multiplication and addition.
-We form the $\mathbf{T}^{2^\ratord}$ via repeated squaring using H-matrix multiplication.
-We recommend using $\ratord=1$ or $2$.
-Larger $\ratord$ is unnecessary and may make the method unstable.
-The approximate spectral projector $\mathbf{\Pi}_\ratord$ is formed from $\mathbf{T}^{2^\ratord}$ per~\eqref{eq:rational343}, using fast H-matrix operations to perform the required matrix-matrix addition and matrix inversion.
-Finally, a positive semi-definite approximation $\mathbf{A}_H^{\text{sym}+}$ of $\mathbf{A}_H^{\text{sym}}$ is formed as
-%\begin{equation*}
-%\mathbf{A}_H^{\text{sym}+} := \left(\mathbf{I} - 2\mathbf{\Pi}_\ratord\right)\mathbf{A}_H^{\text{sym}},
-%\end{equation*}
-%where fast H-matrix operations are used to perform the necessary matrix-matrix subtraction and multiplication.
-Note that for any $\ratord \ge 1$, $\Pi_\ratord(\lambda)> 1/2$ for $\lambda \in (a,b)$, so $\mathbf{A}_H^{\text{sym}+}$ is positive semi-definite\footnote{Some indefiniteness may be introduced by truncation error in the H-matrix operations, but the negative eigenvalues caused by this truncation error are typically extremely small.}.
-At the end, $\mathbf{A}_H^{\text{sym}+}$ is symmetrized once again, to correct for any small non-symmetries that are introduced by truncation error in the H-matrix operations.
-Once constructed, we use the symmetric positive semi-definite matrix $\mathbf{A}_H^{\text{sym}+}$ in place of $\mathbf{A}_H$ in further linear algebra operations.
-
-%\section{Recycling Krylov information via low rank updates}
-%\label{sec:recycle_krylov}
-%
-%In applications, often one wants to use the H-matrix approximation $\mathbf{A}_H$ to build preconditioners for solving a sequence of slowly varying linear systems, where the $k$th linear system has the form
-%\begin{equation}
-%\label{eq:Ax_equals_b}
-%\mathbf{B}^{(k)} \mathbf{u}^{(k)} = \mathbf{f}^{(k)}.
-%\end{equation}
-%Each of these linear systems is solved using an iterative method. By ``slowly varying,'' we mean that $\mathbf{B}^{(k+1)}$ does not differ much from $\mathbf{B}^{(k)}$. 
-%For example, in a Newton-Krylov method for solving an optimization problem, one uses a Krylov method such as conjugate gradient to solve a sequence of systems of the form~\eqref{eq:Ax_equals_b}, where $\mathbf{B}^{(k)}$ and $\mathbf{f}^{(k)}$ are the Hessian and the negative gradient of the objective function in the optimization problem, respectively, evaluated at the $k$th Newton iterate.
-%As the optimization solver converges, $\mathbf{B}^{(k)}$ converges to the Hessian at the optimal point, and therefore changes little from one iteration to the next.
-Since $\mathbf{B}^{(k)}$ varies slowly, it is reasonable to assume that Krylov information about $\mathbf{B}^{(k)}$ can be leveraged to update the preconditioner from $\mathbf{B}^{(k)}$ in order to build a preconditioner for $\mathbf{B}^{(k+1)}$.
-%Concretely, let 
-%\begin{equation*}
-%\mathbf{P}^{(k)} \approx \left(\mathbf{B}^{(k)}\right)^{-1}
-%\end{equation*}
-%denote a H-matrix preconditioner used to accelerate the solution of~\eqref{eq:Ax_equals_b} with an iterative method.
-The iterative method for solving the $k$th linear system will typically involve, as a sub-problem, applying $\mathbf{B}^{(k)}$ to a collection of vectors $\{\mathbf{x}_i\}_{i=1}^m$, thereby generating vectors $\{\mathbf{y}_i = \mathbf{B}^{(k)} \mathbf{x}_i\}_{i=1}^m$.
-Instead of rebuilding the H-matrix preconditioner for each successive linear system, we \emph{recycle} the information generated by the iterative method while solving the current linear system, performing a low-rank update to $\mathbf{P}^{(k)}$ to create $\mathbf{P}^{(k+1)}$.
-To do so, we build on the rank-1 update formulas discussed in~\cite[Chapter 6]{NocedalWright99}, which we generalize to many vector ``rank-$m$'' updates.
-In what follows, let $\mathbf{X}$ and $\mathbf{Y}$ denote the $\fedim \times m$ matrices with columns given by the vectors $\mathbf{x}_i$ and $\mathbf{y}_i$, respectively.
-The matrix-vector products with $\mathbf{B}^{(k)}$ performed while solving the $k$th linear system may be summarized by the equation:
-%\begin{equation*}
-%\mathbf{B}^{(k)}\mathbf{X} = \mathbf{Y}.
-%\end{equation*}
-%To form $\mathbf{P}^{(k+1)}$ from $\mathbf{P}^{(k)}$, one may use one of the following three update formulas:
-%\begin{align}
-%	\mathbf{P}^{(k+1)} &= \mathbf{P}^{(k)} + \mathbf{R}\left(\mathbf{Y}^T \mathbf{Y}\right)^{-1} \mathbf{Y}^T \label{eq:res_update} \\
-%	\mathbf{P}^{(k+1)} &= \mathbf{P}^{(k)} + \mathbf{R} \left(\mathbf{R}^T \mathbf{Y}\right)^{-1} \mathbf{R}^T \label{eq:srk_update} \\
-%	\mathbf{P}^{(k+1)} &= \left(\mathbf{I} - \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T\right)\mathbf{P}^{(k)}\left(\mathbf{I} - \mathbf{Y}\boldsymbol{\Theta}\mathbf{X}^T\right) + \mathbf{X}\boldsymbol{\Theta}\mathbf{X}^T, \label{eq:dfp_update}
-%\end{align}
-%where $\mathbf{R} := \mathbf{X}-\mathbf{P}^{(k)}\mathbf{Y}$ and $\boldsymbol{\Theta} := \left(\mathbf{X}^T\mathbf{Y}\right)^{-1}$.
-%We note that for all three cases, the update is exact for $\mathbf{B}^{k}$, in the sense that 
-%\begin{equation*}
-%\mathbf{P}^{(k+1)}\mathbf{Y}=\mathbf{X} = \left(\mathbf{B}^{(k)}\right)^{-1} \mathbf{Y}.
-%\end{equation*}
-%Because the matrices $\mathbf{B}^{k}$ vary slowly, we expect that $\mathbf{P}^{(k+1)}$ is a good approximation for $\left(\mathbf{B}^{(k+1)}\right)^{-1}$ on the space spanned by the vectors $\mathbf{y}_i$.
-One should choose which formula to use based on the properties of the matrix $\mathbf{B}^{(k)}$.
-Formula~\eqref{eq:res_update} does not preserve symmetry, and should therefore be used when $\mathbf{B}^{(k)}$ is not symmetric.
-Formula~\eqref{eq:srk_update} preserves symmetry but not positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric and indefinite.
-Formula~\eqref{eq:dfp_update}, which is used in this paper, preserves both symmetry and positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric positive definite.
-Formulas~\eqref{eq:res_update} and~\eqref{eq:srk_update} perform rank-$m$ updates, while formula~\eqref{eq:dfp_update} performs a rank-$2m$ update.
-%We note that when $m=1$, formulas~\eqref{eq:srk_update} are~\eqref{eq:dfp_update} are known as SR1 (symmetric rank-$1$) and DFP (Davidson-Fletcher-Powell) updates, respectively~\cite[Chapter 6]{NocedalWright99}.
-
 \section*{Acknowledgments}

### 2023-07-04 paper/localpsf.tex
remove commented out parts of paper. Will keep z notation
working through Omar's edits

@@
 
-% CITATIONS TO DO
-%
-% TASKS:
-% 1) Incorporate Ronald Kriemann's suggested citations: (Nick)
-%    - Steffen Börm, Lars Grasedyck and Wolfgang Hackbusch:
-%      "Introduction to Hierarchical Matrices with Applications",
-%       Engineering Analysis with Boundary Elements 27(5), pp. 405--422, 2003. (DONE)
-%    - Lars Grasedyck, Ronald Kriemann and Sabine Le Borne:
-%      "Parallel blackbox H-LU preconditioning for elliptic boundary value problems",
-%       Computing and Visualization in Sciences, Vol. 11, pp 273--291, 2007. (DONE)
-%    - Ronald Kriemann:
-%      "H-LU Factorization on Many-Core Systems",
-%      Computing and Visualization in Science, 16, pp. 105-117, 2015. (DONE)
-%
-% 2) Grants
-%    - Nick and Omar's grants (Omar)
-%    - Tucker and Noemi's grants (DONE)
-%
-% 3) Order citations in ascending order within brackets (Nick)
-%    e.g., [1,2,3] = good
-%          [2,3,1] = bad
-%
-% 4) Find and cite relevant papers from Noemis's group
-%    - Tucker's technical report (Tucker)
-%
-% 5) Find and cite relevant papers from Omar's group
-%    - Toby's papers (Tucker)
-%    - Noemi's 2012 stokes paper (Tucker)
-%    - Omar and Karen's recent review paper (DONE)
-%
-% 6) Look through Nick's previous PC paper for missed citations (DONE)
-
-% For citations we need:
-%    a) bibtex info
-%    b) exactly where in the paper to add the ref
-%    c) any words or sentences to add with the ref
-
 \newcommand{\todo}[1]{\noindent\emph{\textcolor{red}{To do: #1\:}}}
@@
 The method computes impulse responses of the operator at a collection of scattered points, and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel.
-By applying the operator to Dirac combs associated with batches of point sources, we compute a large number of impulse responses using a small number of operator applies.
+By applying the operator to Dirac combs associated with batches of point sources, we compute a large number of impulse responses using a small number of operator applications.
 We solve an ellipsoid packing problem to select as many point sources as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
@@
 The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
-Further matrix computations are performed with H-matrix methods.
+Further matrix computations such as matrix vector products, matrix-matrix additions, matrix factorizations, and the solution of linear systems are performed with H-matrix methods.
 We apply the proposed method to approximate the Hessian operator arising in inverse problems governed by partial differential equations (PDEs).
 While for many ill-posed inverse problems the Hessian of the data misfit term exhibits a low rank structure, and hence a low rank approximation is suitable, for many problems of practical interest the numerical rank of this operator is still large.
-This makes conventional matrix-free Hessian approximation methods expensive in terms of the required number of Hessian operator applies (hence PDE solves).
+This makes conventional matrix-free Hessian approximation methods expensive in terms of the required number of Hessian operator applications (hence PDE solves).
 In contrast, the performance of the proposed method depends on the locality and shapes of the impulse responses rather than the numerical rank.
-Hessian impulse responses typically become more local as the numerical rank increases, which makes this operator a perfect candidate for the proposed approach.
-We illustrate the proposed approach with two specific problems: inversion for the basal sliding coefficient in an ice sheet flow problem and inversion for the initial condition in a diffusive transport problem.
+Hessian impulse responses typically become more local as the numerical rank increases, which makes high rank Hessians perfect candidates for the proposed approach.
+We illustrate the proposed approach with two specific problems: inversion for the basal friction coefficient in an ice sheet flow problem and inversion for the initial condition in a diffusive transport problem.
 To demonstrate the effectiveness of the approximation, we use the proposed approach to build preconditioners to accelerate the solution of the inverse problems.
@@
 These parameters have an effect on the shape of the impulse responses and on the rank of the Hessian.
-The results reveal that PSF-based preconditioners are able to form good approximations of high-rank Hessians using a small number of operator applies.
+The results reveal that PSF-based preconditioners are able to form good approximations of high-rank Hessians using a small number of operator applications.
 \end{abstract}
@@
 \begin{keywords}
- 	data scalability, Hessian, hierarchical matrix, matrix-free, operator approximation, PDE constrained inverse problems, point spread function, product convolution, local translation invariance, impulse response, moment methods
+ 	data scalability, Hessian, hierarchical matrix, matrix-free, operator approximation, PDE constrained inverse problems, point spread function, product convolution, local translation invariance, impulse response, moment methods, high-rank
 \end{keywords}
@@
 Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, and $L^2(\Omega)'$ is the space of real-valued continuous linear functionals on $L^2(\Omega)$.
-Such operators appear, for instance, as Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{AntilEtAl18,BorziSchulz11}, Schur complements in Schur complement methods for solving partial differential equations (PDEs)~\cite{le1997non,saad1999distributed}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics~\cite{ChenStein21,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}.
+Such operators appear, for instance, as Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{AntilEtAl18,BorziSchulz11,Delosreyes15,HinzePinnauUlbrichUlbrich08}, Schur complements in Schur complement methods for solving partial differential equations (PDEs) and Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps)~\cite{ChanMathew94,SmithBjorstadGropp96,ToselliWidlund04}, covariance operators in spatial statistics~\cite{ChenStein21,GentonKeyesTurkiyyah18,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}.
 Here ``matrix-free'' means that we may apply $\Aop$ and its transpose\footnote{Recall that $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for operations with $\Aop^T$.
@@
 but cannot easily access entries of $\Aop$'s integral kernel.
-Evaluating the maps in~\eqref{eq:eval_maps} may require solving a sub problem that involves PDEs, or performing other costly computations.
+Evaluating the maps in~\eqref{eq:eval_maps} may require solving a subproblem that involves PDEs, or performing other costly computations.
 By ``non-negative integral kernel,'' we mean that entries of the integral kernel are non-negative numbers; this is not the same as positive semi-definiteness of $\Aop$.
@@
 
-We use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applies.
+We use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applications.
 The impulse response, $\phi_x$, associated with a point, $x$, is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
@@
 PC and PSF methods have a long history dating back several decades.
-We note the following papers (among many others), \cite{Adorf94,AlgerEtAl19,BigotEscandeWeiss19,EscandeWeiss15,EscandeWeiss22,EscandeWeiss12,FishEtAl96,NagyOleary98,ZhuLiFomelEtAl16}, in which the convolution kernels are constructed from sampling impulse responses of the operator to scattered point sources.
+We note the following papers (among many others) in which the convolution kernels are constructed from sampling impulse responses of the operator to scattered point sources:~\cite{Adorf94,AlgerEtAl19,BigotEscandeWeiss19,EscandeWeiss15,EscandeWeiss22,EscandeWeiss12,FishEtAl96,NagyOleary98,ZhuLiFomelEtAl16}.
 For background on PC and PSF methods, we recommend the following papers: \cite{DenisEtAl15,EscandeWeiss17,GentileCourbinMeylan13}.
@@
 
-\section{Why we need high rank Hessian approximations}
+\section{Why we need more efficient approximations of high rank Hessians}
 \label{sec:hessian}
@@
 In distributed parameter inverse problems governed by PDEs, one seeks to infer an unknown spatially varying parameter field from limited observations of a state variable that depends on the parameter implicitly through the solution of a PDE.
-Conventionally, the inverse problem is formulated using either a deterministic framework~\cite{AsterBorchersThurber18,BorziSchulz11,Vogel02},
+Conventionally, the inverse problem is formulated using either a deterministic framework~\cite{AsterBorchersThurber18,BanksKunisch89,Vogel02},
 or a Bayesian probabilistic framework~\cite{KaipioSomersalo05,Stuart10a,Tarantola05}.
@@
 In the probabilistic framework, Bayes' theorem combines the observations with prior information to form a posterior distribution over the space of all possible parameter fields, and computations are performed to extract statistical information about the parameter from this posterior.
-The Hessian of the objective function in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian.'' The Hessian consists of a data misfit term (the \emph{data misfit Hessian}), which depends on a discrepancy between the observations and the associated model predictions, and a regularization or prior term (the \emph{regularization Hessian}) which does not depend on the observations.
+The Hessian of the objective function with respect to the parameter in the determinstic optimization problem and the Hessian of the negative log posterior in the Bayesian setting are equal or approximately equal under typical noise, regularization, and prior models, so we refer to both of these Hessians as ``the Hessian.'' The Hessian consists of a data misfit term (the \emph{data misfit Hessian}), which depends on a discrepancy between the observations and the associated model predictions, and a regularization or prior term (the \emph{regularization Hessian}) which does not depend on the observations.
 For more details on the Hessian, see~\cite{Alger19,GhattasWillcox21,VillaPetraGhattas21}.
@@
 Owing to the implicit dependence of predicted observations on the parameter, entries of the Hessian are not easily accessible.
-Rather, the Hessian may be applied to a vector via a computational process that involves solving a pair of forward and adjoint PDEs which are linearizations of the original PDE~\cite{GhattasWillcox21,Gunzburger03}.
+Rather, the Hessian may be applied to a vector via a computational process that involves solving a pair of forward and adjoint PDEs which are linearizations of the original PDE~\cite{GhattasWillcox21}
+%,Gunzburger03}
+.
 The most popular matrix-free Hessian approximation methods are based on low rank approximation of either the data misfit Hessian, or the data misfit Hessian preconditioned by the regularization Hessian, e.g.,~\cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15}.
@@
 
-Examples of inverse problems with high rank data misfit Hessians include large-scale ice sheet inverse problems \cite{HartlandEtAl23,IsaacEtAl15}, advection dominated advection-diffusion inverse problems~\cite{AkcelikBirosDraganescuEtAl05}\cite[Chapter 5]{Flath13}, high frequency wave inverse problems~\cite{BuiEtAl13}, and more generally, all inverse problems in which the observations highly inform the parameter.
-The eigenvalues of the data misfit Hessian characterize the sensitivity of the predicted observations to perturbations of the parameter in the corresponding eigenvector directions,
+Examples of inverse problems with high rank data misfit Hessians include large-scale ice sheet inverse problems \cite{HartlandEtAl23,IsaacEtAl15}, advection dominated advection-diffusion inverse problems~\cite{AkcelikBirosDraganescuEtAl05}\cite[Chapter 5]{Flath13}, high frequency wave propagation inverse problems~\cite{BuiEtAl13}, inverse problems governed by high Reynolds number flows, and more generally, all inverse problems in which the observations highly inform the parameter.
+The eigenvalues of the data misfit Hessian characterize how informative the data are about components of the parameter in the corresponding eigenvector directions,
 hence more informative data leads to larger eigenvalues and a larger numerical rank~\cite{AlexanderianGloorGhattas16}\cite[Section 1.4 and Chapter 4]{Alger19}.
@@
 would instead be reconstructed based on the regularization~\cite[Section 4]{AlgerEtAl17}\cite[Chapters 1 and 7]{Vogel02}.
-Hence, low rank approximation methods suffer from a predicament: if the data highly inform the parameter and the regularization is chosen appropriately, then a large number of operator applies are required to form an accurate approximation of the Hessian using low rank approximation methods.
-High rank Hessian approximation methods are needed.
+Hence, low rank approximation methods suffer from a predicament: if the data highly inform the parameter and the regularization is chosen appropriately, then a large number of operator applications are required to form an accurate approximation of the Hessian using low rank approximation methods.
+High rank Hessian approximation methods are thus needed.
 
@@
 Although these methods are promising,
-the required number of operator applies is still large (e.g., hundreds to thousands).
+the required number of operator applications is still large (e.g., hundreds to thousands).
 In this paper, we also form an H-matrix Hessian approximation.
-However, to reduce the required number of operator applies, we first form a PSF approximation of the data misfit Hessian by exploiting locality and non-negative integral kernel properties, then form the H-matrix using classical H-matrix techniques.
+However, to reduce the required number of operator applications, we first form a PSF approximation of the data misfit Hessian by exploiting locality and non-negative integral kernel properties, then form the H-matrix using classical H-matrix techniques.
 Not all data misfit Hessians satisfy the local non-negative integral kernel properties (for example, the wave inverse problem data misfit Hessian has a substantial amount of negative entries in its integral kernel).
@@
 Computing $\phi_x$ in order to determine $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ would be extremely computationally expensive, and defeat the purpose of the proposed method, which is to reduce the computational cost by computing impulse responses in batches.
-Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ indirectly, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions (e.g., 6 total operator applies in two spatial dimensions and 10 in three spatial dimensions).
+Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ indirectly, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions (e.g., 6 total operator applications in two spatial dimensions and 10 in three spatial dimensions).
 This may be motivated by analogy to matrices.
@@
 \paragraph{(1) Computing impulse response moments and batches} Computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ requires applying $\Aop$ to $1$, $\gdim$, and $\gdim(\gdim+1)/2$ vectors, respectively.
-Computing each $\eta_i$ requires applying $\Aop$ to one vector, so computing $\{\eta_i\}_{i=1}^{\nbatch}$ requires $\nbatch$ operator applies.
+Computing each $\eta_i$ requires applying $\Aop$ to one vector, so computing $\{\eta_i\}_{i=1}^{\nbatch}$ requires $\nbatch$ operator applications.
 In total, computing all impulse response moments and batches therefore requires
 \begin{equation*}
-	1 + \gdim + \gdim(\gdim+1)/2 + \nbatch \qquad \text{operator applies.}
+	1 + \gdim + \gdim(\gdim+1)/2 + \nbatch \qquad \text{operator applications.}
 \end{equation*}
-In a typical application one might have $\gdim=2$ and $\nbatch=5$, in which case a modest $11$ operator applies are required.
+In a typical application one might have $\gdim=2$ and $\nbatch=5$, in which case a modest $11$ operator applications are required.
 
@@
 
-In summary, computing all the necessary ingredients to evaluate kernel entries of the PSF approximation requires a handful of operator applies (e.g., $6+\nbatch$ operator applies in two dimensions, or $10+\nbatch$ operator applies in three dimensions, with $\nbatch$ typically in the range 1--25), plus comparatively cheap additional overhead costs, most notably performing ellipsoid intersection tests while choosing sample point batches.
-Once these ingredients are computed, no more operator applies are required, and approximate kernel entries can be evaluated fast.
+In summary, computing all the necessary ingredients to evaluate kernel entries of the PSF approximation requires a handful of operator applications (e.g., $6+\nbatch$ operator applications in two dimensions, or $10+\nbatch$ operator applications in three dimensions, with $\nbatch$ typically in the range 1--25), plus comparatively cheap additional overhead costs, most notably performing ellipsoid intersection tests while choosing sample point batches.
+Once these ingredients are computed, no more operator applications are required, and approximate kernel entries can be evaluated fast.
 Constructing the H-matrix from kernel entries requires a number of elementary operations that scales nearly linear in $N$.
@@
 
-\subsection{Example 1: Inversion for the basal sliding coefficient in an ice sheet flow problem}
+\subsection{Example 1: Inversion for the basal friction coefficient in an ice sheet flow problem}
 \label{sec:ice}
@@
 For this example, we consider a sheet of ice flowing down a mountain (see Figure~\ref{fig:ice_mountain_mesh}).
-Given observations of the tangential component of the ice velocity on the top surface of the ice, we invert for the logarithm of the unknown spatially varying basal sliding Robin coefficient field, which governs the resistance to sliding along the base of the ice sheet.
+Given observations of the tangential component of the ice velocity on the top surface of the ice, we invert for the logarithm of the unknown spatially varying basal friction Robin coefficient field, which governs the resistance to sliding along the base of the ice sheet.
 The setup, which we briefly summarize, follows~\cite{IsaacEtAl15,PetraEtAl12}.
@@
 Here, $\basalfriction$ is the unknown logarithmic
-basal sliding field (large $\basalfriction$ corresponds to high resistance to sliding) defined on the surface $\Gamma_b$.
+basal friction field (large $\basalfriction$ corresponds to high resistance to sliding) defined on the surface $\Gamma_b$.
 The quantity $\bodyforce$ is the body force density due to gravity, $\stokesrobincoeff=10^6$ is a Robin boundary condition constant,  $\normalvec$ is the outward normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary.
@@
 We use $5\%$ noise in all cases, except for Figure~\ref{fig:stokes_reconstructions} and Table \ref{tab:condition_number} where the noise is varied from $1\%$ to $25\%$ and the regularization is determined by the Morozov discrepancy principle for each noise level.
-The true basal sliding
+The true basal friction
 coefficient and velocity fields, which are obtained by
@@
     \end{subfigure}
-	\caption{(Ice sheet) The log basal sliding parameter, with colorscale as in Figure~\ref{fig:true_beta}, computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $5.0\%$ (middle), and $1.0\%$ (right).}
+	\caption{(Ice sheet) The log basal friction parameter, with color scale as in Figure~\ref{fig:true_beta}, computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $5.0\%$ (middle), and $1.0\%$ (right).}
 	\label{fig:stokes_reconstructions} 
@@
 Although the method we presented is not applicable to all Hessians, it is applicable to many Hessians of practical interest.
-For these Hessians, the proposed method offers a \emph{data scalable} alternative to conventional low rank approximation, due to the ability to form high rank approximations of an operator using a small number of operator applies, and thus forward and adjoint PDE solves.
+For these Hessians, the proposed method offers a \emph{data scalable} alternative to conventional low rank approximation, due to the ability to form high rank approximations of an operator using a small number of operator applications, and thus forward and adjoint PDE solves.
 

### 2023-07-04 paper/localpsf.tex
working through Omar's edits

@@
 where $\Aop$ is a symmetric positive semi-definite operator that we approximate with the PSF method to form an H-matrix $\mathbf{A}_H$, and $\mathcal{R}$ is a symmetric positive definite operator that may be easily converted to an H-matrix $\mathbf{R}_H$ without using the PSF method.
-For example, in inverse problems $\mathcal{H}$ is the Hessian, $\Aop$ is the data misfit term in the Hessian which is dense and only available matrix-free, and $\mathcal{R}$ is the regularization term, which is typically an elliptic differential operator that becomes a sparse matrix after discretization.
+For example, in inverse problems $\mathcal{H}$ is the Hessian, $\Aop$ is the data misfit term in the Hessian which is dense and available only matrix-free, and $\mathcal{R}$ is the regularization term, which is typically an elliptic differential operator that becomes a sparse matrix after discretization.
 
@@
 		\hline
-		$\fedim$ & $10^2$--$10^9$ & Number of finite element degrees of freedom \\
+		$\fedim$ & $10^3$--$10^9$ & Number of finite element degrees of freedom \\
 		$\nbatch$ & $1$--$25$ & Number of batches \\
@@
 The multiplicative dependence of $\fedim$ with $\nsamplepts$ is undesirable since $\nsamplepts$ may be large, and reducing this cost is possible with more involved computational geometry methods.
-However, from a practical perspective, the cost for choosing sample points is small compared to other parts of the approximation algorithm, hence such improvements are not pursued here.
+However, from a practical perspective, the cost of choosing sample points is small compared to other parts of the approximation algorithm, and hence such improvements are not pursued here.
 
@@
 In summary, computing all the necessary ingredients to evaluate kernel entries of the PSF approximation requires a handful of operator applications (e.g., $6+\nbatch$ operator applications in two dimensions, or $10+\nbatch$ operator applications in three dimensions, with $\nbatch$ typically in the range 1--25), plus comparatively cheap additional overhead costs, most notably performing ellipsoid intersection tests while choosing sample point batches.
-Once these ingredients are computed, no more operator applications are required, and approximate kernel entries can be evaluated fast.
-Constructing the H-matrix from kernel entries requires a number of elementary operations that scales nearly linear in $N$.
+Once these ingredients are computed, no more operator applications are required, and approximate kernel entries can be evaluated rapidly.
+Constructing the H-matrix from kernel entries requires a number of elementary operations that scales polylog linearly in $N$.
 Using the resulting H-matrix to perform further linear algebra operations also scales nearly linearly in $N$, though the details of these costs depend heavily on the type of H-matrix and operation being performed.
@@
 To reconstruct the unknown parameter fields, denoted $q$, the inverse problems are formulated as nonlinear least squares
-optimization problems, whose cost functions consist of a data misfit term
+optimization problems, whose objective functions consist of a data misfit term
 (between the observations and model output) and a bi-Laplacian
@@
 the discretized gradient, Hessian, and Gauss-Newton Hessian of the
-inverse problem cost function, respectively, evaluated at the current Newton iterate.
+inverse problem objective function, respectively, evaluated at the current Newton iterate.
 To ensure positive definiteness of the Hessian we use

### 2023-07-05 paper/localpsf.tex
edits during zoom meeting with Noemi and Tucker

@@
 While PC approximations are based on an assumption of local translation invariance, the method we propose is based on a more general assumption we call ``local mean displacement invariance'' (Section~\ref{sec:local_mean_displacement_invariance} and Figure~\ref{fig:mean_displacement_invariance}).
+This improves the interpolation of the impulse responses.
+Furthermore, in our previous work~\cite{AlgerEtAl19}, we chose point sources optimally via a sequential adaptive procedure.
+However, in that work each point source required a separate operator application, making the previous method expensive when a large number of impulse responses is desired.
+In this paper, we use a new moment method (Section \ref{sec:intromoments}) which permits computation of many impulse responses (e.g., 50) per operator application.
+Finally, the method we propose never evaluates computed impulse responses outside of their domain of definition.
+This eliminates boundary artifacts which plague conventional PC methods.
 
@@
 However, to reduce the required number of operator applications, we first form a PSF approximation of the data misfit Hessian by exploiting locality and non-negative integral kernel properties, then form the H-matrix using classical H-matrix techniques.
-Not all data misfit Hessians satisfy the local non-negative integral kernel properties (for example, the wave inverse problem data misfit Hessian has a substantial amount of negative entries in its integral kernel).
+By using this new two stage approach, we reduce the number of operator applications to a few dozen at most.
+One limitation of our method is that not all data misfit Hessians satisfy the local non-negative integral kernel properties (for example, the wave inverse problem data misfit Hessian has a substantial amount of negative entries in its integral kernel).
 However, many data misfit Hessians of practical interest do satisfy these properties (either exactly or approximately), and the proposed method is targeted at approximating these Hessians.

### 2023-07-05 paper/localpsf.tex
edits to abstract. split into 2 paragraphs. About to combine figures 1 and 2

@@
 Further matrix computations such as matrix vector products, matrix-matrix additions, matrix factorizations, and the solution of linear systems are performed with H-matrix methods.
-We apply the proposed method to approximate the Hessian operator arising in inverse problems governed by partial differential equations (PDEs).
-While for many ill-posed inverse problems the Hessian of the data misfit term exhibits a low rank structure, and hence a low rank approximation is suitable, for many problems of practical interest the numerical rank of this operator is still large.
+
+	We demonstrate the effectiveness of the proposed method by using it to build preconditioners for the Hessian operator arising in two inverse problems governed by partial differential equations (PDEs): inversion for the basal friction coefficient in an ice sheet flow problem and inversion for the initial condition in a diffusive transport problem.
+	While for many ill-posed inverse problems the Hessian of the data misfit term exhibits a low rank structure, and hence a low rank approximation is suitable, for many problems of practical interest the numerical rank of the Hessian is still large.
 This makes conventional matrix-free Hessian approximation methods expensive in terms of the required number of Hessian operator applications (hence PDE solves).
@@
 Hessian impulse responses typically become more local as the numerical rank increases, which makes high rank Hessians perfect candidates for the proposed approach.
-We illustrate the proposed approach with two specific problems: inversion for the basal friction coefficient in an ice sheet flow problem and inversion for the initial condition in a diffusive transport problem.
-To demonstrate the effectiveness of the approximation, we use the proposed approach to build preconditioners to accelerate the solution of the inverse problems.
-We compare the results with classical regularization-based preconditioning and no preconditioning.
-The numerical results demonstrate the superior performance (measured in number of PDE solves) of the inverse solver when using the proposed PSF-based preconditioner.
+%	To demonstrate the effectiveness of the approximation, w
+%	We use the proposed approach to build preconditioners to accelerate the solution of the inverse problems. 
+%	We compare the results with classical regularization-based preconditioning and no preconditioning. 
+%	The numerical results demonstrate the superior performance (measured in number of PDE solves) of the inverse solver when
+	We observe that using the proposed PSF-based preconditioner clusters the spectrum of the preconditioned Hessian near one, yielding roughly 5X-10X reductions in the number of PDE solves required to solve the inverse problems, as compared to classical regularization-based preconditioning and no preconditioning.
 Finally, we present a comprehensive numerical study for the influence of the diffusion parameters and simulation time (for the diffusive transport problem) on the effectiveness of the approximation.
@@
 However, to reduce the required number of operator applications, we first form a PSF approximation of the data misfit Hessian by exploiting locality and non-negative integral kernel properties, then form the H-matrix using classical H-matrix techniques.
-By using this new two stage approach, we reduce the number of operator applications to a few dozen at most.
+By using this two stage approach, we reduce the number of operator applications to a few dozen at most.
 One limitation of our method is that not all data misfit Hessians satisfy the local non-negative integral kernel properties (for example, the wave inverse problem data misfit Hessian has a substantial amount of negative entries in its integral kernel).

### 2023-07-05 paper/localpsf.tex
editing paper. about to convert\bm to \mathbf

@@
 \begin{figure}
-	\begin{subfigure}{0.32\textwidth}
+	\begin{subfigure}{0.49\textwidth}
 		\centering
-		\includegraphics[scale=0.34]{impulse_batch1.png}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
+		\includegraphics[scale=0.4]{impulse_batch1.png}
+		\caption{One impulse response batch}
+		\label{fig:batches_intro}
+	\end{subfigure}
+	\begin{subfigure}{0.49\textwidth}
 		\centering
-		\includegraphics[scale=0.34]{impulse_batch2.png}
-	\end{subfigure}
-	\begin{subfigure}{0.32\textwidth}
-		\centering
-		\includegraphics[scale=0.34]{impulse_batch3.png}
-	\end{subfigure}
-	\caption{Three batches, $\eta_b$, of normalized impulse responses, $\impulseresponse_{x}$, that arise from applying $\Aop$ to weighted sums of scattered point sources (see Section \ref{sec:get_impulse_response}).
+		\includegraphics[scale=0.8]{mean_displacement_invariance.pdf}
+		\caption{Local mean displacement invariance}
+		\label{fig:mean_displacement_invariance}
+	\end{subfigure}
+%	\begin{subfigure}{0.32\textwidth}
+%		\centering
+%		\includegraphics[scale=0.34]{impulse_batch3.png}
+%	\end{subfigure}
+	\caption{(\ref{fig:batches_intro}) One batch, $\eta_b$, of normalized impulse responses, $\impulseresponse_{x}$, that arise from applying $\Aop$ to a weighted sum of scattered point sources (see Section \ref{sec:get_impulse_response}).
 Here $\Aop$ is the ice sheet inverse problem data misfit Gauss-Newton Hessian described in Section~\ref{sec:numerical_results}.
@@
 The large circle is $\partial \Omega$.
+(\ref{fig:mean_displacement_invariance}) Illustration of impulse responses, $\impulseresponse_x$ and $\impulseresponse_{x'}$, that arise from applying $\Aop$ to point sources at two points, $x$ and $x'$.
+The operator $\Aop$ is locally mean displacement invariant (Section \ref{sec:local_mean_displacement_invariance}) if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$.
+Here $\spatialmean(z)$ denotes the mean (center of mass) of $\phi_z$.
 	}
-	\label{fig:batches_intro}
+	\label{fig:batches_and_mdi}
 \end{figure}
@@
 For background on PC and PSF methods, we recommend the following papers: \cite{DenisEtAl15,EscandeWeiss17,GentileCourbinMeylan13}.
-While PC approximations are based on an assumption of local translation invariance, the method we propose is based on a more general assumption we call ``local mean displacement invariance'' (Section~\ref{sec:local_mean_displacement_invariance} and Figure~\ref{fig:mean_displacement_invariance}).
-This improves the interpolation of the impulse responses.
-Furthermore, in our previous work~\cite{AlgerEtAl19}, we chose point sources optimally via a sequential adaptive procedure.
+The proposed method improves upon existing PC methods in the following ways: (1) While PC approximations are based on an assumption of local translation invariance, the method we propose is based on a more general assumption we call ``local mean displacement invariance'' (Section~\ref{sec:local_mean_displacement_invariance} and Figure~\ref{fig:mean_displacement_invariance}), which improves the interpolation of the impulse responses.
+(2) In our previous work~\cite{AlgerEtAl19}, we chose point sources optimally via a sequential adaptive procedure.
 However, in that work each point source required a separate operator application, making the previous method expensive when a large number of impulse responses is desired.
 In this paper, we use a new moment method (Section \ref{sec:intromoments}) which permits computation of many impulse responses (e.g., 50) per operator application.
-Finally, the method we propose never evaluates computed impulse responses outside of their domain of definition.
+(3) The method we propose never evaluates computed impulse responses outside of their domain of definition.
 This eliminates boundary artifacts which plague conventional PC methods.
 
-\begin{figure}
-	\begin{center}
-		\includegraphics[scale=0.75]{mean_displacement_invariance.pdf}
-	\end{center}
-	\caption{This figure illustrates impulse responses, $\impulseresponse_x$ and $\impulseresponse_{x'}$, that arise from applying $\Aop$ to point sources at two points, $x$ and $x'$.
+%\begin{figure}
+%	\begin{center}
+%		\includegraphics[scale=0.75]{mean_displacement_invariance.pdf}
+%	\end{center}
+%	\caption{This figure illustrates impulse responses, $\impulseresponse_x$ and $\impulseresponse_{x'}$, that arise from applying $\Aop$ to point sources at two points, $x$ and $x'$.
 The operator $\Aop$ is locally mean displacement invariant (Section \ref{sec:local_mean_displacement_invariance}) if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$.
 Here $\spatialmean(z)$ denotes the mean (center of mass) of $\phi_z$.}
-	\label{fig:mean_displacement_invariance}
-\end{figure}
+%	\label{fig:mean_displacement_invariance}
+%\end{figure}
 
@@
 and flow of a contaminant~\cite{PetraStadler11}
-(Section~\ref{sec:adv}). 
+(Section~\ref{sec:adv}). These inverse problems are described in Sections~\ref{sec:ice} and~\ref{sec:adv}, respectively.
 To reconstruct the unknown parameter fields, denoted $q$, the inverse problems are formulated as nonlinear least squares
@@
 We use the framework described in this paper to generate Hessian preconditioners.
-We build H-matrix approximations, $\mathbf{A}_H$, of the data misfit Gauss-Newton Hessian (the term in $\bm{H}_\text{gn}$ that arises from the data misfit).
+We build H-matrix approximations, $\mathbf{A}_H$, of the data misfit Gauss-Newton Hessian (the term in $\mathbf{H}_\text{gn}$ that arises from the data misfit).
 The
@@
 standard sparse H-matrix techniques. Then, H-matrix approximations of the Gauss-Newton Hessian, 
-$$\bm{H}_\text{gn}\approx \widetilde{\bm{H}}:=\mathbf{A}_H + \mathbf{R}_H,$$
+$$\mathbf{H}_\text{gn}\approx \widetilde{\mathbf{H}}:=\mathbf{A}_H + \mathbf{R}_H,$$
 are formed by adding $\mathbf{A}_H$ to $\mathbf{R}_H$ using fast H-matrix arithmetic.
-We modify $\widetilde{\bm{H}}$ to be (approximately) symmetric positive semi-definite via the procedure described in Section~\ref{sec:flip_eigs}.
-We factor $\bm{\preconditioner}$
+We modify $\widetilde{\mathbf{H}}$ to be (approximately) symmetric positive semi-definite via the procedure described in Section~\ref{sec:flip_eigs}.
+We factor $\mathbf{\preconditioner}$
 using fast H-matrix methods, then use the factorization as a
 preconditioner.
-We approximate $\bm{H}_\text{gn}$ rather than $\bm{H}$ because $\bm{H}$ more often has negative values in its integral kernel.
-The numerical results show that~$\widetilde{\bm{H}}$ is a good preconditioner for both~$\bm{H}_\text{gn}$ and~$\bm{H}$.
+We approximate $\mathbf{H}_\text{gn}$ rather than $\mathbf{H}$ because $\mathbf{H}$ more often has negative values in its integral kernel.
+The numerical results show that~$\widetilde{\mathbf{H}}$ is a good preconditioner for both~$\mathbf{H}_\text{gn}$ and~$\mathbf{H}$.
 
@@
 		\includegraphics[scale=0.22]{mtrue_withColorBar.png}
-		\caption{$\basalfriction_\text{true}$.}
+		\caption{$\basalfriction_\text{true}$}
 		\label{fig:true_beta}
@@
 		\includegraphics[scale=0.22]{trueVelocity2_glyphs.png}
-		\caption{$\velocity_\text{true}$.}
+		\caption{$\velocity_\text{true}$}
 		\label{fig:stokes_velocity}
@@
 				Iter & 
-				\#CG & \#Stokes & $\|\bm{g}\|$ & 
-				\#CG & \#Stokes & $\|\bm{g}\|$ & 
-				\#CG & \#Stokes & $\|\bm{g}\|$ \\
+				\#CG & \#Stokes & $\|\mathbf{g}\|$ & 
+				\#CG & \#Stokes & $\|\mathbf{g}\|$ & 
+				\#CG & \#Stokes & $\|\mathbf{g}\|$ \\
 				0 &

### 2023-07-05 paper/localpsf.tex
done converting \bm to \mathbf

@@
 We assume that the finite element basis satisfies the Kronecker property, i.e., $\febasis_i(\fepoint_i)=1$ and $\febasis_i(\fepoint_j)=0$ if $i\neq j$.
-For $u_h \in V_h$ we write $\bm{u} \in \mathbb{R}^\nsamplepts_\massmatrix$ to denote the coefficient vector for $u_h$ with respect to the finite element basis, i.e.,
-$
-u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
+For $u_h \in V_h$ we write $\mathbf{u} \in \mathbb{R}^\nsamplepts_\massmatrix$ to denote the coefficient vector for $u_h$ with respect to the finite element basis, i.e.,
+$
+u_h(x) = \sum_{i=1}^\fedim \mathbf{u}_i \febasis_i(x).
 $
@@
 conjugate gradient (PCG) scheme and a globalizing Armijo line search~\cite{NocedalWright99}. The Newton updates,
-$\bm{\searchdir}$, are obtained by solving 
+$\mathbf{\searchdir}$, are obtained by solving 
 %
@@
 	\label{eq:newton_system}
-	\bm{H} \bm{\searchdir} = - \bm{g} \qquad \text{or} \qquad \bm{H}_\text{gn} \bm{\searchdir} = - \bm{g},
+	\mathbf{H} \mathbf{\searchdir} = - \mathbf{g} \qquad \text{or} \qquad \mathbf{H}_\text{gn} \mathbf{\searchdir} = - \mathbf{g},
 \end{equation}
@@
 wherein we choose the initial guess as the discretization of the constant
-function $q_0$. Here $\bm{g}$, $\bm{H}$ and $\bm{H}_\text{gn}$ are
+function $q_0$. Here $\mathbf{g}$, $\mathbf{H}$ and $\mathbf{H}_\text{gn}$ are
 the discretized gradient, Hessian, and Gauss-Newton Hessian of the
@@
 To ensure positive definiteness of the Hessian we use
-$\bm{H}_\text{gn}$ for the first five iterations, and $\bm{H}$ for all
+$\mathbf{H}_\text{gn}$ for the first five iterations, and $\mathbf{H}$ for all
 subsequent iterations. The Newton iterations are terminated when
-$\|\bm{g}\| < 10^{-6} \|\bm{g}_0\|$, where $\bm{g}_0$ is the gradient evaluated at the initial guess. Systems~\eqref{eq:newton_system} are solved inexactly using an inner PCG
+$\|\mathbf{g}\| < 10^{-6} \|\mathbf{g}_0\|$, where $\mathbf{g}_0$ is the gradient evaluated at the initial guess. Systems~\eqref{eq:newton_system} are solved inexactly using an inner PCG
 iteration, which is terminated early based on the Eisenstat-Walker \cite{EisenstatWalker96} and Steihaug \cite{Steihaug83} conditions.
@@
 Preconditioners shown are the proposed PSF method with five batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
-Columns titled \#CG show the number of PCG iterations used to solve the Newton system for $\bm{\widehat{\basalfriction}}$.
-Columns titled $\|\bm{g}\|$ show the $l^2$ norm of the gradient at $\bm{\basalfriction}$.
+Columns titled \#CG show the number of PCG iterations used to solve the Newton system for $\mathbf{\widehat{\basalfriction}}$.
+Columns titled $\|\mathbf{g}\|$ show the $l^2$ norm of the gradient at $\mathbf{\basalfriction}$.
 Columns titled \#Stokes show the total number of Stokes PDE solves performed in each Newton iteration.
@@
     \end{subfigure}
-	\caption{(Ice sheet) Left: convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
-random entries drawn from the standard Gaussian distribution and $\bm{H}$ is evaluated at the solution of the inverse problem.
+	\caption{(Ice sheet) Left: convergence history for solving $\mathbf{H} \mathbf{x} = \mathbf{b}$ using PCG, where $\mathbf{b}$ has i.i.d.
+random entries drawn from the standard Gaussian distribution and $\mathbf{H}$ is evaluated at the solution of the inverse problem.
 	Results in these figures are shown for the PSF-based preconditioners with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
-The preconditioner is constructed using $\bm{H}_\text{gn}$.
-	Right: generalized eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$.
-Here $\bm{H}$ is the Hessian and $\bm{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\bm{H}_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), the regularization Hessian (REG), or the identity matrix (NONE).}
+The preconditioner is constructed using $\mathbf{H}_\text{gn}$.
+	Right: generalized eigenvalues for generalized eigenvalue problem $\mathbf{H} \mathbf{u}_{k}=\lambda_{k} \mathbf{\preconditioner} \mathbf{u}_{k}$.
+Here $\mathbf{H}$ is the Hessian and $\mathbf{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\mathbf{H}_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), the regularization Hessian (REG), or the identity matrix (NONE).}
 	\label{fig:krylov_convergence}
@@
 
-Next, we build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\bm{\basalfriction}$.
+Next, we build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\mathbf{\basalfriction}$.
 We use PCG to solve a linear system with the Hessian as the coefficient operator and a right hand vector with random independent and identically distributed (i.i.d.) entries drawn from the standard Gaussian distribution.
@@
 $
-\bm{H} \mathbf{u} = \lambda \bm{\preconditioner} \mathbf{u}.
-$ 
-The matrix $\bm{\preconditioner}$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations, the regularization Hessian (REG), or the identity matrix (NONE).
+\mathbf{H} \mathbf{u} = \lambda \mathbf{\preconditioner} \mathbf{u}.
+$ 
+The matrix $\mathbf{\preconditioner}$ is one of the PSF (1), PSF (5), or PSF (25) Gauss-Newton Hessian approximations, the regularization Hessian (REG), or the identity matrix (NONE).
 With the PSF-based preconditioners, the generalized eigenvalues cluster near one, with more batches yielding better clustering.
@@
 			\begin{tabular}{c| c c c c c}
-				noise    & \multicolumn{5}{c}{COND$(\bm{\preconditioner}^{-1} \bm{H})$ } \\ \cline{2-6}
+				noise    & \multicolumn{5}{c}{COND$(\mathbf{\preconditioner}^{-1} \mathbf{H})$ } \\ \cline{2-6}
 				level    & REG     &	NONE  & PSF $(1)$ & PSF $(5)$ & PSF $(25)$ \\ \hline 
@@
 		\end{center}
-	\caption{(Ice sheet) Condition number for $\bm{\preconditioner}^{-1} \bm{H}$ for the PSF-based preconditioners with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), no preconditioner (NONE) and regularization (REG).
+	\caption{(Ice sheet) Condition number for $\mathbf{\preconditioner}^{-1} \mathbf{H}$ for the PSF-based preconditioners with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), no preconditioner (NONE) and regularization (REG).
 All operators are evaluated at the soutions of the inverse problems for
@@
 
-In Figure~\ref{fig:adv_krylov_convergence} (left) we show the convergence of PCG for solving $\bm{H}\bm{x}=\bm{b}$, where $\bm{b}$ has i.i.d.
+In Figure~\ref{fig:adv_krylov_convergence} (left) we show the convergence of PCG for solving $\mathbf{H}\mathbf{x}=\mathbf{b}$, where $\mathbf{b}$ has i.i.d.
 random entries drawn from a standard Gaussian distribution.
-The preconditioners used, $\widetilde{\bm{H}}$, are the PSF-based preconditioners with $1$, $5$, or $25$ batches, the regularization Hessian, and the identity matrix (i.e., no preconditioning).
+The preconditioners used, $\widetilde{\mathbf{H}}$, are the PSF-based preconditioners with $1$, $5$, or $25$ batches, the regularization Hessian, and the identity matrix (i.e., no preconditioning).
 The results show that PCG converges fastest with the PSF-based preconditioners, with more batches yielding faster convergence.
-In Figure~\ref{fig:adv_krylov_convergence} (right), we show the eigenvalues for the generalized eigenvalue problem $\bm{H}\bm{u}=\lambda \widetilde{\bm{H}}\bm{u}$, where the $\widetilde{\bm{H}}$ are the preconditioners stated above.
+In Figure~\ref{fig:adv_krylov_convergence} (right), we show the eigenvalues for the generalized eigenvalue problem $\mathbf{H}\mathbf{u}=\lambda \widetilde{\mathbf{H}}\mathbf{u}$, where the $\widetilde{\mathbf{H}}$ are the preconditioners stated above.
 With the PSF-based preconditioners the eigenvalues cluster near one, and more batches yields better clustering.
@@
 	\end{center}
-	\caption{(Diffusive transport) Number of PCG iterations required to solve the Newton linear system to tolerance $||\widehat{\bm{q}}_j - \widehat{\bm{q}}|| < 10^{-6}||\widehat{\bm{q}}||$, where $\widehat{\bm{q}}_j$ is the $j$th iterate, and $\widehat{\bm{q}}$ is the solution of the Newton linear system.
+	\caption{(Diffusive transport) Number of PCG iterations required to solve the Newton linear system to tolerance $||\widehat{\mathbf{q}}_j - \widehat{\mathbf{q}}|| < 10^{-6}||\widehat{\mathbf{q}}||$, where $\widehat{\mathbf{q}}_j$ is the $j$th iterate, and $\widehat{\mathbf{q}}$ is the solution of the Newton linear system.
 Iteration counts are shown for a variety of different diffusion parameters $\kappa$, simulation times $T$, and preconditioners.}
@@
 	\end{subfigure}
-	\caption{(Diffusive transport) Left: convergence history for solving $\bm{H} \bm{x} = \bm{b}$ using PCG, where $\bm{b}$ has i.i.d.
+	\caption{(Diffusive transport) Left: convergence history for solving $\mathbf{H} \mathbf{x} = \mathbf{b}$ using PCG, where $\mathbf{b}$ has i.i.d.
 random entries drawn from the standard Gaussian distribution.
-Right: generalized eigenvalues for generalized eigenvalue problem $\bm{H} \bm{u}_{k}=\lambda_{k} \bm{\preconditioner} \bm{u}_{k}$.
-Here $\bm{H}$ is the Hessian and the preconditioner $\bm{\preconditioner}$ is the PSF approximation for 1, 5, or 25 batches (PSF (1), PSF (5), and PSF (25), respectively), the regularization Hessian (REG), or the identity matrix (NONE).}
+Right: generalized eigenvalues for generalized eigenvalue problem $\mathbf{H} \mathbf{u}_{k}=\lambda_{k} \mathbf{\preconditioner} \mathbf{u}_{k}$.
+Here $\mathbf{H}$ is the Hessian and the preconditioner $\mathbf{\preconditioner}$ is the PSF approximation for 1, 5, or 25 batches (PSF (1), PSF (5), and PSF (25), respectively), the regularization Hessian (REG), or the identity matrix (NONE).}
 	\label{fig:adv_krylov_convergence}

### 2023-07-05 paper/localpsf.tex
more edits to numerical results section. Update Hartland23 reference to published version

@@
 Hessian impulse responses typically become more local as the numerical rank increases, which makes high rank Hessians perfect candidates for the proposed approach.
-%	To demonstrate the effectiveness of the approximation, w
-%	We use the proposed approach to build preconditioners to accelerate the solution of the inverse problems. 
-%	We compare the results with classical regularization-based preconditioning and no preconditioning. 
-%	The numerical results demonstrate the superior performance (measured in number of PDE solves) of the inverse solver when
-	We observe that using the proposed PSF-based preconditioner clusters the spectrum of the preconditioned Hessian near one, yielding roughly 5X-10X reductions in the number of PDE solves required to solve the inverse problems, as compared to classical regularization-based preconditioning and no preconditioning.
+	We observe that using the proposed PSF-based preconditioner clusters the spectrum of the preconditioned Hessian near one, yielding roughly 5X--10X reductions in the number of PDE solves required to solve the inverse problems, as compared to classical regularization-based preconditioning and no preconditioning.
 Finally, we present a comprehensive numerical study for the influence of the diffusion parameters and simulation time (for the diffusive transport problem) on the effectiveness of the approximation.
@@
 
-Examples of inverse problems with high rank data misfit Hessians include large-scale ice sheet inverse problems \cite{HartlandEtAl23,IsaacEtAl15}, advection dominated advection-diffusion inverse problems~\cite{AkcelikBirosDraganescuEtAl05}\cite[Chapter 5]{Flath13}, high frequency wave propagation inverse problems~\cite{BuiEtAl13}, inverse problems governed by high Reynolds number flows, and more generally, all inverse problems in which the observations highly inform the parameter.
+Examples of inverse problems with high rank data misfit Hessians include large-scale ice sheet inverse problems \cite{Hartland_2023,IsaacEtAl15}, advection dominated advection-diffusion inverse problems~\cite{AkcelikBirosDraganescuEtAl05}\cite[Chapter 5]{Flath13}, high frequency wave propagation inverse problems~\cite{BuiEtAl13}, inverse problems governed by high Reynolds number flows, and more generally, all inverse problems in which the observations highly inform the parameter.
 The eigenvalues of the data misfit Hessian characterize how informative the data are about components of the parameter in the corresponding eigenvector directions,
@@
 Recently there have been improvements in matrix-free H-matrix construction methods in which an operator is applied to structured random vectors, and the response of the operator to those random vectors is processed to construct an H-matrix approximation~\cite{LevittMartinsson22,LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}.
-These methods (which we do not use here) have been used to approximate Hessians in PDE constrained inverse problems~\cite{AmbartsumyanEtAl20,HartlandEtAl23}. 
+These methods (which we do not use here) have been used to approximate Hessians in PDE constrained inverse problems~\cite{AmbartsumyanEtAl20,Hartland_2023}. 
 Although these methods are promising,
@@
 Here, $\basalfriction$ is the unknown logarithmic
-basal friction field (large $\basalfriction$ corresponds to high resistance to sliding) defined on the surface $\Gamma_b$.
-The quantity $\bodyforce$ is the body force density due to gravity, $\stokesrobincoeff=10^6$ is a Robin boundary condition constant,  $\normalvec$ is the outward normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary.
-We employ Glen's flow law~\cite{Glen1955}, $\stress(v, p)= 2\eta
+basal friction field (large $\basalfriction$ corresponds to large resistance to sliding) defined on the surface $\Gamma_b$.
+The quantity $\bodyforce$ is the body force density due to gravity, $\stokesrobincoeff=10^6$ is a Robin boundary condition constant,  $\normalvec$ is the outward unit normal and $\tangentop$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary.
+We employ a Newtonian constitutive law, $\stress(v, p)= 2\eta
 \dot{\strain}(v) -I\pressure$,
-which is a constitutive law for ice that relates the stress
-tensor, $\stress$, to the strain rate tensor,
+where $\stress$ is the stress tensor and
 $\dot{\strain}(v)= \frac{1}{2}\left(
-\nabla\velocity+\nabla\velocity^{\top} \right)$.
-Here $\eta$ is the effective viscosity and $I$ is the unit matrix.
-Glen's exponent has been chosen to be one, which provides for a linear Stokes model.
+\nabla\velocity+\nabla\velocity^{\top} \right)$ is the strain rate tensor~\cite{Glen1955}. Here $\eta$ is the viscosity and $I$ is the identity operator. 
+%Glen's exponent has been chosen to be one, which provides for a linear Stokes model. 
 Note that while the PDE is linear, the parameter-to-solution map, $\basalfriction \mapsto (\velocity, \pressure)$, is nonlinear.
@@
 Note that $\Gamma_b$ is a two-dimensional surface embedded in three dimensions due to the mountain topography.
-We also generate a flattened version of $\Gamma_b$, denoted by $\Omega \subset \mathbb{R}^2$, by ignoring the height coordinate.
+We also generate a flattened version of $\Gamma_b$, denoted by $\Omega \subset \mathbb{R}^2$, by ignoring the height coordinate\footnote{The proposed PSF method requires translating impulse responses.
+This requires either a flat domain, or a notion of local parallel transport.}.
 The parameter $\basalfriction$ is viewed as a function on $\Gamma_b$ for the purpose of solving the Stokes equations, and as a function on $\Omega$ for the purpose of building Hessian approximations and defining the regularization.
@@
 The true basal friction
-coefficient and velocity fields, which are obtained by
+coefficient and resulting velocity fields, which are obtained by
 solving~\eqref{Stokeseqn}, are shown in Figure~\ref{fig:true_beta_u}.
@@
 We compare the proposed method with the most commonly used existing preconditioners: no preconditioning (NONE), and preconditioning by the regularization term in the Hessian (REG).
-The results show that using PSF (5) reduces the total number of Stokes PDE solves substantially compared to regularization preconditioning and no preconditioning.
+The results show that using PSF (5) reduces the total number of Stokes PDE solves to 70, as compared to 908 for regularization preconditioning and 308 for no preconditioning.
 In Figure~\ref{fig:stokes_reconstructions} we show reconstructions for $1\%$, $5\%$, and $25\%$ noise.
@@
 Note that the condition number using PSF-based preconditioners is extremely small (ranging between 1 and 10) and relatively stable over this range of noise levels.
-As expected, PSF (25) outperforms than PSF (5), which outperforms than PSF (1).
-All PSF-based preconditioners outperform no preconditioning and regularization preconditioning for all noise levels.
+As expected, PSF (25) outperforms PSF (5), which outperforms PSF (1).
+All PSF-based preconditioners outperform no preconditioning and regularization preconditioning by several orders of magnitude for all noise levels.
 
@@
 
-\subsection{Example 2: Inversion for the initial condition in an diffusive transport problem}
+\subsection{Example 2: Inversion for the initial condition in a diffusive transport problem}
 \label{sec:adv}
@@
 Here we consider a time-dependent advection-diffusion equation in
-which we seek to infer the unknown spatially varying initial condition, $\ipar$, from noisy observations of the state  
+which we seek to infer the unknown spatially varying initial condition, $\ipar$, from noisy observation of the full state  
 at a final time, $T$. This PDE
@@
 parameter-to-observable map $\iFF$ in this case maps an initial
-condition $\ipar \in L^2(\advdomain)$ to observations of the concentration at a final time, $\concentration(x, T)$,  through solution of the advection-diffusion
+condition $\ipar \in L^2(\advdomain)$ to the concentration field at a final time, $\concentration(x, T)$,  through solution of the advection-diffusion
 equation given by
@@
 In Table~\ref{tab:adv_convergence} we show the number of PCG iterations, $j$, required to solve the Newton linear system to a relative error tolerance of $\|\widehat{\mathbf{q}} - \widehat{\mathbf{q}}_j\| < 10^{-6} \|\widehat{\mathbf{q}}\|$.
-The solution of the Newton system that we compare to, $\widehat{\mathbf{q}}$, is found via another PCG iteration with a relative residual tolerance of $10^{-11}$.
+The solution of the Newton system to which we compare, $\widehat{\mathbf{q}}$, is found via another PCG iteration with a relative residual tolerance of $10^{-11}$.
 We show results for $T$ ranging from $0.5$ to $2.0$ and $\kappa$ ranging from $10^{-4}$ to $10^{-3}$ using the PSF-based preconditioners with $1$, $5$, and $25$ batches, regularization preconditioning, and no preconditioning.

### 2023-07-06 paper/localpsf.tex
more edits. Todo: (1) figure out correct quantitative statements to make (5X-10X may not apply to advection diffusion). (2) shorten by about 1 page

@@
     \end{subfigure}
-	\caption{(Ice sheet) Left: convergence history for solving $\mathbf{H} \mathbf{x} = \mathbf{b}$ using PCG, where $\mathbf{b}$ has i.i.d.
-random entries drawn from the standard Gaussian distribution and $\mathbf{H}$ is evaluated at the solution of the inverse problem.
+	\caption{(Ice sheet) Left: convergence history for solving $\mathbf{H} \mathbf{x} = \mathbf{b}$ using PCG, where $\mathbf{b}$ has i.i.d.\ random entries drawn from the standard Gaussian distribution and $\mathbf{H}$ is evaluated at the solution of the inverse problem.
 	Results in these figures are shown for the PSF-based preconditioners with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
@@
 We show results for $T$ ranging from $0.5$ to $2.0$ and $\kappa$ ranging from $10^{-4}$ to $10^{-3}$ using the PSF-based preconditioners with $1$, $5$, and $25$ batches, regularization preconditioning, and no preconditioning.
-The results show that PSF-based preconditioning outperforms regularization preconditioning and no preconditioning, with more batches yielding better results.
-
-In Figure~\ref{fig:adv_krylov_convergence} (left) we show the convergence of PCG for solving $\mathbf{H}\mathbf{x}=\mathbf{b}$, where $\mathbf{b}$ has i.i.d.
-random entries drawn from a standard Gaussian distribution.
+The results show that PSF-based preconditioning outperforms regularization preconditioning and no preconditioning.
+Adding more batches yields better results, and the impact of adding more batches is more pronounced here than in the ice sheet example.
+For example, in the mid-range values $T=1.0$ and $\kappa=3.2e{-4}$, PSF(1), PSF(5), and PSF(25) require roughly 1.3X, 2.3X, 5.1X fewer PCG iterations, respectively, as compared to no preconditioning (and greater improvements as compared to regularization preconditioning).
+The PSF preconditioners perform best in the high rank regime where $T$ is small, which makes sense given that short simulation times yield more localized impulse responses.
+For example, for $\kappa=3.2e{-4}$, and $T=0.5$, $T=1.0$, and $T=2.0$, using the PSF(5) preconditioner yields 140, 202, and 379 iterations, respectively.
+The performance of the PSF preconditioners as a function of $\kappa$ is mixed.
+Reducing $\kappa$ makes the impulse responses thinner and hence easier to fit in batches, but also increases the complexity of the impulse response shapes, which may reduce the accuracy of the RBF interpolation.
+
+In Figure~\ref{fig:adv_krylov_convergence} (left) we show the convergence of PCG for solving $\mathbf{H}\mathbf{x}=\mathbf{b}$, where $\mathbf{b}$ has i.i.d.\ random entries drawn from a standard Gaussian distribution.
 The preconditioners used, $\widetilde{\mathbf{H}}$, are the PSF-based preconditioners with $1$, $5$, or $25$ batches, the regularization Hessian, and the identity matrix (i.e., no preconditioning).
@@
 	\end{subfigure}
-	\caption{(Diffusive transport) Left: convergence history for solving $\mathbf{H} \mathbf{x} = \mathbf{b}$ using PCG, where $\mathbf{b}$ has i.i.d.
-random entries drawn from the standard Gaussian distribution.
+	\caption{(Diffusive transport) Left: convergence history for solving $\mathbf{H} \mathbf{x} = \mathbf{b}$ using PCG, where $\mathbf{b}$ has i.i.d.\ random entries drawn from the standard Gaussian distribution.
 Right: generalized eigenvalues for generalized eigenvalue problem $\mathbf{H} \mathbf{u}_{k}=\lambda_{k} \mathbf{\preconditioner} \mathbf{u}_{k}$.
-Here $\mathbf{H}$ is the Hessian and the preconditioner $\mathbf{\preconditioner}$ is the PSF approximation for 1, 5, or 25 batches (PSF (1), PSF (5), and PSF (25), respectively), the regularization Hessian (REG), or the identity matrix (NONE).}
+Here $\mathbf{H}$ is the Hessian and the preconditioner, $\mathbf{\preconditioner}$, is the PSF approximation for 1, 5, or 25 batches (PSF (1), PSF (5), and PSF (25), respectively), the regularization Hessian (REG), or the identity matrix (NONE).}
 	\label{fig:adv_krylov_convergence}
@@
 We presented an efficient matrix-free PSF method for approximating operators with locally supported non-negative integral kernels.
-The method only requires access to the operator via application of the operator to a small number of vectors.
+The method requires access to the operator only via application of the operator to a small number of vectors.
 The idea of the method is to compute batches of impulse responses by applying the operator to Dirac combs of scattered point sources, then interpolate these impulse responses to approximate entries of the operator's integral kernel.
@@
 We applied the method to approximate the Gauss-Newton Hessians in an ice sheet flow inverse problem governed by the Stokes PDE, and a diffusive transport inverse problem governed by an advection-diffusion PDE.
-We saw that the proposed approximation method substantially outperforms existing Hessian approximation methods.
+We saw that preconditioners based on the proposed approximation cluster the eigenvalues of the preconditioned Hessian near one, and reduce the required number of PDE solves by factors of roughly 5X--10X.
 Although the method we presented is not applicable to all Hessians, it is applicable to many Hessians of practical interest.

### 2023-07-06 paper/localpsf.tex
Made paper almost achieve the page limit by making figures and tables smaller

@@
 		\centering
-		\includegraphics[scale=0.4]{impulse_batch1.png}
+		\includegraphics[scale=0.3]{impulse_batch1.png}
 		\caption{One impulse response batch}
@@
 		\centering
-		\includegraphics[scale=0.8]{mean_displacement_invariance.pdf}
+		\includegraphics[scale=0.7]{mean_displacement_invariance.pdf}
 		\caption{Local mean displacement invariance}
@@
 standard sparse H-matrix techniques. Then, H-matrix approximations of the Gauss-Newton Hessian, 
-$$\mathbf{H}_\text{gn}\approx \widetilde{\mathbf{H}}:=\mathbf{A}_H + \mathbf{R}_H,$$
+$\mathbf{H}_\text{gn}\approx \widetilde{\mathbf{H}}:=\mathbf{A}_H + \mathbf{R}_H$,
 are formed by adding $\mathbf{A}_H$ to $\mathbf{R}_H$ using fast H-matrix arithmetic.
@@
     \begin{center} 
-	\includegraphics[scale=0.23]{meshHeight_view2_edges.png}
+	\includegraphics[scale=0.2]{meshHeight_view2_edges.png}
 	\end{center}
@@
 	\begin{subfigure}{0.49\textwidth}
-		\includegraphics[scale=0.22]{mtrue_withColorBar.png}
+		\includegraphics[scale=0.2]{mtrue_withColorBar.png}
 		\caption{$\basalfriction_\text{true}$}
@@
 	\begin{subfigure}{0.49\textwidth}
-		\includegraphics[scale=0.22]{trueVelocity2_glyphs.png}
+		\includegraphics[scale=0.2]{trueVelocity2_glyphs.png}
 		\caption{$\velocity_\text{true}$}
@@
 	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[scale=0.20]{mstar_0.25noise_cropped.png}
+		\includegraphics[scale=0.17]{mstar_0.25noise_cropped.png}
 	\end{subfigure}
 	\begin{subfigure}{0.32\textwidth}
-		\includegraphics[scale=0.20]{mstar_0.05noise_cropped.png}
+		\includegraphics[scale=0.17]{mstar_0.05noise_cropped.png}
 	\end{subfigure}
 	\begin{subfigure}{0.32\textwidth}
-	\includegraphics[scale=0.20]{mstar_0.01noise_cropped.png}
+	\includegraphics[scale=0.17]{mstar_0.01noise_cropped.png}
     \end{subfigure}
@@
 \begin{table}
-	{\small
+	{\footnotesize
 		\begin{center}
@@
 We build the PSF (5) preconditioner in the third Gauss-Newton iteration, and reuse it for all subsequent Gauss-Newton and Newton iterations.
-No preconditioning is used in the Gauss-Newton iterations before the PSF (5) preconditioner is built.
+No preconditioning is used in the iterations before the PSF (5) preconditioner is built.
 We compare the proposed method with the most commonly used existing preconditioners: no preconditioning (NONE), and preconditioning by the regularization term in the Hessian (REG).
@@
 		\centering
-		\includegraphics[scale=1.0]{stokes_pcg_convergence.pdf}
+		\includegraphics[scale=0.9]{stokes_pcg_convergence.pdf}
 	\end{subfigure}
@@
 	\begin{center}
-		\includegraphics[scale=1.0]{stokes_geigs_figure.pdf}
+		\includegraphics[scale=0.9]{stokes_geigs_figure.pdf}
 	\end{center}
@@
 	\begin{center}
+			{\footnotesize
 			\begingroup
@@
 			\endgroup
+			}
 		\end{center}
@@
 		\begin{center}
-		\includegraphics[scale=0.14]{true_initial_condition.png}
+		\includegraphics[scale=0.12]{true_initial_condition.png}
 	\end{center}
@@
 		\begin{center}
-		\includegraphics[scale=0.14]{wind_velocity_colorscheme3.png}
+		\includegraphics[scale=0.12]{wind_velocity_colorscheme3.png}
 		\end{center}
@@
 		\begin{center}
-		\includegraphics[scale=0.14]{noisy_observations_T1.0_K3.2e-4.png}
+		\includegraphics[scale=0.12]{noisy_observations_T1.0_K3.2e-4.png}
 		\end{center}
@@
 		\begin{center}
-		\includegraphics[scale=0.14]{reconstructed_initial_condition_T1.0_K3.2e-4.png}
+		\includegraphics[scale=0.12]{reconstructed_initial_condition_T1.0_K3.2e-4.png}
 		\end{center}
@@
 	\begin{center}
+		{\footnotesize
 		\begingroup
@@
 		\endgroup
+		}
 	\end{center}
@@
 			& $T=2.0$ \\
-			$\kappa=1.0$e${-4}$  & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-4.png} 
+			$\kappa=1.0$e${-4}$  & \includegraphics[scale=0.12, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-4.png} 
 %			& \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-4.png}
-			& \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-4.png} \\
+			& \includegraphics[scale=0.12, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-4.png} \\
 %			$\kappa=3.2$e${-4}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K3.2e-4.png} \\
-			$\kappa=1.0$e${-3}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-3.png} 
+			$\kappa=1.0$e${-3}$ & \includegraphics[scale=0.12, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-3.png} 
 %			& \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-3.png} 
-			& \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-3.png}
+			& \includegraphics[scale=0.12, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-3.png}
 		\end{tabular}
@@
 		\centering
-		\includegraphics[scale=1.0]{adv_pcg_convergence.pdf}
+		\includegraphics[scale=0.9]{adv_pcg_convergence.pdf}
 	\end{subfigure}
@@
 		\centering
-		\includegraphics[scale=1.0]{adv_geigs_figure.pdf}
+		\includegraphics[scale=0.9]{adv_geigs_figure.pdf}
 	\end{subfigure}

### 2023-07-06 paper/localpsf.tex
removing less important references to shorten paper

@@
 Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, and $L^2(\Omega)'$ is the space of real-valued continuous linear functionals on $L^2(\Omega)$.
-Such operators appear, for instance, as Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{AntilEtAl18,BorziSchulz11,Delosreyes15,HinzePinnauUlbrichUlbrich08}, Schur complements in Schur complement methods for solving partial differential equations (PDEs) and Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps)~\cite{ChanMathew94,SmithBjorstadGropp96,ToselliWidlund04}, covariance operators in spatial statistics~\cite{ChenStein21,GentonKeyesTurkiyyah18,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}.
+Such operators appear, for instance, as Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{BorziSchulz11,Delosreyes15,HinzePinnauUlbrichUlbrich08},
+%AntilEtAl18
+Schur complements in Schur complement methods for solving partial differential equations (PDEs) and Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps)~\cite{ChanMathew94,SmithBjorstadGropp96,ToselliWidlund04}, covariance operators in spatial statistics~\cite{ChenStein21,GentonKeyesTurkiyyah18,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}.
 Here ``matrix-free'' means that we may apply $\Aop$ and its transpose\footnote{Recall that $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for operations with $\Aop^T$.
@@
 Here, we use the lumped mass matrix in which the $i$th diagonal entry of the lumped mass matrix is the sum of all entries in the $i$th row of the mass matrix.
-Other mass lumping techniques may be used~\cite{Hughes12}.
+Other mass lumping techniques may be used.
+%~\cite{Hughes12}.
 
@@
 Approximation~\eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close.
-Operators satisfying~\eqref{eq:translate1} can be well approximated via low rank CUR approximation~\cite{MahoneyDrineas09}.
+Operators satisfying~\eqref{eq:translate1} can be well approximated via low rank CUR approximation.
+%~\cite{MahoneyDrineas09}.
 However, the required rank in the low rank approximation can be large, which makes algorithms based on~\eqref{eq:translate1} expensive.

### 2023-07-06 paper/localpsf.tex
wordsmithed 5x reduction stuff

@@
 Hessian impulse responses typically become more local as the numerical rank increases, which makes high rank Hessians perfect candidates for the proposed approach.
-	We observe that using the proposed PSF-based preconditioner clusters the spectrum of the preconditioned Hessian near one, yielding roughly 5X--10X reductions in the number of PDE solves required to solve the inverse problems, as compared to classical regularization-based preconditioning and no preconditioning.
+	We observe that using the proposed PSF-based preconditioner clusters the spectrum of the preconditioned Hessian near one, yielding roughly 5X reductions in the number of PDE solves required to solve the inverse problems, as compared to classical regularization-based preconditioning and no preconditioning.
 Finally, we present a comprehensive numerical study for the influence of the diffusion parameters and simulation time (for the diffusive transport problem) on the effectiveness of the approximation.
@@
 The multiplicative dependence of $\fedim$ with $\nsamplepts$ is undesirable since $\nsamplepts$ may be large, and reducing this cost is possible with more involved computational geometry methods.
-However, from a practical perspective, the cost of choosing sample points is small compared to other parts of the approximation algorithm, and hence such improvements are not pursued here.
+However, from a practical perspective, the cost of choosing sample points is small compared to other parts of the algorithm, and hence such improvements are not pursued here.
 
@@
 We show results for $T$ ranging from $0.5$ to $2.0$ and $\kappa$ ranging from $10^{-4}$ to $10^{-3}$ using the PSF-based preconditioners with $1$, $5$, and $25$ batches, regularization preconditioning, and no preconditioning.
-The results show that PSF-based preconditioning outperforms regularization preconditioning and no preconditioning.
+The results show that PSF-based preconditioning outperforms regularization preconditioning and no preconditioning in all cases except one.
+The exception is $T=2.0$ and $\kappa=1.0e{-3}$, in which PSF(1) performs slightly worse than regularization preconditioning but better than no preconditioning.
 Adding more batches yields better results, and the impact of adding more batches is more pronounced here than in the ice sheet example.
-For example, in the mid-range values $T=1.0$ and $\kappa=3.2e{-4}$, PSF(1), PSF(5), and PSF(25) require roughly 1.3X, 2.3X, 5.1X fewer PCG iterations, respectively, as compared to no preconditioning (and greater improvements as compared to regularization preconditioning).
+For example, in the mid-range values $T=1.0$ and $\kappa=3.2e{-4}$, PSF(1), PSF(5), and PSF(25) require 1.3X, 2.3X, 5.1X fewer PCG iterations, respectively, as compared to no preconditioning, and greater improvements as compared to regularization preconditioning.
 The PSF preconditioners perform best in the high rank regime where $T$ is small, which makes sense given that short simulation times yield more localized impulse responses.
-For example, for $\kappa=3.2e{-4}$, and $T=0.5$, $T=1.0$, and $T=2.0$, using the PSF(5) preconditioner yields 140, 202, and 379 iterations, respectively.
-The performance of the PSF preconditioners as a function of $\kappa$ is mixed.
+For example, for $\kappa=3.2e{-4}$ using the PSF(5) preconditioner yields 140, 202, and 379 iterations for $T=0.5$, $1.0$, and $2.0$, respectively.
+The performance of the PSF preconditioners as a function of $\kappa$ does not have as clear of a trend.
 Reducing $\kappa$ makes the impulse responses thinner and hence easier to fit in batches, but also increases the complexity of the impulse response shapes, which may reduce the accuracy of the RBF interpolation.
+The greatest improvements are seen for $T=0.5$ and $\kappa=1e{-3}$, for which PSF(25) requires roughly 10X fewer PCG iterations than no preconditioning, and roughly 20X fewer iterations than regularization preconditioning.
 
@@
 We applied the method to approximate the Gauss-Newton Hessians in an ice sheet flow inverse problem governed by the Stokes PDE, and a diffusive transport inverse problem governed by an advection-diffusion PDE.
-We saw that preconditioners based on the proposed approximation cluster the eigenvalues of the preconditioned Hessian near one, and reduce the required number of PDE solves by factors of roughly 5X--10X.
+We saw that preconditioners based on the proposed approximation cluster the eigenvalues of the preconditioned Hessian near one, and reduce the required number of PDE solves by a factor of roughly 5X.
 Although the method we presented is not applicable to all Hessians, it is applicable to many Hessians of practical interest.

### 2023-07-06 paper/localpsf.tex
5X -> five times

@@
 Hessian impulse responses typically become more local as the numerical rank increases, which makes high rank Hessians perfect candidates for the proposed approach.
-	We observe that using the proposed PSF-based preconditioner clusters the spectrum of the preconditioned Hessian near one, yielding roughly 5X reductions in the number of PDE solves required to solve the inverse problems, as compared to classical regularization-based preconditioning and no preconditioning.
+	We observe that using the proposed PSF-based preconditioner clusters the spectrum of the preconditioned Hessian near one, and allows us to solve the inverse problems using roughly five times fewer PDE solves, as compared to classical regularization-based preconditioning and no preconditioning.
 Finally, we present a comprehensive numerical study for the influence of the diffusion parameters and simulation time (for the diffusive transport problem) on the effectiveness of the approximation.
@@
 Adding more batches yields better results, and the impact of adding more batches is more pronounced here than in the ice sheet example.
-For example, in the mid-range values $T=1.0$ and $\kappa=3.2e{-4}$, PSF(1), PSF(5), and PSF(25) require 1.3X, 2.3X, 5.1X fewer PCG iterations, respectively, as compared to no preconditioning, and greater improvements as compared to regularization preconditioning.
+For example, in the mid-range values $T=1.0$ and $\kappa=3.2e{-4}$, PSF(1), PSF(5), and PSF(25) require 1.3, 2.3, 5.1 times fewer PCG iterations, respectively, as compared to no preconditioning, and greater improvements as compared to regularization preconditioning.
 The PSF preconditioners perform best in the high rank regime where $T$ is small, which makes sense given that short simulation times yield more localized impulse responses.
@@
 Reducing $\kappa$ makes the impulse responses thinner and hence easier to fit in batches, but also increases the complexity of the impulse response shapes, which may reduce the accuracy of the RBF interpolation.
-The greatest improvements are seen for $T=0.5$ and $\kappa=1e{-3}$, for which PSF(25) requires roughly 10X fewer PCG iterations than no preconditioning, and roughly 20X fewer iterations than regularization preconditioning.
+The greatest improvements are seen for $T=0.5$ and $\kappa=1e{-3}$, for which PSF(25) requires roughly 10 times fewer PCG iterations than no preconditioning, and roughly 20 times fewer iterations than regularization preconditioning.
 
@@
 We applied the method to approximate the Gauss-Newton Hessians in an ice sheet flow inverse problem governed by the Stokes PDE, and a diffusive transport inverse problem governed by an advection-diffusion PDE.
-We saw that preconditioners based on the proposed approximation cluster the eigenvalues of the preconditioned Hessian near one, and reduce the required number of PDE solves by a factor of roughly 5X.
+We saw that preconditioners based on the proposed approximation cluster the eigenvalues of the preconditioned Hessian near one, and allow us to solve the inverse problems using roughly five times fewer PDE solves.
 Although the method we presented is not applicable to all Hessians, it is applicable to many Hessians of practical interest.

### 2023-07-06 paper/localpsf.tex
about to update abstract

@@
 \begin{keywords}
- 	data scalability, Hessian, hierarchical matrix, matrix-free, operator approximation, PDE constrained inverse problems, point spread function, product convolution, local translation invariance, impulse response, moment methods, high-rank
+ 	data scalability, Hessian, hierarchical matrix, high-rank, impulse response, local translation invariance, matrix-free, moment methods, operator approximation, PDE constrained inverse problems, point spread function, product convolution
 \end{keywords}
@@
 This Gaussian distribution can be used directly as an approximation of the posterior, or it can be used as a proposal for Markov chain Monte-Carlo methods for drawing samples from the posterior.
-For instance, see~\cite{KimEtAl21,PetraEtAl14} and the references therein.
+For instance, see~\cite{KimEtAl23,PetraEtAl14} and the references therein.
 
@@
 and flow of a contaminant~\cite{PetraStadler11}
-(Section~\ref{sec:adv}). These inverse problems are described in Sections~\ref{sec:ice} and~\ref{sec:adv}, respectively.
-To reconstruct the unknown parameter fields, denoted $q$, the inverse problems are formulated as nonlinear least squares
+(Section~\ref{sec:adv}). These inverse problems are described in detail in their respective sections.
+In both cases, to reconstruct the unknown parameter fields, denoted $q$, the inverse problems are formulated as nonlinear least squares
 optimization problems, whose objective functions consist of a data misfit term

### 2023-07-06 numerical_examples/advection_diffusion/check_eigs_and_pcg_randn_results.py

@@ -1,0 +1,26 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import scipy.sparse.linalg as spla
+
+geigs_psf = np.load('data/geigs_psf.npy')
+geigs_none = np.load('data/geigs_none.npy')
+geigs_reg = np.load('data/geigs_reg.npy')
+
+gg_psf1 = geigs_psf[0,0,2,1,:]
+gg_psf5 = geigs_psf[1,0,2,1,:]
+gg_psf25 = geigs_psf[2,0,2,1,:]
+gg_n = geigs_none[0,2,1,:]
+gg_r = geigs_reg[0,2,1,:]
+
+plt.figure()
+plt.semilogy(gg_psf1)
+plt.semilogy(gg_psf5)
+plt.semilogy(gg_psf25)
+plt.semilogy(gg_n)
+plt.semilogy(gg_r)
+plt.legend(['psf1', 'psf5', 'psf25', 'none', 'reg'])
+plt.title('generalized eigenvalues')
+
+N = len(gg_n)
+b = np.random.randn(N)
+

### 2023-07-06 paper/localpsf.tex
edits to abstract and discussion of numerical results in abstract, numerical results, and conclusion

@@
 	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels.
-The method computes impulse responses of the operator at a collection of scattered points, and interpolates translated and scaled versions of these impulse responses to approximate entries of the integral kernel.
-By applying the operator to Dirac combs associated with batches of point sources, we compute a large number of impulse responses using a small number of operator applications.
-We solve an ellipsoid packing problem to select as many point sources as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap.
-Support ellipsoids are estimated a-priori based on impulse response moments, which are computed by applying the operator to a small number of polynomial functions.
-The interpolation uses a local mean displacement invariance approximation, which generalizes classical local translation invariance.
+The method computes impulse responses of the operator at scattered points, and interpolates these impulse responses to approximate entries of the integral kernel.
+To compute impulse responses efficiently, we apply the operator to Dirac combs associated with batches of point sources, which are chosen by solving an ellipsoid packing problem.
 The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator.
-Further matrix computations such as matrix vector products, matrix-matrix additions, matrix factorizations, and the solution of linear systems are performed with H-matrix methods.
-
-	We demonstrate the effectiveness of the proposed method by using it to build preconditioners for the Hessian operator arising in two inverse problems governed by partial differential equations (PDEs): inversion for the basal friction coefficient in an ice sheet flow problem and inversion for the initial condition in a diffusive transport problem.
+Further matrix computations are then performed with fast H-matrix methods.
+We demonstrate the effectiveness of this PSF-based method by using it to build preconditioners for the Hessian operator arising in two inverse problems governed by partial differential equations (PDEs): inversion for the basal friction coefficient in an ice sheet flow problem and for the initial condition in a diffusive transport problem.
 While for many ill-posed inverse problems the Hessian of the data misfit term exhibits a low rank structure, and hence a low rank approximation is suitable, for many problems of practical interest the numerical rank of the Hessian is still large.
-This makes conventional matrix-free Hessian approximation methods expensive in terms of the required number of Hessian operator applications (hence PDE solves).
-In contrast, the performance of the proposed method depends on the locality and shapes of the impulse responses rather than the numerical rank.
-Hessian impulse responses typically become more local as the numerical rank increases, which makes high rank Hessians perfect candidates for the proposed approach.
-	We observe that using the proposed PSF-based preconditioner clusters the spectrum of the preconditioned Hessian near one, and allows us to solve the inverse problems using roughly five times fewer PDE solves, as compared to classical regularization-based preconditioning and no preconditioning.
-Finally, we present a comprehensive numerical study for the influence of the diffusion parameters and simulation time (for the diffusive transport problem) on the effectiveness of the approximation.
-These parameters have an effect on the shape of the impulse responses and on the rank of the Hessian.
-The results reveal that PSF-based preconditioners are able to form good approximations of high-rank Hessians using a small number of operator applications.
+The Hessian impulse responses on the other hand typically become more local as the numerical rank increases, which benefits the PSF-based method.
+Numerical results reveal that the PSF-based preconditioner clusters the spectrum of the preconditioned Hessian near one, yielding roughly $5\times$--$10\times$ reductions in the required number of PDE solves, as compared to classical regularization-based preconditioning and no preconditioning.
+We also present a comprehensive numerical study for the influence of various parameters (that control the shape of the impulse responses and the rank of the Hessian) on the effectiveness of the approximation.
+The results show that the PSF-based preconditioners are able to form good approximations of high-rank Hessians using only a small number of operator applications.
 \end{abstract}
@@
 We compare the proposed method with the most commonly used existing preconditioners: no preconditioning (NONE), and preconditioning by the regularization term in the Hessian (REG).
-The results show that using PSF (5) reduces the total number of Stokes PDE solves to 70, as compared to 908 for regularization preconditioning and 308 for no preconditioning.
+The results show that using PSF (5) reduces the total number of Stokes PDE solves to 70, as compared to 908 for regularization preconditioning and 308 for no preconditioning, which corresponds to a $5\times$--$10\times$ speedup.
+For problems with a larger physical domain and correspondingly more observations, such as continental scale ice sheet inversion, the speedup will be even greater.
+This is because the rank of the data misfit Hessian will increase, while the locality of the impulse responses will remain the same.
 In Figure~\ref{fig:stokes_reconstructions} we show reconstructions for $1\%$, $5\%$, and $25\%$ noise.
@@
 The results show that PSF-based preconditioning outperforms regularization preconditioning and no preconditioning in all cases except one.
-The exception is $T=2.0$ and $\kappa=1.0e{-3}$, in which PSF(1) performs slightly worse than regularization preconditioning but better than no preconditioning.
+The exception is $T=2.0$ and $\kappa=1.0$e${-3}$, in which PSF(1) performs slightly worse than regularization preconditioning but better than no preconditioning.
 Adding more batches yields better results, and the impact of adding more batches is more pronounced here than in the ice sheet example.
-For example, in the mid-range values $T=1.0$ and $\kappa=3.2e{-4}$, PSF(1), PSF(5), and PSF(25) require 1.3, 2.3, 5.1 times fewer PCG iterations, respectively, as compared to no preconditioning, and greater improvements as compared to regularization preconditioning.
-The PSF preconditioners perform best in the high rank regime where $T$ is small, which makes sense given that short simulation times yield more localized impulse responses.
-For example, for $\kappa=3.2e{-4}$ using the PSF(5) preconditioner yields 140, 202, and 379 iterations for $T=0.5$, $1.0$, and $2.0$, respectively.
+For example, in the mid-range values $T=1.0$ and $\kappa=3.2$e${-4}$, PSF(1), PSF(5), and PSF(25) require $1.3\times$, $2.3\times$, and $5.1\times$ fewer PCG iterations, respectively, as compared to no preconditioning, and greater improvements as compared to regularization preconditioning.
+The PSF preconditioners perform best in the high rank regime where $T$ is small, which makes sense given that short simulation times yield more localized impulse responses (see Figure~\ref{tab:adv_convergence}).
+For example, for $\kappa=3.2$e${-4}$ using the PSF(5) preconditioner yields 140, 202, and 379 iterations for $T=0.5$, $1.0$, and $2.0$, respectively.
 The performance of the PSF preconditioners as a function of $\kappa$ does not have as clear of a trend.
 Reducing $\kappa$ makes the impulse responses thinner and hence easier to fit in batches, but also increases the complexity of the impulse response shapes, which may reduce the accuracy of the RBF interpolation.
-The greatest improvements are seen for $T=0.5$ and $\kappa=1e{-3}$, for which PSF(25) requires roughly 10 times fewer PCG iterations than no preconditioning, and roughly 20 times fewer iterations than regularization preconditioning.
+The greatest improvements are seen for $T=0.5$ and $\kappa=1$e${-3}$, for which PSF(25) requires roughly $10\times$ fewer PCG iterations than no preconditioning, and roughly $20\times$ fewer iterations than regularization preconditioning.
 
@@
 We applied the method to approximate the Gauss-Newton Hessians in an ice sheet flow inverse problem governed by the Stokes PDE, and a diffusive transport inverse problem governed by an advection-diffusion PDE.
-We saw that preconditioners based on the proposed approximation cluster the eigenvalues of the preconditioned Hessian near one, and allow us to solve the inverse problems using roughly five times fewer PDE solves.
+We saw that preconditioners based on the proposed approximation cluster the eigenvalues of the preconditioned Hessian near one, and allow us to solve the inverse problems using roughly $5\times$--$10\times$ fewer PDE solves.
+For larger domains with more observations, the rank of the data misfit Hessian will increase, while the locality of impulse responses will remain the same.
+Hence, we expect the speedup will be even greater for such problems.
 Although the method we presented is not applicable to all Hessians, it is applicable to many Hessians of practical interest.
-For these Hessians, the proposed method offers a \emph{data scalable} alternative to conventional low rank approximation, due to the ability to form high rank approximations of an operator using a small number of operator applications, and thus forward and adjoint PDE solves.
+For these Hessians, the proposed method offers a \emph{data scalable} alternative to conventional low rank approximation, due to the ability to form high rank approximations of an operator using a small number of operator applications, and thus PDE solves.
 

### 2023-07-06 paper/localpsf.tex
edits with Noemi. Paper should be done

@@
 Further matrix computations are then performed with fast H-matrix methods.
-We demonstrate the effectiveness of this PSF-based method by using it to build preconditioners for the Hessian operator arising in two inverse problems governed by partial differential equations (PDEs): inversion for the basal friction coefficient in an ice sheet flow problem and for the initial condition in a diffusive transport problem.
+We demonstrate the effectiveness of this PSF-based method by using it to build preconditioners for the Hessian operator arising in two inverse problems governed by partial differential equations (PDEs): inversion for the basal friction coefficient in an ice sheet flow problem and for the initial condition in an advective-diffusive transport problem.
 While for many ill-posed inverse problems the Hessian of the data misfit term exhibits a low rank structure, and hence a low rank approximation is suitable, for many problems of practical interest the numerical rank of the Hessian is still large.
@@
 \begin{keywords}
- 	data scalability, Hessian, hierarchical matrix, high-rank, impulse response, local translation invariance, matrix-free, moment methods, operator approximation, PDE constrained inverse problems, point spread function, product convolution
+ 	data scalability, Hessian, hierarchical matrix, high-rank, impulse response, local translation invariance, matrix-free, moment methods, operator approximation, PDE constrained inverse problems, point spread function, preconditioning, product convolution
 \end{keywords}
@@
 \begin{figure}
-	\begin{subfigure}{0.49\textwidth}
+	\begin{subfigure}[b]{0.49\textwidth}
 		\centering
@@
 	\end{subfigure}
-	\begin{subfigure}{0.49\textwidth}
+	\begin{subfigure}[b]{0.49\textwidth}
 		\centering
@@
 The large circle is $\partial \Omega$.
-(\ref{fig:mean_displacement_invariance}) Illustration of impulse responses, $\impulseresponse_x$ and $\impulseresponse_{x'}$, that arise from applying $\Aop$ to point sources at two points, $x$ and $x'$.
+(\ref{fig:mean_displacement_invariance}) Illustration of impulse responses, $\impulseresponse_x$ and $\impulseresponse_{x'}$, corresponding to points $x$ and $x'$.
 The operator $\Aop$ is locally mean displacement invariant (Section \ref{sec:local_mean_displacement_invariance}) if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$.
@@
 The moment method that we use estimates the support of $\phi_x$ more accurately than random noise probing in resolution analysis, at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
+Adding more batches yields impulse responses at more points, increasing the accuracy of the approximation at the cost of one operator application per batch.
 
@@
 In distributed parameter inverse problems governed by PDEs, one seeks to infer an unknown spatially varying parameter field from limited observations of a state variable that depends on the parameter implicitly through the solution of a PDE.
-Conventionally, the inverse problem is formulated using either a deterministic framework~\cite{AsterBorchersThurber18,BanksKunisch89,Vogel02},
+Conventionally, the inverse problem is formulated using either a deterministic framework~\cite{BanksKunisch89,Vogel02},
 or a Bayesian probabilistic framework~\cite{KaipioSomersalo05,Stuart10a,Tarantola05}.
@@
 Owing to the implicit dependence of predicted observations on the parameter, entries of the Hessian are not easily accessible.
-Rather, the Hessian may be applied to a vector via a computational process that involves solving a pair of forward and adjoint PDEs which are linearizations of the original PDE~\cite{GhattasWillcox21}
+Rather, the Hessian may be applied to a vector via a computational process that involves solving a pair of forward and adjoint PDEs which are linearizations of the original PDE~\cite{GhattasWillcox21,PetraStadler11}.
 %,Gunzburger03}
-.
 The most popular matrix-free Hessian approximation methods are based on low rank approximation of either the data misfit Hessian, or the data misfit Hessian preconditioned by the regularization Hessian, e.g.,~\cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15}.
@@
 	\item The supports of the impulse responses $\impulseresponse_x$ are contained in localized regions.
-	\item The integral kernel is non-negative in the sense that $$\Aker(y,x) \ge 0$$ for all $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
-The operator $\Aop$ need not be positive semi-definite to use the proposed method, and positive semi-definite operators need not have a non-negative integral kernel.}
+	\item The integral kernel is non-negative\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
+The operator $\Aop$ need not be positive semi-definite to use the proposed method, and positive semi-definite operators need not have a non-negative integral kernel.} in the sense that $$\Aker(y,x) \ge 0 \quad \text{for all} \quad (y,x) \in \Omega \times \Omega.$$
 \end{enumerate}
@@
 in inverse problems governed by PDEs which model steady state ice
-sheet flow~\cite{PetraEtAl12} (Section~\ref{sec:ice}) and transport
-and flow of a contaminant~\cite{PetraStadler11}
+sheet flow~\cite{PetraEtAl12} (Section~\ref{sec:ice}) and advective-diffusive transport of a contaminant~\cite{PetraStadler11}
 (Section~\ref{sec:adv}). These inverse problems are described in detail in their respective sections.
@@
 We solve the ice sheet inverse problem with an inexact Newton preconditioned
-conjugate gradient (PCG) scheme and a globalizing Armijo line search~\cite{NocedalWright99}. The Newton updates,
+conjugate gradient (PCG) scheme and a globalizing Armijo line search~\cite{NocedalWright99}. The Newton search directions,
 $\mathbf{\searchdir}$, are obtained by solving 
@@
 %
-The solution to these equations is the pair $(\velocity, \pressure)$, where $\velocity$ is the ice flow velocity field\footnote{We do not use bold to denote vector or tensor fields to avoid confusion with vectors that arise from finite element discretizations, which are already denoted with bold.}, and $\pressure$ is the pressure field.
+The solution to these equations is the pair $(\velocity, \pressure)$, where $\velocity$ is the ice flow velocity field\footnote{We do not use bold to denote vector or tensor fields to avoid confusion with vectors that arise from finite element discretizations, which are already denoted with bold.} and $\pressure$ is the pressure field.
 Here, $\basalfriction$ is the unknown logarithmic
@@
 $\dot{\strain}(v)= \frac{1}{2}\left(
-\nabla\velocity+\nabla\velocity^{\top} \right)$ is the strain rate tensor~\cite{Glen1955}. Here $\eta$ is the viscosity and $I$ is the identity operator. 
+\nabla\velocity+\nabla\velocity^{\top} \right)$ is the strain rate tensor~\cite{IsaacEtAl15}. Here $\eta$ is the viscosity and $I$ is the identity operator. 
 %Glen's exponent has been chosen to be one, which provides for a linear Stokes model. 
@@
 Note that $\Gamma_b$ is a two-dimensional surface embedded in three dimensions due to the mountain topography.
-We also generate a flattened version of $\Gamma_b$, denoted by $\Omega \subset \mathbb{R}^2$, by ignoring the height coordinate\footnote{The proposed PSF method requires translating impulse responses.
-This requires either a flat domain, or a notion of local parallel transport.}.
+The proposed PSF method involves translating impulse responses.
+Hence it requires either a flat domain, or a notion of local parallel transport.
+We therefore generate a flattened version of $\Gamma_b$, denoted by $\Omega \subset \mathbb{R}^2$, by ignoring the height coordinate.
 The parameter $\basalfriction$ is viewed as a function on $\Gamma_b$ for the purpose of solving the Stokes equations, and as a function on $\Omega$ for the purpose of building Hessian approximations and defining the regularization.
@@
     \end{subfigure}
-	\caption{(Ice sheet) The log basal friction parameter, with color scale as in Figure~\ref{fig:true_beta}, computed from the PDE-constrained optimization problem with noise levels: $25\%$ (left), $5.0\%$ (middle), and $1.0\%$ (right).}
+	\caption{(Ice sheet) The log basal friction parameter, with color scale as in Figure~\ref{fig:true_beta}, computed from the PDE constrained optimization problem with noise levels: $25\%$ (left), $5.0\%$ (middle), and $1.0\%$ (right).}
 	\label{fig:stokes_reconstructions} 
@@
 	}
-	\caption{(Ice sheet) Convergence history for solving the Stokes inverse problem using inexact Newton PCG.
-The Newton iteration is terminated when $||\mathbf{g}|| < 10^{-6} ||\mathbf{g}_0||$, where $\mathbf{g}$ is the gradient at the current iteration, and $\mathbf{g}_0$ is the gradient at the initial guess.
+	\caption{(Ice sheet) Convergence history for solving the Stokes inverse problem using inexact Newton PCG to tolerance $10^{-6}$. 
+%	The Newton iteration is terminated when $||\mathbf{g}|| < 10^{-6} ||\mathbf{g}_0||$, where $\mathbf{g}$ is the gradient at the current iteration, and $\mathbf{g}_0$ is the gradient at the initial guess.
 	Preconditioners shown are the proposed PSF method with five batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
-Columns titled \#CG show the number of PCG iterations used to solve the Newton system for $\mathbf{\widehat{\basalfriction}}$.
-Columns titled $\|\mathbf{g}\|$ show the $l^2$ norm of the gradient at $\mathbf{\basalfriction}$.
-Columns titled \#Stokes show the total number of Stokes PDE solves performed in each Newton iteration.
-This consists of Stokes solves performed during the line search to arrive at the current iterate from the previous iterate, plus one adjoint Stokes solve to compute the gradient at the current iterate, plus two Stokes solves per PCG iteration when solving the Newton system to compute the next search direction.
+Columns \#CG show the number of PCG iterations used to solve the Newton system for $\mathbf{\widehat{\basalfriction}}$.
+Columns $\|\mathbf{g}\|$ show the $l^2$ norm of the gradient at $\mathbf{\basalfriction}$.
+Columns \#Stokes show the total number of Stokes PDE solves performed in each Newton iteration.
+%	This consists of Stokes solves performed during the line search to arrive at the current iterate from the previous iterate, plus one adjoint Stokes solve to compute the gradient at the current iterate, plus two Stokes solves per PCG iteration when solving the Newton system to compute the next search direction.
 	Under PSF (5) and in row Iter 3, we write $6+22$ to indicate that $6$ Stokes solves were used during the standard course of the iteration, and $22$ Stokes solves were used to build the PSF (5) preconditioner.}
@@
 	\caption{(Ice sheet) Left: convergence history for solving $\mathbf{H} \mathbf{x} = \mathbf{b}$ using PCG, where $\mathbf{b}$ has i.i.d.\ random entries drawn from the standard Gaussian distribution and $\mathbf{H}$ is evaluated at the solution of the inverse problem.
-	Results in these figures are shown for the PSF-based preconditioners with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
+	Results in these figures are shown for the PSF-based preconditioners with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
 The preconditioner is constructed using $\mathbf{H}_\text{gn}$.
@@
 
-In Table~\ref{tab:condition_number}, we show the condition number of the preconditioned Hessian for a variety of noise levels ranging from $1\%$ to $25\%$.
+In Table~\ref{tab:condition_number}, we show the condition number of the preconditioned Hessian for noise levels ranging from $1\%$ to $25\%$.
 Note that the condition number using PSF-based preconditioners is extremely small (ranging between 1 and 10) and relatively stable over this range of noise levels.
 As expected, PSF (25) outperforms PSF (5), which outperforms PSF (1).
-All PSF-based preconditioners outperform no preconditioning and regularization preconditioning by several orders of magnitude for all noise levels.
+All PSF-based preconditioners outperform regularization and no preconditioning by several orders of magnitude for all noise levels.
 
@@
 	\begin{center}
-			{\footnotesize
+			{\small
 			\begingroup
@@
 		\end{center}
-	\caption{(Ice sheet) Condition number for $\mathbf{\preconditioner}^{-1} \mathbf{H}$ for the PSF-based preconditioners with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF(25), respectively), no preconditioner (NONE) and regularization (REG).
+	\caption{(Ice sheet) Condition number for $\mathbf{\preconditioner}^{-1} \mathbf{H}$ for the PSF-based preconditioners with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), no preconditioner (NONE) and regularization (REG).
 All operators are evaluated at the soutions of the inverse problems for
@@
 
-\subsection{Example 2: Inversion for the initial condition in a diffusive transport problem}
+\subsection{Example 2: Inversion for the initial condition in an advective-diffusive transport problem}
 \label{sec:adv}
@@
 at a final time, $T$. This PDE
-models diffusive transport in a domain $\advdomain \subset \R^d$, which is
+models advective-diffusive transport in a domain $\advdomain \subset \R^d$, which is
 depicted in Figure~\ref{fig:adv_inversion1}. In this case, the state, $\concentration(x,t)$, could be interpreted as the concentration of a contaminant. The problem description
-below closely follows the one in~\cite{PetraStadler11}. The domain
+below closely follows~\cite{PetraStadler11,VillaPetraGhattas21}. The domain
 boundaries $\partial \advdomain$ include the outer boundaries as well as the
@@
 The results show that PSF-based preconditioning outperforms regularization preconditioning and no preconditioning in all cases except one.
-The exception is $T=2.0$ and $\kappa=1.0$e${-3}$, in which PSF(1) performs slightly worse than regularization preconditioning but better than no preconditioning.
+The exception is $T=2.0$ and $\kappa=1.0$e${-3}$, in which PSF (1) performs slightly worse than regularization preconditioning but better than no preconditioning.
 Adding more batches yields better results, and the impact of adding more batches is more pronounced here than in the ice sheet example.
-For example, in the mid-range values $T=1.0$ and $\kappa=3.2$e${-4}$, PSF(1), PSF(5), and PSF(25) require $1.3\times$, $2.3\times$, and $5.1\times$ fewer PCG iterations, respectively, as compared to no preconditioning, and greater improvements as compared to regularization preconditioning.
+For example, in the mid-range values $T=1.0$ and $\kappa=3.2$e${-4}$, PSF (1), PSF (5), and PSF (25) require $1.3\times$, $2.3\times$, and $5.1\times$ fewer PCG iterations, respectively, as compared to no preconditioning, and greater improvements as compared to regularization preconditioning.
 The PSF preconditioners perform best in the high rank regime where $T$ is small, which makes sense given that short simulation times yield more localized impulse responses (see Figure~\ref{tab:adv_convergence}).
-For example, for $\kappa=3.2$e${-4}$ using the PSF(5) preconditioner yields 140, 202, and 379 iterations for $T=0.5$, $1.0$, and $2.0$, respectively.
+For example, for $\kappa=3.2$e${-4}$ using the PSF (5) preconditioner yields 140, 202, and 379 iterations for $T=0.5$, $1.0$, and $2.0$, respectively.
 The performance of the PSF preconditioners as a function of $\kappa$ does not have as clear of a trend.
 Reducing $\kappa$ makes the impulse responses thinner and hence easier to fit in batches, but also increases the complexity of the impulse response shapes, which may reduce the accuracy of the RBF interpolation.
-The greatest improvements are seen for $T=0.5$ and $\kappa=1$e${-3}$, for which PSF(25) requires roughly $10\times$ fewer PCG iterations than no preconditioning, and roughly $20\times$ fewer iterations than regularization preconditioning.
+The greatest improvements are seen for $T=0.5$ and $\kappa=1$e${-3}$, for which PSF (25) requires roughly $10\times$ and $20\times$ fewer PCG iterations than no preconditioning and regularization preconditioning, respectively.
 
@@
 	\end{center}
-	\caption{(Diffusive transport) (\ref{fig:adv_initial_condition}) True initial condition.
+	\caption{(Advective-diffusive transport) (\ref{fig:adv_initial_condition}) True initial condition.
 (\ref{fig:adv_velocity}) Velocity field.
@@
 	\begin{center}
-		{\footnotesize
+		{\small
 		\begingroup
@@
 	\end{center}
-	\caption{(Diffusive transport) Number of PCG iterations required to solve the Newton linear system to tolerance $||\widehat{\mathbf{q}}_j - \widehat{\mathbf{q}}|| < 10^{-6}||\widehat{\mathbf{q}}||$, where $\widehat{\mathbf{q}}_j$ is the $j$th iterate, and $\widehat{\mathbf{q}}$ is the solution of the Newton linear system.
+	\caption{(Advective-diffusive transport) Number of PCG iterations required to solve the Newton linear system to tolerance $||\widehat{\mathbf{q}}_j - \widehat{\mathbf{q}}|| < 10^{-6}||\widehat{\mathbf{q}}||$, where $\widehat{\mathbf{q}}_j$ is the $j$th iterate, and $\widehat{\mathbf{q}}$ is the solution of the Newton linear system.
 Iteration counts are shown for a variety of different diffusion parameters $\kappa$, simulation times $T$, and preconditioners.}
@@
 	}
-	\caption{(Diffusive transport) Impulse responses for small and large diffusion parameters $\kappa$ and simulation times $T$.}
+	\caption{(Advective-diffusive transport) Impulse responses for small and large diffusion parameters $\kappa$ and simulation times $T$.}
 	\label{tbl:adv_vary_kappa_tf}
@@
 	\end{subfigure}
-	\caption{(Diffusive transport) Left: convergence history for solving $\mathbf{H} \mathbf{x} = \mathbf{b}$ using PCG, where $\mathbf{b}$ has i.i.d.\ random entries drawn from the standard Gaussian distribution.
+	\caption{(Advective-diffusive transport) Left: convergence history for solving $\mathbf{H} \mathbf{x} = \mathbf{b}$ using PCG, where $\mathbf{b}$ has i.i.d.\ random entries drawn from the standard Gaussian distribution.
 Right: generalized eigenvalues for generalized eigenvalue problem $\mathbf{H} \mathbf{u}_{k}=\lambda_{k} \mathbf{\preconditioner} \mathbf{u}_{k}$.
@@
 Point source locations for the impulse response batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulse responses per batch as possible, while ensuring that the corresponding ellipsoids do not overlap.
-We applied the method to approximate the Gauss-Newton Hessians in an ice sheet flow inverse problem governed by the Stokes PDE, and a diffusive transport inverse problem governed by an advection-diffusion PDE.
+We applied the method to approximate the Gauss-Newton Hessians in an ice sheet flow inverse problem governed by a linear Stokes PDE, and an advective-diffusive transport inverse problem governed by an advection-diffusion PDE.
 We saw that preconditioners based on the proposed approximation cluster the eigenvalues of the preconditioned Hessian near one, and allow us to solve the inverse problems using roughly $5\times$--$10\times$ fewer PDE solves.

### 2023-07-06 paper/localpsf.tex
re-read all of Omar's comments. I believe we have addressed them

@@
 In summary, computing all the necessary ingredients to evaluate kernel entries of the PSF approximation requires a handful of operator applications (e.g., $6+\nbatch$ operator applications in two dimensions, or $10+\nbatch$ operator applications in three dimensions, with $\nbatch$ typically in the range 1--25), plus comparatively cheap additional overhead costs, most notably performing ellipsoid intersection tests while choosing sample point batches.
-Once these ingredients are computed, no more operator applications are required, and approximate kernel entries can be evaluated rapidly.
+Once these ingredients are computed, no more operator applications (and thus PDE solves) are required, and approximate kernel entries can be evaluated rapidly.
 Constructing the H-matrix from kernel entries requires a number of elementary operations that scales polylog linearly in $N$.
@@
 We compare the proposed method with the most commonly used existing preconditioners: no preconditioning (NONE), and preconditioning by the regularization term in the Hessian (REG).
-The results show that using PSF (5) reduces the total number of Stokes PDE solves to 70, as compared to 908 for regularization preconditioning and 308 for no preconditioning, which corresponds to a $5\times$--$10\times$ speedup.
+The results show that using PSF (5) reduces the total number of Stokes PDE solves to 70, as compared to 908 for regularization preconditioning and 308 for no preconditioning.
 For problems with a larger physical domain and correspondingly more observations, such as continental scale ice sheet inversion, the speedup will be even greater.
@@
 Adding more batches yields better results, and the impact of adding more batches is more pronounced here than in the ice sheet example.
-For example, in the mid-range values $T=1.0$ and $\kappa=3.2$e${-4}$, PSF (1), PSF (5), and PSF (25) require $1.3\times$, $2.3\times$, and $5.1\times$ fewer PCG iterations, respectively, as compared to no preconditioning, and greater improvements as compared to regularization preconditioning.
+For example, in the mid-range values $T=1.0$ and $\kappa=3.2$e${-4}$, PSF (1), PSF (5), and PSF (25) require $1.3\times$, $2.3\times$, and $5.1\times$ fewer PCG iterations, respectively, as compared to no preconditioning, and exhibit greater improvements as compared to regularization preconditioning.
 The PSF preconditioners perform best in the high rank regime where $T$ is small, which makes sense given that short simulation times yield more localized impulse responses (see Figure~\ref{tab:adv_convergence}).

### 2023-07-06 paper/localpsf.tex
did a complete readover. A coupld minor fixes. Going to submit this version.

@@
 Numerical results reveal that the PSF-based preconditioner clusters the spectrum of the preconditioned Hessian near one, yielding roughly $5\times$--$10\times$ reductions in the required number of PDE solves, as compared to classical regularization-based preconditioning and no preconditioning.
-We also present a comprehensive numerical study for the influence of various parameters (that control the shape of the impulse responses and the rank of the Hessian) on the effectiveness of the approximation.
+We also present a comprehensive numerical study for the influence of various parameters (that control the shape of the impulse responses and the rank of the Hessian) on the effectiveness of the advection-diffusion Hessian approximation.
 The results show that the PSF-based preconditioners are able to form good approximations of high-rank Hessians using only a small number of operator applications.
@@
 Here $\spatialvol_\text{max}$ is the largest value of $\spatialvol(\candidatepoint_j)$ over all points $q$ in the initial set of candidate points, and $\epsilon_V$ is a small threshold parameter (we use $\epsilon_V=10^{-5}$).
-	\item We make sure that all eigenvalues of $\spatialcov(\candidatepoint_i)$ are positive, and the aspect ratio of $E_{\candidatepoint_i}$ (ratio of the largest eigenvalue of $\spatialcov(\candidatepoint_i)$ to the smallest) is bounded by a constant $1/\epsilon_\Sigma$ (we use $1/\epsilon_\Sigma=20$).
+	\item We make sure that all eigenvalues of $\spatialcov(\candidatepoint_i)$ are positive, and the aspect ratio of $E_{\candidatepoint_i}$ (square root of the ratio of the largest eigenvalue of $\spatialcov(\candidatepoint_i)$ to the smallest) is bounded by a constant $1/\epsilon_\Sigma$ (we use $1/\epsilon_\Sigma=20$).
 Negative integral kernel entries due to discretization error can cause $\spatialcov(\candidatepoint_i)$ to be indefinite or highly ill-conditioned.
@@
 The primary computational task in the Cayley shift-and-invert scheme is the solution of shifted linear systems of the form
-$\left(\mathbf{A}_H^{\text{sym}} + \mu_i \mathbf{R}_H\right) x = b,$
+$\left(\mathbf{A}_H^{\text{sym}} + \mu_i \mathbf{R}_H\right) \mathbf{x} = \mathbf{b},$
 for a small number of positive shifts $\mu_i$.
@@
 We compare the proposed method with the most commonly used existing preconditioners: no preconditioning (NONE), and preconditioning by the regularization term in the Hessian (REG).
-The results show that using PSF (5) reduces the total number of Stokes PDE solves to 70, as compared to 908 for regularization preconditioning and 308 for no preconditioning.
+The results show that using PSF (5) reduces the total number of Stokes PDE solves to 70, as compared to 908 for regularization preconditioning and 308 for no preconditioning, a reduction in cost of roughly $5\times$--$10\times$.
 For problems with a larger physical domain and correspondingly more observations, such as continental scale ice sheet inversion, the speedup will be even greater.
@@
 	Right: generalized eigenvalues for generalized eigenvalue problem $\mathbf{H} \mathbf{u}_{k}=\lambda_{k} \mathbf{\preconditioner} \mathbf{u}_{k}$.
-Here $\mathbf{H}$ is the Hessian and $\mathbf{\preconditioner}$ is the PSF approximation constructed using the Gauss-Newton Hessian, $\mathbf{H}_\text{gn}$, for 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), the regularization Hessian (REG), or the identity matrix (NONE).}
+Here $\mathbf{H}$ is the Hessian and the matrices $\mathbf{\preconditioner}$ are the same Hessian approximations used in the left sub-figure, with NONE corresponding to the identity matrix.}
 	\label{fig:krylov_convergence}
@@
 		\end{center}
-	\caption{(Ice sheet) Condition number for $\mathbf{\preconditioner}^{-1} \mathbf{H}$ for the PSF-based preconditioners with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), no preconditioner (NONE) and regularization (REG).
+	\caption{(Ice sheet) Condition number for $\mathbf{\preconditioner}^{-1} \mathbf{H}$ for the PSF-based preconditioners with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), no preconditioner (NONE) and regularization preconditioning (REG).
 All operators are evaluated at the soutions of the inverse problems for
@@
 Reynolds number 50, with boundary conditions $v(x)=(0,1)$ on the left boundary, $v(x)=(0,-1)$ on the right boundary, and $v(x)=(0,0)$ on the top and bottom boundaries, as in~\cite[Section 3]{PetraStadler11}.
-We use a checkerboard image for the initial condition (Figure~\ref{fig:adv_initial_condition}) and add $5\%$ multiplicative noise to generate synthetic observations at the final time, $T$.
+We use a checkerboard image for the initial condition (Figure~\ref{fig:adv_initial_condition}) and add $5\%$ multiplicative noise to generate a synthetic observation at the final time, $T$.
 The initial condition, velocity field, noisy observations, and reconstructed initial condition are shown in Figure~\ref{fig:adv_inversion1}.
@@
 Color indicates magnitude of velocity vector.
-(\ref{fig:adv_observations}) Noisy observations of concentration at final time.
+(\ref{fig:adv_observations}) Noisy observations of concentration at the final time.
 (\ref{fig:adv_reconstruction}) Reconstructed initial condition.}

### 2023-07-06 paper_sisc_first_submission/KrylovConvergencePlot.tex
add subdirectory with frozen paper from first submission

\begin{figure}[H]
\begin{center} 
\begin{tikzpicture}[baseline, trim axis right]
\begin{axis}[width=0.99\textwidth, height=0.75\textwidth,
            ymode=log,
            ymin=1e-11,
            ymax=2,
            xmax=750,
            xtick={0, 100, 200, 300, 400, 500, 600, 700},
            xlabel={$k$, CG iteration},
            ylabel={$||r_{k}||_{2}/||r_{0}||_{2}$, relative residual norm},
			legend pos=north east,
]
\addlegendentry{None}
\addplot[dash dot, line width=0.5mm] table {cg_relres_none.dat};
\addlegendentry{Reg}
\addplot[dotted, line width=0.5mm] table {cg_relres_reg.dat};
\addlegendentry{PSF (1)}
\addplot[blue, line width=0.5mm] table {cg_relres_p1.dat};
\addlegendentry{PSF (5)}
\addplot[red, line width=0.5mm] table {cg_relres_p5.dat};
\addlegendentry{PSF (25)}
\addplot[purple, line width=0.5mm] table {cg_relres_p25.dat};
\end{axis}
\end{tikzpicture}
\end{center}
\end{figure}

### 2023-07-07 submission_letter/cover_letter.tex
add submission letter

\documentclass[11pt]{article}
\setlength{\hoffset}{-1in}       \setlength{\voffset}{-1in}
\setlength{\textwidth}{140mm}    \setlength{\textheight}{240mm}
\setlength{\topmargin}{18mm} \setlength{\oddsidemargin}{35mm}
\setlength{\evensidemargin}{25mm} \setlength{\headheight}{25mm}
\setlength{\headsep}{5mm} \setlength{\marginparwidth}{20mm}

\usepackage{graphicx}
\usepackage[right]{eurosym}

\begin{document}

\pagestyle{empty}
 \begin{figure}
 \begin{minipage}{.5\columnwidth}
 \end{minipage}\hfill
 \begin{minipage}[b]{.48\columnwidth}
\flushright
 \noindent{\sc\large Nick Alger}\vspace{-0.9cm}
 \end{minipage}
 \end{figure}
 \noindent\rule{\columnwidth}{.5pt}\vspace{-2ex}
 \begin{figure}[h]
 \begin{minipage}{.2\columnwidth}
 \end{minipage}\hfill
 \begin{minipage}{.78\columnwidth}
 \flushright\vspace{-2ex}
 {\noindent Oden Institute for Computational Engineering \& Sciences \\ 
 University of Texas at Austin, \\ 
 1 Univ.\ Station, C0200,
 TX 78712 USA \\
 email: {\tt nalger225@gmail.com}\\
 tel: +1 425 736 4111}\\
 \end{minipage}
 \end{figure}
% \vspace{1cm} \noindent \\
\phantom{adafdsf}
\vspace{3ex}

%\noindent
%Jan S. Hesthaven \\
%EPFL SB-DO\\
%PH A2 364 (B\^atiment PH)\\
%Station 3\\
%CH-1015 Lausanne \\
%Switzerland
%{
%\flushright{\today}\\
%}
% \vspace{0.3cm}
\noindent Dear Professor Hans De Sterck:\\[2ex]

We wish to submit the manuscript {\em Point spread function approximation of high rank Hessians with locally supported non-negative integral kernels} by Nick Alger, Tucker Hartland, Noemi Petra, and Omar Ghattas to be considered for publication in SISC in the {\em
  Numerical Algorithms for Scientific Computing} section.
%

In this manuscript, we propose an efficient point spread function (PSF) method for approximating matrix-free operators with locally supported non-negative integral kernels. We improve upon existing PSF methods by using a novel moment method for forming a-priori ellipsoid estimates for the supports of impulse responses. By solving an ellipsoid packing problem, we compute a large number of impulse responses per operator application, making the method highly efficient in terms of operator applications. We also provide an improved method for interpolating impulse responses based on a notion we call ``local mean displacement invariance,'' which generalizes classical local translation invariance. We apply the method to approximate high rank Hessians PDE constrained inverse problems. Our numerical results demonstrate that the proposed method outperforms conventional approaches, and that the method can approximate high rank Hessians with a small number of operator applies. The manuscript has not been submitted elsewhere for
publication.\\[1ex]

\noindent Thank you for your consideration.\\[3ex]

\noindent Sincerely,\\[5ex]
\phantom{dd}\hspace{2cm} Nick Alger
\end{document}

### 2024-01-05 revision/reply_to_reviewer1.tex
clean up

\documentclass[11pt]{article}
\renewcommand*\familydefault{\sfdefault}
%\usepackage{amssymb,amsmath,amsfonts,comment}
%\usepackage{amsmath,amssymb,graphicx,subfigure,psfrag}
\usepackage{amsmath,amssymb,graphicx,subfigure,psfrag,upgreek}
\usepackage{amssymb,mathrsfs}
\usepackage[margin=1in]{geometry}
\usepackage{tikz,pgfplots,graphicx}
\usepackage{color,pdfcolmk}
\usepackage{amsmath}

%\newcommand{\alennote}[1]{\noindent\emph{\textcolor{blue!50!black}{Alen:#1\:}}}
%\newcommand{\gsnote}[1]{\noindent\emph{\textcolor{green}{GS: #1\:}}}
\newcommand{\todo}[1]{\noindent\emph{\textcolor{red}{Todo: #1\:}}}
\newcommand{\note}[1]{\noindent\emph{\vspace{1ex}\textcolor{blue}{Note: #1\:}}\\[1ex]}
\newcommand{\nnote}[1]{\noindent\emph{\vspace{1ex}\textcolor{red}{Note: #1\:}}}
\newcommand{\referee}[1]{\vspace{.1ex}\noindent{\textcolor{blue}{#1}}}
\newcommand{\nnn}{\mathbf{n}}
\newcommand{\fff}{\mathbf{f}}
\newcommand{\uuu}{\mathbf{u}}
\newcommand{\+}{\otimes}
\newcommand{\m}{\mathcal}
\newcommand{\bs}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand{\eps}{\varepsilon}

\newcommand{\gs}[1]{\textcolor{green}{G: #1}}
\newcommand{\np}[1]{\textcolor{blue}{N: #1}}
\newcommand{\tuck}[1]{\textcolor{brown}{T: #1}}
\newcommand{\mauro}[1]{\textcolor{red}{#1}}

\begin{document}

\section*{Reply to the Reviewer 2:}

Thanks for the careful reading and for your helpful comments and
suggestions.  Please find below point-by-point replies (in black) to
your comments and questions (which are reprinted in blue). To give you
an overview of all the changes in the paper, we also provide a
diff-document that highlights the changes between the initial
submission and this re-submission.\\[1ex]

\referee{This article addresses an important topic in large-to-extreme-scale inverse problems. Namely, efficient approximation of the Hessian of the negative-log of the posterior in Bayesian inversion.
Specifically, this work explores utilizing hierarchical off-diagonal low-rank (HODLR) matrix approximations of the Hessian.
The idea is clearly presented and the manuscript is well-written, however, I believe some---I would say minor---adjustments are required before recommending acceptance for publications.
Please see my comments below. \\}

\paragraph{\sc General Comments:}

\begin{enumerate} 
\item \referee{I do not think the terms ``Gaussianization, Gaussianized, ...'' are properly utilized in this work. The term Gaussianization has been actively used, especially in the machine learning literature, to refer to transformation of non-Gaussian data/variable in to a Gaussian. This requires finding a map/transformation between the original distribution and a Gaussian, e.g., an optimal transport, or normalizing flows. The authors, however, use these terms to denote approximating a non-Gaussian distribution with a localized Gaussian, e.g., following a Laplacian approach.}

\noindent{We appreciate drawing our attention to the fact that our usage of ``Gaussianization, Gaussianized...'', would be unclear to many readers. We have updated our usage to ``the Laplace approximation of the posterior''.}

\item \referee{Re: Algorithm 2 (page 22); it would be great to discuss how $r_1, \ldots, r_{L}$ can be auto tuned. What metric can be used to automatically choose the optimal value of $L$ and then choose the rank for each level automatically.}

\noindent{Done, with regard to choosing the ranks $r_{1},\dots,r_{L}$. We are not aware of an automatic way to choose the optimal value of $L$ but have described the components that we believe could be used to develop such means.}
\end{enumerate}

\paragraph{\sc Other comments:}
\begin{enumerate}
\item  \referee{Page 1, Line 25: ``approximation of Hessian applies...''; Hessian of what?}

\noindent{We now use the more descriptive ``discretized objective functional Hessian vector-products''.}

\item \referee{Page 1, Line 28--31: statement too long, maybe try splitting!}

\noindent{Done.}

\item \referee{Page 1, Line 29: ``sampling from Gaussian proposal distributions'' $\rightarrow$ ``sampling Gaussian distributions''.}

\noindent{We have not used the exact suggested text since the sentence phrasing has been changed. We believe that ``...that provide a means to sample Gaussian distributions...'', is as close to the suggested rephrasing but which maintains consistency with the new sentence structure.}

\item \referee{Page 1, Line 30: I am not sure I understand what the authors mean by ``Newton solves by direct linear methods''!}

\noindent{We had intended to describe methods such as LU factorization, which is known as a direct method of linear system solution. We hope that by referencing a relevant factorization that the message intended to be communicated is now made more clear.}

\item \referee{Page 1, Line 46: ``to inverse problems'' ``with inverse problems''.}

\noindent{Done, thanks!}

\item \referee{Page 2, Line 7: simulation... systems plays'' $\rightarrow$ ``simulations ... systems play''.}

\noindent{Done.}

\item \referee{Page 3, Line 17: ``generate approximate samples from a Gaussian posterior distribution for...'' $\rightarrow$ ``generate samples from an approximate Gaussian posterior for...''}

\noindent{Done, but we have kept ``Gaussian posterior distribution for...''}

\item \referee{Page 3, Line 22: ``including for a Greenland ice...'' $\rightarrow$ ``including a Greenland ice...''}

\noindent{Done.}

\item \referee{Page 3, Line 28: Add a reference; ``In this section,'' $\rightarrow$ ``In this section (2.1),''}

\noindent{Done, but with the capitalization scheme used throughout the manuscript, ``In this Section (2.1),''}

\item \referee{Page 3, line 44: Remove ``i.e., the data,'' as it is redundant.}

\noindent{Done. We have added an additional sentence in which we explicitly state that we will later refer to $\bs{d}$ as the data instead of the more verbose observational data.}

\item \referee{Page 3, Line 46--47: ``and mathematically'' $\rightarrow$ ``and is mathematically''}

\noindent{Done.}

\item \referee{Page 4, Lines 12--13: The authors use the term ``parameter-to-PDE-solution map'' which has not been introduced. Only the combined ``parameter-to-observable map'' has been introduced on line $10$ above. I suggest being consistent and either explicitly decompose the parameter-to-observable-map into a solution operator (solution-to-PDE-solution) and an observation operator, or just restrict usage to ``parameter-to-observable-map''. Either one or both of the solution operator and the observation operator can be nonlinear and each is associated with its own difficulties.}

\noindent{Done. We have restricted usage to the ``parameter-to-observable map''.}

\item \referee{Page 4, Line 27: ``it's gradient'' $\rightarrow$ ``its gradient''}

\noindent{Done.}

\item \referee{Page 4, Lines 28--33: This paragraph requires some
  modification. If you can well-approximate the posterior you do not
  necessarily need MCMC; you can even use naive MC by applying MH step
  to samples from this Gaussian. MCMC does not generally require
  approximating the posterior with a Gaussian, however sampling a
  Gaussian approximation can be useful. For example in Hamiltonian
  Monte Carlo (HMC), samples from a Gaussian approximation can be used
  to build a mass matrix that optimizes the performance of the
  sampler; diagonal of the mass matrix are optimally set to posterior
  precisions (or approximations thereof). See e.g., reference
  \cite{neal2011mcmc, attia2017reduced}.}

\noindent{We agree with the reviewer, but our experience is that using
  MCMC is still superior to using MH (or importance sampling)
  directly, in particular in high dimensions.  We adjusted this
  paragraph in the revised manuscript.}

%\todo{Georg, please see if you you'd like to add or change anything here.}

\item \referee{Page 4, Line 38: define $_{\ell_{2}}$ presented in the previous equation}

\noindent{This item has been addressed by removing any usage of $_{\ell_{2}}$.}

\item \referee{Page 5, Line 38: Maybe redefine paragraph titles to add a period at the end; ``Compression'' $\rightarrow$ ``Compression.''}

\noindent{Done.}

\item \referee{Page 5, Line 50: replace the comma with a semicolon.}

\noindent{Done.}

\item \referee{Page 5, Line 60: ``algorithm is made possible''; where? add a reference/citation here!}

\noindent{Done.}

\item \referee{Page 6, Line 5: a period is missing after ``Approximations''}

\noindent{Done.}

\item \referee{Page 6, Line 6: hereafter use a different symbol for the number of oversampling vectors as $\bs{d}$ has been utilized to denote the data, despite being boldface.}

\noindent{Done. We are now using the symbol $q$ to denote the number of oversampling vectors.}

\item \referee{Page 6, Lines 24--30: this sentence is too long; try splitting.}

\noindent{We have reformulated this sentence.}

\item \referee{Page 7, Line 40: the term data-misfit Hessian has already been introduced before; remove ``i.e., the Hessian ... functional'' as it is redundant.}

\noindent{Done.}

\item \referee{Page 7, Line 44: remove ``of it''}

\noindent{Done.}

\item \referee{Page 8, Line 40: ``basis function'' $\rightarrow$ ``basis functions''}

\noindent{Done.}

\item \referee{Page 8, Line 51: please see my comment above about using the term ``Gaussianized''}

\noindent{Noted and addressed.}

\item \referee{Page 9, Line 22: I am not sure what ``means of applying'' here means!}

\noindent{We have clarified the vague ``means of applying'' with ``means of computing square root and inverse square root matrix-vector products.}

\item \referee{Page 9, Lines 33--40: this paragraph is redundant. This material/argument has already been introduced in Section 2.1.}

\noindent{We believe that this material is important as it helps show how the example problem fits into the larger class of inverse problems, therefore we decided to keep it here.}

%\todo{I (Tucker) believe that this argument could be strengthened.}

\item \referee{Page 9, Line 49: ``can mathematically'' $\rightarrow$ ``can be mathematically''}

\noindent{Done.}

\item \referee{Page 10 I suggest using subequations for the model Equations (6)--(9)}

Done.

\item \referee{Page 10, Lines 16--19, 22, 43: in reference to model equations, use $\backslash$eqref instead of $\backslash$ref}

Done.

\item \referee{Page 11, Line 5: see my comment above about using the term ``Gaussianizaiton''. I will not comment on this hereafter.}

\noindent{Noted and addressed.}

\item \referee{Page 11, Lines 21, 28: use $\backslash$eqref in reference to model equations.}

Done.

\item \referee{Page 11, Line 31: ``add $1\%$ relative Gaussian noise...'': $1\%$ of what? mean true observation magnitude, $\ell_{\infty}$, ...etc.; please clarify!}

\noindent{Done.}

\item \referee{Page 11, Line 57: I do not believe ``approximate samples'' is a correct terminology. I suggest using ``samples from the approximate posterior''.}

\noindent{This has been addressed.}

\item \referee{Page 12, Figure 3 caption: something is not quite right here. The caption states that circles show representative (random measurement locations), however, lines 29-30 on page 11 say that observations are uniformly distributed. What am I missing here? }

\noindent{The figure has been adjusted so that the measurement locations are now illustrated as being uniformly distributed.}

\item \referee{Page 14, Line 57: there is an extra comma after ``more data''}

\noindent{Done.}

\item \referee{Page 15, Figure 7 caption (line 21): ``low-rank (LR)'' $\rightarrow$ ``LR'' as the acronym has already been introduced; same on line 24 of the same caption: ``low-rank'' $\rightarrow$ ``LR'' }

\noindent{Done.}

\item \referee{Page 15, Figure 8 caption: same as the comment above on using ``LR''}

\noindent{Done.}

\item \referee{Page 16, Line 5: remove the comma after ``Albany'', and remove ``so-called''}

\noindent{Done.}

\item \referee{Page 16, Line 30: use $\backslash$eqref in reference to Equation (3)}

Done.

\item \referee{Page 20, Line 29: remove the comma before ``associated to finite...''}

\noindent{Done.}

\item \referee{Page 23, Line  5: ``low-rank'' $\rightarrow$ ``LR'' }

\noindent{Done.}

\end{enumerate}
%\bibliographystyle{unsrt}
\bibliographystyle{iopart-num}
\bibliography{reviewer2_references}

%\noindent [1]- Neal, Radford M. "MCMC using Hamiltonian dynamics." Handbook of markov chain monte carlo 2, no. 11 (2011): 2.

%\noindent [2]- Attia, Ahmed, Razvan Stefanescu, and Adrian Sandu. "The reduced-order hybrid Monte Carlo sampling smoother." International Journal for Numerical Methods in Fluids 83, no. 1 (2017): 28-51.
\end{document}

### 2024-01-05 revision/reply_to_reviewer1.tex
aaaa

@@
 
-\section*{Reply to the Reviewer 2:}
+\section*{Reply to the Reviewer 1:}
 
@@
 
-\referee{This article addresses an important topic in large-to-extreme-scale inverse problems.
-Namely, efficient approximation of the Hessian of the negative-log of the posterior in Bayesian inversion.
-Specifically, this work explores utilizing hierarchical off-diagonal low-rank (HODLR) matrix approximations of the Hessian.
-The idea is clearly presented and the manuscript is well-written, however, I believe some---I would say minor---adjustments are required before recommending acceptance for publications.
-Please see my comments below. \\}
+\referee{Many Hessian-based computational algorithms have been
+  developed to solve large-scale inverse problems constrained by
+  partial differential equations (PDEs) in the last decade, especially
+  in the context of inferring infinite-dimensional
+  parameters. However, it remains a critical computational challenge
+  to solve such inverse problems when the Hessian of the data-misfit
+  term does not have fast spectrum decay, or the Hessian is
+  high-rank. This challenge often leads to the need to solve a
+  considerable number of PDEs in many (preconditioned) CG iterations
+  with an inexact Newton-CG algorithm, one of the most advanced
+  optimization algorithms in solving infinite-dimensional inverse
+  problems.\\}
 
-\paragraph{\sc General Comments:}
+  \referee{To address this challenge, this paper proposes an
+  efficient preconditioned CG method with the preconditioner computed
+  as an approximate Hessian at some suitable parameter point, which is
+  demonstrated to significantly reduce the number of required CG
+  iterations to achieve given accuracy. The key novelty of this paper
+  is the development of an efficient hierarchical matrix approximation
+  of the Hessian using only a small number of Hessian matrix-vector
+  products, thus a small number of PDE solves. This novelty is made
+  possible by exploiting the property of the Hessian operator with
+  locally supported non-negative integral kernels. Specifically, the
+  authors propose a point spread function (PSF) approximation of the
+  Hessian by (1) computing the zeroth, first, and second-order
+  impulse response moments of the Hessian by applying its product with
+  constant, linear, and quadratic functions, (2) building local
+  ellipsoid support es- timate of the impulse response functions of
+  the Hessian based on these moments, (3) selecting sample points of
+  the impulse response from a candidate set by a greedy ellipsoid
+  packing algorithm, (4) computing the impulse responses with disjoint
+  ellipsoid in batches by applying the Hessian to Dirac combs, which
+  plays a key role in reducing the total number of Hessian
+  matrix-vector product, (5) approximating any integral kernel entries
+  by a radial basis interpolation based on the computed impulse
+  responses, and (6) building the hierarchical matrix approximation of
+  the Hessian with the radial basis interpolation of the integral
+  kernel entries before applying proper symmetrizing and flipping of
+  the negative eigenvalues to make the approximate matrix symmetric
+  positive semi-definite.\\}
+
+\referee{The PSF approximation in this paper is built for more general
+  operators than Hessian. It demonstrates the effectiveness of the
+  PSF-based method by 1 using it to build preconditioners for the
+  Hessian operator in two challenging inverse problems of basal
+  friction coefficient inversion of ice sheet flow and ini- tial
+  condition inversion of advection-dominated transport. It shows
+  significant reductions (5-10X) in the required number of PDE solves
+  compared to classical regularization-based preconditioning and no
+  preconditioning. The paper also presents a comprehensive numerical
+  study on the influence of various parameters (data noises, \# batchs,
+  diffusion coefficients, terminal times) on the effective- ness and
+  data-scalability of the proposed method, showing that the PSF-based
+  preconditioners can form good approximations of high-rank Hessians
+  using only a small number of operator applications.  The proposed
+  method is very interesting and makes a great contribution to solving
+  large-scale inverse problems with high-rank Hessians. The
+  presentation of the proposed method is concise and illustrative. The
+  numerical results are very convincing. Overall, the paper is very
+  well written.\\}
+
+\paragraph{General Comments:}
+
+\referee{I recommend to accept the paper for publication with minor
+  revisions. I suggest the authors properly address the following
+  questions to help for its better understanding and broader
+  applications.\\}
 
 \begin{enumerate} 
-\item \referee{I do not think the terms ``Gaussianization, Gaussianized, ...'' are properly utilized in this work.
-The term Gaussianization has been actively used, especially in the machine learning literature, to refer to transformation of non-Gaussian data/variable in to a Gaussian.
-This requires finding a map/transformation between the original distribution and a Gaussian, e.g., an optimal transport, or normalizing flows.
-The authors, however, use these terms to denote approximating a non-Gaussian distribution with a localized Gaussian, e.g., following a Laplacian approach.}
+\item \referee{The author mentioned one limitation of the proposed method:
+      the Hessian should have a local non-negative integral kernel,
+      which is not satisfied for wave inverse problems that lead to a
+      substantial amount of negative entries of the Hessian. It would
+      be interesting if the author could elaborate more details with
+      intuition or numerical evidence for this property and the
+      limitation, especially if you also use Gauss-Newton approximate
+      Hessian for the wave inverse problems as in this paper.}
 
-\noindent{We appreciate drawing our attention to the fact that our usage of ``Gaussianization, Gaussianized...'', would be unclear to many readers.
-We have updated our usage to ``the Laplace approximation of the posterior''.}
+\noindent{To do.}
 
-\item \referee{Re: Algorithm 2 (page 22); it would be great to discuss how $r_1, \ldots, r_{L}$ can be auto tuned.
-What metric can be used to automatically choose the optimal value of $L$ and then choose the rank for each level automatically.}
+\item \referee{It is interesting to see the effectiveness of the
+  approximate Hessian used as a preconditioner. It would also be very
+  interesting to see how accurate the approximate Hessian is compared
+  to a full Hessian in terms of the number of batches, ellipsoid
+  sizes, and in particular the total number of impulse responses. Can
+  you say anything about the convergence of the approximation, either
+  numerical or theoretical?}
 
-\noindent{Done, with regard to choosing the ranks $r_{1},\dots,r_{L}$.
-We are not aware of an automatic way to choose the optimal value of $L$ but have described the components that we believe could be used to develop such means.}
+  \noindent{To do.}
+
+ \item \referee{The paper presented a nice complexity analysis of the steps in
+    constructing the approximate Hessian, with the dominating cost
+    arising from the Hessian matrix-vector product. It would be
+    interesting to see the computational time of each step for the
+    numerical examples to provide a good sense of the complexity and
+    to support the 5-10X computational reduction by accounting also
+    the overhead beyond the PDE solves.}
+
+   \noindent{To do.}
+
 \end{enumerate}
 
-\paragraph{\sc Other comments:}
-\begin{enumerate}
-\item  \referee{Page 1, Line 25: ``approximation of Hessian applies...''; Hessian of what?}
+%\bibliographystyle{unsrt}
+%\bibliographystyle{iopart-num}
+%\bibliography{reviewer2_references}
 
-\noindent{We now use the more descriptive ``discretized objective functional Hessian vector-products''.}
-
-\item \referee{Page 1, Line 28--31: statement too long, maybe try splitting!}
-
-\noindent{Done.}
-
-\item \referee{Page 1, Line 29: ``sampling from Gaussian proposal distributions'' $\rightarrow$ ``sampling Gaussian distributions''.}
-
-\noindent{We have not used the exact suggested text since the sentence phrasing has been changed.
-We believe that ``...that provide a means to sample Gaussian distributions...'', is as close to the suggested rephrasing but which maintains consistency with the new sentence structure.}
-
-\item \referee{Page 1, Line 30: I am not sure I understand what the authors mean by ``Newton solves by direct linear methods''!}
-
-\noindent{We had intended to describe methods such as LU factorization, which is known as a direct method of linear system solution.
-We hope that by referencing a relevant factorization that the message intended to be communicated is now made more clear.}
-
-\item \referee{Page 1, Line 46: ``to inverse problems'' ``with inverse problems''.}
-
-\noindent{Done, thanks!}
-
-\item \referee{Page 2, Line 7: simulation... systems plays'' $\rightarrow$ ``simulations ... systems play''.}
-
-\noindent{Done.}
-
-\item \referee{Page 3, Line 17: ``generate approximate samples from a Gaussian posterior distribution for...'' $\rightarrow$ ``generate samples from an approximate Gaussian posterior for...''}
-
-\noindent{Done, but we have kept ``Gaussian posterior distribution for...''}
-
-\item \referee{Page 3, Line 22: ``including for a Greenland ice...'' $\rightarrow$ ``including a Greenland ice...''}
-
-\noindent{Done.}
-
-\item \referee{Page 3, Line 28: Add a reference; ``In this section,'' $\rightarrow$ ``In this section (2.1),''}
-
-\noindent{Done, but with the capitalization scheme used throughout the manuscript, ``In this Section (2.1),''}
-
-\item \referee{Page 3, line 44: Remove ``i.e., the data,'' as it is redundant.}
-
-\noindent{Done. We have added an additional sentence in which we explicitly state that we will later refer to $\bs{d}$ as the data instead of the more verbose observational data.}
-
-\item \referee{Page 3, Line 46--47: ``and mathematically'' $\rightarrow$ ``and is mathematically''}
-
-\noindent{Done.}
-
-\item \referee{Page 4, Lines 12--13: The authors use the term ``parameter-to-PDE-solution map'' which has not been introduced.
-Only the combined ``parameter-to-observable map'' has been introduced on line $10$ above.
-I suggest being consistent and either explicitly decompose the parameter-to-observable-map into a solution operator (solution-to-PDE-solution) and an observation operator, or just restrict usage to ``parameter-to-observable-map''.
-Either one or both of the solution operator and the observation operator can be nonlinear and each is associated with its own difficulties.}
-
-\noindent{Done. We have restricted usage to the ``parameter-to-observable map''.}
-
-\item \referee{Page 4, Line 27: ``it's gradient'' $\rightarrow$ ``its gradient''}
-
-\noindent{Done.}
-
-\item \referee{Page 4, Lines 28--33: This paragraph requires some
-  modification. If you can well-approximate the posterior you do not
-  necessarily need MCMC; you can even use naive MC by applying MH step
-  to samples from this Gaussian. MCMC does not generally require
-  approximating the posterior with a Gaussian, however sampling a
-  Gaussian approximation can be useful. For example in Hamiltonian
-  Monte Carlo (HMC), samples from a Gaussian approximation can be used
-  to build a mass matrix that optimizes the performance of the
-  sampler; diagonal of the mass matrix are optimally set to posterior
-  precisions (or approximations thereof). See e.g., reference
-  \cite{neal2011mcmc, attia2017reduced}.}
-
-\noindent{We agree with the reviewer, but our experience is that using
-  MCMC is still superior to using MH (or importance sampling)
-  directly, in particular in high dimensions.  We adjusted this
-  paragraph in the revised manuscript.}
-
-%\todo{Georg, please see if you you'd like to add or change anything here.}
-
-\item \referee{Page 4, Line 38: define $_{\ell_{2}}$ presented in the previous equation}
-
-\noindent{This item has been addressed by removing any usage of $_{\ell_{2}}$.}
-
-\item \referee{Page 5, Line 38: Maybe redefine paragraph titles to add a period at the end; ``Compression'' $\rightarrow$ ``Compression.''}
-
-\noindent{Done.}
-
-\item \referee{Page 5, Line 50: replace the comma with a semicolon.}
-
-\noindent{Done.}
-
-\item \referee{Page 5, Line 60: ``algorithm is made possible''; where? add a reference/citation here!}
-
-\noindent{Done.}
-
-\item \referee{Page 6, Line 5: a period is missing after ``Approximations''}
-
-\noindent{Done.}
-
-\item \referee{Page 6, Line 6: hereafter use a different symbol for the number of oversampling vectors as $\bs{d}$ has been utilized to denote the data, despite being boldface.}
-
-\noindent{Done. We are now using the symbol $q$ to denote the number of oversampling vectors.}
-
-\item \referee{Page 6, Lines 24--30: this sentence is too long; try splitting.}
-
-\noindent{We have reformulated this sentence.}
-
-\item \referee{Page 7, Line 40: the term data-misfit Hessian has already been introduced before; remove ``i.e., the Hessian ... functional'' as it is redundant.}
-
-\noindent{Done.}
-
-\item \referee{Page 7, Line 44: remove ``of it''}
-
-\noindent{Done.}
-
-\item \referee{Page 8, Line 40: ``basis function'' $\rightarrow$ ``basis functions''}
-
-\noindent{Done.}
-
-\item \referee{Page 8, Line 51: please see my comment above about using the term ``Gaussianized''}
-
-\noindent{Noted and addressed.}
-
-\item \referee{Page 9, Line 22: I am not sure what ``means of applying'' here means!}
-
-\noindent{We have clarified the vague ``means of applying'' with ``means of computing square root and inverse square root matrix-vector products.}
-
-\item \referee{Page 9, Lines 33--40: this paragraph is redundant. This material/argument has already been introduced in Section 2.1.}
-
-\noindent{We believe that this material is important as it helps show how the example problem fits into the larger class of inverse problems, therefore we decided to keep it here.}
-
-%\todo{I (Tucker) believe that this argument could be strengthened.}
-
-\item \referee{Page 9, Line 49: ``can mathematically'' $\rightarrow$ ``can be mathematically''}
-
-\noindent{Done.}
-
-\item \referee{Page 10 I suggest using subequations for the model Equations (6)--(9)}
-
-Done.
-
-\item \referee{Page 10, Lines 16--19, 22, 43: in reference to model equations, use $\backslash$eqref instead of $\backslash$ref}
-
-Done.
-
-\item \referee{Page 11, Line 5: see my comment above about using the term ``Gaussianizaiton''. I will not comment on this hereafter.}
-
-\noindent{Noted and addressed.}
-
-\item \referee{Page 11, Lines 21, 28: use $\backslash$eqref in reference to model equations.}
-
-Done.
-
-\item \referee{Page 11, Line 31: ``add $1\%$ relative Gaussian noise...'': $1\%$ of what? mean true observation magnitude, $\ell_{\infty}$, ...etc.; please clarify!}
-
-\noindent{Done.}
-
-\item \referee{Page 11, Line 57: I do not believe ``approximate samples'' is a correct terminology. I suggest using ``samples from the approximate posterior''.}
-
-\noindent{This has been addressed.}
-
-\item \referee{Page 12, Figure 3 caption: something is not quite right here.
-The caption states that circles show representative (random measurement locations), however, lines 29-30 on page 11 say that observations are uniformly distributed.
-What am I missing here?
-}
-
-\noindent{The figure has been adjusted so that the measurement locations are now illustrated as being uniformly distributed.}
-
-\item \referee{Page 14, Line 57: there is an extra comma after ``more data''}
-
-\noindent{Done.}
-
-\item \referee{Page 15, Figure 7 caption (line 21): ``low-rank (LR)'' $\rightarrow$ ``LR'' as the acronym has already been introduced; same on line 24 of the same caption: ``low-rank'' $\rightarrow$ ``LR'' }
-
-\noindent{Done.}
-
-\item \referee{Page 15, Figure 8 caption: same as the comment above on using ``LR''}
-
-\noindent{Done.}
-
-\item \referee{Page 16, Line 5: remove the comma after ``Albany'', and remove ``so-called''}
-
-\noindent{Done.}
-
-\item \referee{Page 16, Line 30: use $\backslash$eqref in reference to Equation (3)}
-
-Done.
-
-\item \referee{Page 20, Line 29: remove the comma before ``associated to finite...''}
-
-\noindent{Done.}
-
-\item \referee{Page 23, Line  5: ``low-rank'' $\rightarrow$ ``LR'' }
-
-\noindent{Done.}
-
-\end{enumerate}
-%\bibliographystyle{unsrt}
-\bibliographystyle{iopart-num}
-\bibliography{reviewer2_references}
-
-%\noindent [1]- Neal, Radford M. "MCMC using Hamiltonian dynamics." Handbook of markov chain monte carlo 2, no. 11 (2011): 2.
-
-%\noindent [2]- Attia, Ahmed, Razvan Stefanescu, and Adrian Sandu.
-"The reduced-order hybrid Monte Carlo sampling smoother." International Journal for Numerical Methods in Fluids 83, no.
-1 (2017): 28-51.
 \end{document}

### 2024-01-05 revision/reply_to_reviewer2.tex
aaaa

--- 2024-01-05 revision/reply_to_reviewer1.tex
@@
 \usepackage{amsmath}
+\usepackage{hyperref}
 
@@
 
-\section*{Reply to the Reviewer 1:}
+\section*{Reply to the Reviewer 2:}
 
@@
 
-\referee{Many Hessian-based computational algorithms have been
-  developed to solve large-scale inverse problems constrained by
-  partial differential equations (PDEs) in the last decade, especially
-  in the context of inferring infinite-dimensional
-  parameters. However, it remains a critical computational challenge
-  to solve such inverse problems when the Hessian of the data-misfit
-  term does not have fast spectrum decay, or the Hessian is
-  high-rank. This challenge often leads to the need to solve a
-  considerable number of PDEs in many (preconditioned) CG iterations
-  with an inexact Newton-CG algorithm, one of the most advanced
-  optimization algorithms in solving infinite-dimensional inverse
-  problems.\\}
-
-  \referee{To address this challenge, this paper proposes an
-  efficient preconditioned CG method with the preconditioner computed
-  as an approximate Hessian at some suitable parameter point, which is
-  demonstrated to significantly reduce the number of required CG
-  iterations to achieve given accuracy. The key novelty of this paper
-  is the development of an efficient hierarchical matrix approximation
-  of the Hessian using only a small number of Hessian matrix-vector
-  products, thus a small number of PDE solves. This novelty is made
-  possible by exploiting the property of the Hessian operator with
-  locally supported non-negative integral kernels. Specifically, the
-  authors propose a point spread function (PSF) approximation of the
-  Hessian by (1) computing the zeroth, first, and second-order
-  impulse response moments of the Hessian by applying its product with
-  constant, linear, and quadratic functions, (2) building local
-  ellipsoid support es- timate of the impulse response functions of
-  the Hessian based on these moments, (3) selecting sample points of
-  the impulse response from a candidate set by a greedy ellipsoid
-  packing algorithm, (4) computing the impulse responses with disjoint
-  ellipsoid in batches by applying the Hessian to Dirac combs, which
-  plays a key role in reducing the total number of Hessian
-  matrix-vector product, (5) approximating any integral kernel entries
-  by a radial basis interpolation based on the computed impulse
-  responses, and (6) building the hierarchical matrix approximation of
-  the Hessian with the radial basis interpolation of the integral
-  kernel entries before applying proper symmetrizing and flipping of
-  the negative eigenvalues to make the approximate matrix symmetric
-  positive semi-definite.\\}
-
-\referee{The PSF approximation in this paper is built for more general
-  operators than Hessian. It demonstrates the effectiveness of the
-  PSF-based method by 1 using it to build preconditioners for the
-  Hessian operator in two challenging inverse problems of basal
-  friction coefficient inversion of ice sheet flow and ini- tial
-  condition inversion of advection-dominated transport. It shows
-  significant reductions (5-10X) in the required number of PDE solves
-  compared to classical regularization-based preconditioning and no
-  preconditioning. The paper also presents a comprehensive numerical
-  study on the influence of various parameters (data noises, \# batchs,
-  diffusion coefficients, terminal times) on the effective- ness and
-  data-scalability of the proposed method, showing that the PSF-based
-  preconditioners can form good approximations of high-rank Hessians
-  using only a small number of operator applications.  The proposed
-  method is very interesting and makes a great contribution to solving
-  large-scale inverse problems with high-rank Hessians. The
-  presentation of the proposed method is concise and illustrative. The
-  numerical results are very convincing. Overall, the paper is very
-  well written.\\}
-
-\paragraph{General Comments:}
-
-\referee{I recommend to accept the paper for publication with minor
-  revisions. I suggest the authors properly address the following
-  questions to help for its better understanding and broader
-  applications.\\}
+\referee{I have read the paper with pleasure and have the following
+  suggestions for improvement:\\}
 
 \begin{enumerate}
-\item \referee{The author mentioned one limitation of the proposed method:
-      the Hessian should have a local non-negative integral kernel,
-      which is not satisfied for wave inverse problems that lead to a
-      substantial amount of negative entries of the Hessian. It would
-      be interesting if the author could elaborate more details with
-      intuition or numerical evidence for this property and the
-      limitation, especially if you also use Gauss-Newton approximate
-      Hessian for the wave inverse problems as in this paper.}
+\item \referee{Interesting work, generally well-written. Perhaps a
+  small illustrative example with a few figures at the beginning could
+  help understand the procedure intuitively.}
 
@@
 
-\item \referee{It is interesting to see the effectiveness of the
-  approximate Hessian used as a preconditioner. It would also be very
-  interesting to see how accurate the approximate Hessian is compared
-  to a full Hessian in terms of the number of batches, ellipsoid
-  sizes, and in particular the total number of impulse responses. Can
-  you say anything about the convergence of the approximation, either
-  numerical or theoretical?}
+\item \referee{possibly interesting references for estimating Hessians
+  in linearised seismic imaging include:
+  \url{https://doi.org/10.1016/j.acha.2007.06.007},
+  \url{https://doi.org/10.1190/1.2836323}}
 
@@
 
- \item \referee{The paper presented a nice complexity analysis of the steps in
-    constructing the approximate Hessian, with the dominating cost
-    arising from the Hessian matrix-vector product. It would be
-    interesting to see the computational time of each step for the
-    numerical examples to provide a good sense of the complexity and
-    to support the 5-10X computational reduction by accounting also
-    the overhead beyond the PDE solves.}
+\item \referee{How about comparison with other methods, like those
+  proposed in [7],[52]? and you say a bit more? it is hard to conclude
+  that your method will always outperform these methods.}
+
+    \noindent{To do.}
+
+\item \referee{Perhaps it is illustrative to include some 1D blurring
+  examples to illustrate robustness of method to departure from
+  assumptions (positivity, locality) e.g. on 'indefinite' kernels or
+  kernels with multiple modes.}
+
+  \noindent{To do.}
+
+ \item \referee{is code for applying the developed methods on other
+   inverse problems available?}
 

### 2024-01-08 revision/reply_to_reviewer1.tex
started response files

@@
 
+\referee{Thank you for the positive assesment of our work.\\}
+
 \paragraph{General Comments:}
@@
 
-\noindent{To do.}
+\noindent{Nick will try to get an impulse response for the wave
+  problem.}
 
@@
 
-  \noindent{To do.}
+  \noindent{Nick will try to create a plot that shows some error
+    (e.g. Frob norm, cond number) as a function of number of impulse
+    responses.}
 
@@
 
-   \noindent{To do.}
+   \noindent{Nick will time one of the examples in the
+     advection-diffusion.}
 

### 2024-01-08 revision/reply_to_reviewer2.tex
started response files

@@
 
-  \noindent{To do.}
+  \noindent{Noemi will give it a try.}
 
@@
 
-    \noindent{To do.}
+    \noindent{For ref 7: Ask Tucker if we could do HODL on the Hessian
+      coming from ice mountain problem and record the number of
+      mat-vecs needed to achieve a certain accuracy, something like
+      Figure 8 in teh HODLR paper. Nick will do some estimates to
+      compare things with ref 52.}
 
@@
 
-  \noindent{To do.}
+  \noindent{Maybe we can answer with the wave problem discussion from
+    reviewer 1. or even cook up a 2D deblurring problem, such that the
+    Hessian entries will be some modified Gaussians.}
 

### 2024-01-08 revision/reply_to_reviewer2.tex
started response files

@@
 
-  \noindent{Maybe we can answer with the wave problem discussion from
-    reviewer 1. or even cook up a 2D deblurring problem, such that the
-    Hessian entries will be some modified Gaussians.}
+  \noindent{Use a 2D-deblurring problem and cook up a Hessian with
+    entries given by be some modified Gaussian and show that the error
+    increases as more negative entires are present in the kernel.}
 
@@
 
-  \noindent{To do.}
+  \noindent{Ideally we would like to release the code, however this
+    will take time. In the meantime, we will make the code available
+    upon request.}
 

### 2024-01-15 revision/reply_to_reviewer2.tex
minor

@@
 
-    \noindent{For ref 7: Ask Tucker if we could do HODL on the Hessian
-      coming from ice mountain problem and record the number of
-      mat-vecs needed to achieve a certain accuracy, something like
+    \noindent{For ref 7: Ask Tucker if we could do HODLR on the
+      Hessian coming from ice mountain problem and record the number
+      of mat-vecs needed to achieve a certain accuracy, something like
       Figure 8 in teh HODLR paper. Nick will do some estimates to

### 2024-01-15 paper_sisc_first_revision/localpsf.tex
makefile works
Addressed reviewer 2 point 2

--- 2023-07-06 paper/localpsf.tex
@@
 The results show that the PSF-based preconditioners are able to form good approximations of high-rank Hessians using only a small number of operator applications.
+
+	TEST ASDF
 \end{abstract}
@@
 By using this two stage approach, we reduce the number of operator applications to a few dozen at most.
-One limitation of our method is that not all data misfit Hessians satisfy the local non-negative integral kernel properties (for example, the wave inverse problem data misfit Hessian has a substantial amount of negative entries in its integral kernel).
-However, many data misfit Hessians of practical interest do satisfy these properties (either exactly or approximately), and the proposed method is targeted at approximating these Hessians.
+
+One limitation of our method is that not all data misfit Hessians satisfy the local non-negative integral kernel properties.
+We note, in particular, that the wave inverse problem data misfit Hessian has a substantial proportion of negative entries in its integral kernel.
+In this case more specialized techniques based on pseudodifferential operator theory have been developed, e.g., \cite{DemanetEtAl12,HerrmannMoghaddamStolk08,Symes08}.
+However, many data misfit Hessians of practical interest do satisfy the local non-negative integral kernel properties (either exactly or approximately), and the proposed method is targeted at approximating these Hessians.
 

### 2024-01-24 numerical_examples/blur/blur_example.py
blur example working ok

@@ -1,0 +1,136 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import dolfin as dl
+import localpsf.localpsf_cg1_lumped as lpsf1
+
+
+nx = 50
+ny = 50
+mesh = dl.UnitSquareMesh(nx, ny)
+Vh = dl.FunctionSpace(mesh, 'CG', 1)
+dof_coords = Vh.tabulate_dof_coordinates()
+#
+# xx_linear = np.linspace(-1.0, 1.0, nx)
+# yy_linear = np.linspace(-1.0, 1.0, ny)
+#
+# X, Y = np.meshgrid(xx_linear, yy_linear)
+# N = len(X.reshape(-1))
+#
+# plt.figure()
+# plt.scatter(X.reshape(-1), Y.reshape(-1))
+
+# Ricker-type wavelet, https://en.wikipedia.org/wiki/Ricker_wavelet
+def ricker(
+    xx: np.ndarray, # shape=(N, 2), N=nx*ny
+    vol: float,
+    mu: np.ndarray, # shape=(2,)
+    Sigma: np.ndarray, # shape=(2, 2)
+    a: float # how much negative to include. a=0: gaussian
+) -> np.ndarray: # shape=(N,)
+    assert(xx.shape[1] == 2)
+    N = xx.shape[0]
+    assert(Sigma.shape == (2, 2))
+    assert(mu.shape == (2,))
+
+    pp = xx - mu.reshape((1,2))
+    inv_Sigma = np.linalg.inv(Sigma)
+    t_squared_over_sigma_squared = np.einsum('ai,ai->a', pp, np.einsum('ij,aj->ai', inv_Sigma, pp))
+    G = vol / (2.0 * np.pi * np.sqrt(np.linalg.det(Sigma))) * np.exp(-0.5 * t_squared_over_sigma_squared)
+
+    return (1.0 - a * t_squared_over_sigma_squared) * G
+
+# xx = np.vstack([X.reshape(-1), Y.reshape(-1)]).T
+
+vol = 1.0
+mu = np.array([0.5, 0.5])
+Sigma = np.array([[0.01, 0.0], [0.0, 0.0025]])
+a = 0.3
+
+rr = ricker(dof_coords, vol, mu, Sigma, a)
+
+f1 = dl.Function(Vh)
+f1.vector()[:] = rr
[…]

### 2024-01-24 numerical_examples/stokes/preconditioned_spectrum.py

@@ -16,8 +16,9 @@ import scipy.linalg as sla
 import scipy.sparse.linalg as spla
 import scipy.sparse as sps
 
-save_data = True
-save_figures = True
+save_data = False
+save_figures = False
+save_dense_misfit_and_regularization_Hessian = True
 
 stokes_base_dir = get_project_root() / 'numerical_examples' / 'stokes'
 
@@ -165,6 +166,11 @@ if use_dense:
     ee_reg0, _ = sla.eigh(H_dense - R_dense, R_dense)
     ee_reg = ee_reg0[-num_eigs:]
     abs_ee_reg = np.abs(ee_reg)[::-1]
+
+    if save_dense_misfit_and_regularization_Hessian:
+        np.savetxt(str(save_dir) + '/H_dense.dat', H_dense)
+        np.savetxt(str(save_dir) + '/R_dense.dat', R_dense)
+
 else:
     Rinv_linop = spla.LinearOperator((StokesIP.N, StokesIP.N), matvec = lambda x: StokesIP.apply_Rinv_numpy(x))
 

### 2024-01-31 revision/reply_to_reviewer1.tex
minor
frog example is good

--- 2024-01-05 revision/reply_to_reviewer1.tex
@@
 
+\noindent{Thank you for the positive assesment of our work.\\}
+
 \paragraph{General Comments:}
@@
 
-\noindent{To do.}
+\noindent{(1) Three-figure plot: Stokes, Advection/diffusion, wave impulse responses.
+
+(2) showing that the method performs poorly for the ricker example
+
+Nick will try to get an impulse response for the wave
+  problem.}
 
@@
 
-  \noindent{To do.}
+  \noindent{
+  	Frog example with a=1. Error vs #batches, for different tau. Also do a plot with #impulses on the x-axis.
+
+  	Nick will try to create a plot that shows some error
+    (e.g. Frob norm, cond number) as a function of number of impulse
+    responses.}
 

### 2024-01-31 revision/reply_to_reviewer2.tex
Addressed reviewer 2 point 2
frog example is good

@@
 
-  \noindent{To do.}
+  \noindent{Frog example with a=1.
+  	Structure: local non-negative kernel
+(0) Matrix itself column highlighted, show impulse response
+(1) moments. Show one impulse with ellipsoid with widths illustrated
+(2) impulse response batches
+(3) RBF interpolation (?)
+(4) H-matrix
+}
 
@@
 
-  \noindent{Noemi will give it a try.}
+  \noindent{We added text in the paragraph at the end of Section 2 to discuss the (lack of) applicability of the method to wave inverse problems and added these references.}
 

### 2024-01-31 numerical_examples/blur/blur_example.py

@@ -2,6 +2,7 @@ import numpy as np
 import matplotlib.pyplot as plt
 import dolfin as dl
 import localpsf.localpsf_cg1_lumped as lpsf1
+import scipy.linalg as sla
 
 
 nx = 50
@@ -9,15 +10,6 @@ ny = 50
 mesh = dl.UnitSquareMesh(nx, ny)
 Vh = dl.FunctionSpace(mesh, 'CG', 1)
 dof_coords = Vh.tabulate_dof_coordinates()
-#
-# xx_linear = np.linspace(-1.0, 1.0, nx)
-# yy_linear = np.linspace(-1.0, 1.0, ny)
-#
-# X, Y = np.meshgrid(xx_linear, yy_linear)
-# N = len(X.reshape(-1))
-#
-# plt.figure()
-# plt.scatter(X.reshape(-1), Y.reshape(-1))
 
 # Ricker-type wavelet, https://en.wikipedia.org/wiki/Ricker_wavelet
 def ricker(
@@ -37,56 +29,128 @@ def ricker(
     t_squared_over_sigma_squared = np.einsum('ai,ai->a', pp, np.einsum('ij,aj->ai', inv_Sigma, pp))
     G = vol / (2.0 * np.pi * np.sqrt(np.linalg.det(Sigma))) * np.exp(-0.5 * t_squared_over_sigma_squared)
 
-    return (1.0 - a * t_squared_over_sigma_squared) * G
+    # cos_t_over_sigma = np.cos(np.sqrt(t_squared_over_sigma_squared))
+    cos_x_over_sigma = np.cos(pp[:,0] / (np.sqrt(Sigma[0,0]) / 2.0))
+    sin_y_over_sigma = np.sin(pp[:,1] / (np.sqrt(Sigma[1,1]) / 2.0))
+    return (1.0 + a * cos_x_over_sigma * sin_y_over_sigma) * G
+    # return (1.0 + a * cos_x_over_sigma) * G
+    # return cos_x_over_sigma * G
+    # return cos_t_over_sigma * G
+    # return (1.0 + a * cos_t_over_sigma) * G
+    # return (1.0 - a * t_squared_over_sigma_squared) * G
 
 # xx = np.vstack([X.reshape(-1), Y.reshape(-1)]).T
 
+length_scaling = 0.25 # <--- Tucker: experiment with this
 vol = 1.0
[…]

### 2024-02-05 revision/reply_to_reviewer2.tex
frog example is good

@@
 (2) impulse response batches
-(3) RBF interpolation (?)
+(3) RBF interpolation (?), evaluation of entries. Columns we know, and entry we don't know. 
 (4) H-matrix

### 2024-02-05 numerical_examples/blur/blur_example.py

@@ -5,8 +5,8 @@ import localpsf.localpsf_cg1_lumped as lpsf1
 import scipy.linalg as sla
 
 
-nx = 50
-ny = 50
+nx = 63
+ny = 63
 mesh = dl.UnitSquareMesh(nx, ny)
 Vh = dl.FunctionSpace(mesh, 'CG', 1)
 dof_coords = Vh.tabulate_dof_coordinates()
@@ -29,114 +29,71 @@ def ricker(
     t_squared_over_sigma_squared = np.einsum('ai,ai->a', pp, np.einsum('ij,aj->ai', inv_Sigma, pp))
     G = vol / (2.0 * np.pi * np.sqrt(np.linalg.det(Sigma))) * np.exp(-0.5 * t_squared_over_sigma_squared)
 
-    # cos_t_over_sigma = np.cos(np.sqrt(t_squared_over_sigma_squared))
     cos_x_over_sigma = np.cos(pp[:,0] / (np.sqrt(Sigma[0,0]) / 2.0))
     sin_y_over_sigma = np.sin(pp[:,1] / (np.sqrt(Sigma[1,1]) / 2.0))
     return (1.0 + a * cos_x_over_sigma * sin_y_over_sigma) * G
-    # return (1.0 + a * cos_x_over_sigma) * G
-    # return cos_x_over_sigma * G
-    # return cos_t_over_sigma * G
-    # return (1.0 + a * cos_t_over_sigma) * G
-    # return (1.0 - a * t_squared_over_sigma_squared) * G
 
-# xx = np.vstack([X.reshape(-1), Y.reshape(-1)]).T
 
-length_scaling = 0.25 # <--- Tucker: experiment with this
+length_scaling = 1.0 # <--- Tucker: experiment with this
 vol = 1.0
 mu = np.array([0.5, 0.5])
 Sigma = length_scaling * np.array([[0.01, 0.0], [0.0, 0.0025]]) # 0.25*np.array([[0.01, 0.0], [0.0, 0.0025]])
-a_small = 0.0
-a_medium = 1.1 # 0.1
-a_large = 5.0 # <--- Tucker: use a=1
+a = 1.0
 
-rr_small = ricker(dof_coords, vol, mu, Sigma, a_small)
-rr_medium = ricker(dof_coords, vol, mu, Sigma, a_medium)
[…]

### 2024-02-05 numerical_examples/blur/blur_example.py
add kernel and impulse illustration

@@ -4,6 +4,7 @@ import dolfin as dl
 import localpsf.localpsf_cg1_lumped as lpsf1
 import scipy.linalg as sla
 
+save_figures = False
 
 nx = 63
 ny = 63
@@ -24,7 +25,16 @@ def ricker(
     assert(Sigma.shape == (2, 2))
     assert(mu.shape == (2,))
 
-    pp = xx - mu.reshape((1,2))
+    rotation_center = np.array([0.5, 0.5])
+    radial_vector = (mu - rotation_center)
+    theta = 4.0 * np.arctan(radial_vector[1] / radial_vector[0])
+    # xprime = (mu - rotation_center) / np.linalg.norm(mu - rotation_center)
+    # yprime = np.array([xprime[1], -xprime[0]])
+    # Rot_matrix = np.vstack([xprime, yprime]).T
+    # theta = (np.pi / 2.0) * mu[0] * mu[1]
+    Rot_matrix = np.array([[np.cos(theta), -np.sin(theta)], [np.sin(theta), np.cos(theta)]])
+
+    pp = (xx - mu.reshape((1,2))) @ Rot_matrix.T #@ Rot_matrix.T
     inv_Sigma = np.linalg.inv(Sigma)
     t_squared_over_sigma_squared = np.einsum('ai,ai->a', pp, np.einsum('ij,aj->ai', inv_Sigma, pp))
     G = vol / (2.0 * np.pi * np.sqrt(np.linalg.det(Sigma))) * np.exp(-0.5 * t_squared_over_sigma_squared)
@@ -65,48 +75,75 @@ for ii in range(Vh.dim()):
     Ker[:,ii] = ricker(dof_coords, vol, dof_coords[ii,:], Sigma, a)
 
 
-p1 = np.array([0.2, 0.55])
-p2 = np.array([0.5, 0.65])
+# p1 = np.array([0.2, 0.55])
+# p2 = np.array([0.5, 0.65])
+# pp = [p1, p2]
+pp = [
+    np.array([0.2, 0.2]),
+    np.array([0.2, 0.4]),
+    np.array([0.2, 0.6]),
+    np.array([0.2, 0.8]),
+    np.array([0.4, 0.2]),
+    np.array([0.4, 0.4]),
+    np.array([0.4, 0.6]),
+    np.array([0.4, 0.8]),
+    np.array([0.6, 0.2]),
+    np.array([0.6, 0.4]),
[…]

### 2024-02-06 numerical_examples/blur/blur_example.py
unclear how to illustrate h-matrix stuff

@@ -1,19 +1,21 @@
 import numpy as np
 import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
 import dolfin as dl
 import localpsf.localpsf_cg1_lumped as lpsf1
 import scipy.linalg as sla
+from nalger_helper_functions import plot_ellipse
 
-save_figures = False
+save_figures = True
 
-nx = 63
-ny = 63
+nx = 89
+ny = 89
 mesh = dl.UnitSquareMesh(nx, ny)
 Vh = dl.FunctionSpace(mesh, 'CG', 1)
 dof_coords = Vh.tabulate_dof_coordinates()
 
-# Ricker-type wavelet, https://en.wikipedia.org/wiki/Ricker_wavelet
-def ricker(
+# Frog function
+def frog_function(
     xx: np.ndarray, # shape=(N, 2), N=nx*ny
     vol: float,
     mu: np.ndarray, # shape=(2,)
@@ -25,13 +27,7 @@ def ricker(
     assert(Sigma.shape == (2, 2))
     assert(mu.shape == (2,))
 
-    rotation_center = np.array([0.5, 0.5])
-    radial_vector = (mu - rotation_center)
-    theta = 4.0 * np.arctan(radial_vector[1] / radial_vector[0])
-    # xprime = (mu - rotation_center) / np.linalg.norm(mu - rotation_center)
-    # yprime = np.array([xprime[1], -xprime[0]])
-    # Rot_matrix = np.vstack([xprime, yprime]).T
-    # theta = (np.pi / 2.0) * mu[0] * mu[1]
+    theta = np.pi * (mu[0] + mu[1])
     Rot_matrix = np.array([[np.cos(theta), -np.sin(theta)], [np.sin(theta), np.cos(theta)]])
 
     pp = (xx - mu.reshape((1,2))) @ Rot_matrix.T #@ Rot_matrix.T
@@ -50,7 +46,7 @@ mu = np.array([0.5, 0.5])
 Sigma = length_scaling * np.array([[0.01, 0.0], [0.0, 0.0025]]) # 0.25*np.array([[0.01, 0.0], [0.0, 0.0025]])
 a = 1.0
 
-rr = ricker(dof_coords, vol, mu, Sigma, a)
+rr = frog_function(dof_coords, vol, mu, Sigma, a)
 
 f1 = dl.Function(Vh)
 f1.vector()[:] = rr
[…]

### 2024-02-06 numerical_examples/blur/blur_impulse_batch_pictures.py

@@ -1,0 +1,156 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import dolfin as dl
+import localpsf.localpsf_cg1_lumped as lpsf1
+import scipy.linalg as sla
+from nalger_helper_functions import plot_ellipse
+import hlibpro_python_wrapper as hpro
+
+save_figures = True
+
+nx = 63
+ny = 63
+mesh = dl.UnitSquareMesh(nx, ny)
+Vh = dl.FunctionSpace(mesh, 'CG', 1)
+dof_coords = Vh.tabulate_dof_coordinates()
+
+# Frog function
+def frog_function(
+    xx: np.ndarray, # shape=(N, 2), N=nx*ny
+    vol: float,
+    mu: np.ndarray, # shape=(2,)
+    Sigma: np.ndarray, # shape=(2, 2)
+    a: float # how much negative to include. a=0: gaussian
+) -> np.ndarray: # shape=(N,)
+    assert(xx.shape[1] == 2)
+    N = xx.shape[0]
+    assert(Sigma.shape == (2, 2))
+    assert(mu.shape == (2,))
+
+    theta = np.pi * (mu[0] + mu[1])
+    Rot_matrix = np.array([[np.cos(theta), -np.sin(theta)], [np.sin(theta), np.cos(theta)]])
+
+    pp = (xx - mu.reshape((1,2))) @ Rot_matrix.T #@ Rot_matrix.T
+    inv_Sigma = np.linalg.inv(Sigma)
+    t_squared_over_sigma_squared = np.einsum('ai,ai->a', pp, np.einsum('ij,aj->ai', inv_Sigma, pp))
+    G = vol / (2.0 * np.pi * np.sqrt(np.linalg.det(Sigma))) * np.exp(-0.5 * t_squared_over_sigma_squared)
+
+    cos_x_over_sigma = np.cos(pp[:,0] / (np.sqrt(Sigma[0,0]) / 2.0))
+    sin_y_over_sigma = np.sin(pp[:,1] / (np.sqrt(Sigma[1,1]) / 2.0))
+    return (1.0 + a * cos_x_over_sigma * sin_y_over_sigma) * G - 1e-12 # 1e-12: undo hack workaround for hlibpro bug
+
+
+length_scaling = 0.5
+vol = 1.0
+mu = np.array([0.5, 0.5])
[…]

### 2024-02-07 numerical_examples/blur/blur_example.py
making interpolation figure

@@ -16,17 +16,19 @@ def frog_function(
     vol: float,
     mu: np.ndarray, # shape=(2,)
     Sigma: np.ndarray, # shape=(2, 2)
-    a: float # how much negative to include. a=0: gaussian
+    a: float, # how much negative to include. a=0: gaussian
+    use_bump: bool = True,
 ) -> np.ndarray: # shape=(N,)
     assert(xx.shape[1] == 2)
     N = xx.shape[0]
     assert(Sigma.shape == (2, 2))
     assert(mu.shape == (2,))
 
-    theta = np.pi * (mu[0] + mu[1])
+    # theta = np.pi * (mu[0] + mu[1])
+    theta = (np.pi / 2.0) * (mu[0] + mu[1])
     Rot_matrix = np.array([[np.cos(theta), -np.sin(theta)], [np.sin(theta), np.cos(theta)]])
 
-    bump = mu[0]*(1 - mu[0])*mu[1]*(1-mu[1])
+    bump = mu[0]*(1 - mu[0])*mu[1]*(1-mu[1]) if use_bump else 0.0
 
     pp = (xx - mu.reshape((1,2))) @ Rot_matrix.T #@ Rot_matrix.T
     inv_Sigma = np.linalg.inv(Sigma)
@@ -284,23 +286,40 @@ for b in bb:
 
 raise RuntimeError
 
-nx = 89
-length_scaling = 0.0625
+nx = 49
+length_scaling = 1.0 #0.0625
 a = 1.0
 Ker, phi_function, Vh, H, mass_lumps, dof_coords = frog_setup(nx, length_scaling, a)
+
+num_neighbors = 10
 
 psf_object = lpsf1.make_psf_fenics(
     lambda X: H @ X,
     lambda X: H.T @ X,
     Vh, Vh,
     mass_lumps, mass_lumps,
-    num_initial_batches=10,#0,
+    num_initial_batches=20,#0,
     tau = 3.0, display=True,
+    num_neighbors = num_neighbors
 )
 
 Ker_psf = psf_object.psf_object.psf_kernel.cpp_object.eval_integral_kernel_block(dof_coords.T, dof_coords.T)
 err_psf = np.linalg.norm(Ker_psf - Ker) / np.linalg.norm(Ker)
 print('err_psf=', err_psf)
+
+p0 =  np.array([0.4, 0.4])
[…]

### 2024-02-08 numerical_examples/blur/blur_example.py
finished draft of interpolation figure

@@ -6,6 +6,7 @@ import localpsf.localpsf_cg1_lumped as lpsf1
 import scipy.linalg as sla
 from functools import partial
 from nalger_helper_functions import plot_ellipse
+import hlibpro_python_wrapper as hpro
 from tqdm.auto import tqdm
 
 save_figures = True
@@ -377,6 +378,18 @@ for k in range(num_neighbors):
     coords_string = str(sp_k[0]) + '_' + str(sp_k[1])
     plt.savefig('frog_neighbor_impulse_' + coords_string + '_.png', dpi=300, bbox_inches='tight')
 
+#
+
+print('Making row and column cluster trees')
+ct = hpro.build_cluster_tree_from_pointcloud(dof_coords, cluster_size_cutoff=32)
+
+print('Making block cluster trees')
+bct = hpro.build_block_cluster_tree(ct, ct, admissibility_eta=2.0)
+
+bct.visualize('frog_bct.eps')
+
+
+raise RuntimeError
 
 #
 

### 2024-02-08 localpsf/blur.py
update illustrations

@@ -1,0 +1,64 @@
+import numpy as np
+import dolfin as dl
+from localpsf.geometric_sort import geometric_sort
+
+
+# Frog function
+def frog_function(
+    xx: np.ndarray, # shape=(N, 2), N=nx*ny
+    vol: float,
+    mu: np.ndarray, # shape=(2,)
+    Sigma: np.ndarray, # shape=(2, 2)
+    a: float, # how much negative to include. a=0: gaussian
+    use_bump: bool = True,
+) -> np.ndarray: # shape=(N,)
+    assert(xx.shape[1] == 2)
+    N = xx.shape[0]
+    assert(Sigma.shape == (2, 2))
+    assert(mu.shape == (2,))
+
+    # theta = np.pi * (mu[0] + mu[1])
+    theta = (np.pi / 2.0) * (mu[0] + mu[1])
+    Rot_matrix = np.array([[np.cos(theta), -np.sin(theta)], [np.sin(theta), np.cos(theta)]])
+
+    bump = mu[0]*(1 - mu[0])*mu[1]*(1-mu[1]) if use_bump else 0.0
+
+    pp = (xx - mu.reshape((1,2))) @ Rot_matrix.T #@ Rot_matrix.T
+    inv_Sigma = np.linalg.inv(Sigma)
+    t_squared_over_sigma_squared = np.einsum('ai,ai->a', pp, np.einsum('ij,aj->ai', inv_Sigma, pp))
+    G = vol / (2.0 * np.pi * np.sqrt(np.linalg.det(Sigma))) * np.exp(-0.5 * t_squared_over_sigma_squared)
+
+    cos_x_over_sigma = np.cos(pp[:,0] / (np.sqrt(Sigma[0,0]) / 2.0))
+    sin_y_over_sigma = np.sin(pp[:,1] / (np.sqrt(Sigma[1,1]) / 2.0))
+    return bump * (1.0 + a * cos_x_over_sigma * sin_y_over_sigma) * G
+
+
+def frog_setup(
+        nx:             int     = 63,
+        length_scaling: float   = 1.0,
+        a:              float   = 1.0,
+):
+    ny = nx
+    mesh = dl.UnitSquareMesh(nx, ny)
+    Vh = dl.FunctionSpace(mesh, 'CG', 1)
+    dof_coords = Vh.tabulate_dof_coordinates()
+
+    vol = 1.0
+
[…]

### 2024-02-08 localpsf/geometric_sort.py

@@ -24,10 +24,10 @@ def geometric_sort_helper(
 
 def geometric_sort(
     points: np.ndarray, # shape=(num_pts, spatial_dimension)
-) -> np.ndarray[int]: # sort_inds, len=num_pts
+) -> np.ndarray: # sort_inds, len=num_pts
     num_pts, dim = points.shape
     sort_inds: np.ndarray = np.arange(num_pts)
-    geometric_sort_helper( 0, num_pts, 0, dim, points.copy(), sort_inds ) # modifies sort_inds
+    geometric_sort_helper(0, num_pts, 0, dim, points.copy(), sort_inds) # modifies sort_inds
     return sort_inds
 
 

### 2024-02-08 numerical_examples/blur/blur_ellipsoid_and_impulse_batches.py

@@ -1,0 +1,77 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import dolfin as dl
+import localpsf.localpsf_cg1_lumped as lpsf1
+from localpsf.blur import *
+import scipy.linalg as sla
+from functools import partial
+from nalger_helper_functions import plot_ellipse
+import hlibpro_python_wrapper as hpro
+from tqdm.auto import tqdm
+
+
+nx = 63
+length_scaling = 1.0
+a = 1.0
+Ker, phi_function, Vh, H, mass_lumps, dof_coords, kdtree_sort_inds = frog_setup(nx, length_scaling, a)
+
+#### Create PSF approximation
+
+psf_object = lpsf1.make_psf_fenics(
+    lambda X: H @ X,
+    lambda X: H.T @ X,
+    Vh, Vh,
+    mass_lumps, mass_lumps,
+    tau = 3.0, display=True,
+)
+
+#### Plot ellipsoid batch without impulses
+
+bb = [0,1,2]
+
+for b in bb:
+    IRB = psf_object.psf_object.impulse_response_batches
+    fig = plt.figure(figsize=(4,4))
+
+    phi = psf_object.impulse_response_batch(b)
+
+    start = IRB.batch2point_start[b]
+    stop = IRB.batch2point_stop[b]
+    pp = IRB.sample_points[start:stop, :]
+    mu_batch = IRB.sample_mu[start:stop, :]
+    Sigma_batch = IRB.sample_Sigma[start:stop, :, :]
+
+    for k in range(mu_batch.shape[0]):
+        plot_ellipse(mu_batch[k, :], Sigma_batch[k, :, :], n_std_tau=IRB.tau,
+                     facecolor='none', edgecolor='k', linewidth=1)
+
+    plt.gca().set_xticks([])
+    plt.gca().set_yticks([])
+
+    plt.savefig('frog_ellipsoid_batch' + str(b) + '.png', bbox_inches='tight', dpi=400)
+
+#### Plot ellipsoid batch with impulses
+
+for b in bb:
+    IRB = psf_object.psf_object.impulse_response_batches
+    fig = plt.figure(figsize=(4,4))
+
+    phi = psf_object.impulse_response_batch(b)
+
+    start = IRB.batch2point_start[b]
+    stop = IRB.batch2point_stop[b]
+    pp = IRB.sample_points[start:stop, :]
[…]

### 2024-02-08 numerical_examples/blur/blur_example.py

@@ -3,77 +3,19 @@ import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 import dolfin as dl
 import localpsf.localpsf_cg1_lumped as lpsf1
+from localpsf.blur import *
 import scipy.linalg as sla
 from functools import partial
 from nalger_helper_functions import plot_ellipse
 import hlibpro_python_wrapper as hpro
 from tqdm.auto import tqdm
 
-save_figures = True
-
-# Frog function
-def frog_function(
-    xx: np.ndarray, # shape=(N, 2), N=nx*ny
-    vol: float,
-    mu: np.ndarray, # shape=(2,)
-    Sigma: np.ndarray, # shape=(2, 2)
-    a: float, # how much negative to include. a=0: gaussian
-    use_bump: bool = True,
-) -> np.ndarray: # shape=(N,)
-    assert(xx.shape[1] == 2)
-    N = xx.shape[0]
-    assert(Sigma.shape == (2, 2))
-    assert(mu.shape == (2,))
-
-    # theta = np.pi * (mu[0] + mu[1])
-    theta = (np.pi / 2.0) * (mu[0] + mu[1])
-    Rot_matrix = np.array([[np.cos(theta), -np.sin(theta)], [np.sin(theta), np.cos(theta)]])
-
-    bump = mu[0]*(1 - mu[0])*mu[1]*(1-mu[1]) if use_bump else 0.0
-
-    pp = (xx - mu.reshape((1,2))) @ Rot_matrix.T #@ Rot_matrix.T
-    inv_Sigma = np.linalg.inv(Sigma)
-    t_squared_over_sigma_squared = np.einsum('ai,ai->a', pp, np.einsum('ij,aj->ai', inv_Sigma, pp))
-    G = vol / (2.0 * np.pi * np.sqrt(np.linalg.det(Sigma))) * np.exp(-0.5 * t_squared_over_sigma_squared)
-
-    cos_x_over_sigma = np.cos(pp[:,0] / (np.sqrt(Sigma[0,0]) / 2.0))
-    sin_y_over_sigma = np.sin(pp[:,1] / (np.sqrt(Sigma[1,1]) / 2.0))
-    return bump * (1.0 + a * cos_x_over_sigma * sin_y_over_sigma) * G
-
-
-def frog_setup(
-        nx:             int     = 63,
-        length_scaling: float   = 1.0,
-        a:              float   = 1.0,
-):
-    ny = nx
[…]

### 2024-02-08 numerical_examples/blur/blur_impulse_grid_plot.py

@@ -1,0 +1,39 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import dolfin as dl
+import localpsf.localpsf_cg1_lumped as lpsf1
+from localpsf.blur import *
+import scipy.linalg as sla
+from functools import partial
+from nalger_helper_functions import plot_ellipse
+import hlibpro_python_wrapper as hpro
+from tqdm.auto import tqdm
+
+
+nx = 63
+length_scaling = 1.0
+a = 1.0
+Ker, phi_function, Vh, H, mass_lumps, dof_coords, kdtree_sort_inds = frog_setup(nx, length_scaling, a)
+
+#### Impulse response grid plot
+
+num_pts_x = 5
+num_pts_y = 5
+pX, pY = np.meshgrid(np.linspace(0.01,0.99,num_pts_x), np.linspace(0.01,99,num_pts_y)[::-1])
+pp = np.vstack([pX.reshape(-1), pY.reshape(-1)]).T
+
+plt.figure(figsize = (num_pts_x,num_pts_y))
+gs1 = gridspec.GridSpec(num_pts_x, num_pts_y)
+gs1.update(wspace=0.025, hspace=0.05) # set the spacing between axes.
+for k in range(len(pp)):
+    p = pp[k]
+    ii = nearest_ind_func(dof_coords, p)
+    phi = dl.Function(Vh)
+    phi.vector()[:] = Ker[:, ii].copy()
+    plt.subplot(gs1[k])
+    cm = dl.plot(phi, cmap='binary')
+    plt.gca().set_xticks([])
+    plt.gca().set_yticks([])
+
+plt.savefig('frog_impulse_response_grid.png', bbox_inches='tight', dpi=300)

### 2024-02-08 numerical_examples/blur/blur_kernel_and_impulses_figure.py

@@ -1,0 +1,51 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import dolfin as dl
+import localpsf.localpsf_cg1_lumped as lpsf1
+from localpsf.blur import *
+import scipy.linalg as sla
+from functools import partial
+from nalger_helper_functions import plot_ellipse
+import hlibpro_python_wrapper as hpro
+from tqdm.auto import tqdm
+
+
+nx = 63
+length_scaling = 1.0
+a = 1.0
+Ker, phi_function, Vh, H, mass_lumps, dof_coords, kdtree_sort_inds = frog_setup(nx, length_scaling, a)
+
+Ker_reordered = Ker[:, kdtree_sort_inds][kdtree_sort_inds,:]
+dof_coords_reordered = dof_coords[kdtree_sort_inds,:]
+
+#### Plot two impulse responses individually, and kernel with the corresponding columns indicated
+
+p1 = np.array([0.2, 0.45])
+p2 = np.array([0.6, 0.6])
+pp = [p1, p2]
+
+plt.matshow(Ker_reordered, cmap='binary')
+plt.ylim([Ker_reordered.shape[0], 0])
+plt.gca().set_xticks([])
+plt.gca().set_yticks([])
+
+for k in range(len(pp)):
+    p = pp[k]
+    ii = nearest_ind_func(dof_coords_reordered, p)
+    print('k=', k, ', ii=', ii)
+    plt.plot([ii, ii], [0, Ker.shape[0]], 'k', linestyle='dotted', linewidth=2.0)
+
+plt.gcf().set_size_inches(8, 8)
+plt.savefig('frog_kernel_matrix_a1.png', bbox_inches='tight', dpi=400)
+
+for k in range(len(pp)):
+    p = pp[k]
+    ii = nearest_ind_func(dof_coords, p)
+    phi = dl.Function(Vh)
+    phi.vector()[:] = Ker[:, ii].copy()
+    plt.figure(figsize=(3.8, 3.8))
+    cm = dl.plot(phi, cmap='binary')
+    plt.gca().set_xticks([])
+    plt.gca().set_yticks([])
+    plt.savefig('frog_impulse_response' + str(k) + '.png', bbox_inches='tight', dpi=400)

### 2024-02-08 numerical_examples/blur/blur_moments_figure.py

@@ -1,0 +1,71 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import dolfin as dl
+import localpsf.localpsf_cg1_lumped as lpsf1
+from localpsf.blur import *
+import scipy.linalg as sla
+from functools import partial
+from nalger_helper_functions import plot_ellipse
+import hlibpro_python_wrapper as hpro
+from tqdm.auto import tqdm
+
+
+nx = 63
+length_scaling = 1.0
+a = 1.0
+Ker, phi_function, Vh, H, mass_lumps, dof_coords, kdtree_sort_inds = frog_setup(nx, length_scaling, a)
+
+#### Create PSF approximation
+
+psf_object = lpsf1.make_psf_fenics(
+    lambda X: H @ X,
+    lambda X: H.T @ X,
+    Vh, Vh,
+    mass_lumps, mass_lumps,
+    tau = 3.0, display=True,
+)
+
+#### Plot moments
+
+plt.figure(figsize=(4,4))
+dl.plot(psf_object.vol(), cmap='binary', clim=[0.0, None])
+# plt.colorbar(cm)
+# plt.title(r'$V$')
+plt.gca().set_xticks([])
+plt.gca().set_yticks([])
+plt.savefig('frog_vol.png', bbox_inches='tight', dpi=400)
+
+plt.figure(figsize=(4,4))
+cm = dl.plot(psf_object.mu(0), cmap='binary')
+plt.gca().set_xticks([])
+plt.gca().set_yticks([])
+plt.savefig('frog_mu0.png', bbox_inches='tight', dpi=400)
+
+plt.figure(figsize=(4,4))
+dl.plot(psf_object.mu(1), cmap='binary')
+# plt.title(r'$\mu^2$')
+plt.gca().set_xticks([])
+plt.gca().set_yticks([])
+plt.savefig('frog_mu1.png', bbox_inches='tight', dpi=400)
+
+plt.figure(figsize=(4,4))
+dl.plot(psf_object.Sigma(0,0), cmap='binary')
+# plt.title(r'$\Sigma^{11}$')
+plt.gca().set_xticks([])
+plt.gca().set_yticks([])
+plt.savefig('frog_Sigma00.png', bbox_inches='tight', dpi=400)
+
+plt.figure(figsize=(4,4))
+dl.plot(psf_object.Sigma(0,1), cmap='binary')
[…]

### 2024-02-09 paper_sisc_first_revision/localpsf.tex
blur example working ok
unclear how to illustrate h-matrix stuff
put more frog example images into paper
finished draft of interpolation figure
update illustrations
generated columnwise error plots and added to paper

@@
 However, many data misfit Hessians of practical interest do satisfy the local non-negative integral kernel properties (either exactly or approximately), and the proposed method is targeted at approximating these Hessians.
+
+\section{Illustration of the method on a spatially variant blur operator}
+
+Before delving into technical details, we illustrate the method on a spatially variant blurring operator with the following integral kernel:
+$$\Phi(y,x) = (1 - a f(y,x)) g(x) \exp \left(-\frac{1}{2}h^T(y,x) C^{-1} h(y,x)\right).$$ 
+Here, $h(y,x) = R(\theta(x)) (y-x)$, $R(\theta(x))$ is the 2D rotation matrix by angle $\theta(x) = (x^1 + x^2) \pi / 2$, $C = L^2 \operatorname{diag}(c_1, c_2)$, $g(x) = x^1(1-x^1)x^2(1-x^2)$,  and $f(x,y) = \cos\left(p^1(y,x) / \sqrt{c_1 / 2.0}\right) \sin\left(p^2(y,x) / \sqrt{c_2 / 2.0}\right)$ for some constants $a, L, c_1, c_2$.
+We use $a=1.0$, $L^2=1.0$, $c_1=0.01$, and $c_2=0.0025$.
+Using this kernel, the impulse responses, $\phi_x$, are bumpy blobs that are centered roughly at $x$, and are rotated and scaled in a manner that depends on $x$ (see Figure \ref{fig:frog1}).
+The constant $L$ controls how wide the blob is, and $a$ controls the bumpiness.
+With $a=0$, $\phi_x$ is a Gaussian.
+As $a$ increases, $\phi_x$ becomes more bumpy.
+When $a > 1$, the integral kernel contains negative values, which violates the assumptions of our method.
+In the numerical results section we will vary $a$ and $L$, even allowing $a>1$ to test the robustness of the method to violations of the non-negative kernel assumption.
+Note that for this operator, using our method is unnecessary because entries of the kernel are given directly via a formula.
+The purpose of this example is to illustrate and test the method on an operator which is known exactly and may be easily visualized.
+
+\begin{figure}
+	\begin{subfigure}[b]{0.49\textwidth}
+		\centering
+		\includegraphics[scale=0.25]{kernel_impulse_illustration2.pdf}
+		\caption{Kernel and impulse responses}
+		\label{fig:kernel_impulse_illustration}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.49\textwidth}
+		\centering
+		\includegraphics[scale=0.5]{frog_ellipsoid_width.pdf}
+		\caption{Impulse response support ellipsoid}
+		\label{fig:frog_ellipsoid_width}
+	\end{subfigure}
+
+	\caption{(\ref{fig:kernel_impulse_illustration}) asdf. \ref{fig:frog_ellipsoid_width}) asdf2
+	}
+	\label{fig:frog1}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[scale=0.25]{frog_moments2.pdf}	
+	\caption{asdf3
+	}
+	\label{fig:frog2}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[scale=0.3]{frog_batches2.pdf}	
+	\caption{asdf4
+	}
+	\label{fig:frog3}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[scale=0.24]{frog_hmatrix_knn_impulse2.png}	
+	\caption{asdfg
+	}
+	\label{fig:frog4}
+\end{figure}
+
+\begin{figure}
+	\begin{subfigure}[b]{0.30\textwidth}
+		\centering
+		\includegraphics[scale=0.4]{frog_column_errors_psf5.png}
+		\caption{PSF(5)}
+		\label{fig:blur_error_psf5}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.30\textwidth}
+		\centering
+		\includegraphics[scale=0.4]{frog_column_errors_psf10.png}
+		\caption{PSF(10)}
+		\label{fig:blur_error_psf10}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.38\textwidth}
+		\centering
+		\includegraphics[scale=0.4]{frog_column_errors_psf20_colorbar.png}
+		\caption{PSF(20)}
+		\label{fig:blur_error_psf20}
+	\end{subfigure}
+
+	\caption{columnwise error plots for kernel reconstruction
+	}
+	\label{fig:frog5}
+\end{figure}
 
@@
 Alger, Longfei Gao, Mathew Hu, and Rami Nammour for helpful discussions.
+The seismic Hessian impulse response image was provided to us by Mathew Hu.
 We thank Trevor Heise for editing suggestions.

### 2024-02-09 numerical_examples/blur/blur_hmatrix_bct.py

@@ -1,0 +1,39 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import dolfin as dl
+import localpsf.localpsf_cg1_lumped as lpsf1
+from localpsf.blur import *
+import scipy.linalg as sla
+from functools import partial
+from nalger_helper_functions import plot_ellipse
+import hlibpro_python_wrapper as hpro
+from tqdm.auto import tqdm
+
+
+nx = 49
+length_scaling = 1.0 #0.0625
+a = 1.0
+Ker, phi_function, Vh, H, mass_lumps, dof_coords, kdtree_sort_inds = frog_setup(nx, length_scaling, a)
+
+num_neighbors = 10
+
+psf_object = lpsf1.make_psf_fenics(
+    lambda X: H @ X,
+    lambda X: H.T @ X,
+    Vh, Vh,
+    mass_lumps, mass_lumps,
+    num_initial_batches=20,#0,
+    tau = 3.0, display=True,
+    num_neighbors = num_neighbors
+)
+
+#
+
+print('Making row and column cluster trees')
+ct = hpro.build_cluster_tree_from_pointcloud(dof_coords, cluster_size_cutoff=32)
+
+print('Making block cluster trees')
+bct = hpro.build_block_cluster_tree(ct, ct, admissibility_eta=2.0)
+
+bct.visualize('frog_bct.eps')

### 2024-02-09 paper_sisc_first_revision/localpsf.tex
update ellipsoid width illustration. Work on blur section

--- 2024-01-15 paper_sisc_first_revision/localpsf.tex
@@
 
-\section{Preliminaries}
+\section{Overview of the method and illustration on a blur operator}
+
+Before delving into technical details, we provide an overview of the method.
+Along the way, we illustrate steps using a spatially variant blur example, in which the integral kernel is given as follows:
+$$\Phi(y,x) = (1 - a f(y,x)) g(x) \exp \left(-\frac{1}{2}h^T(y,x) C^{-1} h(y,x)\right).$$ 
+Here, $h(y,x) = R(\theta(x)) (y-x)$, $R(\theta(x))$ is the 2D rotation matrix by angle $\theta(x) = (x^1 + x^2) \pi / 2$, $C = L^2 \operatorname{diag}(c_1, c_2)$, $g(x) = x^1(1-x^1)x^2(1-x^2)$,  and $f(x,y) = \cos\left(p^1(y,x) / \sqrt{c_1 / 2.0}\right) \sin\left(p^2(y,x) / \sqrt{c_2 / 2.0}\right)$ for some constants $a, L, c_1, c_2$.
+In this section we use $a=1.0$, $L^2=1.0$, $c_1=0.01$, and $c_2=0.0025$.
+Using this kernel, the impulse response $\phi_x$ is a bumpy blob that is centered near $x$, and is rotated and scaled in a manner that depends on $x$ (see Figure \ref{fig:frog1}).
+Note that for this operator, using our method is unnecessary because entries of the kernel are given directly via a formula.
+The purpose of this example is to illustrate the method using an operator which is known exactly and may be easily visualized.
+
+%The constant $L$ controls how wide the blob is, and $a$ controls the bumpiness.
+With $a=0$, $\phi_x$ is a Gaussian.
+As $a$ increases, $\phi_x$ becomes more bumpy.
+When $a > 1$, the integral kernel contains negative values, which violates the assumptions of our method.
+In the numerical results section we will vary $a$ and $L$, even allowing $a>1$ to test the robustness of the method to violations of the non-negative kernel assumption.
+
+\begin{figure}
+	\begin{subfigure}[b]{0.49\textwidth}
+		\centering
+		\includegraphics[scale=0.2]{kernel_impulse_illustration2.pdf}
+		\caption{Kernel and impulse responses}
+		\label{fig:kernel_impulse_illustration}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.49\textwidth}
+		\centering
+		\includegraphics[scale=0.4]{frog_ellipsoid_width2.pdf}
+		\caption{Impulse response support ellipsoid}
+		\label{fig:frog_ellipsoid_width}
+	\end{subfigure}
+
+	\caption{(\ref{fig:kernel_impulse_illustration}) asdf. \ref{fig:frog_ellipsoid_width}) asdf2
+	}
+	\label{fig:frog1}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[scale=0.2]{frog_moments2.pdf}	
+	\caption{asdf3
+	}
+	\label{fig:frog2}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[scale=0.25]{frog_batches2.pdf}	
+	\caption{asdf4
+	}
+	\label{fig:frog3}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[scale=0.2]{frog_hmatrix_knn_impulse2.png}	
+	\caption{asdfg
+	}
+	\label{fig:frog4}
+\end{figure}
+
+\begin{figure}
+	\begin{subfigure}[b]{0.30\textwidth}
+		\centering
+		\includegraphics[scale=0.35]{frog_column_errors_psf5.png}
+		\caption{PSF(5)}
+		\label{fig:blur_error_psf5}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.30\textwidth}
+		\centering
+		\includegraphics[scale=0.35]{frog_column_errors_psf10.png}
+		\caption{PSF(10)}
+		\label{fig:blur_error_psf10}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.37\textwidth}
+		\centering
+		\includegraphics[scale=0.35]{frog_column_errors_psf20_colorbar.png}
+		\caption{PSF(20)}
+		\label{fig:blur_error_psf20}
+	\end{subfigure}
+
+	\caption{columnwise error plots for kernel reconstruction
+	}
+	\label{fig:frog5}
+\end{figure}
+
+\section{Technical preliminaries}
 

### 2024-02-09 paper_sisc_first_revision/localpsf.tex
editing figures with noemi

@@
 		\centering
-		\includegraphics[scale=0.2]{kernel_impulse_illustration2.pdf}
-		\caption{Kernel and impulse responses}
+		\includegraphics[scale=0.25]{kernel_impulse_illustration2.pdf}
+%		\caption{Kernel and impulse responses}
 		\label{fig:kernel_impulse_illustration}
@@
 		\centering
-		\includegraphics[scale=0.4]{frog_ellipsoid_width2.pdf}
-		\caption{Impulse response support ellipsoid}
+		\includegraphics[scale=0.5]{frog_ellipsoid_width2.pdf}
+%		\caption{Impulse response support ellipsoid}
 		\label{fig:frog_ellipsoid_width}
@@
 
-	\caption{(\ref{fig:kernel_impulse_illustration}) asdf. \ref{fig:frog_ellipsoid_width}) asdf2
+	\caption{left kernel and impulse responses. right impulse response support ellipsoid
 	}
@@
 	\centering
-	\includegraphics[scale=0.2]{frog_moments2.pdf}	
+	\includegraphics[scale=0.25]{frog_moments2.pdf}	
 	\caption{asdf3
@@
 	\centering
-	\includegraphics[scale=0.25]{frog_batches2.pdf}	
+	\includegraphics[scale=0.3]{frog_batches3.pdf}	
 	\caption{asdf4
@@
 	\centering
-	\includegraphics[scale=0.2]{frog_hmatrix_knn_impulse2.png}	
+	\includegraphics[scale=0.2]{frog_hmatrix_knn_impulse3.pdf}	
 	\caption{asdfg
@@
 		\includegraphics[scale=0.35]{frog_column_errors_psf5.png}
-		\caption{PSF(5)}
+%		\caption{PSF(5)}
 		\label{fig:blur_error_psf5}
@@
 		\includegraphics[scale=0.35]{frog_column_errors_psf10.png}
-		\caption{PSF(10)}
+%		\caption{PSF(10)}
 		\label{fig:blur_error_psf10}
 	\end{subfigure}
-	\begin{subfigure}[b]{0.37\textwidth}
-		\centering
-		\includegraphics[scale=0.35]{frog_column_errors_psf20_colorbar.png}
-		\caption{PSF(20)}
+	\begin{subfigure}[b]{0.30\textwidth}
+		\centering
+		\includegraphics[scale=0.35]{frog_column_errors_psf20.png}
+%		\caption{PSF(20)}
 		\label{fig:blur_error_psf20}
 		\end{subfigure}
-
-	\caption{columnwise error plots for kernel reconstruction
+			\begin{subfigure}[b]{0.08\textwidth}
+			\centering
+			\includegraphics[scale=0.35]{frog_colorbar_edited1.pdf}
+			\label{fig:blur_colorbar}
+		\end{subfigure}
+
+	\caption{columnwise error plots for kernel reconstruction left 5, center 10, right 20
 	}

### 2024-02-09 paper_sisc_first_revision/localpsf.tex
editing figures with noemi

@@
 
-\section{Overview of the method and illustration on a blur operator}
-
-Before delving into technical details, we provide an overview of the method.
-Along the way, we illustrate steps using a spatially variant blur example, in which the integral kernel is given as follows:
-$$\Phi(y,x) = (1 - a f(y,x)) g(x) \exp \left(-\frac{1}{2}h^T(y,x) C^{-1} h(y,x)\right).$$ 
-Here, $h(y,x) = R(\theta(x)) (y-x)$, $R(\theta(x))$ is the 2D rotation matrix by angle $\theta(x) = (x^1 + x^2) \pi / 2$, $C = L^2 \operatorname{diag}(c_1, c_2)$, $g(x) = x^1(1-x^1)x^2(1-x^2)$,  and $f(x,y) = \cos\left(p^1(y,x) / \sqrt{c_1 / 2.0}\right) \sin\left(p^2(y,x) / \sqrt{c_2 / 2.0}\right)$ for some constants $a, L, c_1, c_2$.
-In this section we use $a=1.0$, $L^2=1.0$, $c_1=0.01$, and $c_2=0.0025$.
-Using this kernel, the impulse response $\phi_x$ is a bumpy blob that is centered near $x$, and is rotated and scaled in a manner that depends on $x$ (see Figure \ref{fig:frog1}).
-Note that for this operator, using our method is unnecessary because entries of the kernel are given directly via a formula.
-The purpose of this example is to illustrate the method using an operator which is known exactly and may be easily visualized.
-
-%The constant $L$ controls how wide the blob is, and $a$ controls the bumpiness.
-With $a=0$, $\phi_x$ is a Gaussian.
-As $a$ increases, $\phi_x$ becomes more bumpy.
-When $a > 1$, the integral kernel contains negative values, which violates the assumptions of our method.
-In the numerical results section we will vary $a$ and $L$, even allowing $a>1$ to test the robustness of the method to violations of the non-negative kernel assumption.
-
 \begin{figure}
-	\begin{subfigure}[b]{0.49\textwidth}
-		\centering
-		\includegraphics[scale=0.25]{kernel_impulse_illustration2.pdf}
+	\begin{center}
+	\begin{subfigure}[b]{0.58\textwidth}
+		\centering
+		\includegraphics[scale=0.275]{kernel_impulse_illustration2.pdf}
 %		\caption{Kernel and impulse responses}
@@
 	\end{subfigure}
-	\begin{subfigure}[b]{0.49\textwidth}
-		\centering
-		\includegraphics[scale=0.5]{frog_ellipsoid_width2.pdf}
+	\begin{subfigure}[b]{0.41\textwidth}
+		\centering
+		\includegraphics[scale=0.55]{frog_ellipsoid_width2.pdf}
 %		\caption{Impulse response support ellipsoid}
@@
 	\end{subfigure}
+	\end{center}
 
@@
 	\centering
-	\includegraphics[scale=0.25]{frog_moments2.pdf}	
+	\includegraphics[scale=0.28]{frog_moments2.pdf}	
 	\caption{asdf3
@@
 
-\section{Technical preliminaries}
+\section{Preliminaries}
 
@@
 
-\subsection{Example 1: Inversion for the basal friction coefficient in an ice sheet flow problem}
+\subsection{Example 1: Spatially variant blur}
+
+Before delving into technical details, we provide an overview of the method.
+Along the way, we illustrate steps using a spatially variant blur example, in which the integral kernel is given as follows:
+$$\Phi(y,x) = (1 - a f(y,x)) g(x) \exp \left(-\frac{1}{2}h^T(y,x) C^{-1} h(y,x)\right).$$ 
+Here, $h(y,x) = R(\theta(x)) (y-x)$, $R(\theta(x))$ is the 2D rotation matrix by angle $\theta(x) = (x^1 + x^2) \pi / 2$, $C = L^2 \operatorname{diag}(c_1, c_2)$, $g(x) = x^1(1-x^1)x^2(1-x^2)$,  and $f(x,y) = \cos\left(p^1(y,x) / \sqrt{c_1 / 2.0}\right) \sin\left(p^2(y,x) / \sqrt{c_2 / 2.0}\right)$ for some constants $a, L, c_1, c_2$.
+In this section we use $a=1.0$, $L^2=1.0$, $c_1=0.01$, and $c_2=0.0025$.
+Using this kernel, the impulse response $\phi_x$ is a bumpy blob that is centered near $x$, and is rotated and scaled in a manner that depends on $x$ (see Figure \ref{fig:frog1}).
+The constant $L$ controls how wide the blob is, and $a$ controls the bumpiness.
+With $a=0$, $\phi_x$ is a Gaussian.
+As $a$ increases, $\phi_x$ becomes more bumpy.
+When $a > 1$, the integral kernel contains negative values, which violates the assumptions of our method.
+Note that for this operator, using our method is unnecessary because entries of the kernel are given directly via a formula.
+The purpose of this example is to illustrate the method using an operator which is known exactly and may be easily visualized.
+
+\subsection{Example 2: Inversion for the basal friction coefficient in an ice sheet flow problem}
 \label{sec:ice}
@@
 
-\subsection{Example 2: Inversion for the initial condition in an advective-diffusive transport problem}
+\subsection{Example 3: Inversion for the initial condition in an advective-diffusive transport problem}
 \label{sec:adv}

### 2024-02-09 paper_sisc_first_revision/localpsf.tex
incorporated figure references into intro text

@@
 
+\begin{figure}
+	\begin{center}
+		%	\begin{subfigure}[b]{0.58\textwidth}
+		%		\centering
+		\includegraphics[scale=0.35]{kernel_impulse_illustration2.pdf}
+		%%		\caption{Kernel and impulse responses}
+		%		\label{fig:kernel_impulse_illustration}
+		%	\end{subfigure}
+		%	\begin{subfigure}[b]{0.41\textwidth}
+		%		\centering
+		%		\includegraphics[scale=0.55]{frog_ellipsoid_width2.pdf}
+		%%		\caption{Impulse response support ellipsoid}
+		%		\label{fig:frog_ellipsoid_width}
+		%	\end{subfigure}
+	\end{center}
+
+	\caption{left kernel and impulse responses. right impulse response support ellipsoid
+	}
+	\label{fig:frog_kernel_impulse_responses}
+\end{figure}
+
 We use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applications.
@@
 We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with batches of points scattered throughout the domain (see Figure~\ref{fig:batches_intro}).
-Then we interpolate translated and scaled versions of these impulse responses to approximate entries of the operator's integral kernel.
-Picking the batches of points requires us to estimate the supports of the impulse responses $\phi_x$ \emph{before} they are computed.
-Ellipsoid estimates for the supports of all $\phi_x$ are determined a-priori via a moment method that involves applying $\Aop^T$ to a small number of polynomial functions (see Section~\ref{sec:intromoments}).
-We are inspired by resolution analysis in seismic imaging, in which
-$\Aop^T$ is applied to a random noise function, and the width of the support $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$~\cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
-The moment method that we use estimates the support of $\phi_x$ more accurately than random noise probing in resolution analysis, at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
-Adding more batches yields impulse responses at more points, increasing the accuracy of the approximation at the cost of one operator application per batch.
+Batches of impulse responses may be thought of intuitively as sets of ``columns'' of the kernel (Figure \ref{fig:frog_kernel_impulse_responses}).
+To choose the batches, we form ellipsoid estimates for the supports of all $\phi_x$ via a moment method (Figure~\ref{fig:frog_moments_ellipsoid}) that involves applying $\Aop^T$ to a small number of polynomial functions (see Section~\ref{sec:intromoments}).
+We then use a greedy ellipsoid packing algorithm (Figure~\ref{fig:frog_batches}) to maximize the number of impulse responses per batch.
+% which requires us to estimate the supports of the impulse responses $\phi_x$ \emph{before} they are computed. 
+Then we interpolate translated and scaled versions of these impulse responses to approximate entries of the operator's integral kernel (Figure~\ref{fig:hmatrix_neighbors_frog}).
+Adding more batches yields impulse responses at more points, increasing the accuracy of the approximation at the cost of one operator application per batch (Figure~\ref{fig:frog_column_errors}).
+
+\begin{figure}
+	\begin{subfigure}[b]{0.65\textwidth}
+		\centering
+		\includegraphics[scale=0.22]{frog_moments3.pdf}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.34\textwidth}
+		\centering
+		\includegraphics[scale=0.44]{frog_ellipsoid_width2.pdf}
+		%		\label{fig:frog_ellipsoid_width}
+	\end{subfigure}
+	\caption{asdf3
+	}
+	\label{fig:frog_moments_ellipsoid}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[scale=0.3]{frog_batches3.pdf}
+	\caption{asdf4
+	}
+	\label{fig:frog_batches}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[scale=0.2]{frog_hmatrix_knn_impulse3.pdf}	
+	\caption{asdfg
+	}
+	\label{fig:hmatrix_neighbors_frog}
+\end{figure}
+
+\begin{figure}
+	\begin{subfigure}[b]{0.30\textwidth}
+		\centering
+		\includegraphics[scale=0.35]{frog_column_errors_psf5.png}
+		%		\caption{PSF(5)}
+		\label{fig:blur_error_psf5}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.30\textwidth}
+		\centering
+		\includegraphics[scale=0.35]{frog_column_errors_psf10.png}
+		%		\caption{PSF(10)}
+		\label{fig:blur_error_psf10}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.30\textwidth}
+		\centering
+		\includegraphics[scale=0.35]{frog_column_errors_psf20.png}
+		%		\caption{PSF(20)}
+		\label{fig:blur_error_psf20}
+	\end{subfigure}
+	\begin{subfigure}[b]{0.08\textwidth}
+		\centering
+		\includegraphics[scale=0.35]{frog_colorbar_edited1.pdf}
+		%			\label{fig:blur_colorbar}
+	\end{subfigure}
+
+	\caption{columnwise error plots for kernel reconstruction left 5, center 10, right 20
+	}
+	\label{fig:frog_column_errors}
+\end{figure}
 
@@
 In this paper, we use a new moment method (Section \ref{sec:intromoments}) which permits computation of many impulse responses (e.g., 50) per operator application.
-(3) The method we propose never evaluates computed impulse responses outside of their domain of definition.
-This eliminates boundary artifacts which plague conventional PC methods.
+We are inspired by resolution analysis in seismic imaging, in which
+$\Aop^T$ is applied to a random noise function, and the width of the support $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$~\cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
+The moment method that we use estimates the support of $\phi_x$ more accurately than random noise probing in resolution analysis, at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
+(3) {\color{red} The method we propose never evaluates computed impulse responses outside of their domain of definition.
+This eliminates boundary artifacts which plague conventional PC methods.}
 
@@
 However, many data misfit Hessians of practical interest do satisfy the local non-negative integral kernel properties (either exactly or approximately), and the proposed method is targeted at approximating these Hessians.
-
-\begin{figure}
-	\begin{center}
-	\begin{subfigure}[b]{0.58\textwidth}
-		\centering
-		\includegraphics[scale=0.275]{kernel_impulse_illustration2.pdf}
-%		\caption{Kernel and impulse responses}
-		\label{fig:kernel_impulse_illustration}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.41\textwidth}
-		\centering
-		\includegraphics[scale=0.55]{frog_ellipsoid_width2.pdf}
-%		\caption{Impulse response support ellipsoid}
-		\label{fig:frog_ellipsoid_width}
-	\end{subfigure}
-	\end{center}
-
-	\caption{left kernel and impulse responses. right impulse response support ellipsoid
-	}
-	\label{fig:frog1}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[scale=0.28]{frog_moments2.pdf}	
-	\caption{asdf3
-	}
-	\label{fig:frog2}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[scale=0.3]{frog_batches3.pdf}	
-	\caption{asdf4
-	}
-	\label{fig:frog3}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[scale=0.2]{frog_hmatrix_knn_impulse3.pdf}	
-	\caption{asdfg
-	}
-	\label{fig:frog4}
-\end{figure}
-
-\begin{figure}
-	\begin{subfigure}[b]{0.30\textwidth}
-		\centering
-		\includegraphics[scale=0.35]{frog_column_errors_psf5.png}
-%		\caption{PSF(5)}
-		\label{fig:blur_error_psf5}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.30\textwidth}
-		\centering
-		\includegraphics[scale=0.35]{frog_column_errors_psf10.png}
-%		\caption{PSF(10)}
-		\label{fig:blur_error_psf10}
-	\end{subfigure}
-	\begin{subfigure}[b]{0.30\textwidth}
-		\centering
-		\includegraphics[scale=0.35]{frog_column_errors_psf20.png}
-%		\caption{PSF(20)}
-		\label{fig:blur_error_psf20}
-		\end{subfigure}
-			\begin{subfigure}[b]{0.08\textwidth}
-			\centering
-			\includegraphics[scale=0.35]{frog_colorbar_edited1.pdf}
-			\label{fig:blur_colorbar}
-		\end{subfigure}
-
-	\caption{columnwise error plots for kernel reconstruction left 5, center 10, right 20
-	}
-	\label{fig:frog5}
-\end{figure}
 

### 2024-02-09 seismic_impulse_figure.py

@@ -1,0 +1,29 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+plt.figure()
+Hv = np.load("imp_res_shallow.npy")
+plt.imshow(Hv.T, cmap='binary')
+plt.colorbar()
+
+plt.figure()
+Hv = np.load("imp_res_deep.npy")
+# Hv = np.abs(Hv)
+plt.imshow(Hv.T, cmap='binary')
+plt.colorbar()
+
+C = np.ones(Hv.T.shape)
+Lx = np.outer(np.ones(Hv.shape[1]), np.arange(Hv.shape[0]))
+Ly = np.outer(np.arange(Hv.shape[1]), np.ones(Hv.shape[0]))
+Qxx = Lx * Lx
+Qxy = Lx * Ly
+Qyy = Ly * Ly
+
+# plt.imshow(Ly, cmap='binary')
+# plt.colorbar()
+
+vol = np.sum(Hv.T * C)
+mu_x = np.sum(Hv.T * Lx / vol)
+mu_y = np.sum(Hv.T * Ly / vol)
+
+plt.plot(mu_x, mu_y, '.r')

### 2024-02-09 seismic_impulse_figure.py
add seismic example. fix reply to reviewer number bug

@@ -1,20 +1,40 @@
 import numpy as np
 import matplotlib.pyplot as plt
+from nalger_helper_functions import plot_ellipse
+
+Hv1 = np.load("imp_res_shallow.npy")
+Hv2 = np.load("imp_res_deep.npy")
+# phi = Hv1.T / np.max(Hv1) + Hv2.T / np.max(Hv2)
+phi1 = Hv1.T / np.max(Hv1.T)
+phi2 = Hv2.T / np.max(Hv2.T)
 
 plt.figure()
-Hv = np.load("imp_res_shallow.npy")
-plt.imshow(Hv.T, cmap='binary')
+# plt.imshow(phi, cmap='binary', vmax=np.max(phi)/5.0)
+plt.imshow(phi2, cmap='binary', vmax=np.max(phi2))
+plt.xlim(200 - 60, 200 + 60)
 plt.colorbar()
 
-plt.figure()
-Hv = np.load("imp_res_deep.npy")
-# Hv = np.abs(Hv)
-plt.imshow(Hv.T, cmap='binary')
-plt.colorbar()
+plt.gca().set_xticks([])
+plt.gca().set_yticks([])
 
-C = np.ones(Hv.T.shape)
-Lx = np.outer(np.ones(Hv.shape[1]), np.arange(Hv.shape[0]))
-Ly = np.outer(np.arange(Hv.shape[1]), np.ones(Hv.shape[0]))
+plt.savefig('marmousi_impulse_response1.png', dpi=300, bbox_inches='tight')
+
+
+raise RuntimeError
+
+# plt.figure()
+# # plt.imshow(phi, cmap='binary', vmax=np.max(phi)/5.0)
+# plt.imshow(phi2, cmap='binary', vmax=np.max(phi2))
+# plt.xlim(200 - 60, 200 + 60)
+# plt.colorbar()
+
+#
+
+phi = np.abs(phi)
+
+C = np.ones(phi.shape)
+Lx = np.outer(np.ones(  phi.shape[0]),  np.arange(phi.shape[1]))
+Ly = np.outer(np.arange(phi.shape[0]),  np.ones(  phi.shape[1]))
 Qxx = Lx * Lx
 Qxy = Lx * Ly
 Qyy = Ly * Ly
@@ -22,8 +42,17 @@ Qyy = Ly * Ly
[…]

### 2024-02-12 paper_sisc_first_revision/localpsf.tex
added captions

@@
 
-	\caption{left kernel and impulse responses. right impulse response support ellipsoid
+	\caption{Left: Matrix created by evaluating the integral kernel $\Phi$ for $\Aop$ (Equation~\ref{eq:kernel_representation}) at all pairs of mesh vertices.
+This illustration is for the integral kernel given in Equation~\ref{eq:frog_kernel} in Section~\ref{sec:frog}.
+Dark colors indicate large entries and light colors indicate small entries.
+Rows and columns are ordered according to a kd-tree hierarchical clustering.
+Right: Impulse responses
+%	$\phi_{x_1}=\Phi(~\cdot~, x_1)$ and $\phi_{x_1}=\Phi(~\cdot~, x_2)$
+	associated with points $x_1,x_2\in \Omega$, shown by the two dotted vertical lines. Intuitively, one may think of impulse responses as ``columns'' of the integral kernel.
 	}
@@
 \end{figure}
+
+x1 = (0.5, 0.5)
+y = (y1, y2)
+Phi((y1, y2), (x1, x2)) = Phi(y,x)
 
@@
 	\end{subfigure}
-	\caption{asdf3
+	\caption{Left: Impulse response moments: scaling factor ($V$), mean ($\mu$), and covariance ($\Sigma$).
+For each point $x \in \Omega$, the quantity $V(x)$ is the integral of $\phi_x$ over $\Omega$, $\mu(x)$ is the location that $\phi_x$ is centered at, and $\Sigma(x)$ is a matrix with eigenvectors and eigenvalues that characterize the width of $\phi_x$ (see Section~\ref{sec:intromoments}).
+Right: Ellipsoid support for an impulse response.
+This ellipsoid is the set of points within $\tau$ standard deviations of the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$.
+The scaling factor $V(x)$ characterizes the magnitude of $\phi_x$.
 	}
@@
 	\includegraphics[scale=0.3]{frog_batches3.pdf}
-	\caption{asdf4
+	\caption{Illustration of the process to compute one impulse response batch.
+Impulse response moments are first used to form ellipsoid shaped estimates of the supports of impulse responses (Equation~\ref{eq:support_ellipsoid}).
+Then, an ellipsoid packing problem is solved to choose batches of non-overlapping support ellipsoids (Section~\ref{sec:sample_point_selection}).
+%		Note that we do not need to compute the impulse response in order to know its support ellipsoid. 
+	Finally, $\Aop$ is applied to a Dirac comb associated with the points $x_i$, which correspond to the ellipsoids (Section~\ref{sec:get_impulse_response}). The process is repeated to form more batches.
+%		
+%		Impulse response moments are used to choose batches of non-overlapping impulse responses by solving an ellipsoid packing problem using a greedy algorithm. Batches of impulse responses 
+%		
+%		Moments of the impulse responses used to form ellipsoid shaped a-priori estimates for the supports of the impulse responses.
 	}
@@
 	\includegraphics[scale=0.2]{frog_hmatrix_knn_impulse3.pdf}	
-	\caption{asdfg
+	\caption{Left: H-matrix structure for $\Aker$.
+Computing an entry of this matrix requires evaluating the integral kernel, $\Phi(y,x)$, at a pair of points $(y,x) \in \Omega \times \Omega$.
+Center: Kernel evaluation points $x$ and $y$ (black circles),  sample points for the approximation (light gray and black dots), and the $k_n$ sample points, $x'$, that are nearest to $x$ (black dots).
+Right: Known impulse response at $x'$.
+Using radial basis function interpolation, the desired kernel entry is approximated as a weighted linear combination of translated and scaled versions of impulse responses at the points $x'$ (Section~\ref{sec:approximate_kernel_entries}).
+%		the values $\phi_{x’}(y - \spatialmean(x) + \spatialmean(x’))$ for these nearby points $x'$.
+%	Right:  Illustration of the radial basis function interpolation method used to approximate $\Aker(y,x)$ by a weighted linear combination of the values $\frac{\spatialvol(x)}{\spatialvol(x’)}\phi_{x’}(y - \spatialmean(x) + \spatialmean(x’))$ for a small number of sample points near $x$.
+The weights are determined by radial basis function (RBF) interpolation.
+Right: The impulse response centered at $x’$.
 	}
@@
 
-	\caption{columnwise error plots for kernel reconstruction left 5, center 10, right 20
+	\caption{Relative error, $||\Phi(\cdot,x) - \widetilde{\Phi}(\cdot,x)||/||\Phi(\cdot,x)||$, in the approximation of the ``column'' of the integral kernel associated with $x$, using 5 (left), 10 (center) and 20 (right) impulse response batches.
+Sample points are indicated by black dots.
+The error associated with the point $x$ is the shade of the image at location $x$, with white indicating zero error and black indicating $100\%$ error.
+At the sample points, the error is zero.
+The further the point $x$ is from the samples points, the larger the error.
+Adding more batches yields a more accurate approximation.
+%		columnwise error plots for kernel reconstruction left 5, center 10, right 20
 	}
@@
 \subsection{Example 1: Spatially variant blur}
+\label{sec:frog}
 
@@
 Along the way, we illustrate steps using a spatially variant blur example, in which the integral kernel is given as follows:
-$$\Phi(y,x) = (1 - a f(y,x)) g(x) \exp \left(-\frac{1}{2}h^T(y,x) C^{-1} h(y,x)\right).$$ 
+\begin{equation}
+\label{eq:frog_kernel}
+\Phi(y,x) = (1 - a f(y,x)) g(x) \exp \left(-\frac{1}{2}h^T(y,x) C^{-1} h(y,x)\right).
+\end{equation} 
 Here, $h(y,x) = R(\theta(x)) (y-x)$, $R(\theta(x))$ is the 2D rotation matrix by angle $\theta(x) = (x^1 + x^2) \pi / 2$, $C = L^2 \operatorname{diag}(c_1, c_2)$, $g(x) = x^1(1-x^1)x^2(1-x^2)$,  and $f(x,y) = \cos\left(p^1(y,x) / \sqrt{c_1 / 2.0}\right) \sin\left(p^2(y,x) / \sqrt{c_2 / 2.0}\right)$ for some constants $a, L, c_1, c_2$.

### 2024-02-13 numerical_examples/blur/blur_length_scale_experiments.py
addglr and psf  results from frog vs length_scale

@@ -14,9 +14,12 @@ from tqdm.auto import tqdm
 max_batches = 100
 nx = 63
 a = 1.0 # <-- How bumpy? Use 1.0, which is maximum bumpiness without negative numbers
-all_length_scalings = [1.0 / (t**2) for t in [1.0, 2.0, 3.0]][::-1]
+all_length_scalings = [1.0 / (t**2) for t in [1.0, 2.0, 3.0]]
 all_error_levels = [0.2, 0.1, 0.05]
 # length_scaling = 1./(3.0**2) #1.0
+
+np.savetxt('all_length_scalings.txt', all_length_scalings)
+np.savetxt('all_error_levels.txt', all_error_levels)
 
 all_all_num_batches = []
 all_all_num_impulses = []
@@ -33,7 +36,7 @@ for length_scaling in all_length_scalings:
 
     #### Global low rank
 
-    _,ss,_ = np.linalg.svd(Ker)
+    _, ss, _ = np.linalg.svd(Ker)
     all_ss.append(ss)
     norm_ker_fro = np.linalg.norm(Ker)
 
@@ -63,6 +66,12 @@ for length_scaling in all_length_scalings:
         num_neighbors=10
     )
 
+    # print('Making row and column cluster trees')
+    # ct = hpro.build_cluster_tree_from_pointcloud(dof_coords, cluster_size_cutoff=32)
+    #
+    # print('Making block cluster trees')
+    # bct = hpro.build_block_cluster_tree(ct, ct, admissibility_eta=2.0)
+
     all_num_batches = []
     all_num_impulses = []
     all_fro_errors = []
@@ -70,6 +79,14 @@ for length_scaling in all_length_scalings:
         psf_object.add_impulse_response_batch()
         num_batches = psf_object.psf_object.impulse_response_batches.num_batches
         num_impulses = psf_object.psf_object.impulse_response_batches.num_sample_points
+
+        # H_psf_hmatrix, Ker_psf_hmatrix = psf_object.construct_hmatrices(bct)
+        #
+        # Ker_psf = np.zeros(Ker.shape)
+        # for ii in range(Ker_psf.shape[1]):
+        #     ei = np.zeros(Ker_psf.shape[1])
+        #     ei[ii] = 1.0
+        #     Ker_psf[:,ii] = Ker_psf_hmatrix * ei
 
         Ker_psf = psf_object.psf_object.psf_kernel.cpp_object.eval_integral_kernel_block(dof_coords.T, dof_coords.T)
[…]

### 2024-02-14 paper_sisc_first_revision/localpsf.tex
addglr and psf  results from frog vs length_scale
discussion with noemi and tucker, include HODLR

@@
 \end{figure}
-
-x1 = (0.5, 0.5)
-y = (y1, y2)
-Phi((y1, y2), (x1, x2)) = Phi(y,x)
 
@@
 The purpose of this example is to illustrate the method using an operator which is known exactly and may be easily visualized.
+
+Define HODLR. Define randomized HODLR algo. Ref to Tucker's paper and "Fast randomized eigensolver with structured LDL..."
+
+Define randomized SVD, refer to Halko Martinsson Tropp.
+
+Discuss why PSF does better with smaller $L$.
+
+\begin{table}
+	\begin{center}
+		{\small
+		\begingroup
+		\setlength{\tabcolsep}{8pt}
+		\renewcommand{\arraystretch}{1.0}
+		\begin{tabular}{c| c | c c c c c}
+						& Error tol.    &	\#applies PSF   & \#applies HODLR 	& rank \\ 
+			\hline 
+						& 20\%    		& 11  				& 592				& 37  \\ 
+			$L=1$ 		& 10\% 			& 15  				& 772				& 67  \\ 
+						& 5\% 			& 23   				& 924				& 100 \\ 
+			\hline
+						& 20\%    		& 8  				& 852 				& 143 \\ 
+			$L=1/2$ 	& 10\%    		& 10  				& 1144				& 256 \\ 
+						& 5\% 	  		& 12   				& 1404				& 377 \\ 
+			\hline
+						& 20\%    	 	& 7					& 932				& 294  \\ 
+			$L=1/3$ 	& 10\%    		& 8  				& 1264 				& 525  \\ 
+						& 5\% 			& 8   				& 1520 				& 767  \\ 
+		\end{tabular}
+		\endgroup
+	}
+	\end{center}
+	\caption{(Spatially variant blur) Comparison of cost to approximate the blur kernel from Equation \eqref{eq:frog_kernel} using the PSF method, the randomized HODLR (hierarchical off diagonal low rank) method, and GLR (global low rank) approximation using randomized SVD.
+The quantity $L$ scales the width of the impulse responses, hence it influences the rank of the operator.
+Large $L$ means low rank, and small $L$ means high rank.
+The second column (``Error tol'') is the relative error in the approximation of the kernel measured in the Frobenius norm, $||\Phi - \widetilde{\Phi}||_\text{Fro} / ||\Phi||_\text{Fro}$.
+The remaining three columns show the number of operator applies require to achieve the given error tolerance, using the PSF, HODLR, and GLR methods.}
+	\label{tab:frog_psf_vs_hodlr_vs_glr}
+\end{table}
 

### 2024-02-14 localpsf/blur.py

@@ -11,6 +11,7 @@ def frog_function(
     Sigma: np.ndarray, # shape=(2, 2)
     a: float, # how much negative to include. a=0: gaussian
     use_bump: bool = True,
+    rotation_rate: float = 1.0,
 ) -> np.ndarray: # shape=(N,)
     assert(xx.shape[1] == 2)
     N = xx.shape[0]
@@ -18,7 +19,7 @@ def frog_function(
     assert(mu.shape == (2,))
 
     # theta = np.pi * (mu[0] + mu[1])
-    theta = (np.pi / 2.0) * (mu[0] + mu[1])
+    theta = rotation_rate * (np.pi / 2.0) * (mu[0] + mu[1])
     Rot_matrix = np.array([[np.cos(theta), -np.sin(theta)], [np.sin(theta), np.cos(theta)]])
 
     bump = mu[0]*(1 - mu[0])*mu[1]*(1-mu[1]) if use_bump else 0.0
@@ -37,6 +38,7 @@ def frog_setup(
         nx:             int     = 63,
         length_scaling: float   = 1.0,
         a:              float   = 1.0,
+        rotation_rate:  float   = 1.0,
 ):
     ny = nx
     mesh = dl.UnitSquareMesh(nx, ny)
@@ -47,7 +49,7 @@ def frog_setup(
 
     Sigma = length_scaling * np.array([[0.01, 0.0], [0.0, 0.0025]]) # 0.25*np.array([[0.01, 0.0], [0.0, 0.0025]])
 
-    phi_function = lambda yy, x: frog_function(yy, vol, x, Sigma, a)
+    phi_function = lambda yy, x: frog_function(yy, vol, x, Sigma, a, rotation_rate=rotation_rate)
 
     Ker = np.zeros((Vh.dim(), Vh.dim()))
     for ii in range(Vh.dim()):

### 2024-02-14 numerical_examples/blur/blur_error_eigenvectors.py

@@ -1,0 +1,192 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import dolfin as dl
+import localpsf.localpsf_cg1_lumped as lpsf1
+from localpsf.blur import *
+import scipy.linalg as sla
+from functools import partial
+from nalger_helper_functions import plot_ellipse
+import hlibpro_python_wrapper as hpro
+from tqdm.auto import tqdm
+
+
+num_batches = 50
+nx = 89 #63
+a = 1.0 # <-- How bumpy? Use 1.0, which is maximum bumpiness without negative numbers
+length_scaling = 1.0 / (2.0**2)
+tau = 4.0
+num_neighbors=10
+rotation_rate = 1.0
+
+
+Ker, phi_function, Vh, H, mass_lumps, dof_coords, kdtree_sort_inds = frog_setup(nx, length_scaling, a, rotation_rate=rotation_rate)
+
+psf_object = lpsf1.make_psf_fenics(
+    lambda X: H @ X,
+    lambda X: H.T @ X,
+    Vh, Vh,
+    mass_lumps, mass_lumps,
+    num_initial_batches=num_batches,
+    tau=tau, display=True,
+    num_neighbors=num_neighbors
+)
+
+k=510
+col_func = dl.Function(Vh)
+col_func.vector()[:] = Ker[:,k].copy()
+plt.figure()
+cm = dl.plot(col_func)
+plt.colorbar(cm)
+plt.title('col_func, k=' +str(k))
+
+sp = psf_object.psf_object.impulse_response_batches.sample_points.copy()
+
+Ker_psf = psf_object.psf_object.psf_kernel.cpp_object.eval_integral_kernel_block(dof_coords.T, dof_coords.T)
+
+err_psf = np.linalg.norm(Ker - Ker_psf) / np.linalg.norm(Ker_psf)
+print('err_psf=', err_psf)
+
+#
+
+ker_err_func = dl.Function(Vh)
+ker_err_vec = np.linalg.norm(Ker_psf - Ker, axis=0) / np.linalg.norm(Ker, axis=0)
[…]

### 2024-02-14 numerical_examples/blur/blur_num_neighbors_comparison.py

@@ -1,0 +1,82 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.gridspec as gridspec
+import dolfin as dl
+import localpsf.localpsf_cg1_lumped as lpsf1
+from localpsf.blur import *
+import scipy.linalg as sla
+from functools import partial
+from nalger_helper_functions import plot_ellipse
+import hlibpro_python_wrapper as hpro
+from tqdm.auto import tqdm
+
+
+max_batches = 30
+nx = 63
+a = 1.0 # <-- How bumpy? Use 1.0, which is maximum bumpiness without negative numbers
+length_scaling = 1.0 / (2.0**2)
+tau = 4.0
+all_num_neighbors = [1,2,3,5,7,10,15,25,50]
+
+np.savetxt('all_num_neighbors.txt', all_num_neighbors)
+
+Ker, phi_function, Vh, H, mass_lumps, dof_coords, kdtree_sort_inds = frog_setup(nx, length_scaling, a)
+
+all_all_num_batches = []
+all_all_num_impulses = []
+all_all_fro_errors = []
+for num_neighbors in all_num_neighbors:
+    print('num_neighbors=', num_neighbors)
+
+    psf_object = lpsf1.make_psf_fenics(
+        lambda X: H @ X,
+        lambda X: H.T @ X,
+        Vh, Vh,
+        mass_lumps, mass_lumps,
+        num_initial_batches=max_batches,
+        tau=tau, display=True,
+        num_neighbors=num_neighbors
+    )
+
+    all_num_batches = []
+    all_num_impulses = []
+    all_fro_errors = []
+    # for num_batches in range(max_batches):
+    if True:
+        # psf_object.add_impulse_response_batch()
+        num_batches = psf_object.psf_object.impulse_response_batches.num_batches
+        num_impulses = psf_object.psf_object.impulse_response_batches.num_sample_points
+
+        Ker_psf = psf_object.psf_object.psf_kernel.cpp_object.eval_integral_kernel_block(dof_coords.T, dof_coords.T)
+        err_psf = np.linalg.norm(Ker_psf - Ker) / np.linalg.norm(Ker)
+        print('err_psf=', err_psf)
+        print('num_batches=', num_batches)
+        print('num_impulses=', num_impulses)
+
+        all_fro_errors.append(err_psf)
[…]

### 2024-02-14 numerical_examples/blur/blur_length_scale_experiments.py
fixed sqrt bug when calculating GLR FRO rank

@@ -18,6 +18,8 @@ all_length_scalings = [1.0 / (t**2) for t in [1.0, 2.0, 3.0]]
 all_error_levels = [0.2, 0.1, 0.05]
 # length_scaling = 1./(3.0**2) #1.0
 
+do_PSF = False
+
 np.savetxt('all_length_scalings.txt', all_length_scalings)
 np.savetxt('all_error_levels.txt', all_error_levels)
 
@@ -42,7 +44,7 @@ for length_scaling in all_length_scalings:
 
     all_glr_ranks_fro = []
     for rtol in all_error_levels:
-        glr_num_fro = np.sum((norm_ker_fro**2 - np.cumsum(ss**2)) > rtol * norm_ker_fro**2)
+        glr_num_fro = np.sum((norm_ker_fro**2 - np.cumsum(ss**2)) > rtol**2 * norm_ker_fro**2)
         all_glr_ranks_fro.append(glr_num_fro)
         print('rtol=', rtol, ', glr_num_fro=', glr_num_fro)
     all_all_glr_ranks_fro.append(all_glr_ranks_fro)
@@ -54,62 +56,64 @@ for length_scaling in all_length_scalings:
         print('rtol=', rtol, ', glr_num_2=', glr_num_2)
     all_all_glr_ranks_2.append(all_glr_ranks_2)
 
-    #### PSF
-
-    psf_object = lpsf1.make_psf_fenics(
-        lambda X: H @ X,
-        lambda X: H.T @ X,
-        Vh, Vh,
-        mass_lumps, mass_lumps,
-        num_initial_batches=0,
-        tau=3.0, display=True,
-        num_neighbors=10
-    )
-
-    # print('Making row and column cluster trees')
-    # ct = hpro.build_cluster_tree_from_pointcloud(dof_coords, cluster_size_cutoff=32)
-    #
-    # print('Making block cluster trees')
-    # bct = hpro.build_block_cluster_tree(ct, ct, admissibility_eta=2.0)
-
-    all_num_batches = []
-    all_num_impulses = []
-    all_fro_errors = []
-    for num_batches in range(max_batches):
-        psf_object.add_impulse_response_batch()
-        num_batches = psf_object.psf_object.impulse_response_batches.num_batches
-        num_impulses = psf_object.psf_object.impulse_response_batches.num_sample_points
-
-        # H_psf_hmatrix, Ker_psf_hmatrix = psf_object.construct_hmatrices(bct)
[…]

### 2024-02-14 numerical_examples/blur/blur_length_scale_experiments.py
include rsvd data in paper

@@ -1,4 +1,5 @@
 import numpy as np
+import numpy.typing as typ
 import matplotlib.pyplot as plt
 import matplotlib.gridspec as gridspec
 import dolfin as dl
@@ -23,10 +24,50 @@ do_PSF = False
 np.savetxt('all_length_scalings.txt', all_length_scalings)
 np.savetxt('all_error_levels.txt', all_error_levels)
 
+
+def compute_rsvd_applies(A: np.ndarray, tol: float) -> int: # helper function
+    big_Omega = np.random.randn(A.shape[1], A.shape[1])
+    big_Y = A @ big_Omega
+    big_Q, big_R = np.linalg.qr(big_Y)
+    # assert(np.linalg.norm(big_R - np.triu(big_R)) < 1e-10 * np.linalg.norm(big_R)) # make sure there was no pivoting
+
+    big_B = big_Q.T @ A
+
+    def rsvd_error(k: int) -> float:
+        Q = big_Q[:, :k]
+        B = big_B[:k, :]
+        err_rsvd = np.linalg.norm(Q @ B - Ker) / np.linalg.norm(Ker)
+        # print('k=', k, ', err_rsvd=', err_rsvd)
+        return err_rsvd
+
+    k0 = 0
+    k1 = A.shape[1]
+    err0 = rsvd_error(k0)
+    err1 = rsvd_error(k1)
+    k = int((k1 + k0) / 2)
+    while k0 < k and k < k1:
+        err = rsvd_error(k)
+        print('k0=', k0, ', err0=', err0)
+        print('k=', k, ', err=', err)
+        print('k1=', k1, ', err1=', err1)
+        print()
+        if err < tol:
+            k1 = k
+            err1 = err
+        else:
+            k0 = k
+            err0 = err
+        k = int((k1 + k0) / 2)
+
+    rank = k1
+    num_applies_rsvd = 2 * rank
+    return num_applies_rsvd
+
 all_all_num_batches = []
 all_all_num_impulses = []
 all_all_fro_errors = []
 all_ss = []
+all_all_rsvd_applies = []
 all_all_glr_ranks_fro = []
 all_all_glr_ranks_2 = []
 for length_scaling in all_length_scalings:
@@ -41,6 +82,13 @@ for length_scaling in all_length_scalings:
     _, ss, _ = np.linalg.svd(Ker)
     all_ss.append(ss)
     norm_ker_fro = np.linalg.norm(Ker)
[…]

### 2024-02-15 localpsf/blur.py
rsvd data seems correct

@@ -64,3 +64,22 @@ def frog_setup(
     return Ker, phi_function, Vh, H, mass_lumps, dof_coords, kdtree_sort_inds
 
 nearest_ind_func = lambda yy, x: np.argmin(np.linalg.norm(yy - x.reshape((1, -1)), axis=1))
+
+
+def ricker_function(
+    xx: np.ndarray, # shape=(N, 2), N=nx*ny
+    mu: np.ndarray, # shape=(2,)
+    Sigma: np.ndarray, # shape=(2, 2)
+    a: float, # how much negative to include. a=0: gaussian
+) -> np.ndarray: # shape=(N,)
+    assert(xx.shape[1] == 2)
+    N = xx.shape[0]
+    assert(Sigma.shape == (2, 2))
+    assert(mu.shape == (2,))
+
+    pp = xx - mu.reshape((1,2))
+    inv_Sigma = np.linalg.inv(Sigma)
+    t_squared_over_sigma_squared = np.einsum('ai,ai->a', pp, np.einsum('ij,aj->ai', inv_Sigma, pp))
+    G = (2.0 * np.pi * np.sqrt(np.linalg.det(Sigma))) * np.exp(-0.5 * t_squared_over_sigma_squared)
+
+    return (1.0 + a * t_squared_over_sigma_squared) * G

### 2024-02-15 numerical_examples/blur/blur_length_scale_experiments.py

@@ -36,10 +36,11 @@ def compute_rsvd_applies(A: np.ndarray, tol: float) -> int: # helper function
     def rsvd_error(k: int) -> float:
         Q = big_Q[:, :k]
         B = big_B[:k, :]
-        err_rsvd = np.linalg.norm(Q @ B - Ker) / np.linalg.norm(Ker)
+        err_rsvd = np.linalg.norm(Q @ B - A) / np.linalg.norm(A)
         # print('k=', k, ', err_rsvd=', err_rsvd)
         return err_rsvd
 
+    # k0--------k---------k1
     k0 = 0
     k1 = A.shape[1]
     err0 = rsvd_error(k0)
@@ -166,4 +167,6 @@ for length_scaling in all_length_scalings:
 
 
 # all_all_rsvd_applies
-# Out[3]: [[354, 522, 674], [1318, 1918, 2454], [2622, 3734, 4662]]
+# Out[3]: [[354, 520, 674], [1316, 1916, 2456], [2624, 3734, 4660]]
+
+#

### 2024-02-15 revision/reply_to_reviewer2.tex
add kernel and impulse illustration
making interpolation figure
wrote up [52] h matrix discussion / back of envelope calculation

@@
   	Structure: local non-negative kernel
-(0) Matrix itself column highlighted, show impulse response
+(0) Matrix itself column highlighted, show impulse response (kernel\_impulse\_illustration.pdf)
 (1) moments. Show one impulse with ellipsoid with widths illustrated
@@
 (3) RBF interpolation (?), evaluation of entries. Columns we know, and entry we don't know. 
-(4) H-matrix
+(4) H-matrix (structure only)
 }
@@
       of mat-vecs needed to achieve a certain accuracy, something like
-      Figure 8 in teh HODLR paper. Nick will do some estimates to
+      Figure 8 in teh HODLR paper (DONE).
+
+      Nick will do some estimates to
       compare things with ref 52.}

### 2024-02-15 paper_sisc_first_revision/localpsf.tex
fixed sqrt bug when calculating GLR FRO rank
include rsvd data in paper
rsvd data seems correct
wrote up [52] h matrix discussion / back of envelope calculation

@@
 the required number of operator applications is still large (e.g., hundreds to thousands).
-In this paper, we also form an H-matrix Hessian approximation.
+For example, using the method in \cite{LinYing11}, the required number of operator applies to construct an $H^1$ matrix with hierarchical rank $r$ for problems in a 2D domain discretized with a regular grid is $\#\text{levels} \cdot 64 \cdot (r + c)$, where $\#\text{levels}$ is the depth of the hierarchical partitioning, $r$ is the rank of the blocks (hierarchical rank), and $c$ is an oversampling parameter (See \cite[Section 2.4]{LinYing11}).
+On a $64 \times 64$ grid with depth $4$, hierarchical rank $10$, and oversampling parameter $c=5$, the works out to $4 \cdot 64 \cdot (10 + 5) = 3840$ operator applies.
+In Section \ref{sec:frog} see numerically that the randomized hierarchical off-diagonal low rank (HODLR) method in \cite{Martinsson16} requires hundreds to thousands of matrix-vector products to construct approximations of blur operators with modest (e.g., $10\%$) relative error.
+%while the PSF method in this paper can achieve the same level of error in one to two dozen operator applies.
+In this paper, we also form an H-matrix approximation.
 However, to reduce the required number of operator applications, we first form a PSF approximation of the data misfit Hessian by exploiting locality and non-negative integral kernel properties, then form the H-matrix using classical H-matrix techniques.
@@
 	\begin{tabular}{c| c | c c c c c}
-						& Error tol.    &	\#applies PSF   & \#applies HODLR 	& rank \\ 
+				& Error tol.    &	\#applies PSF   & \#applies HODLR 	& \#applies RSVD 	\\%& rank \\ 
 	\hline 
-						& 20\%    		& 11  				& 592				& 37  \\ 
-			$L=1$ 		& 10\% 			& 15  				& 772				& 67  \\ 
-						& 5\% 			& 23   				& 924				& 100 \\ 
+				& 20\%    		& 11  				& 592				& 354				\\%& 111  \\ 
+	$L=1$ 		& 10\% 			& 15  				& 772				& 520				\\%& 178  \\ 
+				& 5\% 			& 23   				& 924				& 674				\\%& 244 \\ 
 	\hline
-						& 20\%    		& 8  				& 852 				& 143 \\ 
-			$L=1/2$ 	& 10\%    		& 10  				& 1144				& 256 \\ 
-						& 5\% 	  		& 12   				& 1404				& 377 \\ 
+				& 20\%    		& 8  				& 852 				& 1316				\\%& 417 \\ 
+	$L=1/2$ 	& 10\%    		& 10  				& 1144				& 1916				\\%& 665 \\ 
+				& 5\% 	  		& 12   				& 1404				& 2456				\\%& 903 \\ 
 	\hline
-						& 20\%    	 	& 7					& 932				& 294  \\ 
-			$L=1/3$ 	& 10\%    		& 8  				& 1264 				& 525  \\ 
-						& 5\% 			& 8   				& 1520 				& 767  \\ 
+				& 20\%    	 	& 7					& 932				& 2624				\\%& 845  \\ 
+	$L=1/3$ 	& 10\%    		& 8  				& 1264 				& 3734				\\%& 1321  \\ 
+				& 5\% 			& 8   				& 1520 				& 4660				\\%& 1759  \\ 
 	\end{tabular}

### 2024-02-15 revision/reply_to_reviewer2.tex
editing response with noemi

--- 2024-01-31 revision/reply_to_reviewer2.tex
@@
 
-  \noindent{Frog example with a=1.
-  	Structure: local non-negative kernel
-(0) Matrix itself column highlighted, show impulse response
-(1) moments. Show one impulse with ellipsoid with widths illustrated
-(2) impulse response batches
-(3) RBF interpolation (?)
-(4) H-matrix
+  \noindent{Thank you for this suggestion!
+We added a 2D blurring kernel example, with figures illustrating the end-to-end PSF method.
+These are shown in the introduction and the numerical results sections (Sections 1 and 7.1).
+We think that adding these illustrations helps the reader better understand the intuition behind the method.
+%  	Frog example with a=1.
+%  	Structure: local non-negative kernel
+%(0) Matrix itself column highlighted, show impulse response (kernel\_impulse\_illustration.pdf)
+%(1) moments. Show one impulse with ellipsoid with widths illustrated
+%(2) impulse response batches
+%(3) RBF interpolation (?), evaluation of entries. Columns we know, and entry we don't know. 
+%(4) H-matrix (structure only)
 }
@@
 
-    \noindent{For ref 7: Ask Tucker if we could do HODLR on the
-      Hessian coming from ice mountain problem and record the number
-      of mat-vecs needed to achieve a certain accuracy, something like
-      Figure 8 in teh HODLR paper. Nick will do some estimates to
-      compare things with ref 52.}
+    \noindent{We added a numerical comparison of our method with the HODLR method in [7] for a blur operator in Table 2 of Section 7.1 in the revised manuscript.
+The results show that the HODLR method requires hundreds to thousands of operator applies to achieve modest accuracies (e.g.
+10\% error) while our method requires only order 10 operator applies.
+In Section 2 of the revised manuscript we also added a discussion of the costs of the method proposed in [52].
+We present back of the envelope calculations based on complexity estimates given in [52], showing that thousands of operator applies may be required.
+%    	For ref 7: Ask Tucker if we could do HODLR on the
+%      Hessian coming from ice mountain problem and record the number
+%      of mat-vecs needed to achieve a certain accuracy, something like
+%      Figure 8 in teh HODLR paper (DONE).
+%      
+%      Nick will do some estimates to
+%      compare things with ref 52.
+  }
 

### 2024-02-15 paper_sisc_first_revision/localpsf.tex
editing response with noemi

@@
 the required number of operator applications is still large (e.g., hundreds to thousands).
-For example, using the method in \cite{LinYing11}, the required number of operator applies to construct an $H^1$ matrix with hierarchical rank $r$ for problems in a 2D domain discretized with a regular grid is $\#\text{levels} \cdot 64 \cdot (r + c)$, where $\#\text{levels}$ is the depth of the hierarchical partitioning, $r$ is the rank of the blocks (hierarchical rank), and $c$ is an oversampling parameter (See \cite[Section 2.4]{LinYing11}).
-On a $64 \times 64$ grid with depth $4$, hierarchical rank $10$, and oversampling parameter $c=5$, the works out to $4 \cdot 64 \cdot (10 + 5) = 3840$ operator applies.
+For example, using the method in \cite{LinYing11}, the required number of operator applies to construct an $H^1$ matrix with hierarchical rank $r$ for problems in a 2D domain discretized with a regular grid is $\#\text{levels} \cdot 64 \cdot (r + c)$, where $\#\text{levels}$ is the depth of the hierarchical partitioning, $r$ is the rank of the blocks (hierarchical rank), and $c$ is an oversampling parameter (see \cite[Section 2.4]{LinYing11}).
+On a $64 \times 64$ grid with depth $4$, hierarchical rank $10$, and oversampling parameter $c=5$, this works out to $4 \cdot 64 \cdot (10 + 5) = 3840$ operator applies.
 In Section \ref{sec:frog} see numerically that the randomized hierarchical off-diagonal low rank (HODLR) method in \cite{Martinsson16} requires hundreds to thousands of matrix-vector products to construct approximations of blur operators with modest (e.g., $10\%$) relative error.
 %while the PSF method in this paper can achieve the same level of error in one to two dozen operator applies.
+Matrix-free H-matrix construction is currently an active area of research, hence these costs may decrease as new algorithms are developed.
 In this paper, we also form an H-matrix approximation.
@@
 The second column (``Error tol'') is the relative error in the approximation of the kernel measured in the Frobenius norm, $||\Phi - \widetilde{\Phi}||_\text{Fro} / ||\Phi||_\text{Fro}$.
-The remaining three columns show the number of operator applies require to achieve the given error tolerance, using the PSF, HODLR, and GLR methods.}
+The remaining three columns show the number of operator applies required to achieve the given error tolerances, using the PSF, HODLR, and GLR methods.}
 	\label{tab:frog_psf_vs_hodlr_vs_glr}

### 2024-02-15 paper_sisc_first_revision/localpsf.tex
added color to frog negative impulse pictures

@@
 One limitation of our method is that not all data misfit Hessians satisfy the local non-negative integral kernel properties.
-We note, in particular, that the wave inverse problem data misfit Hessian has a substantial proportion of negative entries in its integral kernel.
+We note, in particular, that the wave inverse problem data misfit Hessian and Gauss-Newton Hessian have a substantial proportion of negative entries in their integral kernels.
 In this case more specialized techniques based on pseudodifferential operator theory have been developed, e.g., \cite{DemanetEtAl12,HerrmannMoghaddamStolk08,Symes08}.

### 2024-02-15 numerical_examples/blur/blur_ellipsoids_getting_worse.py

@@ -13,7 +13,7 @@ from tqdm.auto import tqdm
 
 nx = 49
 length_scaling = 1.0
-aa = [1.0, 20.0, 17.0]
+aa = [1.0, 20.0, 27.0]
 
 for a in aa:
     Ker, phi_function, Vh, H, mass_lumps, dof_coords, kdtree_sort_inds = frog_setup(nx, length_scaling, a)
@@ -48,18 +48,25 @@ for a in aa:
     rr_1d = phi_function(coords_1d, p)
 
     plt.figure()
-    plt.plot(rr_1d)
+    plt.plot(np.zeros(rr_1d.size), '--', c='gray')
+    plt.plot(rr_1d, c='k', linewidth=2)
     plt.gca().set_xticks([])
+    plt.gca().set_yticks([])
+    plt.xlim(0, len(coords_1d))
+
     # plt.title('rr_1d')
 
     plt.savefig('frog_1d_a=' + str(a) + '.png', bbox_inches='tight', dpi=400)
 
+    vmax = np.max(np.abs(phi.vector()[:]))
+
     plt.figure()
-    cm = dl.plot(phi, cmap='binary')
+    # cm = dl.plot(phi, cmap='binary')
+    cm = dl.plot(phi, cmap='bwr', vmin=-vmax, vmax=vmax)
     plt.gca().set_xticks([])
     plt.gca().set_yticks([])
     # plt.title('frog1')
-    plt.plot(coords_1d[:,0], coords_1d[:,1], '.r')
+    plt.plot(coords_1d[:,0], coords_1d[:,1], c='k', linewidth=1)
     plt.xlim(0.3, 0.7)
     plt.ylim(0.1, 0.9)
 
@@ -68,7 +75,7 @@ for a in aa:
                         [psf_object.Sigma(1,0)(p), psf_object.Sigma(1,1)(p)]])
 
     tau=3.0
-    plot_ellipse(mu_p, Sigma_p, n_std_tau=tau, facecolor='none', edgecolor='k', linewidth=1)
+    plot_ellipse(mu_p, Sigma_p, n_std_tau=tau, facecolor='none', edgecolor='k', linewidth=2)
 
 
     plt.savefig('frog_ellipsoid_a=' + str(a) + '.png', bbox_inches='tight', dpi=400)

### 2024-02-15 numerical_examples/blur/blur_ellipsoids_getting_worse.py
ricker vs frog ellipsoids figure added

@@ -11,7 +11,7 @@ import hlibpro_python_wrapper as hpro
 from tqdm.auto import tqdm
 
 
-nx = 49
+nx = 63
 length_scaling = 1.0
 aa = [1.0, 20.0, 27.0]
 
@@ -163,391 +163,3 @@ for a in aa:
 
 ##
 
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-
-raise RuntimeError
-
-#### Impulse response grid plot
-
-num_pts_x = 5
-num_pts_y = 5
-pX, pY = np.meshgrid(np.linspace(0,1,num_pts_x), np.linspace(0,1,num_pts_y)[::-1])
-pp = np.vstack([pX.reshape(-1), pY.reshape(-1)]).T
-
-plt.figure(figsize = (num_pts_x,num_pts_y))
-gs1 = gridspec.GridSpec(num_pts_x, num_pts_y)
-gs1.update(wspace=0.025, hspace=0.05) # set the spacing between axes.
-for k in range(len(pp)):
-    p = pp[k]
-    ii = nearest_ind_func(dof_coords, p)
-    phi = dl.Function(Vh)
-    phi.vector()[:] = Ker[:, ii].copy()
-    plt.subplot(gs1[k])
-    cm = dl.plot(phi, cmap='binary')
-    plt.gca().set_xticks([])
-    plt.gca().set_yticks([])
-
-plt.savefig('frog_impulse_response_grid.png', bbox_inches='tight', dpi=300)
-
-
-#### Plot two impulse responses individually, and kernel with the corresponding columns indicated
-
-p1 = np.array([0.2, 0.45])
-p2 = np.array([0.6, 0.6])
-pp = [p1, p2]
-
-plt.matshow(Ker, cmap='binary')
-plt.ylim([Ker.shape[0], 0])
-plt.gca().set_xticks([])
-plt.gca().set_yticks([])
-
-for k in range(len(pp)):
-    p = pp[k]
-    ii = nearest_ind_func(dof_coords, p)
-    print('k=', k, ', ii=', ii)
-    plt.plot([ii, ii], [0, Ker.shape[0]], 'k', linestyle='dotted', linewidth=2.0)
-
-plt.gcf().set_size_inches(8, 8)
-plt.savefig('frog_kernel_matrix_a1.png', bbox_inches='tight', dpi=400)
-
-for k in range(len(pp)):
-    p = pp[k]
-    ii = nearest_ind_func(dof_coords, p)
[…]

### 2024-02-16 paper_sisc_first_revision/localpsf.tex
ricker vs frog ellipsoids figure added
added discussion of what happens when there are negative numbers in the integral kernel

@@
 The proposed method may still perform well if these conditions are relaxed slightly.
-It is acceptable if the support of $\impulseresponse_x$ is not perfectly contained in a localized region, so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
-The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$ so long as it is non-negative for the vast majority of pairs of points $(y,x)$, and so long as the negative numbers are comparatively small.
-If these conditions are violated, the method will incur additional error, and depending on the severity of the violation, the proposed method may fail.
+
+%If these conditions are violated, the method will incur additional error, and depending on the severity of the violation, the proposed method may fail.
+
+It is acceptable if the support of $\impulseresponse_x$ is not perfectly contained in a localized region (violating Assumption 2), so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
+In this case the method can still be applied, but the ellipsoid impulse response support estimates will be large.
+Therefore, only a small number of impulse responses can be computed per batch.
+This means that more batches, and hence more operator applies, are needed to form an accurate approximation.
+
+%The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$ (Assumption 3) so long as it is non-negative for the vast majority of pairs of points $(y,x)$, and so long as the negative numbers are comparatively small.
+
+If there are negative numbers in the integral kernel (violating Assumption 3), the ellipsoid estimation procedure may incur errors or fail, leading to poor performance or failure of the PSF method.
+In Figure \ref{fig:robustness_to_negatives} we investigate the robustness of the ellipsoid support estimation procedure to violations of Assumption 3.
+We study two integral kernel examples, both of which are parameterized by a quantity that controls how negative the kernels are.
+We make the following observations:
+\begin{itemize}
+	\item The larger and more numerous the negative numbers are, the more inaccurate the ellipsoid support estimate is.
+	\item The further away from the center of the ellipsoid the negative numbers are, the more influence they have on the quality of the ellipsoid support estimate.
+This is because moment formulas (Equations \ref{eq:define_mean} and \ref{eq:define_cov}) assign more weight to entries in the kernel that are further from the center.
+	\item Negative numbers in the integral kernel affect the method more if they are isolated, and less if they are balanced by nearby positive numbers.
+	\item As kernels become more negative, the ellipsoid estimation performs well up to a certain threshold that depends on the spatial distribution of negative and positive entries.
+After that threshold is crossed, the estimation rapidly transitions to performing poorly and ultimately failing.
+\end{itemize} 
+%The location of the negative numbers also affects the accuracy of the ellipsoid estimates substantially, but in a more complicated manner. 
+For the kernel in the left two columns of Figure \ref{fig:robustness_to_negatives}, negative numbers are interspersed with positive numbers, allowing us to include a large amount of negative numbers in the integral kernel before the method fails.
+For the kernel in the right two columns, the ellipsoid estimate fails with tiny amounts of negative numbers because the negative numbers are far away from the mean and not balanced by positive numbers at similar distances and angles.
+In the bottom two rows, we see that the method rapidly transitions from performing reasonably well to performing poorly with only a small change to the integral kernel.
+
+%Location and magnitude of negatives matter.
+Larger negatives: bad, smaller negatives good.
+Negatives closer to center: good, further from center: bad.
+Correlated negatives: bad, randomly distributed: good.
+Good Good Good Fail quick
+%
+%Non-local: will work, but choose few ellipsoids, so inefficient
+
+\begin{figure}
+	\centering
+	{
+		%	\setlength{\tabcolsep}{0.1em}
+		%	\renewcommand{\arraystretch}{2.0}
+		\begin{tabular}{cccc}
+	\includegraphics[scale=0.3, valign=c]{frog_ellipsoid_a=1.0.png} & 
+	\includegraphics[scale=0.3, valign=c]{frog_1d_a=1.0.png} & 
+	\includegraphics[scale=0.3, valign=c]{ricker_ellipsoid_a=0.0.png} & 
+	\includegraphics[scale=0.3, valign=c]{ricker_1d_a=0.0.png} \\
+	%
+	\includegraphics[scale=0.3, valign=c]{frog_ellipsoid_a=20.0.png} & 
+	\includegraphics[scale=0.3, valign=c]{frog_1d_a=20.0.png} & 
+	\includegraphics[scale=0.3, valign=c]{ricker_ellipsoid_a=0.23.png} & 
+	\includegraphics[scale=0.3, valign=c]{ricker_1d_a=0.23.png} \\
+	%
+	\includegraphics[scale=0.3, valign=c]{frog_ellipsoid_a=27.0.png} & 
+	\includegraphics[scale=0.3, valign=c]{frog_1d_a=27.0.png} & 
+	\includegraphics[scale=0.3, valign=c]{ricker_ellipsoid_a=0.249.png} & 
+	\includegraphics[scale=0.3, valign=c]{ricker_1d_a=0.249.png} 
+	\end{tabular}
+	}
+	\caption{Illustration of the influence of negative numbers in the integral kernel on the robustness of the ellipsoid estimates for the supports of impulse responses.
+Left two columns: spatially variant blurring kernel given in Equation \ref{eq:frog_kernel}.
+Right two columns: Ricker wavelet-type kernel given by $\Aker(y,x) = (1 - a \theta)\exp\left(-\theta/2\right)$, where $\theta=(y-x)^T \Sigma^{-1} (y-x)$, and $\Sigma=\operatorname{diag}\left(0.0025, 0.01\right)$.
+Ordered from top to bottom, the results are obtained with $a\in\{1.0, 20.0, 27.0\}$ for the blur kernel, and $a\in \{0.0, 0.23, 0.249\}$ for the Ricker wavelet-type kernel.
+Columns 1 and 3: impulse responses for the blur and Ricker wavelet-type kernels, respectively, with estimated support ellipsoids indicated by the black ellipses.
+Red and blue represent positive and negative numbers in the integral kernel, respectively.
+Columns 2 and 4: one-dimensional slice along the horizontal line indicated in the two-dimensional plots.
+The dashed gray line is at zero.
+	}
+	\label{fig:robustness_to_negatives}
+\end{figure}
 
@@
 Alger, Longfei Gao, Mathew Hu, and Rami Nammour for helpful discussions.
-The seismic Hessian impulse response image was provided to us by Mathew Hu.
-We thank Trevor Heise for editing suggestions.
+We thank Nicole Aretz and Trevor Heise for editing suggestions.
 We thank Georg Stadler for help with the domain setup for the ice sheet problem.

### 2024-02-16 revision/reply_to_reviewer1.tex
add seismic example. fix reply to reviewer number bug
draft after discussion with Noemi and Tucker

@@
 
-\noindent{(1) Three-figure plot: Stokes, Advection/diffusion, wave impulse responses.
+\noindent{
+	We added a numerical study investigating the robustness of the method to violations of the non-negative kernel assumption on two 2D blurring kernel examples (Figure 7).
+One takeaway message from this figure is that the impact of negative numbers depends strongly on the spatial configuration of these negative numbers---in the first example (columns 1 and 2 in Figure 7), the negative numbers have minimal impact on the ellipsoid impulse response support estimates, whereas in the other example (columns 3 and 4), even a small amount of negative numbers caused the ellipsoid estimate to be poor.
+The distribution of negative numbers for wave inverse problem Hessians is similar to that of the second example, and therefore out method is unlikely to be effective for wave Hessians.
+This is also the case with the Gauss-Newton Hessian.
 
-(2) showing that the method performs poorly for the ricker example
+	Furthermore, although the component of the wave inverse problem Hessian corresponding to reflected waves is local, the component corresponding to transmission waves is not local.
+This non-locality also poses challenges for our method (as discussed in Section 3.2).
+Perhaps this issue caused by transmission waves could be addressed by frequency filtering (thereby removing the non-local component).
+We do not pursue this here.
 
-Nick will try to get an impulse response for the wave
-  problem.}
+	NICK ADD PLOT
+%	
+%	In particular, we note the poor performance for the Ricker wavelet-type kernel (right two columns of Figure 7).
+We observed a similar configuration for the impulse responses for the Hessian and Gauss-Newton Hessian in wave inverse problems.
+
+%	 The configuration of negative numbers in this example is similar The negative numbers in the integral kernel for the wave Hessian (and Gauss-Newton Hessian) are distributed in a manner similar to the Ricker wavelet-type impulse response shown in the right two columns of Figure 7.
+%	
+%	 We also added a discussion of the impact of non-locality and negativity on the performance of the method (Section 3.2).
+%	
+%	The 
+%	
+%	(1) Three-figure plot: Stokes, Advection/diffusion, wave impulse responses. 
+%	
+%(2) showing that the method performs poorly for the ricker example
+%	
+%Nick will try to get an impulse response for the wave
+%  problem.
+}
 
@@
   \noindent{
-  	Frog example with a=1. Error vs #batches, for different tau. Also do a plot with #impulses on the x-axis.
+  	Frog example with a=1. Error vs \#batches, for different tau. Also do a plot with \#impulses on the x-axis.
 

### 2024-02-16 revision/reply_to_reviewer2.tex
draft after discussion with Noemi and Tucker

@@
 
-  \noindent{Use a 2D-deblurring problem and cook up a Hessian with
-    entries given by be some modified Gaussian and show that the error
-    increases as more negative entires are present in the kernel.}
+  \noindent{
+  	We added a numerical study investigating the robustness of the method to violations of the non-negative kernel assumption on two 2D blurring kernel examples (Figure 7).
+We also added a discussion of the impact of non-locality and negativity on the performance of the method (Section 3.2).
+%  	 Figure 7 gives insight into how negative numbers in the integral kernel affect the ellipsoid impulse response support estimatates.
+%  	
+%  	Use a 2D-deblurring problem and cook up a Hessian with
+%    entries given by be some modified Gaussian and show that the error
+%    increases as more negative entires are present in the kernel.
+}
 

### 2024-02-17 revision/reply_to_reviewer1.tex
add marmousi stuff to response

@@
 \usepackage{amsmath}
+%\usepackage{subcaption}
 
@@
 	We added a numerical study investigating the robustness of the method to violations of the non-negative kernel assumption on two 2D blurring kernel examples (Figure 7).
-One takeaway message from this figure is that the impact of negative numbers depends strongly on the spatial configuration of these negative numbers---in the first example (columns 1 and 2 in Figure 7), the negative numbers have minimal impact on the ellipsoid impulse response support estimates, whereas in the other example (columns 3 and 4), even a small amount of negative numbers caused the ellipsoid estimate to be poor.
-The distribution of negative numbers for wave inverse problem Hessians is similar to that of the second example, and therefore out method is unlikely to be effective for wave Hessians.
-This is also the case with the Gauss-Newton Hessian.
+We also added a discussion of the impact of non-locality and negativity on the performance of the method (Section 3.2).
 
-	Furthermore, although the component of the wave inverse problem Hessian corresponding to reflected waves is local, the component corresponding to transmission waves is not local.
-This non-locality also poses challenges for our method (as discussed in Section 3.2).
-Perhaps this issue caused by transmission waves could be addressed by frequency filtering (thereby removing the non-local component).
-We do not pursue this here.
+	One takeaway message from Figure 7 is that the impact of negative numbers depends strongly on the spatial configuration of these negative numbers---in the first example (columns 1 and 2 in Figure 7), the negative numbers have minimal impact on the ellipsoid impulse response support estimates, whereas in the second example (columns 3 and 4), even a small amount of negative numbers causes the impulse response support estimate to be poor.
+The configuration of negative numbers for the integral kernel in wave inverse problem Hessians (including Gauss-Newton Hessians) is similar to that of the second example, and therefore our method is unlikely to be effective for wave Hessians.
 
-	NICK ADD PLOT
+	Although the component of the wave inverse problem Hessian corresponding to reflected waves is local, the component corresponding to transmission waves is non-local, which also poses challenges for our method (as discussed in the revised version of Section 3.2).
+If one applied a high-pass filter to remove the transmission component, in principle the impulse responses could be localized.
+However, the non-negativity would still need to be addressed.
+
+	In Figure \ref{fig:marmousi_impulse} here (but not in the paper due to page limitations), we show an impulse response for the Gauss-Newton Hessian for the Marmousi wave inverse problem.
+Note that the non-local component of the impulse response causes the ellipsoid impulse response support estimate to be effectively the entire domain.
+
+	\begin{figure}
+		\centering
+		\includegraphics[scale=0.735]{marmousi_impulse_response_full_domain_ellipsoid_with_tau_annotations.png}
+		\includegraphics[scale=0.3975]{marmousi_impulse_response_zoomed_in.png}
+
+		\caption{Impulse response for the Gauss-Newton Hessian for the Marmousi wave inverse problem.
+Data provided by Mathew Hu.
+On the left we show the impulse response on the whole domain.
+The color scale is adjusted for better visualization of the non-local transmission components.
+Estimated ellipsoids with $\tau\in \{1.0, 2.0, 3.0\}$  are shown.
+On the right we show a zoomed in impulse response, showcasing the negative numbers in the integral kernel.
+		}
+		\label{fig:marmousi_impulse}
+	\end{figure}
+
+%	Perhaps this issue could be addressed by applying a high-pass filter to remove the non-local transmission component.
+%	 The transmission component is typically low rank and can therefore be addressed with other methods. 
+%	We do not pursue this here.
+
+%	NICK ADD PLOT
 %	
@@
   \noindent{
+  	Figure 6 illustrates how the error decreases with increasing numbers of batches and therefore impulse responses. ALSO TABLE 4 CONDITION NUMBER
+
   	Frog example with a=1. Error vs \#batches, for different tau. Also do a plot with \#impulses on the x-axis.

### 2024-02-17 revision/reply_to_reviewer1.tex
small updates

@@
 	In Figure \ref{fig:marmousi_impulse} here (but not in the paper due to page limitations), we show an impulse response for the Gauss-Newton Hessian for the Marmousi wave inverse problem.
-Note that the non-local component of the impulse response causes the ellipsoid impulse response support estimate to be effectively the entire domain.
+Note that the non-local component of the impulse response shifts the mean far away from the visual ``center'' of the impulse response, and causes the ellipsoid impulse response support estimate to be effectively the entire domain.
 
@@
 		\caption{Impulse response for the Gauss-Newton Hessian for the Marmousi wave inverse problem.
-Data provided by Mathew Hu.
 On the left we show the impulse response on the whole domain.
-The color scale is adjusted for better visualization of the non-local transmission components.
-Estimated ellipsoids with $\tau\in \{1.0, 2.0, 3.0\}$  are shown.
+The color scale is truncated for better visualization of the non-local transmission components.
+The estimated mean of the impulse response is shown as a red dot.
+Estimated support ellipsoids with $\tau\in \{1.0, 2.0, 3.0\}$  are shown as black ellipses.
 On the right we show a zoomed in impulse response, showcasing the negative numbers in the integral kernel.
+The data for this impulse response was kindly provided to us by Mathew Hu.
 		}

### 2024-02-17 paper_sisc_first_revision/localpsf.tex
add marmousi stuff to response
small updates

@@
 %AntilEtAl18
-Schur complements in Schur complement methods for solving partial differential equations (PDEs) and Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps)~\cite{ChanMathew94,SmithBjorstadGropp96,ToselliWidlund04}, covariance operators in spatial statistics~\cite{ChenStein21,GentonKeyesTurkiyyah18,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}.
+Schur complements in Schur complement methods for solving partial differential equations and Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps)~\cite{ChanMathew94,SmithBjorstadGropp96,ToselliWidlund04}, covariance operators in spatial statistics~\cite{ChenStein21,GentonKeyesTurkiyyah18,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}.
 Here ``matrix-free'' means that we may apply $\Aop$ and its transpose\footnote{Recall that $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for operations with $\Aop^T$.
@@
 Evaluating the maps in~\eqref{eq:eval_maps} may require solving a subproblem that involves PDEs, or performing other costly computations.
-By ``non-negative integral kernel,'' we mean that entries of the integral kernel are non-negative numbers; this is not the same as positive semi-definiteness of $\Aop$.
+By ``non-negative integral kernel,'' we mean that entries of the integral kernel (defined in Equation \ref{eq:kernel_representation}) are non-negative numbers; this is not the same as positive semi-definiteness of $\Aop$.
 
@@
 The proposed method improves upon existing PC methods in the following ways: (1) While PC approximations are based on an assumption of local translation invariance, the method we propose is based on a more general assumption we call ``local mean displacement invariance'' (Section~\ref{sec:local_mean_displacement_invariance} and Figure~\ref{fig:mean_displacement_invariance}), which improves the interpolation of the impulse responses.
-(2) In our previous work~\cite{AlgerEtAl19}, we chose point sources optimally via a sequential adaptive procedure.
+(2) In our previous work~\cite{AlgerEtAl19}, we chose point sources in adaptive grid via a sequential procedure; the refinements to the adaptive grid were chosen to reduce the error as much as possible at each step.
 However, in that work each point source required a separate operator application, making the previous method expensive when a large number of impulse responses is desired.
@@
 Alger, Longfei Gao, Mathew Hu, and Rami Nammour for helpful discussions.
-We thank Nicole Aretz and Trevor Heise for editing suggestions.
+We thank Nicole Aretz, Trevor Heise, and the anonymous reviewers for editing suggestions.
 We thank Georg Stadler for help with the domain setup for the ice sheet problem.

### 2024-02-17 paper_sisc_first_revision/localpsf.tex
addressing issues raised by Nicole

@@
 The results show that the PSF-based preconditioners are able to form good approximations of high-rank Hessians using only a small number of operator applications.
-
-	TEST ASDF
 \end{abstract}
@@
 Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, and $L^2(\Omega)'$ is the space of real-valued continuous linear functionals on $L^2(\Omega)$.
-Such operators appear, for instance, as Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{BorziSchulz11,Delosreyes15,HinzePinnauUlbrichUlbrich08},
+By ``non-negative integral kernel,'' we mean that entries of $\Aop$'s integral kernel (defined in Equation \ref{eq:kernel_representation}) are non-negative numbers; this is not the same as positive semi-definiteness of $\Aop$.
+Such operators appear as, for example, Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{BorziSchulz11,Delosreyes15,HinzePinnauUlbrichUlbrich08},
 %AntilEtAl18
 Schur complements in Schur complement methods for solving partial differential equations and Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps)~\cite{ChanMathew94,SmithBjorstadGropp96,ToselliWidlund04}, covariance operators in spatial statistics~\cite{ChenStein21,GentonKeyesTurkiyyah18,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}.
+
 Here ``matrix-free'' means that we may apply $\Aop$ and its transpose\footnote{Recall that $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for operations with $\Aop^T$.
@@
 Evaluating the maps in~\eqref{eq:eval_maps} may require solving a subproblem that involves PDEs, or performing other costly computations.
-By ``non-negative integral kernel,'' we mean that entries of the integral kernel (defined in Equation \ref{eq:kernel_representation}) are non-negative numbers; this is not the same as positive semi-definiteness of $\Aop$.
+Matrix-free does not mean that matrices are never used; our approximation is ultimately converted into a hierarchical matrix (H-matrix), a compressed matrix format that permits fast and scalable linear algebra operations.
 
@@
 
-We use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applications.
+The idea of the method is to use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applications.
 The impulse response, $\phi_x$, associated with a point, $x$, is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
@@
 The ability to rapidly approximate entries of $\Aop$'s integral kernel allows one to approximate discretized versions of $\Aop$ using the full arsenal of tools for matrix approximation that rely on fast access to matrix entries.
-In this work, we form a hierarchical matrix (H-matrix)~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of a discretized version of $\Aop$.
-H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank.
+In this work, we form a hierarchical matrix~\cite{BormGrasedyckHackbusch03,Hackbusch99} approximation of a discretized version of $\Aop$.
+H-matrices are a compressed matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high rank.
 H-matrix methods permit us to perform matrix-vector products cheaply, and perform other useful linear algebra operations that cannot be done easily using the original operator.
 These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion.
-The work and memory required to perform these operations for an $\fedim \times \fedim$ H-matrix scales 
-\emph{nearly linearly} in $N$ (i.e., $o(N^{1+\epsilon})$ for any $\epsilon > 0$).
-The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks~\cite{GrasedyckHackbusch03}.
+The work and memory required to perform these operations for an $\fedim \times \fedim$ H-matrix with rank $\hrank$ blocks scales as $O\left(\hrank^a \fedim \log(\fedim)^b\right)$ where $a,b \in \{0,1,2,3\}$ are constants which depend on the type of H-matrix used and the operation being performed~\cite{GrasedyckHackbusch03}\cite[Section 2.1]{Kriemann13}.
 
@@
 \label{sec:hessian}
+
+While the PSF method in this paper may be used to approximate any operator that has a locally supported non-negative integral kernel, we are primarily motivated by approximation of high-rank Hessians in distributed parameter inverse problems governed by PDEs.
+In this section, we provide a brief background on this topic, and explain why existing Hessian approximation methods are not satisfactory.
 
@@
 In target applications, (1) is the dominant cost because applying $\Aop$ to a vector requires an expensive computational procedure such as solving a PDE, and (1) is the only step that requires applying $\Aop$ to vectors.
-All operations that do not require applications of $\Aop$ to vectors are nearly linear, and therefore scalable, in the size of the problem, $N$.
+All operations that do not require applications of $\Aop$ to vectors are polylog linear, and therefore scalable, in the size of the problem, $N$.
 We now describe these costs in detail. For convenience, Table~\ref{tab:vars} lists variable symbols and their approximate sizes.
@@
 Constructing the H-matrix from kernel entries requires a number of elementary operations that scales polylog linearly in $N$.
-Using the resulting H-matrix to perform further linear algebra operations also scales nearly linearly in $N$, though the details of these costs depend heavily on the type of H-matrix and operation being performed.
+Using the resulting H-matrix to perform further linear algebra operations also scales polylog linearly in $N$, though the details of these costs depend heavily on the type of H-matrix and operation being performed.
 

### 2024-02-17 paper_sisc_first_revision/localpsf.tex
addressing issues raised by Nicole

@@
 It is acceptable if the support of $\impulseresponse_x$ is not perfectly contained in a localized region (violating Assumption 2), so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
-In this case the method can still be applied, but the ellipsoid impulse response support estimates will be large.
-Therefore, only a small number of impulse responses can be computed per batch.
-This means that more batches, and hence more operator applies, are needed to form an accurate approximation.
+In principle, the method can be applied even if the impulse responses are not localized.
+However, in this case only a small number of impulse responses can be computed per batch, which means more batches, and hence more operator applies, are needed to form an accurate approximation.
 
@@
 
-If there are negative numbers in the integral kernel (violating Assumption 3), the ellipsoid estimation procedure may incur errors or fail, leading to poor performance or failure of the PSF method.
+If there are negative numbers in the integral kernel (violating Assumption 3), the ellipsoid estimation procedure may incur errors or fail, leading to poor performance or failure of the method.
 In Figure \ref{fig:robustness_to_negatives} we investigate the robustness of the ellipsoid support estimation procedure to violations of Assumption 3.
@@
 For the kernel in the right two columns, the ellipsoid estimate fails with tiny amounts of negative numbers because the negative numbers are far away from the mean and not balanced by positive numbers at similar distances and angles.
-In the bottom two rows, we see that the method rapidly transitions from performing reasonably well to performing poorly with only a small change to the integral kernel.
+In the bottom two rows, we see the aforementioned threshold effect, in which the method rapidly transitions from performing reasonably well to performing poorly with only a small change to the integral kernel.
 
@@
 	\caption{Illustration of the influence of negative numbers in the integral kernel on the robustness of the ellipsoid estimates for the supports of impulse responses.
-Left two columns: spatially variant blurring kernel given in Equation \ref{eq:frog_kernel}.
+Left two columns: spatially variant blur kernel given in Equation \ref{eq:frog_kernel}.
 Right two columns: Ricker wavelet-type kernel given by $\Aker(y,x) = (1 - a \theta)\exp\left(-\theta/2\right)$, where $\theta=(y-x)^T \Sigma^{-1} (y-x)$, and $\Sigma=\operatorname{diag}\left(0.0025, 0.01\right)$.
-Ordered from top to bottom, the results are obtained with $a\in\{1.0, 20.0, 27.0\}$ for the blur kernel, and $a\in \{0.0, 0.23, 0.249\}$ for the Ricker wavelet-type kernel.
-Columns 1 and 3: impulse responses for the blur and Ricker wavelet-type kernels, respectively, with estimated support ellipsoids indicated by the black ellipses.
+Ordered from top to bottom, the results are obtained with $a\in\{1.0, 20.0, 27.0\}$ for the left two columns, and $a\in \{0.0, 0.23, 0.249\}$ for the right two columns.
+Columns 1 and 3: impulse responses with estimated support ellipsoids indicated by the black ellipses.
 Red and blue represent positive and negative numbers in the integral kernel, respectively.

### 2024-02-17 paper_sisc_first_revision/localpsf.tex
trying to figure out notation for Kronecker points of finite elements

@@
 In this paper, we also form an H-matrix approximation.
-However, to reduce the required number of operator applications, we first form a PSF approximation of the data misfit Hessian by exploiting locality and non-negative integral kernel properties, then form the H-matrix using classical H-matrix techniques.
-By using this two stage approach, we reduce the number of operator applications to a few dozen at most.
-
-One limitation of our method is that not all data misfit Hessians satisfy the local non-negative integral kernel properties.
+However, to reduce the required number of operator applications, we first form a PSF approximation of the data misfit Hessian by exploiting locality and non-negative integral kernel properties, then form the H-matrix using classical techniques.
+Using this two stage approach, we reduce the number of operator applications to a few dozen at most.
+
+Not all data misfit Hessians satisfy the local non-negative integral kernel properties.
 We note, in particular, that the wave inverse problem data misfit Hessian and Gauss-Newton Hessian have a substantial proportion of negative entries in their integral kernels.
-In this case more specialized techniques based on pseudodifferential operator theory have been developed, e.g., \cite{DemanetEtAl12,HerrmannMoghaddamStolk08,Symes08}.
+In this case more specialized techniques have been developed using, eg., pseudodifferential operator theory \cite{DemanetEtAl12,Symes08}, and sparsity in the wavelet domain \cite{HerrmannMoghaddamStolk08}.
 However, many data misfit Hessians of practical interest do satisfy the local non-negative integral kernel properties (either exactly or approximately), and the proposed method is targeted at approximating these Hessians.
@@
 
-The operator $\Aop$ may be applied to distributions if $\Aker$ is sufficiently regular. 
+The operator $\Aop$ may be applied to distributions\footnote{I.e., generalized functions such as the Dirac delta distribution.
+See, for example, \cite[Chapter 5]{ArbogastBona08}.} if $\Aker$ is sufficiently regular.
 Given $\rho \in L^2(\Omega)'$, let $\rho^* \in L^2(\Omega)$ denote the Riesz representative of $\rho$ with respect to the $L^2(\Omega)$ inner product. We have
@@
 \end{equation}
-that is formed by applying $\Aop$ to $\delta_x$, then taking the Riesz representation of the resulting linear functional.
+that is formed by applying $\Aop$ to $\delta_x$ (per the generalized notion of operator ``application'' defined above), then taking the Riesz representation of the resulting linear functional.
 Using~\eqref{eq:action_on_distribution} and the definition of the delta distribution, we see that $\phi_x$ may also be written as the function $\impulseresponse_x(y) = \Aker(y, x)$.
@@
 Left two columns: spatially variant blur kernel given in Equation \ref{eq:frog_kernel}.
-Right two columns: Ricker wavelet-type kernel given by $\Aker(y,x) = (1 - a \theta)\exp\left(-\theta/2\right)$, where $\theta=(y-x)^T \Sigma^{-1} (y-x)$, and $\Sigma=\operatorname{diag}\left(0.0025, 0.01\right)$.
+Right two columns: Ricker wavelet-type kernel given by $\Aker(y,x) = (1 - a \gamma)\exp\left(-\gamma/2\right)$, where $\gamma=(y-x)^T \Sigma^{-1} (y-x)$, and $\Sigma=\operatorname{diag}\left(0.0025, 0.01\right)$.
 Ordered from top to bottom, the results are obtained with $a\in\{1.0, 20.0, 27.0\}$ for the left two columns, and $a\in \{0.0, 0.23, 0.249\}$ for the right two columns.
@@
 In computations, functions are discretized and replaced by finite-dimensional vectors, and operators mapping between infinite-dimensional spaces are replaced by operators mapping between finite-dimensional spaces.
-In this paper we discretize using continuous Galerkin finite elements satisfying the Kronecker property (defined below).
+In this paper we discretize the functions that $\mathcal{A}$ and $\mathcal{A}^T$ are applied to using continuous finite elements satisfying the Kronecker property (defined below).
 With minor modifications, the proposed method could be used with more general finite element methods, or other discretization schemes such as finite differences or finite volumes.
-
-Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of continuous Galerkin finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space under the $L^2$ inner product, and let $\fepoint_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
+These restrictions on discretization only apply to functions $u$ that $\mathcal{A}$ and $\mathcal{A}^T$ are applied to.
+Other functions that arise internally during the process of computing actions of $\mathcal{A}$, such as state variables in a PDE that is solved in a subproblem, may be discretized with any method.
+
+Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of continuous finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space under the $L^2$ inner product, and let $\fepoint_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$.
 We assume that the finite element basis satisfies the Kronecker property, i.e., $\febasis_i(\fepoint_i)=1$ and $\febasis_i(\fepoint_j)=0$ if $i\neq j$.

### 2024-02-17 paper_sisc_first_revision/localpsf.tex
more edits based on Nicole's comments

@@
 \begin{align}
-\spatialmean^i(x) :=& \int_\Omega (\impulseresponse_x(y) / V(x)) y^i ~\mathrm{d}y \label{eq:define_mean} \\
-\spatialcov^{ij}(x) :=& \int_\Omega (\impulseresponse_x(y) / V(x)) \left(y^i - \spatialmean^i(x)\right) \left(y^j - \spatialmean^j(x)\right) ~\mathrm{d}y, \label{eq:define_cov}
+\spatialmean^i(x) :=& \frac{1}{V(x)}\int_\Omega \impulseresponse_x(y) y^i ~\mathrm{d}y \label{eq:define_mean} \\
+\spatialcov^{ij}(x) :=& \frac{1}{V(x)}\int_\Omega \impulseresponse_x(y) \left(y^i - \spatialmean^i(x)\right) \left(y^j - \spatialmean^j(x)\right) ~\mathrm{d}y, \label{eq:define_cov}
 \end{align}
@@
 %
-where $\spatialmean^i(x)$ denotes the $i^\text{th}$ component of the vector $\spatialmean(x)$, and $\spatialcov^{ij}(x)$ denotes the $(i,j)$ entry of the matrix $\spatialcov(x)$.
+where $\spatialmean^i(x)$ and $y^i$ denote the $i^\text{th}$ components of the vectors $\spatialmean(x)$ and $y$, respectively, and $\spatialcov^{ij}(x)$ denotes the $(i,j)$ entry of the matrix $\spatialcov(x)$.
 The vector $\spatialmean(x)\in \mathbb{R}^\gdim$ and the matrix $\spatialcov(x) \in \mathbb{R}^{\gdim \times \gdim}$ are the mean and covariance of the normalized version of $\impulseresponse_x$, respectively.
@@
 Then one can post process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$.
-However, this direct approach is not feasible because we need to know $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ before we compute $\phi_x$ in order choose the point $x$.
+However, this direct approach is not feasible because our algorithm for picking sample points (Section \ref{sec:sample_point_selection} and Figure \ref{fig:frog_batches}) needs to know $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ before we compute $\phi_x$.
 Computing $\phi_x$ in order to determine $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ would be extremely computationally expensive, and defeat the purpose of the proposed method, which is to reduce the computational cost by computing impulse responses in batches.

### 2024-02-17 paper_sisc_first_revision/localpsf.tex
addressed most of Nicole's comments

@@
 %
-where $\tau$ is a fixed constant.
+where $\tau$ is a fixed constant (see Figure \ref{fig:frog_moments_ellipsoid}).
 The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as the normalized version of $\impulseresponse_x$.
@@
 
+Here we provide an overview of the operator approximation algorithm.
+Then, in the subsections of this section (Sections \ref{sec:sample_point_selection}--\ref{sec:flip_eigs}) we present the technical details of the algorithm that were not already covered in Section \ref{sec:prerequisites}.
+
+%\subsection{Algorithm overview}
+%\label{sec:algorithm_overview}
+
 We use~\eqref{eq:vol_mean_var} to compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ by applying $\Aop^T$ to polynomial functions.
-Then we use~\eqref{eq:support_ellipsoid} to form ellipsoid shaped estimates for the support of the $\phi_x$'s, \emph{without} computing them.
-This allows us to compute large numbers of $\impulseresponse_{x_i}$ in ``batches,'' $\eta_b$ (see Figure~\ref{fig:batches_intro}).
+Then we use~\eqref{eq:support_ellipsoid} to form ellipsoid shaped estimates for the support of the $\phi_x$'s, \emph{without} computing them (see Figure \ref{fig:frog_moments_ellipsoid}).
+This allows us to compute large numbers of $\impulseresponse_{x_i}$ in ``batches,'' $\eta_b$ (see Figures~\ref{fig:batches_intro}~and~\ref{fig:frog_batches}).
 We compute one batch, denoted $\eta_b$, by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section~\ref{sec:get_impulse_response}).
@@
 
-Once the batches $\eta_b$ are computed, we approximate the integral kernel $\Aker(y,x)$ at arbitrary points $(y,x)$ by interpolation of translated and scaled versions of the computed $\impulseresponse_{x_i}$ (Section~\ref{sec:approximate_kernel_entries}).
+Once the batches $\eta_b$ are computed, we approximate the integral kernel $\Aker(y,x)$ at arbitrary points $(y,x)$ by interpolation of translated and scaled versions of the computed $\impulseresponse_{x_i}$ (Section~\ref{sec:approximate_kernel_entries} and Figure \ref{fig:hmatrix_neighbors_frog}).
 The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section~\ref{sec:local_mean_displacement_invariance}.
@@
 If $\candidatepoint_i$ passes these checks 
-(i.e., if $\candidatepoint_i$ is sufficiently far from other points in the batch, and $\spatialvol(\candidatepoint_i)$ and $\spatialcov(\candidatepoint_i)$ are acceptable)
+%(i.e., if $\candidatepoint_i$ is sufficiently far from other points in the batch, and $\spatialvol(\candidatepoint_i)$ and $\spatialcov(\candidatepoint_i)$ are acceptable)
 then we add $\candidatepoint_i$ to $\pointbatch_k$. Otherwise we discard $\candidatepoint_i$. This process repeats until there are no more points in $\candidatepts$.  
@@
 We flip the signs of these eigenvalues to be positive instead of negative (i.e., $\lambda \rightarrow |\lambda|$) by performing a low rank update to $\mathbf{A}_H^{\text{sym}}$.
+We observe that the eigenvectors associated with large erroneous negative eigenvalues tend to be directions that are nevertheless ``important'' to $\mathcal{A}$, so flipping the eigenvalues instead of setting them to zero tends to yield better approximations.
 The primary computational task in the Cayley shift-and-invert scheme is the solution of shifted linear systems of the form
@@
 
-\paragraph{(1) Computing impulse response moments and batches} Computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ requires applying $\Aop$ to $1$, $\gdim$, and $\gdim(\gdim+1)/2$ vectors, respectively.
+\paragraph{(1) Computing impulse response moments and batches} Computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ requires applying $\Aop^T$ to $1$, $\gdim$, and $\gdim(\gdim+1)/2$ vectors, respectively.
 Computing each $\eta_i$ requires applying $\Aop$ to one vector, so computing $\{\eta_i\}_{i=1}^{\nbatch}$ requires $\nbatch$ operator applications.

### 2024-02-18 paper_sisc_first_revision/localpsf.tex
add gaussian rbf shape parameter discussion to paper

@@
 where $\rbfweight_i$ are weights, and 
-$\rbf(r):= \exp\left(-C_\text{RBF}^2\frac{r^2}{2 L^2}\right)$
+$\rbf(r):= \exp\left(-\frac{1}{2}\left(\epsilon\frac{r}{L}\right)^2\right)$
 is a Gaussian kernel radial basis function.
-Here $L:=\diam\left(\{x_i\}_{i=1}^{\numnbr}\right)$ is the diameter of the set of sample points used in the interpolation, and $C_\text{RBF}$ is a user-defined shape parameter that controls the width of the kernel function (we use $C_\text{RBF}=3.0$).
+Here $L:=\diam\left(\{x_i\}_{i=1}^{\numnbr}\right)$ is the diameter of the set of sample points used in the interpolation, and $\epsilon$ is a user-defined shape parameter that controls the width of the kernel function.
 The vector of weights, $\rbfweight = (\rbfweight_1, \rbfweight_2, \dots, \rbfweight_{\numnbr})^T$, is found as the solution to the $\numnbr \times \numnbr$ linear system
@@
 Fortunately, the mesh cell containing $z_i$ was determined as a side effect of querying the AABB tree of mesh cells when we checked whether $z_i \in \Omega$.
+
+%\begin{figure}
+%	\begin{subfigure}[b]{0.24\textwidth}
+%		\centering
+%		\includegraphics[scale=0.25]{rbf_cosine_shape_parameter=1.0.pdf}
+%	\end{subfigure}
+%	\begin{subfigure}[b]{0.24\textwidth}
+%		\centering
+%		\includegraphics[scale=0.25]{rbf_cosine_shape_parameter=10.0.pdf}
+%	\end{subfigure}
+%	\begin{subfigure}[b]{0.24\textwidth}
+%		\centering
+%		\includegraphics[scale=0.25]{rbf_cosine_shape_parameter=100.0.pdf}
+%	\end{subfigure}
+%	\begin{subfigure}[b]{0.24\textwidth}
+%		\centering
+%		\includegraphics[scale=0.25]{cosine_RBF_convergence.pdf}
+%	\end{subfigure}
+%	
+%	\caption{asdf
+%	}
+%	\label{fig:RBF_width}
+%\end{figure}
+
+The shape parameter $\epsilon$ mediates a tradeoff between accuracy and stability.
+Small $\epsilon$ is required for RBF interpolation with Gaussian kernels to achieve high accuracy, but small $\epsilon$ also makes RBF interpolation less robust to errors or nonsmoothness in the function being interpolated.
+%This is illustrated in Figure \ref{fig:RBF_width} for a simple one dimensional interpolation problem. 
+If high accuracy is not required (for example, if the method is being used to construct a preconditioner) we recommend a conservative choice of $\epsilon=3.0$ in order to make the method more robust to nonsmoothness of the kernel, discretization error, violations of our assumptions, and issues related to boundaries.
+We use $\epsilon=3.0$ in our numerical results involving Hessians for this reason.
+In our numerical results for blur kernels, we use a smaller value of $\epsilon=0.5$ so that the RBF interpolation accuracy is not a limiting factor as we study convergence of the method.
 

### 2024-02-18 paper_sisc_first_revision/localpsf.tex
Allow the user to vary the RBF shape parameter, rather than having it as a hardcoded constant

@@
 
-The shape parameter $\epsilon$ mediates a tradeoff between accuracy and stability.
+The shape parameter, $\epsilon$, mediates a tradeoff between accuracy and stability.
 Small $\epsilon$ is required for RBF interpolation with Gaussian kernels to achieve high accuracy, but small $\epsilon$ also makes RBF interpolation less robust to errors or nonsmoothness in the function being interpolated.
 %This is illustrated in Figure \ref{fig:RBF_width} for a simple one dimensional interpolation problem. 
-If high accuracy is not required (for example, if the method is being used to construct a preconditioner) we recommend a conservative choice of $\epsilon=3.0$ in order to make the method more robust to nonsmoothness of the kernel, discretization error, violations of our assumptions, and issues related to boundaries.
+If high accuracy is not required (for example, if the method is being used to construct a preconditioner) we recommend a conservative choice of $\epsilon=3.0$ in order to make the method more robust to nonsmoothness of the kernel, discretization error, errors caused by violations of our assumptions, and issues related to boundaries.
 We use $\epsilon=3.0$ in our numerical results involving Hessians for this reason.

### 2024-02-18 localpsf/localpsf_cg1_lumped.py

@@ -381,6 +381,13 @@ class PSFKernel:
     def tau(me):
         return me.IRB.tau
 
+    @property
+    def shape_parameter(me):
+        return me.cpp_object.shape_parameter
+
+    def update_shape_parameter(me, new_shape_parameter: float):
+        me.cpp_object.shape_parameter = new_shape_parameter
+
     def update_tau(me, new_tau: float):
         assert_gt(new_tau, 0.0)
         me.IRB.update_tau(new_tau)
@@ -398,11 +405,12 @@ class PSFKernel:
         return (me.V_out.ndof, me.V_in.ndof)
 
 
-def make_product_convolution_kernel(col_batches: ImpulseResponseBatches) -> PSFKernel:
+def make_product_convolution_kernel(col_batches: ImpulseResponseBatches, shape_parameter: float=3.0) -> PSFKernel:
     cpp_object = hpro.hpro_cpp.ProductConvolutionKernelRBFColsOnly(
         col_batches.cpp_object,
         list(col_batches.V_in.vertices),
         list(col_batches.V_out.vertices))
+    cpp_object.shape_parameter = shape_parameter
     return PSFKernel(col_batches, cpp_object)
 
 
@@ -488,6 +496,7 @@ def make_psf(
         tau: float = 3.0,
         num_neighbors: int = 10,
         max_candidate_points: int = None,
+        shape_parameter: float = 3.0,
         display: bool=False,
 ) -> PSFObject:
     assert_ge(num_initial_batches, 0)
@@ -552,7 +561,7 @@ def make_psf(
         num_initial_batches=num_initial_batches, tau=tau,
         num_neighbors=num_neighbors, max_candidate_points=max_candidate_points)
 
-    psf_kernel: PSFKernel = make_product_convolution_kernel(col_batches)
+    psf_kernel: PSFKernel = make_product_convolution_kernel(col_batches, shape_parameter=shape_parameter)
 
     psf_object = PSFObject(
         psf_kernel, IM, bad_vols, tiny_Sigmas, huge_Sigmas, bad_aspect_Sigmas,
@@ -729,6 +738,7 @@ def make_psf_fenics(
         tau: float = 3.0,
         num_neighbors: int = 10,
         max_candidate_points: int = None,
+        shape_parameter: float = 3.0,
         display: bool=False,
 ) -> PSFObjectFenicsWrapper:
     assert_equal(mass_lumps_in.shape, (V_in.dim(),))
@@ -743,6 +753,6 @@ def make_psf_fenics(
[…]

### 2024-02-18 numerical_examples/blur/blur_tau_experiments.py

@@ -11,23 +11,22 @@ import hlibpro_python_wrapper as hpro
 from tqdm.auto import tqdm
 
 
-max_batches = 45
 nx = 63
 a = 1.0 # <-- How bumpy? Use 1.0, which is maximum bumpiness without negative numbers
 length_scaling = 1.0 / (2.0**2)
 all_tau = [2.0, 2.5, 3.0, 3.5, 4.0]
-
-np.savetxt('all_tau.txt', all_tau)
+all_num_batches = [1, 2, 4, 8, 16, 32, 64, 128, 256, 512]
+shape_parameter = 0.5
+num_neighbors = 10
 
 Ker, phi_function, Vh, H, mass_lumps, dof_coords, kdtree_sort_inds = frog_setup(nx, length_scaling, a)
 
 Ker_reordered = Ker[:, kdtree_sort_inds][kdtree_sort_inds,:]
 dof_coords_reordered = dof_coords[kdtree_sort_inds,:]
 
-all_all_num_batches = []
-all_all_num_impulses = []
-all_all_fro_errors = []
-for tau in all_tau:
+all_num_impulses    = np.zeros((len(all_tau), len(all_num_batches)))
+all_fro_errors      = np.zeros((len(all_tau), len(all_num_batches)))
+for ii, tau in enumerate(all_tau):
     print('tau=', tau)
 
     psf_object = lpsf1.make_psf_fenics(
@@ -37,14 +36,14 @@ for tau in all_tau:
         mass_lumps, mass_lumps,
         num_initial_batches=0,
         tau=tau, display=True,
-        num_neighbors=10
+        num_neighbors=num_neighbors,
+        shape_parameter=shape_parameter,
     )
 
-    all_num_batches = []
-    all_num_impulses = []
-    all_fro_errors = []
-    for num_batches in range(max_batches):
-        psf_object.add_impulse_response_batch()
+    for jj, num_batches in enumerate(all_num_batches):
+        while psf_object.psf_object.impulse_response_batches.num_batches < num_batches:
+            psf_object.add_impulse_response_batch()
+
         num_batches = psf_object.psf_object.impulse_response_batches.num_batches
         num_impulses = psf_object.psf_object.impulse_response_batches.num_sample_points
 
@@ -54,18 +53,16 @@ for tau in all_tau:
         print('num_batches=', num_batches)
         print('num_impulses=', num_impulses)
 
[…]

### 2024-02-18 paper_sisc_first_revision/frog_tau_batches.tex
created tikz plots for frog convergence for different tau

\documentclass{standalone}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{amsmath}

% SIAM Shared Information Template
% This is information that is shared between the main document and any
% supplement. If no supplement is required, then this information can
% be included directly in the main document.

% Packages and macros go here
\usepackage{lipsum}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{pgfplots}
\usepackage{bm}

\begin{document}
	\begin{tikzpicture}[baseline]%, trim axis right]
		\begin{axis}[width=0.49\linewidth, height=0.5\linewidth,
		xmode=log,
		ymode=log,
		ymin=1e-4,
		ymax=1e0,
%		xmax=1000,
		axis x line*=bottom,
		axis y line*=left,
%		xtick={0, 200, 400, 600,800,1000},
		xlabel={\#batches},
%		ylabel={$\|\boldsymbol{\Phi} - \widetilde{\boldsymbol{\Phi}}\|_\text{Fro} / \|\boldsymbol{\Phi}\|_\text{Fro}$},
		title={PSF convergence (batches)},
%		legend style={at={(1.2,1)},anchor=north east,draw=none},
%		legend pos=south west,
%		legend style={font=\small}
		]
%			\addlegendentry{$\tau=2.0$}
			\addplot[black, line width=0.5mm] table {frog_convergence_batches_tau2.0.dat};
%			\addlegendentry{$\tau=2.5$}
			\addplot[blue, line width=0.5mm] table {frog_convergence_batches_tau2.5.dat};
%			\addlegendentry{$\tau=3.0$}
			\addplot[cyan, line width=0.5mm] table {frog_convergence_batches_tau3.0.dat};
%			\addlegendentry{$\tau=3.5$}
			\addplot[red, line width=0.5mm] table {frog_convergence_batches_tau3.5.dat};
%			\addlegendentry{$\tau=4.0$}
			\addplot[purple, line width=0.5mm] table {frog_convergence_batches_tau4.0.dat};
		\end{axis}
	\end{tikzpicture}
\end{document}

### 2024-02-18 paper_sisc_first_revision/frog_tau_impulses.tex
created tikz plots for frog convergence for different tau

--- 2024-02-18 paper_sisc_first_revision/frog_tau_batches.tex
@@
 %		xtick={0, 200, 400, 600,800,1000},
-		xlabel={\#batches},
-%		ylabel={$\|\boldsymbol{\Phi} - \widetilde{\boldsymbol{\Phi}}\|_\text{Fro} / \|\boldsymbol{\Phi}\|_\text{Fro}$},
-		title={PSF convergence (batches)},
-%		legend style={at={(1.2,1)},anchor=north east,draw=none},
-%		legend pos=south west,
-%		legend style={font=\small}
+		xlabel={\#impulse responses},
+		ylabel={$\|\boldsymbol{\Phi} - \widetilde{\boldsymbol{\Phi}}\|_\text{Fro} / \|\boldsymbol{\Phi}\|_\text{Fro}$},
+		title={PSF convergence (impulse responses)},
+		legend style={at={(1.2,1)},anchor=north east,draw=none},
+		legend pos=south west,
+		legend style={font=\small}
 		]
-%			\addlegendentry{$\tau=2.0$}
-			\addplot[black, line width=0.5mm] table {frog_convergence_batches_tau2.0.dat};
-%			\addlegendentry{$\tau=2.5$}
-			\addplot[blue, line width=0.5mm] table {frog_convergence_batches_tau2.5.dat};
-%			\addlegendentry{$\tau=3.0$}
-			\addplot[cyan, line width=0.5mm] table {frog_convergence_batches_tau3.0.dat};
-%			\addlegendentry{$\tau=3.5$}
-			\addplot[red, line width=0.5mm] table {frog_convergence_batches_tau3.5.dat};
-%			\addlegendentry{$\tau=4.0$}
-			\addplot[purple, line width=0.5mm] table {frog_convergence_batches_tau4.0.dat};
+			\addlegendentry{$\tau=2.0$}
+			\addplot[black, line width=0.5mm] table {frog_convergence_impulses_tau2.0.dat};
+			\addlegendentry{$\tau=2.5$}
+			\addplot[blue, line width=0.5mm] table {frog_convergence_impulses_tau2.5.dat};
+			\addlegendentry{$\tau=3.0$}
+			\addplot[cyan, line width=0.5mm] table {frog_convergence_impulses_tau3.0.dat};
+			\addlegendentry{$\tau=3.5$}
+			\addplot[red, line width=0.5mm] table {frog_convergence_impulses_tau3.5.dat};
+			\addlegendentry{$\tau=4.0$}
+			\addplot[purple, line width=0.5mm] table {frog_convergence_impulses_tau4.0.dat};
 		\end{axis}

### 2024-02-18 numerical_examples/blur/blur_tau_experiments.py

@@ -89,3 +89,8 @@ plt.xlabel('#impulse responses')
 plt.ylabel(r'$||\Phi - \widetilde{\Phi}||/||\Phi||$')
 plt.title(r'Convergence for different $\tau$')
 plt.savefig('frog_tau_comparison_by_impulse.pdf', bbox_inches='tight', dpi=300)
+
+for ii in range(len(all_tau)):
+    np.savetxt('frog_convergence_batches_tau'+str(all_tau[ii])+'.dat', np.vstack([all_num_batches[ii,:], all_fro_errors[ii,:]]).T)
+    np.savetxt('frog_convergence_impulses_tau'+str(all_tau[ii])+'.dat', np.vstack([all_num_impulses[ii, :], all_fro_errors[ii, :]]).T)
+

### 2024-02-18 paper_sisc_first_revision/frog_tau_batches.tex
created tikz plots for frog convergence for different tau

@@
 %			\addlegendentry{$\tau=3.5$}
-			\addplot[red, line width=0.5mm] table {frog_convergence_batches_tau3.5.dat};
+			\addplot[purple, line width=0.5mm] table {frog_convergence_batches_tau3.5.dat};
 %			\addlegendentry{$\tau=4.0$}
-			\addplot[purple, line width=0.5mm] table {frog_convergence_batches_tau4.0.dat};
+			\addplot[red, line width=0.5mm] table {frog_convergence_batches_tau4.0.dat};
 		\end{axis}

### 2024-02-18 paper_sisc_first_revision/frog_tau_impulses.tex
created tikz plots for frog convergence for different tau

@@
 			\addlegendentry{$\tau=3.5$}
-			\addplot[red, line width=0.5mm] table {frog_convergence_impulses_tau3.5.dat};
+			\addplot[purple, line width=0.5mm] table {frog_convergence_impulses_tau3.5.dat};
 			\addlegendentry{$\tau=4.0$}
-			\addplot[purple, line width=0.5mm] table {frog_convergence_impulses_tau4.0.dat};
+			\addplot[red, line width=0.5mm] table {frog_convergence_impulses_tau4.0.dat};
 		\end{axis}

### 2024-02-18 paper_sisc_first_revision/localpsf.tex
updated section 7.1

@@
 \section{Preliminaries}
+\label{sec:prelims}
 
@@
 
-\subsection{Example 1: Spatially variant blur}
+\subsection{Example 1: Spatially varying blurring problem}
 \label{sec:frog}
 
-Before delving into technical details, we provide an overview of the method.
-Along the way, we illustrate steps using a spatially variant blur example, in which the integral kernel is given as follows:
-\begin{equation}
-\label{eq:frog_kernel}
-\Phi(y,x) = (1 - a f(y,x)) g(x) \exp \left(-\frac{1}{2}h^T(y,x) C^{-1} h(y,x)\right).
-\end{equation} 
-Here, $h(y,x) = R(\theta(x)) (y-x)$, $R(\theta(x))$ is the 2D rotation matrix by angle $\theta(x) = (x^1 + x^2) \pi / 2$, $C = L^2 \operatorname{diag}(c_1, c_2)$, $g(x) = x^1(1-x^1)x^2(1-x^2)$,  and $f(x,y) = \cos\left(p^1(y,x) / \sqrt{c_1 / 2.0}\right) \sin\left(p^2(y,x) / \sqrt{c_2 / 2.0}\right)$ for some constants $a, L, c_1, c_2$.
-In this section we use $a=1.0$, $L^2=1.0$, $c_1=0.01$, and $c_2=0.0025$.
-Using this kernel, the impulse response $\phi_x$ is a bumpy blob that is centered near $x$, and is rotated and scaled in a manner that depends on $x$ (see Figure \ref{fig:frog1}).
-The constant $L$ controls how wide the blob is, and $a$ controls the bumpiness.
-With $a=0$, $\phi_x$ is a Gaussian.
-As $a$ increases, $\phi_x$ becomes more bumpy.
-When $a > 1$, the integral kernel contains negative values, which violates the assumptions of our method.
-Note that for this operator, using our method is unnecessary because entries of the kernel are given directly via a formula.
-The purpose of this example is to illustrate the method using an operator which is known exactly and may be easily visualized.
+Here we define the PDE-free spacially varying blurring problem, used
+in Sections~\ref{sec:intro} and~\ref{sec:prelims} for visual
+illustration of various stages and aspects of the proposed method.  To
+study the influence of the presence of negative numbers in the
+integral kernel, i.e., the robustness of the proposed method to
+violations of the non-negative kernel assumption discussed in
+Section~\ref{sec:impulse_response_2}, this example is set up so that
+the impulse response, $\phi_x$, corresponding to the integral kernel
+is a bumpy blob that is centered near $x$, and is rotated and scaled
+in a manner that depends on $x$ (see Figure
+\ref{fig:frog_kernel_impulse_responses}). The closed form expression
+of the integral kernel is given by
+%
+%Before delving into technical details, we provide an overview of the
+%method. Along the way, we illustrate steps using a spatially variant
+%blur example, in which the integral kernel is given as follows:
+%\begin{equation}
+%\label{eq:frog_kernel}
+%\Phi(y,x) = (1 - a f(y,x)) x^1(1-x^1)x^2(1-x^2) \exp \left(-\frac{1}{2}h^T(y,x) C^{-1} h(y,x)\right).
+%\end{equation}
+%Here, $h(y,x) = R(\theta(x)) (y-x)$, $R(\theta(x))$ is the 2D rotation
+%matrix by angle $\theta(x) = (x^1 + x^2) \pi / 2$, $C = L^2
+%\operatorname{diag}(c_1, c_2)$, $g(x) = x^1(1-x^1)x^2(1-x^2)$, and
+%$f(x,y) = \cos\left(p^1(y,x) / \sqrt{c_1 / 2.0}\right)
+%\sin\left(p^2(y,x) / \sqrt{c_2 / 2.0}\right)$ for some constants $a,
+%L, c_1, c_2$. In this section we use $a=1.0$, $L^2=1.0$, $c_1=0.01$,
+%and $c_2=0.0025$.
+%
+\begin{align}
+\label{eq:example_frog_kernel}
+\Phi(y,x) = (1 - a f(y,x)) g(x) \exp \left(-\frac{1}{2}(h(y,x)^T
+C^{-1} h(y,x) \right),
+\end{align}
+where $f(y,x) = \cos\left(h^1(y,x) / \sqrt{c_1 /
+  2}\right)\sin\left(h^2(y,x) / \sqrt{c_2 / 2}\right)$, with
+$h^i(y,x)$ the $i^{th}$ component of $R(\theta(x)) (y-x)$, with
+$R(\theta(x))$ a two-dimensional rotation matrix by angle $\theta(x) =
+(x^1 + x^2) \pi / 2$, $g(x) = x^1(1-x^1)x^2(1-x^2)$, and $C = L^2
+\operatorname{diag}(c_1, c_2)$. The constant $L$ controls how wide the
+blob is (hence the support of the impulse response), $a$ controls the
+magnitude of the bumps, and $(c_1, c_2)$ control the width of the
+Gaussian. We note that when $a=0$, $\phi_x$ is a Gaussian, as $a$
+increases, $\phi_x$ becomes more bumpy. For instance, when $a > 1$,
+the integral kernel contains negative values, which violates the
+Assumption 3 (see Section~\ref{sec:conditions_2}).
+
+%The impulse response $\phi_x$ corresponding to this kernel is a bumpy
+%blob that is centered near $x$, and is rotated and scaled in a manner
+%that depends on $x$ (see Figure
+%\ref{fig:frog_kernel_impulse_responses}). The constant $L$ controls
+%how wide the blob is (hence the support of the impulse response), and
+%$a$ controls the magnitude of the bumps.
+
+%Note that for this
+%operator, using our method is unnecessary because entries of the
+%kernel are given directly via a formula. The purpose of this example
+%is to illustrate the method using an operator which is known exactly
+%and may be easily visualized.
 

### 2024-02-19 paper_sisc_first_revision/localpsf.tex
done with section 7.1

@@
 
-Here we define the PDE-free spacially varying blurring problem, used
-in Sections~\ref{sec:intro} and~\ref{sec:prelims} for visual
-illustration of various stages and aspects of the proposed method.  To
-study the influence of the presence of negative numbers in the
-integral kernel, i.e., the robustness of the proposed method to
-violations of the non-negative kernel assumption discussed in
-Section~\ref{sec:impulse_response_2}, this example is set up so that
-the impulse response, $\phi_x$, corresponding to the integral kernel
-is a bumpy blob that is centered near $x$, and is rotated and scaled
-in a manner that depends on $x$ (see Figure
-\ref{fig:frog_kernel_impulse_responses}). The closed form expression
-of the integral kernel is given by
+Here we define a PDE-free spatially varying blurring problem, used in
+Sections~\ref{sec:intro} and~\ref{sec:prelims} for visual illustration
+of various stages and aspects of the proposed method.  To study the
+robustness of the proposed method to violations of the non-negative
+kernel assumption discussed in Section~\ref{sec:impulse_response_2},
+this example is set up so that the impulse response, $\phi_x$,
+corresponding to the integral kernel is a bumpy blob that is centered
+near $x$, and is rotated and scaled in a manner that depends on $x$
+(see Figure \ref{fig:frog_kernel_impulse_responses}). The closed form
+expression of the integral kernel is given by
 %
@@
 \begin{align}
-\label{eq:example_frog_kernel}
+\label{eq:frog_kernel}
 \Phi(y,x) = (1 - a f(y,x)) g(x) \exp \left(-\frac{1}{2}(h(y,x)^T
@@
 (x^1 + x^2) \pi / 2$, $g(x) = x^1(1-x^1)x^2(1-x^2)$, and $C = L^2
-\operatorname{diag}(c_1, c_2)$. The constant $L$ controls how wide the
-blob is (hence the support of the impulse response), $a$ controls the
-magnitude of the bumps, and $(c_1, c_2)$ control the width of the
-Gaussian. We note that when $a=0$, $\phi_x$ is a Gaussian, as $a$
-increases, $\phi_x$ becomes more bumpy. For instance, when $a > 1$,
-the integral kernel contains negative values, which violates the
-Assumption 3 (see Section~\ref{sec:conditions_2}).
+\operatorname{diag}(c_1, c_2)$. The constant $L$ controls the width of
+the blob, and $c_1/c_2$ controls its aspect ratio. The constant $a$
+represents deviation from a Gaussian. When $a=0$, $\phi_x$ is a
+Gaussian, and as $a$ increases, $\phi_x$ becomes non-Gaussian. When $a
+> 1$, the integral kernel contains negative values, which allows us to
+study the robustness of the PSF method to violations of Assumption 3
+(Section~\ref{sec:conditions_2}).
 
@@
 
-Define HODLR. Define randomized HODLR algo. Ref to Tucker's paper and "Fast randomized eigensolver with structured LDL..."
-
-Define randomized SVD, refer to Halko Martinsson Tropp.
-
-Discuss why PSF does better with smaller $L$.
+In Table~\ref{tab:frog_psf_vs_hodlr_vs_glr} we compare the
+computational cost to approximate the blur kernel from
+Equation~\eqref{eq:frog_kernel} using the proposed PSF method, the
+randomized HODLR (hierarchical off diagonal low rank)
+method~\cite{Martinsson16,HartlandEtAl23}, and GLR (global low rank)
+approximation using randomized SVD~\cite{HalkoMartinssonTropp11}. For
+these results we vary the quantity $L$ to scale the width of the
+impulse responses and hence the rank of the operator. For each case,
+we calculate the relative error in the approximation of the kernel
+measured in the Frobenius norm, $||\boldsymbol{\Phi} -
+\widetilde{\boldsymbol{\Phi}}||_\text{Fro} /
+||\boldsymbol{\Phi}||_\text{Fro}$ and show the number of operator
+applications required to achieve 20\%, 10\%, and 5\% relative error by
+each method. The results reveal that superior performance of the PSF
+method as compared to the HODLR and GLR methods for all cases. We note
+that as we increase the rank and decrease the error tolerance, the
+performance of the the HODLR and GLR methods deteriorates.
+
+%Define HODLR. Define randomized HODLR algo. Ref to Tucker's paper and "Fast randomized eigensolver with structured LDL..."
+
+%Define randomized SVD, refer to Halko Martinsson Tropp.
+
+%Discuss why PSF does better with smaller $L$.
 
@@
 	\end{center}
-	\caption{(Spatially variant blur) Comparison of cost to approximate the blur kernel from Equation \eqref{eq:frog_kernel} using the PSF method, the randomized HODLR (hierarchical off diagonal low rank) method, and GLR (global low rank) approximation using randomized SVD.
-The quantity $L$ scales the width of the impulse responses, hence it influences the rank of the operator.
-Large $L$ means low rank, and small $L$ means high rank.
-The second column (``Error tol'') is the relative error in the approximation of the kernel measured in the Frobenius norm, $||\Phi - \widetilde{\Phi}||_\text{Fro} / ||\Phi||_\text{Fro}$.
-The remaining three columns show the number of operator applies required to achieve the given error tolerances, using the PSF, HODLR, and GLR methods.}
+	\caption{(Spatially variant blur) Comparison of cost to
+          approximate the blur kernel from Equation
+          \eqref{eq:frog_kernel} using the PSF method, the randomized
+          HODLR (hierarchical off diagonal low rank) method, and GLR
+          (global low rank) approximation using randomized SVD. The
+          quantity $L$ scales the width of the impulse responses,
+          hence it influences the rank of the operator. Large $L$
+          means low rank, and small $L$ means high rank. The second
+          column (``Error tol'') is the relative error in the
+          approximation of the kernel measured in the Frobenius norm,
+          $||\boldsymbol{\Phi} -
+          \widetilde{\boldsymbol{\Phi}}||_\text{Fro} /
+          ||\boldsymbol{\Phi}||_\text{Fro}$. The remaining three
+          columns show the number of operator applies required to
+          achieve the given error tolerances, using the PSF, HODLR,
+          and GLR methods.}
 	\label{tab:frog_psf_vs_hodlr_vs_glr}

### 2024-02-19 paper_sisc_first_revision/frog_tau_batches.tex
added frog convergence plots to paper and wrote caption

@@
 %			\addlegendentry{$\tau=2.0$}
-			\addplot[black, line width=0.5mm] table {frog_convergence_batches_tau2.0.dat};
+			\addplot[olive, line width=0.5mm] table {frog_convergence_batches_tau2.0.dat};
 %			\addlegendentry{$\tau=2.5$}
@@
 %			\addlegendentry{$\tau=3.5$}
-			\addplot[purple, line width=0.5mm] table {frog_convergence_batches_tau3.5.dat};
+			\addplot[black, line width=0.5mm] table {frog_convergence_batches_tau3.5.dat};
 %			\addlegendentry{$\tau=4.0$}
 			\addplot[red, line width=0.5mm] table {frog_convergence_batches_tau4.0.dat};
+			\addplot[dashed, color=gray, line width=0.5mm] table {frog_convergence_rate_batches.dat};
 		\end{axis}

### 2024-02-19 paper_sisc_first_revision/localpsf.tex
added frog convergence plots to paper and wrote caption

@@
 randomized HODLR (hierarchical off diagonal low rank)
-method~\cite{Martinsson16,HartlandEtAl23}, and GLR (global low rank)
+method~\cite{Martinsson16,Hartland_2023}, and GLR (global low rank)
 approximation using randomized SVD~\cite{HalkoMartinssonTropp11}. For
@@
 \end{table}
+
+\begin{figure}
+	\begin{subfigure}{0.52\textwidth}
+		\begin{center}
+			\includegraphics[scale=1.0]{frog_tau_impulses.pdf}
+		\end{center}
+	\end{subfigure}
+	\begin{subfigure}{0.47\textwidth}
+		\begin{center}
+			\includegraphics[scale=1.0]{frog_tau_batches.pdf}
+		\end{center}
+	\end{subfigure}
+	\caption{Relative error in the PSF approximation of the kernel, $||\boldsymbol{\Phi} - \widetilde{\boldsymbol{\Phi}}||_\text{Fro}/||\boldsymbol{\Phi}||_\text{Fro}$, for different ellipsoid size parameters, $\tau$.
+Left: Relative error vs.
+the total number of impulse responses used in the approximation.
+Right: Relative error vs.
+the number of impulse response batches.
+The dashed gray lines show linear convergence rates, i.e., $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$ on the left, and  $\text{constant} \times \left(\# \text{batches}\right)^{-1}$ on the right.}
+	\label{fig:PSF_tau_convergence}
+\end{figure}
 

### 2024-02-19 numerical_examples/blur/blur_tau_experiments.py

@@ -64,13 +64,15 @@ for ii, tau in enumerate(all_tau):
         np.savetxt('all_num_impulses.txt', all_num_impulses)
         np.savetxt('all_fro_errors.txt', all_fro_errors)
 
-
+fro_errs_batches_theory = 0.15 / np.array(all_num_batches[-1,:])
 
 plt.figure()
 leg = []
 for ii, tau in enumerate(all_tau):
     plt.loglog(all_num_batches[ii,:], all_fro_errors[ii,:])
     leg.append(r'$\tau$='+str(tau))
+plt.loglog(all_num_batches[-1,:], fro_errs_batches_theory, '--', c='gray')
+leg.append(r'$const \times \# batches^{-1}$')
 plt.legend(leg)
 plt.ylim(None, 1e0)
 plt.xlabel('#batches')
@@ -78,11 +80,16 @@ plt.ylabel(r'$||\Phi - \widetilde{\Phi}||/||\Phi||$')
 plt.title(r'Convergence for different $\tau$')
 plt.savefig('frog_tau_comparison_by_batch.pdf', bbox_inches='tight', dpi=300)
 
+
+fro_errs_impulses_theory = 10*0.15 / np.array(all_num_impulses[-1,:])
+
 plt.figure()
 leg = []
 for ii, tau in enumerate(all_tau):
     plt.loglog(all_num_impulses[ii,:], all_fro_errors[ii,:])
     leg.append(r'$\tau$='+str(tau))
+plt.loglog(all_num_impulses[-1,:], fro_errs_impulses_theory, '--', c='gray')
+leg.append(r'$const \times \left(\# impulse~responses\right)^{-1}$')
 plt.legend(leg)
 plt.ylim(None, 1e0)
 plt.xlabel('#impulse responses')
@@ -94,3 +101,5 @@ for ii in range(len(all_tau)):
     np.savetxt('frog_convergence_batches_tau'+str(all_tau[ii])+'.dat', np.vstack([all_num_batches[ii,:], all_fro_errors[ii,:]]).T)
     np.savetxt('frog_convergence_impulses_tau'+str(all_tau[ii])+'.dat', np.vstack([all_num_impulses[ii, :], all_fro_errors[ii, :]]).T)
 
+np.savetxt('frog_convergence_rate_batches.dat', np.vstack([all_num_batches[-1,:], fro_errs_batches_theory]).T)
[…]

### 2024-02-19 paper_sisc_first_revision/localpsf.tex
more editing

@@
 Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, and $L^2(\Omega)'$ is the space of real-valued continuous linear functionals on $L^2(\Omega)$.
-By ``non-negative integral kernel,'' we mean that entries of $\Aop$'s integral kernel (defined in Equation \ref{eq:kernel_representation}) are non-negative numbers; this is not the same as positive semi-definiteness of $\Aop$.
+By ``non-negative integral kernel,'' we mean that entries of $\Aop$'s integral kernel
+%(defined in Equation \ref{eq:kernel_representation})
+are non-negative numbers; this is not the same as positive semi-definiteness of $\Aop$.
 Such operators appear as, for example, Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{BorziSchulz11,Delosreyes15,HinzePinnauUlbrichUlbrich08},
@@
 
-Here ``matrix-free'' means that we may apply $\Aop$ and its transpose\footnote{Recall that $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for operations with $\Aop^T$.
+Here, ``matrix-free'' means that we may apply $\Aop$ and its transpose\footnote{Recall that $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for operations with $\Aop^T$.
 }, $\Aop^T$, to functions,
@@
 Here we provide an overview of the operator approximation algorithm.
-Then, in the subsections of this section (Sections \ref{sec:sample_point_selection}--\ref{sec:flip_eigs}) we present the technical details of the algorithm that were not already covered in Section \ref{sec:prerequisites}.
+Then, in the subsections (Sections \ref{sec:sample_point_selection}--\ref{sec:flip_eigs}) we present technical details of the algorithm that were not already covered in Section \ref{sec:prerequisites}.
 
@@
 %This is illustrated in Figure \ref{fig:RBF_width} for a simple one dimensional interpolation problem. 
-If high accuracy is not required (for example, if the method is being used to construct a preconditioner) we recommend a conservative choice of $\epsilon=3.0$ in order to make the method more robust to nonsmoothness of the kernel, discretization error, errors caused by violations of our assumptions, and issues related to boundaries.
-We use $\epsilon=3.0$ in our numerical results involving Hessians for this reason.
-In our numerical results for blur kernels, we use a smaller value of $\epsilon=0.5$ so that the RBF interpolation accuracy is not a limiting factor as we study convergence of the method.
+For our numerical results involving Hessians in inverse problems governed by PDEs (Sections~\ref{sec:ice}~and~\ref{sec:adv}), high accuracy is not required because the method is used to construct a preconditioner.
+Hence, for these Hessian approximations we use a conservative choice of $\epsilon=3.0$ to ensure robustness.
+For our numerical results for blur kernels (Section~\ref{sec:frog}), we use a smaller value of $\epsilon=0.5$ so that the RBF interpolation accuracy is not a limiting factor as we study convergence of the method.
 
@@
 
-Here we define a PDE-free spatially varying blurring problem, used in
-Sections~\ref{sec:intro} and~\ref{sec:prelims} for visual illustration
-of various stages and aspects of the proposed method.  To study the
-robustness of the proposed method to violations of the non-negative
-kernel assumption discussed in Section~\ref{sec:impulse_response_2},
-this example is set up so that the impulse response, $\phi_x$,
-corresponding to the integral kernel is a bumpy blob that is centered
+Here we define a PDE-free spatially varying blurring problem, in which the impulse response, $\phi_x$,
+is a bumpy blob that is centered
 near $x$, and is rotated and scaled in a manner that depends on $x$
-(see Figure \ref{fig:frog_kernel_impulse_responses}). The closed form
-expression of the integral kernel is given by
+(see Figure \ref{fig:frog_kernel_impulse_responses}). This blurring problem is used in
+Sections~\ref{sec:intro} and~\ref{sec:prelims} to visually illustrate
+various stages and aspects of the proposed method, and robustness (or lack thereof) of the proposed method to violations of the non-negative kernel assumption (Section~\ref{sec:conditions_2}).
+In this section, the blurring problem is used to compare the proposed method to hierarchical off-diagonal low rank (HODLR) and global low rank (GLR) methods, to study convergence of the method, and to investigate the effect of the ellipsoid size parameter $\tau$.
+%To study the
+%robustness of the proposed method to violations of the non-negative
+%kernel assumption discussed in Section~\ref{sec:impulse_response_2},
+The closed form
+expression for the integral kernel is given by
 %
@@
 $R(\theta(x))$ a two-dimensional rotation matrix by angle $\theta(x) =
-(x^1 + x^2) \pi / 2$, $g(x) = x^1(1-x^1)x^2(1-x^2)$, and $C = L^2
+(x^1 + x^2) \pi / 2$, $g(x) = x^1(1-x^1)x^2(1-x^2)$, and with $C = L^2
 \operatorname{diag}(c_1, c_2)$. The constant $L$ controls the width of
@@
 applications required to achieve 20\%, 10\%, and 5\% relative error by
-each method. The results reveal that superior performance of the PSF
+each method. The results reveal superior performance of the PSF
 method as compared to the HODLR and GLR methods for all cases. We note
@@
 the number of impulse response batches.
-The dashed gray lines show linear convergence rates, i.e., $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$ on the left, and  $\text{constant} \times \left(\# \text{batches}\right)^{-1}$ on the right.}
+The dashed gray lines show linear convergence rates, i.e., $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$ on the left, and  $\text{constant} \times \# \text{batches}^{-1}$ on the right.}
 	\label{fig:PSF_tau_convergence}

### 2024-02-19 revision/reply_to_reviewer1.tex
trying to figure out notation for Kronecker points of finite elements
update response to reviewer 1

@@
 
+	We also added more references and discussion regarding the wave inverse problem Hessian at the end of Section 2 in the revised manuscript.
+
 	\begin{figure}
@@
 
- \item \referee{The paper presented a nice complexity analysis of the steps in
-    constructing the approximate Hessian, with the dominating cost
-    arising from the Hessian matrix-vector product. It would be
+ \item \referee{The paper presented a nice complexity analysis of the
+   steps in constructing the approximate Hessian, with the dominating
+   cost arising from the Hessian matrix-vector product. It would be
    interesting to see the computational time of each step for the
-    numerical examples to provide a good sense of the complexity and
-    to support the 5-10X computational reduction by accounting also
-    the overhead beyond the PDE solves.}
+   numerical examples to provide a good sense of the complexity and to
+   support the 5-10X computational reduction by accounting also the
+   overhead beyond the PDE solves.}
 
-   \noindent{Nick will time one of the examples in the
-     advection-diffusion.}
+   \noindent{We agree with the reviewer that there is an ``overhead''
+     beyond the PDE applies. However, for all numerical tests we
+     present in this paper, this cost was small compared to
+     PDE-applies. This is due to the need for a relatively small
+     number of sample points ($m$) in all batches. Since we have not
+     invested time to optimize our codes for performance, it would be
+     perhaps misleading to report profilingtiming information. When
+     $m$ is very large, one can reduce this cost with more involved
+     computational geometry methods. This was beyond the scope of our
+     work.
+
+     In the revised manuscript, following the other reviewer's
+     suggestion, we added a comparison of the computational cost
+     (measured in operator applies) to approximate the spacially
+     varying blurring kernel example using the proposed PSF method,
+     the randomized HODLR (hierarchical off diagonal low rank) method,
+     and GLR (global low rank) approximation using randomized SVD. The
+     results show an overwhelming computational gain for this
+     particular example.}
 

### 2024-02-19 revision/letter_revision.tex
added draft of response letter

--- 2023-07-07 submission_letter/cover_letter.tex
@@
 
-%\noindent
-%Jan S. Hesthaven \\
-%EPFL SB-DO\\
-%PH A2 364 (B\^atiment PH)\\
-%Station 3\\
-%CH-1015 Lausanne \\
-%Switzerland
-%{
-%\flushright{\today}\\
-%}
-% \vspace{0.3cm}
-\noindent Dear Professor Hans De Sterck:\\[2ex]
+\bigskip
 
-We wish to submit the manuscript {\em Point spread function approximation of high rank Hessians with locally supported non-negative integral kernels} by Nick Alger, Tucker Hartland, Noemi Petra, and Omar Ghattas to be considered for publication in SISC in the {\em
-  Numerical Algorithms for Scientific Computing} section.
+\noindent Dear Professor Antil:\\[2ex]
+
+Please find enclosed the revised manuscript ``Point spread function
+approximation of high rank Hessians with locally supported
+non-negative integral kernels'' by Nick Alger, Tucker Hartland, Noemi
+Petra, and Omar Ghattas which we would like to resubmit to SISC in the
+Numerical Algorithms for Scientific Computing section.
+
 %
+This cover letter is accompanied by a point-by-point list of responses
+to the referees' comments. In addition, we have prepared a diff
+document that highlights the differences between the initially
+submitted and the new manuscript, where revisions are indicated by
+color (blue represents new text, and red is deleted text). Please send
+the diff document to the reviewers as well.
 
-In this manuscript, we propose an efficient point spread function (PSF) method for approximating matrix-free operators with locally supported non-negative integral kernels.
-We improve upon existing PSF methods by using a novel moment method for forming a-priori ellipsoid estimates for the supports of impulse responses.
-By solving an ellipsoid packing problem, we compute a large number of impulse responses per operator application, making the method highly efficient in terms of operator applications.
-We also provide an improved method for interpolating impulse responses based on a notion we call ``local mean displacement invariance,'' which generalizes classical local translation invariance.
-We apply the method to approximate high rank Hessians PDE constrained inverse problems.
-Our numerical results demonstrate that the proposed method outperforms conventional approaches, and that the method can approximate high rank Hessians with a small number of operator applies.
-The manuscript has not been submitted elsewhere for
-publication.\\[1ex]
-
-\noindent Thank you for your consideration.\\[3ex]
+We would like to thank the reviewers for their careful reading of our article 
+and their constructive feedback, and many thanks to you for handling the
+manuscript.\\[3ex]
 

### 2024-02-19 paper_sisc_first_revision/localpsf.tex
add discussion of frog convergence figure

@@
 performance of the the HODLR and GLR methods deteriorates.
+
+In Figure \ref{fig:PSF_tau_convergence} we show the convergence of the proposed PSF method on the blur kernel as a function of the total number of impulse responses and the number of batches.
+We show convergence for several ellipsoid size parameters $\tau$, ranging from $2.0$--$4.0$.
+We observe a roughly linear convergence rate until a limit, which depends on $\tau$, is reached.
+I.e., the relative error decreases as $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$ or $\text{constant} \times \# \text{batches}^{-1}$ until the limit is reached, then the error does not decrease any further.
+Increasing $\tau$ lowers this limit, allowing the method to achieve higher accuracy.
+Before this limit is reached, the convergence rate is the same for all $\tau$ as a function the number of impulse responses.
+This means that the convergence is faster for smaller $\tau$ as a function of the number of batches, since smaller $\tau$ means more impulse responses per batch.
 

### 2024-02-19 revision/reply_to_reviewer1.tex
respond to reviewer 1 question 2

@@
   \noindent{
-  	Figure 6 illustrates how the error decreases with increasing numbers of batches and therefore impulse responses. ALSO TABLE 4 CONDITION NUMBER
+  	In Figure 8 in the revised manuscript we added a convergence study for the blur kernel example.
+We show the convergence of the proposed PSF method as a function of the total number of impulse responses and the number of batches, for several ellipsoid size parameters $\tau$ ranging from $2.0$--$4.0$.
+We observe a roughly linear convergence rate until a limit, which depends on $\tau$, is reached.
+The larger $\tau$ is, the smaller the limit is on the achievable error.
+Before this limit, the error decreases at roughly the same rate for all $\tau$ in terms of the number of impulse responses.
+This means that the error decreases faster for smaller $\tau$ as a function of the number of batches, since smaller $\tau$ means more impulse responses per batch.
+Basically, taking $\tau$ larger makes the method converge more slowly, but allows it to eventually achieve a lower level of error.
+We added a discussion of this convergence study in Section 7.1 of the revised manuscript.
 
-  	Frog example with a=1. Error vs \#batches, for different tau. Also do a plot with \#impulses on the x-axis.
+%  	 I.e., the relative error decreases as $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$ or $\text{constant} \times \# \text{batches}^{-1}$ until the limit is reached, then the error does not decrease any further.
+Increasing $\tau$ lowers this limit, allowing the method to achieve higher accuracy.
+Before this limit is reached, the convergence rate is the same for all $\tau$ as a function the number of impulse responses.
 
-  	Nick will try to create a plot that shows some error
-    (e.g. Frob norm, cond number) as a function of number of impulse
-    responses.}
+%  	Figure 6 illustrates how the error decreases with increasing numbers of batches and therefore impulse responses. ALSO TABLE 4 CONDITION NUMBER
+%  	
+%  	Frog example with a=1. Error vs \#batches, for different tau. Also do a plot with \#impulses on the x-axis.
+%  	
+%  	Nick will try to create a plot that shows some error
+%    (e.g. Frob norm, cond number) as a function of number of impulse
+%    responses.
+}
 

### 2024-02-19 revision/reply_to_reviewer1.tex
edits during discussion between Nick and Noemi

@@
 \noindent{
-	We added a numerical study investigating the robustness of the method to violations of the non-negative kernel assumption on two 2D blurring kernel examples (Figure 7).
+	We added a numerical study investigating the robustness of the method to violations of the non-negative kernel assumption on two two-dimensional blurring kernel examples (Figure 7).
 We also added a discussion of the impact of non-locality and negativity on the performance of the method (Section 3.2).
 
-	One takeaway message from Figure 7 is that the impact of negative numbers depends strongly on the spatial configuration of these negative numbers---in the first example (columns 1 and 2 in Figure 7), the negative numbers have minimal impact on the ellipsoid impulse response support estimates, whereas in the second example (columns 3 and 4), even a small amount of negative numbers causes the impulse response support estimate to be poor.
+	One takeaway message from Figure 7 is that the impact of negative numbers depends strongly on the spatial configuration of these negative numbers---in the first example (columns~1 and~2 in Figure~7), the negative numbers have minimal impact on the ellipsoid impulse response support estimates, whereas in the second example (columns~3 and~4), even a small amount of negative numbers causes the impulse response support estimate to be poor.
 The configuration of negative numbers for the integral kernel in wave inverse problem Hessians (including Gauss-Newton Hessians) is similar to that of the second example, and therefore our method is unlikely to be effective for wave Hessians.
@@
 On the right we show a zoomed in impulse response, showcasing the negative numbers in the integral kernel.
-The data for this impulse response was kindly provided to us by Mathew Hu.
+The data for this impulse response was kindly provided to us by Mathew Hu (UT Austin).
 		}
@@
   \noindent{
-  	In Figure 8 in the revised manuscript we added a convergence study for the blur kernel example.
+  	In Figure 8 in the revised manuscript we added a convergence study for one of the blur kernel examples.
 We show the convergence of the proposed PSF method as a function of the total number of impulse responses and the number of batches, for several ellipsoid size parameters $\tau$ ranging from $2.0$--$4.0$.
@@
 This means that the error decreases faster for smaller $\tau$ as a function of the number of batches, since smaller $\tau$ means more impulse responses per batch.
-Basically, taking $\tau$ larger makes the method converge more slowly, but allows it to eventually achieve a lower level of error.
+Basically, with larger $\tau$ the method converge more slowly than with smaller $\tau$, but larger $\tau$ allows the method to eventually achieve a lower level of error.
 We added a discussion of this convergence study in Section 7.1 of the revised manuscript.
+
+  	Although the method can achieve high accuracy, we would like to emphasize that it is targeted at achieving modest accuracies (e.g., $5\%$ error) to build preconditioners for problems that are challenging for existing methods.
+In Table 2 in the revised manuscript, following the other reviewer's suggestion, we added a comparison of the computational cost
+  	(measured in operator applies) to approximate the kernel in one of the spatially
+  	varying blurring kernel examples. We compare the results using the proposed PSF method,
+  	the randomized HODLR (hierarchical off diagonal low rank) method,
+  	and GLR (global low rank) approximation using randomized SVD. The
+  	results show an overwhelming computational gain.
 
@@
    numerical examples to provide a good sense of the complexity and to
-   support the 5-10X computational reduction by accounting also the
-   overhead beyond the PDE solves.}
+   support the 5--10X computational reduction by accounting also the
+   overhead beyond the PDE solves.
+}
 
-   \noindent{We agree with the reviewer that there is an ``overhead''
-     beyond the PDE applies. However, for all numerical tests we
-     present in this paper, this cost was small compared to
-     PDE-applies. This is due to the need for a relatively small
-     number of sample points ($m$) in all batches. Since we have not
-     invested time to optimize our codes for performance, it would be
-     perhaps misleading to report profilingtiming information. When
-     $m$ is very large, one can reduce this cost with more involved
-     computational geometry methods. This was beyond the scope of our
-     work.
-
-     In the revised manuscript, following the other reviewer's
-     suggestion, we added a comparison of the computational cost
-     (measured in operator applies) to approximate the spacially
-     varying blurring kernel example using the proposed PSF method,
-     the randomized HODLR (hierarchical off diagonal low rank) method,
-     and GLR (global low rank) approximation using randomized SVD. The
-     results show an overwhelming computational gain for this
-     particular example.}
+   \noindent{This is a fair point. We agree with the reviewer that there is an overhead
+     beyond the PDE solves. However, since we have not optimized our codes for performance, it would be
+     perhaps misleading to report profiling/timing information. Nevertheless, for all numerical tests we
+     present in this paper, the cost to build the H-matrices (item (2) in Section 6) was small compared to
+     PDE solves (item (1) in Section 6).
+The further H-matrix linear algebra operations (item (3) in Section 6), as implemented in the HLIBPro library, were less costly than the PDE solves, but their costs were not negligible.
+In particular, enforcing positive definiteness of the H-matrix via the procedure presented in Section 5.5 was costly.
+Because of the polylog linear complexity of H-matrix operations, for large-scale problems these H-matrix overhead costs would eventually become insignificant compared to the PDE solves.
+%     
+%     However, for all numerical tests we
+%     present in this paper, this cost was small compared to
+%     PDE-applies. This is due to the need for a relatively small
+%     number of sample points ($m$) in all batches. Since we have not
+%     invested time to optimize our codes for performance, it would be
+%     perhaps misleading to report profiling-timing information. When
+%     $m$ is very large, one can reduce this cost with more involved
+%     computational geometry methods. This was beyond the scope of our
+%     work.
+}
 

### 2024-02-19 paper_sisc_first_revision/localpsf.tex
respond to reviewer 1 question 2
edits during discussion between Nick and Noemi

--- 2024-02-19 paper_sisc_first_revision/localpsf.tex
@@
 However, in Section~\ref{sec:sample_point_selection} we will see that the cost of the proposed method depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids is better), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
-In practice, we find that $\tau=3.0$ yields a reasonable balance between these competing interests, and use $\tau=3.0$ in the numerical results.
+In practice, we find that $\tau=3.0$ yields a reasonable balance between these competing interests, and use $\tau=3.0$ in all numerical results, except for Figure \ref{fig:PSF_tau_convergence}, where we study the effects of varying $\tau$.
 The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is typically conservative.
@@
 %and may be easily visualized.
-
-In Table~\ref{tab:frog_psf_vs_hodlr_vs_glr} we compare the
-computational cost to approximate the blur kernel from
-Equation~\eqref{eq:frog_kernel} using the proposed PSF method, the
-randomized HODLR (hierarchical off diagonal low rank)
-method~\cite{Martinsson16,Hartland_2023}, and GLR (global low rank)
-approximation using randomized SVD~\cite{HalkoMartinssonTropp11}. For
-these results we vary the quantity $L$ to scale the width of the
-impulse responses and hence the rank of the operator. For each case,
-we calculate the relative error in the approximation of the kernel
-measured in the Frobenius norm, $||\boldsymbol{\Phi} -
-\widetilde{\boldsymbol{\Phi}}||_\text{Fro} /
-||\boldsymbol{\Phi}||_\text{Fro}$ and show the number of operator
-applications required to achieve 20\%, 10\%, and 5\% relative error by
-each method. The results reveal superior performance of the PSF
-method as compared to the HODLR and GLR methods for all cases. We note
-that as we increase the rank and decrease the error tolerance, the
-performance of the the HODLR and GLR methods deteriorates.
-
-%Define HODLR. Define randomized HODLR algo. Ref to Tucker's paper and "Fast randomized eigensolver with structured LDL..."
-
-%Define randomized SVD, refer to Halko Martinsson Tropp.
-
-%Discuss why PSF does better with smaller $L$.
 
@@
 	\end{center}
-	\caption{(Spatially variant blur) Comparison of cost to
+	\caption{(Blur) Comparison of cost to
 		approximate the blur kernel from Equation
@@
 
+In Table~\ref{tab:frog_psf_vs_hodlr_vs_glr} we compare the
+computational cost to approximate the blur kernel from
+Equation~\eqref{eq:frog_kernel} using the proposed PSF method, the
+randomized HODLR (hierarchical off diagonal low rank)
+method~\cite{Martinsson16,Hartland_2023}, and GLR (global low rank)
+approximation using randomized SVD~\cite{HalkoMartinssonTropp11}. For
+these results we vary the quantity $L$ to scale the width of the
+impulse responses and hence the rank of the operator. For each case,
+we calculate the relative error in the approximation of the kernel
+measured in the Frobenius norm, $||\boldsymbol{\Phi} -
+\widetilde{\boldsymbol{\Phi}}||_\text{Fro} /
+||\boldsymbol{\Phi}||_\text{Fro}$ and show the number of operator
+applications required to achieve 20\%, 10\%, and 5\% relative error by
+each method. The results reveal superior performance of the PSF
+method as compared to the HODLR and GLR methods for all cases. We note
+that as we increase the rank and decrease the error tolerance, the
+performance of the the HODLR and GLR methods deteriorates.
+
 \begin{figure}
@@
 	\end{subfigure}
-	\caption{Relative error in the PSF approximation of the kernel, $||\boldsymbol{\Phi} - \widetilde{\boldsymbol{\Phi}}||_\text{Fro}/||\boldsymbol{\Phi}||_\text{Fro}$, for different ellipsoid size parameters, $\tau$.
-Left: Relative error vs.
-the total number of impulse responses used in the approximation.
-Right: Relative error vs.
-the number of impulse response batches.
+	\caption{(Blur) Relative error 
+		%		in the PSF approximation of the kernel
+		%	$||\boldsymbol{\Phi} - \widetilde{\boldsymbol{\Phi}}||_\text{Fro}/||\boldsymbol{\Phi}||_\text{Fro}$,
+		for different ellipsoid size parameters, $\tau$, vs. the total number of impulse responses (left), and the number of batches (right). 
+		%	Left: Relative error vs. the total number of impulse responses used in the approximation. Right: Relative error vs. the number of impulse response batches. 
 		The dashed gray lines show linear convergence rates, i.e., $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$ on the left, and  $\text{constant} \times \# \text{batches}^{-1}$ on the right.}
@@
 \end{figure}
+
+In Figure \ref{fig:PSF_tau_convergence} we show the convergence of the proposed PSF method on the blur kernel as a function of the total number of impulse responses (left), and the number of batches (right).
+We show convergence for several ellipsoid size parameters $\tau$, ranging from $2.0$--$4.0$.
+The results in Figure \ref{fig:PSF_tau_convergence} (left) show that the relative error decreases as $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$
+% or $\text{constant} \times \# \text{batches}^{-1}$, 
+suggesting linear convergence. The linear convergence stalls at a limit which depends on $\tau$. Increasing $\tau$ lowers this limit, allowing the method to achieve higher accuracy. 
+In Figure \ref{fig:PSF_tau_convergence} (right), the results show that before this limit is reached, the convergence is faster for smaller $\tau$ in terms of the number of batches.
+This is expected because smaller $\tau$ results in more impulse responses per batch.
+Larger $\tau$ causes the method converge more slowly than smaller $\tau$, but with larger $\tau$ the method stalls at a lower level of error than it does with smaller $\tau$ (see, e.g., $\tau=4.0$ vs.
+$\tau=2.5$).
+
+%----
+%, but allows the method to eventually achieve a lower level of error.
+%
+%----
+%
+%For smaller $\tau$ regime, we observe faster convergence
+%
+%Before this limit is reached, the convergence rate is the same for all $\tau$ as a function the number of impulse responses.
+This means that the convergence is faster for smaller $\tau$ as a function of the number of batches, since smaller $\tau$ means more impulse responses per batch.
+In summary, taking $\tau$ larger makes the method converge more slowly, but allows it to eventually achieve a lower level of error.
+
+%Define HODLR. Define randomized HODLR algo. Ref to Tucker's paper and "Fast randomized eigensolver with structured LDL..."
+
+%Define randomized SVD, refer to Halko Martinsson Tropp.
+
+%Discuss why PSF does better with smaller $L$.
 

### 2024-02-19 revision/reply_to_reviewer2.tex
edits during discussion between Nick and Noemi

@@
 
-    \noindent{We added a numerical comparison of our method with the HODLR method in [7] for a blur operator in Table 2 of Section 7.1 in the revised manuscript.
-The results show that the HODLR method requires hundreds to thousands of operator applies to achieve modest accuracies (e.g.
-10\% error) while our method requires only order 10 operator applies.
+    \noindent{We added a numerical comparison of our method with the HODLR method in [7] for a blur kernel in Table 2 of Section 7.1 in the revised manuscript.
+The results show that the HODLR method requires hundreds to thousands of operator applies to achieve modest accuracies (e.g., 10\% error) while our method requires only order 10 operator applies.
 In Section 2 of the revised manuscript we also added a discussion of the costs of the method proposed in [52].

### 2024-02-19 paper_sisc_first_revision/localpsf.tex
edits during discussion between Nick and Noemi

@@
 Further matrix computations are then performed with fast H-matrix methods.
-We demonstrate the effectiveness of this PSF-based method by using it to build preconditioners for the Hessian operator arising in two inverse problems governed by partial differential equations (PDEs): inversion for the basal friction coefficient in an ice sheet flow problem and for the initial condition in an advective-diffusive transport problem.
+This end-to-end procedure is illustrated on a blur problem.
+We demonstrate the PSF-based method's effectiveness by using it to build preconditioners for the Hessian operator arising in two inverse problems governed by partial differential equations (PDEs): inversion for the basal friction coefficient in an ice sheet flow problem and for the initial condition in an advective-diffusive transport problem.
 While for many ill-posed inverse problems the Hessian of the data misfit term exhibits a low rank structure, and hence a low rank approximation is suitable, for many problems of practical interest the numerical rank of the Hessian is still large.
@@
 The moment method that we use estimates the support of $\phi_x$ more accurately than random noise probing in resolution analysis, at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
-(3) {\color{red} The method we propose never evaluates computed impulse responses outside of their domain of definition.
-This eliminates boundary artifacts which plague conventional PC methods.}
+(3) The method we propose never evaluates computed impulse responses outside of their domain of definition.
+This eliminates ``boundary-artifact'' errors (see \cite[Section 1.1]{AlgerEtAl19}) that plague conventional PC methods.
 

### 2024-02-19 paper_sisc_first_revision/localpsf.tex
edits during discussion between Nick and Noemi

@@
 
-Here we provide an overview of the operator approximation algorithm.
-Then, in the subsections (Sections \ref{sec:sample_point_selection}--\ref{sec:flip_eigs}) we present technical details of the algorithm that were not already covered in Section \ref{sec:prerequisites}.
-
-%\subsection{Algorithm overview}
-%\label{sec:algorithm_overview}
+Before presenting the technical details of the algorithm in Sections \ref{sec:sample_point_selection}--\ref{sec:flip_eigs}, we first provide an overview.
+%Here we provide an overview of the operator approximation algorithm. 
+%Then, in the subsections (Sections \ref{sec:sample_point_selection}--\ref{sec:flip_eigs}) we present technical details of the algorithm that were not already covered in Section \ref{sec:prerequisites}.
 

### 2024-02-19 revision/letter_revision.tex
edits during discussion between Nick and Noemi

@@
 
-Please find enclosed the revised manuscript ``Point spread function
+Please find enclosed our revision to the manuscript ``Point spread function
 approximation of high rank Hessians with locally supported
 non-negative integral kernels'' by Nick Alger, Tucker Hartland, Noemi
-Petra, and Omar Ghattas which we would like to resubmit to SISC in the
+Petra, and Omar Ghattas, which was submitted to SISC in the
 Numerical Algorithms for Scientific Computing section.
@@
 document that highlights the differences between the initially
-submitted and the new manuscript, where revisions are indicated by
+submitted manuscript and the revised manuscript, where revisions are indicated by
 color (blue represents new text, and red is deleted text). Please send

### 2024-02-19 paper_sisc_first_revision/localpsf.tex
addressing consistency issues

@@
 
-The idea of the method is to use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applications.
+The idea of the proposed PSF-based method is to use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applications.
 The impulse response, $\phi_x$, associated with a point, $x$, is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
@@
 The proposed method improves upon existing PC methods in the following ways: (1) While PC approximations are based on an assumption of local translation invariance, the method we propose is based on a more general assumption we call ``local mean displacement invariance'' (Section~\ref{sec:local_mean_displacement_invariance} and Figure~\ref{fig:mean_displacement_invariance}), which improves the interpolation of the impulse responses.
-(2) In our previous work~\cite{AlgerEtAl19}, we chose point sources in adaptive grid via a sequential procedure; the refinements to the adaptive grid were chosen to reduce the error as much as possible at each step.
+(2) In our previous work~\cite{AlgerEtAl19}, we chose point sources in an adaptive grid via a sequential procedure; the refinements to the adaptive grid were chosen to maximally reduce the error at each step.
 However, in that work each point source required a separate operator application, making the previous method expensive when a large number of impulse responses is desired.
@@
 
-While the PSF method in this paper may be used to approximate any operator that has a locally supported non-negative integral kernel, we are primarily motivated by approximation of high-rank Hessians in distributed parameter inverse problems governed by PDEs.
+While the PSF-based method in this paper may be used to approximate any operator that has a locally supported non-negative integral kernel, we are primarily motivated by approximation of high-rank Hessians in distributed parameter inverse problems governed by PDEs.
 In this section, we provide a brief background on this topic, and explain why existing Hessian approximation methods are not satisfactory.
@@
 On a $64 \times 64$ grid with depth $4$, hierarchical rank $10$, and oversampling parameter $c=5$, this works out to $4 \cdot 64 \cdot (10 + 5) = 3840$ operator applies.
-In Section \ref{sec:frog} see numerically that the randomized hierarchical off-diagonal low rank (HODLR) method in \cite{Martinsson16} requires hundreds to thousands of matrix-vector products to construct approximations of blur operators with modest (e.g., $10\%$) relative error.
+In Section \ref{sec:frog} see numerically that the randomized hierarchical off-diagonal low rank (HODLR) method in \cite{Martinsson16} requires hundreds to thousands of matrix-vector products to construct approximations of the integral kernel for a blur problem example with modest (e.g., $10\%$) relative error.
 %while the PSF method in this paper can achieve the same level of error in one to two dozen operator applies.
@@
 It is acceptable if the support of $\impulseresponse_x$ is not perfectly contained in a localized region (violating Assumption 2), so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
-In principle, the method can be applied even if the impulse responses are not localized.
+In principle, the proposed method can be applied even if the impulse responses are not localized.
 However, in this case only a small number of impulse responses can be computed per batch, which means more batches, and hence more operator applies, are needed to form an accurate approximation.
@@
 
-If there are negative numbers in the integral kernel (violating Assumption 3), the ellipsoid estimation procedure may incur errors or fail, leading to poor performance or failure of the method.
+If there are negative numbers in the integral kernel (violating Assumption 3), the ellipsoid estimation procedure may incur errors or fail, leading to poor performance or failure of the proposed method.
 In Figure \ref{fig:robustness_to_negatives} we investigate the robustness of the ellipsoid support estimation procedure to violations of Assumption 3.
@@
 This is because moment formulas (Equations \ref{eq:define_mean} and \ref{eq:define_cov}) assign more weight to entries in the kernel that are further from the center.
-	\item Negative numbers in the integral kernel affect the method more if they are isolated, and less if they are balanced by nearby positive numbers.
+	\item Negative numbers in the integral kernel affect the proposed method more if they are isolated, and less if they are balanced by nearby positive numbers.
 	\item As kernels become more negative, the ellipsoid estimation performs well up to a certain threshold that depends on the spatial distribution of negative and positive entries.
@@
 %The location of the negative numbers also affects the accuracy of the ellipsoid estimates substantially, but in a more complicated manner. 
-For the kernel in the left two columns of Figure \ref{fig:robustness_to_negatives}, negative numbers are interspersed with positive numbers, allowing us to include a large amount of negative numbers in the integral kernel before the method fails.
+For the kernel in the left two columns of Figure \ref{fig:robustness_to_negatives}, negative numbers are interspersed with positive numbers, allowing us to include a large amount of negative numbers in the integral kernel before the proposed method fails.
 For the kernel in the right two columns, the ellipsoid estimate fails with tiny amounts of negative numbers because the negative numbers are far away from the mean and not balanced by positive numbers at similar distances and angles.
-In the bottom two rows, we see the aforementioned threshold effect, in which the method rapidly transitions from performing reasonably well to performing poorly with only a small change to the integral kernel.
+In the bottom two rows, we see the aforementioned threshold effect, in which the proposed method rapidly transitions from performing reasonably well to performing poorly with only a small change to the integral kernel.
 
@@
 	\caption{Illustration of the influence of negative numbers in the integral kernel on the robustness of the ellipsoid estimates for the supports of impulse responses.
-Left two columns: spatially variant blur kernel given in Equation \ref{eq:frog_kernel}.
+Left two columns: Blur kernel given in Equation \ref{eq:frog_kernel}.
 Right two columns: Ricker wavelet-type kernel given by $\Aker(y,x) = (1 - a \gamma)\exp\left(-\gamma/2\right)$, where $\gamma=(y-x)^T \Sigma^{-1} (y-x)$, and $\Sigma=\operatorname{diag}\left(0.0025, 0.01\right)$.
@@
 %This is illustrated in Figure \ref{fig:RBF_width} for a simple one dimensional interpolation problem. 
-For our numerical results involving Hessians in inverse problems governed by PDEs (Sections~\ref{sec:ice}~and~\ref{sec:adv}), high accuracy is not required because the method is used to construct a preconditioner.
+For our numerical results involving Hessians in inverse problems governed by PDEs (Sections~\ref{sec:ice}~and~\ref{sec:adv}), high accuracy is not required because the proposed method is used to construct a preconditioner.
 Hence, for these Hessian approximations we use a conservative choice of $\epsilon=3.0$ to ensure robustness.
-For our numerical results for blur kernels (Section~\ref{sec:frog}), we use a smaller value of $\epsilon=0.5$ so that the RBF interpolation accuracy is not a limiting factor as we study convergence of the method.
+For our numerical results for the blur problem example (Section~\ref{sec:frog}), we use a smaller value of $\epsilon=0.5$ so that the RBF interpolation accuracy is not a limiting factor as we study convergence of the proposed method.
 
@@
 $\mathcal{H} = \Aop + \mathcal{R},$
-where $\Aop$ is a symmetric positive semi-definite operator that we approximate with the PSF method to form an H-matrix $\mathbf{A}_H$, and $\mathcal{R}$ is a symmetric positive definite operator that may be easily converted to an H-matrix $\mathbf{R}_H$ without using the PSF method.
+where $\Aop$ is a symmetric positive semi-definite operator that we approximate with the proposed method to form an H-matrix $\mathbf{A}_H$, and $\mathcal{R}$ is a symmetric positive definite operator that may be easily converted to an H-matrix $\mathbf{R}_H$ without using the proposed method.
 For example, in inverse problems $\mathcal{H}$ is the Hessian, $\Aop$ is the data misfit term in the Hessian which is dense and available only matrix-free, and $\mathcal{R}$ is the regularization term, which is typically an elliptic differential operator that becomes a sparse matrix after discretization.
@@
 \paragraph{(1) Computing impulse response moments and batches} Computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ requires applying $\Aop^T$ to $1$, $\gdim$, and $\gdim(\gdim+1)/2$ vectors, respectively.
+This works out to $3$ applications of $\Aop^T$ in one spatial dimension, $6$ in two dimensions, and $10$ in three dimensions.
 Computing each $\eta_i$ requires applying $\Aop$ to one vector, so computing $\{\eta_i\}_{i=1}^{\nbatch}$ requires $\nbatch$ operator applications.
@@
 where $a,b \in \{0,1,2,3\}$ are constants which depend on the type of H-matrix used and the operation being performed~\cite{GrasedyckHackbusch03}\cite[Section 2.1]{Kriemann13}.
-In the numerical results (Section~\ref{sec:numerical_results}), we use one matrix-matrix addition to add the H-matrix approximation of the data misfit term in the Hessian to the regularization term in the Hessian.
+For our numerical results involving Hessians (Sections~\ref{sec:ice} and~\ref{sec:adv}), we use one matrix-matrix addition to add the H-matrix approximation of the data misfit term in the Hessian to the regularization term in the Hessian.
 Symmetrizing $\mathbf{A}_H$ requires one matrix-matrix addition.
@@
 
-Here we define a PDE-free spatially varying blurring problem, in which the impulse response, $\phi_x$,
+Here we define a PDE-free spatially varying blur problem, in which the impulse response, $\phi_x$,
 is a bumpy blob that is centered
 near $x$, and is rotated and scaled in a manner that depends on $x$
-(see Figure \ref{fig:frog_kernel_impulse_responses}). This blurring problem is used in
+(see Figure \ref{fig:frog_kernel_impulse_responses}). This blur problem is used in
 Sections~\ref{sec:intro} and~\ref{sec:prelims} to visually illustrate
 various stages and aspects of the proposed method, and robustness (or lack thereof) of the proposed method to violations of the non-negative kernel assumption (Section~\ref{sec:conditions_2}).
-In this section, the blurring problem is used to compare the proposed method to hierarchical off-diagonal low rank (HODLR) and global low rank (GLR) methods, to study convergence of the method, and to investigate the effect of the ellipsoid size parameter $\tau$.
+In this section, the blur problem is used to compare the proposed method to hierarchical off-diagonal low rank (HODLR) and global low rank (GLR) methods, to study convergence of the proposed method, and to investigate the effect of the ellipsoid size parameter $\tau$.
 %To study the
@@
 > 1$, the integral kernel contains negative values, which allows us to
-study the robustness of the PSF method to violations of Assumption 3
+study the robustness of the proposed method to violations of Assumption 3
 (Section~\ref{sec:conditions_2}).
@@
 		approximate the blur kernel from Equation
-		\eqref{eq:frog_kernel} using the PSF method, the randomized
+		\eqref{eq:frog_kernel} using the proposed PSF-based method, the randomized
 		HODLR (hierarchical off diagonal low rank) method, and GLR
@@
 computational cost to approximate the blur kernel from
-Equation~\eqref{eq:frog_kernel} using the proposed PSF method, the
+Equation~\eqref{eq:frog_kernel} using the proposed PSF-based method, the
 randomized HODLR (hierarchical off diagonal low rank)
@@
 
-In Figure \ref{fig:PSF_tau_convergence} we show the convergence of the proposed PSF method on the blur kernel as a function of the total number of impulse responses (left), and the number of batches (right).
+In Figure \ref{fig:PSF_tau_convergence} we show the convergence of the proposed method on the blur kernel as a function of the total number of impulse responses (left), and the number of batches (right).
 We show convergence for several ellipsoid size parameters $\tau$, ranging from $2.0$--$4.0$.
@@
 % or $\text{constant} \times \# \text{batches}^{-1}$, 
-suggesting linear convergence. The linear convergence stalls at a limit which depends on $\tau$. Increasing $\tau$ lowers this limit, allowing the method to achieve higher accuracy. 
+suggesting linear convergence. The linear convergence stalls at a limit which depends on $\tau$. Increasing $\tau$ lowers this limit, allowing the proposed method to achieve higher accuracy. 
 In Figure \ref{fig:PSF_tau_convergence} (right), the results show that before this limit is reached, the convergence is faster for smaller $\tau$ in terms of the number of batches.
 This is expected because smaller $\tau$ results in more impulse responses per batch.
-Larger $\tau$ causes the method converge more slowly than smaller $\tau$, but with larger $\tau$ the method stalls at a lower level of error than it does with smaller $\tau$ (see, e.g., $\tau=4.0$ vs.
+Larger $\tau$ causes the proposed method converge more slowly than smaller $\tau$, but with larger $\tau$ the proposed method stalls at a lower level of error than it does with smaller $\tau$ (see, e.g., $\tau=4.0$ vs.
 $\tau=2.5$).
@@
 Note that $\Gamma_b$ is a two-dimensional surface embedded in three dimensions due to the mountain topography.
-The proposed PSF method involves translating impulse responses.
+The proposed PSF-based method involves translating impulse responses.
 Hence it requires either a flat domain, or a notion of local parallel transport.
@@
 %	The Newton iteration is terminated when $||\mathbf{g}|| < 10^{-6} ||\mathbf{g}_0||$, where $\mathbf{g}$ is the gradient at the current iteration, and $\mathbf{g}_0$ is the gradient at the initial guess.
-	Preconditioners shown are the proposed PSF method with five batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
+	Preconditioners shown are the proposed PSF-based method with five batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
 Columns \#CG show the number of PCG iterations used to solve the Newton system for $\mathbf{\widehat{\basalfriction}}$.
@@
     \end{subfigure}
-	\caption{(Ice sheet) Left: convergence history for solving $\mathbf{H} \mathbf{x} = \mathbf{b}$ using PCG, where $\mathbf{b}$ has i.i.d.\ random entries drawn from the standard Gaussian distribution and $\mathbf{H}$ is evaluated at the solution of the inverse problem.
+	\caption{(Ice sheet) Left: Convergence history for solving $\mathbf{H} \mathbf{x} = \mathbf{b}$ using PCG, where $\mathbf{b}$ has i.i.d.\ random entries drawn from the standard Gaussian distribution and $\mathbf{H}$ is evaluated at the solution of the inverse problem.
 	Results in these figures are shown for the PSF-based preconditioners with 1, 5, and 25 batches (PSF (1), PSF (5), and PSF (25), respectively), regularization preconditioning (REG), and no preconditioning (NONE).
 The preconditioner is constructed using $\mathbf{H}_\text{gn}$.
-	Right: generalized eigenvalues for generalized eigenvalue problem $\mathbf{H} \mathbf{u}_{k}=\lambda_{k} \mathbf{\preconditioner} \mathbf{u}_{k}$.
+	Right: Generalized eigenvalues for generalized eigenvalue problem $\mathbf{H} \mathbf{u}_{k}=\lambda_{k} \mathbf{\preconditioner} \mathbf{u}_{k}$.
 Here $\mathbf{H}$ is the Hessian and the matrices $\mathbf{\preconditioner}$ are the same Hessian approximations used in the left sub-figure, with NONE corresponding to the identity matrix.}
@@
 	\end{subfigure}
-	\caption{(Advective-diffusive transport) Left: convergence history for solving $\mathbf{H} \mathbf{x} = \mathbf{b}$ using PCG, where $\mathbf{b}$ has i.i.d.\ random entries drawn from the standard Gaussian distribution.
-Right: generalized eigenvalues for generalized eigenvalue problem $\mathbf{H} \mathbf{u}_{k}=\lambda_{k} \mathbf{\preconditioner} \mathbf{u}_{k}$.
+	\caption{(Advective-diffusive transport) Left: Convergence history for solving $\mathbf{H} \mathbf{x} = \mathbf{b}$ using PCG, where $\mathbf{b}$ has i.i.d.\ random entries drawn from the standard Gaussian distribution.
+Right: Generalized eigenvalues for generalized eigenvalue problem $\mathbf{H} \mathbf{u}_{k}=\lambda_{k} \mathbf{\preconditioner} \mathbf{u}_{k}$.
 Here $\mathbf{H}$ is the Hessian and the preconditioner, $\mathbf{\preconditioner}$, is the PSF approximation for 1, 5, or 25 batches (PSF (1), PSF (5), and PSF (25), respectively), the regularization Hessian (REG), or the identity matrix (NONE).}
@@
 
-We presented an efficient matrix-free PSF method for approximating operators with locally supported non-negative integral kernels.
-The method requires access to the operator only via application of the operator to a small number of vectors.
-The idea of the method is to compute batches of impulse responses by applying the operator to Dirac combs of scattered point sources, then interpolate these impulse responses to approximate entries of the operator's integral kernel.
+We presented an efficient matrix-free PSF-based method for approximating operators with locally supported non-negative integral kernels.
+The proposed method requires access to the operator only via application of the operator to a small number of vectors.
+The idea of the proposed method is to compute batches of impulse responses by applying the operator to Dirac combs of scattered point sources, then interpolate these impulse responses to approximate entries of the operator's integral kernel.
 The interpolation is based on a new principle we call ``local mean displacement invariance,'' which generalizes classical local translation invariance.
@@
 Point source locations for the impulse response batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulse responses per batch as possible, while ensuring that the corresponding ellipsoids do not overlap.
-We applied the method to approximate the Gauss-Newton Hessians in an ice sheet flow inverse problem governed by a linear Stokes PDE, and an advective-diffusive transport inverse problem governed by an advection-diffusion PDE.
+We applied the proposed method to approximate the Gauss-Newton Hessians in an ice sheet flow inverse problem governed by a linear Stokes PDE, and an advective-diffusive transport inverse problem governed by an advection-diffusion PDE.
 We saw that preconditioners based on the proposed approximation cluster the eigenvalues of the preconditioned Hessian near one, and allow us to solve the inverse problems using roughly $5\times$--$10\times$ fewer PDE solves.
@@
 Hence, we expect the speedup will be even greater for such problems.
-Although the method we presented is not applicable to all Hessians, it is applicable to many Hessians of practical interest.
+Although the proposed method is not applicable to all Hessians, it is applicable to many Hessians of practical interest.
 For these Hessians, the proposed method offers a \emph{data scalable} alternative to conventional low rank approximation, due to the ability to form high rank approximations of an operator using a small number of operator applications, and thus PDE solves.

### 2024-02-19 paper_sisc_first_revision/localpsf.tex
about to do one last read before sending to Omar

@@
 
-If there are negative numbers in the integral kernel (violating Assumption 3), the ellipsoid estimation procedure may incur errors or fail, leading to poor performance or failure of the proposed method.
-In Figure \ref{fig:robustness_to_negatives} we investigate the robustness of the ellipsoid support estimation procedure to violations of Assumption 3.
-We study two integral kernel examples, both of which are parameterized by a quantity that controls how negative the kernels are.
-We make the following observations:
-\begin{itemize}
-	\item The larger and more numerous the negative numbers are, the more inaccurate the ellipsoid support estimate is.
-	\item The further away from the center of the ellipsoid the negative numbers are, the more influence they have on the quality of the ellipsoid support estimate.
-This is because moment formulas (Equations \ref{eq:define_mean} and \ref{eq:define_cov}) assign more weight to entries in the kernel that are further from the center.
-	\item Negative numbers in the integral kernel affect the proposed method more if they are isolated, and less if they are balanced by nearby positive numbers.
-	\item As kernels become more negative, the ellipsoid estimation performs well up to a certain threshold that depends on the spatial distribution of negative and positive entries.
-After that threshold is crossed, the estimation rapidly transitions to performing poorly and ultimately failing.
-\end{itemize} 
-%The location of the negative numbers also affects the accuracy of the ellipsoid estimates substantially, but in a more complicated manner. 
-For the kernel in the left two columns of Figure \ref{fig:robustness_to_negatives}, negative numbers are interspersed with positive numbers, allowing us to include a large amount of negative numbers in the integral kernel before the proposed method fails.
-For the kernel in the right two columns, the ellipsoid estimate fails with tiny amounts of negative numbers because the negative numbers are far away from the mean and not balanced by positive numbers at similar distances and angles.
-In the bottom two rows, we see the aforementioned threshold effect, in which the proposed method rapidly transitions from performing reasonably well to performing poorly with only a small change to the integral kernel.
-
-%Location and magnitude of negatives matter.
-Larger negatives: bad, smaller negatives good.
-Negatives closer to center: good, further from center: bad.
-Correlated negatives: bad, randomly distributed: good.
-Good Good Good Fail quick
-%
-%Non-local: will work, but choose few ellipsoids, so inefficient
-
 \begin{figure}
@@
 \end{figure}
+
+If there are negative numbers in the integral kernel (violating Assumption 3), the ellipsoid estimation procedure may incur errors or fail, leading to poor performance or failure of the proposed method.
+In Figure \ref{fig:robustness_to_negatives} we investigate the robustness of the ellipsoid support estimation procedure to violations of Assumption 3.
+We study two integral kernel examples, both of which are parameterized by a quantity that controls how negative the kernels are.
+We make the following observations:
+\begin{itemize}
+	\item The larger and more numerous the negative numbers are, the more inaccurate the ellipsoid support estimate is.
+	\item The further away from the center of the ellipsoid the negative numbers are, the more influence they have on the quality of the ellipsoid support estimate.
+This is because moment formulas (Equations \ref{eq:define_mean} and \ref{eq:define_cov}) assign more weight to entries in the kernel that are further from the center.
+	\item Negative numbers in the integral kernel affect the proposed method more if they are isolated, and less if they are balanced by nearby positive numbers.
+	\item As kernels become more negative, the ellipsoid estimation performs well up to a certain threshold that depends on the spatial distribution of negative and positive entries.
+After that threshold is crossed, the estimation rapidly transitions to performing poorly and ultimately failing.
+\end{itemize} 
+%The location of the negative numbers also affects the accuracy of the ellipsoid estimates substantially, but in a more complicated manner. 
+For the kernel in the left two columns of Figure \ref{fig:robustness_to_negatives}, negative numbers are interspersed with positive numbers, allowing us to include a large amount of negative numbers in the integral kernel before the proposed method fails.
+For the kernel in the right two columns, the ellipsoid estimate fails with tiny amounts of negative numbers because the negative numbers are far away from the mean and not balanced by positive numbers at similar distances and angles.
+In the bottom two rows, we see the aforementioned threshold effect, in which the proposed method rapidly transitions from performing reasonably well to performing poorly with only a small change to the integral kernel.
+
+%Location and magnitude of negatives matter.
+Larger negatives: bad, smaller negatives good.
+Negatives closer to center: good, further from center: bad.
+Correlated negatives: bad, randomly distributed: good.
+Good Good Good Fail quick
+%
+%Non-local: will work, but choose few ellipsoids, so inefficient
 
@@
 
-We use the proposed method to approximate the Newton (or Gauss-Newton) Hessians
+We use the proposed method to approximate the integral kernel in a blur problem (Section~\ref{sec:frog}), and the Newton (or Gauss-Newton) Hessians
 in inverse problems governed by PDEs which model steady state ice
@@
 (Section~\ref{sec:adv}). These inverse problems are described in detail in their respective sections.
-In both cases, to reconstruct the unknown parameter fields, denoted $q$, the inverse problems are formulated as nonlinear least squares
+In both inverse problems, to reconstruct the unknown parameter fields, denoted $q$, the inverse problems are formulated as nonlinear least squares
 optimization problems, whose objective functions consist of a data misfit term

### 2024-02-20 paper_sisc_first_revision/localpsf.tex
reduced to 32 pages

@@
 	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels.
-The method computes impulse responses of the operator at scattered points, and interpolates these impulse responses to approximate entries of the integral kernel.
+The proposed method computes impulse responses of the operator at scattered points, and interpolates these impulse responses to approximate entries of the integral kernel.
 To compute impulse responses efficiently, we apply the operator to Dirac combs associated with batches of point sources, which are chosen by solving an ellipsoid packing problem.
@@
 This end-to-end procedure is illustrated on a blur problem.
-We demonstrate the PSF-based method's effectiveness by using it to build preconditioners for the Hessian operator arising in two inverse problems governed by partial differential equations (PDEs): inversion for the basal friction coefficient in an ice sheet flow problem and for the initial condition in an advective-diffusive transport problem.
+We demonstrate the proposed method's effectiveness by using it to build preconditioners for the Hessian operator arising in two inverse problems governed by partial differential equations (PDEs): inversion for the basal friction coefficient in an ice sheet flow problem and for the initial condition in an advective-diffusive transport problem.
 While for many ill-posed inverse problems the Hessian of the data misfit term exhibits a low rank structure, and hence a low rank approximation is suitable, for many problems of practical interest the numerical rank of the Hessian is still large.
-The Hessian impulse responses on the other hand typically become more local as the numerical rank increases, which benefits the PSF-based method.
-Numerical results reveal that the PSF-based preconditioner clusters the spectrum of the preconditioned Hessian near one, yielding roughly $5\times$--$10\times$ reductions in the required number of PDE solves, as compared to classical regularization-based preconditioning and no preconditioning.
+The Hessian impulse responses on the other hand typically become more local as the numerical rank increases, which benefits the proposed method.
+Numerical results reveal that the preconditioner clusters the spectrum of the preconditioned Hessian near one, yielding roughly $5\times$--$10\times$ reductions in the required number of PDE solves, as compared to classical regularization-based preconditioning and no preconditioning.
 We also present a comprehensive numerical study for the influence of various parameters (that control the shape of the impulse responses and the rank of the Hessian) on the effectiveness of the advection-diffusion Hessian approximation.
-The results show that the PSF-based preconditioners are able to form good approximations of high-rank Hessians using only a small number of operator applications.
+The results show that the proposed method is able to form good approximations of high-rank Hessians using only a small number of operator applications.
 \end{abstract}
@@
 Schur complements in Schur complement methods for solving partial differential equations and Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps)~\cite{ChanMathew94,SmithBjorstadGropp96,ToselliWidlund04}, covariance operators in spatial statistics~\cite{ChenStein21,GentonKeyesTurkiyyah18,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}.
-
 Here, ``matrix-free'' means that we may apply $\Aop$ and its transpose\footnote{Recall that $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for operations with $\Aop^T$.
@@
 via a black box computational procedure, 
-but cannot easily access entries of $\Aop$'s integral kernel.
-Evaluating the maps in~\eqref{eq:eval_maps} may require solving a subproblem that involves PDEs, or performing other costly computations.
-Matrix-free does not mean that matrices are never used; our approximation is ultimately converted into a hierarchical matrix (H-matrix), a compressed matrix format that permits fast and scalable linear algebra operations.
+but cannot easily access entries of $\Aop$'s integral kernel. Evaluating the maps in~\eqref{eq:eval_maps} may require solving a subproblem that involves PDEs, or performing other costly computations. 
+%Matrix-free does not mean that matrices are never used; our approximation is ultimately converted into a hierarchical matrix (H-matrix), a compressed matrix format that permits fast and scalable linear algebra operations.
 
@@
 		%		\centering
-		\includegraphics[scale=0.35]{kernel_impulse_illustration2.pdf}
+		\includegraphics[scale=0.3]{kernel_impulse_illustration2.pdf}
 		%%		\caption{Kernel and impulse responses}
@@
 	\caption{Left: Matrix created by evaluating the integral kernel $\Phi$ for $\Aop$ (Equation~\ref{eq:kernel_representation}) at all pairs of mesh vertices.
-This illustration is for the integral kernel given in Equation~\ref{eq:frog_kernel} in Section~\ref{sec:frog}.
+This illustration is for the integral kernel in Equation~\ref{eq:frog_kernel}.
 Dark colors indicate large entries and light colors indicate small entries.
@@
 
-The idea of the proposed PSF-based method is to use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applications.
+The idea of the proposed method is to use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applications.
 The impulse response, $\phi_x$, associated with a point, $x$, is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
@@
 Batches of impulse responses may be thought of intuitively as sets of ``columns'' of the kernel (Figure \ref{fig:frog_kernel_impulse_responses}).
-To choose the batches, we form ellipsoid estimates for the supports of all $\phi_x$ via a moment method (Figure~\ref{fig:frog_moments_ellipsoid}) that involves applying $\Aop^T$ to a small number of polynomial functions (see Section~\ref{sec:intromoments}).
+To choose the batches, we form ellipsoid estimates for the supports of all $\phi_x$ via a moment method (Figure~\ref{fig:frog_moments_ellipsoid}) that involves applying $\Aop^T$ to a small number of polynomials (see Section~\ref{sec:intromoments}).
 We then use a greedy ellipsoid packing algorithm (Figure~\ref{fig:frog_batches}) to maximize the number of impulse responses per batch.
@@
 Then we interpolate translated and scaled versions of these impulse responses to approximate entries of the operator's integral kernel (Figure~\ref{fig:hmatrix_neighbors_frog}).
-Adding more batches yields impulse responses at more points, increasing the accuracy of the approximation at the cost of one operator application per batch (Figure~\ref{fig:frog_column_errors}).
+Adding more batches yields impulse responses at more points, increasing the approximation accuracy at the cost of one operator application per batch (Figure~\ref{fig:frog_column_errors}).
 
@@
 	\caption{Left: Impulse response moments: scaling factor ($V$), mean ($\mu$), and covariance ($\Sigma$).
-For each point $x \in \Omega$, the quantity $V(x)$ is the integral of $\phi_x$ over $\Omega$, $\mu(x)$ is the location that $\phi_x$ is centered at, and $\Sigma(x)$ is a matrix with eigenvectors and eigenvalues that characterize the width of $\phi_x$ (see Section~\ref{sec:intromoments}).
+For each point $x \in \Omega$, the quantity $V(x)$ is the integral of $\phi_x$ over $\Omega$, $\mu(x)$ is the location that $\phi_x$ is centered at, and $\Sigma(x)$ is a matrix with eigenvectors and eigenvalues that characterize the width of the support of $\phi_x$ about $\mu(x)$ (see Section~\ref{sec:intromoments}).
 Right: Ellipsoid support for an impulse response.
@@
 		\centering
-		\includegraphics[scale=0.35]{frog_column_errors_psf5.png}
+		\includegraphics[scale=0.32]{frog_column_errors_psf5.png}
 		%		\caption{PSF(5)}
@@
 		\centering
-		\includegraphics[scale=0.35]{frog_column_errors_psf10.png}
+		\includegraphics[scale=0.32]{frog_column_errors_psf10.png}
 		%		\caption{PSF(10)}
@@
 		\centering
-		\includegraphics[scale=0.35]{frog_column_errors_psf20.png}
+		\includegraphics[scale=0.32]{frog_column_errors_psf20.png}
 		%		\caption{PSF(20)}
@@
 		\centering
-		\includegraphics[scale=0.35]{frog_colorbar_edited1.pdf}
+		\includegraphics[scale=0.32]{frog_colorbar_edited1.pdf}
 		%			\label{fig:blur_colorbar}
@@
 
-While the PSF-based method in this paper may be used to approximate any operator that has a locally supported non-negative integral kernel, we are primarily motivated by approximation of high-rank Hessians in distributed parameter inverse problems governed by PDEs.
+While the method proposed in this paper may be used to approximate any operator that has a locally supported non-negative integral kernel, we are primarily motivated by approximation of high-rank Hessians in distributed parameter inverse problems governed by PDEs.
 In this section, we provide a brief background on this topic, and explain why existing Hessian approximation methods are not satisfactory.
@@
 		\begin{tabular}{cccc}
-			\includegraphics[scale=0.3, valign=c]{frog_ellipsoid_a=1.0.png} & 
-			\includegraphics[scale=0.3, valign=c]{frog_1d_a=1.0.png} & 
-			\includegraphics[scale=0.3, valign=c]{ricker_ellipsoid_a=0.0.png} & 
-			\includegraphics[scale=0.3, valign=c]{ricker_1d_a=0.0.png} \\
-			%
-			\includegraphics[scale=0.3, valign=c]{frog_ellipsoid_a=20.0.png} & 
-			\includegraphics[scale=0.3, valign=c]{frog_1d_a=20.0.png} & 
-			\includegraphics[scale=0.3, valign=c]{ricker_ellipsoid_a=0.23.png} & 
-			\includegraphics[scale=0.3, valign=c]{ricker_1d_a=0.23.png} \\
-			%
-			\includegraphics[scale=0.3, valign=c]{frog_ellipsoid_a=27.0.png} & 
-			\includegraphics[scale=0.3, valign=c]{frog_1d_a=27.0.png} & 
-			\includegraphics[scale=0.3, valign=c]{ricker_ellipsoid_a=0.249.png} & 
-			\includegraphics[scale=0.3, valign=c]{ricker_1d_a=0.249.png} 
+			\includegraphics[scale=0.25, valign=c]{frog_ellipsoid_a=1.0.png} & 
+			\includegraphics[scale=0.25, valign=c]{frog_1d_a=1.0.png} & 
+			\includegraphics[scale=0.25, valign=c]{ricker_ellipsoid_a=0.0.png} & 
+			\includegraphics[scale=0.25, valign=c]{ricker_1d_a=0.0.png} \\
+			%
+			\includegraphics[scale=0.25, valign=c]{frog_ellipsoid_a=20.0.png} & 
+			\includegraphics[scale=0.25, valign=c]{frog_1d_a=20.0.png} & 
+			\includegraphics[scale=0.25, valign=c]{ricker_ellipsoid_a=0.23.png} & 
+			\includegraphics[scale=0.25, valign=c]{ricker_1d_a=0.23.png} \\
+			%
+			\includegraphics[scale=0.25, valign=c]{frog_ellipsoid_a=27.0.png} & 
+			\includegraphics[scale=0.25, valign=c]{frog_1d_a=27.0.png} & 
+			\includegraphics[scale=0.25, valign=c]{ricker_ellipsoid_a=0.249.png} & 
+			\includegraphics[scale=0.25, valign=c]{ricker_1d_a=0.249.png} 
 		\end{tabular}
@@
 %The location of the negative numbers also affects the accuracy of the ellipsoid estimates substantially, but in a more complicated manner. 
-For the kernel in the left two columns of Figure \ref{fig:robustness_to_negatives}, negative numbers are interspersed with positive numbers, allowing us to include a large amount of negative numbers in the integral kernel before the proposed method fails.
+For the kernel in the left two columns of Figure \ref{fig:robustness_to_negatives}, negative numbers are interspersed with positive numbers, allowing us to include a large amount of negative numbers before the proposed method fails.
 For the kernel in the right two columns, the ellipsoid estimate fails with tiny amounts of negative numbers because the negative numbers are far away from the mean and not balanced by positive numbers at similar distances and angles.
@@
 Constructing the H-matrix from kernel entries requires a number of elementary operations that scales polylog linearly in $N$.
-Using the resulting H-matrix to perform further linear algebra operations also scales polylog linearly in $N$, though the details of these costs depend heavily on the type of H-matrix and operation being performed.
+Using the H-matrix to perform linear algebra operations also scales polylog linearly in $N$, though the details of these costs depend heavily on the type of H-matrix and operation being performed.
 
@@
 randomized HODLR (hierarchical off diagonal low rank)
-method~\cite{Martinsson16,Hartland_2023}, and GLR (global low rank)
-approximation using randomized SVD~\cite{HalkoMartinssonTropp11}. For
+method~\cite{Martinsson16,Hartland_2023} with 8 hierarchical levels, and GLR (global low rank)
+approximation using double-pass randomized SVD~\cite{HalkoMartinssonTropp11}. For
 these results we vary the quantity $L$ to scale the width of the

### 2024-02-20 paper_sisc_first_revision/localpsf.tex
minor edits

@@
 are non-negative numbers; this is not the same as positive semi-definiteness of $\Aop$.
-Such operators appear as, for example, Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{BorziSchulz11,Delosreyes15,HinzePinnauUlbrichUlbrich08},
+Such operators appear, for instance, as Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{BorziSchulz11,Delosreyes15,HinzePinnauUlbrichUlbrich08},
 %AntilEtAl18

### 2024-02-20 paper_sisc_first_revision/localpsf.tex
"Here," instead of "Here "

@@
 We present an efficient \emph{matrix-free} point spread function (PSF) method for approximating operators $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$ that have locally supported non-negative integral kernels.
-Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, and $L^2(\Omega)'$ is the space of real-valued continuous linear functionals on $L^2(\Omega)$.
+Here, $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, and $L^2(\Omega)'$ is the space of real-valued continuous linear functionals on $L^2(\Omega)$.
 By ``non-negative integral kernel,'' we mean that entries of $\Aop$'s integral kernel
@@
 	\caption{(\ref{fig:batches_intro}) One batch, $\eta_b$, of normalized impulse responses, $\impulseresponse_{x}$, that arise from applying $\Aop$ to a weighted sum of scattered point sources (see Section \ref{sec:get_impulse_response}).
-Here $\Aop$ is the ice sheet inverse problem data misfit Gauss-Newton Hessian described in Section~\ref{sec:numerical_results}.
+Here, $\Aop$ is the ice sheet inverse problem data misfit Gauss-Newton Hessian described in Section~\ref{sec:numerical_results}.
 	Black stars are point source locations.
@@
 The operator $\Aop$ is locally mean displacement invariant (Section \ref{sec:local_mean_displacement_invariance}) if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$.
-Here $\spatialmean(z)$ denotes the mean (center of mass) of $\phi_z$.
+Here, $\spatialmean(z)$ denotes the mean (center of mass) of $\phi_z$.
 	}
@@
 Linear functionals $\rho_h \in V_h'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^\nsamplepts_{\massmatrix^{-1}}$, with entries $\boldsymbol{\rho}_i = \rho_h(\febasis_i)$ for $i=1,\dots,\nsamplepts$.
-Here $\massmatrix \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries $\massmatrix_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) \mathrm{d}x$ for $i,j=1,\dots,\fedim$.
+Here, $\massmatrix \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries $\massmatrix_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) \mathrm{d}x$ for $i,j=1,\dots,\fedim$.
 The space $\mathbb{R}^\fedim_\massmatrix$ is $\mathbb{R}^\fedim$ with the inner product $(\mathbf{u},\mathbf{w})_\massmatrix := \mathbf{u}^T \massmatrix \mathbf{w}$, and $\mathbb{R}^\fedim_{\massmatrix^{-1}}$ is the analogous space with $\massmatrix^{-1}$ replacing $\massmatrix$.
@@
 \end{subequations}
-for $i,j=1,\dots, \gdim$. Here $u/w$ denotes pointwise division, $\left(u/w\right)(x) = u(x)/w(x)$, and $u\cdot w$ denotes pointwise multiplication, $(u \cdot w)(x) = u(x)w(x)$.
+for $i,j=1,\dots, \gdim$. Here, $u/w$ denotes pointwise division, $\left(u/w\right)(x) = u(x)/w(x)$, and $u\cdot w$ denotes pointwise multiplication, $(u \cdot w)(x) = u(x)w(x)$.
 
@@
 	\item We make sure that $\spatialvol(\candidatepoint_i)$ is not too small, by checking whether $\spatialvol(\candidatepoint_i) > \epsilon_V \spatialvol_\text{max}$.
-Here $\spatialvol_\text{max}$ is the largest value of $\spatialvol(\candidatepoint_j)$ over all points $q$ in the initial set of candidate points, and $\epsilon_V$ is a small threshold parameter (we use $\epsilon_V=10^{-5}$).
+Here, $\spatialvol_\text{max}$ is the largest value of $\spatialvol(\candidatepoint_j)$ over all points $q$ in the initial set of candidate points, and $\epsilon_V$ is a small threshold parameter (we use $\epsilon_V=10^{-5}$).
 	\item We make sure that all eigenvalues of $\spatialcov(\candidatepoint_i)$ are positive, and the aspect ratio of $E_{\candidatepoint_i}$ (square root of the ratio of the largest eigenvalue of $\spatialcov(\candidatepoint_i)$ to the smallest) is bounded by a constant $1/\epsilon_\Sigma$ (we use $1/\epsilon_\Sigma=20$).
@@
 
-Here we describe how to rapidly evaluate arbitrary entries of an approximation to the integral kernel by performing radial basis function interpolation of translated and scaled versions of nearby known impulse responses.
+Here, we describe how to rapidly evaluate arbitrary entries of an approximation to the integral kernel by performing radial basis function interpolation of translated and scaled versions of nearby known impulse responses.
 In Section~\ref{sec:h_matrix_construction_short} we use this procedure for rapidly evaluating kernel entries to construct the H-matrix approximation of $\mathbf{A}$.
@@
 for $i=1,\dots,\numnbr$, where $\{x_i\}_{i=1}^{\numnbr}$ are the $\numnbr$ nearest sample points to $x$, excluding points $x_i$ for which $z_i \notin \Omega$.
-Here $\numnbr$ is a small user-defined parameter, e.g., $\numnbr=10$.
+Here, $\numnbr$ is a small user-defined parameter, e.g., $\numnbr=10$.
 We find the $\numnbr$ nearest sample points to $x$ by querying a precomputed kd-tree~\cite{Bentley75} of all sample points.
@@
 is a Gaussian kernel radial basis function.
-Here $L:=\diam\left(\{x_i\}_{i=1}^{\numnbr}\right)$ is the diameter of the set of sample points used in the interpolation, and $\epsilon$ is a user-defined shape parameter that controls the width of the kernel function.
+Here, $L:=\diam\left(\{x_i\}_{i=1}^{\numnbr}\right)$ is the diameter of the set of sample points used in the interpolation, and $\epsilon$ is a user-defined shape parameter that controls the width of the kernel function.
 The vector of weights, $\rbfweight = (\rbfweight_1, \rbfweight_2, \dots, \rbfweight_{\numnbr})^T$, is found as the solution to the $\numnbr \times \numnbr$ linear system
@@
 $
-Here we use a diagonal lumped mass matrix, so these matrix-matrix multiplications are trivial.
+Here, we use a diagonal lumped mass matrix, so these matrix-matrix multiplications are trivial.
 If a non-diagonal mass matrix is used, one may form a H-matrix representation of the mass matrix, then perform the matrix-matrix multiplications in \eqref{eq:boldA} using H-matrix methods.
@@
 Although $\mathbf{\Aker}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Aker}_H$ only requires evaluation of $O(\hrank \fedim \log \fedim)$ kernel entries $\mathbf{\Aker}_{ij} = \widetilde{\Aker}(\fepoint_i, \fepoint_j)$ (see \cite{BebendorfRjasanow03}), and these entries are computed via the radial basis function interpolation method described in Section~\ref{sec:approximate_kernel_entries}.
-Here $\hrank$ is the rank of the highest rank block in the H-matrix.
+Here, $\hrank$ is the rank of the highest rank block in the H-matrix.
 We emphasize that the dense matrix $\mathbf{\Aker}$ is never formed.
@@
 wherein we choose the initial guess as the discretization of the constant
-function $q_0$. Here $\mathbf{g}$, $\mathbf{H}$ and $\mathbf{H}_\text{gn}$ are
+function $q_0$. Here, $\mathbf{g}$, $\mathbf{H}$ and $\mathbf{H}_\text{gn}$ are
 the discretized gradient, Hessian, and Gauss-Newton Hessian of the
@@
 
-Here we define a PDE-free spatially varying blur problem, in which the impulse response, $\phi_x$,
+Here, we define a PDE-free spatially varying blur problem, in which the impulse response, $\phi_x$,
 is a bumpy blob that is centered
@@
 $\dot{\strain}(v)= \frac{1}{2}\left(
-\nabla\velocity+\nabla\velocity^{\top} \right)$ is the strain rate tensor~\cite{IsaacEtAl15}. Here $\eta$ is the viscosity and $I$ is the identity operator. 
+\nabla\velocity+\nabla\velocity^{\top} \right)$ is the strain rate tensor~\cite{IsaacEtAl15}. Here, $\eta$ is the viscosity and $I$ is the identity operator. 
 %Glen's exponent has been chosen to be one, which provides for a linear Stokes model. 
@@
 	Right: Generalized eigenvalues for generalized eigenvalue problem $\mathbf{H} \mathbf{u}_{k}=\lambda_{k} \mathbf{\preconditioner} \mathbf{u}_{k}$.
-Here $\mathbf{H}$ is the Hessian and the matrices $\mathbf{\preconditioner}$ are the same Hessian approximations used in the left sub-figure, with NONE corresponding to the identity matrix.}
+Here, $\mathbf{H}$ is the Hessian and the matrices $\mathbf{\preconditioner}$ are the same Hessian approximations used in the left sub-figure, with NONE corresponding to the identity matrix.}
 	\label{fig:krylov_convergence}
@@
 
-Here we consider a time-dependent advection-diffusion equation in
+Here, we consider a time-dependent advection-diffusion equation in
 which we seek to infer the unknown spatially varying initial condition, $\ipar$, from noisy observation of the full state  
@@
 Right: Generalized eigenvalues for generalized eigenvalue problem $\mathbf{H} \mathbf{u}_{k}=\lambda_{k} \mathbf{\preconditioner} \mathbf{u}_{k}$.
-Here $\mathbf{H}$ is the Hessian and the preconditioner, $\mathbf{\preconditioner}$, is the PSF approximation for 1, 5, or 25 batches (PSF (1), PSF (5), and PSF (25), respectively), the regularization Hessian (REG), or the identity matrix (NONE).}
+Here, $\mathbf{H}$ is the Hessian and the preconditioner, $\mathbf{\preconditioner}$, is the PSF approximation for 1, 5, or 25 batches (PSF (1), PSF (5), and PSF (25), respectively), the regularization Hessian (REG), or the identity matrix (NONE).}
 	\label{fig:adv_krylov_convergence}

### 2024-02-20 numerical_examples/blur/blur_length_scale_experiments.py

@@ -17,9 +17,11 @@ nx = 63
 a = 1.0 # <-- How bumpy? Use 1.0, which is maximum bumpiness without negative numbers
 all_length_scalings = [1.0 / (t**2) for t in [1.0, 2.0, 3.0]]
 all_error_levels = [0.2, 0.1, 0.05]
+shape_parameter = 0.5
 # length_scaling = 1./(3.0**2) #1.0
 
-do_PSF = False
+do_PSF = True
+do_others = False
 
 np.savetxt('all_length_scalings.txt', all_length_scalings)
 np.savetxt('all_error_levels.txt', all_error_levels)
@@ -80,35 +82,36 @@ for length_scaling in all_length_scalings:
 
     #### Global low rank
 
-    _, ss, _ = np.linalg.svd(Ker)
-    all_ss.append(ss)
-    norm_ker_fro = np.linalg.norm(Ker)
+    if do_others:
+        _, ss, _ = np.linalg.svd(Ker)
+        all_ss.append(ss)
+        norm_ker_fro = np.linalg.norm(Ker)
 
-    all_rsvd_applies = []
-    for rtol in all_error_levels:
-        rsvd_num_applies = compute_rsvd_applies(Ker, rtol)
-        all_rsvd_applies.append(rsvd_num_applies)
-        print('rtol=', rtol, ', rsvd_num_applies=', rsvd_num_applies)
-    all_all_rsvd_applies.append(all_rsvd_applies)
+        all_rsvd_applies = []
+        for rtol in all_error_levels:
+            rsvd_num_applies = compute_rsvd_applies(Ker, rtol)
+            all_rsvd_applies.append(rsvd_num_applies)
+            print('rtol=', rtol, ', rsvd_num_applies=', rsvd_num_applies)
+        all_all_rsvd_applies.append(all_rsvd_applies)
 
-    all_glr_ranks_fro = []
-    for rtol in all_error_levels:
-        glr_num_fro = np.sum((norm_ker_fro**2 - np.cumsum(ss**2)) > rtol**2 * norm_ker_fro**2)
-        all_glr_ranks_fro.append(glr_num_fro)
-        print('rtol=', rtol, ', glr_num_fro=', glr_num_fro)
-    all_all_glr_ranks_fro.append(all_glr_ranks_fro)
+        all_glr_ranks_fro = []
+        for rtol in all_error_levels:
[…]

### 2024-02-20 paper_sisc_first_revision/localpsf.tex
re-ran length scale experiments with shape_parameter=0.5 instead of 3.0. Updated table. Barely affected anything.

@@
 On a $64 \times 64$ grid with depth $4$, hierarchical rank $10$, and oversampling parameter $c=5$, this works out to $4 \cdot 64 \cdot (10 + 5) = 3840$ operator applies.
-In Section \ref{sec:frog} see numerically that the randomized hierarchical off-diagonal low rank (HODLR) method in \cite{Martinsson16} requires hundreds to thousands of matrix-vector products to construct approximations of the integral kernel for a blur problem example with modest (e.g., $10\%$) relative error.
+In Section \ref{sec:frog}, we see numerically that the randomized hierarchical off-diagonal low rank (HODLR) method in \cite{Martinsson16} requires hundreds to thousands of matrix-vector products to construct approximations of the integral kernel for a blur problem example with modest (e.g., $10\%$) relative error.
 %while the PSF method in this paper can achieve the same level of error in one to two dozen operator applies.
@@
 								& 20\%    		& 11  				& 592				& 354				\\%& 111  \\ 
-				$L=1$ 		& 10\% 			& 15  				& 772				& 520				\\%& 178  \\ 
-				& 5\% 			& 23   				& 924				& 674				\\%& 244 \\ 
+				$L=1$ 			& 10\% 			& 16  				& 772				& 520				\\%& 178  \\ 
+								& 5\% 			& 22   				& 924				& 674				\\%& 244 \\ 
 				\hline
 								& 20\%    		& 8  				& 852 				& 1316				\\%& 417 \\ 
-				$L=1/2$ 	& 10\%    		& 10  				& 1144				& 1916				\\%& 665 \\ 
+				$L=1/2$ 		& 10\%    		& 9  				& 1144				& 1916				\\%& 665 \\ 
 								& 5\% 	  		& 12   				& 1404				& 2456				\\%& 903 \\ 

### 2024-02-20 paper_sisc_first_revision/localpsf.tex
PSF-based armageddon

@@
 	We present an efficient matrix-free point spread function (PSF) method for approximating operators that have locally supported non-negative integral kernels.
-The proposed method computes impulse responses of the operator at scattered points, and interpolates these impulse responses to approximate entries of the integral kernel.
+The PSF-based method computes impulse responses of the operator at scattered points, and interpolates these impulse responses to approximate entries of the integral kernel.
 To compute impulse responses efficiently, we apply the operator to Dirac combs associated with batches of point sources, which are chosen by solving an ellipsoid packing problem.
@@
 This end-to-end procedure is illustrated on a blur problem.
-We demonstrate the proposed method's effectiveness by using it to build preconditioners for the Hessian operator arising in two inverse problems governed by partial differential equations (PDEs): inversion for the basal friction coefficient in an ice sheet flow problem and for the initial condition in an advective-diffusive transport problem.
+We demonstrate the PSF-based method's effectiveness by using it to build preconditioners for the Hessian operator arising in two inverse problems governed by partial differential equations (PDEs): inversion for the basal friction coefficient in an ice sheet flow problem and for the initial condition in an advective-diffusive transport problem.
 While for many ill-posed inverse problems the Hessian of the data misfit term exhibits a low rank structure, and hence a low rank approximation is suitable, for many problems of practical interest the numerical rank of the Hessian is still large.
-The Hessian impulse responses on the other hand typically become more local as the numerical rank increases, which benefits the proposed method.
+The Hessian impulse responses on the other hand typically become more local as the numerical rank increases, which benefits the PSF-based method.
 Numerical results reveal that the preconditioner clusters the spectrum of the preconditioned Hessian near one, yielding roughly $5\times$--$10\times$ reductions in the required number of PDE solves, as compared to classical regularization-based preconditioning and no preconditioning.
 We also present a comprehensive numerical study for the influence of various parameters (that control the shape of the impulse responses and the rank of the Hessian) on the effectiveness of the advection-diffusion Hessian approximation.
-The results show that the proposed method is able to form good approximations of high-rank Hessians using only a small number of operator applications.
+The results show that the PSF-based method is able to form good approximations of high-rank Hessians using only a small number of operator applications.
 \end{abstract}
@@
 
-The idea of the proposed method is to use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applications.
+The idea of the proposed method, which we refer to throughout the paper as the ``PSF-based method'', is to use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applications.
 The impulse response, $\phi_x$, associated with a point, $x$, is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
@@
 
-The proposed operator approximation method may be categorized as a PSF method that is loosely based on ``product convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
+The PSF-based method we propose is a PSF method that is loosely based on ``product convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
 PC and PSF methods have a long history dating back several decades.
@@
 For background on PC and PSF methods, we recommend the following papers: \cite{DenisEtAl15,EscandeWeiss17,GentileCourbinMeylan13}.
-The proposed method improves upon existing PC methods in the following ways: (1) While PC approximations are based on an assumption of local translation invariance, the method we propose is based on a more general assumption we call ``local mean displacement invariance'' (Section~\ref{sec:local_mean_displacement_invariance} and Figure~\ref{fig:mean_displacement_invariance}), which improves the interpolation of the impulse responses.
+The proposed PSF-based method improves upon existing PC and PSF methods in the following ways: (1) While PC and PSF approximations are typically based on an assumption of local translation invariance, the method we propose is based on a more general assumption we call ``local mean displacement invariance'' (Section~\ref{sec:local_mean_displacement_invariance} and Figure~\ref{fig:mean_displacement_invariance}), which improves the interpolation of the impulse responses.
 (2) In our previous work~\cite{AlgerEtAl19}, we chose point sources in an adaptive grid via a sequential procedure; the refinements to the adaptive grid were chosen to maximally reduce the error at each step.
@@
 The moment method that we use estimates the support of $\phi_x$ more accurately than random noise probing in resolution analysis, at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.
-(3) The method we propose never evaluates computed impulse responses outside of their domain of definition.
-This eliminates ``boundary-artifact'' errors (see \cite[Section 1.1]{AlgerEtAl19}) that plague conventional PC methods.
+(3) The PSF-based method we propose never evaluates computed impulse responses outside of their domain of definition.
+This eliminates ``boundary-artifact'' errors (see \cite[Section 1.1]{AlgerEtAl19}) that plague conventional PC and PSF methods.
 
@@
 
-While the method proposed in this paper may be used to approximate any operator that has a locally supported non-negative integral kernel, we are primarily motivated by approximation of high-rank Hessians in distributed parameter inverse problems governed by PDEs.
+While the PSF-based method proposed in this paper may be used to approximate any operator that has a locally supported non-negative integral kernel, we are primarily motivated by approximation of high-rank Hessians in distributed parameter inverse problems governed by PDEs.
 In this section, we provide a brief background on this topic, and explain why existing Hessian approximation methods are not satisfactory.
@@
 In this case more specialized techniques have been developed using, eg., pseudodifferential operator theory \cite{DemanetEtAl12,Symes08}, and sparsity in the wavelet domain \cite{HerrmannMoghaddamStolk08}.
-However, many data misfit Hessians of practical interest do satisfy the local non-negative integral kernel properties (either exactly or approximately), and the proposed method is targeted at approximating these Hessians.
+However, many data misfit Hessians of practical interest do satisfy the local non-negative integral kernel properties (either exactly or approximately), and the PSF-based method we propose is targeted at approximating these Hessians.
 
@@
 The integral kernel, $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$, exists but is not easily accessible.
-In this section we describe how to extend the domain of $\Aop$ to distributions, which allows us to define impulse responses (Section~\ref{sec:impulse_response_2}), we then state the conditions on $\Aop$ that the proposed method requires (Section~\ref{sec:conditions_2}), and detail finite element discretization (Section~\ref{sec:finite_element_kernel}).
+In this section we describe how to extend the domain of $\Aop$ to distributions, which allows us to define impulse responses (Section~\ref{sec:impulse_response_2}), we then state the conditions on $\Aop$ that the PSF-based method requires (Section~\ref{sec:conditions_2}), and detail finite element discretization (Section~\ref{sec:finite_element_kernel}).
 
@@
 	\item The integral kernel is non-negative\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness.
-The operator $\Aop$ need not be positive semi-definite to use the proposed method, and positive semi-definite operators need not have a non-negative integral kernel.} in the sense that $$\Aker(y,x) \ge 0 \quad \text{for all} \quad (y,x) \in \Omega \times \Omega.$$
+The operator $\Aop$ need not be positive semi-definite to use the PSF-based method, and positive semi-definite operators need not have a non-negative integral kernel.} in the sense that $$\Aker(y,x) \ge 0 \quad \text{for all} \quad (y,x) \in \Omega \times \Omega.$$
 \end{enumerate}
-The proposed method may still perform well if these conditions are relaxed slightly.
+The PSF-based method may still perform well if these conditions are relaxed slightly.
 
@@
 It is acceptable if the support of $\impulseresponse_x$ is not perfectly contained in a localized region (violating Assumption 2), so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
-In principle, the proposed method can be applied even if the impulse responses are not localized.
+In principle, the PSF-based method can be applied even if the impulse responses are not localized.
 However, in this case only a small number of impulse responses can be computed per batch, which means more batches, and hence more operator applies, are needed to form an accurate approximation.
@@
 
-If there are negative numbers in the integral kernel (violating Assumption 3), the ellipsoid estimation procedure may incur errors or fail, leading to poor performance or failure of the proposed method.
+If there are negative numbers in the integral kernel (violating Assumption 3), the ellipsoid estimation procedure may incur errors or fail, leading to poor performance or failure of the PSF-based method.
 In Figure \ref{fig:robustness_to_negatives} we investigate the robustness of the ellipsoid support estimation procedure to violations of Assumption 3.
@@
 This is because moment formulas (Equations \ref{eq:define_mean} and \ref{eq:define_cov}) assign more weight to entries in the kernel that are further from the center.
-	\item Negative numbers in the integral kernel affect the proposed method more if they are isolated, and less if they are balanced by nearby positive numbers.
+	\item Negative numbers in the integral kernel affect the ellipsoid estimation method more if they are isolated, and less if they are balanced by nearby positive numbers.
 	\item As kernels become more negative, the ellipsoid estimation performs well up to a certain threshold that depends on the spatial distribution of negative and positive entries.
@@
 %The location of the negative numbers also affects the accuracy of the ellipsoid estimates substantially, but in a more complicated manner. 
-For the kernel in the left two columns of Figure \ref{fig:robustness_to_negatives}, negative numbers are interspersed with positive numbers, allowing us to include a large amount of negative numbers before the proposed method fails.
+For the kernel in the left two columns of Figure \ref{fig:robustness_to_negatives}, negative numbers are interspersed with positive numbers, allowing us to include a large amount of negative numbers before the ellipsoid estimation fails.
 For the kernel in the right two columns, the ellipsoid estimate fails with tiny amounts of negative numbers because the negative numbers are far away from the mean and not balanced by positive numbers at similar distances and angles.
-In the bottom two rows, we see the aforementioned threshold effect, in which the proposed method rapidly transitions from performing reasonably well to performing poorly with only a small change to the integral kernel.
+In the bottom two rows, we see the aforementioned threshold effect, in which the ellipsoid estimation method rapidly transitions from performing reasonably well to performing poorly with only a small change to the integral kernel.
 
@@
 In this paper we discretize the functions that $\mathcal{A}$ and $\mathcal{A}^T$ are applied to using continuous finite elements satisfying the Kronecker property (defined below).
-With minor modifications, the proposed method could be used with more general finite element methods, or other discretization schemes such as finite differences or finite volumes.
+With minor modifications, the PSF-based method could be used with more general finite element methods, or other discretization schemes such as finite differences or finite volumes.
 These restrictions on discretization only apply to functions $u$ that $\mathcal{A}$ and $\mathcal{A}^T$ are applied to.
@@
 under the isomorphism discussed above.
-The proposed method is agnostic to the computational procedure for approximating $\Aop$ with $\mathbf{A}$.
+The PSF-based method is agnostic to the computational procedure for approximating $\Aop$ with $\mathbf{A}$.
 What is important is that we do not have direct access to matrix entries $\mathbf{A}_{ij}$.
 Rather, we have a computational procedure that allows us to compute matrix-vector products $\mathbf{u}\mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{w}\mapsto \mathbf{A}^T\mathbf{w}$, and computing these matrix-vector products is costly.
-The proposed method mitigates this cost by performing as few matrix-vector products as possible.
+The PSF-based method mitigates this cost by performing as few matrix-vector products as possible.
 Of course, matrix entries can be computed via matrix-vector products as $\mathbf{A}_{ij} = \left(\mathbf{A}\mathbf{e}_j\right)_i$,
@@
 Classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Aker}_{ij}$, but these matrix entries are not easily accessible.
-The bulk of the proposed method is therefore dedicated to forming approximations of these matrix entries that can be evaluated rapidly.
+The bulk of the PSF-based method is therefore dedicated to forming approximations of these matrix entries that can be evaluated rapidly.
 
@@
 
-In this section we present two key innovations that the proposed method is based on.
+In this section we present two key innovations that the PSF-based method is based on.
 First, we define moments of the impulse responses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section~\ref{sec:intromoments}).
@@
 However, this direct approach is not feasible because our algorithm for picking sample points (Section \ref{sec:sample_point_selection} and Figure \ref{fig:frog_batches}) needs to know $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ before we compute $\phi_x$.
-Computing $\phi_x$ in order to determine $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ would be extremely computationally expensive, and defeat the purpose of the proposed method, which is to reduce the computational cost by computing impulse responses in batches.
+Computing $\phi_x$ in order to determine $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ would be extremely computationally expensive, and defeat the purpose of the PSF-based method, which is to reduce the computational cost by computing impulse responses in batches.
 Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ indirectly, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions (e.g., 6 total operator applications in two spatial dimensions and 10 in three spatial dimensions).
@@
 The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$.
-However, in Section~\ref{sec:sample_point_selection} we will see that the cost of the proposed method depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids is better), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
+However, in Section~\ref{sec:sample_point_selection} we will see that the cost of the PSF-based method depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids is better), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$.
 In practice, we find that $\tau=3.0$ yields a reasonable balance between these competing interests, and use $\tau=3.0$ in all numerical results, except for Figure \ref{fig:PSF_tau_convergence}, where we study the effects of varying $\tau$.
@@
 These are four different ways to approximate an impulse response by a nearby impulse response, with each successive approximation building upon the previous ones.
-The proposed method uses~\eqref{eq:translate4}, which is the most sophisticated.
+The PSF-based method uses~\eqref{eq:translate4}, which is the most sophisticated.
 Approximation~\eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close.
@@
 %This is illustrated in Figure \ref{fig:RBF_width} for a simple one dimensional interpolation problem. 
-For our numerical results involving Hessians in inverse problems governed by PDEs (Sections~\ref{sec:ice}~and~\ref{sec:adv}), high accuracy is not required because the proposed method is used to construct a preconditioner.
+For our numerical results involving Hessians in inverse problems governed by PDEs (Sections~\ref{sec:ice}~and~\ref{sec:adv}), high accuracy is not required because the PSF-based method is used to construct a preconditioner.
 Hence, for these Hessian approximations we use a conservative choice of $\epsilon=3.0$ to ensure robustness.
-For our numerical results for the blur problem example (Section~\ref{sec:frog}), we use a smaller value of $\epsilon=0.5$ so that the RBF interpolation accuracy is not a limiting factor as we study convergence of the proposed method.
+For our numerical results for the blur problem example (Section~\ref{sec:frog}), we use a smaller value of $\epsilon=0.5$ so that the RBF interpolation accuracy is not a limiting factor as we study convergence of the PSF-based method.
 
@@
 $\mathcal{H} = \Aop + \mathcal{R},$
-where $\Aop$ is a symmetric positive semi-definite operator that we approximate with the proposed method to form an H-matrix $\mathbf{A}_H$, and $\mathcal{R}$ is a symmetric positive definite operator that may be easily converted to an H-matrix $\mathbf{R}_H$ without using the proposed method.
+where $\Aop$ is a symmetric positive semi-definite operator that we approximate with the PSF-based method to form an H-matrix $\mathbf{A}_H$, and $\mathcal{R}$ is a symmetric positive definite operator that may be easily converted to an H-matrix $\mathbf{R}_H$ without using the PSF-based method.
 For example, in inverse problems $\mathcal{H}$ is the Hessian, $\Aop$ is the data misfit term in the Hessian which is dense and available only matrix-free, and $\mathcal{R}$ is the regularization term, which is typically an elliptic differential operator that becomes a sparse matrix after discretization.
@@
 
-The computational cost of the proposed method may be divided into the costs to perform the following tasks: (1) Computing impulse response moments and batches (Lines 1 and 4 in Algorithm \ref{alg:construct_Atilde}); (2) Building the H-matrix (Lines 5 and 6 in Algorithm \ref{alg:construct_Atilde}); (3) Performing linear algebra operations with the H-matrix.
+The computational cost of the PSF-based method may be divided into the costs to perform the following tasks: (1) Computing impulse response moments and batches (Lines 1 and 4 in Algorithm \ref{alg:construct_Atilde}); (2) Building the H-matrix (Lines 5 and 6 in Algorithm \ref{alg:construct_Atilde}); (3) Performing linear algebra operations with the H-matrix.
 This may optionally include the symmetric positive semi-definite modifications described in Section \ref{sec:flip_eigs}.
@@
 
-We use the proposed method to approximate the integral kernel in a blur problem (Section~\ref{sec:frog}), and the Newton (or Gauss-Newton) Hessians
+We use the PSF-based method to approximate the integral kernel in a blur problem (Section~\ref{sec:frog}), and the Newton (or Gauss-Newton) Hessians
 in inverse problems governed by PDEs which model steady state ice
@@
 Sections~\ref{sec:intro} and~\ref{sec:prelims} to visually illustrate
-various stages and aspects of the proposed method, and robustness (or lack thereof) of the proposed method to violations of the non-negative kernel assumption (Section~\ref{sec:conditions_2}).
-In this section, the blur problem is used to compare the proposed method to hierarchical off-diagonal low rank (HODLR) and global low rank (GLR) methods, to study convergence of the proposed method, and to investigate the effect of the ellipsoid size parameter $\tau$.
+various stages and aspects of the PSF-based method, and robustness (or lack thereof) of the PSF-based method to violations of the non-negative kernel assumption (Section~\ref{sec:conditions_2}).
+In this section, the blur problem is used to compare the PSF-based method to hierarchical off-diagonal low rank (HODLR) and global low rank (GLR) methods, to study convergence of the PSF-based method, and to investigate the effect of the ellipsoid size parameter $\tau$.
 %To study the
@@
 > 1$, the integral kernel contains negative values, which allows us to
-study the robustness of the proposed method to violations of Assumption 3
+study the robustness of the PSF-based method to violations of Assumption 3
 (Section~\ref{sec:conditions_2}).
@@
 		approximate the blur kernel from Equation
-		\eqref{eq:frog_kernel} using the proposed PSF-based method, the randomized
+		\eqref{eq:frog_kernel} using the PSF-based method, the randomized
 		HODLR (hierarchical off diagonal low rank) method, and GLR
@@
 computational cost to approximate the blur kernel from
-Equation~\eqref{eq:frog_kernel} using the proposed PSF-based method, the
+Equation~\eqref{eq:frog_kernel} using the PSF-based method, the
 randomized HODLR (hierarchical off diagonal low rank)
@@
 
-In Figure \ref{fig:PSF_tau_convergence} we show the convergence of the proposed method on the blur kernel as a function of the total number of impulse responses (left), and the number of batches (right).
+In Figure \ref{fig:PSF_tau_convergence} we show the convergence of the PSF-based method on the blur kernel as a function of the total number of impulse responses (left), and the number of batches (right).
 We show convergence for several ellipsoid size parameters $\tau$, ranging from $2.0$--$4.0$.
@@
 % or $\text{constant} \times \# \text{batches}^{-1}$, 
-suggesting linear convergence. The linear convergence stalls at a limit which depends on $\tau$. Increasing $\tau$ lowers this limit, allowing the proposed method to achieve higher accuracy. 
+suggesting linear convergence. The linear convergence stalls at a limit which depends on $\tau$. Increasing $\tau$ lowers this limit, allowing the PSF-based method to achieve higher accuracy. 
 In Figure \ref{fig:PSF_tau_convergence} (right), the results show that before this limit is reached, the convergence is faster for smaller $\tau$ in terms of the number of batches.
 This is expected because smaller $\tau$ results in more impulse responses per batch.
-Larger $\tau$ causes the proposed method converge more slowly than smaller $\tau$, but with larger $\tau$ the proposed method stalls at a lower level of error than it does with smaller $\tau$ (see, e.g., $\tau=4.0$ vs.
+Larger $\tau$ causes the PSF-based method converge more slowly than smaller $\tau$, but with larger $\tau$ the PSF-based method stalls at a lower level of error than it does with smaller $\tau$ (see, e.g., $\tau=4.0$ vs.
 $\tau=2.5$).
@@
 Note that $\Gamma_b$ is a two-dimensional surface embedded in three dimensions due to the mountain topography.
-The proposed PSF-based method involves translating impulse responses.
+The PSF-based method involves translating impulse responses.
 Hence it requires either a flat domain, or a notion of local parallel transport.
@@
 %	The Newton iteration is terminated when $||\mathbf{g}|| < 10^{-6} ||\mathbf{g}_0||$, where $\mathbf{g}$ is the gradient at the current iteration, and $\mathbf{g}_0$ is the gradient at the initial guess.
-	Preconditioners shown are the proposed PSF-based method with five batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
+	Preconditioners shown are the PSF-based method with five batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
 Columns \#CG show the number of PCG iterations used to solve the Newton system for $\mathbf{\widehat{\basalfriction}}$.
@@
 No preconditioning is used in the iterations before the PSF (5) preconditioner is built.
-We compare the proposed method with the most commonly used existing preconditioners: no preconditioning (NONE), and preconditioning by the regularization term in the Hessian (REG).
+We compare the PSF-based method with the most commonly used existing preconditioners: no preconditioning (NONE), and preconditioning by the regularization term in the Hessian (REG).
 The results show that using PSF (5) reduces the total number of Stokes PDE solves to 70, as compared to 908 for regularization preconditioning and 308 for no preconditioning, a reduction in cost of roughly $5\times$--$10\times$.
@@
 We presented an efficient matrix-free PSF-based method for approximating operators with locally supported non-negative integral kernels.
-The proposed method requires access to the operator only via application of the operator to a small number of vectors.
-The idea of the proposed method is to compute batches of impulse responses by applying the operator to Dirac combs of scattered point sources, then interpolate these impulse responses to approximate entries of the operator's integral kernel.
+The PSF-based method requires access to the operator only via application of the operator to a small number of vectors.
+The idea of the PSF-based method is to compute batches of impulse responses by applying the operator to Dirac combs of scattered point sources, then interpolate these impulse responses to approximate entries of the operator's integral kernel.
 The interpolation is based on a new principle we call ``local mean displacement invariance,'' which generalizes classical local translation invariance.
@@
 Point source locations for the impulse response batches are chosen using a greedy ellipsoid packing procedure, in which we choose as many impulse responses per batch as possible, while ensuring that the corresponding ellipsoids do not overlap.
-We applied the proposed method to approximate the Gauss-Newton Hessians in an ice sheet flow inverse problem governed by a linear Stokes PDE, and an advective-diffusive transport inverse problem governed by an advection-diffusion PDE.
-We saw that preconditioners based on the proposed approximation cluster the eigenvalues of the preconditioned Hessian near one, and allow us to solve the inverse problems using roughly $5\times$--$10\times$ fewer PDE solves.
+We applied the PSF-based method to approximate the Gauss-Newton Hessians in an ice sheet flow inverse problem governed by a linear Stokes PDE, and an advective-diffusive transport inverse problem governed by an advection-diffusion PDE.
+We saw that preconditioners based on the PSF-based approximation cluster the eigenvalues of the preconditioned Hessian near one, and allow us to solve the inverse problems using roughly $5\times$--$10\times$ fewer PDE solves.
 For larger domains with more observations, the rank of the data misfit Hessian will increase, while the locality of impulse responses will remain the same.
 Hence, we expect the speedup will be even greater for such problems.
-Although the proposed method is not applicable to all Hessians, it is applicable to many Hessians of practical interest.
-For these Hessians, the proposed method offers a \emph{data scalable} alternative to conventional low rank approximation, due to the ability to form high rank approximations of an operator using a small number of operator applications, and thus PDE solves.
+Although the PSF-based method is not applicable to all Hessians, it is applicable to many Hessians of practical interest.
+For these Hessians, the PSF-based method offers a \emph{data scalable} alternative to conventional low rank approximation, due to the ability to form high rank approximations of an operator using a small number of operator applications, and thus PDE solves.
 

### 2024-02-20 paper_sisc_first_revision/localpsf.tex
incorporated Omar's comments

@@
 that as we increase the rank and decrease the error tolerance, the
-performance of the the HODLR and GLR methods deteriorates.
+performance of the HODLR and GLR methods deteriorates.
 
@@
 % or $\text{constant} \times \# \text{batches}^{-1}$, 
-suggesting linear convergence. The linear convergence stalls at a limit which depends on $\tau$. Increasing $\tau$ lowers this limit, allowing the PSF-based method to achieve higher accuracy. 
+suggesting linear convergence. The linear convergence stalls at a limit that depends on $\tau$. Increasing $\tau$ lowers this limit, allowing the PSF-based method to achieve higher accuracy. 
 In Figure \ref{fig:PSF_tau_convergence} (right), the results show that before this limit is reached, the convergence is faster for smaller $\tau$ in terms of the number of batches.
 This is expected because smaller $\tau$ results in more impulse responses per batch.
-Larger $\tau$ causes the PSF-based method converge more slowly than smaller $\tau$, but with larger $\tau$ the PSF-based method stalls at a lower level of error than it does with smaller $\tau$ (see, e.g., $\tau=4.0$ vs.
+Larger $\tau$ causes the PSF-based method to converge more slowly than smaller $\tau$, but with larger $\tau$ the PSF-based method stalls at a lower level of error than it does with smaller $\tau$ (see, e.g., $\tau=4.0$ vs.
 $\tau=2.5$).

### 2024-02-20 paper_sisc_first_revision/localpsf.tex
L -> r_0 in RBF to resolve notation conflict

@@
 This is because moment formulas (Equations \ref{eq:define_mean} and \ref{eq:define_cov}) assign more weight to entries in the kernel that are further from the center.
-	\item Negative numbers in the integral kernel affect the ellipsoid estimation method more if they are isolated, and less if they are balanced by nearby positive numbers.
+	\item Negative numbers affect the ellipsoid estimation method more if they are isolated, and less if they are balanced by nearby positive numbers.
 	\item As kernels become more negative, the ellipsoid estimation performs well up to a certain threshold that depends on the spatial distribution of negative and positive entries.
@@
 where $\spatialmean^i(x)$ and $y^i$ denote the $i^\text{th}$ components of the vectors $\spatialmean(x)$ and $y$, respectively, and $\spatialcov^{ij}(x)$ denotes the $(i,j)$ entry of the matrix $\spatialcov(x)$.
-The vector $\spatialmean(x)\in \mathbb{R}^\gdim$ and the matrix $\spatialcov(x) \in \mathbb{R}^{\gdim \times \gdim}$ are the mean and covariance of the normalized version of $\impulseresponse_x$, respectively.
+The quantities $\spatialmean(x)\in \mathbb{R}^\gdim$ and $\spatialcov(x) \in \mathbb{R}^{\gdim \times \gdim}$ are the mean and covariance of the normalized version of $\impulseresponse_x$, respectively.
 
@@
 where $\rbfweight_i$ are weights, and 
-$\rbf(r):= \exp\left(-\frac{1}{2}\left(\epsilon\frac{r}{L}\right)^2\right)$
+$\rbf(r):= \exp\left(-\frac{1}{2}\left(C_\text{RBF}\frac{r}{r_0}\right)^2\right)$
 is a Gaussian kernel radial basis function.
-Here, $L:=\diam\left(\{x_i\}_{i=1}^{\numnbr}\right)$ is the diameter of the set of sample points used in the interpolation, and $\epsilon$ is a user-defined shape parameter that controls the width of the kernel function.
+Here, $r_0:=\diam\left(\{x_i\}_{i=1}^{\numnbr}\right)$ is the diameter of the set of sample points used in the interpolation, and $C_\text{RBF}$ is a user-defined shape parameter that controls the width of the kernel function.
 The vector of weights, $\rbfweight = (\rbfweight_1, \rbfweight_2, \dots, \rbfweight_{\numnbr})^T$, is found as the solution to the $\numnbr \times \numnbr$ linear system
@@
 
-The shape parameter, $\epsilon$, mediates a tradeoff between accuracy and stability.
-Small $\epsilon$ is required for RBF interpolation with Gaussian kernels to achieve high accuracy, but small $\epsilon$ also makes RBF interpolation less robust to errors or nonsmoothness in the function being interpolated.
+The shape parameter, $C_\text{RBF}$, mediates a tradeoff between accuracy and stability.
+Small $C_\text{RBF}$ is required for RBF interpolation with Gaussian kernels to achieve high accuracy, but small $C_\text{RBF}$ also makes RBF interpolation less robust to errors or nonsmoothness in the function being interpolated.
 %This is illustrated in Figure \ref{fig:RBF_width} for a simple one dimensional interpolation problem. 
 For our numerical results involving Hessians in inverse problems governed by PDEs (Sections~\ref{sec:ice}~and~\ref{sec:adv}), high accuracy is not required because the PSF-based method is used to construct a preconditioner.
-Hence, for these Hessian approximations we use a conservative choice of $\epsilon=3.0$ to ensure robustness.
-For our numerical results for the blur problem example (Section~\ref{sec:frog}), we use a smaller value of $\epsilon=0.5$ so that the RBF interpolation accuracy is not a limiting factor as we study convergence of the PSF-based method.
+Hence, for these Hessian approximations we use a conservative choice of $C_\text{RBF}=3.0$ to ensure robustness.
+For our numerical results for the blur problem example (Section~\ref{sec:frog}), we use a smaller value of $C_\text{RBF}=0.5$ so that the RBF interpolation accuracy is not a limiting factor as we study convergence of the PSF-based method.
 
@@
 In Table~\ref{tab:frog_psf_vs_hodlr_vs_glr} we compare the
-computational cost to approximate the blur kernel from
+cost to approximate the blur kernel from
 Equation~\eqref{eq:frog_kernel} using the PSF-based method, the
 randomized HODLR (hierarchical off diagonal low rank)
-method~\cite{Martinsson16,Hartland_2023} with 8 hierarchical levels, and GLR (global low rank)
+method~\cite{Martinsson16,Hartland_2023} with 8 levels, and GLR (global low rank)
 approximation using double-pass randomized SVD~\cite{HalkoMartinssonTropp11}. For

### 2024-02-20 paper_sisc_first_revision/localpsf.tex
paper down to 32 pages. hooray!
fix comma typo

@@
 
-The idea of the proposed method, which we refer to throughout the paper as the ``PSF-based method'', is to use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applications.
+The idea of the proposed method, which we refer to throughout the paper as the ``PSF-based method,'' is to use \emph{impulse response interpolation} to form a high rank approximation of $\Aop$ using a small number of operator applications.
 The impulse response, $\phi_x$, associated with a point, $x$, is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(w) = \left(\rho^*,w\right)_{L^2(\Omega)}$ for all $w \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$.
@@
 	\centering
-	\includegraphics[scale=0.2]{frog_hmatrix_knn_impulse3.pdf}	
+	\includegraphics[scale=0.18]{frog_hmatrix_knn_impulse3.pdf}	
 	\caption{Left: H-matrix structure for $\Aker$.

### 2024-02-20 paper_sisc_first_revision/localpsf.tex
fix grammar issue in moments figure
fix "samples points" typo

@@
 	\end{subfigure}
-	\caption{Left: Impulse response moments: scaling factor ($V$), mean ($\mu$), and covariance ($\Sigma$).
+	\caption{Left: Impulse response moments.
+Scaling factor ($V$), mean ($\mu$), and covariance ($\Sigma$).
 For each point $x \in \Omega$, the quantity $V(x)$ is the integral of $\phi_x$ over $\Omega$, $\mu(x)$ is the location that $\phi_x$ is centered at, and $\Sigma(x)$ is a matrix with eigenvectors and eigenvalues that characterize the width of the support of $\phi_x$ about $\mu(x)$ (see Section~\ref{sec:intromoments}).
@@
 At the sample points, the error is zero.
-The further the point $x$ is from the samples points, the larger the error.
+The further the point $x$ is from the sample points, the larger the error.
 Adding more batches yields a more accurate approximation.
@@
 
-The PSF-based method we propose is a PSF method that is loosely based on ``product convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
+The PSF-based method we propose is loosely based on ``product convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights.
 PC and PSF methods have a long history dating back several decades.

### 2024-02-20 paper_sisc_first_revision/localpsf.tex
fix "support phi_x" typo

@@
 We are inspired by resolution analysis in seismic imaging, in which
-$\Aop^T$ is applied to a random noise function, and the width of the support $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$~\cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
+$\Aop^T$ is applied to a random noise function, and the width of the support of $\phi_x$ is estimated to be the autocorrelation length of the resultant function near $x$~\cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}.
 The moment method that we use estimates the support of $\phi_x$ more accurately than random noise probing in resolution analysis, at the cost of the additional constraint that $\Aop$ has a non-negative integral kernel.

### 2024-02-20 paper_sisc_first_revision/localpsf.tex
not localized -> widely dispersed
a H-matrix -> an H-matrix

@@
 It is acceptable if the support of $\impulseresponse_x$ is not perfectly contained in a localized region (violating Assumption 2), so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
-In principle, the PSF-based method can be applied even if the impulse responses are not localized.
+In principle, the PSF-based method can be applied even if the impulse responses are widely dispersed.
 However, in this case only a small number of impulse responses can be computed per batch, which means more batches, and hence more operator applies, are needed to form an accurate approximation.
@@
 \end{equation}
-Broadly, we will construct an H-matrix approximation of $\mathbf{A}$ by forming a H-matrix approximation of $\boldsymbol{\Aker}$, then multiplying $\boldsymbol{\Aker}$ by $\massmatrix$ (or a lumped mass version of $\massmatrix$) on the left and right using H-matrix methods.
+Broadly, we will construct an H-matrix approximation of $\mathbf{A}$ by forming an H-matrix approximation of $\boldsymbol{\Aker}$, then multiplying $\boldsymbol{\Aker}$ by $\massmatrix$ (or a lumped mass version of $\massmatrix$) on the left and right using H-matrix methods.
 Classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Aker}_{ij}$, but these matrix entries are not easily accessible.
@@
 Here, we use a diagonal lumped mass matrix, so these matrix-matrix multiplications are trivial.
-If a non-diagonal mass matrix is used, one may form a H-matrix representation of the mass matrix, then perform the matrix-matrix multiplications in \eqref{eq:boldA} using H-matrix methods.
+If a non-diagonal mass matrix is used, one may form an H-matrix representation of the mass matrix, then perform the matrix-matrix multiplications in \eqref{eq:boldA} using H-matrix methods.
 We use H1 matrices in the numerical results, but any other H-matrix format could be used instead. For more details on H-matrices, see~\cite{Hackbusch15}.
@@
 stiffness and mass matrices, which are sparse. Therefore, we form
-H-matrix representations of these matrices and combine them into a H-matrix approximation of the regularization term in the Hessian, $\mathbf{R}_H$, using
+H-matrix representations of these matrices and combine them into an H-matrix approximation of the regularization term in the Hessian, $\mathbf{R}_H$, using
 standard sparse H-matrix techniques. Then, H-matrix approximations of the Gauss-Newton Hessian, 

### 2024-02-20 paper_sisc_first_revision/localpsf.tex
define polylig linear

@@
 In target applications, (1) is the dominant cost because applying $\Aop$ to a vector requires an expensive computational procedure such as solving a PDE, and (1) is the only step that requires applying $\Aop$ to vectors.
-All operations that do not require applications of $\Aop$ to vectors are polylog linear, and therefore scalable, in the size of the problem, $N$.
+All operations that do not require applications of $\Aop$ to vectors are polylog linear (i.e., $O(N \log(N)^b)$ for some $b$), and therefore scalable, in the size of the problem, $N$.
 We now describe these costs in detail. For convenience, Table~\ref{tab:vars} lists variable symbols and their approximate sizes.

### 2024-02-20 paper_sisc_first_revision/localpsf.tex
one readthrough complete

@@
 
-The PSF approximation $\mathbf{A}_H$, and therefore $\mathbf{A}_H + \mathbf{R}_H$, may be non-symmetric and indefinite because of approximation error. 
+The PSF-based approximation $\mathbf{A}_H$, and therefore $\mathbf{A}_H + \mathbf{R}_H$, may be non-symmetric and indefinite because of approximation error. 
 This is undesirable because symmetry and positive semi-definiteness are important properties which should be preserved if possible.
@@
 
-In summary, computing all the necessary ingredients to evaluate kernel entries of the PSF approximation requires a handful of operator applications (e.g., $6+\nbatch$ operator applications in two dimensions, or $10+\nbatch$ operator applications in three dimensions, with $\nbatch$ typically in the range 1--25), plus comparatively cheap additional overhead costs, most notably performing ellipsoid intersection tests while choosing sample point batches.
+In summary, computing all the necessary ingredients to evaluate kernel entries of the PSF-based approximation requires a handful of operator applications (e.g., $6+\nbatch$ operator applications in two dimensions, or $10+\nbatch$ operator applications in three dimensions, with $\nbatch$ typically in the range 1--25), plus comparatively cheap additional overhead costs, most notably performing ellipsoid intersection tests while choosing sample point batches.
 Once these ingredients are computed, no more operator applications (and thus PDE solves) are required, and approximate kernel entries can be evaluated rapidly.
@@
 
-Next, we build PSF (1), PSF (5), and PSF (25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\mathbf{\basalfriction}$.
-We use PCG to solve a linear system with the Hessian as the coefficient operator and a right hand vector with random independent and identically distributed (i.i.d.) entries drawn from the standard Gaussian distribution.
+Next, we build PSF~(1), PSF~(5), and PSF~(25) preconditioners based on the Gauss-Newton Hessian evaluated at the converged solution $\mathbf{\basalfriction}$ (note: $k$ in PSF~($k$) refers to the number of batches; this is not to be confused with the noise levels which range over the same numerical values).
+We use PCG to solve a linear system with the Hessian as the coefficient operator and a right hand side vector with random independent and identically distributed (i.i.d.) entries drawn from the standard Gaussian distribution.
 In Figure~\ref{fig:krylov_convergence} (left) we compare the convergence of PCG for solving this linear system using the PSF (1), PSF (5), PSF (25), REG, and NONE preconditioners.
@@
 Right: Generalized eigenvalues for generalized eigenvalue problem $\mathbf{H} \mathbf{u}_{k}=\lambda_{k} \mathbf{\preconditioner} \mathbf{u}_{k}$.
-Here, $\mathbf{H}$ is the Hessian and the preconditioner, $\mathbf{\preconditioner}$, is the PSF approximation for 1, 5, or 25 batches (PSF (1), PSF (5), and PSF (25), respectively), the regularization Hessian (REG), or the identity matrix (NONE).}
+Here, $\mathbf{H}$ is the Hessian and the preconditioner, $\mathbf{\preconditioner}$, is the PSF-based approximation for 1, 5, or 25 batches (PSF (1), PSF (5), and PSF (25), respectively), the regularization Hessian (REG), or the identity matrix (NONE).}
 	\label{fig:adv_krylov_convergence}

### 2024-02-20 paper_sisc_first_revision/localpsf.tex
re-order examples to put blur last

@@
 
-We use the PSF-based method to approximate the integral kernel in a blur problem (Section~\ref{sec:frog}), and the Newton (or Gauss-Newton) Hessians
+We use the PSF-based method to approximate the Newton (or Gauss-Newton) Hessians
 in inverse problems governed by PDEs which model steady state ice
 sheet flow~\cite{PetraEtAl12} (Section~\ref{sec:ice}) and advective-diffusive transport of a contaminant~\cite{PetraStadler11}
-(Section~\ref{sec:adv}). These inverse problems are described in detail in their respective sections.
-In both inverse problems, to reconstruct the unknown parameter fields, denoted $q$, the inverse problems are formulated as nonlinear least squares
+(Section~\ref{sec:adv}), and to approximate the integral kernel in a blur problem that is not based on PDEs (Section~\ref{sec:frog}).
+These problems are described in detail in their respective sections.
+
+In both PDE-based inverse problems (Sections~\ref{sec:ice} and~\ref{sec:adv}), to reconstruct the unknown parameter fields, denoted $q$, the inverse problems are formulated as nonlinear least squares
 optimization problems, whose objective functions consist of a data misfit term
@@
 
-\subsection{Example 1: Spatially varying blurring problem}
-\label{sec:frog}
-
-Here, we define a PDE-free spatially varying blur problem, in which the impulse response, $\phi_x$,
-is a bumpy blob that is centered
-near $x$, and is rotated and scaled in a manner that depends on $x$
-(see Figure \ref{fig:frog_kernel_impulse_responses}). This blur problem is used in
-Sections~\ref{sec:intro} and~\ref{sec:prelims} to visually illustrate
-various stages and aspects of the PSF-based method, and robustness (or lack thereof) of the PSF-based method to violations of the non-negative kernel assumption (Section~\ref{sec:conditions_2}).
-In this section, the blur problem is used to compare the PSF-based method to hierarchical off-diagonal low rank (HODLR) and global low rank (GLR) methods, to study convergence of the PSF-based method, and to investigate the effect of the ellipsoid size parameter $\tau$.
-%To study the
-%robustness of the proposed method to violations of the non-negative
-%kernel assumption discussed in Section~\ref{sec:impulse_response_2},
-The closed form
-expression for the integral kernel is given by
-%
-%Before delving into technical details, we provide an overview of the
-%method. Along the way, we illustrate steps using a spatially variant
-%blur example, in which the integral kernel is given as follows:
-%\begin{equation}
-%\label{eq:frog_kernel}
-%\Phi(y,x) = (1 - a f(y,x)) x^1(1-x^1)x^2(1-x^2) \exp \left(-\frac{1}{2}h^T(y,x) C^{-1} h(y,x)\right).
-%\end{equation}
-%Here, $h(y,x) = R(\theta(x)) (y-x)$, $R(\theta(x))$ is the 2D rotation
-%matrix by angle $\theta(x) = (x^1 + x^2) \pi / 2$, $C = L^2
-%\operatorname{diag}(c_1, c_2)$, $g(x) = x^1(1-x^1)x^2(1-x^2)$, and
-%$f(x,y) = \cos\left(p^1(y,x) / \sqrt{c_1 / 2.0}\right)
-%\sin\left(p^2(y,x) / \sqrt{c_2 / 2.0}\right)$ for some constants $a,
-%L, c_1, c_2$. In this section we use $a=1.0$, $L^2=1.0$, $c_1=0.01$,
-%and $c_2=0.0025$.
-%
-\begin{align}
-\label{eq:frog_kernel}
-\Phi(y,x) = (1 - a f(y,x)) g(x) \exp \left(-\frac{1}{2}(h(y,x)^T
-C^{-1} h(y,x) \right),
-\end{align}
-where $f(y,x) = \cos\left(h^1(y,x) / \sqrt{c_1 /
-  2}\right)\sin\left(h^2(y,x) / \sqrt{c_2 / 2}\right)$, with
-$h^i(y,x)$ the $i^{th}$ component of $R(\theta(x)) (y-x)$, with
-$R(\theta(x))$ a two-dimensional rotation matrix by angle $\theta(x) =
-(x^1 + x^2) \pi / 2$, $g(x) = x^1(1-x^1)x^2(1-x^2)$, and with $C = L^2
-\operatorname{diag}(c_1, c_2)$. The constant $L$ controls the width of
-the blob, and $c_1/c_2$ controls its aspect ratio. The constant $a$
-represents deviation from a Gaussian. When $a=0$, $\phi_x$ is a
-Gaussian, and as $a$ increases, $\phi_x$ becomes non-Gaussian. When $a
-> 1$, the integral kernel contains negative values, which allows us to
-study the robustness of the PSF-based method to violations of Assumption 3
-(Section~\ref{sec:conditions_2}).
-
-%The impulse response $\phi_x$ corresponding to this kernel is a bumpy
-%blob that is centered near $x$, and is rotated and scaled in a manner
-%that depends on $x$ (see Figure
-%\ref{fig:frog_kernel_impulse_responses}). The constant $L$ controls
-%how wide the blob is (hence the support of the impulse response), and
-%$a$ controls the magnitude of the bumps.
-
-%Note that for this
-%operator, using our method is unnecessary because entries of the
-%kernel are given directly via a formula. The purpose of this example
-%is to illustrate the method using an operator which is known exactly
-%and may be easily visualized.
-
-\begin{table}
-	\begin{center}
-		{\small
-			\begingroup
-			\setlength{\tabcolsep}{8pt}
-			\renewcommand{\arraystretch}{1.0}
-			\begin{tabular}{c| c | c c c c c}
-				& Error tol.    &	\#applies PSF   & \#applies HODLR 	& \#applies RSVD 	\\%& rank \\ 
-				\hline 
-								& 20\%    		& 11  				& 592				& 354				\\%& 111  \\ 
-				$L=1$ 			& 10\% 			& 16  				& 772				& 520				\\%& 178  \\ 
-								& 5\% 			& 22   				& 924				& 674				\\%& 244 \\ 
-				\hline
-								& 20\%    		& 8  				& 852 				& 1316				\\%& 417 \\ 
-				$L=1/2$ 		& 10\%    		& 9  				& 1144				& 1916				\\%& 665 \\ 
-								& 5\% 	  		& 12   				& 1404				& 2456				\\%& 903 \\ 
-				\hline
-								& 20\%    	 	& 7					& 932				& 2624				\\%& 845  \\ 
-				$L=1/3$ 		& 10\%    		& 8  				& 1264 				& 3734				\\%& 1321  \\ 
-								& 5\% 			& 8   				& 1520 				& 4660				\\%& 1759  \\ 
-			\end{tabular}
-			\endgroup
-		}
-	\end{center}
-	\caption{(Blur) Comparison of cost to
-		approximate the blur kernel from Equation
-		\eqref{eq:frog_kernel} using the PSF-based method, the randomized
-		HODLR (hierarchical off diagonal low rank) method, and GLR
-		(global low rank) approximation using randomized SVD. The
-		quantity $L$ scales the width of the impulse responses,
-		hence it influences the rank of the operator. Large $L$
-		means low rank, and small $L$ means high rank. The second
-		column (``Error tol'') is the relative error in the
-		approximation of the kernel measured in the Frobenius norm,
-		$||\boldsymbol{\Phi} -
-		\widetilde{\boldsymbol{\Phi}}||_\text{Fro} /
-		||\boldsymbol{\Phi}||_\text{Fro}$. The remaining three
-		columns show the number of operator applies required to
-		achieve the given error tolerances, using the PSF, HODLR,
-		and GLR methods.}
-	\label{tab:frog_psf_vs_hodlr_vs_glr}
-\end{table}
-
-In Table~\ref{tab:frog_psf_vs_hodlr_vs_glr} we compare the
-cost to approximate the blur kernel from
-Equation~\eqref{eq:frog_kernel} using the PSF-based method, the
-randomized HODLR (hierarchical off diagonal low rank)
-method~\cite{Martinsson16,Hartland_2023} with 8 levels, and GLR (global low rank)
-approximation using double-pass randomized SVD~\cite{HalkoMartinssonTropp11}. For
-these results we vary the quantity $L$ to scale the width of the
-impulse responses and hence the rank of the operator. For each case,
-we calculate the relative error in the approximation of the kernel
-measured in the Frobenius norm, $||\boldsymbol{\Phi} -
-\widetilde{\boldsymbol{\Phi}}||_\text{Fro} /
-||\boldsymbol{\Phi}||_\text{Fro}$ and show the number of operator
-applications required to achieve 20\%, 10\%, and 5\% relative error by
-each method. The results reveal superior performance of the PSF
-method as compared to the HODLR and GLR methods for all cases. We note
-that as we increase the rank and decrease the error tolerance, the
-performance of the HODLR and GLR methods deteriorates.
-
-\begin{figure}
-	\begin{subfigure}{0.52\textwidth}
-		\begin{center}
-			\includegraphics[scale=1.0]{frog_tau_impulses.pdf}
-		\end{center}
-	\end{subfigure}
-	\begin{subfigure}{0.47\textwidth}
-		\begin{center}
-			\includegraphics[scale=1.0]{frog_tau_batches.pdf}
-		\end{center}
-	\end{subfigure}
-	\caption{(Blur) Relative error 
-		%		in the PSF approximation of the kernel
-		%	$||\boldsymbol{\Phi} - \widetilde{\boldsymbol{\Phi}}||_\text{Fro}/||\boldsymbol{\Phi}||_\text{Fro}$,
-		for different ellipsoid size parameters, $\tau$, vs. the total number of impulse responses (left), and the number of batches (right). 
-		%	Left: Relative error vs. the total number of impulse responses used in the approximation. Right: Relative error vs. the number of impulse response batches. 
-		The dashed gray lines show linear convergence rates, i.e., $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$ on the left, and  $\text{constant} \times \# \text{batches}^{-1}$ on the right.}
-	\label{fig:PSF_tau_convergence}
-\end{figure}
-
-In Figure \ref{fig:PSF_tau_convergence} we show the convergence of the PSF-based method on the blur kernel as a function of the total number of impulse responses (left), and the number of batches (right).
-We show convergence for several ellipsoid size parameters $\tau$, ranging from $2.0$--$4.0$.
-The results in Figure \ref{fig:PSF_tau_convergence} (left) show that the relative error decreases as $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$
-% or $\text{constant} \times \# \text{batches}^{-1}$, 
-suggesting linear convergence. The linear convergence stalls at a limit that depends on $\tau$. Increasing $\tau$ lowers this limit, allowing the PSF-based method to achieve higher accuracy. 
-In Figure \ref{fig:PSF_tau_convergence} (right), the results show that before this limit is reached, the convergence is faster for smaller $\tau$ in terms of the number of batches.
-This is expected because smaller $\tau$ results in more impulse responses per batch.
-Larger $\tau$ causes the PSF-based method to converge more slowly than smaller $\tau$, but with larger $\tau$ the PSF-based method stalls at a lower level of error than it does with smaller $\tau$ (see, e.g., $\tau=4.0$ vs.
-$\tau=2.5$).
-
-%----
-%, but allows the method to eventually achieve a lower level of error.
-%
-%----
-%
-%For smaller $\tau$ regime, we observe faster convergence
-%
-%Before this limit is reached, the convergence rate is the same for all $\tau$ as a function the number of impulse responses.
-This means that the convergence is faster for smaller $\tau$ as a function of the number of batches, since smaller $\tau$ means more impulse responses per batch.
-In summary, taking $\tau$ larger makes the method converge more slowly, but allows it to eventually achieve a lower level of error.
-
-%Define HODLR. Define randomized HODLR algo. Ref to Tucker's paper and "Fast randomized eigensolver with structured LDL..."
-
-%Define randomized SVD, refer to Halko Martinsson Tropp.
-
-%Discuss why PSF does better with smaller $L$.
-
-\subsection{Example 2: Inversion for the basal friction coefficient in an ice sheet flow problem}
+\subsection{Example 1: Inversion for the basal friction coefficient in an ice sheet flow problem}
 \label{sec:ice}
@@
 
-\subsection{Example 3: Inversion for the initial condition in an advective-diffusive transport problem}
+\subsection{Example 2: Inversion for the initial condition in an advective-diffusive transport problem}
 \label{sec:adv}
@@
 
+\subsection{Example 3: Spatially varying blurring problem}
+\label{sec:frog}
+
+Here, we define a PDE-free spatially varying blur problem, in which the impulse response, $\phi_x$,
+is a bumpy blob that is centered
+near $x$, and is rotated and scaled in a manner that depends on $x$
+(see Figure \ref{fig:frog_kernel_impulse_responses}). This blur problem is used in
+Sections~\ref{sec:intro} and~\ref{sec:prelims} to visually illustrate
+various stages and aspects of the PSF-based method, and robustness (or lack thereof) of the PSF-based method to violations of the non-negative kernel assumption (Section~\ref{sec:conditions_2}).
+In this section, the blur problem is used to compare the PSF-based method to hierarchical off-diagonal low rank (HODLR) and global low rank (GLR) methods, to study convergence of the PSF-based method, and to investigate the effect of the ellipsoid size parameter $\tau$.
+%To study the
+%robustness of the proposed method to violations of the non-negative
+%kernel assumption discussed in Section~\ref{sec:impulse_response_2},
+The closed form
+expression for the integral kernel is given by
+%
+%Before delving into technical details, we provide an overview of the
+%method. Along the way, we illustrate steps using a spatially variant
+%blur example, in which the integral kernel is given as follows:
+%\begin{equation}
+%\label{eq:frog_kernel}
+%\Phi(y,x) = (1 - a f(y,x)) x^1(1-x^1)x^2(1-x^2) \exp \left(-\frac{1}{2}h^T(y,x) C^{-1} h(y,x)\right).
+%\end{equation}
+%Here, $h(y,x) = R(\theta(x)) (y-x)$, $R(\theta(x))$ is the 2D rotation
+%matrix by angle $\theta(x) = (x^1 + x^2) \pi / 2$, $C = L^2
+%\operatorname{diag}(c_1, c_2)$, $g(x) = x^1(1-x^1)x^2(1-x^2)$, and
+%$f(x,y) = \cos\left(p^1(y,x) / \sqrt{c_1 / 2.0}\right)
+%\sin\left(p^2(y,x) / \sqrt{c_2 / 2.0}\right)$ for some constants $a,
+%L, c_1, c_2$. In this section we use $a=1.0$, $L^2=1.0$, $c_1=0.01$,
+%and $c_2=0.0025$.
+%
+\begin{align}
+\label{eq:frog_kernel}
+\Phi(y,x) = (1 - a f(y,x)) g(x) \exp \left(-\frac{1}{2}(h(y,x)^T
+C^{-1} h(y,x) \right),
+\end{align}
+where $f(y,x) = \cos\left(h^1(y,x) / \sqrt{c_1 /
+	2}\right)\sin\left(h^2(y,x) / \sqrt{c_2 / 2}\right)$, with
+$h^i(y,x)$ the $i^{th}$ component of $R(\theta(x)) (y-x)$, with
+$R(\theta(x))$ a two-dimensional rotation matrix by angle $\theta(x) =
+(x^1 + x^2) \pi / 2$, $g(x) = x^1(1-x^1)x^2(1-x^2)$, and with $C = L^2
+\operatorname{diag}(c_1, c_2)$. The constant $L$ controls the width of
+the blob, and $c_1/c_2$ controls its aspect ratio. The constant $a$
+represents deviation from a Gaussian. When $a=0$, $\phi_x$ is a
+Gaussian, and as $a$ increases, $\phi_x$ becomes non-Gaussian. When $a
+> 1$, the integral kernel contains negative values, which allows us to
+study the robustness of the PSF-based method to violations of Assumption 3
+(Section~\ref{sec:conditions_2}).
+
+%The impulse response $\phi_x$ corresponding to this kernel is a bumpy
+%blob that is centered near $x$, and is rotated and scaled in a manner
+%that depends on $x$ (see Figure
+%\ref{fig:frog_kernel_impulse_responses}). The constant $L$ controls
+%how wide the blob is (hence the support of the impulse response), and
+%$a$ controls the magnitude of the bumps.
+
+%Note that for this
+%operator, using our method is unnecessary because entries of the
+%kernel are given directly via a formula. The purpose of this example
+%is to illustrate the method using an operator which is known exactly
+%and may be easily visualized.
+
+\begin{table}
+	\begin{center}
+		{\small
+			\begingroup
+			\setlength{\tabcolsep}{8pt}
+			\renewcommand{\arraystretch}{1.0}
+			\begin{tabular}{c| c | c c c c c}
+				& Error tol.    &	\#applies PSF   & \#applies HODLR 	& \#applies RSVD 	\\%& rank \\ 
+				\hline 
+				& 20\%    		& 11  				& 592				& 354				\\%& 111  \\ 
+				$L=1$ 			& 10\% 			& 16  				& 772				& 520				\\%& 178  \\ 
+				& 5\% 			& 22   				& 924				& 674				\\%& 244 \\ 
+				\hline
+				& 20\%    		& 8  				& 852 				& 1316				\\%& 417 \\ 
+				$L=1/2$ 		& 10\%    		& 9  				& 1144				& 1916				\\%& 665 \\ 
+				& 5\% 	  		& 12   				& 1404				& 2456				\\%& 903 \\ 
+				\hline
+				& 20\%    	 	& 7					& 932				& 2624				\\%& 845  \\ 
+				$L=1/3$ 		& 10\%    		& 8  				& 1264 				& 3734				\\%& 1321  \\ 
+				& 5\% 			& 8   				& 1520 				& 4660				\\%& 1759  \\ 
+			\end{tabular}
+			\endgroup
+		}
+	\end{center}
+	\caption{(Blur) Comparison of cost to
+		approximate the blur kernel from Equation
+		\eqref{eq:frog_kernel} using the PSF-based method, the randomized
+		HODLR (hierarchical off diagonal low rank) method, and GLR
+		(global low rank) approximation using randomized SVD. The
+		quantity $L$ scales the width of the impulse responses,
+		hence it influences the rank of the operator. Large $L$
+		means low rank, and small $L$ means high rank. The second
+		column (``Error tol'') is the relative error in the
+		approximation of the kernel measured in the Frobenius norm,
+		$||\boldsymbol{\Phi} -
+		\widetilde{\boldsymbol{\Phi}}||_\text{Fro} /
+		||\boldsymbol{\Phi}||_\text{Fro}$. The remaining three
+		columns show the number of operator applies required to
+		achieve the given error tolerances, using the PSF, HODLR,
+		and GLR methods.}
+	\label{tab:frog_psf_vs_hodlr_vs_glr}
+\end{table}
+
+In Table~\ref{tab:frog_psf_vs_hodlr_vs_glr} we compare the
+cost to approximate the blur kernel from
+Equation~\eqref{eq:frog_kernel} using the PSF-based method, the
+randomized HODLR (hierarchical off diagonal low rank)
+method~\cite{Martinsson16,Hartland_2023} with 8 levels, and GLR (global low rank)
+approximation using double-pass randomized SVD~\cite{HalkoMartinssonTropp11}. For
+these results we vary the quantity $L$ to scale the width of the
+impulse responses and hence the rank of the operator. For each case,
+we calculate the relative error in the approximation of the kernel
+measured in the Frobenius norm, $||\boldsymbol{\Phi} -
+\widetilde{\boldsymbol{\Phi}}||_\text{Fro} /
+||\boldsymbol{\Phi}||_\text{Fro}$ and show the number of operator
+applications required to achieve 20\%, 10\%, and 5\% relative error by
+each method. The results reveal superior performance of the PSF
+method as compared to the HODLR and GLR methods for all cases. We note
+that as we increase the rank and decrease the error tolerance, the
+performance of the HODLR and GLR methods deteriorates.
+
+\begin{figure}
+	\begin{subfigure}{0.52\textwidth}
+		\begin{center}
+			\includegraphics[scale=1.0]{frog_tau_impulses.pdf}
+		\end{center}
+	\end{subfigure}
+	\begin{subfigure}{0.47\textwidth}
+		\begin{center}
+			\includegraphics[scale=1.0]{frog_tau_batches.pdf}
+		\end{center}
+	\end{subfigure}
+	\caption{(Blur) Relative error 
+		%		in the PSF approximation of the kernel
+		%	$||\boldsymbol{\Phi} - \widetilde{\boldsymbol{\Phi}}||_\text{Fro}/||\boldsymbol{\Phi}||_\text{Fro}$,
+		for different ellipsoid size parameters, $\tau$, vs. the total number of impulse responses (left), and the number of batches (right). 
+		%	Left: Relative error vs. the total number of impulse responses used in the approximation. Right: Relative error vs. the number of impulse response batches. 
+		The dashed gray lines show linear convergence rates, i.e., $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$ on the left, and  $\text{constant} \times \# \text{batches}^{-1}$ on the right.}
+	\label{fig:PSF_tau_convergence}
+\end{figure}
+
+In Figure \ref{fig:PSF_tau_convergence} we show the convergence of the PSF-based method on the blur kernel as a function of the total number of impulse responses (left), and the number of batches (right).
+We show convergence for several ellipsoid size parameters $\tau$, ranging from $2.0$--$4.0$.
+The results in Figure \ref{fig:PSF_tau_convergence} (left) show that the relative error decreases as $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$
+% or $\text{constant} \times \# \text{batches}^{-1}$, 
+suggesting linear convergence. The linear convergence stalls at a limit that depends on $\tau$. Increasing $\tau$ lowers this limit, allowing the PSF-based method to achieve higher accuracy. 
+In Figure \ref{fig:PSF_tau_convergence} (right), the results show that before this limit is reached, the convergence is faster for smaller $\tau$ in terms of the number of batches.
+This is expected because smaller $\tau$ results in more impulse responses per batch.
+Larger $\tau$ causes the PSF-based method to converge more slowly than smaller $\tau$, but with larger $\tau$ the PSF-based method stalls at a lower level of error than it does with smaller $\tau$ (see, e.g., $\tau=4.0$ vs.
+$\tau=2.5$).
+
+%----
+%, but allows the method to eventually achieve a lower level of error.
+%
+%----
+%
+%For smaller $\tau$ regime, we observe faster convergence
+%
+%Before this limit is reached, the convergence rate is the same for all $\tau$ as a function the number of impulse responses.
+This means that the convergence is faster for smaller $\tau$ as a function of the number of batches, since smaller $\tau$ means more impulse responses per batch.
+In summary, taking $\tau$ larger makes the method converge more slowly, but allows it to eventually achieve a lower level of error.
+
+%Define HODLR. Define randomized HODLR algo. Ref to Tucker's paper and "Fast randomized eigensolver with structured LDL..."
+
+%Define randomized SVD, refer to Halko Martinsson Tropp.
+
+%Discuss why PSF does better with smaller $L$.
+
 \section{Conclusions}

### 2024-02-20 revision/reply_to_reviewer1.tex
update response letters with new numbering of sections, figures, and tables after moving blur section

@@
   \noindent{
-  	In Figure 8 in the revised manuscript we added a convergence study for one of the blur kernel examples.
+  	In Figure 14 in the revised manuscript we added a convergence study for one of the blur kernel examples.
 We show the convergence of the proposed PSF method as a function of the total number of impulse responses and the number of batches, for several ellipsoid size parameters $\tau$ ranging from $2.0$--$4.0$.
@@
 Basically, with larger $\tau$ the method converge more slowly than with smaller $\tau$, but larger $\tau$ allows the method to eventually achieve a lower level of error.
-We added a discussion of this convergence study in Section 7.1 of the revised manuscript.
+We added a discussion of this convergence study in Section 7.3 of the revised manuscript.
 
   	Although the method can achieve high accuracy, we would like to emphasize that it is targeted at achieving modest accuracies (e.g., $5\%$ error) to build preconditioners for problems that are challenging for existing methods.
-In Table 2 in the revised manuscript, following the other reviewer's suggestion, we added a comparison of the computational cost
+In Table 5 in the revised manuscript, following the other reviewer's suggestion, we added a comparison of the computational cost
   	(measured in operator applies) to approximate the kernel in one of the spatially

### 2024-02-20 revision/reply_to_reviewer2.tex
incorporated Omar's comments
update response letters with new numbering of sections, figures, and tables after moving blur section

@@
 
-  \noindent{Thank you for this suggestion!
+  \noindent{
+  	Thank you for this suggestion, which we agree can help the reader better understand the intuition behind the method.
 We added a 2D blurring kernel example, with figures illustrating the end-to-end PSF method.
-These are shown in the introduction and the numerical results sections (Sections 1 and 7.1).
-We think that adding these illustrations helps the reader better understand the intuition behind the method.
-%  	Frog example with a=1.
-%  	Structure: local non-negative kernel
-%(0) Matrix itself column highlighted, show impulse response (kernel\_impulse\_illustration.pdf)
-%(1) moments. Show one impulse with ellipsoid with widths illustrated
-%(2) impulse response batches
-%(3) RBF interpolation (?), evaluation of entries. Columns we know, and entry we don't know. 
-%(4) H-matrix (structure only)
+These are shown in the introduction and the numerical results sections (Sections 1 and 7.3).
 }
@@
 
-    \noindent{We added a numerical comparison of our method with the HODLR method in [7] for a blur kernel in Table 2 of Section 7.1 in the revised manuscript.
+    \noindent{We added a numerical comparison of our method with the HODLR method in [7] for a blur kernel in Table 5 of Section 7.3 in the revised manuscript.
 The results show that the HODLR method requires hundreds to thousands of operator applies to achieve modest accuracies (e.g., 10\% error) while our method requires only order 10 operator applies.

### 2024-02-20 revision/reply_to_reviewer1.tex
add final revision pdfs to their own subdirectory. Going to submit now

@@
 This means that the error decreases faster for smaller $\tau$ as a function of the number of batches, since smaller $\tau$ means more impulse responses per batch.
-Basically, with larger $\tau$ the method converge more slowly than with smaller $\tau$, but larger $\tau$ allows the method to eventually achieve a lower level of error.
+Basically, with larger $\tau$ the method converges more slowly than with smaller $\tau$, but larger $\tau$ allows the method to eventually achieve a lower level of error.
 We added a discussion of this convergence study in Section 7.3 of the revised manuscript.

### 2024-02-20 revision/reply_to_reviewer2.tex
add final revision pdfs to their own subdirectory. Going to submit now

@@
 
-  \noindent{We added text in the paragraph at the end of Section 2 to discuss the (lack of) applicability of the method to wave inverse problems and added these references.}
+  \noindent{We added text in the paragraph at the end of Section 2 to discuss the (lack of) applicability of our method to wave inverse problems, and added these references.}
 

### 2024-02-22 sisc_final_version/localpsf.tex
made directory with sisc final files

--- 2024-02-20 paper_sisc_first_revision/localpsf.tex
@@
 Here, $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, and $L^2(\Omega)'$ is the space of real-valued continuous linear functionals on $L^2(\Omega)$.
-By ``non-negative integral kernel,'' we mean that entries of $\Aop$'s integral kernel
-%(defined in Equation \ref{eq:kernel_representation})
-are non-negative numbers; this is not the same as positive semi-definiteness of $\Aop$.
-Such operators appear, for instance, as Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{BorziSchulz11,Delosreyes15,HinzePinnauUlbrichUlbrich08},
-%AntilEtAl18
-Schur complements in Schur complement methods for solving partial differential equations and Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps)~\cite{ChanMathew94,SmithBjorstadGropp96,ToselliWidlund04}, covariance operators in spatial statistics~\cite{ChenStein21,GentonKeyesTurkiyyah18,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}.
+By ``non-negative integral kernel,'' we mean that entries of $\Aop$'s integral kernel are non-negative numbers; this is not the same as positive semi-definiteness of $\Aop$.
+Such operators appear, for instance, as Hessians in optimization and inverse problems governed by partial differential equations (PDEs)~\cite{BorziSchulz11,Delosreyes15,HinzePinnauUlbrichUlbrich08}, Schur complements in Schur complement methods for solving partial differential equations and Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps)~\cite{ChanMathew94,SmithBjorstadGropp96,ToselliWidlund04}, covariance operators in spatial statistics~\cite{ChenStein21,GentonKeyesTurkiyyah18,GeogaEtAl20,LindgrenRueLindstrom11}, and blurring operators in imaging~\cite{DenisEtAl11,NagyOleary98}.
 Here, ``matrix-free'' means that we may apply $\Aop$ and its transpose\footnote{Recall that $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(w) = \left(\Aop^T w\right)(u)$ for all $u,w \in L^2(\Omega)$, where $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(w)$ is the result of applying that linear functional to $w \in L^2(\Omega)$, and similar for operations with $\Aop^T$.
@@
 but cannot easily access entries of $\Aop$'s integral kernel. Evaluating the maps in~\eqref{eq:eval_maps} may require solving a subproblem that involves PDEs, or performing other costly computations.
-%Matrix-free does not mean that matrices are never used; our approximation is ultimately converted into a hierarchical matrix (H-matrix), a compressed matrix format that permits fast and scalable linear algebra operations.
 
@@
 	\end{subfigure}
-%	\begin{subfigure}{0.32\textwidth}
-%		\centering
-%		\includegraphics[scale=0.34]{impulse_batch3.png}
-%	\end{subfigure}
 	\caption{(\ref{fig:batches_intro}) One batch, $\eta_b$, of normalized impulse responses, $\impulseresponse_{x}$, that arise from applying $\Aop$ to a weighted sum of scattered point sources (see Section \ref{sec:get_impulse_response}).
@@
 	\begin{center}
-		%	\begin{subfigure}[b]{0.58\textwidth}
-		%		\centering
 		\includegraphics[scale=0.3]{kernel_impulse_illustration2.pdf}
-		%%		\caption{Kernel and impulse responses}
-		%		\label{fig:kernel_impulse_illustration}
-		%	\end{subfigure}
-		%	\begin{subfigure}[b]{0.41\textwidth}
-		%		\centering
-		%		\includegraphics[scale=0.55]{frog_ellipsoid_width2.pdf}
-		%%		\caption{Impulse response support ellipsoid}
-		%		\label{fig:frog_ellipsoid_width}
-		%	\end{subfigure}
 	\end{center}
@@
 Rows and columns are ordered according to a kd-tree hierarchical clustering.
-Right: Impulse responses
-%	$\phi_{x_1}=\Phi(~\cdot~, x_1)$ and $\phi_{x_1}=\Phi(~\cdot~, x_2)$
-	associated with points $x_1,x_2\in \Omega$, shown by the two dotted vertical lines. Intuitively, one may think of impulse responses as ``columns'' of the integral kernel.
+Right: Impulse responses associated with points $x_1,x_2\in \Omega$, shown by the two dotted vertical lines.
+Intuitively, one may think of impulse responses as ``columns'' of the integral kernel.
 	}
@@
 We then use a greedy ellipsoid packing algorithm (Figure~\ref{fig:frog_batches}) to maximize the number of impulse responses per batch.
-% which requires us to estimate the supports of the impulse responses $\phi_x$ \emph{before} they are computed. 
 Then we interpolate translated and scaled versions of these impulse responses to approximate entries of the operator's integral kernel (Figure~\ref{fig:hmatrix_neighbors_frog}).
@@
 		\includegraphics[scale=0.44]{frog_ellipsoid_width2.pdf}
-		%		\label{fig:frog_ellipsoid_width}
 	\end{subfigure}
@@
 Then, an ellipsoid packing problem is solved to choose batches of non-overlapping support ellipsoids (Section~\ref{sec:sample_point_selection}).
-%		Note that we do not need to compute the impulse response in order to know its support ellipsoid. 
-	Finally, $\Aop$ is applied to a Dirac comb associated with the points $x_i$, which correspond to the ellipsoids (Section~\ref{sec:get_impulse_response}). The process is repeated to form more batches.
-%		
-%		Impulse response moments are used to choose batches of non-overlapping impulse responses by solving an ellipsoid packing problem using a greedy algorithm. Batches of impulse responses 
-%		
-%		Moments of the impulse responses used to form ellipsoid shaped a-priori estimates for the supports of the impulse responses.
+Finally, $\Aop$ is applied to a Dirac comb associated with the points $x_i$, which correspond to the ellipsoids (Section~\ref{sec:get_impulse_response}).
+The process is repeated to form more batches.
 	}
@@
 Using radial basis function interpolation, the desired kernel entry is approximated as a weighted linear combination of translated and scaled versions of impulse responses at the points $x'$ (Section~\ref{sec:approximate_kernel_entries}).
-%		the values $\phi_{x’}(y - \spatialmean(x) + \spatialmean(x’))$ for these nearby points $x'$.
-%	Right:  Illustration of the radial basis function interpolation method used to approximate $\Aker(y,x)$ by a weighted linear combination of the values $\frac{\spatialvol(x)}{\spatialvol(x’)}\phi_{x’}(y - \spatialmean(x) + \spatialmean(x’))$ for a small number of sample points near $x$.
-The weights are determined by radial basis function (RBF) interpolation.
-Right: The impulse response centered at $x’$.
 	}
@@
 		\includegraphics[scale=0.32]{frog_column_errors_psf5.png}
-		%		\caption{PSF(5)}
 		\label{fig:blur_error_psf5}
@@
 		\includegraphics[scale=0.32]{frog_column_errors_psf10.png}
-		%		\caption{PSF(10)}
 		\label{fig:blur_error_psf10}
@@
 		\includegraphics[scale=0.32]{frog_column_errors_psf20.png}
-		%		\caption{PSF(20)}
 		\label{fig:blur_error_psf20}
@@
 		\includegraphics[scale=0.32]{frog_colorbar_edited1.pdf}
-		%			\label{fig:blur_colorbar}
 	\end{subfigure}
@@
 Adding more batches yields a more accurate approximation.
-%		columnwise error plots for kernel reconstruction left 5, center 10, right 20
 	}
@@
 
-%\begin{figure}
-%	\begin{center}
-%		\includegraphics[scale=0.75]{mean_displacement_invariance.pdf}
-%	\end{center}
-%	\caption{This figure illustrates impulse responses, $\impulseresponse_x$ and $\impulseresponse_{x'}$, that arise from applying $\Aop$ to point sources at two points, $x$ and $x'$.
-The operator $\Aop$ is locally mean displacement invariant (Section \ref{sec:local_mean_displacement_invariance}) if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$.
-Here $\spatialmean(z)$ denotes the mean (center of mass) of $\phi_z$.}
-%	\label{fig:mean_displacement_invariance}
-%\end{figure}
-
 The ability to rapidly approximate entries of $\Aop$'s integral kernel allows one to approximate discretized versions of $\Aop$ using the full arsenal of tools for matrix approximation that rely on fast access to matrix entries.
@@
 Rather, the Hessian may be applied to a vector via a computational process that involves solving a pair of forward and adjoint PDEs which are linearizations of the original PDE~\cite{GhattasWillcox21,PetraStadler11}.
-%,Gunzburger03}
 The most popular matrix-free Hessian approximation methods are based on low rank approximation of either the data misfit Hessian, or the data misfit Hessian preconditioned by the regularization Hessian, e.g.,~\cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15}.
@@
 In Section \ref{sec:frog}, we see numerically that the randomized hierarchical off-diagonal low rank (HODLR) method in \cite{Martinsson16} requires hundreds to thousands of matrix-vector products to construct approximations of the integral kernel for a blur problem example with modest (e.g., $10\%$) relative error.
-%while the PSF method in this paper can achieve the same level of error in one to two dozen operator applies.
 Matrix-free H-matrix construction is currently an active area of research, hence these costs may decrease as new algorithms are developed.
@@
 
-%If these conditions are violated, the method will incur additional error, and depending on the severity of the violation, the proposed method may fail.
-
 It is acceptable if the support of $\impulseresponse_x$ is not perfectly contained in a localized region (violating Assumption 2), so long as the bulk of the ``mass'' of $\impulseresponse_x$ is contained in a localized region.
@@
 
-%The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$ (Assumption 3) so long as it is non-negative for the vast majority of pairs of points $(y,x)$, and so long as the negative numbers are comparatively small.
-
 \begin{figure}
@@
 	{
-		%	\setlength{\tabcolsep}{0.1em}
-		%	\renewcommand{\arraystretch}{2.0}
 		\begin{tabular}{cccc}
@@
 \end{itemize} 
-%The location of the negative numbers also affects the accuracy of the ellipsoid estimates substantially, but in a more complicated manner. 
 For the kernel in the left two columns of Figure \ref{fig:robustness_to_negatives}, negative numbers are interspersed with positive numbers, allowing us to include a large amount of negative numbers before the ellipsoid estimation fails.
@@
 In the bottom two rows, we see the aforementioned threshold effect, in which the ellipsoid estimation method rapidly transitions from performing reasonably well to performing poorly with only a small change to the integral kernel.
-
-%Location and magnitude of negatives matter.
-Larger negatives: bad, smaller negatives good.
-Negatives closer to center: good, further from center: bad.
-Correlated negatives: bad, randomly distributed: good.
-Good Good Good Fail quick
-%
-%Non-local: will work, but choose few ellipsoids, so inefficient
 
@@
 Other mass lumping techniques may be used.
-%~\cite{Hughes12}.
 
@@
 \horzbar & \mathbf{a}_1^T & \horzbar \\
-%\horzbar & \mathbf{a}_2^T & \horzbar \\
 & \vdots & \\
@@
 \mathbf{a}_1^T \mathbf{w} \\
-%\mathbf{a}_2^T \mathbf{w} \\
 \vdots \\
@@
 Operators satisfying~\eqref{eq:translate1} can be well approximated via low rank CUR approximation.
-%~\cite{MahoneyDrineas09}.
 However, the required rank in the low rank approximation can be large, which makes algorithms based on~\eqref{eq:translate1} expensive.
@@
 Before presenting the technical details of the algorithm in Sections \ref{sec:sample_point_selection}--\ref{sec:flip_eigs}, we first provide an overview.
-%Here we provide an overview of the operator approximation algorithm. 
-%Then, in the subsections (Sections \ref{sec:sample_point_selection}--\ref{sec:flip_eigs}) we present technical details of the algorithm that were not already covered in Section \ref{sec:prerequisites}.
 
@@
 If $\candidatepoint_i$ passes these checks 
-%(i.e., if $\candidatepoint_i$ is sufficiently far from other points in the batch, and $\spatialvol(\candidatepoint_i)$ and $\spatialcov(\candidatepoint_i)$ are acceptable)
 then we add $\candidatepoint_i$ to $\pointbatch_k$. Otherwise we discard $\candidatepoint_i$. This process repeats until there are no more points in $\candidatepts$.  
@@
 
-%\begin{figure}
-%	\begin{subfigure}[b]{0.24\textwidth}
-%		\centering
-%		\includegraphics[scale=0.25]{rbf_cosine_shape_parameter=1.0.pdf}
-%	\end{subfigure}
-%	\begin{subfigure}[b]{0.24\textwidth}
-%		\centering
-%		\includegraphics[scale=0.25]{rbf_cosine_shape_parameter=10.0.pdf}
-%	\end{subfigure}
-%	\begin{subfigure}[b]{0.24\textwidth}
-%		\centering
-%		\includegraphics[scale=0.25]{rbf_cosine_shape_parameter=100.0.pdf}
-%	\end{subfigure}
-%	\begin{subfigure}[b]{0.24\textwidth}
-%		\centering
-%		\includegraphics[scale=0.25]{cosine_RBF_convergence.pdf}
-%	\end{subfigure}
-%	
-%	\caption{asdf
-%	}
-%	\label{fig:RBF_width}
-%\end{figure}
-
 The shape parameter, $C_\text{RBF}$, mediates a tradeoff between accuracy and stability.
 Small $C_\text{RBF}$ is required for RBF interpolation with Gaussian kernels to achieve high accuracy, but small $C_\text{RBF}$ also makes RBF interpolation less robust to errors or nonsmoothness in the function being interpolated.
-%This is illustrated in Figure \ref{fig:RBF_width} for a simple one dimensional interpolation problem. 
 For our numerical results involving Hessians in inverse problems governed by PDEs (Sections~\ref{sec:ice}~and~\ref{sec:adv}), high accuracy is not required because the PSF-based method is used to construct a preconditioner.
@@
 	\begin{align}
-	%-\nabla \cdot \stress(v, p)
-	%&=\bodyforce \,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:1}\\
-	%\nabla\cdot \velocity
-	  %&=0\,\,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:2}\\
           -\nabla \cdot \stress(v, p)=\bodyforce \text{ and }
@@
 \nabla\velocity+\nabla\velocity^{\top} \right)$ is the strain rate tensor~\cite{IsaacEtAl15}. Here, $\eta$ is the viscosity and $I$ is the identity operator. 
-%Glen's exponent has been chosen to be one, which provides for a linear Stokes model. 
 Note that while the PDE is linear, the parameter-to-solution map, $\basalfriction \mapsto (\velocity, \pressure)$, is nonlinear.
@@
 The parameter $\basalfriction$ is viewed as a function on $\Gamma_b$ for the purpose of solving the Stokes equations, and as a function on $\Omega$ for the purpose of building Hessian approximations and defining the regularization.
-%This flattening is necessary because our operator approximation method requires adding points on this surface together to get other points on the surface, and this is only well-defined in flat space.
 The observations are generated by adding multiplicative
@@
 	\caption{(Ice sheet) Convergence history for solving the Stokes inverse problem using inexact Newton PCG to tolerance $10^{-6}$. 
-%	The Newton iteration is terminated when $||\mathbf{g}|| < 10^{-6} ||\mathbf{g}_0||$, where $\mathbf{g}$ is the gradient at the current iteration, and $\mathbf{g}_0$ is the gradient at the initial guess.
 	Preconditioners shown are the PSF-based method with five batches (PSF (5)) constructed at the third iteration, regularization preconditioning (REG), and no preconditioning (NONE).
@@
 Columns \#Stokes show the total number of Stokes PDE solves performed in each Newton iteration.
-%	This consists of Stokes solves performed during the line search to arrive at the current iterate from the previous iterate, plus one adjoint Stokes solve to compute the gradient at the current iterate, plus two Stokes solves per PCG iteration when solving the Newton system to compute the next search direction.
 	Under PSF (5) and in row Iter 3, we write $6+22$ to indicate that $6$ Stokes solves were used during the standard course of the iteration, and $22$ Stokes solves were used to build the PSF (5) preconditioner.}
@@
 			& 1.0e-4    & 584 & 317  & 311   & 151   & 56    \\ 
-			%			& $1.8e{-4}$    & 628 & 313  & 192   & 131   & 49    \\   
 			$T=0.5$ & 3.2e-4    & 685 & 311  & 233   & 140   & 44    \\ 
-			%			& $5.6e{-4}$    & 734 & 310  & 334   & 144   & 40    \\  
 			& 1.0e-3    & 702 & 324  & 122   & 71   & 33    \\ 
@@
 			& 1.0e-4    & 634 & 449  & 539   & 288   & 100    \\ 
-			%			& $1.8e{-4}$    & 665 & 449  & 493   & 212   & 84    \\   
 			$T=1.0$ & 3.2e-4    & 681 & 459  & 350   & 202   & 90    \\ 
-			%			& $5.6e{-4}$    & 619 & 482  & 262   & 197   & 70    \\  
 			& 1.0e-3    & 574 & 520  & 266   & 260   & 208    \\ 
@@
 			& 1.0e-4    & 609 & 591  & 548   & 520   & 165    \\ 
-			%			& 1.8e{-4}$    & 576 & 611  & 434   & 426   & 185    \\   
 			$T=2.0$ & 3.2e-4    & 524 & 645  & 318   & 379   & 170    \\ 
-			%			& $5.6e{-4}$    & 464 & 711  & 480   & 289   & 153    \\  
 			& 1.0e-3    & 349 & 786  & 381   & 262   & 158    \\ 
@@
 	{
-		%	\setlength{\tabcolsep}{0.1em}
-		%	\renewcommand{\arraystretch}{2.0}
 		\begin{tabular}{ccc}
 			& $T=0.5$ 
-%			& $T=1.0$ 
 			& $T=2.0$ \\
 			$\kappa=1.0$e${-4}$  & \includegraphics[scale=0.12, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-4.png} 
-%			& \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-4.png}
 			& \includegraphics[scale=0.12, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-4.png} \\
-%			$\kappa=3.2$e${-4}$ & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K3.2e-4.png} & \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K3.2e-4.png} \\
 			$\kappa=1.0$e${-3}$ & \includegraphics[scale=0.12, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-3.png} 
-%			& \includegraphics[scale=0.15, valign=c]{impulse_response_P(0.3,0.6)_T1.0_K1.e-3.png} 
 			& \includegraphics[scale=0.12, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-3.png}
@@
 In this section, the blur problem is used to compare the PSF-based method to hierarchical off-diagonal low rank (HODLR) and global low rank (GLR) methods, to study convergence of the PSF-based method, and to investigate the effect of the ellipsoid size parameter $\tau$.
-%To study the
-%robustness of the proposed method to violations of the non-negative
-%kernel assumption discussed in Section~\ref{sec:impulse_response_2},
 The closed form
 expression for the integral kernel is given by
-%
-%Before delving into technical details, we provide an overview of the
-%method. Along the way, we illustrate steps using a spatially variant
-%blur example, in which the integral kernel is given as follows:
-%\begin{equation}
-%\label{eq:frog_kernel}
-%\Phi(y,x) = (1 - a f(y,x)) x^1(1-x^1)x^2(1-x^2) \exp \left(-\frac{1}{2}h^T(y,x) C^{-1} h(y,x)\right).
-%\end{equation}
-%Here, $h(y,x) = R(\theta(x)) (y-x)$, $R(\theta(x))$ is the 2D rotation
-%matrix by angle $\theta(x) = (x^1 + x^2) \pi / 2$, $C = L^2
-%\operatorname{diag}(c_1, c_2)$, $g(x) = x^1(1-x^1)x^2(1-x^2)$, and
-%$f(x,y) = \cos\left(p^1(y,x) / \sqrt{c_1 / 2.0}\right)
-%\sin\left(p^2(y,x) / \sqrt{c_2 / 2.0}\right)$ for some constants $a,
-%L, c_1, c_2$. In this section we use $a=1.0$, $L^2=1.0$, $c_1=0.01$,
-%and $c_2=0.0025$.
-%
 \begin{align}
@@
 
-%The impulse response $\phi_x$ corresponding to this kernel is a bumpy
-%blob that is centered near $x$, and is rotated and scaled in a manner
-%that depends on $x$ (see Figure
-%\ref{fig:frog_kernel_impulse_responses}). The constant $L$ controls
-%how wide the blob is (hence the support of the impulse response), and
-%$a$ controls the magnitude of the bumps.
-
-%Note that for this
-%operator, using our method is unnecessary because entries of the
-%kernel are given directly via a formula. The purpose of this example
-%is to illustrate the method using an operator which is known exactly
-%and may be easily visualized.
-
 \begin{table}
@@
 			\begin{tabular}{c| c | c c c c c}
-				& Error tol.    &	\#applies PSF   & \#applies HODLR 	& \#applies RSVD 	\\%& rank \\ 
+				& Error tol.    &	\#applies PSF   & \#applies HODLR 	& \#applies RSVD 	\\
 				\hline 
-				& 20\%    		& 11  				& 592				& 354				\\%& 111  \\ 
-				$L=1$ 			& 10\% 			& 16  				& 772				& 520				\\%& 178  \\ 
-				& 5\% 			& 22   				& 924				& 674				\\%& 244 \\ 
+				& 20\%    		& 11  				& 592				& 354				\\
+				$L=1$ 			& 10\% 			& 16  				& 772				& 520				\\
+				& 5\% 			& 22   				& 924				& 674				\\
 				\hline
-				& 20\%    		& 8  				& 852 				& 1316				\\%& 417 \\ 
-				$L=1/2$ 		& 10\%    		& 9  				& 1144				& 1916				\\%& 665 \\ 
-				& 5\% 	  		& 12   				& 1404				& 2456				\\%& 903 \\ 
+				& 20\%    		& 8  				& 852 				& 1316				\\
+				$L=1/2$ 		& 10\%    		& 9  				& 1144				& 1916				\\
+				& 5\% 	  		& 12   				& 1404				& 2456				\\
 				\hline
-				& 20\%    	 	& 7					& 932				& 2624				\\%& 845  \\ 
-				$L=1/3$ 		& 10\%    		& 8  				& 1264 				& 3734				\\%& 1321  \\ 
-				& 5\% 			& 8   				& 1520 				& 4660				\\%& 1759  \\ 
+				& 20\%    	 	& 7					& 932				& 2624				\\
+				$L=1/3$ 		& 10\%    		& 8  				& 1264 				& 3734				\\
+				& 5\% 			& 8   				& 1520 				& 4660				\\
 			\end{tabular}
@@
 	\caption{(Blur) Relative error 
-		%		in the PSF approximation of the kernel
-		%	$||\boldsymbol{\Phi} - \widetilde{\boldsymbol{\Phi}}||_\text{Fro}/||\boldsymbol{\Phi}||_\text{Fro}$,
 		for different ellipsoid size parameters, $\tau$, vs. the total number of impulse responses (left), and the number of batches (right). 
-		%	Left: Relative error vs. the total number of impulse responses used in the approximation. Right: Relative error vs. the number of impulse response batches. 
 		The dashed gray lines show linear convergence rates, i.e., $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$ on the left, and  $\text{constant} \times \# \text{batches}^{-1}$ on the right.}
@@
 The results in Figure \ref{fig:PSF_tau_convergence} (left) show that the relative error decreases as $\text{constant} \times \left(\# \text{impulse responses}\right)^{-1}$
-% or $\text{constant} \times \# \text{batches}^{-1}$, 
 suggesting linear convergence. The linear convergence stalls at a limit that depends on $\tau$. Increasing $\tau$ lowers this limit, allowing the PSF-based method to achieve higher accuracy. 
@@
 $\tau=2.5$).
-
-%----
-%, but allows the method to eventually achieve a lower level of error.
-%
-%----
-%
-%For smaller $\tau$ regime, we observe faster convergence
-%
-%Before this limit is reached, the convergence rate is the same for all $\tau$ as a function the number of impulse responses.
-This means that the convergence is faster for smaller $\tau$ as a function of the number of batches, since smaller $\tau$ means more impulse responses per batch.
-In summary, taking $\tau$ larger makes the method converge more slowly, but allows it to eventually achieve a lower level of error.
-
-%Define HODLR. Define randomized HODLR algo. Ref to Tucker's paper and "Fast randomized eigensolver with structured LDL..."
-
-%Define randomized SVD, refer to Halko Martinsson Tropp.
-
-%Discuss why PSF does better with smaller $L$.
 

### 2024-02-23 arxiv_final_version/localpsf.tex
add folder for arxiv version of paper

--- 2024-02-22 sisc_final_version/localpsf.tex
@@
 % SIAM Article Template
-\documentclass[review,onefignum,onetabnum]{siamart190516}
+\documentclass[onefignum,onetabnum]{siamart190516}
 
@@
 			& $T=2.0$ \\
-			$\kappa=1.0$e${-4}$  & \includegraphics[scale=0.12, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-4.png} 
-			& \includegraphics[scale=0.12, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-4.png} \\
-			$\kappa=1.0$e${-3}$ & \includegraphics[scale=0.12, valign=c]{impulse_response_P(0.3,0.6)_T0.5_K1.e-3.png} 
-			& \includegraphics[scale=0.12, valign=c]{impulse_response_P(0.3,0.6)_T2.0_K1.e-3.png}
+			$\kappa=1.0$e${-4}$  & \includegraphics[scale=0.12, valign=c]{impulse_response_P_0.3,0.6__T0.5_K1.e-4.png} 
+			& \includegraphics[scale=0.12, valign=c]{impulse_response_P_0.3,0.6__T2.0_K1.e-4.png} \\
+			$\kappa=1.0$e${-3}$ & \includegraphics[scale=0.12, valign=c]{impulse_response_P_0.3,0.6__T0.5_K1.e-3.png} 
+			& \includegraphics[scale=0.12, valign=c]{impulse_response_P_0.3,0.6__T2.0_K1.e-3.png}
 		\end{tabular}

### 2024-02-24 presentations/siamuq24-noemi/figures/stokes_geigs_figure.tex
noemis slides

--- 2023-06-01 paper/stokes_geigs_figure.tex
@@
 		%\addlegendentry{REG}
-		\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues_0.05noise.dat};
-		\addplot[dash dot, line width=0.5mm] table[x=k, y=none] {generalizedEigenvalues_0.05noise.dat};
+		\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {Stokes_generalizedEigenvalues_0.05noise.dat};
+		\addplot[dash dot, line width=0.5mm] table[x=k, y=none] {Stokes_generalizedEigenvalues_0.05noise.dat};
 		%\addlegendentry{PSF (1)}
-		\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues_0.05noise.dat};
+		\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {Stokes_generalizedEigenvalues_0.05noise.dat};
 		%\addlegendentry{PSF (5)}
-		\addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues_0.05noise.dat};
+		\addplot[red, line width=0.5mm] table[x=k, y=psf5] {Stokes_generalizedEigenvalues_0.05noise.dat};
 		%\addlegendentry{PSF (25)}
-		\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues_0.05noise.dat};
+		\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {Stokes_generalizedEigenvalues_0.05noise.dat};
 		\end{axis}

### 2024-02-24 presentations/siamuq24-noemi/localpsf_revised_figures/localpsf_mergebackup.tex
noemis slides

% SIAM Article Template
\documentclass[review,onefignum,onetabnum]{siamart190516}

% Information that is shared between the article and the supplement
% (title and author information, macros, packages, etc.) goes into
% ex_shared.tex. If there is no supplement, this file can be included
% directly.

\input{localpsf_shared}

% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={Fast matrix-free approximation of locally translation invariant operators that have locally supported non-negative integral kernels, with application to Hessians in PDE-constrained inverse problems},
  pdfauthor={N. Alger, N. Petra, T. Hartland, and O. Ghattas}
}
\fi

% The next statement enables references to information in the
% supplement. See the xr-hyperref package for details.

\externaldocument{localpsf_supplement}

% FundRef data to be entered by SIAM
%<funding-group specific-use="FundRef">
%<award-group>
%<funding-source>
%<named-content content-type="funder-name"> 
%</named-content> 
%<named-content content-type="funder-identifier"> 
%</named-content>
%</funding-source>
%<award-id> </award-id>
%</award-group>
%</funding-group>

\begin{document}

\maketitle

% REQUIRED
\begin{abstract}
	We present a fast matrix-free method for approximating operators that have locally supported non-negative integral kernels. Such operators arise, for example, as Schur complements in Schur complement methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained inverse problems. Our method computes impulse responses of the operator at a collection of scattered points, then interpolates these known impulse responses to evaluate entries of the integral kernel at other points where the impulse response is not known. 
%	Classical H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products. 
	Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources. By solving an ellipsoid packing problem, we choose as many points as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap. 
	The ability to rapidly evaluate kernel entries allows us to construct a hierarchical matrix (H-matrix) approximation of the operator. Fast H-matrix methods are used to perform further matrix computations including matrix-matrix addition and multiplication, matrix inversion and factorization, and matrix-vector products in nearly-linear complexity.
	We apply the method to approximate the Hessian in a large-scale PDE-constrained inverse problem with highly informative data. Numerical results demonstrate that our method substantially outperforms existing methods. We are able to approximate high rank operators using only a small number of operator applies.
\end{abstract}

% REQUIRED
\begin{keywords}

\end{keywords}

% REQUIRED
\begin{AMS}

\end{AMS}

\section{Introduction}
\label{sec:intro}

We present a fast \emph{matrix free} method for approximating operators, $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$, that have locally supported non-negative integral kernels. Here $\Omega \subset \mathbb{R}^\gdim$ is a bounded domain, $\gdim$ is the spatial dimension, and $L^2(\Omega)'$ is the space of continuous linear functionals on $L^2(\Omega)$ (i.e., the topological dual space). Such operators arise as Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{le1997non,saad1999distributed,TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps) \cite{DTNMAP}, covariance operators in spatial statistics \cite{GeogaEtAl20,ChenStein21,LindgrenRueLindstrom11}, blurring operators in imaging \cite{DenisEtAl11,NagyOleary98}, and Hessians in distributed parameter PDE-constrained optimization and inverse problems \cite{AlgerEtAl19,IsaacEtAl15}.

By ``matrix free,'' we mean that we have a black-box computational procedure through which we may apply $\Aop$ and its transpose\footnote{Recall that the transpose $\Aop^T:L^2(\Omega)\rightarrow L^2(\Omega)'$ is the unique operator satisfying $\left(\Aop u\right)(v) = \left(\Aop v\right)(u)$ for all $u,v \in L^2(\Omega)$, where the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(v)$ is the result of applying that linear functional to $v \in L^2(\Omega)$, and similar for the analogous operations with $\Aop^T$. }, $\Aop^T$, to arbitrary functions,
\begin{equation}
\label{eq:eval_maps}
u \mapsto\Aop u \qquad \text{and} \qquad v \mapsto \Aop^Tv,
\end{equation} 
but we do not have any further access to $\Aop$. In particular, we cannot easily access entries of $\Aop$'s integral kernel. In applications, evaluating the maps in \eqref{eq:eval_maps} may involve iteratively solving a large linear system, timestepping, or performing some other costly computational procedure.

Our motivation is Hessian approximation in distributed parameter inverse problems governed by partial differential equations (PDEs). That is, inverse problems in which one seeks to reconstruct an unknown parameter field from noisy observations of a state variable that depends on the parameter implicitly through the solution of a PDE.
In the deterministic approach to the inverse problem, one solves a regularized minimization problem to find the parameter that best fits the data. The objective function in the minimization problem is defined implicitly, so information about the Hessian of the objective function is only accessible via application of the Hessian to vectors \cite{HESSIANACTION}. 
The most popular matrix-free Hessian approximation methods are based on low rank approximation \cite{BuiEtAl13,CuiEtAl14,FlathEtAl11,PetraEtAl14,SpantiniEtAl15} because there are efficient methods for constructing low rank approximations of a matrix using only matrix-vector products \cite{Cheng05,HalkoMartinssonTropp11}. 
However, low rank approximation methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of operator applies are required to form an accurate ``low rank'' approximation \cite{Alger19,AlgerEtAl17}. High-rank approximation methods are needed.

\begin{figure}
	\begin{subfigure}{0.32\textwidth}
		\centering
		\includegraphics[scale=0.33]{impulse_batch1.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\centering
		\includegraphics[scale=0.33]{impulse_batch2.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\centering
		\includegraphics[scale=0.33]{impulse_batch3.png}
	\end{subfigure}
	\caption{Batches of normalized impulse responses for the Stokes inverse problem Hessian. 
%	Brightness scale differs between the subfigures. 
	Black stars indicate point source locations. Dashed gray ellipses indicate the estimated support ellipsoids. The large circle is the boundary of the domain. 
	}
	\label{fig:batches_intro}
\end{figure}

The central challenge we face is that existing scalable high-rank approximation methods typically require the ability to rapidly evaluate arbitrary entries of $\Aop$'s integral kernel, and we cannot do this because $\Aop$ is only accessible matrix-free. We address this challenge using \emph{impulse response interpolation}. The impulse response, $\phi_x$, associated with a point $x$ is the Riesz representation\footnote{Recall that the Riesz representative of a functional $\rho \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\rho^* \in L^2(\Omega)$ such that $\rho(v) = \left(\rho^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} of the linear functional that results from applying $\Aop$ to a delta distribution (i.e., point source, impulse) centered at $x$. We compute batches of impulse responses by applying $\Aop$ to weighted sums of delta distributions associated with batches of points scattered throughout the domain (see Figure \ref{fig:batches_intro}). Then we interpolate translated versions of these impulse responses to approximate arbitrary entries of the operator's integral kernel. 
%This allows fast evaluation of approximate matrix entries, which are needed by high-rank approximation methods, in particular, classical H-matrix construction methods.

Picking the batches of points requires us to estimate the supports of the $\phi_x$, \emph{before} we compute them. The idea of estimating the supports of the functions $\phi_x$ a-priori was inspired by techniques from resolution analysis in seismic imaging, in which the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $x$, where $\zeta$ is a random noise funtion  \cite{FichtnerLeeuwen15,TrampertFichtnerRitsema13}. Rather than probing $\Aop^T$ with random noise functions, we use polynomial functions (see Section \ref{sec:intromoments}). Our method estimates the support of $\phi_x$ more accurately and reliably than random noise probing at the cost of the additional constraint that $\Aop$ has a positive integral kernel.

Our impulse response interpolation procedure may be categorized as a ``point spread function'' (PSF) method that is loosely based on so-called ``product-convolution'' (PC) approximations, which are approximations of an operator by weighted sums of convolution operators with spatially varying weights. PC and PSF methods have a long history dating back several decades. We note the following papers (among many others),  \cite{Adorf94,AlgerEtAl19,BigotEscandeWeiss19,NagyOleary98,FishEtAl96,EscandeWeiss12,EscandeWeiss15,EscandeWeiss22,ZhuLiFomelEtAl16}, in which the convolution kernels are constructed from sampling impulse responses of the operator to scattered point sources. For background on PC and PSF methods, we recommend the following papers which provide well-written reviews of the topic: \cite{DenisEtAl15,EscandeWeiss17,GentileCourbinMeylan13}. While PC approximations are based on an assumption of local translation invariance, the method we propose is based on an assumption that we call ``local mean displacement invariance'' (see Figure \ref{fig:mean_displacement_invariance}). This is more general than local translation invariance, and includes both low rank approximation and local translation invariance as special cases (see Section \ref{sec:local_mean_displacement_invariance}). Furthermore, PC approximations have well-known issues near boundaries due to so-called ``boundary artifacts,'' caused by the boundary artificially ``cutting off'' impulse responses. We avoid this issue by temporarily excluding impulse responses from the approximation procedure whenever an attempt is made to evaluate the impulse response at a location where it is undefined.

%\begin{figure}
%	\begin{center}
%		\includegraphics[scale=0.4]{bct1.eps}
%	\end{center}
%	\caption{Hierarchical matrix block structure for the Stokes Hessian. The red blocks (tiny) are dense. All other blocks are low rank.
%	}
%	\label{fig:hmatrix_intro}
%\end{figure}

The ability to rapidly evaluate approximations of arbitrary entries of $\Aop$'s integral kernel allows us to form a hierarchical matrix (H-matrix) \cite{Hackbusch99,} approximation of a discretized version of $\Aop$. H-matrices are a matrix format in which the rows and columns of the matrix are re-ordered, then the matrix is recursively subdivided into blocks, in such a way that many off-diagonal blocks are low rank, even though the matrix as a whole may be high-rank. 
%(see Figure \ref{fig:hmatrix_intro} and Appendix \ref{app:h_matrix}).
H-matrix methods permit us to perform matrix-vector products cheaply. Furthermore, one may use H-matrix methods to perform useful linear algebra operations involving the H-matrix that cannot easily be done using the original operator. These operations include matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion. 
%These H-matrix methods are fast and scale well to large problems. 
The work and memory required to perform these operations for an $\fedim \times \fedim$ H-matrix is scales as $O(\fedim \log(\fedim)^a)$, where $a \in \{0,1,2,3\}$. That is, the cost is \emph{nearly-linear} in $N$. The exact cost depends on the type of H-matrix used, the operation being performed, and the rank of the off-diagonal blocks; for more details see \cite{GrasedyckHackbusch03}. The ability to perform these operations permits, for example, fast solution of Newton linear systems in PDE-constrained optimization, fast sampling of ill-conditioned posterior distribution in Bayesian inverse problems, and construction of high rank surrogate models that can be used for uncertainty quantification.

%Classical H-matrix construction methods 
%(described in Appendix \ref{sec:H_matrix_construction}) 
%require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available matrix free.
Recently there have been improvements in matrix-free H-matrix construction methods based on the so-called ``peeling process'' \cite{LinYing11,Martinsson11,Martinsson16,MartinssonTropp20}, which we do not use here. 
These alternative methods have been applied to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{AmbartsumyanEtAl20,HartlandEtAl22}. 
Methods based on the peeling process are asymptotically scalable (typically the number of matrix vector products required scales as $O(\log N)$), but in practice the required number of matrix vector products is large. Improving the peeling process to reduce its cost is currently an area of active research \cite{PEELINGRESEARCH}.
%This is an area of active research, so more efficient and practical matrix-free H-matrix construction methods based on the peeling process may be discovered in the future.
%Matrix-free H-matrix construction methods are currently a topic of considerable research, so more efficient matrix-free H-matrix methods based on the peeling process may be discovered in the future.

\begin{figure}
	\begin{center}
		\includegraphics[scale=0.85]{mean_displacement_invariance.pdf}
	\end{center}
	\caption{An operator is locally mean-displacement invariant if $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y - \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to $x'$. Here $\spatialmean(x)$ is the mean (center of mass) of $\phi_x$, and similar for $\spatialmean(x')$.}
	\label{fig:mean_displacement_invariance}
\end{figure}

\section{Preliminaries}

Let $\Omega \subset \mathbb{R}^\gdim$ be a bounded domain (in applications typically $\gdim=1$, $2$, or $3$). We seek to approximate integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ that may be written in the following form:
\begin{equation}
\label{eq:kernel_representation}
(\Aop u)(v) := \int_\Omega \int_\Omega v(y) \Aker(y,x) u(x) dx dy.
\end{equation}
Here the linear functional $\Aop u \in L^2(\Omega)'$ is the result of applying $\Aop$ to $u\in L^2(\Omega)$, the scalar $\left(\Aop u\right)(v)$ is the result of applying that linear functional to $v \in L^2(\Omega)$.
The function $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel, which exists in principle but is not easily accessible to us. In this section we describe how to extend the action of $\Aop$ to distributions, which allows us to define the impulse responses (Section \ref{sec:impulse_response_2}). Then we state the conditions on $\Aop$ that our method requires (Section \ref{sec:conditions_2}). Finally, we detail finite element discretization (Section \ref{sec:finite_element_kernel}).

\subsection{Distributions and impulse responses}
\label{sec:impulse_response_2}

The action of $\Aop$ may be extended to distributions if the kernel $\Aker$ is sufficiently regular. We derive the extension as follows. Suppose $\rho \in L^2(\Omega)'$, and let $\rho^* \in L^2(\Omega)$ denote the Riesz representative of $\rho$ with respect to the $L^2(\Omega)$ inner product. We have
\begin{subequations}
	\label{eq:extension_to_distributions}
	\begin{align}
		\left(\Aop \genericdistribution^*\right)(v) &= \int_\Omega \int_\Omega v(y) \Aker(y,x) \genericdistribution^*(x) dx ~dy \\
		&= \int_\Omega v(y) \int_\Omega \Aker(y,x) \genericdistribution^*(x) dx ~dy \\
		&= \int_\Omega v(y) \genericdistribution\left(\Aker(y,\cdot)\right) dy, \label{eq:action_on_distribution}
	\end{align}
\end{subequations}
where $\Aker(y,\cdot)$ denotes the function $x \mapsto \Aker(y,x)$.
Now let $\mathcal{D}(\Omega) \subset L^2(\Omega)$ be a suitable space of test functions, and instead suppose $\rho:\mathcal{D}(\Omega) \rightarrow \mathbb{R}$ is a distribution. In this case, the Riesz representative $\rho^*$ may not exist, so the derivation in \eqref{eq:extension_to_distributions} is not valid. However, if $\Aker$ is sufficiently regular such that the function $y \mapsto \rho\left(\Aker(y,~\cdot~)\right)$ is well-defined for almost all $y\in\Omega$, and if this function is in $L^2(\Omega)$, then the right hand side of \eqref{eq:action_on_distribution} is well-defined. Hence, we \emph{define} the action of $\Aop$ on the distribution $\rho$ to be the right hand side of \eqref{eq:action_on_distribution}. For convenience of notation, we denote this action by ``$\Aop \genericdistribution^*$,'' even if $\rho^*$ does not exist.

Let $\delta_x$ denote the delta distribution\footnote{Recall that the delta distribution $\delta_x:\mathcal{D}(\Omega)\rightarrow \mathbb{R}$ is the linear functional defined by $\delta_x(v) = v(x)$ for all $v\in \mathcal{D}(\Omega)$.} (i.e., point source, impulse) centered at the point $x \in \Omega$.
The \emph{impulse response} of $\Aop$ associated with $x$ is the function $\impulseresponse_x:\Omega \rightarrow \mathbb{R}$, 
\begin{equation}
\label{eq:impulse_response_delta_action}
\impulseresponse_x := \left( \Aop \delta_x^* \right)^*,
\end{equation}
that is formed by applying $\Aop$ to $\delta_x$, then taking the Riesz representation of the resulting linear functional. Using \eqref{eq:action_on_distribution} and the definition of the delta distribution, we see that $\phi_x$ may also be written as the function
\begin{equation}
\label{eq:impulse_response_defn}
\impulseresponse_x(y) = \Aker(y, x),
\end{equation} 
which results from keeping $x$ fixed and viewing $\Aker(y,x)$ as a function of $y$.

\subsection{Required conditions}
\label{sec:conditions_2}

We focus on approximating operators that satisfy the following conditions: 
\begin{enumerate}
	\item The kernel $\Aker$ is sufficiently regular such that $\phi_x$ is a well-defined function for all $x\in \Omega$.
	\item The supports of the impulse responses $\impulseresponse_x$ are contained in localized regions.
	\item The integral kernel is non-negative in the sense that $$\Aker(y,x) \ge 0$$ for all $(y,x) \in \Omega \times \Omega$.\footnote{Note that having a non-negative integral kernel is different from positive semi-definiteness. The operator $\Aop$ need not be positive semi-definite to use our method, and positive semi-definite operators need not have a non-negative integral kernel.}
\end{enumerate}
Our method may still perform well if these conditions are relaxed slightly. For example, it is fine if the support of the impulse response $\impulseresponse_x$ is not perfectly contained in a localized region, so long as the bulk of the ``mass'' of the impulse response is contained in a localized region. The integral kernel does not need to be non-negative for all pairs of points $(y,x) \in \Omega \times \Omega$, as long as it is non-negative for the vast majority of pairs of points $(y,x)$, and as long as the negative numbers are comparatively small. If these conditions are violated, our method will incur additional error. If these conditions are violated too much, our method may fail. Not all operators of interest (and in particular not all Hessians) satisfy these conditions, but many operators of practical interest do satisfy these conditions, and our method is highly effective for approximating these operators.

\subsection{Finite element discretization}
\label{sec:finite_element_kernel}

In computations, functions are discretized and replaced by finite dimensional vectors, and operators mapping between infinite dimensional spaces are replaced by operators mapping between finite dimensional spaces. In this paper we discretize using continuous Galerkin finite elements satisfying the Kronecker property (defined below). With minor modifications, our method could be used with more general finite element methods, or other discretization schemes such as finite differences or finite volumes.

Let $\febasis_1, \febasis_2, \dots, \febasis_\fedim$ be a set of continuous Galerkin finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, let $V_h := \Span\left(\febasis_1, \febasis_2, \dots, \febasis_\fedim\right)$ be the corresponding finite element space under the $L^2$ inner product, and let $p_i \in \mathbb{R}^\gdim$, $i=1,\dots, \fedim$ be the Lagrange nodes associated with the functions $\febasis_i$. We assume that the finite element basis satisfies the Kronecker property, i.e., $\febasis_i(p_i)=1$ and $\febasis_i(p_j)=0$ if $i\neq j$. 
For $u_h \in V_h$ we write $\bm{u} \in \mathbb{R}^\nsamplepts_\massmatrix$ to denote the coefficient vector for $u_h$ with respect to the finite element basis, i.e.,
\begin{equation}
\label{eq:fem_coeff_basis}
u_h(x) = \sum_{i=1}^\fedim \bm{u}_i \febasis_i(x).
\end{equation}
Linear functionals $\rho_h \in V_h'$ have coefficient dual vectors $\boldsymbol{\rho}\in \mathbb{R}^\nsamplepts_{\massmatrix^{-1}}$, with entries 
\begin{equation*}
\boldsymbol{\rho}_i = \rho_h(\febasis_i), \qquad i=1,\dots,\nsamplepts. 
\end{equation*}
Here $\massmatrix \in \mathbb{R}^{\fedim \times \fedim}$ denotes the sparse finite element mass matrix which has entries
\begin{equation*}
\massmatrix_{ij}=\int_\Omega \febasis_i(x) \febasis_j(x) dx, \qquad i,j=1,\dots,\fedim,
\end{equation*}
the space $\mathbb{R}^\fedim_\massmatrix$ is $\mathbb{R}^\fedim$ with the inner product $(\mathbf{u},\mathbf{v})_\massmatrix := \mathbf{u}^T \massmatrix \mathbf{v}$, and $\mathbb{R}^\fedim_{\massmatrix^{-1}}$ is the analogous space with $\massmatrix^{-1}$ replacing $\massmatrix$. Direct calculation shows that $\mathbb{R}^\fedim_\massmatrix$ and $\mathbb{R}^\fedim_{\massmatrix^{-1}}$ are isomorphic to $V_h$ and $V_h'$ as Hilbert spaces, respectively.

After discretization, the operator $\Aop:L^2(\Omega) \rightarrow L^2(\Omega)'$ is replaced by an operator $A_h:V_h \rightarrow V_h'$, which becomes an operator
\begin{equation*}
\mathbf{A}:\mathbb{R}^\fedim_\massmatrix \rightarrow \mathbb{R}^\fedim_{\massmatrix^{-1}}
\end{equation*}
under the isomorphism discussed above. Our method is agnostic to the computational procedure for approximating $\Aop$ with $\mathbf{A}$. What is important is that we do not have direct access to matrix entries $\mathbf{A}_{ij}$. Rather, we have a computational procedure that allows us to compute matrix-vector products $\mathbf{u}\mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{v}\mapsto \mathbf{A}^T\mathbf{v}$, and computing these matrix-vector products is costly.
Of course, matrix entries can be computed via matrix-vector products as $\mathbf{A}_{ij} = \left(\mathbf{A}\mathbf{e}_j\right)_i$,
where $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)^T$ is the length $\fedim$ unit vector with one in the $j$th coordinate and zeros elsewhere. But computing the matrix-vector product $\mathbf{e}_j \mapsto \mathbf{A}\mathbf{e}_j$ is costly, and therefore wasteful if we do not use other matrix entries in the $j$th column of $\mathbf{A}$. Hence, methods for approximating $\mathbf{A}$ are computationally intractable if they require accessing scattered matrix entries from many different rows and columns of $\mathbf{A}$.

The operator $A_h:V_h \rightarrow V_h'$ can be written in integral kernel form as
\begin{equation}
(A_h u_h)(v_h) := \int_\Omega \int_\Omega v_h(y) \Aker_h(y,x) u_h(x) dx dy
\end{equation}
for some kernel $\Aker_h$ which we do not know, and which may differ from $\Aker$ due to discretization error. Since the functions in $V_h$ are continuous at $x$, the delta distribution $\delta_x$ is a continuous linear functional on $V_h$, which has a discrete dual vector $\boldsymbol{\delta}_x \in \mathbb{R}^\fedim_{\massmatrix^{-1}}$ with entries $\left(\boldsymbol{\delta}_x\right)_i = \febasis_i(x)$ for $i=1,\dots,\fedim$. Additionally, it is straightforward to verify that the Riesz representation, $\rho_h^* \in V_h$, of a functional $\rho \in V_h'$ has coefficient vector
\begin{equation*}
\boldsymbol{\rho}^* = \massmatrix^{-1} \boldsymbol{\rho}.
\end{equation*}
Therefore, the formula for the impulse response from \eqref{eq:impulse_response_delta_action} becomes 
\begin{equation}
\label{eq:discrete_kernel}
\boldsymbol{\impulseresponse}_x = \left(A_h \delta_x^*\right)^* =  \massmatrix^{-1}\mathbf{A} \massmatrix^{-1} \boldsymbol{\delta}_x,
\end{equation}
and the $(y,x)$ kernel entry of $\Aker_h$ may be written as
\begin{equation}
\label{eq:approx_kernel_entry_81}
	\Aker_h(y,x) = \boldsymbol{\delta}_y^T \boldsymbol{\impulseresponse}_x = \boldsymbol{\delta}_y^T \massmatrix^{-1}\mathbf{A} \massmatrix^{-1} \boldsymbol{\delta}_x.
\end{equation}
Now define $\mathbf{\Aker} \in \mathbb{R}^{\fedim \times \fedim}$ to be the following dense matrix of kernel entries evaluated at all pairs of Lagrange nodes:
\begin{equation}
\label{eq:Akerpcmat_entries}
\mathbf{\Aker}_{ij} := \Aker_h(p_i, p_j).
\end{equation}
Because of the Kronecker property of the finite element basis, we have $\boldsymbol{\delta}_{p_i} = \mathbf{e}_i$. Thus from \eqref{eq:approx_kernel_entry_81} we have
$
	\Aker_h(p_i,p_j) = \left(\massmatrix^{-1}\mathbf{A} \massmatrix^{-1}\right)_{ij},
$
which implies
\begin{equation}
\label{eq:boldA}
	\mathbf{A} = \massmatrix \mathbf{\Aker} \massmatrix.
\end{equation}
Broadly, our method will construct an H-matrix approximation of $\mathbf{A}$ by forming H-matrix approximations of $\boldsymbol{\Aker}$ and $\massmatrix$ (or a lumped mass version of $\massmatrix$), then multiplying these matrices with H-matrix methods per \eqref{eq:boldA}. The central challenge is that classical H-matrix construction methods require access to arbitrary matrix entries $\mathbf{\Aker}_{ij}$, but these matrix entries are not easily accessible. The bulk of our method is therefore dedicated to forming approximations of these matrix entries that can be evaluated rapidly.

\paragraph{Lumped mass matrix} At the continuum level, the integral kernel $\Aker$ is assumed to be non-negative. However, when we compute impulse responses we are effectively operating with the discrete integral kernel shown in \eqref{eq:approx_kernel_entry_81}.
The inverse mass matrices in this formula, $\massmatrix^{-1}$, will typically contain negative numbers. If there are too many negative numbers, or if the negative numbers are too large, our algorithm will be less robust. We therefore recommend replacing the mass matrix $\massmatrix$ with a positive diagonal \emph{lumped mass} approximation. 
In our numerical results, we use the lumped mass matrix constructed by replacing off-diagonal entries of the mass matrix by zero. Other more advanced mass lumping techniques may be used. See \cite{Hughes12} for a discussion of lumped mass matrices.

\section{Key innovations}
\label{sec:prerequisites}

In this section we present two key innovations that our method is based on.
First, we define moments of the impulse responses, $\phi_x$, show how these moments can be computed efficiently, and use these moments to form ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{sec:intromoments}). Our method will use these ellipsoid estimates to choose sets of impulse responses that can be computed in batches. Second, we describe an improved method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}). Our method will interpolate impulse responses using this approximation.

\subsection{Impulse response moments and ellipsoid support estimate}
\label{sec:intromoments}

The impulse response $\impulseresponse_x$ may be interpreted as a scaled probability distribution because of the non-negative integral kernel property. Let $\spatialvol:\Omega \rightarrow \mathbb{R}$,
\begin{equation}
\label{eq:define_vol}
\spatialvol(x) := \int_\Omega \impulseresponse_x(y) dy
\end{equation}
denote the spatially varying scaling factor, and for $i,j=1,\dots,\gdim$ define $\spatialmean:\Omega \rightarrow \mathbb{R}^\gdim$ and $\spatialcov:\Omega \rightarrow \mathbb{R}^{\gdim \times \gdim}$ as follows:
\begin{align}
\spatialmean^i(x) :=& \int_\Omega (\impulseresponse_p(y) / V(x)) y^i ~dy \label{eq:define_mean} \\
\spatialcov^{ij}(x) :=& \int_\Omega (\impulseresponse_p(y) / V(x)) \left(y^i - \spatialmean^i(x)\right) \left(y^j - \spatialmean^j(x)\right) ~dy, \label{eq:define_cov}
\end{align}
where $\spatialmean^i(x)$ denotes the $i^\text{th}$ component of the vector $\spatialmean(x)$, and $\spatialcov^{ij}(x)$ denotes the $(i,j)$ entry of the matrix $\spatialcov(x)$.
The vector $\spatialmean(x)\in \mathbb{R}^\gdim$ and the matrix $\spatialcov(x) \in \mathbb{R}^{\gdim \times \gdim}$ are the mean and covariance of the normalized version of $\impulseresponse_x$, respectively.

The direct approach to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ is to apply $\Aop$ to a point source centered at $x$ to obtain $\impulseresponse_x$, as per \eqref{eq:impulse_response_delta_action}. Then one can post-process $\impulseresponse_x$ to determine $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$. However, this direct approach is not feasible because we need to know $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ before we compute $\phi_x$, in order choose the point $x$. Computing $\phi_x$ in order to determine $V(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ would be extremely computationally expensive, and defeat the purpose of our algorithm, which is to reduce the computational cost by computing impulse responses in batches. Fortunately, it is possible to compute $\spatialvol(x)$, $\spatialmean(x)$, and $\spatialcov(x)$ indirectly, \emph{for all points $x\in\Omega$ simultaneously}, by applying $\Aop^T$ to one constant function, $\gdim$ linear functions, and $\gdim(\gdim+1)/2$ quadratic functions (e.g., 6 total operator applies in two spatial dimensions and 10 in three spatial dimensions). This may be motivated by analogy to matrices. If $\mathbf{B}\in \mathbb{R}^{\fedim \times \fedim}$ is a matrix with $i^\text{th}$ column $\mathbf{b}_i$ and $\mathbf{f} \in \mathbb{R}^\fedim$, then
\begin{equation*}
\mathbf{B}^T \mathbf{f} = \begin{bmatrix}
\horzbar & \mathbf{b}_1^T & \horzbar \\
\horzbar & \mathbf{b}_2^T & \horzbar \\
& \vdots & \\
\horzbar & \mathbf{b}_\fedim^T & \horzbar
\end{bmatrix}
\mathbf{f} = 
\begin{bmatrix}
\mathbf{b}_1^T \mathbf{f} \\
\mathbf{b}_2^T \mathbf{f} \\
\vdots \\
\mathbf{b}_\fedim^T \mathbf{f}
\end{bmatrix}.
\end{equation*}
By computing one matrix-vector product of $\mathbf{B}^T$ with $\mathbf{f}$, we compute the inner product of each column of $\mathbf{B}$ with $\mathbf{f}$ simultaneously. The operator case is analogous, with the impulse response $\phi_x$ taking the place of a matrix column. We have
\begin{equation}
\label{eq:operator_simultaneous_ips}
\left(\Aop^T f\right)^*(x) = \int_\Omega \Aker(y,x) f(y) dy = \left(\phi_x, f\right)_{L^2(\Omega)}.
\end{equation}
By computing one operator application of $\Aop^T$ to $f$, we compute the inner product of each impulse response $\phi_x$ with $f$, for all points $x$ simultaneously.

Let $C$, $L^i$, and $Q^{ij}$ be the following constant, linear, and quadratic functions:
\begin{equation*}
C(x) := 1, \qquad
L^i(x) := x^i, \qquad
Q^{ij}(x) := x^i x^j
\end{equation*}
for $i,j=1,\dots,\gdim$. Using the definition of $\spatialvol$ in \eqref{eq:define_vol} and using \eqref{eq:operator_simultaneous_ips}, we have
\begin{equation*}
	\spatialvol(x) = \int_\Omega \phi_x(y) C(y) ~ dy = \left(\phi_x, C\right)_{L^2(\Omega)} = \left(\mathcal{A}^T C\right)^*(x).
\end{equation*}
Hence, we may compute $\spatialvol(x)$ for all $x$ simultaneously by applying $\Aop^T$ to $C$. Analogous manipulations show that $\spatialmean(x)$ and $\spatialcov(x)$ may be computed for all points $x$ simultaneously by applying $\Aop^T$ to the functions $L^i$ and $Q^{ij}$, respectively. To summarize the results of doing this, we have
\begin{subequations}
	\label{eq:vol_mean_var}
	\begin{align}
	\spatialvol =& \left(\Aop^T C\right)^* \\
	\spatialmean^i =& \left(\Aop^T L^i\right)^* / \spatialvol \\
	\spatialcov^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \spatialvol - \spatialmean^i\cdot \spatialmean^j
	\end{align}
\end{subequations}
for $i,j=1,\dots, \gdim$. Here $f/g$ denotes pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ denotes pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
%The resulting method for computing discretized versions of $\spatialvol$, $\spatialmean$, and $\spatialcov$ is shown in Algorithm \ref{alg:varhpi_mean_cov}.

%\begin{algorithm2e}
%	\SetAlgoNoLine
%	\SetKwInOut{Input}{Input}
%	\SetKwInOut{Output}{Output}
%	{	
%		\Input{Operator $\bm{A}^T$}
%		\Output{$\bm{\spatialvol}\in \mathbb{R}^\fedim$, $\bm{\spatialmean}\in \mathbb{R}^{\gdim \times \fedim}$, and $\bm{\spatialcov} \in \mathbb{R}^{\gdim \times \gdim \times \fedim}$}
%		
%		Form vector $\bm{C} \in \mathbb{R}^\fedim_{\massmatrix}$ by either projecting or interpolating constant function $C(x)=1$ onto $V_h$
%		
%		$\bm{\spatialvol} \gets \massmatrix^{-1}\bm{A}^T \bm{C}$
%		
%		\For{$i=1,2,\dots,\gdim$}{
%			Form vector $\bm{L}^i \in \mathbb{R}^\fedim_{\massmatrix}$ by either projecting or interpolating linear function $L^i(x) = x^i$ onto $V_h$
%			
%			$\bm{\spatialmean}^i \gets \left(\massmatrix^{-1}\bm{A}^T \bm{L}^i\right) / \bm{\spatialvol}$
%		}
%		\For{$i=1,2,\dots,\gdim$}{
%			\For{$j=1,\dots,i$}{
%				Form quadratic function $\bm{Q}^{ij} = x^i x^j$
%				
%				$\bm{\spatialcov}^{ij} \gets \left(\massmatrix^{-1}\bm{A}^T \bm{Q}^{ij}\right) / \bm{\spatialvol} - \bm{\spatialmean}^i\cdot \bm{\spatialmean}^j$
%				
%				$\bm{\spatialcov}^{ji} \gets \bm{\spatialcov}^{ij}$
%				
%			}
%		}
%		
%	}
%	\caption{Compute discretized scaling factor $\mathbf{\spatialvol}$, mean $\boldsymbol{\spatialmean}$, and covariance $\boldsymbol{\spatialcov}$}
%	\label{alg:varhpi_mean_cov}
%\end{algorithm2e}

Next, we make the approximation that the support of $\impulseresponse_x$ is contained within the ellipsoid
\begin{equation}
\label{eq:support_ellipsoid}
E_x := \{x' \in \Omega: (x' - \spatialmean(x))^T \spatialcov(x)^{-1} (x' - \spatialmean(x)) \le \tau^2\},
\end{equation}
where $\tau$ is a fixed constant. The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, i.e., the Gaussian distribution which has the same mean and covariance as the normalized version of $\impulseresponse_x$.

The quantity $\tau$ is a parameter that must be chosen appropriately. The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$. However, in Section \ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (the more ellipsoids the better), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$. In practice, we find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results. The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.

\subsection{Local mean-displacement invariance}
\label{sec:local_mean_displacement_invariance}

Let $x$ and $x'$ be points in $\Omega$ that are ``close'' to each other, and consider the following approximations:
\begin{align}
\phi_x(y) &\approx \phi_{x'}(y) \label{eq:translate1}\\
\phi_x(y) &\approx \phi_{x'}(y-x+x') \label{eq:translate2}\\
\phi_x(y) &\approx \phi_{x'}\left(y-\spatialmean(x)+\spatialmean(x')\right) \label{eq:translate3}\\
\phi_x(y)/\spatialvol(x) &\approx \phi_{x'}\left(y-\spatialmean(x)+\spatialmean(x')\right) / \spatialvol(x'). \label{eq:translate4}
\end{align}
These are four different ways to approximate an impulse response by a nearby impulse response, with each successive approximation building upon the previous ones. Our method uses \eqref{eq:translate4}, which is the most sophisticated of these approximations.

Approximation \eqref{eq:translate1} says that $\phi_x$ can be approximated by $\phi_{x'}$ when $x$ and $x'$ are close. Operators satisfying \eqref{eq:translate1} can be well-approximated via low-rank CUR approximation \cite{CUR,NYSTROM}. However, the required rank in the low rank approximation can be large, which makes algorithms based on \eqref{eq:translate1} expensive.

Operators that satisfy \eqref{eq:translate2} are called ``locally translation-invariant'' because integral kernel entries $\Aker(y,x)$ for such operators are approximately invariant under translation of $x$ and $y$ by the same displacement, i.e., $x \rightarrow x+h$ and $y \rightarrow y+h$. It is straightforward to show that if equality holds in \eqref{eq:translate2}, then $\Aop$ is a convolution operator. Locally translation-invariant operators act like convolutions locally, and can therefore be well-approximated by PC approximations.

Approximation \eqref{eq:translate3} improves upon \eqref{eq:translate1} and \eqref{eq:translate2}, and generalizes both. On one hand, if \eqref{eq:translate1} holds, then $\spatialmean(x) \approx \spatialmean(x')$, and so \eqref{eq:translate3} holds. On the other hand, translating a distribution translates its mean, so if \eqref{eq:translate2} holds, then $\spatialmean(x')-\spatialmean(x) \approx x' - x$, so again \eqref{eq:translate3} holds. But approximation \eqref{eq:translate3} can hold in situations where neither \eqref{eq:translate1} nor \eqref{eq:translate2} holds. For example, because the expected value commutes with affine transformations, \eqref{eq:translate3} will hold when an $\Aop$ is locally translation-invariant with respect to a rotated frame of reference, while \eqref{eq:translate2} will not hold in this case. Additionally, \eqref{eq:translate3} generalizes to operators $\Aop:L^2(\Omega_1) \rightarrow L^2(\Omega_2)'$ that map between function spaces on different domains $\Omega_1$ and $\Omega_2$, and even operators that map between domains with different spatial dimensions. In contrast, \eqref{eq:translate2} does not naturally generalize to operators that map between function spaces on different domains, because the formula $y-x+x'$ requires vectors in $\Omega_2$ and $\Omega_1$ to be added together.  We call \eqref{eq:translate3} ``local mean-displacement invariance,'' and illustrate \eqref{eq:translate3} in Figure \ref{fig:mean_displacement_invariance}.

We use approximation \eqref{eq:translate4}, which is the same as \eqref{eq:translate3}, except for the extra factors of $1/\spatialvol(x)$ on the left hand side and $1/\spatialvol(x')$ on the right hand side. These factors make the approximation more robust if $\spatialvol(x)$ varies widely. Approximation \eqref{eq:translate4} is equivalent to \eqref{eq:translate3}, but with $\phi_x$ replaced by the its normalized version, $\phi_x/\spatialvol(x)$. We call \eqref{eq:translate4} ``normalized local mean-displacement invariance.''

\section{Operator approximation algorithm}
\label{sec:method}

Algorithm \ref{alg:varhpi_mean_cov} from Section \ref{sec:intromoments} allows us to compute $\spatialvol$, $\spatialmean$, and $\spatialcov$ by applying $\Aop^T$ to a small number of polynomial functions, and \eqref{eq:support_ellipsoid} uses $\spatialvol$, $\spatialmean$, and $\spatialcov$ to form an ellipsoid shaped estimate for the support of $\phi_x$, \emph{without} computing $\phi_x$. This allows us to compute large numbers of impulse responses, $\impulseresponse_{x_i}$, in ``batches,'' $\eta_b$ (see Figure \ref{fig:batches_intro}). We compute the impulse response batch $\eta_b$ by applying $\Aop$ to a weighted sum of point sources (Dirac comb) associated with a batch, $S_b$, of points $x_i$ scattered throughout $\Omega$ (Section \ref{sec:get_impulse_response}). The batch of points $S_b$ is chosen via a greedy ellipsoid packing algorithm so that, for $x_i,x_j \in S_b$, the support of $\impulseresponse_{x_i}$ and the support of $\impulseresponse_{x_j}$ do not overlap (or do not overlap much) if $i \neq j$ (Section \ref{sec:sample_point_selection}). Because these supports do not overlap, we can post-process $\eta_b$ to recover the functions $\impulseresponse_{x_i}$ associated with all points $x_i \in S_b$---with one application of $\Aop$, we recover many $\impulseresponse_{x_i}$ (Section \ref{sec:get_impulse_response}). The process is repeated to get more batches, until a desired number of impulse responses is reached.

Once the impulse response batches $\eta_b$ are computed, we approximate the integral kernel $\Aker(y,x)$ at arbitrary points $(y,x)$ by interpolation (Section \ref{sec:approximate_kernel_entries}). The key idea behind the interpolation is the normalized local mean displacement invariance assumption discussed in Section \ref{sec:local_mean_displacement_invariance}. We approximate $\Aker(y,x) = \phi_x(y)$ by a weighted linear combination of the values $\frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}(y - \spatialmean(x) + \spatialmean(x_i))$ for a collection of sample points $x_i$ near $x$. The weights are determined by radial basis function interpolation.

The ability to rapidly evaluate approximate kernel entries $\Aker(y,x)$ allows us to construct an H-matrix approximation, $\boldsymbol{\Aker}_H \approx \mathbf{\Aker}$, using the conventional adaptive cross H-matrix construction method (Section \ref{sec:h_matrix_construction_short}). In this method, one forms low-rank approximations of off-diagonal blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}. We then use H-matrix methods to convert $\boldsymbol{\Aker}_H$ into an H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$.

When $\Aop$ is symmetric positive semi-definite, $\mathbf{A}_H$ may be non-symmetric and indefinite due to errors in the approximation. In this case, one may (optionally) modify the H-matrix representation of $\mathbf{A}_H$ to make it symmetric positive semi-definite using the rational method described in Appendix \ref{sec:make_spd}.

%. The matrix is symmetrized by computing $\left(\mathbf{A}_H+\mathbf{A}_H^T\right)/2$ using H-matrix addition. Negative eigenvalues modified to be positive using a rational method that is described in Appendix \ref{sec:make_spd}.

Schematically the overall H-matrix approximation proceeds as follows:
\begin{equation*}
\Aop \longrightarrow \spatialvol, \spatialmean, \spatialcov \longrightarrow \underbrace{\{x_i\}_{i=1}^r}_{\text{as }S_b\text{'s}} \longrightarrow \underbrace{\{\impulseresponse_{x_i}\}_{i=1}^r}_{\text{as }\eta_b\text{'s}} \longrightarrow \boldsymbol{\Aker}_H \longrightarrow \mathbf{A}_H.
\end{equation*}
The complete algorithm for constructing $\mathbf{A}_H$ is shown in Algorithm \ref{alg:construct_Atilde}. The cost to construct $\mathbf{A}_H$ is discussed in Section \ref{sec:overall_cost}.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	{	
		\Input{Linear operator $\Aop$, parameter $\nbatch$}
		\Output{H-matrix $\mathbf{A}_H$ (optionally, $\mathbf{A}_H^{\text{sym}+}$)}

		Compute $V, \mu$, and $\Sigma$ (Section \ref{sec:intromoments})

%		Compute $\bm{\spatialvol}$, $\bm{\spatialmean}$, and $\bm{\spatialcov}$ using Algorithm \ref{alg:varhpi_mean_cov}.

		\For{$k=1,2,\dots,\nbatch$}{
			Choose a batch of sample points, $\pointbatch_k$, using Algorithm \ref{alg:point_choice}

			Compute $\combresponse_k$ by applying $\Aop$ to the Dirac comb for $\pointbatch_k$ (Section \ref{sec:get_impulse_response})

		}

		Form H-matrix approximation $\boldsymbol{\Aker}_H$ of integral kernel (Section \ref{sec:h_matrix_construction_short})

		Form H-matrix approximation $\mathbf{A}_H$ of $\Aop$ (Section \ref{sec:h_matrix_construction_short})

		(optional) Modify $\mathbf{A}_H$ to make $\mathbf{A}_H^{\text{sym}+}$ (Section \ref{sec:make_spd})

	}
	\caption{Construct H-matrix approximation}
	\label{alg:construct_Atilde}
\end{algorithm2e}

\subsection{Sample point selection via greedy ellipsoid packing}
\label{sec:sample_point_selection}

We choose sample points, $x_i$, in batches $\pointbatch_k$. We use a greedy ellipsoid packing algorithm to choose as many points as possible per batch, while ensuring that there is no overlap between the support ellipsoids, $E_{x_i}$, associated with the sample points within a batch.

We start with a finite set of candidate points $P$ and build $\pointbatch_k$ incrementally with points selected from $P$. For simplicity of explanation, here $\pointbatch_k$ and $P$ are mutable sets that we add points to and remove points from. First we initialize $\pointbatch_k$ as an empty set. Then we select the candidate point $p \in P$ that is the farthest away from all points in previous sample point batches $S_1 \cup S_2 \cup \dots \cup S_{k-1}$. Candidate points for the first batch $S_1$ are chosen randomly from $P$.
Once $p$ is selected, we remove $p$ from $P$. Then we perform the following two checks:
\begin{enumerate}
	\item We check whether $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in \pointbatch_k$.
	\item We make sure that $\spatialvol(p)$ is not too small, by checking whether $\spatialvol(p) > \epsilon \spatialvol_\text{max}$. Here $\spatialvol_\text{max}$ is the largest value of $\spatialvol(q)$ over all points $q$ in the initial set of candidate points, and $\tau$ is a small threshold parameter (we use $\epsilon=10^{-5}$).
\end{enumerate}
If $p$ passes both these checks (i.e., if $p$ is sufficiently far from other points in the batch, and $\spatialvol(p)$ is not too small) then we add $p$ to $\pointbatch_k$. Otherwise we discard $p$. This process repeats until there are no more points in $P$.  This is detailed in Algorithm \ref{alg:point_choice}. We check whether $E_p \cap E_q = \{\}$ using the ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.

We repeat the process to construct several batches of points $\pointbatch_1, \pointbatch_2, \dots$, until the number of batches equals t desired threshold $\nbatch$. In our implementation, for each batch the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, except for points in previously chosen batches.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	\SetKwProg{Fn}{Function}{}{}

	\Input{Finite set of candidate points $P\subset \Omega$, \\spatially varying mean $\spatialmean(x)$ and covariance $\spatialcov(x)$, \\previous sample point batches $\pointbatch_1, \dots, \pointbatch_{k-1}$}
	\Output{Batch of new sample points $\pointbatch_k$}

	Initialize empty new batch of sample points, $\pointbatch_k = \{\}$

	\While{$P$ is not empty}{
		Determine the point $p \in P$ that is farthest from all points in previous sample point batches $\pointbatch_1,\dots,\pointbatch_{k-1}$

		Remove $p$ from $P$

		\If{$E_p \cap E_q \neq \{\}$ for all $q \in \pointbatch_k$}{
			\tcp{$E_p$ and $E_q$ are the ellipsoids defined in \eqref{eq:support_ellipsoid}}

			Add $p$ to $\pointbatch_k$

			Remove all points $p'$ satisfying $\spatialmean(p') \in E_p$ from $P$
		}
	}

    \SetKwFunction{FMain}{point\_is\_acceptable}
	\caption{Choosing one batch of sample points, $\pointbatch_k$}
	\label{alg:point_choice}
\end{algorithm2e}

\subsection{Impulse response batches}
\label{sec:get_impulse_response}

We compute impulse responses $\impulseresponse_{x_i}$ in batches by applying $\Aop$ to Dirac combs. The Dirac comb, $\diraccomb_k$, associated with a batch of sample points, $\pointbatch_k$, is the following weighted sum of Dirac distributions (point sources) centered at the points $x_i \in \pointbatch_k$:
\begin{equation*}
	\diraccomb_k := \sum_{x_i \in \pointbatch_k} \delta_{x_i} / \spatialvol(x_i).
\end{equation*}
We compute the \emph{impulse response batch}, $\eta_k$, by applying $\Aop$ to the Dirac comb:
\begin{equation}
	\label{eq:dirac_comb_H_action}
	\combresponse_k := \left(\Aop \diraccomb_k^*\right)^*
	=\sum_{x_i \in \pointbatch_k} \impulseresponse_{x_i} / \spatialvol(x_i).
\end{equation}
The last equality in \eqref{eq:dirac_comb_H_action} follows from linearity and the definition of $\impulseresponse_{x_i}$ in \eqref{eq:impulse_response_delta_action}. Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\impulseresponse_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\impulseresponse_j$ do not overlap when $i \neq j$, we have (approximately)
\begin{equation}
\label{eq:varphi_eval}
	\impulseresponse_{x_i}(z) =
	\begin{cases}
		\combresponse_k(z) \spatialvol(x_i), & z \in E_{x_i} \\
		0, & \text{otherwise}
	\end{cases}
\end{equation}
for all $x_i \in \pointbatch_k$. By performing one matrix-vector product, $\xi_k \mapsto \left(\Aop \diraccomb_k^*\right)^*$, we recover the impulse responses $\impulseresponse_{x_i}$ associated with every point $x_i \in \pointbatch_k$.

Each point source $\delta_{x_i}$ is scaled by $1/\spatialvol(x_i)$ so that the resulting scaled impulse responses within $\eta_k$ are comparable in magnitude. Note that we are not in danger of dividing by zero, because our ellipsoid packing procedure from Section \ref{sec:sample_point_selection} excludes $x_i$ from consideration as a sample point if $\spatialvol(x_i)$ is smaller than a predetermined threshold. Without this scaling, the portion of $\phi_{x_i}$ outside of $E_{x_i}$, which we neglect, may overwhelm $\phi_{x_j}$ for a nearby point $x_j$ if $\spatialvol(x_i)$ is much larger than $\spatialvol(x_j)$.

\subsection{Approximate integral kernel entries}
\label{sec:approximate_kernel_entries}

Given $(y,x)\in \Omega \times \Omega$, let
\begin{equation*}
	z_i := y - \spatialmean(x) + \spatialmean(x_i), \qquad i=1,\dots,\numnbr,
\end{equation*}
where $\{x_i\}_{i=1}^{\numnbr}$ are the $\numnbr$ nearest sample points to $x$, excluding sample points $x_i$ for which $z_i \notin \Omega$. Here $\numnbr$ is a small user defined parameter; e.g., $\numnbr=10$. Further, let
\begin{equation}
\label{eq:fxyxp}
f_i := \frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}\left(z_i\right), \qquad i=1,\dots,\numnbr,
\end{equation}
Note that $f_i$ is well-defined, because $\phi_{x_i}\left(z_i\right)$ is defined if and only if $z_i \in \Omega$, and $\spatialvol(x_i)>0$ by our sample point picking procedure in Section \ref{sec:sample_point_selection}. We approximate $\Aker(y,x)$ by interpolating the (point,value) pairs
\begin{equation}
\left(x_i, f_i\right), \qquad i=1,\dots,\numnbr,
\end{equation}
at the point $x$. The idea is that $\Aker(y,x) = f_i$ per the discussion in Section \ref{sec:local_mean_displacement_invariance}, so we approximate $\Aker(y,x)$ by interpolating the values $f_i$.

To find the $\numnbr$ nearest sample points to $x$, we query a precomputed k-d tree \cite{Bentley75} of all sample points.  We check whether $z_i \in \Omega$ by querying a precomputed axis aligned bounding box tree (AABB tree) \cite{Ericson04} of the mesh cells used to discretize the problem.

Interpolation is performed using the following radial basis function scheme:
\begin{equation}
\Aker(y,x) \approx \widetilde{\Aker}(y,x) := \sum_{i=1}^{\numnbr} w_i~ b\left(\|x-x_i\|\right),
\end{equation}
where $b_i$ are radial basis functions and $w_i$ are weights. The vector of weights, $w = (w_1, w_2, \dots, w_{\numnbr})^T$, is found as the solution to the $\numnbr \times \numnbr$ linear system
\begin{equation}
Bw = f,
\end{equation}
where $B \in \mathbb{R}^{\numnbr \times \numnbr}$ has entries
$
B_{ij} := b\left(\|x_i - x_j\|\right),
$
and $f \in \mathbb{R}^\numnbr$ has entries $f_i$ from \eqref{eq:fxyxp}. We use polyharmonic spline radial basis functions, which are defined as follows:
\begin{equation}
b_i(r) := \begin{cases}
r^\gdim, & \gdim=1,3,5,\dots \\
r^\gdim \log r, & \gdim=2,4,6,\dots
\end{cases}
\end{equation}
for $r>0$, and $b_i(0):=0$. This yields the smoothest interpolant of the given points and values, in a certain least-squares sense \cite{Duchon77}. For more details on radial basis function interpolation, see \cite{Wendland04}. The interpolation procedure is illustrate in Figure \ref{fig:rbf_interpolation}.

\begin{figure}
	\centering
	\includegraphics[scale=0.2]{interpolation_rbf_boundary.pdf}
	\caption{To compute the kernel entry $\Aker(y,x)$, we use radial basis functions to interpolate the values $\frac{\spatialvol(x)}{\spatialvol(x_i)}\phi_{x_i}(y-\spatialmean(x)-\spatialmean(x_i))$ for the $\numnbr$ nearest sample points $x_i$ to $x$, excluding sample points for which $y-\spatialmean(x)-\spatialmean(x_i)$ is outside of the domain.}
	\label{fig:rbf_interpolation}
\end{figure}

To evaluate $f_i$, we check whether $z_i \in E_{x_i}$ using \eqref{eq:support_ellipsoid}. If $z_i \notin E_{x_i}$, then $z_i$ is outside the estimated support of $\impulseresponse_{x_i}$, so we set $f_i=0$. If $z_i \in E_{x_i}$, we look up the batch index $b$ such that $x_i \in S_b$, and evaluate $f_i$ via the formula
\begin{equation}
f_i = \spatialvol(x)\eta_b\left(z_i\right),
\end{equation}
per \eqref{eq:varphi_eval}.

Note that $z_i$ is generally not a gridpoint of the mesh used to discretize the problem, even if $y$, $x$, and $x_i$ are gridpoints. Hence,
evaluating $\eta_b\left(z_i\right)$ requires determining which mesh cell contains $z_i$, then evaluating finite element basis functions on that mesh cell. Fortunately, the mesh cell containing $z_i$ was already determined as a side effect of querying the AABB tree of mesh cells when we checked whether $z_i \in \Omega$.

\subsection{Hierarchical matrix construction}
\label{sec:h_matrix_construction_short}

We form our H-matrix approximation $\mathbf{A}_H \approx \mathbf{A}$ by forming H-matrix representations $\mathbf{\Aker}_H$ and $\massmatrix_H$ of $\mathbf{\Aker}$ and $\massmatrix$, respectively, then using fast H-matrix methods to multiply these matrices per \eqref{eq:boldA} to form
\begin{equation*}
\mathbf{A}_H = \massmatrix_H \mathbf{\Aker}_H \massmatrix_H.
\end{equation*}
%Once $\mathbf{A}_H$ is formed, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods \cite{BORMHACKBUSCHBOOK}. 
We use H1 matrices in our numerical results, but any of the other H-matrix formats (such as H2, HODLR, HSS, HBS, and others \cite{HMATRIX}) could be used instead. For more details on H-matrices, see \cite{Hackbusch15}.

We form $\mathbf{\Aker}_H$ using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package \cite{Kriemann08}. Although $\mathbf{\Aker}$ is a dense $\fedim \times \fedim$ matrix, constructing $\mathbf{\Aker}_H$ only requires evaluation of $O(\hrank^2 \fedim \log \fedim)$ entries $\mathbf{\Aker}_{ij} = \widetilde{\Aker}(p_i, p_j)$, and these entries are computed via the radial basis function interpolation method described in Section \ref{sec:approximate_kernel_entries}. Here $\hrank$ is the rank of the highest-rank block in the H-matrix. 
A dense representation of $\mathbf{\Aker}$ is never formed. 
%We describe this H-matrix construction process in detail in Appendix \ref{app:h_matrix}. 
We form $\massmatrix_H$ using standard H-matrix methods for sparse matrices implemented within HLIBpro, using the same recursive block partitioning structure as was used for $\mathbf{\Aker}$.

\section{Computational cost}
\label{sec:overall_cost}

The computational cost of our method may be divided into the costs to perform the following tasks: (1) Computing impulse response moments and batches; (2) Building the H-matrix; (3) Performing linear algebra operations with the H-matrix. This may optionally include modifying the H-matrix to make it symmetric positive semi-definite.
In practice, (1) is the dominant cost, because (1) is the only task that requires applying $\Aop$ to vectors, and our method is targeted at applications in which applying $\Aop$ to a vector requires an expensive computational procedure such as solving a large-scale partial differential equation. All operations that do not require applications of $\Aop$ to vectors are nearly-linear, and therefore scalable, in the size of the problem, $N$.
We now describe these costs in detail. For convenience, Table \ref{tab:vars} lists variable symbols and their approximate sizes.

\begin{table}
	\begin{tabular}{lll}
		Symbol & Typical size & Variable name \\
		\hline
		$\fedim$ & $10^2$--$10^9$ & Number of finite element degrees of freedom \\
		$\nbatch$ & $1$--$25$ & Number of impulse response batches \\
		$\hrank$ & $5$--$50$ & H-matrix rank \\
		$\numnbr$ & $5$--$15$ & Number of nearest neighbors for RBF interpolation \\
		$\gdim$ & $1$--$3$ & Spatial dimension \\
%		$\classicalrank$ & $10^1$--$10^4$ & Classical matrix rank \\
		$\nsamplepts$ & $10^1$--$10^4$ & Total number of sample points (all batches) \\
		$|\pointbatch_i|$ & $1$--$500$ & Number of sample points in the $i$th batch \\
		$\ratord$ & $1$--$2$ & Rational function parameter for SPSD modification
	\end{tabular}
	\caption{Symbols used for variables in computational cost estimates, and approximate ranges for their sizes in practice.}
	\label{tab:vars}
\end{table}

\paragraph{(1) Computing impulse response moments and batches} Computing $\spatialvol$, $\spatialmean$, and $\spatialcov$ requires applying $\Aop$ to $1$, $\gdim$, and $\gdim(\gdim+1)/2$ vectors, respectively.
Computing each impulse response batch requires applying $\Aop$ to one vector, so computing the impulse response batches $\{\eta_i\}_{i=1}^{\nbatch}$ requires $\nbatch$ operator applies. In total, computing the impulse response moments and batches therefore requires
\begin{equation*}
	1 + \gdim + \gdim(\gdim+1)/2 + \nbatch \qquad \text{operator applies.}
\end{equation*}
In a typical application one might have $\gdim=2$ and $\nbatch=5$, in which case a modest $11$ operator applies are required.

Computing the impulse response batches also requires choosing sample point batches via the greedy ellipsoid packing algorithm described in Section \ref{sec:sample_point_selection}. Choosing the $i$th batch of sample points may require performing up to $\fedim |\pointbatch_i|$ ellipsoid intersection tests, where $|\pointbatch_i|$ is the number of sample points in the $i$th batch. Choosing all of the sample points therefore requires performing at most
\begin{equation*}
	\fedim \nsamplepts \qquad \text{ellipsoid intersection tests,}
\end{equation*}
where $\nsamplepts$ is the total number of sample points in all batches. The multiplicative dependence of $\fedim$ with $\nsamplepts$ is undesirable since $\nsamplepts$ may be large, and reducing this cost is a target for future work. 
%Several improvements are possible. The cost of picking sample points could be reduced by ``thinning'' the set of candidate points that the sample points are picked from. That is, choosing sample points from a smaller subset of $\fedim_\text{thin} << \fedim$ points. The cost to perform the ellipsoid intersection tests could be reduced by first checking whether the ellipsoid bounding boxes intersect before performing the ellipsoid intersection test of Appendix \ref{sec:fast_ellipsoid_intersection_test}. The cost of picking the $i$th batch could be reduced to $O(\fedim \log |\pointbatch_i|)$ using more advanced computational geometry techniques, such as organizing the ellipsoids associated with points that have already been picked into a dynamic AABB tree, then using this AABB tree to accelerate the required ellipsoid intersection tests. 
However, from a practical perspective, the cost for picking sample points is small compared to other parts of the approximation algorithm, so we do not pursue these improvements in this paper.

\paragraph{(2) Building the H-matrix} The classical H-matrix construction technique
%, which is described in Appendix \ref{app:h_matrix}, 
requires evaluating $O(\hrank^2 \fedim \log \fedim)$ matrix entries of the approximation, where $\hrank$ is the H-matrix rank, i.e,  the maximum rank among the blocks of the H-matrix. To evaluate one matrix entry, first one must find the $\numnbr$ nearest sample points to a given point, where $\numnbr$ is the number of impulses used in the RBF interpolation. This is done using a precomputed KD tree of sample points, and requires $O(\numnbr \log \nsamplepts)$ floating point and logical elementary operations. Second, one must find the mesh cells that the points $\{z_i\}_{i=1}^{\numnbr}$ reside. This is done using an AABB tree of mesh cells, and requires $O(\numnbr \log \fedim)$ elementary operations. Third, we must evaluate finite element basis functions on each of those cells, which requires $O(\numnbr)$ elementary operations. Finally, we must solve a $\numnbr \times \numnbr$ linear system to perform the radial basis function interpolation, which requires $O(\numnbr^3)$ elementary operations, which typically is negligible. Therefore, building the H-matrix requires
\begin{equation*}
	O\left(\left(\hrank^2 \fedim \log \fedim\right)\left(\numnbr \log \fedim + \numnbr^3\right)\right) \qquad \text{elementary operations}.
\end{equation*}
%This follows from the above discussion, and the fact that $\fedim \ge \nsamplepts$.

\paragraph{(3) Performing linear algebra operations with the H-matrix} It is well-known that H-matrix methods for matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, and matrix inversion, and low rank updates require
\begin{equation}
\label{eq:H_matrix_cost_operations}
	O(\hrank^2 \fedim \log(\fedim)^a) \qquad \text{elementary operations},
\end{equation}
where $a \in \{0,1,2,3\}$ \cite{GrasedyckHackbusch03}. The precise cost depends on the type of H-matrix used and the operation being performed. In this paper, we use one matrix-matrix addition to add the H-matrix approximation of the data misfit term in the Hessian to the regularization term in the Hessian. Then, we use one matrix inversion to form the inverse of the Hessian, which is used as a preconditioner for the Newton linear system. 
%We perform low-rank updates to the H-matrix as the Newton iterations proceed, as per the discussion in Appendix \ref{sec:recycle_krylov}.
The rational method for modifying $\mathbf{A}_H$ to be positive semi-definite, described in Appendix \ref{sec:make_spd}, is performed using two matrix-matrix additions, $\ratord+1$ matrix-matrix multiplications, and one matrix inversion, where $2^\ratord$ is the order of the rational function used.

\section{Numerical results}
\label{sec:numerical_results}

%\textbf{Heat equation}
%\begin{itemize}
%	\item Hessian per-column error plots for 1, 5, and 15 batches
%	\item $H-P$ error vs number of batches for interior and whole domain
%	\item $P^{-1}H-I$: error vs rank for $P=R$ and $P=P_\text{pch}$ diffusion time, for 1, 5, and 15 batches
%	\item Krylov iterations to tols 1e-1 and 1e-6 vs diffusion parameter
%	\item GLR vs PCH vs diffusion time
%	\item Krylov method convergence plot: $R$ vs $P_\text{pch}$ vs no preconditioner
%	\item Krylov iterations to tols 1e-6 vs mesh size $h$, PCH vs R vs no preconditioner (hold)
%	\item true parameter (H)
%	\item deterministic reconstruction (H)
%	\item noisy measurements (1\%, 5\%, 10\% noise). $u|_\text{top}$ (H)
%	\item recovered state at top (H)
%	\item mesh scalability of PCH (CC) (H)
%	\item put plot data into data directory
%	\item save function .pvd files in paraview  directory
%\end{itemize}
%
%\begin{itemize}
%	\item GLR vs PCH number of obs (Computational cost in PDE solves) (S)
%	\item GLR vs PCH aspect ratio (CC) (S)
%	\item turn on preconditioner after number of krylov iterations exceeds 15 in an iteration. Rebuild "as needed". (S)
%\end{itemize}
%
%\begin{itemize}
%	\item Stokes Krylov convergence (10 solves for nonlinear forward + 6 hessian matvecs for ellipsoid estimation + 10 matvecs for impulse responses + 30 matvecs for PCH6 CG solve + 100 matvecs for Reg CG solve)
%	\item Stokes error vs num batches (small number of batches; 1, 3, 6, 9 batches?) (50 matvecs for randomized error estimation +)
%	\item Velocity field
%	\item Basal friction field
%	\item Picture of impulse responses
%	\item Mesh negative numbers picture
%\end{itemize}
%
%\begin{itemize}
%	\item mtrue
%	\item utrue (no arrows)
%	\item utrue (yes arrows)
%	\item reconstruction iterations 1,3,4,8: PCH vs REG
%	\item urecovered (arrows, 3D)
%	\item morozov plot
%	\item Newton convergence (iter, numcg, numPDEsolves (num preconditioner build), gradnorm)
%	\item impulse response batches
%	\item 
%\end{itemize}

We approximate the Gauss-Newton Hessian in an inverse problem governed by a PDE which models steady-state ice sheet flow. Briefly, a sheet of ice is sliding down a mountain (see Figure \ref{fig:ice_mountain_mesh}). Given observations of the tangential component of the ice velocity on the top surface of the ice, we invert for the unknown spatially varying basal friction parameter which governs how easily the ice slides along the ground underneath it.

\begin{figure}[H]
	\begin{subfigure}{0.6\textwidth}
		\includegraphics[scale=0.25]{meshHeight_view2_edges.png}
		\caption{Bird's eye view}
	\end{subfigure}
	\begin{subfigure}{0.38\textwidth}
		\includegraphics[scale=0.15]{meshHeight_view0_edges_withColorBar.png}
		\caption{Top-down view}
	\end{subfigure}
	\caption{The ice sheet, discretized by a mesh of tetrahedra. Color indicates the height of the base of the ice sheet (i.e., the mountain topography).}
	\label{fig:ice_mountain_mesh}
\end{figure}

The domain that is filled with ice is denoted by
$\icedomain \subset\mathbb{R}^{3}$. The basal, lateral and top parts of $\partial \icedomain$ are denoted by $\Gamma_{b}$, $\Gamma_{l}$, and $\Gamma_{t}$, respectively.
The governing equations are the linear incompressible Stokes
equations,

\begin{subequations}\label{Stokeseqn}
	\begin{align}
	-\nabla\cdot\bm{\sigma}_{\bm{u}}
	&=\rho\bm{g}\,\,\,\text{ in }\icedomain, \label{Stokeseqn:1}\\
	\nabla\cdot\bm{u}
	&=0\,\,\,\,\,\,\text{ in }\icedomain, \label{Stokeseqn:2}\\
	\bm{\sigma}_{\bm{u}}\bm{n}
	&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{t}, \label{Stokeseqn:3}\\
	\bm{u}\cdot\bm{n} =0 \quad\text{and}\quad \bm{T}\left(\bm{\sigma}_{\bm{u}}\bm{n}+\exp\left(m\right)\bm{u}\right)
	&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{b},\label{Stokeseqn:4} \\
	\left(\bm{\sigma}_{\bm{u}}\bm{n}+\mu\bm{u}\right)&=\bm{0}\,\,\,\,\,\,\text{ on }\Gamma_{l}.\label{Stokeseqn:5}	
	\end{align}
\end{subequations}
The solution to these equations is the tuple $(\bm{u}, \bm{p})$, where $\bm{u}$ is the ice flow velocity, and $\bm{p}$ is the pressure. Here, $m$ is the unknown logarithmic
basal sliding parameter field which is defined on the $2D$ surface $\Gamma_b$. The quantity $\rho\bm{g}$ is the body force density due to gravity, $\bm{n}$ is the outward normal and $\bm{T}$ is the tangential projection operator that restricts a vector field to its tangential component along the boundary. 
We employ Glen's flow law
(CITE GLEN) %\cite{glen1955},
which is a constitutive law for ice that relates the stress
tensor $\bm{\sigma}_{\bm{u}}$ and the strain rate tensor
$\bm{\dot{\varepsilon}}_{\bm{u}}= \frac{1}{2}\left(
\bm{\nabla}\bm{u}+\bm{\nabla}\bm{u}^{\top} \right)$, $\bm{\sigma}_{\bm{u}}= 2\eta
\bm{\dot{\varepsilon}}_{\bm{u}} -\bm{I}p$, where $\eta=\frac{1}{2}A^{-1}$ is the effective viscosity, $\bm{I}$ is the unit matrix,
and $A$ is a flow rate factor. Glen's exponent has been chosen $n=1$, which provides for a linear Stokes model. Note that while the PDE is linear, the parameter-to-solution map, $m \mapsto (\bm{u}, p)$, is nonlinear.

\begin{figure}[H]
    \begin{subfigure}{0.49\textwidth}
		\includegraphics[scale=0.243]{mtrue_withColorBar.png}
		\caption{True $m$.}
	\end{subfigure}
    \begin{subfigure}{0.49\textwidth}
    	\includegraphics[scale=0.18]{trueVelocity_glyphs_withColorBar.png}
    	\caption{True $\bm{u}$.}
    \end{subfigure}
	\caption{True logarithmic basal sliding parameter $m$ (left) and true velocity $\bm{u}$ (right). In the right plot, arrows indicate the direction of $\bm{u}$ and color indicates the magnitude of $\bm{u}$.}
\end{figure}

\begin{figure}[H]
	\begin{subfigure}{0.24\textwidth}
		\includegraphics[scale=0.1]{m1_psf.png}
		%		\caption{}
		%		\label{fig:newtoniter1}
	\end{subfigure}
	\begin{subfigure}{0.24\textwidth}
		\includegraphics[scale=0.1]{m3_psf.png}
		%		\caption{}
		%		\label{fig:newtoniter2}
	\end{subfigure}
	\begin{subfigure}{0.24\textwidth}
		\includegraphics[scale=0.1]{m4_psf.png}
		%		\caption{}
		%		\label{fig:newtoniter3}
	\end{subfigure}
	\begin{subfigure}{0.24\textwidth}
		\includegraphics[scale=0.1]{m6_psf.png}
		%		\caption{}
		%		\label{fig:newtoniter4}
	\end{subfigure}
	%\begin{subfigure}{0.05\textwidth}
	%	\includegraphics[scale=0.1]{mColorBar.png}
	%\end{subfigure}
	\caption{The parameter after $1$ (left), $3$ (second from left), $4$ (third from left) Newton iterates and the final Newton iterate (right). The preconditioner is constructed between Newton iterations 3 and 4.} 
\end{figure}

\begin{table}[H]
	\begin{center}
		\begingroup
		\setlength{\tabcolsep}{3pt}
		\renewcommand{\arraystretch}{1.25}
		\begin{tabular}{c| c c c | c c c | c c c}
			&  \multicolumn{3}{|c|}{PSF (5)} & \multicolumn{3}{|c|}{REG} & \multicolumn{3}{|c}{NONE} \\
			\hline
			Iter & 
			\#CG & \#Stokes & $||g||$ & 
			\#CG & \#Stokes & $||g||$ & 
			\#CG & \#Stokes & $||g||$ \\
			\verb~0~ &
			\verb~1~ & \verb~4~ & \verb~2.0e+7~ &
			\verb~1~ & \verb~4~ & \verb~2.0e+7~ &
			\verb~1~ & \verb~4~ & \verb~2.0e+7~ \\
			\verb~1~ &
			\verb~2~ & \verb~6~ & \verb~5.8e+6~ &
			\verb~2~ & \verb~6~ & \verb~7.9e+6~ &
			\verb~2~ & \verb~6~ & \verb~5.8e+6~ \\
			\verb~2~ &
			\verb~4~ & \verb~10~ & \verb~2.4e+6~ &
			\verb~5~ & \verb~12~ & \verb~3.9e+6~ &
			\verb~4~ & \verb~10~ & \verb~2.4e+6~ \\
			\verb~3~ &
			\verb~2~ & \verb~6+22~ & \verb~5.8e+5~ &
			\verb~13~ & \verb~28~ & \verb~1.6e+6~ &
			\verb~13~ & \verb~28~ & \verb~5.8e+5~ \\
			\verb~4~ &
			\verb~5~ & \verb~12~ & \verb~5.6e+4~ &
			\verb~42~ & \verb~86~ & \verb~4.8e+5~ &
			\verb~33~ & \verb~68~ & \verb~9.1e+4~ \\
			\verb~5~ &
			\verb~10~ & \verb~22~ & \verb~3.5e+3~ &
			\verb~84~ & \verb~170~ & \verb~7.7e+4~ &
			\verb~57~ & \verb~116~ & \verb~6.2e+3~ \\
			\verb~6~ &
			\verb~14~ & \verb~30~ & \verb~2.7e+1~ &
			\verb~125~ & \verb~252~ & \verb~5.0e+3~ &
			\verb~76~ & \verb~154~ & \verb~1.1e+2~ \\
			\verb~7~ &
			\verb~0~ & \verb~2~ & \verb~4.1e-2~ &
			\verb~194~ & \verb~390~ & \verb~7.9e+1~ &
			\verb~117~ & \verb~236~ & \verb~2.3e-1~ \\
			\verb~8~ & 
			\verb~---~ & \verb~---~ & \verb~---~ &
			\verb~0~ & \verb~2~ & \verb~1.6e-1~ &
			\verb~0~ & \verb~2~ & \verb~4.1e-2~ \\
			\hline
			Total & 
			\verb~38~ & \verb~114~ & \verb~---~ &
			\verb~466~ & \verb~950~ & \verb~---~ &
			\verb~303~ & \verb~624~ & \verb~---~ \\
		\end{tabular}
		\endgroup
	\end{center}
	\caption{Convergence history for solving the Stokes inverse problem using the preconditioned inexact Newton-Krylov method. The Newton iteration terminates when the norm of the gradient has decreased from its initial value by a factor of $10^{-8}$. The Gauss-Newton Hessian is used in iterations $0$--$5$, and the Hessian is used in all subsequent iterations. Preconditioners shown are our method with five impulse response batches (PSF (5)), regularization preconditioning (REG), and no preconditioning (NONE). Our preconditioner (PSF (5)) is built at iteration $3$ using the Gauss-Newton Hessian, and re-used for all subsequent iterations. The iterates are $m_0$ (initial guess), $m_1$ (parameter after one Newton step), $m_2$ (parameter after two Newton steps), and so on. The columns titled \#CG show the number of conjugate gradient iterations used to solve the Newton system at $m_k$, for the purpose of computing a search direction towards $m_{k+1}$. The columns titled $||g||$ show the $l^2$ norm of the gradient at $m_k$. The columns titled \#Stokes show the number of Stokes PDE solves performed in each Newton iteration. This consists of Stokes solves for $u$ performed during the linesearch going from $m_{k-1}$ to $m_k$ (including one Stokes solve at $m_k$), plus one adjoint Stokes solve to compute the gradient at $m_k$, plus one incremental forward and one incremental adjoint Stokes solve per CG iteration for solving the Newton system at iteration $k$. In the PSF (5) portion of row 3, we write $6+22$ to indicate that 6 Stokes solves were used during the standard course of the iteration, and 22 additional Stokes solves were used to build our preconditioner. }
\end{table}

\begin{figure}[H]
\begin{subfigure}{0.49\textwidth}
	\centering
	\begin{tikzpicture}[baseline, trim axis right]
	\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
	ymode=log,
	ymin=1e-11,
	ymax=2,
	xmax=750,
	axis x line*=bottom,
	axis y line*=left,
	xtick={0, 200, 400, 600},
	xlabel={CG iteration},
	ylabel={Relative residual norm},
	title={CG convergence},
	legend style={at={(1.1,1)},anchor=north east},
	%legend pos=north east,
	legend style={font=\small}
	]
	\addlegendentry{PSF (1)}
    \addplot[blue, line width=0.5mm] table {cg_relres_p1.dat};
    \addlegendentry{PSF (5)}
    \addplot[red, line width=0.5mm] table {cg_relres_p5.dat};
    \addlegendentry{PSF (25)}
    \addplot[purple, line width=0.5mm] table {cg_relres_p25.dat};
	\addlegendentry{Reg}
    \addplot[dotted, line width=0.5mm] table {cg_relres_reg.dat};
	\addlegendentry{None}
	\addplot[dash dot, line width=0.5mm] table {cg_relres_none.dat};
	\end{axis}
	\end{tikzpicture}
	\caption{}
	\label{fig:cg_convergence}
\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
		\centering
		\begin{tikzpicture}[baseline, trim axis right]
		\begin{axis}[width=0.99\linewidth, height=1.0\linewidth,
		xmode=log,
		ymode=log,
		xlabel={$\gamma$, regularization parameter},
		ylabel={\# of CG iterations},
		axis y line*=right,
		axis x line*=bottom,
		ylabel near ticks, yticklabel pos=right,
		%ylabel near ticks, ylabel shift={-6pt},
		%legend pos=north east,
		title={Regularization robustness}
		%title={CG iterations to achieve $||H x-b||/||b||<10^{-6}$}
		]
		%\addlegendentry{None}
		\addplot[dash dot, line width=0.5mm] table[x=gamma, y=none] {CGitsvsGamma.dat};
		%\addlegendentry{REG}
		\addplot[dotted, line width=0.5mm] table[x=gamma, y=reg] {CGitsvsGamma.dat};
		%\addlegendentry{PSF (1)}
        \addplot[blue, line width=0.5mm] table[x=gamma, y=psf1] {CGitsvsGamma.dat};
        %\addlegendentry{PSF (5)}
        \addplot[red, line width=0.5mm] table[x=gamma, y=psf5] {CGitsvsGamma.dat};
        %\addlegendentry{PSF (25)}
        \addplot[purple, line width=0.5mm] table[x=gamma, y=psf25] {CGitsvsGamma.dat};
		\end{axis}
		\end{tikzpicture}
		\caption{}
		\label{fig:gamma_sweep}
	\end{subfigure}
	\caption{Left (\ref{fig:cg_convergence}) shows the convergence history for solving a linear system of the form $Hx=b$ using the preconditioned conjugate gradient method, where $b$ has random independent and identically distributed (i.i.d) entries drawn from the standard normal distribution. The y-axis is the relative $l^2$ norm of the residual, i.e., $||b-Hx_k||/||b||$, where $x_k$ is the $k$th iterate. Here the regularization parameter $\gamma$ is determined by the Morozov discrepancy principle. Right (\ref{fig:gamma_sweep}) shows the number of preconditioned conjugate gradient iterations required to reduce the relative $l^2$ norm of the residual below a tolerance of $10^{-6}$, for a range of different regularization parameters $\gamma$. Results in these figures are shown for our preconditioner with 1, 5, and 25 batches of impulse responses (PSF (1), PSF (5), and PSF(25), respectively), regularization preconditioning (REG), and no preconditioning (NONE). The preconditioner is constructed using the Gauss-Newton Hessian. For a fair comparison, the Hessian and Gauss-Newton Hessian are always evaluated at the same parameter $m$, which is the optimal point for the inverse problem with $\gamma$ given by the Morozov discrepancy principle.}
\end{figure}

\begin{figure}[H]
	\begin{subfigure}{0.59\textwidth}
	\begin{center}
		\begin{tikzpicture}[baseline, trim axis right]
		\begin{axis}[width=0.99\linewidth, height=0.8\linewidth,
		ymode=log,
		xmin=-50,
		xmax=1450,
		ymax=1.e4,
		xlabel={$k$, generalized eigenvalue \#},
		ylabel={$\lambda_{k}$},
		xtick={0, 300, 600, 900, 1200},
		%legend pos=north east,
		legend style={font=\small},
		ylabel near ticks, ylabel shift={-6pt},
		axis x line*=bottom,
		axis y line*=left,
		legend style={at={(0.95,1)},anchor=north east},
		title={Generalized eigenvalues, $Hu_{k}=\lambda_{k}Pu_{k}$}
		]
		\addlegendentry{PSF (1)}
        \addplot[blue, line width=0.5mm] table[x=k, y=psf1] {generalizedEigenvalues.dat};
        \addlegendentry{PSF (5)}
        \addplot[red, line width=0.5mm] table[x=k, y=psf5] {generalizedEigenvalues.dat};
        \addlegendentry{PSF (25)}
        \addplot[purple, line width=0.5mm] table[x=k, y=psf25] {generalizedEigenvalues.dat};
		\addlegendentry{REG}
		\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {generalizedEigenvalues.dat};
		\end{axis}
		\end{tikzpicture}
	\end{center}
	\caption{}
	\label{fig:generalized_eigenvalues}
\end{subfigure}
\begin{subfigure}{0.39\textwidth}
		\begin{center}
		\begingroup
		\setlength{\tabcolsep}{4pt}
		\renewcommand{\arraystretch}{1.25}
		\begin{tabular}{c| c c}
			& $\kappa(P^{-1}H)$ &	$\kappa(P^{-1}H_{\text{gn}})$ \\
			\hline 
			PSF $(1)$  & \verb~1.75e+1~ & \verb~1.76e+1~ \\
            PSF $(5)$  & \verb~1.39e+1~ & \verb~1.39e+1~ \\
            PSF $(25)$ & \verb~1.03e+1~ & \verb~1.06e+1~ \\
			REG        & \verb~7.02e+3~ & \verb~7.02e+3~ \\
			NONE       & \verb~8.71e+2~ & \verb~9.22e+2~ 
		\end{tabular}
		\endgroup
	\end{center}
	\caption{}
	\label{fig:condition_number}
\end{subfigure}
	\caption{Left (\ref{fig:generalized_eigenvalues}): eigenvalues for generalized eigenvalue problem $H u_k = \lambda_k P_k$. Here $H$ is the Hessian and $P$ is the PSF approximation constructed using the Gauss-Newton Hessian, $H_\text{gn}$, for 1, 5, and 25 batches, or the regularization operator. Right (\ref{fig:condition_number}): condition number, $\kappa$, for the preconditioned Hessian and Gauss-Newton Hessian for different preconditioners. All operators are evaluated at the parameter $m$ that solves the inverse problem for $\gamma$ given by the Morozov discrepancy principle.}
\end{figure}

\section{Conclusions}
\label{sec:conclusions}

Some conclusions here.

Although the method we present in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest. For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high rank approximations of an operator using a small number of matrix-vector products.

\appendix

%\section{Hierarchical matrix details}
%\label{app:h_matrix}
%
%Often, large dense matrices of practical interest may be permuted, then partitioned into blocks recursively, in such a way that many off-diagonal blocks of the matrix are numerically low rank, even if the matrix is high rank. Such matrices are known as hierarchical matrices (H-matrices). Many classes of H-matrices exist (H1, H2, HSS, HBS, and more), and all of these types of H-matrices could be used in conjunction with our product-convolution approximation. Here we use classical H1 matrices. For this section, when we say H-matrix, we are referring to H1 matrices. 
%
%While a dense $\fedim \times \fedim$ matrix traditionally requires $O(\fedim^2)$ memory to store, H-matrices may be stored using $O(\fedim \log \fedim)$ memory, by storing only the low rank factors for the low rank blocks. Recursive algorithms can take advantage of the H-matrix low rank block structure to perform matrix arithmetic fast. Conventional dense matrix algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix-matrix addition require either $O(\fedim^2)$ or $O(\fedim^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(\fedim \log(\fedim)^a)$ time and memory for H-matrices. Here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix. For more details on H-matrices, we recommend \cite{HMATRIXGOOD}. 
%
%
%\subsection{H-matrix construction}
%\label{sec:H_matrix_construction}
%
%The process of constructing an H-matrix representation of $\mathbf{\Aker}$ proceeds as follows. First, we construct hierarchical partitionings of the degrees of freedom for the columns and rows of the matrix (cluster trees, Section \ref{sec:cluster_trees}). Second, we construct a hierarchical partitioning of the blocks of the matrix, in such a way that many of the blocks in the partitioning are expected to be low rank, and the remaining high rank blocks are small (block cluster tree, Section \ref{sec:block_cluster_tree}). Finally, we form low rank approximations of the blocks of the matrix that are expected to be low rank (adaptive cross approximation, Section \ref{sec:adaptive_cross}), and fill in the remaining high rank  blocks with their numerical values. The first two steps require geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix, but these steps do not depend on the particular values of matrix entries. The third step requires us to evaluate $O(\fedim \log \fedim)$ specially-chosen entries of $\mathbf{\Aker}$, and we evaluate these entries using \eqref{eq:Akerpcmat_entries}. 
%
%
%\subsubsection{Cluster trees}
%\label{sec:cluster_trees}
%
%We use recursive hyperplane splitting to hierarchically cluster the degrees of freedom associated with the columns and rows of the matrix into a \emph{column cluster tree} and a \emph{row cluster tree}, respectively. Here we describe construction of the column cluster tree; the row cluster tree is constructed similarly. 
%
%Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\mathbf{\Aker}$ corresponds to the Lagrange node in $\mathbb{R}^\gdim$ associated with the $i\text{th}$ finite element basis function. The columns of the matrix thus correspond to a point cloud in $\mathbb{R}^\gdim$. We split this point cloud into two equally sized \emph{child} point clouds, using a hyperplane which is perpendicular to the coordinate axis direction in which the point cloud is widest (e.g., either the $x$, $y$, or $z$ axis in $3$D). The two child point clouds are split in the same way. This splitting process repeats until the point clouds have less than a preset number of points (we use $32$ points). This hierarchical partitioning of the point cloud into smaller and smaller point clouds corresponds to a hierarchical partitioning of the columns of the matrix into smaller and smaller \emph{clusters} of columns. This hierarchical partitioning of the columns forms a binary tree, which is called the column cluster tree. The root of the tree is the set of all columns, and the leaves of the tree are clusters of columns that are not subdivided any further. 
%
%A depth-first search ordering of the column cluster tree leaves is then generated. When the columns of the matrix are permuted into this depth-first ordering, the columns associated with each cluster in the cluster tree are contiguous.
%
%In the same way, the degrees of freedom associated with the rows of the matrix are hierarchically clustered into another cluster tree, and a depth-first search ordering for the rows is generated. In our examples, the degrees of freedom for the columns coincide with the degrees of freedom for the rows, so the cluster trees for the rows and columns are the same, but this is not required in general.
%
%
%\subsubsection{Block cluster tree} 
%\label{sec:block_cluster_tree}
%
%We partition the matrix into a recursive hierarchy of mostly low rank blocks called the \emph{block cluster tree}. The idea is that a block of the matrix is likely to be low rank if the point cloud associated with the rows of the block is far away from the point cloud associated with the columns of the block. This is reasonable to expect here because of the locality property of $\Aop$. Indeed, locality implies that many blocks of the matrix corresponding to far away point cloud clusters will be rank zero.
%
%After reordering the rows and columns of $\mathbf{\Aker}$ via the depth-first ordering described above, we partition the reordered version of $\mathbf{\Aker}$ into a tree of $2 \times 2$ block matrices recursively. We use a geometric admissibility condition (discussed below) to decide which blocks to subdivide, and use the cluster trees for the rows and columns to determine how to subdivide those blocks. For the first stage of subdivision, let $r_1$ and $r_2$ be the children row clusters for the root of the row cluster tree, let $\colcluster_1$ and $\colcluster_2$ be the children column clusters for the root of the column cluster tree. The matrix $\mathbf{\Aker}$ is partitioned into blocks as follows:
%\begin{equation*}
%	\begin{bmatrix}
%		\mathbf{\Aker}_{11} & \mathbf{\Aker}_{12} \\
%		\mathbf{\Aker}_{21} & \mathbf{\Aker}_{22}
%	\end{bmatrix},
%\end{equation*}
%where $\mathbf{\Aker}_{11}$ denotes the block of $\mathbf{\Aker}$ with rows $r_1$ and columns $\colcluster_1$, $\mathbf{\Aker}_{12}$ denotes the block of $\mathbf{\Aker}$ with rows $r_1$ and columns $\colcluster_2$, and so on for $\mathbf{\Aker}_{21}$ and $\mathbf{\Aker}_{22}$.
%
%We now loop through the four blocks, $\mathbf{\Aker}_{11}$, $\mathbf{\Aker}_{12}$, $\mathbf{\Aker}_{21}$, and $\mathbf{\Aker}_{22}$, and decide which blocks should be subdivided further. For the purpose of explanation, consider $\mathbf{\Aker}_{12}$. If
%\begin{equation}
%	\label{eq:weak_admissibility_cond}
%	\dist\left(r_1, \colcluster_2\right) \ge \weakadmconst \min\left(\diam\left(r_1\right), \diam\left(\colcluster_2\right)\right),
%\end{equation}
%then we mark $\mathbf{\Aker}_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone. Here $\dist\left(r_1, \colcluster_2\right)$ is the Euclidean distance between the axis-aligned bounding box for the point cloud associated with the row cluster $r_1$, and the axis aligned bounding box for the point cloud associated with the column cluster $\colcluster_2$. The quantity $\diam\left(r_1\right)$ is the diameter of the axis aligned bounding box for the point cloud associated with the row cluster $r_1$, and $\diam\left(\colcluster_2\right)$ is the analogous diameter associated with the column cluster $\colcluster_2$. Here the quantity $\weakadmconst$ is a scalar constant; we use $\weakadmconst=2.0$. Basically, if the point clouds associated with $r_1$ and $\colcluster_2$ are far away from each other relative to their diameters, then we expect that the corresponding block of the matrix will be low rank. This process is repeated for the other blocks to determine which blocks are admissible and which are not. For us, the diagonal blocks $\mathbf{\Aker}_{11}$ and $\mathbf{\Aker}_{22}$ are not admissible because the distance between a point cloud and itself is zero. 
%
%Next, we subdivide all blocks that are not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\mathbf{\Aker}$. But now we subdivide a block based on the two child row clusters and two child column clusters for the rows and columns of that block. This subdivision process continues recursively until all blocks are either admissible, or smaller than the predetermined size mentioned above. The resulting hierarchical partitioning of matrix blocks forms a tree, which is called the \emph{block cluster tree}. The root of the tree is the whole matrix, internal nodes in the tree are blocks that are subdivided, and the leaves of the tree are blocks that are either expected to be low rank, or are small.
%
%
%\subsubsection{Adaptive cross low rank approximation of blocks}
%\label{sec:adaptive_cross}
%
%Once the block cluster tree has been constructed, low rank approximations of the admissible (low rank) blocks are formed using the adaptive cross method \cite{ACA}. Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\mathbf{\Aker}$. The idea of the adaptive cross method is to form a low rank approximation of $X$ by sampling a small number of rows and columns of $X$. 
%\begin{itemize}
%	\item Let $C \in \mathbb{R}^{m \times \hrank}$ be a matrix consisting of a subset of $\hrank$ columns of $X$, such that the span of the columns in $C$ approximates the column space of $X$. 
%	\item Let $R\in\mathbb{R}^{\hrank \times m}$ be a subset of the rows of $X$, such that the span of the rows in $R$ approximates the row space of $X$.
%	\item Let $U \in \mathbb{R}^{\hrank \times \hrank}$ be the block of $X$ consisting of the intersection of the rows from $R$ with the columns from $C$.
%\end{itemize}
%Then it is well-established that
%\begin{equation}
%	\label{eq:CUR}
%	X \approx C U^+ R,
%\end{equation}
%where $U^+$ is the pseudoinverse of $U$ \cite{goreinov1997theory,MahoneyDrineas09}. The quality of approximation \eqref{eq:CUR} depends on how well the columns of $C$ approximate the column space of $X$, and how well the rows of $R$ approximate the row space of $X$. 
%
%In the adaptive cross method, ``good'' columns, $C$, and rows, $R$, are chosen via an alternating iterative process. An initial set of columns $C$ is chosen. Keeping $C$ fixed, a set of rows $R$ is chosen so that the so that determinant of the submatrix $U$ within $C$ is as large as possible \cite{GoreinovEtAl10}. Now, keeping $R$ fixed, a set of new columns $C$ is chosen so that the submatrix $U$ within $R$ is as large as possible. This process repeats a small number of times. This results in matrices $R$ and $C$ such that the error in the approximation, \eqref{eq:CUR}, is small. The dominant cost of this procedure is the cost of computing $a \hrank$ columns of $X$ and $a\hrank$ rows of $X$, where $a$ is the number of alternating iterations. There is also a small linear algebra overhead cost for the determinant maximization process that is performed at each step. 
%The key point is that the adaptive cross method allows us to form a rank-$\hrank$ approximation of an $m \times m$ block of the matrix via a process that only requires accessing $O(m\hrank)$ entries of that block.
%
%We use the adaptive cross method to form low rank approximations for each admissible block of $\mathbf{\Aker}$. We directly compute all entries of the small dense blocks of $\mathbf{\Aker}$ that are not admissible. This process requires us to compute $O(\hrank^2 \fedim \log \fedim)$ entries of $\mathbf{\Aker}$, which is relatively cheap compared to the operator actions of $\Aop$ that are required to form the product-convolution approximation.

\section{Ellipsoid intersection test}
\label{sec:fast_ellipsoid_intersection_test}
The procedure for choosing sample points relies on quickly determining whether two ellipsoids intersect. Let $E_1$ and $E_2$ be the ellipsoids defined as
\begin{align*}
	E_1 :=& \{x : (x - \spatialmean_1)^T \spatialcov_1^{-1} (x - \spatialmean_1) \le \tau^2\}, \\
	E_2 :=& \{x : (x - \spatialmean_2)^T \spatialcov_2^{-1} (x - \spatialmean_2) \le \tau^2\}, \\
\end{align*}
where $\spatialmean_1, \spatialmean_2 \in \mathbb{R}^\gdim$, and $\spatialcov_1, \spatialcov_2 \in \mathbb{R}^{\gdim \times \gdim}$ are positive definite. Let $K$ be the following one dimensional convex function:
\begin{equation*}
	K(s) := 1 - \frac{1}{\tau^2} (\spatialmean_1 - \spatialmean_2)^T \left(\frac{1}{1-s}\spatialcov_1 + \frac{1}{s}\spatialcov_2\right)^{-1}(\spatialmean_1 - \spatialmean_2)	
\end{equation*}
In \cite{GilitschenskiHanebeck12} it is shown that $E_1 \cap E_2 = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$ (see Figure \ref{fig:ellipsoid_intersection_test}).

\begin{figure}
	\begin{subfigure}{0.37\textwidth}
		\centering
		\includegraphics[scale=1.0]{ellipsoids_intersect0.pdf}
	\end{subfigure}
	\begin{subfigure}{0.30\textwidth}
		\centering
		\includegraphics[scale=1.0]{ellipsoids_intersect1.pdf}
	\end{subfigure}
	\begin{subfigure}{0.30\textwidth}
		\centering
		\includegraphics[scale=1.0]{ellipsoids_intersect2.pdf}
	\end{subfigure}
	\caption{The ellipsoids intersect if $K(s) \ge 0$ for all $s \in (0,1)$.
	}
	\label{fig:ellipsoid_intersection_test}
\end{figure}

The function $K(s)$ may be evaluated quickly for many $s$ by pre-computing the solution to the generalized eigenvalue problem
\begin{equation*}
	\spatialcov_1 P = \spatialcov_2 P \Lambda,
\end{equation*}
where $P \in \mathbb{R}^{\gdim \times \gdim}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_\gdim)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$. The matrix $P$ simultaneously diagonalizes $\spatialcov_1$ and $\spatialcov_2$, in the sense that $P^T\spatialcov_1 P = \Lambda$, and $P^T\spatialcov_2 P = I$, where $I$ is the $\gdim \times \gdim$ identity matrix. Using this diagonalization, and some algebraic manipulations, we may write $K(s)$ as
\begin{equation}
	\label{eq:Ks_generalized}
	K(s) = 1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2,
\end{equation}
where $v := P^T\left(\spatialmean_1 - \spatialmean_2\right)$.

To check whether $E_1$ and $E_2$ intersect, we compute the generalized eigenvalue decomposition of $\spatialcov_1$ and $\spatialcov_2$, form $v$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's minimization algorithm \cite{Brent71} (any fast 1 dimensional convex optimization routine may be used). If $K(s^*) <0$ at the minimizer $s^*$, then $E_1 \cap E_2 = \{\}$. Otherwise $E_1 \cap E_2 \neq \{\}$.
%The resulting algorithm for checking whether $E_1$ and $E_2$ intersect is summarized in Algorithm \ref{alg:ellipsoid_intersection_test}.

%\begin{algorithm2e}
%	\SetAlgoNoLine
%	\SetKwInOut{Input}{Input}
%	\SetKwInOut{Output}{Output}
%	\SetKwProg{Fn}{Function}{}{}
%	\Input{Ellipsoid $E_1$ with mean $\spatialmean_1$ and covariance $\spatialcov_1/\tau^2$\\
%		Ellipsoid $E_2$ with mean $\spatialmean_2$ and covariance $\spatialcov_2/\tau^2$ 
%	}
%	
%	\Output{Boolean $\text{ellipsoids\_intersect}$ which is true if $E_1 \cap E_2 \neq \{\}$ and false otherwise}
%	
%	
%	Solve generalized eigenvalue problem $\spatialcov_1 P = \spatialcov_2 P \Lambda$
%		
%	$v \gets P^T\left(\spatialmean_1 - \spatialmean_2\right)$
%		
%	$\displaystyle K^* \gets \min_{s \in (0,1)}~1 - \frac{1}{\tau^2} \sum_{i=1}^\gdim \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2$
%		
%	\If{$K^* < 0$}{
%			
%		$\text{ellipsoids\_intersect} \gets \text{False}$
%			
%	}
%	\Else{
%			
%		$\text{ellipsoids\_intersect} \gets \text{True}$
%			
%	}
%	\caption{Determining whether two ellipsoids intersect}
%	\label{alg:ellipsoid_intersection_test}
%\end{algorithm2e}

\subsection{Ellipsoid bounding box test}

One may reduce the average cost of performing ellipsoid intersection tests by first checking whether the coordinate axis aligned bounding boxes for the ellipsoids intersect. If the bounding boxes do not intersect, then the ellipsoids do not intersect, so no further action is necessary. If the bounding boxes intersect, the ellipsoids might intersect, or they might not intersect, so one must perform the complete ellipsoid intersection test described in Section \ref{sec:fast_ellipsoid_intersection_test}.

The smallest coordinate axis aligned box that contains the ellipsoid 
$
E = \{x : (x - \spatialmean)^T \spatialcov^{-1} (x - \spatialmean) \le \tau^2\}
$
is given as follows:
\begin{equation*}
	\prod_{i=1}^\gdim \left[ \spatialmean_i -\tau\sqrt{\Sigma_{ii}}, \spatialmean_i +\tau\sqrt{\Sigma_{ii}} \right],
\end{equation*}
where $\Sigma_{ii}$ is the $i$th diagonal entry of $\Sigma$\footnote{See \url{https://math.stackexchange.com/q/3928964}}. The boxes
\begin{equation*}
	\prod_{i=1}^\gdim \left[a_i^-, a_i^+\right], \quad \text{and} \quad \prod_{i=1}^\gdim \left[b_i^-, b_i^+\right]
\end{equation*}
intersect if and only if $a_i^- \le b_i^+$ and $b_i^- \le a_i^+$ for $i=1,\dots,d$.

\section{Rational positive semi-definite modification}
\label{sec:make_spd}

In many problems of practical interest (e.g., Hessian approximation) $\Aop$ is symmetric positive semi-definite. However, $\mathbf{A}_H$ is generally non-symmetric and indefinite because of approximation error. 
This is undesirable. Symmetry and positive semi-definiteness are important properties which should be preserved if possible. Also, lacking these properties may prevent one from using highly effective algorithms, such as the conjugate gradient method, to perform further useful operations involving $\mathbf{A}_H$.

If $\Aop$ is symmetric, we symmetrize $\mathbf{A}_H$ as follows
\begin{equation*}
\mathbf{A}_H^{\text{sym}} := \frac{1}{2}\left(\mathbf{A}_H + \mathbf{A}_H^T\right).
\end{equation*}
If $\Aop$ is also positive semi-definite, we adapt the high order rational function spectral projection method from \cite[Section 13.8]{HackbuschKress07} to flip (or approximately flip) the negative eigenvalues of $\mathbf{A}_H^{\text{sym}}$ to be positive. In detail, let $\Pi$ be the indicator function for the interval $(a,b)$,
\begin{equation*}
	\Pi(\lambda) = \begin{cases}
		1, & \lambda \in (a,b), \\
		0, & \text{otherwise}.
	\end{cases}
\end{equation*}
Using functional calculus to extend the domain of the function $\Pi$ from scalars to matrices, we define
\begin{equation*}
\mathbf{\Pi} := \Pi(\mathbf{A}_H^{\text{sym}}).
\end{equation*}
The matrix $\mathbf{\Pi}$ is the spectral projector onto the eigenspace of $\mathbf{A}_H^{\text{sym}}$ corresponding to eigenvalues of $\mathbf{A}_H^{\text{sym}}$ that reside in $(a,b)$. If we choose $a < \lambda_\text{min}$ and $b=0$, where $\lambda_\text{min}$ is the smallest eigenvalue of $\mathbf{A}_H^{\text{sym}}$, then $\mathbf{\Pi}$ is the spectral projector onto the eigenspace associated with negative eigenvalues, and $\mathbf{\Pi}\mathbf{A}_H^{\text{sym}}$ is the ``negative component'' of $\mathbf{A}_H^{\text{sym}}$ in which all the negative eigenvalues are the same but all the positive eigenvalues are replaced by zero. Subtracting a negative number from itself twice flips its sign, so the matrix absolute value of $\mathbf{A}_H^{\text{sym}}$ is given by
\begin{equation}
\label{eq:I_Pi_abs}
	|\mathbf{A}_H^{\text{sym}}| = \left(\mathbf{I} - 2\mathbf{\Pi}\right)\mathbf{A}_H^{\text{sym}},
\end{equation}
where $\mathbf{I}$ is the identity matrix of the appropriate size.
By ``matrix absolute value,'' we mean that $|\mathbf{A}_H^{\text{sym}}|$ is the matrix that has the same eigenvectors as $\mathbf{A}_H^{\text{sym}}$, but the eigenvalues of $|\mathbf{A}_H^{\text{sym}}|$ are the absolute values of the eigenvalues of $\mathbf{A}_H^{\text{sym}}$.

%\begin{figure}
%	\centering
%	\includegraphics[scale=0.5]{spd_rational_function.pdf}
%	\caption{Rational function $\Pi_\ratord(\lambda)$ for $a=-1$, $b=0$, and $\ratord \in \{1,2,3\}$. As $\ratord$ increases, $\Pi_\ratord$ approaches the indicator function for the set $(a,b)$. }
%	\label{fig:spd_rat}
%\end{figure}

Computing $\mathbf{\Pi}=\Pi(\mathbf{A}_H^{\text{sym}})$ directly is difficult, but we may approximate $\Pi$ by a high order rational function $\Pi_\ratord \approx \Pi$ which is defined as follows:
\begin{equation*}
	 \Pi_\ratord(\lambda) := \frac{1}{1 + T(\lambda)^{2^\ratord} }, \qquad \text{where} \qquad T(\lambda) := \frac{2 \lambda - (b+a)}{b-a}.
\end{equation*}
Here $\ratord$ is a positive integer; the larger $\ratord$ is, the better the approximation $\Pi \approx \Pi_\ratord$. 
%This is illustrated in Figure \ref{fig:spd_rat}. 
Replacing scalar operations with matrix operations, we have an approximation $\mathbf{\Pi}_\ratord \approx \mathbf{\Pi}$ given by
\begin{equation}
\label{eq:rational343}
\mathbf{\Pi}_\ratord := \Pi_\ratord\left(\mathbf{A}_H^{\text{sym}}\right) = \left(\mathbf{I} + \mathbf{T}^{2^\ratord}\right)^{-1},
\end{equation}
where
\begin{equation*}
\mathbf{T} := T(\mathbf{A}_H^{\text{sym}}) = \left(2 \mathbf{A}_H^{\text{sym}} - (b+a) \mathbf{I}\right)/(b-a).
\end{equation*}
Using $\mathbf{\Pi}_\ratord$ instead of $\mathbf{\Pi}$ in \eqref{eq:I_Pi_abs} allows us to efficiently approximate $|\mathbf{A}_H^{\text{sym}}|$ using H-matrix methods. 
This approximation procedure is as follows. First, we compute $\lambda_\text{min}$ using the Lanczos method, and set $a=\gamma \lambda_\text{min}$ and $b=0$, where $\gamma\ge 1$ is a parameter that must be chosen. 
The method is relatively insensitive to $\gamma$ and numerically we find that the method works well for any $\gamma \in [1,2]$. We use $\gamma=1.5$ in our numerical results. 
Next, we compute $\mathbf{T}$ using fast H-matrix scalar multiplication and addition. We form the $\mathbf{T}^{2^\ratord}$ via repeated squaring, $\mathbf{T}^2 = \mathbf{T} \mathbf{T}$, $\mathbf{T}^4 = \mathbf{T}^2 \mathbf{T}^2$, $\mathbf{T}^8 = \mathbf{T}^4 \mathbf{T}^4$ and so on, where each of these matrix squarings is performed using H-matrix multiplication, which is fast. We recommend using $\ratord=1$ or $2$ (we use $\ratord=1$ in our numerical results). Larger $\ratord$ is unnecessary and may make the method unstable. The approximate spectral projector $\mathbf{\Pi}_\ratord$ is formed from $\mathbf{T}^{2^\ratord}$ per \eqref{eq:rational343}, using fast H-matrix operations to perform the required matrix-matrix addition and matrix inversion. Finally, a positive semi-definite approximation $\mathbf{A}_H^{\text{sym}+}$ of $\mathbf{A}_H^{\text{sym}}$ is formed as
\begin{equation*}
\mathbf{A}_H^{\text{sym}+} := \left(\mathbf{I} - 2\mathbf{\Pi}_\ratord\right)\mathbf{A}_H^{\text{sym}},
\end{equation*}
where fast H-matrix operations are used to perform the necessary matrix-matrix subtraction and multiplication. Note that for any $\ratord \ge 1$, $\Pi_\ratord(\lambda)> 1/2$ for $\lambda \in (a,b)$, so $\mathbf{A}_H^{\text{sym}+}$ is positive semi definite\footnote{Some indefiniteness may be introduced by truncation error in the H-matrix operations, but the negative eigenvalues caused by this truncation error are typically extremely small.}. At the end, $\mathbf{A}_H^{\text{sym}+}$ is symmetrized once again, to correct for any small non-symmetries that are introduced by truncation error in the H-matrix operations. Once constructed, we use the symmetric positive semidefinite matrix $\mathbf{A}_H^{\text{sym}+}$ in place of $\mathbf{A}_H$ in further linear algebra operations.

%\section{Recycling Krylov information via low rank updates}
%\label{sec:recycle_krylov}
%
%In applications, often one wants to use the H-matrix approximation $\mathbf{A}_H$ to build preconditioners for solving a sequence of slowly varying linear systems, where the $k$th linear system has the form
%\begin{equation}
%\label{eq:Ax_equals_b}
%\mathbf{B}^{(k)} \mathbf{u}^{(k)} = \mathbf{f}^{(k)}.
%\end{equation}
%Each of these linear systems is solved using an iterative method. By ``slowly varying,'' we mean that $\mathbf{B}^{(k+1)}$ does not differ much from $\mathbf{B}^{(k)}$. 
%For example, in a Newton-Krylov method for solving an optimization problem, one uses a Krylov method such as conjugate gradient to solve a sequence of systems of the form \eqref{eq:Ax_equals_b}, where $\mathbf{B}^{(k)}$ and $\mathbf{f}^{(k)}$ are the Hessian and the negative gradient of the objective function in the optimization problem, respectively, evaluated at the $k$th Newton iterate. 
%As the optimization solver converges, $\mathbf{B}^{(k)}$ converges to the Hessian at the optimal point, and therefore changes little from one iteration to the next. Since $\mathbf{B}^{(k)}$ varies slowly, it is reasonable to assume that Krylov information about $\mathbf{B}^{(k)}$ can be leveraged to update the preconditioner from $\mathbf{B}^{(k)}$ in order to build a preconditioner for $\mathbf{B}^{(k+1)}$.
%Concretely, let 
%\begin{equation*}
%\mathbf{P}^{(k)} \approx \left(\mathbf{B}^{(k)}\right)^{-1}
%\end{equation*}
%denote a H-matrix preconditioner used to accelerate the solution of \eqref{eq:Ax_equals_b} with an iterative method. The iterative method for solving the $k$th linear system will typically involve, as a sub-problem, applying $\mathbf{B}^{(k)}$ to a collection of vectors $\{\mathbf{x}_i\}_{i=1}^m$, thereby generating vectors $\{\mathbf{y}_i = \mathbf{B}^{(k)} \mathbf{x}_i\}_{i=1}^m$. Instead of rebuilding the H-matrix preconditioner for each successive linear system, we \emph{recycle} the information generated by the iterative method while solving the current linear system, performing a low-rank update to $\mathbf{P}^{(k)}$ to create $\mathbf{P}^{(k+1)}$. To do so, we build on the rank-1 update formulas discussed in \cite[Chapter 6]{NocedalWright99}, which we generalize to many vector ``rank-$m$'' updates. In what follows, let $\mathbf{X}$ and $\mathbf{Y}$ denote the $\fedim \times m$ matrices with columns given by the vectors $\mathbf{x}_i$ and $\mathbf{y}_i$, respectively. The matrix-vector products with $\mathbf{B}^{(k)}$ performed while solving the $k$th linear system may be summarized by the equation:
%\begin{equation*}
%\mathbf{B}^{(k)}\mathbf{X} = \mathbf{Y}.
%\end{equation*}
%To form $\mathbf{P}^{(k+1)}$ from $\mathbf{P}^{(k)}$, one may use one of the following three update formulas:
%\begin{align}
%	\mathbf{P}^{(k+1)} &= \mathbf{P}^{(k)} + \mathbf{R}\left(\mathbf{Y}^T \mathbf{Y}\right)^{-1} \mathbf{Y}^T \label{eq:res_update} \\
%	\mathbf{P}^{(k+1)} &= \mathbf{P}^{(k)} + \mathbf{R} \left(\mathbf{R}^T \mathbf{Y}\right)^{-1} \mathbf{R}^T \label{eq:srk_update} \\
%	\mathbf{P}^{(k+1)} &= \left(\mathbf{I} - \mathbf{X}\boldsymbol{\Theta}\mathbf{Y}^T\right)\mathbf{P}^{(k)}\left(\mathbf{I} - \mathbf{Y}\boldsymbol{\Theta}\mathbf{X}^T\right) + \mathbf{X}\boldsymbol{\Theta}\mathbf{X}^T, \label{eq:dfp_update}
%\end{align}
%where $\mathbf{R} := \mathbf{X}-\mathbf{P}^{(k)}\mathbf{Y}$ and $\boldsymbol{\Theta} := \left(\mathbf{X}^T\mathbf{Y}\right)^{-1}$.
%We note that for all three cases, the update is exact for $\mathbf{B}^{k}$, in the sense that 
%\begin{equation*}
%\mathbf{P}^{(k+1)}\mathbf{Y}=\mathbf{X} = \left(\mathbf{B}^{(k)}\right)^{-1} \mathbf{Y}.
%\end{equation*}
%Because the matrices $\mathbf{B}^{k}$ vary slowly, we expect that $\mathbf{P}^{(k+1)}$ is a good approximation for $\left(\mathbf{B}^{(k+1)}\right)^{-1}$ on the space spanned by the vectors $\mathbf{y}_i$. One should choose which formula to use based on the properties of the matrix $\mathbf{B}^{(k)}$. Formula \eqref{eq:res_update} does not preserve symmetry, and should therefore be used when $\mathbf{B}^{(k)}$ is not symmetric. Formula \eqref{eq:srk_update} preserves symmetry but not positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric and indefinite. Formula \eqref{eq:dfp_update}, which is used in this paper, preserves both symmetry and positive definiteness, and should therefore be used when $\mathbf{B}^{(k)}$ is symmetric positive definite. Formulas \eqref{eq:res_update} and \eqref{eq:srk_update} perform rank-$m$ updates, while formula \eqref{eq:dfp_update} performs a rank-$2m$ update. 
%We note that when $m=1$, formulas \eqref{eq:srk_update} are \eqref{eq:dfp_update} are known as SR1 (symmetric rank-$1$) and DFP (Davidson-Fletcher-Powell) updates, respectively \cite[Chapter 6]{NocedalWright99}.

\section*{Acknowledgments}
We thank J.J. Alger, Longfei Gao, Mathew Hu, and Rami Nammour for helpful discussions. We thank Trevor Heise for editing suggestions. We thank Georg Stadler for domain geometry help.

\FloatBarrier

\bibliographystyle{siamplain}
\bibliography{localpsf}

\end{document}

### 2024-02-24 presentations/siamuq24-noemi/localpsf_revised_figures/localpsf_old.tex
noemis slides

% SIAM Article Template
\documentclass[review,onefignum,onetabnum]{siamart190516}

% Information that is shared between the article and the supplement
% (title and author information, macros, packages, etc.) goes into
% ex_shared.tex. If there is no supplement, this file can be included
% directly.

\input{localpsf_shared}

% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={Fast matrix-free approximation of smoothly varying blur operators, with application to Hessians in PDE-constrained inverse problems with highly informative data},
  pdfauthor={N. Alger, N. Petra, and O. Ghattas}
}
\fi

% The next statement enables references to information in the
% supplement. See the xr-hyperref package for details.

\externaldocument{localpsf_supplement}

% FundRef data to be entered by SIAM
%<funding-group specific-use="FundRef">
%<award-group>
%<funding-source>
%<named-content content-type="funder-name"> 
%</named-content> 
%<named-content content-type="funder-identifier"> 
%</named-content>
%</funding-source>
%<award-id> </award-id>
%</award-group>
%</funding-group>

\begin{document}

\maketitle

% REQUIRED
\begin{abstract}
  We present an efficient matrix-free method for approximating locally translation invariant operators that have locally supported non-negative integral kernels. The idea of the method is to compute impulse responses of the operator at a collection of scattered points, then interpolate these computed impulse responses to approximate impulse responses at other arbitrary points. This results in a product-convolution approximation, which we convert to hierarchical matrix format. Once the approximation is represented in hierarchical matrix format, we perform matrix inversion and other matrix operations using fast hierarchical matrix arithmetic. Impulse responses are computed by applying the operator to a small number of Dirac comb ``batches'' of point sources. A key innovation of our method is a matrix-free procedure for choosing as many points as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap. We apply the method to approximate Hessians in large-scale PDE-constrained inverse problem with highly informative data. Numerical results demonstrate that our method substantially outperforms existing state-of-the-art Hessian approximation methods which are based on low-rank approximation. Our method is able to form high quality approximations of high rank Hessians using only a small number of Hessian matrix-vector products.
%  Within each Dirac comb, the point sources must be spaced sufficiently far apart so that the impulse responses to those point sources do not overlap. We estimate the required spacing of point sources by applying the operator to a small number of constant, linear, and quadratic functions, then postprocessing the results.
\end{abstract}

% REQUIRED
\begin{keywords}
  example, \LaTeX
\end{keywords}

% REQUIRED
\begin{AMS}
  68Q25, 68R10, 68U05
\end{AMS}

\section{Introduction}

We present a fast matrix free method for approximating locally translation-invariant operators, $\mathcal{A}$, that have locally supported non-negative integral kernels. Such operators arise, for example, as Schur complements in reduced space methods for solving partial differential equations (PDEs), and as Hessians in distributed parameter PDE-constrained inverse problems.

Let $\Omega \subset \mathbb{R}^d$ be a bounded domain in dimension $d=1$, $2$, or $3$, and let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be an integral operator of the form
\begin{equation}
\label{eq:kernel_representation}
(\mathcal{A}u)(v) := \int_\Omega \int_\Omega v(y) A(y,x) u(x) dx dy,
\end{equation}
with integral kernel $A:\Omega \times \Omega \rightarrow \mathbb{R}$. We focus on operators of this form which have the following properties:
\begin{description}
\item[Matrix-free:] One cannot easily access kernel entries $A(y,x)$. Access to $\mathcal{A}$ is available only through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors. That is, evaluation of the functions
\begin{equation*}
	u \mapsto\mathcal{A}u \quad \text{and} \quad v \mapsto\mathcal{A}^Tv.
\end{equation*} 
\item[Non-negative kernel:] For all $(y,x) \in \Omega \times \Omega$, we have
\begin{equation*}
	A(y,x) \ge 0.
\end{equation*}
\item[Local support:] $A(y, x)$ is supported near $y=x$.
\item[Local translation invariance:] If $h$ is not too large, then
\begin{equation}
	\label{eq:local_translation_invariance}
	A(y+h, x+h) \approx A(y,x).
\end{equation}
\end{description}
Local translation invariance is illustrated in Figure FIG. 
%We qualify the degree of local translation invariance with the words ``perfectly'', ``highly'', and ``poorly''.
%\begin{itemize}
%\item If equality holds in \eqref{eq:local_translation_invariance} for all $x$, $y$, and $h$, then we say $\mathcal{A}$ is \emph{perfectly translation invariant}. 
%\item If the discrepancy between left hand side and right hand side of \eqref{eq:local_translation_invariance} is small for a large set of points $x$, $y$, and $h$, then we say that $\mathcal{A}$ is \emph{highly locally translation invariant}. 
%\item If the discrepancy between left hand side and right hand side of \eqref{eq:local_translation_invariance} is large, or if the discrepancy is only small for a small set of $x$, $y$, and $h$, then we say that $\mathcal{A}$ is \emph{poorly locally translation invariant}.
%\end{itemize} 
If $\mathcal{A}$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then it is straightforward to show that $\mathcal{A}$ would be a convolution operator, and the convolution kernel, $\varphi_x$, would be given by
\begin{equation}
\label{eq:convolution_kernel}
	\varphi_x(y) := A(y+x,x).
\end{equation}
The function $\varphi_x$ is the result of applying $\mathcal{A}$ to a point source, $\delta_x$, at some fixed point $x \in \Omega$, then translating the resulting impulse response function to re-center it at zero instead of $x$. Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes. We therefore approximate $\mathcal{A}$ by a spatially varying weighted sum of convolution operators of the form
\begin{equation}
\label{eq:product_convolution}
	\left(\mathcal{A}u\right)(v) \approx \left(\widetilde{\mathcal{A}}u\right)(v) := \left( v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
\end{equation}
where $f \cdot g$ denotes pointwise multiplication, $f \ast g$ denotes convolution, and $\left(f, g\right)_{L^2(\Omega)}$ denotes the $L^2$ inner product on $\Omega$, for functions $f,g \in L^2(\Omega)$. Here
\begin{itemize}
	\item The local convolution kernels, $\varphi_i := \varphi_{x_i}$, are given by formula \eqref{eq:convolution_kernel}, but with $x$ replaced by $x_i$.
	\item $\{x_i\}_{i=1}^r \subset \Omega$ is a collection of points scattered throughout the domain.
	\item The functions $w_i$ are spatially varying weighting functions that are used to interpolate the convolution kernels $\varphi_i$.
\end{itemize}
Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution. The more locally translation invariant an operator is (i.e., the smaller the discrepancy between the left hand side and right hand side in \eqref{eq:local_translation_invariance}), the smaller the number of terms that are required in \eqref{eq:product_convolution} to achieve an accurate product-convolution approximation.

\subsection{Overview of the method}

Our operator approximation method is defined by how we compute the convolution kernels, $\varphi_i$, how we choose the points, $x_i$, what weighting functions, $w_i$, that we use, and how we use the resulting product-convolution approximation to perform linear algebra operations such as solving linear systems. In this section we summarize the answers to these questions; we will provide detailed answers to these questions in Section \ref{sec:method} and Section \ref{sec:using_approximation}.

We compute one ``batch'' of $\varphi_i$'s by applying $\mathcal{A}$ to a sum of point sources, $\delta_{x_i},$ at many points $x_i$ (a ``Dirac comb''). That is, we compute
\begin{equation*}
	\eta := \sum_{i=1}^m A(~\cdot~, x_i) = \left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*.
\end{equation*}
Here $\sum_{i=1}^m \delta_{x_i}$ is the Dirac comb distribution, and $\left\langle \mathcal{A}, \sum_{i=1}^m \delta_{x_i} \right\rangle^*$ denotes the Riesz representation of the application of $\mathcal{A}$ to the Dirac comb. In computations, this amounts to computing $\boldsymbol{\eta} = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \left(\sum_{i=1}^m \boldsymbol{\delta}_{x_i}\right)$, where $\boldsymbol{\eta}$, $\mathbf{A}$, and $\boldsymbol{\delta}_{x_i}$ are discretized versions of $\eta$, $A$, and $\delta_{x_i}$, respectively, and $\mathbf{M}$ is a mass matrix. The batch of points $x_i$ are chosen so that the support of $\left\langle \mathcal{A}, \delta_{x_i}\right\rangle^*$ and the support of $\left\langle \mathcal{A}, \delta_{x_j}\right\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$. Since the impulse responses do not overlap, we may post-process $\eta$ to recover $\varphi_i$ for all of the points $x_i$ in the batch---by applying $\mathcal{A}$ to one distribution, we recover many $\varphi_i$. This is illustrated in Figure FIG. The process is repeated to get more batches of $\varphi_i$'s, until the approximation is sufficiently accurate or a maximum number of sample points is reached.

In order to choose the points $x_i$, we need to estimate the support of $\left\langle \mathcal{A}, \delta_{x_i}\right\rangle^*$ \emph{before} we compute it. The supports of the functions $\left\langle \mathcal{A}, \delta_{x}\right\rangle^*$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\mathcal{A}^T$ to a small number of specially chosen functions (three functions if $d=1$, six if $d=2$, ten if $d=3$), then post processing the results (Section \ref{eq:mean_and_covariance_estimation}). This procedure is the only part of the paper that requires non-negativity of the integral kernel.

We choose the weighting functions, $w_i$, to be the solutions to the following optimization problems:
\begin{equation}
\label{eq:wi_optimization_problem_intro}
\begin{aligned}
\min_{w_i \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$ with pure Neumann boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$. Since the Laplacian measures local deviation from flatness [CITE JACOBSEN], the function $w_i$ is the flattest function, in a least-squares sense, that takes the value one at the sample point $x_i$, and the value zero at the other sample points $x_j$, $i \neq j$. Computing the weighting functions $w_i$ requires solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid. We call this interpolation scheme \emph{Poisson interpolation}. The advantage of Poisson interpolation over other interpolation methods, most notably radial basis function interpolation [cite Escande etc], is that Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions. This is illustrated in Figure FIG.

%A small number of terms in \eqref{eq:product_convolution} are required to accurately approximate a highly locally translation invariant operator, while a large number of terms are required to accurately approximate a poorly locally translation invariant operator.

Once the $\varphi_i$ and $w_i$ are computed, we convert $\widetilde{\mathcal{A}}$ to hierarchical matrix (H-matrix) format. Let $\widetilde{A}$ denote the integral kernel associated with $\widetilde{\mathcal{A}}$. While $A(y,x)$ is not easily computable, $\widetilde{A}(y,x)$ is given by the formula
\begin{equation}
\label{eq:kernel_entries}
	\widetilde{A}(y,x) = \sum_{i=1}^r \varphi_i(y-x) w_i(x).
\end{equation}
Formula \eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\widetilde{\mathcal{A}}$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks [CITE]. Once in H-matrix format, fast H-matrix arithmetic is used to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations.

\subsection{Motivation: PDE constrained inverse problems with highly informative data}
\label{sec:PDE_hessian_motivation}

Our motivation for this work is approximation of Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs). That is, inverse problems in which one seeks to reconstruct an unknown parameter field, $m$, from noisy observations, 
\begin{equation*}
y=f(m,u(m)) + \text{noise},
\end{equation*}
which depend on  a state variable $u$. In turn, $u$ depends on $m$ implicitly through the solution of a PDE, 
\begin{equation}
\label{eq:state_pde}
0=g(m,u).
\end{equation}
Here $u(m)$ denotes the solution of the PDE \eqref{eq:state_pde} as a function of $m$.
%
%That is, inverse problems in which one seeks to reconstruct an unknown parameter field, $m$, from noisy measurements,
%\begin{equation*}
%	y = B(m,u) + \xi,
%\end{equation*}
%of a state variable, $u$, which depends on $m$ indirectly through the solution of a PDE,
%\begin{equation}
%\label{eq:state_equation}
%	0=F(m,u).	
%\end{equation}
%Here $\xi$ is the noise. The observations $y$ are known, while $m$, $u$, and $\xi$ are unknown. In the deterministic setting, the inverse problem of finding $m$ is commonly framed as a least squares minimization problem
%\begin{equation}
%\label{eq:minimization_problem}
%	\min_m \quad J_d(m) + R(m),
%\end{equation}
%where
%\begin{equation*}
%	J_d(m) := \frac{1}{2}\|y - B(m,u(m))\|_{\Gamma}^2
%\end{equation*}
%is a data misfit term, $\Gamma$ is the noise covariance, $R(m)$ is a regularization term, and $u(m)$ denotes the solution $u$ of the PDE \eqref{eq:state_equation} as a function of $m$, . 
%
%Approximation of the Hessian,
%\begin{equation*}
%	\mathcal{H} = \mathcal{H}_d + \mathcal{H}_R := \frac{d^2 J_d}{dm^2} + \frac{d^2 R}{dm^2},
%\end{equation*}
In the deterministic approach to inverse problems, one typically finds $m$ as the solution to the minimization problem 
\begin{equation}
\label{eq:minimization_problem}
\min_m \quad \frac{1}{2}\|y - f(m,u(m))\|_W^2 + R(m),
\end{equation}
where $\|\cdot\|_W$ is weighted norm which depends on the noise covariance, and $R(m)$ is a regularization term. Approximation of the Hessian of the objective function, $\mathcal{H}$, allows for fast solution of \eqref{eq:minimization_problem} via Newton-type methods. Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because $\mathcal{H}^{-1}$ locally approximates the Bayesian posterior covariance for $m$. 
We only have access to $\mathcal{H}$ via its action on an arbitrary vectors $v$, i.e., evaluation of the map $v \mapsto \mathcal{H} v$. Evaluating $v \mapsto \mathcal{H} v$ is an expensive process that involves solving two auxiliary linear PDEs.

%The regularization term in the Hessian, $\mathcal{H}_R := \frac{d^2 R}{dm^2}$, is typically an elliptic differential operator, and is therefore easy to work with. The data misfit term, $\mathcal{H}_d = \frac{d^2 J}{dm^2}$, is difficult to work with. The data misfit term $\mathcal{H}_d$ is not directly accessible because it depends on $u$, which depends on $m$ implicitly through the solution of the PDE \eqref{eq:state_equation}. 
%
%We only have access to $\mathcal{H}_d$ via its action on an arbitrary vectors $v$, i.e., evaluation of the map $v \mapsto \mathcal{H} v$. Evaluating $v \mapsto \mathcal{H} v$ is an expensive process that involves solving two auxiliary PDEs. We show how to evaluate $v \mapsto \mathcal{H} v$ in Appendix ADSF.

The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term [CITE LOTS OF STUFF]. Either the Lanczos method or the randomized singular value decomposition [CITE RSVD] are used to perform the low rank approximation using only matrix-vector products. These methods suffer from a ``data predicament''---as more informative data is used in the inversion, the numerical rank of the data misfit term in the Hessian grows, so more matrix-vector products are required to form the aforementioned low rank approximation. The ideal scenario from a scientific perspective (highly informative data) is the worst case scenario from a computational perspective (large computational cost) [CITE MY DISSERTATION]. 
%In our numerical results, we show that our operator approximation method can efficiently approximate the data misfit term in the Hessian in certain PDE constrained inverse problems. 
The method in this paper offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because high-rank approximations of an operator can be formed using a small number of matrix-vector products.

\subsection{Existing work}

In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\varphi_i$ and $w_i$ may be arbitrary functions. Product-convolution approximations have been used in a wide variety of fields going back several decades [CITE BIG LIST].
For background on product-convolution approximations, we recommend reading the following papers: [CITES, INCLUDE ESCANDE SCATTERED].

We build upon the class of product convolution approximations in which the functions, $\varphi_i$, are impulse responses of $\mathcal{A}$ to point sources at a collection of points $x_i$ [CITE BIG LIST]. 
%The primary question in methods of this type are how to choose the points $x_i$, and how to choose the weighting functions $w_i$.
A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the functions $\varphi_i$ with piecewise linear interpolation [CITE NAGY], or splines [CITE SPLINES]. With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation. This is computationally prohibitive in our applications, because computing the functions $\varphi_i$ requires one matrix vector product with $\mathcal{A}$ for each point $x_i$, and these matrix vector products are computationally costly. In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large [CITE, CITE MY PAPER]. But even with adaptive refinement, many matrix vector products with $\mathcal{A}$ may be required. In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\varphi_i$ are computed with each matrix vector product. We do this by taking advantage of the fact that the functions $\varphi_i$ are locally supported; existing methods typically do not require this local support property.

Our method of estimating the support of the functions $\varphi_i$ was inspired by resolution analysis in seismic imaging [CITE]. In resolution analysis, $\mathcal{A}$ is the Hessian for a seismic inverse problem, and the width of $\varphi_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$. If the width of $\varphi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size. In [CITE], the width of $\varphi_p$ is estimated to be the local autocorrelation length of the function $\mathcal{A}^T \zeta$ near $p$, where $\zeta$ is a random noise funtion. In this paper, rather than probing $\mathcal{A}^T$ with random noise functions, we use specific constant, linear, and quadratic functions. Our method estimates the support of $\varphi_p$ more accurately than resolution analysis, but our method requires that $\mathcal{A}$ has a positive integral kernel, while the resolution analysis method does not have this requirement.

%The goal of these existing papers is to pick points $x_i$ so that the required the number of $\varphi_i$ is minimized. Our goal here is to minimize the required number of matrix vector products with $\mathcal{H}$.

%We do this by taking advantage of the local support property of $\mathcal{A}$.

%SVIR low rank: sample columns of matrix

%Pros and cons of different point picking strategies depend on the application of interest.

%In astronomy image deblurring applications, the point source positions typically correspond to locations of stars in an image, which are fixed.

%In [CITE], Escande and Weiss explored the idea of approximating linear operators from scattered impulse responses using a product-convolution scheme of the same form as \eqref{eq:product_convolution}.

%In our previous work [CITE MY PAPER], we choose the point locations with an adaptive grid. However,

%Nagy's paper: breaking domain into regular grid, using piecewise linear interpolation.

Matrix-vector products with product-convolution approximations can be performed using the fast Fourier transform if the problem is discretized on a regular grid [CITE]. The product-convolution approximation can then replace the original operator when using Krylov methods to solve linear systems with the original operator as the coefficient operator [CITE]. However, for complex geometries, problems are rarely discretized using regular grids, so we cannot easily use the fast Fourier transform. Furthermore, in big-data inverse problems the Hessian is typically ill-conditioned, so the required number of Krylov iterations is large. Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are amenable to fast linear algebra operations. Wavelet compression methods have been used for this purpose [CITE WAVELETS]. Here we follow our previous work [CITE MY PAPER], in which we convert the product-convolution approximation to H-matrix format.

Conventional H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products. There are matrix-free methods for H-matrix approximation [CITE LEXING PAPER, CITE MARTINSSON, CITE OTHERS], and these methods have been used to form H-matrix compressions of Hessians in PDE constrained inverse problems [CITE ILONA, CITE NOEMI/TUCKER HODLR]. While these matrix-free H-matrix construction methods are asymptotically scalable in theory, the required number of matrix vector products can be large in practice.

Product-convolution type approximations of Hessians have been used in a seismic inverse problem in [CITE GEORG], and in an advection-diffusion inverse problem in [CITE MY PAPER]. In [DEMANET SEISMIC], matrix probing [DEMANET MATRIX] is used to approximate the Hessian in a seismic inverse problem as the sum of simple pseudodifferential operators. Although it is not explicitly mentioned, the approximation in [DEMANET SEISMIC] could be interpreted as a convolution-product interpolation, which is like a product-convolution approximation, except the order of pointwise multiplications and convolutions is reversed in \eqref{eq:product_convolution}.

%H-matrices for Hessians: Ilonas paper with KAUST people

%Demanet matrix probing can be interpreted as convolution-product.

%Resolution analysis: uses matvecs with random functions to estimate spatial covariance. Was inspiration for us.

%Getting many impulse responses at once (batching).

%Geometry-aware weighting functions vs radial basis functions. Kriging.

%\subsection{Operator access levels}
%
%
%%In order to solve inverse or optimization problems governed by PDEs, we need to solve linear systems with the Hessian as the coefficient matrix [cite cite]. In order to do this efficiently, we have access to only Hessian-vector products. That is, we have the ability to compute the functions $x \mapsto \mathcal{H} x$ and $x \mapsto \mathcal{H}^T x$ for arbitrary vectors $x$. Therefore 
%
%A linear operator may be classified by the way in which we may access the operator. 
%
%
%\begin{itemize}
%	\item \emph{Level 1:} A matrix representation of the operator, $\mathbf{A}$, is stored in memory in dense or sparse format.
%	\item \emph{Level 2:} We can efficiently compute arbitrary matrix entries $\mathbf{A}_{ij}$ at will, but the matrix $\mathbf{A}$ is not stored in memory.
%	\item \emph{Level 3:} Access to $\mathbf{A}$ is only available through an algorithm that computes matrix-vector products $\mathbf{u} \mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{u} \mapsto \mathbf{A}^T \mathbf{u}$.
%\end{itemize}
%In large-scale scientific computing, 
%
%
%As discussed in Section \ref{sec:PDE_hessian_motivation}, in inverse problems governed by PDEs, matrix entries of the Hessian are not easily computable since getting one entry would require computing an entire column of the Hessian, then throwing away all but one entry in the column. Computing a single column costs at least two PDE solves. Many such entries are needed to approximate the Hessian in inverse problems with highly informative data (i.e., for Hessians with large numerical rank) [cite cite].
%
% of the implicit definition of $u(m)$. 
%
%Hierarchical matrices (H, H2, HSS, HBS, HODLR), quantized tensor trains, hierarchical interpolative factorization, STRUMPACK (LBNL), 
%
%Level 1: not relevant in large-scale problems
%State clearly: we do not use Level 1 at all.
%This paper: Level 3 -> Level 2
%H-matrix: solve on level 2
%
%%\subsection{Existing work}
%%
%%
%%\begin{itemize}
%%	\item CITE ESCANDE, others, for a reviews of the subject.
%%	\item Resolution analysis seismic papers: try to figure out size of $\varphi_x$ by matvec with random function, post-processing to get autocorrelation length. They did the constant matvec volume thing. Their goal is to get a sense of local length scales that can be resolved in the inverse problem. They do not build preconditioner
%%	\item Escande Weiss scattered response paper. They interpolate scattered impulse responses using radial basis functions and study how the geometry of the points affects the approximation accuracy. Matrix-free / getting many points at once is not a concern of them. The applications they consider one has direct access to the impulse response, e.g., through formula for the kernel. 
%%	\item Imaging problems: ``matvec'' requires performing an experiment
%%	\item pseudodifferential operator equivalence
%%	\item Laundry list of applications
%%	\item scattered impulse response of escande weiss paper: radial basis functions for approximation in $y$. How to determine if the impulse responses are far enough away from each other?
%%	\item my convolution paper: compute impulse responses one at a time, one impulse response per matvec: too slow! Here we compute many impulse responses using only 1 matvec.
%%	\item uniformity of smoothness condition
%%\end{itemize}

\section{Constructing the approximation}
\label{sec:method}
This section describes how we compute the convolution kernels, $\varphi_i$, and weighting functions, $w_i$, for the product-convolution approximation, $\widetilde{\mathcal{A}} \approx \mathcal{A}$, given in \eqref{eq:product_convolution}.
\begin{itemize}
\item We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}). To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as individual functions.
\item We use a greedy algorithm to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}). 
\item We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\langle\mathcal{A},\delta_{x}\rangle^*$ (Section \ref{eq:mean_and_covariance_estimation}). 
\item For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
\end{itemize}
The complete algorithm for constructing $\widetilde{\mathcal{A}}$ is shown in Algorithm \ref{alg:construct_Atilde}.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	%	\Input{Function $v \mapsto \mathcal{H}^Tv$}

	%	\mbox{}\\

	%	\Output{Spatially varying mean $\mu:\Omega \rightarrow \mathbb{R}^d$, \\ 
	%		Spatially varying covariance $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$}
	%	\Begin{
	{	
		Compute spatially varying mean, $\mu$, and covariance, $\Sigma$, using Algorithm \ref{alg:varhpi_mean_cov}.

		\For{$b=1,2,\dots,\text{max\_batches}$}{
			Choose a batch of sample points, $S_b$, using Algorithm \ref{alg:point_choice}

			Compute $\eta^b$ by applying $\mathcal{A}$ to the dirac comb associated with $S_b$ (Section \ref{sec:get_impulse_response})

			Compute weighting functions $w_i$ using Algorithm \ref{alg:weighting_functions_incremental}. In lines 2 and 3 of Algorithm \ref{alg:weighting_functions_incremental}, only solve for $\psi_i$ and $\theta_i$ associated with new points $x_i$ in $S_b$. The functions $\psi_i$ and $\theta_i$ associated with points in previous batches have already been computed in previous iterations of this for loop.
		}

		%		\Return{$\mu$, $\Sigma$}
	}
	\caption{Construct product-convolution approximation $\widetilde{\mathcal{A}}$}
	\label{alg:construct_Atilde}
\end{algorithm2e}

%The function $\varphi_x(y)$ is undefined when $x + y \notin \Omega$, which can occur even if $x \in \Omega$ and $y \in \Omega$. Extending $\varphi_x$ by zero makes it defined for all $y \in \mathbb{R}^d$, but this extension is only reasonable if $x$ is far enough from $\partial \Omega$ that $\varphi_x(y)$ decays to zero, as a function of $y$, before $y$ reaches $\partial \Omega$. We therefore choose exclude sample points $x_i$ if the ellipsoid associated with that point overlaps the boundary.

\subsection{Estimating impulse response supports}
\label{eq:mean_and_covariance_estimation}

Since $\varphi_x$, and thus $\langle\mathcal{A},\delta_{x}\rangle^*$, is nonnegative, $\langle\mathcal{A},\delta_{x}\rangle^*$ is a scaled probability distribution. We compute the scaling factor, denoted $\alpha(x)$, for all $x$ simultaneously by applying $\mathcal{A}$ to a constant function. We then compute the spatially varying mean, $\mu(x)$, and covariance, $\Sigma(x)$ of the normalized version of $\langle\mathcal{A},\delta_{x}\rangle^*$, for all $x$ simultaneously, by applying $\mathcal{A}^T$ to a small number of linear and quadratic functions. We make the approximation that the support of $\langle\mathcal{A},\delta_{x}\rangle^*$ is contained within the ellipsoid
\begin{equation}
\label{eq:support_ellipsoid}
E_x := \{x' \in \Omega: (x' - \mu(x))^T \Sigma(x)^{-1} (x' - \mu(x)) \le \tau^2\},
\end{equation}
where $\tau$ is a small number of standard deviations; we use $\tau=3$. The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations of the mean for the Gaussian distribution that has the same mean and covariance as the normalized version of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$. The fraction of the mass of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though we find this bound is conservative and far less mass resides in this region in practice.

\begin{defn}
	We define the spatially varying volume, $\alpha$, the scaled impulse response, $\rho_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
	\begin{align*}
		\alpha:\Omega \rightarrow \mathbb{R}, &\qquad \alpha(x) := \int_{\Omega} \langle\mathcal{A}, \delta_x\rangle^*(z) dz, \\
		\rho_x:\Omega \rightarrow \mathbb{R}, &\qquad \rho_x := \langle\mathcal{A}, \delta_x\rangle^* \big/ \alpha(x), \\
		\mu:\Omega \rightarrow \mathbb{R}^d, &\qquad \mu(x) := \int_\Omega z \rho_x(z) dz, \\
		\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\Sigma(x) := \int_\Omega (z - \mu(x))(z - \mu(x))^T \rho_x(z) dz.
	\end{align*}
\end{defn}

%The support of $\langle\mathcal{A},\delta_{x}\rangle^*$ is estimated to be contained within a small number of standard deviations of the mean for a Gaussian distribution with the same mean and covariance as the normalized version of $\langle\mathcal{A},\delta_{x}\rangle^*$.

%Let 
%\begin{equation*}
%	\rho_x := \langle\mathcal{A}, \delta_x\rangle^* \big/ \alpha(x)
%\end{equation*}
%denote the normalized version of $\langle\mathcal{A}, \delta_x\rangle^*$, where $\alpha:\Omega \rightarrow \mathbb{R}$ is the volume function
%\begin{equation*}
%	\alpha(x) := \int_{\Omega} \langle\mathcal{A}, \delta_x\rangle^*(z) dz.
%\end{equation*}
%Note that $\rho_x$ is a probability distribution on $\Omega$.
%
%In Proposition \ref{thm:vol_mean_cov}, we see that $\alpha(x)$ may be computed for all $x$ simultaneously by applying $\mathcal{A}^T$ to a constant function. We also see that the spatially varying mean, $\mu(x)$, and covariance, $\Sigma(x)$, of $\rho_x$, defined as
%\begin{align*}
%	\mu(x) :=& \int_\Omega z \rho_x(z) dz \\
%	\Sigma(x) :=& \int_\Omega (z - \mu(x))(z - \mu(x))^T \rho_x(z) dz
%\end{align*}
%may be computed for all $x$ simultaneously by applying $\mathcal{A}^T$ to $d$ linear functions, and $d(d+1)/2$ quadratic functions, respectively. In Algorithm \ref{alg:varhpi_mean_cov} we describe an algorithm for computing the volume function, mean, and covariance based on the results of Proposition \ref{thm:vol_mean_cov}.

\begin{theorem}
	\label{thm:vol_mean_cov}
	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$, and let $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions, respectively:
	\begin{equation*}
		C(x) := 1, \qquad
		L^i(x) := x^i, \qquad
		Q^{ij}(x) = x^i x^j.
	\end{equation*}
	We have
	\begin{align}
		\alpha =& \left(\mathcal{A}^TC\right)^*, \label{eq:vol_mean_var_thm1}\\
		\mu^i =& \left(\mathcal{A}^T L^i\right)^* / \alpha, \label{eq:vol_mean_var_thm2}\\
		\Sigma^{ij} =& \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot\mu^j, \label{eq:vol_mean_var_thm3}
	\end{align}
	We write $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
\end{theorem}

The proof of Theorem \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
%	\Input{Function $v \mapsto \mathcal{H}^Tv$}

%	\mbox{}\\

%	\Output{Spatially varying mean $\mu:\Omega \rightarrow \mathbb{R}^d$, \\ 
%		Spatially varying covariance $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$}
%	\Begin{
	{	
		Form constant function $C(x)=1$

		$\alpha = \left(\mathcal{A}^T C\right)^*$

		\For{$i=1,2,\dots,d$}{
			Form linear function $L^i(x) = x^i$

			$\mu^i = \left(\mathcal{A}^T L^i\right)^* / \alpha$
		}
		\For{$i=1,2,\dots,d$}{
			\For{$j=1,\dots,i$}{
				Form quadratic function $Q^{ij} = x^i x^j$

				$\Sigma^{ij} = \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot \mu^j$

				$\Sigma^{ij} = \Sigma^{ji}$

			}
		}

%		\Return{$\mu$, $\Sigma$}
	}
	\caption{Compute volume $\alpha$, mean $\mu$, and covariance $\Sigma$}
	\label{alg:varhpi_mean_cov}
\end{algorithm2e}

\subsection{Greedy sample point selection}
\label{sec:greedy_point_selection}

% sample points $x_i$ and $x_j$ should be far enough apart from each other that the supports of $\left\langle\mathcal{A}, \delta_{x_i}\right\rangle^*$ and $\left\langle\mathcal{A}, \delta_{x_j}\right\rangle^*$ do not overlap.

%Since it is too expensive compute the support of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ directly, we instead estimate that the support of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ is contained in the ellipsoid
%\begin{equation}
%\label{eq:support_ellipsoid}
%	E_x := \{x' \in \Omega: (x' - \mu(x))^T \Sigma(x)^{-1} (x' - \mu(x)) \le \tau^2\},
%\end{equation}
%where $\tau$ is a small number of standard deviations; we use $\tau=3$. The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations of the mean for the Gaussian distribution that has the same mean and covariance as $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$. The fraction of the mass of $\left\langle\mathcal{A}, \delta_{x}\right\rangle^*$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though we find this bound is conservative and far less mass resides in this region in practice.

%We call $E_x$ the \emph{unit ellipsoid} with mean $\mu(x)$ and covariance $\Sigma(x)/\tau^2$.

Starting from a finite set of candidate points $P$, we select batches of sample points, $S_b$, using a greedy algorithm. To build a batch $S_b$, first we initialize $S_b$ as an empty set. Then we select a candidate point $p \in P$ that is the farthest away from all points in previous sample point batches. That is, $p$ is a maximizer of the following optimization problem:
\begin{equation*}
\max_{p \in P} \min_{x \in S_1 \cup \dots \cup S_{b-1}} \|p - x\|.
\end{equation*}
Once $p$ is selected, we remove $p$ from $P$. If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in S_b$, and if $p$ is sufficiently far from the boundary, in the sense that $E_p \cap \partial \Omega = \{\}$, then we add $p$ to $S_b$. Otherwise we discard $p$. This process repeats until there are no more points in $P$.  This is detailed in Algorithm \ref{alg:point_choice}. We determine whether $E_p \cap E_q = \{\}$ using the fast ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.

%We select a batches of sample points, $S_b$, from a finite set of candidate points $P$ using a greedy algorithm. Starting from an empty set, we build $S_b$ from candidate points $p$ from $P$, one point at a time. The current candidate point, $p$, is the point in $P$

We repeat the process to construct several batches of points $S_1, S_2, \dots$, until the number of batches exceeds a desired threshold. In our implementation, for each batch the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, minus all points in previously chosen batches.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwProg{Fn}{Function}{}{}

	Start with a finite set of candidate points $P \subset \Omega$

	Initialize new empty batch of sample points, $S_b$

	\While{$P$ is not empty}{
		Determine the point $p \in P$ that is farthest from all points in the previous sample point batches $S_1,\dots,S_{b-1}$

		Remove $p$ from $P$

%		\If{$\text{point\_is\_acceptable}(p, S_b)$}{
		\If{$E_p \cap E_q \neq \{\}$ for all $q \in S_b$ and $E_p \cap \partial \Omega = \{\}$}{

			Add $p$ to $S_b$

			Remove all points $x$ satisfying $\mu(x) \in E_p$ from $P$
		}
	}

    \SetKwFunction{FMain}{point\_is\_acceptable}
%\SetKwProg{Fn}{Function}{:}{}
%\Fn{\FMain{$p$, $S_b$}}{
%	
%	$p\text{\_is\_acceptable} \gets \text{True}$
%	
%	$E_p$ is the ellipsoid associated with $p$ (see \eqref{eq:support_ellipsoid})
%	
%	\For{$q \in S_b$}{
%		
%		$E_q$ is the ellipsoid associated with $q$
%		
%		\If{$E_p \cap E_q \neq \{\}$}{
%			
%			$p\text{\_is\_acceptable} \gets \text{False}$
%			
%			Terminate for loop
%		}
%	}
%	\textbf{return} $p\text{\_is\_acceptable}$ 
%}

	\caption{Choosing one batch of sample points, $S_b$}
	\label{alg:point_choice}
\end{algorithm2e}

\subsection{Batched computation of impulse responses}
\label{sec:get_impulse_response}

The \emph{Dirac comb}, $\xi_b$, associated with a batch of sample points $S_b$ is the sum of Dirac distributions (point sources) centered at the points $x_i \in S_b$. That is,
\begin{equation*}
	\xi^b := \sum_{x \in S_b} \delta_{x_i}.
\end{equation*}
For each batch $S_b$, we compute as the action of $\mathcal{A}$ on the associated Dirac comb:
\begin{equation}
\label{eq:dirac_comb_H_action}
	\eta^b := \langle\mathcal{A}, \xi^b\rangle^*.
\end{equation}
By linearity, $\eta^b$ may be written as
\begin{equation}
\label{eq:phi_b}
	\eta^b = \left\langle\mathcal{A},\sum_{x_i \in S_b} \delta_{x_i}\right\rangle^* = \sum_{x_i \in S_b} \langle\mathcal{A}, \delta_{x_i}\rangle^*.
\end{equation}
Since the points $x_i$ are chosen so that the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ and the support of $\langle\mathcal{A},\delta_{x_j}\rangle^*$ do not overlap (or do not overlap much), we have
\begin{equation}
\label{eq:varphi_eval}
	\varphi_{x_i}(z) = \langle\mathcal{A}, \delta_{x_i}\rangle^*(z+x_i) = \begin{cases}
		\eta^b(z+x_i), & z+x_i \in E_{x_i} \\
		0, & \text{otherwise}.
		\end{cases}
\end{equation}
for all $x_i \in S_b$. By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover $\varphi_{x_i}$ for every point $x_i \in S_b$. To minimize memory usage, the functions $\varphi_i$ are never isolated and stored individually. Instead, we store the functions $\eta^b$, and use formula \eqref{eq:varphi_eval} to perform function evaluations $z \mapsto \varphi_i(z)$ as needed.
%For $x_i \in X$ and $y \in \mathbb{R}^d$, we may evaluate $\varphi_{x_i}(y)$ by finding the batch $X_b$ that $x_i$ is contained in, then using \eqref{eq:varphi_eval}.

%\begin{algorithm2e}
%	\SetAlgoNoLine
%	\SetKwInOut{Input}{Input}
%	\SetKwInOut{Output}{Output}
%	\SetKwProg{Fn}{Function}{}{}
%	\Input{Point source locations $x_1, x_2, \dots, x_m$
%	}
%	\Output{Response $\boldsymbol{\eta}$ of $\mathbf{A}$ to Dirac comb of point sources}
%
%	\BlankLine
%
%	Construct $\boldsymbol{\xi} \in \mathbb{R}^N_{\mathbf{M}^{-1}}, \quad \boldsymbol{\xi}_i = \sum_{j=1}^m \phi_i(x_j)$, \quad $i=1,\dots,N$
%	
%	$\boldsymbol{\eta} \gets \mathbf{M}^{-1} \mathbf{A}^T \mathbf{M}^{-1}\boldsymbol{\xi}$
%
%	\caption{Computing response $\mathbf{A}$ to Dirac comb}
%	\label{alg:dirac_comb_response}
%\end{algorithm2e}

\subsection{Poisson interpolation weighting functions}
\label{sec:weighting_functions}

We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to the following optimization problem:
\begin{equation}
\label{eq:wi_optimization_problem}
\begin{aligned}
\min_{w_i \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$, and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions. Using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense (Theorem \ref{thm:smoothest_interpolant}). Also, these weighting functions sum to one pointwise (Proposition \ref{thm:wi_sum_to_one}).

In Algorithm \ref{alg:weighting_functions_incremental}, we show how to efficiently solve optimization problem \eqref{eq:wi_optimization_problem}, and thus compute $w_i$.  Constructing all $r$ of the functions $w_i$ requires solving the Poisson equation $2r$ times with different right hand side sources, and these Poisson PDE solves can be performed cheaply with multigrid. All of the weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added. However, the dominant computational cost in constructing the weighting functions is the solution of the Poisson PDEs, and these Poisson PDE solves can be performed incrementally---we only need to perform two new Poisson PDE solves when a new point is added, rather than $2(r+1)$ new PDE solves.

It is important to note that functions in $H^2_N(\Omega)$ only have well-defined pointwise values if the spatial dimension is $d=1$, $2$, or $3$. Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-defined in three spatial dimensions or fewer. When inverting for parameter fields defined on four or higher spatial dimensions, the solution to this optimization problem does not exist, and these weighting functions should not be used. Proofs for results in this section are given in Appendix \ref{app:proofs}.

\begin{thm}[Optimal interpolation]
	\label{thm:smoothest_interpolant}
Let $y \in \mathbb{R}^d$, and let $u$ be the function
\begin{equation}
\label{eq:u_defn_in_thm}
u(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \widetilde{A}\left(y+x,x\right)
\end{equation}
where the weighting functions $w_i$ are the solutions to optimization problem \eqref{eq:wi_optimization_problem}. Then $u$ solves the optimization problem
\begin{equation}
\label{eq:uy_optimization_problem}
\begin{aligned}
\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
\text{such that} &\quad u(x_i) = \varphi_i(y), \quad i=1,\dots,q.
\end{aligned}
\end{equation}
\end{thm}

\begin{prop}[Weighting functions sum to one]
	\label{thm:wi_sum_to_one}
	We have
	\begin{equation*}
		\sum_{i=1}^r w_i(x) = 1
	\end{equation*}
	for all $x \in \Omega$.
\end{prop}

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	%	\Input{Function $v \mapsto \mathcal{H}^Tv$}

	%	\mbox{}\\

	%	\Output{Spatially varying mean $\mu:\Omega \rightarrow \mathbb{R}^d$, \\ 
	%		Spatially varying covariance $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$}
	%	\Begin{
	{	
		\For{$i=1,2, \dots, r$}{
			Compute $\psi_i$ as the solution to the Poisson problem
			\begin{equation}
			\label{eq:psi_eq}
			\Delta \psi_i = \delta_{x_i} - C/|\Omega|, \quad \text{in }\Omega,
			\end{equation}
			with Neumann zero boundary conditions, $\nu \cdot \nabla \psi_i = 0$ on $\partial \Omega$, and the average value zero condition, $\int_\Omega \psi_i(x) dx = 0$. Recall $\delta_{x_i}$ is a point source (delta distribution) centered at $x_i$, $C$ is the constant function $C(x)=1$, and $|\Omega|$ is the Lebesgue measure of $\Omega$.

			Compute $\theta_i$ as the solution to the Poisson problem
			\begin{equation}
			\label{eq:theta_eq}
			\Delta \theta_i = \psi_i, \quad \text{in }\Omega,
			\end{equation}
			with Neumann zero boundary conditions, $\nu \cdot \nabla \psi_i = 0$ on $\partial \Omega$, and the average value zero condition, $\int_\Omega \psi_i(x) dx = 0$.

		}

		Form the $r \times r$ matrix $S$ with entries
		\begin{equation}
		\label{eq:defn_of_S_alg}
		S_{ij} = \left(\psi_i, \psi_j\right)_{L^2(\Omega)}.
		\end{equation}

		\For{$i=1,2,\dots,r$}{
			Compute
			\begin{equation}
			\label{eq:definition_of_alpha_alg}
			\alpha = \frac{\mathbf{1}^T S^{-1} e_i}{\mathbf{1}^T S^{-1} \mathbf{1}},
			\end{equation}
			where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one, and $e_i = (0,\dots,0,1,0,\dots,0)$ is the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero.

			Compute 
			\begin{equation}
			\label{eq:definition_of_mu_alg}
			\lambda = -S^{-1} \left(e_i - \alpha \mathbf{1} \right).
			\end{equation}

			Form the weighting function $w_i$ given by
			\begin{equation}
			\label{eq:definition_of_wi_alg}
				w_i(x) = \alpha - \sum_{k=1}^r \lambda_k \theta_k(x).
			\end{equation}
		}
	}
	\caption{Compute weighting functions $w_i$, $i=1,\dots,r$.}
	\label{alg:weighting_functions_incremental}
\end{algorithm2e}

%	
%	The standard weak form for $\psi_j$ in \eqref{eq:psi_eq} is given by the following saddle point system for $\psi_j \in H^1(\Omega)$ and $\rho \in \mathbb{R}$:
%	\begin{alignat}{5}
%		-\int_\Omega \nabla \psi_j \cdot \nabla h ~dx &~+~& \rho \int_\Omega h ~dx &~=~ h(x_j) &\quad \forall h \in H^1(\Omega) \label{eq:lem_delta_eval_feq} \\
%		\int_\Omega \psi_j ~dx & & &~=~ 0. & \nonumber
%	\end{alignat}
%	This follows from integration by parts, using the Neumann zero boundary conditions for $\psi_j$, and enforcing the average value zero constraint for $\psi_j$ with the Lagrange multiplier $\rho$. If $h$ is the constant function $h(x)=1$, the first equation becomes
%	\begin{equation*}
%		\rho \int_\Omega 1 ~ dx = 1,
%	\end{equation*}
%	which implies $\rho = \frac{1}{|\Omega|}$. Now suppose that $h\in H^2_N(\Omega)$. Of course, we still have $h\in H^1(\Omega)$ as well, so equation \eqref{eq:lem_delta_eval_feq} still holds. But now $h$ has additional regularity, which allows us to use integration by parts on the first term in \eqref{eq:lem_delta_eval_feq} to bring all of the derivatives onto $h$. Furthermore, $h$ has Neumann zero boundary conditions, so the boundary term that arises during integration by parts is zero. Performing this integration by parts, and substituting in the value for $\rho$, yields
%	\begin{equation*}
%		\int_\Omega \psi_j \Delta h ~dx + \frac{1}{|\Omega|} \int_\Omega h ~dx = h(x_j),
%	\end{equation*}
%	which is the desired result.

\begin{thm}[Correctness of Algorithm \ref{alg:weighting_functions_incremental}]
\label{thm:wi_alg_is_correct}
	The weighting functions $w_i$ computed by Algorithm \ref{alg:weighting_functions_incremental} solve optimization problem \ref{eq:wi_optimization_problem}.
\end{thm}

%Since the Laplacian measures local deviation from flatness [cite Jacobsen], from \eqref{eq:uy_optimization_problem} we see that the interpolant $x \mapsto u_y(x) = \widetilde{\varphi}_x(y)$ is the ``flattest possible'' function (in a least-squares sense) that has Neumann boundary conditions and interpolates the data $\varphi_{x_i}(y)$.

% For example,
% \begin{itemize}
%     \item In inverse problems where the forward model is time-dependent, computing Hessian matrix-vector products requires solving time-dependent PDEs. In contrast, the Poisson equation is a stationary PDE.
%     \item In inverse problems where the parameter being inverted for resides on a boundary, computing Hessian matrix-vector products requires solving PDEs that are defined on a higher dimensional geometric domain than the Poisson PDEs that must be solved for the interpolation.
%     \item In multi-experiment inverse problems, data from many different ``experiments'' on the same medium are used in the inversion. Performing a Hessian matrix-vector product requires solving two PDEs \emph{per experiment}. The number of Poisson solves used to construct the weighting functions does not depend on the number of experiments.
% \end{itemize}
% The additional cost of Poisson interpolation is, therefore, a worthwhile investment if it allows one to achieve the same level of accuracy as other interpolation methods but with less sample points.

% One may also show that the weighting functions $w_i$ sum to one pointwise.

\section{Using the approximation}
\label{sec:using_approximation}

\subsection{Evaluating kernel entries of the approximation}
\label{sec:eval_matrix_entries}

We now describe how to efficiently evaluate $\widetilde{A}(y,x)$ at an arbitrary point $(x,y) \in \Omega \times \Omega$. From \eqref{eq:H_kernel_entries}, we have
\begin{equation}
\label{eq:eval_y_x}
\widetilde{A}(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x),
\end{equation}
and from \eqref{eq:varphi_eval}, we have
\begin{equation}
\label{eq:kernel_entry_formua}
\varphi_{x_i}(y-x) = \begin{cases}
\eta^{b(i)}(y - x + x_i), & y - x + x_i \in E_{x_i} \\
0, & \text{otherwise},
\end{cases}
\end{equation}
where $b(i)$ denotes the batch index such that $x_i \in S^{b(i)}$. To compute $z=\widetilde{A}(y,x)$, we start with $z=0$, then loop through all batches $b$, and all sample points in each batch. For each sample point, we add one term in the sum in \eqref{eq:eval_y_x} to $z$, using \eqref{eq:kernel_entry_formua} as needed to evaluate $\varphi_{x_i}$. This is shown in Algorithm \ref{alg:eval_y_x}.
%
%\begin{align*}
%	\widetilde{A}(y',x') &= \sum_{i=1}^q w_i(x') \varphi_{x_i}(y'-x') \\
%	&= \sum_{i=1}^q w_i(x') \begin{cases}
%		\eta^{b(i)}(\zeta_i), & \zeta_i \in E_{x_i} \\
%		0, & \text{otherwise}
%	\end{cases} \\
%	&= \sum_{i=1}^q w_i(x')\Psi_{ik},
%\end{align*}
%where $\omega_{ik}:= w_i(x_k') = \left(\mathcal{U}_X^+ \mathcal{U}\right)_{ik}$, and 
%\begin{equation*}
%	\Psi_{ik} := \begin{cases}
%		\eta^{b(i)}(\zeta_{ik}), & \zeta_{ik} \in E_{x_i}\\
%		0, & \text{otherwise}.
%	\end{cases}
%\end{equation*}
%To compute $\omega$, first we construct $\nu \in \mathbb{R}^{r \times N}$,	$\nu_{jk} := u_j(x_k)$,
%then we set
%\begin{equation*}
%	\omega := \mathcal{U}_X^+ \nu.
%\end{equation*}
%To compute $\Psi$, we must evaluate $\eta^{b(i)}(\zeta_{ik})$ for many points $\zeta_{ik}$.
%If the mesh used to discretize the problem is non-uniform, generally $\zeta_{ik}$ will not be a degree of freedom location, even if the points $x_k'$, $y_k'$, and $x_i$ are degree of freedom locations. Thus, evaluating $\eta^{b(i)}(\zeta_{ik})$ requires interpolation. Since it is typically more efficient to perform interpolation at many points all at once, rather than interpolating points one at a time, we precompute the set of indices
%\begin{equation*}
%	\chi^b := \{(i,k): \zeta_{ik} \in E_{x_i},~x_i \in X_b,~ k=1,\dots,N\}
%\end{equation*}
%for which the interpolation $\eta^{b(i)}(\zeta_{ik})$ must be performed. Then we compute the batch of interpolations $\eta^b\left(\zeta_{\chi^b}\right)$, and set 
%\begin{equation*}
%	\Psi_{\chi^b} = \eta^b\left(\zeta_{\chi^b}\right).
%\end{equation*}
%Here $\Psi_{\chi^b}$ is the subset of entries of $\Psi$ corresponding to indices $(i,k) \in \chi^b$, and $\eta^b\left(\zeta_{\chi^b}\right)$ is the interpolation of $\eta^b$ at all of the points $\zeta_{ik}$ corresponding to indices $(i,k) \in \chi^b$. This process is repeated for each batch $b$. All other entries of $\Psi$ are set to zero. The resulting procedure for evaluating $\widetilde{H}(y_k',x_k')$ is summarized in Algorithm \ref{alg:H_entries}.

%, for each batch $X_b$ we predetermine the subset of points 
%\begin{equation*}
%	p_{ik} := y_k - x_k + x_i
%	\end{equation*}
%satisfying $p_{bik} \in E_{x_i}$ for all $x_i \in X_b$ then evaluate $\eta_b(p_{bik})$ for this subsets of points all at once. Let
%\begin{equation*}
%	\chi_b:=\{p_{bik}:p_{bik} \in E_{x_i}\}.
%\end{equation*}
%and define $\Psi \in \mathbb{R}^{q \times N}$, 
%
%so that
%\begin{equation*}
%	\widetilde{H}(y_k,x_k) = \sum_{i=1}^q w_i(x_k) \Psi_{ik}.	
%\end{equation*}
%We have 
%\begin{equation*}
%	\Psi_{ik}(\chi_b) = \sum_{b=1}^B \eta_b(p_{bik})
%\end{equation*}
%where $\eta_b(\chi)$ is the 
%
%In Algorithm asdf, we describe how to evaluate $\widetilde{H}$ at many points using this procedure.
%
\begin{algorithm2e}
	\SetAlgoNoLine

	$z \gets 0$

	\For{$b=1,2,\dots,\text{}max\_batches$}{
		\For{$x_i \in S_b$}{
			\If{$y - x + x_i \in E_{x_i}$}{
				$z \gets z + w_i(x)\eta^b(y-x+x_i)$
			}
		}
	}	
	\caption{Compute approximate kernel entry $z = \widetilde{A}(y,x)$}
	\label{alg:eval_y_x}
\end{algorithm2e}

%\begin{algorithm2e}
%	\SetAlgoNoLine
%	\SetKwInOut{Input}{Input}
%	\SetKwInOut{Output}{Output}
%	\SetKwProg{Fn}{Function}{}{}
%	\Input{Pairs of points $\{(x_k',y_k')\}_{k=1}^n \subset \Omega \times \Omega$}	
%	\Output{Approximate kernel entries $\mathbf{z} \in \mathbb{R}^n$, $\mathbf{z}_k = \widetilde{H}(y_k',x_k')$}
%	
%	Compute $\boldsymbol{\omega} \in \mathbb{R}^{r \times n}$, $\boldsymbol{\omega}_{i,k} = w_i(x_k')$
%	
%	\mbox{}\\
%	
%	\Fn{$\computekernelentries{\{x'_k\}_{k=1}^n}{\{y'_k\}_{k=1}^n}$}
%	{
%		$\boldsymbol{\omega} \gets \computeweightingentries{\{x'_k\}_{k=1}^n}$
%		
%		Construct $\boldsymbol{\zeta} \in \mathbb{R}^{d \times q \times n}$, ~$\boldsymbol{\zeta}[:,i,k] = y_k' - x_k' + \mathbf{X}\left[:,\mathtt{S}[i]\right]$
%
%		\BlankLine
%		
%		\BlankLine
%		
%		$\boldsymbol{\Psi}\gets \mathbf{0}_{q \times n}$
%		
%		\For{$b=1,\dots,N_B$}{
%			$\mathtt{\chi}_b \gets \{\}$
%			
%			\For{$j \in \mathtt{S}_b$}{
%				$i \gets$ index such that $\mathtt{S}[i] = j$
%				
%				$E_j \gets$ unit ellipsoid with mean $\boldsymbol{\mu}[:,j]$ and covariance $\boldsymbol{\Sigma}[:,:,j]/\tau^2$
%				
%				$\mathtt{\chi}_b \gets \mathtt{\chi}_b \cup \{(i,k) : k \in \{1,\dots,N\}, \boldsymbol{\zeta}[:,i,k] \in E_j\}$
%				
%			}
%
%			
%			$\boldsymbol{\Psi}[\chi_b] \gets \interpolate\left(\boldsymbol{\eta}^b, \{\boldsymbol{\zeta}[:,i,k] : (i,k) \in \mathtt{\chi}^b\}\right)$
%		
%		}
%	
%		\BlankLine
%		
%		\BlankLine
%		
%		Construct $\mathbf{z} \in \mathbb{R}^n$, ~$\mathbf{z}_k = \sum_{i=1}^q \boldsymbol{\omega}_{ik} \boldsymbol{\Psi}_{ik}$
%	
%		\Return{$\mathbf{z}$}
%	}
%	\caption{Compute many approximate kernel entries $z_k = \widetilde{A}(y_k',x_k')$}
%	\label{alg:H_entries}
%\end{algorithm2e}

\subsection{Conversion to hierarchical matrix format}

In order to use the product-convolution approximation $\widetilde{\mathcal{A}}$, we convert $\widetilde{\mathcal{A}}$ to hierarchical matrix (H-matrix) format. Many classes of hierarchical matrices exist---H1, uniform H1, H2, HODLR, HSS, HBS, and more [CITES]. We use H1 matrices in our numerical results, but any of the other H-matrix formats could be used instead. The basic ideas behind H-matrices are as follows:
\begin{itemize} 
	\item Often, large dense matrices of practical interest may be permuted, then partitioned into blocks recursively, in such a way that many off-diagonal blocks of the matrix are numerically low rank, even if the matrix is high rank.
	\item Recursive algorithms can take advantage of this low rank block structure to perform matrix arithmetic fast. Conventional algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix addition require either $O(n^2)$ or $O(n^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(n \log^a n)$ time and memory for hierarchical matrices (here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix).
\end{itemize}
For more details on H-matrices, we recommend [CITES].

We convert $\widetilde{\mathcal{A}}$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBpro software package [CITE]. Briefly, the H-matrix construction process that we use proceeds as follows. First, we construct a hierarchical partitionings of the degrees of freedom for the rows and columns of the matrix (cluster trees, Section \ref{sec:cluster_trees}). Second, we construct a hierarchical partitioning of the blocks of the matrix, in such a way that many of the blocks in the partitioning are expected to be low rank, and the remaining high rank blocks are small (block cluster tree, Section \ref{sec:block_cluster_tree}). Finally, we form low rank approximations of the blocks of the matrix that are expected to be low rank (adaptive cross approximation, Section \ref{sec:adaptive_cross}), and fill in the remaining high rank  blocks with their numerical values. The first two steps require geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix, but these steps do not depend on the particular values of matrix entries. The third step requires us to evaluate $O(n \log n)$ specially-chosen matrix entries of $\widetilde{\mathcal{A}}$, which we do using the method described in Section \ref{sec:eval_matrix_entries}. In the remainder of this section we describe each of these steps in more detail.

\subsubsection{Cluster trees}
\label{sec:cluster_trees}

The degrees of freedom associated with the columns of the matrix are hierarchically clustered into a \emph{cluster tree} using recursive hyperplane splitting. Since we use finite elements to discretize the problem, we associate the $i^\text{th}$ column of $\widetilde{\mathcal{A}}$ with the nodal point in $\mathbb{R}^d$ associated with the $i^\text{th}$ finite element basis function. The columns of the matrix thus correspond to a point cloud in $\mathbb{R}^d$. We split this point cloud into two equally sized \emph{child} point clouds, using a hyperplane which is perpendicular to the coordinate axis direction in which the point cloud is widest (e.g., either the $x$, $y$, or $z$ axis in $3$D). The two child point clouds are split in the same way. This splitting process repeats until the point clouds have less than a preset number of points (we use $32$ points). This hierarchical partitioning of the point cloud into smaller and smaller point clouds corresponds to a hierarchical partitioning of the columns of the matrix into smaller and smaller \emph{clusters} of columns. This hierarchical partitioning of the columns forms a binary tree, which we call a cluster tree. The root of the tree is the set of all columns, and the leaves of the tree are clusters of columns that are not subdivided any further.

A depth-first search ordering of the cluster tree leaves is then generated. When the columns of the matrix are permuted into this depth-first ordering, the columns associated with each cluster in the cluster tree are contiguous.

In the same way, the degrees of freedom associated with the rows of the matrix are hierarchically clustered into another cluster tree, and an ordering for the rows is generated. In our examples, the degrees of freedom for the columns coincide with the degrees of freedom for the rows, so the cluster trees for the rows and columns are the same, but this is not required in general.

\subsubsection{Block cluster tree} 
\label{sec:block_cluster_tree}

We partition the matrix into a recursive hierarchy of mostly low rank blocks called the \emph{block cluster tree}. The idea is that a block of the matrix is likely to be low rank if the point cloud associated with the rows of the block is far away from the point cloud associated with the columns of the block. This is reasonable to expect here because of the locality property of $\mathcal{A}$. Indeed, locality implies that many blocks of the matrix corresponding to far away point cloud clusters will be rank zero.

After reordering the rows and columns of $\widetilde{\mathcal{A}}$ via the depth-first ordering described above, we partition the reordered version of $\widetilde{\mathcal{A}}$ into a tree of $2 \times 2$ block matrices recursively. We use a geometric admissibility condition (discussed below) to decide which blocks to subdivide, and use the cluster trees for the rows and columns to determine how to subdivide those blocks. For the first stage of subdivision, let $r_1$ and $r_2$ be the children row clusters for the root of the row cluster tree, let $c_1$ and $c_2$ be the children column clusters for the root of the column cluster tree. The matrix $\widetilde{\mathcal{A}}$ is partitioned into blocks as follows:
\begin{equation*}
	\begin{bmatrix}
		\widetilde{\mathcal{A}}_{11} & \widetilde{\mathcal{A}}_{12} \\
		\widetilde{\mathcal{A}}_{21} & \widetilde{\mathcal{A}}_{22},
	\end{bmatrix}
\end{equation*}
where $\widetilde{\mathcal{A}}_{11}$ denotes the block of $\widetilde{\mathcal{A}}$ with rows $r_1$ and columns $c_1$, $\widetilde{\mathcal{A}}_{12}$ denotes the block of $\widetilde{\mathcal{A}}$ with rows $r_1$ and columns $c_2$, and so on for $\widetilde{\mathcal{A}}_{21}$ and $\widetilde{\mathcal{A}}_{22}$.

We now loop through the four blocks, $\widetilde{\mathcal{A}}_{11}$, $\widetilde{\mathcal{A}}_{12}$, $\widetilde{\mathcal{A}}_{21}$, and $\widetilde{\mathcal{A}}_{22}$, and decide which blocks should be subdivided further. For the purpose of explanation, consider $\widetilde{\mathcal{A}}_{12}$. If
\begin{equation}
	\label{eq:weak_admissibility_cond}
	\dist\left(r_1, c_2\right) \ge \eta \min\left(\diam\left(r_1\right), \diam\left(c_2\right)\right),
\end{equation}
then we mark $\widetilde{\mathcal{A}}_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone. Here $\dist\left(r_1, c_2\right)$ is the Euclidean distance between the axis-aligned bounding box for the point cloud associated with the row cluster $r_1$, and the axis aligned bounding box for the point cloud associated with the column cluster $c_2$. The quantity $\diam\left(r_1\right)$ is the diameter of the axis aligned bounding box for the point cloud associated with the row cluster $r_1$, and $\diam\left(c_2\right)$ is the analogous diameter associated with the column cluster $c_2$. The quantity $\eta$ is a scalar constant; we use $\eta=2.0$. Basically, if the point clouds associated with $r_1$ and $c_2$ are far away from each other relative to their diameters, then we expect that the corresponding block of the matrix will be low rank. This process is repeated for the other blocks to determine which blocks are admissible and which are not. For us, the diagonal blocks $\widetilde{\mathcal{A}}_{11}$ and $\widetilde{\mathcal{A}}_{22}$ are not admissible because the distance between a point cloud and itself is zero.

Next, we subdivide all blocks that are not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\mathcal{\widetilde{A}}$. But now we subdivide a block based on the two child row clusters and two child column clusters for the rows and columns of that block. This subdivision process continues recursively until all blocks are either admissible, or smaller than the predetermined size mentioned above. The resulting hierarchical partitioning of matrix blocks forms a tree (quad-tree in $2$D, or oct-tree in $3$D), which is called the \emph{block cluster tree}. The root of the tree is the whole matrix, internal nodes in the tree are blocks that are subdivided, and the leaves of the tree are blocks that are either expected to be low rank, or are small.

%Note that the cluster trees and block cluster tree are computed using only the geometric information about the degree of freedom locations, and do not depend on the values of the matrix entries of $\widetilde{\mathcal{A}}$.

\subsubsection{Adaptive cross low rank approximation of blocks}
\label{sec:adaptive_cross}

Once the block cluster tree has been constructed, low rank approximations of the admissible (low rank) blocks are formed using the adaptive cross method [CITE]. Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\widetilde{\mathcal{A}}$. The idea of the adaptive cross method is to form a low rank approximation of $X$ by sampling a small number of rows and columns of $X$. 
\begin{itemize}
\item Let $C \in \mathbb{R}^{m \times r}$ be a matrix consisting of a subset of $r$ columns of $X$, such that the span of the columns in $C$ approximates the column space of $X$. 
\item Let $R\in\mathbb{R}^{r \times m}$ be a subset of the rows of $X$, such that the span of the rows in $R$ approximates the row space of $X$.
\item Let $U \in \mathbb{R}^{r \times r}$ be the block of $X$ consisting of the intersection of the rows from $R$ with the columns from $C$.
\end{itemize}
Then it is well-established that
\begin{equation}
\label{eq:CUR}
	X \approx C U^+ R,
\end{equation}
where $U^+$ is the pseudoinverse of $U$ [CITE]. The quality of approximation \eqref{eq:CUR} depends on how well the columns of $C$ approximate the column space of $X$, and how well the rows of $R$ approximate the row space of $X$.

In the adaptive cross method, ``good'' columns, $C$, and rows, $R$, are chosen via an alternating iterative process. An initial set of columns $C$ is chosen. Keeping $C$ fixed, a set of rows $R$ is chosen so that the so that determinant of the submatrix $U$ within $C$ is as large as possible. This may be done via either the maxvol procedure [CITE GOOD SUBMATRIX], or via QR factorization. Now, keeping $R$ fixed, a set of new columns $C$ is chosen so that the submatrix $U$ within $R$ is as large as possible. This process repeats a small number of times. This results in matrices $R$ and $C$ such that the error in the approximation, \eqref{eq:CUR}, is small. The dominant cost of this procedure is the cost of computing $kr$ columns of $X$ and $kr$ rows of $X$, where $k$ is the number of alternating iterations. There is also a small linear algebra overhead cost for the determinant maximization process that is performed at each step. For more details on adaptive cross low rank approximation, see [CITE]. The key point is that the adaptive cross method allows us to form a rank-$r$ approximation of an $m \times m$ block of the matrix via a process that only requires accessing $O(mr)$ entries of that block.

We use the adaptive cross method to form low rank approximations for each admissible block of $\widetilde{\mathcal{A}}$. We directly compute all entries of the small dense blocks of $\widetilde{\mathcal{A}}$ that are not admissible. This process requires us to compute $O(r^2 N \log N)$ entries of $\widetilde{\mathcal{A}}$, which is relatively cheap compared to the matrix-vector products with $\mathcal{A}$ that are required to form the product-convolution approximation $\widetilde{\mathcal{A}}$.

\section{Interpolating impulse response samples old}

We interpolate the functions $\varphi_{x_i}$ associated with the points $x_i \in S$, to generate functions $\varphi_x$ associated with new points $x \notin S$. This is done using a continuous generalization of the interpolative decomposition for matrices, wherein a select number of columns of a matrix are interpolated to fill the whole matrix. Interpolating $\varphi_x$ differs from interpolative low rank approximation of $\mathcal{A}$, which corresponds to interpolating $\langle\mathcal{A},\delta_x\rangle^*$. Because of the shifting that maps $\langle\mathcal{A}, \delta_x\rangle$ to $\varphi_x$, interpolation of $\varphi_{x_i}$ yields an approximation $\widetilde{\mathcal{A}}$ which can be high-rank.

The interpolation procedure requires a basis of smooth functions. We form this basis from smooth eigenmodes of the Laplacian on $\Omega$ with Neumann boundary conditions. The resulting approximation $\mathcal{A}$ may be interpreted as a weighted sum of convolution operators, with smooth spatially varying weighting functions. Kernel entries of the approximation, $\widetilde{A}(y,x)$, may be evaluated efficiently.

\subsection{Interpolative factorization of the SVIR}
\label{sec:interp_fact_svir}

The \emph{spatially varying impulse response}, denoted $T$, is the reinterpretation of $\varphi_x(y)$ as a function of both $y$ and $x$:
\begin{equation*}
	T(y,x) := \varphi_x(y) = A(y+x,x).
\end{equation*}
Since $T(y,x)$ is smooth as a function of $x$ (Condition 1), the continuous analog of the row space of $T$ should be well-approximated by the span of a small number of orthonormal smooth functions, $\{u_k\}_{k=1}^r$. Condition 1, therefore, is equivalent to the existence of a low rank approximation $\widetilde{T}$ of $T$ of the form 
\begin{equation}
	\label{eq:formula_for_A}
	T \approx \widetilde{T} = \Psi U
\end{equation}
where $\Psi:\mathbb{R}^d \times \{1,\dots,r\} \rightarrow \mathbb{R}$ is an unknown semi-discrete kernel with $r$ columns, and $U:\{1,\dots,r\} \times \Omega \rightarrow \mathbb{R}$ is the semi-discrete kernel with $r$ rows, such that the $i^\text{th}$ row of $U$ is the function $u_i$. That is,
\begin{equation*}
	U(i,x) = u_i(x).
\end{equation*}
The multiplication $K_1 K_2$ of kernels $K_1:X \times Y \rightarrow \mathbb{R}$ and $K_2:Y \times Z \rightarrow \mathbb{R}$ is defined as
\begin{equation*}
	\left(K_1 K_2\right)(x,z) := \sum_{y \in Y} K_1(x,y) K_2(y,z)
\end{equation*}
if $Y$ is a discrete set, and
\begin{equation*}
	\left(K_1 K_2\right)(x,z) := \int_Y K_1(x,y) K_2(y,z) dy	
\end{equation*}
if $Y\subset \mathbb{R}^d$ is a domain.

The number of smooth functions, $r$, required to well-approximate $T$ using \eqref{eq:formula_for_A} can be small even if the numerical rank of $\mathcal{A}$ is large. For example, let $\varphi$ be any convolution kernel, and suppose $\mathcal{A}$ is the convolution operator $\left(\mathcal{A}v\right)^* := \varphi \ast v$. The rank of $\mathcal{A}$ can be made arbitrarily large or small by changing the convolution kernel $\varphi$. At the same time, $T=\Psi U$ holds with $r=1$, where $\Psi(x,1):= \varphi(x)$ and $U(1,x):=1$.

We may determine $\Psi$ using a continuous analog of the process used to form the interpolative matrix factorization Let $S:=\{x_j\}_{j=1}^q \subset \Omega$ be the set of $q \ge r$ sample points determined by the method in Section \ref{sec:choosing_sample_points}, and define $T_S:\Omega \times \{1,\dots,q\}\rightarrow\mathbb{R}$ to be the semi-discrete kernel with $i^\text{th}$ column $\varphi_{x_i}$. That is,
\begin{equation*}
	T_S(y,i):= \varphi_{x_i}(y) = T(y,x_i).	
\end{equation*}
Further, let $U_S$ be the $r \times q$ matrix with $(i,j)$ entry given by $u_i(x_j)$, and assume that the points $x_i$ are chosen such that $U_S$ has full row rank. This is not difficult; for example, if the points $x_i$ were chosen uniformly at random then $U_S$ would have full rank almost surely. Indeed, with uniform random points, $\frac{|\Omega|}{q}\left(U_S U_S^T\right)_{ij}$ is a Monte-Carlo approximation of $\langle u_i, u_j \rangle_{L^2(\Omega)}$, where $|\Omega|$ is the total volume of the domain $\Omega$. Thus $\frac{|\Omega|}{q}\left(U_S U_S^T\right) \rightarrow I$ as $q \rightarrow \infty$, where $I$ is the identity matrix. This implies $\sigma_\text{min} \rightarrow \sqrt{q/|\Omega|}$ as $q \rightarrow \infty$, where $\sigma_\text{min}$ is the minimum singular value of $U_S$. We do not use uniform random sampling to choose the points, but the way we choose points does achieve relatively uniform coverage of the domain.

Restricting the second input of both sides of the equation \eqref{eq:formula_for_A} to points in $S$, we have
\begin{equation}
\label{eq:A_I_equals_X_U_I}
\widetilde{T}_S = \Psi U_S.
\end{equation}
Multiplying both sides of \cref{eq:A_I_equals_X_U_I} by the Moore-Penrose pseudoinverse of $U_S$, denoted $U_S^+$, and using the fact that $U_S$ has full row rank, yields
\begin{equation}
\label{eq:formula_for_X}
\widetilde{T}_S U_S^+ = \Psi.
\end{equation}
Substituting the left hand side of \eqref{eq:formula_for_X} for $\Psi$ into \eqref{eq:formula_for_A} yields the following interpolative-type low rank factorization $\widetilde{T}$ of $T$:
\begin{equation}
\label{eq:interpolative_decomposition0}
T \approx \widetilde{T} = T_S U_S^+ U.
\end{equation}
%We will quantify the accuracy of this approximation in Theorem \ref{thm:smooth_basis_err} in Section \ref{sec:error_estimate_adaptive_algorithm}.

%Briefly, the error will be small when (1) $\mathcal{U}$ well-approximates the row space of $\mathcal{T}$, and (2) the smallest singular value of $\mathcal{U}_X$ is not small.

\subsection{Smooth basis of Laplacian eigenmodes}

For the basis of smooth functions $u_i$, we use the dominant eigenfunctions of the inverse Laplacian on $\Omega$ with Neumann boundary conditions. This basis of inverse Laplacian eigenfunctions generalizes the Fourier basis to non-rectangular domains. Let $\Delta_N^{-1}$ denote the solution operator for the following PDE:
\begin{equation}
\label{eq:laplacian_pde}
\begin{cases}
-\Delta u =f, & \text{in } \Omega, \\
n \cdot \nabla u = 0, & \text{on } \partial \Omega, \\
\int_\Omega u dx = 0, &
\end{cases}	
\end{equation}
where $\Delta$ is the Laplacian and $n$ is the normal to $\partial \Omega$. That is, $\Delta_N^{-1}$ is the map from source $f$ to PDE solution $u$. The first smooth basis function used for interpolation is the constant function $u_1(x):=1$. The remaining smooth basis functions, $u_2, u_3, \dots, u_r$, are the eigenfunctions corresponding to the $r-1$ largest eigenvalues of $\Delta_N^{-1}$. These eigenfunctions may be computed efficiently using randomized SVD \cite{HMT11}, Krylov methods such as Lanczos, power iterations with deflation, or other methods. We use randomized SVD. All of these methods require the ability to apply $\Delta_N^{-1}$ to given vectors, which amounts to solving PDE \eqref{eq:laplacian_pde} with the given vector as the right hand side source. Highly efficient multigrid schemes exist for solving this PDE.
%; we use an algebraic multigrid implementation in PyAMG.

\subsection{Action of the approximate operator}

Defining
\begin{equation*}
	W := U_S^+ U,
\end{equation*}
we may write $\widetilde{T}$ as
\begin{equation*}
	\widetilde{T} = T_S W.	
\end{equation*}
Using the definition of $T_S$, we may write the entries of the resulting approximation to $A$ as
\begin{equation}
\label{eq:H_kernel_entries}
	\widetilde{A}(y,x) = \widetilde{T}(y-x,x) = \sum_{i=1}^q w_i(x) \varphi_{x_i}(y-x) 
\end{equation}
where the functions $w_i(x):=W(i,x)$ are the rows of $W$, which are linear combinations of the functions $u_i$.
%Since the Laplacian eigenfunctions $u_i$ are defined on $\Omega$, the weighting functions $w_i$ are defined on all of $\Omega$, so $T$ and $H$ are defined on $\mathbb{R}^d \times \Omega$. However, the error estimate we present in the next section will only be valid for these operators on $\mathbb{R}^d \times \Omega_I$. 
From \eqref{eq:H_kernel_entries}, we derive the action of $\widetilde{\mathcal{A}}$ as
\begin{align*}
	\left(\widetilde{\mathcal{A}}u\right)(v) &= \int_\Omega \int_\Omega v(y) \widetilde{A}(y,x)u(x) dx dy \\
	&= \int_\Omega \int_\Omega v(y) \widetilde{T}(y-x,x)u(x) dx dy \\
	&= \int_\Omega \int_\Omega v(y) \left(\sum_{i=1}^q  \varphi_{x_i}(y-x) w_i(x)\right) u(x) dx dy \\
	&= \int_\Omega v(y) \sum_{i=1}^q \int_\Omega \varphi_{x_i}(y-x) w_i(x) u(x) dx dy\\
	&= \int_\Omega v(y) \left(\sum_{i=1}^q \varphi_{x_i} \ast \left(w_i \cdot u\right)\right)(y) dy,
\end{align*}
which implies
\begin{equation*}
	\left(\widetilde{\mathcal{A}} u	\right)^* = \sum_{i=1}^q \varphi_{x_i} \ast \left(w_i \cdot u\right)
\end{equation*}
Here $f \ast g$ denotes the convolution of the functions $f$ and $g$. Thus, $\widetilde{\mathcal{A}}$ is a weighted sum of convolution operators with smooth spatially varying weighting functions.

\section{Adaptive algorithm}
\label{sec:error_estimate_adaptive_algorithm}

In this section we bound the error in the approximation to $T$ (Theorem \ref{thm:smooth_basis_err}), and use this bound to drive an adaptive algorithm (Algorithm \ref{alg:adaptive}) for constructing the approximation.

\begin{thm}
	\label{thm:smooth_basis_err}
	Let $\sigma_\text{min}$ be the smallest singular value of $U_S$, and let
	\begin{equation*}
		\gamma(q) := \sigma_\text{min} \sqrt{\frac{q}{|\Omega|}},
	\end{equation*}
	where $|\Omega|$ is the measure of $\Omega$. We have
	\begin{equation*}
		\|T - T_S U_S^+ U\|_2 \le \left(1 + \frac{|\Omega|}{\gamma(q)} \right)\|T - TU^TU\|_2,
	\end{equation*}
	where $\sigma_\text{min}$ is the smallest singular value of $U_S$, and $\nor{\mathcal{A}}$ is the induced norm for the operator $\mathcal{A}$.
\end{thm}

\begin{proof}
	Let $\widehat{T} := TU^TU$, and let $\widehat{T}_S:\mathbb{R}^d \times \{1,\dots,q\} \rightarrow \mathbb{R}$ be the semi-discrete kernel defined by
	\begin{equation*}
		\widehat{T}_S(y,i) = \widehat{T}_S(y,x_i).
	\end{equation*}
	Using the triangle inequality yields
	\begin{equation}
		\nor{T - T_S U_S^+ U}_2 \le \nor{T - \widehat{T}_S U_S^+ U}_2 
		+ \nor{\left(T_S - \widehat{T}_S\right) U_S^+ U}_2 \label{eq:two_bound_align}
	\end{equation}
	For the first term, note that $\widehat{T}_S = T U^T U_S$, so 
	\begin{equation*}
		\widehat{T}_S U_S^+ U = T U^T U_S U_S^+ U = T U^T U = \widehat{T}.
	\end{equation*}
	Thus,
	\begin{equation}
	\label{eq:A_nogamma}
	\nor{T - \widehat{T}_S U_S^+ U }_2 = \nor{T - \widehat{T}}_2.
	\end{equation}
	We estimate the second term as:
	\begin{equation*}
		\nor{\left(T_S - \widehat{T}_S\right) U_S^+ U}_2 \le \nor{T_S - \widehat{T}_S}_2 \nor{U_S^+}_2 \nor{U}_2 = \frac{1}{\sigma_\text{min}} \nor{T_S - \widehat{T}_S}_2	
	\end{equation*}	
%	\begin{align*}
%		\nor{\left(T_S - \widehat{T}_S\right) U_S^+ U}_2 &\le \nor{T_S - \widehat{T}_S}_2 \nor{U_S^+ U}_2 \\
%		&\le \sqrt{q}\nor{T_S - \widehat{T}_S}_1 \nor{U_S^+ U}_2 \\
%		&\le \sqrt{q}\nor{T_S - \widehat{T}_S}_1 \nor{U_S^+}_2 \nor{U}_2 = \frac{\sqrt{q}}{\sigma_\text{min}} \nor{T_S - \widehat{T}_S}_1. 
%	\end{align*}
	where we used the facts that $\nor{U_S^+}_2 = 1/\sigma_\text{min}$ and $\nor{U}_2=1$. By H\"older's inequality, we have 
	\begin{equation*}
		\nor{T - \widehat{T}}_{2} \le |\Omega|^{1/2} \nor{T - \widehat{T}}_{2,1}.
	\end{equation*}

	To estimate this term further, let $\delta_{x_i}^\epsilon \in L^1(\Omega)$ be an $\epsilon$-approximation to $\delta_{x_i}$ with norm $\|\delta_{x_i}^\epsilon\|_1=1$, defined as
	\begin{equation*}
		\delta_{x_i}^\epsilon(x) := \begin{cases}
			1 / \vol(B_\epsilon(x_i)), & x \in B_\epsilon(x_i), \\
			0, & \text{otherwise},
		\end{cases}
	\end{equation*}
	where $B_\epsilon(x_i)$ is the ball of radius $\epsilon$ centered at $x_i$, and $\vol(B_\epsilon(x_i))$ is the volume of this ball.
	We have
	\begin{align*}
		\nor{T_S - \widehat{T}_S}_1 &= \max_{i \in 1,2,\dots,q} \|T(~\cdot~,x_i) - \widehat{T}(~\cdot~,x_i)\|_1\\
		&= \|T(~\cdot~,x_k) - \widehat{T}(~\cdot~,x_k)\|_1 \\
		&= \lim_{\epsilon \rightarrow 0} \nor{(T-\widehat{T}) \delta_{x_k}^\epsilon}_1 \\
		&\le \sup_{\|v\|_1=1} \nor{(T - \widehat{T}) v}_1 = \nor{T - \widehat{T}}_1,
	\end{align*}
	where $k$ is the index achieving the maximum in the first line. Thus
	\begin{equation}
	\label{eq:_gamma_A_bound}
	\nor{\left(T_S - \widehat{T}_S\right) U_S^+ U}_2 \le \frac{\sqrt{q}}{\sigma_\text{min}}\nor{T - \widehat{T}}_1.
	\end{equation}
	The proposition follows from substituting bounds \eqref{eq:A_nogamma} and \eqref{eq:_gamma_A_bound} into \eqref{eq:two_bound_align}.
\end{proof}

Theorem \ref{thm:smooth_basis_err} shows that the error in our approximation depends on (1) how well the row space of $\mathcal{U}$ captures the row space of $\mathcal{T}$ (smaller $\|\mathcal{T} - \mathcal{T}\mathcal{U}^T\mathcal{U}\|$ $\implies$ less error), and (2) the minimum singular value of $\mathcal{U}_X$ (larger $\sigma_\text{min}$ $\implies$ less error). We can decrease $\|\mathcal{T} - \mathcal{T}\mathcal{U}^T\mathcal{U}\|$ by including more functions $u_i$ in $\mathcal{U}$. We can increase $\sigma_\text{min}$ by incorporating more batches of sample points $X_b$ in the approximation. But if $\sigma_\text{min}>1$, there are diminishing returns to making $\sigma_\text{min}$ bigger. By analogy to constrained optimization, $\|\mathcal{T} - \mathcal{T}\mathcal{U}^T\mathcal{U}\|$ is like an objective function to be minimized, while ensuring $\sigma_\text{min}$ is not small is like a constraint to be satisfied. We therefore increase the accuracy of the approximation by including more functions in $\mathcal{U}$, adding more batches of sample points as needed to ensure $\sigma_\text{min} > 1$.

The error in the approximation for batches $1,\dots,b$ is estimated with the the error for the next batch
\begin{equation*}
	\mathcal{E} := \frac{\nor{\eta^{b+1} - \widetilde{\eta}^{b+1}}}{\nor{\eta^{b+1}}},
\end{equation*}
where $\eta^{b+1}$ and $\widetilde{\eta}^{b+1}$ are the responses of $\mathcal{H}$ and $\widetilde{\mathcal{H}}$, respectively, to the Dirac comb at the next batch $\mathtt{S}_{b+1}$ of sample points, that were not used in the approximation $\widetilde{\mathcal{H}}$ yet. Entries of $\widetilde{\eta}_{b+1}$ take the form
\begin{equation}
\label{eq:eta_test}
	\widetilde{\eta}^{b+1}(y_k) = \sum_{x_j \in S_{b+1}} \widetilde{H}(y_k,x_j),
\end{equation}
which we compute using Algorithm \ref{alg:H_entries} for all degree of freedom locations $y_k$ for the discretization of the problem.
If $\mathcal{E}$ is larger than a desired tolerance, we compute more eigenmodes of $\Delta_N^{-1}$ and add those modes to the basis $\mathcal{U}$. Adding more functions to $\mathcal{U}$ may change $\sigma_\text{min}$. If $\sigma_\text{min}<1$ after this change, we select new batches of sample points and incorporate them in the approximation until $\sigma_\text{min} \ge 1$ once again. This process repeats until $\mathcal{E}$ is less than the desired tolerance. Each time an additional batch is needed, we compute another batch for testing, and incorporate the batch that was previously used for testing into the approximation. Once the desired accuracy is reached, we incorporate the last batch that was used for testing into the approximation to increase the accuracy further.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
%	\Input{Function $v \mapsto \mathcal{H}v$\\
%		Function $v \mapsto \mathcal{H}^Tv$
%	}

%	\mbox{}\\

%	\Output{Responses $\eta^1, \dots, \eta^b, \eta^\text{test}$ of $\mathcal{H}$ to Dirac combs \\
%		Basis of smooth functions $u_1, \dots, u_r$ \\
%		Interpolation pseudoinverse $\mathcal{U}_X^+$}
	%	\Begin{
	{

		Determine $\Omega_I$ with Algorithm \ref{alg:boundary_region}

		Compute $\mu$ and $\Sigma$ with Algorithm \ref{alg:varhpi_mean_cov}

		Select first sample point batch $S_1$ (Section \ref{sec:choosing_sample_points})

		Compute $\eta^1$ (Section \ref{sec:get_impulse_response})

		Set $S = \{\}$ and $b=0$

%		Form $\mathbf{X} \in \mathbb{R}^{d \times N}$ as the array of all degree of freedom coordinates
%		
%		$\mathtt{C} \gets \mathtt{I}$
%		
%		$b=0$
%		
%		$\mathtt{S}_1 \gets \choosebatch{\mathtt{C}}$ \tcp*{Algorithm \ref{alg:point_choice}}
%		
%		$\mathtt{C} \gets \mathtt{C} \setminus \mathtt{S}_1$
%		
%		$\boldsymbol{\eta}_1 \gets \computediracresponse{\mathtt{S}_1}$ \tcp*{Algorithm \ref{alg:dirac_comb_response}}
%		
%		$\mathtt{S} \gets \{\}$

		\For{$k=1,2,\dots,$}{

			Compute $u_k$ as the $k^\text{th}$ eigenfunction of $\Delta_N^{-1}$

%			\tcp{$U_S$ is the matrix with entries $\left(U_S\right)_{ij} = u_i(x_j)$ for $x_j\in S$}

%			\If{$U_S$ has less columns than rows}{
%				Select more sample point batches until $U_S$ has more columns than rows
%				
%				Compute functions $\eta^b$ for the new sample point batches
%			}

			\While{$|S| < k$}{

				Set $b=b+1$

				Set $S=S\cup S_b$

				Select new sample point batch $S_{b+1}$

				Form $\eta^{b+1}$

%				Form Dirac comb $\xi^b$ associated with $S_b$ and compute $\eta^b=\mathcal{A}\xi^b$ 
%				
%				Add new columns to $U_S$ corresponding to points $x_j \in S_b$
%				
%				Add new columns to $U_S$ corresponding to points $x_j \in S_b$
%				
%				$b \gets b+1$
%			
%				$\mathtt{S} \gets \mathtt{S} \cup \mathtt{S}_b$
%				
%				$\mathtt{S}_{b+1} \gets \choosebatch{\mathtt{C}}$
%				
%				$\mathtt{C} \gets \mathtt{C} \setminus \mathtt{S}_{b+1}$
%				
%				$\boldsymbol{\eta}_{b+1} \gets \computediracresponse{\mathtt{S}_{b+1}}$

			}

			Compute $\widetilde{\eta}^{b+1}$ using formulas \eqref{eq:eta_test} and \eqref{eq:kernel_entry_formua}.

			\If{$\displaystyle\frac{\nor{\eta^{b+1} - \widetilde{\eta}^{b+1}}}{\nor{\eta^{b+1}}}$ is sufficiently small}{

				Terminate for loop

			}

		}

		Set $S = S \cup S_{b+1}$

		Add new columns to $U_S$ corresponding to points $x_j \in S_b$

%		$\mathtt{Q} \gets \mathtt{Q} \cup \mathtt{Q}_\text{test}$

%		\Return{$\mu, \Sigma$, $\eta^1, \dots, \eta^b, \eta^\text{test}$, $u_1, \dots, u_r$, $\mathcal{U}_X^+$}

	}
	\caption{Adaptive approximation of $\mathcal{H}$}
	\label{alg:adaptive}
\end{algorithm2e}

\section{Hierarchical matrix construction}

\section{Numerical results}

\section{Conclusions}
\label{sec:conclusions}

Some conclusions here.

\appendix

\section{Fast ellipsoid intersection test}
\label{sec:fast_ellipsoid_intersection_test}
The procedure for choosing sample points relies on quickly determining whether two ellipsoids intersect. Let $E_p$ and $E_q$ be the ellipsoids defined as
\begin{align*}
	E_p :=& \{x : (x - \mu_p)^T \Sigma_p^{-1} (x - \mu_p) \le \tau^2\} \\
	E_q :=& \{x : (x - \mu_q)^T \Sigma_q^{-1} (x - \mu_q) \le \tau^2\}, \\
\end{align*}
where $\mu_p, \mu_q \in \mathbb{R}^d$, and $\Sigma_p, \Sigma_q \in \mathbb{R}^{d \times d}$ are positive definite. Let $K$ be the following one dimensional convex function:
\begin{equation*}
	K(s) := 1 - \frac{1}{\tau^2} (\mu_p - \mu_q)^T \left(\frac{1}{1-s}\Sigma_p + \frac{1}{s}\Sigma_q\right)^{-1}(\mu_p - \mu_q)	
\end{equation*}
In CITE it is shown that $E_p \cap E_q = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$. We check whether $E_p$ and $E_q$ intersect by minimizing $K(s)$ on $(0,1)$. If $K(s^*) <0$ at the minimizer $s^*$, then $E_p \cap E_q = \{\}$. Otherwise $E_p \cap E_q \neq \{\}$.

The function $K(s)$ may be evaluated quickly for many $s$ by pre-computing the solution to the generalized eigenvalue problem
\begin{equation*}
	\Sigma_p \Phi = \Sigma_q \Phi \Lambda,
\end{equation*}
where $\Phi \in \mathbb{R}^{d \times d}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_d)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$. The matrix $\Phi$ simultaneously diagonalizes $\Sigma_p$ and $\Sigma_q$, in the sense that $\Phi^T\Sigma_p\Phi = \Lambda$, and $\Phi^T\Sigma_q\Phi = I$, where $I$ is the $d \times d$ identity matrix. Using this diagonalization, and some algebraic manipulations, we may write $K(s)$ as
\begin{equation}
\label{eq:Ks_generalized}
K(s) = 1 - \frac{1}{\tau^2} \sum_{i=1}^d \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2,
\end{equation}
where $v := \Phi^T\left(\mu_p - \mu_q\right)$. We compute the generalized eigenvalue decomposition of $\Sigma_p$ and $\Sigma_q$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's algorithm (any fast 1 dimensional convex optimization routine may be used). The resulting algorithm for checking whether $E_p$ and $E_q$ intersect is summarized in Algorithm \ref{alg:ellipsoid_intersection_test}.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	\SetKwProg{Fn}{Function}{}{}
	\Input{Ellipsoid $E_p$ with mean $\mu_p$ and covariance $\Sigma_p/\tau^2$\\
		Ellipsoid $E_q$ with mean $\mu_q$ and covariance $\Sigma_q/\tau^2$ 
	}

	\Output{Boolean which is true if $E_p \cap E_q \neq \{\}$ and false otherwise}

	\mbox{}\\

	\Fn{$\ellipsoidsintersect{E_p}{E_q}$}
	{
		Solve generalized eigenvalue problem $\Sigma_p \Phi = \Sigma_q \Phi \Lambda$

		$v \gets \Phi^T\left(\mu_p - \mu_q\right)$

		$\displaystyle K^* \gets \min_{s \in (0,1)}~1 - \frac{1}{\tau^2} \sum_{i=1}^d \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2$

		\If{$K^* < 0$}{

			\Return{False}

		}
		\Else{

			\Return{True}

		}
	}
	\caption{Determining whether two ellipsoids intersect}
	\label{alg:ellipsoid_intersection_test}
\end{algorithm2e}

\section{Additional theory}
\label{app:proofs}

These results rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}. We prove correctness of Algorithm \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}. Theorem \ref{thm:wi_alg_is_correct} relies on a prerequisite result which we present in Lemma \ref{lem:psi_delta_eval}.

\begin{lem}[Preparation for Theorem \ref{thm:smoothest_interpolant}]
	\label{lem:T_is_linear}
	Let $f \in \mathbb{R}^r$. There exists a unique solution to the following optimization problem:
	\begin{equation}
	\label{eq:generic_optimization_initial}
	\begin{aligned}
	\min_{u \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r.
	\end{aligned}
	\end{equation}
	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
	\begin{equation*}
	\mathcal{T}(f) := u^*,
	\end{equation*}
	where $u^*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
\end{lem}
\begin{proof}
	asdf
\end{proof}

\begin{proof}[Proof of Theorem \ref{thm:smoothest_interpolant}]
	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero. The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
	\begin{align*}
	\mathcal{T}(f) &= \varphi_1(y) \mathcal{T}(e_1) + \varphi_2(y) \mathcal{T}(e_2) + \dots +\varphi_r(y) \mathcal{T}(e_r) \\
	&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
	&= u.
	\end{align*}
	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $u$ in \eqref{eq:u_defn_in_thm}.
\end{proof}

\begin{proof}[Proof of Proposition \ref{thm:wi_sum_to_one}]
	By linearity of $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), we have
	\begin{equation*}
	\sum_{i=1}^r w_i(x) = \sum_{i=1}^r \mathcal{T}(e_i)(x) = \mathcal{T}\left(\sum_{i=1}^r e_i\right)(x) = \mathcal{T}(\mathbf{1})(x) = 1,
	\end{equation*}
	where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one. The last equality ($\mathcal{T}(\mathbf{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \mathbf{1}$, and yields the minimum possible value of zero for the objective function.
\end{proof}

\begin{lem}[Preparation for Theorem \ref{thm:wi_alg_is_correct}]
	\label{lem:psi_delta_eval}
	We have
	\begin{equation}
	\label{eq:inline_wi_lemma}
	h(x_j) = \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \avg(h)
	\end{equation}
	for all $h\in H^2_N(\Omega)$, where
	\begin{equation*}
	\avg(h) :=	 \frac{1}{|\Omega|} \int_\Omega h ~dx,
	\end{equation*}
	and $|\Omega| = \int_\Omega 1 ~dx$ is the Lebesgue measure of $\Omega$.
\end{lem}

\begin{proof}
	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\psi_j$ onto $h$, and using the Neumann zero boundary conditions for $\psi_j$ and $h$ to eliminate boundary integral terms.
\end{proof}

\begin{proof}[Proof of Theorem \ref{alg:weighting_functions_incremental}]
	For notational convenience, in this proof we write $w$, $\lambda$, and $e$ instead of $w_i$, $\lambda_i$, and $e_i$, respectively. Here subscripts for $\lambda$, and $e$ instead denote components of those vectors. For example, here $\lambda_j$ denotes the $j^\text{th}$ component of the vector $\lambda$, which corresponds to the $j^\text{th}$ component of the vector $\lambda_i$ in Algorithm \ref{alg:weighting_functions_incremental}.

	From standard optimization theory, the solution to optimization problem \ref{eq:wi_optimization_problem} is the stationary point of the Lagrangian,
	\begin{equation*}
	\mathcal{L}(w,\lambda) = \frac{1}{2}\|\Delta w\|^2 + \sum_{j=1}^r \lambda_j\left(w(x_j) - e_j\right).
	\end{equation*}
	The plan of this proof is to show that $\nabla \mathcal{L}(w,\lambda)=0$, which will verify that $w$ solves optimization problem \ref{eq:wi_optimization_problem}.

	Using the definition of $w_i$ in \eqref{eq:definition_of_wi_alg}, we may write the gradient of $\mathcal{L}$ with respect to $\lambda_j$ as
	\begin{equation*}
	\nabla_{\lambda_j} \mathcal{L}(w, \lambda) = w(x_j) - e_j
	= \alpha - \sum_{k=1}^r \lambda_k \theta_k(x_j) - e_j.
	\end{equation*}
	Using Lemma \ref{lem:psi_delta_eval} with $h=\theta_k$, the equation for $\theta_k$ in \eqref{eq:theta_eq}, and the definition of $S$ in \eqref{eq:defn_of_S_alg}, we have
	\begin{align*}
	\sum_{k=1}^r \lambda_k \theta_k(x_j) &= \sum_{k=1}^r \lambda_k \left(\psi_j, \Delta \theta_k\right)_{L^2(\Omega)}\\
	&= \sum_{k=1}^r \lambda_k \left(\psi_j, \psi_k\right)_{L^2(\Omega)} = \left(S \lambda\right)_j.
	\end{align*}
	Combining all of these components of $\nabla_\lambda \mathcal{L}$ into a single vector and using the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} yields
	\begin{align*}
	\nabla_\lambda \mathcal{L}(w, \lambda) &= \alpha \mathbf{1} - e - S \lambda \\
	&= \alpha \mathbf{1} - e - S \left(-S^{-1}\left(e - \alpha \mathbf{1}\right)\right) = 0.
	\end{align*}

	Taking the directional derivative of $\mathcal{L}$ with respect to $w$ in direction $h$ and using Lemma \ref{lem:psi_delta_eval} yields
	\begin{align}
	\nabla_w \mathcal{L}(w, \lambda)h &= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j h(x_j) \nonumber \\
	&= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j \left(\psi_j, \Delta h\right)_{L^2(\Omega)} + \lambda_j \avg(h) \nonumber \\
	&= \left(\Delta w + \sum_{j=1}^r \lambda_j \psi_j,~ \Delta h\right)_{L^2(\Omega)} + \avg(h) \mathbf{1}^T \lambda. \label{eq:L_w_38}
	\end{align}
	Using the definition of $w$ in Algorithm \ref{alg:weighting_functions_incremental}, and the fact that the Laplacian of a constant function is zero, we have
	\begin{equation*}
	\Delta w = -\sum_{j=1}^r \lambda_j \Delta \theta_j 
	= -\sum_{j=1}^r \lambda_j \psi_j,
	\end{equation*}
	which implies that the first term in \eqref{eq:L_w_38} is zero. By the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} and $\alpha$ in \eqref{eq:definition_of_alpha_alg}, we have
	\begin{align*}
	\mathbf{1}^T \lambda &= -\mathbf{1}^T S^{-1} \left(e - \alpha \mathbf{1}\right) \\
	&= -\mathbf{1}^T S^{-1} e + \alpha \mathbf{1}^T S^{-1} \mathbf{1} \\
	&= -\mathbf{1}^T S^{-1} e + \frac{\mathbf{1}^T S^{-1} e}{\mathbf{1}^T S^{-1} \mathbf{1}} \mathbf{1}^T S^{-1} \mathbf{1} = 0,
	\end{align*}
	which implies that the second term in \eqref{eq:L_w_38} is zero. Thus $\nabla_w \mathcal{L}(w, \lambda) = 0$, so Algorithm \ref{alg:weighting_functions_incremental} produces the solution to optimization problem \ref{eq:wi_optimization_problem}.
\end{proof}

%\begin{proof}[Proof of Proposition \ref{prop:boundary_source}]
%	Let $w \in L^2(\Omega)$ be arbitrary. We have
%	\begin{align*}
%		\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle(w) =& \int_\Omega w(y) \int_{\partial \Omega} A(y,x) dy dx \\
%		=& \int_\Omega w(x) \int_{\partial \Omega} \varphi_x(y-x) dy dx = \left(w, z\right)_{L^2(\Omega)},
%	\end{align*}
%	where
%	\begin{equation*}
%		z(x) := \int_{\partial \Omega} \varphi_x(y-x) dy,	
%	\end{equation*}
%	which implies
%	\begin{equation*}
%		\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle^*(x) = z = \int_{\partial \Omega} \varphi_x(y-x) dy.
%	\end{equation*}
%	Since $\varphi_x$ is a non-negative function, the result follows from the fact that the integral of a non-negative function over a set of positive measure is zero if and only if the function is zero almost everywhere on that set.
%\end{proof}

\begin{proof}[Proof of Theorem \ref{thm:vol_mean_cov}]
	Let $w \in L^2$ be arbitrary. We have
	\begin{align*}
		\left(w, \alpha\right)_{L^2(\Omega)} =& \int_\Omega w(x) \int_{\Omega} \langle\mathcal{A},\delta_x\rangle^*(z) dz dx\\
		=& \int_\Omega w(x) \int_{\Omega} A(z,x) dz dx \\
		=& \int_\Omega \int_\Omega w(x) A(z,x) C(z) dz dx \\
		=& \left(\mathcal{A}^T C\right)(w),
	\end{align*}
	which implies \eqref{eq:vol_mean_var_thm1}.

	Using similar techniques, we have
	\begin{align*}
		\left(w, \alpha \cdot \mu^i\right)_{L^2(\Omega)} =& \int_\Omega w(x) \alpha(x) \int_{\Omega} z^i \langle\mathcal{A},\delta_x\rangle^*(z)/\alpha(x) dz dx\\
		=& \int_\Omega \int_\Omega w(x) L^i(z) A(z,x) dz dx \\
		=& \left(\mathcal{A}^T L^i\right)(w),
	\end{align*}
	which implies \eqref{eq:vol_mean_var_thm2}.

	To prove \eqref{eq:vol_mean_var_thm3} we use the identity
	\begin{equation*}
		\Sigma(x) = \int_\Omega zz^T \rho_x(z) dz - \mu(x) \mu(x)^T.
	\end{equation*}
	We have
	\begin{align*}
		\left(w, \alpha \cdot \left(\Sigma^{ij}+\mu^i \cdot \mu^j\right)\right)_{L^2(\Omega)} =& \int_\Omega w(x) \alpha(x) \int_\Omega z^i z^j \langle\mathcal{A},\delta_x\rangle^*(z) / \alpha(x) dz dx \\
		=& \int_\Omega \int_\Omega w(x) Q^{ij}(z) A(z,x) dz dx \\
		=& \left(\mathcal{A}Q^{ij}\right)(w),
	\end{align*}
	which implies \eqref{eq:vol_mean_var_thm3}.

\end{proof}

\section{Discretization}
\label{sec:discretization}

Let $\phi_i$, $i=1,\dots,N$ be a set of finite element basis functions, and let
\begin{equation*}
V_h = \Span\left(\phi_1, \phi_2, \dots, \phi_N\right).
\end{equation*}
be the corresponding  finite element space approximating $L^2(\Omega)$. Here $V_h$ inherits the $L^2$ inner product, and $h$ denotes the mesh size parameter. Functions $u\in L^2(\Omega)$ are approximated by functions $u_h \in V_h$. In turn, functions $u_h\in V_h$ are represented in computations by length $N$ arrays $\mathbf{u}$, such that the array entries of $\mathbf{u}$ are the coordinates of $u_h$ in the finite element basis. That is,
\begin{equation*}
u(x) \approx u_h(x) = \sum_{i=1}^N \mathbf{u}_i \phi_i(x).
\end{equation*}
We say that the coefficient arrays $\mathbf{u}$ reside $\mathbb{R}^N_\mathbf{M}$, which is the space $\mathbb{R}^N$ with the matrix-weighted inner product $\left(\mathbf{u},\mathbf{v}\right)_\mathbf{M} := \left(\mathbf{M} \mathbf{u}\right)^T \mathbf{v}$, where
$\mathbf{M} \in \mathbb{R}^{N \times N}$ is the \emph{mass matrix} which has entries
\begin{equation*}
\mathbf{M}_{ij} = \left(\phi_i, \phi_j\right)_{L^2(\Omega)} = \int_{\Omega} \phi_i(x) \phi_j(x) dx.
\end{equation*}
Direct calculation shows that $V_h$ is isometrically isomorphic $\mathbb{R}^N_\mathbf{M}$, in the sense that functions $u_h \in V_h$ are in bijective correspondence to coefficient vectors $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, and $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u},\mathbf{v}\right)_\mathbf{M}$. Likewise, the dual space of continuous linear functionals $\psi_h:V_h \rightarrow \mathbb{R}$, denoted $V_h'$, is isometrically isomorphic to $\mathbb{R}^N_{\mathbf{M}^{-1}}$. All operations on functions in $V_h$ and linear functionals in $V_h'$ may therefore be performed on coefficient vectors in $\mathbb{R}^N_\mathbf{M}$ and coefficient dual vectors in $\mathbb{R}^N_{\mathbf{M}^{-1}}$. We summarize coefficient vector versions of the function space operation that are used in this paper in Appendix \ref{app:discretized_operations}.

\subsection{Discretized operations}
\label{app:discretized_operations}

As described in Section \ref{sec:discretization}, a function $u_h$ in a finite element space $V_h$ may be written in terms of its vector of coefficients, $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, with respect the finite element basis $\{\phi_i\}_{i=1}^N$. Recall that $\mathbf{M}$ is the mass matrix. We have
\begin{equation*}
	u(x) \approx u_h(x) = \sum_{i=1}^N \mathbf{u}_i \phi_i(x).
\end{equation*}
The finite element function space $V_h$ is isometrically isomorphic to the coefficient space $\mathbb{R}^N_\mathbf{M}$, and the dual space $V_h'$ is isometrically isomorphic to $\mathbb{R}^N_{\mathbf{M}^{-1}}$. In computations, we perform operations with coefficient vectors in $\mathbb{R}^N_\mathbf{M}$ and coefficient dual vectors in $\mathbb{R}^N_{\mathbf{M}^{-1}}$, rather than functions in $V_h$ and functionals in $V_h'$. Here we describe how to perform common function space operations in terms of coefficient vectors and coefficient dual vectors.

\begin{description}
	\item[Coefficients of a functional:] A linear functional $\psi_h \in V_h'$ has coefficient dual vector $\boldsymbol{\psi} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$, which has entries
	\begin{equation*}
		\boldsymbol{\psi}_i = \psi_h(\phi_i), \quad i=1,\dots,N.
	\end{equation*}
	The action of a linear functional $\psi_h \in V_h'$ on a function $u_h \in V_h$ is given by
	\begin{equation*}
		\psi_h(v_h) = \mathbf{v}^T \boldsymbol{\psi}.
	\end{equation*}
	\item[Riesz representation of a functional:] Let $\psi_h \in V_h'$. We have
	\begin{equation*} 
		 \psi_h^* \in V_h: \quad \left(\psi_h^*, v_h\right)_{L^2(\Omega)} = \psi_h(v_h) \quad\forall v_h \in V_h \quad \Leftrightarrow \quad \boldsymbol{\psi}^* = \mathbf{M}^{-1} \boldsymbol{\psi}.
	\end{equation*}
	The function $\psi_h^*$ is the Riesz representation of the functional $\psi_h$, and $\boldsymbol{\psi}^* \in \mathbb{R}^N_\mathbf{M}$ is the coefficient vector of $\psi_h^*$.
	\item[$L^2$ projection:] Let $f \in L^2(\Omega)$, and define $\mathbf{f}^* \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ to be the vector with entries
	\begin{equation*}
		\mathbf{f}^*_i = \int_\Omega f(x) \phi_i(x) dx, \quad i=1,\dots,N.
	\end{equation*}
	We have
	\begin{equation*}
		f_h \in V_h, \quad \left(f_h, v_h\right)_{L^2(\Omega)} = \left(f, v_h\right)_{L^2(\Omega)} \quad\forall v_h \in V_h \quad \Leftrightarrow \quad \mathbf{f} = \mathbf{M}^{-1}\mathbf{f}^*.	
	\end{equation*}
	The function $f_h$ is the $L^2$ projection of $f$ onto $V_h$, and $\mathbf{f} \in \mathbb{R}^N_\mathbf{M}$ is the coefficient vector of $f_h$.
	\item[Matrix representation of an operator:] Let $\mathcal{B}_h : V_h \rightarrow V_h'$ be a linear operator, and let $\mathbf{B} \in \mathbf{R}^{N \times N}$ be the matrix with entries
	\begin{equation*}
		\mathbf{B}_{ij} = \left(\mathcal{B}_h \phi_j\right)(\phi_i).
	\end{equation*}
	The matrix $\mathbf{B}$ maps $\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}$ by matrix multiplication, and
	\begin{equation*}
		\psi_h = \mathcal{B}_h u_h \quad \Leftrightarrow \quad \boldsymbol{\psi} = \mathbf{B} \mathbf{u}.
	\end{equation*}
	\item[Transpose:] The matrix $\mathbf{B}^T$ maps $\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}$ via matrix multiplication, and
	\begin{equation*}
		\psi_h = \mathcal{B}_h^T u_h \quad \Leftrightarrow \quad \boldsymbol{\psi} = \mathbf{B}^T \mathbf{u}.
	\end{equation*}
%	\item[Implicitly defined matrix representation:]
%	Let $\mathcal{A}_h:V_h \rightarrow V_h'$ be a finite dimensional approximation of $\mathcal{A}:L^2(\Omega) \rightarrow L^2(\Omega)'$, and let $\mathbf{A}$ be the matrix representation of $\mathcal{A}_h$. The matrix $\mathbf{A}$ is implicitly defined, in the sense that we have a algorithm to compute the matrix-vector products $\mathbf{u} \mapsto \mathbf{A}\mathbf{u}$ and $\mathbf{A}^T \mathbf{u}$ for arbitrary $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, but we do not have direct access to matrix entries $\mathbf{A}_{ij}$. In principle, all matrix entries of $\mathbf{A}$ could be computed by applying $\mathbf{A}$ to each column of the identity matrix, but this is far too expensive to be practical.
	\item[Distributions:] Let $\mu:C\left(\overline{\Omega}\right)\rightarrow \mathbb{R}$ be a distribution. If $\mu(\phi_i)$ is well-defined for all basis vectors $\phi_i$, then the restriction of $\mu$ to domain $V_h$ is a linear functional $\mu_h\in V_h'$. The functional $\mu_h$ has a coefficient dual vector $\boldsymbol{\mu} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ with entries $\boldsymbol{\mu}_i=\mu(\phi_i)$ for $i=1,\dots,N$. 
	\item[Action of an operator on a distribution:] Since $V_h$ is finite-dimensional, the restriction, $\mu_h \in V_h$, of a distribution, $\mu$, has a a Riesz representation $\mu_h^* \in V_h$. Let $B_h$ be the integral kernel associated with an operator $\mathcal{B}_h:V_h \rightarrow V_h'$. For $w_h \in V_h$, we have
	\begin{align*}
		\langle \mathcal{B}_h, \mu_h \rangle(w) &= \int_\Omega w_h(y) \mu_h\left(B_h(y, \cdot)\right) dy \\
		&= \int_\Omega w_h(y) \left(\int_\Omega B_h(y,x) \mu_h^*(x) dx\right) dy \\
		&= \int_\Omega \int_\Omega w_h(y) B_h(y,x) \mu_h^*(x) dx dy = \left(\mathcal{B}_h \mu_h^*\right)(w).
	\end{align*}
	Going from the first to the second line we used the definition of the Riesz representation. The action of $\mathcal{B}_h$ on $\mu_h$ in the sense of distributions is therefore equal to the action of $\mathcal{B}_h$ on $\mu^*_h$ in the conventional sense. We have
	\begin{equation*}
		\psi_h = \langle \mathcal{B}_h, \mu_h \rangle \quad \Leftrightarrow \quad \psi_h = \mathcal{B}_h \mu_h^* \quad \Leftrightarrow \quad \boldsymbol{\psi} = \mathbf{B} \mathbf{M}^{-1} \boldsymbol{\mu}.
	\end{equation*}
\end{description}

\subsection{Discretized mean and covariance estimation} In computations, $\alpha$, $\mu^{i}$, and $\Sigma^{ij}$ are replaced with finite element approximations $\alpha_h, \mu_h^i, \Sigma_h^{ij}\in V_h$, respectively, which have coefficient vectors $\boldsymbol{\alpha}, \boldsymbol{\mu}^{ij}, \boldsymbol{\Sigma}^{ij}\in\mathbb{R}^N_\mathbf{M}$, respectively. The functions $C$, $L^i$, and $Q^{ij}$ are replaced with their $L^2$ projections onto $V_h$, which are denoted by $C_h$, $L_h^i$, and $Q_h^{ij}$, respectively. The corresponding coefficient vectors are $\mathbf{C}$, $\mathbf{L}^i$, and $\mathbf{Q}^{ij}$, respectively. The discretized versions of \eqref{eq:vol_mean_var_thm1}, \eqref{eq:vol_mean_var_thm2}, and \eqref{eq:vol_mean_var_thm3} are
\begin{align*}
\boldsymbol{\alpha} &= \mathbf{M}^{-1}\mathbf{A}^T\mathbf{C} \\
\boldsymbol{\mu}^i &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{L}^i\right) / \boldsymbol{\alpha}\\
\boldsymbol{\Sigma}^{ij} &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{Q}^{ij}\right) / \boldsymbol{\alpha} - \boldsymbol{\mu}^i\cdot \boldsymbol{\mu}^j,
\end{align*}
respectively. Here $\mathbf{f} \cdot \mathbf{g}$ and $\mathbf{f} / \mathbf{g}$ denote the element-wise multiplication and division of vectors, respectively.

\subsection{Discretized sampling of impulse responses}
In computations, the Dirac comb $\xi^b$ is replaced by its restriction to $V_h$, denoted $\xi^b \in V_h'$. The response $\eta^b$ is replaced by the finite element approximation $\eta_h^b \in V_h$. The coefficient dual vector for $\xi_h^b$ is the vector $\boldsymbol{\xi}^b \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has entries
\begin{equation*}
\boldsymbol{\xi}^b_i = \sum_{x_j \in S_b} \phi_i(x_j), \quad i=1,\dots,N.
\end{equation*}
The coefficient vector $\boldsymbol{\eta}^b \in \mathbb{R}^N_\mathbf{M}$ for $\eta_h^b$ is given by
\begin{equation*}
\boldsymbol{\eta}^b = \mathbf{M}^{-1} \mathbf{A}^T \mathbf{M}^{-1}\boldsymbol{\xi}^b.
\end{equation*}

\section{old stuff}

\subsection{Background: distributions}

Let $\overline{\Omega}$ be the closure of $\Omega$, and let $C(\overline{\Omega})$ be the space of continuous functions mapping $\overline{\Omega}\rightarrow \mathbb{R}$. The action of $\mathcal{A}$ is extended to distributions $\mu:C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ via the formula
\begin{equation*}
\langle\mathcal{A},\mu\rangle(w) := \int_\Omega w(y) \mu\left(A(y, \cdot)\right) dy, 
\end{equation*}
where $A(y,\cdot)$ is the function $x \mapsto A(y,x)$. This is derived formally as follows:
\begin{align*}
\langle\mathcal{A},\mu\rangle(w) &= \int_\Omega \int_\Omega w(y) A(y,x) \textrm{``} \mu(x) \textrm{''} dx dy \\
&= \int_\Omega w(y) \int_\Omega A(y,x) \textrm{``} \mu(x) \textrm{''} dx \\
&= \int_\Omega w(y) \mu\left(A(y,\cdot)\right) dx.
\end{align*}
For example, the delta distribution $\delta_x$ is defined by $\delta_x(v) = v(x)$, the action of $\mathcal{A}$ on $\delta_x$ is given by
\begin{equation*}
\langle\mathcal{A},\delta_x\rangle(w)	= \int_\Omega w(y) A(y,x) dy,
\end{equation*}
and so the Riesz representation of $\langle\mathcal{A},\delta_x\rangle$ is given by $\langle\mathcal{A},\delta_x\rangle^* = A(~\cdot~,x)$. That is, $\langle\mathcal{A},\delta_x\rangle^*$ is the function $y \mapsto A(y,x)$.

\begin{defn}
	\label{def:stiffness_mass}
	We define $\mathcal{K}:H^2(\Omega)\rightarrow L^2(\Omega)'$ and $\mathcal{M}:L^2(\Omega) \rightarrow L^2(\Omega)'$ to be the following linear operators:
	\begin{align*}
	\left(\mathcal{K}u\right)(v) :=& \int_\Omega v \Delta u ~dx \\
	\left(\mathcal{M}u\right)(v) :=& \int_\Omega u v ~dx,
	\end{align*}
	and $\mathcal{P}:H^2(\Omega)\rightarrow \mathbb{R}^r$ to be the following pointwise observation operator:
	\begin{equation*}
	(\mathcal{P}u)_i = u(x_i)
	\end{equation*}
	for $i=1,\dots,r$.
\end{defn}

\begin{prop}
	\label{prop:solution_operator_is_linear}
	Let $f \in \mathbb{R}^r$ be arbitrary. The optimization problem
	\begin{equation}
	\label{eq:generic_optimization_initial_old}
	\begin{aligned}
	\min_{u \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r
	\end{aligned}
	\end{equation}
	may be rewritten as follows:
	\begin{equation}
	\label{eq:generic_optimization}
	\begin{aligned}
	\min_{u \in H^2(\Omega)} &\quad \frac{1}{2} \left(\mathcal{K}^T\mathcal{M}^{-1}\mathcal{K}u\right)(u) \\
	\text{such that} &\quad 
	\mathcal{P} u = f.
	\end{aligned}
	\end{equation}
\end{prop}

\begin{prop}
	There exists a unique solution to optimization problem \eqref{eq:generic_optimization}. The solution to \eqref{eq:generic_optimization}, $u^*$, and the adjoint variable for enforcing the constraint, $\lambda^*$, solve the following Karun-Kush-Tucker (KKT) system:
	\begin{equation}
	\label{eq:weighting_kkt}
	\begin{bmatrix}
	\mathcal{K}^T\mathcal{M}^{-1}\mathcal{K} & \mathcal{P}^T \\
	\mathcal{P} & 0
	\end{bmatrix}
	\begin{bmatrix}
	u^* \\ \lambda^*
	\end{bmatrix}
	=
	\begin{bmatrix}
	0 \\ f
	\end{bmatrix},
	\end{equation}
	where we use block-matrix notation to denote block-operators.
\end{prop}

\begin{cor}
	\label{cor:linearity_of_solution_operator}
	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization}	as a function of $f$. That is, 
	\begin{equation*}
	\mathcal{T}(f) := u^*,
	\end{equation*}
	where $u^*$ is the minimizer of \eqref{eq:generic_optimization}. The operator $\mathcal{T}$ is a linear operator.
\end{cor}

\footnote{The reason for this dependence on spatial dimension is related to the fact that the fundamental solution of the biharmonic equation is continuous in $3$ or less dimensions, but singular in $4$ or more dimensions. In the unusual scenario where one is inverting for a parameter field in $4$ or more spatial dimensions, one could use a modified Poisson interpolation where $\Delta$ is replaced with $\Delta^\alpha$ in the objective function of \eqref{eq:wi_optimization_problem}, where $\alpha>1$ is chosen so that the fundamental solution of $\Delta^{2\alpha}$ is continuous.}

% $\left(\mathcal{A}u\right)(v) = \left\langle v, \varphi_p \ast u\right\rangle_{L^2(\Omega)}$. Here $f \ast g$ denotes convolution of the functions $f$ and $g$, and the convolution kernel,

%Access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$.
%Integral kernel representation \eqref{eq:kernel_representation} is assumed to exist,
%e.g., by the Schwarz kernel theorem,
%but o
%One cannot easily access entries $A(y,x)$ of the kernel.

% is non-negative in the sense that  $A(y,x) \ge 0$ for all $(y,x) \in \Omega \times \Omega$. 
%Let $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ be the corresponding linear operator generated by currying. That is, for each $u\in L^2(\Omega)$, the linear functional $\mathcal{A}u:L^2(\Omega)\rightarrow \mathbb{R}$ is 
%We approximate the operator $\mathcal{A}:L^2(\Omega)\rightarrow L^2(\Omega)'$ defined by
%\begin{equation*}
%	(\mathcal{A}u)(v) := a(v,u).
%\end{equation*}
%We seek to approximate operators $\mathcal{A}$ in the ``matrix free'' setting, in which access to $\mathcal{A}$ is available through application of $\mathcal{A}$ and $\mathcal{A}^T$ to arbitrary vectors, $u \mapsto\mathcal{A}u$ and $v \mapsto\mathcal{A}^Tv$. Here ($\mathcal{A}^Tv)(u) := (\mathcal{A}u)(v)$. 
%Integral kernel representation \eqref{eq:kernel_representation} is assumed to exist,
%e.g., by the Schwarz kernel theorem,
%but one cannot easily access entries $A(y,x)$ of the kernel. We assume that the impulse response of $\mathcal{A}$ is locally supported in the sense that the function 
%$y \mapsto A(y, x)$ is supported in a region near $y=x$, and we assume that $\mathcal{A}$ is locally translation invariant in the sense that
%\begin{equation*}
%A(y+h, x+h) \approx A(y,x),
%\end{equation*}
%so long as $h$ is not too large.

The \emph{point spread function} (PSF) of $\mathcal{A}$ at $x\in \Omega$ is
\begin{equation*}
\varphi_x(y) := \langle \mathcal{A},\delta_x\rangle^*(y+x) = A(y+x,x).
\end{equation*}
Here $\langle \mathcal{A}, \delta_x \rangle \in L^2(\Omega)'$ is the result of applying $\mathcal{A}$ to a point source $\delta_x$ (delta distribution) centered at $x$, and $\langle\mathcal{A},\delta_x\rangle^* \in L^2(\Omega)$ is the Riesz representation of $\langle\mathcal{A},\delta_x\rangle$. The PSF, $\varphi_x$, is the result of translating $\langle\mathcal{A},\delta_x\rangle^*$ to re-center it at zero instead of $x$. Recall the Riesz representation of a linear functional $\psi \in L^2(\Omega)'$ is the unique vector $\psi^* \in L^2(\Omega)$ satisfying $\left(\psi^*, v\right)_{L^2(\Omega)} = \psi(v)$ for all $v \in \L^2(\Omega)$.

%In this paper we present an efficient matrix-free method for approximating operators $\mathcal{A}$ with a product-convolution approximation, $\widetilde{A}$, of the form
%\begin{equation*}
%	\left(\mathcal{A}u\right) \approx \left(\widetilde{\mathcal{A}}u\right) := \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right)
%\end{equation*}
%
%
%PSFs that satisfy the following properties:
%\begin{enumerate}
%	\item $\varphi_x$ varies smoothly as a function of $x$.
%	\item $\varphi_x(y)$ is zero (or small) when $|y|$ is large.
%	\item $\varphi_x$ is a non-negative (or mostly non-negative) function.
%\end{enumerate}
%If $\varphi_x$ was constant in $x$, then $\mathcal{A}$ would be the convolution operator $\left(\mathcal{A}v\right)^* = \varphi_0 \ast v$. Operators satisfying property 1 therefore act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies smoothly as $x$ changes. Operators satisfying property 2 are local, in the sense that sources at point $x$ have no impact on points far away from $x$. Operators satisfying property 3 blur inputs: their action on a point source consists of scaling the source, then spreading the energy of the scaled point source across a region nonuniformly. We call operators satisfying these properties \emph{smoothly varying blur operators}.

% Computing the functions $w_i$ requires solving two Poisson PDEs for each point $x_i$, and these Poisson PDEs may be solved cheaply with multigrid.

%Specifically, since the integral kernel for $\mathcal{A}$ is non-negative, $\langle\mathcal{A},\delta_x\rangle^*$ may be interpreted as a scaled probability distribution. We compute the mean and covariance of this probability distribution for all points $x \in \Omega$ simultaneously, by applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions. We then estimate the support of $\langle\mathcal{A},\delta_x\rangle^*$ to be contained within a small number of standard deviations of the mean of this probability distribution.

%We interpret $\langle\mathcal{A},\delta_{x_i}\rangle^*$ as a scaled probability distribution, and estimate the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ to be contained within a small number of standard deviations of the mean of this probability distribution (this is why we require a non-negative kernel).  The mean and covariance of $\langle\mathcal{A},\delta_{x}\rangle^*$ can be simultaneously computed for all points $x \in \Omega$ by applying $\mathcal{A}^T$ to a small number of constant, linear, and quadratic functions.

%The functions $w_i$ are the defined as the smoothest functions, in a certain least-squares sense, that interpolate the points $x_i$. Computing the functions $w_i$ requires solving two Poisson PDEs for each point $x_i$, and these Poisson PDEs may be solved cheaply with multigrid.

%The idea of our method is to sample $\varphi_{x}$ at a collection of points $x_i$ by applying $\mathcal{A}$ to a sum of point sources at the points $x_i$ (a ``Dirac comb''). The points $x_i$ are chosen so that the support of $\langle\mathcal{A},\delta_{x_i}\rangle^*$ and the support of $\langle\mathcal{A},\delta_{x_j}\rangle^*$ do not overlap (or do not overlap much) if $i \neq j$. Since the point sources do not overlap, by applying $\mathcal{A}$ to one vector, we recover $\varphi_{x_i}$ for many points $x_i$. The process is repeated to get $\varphi_{x_i}$ at more points, until an error estimate is satisfied. The $\varphi_{x_i}$ are interpolated smoothly to construct $\varphi_x$ for arbitrary $x$, which yields an approximation $\widetilde{A}(y,x)$ to the kernel $A(y,x)$, and an approximation $\widetilde{\mathcal{A}}$ to the operator $\mathcal{A}$. While $A(y,x)$ is inaccessible, we can evaluate $\widetilde{A}(y,x)$ at arbitrary $(y,x)$ points. This allows us to construct a hierarchical matrix representation of $\widetilde{\mathcal{A}}$, which allows us to invert $\widetilde{\mathcal{A}}$, or perform other useful matrix operations, using fast hierarchical matrix arithmetic.

%Properties 1 and 2 are essential to our method. Property 3 is only required in the subprocedure for estimating the support of $\langle\mathcal{A},\delta_x\rangle^*$, and may be dropped if other methods for estimating the support are available.

The approximation $\widetilde{A}(y,x)$ may not be accurate if both $x$ and $y$ are near the boundary. Other approximation methods, if available, should be considered for the block of the operator associated with boundary-boundary interactions. Nevertheless, in our numerical results we see that approximation works well for the the inverse problem considered, without any modifications to the boundary-boundary block.

The solution to this system of equations is found by the following process:
\begin{enumerate}
	\item Compute 
	\begin{equation*}
	\Psi := K^+ E^T
	\end{equation*}
	by solving Neumann poisson problems for point source right hand sides, with point sources located at the points $x_i$, $i=1,\dots,q$.
	\item Compute 
	\begin{equation*}
	\Theta := K^+ M \Psi
	\end{equation*}
	by solving Neumann Poisson problems with the columns of $\Psi$ as distributed sources.
	\item Form 
	\begin{equation*}
	S = \Psi^T M \Psi
	\end{equation*}.
	\item Compute \begin{equation*}
	\alpha = \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{y}.
	\end{equation*}
	\item Compute 
	\begin{equation*}
	\boldsymbol{\lambda} = -S^{-1} \left(\mathbf{y} - \alpha \mathbf{1} \right).
	\end{equation*}
	\item Compute
	\begin{equation*}
	\mathbf{u} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c}
	\end{equation*}
\end{enumerate}

Consider the Pine island glacier geometry shown in Figure REF. Regions of the glacier are separated by thin inlets of ocean. With radial basis function interpolation, points on one side of the inlet inappropriately influence the interpolation on the other side of the inlet. With Poisson interpolation, the PSF samples only influence the interpolation on their own side of the inlet.

\subsection{Determination of the boundary region}
\label{sec:boundary}

Define $\Omega_B \subset \Omega$ to be the set of points that are too close to $\partial \Omega$, in the sense that $x \in \Omega_B$ if $\varphi_x(y-x) \neq 0$ for some $y \in \partial \Omega$. Further, define $\Omega_I := \Omega \setminus \Omega_B$ as the set of points that are in the interior of $\Omega$ and are not too close to $\partial B$. In Proposition \ref{prop:boundary_source} we show that $\Omega_B$ can be determined by applying $\mathcal{A}^T$ to a boundary source distribution and post-processing the result. The resulting algorithm for determining $\Omega_B$ and $\Omega_I$ in the discretized setting is shown in Algorithm \ref{alg:boundary_region}.

\begin{prop}
	\label{prop:boundary_source}
	Let $\delta_{\partial \Omega}: C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ be the boundary source distribution defined by
	\begin{equation}
	\label{eq:boundary_source}
	\delta_{\partial \Omega}(v) = \int_{\partial \Omega} v(y) dy
	\end{equation}
	for all $v \in C(\overline{\Omega})$. We have $\varphi_x(y-x) = 0$ for almost every $y \in \partial \Omega$ if and only if $\langle\mathcal{A}^T, \delta_{\partial \Omega}\rangle^*(x) = 0$.
\end{prop}

The proof of Proposition \ref{prop:boundary_source} is shown in Appendix \ref{app:proofs}.

%Straightforward analysis shows that $\varphi_x(y-x) = 0$ for $y \in \partial \Omega$ almost everywhere if and only if $\left(\mathcal{A}^T \delta_{\partial \Omega}\right)(x) = 0$, where $\delta_{\partial \Omega}$ is the uniform boundary source distribution. That is, the distribution that acts on a test function $v$ by performing a surface integral of $v$ over the boundary of the domain:
%%\begin{equation}
%%\label{eq:boundary_source}
%%	\delta_{\partial \Omega}(v) = \int_{\partial \Omega} v(y) dy.
%%\end{equation}
%We therefore compute $b  = \mathcal{A}^T \delta_{\partial \Omega}$, then identify $\Omega_B$ as the set of points $x \in \Omega$ for which $|b(x)|$ is larger than a small tolerance $\epsilon$. This is described in Algorithm \ref{alg:boundary_region}.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	{

		Construct boundary source $\boldsymbol{\delta}_{\partial \Omega} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$, $\displaystyle \left(\boldsymbol{\delta}_{\partial \Omega}\right)_i = \int_{\partial \Omega} \phi_i(s) ds$, $i=1,\dots,N$

		%		\For{$i=1,\dots,N$}{
		%			$\displaystyle \mathbf{b}_i \gets \int_{\partial \Omega} \phi_i(s) ds$
		%		}

		$\mathbf{b} = \mathbf{M}^{-1}\mathbf{A}^T\mathbf{M}^{-1}\boldsymbol{\delta}_{\partial \Omega}$

		$\Omega_B = \{x \in \Omega : |b_h(x)| \ge \epsilon\}$

		$\Omega_I = \Omega \setminus \Omega_B$

	}
	\caption{Compute boundary region $\Omega_B$ and interior region $\Omega_I$}
	\label{alg:boundary_region}
\end{algorithm2e}

One should only expect $\widetilde{A}(y,x)$ to be accurate for $x \in \Omega_I$, $y \in \Omega$. If $\mathcal{A}$ is symmetric then one may overwrite potentially inaccurate entries $\widetilde{A}(y,x)$, $x \in \Omega_b$, $y \in \Omega_I$, with the accurate entries $\widetilde{A}(x,y)$. If $\mathcal{A}$ is nonsymmetric, one may form an accurate approximation of these entries by applying our procedure to $\mathcal{A}^T$. Our method will furnish an approximation $\widetilde{A}(y,x)$ for $x \in \Omega_B$, $y \in \Omega_B$ (the block of $\widetilde{H}$ corresponding to boundary-boundary interactions), but this part of the approximation is unlikely to satisfy Condition 1 ($\varphi_x$ varies smoothly as a function of $x$).

\subsection{Computation of Poisson weighting functions}
\label{sec:computation_of_weighting_functions}

%Radial basis function interpolation of $\varphi_x(y)$ is cheaper per sample point because radial basis functions are given by analytic formulas. However, the cost of these Poisson PDE solves is typically cheap compared to the cost of computing the impulse response samples, $\varphi_{x_i}$. Computing the impulse response samples, $\varphi_{x_i}$, requires applying the Hessian to vectors, which in turn requires performing PDE solves of the linearized forward and ajoint equations for the inverse problem. These forward and adjoint PDE solves are typically much more expensive than Poisson PDE solves, because the Poisson equation is well-studied and easy to solve with multigrid.

After discretization, optimization problem \eqref{eq:wi_optimization_problem} becomes
\begin{equation}
\label{eq:discrete_poisson_optimization_1}
\begin{aligned}
\min_{\mathbf{w}} &\quad \frac{1}{2}\mathbf{w}^T K^T M^{-1} K \mathbf{w} \\
\text{such that} &\quad E \mathbf{w} = \mathbf{d}.
\end{aligned}
\end{equation}
Here $\mathbf{w}$ is the discretization of $w_i$, $K$ is the stiffness matrix for the discretized Poisson problem with pure Neumann boundary conditions, $M$ is the mass matrix, and $\mathbf{d}=(0,\dots,0,1,0,\dots,0)$ is the length-$q$ vector with $i^\text{th}$ entry one and all other entries zero. Furthermore, $E$ is the pointwise observation matrix defined by $\left(E\mathbf{u}\right)_j = u(x_j)$, where $u$ denotes the function represented by the vector $\mathbf{u}$. Here we dropped the subscripts $i$ from variables for notational convenience. One must solve $q$ optimization problems of the form \eqref{eq:discrete_poisson_optimization_1}; one for each $i=1,\dots,q$.

In the remainder of this section we will derive an efficient algorithm for solving \eqref{eq:discrete_poisson_optimization_1}. The algorithm (Algorithm REF) is based on explicit formula for the solution of \eqref{eq:discrete_poisson_optimization_1} which we present in Proposition \ref{prop:poisson_interpolation_formula}. The formulas in Proposition \ref{prop:poisson_interpolation_formula} rely on certain matrices that we define in Definition \ref{defn:poisson_interpolation_matrices}. The primary computational cost of the algorithm is the solution of Poisson problems with pure Neumann boundary conditions.

\begin{defn}
	\label{defn:poisson_interpolation_matrices}
	Let 
	\begin{align*}
	\Psi :=& K^+ E^T, \\
	\Theta :=& K^+ M \Psi \\
	S :=& \Psi^T M \Psi,
	\end{align*}
	where $K^+$ is the Moore-Penrose pseudoinverse of $K$. Further, let $\mathbf{c}$ dente the discretization of the constant function $c(x)=1$ on $\Omega$, and let $\mathbf{1}:=(1,\dots,1)$ denote the vector of length $q$ with all entries equal to one.
\end{defn}

\begin{remark}
	Computing $\mathbf{g} = K^+\mathbf{f}$ is equivalent to solving the discretized Poisson problem with $f$ as the right hand side source, with pure Neumann boundary conditions, and with the constraint that the solution has average value equal to zero. That is,
	\begin{equation}
	\label{eq:pure_neumann_poisson}
	\mathbf{g} = K^+ \mathbf{f} \quad \Leftrightarrow \quad 
	\begin{cases}
	\Delta g = f, & \text{in }\Omega,\\
	\nu \cdot \nabla g = 0, & \text{on }\partial \Omega,\\
	\int_\Omega g(x) dx = 0,
	\end{cases}
	\end{equation}
	where $\nu$ denotes the normal to the boundary.

	Hence, the $i^\text{th}$ column of $\Psi$ is the result of solving the discretized pure Neumann Poisson problem, with a point source at location $x_i$. The $i^\text{th}$ column of $\Theta$ is the result of solving the discretized pure Neumann Poisson problem with distributed source given by the $i^\text{th}$ column of $\Psi$.
\end{remark}

\begin{prop}
	\label{prop:poisson_interpolation_formula}
	The solution to optimization problem \eqref{eq:discrete_poisson_optimization_1} is given by
	\begin{equation*}
	\mathbf{w} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c},
	\end{equation*}
	where
	\begin{align*}
	\alpha =& \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{d} \\
	\boldsymbol{\lambda} =& -S^{-1} \left(\mathbf{d} - \alpha \mathbf{1} \right). \\
	\end{align*}
\end{prop}

\begin{proof}
	Let
	\begin{equation*}
	v := \mathcal{K} w_i,
	\end{equation*}
	and for the remainder of this proof let $e := e_i$.
	Since the null-space of the Laplacian with pure Neumann boundary conditions is the set of all constant functions, we have
	\begin{equation}
	\label{eq:definition_of_w}
	w_i = \mathcal{K}^+ v + \alpha C
	\end{equation}
	for some scalar $\alpha$.

	Writing optimization problem \eqref{eq:wi_optimization_problem} in terms of $v$ and $\alpha$ instead of $w_i$ yields the following equivalent optimization problem:
	\begin{equation}
	\label{eq:discrete_poisson_optimization_2}
	\begin{aligned}
	\min_{v \in L^2(\Omega)', \alpha \in \mathbb{R}} &\quad \frac{1}{2} v\left(\mathcal{M}^{-1} v\right) \\
	\text{such that} &\quad \mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} = e.
	\end{aligned}
	\end{equation}
	The Lagrangian for optimization problem \eqref{eq:discrete_poisson_optimization_2} is
	\begin{equation*}
	\mathcal{L} = \frac{1}{2}v\left( \mathcal{M}^{-1} v\right) + \lambda^T\left(\mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} - e\right).
	\end{equation*}
	The solution to this optimization problem is the stationary point of the Lagrangian, i.e., the point at which the gradient of the Lagrangian is zero:
	\begin{equation*}
	0 = \nabla \mathcal{L} = 
	\begin{bmatrix}
	\nabla_{v} \mathcal{L} \\
	\nabla_{\lambda} \mathcal{L} \\
	\nabla_\alpha \mathcal{L}
	\end{bmatrix} 
	= \begin{bmatrix}
	\mathcal{M}^{-1} v + \left(\mathcal{K}^+\right)^T \mathcal{P}^T \lambda \\
	\mathcal{P}\mathcal{K}^+v + \alpha \mathbf{1} - e \\
	\mathbf{1}^T \lambda
	\end{bmatrix}.
	\end{equation*}
	This equation may be rewritten as the following block $3 \times 3$ linear system:
	\begin{equation*}
	\begin{bmatrix}
	\mathcal{M}^{-1} & \left(\mathcal{K}^+\right)^T \mathcal{P}^T & 0 \\
	\mathcal{P}\mathcal{K}^+ & 0 & \mathbf{1} \\
	0 & \mathbf{1}^T & 0
	\end{bmatrix}
	\begin{bmatrix}
	v \\ \lambda \\ \alpha
	\end{bmatrix}
	=
	\begin{bmatrix}
	0 \\ e \\ 0
	\end{bmatrix}.
	\end{equation*}
	Performing block Gaussian elimination on this system allows us to reduce the system to the following block triangular form
	\begin{equation*}
	\begin{bmatrix}
	\mathcal{M}^{-1} & \left(\mathcal{K}^+\right)^T \mathcal{P}^T & 0 \\
	& -S & \mathbf{1} \\
	0 & 0 & \mathbf{1}^T S^{-1} \mathbf{1}
	\end{bmatrix}
	\begin{bmatrix}
	v \\ \lambda \\ \alpha
	\end{bmatrix}
	=
	\begin{bmatrix}
	0 \\ e \\ \mathbf{1}^T S^{-1} e
	\end{bmatrix}
	\end{equation*}
	where we recall that $S := \Psi^T \mathcal{M} \Psi = \mathcal{P}\mathcal{K}^+ \mathcal{M} \left(\mathcal{K}^+\right)^T \mathcal{P}^T$. From this block triangular system, we read off the solution as:
	\begin{align*}
	\alpha =& \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{d} \\
	\boldsymbol{\lambda} =& -S^{-1} \left(\mathbf{d} - \alpha \mathbf{1} \right) \\
	\mathbf{v} =& -M K^+ E^T \boldsymbol{\lambda}
	\end{align*}
	Substituting these results into the definition of $\mathbf{w}$ in \eqref{eq:definition_of_w}, we have
	\begin{equation*}
	\mathbf{w} = -K^+ M K^+ E^T \boldsymbol{\lambda} + \alpha \mathbf{c} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c},
	\end{equation*}
	as required.
\end{proof}

We now design an incremental algorithm for constructing interpolants $\mathbf{u}$, in which we can add points $x_i$ more efficiently. Let
\begin{align*}
\Psi :=& K^+ E^T \\
\Theta :=& K^+ M \Psi
\end{align*}
so that 
\begin{equation*}
S = \Psi^T M \Psi
\end{equation*}
and
\begin{equation*}
\mathbf{u} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c}
\end{equation*}
The columns of $\Psi$ are the responses of $A^+$ to point sources (delta distributions) centered at the sample points $x_i$. The columns of $\Theta$ are the responses of $A^+M^{-1} A$ to these point sources. The matrices $\Psi$ and $\Theta$ may be constructed incrementally as new points are added.

\section*{Acknowledgments}
We thank J.J. Alger for helpful discussions

\bibliographystyle{siamplain}
\bibliography{references}
\end{document}

### 2024-02-24 presentations/siamuq24-noemi/localpsf_revised_figures/localpsf_old2.tex
noemis slides

% SIAM Article Template
\documentclass[review,onefignum,onetabnum]{siamart190516}

% Information that is shared between the article and the supplement
% (title and author information, macros, packages, etc.) goes into
% ex_shared.tex. If there is no supplement, this file can be included
% directly.

\input{localpsf_shared}

% Optional PDF information
\ifpdf
\hypersetup{
  pdftitle={Fast matrix-free approximation of smoothly varying blur operators, with application to Hessians in PDE-constrained inverse problems with highly informative data},
  pdfauthor={N. Alger, N. Petra, T. Hartland, and O. Ghattas}
}
\fi

% The next statement enables references to information in the
% supplement. See the xr-hyperref package for details.

\externaldocument{localpsf_supplement}

% FundRef data to be entered by SIAM
%<funding-group specific-use="FundRef">
%<award-group>
%<funding-source>
%<named-content content-type="funder-name"> 
%</named-content> 
%<named-content content-type="funder-identifier"> 
%</named-content>
%</funding-source>
%<award-id> </award-id>
%</award-group>
%</funding-group>

\begin{document}

\maketitle

% REQUIRED
\begin{abstract}
  We present an efficient matrix-free method for approximating locally translation invariant operators that have locally supported non-negative integral kernels. Such operators arise, for example, as Schur complements in Schur complements methods for solving partial differential equations (PDEs), Poincare-Steklov operators in domain decomposition methods, covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems. The method computes impulse responses of the operator at a collection of scattered points, then interpolates these computed impulse responses to form a product-convolution approximation of the operator. The product-convolution approximation is converted to hierarchical matrix format in order to solve linear systems and perform other matrix operations efficiently. Impulse responses are computed by applying the operator to a small number of Dirac combs associated with ``batches'' of point sources. The key innovation of our method is a matrix-free procedure for choosing as many points as possible per batch, while ensuring that the supports of the impulse responses within each batch do not overlap. We apply the method to approximate Hessians in large-scale PDE-constrained inverse problems with highly informative data. Numerical results demonstrate that our method substantially outperforms existing state-of-the-art Hessian approximation methods which are based on low-rank approximation. Our method is able to form high quality approximations of high rank Hessians using only a small number of Hessian matrix-vector products.
\end{abstract}

% REQUIRED
\begin{keywords}
  example, \LaTeX
\end{keywords}

% REQUIRED
\begin{AMS}
  68Q25, 68R10, 68U05
\end{AMS}

\section{Introduction}
\label{sec:intro}

We present a fast matrix free method for approximating locally translation-invariant operators, $\Aop$, that have locally supported non-negative integral kernels. Such operators arise throughout computational science. Examples include Schur complements in Schur complements methods for solving partial differential equations (PDEs) \cite{TOBYPAPER}, Poincare-Steklov operators in domain decomposition methods (e.g., Dirichlet-to-Neumann maps), covariance operators in spatial statistics, blurring operators in imaging, and Hessians in distributed parameter PDE-constrained optimization and inverse problems.

Let $\Omega \subset \mathbb{R}^d$ be a bounded domain in dimension $d=1$, $2$, or $3$. We consider integral operators $\Aop:L^2(\Omega)\rightarrow L^2(\Omega)'$ of the form
\begin{equation}
\label{eq:kernel_representation}
(\Aop u)(v) := \int_\Omega \int_\Omega v(y) \Aker(y,x) u(x) dx dy,
\end{equation}
where $\Aker:\Omega \times \Omega \rightarrow \mathbb{R}$ is the integral kernel. Here $\Aop u \in L^2(\Omega)'$ is the linear functional that results from applying $\Aop$ to $u\in L^2(\Omega)$, and $\left(\Aop u\right)(v)$ is the scalar that results from applying that linear functional to $v \in L^2(\Omega)$. We further narrow our focus to operators which possess the following four properties:
\begin{description}
\item[1) Matrix-free:] One cannot easily evaluate kernel entries $\Phi(y,x)$. Instead, access to $\Aop$ is available through ``operator actions.'' That is, we have a black-box computational procedure (which is typically expensive) through which we may evaluate the maps
\begin{equation*}
	u \mapsto\Aop u \quad \text{and} \quad v \mapsto\Aop^Tv
\end{equation*} 
for arbitrary functions $u$ and $v$. After discretization (see Appendix \ref{app:discretized_operations}), $\Aop$ becomes a dense matrix, $\mathbf{A}$, that is typically too large to be built and stored. This property is called ``matrix-free'' because at the discrete level it means that one can perform matrix-vector products of $\mathbf{A}$ and $\mathbf{A}^T$ with arbitrary vectors, but one cannot easily\footnote{One could compute $\mathbf{v} = \mathbf{A}\mathbf{e}_j$, where $\mathbf{e}_j=(0,\dots,0,1,0,\dots,0)$ is the unit vector which has $j^\text{th}$ component equal to one and all other components equal to zero, then extract $\mathbf{A}_{ij} = \mathbf{v}_i$. However, this process is wasteful because one computes the entire vector $\mathbf{v}$, then discards all but one component.} access matrix entries $\mathbf{A}_{ij}$. In applications, applying $\mathbf{A}$ to a vector may involve iteratively solving a large linear system, timestepping, or performing some other nontrivial computational procedure. For example, in reduced space approaches to PDE-constrained optimization and inverse problems, Hessian information is available only through the application of the Hessian to vectors, and one such Hessian application requires solving two PDEs \cite{HESSIANADJOINT}. 
\item[2) Local support:] Let $\phi_x$ be the following \emph{impulse response}:
\begin{equation}
\label{eq:impulse_response_defn}
\phi_x(y) := \Phi(y, x).
\end{equation} 
We say that $\Aop$ has local support if, for all $x\in \Omega$, the support of $\phi_x$ is contained (or approximately contained) in a neighborhood of $x$. The smaller these supports are, the better our algorithm will perform.  Let $\delta_x$ denote the delta distribution (point source) centered at $x$. The function $\phi_x$ is called the impulse response because straightforward analysis shows that it may be expressed as follows:
\begin{equation}
\label{eq:impulse_response_delta_action}
\phi_x = \left( \Aop \delta_x^* \right)^*,
\end{equation}
where $\Aop \delta_x^* \in L^2(\Omega)'$ is an abuse of notation that denotes the result of applying the operator $\Aop$ to the distribution $\delta_x$, and $\left( \Aop \delta_x^* \right)^* \in L^2(\Omega)$ denotes the Riesz representation of $\Aop \delta_x^*$ with respect to the $L^2$ inner product.\footnote{Recall that the Riesz representative of a functional $\psi \in L^2(\Omega)'$ with respect to the $L^2$ inner product is the unique function $\psi^* \in L^2(\Omega)$ such that $\psi(v) = \left(\psi^*,v\right)_{L^2(\Omega)}$ for all $v \in L^2(\Omega)$.} Note that while the domain of $\Aop$ is defined as $L^2(\Omega)$, which does not contain $\delta_x$, if $\Phi$ is sufficiently regular then the action of $\Aop$ may be extended to distributions. This is described in Appendix \ref{app:distributions}. 
%The action of $\mathcal{A}$ on a distribution $\mu$ is written as $\langle \mathcal{A}, \mu \rangle$. We use this angle bracket notation to distinguish the action of $\mathcal{A}$ on distributions from the action of $\mathcal{A}$ on functions.
%We take the Riesz representative of the linear functional $\left\langle \mathcal{A}, \delta_x \right\rangle \in L^2(\Omega)'$ to convert it into a function $\left\langle \mathcal{A}, \delta_x \right\rangle^* \in L^2(\Omega)$. 
%After discretization, the process of applying $\Aop$ to the distribution $\delta_x$, then forming the Riesz representative of the resulting linear functional, amounts to computing $\boldsymbol{\phi}_x = \mathbf{M}^{-1}\mathbf{A} \mathbf{M}^{-1} \boldsymbol{\delta}_x$, where $\boldsymbol{\phi}_x$, $\mathbf{A}$, and $\boldsymbol{\delta}_x$ are discretized versions of $\phi_x$, $A$ and $\delta_x$, respectively, and $\mathbf{M}$ is a discretization of the $L^2$ inner product bilinear form (e.g., a mass matrix in the finite element context). 
From \eqref{eq:impulse_response_delta_action}, we see that this local support property means that the response of $\Aop$ to a point source at $x$ is zero (or small) at points $y$ that are far from $x$.
\item[3) Local translation invariance:] We say that $\Aop$ is locally translation invariant if
\begin{equation}
	\label{eq:local_translation_invariance}
	\Phi(y+h, x+h) \approx \Phi(y,x)
\end{equation}
when $h$ is not too large. One can imagine the kernel entry $\Phi(y,x)$ as representing the strength of an interaction between a source at point $x$ and a target at point $y$. Local translation invariance means that the interaction strength remains roughly the same if the source and target points are both translated by the same vector. This is illustrated in Figure FIG.
\item[4) Non-negative kernel:] For all $(y,x) \in \Omega \times \Omega$, we have
\begin{equation*}
\Phi(y,x) \ge 0.
\end{equation*}
\end{description}

Our overall strategy is to form a product-convolution approximation, $\AopPc$, of $\Aop$ using only operator actions, then convert $\AopPc$ into a hierarchical matrix (H-matrix), which we denote by $\AmatPc$. Schematically,
\begin{equation*}
\Aop \longrightarrow \underbrace{\AopPc}_{\substack{\text{product}\\ \text{convolution}}} \longrightarrow \underbrace{\AmatPc}_{\text{H-matrix}}.
\end{equation*}
We will define product-convolution approximations in Section \ref{sec:prodconv_intro}. H-matrices are discussed in Appendix \ref{app:h_matrix}. Our method is summarized in Section \ref{sec:overview_intro}, and described in detail in Section \ref{sec:method}.

The novel contributions of our method are in the way that we form the product-convolution approximation ($\Aop \rightarrow \AopPc$). This is the part of the method where we use the four properties of $\Aop$ described above. Fast conversion of $\AopPc$ to H-matrix format ($\AopPc \rightarrow \AmatPc$) is possible because integral kernel entries for $\AopPc$ are easily accessible (unlike integral kernel entries for $\Aop$). Once in H-matrix format, one may use H-matrix methods to perform further useful linear algebra operations involving $\AmatPc$, including matrix-matrix addition, matrix-matrix multiplication, matrix-vector products, matrix factorization, and matrix inversion. These H-matrix methods are fast and scale well to large problems.

\subsection{Motivation: inverse problems with highly informative data}
\label{sec:PDE_hessian_motivation}

Our motivation for this work is approximation of Hessians in distributed parameter inverse problems governed by partial differential equations (PDEs). That is, inverse problems in which one seeks to reconstruct an unknown parameter field, $m$, from noisy observations, 
\begin{equation*}
y=f(m,u(m)) + \text{noise},
\end{equation*}
which depend on  a state variable $u$. In turn, $u$ depends on $m$ implicitly through the solution of a PDE, which we write generically as follows:
\begin{equation}
\label{eq:state_pde}
0=g(m,u).
\end{equation}
Here $u(m)$ denotes the solution of the PDE \eqref{eq:state_pde} as a function of $m$. In the deterministic approach to inverse problems, one typically finds $m$ as the solution to the minimization problem 
\begin{equation}
\label{eq:minimization_problem}
\min_m \quad \frac{1}{2}\|y - f(m,u(m))\|_W^2 + R(m),
\end{equation}
where $\|\cdot\|_W$ is weighted norm which depends on the noise covariance, and $R(m)$ is a regularization term. The objective function in optimization problem \eqref{eq:minimization_problem} is defined indirectly via the implicit function theorem. As a result, the Hessian of the objective function, $\mathcal{H}$, is only accessible matrix-free. In particular, applying $\mathcal{H}$ to a vector requires solving two PDEs \cite{HESSIANACTION}. More accessible approximations of $\mathcal{H}$ are highly desirable, because they allow for fast solution of \eqref{eq:minimization_problem} via Newton-type methods. More accessible Hessian approximations are also central to many methods for uncertainty quantification in Bayesian statistical approaches to the inverse problem, because $\mathcal{H}^{-1}$ locally approximates the Bayesian posterior covariance for $m$. 
%We only have access to $\mathcal{H}$ via its action on an arbitrary vectors $v$, i.e., evaluation of the map $v \mapsto \mathcal{H} v$. Evaluating $v \mapsto \mathcal{H} v$ is an expensive process that involves solving two auxiliary linear PDEs.

The most popular existing Hessian approximation methods are based on forming a low rank approximation of the data misfit term in the Hessian, or the data misfit term preconditioned by the prior term \cite{CCGOPAPERS}. Conventionally, either the Lanczos method or the randomized singular value decomposition \cite{HMTRANDOM} are used to perform the low rank approximation using only matrix-vector products. These methods suffer from a ``data predicament''---if the data are highly informative about the unknown parameter, then the numerical rank of the data misfit term in the Hessian is large, so a large number of matrix-vector products are required to form the aforementioned low rank approximation. The ideal scenario from a scientific perspective (highly informative data) is therefore the worst case scenario from a computational perspective (large computational cost) \cite{MYDISSERTATION}. Although the method we present in this paper is not applicable to all Hessians, it is applicable to several Hessians of practical interest. For these Hessians, our method offers a \emph{data-scalable} alternative to conventional low-rank Hessian approximation methods, because our method can form high-rank approximations of an operator using a small number of matrix-vector products.

\subsection{Product-convolution approximation}
\label{sec:prodconv_intro}

If $\Aop$ were perfectly translation invariant (i.e., if equality held in \eqref{eq:local_translation_invariance} for all $x$, $y$, $h$) then it is straightforward to show that $\Aop$ would be the convolution operator $u \mapsto \left(\varphi_x \ast u\right)^*$, where the convolution kernel is given by
\begin{equation}
\label{eq:convolution_kernel}
	\varphi_x(y) := \phi_x(y+x).
\end{equation}
The convolution kernel $\varphi_x$ is the result of translating the impulse response $\phi_x$ to re-center it at zero instead of $x$. Locally translation invariant operators act like convolution operators locally, but the convolution kernel, $\varphi_x$, varies as $x$ changes. We therefore approximate $\Aop$ by a spatially varying weighted sum of convolution operators, $\AopPc \approx \Aop$, of the form
\begin{equation}
\label{eq:product_convolution}
	\left(\AopPc~u\right)(v) := \left( v,~ \sum_{i=1}^r \varphi_i \ast \left(w_i \cdot u\right) \right)_{L^2(\Omega)},
\end{equation}
where $f \cdot g$ denotes pointwise multiplication, $f \ast g$ denotes convolution, and $\left(f, g\right)_{L^2(\Omega)}$ denotes the $L^2$ inner product on $\Omega$, for functions $f,g \in L^2(\Omega)$. Approximations of the form \eqref{eq:product_convolution} are known as \emph{product-convolution} approximations, because the action of each term in the sum consists of a pointwise product, followed by a convolution. To avoid nested subscripts, we write 
\begin{equation*}
\varphi_i := \varphi_{x_i} \quad \text{and} \quad \phi_i := \phi_{x_i},
\end{equation*}
with $\varphi_{x_i}$ defined in \eqref{eq:convolution_kernel} and $\phi_{x_i}$ defined in \eqref{eq:impulse_response_defn}, to denote local convolution kernels and impulse responses corresponding to a collection of points $\{x_i\}_{i=1}^r \subset \Omega$ scattered throughout the domain. The functions $w_i$ are spatially varying weighting functions that are used to interpolate the convolution kernels $\varphi_i$. Our operator approximation method is defined by how we compute the convolution kernels, $\varphi_i$, how we choose the points, $x_i$, and what weighting functions, $w_i$, that we use. In Section \ref{sec:overview_intro} we summarize the answers to these questions; we will provide detailed answers to these questions in Section \ref{sec:method}.

The more locally translation invariant an operator is (i.e., the smaller the discrepancy between the left hand side and right hand side in \eqref{eq:local_translation_invariance}), the smaller the number of terms that are required in \eqref{eq:product_convolution} to achieve an accurate product-convolution approximation.

\subsection{Overview of the method}
\label{sec:overview_intro}

We compute one ``batch'' of $\varphi_i$'s by applying $\Aop$ to a sum of point sources (Dirac comb) associated with a collection of points $x_i$ scattered throughout $\Omega$. The batch of points $x_i$ are chosen so that the support of $\phi_i$ and the support of $\phi_j$ do not overlap (or do not overlap much) if $i \neq j$. Because these supports do not overlap, we can post-process the response of $\Aop$ to the Dirac comb to recover the functions $\phi_i$, and therefore the functions $\varphi_i$, associated with all points $x_i$ in the batch---with one application of $\Aop$, we recover many $\varphi_i$. This is illustrated in Figure FIG. The process is repeated to get more batches of $\varphi_i$'s, until a desired number of batches is reached.

%First, we compute
%\begin{equation}
%\label{eq:eta_intro}
%\eta := \left\langle \mathcal{A}, \xi \right\rangle^*
%\end{equation}
%where 
%\begin{equation*}
%	\xi := \sum_{i=1}^l \delta_{x_i}
%\end{equation*}
%is the Dirac comb associated with point sources centered at a collection of scattered points $x_i$. The batch of points $x_i$ are chosen so that the support of $\phi_i$ and the support of $\phi_j$ do not overlap (or do not overlap much) if $i \neq j$. By linearity, and the expression for $\phi_x$ in \eqref{eq:impulse_response_delta_action}, we have
%\begin{equation*}
%\eta = \sum_{i=1}^l \phi_i.
%\end{equation*}
%We may therefore post-process $\eta$ 
%
%to recover $\phi_i$ for all of the points $x_i$ in the batch. Since the functions $\varphi_i$ are translated versions of the functions $\phi_i$, we also recover $\varphi_i$ for all points $x_i$ in the batch. This is illustrated in Figure FIG. The process is repeated to get more batches of $\varphi_i$'s, until a desired number of batches is reached.

In order to choose the points $x_i$, we need to estimate the support of $\phi_i$ \emph{before} we compute it. The supports of the functions $\phi_x$ are estimated a-priori, for all $x \in \Omega$ simultaneously, via a procedure that involves applying $\Aop^T$ to a small number of specially chosen functions (three functions if $d=1$, six if $d=2$, ten if $d=3$), then post processing the results (Section \ref{eq:mean_and_covariance_estimation}). This procedure is the only part of the paper that requires non-negativity of the integral kernel.

We choose the weighting functions, $w_i$, to be the solutions to the following optimization problems:
\begin{equation}
\label{eq:wi_optimization_problem_intro}
\begin{aligned}
\min_{w_i \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,q.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$, $H^2_N(\Omega)$ is the space of all functions in the Sobolev space $H^2(\Omega)$ with Neumann zero boundary conditions, and the right hand side of the constraint, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$. Since the Laplacian measures local deviation from flatness \cite{JACOBSEN}, the function $w_i$ is the flattest function, in a least-squares sense, that takes the value one at the sample point $x_i$, and the value zero at the other sample points $x_j$, $i \neq j$. Computing the weighting functions $w_i$ requires solving two Poisson PDEs per weighting function, and these PDEs are solved cheaply using multigrid. We call this interpolation scheme \emph{Poisson interpolation}. The advantage of Poisson interpolation over other interpolation methods, most notably radial basis function interpolation \cite{ESCANDE}, is that Poisson interpolation incorporates information about the domain geometry when constructing the weighting functions. This is illustrated in Figure FIG.

Once the $\varphi_i$ and $w_i$ are computed, we convert $\AopPc$ to hierarchical matrix (H-matrix) format. Let $\Phi_\text{pc}$ denote the integral kernel associated with $\AopPc$. While $\Phi(y,x)$ is not easily computable, $\Phi_\text{pc}(y,x)$ is given by the formula
\begin{equation}
\label{eq:kernel_entries}
	\Phi_\text{pc}(y,x) = \sum_{i=1}^r \varphi_i(y-x) w_i(x).
\end{equation}
This follows from~\eqref{eq:product_convolution} and the fact that the convolution operator $u \mapsto \varphi_i \ast u$ has $(y,x)$ kernel entry given by $\varphi(y-x)$.
Formula~\eqref{eq:kernel_entries} allows us to construct a H-matrix representation of $\AopPc$ using the conventional adaptive cross H-matrix construction method, in which one forms low-rank approximations of blocks of the matrix by sampling rows and columns of those blocks \cite{HACA}. Once in H-matrix format, fast H-matrix arithmetic is used to invert $\AopPc$, or perform other useful matrix operations.

Often we are interested in approximating operators $\Aop$ that are symmetric positive semi-definite. Unfortunately, $\AopPc$ is, in general, non-symmetric and indefinite due to errors in the product convolution approximation. In this case, we modify the H-matrix representation of $\AopPc$ to make it positive definite, by symmetrizing it, then applying a specially chosen rational matrix function to the symmetrized H-matrix.

\subsection{Existing work}

In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\varphi_i$ and $w_i$ may be arbitrary functions. Product-convolution approximations have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}.
For background on product-convolution approximations, we recommend reading the following papers: \cite{PRODCONVGOOD}.

We build upon the class of product convolution approximations in which the functions, $\varphi_i$, are impulse responses of $\Aop$ to point sources at a collection of points $x_i$ \cite{PRODCONVLIST}. A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the functions $\varphi_i$ with piecewise linear interpolation \cite{NAGY}, or splines \cite{PRODCONVSPLINES}. With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation.
%This is computationally prohibitive in our applications, because computing the functions $\varphi_i$ requires one matrix vector product with $\Aop$ for each point $x_i$, and these matrix vector products are computationally costly. 
In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}. But even with adaptive refinement, many matrix vector products with $\Aop$ may be required. In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\varphi_i$ are computed with each matrix vector product.

%We do this by taking advantage of the fact that the functions $\varphi_i$ are locally supported; existing methods typically do not require this local support property.

Our method of estimating the support of the functions $\varphi_i$ was inspired by resolution analysis in seismic imaging \cite{RESOLUTION}. In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\varphi_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$. If the width of $\varphi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size. In \cite{RESOLUTION}, the width of $\varphi_p$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion. In this paper, rather than probing $\Aop^T$ with random noise functions, we use specific constant, linear, and quadratic functions. Our method estimates the support of $\varphi_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.

Matrix-vector products with product-convolution approximations can be performed using the fast Fourier transform if the problem is discretized on a regular grid \cite{PRODCONVFFT}. The product-convolution approximation can then replace the original operator when using Krylov methods to solve linear systems with the original operator as the coefficient operator \cite{PRODCONVKRYLOV}. However, for complex geometries, problems are rarely discretized using regular grids, so we cannot easily use the fast Fourier transform. Furthermore, in big-data inverse problems the required number of Krylov iterations is large \cite{MYDISSERTATION,OTHERS}. Thus, after constructing the product-convolution approximation, it is desirable to convert the product-convolution approximation to other matrix formats that are amenable to fast linear algebra operations. Wavelet compression methods have been used for this purpose \cite{PRODCONVWAVELETS}. Here we follow our previous work \cite{PRODCONVMYPAPER}, in which we convert the product-convolution approximation to H-matrix format.

Conventional H-matrix construction methods require access to matrix entries of the matrix being approximated, and therefore cannot be used to efficiently form H-matrix approximations of operators that are only available through matrix-vector products. There are matrix-free methods for H-matrix approximation \cite{LEXINGPEELINGPROCESS}, and these methods have been used to form H-matrix representations of Hessians in PDE constrained inverse problems \cite{ILONAHMATRIX,NOEMIHMATRIX}. While these matrix-free H-matrix construction methods are asymptotically scalable in theory, the required number of matrix vector products can be large in practice.

Product-convolution type approximations of Hessians have been used in a seismic inverse problem in \cite{GEORGSEISMICPRODCONV}, and in an advection-diffusion inverse problem in \cite{PRODCONVMYPAPER}. In \cite{DEMANETSEISMIC}, matrix probing \cite{MATRIXPROBING} is used to approximate the Hessian in a seismic inverse problem as the sum of simple pseudodifferential operators. Although it is not explicitly mentioned, the approximation in \cite{DEMANETSEISMIC} could be interpreted as a convolution-product interpolation, which is like a product-convolution approximation, except the order of pointwise multiplications and convolutions is reversed in \eqref{eq:product_convolution}.

\section{Constructing the product-convolution H-matrix approximation}
\label{sec:method}
In this section we describe how we choose the sample points, $x_i$ (Section \ref{sec:sample_point_selection}), how we form the convolution kernels, $\varphi_i$ (Section \ref{sec:get_impulse_response}), and how we form the weighting functions, $w_i$ (Section \ref{sec:weighting_functions}), for the product-convolution approximation, $\AopPc \approx \Aop$, given in \eqref{eq:product_convolution}. We also describe how we convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), and (optionally) how we modify the resulting H-matrix, $\AmatPc$, to force it to be symmetric positive semi-definite (Section \ref{sec:make_spd}). The complete algorithm for constructing $\AmatPc$ is shown in Algorithm \ref{alg:construct_Atilde}. 
%The method for evaluating kernel entries of $\Phi_\text{pc}$ is shown in Section \ref{sec:eval_matrix_entries}.

%\begin{itemize}
%\item We extract $\varphi_{x_i}$ for many points $x_i$ by applying $\mathcal{A}$ to Dirac combs associated with ``batches'' of points (Section \ref{sec:get_impulse_response}). To reduce memory usage, all kernels $\varphi_i$ associated with a given batch, $b$, are stored in a single function, $\eta^b$, rather than as distinct functions.
%\item We use a greedy algorithm to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other, and are not too close to $\partial \Omega$ (Section \ref{sec:greedy_point_selection}). 
%\item We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\psi_x$ (Section \ref{eq:mean_and_covariance_estimation}). 
%\item For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$; constructing the weighting functions requires solving two Poisson PDEs per point $x_i$ (Section \ref{sec:weighting_functions}).
%\end{itemize}
%The complete algorithm for constructing $\AopPc$ is shown in Algorithm \ref{alg:construct_Atilde}. The method for evaluating kernel entries of $\Phi_\text{pc}$ is shown in Section \ref{sec:eval_matrix_entries}. The method for converting $\AopPc$ to H-matrix format is shown in Section \ref{sec:H_matrix_conversion}. The method for modifying the resulting H-matrix to make it symmetric positive semi-definite is shown in Section \ref{sec:make_spd}.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	{	
		\Input{Linear operator $\Aop$, parameter $\text{max\_batches}$}
		\Output{H-matrix $\AmatPc$ (optionally, $\AmatPcSymPlus$)}

		Compute $\alpha$, $\mu$, and $\Sigma$ using Algorithm \ref{alg:varhpi_mean_cov}.

		\For{$b=1,2,\dots,\text{max\_batches}$}{
			Choose a batch of sample points, $S_b$, using Algorithm \ref{alg:point_choice}

			Compute $\combresponse_b$ by applying $\Aop$ to the dirac comb for $S_b$ (Section \ref{sec:get_impulse_response})

		}

		Form convolution kernels, $\varphi_{x_i}$, on regular grids (Section \ref{sec:convolution_kernel_regular_grid})

		Modify convolution kernels to address boundary artifacts (Section \ref{sec:boundary_extension})

		Compute all weighting functions $w_i$ using Algorithm \ref{alg:weighting_functions_incremental}
%		In lines 2 and 3 of Algorithm \ref{alg:weighting_functions_incremental}, only solve for $\psi_i$ and $\theta_i$ associated with new points $x_i$ in $S_b$. The functions $\psi_i$ and $\theta_i$ associated with points in previous batches have already been computed in previous iterations of this for loop.

		Form H-matrix $\AmatPc$ (Section \ref{sec:H_matrix_conversion})

		(optional) Modify $\AmatPc$ to make $\AmatPcSymPlus$ (Section \ref{sec:make_spd})

	}
	\caption{Construct product-convolution H-matrix}
	\label{alg:construct_Atilde}
\end{algorithm2e}

\subsection{Sample point selection}
\label{sec:sample_point_selection}

We choose sample points, $x_i$, in batches. A greedy algorithm is used to choose as many points as possible per batch, while ensuring that the points in each batch are not too close to each other. We ensure that the points $x_i$ are well separated from each other by forming a-priori estimates of the supports of the functions $\phi_x$ (Section \ref{eq:mean_and_covariance_estimation}). The support of $\phi_x$ is estimated to be contained within an ellipsoid, so the process of choosing batches of points is an ellipsoid packing problem (Section \ref{sec:greedy_point_selection}). 
%In the next section (Section \ref{sec:get_impulse_response}), we will extract $\varphi_{x_i}$ for all the points $x_i$ in a given batch by applying $\Aop$ to the Dirac comb associated with that batch.

\subsubsection{Estimating impulse response supports}
\label{eq:mean_and_covariance_estimation}

Because of the non-negative kernel property, for each $x \in \Omega$, the impulse response $\phi_x$ is a non-negative function $y \mapsto \phi_x(y)$. Hence, $\phi_x$ is a scaled probability distribution. Let $\alpha$ denote the scaling factor, let $\widehat{\phi}_x := \phi_x / \alpha(x)$ denote the normalized version of $\phi_x$, and let $\mu(x)$ and $\Sigma(x)$ denote the mean and covariance of $\widehat{\phi}_x$, respectively. We make the approximation that the support of $\phi_x$ is contained within the ellipsoid
\begin{equation}
\label{eq:support_ellipsoid}
E_x := \{x' \in \Omega: (x' - \mu(x))^T \Sigma(x)^{-1} (x' - \mu(x)) \le \tau^2\},
\end{equation}
where $\tau$ is a fixed constant. 
%The ellipsoid $E_{x}$ is the set of points within $\tau$ standard deviations of the mean of the Gaussian approximation of the normalized version of $\langle \mathcal{A}, \delta_{x}\rangle^{*}$ which has the same mean and covariance. 
The ellipsoid $E_x$ is the set of points within $\tau$ standard deviations from the mean of the Gaussian distribution with mean $\mu(x)$ and covariance $\Sigma(x)$, i.e., the Gaussian distribution which has the same mean and covariance as $\widehat{\phi}_x$ (see Figure FIG). The quantity $\tau$ is a parameter that must be chosen by the user of the algorithm. The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\phi_x$. However, the cost of our algorithm will depend on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$. We find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results. The fraction of the ``mass'' of $\phi_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region. 
% a good balance between these competing goals. 
%The bigger $\tau$ is, the smaller the overlap between impulse responses within a batch will be, while the smaller $\tau$ is, the more impulse responses we can fit within each batch.

The ellipsoid $E_x$ is not immediately available. To use $E_x$, we must first compute the scalar $\alpha(x)$, the length-$d$ vector $\mu(x)$, and the $d \times d$ matrix $\Sigma(x)$ (recall $d \in \{1,2,3\}$ is the spatial dimension of $\Omega$). The direct approach to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$ is to apply $\Aop$ to a point source centered at $x$ to get $\phi_x$, as per \eqref{eq:impulse_response_delta_action}. Then one can post-process $\phi_x$ to determine $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$. But this direct approach is computationally infeasible because and it requires one operator action of $\Aop$ per point $x$ that we wish to evaluate $E_x$ at, and we will need to evaluate $E_x$ at a large number of points. Fortunately, it is possible to compute $\alpha(x)$, $\mu(x)$, and $\Sigma(x)$, \emph{for all points $x$ simultaneously}, by applying $\Aop^T$ to one constant function, $d$ linear functions, and $d(d+1)/2$ quadratic functions. We present this efficient method of computing $\alpha$, $\mu$, and $\Sigma$ in Theorem \ref{thm:vol_mean_cov} and summarize the method in Algorithm \ref{alg:varhpi_mean_cov}.

%In Algorithm \ref{alg:varhpi_mean_cov} we show how to compute  Definitions required for Algorithm \ref{alg:varhpi_mean_cov} are given in Definition \ref{defn:alpha_rho_mu_sigma}, Definition \ref{defn:C_L_Q}, and Definition \ref{defn:pointwise}. The theoretical basis for Algorithm \ref{alg:varhpi_mean_cov} is shown in Theorem \ref{thm:vol_mean_cov}.

%We compute the scaling factor, denoted $\alpha(x)$, for all $x$ simultaneously by applying $\mathcal{A}^T$ to a constant function. We compute the spatially varying mean, $\mu(x)$, and covariance, $\Sigma(x)$ of the normalized version of $\phi_x$, for all $x$ simultaneously, by applying $\mathcal{A}^T$ to a small number of linear and quadratic functions. Definitions of $\alpha$, $\mu$, and $\Sigma$ are given in Definition \ref{defn:alpha_rho_mu_sigma}. The procedure for computing these quantities is shown in Algorithm \ref{alg:varhpi_mean_cov}, and the theoretical basis for this procedure is shown in Theorem \ref{thm:vol_mean_cov}.

\begin{theorem}[Impulse response moments]
	\label{thm:vol_mean_cov}
	Let $\alpha:\Omega \rightarrow \mathbb{R}$, $\mu:\Omega \rightarrow \mathbb{R}^d$, and $\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}$, be the following functions:
	\begin{equation*}
	\alpha(x) := \int_{\Omega} \phi_x(y) dy, \qquad
	\mu(x) := \mathbb{E}(\widehat{\phi}_x), \qquad
	\Sigma(x) := \Var(\widehat{\phi}_x),
	\end{equation*}
	where $\mathbb{E}(\widehat{\phi}_x)$ denotes the mean of $\widehat{\phi}_x$, and $\Var(\widehat{\phi}_x)$ denotes the variance of $\widehat{\phi}_x$.	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$. Let $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions:
	\begin{equation*}
	C(x) := 1, \qquad
	L^i(x) := x^i, \qquad
	Q^{ij}(x) = x^i x^j.
	\end{equation*}
	Let $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$. The following results hold:
	\begin{subequations}
		\label{eq:vol_mean_var}
	\begin{align}
	\alpha =& \left(\Aop^T C\right)^* \\
	\mu^i =& \left(\Aop^T L^i\right)^* / \alpha \\
	\Sigma^{ij} =& \left(\Aop^T Q^{ij}\right)^* / \alpha - \mu^i\cdot \mu^j
	\end{align}
	\end{subequations}
	 for $i=1,\dots, d$, $j=1,\dots,d$.
\end{theorem}

The proof of Theorem \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}. The proof is straightforward, and follows directly from the definitions.

%
%\begin{defn}[Constant, linear, and quadratic functions]	
%	\label{defn:C_L_Q}
%	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$. We define $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions, respectively:
%	\begin{equation*}
%	C(x) := 1, \qquad
%	L^i(x) := x^i, \qquad
%	Q^{ij}(x) = x^i x^j.
%	\end{equation*}
%\end{defn}
%
%\begin{defn}[Pointwise operations]
%	\label{defn:pointwise}
%	We write $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
%\end{defn}

%\begin{theorem}
%	\label{thm:vol_mean_cov}
%	Let  $x^i$ denote the $i^\text{th}$ component of $x$, i.e., $x = \left(x^1, x^2, \dots, x^d\right)$, and let $C$, $\{L^i\}_{i=1}^d$, and ${\{Q^{ij}\}_{i=1}^d}_{j=1}^d$ be the following constant, linear, and quadratic functions, respectively:
%	\begin{equation*}
%		C(x) := 1, \qquad
%		L^i(x) := x^i, \qquad
%		Q^{ij}(x) = x^i x^j.
%	\end{equation*}
%	We have
%	\begin{align}
%		\alpha =& \left(\mathcal{A}^TC\right)^*, \label{eq:vol_mean_var_thm1}\\
%		\mu^i =& \left(\mathcal{A}^T L^i\right)^* / \alpha, \label{eq:vol_mean_var_thm2}\\
%		\Sigma^{ij} =& \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot\mu^j, \label{eq:vol_mean_var_thm3}
%	\end{align}
%	We write $f/g$ to denote the pointwise division of functions, $\left(f/g\right)(x) = f(x)/g(x)$, and $f\cdot g$ to denote the pointwise multiplication of functions, $(f\cdot g)(x) = f(x)g(x)$.
%\end{theorem}
%
%The proof of Theorem \ref{thm:vol_mean_cov} is shown in Appendix \ref{app:proofs}.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	{	
		\Input{Operator $\mathcal{A}$}
		\Output{$\alpha$, $\mu$, and $\Sigma$}

		\tcp{Compute scaling factor $\alpha$}

		Form constant function $C(x)=1$

		Compute $\alpha = \left(\mathcal{A}^T C\right)^*$

		\tcp{Compute mean $\mu$}
		\For{$i=1,2,\dots,d$}{
			Form linear function $L^i(x) = x^i$

			Compute $\mu^i = \left(\mathcal{A}^T L^i\right)^* / \alpha$
		}
		\tcp{Compute covariance $\Sigma$}
		\For{$i=1,2,\dots,d$}{
			\For{$j=1,\dots,i$}{
				Form quadratic function $Q^{ij}(x) = x^i x^j$

				Compute $\Sigma^{ij} = \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot \mu^j$

				Set $\Sigma^{ji} = \Sigma^{ij}$

			}
		}

	}
	\caption{Compute scaling factor $\alpha$, mean $\mu$, and covariance $\Sigma$}
	\label{alg:varhpi_mean_cov}
\end{algorithm2e}

\subsubsection{Greedy ellipsoid packing}
\label{sec:greedy_point_selection}

We use a greedy ellipsoid packing algorithm to choose sets of sample points, $S_b$, such that there is no overlap between the ellipsoids, $E_{x_i}$, associated with the sample points, $x_i$, within a batch. These sample point batches will be used in Section \ref{sec:get_impulse_response} to compute many impulse responses of $\Aop$ using only a small number of operator actions of $\Aop$. The support of $\phi_x$ is (approximately) contained in the ellipsoid $E_x$, so by applying $\mathcal{A}$ to the Dirac comb associated with all sample points in a batch, we will recover the impulse responses for all sample points in that batch.

We start with a finite set of candidate points $P$. To build a batch of sample points, $S_b$, first we initialize $S_b$ as an empty set. Then we select the point $p \in P$ that is the farthest away from all points in previous sample point batches. That is, $p$ is a maximizer of the following optimization problem:
\begin{equation*}
\max_{p \in P} \min_{x \in S_1 \cup \dots \cup S_{b-1}} \|p - x\|.
\end{equation*}
Candidate points for the first batch are chosen arbitrarily from $P$.
Once $p$ is selected, we remove $p$ from $P$. If $p$ is sufficiently far from all of the previously chosen points in the current batch, in the sense that $E_p \cap E_q = \{\}$ for all $q \in S_b$,
%and if $p$ is sufficiently far from the boundary, in the sense that $E_p \cap \partial \Omega = \{\}$, 
then we add $p$ to $S_b$. Otherwise we discard $p$. 
%Recall that $E_p$ and $E_q$ are ellipsoids associated with points $p$ and $q$, respectively, as defined in \eqref{eq:support_ellipsoid}. 
This process repeats until there are no more points in $P$.  This is detailed in Algorithm \ref{alg:point_choice}. We determine whether $E_p \cap E_q = \{\}$ using the fast ellipsoid intersection test described in Appendix \ref{sec:fast_ellipsoid_intersection_test}.

We repeat the process to construct several batches of points $S_1, S_2, \dots$, until the number of batches exceeds a desired threshold. In our implementation, for each batch the set of candidate points $P$ is initialized as the set of all Lagrange nodes for the finite element basis functions used to discretize the problem, minus all points in previously chosen batches.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	\SetKwProg{Fn}{Function}{}{}

	\Input{Finite set of candidate points $P\subset \Omega$, \\mean $\mu(x)$ and covariance $\Sigma(x)$, \\domain boundary information $\partial \Omega$, \\previous sample point batches $S_1, \dots, S_{b-1}$}
	\Output{Batch of new sample points $S_b$}

	Initialize empty new batch of sample points, $S_b = \{\}$

	\While{$P$ is not empty}{
		Determine the point $p \in P$ that is farthest from all points in the previous sample point batches $S_1,\dots,S_{b-1}$

		Remove $p$ from $P$

		\If{$E_p \cap E_q \neq \{\}$ for all $q \in S_b$ and $E_p \cap \partial \Omega = \{\}$}{
			\tcp{$E_p$ and $E_q$ are the ellipsoids defined in \eqref{eq:support_ellipsoid}}

			Add $p$ to $S_b$

			Remove all points $p'$ satisfying $\mu(p') \in E_p$ from $P$
		}
	}

    \SetKwFunction{FMain}{point\_is\_acceptable}
	\caption{Choosing one batch of sample points, $S_b$}
	\label{alg:point_choice}
\end{algorithm2e}

\subsection{Convolution kernels}
\label{sec:get_impulse_response}

We compute convolution kernels $\varphi_{x_i}$ in batches by applying $\Aop$ to a small number Dirac combs. Each application of $\Aop$ to a Dirac comb yields one batch of impulse responses, $\phi_{x_i}$, which yields one batch of convolution kernels, $\varphi_{x_i}$.
%The batches of impulse responses, $\eta^b$, will ultimately be interpolated with weighting functions described in Section \ref{sec:weighting_functions} to form the desired approximation $\widetilde{\mathcal{A}}\approx \mathcal{A}$.  
The Dirac comb, $\dcomb_b$, associated with a batch of sample points, $S_b$, is the sum of Dirac distributions (point sources) centered at the points $x_i \in S_b$. That is,
\begin{equation*}
	\dcomb_b := \sum_{x_i \in S_b} \delta_{x_i}.
\end{equation*}
For each batch $S_b$, we compute the action of $\mathcal{A}$ on the associated Dirac comb:
\begin{equation}
\label{eq:dirac_comb_H_action}
	\combresponse_b := \left(\mathcal{A} \dcomb_b^*\right)^*.
\end{equation}
By linearity, the resulting function $\combresponse_b$ may be written as
\begin{equation}
\label{eq:phi_b}
	\combresponse_b = \left\langle\mathcal{A},\sum_{x_i \in S_b} \delta_{x_i}\right\rangle^* = \sum_{x_i \in S_b} \langle\mathcal{A}, \delta_{x_i}\rangle^* = \sum_{x_i \in S_b} \phi_{x_i}.
\end{equation}
Since the points $x_i$ are chosen so that the ellipsoid $E_{x_i}$ that (approximately) supports $\phi_i$, and the ellipsoid $E_{x_j}$ that (approximately) supports $\phi_j$ do not overlap when $i \neq j$, from \eqref{eq:convolution_kernel} and \eqref{eq:phi_b} we have (approximately)
\begin{equation}
\label{eq:varphi_eval}
	\varphi_{x_i}(z) = \phi_i(z+x_i) = \begin{cases}
		\combresponse_b(z+x_i), & z+x_i \in E_{x_i} \\
		0, & \text{otherwise}
		\end{cases}
\end{equation}
for all $x_i \in S_b$. By performing one matrix-vector product, \eqref{eq:dirac_comb_H_action}, we recover the convolution kernels $\varphi_{x_i}$ associated with every point $x_i \in S_b$.

\subsubsection{Interpolation onto local regular grids}
\label{sec:convolution_kernel_regular_grid}

To convert $\AopPc$ to H-matrix format (Section \ref{sec:H_matrix_conversion}), we will need to evaluate $\varphi_{x_i}(z)$ for a large number of points $z$. Per \eqref{eq:varphi_eval}, evaluating $\varphi_{x_i}(z)$ requires evaluating $\combresponse_b(z+x_i)$. In computations, $\combresponse_b$ will be a finite element function formed on an irregular mesh, and $z+x_i$ will typically not be a gridpoint of the mesh. Evaluating $\combresponse_b(z+x_i)$ therefore requires finding which mesh cell that $z+x_i$ is in, then evaluating several polynomial functions to evaluate the finite element function on that cell. Although this procedure is asymptotically scalable, it is slow in practice \cite{FEMEVALSLOW}.

To speed up evaluation of $\varphi_{x_i}$, we interpolate $\varphi_{x_i}$ onto a regular rectilinear grid on a coordinate axis aligned box that just barely contains the ellipsoid $E_{x_i}$. We transfer the function $\varphi_{x_i}$ from the irregular mesh to the regular grid by evaluating $\varphi_{x_i}$ at all gridpoints in the regular grid, using the slow process described in the previous paragraph. This is an upfront cost which is done ``offline.'' After $\varphi_{x_i}$ has been transferred to the regular grid, we use fast regular grid interpolation to perform all subsequent ``online'' evaluations of $\varphi_{x_i}$. This yields a substantial overall speedup because the number of finite element function evaluations required to transfer the functions $\varphi_{x_i}$ to the regular grid is $O(N)$, while the number of kernel evaluations required to convert $\AopPc$ to H-matrix format is $O(k N \log N)$. Here $N$ is the number of finite element degrees of freedom, and $k$ is the H-matrix rank. Furthermore, transferring $\varphi_{x_i}$ to a regular grid requires evaluating $\varphi_{x_i}$ at a predetermined and highly structured collection of points. This structure allows for faster determination of which mesh cells that points are in, and allows one to evaluate the finite element function at all regular gridpoints in parallel. In contrast, constructing the H-matrix requires evaluating $\varphi_{x_i}$ at an unpredictable sequence of scattered of points, which is an expensive task for finite element function evaluation, but a cheap task for regular grid interpolation. 
Having $\varphi_{x_i}$ stored on a regular grid also simplifies the implementation of the method that we use to deal with boundary artifacts, which will be described in Section \ref{sec:boundary_extension}.

We incur interpolation error when transferring $\varphi_{x_i}$ from the irregular mesh to the regular grid. To limit this interpolation error, the density of the regular grid should be as dense or denser than the local density of the finite element mesh. In our numerical results, the distance between neighboring points in the regular grid is chosen to be half the minimum distance between any pair of finite element Lagrange nodes that are contained in $E_{x_i}$. The grid transfer procedure is illustrated in Figure FIG.

When $x_i$ is near a boundary, it may occur that some gridpoints for the regular grid are in locations where $\varphi_{x_i}$ is undefined. In this case, we store a placeholder value for $\varphi_{x_i}$ at those gridpoints. We use the IEEE standard value of ``NaN'' (not a number) for this placeholder. In the next section (Section \ref{sec:boundary_extension}) we replace these NaN values with more appropriate values.

%To minimize memory usage, the functions $\varphi_i$ are not stored individually. Instead, we store the functions $\eta^b$, and use formula \eqref{eq:varphi_eval} to perform function evaluations $z \mapsto \varphi_i(z)$ as needed.

\subsubsection{Addressing boundary artifacts}
\label{sec:boundary_extension}

Evaluating the kernel for our product-convolution approximation, $\AkerPc(y,x)$, requires evaluating $\varphi_{x_i}(y-x)$. But $\varphi_{x_i}(y-x)$ is undefined  when $y - x + x_i$ is outside $\Omega$, which may occur even if $x \in \Omega$ and $y \in \Omega$. The default approach to deal with this problem is to extend $\varphi_{x_i}$ by zero wherever it is undefined. But this leads to substantial errors, known as boundary artifacts, which are especially significant when the sample point $x_i$ is near the boundary of the domain. In this case, the impulse response is artificially chopped off by the domain boundary. See Figure FIG for an illustration.

While $\varphi_{x_i}$ may be undefined at a point $z$, often there is a sample point $x_j$ that is near $x_i$, but further away from the boundary than $x_i$, such that $\varphi_{x_j}$ is well defined at $z$. To address boundary artifacts, we fill in missing/undefined portions of convolution kernels using information from other convolution kernels associated with the $k$ nearest sample points (we use $k=8$ in our numerical results). Kernels associated with nearer sample points are given precedence over kernels associated with farther away sample points. 
%After the convolution kernel has been extended using information from neighboring convolution kernels, it is then further extended by zero.

In detail, let $p_1, p_2, \dots, p_k$ be the $k$ nearest sample points to $x_i$, ordered by nearness, including $x_i$ itself. That is, $p_1=x_i$, $p_2$ is the nearest neighbor to $x_i$, $p_3$ is the second nearest neighbor to $x_i$, and so on. Also, recall from the previous section (Section \ref{sec:convolution_kernel_regular_grid}) that the functions $\varphi_{x_i}$ are stored on regular rectilinear grids. To extend $\varphi_{x_i}$, we form an expanded regular rectilinear grid that has the same grid spacing as the grid used for $\varphi_{x_i}$, and contains that grid as a subgrid, but is large enough to contain the boxes associated with all of the neighbor kernels, $\varphi_{p_1}, \varphi_{p_2}, \dots, \varphi_{p_k}$ (recall that the functions $\varphi_{x_i}$ are all centered at zero). We initialize an extended convolution kernel, $\varphi_{x_i}^E$, to zero on this expanded grid. Then we interpolate the functions $\varphi_{p_j}$ onto this expanded grid in reverse order. First we interpolate $\varphi_{p_k}$ onto the grid, then $\varphi_{p_{k-1}}$, and so on until the last step in which we interpolate the original kernel $\varphi_{p_1} = \varphi_{x_i}$. At each step, we overwrite existing values of $\varphi_{x_i}^E$ with new values from the kernel currently being interpolated, except when the new value of $\varphi_{p_j}$ is undefined (``NaN''). At gridpoints where the kernel being interpolated is undefined, we leave $\varphi_{x_i}^E$ unchanged. In this way, we fill in undefined values of $\varphi_{x_i}$ with values from successively further and further away kernels. This process is illustrated in Figure FIG. To avoid notational clutter, and since we will henceforth only work with the extended kernel, $\varphi_{x_i}^E$, in the remainder of this paper the symbol $\varphi_{x_i}$ refers to $\varphi_{x_i}^E$.

\subsection{Poisson weighting functions}
\label{sec:weighting_functions}

%For the weighting functions, we use the smoothest functions (in a least-squares sense) that interpolate the points $x_i$. Constructing the weighting functions requires solving two Poisson PDEs per sample point $x_i$.

We choose the $i^\text{th}$ weighting function, $w_i(x)$, to be the solution to the following optimization problem:
\begin{equation}
\label{eq:wi_optimization_problem}
\begin{aligned}
\min_{w_i \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta w_i}_{L^2(\Omega)}^2 \\
\text{such that} &\quad w_i(x_j) = \delta_{ij}, \quad j=1,\dots,r.
\end{aligned}
\end{equation}
Here $\Delta$ is the Laplacian on $\Omega$, $\delta_{ij}$, is the Kronecker delta, which takes value one if $i=j$ and zero if $i \neq j$, and $H^2_N(\Omega)$ is the space of all functions in $H^2(\Omega)$ with Neumann zero boundary conditions. Using these weighting functions yields the ``flattest possible'' interpolation of the convolution kernels $\varphi_i$, in a least-squares sense (Theorem \ref{thm:smoothest_interpolant}). Also, these weighting functions sum to one pointwise (Proposition \ref{thm:wi_sum_to_one}).

For each $i=1,\dots,r$, optimization problem \eqref{eq:wi_optimization_problem} is an example of the following generic omtimization problem:
\begin{equation}
\label{eq:wi_optimization_problem_generic}
\begin{aligned}
\min_{v \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta v}_{L^2(\Omega)}^2 \\
\text{such that} &\quad v(x_j) = \interpolatedvalues_j, \quad j=1,\dots,r,
\end{aligned}
\end{equation}
where $\interpolatedvalues_j \in \mathbb{R}$, $j=1,\dots,r$ are now arbitrary function values that $v$ must take at the points $x_j$. We may efficiently solve optimization problem \eqref{eq:wi_optimization_problem_generic} via the following procedure:
\begin{enumerate}
\item For each sample point $x_i$, we compute a function $\psi_i$ by solving the following Neumann Poisson problem:
\begin{equation}
\label{eq:psi_eq}
\begin{cases}
\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
\end{cases}
\end{equation}
with condition $\int_\Omega \psi_i(x) dx = 0$. Here $\delta_{x_i}$ is the point source centered at $x_i$, $|\Omega|$ is the measure of the domain $\Omega$, and $\nu$ is the normal vector to $\partial \Omega$.
\item For each function $\psi_i$, we compute a function $\secondgreens_i$ by solving another Neumann Poisson problem:
\begin{equation}
\label{eq:theta_eq}
\begin{cases}
\Delta \secondgreens_i = \psi_i, & \text{in }\Omega, \\
\nu \cdot \nabla \secondgreens_i = 0 & \text{on } \partial \Omega,
\end{cases}
\end{equation}
with condition $\int_\Omega \secondgreens_i(x) dx = 0$.
%\item For each sample point $x_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\delta_{x_i}$ (a point source centered at $x_i$) as the right hand side source. We call the solution to this first Poisson problem $\psi_i$.
%\item For each function $\psi_i$, we solve the Poisson equation with pure Neumann boundary conditions, using $\psi_i$ as the right hand side source. We call the solution to this second Poisson problem $\theta_i$.
\item The solution to optimization problem \eqref{eq:wi_optimization_problem_generic}, which we denote by $v_*$, is formed as follows:
\begin{equation}
\label{eq:optimal_v}
v_*(x) := \beta + \sum_{i=1}^r c_i \secondgreens_i(x),
\end{equation}
where the constants $c_i \in \mathbb{R}$, $i=1,\dots,r$, and $\beta \in \mathbb{R}$ are the solution to the following linear system:
\begin{equation}
	\label{eq:interpolation_linear_system}
	\begin{bmatrix}
	S & \mathbf{1} \\ \mathbf{1}^T & 0
	\end{bmatrix}
	\begin{bmatrix}
	\mathbf{c} \\ \beta
	\end{bmatrix}
	=
	\begin{bmatrix}
	\mathbf{\interpolatedvalues} \\ 0
	\end{bmatrix}.
\end{equation}
Here $\mathbf{c} \in \mathbb{R}^r$ is the vector with $i$th component $c_i$, $\mathbf{1} \in \mathbf{R}^r$ is the column vector of all ones, $\mathbf{\interpolatedvalues} \in \mathbb{R}^r$ is the vector with $i$th component $\interpolatedvalues_i$, and $S \in \mathbb{R}^{r \times r}$ is the matrix with entries given by
\begin{equation}
\label{eq:S_matrix}
S_{ij} := \int_\Omega \firstgreens_i(x) \firstgreens_j(x) dx.
\end{equation}
\end{enumerate}
In Theorem \ref{thm:wi_no_breakdown}, we prove that this procedure does not break down, so long as there is no repetition in the points $x_i$. Linear system \eqref{eq:interpolation_linear_system} is the KKT system that arises if one solves optimization problem \eqref{eq:wi_optimization_problem_generic} on the $r+1$-dimensional subspace spanned by the functions $\firstgreens_i$, $i=1,\dots,r$, and the constant function $C(x)=1$, rather than solving the optimization problem on the infinite dimensional space $H^2_N(\Omega)$. In Theorem \ref{thm:wi_alg_is_correct}, we prove that the function $v_*$ produced by this process solves optimization problem \eqref{eq:wi_optimization_problem_generic} on $H^2_N(\Omega)$.

Constructing all $r$ of the functions $w_i$ thus requires solving the Poisson equation $2r$ times with different right hand side sources to form the functions $\firstgreens_i$ and $\secondgreens_i$, then post-processing these functions. This is shown in Algorithm \ref{alg:weighting_functions_incremental}. The Poisson PDE solves, which are the most expensive part of this procedure, can be performed cheaply with multigrid.

%All of the weighting functions, $\{w_i\}_{i=1}^r$, will change if a new point, $x_{r+1}$, is added. However, the dominant computational cost in constructing the weighting functions is the solution of the Poisson PDEs, and these Poisson PDE solves can be performed incrementally---we only need to compute functions $\psi_i$ and $\theta_i$ associated with the new points. 
%, rather than $2(r+1)$ new PDE solves.

It is important to note that functions in $H^2_N(\Omega)$ only have well-defined pointwise values if the spatial dimension of the domain is $d=1$, $2$, or $3$. Hence optimization problem \eqref{eq:wi_optimization_problem} is only well-posed in three spatial dimensions or fewer. In four or higher spatial dimensions, one could construct appropriate weighting functions using radial basis functions, though we do not discuss this here.

\begin{thm}[Procedure for computing $v_*$ does not break down]
	\label{thm:wi_no_breakdown}
	If all the sample points $x_i$, $i=1,\dots,r$, are distinct, then PDE \eqref{eq:psi_eq} is uniquely solvable for $\firstgreens_i$, PDE \eqref{eq:theta_eq} is uniquely solvable for $\secondgreens_i$, and linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\mathbf{c}$ and $\beta$. That is, the procedure for computing $v_*$ does not break down. Furthermore, $v_* \in H^2_N(\Omega)$.
\end{thm}

\begin{proof}
	Existence and uniqueness for the solutions to PDEs \eqref{eq:psi_eq} and \eqref{eq:theta_eq} follow from the standard elliptic PDE theory. By standard elliptic regularity theory, $\firstgreens_i \in H^1(\Omega)$, so $\firstgreens_i \in L^2(\Omega)$, and so all matrix entries $S_{ij}$ are finite.

	Using the definition of $S$ and basic linear algebra, it is straightforward to show that if $S$ were singular, there would exist constants $a_1, a_2, \dots a_r$, which are not all zero, such that
	\begin{equation*}
		0 = a_1 \firstgreens_1 + a_2 \firstgreens_2 + \dots a_r \firstgreens_r.
	\end{equation*}
	Applying the Laplacian to this equation yields
	\begin{equation*}
		0 = a_1 \delta_{x_1} + a_2 \delta_{x_2} + \dots a_r \delta_{x_r} - \left(\sum_{i=1}^r a_i\right)\bigg/|\Omega|,
	\end{equation*}
	which is impossible because a nonzero linear combination of delta distributions at different points cannot equal a constant function. Thus we must conclude that $S$ is non-singular. By block matrix factorization, non-singularity of $S$ implies that linear system \eqref{eq:interpolation_linear_system} is uniquely solvable for $\mathbf{c}$ and $\beta$.

	By standard elliptic regularity theory, the functions $\secondgreens_i$ are in $H^3(\Omega)$, and therefore they are in $H^2(\Omega)$. By construction the functions $\secondgreens_i$ have Neumann zero boundary conditions, so $\secondgreens_i \in H^2_N(\Omega)$. Since $v_*$ is a linear combination of the functions $\secondgreens_i$ and the constant function $C(x)=1$, $v_* \in H^2_N(\Omega)$.
\end{proof}

\begin{lem}[Preparation for Theorem \ref{thm:wi_alg_is_correct}]
	\label{lem:psi_delta_eval}
	We have
	\begin{equation}
	\label{eq:inline_wi_lemma}
	h(x_j) = \left(\firstgreens_j, \Delta h\right)_{L^2(\Omega)} + \avg(h)
	\end{equation}
	for all $h\in H^2_N(\Omega)$, where
	\begin{equation*}
	\avg(h) :=	 \frac{1}{|\Omega|} \int_\Omega h ~dx,
	\end{equation*}
	and $|\Omega| = \int_\Omega 1 ~dx$ is the Lebesgue measure of $\Omega$.
\end{lem}

\begin{proof}
	The result follows from multiplying \eqref{eq:psi_eq} by $h$, integrating both sides of the equation over $\Omega$, performing integration by parts to move derivatives from $\firstgreens_j$ onto $h$, and using the Neumann zero boundary conditions for $\firstgreens_j$ and $h$ to eliminate boundary integral terms.
\end{proof}

\begin{thm}[Solution of optimization problem \ref{eq:wi_optimization_problem_generic}]
	\label{thm:wi_alg_is_correct}
	The function $v_*$ given in \eqref{eq:optimal_v} is the unique solution to optimization problem \ref{eq:wi_optimization_problem_generic}.
\end{thm}

\begin{proof}
	Let
	\begin{equation*}
		X := \{v \in H^2_N(\Omega) : v(x_i) = \interpolatedvalues_i \text{ for }i=1,\dots,r\}
	\end{equation*}
	denote the feasible set for optimization problem \ref{eq:wi_optimization_problem_generic}, and let 
	\begin{equation*}
	X_0 := \{v \in H^2_N(\Omega) : v(x_i) = 0 \text{ for }i=1,\dots,r\}
	\end{equation*}	
	denote the tangent space to $X$. Note that pointwise values of functions in $H_N^2(\Omega)$ are well-defined since the Sobolev imbedding theorem (see, e.g., Theorem 7.19 in \cite{Arbogast}) implies that $H^2(\Omega)$ is continuously imbedded in the space of continuous functions $C^0(\Omega)$ when $d < 4$, and we are considering $d \in \{1,2,3\}$. The feasible set $X$ is therefore a closed affine subspace of $H^2_N(\Omega)$, the tangent space $X_0$ is a closed subspace of $H^2_N(\Omega)$, and optimization problem \eqref{eq:wi_optimization_problem_generic} is well-defined.

	The proof now proceeds in three steps. First, we show that $v_*$ is feasible ($v_* \in X$). Second, we show that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} if it satisfies a certain variational equation. Third, we verify that $v_*$ satisfies this variational equation.

	\paragraph{Feasibility} To show that $v_*$ is feasible (i.e., $v_* \in X$), we must show that $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,r$. We have
	\begin{align*}
	v_*(x_i) &= \beta + \sum_{j=1}^r c_j \secondgreens_j(x_i) \\
	&= \beta + \sum_{j=1}^r c_j \left(\firstgreens_i, \Delta \secondgreens_j\right)_{L^2(\Omega)} + \avg\left(\secondgreens_j\right) \\
	&= \beta + \sum_{j=1}^r c_j \left(\firstgreens_i, \firstgreens_j\right)_{L^2(\Omega)} \\
	&= \beta + \sum_{j=1}^r c_j S_{ij}
	\end{align*}
	In the first line we used the definition of $v_*$ in \eqref{eq:optimal_v}. In the second line we used Lemma \ref{lem:psi_delta_eval}. In the third line we used the facts that, by construction, $\Delta \secondgreens_j = \firstgreens_j$ and $\int_\Omega \secondgreens_j(x) dx = 0$. In the fourth line we used the definition of $S$ in \eqref{eq:S_matrix}. Hence the condition $v_*(x_i) = \interpolatedvalues_i$ for $i=1,\dots,r$ may be written in matrix form as
	\begin{equation*}
	S \mathbf{c} + \beta \mathbf{1} = \mathbf{\interpolatedvalues},
	\end{equation*}
	and this must hold because it is the first block row of system \eqref{eq:interpolation_linear_system} which is solved to compute $\mathbf{c}$ and $\beta$. So $v_* \in X$.

	\paragraph{Optimality} Since $X$ is an affine subspace, and $X_0$ is the tangent space to $X$, any $v \in X$ may be written in the form $v = v_* + p$ for some $p \in X_0$. We may therefore reframe optimization problem \ref{eq:wi_optimization_problem_generic} as 
	\begin{equation}
	\label{eq:p_optimization}
		\min_{p \in X_0} J(p),
	\end{equation}
	where
	\begin{equation*}
		J(p) := \frac{1}{2}||\Delta (v_* + p)||^2.
	\end{equation*}
	Showing that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} is equivalent to showing that $0$ is the unique minimizer of \eqref{eq:p_optimization}.

	By direct expansion, we have
	\begin{equation}
		\label{eq:Ju_expanded}
		J(p) = J(0) + \left(\Delta u, \Delta p\right)_{L^2(\Omega)} + \frac{1}{2}||\Delta p||^2.
	\end{equation}
	Since $p$ has Neumann zero boundary conditions, if $\Delta p = 0$, then $p$ is constant on each connected component of $\Omega$. Furthermore, since $p(x_i)$ is zero for at least one point $x_i$ per connected component of $\Omega$, if $\Delta p = 0$ then $p=0$. Hence, by contrapositive we have
	\begin{align*}
	\label{eq:contrapositive_laplacian_zero}
		p \neq 0 \quad \implies \quad& 0 < ||\Delta p||_{L^2(\Omega)}\\
		\implies \quad& J(0) < J(p) + \left(\Delta v_*, \Delta p\right)_{L^2(\Omega)}.
	\end{align*}
	If $v_*$ satisfies the following variational equation,
	\begin{equation}
	\label{eq:variational_up}
		(\Delta v_*, \Delta p)_{L^2(\Omega)} = 0 \quad \text{for all~} p \in X_0,
	\end{equation} 
	then
	\begin{equation*}
		p \neq 0 \quad \implies \quad J(0) < J(p).
	\end{equation*}
	Showing that $v_*$ is the unique minimizer of \eqref{eq:wi_optimization_problem_generic} therefore reduces to showing that $v_*$ satisfies variational equation \eqref{eq:variational_up}.

	\paragraph{Variational equation} To show that $v_*$ satiafies variational equation \eqref{eq:variational_up}, notice that
	\begin{align*}
	(\Delta v_*, \Delta p)_{L^2(\Omega)} =& \sum_{i=1}^r c_i(\Delta \secondgreens_i, \Delta p)_{L^2(\Omega)} \\
	=& \sum_{i=1}^r c_i \left(p(x_i) - \avg\left(p\right)\right) \\
	=& -\avg\left(p\right) \sum_{i=1}^r c_i. 
	\end{align*}
	In the first line we used the definition of $\secondgreens_i$ and the fact that the Laplacian of a constant is zero. In the second line we used Lemma \ref{lem:psi_delta_eval}. In the third line we used the fact that $p(x_i)=0$ because $p \in X_0$. Hence variational equation \eqref{eq:variational_up} holds if $\sum_{i=1}^r c_i = 0$, which may be written in matrix form as $\mathbf{1}^T \mathbf{c} = 0$. This must hold because it is the second block row in system \eqref{eq:interpolation_linear_system}.
\end{proof}

\begin{thm}[Poisson weighting functions are optimal]
	\label{thm:smoothest_interpolant}
Let $y \in \mathbb{R}^d$, $d \in \{1,2,3\}$, and let $\horizinterpolant_y$ be the function
\begin{equation}
\label{eq:u_defn_in_thm}
\horizinterpolant_y(x) := \sum_{i=1}^r \varphi_i(y)w_i(x) = \AkerPc\left(y+x,x\right),
\end{equation}
where the weighting functions $w_i$ are the solutions to optimization problem \eqref{eq:wi_optimization_problem}. Then for each $y \in \Omega$, the function $\horizinterpolant_y$ solves the optimization problem
\begin{equation}
\label{eq:uy_optimization_problem}
\begin{aligned}
\min_{\horizinterpolant_y \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \horizinterpolant_y}_{L^2(\Omega)}^2 \\
\text{such that} &\quad \horizinterpolant_y(x_i) = \varphi_i(y), \quad i=1,\dots,q.
\end{aligned}
\end{equation}
\end{thm}

\begin{prop}[Poisson weighting functions sum to one]
	\label{thm:wi_sum_to_one}
	We have
	\begin{equation*}
		\sum_{i=1}^r w_i(x) = 1
	\end{equation*}
	for all $x \in \Omega$.
\end{prop}

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	{	
		\Input{Sample points $\{x_i\}_{i=1}^r$, domain geometry $\Omega$}
		\Output{Weighting functions $\{w_i\}_{i=1}^r$}

%		Initialize constant function $C(x) = 1$

%		Compute Lebesgue measure of the domain, $|\Omega|$

%		\tcp{Compute Poisson impulse responses $\{\psi_i\}_{i=1}^r$ and Poisson squared impulse responses $\{\theta_i\}_{i=1}^r$}

		\For{$k=1,2, \dots, r$}{
%			Initialize $\delta_{x_i}$ as the point source (delta distribution) centered at $x_i$.

			Solve the Neumann Poisson problem \eqref{eq:psi_eq} for $\firstgreens_k$.

%			Compute $\psi_i$ by solving the Neumann Poisson problem
%			\begin{equation}
%			\label{eq:psi_eq}
%			\begin{cases}
%			\Delta \psi_i = \delta_{x_i} - 1/|\Omega|, & \text{in }\Omega, \\
%			\nu \cdot \nabla \psi_i = 0 & \text{on } \partial \Omega,
%			\end{cases}
%			\end{equation}
%			with condition $\int_\Omega \psi_i(x) dx = 0$, where $\nu$ is the normal vector to $\partial \Omega$.

			Solve the Neumann Poisson problem \eqref{eq:theta_eq} for $\secondgreens_k$.

%			Compute $\theta_i$ by solving the Neumann Poisson problem
%			\begin{equation}
%			\label{eq:theta_eq}
%						\begin{cases}
%			\Delta \theta_i = \psi_i, & \text{in }\Omega, \\
%			\nu \cdot \nabla \theta_i = 0 & \text{on } \partial \Omega,
%			\end{cases}
%			\end{equation}
%			with condition $\int_\Omega \theta_i(x) dx = 0$, where $\nu$ is the normal vector to $\partial \Omega$.

		}

		Form the $r \times r$ matrix $S$ given in \eqref{eq:S_matrix}.

%		\tcp{Compute weighting functions $\{w_i\}_{i=1}^r$}

		\For{$i=1,2,\dots,r$}{
			Solve the linear system
			\begin{equation*}
			\begin{bmatrix}
			S & \mathbf{1} \\ \mathbf{1}^T & 0
			\end{bmatrix}
			\begin{bmatrix}
			\mathbf{c}^{(i)} \\ \beta^{(i)}
			\end{bmatrix}
			=
			\begin{bmatrix}
			\mathbf{e}^{(i)} \\ 0
			\end{bmatrix}.
			\end{equation*}
			for $\mathbf{c}^{(i)}=(c_1^{(i)}, c_2^{(i)}, \dots, c_r^{(i)})$ and $\beta^{(i)}$. Here $\mathbf{1} = (1,1,\dots,1) \in \mathbb{R}^r$, and $\mathbf{e}^{(i)} = (0,\dots,0,1,0,\dots,0)\in \mathbb{R}^r$ is the unit vector with $i^\text{th}$ component equal to one and all other components equal to zero.

			Form the weighting function $w_i$ given by
			\begin{equation*}
			w_i(x) = \beta^{(i)} + \sum_{k=1}^r c_k^{(i)} \secondgreens_k(x).
			\end{equation*}

		}
	}
	\caption{Compute weighting functions $w_i$, $i=1,\dots,r$.}
	\label{alg:weighting_functions_incremental}
\end{algorithm2e}

%\subsection{Evaluating kernel entries of the approximation}
%\label{sec:eval_matrix_entries}
%
%We now describe how to efficiently evaluate $\widetilde{\Phi}(y,x)$ at an arbitrary point $(x,y) \in \Omega \times \Omega$. From \eqref{eq:kernel_entries}, we have
%\begin{equation}
%\label{eq:eval_y_x_000}
%\AkerPc(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x),
%\end{equation}
%and from \eqref{eq:varphi_eval}, we have
%\begin{equation}
%\label{eq:kernel_entry_formua}
%\varphi_{x_i}(y-x) = \begin{cases}
%\eta^{b(i)}(y - x + x_i), & y - x + x_i \in E_{x_i} \\
%0, & \text{otherwise},
%\end{cases}
%\end{equation}
%where $b(i)$ denotes the batch index such that $x_i \in S^{b(i)}$. To compute $z=\AkerPc(y,x)$, we start with $z=0$, then loop through all batches $b$, and all sample points in each batch. For each sample point, we add one term in the sum in \eqref{eq:eval_y_x} to $z$, using \eqref{eq:kernel_entry_formua} as needed to evaluate $\varphi_{x_i}$. This is shown in Algorithm \ref{alg:eval_y_x}.
%
%
%\begin{algorithm2e}
%	\SetAlgoNoLine
%	\SetKwInOut{Input}{Input}
%	\SetKwInOut{Output}{Output}
%	
%	\Input{points $x \in \Omega$, $y \in \Omega$, \\sample point batches $S_1, \dots, S_\text{max\_batches}$, \\Dirac comb responses $\{\eta^b\}_{b=1}^{\text{max\_batches}}$, \\weighting functions $\{w_i\}_{i=1}^r$}
%	\Output{Kernel entry $z = \widetilde{A}(y,x)$}
%
%	$z \gets 0$
%	
%	\For{$b=1,2,\dots,\text{}max\_batches$}{
%		\For{$x_i \in S_b$}{
%			\If{$y - x + x_i \in E_{x_i}$}{
%				$z \gets z + w_i(x)\eta^b(y-x+x_i)$
%			}
%		}
%	}	
%	\caption{Compute approximate kernel entry $z = \widetilde{A}(y,x)$}
%	\label{alg:eval_y_x}
%\end{algorithm2e}

\subsection{Discretization and hierarchical matrix construction}
\label{sec:H_matrix_conversion}

Here we discuss how we construct an H-matrix representation of our product-convolution approximation, after discretizing the problem with finite elements.

\subsubsection{Finite element discretization of the integral kernel}

Let $f_1, f_2, \dots, f_N$ be a set of finite element basis functions used to discretize the problem on a mesh with mesh size parameter $h$, and let $V_h := \Span\left(f_1, f_2, \dots, f_N\right)$ be the corresponding finite element space. Further, let $p_i \in \mathbb{R}^d$, $i=1,\dots, N$ be the Lagrange nodes associated with the functions $f_i$. That is, the points such that $f_i(p_j)$ equals one if $i=j$, and zero otherwise.

In Section \ref{sec:intro} we noted that the original integral kernel, $\Aker$, is inaccessible, while in Equation \eqref{eq:kernel_entries}  we saw that the integral kernel for our product-convolution approximation, $\AkerPc \approx \Aker$, is accessible, and may be cheaply evaluated at arbitrary pairs of points $(y,x) \in \Omega \times \Omega$ via the following formula:
\begin{equation}
\label{eq:eval_y_x}
\AkerPc(y,x) = \sum_{i=1}^r w_i(x) \varphi_{x_i}(y-x).
\end{equation}
We now define a further approximation, $\AkerPcMesh \approx \AkerPc$, as follows:
\begin{equation}
\label{eq:defn_of_Akerpcmesh}
\AkerPcMesh(y,x) := \sum_{i=1}^N \sum_{j=1}^N \AkerPc(p_i,p_j) f_i(y) f_j(x).
\end{equation}
The kernel $\AkerPcMesh$ is the interpolation of $\AkerPc$ onto $V_h \otimes V_h$. Error in the overall kernel approximation, $\AkerPcMesh \approx \Aker$, arises both because of error in the product-convolution approximation, and because of finite element discretization error incurred from interpolation onto $V_h \otimes V_h$.

Replacing $\Aker$ with $\AkerPcMesh$ within the definition of $\Aop$ in \eqref{eq:kernel_representation} yields the following operator approximation, $\AopPcMesh$:
\begin{equation}
\label{eq:kernel_representation_pcmesh}
(\AopPcMesh u)(v) := \int_\Omega \int_\Omega v(y) \AkerPcMesh(y,x) u(x) dx dy.
\end{equation}
While $\AopPcMesh$ is well-defined as a mapping $L^2(\Omega) \rightarrow L^2(\Omega)'$, we are interested in the restriction of $\AopPcMesh$ to functions in the finite element space. That is, $\AopPcMesh : V_h \rightarrow V_h'$.
%Error in the approximation $\Aop \approx \AopPcMesh$ is a combination of error in the product-convolution approximation, and error due to finite element discretization.
Let $u_h \in V_h$ and $v_h \in V_h$, and let $\mathbf{u}$ and $\mathbf{v}$ be the coefficient vectors for $u_h$ and $v_h$, respectively, with respect to the finite element basis $f_1, \dots, f_N$. I.e., 
\begin{equation}
\label{eq:fem_coeff_basis}
u_h(x) = \sum_{i=1}^N \mathbf{u}_i f_i(x),
\end{equation}
and similar for $v_h$. Further, let $\mathbf{M} \in \mathbb{R}^{N \times N}$ be the finite element mass matrix,
% which has entries
%\begin{equation}
%\label{eq:mass_matrix_defn}
%\mathbf{M}_{ij} := \int_\Omega f_i(x) f_j(x) dx,
%\end{equation}
and let $\AkerPcMat \in \mathbb{R}^{N \times N}$ be following matrix of approximate kernel entries:
\begin{equation}
\label{eq:Akerpcmat_entries}
\left(\AkerPcMat\right)_{ij} := \AkerPcMesh(p_i, p_j).
\end{equation}
Using \eqref{eq:defn_of_Akerpcmesh}, \eqref{eq:kernel_representation_pcmesh}, \eqref{eq:fem_coeff_basis}, and \eqref{eq:Akerpcmat_entries}, the definition of the mass matrix, and performing algebraic manipulations, it is straightforward but tedious to show that
\begin{equation*}
\left(\AopPcMesh u_h \right)(v_h) = \mathbf{v}^T \mathbf{M} \AkerPcMat \mathbf{M} \mathbf{u}.
\end{equation*}
% PUT INTO APPENDIX!!!!
%\begin{align*}
%\left(\AopPcMesh u_h \right)(v_h) =& \int_\Omega \int_\Omega v_h(y) \AkerPcMesh(y,x) u_h(x) dx dy \\
%=& \int_\Omega \int_\Omega \left(\sum_{i=1}^N \mathbf{v}_i f_i(y)\right) \left(\sum_{j=1}^N \sum_{k=1}^N \AkerPc(p_j,p_k) f_j(y) f_k(x)\right) \left(\sum_{l=1}^N \mathbf{u}_l f_l(x)\right) dx dy \\
%=& \sum_{i,j,k,l=1}^N \mathbf{v}_i \left(\int_\Omega f_i(y) f_j(y) dy\right) \AkerPcMesh(p_j, p_k) \left(\int_\Omega f_k(x) f_l(x) dx\right) \mathbf{u}_l \\
%=& \sum_{i,j,k,l=1}^N \mathbf{v}_i \mathbf{M}_{ij} \left(\AkerPcMat\right)_{jk} \mathbf{M}_{kl} \mathbf{u}_l \\
%=& \mathbf{v}^T \mathbf{M} \AkerPcMat \mathbf{M} \mathbf{u},
%\end{align*}
The finite element matrix representation of $\AopPcMesh$, which we denote by $\AmatPc$, is therefore given by
\begin{equation}
\label{eq:Amatpc_defn}
\AmatPc := \mathbf{M} \AkerPcMat \mathbf{M}.
\end{equation}

\subsubsection{Hierarchical matrix construction}

We form an H-matrix representation of $\AmatPc$ by converting $\AkerPcMat$ and $\mathbf{M}$ to H-matrix format, then using fast H-matrix methods to multiply these matrices according to \eqref{eq:Amatpc_defn}. Once $\AmatPc$ is in H-matrix format, useful matrix operations such as matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, and matrix inversion may be performed using fast scalable H-matrix methods \cite{BORMHACKBUSCHBOOK}. We use H1 matrices in our numerical results, but any of the other H-matrix formats (such as H2, HODLR, HSS, HBS, and others \cite{HMATRIX}) could be used instead. For more details on H-matrices, we recommend \cite{HMATRIXGOOD}. 
%We discuss H-matrices in greater detail in Appendix \ref{app:h_matrix}.

We convert $\AkerPcMat$ into H1 matrix format using the standard geometrical clustering/adaptive cross method implemented within the HLIBPro software package \cite{HLIBPRO}. Although $\AkerPcMat$ is a dense $N \times N$ matrix, constructing the H-matrix approximation of $\AkerPcMat$ only requires evaluation of $O(k^2 N \log N)$ entries of $\AkerPc$, and these entries are computed via the formula in Equation \eqref{eq:eval_y_x}. Here $k$ is the rank of the highest-rank block in the H-matrix. A dense representation of $\AkerPcMat$ is never formed. We describe this H-matrix construction process in detail in Appendix \ref{app:h_matrix}. We convert $\mathbf{M}$ to H-matrix format using standard H-matrix methods for sparse matrices implemented within HLIBPro, using the same recursive block partitioning structure as was used for $\AkerPcMat$.

\subsection{Rational positive semi-definite modification}
\label{sec:make_spd}

In many problems of practical interest (e.g., Hessian approximation) $\Amat$ is symmetric positive semi-definite. However, $\AmatPc$ is generally non-symmetric and indefinite because of errors in the product-convolution approximation. 
This is undesirable. Symmetry and positive semi-definiteness are important properties which should be preserved if possible. Also, lacking these properties may prevent one from using highly effective algorithms, such as the conjugate gradient method, to perform further useful operations involving $\AmatPc$.
In this section we describe a method for modifying $\AmatPc$ to make it symmetric positive semi-definite.

First, we use fast H-matrix addition to symmetrize $\AmatPc$:
\begin{equation*}
\AmatPcSym := \frac{1}{2}\left(\AmatPc + \AmatPc^T\right).
\end{equation*}
Then we modify $\AmatPcSym$ to make it positive semi-definite by forming a rational matrix function of the following form: 
\begin{equation}
\label{eq:rational_matrix_function}
\AmatPcSymPlus := c_0 \mathbf{I} + c_1 \AmatPcSym + c_2 \left(\AmatPcSym + \mu \mathbf{I}\right)^{-1},
\end{equation}
where $\mathbf{I}$ is the identity matrix which has the same shape as $\AmatPcSym$. Approximation \eqref{eq:rational_matrix_function} may be written as $\AmatPcSymPlus = \ratfct(\AmatPcSym)$, where $\ratfct$ is the following rational function:
\begin{equation}
\label{eq:rational_function_scalar}
\ratfct(\lambda) := c_0 + c_1 \lambda + \frac{c_2}{\lambda+\mu}.
\end{equation}
The scalars $c_0$, $c_1$, $c_2$, and $\mu$ are chosen so that the moderate and large positive eigenvalues of $\AmatPcSymPlus$ approximately equal the corresponding moderate and large positive eigenvalues of $\AmatPcSym$, while the negative and small positive eigenvalues of $\AmatPcSym$ are modified so that $\AmatPcSymPlus$ is positive semi-definite. Later in this section we will explain how $c_0$, $c_1$, $c_2$, and $\mu$ are chosen and how our choice of these scalars affects the spectral properties of $\AmatPcSymPlus$. In Figure FIG we illustrate $\ratfct(\lambda)$. Forming $\AmatPcSymPlus$ requires computing an H-matrix inverse, $\left(\AmatPcSym + \mu \mathbf{I}\right)^{-1}$. While explicit computation of matrix inverses should usually be avoided, here it is computationally acceptable because $\AmatPcSym + \mu \mathbf{I}$ is an H-matrix. Inversion of H-matrices is a routine procedure which only requires $O(N \log(N)^2)$ work \cite{Hmatrixinverse}.

We choose $c_0, c_1, c_2$, and $\mu$, to be the solution to the following system of equations:
\begin{subequations}
	\label{eq:four_eq_system}
\begin{align}
\ratfct(0) &= 0 \label{eq:desired_property_1} \\
\ratfct'(0) &= 0 \label{eq:desired_property_2} \\
\ratfct(\lambda_\text{min}) &= |\lambda_\text{min}| \label{eq:desired_property_3} \\
\ratfct(\lambda_\text{max}) &= \lambda_\text{max}, \label{eq:desired_property_4}
\end{align}
\end{subequations}
where $\lambda_\text{min}$ and $\lambda_\text{max}$ are the smallest and largest eigenvalues of $\AmatPcSym$, respectively. We also require $\mu > |\lambda_\text{min}|$, so that the pole of $\ratfct$ is not in $[\lambda_\text{min},\lambda_\text{max}]$. Note that $\lambda_\text{min} < 0$ since $\AmatPcSym$ is indefinite. 
We estimate $\lambda_\text{min}$ and $\lambda_\text{max}$ using the implicitly restarted Lanczos method \cite{lanczos,scipy}. We describe how we solve system \eqref{eq:four_eq_system} in Appendix \ref{app:solve_rational_system}. Solving \eqref{eq:four_eq_system} reduces to solving a smooth one-dimensional optimization problem, which is computationally trivial.

Together, equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, \eqref{eq:desired_property_3},  and \eqref{eq:desired_property_4}, and the condition $\mu > |\lambda_\text{min}|$, imply that $\ratfct(0)=0$ is the unique minimum of $\ratfct$ on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that $\ratfct$ is non-negative on $[\lambda_\text{min},\lambda_\text{max}]$, which implies that  $\AmatPcSymPlus$ is positive semi-definite. Furthermore, \eqref{eq:desired_property_1} implies that the null space of $\AmatPcSymPlus$ contains the null-space of $\AmatPcSym$. This is important when the operator $\mathcal{A}$ has a large (or infinite) dimensional null space, or a spectrum that clusters at zero. Hessians in ill-posed distributed parameter inverse problems typically have both of these properties. Equation \eqref{eq:desired_property_4} forces the moderate and large positive eigenvalues of $\AmatPcSymPlus$ to be close to the corresponding moderate and large positive eigenvalues of $\AmatPcSym$. I.e., $\ratfct$ modifies the ``important'' part of the spectrum of $\AmatPcSym$ as little as possible. Equation \eqref{eq:desired_property_3} ensures that the negative eigenvalues of $\AmatPcSym$, which result from error in the product-convolution approximation, do not get amplified in magnitude by the function $\ratfct$. There is a tradeoff: if we choose a larger value for $s(\lambda_\text{min})$, then the positive eigenvalues of $\AmatPcSymPlus$ will be closer to the corresponding positive eigenvalues of $\AmatPcSym$ (good), but the erroneous negative eigenvalues of $\AmatPcSym$ will be amplified in magnitude (bad). This is shown in Figure FIG. We find that $\ratfct(\lambda_\text{min})=|\lambda_\text{min}|$ is an appropriate balance between these competing interests.

\subsection{Computational cost}

c\begin{itemize}
	\item computational complexity in terms of PDE solves:
	\item cost of global low rank approximation (data scalability)
	\item cost to build approximation
	\item cost without compression to do a solve
	\item cost with compression to do a solve
	\item computational complexity for H-matrix alone vs. prod-conv + h-matrix $O(C k^2 \log N)$ PDE solves, C is big, k is big. Here $10$
	\item HODLR: C is smaller, but k is bigger, vs. H1 peeling process
	\item H-matrix operations cost for our method solve
	\item irregular grid
\end{itemize}

\section{Rational positive definite modification}
WRITE THIS

\section{Numerical results}
\label{sec:numerical_results}

\textbf{Heat equation}
\begin{itemize}
	\item Hessian per-column error plots for 1, 5, and 15 batches
	\item $H-P$ error vs number of batches for interior and whole domain
	\item $P^{-1}H-I$: error vs rank for $P=R$ and $P=P_\text{pch}$ diffusion time, for 1, 5, and 15 batches
	\item Krylov iterations to tols 1e-1 and 1e-6 vs diffusion parameter
	\item GLR vs PCH vs diffusion time
	\item Krylov method convergence plot: $R$ vs $P_\text{pch}$ vs no preconditioner
	\item Krylov iterations to tols 1e-6 vs mesh size $h$, PCH vs R vs no preconditioner (hold)
	\item true parameter (H)
	\item deterministic reconstruction (H)
	\item noisy measurements (1\%, 5\%, 10\% noise). $u|_\text{top}$ (H)
	\item recovered state at top (H)
	\item mesh scalability of PCH (CC) (H)
	\item put plot data into data directory
	\item save function .pvd files in paraview  directory
\end{itemize}

\begin{itemize}
	\item GLR vs PCH number of obs (Computational cost in PDE solves) (S)
	\item GLR vs PCH aspect ratio (CC) (S)
	\item turn on preconditioner after number of krylov iterations exceeds 15 in an iteration. Rebuild "as needed". (S)
\end{itemize}

Some numerical results here.

\section{Conclusions}
\label{sec:conclusions}

Some conclusions here.

\appendix

\section{Hierarchical matrix details}
\label{app:h_matrix}

Often, large dense matrices of practical interest may be permuted, then partitioned into blocks recursively, in such a way that many off-diagonal blocks of the matrix are numerically low rank, even if the matrix is high rank. Such matrices are known as hierarchical matrices (H-matrices). Many classes of H-matrices exist (H1, H2, HSS, HBS, and more), and all of these types of H-matrices could be used in conjunction with our product-convolution approximation. Here we use classical H1 matrices. For this section, when we say H-matrix, we are referring to H1 matrices.

While a dense $N \times N$ matrix traditionally requires $O(N^2)$ memory to store, H-matrices may be stored using $O(N \log N)$ memory, by storing only the low rank factors for the low rank blocks. Recursive algorithms can take advantage of the H-matrix low rank block structure to perform matrix arithmetic fast. Conventional dense matrix algorithms for matrix inversion, matrix factorization, matrix-vector products, matrix-matrix products, and matrix-matrix addition require either $O(N^2)$ or $O(N^3)$ time and memory, while the aforementioned recursive algorithms can perform these matrix operations in $O(N \log(N)^a)$ time and memory for H-matrices. Here $a=0, 1, 2$, or $3$ depending on the operation and type of H-matrix. For more details on H-matrices, we recommend \cite{HMATRIXGOOD}.

\subsection{H-matrix construction}

%The H-matrix construction method uses geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix (here, these locations are the finite element Lagrange nodes $p_i$, $i=1,\dots,N$) to construct a recursive block partitioning of $\AkerPcMat$, such that all blocks at the finest level of the partitioning are either expected to be low rank, or are small. Then, the method forms low rank approximations of the blocks that are expected to be low rank by sampling specially chosen rows and columns of those blocks. The small high rank blocks are formed directly by evaluating their entries.

In detail, the process of constructing an H-matrix representation of $\AkerPcMat$ proceeds as follows. First, we construct hierarchical partitionings of the degrees of freedom for the columns and rows of the matrix (cluster trees, Section \ref{sec:cluster_trees}). Second, we construct a hierarchical partitioning of the blocks of the matrix, in such a way that many of the blocks in the partitioning are expected to be low rank, and the remaining high rank blocks are small (block cluster tree, Section \ref{sec:block_cluster_tree}). Finally, we form low rank approximations of the blocks of the matrix that are expected to be low rank (adaptive cross approximation, Section \ref{sec:adaptive_cross}), and fill in the remaining high rank  blocks with their numerical values. The first two steps require geometric information about the spatial locations of the degrees of freedom associated with the rows and columns of the matrix, but these steps do not depend on the particular values of matrix entries. The third step requires us to evaluate $O(N \log N)$ specially-chosen entries of $\AkerPcMat$, and we evaluate these entries using \eqref{eq:Akerpcmat_entries}.

\subsubsection{Cluster trees}
\label{sec:cluster_trees}

We use recursive hyperplane splitting to hierarchically cluster the degrees of freedom associated with the columns and rows of the matrix into a \emph{column cluster tree} and a \emph{row cluster tree}, respectively. Here we describe construction of the column cluster tree; the row cluster tree is constructed similarly.

Since we use finite elements to discretize the problem, the $i^\text{th}$ column of $\AkerPcMat$ corresponds to the Lagrange node in $\mathbb{R}^d$ associated with the $i^\text{th}$ finite element basis function. The columns of the matrix thus correspond to a point cloud in $\mathbb{R}^d$. We split this point cloud into two equally sized \emph{child} point clouds, using a hyperplane which is perpendicular to the coordinate axis direction in which the point cloud is widest (e.g., either the $x$, $y$, or $z$ axis in $3$D). The two child point clouds are split in the same way. This splitting process repeats until the point clouds have less than a preset number of points (we use $32$ points). This hierarchical partitioning of the point cloud into smaller and smaller point clouds corresponds to a hierarchical partitioning of the columns of the matrix into smaller and smaller \emph{clusters} of columns. This hierarchical partitioning of the columns forms a binary tree, which is called the column cluster tree. The root of the tree is the set of all columns, and the leaves of the tree are clusters of columns that are not subdivided any further.

A depth-first search ordering of the column cluster tree leaves is then generated. When the columns of the matrix are permuted into this depth-first ordering, the columns associated with each cluster in the cluster tree are contiguous.

In the same way, the degrees of freedom associated with the rows of the matrix are hierarchically clustered into another cluster tree, and a depth-first search ordering for the rows is generated. In our examples, the degrees of freedom for the columns coincide with the degrees of freedom for the rows, so the cluster trees for the rows and columns are the same, but this is not required in general.

\subsubsection{Block cluster tree} 
\label{sec:block_cluster_tree}

We partition the matrix into a recursive hierarchy of mostly low rank blocks called the \emph{block cluster tree}. The idea is that a block of the matrix is likely to be low rank if the point cloud associated with the rows of the block is far away from the point cloud associated with the columns of the block. This is reasonable to expect here because of the locality property of $\mathcal{A}$. Indeed, locality implies that many blocks of the matrix corresponding to far away point cloud clusters will be rank zero.

After reordering the rows and columns of $\AkerPcMat$ via the depth-first ordering described above, we partition the reordered version of $\AkerPcMat$ into a tree of $2 \times 2$ block matrices recursively. We use a geometric admissibility condition (discussed below) to decide which blocks to subdivide, and use the cluster trees for the rows and columns to determine how to subdivide those blocks. For the first stage of subdivision, let $r_1$ and $r_2$ be the children row clusters for the root of the row cluster tree, let $c_1$ and $c_2$ be the children column clusters for the root of the column cluster tree. The matrix $\AkerPcMat$ is partitioned into blocks as follows:
\begin{equation*}
\begin{bmatrix}
\left(\AkerPcMat\right)_{11} & \left(\AkerPcMat\right)_{12} \\
\left(\AkerPcMat\right)_{21} & \left(\AkerPcMat\right)_{22},
\end{bmatrix}
\end{equation*}
where $\left(\AkerPcMat\right)_{11}$ denotes the block of $\AkerPcMat$ with rows $r_1$ and columns $c_1$, $\left(\AkerPcMat\right)_{12}$ denotes the block of $\AkerPcMat$ with rows $r_1$ and columns $c_2$, and so on for $\left(\AkerPcMat\right)_{21}$ and $\left(\AkerPcMat\right)_{22}$.

We now loop through the four blocks, $\left(\AkerPcMat\right)_{11}$, $\left(\AkerPcMat\right)_{12}$, $\left(\AkerPcMat\right)_{21}$, and $\left(\AkerPcMat\right)_{22}$, and decide which blocks should be subdivided further. For the purpose of explanation, consider $\left(\AkerPcMat\right)_{12}$. If
\begin{equation}
\label{eq:weak_admissibility_cond}
\dist\left(r_1, c_2\right) \ge \weakadmconst \min\left(\diam\left(r_1\right), \diam\left(c_2\right)\right),
\end{equation}
then we mark $\left(\AkerPcMat\right)_{12}$ as \emph{admissible} (expected to be low rank) and leave it alone. Here $\dist\left(r_1, c_2\right)$ is the Euclidean distance between the axis-aligned bounding box for the point cloud associated with the row cluster $r_1$, and the axis aligned bounding box for the point cloud associated with the column cluster $c_2$. The quantity $\diam\left(r_1\right)$ is the diameter of the axis aligned bounding box for the point cloud associated with the row cluster $r_1$, and $\diam\left(c_2\right)$ is the analogous diameter associated with the column cluster $c_2$. The quantity $\weakadmconst$ is a scalar constant; we use $\weakadmconst=2.0$. Basically, if the point clouds associated with $r_1$ and $c_2$ are far away from each other relative to their diameters, then we expect that the corresponding block of the matrix will be low rank. This process is repeated for the other blocks to determine which blocks are admissible and which are not. For us, the diagonal blocks $\left(\AkerPcMat\right)_{11}$ and $\left(\AkerPcMat\right)_{22}$ are not admissible because the distance between a point cloud and itself is zero.

Next, we subdivide all blocks that are not admissible and are larger than a predetermined size (we use size $32 \times 32$), using the same process that was used to subdivide $\AkerPcMat$. But now we subdivide a block based on the two child row clusters and two child column clusters for the rows and columns of that block. This subdivision process continues recursively until all blocks are either admissible, or smaller than the predetermined size mentioned above. The resulting hierarchical partitioning of matrix blocks forms a tree, which is called the \emph{block cluster tree}. The root of the tree is the whole matrix, internal nodes in the tree are blocks that are subdivided, and the leaves of the tree are blocks that are either expected to be low rank, or are small.

\subsubsection{Adaptive cross low rank approximation of blocks}
\label{sec:adaptive_cross}

Once the block cluster tree has been constructed, low rank approximations of the admissible (low rank) blocks are formed using the adaptive cross method \cite{ACA}. Let $X \in \mathbb{R}^{m \times m}$ be an admissible block of $\AkerPcMat$. The idea of the adaptive cross method is to form a low rank approximation of $X$ by sampling a small number of rows and columns of $X$. 
\begin{itemize}
	\item Let $C \in \mathbb{R}^{m \times r}$ be a matrix consisting of a subset of $r$ columns of $X$, such that the span of the columns in $C$ approximates the column space of $X$. 
	\item Let $R\in\mathbb{R}^{r \times m}$ be a subset of the rows of $X$, such that the span of the rows in $R$ approximates the row space of $X$.
	\item Let $U \in \mathbb{R}^{r \times r}$ be the block of $X$ consisting of the intersection of the rows from $R$ with the columns from $C$.
\end{itemize}
Then it is well-established that
\begin{equation}
\label{eq:CUR}
X \approx C U^+ R,
\end{equation}
where $U^+$ is the pseudoinverse of $U$ \cite{CUR}. The quality of approximation \eqref{eq:CUR} depends on how well the columns of $C$ approximate the column space of $X$, and how well the rows of $R$ approximate the row space of $X$.

In the adaptive cross method, ``good'' columns, $C$, and rows, $R$, are chosen via an alternating iterative process. An initial set of columns $C$ is chosen. Keeping $C$ fixed, a set of rows $R$ is chosen so that the so that determinant of the submatrix $U$ within $C$ is as large as possible. This may be done via either the maxvol procedure \cite{GOODSUBMATRIX}, or via QR factorization. Now, keeping $R$ fixed, a set of new columns $C$ is chosen so that the submatrix $U$ within $R$ is as large as possible. This process repeats a small number of times. This results in matrices $R$ and $C$ such that the error in the approximation, \eqref{eq:CUR}, is small. The dominant cost of this procedure is the cost of computing $kr$ columns of $X$ and $kr$ rows of $X$, where $k$ is the number of alternating iterations. There is also a small linear algebra overhead cost for the determinant maximization process that is performed at each step. For more details on adaptive cross low rank approximation, see \cite{ACA}. The key point is that the adaptive cross method allows us to form a rank-$r$ approximation of an $m \times m$ block of the matrix via a process that only requires accessing $O(mr)$ entries of that block.

We use the adaptive cross method to form low rank approximations for each admissible block of $\AkerPcMat$. We directly compute all entries of the small dense blocks of $\AkerPcMat$ that are not admissible. This process requires us to compute $O(r^2 N \log N)$ entries of $\AkerPcMat$, which is relatively cheap compared to the operator actions of $\Aop$ that are required to form the product-convolution approximation.

\section{Fast ellipsoid intersection test}
\label{sec:fast_ellipsoid_intersection_test}
The procedure for choosing sample points relies on quickly determining whether two ellipsoids intersect. Let $E_p$ and $E_q$ be the ellipsoids defined as
\begin{align*}
	E_p :=& \{x : (x - \mu_p)^T \Sigma_p^{-1} (x - \mu_p) \le \tau^2\} \\
	E_q :=& \{x : (x - \mu_q)^T \Sigma_q^{-1} (x - \mu_q) \le \tau^2\}, \\
\end{align*}
where $\mu_p, \mu_q \in \mathbb{R}^d$, and $\Sigma_p, \Sigma_q \in \mathbb{R}^{d \times d}$ are positive definite. Let $K$ be the following one dimensional convex function:
\begin{equation*}
	K(s) := 1 - \frac{1}{\tau^2} (\mu_p - \mu_q)^T \left(\frac{1}{1-s}\Sigma_p + \frac{1}{s}\Sigma_q\right)^{-1}(\mu_p - \mu_q)	
\end{equation*}
In \cite{ELLIPSOIDINTERSECT} it is shown that $E_p \cap E_q = \{\}$ if and only if $K(s) < 0$ for some $s\in (0,1)$. We check whether $E_p$ and $E_q$ intersect by minimizing $K(s)$ on $(0,1)$. If $K(s^*) <0$ at the minimizer $s^*$, then $E_p \cap E_q = \{\}$. Otherwise $E_p \cap E_q \neq \{\}$.

The function $K(s)$ may be evaluated quickly for many $s$ by pre-computing the solution to the generalized eigenvalue problem
\begin{equation*}
	\Sigma_p P = \Sigma_q P \Lambda,
\end{equation*}
where $P \in \mathbb{R}^{d \times d}$ is the matrix of generalized eigenvectors (which may be non-orthogonal), and $\Lambda = \diag(\lambda_1,\lambda_2,\dots,\lambda_d)$ is the diagonal matrix of generalized eigenvalues $\lambda_i$. The matrix $P$ simultaneously diagonalizes $\Sigma_p$ and $\Sigma_q$, in the sense that $P^T\Sigma_p P = \Lambda$, and $P^T\Sigma_q P = I$, where $I$ is the $d \times d$ identity matrix. Using this diagonalization, and some algebraic manipulations, we may write $K(s)$ as
\begin{equation}
\label{eq:Ks_generalized}
K(s) = 1 - \frac{1}{\tau^2} \sum_{i=1}^d \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2,
\end{equation}
where $v := \Phi^T\left(\mu_p - \mu_q\right)$. We compute the generalized eigenvalue decomposition of $\Sigma_p$ and $\Sigma_q$, then minimize $K(s)$ in the form \eqref{eq:Ks_generalized} on the interval $(0,1)$ using Brent's algorithm \cite{BRENT} (any fast 1 dimensional convex optimization routine may be used). The resulting algorithm for checking whether $E_p$ and $E_q$ intersect is summarized in Algorithm \ref{alg:ellipsoid_intersection_test}.

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	\SetKwProg{Fn}{Function}{}{}
	\Input{Ellipsoid $E_p$ with mean $\mu_p$ and covariance $\Sigma_p/\tau^2$\\
		Ellipsoid $E_q$ with mean $\mu_q$ and covariance $\Sigma_q/\tau^2$ 
	}

	\Output{Boolean $\text{ellipsoids\_intersect}$ which is true if $E_p \cap E_q \neq \{\}$ and false otherwise}

	Solve generalized eigenvalue problem $\Sigma_p \Phi = \Sigma_q \Phi \Lambda$

	$v \gets \Phi^T\left(\mu_p - \mu_q\right)$

	$\displaystyle K^* \gets \min_{s \in (0,1)}~1 - \frac{1}{\tau^2} \sum_{i=1}^d \frac{s(1-s)}{1 + s (\lambda_i - 1)}v_i^2$

	\If{$K^* < 0$}{

		$\text{ellipsoids\_intersect} \gets \text{False}$

	}
	\Else{

		$\text{ellipsoids\_intersect} \gets \text{True}$

	}
	\caption{Determining whether two ellipsoids intersect}
	\label{alg:ellipsoid_intersection_test}
\end{algorithm2e}

\section{Solving for rational function scalars}
\label{app:solve_rational_system}

Using the definition of $\ratfct$ in \eqref{eq:rational_function_scalar}, and basic calculus and linear algebra, we may write Equations \eqref{eq:desired_property_1}, \eqref{eq:desired_property_2}, and \eqref{eq:desired_property_4} in the following matrix form:
\begin{equation}
\label{eq:rational_3_linear_system}
\begin{bmatrix} 1 & 0 & 1/\mu \\ 0 & 1 & -1/\mu^2 \\ 1 & \lambda_\text{max} & 1/(\lambda_\text{max}+\mu) \end{bmatrix} \begin{bmatrix} c_0 \\ c_1 \\ c_2 \end{bmatrix} = \begin{bmatrix} 0 \\ 0 \\ \lambda_\text{max} \end{bmatrix},
\end{equation}
and we may write Equation \eqref{eq:desired_property_3} as follows: 
\begin{equation}
\label{eq:rational_extra_eq}
c_0 + c_1 \lambda_\text{min} + \frac{c_2}{\lambda_\text{min}+\mu} - |\lambda_\text{min}| = 0.
\end{equation}
Let 
\begin{equation*}
g(\mu) := c_0(\mu) + c_1(\mu) \lambda_\text{min} + \frac{c_2(\mu)}{\lambda_\text{min}+\mu} - |\lambda_\text{min}|
\end{equation*}
denote the left hand side of \eqref{eq:rational_extra_eq}, but with $c_0(\mu)$, $c_1(\mu)$, and $c_2(\mu)$ being the solution of \eqref{eq:rational_3_linear_system} for the given value of $\mu$. I.e., within $g(\mu)$ the quantities $c_0$, $c_1$, and $c_2$ are no longer free variables; they now depend on $\mu$ implicitly through the solution of a linear system. We solve the combined system, \eqref{eq:rational_3_linear_system} and $\eqref{eq:rational_extra_eq}$, by solving the one-dimensional nonlinear root finding problem
\begin{equation}
\label{eq:g_mu_zero}
g(\mu)=0
\end{equation}
with bound $\mu > |\lambda_\text{min}|$ and initial guess $\mu_0 = 2 |\lambda_\text{min}|$. Each evaluation of $g$ requires solving the $3 \times 3$ linear system \eqref{eq:rational_3_linear_system}, and once the solution $\mu$ is found, we solve \eqref{eq:rational_3_linear_system} one more time to get the final values of $c_0$, $c_1$, and $c_2$.

\section{Proofs}
\label{app:proofs}

%These results rely on straightforward prerequisites that we present in Lemma \ref{lem:T_is_linear}. We prove correctness of Algorithm \ref{alg:weighting_functions_incremental} in Theorem \ref{thm:wi_alg_is_correct}. Theorem \ref{thm:wi_alg_is_correct} relies on a prerequisite result which we present in Lemma \ref{lem:psi_delta_eval}.

\begin{lem}[Linearity of weighting function solution operator]
	\label{lem:T_is_linear}
	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization_initial}	as a function of $f$. That is, 
		\begin{equation}
		\label{eq:opt_soln_operator}
		\mathcal{T}(f) := \horizinterpolant_*,
		\end{equation}
		where $\horizinterpolant_*$ is the minimizer of \eqref{eq:generic_optimization_initial}. The operator $\mathcal{T}$ is a linear operator.
\end{lem}

\begin{proof}
	asdf
\end{proof}

\begin{proof}[Proof of Theorem \ref{thm:smoothest_interpolant}]
	Let $f = (\varphi_1(y), \varphi_2(y), \dots, \varphi_r(y))$ and let $e_i = (0,\dots,0,1,0,\dots,0)$ be the length $r$ vector with $i^\text{th}$ component equal to one and all other components equal to zero. The solution to optimization problem \eqref{eq:uy_optimization_problem} is given by
	\begin{align*}
	\mathcal{T}(f) &= \varphi_1(y) \mathcal{T}(e_1) + \varphi_2(y) \mathcal{T}(e_2) + \dots +\varphi_r(y) \mathcal{T}(e_r) \\
	&= \varphi_1(y) w_1 + \varphi_2(y) w_2 + \dots +\varphi_r(y) w_r \\
	&= \horizinterpolant_y.
	\end{align*}
	In the first line we used linearity of the solution map, $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), in the second line we used the fact that $w_i = \mathcal{T}(e_i)$, and in the third line we used the definition of $\horizinterpolant_y$ in \eqref{eq:u_defn_in_thm}.
\end{proof}

\begin{proof}[Proof of Proposition \ref{thm:wi_sum_to_one}]
	By linearity of $\mathcal{T}$ (Lemma \ref{lem:T_is_linear}), we have
	\begin{equation*}
	\sum_{i=1}^r w_i(x) = \sum_{i=1}^r \mathcal{T}(e_i)(x) = \mathcal{T}\left(\sum_{i=1}^r e_i\right)(x) = \mathcal{T}(\mathbf{1})(x) = 1,
	\end{equation*}
	where $\mathbf{1} = (1,1,\dots,1)$ is the length $r$ vector which has all components equal to one. The last equality ($\mathcal{T}(\mathbf{1})(x) = 1$) holds because the constant function $C \in H^2(\Omega)$, $C(x)=1$, satisfies the constraint of optimization problem \eqref{eq:generic_optimization_initial} with $f = \mathbf{1}$, and yields the minimum possible value of zero for the objective function.
\end{proof}

\begin{defn}[Moments of $\phi_x$]
	\label{defn:alpha_rho_mu_sigma}
	We define the spatially varying scaling factor, $\alpha$, the scaled impulse response, $\widehat{\phi}_x$, the spatially varying mean, $\mu$, and the spatially varying covariance, $\Sigma$, as follows:
	\begin{align*}
	\alpha:\Omega \rightarrow \mathbb{R}, &\qquad \alpha(x) := \int_{\Omega} \phi_x(z) dz, \\
	\widehat{\phi}_x:\Omega \rightarrow \mathbb{R}, &\qquad \widehat{\phi}_x := \phi_x \big/ \alpha(x), \\
	\mu:\Omega \rightarrow \mathbb{R}^d, &\qquad \mu(x) := \int_\Omega z \widehat{\phi}_x(z) dz, \\
	\Sigma:\Omega \rightarrow \mathbb{R}^{d \times d}, &\qquad\Sigma(x) := \int_\Omega (z - \mu(x))(z - \mu(x))^T \widehat{\phi}_x(z) dz.
	\end{align*}
\end{defn}

\begin{proof}[Proof of Theorem \ref{thm:vol_mean_cov}]
	Let $v \in L^2$ be arbitrary. We have
	\begin{align*}
		\left(v, \alpha\right)_{L^2(\Omega)} =& \int_\Omega v(x) \int_{\Omega} \langle\mathcal{A},\delta_x\rangle^*(z) dz dx\\
		=& \int_\Omega v(x) \int_{\Omega} A(z,x) dz dx \\
		=& \int_\Omega \int_\Omega v(x) A(z,x) C(z) dz dx \\
		=& \left(\mathcal{A}^T C\right)(v),
	\end{align*}
	which implies $\alpha = \left(\mathcal{A}^TC\right)^*$.

	Using similar techniques, we have
	\begin{align*}
		\left(v, \alpha \cdot \mu^i\right)_{L^2(\Omega)} =& \int_\Omega v(x) \alpha(x) \int_{\Omega} z^i \langle\mathcal{A},\delta_x\rangle^*(z)/\alpha(x) dz dx\\
		=& \int_\Omega \int_\Omega v(x) L^i(z) A(z,x) dz dx \\
		=& \left(\mathcal{A}^T L^i\right)(v),
	\end{align*}
	which implies $\mu^i = \left(\mathcal{A}^T L^i\right)^* / \alpha$.

	Using the identity
	\begin{equation*}
		\Sigma(x) = \int_\Omega zz^T \widehat{\phi}_x(z) dz - \mu(x) \mu(x)^T,
	\end{equation*}
	we have
	\begin{align*}
		\left(v, \alpha \cdot \left(\Sigma^{ij}+\mu^i \cdot \mu^j\right)\right)_{L^2(\Omega)} =& \int_\Omega v(x) \alpha(x) \int_\Omega z^i z^j \langle\mathcal{A},\delta_x\rangle^*(z) / \alpha(x) dz dx \\
		=& \int_\Omega \int_\Omega v(x) Q^{ij}(z) A(z,x) dz dx \\
		=& \left(\mathcal{A}Q^{ij}\right)(v),
	\end{align*}
	which implies $\Sigma^{ij} = \left(\mathcal{A}^T Q^{ij}\right)^* / \alpha - \mu^i\cdot\mu^j$.

\end{proof}

\section{Background: distributions}
\label{app:distributions}

Let $\overline{\Omega}$ be the closure of $\Omega$, and let $C(\overline{\Omega})$ be the space of continuous functions mapping $\overline{\Omega}\rightarrow \mathbb{R}$. The action of $\mathcal{A}$ is extended to distributions $\mu:C\left(\overline{\Omega}\right) \rightarrow \mathbb{R}$ via the formula
\begin{equation*}
\langle\mathcal{A},\mu\rangle(v) := \int_\Omega v(y) \mu\left(A(y, \cdot)\right) dy, 
\end{equation*}
where $A(y,\cdot)$ is the function $x \mapsto A(y,x)$. This is derived formally as follows:
\begin{align*}
\langle\mathcal{A},\mu\rangle(v) &= \int_\Omega \int_\Omega v(y) A(y,x) \textrm{``} \mu(x) \textrm{''} dx dy \\
&= \int_\Omega v(y) \int_\Omega A(y,x) \textrm{``} \mu(x) \textrm{''} dx \\
&= \int_\Omega v(y) \mu\left(A(y,\cdot)\right) dx.
\end{align*}
For example, the delta distribution $\delta_x$ is defined by $\delta_x(v) = v(x)$, and the action of $\mathcal{A}$ on $\delta_x$ is given by
\begin{equation}
\label{eq:appendix_impulse_response}
\langle\mathcal{A},\delta_x\rangle(v)	= \int_\Omega v(y) A(y,x) dy.
\end{equation}
Recall that the impulse response of $\mathcal{A}$ to $\delta_x$, denoted $\phi_x$, is the Riesz representation of $\langle\mathcal{A},\delta_x\rangle$. From \eqref{eq:appendix_impulse_response} and the definition of the Riesz representation, we have $\phi_x = \langle\mathcal{A},\delta_x\rangle^* = A(~\cdot~,x)$. That is, $\phi_x$ is the function $y \mapsto A(y,x)$.

\section{Discretization}
\label{sec:discretization}

Let $f_i$, $i=1,\dots,N$ be a set of finite element basis functions, let $h$ denote the mesh size parameter for the finite element mesh, and let
\begin{equation*}
V_h = \Span\left(f_1, f_2, \dots, f_N\right)
\end{equation*}
be the corresponding  finite element space with the $L^2$ inner product. Functions $u\in L^2(\Omega)$ are approximated by functions $u_h \in V_h$. In turn, functions $u_h\in V_h$ are represented in computations by length $N$ arrays $\mathbf{u}$, such that the array entries of $\mathbf{u}$ are the coordinates of $u_h$ in the finite element basis. That is,
\begin{equation}
\label{eq:finite_bijection}
u_h(x) = \sum_{i=1}^N \mathbf{u}_i f_i(x).
\end{equation}

We write $\mathbf{M} \in \mathbb{R}^{N \times N}$ to denote the finite element \emph{mass matrix}, which has entries
\begin{equation*}
\mathbf{M}_{ij} = \left(f_i, f_j\right)_{L^2(\Omega)} = \int_{\Omega} f_i(x) f_j(x) dx,
\end{equation*}
and we write $\mathbb{R}^N_\mathbf{M}$ to denote the space $\mathbb{R}^N$ with the matrix-weighted inner product
\begin{equation*}
\left(\mathbf{u},\mathbf{v}\right)_\mathbf{M} := \mathbf{v}^T\mathbf{M} \mathbf{u}.
\end{equation*}
Direct calculation shows that $V_h$ is isometrically isomorphic $\mathbb{R}^N_\mathbf{M}$. That is, equation \eqref{eq:finite_bijection} establishes a bijection between functions $u_h \in V_h$ and coefficient vectors $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, and this bijection preserves the inner product in the sense that $\left(u_h, v_h\right)_{L^2(\Omega)} = \left(\mathbf{u},\mathbf{v}\right)_\mathbf{M}$. Likewise, the dual space, $V_h'$, is isometrically isomorphic to $\mathbb{R}^N_{\mathbf{M}^{-1}}$.

Here vectors in $\mathbf{u} \in \mathbb{R}^N_{\mathbf{M}}$ and $\boldsymbol{\sigma} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ are viewed as column vectors, and $\mathbf{u}^T$ and $\boldsymbol{\sigma}^T$ are the row vectors that result from taking the standard matrix transposes of $\mathbf{u}$ and $\boldsymbol{\sigma}$, respectively. The discrete dual of a vector $\mathbf{v}$ with respect to the $\mathbf{M}$ inner product is written as $\mathbf{v}^*$, and $\mathbf{v}^*$ is viewed as a column vector.
%These isomorphisms allow us to identify functions $u_h \in V_h$ with their corresponding coefficient arrays $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, and linear functionals $\sigma_h \in V_h'$ with their corresponding coefficient arrays $\boldsymbol{\sigma} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$.

\subsection{Discretized operations}
\label{app:discretized_operations}

Here we list the function space operations on $V_h$ and $V_h'$ that are required for this paper, and show how to perform the corresponding matrix and vector operations on $\mathbb{R}^N_{\mathbf{M}}$ and $\mathbb{R}^N_{\mathbf{M}^{-1}}$.

%We summarize coefficient vector versions of the function space operation that are used in this paper in Appendix \ref{app:discretized_operations}.
%
%As described in Section \ref{sec:discretization}, a function $u_h$ in a finite element space $V_h$ may be written in terms of its vector of coefficients, $\mathbf{u} \in \mathbb{R}^N_\mathbf{M}$, with respect the finite element basis $\{\phi_i\}_{i=1}^N$. Recall that $\mathbf{M}$ is the mass matrix. We have
%\begin{equation*}
%	u(x) \approx u_h(x) = \sum_{i=1}^N \mathbf{u}_i \phi_i(x).
%\end{equation*}
%The finite element function space $V_h$ is isometrically isomorphic to the coefficient space $\mathbb{R}^N_\mathbf{M}$, and the dual space $V_h'$ is isometrically isomorphic to $\mathbb{R}^N_{\mathbf{M}^{-1}}$. In computations, we perform operations with coefficient vectors in $\mathbb{R}^N_\mathbf{M}$ and coefficient dual vectors in $\mathbb{R}^N_{\mathbf{M}^{-1}}$, rather than functions in $V_h$ and functionals in $V_h'$. Here we describe how to perform common function space operations in terms of coefficient vectors and coefficient dual vectors.

\begin{description}
	\item[Coefficients of a functional:] The coefficient dual vector, $\boldsymbol{\sigma} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$, corresponding to a linear functional $\sigma_h \in V_h'$ has components given by
	\begin{equation*}
		\boldsymbol{\sigma}_i = \sigma_h(f_i), \quad i=1,\dots,N.
	\end{equation*}
	\item[Action of a functional:] The action of a linear functional $\sigma_h \in V_h'$ on a function $u_h \in V_h$ may be computed as follows:
	\begin{equation*}
		\sigma_h(v_h) = \boldsymbol{\sigma}^T \mathbf{u}.
	\end{equation*}
	\item[Riesz representation:] Let $\sigma_h \in V_h'$. The Riesz representation of $\sigma_h$ is the unique  function $\sigma_h^* \in V_h$ satisfying 
	\begin{equation*} 
	\left(\sigma_h^*, v_h\right)_{L^2(\Omega)} = \sigma_h(v_h) \quad\forall v_h \in V_h.
	\end{equation*}
	The coefficient dual vector $\boldsymbol{\sigma}^*$ corresponding to $\sigma_h$ is given by
	\begin{equation*} 
		\boldsymbol{\sigma}^* = \mathbf{M}^{-1} \boldsymbol{\sigma}.
	\end{equation*}
	\item[Finite element $L^2$ projection:] Let $g \in L^2(\Omega)$. The orthogonal projection of $g$ onto $V_h$ with respect to the $L^2$ inner product is the unique function $g^\pi_h \in V_h$ satisfying
	\begin{equation*}
		\left(g^\pi_h, v_h\right)_{L^2(\Omega)} = \left(g, v_h\right)_{L^2(\Omega)} \quad\forall v_h \in V_h.
	\end{equation*}
	 Let $\left(\mathbf{g}^\pi\right)^* \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ be the vector with components
	\begin{equation*}
		\left(\mathbf{g}^\pi\right)^*_i = \int_\Omega g(x) f_i(x) dx, \quad i=1,\dots,N.
	\end{equation*}
	The coefficient vector $\mathbf{g}^\pi \in \mathbb{R}^N_{\mathbf{M}}$ corresponding to $g^\pi_h$ is given by
	\begin{equation*}
		\mathbf{g}^\pi = \mathbf{M}^{-1}\left(\mathbf{g}^\pi\right)^*.
	\end{equation*}
	\item[Finite element interpolation:] Let $p_i \in \mathbb{R}^d$, $i=1,\dots,N$, be the Lagrange nodes associated with the finite element basis functions $f_i$. The finite element interpolant of a function $g \in C\left(\overline{\Omega}\right)$ is the function $g^I \in V_h$ given by
	\begin{equation*}
		g_h^I(x) = \sum_{i=1}^N g(p_i) f_i(x).
	\end{equation*}
	The coefficient vector, $\mathbf{g}$, for $g_h^I$ has components 
	\begin{equation*}
	\mathbf{g}_i^I = g(p_i), \quad i=1.,\dots,N.
	\end{equation*}
	\item[Projection vs. interpolation:]
	Typically, $g^\pi$ is a more accurate approximation of $g$ than $g^I$, but $g^I$ is cheaper to compute than $g^\pi$ because computing $g^\pi$ requires solving a linear system, while computing $g^I$ does not. If the finite element basis satisfies the following Kronecker property,
	\begin{equation*}
		f_i(p_j) = \begin{cases}1 & i=j \\
		0 & \text{otherwise},\end{cases}
	\end{equation*}
	then $g^I(p_i) = g(p_i)$ for $i=1,\dots,N$ (i.e., the interpolant equals the original function at the Lagrange nodes). Interpolation is typically sufficiently accurate for practical purposes if the Kronecker property is satisfied. Projection should be used if the Kronecker property is not satisfied.
	\item[Matrix representation of an operator:] Let $\mathcal{B}_h : V_h \rightarrow V_h'$ be a linear operator, and let $\mathbf{B} \in \mathbf{R}^{N \times N}$ be the matrix with entries
	\begin{equation*}
		\mathbf{B}_{ij} = \left(\mathcal{B}_h f_j\right)(f_i).
	\end{equation*}
	The matrix $\mathbf{B}$ maps $\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}$ by matrix multiplication, and
	\begin{equation*}
		\sigma_h = \mathcal{B}_h u_h \quad \Leftrightarrow \quad \boldsymbol{\sigma} = \mathbf{B} \mathbf{u}.
	\end{equation*}
	\item[Transpose:] The matrix $\mathbf{B}^T$ maps $\mathbb{R}^N_\mathbf{M} \rightarrow \mathbb{R}^N_{\mathbf{M}^{-1}}$ via matrix multiplication, and
	\begin{equation*}
		\sigma_h = \mathcal{B}_h^T u_h \quad \Leftrightarrow \quad \boldsymbol{\sigma} = \mathbf{B}^T \mathbf{u}.
	\end{equation*}

	\item[Distributions:] Let $\mu:C\left(\overline{\Omega}\right)\rightarrow \mathbb{R}$ be a distribution. If $\mu(f_i)$ is well-defined for $i=1,\dots,N$, then the restriction of $\mu$ to domain $V_h$ is a linear functional $\mu_h\in V_h'$. The functional $\mu_h$ has a coefficient dual vector $\boldsymbol{\mu} \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ with entries 
	\begin{equation*}
	\boldsymbol{\mu}_i=\mu(f_i), \quad i=1,\dots,N.
	\end{equation*} 
	For example, let $\delta_p$ be the Dirac distribution centered at a point $p\in\Omega$. If the finite element basis functions, $f_i$, are continuous at $p$, then restricting the domain of $\delta_p$ to the space $V_h$ yields a linear functional in $V_h'$. The corresponding coefficient dual vector, $\boldsymbol{\delta}_p$, has entries 
	\begin{equation*}
	\left(\boldsymbol{\delta}_p\right)_i = f_i(p), \quad i=1,\dots,N.
	\end{equation*}

	\item[Action of an operator on a distribution:] Since $V_h$ is finite-dimensional, the restriction, $\mu_h \in V_h$, of a distribution, $\mu$, to the space $V_h$ has a a Riesz representation $\mu_h^* \in V_h$. Let $B_h$ be the integral kernel associated with an operator $\mathcal{B}_h:V_h \rightarrow V_h'$. For $v_h \in V_h$, we have
	\begin{align*}
		\langle \mathcal{B}_h, \mu_h \rangle(v) &= \int_\Omega v_h(y) \mu_h\left(B_h(y, \cdot)\right) dy \\
		&= \int_\Omega v_h(y) \left(\int_\Omega B_h(y,x) \mu_h^*(x) dx\right) dy \\
		&= \int_\Omega \int_\Omega v_h(y) B_h(y,x) \mu_h^*(x) dx dy = \left(\mathcal{B}_h \mu_h^*\right)(v).
	\end{align*}
	Going from the first to the second line we used the definition of the Riesz representation. The action of $\mathcal{B}_h$ on $\mu_h$ in the sense of distributions is therefore equal to the action of $\mathcal{B}_h$ on $\mu^*_h$ in the conventional sense. We have
	\begin{equation*}
		\sigma_h = \langle \mathcal{B}_h, \mu_h \rangle \quad \Leftrightarrow \quad \sigma_h = \mathcal{B}_h \mu_h^* \quad \Leftrightarrow \quad \boldsymbol{\sigma} = \mathbf{B} \mathbf{M}^{-1} \boldsymbol{\mu}.
	\end{equation*}
\end{description}

\subsection{Discretized mean and covariance estimation}

\begin{algorithm2e}
	\SetAlgoNoLine
	\SetKwInOut{Input}{Input}
	\SetKwInOut{Output}{Output}
	{	
		\Input{Operator $\mathbf{A}$}
		\Output{$\boldsymbol{\alpha}\in \mathbb{R}^N$, $\boldsymbol{\mu}\in \mathbb{R}^{d \times N}$, and $\boldsymbol{\Sigma} \in \mathbb{R}^{d \times d \times N}$}

		Form vector $\mathbf{C} \in \mathbb{R}^N_{\mathbf{M}}$ by either projecting or interpolating constant function $C(x)=1$ onto $V_h$

		$\boldsymbol{\alpha} = \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{C}\right)^*$

		\For{$i=1,2,\dots,d$}{
			Form vector $\mathbf{L}^i \in \mathbb{R}^N_{\mathbf{M}}$ by either projecting or interpolating linear function $L^i(x) = x^i$ onto $V_h$

			$\mathbf{\mu}^i = \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{L}^i\right)^* / \boldsymbol{\alpha}$
		}
		\For{$i=1,2,\dots,d$}{
			\For{$j=1,\dots,i$}{
				Form quadratic function $\mathbf{Q}^{ij} = x^i x^j$

				$\boldsymbol{\Sigma}^{ij} = \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{Q}^{ij}\right)^* / \boldsymbol{\alpha} - \boldsymbol{\mu}^i\cdot \boldsymbol{\mu}^j$

				$\boldsymbol{\Sigma}^{ij} = \boldsymbol{\Sigma}^{ji}$

			}
		}

	}
	\caption{Discretized version of Algorithm \ref{alg:varhpi_mean_cov}}
	\label{alg:varhpi_mean_cov_discrete}
\end{algorithm2e}

In computations, $\alpha$, $\mu^{i}$, and $\Sigma^{ij}$ are replaced with finite element approximations $\alpha_h, \mu_h^i, \Sigma_h^{ij}\in V_h$, respectively, which have coefficient vectors $\boldsymbol{\alpha}, \boldsymbol{\mu}^{ij}, \boldsymbol{\Sigma}^{ij}\in\mathbb{R}^N_\mathbf{M}$, respectively. The functions $C$, $L^i$, and $Q^{ij}$ are replaced with either finite element projections onto $V_h$ (if the finite element basis does not satisfy the Kronecker property), or finite element interpolations onto $V_h$ (if the finite element basis satisfies the Kronecker property). In our numerical examples we use finite element basis functions that satisfy the Kronecker property, and we use finite element interpolations. The coefficient vectors corresponding to either the interpolations or projections are denoted $\mathbf{C}$, $\mathbf{L}^i$, and $\mathbf{Q}^{ij}$, respectively. The discretized versions of \eqref{eq:vol_mean_var_thm1}, \eqref{eq:vol_mean_var_thm2}, and \eqref{eq:vol_mean_var_thm3} are
\begin{align*}
\boldsymbol{\alpha} &= \mathbf{M}^{-1}\mathbf{A}^T\mathbf{C} \\
\boldsymbol{\mu}^i &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{L}^i\right) / \boldsymbol{\alpha}\\
\boldsymbol{\Sigma}^{ij} &= \left(\mathbf{M}^{-1}\mathbf{A}^T \mathbf{Q}^{ij}\right) / \boldsymbol{\alpha} - \boldsymbol{\mu}^i\cdot \boldsymbol{\mu}^j,
\end{align*}
respectively. Here $\mathbf{f} \cdot \mathbf{g}$ and $\mathbf{f} / \mathbf{g}$ denote the component-wise multiplication and division of vectors, respectively.

\subsection{Discretized sampling of impulse responses}
In computations, the Dirac comb coefficient dual vector for $\dcomb_h^b$ is the vector $\boldsymbol{\dcomb}^b \in \mathbb{R}^N_{\mathbf{M}^{-1}}$ which has entries
\begin{equation*}
\boldsymbol{\dcomb}^b_i = \sum_{x_j \in S_b} f_i(x_j), \quad i=1,\dots,N.
\end{equation*}
The Dirac comb impulse response coefficient vector for $\combresponse^h_b$ is the vector $\boldsymbol{\combresponse}_b \in \mathbb{R}^N_\mathbf{M}$ given by
\begin{equation*}
\boldsymbol{\combresponse}_b = \mathbf{M}^{-1} \mathbf{A}^T \mathbf{M}^{-1}\boldsymbol{\dcomb}^b.
\end{equation*}

\section{DISORGANIZED OLD STUFF}

\begin{proof}[Proof of Theorem \ref{alg:weighting_functions_incremental}]
	For notational convenience, in this proof we write $w$, $\lambda$, and $e$ instead of $w_i$, $\lambda_i$, and $e_i$, respectively. Here subscripts for $\lambda$, and $e$ instead denote components of those vectors. For example, here $\lambda_j$ denotes the $j^\text{th}$ component of the vector $\lambda$, which corresponds to the $j^\text{th}$ component of the vector $\lambda_i$ in Algorithm \ref{alg:weighting_functions_incremental}.

	From standard optimization theory, the solution to optimization problem \ref{eq:wi_optimization_problem} is the stationary point of the Lagrangian,
	\begin{equation*}
	\mathcal{L}(w,\lambda) = \frac{1}{2}\|\Delta w\|^2 + \sum_{j=1}^r \lambda_j\left(w(x_j) - e_j\right).
	\end{equation*}
	The plan of this proof is to show that $\nabla \mathcal{L}(w,\lambda)=0$, which will verify that $w$ solves optimization problem \ref{eq:wi_optimization_problem}.

	Using the definition of $w_i$ in \eqref{eq:definition_of_wi_alg}, we may write the gradient of $\mathcal{L}$ with respect to $\lambda_j$ as
	\begin{equation*}
	\nabla_{\lambda_j} \mathcal{L}(w, \lambda) = w(x_j) - e_j
	= \beta - \sum_{k=1}^r \lambda_k \secondgreens_k(x_j) - e_j.
	\end{equation*}
	Using Lemma \ref{lem:psi_delta_eval} with $h=\secondgreens_k$, the equation for $\secondgreens_k$ in \eqref{eq:theta_eq}, and the definition of $S$ in \eqref{eq:defn_of_S_alg}, we have
	\begin{align*}
	\sum_{k=1}^r \lambda_k \secondgreens_k(x_j) &= \sum_{k=1}^r \lambda_k \left(\firstgreens_j, \Delta \secondgreens_k\right)_{L^2(\Omega)}\\
	&= \sum_{k=1}^r \lambda_k \left(\firstgreens_j, \firstgreens_k\right)_{L^2(\Omega)} = \left(S \lambda\right)_j.
	\end{align*}
	Combining all of these components of $\nabla_\lambda \mathcal{L}$ into a single vector and using the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} yields
	\begin{align*}
	\nabla_\lambda \mathcal{L}(w, \lambda) &= \beta \mathbf{1} - e - S \lambda \\
	&= \beta \mathbf{1} - e - S \left(-S^{-1}\left(e - \beta \mathbf{1}\right)\right) = 0.
	\end{align*}

	Taking the directional derivative of $\mathcal{L}$ with respect to $w$ in direction $h$ and using Lemma \ref{lem:psi_delta_eval} yields
	\begin{align}
	\nabla_w \mathcal{L}(w, \lambda)h &= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j h(x_j) \nonumber \\
	&= \left(\Delta w, \Delta h\right)_{L^2(\Omega)} + \sum_{j=1}^r \lambda_j \left(\firstgreens_j, \Delta h\right)_{L^2(\Omega)} + \lambda_j \avg(h) \nonumber \\
	&= \left(\Delta w + \sum_{j=1}^r \lambda_j \firstgreens_j,~ \Delta h\right)_{L^2(\Omega)} + \avg(h) \mathbf{1}^T \lambda. \label{eq:L_w_38}
	\end{align}
	Using the definition of $w$ in Algorithm \ref{alg:weighting_functions_incremental}, and the fact that the Laplacian of a constant function is zero, we have
	\begin{equation*}
	\Delta w = -\sum_{j=1}^r \lambda_j \Delta \secondgreens_j 
	= -\sum_{j=1}^r \lambda_j \firstgreens_j,
	\end{equation*}
	which implies that the first term in \eqref{eq:L_w_38} is zero. By the definition of $\lambda$ in \eqref{eq:definition_of_mu_alg} and $\beta$ in \eqref{eq:definition_of_alpha_alg}, we have
	\begin{align*}
	\mathbf{1}^T \lambda &= -\mathbf{1}^T S^{-1} \left(e - \beta \mathbf{1}\right) \\
	&= -\mathbf{1}^T S^{-1} e + \beta \mathbf{1}^T S^{-1} \mathbf{1} \\
	&= -\mathbf{1}^T S^{-1} e + \frac{\mathbf{1}^T S^{-1} e}{\mathbf{1}^T S^{-1} \mathbf{1}} \mathbf{1}^T S^{-1} \mathbf{1} = 0,
	\end{align*}
	which implies that the second term in \eqref{eq:L_w_38} is zero. Thus $\nabla_w \mathcal{L}(w, \lambda) = 0$, so Algorithm \ref{alg:weighting_functions_incremental} produces the solution to optimization problem \ref{eq:wi_optimization_problem}.
\end{proof}

\begin{defn}[Weighting function optimization problem]
	Let $\Omega \subset \mathbb{R}^d$, $d\in\{1,2,3\}$, be an open bounded domain with, let $x_i \in \Omega$ for $i=1,\dots,r$, and let $f \in \mathbb{R}^r$. We consider the optimization problem
	\begin{equation}
	\label{eq:generic_optimization_initial}
	\begin{aligned}
	\min_{\gamma \in H^2_N(\Omega)} &\quad \frac{1}{2}\norm{\Delta \gamma}_{L^2(\Omega)}^2 \\
	\text{such that} &\quad \gamma(x_i) = f_i, \quad i=1,\dots,r.
	\end{aligned}
	\end{equation}
	A minimizer, $\gamma^*$, of \eqref{eq:generic_optimization_initial} is a function defined on $\Omega$ that interpolates the data pairs $(x_i,f_i)$, and is as flat as possible in a least squares sense. 
	We denote the objective function for optimization problem \eqref{eq:generic_optimization_initial} by 
	\begin{equation}
	\label{eq:wi_objective}
	J(\gamma) := \frac{1}{2}||\Delta \gamma||^2_{L^2(\Omega)}.
	\end{equation}
	We denote the feasible set for optimization problem \eqref{eq:generic_optimization_initial} by 
	\begin{equation}
	\label{eq:wi_feasible_set}
	X := \{u \in H^2_N : u(x_i)=f_i~\text{for}~ i=1,\dots,r\}.
	\end{equation}
	We denote the tangent space to the feasible set by
	\begin{equation}
	\label{eq:wi_tangent_space}
	X_0 := \{u \in H^2_N : u(x_i)=0~\text{for}~ i=1,\dots,r\}.
	\end{equation}
\end{defn}

\begin{lem}[Elliptic regularity on $X_0$]
	\label{lem:elliptic_regularity_x0}
	If $\partial \Omega$ is $C^{1,1}$, then there exists a constant $C > 0$ such that
	\begin{equation*}
	||u||_{H^2(\Omega)} \le C ||\Delta u||_{L^2(\Omega)}
	\end{equation*}
	for all $u \in X_0$.
\end{lem}

\begin{proof}
	Let $\widetilde{u}$ be version of $u$ that has been shifted up or down by a constant such that $\int_\Omega \overline{u} = 0$. We have
	\begin{align*}
	||D^1 u||_{\left(L^2(\Omega)\right)^d}^2 + ||D^2 u||_{\left(L^2(\Omega)\right)^{d \times d}}^2 &= ||D^1 \widetilde{u}||_{\left(L^2(\Omega)\right)^d}^2 + ||D^2 \widetilde{u}||_{\left(L^2(\Omega)\right)^{d \times d}}^2\\
	&\le ||\widetilde{u}||_{H^2(\Omega)}^2 \\
	&\le C_1 ||\Delta \widetilde{u}||_{L^2(\Omega)}^2 \\
	&= C_1 ||\Delta u||_{L^2(\Omega)}^2
	\end{align*}
	for some constant $C_1$. In the first and fourth lines we used the fact that the derivative of a constant is zero, in the second line we used the definition of the $H^2$ norm, and in the third line we used the standard elliptic regularity result for pure Neumann problems (see, e.g., Theorem 8.13 (b) in \cite{Arbogast}).

	To complete the proof, we will now show that there exists a constant $C_2$ such that $||u||_{L^2(\Omega)}^2 \le C_2 ||\Delta u||_{L^2(\Omega)}^2$ for all $u \in X_0$. For this, we use a standard compactness argument, which closely follows the proof of Theorem 8.12 in \cite{ARBOGAST}. Suppose for the sake of contradiction that this result does not hold. Then there exists a sequence $\{u_k\}_{k=1}^\infty \subset X_0$ such that
	$||u_k||_{L^2(\Omega)}=1$ and $||\Delta u_k||_{L^2(\Omega)} \le 1/k$.
	By the result already proven, the sequence $u_k$ is bounded in $H^2(\Omega)$ as follows:
	\begin{align*}
	||u_k||_{H^2(\Omega)}^2 &= ||u_k||_{L^2(\Omega)}^2 + ||D^1 u||_{\left(L^2(\Omega)\right)^d}^2 + ||D^2 u||_{\left(L^2(\Omega)\right)^{d \times d}}^2 \\
	&\le ||u_k||_{L^2(\Omega)}^2 + C_1 ||\Delta u||_{L^2(\Omega)}^2 \le 1 + C_1.
	\end{align*}
	Therefore, by the Banach-Alaoglu theorem, there exists a subsequence $u_{k_j}$ and function $u \in X_0$ such that
	\begin{equation*}
	u_{k_j} \overset{w}{\rightharpoonup} u \quad \text{weakly in }H^2(\Omega).
	\end{equation*}
	Let $f \in L^2(\Omega)$. By the Cauchy-Schwarz inequality, the linear functional $v \mapsto \int_\Omega f(x) \Delta v(x) dx$ is bounded in $H^2(\Omega)$, and is therefore an element of $\left(H^2(\Omega)\right)^*$. Weak convergence therefore implies
	\begin{equation*}
	\int_\Omega f(x) \Delta u_k(x) dx \rightarrow \int_\Omega f(x) \Delta u(x) dx.
	\end{equation*}
	By the assumption $||\Delta u_k||_{L^2(\Omega)} \le 1/k$ and Cauchy-Schwarz, we also have 
	\begin{equation*}
	\int_\Omega f(x) \Delta u_k(x) dx \rightarrow 0,
	\end{equation*}
	so we must have
	\begin{equation*}
	\int_\Omega f(x) \Delta u(x) dx = 0.
	\end{equation*}
	Since this holds for all $f \in L^2(\Omega)$, by the fundamental lemma of the calculus of variations, $\Delta u(x) = 0$ almost everywhere in $\Omega$. This, combined with the fact that $u$ has Neumann zero boundary conditions, implies that $u$ is a constant function. Since $u$ has value zero at least one point in each connected component of $\Omega$, we have $u=0$. Note that pointwise values of $u$ are well-defined since the Sobolev embedding theorem imples that $H^2(\Omega)$ is continuously imbedded in the space of continuous functions $C^0(\Omega)$ when $d < 4$, and we are considering $d \in \{1,2,3\}$ (see, e.g., Theorem 7.19 in \cite{Arbogast}).

	On the other hand, by the Rellich-Kondrachov theorem, $H^2(\Omega)$ is compactly imbedded in $L^2(\Omega)$, so there must exist a sub-sub-sequence, $u_{k_{j_i}}$ such that 
	\begin{equation*}
	u_{k_{j_i}} \rightarrow u
	\end{equation*}
	strongly in $L^2(\Omega)$. By assumption $||u_{k_{j_i}}||_{L^2(\Omega)}=1$, so $||u||_{L^2(\Omega)}=1$. This contradicts the result $u=0$ that we previously derived, so we must conclude that there exists a constant $C_2 > 0$ such that
	\begin{equation*}
	||u||_{L^2(\Omega)} \le C_2 ||\Delta u||_{L^2(\Omega)}.
	\end{equation*}
	for all $u \in X_0$. Combining this with out previous result, we have
	\begin{equation*}
	||u||^2_{H^2(\Omega)} \le (C_1 + C_2)||\Delta u||^2
	\end{equation*}
	for all $u \in X_0$, which completes the proof.
\end{proof}

\begin{lem}[Well-posedness of weighting function optimization problem]
	\label{lem:wi_existence_and_uniqueness}
	There exists a unique solution to optimization problem \eqref{eq:generic_optimization_initial}.
\end{lem}

\begin{proof}
	By the Sobolev Imbedding Theorem (see, e.g., Theorem 7.20 in \cite{ARBOGAST}), $H^2_N(\Omega)$ is continuously imbedded in $C^0(\Omega)$, so pointwise evaluation $u \mapsto u(x_i)$ of functions at a point $x_i \in \Omega$ is a continuous linear functional on $H^2_N$. The set 
	\begin{equation}
	\label{eq:inverse_image_of_point}
	\{u \in H^2_N : u(x_i)=f_i\}
	\end{equation}
	is therefore a closed affine set in $H^2_N$, because it is the inverse image of a point under a continuous linear functional. Hence the feasible set $X$, defined in \eqref{eq:wi_feasible_set}, is a closed affine set in $H^2_N$, because it is the intersection of the sets in \eqref{eq:inverse_image_of_point} for $i=1,\dots,r$.

	The feasible set $X$ is convex because it is affine, and it is non-empty because interpolating the data pairs $\{(x_i, f_i)\}_{i=1}^r$ using any smooth interpolation method yields an element of $X$. Furthermore, the objective function, 
	\begin{equation}
	J(\gamma) := \frac{1}{2}||\Delta \gamma||^2_{L^2(\Omega)},
	\end{equation}
	is convex because it is the composition of a convex function with a linear map. Furthermore, the elliptic regularity result in Lemma \ref{lem:elliptic_regularity_x0} implies
	\begin{equation*}
	J''(\gamma)v^2 = ||\Delta v||^2_{L^2(\Omega)} \ge C ||v||_{H^2(\Omega)}^2
	\end{equation*}
	for all $v \in X_0$, which implies that $J$ is strongly convex on $X$.
\end{proof}

\begin{defn}
	\label{def:stiffness_mass}
	We define $\mathcal{K}:H^2(\Omega)\rightarrow L^2(\Omega)'$ and $\mathcal{M}:L^2(\Omega) \rightarrow L^2(\Omega)'$ to be the following linear operators:
	\begin{align*}
	\left(\mathcal{K}u\right)(v) :=& \int_\Omega v \Delta u ~dx \\
	\left(\mathcal{M}u\right)(v) :=& \int_\Omega u v ~dx,
	\end{align*}
	and $\mathcal{P}:H^2(\Omega)\rightarrow \mathbb{R}^r$ to be the following pointwise observation operator:
	\begin{equation*}
	(\mathcal{P}u)_i = u(x_i)
	\end{equation*}
	for $i=1,\dots,r$.
\end{defn}

\begin{prop}
	\label{prop:solution_operator_is_linear}
	Let $f \in \mathbb{R}^r$ be arbitrary. The optimization problem
	\begin{equation}
	\label{eq:generic_optimization_initial_old}
	\begin{aligned}
	\min_{u \in H^2(\Omega)} &\quad \frac{1}{2}\norm{\Delta u}_{L^2(\Omega)}^2 \\
	\text{such that} &\quad u(x_i) = f_i, \quad i=1,\dots,r
	\end{aligned}
	\end{equation}
	may be rewritten as follows:
	\begin{equation}
	\label{eq:generic_optimization}
	\begin{aligned}
	\min_{u \in H^2(\Omega)} &\quad \frac{1}{2} \left(\mathcal{K}^T\mathcal{M}^{-1}\mathcal{K}u\right)(u) \\
	\text{such that} &\quad 
	\mathcal{P} u = f.
	\end{aligned}
	\end{equation}
\end{prop}

\begin{prop}
	There exists a unique solution to optimization problem \eqref{eq:generic_optimization}. The solution to \eqref{eq:generic_optimization}, $u^*$, and the adjoint variable for enforcing the constraint, $\lambda^*$, solve the following Karun-Kush-Tucker (KKT) system:
	\begin{equation}
	\label{eq:weighting_kkt}
	\begin{bmatrix}
	\mathcal{K}^T\mathcal{M}^{-1}\mathcal{K} & \mathcal{P}^T \\
	\mathcal{P} & 0
	\end{bmatrix}
	\begin{bmatrix}
	u^* \\ \lambda^*
	\end{bmatrix}
	=
	\begin{bmatrix}
	0 \\ f
	\end{bmatrix},
	\end{equation}
	where we use block-matrix notation to denote block-operators.
\end{prop}

\begin{cor}
	\label{cor:linearity_of_solution_operator}
	Let $\mathcal{T}:\mathbb{R}^r \rightarrow H^2(\Omega)$ denote the solution operator for optimization problem \eqref{eq:generic_optimization}	as a function of $f$. That is, 
	\begin{equation*}
	\mathcal{T}(f) := u^*,
	\end{equation*}
	where $u^*$ is the minimizer of \eqref{eq:generic_optimization}. The operator $\mathcal{T}$ is a linear operator.
\end{cor}

\footnote{The reason for this dependence on spatial dimension is related to the fact that the fundamental solution of the biharmonic equation is continuous in $3$ or less dimensions, but singular in $4$ or more dimensions. In the unusual scenario where one is inverting for a parameter field in $4$ or more spatial dimensions, one could use a modified Poisson interpolation where $\Delta$ is replaced with $\Delta^\alpha$ in the objective function of \eqref{eq:wi_optimization_problem}, where $\alpha>1$ is chosen so that the fundamental solution of $\Delta^{2\alpha}$ is continuous.}

The \emph{point spread function} (PSF) of $\mathcal{A}$ at $x\in \Omega$ is
\begin{equation*}
\varphi_x(y) := \langle \mathcal{A},\delta_x\rangle^*(y+x) = A(y+x,x).
\end{equation*}
Here $\langle \mathcal{A}, \delta_x \rangle \in L^2(\Omega)'$ is the result of applying $\mathcal{A}$ to a point source $\delta_x$ (delta distribution) centered at $x$, and $\langle\mathcal{A},\delta_x\rangle^* \in L^2(\Omega)$ is the Riesz representation of $\langle\mathcal{A},\delta_x\rangle$. The PSF, $\varphi_x$, is the result of translating $\langle\mathcal{A},\delta_x\rangle^*$ to re-center it at zero instead of $x$. Recall the Riesz representation of a linear functional $\psi \in L^2(\Omega)'$ is the unique vector $\psi^* \in L^2(\Omega)$ satisfying $\left(\psi^*, v\right)_{L^2(\Omega)} = \psi(v)$ for all $v \in \L^2(\Omega)$.

The approximation $\widetilde{A}(y,x)$ may not be accurate if both $x$ and $y$ are near the boundary. Other approximation methods, if available, should be considered for the block of the operator associated with boundary-boundary interactions. Nevertheless, in our numerical results we see that approximation works well for the the inverse problem considered, without any modifications to the boundary-boundary block.

The solution to this system of equations is found by the following process:
\begin{enumerate}
	\item Compute 
	\begin{equation*}
	\Psi := K^+ E^T
	\end{equation*}
	by solving Neumann poisson problems for point source right hand sides, with point sources located at the points $x_i$, $i=1,\dots,q$.
	\item Compute 
	\begin{equation*}
	\Theta := K^+ M \Psi
	\end{equation*}
	by solving Neumann Poisson problems with the columns of $\Psi$ as distributed sources.
	\item Form 
	\begin{equation*}
	S = \Psi^T M \Psi
	\end{equation*}.
	\item Compute \begin{equation*}
	\alpha = \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{y}.
	\end{equation*}
	\item Compute 
	\begin{equation*}
	\boldsymbol{\lambda} = -S^{-1} \left(\mathbf{y} - \alpha \mathbf{1} \right).
	\end{equation*}
	\item Compute
	\begin{equation*}
	\mathbf{u} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c}
	\end{equation*}
\end{enumerate}

Consider the Pine island glacier geometry shown in Figure REF. Regions of the glacier are separated by thin inlets of ocean. With radial basis function interpolation, points on one side of the inlet inappropriately influence the interpolation on the other side of the inlet. With Poisson interpolation, the PSF samples only influence the interpolation on their own side of the inlet.

\subsection{Computation of Poisson weighting functions}
\label{sec:computation_of_weighting_functions}

After discretization, optimization problem \eqref{eq:wi_optimization_problem} becomes
\begin{equation}
\label{eq:discrete_poisson_optimization_1}
\begin{aligned}
\min_{\mathbf{w}} &\quad \frac{1}{2}\mathbf{w}^T K^T M^{-1} K \mathbf{w} \\
\text{such that} &\quad E \mathbf{w} = \mathbf{d}.
\end{aligned}
\end{equation}
Here $\mathbf{w}$ is the discretization of $w_i$, $K$ is the stiffness matrix for the discretized Poisson problem with pure Neumann boundary conditions, $M$ is the mass matrix, and $\mathbf{d}=(0,\dots,0,1,0,\dots,0)$ is the length-$q$ vector with $i^\text{th}$ entry one and all other entries zero. Furthermore, $E$ is the pointwise observation matrix defined by $\left(E\mathbf{u}\right)_j = u(x_j)$, where $u$ denotes the function represented by the vector $\mathbf{u}$. Here we dropped the subscripts $i$ from variables for notational convenience. One must solve $q$ optimization problems of the form \eqref{eq:discrete_poisson_optimization_1}; one for each $i=1,\dots,q$.

In the remainder of this section we will derive an efficient algorithm for solving \eqref{eq:discrete_poisson_optimization_1}. The algorithm (Algorithm REF) is based on explicit formula for the solution of \eqref{eq:discrete_poisson_optimization_1} which we present in Proposition \ref{prop:poisson_interpolation_formula}. The formulas in Proposition \ref{prop:poisson_interpolation_formula} rely on certain matrices that we define in Definition \ref{defn:poisson_interpolation_matrices}. The primary computational cost of the algorithm is the solution of Poisson problems with pure Neumann boundary conditions.

\begin{defn}
	\label{defn:poisson_interpolation_matrices}
	Let 
	\begin{align*}
	\Psi :=& K^+ E^T, \\
	\Theta :=& K^+ M \Psi \\
	S :=& \Psi^T M \Psi,
	\end{align*}
	where $K^+$ is the Moore-Penrose pseudoinverse of $K$. Further, let $\mathbf{c}$ dente the discretization of the constant function $c(x)=1$ on $\Omega$, and let $\mathbf{1}:=(1,\dots,1)$ denote the vector of length $q$ with all entries equal to one.
\end{defn}

\begin{remark}
	Computing $\mathbf{g} = K^+\mathbf{f}$ is equivalent to solving the discretized Poisson problem with $f$ as the right hand side source, with pure Neumann boundary conditions, and with the constraint that the solution has average value equal to zero. That is,
	\begin{equation}
	\label{eq:pure_neumann_poisson}
	\mathbf{g} = K^+ \mathbf{f} \quad \Leftrightarrow \quad 
	\begin{cases}
	\Delta g = f, & \text{in }\Omega,\\
	\nu \cdot \nabla g = 0, & \text{on }\partial \Omega,\\
	\int_\Omega g(x) dx = 0,
	\end{cases}
	\end{equation}
	where $\nu$ denotes the normal to the boundary.

	Hence, the $i^\text{th}$ column of $\Psi$ is the result of solving the discretized pure Neumann Poisson problem, with a point source at location $x_i$. The $i^\text{th}$ column of $\Theta$ is the result of solving the discretized pure Neumann Poisson problem with distributed source given by the $i^\text{th}$ column of $\Psi$.
\end{remark}

\begin{prop}
	\label{prop:poisson_interpolation_formula}
	The solution to optimization problem \eqref{eq:discrete_poisson_optimization_1} is given by
	\begin{equation*}
	\mathbf{w} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c},
	\end{equation*}
	where
	\begin{align*}
	\alpha =& \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{d} \\
	\boldsymbol{\lambda} =& -S^{-1} \left(\mathbf{d} - \alpha \mathbf{1} \right). \\
	\end{align*}
\end{prop}

\begin{proof}
	Let
	\begin{equation*}
	v := \mathcal{K} w_i,
	\end{equation*}
	and for the remainder of this proof let $e := e_i$.
	Since the null-space of the Laplacian with pure Neumann boundary conditions is the set of all constant functions, we have
	\begin{equation}
	\label{eq:definition_of_w}
	w_i = \mathcal{K}^+ v + \alpha C
	\end{equation}
	for some scalar $\alpha$.

	Writing optimization problem \eqref{eq:wi_optimization_problem} in terms of $v$ and $\alpha$ instead of $w_i$ yields the following equivalent optimization problem:
	\begin{equation}
	\label{eq:discrete_poisson_optimization_2}
	\begin{aligned}
	\min_{v \in L^2(\Omega)', \alpha \in \mathbb{R}} &\quad \frac{1}{2} v\left(\mathcal{M}^{-1} v\right) \\
	\text{such that} &\quad \mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} = e.
	\end{aligned}
	\end{equation}
	The Lagrangian for optimization problem \eqref{eq:discrete_poisson_optimization_2} is
	\begin{equation*}
	\mathcal{L} = \frac{1}{2}v\left( \mathcal{M}^{-1} v\right) + \lambda^T\left(\mathcal{P}\mathcal{K}^+ v + \alpha \mathbf{1} - e\right).
	\end{equation*}
	The solution to this optimization problem is the stationary point of the Lagrangian, i.e., the point at which the gradient of the Lagrangian is zero:
	\begin{equation*}
	0 = \nabla \mathcal{L} = 
	\begin{bmatrix}
	\nabla_{v} \mathcal{L} \\
	\nabla_{\lambda} \mathcal{L} \\
	\nabla_\alpha \mathcal{L}
	\end{bmatrix} 
	= \begin{bmatrix}
	\mathcal{M}^{-1} v + \left(\mathcal{K}^+\right)^T \mathcal{P}^T \lambda \\
	\mathcal{P}\mathcal{K}^+v + \alpha \mathbf{1} - e \\
	\mathbf{1}^T \lambda
	\end{bmatrix}.
	\end{equation*}
	This equation may be rewritten as the following block $3 \times 3$ linear system:
	\begin{equation*}
	\begin{bmatrix}
	\mathcal{M}^{-1} & \left(\mathcal{K}^+\right)^T \mathcal{P}^T & 0 \\
	\mathcal{P}\mathcal{K}^+ & 0 & \mathbf{1} \\
	0 & \mathbf{1}^T & 0
	\end{bmatrix}
	\begin{bmatrix}
	v \\ \lambda \\ \alpha
	\end{bmatrix}
	=
	\begin{bmatrix}
	0 \\ e \\ 0
	\end{bmatrix}.
	\end{equation*}
	Performing block Gaussian elimination on this system allows us to reduce the system to the following block triangular form
	\begin{equation*}
	\begin{bmatrix}
	\mathcal{M}^{-1} & \left(\mathcal{K}^+\right)^T \mathcal{P}^T & 0 \\
	& -S & \mathbf{1} \\
	0 & 0 & \mathbf{1}^T S^{-1} \mathbf{1}
	\end{bmatrix}
	\begin{bmatrix}
	v \\ \lambda \\ \alpha
	\end{bmatrix}
	=
	\begin{bmatrix}
	0 \\ e \\ \mathbf{1}^T S^{-1} e
	\end{bmatrix}
	\end{equation*}
	where we recall that $S := \Psi^T \mathcal{M} \Psi = \mathcal{P}\mathcal{K}^+ \mathcal{M} \left(\mathcal{K}^+\right)^T \mathcal{P}^T$. From this block triangular system, we read off the solution as:
	\begin{align*}
	\alpha =& \left(\mathbf{1}^T S^{-1} \mathbf{1}\right)^{-1}\mathbf{1}^T S^{-1} \mathbf{d} \\
	\boldsymbol{\lambda} =& -S^{-1} \left(\mathbf{d} - \alpha \mathbf{1} \right) \\
	\mathbf{v} =& -M K^+ E^T \boldsymbol{\lambda}
	\end{align*}
	Substituting these results into the definition of $\mathbf{w}$ in \eqref{eq:definition_of_w}, we have
	\begin{equation*}
	\mathbf{w} = -K^+ M K^+ E^T \boldsymbol{\lambda} + \alpha \mathbf{c} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c},
	\end{equation*}
	as required.
\end{proof}

We now design an incremental algorithm for constructing interpolants $\mathbf{u}$, in which we can add points $x_i$ more efficiently. Let
\begin{align*}
\Psi :=& K^+ E^T \\
\Theta :=& K^+ M \Psi
\end{align*}
so that 
\begin{equation*}
S = \Psi^T M \Psi
\end{equation*}
and
\begin{equation*}
\mathbf{u} = -\Theta \boldsymbol{\lambda} + \alpha \mathbf{c}
\end{equation*}
The columns of $\Psi$ are the responses of $A^+$ to point sources (delta distributions) centered at the sample points $x_i$. The columns of $\Theta$ are the responses of $A^+M^{-1} A$ to these point sources. The matrices $\Psi$ and $\Theta$ may be constructed incrementally as new points are added.

\section*{Acknowledgments}
We thank J.J. Alger, Longfei Gao, and Rami Nammour for helpful discussions.

\bibliographystyle{siamplain}
\bibliography{references}
\end{document}

### 2024-02-24 presentations/siamuq24-noemi/talknoemisiamuq24.tex
noemis slides

% Tikz plot for UQ. This is based on the tikz code of
% Fabian Schuh for the tikz examples website
\documentclass[10pt,final,xcolor=dvipsnames]{beamer}
%\usepackage{etex}

\usetheme{Madrid}

\usepackage{xspace}
%\usepackage{enumitem}
\usepackage[mathscr]{euscript}
\usepackage{subfigure}
\hypersetup{colorlinks,linkcolor=,urlcolor=blue}

% \usetheme{Boadilla}
% \usecolortheme{seahorse}

% the macros, packages, etc:
\usepackage[absolute,overlay]{textpos}
\setbeamertemplate{items}[ball]
\setbeamertemplate{blocks}[rounded][shadow=true]
\setbeamertemplate{navigation symbols}{}
\newenvironment{reference}[2]{%
  \begin{textblock*}{\textwidth}(#1,#2)
  \scriptsize{\it\color{black}}}{\end{textblock*}}
\setbeamertemplate{blocks}[rounded]% [shadow=false]
\newenvironment{boxalertenv}{\begin{altenv}%
      {\usebeamertemplate{alerted text begin}\usebeamercolor[fg]{alerted text}\usebeamerfont{alerted text}\colorbox{bg}}
      {\usebeamertemplate{alerted text end}}{\color{.}}{}}{\end{altenv}}

\usepackage{graphicx}
\usepackage{multirow,color}
\usepackage{bbm}
\usepackage{amsfonts,amsmath,amssymb,amsbsy,amsthm}
\usepackage{movie15}
\definecolor{themec}{RGB}{51,108,121}

%\usepackage{movie15}
%\usepackage{multimedia}
% custom commands (add your own macros here)

\definecolor{bronze}{rgb}{0.8, 0.5, 0.2}
\definecolor{goldenyellow}{rgb}{1.0, 0.87, 0.0}
\definecolor{amber}{rgb}{1.0, 0.75, 0.0}
\definecolor{azure}{rgb}{0.94, 1.0, 1.0}
\definecolor{bittersweet}{rgb}{1.0, 0.44, 0.37}
\definecolor{brass}{rgb}{0.71, 0.65, 0.26}
\definecolor{camel}{rgb}{0.76, 0.6, 0.42}
\setbeamercolor{block body}{bg=camel!30}
\usefonttheme[onlymath]{serif}
\usepackage{color}

\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{amsfonts,amsmath,amssymb,amsbsy,amsthm}
%\usecolortheme{tango}

%\usefonttheme{serif}     % Font theme: serif
%\usefonttheme[stillsansseriftext]{serif} 
%\renewcommand*\familydefault{\sfdefault}

\usepackage{pgf}
%\usepackage{pgfplots}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{verbatim}
%\usepackage{movie15}
\usepackage{tikz,pgfplots}

%\usepackage[active,tightpage]{preview}
%\PreviewEnvironment{tikzpicture}
\usetikzlibrary{arrows,automata}
\usetikzlibrary{positioning}
\usetikzlibrary{fit}

\input{defjcp14.tex}
\newcommand{\iparb}{\bar{\ipar}}
\newcommand{\euclidnorm}[1]{\left\| {#1} \right\|_2}
\newcommand{\incu}{\upupsilon}
\newcommand{\incp}{\uprho}
\newcommand{\QoIquad}{\QoI_\text{quad}}
\newcommand{\tcblue}[1]{\textcolor{blue}{#1}}
\newcommand{\betad}{\bs{\beta}}

\newcommand{\preconditioner}{\widetilde{H}}
\newcommand{\impulseresponse}{\phi}
\newcommand{\spatialmean}{\mu}
\newcommand{\Aop}{\mathcal{A}}

\tikzset{
    state/.style={
           rectangle,
           rounded corners,
           draw=black, very thick,
           minimum height=3em,
           inner sep=1pt,
           text centered,
           },
}

\newcommand{\Aker}{\Phi}
%\DeclareMathOperator{\diag}{diag}
%\newcommand{\Hmatrix}{\bs{H}}
\newcommand{\Hm}{\mathcal{H}}

% argument #1: any options
\newenvironment{customlegend}[1][]{%
    \begingroup
    % inits/clears the lists (which might be populated from previous
    % axes):
    \csname pgfplots@init@cleared@structures\endcsname
    \pgfplotsset{#1}%
}{%
    % draws the legend:
    \csname pgfplots@createlegend\endcsname
    \endgroup
}
\def\addlegendimage{\csname pgfplots@addlegendimage\endcsname}

% title, authors, etc:
\title[Bayesian ice sheet inverse
  problems]{Exploiting Low-Dimensional Structure in Bayesian Inverse
  Problems Governed by Ice Sheet Flow Models\thanks{Research supported
    partially by NSF-DMS-CAREER-1654311 and NSF-OAC-1550547.}}
%\author[{Noemi
%    Petra}]{No\'{e}mi~Petra\inst{1}}

\author[Noemi Petra]{{No\'{e}mi~Petra},\inst{1}\\[2ex]
  {\small \textcolor{themec}{Joint work with:}}
  \\
  {\small Nick Alger},\inst{2}
  {\small Tucker Hartland},\inst{3}
  {\small Omar Ghattas}\inst{2}
}

\institute[UC Merced]{%
  {Department of Applied Mathematics, University of California, Merced}\\\smallskip
  \inst{2}{Oden Institute for Computational Engineering and Sciences, The University of Texas at Austin}\\\smallskip
  \inst{3}{Center for Applied Scientific Computing (CASC), Lawrence Livermore National Laboratory}
}

\date[February 27, 2024]
     {\footnotesize SIAM Conference on Uncertainty Quantification (UQ24) \\
       Tuesday, February 27}

 \begin{document}

 % -------------------------------------------------------------------
%%\AtBeginSubsection[]  % "Beamer, do the following at the start of every section"
%\AtBeginSection[]  % "Beamer, do the following at the start of every section"
%{
%  \begin{frame}<beamer>
%    \frametitle{Outline} % make a frame titled "Outline"
%    \small
%  %\tableofcontents[currentsubsection]  % show TOC and highlight current section
%  \tableofcontents[currentsection]  % show TOC and highlight current section
%\end{frame}
%}
%----------- titlepage ----------------------------------------------%
\begin{frame}[plain]
  \titlepage
\end{frame}
%% ----------------------------------------------------------------------------
%\begin{frame}
%  \frametitle{Acknowledgements}
%
%  {\bf Various parts of this research is joint work with:}
%  \vspace{0.02in}
%  \begin{itemize}
%  \item Omar Ghattas (UT Austin)
%  \item Toby Isaac (Georgia Tech)
%  \item Georg Stadler (NYU)
%  \item Umberto Villa (Washington University, St. Louis)
%  \item Hongyu (Alice) Zhu (United Technologies)
%  \end{itemize}
%
%  \begin{itemize}
%  \item Olalekan (Lekan) Babaniyi (Rochester Institute of Technology)
%  \item Tucker Hartland and Radoslav Vuchkov (UC Merced) 
%  \item Ruanui (Ru) Nicholson (University of Auckland)
%  \end{itemize}
%
%  \vspace{0.05in}
%  {\bf Funding:}
%  \vspace{0.02in}
%  \begin{itemize}
%  \item Partial funding for this research was provided by the
%    U.S. National Science Foundation, Software Infrastructure for
%    Sustained Innovation (SI2: SSE \& SSI) Program under grants
%    ACI-1550593, and ACI-1550547 and by the NSF Division of
%    Mathematical Sciences under the CAREER grant 1654311.
%  \end{itemize}
%\end{frame}
%% ----------------------------------------------------------------------------
\begin{frame}
  \frametitle{Outline}

  \begin{itemize}
  \item Motivation
    \vspace{0.02in}
    \begin{itemize}
    \item Why we need predictive models with quantified uncertainties
      to accurately anticipate future sea level rise?
      \vspace{0.02in}
    \item Why we need more efficient approximations of high rank
      Hessians?
    \end{itemize}
    \vspace{0.02in}
  \item Bayesian inverse problems governed by PDEs (preliminaries)
    \vspace{0.02in}
  \item Exploiting low-dimensional structure in Bayesian inverse problems governed by PDEs
    \vspace{0.02in}
    \begin{itemize}
    \item (Global) low rank approximation
      \vspace{0.02in}
    \item Hierarchical off-diagonal low rank (HODLR) approximation
      \vspace{0.02in}
    \item Point spread function approximation combined with
      hierarchical matrix (H-matrix) approximation
    \end{itemize}
    \vspace{0.02in}
  \item Example: Inversion for the basal friction
    coefficient in an ice sheet flow problem
    \vspace{0.02in}
  \item Conclusions
  \end{itemize}

\end{frame}
%-------------------------------------------------------------------
\begin{frame}
  \frametitle{Need predictive models with quantified uncertainties to accurately anticipate future sea
   level rise.}
\framesubtitle{Sea level projections from the IPCC 6th Assessment Report, 2021}
 \vspace{0.2in}
  \begin{columns}
    \begin{column}{0.7\paperwidth}
       \centering\includegraphics[width=0.85\columnwidth]{extraplots/venezia}
    \end{column}
    \hspace{-0.4in}
    \begin{column}{0.3\paperwidth}
    \vspace{0.1in}
      \centering\includegraphics[width=0.85\columnwidth]{extraplots/ipcc-ar6-wgi-2021.png}
    \end{column}
  \end{columns}

  %\centering\includegraphics[width=0.55\columnwidth]{extraplots/nasa-sea-level-projection-tool.png}

 \vspace{0.1in}
%\begin{block}{IPCC Fourth Assessment Report: Climate Change 2007}
%  {\small
%    {\em [T]he uncertainty in the projections of the land ice
%      contributions [to sea level rise] is dominated by the various
%      uncertainties in the land ice models themselves
%      ... rather than in the temperature projections.''} %[10.A.6]
%  }
 %\end{block}
 \begin{itemize}
 \item Several port cities will be at risk from coastal
   flooding in the future.
 \item Ice flowing from ice sheets to ocean is primary contributor
   to sea level rise.
 \end{itemize}

 \begin{itemize}
  \item [] \scriptsize{Details in: Masson-Delmotte, V. et al. ``IPCC,
    2021: Summary for Policymakers. In: Climate Change 2021: The
    Physical Science Basis. Contribution of Working Group I to the
    Sixth Assessment Report of the Intergovernmental Panel on Climate
    Change'', Cambridge University Press. In Press.}
  \end{itemize}

\end{frame}
%-----------
%--------------------------------------------------------
\begin{frame}
  %\frametitle{Motivations: Infer basal sliding in an ice sheet model and characterize the associated uncertainties}
  \frametitle{Challenges and the need to exploit problem structure}

  Severe mathematical and computational challenges place significant
  barriers on improving predictability of ice sheet flow models, e.g.,

  \begin{minipage}{.6\columnwidth}
    \begin{itemize}
    \item complex and very high-aspect ratio (thin) geometry,
    \item highly nonlinear and anisotropic rheology,
    \item extremely ill-conditioned and large-scale linear and
      nonlinear algebraic systems that arise upon discretization,
    \item uncertain basal sliding parameter, basal topography,
      geothermal heat flux, and rheology,
    \item modeling error, etc.
    \end{itemize}

  \end{minipage}\hfill
  \begin{minipage}{.4\columnwidth}
    \begin{figure}\centering
      \includegraphics[width=0.99\columnwidth]{extraplots/antarctica_uobs_rotsurface_larger}\\
      \vspace{0.2in}
      \centering\includegraphics[width=0.65\columnwidth]{extraplots/spec_ppmisfit_hess_coarseandfine_new.pdf}
    \end{figure}
  \end{minipage}

  \vspace{0.1in}
  \begin{itemize}
  \item [] \scriptsize{Details in: T. Isaac, N. Petra, G. Stadler, and
    O. Ghattas. ``Scalable and efficient algorithms for the
    propagation of uncertainty from data through inference to
    prediction for large-scale problems, with application to flow of
    the Antarctic ice sheet'', Journal of Computational Physics, 296,
    348-368 (2015). {\it Selected as the 2019 SIAM Activity Group on
    Computational Science and Engineering Best Paper.}}
  \end{itemize}

  %``To address the challenges of learning models from data in the
  %large-scale setting of partial differential equations, we must
  %exploit the low-dimensional structure of the map from inputs to
  %outputs.'' (O. Ghattas and K. Willcox, ``Learning physics-based
  %models from data: perspectives from inverse problems and model
  %reduction'', Acta Numerica (2021).)

\end{frame}
%-------------------------------------------------------------------
\begin{frame}
  %\frametitle{Motivations: Infer basal sliding in an ice sheet model and characterize the associated uncertainties}
  \frametitle{How can we address (some of) these challenges?}

  %To address some of these challenges, we
  %\vspace{0.05in}
  \begin{minipage}{.6\columnwidth}
    \begin{itemize}
    \item Learn models from data, i.e., infer unknown/uncertain
      parameters from available data, e.g., satellite measurements of
      surface ice flow velocity ({\bf statistical inverse problems
        governed by PDEs})
    \item Apply/adapt/design fast, mesh-independent, structure
      exploiting, inner-product- and additional uncertainty-aware
      methods ({\bf scalable, robust and efficient algorithms})
    \item Collect data in an optimal way in order to minimize the
      uncertainty in the inferred parameters or in some predictive
      quantity of interest ({\bf optimal experimental design}).
    \end{itemize}
  \end{minipage}\hfill
  \begin{minipage}{.4\columnwidth}
    \begin{figure}\centering
      \includegraphics[width=0.99\columnwidth]{extraplots/antarctica_uobs_rotsurface_larger}\\
      \vspace{0.2in}
      \centering\includegraphics[width=0.65\columnwidth]{extraplots/spec_ppmisfit_hess_coarseandfine_new.pdf}
    \end{figure}
  \end{minipage}

  \vspace{0.15in}
  \begin{itemize}
  \item [] \scriptsize{Details in: O. Ghattas and
    K. Willcox. ``Learning physics-based models from data:
    perspectives from inverse problems and model reduction'', Acta
    Numerica, Cambridge University Press (2021).}
  \end{itemize}

  %``To address the challenges of learning models from data in the
  %large-scale setting of partial differential equations, we must
  %exploit the low-dimensional structure of the map from inputs to
  %outputs.'' (O. Ghattas and K. Willcox, ``Learning physics-based
  %models from data: perspectives from inverse problems and model
  %reduction'', Acta Numerica (2021).)

\end{frame}
%% -------------------------------------------------------------------
%\begin{frame}
%  \frametitle{The forward problem}
%  \framesubtitle{Nonlinear Stokes ice sheet model (for viscous, shear-thinning, incompressible fluid)}
%
%  Invoking the balance of mass and linear momentum:
%  %  \small
%  \vskip 0.1in
%  \begin{minipage}{.4\columnwidth}
%    \begin{alignat}{2}\nonumber
%      % mass
%      % momentum
%      - \gbf{\nabla} \cdot \left[ 2\gamma(\gbf{u}, n) \, \edot_{\gbf{u}}
%      -\gbf{I}p \right] &= \rho \gbf{g} \ & &\text{ in }
%      \Omega \nonumber\\
%      \gbf{\nabla} \cdot \gbf{u} &= 0 & &\text{ in }
%      \Omega \nonumber\\
%      \gbf{\sigma}_{\gbf{u}} \gbf{n} &= \gbf{0}  & &\text{ on }
%      \Gamma_{\!\scriptsize t}  \nonumber\\
%      \gbf{u}\cdot \gbf{n} = 0, \, \gbf T \gbf{\sigma}_{\gbf{u}} \gbf{n} +
%      {\exp(\beta)}\gbf{T}\gbf{u} &= \gbf{0}
%      & &\text{ on } \Gamma_{\!\scriptsize
%        b}\nonumber
%    \end{alignat}
%%\gbf{u}\cdot \gbf{n} = 0, \, \gbf T \gbf{\sigma}_{\gbf{u}} \gbf{n} +
%%      \tcr{\exp(\beta)} |\gbf{T}\gbf{u}|^{m-1}\gbf{T}\gbf{u} &= \gbf{0}
%%      & &\text{ on } \Gamma_{\!\scriptsize
%%        b}\nonumber
%%%    \hskip 0.75in + \; \text{additional B.C.s}
%  \end{minipage}\hfill
%  \begin{minipage}{.4\columnwidth}
%    \begin{figure}\centering
%      \includegraphics[width=0.99\columnwidth]{../extraplots/ymir/antarctica_uobs_rotsurface_larger}
%    \end{figure}
%  \end{minipage}
%  \\[3.5ex]
%  \begin{minipage}{0.565\columnwidth}
%    \begin{center}
%      \begin{minipage}{\columnwidth}
%	\vfill
%	\begin{itemize}
%	\item $\gbf u$ ice flow velocity, $p$ pressure
%	\item $\gbf{\sigma}_{\gbf{u}} = -\gbf{I}p + 2\gamma(\gbf{u},
%          n)\edot_{\gbf{u}}$ stress tensor
%	\item $\edot_{\gbf{u}} = \frac12 (\gbf{\nabla u} + \gbf{\nabla
%          u}^T)$ strain rate tensor
%        \item $\gamma(\gbf u, {n}) = \frac12 {A}^{-\frac{1}{{n}}} \;
%          \secinve^{\frac{1-{n}}{2{n}}}$ effective viscosity
%        \item $\secinve = \frac12 \mathrm{tr}(\edot_{\gbf{u}}^2)$
%          second invariant of the strain rate tensor
%	\end{itemize}
%        \vfill
%      \end{minipage}
%      %}
%    \end{center}
%  \end{minipage}\hfill
%  \begin{minipage}{0.43\columnwidth}
%    \begin{itemize}
%    \item $\rho$ density, $\gbf g$ gravity
%    \item $\gbf n$ unit normal vector
%    \item ${\beta}$ log basal sliding coefficient
%      %\item $m$ the basal sliding exponent
%    \item $ \gbf{T} = \gbf I - \gbf n \otimes \gbf  n$
%      tangential operator
%    \item $\Gamma_{\!\scriptsize t}$ and $\Gamma_{\!\scriptsize b}$
%      top and base boundaries
%    \end{itemize}
%    \vfill
%  \end{minipage}\\[1ex]
%
%\end{frame}
%
%%-----------------------------------------------------------
\begin{frame}
  \frametitle{The inverse problem}
  %\framesubtitle{\hip Model}

  %\begin{itemize}
  %  \uncover<1-2>{
  %\item \alert{The forward problem}: Given some input parameter field
  %  (coefficient, right hand side, initial condition, boundary
  %  condition, etc.)  $\tcr{\ipar}$ solve
  %  \begin{align*}
  %    r(\tcb{\istate}, \tcr{\ipar}) = 0,
  %  \end{align*}
  %  for $\tcb{\istate}$, where $r: \mathcal{V}\times\iparspace \to
  %  \mathcal{V}^*$ represents the PDE problem with $\mathcal{V}$ a
  %  suitable Hilbert space of functions defined on a bounded domain
  %  $\D \subset \R^d$, and $\iparspace \subseteq L^2(\D)$.
  %  }
  %  \uncover<2-2>{
    % \item \alert{The inverse problem}
  %consists of
  Use available observations/data $\obs$ to infer the
  values of the unknown parameter field $\tcr{\ipar}$ that
  characterize a physical process modeled by PDEs, i.e.,
  %Mathematically
      %this inverse relationship is expressed as
      \begin{align*}
        \obs = \ff(\tcr{\ipar}) + \vec{\eta}.
      \end{align*}
      \vspace{-0.2in}
      \begin{itemize}
      \item The map $\ff: \iparspace \to \mathbb{R}^q$ is the so-called
        {\it parameter-to-observable} map.
      \item Evaluations of $\ff$ involve the solution of the Stokes
        PDE given $\tcr{\ipar}$, followed by the application of an
        observation operator $\mathcal{B}: \mathcal{V} \to
        \mathbb{R}^q$ to extract the observations from the state.
      \item $\vec{\eta}$ accounts for noisy measurements and model
        errors and is modeled as $\vec{\eta} \sim
        \GM{\vec{0}}{\ncov}$, i.e., a centered Gaussian at $\vec{0}$
        with covariance $\ncov$.
      \end{itemize}
   % }
  %\end{itemize}
\end{frame}
%-----------------------------------------------------------
\begin{frame}
  \frametitle{Bayesian formulation of the inverse problem}

   Describes probability of all models that are consistent with the
   observations/data and any prior knowledge about the parameters:
    %\begin{block}
    \begin{equation*}
      d\mupost \propto \exp\Big\{-\half \| \ff(\tcr{\ipar}) -
      \vec{d} \|^2_{\ncov^{-1}} -\half\| \tcr{\ipar} - \iparpr\|^2_{\Cprior^{-1}} \Big\}.
    \end{equation*}
    %\end{block}
    \begin{itemize}
    \item The first term in the exponential is the negative log-likelihood.
    \item The second term represent the negative log-prior (e.g.,
      Gaussian prior, i.e., $\tcr{\ipar} \sim \GM{\iparpr}{\Cprior}$).
    \end{itemize}

    \begin{block}{Goal:}
      \begin{itemize}
      \item characterize the posterior statistically (MAP point, mean, covariance, etc.)
      \item for functions $\bs m$ (large vectors after discretization)
      \item for expensive $\ff(\cdot)$
      \item exploit connection to PDE-constrained optimization
      \end{itemize}
    \end{block}
\end{frame}
%--------------------------------------------------------------------
\begin{frame}
  \frametitle{Inverse problems governed by PDEs}
  %\framesubtitle{\hip Model}
  \begin{itemize}
    \uncover<1-2>{
    \item The \emph{maximum a posteriori} (MAP) point $\iparmap$ is
    defined as the parameter field that maximizes the posterior
    distribution: %It can be obtained by solving the following
    %deterministic optimization problem
    \begin{align*}
      \iparmap :&= \argmin_{\tcr{\ipar} \in \iparspace} (- \log
      d\mupost(\tcr{\ipar}) ) \\
      &= \argmin_{\tcr{\ipar} \in \iparspace} \half \|
      \ff(\tcr{\ipar}) - \vec{d} \|^2_{\ncov^{-1}} + \half\| \tcr{\ipar} -
      \iparpr\|^2_{\Cprior^{-1}} \\
      &=\argmin_{\tcr{\ipar} \in \iparspace} \half \|
      \mathcal{B}(\tcb{\istate}) - \vec{d} \|^2_{\ncov^{-1}} + \half\| \tcr{\ipar} -
      \iparpr\|^2_{\Cprior^{-1}},
    \end{align*}
    where $\tcb{\istate}$ solves the forward Stokes (PDE) problem. %$r(\tcb{\istate}, \tcr{\ipar}) = 0$.
    %We note that, the prior information plays the role of Tikhonov
    %regularization in \eqref{eq:optpb}; in fact the deterministic
    %optimization problem \eqref{eq:optpb} is the same as
    %\eqref{eq:map_continuum} for the choice $\R(\ipar) = \half\| m -
    %\iparpr\|^2_{\Cprior^{-1}}$.  The Hessian $\mathcal{H}(\iparmap)$
    %of the negative log-posterior evaluated at $\iparmap$ plays a
    %fundamental role in quantifying the uncertainty in the inferred
    %parameter. In particular, this indicates which directions in the
    %parameter space are most informed by the data
    %\cite{Bui-ThanhGhattasMartinEtAl13}.
    }
    \uncover<2-2>{
    \item When $\iFF$ is
    linear, due to the particular choice of prior and noise model, the
    posterior measure is Gaussian, $\GM{\iparmap}{\Cpost}$
    \begin{equation*}
      \iparmap = \Cpost(\iFFadj\ncov^{-1}\obs + \Cprior^{-1}\iparpr),
      \qquad 
      \Cpost = \mathcal{H}^{-1} = (\iFFadj \ncov^{-1} \iFF + \Cprior^{-1})^{-1},
    \end{equation*}
    where $\iFFadj:\R^q \to \iparspace$ is the adjoint of $\iFF$, and
    $\mathcal{H}$ is the Hessian (second derivative) of the
    negative-log posterior.
    %\vspace{0.15in}
    \item {\bf Note:} In the general case of nonlinear
      parameter-to-observable map $\iFF$ the posterior distribution is
      not Gaussian. %In this case
      %one would need to use sampling to characterize the posterior.
     }
  \end{itemize}
\end{frame}
%%%% --------------------------------------------------------------------
%\begin{frame}
%  \frametitle{Data$\rightarrow$Inference$\rightarrow$Prediction}
%
%  \begin{itemize}
%  \item While one can formulate a {\bf data-to-prediction framework}
%    to {\bf quantify uncertainties from data to inferred model
%      parameters to predictions} with an underlying model of
%    non-Newtonian ice sheet flow, attempting to execute this framework
%    for the Antarctic ice sheet (or other large-scale complex models)
%    is intractable for high-dimensional parameter fields using current
%    algorithms.
%    \vspace{0.05in}
%    
%  \item Yet, quantifying the uncertainties in predictions of ice sheet
%    models is essential if these models are to play a significant role
%    in projections of future sea level.
%    \vspace{0.05in}
%    
%  \item {\bf Goal:} design an integrated framework and efficient,
%    scalable algorithms (under Gaussian approximations of the
%    posterior and prediction) for carrying out this data-to-prediction
%    process.
%    \vspace{0.05in}
%    
%    %Two key ideas are needed to produce such scalable
%    %algorithms. First, we use Gaussian approximations of both the
%    %posterior pdf that results from Bayesian solution of the inverse
%    %problem of inferring ice sheet parameter fields from satellite
%    %observations of surface velocity, as well as the pdf resulting
%    %from propagating the uncertain parameter fields through the
%    %forward ice sheet model to yield predictions of present-day mass
%    %flux into the ocean. This is accomplished by linearizing the
%    %parameter-to-observable map as well as the parameter-to-prediction
%    %map around the maximum a posterior point.
%    
%  \item {\bf Scalable:} the cost-measured in number of PDE solves-is
%    independent of not only the number of processor cores, but
%    importantly the state variable dimension, the parameter dimension,
%    and the data dimension.
%  \end{itemize}
%  
%  \begin{itemize}
%  \item [] \scriptsize{Details in: T. Isaac, N. Petra, G. Stadler, and
%    O. Ghattas. ``Scalable and efficient algorithms for the
%    propagation of uncertainty from data through inference to
%    prediction for large-scale problems, with application to flow of
%    the Antarctic ice sheet'', Journal of Computational Physics, 296,
%    348-368 (2015).}
%  \end{itemize}
%\end{frame}
%
%------------------------
%---------------------------------------------------------------------------------%
\begin{frame}
  \frametitle{The Hessian (of the negative log posterior) plays a
    critical role in inverse problems}
  \begin{itemize}
  \item Its spectral properties characterize the degree of
    ill-posedness.
  \item The Hessian drives Newton-type optimization algorithms for
    solving the inverse problem.
  \item The inverse of the Hessian locally characterizes the uncertainty
    in the solution of the inverse problem (under the Gaussian
    assumption, it is precisely the posterior covariance matrix).
  \item \alert{Goal:} rapidly perform linear algebraic operations, i.e.,
    manipulation of the Hessian (and its square root and inverse)
    actions needed by sampling or CG solvers, hence \alert{seek
      approaches to approximate the Hessian(-applies)}.
  \item These approximations can then be used as
    \alert{pre-conditioners}, and to \alert{build MCMC proposals} based
    on local Gaussian approximations.
  \end{itemize}
\end{frame}
%---------------------------------------------------------------------------------%
\begin{frame}
\frametitle{Scalability of an ice sheet inverse solver}
\framesubtitle{Inexact Newton-CG}

  \only<1>{
\begin{table}
\begin{center}
  \small
  \label{tbl:performance}
  \begin{tabular}{|r|r|r|r|r|r|}
    \hline
        {\bf \#sdof}    &   {\bf \#pdof}   &    {\bf  \#N}   &  {\bf  \#CG}
        & {\bf  avgCG}   &  {\bf \#Stokes} \\
        \hline
        %69101 &   7485 &  36 &  2534 &    70 & 1609 &     6713 \\
        95,796 &  10,371 &  42 &  2718 &    65         &     7031 \\
                                              % 1553 &    
        233,834 &  25,295 &  39 &  2342 &    60        &     6440 \\
                                              % 1717 &
        848,850 &  91,787 &  39 &  2577 &    66        &     6856 \\
                                              % 1663 &
        3,372,707 & 364,649 &  39 &  2211 &    57       &     6193 \\
                                              % 1732 
        %5667771 & 365709 &  43 &  2139 &    50 &     1978 &     6299\\
        22,570,303 & 1,456,225 &  40 &  1923 &    48     &     5376\\
                                              % 1490
        \hline
  \end{tabular}
\end{center}
\end{table}
  }
  \only<2>{
\begin{table}
\begin{center}
  \small
  %\caption{Scalability of inverse solver}
  \label{tbl:performance}
  \begin{tabular}{|r|r|r|r|r|r|}
    \hline
        {\bf \#sdof}    &   {\bf \#pdof}   &    {\bf  \#N}   &  {\bf  \#CG}
        & {\bf  avgCG}   &  {\bf \#Stokes} \\
        \hline
        %69101 &   7485 &  36 &  2534 &    70 & 1609 &     6713 \\
        95,796 &  10,371 &  42 &  2718 &    65         &     {\bf\alert{7031}} \\
                                              % 1553 &    
        233,834 &  25,295 &  39 &  2342 &    60        &     {\bf\alert{6440}} \\
                                              % 1717 &
        848,850 &  91,787 &  39 &  2577 &    66        &     {\bf\alert{6856}} \\
                                              % 1663 &
        3,372,707 & 364,649 &  39 &  2211 &    57       &     {\bf\alert{6193}} \\
                                              % 1732 
        %5667771 & 365709 &  43 &  2139 &    50 &     1978 &     6299\\
        22,570,303 & 1,456,225 &  40 &  1923 &    48     &     {\bf\alert{5376}}\\
                                              % 1490
        \hline
  \end{tabular}
\end{center}
\end{table}
  }

\small
\begin{itemize}
\item {\bf \#sdof}: number of degrees of freedom for the
  state variables;
\item {\bf \#pdof}: number of degrees of freedom for the
  inversion parameter field;
\item {\bf \#N}: number of Newton iterations;
\item {\bf \#CG}, {\bf avgCG}: total and average (per Newton
  iteration) number of CG iterations;
\item {\bf \#Stokes}: total number of
  linear(ized) Stokes solves (from forward, adjoint, and incremental
  forward and adjoint problems)
\end{itemize}

\begin{itemize}
\item [] \scriptsize{Details in: T. Isaac, N. Petra, G. Stadler, and
  O. Ghattas. {\em Scalable and efficient algorithms for the
    propagation of uncertainty from data through inference to
    prediction for large-scale problems, with application to flow of
    the Antarctic ice sheet}, Journal of Computational Physics, 296,
  348-368 (2015).}
\end{itemize}

\end{frame}

%---------------------------------------------------------------------------------%
\begin{frame}
\frametitle{Low-rank approximation of the Hessian of the likelihood}
% (Gauss-Newton)

\framesubtitle{Under the assumption of Gaussian noise and linearized
  parameter-to-observable map}

\begin{itemize}

\item Invoke low-rank approximation  given by the inverse of the Hessian:
  \begin{minipage}{0.9\textwidth}
  \begin{block}{}
    %\vspace{-0.5cm} 
    \[
    \postcov = \Hmatrix^{-1} =
    \left( \prcov \Hmatrix_{\text{misfit}} + \bs{I} \right)^{-1} \prcov
    \approx \prcov - \bs{W}_r \bs{D}_r \bs{W}^T_r,
    %\left ( {\bs{F}^T\ncov^{-1}\bs{F}} + \prcov^{-1} \right)^{-1}
    %\approx \prcov^{1/2} (\bs{V}_r \bs{\Lambda}_r \bs{V}^T_r + \bs{I})^{-1}
    %\prcov^{1/2} 
    % - \bs{V}_r \bs{D}_r \bs{V}^T_r
    \]
  \end{block}
  \end{minipage}

  \vspace{0.15in} where $\bs{W}_r$ contains the largest $r$
  eigenvectors of $\prcov \Hmatrix_{\text{misfit}}$, and $\bs{D}_r =
  \diag(\lambda_i/(\lambda_i +1))\in \R^{r \times r}$, where
  $\lambda_i$ are eigenvalues of $\prcov \Hmatrix_{\text{misfit}}$.
%  the eigenvalue problem 
%$\bs{F}^T\ncov^{-1}\bs{F} \bs{v}_i = \lambda_i \prcov^{-1} \bs{v}_i $
  %, \bs{\Lambda}_r = \diag(\lambda_i)_{i=1}^r$
%  are truncated
%  generalized eigenvectors of the data misfit
%  Hessian, $\bs{D}_r =\diag(\lambda_i/(\lambda_i+1))$, and used
%  the Sherman-Morrison-Woodbury th. % to invert/factor.

%\item Spectrum of the prior-preconditioned likelihood Hessian for the
%  Arolla glacier (139 parameters, left) and
%  Antarctica (1.19M parameters, right) 
  %
  \end{itemize}
\begin{figure}[htb]
    \centering\includegraphics[width=0.35\textwidth]{extraplots/spectrum.pdf}
    \hspace{0.2in}
    \centering\includegraphics[width=0.35\textwidth]{extraplots/spec_ppmisfit_hess_coarseandfine_new.pdf}
  \end{figure}
  %
%\end{itemize}
\begin{center}
  The spectrum of the prior-preconditioned likelihood Hessian for the
  Arolla glacier (139 parameters, left) and Antarctica (1.19M
  parameters, right).
  \end{center}
\end{frame}

%---------------------------------------------------------------------------------%
\begin{frame}
\frametitle{MCMC sampling: stochastic Newton}
\framesubtitle{Performance results / Convergence diagnostics}
%\vspace{0.1in}
\small

\newcommand{\proposal}{\bs y}
\newcommand{\params}{\bs m}

\only<1>{
\begin{table}[t]
  \begin{center}
    %\def\firstrowcolor{}
    %\def\secondrowcolor{}
    %\only<1>{\def\firstrowcolor{\rowcolor{gray}}}
    %\only<2>{\def\secondrowcolor{\rowcolor{red}}}
    \begin{tabular}{|l|c|c|c|c|c|c|c|}
      \hline\hline
          {} & {\bfseries MPSRF} & {\bfseries IAT} & {\bfseries ESS} & {\bfseries
            MSJ} & {\bfseries ARR} & {\bfseries $\#$Stokes} & {\bfseries time (s)}\\
          \hline%\hline
          %\multicolumn{8}{|c|}{small noise case (more Gaussian) }\\
          %\hline
          %ind. sampler & 1.210 & 253 & 2075  & 1456 & 41 & 2783  & 139\\
          %frozen Hessian  & 1.001 & 6 & 84004 & 1390 & 40 & 72     & 4  \\
          %dynamic Hessian & 1.073 & 125 & 4032  & 565  & 17 & 1375 & 69\\
          %\hline
          %\multicolumn{8}{|c|}{large noise case (more non-Gaussian)}\\
          %\hline
          %SNIS  & 1.507 & 435 & 1207 & 280 & 9  & 4350 &  218\\
          %SNMAP  & 1.045 & 80 & 6563 & 190  & 6  & 960  & 48 \\
          SN & 1.348 & 600 & 875  & 64   & 2  & {\bf \alert{8400}} & 420 \\
          \hline\hline
    \end{tabular}
  \end{center}
\end{table}
}
\vspace{-0.05in}

\begin{columns}
  \begin{column}{0.45\columnwidth}
    \begin{itemize}
    \item {\bfseries MPSRF:} multivariate potential scale reduction factor
      \vspace{-0.02in}
    \item {\bfseries IAT:} integrated autocorrelation time
      \vspace{-0.02in}
    \item {\bfseries ESS:} effecitive sample size
      \vspace{-0.02in}
    \item {\bfseries MSJ:} mean squared jump distance
    \end{itemize}
  \end{column}
  \begin{column}{0.45\columnwidth}
    \begin{itemize}
    \item {\bfseries ARR:} average rejection rate
      \vspace{-0.02in}
    \item {\bfseries $\#$Stokes:} $\#$ of Stokes solves per independent sample
      \vspace{-0.02in}
    \item {\bfseries time:} time per independent sample
    \end{itemize}
  \end{column}
\end{columns}

\begin{columns}
  \hspace{0.26in}
  \begin{column}{0.95\columnwidth}
    \begin{itemize}
    \item {\bfseries Statistics:} 21 parallel chains (each 25k); $\#$
      samples: 525k; dof: 139; rank Hessian: 15
    %\item SNIS, SNMAP and SN are variants of the stochastic Newton
    %  MCMC method
    \end{itemize}
  \end{column}
  \begin{column}{0.5\columnwidth}
  \end{column}
\end{columns}

\vspace{0.05in}

%\vspace{-0.05in}
 \begin{align*}
   %\tcdb{q(\params_k,\proposal) =
     \mbox{ Proposal density: } \frac { \det \bs H^{1/2} }{(2\pi)^{n/2} }
      \exp \left( - \frac 12 \left( \proposal - \bs{\beta}_k + \bs H^{-1} \bs
      g \right) ^T \bs H \left( \proposal - \bs{\beta}_k + \bs H^{-1} \bs g \right)
      \right)
      %}
 \end{align*}

\begin{itemize}
\item [] \scriptsize{Details in: N. Petra, J. Martin, G. Stadler,
  O. Ghattas. {\em A computational framework for
    infinite-dimensional Bayesian inverse problems: Part
    II. Stochastic Newton MCMC with application to ice sheet inverse
    problems}, SIAM Journal on Scientific Computing, 2014}
\end{itemize}

\end{frame}
%---------------------------------------------------------------------
\begin{frame}
  \frametitle{Motivation to go beyond ``{\it global} low-rank approximation''}

    \begin{columns}
    \begin{column}{0.6\paperwidth}
        \begin{itemize}
  \item Existing Hessian approximations are based on low-rank approximation
    methods, which require computing twice as many linearized forward or
    adjoint partial differential equation (PDE) solves as the numerical
    rank of the Hessian.
    \vspace{0.05in}
  \item These methods are inefficient when the numerical rank of the Hessian
    is large, as is the case in continental scale ice sheet inverse
    problems.

  \end{itemize}
    \end{column}
    \begin{column}{0.35\paperwidth}
    \vspace{0.1in}
      \centering\includegraphics[width=0.95\columnwidth]{extraplots/spec_ppmisfit_hess_coarseandfine_new.pdf}
    \end{column}
    \end{columns}

    \vspace{0.1in}
    \begin{center}
      \alert{Our goal is to use additional structure combined with
        hierarchical matrix compression to reduce the
        computational cost of solving the (Bayesian) ice sheet inverse
        problem.}
    \end{center}
\end{frame}
%---------------------------------------------------------------------------
\begin{frame}
 \frametitle{Exploiting the problem structure}
 \framesubtitle{Local sensitivities}
 \center

 \begin{tikzpicture}
   \fill[red!90,nearly transparent] (5,0) -- (7,.5) -- (3,.5) -- (1,0) -- cycle;
   \draw [fill=red!40!white,opacity=1] (3,2.25) -- (3.5,2.25) -- (3.25,.25) -- cycle;
   \draw [fill=red] (3.25,2.25) circle (.25cm and 0.07cm);
   \draw (3.25,2.25) -- (2,3);
   \draw [thick] (2,3) node[above]{sensitivities, $\frac{\partial u}{\partial\beta_j}$};
   \draw[red,dashed] (7,.5) -- (3,.5) -- (1,0);
   \draw [fill=blue!40!white,opacity=1] (4.25,.25) -- (4.5,2.25) -- (4.75,.25) -- cycle;
   \draw [fill=blue] (4.5,.25) circle (.25cm and 0.07cm);
   \draw (4.5,.25) -- (2,-.5);
   \draw [thick] (2,-.5) node[below]{sensitivities, $\frac{\partial u_i}{\partial\beta}$};
   \draw (1,0) -- (5,0) -- (5,2) -- (1,2) -- (1,0);
   \draw  (5,2) -- (7,2.5) -- (3,2.5) -- (1,2);
   \draw  (7,2.5) -- (7,.5) -- (5,0);
   \draw  (5,2.25) -- (6,3) ;
   \draw [thick] (6,3) node[above]{measurements, $\boldsymbol{d}$};
   \draw[red]   (5.5,.25) -- (5.5,.125) ;
   \draw  (5.5,.125) -- (5.5,-.5) ;
   \draw [thick] (5.5,-.5) node[below]{parameter, $\beta(x)$};

   \draw[dashed] (3,.5) -- (3,2);
   \draw[blue,dashed] (3,2) -- (3,2.5);
   \fill[blue!90,nearly transparent] (5,2) -- (7,2.5) -- (3,2.5) -- (1,2) -- cycle;
 \end{tikzpicture}
%----------------------------------------------------------------------
\end{frame}
\begin{frame}
  \frametitle{Relevant literature}
  \scriptsize
  \vspace{-0.04in}
  \begin{itemize}
  \item {\bf (Global) low rank approximation}
    \vspace{0.02in}
    \begin{enumerate}
      \scriptsize
    \item Flath, Wilcox, Ak\c{c}elik, Hill, Van Bloemen Waanders, and
      Ghattas, {\it Fast algorithms for Bayesian uncertainty
        quantification in large-scale linear inverse problems based on
        low-rank partial Hessian approximations}, SISC, 2011.
    \item Halko, Martinsson, and Tropp, {\it Finding structure with
      randomness: Probabilistic algorithms for constructing
      approximate matrix decompositions}, SIAM review, 2011.
    \item Ghattas and Willcox, {\it Learning physics-based models from
      data: perspectives from inverse problems and model reduction},
      Acta Numerica, 2021.
    \end{enumerate}
    \vspace{-0.02in}
  \item {\bf Hierarchical (H-)matrix approximations}
    \vspace{0.02in}
    \begin{enumerate}
      \scriptsize
    \item Hackbusch, {\it Hierarchical matrices: algorithms and
      analysis}, Springer, 2015.
    \item Martinsson, {\it Fast direct solvers for elliptic PDEs},
      SIAM, 2020.
    \item Lin, Lu, and Ying, {\it Fast construction of hierarchical
      matrix representation from matrix-vector multiplication},
      JCP,2011.
    \item Ambartsumyan, Boukaram, Bui-Thanh, Ghattas, Keyes, Stadler,
      Turkiyyah, and Zampini, {\it Hierarchical matrix approximations
        of Hessians arising in inverse problems governed by PDEs},
      SISC, 2020.
    \item Hartland, Stadler, Perego, Liegeois, and Petra, {\it
      Hierarchical off-diagonal low-rank approximation of Hessians in
      inverse problems, with application to ice sheet model
      initialization}, Inverse Problems, 2023.
    \item Chen, Anitescu, {\it calable Physics-based Maximum
      Likelihood Estimation using Hierarchical Matrices}, JUQ, 2023.
    \end{enumerate}
    \vspace{-0.02in}
  \item {\bf Point spread function approximation combined with
    hierarchical matrix (H-matrix) approximation}
    \vspace{0.02in}
    \begin{enumerate}
      \scriptsize
    \item Gentile, Courbin, and Meylan, {\it Interpolating point
      spread function anisotropy} Astronomy \& Astrophysics, 2013.
    \item Escande and Weiss, {\it Approximation of integral operators using product-convolution
expansions}, Journal of Mathematical Imaging and Vision, 2017.
    \item Alger, {\it Data-scalable Hessian preconditioning for
      distributed parameter PDE-constrained inverse problems}, PhD
      thesis, The University of Texas at Austin, 2019.
    \end{enumerate}
  \end{itemize}
\end{frame}
%---------------------------------------------------------------------
\begin{frame}
  \frametitle{Hierarchical matrices ($\mathcal{H}$-matrices)}

    \begin{columns}
    \begin{column}{0.7\paperwidth}
       \begin{itemize}
		\item Matrix is reordered and subdivided recursively
		\item Many off-diagonal blocks are low-rank
		\item Overall matrix may be high rank
		\item $O(N \left(\log N\right)^a)$ complexity matrix operations, $a \in \{0,1,2,3\}$
		\begin{itemize}
			\item matrix-vector products, matrix-matrix addition, matrix-matrix multiplication, matrix factorization, matrix inversion, ...
		\end{itemize}
	\end{itemize}
    \end{column}
    \hspace{-0.4in}
    \begin{column}{0.3\paperwidth}
    \vspace{0.1in}
      \centering\includegraphics[width=0.85\textwidth]{extraplots/heat_inverse_problem_Hfull_hmatrix.eps}
    \end{column}
    \end{columns}

    \vspace{0.1in}
        \begin{itemize}
        \item We use H1 matrices in the numerical results via the
          HLIBpro software package, but any other H-matrix format
          could be used instead.
        \item Details in:
          \scriptsize
          \begin{itemize}
        \item W. Hackbusch, ``A sparse matrix arithmetic based on
          $\mathcal{H}$-matrices. Part I: Introduction to $\mathcal{H}$-matrices'', Computing,
          62 (1999), pp. 89-108.
        \item  R. Kriemann,
          ``HLIBpro user manual'', Max-Planck-Institute for
          Mathematics in the Sciences, Leipzig, 48 (2008).
          \end{itemize}
        \end{itemize}

\end{frame}
%---------------------------------------------------------------------
\begin{frame}
	\frametitle{Hierarchical matrix versus matrix free}
	\begin{itemize}
	\item Classical methods for building $\mathcal{H}$-matrix require matrix entries $\mathbf{H}_{ij}$
          \vspace{0.05in}
	\item New algebraic methods based on ``peeling process'' can build $\mathcal{H}$-matrix from matrix-vector products
          \vspace{0.05in}
	\item \textbf{Problem:} peeling process better than low rank, but \textbf{still expensive} (see for instance [*] below)
          \vspace{0.05in}
	\item Here we build the $\mathcal{H}$-matrix faster by taking advantage of the problem structure (details in [**])
	  \begin{itemize}
	  \item \textbf{Local sensitivities}
	  \item \textbf{Local mean-displacement invariance}
	  \item \textbf{Non-negative impulse responses}
	  \end{itemize}
	\end{itemize}

        \vspace{0.1in}
        \scriptsize
        \begin{itemize}
        \item [*] {T. Hartland, G. Stadler, M. Perego,
          K. Liegeois, N. Petra. ``Hierarchical off-diagonal low-rank
          approximation of Hessians in inverse problems, with application to
          ice sheet model initialization'', Inverse Problems, 39 (8), 2023. }
          \item [**] {Details in: N. Alger, T. Hartland, N. Petra,
            O. Ghattas. ``Point spread function approximation of high
            rank Hessians with locally supported non-negative integral
            kernels'', To appear in SIAM Journal on Scientific
            Computing (SISC).}
     \end{itemize}

\end{frame}
%---------------------------------------------------------------------------------%
\begin{frame}
  \frametitle{Fast high-rank Hessian approximation for Bayesian ice
    sheet inverse problems}
      \begin{itemize}
      \item {\bf Stage 1: Hessian approximation via product-convolution
        approximation:}
        %leads to fast access to matrix entries, which
        %makes
        %the approximation well-suited for hierarchical matrix compression.
        %to reduce the computational cost of the
        %{$\mathcal{H}$-matrice compression, we first approximate the Hessian
        %  using a {\note product-convolution approximation approach} that
        %  takes advantage of the Hessian's local translation invariance.
        \begin{itemize}
        \item Compute {\bf the impulse responses} of the Hessian operator at
          many points by applying the operator to a small number of Dirac
          combs of point sources.
          %\item Within each Dirac comb, the point sources must be spaced
          %  sufficiently far apart so that the impulse responses to those
          %  point sources do not overlap with each other or with the boundary.
        \item Estimate the required spacing of point sources by applying the
          operator to a small number of constant, linear, and quadratic
          functions.
        \item Once computed, the impulse responses are interpolated to approximate arbitrary Hessian entries.
        \end{itemize}
      \item {\bf Stage 2: $\mathcal{H}$-matrix compression:}
        \begin{itemize}
        \item Convert the product-convolution Hessian approximation to
          hierarchical matrix format.
        \item Invert the compressed matrix using fast hierarchical matrix
          arithmetic.
        \end{itemize}

        \item {\bf Stage 3: Hessian-approximation via
          $\mathcal{H}$-matrix compression:}
          \begin{itemize}
            \item Use this approximation to precondition linear systems involving the Hessian (e.g., solve the Newton system to
              compute the MAP point) and/or draw samples from the
              posterior.
          \end{itemize}
      \end{itemize}
\end{frame}

%-------------------------------------------------------------
\begin{frame}
  \frametitle{Stage1: Hessian approximation via PSF}
  \framesubtitle{Intuitively, one may think of
    impulse responses as ``columns'' of the integral kernel.}

  \begin{figure}
    \begin{center}
      \includegraphics[scale=0.25]{localpsf_revised_figures/kernel_impulse_illustration2.pdf}
    \end{center}
    %\caption{Left: Matrix created by evaluating the integral
    %  kernel $\Phi$ for $\Aop$
    %  (Equation~\ref{eq:kernel_representation}) at all pairs of
    %  mesh vertices. This illustration is for the integral kernel
    %  given in Equation~\ref{eq:frog_kernel} in
    %  Section~\ref{sec:frog}. Dark colors indicate large entries
    %  and light colors indicate small entries. Rows and columns
    %  are ordered according to a kd-tree hierarchical
    %  clustering. Right: Impulse responses associated with points
    %  $x_1,x_2\in \Omega$, shown by the two dotted vertical
    %  lines. Intuitively, one may think of impulse responses as
    %  ``columns'' of the integral kernel.  }
    %\label{fig:frog_kernel_impulse_responses}
  \end{figure}

  \begin{center}
    Left: Integral kernel matrix. Right: Impulse responses associated
    with points $x_1,x_2\in \Omega$, shown by the two dotted vertical
    lines.
  \end{center}
\end{frame}
%-------------------------------------------------------------
\begin{frame}
  \frametitle{Stage1: Hessian approximation via PSF}

  \def \pos {0.5\columnwidth}
  \begin{figure}[ht]\centering
    \begin{tikzpicture}
      \only<1>{
        \node (1) at (0*\pos-0.4*\pos,  0*\pos){\includegraphics[scale=0.30]{localpsf_revised_figures/frog_moments3.pdf}};
      }
      \only<2>{
        \node (1) at (0*\pos-0.4*\pos,  0*\pos){\includegraphics[scale=0.30]{localpsf_revised_figures/frog_moments3.pdf}};
        \node (2) at (0*\pos-1.1*\pos,  0*\pos-0.28*\pos){\includegraphics[scale=0.30]{localpsf_revised_figures/frog_ellipsoid_width2.pdf}};
      }
    %\begin{subfigure}[b]{0.34\textwidth}
      %\only<2>{\centering \includegraphics[scale=0.3]{localpsf_revised_figures/frog_ellipsoid_width2.pdf}}
    %\end{subfigure}
    %\caption{Left: Impulse response moments: scaling factor
    %  ($V$), mean ($\mu$), and covariance ($\Sigma$). For each
    %  point $x \in \Omega$, the quantity $V(x)$ is the integral
    %  of $\phi_x$ over $\Omega$, $\mu(x)$ is the location that
    %  $\phi_x$ is centered at, and $\Sigma(x)$ is a matrix with
    %  eigenvectors and eigenvalues that characterize the width of
    %  $\phi_x$ (see Section~\ref{sec:intromoments}). Right:
    %  Ellipsoid support for an impulse response.  This ellipsoid
    %  is the set of points within $\tau$ standard deviations of
    %  the mean of the Gaussian distribution with mean
    %  $\spatialmean(x)$ and covariance $\spatialcov(x)$. The
    %  scaling factor $V(x)$ characterizes the magnitude of
      %  $\phi_x$.  } \label{fig:frog_moments_ellipsoid}
      \end{tikzpicture}
  \end{figure}

  \begin{center}
    \only<1> {Impulse response moments: scaling factor ($V$), mean
    ($\mu$), and covariance ($\Sigma$).}
    \only<2> {Impulse response moments: scaling factor ($V$), mean
    ($\mu$), and covariance ($\Sigma$). Bottom left: Ellipsoid support
    for an impulse response.}
  \end{center}
\end{frame}
  %-------------------------------------------------------------
\begin{frame}
  \frametitle{Stage1: Hessian approximation via PSF}
  \framesubtitle{Illustration of the process to compute one impulse
    response batch}
\begin{figure}
	\centering \includegraphics[scale=0.3]{localpsf_revised_figures/frog_batches3.pdf}
	%\caption{Illustration of the process to compute one impulse
        %  response batch. Impulse response moments are first used to
        %  form ellipsoid shaped estimates of the supports of impulse
        %  responses (Equation~\ref{eq:support_ellipsoid}). Then, an
        %  ellipsoid packing problem is solved to choose batches of
        %  non-overlapping support ellipsoids
        %  (Section~\ref{sec:sample_point_selection}).  Finally,
        %  $\Aop$ is applied to a Dirac comb associated with the
        %  points $x_i$, which correspond to the ellipsoids
        %  (Section~\ref{sec:get_impulse_response}). The process is
        %  repeated to form more batches.}
	%\label{fig:frog_batches}
\end{figure}

\begin{enumerate}
\item Impulse response moments are first used to form ellipsoid shaped
  estimates of the supports of impulse responses
\item Then, an ellipsoid
  packing problem is solved to choose batches of non-overlapping
  support ellipsoids.
\item Finally,  $\Aop$ is applied to a Dirac comb associated with the
  points $x_i$, which correspond to the ellipsoids
\item  The process is
  repeated to form more batches.
\end{enumerate}

\end{frame}
%-------------------------------------------------------------
\begin{frame}
  \frametitle{Stage1: Hessian approximation via PSF}
\framesubtitle{Kernel entry approximation via radial basis function interpolation}
\begin{figure}
	\centering \includegraphics[scale=0.2]{localpsf_revised_figures/frog_hmatrix_knn_impulse3.pdf}
	%\caption{Left: H-matrix structure for $\Aker$. Computing an
        %  entry of this matrix requires evaluating the integral
        %  kernel, $\Phi(y,x)$, at a pair of points $(y,x) \in \Omega
        %  \times \Omega$.  Center: Kernel evaluation points $x$ and
        %  $y$ (black circles), sample points for the approximation
        %  (light gray and black dots), and the $k_n$ sample points,
        %  $x'$, that are nearest to $x$ (black dots). Right: Known
        %  impulse response at $x'$. Using radial basis function
        %  interpolation, the desired kernel entry is approximated as
        %  a weighted linear combination of translated and scaled
        %  versions of impulse responses at the points $x'$
        %  (Section~\ref{sec:approximate_kernel_entries}).
        %  } \label{fig:hmatrix_neighbors_frog}
\end{figure}

\begin{center}
  Left: H-matrix structure for $\Aker$. Center: Kernel evaluation
  points $x$ and $y$ (black circles), sample points for the
  approximation (light gray and black dots), and the $k_n$ sample
  points, $x'$, that are nearest to $x$ (black dots). Right: Known
  impulse response at $x'$.
\end{center}

%\begin{center}
%  Using radial basis function interpolation, the desired kernel entry
%  is approximated as a weighted linear combination of translated and
%  scaled versions of impulse responses at the points $x'$
%\end{center}
\end{frame}

%---------------------------------------------------------------------
\begin{frame}
  \frametitle{Stage1: Hessian approximation via PSF}
  \framesubtitle{Quality of the approximation}

  \begin{figure}
    \centering
    \includegraphics[scale=0.35]{localpsf_revised_figures/frog_column_errors_psf5.png}
    \hfill
    \includegraphics[scale=0.35]{localpsf_revised_figures/frog_column_errors_psf10.png}
    \hfill
    \includegraphics[scale=0.35]{localpsf_revised_figures/frog_column_errors_psf20.png}
    \hfill
    \includegraphics[scale=0.35]{localpsf_revised_figures/frog_colorbar_edited1.pdf}
    %\caption{Relative error, $||\Phi(\cdot,x) -
    %  \widetilde{\Phi}(\cdot,x)||/||\Phi(\cdot,x)||$, in the
    %  approximation of the ``column'' of the integral kernel
    %  associated with $x$, using 5 (left), 10 (center) and 20
    %  (right) impulse response batches. Sample points are
    %  indicated by black dots. The error associated with the point
    %  $x$ is the shade of the image at location $x$, with white
    %  indicating zero error and black indicating $100\%$ error. At
    %  the sample points, the error is zero. The further the point
    %  $x$ is from the samples points, the larger the error. Adding
    %  more batches yields a more accurate approximation.  }
  \end{figure}

  \begin{center}
    Relative error, $||\Phi(\cdot,x) -
      \widetilde{\Phi}(\cdot,x)||/||\Phi(\cdot,x)||$, in the
      approximation of the ``column'' of the integral kernel
      associated with $x$, using 5 (left), 10 (center) and 20
      (right) impulse response batches.
  \end{center}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
  \frametitle{Ellipsoid estimates for the supports
    of impulse responses} \framesubtitle{Blur kernel (left two
    columns) and a Ricker wavelet-type kernel (right two columns)}
  \begin{figure}
    \centering
	{
	  \begin{tabular}{cccc}
	    \includegraphics[scale=0.2]{localpsf_revised_figures/frog_ellipsoid_a=1.0.png} & 
	    \includegraphics[scale=0.2]{localpsf_revised_figures/frog_1d_a=1.0.png} & 
	    \includegraphics[scale=0.2]{localpsf_revised_figures/ricker_ellipsoid_a=0.0.png} & 
	    \includegraphics[scale=0.2]{localpsf_revised_figures/ricker_1d_a=0.0.png} \\
	    %
	    \includegraphics[scale=0.2]{localpsf_revised_figures/frog_ellipsoid_a=20.0.png} & 
	    \includegraphics[scale=0.2]{localpsf_revised_figures/frog_1d_a=20.0.png} & 
	    \includegraphics[scale=0.2]{localpsf_revised_figures/ricker_ellipsoid_a=0.23.png} & 
	    \includegraphics[scale=0.2]{localpsf_revised_figures/ricker_1d_a=0.23.png} \\
	    %
	    \includegraphics[scale=0.2]{localpsf_revised_figures/frog_ellipsoid_a=27.0.png} & 
	    \includegraphics[scale=0.2]{localpsf_revised_figures/frog_1d_a=27.0.png} & 
	    \includegraphics[scale=0.2]{localpsf_revised_figures/ricker_ellipsoid_a=0.249.png} & 
	    \includegraphics[scale=0.2]{localpsf_revised_figures/ricker_1d_a=0.249.png} 
	  \end{tabular}
	}
	%\caption{Illustration of the influence of negative numbers in
        %  the integral kernel on the robustness of the ellipsoid
        %  estimates for the supports of impulse responses. Left two
        %  columns: spatially variant blur kernel given in Equation
        %  \ref{eq:frog_kernel}. Right two columns: Ricker wavelet-type
        %  kernel given by $\Aker(y,x) = (1 - a
        %  \gamma)\exp\left(-\gamma/2\right)$, where $\gamma=(y-x)^T
        %  \Sigma^{-1} (y-x)$, and
        %  $\Sigma=\operatorname{diag}\left(0.0025,
        %  0.01\right)$. Ordered from top to bottom, the results are
        %  obtained with $a\in\{1.0, 20.0, 27.0\}$ for the left two
        %  columns, and $a\in \{0.0, 0.23, 0.249\}$ for the right two
        %  columns. Columns 1 and 3: impulse responses with estimated
        %  support ellipsoids indicated by the black ellipses. Red and
        %  blue represent positive and negative numbers in the integral
        %  kernel, respectively. Columns 2 and 4: one-dimensional slice
        %  along the horizontal line indicated in the two-dimensional
        %  plots. The dashed gray line is at zero.  }
	%\label{fig:robustness_to_negatives}
  \end{figure}

  \begin{center}
    Spatially variant blur kernel (left two columns) and a Ricker
    wavelet-type kernel (right two columns) for progressively more
    significant negative numbers in the integral kernel (top to
    bottom).
  \end{center}

\end{frame}

%---------------------------------------------------------------------
\begin{frame}
  \frametitle{Relative error in the PSF approximation of the kernel}
  %\framesubtitle{Relative error in the PSF approximation of the kernel}

  \begin{figure}
	\includegraphics[scale=0.96]{localpsf_revised_figures/frog_tau_impulses.pdf}
	\includegraphics[scale=0.96]{localpsf_revised_figures/frog_tau_batches.pdf}
  \end{figure}

  \begin{center}
    Left: Relative error vs. the total number of impulse responses
    used in the approximation. Right: Relative error vs. the number of
    impulse response batches.The dashed gray lines show linear
    convergence rates.
  \end{center}

\end{frame}
%---------------------------------------------------------------------
%\begin{frame}
%  \frametitle{Exploiting the problem structure}
%  \framesubtitle{Local mean-displacement invariance}
%  
%  \begin{figure}
%	\centering\includegraphics[scale=0.75]{figures/mean_displacement_invariance.pdf}
%  \end{figure}
%
%  \begin{center}
%    Impulse responses that arise from applying $\Aop$ to point sources \\at two points, $x$ and $x'$.
%  \end{center}
%   %\footnotesize
%   \begin{itemize}
%   \item The operator $\Aop$ is locally mean displacement invariant if
%     $\impulseresponse_{x}(y) \approx \impulseresponse_{x'}\left(y -
%     \spatialmean(x) + \spatialmean(x')\right)$ when $x$ is close to
%     $x'$.
%   \item Here $\spatialmean(z)$ denotes the mean (center of mass) of $\phi_z$.
%   \end{itemize}
%
%\end{frame}
%---------------------------------------------------------------------
\begin{frame}
  \frametitle{Example: Ice sheet inverse problem}

  \begin{figure}
	\centering\includegraphics[scale=0.22]{figures/meshHeight_view2_edges.png}\\
	\includegraphics[scale=0.18]{figures/mtrue_withColorBar.png}
	\includegraphics[scale=0.18]{figures/trueVelocity2_glyphs.png}
  \end{figure}
  \begin{center}
    The ice sheet geometry and discretization (top), the ``true''
    parameter (bottom left) and the corresponding velocity field
    (bottom right).
  \end{center}
\end{frame}
%---------------------------------------------------------------------
\begin{frame}
  \frametitle{Example: Ice sheet inverse problem}
  \framesubtitle{Impulse responses}

  \begin{figure}
	\includegraphics[width=0.32\columnwidth]{figures/impulse_batch1.png}
	\includegraphics[width=0.32\columnwidth]{figures/impulse_batch2.png}
	\includegraphics[width=0.32\columnwidth]{figures/impulse_batch3.png}
  \end{figure}

   %\footnotesize
   \begin{itemize}
   \item Black stars are point source locations.
   \item Shading shows the magnitude of the normalized impulse
     responses (darker means larger function values).
   \item Dashed gray ellipses are estimated impulse
     response support ellipsoids based on the moment method.
   \end{itemize}

\end{frame}
%---------------------------------------------------------------------
\begin{frame}
  \frametitle{Example: Ice sheet inverse problem}

  \begin{figure}
    %[width=0.8\columnwidth]
    \centering\includegraphics[scale=0.6]{extraplots/reconstructions-ice.png}
	%\centering\includegraphics[scale=0.18]{figures/mstar_0.25noise_cropped.png}
	%\includegraphics[scale=0.18]{figures/mstar_0.05noise_cropped.png}
	%\includegraphics[scale=0.18]{figures/mstar_0.01noise_cropped.png}
  \end{figure}
  \begin{center}
    The log basal sliding parameter computed from the PDE-constrained
    optimization problem with noise levels: $25\%$ (left), $5.0\%$
    (middle), and $1.0\%$ (right).
  \end{center}
\end{frame}
%---------------------------------------------------------------------
\begin{frame}
  \frametitle{Example: Ice sheet inverse problem}
  \framesubtitle{Convergence history for solving the Stokes inverse problem using inexact Newton PCG}

  \only<1>{
  \begin{table}
	{\small
		\begin{center}
			\begingroup
			\setlength{\tabcolsep}{3pt}
			\renewcommand{\arraystretch}{1.1}
			\begin{tabular}{c| c c c | c c c | c c c}
				&  \multicolumn{3}{|c|}{PSF (5)} & \multicolumn{3}{|c|}{REG} & \multicolumn{3}{|c}{NONE} \\
				\hline
				Iter & 
				\#CG & \#Stokes & $\|\mathbf{g}\|$ & 
				\#CG & \#Stokes & $\|\mathbf{g}\|$ & 
				\#CG & \#Stokes & $\|\mathbf{g}\|$ \\
				0 &
				1 & 4 & 1.9e+7 &
				3 & 8 & 1.9e+7 &
				1 & 4 & 1.9e+7 \\
				1 &
				2 & 6  & 6.1e+6 &
				8 & 18 & 8.4e+6 &
				2 & 6  & 6.1e+6 \\
				2 &
				4 & 10 & 2.6e+6 &
				16 & 34 & 4.1e+6 &
				4 & 10 & 2.6e+6 \\
				3 &
				2 & 6+22 & 6.9e+5 &
				34 & 70 & 1.8e+6 &
				14 & 30 & 6.9e+5 \\
				4 &
				3 & 8 & 4.4e+4 &
				52 & 106 & 5.6e+5 &
				29 & 60 & 1.3e+5 \\
				5 &
				5 & 12 & 2.2e+3 &
				79 & 160 & 9.4e+4 &
				38 & 78 & 1.0e+4 \\
				6 &
				0 & 2 & 1.1e+1 &
				102 & 206 & 6.5e+3 &
				58 & 118 & 1.8e+2 \\
				7 &
				--- & --- & --- &
				151 & 304 & 1.2e+2 &
				0 & 2 & 5.5e-1 \\
				8 & 
				--- & --- & --- &
				0 & 2 & 2.9e-1 &
				--- & --- & --- \\
				\hline
				Total & 
				17 & 70 & --- &
				445 & 908 & --- &
				146 & 308 & --- \\
			\end{tabular}
			\endgroup
		\end{center}
	}
\end{table}
  }

  \only<2>{
    \begin{table}
	{\small
		\begin{center}
			\begingroup
			\setlength{\tabcolsep}{3pt}
			\renewcommand{\arraystretch}{1.1}
			\begin{tabular}{c| c c c | c c c | c c c}
				&  \multicolumn{3}{|c|}{PSF (5)} & \multicolumn{3}{|c|}{REG} & \multicolumn{3}{|c}{NONE} \\
				\hline
				Iter & 
				\#CG & \#Stokes & $\|\mathbf{g}\|$ & 
				\#CG & \#Stokes & $\|\mathbf{g}\|$ & 
				\#CG & \#Stokes & $\|\mathbf{g}\|$ \\
				0 &
				1 & 4 & 1.9e+7 &
				3 & 8 & 1.9e+7 &
				1 & 4 & 1.9e+7 \\
				1 &
				2 & 6  & 6.1e+6 &
				8 & 18 & 8.4e+6 &
				2 & 6  & 6.1e+6 \\
				2 &
				4 & 10 & 2.6e+6 &
				16 & 34 & 4.1e+6 &
				4 & 10 & 2.6e+6 \\
				3 &
				2 & 6+22 & 6.9e+5 &
				34 & 70 & 1.8e+6 &
				14 & 30 & 6.9e+5 \\
				4 &
				3 & 8 & 4.4e+4 &
				52 & 106 & 5.6e+5 &
				29 & 60 & 1.3e+5 \\
				5 &
				5 & 12 & 2.2e+3 &
				79 & 160 & 9.4e+4 &
				38 & 78 & 1.0e+4 \\
				6 &
				0 & 2 & 1.1e+1 &
				102 & 206 & 6.5e+3 &
				58 & 118 & 1.8e+2 \\
				7 &
				--- & --- & --- &
				151 & 304 & 1.2e+2 &
				0 & 2 & 5.5e-1 \\
				8 & 
				--- & --- & --- &
				0 & 2 & 2.9e-1 &
				--- & --- & --- \\
				\hline
				Total & 
				17 & {\bf \alert{70}} & --- &
				445 & {\bf \alert{908}} & --- &
				146 & {\bf \alert{308}} & --- \\
			\end{tabular}
			\endgroup
		\end{center}
	}
\end{table}  
  }

  \footnotesize
  \begin{itemize}
  \item \#CG:  the number of PCG iterations used to solve the Newton system.
  \item \#Stokes:the total number of Stokes PDE solves performed in each Newton iteration.
  \item $\|\mathbf{g}\|$: the $l^2$ norm of the gradient.
  \end{itemize}
\end{frame}
%---------------------------------------------------------------------
\begin{frame}
  \frametitle{Example: Ice sheet inverse problem}
  \framesubtitle{Convergence history and eigenvalue clustering}

  \begin{figure}
    \begin{center}
    \includegraphics[width=0.85\columnwidth]{extraplots/bbb}
    \end{center}
  \end{figure}
  \begin{center}
    Left: convergence history for solving $\mathbf{H} \mathbf{x} = \mathbf{b}$
    using PCG, where $\mathbf{b}$ has i.i.d. random entries drawn from the
    standard Gaussian distribution and $\mathbf{H}$ is evaluated at the
    solution of the inverse problem. Right: generalized eigenvalues
    for generalized eigenvalue problem $\mathbf{H} \mathbf{u}_{k}=\lambda_{k}
    \mathbf{\preconditioner} \mathbf{u}_{k}$.
  \end{center}
\end{frame}
%---------------------------------------------------------------------
\begin{frame}
  \frametitle{Example: Ice sheet inverse problem}

  \begin{table}
    \begin{center}
      \begingroup
      \setlength{\tabcolsep}{4pt}
      \renewcommand{\arraystretch}{1.25}
      \begin{tabular}{c| c c c c c}
	noise    & \multicolumn{5}{c}{COND$(\mathbf{\preconditioner}^{-1} \mathbf{H})$ } \\ \cline{2-6}
	level    & REG     &	NONE  & PSF $(1)$ & PSF $(5)$ & PSF $(25)$ \\ \hline 
	$25\%$   & 1.01e+3 & 2.96e+3  & 1.34e+0   & 1.30e+0   & 1.18e+0    \\ 
	$11\%$   & 7.40e+3 & 1.05e+3  & 2.27e+0   & 1.55e+0   & 1.31e+0    \\   
	$5.0\%$  & 3.29e+4 & 4.96e+2  & 5.61e+0   & 3.06e+0   & 1.92e+0    \\ 
	$2.2\%$  & 1.66e+5 & 8.89e+2  & 1.58e+1   & 8.07e+0   & 4.03e+0    \\  
	$1.0\%$  & 5.36e+5 & 1.61e+3  & 7.17e+1   & 1.93e+1   & 9.19e+0    \\   
      \end{tabular}
      \endgroup
    \end{center}
  \end{table}

  \begin{center}
  Condition number for $\mathbf{\preconditioner}^{-1} \mathbf{H}$ for
  the PSF-based preconditioners with 1, 5, and 25 batches (PSF (1),
  PSF (5), and PSF(25), respectively), no preconditioner (NONE) and
  regularization (REG).
  \end{center}  
\end{frame}

%------------------------------------------------------------------------------------------------------------
\begin{frame}
  \frametitle{Comparison of cost to approximate the blur kernel}

  \begin{table}
    \begin{center}
      {\small
	\begingroup
	\setlength{\tabcolsep}{8pt}
	\renewcommand{\arraystretch}{1.0}
	\begin{tabular}{c| c | c c c c c}
	  & {\bf Error}    &	{\bf \#applies PSF}   & {\bf \#applies HODLR} 	& {\bf \#applies RSVD} 	\\%& rank \\ 
	  \hline 
	  & 20\%    		& 11  				& 592				& 354				\\%& 111  \\ 
	  $L=1$ 		& 10\% 			& 15  				& 772				& 520				\\%& 178  \\ 
	  & 5\% 			& 23   				& 924				& 674				\\%& 244 \\ 
	  \hline
	  & 20\%    		& 8  				& 852 				& 1316				\\%& 417 \\ 
	  $L=1/2$ 	& 10\%    		& 10  				& 1144				& 1916				\\%& 665 \\ 
	  & 5\% 	  		& 12   				& 1404				& 2456				\\%& 903 \\ 
	  \hline
	  & 20\%    	 	& 7					& 932				& 2624				\\%& 845  \\ 
	  $L=1/3$ 	& 10\%    		& 8  				& 1264 				& 3734				\\%& 1321  \\ 
	  & 5\% 			& 8   				& 1520 				& 4660				\\%& 1759  \\ 
	\end{tabular}
	\endgroup
      }
    \end{center}
%\caption{(Spatially variant blur) Comparison of cost to
%  approximate the blur kernel from Equation
%  \eqref{eq:frog_kernel} using the PSF method, the randomized
%  HODLR (hierarchical off diagonal low rank) method, and GLR
%  (global low rank) approximation using randomized SVD. The
%  quantity $L$ scales the width of the impulse responses,
%  hence it influences the rank of the operator. Large $L$
%  means low rank, and small $L$ means high rank. The second
%  column (``Error tol'') is the relative error in the
%  approximation of the kernel measured in the Frobenius norm,
%  $||\Phi - \widetilde{\Phi}||_\text{Fro} /
%  ||\Phi||_\text{Fro}$. The remaining three columns show the
%  number of operator applies required to achieve the given
%  error tolerances, using the PSF, HODLR, and GLR methods.}
%    \label{tab:frog_psf_vs_hodlr_vs_glr}
  \end{table}

  \begin{itemize}
    %\item Comparison of cost to approximate a blur kernel using the
    %  PSF method, the randomized HODLR (hierarchical off diagonal low
    %  rank) method, and GLR (global low rank) approximation using
    %  randomized SVD.
    %\item PSF, HODLR, and GLR
  \item {$L$}: scales the width of the impulse responses (influences the rank)
    %, hence it influences the rank of the operator.
  \item {\bf Error}: is the relative error in the approximation of the
    kernel measured in the Frobenius norm.
  \item {\bf \#applies}: number of operator applies required to achieve
    the given error tolerances, using the {\bf PSF}, {\bf HODLR} (hierarchical
    off diagonal low rank), and {\bf GLR} (global low rank) methods.
  \end{itemize}

\end{frame}
 %--------------------------------------------------------------------------
%\begin{frame}
%  \frametitle{Summary and future work}
%
%  \begin{itemize}
%    %\setlength\itemsep{2em}
%  \item Hessian approximations are essential for efficient solution
%    of inverse problems governed by partial differential equations.
%    \vspace{0.05in}
%  \item Low-rank approximations of the Hessian become prohibitive as
%    the data becomes more informative (as is the case for ice sheet
%    inverse problems).
%    \vspace{0.05in}
%  \item Local point spread function interpolation combined with
%    Hierarchical matrix representations yield a highly efficient
%    Hessian approximation.
%
%  \item {\bf Future work:} apply this PSF-based method to the
%    Antarctic ice sheet inverse problem and full Bayesian inversion.
%  \end{itemize}
%
%  \vspace{0.1in}
%  \begin{itemize}
%  \item [] {Details in: N. Alger, T. Hartland, N. Petra,
%    O. Ghattas. ``Point spread function approximation of high rank
%    Hessians with locally supported non-negative integral kernels'',
%    To appear in SIAM Journal on Scientific Computing (SISC).}
%  \end{itemize}
%
%\end{frame}
%--------------------------------------------------------------------------
%\begin{frame}
%  \frametitle{Sessions here at UQ24:}
%  \begin{enumerate}
%    %\setlength\itemsep{2em}
%  \item aaa
%  \end{enumerate}
%
%\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
  \frametitle{The taxonomy of rank-structured matrices}

  \begin{table}
    \begin{center}

      \label{tbl:performance}
      \begin{tabular}{|r|r|r|}
        \hline
        & {\it Weak admissibility}       & {\it Strong admissibility}  \\
        & Simpler data structures. & Lower ranks.\\
        & Good for 1D and 2D.      & Longer interaction lists. \\
        \hline
            {\it General basis matrices:}     & HODLR              & $\Hm$-matrices  \\
            Lower ranks.               &                    & \\
            \hline
                {\it Nested basis matrices:}      & HBS/HSS                & $\Hm^2$-matrices \\
                Enable $O(N)$ complexity.  &                    &\\
                \hline
      \end{tabular}
    \end{center}
  \end{table}

  \vspace{0.1in}
  \small
  \vspace{0.02in}
  \begin{itemize}
  \item HODLR: Hierarchical off-diagonal low-rank matrices
  \item HBS: Hierarchical block separable matrices
  \item HSS: Hierarchically semiseparable matrices

    \vspace{0.05in}
  \item Weak admissibility: all the submatrices except those on the
    diagonal become low-rank.
  \item Strong admissibility: the off-diagonal block is compressed
    only if adjacent blocks are ``well''-separated.
  \end{itemize}

  \vspace{0.1in}
  \scriptsize
  Details in:
  \begin{itemize}
    \scriptsize
  \item [] Per-Gunnar Martinsson, {\it Fast direct solvers for
    elliptic PDEs}, SIAM, 2020.
  \item [] J. Ballani and D. Kressner, {\it Matrices with hierarchical
    low-rank structures, Springer, 2016.}
  \end{itemize}
\end{frame}
%--------------------------------------------------------------------------
\begin{frame}
  \frametitle{References}

  \begin{itemize}
    \setlength\itemsep{2em}
  \item N. Alger, T. Hartland, N. Petra, O. Ghattas. ``Point spread
    function approximation of high rank Hessians with locally
    supported non-negative integral kernels'', To appear in {\bf SIAM
      Journal on Scientific Computing (SISC)}
    (\url{http://arxiv.org/abs/2307.03349}).

    \item T. Hartland, G. Stadler, M. Perego, K. Liegeois,
      N. Petra. ``Hierarchical off-diagonal low-rank approximation of
      Hessians in inverse problems, with application to ice sheet
      model initialization'', {\bf Inverse Problems}, 39 (8), 2023.

    \item T. Isaac, N. Petra, G. Stadler, and O. Ghattas. ``Scalable
      and efficient algorithms for the propagation of uncertainty from
      data through inference to prediction for large-scale problems,
      with application to flow of the Antarctic ice sheet'', {\bf
        Journal of Computational Physics (JCP)}, 296, 348-368 (2015).
  \end{itemize}

\end{frame}
%---------------------------------------------------------------------

\end{document}

### 2024-03-01 presentations/siampp24-tucker/figures/adv_geigs_figure.tex
adding files for my SIAM PP24 talk

\documentclass{standalone}
\usepackage{tikz}
\usepackage{pgfplots}

\begin{document}
	\begin{tikzpicture}[baseline]%[baseline, trim axis right]
		\begin{axis}[width=0.49\linewidth, height=0.5\linewidth,
		ymode=log,
		xmin=-50,
		xmax=8100,
%		ymax=1.e6,
		title={Generalized eigenvalues},
		xlabel={$k$, generalized eigenvalue \#},
		ylabel={$\lambda_{k}$},
		xtick={0, 2000, 4000, 6000, 8000},
		ytick={1e-2, 1e0, 1e2, 1e4, 1e6},
		%legend pos=north east,
		legend style={font=\small},
		ylabel near ticks, ylabel shift={-6pt},
		axis x line*=bottom,
		axis y line*=right,
		%legend style={at={(0.95,1)},anchor=north east},
		]
		%\addlegendentry{REG}
		\addplot[dotted, line width=0.5mm] table[x=k, y=reg] {advDiff_generalizedEigenvalues.dat};
		\addplot[dash dot, line width=0.5mm] table[x=k, y=none] {advDiff_generalizedEigenvalues.dat};
		%\addlegendentry{PSF (1)}
		\addplot[blue, line width=0.5mm] table[x=k, y=psf1] {advDiff_generalizedEigenvalues.dat};
		%\addlegendentry{PSF (5)}
		\addplot[red, line width=0.5mm] table[x=k, y=psf5] {advDiff_generalizedEigenvalues.dat};
		%\addlegendentry{PSF (25)}
		\addplot[purple, line width=0.5mm] table[x=k, y=psf25] {advDiff_generalizedEigenvalues.dat};
		\end{axis}
	\end{tikzpicture}
\end{document}

### 2024-03-01 presentations/siampp24-tucker/localpsf_revised_figures/MorozovDiscrepancyPlot.tex
adding files for my SIAM PP24 talk

\begin{figure}[H]
\begin{center} 
\begin{tikzpicture}[baseline, trim axis right]
\begin{axis}[width=0.75\textwidth, height=0.75\textwidth,
            xmode=log,
            ymode=log,
            grid=both,
            xlabel={$\gamma$, regularization parameter},
			legend pos=north west,
]
\addlegendentry{data misfit}
\addplot[color=blue, mark=o, line width=0.5mm] table[x=morozov_gammas, y=morozov_discrepancies] {MorozovData.dat};
\addlegendentry{norm of noise}
\addplot[dash dot, line width=0.5mm] table[x=morozov_gammas, y=noise_datanorm] {MorozovData.dat};
\end{axis}
\end{tikzpicture}
\end{center}
\end{figure}

### 2024-03-01 presentations/siampp24-tucker/localpsf_revised_figures/frog_tau_impulses.tex
adding files for my SIAM PP24 talk

--- 2024-02-18 paper_sisc_first_revision/frog_tau_impulses.tex
@@
 		title={PSF convergence (impulse responses)},
-		legend style={at={(1.2,1)},anchor=north east,draw=none},
-		legend pos=south west,
-		legend style={font=\small}
+		legend style={at={(0.5,0.5)},anchor=north east,draw=none},
+%		legend pos=south west,
+		legend style={font=\small,fill=none}
 		]
 			\addlegendentry{$\tau=2.0$}
-			\addplot[black, line width=0.5mm] table {frog_convergence_impulses_tau2.0.dat};
+			\addplot[olive, line width=0.5mm] table {frog_convergence_impulses_tau2.0.dat};
 			\addlegendentry{$\tau=2.5$}
@@
 			\addlegendentry{$\tau=3.5$}
-			\addplot[purple, line width=0.5mm] table {frog_convergence_impulses_tau3.5.dat};
+			\addplot[black, line width=0.5mm] table {frog_convergence_impulses_tau3.5.dat};
 			\addlegendentry{$\tau=4.0$}
 			\addplot[red, line width=0.5mm] table {frog_convergence_impulses_tau4.0.dat};
+%			\addlegendentry{$\text{const} \times \left(\# \text{impulse responses}\right)$}
+			\addplot[dashed, color=gray, line width=0.5mm] table {frog_convergence_rate_impulses.dat};
 		\end{axis}

### 2024-03-01 presentations/siampp24-tucker/talk.tex
adding files for my SIAM PP24 talk

% Tikz plot for UQ. This is based on the tikz code of
% Fabian Schuh for the tikz examples website
\documentclass[10pt,final,xcolor=dvipsnames,aspect ratio=169]{beamer}
%\usepackage{etex}

\usetheme{Madrid}

\usepackage{xspace}
%\usepackage{enumitem}
\usepackage[mathscr]{euscript}
\usepackage{subfigure}
\hypersetup{colorlinks,linkcolor=,urlcolor=blue}

% \usetheme{Boadilla}
% \usecolortheme{seahorse}

% the macros, packages, etc:
\usepackage[absolute,overlay]{textpos}
\setbeamertemplate{items}[ball]
\setbeamertemplate{blocks}[rounded][shadow=true]
\setbeamertemplate{navigation symbols}{}
\newenvironment{reference}[2]{%
  \begin{textblock*}{\textwidth}(#1,#2)
  \scriptsize{\it\color{black}}}{\end{textblock*}}
\setbeamertemplate{blocks}[rounded]% [shadow=false]
\newenvironment{boxalertenv}{\begin{altenv}%
      {\usebeamertemplate{alerted text begin}\usebeamercolor[fg]{alerted text}\usebeamerfont{alerted text}\colorbox{bg}}
      {\usebeamertemplate{alerted text end}}{\color{.}}{}}{\end{altenv}}

\usepackage{graphicx}
\usepackage{multirow,color}
\usepackage{bbm}
\usepackage{amsfonts,amsmath,amssymb,amsbsy,amsthm}
\usepackage{movie15}
\definecolor{themec}{RGB}{51,108,121}

%\usepackage{movie15}
%\usepackage{multimedia}
% custom commands (add your own macros here)

\definecolor{bronze}{rgb}{0.8, 0.5, 0.2}
\definecolor{goldenyellow}{rgb}{1.0, 0.87, 0.0}
\definecolor{amber}{rgb}{1.0, 0.75, 0.0}
\definecolor{azure}{rgb}{0.94, 1.0, 1.0}
\definecolor{bittersweet}{rgb}{1.0, 0.44, 0.37}
\definecolor{brass}{rgb}{0.71, 0.65, 0.26}
\definecolor{camel}{rgb}{0.76, 0.6, 0.42}
\setbeamercolor{block body}{bg=camel!30}
\usefonttheme[onlymath]{serif}
\usepackage{color}

\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{tikz}
\usepackage{amsfonts,amsmath,amssymb,amsbsy,amsthm}
%\usecolortheme{tango}

%\usefonttheme{serif}     % Font theme: serif
%\usefonttheme[stillsansseriftext]{serif} 
%\renewcommand*\familydefault{\sfdefault}

\usepackage{pgf}
%\usepackage{pgfplots}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{tikz}
\usepackage{verbatim}
%\usepackage{movie15}
\usepackage{tikz,pgfplots}
\setbeamertemplate{caption}{\raggedright\insertcaption\par}
%\usepackage[active,tightpage]{preview}
%\PreviewEnvironment{tikzpicture}
\usetikzlibrary{arrows,automata}
\usetikzlibrary{positioning}
\usetikzlibrary{fit}

\input{defjcp14.tex}
\newcommand{\iparb}{\bar{\ipar}}
\newcommand{\euclidnorm}[1]{\left\| {#1} \right\|_2}
\newcommand{\incu}{\upupsilon}
\newcommand{\incp}{\uprho}
\newcommand{\QoIquad}{\QoI_\text{quad}}
\newcommand{\tcblue}[1]{\textcolor{blue}{#1}}
\newcommand{\betad}{\bs{\beta}}

\newcommand{\preconditioner}{\widetilde{H}}
\newcommand{\impulseresponse}{\phi}
\newcommand{\spatialmean}{\mu}
\newcommand{\Aop}{\mathcal{A}}
\newcommand{\btVFill}{\vskip0pt plus 1filll}
\tikzset{
    state/.style={
           rectangle,
           rounded corners,
           draw=black, very thick,
           minimum height=3em,
           inner sep=1pt,
           text centered,
           },
}

\newcommand{\Aker}{\Phi}
%\DeclareMathOperator{\diag}{diag}
%\newcommand{\Hmatrix}{\bs{H}}
\newcommand{\Hm}{\mathcal{H}}

% argument #1: any options
\newenvironment{customlegend}[1][]{%
    \begingroup
    % inits/clears the lists (which might be populated from previous
    % axes):
    \csname pgfplots@init@cleared@structures\endcsname
    \pgfplotsset{#1}%
}{%
    % draws the legend:
    \csname pgfplots@createlegend\endcsname
    \endgroup
}
\def\addlegendimage{\csname pgfplots@addlegendimage\endcsname}

% title, authors, etc:
\title[Local PSF high-rank Hessian approximation]{Local Point Spread Function Approximation of High-Rank Hessians in Inverse Problems Governed by	PDEs\thanks{Research supported
    partially by NSF-DMS-CAREER-1654311 and NSF-OAC-1550547.}}
%\author[{Noemi
%    Petra}]{No\'{e}mi~Petra\inst{1}}

\author[Tucker Hartland]{{Tucker Hartland}\inst{1}\\[2ex]
  {\small \textcolor{themec}{Joint work with:}}
  \\
  {\small Nick Alger}\inst{2},
  {\small No{\'e}mi Petra}\inst{3},
  {\small Omar Ghattas}\inst{2}
}

\institute[LLNL]{%
  \inst{1}{Center for Applied Scientific Computing, Lawrence Livermore National Laboratory}\\\smallskip
  \inst{2}{Oden Institute for Computational Engineering and Sciences, The University of Texas at Austin}\\\smallskip
  \inst{3}{Department of Applied Mathematics, University of California, Merced}
}

\date[March 5, 2024]
     {\footnotesize 
     	MS11
     	HPC Algorithms for Inverse Problems and Digital Twins \\
     	SIAM Conference on Parallel Processing for Scientific Computing (PP24) \\
       Tuesday, March 5}

 \begin{document}

 % -------------------------------------------------------------------
%%\AtBeginSubsection[]  % "Beamer, do the following at the start of every section"
%\AtBeginSection[]  % "Beamer, do the following at the start of every section"
%{
%  \begin{frame}<beamer>
%    \frametitle{Outline} % make a frame titled "Outline"
%    \small
%  %\tableofcontents[currentsubsection]  % show TOC and highlight current section
%  \tableofcontents[currentsection]  % show TOC and highlight current section
%\end{frame}
%}
%----------- titlepage ----------------------------------------------%
\begin{frame}[plain]
  \titlepage
\end{frame}

%-------------------------------------------------------------------
\begin{frame}
	\frametitle{Need predictive models with quantified uncertainties to accurately anticipate future sea
		level rise.}
	%\framesubtitle{Sea level projections from the IPCC 6th Assessment Report, 2021}
	%\vspace{0.2in}
	\begin{columns}
		\begin{column}{0.7\paperwidth}
			\centering\includegraphics[width=0.75\columnwidth]{extraplots/venezia}
		\end{column}
		\hspace{-0.4in}
		\begin{column}{0.3\paperwidth}
			\vspace{0.1in}
			\centering\includegraphics[width=0.75\columnwidth]{extraplots/ipcc-ar6-wgi-2021.png}
		\end{column}
	\end{columns}

	%\centering\includegraphics[width=0.55\columnwidth]{extraplots/nasa-sea-level-projection-tool.png}

	\vspace{0.1in}
	%\begin{block}{IPCC Fourth Assessment Report: Climate Change 2007}
	%  {\small
	%    {\em [T]he uncertainty in the projections of the land ice
	%      contributions [to sea level rise] is dominated by the various
	%      uncertainties in the land ice models themselves
	%      ... rather than in the temperature projections.''} %[10.A.6]
	%  }
	%\end{block}
	\begin{itemize}
		\item Several port cities will be at risk from coastal
		flooding in the future.
		\item Ice flowing from ice sheets to ocean is primary contributor
		to sea level rise.
	\end{itemize}

	\begin{itemize}
		\item [] \scriptsize{Details in: Masson-Delmotte, V. et al. ``IPCC,
			2021: Summary for Policymakers. In: Climate Change 2021: The
			Physical Science Basis. Contribution of Working Group I to the
			Sixth Assessment Report of the Intergovernmental Panel on Climate
			Change'', Cambridge University Press. In Press.}
	\end{itemize}

\end{frame}

\begin{frame}
	\frametitle{Motivation: inversion for continental scale ice sheet models}

	\vspace{0.2in}
	\begin{columns}
		\begin{column}{0.47\paperwidth}
			\centering\includegraphics[width=0.7\columnwidth]{extraplots/ScienceRignot-crop.pdf}
			%\vspace{-0.2in}
			\vspace{-0.1in}
			\begin{center}
				{\scriptsize Observed surface flow velocity from InSAR \\
					\vspace{-0.05in}
					(Rignot et.\ al, 2011)}
			\end{center}
		\end{column}
		\hspace{-0.4in}
		\begin{column}{0.51\paperwidth}
			\vspace{0.15in}

			\centering\includegraphics[width=0.8\columnwidth]{extraplots/beta_cube_cut.png}

			\vspace{0.1in}
			\begin{center}
				{\scriptsize Antarctic ice sheet inversion \\ for the basal friction parameter field \\
					\vspace{-0.05in}
					from InSAR surface velocities}
			\end{center}
		\end{column}
	\end{columns}

	%\vspace{0.3in}
	\btVFill

	\scriptsize{Details in: T. Isaac, N. Petra, G. Stadler, and
		O. Ghattas. {\em Scalable and efficient algorithms for the
			propagation of uncertainty from data through inference to
			prediction for large-scale problems, with application to flow of
			the Antarctic ice sheet}, Journal of Computational Physics, 296,
		348-368 (2015).}
	%	\begin{block}{IPCC Fourth Assessment Report: Climate Change 2007}
	%		{\small
	%			{\em [T]he uncertainty in the projections of the land ice
	%				contributions [to sea level rise] is dominated by the various
	%				uncertainties in the land ice models themselves
	%				... rather than in the temperature projections.} %[10.A.6]
	%		}
	%	\end{block}
\end{frame}

\begin{frame}
	\frametitle{Outline}
	TODO: edit!!!

	\begin{itemize}
		\item Motivation
		\vspace{0.02in}
		\begin{itemize}
			\item Why we need predictive models with quantified uncertainties
			to accurately anticipate future sea level rise?
			\vspace{0.02in}
			\item Why we need more efficient approximations of high rank
			Hessians?
		\end{itemize}
		\vspace{0.02in}
		\item Bayesian inverse problems governed by PDEs (preliminaries)
		\vspace{0.02in}
		\item Exploiting low-dimensional structure in Bayesian inverse problems governed by PDEs
		\vspace{0.02in}
		\begin{itemize}
			\item (Global) low rank approximation
			\vspace{0.02in}
			\item Hierarchical off-diagonal low rank (HODLR) approximation
			\vspace{0.02in}
			\item Point spread function approximation combined with
			hierarchical matrix (H-matrix) approximation
		\end{itemize}
		\vspace{0.02in}
		\item Example: Inversion for the basal friction
		coefficient in an ice sheet flow problem
		\vspace{0.02in}
		\item Conclusions
	\end{itemize}

\end{frame}

\begin{frame}
	\frametitle{Ice sheet dynamics}
	\framesubtitle{Stokes model}	
	\begin{block}{Balance of mass and linear momentum}
		\[
		\begin{aligned}
		% balance of linear momentum:
		- \gbf{\nabla} \cdot \gbf{\sigma}
		%\left[
		% \eta(\gbf{u}) \, \edot
		%-\gbf{I}p \right]
		&= \rho\, \gbf{g},
		&&[
		\gbf{\sigma}=\eta(\gbf{u}) \, \edot - \gbf{I}p]
		%\edot = \tfrac 12  (\gbf{\nabla u} + \gbf{\nabla  u}^{\top})]
		\hspace{-2em}
		\\
		% balance of mass:
		\gbf{\nabla} \cdot \gbf{u} &= 0, &&[\edot =\tfrac 12 (\gbf{\nabla u}+\gbf{\nabla u}^{\top})]		
		%\vspace{-2em} \\
		% conservation of energy:
		%\rho c \left(\frac{\partial \theta}{\partial t} + \gbf{u} \cdot %\gbf{\nabla}
		%\theta\right)  - \gbf{\nabla} \cdot (K \gbf{\nabla} \theta)
		%&= 2 \, \eta \, \mathrm{tr}(\edot^2)
		\end{aligned}
		\]
	\end{block}
	\vspace{0.2cm}
	\begin{minipage}{.49\columnwidth}
		\alert{Mathematical challenges:}
		\begin{itemize}
			\small
			\item nonlinear PDE model
			\item complex domain geometries
			%\item multiphysics couplings
			\item poorly conditioned linear systems
			\item uncertain \alert{basal boundary conditions}, topography,
			heat flux
			\item observational data: InSAR, laser altimetry, GRACE
			satellite, ice cores, radar
		\end{itemize}
	\end{minipage}
	\begin{minipage}{.48\columnwidth}
		\alert{Inference of basal bdry. cond.}
		\begin{itemize}
			\small
			\item[] inference of effective basal friction field $\exp(\tcr{\beta}({\bs x}))$ in
			Robin boundary condition
			\begin{equation*}
			{\bf T}(\bs \sigma {\bs n} + \exp\left(\tcr{\beta}\right){\bs u}) = \bs 0
			\end{equation*}
			($\bf T$ is tangential projector)
			\item critical for climate simulations % (``spin-up'')
			%\item effective friction/sliding coefficient $\beta(x)$
		\end{itemize}
	\end{minipage}
\end{frame}

% -------------------------------------------------------------------
\begin{frame}[t]
	\frametitle{Inverse problem formulation}
	\framesubtitle{PDE-constrained optimization by Newton's method}
	\textbf{\underline{Goal}:} determine the discretized parameter $\bs{\beta}^{\star}$, which minimizes the (regularized) misfit between data $\bs{d}$ and model-predictions $\bs{\mathcal{F}}(\bs{\beta})$ 
	\begin{align*}
	\bs{\beta}^{\star}&= \arg \;
	\min_{\bs \beta} \;J(\bs{\beta}):= \underbrace{\frac{1}{2} \left\|  \bs{\mathcal{F}}(\bs{\beta}) -\obs
		\right\|^2_{\ncov^{-1}}
		+ \frac{1}{2} \left\|\bs{\beta} - \overline{\bs{\beta}}\right\|^2_{\prcov^{-1}}}_{\text{data-misfit }+\text{ regularization}} \\	
	\onslide<2->{\implies &\bs{\nabla_{\beta}}J(\bs{\beta}^{\star})=\bs{0}} 
	\end{align*}
	\uncover<3->{%
		\textbf{\underline{``Outer'' optimization strategy}:} given a minimizer estimate $\bs{\beta}^{(k)}$, update the estimate by finding the direction $\bs{\hat{\beta}}$ for which the linearization of $\bs{\nabla_{\beta}}J$ about $\bs{\beta}^{(k)}$ vanishes
		\only<3-4>
		{ 
			\begin{align*}
			\bs{H}^{(k)}\,\bs{\hat{\beta}}=-\bs{g}^{(k)}
			\end{align*}
		}
		\only<5->{\newline %
		}
		%
	}
	\only<4>{%
		\textbf{\underline{Notation}:}
		\begin{align*}
		\bs{g}^{(k)}_{i}&=\left[\bs{\nabla_{\beta}}J(\bs{\beta}^{(k)})\right]_{i}\quad\quad \text{ gradient (adjoint-based)},\\
		\bs{H}^{(k)}_{i,j}&=\left[\bs{\nabla}^{2}_{\bs{\beta},\bs{\beta}}J(\bs{\beta}^{(k)})\right]_{i,j}\quad\text{ Hessian (adjoint-based)}. 
		\end{align*}
	}
	\only<5->{%
		\hspace{-0.5cm} \textbf{\underline{Newton system and update}:} 
		\begin{align*}
		\bs{H}^{(k)}\bs{\widehat{\beta}}&=-\bs{g}^{(k)}\\
		\bs{\beta}^{(k+1)}&=\bs{\beta}^{(k)}+\alpha\,\bs{\widehat{\beta}}
		\end{align*}
	}
	\only<6->{%
		\begin{center} 
			\textbf{\underline{\alert{Challenge}}:} computationally efficient iterative (Krylov-subspace) methods to solve the ``inner'' Newton linear systems for PDE-constrained optimization problems.
		\end{center}
	}
\end{frame}

%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{The Hessian plays a critical role in inverse problems}

	\begin{alignat*}{2}
	\bs{\beta}^{\star}&=  \arg \;
	\min_{\bs \beta} J(\bs{\beta}):= \frac{1}{2} \left\|  \bs{\mathcal{F}}(\bs{\beta}) -\obs
	\right\|^2_{\ncov^{-1}}
	+ \frac{1}{2} \left\|\bs{\beta} - \overline{\bs{\beta}}\right\|^2_{\prcov^{-1}}
	\end{alignat*}

	\begin{itemize}
		%\item<1-> Its spectral properties characterize the degree of
		%ill-posedness.
		%\vspace{0.1in}
		\item<2-> The Hessian, $\bs{H}=\bs{\nabla}^{2}_{\bs{\beta},\bs{\beta}}f(\bs{\beta})$, drives Newton-based optimization algorithms.% for
		%solving the inverse problem.
		\vspace{0.1in}
		\item<3-> The inverse of the Hessian locally characterizes uncertainty
		in the solution of the inverse problem.
		\item[]<3-> \scriptsize{Details in: J. Martin, L.C. Wilcox, C. Burstedde and O. Ghattas, {\em A stochastic Newton MCMC method for large-scale statistical inverse problems with application to seismic inversion}, SIAM Journal on Scientific Computing (2012).}		
		%(under the Gaussian
		%assumption, it is precisely the posterior covariance matrix).
		\vspace{0.1in}
		\item<4-> \alert{Challenge}: the Hessian is only available as a matrix-free operator; each Hessian-vector product requires the solution of $2$ linearized PDEs.
		\vspace{0.1in}
		\item<5-> \alert{Goal}: rapidly perform linear algebraic operations, i.e.,
		manipulation of the Hessian, hence \alert{seek
			means to approximate the Hessian}.
		\vspace{0.1in}
		\item<6-> These approximations can be used e.g., as
		\alert{preconditioners}, and to build MCMC proposal distributions.
	\end{itemize}
\end{frame}

%---------------------------------------------------------------------------------%

\begin{frame}[t]
	\frametitle{Exploiting problem structure}
	\framesubtitle{Low-rank (LR) properties of the data-misfit Hessian}
	\begin{columns}
		\begin{column}{0.49\paperwidth}
			\begin{itemize}
				\item<1-> Invoke low-rank (LR) approximation to estimate the inverse of the Hessian:
				\begin{minipage}{0.9\textwidth}
					\begin{block}{}
						\vspace{-0.4cm} 
						\begin{align*} 
						\Hmatrix = \underbrace{\Hmisfit}_{\text{data-misfit}} + \underbrace{\prcov^{-1}}_{\text{prior regularizer}} \\
						\Hmatrix^{-1} \approx \prcov^{1/2} (\bs{V}_r \bs{\Lambda}_r \bs{V}^{\top}_r + \bs{I})^{-1}
						\prcov^{\top/2} 
						% - \bs{V}_r \bs{D}_r \bs{V}^T_r
						\end{align*}
					\end{block}
				\end{minipage}

				\vspace{0.1in}
				where $\bs{V}_r$ and $\bs{\Lambda}_r$ are the eigenvectors/values of
				%  the eigenvalue problem 
				$\Hmisfit \bs{v}_i = \lambda_i \prcov^{-1} \bs{v}_i $
				%, \bs{\Lambda}_r = \diag(\lambda_i)_{i=1}^r$
				%  are truncated
				%  generalized eigenvectors of the data misfit
				%  Hessian, $\bs{D}_r =\diag(\lambda_i/(\lambda_i+1))$, and used
				%  the Sherman-Morrison-Woodbury th. % to invert/factor.perwidth}
			\end{itemize} 
		\end{column}
		\begin{column}{0.49\paperwidth}
			\begin{itemize}
				\item<2-> Spectrum of the prior-preconditioned data-misfit Hessian for Antarctica

				\begin{figure}[htb]
					%\centering\includegraphics[width=0.35\textwidth]{extraplots/spectrum.pdf}
					%\hspace{0.2in}
					\centering\includegraphics[width=0.5\columnwidth]{extraplots/spec_ppmisfit_hess_coarseandfine_new.pdf}
				\end{figure}
			\end{itemize}    	
		\end{column}
	\end{columns}
	\btVFill
	\only<3->
	{
		\begin{center} \textbf{\underline{Take-away messages}}
			\vspace{0.025in}

			\onslide<4->{The prior-preconditioned data-misfit Hessian has a mesh-independent rate of spectral decay.}

			\vspace{0.025in}
			\onslide<5->{
				For many problems of practical interest, the rank of a ``low-rank'' approximation is excessively large.}
		\end{center} 
		\vspace{0.1in}
	}

	\scriptsize{Details in: T. Isaac, N. Petra, G. Stadler, and
		O. Ghattas. {\em Scalable and efficient algorithms for the
			propagation of uncertainty from data through inference to
			prediction for large-scale problems, with application to flow of
			the Antarctic ice sheet}, Journal of Computational Physics, 296,
		348-368 (2015).}	 
\end{frame}

%% -------------------------------------------------------------------
\begin{frame}[t]
	\frametitle{Exploiting localized sensitivity problem structure}
	% \framesubtitle{Localized sensitivities of the parameter-to-PDE solution mapping}
	\begin{columns}
		\begin{column}{0.69\columnwidth}
			\only<1-2>
			{
			\begin{figure}
				\begin{center} 
					\begin{tikzpicture}[scale=0.9]
					\draw[red,dashed] (7,.5) -- (3,.5) -- (1,0);
					\draw[dashed] (3,.5) -- (3,2);
					\draw[blue,dashed] (3,2) -- (3,2.5);
					\fill[red!90,nearly transparent] (5,0) -- (7,.5) -- (3,.5) -- (1,0) -- cycle;
					\draw [fill=red!40!white,opacity=1] (3,2.25) -- (3.5,2.25) -- (3.28,.25) -- (3.22,0.25) -- cycle;
					\draw (3.25,0.25) -- (2.25, -0.5);
					\draw [thick] (2.25, -0.5) node[below]{perturbation $\tilde{\beta}$};
					\draw [fill=red] (3.25,2.25) circle (.25cm and 0.07cm);
					\draw [fill=red] (3.25,0.25) circle (.025cm and 0.007cm);
					\draw (3.25,2.25) -- (2,3);
					\draw [thick] (2,3) node[above]{sensitivity cone, $\frac{\delta \bs{u}}{\delta \beta}(\beta)(\tilde{\beta})$};
					%\draw [fill=blue!40!white,opacity=1] (4.25,.25) -- (4.5,2.25) -- (4.75,.25) -- cycle;
					%\draw [fill=blue] (4.5,.25) circle (.25cm and 0.07cm);
					%\draw (4.5,.25) -- (2,-.5);
					%\draw [thick] (2,-.5) node[below]{sensitivity, $\frac{\partial \beta }{\partial\bm{u}}$};
					\draw (1,0) -- (5,0) -- (5,2) -- (1,2) -- (1,0);
					\draw  (5,2) -- (7,2.5) -- (3,2.5) -- (1,2);
					\draw  (7,2.5) -- (7,.5) -- (5,0);
					\draw  (5,2.25) -- (6,3) ;
					\draw [thick] (6,3) node[above]{$\Gamma_{t}$};
					%\draw[red]   (5.5,.25) -- (5.5,.125) ;
					\draw  (4.75,.125) -- (5.75,-.5) ;
					\draw [thick] (5.75,-.5) node[below]{$\Gamma_{b}$};
					\fill[blue!90,nearly transparent] (5,2) -- (7,2.5) -- (3,2.5) -- (1,2) -- cycle;
					\end{tikzpicture}
				\end{center}
				\caption{Ice-sheet geometry illustration.} %The PDE solution $\bs{u}$, is only meaningfully impacted in a narrow cone about the support of the localized parameter perturbation $\phi_{i}$}
				%\label{fig:sensitivitycone} 
			\end{figure}
		}
	\only<3>
	{
			\begin{figure} 
			\begin{center} 	
				\begin{tikzpicture}[baseline, trim axis right]
				\begin{axis}[
				grid=major, width=8cm, height=4.5cm,
				xmode=log, %ymode=log,
				xlabel = $\|\bs{H}_{\text{misfit}}-\bs{\tilde{H}}_{\text{misfit}}\|_{2}/\|\bs{H}_{\text{misfit}}\|_{2}\text{, approximation error}$,
				ylabel = {Hessian-vector products},
				xmin=3.1e-4, xmax = 1.e-1,
				ymin=5.e1, ymax = 5.e3,
				%ytick={2.e3, 
				%		4.e3, 6e3, 8.e3, 1.e4},
				legend style={font=\small, nodes=left, nodes={scale=0.65, transform shape}},
				%xtick={1.e4, 1.e6, 1.e8, 1.e10, 
				%1.e12, 1.e14, 1.e16},
				%legend style={font=/tiny, nodes=right},
				legend pos=north east,
				label style={font=\small},
				%title = $\text{Block spectra of the Hessian}$
				]
				\addlegendentry{LR}
				\addplot [color=black, line width = 1.25pt]
				table{HODLRdata/GISLRcompression.dat};
				\addlegendentry{HODLR}
				\addplot [color=blue, line width = 1.25pt]
				table{HODLRdata/GISHODLRcompression.dat};
				\end{axis}
				\end{tikzpicture}	
				\caption{The cost to compress the Greenland ice-sheet data-misfit Hessian $\bs{H}_{\text{misfit}}$, into the global low-rank (LR) and HODLR formats $\bs{\tilde{H}}_{\text{misfit}}$.}
			\end{center}
		\end{figure}

}
			\vspace{-0.1in}
		\end{column}
		\begin{column}{0.25\columnwidth}
			\only<2->
			{
				\begin{figure} 	
					\begin{tikzpicture}
					\begin{scope}[yscale=0.2, xscale=0.2]
					\pgfmathtruncatemacro{\L}{3}
					\pgfmathtruncatemacro{\scalefac}{16/(2^\L)}
					\pgfmathtruncatemacro{\shiftval}{2^(\L-1)}
					\begin{scope}[yscale=-1, xscale=1]
					\begin{scope}[yshift=-\shiftval, xshift=-\shiftval]
					\begin{scope}[yscale=\scalefac, xscale=\scalefac]
					\pgfmathtruncatemacro{\msize}{2^\L}
					%\pgfmathtruncatemacro{\gridstep}{2^(-\L)}
					\pgfmathtruncatemacro{\colstep}{40/(\L-1)}
					%\draw[step=\gridstep cm, gray, very thin] (0, 0) grid (\msize, \msize);
					\foreach \l in {1,...,\L}
					{
						\pgfmathtruncatemacro{\delI}{\msize*2^(-\l)}
						\pgfmathtruncatemacro{\DelI}{2*\delI}
						%\pgfmathtruncatemacro{\colorl}{50+(\l-1)*\colstep};
						\pgfmathtruncatemacro{\colorl}{75}
						\pgfmathtruncatemacro{\maxi}{2^(\l-1)}
						\foreach \i in {1,...,\maxi}
						{
							\pgfmathtruncatemacro{\a}{(\i-1)*\DelI}
							\pgfmathtruncatemacro{\b}{(\i-1)*\DelI+\delI}
							\pgfmathtruncatemacro{\c}{\i*\DelI}
							\ifnum\l=1
							\filldraw[fill=teal] (\a, \b) rectangle (\b, \c);
							\filldraw[fill=teal] (\b, \a) rectangle (\c, \b);
							\else
							\ifnum\l=2 
							\filldraw[fill=olive] (\a, \b) rectangle (\b, \c);
							\filldraw[fill=olive] (\b, \a) rectangle (\c, \b);
							\else
							\filldraw[fill=gray] (\a, \b) rectangle (\b, \c);
							\filldraw[fill=gray] (\b, \a) rectangle (\c, \b);
							\fi
							\fi 
						}
					}
					% diagonal blocks
					\pgfmathtruncatemacro{\delI}{\msize*2^(-\L)}
					\pgfmathtruncatemacro{\DelI}{2*\delI}
					\pgfmathtruncatemacro{\colorl}{20+(\L-1)*\colstep}
					\pgfmathtruncatemacro{\maxi}{2^(\L-1)}
					\foreach \i in {1,...,\maxi}
					{
						\pgfmathtruncatemacro{\a}{(\i-1)*\DelI}
						\pgfmathtruncatemacro{\b}{(\i-1)*\DelI+\delI}
						\pgfmathtruncatemacro{\c}{\i*\DelI}
						\filldraw[fill=violet] (\a, \a) rectangle (\b, \b);
						\filldraw[fill=violet] (\b, \b) rectangle (\c, \c);
					}
					\end{scope}
					\end{scope}
					\end{scope}
					\end{scope}
					\end{tikzpicture}
					\caption{HODLR matrix partition structure}
				\end{figure} 
			}

		\end{column}
	\end{columns}

	\begin{itemize} 
		%\item[$\star$]<1-> \textbf{Observation:} a variation of $\beta$ at $\bs{x}_{j}$ will impact $\bs{u}$ in a neighborhood of $\bs{x}_{j}$.
		%  \vspace{0.1in}
		\item[$\star$] Ice-sheet inverse problem Hessians have numerically low-rank off-diagonal blocks 
		\onslide<2->{$\implies$ Hessians can be efficiently approximated with HODLR matrices.}
	\end{itemize}
	\begin{itemize} 
		\item[] \scriptsize{T. Hartland, G. Stadler, M. Perego, K. Liegeois and N. Petra. {\em Hierarchical off-diagonal low-rank approximation of Hessians in inverse problems, with application to ice-sheet model initialization}, Inverse Problems, 2023.}
		%  \vspace{0.1in}
		%\item[$\star$]<3-> \textbf{Computational strategy:} generate HODLR approximations.
	\end{itemize} 
\end{frame}

%% -------------------------------------------------------------------
\begin{frame}
	\frametitle{Hierarchical ($\mathcal{H}$-) matrices}
	\framesubtitle{Construction from matrix-free operators}
	\begin{itemize}
		\setlength\itemsep{1.5em}
		\item<1-> Classic methods for building $\mathcal{H}$-matrices \only<9>{\textbf{\underline{require the evaluation of matrix entries}}}\only<1-8>{require the evaluation of matrix entries} $\bs{H}_{i,j}$
		\item<2-> Purely algebraic methods based on \textit{peeling} can generate $\mathcal{H}$-matrices of matrix-free operators via the application of specially chosen (e.g., random) vectors

		{\scriptsize{P.G. Martinsson.
			{\em Compressing rank-structured matrices via randomized sampling},
			SIAM Journal on Scientific Computing, (2016).}
		}
		%\begin{itemize}
		%	\item {\scriptsize T. Hartland, G. Stadler, M. Perego, K. Li\'{e}geois and N. Petra. \emph{Hierarchical off-diagonal approximation of Hessians in inverse problems}. To be submitted.}
		%	\item Tucker Hartland is giving a talk about this later
		%	\item 3:00-3:25 Thursday, MS140, Room: Augusta F - 7th Floor
		%\end{itemize} 
		\item<3-> \alert{Challenge:} peeling process can be better than low rank, but \textbf{still expensive} \onslide<4->{and expense grows with the complexity of hierarchical matrix partitioning}
		\item<5-> \alert{Approach:} speed up the $\mathcal{H}$-matrix construction by
		exploiting the following problem structure
		\begin{itemize}
			\vspace{0.1in}
			\setlength\itemsep{0.5em}
			\item<6-> \textbf{Localized sensitivities}
			\item<7-> \textbf{Scaled local mean-displacement invariance}
			\item<8-> \textbf{Non-negative impulse responses}
		\end{itemize}
	\end{itemize}
\end{frame}

%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{Local mean-displacement invariance}
	\begin{center}
		\textbf{Impulse response $\phi_x$:}
		\begin{align*}
		\phi_x := H_d~ \delta_x =& \text{ action of Hessian operator on delta distribution at x} \\
		\mu(x) :=& \text{ center of mass (mean) of }\phi_x
		\end{align*}

		\vspace{1em}

		\includegraphics[width=0.35\columnwidth]{mean_displacement_invariance.pdf}

		\vspace{1em}

		\textbf{Local mean-displacement invariance:}
		$$\phi_x(y) \approx \phi_{x'}(y-\mu(x) + \mu(x'))$$

		%\includegraphics[width=0.6\columnwidth]{Hess1_ordered.png}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Hessian impulse responses (Pine Island Glacier)}
	\center
	\only<1>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi0_minimal.png}
		\end{center}
	}
	\only<2>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi1_minimal.png}
		\end{center}
	}
	\only<3>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi2_minimal.png}
		\end{center}
	}
	\only<4>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi3_minimal.png}
		\end{center}
	}
	\only<5>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi4_minimal.png}
		\end{center}
	}
	\only<6>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi5_minimal.png}
		\end{center}
	}
	\only<7>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi6_minimal.png}
		\end{center}
	}
	\only<8>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi7_minimal.png}
		\end{center}
	}
	\only<9>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi8_minimal.png}
		\end{center}
	}
	\only<10>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi9_minimal.png}
		\end{center}
	}
	\only<11>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi10_minimal.png}
		\end{center}
	}
	\only<12>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi11_minimal.png}
		\end{center}
	}
	\only<13>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi12_minimal.png}
		\end{center}
	}
	\only<14>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi13_minimal.png}
		\end{center}
	}
	\only<15>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi14_minimal.png}
		\end{center}
	}
	\only<16>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi15_minimal.png}
		\end{center}
	}
	\only<17>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi16_minimal.png}
		\end{center}
	}
	\only<18>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi17_minimal.png}
		\end{center}
	}
	\only<19>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi18_minimal.png}
		\end{center}
	}
	\only<20>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi19_minimal.png}
		\end{center}
	}
	\only<21>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi20_minimal.png}
		\end{center}
	}
	\only<22>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi21_minimal.png}
		\end{center}
	}
	\only<23>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi22_minimal.png}
		\end{center}
	}
	\only<24>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi23_minimal.png}
		\end{center}
	}
	\only<25>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi24_minimal.png}
		\end{center}
	}
	\only<26>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi25_minimal.png}
		\end{center}
	}
	\only<27>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi26_minimal.png}
		\end{center}
	}
	\only<28>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi27_minimal.png}
		\end{center}
	}
	\only<29>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi28_minimal.png}
		\end{center}
	}
	\only<30>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi29_minimal.png}
		\end{center}
	}
	\only<31>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi30_minimal.png}
		\end{center}
	}
	\only<32>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi31_minimal.png}
		\end{center}
	}
	\only<33>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi32_minimal.png}
		\end{center}
	}
	\only<34>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi33_minimal.png}
		\end{center}
	}
	\only<35>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi34_minimal.png}
		\end{center}
	}
	\only<36>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi35_minimal.png}
		\end{center}
	}
	\only<37>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi36_minimal.png}
		\end{center}
	}
	\only<38>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi37_minimal.png}
		\end{center}
	}
	\only<39>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi38_minimal.png}
		\end{center}
	}
	\only<40>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi39_minimal.png}
		\end{center}
	}
	\only<41>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi40_minimal.png}
		\end{center}
	}
	\only<42>{
		\begin{center}
			\includegraphics[width=0.70\columnwidth]{reordered_phi/phi41_minimal.png}
		\end{center}
	}
	\only<43>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi42_minimal.png}
		\end{center}
	}
	\only<44>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi43_minimal.png}
		\end{center}
	}
	\only<45>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi44_minimal.png}
		\end{center}
	}
	\only<46>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi45_minimal.png}
		\end{center}
	}
	\only<47>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi46_minimal.png}
		\end{center}
	}
	\only<48>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi47_minimal.png}
		\end{center}
	}
	\only<49>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi48_minimal.png}
		\end{center}
	}
	\only<50>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi49_minimal.png}
		\end{center}
	}
	\only<51>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi50_minimal.png}
		\end{center}
	}
	\only<52>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi51_minimal.png}
		\end{center}
	}
	\only<53>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi52_minimal.png}
		\end{center}
	}
	\only<54>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi53_minimal.png}
		\end{center}
	}
	\only<55>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi54_minimal.png}
		\end{center}
	}
	\only<56>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi55_minimal.png}
		\end{center}
	}
	\only<57>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi56_minimal.png}
		\end{center}
	}
	\only<58>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi57_minimal.png}
		\end{center}
	}
	\only<59>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi58_minimal.png}
		\end{center}
	}
	\only<60>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi59_minimal.png}
		\end{center}
	}
	\only<61>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi60_minimal.png}
		\end{center}
	}
	\only<62>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi61_minimal.png}
		\end{center}
	}
	\only<63>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi62_minimal.png}
		\end{center}
	}
	\only<64>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi63_minimal.png}
		\end{center}
	}
	\only<65>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi64_minimal.png}
		\end{center}
	}
	\only<66>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi65_minimal.png}
		\end{center}
	}
	\only<67>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi66_minimal.png}
		\end{center}
	}
	\only<68>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi67_minimal.png}
		\end{center}
	}
	\only<69>{
		\begin{center}
			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi68_minimal.png}
		\end{center}
	}

\end{frame}

\begin{frame}
	\frametitle{Impulse responses are ``columns'' of the integral kernel}
	\begin{center}

		\includegraphics[width=0.5\columnwidth]{kernel_impulse_illustration2.pdf}

	\end{center}
\end{frame}

\begin{frame}[t]
	\frametitle{Framework for approximate Hessian entry evaluation}
	\framesubtitle{Illustration of framework by impulse response local-translation invariance}
	\only<1-2>{
		Consider
		\begin{align*}
		\bs{v}^{\top}\bs{K}\bs{u}&=a(u,v):=
		\int_{\Omega}\bs{\nabla}u\cdot\bs{\nabla}v\mathrm{d}x
		\end{align*}
	}
	\begin{align*}
	\onslide<2->{\bs{v}^{\top}\Hmisfit\bs{u}=
		(H_{\text{\tiny misfit}}\,u)(v)&:=
		\int_{\Omega}\int_{\Omega}v(y)\,\Phi(y,x)\,u(x)\mathrm{d}x\,\mathrm{d}y \\}
	\onslide<3->{\Phi(y,x)&\approx \Phi(y+s,x+s) \text{ for small } s \quad\quad (\text{local-translation invariance}) \\}
	\onslide<4->{\underbrace{{\color<6->{teal}{\phi_{x}}}(y)}_{\text{impulse response}}&:=\Phi(y,x)=({\color<7->{magenta}{H_{\text{\tiny misfit}}\,\delta_{x}^{*}}})^{*}\footnotemark\\}
	\onslide<5->{
		\implies \color<5->{purple}{(H_{\text{\tiny misfit}}u)(v)}&\approx 
		\int_{\Omega}v(y)\left(\sum_{i}\int_{\Omega_{i}}{\color<6->{teal}{\phi_{x_{i}^{\prime}}}}(y-x+x_{i}^{\prime})u(x)\mathrm{d}x\right)\mathrm{d}y}	
	\end{align*}
	\only<4->
	{%
		\footnotetext[1]{$\phantom{e}^{*}$ indicates the Riesz representative, that is $\ell^{*}\in L^{2}(\Omega)$ is the Riesz representative of a linear functional $\ell: L^{2}(\Omega)\rightarrow\mathbb{R}$, provided $\ell(w)=(\ell^{*},w)_{L^{2}(\Omega)}$ for each $w\in L^{2}(\Omega)$. $\delta_{x}$ indicates the Dirac distribution with mass concentrated at $x$.}   
	}
	\only<5->
	{%
		\begin{itemize}
			\item[$\star$] \textbf{Key idea:} \only<5->{approximate {\color<5->{purple}{Hessian entries}}} \only<6->{by interpolating a fixed number of {\color{teal}{impulse response functions}},} 
			\only<7->
			{each of which is determined by one {\color{magenta}{Hessian-apply}}.}
		\end{itemize}
	}

\end{frame}

%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{Hessian approximation method: big idea}
	\begin{itemize}
		\item {\bf Step 1:} Compute ``batches'' of impulse responses by applying Hessian to Dirac combs
		\item {\bf Step 2:} Interpolate known impulse responses to approximate unknown Hessian entries $\mathbf{H}_{ij}$
		\item {\bf Step 3:} Convert to $\mathcal{H}$-matrix to do linear algebra
	\end{itemize}
	\includegraphics[width=0.45\columnwidth]{impulse_batch1.png}  \includegraphics[width=0.45\columnwidth]{impulse_batch2.png}
	*Impulse response batches from Ice Mountain
\end{frame}
%---------------------------------------------------------------------------------%
%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{Technical details}
	{\Large
		\begin{itemize}
			\setlength\itemsep{2em}
			\item How do we choose the impulse response points?
			\begin{itemize}
				\item {\large How do we make sure they don't overlap?}
			\end{itemize}
			\item How do we interpolate the impulse responses? 
			\begin{itemize}
				\item {\large What about boundary issues?}
			\end{itemize}
		\end{itemize}
	}
\end{frame}
%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{How to choose impulse response points?}
	One hessian matrix-vector product $\rightarrow$ many impulse responses
	\begin{center}
		\includegraphics[width=0.5\columnwidth]{impulse_batch2.png} 
	\end{center}
	\begin{itemize}
		\item \textbf{Goal:} choose as many points as possible, such that the impulse response supports don't overlap
		\item \textbf{Dilemma:} How can we know the impulse response supports before we compute them?
	\end{itemize}
\end{frame}
%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{Matrix analogy: getting all row sums}

	\textbf{Matrix:} let $\mathbf{A} \in \mathbb{R}^{N \times N}$. Then

	$$\mathbf{A}^T \begin{bmatrix}
	1 \\ 1 \\ \vdots \\ 1
	\end{bmatrix} = \begin{bmatrix}
	\text{sum of }\mathbf{A} \text{ col }1 \\
	\text{sum of }\mathbf{A} \text{ col }2 \\
	\vdots \\
	\text{sum of }\mathbf{A} \text{ col }N
	\end{bmatrix}$$

	Apply matrix to vector of ones $\rightarrow$ get row sums for all rows

	\vspace{3em}

	\textbf{Operator:} let $C(x)=1$ be the constant function. Then

	$$(H_d^T C)^*(y) = \int_\Omega \left(H_d \delta_y\right)(x) dx$$

	Apply Hessian to constant function $\rightarrow$ get volumes of every impulse response
\end{frame}
%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{Mean and standard deviations of impulse responses}
	\begin{itemize}
		\setlength\itemsep{2em}
		\item Let $V(x)$, $\mu(x)$, and $\Sigma(x)$ be the ``volume'', ``mean'', and ``variance'' of $\phi_x$
		\item Let $C$, $L^i$, and $Q^{ij}$ be the following functions:
		\begin{equation*}
		C(x) := 1, \qquad
		L^i(x) := x^i, \qquad
		Q^{ij}(x) = x^i x^j
		\end{equation*}
		\item Then
		\begin{align*}
		V =& \left(H_d^T C\right)^* \\
		\mu^i =& \left(H_d^T L^i\right)^* / V \\
		\Sigma^{ij} =& \left(H_d^T Q^{ij}\right)^* / V - \mu^i\cdot \mu^j
		\end{align*}
		\item Apply Hessian to constant, linear, and quadratic functions $\rightarrow$ get estimates of support for every impulse response
	\end{itemize}

\end{frame}
%---------------------------------------------------------------------------------%
\begin{frame}
	\frametitle{Impulse response support ellipsoids}
	\begin{itemize}
		\item $\phi_x$ is \textbf{approximately supported} in the ellipsoid
		$$E = \{y: (y - \mu(x))^T\Sigma(x)^{-1}(y - \mu(x)) \le \tau^2\}$$

		\item Picking impulse response points becomes an \textbf{ellipsoid packing problem}

		\item Pack ellipsoids using \textbf{greedy algorithm}
	\end{itemize}
	\begin{center}
		\includegraphics[width=0.5\columnwidth]{impulse_batch2.png} 
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Radial basis function interpolation}
	\begin{center}
		\includegraphics[width=0.8\textwidth]{interpolation_rbf.pdf}
	\end{center}
	\begin{itemize}
		\item Interpolate impulse responses using polyharmonic spline radial basis functions.
		\item Use only $k$-nearest neighbors (must solve $k \times k$ linear system)
	\end{itemize}
\end{frame}

\begin{frame}
	\frametitle{Computational cost}
	\begin{itemize}
		\setlength\itemsep{2em}
		\item \textbf{Hessian-vector products:} 
		$$6 + n_\text{batches}$$
		E.g., $11$ Hessian-vector products for $5$ batches of impulse responses
		\item \textbf{Ellipsoid intersection tests:}
		$$O(Nm)$$
		where $m$ is total number of impulse responses in all batches
		\item \textbf{Elementary operations} to build and use the H-matrix:
		$$O(N \log N)$$ 
	\end{itemize}
\end{frame}

%---------------------------------------------------------------------------------%

%\begin{frame}
%	\frametitle{Boundary considerations (1)}
%	\begin{center}
%		\includegraphics[width=0.8\textwidth]{interpolation_rbf_boundary.pdf}
%	\end{center}
%	\begin{itemize}
%		\item If $p_i + y - x$ is outside the domain, don't use $i$th impulse response for $H(y,x)$
%	\end{itemize}
%\end{frame}
%---------------------------------------------------------------------------------%

%\begin{frame}
%	\frametitle{Boundary considerations (2)}
%	\begin{center}
%		\includegraphics[width=0.8\textwidth]{interpolation_rbf_adjoint.pdf}
%	\end{center}
%	\begin{itemize}
%		\item Take advantage of symmetry
%	\end{itemize}
%\end{frame}
%---------------------------------------------------------------------------------%

%---------------------------------------------------------------------------------%
%\begin{frame}
%	\frametitle{How to choose impulse response points?}
%	One hessian matrix-vector product $\rightarrow$ many impulse responses
%	\begin{center}
%		\includegraphics[width=0.5\columnwidth]{IRB1.png} 
%	\end{center}
%	\begin{itemize}
%		\item \textbf{Goal:} choose as many points as possible, such that the impulse response supports don't overlap
%		\item \textbf{Dilemma:} How can we know the impulse response supports before we compute them?
%	\end{itemize}
%\end{frame}
%---------------------------------------------------------------------------------%

%---------------------------------------------------------------------------------%

%---------------------------------------------------------------------------------%

%---------------------------------------------------------------------------------%
%\begin{frame}
%	\frametitle{Ice Mountain: Setup}
%	\begin{figure}
%		\begin{center} 
%			\begin{subfigure}{0.99\textwidth}
%				\begin{center} 
%					\includegraphics[scale=0.2]{meshHeight_view2_edges.png} \hfil
%				\end{center}
%				\caption{Ice sheet model geometry}
%				\label{fig:ice_mountain_mesh}
%			\end{subfigure} \\
%			\begin{subfigure}{0.49\textwidth}
%				\includegraphics[scale=0.2]{mtrue_withColorBar.png}
%				\caption{$\beta_\text{true}$}
%				\label{fig:true_beta}
%			\end{subfigure}
%			\begin{subfigure}{0.49\textwidth}
%				\includegraphics[scale=0.2]{trueVelocity2_glyphs.png}
%				\caption{$v_\text{true}$}
%				\label{fig:stokes_velocity}
%			\end{subfigure}
%		\end{center}
%	\end{figure} 
%\end{frame}
\begin{frame}
	\frametitle{Example: ``Ice mountain'' inverse problem}

	\begin{figure}
		\centering\includegraphics[scale=0.22]{figures/meshHeight_view2_edges.png}\\
		\includegraphics[scale=0.18]{figures/mtrue_withColorBar.png}
		\includegraphics[scale=0.18]{figures/trueVelocity2_glyphs.png}
	\end{figure}
	\begin{center}
		The ice sheet geometry and discretization (top), the ``true''
		parameter (bottom left) and the corresponding velocity field
		(bottom right).
	\end{center}
\end{frame}
%---------------------------------------------------------------------
\begin{frame}
	\frametitle{Example: ``Ice mountain'' inverse problem}
	\framesubtitle{Impulse responses}

	\begin{figure}
		\includegraphics[width=0.32\columnwidth]{figures/impulse_batch1.png}
		\includegraphics[width=0.32\columnwidth]{figures/impulse_batch2.png}
		\includegraphics[width=0.32\columnwidth]{figures/impulse_batch3.png}
	\end{figure}

	%\footnotesize
	\begin{itemize}
		\item Black stars are point source locations.
		\item Shading shows the magnitude of the normalized impulse
		responses.
		\item Dashed gray ellipses are estimated impulse
		response support ellipsoids based on the moment method.
	\end{itemize}

\end{frame}
%---------------------------------------------------------------------
\begin{frame}
	\frametitle{Example: Ice sheet inverse problem}

	\begin{figure}
		%[width=0.8\columnwidth]
		\centering\includegraphics[scale=0.6]{extraplots/reconstructions-ice.png}
		%\centering\includegraphics[scale=0.18]{figures/mstar_0.25noise_cropped.png}
		%\includegraphics[scale=0.18]{figures/mstar_0.05noise_cropped.png}
		%\includegraphics[scale=0.18]{figures/mstar_0.01noise_cropped.png}
	\end{figure}
	\begin{center}
		The log basal sliding parameter computed from the PDE-constrained
		optimization problem with noise levels: $25\%$ (left), $5.0\%$
		(middle), and $1.0\%$ (right).
	\end{center}
\end{frame}
%---------------------------------------------------------------------
\begin{frame}
	\frametitle{Example: Ice sheet inverse problem}
	\framesubtitle{Convergence history for solving the Stokes inverse problem using inexact Newton PCG}

	\only<1>{
		\begin{table}
			{\small
				\begin{center}
					\begingroup
					\setlength{\tabcolsep}{3pt}
					\renewcommand{\arraystretch}{1.1}
					\begin{tabular}{c| c c c | c c c | c c c}
						&  \multicolumn{3}{|c|}{PSF (5)} & \multicolumn{3}{|c|}{REG} & \multicolumn{3}{|c}{NONE} \\
						\hline
						Iter & 
						\#CG & \#Stokes & $\|\mathbf{g}\|$ & 
						\#CG & \#Stokes & $\|\mathbf{g}\|$ & 
						\#CG & \#Stokes & $\|\mathbf{g}\|$ \\
						0 &
						1 & 4 & 1.9e+7 &
						3 & 8 & 1.9e+7 &
						1 & 4 & 1.9e+7 \\
						1 &
						2 & 6  & 6.1e+6 &
						8 & 18 & 8.4e+6 &
						2 & 6  & 6.1e+6 \\
						2 &
						4 & 10 & 2.6e+6 &
						16 & 34 & 4.1e+6 &
						4 & 10 & 2.6e+6 \\
						3 &
						2 & 6+22 & 6.9e+5 &
						34 & 70 & 1.8e+6 &
						14 & 30 & 6.9e+5 \\
						4 &
						3 & 8 & 4.4e+4 &
						52 & 106 & 5.6e+5 &
						29 & 60 & 1.3e+5 \\
						5 &
						5 & 12 & 2.2e+3 &
						79 & 160 & 9.4e+4 &
						38 & 78 & 1.0e+4 \\
						6 &
						0 & 2 & 1.1e+1 &
						102 & 206 & 6.5e+3 &
						58 & 118 & 1.8e+2 \\
						7 &
						--- & --- & --- &
						151 & 304 & 1.2e+2 &
						0 & 2 & 5.5e-1 \\
						8 & 
						--- & --- & --- &
						0 & 2 & 2.9e-1 &
						--- & --- & --- \\
						\hline
						Total & 
						17 & 70 & --- &
						445 & 908 & --- &
						146 & 308 & --- \\
					\end{tabular}
					\endgroup
				\end{center}
			}
		\end{table}
	}

	\only<2>{
		\begin{table}
			{\small
				\begin{center}
					\begingroup
					\setlength{\tabcolsep}{3pt}
					\renewcommand{\arraystretch}{1.1}
					\begin{tabular}{c| c c c | c c c | c c c}
						&  \multicolumn{3}{|c|}{PSF (5)} & \multicolumn{3}{|c|}{REG} & \multicolumn{3}{|c}{NONE} \\
						\hline
						Iter & 
						\#CG & \#Stokes & $\|\mathbf{g}\|$ & 
						\#CG & \#Stokes & $\|\mathbf{g}\|$ & 
						\#CG & \#Stokes & $\|\mathbf{g}\|$ \\
						0 &
						1 & 4 & 1.9e+7 &
						3 & 8 & 1.9e+7 &
						1 & 4 & 1.9e+7 \\
						1 &
						2 & 6  & 6.1e+6 &
						8 & 18 & 8.4e+6 &
						2 & 6  & 6.1e+6 \\
						2 &
						4 & 10 & 2.6e+6 &
						16 & 34 & 4.1e+6 &
						4 & 10 & 2.6e+6 \\
						3 &
						2 & 6+22 & 6.9e+5 &
						34 & 70 & 1.8e+6 &
						14 & 30 & 6.9e+5 \\
						4 &
						3 & 8 & 4.4e+4 &
						52 & 106 & 5.6e+5 &
						29 & 60 & 1.3e+5 \\
						5 &
						5 & 12 & 2.2e+3 &
						79 & 160 & 9.4e+4 &
						38 & 78 & 1.0e+4 \\
						6 &
						0 & 2 & 1.1e+1 &
						102 & 206 & 6.5e+3 &
						58 & 118 & 1.8e+2 \\
						7 &
						--- & --- & --- &
						151 & 304 & 1.2e+2 &
						0 & 2 & 5.5e-1 \\
						8 & 
						--- & --- & --- &
						0 & 2 & 2.9e-1 &
						--- & --- & --- \\
						\hline
						Total & 
						17 & {\bf \alert{70}} & --- &
						445 & {\bf \alert{908}} & --- &
						146 & {\bf \alert{308}} & --- \\
					\end{tabular}
					\endgroup
				\end{center}
			}
		\end{table}  
	}

	\footnotesize
	\begin{itemize}
		\item \#CG:  the number of PCG iterations used to solve the Newton system.
		\item \#Stokes:the total number of Stokes PDE solves performed in each Newton iteration.
		\item $\|\mathbf{g}\|$: the $l^2$ norm of the gradient.
	\end{itemize}
\end{frame}
%---------------------------------------------------------------------
\begin{frame}
	\frametitle{Example: ``Ice mountain'' inverse problem}
	\framesubtitle{Convergence history and eigenvalue clustering}

	\begin{figure}
		\begin{center}
			\includegraphics[width=0.75\columnwidth]{extraplots/bbb}
		\end{center}
	\end{figure}
	%\begin{center}
		Left: convergence history for solving $\mathbf{H} \mathbf{x} = \mathbf{b}$
		using PCG, where $\mathbf{b}$ has i.i.d. random entries and $\mathbf{H}$ is evaluated at the
		solution of the inverse problem. \\ Right: generalized eigenvalues
		for generalized eigenvalue problem $\mathbf{H} \mathbf{u}_{k}=\lambda_{k}
		\mathbf{\preconditioner} \mathbf{u}_{k}$.
	%\end{center}
\end{frame}
%---------------------------------------------------------------------
\begin{frame}
	\frametitle{Example: ``Ice mountain'' inverse problem}
	\framesubtitle{Condition number of preconditioned Hessian}

	\begin{table}
		\begin{center}
			\begingroup
			\setlength{\tabcolsep}{4pt}
			\renewcommand{\arraystretch}{1.25}
			\begin{tabular}{c| c c c c c}
				noise    & \multicolumn{5}{c}{COND$(\mathbf{\preconditioner}^{-1} \mathbf{H})$ } \\ \cline{2-6}
				level    & REG     &	NONE  & PSF $(1)$ & PSF $(5)$ & PSF $(25)$ \\ \hline 
				$25\%$   & 1.01e+3 & 2.96e+3  & 1.34e+0   & 1.30e+0   & 1.18e+0    \\ 
				$11\%$   & 7.40e+3 & 1.05e+3  & 2.27e+0   & 1.55e+0   & 1.31e+0    \\   
				$5.0\%$  & 3.29e+4 & 4.96e+2  & 5.61e+0   & 3.06e+0   & 1.92e+0    \\ 
				$2.2\%$  & 1.66e+5 & 8.89e+2  & 1.58e+1   & 8.07e+0   & 4.03e+0    \\  
				$1.0\%$  & 5.36e+5 & 1.61e+3  & 7.17e+1   & 1.93e+1   & 9.19e+0    \\   
			\end{tabular}
			\endgroup
		\end{center}
	\end{table}

	%\begin{center}
		Condition number for $\mathbf{\preconditioner}^{-1} \mathbf{H}$ for
		the PSF-based preconditioners with 1, 5, and 25 batches (PSF (1),
		PSF (5), and PSF(25), respectively), no preconditioner (NONE) and
		regularization (REG).
	%\end{center}  
\end{frame}

\begin{frame}
	\frametitle{Summary}

	\begin{itemize}
		\setlength\itemsep{1.5em}
		\item Hessian approximations or preconditioners are essential for solution of Bayesian inverse problems governed by
		partial differential equations.
		\vspace{0.05in}
		\item Low-rank approximations of the Hessian become
		prohibitive as the data becomes more informative, as is the case
		for ice sheet inverse problems.
		\vspace{0.05in}
		\item Local point spread function interpolation combined with Hierarchical matrix representations promise a more efficient
		Hessian approximation.
	\end{itemize}
	\vspace{2em}
	{\small Alger, N., Hartland, T., Petra, N., Ghattas, O. (2024). Point spread function approximation of high rank Hessians with locally supported non-negative integral kernels. To appear in SISC.}
\end{frame}

\end{document}

### 2024-03-05 presentations/siampp24-tucker/talk.tex
many updates to SIAMPP24 slides

@@
 
-\begin{frame}
+\begin{frame}[t]
 	\frametitle{Outline}
-	TODO: edit!!!
-
-	\begin{itemize}
+
+	\begin{itemize}
+		\setlength\itemsep{1.0em}
 		\item Motivation
@@
 		\begin{itemize}
-			\item Why we need predictive models with quantified uncertainties
-			to accurately anticipate future sea level rise?
+			\setlength\itemsep{1.0em}
+			\item Predictive continental scale ice sheet models for anticipating future sea level rise with quantified uncertainties
 			\vspace{0.02in}
-			\item Why we need more efficient approximations of high rank
-			Hessians?
-		\end{itemize}
+			\item The need to go beyond ``low rank'' Hessian approximations
+		\end{itemize}
+		%\vspace{0.02in}
+		%\item Bayesian inverse problems governed by PDEs (preliminaries)
 		\vspace{0.02in}
-		\item Bayesian inverse problems governed by PDEs (preliminaries)
+		\item Exploiting structured low-rank properties of (Bayesian) inverse problems governed by PDEs
 		\vspace{0.02in}
-		\item Exploiting low-dimensional structure in Bayesian inverse problems governed by PDEs
-		\vspace{0.02in}
-		\begin{itemize}
+		\begin{itemize}
+			\setlength\itemsep{1.0em}
 			\item (Global) low rank approximation
@@
 			\vspace{0.02in}
-			\item Point spread function approximation combined with
-			hierarchical matrix (H-matrix) approximation
+			\item A local point spread function method as a means for efficient
+			$\mathcal{H}$ierarchical-matrix approximation
 		\end{itemize}
@@
 		\item Example: Inversion for the basal friction
-		coefficient in an ice sheet flow problem
+		coefficient field in an ice sheet flow problem
 		\vspace{0.02in}
@@
 				%  the eigenvalue problem 
-				$\Hmisfit \bs{v}_i = \lambda_i \prcov^{-1} \bs{v}_i $
+				$\prcov^{\top/2}\Hmisfit\prcov^{1/2}$
 				%, \bs{\Lambda}_r = \diag(\lambda_i)_{i=1}^r$
@@
 					\end{tikzpicture}
-					\caption{HODLR matrix partition structure}
+					\caption{HODLR matrix structure}
 				\end{figure} 
@@
 %% -------------------------------------------------------------------
-\begin{frame}
+\begin{frame}[t]
 	\frametitle{Hierarchical ($\mathcal{H}$-) matrices}
 	\framesubtitle{Construction from matrix-free operators}
-	\begin{itemize}
-		\setlength\itemsep{1.5em}
-		\item<1-> Classic methods for building $\mathcal{H}$-matrices \only<9>{\textbf{\underline{require the evaluation of matrix entries}}}\only<1-8>{require the evaluation of matrix entries} $\bs{H}_{i,j}$
+	\begin{columns}
+	\begin{column}{0.59\columnwidth}
+	\begin{itemize}
+		\setlength\itemsep{1.0em}
+		\item<1-> Classic methods for building $\mathcal{H}$-matrices \\ \only<8>{\textbf{\underline{require evaluation of matrix entries}}}\only<1-7>{require evaluation of matrix entries} $\bs{H}_{i,j}$
 		\item<2-> Purely algebraic methods based on \textit{peeling} can generate $\mathcal{H}$-matrices of matrix-free operators via the application of specially chosen (e.g., random) vectors
-
 		{\scriptsize{P.G. Martinsson.
@@
 		}
+	\end{itemize}
 		%\begin{itemize}
@@
 		%\end{itemize} 
+\end{column}
+\begin{column}{0.4\columnwidth}
+\begin{figure} 
+\begin{center} 
+\includegraphics[scale=0.16]{Hmatrix_Structure.png}
+\caption{$\mathcal{H}$-matrix structure}
+\end{center}
+\end{figure} 
+\btVFill
+\end{column}
+\end{columns}
+\begin{itemize}
+	%\vspace{0.1in}
 		\item<3-> \alert{Challenge:} peeling process can be better than low rank, but \textbf{still expensive} \onslide<4->{and expense grows with the complexity of hierarchical matrix partitioning}
 \item<5-> \alert{Approach:} speed up the $\mathcal{H}$-matrix construction by
-		exploiting the following problem structure
-		\begin{itemize}
-			\vspace{0.1in}
-			\setlength\itemsep{0.5em}
-			\item<6-> \textbf{Localized sensitivities}
-			\item<7-> \textbf{Scaled local mean-displacement invariance}
-			\item<8-> \textbf{Non-negative impulse responses}
-		\end{itemize}
-	\end{itemize}
+exploiting additional problem structure: \only<6->{\textbf{localized sensitivities}} \only<7->{and (local mean-)\textbf{displacement invariance}}
+\end{itemize}
+
 \end{frame}
@@
 		\begin{align*}
-		\phi_x := H_d~ \delta_x =& \text{ action of Hessian operator on delta distribution at x} \\
+		\phi_x := H_{\text{\tiny misfit}}~ \delta_x =& \text{ action of data-misfit Hessian operator on delta distribution at x} \\
 		\mu(x) :=& \text{ center of mass (mean) of }\phi_x
@@
 		%\includegraphics[width=0.6\columnwidth]{Hess1_ordered.png}
-	\end{center}
-\end{frame}
-
-\begin{frame}
-	\frametitle{Hessian impulse responses (Pine Island Glacier)}
-	\center
-	\only<1>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi0_minimal.png}
-		\end{center}
-	}
-	\only<2>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi1_minimal.png}
-		\end{center}
-	}
-	\only<3>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi2_minimal.png}
-		\end{center}
-	}
-	\only<4>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi3_minimal.png}
-		\end{center}
-	}
-	\only<5>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi4_minimal.png}
-		\end{center}
-	}
-	\only<6>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi5_minimal.png}
-		\end{center}
-	}
-	\only<7>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi6_minimal.png}
-		\end{center}
-	}
-	\only<8>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi7_minimal.png}
-		\end{center}
-	}
-	\only<9>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi8_minimal.png}
-		\end{center}
-	}
-	\only<10>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi9_minimal.png}
-		\end{center}
-	}
-	\only<11>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi10_minimal.png}
-		\end{center}
-	}
-	\only<12>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi11_minimal.png}
-		\end{center}
-	}
-	\only<13>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi12_minimal.png}
-		\end{center}
-	}
-	\only<14>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi13_minimal.png}
-		\end{center}
-	}
-	\only<15>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi14_minimal.png}
-		\end{center}
-	}
-	\only<16>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi15_minimal.png}
-		\end{center}
-	}
-	\only<17>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi16_minimal.png}
-		\end{center}
-	}
-	\only<18>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi17_minimal.png}
-		\end{center}
-	}
-	\only<19>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi18_minimal.png}
-		\end{center}
-	}
-	\only<20>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi19_minimal.png}
-		\end{center}
-	}
-	\only<21>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi20_minimal.png}
-		\end{center}
-	}
-	\only<22>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi21_minimal.png}
-		\end{center}
-	}
-	\only<23>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi22_minimal.png}
-		\end{center}
-	}
-	\only<24>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi23_minimal.png}
-		\end{center}
-	}
-	\only<25>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi24_minimal.png}
-		\end{center}
-	}
-	\only<26>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi25_minimal.png}
-		\end{center}
-	}
-	\only<27>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi26_minimal.png}
-		\end{center}
-	}
-	\only<28>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi27_minimal.png}
-		\end{center}
-	}
-	\only<29>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi28_minimal.png}
-		\end{center}
-	}
-	\only<30>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi29_minimal.png}
-		\end{center}
-	}
-	\only<31>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi30_minimal.png}
-		\end{center}
-	}
-	\only<32>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi31_minimal.png}
-		\end{center}
-	}
-	\only<33>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi32_minimal.png}
-		\end{center}
-	}
-	\only<34>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi33_minimal.png}
-		\end{center}
-	}
-	\only<35>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi34_minimal.png}
-		\end{center}
-	}
-	\only<36>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi35_minimal.png}
-		\end{center}
-	}
-	\only<37>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi36_minimal.png}
-		\end{center}
-	}
-	\only<38>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi37_minimal.png}
-		\end{center}
-	}
-	\only<39>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi38_minimal.png}
-		\end{center}
-	}
-	\only<40>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi39_minimal.png}
-		\end{center}
-	}
-	\only<41>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi40_minimal.png}
-		\end{center}
-	}
-	\only<42>{
-		\begin{center}
-			\includegraphics[width=0.70\columnwidth]{reordered_phi/phi41_minimal.png}
-		\end{center}
-	}
-	\only<43>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi42_minimal.png}
-		\end{center}
-	}
-	\only<44>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi43_minimal.png}
-		\end{center}
-	}
-	\only<45>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi44_minimal.png}
-		\end{center}
-	}
-	\only<46>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi45_minimal.png}
-		\end{center}
-	}
-	\only<47>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi46_minimal.png}
-		\end{center}
-	}
-	\only<48>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi47_minimal.png}
-		\end{center}
-	}
-	\only<49>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi48_minimal.png}
-		\end{center}
-	}
-	\only<50>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi49_minimal.png}
-		\end{center}
-	}
-	\only<51>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi50_minimal.png}
-		\end{center}
-	}
-	\only<52>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi51_minimal.png}
-		\end{center}
-	}
-	\only<53>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi52_minimal.png}
-		\end{center}
-	}
-	\only<54>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi53_minimal.png}
-		\end{center}
-	}
-	\only<55>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi54_minimal.png}
-		\end{center}
-	}
-	\only<56>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi55_minimal.png}
-		\end{center}
-	}
-	\only<57>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi56_minimal.png}
-		\end{center}
-	}
-	\only<58>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi57_minimal.png}
-		\end{center}
-	}
-	\only<59>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi58_minimal.png}
-		\end{center}
-	}
-	\only<60>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi59_minimal.png}
-		\end{center}
-	}
-	\only<61>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi60_minimal.png}
-		\end{center}
-	}
-	\only<62>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi61_minimal.png}
-		\end{center}
-	}
-	\only<63>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi62_minimal.png}
-		\end{center}
-	}
-	\only<64>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi63_minimal.png}
-		\end{center}
-	}
-	\only<65>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi64_minimal.png}
-		\end{center}
-	}
-	\only<66>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi65_minimal.png}
-		\end{center}
-	}
-	\only<67>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi66_minimal.png}
-		\end{center}
-	}
-	\only<68>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi67_minimal.png}
-		\end{center}
-	}
-	\only<69>{
-		\begin{center}
-			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi68_minimal.png}
-		\end{center}
-	}
-
-\end{frame}
-
-\begin{frame}
-	\frametitle{Impulse responses are ``columns'' of the integral kernel}
-	\begin{center}
-
-		\includegraphics[width=0.5\columnwidth]{kernel_impulse_illustration2.pdf}
-
 	\end{center}
@@
 
-%---------------------------------------------------------------------------------%
-\begin{frame}
+\begin{frame}[t]
+	\frametitle{Hessian impulse responses (Pine Island Glacier)}
+	\center
+	\only<1>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi0_minimal.png}
+		\end{center}
+	}
+	\only<2>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi1_minimal.png}
+		\end{center}
+	}
+	\only<3>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi2_minimal.png}
+		\end{center}
+	}
+	\only<4>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi3_minimal.png}
+		\end{center}
+	}
+	\only<5>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi4_minimal.png}
+		\end{center}
+	}
+	\only<6>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi5_minimal.png}
+		\end{center}
+	}
+	\only<7>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi6_minimal.png}
+		\end{center}
+	}
+	\only<8>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi7_minimal.png}
+		\end{center}
+	}
+	\only<9>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi8_minimal.png}
+		\end{center}
+	}
+	\only<10>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi9_minimal.png}
+		\end{center}
+	}
+	\only<11>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi10_minimal.png}
+		\end{center}
+	}
+	\only<12>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi11_minimal.png}
+		\end{center}
+	}
+	\only<13>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi12_minimal.png}
+		\end{center}
+	}
+	\only<14>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi13_minimal.png}
+		\end{center}
+	}
+	\only<15>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi14_minimal.png}
+		\end{center}
+	}
+	\only<16>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi15_minimal.png}
+		\end{center}
+	}
+	\only<17>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi16_minimal.png}
+		\end{center}
+	}
+	\only<18>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi17_minimal.png}
+		\end{center}
+	}
+	\only<19>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi18_minimal.png}
+		\end{center}
+	}
+	\only<20>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi19_minimal.png}
+		\end{center}
+	}
+	\only<21>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi20_minimal.png}
+		\end{center}
+	}
+	\only<22>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi21_minimal.png}
+		\end{center}
+	}
+	\only<23>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi22_minimal.png}
+		\end{center}
+	}
+	\only<24>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi23_minimal.png}
+		\end{center}
+	}
+	\only<25>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi24_minimal.png}
+		\end{center}
+	}
+	\only<26>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi25_minimal.png}
+		\end{center}
+	}
+	\only<27>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi26_minimal.png}
+		\end{center}
+	}
+	\only<28>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi27_minimal.png}
+		\end{center}
+	}
+	\only<29>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi28_minimal.png}
+		\end{center}
+	}
+	\only<30>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi29_minimal.png}
+		\end{center}
+	}
+	\only<31>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi30_minimal.png}
+		\end{center}
+	}
+	\only<32>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi31_minimal.png}
+		\end{center}
+	}
+	\only<33>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi32_minimal.png}
+		\end{center}
+	}
+	\only<34>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi33_minimal.png}
+		\end{center}
+	}
+	\only<35>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi34_minimal.png}
+		\end{center}
+	}
+	\only<36>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi35_minimal.png}
+		\end{center}
+	}
+	\only<37>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi36_minimal.png}
+		\end{center}
+	}
+	\only<38>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi37_minimal.png}
+		\end{center}
+	}
+	\only<39>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi38_minimal.png}
+		\end{center}
+	}
+	\only<40>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi39_minimal.png}
+		\end{center}
+	}
+	\only<41>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi40_minimal.png}
+		\end{center}
+	}
+	\only<42>{
+		\begin{center}
+			\includegraphics[width=0.70\columnwidth]{reordered_phi/phi41_minimal.png}
+		\end{center}
+	}
+	\only<43>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi42_minimal.png}
+		\end{center}
+	}
+	\only<44>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi43_minimal.png}
+		\end{center}
+	}
+	\only<45>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi44_minimal.png}
+		\end{center}
+	}
+	\only<46>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi45_minimal.png}
+		\end{center}
+	}
+	\only<47>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi46_minimal.png}
+		\end{center}
+	}
+	\only<48>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi47_minimal.png}
+		\end{center}
+	}
+	\only<49>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi48_minimal.png}
+		\end{center}
+	}
+	\only<50>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi49_minimal.png}
+		\end{center}
+	}
+	\only<51>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi50_minimal.png}
+		\end{center}
+	}
+	\only<52>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi51_minimal.png}
+		\end{center}
+	}
+	\only<53>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi52_minimal.png}
+		\end{center}
+	}
+	\only<54>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi53_minimal.png}
+		\end{center}
+	}
+	\only<55>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi54_minimal.png}
+		\end{center}
+	}
+	\only<56>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi55_minimal.png}
+		\end{center}
+	}
+	\only<57>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi56_minimal.png}
+		\end{center}
+	}
+	\only<58>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi57_minimal.png}
+		\end{center}
+	}
+	\only<59>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi58_minimal.png}
+		\end{center}
+	}
+	\only<60>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi59_minimal.png}
+		\end{center}
+	}
+	\only<61>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi60_minimal.png}
+		\end{center}
+	}
+	\only<62>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi61_minimal.png}
+		\end{center}
+	}
+	\only<63>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi62_minimal.png}
+		\end{center}
+	}
+	\only<64>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi63_minimal.png}
+		\end{center}
+	}
+	\only<65>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi64_minimal.png}
+		\end{center}
+	}
+	\only<66>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi65_minimal.png}
+		\end{center}
+	}
+	\only<67>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi66_minimal.png}
+		\end{center}
+	}
+	\only<68>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi67_minimal.png}
+		\end{center}
+	}
+	\only<69>{
+		\begin{center}
+			\includegraphics[width=0.7\columnwidth]{reordered_phi/phi68_minimal.png}
+		\end{center}
+	}
+
+\end{frame}
+
+\begin{frame}
+	\frametitle{Impulse responses are ``columns'' of the integral kernel}
+	\begin{center}
+
+		\includegraphics[width=0.5\columnwidth]{kernel_impulse_illustration2.pdf}
+
+	\end{center}
+\end{frame}
+
+%---------------------------------------------------------------------------------%
+\begin{frame}[t]
 	\frametitle{Hessian approximation method: big idea}
@@
 		\item {\bf Step 2:} Interpolate known impulse responses to approximate unknown Hessian entries $\mathbf{H}_{ij}$
-		\item {\bf Step 3:} Convert to $\mathcal{H}$-matrix to do linear algebra
-	\end{itemize}
-	\includegraphics[width=0.45\columnwidth]{impulse_batch1.png}  \includegraphics[width=0.45\columnwidth]{impulse_batch2.png}
+		\item {\bf Step 3:} Convert to $\mathcal{H}$-matrix to do linear algebra (with low computational complexity)
+	\end{itemize}
+    \begin{figure} 
+	\includegraphics[width=0.375\columnwidth]{impulse_batch1.png} \hfil  
+	\includegraphics[width=0.375\columnwidth]{impulse_batch2.png}
+	\end{figure} 
 	*Impulse response batches from Ice Mountain
@@
 %---------------------------------------------------------------------------------%
-%---------------------------------------------------------------------------------%
-\begin{frame}
-	\frametitle{Technical details}
-	{\Large
-		\begin{itemize}
-			\setlength\itemsep{2em}
-			\item How do we choose the impulse response points?
-			\begin{itemize}
-				\item {\large How do we make sure they don't overlap?}
-			\end{itemize}
-			\item How do we interpolate the impulse responses? 
-			\begin{itemize}
-				\item {\large What about boundary issues?}
-			\end{itemize}
-		\end{itemize}
-	}
-\end{frame}
-%---------------------------------------------------------------------------------%
-\begin{frame}
+
+%---------------------------------------------------------------------------------%
+\begin{frame}[t]
 	\frametitle{How to choose impulse response points?}
-	One hessian matrix-vector product $\rightarrow$ many impulse responses
-	\begin{center}
-		\includegraphics[width=0.5\columnwidth]{impulse_batch2.png} 
-	\end{center}
-	\begin{itemize}
-		\item \textbf{Goal:} choose as many points as possible, such that the impulse response supports don't overlap
-		\item \textbf{Dilemma:} How can we know the impulse response supports before we compute them?
-	\end{itemize}
+
+	\begin{columns}
+		\begin{column}{0.5\columnwidth}
+			One Hessian-vector product can yield many impulse responses (with small region of support)
+		\begin{center}
+			\includegraphics[width=0.75\columnwidth]{impulse_batch2.png} 
+		\end{center}
+	    \end{column}
+    \begin{column}{0.49\columnwidth}
+    	\only<2>
+    	{ 
+		\begin{itemize}
+			\item \textbf{Goal:} determine as many points as possible, such that the associated impulse response supports do not overlap
+			\item \textbf{Dilemma:} how can we know the impulse response supports before we compute them?
+		\end{itemize}
+	    }
+	\end{column} 
+	\end{columns}
+
 \end{frame}
@@
 
-	$$\mathbf{A}^T \begin{bmatrix}
+	$$\mathbf{A}^{\top} \begin{bmatrix}
 	1 \\ 1 \\ \vdots \\ 1
@@
 
-	Apply matrix to vector of ones $\rightarrow$ get row sums for all rows
+	Sum of each row obtained by transpose apply to vector of ones
 
@@
 
-	$$(H_d^T C)^*(y) = \int_\Omega \left(H_d \delta_y\right)(x) dx$$
-
-	Apply Hessian to constant function $\rightarrow$ get volumes of every impulse response
+	$$(H_{\text{\tiny misfit}}^{\dagger} C)^*(y) = \int_\Omega \left(H_{\text{\tiny misfit}} \delta_y\right)(x) \,\mathrm{d}x=\int_{\Omega}\phi_{y}(x)\,\mathrm{d}x$$
+
+	Volume of each impulse response obtained by adjoint Hessian operator apply to constant function
 \end{frame}
@@
 		\begin{align*}
-		V =& \left(H_d^T C\right)^* \\
-		\mu^i =& \left(H_d^T L^i\right)^* / V \\
-		\Sigma^{ij} =& \left(H_d^T Q^{ij}\right)^* / V - \mu^i\cdot \mu^j
+		V =& \left(H_{\text{\tiny misfit}}^{\dagger} C\right)^* \\
+		\mu^i =& \left(H_{\text{\tiny misfit}}^{\dagger} L^i\right)^* / V \\
+		\Sigma^{ij} =& \left(H_{\text{\tiny misfit}}^{\dagger} Q^{ij}\right)^* / V - \mu^i\cdot \mu^j
 		\end{align*}
-		\item Apply Hessian to constant, linear, and quadratic functions $\rightarrow$ get estimates of support for every impulse response
+		\item Estimates of impulse response supports obtained from 
+		first few moments (adjoint Hessian apply to constant, linear, and quadratic functions)
 	\end{itemize}
@@
 	\begin{center}
-		\includegraphics[width=0.5\columnwidth]{impulse_batch2.png} 
-	\end{center}
-\end{frame}
-
-\begin{frame}
-	\frametitle{Radial basis function interpolation}
-	\begin{center}
-		\includegraphics[width=0.8\textwidth]{interpolation_rbf.pdf}
-	\end{center}
-	\begin{itemize}
-		\item Interpolate impulse responses using polyharmonic spline radial basis functions.
-		\item Use only $k$-nearest neighbors (must solve $k \times k$ linear system)
-	\end{itemize}
+		\includegraphics[width=0.375\columnwidth]{impulse_batch2.png} 
+	\end{center}
 \end{frame}
@@
 		\item \textbf{Hessian-vector products:} 
-		$$6 + n_\text{batches}$$
-		E.g., $11$ Hessian-vector products for $5$ batches of impulse responses
+		$$1+\text{dim}(\Omega)(\text{dim}(\Omega)+3)/2 + n_\text{batches}$$
+		E.g., $11$ Hessian-vector products for $5$ batches of impulse responses and $\text{dim}(\Omega)=2$
 		\item \textbf{Ellipsoid intersection tests:}
@@
 		where $m$ is total number of impulse responses in all batches
-		\item \textbf{Elementary operations} to build and use the H-matrix:
+		\item \textbf{Elementary operations} to build and use the $\mathcal{H}$-matrix:
 		$$O(N \log N)$$ 
@@
 \begin{frame}
-	\frametitle{Example: Ice sheet inverse problem}
+	\frametitle{Example: ``Ice mountain'' inverse problem}
 
@@
 \begin{frame}
-	\frametitle{Example: Ice sheet inverse problem}
+	\frametitle{Example: ``Ice mountain'' inverse problem}
 	\framesubtitle{Convergence history for solving the Stokes inverse problem using inexact Newton PCG}
@@
 		\item \#CG:  the number of PCG iterations used to solve the Newton system.
-		\item \#Stokes:the total number of Stokes PDE solves performed in each Newton iteration.
-		\item $\|\mathbf{g}\|$: the $l^2$ norm of the gradient.
+		\item \#Stokes: the total number of Stokes PDE solves performed in each Newton iteration.
+		\item $\|\mathbf{g}\|$: the Euclidean norm of the gradient.
 	\end{itemize}
@@
 		prohibitive as the data becomes more informative, as is the case
-		for ice sheet inverse problems.
+		for continental-scale ice sheet inverse problems.
 		\vspace{0.05in}
-		\item Local point spread function interpolation combined with Hierarchical matrix representations promise a more efficient
+		\item Local point spread function interpolation combined with hierarchical matrix representations promise a more efficient
 		Hessian approximation.

### 2024-03-06 presentations/siampp24-tucker/talk.tex
small changes

@@
 \title[Local PSF high-rank Hessian approximation]{Local Point Spread Function Approximation of High-Rank Hessians in Inverse Problems Governed by	PDEs\thanks{Research supported
-    partially by NSF-DMS-CAREER-1654311 and NSF-OAC-1550547.}}
+    partially by NSF-DMS-CAREER-1654311, NSF-DMS-184026 and NSF-OAC-1550547.}}
 %\author[{Noemi
@@
 		\vspace{0.02in}
-		\item Exploiting structured low-rank properties of (Bayesian) inverse problems governed by PDEs
+		\item Exploiting properties of (Bayesian) inverse problems governed by PDEs
 		\vspace{0.02in}
@@
 			\setlength\itemsep{1.0em}
-			\item (Global) low rank approximation
+			\item Global low rank structure
 			\vspace{0.02in}
-			\item Hierarchical off-diagonal low rank (HODLR) approximation
+			\item Hierarchical off-diagonal low rank
+		\end{itemize}
 				\vspace{0.02in}
 	\item A local point spread function method as a means for efficient
-			$\mathcal{H}$ierarchical-matrix approximation
-		\end{itemize}
+	hierarchical-matrix approximation
 		\vspace{0.02in}
